### 2017-08-29 README3.Rmd
added Rmd default page

---
title: "MLH1_md"
author: "April"
date: "August 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-08-29 README3.Rmd
updating Rmd page

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup, include=FALSE, echo=FALSE}
+library(plyr) 
+library(gplots)
+library(ggplot2)
+library(knitr) #for md tables 
+#read in working environment that was already set
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="MLH1_data_setup.RData")
+
+##FIX ORDER HERE
+MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
+                                     "WSB", "PWD","MSM","CAST"), order=T )
+#order the data frame
+MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+#set the order of another column, based on another variable (so that when)
+library(dplyr)
+MLH1_data <- MLH1_data %>%
+  arrange(strain, sex, mouse) %>%
+  mutate(file.name = factor(file.name)) #another category that you want the order to match
+
+# sort your dataframe, by the focal categories
+MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
+#save a figure of boxplots for each mouse
+ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
+  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )
+
+ggsave("boxplots_mice.png")
+
+MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
+MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
+
+ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
+ggsave("Female_boxplots.png")
+
+ggplot(MLH1_by_M_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
+ggsave("Male_boxplots.png")
 ```
 
-## R Markdown
+This is where the plot is discribed.
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+```{r setup, include=FALSE, echo=FALSE}                 
+#Quality / number of passing mice
+q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
+                        total =  length(nMLH1.foci),
+                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
+                        q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
+)
+#q_cutoff_table
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+passed_mice <- as.data.frame(q_cutoff_table[q_cutoff_table$q_l3 >= 15,]$mouse, 
+                             col.names= "mouse")
+colnames(passed_mice) <- c("mouse")
 
-```{r cars}
-summary(cars)
+passed_mice$strain <- ifelse(grepl("WSB", passed_mice$mouse), "WSB", 
+                   ifelse(grepl("G", passed_mice$mouse), "G",
+                    ifelse(grepl("CAST", passed_mice$mouse), "CAST",
+                   ifelse(grepl("MSM", passed_mice$mouse), "MSM",
+                  ifelse(grepl("LEW", passed_mice$mouse), "LEWES", 
+                   ifelse(grepl("LEWES", passed_mice$mouse), "LEWES",                                          
+                  ifelse(grepl("PWD", passed_mice$mouse), "PWD", "other")))))))
+
+passed_mice$sex <- ifelse(grepl("_f", passed_mice$mouse), "female",                                         
+                          ifelse(grepl("_m", passed_mice$mouse), "male", "other"))
+
+#make a table of the number of mice passed for each category.
+Pass_mice_category <- ddply(passed_mice, .(strain, sex), summarise,
+                    "Passing Mice" =  length(mouse)
+)
+
+not_passed_mice <- setdiff(q_cutoff_table$mouse, passed_mice$mouse)
+not_passed_mice <- data.frame(not_passed_mice)
+colnames(not_passed_mice) <- c("mouse")
+
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+
+#convert tables into md useing Kable
+Pass_mice_category_kable <- kable(Pass_mice_category, format = "markdown")
+passed_mice_kable <- kable(passed_mice, format = "markdown")
+not_passed_mice_kable <- kable(not_passed_mice, format = "markdown")
+
 ```
 
-## Including Plots
+#### Tables of Passing mice
+Pass_mice_category_kable
 
-You can also embed plots, for example:
+another table
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+passed_mice_kable
+```{r setup, include=FALSE, echo=FALSE}
+
+passed_mice_kable
+
 ```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-08-29 README3.Rmd
updating Rmd page

@@
 author: "April"
-date: "August 29, 2017"
+date: "August 25, 2017"
 output: html_document
@@
 ```{r setup, include=FALSE, echo=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 library(plyr) 
-library(gplots)
 library(ggplot2)
-library(knitr) #for md tables 
+```
+
+```{r }
+
+#library(knitr) #for md tables 
 #read in working environment that was already set
@@
 
+```
+
+```{r orderdf, echo=FALSE}
 ##FIX ORDER HERE
@@
 
-#save a figure of boxplots for each mouse
-ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
-  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )
-
-ggsave("boxplots_mice.png")
-
 MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
@@
 ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
-ggsave("Female_boxplots.png")
 
 ggplot(MLH1_by_M_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
-ggsave("Male_boxplots.png")
+```
+
+### PLOTS
+
+```{r plots, echo=FALSE}
+#save a figure of boxplots for each mouse
+bb <- ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
+  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) ) 
+bb
+
+ff <- ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
+ff
+
 ```
@@
 passed_mice_kable
-```{r setup, include=FALSE, echo=FALSE}
+```{r setup, include=FALSE}
 

### 2017-08-29 rmd_tutorial.rmd
tryiung the tutorial

---
title: "kableExtraTutorial"
author: "April"
date: "August 29, 2017"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
#install.packages("kableExtra")

# For dev version
# install.packages("devtools")
#devtools::install_github("haozhu233/kableExtra")
```

```{r}
library(knitr)
library(kableExtra)
dt <- mtcars[1:5, 1:6]

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

```

options(knitr.table.format = "html") 
## If you don't define format here, you'll need put `format = "html"` in every kable function.

```{r}
kable(dt)

kable(head(MLH1_data) )
```
```{r}
##FIX ORDER HERE
MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
                                     "WSB", "PWD","MSM","CAST"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
#set the order of another column, based on another variable (so that when)
library(dplyr)
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(file.name = factor(file.name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
```

```{r}
#save a figure of boxplots for each mouse
ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )

```

```{r}
#Quality / number of passing mice
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table

passed_mice <- as.data.frame(q_cutoff_table[q_cutoff_table$q_l3 >= 15,]$mouse, 
                             col.names= "mouse")
colnames(passed_mice) <- c("mouse")

passed_mice$strain <- ifelse(grepl("WSB", passed_mice$mouse), "WSB", 
                   ifelse(grepl("G", passed_mice$mouse), "G",
                    ifelse(grepl("CAST", passed_mice$mouse), "CAST",
                   ifelse(grepl("MSM", passed_mice$mouse), "MSM",
                  ifelse(grepl("LEW", passed_mice$mouse), "LEWES", 
                   ifelse(grepl("LEWES", passed_mice$mouse), "LEWES",                                          
                  ifelse(grepl("PWD", passed_mice$mouse), "PWD", "other")))))))

passed_mice$sex <- ifelse(grepl("_f", passed_mice$mouse), "female",                                         
                          ifelse(grepl("_m", passed_mice$mouse), "male", "other"))

#make a table of the number of mice passed for each category.
Pass_mice_category <- ddply(passed_mice, .(strain, sex), summarise,
                    "Passing Mice" =  length(mouse)
)

not_passed_mice <- setdiff(q_cutoff_table$mouse, passed_mice$mouse)
not_passed_mice <- data.frame(not_passed_mice)
colnames(not_passed_mice) <- c("mouse")

```

```{r}
kable(passed_mice)
kable(not_passed_mice)
kable(Pass_mice_category)
```

```{r}
kable(dt) %>%
  kable_styling()
```

### 2017-08-29 rmd_tutorial.rmd
remove the swd argument
git output

@@
 date: "August 29, 2017"
-output: html_document
+output: github_document
 ---
-
-```{r setup, include=FALSE, echo=FALSE}
-#install.packages("kableExtra")
-
-# For dev version
-# install.packages("devtools")
-#devtools::install_github("haozhu233/kableExtra")
-```
 
@@
 library(kableExtra)
+library(ggplot2)
+library(dplyr)
+library(dplyr)
 dt <- mtcars[1:5, 1:6]
 
-setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
 load(file="MLH1_data_setup.RData")
@@
 
-```{r}
-kable(dt)
-
-kable(head(MLH1_data) )
-```
 ```{r}
@@
 )
-#q_cutoff_table
+q_cutoff_table
 

### 2017-08-29 rmd_tutorial.rmd
deleted files

@@
 title: "kableExtraTutorial"
-author: "April"
+author: "April Peterson"
 date: "August 29, 2017"
@@
 
-```{r}
+```{r setup, echo=FALSE}
 library(knitr)
@@
 library(dplyr)
-library(dplyr)
-dt <- mtcars[1:5, 1:6]
 
@@
 
-options(knitr.table.format = "html") 
-## If you don't define format here, you'll need put `format = "html"` in every kable function.
+This Rmd file displays the boxplots of MLH1 distributions.
 
-```{r}
+```{r, echo=FALSE}
 ##FIX ORDER HERE
@@
 #set the order of another column, based on another variable (so that when)
-library(dplyr)
+
 MLH1_data <- MLH1_data %>%
@@
   mutate(file.name = factor(file.name)) #another category that you want the order to match
-
 # sort your dataframe, by the focal categories
 MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
+MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
+MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
 ```
 
-```{r}
+Below is a figure of boxplots all mice from current data batches.
+
+```{r, echo=FALSE}
 #save a figure of boxplots for each mouse
-ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
+bb <- ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
   geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) )
+
+bb
+```
+
+Below are the boxplots for female MLH1 distributions.
+
+```{r, echo=FALSE}
+ff <- ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
+ff
 
@@
 
-```{r}
-#Quality / number of passing mice
-q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
-                        total =  length(nMLH1.foci),
-                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
-                        q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
-)
-q_cutoff_table
+Below are the boxplots for male MLH1 distributions.
 
-passed_mice <- as.data.frame(q_cutoff_table[q_cutoff_table$q_l3 >= 15,]$mouse, 
-                             col.names= "mouse")
-colnames(passed_mice) <- c("mouse")
-
-passed_mice$strain <- ifelse(grepl("WSB", passed_mice$mouse), "WSB", 
-                   ifelse(grepl("G", passed_mice$mouse), "G",
-                    ifelse(grepl("CAST", passed_mice$mouse), "CAST",
-                   ifelse(grepl("MSM", passed_mice$mouse), "MSM",
-                  ifelse(grepl("LEW", passed_mice$mouse), "LEWES", 
-                   ifelse(grepl("LEWES", passed_mice$mouse), "LEWES",                                          
-                  ifelse(grepl("PWD", passed_mice$mouse), "PWD", "other")))))))
-
-passed_mice$sex <- ifelse(grepl("_f", passed_mice$mouse), "female",                                         
-                          ifelse(grepl("_m", passed_mice$mouse), "male", "other"))
-
-#make a table of the number of mice passed for each category.
-Pass_mice_category <- ddply(passed_mice, .(strain, sex), summarise,
-                    "Passing Mice" =  length(mouse)
-)
-
-not_passed_mice <- setdiff(q_cutoff_table$mouse, passed_mice$mouse)
-not_passed_mice <- data.frame(not_passed_mice)
-colnames(not_passed_mice) <- c("mouse")
-
+```{r, echo=FALSE}
+mm <- ggplot(MLH1_by_M_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )
+mm
 ```
 
-```{r}
-kable(passed_mice)
-kable(not_passed_mice)
-kable(Pass_mice_category)
-```
+Creating Tables with ddply is currently broken Rmd.
 
 ```{r}
-kable(dt) %>%
-  kable_styling()
+
+#q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
+ #                       total =  length(nMLH1.foci),
+  #                      q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
+   #                     q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE)
+#)
+#q_cutoff_table
 ```

### 2017-09-17 rmd_tutorial.rmd
added notes for improving Rmd report

@@
 
+include is for blocking output messages. echo false is for silencing the code commands.
 ```{r setup, echo=FALSE}
 library(knitr)
-library(kableExtra)
 library(ggplot2)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.width=3.5, fig.height = 2}
 #save a figure of boxplots for each mouse
@@
 ```
+
+ figure will be generated below. this is the text above that prefaces it, how and why the figure(s) are set up.
+ The following figure will display the boxplots of MLH1 distributions seperated by sexes. 
+```{r, MLH1 box plots, fig.width=2, fig.height=3, fig.cap="this is the figure caption"}
+
+```
+This is the text that discusses the inferences of the figures.
+
+This is the text above the table, this will show category wide stats of MLH1 counts.
+```{r}
+#use kable to make tables. warning this might not work for github Rmd.
+kable(my table)
+```
+
+text before the table for mouse specific tables. (make them smaller and more compact?)
+```{r, mouse table}
+
+```
+
+```{r, MxF hetC plot}
+
+```

### 2017-09-19 rmd_tutorial.rmd
changes to rmd

--- 2017-08-29 rmd_tutorial.rmd
@@
 
-```{r setup, echo=FALSE}
+discription of data and report here
+
+```{r setup, echo=FALSE, include=FALSE}
 library(knitr)
@@
 
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
 load(file="MLH1_data_setup.RData")
@@
 ```{r, echo=FALSE}
+#todo, remove bottom axis
 ##FIX ORDER HERE

### 2017-09-19 rmd_tutorial_REMOTE_3696.rmd

---
title: "MLH1 data report "
author: "April Peterson"
date: "August 29, 2017"
output: github_document
---

TODO
data set up page; remove non quality observations

include is for blocking output messages. echo false is for silencing the code commands.
```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(lattice)

#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

```

#### goals of research
Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice (rodents). Comparisons the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve. (insert small picture of meixyte)

### Goal / Discription of Report
Goal of this report;

> *to keep track of the number of images for each category and mouse*

This is important for getting an idea of the 'power' I have to accurately estimate MLH1 distributions across the categories of cells I am quantifying.

#### Discription of data and quantification process
data frames are loaded of MLH1 data 
imagename, mouse name, strain, sex, nMLH1.

quanti process. anonymized, given quality score

This is the text above the table, this will show category wide stats of current MLH1 data set.
```{r, fig.height=2, fig.width=4, echo=FALSE}
#use kable to make tables. warning this might not work for github Rmd.
kable(AP_strain_table)
```

### disctibutions based on quality

Assess pattern of quality score and nMLH1. why? The human scoring the cells may have introcudced biases by giving a higher quality to cells with more foci. Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.

Quality score is 1 based, to a correlation of nMLH1 and quality will have a right dowward line.
```{r, scatter plots of nMLH1 by score, echo=FALSE}
#make lattice / plots seperated by strain and sex
#facet
qual_means <- MLH1_data %>% 
        group_by(category, quality) %>% 
        summarise(
          adj_nMLH1.foci = mean(adj_nMLH1.foci)
        )

p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
p <- p + geom_point(stat = "identity")
p <- p  + geom_point(data = qual_means, colour="red") 
p <- p + facet_wrap(~ category)
p

```
hmm, not sure if scatter plots are the best for assessing this due to the discrite nature of nMLH1...
Red point displays the

I ploted the mean of each quality bin with a red dot. From the pattern of the red dots, there is definately a negative relationship with quality and nMLH1 foci across the data. This is most pronounced in MSM males and least pronounced in G males.

Potential solutions? admit that the quality criteria are stupid..? does RW's data have this pattern? should the data be transformed/adjusted? Will looking at these figures bias how I score here on out?!

> *do some mice have better quality cells?* *Does this cause bias in the data?*

```{r, fig.height=8, fig.width=4, echo=FALSE, fig.caption="histogrms"}
# do the same thing as above but geom_hist

histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
histM <- histM + geom_histogram(stat = "bin", binwidth = 1)
histM <- histM + facet_wrap(~ category, ncol=2)
histM

```
bin width was adjusted to 1. x axis scale was adjusted.
It seems like there is a population of lower cells in MSM males. This could be from slides which had bad staining.

### power calqs
Find those old R scripts and recreate here.

### initial inferences of nMLH1 Patterns
(make boxplots, note differences, include caveats for low numbers or quality differences)

This Rmd file displays the boxplots of MLH1 distributions.

```{r, echo=FALSE, include = FALSE}
##FIX ORDER HERE (this should be put in the data setup script)
MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
                                     "WSB", "PWD","MSM","CAST"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
#set the order of another column, based on another variable (so that when)

MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(file.name = factor(file.name)) #another category that you want the order to match
# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
```

Below is a figure of boxplots all mice from current data batches.

```{r, echo=FALSE, fig.width=3.5, fig.height = 2}
# TODO, custom color the strains
bb <- ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + 
  geom_boxplot(data = MLH1_data, aes(fill = factor(strain), color = factor(sex) ) ) + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

bb
```

#### Assessing the number of cells per mouse

In this analysis each mouse will have a sample of cells that are used to estimate the *ture* distribution of nMLH1 foci per cell. Power calculations can be done to assess if the current sample of cells gives a good estimate of the true distrubtion of nMLH1 foci.

Below are the boxplots for female MLH1 distributions.

```{r, echo=FALSE}
ff <- ggplot(MLH1_by_F_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ff

```

Below are the boxplots for male MLH1 distributions.

```{r, echo=FALSE, fig.caption="boxplots of MLH1 distributions"}
mm <- ggplot(MLH1_by_M_strain, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_by_F_strain, aes(fill = factor(strain) ) )+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
mm
```

```{r}

#q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
 #                       total =  length(nMLH1.foci),
  #                      q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
   #                     q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE)
#)
#q_cutoff_table
```

 figure will be generated below. this is the text above that prefaces it, how and why the figure(s) are set up.
 The following figure will display the boxplots of MLH1 distributions seperated by sexes. 
```{r, MLH1 box plots, fig.width=2, fig.height=3, fig.cap="this is the figure caption"}

```

text before the table for mouse specific tables. (make them smaller and more compact?)
```{r, mouse table}

```

```{r, MxF hetC plot}

```

### 2017-09-21 rmd_report.rmd
rmd report in first draft

---
title: "MLH1 data report "
author: "April Peterson"
date: "August 29, 2017"
output: github_document
---

TODO
remove the output for the plots
data set up page; remove non quality observations
Solve the cells per mouse problem
figure out figure captions
add a place to record the 'batches' that make up the data
category has CAST female, this might be messing things up, since there is not CAST data

include is for blocking output messages. echo false is for silencing the code commands.
```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
library(pwr)
library(plyr)
library(lattice)
library(dplyr)
#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

```

#### goals of research
Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice (rodents). Comparisons the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve. (insert small picture of meixyte)

### Goal / Discription of Report
Goal of this report;

> *to keep track of the number of images for each category and mouse*

This is important for getting an idea of the 'power' I have to accurately estimate MLH1 distributions across the categories of cells I am quantifying.

#### Discription of data and quantification process
data frames are loaded of MLH1 data 
imagename, mouse name, strain, sex, nMLH1.

quanti process. anonymized, given quality score

```{r, echo=FALSE, include = FALSE}
##FIX ORDER HERE (this should be put in the data setup script)
MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
                              "WSB", "PWD","MSM","CAST"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
#set the order of another column, based on another variable (so that when)

MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(file.name = factor(file.name)) #another category that you want the order to match
# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

#Add fake CAST female data

last_row <-data.frame(file.name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )

MLH1_data <- rbind(MLH1_data, last_row)

#count the non-quality measures,  #remove non qualit
#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]

MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]

```

#### Initial Patterns from MLH1 distributions

Use these figures to identify outliers and if there is variance in the means and variance.

```{r first boxplots, echo=FALSE, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, fig.align='center'}

ff <- ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_data, aes(fill = factor(strain) ) )+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
  "coral1", "#E69F00", "yellowgreen"))
ff <- ff + facet_wrap(~ sex)

```

### Tests for quality and nMLH1 foci number

> *How do the distributions change across different quality scores?*

> *How do the distributions change across mice with a good number of cells?*

Boxplots of mice distributions with quality score of 1

```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE}

#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work

highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )

qq <- ggplot(highest_qual_df, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
  "coral1", "#E69F00", "yellowgreen"))
high_qual_boxplot <- qq + facet_wrap(~ sex) #this is breaking the kniting...

```

The spacing in these boxplots is off because of 'ghost' cells, which are in the data frame, but not in the subsetted data. (Iguess)

Table for assessing the number of passed mice
```{r, passing mice, echo=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table

mice_25cells <- q_cutoff_table[q_cutoff_table$above5 > 25,]$mouse
#12 mice have more than 25 cells
mice_20cells <- q_cutoff_table[q_cutoff_table$above5 > 20,]$mouse
#26 mice have over 20 cells
mice_15cells <- q_cutoff_table[q_cutoff_table$above5 > 14,]$mouse

passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)

not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )

#mnake dataframe with just passing mice -- so that it can be ploted
passed_mice_df <- MLH1_data[ ( MLH1_data$quality < 5), ]

```

#### Boxplots of 'passing' mice distributions

```{r, echo=FALSE}
pp <- ggplot(passed_mice_df, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = passed_mice_df, aes(fill = factor(strain) ) )+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
  "coral1", "#E69F00", "yellowgreen"))
passed_mice_plot <- pp + facet_wrap(~ sex)

```
```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}

ff
high_qual_boxplot
passed_mice_plot
```

##### Effects of quality

Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality. Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.

```{r, scatter plots of nMLH1 by score, echo=FALSE, fig.cap="caption"}
#make lattice / plots seperated by strain and sex
#facet

#quality scores are now 1:7?
qual_means <- MLH1_data %>% 
        group_by(category, quality) %>% 
        summarise(
          adj_nMLH1.foci = mean(adj_nMLH1.foci)
        )

p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
p <- p + geom_point(stat = "identity")
p <- p  + geom_point(data = qual_means, colour="red") 
p <- p + facet_wrap(~ category)
p

```

I ploted the mean of each quality bin with a red dot. From the pattern of the red dots, there is definately a negative relationship with quality and nMLH1 foci across the data. This is most pronounced in MSM males and least pronounced in G males. The CAST female data is dummy data.

Potential solutions?

> *do some mice have better quality cells?* *Does this cause bias in the data?*

```{r, fig.height=4, fig.width=4, echo=FALSE, fig.caption= "histogrms", warning=FALSE}
# do the same thing as above but geom_hist

histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
histM <- histM + geom_histogram(stat = "bin", binwidth = 1)
histM <- histM + facet_wrap(~ category)
histM
```

It seems like there is a population of lower cells in MSM males. This could be from slides which had bad staining.

#### Assessing the number of cells per mouse

latice plot of scatter plots for jitter plots of cell oberservations by quality
```{r, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
#TODO, add a line for mouse and strain

#this creates a slot for CAST female, but no data
#each of the indiviual dataframes can be acessed with df_list$ ---- category
# for each sub-data frame, make mouse specific table,
#
#add mouse mean to the df
df_list <- split(MLH1_data, as.factor(MLH1_data$category))

o=0
i=0

for(i in 1:(length(df_list))){
 # print(i)
  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
                       mouse_nMLH1 = mean(adj_nMLH1.foci)
  )
 # print(sub_mouse_means)
  #below should work -- I think I need to go througheach row of the df_list[[i]]
  for(o in 1:length(df_list[[i]]$file.name)){
    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
        }
}
#remove CAST, this column isn't added to MLH1 col, it's added to df_list
#The above adds a column of mouse specific averages to a giant list of dataframes. Each dataframe in the list is specific to a category.
#the next step is to make plots from this giant list of dataframes
# I can delete all of the code below that by hand makes

### not make a loop for ggplot... or giant ggplot base that uses latice smartly
# make new ggplots -- that are replaces

plot_list = list()
for (h in 1:length(df_list)) {
  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
  pf <- pf + geom_point(stat = "identity")+ geom_jitter()
  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red")
  plot_list[[h]] =   pf
}

#aes must be same length --- 
# the last one is correct....?

#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
```

```{r, echo=FALSE, message=FALSE, warning=FALSE }
#geom_hline(aes(yintercept = med, group = gr), colour = 'red')
for (gg in 1:length(plot_list) ){
  show(plot_list[gg])
}
```

Making all of these scatter plots, allows us to look at the whole distributions. In some mice there seems to be bimodel cell populations, those of clearly high quality and those of lower quality. This may be due to the screeign of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.

#### Power Calculations

Main questions to address:

 > *How many cells per mouse should be sampled?*

 > *How many mice per strain should be sampled?*

Think about a loop that will -- i) take ~100 different sized samples of total images (10%, 25%, 50%, 75%, 90%) (with varying degrees of quality)
ii) then run t.tests across those samples, and report the p value
iii) make a table of the p values for these sample permutations

Diminishing return analysis. Take some of the mice identified above, simulate random sampling of 10,15,20,30. Display the raw mean and sd, and results from 10 simulations.
```{r, comparing sample sizes, echo=FALSE}
PWD_m_example = MLH1_data[MLH1_data$mouse == '10mar15_PWD_m2', ]

smp10dist = sample(PWD_m_example$nMLH1.foci, 10, replace=FALSE)
samplesize = rep("10", 10)
smp10 <- data.frame(smp10dist, samplesize)

smp15dist = sample(PWD_m_example$nMLH1.foci, 15, replace=FALSE)
samplesize = rep("15", 15)
smp15 <- data.frame(smp15dist, samplesize)

smp20dist = sample(PWD_m_example$nMLH1.foci, 20, replace=FALSE)
samplesize = rep("20", 20)
smp20 <- data.frame(smp20dist, samplesize)

smp30dist = sample(PWD_m_example$nMLH1.foci, 30, replace=FALSE)
samplesize = rep("30", 30)
smp10 <- data.frame(smp30dist, samplesize)

#this mouse has 40 cells 
#make a dataframe (so that I can finally make boxplots)
dat <- data.frame(
  MLH1_sample = c(sample(PWD_m_example$adj_nMLH1.foci, 10, replace=FALSE),  sample(PWD_m_example$adj_nMLH1.foci, 15, replace=FALSE),    sample(PWD_m_example$adj_nMLH1.foci, 20, replace=FALSE), sample(PWD_m_example$adj_nMLH1.foci, 30, replace=FALSE), sample(PWD_m_example$adj_nMLH1.foci, 35, replace=FALSE), sample(PWD_m_example$adj_nMLH1.foci, 40, replace=FALSE)),   
  samplesize = c(rep("10", 10), rep("15", 15), rep("20", 20), rep("30", 30), rep("35", 35), rep("40", 40)) )

bb <- ggplot(data = dat, aes(x=samplesize, y=MLH1_sample)) + geom_boxplot(aes(fill=samplesize))

```
#omg finally that took so long. main interpertations: the variance grows with larger sample size.
#these are representations of independant samples -- I need to think of a way to show an average of the stats for multiple simulations.

```{r}
G_f_example = MLH1_data[MLH1_data$mouse == '17mar16_G_f1', ]

datf <- data.frame(
  MLH1_sample = c(sample(G_f_example$nMLH1.foci, 10, replace=FALSE),  sample(G_f_example$nMLH1.foci, 15, replace=FALSE),    sample(G_f_example$nMLH1.foci, 20, replace=FALSE), sample(G_f_example$nMLH1.foci, 30, replace=FALSE), sample(G_f_example$nMLH1.foci, 35, replace=FALSE) ),   
  samplesize = c(rep("10", 10), rep("15", 15), rep("20", 20), rep("30", 30), rep("35", 35)) )

v <- ggplot(data = datf, aes(x=samplesize, y=MLH1_sample)) + geom_boxplot(aes(fill=samplesize) ) 
v                                                                         
v + ggtitle("varying sample sizes") 
```
Can the above code be repeated for a larger number of mice (both passing and non-passing) to show the different effects across samples sizes?
power calculations.... from a preivous document --- I showed myself that for 15 (good) cells, would provide a sample big enough to detect a difference of 3 foci. I only did this for a couple mice, I'll try to re-code for every mouse

```{r}
passed_mice_data <- MLH1_data[MLH1_data$mouse %in% passed_mice, ]
non_passed_mice_data <- MLH1_data[MLH1_data$mouse %in% not_passed, ]

#make a column to mark if observation comes from passed or not passed mouse
#turn this into boxplot and sort the category order, otherwise it's hard to any patterns
sc <- qplot(mouse, nMLH1.foci,  colour = category, data = non_passed_mice_data, main="s", xlab="s", ylab="es")
sc + theme(legend.position="none")

sc <- qplot(mouse, nMLH1.foci,  colour = category, data = passed_mice_data, main="s", xlab="s", ylab="es")
sc + theme(legend.position="none")

# not working not sure of the purpose
#count =1
#for(i in MLH1_data$file.name){
#  templist= strsplit(i, split="_")[[1]]
#  if(MLH1_data$mouse %in% passed_mice) 
#  MLH1_data$mouse[count] <- c
#  count= count +1
#}
```

The code chunck below makes a table of t.test results. Each category has 2 rows. When I tried changing equal variance to false, the code breaks.
 c("strain", "sex"),

```{r}
#for the record, this is a way to make a table with pwr and t.test results
category_pwr_stat <- ddply(MLH1_data, "category", function(x) {
Nmice <- length( unique(x$mouse) )
mean.count <- mean(x$adj_nMLH1.foci)
sd.count <- sd(x$adj_nMLH1.foci)
cv <- sd.count/mean.count

ttest <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$p.value
ttestCI.lowr <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$conf.int[1]
ttestCI.upr <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$conf.int[2]
ttest.est <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$estimate

pwr5mice <- pwr.norm.test(d = (3/sd.count), n = 20, sig.level = 0.05, alternative = "two")$power
data.frame(Nmice=Nmice, cv.count = cv, ttest=ttest, CIlower = ttestCI.lowr, 
           CIupr = ttestCI.upr,
 pwr = pwr5mice, ttest.est=ttest.est)
})
category_pwr_stat
```

```{r}
plot(category_pwr_stat$category, category_pwr_stat$ttest.est)
```
I am not sure how to interpert these power calqs. The estimates are all overlapping.. I guess I shouldn't be suprised by this?

```{r, fig.height=2, fig.width=4, echo=FALSE}
#use kable to make tables. warning this might not work for github Rmd.
kable(AP_strain_table)
```

### 2017-09-21 setup_data.R

@@ -1,0 +1,210 @@
+##R script for making R.data files 
+## read in data files and adjust with additional columns
+## input: csv file of anon MLH1 counts
+## output: .rdata saved workspace with dataframe and other stat tables (save tables)
+#import MLH1 measures across data sets; AP, BD's and Cast female Koehler
+library(plyr)
+
+########################
+# Setup main dataframe #
+########################
+
+#TODO: remove observations without quality scores.
+#Write a make file that will merge all of the batch files and record the batch
+
+
+#setwd("C:/Users/alpeterson7/Documents/MLH1data/Results/figures/")
+MLH1_data = read.csv("full_anon_data_cleaned.csv")
+
+MLH1_data$category <- ifelse(grepl("_WSB_m", MLH1_data$file.name), "WSB male", 
+                      ifelse(grepl("_WSB_f", MLH1_data$file.name), "WSB female",
+                      ifelse(grepl("_G_f", MLH1_data$file.name), "G female",
+                      ifelse(grepl("_G_m", MLH1_data$file.name), "G male",
+                   ifelse(grepl("_CAST_m", MLH1_data$file.name), "CAST male",
+                    ifelse(grepl("_CAST_f", MLH1_data$file.name), "CAST female",
+                    ifelse(grepl("_MSM_m", MLH1_data$file.name), "MSM male",
+                  ifelse(grepl("_MSM_f", MLH1_data$file.name), "MSM female",
+                   ifelse(grepl("_LEW_f", MLH1_data$file.name), "LEWES female", 
+                   ifelse(grepl("_LEW_m", MLH1_data$file.name), "LEWES male",
+                   ifelse(grepl("_LEWES_m", MLH1_data$file.name), "LEWES male",                                       
+                 ifelse(grepl("_PWD_m", MLH1_data$file.name), "PWD male",     
+                  ifelse(grepl("_PWD_f", MLH1_data$file.name), "PWD female", "other")))))))))))))
+
+#set the order of categories (female, male) (cast, dom, musc)
+MLH1_data$category<- factor(MLH1_data$category,levels =c( "G female", "G male", 
+                      "WSB female", "WSB male", "LEWES female", 'LEWES male', 
[…]

### 2017-09-21 rmd_report.rmd
knitd Rmd report

@@
 
-bb <- ggplot(data = dat, aes(x=samplesize, y=MLH1_sample)) + geom_boxplot(aes(fill=samplesize))
+bb <- ggplot(data = dat, aes(x=samplesize, y=MLH1_sample)) + #geom_boxplot(aes(fill=samplesize))
 
 ```
-#omg finally that took so long. main interpertations: the variance grows with larger sample size.
-#these are representations of independant samples -- I need to think of a way to show an average of the stats for multiple simulations.
 
@@
 v <- ggplot(data = datf, aes(x=samplesize, y=MLH1_sample)) + geom_boxplot(aes(fill=samplesize) ) 
-v                                                                         
-v + ggtitle("varying sample sizes") 
+#v                                                                         
+#v + ggtitle("varying sample sizes") 
 ```
@@
 sc <- qplot(mouse, nMLH1.foci,  colour = category, data = non_passed_mice_data, main="s", xlab="s", ylab="es")
-sc + theme(legend.position="none")
+#sc + theme(legend.position="none")
 
 sc <- qplot(mouse, nMLH1.foci,  colour = category, data = passed_mice_data, main="s", xlab="s", ylab="es")
-sc + theme(legend.position="none")
+#sc + theme(legend.position="none")
 
@@
 })
-category_pwr_stat
+#category_pwr_stat
 ```
 
-```{r}
-plot(category_pwr_stat$category, category_pwr_stat$ttest.est)
-```
 I am not sure how to interpert these power calqs. The estimates are all overlapping.. I guess I shouldn't be suprised by this?

### 2017-09-22 rmd_report.rmd
edits to report, sampling simulations

@@
 
+```{r}
+dataset <- cbind (idorg,x,y)
+attempt <- list()
+for (i in 1:6)  {
+  attempt[[i]] <- dataset[sample(1:nrow(dataset), 3, replace=FALSE),]
+
+}
+attempt[6][[1]][,'idorg']# 6th mini df, the column 'idorg'
+
+# apply functions across dataframe for mice, #create a dataframe from the variables made #isolate data set for mice with enough cells (or find a way to code around the errors)
+
+df_sample_size_test <-  ddply(MLH1_data, .(mouse), function(x) {
+#for each mouse in big data set, make samples of different sizes 10 times
+              ncell <- length(unique(x$file.name))       
+              smpl10 <- mean( sample(x$adj_nMLH1.foci, ncell*0.1, replace=FALSE) )
+             # smpl10x10 <- mean( sample(x$adj_nMLH1.foci, ncell*0.1, replace=FALSE) )
+              attempt = list()
+              for (i in 1:10)  {
+                  attempt[[i]] <- dataset[sample(1:nrow(x$adj_nMLH1.foci), 10, replace=FALSE),]
+              }
+              smple_10x10 = mean(attempt[1:10][[1]][,'adj_nMLH1.foci'])
+
+      data.frame(ncell=ncell, mean10 = smpl10, s10x10 = smple_10x10)
+})
+attempt[6][[1]][,'idorg']# 6th mini df, the column 'idorg'
+
+PWD_m_example = MLH1_data[MLH1_data$mouse == '10mar15_PWD_m2', ]
+
+dataset <- cbind (idorg,x,y)
+ attempt = c()
+              for (i in 1:3)  {
+                  attempt[[i]] <- sample(PWD_m_example$nMLH1.foci, 10, replace=FALSE)
+              }
+              smple_10x10 = mean(attempt[1:10][[1]][,'nMLH1.foci'])#col isn't name ... don't need the name
+
+idorg <- c(1,2,3,4,5)
+x <- c(14,20,21,16,17)
+y <- c(31,21,20,50,13)
+
+#loop below simulates sampling 6 times, from the dataset. a mini df is made out of the samples
+#
+dataset <- cbind (idorg,x,y)
+attempt <- list()
+for (i in 1:6)  {
+  attempt[[i]] <- dataset[sample(1:nrow(dataset), 3, replace=FALSE),]
+
+}
+attempt[6][[1]][,'idorg']# 6th mini df, the column 'idorg'
+
+category_pwr_stat <- ddply(MLH1_data, "category", function(x) {
+Nmice <- length( unique(x$mouse) )
+mean.count <- mean(x$adj_nMLH1.foci)
+sd.count <- sd(x$adj_nMLH1.foci)
+cv <- sd.count/mnea.count
+
+ttest.p <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$p.value
+ttestCI.lowr <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$conf.int[1]
+ttestCI.upr <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$conf.int[2]
+ttest.est <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$estimate
+
+pwr5mice <- pwr.norm.test(d = (3/sd.count), n = 20, sig.level = 0.05, alternative = "two")$power
+
+data.frame(Nmice=Nmice, cv = cv, ttest=ttest.p, CIlower = ttestCI.lowr, 
+           CIupr = ttestCI.upr,
+ pwr = pwr5mice, ttest.est=ttest.est)
+})
+
+```
+
 Diminishing return analysis. Take some of the mice identified above, simulate random sampling of 10,15,20,30. Display the raw mean and sd, and results from 10 simulations.

### 2017-09-25 rmd_report.rmd
updated report md file

--- 2017-09-21 rmd_report.rmd
@@
 
-TODO
-remove the output for the plots
-data set up page; remove non quality observations
-Solve the cells per mouse problem
-figure out figure captions
-add a place to record the 'batches' that make up the data
-category has CAST female, this might be messing things up, since there is not CAST data
-
-include is for blocking output messages. echo false is for silencing the code commands.
 ```{r setup, echo=FALSE, include = FALSE}
@@
 
-#### goals of research
+#### Research Discription
 Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice (rodents).
@@
 
-### Goal / Discription of Report
-Goal of this report;
+#### Discription of Report
 
-> *to keep track of the number of images for each category and mouse*
+- Display plots and figures of distributions of MLH1 counts to assess intial overall patterns.
 
-This is important for getting an idea of the 'power' I have to accurately estimate MLH1 distributions across the categories of cells I am quantifying.
+-  *Report basic statistics of results and to help keep track of the number of images for each category and mouse*
 
-#### Discription of data and quantification process
-data frames are loaded of MLH1 data 
-imagename, mouse name, strain, sex, nMLH1.
+#### Discription of quantification process
 
-quanti process. anonymized, given quality score
+Aquired microscope cell images are quantified in batches after being anonimzed.
+The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified.
+A quality score is given; 1 to 5 (best to worst).
 
@@
 #Add fake CAST female data
-
 last_row <-data.frame(file.name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )
@@
 
+#### Mice table
+
+Table of the number of mice used and MLH1 stats. (made with kable).
+
+```{r, fig.height=2, fig.width=4, echo=FALSE}
+#this table is saved in the .RData file
+#use kable to make tables. warning this might not work for github Rmd.
+kable(AP_strain_table)
+```
+
 #### Initial Patterns from MLH1 distributions
 
-Use these figures to identify outliers and if there is variance in the means and variance.
+```{r first boxplots, echo=FALSE, fig.align='center'}
 
-```{r first boxplots, echo=FALSE, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, fig.align='center'}
+ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci)) + geom_boxplot(data = MLH1_data, aes(fill = factor(strain) ) ) + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen"))
+ff <- ff + facet_wrap(~ sex, scales="free")
+ff
+```
 
-ff <- ggplot(MLH1_data, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = MLH1_data, aes(fill = factor(strain) ) )+ theme(axis.title.x=element_blank(),
+After taking the data from 2 highest cell quality, some but not all mouse means converge.
+
+```{r, histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
+
+histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
+histM <- histM + geom_histogram(stat = "bin", binwidth = 1)
+histM <- histM + facet_wrap(~ category)
+histM
+```
+
+```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE}
+
+#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
+
+highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
+
+qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
+  theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
@@
   "coral1", "#E69F00", "yellowgreen"))
-ff <- ff + facet_wrap(~ sex)
-
+high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
 ```
 
-### Tests for quality and nMLH1 foci number
+```{r, passing mice, echo=FALSE}
+q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
+                        total =  length(adj_nMLH1.foci),
+                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
+                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
+                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
+)
+#q_cutoff_table
+passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
+
+not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
+
+#mnake dataframe with just passing mice -- so that it can be ploted
+passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
+```
+
+#### Comparison of MLH1 distributions
+
+```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
+ff
+high_qual_boxplot
+```
+
+After taking the data from 2 highest cell quality, some but not all mouse means converge.
+For some mice (mostly female) the majority of cells are excluded.
+This is especially true for PWD and WSB females.
+
+#### Tests for quality and nMLH1 foci number
 
@@
 
-Boxplots of mice distributions with quality score of 1
-
-```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE}
-
-#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
-
-highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
-
-qq <- ggplot(highest_qual_df, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
-  "coral1", "#E69F00", "yellowgreen"))
-high_qual_boxplot <- qq + facet_wrap(~ sex) #this is breaking the kniting...
-
-```
-
-The spacing in these boxplots is off because of 'ghost' cells, which are in the data frame, but not in the subsetted data. (Iguess)
-
-Table for assessing the number of passed mice
-```{r, passing mice, echo=FALSE}
-q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
-                        total =  length(nMLH1.foci),
-                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
-                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
-                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
-)
-#q_cutoff_table
-
-mice_25cells <- q_cutoff_table[q_cutoff_table$above5 > 25,]$mouse
-#12 mice have more than 25 cells
-mice_20cells <- q_cutoff_table[q_cutoff_table$above5 > 20,]$mouse
-#26 mice have over 20 cells
-mice_15cells <- q_cutoff_table[q_cutoff_table$above5 > 14,]$mouse
-
-passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
-
-not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
-
-#mnake dataframe with just passing mice -- so that it can be ploted
-passed_mice_df <- MLH1_data[ ( MLH1_data$quality < 5), ]
-
-```
-
-#### Boxplots of 'passing' mice distributions
-
-```{r, echo=FALSE}
-pp <- ggplot(passed_mice_df, aes(x = mouse, y = nMLH1.foci)) + geom_boxplot(data = passed_mice_df, aes(fill = factor(strain) ) )+ theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
-  "coral1", "#E69F00", "yellowgreen"))
-passed_mice_plot <- pp + facet_wrap(~ sex)
-
-```
-```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
-
-ff
-high_qual_boxplot
-passed_mice_plot
-```
-
 ##### Effects of quality
 
-Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality. Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
+Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
+Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
+(CAST female data is fake)
 
@@
 
-Potential solutions?
+#### Assessing the Distributions by mouse
 
-> *do some mice have better quality cells?* *Does this cause bias in the data?*
-
-```{r, fig.height=4, fig.width=4, echo=FALSE, fig.caption= "histogrms", warning=FALSE}
-# do the same thing as above but geom_hist
-
-histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
-histM <- histM + geom_histogram(stat = "bin", binwidth = 1)
-histM <- histM + facet_wrap(~ category)
-histM
-```
-
-It seems like there is a population of lower cells in MSM males. This could be from slides which had bad staining.
-
-#### Assessing the number of cells per mouse
-
-latice plot of scatter plots for jitter plots of cell oberservations by quality
+latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.
 ```{r, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE }
+```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE }
 #geom_hline(aes(yintercept = med, group = gr), colour = 'red')
@@
   show(plot_list[gg])
+  #invisible(lapply(obj,function(x) plot(x,main="some plot")))
 }
@@
 
-Making all of these scatter plots, allows us to look at the whole distributions.
+Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
+The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
+
+PWD females and a very large range. This is driven by data from 17apr15_PWD_f4. These data should be looked at more closely.
+
 In some mice there seems to be bimodel cell populations, those of clearly high quality and those of lower quality.
-This may be due to the screeign of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.
+This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.
 
@@
 
-Think about a loop that will -- i) take ~100 different sized samples of total images (10%, 25%, 50%, 75%, 90%) (with varying degrees of quality)
+Outline discription of the loop for comparing sample size effects
+
+i) take ~100 independant samples of total images across different sample sizes (10%, 25%, 50%, 75%, 90%)
 ii) then run t.tests across those samples, and report the p value 
 iii) make a table of the p values for these sample permutations
+iv) plot the p values or mean p values across sample sizes
 
-Diminishing return analysis. Take some of the mice identified above, simulate random sampling of 10,15,20,30. Display the raw mean and sd, and results from 10 simulations.
-```{r, comparing sample sizes, echo=FALSE}
-PWD_m_example = MLH1_data[MLH1_data$mouse == '10mar15_PWD_m2', ]
+Currently this section is under construction.
 
-smp10dist = sample(PWD_m_example$nMLH1.foci, 10, replace=FALSE)
-samplesize = rep("10", 10)
-smp10 <- data.frame(smp10dist, samplesize)
-
-smp15dist = sample(PWD_m_example$nMLH1.foci, 15, replace=FALSE)
-samplesize = rep("15", 15)
-smp15 <- data.frame(smp15dist, samplesize)
-
-smp20dist = sample(PWD_m_example$nMLH1.foci, 20, replace=FALSE)
-samplesize = rep("20", 20)
-smp20 <- data.frame(smp20dist, samplesize)
-
-smp30dist = sample(PWD_m_example$nMLH1.foci, 30, replace=FALSE)
-samplesize = rep("30", 30)
-smp10 <- data.frame(smp30dist, samplesize)
-
-#this mouse has 40 cells 
-#make a dataframe (so that I can finally make boxplots)
-dat <- data.frame(
-  MLH1_sample = c(sample(PWD_m_example$adj_nMLH1.foci, 10, replace=FALSE),  sample(PWD_m_example$adj_nMLH1.foci, 15, replace=FALSE),    sample(PWD_m_example$adj_nMLH1.foci, 20, replace=FALSE), sample(PWD_m_example$adj_nMLH1.foci, 30, replace=FALSE), sample(PWD_m_example$adj_nMLH1.foci, 35, replace=FALSE), sample(PWD_m_example$adj_nMLH1.foci, 40, replace=FALSE)),
-  samplesize = c(rep("10", 10), rep("15", 15), rep("20", 20), rep("30", 30), rep("35", 35), rep("40", 40)) )
-
-bb <- ggplot(data = dat, aes(x=samplesize, y=MLH1_sample)) + #geom_boxplot(aes(fill=samplesize))
-
-```
-
-```{r}
-G_f_example = MLH1_data[MLH1_data$mouse == '17mar16_G_f1', ]
-
-datf <- data.frame(
-  MLH1_sample = c(sample(G_f_example$nMLH1.foci, 10, replace=FALSE),  sample(G_f_example$nMLH1.foci, 15, replace=FALSE),    sample(G_f_example$nMLH1.foci, 20, replace=FALSE), sample(G_f_example$nMLH1.foci, 30, replace=FALSE), sample(G_f_example$nMLH1.foci, 35, replace=FALSE) ),
-  samplesize = c(rep("10", 10), rep("15", 15), rep("20", 20), rep("30", 30), rep("35", 35)) )
-
-v <- ggplot(data = datf, aes(x=samplesize, y=MLH1_sample)) + geom_boxplot(aes(fill=samplesize) ) 
-#v                                                                         
-#v + ggtitle("varying sample sizes") 
-```
-Can the above code be repeated for a larger number of mice (both passing and non-passing) to show the different effects across samples sizes?
-power calculations....
-from a preivous document --- I showed myself that for 15 (good) cells, would provide a sample big enough to detect a difference of 3 foci.
-I only did this for a couple mice, I'll try to re-code for every mouse
-
-```{r}
+```{r, include=FALSE}
 passed_mice_data <- MLH1_data[MLH1_data$mouse %in% passed_mice, ]
@@
 ```
-
-The code chunck below makes a table of t.test results. Each category has 2 rows. When I tried changing equal variance to false, the code breaks.
- c("strain", "sex"),
-
-```{r}
-#for the record, this is a way to make a table with pwr and t.test results
-category_pwr_stat <- ddply(MLH1_data, "category", function(x) {
-Nmice <- length( unique(x$mouse) )
-mean.count <- mean(x$adj_nMLH1.foci)
-sd.count <- sd(x$adj_nMLH1.foci)
-cv <- sd.count/mean.count
-
-ttest <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$p.value
-ttestCI.lowr <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$conf.int[1]
-ttestCI.upr <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$conf.int[2]
-ttest.est <- t.test(x$adj_nMLH1.foci, (20), alternative = "two", var.equal=TRUE)$estimate
-
-pwr5mice <- pwr.norm.test(d = (3/sd.count), n = 20, sig.level = 0.05, alternative = "two")$power
-data.frame(Nmice=Nmice, cv.count = cv, ttest=ttest, CIlower = ttestCI.lowr, 
-           CIupr = ttestCI.upr,
- pwr = pwr5mice, ttest.est=ttest.est)
-})
-#category_pwr_stat
-```
-
-I am not sure how to interpert these power calqs. The estimates are all overlapping.. I guess I shouldn't be suprised by this?
-
-```{r, fig.height=2, fig.width=4, echo=FALSE}
-#use kable to make tables. warning this might not work for github Rmd.
-kable(AP_strain_table)
-```

### 2017-09-25 junkcode.R

@@ -1,0 +1,168 @@
+```{r, echo=FALSE}
+pp <- ggplot(passed_mice_df, aes(x = as.factor(mouse), y = adj_nMLH1.foci)) + geom_boxplot(data = passed_mice_df, aes(fill = factor(strain) ) )+ ggtitle("Passing mice") + theme(axis.title.x=element_blank(),
+                                                                                                                                                                                 axis.text.x=element_blank(),
+                                                                                                                                                                                 axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
+                                                                                                                                                                                                                                            "coral1", "#E69F00", "yellowgreen"))
+passed_mice_plot <- pp + facet_wrap(~ sex, scales="free")
+
+```
+The passing mice plots, are not very different --- it might not be 
+
+
+```{r, include=FALSE}
+
+idorg <- c(1,2,3,4,5)
+x <- c(14,20,21,16,17)
+y <- c(31,21,20,50,13)
+
+#dataset <- cbind (idorg,x,y)
+#attempt <- list()
+#for (i in 1:6)  {
+#  attempt[[i]] <- dataset[sample(1:nrow(dataset), 3, replace=FALSE),]#
+#
+#}
+#attempt[6][[1]][,'idorg']# 6th mini df, the column 'idorg'
+
+# apply functions across dataframe for mice, #create a dataframe from the variables made #isolate data set for mice with enough cells (or find a way to code around the errors)
+
+df_sample_size_test <-  ddply(MLH1_data, .(mouse), function(x) {
+  #for each mouse in big data set, make samples of different sizes 10 times
+  ncell <- length(unique(x$file.name))
+  smpl10 <- ifelse(ncell < 5, 'NA',  mean( sample(x$adj_nMLH1.foci, ncell*0.1, replace=FALSE) ) )
[…]

### 2017-10-12 rmd_report.rmd
playing with permutation code

@@
 MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
-
 ```
@@
 Currently this section is under construction.
+```{r}
+# use q_cut_off table, to find the mice to use
+listOmice <- q_cutoff_table$mouse[ (q_cutoff_table$total >= 25) ]
+#mke new df with from 
+df.pasd.mice <- MLH1_data[MLH1_data$mouse %in% listOmice, ]
+
+#write function --- for taking 100 samples of 10
+
+permut.df <- function(df){
+  #semperate by mouse
+    mouse.data    <- ddply(df, c("mouse"), summarise,
+       df <- data.frame(
+              samp = sample(nMLH1.foci,5)
+            ) 
+        )
+
+  return(mouse.data)
+}
+
+mouse.data <- ddply(df.pasd.mice, c("mouse"), summarise,
+
+  m.df <- data.frame(sample(nMLH1.foci, 5, replace = FALSE) )
+  )
+
+# this creates dataframe, which is a subset of larger dataframe ... not sure se_alone, but samples 40 of 
+this =ddply (df.pasd.mice, .(mouse), subset, sample(seq_along(nMLH1.foci)<=5))
+
+ss <- sample(df.pasd.mice$nMLH1.foci)
+
+Table_BD_mouse <- ddply(BDMLH1_data, c("ANIMAL_ID", "Cross"), summarise,
+                       N  = length(nMLH1_foci),
+                       mean_co = format(round( mean(nMLH1_foci), 3), nsmall = 3),
+                       var = format(round( var(nMLH1_foci),3), nsmall = 3),
+                       sd   = round(sd(nMLH1_foci),3),
+                       se   = round(sd / sqrt(N),3)
+                       #text=paste(Cross, collapse=""))
+)
+
+```
 

### 2017-10-12 rmd_report.rmd
working on simulation things

--- 2017-09-25 rmd_report.rmd
@@
 ```{r first boxplots, echo=FALSE, fig.align='center'}
-
 ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci)) + geom_boxplot(data = MLH1_data, aes(fill = factor(strain) ) ) + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),

### 2017-10-12 rmd_report.rmd
learning how to permute samples

@@
 ```{r}
+
+colnms = c("mouse", "rep")
+
+#makes dfs, with the samples 
+permut.mouse.numb <- function(x, rep_size, samp_size){
+  new.df = matrix(nrow=rep_size, ncol=samp_size)
+for(i in 1:rep_size) {
+  new.df[i,] = sample(x, samp_size)
+  }
+return(as.data.frame(new.df, colnames(colnms) ) )
+}
+
+ply_data <- df.pasd.mice[df.pasd.mice$mouse == "8oct14_PWD_f8",]
+
+#make this the function that I apply over the mice in the data
+t.test(sample(ply_data$nMLH1.foci, 5), sample(ply_data$nMLH1.foci, 5))$p.value
+t.test(sample(ply_data$nMLH1.foci, 13), sample(ply_data$nMLH1.foci, 13))$p.value
+
+#this works
+vv <-permut.mouse.numb(df.pasd.mice$nMLH1.foci,5,2)
+
+##
+## I think this does it! this makes a list of 23 dfs (for each mouse in the large df)
+kk<-dlply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,5) )
+
+jj<-ddply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,10) )
+
+#t.test and p values
+#maybe I should just choose 2 samples, do a t.test, do that 100 times..?
+#below runs t.tests on 
+bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2 samples of size 10
+
+#figure out a way to plot/visualize these pvalues
+#2 samples are good for t.tests, but I need to repeat 100 times
+#I think I need to ... add the replicates that can be xxply through
+
+## for 1:100, cal permt(10,2) then call 
+ll=c()
+ty <- for(i in 1:10){
+  print(i)
+  jj <-ddply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,10) )
+
+  bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )
+  #the t.test$p.values are not exactly what they should be
+#put p values into a list? (100 pvalues )
+  ll[i] = bb$V1[i]#this returns a list of p values for different mice? one mouse?
+}
+
+#for
+
+gg<-t.test(kk$`10mar15_PWD_m2`[1,2:5], kk$`10mar15_PWD_m2`$V3)  
+#jj$`10mar15_PWD_m2`
+#jj[1,1:5]
+#$`10mar15_PWD_m2`
+#[1] 0.6071011
+
+##then do t.test to test the distributions 
+## data is the data frames, by mouse, function is t.test
+
 # use q_cut_off table, to find the mice to use
@@
 #write function --- for taking 100 samples of 10
-
 permut.df <- function(df){

### 2017-10-13 rmd_report.rmd
updated from work

@@
 ply_data <- df.pasd.mice[df.pasd.mice$mouse == "8oct14_PWD_f8",]
+#make this the function that I apply over the mice in the data
 
-#make this the function that I apply over the mice in the data
+#simplify the t.test
 t.test(sample(ply_data$nMLH1.foci, 5), sample(ply_data$nMLH1.foci, 5))$p.value
 t.test(sample(ply_data$nMLH1.foci, 13), sample(ply_data$nMLH1.foci, 13))$p.value
+
+#repeat the sampling 100 times
+
+bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2
+
+#try with do
+blah <- df.pasd.mice %>% group_by(mouse) %>% do(head(.))
+
+plist=data.frame(pv=as.numeric(), mean1=as.numeric(), mean2=as.numeric())
+
+hundred <- function(x){
+  plist=data.frame(pv=as.numeric(), mean1=as.numeric(), mean2=as.numeric())
+  for(eye in 1:100){
+    samp1 = sample(x, 5)
+    samp2 = sample(x, 5)
+    plist$pv[eye] <- t.test(samp1,samp2)$p.value
+    plist$mean1[eye] <- mean(samp1)
+    plist$mean2[eye] <- mean(samp2)
+  }
+  return(plist)
+}
+
+mm <- hundred(ply_data$nMLH1.foci)
 

### 2017-10-22 rmd_report.rmd
still working on pvalue comparison

@@
 load(file="MLH1_data_setup.RData")
-
 ```
@@
 colnms = c("mouse", "rep")
-
-#makes dfs, with the samples 
-permut.mouse.numb <- function(x, rep_size, samp_size){
-  new.df = matrix(nrow=rep_size, ncol=samp_size)
-for(i in 1:rep_size) {
-  new.df[i,] = sample(x, samp_size)
-  }
-return(as.data.frame(new.df, colnames(colnms) ) )
-}
-
 ply_data <- df.pasd.mice[df.pasd.mice$mouse == "8oct14_PWD_f8",]
@@
 #repeat the sampling 100 times
-
 bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2
@@
 } 
-
 mm <- hundred(ply_data$nMLH1.foci)
 
-#this works
-vv <-permut.mouse.numb(df.pasd.mice$nMLH1.foci,5,2)
+#makes dfs, with the samples. // sampling from MLH1
+permut.mouse.numb <- function(x, rep_size, samp_size){
+  new.df = matrix(nrow=rep_size, ncol=samp_size)
+for(i in 1:rep_size) {
+  new.df[i,] = sample(x, samp_size)
+  }
+return(as.data.frame(new.df, colnames(colnms) ) )
+  #return sampled
+}
+#can I integrate the sampling and t.test here?
 
-##
+#function for
+
+#this works. sampling x permutations, y sample size
+vv <- permut.mouse.numb(df.pasd.mice$nMLH1.foci,5,4)
+
 ## I think this does it! this makes a list of 23 dfs (for each mouse in the large df)
+##this makes dfs for doing the t.tests
 kk<-dlply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,5) )
+#applies permut function over mouse -- then run a t.test and return pvalue
+jj<-ddply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,20) )
 
-jj<-ddply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,10) )
+#
+pval <- function()
 
@@
 #below runs t.tests on
-bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2 samples of size 10
+#this is making a new tab;e, 
+bb <- ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2 samples of size 20
+
+# figure out a way to mush dd into a matrix
+
+repeats = 10
+new.p.matrix = matrix(ncol=length(unique(jj$mouse)), nrow = repeats)
+for(e in 1:repeats){
+  new.p.matrix[e]<- ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )
+
+  }
+
+powrange <- seq(0.4, .9, .1)
+n <- sapply(powrange, function(i) power.prop.test(p1=0.4,p2=0.24,power=i,sig.level=0.05)$n)
+plot(powrange, n)
+
+bb$mouse[(bb$V1 < 0.05)]
+#mice which had significantly difference in samples of size 20
+#16jan16_G_f2  20feb16_G_f3  3nov15_WSB_f1
+#17mar16_G_f1
+#16jan16_G_f2   20dec16_LEW_m3
+#30sep16_MSM_f2
+#22jun15_G_m2
+
+#figure out how to permute these tests. repeat the above... then return distributions of p values
+
+#10
+#20feb16_G_f3   13nov16_MSM_m1
 
@@
 this =ddply (df.pasd.mice, .(mouse), subset, sample(seq_along(nMLH1.foci)<=5))
-
-ss <- sample(df.pasd.mice$nMLH1.foci)
+#ss <- sample(df.pasd.mice$nMLH1.foci)
 

### 2017-10-22 test4_pvalue.R

@@ -1,0 +1,31 @@
+
+
+library(knitr)
+library(ggplot2)
+library(pwr)
+library(plyr)
+library(lattice)
+library(dplyr)
+
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="MLH1_data_setup.RData")
+
+
+q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
+                        total =  length(adj_nMLH1.foci),
+                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
+                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
+                        # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
+)
+#q_cutoff_table
+passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
+
+not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
+
+#mnake dataframe with just passing mice -- so that it can be ploted
+passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
+
+
+#make a dataframe of many p values
+
+

### 2017-10-27 rmd_report.rmd
changes to report file

@@
 ```{r}
-
 colnms = c("mouse", "rep")

### 2017-10-27 rmd_report.rmd
refining make file

@@
 library(dplyr)
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
 load(file="MLH1_data_setup.RData")

### 2017-10-27 setup_data.R

@@ -52,11 +52,9 @@ change <- MLH1_data[MLH1_data$quality == "",]
 MLH1_data$category <- as.factor(MLH1_data$category)
 
 MLH1_data$nMLH1.foci <- as.numeric(MLH1_data$nMLH1.foci) #make these numeric just in case there are other characters
-
-MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
-
 #add a column with male adjusted MLH1 values (+1 to all males)
 MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
+MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
 
 count =1
 for(i in MLH1_data$Original.Name){

### 2017-11-02 rmd_report.rmd
cleaned up setup file.

@@
 
-```{r, echo=FALSE, include = FALSE}
-##FIX ORDER HERE (this should be put in the data setup script)
-MLH1_data$strain<- factor(MLH1_data$strain,levels =c("G", "LEWES",
-                              "WSB", "PWD","MSM","CAST"), order=T )
+```{r}
 #order the data frame
@@
   arrange(strain, sex, mouse) %>%
-  mutate(file.name = factor(file.name)) #another category that you want the order to match
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
 # sort your dataframe, by the focal categories
@@
 #Add fake CAST female data
-last_row <-data.frame(file.name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )
+last_row <-data.frame(Batch = c("100", "100"), X= c("",""), Original.Name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )
 
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen"))
+        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "black", "grey", "lightgrey"))
 ff <- ff + facet_wrap(~ sex, scales="free")
@@
         axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
-  "coral1", "#E69F00", "yellowgreen"))
+  "coral1", "#E69F00", "yellowgreen","black", "grey"  ))
 high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
@@
   #below should work -- I think I need to go througheach row of the df_list[[i]]
-  for(o in 1:length(df_list[[i]]$file.name)){
+  for(o in 1:length(df_list[[i]]$Original.Name)){
     df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
@@
 #geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
+#for (gg in 1:length(plot_list) ){
+#  show(plot_list[gg])
   #invisible(lapply(obj,function(x) plot(x,main="some plot")))
-}
+#}
 ```
@@
 Currently this section is under construction.
-```{r}
-colnms = c("mouse", "rep")
-ply_data <- df.pasd.mice[df.pasd.mice$mouse == "8oct14_PWD_f8",]
-#make this the function that I apply over the mice in the data
-
-#simplify the t.test
-t.test(sample(ply_data$nMLH1.foci, 5), sample(ply_data$nMLH1.foci, 5))$p.value
-t.test(sample(ply_data$nMLH1.foci, 13), sample(ply_data$nMLH1.foci, 13))$p.value
-
-#repeat the sampling 100 times
-bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2
-
-#try with do
-blah <- df.pasd.mice %>% group_by(mouse) %>% do(head(.))
-
-plist=data.frame(pv=as.numeric(), mean1=as.numeric(), mean2=as.numeric())
-
-hundred <- function(x){
-  plist=data.frame(pv=as.numeric(), mean1=as.numeric(), mean2=as.numeric())
-  for(eye in 1:100){
-    samp1 = sample(x, 5)
-    samp2 = sample(x, 5)
-    plist$pv[eye] <- t.test(samp1,samp2)$p.value
-    plist$mean1[eye] <- mean(samp1)
-    plist$mean2[eye] <- mean(samp2)
-  }
-  return(plist)
-} 
-mm <- hundred(ply_data$nMLH1.foci)
-
-#makes dfs, with the samples. // sampling from MLH1
-permut.mouse.numb <- function(x, rep_size, samp_size){
-  new.df = matrix(nrow=rep_size, ncol=samp_size)
-for(i in 1:rep_size) {
-  new.df[i,] = sample(x, samp_size)
-  }
-return(as.data.frame(new.df, colnames(colnms) ) )
-  #return sampled
-}
-#can I integrate the sampling and t.test here?
-
-#function for
-
-#this works. sampling x permutations, y sample size
-vv <- permut.mouse.numb(df.pasd.mice$nMLH1.foci,5,4)
-
-## I think this does it! this makes a list of 23 dfs (for each mouse in the large df)
-##this makes dfs for doing the t.tests
-kk<-dlply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,5) )
-#applies permut function over mouse -- then run a t.test and return pvalue
-jj<-ddply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,20) )
-
-#
-pval <- function()
-
-#t.test and p values
-#maybe I should just choose 2 samples, do a t.test, do that 100 times..?
-#below runs t.tests on
-#this is making a new tab;e, 
-bb <- ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2 samples of size 20
-
-# figure out a way to mush dd into a matrix
-
-repeats = 10
-new.p.matrix = matrix(ncol=length(unique(jj$mouse)), nrow = repeats)
-for(e in 1:repeats){
-  new.p.matrix[e]<- ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )
-
-  }
-
-powrange <- seq(0.4, .9, .1)
-n <- sapply(powrange, function(i) power.prop.test(p1=0.4,p2=0.24,power=i,sig.level=0.05)$n)
-plot(powrange, n)
-
-bb$mouse[(bb$V1 < 0.05)]
-#mice which had significantly difference in samples of size 20
-#16jan16_G_f2  20feb16_G_f3  3nov15_WSB_f1
-#17mar16_G_f1
-#16jan16_G_f2   20dec16_LEW_m3
-#30sep16_MSM_f2
-#22jun15_G_m2
-
-#figure out how to permute these tests. repeat the above... then return distributions of p values
-
-#10
-#20feb16_G_f3   13nov16_MSM_m1
-
-#figure out a way to plot/visualize these pvalues
-#2 samples are good for t.tests, but I need to repeat 100 times
-#I think I need to ... add the replicates that can be xxply through
-
-## for 1:100, cal permt(10,2) then call 
-ll=c()
-ty <- for(i in 1:10){
-  print(i)
-  jj <-ddply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,10) )
-
-  bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )
-  #the t.test$p.values are not exactly what they should be
-#put p values into a list? (100 pvalues )
-  ll[i] = bb$V1[i]#this returns a list of p values for different mice? one mouse?
-}
-
-#for
-
-gg<-t.test(kk$`10mar15_PWD_m2`[1,2:5], kk$`10mar15_PWD_m2`$V3)  
-#jj$`10mar15_PWD_m2`
-#jj[1,1:5]
-#$`10mar15_PWD_m2`
-#[1] 0.6071011
-
-##then do t.test to test the distributions 
-## data is the data frames, by mouse, function is t.test
-
-# use q_cut_off table, to find the mice to use
-listOmice <- q_cutoff_table$mouse[ (q_cutoff_table$total >= 25) ]
-#mke new df with from 
-df.pasd.mice <- MLH1_data[MLH1_data$mouse %in% listOmice, ]
-
-#write function --- for taking 100 samples of 10
-permut.df <- function(df){
-  #semperate by mouse
-    mouse.data    <- ddply(df, c("mouse"), summarise,
-       df <- data.frame(
-              samp = sample(nMLH1.foci,5)
-            ) 
-        )
-
-  return(mouse.data)
-}
-
-mouse.data <- ddply(df.pasd.mice, c("mouse"), summarise,
-
-  m.df <- data.frame(sample(nMLH1.foci, 5, replace = FALSE) )
-  )
-
-# this creates dataframe, which is a subset of larger dataframe ... not sure se_alone, but samples 40 of 
-this =ddply (df.pasd.mice, .(mouse), subset, sample(seq_along(nMLH1.foci)<=5))
-#ss <- sample(df.pasd.mice$nMLH1.foci)
-
-Table_BD_mouse <- ddply(BDMLH1_data, c("ANIMAL_ID", "Cross"), summarise,
-                       N  = length(nMLH1_foci),
-                       mean_co = format(round( mean(nMLH1_foci), 3), nsmall = 3),
-                       var = format(round( var(nMLH1_foci),3), nsmall = 3),
-                       sd   = round(sd(nMLH1_foci),3),
-                       se   = round(sd / sqrt(N),3)
-                       #text=paste(Cross, collapse=""))
-)
-
-```
 

### 2017-11-02 Func_addCategory.R

@@ -44,7 +44,3 @@ add_category <- function(oldframe){
   return(dframe)
 }
 #tried writing this with mutate, couldn't get it to work
-
-
-
-

### 2017-11-02 meta_data_analysis.R

@@ -10,29 +10,21 @@ library(plyr)
 library(lattice)
 library(dplyr)
 
-
-#age, mat and pat age, Gough Line, staining batch
-
-#count the number of mice with missing meta-data
-
+#age, mat and pat age, Gough Line, (staining batch for cells)
 
 #load meta data csv file 
 #finalize the coding (NAs, ect)
-meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1data/data/ALP_MouseMetadata2.csv")
+meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1data/data/mouseDOBs.csv")
+
 save.image("play_data_for_meta.RData")
 #data needs to be cleaned up (all NA's removed before calculating )
-
 
 #mouse column already included,
 #change category, sex -- there is no file name...
 #using the functions will be a little tricky
 
-
-
 meta_data$mouse <- as.character(meta_data$mouse)
 #format euth date (from mouse)
-ply_set <- meta_data[5:9,]
-
 #all NA's need to be removed
 
 for( t in 1:length(meta_data$mouse)){
@@ -42,14 +34,16 @@ for( t in 1:length(meta_data$mouse)){
   fomt_euth_nrm <- as.Date(fomt_euth, "%Y-%m-%d")
   vv <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="weeks" )#whoo, this works
   hh <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="hours" )
-  meta_data$diffy_weeks[t] <- vv
-  meta_data$diffy_hours[t] <- hh
-  meta_data$calq_age[t] <- fomt_euth
+  meta_data$age_weeks[t] <- vv
+  meta_data$age_hours[t] <- hh
 
+  #mat age, DOB -Mat_dob
+  mat_dob <- as.Date(meta_data$maternal.age..DOB.[t], format= '%m/%d/%Y')
+#  hh <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="hours" )
+  gg <- as.numeric(difftime(as.Date(meta_data$DOB[t], '%m/%d/%Y'),mat_dob), units="weeks" )
+  meta_data$mat_age_wk[t] <- gg
 }
 
-mouse[1]
[…]

### 2017-11-02 pvalue_code_chunk.R

@@ -1,0 +1,159 @@
+
+#pieces from Rmd Report
+
+```{r}
+colnms = c("mouse", "rep")
+ply_data <- df.pasd.mice[df.pasd.mice$mouse == "8oct14_PWD_f8",]
+#make this the function that I apply over the mice in the data
+
+#simplify the t.test
+t.test(sample(ply_data$nMLH1.foci, 5), sample(ply_data$nMLH1.foci, 5))$p.value
+t.test(sample(ply_data$nMLH1.foci, 13), sample(ply_data$nMLH1.foci, 13))$p.value
+
+#repeat the sampling 100 times
+bb <-ddply(jj, .(mouse), function(x) t.test(x[1,2:5],x[2,2:5])$p.value )#t.test for 2 
+
+#try with do
+blah <- df.pasd.mice %>% group_by(mouse) %>% do(head(.))
+
+plist=data.frame(pv=as.numeric(), mean1=as.numeric(), mean2=as.numeric())
+
+hundred <- function(x){
+  plist=data.frame(pv=as.numeric(), mean1=as.numeric(), mean2=as.numeric())
+  for(eye in 1:100){
+    samp1 = sample(x, 5)
+    samp2 = sample(x, 5)
+    plist$pv[eye] <- t.test(samp1,samp2)$p.value
+    plist$mean1[eye] <- mean(samp1)
+    plist$mean2[eye] <- mean(samp2)
+  }
+  return(plist)
+} 
+mm <- hundred(ply_data$nMLH1.foci)
+
+
+#makes dfs, with the samples. // sampling from MLH1
+permut.mouse.numb <- function(x, rep_size, samp_size){
+  new.df = matrix(nrow=rep_size, ncol=samp_size)
+  for(i in 1:rep_size) {
+    new.df[i,] = sample(x, samp_size)
+  }
+  return(as.data.frame(new.df, colnames(colnms) ) )
+  #return sampled
+}
+#can I integrate the sampling and t.test here?
+
+#function for 
+
+#this works. sampling x permutations, y sample size
+vv <- permut.mouse.numb(df.pasd.mice$nMLH1.foci,5,4)
+
+## I think this does it! this makes a list of 23 dfs (for each mouse in the large df)
+##this makes dfs for doing the t.tests
+kk<-dlply(df.pasd.mice, .(mouse), function(x) permut.mouse.numb(x$nMLH1.foci,2,5) )
+#applies permut function over mouse -- then run a t.test and return pvalue
[…]

### 2017-11-02 setup_data.R

@@ -2,7 +2,10 @@
 ## read in data files and adjust with additional columns
 ## input: csv file of anon MLH1 counts
 ## output: .rdata saved workspace with dataframe and other stat tables (save tables)
+## df of MLH1 should go from 12cols to 17cols
 #import MLH1 measures across data sets; AP, BD's and Cast female Koehler
+# 1. MLH1_data, tables: AP strains
+library(dplyr)
 library(plyr)
 
 ########################
@@ -12,7 +15,28 @@ library(plyr)
 #TODO: remove observations without quality scores.
 #Write a make file that will merge all of the batch files and record the batch
 setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-MLH1_data = read.csv("data/AnonData.csv", header=TRUE)
+MLH1_data = read.csv("data/AnonData.csv", header=TRUE )
+
+original_DF = MLH1_data
+original_length <- length(MLH1_data$Original.Name)
+
+MLH1_data$n <- as.character(MLH1_data$n)
+MLH1_data$nMLH1.foci<- as.character(MLH1_data$nMLH1.foci)
+MLH1_data$quality<- as.character(MLH1_data$quality)
+
+MLH1_data$n <- as.numeric(MLH1_data$n)
+MLH1_data$nMLH1.foci<- as.numeric(MLH1_data$nMLH1.foci)
+MLH1_data$quality<- as.numeric(MLH1_data$quality)
+
+#make list of rows that need more data
+missing_data_DF <- MLH1_data[(MLH1_data$nMLH1.foci == "" | MLH1_data$quality == ""),]
+
+#remove X's and NAs
+MLH1_data <- MLH1_data[ !grepl("X", MLH1_data$X) , ]
+MLH1_data <- MLH1_data[!(is.na(MLH1_data$nMLH1.foci) | MLH1_data$nMLH1.foci==""), ]
+
+MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "X",]
+MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]
 
 source("Func_addCategory.R")
 MLH1_data <- add_category(MLH1_data)
@@ -20,7 +44,7 @@ MLH1_data <- add_category(MLH1_data)
 MLH1_data$category<- factor(MLH1_data$category,levels =c( "G female", "G male", 
                       "WSB female", "WSB male", "LEWES female", 'LEWES male', 
                       "PWD female", "PWD male", "MSM female", "MSM male",
[…]

### 2017-11-02 tables.R

@@ -1,0 +1,93 @@
+# tables for MLH1 data
+# input:
+# output: dplyr tables (strain/category) and mouse
+#
+
+Table_BD_mouse <- ddply(BDMLH1_data, c("ANIMAL_ID", "Cross"), summarise,
+                        N  = length(nMLH1_foci),
+                        mean_co = format(round( mean(nMLH1_foci), 3), nsmall = 3),
+                        var = format(round( var(nMLH1_foci),3), nsmall = 3),
+                        sd   = round(sd(nMLH1_foci),3),
+                        se   = round(sd / sqrt(N),3)
+                        #text=paste(Cross, collapse=""))
+)
+
+#strain averages
+Table_BD_strain <- ddply(BDMLH1_data, c("Cross"), summarise,
+                         Nmice = length(unique(ANIMAL_ID)),
+                         Ncells  = length(nMLH1_foci),
+                         mean_co = format(round (mean(nMLH1_foci), 3), nsmall=3),
+                         var = format(round( var(nMLH1_foci), 3), nsmall=3),
+                         sd   = round( sd(nMLH1_foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3)
+                         #text=paste(Cross, collapse=""))
+)#format(round(   var(nMLH1.foci),3), nsmall=3),
+
+
+dataset <- rep("BD", length(Table_BD_strain$Cross) )
+
+#everything should be male unless, 'female' listed in name
+sex <- ifelse(grepl("Female", Table_BD_strain$Cross), "female", "male") 
+
+Table_BD_strain <- cbind(Table_BD_strain, sex, dataset)# subsp,
+#change name
+Table_BD_strain$Cross[Table_BD_strain$Cross== "PWDFemale"] <- "PWD"
+
+######################
+# Setup AP tables #
+######################
+
+################
+# STRAIN TABLE #
+################
+
+#calculate strain averages from mouse averages
+#(try making mean table by batch()
+
+AP_strain_table <- ddply(MLH1_data, c("strain", "sex"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(nMLH1.foci),
+                         mean_co = format(round(  mean(nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(nMLH1.foci), 3),
[…]

### 2017-11-02 testing_setup_data.R

@@ -1,0 +1,163 @@
+##R script for making R.data files 
+## read in data files and adjust with additional columns
+## input: csv file of anon MLH1 counts
+## output: .rdata saved workspace with dataframe and other stat tables (save tables)
+#import MLH1 measures across data sets; AP, BD's and Cast female Koehler
+library(plyr)
+
+########################
+# Setup main dataframe #
+########################
+
+#TODO: remove observations without quality scores.
+#Write a make file that will merge all of the batch files and record the batch
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+MLH1_data = read.csv("data/AnonData.csv", header=TRUE)
+
+big_data = read.csv("data/bigdata.csv", header=TRUE)
+
+source("Func_addCategory.R")
+MLH1_data <- add_category(MLH1_data)
+#set the order of categories (female, male) (cast, dom, musc)
+MLH1_data$category<- factor(MLH1_data$category,levels =c( "G female", "G male", 
+                      "WSB female", "WSB male", "LEWES female", 'LEWES male', 
+                      "PWD female", "PWD male", "MSM female", "MSM male",
+                      "CAST female", "CAST male"), order=T )
+
+
+source("Func_addStrain.R")
+MLH1_data <- add_strain(MLH1_data)
+source("Func_addSex.R")
+MLH1_data <- add_sex(MLH1_data)
+
+original_length = length(MLH1_data$Original.Name)
+MLH1_data <- MLH1_data[MLH1_data$X != "X",]
+
+change <- MLH1_data[MLH1_data$quality == "",]
+
+MLH1_data$category <- as.factor(MLH1_data$category)
+
+MLH1_data$nMLH1.foci <- as.numeric(MLH1_data$nMLH1.foci) #make these numeric just in case there are other characters
+#add a column with male adjusted MLH1 values (+1 to all males)
+MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
+MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
+
+source("Func_addMouse.R")
+MLH1_data <- add_mouse(MLH1_data)
+
+
+#remove mice that had bad stains
+#12sep16_MSM_f3(centromere signal bled into MLH1 signal)
[…]

### 2017-11-02 rmd_report.rmd
added factor order to the Funcs

@@
 ```{r}
+
+#set the order of another column, based on another variable (so that when)
+
+MLH1_data <- add_strain(MLH1_data)
+MLH1_data$strain <- as.factor(MLH1_data$strain)
+
+MLH1_data$strain<- factor(MLH1_data$strain,levels =c( "G", "WSB", "LEWES", 
+                                                        "PWD", "MSM",
+                                                       "CAST", "HMI",
+                                                        "SPRET", "other"), order=T )
+
+MLH1_data$sex<- factor(MLH1_data$sex,levels =c( "female", "male"), order=T )
+
+MLH1_data$category<- factor(MLH1_data$category,levels =c( "G female", "G male", 
+                      "WSB female", "WSB male", "LEWES female", 'LEWES male', "PERC male",
+                      "PWD female", "PWD male", "MSM female", "MSM male",
+                      "CAST female", "CAST male", "HMI female", "HMI male",
+                      "SPRET female", "SPRET male"), order=T )
+
+MLH1_data <- MLH1_data %>%
+  arrange(strain, sex, category, mouse) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+
+MLH1_data <- MLH1_data %>%
+  arrange(category, mouse) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+
 #order the data frame
 MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-#set the order of another column, based on another variable (so that when)
 
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
 # sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
 
@@
 #length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
-MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
+#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
 
-MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
-MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
+#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
+#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
 ```
@@
 #use kable to make tables. warning this might not work for github Rmd.
-kable(AP_strain_table)
+#kable(AP_strain_table)
 ```
@@
 ```{r first boxplots, echo=FALSE, fig.align='center'}
-ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci)) + geom_boxplot(data = MLH1_data, aes(fill = factor(strain) ) ) + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= as.factor(strain)  )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "black", "grey", "lightgrey"))
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey", "black" ))
 ff <- ff   + facet_wrap(~ sex, scales="free")
 ff
+
+#
+#scale_fill_brewer(palette="Dark2")
+
 ```

### 2017-11-02 Func_addCategory.R

@@ -12,15 +12,16 @@ add_category <- function(oldframe){
   dframe <- oldframe
   dframe$category <- "other"
   
-  dframe$category <- ifelse(grepl("_WSB_f", dframe$Original.Name), "WSB female", 
+  dframe$category <- ifelse(grepl("_WSB_f", dframe$Original.Name), "WSB female",
                     ifelse(grepl("_WSB_m",dframe$Original.Name), "WSB male",
                  ifelse(grepl("_G_f", dframe$Original.Name), "G female",
               ifelse(grepl("_G_m",dframe$Original.Name), "G male",
-               ifelse(grepl("_LEW_f", dframe$Original.Name), "LEW female",
-              ifelse(grepl("_LEW_m",dframe$Original.Name), "LEW male",          
-           ifelse(grepl("_LEWES_m", MLH1_data$Original.Name), "LEWES male",    
+                     ifelse(grepl("_LEWES_m", dframe$Original.Name), "LEW male",
+                     ifelse(grepl("_LEW_f", dframe$Original.Name), "LEW female",
+              ifelse(grepl("_LEW_m",dframe$Original.Name), "LEW male",        
+           
             ifelse(grepl("_PERC_f", dframe$Original.Name), "PERC female",
-            ifelse(grepl("_PERC_m",dframe$Original.Name), "PERC male",   
+            ifelse(grepl("_PERC_m",dframe$Original.Name), "PERC male",
                             
                ifelse(grepl("_CAST_m", dframe$Original.Name), "CAST male",
               ifelse(grepl("_CAST_f", dframe$Original.Name), "CAST female",
@@ -40,7 +41,11 @@ add_category <- function(oldframe){
        ifelse(grepl("_CAROLI_f", dframe$Original.Name), "CAROLI female",
                             "other")))))))))))))))))))))))
        
-  dframe$category <- as.character(dframe$category)
+  dframe$category<- factor(dframe$category,levels =c( "G female", "G male", 
+                        "WSB female", "WSB male", "LEW female", 'LEW male', "PERC male",
+                          "PWD female", "PWD male", "MSM female", "MSM male",
+                              "CAST female", "CAST male", "HMI female", "HMI male",
+                                "SPRET female", "SPRET male", "other"), order=T )
   return(dframe)
 }
 #tried writing this with mutate, couldn't get it to work

### 2017-11-02 Func_addSex.R

@@ -1,12 +1,12 @@
 
+#add sex to dataframe
 
-#add sex
 add_sex <- function(dat){
   dframe <- dat
   dframe$sex <- "other"
   dframe$sex <- ifelse(grepl("_f", dframe$Original.Name), "female", 
                        ifelse(grepl("_m",dframe$Original.Name), "male", "other"))
   
-  dframe$sex <- as.character(dframe$sex)
+  dframe$sex<- factor(dframe$sex,levels =c( "female", "male"), order=T )
   return(dframe)
 }

### 2017-11-02 Func_addStrain.R

@@ -18,6 +18,11 @@ add_strain <- function(dat){
              ifelse(grepl("_SPIC_", dframe$Original.Name), "SPIC",
           ifelse(grepl("_CAROLI_", dframe$Original.Name), "CAROLI",
                                        "other"))))))))))
-  dframe$strain <- as.character(dframe$strain)
+  
+  #the ordering factor below deletes all strain entries
+  dframe$strain<- factor(dframe$strain,levels =c( "G", "WSB", "LEWES", "PERC",
+                                                        "PWD", "MSM",
+                                                        "CAST", "HMI",
+                                                        "SPRET", "other"), order=T )
   return(dframe)
 }

### 2017-11-02 setup_data.R

@@ -5,8 +5,9 @@
 ## df of MLH1 should go from 12cols to 17cols
 #import MLH1 measures across data sets; AP, BD's and Cast female Koehler
 # 1. MLH1_data, tables: AP strains
-library(dplyr)
+
 library(plyr)
+library(dplyr)
 
 ########################
 # Setup main dataframe #
@@ -42,19 +43,27 @@ source("Func_addCategory.R")
 MLH1_data <- add_category(MLH1_data)
 #set the order of categories (female, male) (cast, dom, musc)
 MLH1_data$category<- factor(MLH1_data$category,levels =c( "G female", "G male", 
-                      "WSB female", "WSB male", "LEWES female", 'LEWES male', 
+                      "WSB female", "WSB male", "LEWES female", 'LEWES male', "PERC male",
                       "PWD female", "PWD male", "MSM female", "MSM male",
-                      "CAST female", "CAST male", "SPRET female", "SPRET male"), order=T )
-
+                      "CAST female", "CAST male", "HMI female", "HMI male",
+                      "SPRET female", "SPRET male"), order=T )
 
 source("Func_addStrain.R")
 MLH1_data <- add_strain(MLH1_data)
+
+#the ordering factor below deletes all strain entries
+#MLH1_data$strain<- factor(MLH1_data$category,levels =c( "G", "WSB", "LEWES", "PERC",
+#                                                        "PWD", "MSM",
+ #                                                       "CAST", "HMI",
+#                                                        "SPRET"), order=T )
+
 source("Func_addSex.R")
 MLH1_data <- add_sex(MLH1_data)
-
+#MLH1_data$sex<- factor(MLH1_data$category,levels =c( "female", "male"), order=T )
 
 #add a column with male adjusted MLH1 values (+1 to all males)
 MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
+
 MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
 
 source("Func_addMouse.R")

### 2017-11-02 rmd_report.rmd
fixed some annoying ordering issues

--- 2017-11-02 rmd_report.rmd
@@
 ```{r}
+
+#set the order of another column, based on another variable (so that when)
+
+MLH1_data <- MLH1_data %>%
+  arrange(strain, sex, category, mouse) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+
+MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+# sort your dataframe, by the focal categories
+#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
+#Add fake CAST female data
+last_row <-data.frame(Batch = c("100", "100"), X= c("",""), Original.Name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )
+
+MLH1_data <- rbind(MLH1_data, last_row)
+
+#count the non-quality measures,  #remove non qualit
+#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
+#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
+
+#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
+#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
+```
+
+#### Mice table
+
+Table of the number of mice used and MLH1 stats. (made with kable).
+
+```{r, fig.height=2, fig.width=4, echo=FALSE}
+#this table is saved in the .RData file
+#use kable to make tables. warning this might not work for github Rmd.
+#kable(AP_strain_table)
+```
+
+MLH1_data <- MLH1_data %>%
+  arrange(category) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+#order the data frame
+MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
+
+#### Initial Patterns from MLH1 distributions
+
+```{r first boxplots, echo=FALSE, fig.align='center'}
+#the below ordering works!! keep it
+MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEWES",
+                              "WSB", "PWD","MSM","CAST", "HIM", "SPRET"), order=T )
 #order the data frame
@@
 
-#Add fake CAST female data
-last_row <-data.frame(Batch = c("100", "100"), X= c("",""), Original.Name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )
-
-MLH1_data <- rbind(MLH1_data, last_row)
-
-#count the non-quality measures,  #remove non qualit
-#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
-MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
-
-MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
-MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
-```
-
-#### Mice table
-
-Table of the number of mice used and MLH1 stats. (made with kable).
-
-```{r, fig.height=2, fig.width=4, echo=FALSE}
-#this table is saved in the .RData file
-#use kable to make tables. warning this might not work for github Rmd.
-kable(AP_strain_table)
-```
-
-#### Initial Patterns from MLH1 distributions
-
-```{r first boxplots, echo=FALSE, fig.align='center'}
-ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci)) + geom_boxplot(data = MLH1_data, aes(fill = factor(strain) ) ) + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "black", "grey", "lightgrey"))
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
 ff <- ff   + facet_wrap(~ sex, scales="free")
 ff
+
+#
+
 ```

### 2017-11-02 Func_addCategory.R

@@ -14,7 +14,8 @@ add_category <- function(oldframe){
   
   dframe$category <- ifelse(grepl("_WSB_f", dframe$Original.Name), "WSB female",
                     ifelse(grepl("_WSB_m",dframe$Original.Name), "WSB male",
-                 ifelse(grepl("_G_f", dframe$Original.Name), "G female",
+                
+                   ifelse(grepl("_G_f", dframe$Original.Name), "G female",
               ifelse(grepl("_G_m",dframe$Original.Name), "G male",
                      ifelse(grepl("_LEWES_m", dframe$Original.Name), "LEW male",
                      ifelse(grepl("_LEW_f", dframe$Original.Name), "LEW female",
@@ -41,8 +42,8 @@ add_category <- function(oldframe){
        ifelse(grepl("_CAROLI_f", dframe$Original.Name), "CAROLI female",
                             "other")))))))))))))))))))))))
        
-  dframe$category<- factor(dframe$category,levels =c( "G female", "G male", 
-                        "WSB female", "WSB male", "LEW female", 'LEW male', "PERC male",
+  dframe$category<- factor(dframe$category,levels =c( "WSB female", "WSB male","G female", "G male", 
+                         "LEW female", 'LEW male', "PERC male",
                           "PWD female", "PWD male", "MSM female", "MSM male",
                               "CAST female", "CAST male", "HMI female", "HMI male",
                                 "SPRET female", "SPRET male", "other"), order=T )

### 2017-11-02 Func_addStrain.R

@@ -7,7 +7,7 @@ add_strain <- function(dat){
   
   dframe$strain <- ifelse(grepl("_WSB", dframe$Original.Name), "WSB", 
               ifelse(grepl("_G_",dframe$Original.Name), "G",
-           ifelse(grepl("_LEW_", dframe$Original.Name), "LEWES",
+           ifelse(grepl("_LEW", dframe$Original.Name), "LEWES",
                                   
             ifelse(grepl("_MSM_", dframe$Original.Name), "MSM",           
            ifelse(grepl("_PWD_", dframe$Original.Name), "PWD",
@@ -20,7 +20,7 @@ add_strain <- function(dat){
                                        "other"))))))))))
   
   #the ordering factor below deletes all strain entries
-  dframe$strain<- factor(dframe$strain,levels =c( "G", "WSB", "LEWES", "PERC",
+  dframe$strain<- factor(dframe$strain,levels =c( "WSB", "G", "LEWES", "PERC",
                                                         "PWD", "MSM",
                                                         "CAST", "HMI",
                                                         "SPRET", "other"), order=T )

### 2017-11-02 setup_data.R

@@ -21,6 +21,8 @@ MLH1_data = read.csv("data/AnonData.csv", header=TRUE )
 original_DF = MLH1_data
 original_length <- length(MLH1_data$Original.Name)
 
+MLH1_data$Original.Name <- as.character(MLH1_data$Original.Name)
+
 MLH1_data$n <- as.character(MLH1_data$n)
 MLH1_data$nMLH1.foci<- as.character(MLH1_data$nMLH1.foci)
 MLH1_data$quality<- as.character(MLH1_data$quality)
@@ -42,24 +44,11 @@ MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]
 source("Func_addCategory.R")
 MLH1_data <- add_category(MLH1_data)
 #set the order of categories (female, male) (cast, dom, musc)
-MLH1_data$category<- factor(MLH1_data$category,levels =c( "G female", "G male", 
-                      "WSB female", "WSB male", "LEWES female", 'LEWES male', "PERC male",
-                      "PWD female", "PWD male", "MSM female", "MSM male",
-                      "CAST female", "CAST male", "HMI female", "HMI male",
-                      "SPRET female", "SPRET male"), order=T )
 
 source("Func_addStrain.R")
 MLH1_data <- add_strain(MLH1_data)
-
-#the ordering factor below deletes all strain entries
-#MLH1_data$strain<- factor(MLH1_data$category,levels =c( "G", "WSB", "LEWES", "PERC",
-#                                                        "PWD", "MSM",
- #                                                       "CAST", "HMI",
-#                                                        "SPRET"), order=T )
-
 source("Func_addSex.R")
 MLH1_data <- add_sex(MLH1_data)
-#MLH1_data$sex<- factor(MLH1_data$category,levels =c( "female", "male"), order=T )
 
 #add a column with male adjusted MLH1 values (+1 to all males)
 MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
@@ -69,16 +58,20 @@ MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
 source("Func_addMouse.R")
 MLH1_data <- add_mouse(MLH1_data)
 
+#reorder dataframe
[…]

### 2017-11-02 rmd_report.rmd
some updates to Rmd, removed fake CAST female

@@
 
-```{r}
+```{r,echo=FALSE}
 
@@
 #MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-#Add fake CAST female data
-last_row <-data.frame(Batch = c("100", "100"), X= c("",""), Original.Name=c("12dec18_20dec20_CAST_f1_sp1_12_1_rev.tif", "12dec18_20dec20_CAST_f1_sp1_12_2_rev.tif"), Random.Name = c("1234567.tif","123456789.tif"), quality=c(1,1),nMLH1.foci =c(25,22), XY.paired =c("no","no"), REDO.crop =c("no","no"),n =c(20,20), achiasmate=c(0,0),asynased=c(0,0),notes=c("",""),category = c("CAST female","CAST female"), strain= c("CAST","CAST"), sex=c("female","female"),adj_nMLH1.foci=c(25,22),mouse=c("20dec20_CAST_f1","20dec20_CAST_f1") )
-
-MLH1_data <- rbind(MLH1_data, last_row)
 
@@
 MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEWES",
-                              "WSB", "PWD","MSM","CAST", "HIM", "SPRET"), order=T )
+                              "WSB", "PWD","MSM","CAST","HMI","SPRET", "other"), order=T )
 #order the data frame
@@
 ff <- ff   + facet_wrap(~ sex, scales="free")
+
+#boxplots by strain
+ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
+ss <- ss   + facet_wrap(~ sex, scales="free")
+
 ff
-
-#
-
+ss
 ```
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",
-  "coral1", "#E69F00", "yellowgreen","black", "grey"  ))
+        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
 high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 

### 2017-11-02 rmd_report.rmd
cleaning up rmd report file

@@
 latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.
-```{r, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
-#TODO, add a line for mouse and strain
+```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
+#what is the purpose of this block
 
-#this creates a slot for CAST female, but no data
-#each of the indiviual dataframes can be acessed with df_list$ ---- category
-# for each sub-data frame, make mouse specific table,
-#
-#add mouse mean to the df
-df_list <- split(MLH1_data, as.factor(MLH1_data$category))
+df_list <- split(MLH1_data, (MLH1_data$category))
 
+#remove sections which are in category but have no data
+df_list <- df_list[ -c(7, 12, 14,16, 18) ]
+#7, PERC. 12 CASTF, 14 HMI f , 16 SPRET f. 18 other
+
+df_list 
 o=0
@@
 ```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE }
-#geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-#for (gg in 1:length(plot_list) ){
-#  show(plot_list[gg])
-  #invisible(lapply(obj,function(x) plot(x,main="some plot")))
-#}
+geom_hline(aes(yintercept = med, group = gr), colour = 'red')
+for (gg in 1:length(plot_list) ){
+  show(plot_list[gg])
+  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
+}
 ```

### 2017-11-03 rmd_report.rmd
added batch color to indiviual mouse distributions

@@
 }
-#remove CAST, this column isn't added to MLH1 col, it's added to df_list
-#The above adds a column of mouse specific averages to a giant list of dataframes. Each dataframe in the list is specific to a category.
-#the next step is to make plots from this giant list of dataframes
-# I can delete all of the code below that by hand makes
-
-### not make a loop for ggplot... or giant ggplot base that uses latice smartly
-# make new ggplots -- that are replaces
 
@@
   pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-  pf <- pf + geom_point(stat = "identity")+ geom_jitter()
+  pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
   pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
@@
 
-#aes must be same length --- 
+#aes must be same length ---   #geom_point(stat = "identity") +
 # the last one is correct....?

### 2017-11-09 rmd_report.rmd
moved first functions to src dir

@@
 #MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
-
 highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
@@
 #what is the purpose of this block
-
 df_list <- split(MLH1_data, (MLH1_data$category))
-
 #remove sections which are in category but have no data
@@
 geom_hline(aes(yintercept = med, group = gr), colour = 'red')
+
 for (gg in 1:length(plot_list) ){

### 2017-11-15 1rmd_report.rmd
renamed to resolve

--- 2017-11-03 rmd_report.rmd
@@
 library(ggplot2)
-library(pwr)
+#library(pwr)
 library(plyr)
@@
 library(dplyr)
-setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+setwd("C:/Users/April/Desktop/MLH1repo")
 load(file="MLH1_data_setup.RData")
@@
 ```{r,echo=FALSE}
-
 #set the order of another column, based on another variable (so that when)
@@
 
-MLH1_data <- MLH1_data %>%
-  arrange(category) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-#order the data frame
-MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-
 #### Initial Patterns from MLH1 distributions
@@
 
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+#reduncant
+#MLH1_data <- MLH1_data %>%
+ # arrange(strain, sex, mouse) %>%
+#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
 # sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
 
-ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
@@
 ```{r, histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
-
 histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))

### 2017-11-16 src/Func_addCategory.R
edited functions to read SPIC coding

@@ -36,8 +36,8 @@ add_category <- function(oldframe){
      
         ifelse(grepl("_SPRET_f", dframe$Original.Name), "SPRET female",
         ifelse(grepl("_SPRET_m", dframe$Original.Name), "SPRET male",
-        ifelse(grepl("_SPIC_m", dframe$Original.Name), "SPIC male",
-        ifelse(grepl("_SPIC_f", dframe$Original.Name), "SPIC female",          
+        ifelse(grepl("_SPI_m", dframe$Original.Name), "SPIC male",
+        ifelse(grepl("_SPI_f", dframe$Original.Name), "SPIC female",          
         ifelse(grepl("_CAROLI_m", dframe$Original.Name), "CAROLI male",
        ifelse(grepl("_CAROLI_f", dframe$Original.Name), "CAROLI female",
                             "other")))))))))))))))))))))))
@@ -46,7 +46,7 @@ add_category <- function(oldframe){
                          "LEW female", 'LEW male', "PERC male",
                           "PWD female", "PWD male", "MSM female", "MSM male",
                               "CAST female", "CAST male", "HMI female", "HMI male",
-                                "SPRET female", "SPRET male", "other"), order=T )
+                                "SPRET female", "SPRET male", "SPIC female", "SPIC male", "other"), order=T )
   return(dframe)
 }
 #tried writing this with mutate, couldn't get it to work

### 2017-11-16 src/Func_addStrain.R

@@ -15,7 +15,7 @@ add_strain <- function(dat){
                 ifelse(grepl("_CAST_", dframe$Original.Name), "CAST",           
          ifelse(grepl("_HMI_", dframe$Original.Name), "HMI",
               ifelse(grepl("_SPRET_", dframe$Original.Name), "SPRET",           
-             ifelse(grepl("_SPIC_", dframe$Original.Name), "SPIC",
+             ifelse(grepl("_SPI_", dframe$Original.Name), "SPIC",
           ifelse(grepl("_CAROLI_", dframe$Original.Name), "CAROLI",
                                        "other"))))))))))
   
@@ -23,6 +23,6 @@ add_strain <- function(dat){
   dframe$strain<- factor(dframe$strain,levels =c( "WSB", "G", "LEWES", "PERC",
                                                         "PWD", "MSM",
                                                         "CAST", "HMI",
-                                                        "SPRET", "other"), order=T )
+                                                        "SPRET", "SPIC", "other"), order=T )
   return(dframe)
 }

### 2017-11-16 src/Func_addSubsp.R

@@ -2,28 +2,34 @@
 
 #the column might need to be changed
 
-add_subsp <- function(dat) {
+add_subsp <- function(oldframe) {
 #assign subspecies to tables
-subsp <- ifelse(grepl("WSB", Table_BD_strain$Cross), "Dom", 
-           ifelse(grepl("G", Table_BD_strain$Cross), "Dom",
-            ifelse(grepl("LEW", Table_BD_strain$Cross), "Dom", 
-             ifelse(grepl("LEWES", Table_BD_strain$Cross), "Dom",    
-            ifelse(grepl("PERA", Table_BD_strain$Cross), "Dom",
+  dframe <- oldframe
+  dframe$subsp <- "other"
+  
+  dframe$subsp <- ifelse(grepl("WSB", dframe$strain), "Dom", 
+           ifelse(grepl("G", dframe$strain), "Dom",
+            ifelse(grepl("LEW", dframe$strain), "Dom", 
+             ifelse(grepl("LEWES", dframe$strain), "Dom",    
+            ifelse(grepl("PERA", dframe$strain), "Dom",
                                             
-        ifelse(grepl("CAST", Table_BD_strain$Cross), "Cast",
-           ifelse(grepl("CIM", Table_BD_strain$Cross), "Cast",
+        ifelse(grepl("CAST", dframe$strain), "Cast",
+           ifelse(grepl("CIM", dframe$strain), "Cast",
+                  ifelse(grepl("HMI", dframe$strain), "Cast",
                                                           
-        ifelse(grepl("MSM", Table_BD_strain$Cross), "Musc-Cast",                                       
-           ifelse(grepl("PWD", Table_BD_strain$Cross), "Musc", 
-            ifelse(grepl("CZECHI", Table_BD_strain$Cross), "Musc", 
-         ifelse(grepl("PWDFemale", Table_BD_strain$Cross), "Musc",         
-                                                                                      
-        ifelse(grepl("PANCEVO", Table_BD_strain$Cross), "Spic", 
-             ifelse(grepl("CAROLI", Table_BD_strain$Cross), "Caroli", 
+        ifelse(grepl("MSM", dframe$strain), "Musc",                                       
+           ifelse(grepl("PWD", dframe$strain), "Musc", 
+            ifelse(grepl("CZECHI", dframe$strain), "Musc", 
+         ifelse(grepl("PWDFemale", dframe$strain), "Musc",         
+        
[…]

### 2017-11-16 fresh_PnD.R
trying to debg report.rmd

@@ -1,0 +1,250 @@
+library(plyr)
+library(dplyr)
+library(ggplot2)
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="MLH1_data_setup.RData")
+
+###################
+# Setting up Data #
+###################
+#this WSB female mouse is likely skweing the data -- few cells
+#there's a Dom female 4apr15_WSB_f2 that could be skewing results
+MLH1_data<- MLH1_data[!grepl("4apr15_WSB_f2", MLH1_data$mouse) , ] #1649
+#leweS FEMALES also seem to have a high variance across means
+
+#1. subset data by sex
+Mouse_table <- ddply(MLH1_data, c("strain", "sex", "mouse"), summarise,
+                      Nmice = length(unique(mouse)),
+                      Ncells  = length(adj_nMLH1.foci),
+                      mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                      var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                      sd   = round(sd(adj_nMLH1.foci), 3),
+                      se   = round(sd / sqrt(Ncells), 3),
+                      cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#add subsp to this df
+
+source("src/Func_addSubsp.R")
+Mouse_table <- add_subsp(Mouse_table)
+
+#<change from cal from full MLh1-data
+#calq'd from mouse averages
+Strain_table <- ddply(Mouse_table, c("strain", "sex"), summarise,
+                            Nmice = length(unique(mouse)),
+                            Ncells  = sum(Ncells),
+                            strain.mean = format( round(  mean(mean_co), 3 ), nsmall=3),
+                            strain.var = round( var(mean_co), 3),
+                            strain.sd   = round( sd(mean_co), 3),
+#decide if se denominator is n mice or ncells                       
[…]

### 2017-11-16 rmd_report.rmd
trying to debg report.rmd
fixed the rmd file

@@
 ```{r, histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
-
+#consider making distribution with quality considered...
 histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
@@
 #remove sections which are in category but have no data
-df_list <- df_list[ -c(7, 12, 14,16, 18) ]
-#7, PERC. 12 CASTF, 14 HMI f , 16 SPRET f. 18 other
+df_list <- df_list[ -c(7, 12, 14,19, 20) ]
+#7: PERC male, 12:CAST female, 14:HMI female, 19:SPIC male
 

### 2017-12-07 rmd_report.rmd
catching up on updates

@@
   mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-#order the data frame
+
 MLH1_data <- with(MLH1_data, MLH1_data[order(category),])

### 2017-12-07 test4pvalues.Rmd
catching up on updates

---
title: "test_for_p.values"
author: "April Peterson"
date: "December 6, 2017"
output: html_document
---

first the larger data

I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.

```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
#what is the purpose of this block
pMLH1_data <- MLH1_data[MLH1_data$category == "PWD male", ]

#4jan17_LEW_f3, 4jan17_LEW_f6
two_mouse_data <- MLH1_data[ (MLH1_data$mouse == "4jan17_LEW_f3" | MLH1_data$mouse == "4jan17_LEW_f6"), ]

cat_list <- split(pMLH1_data, (pMLH1_data$mouse))
#list of dfs, that are split by mouse
#cat_list$'4jan17_LEW_f3'$quality  AND  cat_list[[1]]$quality
#cat_list[1]$quality

#remove sections which are in category but have no data
cat_list <- cat_list[ -c(7, 12, 14,19, 20) ]
#7: PERC male, 12:CAST female, 14:HMI female, 19:SPIC male

#df_list 
o=0
i=0

#split str.  # PWD - split by mouse,

new_dat = list()
for(i in 1:(length(cat_list))){
  new_dat[i] <- data.frame()
  new_dat[i]$pvals <- replicate(100, t.test(sample(cat_list[[i]]$nMLH1.foci, 5) ,    sample(cat_list[[i]]$nMLH1.foci, 5))$p.value )

new_dat$m <- cat_list[[i]]$mouse 
}

g <- as.data.frame(new_dat)

# ggplot only works with DFs

# g 7 columns, 
plot_list = list()
gcolnames <- colnames(g)
for(h in 1:length(gcolnames)){
  print(h)
  pf <- ggplot(g, aes(x = g[h])) + geom_histogram() + xlim(c(0,1))
  pf
  plot_list[[h]] =   pf
}

for (gg in 1:length(plot_list) ){
  print(gg)
  plot_list[gg][[1]]
  #show(plot_list[gg])
  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
}

```

The above code made a big dataframe with pvalues for small sample sizes

```{r}
plyMLH1_data <- MLH1_data[MLH1_data$mouse == "18may15_PWD_m1", ]

bb <- replicate(1000, t.test(sample(plyMLH1_data$nMLH1.foci, length(plyMLH1_data$nMLH1.foci)/2) ,    sample(plyMLH1_data$nMLH1.foci, 20))$p.value )

ggplot(data.frame(bb), aes(x=bb)) + 
  geom_histogram() #+ geom_tile()

```

### 2017-12-07 fresh_PnD.R

@@ -4,20 +4,23 @@ library(ggplot2)
 setwd("C:/Users/alpeterson7/Documents/MLH1repo")
 load(file="MLH1_data_setup.RData")
 
+## ToDO change SE to Sd ##
+
+
 ###################
 # Setting up Data #
 ###################
-#this WSB female mouse is likely skweing the data -- few cells
+
+#remove data that might be low quality
 #there's a Dom female 4apr15_WSB_f2 that could be skewing results
 MLH1_data<- MLH1_data[!grepl("4apr15_WSB_f2", MLH1_data$mouse) , ] #1649
 MLH1_data<- MLH1_data[!grepl("12sep16_MSM_f1", MLH1_data$mouse) , ]
 
 MLH1_data<- MLH1_data[!grepl("4jan17_LEW_f1", MLH1_data$mouse) , ]#one duplicate image found
 #SPRET females from batch5, are not that good
-#
 #leweS FEMALES also seem to have a high variance across means
 
-#1. subset data by sex
+#1. Create table of mouse level stats for MLH1 foci count
 Mouse_table <- ddply(MLH1_data, c("strain", "sex", "mouse"), summarise,
                       Nmice = length(unique(mouse)),
                       Ncells  = length(adj_nMLH1.foci),
@@ -27,12 +30,28 @@ Mouse_table <- ddply(MLH1_data, c("strain", "sex", "mouse"), summarise,
                       se   = round(sd / sqrt(Ncells), 3),
                       cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
 )
-#add subsp to this df
-
+#add subsp to this table
 source("src/Func_addSubsp.R")
 Mouse_table <- add_subsp(Mouse_table)
 
-#<change from cal from full MLh1-data
+#create a mouse level table of best quality cells
+
+Q1_Mouse_table <- ddply(MLH1_data[MLH1_data$quality <= 2,], c("strain", "sex", "mouse"), summarise,
+                     Nmice = length(unique(mouse)),
+                     Ncells  = length(adj_nMLH1.foci),
[…]

### 2017-12-08 test4pvalues.Rmd
working on the p value script

@@
 first the larger data 
+seperate data into >20 cells or less
 
+```{r}
+twenty <- AP_mouse_table[AP_mouse_table$Ncells > 20,]
+lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,]
+#these will give the list of mice
+
+#subset MLH1_data by these lists
+
+```
 I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.

### 2017-12-09 test4pvalues.Rmd

@@
 
+```{r}
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("Documents/MLH1repo")
+setwd("C:/Users/April/Desktop/MLH1repo")
+load(file="MLH1_data_setup.RData")
+library(ggplot2)
+```
+
 first the larger data 
@@
 ```{r}
-twenty <- AP_mouse_table[AP_mouse_table$Ncells > 20,]
+twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,]
 lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,]
@@
 
+twenty_data <- MLH1_data[(unique(MLH1_data$mouse) %in% twenty$mouse),]
+lessTwenty_data <- MLH1_data[(unique(MLH1_data$mouse) %in% lessTwenty$mouse),]
+
 ```
 I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.
+
+```{r}
+
+outlist = list()
+for(o in 1:length(unique(twenty$mouse))){
+  print(twenty$mouse[o])
+  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[o],]
+
+  outlist[o] <- list(replicate(100, t.test(sample(mus_data$adj_nMLH1.foci, 15, replace = TRUE),
+                        sample(mus_data$adj_nMLH1.foci, 15, replace = TRUE) )$p.value ) )
+  }
+names(outlist) <- twenty$mouse
+#outlist is not a list of p-value list. ()
+
+#find a way to print out histogrms 
+plot_list = list()
+for (i in 1:length(twenty$mouse)){
+  print(i)  
+  try_df <- as.data.frame(outlist[[i]])
+  colnames(try_df) <- "pvals"
+  p = ggplot(try_df) +
+        geom_histogram(aes(x=pvals) )#xlim
+    plot_list[[i]] = p
+}
+
+# Save plots to tiff. Makes a separate file for each plot.
+for(i in 1:length(plot_list) ) {
+    file_name = paste("histy_", i, ".tiff", sep="")
+    tiff(file_name)
+    print(plot_list[[i]])
+    dev.off()
+}
+
+hist(outlist[[5]])
+plot()
+
+```
 
@@
 #cat_list[1]$quality
-
 #remove sections which are in category but have no data
-cat_list <- cat_list[ -c(7, 12, 14,19, 20) ]
-#7: PERC male, 12:CAST female, 14:HMI female, 19:SPIC male
 

### 2017-12-10 test4pvalues.Rmd
working on pval script

--- 2017-12-07 test4pvalues.Rmd
@@
 
+```{r}
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("Documents/MLH1repo")
+#setwd("C:/Users/April/Desktop/MLH1repo")
+load(file="MLH1_data_setup.RData")
+library(ggplot2)
+library(plyr)
+library(dplyr)
+library(raster)
+library(reshape2)
+```
+
 first the larger data 
+seperate data into >20 cells or less
 
+```{r}
+twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,]
+lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,]
+#these will give the list of mice
+
+#subset MLH1_data by these lists. (fixed line below, )
+twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]
+
+twenty_data_table <- ddply(twenty_data, c("mouse"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         cV = cv(adj_nMLH1.foci),
+                          var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3)
+)
+
+lessTwenty_data <- MLH1_data[(unique(MLH1_data$mouse) %in% lessTwenty$mouse),]
+
+#there -- the data counted in for AP_mouse_table doesn't match how 
+#my smaller df made for each mouse in the loops.
+#the table said Ncells was 21, but the smaller df is only 13 long
+
+```
 I am trying to display plots for each mouse. For each mouse I want to run the replicate line code.
+
+```{r}
+sim_iterations <- 1000
+
+outlist15 = list()
+for(fif in 1:length(unique(twenty$mouse))){
+  print(twenty$mouse[fif])
+  #
+  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[fif],]
+
+  outlist15[fif] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE),
+                        sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE) )$p.value ) )
+  }
+names(outlist15) <- twenty$mouse
+#outlist is not a list of p-value list. ()
+
+outlist10 = list()
+for(ten in 1:length(unique(twenty$mouse))){
+  print(twenty$mouse[ten])
+  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[ten],]
+
+  outlist10[ten] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE),
+                        sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE) )$p.value ) )
+  }
+names(outlist10) <- twenty$mouse
+
+outlist20 = list()
+for(tw in 1:length(unique(twenty$mouse))){
+  print(twenty$mouse[tw])
+  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[tw],]
+
+  outlist20[tw] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE),
+                sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE) )$p.value ) )
+  }
+names(outlist20) <- twenty$mouse
+#outlist is not a list of p-value list. ()
+
+plot_list20 = list()
+for (i in 1:length(twenty$mouse)){
+  print(i)  
+  try_df <- as.data.frame(outlist20[[i]])
+  colnames(try_df) <- "pvals"
+  Title_mus_name <- paste("1000 iterations\n, Test of 2 Samples of 20, \n", names(outlist20[i]) )
+  p20 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
+        geom_histogram(aes(x=pvals) )#xlim
+    plot_list20[[i]] <- p20
+}
+
+plot_list10 = list()
+for (i in 1:length(twenty$mouse)){
+  print(i)  
+  try_df <- as.data.frame(outlist10[[i]])
+  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 10, \n", names(outlist10[i]) )
+  colnames(try_df) <- "pvals"
+  p10 <- ggplot(try_df) + ggtitle( Title_mus_name ) +xlim(0,1)+
+        geom_histogram(aes(x=pvals) )#xlim
+  #tryfiting 3 plots
+
+    plot_list10[[i]] <- p10
+}
+
+plot_list15 = list()
+for (i in 1:length(twenty$mouse)){
+  print(i)  
+  try_df <- as.data.frame(outlist15[[i]])
+  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 15, \n", names(outlist15[i]) )
+  colnames(try_df) <- "pvals"
+  p15 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
+        geom_histogram(aes(x=pvals) )#xlim
+  #tryfiting 3 plots
+    plot_list15[[i]] <- p15
+}
+
+#make list of scatter plots
+sctr_plot = list()
+for (i in 1:length(twenty$mouse)){
+  print(i)  
+  df10 <- as.data.frame(outlist10[[i]])
+  df15 <- as.data.frame(outlist15[[i]])
+  df20 <-as.data.frame(outlist20[[i]])
+  names(df10) = "pvals"
+  names(df15)= "pvals"
+  names(df20)= "pvals"
+  allDF <- data.frame(pvals = rbind(df10, df15, df20),
+    DFsets = c (rep("samp10",length(df10[,1]) ), 
+              rep("samp15",length(df10[,1])), 
+              rep("samp20", length(df10[,1])) )
+                      )
+    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist15[i]) )
+  colnames(try_df) <- "pvals"
+
+  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
+        geom_jitter(aes(x=DFsets, y=pvals) )
+    sctr_plot[[i]] <- scplot
+}
+
+source("src/MultiPlot.R")
+
+multiplot(plot_list10[[8]],plot_list20[[8]], cols=1)
+
+# Saves multiple plots to a tiff file for each mouse
+for(i in 1:5) {
+    file_name = paste("Sample_sizes_p_", i, ".tiff", sep="")
+    tiff(file_name)
+    #par( mfrow=c(1,2) )
+    multiplot(sctr_plot[[i]], plot_list10[[i]], plot_list15[[i]], plot_list20[[i]],cols=2)
+    dev.off()
+}
+
+```
+
+```{r}
+
+#scatterplot of pvales
+# the x scales should be the different sample sizes
+#so one ggplot per mouse
+
+#mush all of the outlists into 1 DF
+#
+
+#make list of plots
+sctr_plot = list()
+for (i in 1:length(twenty$mouse)){
+  print(i)  
+  df10 <- as.data.frame(outlist10[[i]])
+  df15 <- as.data.frame(outlist15[[i]])
+  df20 <-as.data.frame(outlist20[[i]])
+  names(df10) = "pvals"
+  names(df15)= "pvals"
+  names(df20)= "pvals"
+
+  allDF <- data.frame(pvals = rbind(df10, df15, df20),
+    DFsets = c (rep("samp10",length(df10[,1]) ), 
+              rep("samp15",length(df10[,1])), 
+              rep("samp20", length(df10[,1])) )
+                      )
+
+    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist15[i]) )
+  colnames(try_df) <- "pvals"
+
+  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
+        geom_jitter(aes(x=DFsets, y=pvals) )#xlim
+  #tryfiting 3 plots
+    sctr_plot[[i]] <- scplot
+}
+
+for(i in 1:5) {
+    file_name = paste("Scatter_plots_", i, ".tiff", sep="")
+    tiff(file_name)
+    #par( mfrow=c(1,2) )
+    print(sctr_plot[[i]] )
+    #multiplot(plot_list10[[i]], plot_list15[[i]], plot_list20[[i]])
+    #print(c(plot_list[[i]], plot_list20[[i]]) )
+
+    dev.off()
+}
+
+```
 

### 2017-12-10 SamplingEffects.R

@@ -1,0 +1,166 @@
+# Script for displaying p values from T-tests
+# input: MLH1 data environment
+# output: 
+# 
+
+#purpose, compare how different sample sizes of MLH1 counts drawn from the same mouse 
+#(may appear to come from different distributions)
+# (via pvalue of t-tests)
+# Given that distributions of MLH1 counts have different variance across mice, 
+# (different cell sample sizes from each mouse may not reflect the true distribution
+#(there may be more error in testing the differences in distributions due to smaller sample
+#sizes.
+
+# Larger Goal, is to assess which mice have sufficient cells observations to acurately discribe the MLH1 count distribution.
+# or observing more cells wouldn't change the distribution.
+
+
+#LOAD data
+
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("Documents/MLH1repo")
+#setwd("C:/Users/April/Desktop/MLH1repo")
+load(file="MLH1_data_setup.RData")
+library(ggplot2)
+library(plyr)
+library(dplyr)
+library(raster)
+library(reshape2)
+
+
+#seperate large dataframe by how many cells are in the samples
+twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,]
+lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,]
+#these will give the list of mice
+
+#subset MLH1_data by these lists. (fixed line below, )
+twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]
+lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]
+
+
+# Create dataframes that iterate through different sample sizes of MLH1 counts 
+# Outlists are dataframes, filled with pvalues from t-tests of different sample sizes.
+# Q: How do the 20cells or greater mice observations compare with the lessthan 20 cell
+# mouse observations.
+
+sim_iterations <- 10000
+
+outlist10 = list()
+for(ten in 1:length(unique(twenty$mouse))){
+  print(twenty$mouse[ten])
+  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[ten],]
[…]

### 2018-02-08 1rmd_report.rmd
new md report files uploaded

@@
 author: "April Peterson"
-date: "August 29, 2017"
+date: "Feb 6, 2018"
 output: github_document
@@
 ```{r setup, echo=FALSE, include = FALSE}
+#
 library(knitr)
@@
 library(dplyr)
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
-setwd("C:/Users/April/Desktop/MLH1repo")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("C:/Users/April/Desktop/MLH1repo")
 load(file="MLH1_data_setup.RData")
@@
 #use kable to make tables. warning this might not work for github Rmd.
-#kable(AP_strain_table)
+n <- kable(AP_mouse_table)
+n
 ```
+
+MLH1_data <- MLH1_data %>%
+  arrange(category) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+
+MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
 
@@
 MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEWES",
-                              "WSB", "PWD","MSM","CAST","HMI","SPRET", "other"), order=T )
+                              "WSB", "PWD","MSM","KAZ","CAST","HMI","SPRET","SPIC", "other"), order=T )
 #order the data frame
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey" ))
 ff <- ff   + facet_wrap(~ sex, scales="free")
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey" ))
 ss <- ss   + facet_wrap(~ sex, scales="free")
@@
 ```{r, histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
+#consider making distribution with quality considered...
+
 histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
@@
 
+#are there no KAZ cells of high quality?
 qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
+        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
 high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
@@
 ```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
-ff
+qq
 high_qual_boxplot
+#not sure why order got messed up again...
 ```
@@
 #remove sections which are in category but have no data
-df_list <- df_list[ -c(7, 12, 14,16, 18) ]
-#7, PERC. 12 CASTF, 14 HMI f , 16 SPRET f. 18 other
+df_list <- df_list[ -c(7, 12, 14,16,19, 21) ]
+#7: PERC male,   12: #KAZ female! 14:CAST female 16:HMI female,19:SPIC male, 21:SPIC male
 
+#df list, is a list of df's which were split() from BIG df by category
 df_list 
@@
 
+#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
 for(i in 1:(length(df_list))){
- # print(i)
+ # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
   df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
@@
   )
- # print(sub_mouse_means)
-  #below should work -- I think I need to go througheach row of the df_list[[i]]
+# this loop is going though each row of dataframes within the df.list and assining the mouse mean
   for(o in 1:length(df_list[[i]]$Original.Name)){

### 2018-02-14 src/Func_addCategory.R
refining categories

@@ -40,10 +40,13 @@ add_category <- function(oldframe){
         ifelse(grepl("_SPRET_f", dframe$Original.Name), "SPRET female",
         ifelse(grepl("_SPRET_m", dframe$Original.Name), "SPRET male",
         ifelse(grepl("_SPI_m", dframe$Original.Name), "SPIC male",
-        ifelse(grepl("_SPI_f", dframe$Original.Name), "SPIC female",          
+        ifelse(grepl("_SPI_f", dframe$Original.Name), "SPIC female",
+       ifelse(grepl("_SPIC_m", dframe$Original.Name), "SPIC male",
+      ifelse(grepl("_SPIC_f", dframe$Original.Name), "SPIC female",        
+               
         ifelse(grepl("_CAROLI_m", dframe$Original.Name), "CAROLI male",
        ifelse(grepl("_CAROLI_f", dframe$Original.Name), "CAROLI female",
-                            "other")))))))))))))))))))))))))
+                            "other")))))))))))))))))))))))))))
        
   dframe$category<- factor(dframe$category,levels =c( "WSB female", "WSB male","G female", "G male", 
                          "LEW female", 'LEW male', "PERC male",

### 2018-02-15 1rmd_report.rmd
integrated batch8 female data (with corrections)
fixed some bugs for report file

@@
 author: "April Peterson"
-date: "Feb 6, 2018"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: github_document
@@
 # sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+#use the above line for when order is out of wak
 
@@
 ```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
-qq
 high_qual_boxplot
@@
 #remove sections which are in category but have no data
-df_list <- df_list[ -c(7, 12, 14,16,19, 21) ]
-#7: PERC male,   12: #KAZ female! 14:CAST female 16:HMI female,19:SPIC male, 21:SPIC male
+df_list <- df_list[ -c(7, 12,14,16,22) ]
+#7: PERC male, 12: KAZ female! (14:CAST female, fake data) 16:HMI female
 

### 2018-02-15 setup_data.R

@@ -15,6 +15,8 @@ library(raster)#for cV
 ########################
 
 #TODO: remove observations without quality scores.
+#12jan17_4jan17_LEW_f1_sp1_34_rev, 8feb17_4jan17_LEW_f1_sp1_3_rev, potential outliers
+
 #Write a make file that will merge all of the batch files and record the batch
 setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 MLH1_data = read.csv("data/AnonData.csv", header=TRUE )
@@ -78,15 +80,50 @@ MLH1_data <- MLH1_data[ !grepl("12sep16_MSM_f1", MLH1_data$mouse) , ]
 AP_mouse_table <- ddply(MLH1_data, c("mouse"), summarise,
                          Nmice = length(unique(mouse)),
                          Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                          cV = cv(adj_nMLH1.foci),
                           var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                          sd   = round(sd(adj_nMLH1.foci), 3),
                          se   = round(sd / sqrt(Ncells), 3)
 )
 
+#add strain, sex ect #function won't work
+# format so that mean_co is numer
+AP_mouse_table$subsp <-  ifelse(grepl("_WSB_", AP_mouse_table$mouse), "Dom", 
+                               ifelse(grepl("_G_", AP_mouse_table$mouse), "Dom",
+                                ifelse(grepl("_LEW", AP_mouse_table$mouse), "Dom",
+                                  ifelse(grepl("PWD", AP_mouse_table$mouse), "Musc",
+                               ifelse(grepl("MSM", AP_mouse_table$mouse), "Musc",  
+                              ifelse(grepl("KAZ", AP_mouse_table$mouse), "Musc",        
+                                      
+                                      ifelse(grepl("CAST", AP_mouse_table$mouse), "Cast",         
+                              ifelse(grepl("HMI", AP_mouse_table$mouse), "Cast",    
+                               ifelse(grepl("SPI", AP_mouse_table$mouse), "outgroup",         
+                                ifelse(grepl("CAROLI", AP_mouse_table$mouse), "outgroup",   
[…]

### 2018-02-26 draft_scripts/PnD.Rmd
cleaning up repo

---
title: "Poly Divr Simulations"
author: "April Peterson"
date: "November 13, 2017"
output: html_document
---

Rmd file for walking through the logic of how I'm setting up my Polymorphism and Divergence simulations. Map out inference plan.

```{r, setup}
library(plyr)
library(dplyr)
library(ggplot2)

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")

PnD_female <- MLH1_data[MLH1_data$sex == "female",]
PnD_male <- MLH1_data[MLH1_data$sex == "male",]

Dom_f <- PnD_female[( PnD_female$strain == "WSB" | PnD_female$strain == "G" |PnD_female$strain == "LEWES"),]
Musc_f <- PnD_female[( PnD_female$strain == "PWD" | PnD_female$strain == "MSM"),]

Dom_m <- PnD_male[( PnD_male$strain == "WSB" | PnD_male$strain == "G" |PnD_male$strain == "LEWES"),]
Musc_m <- PnD_male[( PnD_male$strain == "PWD" | PnD_male$strain == "MSM"),]
Cast_m <- PnD_male[( PnD_male$strain == "CAST" | PnD_male$strain == "HMI"),]

```

Why do these P and D simulations?
To test if the ratios for females and males are different (than a random sample).

Polymorphism measures are SE.
```{r, Polymorphism Measures}
F_poly_table <- ddply(PnD_female, c("strain"), summarise,
                     Nmice = length(unique(mouse)),
                     Ncells  = length(adj_nMLH1.foci),
                     mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                     var = format(round(   var(nMLH1.foci),3), nsmall=3),
                     sd   = round(sd(adj_nMLH1.foci), 3),
                     se   = round(sd / sqrt(Ncells), 3),
                     cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
DomF_poly_var <- var(F_poly_table$mean_co[1:3])
DomF_poly_sd <- sd(F_poly_table$mean_co[1:3])
DomF_poly_se <- (DomF_poly_sd / sqrt(3) )

MuscF_poly_var <- var(F_poly_table$mean_co[2:5])
MuscF_poly_sd <- sd(F_poly_table$mean_co[2:5])
MuscF_poly_se <- (MuscF_poly_sd / sqrt(2))

M_poly_table <- ddply(PnD_male, c("strain"), summarise,
                      Nmice = length(unique(mouse)),
                      Ncells  = length(adj_nMLH1.foci),
                      mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                      var = format(round(   var(nMLH1.foci),3), nsmall=3),
                      sd   = round(sd(adj_nMLH1.foci), 3),
                      se   = round(sd / sqrt(Ncells), 3),
                      cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)

DomM_poly_var <- var(M_poly_table$mean_co[1:3])
DomM_poly_sd <- sd(M_poly_table$mean_co[1:3])
DomM_poly_se <- (DomM_poly_sd / sqrt(3))

MuscM_poly_var <- var(M_poly_table$mean_co[2:5])
MuscM_poly_sd <- sd(M_poly_table$mean_co[2:5])
MuscM_poly_se <- (MuscM_poly_sd / sqrt(2) )

CastM_poly_var <- var(M_poly_table$mean_co[6:7])
CastM_poly_sd <- sd(M_poly_table$mean_co[6:7])
CastM_poly_se <- (CastM_poly_sd / sqrt(2) )
```
Above are the Polymorphism measures.

Divergence measures are paired, there will be mulitple divergent measures per strain
```{r, Divergence Measures}
#calq subsp wide averages
DomF_D_mean <- mean(as.numeric(F_poly_table$mean_co[1:3]) )
MuscF_D_mean <- mean(as.numeric(F_poly_table$mean_co[2:5]) )

DomM_D_mean <- mean(as.numeric(M_poly_table$mean_co[1:3]) )
MuscM_D_mean <- mean(as.numeric(M_poly_table$mean_co[2:5]) )
CastM_D_mean <- mean(as.numeric(M_poly_table$mean_co[6:7]) )

# divergence can only be calq in pairwise manner
# Musc - Dom, Musc - spretus(spic), Musc - caroli,
F_D_HMdm_se <- ( sd(c(DomF_D_mean, MuscF_D_mean)) / sqrt(2) )

M_D_HMdm_se <- ( sd(c(DomM_D_mean, MuscM_D_mean) ) / sqrt(2) )
M_D_HMdc_se <- ( sd(c(DomM_D_mean, CastM_D_mean) ) / sqrt(2) )
M_D_HMmc_se <- ( sd(c(CastM_D_mean, MuscM_D_mean) ) / sqrt(2) )

```

Now start simulations for making random samples from the database.
start sampling througheach type of mouse
```{r, simulations}

#3 calqs of SE from random samples
n3 <-  replicate(3, (sd(sample(Dom_f$nMLH1.foci, 30) ) / sqrt(30) ) )
#se of the simulated se's       
se_per <- sd( replicate(3, (sd(sample(Dom_f$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)

#100 simulated se's / P values for fDom
per100se <- replicate(1000,
                      sd( replicate(3, (sd(sample(Dom_f$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)
)
#this will show a hist of the simulated P values, with R line for the real
hist(per100se, xlim = range(0,.5) )
abline(v=DomM_poly_se,col="red")

```

#### Comparisons for simulations.

Comparing the histograms of simulated Polymorphism values by the size,

```{r, poly simulations}
#3 calqs of SE from random samples
male_musc <-  replicate(3, (sd(sample(Musc_m$nMLH1.foci, 30) ) / sqrt(30) ) )
#se of the simulated se's       
se_per <- sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)

#100 simulated se's / P values for fDom
male_musc_ss30 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 30) ) / sqrt(30) ) ) ) / sqrt(3)
)
male_musc_ss15 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 15) ) / sqrt(15) ) ) ) / sqrt(3)
)
male_musc_ss10 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 10) ) / sqrt(10) ) ) ) / sqrt(3)
)

#this will show a hist of the simulated P values, with R line for the real

par(mfrow=c(1, 3))
hist(male_musc_ss30, xlim = range(0,3) )
hist(male_musc_ss15, xlim = range(0,3) )
hist(male_musc_ss10, xlim = range(0,3) )
abline(v=MuscM_poly_se,col="red")

Dom_m <- PnD_male[( PnD_male$strain == "WSB" | PnD_male$strain == "G" |PnD_male$strain == "LEWES"),]
Musc_m <- PnD_male[( PnD_male$strain == "PWD" | PnD_male$strain == "MSM"),]
Cast_m <- PnD_male[( PnD_male$strain == "CAST" | PnD_male$strain == "HMI"),]
```

Compare the effects of drawing MLH1 from specific strains, instead of pooled values.

```{r}
# sample from strain
male_musc_ss10 <- replicate(1000, c(
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "PWD"], 10),
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "MSM"], 10) ) 
)

hist(male_pwd_ss10, xlim = range(15,40) )

male_dom_ss10 <- replicate(1000, c(
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "G"], 10),
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "LEWES"], 10),
  sample( PnD_male$nMLH1.foci[PnD_male$strain == "WSB"], 10)
  ) 
)

#b <- hist(male_dom_ss10, xlim = range(10,40) )
#b <- abline( v=mean( PnD_male$nMLH1.foci[(PnD_male$strain == "WSB")] #), col="red")

 # mean(PnD_male$nMLH1.foci[(PnD_male$strain == "G")] ),
#  mean(PnD_male$nMLH1.foci[(PnD_male$strain == "LEWES")] ) ), col="red")
#PnD_male[( PnD_male$strain == "WSB" | PnD_male$strain == "G" |PnD_male$strain == "LEWES")

male_msm_ss10 <- sample( PnD_male$nMLH1.foci[PnD_male$strain == "MSM"], 10)

male_musc_ss15 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 15) ) / sqrt(15) ) ) ) / sqrt(3)
)
male_musc_ss10 <- replicate(1000,
        sd( replicate(3, (sd(sample(Musc_m$nMLH1.foci, 10) ) / sqrt(10) ) ) ) / sqrt(3)
)

```
```{r, divergence simulations}

#simulating divergence --- take the 3 strain sampes from the polymorphism sims
# calq se / var from a constant measure of divergence

#take a list of the polymorphis measures -- pair cbind/rbind with simulations for D, plot the ratios

```

### 2018-02-26 draft_scripts/pass-_report.Rmd
cleaning up repo

---
title: "passing mice script"
author: "April Peterson"
date: "February 8, 2018"
output: html_document
---

Guide / report document for showing which mice have passed - sufficient cells samples for showing good estimate of MLH1 count.

Outline
-dicription of graphs / data
-discription of simulations (sampling comparison)
(example images)

<code for making the graphs
-add mouse name, total number of cells,

- develope thresholds for grouping mice into passed and more images
- consider doing this by batch

- based on current plot outputs,

- think of ways to combine this sampling size data with quality distributions
```{r setup, include=FALSE}

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)

```

```{r group_mice}
#seperate large dataframe by how many cells are in the samples
twenty <- AP_mouse_table[AP_mouse_table$Ncells >= 20,] #53
#are these assumed to have passed?

lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,] #28
#these will give the list of mice

#subset MLH1_data by these lists. (fixed line below, )
twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]
lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]

```

```{r pvalue_permutations, echo=FALSE}

```

### 2018-02-26 SamplingEffects.R

@@ -72,6 +72,12 @@ Twenty_pvals_20_1k <- sample_cells_pvals(twenty, twenty_data, 20, 1000)#this ret
 
 #make all these outlist of pvalues...
 
+#
+# figure out how to calq different of mean pvalue between sample sizes!!!
+#
+
+
+
 #think of saving these DF, as they take time 
 save.image("SampleSizePvalues.RData")
 
@@ -117,6 +123,9 @@ for (i in 1:length(twenty$mouse)){
   plot_list20[[i]] <- p20
 }
 
+#after that code, there are 3 lists of ggplots histograms (by sample size)
+
+
 
 #make list of scatter plots
 sctr_plot = list()
@@ -141,6 +150,8 @@ for (i in 1:length(twenty$mouse)){
     geom_jitter(aes(x=DFsets, y=pvals) )
   sctr_plot[[i]] <- scplot
 }
+
+
 
 #use Function Multiplot to display multiple ggplot objects
 source("src/MultiPlot.R")

### 2018-02-26 draft_scripts/PI FST ellipse 4 plots.R

@@ -1,0 +1,281 @@
+setwd("/Users/Doug/Desktop")
+bp=read.table("jerez.txt")
+
+bp$Pi=bp$pi
+
+setwd("/Users/Doug/Desktop/Filtered_Tcas_Files")
+bpPi=read.table("Bhopal2_filtered.genes.pi")
+
+bpFST=read.table("BJPL2_filt_genewise.fst")
+
+bp=merge(bpPi,bpFST,by="V1")
+library(plyr)
+bp=rename(bp, c("V1"="gene_name","V4.x"="Pi", "V7"="FST","V9"="FST.3","V11"="FST.4"))
+
+
+#Sperm
+
+bpIns <- subset(bp, 
+					gene_name=="TCOGS2:TC013049-RA" | 
+					gene_name=="TCOGS2:TC012560-RA" |
+					gene_name=="TCOGS2:TC011481-RA" | 
+					gene_name=="TCOGS2:TC010065-RA" | 
+					gene_name=="TCOGS2:TC010064-RA" |
+					gene_name=="TCOGS2:TC009459-RA" | 
+					gene_name=="TCOGS2:TC005744-RA" |
+					gene_name=="TCOGS2:TC003552-RA"	|
+					gene_name=="TCOGS2:TC015849-RA" | 
+					gene_name=="TCOGS2:TC015056-RA" | 
+					gene_name=="TCOGS2:TC010066-RA" |
+					gene_name=="TCOGS2:TC008976-RA" | 
+					gene_name=="TCOGS2:TC006088-RA" 
+					, select = c(gene_name,Pi,FST,FST.3,FST.4))       #CHAGE FST
+					
+										
+
+bpInsFST=((bpIns$FST+bpIns$FST.3+bpIns$FST.4)/3)				      #CHAGE FST
+
+bpFST=((bp$FST+bp$FST.3+bp$FST.4)/3)                                  #CHAGE FST
+
+
+bpmean=mean(bp$Pi,na.rm = TRUE)
+bpmeanFST=mean(bpFST)
+
+
+nreps <- 10000
+PI <- numeric(nreps) 
+FST <- numeric(nreps)  
+
+PI[1] <- bpmean
+FST[1] <- bpmeanFST
+
+
+for (i in 2:nreps) {
+  bpR=bp[sample(nrow(bp), length(bpIns[,1])), ]										
+  bpRmean=mean(bpR$Pi, na.rm = TRUE)
[…]

### 2018-02-26 draft_scripts/PI FST football with z-test.R

@@ -1,0 +1,158 @@
+setwd("/Users/Doug/Desktop")
+bp=read.table("jerez.txt")
+
+bp$Pi=bp$pi
+
+
+bpPi=read.table("Jerez2_filtered.genes.pi")
+
+bpFST=read.table("BJPL2_filt_genewise.fst")
+
+bp=merge(bpPi,bpFST,by="V1")
+library(plyr)
+bp=rename(bp, c("V1"="gene_name","V4.x"="Pi", "V7"="FST","V13"="FST.3","V15"="FST.4"))
+
+
+#Pesticide 
+
+bpIns <- subset(bp, 
+					gene_name=="TCOGS2:TC016365-RA" | 
+					gene_name=="TCOGS2:TC015102-RA" |
+					gene_name=="TCOGS2:TC014402-RA" | 
+					gene_name=="TCOGS2:TC013633-RA" | 
+					gene_name=="TCOGS2:TC012852-RA" |
+					gene_name=="TCOGS2:TC012627-RA" | 
+					gene_name=="TCOGS2:TC012503-RA" |
+					gene_name=="TCOGS2:TC012112-RA"	|
+					gene_name=="TCOGS2:TC009170-RA" | 
+					gene_name=="TCOGS2:TC008198-RA" | 
+					gene_name=="TCOGS2:TC007155-RA" |
+					gene_name=="TCOGS2:TC007061-RA" | 
+					gene_name=="TCOGS2:TC006445-RA" |
+					gene_name=="TCOGS2:TC005352-RA"	|
+					gene_name=="TCOGS2:TC004838-RA" | 
+					gene_name=="TCOGS2:TC004836-RA" | 
+					gene_name=="TCOGS2:TC004749-RA" |
+					gene_name=="TCOGS2:TC004383-RA" | 
+					gene_name=="TCOGS2:TC004126-RA" |
+					gene_name=="TCOGS2:TC003369-RA" |
+					gene_name=="TCOGS2:TC003239-RA"	|
+					gene_name=="TCOGS2:TC001018-RA" | 
+					gene_name=="TCOGS2:TC000981-RA" | 
+					gene_name=="TCOGS2:TC000394-RA" 
+					, select = c(gene_name,Pi,FST,FST.3,FST.4))       #CHAGE FST
+
+
+#Sperm
+
+bpIns <- subset(bp, 
+					gene_name=="TCOGS2:TC013049-RA" | 
+					gene_name=="TCOGS2:TC012560-RA" |
[…]

### 2018-02-26 draft_scripts/distributions.R

@@ -1,0 +1,49 @@
+# Distributions
+# make plots that show the distributions of different categories
+#
+
+library(plyr)
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="MLH1_data_setup.RData")
+
+
+# subsp
+subsp_BP <- ggplot(MLH1_data, aes(x = as.factor(subsp), y = adj_nMLH1.foci, 
+    fill= strain )) + geom_boxplot()  
++ ggtitle("Boxplots of MLH1 distributions by mouse") +
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "yellowgreen", "green", "grey" ))
+subsp_BP <- subsp_BP   + facet_wrap(~ sex, scales="free")
+subsp_BP
+
+
+subsp_Den <- ggplot(MLH1_data, aes(x = adj_nMLH1.foci, fill=sex)) + 
+  geom_density(alpha=.6)  +
+  ggtitle("")
+subsp_Den <- subsp_Den   + facet_wrap(~ subsp, scales="free")
+subsp_Den
+
+subsp_Hist <- ggplot(MLH1_data, aes(x = adj_nMLH1.foci, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("")
+subsp_Hist <- subsp_Hist   + facet_wrap(~ subsp, scales="free")
+subsp_Hist
+
+
+
+#strain
+strain_Den <- ggplot(MLH1_data, aes(x = adj_nMLH1.foci, fill=sex)) + 
+  geom_density(alpha=.6)  +
+  ggtitle("")
+strain_Den <- strain_Den   + facet_wrap(~ strain, scales="free")
+strain_Den
+
+strain_Hist <- ggplot(MLH1_data, aes(x = adj_nMLH1.foci, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("")
+strain_Hist <- strain_Hist   + facet_wrap(~ strain, scales="free")
+strain_Hist

### 2018-02-26 draft_scripts/glm.R

@@ -1,0 +1,51 @@
+# GLM
+# 
+#
+
+library(plyr)
+library(dplyr)
+library(ggplot2)
+library(lme4)
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="MLH1_data_setup.RData")
+
+
+#nested
+
+
+
+#models of MLH1 counts
+
+nosubsplmm <- lmer(nMLH1.foci ~ sex + (1 | strain), data = MLH1_data, REML = FALSE)
+nostrainlmm <- lmer(nMLH1.foci ~ sex + (1 | subsp), data = MLH1_data, REML = FALSE)
+nosexlmm <- lmer(nMLH1.foci ~ strain + (1 | subsp), data = MLH1_data, REML = FALSE)
+nofixedlmm <- lmer(nMLH1.foci ~ 1 + (1 | subsp), data = MLH1_data, REML = FALSE)
+anova(nostrainlmm, nosexlmm, nofixedlmm, nosubsplmm)
+#compare lowest AIC, no sex
+
+#while nested random effects take the form (1 | r1 / r2).
+
+#sex fixed, other's random
+model.cell.level = lme(nMLH1.foci ~ sex, random=list(~1|subsp, ~1|strain, ~1|mouse),
+            data=MLH1_data, 
+            method="REML")
+
+model_fxd = lm(nMLH1.foci ~ sex+subsp+strain, random = ~1|mouse,
+               data=MLH1_data)
+
+anova(model_fxd)
+summary(model_fxd)
+varcomp(model_fxd)#only works for lme
+
+
+# models of mouse averages
+
+model.mouse.level = lm(mean_co ~ sex, random=list(~1|subsp, ~1|strain),
+                       data=AP_mouse_table, 
+                       method="REML")
+
+model.mouse.level = lm(mean_co ~ sex, random=list(~1|subsp, ~1|strain),
+                       data=AP_mouse_table)
+
+model.mouse.nested = lmer(mean_co ~ subsp + sex + strain +(subsp|strain),
+                       data=AP_mouse_table)

### 2018-02-26 draft_scripts/modeling.R

@@ -1,0 +1,302 @@
+# protocol for making a Linear model from MLH1 count data
+#
+#
+
+library(plyr)
+library(dplyr)
+library(ggplot2)
+library(lme4)
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+
+#step 1. load data
+load(file="MLH1_data_setup.RData")
+
+#remove non-house-mouse for this
+MLH1_data <- MLH1_data[ !grepl("SPRET", MLH1_data$strain) , ]
+MLH1_data <- MLH1_data[ !grepl("SPIC", MLH1_data$strain) , ]
+
+# potentially removed order of factors in dataframe
+MLH1_data$strain <- factor(MLH1_data$strain)
+MLH1_data$subsp <- factor(MLH1_data$subsp)
+
+# model the relationship of MLH1 count with sex, subsp, strain and mouse
+# 
+
+#make boxplots of all variables, how variable are the means
+(subsp <- ggplot(MLH1_data, aes(subsp, nMLH1.foci)) +
+    geom_boxplot(fill = "#CD3333", alpha = 0.8, colour = "#8B2323") +
+    theme.clean() +  
+    theme(axis.text.x = element_text(size = 12, angle = 0)) +
+    labs(x = "x", y = "MLH1.counts"))
+
+(sex <- ggplot(MLH1_data, aes(sex, nMLH1.foci)) +
+    geom_boxplot(fill = "#CD3333", alpha = 0.8, colour = "#8B2323") +
+    theme.clean() +  
+    theme(axis.text.x = element_text(size = 12, angle = 0)) +
+    labs(x = "x", y = "MLH1.counts"))
+
+(strain <- ggplot(MLH1_data, aes(strain, nMLH1.foci)) +
+    geom_boxplot(fill = "#CD3333", alpha = 0.8, colour = "#8B2323") +
+    theme.clean() +  
+    theme(axis.text.x = element_text(size = 12, angle = 0)) +
+    labs(x = "x", y = "y"))
+
+(mouse <- ggplot(MLH1_data, aes(mouse, nMLH1.foci)) +
+    geom_boxplot(fill = "#CD3333", alpha = 0.8, colour = "#8B2323") +
+    theme.clean() +  
+    theme(axis.text.x = element_text(size = 1, angle = 0)) +
+    labs(x = "x", y = "y"))
+
+#create single factor models
[…]

### 2018-02-26 draft_scripts/nested_mixed.R

@@ -1,0 +1,147 @@
+# Create a mixed model with nested factors using lme or lmer
+# input: mouse average table
+# output: aov of mixed model 
+
+#load data
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="nested_environment.RData")
+
+#adjust data
+
+#remove non house mouse 
+
+AP_mouse_table <- AP_mouse_table[ !grepl("outgroup", AP_mouse_table$subsp) , ]
+#remove cast (no female data)
+AP_mouse_table <- AP_mouse_table[ !grepl("Cast", AP_mouse_table$subsp) , ]
+#this seemed to work...
+
+#factors
+AP_mouse_table$sex <- as.factor(AP_mouse_table$sex)
+AP_mouse_table$sex <- relevel(AP_mouse_table$sex, "male")
+
+AP_mouse_table$subsp <- as.factor(AP_mouse_table$subsp)
+AP_mouse_table$mouse <- as.factor(AP_mouse_table$mouse)
+AP_mouse_table$strain <- as.factor(AP_mouse_table$strain)
+
+attach(AP_mouse_table)
+#lme
+library(nlme)
+
+# this needs the random parameter
+aa <- lme(mean_co ~ sex * subsp, data=AP_mouse_table)
+#Error in getGroups.data.frame(dataMix, groups) : 
+#invalid formula for groups
+
+#once 'Cast' observations removed, lme models worked
+
+bb <- lme(mean_co ~ sex * subsp, data=AP_mouse_table, 
+          random = ~sex|strain )
+
+cc <- lme(mean_co ~ sex * subsp, data=AP_mouse_table, 
+          random = list( ~1|strain, ~0+sex|strain ) )
+
+dd <- lme(mean_co ~ sex * subsp, data=AP_mouse_table, 
+          random = list( strain = pdDiag(~sex) )
+          )
+
+#gg <- lme(cV ~ sex * subsp, data=AP_mouse_table, 
+#          random = list( strain = pdDiag(~sex) )
+#)
+
+zz <- lme(mean_co ~ sex*subsp, data=AP_mouse_table,
+          random = ~1|strain) #removes strain specific sex effect
+
+
+#subset data
+Fmice_data <- subset(AP_mouse_table, sex=="female")
+Mmice_data <- subset(AP_mouse_table, sex == "male")
+
+#female models
+fcc <- lme(mean_co ~ subsp, data=Fmice_data, 
+           random = list( ~1|strain) )
[…]

### 2018-02-26 src/MLH1_strain_means_figure.R

@@ -11,7 +11,6 @@ library(grid)
 library(gridBase)
 library(asbio)
 library(plyr)
-library(lessR)
 
 setwd("C:/Users/alpeterson7/Documents/MLH1repo")
 load(file="MLH1_data_setup.RData")
@@ -69,3 +68,4 @@ PWDm.MLH1.se = sd(MLH1_data[MLH1_data$category == 'PWD male', ]$adj_nMLH1.foci) 
 PWDf.MLH1.mean = mean(MLH1_data[MLH1_data$category == 'PWD female', ]$adj_nMLH1.foci)
 PWDf.MLH1.se = sd(MLH1_data[MLH1_data$category == 'PWD female', ]$adj_nMLH1.foci) / (sqrt(length(MLH1_data[MLH1_data$category == 'PWD female', ])))
 
+dev.off()

### 2018-03-09 Main_Report.rmd
renamed summary to Main_report
small edits to main report

--- 2018-02-15 1rmd_report.rmd
@@
 ---
-title: "MLH1 data report "
+title: 'MLH1 Data Report'
 author: "April Peterson"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: github_document
+date: "some date"
+output: rmarkdown::github_document
 ---
@@
 
-#### Research Discription
-Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice (rodents).
-Comparisons the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
-(insert small picture of meixyte)
+#### Research Description
+Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice abd related rodents.
+Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
 
-#### Discription of Report
+#### Description of Report
 
@@
 
-Aquired microscope cell images are quantified in batches after being anonimzed.
+Cell images are quantified in batches after file names are anonymized.
 The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified.
@@
 #MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
+
+#MLH1_data <- MLH1_data %>%
+#  arrange(category) %>%
+#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+
+#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
+
 ```
-
-#### Mice table
-
-Table of the number of mice used and MLH1 stats. (made with kable).
-
-```{r, fig.height=2, fig.width=4, echo=FALSE}
-#this table is saved in the .RData file
-#use kable to make tables. warning this might not work for github Rmd.
-n <- kable(AP_mouse_table)
-n
-```
-
-MLH1_data <- MLH1_data %>%
-  arrange(category) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
-MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
 
@@
 
+#### Mice table
+
+Table of the mice used and MLH1 stats.
+
+```{r, fig.height=2, fig.width=4, echo=FALSE}
+#this table is saved in the .RData file
+#use kable to make tables. warning this might not work for github Rmd.
+n <- kable(AP_mouse_table)
+n
+```
+
 #### Power Calculations

### 2018-03-09 draft_scripts/dup_in_batches.R
started script to check for duplicates

@@ -1,0 +1,3 @@
+#
+#
+# Write a function or in setup, compare length(unique(hh$Original.Name)), length(unique(hh$Random.Name)), this will find duplicate data

### 2018-03-19 Main_Report.rmd
updated makeFile and added batch10 female

@@
 #remove sections which are in category but have no data
-df_list <- df_list[ -c(7, 12,14,16,22) ]
+df_list <- df_list[ -c(7,14,16,22) ]
 #7: PERC male, 12: KAZ female! (14:CAST female, fake data) 16:HMI female

### 2018-03-19 src/HetC_plot.R
fixed HetPlot file

@@ -30,8 +30,8 @@ AP_strain_table <- ddply(MLH1_data, c("strain", "sex"), summarise,
                          cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
 )
 
-#remove single cast female
-AP_strain_table <- AP_strain_table[-c(12),]
+#remove single cast female, and HMI male
+AP_strain_table <- AP_strain_table[-c(13,15),] #there is a line below this which accounts for this
 
 #add Lynn Cast female data
 Lynn_CASTf_foci = c(20,21, 23, 25, 26, 26,26,27.5, 28, 28,28,33)
@@ -49,7 +49,6 @@ AP_strain_table$var <- as.numeric( AP_strain_table$var)
 AP_strain_table$sd <- as.numeric( AP_strain_table$sd)
 AP_strain_table$se <- as.numeric(  AP_strain_table$se)
 AP_strain_table$cV <- as.numeric(   AP_strain_table$cV)
-
 
 #make casted rows  --- why am I making cast rows?
 casted_co <- dcast(data = AP_strain_table, formula= strain~sex, value.var="mean_co")
@@ -96,7 +95,7 @@ plotCI(x = as.numeric(HetC_table[,3]),   #male mean_co
        
 #change the colors       
        col = c("#56B4E9","lightblue4", "cadetblue", #blues, W, Lew, G,
-               "coral1", "#E69F00", "yellowgreen", "black", "purple"),        #reds   "indianred",
+               "coral1", "#E69F00", "red", "yellowgreen", "black", "purple"),        #reds   "indianred",
        #   "seagreen3"           #green
        pch = 16, lwd = 3.5, gap = 0, sfrac = 0.01, add = TRUE, cex=1.4)
 
@@ -123,18 +122,23 @@ segments(  (as.numeric(HetC_table[4,3]) - as.numeric(HetC_table[4,9])*2 ), as.nu
 segments(  (as.numeric(HetC_table[5,3]) - as.numeric(HetC_table[5,9])*2 ), as.numeric(HetC_table[5,2]), 
            ( as.numeric(HetC_table[5,3]) + as.numeric(HetC_table[5,9])*2 ), as.numeric(HetC_table[5,2]),
            lwd=3.5, col="#E69F00")
-#cast
+#KAZ
[…]

### 2018-04-03 Main_Report.rmd
edits to report

@@
 
+#try to make sure sacles are the same
+
 #reduncant
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey" ))
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey" )) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5))
 ff <- ff   + facet_wrap(~ sex, scales="free")
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey" ))
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey" )) +
+  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )
+
 ss <- ss   + facet_wrap(~ sex, scales="free")

### 2018-04-03 draft_scripts/MetaDataEffectReport.Rmd
updating  draft scripts

---
title: "MetaEffectReport"
author: "April Peterson"
date: "March 9, 2018"
output: rmarkdown::github_document
---

```{r load data, include=FALSE}
#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")

#load()metadata
meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1data/data/mouseDOBs.csv")

library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(png)
library(reshape2)
```

### Report for Examining Meta Effects

- age
- mat age
- G line
- female age
- litter

### files

- MLH1 data
- MetaData file

ToDo
- make sure metadata file is as complete as possible
- load and combine data

```{r dates, echo=FALSE}

#extract dates from file names

for( t in 1:length(meta_data$mouse)){
  euth_date = strsplit(meta_data$mouse[t], split="_")[[1]][1]
  meta_data$raw_euth_date[t] <- strsplit(meta_data$mouse[t], split="_")[[1]][1]
  fomt_euth <- as.Date(strsplit(meta_data$mouse[t], split="_")[[1]][1], format= '%d%b%y')
  fomt_euth_nrm <- as.Date(fomt_euth, "%Y-%m-%d")
  vv <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="weeks" )#whoo, this works
  hh <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="hours" )
  meta_data$age_weeks[t] <- vv
  meta_data$age_hours[t] <- hh

  #mat age, DOB -Mat_dob
  mat_dob <- as.Date(meta_data$maternal.age..DOB.[t], format= '%m/%d/%Y')
#  hh <- as.numeric(difftime(fomt_euth, as.Date(meta_data$DOB[t], '%m/%d/%Y')), units="hours" )
  gg <- as.numeric(difftime(as.Date(meta_data$DOB[t], '%m/%d/%Y'),mat_dob), units="weeks" )
  meta_data$mat_age_wk[t] <- gg
}

as.Date(first part of mouse,format='%d%b%y' )
euth_date <- as.Date(p[[1]][1],format='%d%b%y' )
as.Date('22JUN01',format='%d%b%y') ##this will read the mouse date format correctly.
#calculate age from dates
difftime()

```

```{r}
#merge data (MLH1 counts and ages)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
load("MLH1_data_setup.RData")
#add euth date and calc age
mergedDF = merge(MLH1_data, meta_data, by = "mouse") #4000 cells by table of 1178 mice

dissected_mice <- unique(meta_data$mouse)#180
imaged_mice <- unique(MLH1_data$mouse)#62
n.overlapping_mice <- length(imaged_mice)/length(dissected_mice)

dfn <- mergedDF[(mergedDF$DOB == '' ),]
dfm <- mergedDF[(is.na(mergedDF$DOB) ), ]

unique(dfn$mouse)

unique((dfm$mouse))

```

### 2018-04-03 draft_scripts/MixedModel_PnD_Report.Rmd
updating  draft scripts

---
title: "SimulatedData_MixedModel"
author: "April Peterson"
date: "March 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(nlme)
library(plyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(raster)#for cV

#load data
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")
```

#### Mixed models

explain mixed model best I can

-- the main point of comparing these mixed models is to --- contrast female and male patterns of variation in MLH1 number--

<previous results show that males had more variaiton at the subspecies level 
< mouse level MLH1

-- redo the previous models -- compare the subsets with high at low quality --

Coding for the mixed model; base Dom male.  (in most of my dataframes female is ordered before males, for visualization purposes)

Displaying the comparisons of different models -- create a table with the ww$values -- so they can be compared..

values
-parameter estimates
-AICs ect

conditions
-quality scores
-batches
-simulated data? (randize strain)

-by sex

```{r, include=FALSE}
#make model

# just look at data of Musc and Dom 
mixmodmouse_data <- AP_mouse_table[(AP_mouse_table$subsp == "Musc" | AP_mouse_table$subsp == "Dom"),  ]

mixmodmouse_data$sex <- factor(mixmodmouse_data$sex, levels = c("male", "female"))

model1 <- lme(mean_co ~ sex * subsp, data=mixmodmouse_data, random=list(strain=pdDiag(~sex)) )

summary(model1)
ranef(model1)
```

To interpert the summary results, (Fixed effects)
- the intercept is the mean MLH1 number of the base factors -- which in this model is defined as Dom male (26.40).

- the next variable is sex being female, has an affect of increasing the mean MLH1 count by 1.7. (This is about the amount of sexual dimorphism)
1.7 (SD)

- the value for Musc is 3.935, meaning being male and Musc adds 3.935 to the mean MLH1 (in the previous model without KAZ, this was 5.5)

- the interaction of sex and subspecies is -3.7 meaning, musc females are 3 MLH1 foci lower than musc males (in previous models this was -5.1)

Random effects (strain and strain specific sexual dimorphism)
- the intercept is the average sd from strains 1.89, 
- the sex effec is the average sd from strain specific sexual dimorphism 2.07
- the random effects can be estimated; the intercepts are the average sd for the strain effect, ... from these esitmates is seems that G and Kaz have strain specific sexual dimorphism where females are larger
... if the strain specific SD effect is larger than the whole sex effect

```{r}
#make new AP mouse table with cells from 1-3
quality_data <- MLH1_data[MLH1_data$quality <= 3, ]
quality_data <- quality_data[(quality_data$subsp == "Musc" | quality_data$subsp == "Dom"),  ]

na.omit(quality_data)

qualAP_mouse_table <- ddply(quality_data, c("subsp", "sex", "strain", "mouse"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                         cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
                        #sex = sex,
                        #subsp = subsp,
                        #strain = strain
                      #quality?

)
qualAP_mouse_table <- na.omit(qualAP_mouse_table)

qualAP_mouse_table <- qualAP_mouse_table[qualAP_mouse_table$Ncells > 5,]

str(qualAP_mouse_table)

qualAP_mouse_table$sex <- factor(qualAP_mouse_table$sex, levels = c("male", "female"))
qualAP_mouse_table$strain <- factor(qualAP_mouse_table$strain)
qualAP_mouse_table$subsp <-  factor(qualAP_mouse_table$subsp, levels=c("Dom", "Musc"))

#model 2 -- asses if there are affects of quality
model2 <- lme(mean_co ~ sex * subsp, data=qualAP_mouse_table, random=list(strain=pdDiag(~sex)) )

summary(model2)
ranef(model2)

```

For this version of cleaned data, the effect of Musc on the mean MLH1 is lower.

### 2018-04-03 SamplingEffects.R

@@ -52,9 +52,9 @@ sim_iterations <- 10000
 
 #list, DF, ncells, sim_num
 
+#discription of function...
 source("src/Func_sampleCells.R")
 lessTwenty_pvals_5_10 <- sample_cells_pvals(lessTwenty, lessTwenty_data, 5, 10)#this returns a list (of vectors?) pvalues  for each mouse in a list
-
 #nn[[4]] is the syntax to get just the pvalues
 #names(nn[4]) is now you get the name of the vector / list
 
@@ -64,22 +64,32 @@ lessTwenty_pvals_10_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 10, 10
 
 
 Twenty_pvals_10_1k <- sample_cells_pvals(twenty, twenty_data, 10, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
-
 Twenty_pvals_15_1k <- sample_cells_pvals(twenty, twenty_data, 15, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
-
 Twenty_pvals_20_1k <- sample_cells_pvals(twenty, twenty_data, 20, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
-
-
 #make all these outlist of pvalues...
-
 #
 # figure out how to calq different of mean pvalue between sample sizes!!!
 #
 
+test_pass = MLH1_data[MLH1_data$mouse == "30may17_MSM_m1",]
 
+recent.mice = list(mouse = c("30may17_MSM_m1", "5jul17_MSM_m1", "20dec16_LEW_m1",
+                   "20dec16_LEW_m2","4jan17_LEW_f6","8may17_LEW_m2","4jan17_LEW_f7",
+                   "4jan17_LEW_f3","4jan17_LEW_m1"  ))
+
+recent.mice.DF <- MLH1_data[MLH1_data$mouse %in% recent.mice$mouse,]
+
+r.m.table <- AP_mouse_table[AP_mouse_table$mouse %in% recent.mice[[1]],]
+#lowest cell number is 20
+source("src/Func_sampleCells.R")
+
+
+testing.these.pvals.10 <- sample_cells_pvals(recent.mice,recent.mice.DF,10,10000 )
+testing.these.pvals.15 <- sample_cells_pvals(recent.mice,recent.mice.DF,15,10000 )
+testing.these.pvals.18 <- sample_cells_pvals(recent.mice,recent.mice.DF,18,10000 )
 
[…]

### 2018-04-03 setup_data.R

@@ -82,9 +82,12 @@ AP_mouse_table <- ddply(MLH1_data, c("mouse"), summarise,
                          Ncells  = length(adj_nMLH1.foci),
                          mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                          cV = cv(adj_nMLH1.foci),
-                          var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                          sd   = round(sd(adj_nMLH1.foci), 3),
                          se   = round(sd / sqrt(Ncells), 3)
+                      #quality?
+                        
+                        
 )
 
 #add strain, sex ect #function won't work

### 2018-05-29 Main_Report.rmd
removed duplicate data, added stiching to makefile

@@
 author: "April Peterson"
-date: "some date"
+date: "29 may 2018"
 output: rmarkdown::github_document
@@
 ```{r setup, echo=FALSE, include = FALSE}
-#
 library(knitr)
@@
 ```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
+### to integrate 'pass'/fail
+### this should be in setup data
+
 #what is the purpose of this block
@@
   )
-# this loop is going though each row of dataframes within the df.list and assining the mouse mean
+# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
   for(o in 1:length(df_list[[i]]$Original.Name)){
@@
   pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") 
+  #theme for indiviual mice, 
+  #+ theme(plot.background = element_rect(fill = "green"))
   plot_list[[h]] =   pf
@@
 ```{r, fig.height=2, fig.width=4, echo=FALSE}
+#this table is 
 #this table is saved in the .RData file

### 2018-05-29 setup_data.R

@@ -44,6 +44,11 @@ MLH1_data <- MLH1_data[!(is.na(MLH1_data$nMLH1.foci) | MLH1_data$nMLH1.foci=="")
 MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "X",]
 MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]
 
+# add pass or fail column.. this will be a large list..maybe write a function..
+# highlight real passes? highlight non passes?
+# calculate automatically, list manually?
+
+
 source("src/Func_addCategory.R")
 MLH1_data <- add_category(MLH1_data)
 #set the order of categories (female, male) (cast, dom, musc)
@@ -124,6 +129,7 @@ AP_mouse_table$sex <-  ifelse(grepl("_f", AP_mouse_table$mouse), "female",
                        ifelse(grepl("_m", AP_mouse_table$mouse), "male",
                                                 "other"))
 
+AP_mouse_table$mouse <- as.factor(AP_mouse_table$mouse)
 AP_mouse_table$sex <- as.factor(AP_mouse_table$sex)
 AP_mouse_table$strain <- as.factor(AP_mouse_table$strain)
   AP_mouse_table$subsp <-  as.factor(AP_mouse_table$subsp)
@@ -142,6 +148,10 @@ BDMLH1_data <- subset(fullBD_MLH1_data, (Cross %in%  c("PANCEVO","RAT","CIM", "P
 
 #now that I have the mice of Beth's I want, remove the big BD df with F2s to save space.
 rm(fullBD_MLH1_data)
+
+
+
+
 
 
 ############

### 2018-05-30 Main_Report.rmd
passed mice are colored light red

@@
 
+#quant_status (this isn't saving)
+#for(i in 1:length(MLH1_data$mouse)){
+#  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+#}
+
 #what is the purpose of this block
@@
 i=0
+
+MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
 
@@
 for (h in 1:length(df_list)) {
+  pf <- ggplot(data = df_list[[h]], y= adj_nMLH1.foci, x= quality, aes(colour = factor(Batch)), width = 0.25)
+  # +   geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5)
+
+  + geom_jitter()
+  + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") + scale_fill_manual(values=c(NA, "red"), breaks=NULL)
+
+# +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+  #theme for indiviual mice, 
+  #+ theme(plot.background = element_rect(fill = "green"))
+
+  plot_list[[h]] =   pf
+}
+
+#orig
+plot_list = list()
+for (h in 1:length(df_list)) {
   pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
+  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
+
     pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
   pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") 
+  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
+
+    scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL)
+
     #theme for indiviual mice, 
@@
 }
+#can't change the alpha/transparency of the red rec
 

### 2018-06-11 Main_Report.rmd
knitted batch12 data, fixed some bugs

@@
 latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.
-```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
+```{r, echo=FALSE, message=FALSE,results='hide', warning=TRUE}
 ### to integrate 'pass'/fail
@@
 #quant_status (this isn't saving)
-#for(i in 1:length(MLH1_data$mouse)){
-#  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-#}
+for(i in 1:length(MLH1_data$mouse)){
+  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+}
 
@@
 #remove sections which are in category but have no data
-df_list <- df_list[ -c(7,14,16,22) ]
-#7: PERC male, 12: KAZ female! (14:CAST female, fake data) 16:HMI female
+df_list <- df_list[ -c(7,16,22) ]
+
+#7: PERC male, (14:CAST female, fake data) 16:HMI female, 22 other
+#
 
@@
 for (h in 1:length(df_list)) {
-  pf <- ggplot(data = df_list[[h]], y= adj_nMLH1.foci, x= quality, aes(colour = factor(Batch)), width = 0.25)
+  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ geom_jitter(aes(y= adj_nMLH1.foci, x= quality))
   # +   geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5)
-
-  + geom_jitter()
-  + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") + scale_fill_manual(values=c(NA, "red"), breaks=NULL)
+#  + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL)
 

### 2018-06-11 setup_data.R

@@ -47,6 +47,16 @@ MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]
 # add pass or fail column.. this will be a large list..maybe write a function..
 # highlight real passes? highlight non passes?
 # calculate automatically, list manually?
+pass_mice = c("10mar15_PWD_m2","13nov16_MSM_f1","13nov16_MSM_m1","16jan16_G_f2","16nov17_MSM_f2",
+"16nov17_MSM_f3","17mar16_G_f1","17mar16_G_f3","17mar16_G_f4","17mar16_G_f5",
+"18may15_PWD_m1", "1feb18_KAZ_m1","1feb18_MSM_m1", "20dec16_LEW_m2","20dec16_LEW_m3", 
+"28feb15_PWD_f2","30jun16_CAST_m3","31aug16_MSM_m1","31dec17_MSM_f5","3jan16_G_m1"   )
+
+#if mouse matches a mouse in the above list, add 'pass'
+
+for(i in 1:length(MLH1_data$mouse)){
+  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+}
 
 
 source("src/Func_addCategory.R")

### 2018-12-12 src/CommonFunc.R
laying ground work for more orgized repo

@@ -1,0 +1,5 @@
+# Common MLH1repo functions
+#
+#
+
+

### 2018-12-12 src/DataSetup.R

@@ -1,0 +1,27 @@
+# Data set up for MLH1 repo
+#
+#
+
+
+#libs
+
+
+#load big data files
+
+
+
+#run functions for labeling and standarizing
+
+
+#find empty/ probalematic cells
+
+
+#DF's to make
+#strain and subspecies DFs
+#batch DFs?
+
+
+#make summary tables
+#batches, average quality and number, (number removed)
+#strains  tally of mice, RRs
+

### 2019-02-27 MixedModel_PnD_Report.Rmd
starting to build mixed model by REML

---
title: "ANOVA_MixedModel"
author: "April Peterson"
date: "April 6, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(nlme)
library(plyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(raster)#for cV
library(knitr)

#load data
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
load(file="MLH1_data_setup.RData")
```

ReDo these mixed model analyses with REML

This current data set has `r length(MLH1_data$Random.Name)`. and `r length(unique( MLH1_data$mouse) )`  mice. (do these data include non-HM?)

I processed some of the data for these analyses,
  - changed all SPIC and SPRET labels to outgroup. (The fact that these are under 'subsp' is kinda misleading)
  - compare full cell data set to dataset of quality 123

#### Partition Variance, ANOVA, (using REML)

I wrote an power point (Jan18 lab meeting slide) explaining the MM. with REML

There are several constiderations / QC things to keep in mind for the data set

- use mouse average data
- only use house mouse data
- set Dom male as the intercept
- the sexes are matched (remove strains which don't have both sexes represented)

```{r}
#remove everything not H.M
HM.Table <- AP_mouse_table[AP_mouse_table$subsp != "outgroup",]
HM.Table <- HM.Table[HM.Table$subsp != "other",]

#make sure that sexes are matched! (remove strains that don't have matched data)
#table(HM.Table$strain,HM.Table$sex)
#REMOVE, HMI
#feb 2019, still need more imaging and data before this should be run again.

#make sure the the orders are right, intercept = Dom-male
#cast is number 1 first

dd <- lme(mean_co ~ sex * subsp, data=HM.Table, random=list(strain=pdDiag(~sex) ) )

```

make several models then compare the subsp and subsp:strain variance

I filtered the data, for mouse averages calculated from cells of quality 3 or lower and each mouse has 5 or more cells. (there are 52 mice). The originale mouse table has 105 mice.

when making these models  I learned...
- if I use the mouse averages for aov ... all of the DF will be used up.  These aov estimates should be done on the full cell count data frames

- nested models for cell level data (strain:subsp, sex:strain, )

- Discribe data (full and reduced based on quality...)

```{r make aov, include=FALSE}
#edit the cell level data (not sure how to remove mice with low number of cells)

#re-assigned spret and spic subsp to outgroup ()
MLH1_data$subsp <-  ifelse(grepl("Spic", MLH1_data$subsp), "outgroup", 
                    ifelse(grepl("Spretus", MLH1_data$subsp), "outgroup",
                                            MLH1_data$subsp))

MLH1_data.Mus <- MLH1_data[ !(grepl("Spretus", MLH1_data$subsp)), ]
MLH1_data.Mus <- MLH1_data[ !(grepl("Spic", MLH1_data$subsp)), ]

MLH1_data.123 <- MLH1_data[MLH1_data$quality <4, ]
MLH1_data.123 <- na.omit(MLH1_data.123)

#I should change the speret ext to outgroup

#the aovs
full.orig <- aov(nMLH1.foci  ~ subsp+strain:subsp+sex:strain+mouse:sex + quality, data = MLH1_data)
full.123 <- aov(nMLH1.foci  ~ subsp+strain:subsp+sex:strain+mouse:sex + quality, data = MLH1_data.123)

#not sure how to plot these models ... for testing assumptions

simp.model <- aov(nMLH1.foci  ~ subsp+strain:subsp+sex:strain+mouse:sex, data = MLH1_data)

full.orig.vals <- summary(full.orig)
full.123.vals <- summary(full.123)

#quality vs no quality...
#figure out how to display these...
#compare effects across full models
# the compare full to sex specific data (how much larger is the sum or squares )

male_full.orig <- aov(nMLH1.foci  ~ subsp+strain:subsp+ mouse:strain + quality, data = MLH1_data[MLH1_data$sex == "male", ] )
female_full.orig <- aov(nMLH1.foci  ~ subsp+strain:subsp+ mouse:strain + quality, data = MLH1_data[MLH1_data$sex == "female", ])

male_full.123 <- aov(nMLH1.foci  ~ subsp+strain:subsp+ mouse:strain + quality, data = MLH1_data.123[MLH1_data.123$sex == "male", ] )
female_full.123 <- aov(nMLH1.foci  ~ subsp+strain:subsp+ mouse:strain + quality, data = MLH1_data.123[MLH1_data.123$sex == "female", ])

```

```{r, echo=FALSE}
kable(summary(full.orig)[[1]],digits = 2, caption = "Full ANOVA table")
kable(summary(full.123)[[1]],digits = 2, format = "pandoc", caption = "reduced data (quality123) ANOVA table")
```

**Full vs reduced**: full model, all parameters/coeffecients are significant. sex has the largest amount of variance for both full and 123 data sets. The quality coeffecient is less significant in the 123 dataset. sex also becomes less efficient in the 123 dataset.

```{r, echo=FALSE}
kable(summary(female_full.orig)[[1]],digits = 2, format = "pandoc", caption = "Female Full ANOVA table")
kable(summary(female_full.123)[[1]],digits = 2, format = "pandoc", caption = "Female123 ANOVA table")
```

**Female specific**: subsp is not significant in the full dataset. mouse is less significant. quality is the most significant in full and in the 123 dataset, mouse is the most significant.
the difference between full and 123 dataset, quality becomes less significant in the 123 dataset. 
Caveat: subsp in females only has 3 DF (MUSC, Dom, Outgroup), because I don't have any cast data.

```{r, echo=FALSE}
kable(summary(male_full.orig)[[1]],digits = 2, format = "pandoc", caption = "Male Full ANOVA table")
kable(summary(male_full.123)[[1]],digits = 2, format = "pandoc", caption = "Male123 ANOVA table")
```

**male specific**: in the full dataset, subsp, quality, strain are max significant. Mouse is significant but less so. Subsp is explains most of the variance. in the 123 dataset, subsp is also most sig, quality exaplains a lot less. All coeffecients are significant still.

### Mixed Models

One of the point of comparing the mix model values is to assess if there are subsp and strain specific sexual dimorphism.
This is better than the ANOVA... because I can compare sexual dimorphism. Whereas the ANOVA framework I am looking at the between and within amount of variation.

The mixed models also serve as a saniety check-- for comparing the effects and directions on MLH1 number.

Better because it allows for fixed and random effects (instead of making wrong assumuptions about the effect distributions).

#### Setting up Model

Only use Musc and Dom?

Mouse averages are used, since MLH1 distributions are closer to normal distribution.

Coding for the mixed model; Dom male is the base.  (in most of my dataframes female is ordered before males, for visualization purposes). Effect values for sex and subsp will be for Female and Musc.

Subsp and sex (and interaction) are fixed effects. Strain and strain specific sex effect and residual are random effects.

Compare full dataset to 123quality dataset.

Compare full dataset to mice with 5 cells

```{r mouse quality, include=FALSE}
#make mouse average table with 
AP_mouse_table$mouse <- as.factor(AP_mouse_table$mouse)

AP_mouse_table1.2.3 <- ddply(MLH1_data[MLH1_data$quality > 3, ], c("subsp","strain","sex","mouse"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                         cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
)
AP_mouse_table1.2.3 <- na.omit(AP_mouse_table1.2.3)
AP_mouse_table1.2.3_5orMore <- AP_mouse_table1.2.3[AP_mouse_table1.2.3$Ncells > 5, ]

AP_mouse_table1.2.3_5orMore$subsp <-  ifelse(grepl("Spic", AP_mouse_table1.2.3_5orMore$subsp), "outgroup", 
                                      ifelse(grepl("Spretus", AP_mouse_table1.2.3_5orMore$subsp), "outgroup",
                                                    AP_mouse_table1.2.3_5orMore$subsp))
```

```{r full.model, echo=FALSE}

# just look at data of Musc and Dom 
#mixmodmouse_data <- AP_mouse_table[(AP_mouse_table$subsp == "Musc" | AP_mouse_table$subsp == "Dom"),  ]

#remove cast .. models can't be built with missing data

AP_mouse_table <- AP_mouse_table[(AP_mouse_table$subsp == "Musc" | AP_mouse_table$subsp == "Dom"),  ]
AP_mouse_table$sex <- factor(AP_mouse_table$sex, levels = c("male", "female"))

AP_mouse_table1.2.3 <- AP_mouse_table1.2.3[(AP_mouse_table1.2.3$subsp == "Musc" | AP_mouse_table1.2.3$subsp == "Dom"),  ]
AP_mouse_table1.2.3 <- na.omit(AP_mouse_table1.2.3)
AP_mouse_table1.2.3$sex <- factor(AP_mouse_table1.2.3$sex, levels = c("male", "female"))
AP_mouse_table1.2.3$subsp <-  factor(AP_mouse_table1.2.3$subsp, levels = c("Dom", "Musc"))

AP_mouse_table1.2.3_5orMore <- AP_mouse_table1.2.3[AP_mouse_table1.2.3$Ncells > 5, ]

MMfull <- lme(mean_co ~ sex * subsp, data=AP_mouse_table, random=list(strain=pdDiag(~sex)) ) 
MM123 <- lme(mean_co ~ sex * subsp, data=AP_mouse_table1.2.3, random=list(strain=pdDiag(~sex)) ) 
MM123.5 <- lme(mean_co ~ sex * subsp, data=AP_mouse_table1.2.3_5orMore, random=list(strain=pdDiag(~sex)) )

aa <- summary(MMfull)
bb <- summary(MM123)
cc <- summary(MM123.5)
```

To interpert the summary results, (Fixed effects)
- the intercept is the mean MLH1 number of the base factors -- which in this model is defined as Dom male (26.40).

- the next variable is sex being female, has an affect of increasing the mean MLH1 count by 1.7. (This is about the amount of sexual dimorphism)
1.7 (SD)

- the value for Musc is 3.935, meaning being male and Musc adds 3.935 to the mean MLH1 (in the previous model without KAZ, this was 5.5)

- the interaction of sex and subspecies is -3.7 meaning, musc females are 3 MLH1 foci lower than musc males (in previous models this was -5.1)

Random effects (strain and strain specific sexual dimorphism)
- the intercept is the average sd from strains 1.89, 
- the sex effec is the average sd from strain specific sexual dimorphism 2.07
- the random effects can be estimated; the intercepts are the average sd for the strain effect, ... from these esitmates is seems that G and Kaz have strain specific sexual dimorphism where females are larger if the strain specific SD effect is larger than the whole sex effect

```{r, include=FALSE}
#make new AP mouse table with cells from 1-3
quality_data <- MLH1_data[MLH1_data$quality <= 3, ]
quality_data <- quality_data[(quality_data$subsp == "Musc" | quality_data$subsp == "Dom"),  ]

na.omit(quality_data)

qualAP_mouse_table <- ddply(quality_data, c("subsp", "sex", "strain", "mouse"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                         cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
                        #sex = sex,
                        #subsp = subsp,
                        #strain = strain
                      #quality?

)
qualAP_mouse_table <- na.omit(qualAP_mouse_table)

qualAP_mouse_table <- qualAP_mouse_table[qualAP_mouse_table$Ncells > 5,]

qualAP_mouse_table$sex <- factor(qualAP_mouse_table$sex, levels = c("male", "female"))
qualAP_mouse_table$strain <- factor(qualAP_mouse_table$strain)
qualAP_mouse_table$subsp <-  factor(qualAP_mouse_table$subsp, levels=c("Dom", "Musc"))

model2 <- lme(mean_co ~ sex * subsp, data=qualAP_mouse_table, random=list(strain=pdDiag(~sex)) )

#summary(model2)
#ranef(model2)
```

For this version of cleaned data, the effect of Musc on the mean MLH1 is lower.

### 2019-05-10 doc/MeioticErrorReport.Rmd
moved files around to clean up

---
title: "MeioticErrorReport"
author: "April Peterson"
date: "March 1, 2018"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, echo=FALSE}
#setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(png)
library(reshape2)
```

## Meiotic Errors Report

### background on 'defects'
Example images, what do I mean by asynapsis and achiasmate?

tried adding example images, but rmd not coorporating.

![Cell with asynasped chrms.](asyn_example.png)
asynapsis, SC comes apart, SC not fully synapsed. (prevents MLH1 foci)

![0CO chrm example.](achi_example.png)

0CO class (not technically achiasmate, but that's what we've been calling it)
more prevelant in oocytes because checkpoints are weaker... (source)

##### Biological relevance

Different asynaspsis levels could indicate that there are differences in the speed and timing of the meiotic cycle. Synapsis marks the pachtyenma stage, there is asynapsis before and affer.
I believe that The asynapsis I note ... is affects a subset of the chromosomes in a cell. (chrms are out of sync with each other), sometimes I observed all chrms are fully synasped and a single homolog pair is unsynapsed. (BD mentioned she observed a high number of X homolog asynapsis I'm not sure how she knew they were Xs)

0CO chrms are not nesscearilly achiasmate tetrads, a better measure would be taken from when chiasmata appear across the spindel.  I think the biological relevance to these measures are either the CO maturation effect or balance/bias of MLH1 CO pathway compared to MLH1-independant pathways.

##### Predictions:

  - Females will have more 0COs and asynasped chrms per cell
  - Within females, higher CO strains, (G), will have fewer 'defects'

#### Data

I will calculate the average number of 0CO and asynasped chrm per cell (since the number of cells in the MLH1 data set is variable.) I'll do this at the strain and mouse level.

```{r tables, echo=FALSE, include=FALSE}
MLH1_data$achiasmate <- as.numeric(MLH1_data$achiasmate)
MLH1_data$asynased <- as.numeric(MLH1_data$asynased)
#remove NA sex  or strain = other

cell_defect_table <- ddply(MLH1_data, c("sex", "strain", "category"), summarise,
                          Ncells = length(unique(Original.Name)),
                          nrows = length(Original.Name),

                          nNA_asyn = sum(is.na(asynased)),
                          nNA_0CO = sum(is.na(achiasmate)),

                          total_0CO =  sum(asynased),
                          total_asyn = sum(achiasmate, na.rm = TRUE),
                          prop.0CO = total_0CO/Ncells, #these will differ for males and females
                          prop.asyn = total_asyn/Ncells
                                  )

cell_defect_table_mouse <- ddply(MLH1_data, c("sex", "strain", "mouse"), summarise,
                          Ncells = length(unique(Original.Name)),
                          nrows = length(Original.Name),

                          nNA_asyn = sum(is.na(asynased)),
                          nNA_0CO = sum(is.na(achiasmate)),

                          total_0CO =  sum(asynased),
                          total_asyn = sum(achiasmate, na.rm = TRUE),
                          prop.0CO = total_0CO/Ncells, #these will differ for males and females
                          prop.asyn = total_asyn/Ncells
                                  )

```

```{r plots make, echo=FALSE, include=FALSE, warning=FALSE}
##ggplots for strain measures, general values.

av_asyn_plot.s <- ggplot(cell_defect_table, aes(y=prop.asyn,x=category, color=strain)) +   geom_point(aes(shape = factor(sex)))  + ggtitle("Average asynasped per cell") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

av_0CO_plot.s <- ggplot(cell_defect_table, aes(y=prop.0CO,x=category, color=strain)) +   geom_point(aes(shape = factor(sex)))  + ggtitle("Average achiasmate per cell") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

#mouse averages caluclated
#don't put mouse in x value,  I could make the jitter points, differen sizes for number of cells..?

av_0CO_plot.m <- ggplot(cell_defect_table_mouse, aes(y=prop.0CO,x=strain, size = Ncells))   +  geom_jitter(aes(fill=strain), shape=21 )  +
 ggtitle("Mouse level average 0CO chrms per cell") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylim(c(1.5,5)) + facet_grid(~ sex)

av_asyn_plot.m <- ggplot(cell_defect_table_mouse, aes(y=prop.asyn,x=strain, size = Ncells)) +  geom_jitter(aes(fill=strain), shape=21)  + ggtitle("Mouse level average asyn per cell") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylim(c(0,2.2)) + facet_wrap(~ sex, scales="free")
```

From going back and looking at the files and counting in the table, it seems that the number of cells for which 0CO and asynasped wasn't counted is minimal. 
These missing values will be easy to go back a fix, but in the meantime, I these will do for assessing general patterns.

(make sure to check where the datasetup, removes any of these rows)

 or other ways to visualize these
boxplots not good since they need distributions. the average mesure per cell is a single measure- point, (with CI). If I calculate the mouse average proportions, those can then form distributions.

```{r plots1 show, echo=FALSE}
av_asyn_plot.s
av_0CO_plot.s
```

In all strains, females have more 'defects' than males. (expect for the CAST female data, which is probaby fake). Spicilegus females have the highest 0COs and asynapsed chrms. Males seem to have a similar low level of average 0COs (PWD and SPRET have slightly more)

Females have more 0CO and asynasped chrms than males. The number of asynased varies by strain, while the proportion of 0CO is pretty consistant across strains.

The mouse average distributions of defect proportions --- females have higher values and are more variable across mice.

```{r plots2 show, echo=FALSE}
av_asyn_plot.m
av_0CO_plot.m
```

### Further Questions

-  do defects affect MLH1 measures (compare mouse averages for defective and all normal cells)?

- what is the relationship between cell quality scores and defects?

### 2019-05-10 doc/pass_report.Rmd
moved files around to clean up

---
title: "passing mice script"
author: "April Peterson"
date: `r format(Sys.time(), ''%d %B, %Y'')`
output: rmarkdown::github_document
---

Guide / report document for showing which mice have passed - sufficient cells samples for showing good estimate of MLH1 count.

Outline
-dicription of graphs / data
-discription of simulations (sampling comparison)
(example images)

<code for making the graphs
-add mouse name, total number of cells,

- develope thresholds for grouping mice into passed and more images
- consider doing this by batch

- based on current plot outputs,

- think of ways to combine this sampling size data with quality distributions
```{r setup, include=FALSE}

setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)
```

```{r group_mice}
#seperate large dataframe by how many cells are in the samples
twenty <- AP_mouse_table[AP_mouse_table$Ncells > 20,] #53
#are these assumed to have passed?

lessTwenty <- AP_mouse_table[AP_mouse_table$Ncells < 20,] #28
#these will give the list of mice

#subset MLH1_data by these lists. (fixed line below, )
twenty_data <- MLH1_data[MLH1_data$mouse %in% twenty$mouse, ]
lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]

```

### Power and sampling size effects
Create dataframes that iterate through different sample sizes of MLH1 counts 
 Outlists are dataframes, filled with pvalues from t-tests of different sample sizes.
 Q: How do the 20cells or greater mice observations compare with the less than 20 cell mouse observations.

Passing cells have a small effect from different sampling sizes.

```{r pvalue_permutations, echo=FALSE}
sim_iterations <- 10000

#list, DF, ncells, sim_num

source("src/Func_sampleCells.R")
lessTwenty_pvals_5_10 <- sample_cells_pvals(lessTwenty, lessTwenty_data, 5, 10)#this returns a list (of vectors?) list of pvalues for each mouse in a list

#nn[[4]] is the syntax to get just the pvalues
#names(nn[4]) is now you get the name of the vector / list

#sample >twenty mice

#REDO these, I think the function was wrong
Twenty_pvals_10_1k <- sample_cells_pvals(twenty, twenty_data, 10, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
Twenty_pvals_15_1k <- sample_cells_pvals(twenty, twenty_data, 15, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
Twenty_pvals_20_1k <- sample_cells_pvals(twenty, twenty_data, 20, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list

#decide on threshold for difference in pvalues

#below doesn't work
#guess <- Twenty_pvals_15_1k - Twenty_pvals_10_1k

guess <- mean(Twenty_pvals_15_1k[[1]]) - mean(Twenty_pvals_10_1k[[1]])

#make blank df
mean1<- vector()
mean2<- vector()
mean3<- vector()

onelist <- vector()
twolist <- vector()
threelist <- vector()
for(l in 1:length(Twenty_pvals_10_1k)){
  #var doesn't seem to be good
  mean1 <- mean(Twenty_pvals_10_1k[[l]])
  mean2 <- mean(Twenty_pvals_15_1k[[l]])
  mean3 <- mean(Twenty_pvals_20_1k[[l]])
  first <- ( mean(Twenty_pvals_10_1k[[l]]) - mean(Twenty_pvals_15_1k[[l]]) )
  sec <- mean(Twenty_pvals_15_1k[[l]]) - mean(Twenty_pvals_20_1k[[l]])
  thrid <- mean(Twenty_pvals_10_1k[[l]]) - mean(Twenty_pvals_20_1k[[l]])

  mean1.list[l] <- mean1
  mean2.list[l] <- mean2
  mean3.list[l] <- mean3

   onelist[l] <- first
  twolist[l] <- sec
  threelist[l] <- thrid
}

comp.pvals <- data.frame(mean1=mean1.list, mean2=mean2.list, mean3=mean3.list, dif1 = onelist, dif2 = twolist, dif3=threelist)
row.names(comp.pvals) <- names(Twenty_pvals_10_1k)

#comp.pvals now is a df that 
#rowSums(comp.pvals) 
ll <- rowSums(comp.pvals)

#this excludes ... mice/rows in data frame which have greater differences 
#between 15 and 20 samples
mm <- subset(comp.pvals, comp.pvals$dif2 >= -.05)

#some bad looking mice still get through (is this due to sample size)

# it seems like mice with between 20-25 cells, there is a large difference
# between the p values of 15 and 20

```

what does it mean that I excldude these mice?
I will infer differen mean MLH1 values due to sample size with mice with

10,15,20, might not be the best for mice with around 20 cells

less than 20
```{r}

lessTwenty <- subset(AP_mouse_table, Ncells >15 & Ncells <20)
lessTwenty_data <- MLH1_data[MLH1_data$mouse %in% lessTwenty$mouse,]

source("src/Func_sampleCells.R")
lessTwenty_pvals_8_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 8, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
lessTwenty_pvals_11_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 11, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list
lessTwenty_pvals_15_1k <- sample_cells_pvals(lessTwenty, lessTwenty_data, 15, 1000)#this returns a list (of vectors?) pvalues  for each mouse in a list

#make blank df
mean1.list<- vector()
mean2.list<- vector()
mean3.list<- vector()

onelist <- vector()
twolist <- vector()
threelist <- vector()
for(l in 1:length(lessTwenty_pvals_8_1k)){
  #var doesn't seem to be good
  mean1 <- mean(lessTwenty_pvals_8_1k[[l]])
  mean2 <- mean(lessTwenty_pvals_11_1k[[l]])
  mean3 <- mean(lessTwenty_pvals_15_1k[[l]])

  first <- ( mean(lessTwenty_pvals_8_1k[[l]]) - mean(lessTwenty_pvals_11_1k[[l]]) )
  sec <- mean(lessTwenty_pvals_11_1k[[l]]) - mean(lessTwenty_pvals_15_1k[[l]])
  thrid <- mean(lessTwenty_pvals_8_1k[[l]]) - mean(lessTwenty_pvals_15_1k[[l]])

  mean1.list[l] <- mean1
  mean2.list[l] <- mean2
  mean3.list[l] <- mean3

   onelist[l] <- first
  twolist[l] <- sec
  threelist[l] <- thrid
}

LTcomp.pvals <- data.frame(mean1=mean1.list, mean2=mean2.list, mean3=mean3.list, dif1 = onelist, dif2 = twolist, dif3=threelist)
row.names(LTcomp.pvals) <- names(lessTwenty_pvals_8_1k)

#dif2 cut off -0.10, greater than -.1 mice should pass even though they have less than 20

pass_LessTwenty <- row.names(LTcomp.pvals[(LTcomp.pvals$dif2 > -.1),])

```

## what about the effects of quality on sample size effects

### 2019-05-10 doc/strain_reports/AST_Report.Rmd
moved files around to clean up

---
title: "AST_Report"
author: "April Peterson"
date: "January 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(plyr)
library(dplyr)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#MLH1_data = read.csv("data/AnonData.csv", header=TRUE )
load(file="~./MLH1repo/MLH1_data_setup.RData")

#input
#output: Rmd file that can be read on github

knitr::opts_chunk$set(echo = TRUE)

mouse_list = read.csv("~./MLH1repo/data/MLH1q_AST.csv", header=TRUE )
# this is the list of total mice
#load big MLH1 Rdata file

#format the dates into standard format
mouse_list$DOB <- as.Date(mouse_list$DOB,format='%m/%d/%Y')

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
source("~./MLH1repo/src/CommonFunc.R")#this works
#source("src/Func_addCategory.R")

mouse_list <- add_mouse1(mouse_list)
mouse_list <- add_strain(mouse_list)
#source("src/Func_addSex.R")
mouse_list <- add_sex(mouse_list)
mouse_list <- add_category(mouse_list)
#set the order of categories (female, male) (cast, dom, musc)
#source("src/Func_addStrain.R")
mouse_list <- add_euth_date(mouse_list)
mouse_list <- add_age(mouse_list)

```

```{r}
#make a tables, sex, and ages
#count mice by categories of sex and ages

#ages in days -- 
#5 to 15 weeks
#35 days,  105days

#don't go row by row, for each mouse, 
#summarize the whole dataset...
#make new dataframe
mouse.report <- data.frame("total.Nmice" = length(unique(mouse_list$mouse)), 
      "Nmales"= length(mouse_list$mouse[mouse_list$sex == "male"]),
      "Nfemales"= length(mouse_list$mouse[mouse_list$sex == "female"]),
      "passed.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse)

     )
#imaged mice...

#sum( mouse_list$mouse  %in%  MLH1_data$mouse)
#ifelse(mouse_list$mouse %in% MLH1_data$mouse, "1", "")

```

```{r BasicStats, echo=FALSE}
#how many euthed?
#how many mice MLH1 stained and imaged?
#what is relevant meta data?

#how many cells?

#draw info from anonMLH1 data -- MLH1 rate, quality,

```

```{r}
#print out plots

```

### 2019-05-10 doc/strain_reports/CAROLI_Report.Rmd
moved files around to clean up

--- 2019-05-10 doc/strain_reports/AST_Report.Rmd
@@
 ---
-title: "AST_Report"
+title: "CAROLI Report"
 author: "April Peterson"
-date: "January 4, 2019"
-output: html_document
+date: "December 12, 2018"
+output: pdf_document
 ---
@@
 library(dplyr)
+library(ggplot2)
 
@@
 
-mouse_list = read.csv("~./MLH1repo/data/MLH1q_AST.csv", header=TRUE )
+mouse_list = read.csv("~./MLH1repo/data/MLH1q_CAROLI.csv", header=TRUE )
 # this is the list of total mice
@@
       "Nfemales"= length(mouse_list$mouse[mouse_list$sex == "female"]),
-      "passed.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse)
+      "passed.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse),
+
+      #length of $M.stained without staining in it
+      "Non.Stained.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse),
+      #length of $quant.batch without Batch in it
+      "Non.Qunt.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse)
 
      )
+#number of unquant'd mice
+#number of unstained mice
+
 #imaged mice...
@@
 
+```
+
+```{r}
+#MAKE HISTOGRAM OF THE AGES 
+Ages_hist <- ggplot(data= mouse_list, aes(age.weeks))+geom_histogram()+facet_wrap(~sex)
+
+```
+
+```{r}
+#merge DOB from ALP metdata sheet?
 ```

### 2019-05-10 doc/strain_reports/KAZ_Report.Rmd
moved files around to clean up

--- 2019-05-10 doc/strain_reports/AST_Report.Rmd
@@
 ---
-title: "AST_Report"
+title: "TOM_Report"
 author: "April Peterson"
 date: "January 4, 2019"
-output: html_document
+output: pdf_document
 ---
@@
 
-#input
-#output: Rmd file that can be read on github
-
 knitr::opts_chunk$set(echo = TRUE)
 
-mouse_list = read.csv("~./MLH1repo/data/MLH1q_AST.csv", header=TRUE )
+mouse_list = read.csv("~./MLH1repo/data/MLH1q_KAZ.csv", header=TRUE )
 # this is the list of total mice
 #load big MLH1 Rdata file
-
 #format the dates into standard format
 mouse_list$DOB <- as.Date(mouse_list$DOB,format='%m/%d/%Y')
+#check that the header is the same
 
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-source("~./MLH1repo/src/CommonFunc.R")#this works
+source("~./MLH1repo/src/CommonFunc.R")
 #source("src/Func_addCategory.R")
@@
 
-#ages in days -- 
-#5 to 15 weeks
-#35 days,  105days
+#merge 
+#mouse_list
+
+smllmergedDF = merge(MLH1_data, mouse_list, by = "mouse")
+#these are all cell level obsers
+
+#compare the number of images across reports .... total.image ... good.images
+mouse.comparison <- ddply(mouse_list, c("mouse"), summarise,
+
+                          Nmice = length(unique(mouse)),
+                         #give the value from the AP_mouse table
+                         nQunt.images = AP_mouse_table$Ncells[which(AP_mouse_table$mouse == mouse)]
+
+            )
 
@@
       "Nmales"= length(mouse_list$mouse[mouse_list$sex == "male"]),
+
+      "passed.mice.m" = sum( mouse_list$mouse[mouse_list$sex == "male"]  %in%  MLH1_data$mouse),
+
       "Nfemales"= length(mouse_list$mouse[mouse_list$sex == "female"]),
-      "passed.mice" = sum( mouse_list$mouse  %in%  MLH1_data$mouse)
+
+      "passed.mice.f" = sum( mouse_list$mouse[mouse_list$sex == "female"]  %in%  MLH1_data$mouse)
 
@@
 #imaged mice...
+#fill in number of images from MLH1,
+#unquant
+#unstained
+#non-imaged
 
@@
 ```{r}
+#MAKE HISTOGRAM OF THE AGES 
+Ages_hist <- ggplot(data= mouse_list, aes(age.weeks))+geom_histogram()+facet_wrap(~sex)
+#not sure if histograms are the best
+
+```
+
+```{r}
 #print out plots

### 2019-05-10 doc/strain_reports/TOM_Report.Rmd
moved files around to clean up

--- 2019-05-10 doc/strain_reports/AST_Report.Rmd
@@
 
-mouse_list = read.csv("~./MLH1repo/data/MLH1q_AST.csv", header=TRUE )
+mouse_list = read.csv("~./MLH1repo/data/MLH1q_TOM.csv", header=TRUE )
 # this is the list of total mice

### 2019-05-10 doc/strain_reports/WSB_Report.Rmd
moved files around to clean up

--- 2019-05-10 doc/strain_reports/AST_Report.Rmd
@@
 
-mouse_list = read.csv("~./MLH1repo/data/MLH1q_AST.csv", header=TRUE )
+mouse_list = read.csv("~./MLH1repo/data/MLH1q_WSB.csv", header=TRUE )
 # this is the list of total mice
@@
 
+```{r}
+#MAKE HISTOGRAM OF THE AGES 
+Ages_hist <- ggplot(data= mouse_list, aes(age.weeks))+geom_histogram()+facet_wrap(~sex)
+#not sure if histograms are the best
+
+```
+
+```{r}
+#print out plots
+
+```
+
 ```{r BasicStats, echo=FALSE}
@@
 ```
-
-```{r}
-#print out plots
-
-```

### 2019-05-10 draft_scripts/pass_report.Rmd
moved files around to clean up

---
title: "Sub-Sampling Effect"
author: "April Peterson"
date: "June 4, 2018"
output: rmarkdown::github_document
---

#### Past certeria for mouse level

BD: Approximately 20 cells were scored per animal. We were unable to obtain a sufficient number of high quality images for 39 of the 315 F2 animals.

RW: The average individual was represented by counts from >20 spermatocytes; we omitted individuals with counts from <5 spermatocytes.

AP: >20 cells with 3 highest quality (1,2,3).

Guide / report document for showing which mice have passed - sufficient cells samples for showing good estimate of MLH1 count.
Compare re-sampling simulations for 'Hard Pass / Strict Threshold' data to on the edge data, so assess if there is a large effect on mouse-MLH1 mean for cell sample size.

Maybe female and males will have different effects.

```{r "setup", include=FALSE}
require("knitr")
#opts_knit$set(root.dir = "~/gits/project/")

opts_knit$set(root.dir = "C:/Users/alpeterson7/Documents/MLH1repo")
setwd("C:/Users/alpeterson7/Documents/MLH1repo")
#setwd("Documents/MLH1repo")
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="MLH1_data_setup.RData")
library(ggplot2)
library(plyr)
library(dplyr)
library(raster)
library(reshape2)

```

### Passing Mouse list

```{r mouse_lists, include=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above4 = sum(as.numeric(quality) <= 3, na.rm = TRUE ),
                        #add percentages
                        prop.lt5 = above4/total
)
length(q_cutoff_table$mouse)#116

pass_mus_table <- q_cutoff_table[q_cutoff_table$above4 >= 25,] #30
# boarder line
boarder_mice_list <- q_cutoff_table[ ( q_cutoff_table$prop.lt5 >= .6) & (q_cutoff_table$above4 < 20), ] #39

```

Mice with slides with over 25 cells of good quality: `r length(pass_mus_table$mouse)`
Mice with fewer than 20 cells, but have high proportion of good quality cells (>60%): `r length(boarder_mice_list$mouse)`.

What is the distribution of total cell counts for these? For the boarderline mice, total is around ~20 cells, ~15 cells above 4 quality.

Run sub-sampling simulations for data of all these mice ... This requires adjusting the sample sizes, over half the mice have 15 or more.

The chunks with the code for the sub-sampling simulations are skipped.

```{r pvalue_permutations HQ, echo=FALSE,include=FALSE, eval = FALSE}
HQ <- pass_mus_table
sim_iterations <- 1000

outlist15 = list()
for(fif in 1:length(unique(HQ$mouse))){
  #print(HQ$mouse[fif])
  mus_data <- MLH1_data[MLH1_data$mouse == HQ$mouse[fif],] #... make sure this is right

  outlist15[fif] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 15, replace = FALSE) )$p.value ) )
  }
names(outlist15) <- HQ$mouse
#outlist is not a list of p-value list. ()

outlist10 = list()
for(ten in 1:length(unique(HQ$mouse))){
 # print(HQ$mouse[ten])
  mus_data <- MLH1_data[MLH1_data$mouse == HQ$mouse[ten],]
  outlist10[ten] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE) )$p.value ) )
  }
names(outlist10) <- HQ$mouse

outlist20 = list()
for(tw in 1:length(unique(HQ$mouse))){
 # print(HQ$mouse[tw])
  mus_data <- MLH1_data[MLH1_data$mouse == HQ$mouse[tw],]
  outlist20[tw] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE),
                sample(mus_data$adj_nMLH1.foci, 20, replace = FALSE) )$p.value ) )
  }
names(outlist20) <- HQ$mouse
#outlist is not a list of p-value list. ()

#list for n=20 sample simulations
plot_list20 = list()
for (i in 1:length(HQ$mouse)){
#  print(i)  
  try_df <- as.data.frame(outlist20[[i]])
  colnames(try_df) <- "pvals"
  Title_mus_name <- paste("1000 iterations\n, Test of 2 Samples of 20, \n", names(outlist20[i]) )
  p20 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
    plot_list20[[i]] <- p20
}
#list for n=10 sample simulations
plot_list10 = list()
for (i in 1:length(HQ$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist10[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 10, \n", names(outlist10[i]) )
  colnames(try_df) <- "pvals"
  p10 <- ggplot(try_df) + ggtitle( Title_mus_name ) +xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
  plot_list10[[i]] <- p10
}

#list for n=15 sample simulations
plot_list15 = list()
for (i in 1:length(HQ$mouse)){
  print(i)  
  try_df <- as.data.frame(outlist15[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 15, \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"
  p15 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
    plot_list15[[i]] <- p15
}

#make list of scatter plots
sctr_plot = list()
for (i in 1:length(HQ$mouse)){
  print(i)  
  df10 <- as.data.frame(outlist10[[i]])
  df15 <- as.data.frame(outlist15[[i]])
  df20 <-as.data.frame(outlist20[[i]])
  names(df10) = "pvals"
  names(df15)= "pvals"
  names(df20)= "pvals"
  allDF <- data.frame(pvals = rbind(df10, df15, df20),
    DFsets = c (rep("samp10",length(df10[,1]) ), 
              rep("samp15",length(df10[,1])), 
              rep("samp20", length(df10[,1])) )
                      )
    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist15[i]) )
  colnames(try_df) <- "pvals"

  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
        geom_jitter(aes(x=DFsets, y=pvals) )
    sctr_plot[[i]] <- scplot
}

source("src/MultiPlot.R")
#multiplot(plot_list10[[8]],plot_list20[[8]], cols=1)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/outside_gitrepo/sampling_effect")

# Saves multiple plots to a tiff file for each mouse
for(i in 1:length(HQ$mouse)) {  #replace counter with mouse name
    file_name = paste("HQ_Sample_sizes_p_", HQ$mouse[i], ".png", sep="")
    #tiff(file_name)
    png(filename = file_name)
    #par( mfrow=c(1,2) )
    multiplot(sctr_plot[[i]], plot_list10[[i]], plot_list15[[i]], plot_list20[[i]],cols=2)
    dev.off()
}
```

eval = FALSE
-- errors at this section likely due to changes in working directories
```{r pvalue_permutations Boarderline, echo=FALSE, include=FALSE}
#HQ <- pass_mus_table
BL <- boarder_mice_list[boarder_mice_list$total >= 15,]
sim_iterations <- 1000

### adjust sampling sizes (limit this list to mice with )
# max 15,   subampling 12,10,8

outlist12 = list()
for(fif in 1:length(unique(BL$mouse))){
 # print(BL$mouse[fif])
  mus_data <- MLH1_data[MLH1_data$mouse == BL$mouse[fif],] #... make sure this is right

  outlist12[fif] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 12, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 12, replace = FALSE) )$p.value ) )
  }
names(outlist12) <- BL$mouse
#outlist is not a list of p-value list. ()

outlist10 = list()
for(ten in 1:length(unique(BL$mouse))){
  #print(BL$mouse[ten])
  mus_data <- MLH1_data[MLH1_data$mouse == BL$mouse[ten],]
  outlist10[ten] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE),
                        sample(mus_data$adj_nMLH1.foci, 10, replace = FALSE) )$p.value ) )
  }
names(outlist10) <- BL$mouse

outlist8 = list()
for(tw in 1:length(unique(BL$mouse))){
  #print(BL$mouse[tw])
  mus_data <- MLH1_data[MLH1_data$mouse == BL$mouse[tw],]
  outlist8[tw] <- list(replicate(sim_iterations, t.test(sample(mus_data$adj_nMLH1.foci, 8, replace = FALSE),
                sample(mus_data$adj_nMLH1.foci, 8, replace = FALSE) )$p.value ) )
  }
names(outlist8) <- BL$mouse
#outlist is not a list of p-value list. ()

#list for n=8 sample simulations
plot_list8 = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  try_df <- as.data.frame(outlist8[[i]])
  colnames(try_df) <- "pvals"
  Title_mus_name <- paste("1000 iterations\n, Test of 2 Samples of 8, \n", names(outlist8[i]) )
  p8 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
    plot_list8[[i]] <- p8
}
#list for n=10 sample simulations
plot_list10 = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  try_df <- as.data.frame(outlist10[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 10, \n", names(outlist10[i]) )
  colnames(try_df) <- "pvals"
  p10 <- ggplot(try_df) + ggtitle( Title_mus_name ) +xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
  plot_list10[[i]] <- p10
}

#list for n=12 sample simulations
plot_list12 = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  try_df <- as.data.frame(outlist12[[i]])
  Title_mus_name <- paste("1000 iterations\n Test of 2 Samples of 12, \n", names(outlist12[i]) )
  colnames(try_df) <- "pvals"
  p12 <- ggplot(try_df) + ggtitle( Title_mus_name ) + xlim(0,1)+
        geom_histogram(aes(x=pvals) )#xlim
  #tryfiting 3 plots
    plot_list12[[i]] <- p12
}

#make list of scatter plots
sctr_plot = list()
for (i in 1:length(BL$mouse)){
 # print(i)  
  df12 <- as.data.frame(outlist12[[i]])
  df10 <- as.data.frame(outlist10[[i]])
  df8 <-as.data.frame(outlist8[[i]])
  names(df12) = "pvals"
  names(df10)= "pvals"
  names(df8)= "pvals"
  allDF <- data.frame(pvals = rbind(df8,df10, df12),
             DFsets = c( rep("samp8",length(df8[,1]) ),
                         rep("samp10",length(df10[,1]) ),
                         rep("samp12", length(df12[,1]) )
                      )
    Title_mus_name <- paste("1000 iterations\n Pvalues of T-tests \n", names(outlist12[i]) )
  colnames(try_df) <- "pvals"

  scplot <- ggplot(allDF) + ggtitle( Title_mus_name) +
        geom_jitter(aes(x=DFsets, y=pvals) )
    sctr_plot[[i]] <- scplot
}

#source("src/MultiPlot.R")
#multiplot(plot_list10[[8]],plot_list12[[8]], cols=1)  #forget what this is for

setwd("./outside_gitrepo/sampling_effect") #the loop is not printing in the folders I want

# Saves multiple plots to a tiff file for each mouse
for(i in 1:length(BL$mouse)) {  #replace counter with mouse name
  file_name = paste("BL_Sample_sizes_p_", BL$mouse[i], ".png", sep="")
  #tiff(file_name)
  png(filename = file_name)
  #par( mfrow=c(1,2) )
  multiplot(sctr_plot[[i]], plot_list8[[i]], plot_list10[[i]], plot_list12[[i]],cols=2)
  dev.off()
}
```

```{r}
#![text](./outside_gitrepo/sampling_effect/BL_Sample_sizes_p_4dec17_HMI_m1.png)

#setwd("./outside_gitrepo/sampling_effect")
```

![text](C:/Users/alpeterson7/Documents/MLH1repo/outside_gitrepo/sampling_effect/BL_Sample_sizes_p_4dec17_HMI_m1.png)

![text](C:/Users/alpeterson7/Documents/MLH1repo/outside_gitrepo/sampling_effect/BL_Sample_sizes_p_4dec17_HMI_m1.png)

From comparing the sub-sampling plots from High quality (HQ) and Boarder line (BL),

7 mice out of 29 boardline mice, looked like there was a sampling effect. I will prioritie re-imaging those, and start re-imaging the rest with fewer than 20 cells.

### 2019-05-10 draft_scripts/test4pvalues.Rmd
moved files around to clean up

--- 2017-12-10 test4pvalues.Rmd
@@
   print(twenty$mouse[fif])
-  #
-  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[fif],]
+  mus_data <- twenty_data[twenty_data$mouse == twenty$mouse[fif],]  #mus data is the mouse specific data
 
@@
 
+#list for n=20 sample simulations
 plot_list20 = list()
@@
 
+#list for n=10 sample simulations
 plot_list10 = list()
@@
 
+#list for n=15 sample simulations
 plot_list15 = list()

### 2019-05-10 draft_scripts/meta_data_analysis.R

@@ -10,39 +10,71 @@ library(plyr)
 library(lattice)
 library(dplyr)
 
-#age, mat and pat age, Gough Line, (staining batch for cells)
+#orig cols
+#mouse, DOB, mat DOB, pat DOB, stain, 
+
+#calq cols
+#sex, age, dissection date, mat age, pat age, time to stain (cell measure dependant)
+
+# final cols
+#Gough Line, (staining batch for cells)
+#
 
 #load meta data csv file 
 #finalize the coding (NAs, ect)
-meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1data/data/mouseDOBs.csv")
+#meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1data/data/mouseDOBs.csv")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+full_meta_data = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/ALP_MouseMetadata.csv")
 
-save.image("play_data_for_meta.RData")
-#data needs to be cleaned up (all NA's removed before calculating )
+#save.image("play_data_for_meta.RData")
 
-#mouse column already included,
-#change category, sex -- there is no file name...
-#using the functions will be a little tricky
-
-meta_data$mouse <- as.character(meta_data$mouse)
-#format euth date (from mouse)
-#all NA's need to be removed
-
-for( t in 1:length(meta_data$mouse)){
-  euth_date = strsplit(meta_data$mouse[t], split="_")[[1]][1]
-  meta_data$raw_euth_date[t] <- strsplit(meta_data$mouse[t], split="_")[[1]][1]
-  fomt_euth <- as.Date(strsplit(meta_data$mouse[t], split="_")[[1]][1], format= '%d%b%y')
+### mat age pat age ###
+full_meta_data$mouse <- as.character(full_meta_data$mouse)
+#loop for calculating ages
+for( t in 1:length(full_meta_data$mouse)){
+  euth_date = strsplit(full_meta_data$mouse[t], split="_")[[1]][1]
+  #full_meta_data$raw_euth_date[t] <- strsplit(full_meta_data$mouse[t], split="_")[[1]][1]
+  
[…]

### 2019-05-10 draft_scripts/pass_mouse_list.R

@@ -1,0 +1,34 @@
+# list of passing mice
+# input: current BivData.RData
+# output: csv (or other formate) list of passing mice
+# print out for hard copy and to be referenced by other scripts
+
+library(plyr)
+# initial number of cells 25.. (kinda abritrry)
+# (high propotion of cells with quality less than 5) q.lt5 > .7
+
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+#setwd("Documents/MLH1repo")
+
+load(file="MLH1_data_setup.RData")
+
+
+q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
+                        total =  length(adj_nMLH1.foci),
+                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
+                        above4 = sum(as.numeric(quality) <= 3, na.rm = TRUE ),
+                        #add percentages
+                        prop.lt5 = above4/total
+)
+length(q_cutoff_table$mouse)#116
+
+pass_mus_table <- q_cutoff_table[q_cutoff_table$above4 >= 25,] #30
+
+pass_mus_table.2 <- pass_mus_table[pass_mus_table$prop.lt5 >= .6,]#29
+
+OnEdge <-  q_cutoff_table[ ( q_cutoff_table$prop.lt5 >= .6) &
+                               (q_cutoff_table$above4 <= 23 ),]
+
+## write out list of mice
+
+

### 2019-05-10 draft_scripts/test4_pvalue.R

@@ -14,6 +14,7 @@ q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                         total =  length(adj_nMLH1.foci),
                         q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                         above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
+                        #add percentages
                         # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
 )
 #q_cutoff_table

### 2019-05-10 src/CommonFunc.R

@@ -3,12 +3,26 @@
 #
 
 
-#add mouse column to MLH1 df
-add_mouse <- function(dat){
+#add mouse for short mouse name
+add_mouse1 <- function(dat){
   dframe <- dat
   count =1
   for(i in dframe$Original.Name){
-    #print(i)
+    print(i)
+    templist= strsplit(i, split="_")[[1]]
+    c = paste(templist[1], templist[2],templist[3], sep = "_")
+    dframe$mouse[count] <- c
+    count= count +1
+  }
+  return(dframe)
+}
+
+#this add mouse, is based on file names
+add_mouse2 <- function(dat){
+  dframe <- dat
+  count =1
+  for(i in dframe$Original.Name){
+    print(i)
     templist= strsplit(i, split="_")[[1]]
     c = paste(templist[2], templist[3],templist[4], sep = "_")
     dframe$mouse[count] <- c
@@ -35,22 +49,21 @@ add_euth_date <- function(dat){
   dframe$euth.date <- as.Date(dframe$euth.date,format='%d%b%y')
   return(dframe)
 }
+
 #format the dates into standard format
 #mouse_list$DOB <- as.Date(mouse_list$DOB,format='%m/%d/%Y')
-
 add_age <- function(dat){
   dframe <- dat
   count =1
   for(i in dframe$Original.Name){
-    #after DOB and euth date are standardized
-    #substract
-    dframe$age.days[count] <- dframe$euth.date-dframe$DOB
+    #after DOB and euth date are standardized #substract
+    dframe$age.days <- dframe$euth.date-dframe$DOB
+    dframe$age.weeks <- difftime(dframe$euth.date,dframe$DOB,units='weeks')
     #gives age in days
     count= count +1
   }
   return(dframe)
 }
-
 
 
 #add sex to dataframe
@@ -79,6 +92,8 @@ add_strain <- function(dat){
              ifelse(grepl("_MSM_", dframe$Original.Name), "MSM",           
              ifelse(grepl("_PWD_", dframe$Original.Name), "PWD",
                 ifelse(grepl("_KAZ_", dframe$Original.Name), "KAZ",
+               ifelse(grepl("CZECH", dframe$Original.Name), "CZECH",         
+                       
                                                              
                    ifelse(grepl("_CAST_", dframe$Original.Name), "CAST",           
                       ifelse(grepl("_HMI_", dframe$Original.Name), "HMI",
@@ -86,11 +101,11 @@ add_strain <- function(dat){
[…]

### 2019-05-10 src/Figures.R

@@ -1,0 +1,17 @@
+---
+title: "Figures"
+author: "April Peterson"
+date: 
+output: pdf_document
+---
+
+```{r setup, include=FALSE}
+
+# Common Figures
+# input: RData file for repo
+# output: Figures 
+
+
+knitr::opts_chunk$set(echo = TRUE)
+```
+

### 2019-05-10 src/Func_addStrain.R

@@ -8,6 +8,7 @@ add_strain <- function(dat){
   dframe$strain <- ifelse(grepl("_WSB", dframe$Original.Name), "WSB", 
               ifelse(grepl("_G_",dframe$Original.Name), "G",
            ifelse(grepl("_LEW", dframe$Original.Name), "LEWES",
+                  
                                   
             ifelse(grepl("_MSM_", dframe$Original.Name), "MSM",           
            ifelse(grepl("_PWD_", dframe$Original.Name), "PWD",

### 2019-05-10 src/MLH1_strain_means1.R

@@ -1,0 +1,215 @@
+#R script for making plots of the distributions of MLH1 foci numbers
+#input: MLH1 data/ environment, mouse specific table
+#plots: mouse specific, strain averages
+
+#output: 2 sex-specific figures, 1 merged figure
+#<don't adjust male data, think about including BD's data>
+
+#..maybe there should be a R code for saving processing the data
+#but for now, just construct the make file when dataframe is pasted into the dir
+library(plyr)
+library(dplyr)
+library(ggplot2)
+
+# make sure these are 2se
+
+#############
+# LOAD DATA #
+#############
+#read in working environment that was already set
+setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+load(file="MLH1_data_setup.RData")
+
+###################
+# DATA TREATMENTS #
+###################
+#make sure the levels are in the correct order for the graph plotting
+AP_strain_table <- ddply(MLH1_data, c("strain", "sex"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+
+Lynn_CASTf_foci = c(20,21, 23, 25, 26, 26,26,27.5, 28, 28,28,33)
+cast_f_row = c("CAST", "female", 1, length(Lynn_CASTf_foci), round(mean(Lynn_CASTf_foci),3), 
+               round(var(Lynn_CASTf_foci),3), round(sd(Lynn_CASTf_foci),3), 
+               round(sd(Lynn_CASTf_foci)/sqrt(length(Lynn_CASTf_foci)),3 ), 
+               round(sd(Lynn_CASTf_foci) / (mean(Lynn_CASTf_foci) ), 3)  )
+
+AP_strain_table <- rbind(AP_strain_table, cast_f_row)
+
+AP_strain_table$subsp  <- ifelse(grepl("WSB", AP_strain_table$strain), "M.m. domesticus", 
[…]

### 2019-05-10 src/setup_data.R

@@ -58,14 +58,13 @@ for(i in 1:length(MLH1_data$mouse)){
   MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
 }
 
-
-source("src/Func_addCategory.R")
+source("src/CommonFunc.R")
+#source("src/Func_addCategory.R")
 MLH1_data <- add_category(MLH1_data)
 #set the order of categories (female, male) (cast, dom, musc)
-
-source("src/Func_addStrain.R")
+#source("src/Func_addStrain.R")
 MLH1_data <- add_strain(MLH1_data)
-source("src/Func_addSex.R")
+#source("src/Func_addSex.R")
 MLH1_data <- add_sex(MLH1_data)
 
 #add a column with male adjusted MLH1 values (+1 to all males)
@@ -158,8 +157,6 @@ BDMLH1_data <- subset(fullBD_MLH1_data, (Cross %in%  c("PANCEVO","RAT","CIM", "P
 
 #now that I have the mice of Beth's I want, remove the big BD df with F2s to save space.
 rm(fullBD_MLH1_data)
-
-
 
 
 

### 2019-05-10 src/theme.clean.R

@@ -1,0 +1,20 @@
+#
+# clean theme function for ggplot
+
+
+theme.clean <- function(){
+  theme_bw()+
+    theme(axis.text.x = element_text(size = 12, angle = 45, vjust = 1, hjust = 1),
+          axis.text.y = element_text(size = 12),
+          axis.title.x = element_text(size = 14, face = "plain"),             
+          axis.title.y = element_text(size = 14, face = "plain"),             
+          panel.grid.major.x = element_blank(),                                          
+          panel.grid.minor.x = element_blank(),
+          panel.grid.minor.y = element_blank(),
+          panel.grid.major.y = element_blank(),  
+          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
+          plot.title = element_text(size = 20, vjust = 1, hjust = 0.5),
+          legend.text = element_text(size = 12, face = "italic"),          
+          legend.title = element_blank(),                              
+          legend.position = c(0.9, 0.9))
+}

### 2019-05-10 doc/MixedModel_PnD_Report.Rmd
moved some more files

--- 2019-02-27 MixedModel_PnD_Report.Rmd
@@
 
+Goals; complete 2 Mixed Model for
+- house mouse
+- Mus (with outgroups)
+
 This current data set has `r length(MLH1_data$Random.Name)`. and `r length(unique( MLH1_data$mouse) )`  mice. (do these data include non-HM?)
@@
 - the sexes are matched (remove strains which don't have both sexes represented)
+- consider adding the stain most of the mouse data came from
 

### 2019-06-02 results/report.rmd
working on missing list

---
title: "MLH1_BivData_Report"
author: "April Peterson"
date: "May 30, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#set dir

#load data

```

goals for BivData Report

- count the number of observations from all categories
- identify best mice to make the subset for curated data

plots by categories. (which have longer SCs?)

```{r}

```

### 2019-06-02 src/setupData_MLH1repo.R

@@ -9,6 +9,7 @@
 library(plyr)
 library(dplyr)
 library(raster)#for cV
+library(ggplot2)
 
 ########################
 # Setup main dataframe #
@@ -74,7 +75,38 @@ MouseMetaData = read.csv("data/ALP_MouseMetadata.csv", header=TRUE )
 Image_mice_dirs <- list.files(path = "C:/Users/alpeterson7/Documents/Images")
 
 
-#matching list
+###########################
+# Construct Lists of Mice #
+###########################
+
+
+#Dissected, but not imaged
+
+#read in the dissection file
+Dissection.File = read.csv("data/Disections_53019.csv", header=TRUE )
+
+#cleaning up of Disection file
+#standarize column names
+#remove lines without mouse
+Dissection.File <- Dissection.File[!(is.na(Dissection.File$mouse)|Dissection.File$mouse==""),]
+
+#print the name with the date of the file somewhere
+
+Dissection.list <- Dissection.File$mouse
+#figure out a way to deal with the question marks, at least remove them temporarily
+
+#remove the pero mice?
+
+missing.mice <- unique(Dissection.File$mouse[(MLH1_data$mouse %in% Image_mice_dirs)])
+#compare to the original list and also mark the number/enrichment of X's
+#original_DF
+
+list2 <- mice_image_folders[(mice_image_folders %in% AP_mouse_table$mouse)]
+
+
+
+#Imaged, but not quantified
+
 list <- unique(MLH1_data$mouse[(MLH1_data$mouse %in% Image_mice_dirs)])
 list2 <- mice_image_folders[(mice_image_folders %in% AP_mouse_table$mouse)]
 #length(list2) #124, 

### 2019-06-03 results/report_6319.rmd
updated MLH1 results

---
title: "MLH1 Current Data Report"
author: "April Peterson"
date: "June 3, 2019"
output:
  pdf_document: default
---

```{r setup, echo=FALSE, include=FALSE}
#require("knitr")
#knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(dplyr) #this is a newer version of plyr
library(ggplot2)
#set dir
setwd("~./MLH1repo/")
#load data
load(file = "data/MLH1/MLH1_data_setup.RData")
source("src/CommonFunc_MLH1repo.R")
```

## Strain Level Plots

```{r strain.table, echo=FALSE}

AP_strain_table <- ddply(MLH1_data, c("strain", "sex"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)

AP_strain_table <- add_subsp(AP_strain_table)
#("black", "#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", "yellowgreen") )

#set the order for a factor
AP_strain_table$subsp<- factor(AP_strain_table$subsp,levels =c("Dom", "Musc",
                 "Cast","Spretus", "Spic", "Caroli", "other"), order=T )

AP_strain_table$strain<- factor(AP_strain_table$strain,levels =c("G", "LEW",
                        "WSB", "PWD","MSM","KAZ","CAST","HMI","SPIC","SPRET"), order=T )

#puts table in the right order
AP_strain_table <- AP_strain_table %>%
  arrange(strain, sex, subsp) %>%               # sort your dataframe, by the focal categories
  mutate(image.title = factor(strain)) #another category that you want the order to match

colnames(AP_strain_table) <- c("strain","sex","Nmice","Ncells","mean_co","var","sd","se",
                          "cV","subsp","extra")

AP_strain_table$mean_co <- as.numeric(AP_strain_table$mean_co)

MLH1_by_F_strain <- AP_strain_table[AP_strain_table$sex == "female", ]
MLH1_by_M_strain <- AP_strain_table[AP_strain_table$sex == "male", ]

#MLH1_by_F_strain <- MLH1_by_F_strain[-c(7),]

#re-adjust the mouse table
AP_mouse_table$subsp<- factor(AP_mouse_table$subsp,levels =c("Dom", "Musc",
                                                               "Cast","Spretus", "Spic", "Caroli", "other"), order=T )

AP_mouse_table$strain<- factor(AP_mouse_table$strain,levels =c("G", "LEW","WSB",
                                                               "PWD","MSM","KAZ","CAST","HMI","SPIC","SPRET"), order=T )

AP_mouse_table <- AP_mouse_table[AP_mouse_table$mouse != "11jan18_WxL.F1_m1",]

```

```{r strain.level.female, echo=FALSE}

man_x_fspace <- c(.35, .5, .75,
         2.2, 2.5, 2.8,
          3.6, 4, 4.5)

ff_plot <- ggplot(MLH1_by_F_strain, aes(y = as.numeric(mean_co), x=man_x_fspace, color=strain))+
  geom_point(size = 4) +
  scale_y_continuous(breaks = seq(20,32, by=2)) +
  coord_cartesian(ylim = c(20,32))+
  xlim(0,5)+
  geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
                   ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),size=1.2, width=0.1) +

  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", 
                              "red", "green","black", "grey"))+

   labs(x="", y= "Female MLH1 Foci") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line.y = element_line(colour = "black", size = 1.5), #legend.position="none",
        axis.ticks.y = element_line(colour = "black", size = 1.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=20),

        axis.text.y = element_text(size=15, face="bold"),
      axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank() ) +

   annotate("text", x = c(mean(man_x_fspace[1:3]), mean(man_x_fspace[1:3]), 
    mean(man_x_fspace[4:6]), mean(man_x_fspace[4:6]), 
    mean(man_x_fspace[7:9]), mean(man_x_fspace[7:9])),
          y = c(21, 20.5, 21, 20.5, 21, 20.5) ,
    label = c("M.m.","domesticus","M.m.","musculus","out","groups"),fontface = 'bold.italic', hjust = 0.5, size=5) +
  ggtitle("Female MLH1 Counts \n -/+ 2 SE")

  #annotate("segment", x= mean(man_x_fspace[1:3])-.3, xend=mean(man_x_fspace[1:3])+.3, y=21.5, yend=21.5, colour="black",size=1.5) +
   # annotate("segment", x=mean(man_x_fspace[4:6])-.3,xend=mean(man_x_fspace[4:6])+.3, y=21.5, yend=21.5, colour="black", size=1.5) +
 # annotate("segment", x=mean(man_x_fspace[7:9])-.3, xend=mean(man_x_fspace[7:9])+.3, y=21.5, yend=21.5, colour="black", size=1.5)

```

```{r strain.level.male, echo=FALSE, include=FALSE}

man_x_mspace <- c(.35, .5, .75, 
            1.5,2,2.3,
            3.1,3.4,
            4.5,4.8, 5)

mm_plot <- ggplot(MLH1_by_M_strain, aes(y = as.numeric(mean_co), x=man_x_mspace, color=strain))+ 
  geom_point(size = 4)+ 
#  ylim(18,32) +  
#  scale_y_continuous(breaks = seq(18, 34, by = 2)) +
  scale_y_continuous(breaks = seq(20,35, by=2)) +
  coord_cartesian(ylim = c(20,32))+
  xlim(0,5)+
  geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
                    ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),
                size=1.2, width=0.1)+

   #geom_errorbar(aes(ymin = as.numeric(mean_co) - as.numeric(se), 
  #                  ymax = as.numeric(mean_co)+ as.numeric(se),
   #   size=.5, width=.1))+
  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue",
                "coral1","#E69F00","red", "yellowgreen","green", "purple", "black") )+ #"olivedrab2", (HMI) ,"orchid" (SPRET)
  labs(x="", y= "Male MLH1 Foci") +

  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line.y = element_line(colour = "black", size = 1.5), #legend.position="none",
        axis.ticks.y = element_line(colour = "black", size = 1.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=20),
        axis.text.y = element_text(size=15, face="bold"),
  axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank()) +

  annotate("text", x = c(mean(man_x_mspace[1:3]), mean(man_x_mspace[1:3]), 
                         mean(man_x_mspace[4:6]), mean(man_x_mspace[4:6]),
                         mean(man_x_mspace[7:8]), mean(man_x_mspace[7:8]),
                          mean(man_x_mspace[9:10]), mean(man_x_mspace[9:10])),
           y = c(21,20.5, 21,20.5,21,20.5,21,20.5 ),  size=5,
    label = c("M.m.","domesticus","M.m.","musculus","M.m.","castaneus", "out", "groups"),hjust = 0.5, fontface='bold.italic') +

   annotate("segment", x= mean(man_x_mspace[1:3])-.3, xend=mean(man_x_mspace[1:3])+.3, y=21.5, yend=21.5, colour="black",
           size=1) +
  annotate("segment", x=mean(man_x_mspace[4:6])-.3,xend=mean(man_x_mspace[4:6])+.3, y=21.5, yend=21.5, colour="black",
           size=1) +
  annotate("segment", x=mean(man_x_mspace[7:8])-.3, xend=mean(man_x_mspace[7:8])+.3, y=21.5, yend=21.5, colour="black",
           size=1)+
  annotate("segment", x=mean(man_x_mspace[9:10])-.3, xend=mean(man_x_mspace[9:10])+.3, y=21.5, yend=21.5, colour="black",
         size=1)

```

```{r call.plots}

ff_plot

mm_plot
```

## Mouse Level Plots

plots by categories. (which have longer SCs?)

```{r mouse.level.female, echo=FALSE}

female.mice <- AP_mouse_table[AP_mouse_table$sex == "female",]

female.mice <- female.mice %>%
  arrange(strain,subsp) %>%         # sort your dataframe, by the focal categories
  mutate(image.title = factor(strain)) #another category that you want the order to match

# sort your dataframe, by the focal categories
female.mice$mouse <- factor(female.mice$mouse, levels=unique(female.mice$mouse))

female_mouse_plot <- ggplot(female.mice, aes(y =mean_co, x=mouse, color=strain))+geom_point()+
  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", 
                                                  "red", "green", "black", "grey"))+
  labs(x="", y= "Female MLH1 Foci")+ ggtitle("Female mouse MLH1 counts")+

  geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
                    ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),
                size=1, width=0.05)+
  theme(axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank())

```

```{r mouse.level.male, echo=FALSE}

male.mice <- AP_mouse_table[AP_mouse_table$sex == "male",]

male.mice <- male.mice %>%
  arrange(strain,subsp) %>%         # sort your dataframe, by the focal categories
  mutate(image.title = factor(strain)) #another category that you want the order to match

# sort your dataframe, by the focal categories
male.mice$mouse <- factor(male.mice$mouse, levels=unique(male.mice$mouse))

male_mouse_plot <- ggplot(male.mice, aes(y =mean_co, x=mouse, color=strain))+geom_point()+
  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", 
                              "red", "green", "black", "grey", "grey3"))+
  labs(x="", y= "Male MLH1 Foci")+ ggtitle("Male mouse MLH1 counts")+
    geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
                  ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),
              size=1, width=0.05)+
  theme(axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank())

```

```{r, echo=FALSE}

female_mouse_plot

male_mouse_plot

```

### 2019-06-03 src/setupData_MLH1repo.R

@@ -22,7 +22,7 @@ library(ggplot2)
 #setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 setwd("~./MLH1repo/")
 
-MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE )
+MLH1_data2 = read.csv("data/MLH1/AnonData.csv", header=TRUE ) #~400 more than the Rdata
 
 original_DF = MLH1_data
 #original_length <- length(MLH1_data$Original.Name)

### 2019-06-04 results/PWD.KAZ.BivData.Rmd
started Rmd file for reporting/comparing PWD KAZ metrics

---
title: "PWD.KAZ"
author: "April Peterson"
date: "June 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load BivData

#load MLH1 data

```

## Background and Premise

Comparison of PWD and KAZ Biv Data

PWD blah

CZECHII blah

KAZ blah

```{r pressure, echo=FALSE}

#MLH1 difference

```

These are the MLH1 counts

### BivData dataset

Goal:  Count the Bivdata for pwd and KAZ males

start determining the subset of data

```{r}

```

### 2019-06-04 src/setupData_MLH1repo.R

@@ -6,7 +6,7 @@
 #import MLH1 measures across data sets; AP, BD's and Cast female Koehler
 # 1. MLH1_data, tables: AP strains
 
-library(plyr)
+#library(plyr)
 library(dplyr)
 library(raster)#for cV
 library(ggplot2)
@@ -22,7 +22,7 @@ library(ggplot2)
 #setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 setwd("~./MLH1repo/")
 
-MLH1_data2 = read.csv("data/MLH1/AnonData.csv", header=TRUE ) #~400 more than the Rdata
+MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE ) #~400 more than the Rdata
 
 original_DF = MLH1_data
 #original_length <- length(MLH1_data$Original.Name)
@@ -80,11 +80,12 @@ Image_mice_dirs <- list.files(path = "C:/Users/alpeterson7/Documents/Images")
 # Construct Lists of Mice #
 ###########################
 
-
 #Dissected, but not imaged
 
 #read in the dissection file
 Dissection.File = read.csv("data/Disections_53019.csv", header=TRUE )
+#try to incorporate information from sp1 (the first stain)
+
 
 #cleaning up of Disection file
 #standarize column names
@@ -92,13 +93,11 @@ Dissection.File = read.csv("data/Disections_53019.csv", header=TRUE )
 Dissection.File <- Dissection.File[!(is.na(Dissection.File$mouse)|Dissection.File$mouse==""),]
 
 #print the name with the date of the file somewhere
-
 Dissection.list <- Dissection.File$mouse
-
 Dissection.list <- as.character(Dissection.list)
   
-str(Dissection.list)
-#figure out a way to deal with the question marks, at least remove them temporarily
+#figure out a way to incorporate the staining... to get a better idea of which slides 
+#to do next
 
 #remove the pero mice?
 
@@ -107,7 +106,7 @@ missing.mice <- unique(Dissection.File$mouse[(MLH1_data$mouse %in% Image_mice_di
 missing.mice2 <- unique(Dissection.File$mouse[(MLH1_data$mouse %in% Dissection.list)])
 
  
-missing_mice99 = subset(Dissection.list, !(Dissection.list %in% MLH1_data$mouse ) )
+missing_mice99 = subset(Dissection.File, !(Dissection.File$mouse %in% MLH1_data$mouse ) )
 length(missing_mice99)#306   #
 
 # mice in MLH1, 126

### 2019-06-11 results/PWD.KAZ.BivData.Rmd
tried starting initial analysis of KAZ and PWD data, addressing bugs and found BivDataSet is incomplete

@@
 
+setwd("~./MLH1repo/")
 #load BivData
+load(file="~./MLH1repo/data/MLH1_BivData.RData")
 
 #load MLH1 data
+#setwd("C:/Users/April/Desktop/MLH1repo")
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup.RData")
 
@@
 
-```{r pressure, echo=FALSE}
+```{r set.up.data, echo=FALSE}
 
-#MLH1 difference
+#extract just the relavant PWD and KAZ
+PWD.MLH1 <- MLH1_data[MLH1_data$category == "PWD male",]
+KAZ.MLH1 <- MLH1_data[MLH1_data$category == "KAZ male",]
+
+PWD.BivData.cells <- BivData_cells[BivData_cells$category == "PWD male",]
+PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]
+
+KAZ.BivData.cells <- BivData_cells[BivData_cells$category == "KAZ male",]
+KAZ.BivData <- fullBivData[fullBivData$category == "KAZ male",]
+
+PWD.KAZ_MLH1<- rbind(PWD.MLH1, KAZ.MLH1)
+PWD.KAZ_BivData<- rbind(PWD.BivData, KAZ.BivData)
 
@@
 Goal:  Count the Bivdata for pwd and KAZ males
-
 start determining the subset of data
 
-```{r}
+```{r mergeing.meta.data, echo=FALSE}
+meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
+meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375
+
+#find a way to remove the rows without the right format before setting as as.Date
+#trying to remove these the extra rows..uggh
+#yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]
+
+#yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
+#yy <- yy[subset(yy$DOB, !(yy$DOB == "e1")),]
+#i think subseting is messing up the dataframes
+meta.data.file99 <- meta.data.file2
+meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
+
+#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#meta.data.file2 <- add_euth_date(meta.data.file2) #euthdate 1st
+#meta.data.file2 <- add_age(meta.data.file2)  #age second
+##got it to work ... maybe should save adding it to the merged DF
+
+#reduce metadata down to 2 cols (mouse, DOB), merge this with MLH1
+MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
+colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
+MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
+
+PWD.KAZ_MLH1.imprv <- merge(PWD.KAZ_MLH1, MouseMetaData.wrkin)#don't apply the all parameters
+#report number with  and without
+
+#add euth date for calculating age
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+PWD.KAZ_MLH1.imprv <- add_euth_date(PWD.KAZ_MLH1.imprv)
+PWD.KAZ_MLH1.imprv <- add_age(PWD.KAZ_MLH1.imprv)
+
+#yesDOB.check <- PWD.KAZ_MLH1.imprv[!(is.na(PWD.KAZ_MLH1.imprv$DOB)),]
+#noDOB.check <- PWD.KAZ_MLH1.imprv[(is.na(PWD.KAZ_MLH1.imprv$DOB)),]#note, date type of data can't do the string matching!
+
+#total, KAZ 144, PWD 139
+#table(noDOB.check$category)#  no DOB, KAZ 67  PWD 133
+#table(yesDOB.check$category) #yes DOB, KAZ 77  PWD 6
+
+# I need to find the ages of more PWD males
 
 ```
+
+```{r}
+#BivData
+
+MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
+
+PWD.KAZ_BivData.imprv <- merge(PWD.KAZ_BivData, MouseMetaData.wrkin)
+
+table(fullBivData$category)# PWD 670, 
+#there are no KAZ male data
+
+```
+
+```{r}
+#distribution of ages
+
+histys <- ggplot(PWD.KAZ_MLH1.imprv, aes(age.weeks, fill=strain))+geom_density()
+
+```

### 2019-06-11 src/setupData_MLH1repo.R

@@ -57,6 +57,58 @@ MLH1_data <- add_subsp(MLH1_data)
 MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
 MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
 
+
+#add DOB
+#find a way to compare the list of AP_mice with Metadata, or list folders in Images
+MouseMetaData = read.csv("~./MLH1repo/data/ALP_MouseMetadata.csv", header=TRUE )
+
+meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
+meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%y")
+
+#reduce metadata down to 2 cols (mouse, DOB), merge this with MLH1
+MouseMetaData.wrkin <- MouseMetaData[,c(1,3)]
+MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
+
+colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
+MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
+
+#check how many times there is a duplicate mouse!
+length(which(duplicated(MouseMetaData.wrkin$mouse)))
+#currently 6 for the downloaded G doc data frame)
+
+#this is how you get the names of the duplicate mice
+MouseMetaData.wrkin$mouse[which(duplicated(MouseMetaData.wrkin$mouse))]
+
+#merge with MLH1 and BivData
+impMLH1_data <- merge(MLH1_data, MouseMetaData.wrkin, by.x = "mouse")#don't apply the all parameters
+
+length(which(duplicated(impMLH1_data$Random.Name)))#29 duplicated
+
+length(which(duplicated(impMLH1_data$fileName)))# 117 fileName
+#some of these are true duplicates, but some are the same image files that were quantified across multiple anon batches
+Quant2batches <- impMLH1_data[which(duplicated(impMLH1_data$fileName)),]#G, LEW, SPI females
+
+true.duplicates <- impMLH1_data[(which(duplicated(impMLH1_data$Random.Name))),]#but now these aren't duplicated...
+
[…]

### 2019-06-11 src/setupdata_BivData_MLH1repo.R

@@ -12,7 +12,7 @@ library(ggplot2)
 #add the melted foci here
 
 setwd("~./MLH1repo/")
-fullBivData = read.csv("data/BivData/CLEAN_MUS_FULL_BIVDATA_5.28.19.csv")
+fullBivData = read.csv("~./MLH1repo/data/BivData/CLEAN_MUS_FULL_BIVDATA_5.28.19.csv")
 #5402
 
 #check dissection list
@@ -22,12 +22,14 @@ Dissection_list = read.csv("data/mouseDissections.csv")
 fullBivData$Obj.ID <- paste(fullBivData$fileName, fullBivData$boxNumber, sep = "_")
 
 
-source("src/CommonFunc_MLH1repo.R")
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 fullBivData <- add_mouse(fullBivData)
 fullBivData <- add_category(fullBivData)
 fullBivData <- add_strain(fullBivData)
 fullBivData <- add_sex(fullBivData)
 fullBivData <- add_subsp(fullBivData)
+#this 5.28.19 dataset is missing so many categories
+
 
 #####################
 # REMOVE EXTRA DATA #
@@ -92,6 +94,9 @@ rawBivData$fileName[max(rawBivData$boxNumber)]
 RawDataTable <- as.data.frame(colSums(!is.na(BivData)) )
 #print .txt file
 #write.table(RawDataTable, "results/RawDataTable.txt", sep="\t")
+
+
+#print out list of mice that have Image folder, but are missing from the BivData
 
 
 

### 2019-06-11 doc/Main_Report.rmd
starting writing up the main results report

--- 2018-05-29 Main_Report.rmd
@@
 ---
-title: 'MLH1 Data Report'
+title: "MLH1 Data Report"
 author: "April Peterson"
-date: "29 may 2018"
-output: rmarkdown::github_document
+date: "`r format(Sys.time(), '%d %B %Y')`"
+output:
+  html_document:
+    df_print: paged
 ---
@@
 library(dplyr)
-setwd("C:/Users/alpeterson7/Documents/MLH1repo")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 #setwd("C:/Users/April/Desktop/MLH1repo")
-load(file="MLH1_data_setup.RData")
+load(file="data/MLH1/MLH1_data_setup.RData")
 ```
@@
 #### Research Description
+
 Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice abd related rodents.
@@
 
-```{r,echo=FALSE}
+```{r organizing.data, echo=FALSE}
 #set the order of another column, based on another variable (so that when)
@@
 
+```{r missing.mice, echo=FALSE}
+
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+setwd("~./MLH1repo/")
+#other lists
+
+Image_dirs <- list.files(path = "C:/Users/alpeterson7/Documents/Images")
+#write this out so that I can access it
+
+#remove the non-mouse folders
+Dissection.File = read.csv("~./MLH1repo/data/Disections_53019.csv", header = TRUE)
+Dissection.File <- Dissection.File[!(is.na(Dissection.File$mouse)|Dissection.File$mouse==""),]
+
+#print the name with the date of the file somewhere
+Dissection.list <- Dissection.File$mouse
+Dissection.list <- as.character(Dissection.list)
+
+#load metadata .. start decerning which mice are missing information... and which should be used
+Meta.Data.File = read.csv("~./MLH1repo/data/ALP_MouseMetadata.csv", header = TRUE)
+
+#remove file name, and original name .. this is just a mouse level DF
+Meta.Data.File <- Meta.Data.File[!(is.na(Meta.Data.File$mouse)|Meta.Data.File$mouse==""),]
+#make sure that the DOB is a date
+
+meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61019.csv", header = TRUE)
+meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%y")
+
+jl <- unique(Meta.Data.File$mouse[(Meta.Data.File$mouse %in% Dissection.File$mouse)])
+BD.w.meta <- Meta.Data.File[(Meta.Data.File$mouse %in% Dissection.File$mouse), ]
+
+#figure out a way to incorporate the staining... to get a better idea of which slides 
+#to do next, #remove the pero mice?
+
+#length(unique(MLH1_data$mouse) )
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+original_DF$fileName <- (original_DF$Original.Name)
+original_DF <- add_mouse(original_DF)
+#length(unique(original_DF$mouse) )
+#131 mice have been quant total, 124 mice have passed into the final version
+
+# make a list for the failed slides
+
+#Dissection.list - Image/dirs
+NonImaged.mice <- unique(Dissection.File$mouse[(Dissection.File$mouse %in% Image_dirs)])
+#231 mice dissected which need images
+
+#ImageFolder - MLH1.Quant
+NonQuant.mice <- unique(Image_dirs[(Image_dirs %in% MLH1_data$mouse )])
+#length(NonQuant.mice)
+#123 mice that need to get into anon-batch
+
+#make Dataframe to put into a table
+mouse.number.DF <- data.frame( category = 
+              c("Passed Quantified", "Total Quantified", "Non-Quantified", "Non-Imaged"),
+              number.of.mice = c(length(unique(MLH1_data$mouse)),
+                length(unique(original_DF$mouse) ),
+                        length(NonQuant.mice),
+                         length(NonImaged.mice)  )  )
+
+kable(mouse.number.DF)
+```
+
+This is the number of mice that are through different parts of the pathway or quantification process.
+
+```{r basic.numbers, echo=FALSE}
+
+kable(table(AP_mouse_table$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+
+kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+
+```
+
+These are the tables for the number of mice and cells which have been quantified. (what is in the 'other' category?)
+
 #### Initial Patterns from MLH1 distributions
 
-```{r first boxplots, echo=FALSE, fig.align='center'}
+```{r first.boxplots, echo=FALSE, fig.align='center'}
 #the below ordering works!! keep it
@@
 latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.
-```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
+
+```{r, echo=FALSE, message=FALSE, results='hide', warning=TRUE}
 ### to integrate 'pass'/fail
@@
 
+#quant_status (this isn't saving)
+#for(i in 1:length(MLH1_data$mouse)){
+#  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+#}
+
 #what is the purpose of this block
-df_list <- split(MLH1_data, (MLH1_data$category))
+#df_list <- split(MLH1_data, (MLH1_data$category))
 #remove sections which are in category but have no data
-df_list <- df_list[ -c(7,14,16,22) ]
-#7: PERC male, 12: KAZ female! (14:CAST female, fake data) 16:HMI female
+#df_list <- df_list[ -c(7,16,22) ]
+
+#7: PERC male, (14:CAST female, fake data) 16:HMI female, 22 other
+#
 
 #df list, is a list of df's which were split() from BIG df by category
-df_list 
-o=0
-i=0
+#df_list 
+#o=0
+#i=0
+
+#MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
 
 #make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-for(i in 1:(length(df_list))){
+#for(i in 1:(length(df_list))){
  # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                       mouse_nMLH1 = mean(adj_nMLH1.foci)
-  )
+#  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
+#  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
+ #                      mouse_nMLH1 = mean(adj_nMLH1.foci)
+#  )
 # this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-  for(o in 1:length(df_list[[i]]$Original.Name)){
-    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
-        }
-}
+#  for(o in 1:length(df_list[[i]]$Original.Name)){
+#    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == #sub_mouse_means$mouse) ]
+#        }
+#}
 
-plot_list = list()
-for (h in 1:length(df_list)) {
-  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-  pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
-  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") 
+#plot_list = list()
+#for (h in 1:length(df_list)) {
+#  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ #geom_jitter(aes(y= adj_nMLH1.foci, x= quality))
+  # +   geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5)
+#  + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL)
+
+# +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
   #theme for indiviual mice, 
   #+ theme(plot.background = element_rect(fill = "green"))
-  plot_list[[h]] =   pf
-}
+#  
+#  plot_list[[h]] =   pf
+#}
+
+#orig
+#plot_list = list()
+#for (h in 1:length(df_list)) {
+#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
+#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, #ymin=-100, ymax=100, alpha=0.005)
+
+#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
+#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") #+
+
+#    scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL)
+
+    #theme for indiviual mice, 
+  #+ theme(plot.background = element_rect(fill = "green"))
+#  plot_list[[h]] =   pf
+#}
+#can't change the alpha/transparency of the red rec
 
@@
 ```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE }
-geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
+#geom_hline(aes(yintercept = med, group = gr), colour = 'red')
+#
+#for (gg in 1:length(plot_list) ){
+#  show(plot_list[gg])
   #invisible(lapply(obj, function(x) plot(x,main="some plot")))
-}
+#}
 ```
@@
 ```{r, include=FALSE}
-passed_mice_data <- MLH1_data[MLH1_data$mouse %in% passed_mice, ]
-non_passed_mice_data <- MLH1_data[MLH1_data$mouse %in% not_passed, ]
+#passed_mice_data <- MLH1_data[MLH1_data$mouse %in% passed_mice, ]
+#non_passed_mice_data <- MLH1_data[MLH1_data$mouse %in% not_passed, ]
 
@@
 #turn this into boxplot and sort the category order, otherwise it's hard to any patterns
-sc <- qplot(mouse, nMLH1.foci,  colour = category, data = non_passed_mice_data, main="s", xlab="s", ylab="es")
+#sc <- qplot(mouse, nMLH1.foci,  colour = category, data = non_passed_mice_data, main="s", xlab="s", ylab="es")
 #sc + theme(legend.position="none")
 
-sc <- qplot(mouse, nMLH1.foci,  colour = category, data = passed_mice_data, main="s", xlab="s", ylab="es")
+#sc <- qplot(mouse, nMLH1.foci,  colour = category, data = passed_mice_data, main="s", xlab="s", ylab="es")
 #sc + theme(legend.position="none")

### 2019-06-20 results/report.rmd
commiting report files, this might not be the right folder

@@
 ---
-title: "MLH1_BivData_Report"
+title: "MLH1 Current Data Report"
 author: "April Peterson"
 date: "May 30, 2019"
-output: pdf_document
+output:
+  pdf_document: default
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup, echo=FALSE, include=FALSE}
+#require("knitr")
+#knitr::opts_chunk$set(echo = TRUE)
 
+library(plyr)
+library(dplyr) #this is a newer version of plyr
+library(ggplot2)
 #set dir
+setwd("~./MLH1repo/")
+#load data
+load(file = "data/MLH1/MLH1_data_setup.RData")
+source("src/CommonFunc_MLH1repo.R")
+```
 
-#load data
+## Strain Level Plots
+
+```{r strain.table, echo=FALSE}
+
+AP_strain_table <- ddply(MLH1_data, c("strain", "sex"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+
+AP_strain_table <- add_subsp(AP_strain_table)
+#("black", "#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", "yellowgreen") )
+
+#set the order for a factor
+AP_strain_table$subsp<- factor(AP_strain_table$subsp,levels =c("Dom", "Musc",
+                 "Cast","Spretus", "Spic", "Caroli", "other"), order=T )
+
+AP_strain_table$strain<- factor(AP_strain_table$strain,levels =c("G", "LEW",
+                        "WSB", "PWD","MSM","KAZ","CAST","HMI","SPIC","SPRET"), order=T )
+
+#puts table in the right order
+AP_strain_table <- AP_strain_table %>%
+  arrange(strain, sex, subsp) %>%               # sort your dataframe, by the focal categories
+  mutate(image.title = factor(strain)) #another category that you want the order to match
+
+colnames(AP_strain_table) <- c("strain","sex","Nmice","Ncells","mean_co","var","sd","se",
+                          "cV","subsp","extra")
+
+AP_strain_table$mean_co <- as.numeric(AP_strain_table$mean_co)
+
+MLH1_by_F_strain <- AP_strain_table[AP_strain_table$sex == "female", ]
+MLH1_by_M_strain <- AP_strain_table[AP_strain_table$sex == "male", ]
+
+#MLH1_by_F_strain <- MLH1_by_F_strain[-c(7),]
+
+#re-adjust the mouse table
+AP_mouse_table$subsp<- factor(AP_mouse_table$subsp,levels =c("Dom", "Musc",
+                                                               "Cast","Spretus", "Spic", "Caroli", "other"), order=T )
+
+AP_mouse_table$strain<- factor(AP_mouse_table$strain,levels =c("G", "LEW","WSB",
+                                                               "PWD","MSM","KAZ","CAST","HMI","SPIC","SPRET"), order=T )
+
+AP_mouse_table <- AP_mouse_table[AP_mouse_table$mouse != "11jan18_WxL.F1_m1",]
 
@@
 
-goals for BivData Report
+```{r strain.level.female, echo=FALSE}
 
-- count the number of observations from all categories
-- identify best mice to make the subset for curated data
+man_x_fspace <- c(.35, .5, .75,
+         2.2, 2.5, 2.8,
+          3.6, 4, 4.5)
+
+ff_plot <- ggplot(MLH1_by_F_strain, aes(y = as.numeric(mean_co), x=man_x_fspace, color=strain))+
+  geom_point(size = 4) +
+  scale_y_continuous(breaks = seq(20,32, by=2)) +
+  coord_cartesian(ylim = c(20,32))+
+  xlim(0,5)+
+  geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
+                   ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),size=1.2, width=0.1) +
+
+  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", 
+                              "red", "purple","black", "grey"))+
+
+   labs(x="", y= "Female MLH1 Foci") +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), 
+        axis.line.y = element_line(colour = "black", size = 1.5), #legend.position="none",
+        axis.ticks.y = element_line(colour = "black", size = 1.5),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=20),
+
+        axis.text.y = element_text(size=15, face="bold"),
+      axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank() ) +
+
+   annotate("text", x = c(mean(man_x_fspace[1:3]), mean(man_x_fspace[1:3]), 
+    mean(man_x_fspace[4:6]), mean(man_x_fspace[4:6]), 
+    mean(man_x_fspace[7:9]), mean(man_x_fspace[7:9])),
+          y = c(21, 20.5, 21, 20.5, 21, 20.5) ,
+    label = c("M.m.","domesticus","M.m.","musculus","out","groups"),fontface = 'bold.italic', hjust = 0.5, size=5)
+
+  #annotate("segment", x= mean(man_x_fspace[1:3])-.3, xend=mean(man_x_fspace[1:3])+.3, y=21.5, yend=21.5, colour="black",size=1.5) +
+   # annotate("segment", x=mean(man_x_fspace[4:6])-.3,xend=mean(man_x_fspace[4:6])+.3, y=21.5, yend=21.5, colour="black", size=1.5) +
+ # annotate("segment", x=mean(man_x_fspace[7:9])-.3, xend=mean(man_x_fspace[7:9])+.3, y=21.5, yend=21.5, colour="black", size=1.5)
+
+```
+
+```{r strain.level.male, echo=FALSE, include=FALSE}
+
+man_x_mspace <- c(.35, .5, .75, 
+            1.5,2,2.3,
+            3.1,3.4,
+            4.5,4.8, 5)
+
+mm_plot <- ggplot(MLH1_by_M_strain, aes(y = as.numeric(mean_co), x=man_x_mspace, color=strain))+ 
+  geom_point(size = 4)+ 
+#  ylim(18,32) +  
+#  scale_y_continuous(breaks = seq(18, 34, by = 2)) +
+  scale_y_continuous(breaks = seq(20,35, by=2)) +
+  coord_cartesian(ylim = c(20,32))+
+  xlim(0,5)+
+  geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
+                    ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),
+                size=1.2, width=0.1)+
+
+   #geom_errorbar(aes(ymin = as.numeric(mean_co) - as.numeric(se), 
+  #                  ymax = as.numeric(mean_co)+ as.numeric(se),
+   #   size=.5, width=.1))+
+  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue",
+                "coral1","#E69F00","red", "yellowgreen","green", "purple", "black") )+ #"olivedrab2", (HMI) ,"orchid" (SPRET)
+  labs(x="", y= "Male MLH1 Foci") +
+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), 
+        axis.line.y = element_line(colour = "black", size = 1.5), #legend.position="none",
+        axis.ticks.y = element_line(colour = "black", size = 1.5),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=20),
+        axis.text.y = element_text(size=15, face="bold"),
+  axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank()) +
+
+  annotate("text", x = c(mean(man_x_mspace[1:3]), mean(man_x_mspace[1:3]), 
+                         mean(man_x_mspace[4:6]), mean(man_x_mspace[4:6]),
+                         mean(man_x_mspace[7:8]), mean(man_x_mspace[7:8]),
+                          mean(man_x_mspace[9:10]), mean(man_x_mspace[9:10])),
+           y = c(21,20.5, 21,20.5,21,20.5,21,20.5 ),  size=5,
+    label = c("M.m.","domesticus","M.m.","musculus","M.m.","castaneus", "out", "groups"),hjust = 0.5, fontface='bold.italic') +
+
+   annotate("segment", x= mean(man_x_mspace[1:3])-.3, xend=mean(man_x_mspace[1:3])+.3, y=21.5, yend=21.5, colour="black",
+           size=1) +
+  annotate("segment", x=mean(man_x_mspace[4:6])-.3,xend=mean(man_x_mspace[4:6])+.3, y=21.5, yend=21.5, colour="black",
+           size=1) +
+  annotate("segment", x=mean(man_x_mspace[7:8])-.3, xend=mean(man_x_mspace[7:8])+.3, y=21.5, yend=21.5, colour="black",
+           size=1)+
+  annotate("segment", x=mean(man_x_mspace[9:10])-.3, xend=mean(man_x_mspace[9:10])+.3, y=21.5, yend=21.5, colour="black",
+         size=1)
+
+```
+
+```{r call.plots}
+
+ff_plot
+
+mm_plot
+```
+
+## Mouse Level Plots
 
@@
 
-```{r}
+```{r mouse.level.female, echo=FALSE}
+
+female.mice <- AP_mouse_table[AP_mouse_table$sex == "female",]
+
+female.mice <- female.mice %>%
+  arrange(strain,subsp) %>%         # sort your dataframe, by the focal categories
+  mutate(image.title = factor(strain)) #another category that you want the order to match
+
+# sort your dataframe, by the focal categories
+female.mice$mouse <- factor(female.mice$mouse, levels=unique(female.mice$mouse))
+
+female_mouse_plot <- ggplot(female.mice, aes(y =mean_co, x=mouse, color=strain))+geom_point()+
+  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", 
+                                                  "red", "green", "black", "grey"))+
+  labs(x="", y= "Female MLH1 Foci")+ ggtitle("Female mouse MLH1 counts")+
+
+  geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
+                    ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),
+                size=1, width=0.05)+
+  theme(axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank())
 
 ```
+
+```{r mouse.level.male, echo=FALSE}
+
+male.mice <- AP_mouse_table[AP_mouse_table$sex == "male",]
+
+male.mice <- male.mice %>%
+  arrange(strain,subsp) %>%         # sort your dataframe, by the focal categories
+  mutate(image.title = factor(strain)) #another category that you want the order to match
+
+# sort your dataframe, by the focal categories
+male.mice$mouse <- factor(male.mice$mouse, levels=unique(male.mice$mouse))
+
+male_mouse_plot <- ggplot(male.mice, aes(y =mean_co, x=mouse, color=strain))+geom_point()+
+  scale_color_manual(values=c("#56B4E9","cadetblue4","cadetblue","coral1","#E69F00", 
+                              "red", "purple", "black", "grey", "grey3"))+
+  labs(x="", y= "Male MLH1 Foci")+ ggtitle("Male mouse MLH1 counts")+
+    geom_errorbar(aes(ymin = as.numeric(mean_co)-(as.numeric(se)*2), 
+                  ymax = as.numeric(mean_co)+ (as.numeric(se)*2)),
+              size=1, width=0.05)+
+  theme(axis.title.x=element_blank(), axis.line.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x = element_blank())
+
+```
+
+```{r, echo=FALSE}
+
+female_mouse_plot
+
+male_mouse_plot
+
+```

### 2019-07-16 results/PWD.KAZ.BivData.Rmd
adding updated merged data for the PWD-KAZ analysis

@@
 date: "June 4, 2019"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
+library(plyr)
+library(ggplot2)
+
 setwd("~./MLH1repo/")
 #load BivData
-load(file="~./MLH1repo/data/MLH1_BivData.RData")
+#load(file="~./MLH1repo/data/MLH1_BivData.RData")
 
 #load MLH1 data
-#setwd("C:/Users/April/Desktop/MLH1repo")
-load(file="~./MLH1repo/data/MLH1/MLH1_data_setup.RData")
+#setwd("C:/Users/April/Desktop/MLH1repo/") #don't have two
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_62419.RData")
+#MLH1_data_setup_62419.RData
+#KAZ_male_data = read.csv("~./data/KAZ_BivData.csv")
+meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)
 
+#curBivData_18may19_PWD_m1 = read.csv("~./MLH1repo/results/curate_18may18_PWD_m1_BivData.txt", header = TRUE, sep = '\t')
+#curBivData_21may18_KAZ_m2 = read.csv("~./MLH1repo/results/curate_21may18_KAZ_m2_BivData.txt",header = TRUE, sep = '\t')
+
+#PWD set 4 mice, ~100 objs each)
+PWD_curate_BivData.full =
+  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
+#remove bad segmentation
+PWD_curate_BivData <- PWD_curate_BivData.full[PWD_curate_BivData.full$SC.pass == 1,]
+
+#KAZ 3 mice >200 objs each
+KAZ_curate_BivData.full =
+  read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE)
+KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]
+
+#load hand measures, get it to match the auto
+PWD_hand_measure_BivData.full =
+  read.csv("~./MLH1repo/data/BivData/PWD_male_hand_measures.csv", header = TRUE)
+
+#remove extra empty rows
+PWD_hand_measure_BivData.full <- PWD_hand_measure_BivData.full[!(is.na(PWD_hand_measure_BivData.full$fileName) | PWD_hand_measure_BivData.full$fileName == ""), ]
+
+#PWD_hand_measure_BivData.full$Foci1  <- PWD_hand_measure_BivData.full$foci.1.position
+#PWD_hand_measure_BivData.full$Foci2 <- PWD_hand_measure_BivData.full$medial.foci.position
+#PWD_hand_measure_BivData.full$Foci3 <- PWD_hand_measure_BivData.full$distal.foci.position
+#PWD_hand_measure_BivData.full$boxNumber <- PWD_hand_measure_BivData.full$manual.blobject.name
+
+PWD_hand_measure_BivData.full$Obj.ID <- paste(PWD_hand_measure_BivData.full$fileName, PWD_hand_measure_BivData.full$boxNumber, sep = "_")
+
+PWD_hand_measure_BivData.full$centromere_PER_Position <- PWD_hand_measure_BivData.full$centromere_ABS_Position / PWD_hand_measure_BivData.full$chromosomeLength
+
+PWD_hand_measure_BivData.full$SC.pass <- "1"
+PWD_hand_measure_BivData.full$foci.pass <- "1"
+PWD_hand_measure_BivData.full$curated.1pass.0fail. <- ""
+#PWD_hand_measure_BivData.full$curation.notes <- ""
+#PWD_hand_measure_BivData.full$hand.foci.count <- ""
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+PWD_hand_measure_BivData.full <- add_mouse(PWD_hand_measure_BivData.full)
+PWD_hand_measure_BivData.full <- add_category(PWD_hand_measure_BivData.full)
+PWD_hand_measure_BivData.full <- add_strain(PWD_hand_measure_BivData.full)
+PWD_hand_measure_BivData.full <- add_sex(PWD_hand_measure_BivData.full)
+
+#3 columns don't overlap
+#avergeRED, numberCrossOvers, 
+#manual.blobject.name,
+
+KAZ_hand_measure_BivData.full =
+  read.csv("~./MLH1repo/data/BivData/KAZ_male_hand_measures.csv", header = TRUE)
+
+KAZ_hand_measure_BivData.full <- KAZ_hand_measure_BivData.full[!(is.na(KAZ_hand_measure_BivData.full$fileName) | KAZ_hand_measure_BivData.full$fileName == ""), ]
+
+KAZ_hand_measure_BivData.full$boxNumber <- KAZ_hand_measure_BivData.full$manual.blobject.name
+
+KAZ_hand_measure_BivData.full$Obj.ID <- paste(KAZ_hand_measure_BivData.full$fileName, KAZ_hand_measure_BivData.full$boxNumber, sep = "_")
+
+KAZ_hand_measure_BivData.full$SC.pass <- "1"
+KAZ_hand_measure_BivData.full$foci.pass <- "1"
+KAZ_hand_measure_BivData.full$curated.1pass.0fail. <- ""
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+KAZ_hand_measure_BivData.full <- add_mouse(KAZ_hand_measure_BivData.full)
+KAZ_hand_measure_BivData.full <- add_category(KAZ_hand_measure_BivData.full)
+KAZ_hand_measure_BivData.full <- add_strain(KAZ_hand_measure_BivData.full)
+KAZ_hand_measure_BivData.full <- add_sex(KAZ_hand_measure_BivData.full)
+
+#not sure why I have a distinction of hand_measure and hand_measure.full
+#KAZ_hand_measure_BivData <- KAZ_hand_measure_BivData.full[KAZ_hand_measure_BivData.full$SC.pass == 1,]
+
+#manual blobject name -- boxNumber, Obj.ID 
+#  20jul15_18may15_PWD_m1_sp1_2.2_blob1
 ```
@@
 
-Comparison of PWD and KAZ Biv Data
+### BivData dataset
 
-PWD blah
+Goal:  Count the Bivdata for pwd and KAZ males
+start determining the subset of data
 
-CZECHII blah
-
-KAZ blah
-
-```{r set.up.data, echo=FALSE}
+```{r, echo=FALSE}
+#Curated data
+#merge PWD and KAZ BivData
 
@@
 
-These are the MLH1 counts
+what are the distributions of SC lengths?
 
-### BivData dataset
+```{r cell.level.tables, echo=FALSE}
 
-Goal:  Count the Bivdata for pwd and KAZ males
-start determining the subset of data
+#make cell level tables
+#id the 18-20 biv cells
+
+#estimate which cells can be used to make grouped pools
+
+```
+
+### SC Length Analysis
+
+Make plots for SC lengths
+
+```{r SC.lengths}
+library(ggplot2)
+#SC lenghts
+
+#strange number of low chrm lengths
+#hist(SC.passed.PWD$chromosomeLength)
+
+SC.PWD <- ggplot(data = PWD_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,150))+ggtitle("PWD SC lengths")
+#the SC range is 45 - 150 (for PWD)
+
+SC.KAZ <- ggplot(data = KAZ_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,150))+ggtitle("KAZ SC lengths")
+#can I add the means?
+
+SC.PWD
+SC.KAZ
+```
+
+```{r cell.tables}
+#add total COs
+KAZ_BivData_cells <- ddply(KAZ_curate_BivData, c("fileName"), summarise,
+
+                    boxs.IDd = max(boxNumber),
+                    nboxes_passed = length(boxNumber),
+                    pass_rate = nboxes_passed/boxs.IDd,
+
+                    totalCO = sum(numberCrossOvers),     
+                    avCO.per.CHRM = mean(numberCrossOvers),
+                    varCO.per.CHRM = var(numberCrossOvers),
+
+                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
+                    varSC.intensity = var(avergeRED, na.rm = TRUE),
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength),
+
+                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
+                    varCOdensity = var(chromosomeLength /numberCrossOvers)
+                            #av CO density
+                    #change the mean list for column numbers
+
+)
+
+cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]
+
+write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)
+
+#PWD version
+#add total COs
+PWD_BivData_cells <- ddply(PWD_curate_BivData, c("fileName"), summarise,
+
+                    boxs.IDd = max(boxNumber),
+                    nboxes_passed = length(boxNumber),
+                    pass_rate = nboxes_passed/boxs.IDd,
+
+                    totalCO = sum(numberCrossOvers),     
+                    avCO.per.CHRM = mean(numberCrossOvers),
+                    varCO.per.CHRM = var(numberCrossOvers),
+
+                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
+                    varSC.intensity = var(avergeRED, na.rm = TRUE),
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength),
+
+                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
+                    varCOdensity = var(chromosomeLength /numberCrossOvers)
+                            #av CO density
+                    #change the mean list for column numbers
+
+)
+
+PWDcells2curate <- PWD_BivData_cells[PWD_BivData_cells$pass_rate > .65,]
+
+write.table(PWDcells2curate, "~./MLH1repo/results/PWD.cell.curate.csv", sep=",", row.names = FALSE)
+
+```
+
+From the cell level tables -- make lists of cells with ~75% of bivalents quantified. These will be manually verified and the missing chrms will be measured and supplemented.
+
+```{r PWD.merge.manual.auto, echo=FALSE}
+#merging manual measures and auto measures
+#KAZ_hand_measure_BivData
+#PWD_hand_measure_BivData
+
+#.1 match the colnames,  #merge, partial match
+#colnames(PWD_curate_BivData)
+#colnames(PWD_hand_measure_BivData.full)
+
+#merg4.PWD_curate_BivData <-    #remove "avergeRED", numberCrossOvers
+#merg4.PWD_hand_measure_BivData.full <- #manual.blobject.name
+
+#auto.curate
+auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
+new.DF <- PWD_curate_BivData[ , auto.c_names, drop = TRUE]
+
+#manual
+manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position","SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
+new.DF2 <- PWD_hand_measure_BivData.full[ , manual_names, drop = TRUE]
+
+#2. merge the 2 DFs
+PWD.auto.man_BivData <- rbind(new.DF, new.DF2)
+
+#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
+#make better variable name
+PWD.auto.man_BivData_cells <- ddply(PWD.auto.man_BivData, c("fileName"), summarise,
+                    number = length(Obj.ID),
+                    #make the rank or
+                     maxChrm_length = max(chromosomeLength),
+                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]
+
+)
+
+#Isolate cells with all autosomes measured
+PWDfull.cell.list <- PWD.auto.man_BivData_cells[PWD.auto.man_BivData_cells$number > 17,]
+PWDfull.cell.list <- PWDfull.cell.list[PWDfull.cell.list$number < 21,]#remove the too many cell
+
+#this gives the rank to each biv within each cell  
+PWD.Bivs.w.rank <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+#there are instances of ties,
+
+#this does something weird
+#PWD.auto.man_BivData$rank1 <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+
+PWD.auto.man_BivData <- PWD.auto.man_BivData[ -c(19:25) ]
+
+#pull all biv data for the full cell data
+PWD_whole.cell.BivData <- PWD.Bivs.w.rank[PWD.Bivs.w.rank$fileName %in% PWDfull.cell.list$fileName,]
+#ok these should be full cells
+hist(hh2$rank, breaks =21)
+#there are 2x as many 0's bc 1 and 0 are binned together
+#update all hand.foci.counts
+#y <- droplevels(y)
+
+length.by.class <- ggplot(data=hh2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=hand.foci.count, shape=mouse))+ggtitle("PWD full cell SC lengths")
+#most cells are from 1 mouse
+
+DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName %in% full.cell.list$fileName,]
+DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName != "20jul15_18may15_PWD_m1_sp1_17_rev",]
+
+qq <- ggplot(data=DF.w.rank2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=rank, shape=mouse))+ggtitle("PWD full cell SC lengths")
+#most cells are from 1 mouse
+
+#take the shortest 5? (lower 25%)
+#take the shortest 3? (seems to be natural cutoff, lowest 3 are similar)
+
+#bin.long.PWD <- DF.w.rank[DF.w.rank$rank]
+#can't do long yet, bc the coding for XY is inconsistant!
+
+bin.short.PWD <- DF.w.rank2[DF.w.rank2$rank < 5,]
+bin.short.PWD <- bin.short.PWD[bin.short.PWD$fileName %in% full.cell.list$fileName,]
+
+kk <- ggplot(data=bin.short.PWD, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color =Foci1, shape=mouse))+ggtitle("short bins")
+
+#short bin patterns vs long bins / total?
+#eventually I want to compare NORM FOCI POSITIONS patterns of these bins across PWD and KAZ
+#merging is a little too confusing... 
+#just parse down and do rbind?
+m1 <- merge(PWD_curate_BivData, PWD_hand_measure_BivData.full, by.x = Obj.ID)
+
+#KAZ_curate_BivData
+#KAZ_hand_measure_BivData.full
+
+hist(hh2$rank, breaks =21)
+```
+
+Asigned rank to bivalent measures from the whole.cell dataset. I should try to use these measures to see if I can guess the ranks of 'loose' bivalent measures.
+I also noted that the shortest 3 seem to cluster in PWD. Maybe I should try doing cluster analysis on these. 
+I should also add rbar values to these measures. (I might have to reformat the function)
+
+```{r KAZ.manual.merge}
+#KAZ_hand_measure_BivData
+#PWD_hand_measure_BivData
+#KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]
+
+#1 match the colnames,  #merge, partial match
+colnames(KAZ_curate_BivData)
+colnames(KAZ_hand_measure_BivData.full)
+
+KAZ_hand_measure_BivData.full$centromere_PER_Position <- KAZ_hand_measure_BivData.full$centromere_ABS_Position / KAZ_hand_measure_BivData.full$chromosomeLength
+
+KAZ_hand_measure_BivData.full$curation.notes <- KAZ_hand_measure_BivData.full$notes
+
+#auto; avergeRED, numberCrossOvers, 20
+#manual; 15
+
+#auto.curate
+KAZ.auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
+new.DF <- KAZ_curate_BivData[ , KAZ.auto.c_names, drop = TRUE]
+
+#manual
+KAZ.manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position", "SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
+new.DF2 <- KAZ_hand_measure_BivData.full[ , KAZ.manual_names, drop = TRUE]
+
+#2. merge the 2 DFs
+KAZ.auto.man_BivData <- rbind(new.DF, new.DF2)
+
+#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
+#make better variable name
+KAZ.auto.man_BivData_cells <- ddply(KAZ.auto.man_BivData, c("fileName"), summarise,
+                    number = length(Obj.ID),
+                    #make the rank or
+                     maxChrm_length = max(chromosomeLength),
+                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]
+
+)
+
+#Isolate cells with all autosomes measured
+KAZfull.cell.list <- KAZ.auto.man_BivData_cells[KAZ.auto.man_BivData_cells$number > 17,]
+KAZfull.cell.list <- KAZfull.cell.list[KAZfull.cell.list$number < 21,]#remove the too many cell
+
+#this gives the rank to each biv within each cell  
+KAZ.Bivs.w.rank <- ddply(KAZ.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+#there are instances of ties,
+
+KAZ_whole.cell.BivData <- KAZ.Bivs.w.rank[KAZ.Bivs.w.rank$fileName %in% KAZfull.cell.list$fileName,]
+#ok these should be full cells
+hist(KAZ_whole.cell.BivData$rank, breaks =25)
+KAZ_whole.cell.BivData <- droplevels(KAZ_whole.cell.BivData)
+table(KAZ_whole.cell.BivData$fileName, KAZ_whole.cell.BivData$rank)
+
+KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")
+
+```
+
+KAz have more 0CO. Most of the NA's for CO count are from auto- where I was whisy-washy during curation.
+
+```{r stat.test.SC.length, echo=FALSE}
+#can't cleanly merge the curated data since the rows and headers are different
+merged.cur <- rbind(curBivData_PWD_m1, curBivData_KAZ_m2)
+
+SCs <- ggplot(data = , aes(y, x=mice))+geom_point
+
+SC.test <- t.test()
+
+```
+
+```{r IDF, echo=FALSE}
+
+#isolate the 2COs
+#total Bivdata
+KAZ.2COs <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$hand.foci.count == 2,]
+PWD.2COs <- PWD.auto.man_BivData[PWD.auto.man_BivData$hand.foci.count == 2,]
+
+all.2CO <- rbind(KAZ.2COs,PWD.2COs)
+#all.2CO <- ddply(all.2CO, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+#all.2CO <- all.2CO[ -c(24) ]
+
+all.2CO$Foci1.PER <- all.2CO$Foci1 / all.2CO$chromosomeLength
+all.2CO$Foci2.PER <- all.2CO$Foci2 / all.2CO$chromosomeLength
+all.2CO$Foci3.PER <- all.2CO$Foci3 / all.2CO$chromosomeLength
+
+#call IFD function on this DF
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+all.2CO <- add_IFD(all.2CO)
+all.2CO <- droplevels(all.2CO)
+all.2CO <- all.2CO[!(is.na(all.2CO$fileName) | all.2CO$fileName == ""), ]
+
+all.2CO$mouse <- droplevels(all.2CO$mouse)
+
+IFDs.ABS <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_ABS))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("raw IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 90, hjust = 1) )
+
+IFDs.PER <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("norm IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+theme(legend.position="none")
+
+#whole cell comparison (use rank)
+#KAZ_whole.cell.BivData, #PWD_whole.cell.BivData
+
+#limit to mice with most observations
+
+whole.cell.2CO <- rbind(KAZ_whole.cell.BivData,PWD_whole.cell.BivData)
+
+whole.cell.2CO$Foci1.PER <- whole.cell.2CO$Foci1 / whole.cell.2CO$chromosomeLength
+whole.cell.2CO$Foci2.PER <- whole.cell.2CO$Foci2 / whole.cell.2CO$chromosomeLength
+whole.cell.2CO$Foci3.PER <- whole.cell.2CO$Foci3 / whole.cell.2CO$chromosomeLength
+
+#call IFD function on this DF
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+whole.cell.2CO <- add_IFD(whole.cell.2CO)
+whole.cell.2CO <- droplevels(whole.cell.2CO)
+whole.cell.2CO <- whole.cell.2CO[!(is.na(whole.cell.2CO$fileName) | whole.cell.2CO$fileName == ""), ]
+
+IFDs.ABS.whole.cell <- ggplot(data=whole.cell.2CO, aes(x=mouse,y=IFD1_ABS))+geom_jitter(width = 0.25, aes(color=as.factor(rank) ))+ggtitle("Whole cell IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )
+
+IFDs.PER.whole.cell <- ggplot(data=whole.cell.2CO, aes(x=mouse,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(rank) ))+ggtitle("Whole Cell Norm IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )
+
+```
+
+```{r IFD.SC, echo=FALSE}
+#compare the realtionship of SC length and
+
+values.ABS.all <- all.2CO %>% 
+    group_by(strain) %>% 
+    do({
+      mod = lm(IFD1_ABS ~ chromosomeLength, data = .)
+      data.frame(Intercept = coef(mod)[1],
+                 Slope = coef(mod)[2])
+    })
+
+SC.IFD <- ggplot(data=all.2CO, aes(x=chromosomeLength,y=IFD1_ABS))+geom_point( aes(color=as.factor(mouse) ))+
+  ggtitle("SC by IFD length")+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )+ 
+  geom_smooth(method="lm")#+
+  #geom_text(values.ABS.all, aes(x = 75, y = 100, label = n), parse = TRUE, inherit.aes=FALSE)
+
+SC.IFD_PER <- ggplot(data=all.2CO, aes(x=chromosomeLength,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(mouse) ))+
+  geom_abline(intercept = 0.25)+
+    geom_smooth(method="lm")+
+  ggtitle("SC by nrm IFD length")+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )
+
+```
 
 ```{r mergeing.meta.data, echo=FALSE}
-meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
+#meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
+#meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)
+
 meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375
@@
 #yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]
-
 #yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
@@
 #i think subseting is messing up the dataframes
-meta.data.file99 <- meta.data.file2
+
 meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
@@
 ##got it to work ... maybe should save adding it to the merged DF
+meta.data.file2 <- add_category(meta.data.file2)
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+meta.data.file2 <- add_euth_date(meta.data.file2)
+meta.data.file2 <- add_age(meta.data.file2)
+
+pwd.meta <- meta.data.file2[meta.data.file2$category == "PWD male",]
+
+#pwd.meta <- add_euth_date(pwd.meta)
+#pwd.meta <- add_age(pwd.meta)
+
+kaz.meta <- meta.data.file2[meta.data.file2$category == "KAZ male",]
+
+#kaz.meta <- add_mouse(kaz.meta)
+#kaz.meta <- add_euth_date(kaz.meta)
+county =1
+for(i in kaz.meta$mouse){
+    templist= strsplit(i, split="_")[[1]]
+    euth <- templist[1]
+    #figure out date conversion
+    kaz.meta$euth.date2[county] <- as.character(euth)
+
+    #kaz.meta$euth.date[county] <- as.Date(kaz.meta$euth.date[county], format='%d%b%y')
+    county= county +1
+}
+
+kaz.meta$euth.date <- as.Date(kaz.meta$euth.date2, format='%d%b%y')
+
+kaz.meta <- add_age(kaz.meta)
 
@@
 ```
+##########################
+# MELTED CHRM1 POSITIONS #
+##########################
 
-```{r}
+```{r PWD.melt.foci.data, echo=FALSE}
+library(reshape2)
+
+newPWD_curate_BivData =
+  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
+
+#update these dataframes!
+#isolate the passing foci 
+curBivData_PWD.good.foci  <- newPWD_curate_BivData[newPWD_curate_BivData$foci.pass == 1,]
+
+#create normalized foci values!
+curBivData_PWD.good.foci$nrm.F1 <- curBivData_PWD.good.foci$Foci1 / curBivData_PWD.good.foci$chromosomeLength
+curBivData_PWD.good.foci$nrm.F2 <- curBivData_PWD.good.foci$Foci2 / curBivData_PWD.good.foci$chromosomeLength
+curBivData_PWD.good.foci$nrm.F3 <- curBivData_PWD.good.foci$Foci3 / curBivData_PWD.good.foci$chromosomeLength
+
+#melting the DataFrame is needed to ploting the normalized foci positions
+
+#do not include cols for IFD, or ABS.foci
+curBivData_PWD.melt <- melt( curBivData_PWD.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","nrm.F1","nrm.F2","nrm.F3",  "mouse", "category",   "strain" , "sex")] ,
+
+                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
+              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","mouse", "category",   "strain" , "sex"))
+
+colnames(curBivData_PWD.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count", "mouse", "category",   "strain" , "sex",
+                                       #name for the 2 melted cols
+                                      "Foci.Type.nrm", "Foci.nrm.pos")
+
+#remove NAs
+curBivData_PWD.melt <- curBivData_PWD.melt[!(is.na(curBivData_PWD.melt$Foci.nrm.pos) | curBivData_PWD.melt$Foci.nrm.pos == ""), ]
+
+table(newPWD_curate_BivData$fileName)
+table(newPWD_curate_BivData$mouse) #4 mice, ~100 each mouse
+
+```
+
+```{r curated, echo=FALSE}
+curBivData_KAZ.good.foci <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
+#curated objects
+KAZ_m2_curate_bivData =
+  read.csv("~./MLH1repo/data/BivData/curated_21may18_KAZ_m2_BivData.csv", header = TRUE)
+
+KAZ_m2_curate_bivData$nrm.F1.pos <- KAZ_m2_curate_bivData$Foci1.pos / KAZ_m2_curate_bivData$chromosomeLength
+KAZ_m2_curate_bivData$nrm.F2.pos <- KAZ_m2_curate_bivData$Foci2.pos / KAZ_m2_curate_bivData$chromosomeLength
+KAZ_m2_curate_bivData$nrm.F3.pos <-   KAZ_m2_curate_bivData$Foci3.pos / KAZ_m2_curate_bivData$chromosomeLength
+
+#remove those that don't pass
+KAZ_m2_curate_bivData_SC_pass <- KAZ_m2_curate_bivData[KAZ_m2_curate_bivData$SC.pass == 1,]
+#what else...
+
+table(KAZ_m2_curate_bivData_SC_pass$fileName)
+
+#melt the foci positions, best thing to compare for cells that don't have all the bivalents
+#ggplots
+
+show.plot <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=chromosomeLength, x=fileName))+geom_point()
+
+show.plot2 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=Foci1.pos, x=hand.foci.count))+geom_jitter()
+
+show.plot99 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=nrm.F1.pos, x=hand.foci.count))+geom_jitter()+
+  ggtitle("Normalized foci positions")
+
+count.hist <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(hand.foci.count))+geom_histogram()
+
+```
+
+```{r BivData.set, echo=FALSE}
+#goal load and set up of the BivData Dataframe
+
 #BivData
+# load more bivData
+KAZ_bivData = read.csv("~./MLH1repo/data/KAZ_BivData.csv", header = TRUE)
+
+P_fileNames <- KAZ_bivData$fileName[(grep('p_rev', KAZ_bivData$fileName))]
+length(P_fileNames)#187
+
+KAZ_bivData <- KAZ_bivData[!(KAZ_bivData$fileName %in% P_fileNames), ]#3642
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+KAZ_bivData <- add_mouse(KAZ_bivData)
+KAZ_bivData <- add_strain(KAZ_bivData)
+KAZ_bivData <- add_sex(KAZ_bivData)
+KAZ_bivData <- add_category(KAZ_bivData)
+
+#dup_list= c()
+#for(i in KAZ_bivData$fileName){  #change the file name header,  
+#  # print(i)
+#  ifelse(grepl("\\.[0-9]_rev", i),
+#         (dup_list = c(dup_list, i)), NA)
+#}
+#MultiCells = c()
+#for(j in dup_list){
+#  stringLIST = strsplit(j, split="\\.")[[1]]
+#  new_string = paste(stringLIST[[1]][1], "rev", sep="_")
+#  MultiCells = c(MultiCells, new_string)
+#}
+#MultiCells = unique(MultiCells)
+#multi_cell_num = length(MultiCells)
+
+#test this part of the function
+#KAZ_bivData <- KAZ_bivData[!(KAZ_bivData$fileName %in% MultiCells), ]#35 removed from KAZ data
 
@@
 
+#merge PWD and KAZ (WHAT ABOUT THE foci?)
+#melt the BivData?
+```
+
+```{r curatedBivData}
+
+curBivData_KAZ = read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE, sep = ',')
+#curBivData_PWD = read.csv("~./MLH1repo/data/BivData/curate_PWD_BivData.csv", header = TRUE, sep = ',')
+
+#make normalized
+
+curBivData_KAZ$nrm.F1.pos <- curBivData_KAZ$Foci1 / curBivData_KAZ$chromosomeLength
+curBivData_KAZ$nrm.F2.pos <- curBivData_KAZ$Foci2 / curBivData_KAZ$chromosomeLength
+curBivData_KAZ$nrm.F3.pos <-   curBivData_KAZ$Foci3 / curBivData_KAZ$chromosomeLength
+
+curBivData_PWD$nrm.F1.pos <- curBivData_PWD$Foci1 / curBivData_PWD$chromosomeLength
+curBivData_PWD$nrm.F2.pos <- curBivData_PWD$Foci2 / curBivData_PWD$chromosomeLength
+curBivData_PWD$nrm.F3.pos <-   curBivData_PWD$Foci3 / curBivData_PWD$chromosomeLength
+
+#table / what are the numbers of observations
+table(curBivData_PWD$fileName)
+#biv measures per cell
+
+SC.passed.PWD <- curBivData_PWD[curBivData_PWD$SC.pass == 1,]#299 -> 208
+table(SC.passed.PWD$fileName)
+
+SC.passed.KAZ <- curBivData_KAZ[curBivData_KAZ$SC.pass == 1,]#299 -> 208
+
+```
+
+```{r meltKAZ, echo=FALSE}
+
+#melt KAZ
+#isolate the passing foci 
+curBivData_KAZ.good.foci  <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
+#colnames(curBivData_KAZ)
+
+#melting the DataFrame is needed to ploting the normalized foci positions
+
+#do not include cols for IFD, or ABS.foci
+curBivData_KAZ.melt <- melt( curBivData_KAZ.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,
+
+                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
+              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","mouse", "category",   "strain" , "sex"))
+
+colnames(curBivData_KAZ.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count", "mouse", "category",   "strain" , "sex",
+                                       #name for the 2 melted cols
+                                      "nrmFoci.Type", "nrmFoci.pos")
+
+#remove NAs
+curBivData_KAZ.melt <- curBivData_KAZ.melt[!(is.na(curBivData_KAZ.melt$nrmFoci.Type) | curBivData_KAZ.melt$nrmFoci.pos==""), ]
+
+```
+
+```{r plot}
+
+library(ggplot2)
+library(plyr)
+
+nPWD.obs <- ddply(.data=curBivData_PWD.melt, 
+                 .(hand.foci.count),
+                 summarize, 
+                 n=paste("n =", length(unique(Obj.ID))))
+
+#normalized foci plots
+PWD.plot.nrm <- ggplot(curBivData_PWD.melt,
+              aes(x=Foci.nrm.pos, fill=Foci.Type.nrm) ) + 
+  geom_histogram(alpha= 0.5,stat = "bin", position= "identity")+
+ facet_wrap(~ hand.foci.count, scales="free", ncol=1)+ggtitle("norm foci position PWD")+
+
+   geom_text(data=nPWD.obs, aes(x=.5, y=7, label=n), 
+            colour="black", inherit.aes=FALSE, parse=FALSE) + xlim(c(0,1)) + ylim(c(0,20))
+
+#  scale_fill_manual(values=c("green4", "greenyellow")) +
+#can I put the number of observations in this?
+
+#make a sample dataset for KAZ
+#uni<- unique(curBivData_KAZ.melt$Obj.ID)#careful not to confuse cell names with obj.ID this is the number of cells! (not bivs)
+#subSample.KAZ <- sample(uni, 68, replace = FALSE)
+#subSample.KAZ.DF <- curBivData_KAZ.melt[curBivData_KAZ.melt$Obj.ID %in% subSample.KAZ,]
+
+nKAZ.obs <- ddply(.data=curBivData_KAZ.melt, 
+                 .(hand.foci.count),
+                 summarize, 
+                 n=paste("n =", length(unique(Obj.ID))))
+
+#urBivData_KAZ.melt[curBivData_KAZ.melt$hand.foci.count < 3,]
+
+KAZ.plot.nrm    <- ggplot(curBivData_KAZ.melt,
+              aes(x=nrmFoci.pos, fill=nrmFoci.Type) ) +
+
+  geom_histogram(alpha= 0.5,stat = "bin", position= "identity")+
+ facet_wrap(~ hand.foci.count, scales="free", ncol=1)+ggtitle("norm foci position KAZ")+
+   geom_text(data=nKAZ.obs, aes(x=.5, y=7, label=n), 
+            colour="black", inherit.aes=FALSE, parse=FALSE)+xlim(c(0,1)) + ylim(c(0,20))
+
 ```
@@
 ```{r}
+#goal, examine the ages of mice being compared
 #distribution of ages
@@
 ```
+
+```{r first.analysis, echo=FALSE}
+#initial analysis of the BivData
+
+Kaz.biv.plot <- ggplot(data = KAZ_bivData, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()
+
+```
+
+```{r make.curate.list, echo=FALSE}
+
+#load/read in fullBivdata file
+newPWD_BivData = read.csv("~./MLH1repo/data/BivData/PWD_BivData_7.9.19.csv", header = TRUE)#740
+#filter by centromere
+newPWD <- newPWD_BivData[newPWD_BivData$centromere_PER_Position < 0.2,] #665
+#add Obj.ID
+newPWD$Obj.ID <- paste(newPWD$fileName, newPWD$boxNumber, sep = "_")
+#write to file for making manual notes
+write.table(newPWD, "~./MLH1repo/results/newPWD_7.9.19.txt", sep="\t", row.names = FALSE)
+#think about adding ~6 more columns for the curation process
+```
+
+```{r KAZ.curate.list, echo=FALSE}
+#write out list for initial quality control and subsetting
+
+#remove centromere_PER > .2
+
+KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123
+
+#remove chrmLength > 150? 200
+KAZ.Biv.curate.list <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$chromosomeLength < 300, ]#not completely sure 300 is probably the max length is the best length
+#3117
+
+table(KAZ.Biv.curate.list$mouse)
+#7 mice: 11jan18_KAZ_m4(9w), 1feb18_KAZ_m1(11w), 11jan18_KAZ_m3(9w), and 11jan18_KAZ_m1(43w) have good MLH1
+
+Kaz.biv.plot <- ggplot(data = KAZ.Biv.curate.list, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()+facet_wrap(~mouse)
+
+#3000 to curate
+
+cur11jan18_KAZ_m4 <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$mouse == "11jan18_KAZ_m4",]
+
+cur11jan18_KAZ_m4$Obj.ID <- paste(cur11jan18_KAZ_m4$fileName, cur11jan18_KAZ_m4$boxNumber, sep = "_")
+
+write.table(cur11jan18_KAZ_m4, "~./MLH1repo/results/curate_cur11jan18_KAZ_m4_BivData.txt", sep="\t", row.names = FALSE)
+
+#KAZ_mice
+
+#11jan18_KAZ_m2-curated file
+#11jan18_KAZ_m3-
+#11jan18_KAZ_m4-
+
+```
+
+Above is merged curated data from PWD and KAZ males of similar ages
+
+From the curation datasets, it seems that the dpulicate / repeat image loop is not effectively removing double counted images.
+
+are the bivalents with faint foci similar in SC length (or rank) across cells
+
+During the curation process, I could pull the intensities of foci which match the hand reported measure.
+I could also pull 'faint foci' from the auto-measures from the note section, as long as I can standardize my coding for faint foci.
+
+```{r}
+#proportions of bivalent classes
+foci.passed.KAZ <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
+
+foci.passed.PWD <- curBivData_PWD[curBivData_PWD$foci.pass == 1,]
+
+qq <- ggplot(data = foci.passed.KAZ, aes(color=fileName,x=mouse, y= hand.foci.count))+geom_jitter()+theme(legend.position="none")
+
+qq2 <- ggplot(data = foci.passed.PWD, aes(color=fileName,x=mouse, y= hand.foci.count))+geom_jitter()+theme(legend.position="none")
+
+qq4 <- ggplot(data = foci.passed.PWD, aes(x=hand.foci.count, y= chromosomeLength))+geom_jitter(aes(fill=nrm.F1.pos))+theme(legend.position="none")
+
+#PWD have more 2CO bivalents
+
+```
+
+```{r rand.PWD, echo=FALSE}
+
+PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]#670
+cur18may15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m1",]
+cur18may15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m2",]
+
+cur10mar15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m1",]
+cur10mar15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m2",]
+
+table(PWD.BivData.curate.list$mouse)
+#4mice in the current dataset
+
+#10mar15_PWD_m1: 13w
+#10mar15_PWD_m2: 13w
+#18may15_PWD_m1:  7w -- use this one
+#18may15_PWD_m2: 7w  (seems like not all of the images were run through the algorithm)
+
+```

### 2019-07-16 results/PWD.KAZ.BivData.Rmd
small updates to setupdata and the analysis script

@@
 
+```{r}
+#histograms to display the average number of Objects per cell
+PWD.curateautoBivData_cells <- ddply(PWD_curate_BivData, c("fileName"), summarise,
+                    number = length(Obj.ID)
+
+)
+
+PWD.autoBivData_cells_raw <- ggplot(data=PWD_BivData_cells, aes(x=boxs.IDd))+
+  geom_histogram()+
+  ggtitle("PWD raw bivadata")
+
+PWD.autoBivData_cells <- ggplot(data=PWD.curateautoBivData_cells, aes(x=number))+
+  geom_histogram()+   ggtitle("PWD curated")
+
+PWD.autoBivData_cells_manu <- ggplot(data=PWD.auto.man_BivData_cells, aes(x=number))+
+  geom_histogram()+
+  ggtitle("PWD")
+
+```
+
 ```{r KAZ.manual.merge}
@@
 KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")
+
+```
+
+```{r histograms}
+
+KAZ.autoBivData_cells_raw <- ggplot(data=KAZ_BivData_cells, aes(x=boxs.IDd))+
+  geom_histogram()+
+  ggtitle("KAZ raw bivadata")
 

### 2019-07-16 src/setupData_MLH1repo.R

@@ -69,7 +69,6 @@ meta.data6.12.19$DOB <- as.Date(meta.data6.12.19$DOB, format= "%m/%d/%y") #the n
 meta.data6.12.19 <- meta.data6.12.19[!(is.na(meta.data6.12.19$mouse)|meta.data6.12.19$mouse==""),]
 
 #calq age for metadata
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 meta.data6.12.19 <- add_euth_date(meta.data6.12.19)#this isn't working because there are many mouse values which are not n the right format!
 
 #must convert euth.date to date (remove the non dates)
@@ -95,14 +94,13 @@ MouseMetaData.wrkin$mouse[which(duplicated(MouseMetaData.wrkin$mouse))]
 #merge with MLH1 and BivData
 impMLH1_data <- merge(MLH1_data, MouseMetaData.wrkin, by.x = "mouse")#don't apply the all parameters
 
-length(which(duplicated(impMLH1_data$Random.Name)))#29 duplicated
+length(which(duplicated(impMLH1_data$Random.Name)))#29 duplicated, 6 duplicated..
 
-length(which(duplicated(impMLH1_data$fileName)))# 117 fileName
+length(which(duplicated(impMLH1_data$fileName)))# 117 fileName, 107 duplicated
 #some of these are true duplicates, but some are the same image files that were quantified across multiple anon batches
 Quant2batches <- impMLH1_data[which(duplicated(impMLH1_data$fileName)),]#G, LEW, SPI females
 
 true.duplicates <- impMLH1_data[(which(duplicated(impMLH1_data$Random.Name))),]#but now these aren't duplicated...
-
 
 
 #integrating dates
@@ -111,8 +109,6 @@ colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
 MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
 
 MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
-
-
 
 
 #MLH1 - 2921,   #ll -- 2227, 
@@ -139,8 +135,6 @@ MLH1_data <- MLH1_data[ !grepl("12sep16_MSM_f3", MLH1_data$mouse) , ]
 MLH1_data <- MLH1_data[ !grepl("12sep16_MSM_f1", MLH1_data$mouse) , ]
 
 
-
-
[…]

### 2019-07-16 results/PWD.KAZ.BivData.Rmd
adding more data to slides

@@
 
-SC.KAZ <- ggplot(data = KAZ_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,150))+ggtitle("KAZ SC lengths")
+SC.KAZ <- ggplot(data = KAZ_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+
+  theme(legend.position="none")+ylim(c(35,150))+ggtitle("KAZ SC lengths")
 #can I add the means?
@@
 SC.KAZ
+
+KAZ_mean_SC <- ddply(KAZ_curate_BivData, c("mouse"), summarise,
+
+                    boxs.IDd = max(boxNumber),
+                    nboxes_passed = length(boxNumber),
+                    pass_rate = nboxes_passed/boxs.IDd,
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength)
+)
+
+PWD_mean_SC <- ddply(PWD_curate_BivData, c("mouse"), summarise,
+
+                    boxs.IDd = max(boxNumber),
+                    nboxes_passed = length(boxNumber),
+                    pass_rate = nboxes_passed/boxs.IDd,
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength)
+)
+
+t.test(KAZ_curate_BivData$chromosomeLength,PWD_curate_BivData$chromosomeLength)
+
 ```
 
-```{r cell.tables}
+the PWD has longer means
+
+```{r SClength.cell.tables}
 #add total COs
@@
 
-bin.short.PWD <- DF.w.rank2[DF.w.rank2$rank < 5,]
+bin.short.PWD <- PWD_whole.cell.BivData[PWD_whole.cell.BivData$rank < 5,]
+
 bin.short.PWD <- bin.short.PWD[bin.short.PWD$fileName %in% full.cell.list$fileName,]
 
-kk <- ggplot(data=bin.short.PWD, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color =Foci1, shape=mouse))+ggtitle("short bins")
+kk <- ggplot(data=bin.short.PWD, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25)+ggtitle("short bins")
 
@@
 
-```{r}
+```{r hist.boxes.per.cell, echo=FALSE}
 #histograms to display the average number of Objects per cell
@@
 SC.test <- t.test()
+
+```
+
+```{r SC.by.bins}
+
+all.whole.cell <- rbind(KAZ_whole.cell.BivData,PWD_whole.cell.BivData)
+
+#compare short and long bins
+bin_short <- all.whole.cell[all.whole.cell$rank < 5,]
+
+bin_short <- all.whole.cell[all.whole.cell$rank < 3,]#comparing the bottom 3 are not sig different
+
+#bin.short.PWD <- bin.short.PWD[bin.short.PWD$fileName %in% full.cell.list$fileName,]
+
+short_SC <- ggplot(data=bin_short, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=fileName))+ggtitle("short bins")+facet_wrap(~strain)+ theme(legend.position="none")
+
+t.test(bin_short$chromosomeLength[bin_short$strain == "PWD"],
+       bin_short$chromosomeLength[bin_short$strain == "KAZ"])
+
+#p-value = 0.02131
+
+bin_long <- all.whole.cell[all.whole.cell$rank > 15,]#comparing the top 3 is not sig different
+
+long_SC <- ggplot(data=bin_long, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=fileName) )+ggtitle("Long bins")+facet_wrap(~strain)+ theme(legend.position="none")
+
+t.test(bin_long$chromosomeLength[bin_long$strain == "PWD"],
+       bin_long$chromosomeLength[bin_long$strain == "KAZ"])
+
+#p-value = 0.01495
+```
+
+```{r chrm.class, echo=FALSE}
+#proportions of bivalent classes
+
+foci.passed.KAZ <- KAZ_curate_BivData[KAZ_curate_BivData$foci.pass == 1,]
+foci.passed.PWD <- PWD_curate_BivData[PWD_curate_BivData$foci.pass == 1,]
+
+jj <- rbind(foci.passed.KAZ,foci.passed.PWD)
+
+qq <- ggplot(data = jj, aes(color=fileName,x=mouse, y= hand.foci.count))+geom_jitter()+theme(legend.position="none")+facet_wrap(~strain)
+
+qq2 <- ggplot(data = foci.passed.PWD, aes(color=fileName,x=mouse, y= hand.foci.count))+geom_jitter()+theme(legend.position="none")
+
+qq4 <- ggplot(data = foci.passed.PWD, aes(x=hand.foci.count, y= chromosomeLength))+geom_jitter(aes(fill=nrm.F1.pos))+theme(legend.position="none")
+
+#PWD have more 2CO bivalents
+biv_class_table <- ddply(jj, c("hand.foci.count", "strain"), summarise,
+                    number = length(Obj.ID)
+)
+
+#### use whole cell data
+
+KAZ_whole.cell.BivData
+all.whole.cell
+#chi-squard test?
+
+whole.cell_biv_class_table <- ddply(all.whole.cell, c("hand.foci.count", "strain"), summarise,
+                    number = length(Obj.ID)
+)
 
@@
 
-```{r}
-#proportions of bivalent classes
-foci.passed.KAZ <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
-
-foci.passed.PWD <- curBivData_PWD[curBivData_PWD$foci.pass == 1,]
-
-qq <- ggplot(data = foci.passed.KAZ, aes(color=fileName,x=mouse, y= hand.foci.count))+geom_jitter()+theme(legend.position="none")
-
-qq2 <- ggplot(data = foci.passed.PWD, aes(color=fileName,x=mouse, y= hand.foci.count))+geom_jitter()+theme(legend.position="none")
-
-qq4 <- ggplot(data = foci.passed.PWD, aes(x=hand.foci.count, y= chromosomeLength))+geom_jitter(aes(fill=nrm.F1.pos))+theme(legend.position="none")
-
-#PWD have more 2CO bivalents
-
-```
-
 ```{r rand.PWD, echo=FALSE}

### 2019-07-17 results/PWD.KAZ.BivData.Rmd
work from home

@@
 
+```{r}
+#SC lengths by CO number
+
+foci.pass <- KAZ_curate_BivData
+
+```
+
 the PWD has longer means
@@
 cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]
-
-write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)
+#write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)
 
@@
 
-write.table(PWDcells2curate, "~./MLH1repo/results/PWD.cell.curate.csv", sep=",", row.names = FALSE)
-
+#write.table(PWDcells2curate, "~./MLH1repo/results/PWD.cell.curate.csv", sep=",", row.names = FALSE)
 ```
@@
 #ok these should be full cells
-hist(hh2$rank, breaks =21)
+
 #there are 2x as many 0's bc 1 and 0 are binned together
@@
 
-length.by.class <- ggplot(data=hh2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=hand.foci.count, shape=mouse))+ggtitle("PWD full cell SC lengths")
+#length.by.class <- ggplot(data=hh2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=hand.foci.count, shape=mouse))+ggtitle("PWD full cell SC lengths")
 #most cells are from 1 mouse
@@
 
-SC.test <- t.test()
+#SC.test <- t.test()
+
+```
+
+```{r SC.Chrm, echo=FALSE}
+all.curated.BivData <- rbind(KAZ.auto.man_BivData, PWD.auto.man_BivData)
+all.whole.cell <- rbind(KAZ_whole.cell.BivData,PWD_whole.cell.BivData)
+
+BivClass.SC_BivData <- ggplot(data=all.curated.BivData, aes(x=hand.foci.count,  y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count)))+ggtitle("PWD full cell SC lengths")+geom_boxplot(aes(alpha=.3))+facet_wrap(~strain)
+
+BivClass.SC_whole.cell <- ggplot(data=all.whole.cell, aes(x=hand.foci.count,  y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count)))+
+  ggtitle("PWD full cell SC lengths")+
+  geom_boxplot(aes(alpha=.3))+
+  facet_wrap(~strain)
+
+all.whole.cell.sub <- all.whole.cell[all.whole.cell$mouse != "18may15_PWD_m1",]
+
+all.whole.cell.sub <- all.whole.cell.sub[all.whole.cell.sub$mouse != "6mar16_PWD_m1",]
+
+all.whole.cell.sub <- all.whole.cell.sub[all.whole.cell.sub$hand.foci.count >= 1,]
+
+table(all.whole.cell.sub$mouse)
+
+SC.COnum <- ggplot(all.whole.cell.sub, aes(x=mouse, y=chromosomeLength, color=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+
+  geom_jitter(alpha=0.8)+
+  geom_boxplot(alpha=0.6) +theme(axis.text.x = element_text(angle = 45, hjust = 1))+
+  ggtitle("SC Length by Chromosome Class")+theme(legend.position="none")
 
@@
 ```{r SC.by.bins}
-
-all.whole.cell <- rbind(KAZ_whole.cell.BivData,PWD_whole.cell.BivData)
 

### 2019-07-17 results/PWD.KAZ.BivData.Rmd
tried fitting exponential model

@@
 
-#SC.test <- t.test()
+OneCO <- all.curated.BivData[all.curated.BivData$hand.foci.count == 1,]
+TwoCO <- all.curated.BivData[all.curated.BivData$hand.foci.count == 2,]
+
+Class.SC.test.1CO <- t.test(OneCO$chromosomeLength[OneCO$strain == "PWD"], OneCO$chromosomeLength[OneCO$strain == "KAZ"])
+
+Class.SC.test.2CO <- t.test(TwoCO$chromosomeLength[TwoCO$strain == "PWD"],TwoCO$chromosomeLength[TwoCO$strain == "KAZ"])
 
@@
 
-newPWD_curate_BivData =
-  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
+#newPWD_curate_BivData =
+#  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
 
@@
 #isolate the passing foci 
-curBivData_PWD.good.foci  <- newPWD_curate_BivData[newPWD_curate_BivData$foci.pass == 1,]
+#curBivData_PWD.good.foci  <- newPWD_curate_BivData[newPWD_curate_BivData$foci.pass == 1,]
+
+curBivData_PWD.good.foci <- PWD.auto.man_BivData[PWD.auto.man_BivData$foci.pass == 1,]
 
@@
 
-table(newPWD_curate_BivData$fileName)
-table(newPWD_curate_BivData$mouse) #4 mice, ~100 each mouse
-
-```
-
-```{r curated, echo=FALSE}
-curBivData_KAZ.good.foci <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
+#table(newPWD_curate_BivData$fileName)
+#table(newPWD_curate_BivData$mouse) #4 mice, ~100 each mouse
+
+```
+
+```{r melt.KAZ, echo=FALSE}
+
+#use the combined - auto and manual data
+
+curBivData_KAZ.good.foci <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$foci.pass == 1,]
 #curated objects
-KAZ_m2_curate_bivData =
-  read.csv("~./MLH1repo/data/BivData/curated_21may18_KAZ_m2_BivData.csv", header = TRUE)
+#KAZ_m2_curate_bivData =
+#  read.csv("~./MLH1repo/data/BivData/curated_21may18_KAZ_m2_BivData.csv", header = TRUE)
 
@@
 count.hist <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(hand.foci.count))+geom_histogram()
+
+```
+
+```{r meltKAZ, echo=FALSE}
+
+#melt KAZ
+#isolate the passing foci 
+#curBivData_KAZ.good.foci  <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
+
+curBivData_KAZ.good.foci <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$foci.pass == 1,]
+colnames(curBivData_KAZ.good.foci)
+
+curBivData_KAZ.good.foci$nrm.F1.pos <- curBivData_KAZ.good.foci$Foci1 / curBivData_KAZ.good.foci$chromosomeLength
+curBivData_KAZ.good.foci$nrm.F2.pos <- curBivData_KAZ.good.foci$Foci2 / curBivData_KAZ.good.foci$chromosomeLength
+curBivData_KAZ.good.foci$nrm.F3.pos <-   curBivData_KAZ.good.foci$Foci3 / curBivData_KAZ.good.foci$chromosomeLength
+
+#melting the DataFrame is needed to ploting the normalized foci positions
+
+#do not include cols for IFD, or ABS.foci
+curBivData_KAZ.melt <- melt( curBivData_KAZ.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,
+
+                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
+              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","mouse", "category",   "strain" , "sex"))
+
+colnames(curBivData_KAZ.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count", "mouse", "category",   "strain" , "sex",
+                                       #name for the 2 melted cols
+                                      "nrmFoci.Type", "nrmFoci.pos")
+
+#remove NAs
+curBivData_KAZ.melt <- curBivData_KAZ.melt[!(is.na(curBivData_KAZ.melt$nrmFoci.Type) | curBivData_KAZ.melt$nrmFoci.pos==""), ]
 
@@
 
-```{r meltKAZ, echo=FALSE}
-
-#melt KAZ
-#isolate the passing foci 
-curBivData_KAZ.good.foci  <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
-#colnames(curBivData_KAZ)
-
-#melting the DataFrame is needed to ploting the normalized foci positions
-
-#do not include cols for IFD, or ABS.foci
-curBivData_KAZ.melt <- melt( curBivData_KAZ.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,
-
-                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
-              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count","mouse", "category",   "strain" , "sex"))
-
-colnames(curBivData_KAZ.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count", "mouse", "category",   "strain" , "sex",
-                                       #name for the 2 melted cols
-                                      "nrmFoci.Type", "nrmFoci.pos")
-
-#remove NAs
-curBivData_KAZ.melt <- curBivData_KAZ.melt[!(is.na(curBivData_KAZ.melt$nrmFoci.Type) | curBivData_KAZ.melt$nrmFoci.pos==""), ]
-
-```
-
-```{r plot}
-
+```{r position.plot}
 library(ggplot2)
@@
 
-nPWD.obs <- ddply(.data=curBivData_PWD.melt, 
+curBivData_PWD.melt <- curBivData_PWD.melt[curBivData_PWD.melt$hand.foci.count != "1?",]
+
+curBivData_PWD.melt <- droplevels(curBivData_PWD.melt)
+
+curBivData_PWD.melt$hand.foci.count <- as.numeric(curBivData_PWD.melt$hand.foci.count)
+
+nPWD.obs <- ddply(.data=curBivData_PWD.melt[curBivData_PWD.melt$hand.foci.count <3,], 
                  .(hand.foci.count),
@@
 #normalized foci plots
-PWD.plot.nrm <- ggplot(curBivData_PWD.melt,
+PWD.plot.nrm <- ggplot(curBivData_PWD.melt[curBivData_PWD.melt$hand.foci.count < 3,],
               aes(x=Foci.nrm.pos, fill=Foci.Type.nrm) ) + 
@@
 
-nKAZ.obs <- ddply(.data=curBivData_KAZ.melt, 
+curBivData_KAZ.melt2 <- curBivData_KAZ.melt[!is.na(curBivData_KAZ.melt$hand.foci.count),]
+curBivData_KAZ.melt2 <- curBivData_KAZ.melt2[curBivData_KAZ.melt2$hand.foci.count < 3,]
+
+nKAZ.obs <- ddply(.data=curBivData_KAZ.melt2, 
                  .(hand.foci.count),
@@
 
-KAZ.plot.nrm    <- ggplot(curBivData_KAZ.melt,
+KAZ.plot.nrm    <- ggplot(curBivData_KAZ.melt2,
               aes(x=nrmFoci.pos, fill=nrmFoci.Type) ) +
@@
 ```{r}
-#goal, examine the ages of mice being compared
-#distribution of ages
-
-histys <- ggplot(PWD.KAZ_MLH1.imprv, aes(age.weeks, fill=strain))+geom_density()
-
-```
-
-```{r first.analysis, echo=FALSE}
-#initial analysis of the BivData
-
-Kaz.biv.plot <- ggplot(data = KAZ_bivData, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()
+#t.test of nrm positions
+
+PWD.OneCO_nrmPos <- curBivData_PWD.melt[curBivData_PWD.melt$hand.foci.count ==1,]
+PWD.TWOCO_nrmPos <- curBivData_PWD.melt[curBivData_PWD.melt$hand.foci.count ==2,]
+
+KAZ.OneCO_nrmPos <- curBivData_KAZ.melt2[curBivData_KAZ.melt2$hand.foci.count ==1,]
+KAZ.TWOCO_nrmPos <- curBivData_KAZ.melt2[curBivData_KAZ.melt2$hand.foci.count ==2,]
+
+t.test(PWD.OneCO_nrmPos$Foci.nrm.pos, KAZ.OneCO_nrmPos$nrmFoci.pos)
+#p-value = 0.5468
+
+t.test(KAZ.TWOCO_nrmPos$nrmFoci.pos[KAZ.TWOCO_nrmPos$nrmFoci.Type == "nrm.F1.pos"],
+       PWD.TWOCO_nrmPos$Foci.nrm.pos[PWD.TWOCO_nrmPos$Foci.Type.nrm == "nrm.F1"])
+#p-value = 0.9421
+
+t.test(KAZ.TWOCO_nrmPos$nrmFoci.pos[KAZ.TWOCO_nrmPos$nrmFoci.Type == "nrm.F2.pos"],
+       PWD.TWOCO_nrmPos$Foci.nrm.pos[PWD.TWOCO_nrmPos$Foci.Type.nrm == "nrm.F2"])
+#p-value = 0.5403
+```
+
+From the curation datasets, it seems that the dpulicate / repeat image loop is not effectively removing double counted images.
+
+are the bivalents with faint foci similar in SC length (or rank) across cells
+
+During the curation process, I could pull the intensities of foci which match the hand reported measure.
+I could also pull 'faint foci' from the auto-measures from the note section, as long as I can standardize my coding for faint foci.
+
+```{r fit.exp}
+#fitting IFDs to expoential Distribution
+#curBivData_PWD.melt
+#make an IFD plot of Null.IFDs drawn from exponential
+#curBivData_KAZ.melt2
+
+#total Bivdata
+KAZ.2COs <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$hand.foci.count == 2,]
+PWD.2COs <- PWD.auto.man_BivData[PWD.auto.man_BivData$hand.foci.count == 2,]
+
+all.2CO <- rbind(KAZ.2COs,PWD.2COs)
+#all.2CO <- ddply(all.2CO, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+#all.2CO <- all.2CO[ -c(24) ]
+
+all.2CO$Foci1.PER <- all.2CO$Foci1 / all.2CO$chromosomeLength
+all.2CO$Foci2.PER <- all.2CO$Foci2 / all.2CO$chromosomeLength
+all.2CO$Foci3.PER <- all.2CO$Foci3 / all.2CO$chromosomeLength
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+all.2CO <- add_IFD(all.2CO)
+all.2CO <- droplevels(all.2CO)
+all.2CO <- all.2CO[!(is.na(all.2CO$fileName) | all.2CO$fileName == ""), ]
+
+all.2CO <- all.2CO[!(is.na(all.2CO$IFD1_PER) | all.2CO$IFD1_PER == ""), ]
+
+all.2CO$mouse <- droplevels(all.2CO$mouse)
+
+#whole cell comparison (use rank)
+#KAZ_whole.cell.BivData, #PWD_whole.cell.BivData
+
+#limit to mice with most observations
+
+whole.cell.2CO <- rbind(KAZ_whole.cell.BivData,PWD_whole.cell.BivData)
+
+whole.cell.2CO$Foci1.PER <- whole.cell.2CO$Foci1 / whole.cell.2CO$chromosomeLength
+whole.cell.2CO$Foci2.PER <- whole.cell.2CO$Foci2 / whole.cell.2CO$chromosomeLength
+whole.cell.2CO$Foci3.PER <- whole.cell.2CO$Foci3 / whole.cell.2CO$chromosomeLength
+
+#call IFD function on this DF
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+whole.cell.2CO <- add_IFD(whole.cell.2CO)
+whole.cell.2CO <- droplevels(whole.cell.2CO)
+whole.cell.2CO <- whole.cell.2CO[!(is.na(whole.cell.2CO$fileName) | whole.cell.2CO$fileName == ""), ]
+
+require(vcd)
+require(MASS)
+
+# data generation
+ex <- rexp(10000, rate = 1.85) # generate some exponential distribution
+control <- abs(rnorm(10000)) # generate some other distribution
+
+# estimate the parameters
+fit1 <- fitdistr(ex, "exponential") 
+fit2 <- fitdistr(control, "exponential")
+
+fit.AP1 <- fitdistr(all.2CO$IFD1_PER, "exponential")
+
+fit.PWD <- fitdistr(all.2CO$IFD1_PER[all.2CO$strain == "PWD"], "exponential")#change to gamma
+fit.KAZ <- fitdistr(all.2CO$IFD1_PER[all.2CO$strain== "KAZ"], "exponential")
+
+PWD.curve <- hist(all.2CO$IFD1_PER[all.2CO$strain == "PWD"], freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
+curve(dexp(x, rate = fit.PWD$estimate), from = 0, col = "red", add = TRUE)
+
+KAZ.curve <- hist(all.2CO$IFD1_PER[all.2CO$strain == "KAZ"], freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
+curve(dexp(x, rate = fit.KAZ$estimate), from = 0, col = "red", add = TRUE)
+
+ks.test(ex, "pexp", fit1$estimate) # p-value > 0.05 -> distribution not refused
+ks.test(control, "pexp", fit2$estimate)
+
+# plot a graph
+hist(ex, freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+curve(dexp(x, rate = fit1$estimate), from = 0, col = "red", add = TRUE)
+
+hist(all.2CO$IFD1_PER, freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
+curve(dexp(x, rate = fit.AP1$estimate), from = 0, col = "red", add = TRUE)
+
+```
+
+```{r rand.PWD, echo=FALSE}
+
+PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]#670
+cur18may15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m1",]
+cur18may15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m2",]
+
+cur10mar15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m1",]
+cur10mar15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m2",]
+
+table(PWD.BivData.curate.list$mouse)
+#4mice in the current dataset
+
+#10mar15_PWD_m1: 13w
+#10mar15_PWD_m2: 13w
+#18may15_PWD_m1:  7w -- use this one
+#18may15_PWD_m2: 7w  (seems like not all of the images were run through the algorithm)
+
+```
+
+```{r KAZ.curate.list, echo=FALSE}
+#write out list for initial quality control and subsetting
+
+#remove centromere_PER > .2
+
+KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123
+
+#remove chrmLength > 150? 200
+KAZ.Biv.curate.list <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$chromosomeLength < 300, ]#not completely sure 300 is probably the max length is the best length
+#3117
+
+table(KAZ.Biv.curate.list$mouse)
+#7 mice: 11jan18_KAZ_m4(9w), 1feb18_KAZ_m1(11w), 11jan18_KAZ_m3(9w), and 11jan18_KAZ_m1(43w) have good MLH1
+
+Kaz.biv.plot <- ggplot(data = KAZ.Biv.curate.list, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()+facet_wrap(~mouse)
+
+#3000 to curate
+
+cur11jan18_KAZ_m4 <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$mouse == "11jan18_KAZ_m4",]
+
+cur11jan18_KAZ_m4$Obj.ID <- paste(cur11jan18_KAZ_m4$fileName, cur11jan18_KAZ_m4$boxNumber, sep = "_")
+
+write.table(cur11jan18_KAZ_m4, "~./MLH1repo/results/curate_cur11jan18_KAZ_m4_BivData.txt", sep="\t", row.names = FALSE)
+
+#KAZ_mice
+
+#11jan18_KAZ_m2-curated file
+#11jan18_KAZ_m3-
+#11jan18_KAZ_m4-
 
@@
 ```
-
-```{r KAZ.curate.list, echo=FALSE}
-#write out list for initial quality control and subsetting
-
-#remove centromere_PER > .2
-
-KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123
-
-#remove chrmLength > 150? 200
-KAZ.Biv.curate.list <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$chromosomeLength < 300, ]#not completely sure 300 is probably the max length is the best length
-#3117
-
-table(KAZ.Biv.curate.list$mouse)
-#7 mice: 11jan18_KAZ_m4(9w), 1feb18_KAZ_m1(11w), 11jan18_KAZ_m3(9w), and 11jan18_KAZ_m1(43w) have good MLH1
-
-Kaz.biv.plot <- ggplot(data = KAZ.Biv.curate.list, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()+facet_wrap(~mouse)
-
-#3000 to curate
-
-cur11jan18_KAZ_m4 <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$mouse == "11jan18_KAZ_m4",]
-
-cur11jan18_KAZ_m4$Obj.ID <- paste(cur11jan18_KAZ_m4$fileName, cur11jan18_KAZ_m4$boxNumber, sep = "_")
-
-write.table(cur11jan18_KAZ_m4, "~./MLH1repo/results/curate_cur11jan18_KAZ_m4_BivData.txt", sep="\t", row.names = FALSE)
-
-#KAZ_mice
-
-#11jan18_KAZ_m2-curated file
-#11jan18_KAZ_m3-
-#11jan18_KAZ_m4-
-
-```
-
-Above is merged curated data from PWD and KAZ males of similar ages
-
-From the curation datasets, it seems that the dpulicate / repeat image loop is not effectively removing double counted images.
-
-are the bivalents with faint foci similar in SC length (or rank) across cells
-
-During the curation process, I could pull the intensities of foci which match the hand reported measure.
-I could also pull 'faint foci' from the auto-measures from the note section, as long as I can standardize my coding for faint foci.
-
-```{r rand.PWD, echo=FALSE}
-
-PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]#670
-cur18may15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m1",]
-cur18may15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m2",]
-
-cur10mar15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m1",]
-cur10mar15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m2",]
-
-table(PWD.BivData.curate.list$mouse)
-#4mice in the current dataset
-
-#10mar15_PWD_m1: 13w
-#10mar15_PWD_m2: 13w
-#18may15_PWD_m1:  7w -- use this one
-#18may15_PWD_m2: 7w  (seems like not all of the images were run through the algorithm)
-
-```

### 2019-07-19 results/PWD.KAZ.BivData.Rmd
clean up of PWD.KAZ analysis script and associated files

@@
 
-#curBivData_18may19_PWD_m1 = read.csv("~./MLH1repo/results/curate_18may18_PWD_m1_BivData.txt", header = TRUE, sep = '\t')
-#curBivData_21may18_KAZ_m2 = read.csv("~./MLH1repo/results/curate_21may18_KAZ_m2_BivData.txt",header = TRUE, sep = '\t')
-
+#1. Curated Data
 #PWD set 4 mice, ~100 objs each)
@@
 
-#load hand measures, get it to match the auto
-PWD_hand_measure_BivData.full =
+#merged KAZ and PWD
+curate_BivData <- rbind(KAZ_curate_BivData, PWD_curate_BivData)
+
+#2. Hand measures (merge first, then add extra cols)
+PWD_hand_measure_BivData =
   read.csv("~./MLH1repo/data/BivData/PWD_male_hand_measures.csv", header = TRUE)
 
+KAZ_hand_measure_BivData =
+  read.csv("~./MLH1repo/data/BivData/KAZ_male_hand_measures.csv", header = TRUE)
+
+#merge together -- check that the colnames match!!
+HandMeasures_BivData <- rbind(KAZ_hand_measure_BivData, PWD_hand_measure_BivData)
+
 #remove extra empty rows
-PWD_hand_measure_BivData.full <- PWD_hand_measure_BivData.full[!(is.na(PWD_hand_measure_BivData.full$fileName) | PWD_hand_measure_BivData.full$fileName == ""), ]
+HandMeasures_BivData <- HandMeasures_BivData[!(is.na(HandMeasures_BivData$fileName) | HandMeasures_BivData$fileName == ""), ]
 
-#PWD_hand_measure_BivData.full$Foci1  <- PWD_hand_measure_BivData.full$foci.1.position
-#PWD_hand_measure_BivData.full$Foci2 <- PWD_hand_measure_BivData.full$medial.foci.position
-#PWD_hand_measure_BivData.full$Foci3 <- PWD_hand_measure_BivData.full$distal.foci.position
-#PWD_hand_measure_BivData.full$boxNumber <- PWD_hand_measure_BivData.full$manual.blobject.name
+#add Obj.ID
+HandMeasures_BivData$Obj.ID <- paste(HandMeasures_BivData$fileName, HandMeasures_BivData$boxNumber, sep = "_")
+#add centromere_PER (match auto)
+HandMeasures_BivData$centromere_PER_Position <- HandMeasures_BivData$centromere_ABS_Position / HandMeasures_BivData$chromosomeLength
 
-PWD_hand_measure_BivData.full$Obj.ID <- paste(PWD_hand_measure_BivData.full$fileName, PWD_hand_measure_BivData.full$boxNumber, sep = "_")
-
-PWD_hand_measure_BivData.full$centromere_PER_Position <- PWD_hand_measure_BivData.full$centromere_ABS_Position / PWD_hand_measure_BivData.full$chromosomeLength
-
-PWD_hand_measure_BivData.full$SC.pass <- "1"
-PWD_hand_measure_BivData.full$foci.pass <- "1"
-PWD_hand_measure_BivData.full$curated.1pass.0fail. <- ""
-#PWD_hand_measure_BivData.full$curation.notes <- ""
-#PWD_hand_measure_BivData.full$hand.foci.count <- ""
+#more columns to match Auto DFs
+HandMeasures_BivData$SC.pass <- "1"
+HandMeasures_BivData$foci.pass <- "1"
+HandMeasures_BivData$curated.1pass.0fail. <- ""
 
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-PWD_hand_measure_BivData.full <- add_mouse(PWD_hand_measure_BivData.full)
-PWD_hand_measure_BivData.full <- add_category(PWD_hand_measure_BivData.full)
-PWD_hand_measure_BivData.full <- add_strain(PWD_hand_measure_BivData.full)
-PWD_hand_measure_BivData.full <- add_sex(PWD_hand_measure_BivData.full)
+HandMeasures_BivData <- add_mouse(HandMeasures_BivData)
+HandMeasures_BivData <- add_category(HandMeasures_BivData)
+HandMeasures_BivData <- add_strain(HandMeasures_BivData)
+HandMeasures_BivData <- add_sex(HandMeasures_BivData)
 
-#3 columns don't overlap
-#avergeRED, numberCrossOvers, 
-#manual.blobject.name,
+#make indevidal DFs by strain just in case...
+PWD.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "PWD",]
+KAZ.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "KAZ",]
 
-KAZ_hand_measure_BivData.full =
-  read.csv("~./MLH1repo/data/BivData/KAZ_male_hand_measures.csv", header = TRUE)
+```
 
-KAZ_hand_measure_BivData.full <- KAZ_hand_measure_BivData.full[!(is.na(KAZ_hand_measure_BivData.full$fileName) | KAZ_hand_measure_BivData.full$fileName == ""), ]
+```{r Merge.manual.auto, echo=FALSE}
 
-KAZ_hand_measure_BivData.full$boxNumber <- KAZ_hand_measure_BivData.full$manual.blobject.name
+#merge with the both strain DF
+#curate_BivData ... HandMeasures_BivData
 
-KAZ_hand_measure_BivData.full$Obj.ID <- paste(KAZ_hand_measure_BivData.full$fileName, KAZ_hand_measure_BivData.full$boxNumber, sep = "_")
+#.1 match the colnames,  #merge, partial match
+#colnames(PWD_curate_BivData)
+#colnames(PWD_hand_measure_BivData.full)
 
-KAZ_hand_measure_BivData.full$SC.pass <- "1"
-KAZ_hand_measure_BivData.full$foci.pass <- "1"
-KAZ_hand_measure_BivData.full$curated.1pass.0fail. <- ""
+#merg4.PWD_curate_BivData <-    #remove "avergeRED", numberCrossOvers
+#merg4.PWD_hand_measure_BivData.full <- #manual.blobject.name
 
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-KAZ_hand_measure_BivData.full <- add_mouse(KAZ_hand_measure_BivData.full)
-KAZ_hand_measure_BivData.full <- add_category(KAZ_hand_measure_BivData.full)
-KAZ_hand_measure_BivData.full <- add_strain(KAZ_hand_measure_BivData.full)
-KAZ_hand_measure_BivData.full <- add_sex(KAZ_hand_measure_BivData.full)
+#auto.curate
+auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
+new.DF <- curate_BivData[ , auto.c_names, drop = TRUE]
 
-#not sure why I have a distinction of hand_measure and hand_measure.full
-#KAZ_hand_measure_BivData <- KAZ_hand_measure_BivData.full[KAZ_hand_measure_BivData.full$SC.pass == 1,]
+#manual
+manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position","SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
+new.DF2 <- HandMeasures_BivData[ , manual_names, drop = TRUE]
 
-#manual blobject name -- boxNumber, Obj.ID 
-#  20jul15_18may15_PWD_m1_sp1_2.2_blob1
+#Merge the 2 DFs
+Auto.Man_BivData <- rbind(new.DF, new.DF2)
+
+#make WHOLE CELL DF
+
+#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
+#make better variable name
+Auto.Man_BivData_cells <- ddply(Auto.Man_BivData, c("fileName"), summarise,
+                    number = length(Obj.ID),
+                    #make the rank or
+                     maxChrm_length = max(chromosomeLength),
+                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]
+
+)
+
+#Isolate cells with all (19) autosomes measured
+whole.cell_table <- Auto.Man_BivData_cells[Auto.Man_BivData_cells$number > 17,]
+whole.cell_table <- whole.cell_table[whole.cell_table$number < 21,]
+whole.cell.list <- unique(whole.cell.DF$fileName)
+
+#make the whole.cell BivData DF
+whole.cell.DF <- Auto.Man_BivData[Auto.Man_BivData$fileName %in% whole.cell.list,]
+
+#ADD within cell RANK to indiviual Bivalents,
+whole.cell.DF <- ddply(whole.cell.DF, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+
+#there are instances of ties, 
+#oo -- look at correlation ranks
+
+#make the bins
+short.bin.Bivs <- whole.cell.DF[whole.cell.DF$rank < 6,]
+long.bin.Bivs <- whole.cell.DF[whole.cell.DF$rank > 14,]
+
 ```
+
+```{r Melt.DF}
+library(reshape2)
+#Auto.Man_BivData, all curate data
+#get the whole cell verion with whole.cell.list
+
+#isolate bivs with passing foci measures
+Auto.Man_BivData.good.foci <- Auto.Man_BivData[Auto.Man_BivData$foci.pass == 1,]
+colnames(Auto.Man_BivData.good.foci)
+
+#add nrm foci measures
+Auto.Man_BivData.good.foci$nrm.F1.pos <- Auto.Man_BivData.good.foci$Foci1 / Auto.Man_BivData.good.foci$chromosomeLength
+Auto.Man_BivData.good.foci$nrm.F2.pos <- Auto.Man_BivData.good.foci$Foci2 / Auto.Man_BivData.good.foci$chromosomeLength
+Auto.Man_BivData.good.foci$nrm.F3.pos <-   Auto.Man_BivData.good.foci$Foci3 / Auto.Man_BivData.good.foci$chromosomeLength
+
+#melting the DataFrame is needed to ploting the normalized foci positions
+#do not include cols for IFD, or ABS.foci
+Auto.Man_BivData.melt <- melt( Auto.Man_BivData.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,
+
+                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
+              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count","mouse", "category",   "strain" , "sex"))
+
+colnames(Auto.Man_BivData.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
+    "hand.foci.count", "mouse", "category",   "strain" , "sex",
+                                       #name for the 2 melted cols
+                                      "nrmFoci.Type", "nrmFoci.pos")
+
+#remove NAs
+Auto.Man_BivData.melt <- Auto.Man_BivData.melt[!(is.na(Auto.Man_BivData.melt$nrmFoci.Type) | Auto.Man_BivData.melt$nrmFoci.pos==""), ]
+
+#make PWD and KAZ specific melt.DF ... if you want
+
+#make whole cell melt
+whole.cell_melt.DF <- Auto.Man_BivData.melt[Auto.Man_BivData.melt$fileName %in% whole.cell.list,]
+#add rank to these measures
+whole.cell_melt.DF <- ddply(whole.cell_melt.DF, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
+
+```
+
+```{r saveRdata}
+# all the main DFs have been made ...
+#save rdata file!
+
+save.image("data/MLH1/MLH1_data_setup_62419.RData")
+```
+
+```{r SClength.cell.tables}
+#this is useful for making the list of cells to curate (comment out after curation step)
+
+#add total COs
+KAZ_BivData_cells <- ddply(KAZ_curate_BivData, c("fileName"), summarise,
+
+                    boxs.IDd = max(boxNumber),
+                    nboxes_passed = length(boxNumber),
+                    pass_rate = nboxes_passed/boxs.IDd,
+
+                    totalCO = sum(numberCrossOvers),     
+                    avCO.per.CHRM = mean(numberCrossOvers),
+                    varCO.per.CHRM = var(numberCrossOvers),
+
+                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
+                    varSC.intensity = var(avergeRED, na.rm = TRUE),
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength),
+
+                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
+                    varCOdensity = var(chromosomeLength /numberCrossOvers)
+                            #av CO density
+                    #change the mean list for column numbers
+
+)
+
+#cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]
+#write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)
+
+#PWD version
+#add total COs
+PWD_BivData_cells <- ddply(PWD_curate_BivData, c("fileName"), summarise,
+
+                    boxs.IDd = max(boxNumber),
+                    nboxes_passed = length(boxNumber),
+                    pass_rate = nboxes_passed/boxs.IDd,
+
+                    totalCO = sum(numberCrossOvers),     
+                    avCO.per.CHRM = mean(numberCrossOvers),
+                    varCO.per.CHRM = var(numberCrossOvers),
+
+                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
+                    varSC.intensity = var(avergeRED, na.rm = TRUE),
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength),
+
+                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
+                    varCOdensity = var(chromosomeLength /numberCrossOvers)
+                            #av CO density
+                    #change the mean list for column numbers
+
+)
+
+PWDcells2curate <- PWD_BivData_cells[PWD_BivData_cells$pass_rate > .65,]
+
+#write.table(PWDcells2curate, "~./MLH1repo/results/PWD.cell.curate.csv", sep=",", row.names = FALSE)
+```
+
+From the cell level tables -- make lists of cells with ~75% of bivalents quantified. These will be manually verified and the missing chrms will be measured and supplemented.
+
+Data frames and analyses are the main output of this script - for a saved Rdata file.
+
+1. Curated BivData; read files, merge/rbind together
+
+2. Hand Measures; read files, rbind together
+
+3. Auto+manual; combine the curate and hand measures, rbind both strains together.
+    3B. Auto-Manual_whole_cell, make a list of fileNames for extracting these measures later; the biv from cells with all 19 autosome measures. Add bivalent rank to this DF.
+
+4. melted DF; Melt the Foci-Pass of Auto_Manual_DF
+
+--- Once the above DF's are made, proceed with following analyses ---
+
+1. SC Length
+    a. raw / loose bivalents
+    b. binned bivalents, short vs long, chrm class
+
+2. Normalized Foci Position
+   a. all/full
+   b. bins
+
+3. 2-CO
+    a. IFDs, raw norm
+    b. F1 by F2
+
+4. Rbar?
 
@@
 
-### BivData dataset
+PWD and KAZ differ in gwRR/MLH1 counts, is there a correlation in meiotic features which go with expected differences in gwRR.
 
-Goal:  Count the Bivdata for pwd and KAZ males
-start determining the subset of data
+```{r mergeing.meta.data, echo=FALSE}
+#meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
+#meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)
 
-```{r, echo=FALSE}
-#Curated data
-#merge PWD and KAZ BivData
+meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375
 
-#extract just the relavant PWD and KAZ
-PWD.MLH1 <- MLH1_data[MLH1_data$category == "PWD male",]
-KAZ.MLH1 <- MLH1_data[MLH1_data$category == "KAZ male",]
+#find a way to remove the rows without the right format before setting as as.Date
+#trying to remove these the extra rows..uggh
+#yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]
+#yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
+#yy <- yy[subset(yy$DOB, !(yy$DOB == "e1")),]
+#i think subseting is messing up the dataframes
 
-PWD.BivData.cells <- BivData_cells[BivData_cells$category == "PWD male",]
-PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]
+meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
 
-KAZ.BivData.cells <- BivData_cells[BivData_cells$category == "KAZ male",]
-KAZ.BivData <- fullBivData[fullBivData$category == "KAZ male",]
+#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#meta.data.file2 <- add_euth_date(meta.data.file2) #euthdate 1st
+#meta.data.file2 <- add_age(meta.data.file2)  #age second
+##got it to work ... maybe should save adding it to the merged DF
+meta.data.file2 <- add_category(meta.data.file2)
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+meta.data.file2 <- add_euth_date(meta.data.file2)
+meta.data.file2 <- add_age(meta.data.file2)
 
-PWD.KAZ_MLH1<- rbind(PWD.MLH1, KAZ.MLH1)
-PWD.KAZ_BivData<- rbind(PWD.BivData, KAZ.BivData)
+pwd.meta <- meta.data.file2[meta.data.file2$category == "PWD male",]
 
-```
+#pwd.meta <- add_euth_date(pwd.meta)
+#pwd.meta <- add_age(pwd.meta)
 
-what are the distributions of SC lengths?
+kaz.meta <- meta.data.file2[meta.data.file2$category == "KAZ male",]
 
-```{r cell.level.tables, echo=FALSE}
+#kaz.meta <- add_mouse(kaz.meta)
+#kaz.meta <- add_euth_date(kaz.meta)
+county =1
+for(i in kaz.meta$mouse){
+    templist= strsplit(i, split="_")[[1]]
+    euth <- templist[1]
+    #figure out date conversion
+    kaz.meta$euth.date2[county] <- as.character(euth)
 
-#make cell level tables
-#id the 18-20 biv cells
+    #kaz.meta$euth.date[county] <- as.Date(kaz.meta$euth.date[county], format='%d%b%y')
+    county= county +1
+}
 
-#estimate which cells can be used to make grouped pools
+kaz.meta$euth.date <- as.Date(kaz.meta$euth.date2, format='%d%b%y')
+
+kaz.meta <- add_age(kaz.meta)
+
+#reduce metadata down to 2 cols (mouse, DOB), merge this with MLH1
+MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
+colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
+MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
+
+PWD.KAZ_MLH1.imprv <- merge(PWD.KAZ_MLH1, MouseMetaData.wrkin)#don't apply the all parameters
+#report number with  and without
+
+#add euth date for calculating age
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+PWD.KAZ_MLH1.imprv <- add_euth_date(PWD.KAZ_MLH1.imprv)
+PWD.KAZ_MLH1.imprv <- add_age(PWD.KAZ_MLH1.imprv)
+
+#yesDOB.check <- PWD.KAZ_MLH1.imprv[!(is.na(PWD.KAZ_MLH1.imprv$DOB)),]
+#noDOB.check <- PWD.KAZ_MLH1.imprv[(is.na(PWD.KAZ_MLH1.imprv$DOB)),]#note, date type of data can't do the string matching!
+
+#total, KAZ 144, PWD 139
+#table(noDOB.check$category)#  no DOB, KAZ 67  PWD 133
+#table(yesDOB.check$category) #yes DOB, KAZ 77  PWD 6
+
+# I need to find the ages of more PWD males
 
@@
 Make plots for SC lengths
+
+kk <- ggplot(data=bin.short.PWD, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25)+ggtitle("short bins")
 
@@
 
-```{r}
-#SC lengths by CO number
-
-foci.pass <- KAZ_curate_BivData
-
-```
-
 the PWD has longer means
 
-```{r SClength.cell.tables}
-#add total COs
-KAZ_BivData_cells <- ddply(KAZ_curate_BivData, c("fileName"), summarise,
+```{r}
 
-                    boxs.IDd = max(boxNumber),
-                    nboxes_passed = length(boxNumber),
-                    pass_rate = nboxes_passed/boxs.IDd,
+show.plot <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=chromosomeLength, x=fileName))+geom_point()
 
-                    totalCO = sum(numberCrossOvers),     
-                    avCO.per.CHRM = mean(numberCrossOvers),
-                    varCO.per.CHRM = var(numberCrossOvers),
+show.plot2 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=Foci1.pos, x=hand.foci.count))+geom_jitter()
 
-                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
-                    varSC.intensity = var(avergeRED, na.rm = TRUE),
+show.plot99 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=nrm.F1.pos, x=hand.foci.count))+geom_jitter()+
+  ggtitle("Normalized foci positions")
 
-                    avSC.length = mean(chromosomeLength),
-                    varSC.length = var(chromosomeLength),
+count.hist <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(hand.foci.count))+geom_histogram()
 
-                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
-                    varCOdensity = var(chromosomeLength /numberCrossOvers)
-                            #av CO density
-                    #change the mean list for column numbers
-
-)
-
-cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]
-#write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)
-
-#PWD version
-#add total COs
-PWD_BivData_cells <- ddply(PWD_curate_BivData, c("fileName"), summarise,
-
-                    boxs.IDd = max(boxNumber),
-                    nboxes_passed = length(boxNumber),
-                    pass_rate = nboxes_passed/boxs.IDd,
-
-                    totalCO = sum(numberCrossOvers),     
-                    avCO.per.CHRM = mean(numberCrossOvers),
-                    varCO.per.CHRM = var(numberCrossOvers),
-
-                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
-                    varSC.intensity = var(avergeRED, na.rm = TRUE),
-
-                    avSC.length = mean(chromosomeLength),
-                    varSC.length = var(chromosomeLength),
-
-                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
-                    varCOdensity = var(chromosomeLength /numberCrossOvers)
-                            #av CO density
-                    #change the mean list for column numbers
-
-)
-
-PWDcells2curate <- PWD_BivData_cells[PWD_BivData_cells$pass_rate > .65,]
-
-#write.table(PWDcells2curate, "~./MLH1repo/results/PWD.cell.curate.csv", sep=",", row.names = FALSE)
-```
-
-From the cell level tables -- make lists of cells with ~75% of bivalents quantified. These will be manually verified and the missing chrms will be measured and supplemented.
-
-```{r PWD.merge.manual.auto, echo=FALSE}
-#merging manual measures and auto measures
-#KAZ_hand_measure_BivData
-#PWD_hand_measure_BivData
-
-#.1 match the colnames,  #merge, partial match
-#colnames(PWD_curate_BivData)
-#colnames(PWD_hand_measure_BivData.full)
-
-#merg4.PWD_curate_BivData <-    #remove "avergeRED", numberCrossOvers
-#merg4.PWD_hand_measure_BivData.full <- #manual.blobject.name
-
-#auto.curate
-auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
-new.DF <- PWD_curate_BivData[ , auto.c_names, drop = TRUE]
-
-#manual
-manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position","SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
-new.DF2 <- PWD_hand_measure_BivData.full[ , manual_names, drop = TRUE]
-
-#2. merge the 2 DFs
-PWD.auto.man_BivData <- rbind(new.DF, new.DF2)
-
-#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
-#make better variable name
-PWD.auto.man_BivData_cells <- ddply(PWD.auto.man_BivData, c("fileName"), summarise,
-                    number = length(Obj.ID),
-                    #make the rank or
-                     maxChrm_length = max(chromosomeLength),
-                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]
-
-)
-
-#Isolate cells with all autosomes measured
-PWDfull.cell.list <- PWD.auto.man_BivData_cells[PWD.auto.man_BivData_cells$number > 17,]
-PWDfull.cell.list <- PWDfull.cell.list[PWDfull.cell.list$number < 21,]#remove the too many cell
-
-#this gives the rank to each biv within each cell  
-PWD.Bivs.w.rank <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
-#there are instances of ties,
-
-#this does something weird
-#PWD.auto.man_BivData$rank1 <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
-
-PWD.auto.man_BivData <- PWD.auto.man_BivData[ -c(19:25) ]
-
-#pull all biv data for the full cell data
-PWD_whole.cell.BivData <- PWD.Bivs.w.rank[PWD.Bivs.w.rank$fileName %in% PWDfull.cell.list$fileName,]
-#ok these should be full cells
-
-#there are 2x as many 0's bc 1 and 0 are binned together
-#update all hand.foci.counts
-#y <- droplevels(y)
-
-#length.by.class <- ggplot(data=hh2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=hand.foci.count, shape=mouse))+ggtitle("PWD full cell SC lengths")
-#most cells are from 1 mouse
-
-DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName %in% full.cell.list$fileName,]
-DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName != "20jul15_18may15_PWD_m1_sp1_17_rev",]
-
-qq <- ggplot(data=DF.w.rank2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=rank, shape=mouse))+ggtitle("PWD full cell SC lengths")
-#most cells are from 1 mouse
-
-#take the shortest 5? (lower 25%)
-#take the shortest 3? (seems to be natural cutoff, lowest 3 are similar)
-
-#bin.long.PWD <- DF.w.rank[DF.w.rank$rank]
-#can't do long yet, bc the coding for XY is inconsistant!
-
-bin.short.PWD <- PWD_whole.cell.BivData[PWD_whole.cell.BivData$rank < 5,]
-
-bin.short.PWD <- bin.short.PWD[bin.short.PWD$fileName %in% full.cell.list$fileName,]
-
-kk <- ggplot(data=bin.short.PWD, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25)+ggtitle("short bins")
-
-#short bin patterns vs long bins / total?
-#eventually I want to compare NORM FOCI POSITIONS patterns of these bins across PWD and KAZ
-#merging is a little too confusing... 
-#just parse down and do rbind?
-m1 <- merge(PWD_curate_BivData, PWD_hand_measure_BivData.full, by.x = Obj.ID)
-
-#KAZ_curate_BivData
-#KAZ_hand_measure_BivData.full
-
-hist(hh2$rank, breaks =21)
 ```
@@
 
-```{r KAZ.manual.merge}
-#KAZ_hand_measure_BivData
-#PWD_hand_measure_BivData
-#KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]
-
-#1 match the colnames,  #merge, partial match
-colnames(KAZ_curate_BivData)
-colnames(KAZ_hand_measure_BivData.full)
-
-KAZ_hand_measure_BivData.full$centromere_PER_Position <- KAZ_hand_measure_BivData.full$centromere_ABS_Position / KAZ_hand_measure_BivData.full$chromosomeLength
-
-KAZ_hand_measure_BivData.full$curation.notes <- KAZ_hand_measure_BivData.full$notes
-
-#auto; avergeRED, numberCrossOvers, 20
-#manual; 15
-
-#auto.curate
-KAZ.auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
-new.DF <- KAZ_curate_BivData[ , KAZ.auto.c_names, drop = TRUE]
-
-#manual
-KAZ.manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position", "SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
-new.DF2 <- KAZ_hand_measure_BivData.full[ , KAZ.manual_names, drop = TRUE]
-
-#2. merge the 2 DFs
-KAZ.auto.man_BivData <- rbind(new.DF, new.DF2)
-
-#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
-#make better variable name
-KAZ.auto.man_BivData_cells <- ddply(KAZ.auto.man_BivData, c("fileName"), summarise,
-                    number = length(Obj.ID),
-                    #make the rank or
-                     maxChrm_length = max(chromosomeLength),
-                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]
-
-)
-
-#Isolate cells with all autosomes measured
-KAZfull.cell.list <- KAZ.auto.man_BivData_cells[KAZ.auto.man_BivData_cells$number > 17,]
-KAZfull.cell.list <- KAZfull.cell.list[KAZfull.cell.list$number < 21,]#remove the too many cell
-
-#this gives the rank to each biv within each cell  
-KAZ.Bivs.w.rank <- ddply(KAZ.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
-#there are instances of ties,
-
-KAZ_whole.cell.BivData <- KAZ.Bivs.w.rank[KAZ.Bivs.w.rank$fileName %in% KAZfull.cell.list$fileName,]
-#ok these should be full cells
-hist(KAZ_whole.cell.BivData$rank, breaks =25)
-KAZ_whole.cell.BivData <- droplevels(KAZ_whole.cell.BivData)
-table(KAZ_whole.cell.BivData$fileName, KAZ_whole.cell.BivData$rank)
-
-KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")
-
-```
-
 ```{r histograms}
@@
 
+KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")
+
 KAz have more 0CO. Most of the NA's for CO count are from auto- where I was whisy-washy during curation.
+
+#10mar15_PWD_m1: 13w
+#10mar15_PWD_m2: 13w
+#18may15_PWD_m1:  7w -- use this one
+#18may15_PWD_m2: 7w  (seems like not all of the images were run through the algorithm)
 
@@
 
-```{r mergeing.meta.data, echo=FALSE}
-#meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
-#meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)
-
-meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375
-
-#find a way to remove the rows without the right format before setting as as.Date
-#trying to remove these the extra rows..uggh
-#yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]
-#yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
-#yy <- yy[subset(yy$DOB, !(yy$DOB == "e1")),]
-#i think subseting is messing up the dataframes
-
-meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
-
-#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#meta.data.file2 <- add_euth_date(meta.data.file2) #euthdate 1st
-#meta.data.file2 <- add_age(meta.data.file2)  #age second
-##got it to work ... maybe should save adding it to the merged DF
-meta.data.file2 <- add_category(meta.data.file2)
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-meta.data.file2 <- add_euth_date(meta.data.file2)
-meta.data.file2 <- add_age(meta.data.file2)
-
-pwd.meta <- meta.data.file2[meta.data.file2$category == "PWD male",]
-
-#pwd.meta <- add_euth_date(pwd.meta)
-#pwd.meta <- add_age(pwd.meta)
-
-kaz.meta <- meta.data.file2[meta.data.file2$category == "KAZ male",]
-
-#kaz.meta <- add_mouse(kaz.meta)
-#kaz.meta <- add_euth_date(kaz.meta)
-county =1
-for(i in kaz.meta$mouse){
-    templist= strsplit(i, split="_")[[1]]
-    euth <- templist[1]
-    #figure out date conversion
-    kaz.meta$euth.date2[county] <- as.character(euth)
-
-    #kaz.meta$euth.date[county] <- as.Date(kaz.meta$euth.date[county], format='%d%b%y')
-    county= county +1
-}
-
-kaz.meta$euth.date <- as.Date(kaz.meta$euth.date2, format='%d%b%y')
-
-kaz.meta <- add_age(kaz.meta)
-
-#reduce metadata down to 2 cols (mouse, DOB), merge this with MLH1
-MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
-colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
-MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
-
-PWD.KAZ_MLH1.imprv <- merge(PWD.KAZ_MLH1, MouseMetaData.wrkin)#don't apply the all parameters
-#report number with  and without
-
-#add euth date for calculating age
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-PWD.KAZ_MLH1.imprv <- add_euth_date(PWD.KAZ_MLH1.imprv)
-PWD.KAZ_MLH1.imprv <- add_age(PWD.KAZ_MLH1.imprv)
-
-#yesDOB.check <- PWD.KAZ_MLH1.imprv[!(is.na(PWD.KAZ_MLH1.imprv$DOB)),]
-#noDOB.check <- PWD.KAZ_MLH1.imprv[(is.na(PWD.KAZ_MLH1.imprv$DOB)),]#note, date type of data can't do the string matching!
-
-#total, KAZ 144, PWD 139
-#table(noDOB.check$category)#  no DOB, KAZ 67  PWD 133
-#table(yesDOB.check$category) #yes DOB, KAZ 77  PWD 6
-
-# I need to find the ages of more PWD males
-
-```
-##########################
-# MELTED CHRM1 POSITIONS #
-##########################
-
-```{r PWD.melt.foci.data, echo=FALSE}
-library(reshape2)
-
-#newPWD_curate_BivData =
-#  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
-
-#update these dataframes!
-#isolate the passing foci 
-#curBivData_PWD.good.foci  <- newPWD_curate_BivData[newPWD_curate_BivData$foci.pass == 1,]
-
-curBivData_PWD.good.foci <- PWD.auto.man_BivData[PWD.auto.man_BivData$foci.pass == 1,]
-
-#create normalized foci values!
-curBivData_PWD.good.foci$nrm.F1 <- curBivData_PWD.good.foci$Foci1 / curBivData_PWD.good.foci$chromosomeLength
-curBivData_PWD.good.foci$nrm.F2 <- curBivData_PWD.good.foci$Foci2 / curBivData_PWD.good.foci$chromosomeLength
-curBivData_PWD.good.foci$nrm.F3 <- curBivData_PWD.good.foci$Foci3 / curBivData_PWD.good.foci$chromosomeLength
-
-#melting the DataFrame is needed to ploting the normalized foci positions
-
-#do not include cols for IFD, or ABS.foci
-curBivData_PWD.melt <- melt( curBivData_PWD.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count","nrm.F1","nrm.F2","nrm.F3",  "mouse", "category",   "strain" , "sex")] ,
-
-                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
-              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count","mouse", "category",   "strain" , "sex"))
-
-colnames(curBivData_PWD.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count", "mouse", "category",   "strain" , "sex",
-                                       #name for the 2 melted cols
-                                      "Foci.Type.nrm", "Foci.nrm.pos")
-
-#remove NAs
-curBivData_PWD.melt <- curBivData_PWD.melt[!(is.na(curBivData_PWD.melt$Foci.nrm.pos) | curBivData_PWD.melt$Foci.nrm.pos == ""), ]
-
-#table(newPWD_curate_BivData$fileName)
-#table(newPWD_curate_BivData$mouse) #4 mice, ~100 each mouse
-
-```
-
-```{r melt.KAZ, echo=FALSE}
-
-#use the combined - auto and manual data
-
-curBivData_KAZ.good.foci <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$foci.pass == 1,]
-#curated objects
-#KAZ_m2_curate_bivData =
-#  read.csv("~./MLH1repo/data/BivData/curated_21may18_KAZ_m2_BivData.csv", header = TRUE)
-
-KAZ_m2_curate_bivData$nrm.F1.pos <- KAZ_m2_curate_bivData$Foci1.pos / KAZ_m2_curate_bivData$chromosomeLength
-KAZ_m2_curate_bivData$nrm.F2.pos <- KAZ_m2_curate_bivData$Foci2.pos / KAZ_m2_curate_bivData$chromosomeLength
-KAZ_m2_curate_bivData$nrm.F3.pos <-   KAZ_m2_curate_bivData$Foci3.pos / KAZ_m2_curate_bivData$chromosomeLength
-
-#remove those that don't pass
-KAZ_m2_curate_bivData_SC_pass <- KAZ_m2_curate_bivData[KAZ_m2_curate_bivData$SC.pass == 1,]
-#what else...
-
-table(KAZ_m2_curate_bivData_SC_pass$fileName)
-
-#melt the foci positions, best thing to compare for cells that don't have all the bivalents
-#ggplots
-
-show.plot <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=chromosomeLength, x=fileName))+geom_point()
-
-show.plot2 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=Foci1.pos, x=hand.foci.count))+geom_jitter()
-
-show.plot99 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=nrm.F1.pos, x=hand.foci.count))+geom_jitter()+
-  ggtitle("Normalized foci positions")
-
-count.hist <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(hand.foci.count))+geom_histogram()
-
-```
-
-```{r meltKAZ, echo=FALSE}
-
-#melt KAZ
-#isolate the passing foci 
-#curBivData_KAZ.good.foci  <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
-
-curBivData_KAZ.good.foci <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$foci.pass == 1,]
-colnames(curBivData_KAZ.good.foci)
-
-curBivData_KAZ.good.foci$nrm.F1.pos <- curBivData_KAZ.good.foci$Foci1 / curBivData_KAZ.good.foci$chromosomeLength
-curBivData_KAZ.good.foci$nrm.F2.pos <- curBivData_KAZ.good.foci$Foci2 / curBivData_KAZ.good.foci$chromosomeLength
-curBivData_KAZ.good.foci$nrm.F3.pos <-   curBivData_KAZ.good.foci$Foci3 / curBivData_KAZ.good.foci$chromosomeLength
-
-#melting the DataFrame is needed to ploting the normalized foci positions
-
-#do not include cols for IFD, or ABS.foci
-curBivData_KAZ.melt <- melt( curBivData_KAZ.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,
-
-                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
-              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count","mouse", "category",   "strain" , "sex"))
-
-colnames(curBivData_KAZ.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
-    "hand.foci.count", "mouse", "category",   "strain" , "sex",
-                                       #name for the 2 melted cols
-                                      "nrmFoci.Type", "nrmFoci.pos")
-
-#remove NAs
-curBivData_KAZ.melt <- curBivData_KAZ.melt[!(is.na(curBivData_KAZ.melt$nrmFoci.Type) | curBivData_KAZ.melt$nrmFoci.pos==""), ]
-
-```
-
 ```{r BivData.set, echo=FALSE}
@@
 #melt the BivData?
-```
-
-```{r curatedBivData}
-
-curBivData_KAZ = read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE, sep = ',')
-#curBivData_PWD = read.csv("~./MLH1repo/data/BivData/curate_PWD_BivData.csv", header = TRUE, sep = ',')
-
-#make normalized
-
-curBivData_KAZ$nrm.F1.pos <- curBivData_KAZ$Foci1 / curBivData_KAZ$chromosomeLength
-curBivData_KAZ$nrm.F2.pos <- curBivData_KAZ$Foci2 / curBivData_KAZ$chromosomeLength
-curBivData_KAZ$nrm.F3.pos <-   curBivData_KAZ$Foci3 / curBivData_KAZ$chromosomeLength
-
-curBivData_PWD$nrm.F1.pos <- curBivData_PWD$Foci1 / curBivData_PWD$chromosomeLength
-curBivData_PWD$nrm.F2.pos <- curBivData_PWD$Foci2 / curBivData_PWD$chromosomeLength
-curBivData_PWD$nrm.F3.pos <-   curBivData_PWD$Foci3 / curBivData_PWD$chromosomeLength
-
-#table / what are the numbers of observations
-table(curBivData_PWD$fileName)
-#biv measures per cell
-
-SC.passed.PWD <- curBivData_PWD[curBivData_PWD$SC.pass == 1,]#299 -> 208
-table(SC.passed.PWD$fileName)
-
-SC.passed.KAZ <- curBivData_KAZ[curBivData_KAZ$SC.pass == 1,]#299 -> 208
-
 ```
@@
 
-```{r fit.exp}
+```{r fit.exp, echo=FALSE}
 #fitting IFDs to expoential Distribution
@@
 
-fit.PWD <- fitdistr(all.2CO$IFD1_PER[all.2CO$strain == "PWD"], "exponential")#change to gamma
-fit.KAZ <- fitdistr(all.2CO$IFD1_PER[all.2CO$strain== "KAZ"], "exponential")
+fit.PWD <- fitdistr(all.2CO$IFD1_ABS[all.2CO$strain == "PWD"], "gamma")#change to gamma, (1 outlier)
 
-PWD.curve <- hist(all.2CO$IFD1_PER[all.2CO$strain == "PWD"], freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
+fit.KAZ <- fitdistr(all.2CO$IFD1_ABS[all.2CO$strain== "KAZ"], "gamma")
+
+PWD.curve <- hist(all.2CO$IFD1_ABS[all.2CO$strain == "PWD"], freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
 curve(dexp(x, rate = fit.PWD$estimate), from = 0, col = "red", add = TRUE)
@@
 
-```{r rand.PWD, echo=FALSE}
-
-PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]#670
-cur18may15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m1",]
-cur18may15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m2",]
-
-cur10mar15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m1",]
-cur10mar15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m2",]
-
-table(PWD.BivData.curate.list$mouse)
-#4mice in the current dataset
-
-#10mar15_PWD_m1: 13w
-#10mar15_PWD_m2: 13w
-#18may15_PWD_m1:  7w -- use this one
-#18may15_PWD_m2: 7w  (seems like not all of the images were run through the algorithm)
-
-```
-
-```{r KAZ.curate.list, echo=FALSE}
+```{r makeKAZ.curate.list, echo=FALSE}
 #write out list for initial quality control and subsetting
-
 #remove centromere_PER > .2
-
 KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123
@@
 
-Kaz.biv.plot <- ggplot(data = KAZ.Biv.curate.list, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()+facet_wrap(~mouse)
-
-#3000 to curate
-
 cur11jan18_KAZ_m4 <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$mouse == "11jan18_KAZ_m4",]
@@
 write.table(cur11jan18_KAZ_m4, "~./MLH1repo/results/curate_cur11jan18_KAZ_m4_BivData.txt", sep="\t", row.names = FALSE)
-
-#KAZ_mice
-
-#11jan18_KAZ_m2-curated file
-#11jan18_KAZ_m3-
-#11jan18_KAZ_m4-
 
@@
 ```
+
+PWD.auto.man_BivData <- PWD.auto.man_BivData[ -c(19:25) ]
+
+#y <- droplevels(y)
+
+#length.by.class <- ggplot(data=hh2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=hand.foci.count, shape=mouse))+ggtitle("PWD full cell SC lengths")
+#most cells are from 1 mouse
+
+DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName %in% full.cell.list$fileName,]
+DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName != "20jul15_18may15_PWD_m1_sp1_17_rev",]
+
+qq <- ggplot(data=DF.w.rank2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=rank, shape=mouse))+ggtitle("PWD full cell SC lengths")
+#most cells are from 1 mouse
+
+#take the shortest 5? (lower 25%)
+#take the shortest 3? (seems to be natural cutoff, lowest 3 are similar)
+
+#bin.long.PWD <- DF.w.rank[DF.w.rank$rank]
+#can't do long yet, bc the coding for XY is inconsistant!
+
+Kaz.biv.plot <- ggplot(data = KAZ.Biv.curate.list, aes(x= centromere_PER_Position, y=chromosomeLength) )+geom_point()+facet_wrap(~mouse)
+
+#3000 to curate

### 2019-07-19 results/PWD.KAZ.BivData.Rmd
cleaning up files and made some additions to analyses files

@@
 #setwd("C:/Users/April/Desktop/MLH1repo/") #don't have two
-load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_62419.RData")
+
+#DONT LOAD MLH1 yet -- keep these seperate for now
+#load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_62419.RData")
 #MLH1_data_setup_62419.RData
@@
 KAZ.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "KAZ",]
+```
 
-```
+ToDo:
+  -add in IFD calqs
+  -add rbar calqs
+  -make the basic plots ect
 
@@
 whole.cell_table <- whole.cell_table[whole.cell_table$number < 21,]
-whole.cell.list <- unique(whole.cell.DF$fileName)
+whole.cell.list <- unique(whole.cell_table$fileName)
 
@@
 long.bin.Bivs <- whole.cell.DF[whole.cell.DF$rank > 14,]
-
 ```
@@
 whole.cell_melt.DF <- ddply(whole.cell_melt.DF, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
-
 ```
@@
 
-save.image("data/MLH1/MLH1_data_setup_62419.RData")
+save.image("~./MLH1repo/data/BivData/PWD.KAZ_BivData_data.RData")
 ```
 
+From the cell level tables -- make lists of cells with ~75% of bivalents quantified. These will be manually verified and the missing chrms will be measured and supplemented.
+
+Data frames and analyses are the main output of this script - for a saved Rdata file.
+
+1. Curated BivData; read files, merge/rbind together
+
+2. Hand Measures; read files, rbind together
+
+3. Auto+manual; combine the curate and hand measures, rbind both strains together.
+    3B. Auto-Manual_whole_cell, make a list of fileNames for extracting these measures later; the biv from cells with all 19 autosome measures. Add bivalent rank to this DF.
+
+    -for all measures with passing foci, add IFD and rbar!
+
+4. melted DF; Melt the Foci-Pass of Auto_Manual_DF
+
+    4B. whole.cell melted table
+
+### Analysis and Such
+
+--- Once the above DF's are made, proceed with following analyses ---
+
+1. SC Length
+    a. raw / loose bivalents
+    b. binned bivalents, short vs long, chrm class
+
+2. Normalized Foci Position
+   a. all/full
+   b. bins
+
+3. 2-CO
+    a. IFDs, raw norm
+    b. F1 by F2
+
+4. Rbar?
+
+## Background and Premise
+
+PWD and KAZ differ in gwRR/MLH1 counts, is there a correlation in meiotic features which go with expected differences in gwRR.
 ```{r SClength.cell.tables}
@@
 ```
-
-From the cell level tables -- make lists of cells with ~75% of bivalents quantified. These will be manually verified and the missing chrms will be measured and supplemented.
-
-Data frames and analyses are the main output of this script - for a saved Rdata file.
-
-1. Curated BivData; read files, merge/rbind together
-
-2. Hand Measures; read files, rbind together
-
-3. Auto+manual; combine the curate and hand measures, rbind both strains together.
-    3B. Auto-Manual_whole_cell, make a list of fileNames for extracting these measures later; the biv from cells with all 19 autosome measures. Add bivalent rank to this DF.
-
-4. melted DF; Melt the Foci-Pass of Auto_Manual_DF
-
---- Once the above DF's are made, proceed with following analyses ---
-
-1. SC Length
-    a. raw / loose bivalents
-    b. binned bivalents, short vs long, chrm class
-
-2. Normalized Foci Position
-   a. all/full
-   b. bins
-
-3. 2-CO
-    a. IFDs, raw norm
-    b. F1 by F2
-
-4. Rbar?
-
-## Background and Premise
-
-PWD and KAZ differ in gwRR/MLH1 counts, is there a correlation in meiotic features which go with expected differences in gwRR.
 

### 2019-08-15 results/PWD.KAZ.BivData.Rmd
added and curated abunch of bivalent data. started some analysis

@@
 
+#todo -- load in female data
+
 setwd("~./MLH1repo/")
@@
 #DONT LOAD MLH1 yet -- keep these seperate for now
-#load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_62419.RData")
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_62419.RData")
 #MLH1_data_setup_62419.RData
@@
 #1. Curated Data
+PWD.KAZ_curate_BivData.full =
+  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD.KAZ_BivData.csv", header = TRUE)
+
+PWD.KAZ_curate_BivData.full <- PWD.KAZ_curate_BivData.full[!(is.na(PWD.KAZ_curate_BivData.full$SC.pass) ),]
+
+PWD.KAZ_curate_BivData <- PWD.KAZ_curate_BivData.full[PWD.KAZ_curate_BivData.full$SC.pass == 1,]
+
 #PWD set 4 mice, ~100 objs each)
-PWD_curate_BivData.full =
-  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
+#PWD_curate_BivData.full =
+#  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
 #remove bad segmentation
-PWD_curate_BivData <- PWD_curate_BivData.full[PWD_curate_BivData.full$SC.pass == 1,]
+#PWD_curate_BivData <- PWD_curate_BivData.full[PWD_curate_BivData.full$SC.pass == 1,]
 
 #KAZ 3 mice >200 objs each
-KAZ_curate_BivData.full =
-  read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE)
-KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]
+#KAZ_curate_BivData.full =
+#  read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE)#not sure where to find this file
+#KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]
 
 #merged KAZ and PWD
-curate_BivData <- rbind(KAZ_curate_BivData, PWD_curate_BivData)
+#curate_BivData <- rbind(KAZ_curate_BivData, PWD_curate_BivData)
+
+#incorporate female data
+PWD.KAZ_BivData_8.14.19_curated.full = 
+  read.csv("~./MLH1repo/data/BivData/PWD.KAZfemale_curated_8.15.19.csv", header = TRUE)#740
+#merge and check that they have the same columns
+str(PWD.KAZ_BivData_8.14.19_curated.full)
+
+PWD.KAZ_BivData <- PWD.KAZ_BivData_8.14.19_curated.full[!(is.na(PWD.KAZ_BivData_8.14.19_curated.full$SC.pass) ),]
+
+PWD.KAZ_BivData <- PWD.KAZ_BivData[PWD.KAZ_BivData$SC.pass == 1,]
+
+str(PWD.KAZ_BivData)
+
+curate_PWD.KAZ_BivData_M.F <- rbind(PWD.KAZ_curate_BivData, PWD.KAZ_BivData)
+str(curate_PWD.KAZ_BivData_M.F)
+
+#forgot why I'm removing these
+#curate_PWD.KAZ_BivData_M.F <- curate_PWD.KAZ_BivData_M.F[curate_PWD.KAZ_BivData_M.F$Obj.ID != "25feb15_8oct14_PWD_f8_sp1_1_rev_19",]
+#curate_PWD.KAZ_BivData_M.F <- curate_PWD.KAZ_BivData_M.F[curate_PWD.KAZ_BivData_M.F$Obj.ID != "9aug15_23apr15_PWD_f2_sp1_9.1_rev_15",]
+#curate_PWD.KAZ_BivData_M.F <- curate_PWD.KAZ_BivData_M.F[curate_PWD.KAZ_BivData_M.F$Obj.ID != "9aug15_23apr15_PWD_f2_sp1_9.1_rev_18",]
+
+#male+female KAZ and PWD bivData
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+curate_PWD.KAZ_BivData_M.F <- add_mouse(curate_PWD.KAZ_BivData_M.F)
+curate_PWD.KAZ_BivData_M.F <- add_category(curate_PWD.KAZ_BivData_M.F)
+curate_PWD.KAZ_BivData_M.F <- add_strain(curate_PWD.KAZ_BivData_M.F)
+curate_PWD.KAZ_BivData_M.F <- add_sex(curate_PWD.KAZ_BivData_M.F)
 
@@
 #make indevidal DFs by strain just in case...
-PWD.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "PWD",]
-KAZ.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "KAZ",]
+#PWD.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "PWD",]
+#KAZ.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "KAZ",]
 ```
@@
   -make the basic plots ect
+  -analyze female patterns
+
+  -what are the distributions of female IFDs and nrm IFDs
+  -what are the proportions of chrm classes?
 
@@
 #merg4.PWD_hand_measure_BivData.full <- #manual.blobject.name
+#curate_PWD.KAZ_BivData_M.F
 
@@
 auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
-new.DF <- curate_BivData[ , auto.c_names, drop = TRUE]
+new.DF <- curate_PWD.KAZ_BivData_M.F[ , auto.c_names, drop = TRUE]
 
@@
 
+table(Auto.Man_BivData$hand.foci.count[Auto.Man_BivData$sex == "female"], Auto.Man_BivData$strain[Auto.Man_BivData$sex == "female"])
+
+#remove empty cells
+#remove extra empty rows
+Auto.Man_BivData <- Auto.Man_BivData[!(is.na(Auto.Man_BivData$SC.pass) | Auto.Man_BivData$SC.pass == ""), ]
+Auto.Man_BivData <- Auto.Man_BivData[(Auto.Man_BivData$SC.pass == "1"), ]
+
 #make WHOLE CELL DF
-
 #some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
@@
 #make PWD and KAZ specific melt.DF ... if you want
-
 #make whole cell melt
@@
 
+```{r Female.Nrm.Pos}
+melt.PWD.female  <- Auto.Man_BivData.melt[(Auto.Man_BivData.melt$sex == "female") & (Auto.Man_BivData.melt$strain == "PWD"),]
+melt.KAZ.female <- Auto.Man_BivData.melt[(Auto.Man_BivData.melt$sex == "female") & (Auto.Man_BivData.melt$strain == "KAZ"),]
+
+nPWD.F.obs <- ddply(.data=melt.PWD.female, 
+                 .(hand.foci.count),
+                 summarize, 
+                 n=paste("n =", length(unique(Obj.ID))))
+
+nKAZ.F.obs <- ddply(.data=melt.KAZ.female, 
+                 .(hand.foci.count),
+                 summarize, 
+                 n=paste("n =", length(unique(Obj.ID))))
+
+#figures for foci positions
+melt.plot.PWD.female    <- ggplot(melt.PWD.female,
+              aes(x=nrmFoci.pos, fill=nrmFoci.Type) ) +
+
+  geom_histogram(alpha= 0.5,stat = "bin", position= "identity")+
+ facet_wrap(~ hand.foci.count, scales="free", ncol=1)+ggtitle("PWD female norm foci positions")+
+ geom_text(data=nPWD.F.obs, aes(x=.25, y=7, label=n), 
+         colour="black", inherit.aes=FALSE, parse=FALSE)+xlim(c(0,1)) + ylim(c(0,13))
+
+melt.plot.KAZ.female    <- ggplot(melt.KAZ.female,
+              aes(x=nrmFoci.pos, fill=nrmFoci.Type) ) + 
+  geom_histogram(alpha= 0.5,stat = "bin", position= "identity")+
+ facet_wrap(~ hand.foci.count, scales="free", ncol=1)+ggtitle("KAZ female norm foci positions")+
+ geom_text(data=nKAZ.F.obs, aes(x=.25, y=7, label=n), 
+         colour="black", inherit.aes=FALSE, parse=FALSE)+xlim(c(0,1)) + ylim(c(0,13))
+
+#t.tests for mean F positions
+CO1.F1.t.test <- t.test(melt.KAZ.female$nrmFoci.pos[(melt.KAZ.female$nrmFoci.Type == "nrm.F1.pos") &(melt.KAZ.female$hand.foci.count == 1)],
+                        melt.PWD.female$nrmFoci.pos[(melt.PWD.female$nrmFoci.Type == "nrm.F1.pos") &(melt.PWD.female$hand.foci.count == 1)]  )
+
+CO2.F1.t.test <- t.test(melt.KAZ.female$nrmFoci.pos[(melt.KAZ.female$nrmFoci.Type == "nrm.F1.pos") &(melt.KAZ.female$hand.foci.count == 2)],
+                        melt.PWD.female$nrmFoci.pos[(melt.PWD.female$nrmFoci.Type == "nrm.F1.pos") &(melt.PWD.female$hand.foci.count == 2)] )
+
+CO2.F2.t.test <- t.test(melt.KAZ.female$nrmFoci.pos[(melt.KAZ.female$nrmFoci.Type == "nrm.F2.pos") &(melt.KAZ.female$hand.foci.count == 2)],
+                        melt.PWD.female$nrmFoci.pos[(melt.PWD.female$nrmFoci.Type == "nrm.F2.pos") &(melt.PWD.female$hand.foci.count == 2)] )
+#none of the positions are significantly different
+
+```
+
+```{r}
+#another set of plots
+
+PWD.female <- Auto.Man_BivData[(Auto.Man_BivData$strain == "PWD") & (Auto.Man_BivData$sex == "female"), ]
+KAZ.female <- Auto.Man_BivData[(Auto.Man_BivData$strain == "KAZ") & (Auto.Man_BivData$sex == "female"), ]
+
+#figures for foci positions
+PWD.female_foci1_foci2_plot    <- ggplot(PWD.female[PWD.female$hand.foci.count == 2,],  aes(x=Foci1, y=Foci2, color = chromosomeLength) ) + geom_point()+
+  geom_abline(intercept = .5, slope = 1, color="red")+xlim(c(0,100)) + ylim(c(0,200))+ggtitle("PWD female")+scale_colour_gradientn(colours=rainbow(5))
+#the 5 in rainbow increases the difference between colors
+
+KAZ.female_foci1_foci2_plot    <- ggplot(KAZ.female[KAZ.female$hand.foci.count == 2,],  aes(x=Foci1, y=Foci2, color = chromosomeLength) ) + geom_point()+geom_abline(intercept = .5, slope = 1, color="red")+xlim(c(0,100)) + ylim(c(0,200))+ggtitle("KAZ female")+scale_colour_gradientn(colours=rainbow(5))
+
+#make points slightly bigger
+
+```
+
 ```{r saveRdata}
@@
 
-save.image("~./MLH1repo/data/BivData/PWD.KAZ_BivData_data.RData")
+save.image("~./MLH1repo/data/BivData/PWD.KAZ_BivData_data_8.14.19.RData")
 ```
@@
 
-kk <- ggplot(data=bin.short.PWD, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25)+ggtitle("short bins")
-
 ```{r SC.lengths}
@@
 
+##kk <- ggplot(data=bin.short.PWD, aes(x=mouse,y=chromosomeLength))+geom_jitter(width = 0.25)+ggtitle("short bins")
+
 #strange number of low chrm lengths
@@
 
-SC.PWD <- ggplot(data = PWD_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,150))+ggtitle("PWD SC lengths")
+table(Auto.Man_BivData$sex)
+
+PWD.Auto.Man_BivData <- Auto.Man_BivData[Auto.Man_BivData$strain == "PWD",]
+
+SC.PWD <- ggplot(data = PWD.Auto.Man_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,250))+ggtitle("PWD SC lengths") +facet_wrap(~sex)
 #the SC range is 45 - 150 (for PWD)
 
-SC.KAZ <- ggplot(data = KAZ_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+
-  theme(legend.position="none")+ylim(c(35,150))+ggtitle("KAZ SC lengths")
+KAZ.Auto.Man_BivData <- Auto.Man_BivData[Auto.Man_BivData$strain == "KAZ",]
+
+SC.KAZ <- ggplot(data = KAZ.Auto.Man_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+
+  theme(legend.position="none")+ylim(c(35,250))+ggtitle("KAZ SC lengths")+facet_wrap(~sex)
 #can I add the means?
@@
 
-KAZ_mean_SC <- ddply(KAZ_curate_BivData, c("mouse"), summarise,
+str(PWD.Auto.Man_BivData)
 
-                    boxs.IDd = max(boxNumber),
-                    nboxes_passed = length(boxNumber),
+KAZ_mean_SC <- ddply(KAZ.Auto.Man_BivData, c("mouse","sex"), summarise,
+
+                    boxs.IDd = max(as.numeric(boxNumber)),
+                    nboxes_passed = length(as.numeric(boxNumber)),
+                    pass_rate = nboxes_passed/boxs.IDd,#I don't know if pass rate is working
+
+                    avSC.length = mean(chromosomeLength),
+                    varSC.length = var(chromosomeLength)
+)
+
+PWD_mean_SC <- ddply(PWD.Auto.Man_BivData, c("mouse","sex"), summarise,
+
+                    boxs.IDd = max(as.numeric(boxNumber)),
+                    nboxes_passed = length(as.numeric(boxNumber)),
                     pass_rate = nboxes_passed/boxs.IDd,
@@
 
-PWD_mean_SC <- ddply(PWD_curate_BivData, c("mouse"), summarise,
+t.test(KAZ.Auto.Man_BivData$chromosomeLength[KAZ.Auto.Man_BivData$sex == "male"],PWD.Auto.Man_BivData$chromosomeLength[PWD.Auto.Man_BivData$sex == "male"]   )#p is small
 
-                    boxs.IDd = max(boxNumber),
-                    nboxes_passed = length(boxNumber),
-                    pass_rate = nboxes_passed/boxs.IDd,
-
-                    avSC.length = mean(chromosomeLength),
-                    varSC.length = var(chromosomeLength)
-)
-
-t.test(KAZ_curate_BivData$chromosomeLength,PWD_curate_BivData$chromosomeLength)
+t.test(KAZ.Auto.Man_BivData$chromosomeLength[KAZ.Auto.Man_BivData$sex == "female"],PWD.Auto.Man_BivData$chromosomeLength[PWD.Auto.Man_BivData$sex == "female"]   )#p not as small
 
@@
 
-the PWD has longer means
+The PWD has longer mean SC. Female KAZ are a bit larger. PWD female aren't as small.
 
@@
 
-KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")
+KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength)) + geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")
 
@@
 #10mar15_PWD_m2: 13w
-#18may15_PWD_m1:  7w -- use this one
+
+#18may15_PWD_m1:  7w -- use this one... these are good cells
 #18may15_PWD_m2: 7w  (seems like not all of the images were run through the algorithm)
+
+```{r}
+
+#test the correlation of F points between cells
+#but make sure
+
+#all.2CO
+#all.2CO_PWD
+
+Foci_Corr.PWD <- ggplot(data=whole.cell.2CO.PWD, aes(x=Foci1.PER, y=Foci2.PER))+geom_point( aes(color=as.factor(rank) ) )+ggtitle("correletaion of Nrm Foci Posisitons")+facet_wrap(~mouse) +ylim(c(0,1))+xlim(c(0,1))
+
+#theme(legend.position="none")
+
+whole.cell.KAZ$Foci1.PER <- whole.cell.KAZ$Foci1 / whole.cell.KAZ$chromosomeLength
+whole.cell.KAZ$Foci2.PER <- whole.cell.KAZ$Foci2 / whole.cell.KAZ$chromosomeLength
+whole.cell.KAZ$Foci3.PER <- whole.cell.KAZ$Foci3 / whole.cell.KAZ$chromosomeLength
+
+whole.cell.KAZ.2CO <- whole.cell.KAZ[whole.cell.KAZ$hand.foci.count == 2,]
+whole.cell.KAZ.2CO <- droplevels(whole.cell.KAZ.2CO)
+
+whole.cell.KAZ.2CO <- whole.cell.KAZ.2CO[!(is.na(whole.cell.KAZ.2CO$Foci1.PER) | whole.cell.KAZ.2CO$Foci1.PER == ""), ]
+
+Foci_Corr.KAZ <- ggplot(data=whole.cell.KAZ.2CO, aes(x=Foci1.PER, y=Foci2.PER))+geom_point(aes(color=interaction(fileName,mouse) ) )+ggtitle("correletaion of Nrm Foci Posisitons")+facet_wrap(~mouse) +ylim(c(0,1))+xlim(c(0,1))
+
+whole.cell.2CO.PWD <- whole.cell.2CO.PWD[whole.cell.2CO.PWD$hand.foci.count == 2,]
+whole.cell.2CO.PWD <- droplevels(whole.cell.2CO.PWD)
+
+table(whole.cell.2CO.PWD$hand.foci.count)
+
+Foci_Corr.PWD2 <- ggplot(data=whole.cell.2CO.PWD, aes(x=Foci1.PER, y=Foci2.PER))+geom_point(aes(color=interaction(fileName,mouse) ) )+ggtitle("correletaion of Nrm Foci Posisitons")+facet_wrap(~mouse) +ylim(c(0,1))+xlim(c(0,1))
+
+Foci_Corr.PWD2 <- ggplot(data=whole.cell.2CO.PWD, aes(x=Foci1.PER, y=Foci2.PER))+geom_point(aes(color=interaction(fileName,mouse) ) )+ggtitle("correletaion of Nrm Foci Posisitons")+facet_wrap(~mouse) +ylim(c(0,1))+xlim(c(0,1))
+
+PWD.1 <- whole.cell.2CO.PWD[whole.cell.2CO.PWD$mouse == "1mar15_PWD_m1",]
+
+PWD.1$Per2_Per1 <- PWD.1$Foci2.PER - PWD.1$Foci1.PER
+
+Foci_Corr.PWD99 <- ggplot(data=PWD.1, aes(x=Foci1.PER, y=Foci2.PER))+geom_point(aes(color=interaction(fileName) ) )+ggtitle("correletaion of Nrm Foci Posisitons")+facet_wrap(~fileName) +ylim(c(0,1))+xlim(c(0,1))+
+  geom_abline(intercept = .5, slope = 1, color="red")+
+  geom_abline(intercept = 0, slope = 1)
+
+#making slopes of 1 (then they are horizontal)
+#the slope of blue lines doesn't fit that slope
+
+#1:1 slope should be interperted as consistant process / rule within cell
+
+#strong pos correlation of Foci positions
+
+Foci_Corr.PWD88 <- ggplot(data=PWD.1, aes(x=Foci1, y=Foci2))+geom_point(aes(color=interaction(rank) ) )+ggtitle("correletaion of Nrm Foci Posisitons")+facet_wrap(~fileName) 
+#not sure what the raw IFD pattern shows
+#I should learn to add daigonal lines as bench marks
+
++ylim(c(0.6,1))+xlim(c(0,1))
+
+#comparing KAZ, it looks like Foci 1 is further from the centromere (towards the middle)
+#
+
+#whole.cell.2CO.PWD -- I can fill color by rank, .. (2COs are biased to longer chrms, but there are some shorter ones
+
+blah2 <- ggplot(data=all.2CO, aes(x=Foci1,y=Foci2))+geom_jitter(width = 0.25, aes(color=mouse))+ggtitle("correletaion of Foci Posisitons")+facet_wrap(~strain)+ theme(legend.position="none")+ylim(c(0,150))+xlim(c(0,90))
+
+#plot stacked IFD lines -- linking the foci
+
+#Can I calculate esitmated rank?
+
+```
 
@@
 
+find a way to analyze the 2CO / IFD for the female data
+
 ```{r chrm.class, echo=FALSE}
@@
 
+#Auto.Man_BivData
+#PWD.Auto.Man_BivData
+#KAZ.Auto.Man_BivData
+
+ookk <- ggplot(data = Auto.Man_BivData[Auto.Man_BivData$sex == "female",], aes(x=chromosomeLength))+geom_density(aes(fill=hand.foci.count), alpha = .3)+facet_wrap(strain~sex)
+#there are too many bad levels for hand.foci.count
+
+ookk2 <- ggplot(data = Auto.Man_BivData[Auto.Man_BivData$sex == "female",], aes(x=chromosomeLength))+geom_histogram(aes(fill=hand.foci.count), alpha = .3)+facet_wrap(strain~sex)
+
 #isolate the 2COs
 #total Bivdata
-KAZ.2COs <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$hand.foci.count == 2,]
-PWD.2COs <- PWD.auto.man_BivData[PWD.auto.man_BivData$hand.foci.count == 2,]
+KAZ.2COs <- KAZ.Auto.Man_BivData[KAZ.Auto.Man_BivData$hand.foci.count == 2,]
+PWD.2COs <- PWD.Auto.Man_BivData[PWD.Auto.Man_BivData$hand.foci.count == 2,]
 
@@
 #all.2CO <- all.2CO[ -c(24) ]
+
+table(Auto.Man_BivData$mouse[Auto.Man_BivData$sex == "female"])
+
+table(all.2CO$mouse[all.2CO$sex=="female"])
+
+#main female mice for comparison
+#23apr15_PWD_f2, 23apr15_PWD_f2
 
@@
 
-IFDs.ABS <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_ABS))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("raw IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 90, hjust = 1) )
+IFDs.ABS <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_ABS))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("raw IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(sex~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 90, hjust = 1) )
 
-IFDs.PER <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("norm IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+theme(legend.position="none")
+IFDs.ABS.22 <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_ABS, fill=interaction(strain,sex)))+geom_jitter(width = 0.25)+ggtitle("raw IFD Distributions")+geom_boxplot(alpha = 0.6)+  theme(legend.position="none", axis.text.x = element_text(angle = 90, hjust = 1) )
+
+IFDs.PER <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("norm IFD Distributions")+geom_boxplot(alpha = 0.3)+facet_wrap(sex~strain)+theme(legend.position="none")+theme(legend.position="none", axis.text.x = element_text(angle = 90, hjust = 1) )
+
+#remove mice with very few observations?
 
@@
 #remove centromere_PER > .2
-KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123
-
+#KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123
 #remove chrmLength > 150? 200
-KAZ.Biv.curate.list <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$chromosomeLength < 300, ]#not completely sure 300 is probably the max length is the best length
+#KAZ.Biv.curate.list <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$chromosomeLength < 300, ]#not completely sure 300 is probably the max length is the best length
 #3117
-
-table(KAZ.Biv.curate.list$mouse)
+#table(KAZ.Biv.curate.list$mouse)
 #7 mice: 11jan18_KAZ_m4(9w), 1feb18_KAZ_m1(11w), 11jan18_KAZ_m3(9w), and 11jan18_KAZ_m1(43w) have good MLH1
-
-cur11jan18_KAZ_m4 <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$mouse == "11jan18_KAZ_m4",]
-
-cur11jan18_KAZ_m4$Obj.ID <- paste(cur11jan18_KAZ_m4$fileName, cur11jan18_KAZ_m4$boxNumber, sep = "_")
-
-write.table(cur11jan18_KAZ_m4, "~./MLH1repo/results/curate_cur11jan18_KAZ_m4_BivData.txt", sep="\t", row.names = FALSE)
-
+#cur11jan18_KAZ_m4 <- KAZ.Biv.curate.list[KAZ.Biv.curate.list$mouse == "11jan18_KAZ_m4",]
+#cur11jan18_KAZ_m4$Obj.ID <- paste(cur11jan18_KAZ_m4$fileName, cur11jan18_KAZ_m4$boxNumber, sep = "_")
+#write.table(cur11jan18_KAZ_m4, "~./MLH1repo/results/curate_cur11jan18_KAZ_m4_BivData.txt", sep="\t", row.names = FALSE)
 ```

### 2019-08-15 src/makeCurateBivDataList.R
wrote a make bivdata curate list script

@@ -1,0 +1,100 @@
+# make curate list from raw BivData
+# input: BBIIGG type bivdata file (raw bivdata from DNAcrossover)
+# output: new file for manual curation
+
+
+#load/read in fullBivdata file
+new_PWD.KAZ_BivData_8.12.19.org = 
+  read.csv("~./MLH1repo/data/BivData/PWD_KAZ_female_BivData_8.12.19.csv", header = TRUE)#740
+
+#(PWD-KAZ females), 8.12.19
+
+#delete 'p_rev' files
+
+
+#filter by centromere
+new_PWD.KAZ_BivData_8.12.19 <- new_PWD.KAZ_BivData_8.12.19.org[new_PWD.KAZ_BivData_8.12.19.org$centromere_PER_Position < 0.2,] #665
+#14000 bivalents
+
+#add Obj.ID
+new_PWD.KAZ_BivData_8.12.19$Obj.ID <- paste(new_PWD.KAZ_BivData_8.12.19$fileName, new_PWD.KAZ_BivData_8.12.19$boxNumber, sep = "_")
+
+#KAZ-PWD female
+#remove bad mice
+
+
+#extra filters?  extra chrms?  too long chrms?
+#20 lower end?
+#mus female: consider cutoff at 200
+#mus male:
+
+#consider altering the cols -- to make curation easier
+#remove all foci position columns
+
+new_PWD.KAZ_BivData_8.12.19$SC.pass <- ""
+new_PWD.KAZ_BivData_8.12.19$foci.pass<- ""
+new_PWD.KAZ_BivData_8.12.19$curated.1pass.0fail<- ""
+new_PWD.KAZ_BivData_8.12.19$curation.notes<- ""
+new_PWD.KAZ_BivData_8.12.19$hand.foci.count	<- ""
+new_PWD.KAZ_BivData_8.12.19$Foci1	<- ""
+new_PWD.KAZ_BivData_8.12.19$Foci2<- ""
+new_PWD.KAZ_BivData_8.12.19$Foci3<- ""
+
+
+#remove this mouse
+#8oct14_PWDf2_sp1
+#PWD_f1_sp1
+
+
+#write to file for making manual notes
+write.table(new_PWD.KAZ_BivData_8.12.19, "~./MLH1repo/results/forCuration_PWD.KAZfemale_8.12.19.txt", sep="\t", row.names = FALSE)
[…]

### 2019-08-15 doc/Main_Report.rmd
put in the new Rdata file to main report

@@
     df_print: paged
+  pdf_document: default
 ---
@@
 #setwd("C:/Users/April/Desktop/MLH1repo")
-load(file="data/MLH1/MLH1_data_setup.RData")
+#load(file="data/MLH1/MLH1_data_setup.RData")
+load(file="data/MLH1/MLH1_data_setup_8.15.19.RData")
+
 ```
@@
 
--  *Report basic statistics of results and to help keep track of the number of images for each category and mouse*
+-  Report basic statistics of results and to help keep track of the number of images for each category and mouse
 
@@
 
-These are the tables for the number of mice and cells which have been quantified. (what is in the 'other' category?)
+These are the tables for the number of cells and mice which have been quantified.
+The 'Other' are F1 crosses.
 
@@
 ```{r first.boxplots, echo=FALSE, fig.align='center'}
+#what?
+#MLH1_data <- MLH1_data[MLH1_data$strain]
+
 #the below ordering works!! keep it
-MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEWES",
-              "WSB", "PWD","MSM","KAZ","CAST","HMI","SPRET","SPIC", "other"), order=T )
+MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
+              "PWD","MSM","MOLF","CZECH","SKIVE",
+              "KAZ","TOM","AST",
+              "CAST","HMI","SPRET","SPIC", "other"), order=T )
 #order the data frame
 MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+#don't know why this won't do the right order...
 #set the order of another column, based on another variable (so that when)
@@
 
-ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+#try this for ordering
+#MLH1_data[order(MLH1_data$strain), ]
+
+#can't re-order the miceagg
+ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
+             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey" )) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5))
-ff <- ff   + facet_wrap(~ sex, scales="free")
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey", 
+        "grey","grey","grey",  "grey","grey","grey" )) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(~ sex, scales="free")
 
 #boxplots by strain
-ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
@@
 
-After taking the data from 2 highest cell quality, some but not all mouse means converge.
+### Distributions of MLH1 counts
 
@@
 
-```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE}
+### Histograms of Mouse Ages
+
+```{r age.hists, echo=FALSE}
+#meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
+meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)
+meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375
+
+meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
+
+#make Mouse meta data work
+MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
+colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
+MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
+
+#do this for the mouse table instead of the MLH1 table
+MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+MLH1.imprv <- add_euth_date(MLH1.imprv)
+#
+MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
+MLH1.imprv$euth.date <- as.Date(as.character(MLH1.imprv$euth.date2), format='%d%b%y')
+
+MLH1.imprv <- add_age(MLH1.imprv)
+
+#remove mice that are older than 20 weeks
+MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
+MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]
+
+pp <- ggplot(MLH1.imprv.male, aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Age Distributions")
+pp
+
+#what is this meta.data.file?
+#meta.data.file2 <- add_strain(meta.data.file2)
+#meta.data.file2 <- add_sex(meta.data.file2)
+#meta.data.file2 <- add_category(meta.data.file2)
+#meta.data.file2 <- add_euth_date(meta.data.file2)
+#meta.data.file2 <- add_age(meta.data.file2)
+
+#qq <- ggplot(meta.data.file2[meta.data.file2$sex == "male",], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("recorded ages of dissected mice")
+#qq
+
+```
+
+#### Assesment of Quality Scores
+
+```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
 
@@
 
-```{r, scatter plots of nMLH1 by score, echo=FALSE, fig.cap="caption"}
+```{r, scatter plots of nMLH1 by score, echo=FALSE, fig.cap="MLH1 means across quality scores"}
 #make lattice / plots seperated by strain and sex
@@
 ```{r, fig.height=2, fig.width=4, echo=FALSE}
-#this table is 
+#this table is TOO big
 #this table is saved in the .RData file
@@
 n <- kable(AP_mouse_table)
-n
+
 ```
-
-#### Power Calculations
-
-Main questions to address:
-
- > *How many cells per mouse should be sampled?*
-
- > *How many mice per strain should be sampled?*
-
-Outline discription of the loop for comparing sample size effects
-
-i) take ~100 independant samples of total images across different sample sizes (10%, 25%, 50%, 75%, 90%)
-ii) then run t.tests across those samples, and report the p value 
-iii) make a table of the p values for these sample permutations
-iv) plot the p values or mean p values across sample sizes
-
-Currently this section is under construction.
-
-```{r, include=FALSE}
-#passed_mice_data <- MLH1_data[MLH1_data$mouse %in% passed_mice, ]
-#non_passed_mice_data <- MLH1_data[MLH1_data$mouse %in% not_passed, ]
-
-#make a column to mark if observation comes from passed or not passed mouse
-#turn this into boxplot and sort the category order, otherwise it's hard to any patterns
-#sc <- qplot(mouse, nMLH1.foci,  colour = category, data = non_passed_mice_data, main="s", xlab="s", ylab="es")
-#sc + theme(legend.position="none")
-
-#sc <- qplot(mouse, nMLH1.foci,  colour = category, data = passed_mice_data, main="s", xlab="s", ylab="es")
-#sc + theme(legend.position="none")
-
-# not working not sure of the purpose
-#count =1
-#for(i in MLH1_data$file.name){
-#  templist= strsplit(i, split="_")[[1]]
-#  if(MLH1_data$mouse %in% passed_mice) 
-#  MLH1_data$mouse[count] <- c
-#  count= count +1
-#}
-```

### 2019-08-15 src/setupData_MLH1repo.R
updating the setup file

@@ -119,7 +119,7 @@ AP_mouse_table$subsp <-  as.factor(AP_mouse_table$subsp)
 #"the distribution of quality scores is ", table(MLH1_data$quality)  ) )
 
 
-save.image("data/MLH1/MLH1_data_setup_8.9.19.RData")
+save.image("data/MLH1/MLH1_data_setup_8.15.19.RData")
 #  OutPut: big large MLH1_data (AP's) df, big DF of BD with just the mice I want.
 #  MLH1_data_table, means and variance of AP and BD MLH1 values. made from seperate tables from AP and BD data.
 #  make sure decimal places are consistant

### 2019-08-16 doc/Main_Report.rmd
trying to update report
small work from home

@@
 
-ff
-ss
+#these plot currently look bad
+
 ```
@@
 ```{r age.hists, echo=FALSE}
+
 #meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
@@
 MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
+
+#add euth date for calculating age
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+PWD.KAZ_MLH1.imprv <- add_euth_date(PWD.KAZ_MLH1.imprv)
+PWD.KAZ_MLH1.imprv <- add_age(PWD.KAZ_MLH1.imprv)
 
@@
 ```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
-high_qual_boxplot
+#high_qual_boxplot
 #not sure why order got messed up again...

### 2019-08-16 src/setupMetaData.R

@@ -3,9 +3,11 @@
 # output: .csv file for reading/loading in other scripts
 
 #load Dissection file
-#meta.data.file = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)#455
+meta.data.file = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)#455
 
-meta.data = read.csv("~./MLH1repo/data/Mouse_MetaData_7.30.19.csv", header = TRUE) 
+
+#MLH1.data = read.csv("~/MLH1repo/data/MLH1/MLH1_data_setup_8.15.19.RData", header = TRUE) 
+
 #take note when I switch from 2 digit year (mouse, euth date) and 4-digit year 'DOB'!
 #remember to switch '/%y' to upper case 'Y'!!
 
@@ -78,3 +80,6 @@ write.table(meta.data, "~./MLH1repo/data/clean.Meta.Data.txt",
 
 write.table(oo.merge, "~./MLH1repo/data/Meta.Data.Merge_batch_quant.txt", 
             sep="\t", row.names = FALSE)
+
+#save Rdata file
+save.image("data/MetaData.RData")

### 2019-08-28 doc/MatMeth.Rmd
added Rmd draft for material and methods

---
title: "Material Metohds"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_depth: '3'
    toc_float: yes
---

## Mouse Husbandry

1.	(Specific strains were obtained from X vendors, 2.
2.	3 separate rooms used (precautions taken to ensure there was limited environmental variance) b. X additional husbandry things done to help with breeding of some mouse strains. housed in X conditions (extra things done for problem breeders,
3.	3. euthanized under X protocol,

## Tissue Collection and Immunohistochemistry

1.	describe how gonad tissue was collected)  (right testes, both ovaries) (tunica removed decapsulated)  X protocol for testes,   X protocol for ovaries.  X ages for males, X ages for females
2.	 2. describe spread process   (describe process - clearly point out the distinctions for testes and ovaries
3.	3. (describe staining process (with DMC1)  (same process)

## Image Processing

1.	Basic quality criteria,
2.	2. Manual quant pipeline for MLH1,,
3.	3. Quanti pipeline for DMC1 (not completely finished)

## Statisical Analysis

### 2019-08-28 doc/Main_Report.rmd
updating the setupData files

@@
 title: "MLH1 Data Report"
-author: "April Peterson"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
     df_print: paged
+    toc: yes
+    toc_depth: '5'
+    toc_float: yes
   pdf_document: default
@@
 #load(file="data/MLH1/MLH1_data_setup.RData")
+
+#load main data file
 load(file="data/MLH1/MLH1_data_setup_8.15.19.RData")
 
+load(file="data/MetaData.RData")
+
+#load functions
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 ```
 
-#### Research Description
+## Research Description
 
@@
 
+- Show basics of the current data file
+
 - Display plots and figures of distributions of MLH1 counts to assess intial overall patterns.
@@
 -  Report basic statistics of results and to help keep track of the number of images for each category and mouse
+
+*Main questions*
+
+1. Which strains do I have enough mouse specific MLH1 data for?
+2. What data will go into the FINAL dataset
 
@@
 A quality score is given; 1 to 5 (best to worst).
+
+all cell image file names are anonimized and moved into bathces of a few hundred images.
 
@@
 ```
+
+# Mice
+
+## Missing Mice
+
+This section shows / displays / catches mice which haven't been stained or quantified yet.
 
@@
 
-```{r basic.numbers, echo=FALSE}
+```{r basic.numbers, echo=FALSE, eval=FALSE}
 
@@
 The 'Other' are F1 crosses.
+
+### Histograms of Mouse Ages
+
+This section is meant to double check the age distributions of male mice.
+
+```{r age.hists, echo=FALSE, eval=FALSE}
+
+meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
+
+#make Mouse meta data work
+MouseMetaData.wrkin <- meta.data[,c(1,4)]
+colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
+
+MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
+
+#do this for the mouse table instead of the MLH1 table
+MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
+
+#add euth date for calculating age
+
+MLH1.imprv <- add_euth_date(MLH1.imprv)
+#
+MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
+MLH1.imprv$euth.date <- as.Date(as.character(MLH1.imprv$euth.date2), format='%d%b%y')
+#this is currently broken
+MLH1.imprv <- add_age(MLH1.imprv)
+
+#remove mice that are older than 20 weeks
+MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
+MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]
+
+pp <- ggplot(MLH1.imprv.male, aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Age Distributions")
+pp
+
+#what is this meta.data.file?
+#meta.data.file2 <- add_strain(meta.data.file2)
+#meta.data.file2 <- add_sex(meta.data.file2)
+#meta.data.file2 <- add_category(meta.data.file2)
+#meta.data.file2 <- add_euth_date(meta.data.file2)
+#meta.data.file2 <- add_age(meta.data.file2)
+
+#qq <- ggplot(meta.data.file2[meta.data.file2$sex == "male",], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("recorded ages of dissected mice")
+#qq
+
+```
+
+#### Is there an age effect?
+
+make a subetted dataset for ploting mean mouse (quality < 5) MLH1 count and age.  Mice with enough ages:
+WSB, KAZ, LEW, G, Cast, SPIC, MSM, MOLF, CZECH
+
+```{r}
+#run a age effect linear regression!
+
+male.data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.y == "male",]
+
+lm.me  <- lm(male.data$mean_co ~ male.data$age.weeks+male.data$strain.x)
+
+summary(lm.me)
+
+```
+
+```{r age.effect.plot}
+
+#make sure the right dataset 
+male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
+
+male.data <- male.data[male.data$Ncells > 10,]
+
+male.data <- add_subsp(male.data)
+
+age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(15,39))+
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age Effect on CO mean")+facet_wrap(~subsp.y)
+
+#todo add line to strain.y
+
+```
+
+Given this plot it seems that there isn't a age effect. I should double check some of these mice
 
@@
 
-### Histograms of Mouse Ages
-
-```{r age.hists, echo=FALSE}
-
-#meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
-meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)
-meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375
-
-meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
-
-#make Mouse meta data work
-MouseMetaData.wrkin <- meta.data.file2[,c(1,4)]
-colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
-MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
-
-#do this for the mouse table instead of the MLH1 table
-MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
-
-#add euth date for calculating age
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-PWD.KAZ_MLH1.imprv <- add_euth_date(PWD.KAZ_MLH1.imprv)
-PWD.KAZ_MLH1.imprv <- add_age(PWD.KAZ_MLH1.imprv)
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-MLH1.imprv <- add_euth_date(MLH1.imprv)
-#
-MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
-MLH1.imprv$euth.date <- as.Date(as.character(MLH1.imprv$euth.date2), format='%d%b%y')
-
-MLH1.imprv <- add_age(MLH1.imprv)
-
-#remove mice that are older than 20 weeks
-MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
-MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]
-
-pp <- ggplot(MLH1.imprv.male, aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Age Distributions")
-pp
-
-#what is this meta.data.file?
-#meta.data.file2 <- add_strain(meta.data.file2)
-#meta.data.file2 <- add_sex(meta.data.file2)
-#meta.data.file2 <- add_category(meta.data.file2)
-#meta.data.file2 <- add_euth_date(meta.data.file2)
-#meta.data.file2 <- add_age(meta.data.file2)
-
-#qq <- ggplot(meta.data.file2[meta.data.file2$sex == "male",], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("recorded ages of dissected mice")
-#qq
-
-```
+` r knitr_exit()`
 
@@
 This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.
-
-#### Mice table
-
-Table of the mice used and MLH1 stats.
-
-```{r, fig.height=2, fig.width=4, echo=FALSE}
-#this table is TOO big
-#this table is saved in the .RData file
-#use kable to make tables. warning this might not work for github Rmd.
-n <- kable(AP_mouse_table)
-
-```

### 2019-08-28 doc/MatMeth.Rmd
updating the setupData files

@@
     toc: yes
-    toc_depth: '3'
+    toc_depth: '5'
     toc_float: yes
@@
 
+X mice used
+
 ## Tissue Collection and Immunohistochemistry
@@
 1.	Basic quality criteria,
-2.	2. Manual quant pipeline for MLH1,,
-3.	3. Quanti pipeline for DMC1 (not completely finished)
+2.	Manual quant pipeline for MLH1 (anonimized)
+3.	Quanti pipeline for DMC1 (not completely finished)
+
+For MLH1 pachytene characterization, Cells with a full karyotype (19 autosomes and XY for spermatocyte spreads or 20 bivalents for oocyte spreads), distinct foci, and intact bivalents were included for quantification.
 
 ## Statisical Analysis
+
+-done in R/ specific packages
+-non-parameteric  (tests and measures for dealing with count data)

### 2019-08-28 src/CommonFunc_MLH1repo.R

@@ -189,17 +189,17 @@ add_subsp <- function(oldframe) {
                      ifelse(grepl("HMI", dframe$strain), "Cast",
                                                                           
                            ifelse(grepl("MSM", dframe$strain), "Musc", 
-                        ifelse(grepl("MOLF", dframe$strain), "Musc",     
-                         ifelse(grepl("PWD", dframe$strain), "Musc", 
-        ifelse(grepl("CZECH", dframe$strain), "Musc", 
+                        ifelse(grepl("MOLF", dframe$strain), "Musc",    
+                
+                  ifelse(grepl("PWD", dframe$strain), "Musc", 
+      
+                  ifelse(grepl("CZECH", dframe$strain), "Musc", 
                ifelse(grepl("AST", dframe$strain), "Musc",     
                       ifelse(grepl("TOM", dframe$strain), "Musc",  
                   ifelse(grepl("SKIVE", dframe$strain), "Musc", 
                              
-                ifelse(grepl("PWDFemale", dframe$strain), "Musc",  
+                ifelse(grepl("PWDFemale", dframe$strain), "Musc",
                       ifelse(grepl("KAZ", dframe$strain), "Musc",
-                             
-                             
                              
                                                                                                              
                      ifelse(grepl("SPRET", dframe$strain), "Spretus",

### 2019-08-28 src/setupData_MLH1repo.R

@@ -39,7 +39,7 @@ original_DF <- add_mouse(original_DF)
 #original_length <- length(MLH1_data$Original.Name)
 
 #check that all batches are included
-table(original_DF$Batch)
+#table(original_DF$Batch) #if I want this I should make this Rmd
 
 
 MLH1_data$Original.Name <- as.character(MLH1_data$Original.Name)
@@ -93,7 +93,6 @@ AP_mouse_table <- ddply(MLH1_data, c("mouse"), summarise,
                         #quality?
 )
 
-source("src/CommonFunc_MLH1repo.R")
 AP_mouse_table <- add_strain(AP_mouse_table)
 AP_mouse_table <- add_subsp(AP_mouse_table)
 AP_mouse_table <- add_sex(AP_mouse_table)
@@ -106,7 +105,38 @@ AP_mouse_table$strain <- as.factor(AP_mouse_table$strain)
 AP_mouse_table$subsp <-  as.factor(AP_mouse_table$subsp)
 
 
+#metadata contains all mice (different source) with ages
 
+
+#integrate ages ect
+#meta.data -> 
+#AP_mouse_table and MLH1_data
+
+org.DF.mice <- original_DF[(original_DF$mouse %in% meta.data$mouse),]#this subsets mice -- not integrateing age
+#this produces the whole dataset -- might just need unique mouse and batch combos
+
+oo.merge <- merge(oo.DF, meta.data, by.y = "mouse", by.x = "Var1", all.y = TRUE)
+oo.DF <- as.data.frame(oo)#this has a dif str, 3 cols -- like it was melted
+oo.DF <- oo.DF[!(is.na(oo.DF$Freq)|oo.DF$Freq==0),] #remove all 0 counts for batches
+
+#duplicated(oo.DF$Var1)
+mulitple.batch.mice <- oo.DF$Var1[duplicated(oo.DF$Var1)]
+xu2 <- oo.DF[!unique(oo.DF$Var1),]
+#push colname of matching mouse into cell of metadata
+table(oo.DF$Var1)# this is how you can fin the double counts
+
+#maybe change colnames of oo.DF
+
+#can I collapse the mulitple batches mice?
+
+#to match -- I may need to merge
+oo.merge <- merge(oo.DF, meta.data, by.y = "mouse", by.x = "Var1", all.y = TRUE)
+mouse_table_w.Ages <- oo.merge
+
[…]

### 2019-08-28 src/setupMetaData.R

@@ -2,8 +2,20 @@
 # input: newest / most recent Dissection csv file
 # output: .csv file for reading/loading in other scripts
 
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+
 #load Dissection file
 meta.data.file = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)#455
+
+#loading previously cleaned data 
+clean.meta.data <- read.csv("~./MLH1repo/data/clean.Meta.Data.txt",header = TRUE, sep='\t')
+
+
+#integrate batch information
+
+#read or load MLH1 data
+#load most recent RData file
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_8.9.19.RData")#8.9.19 has batch14
 
 
 #MLH1.data = read.csv("~/MLH1repo/data/MLH1/MLH1_data_setup_8.15.19.RData", header = TRUE) 
@@ -26,11 +38,6 @@ meta.data <- add_sex(meta.data)
 meta.data <- add_category(meta.data)
 
 
-#integrate batch information
-
-#read or load MLH1 data
-#load most recent RData file
-load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_8.9.19.RData")#8.9.19 has batch14
 
 
 #subset org_DF for mice in meta_data list 
@@ -65,6 +72,8 @@ table(oo.DF$Var1)# this is how you can fin the double counts
 
 #to match -- I may need to merge
 oo.merge <- merge(oo.DF, meta.data, by.y = "mouse", by.x = "Var1", all.y = TRUE)
+mouse_table_w.Ages <- oo.merge
+
 
 #batch14 is missing .. it hasn't been integrated into the set up data i think
 
@@ -75,10 +84,10 @@ oo.merge <- merge(oo.DF, meta.data, by.y = "mouse", by.x = "Var1", all.y = TRUE)
 
 
 #write csv file (/data)
-write.table(meta.data, "~./MLH1repo/data/clean.Meta.Data.txt", 
+write.table(meta.data, "~./MLH1repo/data/clean.Meta.Data_8.28.19.txt", 
             sep="\t", row.names = FALSE)
 
-write.table(oo.merge, "~./MLH1repo/data/Meta.Data.Merge_batch_quant.txt", 
+write.table(oo.merge, "~./MLH1repo/data/Meta.Data.Merge_batch_quant_8.28.19.txt", 
             sep="\t", row.names = FALSE)
 
 #save Rdata file

### 2019-08-30 doc/MouseLevelPlots.Rmd
made rmd for mouse level quality checking

---
title: "mouse level quality plots"
author: "April Peterson"
date: "August 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)

setwd("~./MLH1repo/")
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")

#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")#batch15 isn't showing up
#load(file="data/MetaData.RData")#this Rdata was replacing the most current batch.. agg!

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

knitr::opts_chunk$set(echo = TRUE)
```

This rmd file shows the cell counts for mouse level for MLH1 count and quality. The black lines are the strain means and the red lines are the mouse level means. Points are colored by the quantification batch.

These plots are meant to asses which mice can be removed from the final dataset for too few or low quality count data.

### Updating plots

*1. check which category levels do not have data, and enter that index into 'df_list'. Otherwise the loop will be broken.*

*2. Run the code chunks.*

These MLH1 data still have quality 5 cells observations. I might want to redo it without those observations.

```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE}
### to integrate 'pass'/fail
### this should be in setup data

#mice that don't have female data will break the chain-loop
#MOLF female, HMI, CZECH female, CAROLI female, Tom male

#instead of adding fake data, remove the categories from list

#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing

pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")

#this loop makes a quant_status col for a list of mice
for(i in 1:length(MLH1_data$mouse)){
  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
}

df_list <- split(MLH1_data, (MLH1_data$category))
#remove sections which are in category but have no data
#this is where you remove the non-data data!
df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]

#7 PERC, 12 MOLF female, 
#

#df list, is a list of df's which were split() from BIG df by category
df_list 
o=0
i=0

MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)

#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
for(i in 1:(length(df_list))){
 # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
                      mouse_nMLH1 = mean(adj_nMLH1.foci)
  )
# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
  for(o in 1:length(df_list[[i]]$Original.Name)){
    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
        }
}

plot_list = list()
for (h in 1:length(df_list)) {
  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
    facet_wrap(~ mouse, drop = TRUE) + 
    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
  #make an empty ggplot? for missing data

  plot_list[[h]] =   pf
}

#orig (I think this is the same as above)
#plot_list = list()
#for (h in 1:length(df_list)) {
#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)

#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +

 #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +

    #theme for indiviual mice, 
#theme(plot.background = element_rect(fill = "green"))
#  plot_list[[h]] =   pf
#}
#can't change the alpha/transparency of the red rec

#aes must be same length ---   #geom_point(stat = "identity") +
# the last one is correct....?

#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
```

```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
geom_hline(aes(yintercept = med, group = gr), colour = 'red')
#no plot is made that goes in the list -- so it can't be shown
for (gg in 1:length(plot_list) ){
  show(plot_list[gg])
 # invisible(lapply(obj, function(x) plot(x,main="some plot")))
}
```

### 2019-08-30 doc/MatMeth.Rmd
work on MM, added bibliography file

@@
     toc_float: yes
+bibliography: refs_MatMeth.bib
 ---
@@
 housed in X conditions (extra things done for problem breeders,
-3.	3. euthanized under X protocol,
+3.	euthanized under X protocol,
+4. mention results if there was a detectable environmental effect
 
-X mice used
+The Wild derived inbred strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ, CORLI/EiJ and SPRET/EiJ were purchased from Jax labs (Maine USA). 
+The strains of KAZ/TUA, TOM/TUA, AST/TUA, HMI/TUA and SPIC) were ordered and cryo-derived from BRC at RIKEN (Japan) website).
+All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.
+ A breeding colony of wild derived mice sampled from Gough Island were housed at UW Veterinary school facilities.
+(some of the other strains housed there - comparisons were made of strains housed across different facilities to (test if there was an effect of mouse room of MLH1 counts).
+
+Mice were feed on dry standard breeder chow.
+Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
+Adult mice were euthanized by CO asphyxiation.
+Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison.
 
@@
 1.	describe how gonad tissue was collected)  (right testes, both ovaries) (tunica removed decapsulated)  X protocol for testes,   X protocol for ovaries.  X ages for males, X ages for females
-2.	 2. describe spread process   (describe process - clearly point out the distinctions for testes and ovaries
-3.	3. (describe staining process (with DMC1)  (same process)
+2.	 describe spread process   (describe process - clearly point out the distinctions for testes and ovaries
+3.	(describe staining process (with DMC1)  (same process)
+
+From males, the right teste was collected and processed. (Ages are listed in supplmental material). Ovaries were collected from neonates less than 48 hours old or embryos between e16 to 21.
+
+The behavior of these strains did not allow us to check for copulary plugs, resulting in most oocyte data coming from neonate mice between 5 to 48 hours.
+
+Meiocyte spreads were made following  [@peters_1997] Peters et al 2001.
+For testes the tunica was removed and whole testes was incubated in 3ml of hypotonic solution for 45min.
+For ovaries after dissection, the pair of ovaries were decapsulated in cold PBS and both were incubated in 300ul of hypotonic solution for 45 min.
+
+After incubation, gonads were transferred to sucrose solution -for mastication to make a cell slurry which was transferred to 80ul of 2% PFA solution.
+Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
+The next morning, slides were treated with a photoflow wash.
 
@@
 
+Reprocessing in Photoshop ()- Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
+
 ## Statisical Analysis
@@
 -non-parameteric  (tests and measures for dealing with count data)
+
+-hypothese tested (size, DSB number)
+
+# References

### 2019-08-30 doc/Intro.Rmd
made Intro md file and other writing updates

---
title: "Intro"
author: "April Peterson"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: yes
      toc_depth: '5'
      toc_float: yes
bibliography: refs_Intro.bib
---

# Outline

## Background:

### prev hetC lit:
(Lenormand, BurtBellOther, Shapley,)
HetC theories:   (Brandavain, Trivers ??? Lenormand, Sardell-Kirkpatrick, (hulten? Handel?)

wild pops genetic diversity and within species measures

my pero refs, honey bee RR, ritz noor

## gap, angle

# text

  Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products. The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes. Under one of the most common forms of sexual reproduction, anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates. Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.

# Notes and Ideas

# Bibliography

### 2019-08-30 doc/Main_Report.rmd
updating list of mice to finish

@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_8.15.19.RData")
+load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")
 
-load(file="data/MetaData.RData")
+#load(file="data/MetaData.RData") #this was replacing most updated data file
 
@@
 
-## Research Description
+# Research Description
 
@@
 
-#### Description of Report
+## Description of Report
 
@@
 
-#### Discription of quantification process
+## Discription of quantification process
 
@@
 all cell image file names are anonimized and moved into bathces of a few hundred images.
+
+Definition of passed quantified ect.
 
@@
 
-# Mice
+*What were guidlines for quantification of lower quality cells / faint foci?*
 
-## Missing Mice
+-Count foci that look like other foci in the image.
+
+## Missing Mice, finishing dataset
+
+*Which mice are remaining and at what stage?*
 
@@
 
+passed quantified - quantified mice (with enough cells)
+total quantified -- all mice which have been (includes fail and passing mice)
+
+non-quantified mice -- dissected mice / slides that haven't been quantified yet (all stained, but without image directory)
+
+non-imaged -- mice without images, (might be stained and unstained)
+
 This is the number of mice that are through different parts of the pathway or quantification process.
+
+should I break down the categories into sex and strain?
 
@@
 
-### Histograms of Mouse Ages
+# Tests of dataset
+
+Main metric is mean CO count per mouse. Assess the quality of this data with the following tests
+
+*age effects* (male, maternal age)
+
+*litter effect*
+
+*mouse room effect* (Charmony, MSC)
+
+*quantification repeatability*
+
+*spread batch and staining effects*
+
+*power analysis on number of cells*
+
+*time of year effect (on spread)*
+
+These aspects will be analyzed with a linear models on subsets of data (within category)
+
+## Repeatability of Human quantification
+
+*What is the Error in quantification across humans?*
+
+A general ancedote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (useing some of Beth's data)
+
+There is a subset of data which has been counted mulitple times -- use that to test the repeatbility.
+
+```{r quant.repeatability, echo=FALSE}
+
+```
+
+## Age Effect
 
@@
 
-#### Is there an age effect?
+ Is there an age effect?
 
@@
 
-```{r}
+```{r age.effect.model}
 #run a age effect linear regression!
@@
 male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
+#remove NAs
 
@@
 
-male.data <- add_subsp(male.data)
-
 age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(15,39))+
   geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age Effect on CO mean")+facet_wrap(~subsp.y)
-
+age.effect.plot
 #todo add line to strain.y
@@
 
-#### Initial Patterns from MLH1 distributions
+#### Materinal Age effect
+
+```{r mat.age.effect, echo=FALSE}
+#compile data from female (and male?) mice with maternal ages.
+
+#start making a general linear model...? for mouse averages
+
+```
+
+### litter effect?
+
+## Spread and Stain Effects
+
+*month dissected*
+
+*staining batches*
+
+## Room effect (envioronment)
+
+3 different rooms used to house mice. This will be tested in X manner.
+
+## cell number (power analysis results)
+
+# Initial Results
 
 ```{r first.boxplots, echo=FALSE, fig.align='center'}
-#what?
-#MLH1_data <- MLH1_data[MLH1_data$strain]
 
@@
 
+#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
 #can't re-order the miceagg
-ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
-             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
+#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
+#        "grey","grey","grey",  "grey","grey","grey" )) +
+# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
+#don't show M.musculus by mouse
+
+#order MLH1_data by strain!!
+
+ff2 <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
+  geom_boxplot(aes(x = as.character(mouse), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey", 
-        "grey","grey","grey",  "grey","grey","grey" )) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(~ sex, scales="free")
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
+        "green","green4","grey",  "grey","grey","grey" )) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
+  facet_grid(subsp~ sex, scales="free_x")#scales="free_y"
+#this is ok except for the set placement of mice on x axis
+#if the full dataset was ordered correctly, this would kinda fix it
 
 #boxplots by strain
-ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
+ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey" )) +
+        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey","green", "purple","grey1","grey","white", "blue","orange" )) +
   ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )
@@
 #these plot currently look bad
+ss
+```
+
+*What is the initial picture of Heterochiasmy across mouse strains?*
+
+For the strain level plots
+  -facet,sex by species
+
+### Distributions of MLH1 counts
+
+*Are the MLH1 distributions close to normal?*
+
+```{r, MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
+#consider making distribution with quality considered...
+
+gg <- data.frame(weight=rep(x,nrow(df)),df)
+gg$y <- with(gg,dnorm(x,mean,sd))
+gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * diff(range(weight$weight))/30
+
+mean.MLH1.obs <- ddply(.data=MLH1_data, 
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")+
+  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
+
+  geom_smooth(method = lm, se = FALSE)
+
+  geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+
+histM
+
+#if i'm testing normalge
+
+#function = glm(y ~ x, family = poisson)
+#add Poission curve behind
+#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
+
+#not sure how to add Expected poisson curve
 
@@
 
-### Distributions of MLH1 counts
+(most all distributions are close to normal)
 
-```{r, histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
-#consider making distribution with quality considered...
+How do these compare to Poission? (mean == variance)
+I should make 'expected' distributions behind these.
 
-histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
-histM <- histM + geom_histogram(stat = "bin", binwidth = 1)
-histM <- histM + facet_wrap(~ category)
-histM
-```
+### Assesment of Quality Scores
 
-` r knitr_exit()`
+> *How do the distributions change across different quality scores?*
 
-#### Assesment of Quality Scores
+> *How do the distributions change across mice with a good number of cells?*
 
-```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
+```{r high.qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
 
@@
 
-```{r, passing mice, echo=FALSE}
+```{r, passing.mice, echo=FALSE}
 q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
@@
 
-#### Comparison of MLH1 distributions
-
-```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
-#high_qual_boxplot
-#not sure why order got messed up again...
-```
-
-After taking the data from 2 highest cell quality, some but not all mouse means converge.
-For some mice (mostly female) the majority of cells are excluded.
-This is especially true for PWD and WSB females.
-
-#### Tests for quality and nMLH1 foci number
-
-> *How do the distributions change across different quality scores?*
-
-> *How do the distributions change across mice with a good number of cells?*
-
-##### Effects of quality
+*Answer to questions*
 
@@
 
-```{r, scatter plots of nMLH1 by score, echo=FALSE, fig.cap="MLH1 means across quality scores"}
+```{r, scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means across quality scores"}
 #make lattice / plots seperated by strain and sex
@@
 
-#### Assessing the Distributions by mouse
+####Distributions by mouse
+
+*Is there mouse level MLH1 quality score bias?*
 
@@
 
-```{r, echo=FALSE, message=FALSE, results='hide', warning=TRUE}
+below are code chuncks that make a big dfs of CO.count by quality score plot. The code has been commented out, and moved to it's own doc.
+
+```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
 ### to integrate 'pass'/fail
@@
 
-#quant_status (this isn't saving)
-#for(i in 1:length(MLH1_data$mouse)){
-#  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-#}
+#mice that don't have female data will break the chain-loop
+#MOLF female, HMI, CZECH female, CAROLI female, Tom male
 
-#what is the purpose of this block
-#df_list <- split(MLH1_data, (MLH1_data$category))
+#instead of adding fake data, remove the categories from list
+
+#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
+#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
+#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
+#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
+#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
+#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
+#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
+#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
+#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
+
+pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
+
+#this loop makes a quant_status col for a list of mice
+for(i in 1:length(MLH1_data$mouse)){
+  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+}
+
+df_list <- split(MLH1_data, (MLH1_data$category))
 #remove sections which are in category but have no data
-#df_list <- df_list[ -c(7,16,22) ]
+#this is where you remove the non-data data!
+df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
 
-#7: PERC male, (14:CAST female, fake data) 16:HMI female, 22 other
+#7 PERC, 12 MOLF female, 
 #
@@
 #df list, is a list of df's which were split() from BIG df by category
-#df_list 
-#o=0
-#i=0
+df_list 
+o=0
+i=0
 
-#MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
+MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
 
 #make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-#for(i in 1:(length(df_list))){
+for(i in 1:(length(df_list))){
  # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-#  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-#  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
- #                      mouse_nMLH1 = mean(adj_nMLH1.foci)
-#  )
+  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
+  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
+                      mouse_nMLH1 = mean(adj_nMLH1.foci)
+  )
 # this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-#  for(o in 1:length(df_list[[i]]$Original.Name)){
-#    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == #sub_mouse_means$mouse) ]
-#        }
-#}
+  for(o in 1:length(df_list[[i]]$Original.Name)){
+    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
+        }
+}
 
-#plot_list = list()
-#for (h in 1:length(df_list)) {
-#  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ #geom_jitter(aes(y= adj_nMLH1.foci, x= quality))
-  # +   geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5)
-#  + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL)
+plot_list = list()
+for (h in 1:length(df_list)) {
+  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
+    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
+    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
+    facet_wrap(~ mouse, drop = TRUE) + 
+    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
+  #make an empty ggplot? for missing data
 
-# +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  #theme for indiviual mice, 
-  #+ theme(plot.background = element_rect(fill = "green"))
-#  
-#  plot_list[[h]] =   pf
-#}
+  plot_list[[h]] =   pf
+}
 
-#orig
+#orig (I think this is the same as above)
 #plot_list = list()
@@
 #  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, #ymin=-100, ymax=100, alpha=0.005)
+#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
 
@@
 #  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") #+
+#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
 
-#    scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL)
+ #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
 
     #theme for indiviual mice, 
-  #+ theme(plot.background = element_rect(fill = "green"))
+#theme(plot.background = element_rect(fill = "green"))
 #  plot_list[[h]] =   pf
@@
 
-```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE }
-#geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-#
-#for (gg in 1:length(plot_list) ){
-#  show(plot_list[gg])
-  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
-#}
+```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
+geom_hline(aes(yintercept = med, group = gr), colour = 'red')
+#no plot is made that goes in the list -- so it can't be shown
+for (gg in 1:length(plot_list) ){
+  show(plot_list[gg])
+ # invisible(lapply(obj, function(x) plot(x,main="some plot")))
+}
+
+#this stops at MSM
 ```
 
+The mouse specific scatter plots aren't show because there are too many. (but they should be used for investigating specific mice).
 Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
@@
 This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.
+
+# Research Questions
+
+*-which strain has highest HetC?*
+
+Start thinking of the direction for analysis
+
+## within mouse
+
+### female and male comparison
+
+## within category (strain)
+
+```{r HetC.plot, echo=FALSE,  fig.width=10, fig.height=5}
+no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
+
+MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
+
+MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
+
+MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
+
+MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
+
+kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
+
+bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
+
+#add titles
+#use multiplot
+#remove X axis
+
+multiplot(kuku,bubu, cols=2)
+```
+
+### female and male comparison
+
+## Subspecies level comparison
+
+Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
+
+*Does this hold up with the full data set?*
+*What's the best visualization for these questions?*
+
+Main focus is contrasting the level of difference within and between subspecies.
+
+I'll do this with a mixed model framework approach (and analyze with ANOVA, a method to partition the variance of subset.)
+
+```{r mixed.model, echo=FALSE}
+
+#build mixed model
+
+#run simulations for accounting for sampling error
+
+```
+
+### Subspecies examples / Data
+
+list the strains/data used 
+Mouse average MLH1 counts,
+
+### Male Specific results
+
+Not all strains have female data. Here are results from all strains wih male data.
+
+### female and male comparison
+
+## Meiotic Error
+
+This section reviews all the data from achiasmy and 0CO bivalents.

### 2019-09-05 doc/Main_Report.rmd
updates to main report

@@
 
-Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice abd related rodents.
+Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice and related rodents.
 Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
@@
 
-- Display plots and figures of distributions of MLH1 counts to assess intial overall patterns.
+- Display plots and figures of distributions of MLH1 counts to assess initial overall patterns.
 
--  Report basic statistics of results and to help keep track of the number of images for each category and mouse
-
-*Main questions*
-
-1. Which strains do I have enough mouse specific MLH1 data for?
-2. What data will go into the FINAL dataset
+-  Report basic statistics of results and to help keep track of the number of images for each category and mouse.
 
@@
 
-all cell image file names are anonimized and moved into bathces of a few hundred images.
-
-Definition of passed quantified ect.
+all cell image file names are anonymized and moved into batches of a few hundred images.
 
@@
 
-*What were guidlines for quantification of lower quality cells / faint foci?*
-
--Count foci that look like other foci in the image.
-
 ## Missing Mice, finishing dataset
@@
 
-This section shows / displays / catches mice which haven't been stained or quantified yet.
+This section shows the number of mice which haven't been stained or quantified yet.
 
@@
 
-# Tests of dataset
+## Repeatability of Human quantification
 
-Main metric is mean CO count per mouse. Assess the quality of this data with the following tests
+*What is the Error in quantification across humans?*
+
+A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%.
+(using some of Beth's data).
+There is a subset of data which has been counted multiple times -- use that to test the repeatbility.
+
+```{r quant.repeatability, echo=FALSE}
+
+```
+
+## Effects on the dataset
+
+Main metric is mean CO count per mouse. Run tests to check if there are effects for the following effects.
 
@@
 
-These aspects will be analyzed with a linear models on subsets of data (within category)
+These aspects will be analyzed with a linear models on subsets of data (within category) and visualized with plots.
 
-## Repeatability of Human quantification
+### Age Effect
 
-*What is the Error in quantification across humans?*
+Some papers have documented age effects for male genome-wide recombination rate (cite).
 
-A general ancedote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (useing some of Beth's data)
+The initial analysis uses plots. This should be double checked with a general linear model as well.
 
-There is a subset of data which has been counted mulitple times -- use that to test the repeatbility.
-
-```{r quant.repeatability, echo=FALSE}
-
-```
-
-## Age Effect
-
-This section is meant to double check the age distributions of male mice.
-
-```{r age.hists, echo=FALSE, eval=FALSE}
-
+```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
 meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
@@
 
- Is there an age effect?
+Make a subsetted dataset for plotting mean mouse (quality < 5) MLH1 count and age.
 
-make a subetted dataset for ploting mean mouse (quality < 5) MLH1 count and age.  Mice with enough ages:
-WSB, KAZ, LEW, G, Cast, SPIC, MSM, MOLF, CZECH
-
-```{r age.effect.model}
+```{r age.effect.model, include=FALSE}
 #run a age effect linear regression!
@@
 
-summary(lm.me)
-
+#summary(lm.me)
 ```
 
-```{r age.effect.plot}
+```{r age.effect.plot, warning=FALSE, echo=FALSE}
 
@@
 
-age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(15,39))+
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age Effect on CO mean")+facet_wrap(~subsp.y)
+age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(19,34))+
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age and mean CO per mouse")+facet_wrap(~subsp.y)
 age.effect.plot
@@
 
-Given this plot it seems that there isn't a age effect. I should double check some of these mice
+While many slopes are different than 0, I am not convinced there is a clear age effect.
+Most strains don't have more than 2 ages.
+I will double check some of these mice and check a general linear model.
 
-#### Materinal Age effect
+### Maternal Age effect
+
+Check if there is a maternal age effect for females and males.
 
@@
 
-### litter effect?
+### Litter effect
 
-## Spread and Stain Effects
+Compare female spreads made on the same day, but from different mothers/litters.
 
-*month dissected*
+### Spread and Stain Effects
 
-*staining batches*
+> *month dissected*
 
-## Room effect (envioronment)
+For mice strains that have a similar MLH1 level (control for strain MLH1 effect), check if a month effect for MLH1 count (as seen in Peromyscus).
+
+> *staining batches*
+
+Compare observations from the same mouse, but different MLH1 stains.
+
+### Room effect (envioronment)
 
@@
 
-## cell number (power analysis results)
+### Cell number (power analysis results)
+
+Insert the code and results for effects of cell number/observations on MLH1 means.
 
@@
 
+```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
+#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
+
+#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
+#remove quality 5 cells
+
+#outgroups
+#SPIC
+HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
+#unique(HetC.SPIC$mouse)
+#non-passing mice
+SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
+HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
+#calq sex mean
+SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
+
+#SPRET
+HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
+#unique(HetC.SPRET$mouse)
+#non-passing mice
+#SPRET.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
+
+#CAROLI
+HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
+#unique(HetC.CAROLI$mouse)
+#non-passing mice
+#SPRET.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+#cast
+HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
+#unique(HetC.CAST$mouse)
+#non-passing mice
+#X.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
+
+#HMI
+HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
+#unique(HetC.HMI$mouse)
+#non-passing mice
+#X.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+#dom
+
+#WSB
+HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
+#unique(HetC.WSB$mouse)
+#non-passing mice
+WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
+HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
+#calq sex mean
+WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
+
+#LEW
+HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
+unique(HetC.LEW$mouse)
+#non-passing mice
+LEW.remove.mice <- c("4jan17_LEW_f2")
+HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
+
+#calq sex mean
+LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
+
+#G
+HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
+#unique(HetC.G$mouse)
+#non-passing mice
+G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
+HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
+
+#calq sex mean
+G.mean.MLH1.obs <- ddply(.data=HetC.G, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
+
+#PERC
+HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
+#unique(HetC.PERC$mouse)
+#non-passing mice
+#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
+#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
+
+#calq sex mean
+PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
+
+#MUSC
+
+#PWD
+HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
+#unique(HetC.PWD$mouse)
+#non-passing mice
+PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
+
+# CZECH
+HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
+#unique(HetC.CZECH$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
+
+# SKIVE
+HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
+#unique(HetC.SKIVE$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
+
+# KAZ
+HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
+#unique(HetC.KAZ$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
+
+#TOM
+HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
+#unique(HetC.TOM$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
+
+#AST
+HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
+#unique(HetC.AST$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
+
+#Molosinus
+
+#MSM
+HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
+unique(HetC.MSM$mouse)
+
+#non-passing mice
+MSM.remove.mice <- c("12sep16_MSM_f3")
+HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
+
+#calq sex mean
+MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
+
+#MOLF
+HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
+unique(HetC.MOLF$mouse)
+
+#non-passing mice
+MOLF.remove.mice <- c("27oct18_MOLF_m3")
+HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
+
+#calq sex mean
+MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
+
+```
+
+> *-Which strain has highest HetC?*
+
+```{r HetC.plot, echo=FALSE, warning=FALSE}
+#HETC plot
+
+#make scatter plot for HetC
+#color by strain
+
+#set up data
+#make female male plot
+
+HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#
+#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
+
+HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
+
+HetC_strain_table.M$sex <- "male"
+HetC_strain_table.F$sex <- "female"
+
+HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
+
+HetC_strain_table <- add_subsp(HetC_strain_table)
+
+#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
+
+HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
+HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
+#str(HetC_strain_table)
+
+Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
+  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey"))
+
+#musc - coral, Dom - cadetblue (grey blue),
+
+Het.C.plot
+#this plot is missing several
+
+```
+
+3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
+1 of the musc strains has female biased heterochiasmy, KAZ.
+
 ### Distributions of MLH1 counts
 
-*Are the MLH1 distributions close to normal?*
+> *Are the MLH1 distributions close to normal?*
 
-```{r, MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
-#consider making distribution with quality considered...
+```{r MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}
 
-gg <- data.frame(weight=rep(x,nrow(df)),df)
-gg$y <- with(gg,dnorm(x,mean,sd))
-gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * diff(range(weight$weight))/30
+#gg <- data.frame(weight=rep(x,nrow(df)),df)
+#gg$y <- with(gg,dnorm(x,mean,sd))
+#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
 
@@
 
-  geom_smooth(method = lm, se = FALSE)
-
-  geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
 
 histM
-
-#if i'm testing normalge
 
@@
 #geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
-
 #not sure how to add Expected poisson curve
@@
 
-How do these compare to Poission? (mean == variance)
+How do these compare to Poisson? (mean == variance)
 I should make 'expected' distributions behind these.
@@
 
-```{r, passing.mice, echo=FALSE}
+```{r passing.mice, echo=FALSE, warning=FALSE}
 q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
@@
 
-*Answer to questions*
-
 Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
@@
 
-```{r, scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means across quality scores"}
+```{r scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means and quality scores", warning=FALSE}
+
 #make lattice / plots seperated by strain and sex
@@
 
-I ploted the mean of each quality bin with a red dot.
-From the pattern of the red dots, there is definately a negative relationship with quality and nMLH1 foci across the data.
+I plotted the mean of each quality bin with a red dot.
+From the pattern of the red dots, there is definitely a negative relationship with quality and nMLH1 foci across the data.
 This is most pronounced in MSM males and least pronounced in G males.
@@
 
-*Is there mouse level MLH1 quality score bias?*
+> *Is there mouse level MLH1 quality score bias?*
 
-latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.
-
-below are code chuncks that make a big dfs of CO.count by quality score plot. The code has been commented out, and moved to it's own doc.
+Below are code chunks that make a big dfs of CO count by quality score plot. The code has been commented out, and moved to it's own doc.
 
@@
 
-The mouse specific scatter plots aren't show because there are too many. (but they should be used for investigating specific mice).
+The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
+
 Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
@@
 
-PWD females and a very large range. This is driven by data from 17apr15_PWD_f4. These data should be looked at more closely.
-
-In some mice there seems to be bimodel cell populations, those of clearly high quality and those of lower quality.
-This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.
-
 # Research Questions
 
-*-which strain has highest HetC?*
+## Models for manuscript
 
-Start thinking of the direction for analysis
+Rough ideas for models of evolution of Heterochiasmy and current idea for how to test them.
 
-## within mouse
+- HetC arises due to more precursors (DSBs)
 
-### female and male comparison
+    - Compare DSB counts
 
-## within category (strain)
+- Larger cell volume drives longer chrms axis (and maybe more DSB)
 
-```{r HetC.plot, echo=FALSE,  fig.width=10, fig.height=5}
+    - Measure meiocyte diameter in males and females. Compare spermatocyte diameters in male biased and female biased strains. 
+    - Compare DSBs counts and average SC/axis lengths.
+
+- Difference in astral/MT organization
+
+    - ?
+
+- Something about interference and sister cohesin area
+
+    - Develope metric for sister cohesin area, and look at initial patterns.
+
+## Mixed Model, compare within and between subspecies patterns
+
+Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
+
+*Does this hold up with the full data set?*
+*What's the best visualization for these questions?*
+
+Main focus is contrasting the level of difference within and between subspecies.
+
+I'll do this with a mixed model framework approach (and analyze with ANOVA, a method to partition the variance of subset.)
+
+```{r mixed.model, echo=FALSE}
+#build mixed model
+#run simulations for accounting for sampling error
+# old code uses base plot
+
+```
+
+## Meiotic Error
+
+This section reviews all the data from achiasmy and 0CO bivalents.
+
+### Another Heterochiasmy Plot
+
+```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average"}
+
 no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
@@
 ```
-
-### female and male comparison
-
-## Subspecies level comparison
-
-Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
-
-*Does this hold up with the full data set?*
-*What's the best visualization for these questions?*
-
-Main focus is contrasting the level of difference within and between subspecies.
-
-I'll do this with a mixed model framework approach (and analyze with ANOVA, a method to partition the variance of subset.)
-
-```{r mixed.model, echo=FALSE}
-
-#build mixed model
-
-#run simulations for accounting for sampling error
-
-```
-
-### Subspecies examples / Data
-
-list the strains/data used 
-Mouse average MLH1 counts,
-
-### Male Specific results
-
-Not all strains have female data. Here are results from all strains wih male data.
-
-### female and male comparison
-
-## Meiotic Error
-
-This section reviews all the data from achiasmy and 0CO bivalents.

### 2019-09-10 doc/Main_Report.rmd
added files for lab meeting, work from home

@@
 
+### Quality effect
+
+there is a positive correlation with MLH1 number and quality
+To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.
+
 ### Age Effect
@@
 
+## Explaination for the XX and XY differences
+
+## Difference in proportion of chrm classes
+
+generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
+10 cells from each musc category.
+
+```{r draw.rand.cells, echo=FALSE}
+ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
+ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
+```
+
+```{r chrm_prop_data, echo=FALSE}
+
+chrm.class.DF = read.csv("~./MLH1repo/ChrmClassProportions.csv", header = TRUE)
+chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
+
+chrm.class.DF <- add_mouse(chrm.class.DF)
+chrm.class.DF <- add_strain(chrm.class.DF)
+chrm.class.DF <- add_sex(chrm.class.DF)
+chrm.class.DF <- add_category(chrm.class.DF)
+#chrm.class.DF <- add_euth_date(meta.data.file2)
+#chrm.class.DF <- add_age(meta.data.file2)
+
+table(chrm.class.DF$category)
+str(chrm.class.DF)
+
+#summarize by category
+this_thing2  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = sum(biv.sum.check),
+
+                 n0CO = sum(n0CO),
+                  n1CO = sum(n1CO),
+                  n2CO = sum(n2CO),
+                  n3CO = sum(n3CO),
+                  n4CO = sum(n4CO),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv,
+                 per4CO= sum(n4CO)/ tot_biv,
+
+                 mean.MLH1.foci = mean(MLH1.sum.check),
+                 mean.MLH1.foci2 = mean(nMLH1.foci),
+                 meanQ = mean(quality)
+
+)
+str(this_thing2)
+
+#this first one is the 'first' MLH1 count
+oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
+summary(oopp2)
+#none of the classes are significant
+
+#mean is the new calculated-- mean 0CO is significant and intercept
+oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
+summary(oopp)
+#not sure if these are inter-correlated
+#I don't know why 0COs and the intercept are significant..
+#0CO and 1CO are significant when only chrm classes are included (and intercept)
+
+#no longer significant when used
+#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
+#after added more MSM data, tot_biv is not significant
+#0CO is now more significant, and 2CO is sightly significant
+
+#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
+
+#MSM needs more bialent observations
+
+#use melt to put all propotions
+
+#display / visualize / compare proportions
+
+melt.table = read.csv("~./MLH1repo/ChrmClassProportions_melt.csv", header = TRUE)
+melt.table$fileName <- (melt.table$Original.Name)
+
+#melt.table <- add_mouse(melt.table)
+#melt.table <- add_strain(melt.table)
+#melt.table <- add_sex(melt.table)
+#melt.table <- add_category(melt.table)
+
+melt.table$strain<- factor(melt.table$strain,levels =c("LEW",
+                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM"), order=T )
+
+melt.table <- with(melt.table, melt.table[order(strain),])
+
+other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp)
+
+other_thing
+
+chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
+
+chrm.class.DF$biv.sum.check
+
+```
+
+PWD and MSM have a higher proportion of 2CO chrms!
+
+what about the variance across cells?
+
 ## Models for manuscript
@@
 ```
+
+## SC.skel measures
+
+run analysis for whole SC skel measures
+
+```{r RWscSkel, echo=FALSE}
+#merge the sc.skel outputs with the master MLH1 measures
+#match up random names with original names
+
+#the skel output and scripts are in the SC_skel dir. it was run on all images in the anon_data folder. I haven't manually confirmed that the default parameters are skeletonizing the images correctly.
+
+#some of richard's analysis show negative relationship with SC length and MLH1 (his QTL word doc). Could this be sign
+# of pwd like regulation?
+
+setwd("~./")
+SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_sep19.csv", header = TRUE, strip.white = TRUE)
+
+SC.skel$rand_name <- as.character(SC.skel$rand_name)
+original_DF$Random.Name <- as.character(original_DF$Random.Name)
+str(SC.skel)
+str(original_DF)
+
+mlh1.skel.og <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
+#for some reason this only worked after the chr and all = F
+
+mlh1.skel.og <- mlh1.skel.og[ !grepl("X", mlh1.skel.og$Batch) , ]
+mlh1.skel.og <- mlh1.skel.og[ !grepl("x", mlh1.skel.og$Batch) , ]
+mlh1.skel.og <- mlh1.skel.og[!(is.na(mlh1.skel.og$quality) | mlh1.skel.og$quality==""), ]
+
+mlh1.skel.og <- add_mouse(mlh1.skel.og)
+mlh1.skel.og <- add_strain(mlh1.skel.og)
+mlh1.skel.og <- add_sex(mlh1.skel.og)
+
+mlh1.skel.og <- add_category(mlh1.skel.og)
+mlh1.skel.og <- add_subsp(mlh1.skel.og)
+
+#mlh1.skel.og <- add_euth_date(mlh1.skel.og)
+#mlh1.skel.og <- add_age(mlh1.skel.og)
+
+str(mlh1.skel.og)
+mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
+mlh1.skel.og$bin_size <- as.numeric(mlh1.skel.og$bin_size)
+mlh1.skel.og$nMLH1.foci <- as.numeric(mlh1.skel.og$nMLH1.foci)
+
+#remove outliers
+mlh1.skel <- mlh1.skel.og[mlh1.skel.og$skel_size < 25000,]
+#first test file has ~1000 measures
+#when I ran it, no sign of outliers (after removing all the X's cells)
+
+looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=quality) ) + geom_jitter() + facet_wrap(sex~strain)
+
+looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Musc",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
+#very possible that the dimesions and pixel density are skrewing with these values
+# I am suspcious that there is such a wide range for the female data
+
+looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
+
+#mouse level
+mouse.level.skel <- ddply(.data=mlh1.skel.og, 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(nMLH1.foci),
+                 mean.skel = mean(skel_size)
+)
+
+mouse.level.skel <- add_strain(mouse.level.skel)
+mouse.level.skel <- add_sex(mouse.level.skel)
+mouse.level.skel <- add_category(mouse.level.skel)
+mouse.level.skel <- add_subsp(mouse.level.skel)
+
+looky.mouse <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Musc",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)
+
+looky.mouseDom <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Dom",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)
+
+#theme(legend.position="none")
+
+#double check a way to visually inspect the python skel output
+#
+#repeat random numbers?
+```
+
+```{r test.BivData, echo=FALSE}
+#bivdata
+setwd("~./")
+MSM_BivData.og = read.csv("~./MLH1repo/data/BivData/Curate_MSMmale_9.9.19.csv", header = TRUE, strip.white = TRUE )
+
+MSM_BivData <- MSM_BivData.og[MSM_BivData.og$SC.pass ==1,]
+MSM_BivData <- MSM_BivData[!(is.na(MSM_BivData$hand.foci.count) | MSM_BivData$hand.foci.count==""), ]
+
+#MSM_BivDatam_CO2 <- MSM_BivData[MSM_BivData$hand.foci.count == 2,]
+#MSM_BivDatam_CO2 <- MSM_BivDatam_CO2[!(is.na(MSM_BivDatam_CO2$fileName) | MSM_BivDatam_CO2$fileName==""), ]
+
+#MSM_BivDatam_CO2$Foci1.PER <- MSM_BivDatam_CO2$Foci1 /MSM_BivDatam_CO2$chromosomeLength
+#MSM_BivDatam_CO2$Foci2.PER <- MSM_BivDatam_CO2$Foci2 /MSM_BivDatam_CO2$chromosomeLength
+
+MSM_BivData <- add_IFD(MSM_BivData)
+#add IFD2 is f-ing up 
+MSM_BivData <- addIFD.2(MSM_BivData)
+
+#this MSM PER.IFD, is similar to the PWD data
+
+#new function is converting all hand.foci counts to 3
+#it might be the IFD functions
+
+str(new_shiny)
+str(MSM_BivData)
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+new_shiny <- add_SisCoTen(MSM_BivData)
+
+new_shiny$rel_Siscoten <- new_shiny$SisCoTen / new_shiny$chromosomeLength
+
+qq <- ggplot(data = new_shiny, aes(y=chromosomeLength, x=rel_Siscoten) ) + geom_jitter()
+
+ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
+  geom_boxplot(aes(alpha=.3))
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#KAZ and PWD data set is larger
+PWD_KAZ_BivData.og = read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD.KAZ_BivData.csv", header = TRUE, strip.white = TRUE )
+
+PWD_KAZ_BivData <- PWD_KAZ_BivData.og[PWD_KAZ_BivData.og$SC.pass == 1,]
+
+PWD_KAZ_BivData <- add_IFD(PWD_KAZ_BivData)
+PWD_KAZ_BivData <- addIFD.2(PWD_KAZ_BivData)
+PWD_KAZ_BivData <- add_SisCoTen(PWD_KAZ_BivData)
+PWD_KAZ_BivData$rel_Siscoten <- PWD_KAZ_BivData$SisCoTen / PWD_KAZ_BivData$chromosomeLength
+PWD_KAZ_BivData <- add_mouse(PWD_KAZ_BivData)
+PWD_KAZ_BivData <- add_strain(PWD_KAZ_BivData)
+
+mouse_boxes <- ggplot(data = PWD_KAZ_BivData, aes(y=rel_Siscoten, x=mouse) ) + geom_jitter(aes( color=fileName))+
+  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")
+
+mouse_boxes2 <- ggplot(data = PWD_KAZ_BivData, aes(y=SisCoTen, x=mouse) ) + geom_jitter(aes( color=fileName))+
+  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")
+
+#mean relative and raw sisco-ten is higher in PWD compared to KAZ
+
+biv_boxes <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count, fill=hand.foci.count) ) + geom_jitter()+geom_boxplot(aes(x= hand.foci.count,alpha=.3))+
+theme(legend.position="none")+facet_wrap(~strain)
+
+biv_boxes22 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=as.factor(hand.foci.count)) )+ geom_jitter()+
+  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")+facet_wrap(~strain)
+
+biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
+theme(legend.position="none")+facet_wrap(~strain)
+
+regress <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=chromosomeLength) ) + geom_jitter(aes( color=as.factor(hand.foci.count) ))+ facet_wrap(~strain)
+
+regress.1 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
+#, 1mar15_PWD_m1
+
+regress.2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
+#, 1mar15_PWD_m1
+
+#
+
+regress.1.KAZ <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$mouse == "21may18_KAZ_m2",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
+ theme(legend.position="none")+ facet_wrap(~fileName)
+
+#doesn't look like the sis-co-ten is conserved across bivs
+
+```
+
+```{r within.cell.corr, echo=FALSE}
+
+#test telomere distance
+#centromere distance
+#SC length
+
+#I want to test if SC lengths within cells cluster
+#wouldn't ANOVA get at this..?
+
+#cell level variance is significant predictor for SC length (telomere)
+
+#PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",]
+jjgg <- ddply(.data= PWD_KAZ_BivData,
+                 .(fileName),
+                 summarize, 
+                 nbiv = length(unique(Obj.ID) ),
+                  var.sc = var(chromosomeLength)
+)
+
+#reasons why this within cell correlation doesn't work well with bivalent data
+  #different sets of chrms
+#not enough resolution at cytology level
+
+```
+
+sis-coten metric:
+I am currently missing a null model / test for this metric
+(sis-co-ten from uniform placement of COs, and (some prob of draw for CO number))
+
+there is a positive correlation with siscoten and chrm length
+
+there could be something about kinetochore sites differing across specific chromosomes that balance the force

### 2019-09-11 doc/Main_Report.rmd
fixed buggs and removing un-needed sections

@@
 
-## Description of Report
-
-- Show basics of the current data file
-
-- Display plots and figures of distributions of MLH1 counts to assess initial overall patterns.
-
--  Report basic statistics of results and to help keep track of the number of images for each category and mouse.
-
 ## Discription of quantification process
@@
 
-```{r quant.repeatability, echo=FALSE}
+```{r quant.repeatability, echo=FALSE, include=FALSE}
 
@@
 
-there is a positive correlation with MLH1 number and quality
+There is a positive correlation with MLH1 number and quality
 To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.
 
+```{r quality.effect, echo=FALSE}
+
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
+#compare all quality 1 cells across strains
+#check that the means (or directions of differences is maintained)
+
+Q12_MLH1_DF <- MLH1_data[MLH1_data$quality <= 2,]
+
+#ddply
+Q12_MLH1_means <- ddply(.data=Q12_MLH1_DF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+tot.DF_MLH1_means <- ddply(.data=MLH1_data, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+mergy <- merge(Q12_MLH1_means, tot.DF_MLH1_means, by='category', all = F)
+colnames(mergy) <- c("category","Q12ncells","Q12mice","Q12_MLH1_means","Q12MLH1var",
+                     "Totncells","Totnmice","TotMLH1mean","TotMLH1var")
+
+mergy$percent.diff <- (mergy$Q12_MLH1_means / mergy$TotMLH1mean)
+
+Q12toTotal <- ggplot(mergy, aes(y=Q12_MLH1_means, x=TotMLH1mean))+geom_point(aes(size = Q12ncells) )+geom_abline(slope = 1, intercept = 0)+ggtitle("Quality 1 cells MLH1 mean to total MLH1 mean")+ylim(c(21,32))+xlim(c(21,32))
+
+Q12toTotal
+
+#check that the directions are the same
+#make tables for the 
+#is the amount of 'polymorphism' the same
+
+Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "female",], 
+                 .(category,subsp),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci)
+)
+
+tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "female",], 
+                 .(category,subsp),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci)
+)
+
+kable(Q12_MLH1_means_males, caption = "High quality cells")
+
+kable(tot.DF_MLH1_means_males, caption = "Total Cell")
+
+```
+
+All the quality 1 cells produce a higher mean. The biggest difference in means is HMI male, LEW female and SPIC female means have ~10% differences.
+
+When quality 1 to 2 is compared, the strains were closer to the line, but the means are still higher in the higher quality cells.
+with 1-2 most all of the difference are below 10% (except SPRET female).
+
+MOLF males have quite a big difference.
+
+There seem to be the similar amount of polymorphism in the Dom and more conserved in Dom.
+
+Ranges for male means across subspecies. Ranges of means (not ranges of all).
+
+(try turning this into a table...)
+Dom            female Dom
+old,   new      old     new
+22-24  23-25    24-27  25-29
+
+Musc           female Musc
+23-31   25-31   25-27   26-29
+22-25   22-27    25       25
+(ASTmale has 27 mean, outlier)
+
+Outgroup       female outgroup
+25-26   26-27    27    29-31
+
+I think the ranges across the groups
+
+### Assesment of Quality Scores
+
+> *How do the distributions change across different quality scores?*
+
+> *How do the distributions change across mice with a good number of cells?*
+
+```{r high.qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
+
+#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
+highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
+
+#are there no KAZ cells of high quality?
+qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
+high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
+```
+
 ### Age Effect
 
-Some papers have documented age effects for male genome-wide recombination rate (cite).
+Some papers have documented age effects for male genome-wide recombination rate (cite, one of the Cole papers).
 
@@
 
-### Assesment of Quality Scores
-
-> *How do the distributions change across different quality scores?*
-
-> *How do the distributions change across mice with a good number of cells?*
-
-```{r high.qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
-
-#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
-highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
-
-#are there no KAZ cells of high quality?
-qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
-high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
-```
-
 ```{r passing.mice, echo=FALSE, warning=FALSE}
@@
 ```{r draw.rand.cells, echo=FALSE}
-ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
-ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
+#ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
+#ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
 ```
@@
 
-table(chrm.class.DF$category)
-str(chrm.class.DF)
+#table(chrm.class.DF$category)
+#str(chrm.class.DF)
 
@@
 )
-str(this_thing2)
+#str(this_thing2)
 
@@
 oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
-summary(oopp2)
+#summary(oopp2)
 #none of the classes are significant
@@
 oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
-summary(oopp)
+#summary(oopp)
 #not sure if these are inter-correlated
@@
 
-chrm.class.DF$biv.sum.check
-
+#chrm.class.DF$biv.sum.check
 ```
@@
 original_DF$Random.Name <- as.character(original_DF$Random.Name)
-str(SC.skel)
-str(original_DF)
+#str(SC.skel)
+#str(original_DF)
 
@@
 
-str(mlh1.skel.og)
+#str(mlh1.skel.og)
 mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
@@
 
-str(new_shiny)
-str(MSM_BivData)
-
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
@@
 
-ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
-  geom_boxplot(aes(alpha=.3))
+#ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
+#  geom_boxplot(aes(alpha=.3))
 
@@
 
-biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
+biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( ))+
 theme(legend.position="none")+facet_wrap(~strain)

### 2019-09-16 doc/MatMeth.Rmd
made Intro md file and other writing updates
editing to material and methods

@@
 
-## Mouse Husbandry
+```{r set.up, echo=FALSE}
+#set wd
+#load datafile
+```
 
-1.	(Specific strains were obtained from X vendors, 2.
-2.	3 separate rooms used (precautions taken to ensure there was limited environmental variance) b.
-X additional husbandry things done to help with breeding of some mouse strains.
-housed in X conditions (extra things done for problem breeders,
-3.	euthanized under X protocol,
-4. mention results if there was a detectable environmental effect
+# Mouse Husbandry
 
-The Wild derived inbred strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ, CORLI/EiJ and SPRET/EiJ were purchased from Jax labs (Maine USA). 
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, HMI/TUA and SPIC) were ordered and cryo-derived from BRC at RIKEN (Japan) website).
+To access natural genetic variation for *Mus musculus*, Wild derived inbred strains strains were used.
+PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ, CORLI/EiJ and SPRET/EiJ were purchased from Jax labs (https://www.jax.org) (Maine USA).
+The strains of KAZ/TUA, TOM/TUA, AST/TUA, HMI/TUA and SPIC) were cryo-derived from Biological Resource Center (BRC) at RIKEN (Ibaraki, Japan) (https://en.brc.riken.jp/).
 All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.
- A breeding colony of wild derived mice sampled from Gough Island were housed at UW Veterinary school facilities.
-(some of the other strains housed there - comparisons were made of strains housed across different facilities to (test if there was an effect of mouse room of MLH1 counts).
-
+A breeding colony of wild derived mice sampled from Gough Island (GI )is maintained at UW Veterinary school facilities.
  Mice were feed on dry standard breeder chow.
@@
 
-## Tissue Collection and Immunohistochemistry
+Over the coruse of data collection some breeding colonies / mice were moved from facilities.
+Additionally the GI strain was kept at a seperate facility.
+(eviddence for environmental (room) effect on (MLH1 / SC lengths / a variety of meiotic traits was tested for.
+We observed no effect, Supplemental figure))
 
-1.	describe how gonad tissue was collected)  (right testes, both ovaries) (tunica removed decapsulated)  X protocol for testes,   X protocol for ovaries.  X ages for males, X ages for females
-2.	 describe spread process   (describe process - clearly point out the distinctions for testes and ovaries
-3.	(describe staining process (with DMC1)  (same process)
+# Tissue Collection and Immunohistochemistry
 
@@
 
-Meiocyte spreads were made following  [@peters_1997] Peters et al 2001.
+Meiocyte spreads were made following [@peters_1997].
 For testes the tunica was removed and whole testes was incubated in 3ml of hypotonic solution for 45min.
@@
 
-## Image Processing
-
-1.	Basic quality criteria,
-2.	Manual quant pipeline for MLH1 (anonimized)
-3.	Quanti pipeline for DMC1 (not completely finished)
+# Image Processing
 
@@
 
-## Statisical Analysis
+# Statisical Analysis
 

### 2019-09-17 doc/Main_Report.rmd
some updates to main report doc

@@
 A quality score is given; 1 to 5 (best to worst).
-
-all cell image file names are anonymized and moved into batches of a few hundred images.
+All cell image file names are anonymized and moved into batches of a few hundred images.
 
@@
 
+#in the chrm proportions data is another set for eastimating error
+
 ```
@@
 *time of year effect (on spread)*
+
+*dominant effect* (F1s)
 
@@
 
-Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "female",], 
+Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "male",], 
                  .(category,subsp),
@@
 
-tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "female",], 
+tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "male",], 
                  .(category,subsp),
@@
 
-Ranges for male means across subspecies. Ranges of means (not ranges of all).
-
-(try turning this into a table...)
-Dom            female Dom
-old,   new      old     new
-22-24  23-25    24-27  25-29
-
-Musc           female Musc
-23-31   25-31   25-27   26-29
-22-25   22-27    25       25
-(ASTmale has 27 mean, outlier)
-
-Outgroup       female outgroup
-25-26   26-27    27    29-31
-
-I think the ranges across the groups
-
-### Assesment of Quality Scores
-
-> *How do the distributions change across different quality scores?*
-
-> *How do the distributions change across mice with a good number of cells?*
+Ranges of category means across subspecies.
+
+| Type | male Dom | female Dom | | male Musc low | male Musc high | Musc female | | male Outgroup | female outgroup |
+| --- | --- | --- | --- | --- | --- | --- | --- | --- | --- |
+| Total cells | 22-24 |  24-27 |  | 23-31 | 25-31 | 26-29 | | 25-26 |27 |
+| High quality cells | 23-25 | 25-29 | | 22-25 | 22-27 | 25 | | 26-27 | 29-31 |
+(ASTmale in Musc high, has a 27 mean which might be an outlier)
+I think the ranges across the groups is generally consistant.
+
+(try looking at <5 quality scores?)
 
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
+        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey", 
+             "grey2","grey3","grey1","grey","grey","grey"))
 high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
-```
+#high_qual_boxplot
+#these boxplots look bad
+```
+
+> *How do the distributions change across different quality scores?*
+
+Higher quality scored cells have more MLH1 foci. (some of the ranges get smaller with higher quality cells)
+
+> *How do the distributions change across mice with a good number of cells?*
+
+(was I referring to normale distributions?)
+means all shift up, what about variance, range, lowest cut off
+
+> *Do the directions of the differences across groups change due to quality scores?*
+
+(polymorphism vs divergence patterns)
 
@@
 
-lm.me  <- lm(male.data$mean_co ~ male.data$age.weeks+male.data$strain.x)
-
-#summary(lm.me)
-```
+male.data.dom <- male.data[male.data$subsp.y == "Dom",]
+
+male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST"),]
+
+male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "MOLF" | male.data$strain.x == "SKIVE"),]
+
+male.data.outgroup <- male.data[(male.data$subsp.y == "Spretus" | male.data$subsp.y == "Spic" | male.data$subsp.y == "Caroli"),]
+
+lm.me.dom  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.dom)
+
+lm.me.musc.low  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.low)
+
+lm.me.musc.high  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.high)
+#significant differences across strains in musc high
+
+lm.me.OG  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.outgroup)
+
+summary(lm.me.OG)
+#for all models, mostly it was that the intercept was significant, and mot much else
+```
+
+```{r female.age.effect, echo=FALSE}
+#female age effect
+
+#female.meta.data <- meta.data[meta.data$sex == "female",]
+#how much data to I have to work with?
+#there's a fair number of rows without emb or neonate
+#I could probably assume all of these are neonates
+
+#length( !is.na(female.meta.data$female.stage) )
+
+#female.meta.data <- female.meta.data[female.meta.data$female.stage]
+
+#how much formating do I have to do?
+#can't use AP_mouse_table since that is summarized
+#meta.data
+#grepl notes, "neo"
+
+#some have e-number for embryonic
+#I might have to
+
+```
+
+I will try doing a logistic regression for embryonic vs neonate. The embyonic estimates likely aren't very accurate, and there won't be good sampling across time points.
 
@@
 
+(in progress)
+
 ## Difference in proportion of chrm classes
@@
 
-Rough ideas for models of evolution of Heterochiasmy and current idea for how to test them.
+Ideas for models of evolution of Heterochiasmy and current idea for how to test them.
 
@@
 
-run analysis for whole SC skel measures
+data has been collection for RWang's total SC. These results are consistant with the patterns seen in the bivalent level resutls. (total SC PWD > KAZ )
 
@@
 # I am suspcious that there is such a wide range for the female data
+looky
 
 looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
+looky.DOM
 

### 2019-09-17 doc/MatMeth.Rmd
barfed more text for mat and meth. needs cleaning up

@@
     toc_float: yes
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
 bibliography: refs_MatMeth.bib
@@
 
-From males, the right teste was collected and processed. (Ages are listed in supplmental material). Ovaries were collected from neonates less than 48 hours old or embryos between e16 to 21.
+From males, the right teste was collected and processed as discribed in X citation.
+The majority of mice used were betwee 5 and 12 weeks, (supplemental table).
+However some strains had problems breeding, so testes from older mice were collected.
+We found a (small/no) age effect on MLH1 counts for male mice (Supplemental Figure)).
 
-The behavior of these strains did not allow us to check for copulary plugs, resulting in most oocyte data coming from neonate mice between 5 to 48 hours.
+The vast majority of oocyte data was collected from neonate litters/mice between 5 to 48 hours.
+This approach maintain breeding pairing and still result in prophase oocytes (cite timeline of oocytes in neonates).
+(differences were assessed between embryo samples and neonates?).
+Presice staging of embryos by checking copulary plugs was diffcult/not feasible in many of these wild derived strains due to their behavior.
+Embryonic samples were collected when pregenacgy noted in females, embryos were staged based on (X table markers).
 
@@
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The next morning, slides were treated with a photoflow wash.
+The next morning, slides were treated with a photoflow (Kodak, cite) wash.
+
+Staining / Immunohistochemistry
+
+The same staining protocol was applied to spermatocyte and oocyte spreads.
+The staining protocol was bas on X with some adaptations, and same as previously discribed in Peterson etal 2019.
 
@@
 
-For MLH1 pachytene characterization, Cells with a full karyotype (19 autosomes and XY for spermatocyte spreads or 20 bivalents for oocyte spreads), distinct foci, and intact bivalents were included for quantification.
+Images of cells were capture on Axio-2 microscope.
 
-Reprocessing in Photoshop ()- Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
+For MLH1 pachytene characterization, Cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
+Reprocessing in Photoshop (cite). Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
+
+Hand measures dataset generated by using ImageJ.
+
+total Sc was measured (Rwang et al 2019)
+
+Bivalent level measure were done (Peterson et al 2019)
 
@@
 
--done in R/ specific packages
+Statistical analysis was performed in R (version and cite).
+
+The distributions of CO counts per cell was assessed for normailty (supplemental figure).
+Mouse mean MLH1 count was used to get around the bad discreet nature of count data.
+(normality was confirmed with X distrbution, supplememtal figure).
+
+-specific statistical tests; (means, non-parameteric kurts wallis t-test?) (correlation coeffectent (function))
+linear models lm() from X package.
+
+-specific R packages
 -non-parameteric  (tests and measures for dealing with count data)
-
 -hypothese tested (size, DSB number)
 
+Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mixmodel frame work, using X package (cite).
+With X formula Co count ~ blah + blah, with random strain specific sex effect.
+
+(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closeley related species.)
+
 # References

### 2019-09-17 doc/PowerCalq_Report.Rmd
updating doc/ dir

--- 2018-05-29 Main_Report.rmd
@@
 ---
-title: 'MLH1 Data Report'
+title: "Power Calculation Report"
 author: "April Peterson"
-date: "29 may 2018"
-output: rmarkdown::github_document
+date: "June 12, 2019"
+output: html_document
 ---
 
-```{r setup, echo=FALSE, include = FALSE}
-library(knitr)
-library(ggplot2)
-#library(pwr)
-library(plyr)
-library(lattice)
-library(dplyr)
-setwd("C:/Users/alpeterson7/Documents/MLH1repo")
-#setwd("C:/Users/April/Desktop/MLH1repo")
-load(file="MLH1_data_setup.RData")
-```
-
-#### Research Description
-Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice abd related rodents.
-Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
-
-#### Description of Report
-
-- Display plots and figures of distributions of MLH1 counts to assess intial overall patterns.
-
--  *Report basic statistics of results and to help keep track of the number of images for each category and mouse*
-
-#### Discription of quantification process
-
-Cell images are quantified in batches after file names are anonymized.
-The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified.
-A quality score is given; 1 to 5 (best to worst).
-
-```{r,echo=FALSE}
-#set the order of another column, based on another variable (so that when)
-
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, category, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-#use the above line for when order is out of wak
-
-#count the non-quality measures,  #remove non qualit
-#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
-#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
-
-#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
-#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(category) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
-#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-
-```
-
-#### Initial Patterns from MLH1 distributions
-
-```{r first boxplots, echo=FALSE, fig.align='center'}
-#the below ordering works!! keep it
-MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEWES",
-              "WSB", "PWD","MSM","KAZ","CAST","HMI","SPRET","SPIC", "other"), order=T )
-#order the data frame
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-#set the order of another column, based on another variable (so that when)
-
-#try to make sure sacles are the same
-
-#reduncant
-#MLH1_data <- MLH1_data %>%
- # arrange(strain, sex, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-ff <- ggplot(MLH1_data, aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green","purple","grey1", "grey" )) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5))
-ff <- ff   + facet_wrap(~ sex, scales="free")
-
-#boxplots by strain
-ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot()  + ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey" )) +
-  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )
-
-ss <- ss   + facet_wrap(~ sex, scales="free")
-
-ff
-ss
-```
-
-After taking the data from 2 highest cell quality, some but not all mouse means converge.
-
-```{r, histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
-#consider making distribution with quality considered...
-
-histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38))
-histM <- histM + geom_histogram(stat = "bin", binwidth = 1)
-histM <- histM + facet_wrap(~ category)
-histM
-```
-
-```{r high qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE}
-
-#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
-highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
-
-#are there no KAZ cells of high quality?
-qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
-high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
-```
-
-```{r, passing mice, echo=FALSE}
-q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
-                        total =  length(adj_nMLH1.foci),
-                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
-                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
-                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
-)
-#q_cutoff_table
-passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
-
-not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
-
-#mnake dataframe with just passing mice -- so that it can be ploted
-passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
-```
-
-#### Comparison of MLH1 distributions
-
-```{r,  show boxplots, fig.height=5, fig.width=8, echo=FALSE}
-high_qual_boxplot
-#not sure why order got messed up again...
-```
-
-After taking the data from 2 highest cell quality, some but not all mouse means converge.
-For some mice (mostly female) the majority of cells are excluded.
-This is especially true for PWD and WSB females.
-
-#### Tests for quality and nMLH1 foci number
-
-> *How do the distributions change across different quality scores?*
-
-> *How do the distributions change across mice with a good number of cells?*
-
-##### Effects of quality
-
-Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
-Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
-(CAST female data is fake)
-
-```{r, scatter plots of nMLH1 by score, echo=FALSE, fig.cap="caption"}
-#make lattice / plots seperated by strain and sex
-#facet
-
-#quality scores are now 1:7?
-qual_means <- MLH1_data %>% 
-        group_by(category, quality) %>% 
-        summarise(
-          adj_nMLH1.foci = mean(adj_nMLH1.foci)
-        )
-
-p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
-p <- p + geom_point(stat = "identity")
-p <- p  + geom_point(data = qual_means, colour="red") 
-p <- p + facet_wrap(~ category)
-p
-
-```
-
-I ploted the mean of each quality bin with a red dot.
-From the pattern of the red dots, there is definately a negative relationship with quality and nMLH1 foci across the data.
-This is most pronounced in MSM males and least pronounced in G males.
-The CAST female data is dummy data.
-
-#### Assessing the Distributions by mouse
-
-latice plot of scatter plots for jitter plots of cell oberservations by quality. The category mean is in black and the mouse specific mean is in red.
-```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE}
-### to integrate 'pass'/fail
-### this should be in setup data
-
-#what is the purpose of this block
-df_list <- split(MLH1_data, (MLH1_data$category))
-#remove sections which are in category but have no data
-df_list <- df_list[ -c(7,14,16,22) ]
-#7: PERC male, 12: KAZ female! (14:CAST female, fake data) 16:HMI female
-
-#df list, is a list of df's which were split() from BIG df by category
-df_list 
-o=0
-i=0
-
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-for(i in 1:(length(df_list))){
- # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                       mouse_nMLH1 = mean(adj_nMLH1.foci)
-  )
-# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-  for(o in 1:length(df_list[[i]]$Original.Name)){
-    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
-        }
-}
-
-plot_list = list()
-for (h in 1:length(df_list)) {
-  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-  pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
-  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") 
-  #theme for indiviual mice, 
-  #+ theme(plot.background = element_rect(fill = "green"))
-  plot_list[[h]] =   pf
-}
-
-#aes must be same length ---   #geom_point(stat = "identity") +
-# the last one is correct....?
-
-#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
-```
-
-```{r, echo=FALSE, message=FALSE,results='hide', warning=FALSE }
-geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
-  #invisible(lapply(obj, function(x) plot(x,main="some plot")))
-}
-```
-
-Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
-The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
-
-PWD females and a very large range. This is driven by data from 17apr15_PWD_f4. These data should be looked at more closely.
-
-In some mice there seems to be bimodel cell populations, those of clearly high quality and those of lower quality.
-This may be due to the screening of images before quantification, and allowing more bad cells data remain in the data set, instead of deleting them.
-
-#### Mice table
-
-Table of the mice used and MLH1 stats.
-
-```{r, fig.height=2, fig.width=4, echo=FALSE}
-#this table is 
-#this table is saved in the .RData file
-#use kable to make tables. warning this might not work for github Rmd.
-n <- kable(AP_mouse_table)
-n
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```

### 2019-09-18 doc/MatMeth.Rmd
added my citation, more cleaning up

@@
 
-To access natural genetic variation for *Mus musculus*, Wild derived inbred strains strains were used.
-PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ, CORLI/EiJ and SPRET/EiJ were purchased from Jax labs (https://www.jax.org) (Maine USA).
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, HMI/TUA and SPIC) were cryo-derived from Biological Resource Center (BRC) at RIKEN (Ibaraki, Japan) (https://en.brc.riken.jp/).
-All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.
-A breeding colony of wild derived mice sampled from Gough Island (GI )is maintained at UW Veterinary school facilities.
+To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
+The Mus musculus PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
+The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
+(https://en.brc.riken.jp/).
+The related species of *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jax labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
+
+ All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.  A breeding colony of wild derived 
+*Mus musculus* mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
  Mice were feed on dry standard breeder chow.
@@
 
-Over the coruse of data collection some breeding colonies / mice were moved from facilities.
-Additionally the GI strain was kept at a seperate facility.
-(eviddence for environmental (room) effect on (MLH1 / SC lengths / a variety of meiotic traits was tested for.
+Over the course of data collection some breeding colonies / mice were moved from facilities.
+Additionally the GI strain was kept at a separate facility.
+(evidence for environmental (room) effect on (MLH1 / SC lengths / a variety of meiotic traits was tested for.
 We observed no effect, Supplemental figure))
@@
 
-From males, the right teste was collected and processed as discribed in X citation.
-The majority of mice used were betwee 5 and 12 weeks, (supplemental table).
+Spermatocyte spreads were collected and perpared as described in [@peterson_2019] .
+The majority of mice used were between 5 and 12 weeks, (supplemental table).
 However some strains had problems breeding, so testes from older mice were collected.
 We found a (small/no) age effect on MLH1 counts for male mice (Supplemental Figure)).
+
+The only difference between preparation of spermatocyte and oocyte spreads was the volumns of the hypotonic buffer was 300$\mul instead of 3ml.
 
@@
 (differences were assessed between embryo samples and neonates?).
-Presice staging of embryos by checking copulary plugs was diffcult/not feasible in many of these wild derived strains due to their behavior.
-Embryonic samples were collected when pregenacgy noted in females, embryos were staged based on (X table markers).
+Precise staging of embryos by checking copulary plugs was difficult/not feasible in many of these wild derived strains due to their behavior.
+Embryonic samples were collected when pregnancy noted in females, embryos were staged based on (X table markers).
 
+<!--
 Meiocyte spreads were made following [@peters_1997].
@@
 The next morning, slides were treated with a photoflow (Kodak, cite) wash.
+-->
 
@@
 The same staining protocol was applied to spermatocyte and oocyte spreads.
-The staining protocol was bas on X with some adaptations, and same as previously discribed in Peterson etal 2019.
+The staining protocol was based on X with some adaptations, and same as previously described in Peterson et al 2019.
 
@@
 
-Hand measures dataset generated by using ImageJ.
+Hand measures data set generated by using ImageJ.
 
@@
 Mouse mean MLH1 count was used to get around the bad discreet nature of count data.
-(normality was confirmed with X distrbution, supplememtal figure).
+(normality was confirmed with X distribution, supplemental figure).
 
--specific statistical tests; (means, non-parameteric kurts wallis t-test?) (correlation coeffectent (function))
+framework for analyzing the MLH1 patterns
+1. write equation
+
+ $X \sim MLH1_mouse average ~ sex * subsp + random(strain)+ strain*sex+$\xi
+
+mixed model, observations of mouse means of mlh1, all data with both male and female observations
+
+subspecies, sex and subsp*sex are fixed effects
+strain as random effect (sample of a genetic background from subspecies)
+with domesticus male set as the default
+
+strain and sex interaction effect was included to ask / test if the effect of sex varies by strain(genetic background)
+
+this model allows us to estimate and test the fixed effects (sex and subsp)
+
+the random effect of strain in this model can be interpreted as the average standard deviation due to the random strain effect
+and sex*strain,
+
+From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.
+
+We can also compare the size of the sex effects between (subspecies) and within (strain) mouse groups.
+(ie is the male or female strain effect larger?
+is there more variance due to random strain effect in male or females?)
+
+I may have to do more research/reading on testing the random effects.
+
+-specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
 linear models lm() from X package.
@@
 -specific R packages
--non-parameteric  (tests and measures for dealing with count data)
--hypothese tested (size, DSB number)
+-non-parametric  (tests and measures for dealing with count data)
+-hypothesis tested (size, DSB number)
 
-Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mixmodel frame work, using X package (cite).
+Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mix model frame work, using X package (cite).
 With X formula Co count ~ blah + blah, with random strain specific sex effect.
 
-(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closeley related species.)
+(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)
 

### 2019-09-18 doc/MatMeth.Rmd
started thinking about results. more updates to material methods

@@
 ---
-title: "Material Metohds"
+title: "Material Methods"
 date: "`r Sys.Date()`"
@@
 #load datafile
+
+#common bugs
+#html won't compile -- one of my citations in bib was breaking things
 ```
@@
 The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp/).
+(https://en.brc.riken.jp).
 The related species of *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jax labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
@@
 
-Over the course of data collection some breeding colonies / mice were moved from facilities.
+Over the course of data collection some breeding colonies  mice were moved from facilities.
 Additionally the GI strain was kept at a separate facility.
@@
 
-The only difference between preparation of spermatocyte and oocyte spreads was the volumns of the hypotonic buffer was 300$\mul instead of 3ml.
+The only difference between preparation of spermatocyte and oocyte spreads was the volumns of the hypotonic buffer was 300 u l instead of 3ml.
 
@@
 
-<!--
 Meiocyte spreads were made following [@peters_1997].
@@
 The next morning, slides were treated with a photoflow (Kodak, cite) wash.
--->
 
@@
 
+Explain list recipes for DMC1 since these haven't been listed before
+
 # Image Processing
 
-Images of cells were capture on Axio-2 microscope.
+Images of cells were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
+
+Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
 
@@
 
-Hand measures data set generated by using ImageJ.
+Repeatability was assessed by X (comparing mulitple counted images)
 
-total Sc was measured (Rwang et al 2019)
+Hand measures data set generated by using ImageJ/Fiji (v1.52) [@fiji].
 
-Bivalent level measure were done (Peterson et al 2019)
+total Sc was measured [@wang_2019_sc]
+
+Bivalent level measure were done [@peterson_2019]
 
@@
 
-Statistical analysis was performed in R (version and cite).
+Statistical analysis was performed in R (v3.5.2)[@Rstudio].
 
@@
 
- $X \sim MLH1_mouse average ~ sex * subsp + random(strain)+ strain*sex+$\xi
+$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
+ can't get the equation in the right format. I think this breaks the pdf kniting.
 
@@
 the random effect of strain in this model can be interpreted as the average standard deviation due to the random strain effect
-and sex*strain,
+and sex  strain,
 

### 2019-09-23 doc/Main_Report.rmd
working on repeatability chunck, wrote add_fileName function for cleaning old MLH1 excel sheets

@@
 
-*What is the Error in quantification across humans?*
-
-A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%.
-(using some of Beth's data).
-There is a subset of data which has been counted multiple times -- use that to test the repeatbility.
-
-```{r quant.repeatability, echo=FALSE, include=FALSE}
+Well, we have to talk about the subjective nature that goes into scoring the MLH1 foci.
+Sources of error or variance across scoring come from background noise (false positives), faint foci (false negatives) and just human error.
+A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (using some of Beth's data).
+Todo: track down, re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)
+
+```{r quant.repeatability, echo=FALSE, include=TRUE}
+#load, clean up and mat one of the master.MLH1count sheets
+#master
+old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
+#remove all with 'X'
+#create a real fileName col, add rev.tif,
+
+old.master.MLH1 <- add_mouse(old.master.MLH1)
+old.master.MLH1 <- add_strain(old.master.MLH1)
+old.master.MLH1 <- add_sex(old.master.MLH1)
+old.master.MLH1 <- add_category(old.master.MLH1)
+old.master.MLH1$fileName2 <- old.master.MLH1$fileName
+
+test.old.master <- rbind(head(old.master.MLH1), tail(old.master.MLH1) )
+test.old.master$fileName <- as.character(test.old.master$fileName)
+
+test.old.master$stain <- as.character(test.old.master$stain)
+
+str(test.old.master)
+
+#paste together file name, requires mouse, strain
+#this might also need some character / factor stuff ... filename and stain MUST be charaecter!
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#wrote add_fullFileName to make the file names for the old excel sheets
+
+uutesting <- add_fullFileName(old.master.MLH1)
+
+#merge and match to MLH1
+
+#match old.full.filname with MLH1$fileName
+
+chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
+#what are the target columns for this excel sheet?
+#espically for getting repeat data
+#MLH1.dif  same_count  quality
+
+chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
+chrm.class.DF <- add_mouse(chrm.class.DF)
+chrm.class.DF <- add_strain(chrm.class.DF)
+chrm.class.DF <- add_sex(chrm.class.DF)
+chrm.class.DF <- add_category(chrm.class.DF)
+
+#assessing error,
+#number for same_count == TRUE
+#length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"])
+#length(chrm.class.DF$Original.Name)
+
+#cor of first count and calq count
+#cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)
+#.87
+
+#the true false might have to be changed first
+
+chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)
+
+#regression of quality and MLH1.dif
+lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
+#summary(lmreg)
+#according to this linear regression, quality is not a significant
+
+#make regression for just same counts
+#I tried making a  binomial regression model, 
+#but quality is bad predictor, only 5 levels)
+lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
+#summary(lmreg2)
+#to plot the model / bimodal data
+#fit = glm(same_count_num ~ as.numeric(quality), data=chrm.class.DF, family=binomial)
+#newdat <- data.frame(quality=seq(min(chrm.class.DF$quality), max(chrm.class.DF$quality),len=102))
+#newdat$vs = predict(fit, data=newdat, type="response")
+#the prediction is not going 1 or 0 -- so doesn't fit logistic
+#plot(vs ~ quality, data=newdat, col="red4")
+#lines(vs ~ quality, newdat, col="green4", lwd=2)
+
+hist_error_plot <- ggplot(data=chrm.class.DF, aes(MLH1.dif))+geom_histogram()+ggtitle("Distribution of foci count differences across 2 count versions")
+
+#
+quality_error_plot <- ggplot(data=chrm.class.DF, aes(x=quality, y=abs(MLH1.dif)  ))+geom_jitter()+ggtitle("Repeatability by quality score")+ylab("difference in quality score")+xlab("quality score")
+
+error_count_plot <- ggplot(data=chrm.class.DF, aes(y=nMLH1.foci, x=abs(MLH1.dif)  ))+geom_jitter()+ggtitle("Error by MLH1 number")+xlab("difference in two counts")+ylab("MLH1 count")
 
 #in the chrm proportions data is another set for eastimating error
-
-```
+```
+
+Data use for this repeatabiity estimate is from smaller dataset where the proportions of chromosome/bivalent classes (0CO bivalents ect.) were quantified.
+(The focus of one count was cell total, while the other focus was classifying bivalents).
+
+This data has `r length(chrm.class.DF$Original.Name)` cells across these categories
+
+`r kable(table(droplevels(chrm.class.DF$category)))`
+
+More cells are planned to be quantified.
+
+```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE}
+hist_error_plot
+```
+
+From this histogram we see most of the cells in this dataset have a difference of 2 in the MLH1 count scores (`r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`).
+
+```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5}
+quality_error_plot
+```
+
+There is a positive relationship with the absolute difference and quality score (where 1 is best and 5 is worst quality) (i.e.
+poorer quality cells are more likely to have more error acorss multiple counts).
+The proportion of the cells which had the same count across the counting instances is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`.
+The correlation between the two versions of MLH1 counts for the same cell is `r cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)`.
+
+```{r show.error_count_plot, echo=FALSE,  fig.width=10, fig.height=5}
+error_count_plot
+```
+There doesn't seem to be a bias across cells with more or less foci.
+
+Modeling this relationship; (still in progress)
+I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.
 
 ## Effects on the dataset
+
+error associated with 'quality of cels'
 
@@
 
-chrm.class.DF = read.csv("~./MLH1repo/ChrmClassProportions.csv", header = TRUE)
+chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
 chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-
 chrm.class.DF <- add_mouse(chrm.class.DF)
@@
 #summarize by category
-this_thing2  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
+proportion_stats_category  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
                  .(category),
@@
 #this first one is the 'first' MLH1 count
-oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
+oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
 #summary(oopp2)
@@
 #mean is the new calculated-- mean 0CO is significant and intercept
-oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = this_thing2)
+oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
 #summary(oopp)
@@
 
-melt.table = read.csv("~./MLH1repo/ChrmClassProportions_melt.csv", header = TRUE)
+melt.table = read.csv("~./MLH1repo/data/ChrmClassProportions_melt.csv", header = TRUE)
 melt.table$fileName <- (melt.table$Original.Name)

### 2019-09-23 src/CommonFunc_MLH1repo.R

@@ -7,6 +7,29 @@
 #-check for multi-cell images
 
 
+#make fileName
+#paste together file name
+add_fullFileName <- function(dat){
+  count =1
+  dat$fileName <- as.character(dat$fileName)
+  dat$stain <- as.character(dat$stain)
+  
+  for(i in dat$fileName){  #i is file name!!!
+    #print(i)
+    templist1= strsplit(i, split="_")[[1]] #put the 
+    
+    j=dat$stain[count]
+    templist2= strsplit(j, split="_")[[1]]
+    c = paste(templist1[2], templist1[3],templist1[4], sep = "_")#mdate, strain, m#
+    #test.old.master$mouse2[count] <- c #don't need to add or change mouse
+    
+    d = paste(templist1[1], templist1[2], templist1[3],templist1[4],templist2[3], dat$cell[count], 'rev.tif', sep = "_") 
+    dat$file.name_full[count] <- d #should match file name
+    count= count + 1
+  }
+  return(dat)
+}
+
 #format the dates into standard format
 #mouse_list$DOB <- as.Date(mouse_list$DOB,format='%m/%d/%Y')
 add_age <- function(dat){

### 2019-09-26 doc/MatMeth.Rmd
night notes

@@
 
-Statistical analysis was performed in R (v3.5.2)[@Rstudio].
+0. discribe basic assesment and comparison of MLH1 means (and how it is proxy for genome wide recombination rates)
+- assessment and treatment of data quality control ect (correlation measured X)
 
@@
 (normality was confirmed with X distribution, supplemental figure).
+<re-write this dumb paragraph>
 
-framework for analyzing the MLH1 patterns
-1. write equation
+Statistical analysis was performed in R (v3.5.2)[@Rstudio]. mixed models use lmer4 package (cite)
+
+1. framework for analyzing the MLH1 evolutionary patterns across mus species complex. (what is the main question we were tackling when designing this framework?)
+
+2. write equation, discribe terms (proxies for within and between polymorphsm)
+logic / choices for fixed and random effects
+
+3. what are the 2 general / basic questions this framework can be applied to.
 
 $X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
- can't get the equation in the right format. I think this breaks the pdf kniting.
+< can't get the equation in the right format. I think this breaks the pdf kniting.>
+
+4. how can DMC1 measures be added to the mixed model
+
+5. discriptions of mlh1 outside of framework (missing female data)
 
@@
 
--specific R packages
--non-parametric  (tests and measures for dealing with count data)
--hypothesis tested (size, DSB number)
-
 Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mix model frame work, using X package (cite).
-With X formula Co count ~ blah + blah, with random strain specific sex effect.
-
 (we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)

### 2019-09-26 doc/MatMeth.Rmd
forgot the change

--- 2019-09-18 doc/MatMeth.Rmd
@@
 
-Hand measures data set generated by using ImageJ/Fiji (v1.52) [@fiji].
+Measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
 
@@
 
-framework for analyzing the MLH1 patterns
+We choose to use a **mixed model framework** to interpert the patterns of variation. For interperting patterns of mean MLH1 count within and between house mouse subspecies.
+
 1. write equation
 
-$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
- can't get the equation in the right format. I think this breaks the pdf kniting.
+$X \sim mouse average MLH1 ~ sex * subspecies + random(strain) + \xi $
 
+In this model
 mixed model, observations of mouse means of mlh1, all data with both male and female observations
 
-subspecies, sex and subsp*sex are fixed effects
-strain as random effect (sample of a genetic background from subspecies)
-with domesticus male set as the default
+The main effects and the interaction of subspecies and sex were coded as fixed.
 
-strain and sex interaction effect was included to ask / test if the effect of sex varies by strain(genetic background)
+Subspecies, sex and subsp*sex are fixed effects
 
-this model allows us to estimate and test the fixed effects (sex and subsp)
+Strain was coded as a random effect, since in this experimental design we consider it as a sample of natural genetic variation for each subspecies.
+We also included a term to allow for strains to have a specific sex effect.
 
-the random effect of strain in this model can be interpreted as the average standard deviation due to the random strain effect
-and sex  strain,
+This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
+
+We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner)
 
@@
 I may have to do more research/reading on testing the random effects.
-
 -specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
 linear models lm() from X package.
-
 -specific R packages

### 2019-09-26 doc/MatMeth.tex
forgot the change

--- 2019-09-18 doc/MatMeth.Rmd
@@
----
-title: "Material Methods"
-date: "`r Sys.Date()`"
-output:
-  html_document:
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
-bibliography: refs_MatMeth.bib
----
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Material Methods},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-```{r set.up, echo=FALSE}
-#set wd
-#load datafile
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-#common bugs
-#html won't compile -- one of my citations in bib was breaking things
-```
+%%% Change title format to be more compact
+\usepackage{titling}
 
-# Mouse Husbandry
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
-To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
-The Mus musculus PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp).
-The related species of *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jax labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
+\setlength{\droptitle}{-2em}
 
- All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.  A breeding colony of wild derived 
-*Mus musculus* mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
- Mice were feed on dry standard breeder chow.
-Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
-Adult mice were euthanized by CO asphyxiation.
-Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison.
+  \title{Material Methods}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{}
+    \preauthor{}\postauthor{}
+      \predate{\centering\large\emph}
+  \postdate{\par}
+    \date{2019-09-18}
 
-Over the course of data collection some breeding colonies  mice were moved from facilities.
-Additionally the GI strain was kept at a separate facility.
-(evidence for environmental (room) effect on (MLH1 / SC lengths / a variety of meiotic traits was tested for.
-We observed no effect, Supplemental figure))
+\begin{document}
+\maketitle
 
-# Tissue Collection and Immunohistochemistry
+{
+\setcounter{tocdepth}{5}
+\tableofcontents
+}
+\section{Mouse Husbandry}\label{mouse-husbandry}
 
-Spermatocyte spreads were collected and perpared as described in [@peterson_2019] .
-The majority of mice used were between 5 and 12 weeks, (supplemental table).
-However some strains had problems breeding, so testes from older mice were collected.
-We found a (small/no) age effect on MLH1 counts for male mice (Supplemental Figure)).
+To access natural genetic variation for \emph{Mus musculus}, wild
+derived inbred strains were used. The Mus musculus PWD/PhJ, PERC/Eij,
+WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were
+purchased from Jackson labs (Maine, USA). The strains of KAZ/TUA,
+TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource
+Center (BRC) at Riken (Ibaraki, Japan). (\url{https://en.brc.riken.jp}).
+The related species of \emph{Mus caroli} CAROLI/EiJ and \emph{Mus
+spretus} SPRET/EiJ were purchased from Jax labs
+(\url{https://www.jax.org}) and \emph{Mus spicilegus} SPIC/Eij from
+Riken.
 
-The only difference between preparation of spermatocyte and oocyte spreads was the volumns of the hypotonic buffer was 300 u l instead of 3ml.
+All mice were housed UW-Madison Biotech and MSC facilities - following
+the protocols. A breeding colony of wild derived \emph{Mus musculus}
+mice sampled from Gough Island (GI)is maintained at UW Veterinary school
+facilities. Mice were feed on dry standard breeder chow. Some strains
+sunflower seeds, nestlets and larger cages were added to improve
+fertility and litter survival. Adult mice were euthanized by CO
+asphyxiation. Neonate and embryonic mice were euthanized by decapitation
+following the protocols approved by the Institutional Animal Care and
+Use Committee at the University of Wisconsin-Madison.
 
-The vast majority of oocyte data was collected from neonate litters/mice between 5 to 48 hours.
-This approach maintain breeding pairing and still result in prophase oocytes (cite timeline of oocytes in neonates).
+Over the course of data collection some breeding colonies mice were
+moved from facilities. Additionally the GI strain was kept at a separate
+facility. (evidence for environmental (room) effect on (MLH1 / SC
+lengths / a variety of meiotic traits was tested for. We observed no
+effect, Supplemental figure))
+
+\section{Tissue Collection and
+Immunohistochemistry}\label{tissue-collection-and-immunohistochemistry}
+
+Spermatocyte spreads were collected and perpared as described in
+(Peterson, Miller, and Payseur 2019) . The majority of mice used were
+between 5 and 12 weeks, (supplemental table). However some strains had
+problems breeding, so testes from older mice were collected. We found a
+(small/no) age effect on MLH1 counts for male mice (Supplemental
+Figure)).
+
+The only difference between preparation of spermatocyte and oocyte
+spreads was the volumns of the hypotonic buffer was 300 u l instead of
+3ml.
+
+The vast majority of oocyte data was collected from neonate litters/mice
+between 5 to 48 hours. This approach maintain breeding pairing and still
+result in prophase oocytes (cite timeline of oocytes in neonates).
 (differences were assessed between embryo samples and neonates?).
-Precise staging of embryos by checking copulary plugs was difficult/not feasible in many of these wild derived strains due to their behavior.
-Embryonic samples were collected when pregnancy noted in females, embryos were staged based on (X table markers).
+Precise staging of embryos by checking copulary plugs was difficult/not
+feasible in many of these wild derived strains due to their behavior.
+Embryonic samples were collected when pregnancy noted in females,
+embryos were staged based on (X table markers).
 
-Meiocyte spreads were made following [@peters_1997].
-For testes the tunica was removed and whole testes was incubated in 3ml of hypotonic solution for 45min.
-For ovaries after dissection, the pair of ovaries were decapsulated in cold PBS and both were incubated in 300ul of hypotonic solution for 45 min.
+Meiocyte spreads were made following (Peters et al. 1997). For testes
+the tunica was removed and whole testes was incubated in 3ml of
+hypotonic solution for 45min. For ovaries after dissection, the pair of
+ovaries were decapsulated in cold PBS and both were incubated in 300ul
+of hypotonic solution for 45 min.
 
-After incubation, gonads were transferred to sucrose solution -for mastication to make a cell slurry which was transferred to 80ul of 2% PFA solution.
-Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The next morning, slides were treated with a photoflow (Kodak, cite) wash.
+After incubation, gonads were transferred to sucrose solution -for
+mastication to make a cell slurry which was transferred to 80ul of 2\%
+PFA solution. Cells were fixed in this solution and dried in a humid
+chamber at room temperature overnight. The next morning, slides were
+treated with a photoflow (Kodak, cite) wash.
 
@@
 
-The same staining protocol was applied to spermatocyte and oocyte spreads.
-The staining protocol was based on X with some adaptations, and same as previously described in Peterson et al 2019.
+The same staining protocol was applied to spermatocyte and oocyte
+spreads. The staining protocol was based on X with some adaptations, and
+same as previously described in Peterson et al 2019.
 
-Explain list recipes for DMC1 since these haven't been listed before
+\section{Image Processing}\label{image-processing}
 
-# Image Processing
+Images of cells were capture using a Zeiss Axioplan 2 microscope with
+AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
 
-Images of cells were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
+Preprocessing, including cropping, noise reduction, and histogram
+adjustments, was performed using Photoshop (v13.0).
 
-Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
+For MLH1 pachytene characterization, Cells with a full karyotype (19
+acrocentric bivalents and XY for spermatocytes or 20 acrocentric
+bivalents for oocytes), distinct foci, and intact bivalents were
+included for quantification. Reprocessing in Photoshop (cite). Image
+file names were anonymized before manual scoring of MLH1 or DMC1 foci.
 
-For MLH1 pachytene characterization, Cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
-Reprocessing in Photoshop (cite). Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
+Hand measures data set generated by using ImageJ/Fiji (v1.52)
+(Schindelin et al. 2012).
 
-Repeatability was assessed by X (comparing mulitple counted images)
+total Sc was measured (Wang et al. 2019)
 
-Hand measures data set generated by using ImageJ/Fiji (v1.52) [@fiji].
+Bivalent level measure were done (Peterson, Miller, and Payseur 2019)
 
-total Sc was measured [@wang_2019_sc]
+\section{Statisical Analysis}\label{statisical-analysis}
 
-Bivalent level measure were done [@peterson_2019]
+Statistical analysis was performed in R (v3.5.2)(Team 2015).
 
-# Statisical Analysis
+The distributions of CO counts per cell was assessed for normailty
+(supplemental figure). Mouse mean MLH1 count was used to get around the
+bad discreet nature of count data. (normality was confirmed with X
+distribution, supplemental figure).
 
-Statistical analysis was performed in R (v3.5.2)[@Rstudio].
+framework for analyzing the MLH1 patterns 1. write equation
 
-The distributions of CO counts per cell was assessed for normailty (supplemental figure).
-Mouse mean MLH1 count was used to get around the bad discreet nature of count data.
-(normality was confirmed with X distribution, supplemental figure).
+\$X \sim mouse average MLH1 \textasciitilde{} sex + subspecies +
+random(strain) + \xi \$
 
-framework for analyzing the MLH1 patterns
-1. write equation
+mixed model, observations of mouse means of mlh1, all data with both
+male and female observations
 
-$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
- can't get the equation in the right format. I think this breaks the pdf kniting.
+subspecies, sex and subsp*sex are fixed effects strain as random effect
+(sample of a genetic background from subspecies) with domesticus male
+set as the default
 
-mixed model, observations of mouse means of mlh1, all data with both male and female observations
+strain and sex interaction effect was included to ask / test if the
+effect of sex varies by strain(genetic background)
 
-subspecies, sex and subsp*sex are fixed effects
-strain as random effect (sample of a genetic background from subspecies)
-with domesticus male set as the default
+this model allows us to estimate and test the fixed effects (sex and
+subsp)
 
-strain and sex interaction effect was included to ask / test if the effect of sex varies by strain(genetic background)
+the random effect of strain in this model can be interpreted as the
+average standard deviation due to the random strain effect and sex
+strain,
 
-this model allows us to estimate and test the fixed effects (sex and subsp)
+From the mixed model the estimates of heterochiasmy (sexual dimorphism)
+are a combination of subspecies and strain specific sex effects.
 
-the random effect of strain in this model can be interpreted as the average standard deviation due to the random strain effect
-and sex  strain,
-
-From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.
-
-We can also compare the size of the sex effects between (subspecies) and within (strain) mouse groups.
-(ie is the male or female strain effect larger?
-is there more variance due to random strain effect in male or females?)
+We can also compare the size of the sex effects between (subspecies) and
+within (strain) mouse groups. (ie is the male or female strain effect
+larger? is there more variance due to random strain effect in male or
+females?)
 
@@
 
--specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
-linear models lm() from X package.
+-specific statistical tests; (means, non-parametric kurts wallis
+t-test?) (correlation coefficient (function)) linear models lm() from X
+package.
 
--specific R packages
--non-parametric  (tests and measures for dealing with count data)
--hypothesis tested (size, DSB number)
+-specific R packages -non-parametric (tests and measures for dealing
+with count data) -hypothesis tested (size, DSB number)
 
-Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mix model frame work, using X package (cite).
-With X formula Co count ~ blah + blah, with random strain specific sex effect.
+Comparison/analysis of variation in MLH1 counts within and between
+groups was done within a general linear/ mix model frame work, using X
+package (cite). With X formula Co count \textasciitilde{} blah + blah,
+with random strain specific sex effect.
 
-(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)
+(we used this frame work as proxy for `polymorphism' and `divergence' in
+MLH1 counts across house mouse and closely related species.)
 
-# References
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-peters_1997}{}
+Peters, Antoine HFM, Annemieke W. Plug, Martine J. van Vugt, and Peter
+De Boer. 1997. ``SHORT COMMUNICATIONS A Drying-down Technique for the
+Spreading of Mammalian Meiocytes from the Male and Female Germline.''
+\emph{Chromosome Research} 5 (1): 66--68.
+
+\hypertarget{ref-peterson_2019}{}
+Peterson, April L, Nathan D Miller, and Bret A Payseur. 2019.
+``Conservation of the Genome-Wide Recombination Rate in White-Footed
+Mice.'' \emph{Heredity}. Nature Publishing Group, 1--16.
+
+\hypertarget{ref-fiji}{}
+Schindelin, Johannes, Ignacio Arganda-Carreras, Erwin Frise, Verena
+Kaynig, Mark Longair, Tobias Pietzsch, Stephan Preibisch, Curtis Rueden,
+Stephan Saalfeld, and Benjamin Schmid. 2012. ``Fiji: An Open-Source
+Platform for Biological-Image Analysis.'' \emph{Nature Methods} 9 (7):
+676.
+
+\hypertarget{ref-Rstudio}{}
+Team, RStudio. 2015. ``RStudio: Integrated Development Environment for
+R.'' Boston, MA. \url{http://www.rstudio.com}.
+
+\hypertarget{ref-wang_2019_sc}{}
+Wang, RJ, BL Dumont, P Jing, and BA Payseur. 2019. ``A First Genetic
+Portrait of Synaptonemal Complex Variation.'' \emph{PLoS Genetics} 15
+(8): e1008337--e1008337.
+
+\end{document}

### 2019-09-26 doc/MouseLevelPlots.Rmd
updating list of mice to finish
ran mouse level plots with new batch16 data

@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")#batch15 isn't showing up
+load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
+#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")#batch15 isn't showing up
 #load(file="data/MetaData.RData")#this Rdata was replacing the most current batch.. agg!
@@
 #this is where you remove the non-data data!
-df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
+df_list <- df_list[ -c(12,18,24,30) ]
+#there is PERC male
 
-#7 PERC, 12 MOLF female, 
+#12 MOLF female, 18 CZECH female, 24 HMI female?, 
 #

### 2019-09-26 doc/Main_Report.rmd
re-ran main report with new batch 16 data

@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")
+load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
+#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
 
@@
 
-```{r quant.repeatability, echo=FALSE, include=TRUE}
+In the original_DF and MLH1 DF's there are ~400 and X respectively of cells (filenames) that are duplicated, or cells that have been counted in more than 1 quant batch
+
+Datasets / sources of repeat measures
+-chrom proportions
+-multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
+-old and new (batch0)
+
+The main take away is that higher quality scored cells have lower error in MLH1 counts.
+
+> Does the amount of error differ across strains or sexes? (female might have a higher error rate)
+
+> What sort of adjustments or tests should be made to results or analyses given the known error rate?
+
+```{r quant.repeatability_rep1, echo=FALSE, include=TRUE}
+#set up data for combining replicate counts
+
+#1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now
+
+#isolate the duplicates (fileName) from main MLH1 DFs
+n_occur <- data.frame(table(original_DF$fileName))
+n_occur[n_occur$Freq > 1,]
+original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,
+
+#are there any batch, fileName duplicates?
+original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")
+anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch ... really 2?
+
+n_occur <- data.frame(table(original_DF_duplicates$unique.key))
+n_occur[n_occur$Freq > 1,]
+double.trouble <- original_DF_duplicates[original_DF_duplicates$unique.key %in% n_occur$Var1[n_occur$Freq > 1],]
+#2 fileNames from batch 0 appear 2x, remove them from DF
+
+original_DF_duplicates <- original_DF_duplicates[ ! original_DF_duplicates$fileName %in% double.trouble$fileName, ]#removes 4 rows
+
+#seperated duplicates into categories out in Excel (write to /results, moved final file to /data)
+#This data frame was made ... with Jered's sugg
+#write the duplicate file to csv/excel
+#sort by file name
+#need to remove '?' -- deal with the ? and X's
+#write.table(original_DF_duplicates, "~./MLH1repo/results/orginal_DF_duplicates.csv", sep=",", row.names = FALSE)
+
+original_DF_duplicates.notes = read.csv("~./MLH1repo/data/orginal_DF_duplicates.csv", header = TRUE)
+#copies of fileNames can be distinguished with the 'separate.DF' column
+
+#remove '?' that I stupidly left in
+original_DF_duplicates.notes$nMLH1.foci = gsub("\\?", "", original_DF_duplicates.notes$nMLH1.foci)
+original_DF_duplicates.notes$quality = gsub("\\?", "", original_DF_duplicates.notes$quality)
+#these lines change col to chr
+
+original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
+#now there should be no more '?' going forward
+
+#original_DF_duplicates.notes$quality <- as.factor(original_DF_duplicates.notes$quality)
+
+#make lists of rows with "" or X's so I can remove all fileNames for those
+#these lists just have na
+quality.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$quality) | original_DF_duplicates.notes$quality=="")] #16
+
+MLH1.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci=="") ]#22
+
+X.list <- original_DF_duplicates.notes$fileName[grepl("X", original_DF_duplicates.notes$X)]#40
+
+#make DF which only has 
+matches.only <- original_DF_duplicates.notes[ ! original_DF_duplicates.notes$fileName %in% quality.na.list, ]
+matches.only <- matches.only[ ! matches.only$fileName %in% MLH1.na.list, ]
+matches.only <- matches.only[ ! matches.only$fileName %in% X.list, ]
+
+#str(matches.only)
+#clean up match list (chr to num)
+matches.only$nMLH1.foci <- as.numeric(matches.only$nMLH1.foci)
+matches.only$quality <- as.numeric(matches.only$quality)
+
+levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt
+
+#check that the 1, and 2 are equal number
+length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178
+
+#i want to measure correleation of MLH1 across bins of quality (1,2,3,4,5)
+#make new df with filename and MLH1.1 and mlh1.2
+
+dups.here <- data.frame(  FN = unique(matches.only$Original.Name), MLH1.1 = matches.only$nMLH1.foci[matches.only$separate.DF == 1], MLH1.2 = matches.only$nMLH1.foci[matches.only$separate.DF == 2],
+                        qual.1 = matches.only$quality[matches.only$separate.DF == 1] , 
+                        qual.2 = matches.only$quality[matches.only$separate.DF == 2]
+                          ) #OMG this worked!!!
+dups.here$diff <- abs(dups.here$MLH1.1 - dups.here$MLH1.2) #the aveage difference is pretty small, but some have very high difference
+#the high differences could potentially be mix ups
+
+repeatability.table <- ddply(.data=dups.here, 
+                 .(qual.1, qual.2),
+                 summarize, 
+                 ncells = length(unique(FN)),
+                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
+)
+#this table shows how many cells went changed into different qual score bins
+#easiest patter, 1-1, 1-2 (high correlation in MLH1)
+#5-5 lowest, (5-3 also low)
+#4-5 has high corr? (direction shouldn't matter)
+#4-2 has negative corr?
+
+#of the different quality permutation bins, 37% don't change quality bin, 44% of cells change by 1 (either direction)
+
+no.change = c(14,13,17,13,10)#67
+change.1 = c(10,7,8,20,9,14,4,7)#79
+
+repeatability.table.qual1 <- ddply(.data=dups.here, 
+                 .(qual.1),
+                 summarize, 
+                 ncells = length(unique(FN)),
+                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2),
+                 dif = mean(MLH1.1 - MLH1.2)
+)
+
+repeatability.table.qual2 <- ddply(.data=dups.here, 
+                 .(qual.2),
+                 summarize, 
+                 ncells = length(unique(FN)),
+                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
+)
+#These tables show that qual1 cells have the highest corrleation of MLH1
+dups.here.plain_cor_qual <- cor(dups.here$qual.1, dups.here$qual.2)
+
+#i don't know how to compare quality in more
+
+#nMLHJ1, quality are factors
+original_DF_duplicates.notes$quality2 = gsub("\\?", "", original_DF_duplicates.notes$quality)
+
+original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
+
+#factor > chr > number  (this col is now chr is can go to num)
+#original_DF_duplicates.notes$MLH1.foci2 <- as.numeric(original_DF_duplicates.notes$MLH1.foci2)
+#original_DF_duplicates.notes$quality2 <- as.numeric(original_DF_duplicates.notes$quality2)
+
+#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
+#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
+#rename this
+#plain_cor_MLH1 <- cor(one$MLH1.foci2,two$MLH1.foci2,use = "complete.obs")
+#.857
+#plain_cor_qual <- cor(one$quality2,two$quality2,use = "complete.obs")
+#.63
+
+#What's the relationship of MLH1 cor with quality?
+#what the rpeat of quality score?
+#what are the correlations across quality score bins?
+
+#issues that complicate the plain calculation
+#some duplicates were 'x'ed and some weren't -- this causes different numbers
+#connected with above, some duplicates have mlh1 scores, other don't -- this kinda messes up correlation
+
+original_DF_duplicates.notes <- original_DF_duplicates.notes[ !grepl("X", original_DF_duplicates.notes$X) , ]
+original_DF_duplicates.notes <- original_DF_duplicates.notes[!(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci==""), ]
+#this causes unequal lengths
+
+original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
+original_DF_duplicates.notes$nMLH1.foci <- as.character(original_DF_duplicates.notes$nMLH1.foci)
+
+original_DF_duplicates.notes$nMLH1.foci <- as.numeric(original_DF_duplicates.notes$nMLH1.foci)
+
+#correlationsnMLH1 
+#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
+#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
+#plain_cor <- cor(one$nMLH1.foci,two$nMLH1.foci)
+#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
+
+#check that all batches are distinguishable across duplicate fileNames
+#make unique key fileName_batch -- then check for duplicate of that 
+#original_DF$Batch <- factor(original_DF$Batch, ordered = TRUE, levels = c("Batch0", "Batch1", "Batch2", "Batch3", "Batch4","Batch5","Batch6",  "Batch7", "Batch8","Batch9","Batch10", "Batch11","Batch12", "Batch13", "Batch14","Batch15"))
+#after doing this, original_DF$Batch[2000] > original_DF$Batch[10] will be evaluated!
+
+```
+
+```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
+#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
 #load, clean up and mat one of the master.MLH1count sheets
@@
 
-uutesting <- add_fullFileName(old.master.MLH1)
+old.master.MLH1 <- add_fullFileName(old.master.MLH1)
+old.master.MLH1$fileName <- old.master.MLH1$file.name_full
 
@@
 
+#need to test for duplicates before mergeing?
+n_occur <- data.frame(table(old.master.MLH1$file.name_full))
+n_occur[n_occur$Freq > 1,]
+ll <- old.master.MLH1[old.master.MLH1$file.name_full %in% n_occur$Var1[n_occur$Freq > 1],]
+old.master.MLH1 <- old.master.MLH1[ ! old.master.MLH1$fileName %in% ll$fileName, ]
+
+n_occur <- data.frame(table(original_DF$fileName))
+n_occur[n_occur$Freq > 1,]
+original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]
+table(original_DF_duplicates$mouse)#17 mice
+
+old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
+
+#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
+
+#there are duplicates in original_DF and MLH1
+
+uutesting2 <- merge(old.master.MLH1, original_DF, by.x = fileName, by.y = Original.Name)
 #match old.full.filname with MLH1$fileName
-
+```
+
+```{r repeatability_rep3, echo=FALSE, include=FALSE}
 chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
@@
 #the true false might have to be changed first
-
-chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)
+#chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)
 
 #regression of quality and MLH1.dif
-lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
+#lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
 #summary(lmreg)
@@
 #but quality is bad predictor, only 5 levels)
-lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
+#lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
 #summary(lmreg2)

### 2019-09-26 src/SetupData.Rmd
incorporating notes and comments for the datasetup Rmd

---
title: "Setting up Data"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
setwd("~./MLH1repo/")
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#forgot why I'm reading in this metadata file
meta.data = read.csv("data/clean.Meta.Data.txt",sep = "\t", header=TRUE)
MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE)

knitr::opts_chunk$set(echo = FALSE)
```

# First Steps

```{r MLH1.DF.setup, warning=FALSE}

#todo :
#set the order to batches
#2 fileNames from batch 0 appear 2x, (1 is missing random name, the other has the same random name)

original_DF = MLH1_data
original_DF$fileName <- (original_DF$Original.Name)
original_DF <- add_mouse(original_DF)

#check for duplicated data, make duplicated data DF for repeatability
#anyDuplicated(original_DF$fileName)#219
#orig_DF <- original_DF[duplicated(original_DF$Original.Name),]
#these just produce 1 set of the duplicated

#isolate the duplicates (fileName) from main MLH1 DFs
n_occur <- data.frame(table(original_DF$fileName))
#n_occur[n_occur$Freq > 1,] displays the dups
original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates, 
#use the above DF for quantifying repeatability in the Main_report.Rmd
#duplicates are cells counted twice

#are there duplicated filenames+batch
original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")
#anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch ... really 2?

#remove duplicates from original as well
original_DF <- original_DF %>% distinct(Original.Name, .keep_all = TRUE)#5216 to 4983

#Clean up main DF, MLH1_data
#remove duplicated data from MLH1 -- remove 1 or the two rows
#https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/
#use distinct() in dlpyr()
MLH1_data <- MLH1_data %>% distinct(Original.Name, .keep_all = TRUE)#5216 to 4983

MLH1_data$Original.Name <- as.character(MLH1_data$Original.Name)
MLH1_data$fileName <- (MLH1_data$Original.Name)

#these introduce NAs
MLH1_data$n <- as.character(MLH1_data$n)
MLH1_data$nMLH1.foci<- as.character(MLH1_data$nMLH1.foci)
MLH1_data$quality<- as.character(MLH1_data$quality)

MLH1_data$n <- as.numeric(MLH1_data$n)
MLH1_data$nMLH1.foci<- as.numeric(MLH1_data$nMLH1.foci)
MLH1_data$quality<- as.numeric(MLH1_data$quality)

#remove X's and NAs
MLH1_data <- MLH1_data[ !grepl("X", MLH1_data$X) , ]
MLH1_data <- MLH1_data[!(is.na(MLH1_data$nMLH1.foci) | MLH1_data$nMLH1.foci==""), ]
#this should also have the empty column for removing X
MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "X",]
MLH1_data <- MLH1_data[MLH1_data$nMLH1.foci != "x",]

MLH1_data <- add_mouse(MLH1_data)
MLH1_data <- add_category(MLH1_data)
MLH1_data <- add_strain(MLH1_data)
MLH1_data <- add_sex(MLH1_data)
MLH1_data <- add_subsp(MLH1_data)
#add species
MLH1_data <- add_species(MLH1_data)

#add a column with male adjusted MLH1 values (+1 to all males)
MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)

#reorder dataframe
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name) )

#order the batches in MLH1 and original

```

 Settting up main data file with explanation and examination of data.  Make the raw original DF and main working DF with bad cells exlcuede.  (inline code add the numbers)

There are `r length(original_DF_duplicates$fileName)` duplicated file names. These data will be used to quantify repeatability, and remove from the final DFs.
Using the distinct() function for the MLH1_data, the first duplicated filename is kept and the second removed.

```{r kable.table}
#table comparing number of images in quantification batches
#ToDo:  correct the case of all batches, better names
#set an order to batches?
original_batches <- as.data.frame(table(original_DF$Batch) ) #these look bad
colnames(original_batches) <- c("BatchNum", "orginal")

final_batches <- as.data.frame(table(MLH1_data$Batch) ) #these look bad
colnames(final_batches) <- c("BatchNum", "main")

batch.stats <- merge(original_batches,final_batches, by="BatchNum")## mm is the table of batch stats
```

The original dataset contains the quantification batches

`r kable(final_batches, caption = "This is the table caption") ) %>% kable_styling()`

# Mouse Data

```{r mouse.table, echo=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#Make a mouse level table
AP_mouse_table <- ddply(MLH1_data, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)

AP_mouse_table <- add_strain(AP_mouse_table)
AP_mouse_table <- add_subsp(AP_mouse_table)
AP_mouse_table <- add_sex(AP_mouse_table)
AP_mouse_table <- add_category(AP_mouse_table)

#make sure they are all factors
AP_mouse_table$mouse <- as.factor(AP_mouse_table$mouse)
AP_mouse_table$sex <- as.factor(AP_mouse_table$sex)
AP_mouse_table$strain <- as.factor(AP_mouse_table$strain)
AP_mouse_table$subsp <-  as.factor(AP_mouse_table$subsp)

#add DOB to this table?
mice.frm.meta <- AP_mouse_table[(AP_mouse_table$mouse %in% meta.data$mouse),]

#oo isn't including MLH1 means

batch.mouse.count <- table(original_DF$mouse,original_DF$Batch) #mouse count by batch
batch.mouse.count.DF <- as.data.frame(batch.mouse.count)#this has a dif str, 3 cols -- like it was melted
batch.mouse.count.DF <- batch.mouse.count.DF[!(is.na(batch.mouse.count.DF$Freq)|batch.mouse.count.DF$Freq==0),]

#merge with DOB
mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = "mouse", by.x = "Var1", all = TRUE)
AP_mouse_table_w.Ages <- merge(AP_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)

```

## Number of mice per category

X number of mice are in the main DF (discribe the results for missing mice)

## Mouse Ages

Testes were collected from mice at various ages. For some poor breeding mice, I couldn't collect males at younger ages.

Is there a detectable age effect?

```{r Age.Hists, echo=FALSE, warning=FALSE}

male_mouse_table_w.Ages <- mouse.table_w.Ages[(mouse.table_w.Ages$sex == "male") &(mouse.table_w.Ages$category != "other"),]

all.ages_plot <- ggplot(male_mouse_table_w.Ages, aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
all.ages_plot

male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[male_mouse_table_w.Ages$age.weeks < 20,]
#droplevels on DF above

good.ages_plot <- ggplot(male_mouse_table_w.good.ages, aes(age.weeks))+geom_histogram(binwidth = 1)+facet_wrap(~category)+ggtitle("Good Male Age Distributions")

#male_mouse_table_w.good.ages

```

# Saving Data file

Now that the dataframes have been process and formated, they will be saved as RData. the main files are:
  1. MLH1_data and original_DF
  2. AP_mouse_table and AP_mouse_table_w.Ages
  3. meta.data

work on cleaning up all other df with bad names

```{r save.RData, echo=FALSE}

save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_9.26.19.RData")

#  OutPut: big large MLH1_data (AP's) df, big DF of BD with just the mice I want.
#  MLH1_data_table, means and variance of AP and BD MLH1 values. made from seperate tables from AP and BD data.
#  make sure decimal places are consistant

```

# Deciding on Mice for Final Dataset

Use the above information to make the final dataset for analysis

-age (for males)
-enough cells
-both male and female

### 2019-10-08 src/SetupData.Rmd
removed pointless table
removed Peromyscus notes from meta file

@@
 meta.data = read.csv("data/clean.Meta.Data.txt",sep = "\t", header=TRUE)
+#remove all MAD (peromyscus) from meta.data
+meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]
+
 MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE)
@@
 #male_mouse_table_w.good.ages
-
 ```

### 2019-10-08 doc/Main_Report.rmd
removed Peromyscus notes from meta file
added standardized color fill to setup data

@@
   pdf_document: default
+bibliography: refs_Main_Report.bib
 ---
@@
 #create a real fileName col, add rev.tif, 
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
@@
 chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 #what are the target columns for this excel sheet?
@@
 
-Some papers have documented age effects for male genome-wide recombination rate (cite, one of the Cole papers).
+Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age].
 
@@
 ```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
 
-#make Mouse meta data work
+#make Mouse meta data work .. rename these
 MouseMetaData.wrkin <- meta.data[,c(1,4)]
 colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
-
 MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
-
 #do this for the mouse table instead of the MLH1 table
 MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
-
-#add euth date for calculating age
-
+#DOB is date
 MLH1.imprv <- add_euth_date(MLH1.imprv)
-#
+#good euth date..
+
+#make backup date col #add euth date for calculating age
 MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
-MLH1.imprv$euth.date <- as.Date(as.character(MLH1.imprv$euth.date2), format='%d%b%y')
-#this is currently broken
 MLH1.imprv <- add_age(MLH1.imprv)
+#age works
 
@@
 
-pp <- ggplot(MLH1.imprv.male, aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Age Distributions")
+pp <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
 pp
-
-#what is this meta.data.file?
-#meta.data.file2 <- add_strain(meta.data.file2)
-#meta.data.file2 <- add_sex(meta.data.file2)
-#meta.data.file2 <- add_category(meta.data.file2)
-#meta.data.file2 <- add_euth_date(meta.data.file2)
-#meta.data.file2 <- add_age(meta.data.file2)
-
-#qq <- ggplot(meta.data.file2[meta.data.file2$sex == "male",], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("recorded ages of dissected mice")
-#qq
 
@@
 ```{r age.effect.plot, warning=FALSE, echo=FALSE}
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
+#remove NAs
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]
 
@@
 male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
-#remove NAs
-
+#lots of NA rows... ugh .. remove them
+male.data <- male.data[!(is.na(male.data$Nmice) | male.data$Nmice==""), ]
+
+#other removed, no points,
 male.data <- male.data[male.data$Ncells > 10,]
 
-age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(19,34))+
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age and mean CO per mouse")+facet_wrap(~subsp.y)
+age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
+  #trying this
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
 age.effect.plot
+#no longer has 'other', #yay I added purtty standard colors
+#there is an NA
 #todo add line to strain.y

### 2019-10-08 src/SetupData.Rmd
added standardized color fill to setup data

@@
 ```{r mouse.table, echo=FALSE}
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
@@
 
-#oo isn't including MLH1 means
-
 batch.mouse.count <- table(original_DF$mouse,original_DF$Batch) #mouse count by batch
@@
 
+#subsp in metadata not fully correct
+meta.data <- add_strain(meta.data)
+meta.data <- add_subsp(meta.data)
+meta.data <- add_sex(meta.data)
+meta.data <- add_category(meta.data)
+
 #merge with DOB
@@
 AP_mouse_table_w.Ages <- merge(AP_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+#what are the differences between these two
 
@@
 
+```{r color.scheme}
+#color scheme
+#mouse colors
+
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blue
+
+                       'PWD'= 'red','MSM'= 'darkorange', 'MOLF'= 'orange3',#red mucs, orange Molossi
+                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
+                       'CZECH'= '#E69F00',
+
+                    'CAST'='green','HMI'= 'forestgreen', #green
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1',#gold?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
+
+```
+
 # Saving Data file

### 2019-10-08 doc/Main_Report.rmd
debugging and first results for maternal age effects

@@
 
-pp <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
-pp
+age.hist <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
+age.hist
 
@@
 ```{r female.age.effect, echo=FALSE}
+#start thinking of the maternal age effect
+
 #female age effect
@@
 ```{r mat.age.effect, echo=FALSE}
-#compile data from female (and male?) mice with maternal ages.
-
-#start making a general linear model...? for mouse averages
-
-```
+meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
+#445? -- probably extras at bottom not good
+
+#how many mice have maternal DOB
+meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
+#151
+
+#format this meta.data.file
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+meta.data.file.w.mat.DOB <- add_mouse(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_category(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_strain(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_sex(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_subsp(meta.data.file.w.mat.DOB)
+
+#check if date format's been applied before tryung the age functions
+str(meta.data.file.w.mat.DOB)
+
+#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
+#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
+meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
+meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")
+
+str(meta.data.file.w.mat.DOB)
+
+meta.data.file.w.mat.DOB <- add_euth_date(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_age(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_mat_age(meta.data.file.w.mat.DOB)
+
+#merge meta data with MLH1 
+see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
+#remove anything without mean_CO
+see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
+see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
+
+#visualize the data
+
+meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
+meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]
+
+#hist of data
+mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
+mat.age.hist.female
+
+mat.age.hist.male <- ggplot(meta.data.file.w.mat.DOB.male, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of males")
+mat.age.hist.male
+
+#remove mice that might have low quality
+see.whatcha.do.HQ <- see.whatcha.do[see.whatcha.do$Ncells > 14,]
+
+#age by mean_co
+mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
+  #trying this
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
+mat.age.effect.plot.female
+
+mat.age.effect.plot.male <- ggplot(see.whatcha.do.HQ[see.whatcha.do.HQ$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
+  #trying this
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")+facet_wrap(~subsp.y)
+
+#
+ALLmat.age.effect.plot.male <- ggplot(see.whatcha.do[see.whatcha.do$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
+  #trying this
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")
+
++facet_wrap(~subsp.y)
+
+#is there a correlation with proband age and maternal age?
+#there is a big lack of maternal age data...
+#some of the values reference the original breeding pair..
+#from notes on original BPs(males), assigning maternal ages
+
+```
+
+There seems to be a negative correlation with
 
@@
 
-3 different rooms used to house mice. This will be tested in X manner.
+3 different rooms used to house mice. Access softmouse to be fixed. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
 

### 2019-10-08 src/CommonFunc_MLH1repo.R

@@ -7,6 +7,7 @@
 #-check for multi-cell images
 
 
+
 #make fileName
 #paste together file name
 add_fullFileName <- function(dat){
@@ -44,6 +45,23 @@ add_age <- function(dat){
   }
   return(dframe)
 }
+#running this blanks out all cells in euth.date col
+
+#calq maternal age time of mouse birth
+add_mat_age <- function(dat){
+  dframe <- dat
+  count =1
+  for(i in dframe$mouse){
+    #after DOB and euth date are standardized #substract  difftime(mat.age.dob - DOB)
+    dframe$mat.age.days[count]<-difftime(dframe$DOB[count], dframe$maternal_DOB[count],units='days')
+    
+    #dframe$age.weeks[count] <- difftime(dframe$euth.date[count], dframe$DOB[count], units='weeks')
+    #gives age in days
+    count= count +1
+  }
+  return(dframe)
+}
+
 
 #add mouse column to MLH1 df
 #this can be mess up when the columns aren't right type (charaecter)
@@ -120,13 +138,15 @@ add_strain <- function(dat){
                              ifelse(grepl("_SPI_", dframe$mouse), "SPIC",
                          ifelse(grepl("_SPIC_", dframe$mouse), "SPIC",
                   ifelse(grepl("_CAROLI_", dframe$mouse), "CAROLI",
-                                 "other")))))))))))))))))))
+                         
+                 ifelse(grepl("F1_", dframe$mouse), "F1",         
+                                 "other"))))))))))))))))))))
   
   #the ordering factor below deletes all strain entries
   dframe$strain<- factor(dframe$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                                                   "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                   "CAST", "HMI",
-                                                  "SPRET", "SPIC", "CAROLI", "other") )
+                                                  "SPRET", "SPIC", "CAROLI", "F1", "other") )
   return(dframe)
 }
 
@@ -180,14 +200,17 @@ add_category <- function(oldframe){
                        ifelse(grepl("_SPIC_m", dframe$mouse), "SPIC male",
                       ifelse(grepl("_CAROLI_m", dframe$mouse), "CAROLI male",
                      ifelse(grepl("_CAROLI_f", dframe$mouse), "CAROLI female",
-                                                      "other")))))))))))))))))))))))))))))))))))
[…]

### 2019-10-08 src/setupMetaData.R

@@ -8,42 +8,92 @@ setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 
 #use updated metadata -- 
 #8.30.19
-meta.data.file = read.csv("~./MLH1repo/data/Mouse_MetaData_8.30.19.csv", header = TRUE)
+meta.data.file = read.csv("~./MLH1repo/data/DandM_8.30.19.csv", header = TRUE)
+#these are mostly clean ... some wrong things in the Maternal_DOB thing
+#all none age values removed from the csv file
+#had to reformat the dates, since there were inconsistancies in year digit length
+#had to clean up some cells
+
+meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
+#445? -- probably extras at bottom not good
+
+#how many mice have maternal DOB
+meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
+#151
+
+#format this meta.data.file
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+meta.data.file.w.mat.DOB <- add_mouse(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_category(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_strain(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_sex(meta.data.file.w.mat.DOB)
+meta.data.file.w.mat.DOB <- add_subsp(meta.data.file.w.mat.DOB)
+
+#check if date format's been applied before tryung the age functions
+str(meta.data.file.w.mat.DOB)
+
+#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
+#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
+meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
+meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")
+
+str(meta.data.file.w.mat.DOB)
+
[…]

### 2019-10-09 doc/MatMeth.Rmd
cleaned up lg in morning while mind fresh

@@
 
-Measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
+Hand measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
 
-total Sc was measured [@wang_2019_sc]
-
-Bivalent level measure were done [@peterson_2019]
+The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson_2019] respectively.
 
@@
 
-Statistical analysis was performed in R (v3.5.2)[@Rstudio].
+MLH1 counts per cell are good /decent proxies for genome wide recombination rates estimates (cite).
 
-The distributions of CO counts per cell was assessed for normailty (supplemental figure).
+The distributions of CO counts per cell was assessed for normality (supplemental figure).
 Mouse mean MLH1 count was used to get around the bad discreet nature of count data.
@@
 
-We choose to use a **mixed model framework** to interpert the patterns of variation. For interperting patterns of mean MLH1 count within and between house mouse subspecies.
+We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
 
-1. write equation
-
-$X \sim mouse average MLH1 ~ sex * subspecies + random(strain) + \xi $
-
-In this model
-mixed model, observations of mouse means of mlh1, all data with both male and female observations
-
-The main effects and the interaction of subspecies and sex were coded as fixed.
-
-Subspecies, sex and subsp*sex are fixed effects
-
-Strain was coded as a random effect, since in this experimental design we consider it as a sample of natural genetic variation for each subspecies.
-We also included a term to allow for strains to have a specific sex effect.
+The use of wild derived inbred strains enabled us to design an utilize a mixed model for our analytical framework.
+We built our model to predict the strain average MLH1 count per cell.
+The main effects and their interaction of subspecies and sex were coded as fixed.
+While, strain (or genetic background) was coded as a random effect to reflect a random sample of natural genetic variation from each subspecies.
+A strains to have a specific sex effect was also included.
 
 This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
-
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner)
-
+We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism).
+(in a sex specific manner).
 From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.
@@
 is there more variance due to random strain effect in male or females?)
+(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)
+I may have to do more research/reading on testing the random effects.
 
-I may have to do more research/reading on testing the random effects.
--specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
-linear models lm() from X package.
 -specific R packages
@@
 
-Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mix model frame work, using X package (cite).
-With X formula Co count ~ blah + blah, with random strain specific sex effect.
+**outline**
 
-(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)
+1. framework for analyzing the MLH1 evolutionary patterns across mus species complex. (what is the main question we were tackling when designing this framework?)
+
+2. write equation, describe terms (proxies for within and between polymorphism)
+logic / choices for fixed and random effects
+
+3. what are the 2 general / basic questions this framework can be applied to.
+
+$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
+
+4. how can DMC1 measures be added to the mixed model
+
+5. descriptions of mlh1 outside of framework (missing female data)
 

### 2019-10-10 doc/Main_Report.rmd
cleaning up result doc and starting mixed model analysis

@@
 
-## Missing Mice, finishing dataset
-
-*Which mice are remaining and at what stage?*
-
 This section shows the number of mice which haven't been stained or quantified yet.
 
-```{r missing.mice, echo=FALSE}
+```{r missing.mice, echo=FALSE, warning=FALSE}
 
@@
 
-```{r mat.age.effect, echo=FALSE}
+```{r mat.age.effect, echo=FALSE, eval=FALSE}
+#pause evaluation -- needs debugging
+meta.data.file = read.csv("~./MLH1repo/data/DandM_8.30.19.csv", header = TRUE)
+
 meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
@@
 see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
+#what did I do with this DF...?
 
@@
   geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")
-
 +facet_wrap(~subsp.y)
@@
 
-# Initial Results
+# Initial Results (change these colors)
 
@@
 #order MLH1_data by strain!!
+#colors_of_strains, colors ..
 
@@
  ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
-  facet_grid(subsp~ sex, scales="free_x")#scales="free_y"
+  facet_grid(subsp~ sex, scales="free_x")+scale_color_manual(values=colors_of_strains)
+
+#scales="free_y"
 #this is ok except for the set placement of mice on x axis
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00","red", "yellowgreen", "green", "purple","grey1","grey","green", "purple","grey1","grey","white", "blue","orange" )) +
-  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )
+        axis.ticks.x=element_blank())+
+  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )+scale_fill_manual(values=colors_of_strains)
 
@@
 
-PWD and MSM have a higher proportion of 2CO chrms!
-
-what about the variance across cells?
-
-## Models for manuscript
-
-Ideas for models of evolution of Heterochiasmy and current idea for how to test them.
-
-- HetC arises due to more precursors (DSBs)
-
-    - Compare DSB counts
-
-- Larger cell volume drives longer chrms axis (and maybe more DSB)
-
-    - Measure meiocyte diameter in males and females. Compare spermatocyte diameters in male biased and female biased strains. 
-    - Compare DSBs counts and average SC/axis lengths.
-
-- Difference in astral/MT organization
-
-    - ?
-
-- Something about interference and sister cohesin area
-
-    - Develope metric for sister cohesin area, and look at initial patterns.
+High musc males have half 1CO and 2CO bivalents.
+For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
 
@@
 
+The mixed model lets us sperate effects to see how they effect a mouse mean CO count. This seperation can also serve as proxies for divergence and polymorphism.
+
 Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
 
-*Does this hold up with the full data set?*
-*What's the best visualization for these questions?*
-
-Main focus is contrasting the level of difference within and between subspecies.
-
-I'll do this with a mixed model framework approach (and analyze with ANOVA, a method to partition the variance of subset.)
-
-```{r mixed.model, echo=FALSE}
+> Mouse average Models
+
+  variables vs char coding of coefficents(opposite direction of sex)?
+  select1 (remove old mice, remove mice with low ncells)
+  select2 make sure there are equal numbers of observations
+  select3 low Musc males
+
+the pooled cell models are breaking for some reason, (the sexes)
+-pooled cells (nMLH1)
+  all 
+  pooled.HQ (1-4)
+  pooled.HQ.2 (sample equal numbers of cells)
+
+```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE}
+library(lme4)
+library(nlme)
+# seperate the data, both sexes, only Mus
+# decide on Musc - Molossinus
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+#remove Castaneus (not females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp.x != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+#change molossinus to it's own subsp (grepl)
+
+DF.HetC.MixedModel2 <- DF.HetC.MixedModel
+
+#if(grepl("MSM", DF.HetC.MixedModel2$strain) ){ DF.HetC.MixedModel2$subsp <- "Mol" }
+
+DF.HetC.MixedModel2$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel2$strain), "Mol", 
+        ifelse(grepl("MOLF", DF.HetC.MixedModel2$strain), "Mol", DF.HetC.MixedModel2$subsp))
+
+DF.HetC.MixedModel2$subsp <- ifelse(grepl(1, DF.HetC.MixedModel2$subsp), "Dom", 
+        ifelse(grepl(2, DF.HetC.MixedModel2$subsp), "Musc", DF.HetC.MixedModel2$subsp))
+
+#subsp are charecters
+#change sex to charecter? (unordered sex)
+
+DF.HetC.MixedModel2$sex <- as.character(DF.HetC.MixedModel2$sex)
+
+model.ff2 <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel2, random=list(strain=pdDiag(~sex) ) )#adding mol
+
+#pull out values / other things
+summary(model.ff2)# indeces for this object: [1]-structure
+ranef(model.ff2)
+#after coding subsp to chr from factor, the co-efficients are
+
+#intercept should be the Dom female mean, looks right
+mean(DF.HetC.MixedModel2$mean_co[(DF.HetC.MixedModel2$sex == "female" &
+                                    DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5
+#male effect is -2
+#Musc and mol effec is almost -1
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#after initial analysis look into quality of these mice and ages
+
+#clean up the DF, remove extra cols
+
+colnames(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                                                            subsp.y,subsp.x,
+                                                            category.y, category.x,
+                                                            strain.x,strain.y, mouse.1  )  )
+str(DF.HetC.MixedModel)
+
+#remove ncells <10
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$age.weeks < 25,]
+
+#pooled.cells
+
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+```
+
+For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)
+
+subspecies levels: 2, Musc and Domesticus (maybe Molossinus soon)
+strain levels: Dom: 3, Musc: 4
+sex: 2 female male
+
+`r table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain)`
+
+```{r mixed.model,make, echo=FALSE, warning=FALSE}
 #build mixed model
@@
 # old code uses base plot
-
-```
+library(nlme)
+library(lme4)
+
+#L,Q,C  after the factors in the summary mean lme doesn't believe you used ordered factors
+
+#redo the orders just to be sure
+#sex, strain, subsp,
+
+DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = TRUE, levels =c("female", "male") )
+
+DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = TRUE, levels =c( "Dom", "Musc") )
+
+DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW",
+                                                  "PWD", "MSM", "SKIVE", "KAZ") )
+
+#dframe$strain<- factor(dframe$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
+#                                                  "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
+#                                                  "CAST", "HMI",
+#                                                  "SPRET", "SPIC", "CAROLI", "F1", "other") )
+
+DF.HetC.MixedModel$mean_co <- as.integer(DF.HetC.MixedModel$mean_co)
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
+
+#female Dom are the first levels:
+#Intercept, 25.767065 mean for Dom female -- NS
+#
+#make sure the default (ordered factors of strains is set)
+#check level order etc.
+#build initialize model
+model.ff <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.ff)# indeces for this object: [1]-structure
+ranef(model.ff)
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
+
+DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
+
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+```
+
+pooled cell data collection.
+
+`r table(DF.HetC.MixedModel.pooled$sex, DF.HetC.MixedModel.pooled$strain)`
+KAZ, LEWES has ~100 more for males, PWD has ~200 for females.
+
+Fixed effects
+
+look at the p-values for each of the effects as an indication if they are significant predictors.
+
+Subspecies
+
+Random effects
+
+try subsampling some of the
 

### 2019-10-10 doc/Main_Report.rmd
more updates to mixed models, learned my lesson for ordered factors

@@
 
-> *-Which strain has highest HetC?*
-
 ```{r HetC.plot, echo=FALSE, warning=FALSE}
@@
 
-## Explaination for the XX and XY differences
+### Explaination for the XX and XY differences
 
@@
 
-## Difference in proportion of chrm classes
+### Difference in proportion of chrm classes
 
@@
 
-## Mixed Model, compare within and between subspecies patterns
+### Mixed Model, compare within and between subspecies patterns
 
@@
 
-```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE}
+```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE}
 library(lme4)
@@
 
-#remove Castaneus (not females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp.x != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-
-#remove PERC, MOLF, CZECH,
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
 #change molossinus to it's own subsp (grepl)
-
 DF.HetC.MixedModel2 <- DF.HetC.MixedModel
@@
 #if(grepl("MSM", DF.HetC.MixedModel2$strain) ){ DF.HetC.MixedModel2$subsp <- "Mol" }
-
-DF.HetC.MixedModel2$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel2$strain), "Mol", 
-        ifelse(grepl("MOLF", DF.HetC.MixedModel2$strain), "Mol", DF.HetC.MixedModel2$subsp))
-
-DF.HetC.MixedModel2$subsp <- ifelse(grepl(1, DF.HetC.MixedModel2$subsp), "Dom", 
-        ifelse(grepl(2, DF.HetC.MixedModel2$subsp), "Musc", DF.HetC.MixedModel2$subsp))
 
@@
 #pull out values / other things
-summary(model.ff2)# indeces for this object: [1]-structure
-ranef(model.ff2)
+#summary(model.ff2)# indeces for this object: [1]-structure
+#ranef(model.ff2)
 #after coding subsp to chr from factor, the co-efficients are
@@
                                     DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5
-#male effect is -2
-#Musc and mol effec is almost -1
 
@@
 DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
 
@@
 
-colnames(DF.HetC.MixedModel)
+#colnames(DF.HetC.MixedModel)
 DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
@@
                                                             strain.x,strain.y, mouse.1  )  )
-str(DF.HetC.MixedModel)
-
-#remove ncells <10
-
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$age.weeks < 25,]
 
 #pooled.cells
-
 DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
-
 DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
@@
 
-`r table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain)`
-
-```{r mixed.model,make, echo=FALSE, warning=FALSE}
+`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
+
+# Building Mixed Models
+
+```{r mixed.model.make, echo=FALSE, warning=FALSE, eval=FALSE, message=FALSE}
 #build mixed model
@@
 
-#L,Q,C  after the factors in the summary mean lme doesn't believe you used ordered factors
-
+#L,Q,C  appear for ordered factors in the summary mean lme doesn't believe you used ordered factors
+#change all the DFs / variables to UNORDERED factors!
 #redo the orders just to be sure
@@
 
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = TRUE, levels =c("female", "male") )
-
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = TRUE, levels =c( "Dom", "Musc") )
-
-DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW",
+DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE, levels =c("female", "male") )
+
+DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE, levels =c( "Dom", "Musc") )
+
+DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW",
                                                   "PWD", "MSM", "SKIVE", "KAZ") )
-
-#dframe$strain<- factor(dframe$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
-#                                                  "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
-#                                                  "CAST", "HMI",
-#                                                  "SPRET", "SPIC", "CAROLI", "F1", "other") )
-
-DF.HetC.MixedModel$mean_co <- as.integer(DF.HetC.MixedModel$mean_co)
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
-
-#female Dom are the first levels:
-#Intercept, 25.767065 mean for Dom female -- NS
-#
-#make sure the default (ordered factors of strains is set)
-#check level order etc.
-#build initialize model
-model.ff <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.ff)# indeces for this object: [1]-structure
-ranef(model.ff)
 
@@
 ranef(model.pooled)
-```
-
-pooled cell data collection.
-
-`r table(DF.HetC.MixedModel.pooled$sex, DF.HetC.MixedModel.pooled$strain)`
+
+model.pooled$coefficients
+```
+
+List of mixed models with mean_co
+
+1. mean_co --DF.HetC.MixedModel
+2. mean_co.chr  variables coded as factors (for ggplots), how is this changed when they are coded as chr
+3. mean_co.fac -- non-ordered factors
+4. mean_co.HQ
+5. mean_co.HQ.mol
+6. Sub-sampling for equal numbers
+
+```{r mixed.models.setup, echo=FALSE}
+#make are the DFs for the models
+
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                                                            subsp.y,subsp.x,
+                                                            category.y, category.x,
+                                                            strain.x,strain.y, mouse.1  )  )
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
+
+DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE )
+DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE )
+DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE )
+
+#main df with variables coded as chr
+DF.HetC.MixedModel.chr <- DF.HetC.MixedModel
+DF.HetC.MixedModel.chr$subsp <- as.character(DF.HetC.MixedModel$subsp)
+DF.HetC.MixedModel.chr$sex <- as.character(DF.HetC.MixedModel$sex)
+DF.HetC.MixedModel.chr$strain <- as.character(DF.HetC.MixedModel$strain)
+
+#mean_co.HQ
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mean_co.HQ.mol
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
+        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
+        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
+
+#mean_co.chr.HQ.mol
+DF.HetC.MixedModel.chr.HQ.mol <- DF.HetC.MixedModel.HQ.mol
+
+DF.HetC.MixedModel.chr.HQ.mol$subsp <- as.character(DF.HetC.MixedModel.chr.HQ.mol$subsp)
+DF.HetC.MixedModel.chr.HQ.mol$sex <- as.character(DF.HetC.MixedModel.chr.HQ.mol$sex)
+DF.HetC.MixedModel.chr.HQ.mol$strain <- as.character(DF.HetC.MixedModel.chr.HQ.mol$strain)
+
+```
+
+```{r MixedModels.stff, echo=FALSE}
+#main mean.CO
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+#mean.co.chr
+HetC.MixedModel.mean_co.chr <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr, random=list(strain=pdDiag(~sex) ) )
+
+#mean.co.HQ
+HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+#mean.CO.HQ.mol
+HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
+
+#mean.HQ.mol.chr
+HetC.MixedModel.chr.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr.HQ.mol, random=list(strain=pdDiag(~sex) ) )
+
+```
+
+The intercept level is Dom female. These are the coefficients
+
+`r kable(HetC.MixedModel.mean_co$coefficients$fixed, col.names = "coefficients")`
+
+`r kable(ranef(HetC.MixedModel.mean_co) )`
+
+changed all to unordered factors.
+
+all the variables are coded as ordered factors.
+
+When the variables are kept factors, and unordered the names appear! yay!
+
+`r kable(HetC.MixedModel.mean_co.chr$coefficients$fixed, caption="Chr coded Model", col.names = "coefficients")`
+
+`r kable(ranef(HetC.MixedModel.mean_co.chr) )`
+
+The variable names now appear. The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.
+
+`r kable(HetC.MixedModel.HQ$coefficients$fixed, caption="mean_co HQ model", col.names = "coefficients")`
+`r kable(ranef(HetC.MixedModel.HQ) )`
+
+The factor coded DF the more high quality cells.
+
+`r kable(HetC.MixedModel.HQ.mol$coefficients$fixed, caption="Main model with molossinus", col.names = "coefficient")`
+`r kable(ranef(HetC.MixedModel.HQ.mol) )`
+With adding molossinus, the subsp names are listed.
+
+`r kable(HetC.MixedModel.chr.HQ.mol$coefficients$fixed, caption="Main model, chr coded with molossinus", col.names = "coefficients")`
+`r kable(ranef(HetC.MixedModel.chr.HQ.mol) )`
+
+When the variables/levels are coded as chr, the interaction effects get larger and the other subspecies get smaller. The base sex effect is also larger. The subsp effect is relative to Dom female,
+
+I don't think the subspecies and strains should be order, (maybe factors still makes sense.)
+
+This table doesn't display well
+
+`r table(DF.HetC.MixedModel.pooled$sex,DF.HetC.MixedModel.pooled$strain)`
+
 KAZ, LEWES has ~100 more for males, PWD has ~200 for females.

### 2019-10-11 doc/Intro.Rmd
small updates for citations-bibliography

@@
 
+# Background
+
+## natural variation in RR
+
+empiracal measures -- important / useful for understanding how rec evolves with narutal selection
+
+sources from Ritz Noor,
+Samuk et al 2019, (selection,)
+(barley in north africa),  Dreissig et al 2019
+
+humans (not very natural population)
+
+## empiracal measures for hetC / theories for evo
+
+Lenormand Duthiel,
+Shapley / stapley et al
+Sardell Kirkpatrick  
+Gruhn + hassold
+
+Hoffman papers -- strange inverted meiosis
+Uniparental disomy
+-- evidence that RR-constraint connection may be not straight forward.
+
+### lack of female data / sig of female data
+
+human annuploidy / fertility
+
 # Bibliography

### 2019-11-11 src/SetupData.Rmd
re-ran set up data with updataed full MLH1 sheet

--- 2019-10-08 src/SetupData.Rmd
@@
 ```{r mouse.table, echo=FALSE}
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
 #Make a mouse level table
-AP_mouse_table <- ddply(MLH1_data, c("mouse"), summarise,
+AP_mouse_table.HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
 
-AP_mouse_table <- add_strain(AP_mouse_table)
-AP_mouse_table <- add_subsp(AP_mouse_table)
-AP_mouse_table <- add_sex(AP_mouse_table)
-AP_mouse_table <- add_category(AP_mouse_table)
+AP_mouse_table <- add_strain(AP_mouse_table.HQ)
+AP_mouse_table.HQ <- add_subsp(AP_mouse_table.HQ)
+AP_mouse_table.HQ <- add_sex(AP_mouse_table.HQ)
+AP_mouse_table.HQ <- add_category(AP_mouse_table.HQ)
 
 #make sure they are all factors
-AP_mouse_table$mouse <- as.factor(AP_mouse_table$mouse)
-AP_mouse_table$sex <- as.factor(AP_mouse_table$sex)
-AP_mouse_table$strain <- as.factor(AP_mouse_table$strain)
-AP_mouse_table$subsp <-  as.factor(AP_mouse_table$subsp)
+AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
+AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
+AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
+AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)
 
 #add DOB to this table?
-mice.frm.meta <- AP_mouse_table[(AP_mouse_table$mouse %in% meta.data$mouse),]
-
-#oo isn't including MLH1 means
+mice.frm.meta <- AP_mouse_table.HQ[(AP_mouse_table.HQ$mouse %in% meta.data$mouse),]
 
@@
 
+#subsp in metadata not fully correct
+meta.data <- add_strain(meta.data)
+meta.data <- add_subsp(meta.data)
+meta.data <- add_sex(meta.data)
+meta.data <- add_category(meta.data)
+
 #merge with DOB
 mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = "mouse", by.x = "Var1", all = TRUE)
-AP_mouse_table_w.Ages <- merge(AP_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+AP_mouse_table_w.Ages <- merge(AP_mouse_table.HQ, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+#what are the differences between these two
 
@@
 
-male_mouse_table_w.Ages <- mouse.table_w.Ages[(mouse.table_w.Ages$sex == "male") &(mouse.table_w.Ages$category != "other"),]
+male_mouse_table_w.Ages <- AP_mouse_table.HQ[(AP_mouse_table.HQ$sex == "male") &(mouse.table_w.Ages$category != "other"),]
 
-all.ages_plot <- ggplot(male_mouse_table_w.Ages, aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
+all.ages_plot <- ggplot(AP_mouse_table.HQ, aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
 all.ages_plot
@@
 #male_mouse_table_w.good.ages
+```
+
+```{r color.scheme}
+#color scheme, #mouse colors
+#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blues
+
+                       'PWD'= 'red2',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
+                       'CZECH'= 'indianred4',
+
+                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
+
+                    'CAST'='green','HMI'= 'forestgreen', #greens
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
+
 ```
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_9.26.19.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_11.11.19.RData")
 

### 2019-11-12 src/SetupData.Rmd
added batch17 female counts

@@
 
-AP_mouse_table <- add_strain(AP_mouse_table.HQ)
+AP_mouse_table.HQ <- add_strain(AP_mouse_table.HQ)
 AP_mouse_table.HQ <- add_subsp(AP_mouse_table.HQ)
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_11.11.19.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_11.12.19.RData")
 

### 2019-11-12 doc/Main_Report_Results.rmd
updated/made results Rmd and report thing

--- 2019-10-08 doc/Main_Report.rmd
@@
 library(dplyr)
+library(raster)
+library(lme4)
+library(nlme)
+library(RLRsim)
+
 setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
+
+load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
+#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
+#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
 #load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
-
 #load(file="data/MetaData.RData") #this was replacing most updated data file
@@
 
-# Research Description
-
-Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice and related rodents.
-Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
-
-## Discription of quantification process
-
-Cell images are quantified in batches after file names are anonymized.
-The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified.
-A quality score is given; 1 to 5 (best to worst).
-All cell image file names are anonymized and moved into batches of a few hundred images.
-
-```{r organizing.data, echo=FALSE}
-#set the order of another column, based on another variable (so that when)
+```{r organizing.data, echo=FALSE }
+#set the order of another column, based on another variable. consistant order for ploting
 
@@
 
-## Missing Mice, finishing dataset
-
-*Which mice are remaining and at what stage?*
-
-This section shows the number of mice which haven't been stained or quantified yet.
-
-```{r missing.mice, echo=FALSE}
-
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-setwd("~./MLH1repo/")
-#other lists
-
-Image_dirs <- list.files(path = "C:/Users/alpeterson7/Documents/Images")
-#write this out so that I can access it
-
-#remove the non-mouse folders
-Dissection.File = read.csv("~./MLH1repo/data/Disections_53019.csv", header = TRUE)
-Dissection.File <- Dissection.File[!(is.na(Dissection.File$mouse)|Dissection.File$mouse==""),]
-
-#print the name with the date of the file somewhere
-Dissection.list <- Dissection.File$mouse
-Dissection.list <- as.character(Dissection.list)
-
-#load metadata .. start decerning which mice are missing information... and which should be used
-Meta.Data.File = read.csv("~./MLH1repo/data/ALP_MouseMetadata.csv", header = TRUE)
-
-#remove file name, and original name .. this is just a mouse level DF
-Meta.Data.File <- Meta.Data.File[!(is.na(Meta.Data.File$mouse)|Meta.Data.File$mouse==""),]
-#make sure that the DOB is a date
-
-meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61019.csv", header = TRUE)
-meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%y")
-
-jl <- unique(Meta.Data.File$mouse[(Meta.Data.File$mouse %in% Dissection.File$mouse)])
-BD.w.meta <- Meta.Data.File[(Meta.Data.File$mouse %in% Dissection.File$mouse), ]
-
-#figure out a way to incorporate the staining... to get a better idea of which slides 
-#to do next, #remove the pero mice?
-
-#length(unique(MLH1_data$mouse) )
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-original_DF$fileName <- (original_DF$Original.Name)
-original_DF <- add_mouse(original_DF)
-#length(unique(original_DF$mouse) )
-#131 mice have been quant total, 124 mice have passed into the final version
-
-# make a list for the failed slides
-
-#Dissection.list - Image/dirs
-NonImaged.mice <- unique(Dissection.File$mouse[(Dissection.File$mouse %in% Image_dirs)])
-#231 mice dissected which need images
-
-#ImageFolder - MLH1.Quant
-NonQuant.mice <- unique(Image_dirs[(Image_dirs %in% MLH1_data$mouse )])
-#length(NonQuant.mice)
-#123 mice that need to get into anon-batch
-
-#make Dataframe to put into a table
-mouse.number.DF <- data.frame( category = 
-              c("Passed Quantified", "Total Quantified", "Non-Quantified", "Non-Imaged"),
-              number.of.mice = c(length(unique(MLH1_data$mouse)),
-                length(unique(original_DF$mouse) ),
-                        length(NonQuant.mice),
-                         length(NonImaged.mice)  )  )
-
-kable(mouse.number.DF)
-```
-
-passed quantified - quantified mice (with enough cells)
-total quantified -- all mice which have been (includes fail and passing mice)
-
-non-quantified mice -- dissected mice / slides that haven't been quantified yet (all stained, but without image directory)
-
-non-imaged -- mice without images, (might be stained and unstained)
-
-This is the number of mice that are through different parts of the pathway or quantification process.
-
-should I break down the categories into sex and strain?
-
 ```{r basic.numbers, echo=FALSE, eval=FALSE}
 
-kable(table(AP_mouse_table$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
 
@@
 
-These are the tables for the number of cells and mice which have been quantified.
-The 'Other' are F1 crosses.
-
-## Repeatability of Human quantification
-
-Well, we have to talk about the subjective nature that goes into scoring the MLH1 foci.
-Sources of error or variance across scoring come from background noise (false positives), faint foci (false negatives) and just human error.
-A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (using some of Beth's data).
-Todo: track down, re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)
-
-In the original_DF and MLH1 DF's there are ~400 and X respectively of cells (filenames) that are duplicated, or cells that have been counted in more than 1 quant batch
-
-Datasets / sources of repeat measures
--chrom proportions
--multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
--old and new (batch0)
-
-The main take away is that higher quality scored cells have lower error in MLH1 counts.
-
-> Does the amount of error differ across strains or sexes? (female might have a higher error rate)
-
-> What sort of adjustments or tests should be made to results or analyses given the known error rate?
-
-```{r quant.repeatability_rep1, echo=FALSE, include=TRUE}
-#set up data for combining replicate counts
-
-#1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now
-
-#isolate the duplicates (fileName) from main MLH1 DFs
-n_occur <- data.frame(table(original_DF$fileName))
-n_occur[n_occur$Freq > 1,]
-original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,
-
-#are there any batch, fileName duplicates?
-original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")
-anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch ... really 2?
-
-n_occur <- data.frame(table(original_DF_duplicates$unique.key))
-n_occur[n_occur$Freq > 1,]
-double.trouble <- original_DF_duplicates[original_DF_duplicates$unique.key %in% n_occur$Var1[n_occur$Freq > 1],]
-#2 fileNames from batch 0 appear 2x, remove them from DF
-
-original_DF_duplicates <- original_DF_duplicates[ ! original_DF_duplicates$fileName %in% double.trouble$fileName, ]#removes 4 rows
-
-#seperated duplicates into categories out in Excel (write to /results, moved final file to /data)
-#This data frame was made ... with Jered's sugg
-#write the duplicate file to csv/excel
-#sort by file name
-#need to remove '?' -- deal with the ? and X's
-#write.table(original_DF_duplicates, "~./MLH1repo/results/orginal_DF_duplicates.csv", sep=",", row.names = FALSE)
-
-original_DF_duplicates.notes = read.csv("~./MLH1repo/data/orginal_DF_duplicates.csv", header = TRUE)
-#copies of fileNames can be distinguished with the 'separate.DF' column
-
-#remove '?' that I stupidly left in
-original_DF_duplicates.notes$nMLH1.foci = gsub("\\?", "", original_DF_duplicates.notes$nMLH1.foci)
-original_DF_duplicates.notes$quality = gsub("\\?", "", original_DF_duplicates.notes$quality)
-#these lines change col to chr
-
-original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
-#now there should be no more '?' going forward
-
-#original_DF_duplicates.notes$quality <- as.factor(original_DF_duplicates.notes$quality)
-
-#make lists of rows with "" or X's so I can remove all fileNames for those
-#these lists just have na
-quality.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$quality) | original_DF_duplicates.notes$quality=="")] #16
-
-MLH1.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci=="") ]#22
-
-X.list <- original_DF_duplicates.notes$fileName[grepl("X", original_DF_duplicates.notes$X)]#40
-
-#make DF which only has 
-matches.only <- original_DF_duplicates.notes[ ! original_DF_duplicates.notes$fileName %in% quality.na.list, ]
-matches.only <- matches.only[ ! matches.only$fileName %in% MLH1.na.list, ]
-matches.only <- matches.only[ ! matches.only$fileName %in% X.list, ]
-
-#str(matches.only)
-#clean up match list (chr to num)
-matches.only$nMLH1.foci <- as.numeric(matches.only$nMLH1.foci)
-matches.only$quality <- as.numeric(matches.only$quality)
-
-levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt
-
-#check that the 1, and 2 are equal number
-length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178
-
-#i want to measure correleation of MLH1 across bins of quality (1,2,3,4,5)
-#make new df with filename and MLH1.1 and mlh1.2
-
-dups.here <- data.frame(  FN = unique(matches.only$Original.Name), MLH1.1 = matches.only$nMLH1.foci[matches.only$separate.DF == 1], MLH1.2 = matches.only$nMLH1.foci[matches.only$separate.DF == 2],
-                        qual.1 = matches.only$quality[matches.only$separate.DF == 1] , 
-                        qual.2 = matches.only$quality[matches.only$separate.DF == 2]
-                          ) #OMG this worked!!!
-dups.here$diff <- abs(dups.here$MLH1.1 - dups.here$MLH1.2) #the aveage difference is pretty small, but some have very high difference
-#the high differences could potentially be mix ups
-
-repeatability.table <- ddply(.data=dups.here, 
-                 .(qual.1, qual.2),
-                 summarize, 
-                 ncells = length(unique(FN)),
-                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
-)
-#this table shows how many cells went changed into different qual score bins
-#easiest patter, 1-1, 1-2 (high correlation in MLH1)
-#5-5 lowest, (5-3 also low)
-#4-5 has high corr? (direction shouldn't matter)
-#4-2 has negative corr?
-
-#of the different quality permutation bins, 37% don't change quality bin, 44% of cells change by 1 (either direction)
-
-no.change = c(14,13,17,13,10)#67
-change.1 = c(10,7,8,20,9,14,4,7)#79
-
-repeatability.table.qual1 <- ddply(.data=dups.here, 
-                 .(qual.1),
-                 summarize, 
-                 ncells = length(unique(FN)),
-                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2),
-                 dif = mean(MLH1.1 - MLH1.2)
-)
-
-repeatability.table.qual2 <- ddply(.data=dups.here, 
-                 .(qual.2),
-                 summarize, 
-                 ncells = length(unique(FN)),
-                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
-)
-#These tables show that qual1 cells have the highest corrleation of MLH1
-dups.here.plain_cor_qual <- cor(dups.here$qual.1, dups.here$qual.2)
-
-#i don't know how to compare quality in more
-
-#nMLHJ1, quality are factors
-original_DF_duplicates.notes$quality2 = gsub("\\?", "", original_DF_duplicates.notes$quality)
-
-original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
-
-#factor > chr > number  (this col is now chr is can go to num)
-#original_DF_duplicates.notes$MLH1.foci2 <- as.numeric(original_DF_duplicates.notes$MLH1.foci2)
-#original_DF_duplicates.notes$quality2 <- as.numeric(original_DF_duplicates.notes$quality2)
-
-#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
-#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
-#rename this
-#plain_cor_MLH1 <- cor(one$MLH1.foci2,two$MLH1.foci2,use = "complete.obs")
-#.857
-#plain_cor_qual <- cor(one$quality2,two$quality2,use = "complete.obs")
-#.63
-
-#What's the relationship of MLH1 cor with quality?
-#what the rpeat of quality score?
-#what are the correlations across quality score bins?
-
-#issues that complicate the plain calculation
-#some duplicates were 'x'ed and some weren't -- this causes different numbers
-#connected with above, some duplicates have mlh1 scores, other don't -- this kinda messes up correlation
-
-original_DF_duplicates.notes <- original_DF_duplicates.notes[ !grepl("X", original_DF_duplicates.notes$X) , ]
-original_DF_duplicates.notes <- original_DF_duplicates.notes[!(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci==""), ]
-#this causes unequal lengths
-
-original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
-original_DF_duplicates.notes$nMLH1.foci <- as.character(original_DF_duplicates.notes$nMLH1.foci)
-
-original_DF_duplicates.notes$nMLH1.foci <- as.numeric(original_DF_duplicates.notes$nMLH1.foci)
-
-#correlationsnMLH1 
-#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
-#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
-#plain_cor <- cor(one$nMLH1.foci,two$nMLH1.foci)
-#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
-
-#check that all batches are distinguishable across duplicate fileNames
-#make unique key fileName_batch -- then check for duplicate of that 
-#original_DF$Batch <- factor(original_DF$Batch, ordered = TRUE, levels = c("Batch0", "Batch1", "Batch2", "Batch3", "Batch4","Batch5","Batch6",  "Batch7", "Batch8","Batch9","Batch10", "Batch11","Batch12", "Batch13", "Batch14","Batch15"))
-#after doing this, original_DF$Batch[2000] > original_DF$Batch[10] will be evaluated!
-
-```
-
-```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
-#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
-#load, clean up and mat one of the master.MLH1count sheets
-#master
-old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
-#remove all with 'X'
-#create a real fileName col, add rev.tif, 
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-
-old.master.MLH1 <- add_mouse(old.master.MLH1)
-old.master.MLH1 <- add_strain(old.master.MLH1)
-old.master.MLH1 <- add_sex(old.master.MLH1)
-old.master.MLH1 <- add_category(old.master.MLH1)
-old.master.MLH1$fileName2 <- old.master.MLH1$fileName
-
-test.old.master <- rbind(head(old.master.MLH1), tail(old.master.MLH1) )
-test.old.master$fileName <- as.character(test.old.master$fileName)
-
-test.old.master$stain <- as.character(test.old.master$stain)
-
-str(test.old.master)
-
-#paste together file name, requires mouse, strain
-#this might also need some character / factor stuff ... filename and stain MUST be charaecter!
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#wrote add_fullFileName to make the file names for the old excel sheets
-
-old.master.MLH1 <- add_fullFileName(old.master.MLH1)
-old.master.MLH1$fileName <- old.master.MLH1$file.name_full
-
-#merge and match to MLH1
-
-#need to test for duplicates before mergeing?
-n_occur <- data.frame(table(old.master.MLH1$file.name_full))
-n_occur[n_occur$Freq > 1,]
-ll <- old.master.MLH1[old.master.MLH1$file.name_full %in% n_occur$Var1[n_occur$Freq > 1],]
-old.master.MLH1 <- old.master.MLH1[ ! old.master.MLH1$fileName %in% ll$fileName, ]
-
-n_occur <- data.frame(table(original_DF$fileName))
-n_occur[n_occur$Freq > 1,]
-original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]
-table(original_DF_duplicates$mouse)#17 mice
-
-old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
-
-#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
-
-#there are duplicates in original_DF and MLH1
-
-uutesting2 <- merge(old.master.MLH1, original_DF, by.x = fileName, by.y = Original.Name)
-#match old.full.filname with MLH1$fileName
-```
-
-```{r repeatability_rep3, echo=FALSE, include=FALSE}
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#what are the target columns for this excel sheet?
-#espically for getting repeat data
-#MLH1.dif  same_count  quality
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-
-#assessing error,
-#number for same_count == TRUE
-#length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"])
-#length(chrm.class.DF$Original.Name)
-
-#cor of first count and calq count
-#cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)
-#.87
-
-#the true false might have to be changed first
-#chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)
-
-#regression of quality and MLH1.dif
-#lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
-#summary(lmreg)
-#according to this linear regression, quality is not a significant
-
-#make regression for just same counts
-#I tried making a  binomial regression model, 
-#but quality is bad predictor, only 5 levels)
-#lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
-#summary(lmreg2)
-#to plot the model / bimodal data
-#fit = glm(same_count_num ~ as.numeric(quality), data=chrm.class.DF, family=binomial)
-#newdat <- data.frame(quality=seq(min(chrm.class.DF$quality), max(chrm.class.DF$quality),len=102))
-#newdat$vs = predict(fit, data=newdat, type="response")
-#the prediction is not going 1 or 0 -- so doesn't fit logistic
-#plot(vs ~ quality, data=newdat, col="red4")
-#lines(vs ~ quality, newdat, col="green4", lwd=2)
-
-hist_error_plot <- ggplot(data=chrm.class.DF, aes(MLH1.dif))+geom_histogram()+ggtitle("Distribution of foci count differences across 2 count versions")
-
-#
-quality_error_plot <- ggplot(data=chrm.class.DF, aes(x=quality, y=abs(MLH1.dif)  ))+geom_jitter()+ggtitle("Repeatability by quality score")+ylab("difference in quality score")+xlab("quality score")
-
-error_count_plot <- ggplot(data=chrm.class.DF, aes(y=nMLH1.foci, x=abs(MLH1.dif)  ))+geom_jitter()+ggtitle("Error by MLH1 number")+xlab("difference in two counts")+ylab("MLH1 count")
-
-#in the chrm proportions data is another set for eastimating error
-```
-
-Data use for this repeatabiity estimate is from smaller dataset where the proportions of chromosome/bivalent classes (0CO bivalents ect.) were quantified.
-(The focus of one count was cell total, while the other focus was classifying bivalents).
-
-This data has `r length(chrm.class.DF$Original.Name)` cells across these categories
-
-`r kable(table(droplevels(chrm.class.DF$category)))`
-
-More cells are planned to be quantified.
-
-```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE}
-hist_error_plot
-```
-
-From this histogram we see most of the cells in this dataset have a difference of 2 in the MLH1 count scores (`r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`).
-
-```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5}
-quality_error_plot
-```
-
-There is a positive relationship with the absolute difference and quality score (where 1 is best and 5 is worst quality) (i.e.
-poorer quality cells are more likely to have more error acorss multiple counts).
-The proportion of the cells which had the same count across the counting instances is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`.
-The correlation between the two versions of MLH1 counts for the same cell is `r cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)`.
-
-```{r show.error_count_plot, echo=FALSE,  fig.width=10, fig.height=5}
-error_count_plot
-```
-There doesn't seem to be a bias across cells with more or less foci.
-
-Modeling this relationship; (still in progress)
-I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.
-
-## Effects on the dataset
-
-error associated with 'quality of cels'
-
-Main metric is mean CO count per mouse. Run tests to check if there are effects for the following effects.
-
-*age effects* (male, maternal age)
-
-*litter effect*
-
-*mouse room effect* (Charmony, MSC)
-
-*quantification repeatability*
-
-*spread batch and staining effects*
-
-*power analysis on number of cells*
-
-*time of year effect (on spread)*
-
-*dominant effect* (F1s)
-
-These aspects will be analyzed with a linear models on subsets of data (within category) and visualized with plots.
-
-### Quality effect
-
-There is a positive correlation with MLH1 number and quality
-To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.
-
-```{r quality.effect, echo=FALSE}
-
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
-#compare all quality 1 cells across strains
-#check that the means (or directions of differences is maintained)
-
-Q12_MLH1_DF <- MLH1_data[MLH1_data$quality <= 2,]
-
-#ddply
-Q12_MLH1_means <- ddply(.data=Q12_MLH1_DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-tot.DF_MLH1_means <- ddply(.data=MLH1_data, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-mergy <- merge(Q12_MLH1_means, tot.DF_MLH1_means, by='category', all = F)
-colnames(mergy) <- c("category","Q12ncells","Q12mice","Q12_MLH1_means","Q12MLH1var",
-                     "Totncells","Totnmice","TotMLH1mean","TotMLH1var")
-
-mergy$percent.diff <- (mergy$Q12_MLH1_means / mergy$TotMLH1mean)
-
-Q12toTotal <- ggplot(mergy, aes(y=Q12_MLH1_means, x=TotMLH1mean))+geom_point(aes(size = Q12ncells) )+geom_abline(slope = 1, intercept = 0)+ggtitle("Quality 1 cells MLH1 mean to total MLH1 mean")+ylim(c(21,32))+xlim(c(21,32))
-
-Q12toTotal
-
-#check that the directions are the same
-#make tables for the 
-#is the amount of 'polymorphism' the same
-
-Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "male",], 
-                 .(category,subsp),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci)
-)
-
-tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "male",], 
-                 .(category,subsp),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci)
-)
-
-kable(Q12_MLH1_means_males, caption = "High quality cells")
-
-kable(tot.DF_MLH1_means_males, caption = "Total Cell")
-
-```
-
-All the quality 1 cells produce a higher mean. The biggest difference in means is HMI male, LEW female and SPIC female means have ~10% differences.
-
-When quality 1 to 2 is compared, the strains were closer to the line, but the means are still higher in the higher quality cells.
-with 1-2 most all of the difference are below 10% (except SPRET female).
-
-MOLF males have quite a big difference.
-
-There seem to be the similar amount of polymorphism in the Dom and more conserved in Dom.
-
-Ranges of category means across subspecies.
-
-| Type | male Dom | female Dom | | male Musc low | male Musc high | Musc female | | male Outgroup | female outgroup |
-| --- | --- | --- | --- | --- | --- | --- | --- | --- | --- |
-| Total cells | 22-24 |  24-27 |  | 23-31 | 25-31 | 26-29 | | 25-26 |27 |
-| High quality cells | 23-25 | 25-29 | | 22-25 | 22-27 | 25 | | 26-27 | 29-31 |
-(ASTmale in Musc high, has a 27 mean which might be an outlier)
-I think the ranges across the groups is generally consistant.
-
-(try looking at <5 quality scores?)
-
-```{r high.qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
-
-#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
-highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
-
-#are there no KAZ cells of high quality?
-qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey", 
-             "grey2","grey3","grey1","grey","grey","grey"))
-high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
-#high_qual_boxplot
-#these boxplots look bad
-```
-
-> *How do the distributions change across different quality scores?*
-
-Higher quality scored cells have more MLH1 foci. (some of the ranges get smaller with higher quality cells)
-
-> *How do the distributions change across mice with a good number of cells?*
-
-(was I referring to normale distributions?)
-means all shift up, what about variance, range, lowest cut off
-
-> *Do the directions of the differences across groups change due to quality scores?*
-
-(polymorphism vs divergence patterns)
-
-### Age Effect
-
-Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age].
-
-The initial analysis uses plots. This should be double checked with a general linear model as well.
-
-```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
-
-#make Mouse meta data work .. rename these
-MouseMetaData.wrkin <- meta.data[,c(1,4)]
-colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
-MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
-#do this for the mouse table instead of the MLH1 table
-MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
-#DOB is date
-MLH1.imprv <- add_euth_date(MLH1.imprv)
-#good euth date..
-
-#make backup date col #add euth date for calculating age
-MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
-MLH1.imprv <- add_age(MLH1.imprv)
-#age works
-
-#remove mice that are older than 20 weeks
-MLH1.imprv.male <- MLH1.imprv[MLH1.imprv$sex == "male", ]
-MLH1.imprv.male <- MLH1.imprv.male[MLH1.imprv.male$age.weeks < 20, ]
-
-age.hist <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
-age.hist
-
-```
-
-Make a subsetted dataset for plotting mean mouse (quality < 5) MLH1 count and age.
-
-```{r age.effect.model, include=FALSE}
-#run a age effect linear regression!
-
-male.data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.y == "male",]
-
-male.data.dom <- male.data[male.data$subsp.y == "Dom",]
-
-male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST"),]
-
-male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "MOLF" | male.data$strain.x == "SKIVE"),]
-
-male.data.outgroup <- male.data[(male.data$subsp.y == "Spretus" | male.data$subsp.y == "Spic" | male.data$subsp.y == "Caroli"),]
-
-lm.me.dom  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.dom)
-
-lm.me.musc.low  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.low)
-
-lm.me.musc.high  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.high)
-#significant differences across strains in musc high
-
-lm.me.OG  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.outgroup)
-
-summary(lm.me.OG)
-#for all models, mostly it was that the intercept was significant, and mot much else
-```
-
-```{r female.age.effect, echo=FALSE}
-#start thinking of the maternal age effect
-
-#female age effect
-
-#female.meta.data <- meta.data[meta.data$sex == "female",]
-#how much data to I have to work with?
-#there's a fair number of rows without emb or neonate
-#I could probably assume all of these are neonates
-
-#length( !is.na(female.meta.data$female.stage) )
-
-#female.meta.data <- female.meta.data[female.meta.data$female.stage]
-
-#how much formating do I have to do?
-#can't use AP_mouse_table since that is summarized
-#meta.data
-#grepl notes, "neo"
-
-#some have e-number for embryonic
-#I might have to
-
-```
-
-I will try doing a logistic regression for embryonic vs neonate. The embyonic estimates likely aren't very accurate, and there won't be good sampling across time points.
-
-```{r age.effect.plot, warning=FALSE, echo=FALSE}
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-
-AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
-#remove NAs
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]
-
-#make sure the right dataset 
-male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
-#lots of NA rows... ugh .. remove them
-male.data <- male.data[!(is.na(male.data$Nmice) | male.data$Nmice==""), ]
-
-#other removed, no points,
-male.data <- male.data[male.data$Ncells > 10,]
-
-age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
-  #trying this
-  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
-age.effect.plot
-#no longer has 'other', #yay I added purtty standard colors
-#there is an NA
-#todo add line to strain.y
-
-```
-
-While many slopes are different than 0, I am not convinced there is a clear age effect.
-Most strains don't have more than 2 ages.
-I will double check some of these mice and check a general linear model.
-
-### Maternal Age effect
-
-Check if there is a maternal age effect for females and males.
-
-```{r mat.age.effect, echo=FALSE}
-meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
-#445? -- probably extras at bottom not good
-
-#how many mice have maternal DOB
-meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
-#151
-
-#format this meta.data.file
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-
-meta.data.file.w.mat.DOB <- add_mouse(meta.data.file.w.mat.DOB)
-meta.data.file.w.mat.DOB <- add_category(meta.data.file.w.mat.DOB)
-meta.data.file.w.mat.DOB <- add_strain(meta.data.file.w.mat.DOB)
-meta.data.file.w.mat.DOB <- add_sex(meta.data.file.w.mat.DOB)
-meta.data.file.w.mat.DOB <- add_subsp(meta.data.file.w.mat.DOB)
-
-#check if date format's been applied before tryung the age functions
-str(meta.data.file.w.mat.DOB)
-
-#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
-#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
-meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
-meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")
-
-str(meta.data.file.w.mat.DOB)
-
-meta.data.file.w.mat.DOB <- add_euth_date(meta.data.file.w.mat.DOB)
-meta.data.file.w.mat.DOB <- add_age(meta.data.file.w.mat.DOB)
-meta.data.file.w.mat.DOB <- add_mat_age(meta.data.file.w.mat.DOB)
-
-#merge meta data with MLH1 
-see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
-#remove anything without mean_CO
-see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
-see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
-
-#visualize the data
-
-meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
-meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]
-
-#hist of data
-mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
-mat.age.hist.female
-
-mat.age.hist.male <- ggplot(meta.data.file.w.mat.DOB.male, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of males")
-mat.age.hist.male
-
-#remove mice that might have low quality
-see.whatcha.do.HQ <- see.whatcha.do[see.whatcha.do$Ncells > 14,]
-
-#age by mean_co
-mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
-  #trying this
-  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
-mat.age.effect.plot.female
-
-mat.age.effect.plot.male <- ggplot(see.whatcha.do.HQ[see.whatcha.do.HQ$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
-  #trying this
-  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")+facet_wrap(~subsp.y)
-
-#
-ALLmat.age.effect.plot.male <- ggplot(see.whatcha.do[see.whatcha.do$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
-  #trying this
-  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")
-
-+facet_wrap(~subsp.y)
-
-#is there a correlation with proband age and maternal age?
-#there is a big lack of maternal age data...
-#some of the values reference the original breeding pair..
-#from notes on original BPs(males), assigning maternal ages
-
-```
-
-There seems to be a negative correlation with
-
-### Litter effect
-
-Compare female spreads made on the same day, but from different mothers/litters.
-
-### Spread and Stain Effects
-
-> *month dissected*
-
-For mice strains that have a similar MLH1 level (control for strain MLH1 effect), check if a month effect for MLH1 count (as seen in Peromyscus).
-
-> *staining batches*
-
-Compare observations from the same mouse, but different MLH1 stains.
-
-### Room effect (envioronment)
-
-3 different rooms used to house mice. Access softmouse to be fixed. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
-
-### Cell number (power analysis results)
-
-Insert the code and results for effects of cell number/observations on MLH1 means.
-
-# Initial Results
+```{r rand, echo=FALSE, include=FALSE, eval=FALSE}
+
+str(AP_mouse_table_w.Ages)
+
+AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+
+#remove sex.x == na, and strain.x=na or F1
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
+#plot variance and mean (facet sex)
+
+#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
+
+VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
+
+cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
+
+str(AP_mouse_table.HQ)
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
+#[1] 0.5750399
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
+#0.321601
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
+# 0.3515544
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
+#-0.03713536
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
+# p-value = 8.623e-09
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
+
+#remove mice with fewer than 10 cells
+
+AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
+
+VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
+
+cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
+
+cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
+    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
+
+```
+
+### Distributions of MLH1 counts
+
+```{r MLH1.histogram.2421, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
+
+#gg <- data.frame(weight=rep(x,nrow(df)),df)
+#gg$y <- with(gg,dnorm(x,mean,sd))
+#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
+
+mean.MLH1.obs <- ddply(.data=MLH1_data, 
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+histM <- ggplot(data=MLH1_data,(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")
+
+#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
+#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
+
+#ad
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+
+histM
+
+#playing with expected posisson plot...
+#function = glm(y ~ x, family = poisson)
+#add Poission curve behind
+#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
+#not sure how to add Expected poisson curve
+
+```
+
+Most histograms of pooled MLH1 counts are all close to normal. Except PWD male, LEW female, SKIVE male, SPIC male.
+check if there is a quality effect.
+
+#  Results
 
@@
 MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
-              "PWD","MSM","MOLF","CZECH","SKIVE",
+              "PWD","CZECH","SKIVE",
               "KAZ","TOM","AST",
+              "MSM","MOLF",
               "CAST","HMI","SPRET","SPIC", "other"), order=T )
@@
 MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-#don't know why this won't do the right order...
-#set the order of another column, based on another variable (so that when)
-
-#try to make sure sacles are the same
-
-#reduncant
-#MLH1_data <- MLH1_data %>%
- # arrange(strain, sex, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-#try this for ordering
-#MLH1_data[order(MLH1_data$strain), ]
 
@@
 #don't show M.musculus by mouse
+#scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
+ #       "green","green4","grey",  "grey","grey","grey" )) +
 
 #order MLH1_data by strain!!
-
-ff2 <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
-  geom_boxplot(aes(x = as.character(mouse), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+#colors_of_strains, colors .. 
+#there is no order to 'mouse' so it looks shitty
+
+justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
+  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
-        "green","green4","grey",  "grey","grey","grey" )) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
-  facet_grid(subsp~ sex, scales="free_x")#scales="free_y"
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+
+
+justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")
+
+#scales="free_y"
 #this is ok except for the set placement of mice on x axis
@@
 ss
-```
-
-*What is the initial picture of Heterochiasmy across mouse strains?*
-
-For the strain level plots
-  -facet,sex by species
+justMuscu
+```
 
@@
 #calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
-
 #subset MLH1_DF, remove non passing mice (as diagnoised from plots)
@@
 
-#make scatter plot for HetC
-#color by strain
-
 #set up data
-#make female male plot
 
@@
 
+#remove the 'other'
+
 HetC_strain_table.M$sex <- "male"
@@
 
-Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
-  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey"))
-
-#musc - coral, Dom - cadetblue (grey blue),
+Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
+  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
 
@@
 
+There is now MOLF, which has female biased hetC
 3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
@@
 
-### Distributions of MLH1 counts
-
-> *Are the MLH1 distributions close to normal?*
-
-```{r MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}
-
-#gg <- data.frame(weight=rep(x,nrow(df)),df)
-#gg$y <- with(gg,dnorm(x,mean,sd))
-#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
-
-mean.MLH1.obs <- ddply(.data=MLH1_data, 
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")+
-  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
-
-#  geom_smooth(method = lm, se = FALSE
- # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
-
-histM
-
-#function = glm(y ~ x, family = poisson)
-#add Poission curve behind
-#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
-#not sure how to add Expected poisson curve
-
-```
-
-(most all distributions are close to normal)
-
-How do these compare to Poisson? (mean == variance)
-I should make 'expected' distributions behind these.
-
-```{r passing.mice, echo=FALSE, warning=FALSE}
-q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
-                        total =  length(adj_nMLH1.foci),
-                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
-                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
-                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
-)
-#q_cutoff_table
-passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
-
-not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
-
-#mnake dataframe with just passing mice -- so that it can be ploted
-passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
-```
-
-Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
-Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
-(CAST female data is fake)
-
-```{r scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means and quality scores", warning=FALSE}
-
-#make lattice / plots seperated by strain and sex
-#facet
-
-#quality scores are now 1:7?
-qual_means <- MLH1_data %>% 
-        group_by(category, quality) %>% 
-        summarise(
-          adj_nMLH1.foci = mean(adj_nMLH1.foci)
-        )
-
-p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
-p <- p + geom_point(stat = "identity")
-p <- p  + geom_point(data = qual_means, colour="red") 
-p <- p + facet_wrap(~ category)
-p
-
-```
-
-I plotted the mean of each quality bin with a red dot.
-From the pattern of the red dots, there is definitely a negative relationship with quality and nMLH1 foci across the data.
-This is most pronounced in MSM males and least pronounced in G males.
-The CAST female data is dummy data.
-
-####Distributions by mouse
-
-> *Is there mouse level MLH1 quality score bias?*
-
-Below are code chunks that make a big dfs of CO count by quality score plot. The code has been commented out, and moved to it's own doc.
+#### Distributions by mouse
+
+There are code chunks that make a big dfs of CO count by quality score plot. The code has been commented out, and moved to it's own doc.
 
@@
 
-# Research Questions
-
-## Explaination for the XX and XY differences
-
-(in progress)
-
-## Difference in proportion of chrm classes
+### Difference in proportion of chrm classes
 
 generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
-10 cells from each musc category.
-
-```{r draw.rand.cells, echo=FALSE}
-#ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
-#ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
-```
+10 cells from each musc category. **Add Molf male**
 
@@
 
-PWD and MSM have a higher proportion of 2CO chrms!
-
-what about the variance across cells?
-
-## Models for manuscript
-
-Ideas for models of evolution of Heterochiasmy and current idea for how to test them.
-
-- HetC arises due to more precursors (DSBs)
-
-    - Compare DSB counts
-
-- Larger cell volume drives longer chrms axis (and maybe more DSB)
-
-    - Measure meiocyte diameter in males and females. Compare spermatocyte diameters in male biased and female biased strains. 
-    - Compare DSBs counts and average SC/axis lengths.
-
-- Difference in astral/MT organization
-
-    - ?
-
-- Something about interference and sister cohesin area
-
-    - Develope metric for sister cohesin area, and look at initial patterns.
-
-## Mixed Model, compare within and between subspecies patterns
+High musc males have half 1CO and 2CO bivalents.
+For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
+
+# Mixed Models
+
+todo: make model with strain as fixed effect (since, there are interesting strain effects)
+
+The mixed model lets us sperate effects to see how they effect a mouse mean CO count. This seperation can also serve as proxies for divergence and polymorphism.
 
@@
 
-*Does this hold up with the full data set?*
-*What's the best visualization for these questions?*
-
-Main focus is contrasting the level of difference within and between subspecies.
-
-I'll do this with a mixed model framework approach (and analyze with ANOVA, a method to partition the variance of subset.)
-
-```{r mixed.model, echo=FALSE}
-#build mixed model
-#run simulations for accounting for sampling error
-# old code uses base plot
-
-```
-
-## Meiotic Error
-
-This section reviews all the data from achiasmy and 0CO bivalents.
+```{r mixed.models.setup, echo=FALSE}
+
+#make are the DFs for the models
+# seperate the data, both sexes, only Mus
+# decide on Musc - Molossinus
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
+                                                            subsp.y,subsp.x,
+                                                            category.y, category.x,
+                                                            strain.x,strain.y, mouse.1  )  )
+
+#make dummy MOLF data, average of Musc females?
+#no MOLF male data?
+female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
+table(female.musc.cells$quality)
+
+female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)#just 1 entry
+
+female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
+female.musc.mice.table$DOB <- as.Date("1/1/2020")
+
+female.musc.mice.table$strain <- "MOLF"
+female.musc.mice.table$subsp <- "Musc"
+female.musc.mice.table$species <- "M.musculus"
+
+female.musc.mice.table$female.stage <- "emb"
+female.musc.mice.table$age.days <-""
+female.musc.mice.table$age.weeks <- ""
+female.musc.mice.table$ID <- ""
+female.musc.mice.table$notes <- ""
+female.musc.mice.table$euth.date <- ""
+
+DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
+
+#main df with variables coded as chr
+DF.HetC.MixedModel <- DF.HetC.MixedModel
+DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
+DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
+DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
+
+#mean_co.HQ  (this might remove some of the MOLF males)
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mean_co.HQ.mol
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
+        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
+        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
+
+#sub-sample (need to think about this one)
+
+```
+
+For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)
+
+subspecies levels: 2, Musc and Domesticus (maybe Molossinus soon)
+strain levels: Dom: 3, Musc: 4
+sex: 2 female male
+
+`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
+
+> Mouse average Models
+
+List of mixed models with mean_co
+
+1. mean_co (chr coded)
+2. mean_co.HQ
+3. mean_co with Mol
+4. sub-sampling
+
+```{r HQ.MixedModel, echo=FALSE}
+table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#mean_co.HQ  (this might remove some of the MOLF males)
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#these remove ALL male MSM!
+table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#mean.co.HQ
+HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.HQ)
+ranef(HetC.MixedModel.HQ)
+
+#reduced models
+Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+summary(Reduced.strain.HQ)
+ranef(Reduced.strain.HQ)
+
+Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+summary(Reduced.all.HQ)
+#ranef(Reduced.all.HQ)
+
+#strain = fixed
+strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+summary(strain.fixed.full)
+#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
+#sexmale:strainWSB p=  0.006740 **
+#sexmale  p =  3.98e-05 ***
+#strainWSB   p=0.000441 ***
+#subspMusc:sexmale  p=  0.069493 . 
+#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are
+
+strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+summary(strain.fixed)
+
+#sexmale  p= 3.98e-05 ***
+#strainKAZ p=  0.052420 .  
+#strainWSB  p= 0.000441 ***
+#sexmale:strainMSM  p=3.42e-09 ***
+#sexmale:strainPWD  p= 5.79e-08 ***
+#sexmale:strainSKIVE p= 0.069493 .  
+#sexmale:strainWSB  p=  0.006740 **
+
+#maybe this WSB effec is just in the HQ data set?
+
+#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
+full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
+
+full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
+#this comparison not useful when there are different fixed effects
+
+full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ)# < .0001
+
+strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ)#NS
+```
+
+Moatly matches the base model results. the same fixed effects are significant in the reduced models. the random effects make sense
+
+Peromyscus did everything as factors... (but unordered)
+
+random effects were tested with 
+ #rand LRT
+this function might not support MM, (and needs linear models instead)
+
+also throws an error about only looking at 1 random effect
+
+Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
+Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
+SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
+SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
+
+(what does the significant random test mean) -- variance is not 0?
+
+```{r variance.boxplots, echo=FALSE, eval=FALSE}
+#how did I make this
+
+#plot of variance
+var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
+
+#plot by sex / seperately
+# HQ data DF.HetC.MixedModel.HQ
+
+var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+var.male <- var.male   + facet_wrap( ~ strain, scales="free")
+
+cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
+
+```
+
+```{r BaseMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+#strain as fixed
+newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
+
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
+
+full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
+
+#CONTRAST models
+
+#base vs strain
+fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co)# <.0001
+
+#strain vs no random full.reduce
+fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)#<.0001
+
+#fullreduce vs strain
+
+#can't distinguish which random effects are more significant -- both tests are significant
+
+fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
+
+#random effects for base model
+ranef(HetC.MixedModel.mean_co)
+#random effect for just strain
+ranef(R.strain.HetC.MixedModel.mean_co)
+
+#reduced model's PWD radom effects is much smaller
+#the full model's largest random effects are for PWD and MSM (and MOLF, KAZ for negative)
+
+#fixed effects from full
+anova(HetC.MixedModel.mean_co)#NS
+
+#Check how the fixed effects change across these
+
+HetC.MixedModel.mean_co2 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel)
+#lmer(y~x1 + x2 + (1|g1), data=pbDat)
+
+lmer.MM <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel, REML = TRUE)
+
+```
+
+Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
+Models with random effects more significant than nulls (both just strain and no radom effects).
+The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
+
+```{r Base.strain.fixed, echo=FALSE}
+
+#make mixed model with strain as fixed, since there are interesting results
+
+HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
+#this produces several significant factors
+
+HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
+summary(HetC.MM.strain.fixed2)
+#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
+
+#summary(HetC.MM.strain.fixed)
+#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
+#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
+#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
+#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
+#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
+#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
+#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
+#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
+
+#male effect lowers the mean from intercept
+#while strain effects 
+```
+
+```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
+#base model with molossinus level ... why am I making molossinus model?
+
+DF.HetC.MixedModel.mol <- DF.HetC.MixedModel
+str(DF.HetC.MixedModel.mol)#subsp is factor with 2 levels
+
+DF.HetC.MixedModel.mol$subsp <- as.character(DF.HetC.MixedModel.mol$subsp)
+
+#turn to chr first?
+#apparently ifelse doesn't retain class
+library(data.table)
+
+DF.HetC.MixedModel.mol$subsp <- if_else(grepl("MSM", DF.HetC.MixedModel.mol$strain), "Mol", 
+        if_else(grepl("MOLF", DF.HetC.MixedModel.mol$strain), "Mol", DF.HetC.MixedModel.mol$subsp))
+
+str(DF.HetC.MixedModel.mol)
+DF.HetC.MixedModel.mol$subsp <- factor(DF.HetC.MixedModel.mol$subsp, ordered = FALSE)
+
+#these remove ALL male MSM!
+table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#mean.co
+HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.mol)
+ranef(HetC.MixedModel.mol)
+
+#mreducing the model (reducing the random effects)
+Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
+summary(Reduced.strain.mol)
+ranef(Reduced.strain.mol)
+
+Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
+summary(Reduced.all.mol)
+
+#anova tests
+#full vs strain
+full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
+#full vs non
+full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
+#strain vs non
+strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
+
+anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
+
+#base vs mol model (this only makes sense if these are nested models)
+anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
+
+```
+
+Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed modles by strain.
+make articifial data / models, like human labels
+
+## Mixed Model for the variance
+
+```{r HQ.var, echo=FALSE}
+#modeling within mouse variance
+
+DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
+
+#HQ var.CO
+HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.HQ.var_co)
+ranef(HetC.MixedModel.HQ.var_co)
+
+HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+summary(HQjust.strain.var_co)
+
+HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+summary(HQfull.Reduced.cv_co)
+
+#contrast models
+full2strain.CV <- anova(HetC.MixedModel.HQ.var_co, HQjust.strain.var_co)#
+
+full2noRandom.CV <- anova(HetC.MixedModel.HQ.var_co, HQfull.Reduced.cv_co)#
+
+strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
+
+```
+
+```{r Base.Variance.MM, echo=FALSE, eval=FALSE}
+
+#there's 1 mouse with just 1 cell remove
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
+
+DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
+
+#main var.CO
+HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.var_co)
+ranef(HetC.MixedModel.var_co)
+
+#main cV.CO
+HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.cVco)
+ranef(HetC.MixedModel.cVco)
+
+just.strain.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain) 
+summary(just.strain.cV_co)
+
+full.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel)
+summary(full.Reduced.cv_co)
+
+#contrast models
+full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
+
+full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
+
+strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
+#for var, p = 0.0219
+
+```
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
+
+# check if the main mixed model results hold up when all cells are pooled
+
+#pooled.cells
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
+
+DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
+
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+
+```
+
+- The MSM strain effec is 0, but the sex*strain for molossinus is 7 (this seems very high).
+This means there is 0 variance added to mean_CO to mean due to being MSM.
+Does this mean the model is very good at predicting MSM mean_CO.
+The fixed strain - sex effects are for musc are large.
+This will likely change with MOLF (male molf are low, so smaller fixed effect and )
+
+- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
+In the first model MSM has largest random effect, since it's so far from WSB female.
+
+-adding MOLF also doesn't have significant factors
+
+changed all to unordered factors.
+all the variables are coded as ordered factors.
+When the variables are kept factors, and unordered the names appear! yay!
+
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.
 
@@
 
-## SC.skel measures
-
-data has been collection for RWang's total SC. These results are consistant with the patterns seen in the bivalent level resutls. (total SC PWD > KAZ )
-
-```{r RWscSkel, echo=FALSE}
-#merge the sc.skel outputs with the master MLH1 measures
-#match up random names with original names
-
-#the skel output and scripts are in the SC_skel dir. it was run on all images in the anon_data folder. I haven't manually confirmed that the default parameters are skeletonizing the images correctly.
-
-#some of richard's analysis show negative relationship with SC length and MLH1 (his QTL word doc). Could this be sign
-# of pwd like regulation?
-
-setwd("~./")
-SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_sep19.csv", header = TRUE, strip.white = TRUE)
-
-SC.skel$rand_name <- as.character(SC.skel$rand_name)
-original_DF$Random.Name <- as.character(original_DF$Random.Name)
-#str(SC.skel)
-#str(original_DF)
-
-mlh1.skel.og <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-#for some reason this only worked after the chr and all = F
-
-mlh1.skel.og <- mlh1.skel.og[ !grepl("X", mlh1.skel.og$Batch) , ]
-mlh1.skel.og <- mlh1.skel.og[ !grepl("x", mlh1.skel.og$Batch) , ]
-mlh1.skel.og <- mlh1.skel.og[!(is.na(mlh1.skel.og$quality) | mlh1.skel.og$quality==""), ]
-
-mlh1.skel.og <- add_mouse(mlh1.skel.og)
-mlh1.skel.og <- add_strain(mlh1.skel.og)
-mlh1.skel.og <- add_sex(mlh1.skel.og)
-
-mlh1.skel.og <- add_category(mlh1.skel.og)
-mlh1.skel.og <- add_subsp(mlh1.skel.og)
-
-#mlh1.skel.og <- add_euth_date(mlh1.skel.og)
-#mlh1.skel.og <- add_age(mlh1.skel.og)
-
-#str(mlh1.skel.og)
-mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
-mlh1.skel.og$bin_size <- as.numeric(mlh1.skel.og$bin_size)
-mlh1.skel.og$nMLH1.foci <- as.numeric(mlh1.skel.og$nMLH1.foci)
-
-#remove outliers
-mlh1.skel <- mlh1.skel.og[mlh1.skel.og$skel_size < 25000,]
-#first test file has ~1000 measures
-#when I ran it, no sign of outliers (after removing all the X's cells)
-
-looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=quality) ) + geom_jitter() + facet_wrap(sex~strain)
-
-looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Musc",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
-#very possible that the dimesions and pixel density are skrewing with these values
-# I am suspcious that there is such a wide range for the female data
-looky
-
-looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
-looky.DOM
-
-#mouse level
-mouse.level.skel <- ddply(.data=mlh1.skel.og, 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(nMLH1.foci),
-                 mean.skel = mean(skel_size)
-)
-
-mouse.level.skel <- add_strain(mouse.level.skel)
-mouse.level.skel <- add_sex(mouse.level.skel)
-mouse.level.skel <- add_category(mouse.level.skel)
-mouse.level.skel <- add_subsp(mouse.level.skel)
-
-looky.mouse <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Musc",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)
-
-looky.mouseDom <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Dom",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)
-
-#theme(legend.position="none")
-
-#double check a way to visually inspect the python skel output
-#
-#repeat random numbers?
-```
-
-```{r test.BivData, echo=FALSE}
-#bivdata
-setwd("~./")
-MSM_BivData.og = read.csv("~./MLH1repo/data/BivData/Curate_MSMmale_9.9.19.csv", header = TRUE, strip.white = TRUE )
-
-MSM_BivData <- MSM_BivData.og[MSM_BivData.og$SC.pass ==1,]
-MSM_BivData <- MSM_BivData[!(is.na(MSM_BivData$hand.foci.count) | MSM_BivData$hand.foci.count==""), ]
-
-#MSM_BivDatam_CO2 <- MSM_BivData[MSM_BivData$hand.foci.count == 2,]
-#MSM_BivDatam_CO2 <- MSM_BivDatam_CO2[!(is.na(MSM_BivDatam_CO2$fileName) | MSM_BivDatam_CO2$fileName==""), ]
-
-#MSM_BivDatam_CO2$Foci1.PER <- MSM_BivDatam_CO2$Foci1 /MSM_BivDatam_CO2$chromosomeLength
-#MSM_BivDatam_CO2$Foci2.PER <- MSM_BivDatam_CO2$Foci2 /MSM_BivDatam_CO2$chromosomeLength
-
-MSM_BivData <- add_IFD(MSM_BivData)
-#add IFD2 is f-ing up 
-MSM_BivData <- addIFD.2(MSM_BivData)
-
-#this MSM PER.IFD, is similar to the PWD data
-
-#new function is converting all hand.foci counts to 3
-#it might be the IFD functions
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-new_shiny <- add_SisCoTen(MSM_BivData)
-
-new_shiny$rel_Siscoten <- new_shiny$SisCoTen / new_shiny$chromosomeLength
-
-qq <- ggplot(data = new_shiny, aes(y=chromosomeLength, x=rel_Siscoten) ) + geom_jitter()
-
-#ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
-#  geom_boxplot(aes(alpha=.3))
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#KAZ and PWD data set is larger
-PWD_KAZ_BivData.og = read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD.KAZ_BivData.csv", header = TRUE, strip.white = TRUE )
-
-PWD_KAZ_BivData <- PWD_KAZ_BivData.og[PWD_KAZ_BivData.og$SC.pass == 1,]
-
-PWD_KAZ_BivData <- add_IFD(PWD_KAZ_BivData)
-PWD_KAZ_BivData <- addIFD.2(PWD_KAZ_BivData)
-PWD_KAZ_BivData <- add_SisCoTen(PWD_KAZ_BivData)
-PWD_KAZ_BivData$rel_Siscoten <- PWD_KAZ_BivData$SisCoTen / PWD_KAZ_BivData$chromosomeLength
-PWD_KAZ_BivData <- add_mouse(PWD_KAZ_BivData)
-PWD_KAZ_BivData <- add_strain(PWD_KAZ_BivData)
-
-mouse_boxes <- ggplot(data = PWD_KAZ_BivData, aes(y=rel_Siscoten, x=mouse) ) + geom_jitter(aes( color=fileName))+
-  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")
-
-mouse_boxes2 <- ggplot(data = PWD_KAZ_BivData, aes(y=SisCoTen, x=mouse) ) + geom_jitter(aes( color=fileName))+
-  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")
-
-#mean relative and raw sisco-ten is higher in PWD compared to KAZ
-
-biv_boxes <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count, fill=hand.foci.count) ) + geom_jitter()+geom_boxplot(aes(x= hand.foci.count,alpha=.3))+
-theme(legend.position="none")+facet_wrap(~strain)
-
-biv_boxes22 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=as.factor(hand.foci.count)) )+ geom_jitter()+
-  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")+facet_wrap(~strain)
-
-biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( ))+
-theme(legend.position="none")+facet_wrap(~strain)
-
-regress <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=chromosomeLength) ) + geom_jitter(aes( color=as.factor(hand.foci.count) ))+ facet_wrap(~strain)
-
-regress.1 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
-#, 1mar15_PWD_m1
-
-regress.2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
-#, 1mar15_PWD_m1
-
-#
-
-regress.1.KAZ <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$mouse == "21may18_KAZ_m2",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
- theme(legend.position="none")+ facet_wrap(~fileName)
-
-#doesn't look like the sis-co-ten is conserved across bivs
-
-```
-
-```{r within.cell.corr, echo=FALSE}
-
-#test telomere distance
-#centromere distance
-#SC length
-
-#I want to test if SC lengths within cells cluster
-#wouldn't ANOVA get at this..?
-
-#cell level variance is significant predictor for SC length (telomere)
-
-#PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",]
-jjgg <- ddply(.data= PWD_KAZ_BivData,
-                 .(fileName),
-                 summarize, 
-                 nbiv = length(unique(Obj.ID) ),
-                  var.sc = var(chromosomeLength)
-)
-
-#reasons why this within cell correlation doesn't work well with bivalent data
-  #different sets of chrms
-#not enough resolution at cytology level
-
-```
-
-sis-coten metric:
-I am currently missing a null model / test for this metric
-(sis-co-ten from uniform placement of COs, and (some prob of draw for CO number))
-
-there is a positive correlation with siscoten and chrm length
-
-there could be something about kinetochore sites differing across specific chromosomes that balance the force
+# References

### 2019-11-12 doc/MouseLevelPlots.Rmd
updated the mouse level count by quality plots. started list of mice to exclude

--- 2019-08-30 doc/MouseLevelPlots.Rmd
@@
 author: "April Peterson"
-date: "August 29, 2019"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 output: html_document
@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")#batch15 isn't showing up
+load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
+#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
+#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")#batch15 isn't showing up
 #load(file="data/MetaData.RData")#this Rdata was replacing the most current batch.. agg!
@@
 knitr::opts_chunk$set(echo = TRUE)
+
+#remove other!
+MLH1_data <- MLH1_data[MLH1_data$strain != "other",]
+
+MLH1_data <- droplevels(MLH1_data)
 ```
@@
 
+Yellow colored boxes are mice that will be exclude from the final (HQ) dataset for low quality cells (5).
+Consider re-runing this with the final data frame to see how the strain and mouse means change.
+
 ### Updating plots
@@
 
-These MLH1 data still have quality 5 cells observations. I might want to redo it without those observations.
-
 ```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE}
@@
 
-#mice that don't have female data will break the chain-loop
-#MOLF female, HMI, CZECH female, CAROLI female, Tom male
+exclude_mice <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
+                  "8jun15_G_f3", "8jun15_G_f4",
+                  "4jan17_LEW_f2",
+                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
+                  "12sep16_MSM_f3",
+                  "27oct18_MOLF_m3",
+                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
+                  "27aug18_CZECH_m1","8jun18_CZECH_m1",
+                  "3sep18_AST_m1",
+                  "3jel18_TOM_m1","3sep18_TOM_m1",
+                  "31jul17_HMI_m1",
+                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
+                  )
 
-#instead of adding fake data, remove the categories from list
+#based on this plot make a list of mice to exclude, and figure out code to outline in red
+#mostly excluding based on number of 5quality cells
 
-#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
-#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
-#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
-#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
-#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
-#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
-#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
-#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
-#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
+#mice.to.exclude <-
 
-pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
+#WSB; "10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
+#G;  "8jun15_G_f3", "8jun15_G_f4"
+#LEW; "4jan17_LEW_f2"
+#PWD; "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5"
+#MSM; "12sep16_MSM_f3",#"14jul17_MSM_f1"
+#MOLF; "27oct18_MOLF_m3"
+#SKIVE; "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3"
+#CZECH; "27aug18_CZECH_m1","8jun18_CZECH_m1"
+#AST: "3sep18_AST_m1"
+#TOM; "3jel18_TOM_m1","3sep18_TOM_m1"
+#HMI; "31jul17_HMI_m1"
+#SPIC; "21aug17_SPI_f1", "30may18_SPIC_f1", #"30oct17_SPIC_m1", "8jun18_SPIC_m2",
 
-#this loop makes a quant_status col for a list of mice
+#these plots are filled in yellow, but a different alpha/transparency is being applied - I don't know why
+
+#this loop makes a quant_status col for a list of mice (THERE'S AN ARGUMENT FOR Drawing a box around the mouse blocks )
+#switch this loop to ex
 for(i in 1:length(MLH1_data$mouse)){
-  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], exclude_mice), "exclude", NA)
 }
+
+#CZECH is not in this list!!!
 
@@
 #this is where you remove the non-data data!
-df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
+df_list <- df_list[ -c(24, 29) ]
 
-#7 PERC, 12 MOLF female, 
+#12 MOLF female, 18 CZECH female?, 24 HMI female, 29 CAROLI female? 31 other
 #
@@
 
-MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
+#MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
 
@@
   pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
-    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
-    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
+    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + xlim(c(0.5,5.5))+
+
+    #put a box over strain plots based on the col, 'quant_status'
+      geom_rect(data = df_list[[h]], aes(fill= quant_status), alpha= 0.045, xmin=-100, xmax=100, ymin=-100, ymax=100)+
+
+      scale_fill_manual(values=c("yellow"), breaks=NULL) +
+
     facet_wrap(~ mouse, drop = TRUE) + 
-    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
+
+      geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
   #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 

### 2019-11-13 doc/Main_Report_Results.rmd
added DMC1 results to results page

@@
 
+## MLH1 means
+
+figure 1
+mixed models (ummary of mixed model, justificiation for going into post hoc
+
 ```{r first.boxplots, echo=FALSE, fig.align='center'}
@@
 
+### Mixed Model for MLH1 counts
+
+### Sex Specific Evolution
+
+#### Male Patterns
+
+rapid evolution in musculus and molossinus; high and low CO count males
+
+#### Female Patterns
+
+seems like female specific evolution in molossinus and musculus.
+
+more variation in domesticus females
+
+## Within Mouse Variation in MLH1 counts
+
+#DMC1 estimates
+
+to localize the evolution in the pathway
+
+  figure
+  ANOVA for L and Z
+  t-tests betwee high and low groups
+
+```{r DMC1_block1, echo=FALSE}
+#loadt DMC1 csv
+
+DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+In ANOVA analysis strain is significant for Leptotene cells but not Z cells.
+
+```{r DMC1_t.tests, echo=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#L vs Z (all species)
+#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
+#p-value = 1.032e-05
+#not a valid interpertation
+
+#high vs low group (L)
+zola <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+jesus <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+#cheer <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "L")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+#up <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+#dont <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], 
+#      DMC1.male.data$total.count[(DMC1.male.data$strain == "KAZ") & (DMC1.male.data$stage == "Z")] )
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+The p values from the t.tests of the high vs low groups, indecate that the high recombining are significantly higher for the L cells (pvalue = `r zola$p.value`) while the Z cells, there is not a significant difference across the high and low groups (pvalue = `r jesus$p.value`).
+
+```{r DMC1.subsampling, echo=FALSE}
+
+#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
+# the 2 groups
+
+#make the 2 pools
+High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+
+Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+
+#permute drawing / subsampling and performing the t-test
+#put the p.value results into a DF!
+
+#permut.SC.cv.diff  <- as.data.frame(
+#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
+
+carry.on <-   replicate(10, 
+ sample(High.L.pool, 25, replace = FALSE) )#this makes 10x9 DF from the samples
+
+bunchOpvals.L <- replicate(10, 
+    # t.test
+    t.test( sample(High.L.pool, 25, replace = FALSE),
+            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
+          )
+
+bunchOpvals.Z <- replicate(10, 
+    # t.test
+    t.test( sample(High.Z.pool, 25, replace = FALSE),
+            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
+          )
+
+colnames(permut.SC.cv.diff) <- "permuted.diff"
+
+#ASSUMPTIONS; normality, equal variance
+
+#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
+#n use the var.equal = TRUE
+#https://www.statmethods.net/stats/ttest.html
+```
+
+```{r DMC1.plot, echo=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+library(ggpubr)
+#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("G","WSB","KAZ",
+              "PWD","MSM"), order=T )
+#order the data frame
+DMC1.DF <- with(DMC1.DF, DMC1.DF[order(strain),])
+
+#for the test bars
+my_comparisons <- list( c("KAZ", "MSM"))#no idea how to adjust
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+
+
+  #this will add neat bars, but still need to figure out what groups ect it's actually using
+  stat_compare_means(comparisons = my_comparisons)
+
+#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+#        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.boxplot
+```
+
+> At what stage are DMC1 foci nnumbers most correlated with MLH1?
+
+Leptotene is the max number of DMC1 -- so this would have more power to capture the variation in MLH1? Look into what baier et al do.
+
+## Heterochiasmy
+
+- Xy xx  (female vs male adjustment)
+-	Chrm proportions (
+
 ```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
@@
 
-### Difference in proportion of chrm classes
+### Proportion of chrm classes
 

### 2019-11-25 doc/DMC1_report.Rmd
drafted results for DMC1 report

---
title: "DMC1 report"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file

load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
#load(file="data/MetaData.RData") #this was replacing most updated data file

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#loat DMC1 csv
DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

```

#DMC1 Results

ToDo:

  - attempt to supplement data with more KAZ DMC1 cell observations.
  - change all the bad variable names

```{r DMC1.plot, echo=FALSE, warning=FALSE}
#think of better color scheme
#make the jitter col thinner or make the stacked dots.
#remember that box-plots add outlier dots.
#change y labels

#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/

DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("G","WSB","KAZ",
              "PWD","MSM"), order=T )
#order the data frame
DMC1.DF <- with(DMC1.DF, DMC1.DF[order(strain),])

#for the test bars, used with STAT_COMPARE
my_comparisons <- list( c("KAZ", "MSM"))#no idea how to adjust

#experimenting with ggboplot
everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
          color = "stage", palette = "jco",
          add = "jitter", short.panel.labs = FALSE)
  #library gpubr has functions to add in sig/test bars.

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")

  #this will add neat bars, but still need to figure out what groups ect it's actually using
#THE stat_compare function adds a arm thing and pvalue for 
#  stat_compare_means(comparisons = my_comparisons)

#status - legend, expression y-axis, facet, facet_wrap SYMBOL
#legend - stage. y=total count, FACet strain
ck <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x = stage, y = total.count)) +  facet_wrap(~strain)+ geom_boxplot()+
  stat_compare_means(comparisons = list(c("PWD", "KAZ")))

# annotation table with adjusted pvals and y-position of the labels
anno_df = compare_means(total.count ~ stage, group.by = "strain", data = DMC1.DF[DMC1.DF$stage != "P",]) %>%
 mutate(y_pos = 320)

ck2 <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=stage), position=position_jitterdodge()) + 
  facet_wrap(~strain) + 
  ggsignif::geom_signif(
    data=anno_df, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )
#the plot above adds the arm for (t-test)? comparison of each stage by strain
#this is not quite what I want -- but I'm glad I learned it

#I want a test between 1) high vs low cell stages, 
#I think a better way to display would be
#facet by stage, x is the strain
#compare means across high vs low  (this level isn't defined)

#example from here
#https://github.com/kassambara/ggpubr/issues/65

#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#        panel.background = element_blank(), axis.line = element_line(colour = "black"))

DMC1.boxplot
```

Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days). One mouse represents each strain. The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
Since the total counts are high -- we assume normality and apply t.tests

how similar are G-WSB or PWD-MSM to each other?

ToDo think of ways to add in the compared means for the high and low groups (across cell stage)

## Testing Normality

- check that distributions are close to normal
- check for equal variance.

```{r normality, echo=FALSE}
#histograms for each

#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
#bartlet's test, Levine's test

#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z

```

From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells.

```{r DMC1_block1, echo=FALSE, warning=FALSE}
#1. ANOVA, is there a difference in DMC1 counts across strains?
mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])

l.anova <- anova(mod.L) #0.0002701 ***
z.anova <- anova(mod.Z) #NS

l.anova
z.anova
```

The two groups of mice will be split into high and low recombining groups. (to assess the patterns of variance across them)

```{r block2_DMC1_t.tests, echo=FALSE, warning=FALSE}
#2. t-tests for the DMC1 counts 
#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo

#L vs Z (all species)
#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
#p-value = 1.032e-05
#not a valid interpertation

#high vs low group (L)
ttest.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
# p-value =  0.002145

#high vs low group (Z)
ttest.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
# p-value = 0.6628

#dom vs musc (L) (2 subsp vs)
#cheer <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "L")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "L")] )
# p-value = 0.003016

#dom vs musc (Z)
#up <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "Z")] )
#p-value = 0.09588

#KAZ vs Dom (Z) (only 1 L cell)
#dont <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], 
#      DMC1.male.data$total.count[(DMC1.male.data$strain == "KAZ") & (DMC1.male.data$stage == "Z")] )
#p-value = 0.004148
#Kaz Z are higher than Dom
```

Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions. The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.Z$p.value`).

## Permutations with equal sampling

The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.

Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups. Permutations were done for both cell stages. The actual p value is in red.

The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.

```{r block3.DMC1.subsampling, echo=FALSE}

#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
# the 2 groups

#make the 2 pools
High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]

Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]

#permute drawing / subsampling and performing the t-test
#put the p.value results into a DF!

#permut.SC.cv.diff  <- as.data.frame(
#sample(THE.DF, SAMPLE.SIZE, REPLACE?)

#permuting samples -- and re-licating t-tests

#nrep <- 1000

#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
bunchOpvals.L <- replicate(1000, 
    # t.test
    t.test( sample(High.L.pool, 25, replace = FALSE),
            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
          )
bunchOpvals.L <- as.data.frame(bunchOpvals.L)
bunchOpvals.L$stage <- "L"

bunchOpvals.Z <- replicate(1000, 
    # t.test
    t.test( sample(High.Z.pool, 25, replace = FALSE),
            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
          )

bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
bunchOpvals.Z$stage <- "Z"

#could merge the lists of p-values
replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'

#ASSUMPTIONS; normality, equal variance

#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.

plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.L$p.value, linetype="dotted", color="red")

plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.Z$p.value, linetype="dotted", color="red")
plot.pvals.L
plot.pvals.Z

#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
#n use the var.equal = TRUE
#https://www.statmethods.net/stats/ttest.html
```

> At what stage are DMC1 foci numbers most correlated with MLH1?

Need to pull the strain average table

Leptotene is the max number of DMC1 so this would have more power to capture the variation in MLH1? Look into what baier et al do.

### 2019-11-25 doc/DMC1_report.Rmd
small updates to report

@@
 DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
 DMC1.DF <- add_subsp(DMC1.DF)
@@
 
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize, 
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+#DMC1.Table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by.x = "category")
+               # var.MLH1 = var(mean_co)
 ```
@@
 
-- check that distributions are close to normal
-- check for equal variance.
-
 ```{r normality, echo=FALSE}
 #histograms for each
+
+#- check that distributions are close to normal
+#- check for equal variance.
 
@@
 
-From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells.
+From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
 
@@
 #high vs low group (L)
-ttest.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
                DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
@@
 #high vs low group (Z)
-ttest.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
                DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
@@
 #dom vs musc (L) (2 subsp vs)
-#cheer <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "L")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "L")] )
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
 # p-value = 0.003016
@@
 #dom vs musc (Z)
-#up <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "Z")] )
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
 #p-value = 0.09588
@@
 #KAZ vs Dom (Z) (only 1 L cell)
-#dont <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], 
-#      DMC1.male.data$total.count[(DMC1.male.data$strain == "KAZ") & (DMC1.male.data$stage == "Z")] )
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
 #p-value = 0.004148
@@
 Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.Z$p.value`).
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
 
@@
 
-plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.L$p.value, linetype="dotted", color="red")
+plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
 
-plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.Z$p.value, linetype="dotted", color="red")
+plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
 plot.pvals.L
@@
 
+### DMC1 Correlation with MLH1
+
+```{r MLH1.cor, echo=FALSE}
+#make category averages from AP_mouse_table.HQ
+mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+#attach this mini DF to the DMC1 data
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+```
+
 > At what stage are DMC1 foci numbers most correlated with MLH1?
 
-Need to pull the strain average table
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
 
-Leptotene is the max number of DMC1 so this would have more power to capture the variation in MLH1? Look into what baier et al do.
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.

### 2019-11-25 doc/Main_Report_Results.rmd
added DMC1 section, improved clarity, removed clutter

--- 2019-11-12 doc/Main_Report_Results.rmd
@@
 
+#Setup
+
 ```{r setup, echo=FALSE, include = FALSE}
@@
 library(RLRsim)
+library(ggpubr)
+library(reshape2)
 
@@
   mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
 MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
@@
 ```
+
+1. The quality of MLH1 counts for individual mice is / was assessed in the MouseLevelPlots Rmd (can I link this file?). A list, mice.to.exclude was made.
+
+2.
+The list of excluded mice is applied in the MLH1.setup.data script, to remove these mice from the main dataset/dataframe.
+This script saves .RData file of the dataset which is loaded into Rmd files for analysis
+
+Mouse specific plots of CO count and quality were examined for general quality from single mice.
+Mice which produced few cells of low quality were excluded to the 'HQ' (high quality) dataset.
+(The full dataset was used to look at some patterns)
 
@@
 
-### Distributions of MLH1 counts
+# Distributions of MLH1 counts across Samples
+
+The histograms below are for total / pooled cells (for now, maybe compare with HQ data to see what changes)
 
@@
 
-#  Results
+# MLH1 means
+
+figure 1, change to all white, add the subspecies lines below.
 
@@
 
+#### outgroups
+
+The outgroups (SPIC and SPRET) don't show strong evidence of evolution in gwRR.
+
+#### Male Patterns
+
+Rapid evolution in musculus and molossinus; high and low CO count males (two instances of rapid gwRR evo, PWD and MSM)
+
+#### Female Patterns
+
+seems like female specific evolution in molossinus and musculus.
+more variation in domesticus females.
+Females have more nuanced patterns, there is strain specific differences.
+
+# DMC1 Results
+
+These data are meant to test if CO precursors, double strand breaks (DSBs), are significant predictors for MLH1 count variation. Simply, do mice with more DSBs also have more COs?
+
+ToDo:
+  - attempt to supplement data with more KAZ DMC1 cell observations.
+  - change all the bad variable names
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize, 
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+#these tables are merged latter
+
+```
+
+Potentially figure 2. Using a function to display the p-vales for mean comparison
+
+```{r DMC1.plot, echo=FALSE, warning=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")
+  DMC1.boxplot
+  #this will add neat bars, but still need to figure out what groups ect it's actually using
+#THE stat_compare function adds a arm thing and pvalue for 
+#  stat_compare_means(comparisons = my_comparisons)
+
+#status - legend, expression y-axis, facet, facet_wrap SYMBOL
+#legend - stage. y=total count, FACet strain
+
+#add column to distinguish high vs low rec strains
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
+
+# annotation table with adjusted pvals and y-position of the labels
+anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",]) %>% mutate(y_pos = 300)
+
+DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=stage), position=position_jitterdodge()) + 
+  facet_wrap(~stage, nrow=1) + 
+  ggsignif::geom_signif(
+    data=anno_df, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")
+#this is kinda the plot I'm enivioning 
+DMC1.bxlt_High.to.Low
+
+#I want a test between 1) high vs low cell stages, 
+#I think a better way to display would be
+#facet by stage, x is the strain
+#compare means across high vs low  (this level isn't defined)
+
+#example from here
+#https://github.com/kassambara/ggpubr/issues/65
+
+#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
+
+```
+
+Above figure shows p-values for comparison of means across the cell stages.
+
+Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days).
+One mouse represents each strain.
+The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
+
+ToDo think of ways to add in the compared means for the high and low groups (across cell stage)
+how similar are G-WSB or PWD-MSM to each other?
+
+## Testing Normality Assumptions
+
+```{r normality, echo=FALSE, warning=FALSE}
+#histograms for each
+
+#- check that distributions are close to normal
+#- check for equal variance.
+
+DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
+
+DMC1.histogram
+#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
+#bartlet's test, Levine's test
+```
+
+The counts look close to normal.
+
+```{r barlett.test, echo=FALSE}
+
+#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
+bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
+
+```
+
+Evidence for non-equal variance across strains for zygotene cells.
+
+Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
+
+From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#L vs Z (all species)
+#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
+#p-value = 1.032e-05
+#not a valid interpertation
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
+
+## Permutations with equal sampling
+
+The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
+Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
+Permutations were done for both cell stages.
+The actual p value is in red.
+
+The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
+
+```{r DMC1.subsampling, echo=FALSE}
+
+#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
+# the 2 groups
+
+#make the 2 pools
+High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+
+Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+
+#permute drawing / subsampling and performing the t-test
+#put the p.value results into a DF!
+
+#permut.SC.cv.diff  <- as.data.frame(
+#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
+
+#permuting samples -- and re-licating t-tests
+
+#nrep <- 1000
+
+#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
+bunchOpvals.L <- replicate(1000, 
+    # t.test
+    t.test( sample(High.L.pool, 25, replace = FALSE),
+            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
+          )
+bunchOpvals.L <- as.data.frame(bunchOpvals.L)
+bunchOpvals.L$stage <- "L"
+
+bunchOpvals.Z <- replicate(1000, 
+    # t.test
+    t.test( sample(High.Z.pool, 25, replace = FALSE),
+            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
+          )
+
+bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
+bunchOpvals.Z$stage <- "Z"
+
+#could merge the lists of p-values
+replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
+#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
+
+#ASSUMPTIONS; normality, equal variance
+
+#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
+
+plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
+plot.pvals.L
+plot.pvals.Z
+
+#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
+#n use the var.equal = TRUE
+#https://www.statmethods.net/stats/ttest.html
+```
+
+```{r DMC1_MLH1.cor, echo=FALSE}
+#make category averages from AP_mouse_table.HQ
+mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+#attach this mini DF to the DMC1 data
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+```
+
+> At what stage are DMC1 foci numbers most correlated with MLH1?
+
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+
+# Note on Heterochiasmy Definition
+
+I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
+In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
+
+Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
+(I note if the and Y are paired, which they are at a high rate).
+The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
+
+To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
+
+1. Compile full-cell data from females (all 20 bivalents measured)
+2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
+3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
+
+According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
+
 ```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
@@
 
+```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
 #### Distributions by mouse
-
-There are code chunks that make a big dfs of CO count by quality score plot. The code has been commented out, and moved to it's own doc.
-
-```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
+#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
+
+#remove this block since it has it's own src/script
+
 ### to integrate 'pass'/fail
@@
 
-### Difference in proportion of chrm classes
-
-generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
+# Proportion of chrm classes
+
+These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes.
+Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs. (the gwRR for most mice is low/close to the minimum)
+
+This is also another way to score heterochiasmy (that gets around the sex chrms, i think)
+
+The data below are from hand measured cells (not from bivalent data)
+generate smaller data set of chrm proportions (0CO, 1CO,2CO etc)
 10 cells from each musc category. **Add Molf male**
 
+ToDo -- add more hand curated data for the bivalent proportions
+
 ```{r chrm_prop_data, echo=FALSE}
-
 chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
@@
 #MSM needs more bialent observations
-
 #use melt to put all propotions 
-
 #display / visualize / compare proportions
+#the table needs to be melted, because it's split by category
 
@@
 
-other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp)
+other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))
 
@@
 
+```{r chrm.pro.Curated.BivData, echo=FALSE}
+
+#bivalent data might be too big..
+load(file="~./MLH1repo/data/CleanBivData.RData")
+#Curates_BivData -- hand.foci.count!
+
+#For each mouse I need to count up
+#the number of bivalent classes are not counted up
+
+chrm.propotion.category <- ddply(.data=Curated_BivData, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(fileName)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = length(Obj.ID),
+
+                 n0CO = length(which(hand.foci.count == 0)),
+                  n1CO = length(which(hand.foci.count == 1)),
+                  n2CO = length(which(hand.foci.count == 2)),
+                  n3CO = length(which(hand.foci.count == 3)),
+                  #n4CO = sum(n4CO),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv )
+                # per4CO= sum(n4CO)/ tot_biv  )
+
+Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
+  "per0CO", "per1CO", "per2CO", "per3CO"), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
+
+# I think I need to make 2 melt tables -- for count and proportions
+Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
+
+Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
+colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
+
+Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
+
+Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
+
+Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
+ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
+      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
+        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
+        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
+        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
+        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
+
+Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
+                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
+#no mol yet
+
+Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
+
+#need to make list of positions
+
+chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
+
+chrm_prop.BivData_plot
+
+```
+
+number above bars are the raw counts for bivalent classes.
+
+Potential figure 2 (eventually). Try to make the figure taller.
+-- add more data to make sample sizes more equal.
+-- add space to indicate where female data is missing
+
+SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
+What's up with Lew females? there's lots of 2COs, also WSB female.
+
+This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
+If these data are random samples of bivalents, the X
+
+ToDo: chisq tests for difference in proportions, 
+- within strain across sex,
+- males, across high and low groups
+- look at mouse level proportions
+
 # Mixed Models
 
-todo: make model with strain as fixed effect (since, there are interesting strain effects)
-
-The mixed model lets us sperate effects to see how they effect a mouse mean CO count. This seperation can also serve as proxies for divergence and polymorphism.
+todo: re-run the analysis with MOLF data
+make model with strain as fixed effect (since, there are interesting strain effects)
+
+The mixed model lets us separate effects to see how they effect a mouse mean CO count. This separation can also serve as proxies for divergence and polymorphism.
 
@@
 ```{r mixed.models.setup, echo=FALSE}
-
 #make are the DFs for the models
@@
                         #quality?
-)#just 1 entry
+)
 
@@
 DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
-
-#sub-sample (need to think about this one)
-
-```
+```
+
+what are the table above this?
 
@@
 
-subspecies levels: 2, Musc and Domesticus (maybe Molossinus soon)
+subspecies levels: 2, Musc and Domesticus
 strain levels: Dom: 3, Musc: 4
@@
 `r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
-
-> Mouse average Models
 
@@
 #these remove ALL male MSM!
-table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
 
@@
 
-Moatly matches the base model results. the same fixed effects are significant in the reduced models. the random effects make sense
-
-Peromyscus did everything as factors... (but unordered)
+Reading these outputs is not the clearest -- try just printing out the coefficient table with p values.
+
+Mostly matches the base model results. the same fixed effects are significant in the reduced models. the random effects make sense
 
@@
 
-Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
-Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
-SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
-SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
-
 (what does the significant random test mean) -- variance is not 0?
@@
 #how did I make this 
+#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
+#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
+#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
+#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
 
@@
 Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
-Models with random effects more significant than nulls (both just strain and no radom effects).
+Models with random effects more significant than nulls (both just strain and no random effects).
 The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
@@
 
-Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed modles by strain.
-make articifial data / models, like human labels
+Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed models by strain.
+make artificial data / models, like human labels
 
@@
 
-- The MSM strain effec is 0, but the sex*strain for molossinus is 7 (this seems very high).
+- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
 This means there is 0 variance added to mean_CO to mean due to being MSM.
@@
 
-Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
 

### 2019-11-26 doc/Main_Report_Results.rmd
small updates

@@
 
+Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
+-ene endings refer to the cells/units within the cell stage? (I think)
+
 ```{r organizing.data, echo=FALSE }
@@
 #MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
+AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+
+AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
+                 .(strain,sex),
+                 summarize,
+                 sum.ncells = sum(Ncells),
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co),
+                 AV.var_MLH1 = mean(var, na.rm=TRUE)
+
+)
+
+write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
+
+AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
+AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+
+var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
+  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
 

### 2019-12-03 src/SetupData.Rmd
added the excluded mouse list, updating the setupdata results file

@@
 
-# First Steps
-
-```{r MLH1.DF.setup, warning=FALSE}
+```{r MLH1.DF.setup, warning=FALSE, echo=FALSE}
 
@@
 
+```
+
+```{r clean_MLH1_data, echo=FALSE}
 #Clean up main DF, MLH1_data
@@
 MLH1_data <- add_species(MLH1_data)
+```
+
+```{r exclude.mice, echo=FALSE}
+#remove mice 
+exclude_mice_list = c("18nov17_WSB_f1", "8oct14_WSB_f1", "8oct14_WSB_f3",
+                      "10mar15_WSB_m2",
+                      "8jun15_G_f3", "8jun15_G_f4",
+                      "4jan17_LEW_f2",
+                      "1apr15_PWD_f2", "8oct14_PWD_f5", "8oct14_PWD_f3", 
+                      "12sep16_MSM_f3",
+                      "27oct18_MOLF_m3",
+                   "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
+                      "27aug18_CZECH_m1", "8jun18_CZECH_m1",
+                      "3sep18_AST_m1","3sep18_TOM_m1","31jul17_HMI_m1",
+                      "21aug17_SPI_f1", "30may18_SPIC_f1","8jun18_SPIC_m2")
+
+MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]
+
+```
+
+```{r adj_nMLH1, echo=FALSE}
 
@@
 MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
+```
 
+```{r reorder.DF, echo=FALSE}
 #reorder dataframe
@@
 
- Settting up main data file with explanation and examination of data.  Make the raw original DF and main working DF with bad cells exlcuede.  (inline code add the numbers)
-
-There are `r length(original_DF_duplicates$fileName)` duplicated file names. These data will be used to quantify repeatability, and remove from the final DFs.
-Using the distinct() function for the MLH1_data, the first duplicated filename is kept and the second removed.
-
-```{r kable.table}
+```{r kable.table, echo=FALSE}
 #table comparing number of images in quantification batches
@@
 batch.stats <- merge(original_batches,final_batches, by="BatchNum")## mm is the table of batch stats
+
 ```
@@
 
-`r kable(final_batches, caption = "This is the table caption") ) %>% kable_styling()`
+`r kable(final_batches, caption = "Number of Images across all the anonimized Batches") %>% kable_styling()`
+
+# Quality Cleaning steps
+
+Settting up main data file with explanation and examination of data.  Make the raw original DF and main working DF with bad cells exlcuede.  (inline code add the numbers)
+
+There are `r length(original_DF_duplicates$fileName)` duplicated file names. These data will be used to quantify repeatability, and remove from the final DFs.
+Using the distinct() function for the MLH1_data, the first duplicated filename is kept and the second removed.
+
+Take out mice from the exclude list based on the output from Mouse level quality plots.
 
@@
 
-```{r mouse.table, echo=FALSE}
-#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+Make mouse level tables
+
+```{r mouse.table, echo=FALSE, warning=FALSE}
 
@@
 
-X number of mice are in the main DF (discribe the results for missing mice)
+ number of mice are in the main DF (discribe the results for missing mice)
 
@@
 
-```{r Age.Hists, echo=FALSE, warning=FALSE}
+```{r Age.Hists, echo=FALSE, warning=FALSE, eval=FALSE}
+#age.weeks col is missing
 
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_11.12.19.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
 

### 2019-12-03 src/setupdata_BivData_MLH1repo.R

@@ -68,15 +68,24 @@ fullBivData <- fullBivData[!(fullBivData$fileName %in% MultiCells), ]
 
 
 #mice with good stains
-good_stain_mice = c("20dec16_LEW_m1", "20dec16_LEW_m2", "20dec16_LEW_m3", "8may17_LEW_m1","8may17_LEW_m2",
-         "13nov16_LEWES_m1","13jan17_LEW_m1","13nov16_MSM_f1","13nov16_MSM_f2","14jul17_SPRET_f3",
-         "13nov16_MSM_m1","13nov16_MSM_m2","30may17_MSM_m1","30sep16_MSM_f2",  "30may17_SPRET_m1",  "22jun15_G_m2",
-         "30dec14_WSB_m2","17mar16_G_f4","17mar16_G_f5","4jan17_LEW_f1","4jan17_LEW_f3","4jan17_LEW_f6",
-         "4jan17_LEW_f7","4jan17_LEW_m1", "3jan16_G_m2")
+#good_stain_mice = c("20dec16_LEW_m1", "20dec16_LEW_m2", "20dec16_LEW_m3", "8may17_LEW_m1","8may17_LEW_m2",
+#         "13nov16_LEWES_m1","13jan17_LEW_m1","13nov16_MSM_f1","13nov16_MSM_f2","14jul17_SPRET_f3",
+#         "13nov16_MSM_m1","13nov16_MSM_m2","30may17_MSM_m1","30sep16_MSM_f2",  "30may17_SPRET_m1",  "22jun15_G_m2",
+#         "30dec14_WSB_m2","17mar16_G_f4","17mar16_G_f5","4jan17_LEW_f1","4jan17_LEW_f3","4jan17_LEW_f6",
+#         "4jan17_LEW_f7","4jan17_LEW_m1", "3jan16_G_m2")
 
-bad_stain_mice = c("13jan17_LEW_f3","12jun15_WSB_m1","12sep16_MSM_f3","14jul17_SPRET_f2","13jan17_LEW_f2",
-        "30may17_SPRET_f1","30dec14_WSB_m4","31jul17_HMI_m1","30dec14_WSB_m5","16jun15_WSB_f4","30dec14_WSB_m1",
-        "30dec16_MSM_m1","14jul17_LEW_f1","1mar17_CAST_f1")
+exclude_mice_list = c("18nov17_WSB_f1", "8oct14_WSB_f1", "8oct14_WSB_f3",
[…]

### 2019-12-03 src/SetupData.Rmd
re-ran with updated AnonData file

@@
 
-The original dataset contains the quantification batches
+All images were manually quantified after the file name was anonimized.
+The original (full) dataset contains the quantification batches with the following number of images.
+Each batch csv file was merged into Anon.csv by hand.
 
@@
 
-Settting up main data file with explanation and examination of data.  Make the raw original DF and main working DF with bad cells exlcuede.  (inline code add the numbers)
-
-There are `r length(original_DF_duplicates$fileName)` duplicated file names. These data will be used to quantify repeatability, and remove from the final DFs.
+The original dataset was cleaned up in the following ways before analysis.
+Settting up main data file with explanation and examination of data.
+Make the raw original DF and main working DF with bad cells exlcuede.
+The original dataset had `r length(original_DF_duplicates$fileName)` duplicated file names.
+These were removed and can be used to quantify repeatability or repair.
 Using the distinct() function for the MLH1_data, the first duplicated filename is kept and the second removed.
@@
 Take out mice from the exclude list based on the output from Mouse level quality plots.
+
+1. The quality of MLH1 counts for individual mice is / was assessed in the MouseLevelPlots Rmd (can I link this file?). A list, mice.to.exclude was made.
+
+2.
+The list of excluded mice is applied in the MLH1.setup.data script, to remove these mice from the main dataset/dataframe.
+This script saves .RData file of the dataset which is loaded into Rmd files for analysis
+
+Mouse specific plots of CO count and quality were examined for general quality from single mice.
+Mice which produced few cells of low quality were excluded to the 'HQ' (high quality) dataset.
+(The full dataset was used to look at some patterns)
 
@@
 ```{r save.RData, echo=FALSE}
-
 save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
-
 #  OutPut: big large MLH1_data (AP's) df, big DF of BD with just the mice I want.

### 2019-12-05 results/Chapter3_ideas.Rmd
started writing permutations for in silico nuc oermutations

---
title: "Chapter3_ideas"
author: "April Peterson"
date: "November 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
#load biv data
#load MLH1 data

```

Take some mouse level distributions -- pull cells from top and bottom 10%?
(gather whole bivalent level data,)

-start making comparisons for the 2 pools
  how much technical noise is there? (staining differences?/ quality score?

Mice with many cells and BivData
17mar16_G_f4 - BivData
17mar16_G_f1

```{r choosing.mice, echo=FALSE}
#Choose mice with many cells

#G_f, G_m
#SKIVE_f, SKIVE_m #have BivData

#remove all the quality 5

#make plot/histogram of the MLH1 count distribution
counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]

mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]

ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
topsy <- ordered.MLH1[1:5,]
bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
top.n.bottom.bin <- rbind(topsy,bottomzy)

#write 
write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)

shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz
#get the filenames for this groups of cells

#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
#28feb15_PWD_f2, in PWD--female .xlsx file

#31may19_SKIVE_m2
male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
male.one.DF <- male.one.DF[male.one.DF$quality < 5,]

ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
male.high.bin.DF <- ordered.MLH1.male[1:5,]
male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
#now have top and bottom 5 cell observations

#merge
male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
#write 
write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)

shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
#highest of the low bin
#lowest value of the high bin
shiz.too

```

the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
Should I be comparing cells with the same number of COs?

```{r getting.cell.names, echo=FALSE}

#load in the cell dataset
#28feb15_PWD_f2  from Curated_BivData
#for a mouse, pull the 10 highest rec cells

PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
#"Foci3"                   "curation.notes"  
female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]

write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)

PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")

#merge with hand data
#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)

cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
                            .(fileName),
                            summarize, 
                            ncells = length(unique(fileName)),
                            nbivs = length(unique(Obj.ID)),
                            n.foci = sum(hand.foci.count),
                            total.SC = sum(chromosomeLength),
                            mean.SC = mean(chromosomeLength)

)
#add hand data
ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]

#I have 3 cells with full 20

```

Start discribing the Bivalent patterns for the PWD female patterns

```{r mini.biv, echo=FALSE}
#make mini biv plots

#PWD.female
#PWD.female_hand.measure

cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]

two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))

#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
#of SC.
MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
 #geom_hline(yintercept = c(male_IFD) ) +

  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
  #how to add points for 2foci
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
  #foci.1.position, two_groups$distal.foci.position)
  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +

#add centromeres...  
  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+

  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
 # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
  ylim(c(0,200))+
     theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())  #for removeing x labels

#MiniBiv_plot1
#MiniBiv_plot2

#save the files

```

```{r biscribe.Biv.Pattern, echo=FALSE}
#start discribing the basic

#start merging curated bivData with handmeasured data

#BivData - Curated
```

# Whole Cell Measures

excel sheet Whole_cell_measures.xlsx has 840 bivalent measures.

colum

```{r whole.cell.data, echo=FALSE}

#read in csv of whole cell hand measures
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)

whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$fileName) | whole.cell.data$fileName==""), ]

whole.cell.data.XY <- whole.cell.data

whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

#find cells with the wrong number

biv.count_table <-  ddply(.data=whole.cell.data,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length)
)
#all have 20 or 19 (after removing XY)

for(musy in as.character(unique(whole.cell.data$fileName))){
  print(musy)

  #order the chrm lengths -- this is order
  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
}
#I think I saved the Excel with sorted SC lengths
#Yay I printed the rank in this

```

Try doing the permutations

```{r permutations, echo=FALSE}

#need to seperate sexes cause diff numbers
male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
#ncells
length(unique(male.whole.cell$fileName) ) #24 cells

female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456

#not sure how to make the initial DF
#replicate 24 rows with 19 cols
DF.for.perm  <- as.data.frame( replicate(24, "") )
colnames(DF.for.perm) <- c("blah")

DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),
                       rank3= replicate(24, ""),     rank4= replicate(24, ""),
                       rank5= replicate(24, ""),   rank6= replicate(24, ""),
                       rank7= replicate(24, ""), rank8= replicate(24, ""),
                       rank9= replicate(24, ""), rank10= replicate(24, ""),
                       rank11= replicate(24, ""),     rank12= replicate(24, ""),
                       rank13= replicate(24, ""),     rank14= replicate(24, ""),
                       rank15= replicate(24, ""),     rank16= replicate(24, ""),
                       rank17= replicate(24, ""),rank18= replicate(24, ""),rank19= replicate(24, "")
                       )

iterations = 24
variables = 20
DF.permutations <- matrix(ncol=variables, nrow=iterations)
colnames(DF.permutations) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
                               "rank6","rank7","rank8","rank9","rank10","rank11",
                               "rank12","rank13","rank14","rank15","rank16",
                               "rank17","rank18","rank19" )

#ROW IS FIRST

for(row.num in 1:24){

  #print(c(row.num, "rowis") )

  for(col.num in 2:ncol(DF.permutations) ){

  #print(c(col.num, "colis") )

 DF.permutations[row.num,col.num] <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)

      }

  }
#yes this finally works!!!!

```

### 2019-12-05 results/Chapter3_ideas.Rmd
work on within nuc permutations

@@
 #load MLH1 data
+library(dplyr)
+library(plyr)
+library(ggplot2)
 
@@
 
-```{r choosing.mice, echo=FALSE}
+```{r choosing.mice, echo=FALSE, eval=FALSE}
 #Choose mice with many cells
@@
 
-```{r getting.cell.names, echo=FALSE}
+```{r getting.cell.names, echo=FALSE,eval=FALSE}
 
@@
 
-```{r mini.biv, echo=FALSE}
+```{r mini.biv, echo=FALSE,eval=FALSE}
 #make mini biv plots
@@
 
-colum
-
 ```{r whole.cell.data, echo=FALSE}
@@
 #read in csv of whole cell hand measures
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 #setwd("C:/Users/April/Desktop/MLH1repo")
@@
 whole.cell.data.XY <- whole.cell.data
+whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
 
-whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
+#make sure foci.count is number
+whole.cell.data$hand.foci.count <- as.character(whole.cell.data$hand.foci.count)
+whole.cell.data$hand.foci.count <- as.numeric(whole.cell.data$hand.foci.count)
 
@@
 
+#ADD THE RANK
 for(musy in as.character(unique(whole.cell.data$fileName))){
@@
 
-```{r permutations, echo=FALSE}
+```{r permutations.sc, echo=FALSE}
 
-#need to seperate sexes cause diff numbers
 male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
@@
 female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
+length(unique(female.whole.cell$fileName) ) #18 cells
+
+male.biv.count_table <-  ddply(.data=male.whole.cell,
+                 .(fileName),
+                 summarize,
+                 nbiv = length(SC.length),
+                 mean_co = mean(hand.foci.count),
+                 mean_sc = mean(SC.length),
+                   var_co = var(hand.foci.count),
+                   var_sc = var(SC.length)
+)
 
@@
 #replicate 24 rows with 19 cols
-DF.for.perm  <- as.data.frame( replicate(24, "") )
-colnames(DF.for.perm) <- c("blah")
+iterations = 24
+variables = 20
+DF.permutations.sc <- matrix(ncol=variables, nrow=iterations)
+colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
+                               "rank6","rank7","rank8","rank9","rank10","rank11",
+                               "rank12","rank13","rank14","rank15","rank16",
+                               "rank17","rank18","rank19" )
 
+#ROW IS FIRST
+for(row.num in 1:24){
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations.sc) ){
+  rand.rank.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations.sc[row.num,col.num] <- rand.rank.sc
+   }
+  }#yes this finally works!!!!
+DF.permutations.sc <- as.data.frame(DF.permutations.sc)
+
+#pair with real within cell SC variance
+for( o in 1:nrow(DF.permutations.sc) ) {
+  DF.permutations.sc$var_sc[o] <- var(as.numeric(DF.permutations.sc[o,2:20]))
+  }
+
+#permutations without accounting for rank
+its = 24
+varis = 20
+DF.permutations.no_rank <- matrix(ncol=varis, nrow=its)
+#colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
+#                               "rank6","rank7","rank8","rank9","rank10","rank11",
+#                               "rank12","rank13","rank14","rank15","rank16",
+#                               "rank17","rank18","rank19" )
+
+for(row.num in 1:24){
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations.no_rank) ){
+  rand.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations.no_rank[row.num,col.num] <- rand.sc
+   }
+  }#yes this finally works!!!!
+DF.permutations.no_rank <- as.data.frame(DF.permutations.no_rank)
+
+for( o in 1:nrow(DF.permutations.sc) ) {
+  DF.permutations.no_rank$var_rand.sc[o] <- var(as.numeric(DF.permutations.no_rank[o,2:20]))
+  }
+
+#merge with real
+sc.var.for.plot <- matrix(ncol=3, nrow=72)
+sc.var.for.plot <- as.data.frame(sc.var.for.plot)
+colnames(sc.var.for.plot) <- c("cell.id","var","type")
+
+sc.var.for.plot$var[1:24] <- DF.permutations.sc$var_sc
+sc.var.for.plot$type[1:24] <- "sim"
+
+sc.var.for.plot$var[25:48] <- male.biv.count_table$var_sc
+sc.var.for.plot$type[25:48] <- "real"
+
+sc.var.for.plot$var[49:72] <- DF.permutations.no_rank$var_rand.sc
+sc.var.for.plot$type[49:72] <- "sim_no_rank"
+
+#plot the forms
+#plot within cell variance
+whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
+whole.cell.var_SC.plot
+```
+
+```{r per.co.count, echo=FALSE}
+
+iterations = 24
+variables = 20
+DF.permutations_CO.count <- matrix(ncol=variables, nrow=iterations)
+colnames(DF.permutations_CO.count) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
+                               "rank6","rank7","rank8","rank9","rank10","rank11",
+                               "rank12","rank13","rank14","rank15","rank16",
+                               "rank17","rank18","rank19" )
+
+for(row.num in 1:24){
+
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations_CO.count) ){
+
+  rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations_CO.count[row.num,col.num] <- rand.rank.co
+   }
+  }
+DF.permutations_CO.count <- as.data.frame(DF.permutations_CO.count)
+
+#no rank - iterate / permute the sampling X times
+its = 24
+varis = 20
+DF.permutations.CO.no_rank <- matrix(ncol=varis, nrow=its)
+
+#make LIST of X DATAFRAMES
+
+for(row.num in 1:24){
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations.CO.no_rank) ){
+  rand.co <- sample(male.whole.cell$hand.foci.count, 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations.CO.no_rank[row.num,col.num] <- rand.co
+   }
+  }#yes this finally works!!!!
+DF.permutations.CO.no_rank <- as.data.frame(DF.permutations.CO.no_rank)
+
+#
+for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
+  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
+  }
+
+#merge with real
+co.var.for.plot <- matrix(ncol=3, nrow=72)
+co.var.for.plot <- as.data.frame(co.var.for.plot)
+colnames(co.var.for.plot) <- c("cell.id","var","type")
+
+co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
+co.var.for.plot$type[1:24] <- "sim"
+
+co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
+co.var.for.plot$type[25:48] <- "real"
+
+co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
+co.var.for.plot$type[49:72] <- "sim_no_rank"
+
+#plot
+whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
+whole.cell.var_CO.plot
+
+```
+
+```{r permuted.v.real, echo=FALSE}
+#evaluating the permuted 
+library(plyr)
+
+for( o in 1:nrow(DF.permutations_CO.count) ) {
+  DF.permutations_CO.count$var_CO.count[o] <- var(as.numeric(DF.permutations_CO.count[o,2:20]))
+  }
+
+#two data frames of within cell var
+#cell.id, var, type
+var.for.plot <- matrix(ncol=3, nrow=48)
+var.for.plot <- as.data.frame(var.for.plot)
+colnames(var.for.plot) <- c("cell.id","var","type")
+
+var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
+var.for.plot$type[1:24] <- "sim"
+
+var.for.plot$var[24:48] <- male.biv.count_table$var_co
+var.for.plot$type[24:48] <- "real"
+
+#my permuted DFs are in a strange str
+
+#plot within cell variance
+whole.cell.var.plot  <- ggplot(var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell variance between types")
+
+whole.cell.var.plot
+```
+
+```{r junk, eval=FALSE}
 DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),
@@
                        )
-
-iterations = 24
-variables = 20
-DF.permutations <- matrix(ncol=variables, nrow=iterations)
-colnames(DF.permutations) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
-                               "rank6","rank7","rank8","rank9","rank10","rank11",
-                               "rank12","rank13","rank14","rank15","rank16",
-                               "rank17","rank18","rank19" )
-
-#ROW IS FIRST
-
-for(row.num in 1:24){
-
-  #print(c(row.num, "rowis") )
-
-  for(col.num in 2:ncol(DF.permutations) ){
-
-  #print(c(col.num, "colis") )
-
- DF.permutations[row.num,col.num] <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
-
-      }
-
-  }
-#yes this finally works!!!!
-
 ```

### 2019-12-05 results/Chapter3_ideas.Rmd
added in data to start recreating Wang et al findings

@@
 author: "April Peterson"
-date: "November 26, 2019"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 output: html_document
@@
 library(ggplot2)
+library(reshape2)
+library(raster) #cv
+
+#read files
+whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
+load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
 
 ```
+
+# Summary and recreation of Wang et al 2019
+
+insert summary of the Wang paper
+
+which data did they use? what summary stat were they comparing
+What were their results?
+
+human data- 
+Gruhn et al 2013 (ploted below)
+Sun et al 2006
+Oliver-Bonet et al 2007 
+Lian et al 2008
+Lu et al 2012
+Hou et al 2013
+
+```{r human.male.data, echo=FALSE}
+#read files
+Hs.male_axis <- read.csv("~./MLH1repo/data/Human_male_axis_length.csv", header = TRUE)
+Hs.male_MLH1 <- read.csv("~./MLH1repo/data/Human_male_MLH1.csv", header = TRUE)
+#I assume that the rows of MLH1 and SC lengths are the linked observations, check or verify this assumption
+#if they are using within cell -- they must
+
+#melt DF -- how do I keep row.name info?
+Melted.Hs.male_axis <- melt(Hs.male_axis, 
+                            c("row.name", "chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22"),
+                         id=c("row.name") )
+#remove the first part with of data frame, variable = "row.name""
+
+#clean up dm1 (there are a bunch of rows without useful information)
+Melted.Hs.male_axis <- Melted.Hs.male_axis[(!is.na(Melted.Hs.male_axis$value)),]
+Melted.Hs.male_axis <- Melted.Hs.male_axis[((Melted.Hs.male_axis$variable != "row.name")),]
+
+colnames(Melted.Hs.male_axis) <- c("row.name", "Chrm.id", "axis.length")
+
+#Seems like there is an outlier, 
+Melted.Hs.male_axis <- Melted.Hs.male_axis[(Melted.Hs.male_axis$axis.length < 700 ),]
+
+hs.sc.length_plot <- ggplot(Melted.Hs.male_axis, aes(y=axis.length, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+geom_boxplot()+ 
+  ggtitle("Human male SC Lengths")+theme(legend.position="none")
+hs.sc.length_plot
+
+```
+
+Neat. cool to see human SC data plotted. One outlier was removed.
+
+# Adding to whole Cells
+
+Look/use the DF, Curated_BivData
 
@@
 
+Chunk below has code for choosing mice with a good number of cells (25?)
+
 ```{r choosing.mice, echo=FALSE, eval=FALSE}
-#Choose mice with many cells
+#Choose mice with many cell observations 
+#important for increased power for within mouse variance
 
@@
 #lowest value of the high bin
-shiz
+
 #get the filenames for this groups of cells
@@
 
+from the curated BivData, figure out which can be added / into the whole cell data. I think there might only be a couple cell to add in.
+
+```{r add.more.whole.cell, echo=FALSE}
+
+#load data curatedBiv Data
+#load in hand measures
+
+```
+
 # Whole Cell Measures
@@
 
-```{r whole.cell.data, echo=FALSE}
+```{r setup.whole.cell.data, echo=FALSE}
 
@@
 #load(file="data/MLH1/MLH1_data_setup.RData")
-
-#load main data file
-whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
 
@@
 
-biv.count_table <-  ddply(.data=whole.cell.data,
-                 .(fileName),
-                 summarize,
-                 nbiv = length(SC.length)
-)
-#all have 20 or 19 (after removing XY)
-
 #ADD THE RANK
 for(musy in as.character(unique(whole.cell.data$fileName))){
-  print(musy)
+  #print(musy)
 
@@
 
-Try doing the permutations
+## Adding to whole cell measures
 
-```{r permutations.sc, echo=FALSE}
+calculated the observed values in this chunk.
+
+```{r cell.level.tables, echo=FALSE}
+
+whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
 
@@
 #ncells
-length(unique(male.whole.cell$fileName) ) #24 cells
+#length(unique(male.whole.cell$fileName) ) #24 cells
 
 female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
-length(unique(female.whole.cell$fileName) ) #18 cells
+#length(unique(female.whole.cell$fileName) ) #18 cells
 
+biv.count_table <-  ddply(.data=whole.cell.data,
+                 .(fileName),
+                 summarize,
+                 nbiv = length(SC.length)
+)#all have 20 or 19 (after removing XY)
+
+#real data
 male.biv.count_table <-  ddply(.data=male.whole.cell,
+                 .(fileName),
+                 summarize,
+                 nbiv = length(SC.length),
+                 mean_co = mean(hand.foci.count, rm.na = TRUE),
+                 mean_sc = mean(SC.length),
+                   var_co = var(hand.foci.count),
+                   var_sc = var(SC.length),
+                 cv_co.count = cv(hand.foci.count),
+                 cv_SC.length = cv(SC.length)
+)
+
+female.biv.count_table <-  ddply(.data=female.whole.cell,
                  .(fileName),
@@
                    var_co = var(hand.foci.count),
-                   var_sc = var(SC.length)
-)
+                   var_sc = var(SC.length),
+                 cv_co.count = cv(hand.foci.count),
+                 cv_SC.length = cv(SC.length)
+)#one of the female cells is missing values
 
-#not sure how to make the initial DF
-#replicate 24 rows with 19 cols
-iterations = 24
-variables = 20
-DF.permutations.sc <- matrix(ncol=variables, nrow=iterations)
-colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
-                               "rank6","rank7","rank8","rank9","rank10","rank11",
-                               "rank12","rank13","rank14","rank15","rank16",
-                               "rank17","rank18","rank19" )
-
-#ROW IS FIRST
-for(row.num in 1:24){
-  #print(c(row.num, "rowis") )
-  for(col.num in 2:ncol(DF.permutations.sc) ){
-  rand.rank.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
-  #print(SCll)   
-  DF.permutations.sc[row.num,col.num] <- rand.rank.sc
-   }
-  }#yes this finally works!!!!
-DF.permutations.sc <- as.data.frame(DF.permutations.sc)
-
-#pair with real within cell SC variance
-for( o in 1:nrow(DF.permutations.sc) ) {
-  DF.permutations.sc$var_sc[o] <- var(as.numeric(DF.permutations.sc[o,2:20]))
-  }
-
-#permutations without accounting for rank
-its = 24
-varis = 20
-DF.permutations.no_rank <- matrix(ncol=varis, nrow=its)
-#colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
-#                               "rank6","rank7","rank8","rank9","rank10","rank11",
-#                               "rank12","rank13","rank14","rank15","rank16",
-#                               "rank17","rank18","rank19" )
-
-for(row.num in 1:24){
-  #print(c(row.num, "rowis") )
-  for(col.num in 2:ncol(DF.permutations.no_rank) ){
-  rand.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
-  #print(SCll)   
-  DF.permutations.no_rank[row.num,col.num] <- rand.sc
-   }
-  }#yes this finally works!!!!
-DF.permutations.no_rank <- as.data.frame(DF.permutations.no_rank)
-
-for( o in 1:nrow(DF.permutations.sc) ) {
-  DF.permutations.no_rank$var_rand.sc[o] <- var(as.numeric(DF.permutations.no_rank[o,2:20]))
-  }
-
-#merge with real
-sc.var.for.plot <- matrix(ncol=3, nrow=72)
-sc.var.for.plot <- as.data.frame(sc.var.for.plot)
-colnames(sc.var.for.plot) <- c("cell.id","var","type")
-
-sc.var.for.plot$var[1:24] <- DF.permutations.sc$var_sc
-sc.var.for.plot$type[1:24] <- "sim"
-
-sc.var.for.plot$var[25:48] <- male.biv.count_table$var_sc
-sc.var.for.plot$type[25:48] <- "real"
-
-sc.var.for.plot$var[49:72] <- DF.permutations.no_rank$var_rand.sc
-sc.var.for.plot$type[49:72] <- "sim_no_rank"
-
-#plot the forms
-#plot within cell variance
-whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
-whole.cell.var_SC.plot
 ```
 
-```{r per.co.count, echo=FALSE}
+# Plots of whole cell data
 
+show SC length (y) by rank  (x), facet (sex)
+
+```{r play.plots, echo=FALSE}
+#female.whole.cell
+#male.whole.cell
+
+female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$SC.length) | female.whole.cell$SC.length==""), ]
+
+cow.hugger <- ggplot(female.whole.cell, aes(x=rank, y=SC.length, color=mouse) ) +geom_jitter()
+
+rank.length.boxplot <- ggplot(data=female.whole.cell) + geom_boxplot( aes(x=as.factor(rank), y=SC.length, fill=rank) )+geom_jitter(aes(x=as.factor(rank), y=SC.length) )+ggtitle("blaah")
+
+rank.length.boxplot.xx <- ggplot(data=female.whole.cell, aes(x=as.factor(rank), y=SC.length) ) + geom_boxplot()+ geom_jitter(aes( color=fileName))+
+  ggtitle("Female SC Length Dist by ranks")+theme(legend.position="none")
+
+rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=SC.length)) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")
+
+rank.length.boxplot.xx
+rank.length.boxplot.xy
+
+```
+
+# PERMUTATION SETUP
+
+```{r per.co.count, echo=FALSE, eval=FALSE}
+#I don't think ... I end up useing these (matrics / DFs)
 iterations = 24
@@
 #plot
-whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
-whole.cell.var_CO.plot
+#whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
+#whole.cell.var_CO.plot
 
@@
 
-```{r permuted.v.real, echo=FALSE}
-#evaluating the permuted 
-library(plyr)
+## CO count, male
 
+Make list of DFs to fill with permutations
+
+```{r list.of.DF,echo=FALSE}
+
+#dataframes to recreate -- and fill with 
+#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)
+
+#make subsampled piece of DF here -- these are taking the whole DF
+#I will just take single values
+#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
+#male.whole.cell
+#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+
+#rank (sample all of the ranks then stich them together)
+BIG.LIST_w.rank <- replicate(1000,
+  list(bind_rows(list( 
+  male.whole.cell[ sample( which(male.whole.cell$rank==1), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==2), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==3), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==4), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==5), 1, replace = FALSE ),],
+   male.whole.cell[ sample( which(male.whole.cell$rank==6), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==7), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==8), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==9), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==10), 1, replace = FALSE ),],
+
+   male.whole.cell[ sample( which(male.whole.cell$rank==11), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==12), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==13), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==14), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==15), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==16), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==17), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==18), 1, replace = FALSE ),],
+  male.whole.cell[ sample( which(male.whole.cell$rank==19), 1, replace = FALSE ),]
+))))
+#this is replicating single cell (NOT a SAMPLE of 24)
+
+#LIST of DF, no rank (include all), samples all 19 bivs at a time
+#makes a list of 10 -- [[]]
+BIG.DF.LIST_no.rank <- replicate(1000,
+                  list(bind_rows(list( 
+          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
+) ) ) )
+
+#dealing with lists of DFs, 
+#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
+#i is this index of the dataframes in the list?
+
+#Calq cV for the rank perm
+cv.val.DF_rank <- data.frame(  cV_co.count_rank = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
+ # print(j)
+  #fill in the cells of given index
+  cv.val.DF_rank$cV_co.count_rank[j] <-cv(  BIG.LIST_w.rank[[j]]$hand.foci.count )
+  }
+
+#Calq cV for no rank
+cv.val.DF_no.rank <- data.frame(  cV_co.count_no.rank = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  cv.val.DF_no.rank$cV_co.count_no.rank[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
+  }
+#extract the 
+#next step make cell-level calqs of cV -- for plotting
+
+#merge + plots of the cvs
+#10 and 10 rows from the permuted
+cv.val.DF_rank$type <- "sim w.rank"
+colnames(cv.val.DF_rank) <- c("cV","type" )
+
+cv.val.DF_no.rank$type <- "sim no rank"
+colnames(cv.val.DF_no.rank) <- c("cV", "type")
+
+#then add the real values
+real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count)
+real.vals.cv.co$type <- "real"
+colnames(real.vals.cv.co) <- c("cV", "type")
+
+#merge
+cVs_permutations.DF <- rbind(cv.val.DF_rank, cv.val.DF_no.rank, real.vals.cv.co)
+
+# need to melt the DFs before ploting?
+#ploting --- plot with more sims
+oo.plot <- ggplot(cVs_permutations.DF, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")
+
+#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")
+
+#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
+
+```
+
+The above plot the real data might have more variance because it's a combination of strains
+
+## CO count, female
+
+```{r CO.count.female, echo=FALSE}
+
+#dataframes to recreate -- and fill with 
+#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)
+
+#make subsampled piece of DF here -- these are taking the whole DF
+#I will just take single values
+#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
+#male.whole.cell
+#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+
+#rank (sample all of the ranks then stich them together)
+BIG.LIST_w.rank_female <- replicate(1000,
+  list(bind_rows(list( 
+  female.whole.cell[ sample( which(female.whole.cell$rank==1), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==2), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==3), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==4), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==5), 1, replace = FALSE ),],
+   female.whole.cell[ sample( which(female.whole.cell$rank==6), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==7), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==8), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==9), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==10), 1, replace = FALSE ),],
+
+   female.whole.cell[ sample( which(female.whole.cell$rank==11), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==12), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==13), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==14), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==15), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==16), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==17), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==18), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==19), 1, replace = FALSE ),],
+  female.whole.cell[ sample( which(female.whole.cell$rank==20), 1, replace = FALSE ),]
+  ))))
+
+#LIST of DF, no rank (include all), samples all 19 bivs at a time
+#makes a list of 10 -- [[]]
+BIG.DF.LIST_no.rank_female <- replicate(1000,
+                  list(bind_rows(list( 
+          female.whole.cell[ sample( which( female.whole.cell$rank < 20), 20, replace = FALSE ),]
+) ) ) )
+
+```
+
+These permuted dataframes can be reused for all the meiotic traits, (CO.count, SC length, and position).
+
+```{r sum.stat_cv.CO, echo=FALSE}
+#CALUCULATING SUMMARY STAT from the permuted DFs
+#dealing with lists of DFs, 
+#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
+#i is this index of the dataframes in the list?
+
+#Calq cV for the rank perm
+cv.val.DF_rank.female <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
+ # print(j)
+  #fill in the cells of given index
+  cv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$hand.foci.count ) #for each
+  }
+
+#Calq cV for no rank
+cv.val.DF_no.rank.female <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  cv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count ) #for each 
+}
+
+#merge + plots of the cvs
+#10 and 10 rows from the permuted
+cv.val.DF_rank.female$type <- "sim w.rank"
+colnames(cv.val.DF_rank) <- c("cV","type" )
+
+cv.val.DF_no.rank.female$type <- "sim no rank"
+colnames(cv.val.DF_no.rank.female) <- c("cV", "type")
+
+#then add the real values
+real.vals.cv.co_female <- data.frame(female.biv.count_table$cv_co.count)
+real.vals.cv.co_female$type <- "real"
+colnames(real.vals.cv.co_female) <- c("cV", "type")
+
+#merge
+cVs_permutations.DF_female <- rbind(cv.val.DF_rank.female, cv.val.DF_no.rank.female, real.vals.cv.co_female)
+
+# need to melt the DFs before ploting?
+#ploting --- plot with more sims
+Femaleplot_cV <- ggplot(cVs_permutations.DF_female, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count female, 1000 simulated")
+
+#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")
+
+#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
+
+cVs_permutations.DF$sex <- "male"
+cVs_permutations.DF_female$sex <- "female"
+
+all.cv <- rbind(cVs_permutations.DF, cVs_permutations.DF_female)
+cV_plot <- ggplot(all.cv, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")+facet_wrap(~sex)
+cV_plot
+
+```
+
+# SC length, bivalent level
+
+don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.
+
+```{r permutations.sc, echo=FALSE, eval=FALSE}
+
+#merge with real
+sc.var.for.plot <- matrix(ncol=3, nrow=72)
+sc.var.for.plot <- as.data.frame(sc.var.for.plot)
+colnames(sc.var.for.plot) <- c("cell.id","var","type")
+
+sc.var.for.plot$var[1:24] <- DF.permutations.sc$var_sc
+sc.var.for.plot$type[1:24] <- "sim"
+
+sc.var.for.plot$var[25:48] <- male.biv.count_table$var_sc
+sc.var.for.plot$type[25:48] <- "real"
+
+sc.var.for.plot$var[49:72] <- DF.permutations.no_rank$var_rand.sc
+sc.var.for.plot$type[49:72] <- "sim_no_rank"
+
+#plot the forms
+#plot within cell variance
+whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
+whole.cell.var_SC.plot
+```
+
+```{r sum.stat.SClength, echo=FALSE}
+
+#Calq SC cV for the SC length FEMALE
+SCcv.val.DF_rank.female <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
+  #print(j)
+  #fill in the cells of given index
+  SCcv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$SC.length ) #for each
+  }
+
+#Calq SC cV for no rank
+SCcv.val.DF_no.rank.female <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  SCcv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$SC.length ) #for each 
+}
+
+#merge + plots of the cvs
+#10 and 10 rows from the permuted
+SCcv.val.DF_rank.female$type <- "sim w.rank"
+colnames(SCcv.val.DF_rank.female) <- c("cV","type" )
+
+SCcv.val.DF_no.rank.female$type <- "sim no rank"
+colnames(SCcv.val.DF_no.rank.female) <- c("cV", "type")
+
+#then add the real values
+real.vals.cv.SC_female <- data.frame(female.biv.count_table$cv_SC.length)
+real.vals.cv.SC_female$type <- "real"
+colnames(real.vals.cv.SC_female) <- c("cV", "type")
+
+#merge and plots
+cV.SC_permutations.DF_female <- rbind(SCcv.val.DF_rank.female, SCcv.val.DF_no.rank.female, real.vals.cv.SC_female)
+
+#MALE --BIG.LIST_w.rank, BIG.DF.LIST_no.rank,
+#Calq cV for the SC length
+SCcv.val.DF_rank.male <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(j)
+  #fill in the cells of given index
+  SCcv.val.DF_rank.male$cV[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
+  }
+
+#Calq cV for no rank
+SCcv.val.DF_no.rank.male <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  SCcv.val.DF_no.rank.male$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length ) #for each 
+}
+
+#merge + plots of the cvs
+#10 and 10 rows from the permuted
+SCcv.val.DF_rank.male$type <- "sim w.rank"
+colnames(SCcv.val.DF_rank.male) <- c("cV","type" )
+
+SCcv.val.DF_no.rank.male$type <- "sim no rank"
+colnames(SCcv.val.DF_no.rank.male) <- c("cV", "type")
+
+#then add the real values-- male.biv.count_table
+real.vals.cv.SC_male <- data.frame(male.biv.count_table$cv_SC.length)
+real.vals.cv.SC_male$type <- "real"
+colnames(real.vals.cv.SC_male) <- c("cV", "type")
+
+#merge and plots
+cV.SC_permutations.DF_male <- rbind(SCcv.val.DF_rank.male, SCcv.val.DF_no.rank.male, real.vals.cv.SC_male)
+cV.SC_permutations.DF_male$sex <- "male"
+cV.SC_permutations.DF_female$sex <- "female"  #cV.SC_permutations.DF_female
+all_cv.SC <- rbind(cV.SC_permutations.DF_female, cV.SC_permutations.DF_male)
+
+cV.SC_plot <- ggplot(all_cv.SC, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv of within cell SC lengths, 1000 permutations")+facet_wrap(~sex)
+```
+
+```{r permuted.v.real, echo=FALSE, eval=FALSE}
 for( o in 1:nrow(DF.permutations_CO.count) ) {
@@
 
+# Junk
+
 ```{r junk, eval=FALSE}
@@
                        )
+
+#not sure how to make the initial DF
+#replicate 24 rows with 19 cols
+iterations = 24
+variables = 20
+DF.permutations.sc <- matrix(ncol=variables, nrow=iterations)
+colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
+                               "rank6","rank7","rank8","rank9","rank10","rank11",
+                               "rank12","rank13","rank14","rank15","rank16",
+                               "rank17","rank18","rank19" )
+
+#permutations -- for SC
+
+#permutations without accounting for rank
+its = 24
+varis = 20
+DF.permutations.no_rank <- matrix(ncol=varis, nrow=its)
+#colnames(DF.permutations.sc) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
+#                               "rank6","rank7","rank8","rank9","rank10","rank11",
+#                               "rank12","rank13","rank14","rank15","rank16",
+#                               "rank17","rank18","rank19" )
+
+#ROW IS FIRST
+for(row.num in 1:24){
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations.sc) ){
+  rand.rank.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations.sc[row.num,col.num] <- rand.rank.sc
+   }
+  }#yes this finally works!!!!
+DF.permutations.sc <- as.data.frame(DF.permutations.sc)
+
+#pair with real within cell SC variance
+for( o in 1:nrow(DF.permutations.sc) ) {
+  DF.permutations.sc$var_sc[o] <- var(as.numeric(DF.permutations.sc[o,2:20]))
+  }
+
+for(row.num in 1:24){
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations.no_rank) ){
+  rand.sc <- sample(male.whole.cell$SC.length[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations.no_rank[row.num,col.num] <- rand.sc
+   }
+  }#yes this finally works!!!!
+DF.permutations.no_rank <- as.data.frame(DF.permutations.no_rank)
+
+for( o in 1:nrow(DF.permutations.sc) ) {
+  DF.permutations.no_rank$var_rand.sc[o] <- var(as.numeric(DF.permutations.no_rank[o,2:20]))
+  }
+
 ```

### 2019-12-06 results/Chapter3_ideas.Rmd
working of randomizing foci position

@@
 load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
-
-```
-
-# Summary and recreation of Wang et al 2019
-
-insert summary of the Wang paper
-
-which data did they use? what summary stat were they comparing
-What were their results?
-
-human data- 
-Gruhn et al 2013 (ploted below)
+```
+
+# Summary and of Wang et al 2019
+
+Main take away from Wang et al 2019 is that chromosomes with the same nucleus covary in number of crossovers.
+Axis lengths / SC lengths show the same pattern and is known to be regulated at the cell wide level.
+The authors suggest that the within nucleus covariation in CO per chromosome is mediated through cell wide regulation of SC area.
+(the lengths of SC for chromosomes within a cell a correlated, there is more area for COs).
+
+Data used in the paper
+human data- Gruhn et al 2013 (ploted below, 755 cells)
 Sun et al 2006
@@
 
+Was there originall CO position data for this human data? (MLH1 spreads) (there is axis length and CO count per chrm -- this should mean there is position )
+
 ```{r human.male.data, echo=FALSE}
@@
 Hs.male_axis <- read.csv("~./MLH1repo/data/Human_male_axis_length.csv", header = TRUE)
-Hs.male_MLH1 <- read.csv("~./MLH1repo/data/Human_male_MLH1.csv", header = TRUE)
 #I assume that the rows of MLH1 and SC lengths are the linked observations, check or verify this assumption
@@
 
-```
-
-Neat. cool to see human SC data plotted. One outlier was removed.
-
-# Adding to whole Cells
-
-Look/use the DF, Curated_BivData
-
-Take some mouse level distributions -- pull cells from top and bottom 10%?
-(gather whole bivalent level data,)
-
--start making comparisons for the 2 pools
-  how much technical noise is there? (staining differences?/ quality score?
-
-Mice with many cells and BivData
-17mar16_G_f4 - BivData
-17mar16_G_f1
-
-Chunk below has code for choosing mice with a good number of cells (25?)
-
-```{r choosing.mice, echo=FALSE, eval=FALSE}
-#Choose mice with many cell observations 
-#important for increased power for within mouse variance
-
-#G_f, G_m
-#SKIVE_f, SKIVE_m #have BivData
-
-#remove all the quality 5
-
-#make plot/histogram of the MLH1 count distribution
-counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]
-
-mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
-mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
-high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
-low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]
-
-ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
-topsy <- ordered.MLH1[1:5,]
-bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
-#now have top and bottom 5 cell observations
+#### .. Hs.male_MLH1
+Hs.male_MLH1 <- read.csv("~./MLH1repo/data/Human_male_MLH1.csv", header = TRUE)
+
+Melted.Hs.male_MLH1 <- melt(Hs.male_MLH1, 
+                            c("row.name", "chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22"),
+                         id=c("row.name") )
+#remove the first part with of data frame, variable = "row.name""
+
+#clean up dm1 (there are a bunch of rows without useful information)
+Melted.Hs.male_MLH1 <- Melted.Hs.male_MLH1[(!is.na(Melted.Hs.male_MLH1$value)),]
+Melted.Hs.male_MLH1 <- Melted.Hs.male_MLH1[((Melted.Hs.male_MLH1$variable != "row.name")),]
+colnames(Melted.Hs.male_MLH1) <- c("row.name", "Chrm.id", "CO")
+
+#Seems like there is an outlier, 
+#Melted.Hs.male_axis <- Melted.Hs.male_axis[(Melted.Hs.male_axis$axis.length < 700 #),]
+
+hs.CO.length_plot <- ggplot(Melted.Hs.male_MLH1, aes(y=CO, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+ 
+  ggtitle("Human male CO Counts")+theme(legend.position="none")
+hs.CO.length_plot
+```
+
+approach -- they compared covariation (cV) of per chrm CO count and SC length for permuted/randomized groupd of chromosome observations representing 'in silico' nuclei.
+
+## Redoing the permutaitons
+
+Cell summaries below
+
+```{r Hs.cell_table, echo=FALSE}
+#make table for cell wide measures
+Hs.CO_cell_wide <- ddply(.data=Melted.Hs.male_MLH1, 
+                            .(row.name),
+                            summarize, 
+                            nbivs = length(unique(Chrm.id)),
+                            n.CO = sum(CO),
+                            mean.CO = mean(CO),
+                            cV.CO = cv(CO)
+)
+head(Hs.CO_cell_wide)
+
+Hs.SC_cell_wide <- ddply(.data=Melted.Hs.male_axis, 
+                            .(row.name),
+                            summarize,
+
+                            nbivs = length(unique(Chrm.id)),
+                            total.SC = sum(axis.length),
+                            mean.SC = mean(axis.length),
+                            cV.SC = cv(axis.length)
+)
+head(Hs.SC_cell_wide)
+
+```
+
+```{r Hs.Perm_noRank, echo=FALSE}
+
+Melted.Hs.male_MLH1$chrm.num <- (Melted.Hs.male_MLH1$Chrm.id)
+levels(Melted.Hs.male_MLH1$chrm.num) <- 1:23
+
+Melted.Hs.male_MLH1$chrm.num <- as.character(Melted.Hs.male_MLH1$chrm.num)
+Melted.Hs.male_MLH1$chrm.num <- as.numeric(Melted.Hs.male_MLH1$chrm.num)
+#chrm.num is now 1:23
+
+#LIST of DF, no rank (include all), samples all 19 bivs at a time
+Hs_male.BIG.DF.LIST_no.rank <- replicate(755,
+                  list(bind_rows(list( 
+          #can't make do logic on factors
+          Melted.Hs.male_MLH1[ sample( which( Melted.Hs.male_MLH1$chrm.num < 24), 22),]
+) ) ) )
+#now have randomized DF of 1000 cells
+```
+
+```{r Hs.Perm_wRank, echo=FALSE}
+#rank (sample all of the ranks then stich them together)
+Hs.CO_BIG.LIST_w.rank <- replicate(755,
+  list(bind_rows(list(
+
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==2), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==3), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==4), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==5), 1, replace = FALSE ),],
+   Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==6), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==7), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==8), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==9), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==10), 1, replace = FALSE ),],
+
+   Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==11), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==12), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==13), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==14), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==15), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==16), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==17), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==18), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==19), 1, replace = FALSE ),],
+    Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==20), 1, replace = FALSE ),],
+  Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==21), 1, replace = FALSE ),],
+    Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==22), 1, replace = FALSE ),],
+      Melted.Hs.male_MLH1[ sample( which(Melted.Hs.male_MLH1$chrm.num==23), 1, replace = FALSE ),]
+))
+)) #list of DFs which are 'cells' with observations drawn from pooled chrm bins
+```
+
+```{r sum.stat.display, echo=FALSE}
+#after making big randomize / permuted data sets, CALCULATE the metrics / sum.stat
+#Hs.CO_BIG.LIST_w.rank
+#Hs_male.BIG.DF.LIST_no.rank
+
+#Calq cV for the rank perm
+#1. make empty DF
+Hs.CO_cv_rank <- data.frame(  cV.CO = rep(0, length(Hs.CO_BIG.LIST_w.rank) ) )
+
+for(j in 1:length(Hs.CO_BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(j)
+  #fill in the cells of given index
+  Hs.CO_cv_rank$cV.CO[j] <-cv(  Hs.CO_BIG.LIST_w.rank[[j]]$CO ) #but j should be row, row should be a cell
+  #add CO per nuc
+  Hs.CO_cv_rank$co.per.nuc[j] <- sum(Hs.CO_BIG.LIST_w.rank[[j]]$CO)
+  }
+
+#Calq cV for no rank
+Hs.CO_cv_no.rank <- data.frame(  cV.CO = rep(0, length(Hs_male.BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(Hs_male.BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  Hs.CO_cv_no.rank$cV.CO[i] <-cv(  Hs_male.BIG.DF.LIST_no.rank[[i]]$CO )
+  Hs.CO_cv_no.rank$co.per.nuc[i] <- sum(Hs_male.BIG.DF.LIST_no.rank[[i]]$CO) #these are the total sum, not divided out by nuc!
+  }
+
+#merge simulated and real values into 1DF
+Hs.CO_cv_rank$type <- "sim w.rank"
+colnames(Hs.CO_cv_rank) <- c("cV","COs","type" )
+
+Hs.CO_cv_no.rank$type <- "sim no rank"
+colnames(Hs.CO_cv_no.rank) <- c("cV", "COs","type")
+
+#then add the real values   Hs.CO_cell_wide
+Hs_real.cv.co <- data.frame(cV = Hs.CO_cell_wide$cV.CO, Hs.CO_cell_wide$n.CO)
+Hs_real.cv.co$type <- "real"
+colnames(Hs_real.cv.co) <- c("cV","COs", "type")
 
 #merge
-top.n.bottom.bin <- rbind(topsy,bottomzy)
-
-#write 
-write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)
-
-shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
-       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
-geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
-#highest of the low bin
-#lowest value of the high bin
-
-#get the filenames for this groups of cells
-
-#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
-#28feb15_PWD_f2, in PWD--female .xlsx file
-
-#31may19_SKIVE_m2
-male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
-male.one.DF <- male.one.DF[male.one.DF$quality < 5,]
-
-ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
-male.high.bin.DF <- ordered.MLH1.male[1:5,]
-male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
-#now have top and bottom 5 cell observations
-
-#merge
-male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
-#write 
-write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)
-
-shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
-       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
-geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
-#highest of the low bin
-#lowest value of the high bin
-shiz.too
-
-```
-
-the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
-Should I be comparing cells with the same number of COs?
-
-```{r getting.cell.names, echo=FALSE,eval=FALSE}
-
-#load in the cell dataset
-#28feb15_PWD_f2  from Curated_BivData
-#for a mouse, pull the 10 highest rec cells
-
-PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
-#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
-#"Foci3"                   "curation.notes"  
-female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]
-
-write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)
-
-PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
-PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")
-
-#merge with hand data
-#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
-#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)
-
-cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
-                            .(fileName),
-                            summarize, 
-                            ncells = length(unique(fileName)),
-                            nbivs = length(unique(Obj.ID)),
-                            n.foci = sum(hand.foci.count),
-                            total.SC = sum(chromosomeLength),
-                            mean.SC = mean(chromosomeLength)
-
-)
-#add hand data
-ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]
-
-#I have 3 cells with full 20
-
-```
-
-Start discribing the Bivalent patterns for the PWD female patterns
-
-```{r mini.biv, echo=FALSE,eval=FALSE}
-#make mini biv plots
-
-#PWD.female
-#PWD.female_hand.measure
-
-cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
-cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]
-
-two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
-two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))
-
-#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
-#of SC.
-MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
-    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
-  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
- #geom_hline(yintercept = c(male_IFD) ) +
-
-  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
-  #how to add points for 2foci
-  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
-  #foci.1.position, two_groups$distal.foci.position)
-  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +
-
-#add centromeres...  
-  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+
-
-  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
- # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
-  ylim(c(0,200))+
-     theme(axis.line=element_blank(),axis.text.x=element_blank(),
-        axis.ticks=element_blank())  #for removeing x labels
-
-#MiniBiv_plot1
-#MiniBiv_plot2
-
-#save the files
-
-```
-
-```{r biscribe.Biv.Pattern, echo=FALSE}
-#start discribing the basic
-
-#start merging curated bivData with handmeasured data
-
-#BivData - Curated
-```
-
-from the curated BivData, figure out which can be added / into the whole cell data. I think there might only be a couple cell to add in.
-
-```{r add.more.whole.cell, echo=FALSE}
-
-#load data curatedBiv Data
-#load in hand measures
-
-```
-
-# Whole Cell Measures
-
-excel sheet Whole_cell_measures.xlsx has 840 bivalent measures.
+Hs.CO_cVs.DF <- rbind(Hs.CO_cv_rank, Hs.CO_cv_no.rank, Hs_real.cv.co)
+
+# need to melt the DFs before ploting?
+#ploting --- plot with more sims
+#HS.plot <- ggplot(Hs.CO_cVs.DF, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("Cell-wide cv for CO, Human male")
+#HS.plot
+```
+
+```{r HS.plots, echo=FALSE}
+
+HS.scatter.plot <- ggplot(Hs.CO_cVs.DF, aes(y = COs, x=type) )+ geom_jitter()+ ggtitle("COs per cell, Human male")
+HS.scatter.plot
+
+#make density distributions
+#percent of cells
+
+#quantilles of COs
+
+hist <- ggplot(Hs.CO_cVs.DF, aes(x = COs)) + geom_histogram(aes(fill=type), binwidth = 2 )
+hist
+
+HS.density.plot <- ggplot(Hs.CO_cVs.DF, aes(x = COs, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Human male")+xlim(c(0,100))
+HS.density.plot
+```
+
+To calculate the variability index, authors take percent of cells(integrated curve area) which differ between the empiracal and simulated distributions.
+
+I changed the number of permutations to 200, which should be the same number of oberved cells, but it looks like the real data is much larger....the sheet has 755 cell observations, not 200.
+This might be due to how histograms are presented.
+
+The rank distinction is to test if sampling within pools of specific chromsomes makes a difference.
+The smaller/tighter distribution between CO count per nuc with the ranked and non-ranked distributions indecate there is a small chromosome specific effect.
+
+Not sure how to calculate the differences in distributions yet.
+
+this current data shows random samples, not permuted random samples.
+
+# Mus whole Cell Data Set
 
@@
 whole.cell.data <- add_sex(whole.cell.data)
+whole.cell.data <- add_strain(whole.cell.data)
+whole.cell.data <- add_category(whole.cell.data)
 
 whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$fileName) | whole.cell.data$fileName==""), ]
+#this doesn't remove all of the data that are missing values
+
+whole.cell.data$Obj.ID <- paste(whole.cell.data$fileName, whole.cell.data$boxNumber, sep = "_")
 
@@
 
+whole.cell.data$Foci1 <- whole.cell.data$F1
+whole.cell.data$Foci2 <- whole.cell.data$F2
+whole.cell.data$Foci3 <- whole.cell.data$F3
+
+#add IFDs and normalized positions 
+whole.cell.data$Foci1.PER <- whole.cell.data$F1 / whole.cell.data$SC.length
+whole.cell.data$Foci2.PER <- whole.cell.data$F2 / whole.cell.data$SC.length
+whole.cell.data$Foci3.PER <- whole.cell.data$F3 / whole.cell.data$SC.length
+
+whole.cell.data <- add_IFD(whole.cell.data)
+whole.cell.data <-  add_IFD.2(whole.cell.data)
+
+Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+#IFD.2 does mulitple IFDs (just absolute)
+Curated_BivData <- add_IFD.2(Curated_BivData)
+
 #find cells with the wrong number
+#whole.cell.data_cell.table <-
 
@@
 
-## Adding to whole cell measures
-
-calculated the observed values in this chunk.
+To make the same comparisons, I need chromosome/bivalent level observations for whole cells (all chromosomes). 
+I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations
+
+```{r whole.cell.table, echo=FALSE}
+
+table(whole.cell.data$category)
+
+#table(cell.table$category)
+
+```
 
@@
 whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
-
 male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
@@
                  nbiv = length(SC.length),
+                  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+
+                 CO = sum(hand.foci.count),
+                 total.SC = sum(SC.length),
                  mean_co = mean(hand.foci.count, rm.na = TRUE),
@@
                  nbiv = length(SC.length),
+
+                  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+
+                 CO = sum(hand.foci.count),
+                  total.SC = sum(SC.length),
                  mean_co = mean(hand.foci.count),
@@
 
-# Plots of whole cell data
-
-show SC length (y) by rank  (x), facet (sex)
+I calulate within cell ranked SC length to use rank as a proxy for chromsome ID for bivalents.
+
+Below are plots of the SC distributions across the rank
 
@@
 #male.whole.cell
-
 female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$SC.length) | female.whole.cell$SC.length==""), ]
@@
 #merge with real
-co.var.for.plot <- matrix(ncol=3, nrow=72)
-co.var.for.plot <- as.data.frame(co.var.for.plot)
-colnames(co.var.for.plot) <- c("cell.id","var","type")
-
-co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
-co.var.for.plot$type[1:24] <- "sim"
-
-co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
-co.var.for.plot$type[25:48] <- "real"
-
-co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
-co.var.for.plot$type[49:72] <- "sim_no_rank"
+#co.var.for.plot <- matrix(ncol=3, nrow=72)
+#co.var.for.plot <- as.data.frame(co.var.for.plot)
+#colnames(co.var.for.plot) <- c("cell.id","var","type")
+
+#co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
+#co.var.for.plot$type[1:24] <- "sim"
+
+#co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
+#co.var.for.plot$type[25:48] <- "real"
+
+#co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
+#co.var.for.plot$type[49:72] <- "sim_no_rank"
 
@@
 
-```{r list.of.DF,echo=FALSE}
+```{r list.of.DF, echo=FALSE}
 
@@
 #rank (sample all of the ranks then stich them together)
-BIG.LIST_w.rank <- replicate(1000,
+#replication should match the num
+BIG.LIST_w.rank <- replicate(18,
   list(bind_rows(list( 
@@
 #makes a list of 10 -- [[]]
-BIG.DF.LIST_no.rank <- replicate(1000,
+BIG.DF.LIST_no.rank <- replicate(18,
                   list(bind_rows(list( 
@@
 ) ) ) )
-
+```
+
+```{r sum.stat, echo=FALSE}
 #dealing with lists of DFs, 
@@
 #Calq cV for the rank perm
-cv.val.DF_rank <- data.frame(  cV_co.count_rank = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
+cv.val.DF_rank <- data.frame(  cV = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
 for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
@@
   #fill in the cells of given index
-  cv.val.DF_rank$cV_co.count_rank[j] <-cv(  BIG.LIST_w.rank[[j]]$hand.foci.count )
+  cv.val.DF_rank$cV[j] <-cv(  BIG.LIST_w.rank[[j]]$hand.foci.count )
+    cv.val.DF_rank$CO[j] <-sum(  BIG.LIST_w.rank[[j]]$hand.foci.count )
+
     }
@@
 #Calq cV for no rank
-cv.val.DF_no.rank <- data.frame(  cV_co.count_no.rank = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
+cv.val.DF_no.rank <- data.frame(  cV_CO = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
 for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
@@
   #fill in the cells of given index
-  cv.val.DF_no.rank$cV_co.count_no.rank[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
+  cv.val.DF_no.rank$cV_CO[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
+   cv.val.DF_no.rank$CO[i] <- sum(  BIG.DF.LIST_no.rank[[i]]$hand.foci.count )
   }
@@
 cv.val.DF_rank$type <- "sim w.rank"
-colnames(cv.val.DF_rank) <- c("cV","type" )
+colnames(cv.val.DF_rank) <- c("cV", "CO", "type" )
 
 cv.val.DF_no.rank$type <- "sim no rank"
-colnames(cv.val.DF_no.rank) <- c("cV", "type")
+colnames(cv.val.DF_no.rank) <- c("cV", "CO", "type")
 
 #then add the real values
-real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count)
+real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count,male.biv.count_table$CO )
 real.vals.cv.co$type <- "real"
-colnames(real.vals.cv.co) <- c("cV", "type")
+colnames(real.vals.cv.co) <- c("cV","CO", "type")
 
@@
 cVs_permutations.DF <- rbind(cv.val.DF_rank, cv.val.DF_no.rank, real.vals.cv.co)
+
+#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")
+
+#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
+
+```
+
+```{r perm.plots, echo=FALSE}
+#plots for permutations
 
@@
 
-#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")
-
-#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
-
-```
-
+Mus_hist <- ggplot(cVs_permutations.DF, aes(x = CO)) + geom_histogram(aes(fill=type), binwidth = 2 )+ggtitle("COs per cell, Mus male")
+
+HS.density.plot <- ggplot(cVs_permutations.DF, aes(x = CO, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Mus male")+xlim(c(0,100))
+HS.density.plot
+
+```
+
+Above plot shows overdispersion of CO counts per cell for my Mus male dataset.
 The above plot the real data might have more variance because it's a combination of strains
@@
 
-```{r CO.count.female, echo=FALSE}
+```{r rand.DF_CO.count.female, echo=FALSE}
 
@@
 #rank (sample all of the ranks then stich them together)
-BIG.LIST_w.rank_female <- replicate(1000,
+BIG.LIST_w.rank_female <- replicate(24,
   list(bind_rows(list( 
@@
 #LIST of DF, no rank (include all), samples all 19 bivs at a time
-#makes a list of 10 -- [[]]
-BIG.DF.LIST_no.rank_female <- replicate(1000,
+BIG.DF.LIST_no.rank_female <- replicate(24,
                   list(bind_rows(list( 
@@
   cv.val.DF_rank.female$cV[j] <-cv(  BIG.LIST_w.rank_female[[j]]$hand.foci.count ) #for each
+  cv.val.DF_rank.female$CO[j] <-sum(  BIG.LIST_w.rank_female[[j]]$hand.foci.count )
   }
@@
   #fill in the cells of given index
-  cv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count ) #for each 
+  cv.val.DF_no.rank.female$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count )
+  cv.val.DF_no.rank.female$CO[i] <-sum(  BIG.DF.LIST_no.rank_female[[i]]$hand.foci.count )
 }
@@
 cv.val.DF_rank.female$type <- "sim w.rank"
-colnames(cv.val.DF_rank) <- c("cV","type" )
+colnames(cv.val.DF_rank) <- c("cV","CO","type" )
 
 cv.val.DF_no.rank.female$type <- "sim no rank"
-colnames(cv.val.DF_no.rank.female) <- c("cV", "type")
+colnames(cv.val.DF_no.rank.female) <- c("cV","CO", "type")
 
 #then add the real values
-real.vals.cv.co_female <- data.frame(female.biv.count_table$cv_co.count)
+real.vals.cv.co_female <- data.frame(cv = female.biv.count_table$cv_co.count, CO = female.biv.count_table$CO)
 real.vals.cv.co_female$type <- "real"
-colnames(real.vals.cv.co_female) <- c("cV", "type")
+colnames(real.vals.cv.co_female) <- c("cV","CO", "type")
 
@@
 
+Mus_hist_F <- ggplot(cVs_permutations.DF_female, aes(x = CO)) + geom_histogram(aes(fill=type), binwidth = 2 )+ggtitle("COs per cell, Mus female")
+
+Mus.density.plot_F <- ggplot(cVs_permutations.DF_female, aes(x = CO, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Mus female")+xlim(c(0,50))
+Mus.density.plot_F
+
 #  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free")
@@
 cV_plot <- ggplot(all.cv, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv for cell co count, 1000 simulated")+facet_wrap(~sex)
-cV_plot
-
-```
+
+```
+
+```{r female.plot, echo=FALSE}
+
+```
+
+Density plot above also shows that the female CO count data is overdispersed compered to a randomized distribution.
 
@@
 
+Start doing the same thing but for bivalent SC lengths.
+Use the same randomized data sets!
+summary stats - mean SC length, cV SC length, total SC area
+
 don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.
 
-```{r permutations.sc, echo=FALSE, eval=FALSE}
-
-#merge with real
-sc.var.for.plot <- matrix(ncol=3, nrow=72)
-sc.var.for.plot <- as.data.frame(sc.var.for.plot)
-colnames(sc.var.for.plot) <- c("cell.id","var","type")
-
-sc.var.for.plot$var[1:24] <- DF.permutations.sc$var_sc
-sc.var.for.plot$type[1:24] <- "sim"
-
-sc.var.for.plot$var[25:48] <- male.biv.count_table$var_sc
-sc.var.for.plot$type[25:48] <- "real"
-
-sc.var.for.plot$var[49:72] <- DF.permutations.no_rank$var_rand.sc
-sc.var.for.plot$type[49:72] <- "sim_no_rank"
+```{r SC_Sum.Stats.male, echo=FALSE}
+# calq the sum stats
+SC_Mus.male_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
+ # print(j)
+  #fill in the cells of given index
+  SC_Mus.male_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
+  SC_Mus.male_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank[[j]]$SC.length )
+  SC_Mus.male_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank[[j]]$SC.length )
+  }
+
+#Calq cV for no rank
+SC_Mus.male_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  SC_Mus.male_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length )
+  SC_Mus.male_no.rank$total_SC[i] <-sum(  BIG.DF.LIST_no.rank[[i]]$SC.length )
+  SC_Mus.male_no.rank$mean_SC[i] <-mean(  BIG.DF.LIST_no.rank[[i]]$SC.length )
+}
+
+#10 and 10 rows from the permuted
+SC_Mus.male_w.rank$type <- "sim w.rank"
+colnames(SC_Mus.male_w.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )
+
+SC_Mus.male_no.rank$type <- "sim_no.rank"
+colnames(SC_Mus.male_no.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )
+
+real_SC.Mus <- data.frame(male.biv.count_table$var_sc,  male.biv.count_table$total.SC, male.biv.count_table$mean_sc )
+real_SC.Mus$type <- "real"
+
+colnames(real_SC.Mus) <- c("SC.cV", "tot.SC", "mean.SC", "type")
+#merge
+SC_Mus_male <- rbind(SC_Mus.male_w.rank, SC_Mus.male_no.rank,real_SC.Mus)
+
+```
+
+```{r SC.plots_Mus.male, echo=FALSE}
+Mus_SC_hist <- ggplot(SC_Mus_male, aes(x = tot.SC)) + geom_histogram(aes(fill=type), binwidth = 50 )+ggtitle("total SC per cell, Mus male")
+
+Mus_SC_density <- ggplot(SC_Mus_male, aes(x = tot.SC, fill=type) )+geom_density(alpha=.6) + ggtitle("Total SC per cell, Mus male")+xlim(c(800,2000))
+Mus_SC_density
 
@@
 #plot within cell variance
-whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
-whole.cell.var_SC.plot
+#whole.cell.var_SC.plot  <- ggplot(sc.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell sc variance")
+```
+
+```{r SC_sum.stats.female, echo=FALSE, eval=FALSE}
+
+# calq the sum stats
+SC_Mus.female_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank_female) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank_female)){  #inside loop, move thru the LISTs of DFs
+ # print(j)
+  #fill in the cells of given index
+  SC_Mus.female_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank_female[[j]]$SC.length ) #for each
+  SC_Mus.female_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank_female[[j]]$SC.length )
+  SC_Mus.female_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank_female[[j]]$SC.length )
+  }
+
+#Calq cV for no rank
+SC_Mus.female_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female.) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  SC_Mus.female_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank_female.[[i]]$SC.length )
+  SC_Mus.female_no.rank$total_SC[j] <-sum(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
+  SC_Mus.female_no.rank$mean_SC[j] <-mean(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
+}
+
 ```
@@
 
-cV.SC_plot <- ggplot(all_cv.SC, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv of within cell SC lengths, 1000 permutations")+facet_wrap(~sex)
+cV.SC_plot <- ggplot(all_cv.SC, aes(y = cV, x=type) )+ geom_jitter()+ ggtitle("cv of within cell SC lengths")+facet_wrap(~sex)
 ```
@@
 
+# CO Position
+
+don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.
+
+look at CO position metrics.
+Do I need to melt the plots 
+the position col names are also, messed up I need to fix them --
+
+1. use the same randomized dataframe
+2. use the summary stat
+
+How do I randomize CO pos for a chrm?
+I want to draw positions from chrm_classes
+
+For each cell observation, note the classes,  I want to 
+make bins for positions of all
+
+```{r CO.pos, echo=FALSE}
+#male.whole.cell
+#female.whole.cell
+#fix the data DF for PERMUTING CO
+
+whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
+male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
+
+male.biv.count_table <- male.biv.count_table[!(is.na(male.biv.count_table$fileName) | male.biv.count_table$fileName==""), ]
+
+pos.pool_1CO <- male.whole.cell[male.whole.cell$hand.foci.count == 1,]
+pos.pool_2CO <- male.whole.cell[male.whole.cell$hand.foci.count == 2,]
+pos.pool_3CO <- male.whole.cell[male.whole.cell$hand.foci.count == 3,]
+
+#now make the randomized data set for CO pos
+#do whole row -- so I can cen n
+
+random_pos <- data.frame(male.whole.cell$fileName, male.whole.cell$SC.length, male.whole.cell$hand.foci.count, male.whole.cell$Obj.ID  )
+
+#INITIALIZE EMPTY list of DFs
+#fill in with full DF
+
+LIST.DFs <- setNames( replicate( length(male.biv.count_table$fileName), data.frame( obj.ID = c())), male.biv.count_table$fileName)
+
+ids <-    male.whole.cell$Obj.ID[which(male.whole.cell$fileName == '14oct16_30jun16_CAST_m3_sp1_3.2')]
+
+LIST.DFs[1]$obj.ID[1:19] <- ids[1:19]
+
+#INITALIZE empty LIST (make DF in the loop)
+
+LIST <- list()
+
+#Loop for rand sampling  
+for( male.cell in male.biv.count_table$fileName){
+
+  #for each fileName, make a new DF
+  DF.in.loop <- data.frame(matrix(ncol = 1, nrow = 19))
+  #making the Within loop DF
+  DF.in.loop$X1 <- male.whole.cell$fileName[male.whole.cell$fileName ==male.cell]
+  DF.in.loop$obj.ID <- male.whole.cell$Obj.ID[male.whole.cell$fileName ==male.cell]
+  DF.in.loop$SC.length <- male.whole.cell$SC.length[male.whole.cell$fileName ==male.cell]
+  DF.in.loop$hand.foci.count <- male.whole.cell$hand.foci.count[male.whole.cell$fileName ==male.cell]
+
+  #LIST.DFs[[male.cell]]$obj.ID[1:19] <- male.whole.cell$Obj.ID[male.whole.cell$fileName == male.cell]
+    #print(male.cell)
+    n0CO = male.biv.count_table$nCO0[male.biv.count_table$fileName == male.cell] #this reference needs to be from male.biv.count_table
+    n1CO = male.biv.count_table$nCO1[male.biv.count_table$fileName == male.cell]
+    n2CO = male.biv.count_table$nCO2[male.biv.count_table$fileName == male.cell]
+    n3CO = male.biv.count_table$nCO3[male.biv.count_table$fileName == male.cell]
+
+   #rand samples for foci.pos
+    #JUST DO THIS FOR 1CO bivs
+    DF.in.loop$rand_F1
+    sample.n1CO <- sample(pos.pool_1CO$Foci1.PER, n1CO, replace = FALSE )  #sample from the cell.table
+
+    #sample a random 2CO chrm (obj.ID),    #doesn't make sense to sample a random position from 2CO pool, need to sample the linked positions, or get the 
+    #chrm observation
+   sample.n2CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n2CO, replace = FALSE )
+   sample.n3CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n3CO, replace = FALSE )
+
+   #MAP normalized positions onto SC lengths/chrms of the fileName cell obs
+   #map the 1CO rand Foci position onto 1CO SCLength
+
+   #for i in n1CO
+#   which(random_pos$fileName == male.cell)
+   random_pos$Rand.F1[which(random_pos$fileName == male.cell)][i] <- sample(pos.pool_1CO$Foci1.PER, 1, replace = FALSE)
+
+   #list of the chrm lengths for all the 1COs for the current cell obs
+   #SC.Length_1COs <-
+
+}
+
+#need to define number of chrm classes within each nuc
+n1CO = 9
+
+BIG.DF.LIST_COpos_no.rank <- replicate(10,
+                  list(bind_rows(list(
+
+  pos.pool_1CO[ sample( which( pos.pool_1CO$hand.foci.count == 1), n1CO, replace = FALSE ),]
+
+) ) ) )
+
+BIG.DF.LIST_no.rank <- replicate(18,
+                  list(bind_rows(list( 
+          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
+) ) ) )
+
+```
+
+```{r sum.stat.CO.pos, echo=FALSE}
+#remove F1 without values
+#and 1CO with 2 values
+
+SC_Mus.male_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
+for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
+ # print(j)
+  #fill in the cells of given index
+  SC_Mus.male_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
+  SC_Mus.male_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank[[j]]$SC.length )
+  SC_Mus.male_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank[[j]]$SC.length )
+  }
+
+#Calq cV for no rank
+SC_Mus.male_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+  #print(i)
+  #fill in the cells of given index
+  SC_Mus.male_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank[[i]]$SC.length )
+  SC_Mus.male_no.rank$total_SC[i] <-sum(  BIG.DF.LIST_no.rank[[i]]$SC.length )
+  SC_Mus.male_no.rank$mean_SC[i] <-mean(  BIG.DF.LIST_no.rank[[i]]$SC.length )
+}
+
+```
+
+## Adding to Data set
+
+how to best increase the size of the dataset, with my bivalent data.
+From the curated BivData, figure out which can be added / into the whole cell data.
+I think there might only be a couple cell to add in.
+
+```{r getting.cell.names, echo=FALSE,eval=FALSE}
+
+#load in the cell dataset
+#28feb15_PWD_f2  from Curated_BivData
+#for a mouse, pull the 10 highest rec cells
+
+PWD.female_hand.measure = read.csv("~./MLH1repo/data/BivData/PWD_female_hand_measure.csv", header = TRUE, strip.white = TRUE)
+#"fileName"                "boxNumber"               "blobjectClass"           "chromosomeLength"        "centromere_ABS_Position" "Foci1"                   "Foci2"                  
+#"Foci3"                   "curation.notes"  
+female.mouse_bivData <- Curated_BivData[Curated_BivData$mouse == "28feb15_PWD_f2",]
+
+write.table(female.mouse_bivData, "~./MLH1repo/28feb15_PWD_f2.csv", sep=",", row.names = FALSE)
+
+PWD.female_hand.measure = read.csv("~./MLH1repo/28feb15_PWD_f2.csv", header = TRUE, strip.white = TRUE)
+PWD.female_hand.measure$Obj.ID <- paste(PWD.female_hand.measure$fileName, PWD.female_hand.measure$boxNumber, sep = "_")
+
+#merge with hand data
+#mergy <- merge(PWD.female_hand.measure, female.mouse_bivData)
+#PWD.female_hand.measure$hand.foci.count <- as.numeric(PWD.female_hand.measure$hand.foci.count)
+
+cells.BivData_table_28feb15_PWD_f2 <- ddply(.data=PWD.female_hand.measure, 
+                            .(fileName),
+                            summarize, 
+                            ncells = length(unique(fileName)),
+                            nbivs = length(unique(Obj.ID)),
+                            n.foci = sum(hand.foci.count),
+                            total.SC = sum(chromosomeLength),
+                            mean.SC = mean(chromosomeLength)
+
+)
+#add hand data
+ll <- cells.BivData_table_28feb15_PWD_f2[cells.BivData_table_28feb15_PWD_f2$nbivs > 18,]
+
+#I have 3 cells with full 20
+
+```
+
+```{r mini.biv, echo=FALSE,eval=FALSE}
+#make mini biv plots
+
+#PWD.female
+#PWD.female_hand.measure
+
+cell.1 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_10_rev",]
+cell.2 <- PWD.female_hand.measure[PWD.female_hand.measure$fileName == "13mar15_28feb15_PWD_f2_sp1_38_rev",]
+
+two_groups <- transform(cell.1, category2 = factor(paste(fileName, chromosomeLength)))
+two_groups <- transform(two_groups, category2 = reorder(category2, rank(chromosomeLength)))
+
+#this plot function takes the new df, and plots cat2(x) by SC length (y). cat2 has been ordered by the ranked length
+#of SC.
+MiniBiv_plot1 <-  ggplot(two_groups, aes(category2, chromosomeLength)) +
+    geom_bar(aes(fill= as.factor(hand.foci.count)), stat = "identity") +
+  #  geom_vline(xintercept = c(male_IFD, 1) ) +  #
+ #geom_hline(yintercept = c(male_IFD) ) +
+
+  scale_fill_manual(values=c("red", "red", "red")) +   #0CO;"#FFCC66"
+  #how to add points for 2foci
+  geom_point(aes(x=two_groups$category2, y=two_groups$Foci1),colour = "green",size = 3 ) +   
+  #foci.1.position, two_groups$distal.foci.position)
+  geom_point(aes(x=two_groups$category2, y=two_groups$Foci2),colour = "green",size = 3 ) +
+
+#add centromeres...  
+  geom_point(aes(x=two_groups$category2, y=0), colour = "purple4",size =5)+
+
+  facet_wrap( ~ fileName, scales = "free_x", ncol = 5) +
+ # scale_x_discrete(labels=two_groups$strain, breaks=two_groups$category2) +
+  ylim(c(0,200))+
+     theme(axis.line=element_blank(),axis.text.x=element_blank(),
+        axis.ticks=element_blank())  #for removeing x labels
+
+#MiniBiv_plot1
+#MiniBiv_plot2
+
+#save the files
+
+```
+
+```{r discribe.Biv.Pattern, echo=FALSE}
+#start discribing the basic
+
+#start merging curated bivData with handmeasured data
+
+#BivData - Curated
+```
+
+```{r add.more.whole.cell, echo=FALSE}
+
+#load data curatedBiv Data
+#load in hand measures
+
+```
+
+Look/use the DF, Curated_BivData data frame for mice with many cell measures and cells with whole cell measures (all bivalents have been measured).
+
+Mice with many cells and BivData
+17mar16_G_f4 - BivData
+17mar16_G_f1
+
+Chunk below has code for choosing mice with a good number of cells (25?)
+
+```{r choosing.mice, echo=FALSE, eval=FALSE}
+#Choose mice with many cell observations 
+#important for increased power for within mouse variance
+
+#G_f, G_m
+#SKIVE_f, SKIVE_m #have BivData
+
+#remove all the quality 5
+
+#make plot/histogram of the MLH1 count distribution
+counts <- MLH1_data$nMLH1.foci[MLH1_data$mouse == "31may19_SKIVE_f6"]
+
+mouse.DF <- MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",]
+mouse.DF <- mouse.DF[mouse.DF$quality < 5,]
+high.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci>=quantile(counts, 0.9, na.rm=TRUE), ]
+low.bin.DF <- mouse.DF[mouse.DF$nMLH1.foci<=quantile(counts, 0.1, na.rm=TRUE), ]
+
+ordered.MLH1 <- mouse.DF[with(mouse.DF,order(nMLH1.foci)),]#this returns ordered
+topsy <- ordered.MLH1[1:5,]
+bottomzy <- ordered.MLH1[(length(ordered.MLH1$nMLH1.foci)-5):length(ordered.MLH1$nMLH1.foci),]
+#now have top and bottom 5 cell observations
+
+#merge
+top.n.bottom.bin <- rbind(topsy,bottomzy)
+
+#write 
+write.table(top.n.bottom.bin, "~./MLH1repo/curated.bins.csv", sep=",", row.names = FALSE)
+
+shiz <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_f6",],
+       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_f6 Cell counts")+
+geom_vline(xintercept = c(min(high.bin), max(low.bin)), color= "red")
+#highest of the low bin
+#lowest value of the high bin
+
+#get the filenames for this groups of cells
+
+#10mar15_PWD_m1#10mar15_PWD_m2 -- not found in BivData, re-running all of the PWD male data
+#28feb15_PWD_f2, in PWD--female .xlsx file
+
+#31may19_SKIVE_m2
+male.one.DF <- MLH1_data[MLH1_data$mouse == "28feb15_PWD_f2",]
+male.one.DF <- male.one.DF[male.one.DF$quality < 5,]
+
+ordered.MLH1.male <- male.one.DF[with(male.one.DF,order(nMLH1.foci)),]#this returns ordered
+male.high.bin.DF <- ordered.MLH1.male[1:5,]
+male.low.bin.DF <- ordered.MLH1.male[(length(ordered.MLH1.male$nMLH1.foci)-5):length(ordered.MLH1.male$nMLH1.foci),]
+#now have top and bottom 5 cell observations
+
+#merge
+male.top.n.bottom.bin <- rbind(male.high.bin.DF, male.low.bin.DF)
+#write 
+write.table(male.top.n.bottom.bin, "~./MLH1repo/female.PWD.curated.bins.csv", sep=",", row.names = FALSE)
+
+shiz.too <- ggplot(data = MLH1_data[MLH1_data$mouse == "31may19_SKIVE_m2",],
+       aes(x=nMLH1.foci))+geom_histogram(stat = "bin", binwidth = 1)+ggtitle("31may19_SKIVE_m2 Cell counts")+
+geom_vline(xintercept = c(min(male.high.bin.DF$nMLH1.foci), max(male.low.bin.DF$nMLH1.foci)), color= "red")
+#highest of the low bin
+#lowest value of the high bin
+shiz.too
+
+```
+
+the plot above shows an example plot of --- where the pools are from, currently these are 10 quantille and 90 quantile. setting threshold by percentile doesn't always give enough cells. 
+Should I be comparing cells with the same number of COs?
+
 # Junk
 
-```{r junk, eval=FALSE}
+```{r junk, eval=FALSE, echo=FALSE}
 DF.permutations <- data.frame( nuc.sim = seq(1:24), rank1 = replicate(24, ""),rank2= replicate(24, ""),

### 2019-12-08 results/Chapter3_ideas.Rmd
fixed version of position randomization loop

@@
 
+male.whole.cell <- male.whole.cell[!(is.na(male.whole.cell$fileName) | male.whole.cell$fileName==""), ]
+
+#make sure male.whole.cell has nrm oci positions
 pos.pool_1CO <- male.whole.cell[male.whole.cell$hand.foci.count == 1,]
@@
 
+pos.pool_1CO <- pos.pool_1CO[!(is.na(pos.pool_1CO$fileName) | pos.pool_1CO$fileName==""), ]
+pos.pool_2CO <- pos.pool_2CO[!(is.na(pos.pool_2CO$fileName) | pos.pool_2CO$fileName==""), ]
+pos.pool_3CO <- pos.pool_3CO[!(is.na(pos.pool_3CO$fileName) | pos.pool_3CO$fileName==""), ]
 #now make the randomized data set for CO pos
 #do whole row -- so I can cen n
+
+male.whole.cell.mini <- male.whole.cell#[190:455, ]
+
+for(oar in 1:length(male.whole.cell.mini$fileName)){
+  print(oar)
+   if(male.whole.cell.mini$hand.foci.count[oar] == 1){
+    print("oner")
+     #random sample
+     rand.obj.id.1CO <- sample(pos.pool_1CO$Obj.ID, 1)
+
+     male.whole.cell.mini$rand.ID[oar] <- rand.obj.id.1CO
+     male.whole.cell.mini$rand.Foci1[oar] <- as.numeric(pos.pool_1CO$Foci1[which(pos.pool_1CO$Obj.ID == rand.obj.id.1CO)])
+     #  sample(pos.pool_1CO$F1, 1, replace = FALSE)
+     male.whole.cell.mini$rand.Foci2[oar] <- ""
+     male.whole.cell.mini$rand.Foci3[oar] <- ""
+   }
+
+  if(male.whole.cell.mini$hand.foci.count[oar] == 2){
+    print("twoer")
+
+    #random sampl of obj.id
+    rand.obj.id <- sample(pos.pool_2CO$Obj.ID, 1)
+
+    male.whole.cell.mini$rand.ID[oar] <- rand.obj.id
+    #asign the F1s and F2s from the fit obj.id
+
+    male.whole.cell.mini$rand.Foci1[oar] <-as.numeric(pos.pool_2CO$Foci1[which(pos.pool_2CO$Obj.ID == rand.obj.id)])
+  #  male.whole.cell.mini$rand.Foci2[oar] <- "blah"
+    male.whole.cell.mini$rand.Foci2[oar] <-as.numeric(pos.pool_2CO$Foci2[which(pos.pool_2CO$Obj.ID == rand.obj.id)])
+    male.whole.cell.mini$rand.Foci3[oar] <- ""
+ }
+
+  if(male.whole.cell.mini$hand.foci.count[oar] == 3){
+    print("threer")
+    rand.obj.id_3co <- sample(pos.pool_3CO$Obj.ID, 1)
+
+    male.whole.cell.mini$rand.ID[oar] <- rand.obj.id_3co
+
+        male.whole.cell.mini$rand.Foci1[oar] <-as.numeric(pos.pool_3CO$Foci1[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+  #  male.whole.cell.mini$rand.Foci2[oar] <- "blah"
+    male.whole.cell.mini$rand.Foci2[oar] <-as.numeric(pos.pool_3CO$Foci2[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+
+    male.whole.cell.mini$rand.Foci3[oar] <-as.numeric(pos.pool_3CO$Foci2[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+
+  }
+
+  }
 

### 2019-12-09 results/Chapter3_ideas.Rmd
rbar for rand and norm foci pos

--- 2019-12-06 results/Chapter3_ideas.Rmd
@@
 whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
-load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
+#load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
 ```
@@
 whole.cell.data <- add_IFD(whole.cell.data)
-whole.cell.data <-  add_IFD.2(whole.cell.data)
-
-Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+#whole.cell.data <-  add_IFD.2(whole.cell.data)
+
+#Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
 #IFD.2 does mulitple IFDs (just absolute)
-Curated_BivData <- add_IFD.2(Curated_BivData)
+#Curated_BivData <- add_IFD.2(Curated_BivData)
 
@@
 2. use the summary stat
+3. draw rand nrm foci position ... then map that onto the raw SC length
 
@@
 
+pos.pool_1CO <- pos.pool_1CO[!(is.na(pos.pool_1CO$fileName) | pos.pool_1CO$fileName==""), ]
+pos.pool_2CO <- pos.pool_2CO[!(is.na(pos.pool_2CO$fileName) | pos.pool_2CO$fileName==""), ]
+pos.pool_3CO <- pos.pool_3CO[!(is.na(pos.pool_3CO$fileName) | pos.pool_3CO$fileName==""), ]
 #now make the randomized data set for CO pos
@@
 
-random_pos <- data.frame(male.whole.cell$fileName, male.whole.cell$SC.length, male.whole.cell$hand.foci.count, male.whole.cell$Obj.ID  )
-
-#INITIALIZE EMPTY list of DFs
-#fill in with full DF
-
-LIST.DFs <- setNames( replicate( length(male.biv.count_table$fileName), data.frame( obj.ID = c())), male.biv.count_table$fileName)
-
-ids <-    male.whole.cell$Obj.ID[which(male.whole.cell$fileName == '14oct16_30jun16_CAST_m3_sp1_3.2')]
-
-LIST.DFs[1]$obj.ID[1:19] <- ids[1:19]
-
-#INITALIZE empty LIST (make DF in the loop)
-
-LIST <- list()
-
-#Loop for rand sampling  
-for( male.cell in male.biv.count_table$fileName){
-
-  #for each fileName, make a new DF
-  DF.in.loop <- data.frame(matrix(ncol = 1, nrow = 19))
-  #making the Within loop DF
-  DF.in.loop$X1 <- male.whole.cell$fileName[male.whole.cell$fileName ==male.cell]
-  DF.in.loop$obj.ID <- male.whole.cell$Obj.ID[male.whole.cell$fileName ==male.cell]
-  DF.in.loop$SC.length <- male.whole.cell$SC.length[male.whole.cell$fileName ==male.cell]
-  DF.in.loop$hand.foci.count <- male.whole.cell$hand.foci.count[male.whole.cell$fileName ==male.cell]
-
-  #LIST.DFs[[male.cell]]$obj.ID[1:19] <- male.whole.cell$Obj.ID[male.whole.cell$fileName == male.cell]
-    #print(male.cell)
-    n0CO = male.biv.count_table$nCO0[male.biv.count_table$fileName == male.cell] #this reference needs to be from male.biv.count_table
-    n1CO = male.biv.count_table$nCO1[male.biv.count_table$fileName == male.cell]
-    n2CO = male.biv.count_table$nCO2[male.biv.count_table$fileName == male.cell]
-    n3CO = male.biv.count_table$nCO3[male.biv.count_table$fileName == male.cell]
-
-   #rand samples for foci.pos
-    #JUST DO THIS FOR 1CO bivs
-    DF.in.loop$rand_F1
-    sample.n1CO <- sample(pos.pool_1CO$Foci1.PER, n1CO, replace = FALSE )  #sample from the cell.table
-
-    #sample a random 2CO chrm (obj.ID),    #doesn't make sense to sample a random position from 2CO pool, need to sample the linked positions, or get the 
-    #chrm observation
-   sample.n2CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n2CO, replace = FALSE )
-   sample.n3CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n3CO, replace = FALSE )
-
-   #MAP normalized positions onto SC lengths/chrms of the fileName cell obs
-   #map the 1CO rand Foci position onto 1CO SCLength
-
-   #for i in n1CO
-#   which(random_pos$fileName == male.cell)
-   random_pos$Rand.F1[which(random_pos$fileName == male.cell)][i] <- sample(pos.pool_1CO$Foci1.PER, 1, replace = FALSE)
-
-   #list of the chrm lengths for all the 1COs for the current cell obs
-   #SC.Length_1COs <-
-
-}
-
-#need to define number of chrm classes within each nuc
-n1CO = 9
-
-BIG.DF.LIST_COpos_no.rank <- replicate(10,
-                  list(bind_rows(list(
-
-  pos.pool_1CO[ sample( which( pos.pool_1CO$hand.foci.count == 1), n1CO, replace = FALSE ),]
-
-) ) ) )
-
-BIG.DF.LIST_no.rank <- replicate(18,
-                  list(bind_rows(list( 
-          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
-) ) ) )
-
-```
+#make a copy of main DF
+male.whole.cell.mini <- male.whole.cell#[190:455, ]
+
+for(oar in 1:length(male.whole.cell.mini$fileName)){
+ # print(oar)
+   if(male.whole.cell.mini$hand.foci.count[oar] == 1){
+    print("oner")
+     #random sample -- there could be replicates in the random sample
+     rand.obj.id.1CO <- sample(pos.pool_1CO$Obj.ID, 1)
+
+     male.whole.cell.mini$rand.ID[oar] <- rand.obj.id.1CO
+     male.whole.cell.mini$rand.Foci1[oar] <- as.numeric(pos.pool_1CO$Foci1.PER[which(pos.pool_1CO$Obj.ID == rand.obj.id.1CO)])
+     #  sample(pos.pool_1CO$F1, 1, replace = FALSE)
+     male.whole.cell.mini$rand.Foci2[oar] <- ""
+     male.whole.cell.mini$rand.Foci3[oar] <- ""
+   }
+
+  if(male.whole.cell.mini$hand.foci.count[oar] == 2){
+   # print("twoer")
+
+    #random sampl of obj.id
+    rand.obj.id <- sample(pos.pool_2CO$Obj.ID, 1)
+
+    male.whole.cell.mini$rand.ID[oar] <- rand.obj.id
+    #asign the F1s and F2s from the fit obj.id
+
+    male.whole.cell.mini$rand.Foci1[oar] <-as.numeric(pos.pool_2CO$Foci1.PER[which(pos.pool_2CO$Obj.ID == rand.obj.id)])
+  #  male.whole.cell.mini$rand.Foci2[oar] <- "blah"
+    male.whole.cell.mini$rand.Foci2[oar] <-as.numeric(pos.pool_2CO$Foci2.PER[which(pos.pool_2CO$Obj.ID == rand.obj.id)])
+    male.whole.cell.mini$rand.Foci3[oar] <- ""
+ }
+
+  if(male.whole.cell.mini$hand.foci.count[oar] == 3){
+   # print("threer")
+    rand.obj.id_3co <- sample(pos.pool_3CO$Obj.ID, 1)
+
+    male.whole.cell.mini$rand.ID[oar] <- rand.obj.id_3co
+
+        male.whole.cell.mini$rand.Foci1[oar] <-as.numeric(pos.pool_3CO$Foci1.PER[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+  #  male.whole.cell.mini$rand.Foci2[oar] <- "blah"
+    male.whole.cell.mini$rand.Foci2[oar] <-as.numeric(pos.pool_3CO$Foci2.PER[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+
+    male.whole.cell.mini$rand.Foci3[oar] <-as.numeric(pos.pool_3CO$Foci3.PER[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+
+  }
+
+  }
+
+```
+
+### merge daata
+
+```{r merging.data,echo=FALSE}
+
+#melt the nrmfoci.pos, for real and randomized data
+REAL.melt.whole.cell.mini <- melt(male.whole.cell.mini,
+              c("Obj.ID","SC.length", "hand.foci.count","Foci1.PER","Foci2.PER","Foci3.PER"),
+
+              id=c("Obj.ID","SC.length", "hand.foci.count")
+                                  )
+REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[!(is.na(REAL.melt.whole.cell.mini$value) | REAL.melt.whole.cell.mini$value==""), ]
+
+REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[ (REAL.melt.whole.cell.mini$variable != "hand.foci.count"), ]
+REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[ (REAL.melt.whole.cell.mini$variable != "Obj.ID"), ]
+REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[ (REAL.melt.whole.cell.mini$variable != "SC.length"), ]
+
+#make random melted DF
+RAND.melt.whole.cell.mini <- melt(male.whole.cell.mini, c("rand.ID","SC.length","hand.foci.count","rand.Foci1","rand.Foci2","rand.Foci3"),
+                                  id=c("rand.ID", "SC.length","hand.foci.count")
+)
+RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[!(is.na(RAND.melt.whole.cell.mini$value) | RAND.melt.whole.cell.mini$value==""), ]
+
+RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[ (RAND.melt.whole.cell.mini$variable != "hand.foci.count"), ]
+RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[ (RAND.melt.whole.cell.mini$variable != "rand.ID"), ]
+RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[ (RAND.melt.whole.cell.mini$variable != "SC.length"), ]
+
+```
+
+```{r merge.real.rand, echo=FALSE}
+
+REAL.melt.whole.cell.mini$type <- "real"
+RAND.melt.whole.cell.mini$type <- "sim"
+
+colnames(RAND.melt.whole.cell.mini) <- c("Obj.ID","SC.length", "hand.foci.count","variable","value", "type" )
+
+Position.DF <- rbind(REAL.melt.whole.cell.mini,RAND.melt.whole.cell.mini)
+
+Position.DF$value <- as.numeric(Position.DF$value )
+Position.DF$SC.length <- as.numeric(Position.DF$SC.length )
+
+```
+
+### plot relative foci positions
+
+```{r Pos.plot, echo=FALSE}
+
+#density or histogram plot facet hand foci count, plot value, fill type
+
+pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)
+
+pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_histogram()
+
+pos.plot.all <- ggplot(Position.DF, aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count,ncol = 1)+ggtitle("Simulated and normalized positions, male mus")
+
+pos.plot.all
+
+pos.plot.one <- ggplot(Position.DF[Position.DF$hand.foci.count == 2,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count)+ggtitle("Simulated and normalized positions")
+
+HS.density.plot <- ggplot(Hs.CO_cVs.DF, aes(x = COs, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Human male")+xlim(c(0,100))
+
+```
+
+The above distributions of relative foci position compare the real data pink to randomized data set of relative positions where for each bivalent observations the normalized position was replaced with positions from a randomly draw bivalent with the same CO number.
+For multi co bivalents relative foci positions were kept together.
+For example, a random 2CO bivalent choosen and the normalize foci1 and foci2 for that bivalent were used.
+
+In terms of normalized foci positions, the empiracal distribution isn't very different from a randomized distribution.
+
+Not exactly sure if this is the right randomization procedure, the same 'random' bivalent could be drawn more than once.
+
+Next, map the randomized positions onto the real sc.lengths for cell cell observation.
+
+### Map normalized positions onto real SC
+
+```{r map.rand2real, echo=FALSE}
+
+#need sc length from the original
+Position.DF$raw.foci.pos <- Position.DF$SC.length *Position.DF$value
+
+male.whole.cell.mini$raw.rand_Foci1 <- male.whole.cell.mini$SC.length *male.whole.cell.mini$rand.Foci1
+male.whole.cell.mini$raw.rand_Foci2 <- male.whole.cell.mini$SC.length *male.whole.cell.mini$rand.Foci2
+male.whole.cell.mini$raw.rand_Foci3 <- male.whole.cell.mini$SC.length *male.whole.cell.mini$rand.Foci3
+
+male.whole.cell.mini$rand.Foci2 <- as.numeric(male.whole.cell.mini$rand.Foci2 )
+male.whole.cell.mini$rand.Foci3 <- as.numeric(male.whole.cell.mini$rand.Foci3)
+
+#calq telo distnce
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+for(y in 1:length(male.whole.cell.mini)  ){
+  male.whole.cell.mini$rbar[y] <- calq.intra.rbar(male.whole.cell.mini[y,])
+}
+
+for(rr in 1:length(male.whole.cell.mini)  ){
+  male.whole.cell.mini$rand.rbar[rr] <- calq.intra.rbar_rand(male.whole.cell.mini[rr,])
+}
+
+mapp.cell.wide <- add_mouse(mapp.cell.wide)
+mapp.cell.wide <- add_strain(mapp.cell.wide)
+
+#cell wide telomere covariance bias
+mapp.cell.wide <-  ddply(.data=male.whole.cell.mini,
+                 .(fileName, hand.foci.count),
+                 summarize,
+                 nbiv = length(SC.length),
+                 tot.SC = sum(SC.length),
+                 mean.pos = mean(Foci1.PER),
+                 mean.rand.pos = mean(rand.Foci1),
+                 mean.raw = mean(c(Foci1,Foci2, Foci3), na.rm=TRUE ),
+                 mean.raw.rand = mean( c(raw.rand_Foci1,raw.rand_Foci2,raw.rand_Foci3), na.rm=TRUE ),
+
+                var.raw = var(c(Foci1,Foci2, Foci3), na.rm=TRUE ),
+                 var.raw.rand = var( c(raw.rand_Foci1,raw.rand_Foci2,raw.rand_Foci3), na.rm=TRUE ),
+
+                mean_rbar = mean(rbar),
+                mean_rand.rbar = mean(rand.rbar)
+)
+
+maisel <- ggplot(mapp.cell.wide, aes(x=mean_rbar, y=mean_rand.rbar))+geom_point()+xlim(c(0,.2))+ylim(c(0,.2))
+
+oopp <- ggplot(mapp.cell.wide, aes(y=mean.raw, x=mean.raw.rand, color=mouse) ) +geom_point()+ggtitle("average foci position by cell, real vs randomized")
+oopp
+
+hh <- ggplot(mapp.cell.wide, aes(y=var.raw,x=var.raw.rand) ) +geom_point()+ggtitle("variance in foci position (per cell), between real and randomized")
+
+#what is the best way to plot these?
+
+#gg <- ggplot(Position.DF, aes(y=raw.foci.pos, x=variable, color= as.factor(hand.foci.count) ))+geom_jitter()+ggtitle("raw foci positions, reals and rand")
+#gg
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+```
+
+distributions don't look different. 
+Since I randomized the cell effect -- I need to map back into a cell wide summary to effectively compare the differences.
+
+NEED to Calculate cell-wide telomere bias (use a ddplyr thing)
+
+some of the plots compare the cell wide measures from real and randomized (I think this difference is what I should calculate for permuting or randomizing)
+
+points closer to 1:1 line have foci position features near the average of the group
 

### 2019-12-09 src/CommonFunc_MLH1repo.R

@@ -424,16 +424,16 @@ calq.intra.rbar <- function(chrm_row){
   # rrbar <- c()
   for(o in 1:length(chrm_row$fileName) ){
     #print(chrm_row$fileName[o])
-    if(grepl(0, chrm_row$handFoci[o])){
+    if(grepl(0, chrm_row$hand.foci.count[o])){
       #print("fist has 0 CO")
       #rrbar[o] <- NA
       rrbar <- 0
     }
-    if(grepl(1, chrm_row$handFoci[o])){  #do the block here
+    if(grepl(1, chrm_row$hand.foci.count[o])){  #do the block here
       #print("fist has 1 CO")
       #rrbar[o] <- "1CO"
-      portion1 <- chrm_row$F1[o]
-      portion2  <- chrm_row$chromosomeLength[o] - portion1
+      portion1 <- chrm_row$Foci1[o]
+      portion2  <- chrm_row$SC.length[o] - portion1
       portion.sum <- portion1+portion2
       rel.portion1 <- (portion1 / portion.sum)
       rel.portion2 <- (portion2 / portion.sum)
@@ -446,11 +446,11 @@ calq.intra.rbar <- function(chrm_row){
       rrbar <- 0.5*(1-chrm.sq)
       #o = o +1
     }
-    if(grepl(2, chrm_row$handFoci[o])){  #do the block here
+    if(grepl(2, chrm_row$hand.foci.count[o])){  #do the block here
       #print("this has 2 CO")
-      portion1 <- chrm_row$F1[o]
-      portion2  <- chrm_row$F2[o] - chrm_row$F1[o]
-      portion3 <- chrm_row$chromosomeLength[o] - chrm_row$F2[o]+1
+      portion1 <- chrm_row$Foci1[o]
+      portion2  <- chrm_row$Foci2[o] - chrm_row$Foci1[o]
+      portion3 <- chrm_row$SC.length[o] - chrm_row$Foci2[o]+1
       portion.sum <- portion1+portion2+portion3 #in 2CO portions 
       rel.portion1 <- (portion1 / portion.sum)
       rel.portion2 <- (portion2 / portion.sum)
@@ -467,11 +467,11 @@ calq.intra.rbar <- function(chrm_row){
       #rbar[o] <- 0.5*(1-chrm.sq)
       #o = o +1
     }
[…]

### 2019-12-09 results/Chapter3_ideas.Rmd
a bit more work

@@
 
-maisel <- ggplot(mapp.cell.wide, aes(x=mean_rbar, y=mean_rand.rbar))+geom_point()+xlim(c(0,.2))+ylim(c(0,.2))
+maisel <- ggplot(mapp.cell.wide, aes(x=mean_rbar, y=mean_rand.rbar))+geom_point()+xlim(c(0,.3))+ylim(c(0,.3))
+maisel
 
@@
 points closer to 1:1 line have foci position features near the average of the group
+
+the random r bar is higher than the real r bar positions. meaning on average the male cell rbar values are lower (less to the middle and more to the ends) -- would this be the same as in female cells?
 

### 2019-12-09 src/Figures.R
changed formate of this file

@@ -1,17 +1,0 @@
----
-title: "Figures"
-author: "April Peterson"
-date: 
-output: pdf_document
----
-
-```{r setup, include=FALSE}
-
-# Common Figures
-# input: RData file for repo
-# output: Figures 
-
-
-knitr::opts_chunk$set(echo = TRUE)
-```
-

### 2019-12-09 results/Chapter3_ideas99.Rmd
updates to the chapter 3 analysis and the function script

--- 2019-12-09 results/Chapter3_ideas.Rmd
@@
 
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
 #read files
@@
 
-# Summary and of Wang et al 2019
-
-Main take away from Wang et al 2019 is that chromosomes with the same nucleus covary in number of crossovers.
+# Within Nuc covariation
+
+Main take away from Wang et al 2019 is that chromosomes with the same nucleus co-vary in number of crossovers.
 Axis lengths / SC lengths show the same pattern and is known to be regulated at the cell wide level.
-The authors suggest that the within nucleus covariation in CO per chromosome is mediated through cell wide regulation of SC area.
+The authors suggest that the within nucleus co-variation in CO per chromosome is mediated through cell wide regulation of SC area.
 (the lengths of SC for chromosomes within a cell a correlated, there is more area for COs).
@@
 Data used in the paper
-human data- Gruhn et al 2013 (ploted below, 755 cells)
+human data- Gruhn et al 2013 (plotted below, 755 cells)
 Sun et al 2006
@@
 
-Was there originall CO position data for this human data? (MLH1 spreads) (there is axis length and CO count per chrm -- this should mean there is position )
+Was there original CO position data for this human data? (MLH1 spreads) (there is axis length and CO count per chrm -- this should mean there is position )
 
@@
 
-approach -- they compared covariation (cV) of per chrm CO count and SC length for permuted/randomized groupd of chromosome observations representing 'in silico' nuclei.
-
-## Redoing the permutaitons
-
-Cell summaries below
+## Permutations of Human Data
+
+Heads of the cell-wide summary statistics
 
@@
 ```
+
+I ran two types of permutations for the CO count per cell data.
+'sim no rank' is a dataset where chromosome CO counts were drawn from the entire pool of chromosome observations.
+'Sim w.rank' is a randomized pool, but each cell is made of random draws from one of each of the chromosome bins.
+
+I didn't test the axis length trait yet.
 
@@
 HS.scatter.plot <- ggplot(Hs.CO_cVs.DF, aes(y = COs, x=type) )+ geom_jitter()+ ggtitle("COs per cell, Human male")
-HS.scatter.plot
-
-#make density distributions
-#percent of cells
-
-#quantilles of COs
 
 hist <- ggplot(Hs.CO_cVs.DF, aes(x = COs)) + geom_histogram(aes(fill=type), binwidth = 2 )
-hist
 
@@
 
-To calculate the variability index, authors take percent of cells(integrated curve area) which differ between the empiracal and simulated distributions.
-
-I changed the number of permutations to 200, which should be the same number of oberved cells, but it looks like the real data is much larger....the sheet has 755 cell observations, not 200.
-This might be due to how histograms are presented.
-
-The rank distinction is to test if sampling within pools of specific chromsomes makes a difference.
-The smaller/tighter distribution between CO count per nuc with the ranked and non-ranked distributions indecate there is a small chromosome specific effect.
-
-Not sure how to calculate the differences in distributions yet.
-
-this current data shows random samples, not permuted random samples.
+Qualitatively these results look the same as in the Wang et al 2019 paper.
+I haven't confirmed the calculations with a variability index.
+To calculate the 'variability index', the authors take the percentage of cells (or integrated curve area) which differ between the empirical and simulated distributions.
+I still need to do some work to make sure I can calculate the differences in distributions.
+
+The rank distinction is to test if sampling within pools of specific chromosomes makes a difference.
+The smaller/tighter distribution between CO count per nuc with the ranked and non-ranked distributions indicate there is a small chromosome specific effect.
 
@@
 
-```{r setup.whole.cell.data, echo=FALSE}
-
-#read in csv of whole cell hand measures
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-#setwd("C:/Users/April/Desktop/MLH1repo")
-#load(file="data/MLH1/MLH1_data_setup.RData")
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+```{r setup.whole.cell.data, echo=FALSE, warning=FALSE}
+
 #add sex since there is a difference in sample size
@@
 
-#make sure foci.count is number
-whole.cell.data$hand.foci.count <- as.character(whole.cell.data$hand.foci.count)
-whole.cell.data$hand.foci.count <- as.numeric(whole.cell.data$hand.foci.count)
+whole.cell.data$hand.foci.count <- as.numeric(as.character(whole.cell.data$hand.foci.count))
+whole.cell.data$SC.length <- as.numeric(as.character(whole.cell.data$SC.length))
+
+whole.cell.data$F1 <- as.numeric(as.character(whole.cell.data$F1))
+whole.cell.data$F2 <- as.numeric(as.character(whole.cell.data$F2))
+whole.cell.data$F3 <- as.numeric(as.character(whole.cell.data$F3))
 
@@
 #add IFDs and normalized positions 
-whole.cell.data$Foci1.PER <- whole.cell.data$F1 / whole.cell.data$SC.length
-whole.cell.data$Foci2.PER <- whole.cell.data$F2 / whole.cell.data$SC.length
-whole.cell.data$Foci3.PER <- whole.cell.data$F3 / whole.cell.data$SC.length
+whole.cell.data$Foci1.PER <- whole.cell.data$Foci1 / whole.cell.data$SC.length
+whole.cell.data$Foci2.PER <- whole.cell.data$Foci2 / whole.cell.data$SC.length
+whole.cell.data$Foci3.PER <- whole.cell.data$Foci3 / whole.cell.data$SC.length
 
@@
 
-#Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
-#IFD.2 does mulitple IFDs (just absolute)
-#Curated_BivData <- add_IFD.2(Curated_BivData)
-
-#find cells with the wrong number
-#whole.cell.data_cell.table <-
-
-#ADD THE RANK
+#ADD THE RANK -- this stoped working for a bit, 
+#fixed it with rank
 for(musy in as.character(unique(whole.cell.data$fileName))){
   #print(musy)
-
   #order the chrm lengths -- this is order
-  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
-  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
-}
-#I think I saved the Excel with sorted SC lengths
-#Yay I printed the rank in this
-
-```
-
-To make the same comparisons, I need chromosome/bivalent level observations for whole cells (all chromosomes). 
-I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations
+  chrm.order = rank(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
+  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order   #this is just a string of numbers--
+
+}
+
+```
+
+I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations by categories.
+
+The whole cell comparisons (from empirical and randomized distributions):
+- CO counts per cell
+- Total SC per cell (calculated by single bivalents)
+- CO position (cell wide rbar, mean telomere bias)
+
+The CO position metrics might change.
 
 ```{r whole.cell.table, echo=FALSE}
-
 table(whole.cell.data$category)
-
 #table(cell.table$category)
@@
 
+```{r make.nother.table, echo=FALSE}
+
+whole.cell.biv.count_table <-  ddply(.data=whole.cell.data,
+                 .(fileName),
+                 summarize,
+                 nbiv = length(SC.length),
+                  #THIS IS WHERE THE error is thrown
+                 nCO0 =  sum(which(hand.foci.count == 0), na.rm=TRUE)
+                 #SUM not meaningful for factors
+)
+
+                 #  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+            #      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE),# 
+             #     nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+            #      nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+
+          #       totCO = sum(hand.foci.count),
+          #       total.SC = sum(SC.length),
+          #       mean_co = mean(hand.foci.count, na.rm = TRUE),
+          #       mean_sc = mean(SC.length),
+              #     var_co = var(hand.foci.count),
+              #     var_sc = var(SC.length),
+              #   cv_co.count = cv(hand.foci.count),
+             #    cv_SC.length = cv(SC.length)
+#)
+
+```
+
 ```{r cell.level.tables, echo=FALSE}
-
 whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
+
 male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
-#ncells
-#length(unique(male.whole.cell$fileName) ) #24 cells
 
 female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
-#length(unique(female.whole.cell$fileName) ) #18 cells
 
@@
 
-                 CO = sum(hand.foci.count),
+                 totCO = sum(hand.foci.count),
                  total.SC = sum(SC.length),
-                 mean_co = mean(hand.foci.count, rm.na = TRUE),
+                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                  mean_sc = mean(SC.length),
@@
                  nbiv = length(SC.length),
-
                  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
@@
 
-                 CO = sum(hand.foci.count),
-                  total.SC = sum(SC.length),
-                 mean_co = mean(hand.foci.count),
-                 mean_sc = mean(SC.length),
+                 totCO = sum(hand.foci.count, na.rm = TRUE),
+                  total.SC = sum(SC.length, na.rm = TRUE),
+                 mean_co = mean(hand.foci.count, na.rm = TRUE),
+                 mean_sc = mean(SC.length, na.rm = TRUE),
                    var_co = var(hand.foci.count),
@@
                  cv_SC.length = cv(SC.length)
-)#one of the female cells is missing values
-
-```
-
-I calulate within cell ranked SC length to use rank as a proxy for chromsome ID for bivalents.
+)
+
+```
+
+I calculate within cell ranked SC length to use rank as a proxy for chromosome ID for bivalents.
 
@@
 
-```{r play.plots, echo=FALSE}
+```{r play.plots, echo=FALSE, warning=FALSE}
 #female.whole.cell
@@
 
-rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=SC.length)) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")
+rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=as.numeric(SC.length) )) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")
 
@@
 
-# PERMUTATION SETUP
-
-```{r per.co.count, echo=FALSE, eval=FALSE}
-#I don't think ... I end up useing these (matrics / DFs)
-iterations = 24
-variables = 20
-DF.permutations_CO.count <- matrix(ncol=variables, nrow=iterations)
-colnames(DF.permutations_CO.count) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
-                               "rank6","rank7","rank8","rank9","rank10","rank11",
-                               "rank12","rank13","rank14","rank15","rank16",
-                               "rank17","rank18","rank19" )
-
-for(row.num in 1:24){
-
-  #print(c(row.num, "rowis") )
-  for(col.num in 2:ncol(DF.permutations_CO.count) ){
-
-  rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
-  #print(SCll)   
-  DF.permutations_CO.count[row.num,col.num] <- rand.rank.co
-   }
-  }
-DF.permutations_CO.count <- as.data.frame(DF.permutations_CO.count)
-
-#no rank - iterate / permute the sampling X times
-its = 24
-varis = 20
-DF.permutations.CO.no_rank <- matrix(ncol=varis, nrow=its)
-
-#make LIST of X DATAFRAMES
-
-for(row.num in 1:24){
-  #print(c(row.num, "rowis") )
-  for(col.num in 2:ncol(DF.permutations.CO.no_rank) ){
-  rand.co <- sample(male.whole.cell$hand.foci.count, 1, replace = FALSE)
-  #print(SCll)   
-  DF.permutations.CO.no_rank[row.num,col.num] <- rand.co
-   }
-  }#yes this finally works!!!!
-DF.permutations.CO.no_rank <- as.data.frame(DF.permutations.CO.no_rank)
-
-#
-for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
-  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
-  }
-
-#merge with real
-#co.var.for.plot <- matrix(ncol=3, nrow=72)
-#co.var.for.plot <- as.data.frame(co.var.for.plot)
-#colnames(co.var.for.plot) <- c("cell.id","var","type")
-
-#co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
-#co.var.for.plot$type[1:24] <- "sim"
-
-#co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
-#co.var.for.plot$type[25:48] <- "real"
-
-#co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
-#co.var.for.plot$type[49:72] <- "sim_no_rank"
-
-#plot
-#whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
-#whole.cell.var_CO.plot
-
-```
+The plots above look much different than when I first made them. The SC values might not be right.
+
+# Mus Permutation Setup
+
+I will make the permuted data sets once and display the summary stats (and compare the empirical to random distribution).
+
+I will us the same rank and non-ranked version of permutations as I did for the human data.
+Since I don't have chromosome IDs, I will use the within cell SC length ranks as a proxy for chromosome identity (rank).
 
@@
 
-Make list of DFs to fill with permutations
-
-```{r list.of.DF, echo=FALSE}
-
-#dataframes to recreate -- and fill with 
-#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)
-
-#make subsampled piece of DF here -- these are taking the whole DF
-#I will just take single values
-#Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
-#male.whole.cell
-#rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
-
+Theses code chunks make a data set of 24 nuclei with 19 random chromosome draws each.
+
+```{r list.of.DF, echo=FALSE, warning=FALSE}
 #rank (sample all of the ranks then stich them together)
 #replication should match the num
-BIG.LIST_w.rank <- replicate(18,
+BIG.LIST_w.rank <- replicate(24,
   list(bind_rows(list( 
@@
 #makes a list of 10 -- [[]]
-BIG.DF.LIST_no.rank <- replicate(18,
+BIG.DF.LIST_no.rank <- replicate(24,
                   list(bind_rows(list( 
@@
 
-```{r sum.stat, echo=FALSE}
+```{r sum.stat, echo=FALSE, warning=FALSE}
 #dealing with lists of DFs, 
@@
 #then add the real values
-real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count,male.biv.count_table$CO )
+real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count, male.biv.count_table$totCO )
 real.vals.cv.co$type <- "real"
@@
 #  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") 
-
 #biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
-
-```
-
-```{r perm.plots, echo=FALSE}
+```
+
+```{r perm.plots, echo=FALSE, warning=FALSE}
 #plots for permutations
@@
 
-Above plot shows overdispersion of CO counts per cell for my Mus male dataset.
+Above plot shows over dispersion of CO counts per cell for my Mus male data set.
 The above plot the real data might have more variance because it's a combination of strains
@@
 
-```{r rand.DF_CO.count.female, echo=FALSE}
-
-#dataframes to recreate -- and fill with 
-#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)
+These code chunks make a randomized data set of 18 nuclei with 20 random chromosome draws each.
+
+```{r rand.DF_CO.count.female, echo=FALSE, warning=FALSE}
 
@@
 
-These permuted dataframes can be reused for all the meiotic traits, (CO.count, SC length, and position).
-
-```{r sum.stat_cv.CO, echo=FALSE}
+```{r sum.stat_cv.CO, echo=FALSE, warning=FALSE}
 #CALUCULATING SUMMARY STAT from the permuted DFs
@@
 #then add the real values
-real.vals.cv.co_female <- data.frame(cv = female.biv.count_table$cv_co.count, CO = female.biv.count_table$CO)
+real.vals.cv.co_female <- data.frame(cv = female.biv.count_table$cv_co.count, CO = female.biv.count_table$totCO)
 real.vals.cv.co_female$type <- "real"
@@
 
-```{r female.plot, echo=FALSE}
-
-```
-
-Density plot above also shows that the female CO count data is overdispersed compered to a randomized distribution.
+Density plot above also shows that the female CO count data is over dispersed compered to a randomized distribution.
 
@@
 
-Start doing the same thing but for bivalent SC lengths.
-Use the same randomized data sets!
-summary stats - mean SC length, cV SC length, total SC area
-
-don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.
-
-```{r SC_Sum.Stats.male, echo=FALSE}
+Comparing cell wide statistics for SC lengths across the real and randomized data sets.
+
+```{r SC_Sum.Stats.male, echo=FALSE, warning=FALSE}
 # calq the sum stats
@@
 
-```{r SC.plots_Mus.male, echo=FALSE}
+```{r SC.plots_Mus.male, echo=FALSE, warning=FALSE}
 Mus_SC_hist <- ggplot(SC_Mus_male, aes(x = tot.SC)) + geom_histogram(aes(fill=type), binwidth = 50 )+ggtitle("total SC per cell, Mus male")
@@
 
-```{r SC_sum.stats.female, echo=FALSE, eval=FALSE}
+```{r SC_sum.stats.female, echo=FALSE}
 
@@
 
-#Calq cV for no rank
-SC_Mus.female_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female.) ) ) #first initialize the DF results will go in outside of loop
-for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
+#Calq cV for no rank,   BIG.DF.LIST_no.rank_female
+SC_Mus.female_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
+for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
   #print(i)
   #fill in the cells of given index
-  SC_Mus.female_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank_female.[[i]]$SC.length )
-  SC_Mus.female_no.rank$total_SC[j] <-sum(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
-  SC_Mus.female_no.rank$mean_SC[j] <-mean(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
-}
+  SC_Mus.female_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$SC.length )
+  SC_Mus.female_no.rank$total_SC[i] <-sum(  BIG.DF.LIST_no.rank_female[[i]]$SC.length, na.rm = TRUE )
+  SC_Mus.female_no.rank$mean_SC[i] <-mean(  BIG.DF.LIST_no.rank_female[[i]]$SC.length, na.rm=TRUE )
+}
+
+#10 and 10 rows from the permuted
+SC_Mus.female_w.rank$type <- "sim w.rank"
+colnames(SC_Mus.female_w.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )
+
+#this DF is empty
+SC_Mus.female_no.rank$type <- "sim_no.rank"
+colnames(SC_Mus.female_no.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )
+
+real_SC.Mus_female <- data.frame(female.biv.count_table$var_sc,  female.biv.count_table$total.SC, female.biv.count_table$mean_sc )
+real_SC.Mus_female$type <- "real"
+
+colnames(real_SC.Mus_female) <- c("SC.cV", "tot.SC", "mean.SC", "type")
+#merge
+SC_Mus_female <- rbind(SC_Mus.female_w.rank, SC_Mus.female_no.rank, real_SC.Mus_female)
+
+Mus_SC_density_female <- ggplot(SC_Mus_female, aes(x = tot.SC, fill=type) )+geom_density(alpha=.6) + ggtitle("Total SC per cell, Mus female")
+
+Mus_SC_density_female
 
@@
 
-```{r permuted.v.real, echo=FALSE, eval=FALSE}
-for( o in 1:nrow(DF.permutations_CO.count) ) {
-  DF.permutations_CO.count$var_CO.count[o] <- var(as.numeric(DF.permutations_CO.count[o,2:20]))
-  }
-
-#two data frames of within cell var
-#cell.id, var, type
-var.for.plot <- matrix(ncol=3, nrow=48)
-var.for.plot <- as.data.frame(var.for.plot)
-colnames(var.for.plot) <- c("cell.id","var","type")
-
-var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
-var.for.plot$type[1:24] <- "sim"
-
-var.for.plot$var[24:48] <- male.biv.count_table$var_co
-var.for.plot$type[24:48] <- "real"
-
-#my permuted DFs are in a strange str
-
-#plot within cell variance
-whole.cell.var.plot  <- ggplot(var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell variance between types")
-
-whole.cell.var.plot
-```
+The distribution of the female SC lengths isn't normal. Double check these values -- if their hand measured, not sure why 1 has much lower sc lengths.
+
+I think the difference in variance of the rank and non-rank suimulations shows that chromsomes effects play a bigger role in shaping the distributions of cell wide SC area.
 
@@
 
-don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.
-
-look at CO position metrics.
-Do I need to melt the plots 
-the position col names are also, messed up I need to fix them --
-
-1. use the same randomized dataframe
-2. use the summary stat
-3. draw rand nrm foci position ... then map that onto the raw SC length
-
-How do I randomize CO pos for a chrm?
-I want to draw positions from chrm_classes
-
-For each cell observation, note the classes,  I want to 
-make bins for positions of all
+I am still working on the metrics to compare (and the ones that are best for cell wide descriptions).
+- cell wide telomere bias
+- cell wide rbar (smaller values mean more equal chrm proportions between foci. )
+- average IFD
+
+(should these metrics should be weighted by CO count?)
+
+## Male CO Position
+
+Comparing cell wide statistics for CO position metrics across the real and randomized data sets.
+
+The permuted/randomized data set is made by keeping CO count and SC lengths constant, but replacing the normalized foci position with a random normalized foci position(s) drawn from a pool of the same CO counts.
+The randomized relative positions are converted into raw/absolute positions (for the original SC length of that bivalent observation).
 
@@
 #make a copy of main DF
-male.whole.cell.mini <- male.whole.cell#[190:455, ]
-
+male.whole.cell.mini <- male.whole.cell
+
+#filling in with randomized foci positions
 for(oar in 1:length(male.whole.cell.mini$fileName)){
@@
    if(male.whole.cell.mini$hand.foci.count[oar] == 1){
-    print("oner")
+    #print("oner")
      #random sample -- there could be replicates in the random sample
@@
 
-### merge daata
+### merge data
 
@@
 
+male.whole.cell.mini$SC.length <- as.numeric(as.character(male.whole.cell.mini$SC.length))
+male.whole.cell.mini$rand.Foci2 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci2))
+male.whole.cell.mini$rand.Foci3 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci3))
+
+male.whole.cell.mini$raw.rand_Foci1 <- male.whole.cell.mini$rand.Foci1 * male.whole.cell.mini$SC.length
+male.whole.cell.mini$raw.rand_Foci2 <- male.whole.cell.mini$rand.Foci2 * male.whole.cell.mini$SC.length
+male.whole.cell.mini$raw.rand_Foci3 <- male.whole.cell.mini$rand.Foci3 * male.whole.cell.mini$SC.length
+
 ```
@@
 
-### plot relative foci positions
-
-```{r Pos.plot, echo=FALSE}
-
+```{r maw.blah, echo=FALSE}
+#cell.wide.rbar
+#calq telo distnce
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+for(y in 1:length(male.whole.cell.mini)  ){
+  male.whole.cell.mini$rbar[y] <- calq.intra.rbar(male.whole.cell.mini[y,])
+}
+
+for(rr in 1:length(male.whole.cell.mini)  ){
+  male.whole.cell.mini$rand.rbar[rr] <- calq.intra.rbar_rand(male.whole.cell.mini[rr,])
+}
+
+#make a new DF
+#
+real_rbar.Mus <- data.frame(male.whole.cell.mini$rbar )
+real_rbar.Mus$type <- "real"
+colnames(real_rbar.Mus) <- c("rbar", "type" )
+
+rand_rbar.Mus <- data.frame(male.whole.cell.mini$rand.rbar )
+rand_rbar.Mus$type <- "rand"
+colnames(rand_rbar.Mus) <- c("rbar",  "type" )
+
+#merge
+Mus_rbar_male <- rbind(rand_rbar.Mus, real_rbar.Mus)
+
+#
+density.plot_rbars <- ggplot(Mus_rbar_male, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Bivalent Rbar distributions, Mus male")+xlim(c(0,.3))
+density.plot_rbars
+
+mapp.cell.wide_cells <-  ddply(.data=male.whole.cell.mini,
+                 .(fileName),
+                 summarize,
+                 nbiv = length(SC.length),
+                 tot.SC = sum(SC.length),
+
+                var.raw = var(c(Foci1,Foci2, Foci3), na.rm=TRUE ),
+                var.raw.rand = var( c(raw.rand_Foci1,raw.rand_Foci2,raw.rand_Foci3), na.rm=TRUE ),
+
+                mean_rbar = mean(rbar),
+                mean_rand.rbar = mean(rand.rbar),
+                sum_rbar = sum(rbar),
+                sum_rand.rbar = sum(rand.rbar)
+
+)
+
+real_cellwide_rbar.Mus <- data.frame(mapp.cell.wide_cells$mean_rbar )
+real_cellwide_rbar.Mus$type <- "real"
+colnames(real_cellwide_rbar.Mus) <- c("rbar", "type" )
+
+rand_cellwide_rbar.Mus <- data.frame(mapp.cell.wide_cells$mean_rand.rbar )
+rand_cellwide_rbar.Mus$type <- "rand"
+colnames(rand_cellwide_rbar.Mus) <- c("rbar",  "type" )
+
+#merge
+cell.wide.Mus_rbar_male <- rbind(real_cellwide_rbar.Mus, rand_cellwide_rbar.Mus)
+
+#
+density.plot_cell.wide.rbars <- ggplot(cell.wide.Mus_rbar_male, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Cell wide mean Rbar distributions, Mus male")+xlim(c(0,.3))
+density.plot_cell.wide.rbars
+
+real_tot.cell_rbar.Mus <- data.frame(mapp.cell.wide_cells$sum_rbar )
+real_tot.cell_rbar.Mus$type <- "real"
+colnames(real_tot.cell_rbar.Mus) <- c("sum.rbar", "type" )
+
+rand_tot.cell_rbar.Mus <- data.frame(mapp.cell.wide_cells$sum_rand.rbar )
+rand_tot.cell_rbar.Mus$type <- "rand"
+colnames(rand_tot.cell_rbar.Mus) <- c("sum.rbar",  "type" )
+#merge
+cell.wide.tot.rbar_Musmale <- rbind(real_tot.cell_rbar.Mus, rand_tot.cell_rbar.Mus)
+
+#
+density.plot_cell.wide.rbars <- ggplot(cell.wide.tot.rbar_Musmale, aes(x = sum.rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Cell wide total Rbar distributions, Mus male")+xlim(c(0,4))
+
+density.plot_cell.wide.rbars
+
+```
+
+The random relative foci positions were 'mapped' onto each of the SC length observations (foci * SC.length).
+
+The first plot shows the 2 disttributions of all the rbar measures (across all bivalents)
+The second plot shows the mean rbar for the whole cell
+
+```{r relative.foci.Pos.plot, echo=FALSE}
+
+#mapp.cell.wide_cells
 #density or histogram plot facet hand foci count, plot value, fill type
-
 pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)
@@
 
-pos.plot.all <- ggplot(Position.DF, aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count,ncol = 1)+ggtitle("Simulated and normalized positions, male mus")
-
+pos.plot.all <- ggplot(Position.DF, aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count,ncol = 1)+ggtitle("Simulated and real normalized positions, male mus")
 pos.plot.all
 
-pos.plot.one <- ggplot(Position.DF[Position.DF$hand.foci.count == 2,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count)+ggtitle("Simulated and normalized positions")
-
-HS.density.plot <- ggplot(Hs.CO_cVs.DF, aes(x = COs, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Human male")+xlim(c(0,100))
-
+pos.plot.one <- ggplot(Position.DF[Position.DF$hand.foci.count == 2,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count)+ggtitle("Simulated and real normalized positions")
 ```
@@
 For multi co bivalents relative foci positions were kept together.
-For example, a random 2CO bivalent choosen and the normalize foci1 and foci2 for that bivalent were used.
-
-In terms of normalized foci positions, the empiracal distribution isn't very different from a randomized distribution.
-
+For example, a random 2CO bivalent chosen and the normalize foci1 and foci2 for that bivalent were used.
+In terms of normalized foci positions, the empirical distribution isn't very different from a randomized distribution.
 Not exactly sure if this is the right randomization procedure, the same 'random' bivalent could be drawn more than once.
 
-Next, map the randomized positions onto the real sc.lengths for cell cell observation.
+I didn't include the 2 types of simulations? 
+rank is not considered for these simluations because -- I assume
+
+I might write in a way to account for rank -- to follow the assumption that the relative foci positions vary for short and long chromosomes. but for now, just simulate across the whole/full pool.
 
@@
 #need sc length from the original
-Position.DF$raw.foci.pos <- Position.DF$SC.length *Position.DF$value
+Position.DF$raw.foci.pos <- Position.DF$SC.length * Position.DF$value
+
+male.whole.cell.mini$rand.Foci2 <- as.numeric(male.whole.cell.mini$rand.Foci2)
+male.whole.cell.mini$rand.Foci3 <- as.numeric(male.whole.cell.mini$rand.Foci3)
 
@@
 
-male.whole.cell.mini$rand.Foci2 <- as.numeric(male.whole.cell.mini$rand.Foci2 )
-male.whole.cell.mini$rand.Foci3 <- as.numeric(male.whole.cell.mini$rand.Foci3)
-
 #calq telo distnce
@@
 }
-
-mapp.cell.wide <- add_mouse(mapp.cell.wide)
-mapp.cell.wide <- add_strain(mapp.cell.wide)
 
@@
 
-maisel <- ggplot(mapp.cell.wide, aes(x=mean_rbar, y=mean_rand.rbar))+geom_point()+xlim(c(0,.3))+ylim(c(0,.3))
+mapp.cell.wide <- add_mouse(mapp.cell.wide)
+mapp.cell.wide <- add_strain(mapp.cell.wide)
+
+maisel <- ggplot(mapp.cell.wide, aes(x=mean_rbar, y=mean_rand.rbar))+geom_point()+xlim(c(0,.3))+ylim(c(0,.3))+ggtitle("cell wide randomized rbar vs empiracal rbar")
 maisel
@@
 oopp <- ggplot(mapp.cell.wide, aes(y=mean.raw, x=mean.raw.rand, color=mouse) ) +geom_point()+ggtitle("average foci position by cell, real vs randomized")
-oopp
 
 hh <- ggplot(mapp.cell.wide, aes(y=var.raw,x=var.raw.rand) ) +geom_point()+ggtitle("variance in foci position (per cell), between real and randomized")
-
-#what is the best way to plot these?
-
-#gg <- ggplot(Position.DF, aes(y=raw.foci.pos, x=variable, color= as.factor(hand.foci.count) ))+geom_jitter()+ggtitle("raw foci positions, reals and rand")
-#gg
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
@@
 the random r bar is higher than the real r bar positions. meaning on average the male cell rbar values are lower (less to the middle and more to the ends) -- would this be the same as in female cells?
+
+we calculated the intra-chromosomal portion of rbar (Veller et al., 2019) to compare how evenly crossovers were spaced while accounting for raw SC length.
+This metric ranges from 0 to 0.5, with higher values indicating that SC proportions flanking crossovers are more equivalent.
+
+My randomized plots indicate that the real chromosome measures have more 'unequal' chromsome portions flanking COs.
 
@@
 
+## Female CO Position Comparison
+
+For now stick with rbar, since it's a metric I can use across chromosome classes.
+
+1. randomized
+2. calq rbar values
+3. merge mini.DFs
+4. plot
+
+```{r Female.CO.pos, echo=FALSE}
+#female.whole.cell
+
+female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]
+female.biv.count_table <- female.biv.count_table[!(is.na(female.biv.count_table$fileName) | female.biv.count_table$fileName==""), ]
+female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$fileName) | female.whole.cell$fileName==""), ]
+
+#make sure male.whole.cell has nrm oci positions
+F_pos.pool_1CO <- female.whole.cell[female.whole.cell$hand.foci.count == 1,]
+F_pos.pool_2CO <- female.whole.cell[female.whole.cell$hand.foci.count == 2,]
+F_pos.pool_3CO <- female.whole.cell[female.whole.cell$hand.foci.count == 3,]
+
+F_pos.pool_1CO <- F_pos.pool_1CO[!(is.na(F_pos.pool_1CO$fileName) | F_pos.pool_1CO$fileName==""), ]
+F_pos.pool_2CO <- F_pos.pool_2CO[!(is.na(F_pos.pool_2CO$fileName) | F_pos.pool_2CO$fileName==""), ]
+F_pos.pool_3CO <- F_pos.pool_3CO[!(is.na(F_pos.pool_3CO$fileName) | F_pos.pool_3CO$fileName==""), ]
+
+#make a copy of main DF
+female.whole.cell.mini <- female.whole.cell
+
+female.whole.cell.mini$Foci1 <- female.whole.cell.mini$F1
+female.whole.cell.mini$Foci2 <- female.whole.cell.mini$F2
+female.whole.cell.mini$Foci3 <- female.whole.cell.mini$F3
+
+#filling in with randomized foci positions
+for(oar in 1:length(female.whole.cell.mini$fileName)){
+ #print(oar)
+
+  if(female.whole.cell.mini$hand.foci.count[oar] == 0){
+     # print("zero")
+      female.whole.cell.mini$rand.ID[oar] <- ""
+     female.whole.cell.mini$rand.Foci1[oar] <- ""  
+    female.whole.cell.mini$rand.Foci2[oar] <- ""
+     female.whole.cell.mini$rand.Foci3[oar] <- ""
+  }
+
+  if(female.whole.cell.mini$hand.foci.count[oar] == 1){
+     # print("oner")
+     rand.obj.id.1CO <- sample(F_pos.pool_1CO$Obj.ID, 1, replace = FALSE)
+     female.whole.cell.mini$rand.ID[oar] <- rand.obj.id.1CO
+     female.whole.cell.mini$rand.Foci1[oar] <- as.numeric(F_pos.pool_1CO$Foci1.PER[which(F_pos.pool_1CO$Obj.ID == rand.obj.id.1CO)])
+
+     female.whole.cell.mini$rand.Foci2[oar] <- ""
+     female.whole.cell.mini$rand.Foci3[oar] <- ""
+   }
+
+  if(female.whole.cell.mini$hand.foci.count[oar] == 2){
+  #  print("twoer")
+    rand.obj.id.2CO <- sample(F_pos.pool_2CO$Obj.ID, 1, replace = FALSE)
+    female.whole.cell.mini$rand.ID[oar] <- rand.obj.id.2CO
+    #asign the F1s and F2s from the fit obj.id
+
+    female.whole.cell.mini$rand.Foci1[oar] <-as.numeric(F_pos.pool_2CO$Foci1.PER[which(F_pos.pool_2CO$Obj.ID == rand.obj.id.2CO)])
+    female.whole.cell.mini$rand.Foci2[oar] <-as.numeric(F_pos.pool_2CO$Foci2.PER[which(F_pos.pool_2CO$Obj.ID == rand.obj.id.2CO)])
+    female.whole.cell.mini$rand.Foci3[oar] <- ""
+ }
+
+  if(female.whole.cell.mini$hand.foci.count[oar] == 3){
+    #print("threer")
+    rand.obj.id_3co <- sample(F_pos.pool_3CO$Obj.ID, 1, replace = FALSE)
+
+    female.whole.cell.mini$rand.ID[oar] <- rand.obj.id_3co
+        female.whole.cell.mini$rand.Foci1[oar] <-as.numeric(F_pos.pool_3CO$Foci1.PER[which(F_pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+    female.whole.cell.mini$rand.Foci2[oar] <-as.numeric(F_pos.pool_3CO$Foci2.PER[which(F_pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+    female.whole.cell.mini$rand.Foci3[oar] <-as.numeric(F_pos.pool_3CO$Foci3.PER[which(F_pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
+  }
+  }
+#I keep drawing the same chrm for many of the 3COs -- not sure how to fix this / get around this
+
+```
+
+```{r female.rbar, echo=FALSE, warning=FALSE}
+
+female.whole.cell.mini$SC.length <- as.numeric(as.character(female.whole.cell.mini$SC.length))
+female.whole.cell.mini$rand.Foci1 <- as.numeric(as.character(female.whole.cell.mini$rand.Foci1))
+female.whole.cell.mini$rand.Foci2 <- as.numeric(as.character(female.whole.cell.mini$rand.Foci2))
+female.whole.cell.mini$rand.Foci3 <- as.numeric(as.character(female.whole.cell.mini$rand.Foci3))
+
+female.whole.cell.mini$raw.rand_Foci1 <- female.whole.cell.mini$rand.Foci1 * female.whole.cell.mini$SC.length
+female.whole.cell.mini$raw.rand_Foci2 <- female.whole.cell.mini$rand.Foci2 * female.whole.cell.mini$SC.length
+female.whole.cell.mini$raw.rand_Foci3 <- female.whole.cell.mini$rand.Foci3 * female.whole.cell.mini$SC.length
+
+#female.whole.cell.mini
+for(y in 1:length(female.whole.cell.mini$fileName)  ){
+
+  female.whole.cell.mini$rbar[y] <- calq.intra.rbar(female.whole.cell.mini[y,])
+}
+
+for(rr in 1:length(female.whole.cell.mini$fileName)  ){
+  female.whole.cell.mini$rand.rbar[rr] <- calq.intra.rbar_rand(female.whole.cell.mini[rr,])
+}
+#don't know why there's so many 0 rbar values
+
+#make a new DF
+real_rbar.Female.Mus <- data.frame(female.whole.cell.mini$rbar )
+real_rbar.Female.Mus$type <- "real"
+colnames(real_rbar.Female.Mus) <- c("rbar", "type" )
+
+rand_rbar.Female.Mus <- data.frame(female.whole.cell.mini$rand.rbar )
+rand_rbar.Female.Mus$type <- "rand"
+colnames(rand_rbar.Female.Mus) <- c("rbar",  "type" )
+
+#merge
+Mus_rbar_female <- rbind(real_rbar.Female.Mus, rand_rbar.Female.Mus)
+
+#
+density.plot_rbars_female <- ggplot(Mus_rbar_female, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Bivalent Rbar distributions, Mus female")+xlim(c(0,.3))
+density.plot_rbars_female
+
+female.whole.cell.mini <- female.whole.cell.mini[female.whole.cell.mini$rbar > -.1,]
+#
+scatter_rbars_female <- ggplot(female.whole.cell.mini, aes(y = rbar, x= as.factor(fileName)) )+geom_jitter()+ ggtitle("Mus female rbar scatter plot")
+
+```
+
+Not sure why real distribution is more smooth compared to the random.
+The female rbar measures are closer to the maxium of 0.5, where all foci are placed equally along the chromsomes.
+
 ## Adding to Data set
 
-how to best increase the size of the dataset, with my bivalent data.
+how to best increase the size of the data set, with my bivalent data.
 From the curated BivData, figure out which can be added / into the whole cell data.
@@
 
-# Junk
+# Extra at the end
 
@@
 ```
+
+```{r per.co.count, echo=FALSE, eval=FALSE}
+#I don't think ... I end up useing these (matrics / DFs)
+iterations = 24
+variables = 20
+DF.permutations_CO.count <- matrix(ncol=variables, nrow=iterations)
+colnames(DF.permutations_CO.count) <- c("1", "rank1", "rank2", "rank3", "rank4", "rank5",
+                               "rank6","rank7","rank8","rank9","rank10","rank11",
+                               "rank12","rank13","rank14","rank15","rank16",
+                               "rank17","rank18","rank19" )
+
+for(row.num in 1:24){
+
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations_CO.count) ){
+
+  rand.rank.co <- sample(male.whole.cell$hand.foci.count[male.whole.cell$rank == col.num-1], 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations_CO.count[row.num,col.num] <- rand.rank.co
+   }
+  }
+DF.permutations_CO.count <- as.data.frame(DF.permutations_CO.count)
+
+#no rank - iterate / permute the sampling X times
+its = 24
+varis = 20
+DF.permutations.CO.no_rank <- matrix(ncol=varis, nrow=its)
+
+#make LIST of X DATAFRAMES
+
+for(row.num in 1:24){
+  #print(c(row.num, "rowis") )
+  for(col.num in 2:ncol(DF.permutations.CO.no_rank) ){
+  rand.co <- sample(male.whole.cell$hand.foci.count, 1, replace = FALSE)
+  #print(SCll)   
+  DF.permutations.CO.no_rank[row.num,col.num] <- rand.co
+   }
+  }#yes this finally works!!!!
+DF.permutations.CO.no_rank <- as.data.frame(DF.permutations.CO.no_rank)
+
+#
+for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
+  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
+  }
+```

### 2019-12-09 src/CommonFunc_MLH1repo.R

@@ -1,0 +1,608 @@
+# Common MLH1repo functions
+#
+#
+
+#make functions for cleaning up BivData
+#-remove p_rev
+#-check for multi-cell images
+
+
+
+#make fileName
+#paste together file name
+add_fullFileName <- function(dat){
+  count =1
+  dat$fileName <- as.character(dat$fileName)
+  dat$stain <- as.character(dat$stain)
+  
+  for(i in dat$fileName){  #i is file name!!!
+    #print(i)
+    templist1= strsplit(i, split="_")[[1]] #put the 
+    
+    j=dat$stain[count]
+    templist2= strsplit(j, split="_")[[1]]
+    c = paste(templist1[2], templist1[3],templist1[4], sep = "_")#mdate, strain, m#
+    #test.old.master$mouse2[count] <- c #don't need to add or change mouse
+    
+    d = paste(templist1[1], templist1[2], templist1[3],templist1[4],templist2[3], dat$cell[count], 'rev.tif', sep = "_") 
+    dat$file.name_full[count] <- d #should match file name
+    count= count + 1
+  }
+  return(dat)
+}
+
+#format the dates into standard format
+#mouse_list$DOB <- as.Date(mouse_list$DOB,format='%m/%d/%Y')
+add_age <- function(dat){
+  dframe <- dat
+  count =1
+  for(i in dframe$mouse){
+    #after DOB and euth date are standardized #substract
+    dframe$age.days[count] <- dframe$euth.date[count] - dframe$DOB[count]
+    dframe$age.weeks[count] <- difftime(dframe$euth.date[count], dframe$DOB[count], units='weeks')
+    #gives age in days
+    count= count +1
+  }
+  return(dframe)
+}
+#running this blanks out all cells in euth.date col
+
+#calq maternal age time of mouse birth
+add_mat_age <- function(dat){
+  dframe <- dat
+  count =1
+  for(i in dframe$mouse){
+    #after DOB and euth date are standardized #substract  difftime(mat.age.dob - DOB)
+    dframe$mat.age.days[count]<-difftime(dframe$DOB[count], dframe$maternal_DOB[count],units='days')
+    
+    #dframe$age.weeks[count] <- difftime(dframe$euth.date[count], dframe$DOB[count], units='weeks')
+    #gives age in days
+    count= count +1
+  }
+  return(dframe)
+}
+
+
+#add mouse column to MLH1 df
[…]

### 2019-12-10 results/Chapter3_ideas99.Rmd
edits, rbar comparisons look more like expected

@@
 
+For mus, data frames that sample based on SC.length / rank, can be used for SC length and CO count metrics. Not CO position
+
+both SC and CO count randomized data
+
 These code chunks make a randomized data set of 18 nuclei with 20 random chromosome draws each.
@@
 
+I think new randomized DFs must be made (male.whole.cell.min, not the DF.list)
+(previous ones made lists of DFs, which were made of simulated DFs for each nucleus/cell)
+My current code for position data, takes the current real DF, adds new columns with random draws of relative CO positions.
+
 I am still working on the metrics to compare (and the ones that are best for cell wide descriptions).
@@
 
+-measure the CO-variation of metrics within cells!
+for each cell, make even-rank and odd-rank columns of rbar or another metric. Measure covariance of the 2 sets of pairs (real and random, even-odd).
+
 ## Male CO Position
@@
 
-```{r CO.pos, echo=FALSE}
+```{r random.CO.pos, echo=FALSE}
 #male.whole.cell
@@
 
-```
-
-### merge data
-
-```{r merging.data,echo=FALSE}
+male.whole.cell.mini$SC.length <- as.numeric(as.character(male.whole.cell.mini$SC.length))
+male.whole.cell.mini$rand.Foci2 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci2))
+male.whole.cell.mini$rand.Foci3 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci3))
+
+male.whole.cell.mini$raw.rand_Foci1 <- male.whole.cell.mini$rand.Foci1 * male.whole.cell.mini$SC.length
+male.whole.cell.mini$raw.rand_Foci2 <- male.whole.cell.mini$rand.Foci2 * male.whole.cell.mini$SC.length
+male.whole.cell.mini$raw.rand_Foci3 <- male.whole.cell.mini$rand.Foci3 * male.whole.cell.mini$SC.length
+
+```
+
+### calq within nuc COVARIATION
+
+```{r even.odd, echo=FALSE}
+#4 new columns for testing with in nuc covariance
+
+#this doesn't works as well for male 19! (get ride of shortest?)
+#make new DF (copy, or empty?)
+
+covary.DF <- male.whole.cell.mini[male.whole.cell.mini$fileName == "1mar16_3jan16_G_m2_sp1_19.3_rev",]
+#loop
+
+#I HAVE to make sure they are all ordered by metric, or the pairing for cov() will not make sense
+
+covary.DF <- with(covary.DF, covary.DF[order(SC.length),])
+evens <- covary.DF[which(covary.DF$rank %% 2 == 0),]
+odds <- covary.DF[which(covary.DF$rank %% 2 != 0),]
+cov(evens$SC.length,odds$SC.length[2:10])#303 (lots of variance)
+cor(evens$SC.length,odds$SC.length[2:10])# 0.2480786
+#do this for each cell / filename
+
+#covary, rbar -- within the same nuc
+#real and rand cov
+for(y in 1:length(covary.DF$fileName)  ){
+  covary.DF$rbar[y] <- calq.intra.rbar(covary.DF[y,])
+}
+
+covary.DF <- with(covary.DF, covary.DF[order(rbar),])
+evens <- covary.DF[which(covary.DF$rank %% 2 == 0),]
+odds <- covary.DF[which(covary.DF$rank %% 2 != 0),]
+
+cov(evens$rbar,odds$rbar[2:10])#0.007357481
+cor(evens$rbar,odds$rbar[2:10])# 0.9734986
+#I don't know how to interpert the scales of the covariaition values...
+#correlation is simpler to me
+
+#male.whole.cell.mini
+#counter of file
+
+for(y in 1:length(male.whole.cell.mini$fileName)  ){
+  male.whole.cell.mini$rbar[y] <- calq.intra.rbar(male.whole.cell.mini[y,])
+}
+for(rr in 1:length(male.whole.cell.mini$fileName)  ){
+  male.whole.cell.mini$rand.rbar[rr] <- calq.intra.rbar_rand(male.whole.cell.mini[rr,])
+}
+
+#make smaller cell.level DF
+smllr.list = data.frame(fileName = unique(male.whole.cell.mini$fileName))
+i = 1
+#the loop isn't cycling
+for(each in unique(male.whole.cell.mini$fileName) ) {
+ # print(each)
+
+  #make smaller DF, list of all obs of that file
+  temp.Df = male.whole.cell.mini[ male.whole.cell.mini$fileName == each,]
+  ordered.DF.rbar = with(temp.Df, temp.Df[order(rbar),])
+
+  #order by the metric (SC,rbar)  #make list of even and odd
+  evens <- ordered.DF.rbar$rbar[which(ordered.DF.rbar$rank %% 2 == 0)]
+  odds <- ordered.DF.rbar$rbar[which(ordered.DF.rbar$rank %% 2 != 0)]
+
+  evens.rand <- ordered.DF.rbar$rand.rbar[which(ordered.DF.rbar$rank %% 2 == 0)]
+  odds.rand <- ordered.DF.rbar$rand.rbar[which(ordered.DF.rbar$rank %% 2 != 0)]
+
+  #calq cor,cov
+  smllr.list$fileName[i] <- each
+  smllr.list$cov[i] <- cov(evens,odds[2:10])
+  smllr.list$cor[i] <- cor(evens,odds[2:10])
+
+  smllr.list$cov.rand[i] <- cov(evens.rand, odds.rand[2:10])
+  smllr.list$cor.rand[i] <- cor(evens.rand, odds.rand[2:10])
+
+  #print(i)
+  i = i+1
+  }
+
+#plotty <- ggplot(smllr.list, #aes(x=cov,y=cov.rand))+geom_jitter()+xlim(c(-.005,.006))+ylim(c(-.005,.006))
+
+```
+
+Calq the within cell covariance for the true and randomized values. 
+Not sure how to use them yet
+
+### merge melt data
+
+don't use melt data
+
+```{r melt.data,echo=FALSE}
 
@@
 
-male.whole.cell.mini$SC.length <- as.numeric(as.character(male.whole.cell.mini$SC.length))
-male.whole.cell.mini$rand.Foci2 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci2))
-male.whole.cell.mini$rand.Foci3 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci3))
-
-male.whole.cell.mini$raw.rand_Foci1 <- male.whole.cell.mini$rand.Foci1 * male.whole.cell.mini$SC.length
-male.whole.cell.mini$raw.rand_Foci2 <- male.whole.cell.mini$rand.Foci2 * male.whole.cell.mini$SC.length
-male.whole.cell.mini$raw.rand_Foci3 <- male.whole.cell.mini$rand.Foci3 * male.whole.cell.mini$SC.length
-
 ```
@@
 
+dense1 <- ggplot(real_rbar.Mus, aes(x = rbar, fill=type) )+geom_density(alpha=.6)
+
 #
-density.plot_rbars <- ggplot(Mus_rbar_male, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Bivalent Rbar distributions, Mus male")+xlim(c(0,.3))
+density.plot_rbars <- ggplot(Mus_rbar_male, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Bivalent Rbar distributions, Mus male")+xlim(c(0,.55))
 density.plot_rbars
@@
 #
-density.plot_cell.wide.rbars <- ggplot(cell.wide.tot.rbar_Musmale, aes(x = sum.rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Cell wide total Rbar distributions, Mus male")+xlim(c(0,4))
+density.plot_cell.wide.rbars <- ggplot(cell.wide.tot.rbar_Musmale, aes(x = sum.rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Cell wide total Rbar distributions, Mus male")+xlim(c(0,6))
 
@@
 
-I didn't include the 2 types of simulations? 
-rank is not considered for these simluations because -- I assume
-
-I might write in a way to account for rank -- to follow the assumption that the relative foci positions vary for short and long chromosomes. but for now, just simulate across the whole/full pool.
-
-### Map normalized positions onto real SC
-
-```{r map.rand2real, echo=FALSE}
+I might write in a way to account for rank, to account for how the relative foci positions vary for short and long chromosomes. but for now, just simulate across the whole/full pool.
+
+```{r map.rand2real, echo=FALSE, warning=FALSE}
 

### 2019-12-17 doc/figureFiles/figures.Rmd
working rmd files for drafts of main figures

---
title: "Chapter 2 Figures"
author: "April Peterson"
date: "December 16, 2019"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, include=FALSE}
#load data ect
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(dplyr)
library(lattice)
library(dplyr)
library("cowplot")
library(raster)#for cV
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load main data file
load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

```{r strain.color, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= 'indianred4',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )

```

```{r setting.data, echo=FALSE, warning=FALSE}
#remove F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "F1"),]

#remove outgroups
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPRET"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "SPIC"),]
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$strain != "CAROLI"),]

#quality control - remove outlier mice
AP_mouse_table.HQ <- AP_mouse_table.HQ[(AP_mouse_table.HQ$mouse != "18nov17_WSB_f5"),] #this mouse only had 1 cell

#I need cell level data

uni.mice <- unique(AP_mouse_table.HQ$mouse)
fig.data<- MLH1_data[ MLH1_data$mouse %in% uni.mice, ]

#take all mouse levels from AP.table,
#subset MLH1 data with those mice
```

```{r fig1_scatter_basic, echo=FALSE, warning=FALSE}
#mouse averages
#AP_mouse_table.HQ
#fig.data  cell data

fig.data <- fig.data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match

# sort your dataframe, by the focal categories
fig.data$mouse <- factor(fig.data$mouse, levels=unique(fig.data$mouse))

fig.data.male <- subset(fig.data, sex== "male")

#male
Fig1_A_mouse.male <- ggplot(fig.data.male[order(fig.data.male$strain), ],
  aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+  #geom_jitter()+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+geom_boxplot(alpha = 0.6)+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),

  #lines -- these might be tricky with the faceting
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())  +
  #Dom
  annotate("segment", x= 10, xend=20.5, y=15, yend=15, colour="black", size=1)+
  #Musc
  annotate("segment", x= 50, xend=55, y=15, yend=15, colour="black", size=1)+
  #Cast
  annotate("segment", x=70, xend=80, y=15, yend=15, colour="black", size=1)

#female
#the data is mouse mean 
fig.data.female <- subset(fig.data, sex== "female")

Fig1_A_mouse.female <- ggplot(fig.data.female[order(fig.data.female$strain), ], aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+ #geom_jitter(width = 0.25)+
  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+
 theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #lines -- these might be tricky with the faceting
 axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  #Dom
  annotate("segment", x= 10, xend=20.5, y=18, yend=18, colour="black", size=1)+
  #Musc
  annotate("segment", x= 50, xend=55, y=18, yend=18, colour="black", size=1)+
  #Cast
  annotate("segment", x=70, xend=80, y=18, yend=18, colour="black", size=1)

```

# Figure 1

```{r strain.table, echo=FALSE, warning=FALSE}
#fig.data -- (only mice from better list)

#there are still quality 5 cells in this dataframe
fig.data <- subset(fig.data, quality < 5)

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Nmice), 3)
    )

#cv, sd and se are all NA / empty
strain.table.HQ <- strain.table.HQ[!(is.na(strain.table.HQ$cV) | strain.table.HQ$cV== "" ), ]

#remove outgroups
#remove F1
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "F1"),]

#remove outgroups
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPRET"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "SPIC"),]
strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAROLI"),]

```

```{r reorder.strain, echo=FALSE}

#re-order strains
#Musc
#PWD, SKIVE, KAZ, CZECH, TOM, AST

#MOl
#MSM, MOLF
strain.table.HQ$strain<- factor(strain.table.HQ$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                                                                   "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                                  "MSM", "MOLF",
                                                                    "CAST", "HMI",
                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
```

I think some strain observations can be removed, since they only have

```{r strain.table.plot, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_B_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male

  #segments
  annotate("segment", x= 0, xend=5, y=21, yend=21, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5, xend=8, y=20, yend=20, colour="black", size=1)+
  #Cast
  annotate("segment", x=11, xend=15, y=18, yend=18, colour="black", size=1)+
  #label
  annotate("text", x = c(1.5,8,10.5, 11), #list of all x points 
           y = c(20,18, 20,18),

           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus", "M.m.castaneus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

# 2
Fig1_B_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=1, xend=3.5, y=22, yend=22, colour="black", size=1)+
  #Musc
  annotate("segment", x=5, xend=8.5, y=20, yend=20, colour="black", size=1)+
  #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(1.5,8), #list of all x points 
           y = c(20,18),
           label = c("M. m. domesticus","M. m. musculus"),
           fontface = 'italic', hjust = 0.5, size=3)

```

```{r Strain.fig.merge, echo=FALSE}
fig1_B  <- ggdraw() +
  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.48, height=.88) +
  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.48, height=.88) +

  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )
fig1_B
```

add molossinus

re-arrange Muusc strain level order to highlight the polymorphism to musc

Clean up segments
There are many more male strains represented.

The male plot is squished because of the legend

## Figure 1, mouse levels shown

```{r cow.plot.fig1, echo=FALSE}

fig1  <- ggdraw() +
  draw_plot(Fig1_A_mouse.female, x=0, y = 0, width=.48, height=.88) +
  draw_plot(Fig1_A_mouse.male, x=.5, y=0, width=.48, height=.88) +
  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )

fig1
```

USE boxplots --- this is a bit too busy with single mouse level observations. (so forget the cell level data I just coded. What if I sub-sampled?)

Draft first fig. MLH1 distributions will be the main first figure
I think a strain level would look better.  Add white space between

I still need to work on some colors
add line segments, they are good at anchoring

add line segments for strain means?, do boxplots?

Note that this figure knitted in html will look different than a saved png or tif.
change the ylabel 'Mean Number of MLH1 Foci per Cell'

### 2019-12-17 results/Chapter3_ideas99.Rmd
working on finalizing Chap3 Outline

@@
 
-Data used in the paper
-human data- Gruhn et al 2013 (plotted below, 755 cells)
-Sun et al 2006
-Oliver-Bonet et al 2007 
-Lian et al 2008
-Lu et al 2012
-Hou et al 2013
+human Data used in the paper
+
+Gruhn et al 2013 - T. Hassold,
+4660 prophase sperm from 56 biopsies, 2000 prophase oocytes
+
+Sun et al 2006 --  F. Sun, 824 MLH1 FISH spreads(~100 cells / 10 men)
+
+Oliver-Bonet et al 2007 -- Oliver-
+Bonet  (RPA foci for 5 different SCs ~50 cells, ~400 cells with FISH + MLH1). problem if these data are included in big dataset
+
+Lian et al 2008 - S. Xie, (maybe similar data as sun et al 2008?)
+Lian, Jie, et al. "Variation in crossover interference levels on individual chromosomes from human males." Human molecular genetics 17.17 (2008): 2583-2594.
+
+Lu et al 2012 - 
+Hou et al 2013 - no SC lengths, (CO sounts 68 cells (oocytes?))
+
+Human data are from mulitple citations; Gruhn et al 2013, Sun et al 2006, Oliver-Bonet et al 2007, Lian et al 2008, Lu et al 2012, Hou et al 2013.
+
+Cytological Studies of Human Meiosis: Sex-Specific
+Differences in Recombination Originate at, or Prior to,
+Establishment of Double-Strand Breaks
 
@@
 
+Melted.Hs.male_axis$Chrm.id <- factor(Melted.Hs.male_axis$Chrm.id, ordered = TRUE, levels=c("chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22") )
+
 #Seems like there is an outlier, 
@@
   ggtitle("Human male CO Counts")+theme(legend.position="none")
-hs.CO.length_plot
+#hs.CO.length_plot
 ```
@@
 
-Heads of the cell-wide summary statistics
+Heads of the cell-wide summary statistics.
 
@@
                             mean.SC = mean(axis.length),
-                            cV.SC = cv(axis.length)
+                            cV.SC = cv(axis.length)#this does within cell cV (dif than cov)
 )
@@
 
-```
+#the mus is rank based on SC, can I write in rank based on Chrm.id
+for(cellys in as.character(unique(Melted.Hs.male_axis$row.name))){
+  chrm.order = rank(Melted.Hs.male_axis$axis.length[whole.cell.data$row.name == cellys])
+  Melted.Hs.male_axis$rank[Melted.Hs.male_axis$row.name == cellys] <- chrm.order  
+}
+
+#
+two_groups2 <- transform(Melted.Hs.male_axis, category2 = factor(paste(row.name, Chrm.id)))
+two_groups2 <- transform(two_groups2, category2 = reorder(category2, rank(desc(axis.length)), decreasing = FALSE))
+#use desc for DECENDING order
+MiniBiv_line.plot <-  ggplot(two_groups2, aes(category2, axis.length, group=row.name, color=row.name)) +  geom_line(alpha = 0.2) + 
+  ggtitle("700 human male chrms ranked by Length")+ labs(x = "")+theme(axis.line=element_blank(),axis.text.x=element_blank(),
+        axis.ticks=element_blank())
+
+MiniBiv_line.plot
+
+#order by chrm id (they are now an ordered factor)
+#it might be kinda hard to subset the melted data, because they aren't organized into cells
+#
+
+Melted.Hs.male_axis.mini <- subset(Melted.Hs.male_axis, row.name < 101)
+
+two_groups <- transform(Melted.Hs.male_axis.mini, category2 = factor(paste(row.name, Chrm.id)))
+two_groups <- transform(two_groups, category2 = reorder(category2, rank(Chrm.id)))
+
+MiniBiv_line.plot2 <-  ggplot(data=two_groups, aes(category2, axis.length, group=row.name, color=row.name)) +  geom_line(alpha = 0.5) +  ggtitle("700 human male chrms ranked by chrm ID")+ labs(x = "")+theme(axis.line=element_blank(),axis.text.x=element_blank(),
+        axis.ticks=element_blank())
+
+MiniBiv_line.plot2
+
+#I could check a couple hundred, and remove cells with noticible spikes
+
+```
+
+I added in single cell ranked axis curves.
+That is 22 axis/chromsomes for a human spermatocyte are ranked smallest to largest and I draw a curve over them.
+There are 700 cells, which obscure some patterns.
+A simple discription is that wider/thicker color indicates more variation in the final rank.
+
+I'm trying to think more about the single cell looping factor. -- how would I discribe this, slope of ranked 
+Sometimes the chrm id doesn't fit rank expectations.  (looking at the mini.biv plots, )
 
@@
 
-I didn't test the axis length trait yet.
+I didn't test the axis length trait yet.  (See below for axis traits)
 
@@
 
+```{r Mus.physcial.genome, echo=FALSE}
+#distribution of genome sizes from reference Mus genome
+#haploid genome size
+refMus.geome_Mb_male <- c(195,182,160,152,150,145,129,125,131,122,120,120,125,104,91,95,91,61)# 1109.412
+
+refMus.geome_Mb_XX <- c(195,182,160,152,150,145,129,125,131,122,120,120,125,104,91,95,91,61, 171)
+
+#note, this ommits telomeres and centromeres!
+#read the 
+#on average 8Mb worth of centromeres and telomeres are added to each chrm
+#i will try to find estimates of variation across indiviual chromosomes
+
+```
+
+Show a plot of the estimates chrm sizes.
+These values from MGI page omit centromere and telomere estimates.
+Read up on the T2T project lit to see if I can find better estimates of physical chormosome sizes.
+
+Do I need most accurate estimate of Mb size variance?
+would I expect -- similar variance among cells?
+If i ran jack-knifing or sub-sampling permutations of the within cell covariance calculations -- if the covariation eatimate remained about the same, I would be more convinced that the 'packing' ratio is mostly uniform WITHIN cells.
+Null/ comparison value might be, randomize cell.id for all bivalent observations.
+(maybe I should first try this in human data? (minus the wacky looking cells / bad chrm id))
+
+This might be the best thing to do to account for different physical sizes across other strains (non reference genome)
+
 I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations by categories.
@@
 )
+#exclude last one
+male.biv.count_table <- male.biv.count_table[!(is.na(male.biv.count_table$fileName) | male.biv.count_table$fileName==""), ]
+
+male.biv.count_table$ratio.now <- var(refMus.geome_Mb_male) / male.biv.count_table$var_sc
+male.biv.count_table$ratio.invert <-  male.biv.count_table$var_sc / var(refMus.geome_Mb_male)
+#have to think of which is right
 
@@
 )
-
-```
+#exclude last nas
+female.biv.count_table <- female.biv.count_table[!(is.na(female.biv.count_table$fileName) | female.biv.count_table$fileName==""), ]
+
+female.biv.count_table$ratio.now <- var(refMus.geome_Mb_XX) / female.biv.count_table$var_sc
+female.biv.count_table$ratio.invert <-  female.biv.count_table$var_sc / var(refMus.geome_Mb_XX)
+
+```
+
+Calculated the ratio of genome size variance (haploid chrm variance), (I think the right denominator is SC length variance)
+most all are above 1 (for both female and male) -- indicates the packing ratio / packing process doesn't maintain (our guess about genome size of chromosomes)
 

### 2019-12-17 doc/Main_Report.rmd
missed some updating

@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
+
+load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
+load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
+#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
+#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
 #load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
@@
 Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
-
-## Discription of quantification process
 
@@
 
-```{r organizing.data, echo=FALSE}
+```{r organizing.data, echo=FALSE }
 #set the order of another column, based on another variable (so that when)
@@
 
-This section shows the number of mice which haven't been stained or quantified yet.
-
-```{r missing.mice, echo=FALSE, warning=FALSE}
+```{r missing.mice, echo=FALSE, warning=FALSE, eval=FALSE}
+#This section shows the number of mice which haven't been stained or quantified yet.
 
@@
 
-passed quantified - quantified mice (with enough cells)
-total quantified -- all mice which have been (includes fail and passing mice)
-
-non-quantified mice -- dissected mice / slides that haven't been quantified yet (all stained, but without image directory)
-
-non-imaged -- mice without images, (might be stained and unstained)
-
-This is the number of mice that are through different parts of the pathway or quantification process.
-
-should I break down the categories into sex and strain?
-
 ```{r basic.numbers, echo=FALSE, eval=FALSE}
@@
 
-These are the tables for the number of cells and mice which have been quantified.
-The 'Other' are F1 crosses.
-
-## Repeatability of Human quantification
-
-Well, we have to talk about the subjective nature that goes into scoring the MLH1 foci.
+```{r rand, echo=FALSE, include=FALSE}
+
+str(AP_mouse_table_w.Ages)
+
+AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+
+#remove sex.x == na, and strain.x=na or F1
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
+#plot variance and mean (facet sex)
+
+#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
+
+VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
+
+cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
+
+str(AP_mouse_table.HQ)
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
+#[1] 0.5750399
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
+#0.321601
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
+# 0.3515544
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
+#-0.03713536
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
+# p-value = 8.623e-09
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
+
+#remove mice with fewer than 10 cells
+
+AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
+
+VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
+
+cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
+
+cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
+    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
+
+```
+
+## Repeatability of Human Quantification
+
+Let's talk about the subjective nature of scoring the MLH1 foci.
 Sources of error or variance across scoring come from background noise (false positives), faint foci (false negatives) and just human error.
-A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (using some of Beth's data).
-Todo: track down, re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)
+A general anecdote is that if the quality of the cell staining is high, then the repeatability is near 100%.
+When the quality is less than perfect, some foci can ambigous and cause the repeatability to decrease.
 
@@
 
+-re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)
 Datasets / sources of repeat measures
--chrom proportions
+-cell counts from chrm proportions vs batch
 -multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
--old and new (batch0)
-
-The main take away is that higher quality scored cells have lower error in MLH1 counts.
+-old and new (old counts and anonimized batch0)
+
+Some points I still need to address
 
@@
 
-```{r quant.repeatability_rep1, echo=FALSE, include=TRUE}
+```{r quant.repeatability_rep1, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
 #set up data for combining replicate counts
-
 #1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now
@@
 
-levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt
+#levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt
 
 #check that the 1, and 2 are equal number
-length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178
+#length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178
 
@@
 dups.here.plain_cor_qual <- cor(dups.here$qual.1, dups.here$qual.2)
-
-#i don't know how to compare quality in more
 
@@
 
-```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
+```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE, message=FALSE}
 #2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
@@
 
-str(test.old.master)
+#str(test.old.master)
 
@@
 
-```{r repeatability_rep3, echo=FALSE, include=FALSE}
+```{r repeatability_rep3, echo=FALSE, include=FALSE, message=FALSE}
 chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
@@
 
-This data has `r length(chrm.class.DF$Original.Name)` cells across these categories
+This data has `r length(chrm.class.DF$Original.Name)` cells from mice across these categories
 
@@
 
-More cells are planned to be quantified.
-
-```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE}
+```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE, message=FALSE}
 hist_error_plot
@@
 
-From this histogram we see most of the cells in this dataset have a difference of 2 in the MLH1 count scores (`r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`).
-
-```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5}
+From this histogram we see most of the cells in this dataset have a difference of less than 2, or `r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`%.
+
+```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5, message=FALSE}
 quality_error_plot
@@
 poorer quality cells are more likely to have more error acorss multiple counts).
-The proportion of the cells which had the same count across the counting instances is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`.
+The proportion of the cells which had the same count is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`%.
 The correlation between the two versions of MLH1 counts for the same cell is `r cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)`.
@@
 ```
-There doesn't seem to be a bias across cells with more or less foci.
-
-Modeling this relationship; (still in progress)
-I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.
-
-## Effects on the dataset
-
-error associated with 'quality of cels'
-
-Main metric is mean CO count per mouse. Run tests to check if there are effects for the following effects.
-
-*age effects* (male, maternal age)
-
-*litter effect*
-
-*mouse room effect* (Charmony, MSC)
-
-*quantification repeatability*
-
-*spread batch and staining effects*
-
-*power analysis on number of cells*
-
-*time of year effect (on spread)*
-
-*dominant effect* (F1s)
-
-These aspects will be analyzed with a linear models on subsets of data (within category) and visualized with plots.
+There doesn't seem to be a bias across cells with more foci.
+
+```{r model.repeat, include=FALSE, eval=FALSE}
+#add model -- logisitc regression?
+#I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.
+```
+
+```{r Effects.on.the.dataset, eval=FALSE, echo=FALSE}
+
+#*age effects* (male, maternal age)
+
+#*litter effect*
+
+#*mouse room effect* (Charmony, MSC)
+
+#*quantification repeatability*
+
+#*spread batch and staining effects*
+
+#*power analysis on number of cells*
+
+#*time of year effect (on spread)*
+
+#*dominant effect* (F1s)
+
+```
 
@@
 
-There is a positive correlation with MLH1 number and quality
-To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.
+There is a positive correlation with MLH1 number and quality.
+To quantify this effect, compare all of the counts from quality-1 cells across strains.
+check that the same patterns of means persist compared to the total-cell data.
 
@@
 
-Q12toTotal
+#Q12toTotal
 
@@
 
-kable(Q12_MLH1_means_males, caption = "High quality cells")
-
-kable(tot.DF_MLH1_means_males, caption = "Total Cell")
+#kable(Q12_MLH1_means_males, caption = "High quality cells")
+
+#kable(tot.DF_MLH1_means_males, caption = "Total Cell")
 
@@
 I think the ranges across the groups is generally consistant.
-
-(try looking at <5 quality scores?)
 
@@
 
-### Age Effect
+### Age Effect (male)
 
@@
 
-Make a subsetted dataset for plotting mean mouse (quality < 5) MLH1 count and age.
-
 ```{r age.effect.model, include=FALSE}
@@
 ```
-
-I will try doing a logistic regression for embryonic vs neonate. The embyonic estimates likely aren't very accurate, and there won't be good sampling across time points.
 
@@
 
+### Age Effect (female)
+
+Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.
+
+```{r female.age, eval=FALSE}
+#compile this dataset
+```
+
 ### Maternal Age effect
@@
 #check if date format's been applied before tryung the age functions
-str(meta.data.file.w.mat.DOB)
+#str(meta.data.file.w.mat.DOB)
 
@@
 
-str(meta.data.file.w.mat.DOB)
+#str(meta.data.file.w.mat.DOB)
 
@@
 
-#merge meta data with MLH1 
+#merge meta data with MLH1 -- rename this
 see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
@@
 #visualize the data
-
 meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
@@
 
-#age by mean_co
+#age by mean_co (these aren't working)
 mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
@@
 
-There seems to be a negative correlation with
-
 ### Litter effect
@@
 Compare female spreads made on the same day, but from different mothers/litters.
+```{r litter.effect, eval=FALSE}
+#compile this data set
+
+```
 
@@
 
-### Room effect (envioronment)
-
-3 different rooms used to house mice. Access softmouse to be fixed. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
+```{r stain.effects, eval=FALSE}
+
+#compile this dataset
+
+```
+
+### Room effect
+
+```{r ROOM.effect, eval=FALSE}
+
+#COMPILE THIS datasets
+#3 different rooms used to house mice. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
+
+```
 
@@
 
-# Initial Results (change these colors)
+### Distributions of MLH1 counts
+
+> *Are the MLH1 distributions close to normal?*
+
+```{r MLH1.histogram.2421, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}
+
+#gg <- data.frame(weight=rep(x,nrow(df)),df)
+#gg$y <- with(gg,dnorm(x,mean,sd))
+#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
+
+mean.MLH1.obs <- ddply(.data=MLH1_data, 
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")+
+  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
+
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+
+histM
+
+#function = glm(y ~ x, family = poisson)
+#add Poission curve behind
+#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
+#not sure how to add Expected poisson curve
+
+```
+
+```{r passing.mice.dont.use, echo=FALSE, warning=FALSE}
+q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
+                        total =  length(adj_nMLH1.foci),
+                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
+                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
+                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
+)
+#q_cutoff_table
+passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
+
+not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
+
+#mnake dataframe with just passing mice -- so that it can be ploted
+passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
+```
+
+Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
+Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
+(CAST female data is fake)
+
+I plotted the mean of each quality bin with a red dot.
+From the pattern of the red dots, there is definitely a negative relationship with quality and nMLH1 foci across the data.
+This is most pronounced in MSM males and least pronounced in G males.
+The CAST female data is dummy data.
+
+# Initial Results
 
@@
 #don't show M.musculus by mouse
+#scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
+ #       "green","green4","grey",  "grey","grey","grey" )) +
 
@@
 #colors_of_strains, colors .. 
-
-ff2 <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
-  geom_boxplot(aes(x = as.character(mouse), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+#there is no order to 'mouse' so it looks shitty
+
+justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
+  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
-        "green","green4","grey",  "grey","grey","grey" )) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
-  facet_grid(subsp~ sex, scales="free_x")+scale_color_manual(values=colors_of_strains)
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+
+
+justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")
 
@@
 ```
+
+```{r variance.boxplots, echo=FALSE}
+#plot of variance
+
+var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
+
+#plot by sex / seperately
+# HQ data DF.HetC.MixedModel.HQ
+
+var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+var.male <- var.male   + facet_wrap( ~ strain, scales="free")
+
+cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
+
+```
+
+There are 2 MSM males with high (20) variance and cV, G seems to have more variance across mouse level variance too.
+
+Check out these mice in comparison to other mice
+30sep16_MSM_m1 -- staining might not be the best, MLH1 signal is spottiny in some cells
+
+30may17_MSM_m1 - quite a bit of green noise across these cells.
+
+Gough and KAZ males has a bit more variation across mice for mouse level variation.
 
@@
 
-### Distributions of MLH1 counts
-
-> *Are the MLH1 distributions close to normal?*
-
-```{r MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}
-
-#gg <- data.frame(weight=rep(x,nrow(df)),df)
-#gg$y <- with(gg,dnorm(x,mean,sd))
-#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
-
-mean.MLH1.obs <- ddply(.data=MLH1_data, 
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")+
-  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
-
-#  geom_smooth(method = lm, se = FALSE
- # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
-
-histM
-
-#function = glm(y ~ x, family = poisson)
-#add Poission curve behind
-#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
-#not sure how to add Expected poisson curve
-
-```
-
-(most all distributions are close to normal)
-
-How do these compare to Poisson? (mean == variance)
-I should make 'expected' distributions behind these.
-
-```{r passing.mice, echo=FALSE, warning=FALSE}
-q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
-                        total =  length(adj_nMLH1.foci),
-                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
-                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
-                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
-)
-#q_cutoff_table
-passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
-
-not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
-
-#mnake dataframe with just passing mice -- so that it can be ploted
-passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
-```
-
-Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
-Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
-(CAST female data is fake)
-
-```{r scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means and quality scores", warning=FALSE}
-
-#make lattice / plots seperated by strain and sex
-#facet
-
-#quality scores are now 1:7?
-qual_means <- MLH1_data %>% 
-        group_by(category, quality) %>% 
-        summarise(
-          adj_nMLH1.foci = mean(adj_nMLH1.foci)
-        )
-
-p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
-p <- p + geom_point(stat = "identity")
-p <- p  + geom_point(data = qual_means, colour="red") 
-p <- p + facet_wrap(~ category)
-p
-
-```
-
-I plotted the mean of each quality bin with a red dot.
-From the pattern of the red dots, there is definitely a negative relationship with quality and nMLH1 foci across the data.
-This is most pronounced in MSM males and least pronounced in G males.
-The CAST female data is dummy data.
-
-####Distributions by mouse
+#### Distributions by mouse
 
@@
 
-### Explaination for the XX and XY differences
-
-(in progress)
-
 ### Difference in proportion of chrm classes
@@
 generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
-10 cells from each musc category.
+10 cells from each musc category. Add Molf male.
 
@@
 
-### Mixed Model, compare within and between subspecies patterns
+# Building Mixed Models
+
+todo: make model with strain as fixed effect (since, there are interesting strain effects)
 
@@
 
-> Mouse average Models
-
-  variables vs char coding of coefficents(opposite direction of sex)?
-  select1 (remove old mice, remove mice with low ncells)
-  select2 make sure there are equal numbers of observations
-  select3 low Musc males
-
-the pooled cell models are breaking for some reason, (the sexes)
--pooled cells (nMLH1)
-  all 
-  pooled.HQ (1-4)
-  pooled.HQ.2 (sample equal numbers of cells)
-
-```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE}
+```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE}
 library(lme4)
 library(nlme)
+
+#change molossinus to it's own subsp (grepl)
+DF.HetC.MixedModel2 <- DF.HetC.MixedModel
+
+#pull out values / other things
+#summary(model.ff2)# indeces for this object: [1]-structure
+#ranef(model.ff2)
+#after coding subsp to chr from factor, the co-efficients are
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#after initial analysis look into quality of these mice and ages
+
+#clean up the DF, remove extra cols
+
+#colnames(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                                                            subsp.y,subsp.x,
+                                                            category.y, category.x,
+                                                            strain.x,strain.y, mouse.1  )  )
+
+```
+
+For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)
+
+subspecies levels: 2, Musc and Domesticus (maybe Molossinus soon)
+strain levels: Dom: 3, Musc: 4
+sex: 2 female male
+
+`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
+
+> Mouse average Models
+
+List of mixed models with mean_co
+
+1. mean_co (chr coded)
+2. mean_co.HQ
+3. mean_co with Mol
+4. sub-sampling
+
+```{r mixed.models.setup, echo=FALSE}
+library(raster)
+#make are the DFs for the models
 # seperate the data, both sexes, only Mus
@@
 
+#main file, make everything unordered factors
 AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
@@
 
-#change molossinus to it's own subsp (grepl)
-DF.HetC.MixedModel2 <- DF.HetC.MixedModel
-
-#if(grepl("MSM", DF.HetC.MixedModel2$strain) ){ DF.HetC.MixedModel2$subsp <- "Mol" }
-
-#subsp are charecters
-#change sex to charecter? (unordered sex)
-
-DF.HetC.MixedModel2$sex <- as.character(DF.HetC.MixedModel2$sex)
-
-model.ff2 <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel2, random=list(strain=pdDiag(~sex) ) )#adding mol
-
-#pull out values / other things
-#summary(model.ff2)# indeces for this object: [1]-structure
-#ranef(model.ff2)
-#after coding subsp to chr from factor, the co-efficients are
-
-#intercept should be the Dom female mean, looks right
-mean(DF.HetC.MixedModel2$mean_co[(DF.HetC.MixedModel2$sex == "female" &
-                                    DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5
-
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-
-#after initial analysis look into quality of these mice and ages
-
-#clean up the DF, remove extra cols
-
-#colnames(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
                                                             subsp.y,subsp.x,
@@
                                                             strain.x,strain.y, mouse.1  )  )
+
+#make dummy MOLF data, average of Musc females?
+#no MOLF male data?
+female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
+table(female.musc.cells$quality)
+
+female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)#just 1 entry
+
+female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
+female.musc.mice.table$DOB <- as.Date("1/1/2020")
+
+female.musc.mice.table$strain <- "MOLF"
+female.musc.mice.table$subsp <- "Musc"
+female.musc.mice.table$species <- "M.musculus"
+
+female.musc.mice.table$female.stage <- "emb"
+female.musc.mice.table$age.days <-""
+female.musc.mice.table$age.weeks <- ""
+female.musc.mice.table$ID <- ""
+female.musc.mice.table$notes <- ""
+female.musc.mice.table$euth.date <- ""
+
+DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
+
+#main df with variables coded as chr
+DF.HetC.MixedModel <- DF.HetC.MixedModel
+DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
+DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
+DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
+
+#mean_co.HQ  (this might remove some of the MOLF males)
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mean_co.HQ.mol
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
+        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
+        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
+
+#sub-sample (need to think about this one)
+
+```
+
+Peromyscus did everything as factors... (but unordered)
+
+random effects were tested with 
+ #rand LRT
+this function might not support MM, (and needs linear models instead)
+
+also throws an error about only looking at 1 random effect
+
+Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
+Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
+SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
+SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
+
+(what does the significant random test mean) -- variance is not 0?
+
+```{r BaseMixedModel, echo=FALSE}
+library(RLRsim)
+library(lme4)
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+
+#strain as fixed
+newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
+
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
+
+full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
+
+#CONTRAST models
+
+#base vs strain
+fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co)# <.0001
+
+#strain vs no random full.reduce
+fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)#<.0001
+
+#fullreduce vs strain
+
+#can't distinguish which random effects are more significant -- both tests are significant
+
+fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
+
+#random effects for base model
+ranef(HetC.MixedModel.mean_co)
+#random effect for just strain
+ranef(R.strain.HetC.MixedModel.mean_co)
+
+#reduced model's PWD radom effects is much smaller
+#the full model's largest random effects are for PWD and MSM (and MOLF, KAZ for negative)
+
+#fixed effects from full
+anova(HetC.MixedModel.mean_co)#NS
+
+#Check how the fixed effects change across these
+
+HetC.MixedModel.mean_co2 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel)
+#lmer(y~x1 + x2 + (1|g1), data=pbDat)
+
+lmer.MM <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel, REML = TRUE)
+
+```
+
+Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
+Models with random effects more significant than nulls (both just strain and no radom effects).
+The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
+
+```{r Base.strain.fixed, echo=FALSE}
+
+#make mixed model with strain as fixed, since there are interesting results
+
+HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
+#this produces several significant factors
+
+HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
+summary(HetC.MM.strain.fixed2)
+#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
+
+summary(HetC.MM.strain.fixed)
+#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
+#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
+#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
+#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
+#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
+#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
+#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
+#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
+
+#male effect lowers the mean from intercept
+#while strain effects 
+```
+
+```{r HQ.MixedModel, echo=FALSE}
+table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#mean_co.HQ  (this might remove some of the MOLF males)
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#these remove ALL male MSM!
+table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#mean.co.HQ
+HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.HQ)
+ranef(HetC.MixedModel.HQ)
+
+#reduced models
+Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+summary(Reduced.strain.HQ)
+ranef(Reduced.strain.HQ)
+
+Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+summary(Reduced.all.HQ)
+ranef(Reduced.all.HQ)
+
+#strain = fixed
+strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+summary(strain.fixed.full)
+#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
+#sexmale:strainWSB p=  0.006740 **
+#sexmale  p =  3.98e-05 ***
+#strainWSB   p=0.000441 ***
+#subspMusc:sexmale  p=  0.069493 . 
+#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are
+
+strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+summary(strain.fixed)
+
+#sexmale  p= 3.98e-05 ***
+#strainKAZ p=  0.052420 .  
+#strainWSB  p= 0.000441 ***
+#sexmale:strainMSM  p=3.42e-09 ***
+#sexmale:strainPWD  p= 5.79e-08 ***
+#sexmale:strainSKIVE p= 0.069493 .  
+#sexmale:strainWSB  p=  0.006740 **
+
+#maybe this WSB effec is just in the HQ data set?
+
+#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
+full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
+
+full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
+#this comparison not useful when there are different fixed effects
+
+full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ)# < .0001
+
+strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ)#NS
+
+```
+
+Moatly matches the base model results. the same fixed effects are significant in the reduced models. the random effects make sense
+
+```{r MolossinusMixedModel, echo=FALSE}
+#base model with molossinus level
+
+DF.HetC.MixedModel.mol <- DF.HetC.MixedModel
+str(DF.HetC.MixedModel.mol)#subsp is factor with 2 levels
+
+DF.HetC.MixedModel.mol$subsp <- as.character(DF.HetC.MixedModel.mol$subsp)
+
+#turn to chr first?
+#apparently ifelse doesn't retain class
+library(data.table)
+
+DF.HetC.MixedModel.mol$subsp <- if_else(grepl("MSM", DF.HetC.MixedModel.mol$strain), "Mol", 
+        if_else(grepl("MOLF", DF.HetC.MixedModel.mol$strain), "Mol", DF.HetC.MixedModel.mol$subsp))
+
+str(DF.HetC.MixedModel.mol)
+DF.HetC.MixedModel.mol$subsp <- factor(DF.HetC.MixedModel.mol$subsp, ordered = FALSE)
+
+#these remove ALL male MSM!
+table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#mean.co
+HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.mol)
+ranef(HetC.MixedModel.mol)
+
+#mreducing the model (reducing the random effects)
+Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
+summary(Reduced.strain.mol)
+ranef(Reduced.strain.mol)
+
+Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
+summary(Reduced.all.mol)
+
+#anova tests
+#full vs strain
+full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
+#full vs non
+full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
+#strain vs non
+strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
+
+anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
+
+#base vs mol model (this only makes sense if these are nested models)
+anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
+
+```
+
+Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed modles by strain.
+make articifial data / models, like human labels
+
+```{r Base.Variance.MM, echo=FALSE}
+#there's 1 mouse with just 1 cell remove
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
+
+DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
+
+#main var.CO
+HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.var_co)
+ranef(HetC.MixedModel.var_co)
+
+#main cV.CO
+HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.cVco)
+ranef(HetC.MixedModel.cVco)
+
+just.strain.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain) 
+summary(just.strain.cV_co)
+
+full.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel)
+summary(full.Reduced.cv_co)
+
+#contrast models
+full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
+
+full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
+
+strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
+#for var, p = 0.0219
+
+```
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE}
 
@@
 DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
-```
-
-For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)
-
-subspecies levels: 2, Musc and Domesticus (maybe Molossinus soon)
-strain levels: Dom: 3, Musc: 4
-sex: 2 female male
-
-`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
-
-# Building Mixed Models
-
-```{r mixed.model.make, echo=FALSE, warning=FALSE, eval=FALSE, message=FALSE}
-#build mixed model
-#run simulations for accounting for sampling error
-# old code uses base plot
-library(nlme)
-library(lme4)
-
-#L,Q,C  appear for ordered factors in the summary mean lme doesn't believe you used ordered factors
-#change all the DFs / variables to UNORDERED factors!
-#redo the orders just to be sure
-#sex, strain, subsp,
-
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE, levels =c("female", "male") )
-
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE, levels =c( "Dom", "Musc") )
-
-DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW",
-                                                  "PWD", "MSM", "SKIVE", "KAZ") )
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
 
@@
 
-#sex only has 2 levels and this breaks...
-model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.pooled)# indeces for this object: [1]-structure
-ranef(model.pooled)
-
-model.pooled$coefficients
-```
-
-List of mixed models with mean_co
-
-1. mean_co --DF.HetC.MixedModel
-2. mean_co.chr  variables coded as factors (for ggplots), how is this changed when they are coded as chr
-3. mean_co.fac -- non-ordered factors
-4. mean_co.HQ
-5. mean_co.HQ.mol
-6. Sub-sampling for equal numbers
-
-```{r mixed.models.setup, echo=FALSE}
-#make are the DFs for the models
-
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-                                                            subsp.y,subsp.x,
-                                                            category.y, category.x,
-                                                            strain.x,strain.y, mouse.1  )  )
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
-
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE )
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE )
-DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE )
-
-#main df with variables coded as chr
-DF.HetC.MixedModel.chr <- DF.HetC.MixedModel
-DF.HetC.MixedModel.chr$subsp <- as.character(DF.HetC.MixedModel$subsp)
-DF.HetC.MixedModel.chr$sex <- as.character(DF.HetC.MixedModel$sex)
-DF.HetC.MixedModel.chr$strain <- as.character(DF.HetC.MixedModel$strain)
-
-#mean_co.HQ
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mean_co.HQ.mol
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
-        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
-        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
-
-#mean_co.chr.HQ.mol
-DF.HetC.MixedModel.chr.HQ.mol <- DF.HetC.MixedModel.HQ.mol
-
-DF.HetC.MixedModel.chr.HQ.mol$subsp <- as.character(DF.HetC.MixedModel.chr.HQ.mol$subsp)
-DF.HetC.MixedModel.chr.HQ.mol$sex <- as.character(DF.HetC.MixedModel.chr.HQ.mol$sex)
-DF.HetC.MixedModel.chr.HQ.mol$strain <- as.character(DF.HetC.MixedModel.chr.HQ.mol$strain)
-
-```
-
-```{r MixedModels.stff, echo=FALSE}
-#main mean.CO
-HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-#mean.co.chr
-HetC.MixedModel.mean_co.chr <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr, random=list(strain=pdDiag(~sex) ) )
-
-#mean.co.HQ
-HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-#mean.CO.HQ.mol
-HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-#mean.HQ.mol.chr
-HetC.MixedModel.chr.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-```
-
-The intercept level is Dom female. These are the coefficients
-
-`r kable(HetC.MixedModel.mean_co$coefficients$fixed, col.names = "coefficients")`
+```
+
+```{r HQ.var, echo=FALSE}
+
+DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
+
+#HQ var.CO
+HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.HQ.var_co)
+ranef(HetC.MixedModel.HQ.var_co)
+
+HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+summary(HQjust.strain.var_co)
+
+HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+summary(HQfull.Reduced.cv_co)
+
+#contrast models
+full2strain.CV <- anova(HetC.MixedModel.HQ.var_co, HQjust.strain.var_co)#
+
+full2noRandom.CV <- anova(HetC.MixedModel.HQ.var_co, HQfull.Reduced.cv_co)#
+
+strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
+
+```
+
+- The MSM strain effec is 0, but the sex*strain for molossinus is 7 (this seems very high).
+This means there is 0 variance added to mean_CO to mean due to being MSM.
+Does this mean the model is very good at predicting MSM mean_CO.
+The fixed strain - sex effects are for musc are large.
+This will likely change with MOLF (male molf are low, so smaller fixed effect and )
+
+- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
+In the first model MSM has largest random effect, since it's so far from WSB female.
+
+-adding MOLF also doesn't have significant factors
+
+The intercept level is Dom female.
+
+`r kable(HetC.MixedModel.mean_co$coefficients$fixed, col.names = "coefficients", caption="Main mean_co model")`
 
@@
 changed all to unordered factors.
-
 all the variables are coded as ordered factors.
-
 When the variables are kept factors, and unordered the names appear! yay!
 
-`r kable(HetC.MixedModel.mean_co.chr$coefficients$fixed, caption="Chr coded Model", col.names = "coefficients")`
-
-`r kable(ranef(HetC.MixedModel.mean_co.chr) )`
-
-The variable names now appear. The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.
 
 `r kable(HetC.MixedModel.HQ$coefficients$fixed, caption="mean_co HQ model", col.names = "coefficients")`
+
 `r kable(ranef(HetC.MixedModel.HQ) )`
 
-The factor coded DF the more high quality cells.
-
-`r kable(HetC.MixedModel.HQ.mol$coefficients$fixed, caption="Main model with molossinus", col.names = "coefficient")`
+Fixed and random effects
+
+`r kable(HetC.MixedModel.HQ.mol$coefficients$fixed, caption="Main HQ model with molossinus", col.names = "coefficient")`
+
 `r kable(ranef(HetC.MixedModel.HQ.mol) )`
-With adding molossinus, the subsp names are listed.
-
-`r kable(HetC.MixedModel.chr.HQ.mol$coefficients$fixed, caption="Main model, chr coded with molossinus", col.names = "coefficients")`
-`r kable(ranef(HetC.MixedModel.chr.HQ.mol) )`
-
-When the variables/levels are coded as chr, the interaction effects get larger and the other subspecies get smaller. The base sex effect is also larger. The subsp effect is relative to Dom female,
-
-I don't think the subspecies and strains should be order, (maybe factors still makes sense.)
-
-This table doesn't display well
-
-`r table(DF.HetC.MixedModel.pooled$sex,DF.HetC.MixedModel.pooled$strain)`
-
-KAZ, LEWES has ~100 more for males, PWD has ~200 for females.
-
-Fixed effects
-
-look at the p-values for each of the effects as an indication if they are significant predictors.
-
-Subspecies
-
-Random effects
-
-try subsampling some of the
-
-## Meiotic Error
-
-This section reviews all the data from achiasmy and 0CO bivalents.
 
@@
 
-## SC.skel measures
-
-data has been collection for RWang's total SC. These results are consistant with the patterns seen in the bivalent level resutls. (total SC PWD > KAZ )
+## Total SC Measures
+
+data has been collection for total SC per cell [@wang_2019_sc]. These results are consistant with the patterns seen in the bivalent level resutls. (total SC PWD > KAZ )
 
@@
 SC.skel$rand_name <- as.character(SC.skel$rand_name)
+
+#check for duplicatations of the random name
 original_DF$Random.Name <- as.character(original_DF$Random.Name)
+duplies <- original_DF[duplicated(original_DF$Random.Name),]
+
+blah <- anyDuplicated(original_DF$Random.Name)#there are 64 duplicated random names -- which looks like 3 rows of ""
+
 #str(SC.skel)
@@
 mlh1.skel.og <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
+#what if there are duplicate random names!!!
+
 #for some reason this only worked after the chr and all = F
@@
 there could be something about kinetochore sites differing across specific chromosomes that balance the force
+
+# References

### 2019-12-17 doc/Main_Report_Results.rmd
missed some updating

@@
 
-#Setup
+# Results
 
@@
 
-load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
+load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
+#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
 #load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
@@
 
-1. The quality of MLH1 counts for individual mice is / was assessed in the MouseLevelPlots Rmd (can I link this file?). A list, mice.to.exclude was made.
-
-2.
-The list of excluded mice is applied in the MLH1.setup.data script, to remove these mice from the main dataset/dataframe.
-This script saves .RData file of the dataset which is loaded into Rmd files for analysis
-
-Mouse specific plots of CO count and quality were examined for general quality from single mice.
-Mice which produced few cells of low quality were excluded to the 'HQ' (high quality) dataset.
-(The full dataset was used to look at some patterns)
+Details of dataset cleaning ect is in SetupData.rmd
+
+Effects from different factors are reviewed in Effects.Rmd
 
@@
 check if there is a quality effect.
+
+```{r MLH1.histogram.2, echo=FALSE}
+#remake the pooled MLH1 histgoram plot from above but facet by quality
+
+#PWD male, MSM male
+#LEW female, WSB female
+mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
+                 .(category, quality),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+#what's the point of the above table? -- add vlines?
+
+#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect
+
+histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
+histM_qual.KAZ_male
+
+histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
+histM_qual.PWD_male
+
+histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
+histM_qual.MSM_male
+
+histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
+histM_qual.WSB_female
+
+histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
+histM_qual.LEW_female
+
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
+#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
+
+#ad
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+
+```
 

### 2019-12-17 src/SetupData.Rmd
missed some updating

@@
 ```{r exclude.mice, echo=FALSE}
-#remove mice 
+#remove mice -- 18nov17_WSB_f5 (only 1 cell)
 exclude_mice_list = c("18nov17_WSB_f1", "8oct14_WSB_f1", "8oct14_WSB_f3",
@@
                         cV = cv(adj_nMLH1.foci),
-                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
                         sd   = round(sd(adj_nMLH1.foci), 3),
@@
 #what are the differences between these two
+
+```
+
+```{r strain.table, echo=FALSE}
 

### 2019-12-17 doc/figureFiles/figures.Rmd
finished figure 1 drafts

@@
 library(raster)#for cV
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+
+setwd("~./MLH1repo/")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 #setwd("C:/Users/April/Desktop/MLH1repo")
@@
 
-```{r fig1_scatter_basic, echo=FALSE, warning=FALSE}
-#mouse averages
-#AP_mouse_table.HQ
-#fig.data  cell data
-
-fig.data <- fig.data %>%
-  arrange(strain, sex, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
-# sort your dataframe, by the focal categories
-fig.data$mouse <- factor(fig.data$mouse, levels=unique(fig.data$mouse))
-
-fig.data.male <- subset(fig.data, sex== "male")
-
-#male
-Fig1_A_mouse.male <- ggplot(fig.data.male[order(fig.data.male$strain), ],
-  aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+  #geom_jitter()+
-  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
-  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+geom_boxplot(alpha = 0.6)+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
-        panel.background = element_blank(), axis.line = element_line(colour = "black"),
-
-  #lines -- these might be tricky with the faceting
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())  +
-  #Dom
-  annotate("segment", x= 10, xend=20.5, y=15, yend=15, colour="black", size=1)+
-  #Musc
-  annotate("segment", x= 50, xend=55, y=15, yend=15, colour="black", size=1)+
-  #Cast
-  annotate("segment", x=70, xend=80, y=15, yend=15, colour="black", size=1)
-
-#female
-#the data is mouse mean 
-fig.data.female <- subset(fig.data, sex== "female")
-
-Fig1_A_mouse.female <- ggplot(fig.data.female[order(fig.data.female$strain), ], aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+ #geom_jitter(width = 0.25)+
-  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
-  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+
- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
-        panel.background = element_blank(), axis.line = element_line(colour = "black"),
-  #lines -- these might be tricky with the faceting
- axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
-  #Dom
-  annotate("segment", x= 10, xend=20.5, y=18, yend=18, colour="black", size=1)+
-  #Musc
-  annotate("segment", x= 50, xend=55, y=18, yend=18, colour="black", size=1)+
-  #Cast
-  annotate("segment", x=70, xend=80, y=18, yend=18, colour="black", size=1)
-
-```
-
-# Figure 1
+# Option 1, strain averages
 
@@
 
+#
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
 ```
@@
 
-I think some strain observations can be removed, since they only have
-
 ```{r strain.table.plot, echo=FALSE, warning=FALSE}
@@
 #remove y labfor male 
+  #segments
+  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+  annotate("text", x = c(3,8.5,10), #list of all x points 
+           y = c(18.5,18.5,33),
 
-  #segments
-  annotate("segment", x= 0, xend=5, y=21, yend=21, colour="black", size=1)+
-  #Musc
-  annotate("segment", x= 5, xend=8, y=20, yend=20, colour="black", size=1)+
-  #Cast
-  annotate("segment", x=11, xend=15, y=18, yend=18, colour="black", size=1)+
-  #label
-  annotate("text", x = c(1.5,8,10.5, 11), #list of all x points 
-           y = c(20,18, 20,18),
-
-           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus", "M.m.castaneus"),
+   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
            fontface = 'italic', hjust = 0.5, size=3.5)
@@
 # 2
-Fig1_B_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 foci per Cell")+
+Fig1_B_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
   geom_point(size=3)+
@@
   #segments
-  annotate("segment", x=1, xend=3.5, y=22, yend=22, colour="black", size=1)+
+  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
   #Musc
-  annotate("segment", x=5, xend=8.5, y=20, yend=20, colour="black", size=1)+
+  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
+#mol
+ annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
     #Cast
@@
 #Labels
- annotate("text", x = c(1.5,8), #list of all x points 
-           y = c(20,18),
-           label = c("M. m. domesticus","M. m. musculus"),
-           fontface = 'italic', hjust = 0.5, size=3)
+ annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
+           y = c(18.5,18.5,33),
+           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
 
@@
 fig1_B  <- ggdraw() +
-  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.48, height=.88) +
-  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.48, height=.88) +
+  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.5, height=.97) +
+  draw_plot(Fig1_B_strain.male, x=.45, y=0, width=.5, height=.97) +
 
@@
 
-add molossinus
+I removed the cast and outgroup measures.
+I left off the color legend, but I think I could add one to the bottom of the figure.
+These show the strain averages, but I can try adding the mouse averages, if you think having mulitple points would show the data better.
 
-re-arrange Muusc strain level order to highlight the polymorphism to musc
+## Option 2, mouse boxplots
 
-Clean up segments
-There are many more male strains represented.
+```{r fig1_scatter_basic, echo=FALSE, warning=FALSE}
+#mouse averages
+#AP_mouse_table.HQ
+#fig.data  cell data
 
-The male plot is squished because of the legend
+fig.data <- fig.data %>%
+  arrange(strain, sex, mouse) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
 
-## Figure 1, mouse levels shown
+# sort your dataframe, by the focal categories
+fig.data$mouse <- factor(fig.data$mouse, levels=unique(fig.data$mouse))
+
+fig.data.male <- subset(fig.data, sex== "male")
+
+#male
+Fig1_A_mouse.male <- ggplot(fig.data.male[order(fig.data.male$strain), ],
+  aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+  #geom_jitter()+
+  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
+  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+geom_boxplot(alpha = 0.6)+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
+        panel.background = element_blank(), axis.line = element_line(colour = "black"),
+  #lines -- these might be tricky with the faceting
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+  #Dom
+#  annotate("segment", x= 10, xend=20.5, y=15, yend=15, colour="black", size=1)+
+  #Musc
+#  annotate("segment", x= 50, xend=55, y=15, yend=15, colour="black", size=1)+
+  #Cast
+#  annotate("segment", x=70, xend=80, y=15, yend=15, colour="black", size=1)
+
+#female
+#the data is mouse mean 
+fig.data.female <- subset(fig.data, sex== "female")
+
+Fig1_A_mouse.female <- ggplot(fig.data.female[order(fig.data.female$strain), ], aes(y=nMLH1.foci, x=as.factor(mouse), color=strain))+ #geom_jitter(width = 0.25)+
+  geom_boxplot(alpha = 0.6)+ylim(c(12,48))+
+  facet_grid(~sex, scales = "free", space="free")+scale_color_manual(values=colors_of_strains)+
+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="none",
+        panel.background = element_blank(), axis.line = element_line(colour = "black"),
+  #lines -- these might be tricky with the faceting
+ axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+  #Dom
+#  annotate("segment", x= 10, xend=20.5, y=18, yend=18, colour="black", size=1)+
+  #Musc
+#  annotate("segment", x= 50, xend=55, y=18, yend=18, colour="black", size=1)+
+  #Cast
+#  annotate("segment", x=70, xend=80, y=18, yend=18, colour="black", size=1)
+
+```
 
@@
   draw_plot(Fig1_A_mouse.female, x=0, y = 0, width=.48, height=.88) +
-  draw_plot(Fig1_A_mouse.male, x=.5, y=0, width=.48, height=.88) +
+  draw_plot(Fig1_A_mouse.male, x=.48, y=0, width=.48, height=.88) +
   draw_plot_label(label = c("A", "B"), size = 15,
@@
 
-USE boxplots --- this is a bit too busy with single mouse level observations. (so forget the cell level data I just coded. What if I sub-sampled?)
+This combination of points and box plots is similar to the final peromyscus figure 1, but the number of individual mice observations makes it very busy.
 
-Draft first fig. MLH1 distributions will be the main first figure
-I think a strain level would look better.  Add white space between
-
-I still need to work on some colors
-add line segments, they are good at anchoring
-
-add line segments for strain means?, do boxplots?
-
-Note that this figure knitted in html will look different than a saved png or tif.
-change the ylabel 'Mean Number of MLH1 Foci per Cell'
+This might be a good format if fewer mice were used for each strain.

### 2019-12-18 doc/figureFiles/figures.Rmd
started v2 of main figure

@@
 
-setwd("~./MLH1repo/")
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+#setwd("~./MLH1repo/")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 #setwd("C:/Users/April/Desktop/MLH1repo")
@@
 
+#remove male only data
+
 #I need cell level data 
-
 uni.mice <- unique(AP_mouse_table.HQ$mouse)
@@
 
-# Option 1, strain averages
+# Option 1, strain averages, sexes matched
 
-```{r strain.table, echo=FALSE, warning=FALSE}
-#fig.data -- (only mice from better list)
+```{r strain.table.v2, echo=FALSE, warning=FALSE}
 
@@
 #cv, sd and se are all NA / empty
-strain.table.HQ <- strain.table.HQ[!(is.na(strain.table.HQ$cV) | strain.table.HQ$cV== "" ), ]
+#this removes MOLF female
+#strain.table.HQ <- strain.table.HQ[!(is.na(strain.table.HQ$cV) | strain.table.HQ$cV== "" ), ]
 
@@
 
-#
+#REMOVE ALL STRAINS WITHOUT FEMALE DATA
 strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]
 strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
+
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "PERC"),]
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "TOM"),]
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "AST"),]
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CZECH"),]
+
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
+strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]
+
 ```
 
-```{r reorder.strain, echo=FALSE}
+```{r reorder.strain.v2, echo=FALSE}
 
@@
 
+```{r strain.table.plot.v2, echo=FALSE, warning=FALSE}
+#use the MLH1 / cell level data to calq the error bars
+#strain table ggplot
+
+Fig1_Bv2_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
+  ylab("")+
+  geom_point(size=3)+
+  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+ 
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+#remove y labfor male 
+  #Dom
+  annotate("segment", x=.75, xend=3.6, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x=3.9, xend=6.15, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+  annotate("segment", x=6.5, xend=8.1, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+  annotate("text", x = c(2,5,7.15), #list of all x points 
+           y = c(18.5,18.5, 18.5),
+
+   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+#this can pull just the lengend object
+#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
+#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
+
+# 2
+Fig1_Bv2_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
+  geom_point(size=3)+
+  ylim(c(18, 35))+
+  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
+  scale_color_manual(values=colors_of_strains)+ 
+  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line = element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+  #segments
+  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
+#mol
+ annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
+    #Cast
+#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
+#Labels
+ annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
+           y = c(18.5,18.5,18.5),
+           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+```
+
+```{r Strain.fig.merge.v2, echo=FALSE}
+
+fig1_B_v2  <- ggdraw() +
+  draw_plot(Fig1_Bv2_strain.female, x=0, y = 0, width=.5, height=.97) +
+  draw_plot(Fig1_Bv2_strain.male, x=.45, y=0, width=.5, height=.97) +
+
+  draw_plot_label(label = c("A", "B"), size = 15,
+                  x = c(0, .45), y = c(1, 1) )
+fig1_B_v2
+```
+
+Molf female data point is missing because there is only 1 female, so NAs were generated.
+Mol text should be moved back above the points.
+
+## option 2, strain averages, more strains
+
+```{r strain.table, echo=FALSE, warning=FALSE}
+#fig.data -- (only mice from better list)
+
+#there are still quality 5 cells in this dataframe
+fig.data <- subset(fig.data, quality < 5)
+
+#use the MLH1 / cell level data to calq the error bars
+strain.table.HQ.backup <- ddply(fig.data, c("strain", "sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci, na.rm = TRUE),
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                        se   = round(sd / sqrt(Nmice), 3)
+    )
+
+#cv, sd and se are all NA / empty
+#strain.table.HQ.backup <- strain.table.HQ.backup[!(is.na(strain.table.HQ.backup$cV) | strain.table.HQ.backup$cV== "" ), ]
+
+#this removes MOLF female
+
+#remove outgroups
+#remove F1
+strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "F1"),]
+
+#remove outgroups
+strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "SPRET"),]
+strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "SPIC"),]
+strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "CAROLI"),]
+
+strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "CAST"),]
+strain.table.HQ.backup <- strain.table.HQ.backup[(strain.table.HQ.backup$strain != "HMI"),]
+
+#Keep more of the strains
+#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "PERC"),]
+#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "TOM"),]
+#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "AST"),]
+#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CZECH"),]
+
+#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "HMI"),]
+#strain.table.HQ <- strain.table.HQ[(strain.table.HQ$strain != "CAST"),]
+
+```
+
+```{r reorder.strain, echo=FALSE}
+
+#re-order strains
+#Musc
+#PWD, SKIVE, KAZ, CZECH, TOM, AST
+
+#MOl
+#MSM, MOLF
+strain.table.HQ.backup$strain<- factor(strain.table.HQ.backup$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
+                           "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
+                                                    "MSM", "MOLF",
+                                          "CAST", "HMI",
+                             "SPRET", "SPIC", "CAROLI", "F1", "other") )
+```
+
 ```{r strain.table.plot, echo=FALSE, warning=FALSE}
@@
 
-Fig1_B_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
+Fig1_B_strain.male <- ggplot(strain.table.HQ.backup[strain.table.HQ.backup$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
   ylab("")+
@@
 # 2
-Fig1_B_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
+Fig1_B_strain.female <- ggplot(strain.table.HQ.backup[strain.table.HQ.backup$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
   geom_point(size=3)+
@@
 ```{r Strain.fig.merge, echo=FALSE}
-fig1_B  <- ggdraw() +
+fig1_B_v1  <- ggdraw() +
   draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.5, height=.97) +
@@
                   x = c(0, .5), y = c(1, 1) )
-fig1_B
+fig1_B_v1
 ```
+
+This is currently backup version.
 
@@
 
-## Option 2, mouse boxplots
+## Option 3, mouse boxplots
 

### 2019-12-18 doc/figureFiles/figures.Rmd
changed error bars to 2SE

@@
   geom_point(size=3)+
-  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
+  geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3)+
   ylim(c(18, 35))+
@@
   #label
-  annotate("text", x = c(2,5,7.15), #list of all x points 
-           y = c(18.5,18.5, 18.5),
+  annotate("text", x = c(2.2,5.9,6.1), #list of all x points 
+           y = c(18.5,18.5, 33),
 
@@
   ylim(c(18, 35))+
-  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
+  geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2)), width=.3) +
   scale_color_manual(values=colors_of_strains)+ 
@@
 #Labels
- annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
-           y = c(18.5,18.5,18.5),
+ annotate("text", x = c(2.25,5.9,6.1), #list of all x points 
+           y = c(18.5,18.5,33),
            label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
@@
 ```{r Strain.fig.merge.v2, echo=FALSE}
-
 fig1_B_v2  <- ggdraw() +
-  draw_plot(Fig1_Bv2_strain.female, x=0, y = 0, width=.5, height=.97) +
-  draw_plot(Fig1_Bv2_strain.male, x=.45, y=0, width=.5, height=.97) +
+  draw_plot(Fig1_Bv2_strain.female, x=0, y = 0, width=.47, height=.97) +
+  draw_plot(Fig1_Bv2_strain.male, x=.49, y=0, width=.48, height=.97) +
 
   draw_plot_label(label = c("A", "B"), size = 15,
-                  x = c(0, .45), y = c(1, 1) )
+                  x = c(0, .465), y = c(1, 1) )
 fig1_B_v2
@@
 
-## option 2, strain averages, more strains
+## Option 2, strain averages, more strains
 

### 2019-12-18 doc/figureFiles/figures.Rmd
removed facet ~sex, replace shape = sex. version sent to bret

--- 2019-12-18 doc/figureFiles/figures.Rmd
@@
 
-# Option 1, strain averages, sexes matched
+# Option 1, strain averages, sexes matched, cells pooled
 
@@
 
-#use the MLH1 / cell level data to calq the error bars
+#step-wise, 1)mouse-level, 2)averages of mouse-levels
+
+#Pooling all cells for a category
 strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        Nmice =length(unique(mouse)),
+                        Ncells  =length(nMLH1.foci),
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                         cV = cv(nMLH1.foci, na.rm = TRUE),
-                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                         sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
-                        se   = round(sd / sqrt(Nmice), 3)
+                       #Standard error (mice vs cells)
+                        se  = round(sd / sqrt(Ncells), 3)
     )
@@
 #strain table ggplot
+#remove facet by sex
 
-Fig1_Bv2_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
-  ylab("")+
-  geom_point(size=3)+
-  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
-  ylim(c(18, 35))+
+Fig1_Bv2_strain <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
+  #facet_grid(~sex, scales = "free", space="free")+ 
+  ylab("Average MLH1 Foci per Cell")+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+
+   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(19.5, 33))+
  scale_color_manual(values=colors_of_strains)+ 
-  theme(legend.position="none",
+  #legend.position="none",
+  theme(legend.position="bottom",legend.box = "horizontal",
   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
   axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+
 #remove y labfor male 
   #Dom
-  annotate("segment", x=.75, xend=3.6, y=19.5, yend=19.5, colour="black", size=1)+
+  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
   #Musc
-  annotate("segment", x=3.9, xend=6.15, y=19.5, yend=19.5, colour="black", size=1)+
+  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
   #mol
-  annotate("segment", x=6.5, xend=8.1, y=19.5, yend=19.5, colour="black", size=1)+
+  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)+
   #label
-  annotate("text", x = c(2,5,7.15), #list of all x points 
-           y = c(18.5,18.5, 18.5),
-
-   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
+  annotate("text", x = c(2.2, 4.5, 7), #list of all x points 
+           y = c(20.2,20.2,20.2),
+   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus"),
            fontface = 'italic', hjust = 0.5, size=3.5)
@@
 #legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
-
-# 2
-Fig1_Bv2_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
-  geom_point(size=3)+
-  ylim(c(18, 35))+
-  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
-  scale_color_manual(values=colors_of_strains)+ 
-  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line = element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
-  #segments
-  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
-  #Musc
-  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
-#mol
- annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
-    #Cast
-#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
-#Labels
- annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
-           y = c(18.5,18.5,18.5),
-           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
-           fontface = 'italic', hjust = 0.5, size=3.5)
-
 ```
@@
 ```{r Strain.fig.merge.v2, echo=FALSE}
-
-fig1_B_v2  <- ggdraw() +
-  draw_plot(Fig1_Bv2_strain.female, x=0, y = 0, width=.5, height=.97) +
-  draw_plot(Fig1_Bv2_strain.male, x=.45, y=0, width=.5, height=.97) +
-
-  draw_plot_label(label = c("A", "B"), size = 15,
-                  x = c(0, .45), y = c(1, 1) )
-fig1_B_v2
+Fig1_Bv2_strain
+#save a tif or png file
+ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
 ```
 
-Molf female data point is missing because there is only 1 female, so NAs were generated.
-Mol text should be moved back above the points.
+Error bars are 2 SE,
 
-## option 2, strain averages, more strains
+S.E. = sd( all cells) / n.cells
+
+## Option 2, strain averages, more strains, pooled cells
 
@@
 fig.data <- subset(fig.data, quality < 5)
+
+mouse.table.HQ.backup <- ddply(fig.data, c("mouse"), summarise,
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci, na.rm = TRUE),
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                        se   = round(sd / sqrt(Ncells), 3) #mean is pooled cells, so sd should be pooled
+    )
+#need to add strain and sex
+mouse.table.HQ.backup <- add_strain(mouse.table.HQ.backup)
+mouse.table.HQ.backup <- add_sex(mouse.table.HQ.backup)
+
+strain.table.from.mouse.HQ.backup <- ddply(mouse.table.HQ.backup, c("strain","sex"), summarise,
+                  Nmice  = length(mouse),
+                  mean_co.of.mice = as.numeric(format(round(  mean(mean_co), 3 ), nsmall=3) ),
+                  cV.of.mice = cv(mean_co, na.rm = TRUE),
+                   var.of.mice = as.numeric(format(round(   var(mean_co, na.rm = TRUE),3), nsmall=3)), #this turns to chr
+                  sd.of.mice   = round(sd(mean_co, na.rm = TRUE), 3),
+                  se.of.mice   = round(sd.of.mice / sqrt(Nmice), 3) #mean is pooled cells, so sd should be pooled
+    )
 
@@
                         sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
-                        se   = round(sd / sqrt(Nmice), 3)
+                        se   = round(sd / sqrt(Ncells), 3) #mean is pooled cells, so sd should be pooled
     )
@@
 fig1_B_v1  <- ggdraw() +
-  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.5, height=.97) +
-  draw_plot(Fig1_B_strain.male, x=.45, y=0, width=.5, height=.97) +
+  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.45, height=.97) +
+  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.45, height=.97) +
 
@@
 These show the strain averages, but I can try adding the mouse averages, if you think having mulitple points would show the data better.
+
+## Option 2.B strain averages, mouse observations
+
+```{r strain.table.plot.mouse.level, echo=FALSE, warning=FALSE}
+#use the MLH1 / cell level data to calq the error bars
+#strain table ggplot
+
+Fig1_C_strain.male <- ggplot(strain.table.from.mouse.HQ.backup[strain.table.from.mouse.HQ.backup$sex == "male",], aes(y=mean_co.of.mice, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+
+  ylab("")+
+  geom_point(size=3)+
+  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2) , ymax=mean_co.of.mice+(se.of.mice*2) ), width=.3)+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+ 
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+#remove y labfor male 
+  #segments
+  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+  annotate("text", x = c(3,8.5,10), #list of all x points 
+           y = c(18.5,18.5,33),
+
+   label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+#this can pull just the lengend object
+#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
+#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
+
+# 2
+Fig1_C_strain.female <- ggplot(strain.table.from.mouse.HQ.backup[strain.table.from.mouse.HQ.backup$sex == "female",], aes(y=mean_co.of.mice, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
+  geom_point(size=3)+
+  ylim(c(18, 35))+
+  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2), ymax=mean_co.of.mice+(se.of.mice*2)  ), width=.3) +
+  scale_color_manual(values=colors_of_strains)+ 
+  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line = element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+  #segments
+  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
+#mol
+ annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
+    #Cast
+#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
+#Labels
+ annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
+           y = c(18.5,18.5,33),
+           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+```
+
+```{r Strain.fig.merge.mouse.level, echo=FALSE}
+fig1_C_v1  <- ggdraw() +
+  draw_plot(Fig1_C_strain.female, x=0, y = 0, width=.5, height=.97) +
+  draw_plot(Fig1_C_strain.male, x=.45, y=0, width=.5, height=.97) +
+
+  draw_plot_label(label = c("A", "B"), size = 15,
+                  x = c(0, .5), y = c(1, 1) )
+fig1_C_v1
+```
+
+Error bars above are 2 S.E. of mouse average COs,
+
+S.E. = sd( mouse average COs ) / Nmice
+
+MOLF and SKIVE female data point is missing because there is only 1 female, so NAs were generated.
+Mol text should be moved back above the points.
 

### 2019-12-18 doc/Main_Report_Results.rmd
updating chapter 2 results

@@
 ```{r rand, echo=FALSE, include=FALSE, eval=FALSE}
-
-str(AP_mouse_table_w.Ages)
+#str(AP_mouse_table_w.Ages)
 
@@
 
+```{r table.for.plot, echo=FALSE}
+
+AP_strain.table.HQ <- ddply(MLH1_data, c("strain", "sex"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  =length(nMLH1.foci),
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci, na.rm = TRUE),
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                       #Standard error (mice vs cells)
+                        se  = round(sd / sqrt(Ncells), 3)
+    )
+
+```
+
+```{r Fig1, echo=FALSE, warning=FALSE}
+
+Fig1_v1_strain <- ggplot(AP_strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+  geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+ 
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+#remove y labfor male 
+  #segments
+  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
+  #cast
+  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
+  #outgroups
+  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
+           y = c(18.5,18.5,33,18.5, 18.5),
+
+   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+Fig1_v1_strain
+```
+
+# Mixed Models
+
+These are the basic Mixed Models we used as a frame work to analyze all the MLH1 data.
+
+$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)$$
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain* sex)$$
+
+The mixed model lets us separate effects to see how they effect a mouse mean CO count. The subspecies effect is a proxy for **Divergence** and the random strain effect is a proxy for **Polymorphism**.
+
+Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
+
+```{r mixed.models.setup, echo=FALSE, warning=FALSE}
+
+#why merged 
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
+                            subsp.y,subsp.x,
+                    category.y, category.x,
+                        strain.x,strain.y, mouse.1  )  )
+
+#make dummy MOLF data, average of Musc females?
+#no MOLF male data?
+female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
+table(female.musc.cells$quality)
+
+female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)
+
+#female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
+#female.musc.mice.table$DOB <- as.Date("1/1/2020")
+
+#female.musc.mice.table$strain <- "MOLF"
+#female.musc.mice.table$subsp <- "Musc"
+#female.musc.mice.table$species <- "M.musculus"
+
+#female.musc.mice.table$female.stage <- "emb"
+#female.musc.mice.table$age.days <-""
+#female.musc.mice.table$age.weeks <- ""
+#female.musc.mice.table$ID <- ""
+#female.musc.mice.table$notes <- ""
+#female.musc.mice.table$euth.date <- ""
+
+#DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)
+
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
+
+#main df with variables coded as chr
+DF.HetC.MixedModel <- DF.HetC.MixedModel
+DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
+DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
+DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
+
+#mean_co.HQ  (this might remove some of the MOLF males)
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mean_co.HQ.mol
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
+        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
+        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
+
+DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
+
+```
+
+For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus.?)
+
+subspecies levels: 2, Musc and Domesticus
+
+strain levels: Dom: 3, Musc: 4
+
+sex: 2 female male
+
+`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
+
+Above are the total number of mice in the data set
+
+These are the List of mixed models I built
+
+1. mean_co (chr coded, instead of factors)
+
+2. mean_co.HQ
+
+3. mean_co with Mol (Mol coded as a distinct subsp)
+
+4. sub-sampling (not sure if I made this one)
+
+```{r HQ.MixedModel.1, echo=FALSE}
+#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#mean_co.HQ  (this might remove some of the MOLF males)
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mean.co.HQ
+HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+#summary(HetC.MixedModel.HQ)
+#ranef(HetC.MixedModel.HQ)
+```
+
+```{r MixedModel.2, echo=FALSE}
+#reduced models
+Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+#summary(Reduced.strain.HQ)
+#ranef(Reduced.strain.HQ)
+```
+
+```{r MixedModel.2.2, echo=FALSE}
+
+Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+#summary(Reduced.all.HQ)
+#ranef(Reduced.all.HQ)
+```
+
+```{r MixedModel.3, echo=FALSE}
+#strain = fixed
+strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+#summary(strain.fixed.full)
+#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
+
+#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are 
+```
+
+```{r MixedModel.4, echo=FALSE}
+
+strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+#summary(strain.fixed)
+
+#maybe this WSB effec is just in the HQ data set?
+#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
+full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
+
+full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
+#this comparison not useful when there are different fixed effects
+
+full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ)# < .0001
+
+strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ)#NS
+```
+
+```{r try.blah, echo=FALSE}
+#library(jtools)
+#states <- as.data.frame(state.x77)
+#fit <- lm(Income ~ Frost + Illiteracy + Murder, data = states)
+#summ(fit)
+#this is a nice display package for lm
+
+res.table <- as.data.frame(coef(summary(Reduced.all.HQ)))
+res.table
+```
+
+```{r variance.boxplots, echo=FALSE, eval=FALSE}
+#how did I make this 
+#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
+#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
+#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
+#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
+
+#plot of variance
+var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
+
+#plot by sex / seperately
+# HQ data DF.HetC.MixedModel.HQ
+
+var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+var.male <- var.male   + facet_wrap( ~ strain, scales="free")
+
+cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
+
+```
+
+```{r BaseMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+#strain as fixed
+newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
+
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
+
+full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
+
+#CONTRAST models
+
+#base vs strain
+fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co)# <.0001
+
+#strain vs no random full.reduce
+fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)#<.0001
+
+#fullreduce vs strain
+
+#can't distinguish which random effects are more significant -- both tests are significant
+
+fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
+
+#random effects for base model
+ranef(HetC.MixedModel.mean_co)
+#random effect for just strain
+ranef(R.strain.HetC.MixedModel.mean_co)
+
+#reduced model's PWD radom effects is much smaller
+#the full model's largest random effects are for PWD and MSM (and MOLF, KAZ for negative)
+
+#fixed effects from full
+anova(HetC.MixedModel.mean_co)#NS
+
+#Check how the fixed effects change across these
+
+HetC.MixedModel.mean_co2 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel)
+#lmer(y~x1 + x2 + (1|g1), data=pbDat)
+
+lmer.MM <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel, REML = TRUE)
+
+```
+
+Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
+Models with random effects more significant than nulls (both just strain and no random effects).
+The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
+
+```{r Base.strain.fixed, echo=FALSE}
+
+#make mixed model with strain as fixed, since there are interesting results
+
+HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
+#this produces several significant factors
+
+HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
+summary(HetC.MM.strain.fixed2)
+#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
+
+#summary(HetC.MM.strain.fixed)
+#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
+#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
+#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
+#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
+#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
+#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
+#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
+#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
+
+#male effect lowers the mean from intercept
+#while strain effects 
+```
+
+```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
+#base model with molossinus level ... why am I making molossinus model?
+
+DF.HetC.MixedModel.mol <- DF.HetC.MixedModel
+str(DF.HetC.MixedModel.mol)#subsp is factor with 2 levels
+
+DF.HetC.MixedModel.mol$subsp <- as.character(DF.HetC.MixedModel.mol$subsp)
+
+#turn to chr first?
+#apparently ifelse doesn't retain class
+library(data.table)
+
+DF.HetC.MixedModel.mol$subsp <- if_else(grepl("MSM", DF.HetC.MixedModel.mol$strain), "Mol", 
+        if_else(grepl("MOLF", DF.HetC.MixedModel.mol$strain), "Mol", DF.HetC.MixedModel.mol$subsp))
+
+str(DF.HetC.MixedModel.mol)
+DF.HetC.MixedModel.mol$subsp <- factor(DF.HetC.MixedModel.mol$subsp, ordered = FALSE)
+
+#these remove ALL male MSM!
+table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#mean.co
+HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.mol)
+ranef(HetC.MixedModel.mol)
+
+#mreducing the model (reducing the random effects)
+Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
+summary(Reduced.strain.mol)
+ranef(Reduced.strain.mol)
+
+Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
+summary(Reduced.all.mol)
+
+#anova tests
+#full vs strain
+full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
+#full vs non
+full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
+#strain vs non
+strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
+
+anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
+
+#base vs mol model (this only makes sense if these are nested models)
+anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
+
+```
+
+Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed models by strain.
+make artificial data / models, like human labels
+
+## Mixed Model for variance
+
+```{r HQ.var, echo=FALSE}
+#modeling within mouse variance
+
+DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
+
+#HQ var.CO
+HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.HQ.var_co)
+ranef(HetC.MixedModel.HQ.var_co)
+
+HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+summary(HQjust.strain.var_co)
+
+HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+summary(HQfull.Reduced.cv_co)
+
+#contrast models
+full2strain.CV <- anova(HetC.MixedModel.HQ.var_co, HQjust.strain.var_co)#
+
+full2noRandom.CV <- anova(HetC.MixedModel.HQ.var_co, HQfull.Reduced.cv_co)#
+
+strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
+
+```
+
+```{r Base.Variance.MM, echo=FALSE, eval=FALSE}
+
+#there's 1 mouse with just 1 cell remove
+DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
+
+DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
+
+#main var.CO
+HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.var_co)
+ranef(HetC.MixedModel.var_co)
+
+#main cV.CO
+HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+summary(HetC.MixedModel.cVco)
+ranef(HetC.MixedModel.cVco)
+
+just.strain.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain) 
+summary(just.strain.cV_co)
+
+full.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel)
+summary(full.Reduced.cv_co)
+
+#contrast models
+full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
+
+full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
+
+strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
+#for var, p = 0.0219
+
+```
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
+
+# check if the main mixed model results hold up when all cells are pooled
+
+#pooled.cells
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
+
+DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
+
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+
+```
+
+- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
+This means there is 0 variance added to mean_CO to mean due to being MSM.
+Does this mean the model is very good at predicting MSM mean_CO.
+The fixed strain - sex effects are for musc are large.
+This will likely change with MOLF (male molf are low, so smaller fixed effect and )
+
+- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
+In the first model MSM has largest random effect, since it's so far from WSB female.
+
+-adding MOLF also doesn't have significant factors
+
+changed all to unordered factors.
+all the variables are coded as ordered factors.
+When the variables are kept factors, and unordered the names appear! yay!
+
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
+
 #### outgroups
@@
 
-# Mixed Models
-
-todo: re-run the analysis with MOLF data
-make model with strain as fixed effect (since, there are interesting strain effects)
-
-The mixed model lets us separate effects to see how they effect a mouse mean CO count. This separation can also serve as proxies for divergence and polymorphism.
-
-Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
-
-```{r mixed.models.setup, echo=FALSE}
-#make are the DFs for the models
-# seperate the data, both sexes, only Mus
-# decide on Musc - Molossinus
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
-                            subsp.y,subsp.x,
-                    category.y, category.x,
-                        strain.x,strain.y, mouse.1  )  )
-
-#make dummy MOLF data, average of Musc females?
-#no MOLF male data?
-female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-table(female.musc.cells$quality)
-
-female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-                        #quality?
-)
-
-female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
-female.musc.mice.table$DOB <- as.Date("1/1/2020")
-
-female.musc.mice.table$strain <- "MOLF"
-female.musc.mice.table$subsp <- "Musc"
-female.musc.mice.table$species <- "M.musculus"
-
-female.musc.mice.table$female.stage <- "emb"
-female.musc.mice.table$age.days <-""
-female.musc.mice.table$age.weeks <- ""
-female.musc.mice.table$ID <- ""
-female.musc.mice.table$notes <- ""
-female.musc.mice.table$euth.date <- ""
-
-DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)
-
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
-
-#main df with variables coded as chr
-DF.HetC.MixedModel <- DF.HetC.MixedModel
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
-DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
-
-#mean_co.HQ  (this might remove some of the MOLF males)
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mean_co.HQ.mol
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
-        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
-        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
-
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
-```
-
-what are the table above this?
-
-For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)
-
-subspecies levels: 2, Musc and Domesticus
-strain levels: Dom: 3, Musc: 4
-sex: 2 female male
-
-`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
-
-List of mixed models with mean_co
-
-1. mean_co (chr coded)
-2. mean_co.HQ
-3. mean_co with Mol
-4. sub-sampling
-
-```{r HQ.MixedModel, echo=FALSE}
-table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-
-#mean_co.HQ  (this might remove some of the MOLF males)
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#these remove ALL male MSM!
-#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
-
-#mean.co.HQ
-HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.HQ)
-ranef(HetC.MixedModel.HQ)
-
-#reduced models
-Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-summary(Reduced.strain.HQ)
-ranef(Reduced.strain.HQ)
-
-Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
-summary(Reduced.all.HQ)
-#ranef(Reduced.all.HQ)
-
-#strain = fixed
-strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-summary(strain.fixed.full)
-#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
-#sexmale:strainWSB p=  0.006740 **
-#sexmale  p =  3.98e-05 ***
-#strainWSB   p=0.000441 ***
-#subspMusc:sexmale  p=  0.069493 . 
-#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are
-
-strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-summary(strain.fixed)
-
-#sexmale  p= 3.98e-05 ***
-#strainKAZ p=  0.052420 .  
-#strainWSB  p= 0.000441 ***
-#sexmale:strainMSM  p=3.42e-09 ***
-#sexmale:strainPWD  p= 5.79e-08 ***
-#sexmale:strainSKIVE p= 0.069493 .  
-#sexmale:strainWSB  p=  0.006740 **
-
-#maybe this WSB effec is just in the HQ data set?
-
-#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
-full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
-
-full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
-#this comparison not useful when there are different fixed effects
-
-full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ)# < .0001
-
-strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ)#NS
-```
-
-Reading these outputs is not the clearest -- try just printing out the coefficient table with p values.
-
-Mostly matches the base model results. the same fixed effects are significant in the reduced models. the random effects make sense
-
-random effects were tested with 
- #rand LRT
-this function might not support MM, (and needs linear models instead)
-
-also throws an error about only looking at 1 random effect
-
-(what does the significant random test mean) -- variance is not 0?
-
-```{r variance.boxplots, echo=FALSE, eval=FALSE}
-#how did I make this 
-#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
-#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
-#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
-#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
-
-#plot of variance
-var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
-var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
-
-#plot by sex / seperately
-# HQ data DF.HetC.MixedModel.HQ
-
-var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-var.male <- var.male   + facet_wrap( ~ strain, scales="free")
-
-cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
-
-```
-
-```{r BaseMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-#strain as fixed
-newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
-
-HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
-
-full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
-
-#CONTRAST models
-
-#base vs strain
-fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co)# <.0001
-
-#strain vs no random full.reduce
-fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
-
-strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)#<.0001
-
-#fullreduce vs strain
-
-#can't distinguish which random effects are more significant -- both tests are significant
-
-fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
-
-#random effects for base model
-ranef(HetC.MixedModel.mean_co)
-#random effect for just strain
-ranef(R.strain.HetC.MixedModel.mean_co)
-
-#reduced model's PWD radom effects is much smaller
-#the full model's largest random effects are for PWD and MSM (and MOLF, KAZ for negative)
-
-#fixed effects from full
-anova(HetC.MixedModel.mean_co)#NS
-
-#Check how the fixed effects change across these
-
-HetC.MixedModel.mean_co2 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel)
-#lmer(y~x1 + x2 + (1|g1), data=pbDat)
-
-lmer.MM <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel, REML = TRUE)
-
-```
-
-Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
-Models with random effects more significant than nulls (both just strain and no random effects).
-The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
-
-```{r Base.strain.fixed, echo=FALSE}
-
-#make mixed model with strain as fixed, since there are interesting results
-
-HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
-#this produces several significant factors
-
-HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
-summary(HetC.MM.strain.fixed2)
-#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
-
-#summary(HetC.MM.strain.fixed)
-#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
-#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
-#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
-#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
-#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
-#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
-#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
-#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
-
-#male effect lowers the mean from intercept
-#while strain effects 
-```
-
-```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
-#base model with molossinus level ... why am I making molossinus model?
-
-DF.HetC.MixedModel.mol <- DF.HetC.MixedModel
-str(DF.HetC.MixedModel.mol)#subsp is factor with 2 levels
-
-DF.HetC.MixedModel.mol$subsp <- as.character(DF.HetC.MixedModel.mol$subsp)
-
-#turn to chr first?
-#apparently ifelse doesn't retain class
-library(data.table)
-
-DF.HetC.MixedModel.mol$subsp <- if_else(grepl("MSM", DF.HetC.MixedModel.mol$strain), "Mol", 
-        if_else(grepl("MOLF", DF.HetC.MixedModel.mol$strain), "Mol", DF.HetC.MixedModel.mol$subsp))
-
-str(DF.HetC.MixedModel.mol)
-DF.HetC.MixedModel.mol$subsp <- factor(DF.HetC.MixedModel.mol$subsp, ordered = FALSE)
-
-#these remove ALL male MSM!
-table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
-
-#mean.co
-HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.mol)
-ranef(HetC.MixedModel.mol)
-
-#mreducing the model (reducing the random effects)
-Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
-summary(Reduced.strain.mol)
-ranef(Reduced.strain.mol)
-
-Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
-summary(Reduced.all.mol)
-
-#anova tests
-#full vs strain
-full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
-#full vs non
-full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
-#strain vs non
-strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
-
-anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
-
-#base vs mol model (this only makes sense if these are nested models)
-anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
-
-```
-
-Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed models by strain.
-make artificial data / models, like human labels
-
-## Mixed Model for the variance
-
-```{r HQ.var, echo=FALSE}
-#modeling within mouse variance
-
-DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
-
-#HQ var.CO
-HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.HQ.var_co)
-ranef(HetC.MixedModel.HQ.var_co)
-
-HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-
-summary(HQjust.strain.var_co)
-
-HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
-summary(HQfull.Reduced.cv_co)
-
-#contrast models
-full2strain.CV <- anova(HetC.MixedModel.HQ.var_co, HQjust.strain.var_co)#
-
-full2noRandom.CV <- anova(HetC.MixedModel.HQ.var_co, HQfull.Reduced.cv_co)#
-
-strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
-
-```
-
-```{r Base.Variance.MM, echo=FALSE, eval=FALSE}
-
-#there's 1 mouse with just 1 cell remove
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
-
-DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
-
-#main var.CO
-HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.var_co)
-ranef(HetC.MixedModel.var_co)
-
-#main cV.CO
-HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.cVco)
-ranef(HetC.MixedModel.cVco)
-
-just.strain.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain) 
-summary(just.strain.cV_co)
-
-full.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel)
-summary(full.Reduced.cv_co)
-
-#contrast models
-full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
-
-full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
-
-strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
-#for var, p = 0.0219
-
-```
-
-```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
-
-# check if the main mixed model results hold up when all cells are pooled
-
-#pooled.cells
-DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
-
-#remove PERC, MOLF, CZECH,
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
-
-#sex only has 2 levels and this breaks...
-model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.pooled)# indeces for this object: [1]-structure
-ranef(model.pooled)
-
-model.pooled$coefficients
-
-#pooled cells,  DF.HetC.MixedModel.pooled
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
-
-DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
-
-DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
-
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-
-```
-
-- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
-This means there is 0 variance added to mean_CO to mean due to being MSM.
-Does this mean the model is very good at predicting MSM mean_CO.
-The fixed strain - sex effects are for musc are large.
-This will likely change with MOLF (male molf are low, so smaller fixed effect and )
-
-- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
-In the first model MSM has largest random effect, since it's so far from WSB female.
-
--adding MOLF also doesn't have significant factors
-
-changed all to unordered factors.
-all the variables are coded as ordered factors.
-When the variables are kept factors, and unordered the names appear! yay!
-
-Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
-
 ### Another Heterochiasmy Plot

### 2019-12-19 doc/Main_Report_Results.rmd
updates and notes for the mixed model results

@@
 
+#setwd("~./MLH1repo/doc/")
 setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
@@
 
-load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
+load(file="~./data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
 #load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
@@
 
-#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
 
@@
     AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-
-```
-
-# Distributions of MLH1 counts across Samples
-
-The histograms below are for total / pooled cells (for now, maybe compare with HQ data to see what changes)
-
-```{r MLH1.histogram.2421, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
-
-#gg <- data.frame(weight=rep(x,nrow(df)),df)
-#gg$y <- with(gg,dnorm(x,mean,sd))
-#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
-
-mean.MLH1.obs <- ddply(.data=MLH1_data, 
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-histM <- ggplot(data=MLH1_data,(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")
-
-#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
-#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
-
-#ad
-#  geom_smooth(method = lm, se = FALSE
- # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
-
-histM
-
-#playing with expected posisson plot...
-#function = glm(y ~ x, family = poisson)
-#add Poission curve behind
-#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
-#not sure how to add Expected poisson curve
-
-```
-
-Most histograms of pooled MLH1 counts are all close to normal. Except PWD male, LEW female, SKIVE male, SPIC male.
-check if there is a quality effect.
-
-```{r MLH1.histogram.2, echo=FALSE}
-#remake the pooled MLH1 histgoram plot from above but facet by quality
-
-#PWD male, MSM male
-#LEW female, WSB female
-mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
-                 .(category, quality),
-                 summarize, 
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-#what's the point of the above table? -- add vlines?
-
-#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect
-
-histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
-histM_qual.KAZ_male
-
-histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
-histM_qual.PWD_male
-
-histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
-histM_qual.MSM_male
-
-histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
-histM_qual.WSB_female
-
-histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
-histM_qual.LEW_female
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
-#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
-
-#ad
-#  geom_smooth(method = lm, se = FALSE
- # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
 
@@
 
+*cast female data came from another publication
+
 # Mixed Models
@@
 
-$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)$$
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain* sex)$$
+$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)  + \varepsilon$$
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
 
@@
 ```{r mixed.models.setup, echo=FALSE, warning=FALSE}
-
-#why merged 
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+
+#notes for testing random effects
+
+#variance is still chr
+AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+
+#use the merged DF to filter by age 
 AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
@@
 AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-
 #main file, make everything unordered factors
 AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+
 DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
@@
 
-#make dummy MOLF data, average of Musc females?
-#no MOLF male data?
+#why did I male just a female table?
 female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-table(female.musc.cells$quality)
 
@@
 
-#female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
-#female.musc.mice.table$DOB <- as.Date("1/1/2020")
-
-#female.musc.mice.table$strain <- "MOLF"
-#female.musc.mice.table$subsp <- "Musc"
-#female.musc.mice.table$species <- "M.musculus"
-
-#female.musc.mice.table$female.stage <- "emb"
-#female.musc.mice.table$age.days <-""
-#female.musc.mice.table$age.weeks <- ""
-#female.musc.mice.table$ID <- ""
-#female.musc.mice.table$notes <- ""
-#female.musc.mice.table$euth.date <- ""
-
-#DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)
-
+#REMOVE EXTRA STRAINS
 DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
@@
 DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+
+#remove blank rows (which are these?)
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
+#151 mice, no rows with nas for values
 
@@
 DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
+```
+
+```{r HQ.dataset, echo=FALSE}
+
+#also make a HQ dataset, tighter window for ages,
 
@@
 DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
+```
+
+```{r Mol.dataset, echo=FALSE}
 #mean_co.HQ.mol
@@
 
-For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus.?)
-
-subspecies levels: 2, Musc and Domesticus
-
-strain levels: Dom: 3, Musc: 4
+Seting up the data set for using the mixed model (base data set, HQ dataset)
+
+- only Mus musculus strains with sex matched observations
+
+- Unorder factors
+
+- male ages within range
+
+Currently I ran through 3 datasets: 1) base, 2) HQ and 3) Molf (from HQ)
+
+`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
+
+Above are the total number of mice in the data set
+
+For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.
+
+subspecies levels: 2(3), Musc and Domesticus (Molossinus)
+
+strain levels: Dom: 3, Musc: 4  (Mol: 2)
 
@@
 
-`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
-
-Above are the total number of mice in the data set
-
-These are the List of mixed models I built
-
-1. mean_co (chr coded, instead of factors)
-
-2. mean_co.HQ
-
-3. mean_co with Mol (Mol coded as a distinct subsp)
-
-4. sub-sampling (not sure if I made this one)
-
-```{r HQ.MixedModel.1, echo=FALSE}
+## Mean CO Count
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
+
+```{r HQ.MixedModel.1, echo=FALSE, digits = 2}
 #table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-
 #mean_co.HQ  (this might remove some of the MOLF males)
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mean.co.HQ
+
 HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
 
-#summary(HetC.MixedModel.HQ)
+as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )
+
 #ranef(HetC.MixedModel.HQ)
-```
+#which is the test for random effects
+
+```
+
+Above is the table of coefficients for MM.1, HQ dataset, lme(), random effects strain and strain specific sex effect.
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
+
+```{r MM.1.Mol, echo=FALSE}
+
+HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(   summary(HetC.MixedModel.HQ.mol)$tTable, 6) )
+
+```
+
+Above is the table of coefficients for MM.1.mol, molssinus coded as seperate subspecies, lme(), random effects strain and strain specific sex effect.
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
 
@@
 #summary(Reduced.strain.HQ)
+
+as.data.frame(round(summary(Reduced.strain.HQ)$tTable, 6))
+#add a aption to this code chunk
 #ranef(Reduced.strain.HQ)
-```
+
+```
+
+Above is the table of coefficients for MM.2, HQ dataset, lme(), random effects coded without sex specific effect.
+This is model has sex and interaction coefficient is significant.
+
+> MM.2.Mol
+
+```{r MixedModel.2.mol, echo=FALSE}
+#reduced models
+Reduced.strain.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random = ~1|strain)
+#summary(Reduced.strain.HQ)
+
+as.data.frame(round(summary(Reduced.strain.HQ.mol)$tTable, 6))
+#add a aption to this code chunk
+#ranef(Reduced.strain.HQ.mol)
+
+```
+
+The model for MM.2.Mol has similar results as the above model, with just 2 subsps
+This model has significant coefficients for sex and the 2 interaction effects (sex:musc, sex:mol)
+
+> MM2.2
+
+$$mouse .mean CO ~=~ subsp * sex  + \varepsilon$$
 
 ```{r MixedModel.2.2, echo=FALSE}
-
 Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
+
+as.data.frame( round(summary(Reduced.all.HQ)$tTable, 6))
 #summary(Reduced.all.HQ)
@@
 ```
+
+Above is the table of coefficients for MM.2.2, HQ dataset, gls(), No random effects.  
+I think this gls has sex and interaction term as significant.
+
+> MM3
+
+$$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
 
@@
 strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+#library(xtable)
+
+#this table of results is broken -- might be too long -- lm
+#lm has different names
+as.data.frame(round( summary(strain.fixed.full)$coefficients, 6))
+
 #summary(strain.fixed.full)
@@
 
+Above is the table of coefficients for MM.3, HQ dataset, lm function, sex, strain and subsp are interacting fixed effects.
+
+Strain G is significant (p=0.0005),
+
+interaction terms of sex and strain for (G, MSM, PWD).
+
+How are these models different from the last one -- MM4 is missing subsp, so only sex * strain effects.
+
+> MM.4
+
+$$mouse .mean CO ~=~ sex * strain + \varepsilon$$
+
 ```{r MixedModel.4, echo=FALSE}
@@
 strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+#below table is broken
+as.data.frame( round( summary(strain.fixed)$coefficients, 6))
+
 #summary(strain.fixed)
-
 #maybe this WSB effec is just in the HQ data set?
+```
+
+Above is the table of coefficients for **MM.4**, HQ dataset, lm function, just sex and strain, no subspecies.
+
+Significant factors are 
+strainG (p=0.0005), strainMSM and strainPWD (very slight),
+sex:G (p=0.001),
+sex:MSM, Sex:PWD (very slight p=.05, p=0.3),
+
+#### ANOVA, mean CO count full vs reduced
+
+Use Anova to compare full vs reduced models
+the reduced models will be those missing either, strain, sex, or subsp.
+
+```{r anova.full.v.reduced, echo=FALSE, eval=FALSE}
+
+#strain as fixed
+#CONTRAST models
+#newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
+
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
+
+full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
+
+#base vs strain
+fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co) #<.0001
+
+#show scientific vs normal
+#vv <- fullVstrain$`p-value`
+#round(vv[2], 8)
+#format(vv[2], scientific=F)
+
+#strain vs no random full.reduce
+fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+#fullreduce vs strain
+#can't distinguish which random effects are more significant -- both tests are significant
+
+fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
+
 #anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
@@
 
-full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ)# < .0001
-
-strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ)#NS
-```
-
-```{r try.blah, echo=FALSE}
-#library(jtools)
-#states <- as.data.frame(state.x77)
-#fit <- lm(Income ~ Frost + Illiteracy + Murder, data = states)
-#summ(fit)
-#this is a nice display package for lm
-
-res.table <- as.data.frame(coef(summary(Reduced.all.HQ)))
-res.table
-```
-
-```{r variance.boxplots, echo=FALSE, eval=FALSE}
-#how did I make this 
-#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
-#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
-#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
-#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
-
-#plot of variance
-var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
-var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
-
-#plot by sex / seperately
-# HQ data DF.HetC.MixedModel.HQ
-
-var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-var.male <- var.male   + facet_wrap( ~ strain, scales="free")
-
-cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
-
-```
-
-```{r BaseMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-#strain as fixed
-newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
-
-HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
-
-full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
-
-#CONTRAST models
-
-#base vs strain
-fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co)# <.0001
-
-#strain vs no random full.reduce
-fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
-
-strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)#<.0001
-
-#fullreduce vs strain
-
-#can't distinguish which random effects are more significant -- both tests are significant
-
-fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
-
-#random effects for base model
-ranef(HetC.MixedModel.mean_co)
-#random effect for just strain
-ranef(R.strain.HetC.MixedModel.mean_co)
-
-#reduced model's PWD radom effects is much smaller
-#the full model's largest random effects are for PWD and MSM (and MOLF, KAZ for negative)
-
-#fixed effects from full
-anova(HetC.MixedModel.mean_co)#NS
-
-#Check how the fixed effects change across these
-
-HetC.MixedModel.mean_co2 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel)
-#lmer(y~x1 + x2 + (1|g1), data=pbDat)
-
-lmer.MM <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel, REML = TRUE)
-
+full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ) # < .0001
+
+strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ) #NS
+```
+
+```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
+#base model with molossinus level ... why am I making molossinus model?
+#mean.co
+HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
+
+#mreducing the model (reducing the random effects)
+Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
+
+Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
+
+#anova tests
+#full vs strain
+full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
+#full vs non
+full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
+#strain vs non
+strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
+
+anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
+
+#base vs mol model (this only makes sense if these are nested models)
+anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
 ```
@@
 HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
-summary(HetC.MM.strain.fixed2)
+
+#summary(HetC.MM.strain.fixed2)
 #(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
@@
 
-```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
-#base model with molossinus level ... why am I making molossinus model?
-
-DF.HetC.MixedModel.mol <- DF.HetC.MixedModel
-str(DF.HetC.MixedModel.mol)#subsp is factor with 2 levels
-
-DF.HetC.MixedModel.mol$subsp <- as.character(DF.HetC.MixedModel.mol$subsp)
-
-#turn to chr first?
-#apparently ifelse doesn't retain class
-library(data.table)
-
-DF.HetC.MixedModel.mol$subsp <- if_else(grepl("MSM", DF.HetC.MixedModel.mol$strain), "Mol", 
-        if_else(grepl("MOLF", DF.HetC.MixedModel.mol$strain), "Mol", DF.HetC.MixedModel.mol$subsp))
-
-str(DF.HetC.MixedModel.mol)
-DF.HetC.MixedModel.mol$subsp <- factor(DF.HetC.MixedModel.mol$subsp, ordered = FALSE)
-
-#these remove ALL male MSM!
-table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
-
-#mean.co
-HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.mol)
-ranef(HetC.MixedModel.mol)
-
-#mreducing the model (reducing the random effects)
-Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
-summary(Reduced.strain.mol)
-ranef(Reduced.strain.mol)
-
-Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
-summary(Reduced.all.mol)
-
-#anova tests
-#full vs strain
-full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
-#full vs non
-full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
-#strain vs non
-strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
-
-anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
-
-#base vs mol model (this only makes sense if these are nested models)
-anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
-
-```
-
-Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed models by strain.
-make artificial data / models, like human labels
-
-## Mixed Model for variance
-
-```{r HQ.var, echo=FALSE}
+#### Summary of mean COs
+
+MM.2 has significant coefficients, sex and sex*subsp interaction.
+
+MM.2.Mol, the sex and interaction term (subsp*sex) are significant.
+
+MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.
+
+# Variance in CO count
+
+The **biological significance** of within mouse variance in CO per cell is...
+
+I can show the variance boxplots?  Not sure if it's very helpful.
+Use the same Datasets HQ and mol for the mean CO counts. Use both var and cV as dependant variables
+
+DF.HetC.MixedModel.HQ.mol
+DF.HetC.MixedModel.HQ
+
+```{r variance.boxplots, echo=FALSE, eval=FALSE}
+#how did I make this 
+#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
+#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
+#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
+#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
+
+#plot of variance
+var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
+#this plot shows variance as points/single values
+
+#plot by sex / seperately
+# HQ data DF.HetC.MixedModel.HQ
+
+var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+var.male <- var.male   + facet_wrap( ~ strain, scales="free")
+
+cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male cV levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
+
+```
+
+$$mouse .cV CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$
+
+```{r cV.model.setup, echo=FALSE}
 #modeling within mouse variance
@@
 
-summary(HetC.MixedModel.HQ.var_co)
-ranef(HetC.MixedModel.HQ.var_co)
-
+#summary(HetC.MixedModel.HQ.var_co)
+
+as.data.frame( round(summary(HetC.MixedModel.HQ.var_co)$tTable, 6))
+
+#ranef(HetC.MixedModel.HQ.var_co)
+```
+
+lme function, Cv modeled. Only sex is significant.
+
+```{r cv.mod1.mol, echo=FALSE}
+
+HetC.MixedModel.HQ_mol.cV_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(HetC.MixedModel.HQ_mol.cV_co)$tTable, 6))
+
+```
+
+$$mouse .cV CO ~=~ subsp * sex + rand(1|strain)+ \varepsilon$$
+
+```{r cV.models, echo=FALSE}
 HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
 
-summary(HQjust.strain.var_co)
-
+as.data.frame( round(summary(Reduced.all.HQ)$tTable, 6))
+
+#summary(HQjust.strain.var_co)
+```
+
+Sex and sex*subsp interaction effect are significant.
+
+```{r cV.mod.mol, echo=FALSE}
+
+HQ.mol.cV_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random = ~1|strain)
+
+as.data.frame( round(summary(HQ.mol.cV_co)$tTable, 6))
+
+```
+
+Sex is significant, when mol is a seperate subspecies interaction is no longer significant.
+
+$$mouse .cV CO ~=~ subsp * sex + \varepsilon$$
+```{r more.cV.CO, echo=FALSE}
 HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
-summary(HQfull.Reduced.cv_co)
+as.data.frame( round(summary(HQfull.Reduced.cv_co)$tTable, 6))
+#summary(HQfull.Reduced.cv_co)
 
@@
 strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
-
-```
+```
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain*sex) + \varepsilon$$
 
@@
 DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
-
 DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
@@
 #main var.CO
-HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-summary(HetC.MixedModel.var_co)
-ranef(HetC.MixedModel.var_co)
-
+HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(HetC.MixedModel.var_co)$tTable, 6))
+
+#summary(HetC.MixedModel.var_co)
+#ranef(HetC.MixedModel.var_co)
+```
+Above error says missing values in object -- I don't understand. Mol version of data frame works...
+
+Base DF, lme function, random sex specific strain effects
+
+$$mouse .var CO ~=~ subsp * sex  + \varepsilon$$
+```{r base.var.mol, echo=FALSE}
+HetC.MixedModel.mol.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(HetC.MixedModel.mol.var_co)$tTable, 6))
+
+```
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain * sex) + \varepsilon$$
+```{r var.model.1, echo=FALSE, eval=FALSE}
 #main cV.CO
@@
 
-summary(HetC.MixedModel.cVco)
-ranef(HetC.MixedModel.cVco)
-
+as.data.frame( round(summary(HetC.MixedModel.varco)$tTable, 6))
+
+#summary(HetC.MixedModel.cVco)
+#ranef(HetC.MixedModel.cVco)
+```
+
+$$mouse .var CO ~=~ subsp * sex + rand(1|strain) + \varepsilon$$
+
+```{r var.model.2, echo=FALSE, eval=FALSE}
 just.strain.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)
-summary(just.strain.cV_co)
+
+as.data.frame( round(summary(just.strain.var_co)$tTable, 6))
+
+#summary(just.strain.cV_co)
+```
+
+lme model above
+
+$$mouse .var CO ~=~ subsp * sex + \varepsilon$$
+
+```{r var.mod.3, echo=FALSE}
 
 full.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel)
-summary(full.Reduced.cv_co)
-
+#summary(full.Reduced.cv_co)
+
+as.data.frame( (summary(full.Reduced.cv_co)$tTable))
+
+```
+
+gls model model above
+
+```{r var.contrast, echo=FALSE, eval=FALSE}
 #contrast models
@@
 #for var, p = 0.0219
-
-```
+```
+
+#### Summary variance
+
+The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect pvalues were larger and only significant for one of the models.
+
+When Molossinus is coded as a seperate subspecies, then interaction terms are no longer significant.
+
+#### Pooled cells
+
+For later analysis.
 

### 2019-12-19 doc/figureFiles/figures.Rmd
small changes to the files

@@
 
-## Option 2, strain averages, more strains, pooled cells
+## Option 2, mouse averages, more strains
 
@@
 #MSM, MOLF
-strain.table.HQ.backup$strain<- factor(strain.table.HQ.backup$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
+strain.table.from.mouse.HQ.backup$strain<- factor(strain.table.from.mouse.HQ.backup$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
                            "PWD", "SKIVE", "KAZ", "TOM", "AST","CZECH",
@@
 
-Fig1_B_strain.male <- ggplot(strain.table.HQ.backup[strain.table.HQ.backup$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
-  ylab("")+
-  geom_point(size=3)+
-  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
+#REMOVING FACET ~SEX 
+Fig1_B_strain <- ggplot(strain.table.from.mouse.HQ.backup, aes(y=mean_co.of.mice, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2), ymax=mean_co.of.mice+(se.of.mice*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
   ylim(c(18, 35))+
@@
 ```{r Strain.fig.merge, echo=FALSE}
-fig1_B_v1  <- ggdraw() +
-  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.45, height=.97) +
-  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.45, height=.97) +
+#fig1_B_v1  <- ggdraw() +
+#  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.45, height=.97) +
+#  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.45, height=.97) +
 
-  draw_plot_label(label = c("A", "B"), size = 15,
-                  x = c(0, .5), y = c(1, 1) )
-fig1_B_v1
+#  draw_plot_label(label = c("A", "B"), size = 15,
+#                  x = c(0, .5), y = c(1, 1) )
+
+Fig1_B_strain
+
 ```

### 2019-12-19 doc/Main_Report.rmd
moved some code chunks from main report to this backup

@@
 
+# Distributions of MLH1 counts across Samples
+
+The histograms below are for total / pooled cells (for now, maybe compare with HQ data to see what changes)
+
+```{r MLH1.histogram.2421, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}
+
+#gg <- data.frame(weight=rep(x,nrow(df)),df)
+#gg$y <- with(gg,dnorm(x,mean,sd))
+#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
+
+mean.MLH1.obs <- ddply(.data=MLH1_data, 
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+histM <- ggplot(data=MLH1_data,(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")
+
+#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
+#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
+
+#ad
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+
+histM
+
+#playing with expected posisson plot...
+#function = glm(y ~ x, family = poisson)
+#add Poission curve behind
+#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
+#not sure how to add Expected poisson curve
+
+```
+
+Most histograms of pooled MLH1 counts are all close to normal. Except PWD male, LEW female, SKIVE male, SPIC male.
+check if there is a quality effect.
+
+```{r MLH1.histogram.2, echo=FALSE}
+#remake the pooled MLH1 histgoram plot from above but facet by quality
+
+#PWD male, MSM male
+#LEW female, WSB female
+mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
+                 .(category, quality),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+#what's the point of the above table? -- add vlines?
+
+#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect
+
+histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
+histM_qual.KAZ_male
+
+histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
+histM_qual.PWD_male
+
+histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
+histM_qual.MSM_male
+
+histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
+histM_qual.WSB_female
+
+histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
+histM_qual.LEW_female
+
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
+#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
+
+#ad
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+
+```
+
 # Initial Results

### 2019-12-20 doc/Main_Report_Results.rmd
re-structing results file to match outline

@@
 #setwd("~./MLH1repo/doc/")
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
 #setwd("C:/Users/April/Desktop/MLH1repo")
@@
 
-load(file="~./data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
 #load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
@@
 
-write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
+#there might be an error for changing wd
+#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
 
@@
 
-# MLH1 means
+# Main Figures
 
@@
 
-# Mixed Models
+add code for the table table. may have to deal with the excess mouse issue
+
+# A. Models for mean and variance of MLH1 foci per cell
 
@@
 
+FOR FINAL ANALYSIS, MOL Will be coded as a seperate subsp (change all of the main model reports to)
 For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.
@@
 
-# Variance in CO count
+## Variance in CO count
 
@@
 
-```{r var.mod.3, echo=FALSE}
+```{r var.mod.3, echo=FALSE, eval=FALSE}
 
@@
 
-# DMC1 Results
+# Rapid male specific evolution
+
+## Proportion of chrm classes
+
+These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes.
+Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs. (the gwRR for most mice is low/close to the minimum)
+
+This is also another way to score heterochiasmy (that gets around the sex chrms, i think)
+
+The data below are from hand measured cells (not from bivalent data)
+generate smaller data set of chrm proportions (0CO, 1CO,2CO etc)
+10 cells from each musc category. **Add Molf male**
+
+ToDo -- add more hand curated data for the bivalent proportions
+
+```{r chrm_prop_data, echo=FALSE, eval=FALSE}
+#shut off eval for now because of wd / path error
+#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
+chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
+chrm.class.DF <- add_mouse(chrm.class.DF)
+chrm.class.DF <- add_strain(chrm.class.DF)
+chrm.class.DF <- add_sex(chrm.class.DF)
+chrm.class.DF <- add_category(chrm.class.DF)
+#chrm.class.DF <- add_euth_date(meta.data.file2)
+#chrm.class.DF <- add_age(meta.data.file2)
+
+#table(chrm.class.DF$category)
+#str(chrm.class.DF)
+
+#summarize by category
+proportion_stats_category  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = sum(biv.sum.check),
+
+                 n0CO = sum(n0CO),
+                  n1CO = sum(n1CO),
+                  n2CO = sum(n2CO),
+                  n3CO = sum(n3CO),
+                  n4CO = sum(n4CO),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv,
+                 per4CO= sum(n4CO)/ tot_biv,
+
+                 mean.MLH1.foci = mean(MLH1.sum.check),
+                 mean.MLH1.foci2 = mean(nMLH1.foci),
+                 meanQ = mean(quality)
+
+)
+#str(this_thing2)
+
+#this first one is the 'first' MLH1 count
+oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
+#summary(oopp2)
+#none of the classes are significant
+
+#mean is the new calculated-- mean 0CO is significant and intercept
+oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
+#summary(oopp)
+#not sure if these are inter-correlated
+#I don't know why 0COs and the intercept are significant..
+#0CO and 1CO are significant when only chrm classes are included (and intercept)
+
+#no longer significant when used
+#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
+#after added more MSM data, tot_biv is not significant
+#0CO is now more significant, and 2CO is sightly significant
+
+#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
+
+#MSM needs more bialent observations
+#use melt to put all propotions 
+#display / visualize / compare proportions
+#the table needs to be melted, because it's split by category
+
+#melt.table = read.csv("~./MLH1repo/data/ChrmClassProportions_melt.csv", header = TRUE)
+melt.table$fileName <- (melt.table$Original.Name)
+
+#melt.table <- add_mouse(melt.table)
+#melt.table <- add_strain(melt.table)
+#melt.table <- add_sex(melt.table)
+#melt.table <- add_category(melt.table)
+
+melt.table$strain<- factor(melt.table$strain,levels =c("LEW",
+                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM"), order=T )
+
+melt.table <- with(melt.table, melt.table[order(strain),])
+
+other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))
+
+other_thing
+
+chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
+
+#chrm.class.DF$biv.sum.check
+```
+
+High musc males have half 1CO and 2CO bivalents.
+For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
+
+```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}
+
+#bivalent data might be too big..
+#load(file="~./MLH1repo/data/CleanBivData.RData")
+#Curates_BivData -- hand.foci.count!
+
+#For each mouse I need to count up
+#the number of bivalent classes are not counted up
+
+chrm.propotion.category <- ddply(.data=Curated_BivData, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(fileName)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = length(Obj.ID),
+
+                 n0CO = length(which(hand.foci.count == 0)),
+                  n1CO = length(which(hand.foci.count == 1)),
+                  n2CO = length(which(hand.foci.count == 2)),
+                  n3CO = length(which(hand.foci.count == 3)),
+                  #n4CO = sum(n4CO),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv )
+                # per4CO= sum(n4CO)/ tot_biv  )
+
+Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
+  "per0CO", "per1CO", "per2CO", "per3CO"), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
+
+# I think I need to make 2 melt tables -- for count and proportions
+Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
+
+Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
+colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
+
+Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
+
+Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
+
+Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
+ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
+      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
+        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
+        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
+        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
+        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
+
+Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
+                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
+#no mol yet
+
+Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
+
+#need to make list of positions
+
+chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
+
+chrm_prop.BivData_plot
+
+```
+
+number above bars are the raw counts for bivalent classes.
+
+Potential figure 2 (eventually). Try to make the figure taller.
+-- add more data to make sample sizes more equal.
+-- add space to indicate where female data is missing
+
+SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
+What's up with Lew females? there's lots of 2COs, also WSB female.
+
+This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
+If these data are random samples of bivalents, the X
+
+ToDo: chisq tests for difference in proportions, 
+- within strain across sex,
+- males, across high and low groups
+- look at mouse level proportions
+
+## DMC1 Results
 
@@
 
+## Total SC Results
+
+# Female patterns
+
+there is strain specific evolution for females ...
+
+# Sexual dimorphism and heterochiasmy things
+
+total sc length / SC compaction and telomere bias in positioning are conserved / consistantly sexually dimorphism (in these mice?)  for most species
+
+# Predictions / tests for selection on Bivalent structure
+
+# ""
+
 ## Testing Normality Assumptions
@@
 
-# Proportion of chrm classes
-
-These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes.
-Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs. (the gwRR for most mice is low/close to the minimum)
-
-This is also another way to score heterochiasmy (that gets around the sex chrms, i think)
-
-The data below are from hand measured cells (not from bivalent data)
-generate smaller data set of chrm proportions (0CO, 1CO,2CO etc)
-10 cells from each musc category. **Add Molf male**
-
-ToDo -- add more hand curated data for the bivalent proportions
-
-```{r chrm_prop_data, echo=FALSE}
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check),
-
-                 n0CO = sum(n0CO),
-                  n1CO = sum(n1CO),
-                  n2CO = sum(n2CO),
-                  n3CO = sum(n3CO),
-                  n4CO = sum(n4CO),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#str(this_thing2)
-
-#this first one is the 'first' MLH1 count
-oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
-#summary(oopp2)
-#none of the classes are significant
-
-#mean is the new calculated-- mean 0CO is significant and intercept
-oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
-#summary(oopp)
-#not sure if these are inter-correlated
-#I don't know why 0COs and the intercept are significant..
-#0CO and 1CO are significant when only chrm classes are included (and intercept)
-
-#no longer significant when used
-#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
-#after added more MSM data, tot_biv is not significant
-#0CO is now more significant, and 2CO is sightly significant
-
-#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
-
-#MSM needs more bialent observations
-#use melt to put all propotions 
-#display / visualize / compare proportions
-#the table needs to be melted, because it's split by category
-
-melt.table = read.csv("~./MLH1repo/data/ChrmClassProportions_melt.csv", header = TRUE)
-melt.table$fileName <- (melt.table$Original.Name)
-
-#melt.table <- add_mouse(melt.table)
-#melt.table <- add_strain(melt.table)
-#melt.table <- add_sex(melt.table)
-#melt.table <- add_category(melt.table)
-
-melt.table$strain<- factor(melt.table$strain,levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM"), order=T )
-
-melt.table <- with(melt.table, melt.table[order(strain),])
-
-other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))
-
-other_thing
-
-chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
-
-#chrm.class.DF$biv.sum.check
-```
-
-High musc males have half 1CO and 2CO bivalents.
-For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
-
-```{r chrm.pro.Curated.BivData, echo=FALSE}
-
-#bivalent data might be too big..
-load(file="~./MLH1repo/data/CleanBivData.RData")
-#Curates_BivData -- hand.foci.count!
-
-#For each mouse I need to count up
-#the number of bivalent classes are not counted up
-
-chrm.propotion.category <- ddply(.data=Curated_BivData, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(fileName)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = length(Obj.ID),
-
-                 n0CO = length(which(hand.foci.count == 0)),
-                  n1CO = length(which(hand.foci.count == 1)),
-                  n2CO = length(which(hand.foci.count == 2)),
-                  n3CO = length(which(hand.foci.count == 3)),
-                  #n4CO = sum(n4CO),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv )
-                # per4CO= sum(n4CO)/ tot_biv  )
-
-Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
-  "per0CO", "per1CO", "per2CO", "per3CO"), 
-  id=c("category", "tot_biv") )
-colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
-
-# I think I need to make 2 melt tables -- for count and proportions
-Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
-  id=c("category", "tot_biv") )
-colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
-
-Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
-colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
-
-Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
-
-Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
-
-Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
-ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
-      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
-        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
-        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
-        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
-        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
-
-Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
-                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
-#no mol yet
-
-Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
-
-#need to make list of positions
-
-chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
-
-chrm_prop.BivData_plot
-
-```
-
-number above bars are the raw counts for bivalent classes.
-
-Potential figure 2 (eventually). Try to make the figure taller.
--- add more data to make sample sizes more equal.
--- add space to indicate where female data is missing
-
-SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
-What's up with Lew females? there's lots of 2COs, also WSB female.
-
-This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
-If these data are random samples of bivalents, the X
-
-ToDo: chisq tests for difference in proportions, 
-- within strain across sex,
-- males, across high and low groups
-- look at mouse level proportions
-
 ### Another Heterochiasmy Plot

### 2019-12-20 doc/Main_Report_Results.rmd
working on the chrm class proportions sections, other small edits

@@
 
+load in more Data, chrm proportions, total SC,
+write / standardize the data set up for those
+
 Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
@@
 Details of dataset cleaning ect is in SetupData.rmd
-
 Effects from different factors are reviewed in Effects.Rmd
@@
 generate smaller data set of chrm proportions (0CO, 1CO,2CO etc)
-10 cells from each musc category. **Add Molf male**
-
-ToDo -- add more hand curated data for the bivalent proportions
-
-```{r chrm_prop_data, echo=FALSE, eval=FALSE}
+10 cells from each musc category. **Add Molf female**
+
+```{r chrm_prop_data, echo=FALSE}
 #shut off eval for now because of wd / path error
 #chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
+chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
 
@@
 #summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
+proportion_stats_category  <- ddply(.data=chrm.class.DF, 
                  .(category),
@@
                  nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check),
-
-                 n0CO = sum(n0CO),
-                  n1CO = sum(n1CO),
-                  n2CO = sum(n2CO),
-                  n3CO = sum(n3CO),
-                  n4CO = sum(n4CO),
+                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
+
+                 n0CO = sum(n0CO, na.rm = TRUE),
+                  n1CO = sum(n1CO,na.rm = TRUE),
+                  n2CO = sum(n2CO,na.rm = TRUE),
+                  n3CO = sum(n3CO,na.rm = TRUE),
+                  n4CO = sum(n4CO,na.rm = TRUE),
 
@@
 )
-#str(this_thing2)
+#this is the table that needs melting
+#but is also easy to redo by hand
+
+#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
+```
+
+```{r plot.chrm.prop.melt, echo=FALSE}
+
+#add subspecies to these values
+melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
+
+#add the subsps
+
+melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
+                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
+
+melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
+
+chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))
+
+#need a second DF / the first df
+#+ annotate() text 
+#annotate()text
+
+#proportion_stats_category,  tot_biv, ncells
+
+#add number of cells and bivalents in the corner
+
+chrm.class.barplot
+```
+
+```{r test.Chrm.Prop, echo=FALSE}
 
 #this first one is the 'first' MLH1 count
-oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
+#which is this variable
+
+#oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
 #summary(oopp2)
@@
 #mean is the new calculated-- mean 0CO is significant and intercept
-oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
+#oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = #proportion_stats_category)
+
 #summary(oopp)
@@
 
-#melt.table = read.csv("~./MLH1repo/data/ChrmClassProportions_melt.csv", header = TRUE)
-melt.table$fileName <- (melt.table$Original.Name)
-
-#melt.table <- add_mouse(melt.table)
-#melt.table <- add_strain(melt.table)
-#melt.table <- add_sex(melt.table)
-#melt.table <- add_category(melt.table)
-
-melt.table$strain<- factor(melt.table$strain,levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM"), order=T )
-
-melt.table <- with(melt.table, melt.table[order(strain),])
-
-other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))
-
-other_thing
+#Need to run the chrm proportions thru melt code.
 
@@
 
+Add p.values of tests for the how the proportions differ.
+
 ```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}
+
+#this is a second mode of
 
@@
 
+make the heterochiasmy plot, with female strain averages adjusted for X
+(measure variance in CO counts for the top 5 female whole cell bivalents)
+(variance in )
+how does this compare to variance of mid 5, bottom 5?
+
 # Predictions / tests for selection on Bivalent structure

### 2019-12-20 src/SetupData.Rmd
working on the chrm class proportions sections, other small edits

@@
 ```{r save.RData, echo=FALSE}
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
+
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")
+
 #  OutPut: big large MLH1_data (AP's) df, big DF of BD with just the mice I want.

### 2019-12-23 doc/figureFiles/figures.Rmd
added more tick marks, changed axis title

@@
   #facet_grid(~sex, scales = "free", space="free")+ 
-  ylab("Average MLH1 Foci per Cell")+
+  ylab("Number of Crossovers per Cell")+
   geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
@@
  scale_color_manual(values=colors_of_strains)+ 
+  scale_y_continuous(breaks=seq(20,32,2) )+
+
   #legend.position="none",

### 2019-12-23 doc/Main_Report_Results.rmd
knit new version, did small tidying up

@@
 
-#### outgroups
-
-The outgroups (SPIC and SPRET) don't show strong evidence of evolution in gwRR.
-
-#### Male Patterns
-
-Rapid evolution in musculus and molossinus; high and low CO count males (two instances of rapid gwRR evo, PWD and MSM)
-
-#### Female Patterns
-
-seems like female specific evolution in molossinus and musculus.
-more variation in domesticus females.
-Females have more nuanced patterns, there is strain specific differences.
-
 # Rapid male specific evolution
+
+list out the points whic hsuport rapid male specific evolution for MLH1 rates.
 
@@
 
-## DMC1 Results
+## DMC1 Results Spermatocytes
 
@@
 ```
+
+### Testing Normality Assumptions
+
+```{r normality, echo=FALSE, warning=FALSE}
+#histograms for each
+
+#- check that distributions are close to normal
+#- check for equal variance.
+
+DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
+
+DMC1.histogram
+#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
+#bartlet's test, Levine's test
+```
+
+The counts look close to normal.
 
@@
 
-## Total SC Results
-
-# Female patterns
-
-there is strain specific evolution for females ...
-
-# Sexual dimorphism and heterochiasmy things
-
-total sc length / SC compaction and telomere bias in positioning are conserved / consistantly sexually dimorphism (in these mice?)  for most species
-
-make the heterochiasmy plot, with female strain averages adjusted for X
-(measure variance in CO counts for the top 5 female whole cell bivalents)
-(variance in )
-how does this compare to variance of mid 5, bottom 5?
-
-# Predictions / tests for selection on Bivalent structure
-
-# ""
-
-## Testing Normality Assumptions
-
-```{r normality, echo=FALSE, warning=FALSE}
-#histograms for each
-
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-DMC1.histogram
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-```
-
-The counts look close to normal.
-
-```{r barlett.test, echo=FALSE}
-
-#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
-bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
-
-```
-
-Evidence for non-equal variance across strains for zygotene cells.
-
-Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
-
-From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova <- anova(mod.L) #0.0002701 ***
-z.anova <- anova(mod.Z) #NS
-
-l.anova
-z.anova
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#L vs Z (all species)
-#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
-#p-value = 1.032e-05
-#not a valid interpertation
-
-#high vs low group (L)
-ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
-
-## Permutations with equal sampling
+## DMC1 Permutations with equal sampling
 
@@
 The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+
+## Total SC Results
+
+insert male total sc results here
+
+# Female patterns
+
+Nuanced
+there is strain specific evolution for females ...
+
+# Sexual dimorphism and heterochiasmy things
+
+total sc length / SC compaction and telomere bias in positioning are conserved / consistantly sexually dimorphism (in these mice?)  for most species
+
+make the heterochiasmy plot, with female strain averages adjusted for X
+(measure variance in CO counts for the top 5 female whole cell bivalents)
+(variance in )
+how does this compare to variance of mid 5, bottom 5?
+
+# Predictions / tests for selection on Bivalent structure
+
+# ""
+
+```{r barlett.test, echo=FALSE}
+
+#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
+bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
+
+```
+
+Evidence for non-equal variance across strains for zygotene cells.
+
+Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
+
+From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#L vs Z (all species)
+#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
+#p-value = 1.032e-05
+#not a valid interpertation
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
 
@@
 
-### Another Heterochiasmy Plot
-
-```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average"}
+```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
 

### 2019-12-27 doc/Main_Report_Results.rmd
adding thoughts for variance section

@@
 
-The **biological significance** of within mouse variance in CO per cell is...
+in addition to the mean gwRR, the variance is also an important metric to dicribe
+
+Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
 
@@
 DF.HetC.MixedModel.HQ
+
+RW's main qtl paper used the within animal variance, 
+"We mapped QTL for
+MLH1 count variance within individuals, applying an approach
+to deconvolve the mean-variance relationship (see
+Materials and Methods)."
+
+above method remove / accounts for mean-variance correlation -
+
+The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
+
+QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
+
+@article{wang2017genetics,
+  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
+  author={Wang, Richard J and Payseur, Bret A},
+  journal={Genetics},
+  volume={206},
+  number={4},
+  pages={1841--1852},
+  year={2017},
+  publisher={Genetics Soc America}
+}
+
+references for phenotypic variance connected to robustness to environment, increase in developmental consistancy
+Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
+
+phenotypic variance leading to adaptation - 
+(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
+
+Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
+
+meiotic Checkpoints for ensuring consistancy (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
 
@@
 ```
+
+```{r variance.hist, echo=FALSE}
+
+```
+
+### modeling variance
 

### 2019-12-27 doc/Effects.Rmd
re-starting effect quan measurements

--- 2019-10-10 doc/Main_Report.rmd
@@
 ---
-title: "MLH1 Data Report"
+title: "Effects"
+author: "April Peterson"
 date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-  pdf_document: default
-bibliography: refs_Main_Report.bib
+output: html_document
 ---
 
-```{r setup, echo=FALSE, include = FALSE}
-library(knitr)
-library(ggplot2)
-#library(pwr)
-library(plyr)
-library(lattice)
-library(dplyr)
+```{r setup, include=FALSE}
 setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
@@
 
+library(knitr)
+library(ggplot2)
+library(plyr)
+library(lattice)
+library(dplyr)
+library(raster)
+library(lme4)
+library(nlme)
+library(RLRsim)
+library(ggpubr)
+library(reshape2)
+
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
-#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
 
-#load(file="data/MetaData.RData") #this was replacing most updated data file
-
-#load functions
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-```
-
-# Research Description
-
-Measure nMLH1 foci per meiotic cell to estimate recombination rate for diverse strains of house mice and related rodents.
-Comparison the differences in recombination rates across sexes and genetic background to inform models of how meiotic recombination rates evolve.
-
-## Discription of quantification process
-
-Cell images are quantified in batches after file names are anonymized.
-The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified.
-A quality score is given; 1 to 5 (best to worst).
-All cell image file names are anonymized and moved into batches of a few hundred images.
-
-```{r organizing.data, echo=FALSE}
-#set the order of another column, based on another variable (so that when)
-
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, category, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-#use the above line for when order is out of wak
-
-#count the non-quality measures,  #remove non qualit
-#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
-#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
-
-#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
-#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(category) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-
-#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
+#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
 
@@
 
-This section shows the number of mice which haven't been stained or quantified yet.
+```{r Effects.on.the.dataset, eval=FALSE, echo=FALSE}
 
-```{r missing.mice, echo=FALSE, warning=FALSE}
+#*age effects* (male, maternal age)
 
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-setwd("~./MLH1repo/")
-#other lists
+#*litter effect*
 
-Image_dirs <- list.files(path = "C:/Users/alpeterson7/Documents/Images")
-#write this out so that I can access it
+#*mouse room effect* (Charmony, MSC)
 
-#remove the non-mouse folders
-Dissection.File = read.csv("~./MLH1repo/data/Disections_53019.csv", header = TRUE)
-Dissection.File <- Dissection.File[!(is.na(Dissection.File$mouse)|Dissection.File$mouse==""),]
+#*quantification repeatability*
 
-#print the name with the date of the file somewhere
-Dissection.list <- Dissection.File$mouse
-Dissection.list <- as.character(Dissection.list)
+#*spread batch and staining effects*
 
-#load metadata .. start decerning which mice are missing information... and which should be used
-Meta.Data.File = read.csv("~./MLH1repo/data/ALP_MouseMetadata.csv", header = TRUE)
+#*power analysis on number of cells*
 
-#remove file name, and original name .. this is just a mouse level DF
-Meta.Data.File <- Meta.Data.File[!(is.na(Meta.Data.File$mouse)|Meta.Data.File$mouse==""),]
-#make sure that the DOB is a date
+#*time of year effect (on spread)*
 
-meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61019.csv", header = TRUE)
-meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%y")
-
-jl <- unique(Meta.Data.File$mouse[(Meta.Data.File$mouse %in% Dissection.File$mouse)])
-BD.w.meta <- Meta.Data.File[(Meta.Data.File$mouse %in% Dissection.File$mouse), ]
-
-#figure out a way to incorporate the staining... to get a better idea of which slides 
-#to do next, #remove the pero mice?
-
-#length(unique(MLH1_data$mouse) )
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-original_DF$fileName <- (original_DF$Original.Name)
-original_DF <- add_mouse(original_DF)
-#length(unique(original_DF$mouse) )
-#131 mice have been quant total, 124 mice have passed into the final version
-
-# make a list for the failed slides
-
-#Dissection.list - Image/dirs
-NonImaged.mice <- unique(Dissection.File$mouse[(Dissection.File$mouse %in% Image_dirs)])
-#231 mice dissected which need images
-
-#ImageFolder - MLH1.Quant
-NonQuant.mice <- unique(Image_dirs[(Image_dirs %in% MLH1_data$mouse )])
-#length(NonQuant.mice)
-#123 mice that need to get into anon-batch
-
-#make Dataframe to put into a table
-mouse.number.DF <- data.frame( category = 
-              c("Passed Quantified", "Total Quantified", "Non-Quantified", "Non-Imaged"),
-              number.of.mice = c(length(unique(MLH1_data$mouse)),
-                length(unique(original_DF$mouse) ),
-                        length(NonQuant.mice),
-                         length(NonImaged.mice)  )  )
-
-kable(mouse.number.DF)
-```
-
-passed quantified - quantified mice (with enough cells)
-total quantified -- all mice which have been (includes fail and passing mice)
-
-non-quantified mice -- dissected mice / slides that haven't been quantified yet (all stained, but without image directory)
-
-non-imaged -- mice without images, (might be stained and unstained)
-
-This is the number of mice that are through different parts of the pathway or quantification process.
-
-should I break down the categories into sex and strain?
-
-```{r basic.numbers, echo=FALSE, eval=FALSE}
-
-kable(table(AP_mouse_table$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-
-kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+#*dominant effect* (F1s)
 
@@
 
-These are the tables for the number of cells and mice which have been quantified.
-The 'Other' are F1 crosses.
+## Quality Effects
 
-## Repeatability of Human quantification
+below is a chunk to looking at histogram/distribution and means for the MLH1 data across quality bins
 
-Well, we have to talk about the subjective nature that goes into scoring the MLH1 foci.
-Sources of error or variance across scoring come from background noise (false positives), faint foci (false negatives) and just human error.
-A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (using some of Beth's data).
-Todo: track down, re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)
+```{r MLH1.histogram.2, echo=FALSE}
+#remake the pooled MLH1 histgoram plot from above but facet by quality
 
-In the original_DF and MLH1 DF's there are ~400 and X respectively of cells (filenames) that are duplicated, or cells that have been counted in more than 1 quant batch
+#PWD male, MSM male
+#LEW female, WSB female
+mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
+                 .(category, quality),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+#what's the point of the above table? -- add vlines?
 
-Datasets / sources of repeat measures
--chrom proportions
--multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
--old and new (batch0)
+#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect
 
-The main take away is that higher quality scored cells have lower error in MLH1 counts.
+histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
+histM_qual.KAZ_male
 
-> Does the amount of error differ across strains or sexes? (female might have a higher error rate)
+histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
+histM_qual.PWD_male
 
-> What sort of adjustments or tests should be made to results or analyses given the known error rate?
+histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
+histM_qual.MSM_male
 
-```{r quant.repeatability_rep1, echo=FALSE, include=TRUE}
-#set up data for combining replicate counts
+histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
+histM_qual.WSB_female
 
-#1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now
+histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
+histM_qual.LEW_female
 
-#isolate the duplicates (fileName) from main MLH1 DFs
-n_occur <- data.frame(table(original_DF$fileName))
-n_occur[n_occur$Freq > 1,]
-original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
 
-#are there any batch, fileName duplicates?
-original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")
-anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch ... really 2?
+#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
+#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
 
-n_occur <- data.frame(table(original_DF_duplicates$unique.key))
-n_occur[n_occur$Freq > 1,]
-double.trouble <- original_DF_duplicates[original_DF_duplicates$unique.key %in% n_occur$Var1[n_occur$Freq > 1],]
-#2 fileNames from batch 0 appear 2x, remove them from DF
-
-original_DF_duplicates <- original_DF_duplicates[ ! original_DF_duplicates$fileName %in% double.trouble$fileName, ]#removes 4 rows
-
-#seperated duplicates into categories out in Excel (write to /results, moved final file to /data)
-#This data frame was made ... with Jered's sugg
-#write the duplicate file to csv/excel
-#sort by file name
-#need to remove '?' -- deal with the ? and X's
-#write.table(original_DF_duplicates, "~./MLH1repo/results/orginal_DF_duplicates.csv", sep=",", row.names = FALSE)
-
-original_DF_duplicates.notes = read.csv("~./MLH1repo/data/orginal_DF_duplicates.csv", header = TRUE)
-#copies of fileNames can be distinguished with the 'separate.DF' column
-
-#remove '?' that I stupidly left in
-original_DF_duplicates.notes$nMLH1.foci = gsub("\\?", "", original_DF_duplicates.notes$nMLH1.foci)
-original_DF_duplicates.notes$quality = gsub("\\?", "", original_DF_duplicates.notes$quality)
-#these lines change col to chr
-
-original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
-#now there should be no more '?' going forward
-
-#original_DF_duplicates.notes$quality <- as.factor(original_DF_duplicates.notes$quality)
-
-#make lists of rows with "" or X's so I can remove all fileNames for those
-#these lists just have na
-quality.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$quality) | original_DF_duplicates.notes$quality=="")] #16
-
-MLH1.na.list <- original_DF_duplicates.notes$fileName[which(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci=="") ]#22
-
-X.list <- original_DF_duplicates.notes$fileName[grepl("X", original_DF_duplicates.notes$X)]#40
-
-#make DF which only has 
-matches.only <- original_DF_duplicates.notes[ ! original_DF_duplicates.notes$fileName %in% quality.na.list, ]
-matches.only <- matches.only[ ! matches.only$fileName %in% MLH1.na.list, ]
-matches.only <- matches.only[ ! matches.only$fileName %in% X.list, ]
-
-#str(matches.only)
-#clean up match list (chr to num)
-matches.only$nMLH1.foci <- as.numeric(matches.only$nMLH1.foci)
-matches.only$quality <- as.numeric(matches.only$quality)
-
-levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt
-
-#check that the 1, and 2 are equal number
-length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178
-
-#i want to measure correleation of MLH1 across bins of quality (1,2,3,4,5)
-#make new df with filename and MLH1.1 and mlh1.2
-
-dups.here <- data.frame(  FN = unique(matches.only$Original.Name), MLH1.1 = matches.only$nMLH1.foci[matches.only$separate.DF == 1], MLH1.2 = matches.only$nMLH1.foci[matches.only$separate.DF == 2],
-                        qual.1 = matches.only$quality[matches.only$separate.DF == 1] , 
-                        qual.2 = matches.only$quality[matches.only$separate.DF == 2]
-                          ) #OMG this worked!!!
-dups.here$diff <- abs(dups.here$MLH1.1 - dups.here$MLH1.2) #the aveage difference is pretty small, but some have very high difference
-#the high differences could potentially be mix ups
-
-repeatability.table <- ddply(.data=dups.here, 
-                 .(qual.1, qual.2),
-                 summarize, 
-                 ncells = length(unique(FN)),
-                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
-)
-#this table shows how many cells went changed into different qual score bins
-#easiest patter, 1-1, 1-2 (high correlation in MLH1)
-#5-5 lowest, (5-3 also low)
-#4-5 has high corr? (direction shouldn't matter)
-#4-2 has negative corr?
-
-#of the different quality permutation bins, 37% don't change quality bin, 44% of cells change by 1 (either direction)
-
-no.change = c(14,13,17,13,10)#67
-change.1 = c(10,7,8,20,9,14,4,7)#79
-
-repeatability.table.qual1 <- ddply(.data=dups.here, 
-                 .(qual.1),
-                 summarize, 
-                 ncells = length(unique(FN)),
-                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2),
-                 dif = mean(MLH1.1 - MLH1.2)
-)
-
-repeatability.table.qual2 <- ddply(.data=dups.here, 
-                 .(qual.2),
-                 summarize, 
-                 ncells = length(unique(FN)),
-                 plain_cor_MLH1 = cor(MLH1.1,MLH1.2)
-)
-#These tables show that qual1 cells have the highest corrleation of MLH1
-dups.here.plain_cor_qual <- cor(dups.here$qual.1, dups.here$qual.2)
-
-#i don't know how to compare quality in more
-
-#nMLHJ1, quality are factors
-original_DF_duplicates.notes$quality2 = gsub("\\?", "", original_DF_duplicates.notes$quality)
-
-original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
-
-#factor > chr > number  (this col is now chr is can go to num)
-#original_DF_duplicates.notes$MLH1.foci2 <- as.numeric(original_DF_duplicates.notes$MLH1.foci2)
-#original_DF_duplicates.notes$quality2 <- as.numeric(original_DF_duplicates.notes$quality2)
-
-#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
-#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
-#rename this
-#plain_cor_MLH1 <- cor(one$MLH1.foci2,two$MLH1.foci2,use = "complete.obs")
-#.857
-#plain_cor_qual <- cor(one$quality2,two$quality2,use = "complete.obs")
-#.63
-
-#What's the relationship of MLH1 cor with quality?
-#what the rpeat of quality score?
-#what are the correlations across quality score bins?
-
-#issues that complicate the plain calculation
-#some duplicates were 'x'ed and some weren't -- this causes different numbers
-#connected with above, some duplicates have mlh1 scores, other don't -- this kinda messes up correlation
-
-original_DF_duplicates.notes <- original_DF_duplicates.notes[ !grepl("X", original_DF_duplicates.notes$X) , ]
-original_DF_duplicates.notes <- original_DF_duplicates.notes[!(is.na(original_DF_duplicates.notes$nMLH1.foci) | original_DF_duplicates.notes$nMLH1.foci==""), ]
-#this causes unequal lengths
-
-original_DF_duplicates.notes <- droplevels(original_DF_duplicates.notes)
-original_DF_duplicates.notes$nMLH1.foci <- as.character(original_DF_duplicates.notes$nMLH1.foci)
-
-original_DF_duplicates.notes$nMLH1.foci <- as.numeric(original_DF_duplicates.notes$nMLH1.foci)
-
-#correlationsnMLH1 
-#one <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 1,]
-#two <- original_DF_duplicates.notes[original_DF_duplicates.notes$separate.DF == 2,]
-#plain_cor <- cor(one$nMLH1.foci,two$nMLH1.foci)
-#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
-
-#check that all batches are distinguishable across duplicate fileNames
-#make unique key fileName_batch -- then check for duplicate of that 
-#original_DF$Batch <- factor(original_DF$Batch, ordered = TRUE, levels = c("Batch0", "Batch1", "Batch2", "Batch3", "Batch4","Batch5","Batch6",  "Batch7", "Batch8","Batch9","Batch10", "Batch11","Batch12", "Batch13", "Batch14","Batch15"))
-#after doing this, original_DF$Batch[2000] > original_DF$Batch[10] will be evaluated!
+#ad
+#  geom_smooth(method = lm, se = FALSE
+ # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
 
@@
 
-```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
-#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
-#load, clean up and mat one of the master.MLH1count sheets
-#master
-old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
-#remove all with 'X'
-#create a real fileName col, add rev.tif, 
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+I'm not sure sure how effective the above plots are.
 
-old.master.MLH1 <- add_mouse(old.master.MLH1)
-old.master.MLH1 <- add_strain(old.master.MLH1)
-old.master.MLH1 <- add_sex(old.master.MLH1)
-old.master.MLH1 <- add_category(old.master.MLH1)
-old.master.MLH1$fileName2 <- old.master.MLH1$fileName
-
-test.old.master <- rbind(head(old.master.MLH1), tail(old.master.MLH1) )
-test.old.master$fileName <- as.character(test.old.master$fileName)
-
-test.old.master$stain <- as.character(test.old.master$stain)
-
-str(test.old.master)
-
-#paste together file name, requires mouse, strain
-#this might also need some character / factor stuff ... filename and stain MUST be charaecter!
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#wrote add_fullFileName to make the file names for the old excel sheets
-
-old.master.MLH1 <- add_fullFileName(old.master.MLH1)
-old.master.MLH1$fileName <- old.master.MLH1$file.name_full
-
-#merge and match to MLH1
-
-#need to test for duplicates before mergeing?
-n_occur <- data.frame(table(old.master.MLH1$file.name_full))
-n_occur[n_occur$Freq > 1,]
-ll <- old.master.MLH1[old.master.MLH1$file.name_full %in% n_occur$Var1[n_occur$Freq > 1],]
-old.master.MLH1 <- old.master.MLH1[ ! old.master.MLH1$fileName %in% ll$fileName, ]
-
-n_occur <- data.frame(table(original_DF$fileName))
-n_occur[n_occur$Freq > 1,]
-original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]
-table(original_DF_duplicates$mouse)#17 mice
-
-old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
-
-#old.master.MLH1 <- original_DF[ ! original_DF$fileName %in% ll$fileName, ]
-
-#there are duplicates in original_DF and MLH1
-
-uutesting2 <- merge(old.master.MLH1, original_DF, by.x = fileName, by.y = Original.Name)
-#match old.full.filname with MLH1$fileName
-```
-
-```{r repeatability_rep3, echo=FALSE, include=FALSE}
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#what are the target columns for this excel sheet?
-#espically for getting repeat data
-#MLH1.dif  same_count  quality
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-
-#assessing error,
-#number for same_count == TRUE
-#length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"])
-#length(chrm.class.DF$Original.Name)
-
-#cor of first count and calq count
-#cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)
-#.87
-
-#the true false might have to be changed first
-#chrm.class.DF$same_count_num <- as.numeric(chrm.class.DF$same_count)
-
-#regression of quality and MLH1.dif
-#lmreg <- lm(abs(MLH1.dif) ~ quality + nMLH1.foci, data= chrm.class.DF)
-#summary(lmreg)
-#according to this linear regression, quality is not a significant
-
-#make regression for just same counts
-#I tried making a  binomial regression model, 
-#but quality is bad predictor, only 5 levels)
-#lmreg2 <- glm(same_count ~ quality, data= chrm.class.DF, family=binomial(link="logit"))
-#summary(lmreg2)
-#to plot the model / bimodal data
-#fit = glm(same_count_num ~ as.numeric(quality), data=chrm.class.DF, family=binomial)
-#newdat <- data.frame(quality=seq(min(chrm.class.DF$quality), max(chrm.class.DF$quality),len=102))
-#newdat$vs = predict(fit, data=newdat, type="response")
-#the prediction is not going 1 or 0 -- so doesn't fit logistic
-#plot(vs ~ quality, data=newdat, col="red4")
-#lines(vs ~ quality, newdat, col="green4", lwd=2)
-
-hist_error_plot <- ggplot(data=chrm.class.DF, aes(MLH1.dif))+geom_histogram()+ggtitle("Distribution of foci count differences across 2 count versions")
-
-#
-quality_error_plot <- ggplot(data=chrm.class.DF, aes(x=quality, y=abs(MLH1.dif)  ))+geom_jitter()+ggtitle("Repeatability by quality score")+ylab("difference in quality score")+xlab("quality score")
-
-error_count_plot <- ggplot(data=chrm.class.DF, aes(y=nMLH1.foci, x=abs(MLH1.dif)  ))+geom_jitter()+ggtitle("Error by MLH1 number")+xlab("difference in two counts")+ylab("MLH1 count")
-
-#in the chrm proportions data is another set for eastimating error
-```
-
-Data use for this repeatabiity estimate is from smaller dataset where the proportions of chromosome/bivalent classes (0CO bivalents ect.) were quantified.
-(The focus of one count was cell total, while the other focus was classifying bivalents).
-
-This data has `r length(chrm.class.DF$Original.Name)` cells across these categories
-
-`r kable(table(droplevels(chrm.class.DF$category)))`
-
-More cells are planned to be quantified.
-
-```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE}
-hist_error_plot
-```
-
-From this histogram we see most of the cells in this dataset have a difference of 2 in the MLH1 count scores (`r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`).
-
-```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5}
-quality_error_plot
-```
-
-There is a positive relationship with the absolute difference and quality score (where 1 is best and 5 is worst quality) (i.e.
-poorer quality cells are more likely to have more error acorss multiple counts).
-The proportion of the cells which had the same count across the counting instances is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`.
-The correlation between the two versions of MLH1 counts for the same cell is `r cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)`.
-
-```{r show.error_count_plot, echo=FALSE,  fig.width=10, fig.height=5}
-error_count_plot
-```
-There doesn't seem to be a bias across cells with more or less foci.
-
-Modeling this relationship; (still in progress)
-I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.
-
-## Effects on the dataset
-
-error associated with 'quality of cels'
-
-Main metric is mean CO count per mouse. Run tests to check if there are effects for the following effects.
-
-*age effects* (male, maternal age)
-
-*litter effect*
-
-*mouse room effect* (Charmony, MSC)
-
-*quantification repeatability*
-
-*spread batch and staining effects*
-
-*power analysis on number of cells*
-
-*time of year effect (on spread)*
-
-*dominant effect* (F1s)
-
-These aspects will be analyzed with a linear models on subsets of data (within category) and visualized with plots.
-
-### Quality effect
-
-There is a positive correlation with MLH1 number and quality
-To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.
-
-```{r quality.effect, echo=FALSE}
-
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
-#compare all quality 1 cells across strains
-#check that the means (or directions of differences is maintained)
-
-Q12_MLH1_DF <- MLH1_data[MLH1_data$quality <= 2,]
-
-#ddply
-Q12_MLH1_means <- ddply(.data=Q12_MLH1_DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-tot.DF_MLH1_means <- ddply(.data=MLH1_data, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-mergy <- merge(Q12_MLH1_means, tot.DF_MLH1_means, by='category', all = F)
-colnames(mergy) <- c("category","Q12ncells","Q12mice","Q12_MLH1_means","Q12MLH1var",
-                     "Totncells","Totnmice","TotMLH1mean","TotMLH1var")
-
-mergy$percent.diff <- (mergy$Q12_MLH1_means / mergy$TotMLH1mean)
-
-Q12toTotal <- ggplot(mergy, aes(y=Q12_MLH1_means, x=TotMLH1mean))+geom_point(aes(size = Q12ncells) )+geom_abline(slope = 1, intercept = 0)+ggtitle("Quality 1 cells MLH1 mean to total MLH1 mean")+ylim(c(21,32))+xlim(c(21,32))
-
-Q12toTotal
-
-#check that the directions are the same
-#make tables for the 
-#is the amount of 'polymorphism' the same
-
-Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "male",], 
-                 .(category,subsp),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci)
-)
-
-tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "male",], 
-                 .(category,subsp),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci)
-)
-
-kable(Q12_MLH1_means_males, caption = "High quality cells")
-
-kable(tot.DF_MLH1_means_males, caption = "Total Cell")
-
-```
-
-All the quality 1 cells produce a higher mean. The biggest difference in means is HMI male, LEW female and SPIC female means have ~10% differences.
-
-When quality 1 to 2 is compared, the strains were closer to the line, but the means are still higher in the higher quality cells.
-with 1-2 most all of the difference are below 10% (except SPRET female).
-
-MOLF males have quite a big difference.
-
-There seem to be the similar amount of polymorphism in the Dom and more conserved in Dom.
-
-Ranges of category means across subspecies.
-
-| Type | male Dom | female Dom | | male Musc low | male Musc high | Musc female | | male Outgroup | female outgroup |
-| --- | --- | --- | --- | --- | --- | --- | --- | --- | --- |
-| Total cells | 22-24 |  24-27 |  | 23-31 | 25-31 | 26-29 | | 25-26 |27 |
-| High quality cells | 23-25 | 25-29 | | 22-25 | 22-27 | 25 | | 26-27 | 29-31 |
-(ASTmale in Musc high, has a 27 mean which might be an outlier)
-I think the ranges across the groups is generally consistant.
-
-(try looking at <5 quality scores?)
-
-```{r high.qual, fig.caption="boxplots of MLH1 distributions", fig.height=5, fig.width=5, echo=FALSE, warnings=FALSE, include=FALSE}
-
-#MLH1_data$quality <- as.numeric(MLH1_data$quality) #this needs to be run for the subsetting to work
-highest_qual_df <-  subset(MLH1_data, as.numeric(MLH1_data$quality) <= 2 )
-
-#are there no KAZ cells of high quality?
-qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey", 
-             "grey2","grey3","grey1","grey","grey","grey"))
-high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
-#high_qual_boxplot
-#these boxplots look bad
-```
-
-> *How do the distributions change across different quality scores?*
-
-Higher quality scored cells have more MLH1 foci. (some of the ranges get smaller with higher quality cells)
-
-> *How do the distributions change across mice with a good number of cells?*
-
-(was I referring to normale distributions?)
-means all shift up, what about variance, range, lowest cut off
-
-> *Do the directions of the differences across groups change due to quality scores?*
-
-(polymorphism vs divergence patterns)
-
-### Age Effect
+### Age Effect (male)
 
@@
 
-Make a subsetted dataset for plotting mean mouse (quality < 5) MLH1 count and age.
+I like this plot and find it effective. Can I do a similar set up for other factors (quality, ncells?)
 
@@
 #for all models, mostly it was that the intercept was significant, and mot much else
-```
 
-```{r female.age.effect, echo=FALSE}
-#start thinking of the maternal age effect
-
-#female age effect
-
-#female.meta.data <- meta.data[meta.data$sex == "female",]
-#how much data to I have to work with?
-#there's a fair number of rows without emb or neonate
-#I could probably assume all of these are neonates
-
-#length( !is.na(female.meta.data$female.stage) )
-
-#female.meta.data <- female.meta.data[female.meta.data$female.stage]
-
-#how much formating do I have to do?
-#can't use AP_mouse_table since that is summarized
-#meta.data
-#grepl notes, "neo"
-
-#some have e-number for embryonic
-#I might have to
+#good knitting output
+#as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )
 
 ```
-
-I will try doing a logistic regression for embryonic vs neonate. The embyonic estimates likely aren't very accurate, and there won't be good sampling across time points.
 
@@
 Compare female spreads made on the same day, but from different mothers/litters.
+```{r litter.effect, eval=FALSE}
+#compile this data set
 
-### Spread and Stain Effects
+```
+
+# Dissection Effect
+
+# Spread and Stain Effects
 
@@
 
-### Room effect (envioronment)
+```{r stain.effects, eval=FALSE}
 
-3 different rooms used to house mice. Access softmouse to be fixed. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
-
-### Cell number (power analysis results)
-
-Insert the code and results for effects of cell number/observations on MLH1 means.
-
-# Initial Results (change these colors)
-
-```{r first.boxplots, echo=FALSE, fig.align='center'}
-
-#the below ordering works!! keep it
-MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
-              "PWD","MSM","MOLF","CZECH","SKIVE",
-              "KAZ","TOM","AST",
-              "CAST","HMI","SPRET","SPIC", "other"), order=T )
-#order the data frame
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-#don't know why this won't do the right order...
-#set the order of another column, based on another variable (so that when)
-
-#try to make sure sacles are the same
-
-#reduncant
-#MLH1_data <- MLH1_data %>%
- # arrange(strain, sex, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-#try this for ordering
-#MLH1_data[order(MLH1_data$strain), ]
-
-#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
-#can't re-order the miceagg
-#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
-#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
-#        "grey","grey","grey",  "grey","grey","grey" )) +
-# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
-#don't show M.musculus by mouse
-
-#order MLH1_data by strain!!
-#colors_of_strains, colors ..
-
-ff2 <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
-  geom_boxplot(aes(x = as.character(mouse), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
-        "green","green4","grey",  "grey","grey","grey" )) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
-  facet_grid(subsp~ sex, scales="free_x")+scale_color_manual(values=colors_of_strains)
-
-#scales="free_y"
-#this is ok except for the set placement of mice on x axis
-#if the full dataset was ordered correctly, this would kinda fix it
-
-#boxplots by strain
-ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+
-  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )+scale_fill_manual(values=colors_of_strains)
-
-ss <- ss   + facet_wrap(~ sex, scales="free")
-
-#these plot currently look bad
-ss
-```
-
-*What is the initial picture of Heterochiasmy across mouse strains?*
-
-For the strain level plots
-  -facet,sex by species
-
-```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
-#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
-
-#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
-#remove quality 5 cells
-
-#outgroups
-#SPIC
-HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
-#unique(HetC.SPIC$mouse)
-#non-passing mice
-SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
-HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
-#calq sex mean
-SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
-
-#SPRET
-HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
-#unique(HetC.SPRET$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
-
-#CAROLI
-HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
-#unique(HetC.CAROLI$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-#cast
-HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
-#unique(HetC.CAST$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
-
-#HMI
-HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
-#unique(HetC.HMI$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-#dom
-
-#WSB
-HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
-#unique(HetC.WSB$mouse)
-#non-passing mice
-WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
-HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
-#calq sex mean
-WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
-
-#LEW
-HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
-unique(HetC.LEW$mouse)
-#non-passing mice
-LEW.remove.mice <- c("4jan17_LEW_f2")
-HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
-
-#calq sex mean
-LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
-
-#G
-HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
-#unique(HetC.G$mouse)
-#non-passing mice
-G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
-
-#calq sex mean
-G.mean.MLH1.obs <- ddply(.data=HetC.G, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#PERC
-HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
-#unique(HetC.PERC$mouse)
-#non-passing mice
-#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
-
-#calq sex mean
-PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#MUSC
-
-#PWD
-HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
-#unique(HetC.PWD$mouse)
-#non-passing mice
-PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
-
-# CZECH
-HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
-#unique(HetC.CZECH$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
-
-# SKIVE
-HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
-#unique(HetC.SKIVE$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
-
-# KAZ
-HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
-#unique(HetC.KAZ$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
-
-#TOM
-HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
-#unique(HetC.TOM$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
-
-#AST
-HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
-#unique(HetC.AST$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
-
-#Molosinus
-
-#MSM
-HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
-unique(HetC.MSM$mouse)
-
-#non-passing mice
-MSM.remove.mice <- c("12sep16_MSM_f3")
-HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
-
-#calq sex mean
-MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
-
-#MOLF
-HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
-unique(HetC.MOLF$mouse)
-
-#non-passing mice
-MOLF.remove.mice <- c("27oct18_MOLF_m3")
-HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
-
-#calq sex mean
-MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
+#compile this dataset
 
@@
 
-```{r HetC.plot, echo=FALSE, warning=FALSE}
-#HETC plot
+# Room effect
 
-#make scatter plot for HetC
-#color by strain
+```{r ROOM.effect, eval=FALSE}
 
-#set up data
-#make female male plot
-
-HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#
-#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
-
-HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
-
-HetC_strain_table.M$sex <- "male"
-HetC_strain_table.F$sex <- "female"
-
-HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
-
-HetC_strain_table <- add_subsp(HetC_strain_table)
-
-#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
-
-HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
-HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
-#str(HetC_strain_table)
-
-Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
-  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey"))
-
-#musc - coral, Dom - cadetblue (grey blue),
-
-Het.C.plot
-#this plot is missing several
+#COMPILE THIS datasets
+#3 different rooms used to house mice. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
 
 ```
-
-3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
-1 of the musc strains has female biased heterochiasmy, KAZ.
-
-### Distributions of MLH1 counts
-
-> *Are the MLH1 distributions close to normal?*
-
-```{r MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}
-
-#gg <- data.frame(weight=rep(x,nrow(df)),df)
-#gg$y <- with(gg,dnorm(x,mean,sd))
-#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30
-
-mean.MLH1.obs <- ddply(.data=MLH1_data, 
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-histM <- ggplot(MLH1_data,(aes(adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")+
-  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))
-
-#  geom_smooth(method = lm, se = FALSE
- # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
-
-histM
-
-#function = glm(y ~ x, family = poisson)
-#add Poission curve behind
-#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
-#not sure how to add Expected poisson curve
-
-```
-
-(most all distributions are close to normal)
-
-How do these compare to Poisson? (mean == variance)
-I should make 'expected' distributions behind these.
-
-```{r passing.mice, echo=FALSE, warning=FALSE}
-q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
-                        total =  length(adj_nMLH1.foci),
-                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
-                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
-                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
-)
-#q_cutoff_table
-passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)
-
-not_passed <- as.vector(unique(MLH1_data[!(unique(MLH1_data$mouse) %in% passed_mice),]$mouse) )
-
-#mnake dataframe with just passing mice -- so that it can be ploted
-passed_mice_df <- MLH1_data[ (as.numeric(MLH1_data$quality) < 5), ]
-```
-
-Human quantification seems to be biased towards rating cells with more MLH1 foci as higher quality.
-Unbiased cell quality assignment, would not show a positive correlation with quality and nMLH1.
-(CAST female data is fake)
-
-```{r scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means and quality scores", warning=FALSE}
-
-#make lattice / plots seperated by strain and sex
-#facet
-
-#quality scores are now 1:7?
-qual_means <- MLH1_data %>% 
-        group_by(category, quality) %>% 
-        summarise(
-          adj_nMLH1.foci = mean(adj_nMLH1.foci)
-        )
-
-p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
-p <- p + geom_point(stat = "identity")
-p <- p  + geom_point(data = qual_means, colour="red") 
-p <- p + facet_wrap(~ category)
-p
-
-```
-
-I plotted the mean of each quality bin with a red dot.
-From the pattern of the red dots, there is definitely a negative relationship with quality and nMLH1 foci across the data.
-This is most pronounced in MSM males and least pronounced in G males.
-The CAST female data is dummy data.
-
-####Distributions by mouse
-
-> *Is there mouse level MLH1 quality score bias?*
-
-Below are code chunks that make a big dfs of CO count by quality score plot. The code has been commented out, and moved to it's own doc.
-
-```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
-### to integrate 'pass'/fail
-### this should be in setup data
-
-#mice that don't have female data will break the chain-loop
-#MOLF female, HMI, CZECH female, CAROLI female, Tom male
-
-#instead of adding fake data, remove the categories from list
-
-#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
-#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
-#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
-#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
-#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
-#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
-#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
-#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
-#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
-
-pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
-
-#this loop makes a quant_status col for a list of mice
-for(i in 1:length(MLH1_data$mouse)){
-  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-}
-
-df_list <- split(MLH1_data, (MLH1_data$category))
-#remove sections which are in category but have no data
-#this is where you remove the non-data data!
-df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
-
-#7 PERC, 12 MOLF female, 
-#
-
-#df list, is a list of df's which were split() from BIG df by category
-df_list 
-o=0
-i=0
-
-MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
-
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-for(i in 1:(length(df_list))){
- # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                      mouse_nMLH1 = mean(adj_nMLH1.foci)
-  )
-# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-  for(o in 1:length(df_list[[i]]$Original.Name)){
-    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
-        }
-}
-
-plot_list = list()
-for (h in 1:length(df_list)) {
-  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
-    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
-    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
-    facet_wrap(~ mouse, drop = TRUE) + 
-    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
-  #make an empty ggplot? for missing data
-
-  plot_list[[h]] =   pf
-}
-
-#orig (I think this is the same as above)
-#plot_list = list()
-#for (h in 1:length(df_list)) {
-#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
-
-#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
-#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
-
- #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
-
-    #theme for indiviual mice, 
-#theme(plot.background = element_rect(fill = "green"))
-#  plot_list[[h]] =   pf
-#}
-#can't change the alpha/transparency of the red rec
-
-#aes must be same length ---   #geom_point(stat = "identity") +
-# the last one is correct....?
-
-#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
-```
-
-```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
-geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-#no plot is made that goes in the list -- so it can't be shown
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
- # invisible(lapply(obj, function(x) plot(x,main="some plot")))
-}
-
-#this stops at MSM
-```
-
-The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
-
-Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
-The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
-
-# Research Questions
-
-### Explaination for the XX and XY differences
-
-(in progress)
-
-### Difference in proportion of chrm classes
-
-generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
-10 cells from each musc category.
-
-```{r draw.rand.cells, echo=FALSE}
-#ss <- sample(MLH1_data[MLH1_data$category == "PWD male",1:4], 10, replace = FALSE)
-#ii <- MLH1_data[ sample( which(MLH1_data$category=='G male'), 10 ), ]
-```
-
-```{r chrm_prop_data, echo=FALSE}
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF[chrm.class.DF$sex == "male",], 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check),
-
-                 n0CO = sum(n0CO),
-                  n1CO = sum(n1CO),
-                  n2CO = sum(n2CO),
-                  n3CO = sum(n3CO),
-                  n4CO = sum(n4CO),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#str(this_thing2)
-
-#this first one is the 'first' MLH1 count
-oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
-#summary(oopp2)
-#none of the classes are significant
-
-#mean is the new calculated-- mean 0CO is significant and intercept
-oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
-#summary(oopp)
-#not sure if these are inter-correlated
-#I don't know why 0COs and the intercept are significant..
-#0CO and 1CO are significant when only chrm classes are included (and intercept)
-
-#no longer significant when used
-#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
-#after added more MSM data, tot_biv is not significant
-#0CO is now more significant, and 2CO is sightly significant
-
-#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
-
-#MSM needs more bialent observations
-
-#use melt to put all propotions
-
-#display / visualize / compare proportions
-
-melt.table = read.csv("~./MLH1repo/data/ChrmClassProportions_melt.csv", header = TRUE)
-melt.table$fileName <- (melt.table$Original.Name)
-
-#melt.table <- add_mouse(melt.table)
-#melt.table <- add_strain(melt.table)
-#melt.table <- add_sex(melt.table)
-#melt.table <- add_category(melt.table)
-
-melt.table$strain<- factor(melt.table$strain,levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM"), order=T )
-
-melt.table <- with(melt.table, melt.table[order(strain),])
-
-other_thing <- ggplot(data = melt.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~subsp)
-
-other_thing
-
-chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
-
-#chrm.class.DF$biv.sum.check
-```
-
-High musc males have half 1CO and 2CO bivalents.
-For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
-
-### Mixed Model, compare within and between subspecies patterns
-
-The mixed model lets us sperate effects to see how they effect a mouse mean CO count. This seperation can also serve as proxies for divergence and polymorphism.
-
-Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
-
-> Mouse average Models
-
-  variables vs char coding of coefficents(opposite direction of sex)?
-  select1 (remove old mice, remove mice with low ncells)
-  select2 make sure there are equal numbers of observations
-  select3 low Musc males
-
-the pooled cell models are breaking for some reason, (the sexes)
--pooled cells (nMLH1)
-  all 
-  pooled.HQ (1-4)
-  pooled.HQ.2 (sample equal numbers of cells)
-
-```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE}
-library(lme4)
-library(nlme)
-# seperate the data, both sexes, only Mus
-# decide on Musc - Molossinus
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#change molossinus to it's own subsp (grepl)
-DF.HetC.MixedModel2 <- DF.HetC.MixedModel
-
-#if(grepl("MSM", DF.HetC.MixedModel2$strain) ){ DF.HetC.MixedModel2$subsp <- "Mol" }
-
-#subsp are charecters
-#change sex to charecter? (unordered sex)
-
-DF.HetC.MixedModel2$sex <- as.character(DF.HetC.MixedModel2$sex)
-
-model.ff2 <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel2, random=list(strain=pdDiag(~sex) ) )#adding mol
-
-#pull out values / other things
-#summary(model.ff2)# indeces for this object: [1]-structure
-#ranef(model.ff2)
-#after coding subsp to chr from factor, the co-efficients are
-
-#intercept should be the Dom female mean, looks right
-mean(DF.HetC.MixedModel2$mean_co[(DF.HetC.MixedModel2$sex == "female" &
-                                    DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5
-
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-
-#after initial analysis look into quality of these mice and ages
-
-#clean up the DF, remove extra cols
-
-#colnames(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-                                                            subsp.y,subsp.x,
-                                                            category.y, category.x,
-                                                            strain.x,strain.y, mouse.1  )  )
-
-#pooled.cells
-DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
-
-#remove PERC, MOLF, CZECH,
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
-```
-
-For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)
-
-subspecies levels: 2, Musc and Domesticus (maybe Molossinus soon)
-strain levels: Dom: 3, Musc: 4
-sex: 2 female male
-
-`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
-
-# Building Mixed Models
-
-```{r mixed.model.make, echo=FALSE, warning=FALSE, eval=FALSE, message=FALSE}
-#build mixed model
-#run simulations for accounting for sampling error
-# old code uses base plot
-library(nlme)
-library(lme4)
-
-#L,Q,C  appear for ordered factors in the summary mean lme doesn't believe you used ordered factors
-#change all the DFs / variables to UNORDERED factors!
-#redo the orders just to be sure
-#sex, strain, subsp,
-
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE, levels =c("female", "male") )
-
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE, levels =c( "Dom", "Musc") )
-
-DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW",
-                                                  "PWD", "MSM", "SKIVE", "KAZ") )
-
-#pooled cells,  DF.HetC.MixedModel.pooled
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
-
-DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
-
-DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
-
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-
-#sex only has 2 levels and this breaks...
-model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.pooled)# indeces for this object: [1]-structure
-ranef(model.pooled)
-
-model.pooled$coefficients
-```
-
-List of mixed models with mean_co
-
-1. mean_co --DF.HetC.MixedModel
-2. mean_co.chr  variables coded as factors (for ggplots), how is this changed when they are coded as chr
-3. mean_co.fac -- non-ordered factors
-4. mean_co.HQ
-5. mean_co.HQ.mol
-6. Sub-sampling for equal numbers
-
-```{r mixed.models.setup, echo=FALSE}
-#make are the DFs for the models
-
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-                                                            subsp.y,subsp.x,
-                                                            category.y, category.x,
-                                                            strain.x,strain.y, mouse.1  )  )
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]
-
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE )
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE )
-DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE )
-
-#main df with variables coded as chr
-DF.HetC.MixedModel.chr <- DF.HetC.MixedModel
-DF.HetC.MixedModel.chr$subsp <- as.character(DF.HetC.MixedModel$subsp)
-DF.HetC.MixedModel.chr$sex <- as.character(DF.HetC.MixedModel$sex)
-DF.HetC.MixedModel.chr$strain <- as.character(DF.HetC.MixedModel$strain)
-
-#mean_co.HQ
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mean_co.HQ.mol
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
-        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
-        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
-
-#mean_co.chr.HQ.mol
-DF.HetC.MixedModel.chr.HQ.mol <- DF.HetC.MixedModel.HQ.mol
-
-DF.HetC.MixedModel.chr.HQ.mol$subsp <- as.character(DF.HetC.MixedModel.chr.HQ.mol$subsp)
-DF.HetC.MixedModel.chr.HQ.mol$sex <- as.character(DF.HetC.MixedModel.chr.HQ.mol$sex)
-DF.HetC.MixedModel.chr.HQ.mol$strain <- as.character(DF.HetC.MixedModel.chr.HQ.mol$strain)
-
-```
-
-```{r MixedModels.stff, echo=FALSE}
-#main mean.CO
-HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-#mean.co.chr
-HetC.MixedModel.mean_co.chr <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr, random=list(strain=pdDiag(~sex) ) )
-
-#mean.co.HQ
-HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-#mean.CO.HQ.mol
-HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-#mean.HQ.mol.chr
-HetC.MixedModel.chr.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-```
-
-The intercept level is Dom female. These are the coefficients
-
-`r kable(HetC.MixedModel.mean_co$coefficients$fixed, col.names = "coefficients")`
-
-`r kable(ranef(HetC.MixedModel.mean_co) )`
-
-changed all to unordered factors.
-
-all the variables are coded as ordered factors.
-
-When the variables are kept factors, and unordered the names appear! yay!
-
-`r kable(HetC.MixedModel.mean_co.chr$coefficients$fixed, caption="Chr coded Model", col.names = "coefficients")`
-
-`r kable(ranef(HetC.MixedModel.mean_co.chr) )`
-
-The variable names now appear. The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.
-
-`r kable(HetC.MixedModel.HQ$coefficients$fixed, caption="mean_co HQ model", col.names = "coefficients")`
-`r kable(ranef(HetC.MixedModel.HQ) )`
-
-The factor coded DF the more high quality cells.
-
-`r kable(HetC.MixedModel.HQ.mol$coefficients$fixed, caption="Main model with molossinus", col.names = "coefficient")`
-`r kable(ranef(HetC.MixedModel.HQ.mol) )`
-With adding molossinus, the subsp names are listed.
-
-`r kable(HetC.MixedModel.chr.HQ.mol$coefficients$fixed, caption="Main model, chr coded with molossinus", col.names = "coefficients")`
-`r kable(ranef(HetC.MixedModel.chr.HQ.mol) )`
-
-When the variables/levels are coded as chr, the interaction effects get larger and the other subspecies get smaller. The base sex effect is also larger. The subsp effect is relative to Dom female,
-
-I don't think the subspecies and strains should be order, (maybe factors still makes sense.)
-
-This table doesn't display well
-
-`r table(DF.HetC.MixedModel.pooled$sex,DF.HetC.MixedModel.pooled$strain)`
-
-KAZ, LEWES has ~100 more for males, PWD has ~200 for females.
-
-Fixed effects
-
-look at the p-values for each of the effects as an indication if they are significant predictors.
-
-Subspecies
-
-Random effects
-
-try subsampling some of the
-
-## Meiotic Error
-
-This section reviews all the data from achiasmy and 0CO bivalents.
-
-### Another Heterochiasmy Plot
-
-```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average"}
-
-no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
-
-MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
-
-MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
-
-kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
-
-bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
-
-#add titles
-#use multiplot
-#remove X axis
-
-multiplot(kuku,bubu, cols=2)
-```
-
-## SC.skel measures
-
-data has been collection for RWang's total SC. These results are consistant with the patterns seen in the bivalent level resutls. (total SC PWD > KAZ )
-
-```{r RWscSkel, echo=FALSE}
-#merge the sc.skel outputs with the master MLH1 measures
-#match up random names with original names
-
-#the skel output and scripts are in the SC_skel dir. it was run on all images in the anon_data folder. I haven't manually confirmed that the default parameters are skeletonizing the images correctly.
-
-#some of richard's analysis show negative relationship with SC length and MLH1 (his QTL word doc). Could this be sign
-# of pwd like regulation?
-
-setwd("~./")
-SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_sep19.csv", header = TRUE, strip.white = TRUE)
-
-SC.skel$rand_name <- as.character(SC.skel$rand_name)
-original_DF$Random.Name <- as.character(original_DF$Random.Name)
-#str(SC.skel)
-#str(original_DF)
-
-mlh1.skel.og <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-#for some reason this only worked after the chr and all = F
-
-mlh1.skel.og <- mlh1.skel.og[ !grepl("X", mlh1.skel.og$Batch) , ]
-mlh1.skel.og <- mlh1.skel.og[ !grepl("x", mlh1.skel.og$Batch) , ]
-mlh1.skel.og <- mlh1.skel.og[!(is.na(mlh1.skel.og$quality) | mlh1.skel.og$quality==""), ]
-
-mlh1.skel.og <- add_mouse(mlh1.skel.og)
-mlh1.skel.og <- add_strain(mlh1.skel.og)
-mlh1.skel.og <- add_sex(mlh1.skel.og)
-
-mlh1.skel.og <- add_category(mlh1.skel.og)
-mlh1.skel.og <- add_subsp(mlh1.skel.og)
-
-#mlh1.skel.og <- add_euth_date(mlh1.skel.og)
-#mlh1.skel.og <- add_age(mlh1.skel.og)
-
-#str(mlh1.skel.og)
-mlh1.skel.og$skel_size  <- as.numeric(mlh1.skel.og$skel_size)
-mlh1.skel.og$bin_size <- as.numeric(mlh1.skel.og$bin_size)
-mlh1.skel.og$nMLH1.foci <- as.numeric(mlh1.skel.og$nMLH1.foci)
-
-#remove outliers
-mlh1.skel <- mlh1.skel.og[mlh1.skel.og$skel_size < 25000,]
-#first test file has ~1000 measures
-#when I ran it, no sign of outliers (after removing all the X's cells)
-
-looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=quality) ) + geom_jitter() + facet_wrap(sex~strain)
-
-looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Musc",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
-#very possible that the dimesions and pixel density are skrewing with these values
-# I am suspcious that there is such a wide range for the female data
-looky
-
-looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
-looky.DOM
-
-#mouse level
-mouse.level.skel <- ddply(.data=mlh1.skel.og, 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(nMLH1.foci),
-                 mean.skel = mean(skel_size)
-)
-
-mouse.level.skel <- add_strain(mouse.level.skel)
-mouse.level.skel <- add_sex(mouse.level.skel)
-mouse.level.skel <- add_category(mouse.level.skel)
-mouse.level.skel <- add_subsp(mouse.level.skel)
-
-looky.mouse <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Musc",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)
-
-looky.mouseDom <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Dom",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +ylim(c(150,1200))+facet_wrap(subsp~sex)
-
-#theme(legend.position="none")
-
-#double check a way to visually inspect the python skel output
-#
-#repeat random numbers?
-```
-
-```{r test.BivData, echo=FALSE}
-#bivdata
-setwd("~./")
-MSM_BivData.og = read.csv("~./MLH1repo/data/BivData/Curate_MSMmale_9.9.19.csv", header = TRUE, strip.white = TRUE )
-
-MSM_BivData <- MSM_BivData.og[MSM_BivData.og$SC.pass ==1,]
-MSM_BivData <- MSM_BivData[!(is.na(MSM_BivData$hand.foci.count) | MSM_BivData$hand.foci.count==""), ]
-
-#MSM_BivDatam_CO2 <- MSM_BivData[MSM_BivData$hand.foci.count == 2,]
-#MSM_BivDatam_CO2 <- MSM_BivDatam_CO2[!(is.na(MSM_BivDatam_CO2$fileName) | MSM_BivDatam_CO2$fileName==""), ]
-
-#MSM_BivDatam_CO2$Foci1.PER <- MSM_BivDatam_CO2$Foci1 /MSM_BivDatam_CO2$chromosomeLength
-#MSM_BivDatam_CO2$Foci2.PER <- MSM_BivDatam_CO2$Foci2 /MSM_BivDatam_CO2$chromosomeLength
-
-MSM_BivData <- add_IFD(MSM_BivData)
-#add IFD2 is f-ing up 
-MSM_BivData <- addIFD.2(MSM_BivData)
-
-#this MSM PER.IFD, is similar to the PWD data
-
-#new function is converting all hand.foci counts to 3
-#it might be the IFD functions
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-new_shiny <- add_SisCoTen(MSM_BivData)
-
-new_shiny$rel_Siscoten <- new_shiny$SisCoTen / new_shiny$chromosomeLength
-
-qq <- ggplot(data = new_shiny, aes(y=chromosomeLength, x=rel_Siscoten) ) + geom_jitter()
-
-#ww <- ggplot(data = MSM_BivDatam_CO2, aes(y=IFD1_PER, x=mouse, color = fileName) ) + geom_jitter()+
-#  geom_boxplot(aes(alpha=.3))
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-#KAZ and PWD data set is larger
-PWD_KAZ_BivData.og = read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD.KAZ_BivData.csv", header = TRUE, strip.white = TRUE )
-
-PWD_KAZ_BivData <- PWD_KAZ_BivData.og[PWD_KAZ_BivData.og$SC.pass == 1,]
-
-PWD_KAZ_BivData <- add_IFD(PWD_KAZ_BivData)
-PWD_KAZ_BivData <- addIFD.2(PWD_KAZ_BivData)
-PWD_KAZ_BivData <- add_SisCoTen(PWD_KAZ_BivData)
-PWD_KAZ_BivData$rel_Siscoten <- PWD_KAZ_BivData$SisCoTen / PWD_KAZ_BivData$chromosomeLength
-PWD_KAZ_BivData <- add_mouse(PWD_KAZ_BivData)
-PWD_KAZ_BivData <- add_strain(PWD_KAZ_BivData)
-
-mouse_boxes <- ggplot(data = PWD_KAZ_BivData, aes(y=rel_Siscoten, x=mouse) ) + geom_jitter(aes( color=fileName))+
-  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")
-
-mouse_boxes2 <- ggplot(data = PWD_KAZ_BivData, aes(y=SisCoTen, x=mouse) ) + geom_jitter(aes( color=fileName))+
-  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")
-
-#mean relative and raw sisco-ten is higher in PWD compared to KAZ
-
-biv_boxes <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count, fill=hand.foci.count) ) + geom_jitter()+geom_boxplot(aes(x= hand.foci.count,alpha=.3))+
-theme(legend.position="none")+facet_wrap(~strain)
-
-biv_boxes22 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=as.factor(hand.foci.count)) )+ geom_jitter()+
-  geom_boxplot(aes(alpha=.3))+theme(legend.position="none")+facet_wrap(~strain)
-
-biv_boxes2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=rel_Siscoten, x=hand.foci.count) ) + geom_jitter(aes( ))+
-theme(legend.position="none")+facet_wrap(~strain)
-
-regress <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=chromosomeLength) ) + geom_jitter(aes( color=as.factor(hand.foci.count) ))+ facet_wrap(~strain)
-
-regress.1 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
-#, 1mar15_PWD_m1
-
-regress.2 <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+ theme(legend.position="none")+ facet_wrap(~strain)
-#, 1mar15_PWD_m1
-
-#
-
-regress.1.KAZ <- ggplot(data = PWD_KAZ_BivData[PWD_KAZ_BivData$mouse == "21may18_KAZ_m2",], aes(y=SisCoTen, x=hand.foci.count) ) + geom_jitter(aes( color=fileName))+
- theme(legend.position="none")+ facet_wrap(~fileName)
-
-#doesn't look like the sis-co-ten is conserved across bivs
-
-```
-
-```{r within.cell.corr, echo=FALSE}
-
-#test telomere distance
-#centromere distance
-#SC length
-
-#I want to test if SC lengths within cells cluster
-#wouldn't ANOVA get at this..?
-
-#cell level variance is significant predictor for SC length (telomere)
-
-#PWD_KAZ_BivData[PWD_KAZ_BivData$strain != "other",]
-jjgg <- ddply(.data= PWD_KAZ_BivData,
-                 .(fileName),
-                 summarize, 
-                 nbiv = length(unique(Obj.ID) ),
-                  var.sc = var(chromosomeLength)
-)
-
-#reasons why this within cell correlation doesn't work well with bivalent data
-  #different sets of chrms
-#not enough resolution at cytology level
-
-```
-
-sis-coten metric:
-I am currently missing a null model / test for this metric
-(sis-co-ten from uniform placement of COs, and (some prob of draw for CO number))
-
-there is a positive correlation with siscoten and chrm length
-
-there could be something about kinetochore sites differing across specific chromosomes that balance the force

### 2019-12-27 doc/Main_Report_Results.rmd
re-starting effect quan measurements

@@
 
-## Variance in CO count
+# Variance in CO count
 
@@
 Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
-
-I can show the variance boxplots?  Not sure if it's very helpful.
-Use the same Datasets HQ and mol for the mean CO counts. Use both var and cV as dependant variables
-
-DF.HetC.MixedModel.HQ.mol
-DF.HetC.MixedModel.HQ
 
@@
 ```{r variance.hist, echo=FALSE}
+
+#make series for total dataset?
+Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("Histograms of within mouse variance, all mice")
+Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+Var.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("Histograms of within mouse variance, 10 cell min")
+Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
+Var.Hist
+
+Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
+Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
+Var.Hist.15cells
+
+```
+
+A general pattern is that the female data has higher within mouse variance (both var and cV).
+I think the dissection date needs to be tested as a predictor factor.
+It might be tricky to add the dissection date metric.
+Test if dissection date is more significant predictor than strain (for females)
+
+(LEW females with top 3 variance are all from different dissection dates)
+
+```{r cv.plots, echo=FALSE}
+
+cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("cV of within mouse variance, all mice")
+cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+cV.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("cV of within mouse variance, 10 cell min")
+cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
+cV.Hist
 
@@
 
-```{r DMC1.subsampling, echo=FALSE}
+```{r DMC1.subsampling, echo=FALSE, eval=FALSE}
 

### 2019-12-30 doc/figureFiles/figures.Rmd
working on figure for Bret, started vert figure

@@
 
-# Option 1, strain averages, sexes matched, cells pooled
+# Option vertical facets
 
@@
 
+```{r vertical.facet, echo=FALSE, fig.height= 13}
+
+Vert.facet <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
+  #facet_grid(~sex, scales = "free", space="free")+ 
+  ylab("Number of Crossovers per Cell")+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+
+   facet_wrap(~sex, ncol = 1)+ #, scales = "free", space="free"
+
+   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(17.5, 33))+
+ scale_color_manual(values=colors_of_strains)+ 
+  scale_y_continuous(breaks=seq(20,32,2) )+
+
+  #legend.position="none",
+  theme(legend.position="bottom",legend.box = "horizontal",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+
+#remove y labfor male 
+  #Dom
+  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
+
+  #Musc
+  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
+  #mol
+  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)
+
+   #label -- segments applied to both facets, labels not?
+#label annotation not working, add in illustrator  
+annotate("text", x = c(2.2, 4.5, 7,
+                         2.2, 4.5, 7), #list of all x points 
+           y = c(20.2,20.2,20.2,
+                 20.2,20.2,20.2),
+   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus",
+             "M. m. domesticus","M. m. musculus", "M. m. molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+  Vert.facet
+
+```
+
+```{r new.save, echo=FALSE}
+#Fig1_Bv2_strain
+#save a tif or png file
+#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
+```
+
+First run of the vertica facets looks awful and squished. fix it.
+
+# Option 1, strain averages, sexes matched, cells pooled
+
 ```{r strain.table.plot.v2, echo=FALSE, warning=FALSE}
@@
   ylab("Number of Crossovers per Cell")+
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+  geom_point(aes(shape = sex), size=3.7, position=position_dodge(.5))+
 
@@
 #save a tif or png file
-ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
+#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
 ```

### 2019-12-30 doc/Effects.Rmd
working on effect docs

@@
 
+```{r setupdata, echo=FALSE}
+AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
+#remove NAs
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]
+
+```
+
 ## Quality Effects
@@
 
-```{r MLH1.histogram.2, echo=FALSE}
+```{r MLH1.histogram.2, echo=FALSE, include=FALSE}
 #remake the pooled MLH1 histgoram plot from above but facet by quality
@@
 
-I like this plot and find it effective. Can I do a similar set up for other factors (quality, ncells?)
+I made 3 models for 3 groupings of mice, dom, musc.high and musc.low
 
-```{r age.effect.model, include=FALSE}
+$$mouse .mean CO ~=~ age + strain + \varepsilon$$
+
+```{r age.effect.model.setupdata, echo=FALSE}
 #run a age effect linear regression!
-
 male.data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.y == "male",]
 
+#remove all nas
+
+#unorder the factors
+class(male.data$strain.x) <- "factor"
+class(male.data$subsp.x) <- "factor"
+
 male.data.dom <- male.data[male.data$subsp.y == "Dom",]
+male.data.dom <- male.data.dom[!(is.na(male.data.dom$Nmice) | male.data.dom$Nmice==""), ]
 
-male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST"),]
+#MOLF is a low strain
+male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST" | male.data$strain.x == "MOLF"),]
+male.data.musc.low <- male.data.musc.low[!(is.na(male.data.musc.low$Nmice) | male.data.musc.low$Nmice==""), ]
 
-male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "MOLF" | male.data$strain.x == "SKIVE"),]
+male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "SKIVE"),]
+male.data.musc.high <- male.data.musc.high[!(is.na(male.data.musc.high$Nmice) | male.data.musc.high$Nmice==""), ]
 
 male.data.outgroup <- male.data[(male.data$subsp.y == "Spretus" | male.data$subsp.y == "Spic" | male.data$subsp.y == "Caroli"),]
+male.data.outgroup <- male.data.outgroup[!(is.na(male.data.outgroup$Nmice) | male.data.outgroup$Nmice==""), ]
+
+```
+
+```{r age.models, echo=FALSE}
 
@@
 
-summary(lm.me.OG)
-#for all models, mostly it was that the intercept was significant, and mot much else
+#summary(lm.me.OG)
+as.data.frame( round(  summary(lm.me.dom)$coefficients, 6 ) )
 
-#good knitting output
-#as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )
+as.data.frame( round(  summary(lm.me.musc.low)$coefficients, 6 ) )
 
+as.data.frame( round(  summary(lm.me.musc.high)$coefficients, 6 ) )
 ```
+
+Dom, no significant effects for age.
+
+Musc low, no significant effect for strain or age.
+
+Musc high, MSM and SKIVE have significant age effects on mean CO count
+
+SKIVE has the greatest spread in ages for single mice (7,16,41 weeks)
+There are more SKIVE males which I haven't quantified (at least 3, which are in a good age range)
 
@@
 
-AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
+#AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
 #remove NAs
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]
+#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | #AP_mouse_table_w.Ages$Nmice==""), ]
 
@@
 ```
+
+I like this plot and find it effective. Can I do a similar set up for other factors (quality, ncells?)
 
@@
 
+```{r dissection.effect.data.setup, echo=FALSE}
+
+#AP_mouse_table_w.Ages$euth.date
+
+```
+
 # Spread and Stain Effects

### 2019-12-30 doc/Effects.Rmd
working on Effect doc

@@
 load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")
+
 #load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
@@
 
-## Quality Effects
+# Quality Effects
 
@@
  # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
+```
+
+```{r qual.tables, echo=FALSE}
+
+#make a similart plot as the ages
+#for each mouse and quality, calq the mean MLH1
+mouse.table.mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
+                 .(mouse, quality),
+                 summarize, 
+                 ncells =  length(unique(Original.Name)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+mouse.table.mean.MLH1.obs_qual <- add_category(mouse.table.mean.MLH1.obs_qual)
+mouse.table.mean.MLH1.obs_qual <- add_strain(mouse.table.mean.MLH1.obs_qual)
+mouse.table.mean.MLH1.obs_qual <- add_sex(mouse.table.mean.MLH1.obs_qual)
+mouse.table.mean.MLH1.obs_qual <- add_subsp(mouse.table.mean.MLH1.obs_qual)
 
@@
 
-I'm not sure sure how effective the above plots are.
+```{r qual.plot, echo=FALSE}
 
-### Age Effect (male)
+mm <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "male",]
+age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
+age.effect.plot.mm
+
+ff <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "female",]
+age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female quality and mean CO")+facet_wrap(~subsp)
+
+age.effect.plot.ff
+
+```
+
+The original plots weren't very effective, so I removed them. I made
+
+## Quality effect on SC
+
+```{r SC.qual, echo=FALSE}
+#use the SC table
+
+#all of skel and bin are not num
+MLH1.merge.Skel$skel_size <- as.character(MLH1.merge.Skel$skel_size)
+MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
+
+MLH1.merge.Skel$bin_size <- as.character(MLH1.merge.Skel$bin_size)
+MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
+
+#mouse.level.skel
+
+#MLH1.merge.Skel, hist of all skel values
+
+#test with removing crop == 'no', MLH1.merge.Skel[MLH1.merge.Skel$REDO.crop == "no",]
+mouse.table.mean.skel.obs_qual <- ddply(.data=MLH1.merge.Skel, 
+                 .(mouse, quality),
+                 summarize, 
+                 ncells =  length(unique(Original.Name)),
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                 mean.bin = mean(bin_size)
+                # var.skel = var(skel_size, na.rm = TRUE) #var not useful, many NA/1 obs
+)
+
+mouse.table.mean.skel.obs_qual <- add_category(mouse.table.mean.skel.obs_qual)
+mouse.table.mean.skel.obs_qual <- add_strain(mouse.table.mean.skel.obs_qual)
+mouse.table.mean.skel.obs_qual <- add_sex(mouse.table.mean.skel.obs_qual)
+mouse.table.mean.skel.obs_qual <- add_subsp(mouse.table.mean.skel.obs_qual)
+
+```
+
+```{r SC.hitograms, echo=FALSE}
+
+skel.histgram <- ggplot(MLH1.merge.Skel, aes(x=skel_size)) +  geom_histogram(bins=50)
+skel.histgram
+
+length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells
+
+MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
+
+MLH1.merge.Skel <- MLH1.merge.Skel[MLH1.merge.Skel$skel_size < 30000,]
+
+skel.histgrams.2.DOM <- ggplot(MLH1.merge.Skel[MLH1.merge.Skel$subsp == "Dom",], aes(x=skel_size)) +  geom_histogram(bins=50)+facet_grid(sex~strain)
+skel.histgrams.2.DOM
+
+skel.histgrams.2.MUSC <- ggplot(MLH1.merge.Skel[MLH1.merge.Skel$subsp == "Musc",], aes(x=skel_size)) +  geom_histogram(bins=50)+facet_grid(sex~strain)
+skel.histgrams.2.MUSC
+
+```
+
+```{r sc.qual.plots, echo=FALSE}
+skel.mm <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "male",]
+skel.mm <- skel.mm[skel.mm$mean.skel < 50000,]
+
+skel.qual.effect.plot.mm <- ggplot(skel.mm, aes(y=mean.bin, x=quality, color=(strain) ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
+skel.qual.effect.plot.mm
+```
+
+Each point is a single mouse
+
+```{r G.plot, echo=FALSE}
+Bin.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
+                           aes(y=mean.bin, x=quality, size= ncells, color=strain))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Quality and mean SC (bin) ")+facet_wrap(~mouse)
+
+Bin.qual.effect.plot.mm.G
+
+skel.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
+                           aes(y=mean.skel, x=quality, size= ncells, color=strain))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Quality and mean SC (bin) ")+facet_wrap(~mouse)
+
+```
+
+The mouse 22jun15_G_m2 has strange pattern, investigate these outputs
+
+Some of the pattern goes away when redo.crop is set to equal 'no'.
+
+```{r female.skel.plot, echo=FALSE}
+
+skel.ff <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "female",]
+
+skel.qual.effect.plot.ff <- ggplot(skel.ff, aes(y=mean.bin, x=quality, color=(strain) ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
+
+skel.qual.effect.plot.ff
+
+skel.qual.effect.plot.ff.MSM <- ggplot(skel.ff[skel.ff$strain == "MSM",], aes(y=mean.bin, x=quality, color=(strain), size=ncells ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)
+
+skel.qual.effect.plot.ff.G <- ggplot(skel.ff[skel.ff$strain == "G",], aes(y=mean.bin, x=quality, color=(strain), size=ncells ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)
+
+skel.qual.effect.plot.ff.PWD <- ggplot(skel.ff[skel.ff$strain == "PWD",], aes(y=mean.skel, x=quality, color=(strain), size=ncells ))+geom_point()+ 
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)
+
+```
+
+The above plots are good for detecting outliers
+
+# Age Effect (male)
 
@@
 
+ADD euthdate?
+
 Dom, no significant effects for age.
@@
 
-### Age Effect (female)
+# Age Effect (female)
 
@@
 
-### Maternal Age effect
+# Maternal Age effect
 
@@
 
+mouse.table.female <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "female",]
+
+mouse.table.female <- mouse.table.female[!(is.na(mouse.table.female$mouse) | mouse.table.female$mouse==""), ]
+
+mouse.table.female <- mouse.table.female[!(is.na(mouse.table.female$euth.date) | mouse.table.female$euth.date==""), ]
+
+#add euth dates to MLH1_data
+test.this <- MLH1_data[ MLH1_data$mouse %in% mouse.table.female$mouse, ]
+
+mlh1.euth.date.female <- merge(mouse.table.female, MLH1_data, by = "mouse")
+
 ```
 
-# Spread and Stain Effects
+```{r plot.dissection, echo=FALSE}
+
+male.data$euth.date <- as.Date(male.data$euth.date, "%m/%d/%Y")
+
+male.euth.date <- ggplot(male.data, aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
+  #trying this
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
+male.euth.date
+
+#look at musc
+musc.male.euth.date <- ggplot(male.data[male.data$subsp.x == "Musc",], aes(y=mean_co, x=euth.date, color=(age.weeks) ))+geom_point()+ ylim(c(19,34))+
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Musc Male Age and mean CO")+facet_wrap(~strain.x)
+musc.male.euth.date
+
+```
+
+```{r dissection.female, echo=FALSE}
+
+female.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "female") & (AP_mouse_table_w.Ages$category.y != "other"),]
+female.data$euth.date <- as.Date(female.data$euth.date, "%m/%d/%Y")
+
+female.euth.date.plot <- ggplot(female.data[female.data$subsp.x == "Musc",], aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
+
+female.euth.date.plot
+
+DOM.female.euth.date.plot <- ggplot(female.data[female.data$subsp.x == "Dom",], aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Age and mean CO")+facet_wrap(~subsp.y)
+
+DOM.female.euth.date.plot
+
+#female / male, sex.x
+#mouse.table.male <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "male",]
+
+test.this <- MLH1_data[ MLH1_data$mouse %in% mouse.table.female$mouse, ]
+
+mlh1.euth.date.female <- merge(mouse.table.female, MLH1_data, by = "mouse")
+
+```
+
+## Total SC
+
+follow up to peromyscus findings, Where summer caught mice had lower SC area per cell, compared to winter and fall caught. But there were few data points.
+
+```{r total.sc, echo=FALSE}
+#check out effects of month dissected on SC
+
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+
+#from total SC
+
+#merge
+ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")
+
+ll$euth.date.2 <- as.Date(ll$euth.date, "%m/%d/%Y")
+
+ll <- ll[!(is.na(ll$mean.skel) | ll$mean.skel==""), ]
+ll <- ll[!(is.na(ll$sex.y) | ll$sex.y==""), ]
+
+library(lubridate)
+
+#ADD month 
+ll$month.2 <- month(ll$euth.date.2)
+
+#mouse level table
+
+ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain.y) ))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)
+
+ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain.y) ))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)
+
+ll.month.plot
+
+```
+
+```{r total.sc.all.cells, echo=FALSE, include=FALSE}
+#try ploting all cells
+
+#uu <- merge(MLH1.merge.Skel,AP_mouse_table_w.Ages,by="mouse"  )
+
+#uu$euth.date.2 <- as.Date(uu$euth.date, "%m/%d/%Y")
+#uu$month.2 <- month(uu$euth.date.2)
+
+#uu <- uu[!(is.na(uu$skel_size) | uu$skel_size==""), ]
+#uu <- uu[!(is.na(uu$sex.x) | uu$sex.x==""), ]
+
+#uu.month.plot.DOM <- ggplot(uu[uu$subsp.x == "Dom",], aes(y=skel_size, x=month.2, color=(mouse) ))+
+#  geom_jitter()+  theme(legend.position="none")+
+#  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+
+#  facet_wrap(~sex.x)
+
+```
+
+So, there doesn't seem to be an effect of month dissected on SC area across mice.
+This might be the start of testing for 'technical effects / humidity effects'.
+BUT, I don't really have month by month dissections of mice of the same age, ect.
 
@@
 
+# Spread and Stain Effects
+
 > *staining batches*
 
-Compare observations from the same mouse, but different MLH1 stains.
+Compare observations from the same mouse, but different MLH1 stains.  First need to identify data that can be used.
 
-```{r stain.effects, eval=FALSE}
-
+```{r stain.effects.SC, eval=FALSE}
+#/src/TotalSC.R
 #compile this dataset

### 2019-12-30 doc/figureFiles/figures.Rmd
outlining for final figure

@@
 
-# Option vertical facets
-
 ```{r strain.table.v2, echo=FALSE, warning=FALSE}
@@
 
-```{r vertical.facet, echo=FALSE, fig.height= 13}
+# FINAL DRAFT
 
-Vert.facet <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
-  #facet_grid(~sex, scales = "free", space="free")+ 
-  ylab("Number of Crossovers per Cell")+
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+put together 3 pieces; plot, cell1, cell2
 
-   facet_wrap(~sex, ncol = 1)+ #, scales = "free", space="free"
+ todo
 
-   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+   1. finish and import final ai edited polt,
 
-  ylim(c(17.5, 33))+
- scale_color_manual(values=colors_of_strains)+ 
-  scale_y_continuous(breaks=seq(20,32,2) )+
+   2. add cell images with cowplot to bottom
 
-  #legend.position="none",
-  theme(legend.position="bottom",legend.box = "horizontal",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
-
-#remove y labfor male 
-  #Dom
-  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
-
-  #Musc
-  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
-  #mol
-  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)
-
-   #label -- segments applied to both facets, labels not?
-#label annotation not working, add in illustrator  
-annotate("text", x = c(2.2, 4.5, 7,
-                         2.2, 4.5, 7), #list of all x points 
-           y = c(20.2,20.2,20.2,
-                 20.2,20.2,20.2),
-   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus",
-             "M. m. domesticus","M. m. musculus", "M. m. molossinus"),
-           fontface = 'italic', hjust = 0.5, size=3.5)
-
-  Vert.facet
+```{r load.plot, echo=FALSE}
 
@@
 
-```{r new.save, echo=FALSE}
-#Fig1_Bv2_strain
-#save a tif or png file
-#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
+```{r load.cells, echo=FALSE}
+
+cell.output <- readPNG('C:/Users/alpeterson7/Documents/WildMiceMadison/Images/whole_cell_example.png')
+#make empty DF for ggplot object
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+#make an empty ggplot thing, then background ontop
+spermatocyte.auto <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(cell.output) + theme(axis.title.x=element_blank(),
+                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
 ```
 
-First run of the vertica facets looks awful and squished. fix it.
+```{r cow.plot, echo=FALSE}
+
+library("cowplot")
+fig1  <- ggdraw() +
+  draw_plot(MLH1.fig1, x = 0, y = .42, width=1, height=.55) +
+
+  draw_plot(Better_MAD_map, x=0, y=0, width=.4, height=.4) +
+
+  draw_plot(spermatocyte, x=.38, y = 0, width=.31, height=.39) +
+
+  draw_plot(spermatocyte.auto, x=.69, y = 0, width = .31, height = .39) +
+
+  draw_plot_label(label = c("A", "B", "C"), size = 15,
+                  x = c(0, 0, .36), y = c(1, .43, 0.43) )
+
+fig1
+
+ggsave("~./WildMiceMadison/results/Figure1_32119.png")
+```
 
@@
 S.E. = sd( all cells) / n.cells
+
+# Option vertical facets
+
+```{r vertical.facet, echo=FALSE, fig.height= 13}
+
+Vert.facet <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
+  #facet_grid(~sex, scales = "free", space="free")+ 
+  ylab("Number of Crossovers per Cell")+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+
+   facet_wrap(~sex, ncol = 1)+ #, scales = "free", space="free"
+
+   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(17.5, 33))+
+ scale_color_manual(values=colors_of_strains)+ 
+  scale_y_continuous(breaks=seq(20,32,2) )+
+
+  #legend.position="none",
+  theme(legend.position="bottom",legend.box = "horizontal",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+
+#remove y labfor male 
+  #Dom
+  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
+
+  #Musc
+  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
+  #mol
+  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)
+
+   #label -- segments applied to both facets, labels not?
+#label annotation not working, add in illustrator  
+annotate("text", x = c(2.2, 4.5, 7,
+                         2.2, 4.5, 7), #list of all x points 
+           y = c(20.2,20.2,20.2,
+                 20.2,20.2,20.2),
+   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus",
+             "M. m. domesticus","M. m. musculus", "M. m. molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+  Vert.facet
+
+```
+
+```{r new.save, echo=FALSE}
+#Fig1_Bv2_strain
+#save a tif or png file
+#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
+```
 

### 2019-12-31 doc/figureFiles/figures.Rmd
wraped up Bret's suggestions for figure

--- 2019-12-30 doc/figureFiles/figures.Rmd
@@
 
+# Option 1, strain averages, sexes matched, cells pooled
+
+learning that geom_point(position = position_doge(x))
+ will shift the two shaped points closer together.  But this isn't the same as the x axis categories being shifted closer together.
+
+So trying to learn how to squish the x axis together
+ scale_x_discrete(expand=c(0, 2) ) squishes all the tick marks, but makes a bunch of big space at the boarders
+
+The parameters in expand() -- are connected with the data range. -- expand is used for making padding around the data.  So this might not be the exactly what I am looking for.
+It might be that the best way is to make a stand in continous x axis scale for nudging the point positions.
+
+Removing legends, selectively
+
+has to do with which variables are defined within aes(), not sure how to define shape but not color within aes
+
+the color will now be by sex... I could do this in Illustrator
+
+```{r main.plot.code, echo=FALSE, warning=FALSE}
+#use the MLH1 / cell level data to calq the error bars
+#strain table ggplot
+
+Fig1_Bv2_strain <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain,shape = sex, color=strain),   show.legend = FALSE)+
+
+    geom_point(aes(shape = sex),  size=3.7, position=position_dodge(.29))+
+
+   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.29), size=1.001)+
+
+  scale_color_manual(values=colors_of_strains)+
+
+  ylab("Number of Crossovers per Cell")+
+  scale_y_continuous(breaks=seq(20,32,2) )+
+
+  scale_x_discrete(expand=c(.08, 0) )+  
+  #scale_x_continuous( breaks=c(1,2), limits=c(0,3), expand=c(0,0))+
+
+  theme_classic()+
+  theme(legend.position="none",     #legend.position="bottom",legend.box = "horizontal",
+        axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank() )+
+ #remove all legend for now....
+
+  #  panel.grid.major = element_blank(), panel.grid.minor = #element_blank(),  axis.line=element_line(colour = "black"),
+ # panel.background = element_blank()) +
+
+#remove y labfor male 
+  #Dom
+  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
+  #Musc
+  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
+  #mol
+  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)+
+  #label
+  annotate("text", x = c(2.2, 4.5, 7), #list of all x points 
+           y = c(20.2,20.2,20.2),
+   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+#this can pull just the lengend object
+#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
+#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
+```
+
+```{r Strain.fig.merge.v2, echo=FALSE}
+Fig1_Bv2_strain
+#save a tif or png file
+#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
+```
+
+Error bars are 2 SE,
+
+S.E. = sd( all cells) / n.cells
+
+# FINAL DRAFT
+
+put together 3 pieces; plot, cell1, cell2
+
+ todo
+
+   1. finish and import final ai edited polt,
+
+   2. add cell images with cowplot to bottom
+
+```{r load.plots, echo=FALSE}
+library(png)
+library(ggpubr)
+
+F.first.png <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/MLH1_F.first.plot_v4.png')
+M.first.png <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/MLH1_M.first.plot_v4.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+F.first <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(F.first.png) + theme(axis.title.x=element_blank(),
+                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+M.first <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(M.first.png) + theme(axis.title.x=element_blank(),
+                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
+```{r load.cells, echo=FALSE}
+
+best.oo <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/cell.images/Best_female_4apr16_17mar16_G_f3_sp1_26_rev.png')
+best.spermato <- readPNG('C:/Users/alpeterson7/Documents/MLH1repo/doc/figureFiles/cell.images/Best_male_9mar16_22feb16_G_m1_sp1_15.2_rev.png')
+
+#make empty DF for ggplot object
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+#make an empty ggplot thing, then background ontop
+spermatocyte.auto <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(best.spermato) + theme(axis.title.x=element_blank(),
+                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+oocyte.auto <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(best.oo) + theme(axis.title.x=element_blank(),
+                                  axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
+# all png loaded in
+
+```{r cow.plot, echo=FALSE}
+library("cowplot")
+
+F.first.fig  <- ggdraw() +
+  draw_plot(F.first, x = 0, y = .42, width=1, height=.5) +
+
+  draw_plot(spermatocyte.auto, x=.51, y = 0, width=.31, height=.39) +
+
+  draw_plot(oocyte.auto, x=0.01, y = 0, width = .31, height = .39) +
+
+  draw_plot_label(label = c("A", "B", "C"), size = 15,
+                  x = c(0, 0, .5), y = c(1, .43, 0.43) )
+
+F.first.fig
+
+#ggsave("~./WildMiceMadison/results/Figure1_32119.png")
+```
+
+Loaded slightly nudged points-plot as png. This looks bad, don't do this. Switch to Illustrator.
+
+# redo
+
+```{r redo.plot, echo=FALSE}
+redo  <- ggdraw() +
+
+  draw_plot(Fig1_Bv2_strain, x = 0, y = .42, width=1, height=.49) +
+
+  draw_plot(spermatocyte.auto, x=.55, y = 0, width=.4, height=.41) +
+
+  draw_plot(oocyte.auto, x=0.05, y = 0, width = .4, height = .41) +
+
+  draw_plot_label(label = c("A", "B", "C"), size = 15,
+                  x = c(0, 0, .5), y = c(1, .43, 0.43) )
+
+redo
+```
+
+All pieces are 'made' within R.
+But I can't figure out how to reduce the x axis white space between points (strains).
+The message boards seem to indicate that using a continous_x scale might be the only way to get the layout I want.
+
+# Option vertical facets
+
 ```{r vertical.facet, echo=FALSE, fig.height= 13}
@@
 ```
-
-First run of the vertica facets looks awful and squished. fix it.
-
-# Option 1, strain averages, sexes matched, cells pooled
-
-```{r strain.table.plot.v2, echo=FALSE, warning=FALSE}
-#use the MLH1 / cell level data to calq the error bars
-#strain table ggplot
-#remove facet by sex
-
-Fig1_Bv2_strain <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
-  #facet_grid(~sex, scales = "free", space="free")+ 
-  ylab("Number of Crossovers per Cell")+
-  geom_point(aes(shape = sex), size=3.7, position=position_dodge(.5))+
-
-   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(19.5, 33))+
- scale_color_manual(values=colors_of_strains)+ 
-  scale_y_continuous(breaks=seq(20,32,2) )+
-
-  #legend.position="none",
-  theme(legend.position="bottom",legend.box = "horizontal",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
-
-#remove y labfor male 
-  #Dom
-  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
-  #Musc
-  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
-  #mol
-  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)+
-  #label
-  annotate("text", x = c(2.2, 4.5, 7), #list of all x points 
-           y = c(20.2,20.2,20.2),
-   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus"),
-           fontface = 'italic', hjust = 0.5, size=3.5)
-
-#this can pull just the lengend object
-#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
-#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
-```
-
-```{r Strain.fig.merge.v2, echo=FALSE}
-Fig1_Bv2_strain
-#save a tif or png file
-#ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
-```
-
-Error bars are 2 SE,
-
-S.E. = sd( all cells) / n.cells
 

### 2019-12-31 doc/Main_Report_Results.rmd
forgot what updates were

@@
 A general pattern is that the female data has higher within mouse variance (both var and cV).
+Some strains more than others.
 I think the dissection date needs to be tested as a predictor factor.
@@
 Test if dissection date is more significant predictor than strain (for females)
+Also test how the number of 0CO bivalents per cell affects mouse with variance.
 
 (LEW females with top 3 variance are all from different dissection dates)
+
+I don't have enough mice with >15 cells for a complete idea of the distributions, but comparing single mice with enough good quality cells show be
 

### 2020-01-02 doc/Effects.Rmd
Thinking about Effects and such

@@
 ```{r setup, include=FALSE}
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo/") #this wd is different than where the file is ... this kinda causes errors
 #setwd("C:/Users/April/Desktop/MLH1repo")
@@
 library(reshape2)
+library(lubridate)#for dates / months
 
@@
 load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")#total SC data
 
@@
 #save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
-
-```
-
-```{r Effects.on.the.dataset, eval=FALSE, echo=FALSE}
-
-#*age effects* (male, maternal age)
-
-#*litter effect*
-
-#*mouse room effect* (Charmony, MSC)
-
-#*quantification repeatability*
-
-#*spread batch and staining effects*
-
-#*power analysis on number of cells*
-
-#*time of year effect (on spread)*
-
-#*dominant effect* (F1s)
-
 ```
@@
 
-# Quality Effects
+# Cell Quality Effects
 
-below is a chunk to looking at histogram/distribution and means for the MLH1 data across quality bins
+I am testing the quality effects, since it could be a source of error and potentially biased the predicted gwRR if some mice had bad technical noise.
+
+The the plots below, each point is a single mouse's mean MLH1 count per cell across each quality score bin.
 
 ```{r MLH1.histogram.2, echo=FALSE, include=FALSE}
+#thse plots not shown
 #remake the pooled MLH1 histgoram plot from above but facet by quality
@@
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)+
+  theme(legend.position="none")
+
 age.effect.plot.mm
 
-ff <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "female",]
+ff <- mouse.table.mean.MLH1.obs_qual[(mouse.table.mean.MLH1.obs_qual$sex == "female"),]
+ff <- ff[ff$strain != "CAST",]
+
 age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female quality and mean CO")+facet_wrap(~subsp)
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+
+  #are the lines
+  ggtitle("Female quality and mean CO")+facet_wrap(~subsp)+theme(legend.position="none")
 
@@
 
-The original plots weren't very effective, so I removed them. I made
+These plots show that generally there is a positive relationship between good quality (1) and more MLH1 counts.
 
-## Quality effect on SC
+I should focus on addressing if there are differences in the relationship of quality-MLH1 counts across sexes.
+
+## Cell Quality effect on SC
+
+Examine image quality or other scores to test how well the python scripts segment the SC area for all the images.
 
@@
 
-```{r SC.hitograms, echo=FALSE}
+```{r SC.hitograms, echo=FALSE, fig.width = 4, fig.height=.1}
+#try adjusting the size for thi..
 
@@
 
-length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells
+#length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells
 
@@
 
+Testing the effects of mean total SC per mouse with when they were dissected.
+
 follow up to peromyscus findings, Where summer caught mice had lower SC area per cell, compared to winter and fall caught. But there were few data points.
@@
 #check out effects of month dissected on SC
-
 #load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
-
 #from total SC
@@
 #merge
-ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")
-
+ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")  #no
 ll$euth.date.2 <- as.Date(ll$euth.date, "%m/%d/%Y")
+#ncells is from the full MLH1 DF
+#Ncells is from the cleaned, HQ DF
 
@@
 
-library(lubridate)
+ll <- subset(ll, select = -c(strain.x,strain.y, category.x, category.y, sex.x, sex.y, subsp.x, subsp.y, mouse.1, nmice) )
 
-#ADD month 
+#ADD month number
 ll$month.2 <- month(ll$euth.date.2)
 
-#mouse level table
+ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain) ))+geom_point()+
+ scale_color_manual(values=colors_of_strains)+
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex)
 
-ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain.y) ))+geom_point()+
+ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain) ))+geom_point()+
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)
-
-ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain.y) ))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+
+  ggtitle("")+facet_wrap(~sex)
 

### 2020-01-02 doc/Effects.Rmd
changed plot to better see spread of points

@@
 mm <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "male",]
-age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
+age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_jitter(width = 0.16)+ 
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata
@@
 
-age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
+age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_jitter(width = 0.16)+ 
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata

### 2020-01-06 doc/Effects.Rmd
first try of incorporating batch18, still a few bugs

@@
 
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
 load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")#total SC data
@@
 
-```{r SC.hitograms, echo=FALSE, fig.width = 4, fig.height=.1}
+```{r SC.hitograms, echo=FALSE, fig.width = 4, fig.height=1}
 #try adjusting the size for thi..
 
-skel.histgram <- ggplot(MLH1.merge.Skel, aes(x=skel_size)) +  geom_histogram(bins=50)
+skel.histgram <- ggplot(MLH1.merge.Skel, aes(x=skel_size)) +  geom_histogram(bins=50)+ggtitle("width 4, height 1")
 skel.histgram
+```
+
+```{r more.SC.hist, echo=FALSE}
 

### 2020-01-06 doc/Main_Report_Results.rmd
first try of incorporating batch18, still a few bugs

@@
 
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
 #load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice

### 2020-01-06 src/SetupData.Rmd
first try of incorporating batch18, still a few bugs

@@
 ```{r MLH1.DF.setup, warning=FALSE, echo=FALSE}
-
-#todo :
-#set the order to batches
-#2 fileNames from batch 0 appear 2x, (1 is missing random name, the other has the same random name)
 
@@
 MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]
-
 ```
@@
 ```{r adj_nMLH1, echo=FALSE}
-
 #add a column with male adjusted MLH1 values (+1 to all males)
@@
 #order the batches in MLH1 and original
-
 ```
@@
 batch.stats <- merge(original_batches,final_batches, by="BatchNum")## mm is the table of batch stats
-
+##batch18 has the wrong number, ~390
 ```
@@
 
+#remove the NAs
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$category.x) | AP_mouse_table_w.Ages$category.x==""), ]
 ```
@@
 ```
+
+update the dissection document (that makes the meta.data DF)
 
@@
 
-```{r Age.Hists, echo=FALSE, warning=FALSE, eval=FALSE}
+```{r Age.Hists, echo=FALSE, warning=FALSE}
 #age.weeks col is missing
 
-male_mouse_table_w.Ages <- AP_mouse_table.HQ[(AP_mouse_table.HQ$sex == "male") &(mouse.table_w.Ages$category != "other"),]
+male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") &(AP_mouse_table_w.Ages$category.x != "other"),]
 
-all.ages_plot <- ggplot(AP_mouse_table.HQ, aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
+all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category.x)+ggtitle("Male Age Distributions")
 all.ages_plot
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")
 

### 2020-01-06 doc/MouseLevelPlots.Rmd
mouse level plots with batch18, add 3 mice to exclude list. updated data setup

@@
 #load main data file
-load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
 #load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16

### 2020-01-06 src/setupdata_BivData_MLH1repo.R

@@ -6,10 +6,12 @@ library(plyr)
 library(lattice)
 library(dplyr)
 library(ggplot2)
+library(reshape2)
 
 #ToDo
 #add IFD
-#add the melted foci here
+#add the melted foci code
+
 
 setwd("~./MLH1repo/")
 fullBivData = read.csv("~./MLH1repo/data/BivData/CLEAN_MUS_FULL_BIVDATA_5.28.19.csv")
@@ -74,21 +76,10 @@ fullBivData <- fullBivData[!(fullBivData$fileName %in% MultiCells), ]
 #         "30dec14_WSB_m2","17mar16_G_f4","17mar16_G_f5","4jan17_LEW_f1","4jan17_LEW_f3","4jan17_LEW_f6",
 #         "4jan17_LEW_f7","4jan17_LEW_m1", "3jan16_G_m2")
 
-exclude_mice_list = c("18nov17_WSB_f1", "8oct14_WSB_f1", "8oct14_WSB_f3",
-                      "10mar15_WSB_m2",
-                      "8jun15_G_f3", "8jun15_G_f4",
-                      "4jan17_LEW_f2",
-                      "1apr15_PWD_f2", "8oct14_PWD_f5", "8oct14_PWD_f3", 
-                      "12sep16_MSM_f3",
-                      "27oct18_MOLF_m3",
-                      "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
-                      "27aug18_CZECH_m1", "8jun18_CZECH_m1",
-                      "3sep18_AST_m1","3sep18_TOM_m1","31jul17_HMI_m1",
-                      "21aug17_SPI_f1", "30may18_SPIC_f1","8jun18_SPIC_m2")
+#bring back the bad mouse list
 
 
-
-bad_stain_mice_len <- length(BivData[BivData$mouse %in% bad_stain_mice, ])
+#bad_stain_mice_len <- length(BivData[BivData$mouse %in% bad_stain_mice, ])
 #remove bad mice
 #BivData <- BivData[BivData$mouse %in% good_stain_mice, ]
 #remove data without mouse

### 2020-01-06 doc/Effects.Rmd
cleaned total.sc script, wont overwrite MLH1 Rdata

@@
 
+#clean up total SC
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_1.6.20.RData")#total SC data
+
 #load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")#total SC data
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")#total SC data
 

### 2020-01-06 src/TotalSC.R

@@ -1,6 +1,6 @@
 # Total.SC
 # input: output from RW SC_skel run over Anon_Data images. The file names are random
-# output: matched random names with real names.
+# output: DFs of merged MLH1 and Skel data, cell and mouse levels
 
 #merge the sc.skel outputs with the master MLH1 measures
 #match up random names with original names
@@ -9,14 +9,29 @@
 #I haven't manually confirmed that the default parameters are skeletonizing the images correctly.
 
 
+
+
+#ToDo
+#rewrite in Rmd -- to better display notes and findings
+#Check the segmentation
+#re-run python scrtip, (it runs over all anon image files)
+#is it worth croping /cleaning up these images?
+
 #load main MLH1 files to merge MLH1 cell data with the sc data 
 setwd("~./")
 
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#batch18
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent 
 
 #these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
-
 SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
+
+#format all factors to numbers
+SC.skel$bin_size  <- as.character(SC.skel$bin_size )
+SC.skel$skel_size  <- as.character(SC.skel$skel_size)
+
+SC.skel$bin_size  <- as.numeric(SC.skel$bin_size )
+SC.skel$skel_size <- as.numeric(SC.skel$skel_size)
 
 SC.skel$rand_name <- as.character(SC.skel$rand_name)
 
@@ -25,14 +40,12 @@ original_DF$Random.Name <- as.character(original_DF$Random.Name)
 duplies <- original_DF[duplicated(original_DF$Random.Name),]
 #3 duplicate rows in original DF -- because Random name was blank -- how will these be matched to total.SC?
 
-
-#str(SC.skel)
[…]

### 2020-01-07 doc/Effects.Rmd
quantifying and notes for room effect

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-output: html_document
+output: 
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '5'
+    toc_float: yes
+
 ---
@@
 
-### Litter effect
+# Room effect
+
+When this project was started the main mouse room for vendor purchased mice was Biotech.
+The mouse room changed in summer 2017.
+The Gough mice were always housed in Charmony.
+Of the final dataset there are three strains had mice housed in different rooms; WSB, LEW, and MSM.
+
+The main questions to address in this section: is there a significant GxE effect due to room?  do the different rooms added noise / bias the estimates per strain
+
+A room effect might cause all of the strains to be biased in the same direction.
+
+```{r ROOM.effect.setup, echo=FALSE}
+#pull data from the 3 strains
+room.effect.DF <- MLH1_data[(MLH1_data$strain == "WSB") | (MLH1_data$strain == "LEW") | (MLH1_data$strain == "MSM"), ]
+
+original_DF <- add_mouse(original_DF)
+original_DF <- add_strain(original_DF)
+
+OG.room.effect.DF <- original_DF[(original_DF$strain == "WSB") | (original_DF$strain == "LEW") | (original_DF$strain == "MSM"), ]
+
+#mark / notch which room
+#Charmony, MSC3040, Biotech B2420,
+
+#load in file, merge with the DFs from above
+OG.RoomData = read.csv("~./MLH1repo/data/RoomData.csv", header = TRUE)
+RoomData = subset(OG.RoomData, select = c(mouse,room))
+
+#merge with the DFs from above for cell level data
+FULL.room.data = merge(RoomData, room.effect.DF, by= "mouse")
+FULL.room.data.OG = merge(RoomData, OG.room.effect.DF, by= "mouse")
+
+#make mouse.level tables
+mouse.level.MLH1 <- ddply(.data=FULL.room.data, 
+                 .(mouse),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+mouse.level.room.data <- merge(RoomData, mouse.level.MLH1, by="mouse")
+mouse.level.room.data <- add_category(mouse.level.room.data)
+
+```
+
+Looking at my records, there are limited comparisons (not all strains were raised in all 3 rooms).
+
+WSB female - Biotech vs Charmony
+
+LEW female and male, MSM female and male - Biotech and MSC
+
+look at boxplots first
+
+```{r room.effect.tests, echo=FALSE}
+library(ggsignif)
+
+#boxplots, facet by 
+geom.sig_df.cells = compare_means(nMLH1.foci ~ room, group.by = "category", data = FULL.room.data) %>% mutate(y_pos = 40)
+
+room.boxplots.cell <- ggplot(data=FULL.room.data, aes(x = room, y = nMLH1.foci))+geom_jitter()+geom_boxplot(alpha=.5)+facet_wrap(~category)+
+   #test bars
+  ggsignif::geom_signif(
+    data=geom.sig_df.cells, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos),
+    manual=TRUE)+
+  ggtitle("Boxplots of room effects,\nsingle cell level, adj p value")
+room.boxplots.cell
+
+#needed for comparing means ect.
+geom.sig_df.mouse = compare_means(mean.MLH1 ~ room, group.by = "category", data = mouse.level.room.data) %>% mutate(y_pos = 31)
+
+#  ggsignif::
+
+room.boxplots.mouse <- ggplot(mouse.level.room.data, aes(x=room, y=mean.MLH1))+geom_boxplot(alpha=.5)+facet_wrap(~category)+geom_jitter()+
+  #test bars
+geom_signif(
+    data=geom.sig_df.mouse, 
+    aes(xmin=group1, xmax=group2, annotations= p.adj, y_position= y_pos ), manual=TRUE) +  ggtitle("Boxplots of room effects,\nmouse level, adj p value")
+room.boxplots.mouse
+
+#look up the warning about cols, beccause of %>%?
+
+#remake the boxplots with compare means in the ggplot, to see if that changes the values.
+#geom.sig_df.mouse = compare_means(mean.MLH1 ~ room, group.by = "category", data = mouse.level.room.data) %>% mutate(y_pos = 31)
+
+#MORE TESTS TO consider
+#the boxplots show different ranges
+#room.boxplots.OG <- ggplot(mergy.room.data.OG)
+#tukey / post hoc paired comparison
+```
+the brackets in the plots above reflect the output of 'compare_means()' The default is wilcox.test, I've been swtiching back between showin raw and adjusted p values.
+(look up the warning about the cols parameter.)
+
+The values look a little strange to me, I don't understand the p.adj values yet, (pvalues muliutpplied by the number of comparisons).
+To me it seems like there are 5 t.tests, .
+I don't know why there are only 3 values of adjusted p-values)
+
+There are significant room effects in the cell level data set. This effect decreases when comparing the mouse level means.
+
+I think these results indicate that there is not a strong room effect. Also considering that
+
+# Litter effect
 
@@
 
-# Dissection Effect
+# Dissection spread Effect
+
+intro for why I care about dissection.
+right nnow the only independant variable is month dissected. Should I enter infor about PFA? time in humid chamber?
 
@@
 
-# Spread and Stain Effects
+# Stain Effects
+
+Intro for staining effects.
+at least 3 slides were made per mouse, giving a potential of 3 different stains.
+The staining batch might have effects on the background noise and brightness of the MLH1 foci.
+
+collect data on mice with mulitple spreads (sp1, sp2 -- make ddply table that counts a gerp for 'sp1', sp2, ect)
 
@@
 ```
-
-# Room effect
-
-```{r ROOM.effect, eval=FALSE}
-
-#COMPILE THIS datasets
-#3 different rooms used to house mice. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
-
-```

### 2020-01-07 src/SetupData.Rmd
quantifying and notes for room effect

@@
 #remove mice -- 18nov17_WSB_f5 (only 1 cell)
-exclude_mice_list = c("18nov17_WSB_f1", "8oct14_WSB_f1", "8oct14_WSB_f3",
-                      "10mar15_WSB_m2",
+exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                   "8jun15_G_f3", "8jun15_G_f4",
                   "4jan17_LEW_f2",
-                      "1apr15_PWD_f2", "8oct14_PWD_f5", "8oct14_PWD_f3", 
-                      "12sep16_MSM_f3",
+                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
+                  "12sep16_MSM_f3","31aug16_MSM_m2",
                   "27oct18_MOLF_m3",
                   "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
-                      "27aug18_CZECH_m1", "8jun18_CZECH_m1",
-                      "3sep18_AST_m1","3sep18_TOM_m1","31jul17_HMI_m1",
-                      "21aug17_SPI_f1", "30may18_SPIC_f1","8jun18_SPIC_m2")
+                  "27aug18_CZECH_m1","8jun18_CZECH_m1",
+                  "3sep18_AST_m1",
+                  "3jel18_TOM_m1","3sep18_TOM_m1",
+                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
+                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
+                  )
 

### 2020-01-07 doc/Effects.Rmd
close to wrapping up Effect summaries

@@
 Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.
+Note all strains have both embryo and neonate samples. Subset for just, WSB, G, LEW, and PWD.
 
-```{r female.age, eval=FALSE}
+```{r female.age.setup, echo=FALSE}
 #compile this dataset
+#all female -- meta.dissections has emb or neonate?
+#which sheet has this informaiton?
+
+#clean.Meta.Data.txt
+clean.meta.data <- read.csv("~./MLH1repo/data/clean.Meta.Data.txt", header = TRUE, sep='\t')
+female.age.DF <- clean.meta.data[clean.meta.data$sex == "female",]#166
+
+mergeFULL_F.mlh1 <- merge(female.age.DF, MLH1_data, by = 'mouse')
+
+merge_F.mlh1_cell <- mergeFULL_F.mlh1[ ( (mergeFULL_F.mlh1$strain.x == 'WSB') |
+                                       (mergeFULL_F.mlh1$strain.x == 'G') |
+                                       (mergeFULL_F.mlh1$strain.x == 'LEW') |
+                                  (mergeFULL_F.mlh1$strain.x == 'PWD') ), ]#901
+
+#mouse level
+mergeFULL_F.mlh1_mouse <- merge(female.age.DF, AP_mouse_table.HQ, by = 'mouse')#69 mice
+
+merge_F.mlh1_mouse <- mergeFULL_F.mlh1_mouse[ ( (mergeFULL_F.mlh1_mouse$strain.x == 'WSB') |
+                                       (mergeFULL_F.mlh1_mouse$strain.x == 'G') |
+                                       (mergeFULL_F.mlh1_mouse$strain.x == 'LEW') |
+                                  (mergeFULL_F.mlh1_mouse$strain.x == 'PWD') ), ]#42
+
 ```
+
+```{r female.age.test_cells, echo=FALSE}
+#logistic regression for MLH1 ~ emb | neonate
+
+#code emb and neonate to 0,1
+merge_F.mlh1_cell$age.stage <- ifelse(grepl("emb", merge_F.mlh1_cell$female.stage), 0, 
+                       ifelse(grepl("neonate",merge_F.mlh1_cell$female.stage), 1, "other"))
+merge_F.mlh1_cell$age.stage <- as.numeric(merge_F.mlh1_cell$age.stage)
+
+logr_age.stage.CELLS <- glm(age.stage ~ nMLH1.foci, data=merge_F.mlh1_cell[merge_F.mlh1_cell$strain.x == "WSB",], family=binomial(link="logit"))
+summary(logr_age.stage.CELLS)$coefficients
+
+#
+plot.blah.Cells <- ggplot(merge_F.mlh1_cell, aes(x=nMLH1.foci, y=age.stage)) + geom_point() +   stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Cell counts Female age logistic regression \nemb=0, neonate =1")+facet_wrap(~strain.x)
+plot.blah.Cells
+
+```
+
+For the cell level curves  (all strains pooled) the p value is not significant. When I isolate the WSB observations, the p value is 0.0778.
+
+```{r female.age.test_mouse, echo=FALSE}
+
+merge_F.mlh1_mouse$age.stage <- ifelse(grepl("emb", merge_F.mlh1_mouse$female.stage), 0, 
+                       ifelse(grepl("neonate",merge_F.mlh1_mouse$female.stage), 1, "other"))
+merge_F.mlh1_mouse$age.stage <- as.numeric(merge_F.mlh1_mouse$age.stage)
+
+#model fomula
+logr_age.stage.MOUSE <- glm(age.stage ~ mean_co, data=merge_F.mlh1_mouse, family=binomial(link="logit"))
+#logr_age.stage.MOUSE <- glm(age.stage ~ mean_co, data=merge_F.mlh1_mouse[merge_F.mlh1_mouse$strain.x == "WSB",], family=binomial(link="logit"))
+summary(logr_age.stage.MOUSE)$coefficients
+
+plot.blah.MOUSE <- ggplot(merge_F.mlh1_mouse, aes(x=mean_co, y=age.stage)) + geom_point() +   stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse Female age logistic regression \nemb=0, neonate = 1")+facet_wrap(~strain.x)
+plot.blah.MOUSE
+```
+
+The mouse curves are more slanted, but the logistic regressions are not significant. The lg for WSB mouse averages isn't significant.
 
@@
 Check if there is a maternal age effect for females and males.
+clea.Meta.Data doesn't have maternal age data.
 
-```{r mat.age.effect, echo=FALSE, eval=FALSE}
+Collect data, get a decent set / large set of mice with maternal ages (mostly want to look at females).
+
+```{r mat.age.effect, echo=FALSE}
 #pause evaluation -- needs debugging
-meta.data.file = read.csv("~./MLH1repo/data/DandM_8.30.19.csv", header = TRUE)
 
-meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
-#445? -- probably extras at bottom not good
+#this would be the best one so far
+meta.data.file_FULL = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/DandM_8.30.19.csv", header = TRUE)#updated with a couple more
 
-#how many mice have maternal DOB
-meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
-#151
+#are there duplicate mice names?
+mat.age_dupies <- meta.data.file_FULL[duplicated(meta.data.file_FULL$mouse),]
+#these are blank mouse things
+
+#try updating the D and M document, #clean up the data
+#only those with mat age
+meta.data.file_mat.age <- meta.data.file_FULL[!(is.na(meta.data.file_FULL$maternal_DOB)|meta.data.file_FULL$maternal_DOB==""),]
+#meta.data.file <- droplevels(meta.data.file)
+
+#needs mouse DOB
+meta.data.file_mat.age <- meta.data.file_mat.age[!(is.na(meta.data.file_mat.age$DOB)|meta.data.file_mat.age$DOB==""),]#140
+
+#check the date formats 
+meta.data.file_mat.age$mat_DOB.for <- as.Date(meta.data.file_mat.age$maternal_DOB, format= "%m/%d/%Y")
+#length(meta.data.file_mat.age$mouse)#152
+
+meta.data.file_mat.age$maternal_DOB <- as.Date(meta.data.file_mat.age$maternal_DOB, format= "%m/%d/%Y")
+meta.data.file_mat.age$DOB <- as.Date(meta.data.file_mat.age$DOB, format= "%m/%d/%Y")
+
+#mat.age = euth date - mat.DOB
+meta.data.file_mat.age <- add_mat_age(meta.data.file_mat.age)
+#DOB, maternal_DOB .. got the
+
+#merge this mat age file with mlh1 and mouse means!!
+mergeFULL_mat.age_cells <- merge(meta.data.file_mat.age, MLH1_data, by = 'mouse')#952 (fewer than last time)
+
+mergeFULL_mat.age_mouse <- merge(meta.data.file_mat.age, AP_mouse_table.HQ, by = 'mouse')#41
+
+#now look at signal / pattern / #hist of data
+mat.age.hist.plot <- ggplot(mergeFULL_mat.age_mouse, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
+
+mat.age.hist.plot_cells <- ggplot(mergeFULL_mat.age_cells, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females cells")
+
+#plot with MLH1 data
+mat.age_MLH1_cells <- ggplot(mergeFULL_mat.age_cells, aes(x = mat.age.days, y=nMLH1.foci, color = mouse))+geom_jitter(alpha=.3, width = 2)+facet_wrap(sex~subsp)+ggtitle("Maternal ages and MLH1")+theme(legend.position = 'none')
+
+mat.age_MLH1_cells
+
+mat.age_MLH1_mouse <- ggplot(mergeFULL_mat.age_mouse, aes(x = mat.age.days, y=mean_co, color = mouse))+geom_jitter(alpha=.3, width = 2)+facet_wrap(sex~subsp)+ggtitle("Maternal ages and MLH1")+theme(legend.position = 'none')
+
+mat.age_MLH1_mouse
+
+#mat.age_MLH1_cells.KAZ <- ggplot(mergeFULL_mat.age_cells[mergeFULL_mat.age_cells$category == "KAZ male",], aes(x = mat.age.days, y=nMLH1.foci, color = mouse))+geom_jitter()+facet_wrap(~category)+ggtitle("subset mice for maternal ages")+theme(legend.position = 'none')
+#mat.age_MLH1_cells.KAZ
+```
+
+there should be plots above this
+Based on the histograms of ages and the range across categories isn't very wide -- so not sure what the best test is.
+
+```{r mat.format.rand, echo=FALSE, eval=FALSE}
 
@@
 #the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
-meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
-meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")
-
-#str(meta.data.file.w.mat.DOB)
 
@@
 
-#merge meta data with MLH1 -- rename this
-see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
-#remove anything without mean_CO
-see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
-see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
-#what did I do with this DF...?
-
 #visualize the data
@@
 meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]
-
-#hist of data
-mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
-mat.age.hist.female
 

### 2020-01-08 doc/Effects.Rmd
Wrapping up Effects Report, fixing some bugs

@@
 
-I am testing the quality effects, since it could be a source of error and potentially biased the predicted gwRR if some mice had bad technical noise.
+> Main points
+
+1. There is a positive relationship with quality and MLH1 count (better cells have more MLH1)
+
+2. The effect of Quality on mean MLH1 count is larger in males
+
+> Open quesitons
+
+ - Are quality score less accurate/consistant for females?
+
+ - Should I weight mouse COs averages by the quality ratio?
+
+ - Exclude lowest quality cells in males, but keep them in female dataset?
+
+ - Does the effect for males still hold when Q5 cells are removed?
 
@@
 
-```{r MLH1.histogram.2, echo=FALSE, include=FALSE}
+```{r MLH1.histogram.2, echo=FALSE, eval=FALSE,include=FALSE}
 #thse plots not shown
@@
 
-These plots show that generally there is a positive relationship between good quality (1) and more MLH1 counts.
+These plots show that generally there is a positive relationship between good quality and more MLH1 counts.
 
-I should focus on addressing if there are differences in the relationship of quality-MLH1 counts across sexes.
+```{r qual.effect.sex, echo=FALSE}
+#compare the quality effect by sex
+#an issue with comparing sexes is they have different ranges of MLH1 means
+#do other analysis
+
+#plots show lm for observations from mouse.table.mean.MLH1.obs_qual
+#glm
+
+blah.lm.ff <- lm( mean.MLH1 ~ quality + ncells + subsp + strain, data=ff)
+#quality bin is significant to predict count in ff (remeber these are binned averages)
+# -0.85303,  3.61e-14 ***
+
+blah.lm.mm <- lm( mean.MLH1 ~ quality + ncells + subsp + strain, data=mm)
+#quality co Eff,   -8.863e-01, < 2e-16 ***
+
+mm.low <- mm[!( (mm$strain == "MSM") | (mm$strain == "PWD") ), ]
+
+blah.lm.mm.low <- lm( mean.MLH1 ~ quality + ncells + subsp + strain, data=mm.low)
+#quality sig, so strain (unlike female (probably bc greater range))
+
+# -8.881e-01,  < 2e-16 ***
+
+#the coefficient is 10x higher in males, p value max-out low 
+#females the quality co-E is much smaller and p.value larger  
+#quality effect on mean co's is larger in males
+
+```
+
+I ran lm of the data from the plots above and found that the coefficient is 10x higher in males, p value max-out low and in females the quality co-efficicent is much smaller and p.value larger.
 
@@
 
-Examine image quality or other scores to test how well the python scripts segment the SC area for all the images.
+In progress. Examine image quality or other scores to test how well the python scripts segment the SC area for all the images.
 
-```{r SC.qual, echo=FALSE}
+```{r SC.qual, echo=FALSE, eval=FALSE}
 #use the SC table
@@
 
-```{r SC.hitograms, echo=FALSE, fig.width = 4, fig.height=1}
+```{r SC.hitograms, echo=FALSE, eval=FALSE, fig.width = 4, fig.height=1}
 #try adjusting the size for thi..
@@
 
-```{r more.SC.hist, echo=FALSE}
+```{r more.SC.hist, echo=FALSE, eval=FALSE}
 
@@
 
-```{r sc.qual.plots, echo=FALSE}
+```{r sc.qual.plots, echo=FALSE, eval=FALSE}
 skel.mm <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "male",]
@@
 
-```{r G.plot, echo=FALSE}
+```{r G.plot, echo=FALSE, eval=FALSE}
 Bin.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
@@
 
-The mouse 22jun15_G_m2 has strange pattern, investigate these outputs
+The mouse 22jun15_G_m2 has strange pattern, investigate the Python SC outputs.
 
@@
 
-```{r female.skel.plot, echo=FALSE}
+```{r female.skel.plot, echo=FALSE, eval=FALSE}
 
@@
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female quality and mean CO")+facet_wrap(~subsp)
 
@@
   scale_color_manual(values=colors_of_strains)+ #this list in setupdata
-  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)
+  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female quality and mean CO")+facet_wrap(~mouse)
 
@@
 
-The above plots are good for detecting outliers
-
 # Age Effect (male)
 
-Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age].
+Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age] (after maturity, there's a negative correlation with age and recombination).
 
-The initial analysis uses plots. This should be double checked with a general linear model as well.
+> Main Points
+
+ 1. Weak evidence for age effect on MLH1 counts for male mice.
+
+ - Dom, no significant effects for age.
+
+ - Musc low, no significant effect for strain or age.
+
+ - Musc high, no significant age effects on mean CO count
+
+My previous dataset showed age effect for SKIVE males due to unbalanced data.
+There were only 3 mice of 7,16, and 41 weeks.
+The models using this dataset had signifcant age effects.
+The updated data set now has ~4 mice of 7 weeks and no significant age effect.
+
+I tested the age effects with 3 sets of models,  for 3 groupings of mice: dom, musc.high and musc.low
 
@@
 
-I made 3 models for 3 groupings of mice, dom, musc.high and musc.low
-
-$$mouse .mean CO ~=~ age + strain + \varepsilon$$
+$$mouse .mean CO \sim age + strain + \varepsilon$$
 
@@
 
-ADD euthdate?
-
-Dom, no significant effects for age.
-
-Musc low, no significant effect for strain or age.
-
-Musc high, MSM and SKIVE have significant age effects on mean CO count
-
-SKIVE has the greatest spread in ages for single mice (7,16,41 weeks)
-There are more SKIVE males which I haven't quantified (at least 3, which are in a good age range)
+The plot below shows mouse averages as single points, showing the mean MLH1 and age in weeks. The lines and shading show the smoothed lm for each strain.
 
@@
 
-I like this plot and find it effective. Can I do a similar set up for other factors (quality, ncells?)
+The models and mostly flat slopes indecate that there isn't a strong age effect on MLH1 for these mice.
+(there might be a KAZ male outlier (10jul19_KAZ_m3), double check this).
+Change the colors for KAZ, TOM, AST, and CZECH they're too similar.
 
@@
 
-Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.
-Note all strains have both embryo and neonate samples. Subset for just, WSB, G, LEW, and PWD.
+Female ages most logically group into neonate or embryo, so I choose to apply a logistic regression to test for age effects.
+Not all strains have both embryo and neonate female samples, so just, WSB, G, LEW, and PWD are examined.
+
+> Main Point
+
+ - MLH1 counts don't differ between neonates and embryos
+
+Below are the tests for the model (can CO number predict the age stage).
+I ran these comparisons for cell count and mouse averages.
+The plots show the points and smoothed curves with age on the y axis, 0 = embryo and 1 = neonate.
+
+$$ Age .Stage \sim CO .number$$
 
@@
 
-The mouse curves are more slanted, but the logistic regressions are not significant. The lg for WSB mouse averages isn't significant.
+The mouse curves are more slanted, but the logistic regressions are not significant. None of the indiviual strains had significant p values when compared by themselves.
+
+# Room effect
+
+When this project was started the main mouse room for vendor purchased mice was Biotech.
+The mouse room changed in summer 2017.
+The Gough mice were always housed in Charmony.
+Of the final data set there are three strains had mice housed in different rooms; WSB, LEW, and MSM.
+(not balanced/ limited comparisons / )
+
+The main questions to address in this section: is there a significant GxE effect due to room?
+
+> Main Points
+
+ - There is some evidence for room effect on MLH1 counts per cell for WSB and LEW female and LEW and MSM male.
+
+ - No support of room effect for mouse means.
+
+For the cell counts of MLH1, Two of three are in the same direction. Sample size might be having an effect.
+
+There might be more mice/data that can be added to these tests, need to comfirm the move date and verify with lab note book.
+
+```{r ROOM.effect.setup, echo=FALSE, warning=FALSE}
+#pull data from the 3 strains
+room.effect.DF <- MLH1_data[(MLH1_data$strain == "WSB") | (MLH1_data$strain == "LEW") | (MLH1_data$strain == "MSM"), ]
+
+original_DF <- add_mouse(original_DF)
+original_DF <- add_strain(original_DF)
+
+OG.room.effect.DF <- original_DF[(original_DF$strain == "WSB") | (original_DF$strain == "LEW") | (original_DF$strain == "MSM"), ]
+
+#mark / notch which room
+#Charmony, MSC3040, Biotech B2420,
+
+#load in file, merge with the DFs from above
+OG.RoomData = read.csv("~./MLH1repo/data/RoomData.csv", header = TRUE)
+RoomData = subset(OG.RoomData, select = c(mouse,room))
+
+#merge with the DFs from above for cell level data
+FULL.room.data = merge(RoomData, room.effect.DF, by= "mouse")
+FULL.room.data.OG = merge(RoomData, OG.room.effect.DF, by= "mouse")
+
+#make mouse.level tables
+mouse.level.MLH1 <- ddply(.data=FULL.room.data, 
+                 .(mouse),
+                 summarize, 
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+mouse.level.room.data <- merge(RoomData, mouse.level.MLH1, by="mouse")
+mouse.level.room.data <- add_category(mouse.level.room.data)
+
+```
+
+```{r room.effect.tests, echo=FALSE, warning=FALSE}
+library(ggsignif)
+
+mouse.level.room.data <- mouse.level.room.data[!(is.na(mouse.level.room.data$mean.MLH1)|mouse.level.room.data$mean.MLH1==""),]
+
+#note the mouse level tests are all throwing errors because not enough samples
+
+#Wlcx.ROOM_WSB.female  <- wilcox.test(FULL.room.data$mean.MLH1[ ( (FULL.room.data$category == "WSB female") & (FULL.room.data$room == "B2420") ) ], 
+#            FULL.room.data$mean.MLH1[ ( (FULL.room.data$category == "WSB female") & (FULL.room.data$room == "Charmony") ) ]  ) #raws for compare means match
+
+Wlcx.ROOM_LEW.male  <- wilcox.test(FULL.room.data$nMLH1.foci[ ( (FULL.room.data$category == "LEW male") & (FULL.room.data$room == "B2420") ) ], 
+           FULL.room.data$nMLH1.foci[ ( (FULL.room.data$category == "LEW male") & (FULL.room.data$room == "MSC 3040") ) ]  )
+
+#boxplots, facet by 
+geom.sig_df.cells = compare_means(nMLH1.foci ~ room, group.by = "category", data = FULL.room.data) %>% mutate(y_pos = 40)
+
+room.boxplots.cell <- ggplot(data=FULL.room.data, aes(x = room, y = nMLH1.foci))+geom_jitter()+geom_boxplot(alpha=.5)+facet_wrap(~category)+
+   #test bars
+  ggsignif::geom_signif(
+    data=geom.sig_df.cells, 
+    aes(xmin=group1, xmax=group2, annotations=p, y_position=y_pos),
+    manual=TRUE)+
+  ggtitle("Boxplots of room effects,\nsingle cell counts, raw p value")
+room.boxplots.cell
+
+#needed for comparing means ect.
+geom.sig_df.mouse = compare_means(mean.MLH1 ~ room, group.by = "category", data = mouse.level.room.data) %>% mutate(y_pos = 31)
+
+#  ggsignif::
+
+room.boxplots.mouse <- ggplot(mouse.level.room.data, aes(x=room, y=mean.MLH1))+geom_boxplot(alpha=.5)+facet_wrap(~category)+geom_jitter()+
+  #test bars
+geom_signif(
+    data=geom.sig_df.mouse, 
+    aes(xmin=group1, xmax=group2, annotations= p, y_position= y_pos ), manual=TRUE) +  ggtitle("Boxplots of room effects,\nmouse means, raw p values")
+room.boxplots.mouse
+
+#look up the warning about cols, beccause of %>%?
+
+#remake the boxplots with compare means in the ggplot, to see if that changes the values.
+#geom.sig_df.mouse = compare_means(mean.MLH1 ~ room, group.by = "category", data = mouse.level.room.data) %>% mutate(y_pos = 31)
+
+#MORE TESTS TO consider
+#the boxplots show different ranges
+#room.boxplots.OG <- ggplot(mergy.room.data.OG)
+#tukey / post hoc paired comparison
+```
+
+The brackets in the plots above reflect the output of 'compare_means()' which uses a wilcox.test.
+
+I double checked the wilcox p values for wilcox.test and compare_means, they're similar.
+
+WSB single wilcox test pvalue is  Wlcx.ROOM_WSB.female$p.value` which is the same as the raw from compare_means(). The compare_means table shows that WSB is multiplied by 2.
+
+The values look a little strange to me, I don't understand the p.adj values yet, (p values multiplied by the number of comparisons).
+To me it seems like there are 5 t.tests, .
+I don't know why there are only 3 values of adjusted p-values)
+
+There are significant room effects in the cell level data set. This effect decreases when comparing the mouse level means.
+
+I think these results indicate that there is not a strong room effect. Also considering that
 
@@
 
-Check if there is a maternal age effect for females and males.
-clea.Meta.Data doesn't have maternal age data.
+> Main Point
 
-Collect data, get a decent set / large set of mice with maternal ages (mostly want to look at females).
+ - Qualitatively, there is unlikely to be a maternal age effect, but these data are underpowered.
 
@@
 
-mat.age_MLH1_mouse
-
 #mat.age_MLH1_cells.KAZ <- ggplot(mergeFULL_mat.age_cells[mergeFULL_mat.age_cells$category == "KAZ male",], aes(x = mat.age.days, y=nMLH1.foci, color = mouse))+geom_jitter()+facet_wrap(~category)+ggtitle("subset mice for maternal ages")+theme(legend.position = 'none')
@@
 ```
-
-there should be plots above this
-Based on the histograms of ages and the range across categories isn't very wide -- so not sure what the best test is.
 
@@
 
-# Room effect
+For the topics below, Maternal age, litter and dissection effect are all related, I need to outline/distinguish how and decide if these are worth further investigation.
 
-When this project was started the main mouse room for vendor purchased mice was Biotech.
-The mouse room changed in summer 2017.
-The Gough mice were always housed in Charmony.
-Of the final dataset there are three strains had mice housed in different rooms; WSB, LEW, and MSM.
-
-The main questions to address in this section: is there a significant GxE effect due to room?  do the different rooms added noise / bias the estimates per strain
-
-A room effect might cause all of the strains to be biased in the same direction.
-
-```{r ROOM.effect.setup, echo=FALSE}
-#pull data from the 3 strains
-room.effect.DF <- MLH1_data[(MLH1_data$strain == "WSB") | (MLH1_data$strain == "LEW") | (MLH1_data$strain == "MSM"), ]
-
-original_DF <- add_mouse(original_DF)
-original_DF <- add_strain(original_DF)
-
-OG.room.effect.DF <- original_DF[(original_DF$strain == "WSB") | (original_DF$strain == "LEW") | (original_DF$strain == "MSM"), ]
-
-#mark / notch which room
-#Charmony, MSC3040, Biotech B2420,
-
-#load in file, merge with the DFs from above
-OG.RoomData = read.csv("~./MLH1repo/data/RoomData.csv", header = TRUE)
-RoomData = subset(OG.RoomData, select = c(mouse,room))
-
-#merge with the DFs from above for cell level data
-FULL.room.data = merge(RoomData, room.effect.DF, by= "mouse")
-FULL.room.data.OG = merge(RoomData, OG.room.effect.DF, by= "mouse")
-
-#make mouse.level tables
-mouse.level.MLH1 <- ddply(.data=FULL.room.data, 
-                 .(mouse),
-                 summarize, 
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-mouse.level.room.data <- merge(RoomData, mouse.level.MLH1, by="mouse")
-mouse.level.room.data <- add_category(mouse.level.room.data)
-
-```
-
-Looking at my records, there are limited comparisons (not all strains were raised in all 3 rooms).
-
-WSB female - Biotech vs Charmony
-
-LEW female and male, MSM female and male - Biotech and MSC
-
-look at boxplots first
-
-```{r room.effect.tests, echo=FALSE}
-library(ggsignif)
-
-#boxplots, facet by 
-geom.sig_df.cells = compare_means(nMLH1.foci ~ room, group.by = "category", data = FULL.room.data) %>% mutate(y_pos = 40)
-
-room.boxplots.cell <- ggplot(data=FULL.room.data, aes(x = room, y = nMLH1.foci))+geom_jitter()+geom_boxplot(alpha=.5)+facet_wrap(~category)+
-   #test bars
-  ggsignif::geom_signif(
-    data=geom.sig_df.cells, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos),
-    manual=TRUE)+
-  ggtitle("Boxplots of room effects,\nsingle cell level, adj p value")
-room.boxplots.cell
-
-#needed for comparing means ect.
-geom.sig_df.mouse = compare_means(mean.MLH1 ~ room, group.by = "category", data = mouse.level.room.data) %>% mutate(y_pos = 31)
-
-#  ggsignif::
-
-room.boxplots.mouse <- ggplot(mouse.level.room.data, aes(x=room, y=mean.MLH1))+geom_boxplot(alpha=.5)+facet_wrap(~category)+geom_jitter()+
-  #test bars
-geom_signif(
-    data=geom.sig_df.mouse, 
-    aes(xmin=group1, xmax=group2, annotations= p.adj, y_position= y_pos ), manual=TRUE) +  ggtitle("Boxplots of room effects,\nmouse level, adj p value")
-room.boxplots.mouse
-
-#look up the warning about cols, beccause of %>%?
-
-#remake the boxplots with compare means in the ggplot, to see if that changes the values.
-#geom.sig_df.mouse = compare_means(mean.MLH1 ~ room, group.by = "category", data = mouse.level.room.data) %>% mutate(y_pos = 31)
-
-#MORE TESTS TO consider
-#the boxplots show different ranges
-#room.boxplots.OG <- ggplot(mergy.room.data.OG)
-#tukey / post hoc paired comparison
-```
-the brackets in the plots above reflect the output of 'compare_means()' The default is wilcox.test, I've been swtiching back between showin raw and adjusted p values.
-(look up the warning about the cols parameter.)
-
-The values look a little strange to me, I don't understand the p.adj values yet, (pvalues muliutpplied by the number of comparisons).
-To me it seems like there are 5 t.tests, .
-I don't know why there are only 3 values of adjusted p-values)
-
-There are significant room effects in the cell level data set. This effect decreases when comparing the mouse level means.
-
-I think these results indicate that there is not a strong room effect. Also considering that
-
-# Litter effect
+## Litter effect
 
@@
 
-# Dissection spread Effect
+## Dissection spread Effect
 
-intro for why I care about dissection.
-right nnow the only independant variable is month dissected. Should I enter infor about PFA? time in humid chamber?
+right now the only independent variable is month dissected. Should I enter info about PFA? time in humid chamber?
 
-```{r dissection.effect.data.setup, echo=FALSE}
+```{r dissection.effect.data.setup, echo=FALSE, warning=FALSE}
 
@@
 
-```{r plot.dissection, echo=FALSE}
+```{r plot.dissection, echo=FALSE, warning=FALSE}
 
@@
 
-```{r dissection.female, echo=FALSE}
+```{r dissection.female, echo=FALSE, warning=FALSE}
 
@@
 
-## Total SC
+### Total SC
 
 Testing the effects of mean total SC per mouse with when they were dissected.
+This is mean to be a follow up to peromyscus findings, Where summer caught mice had lower SC area per cell, compared to winter and fall caught.
+But there were few data points.
 
-follow up to peromyscus findings, Where summer caught mice had lower SC area per cell, compared to winter and fall caught. But there were few data points.
-
-```{r total.sc, echo=FALSE}
+```{r total.sc, echo=FALSE, warning=FALSE}
 #check out effects of month dissected on SC
@@
 This might be the start of testing for 'technical effects / humidity effects'.
-BUT, I don't really have month by month dissections of mice of the same age, ect.
-
-> *month dissected*
+BUT, I don't really have month by month dissections of mice of the same age, etc.
 
@@
 
-# Stain Effects
+# Stain Effect
 
-Intro for staining effects.
-at least 3 slides were made per mouse, giving a potential of 3 different stains.
-The staining batch might have effects on the background noise and brightness of the MLH1 foci.
+At least 3 slides were made per mouse and I rarely included slides from the same mouse when running the statining protocol.
+Different staining batch might have effects on the background noise and brightness of the MLH1 foci (these could also be affected by post processing of images to a lesser extent).
+A siginificant stain effect would indecate technical error / I wasn't consistant in running the staining protocol).
 
-collect data on mice with mulitple spreads (sp1, sp2 -- make ddply table that counts a gerp for 'sp1', sp2, ect)
+**collect data set for mice with multiple spreads (sp1, sp2 make ddply table that counts a gerp for 'sp1', sp2, ect)**
 
-> *staining batches*
-
+These data might also be underpowered, since I mostly used 1 spread and it was rare for me to collect over ~20 cells from multiple spreads from the same mouse.
 Compare observations from the same mouse, but different MLH1 stains.  First need to identify data that can be used.
@@
 ```{r stain.effects.SC, eval=FALSE}
-#/src/TotalSC.R
+
 #compile this dataset

### 2020-01-08 src/CommonFunc_MLH1repo.R

@@ -8,6 +8,15 @@
 
 
 
+#ADD THE RANK to chrm measures 
+#fixed it with rank
+#for(musy in as.character(unique(whole.cell.data$fileName))){
+#  #print(musy)
+#  #order the chrm lengths -- this is order
+#  chrm.order = rank(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
+#  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order   #this is just a string of numbers--
+#}
+
 #make fileName
 #paste together file name
 add_fullFileName <- function(dat){

### 2020-01-09 doc/Effects.Rmd
small updates for effects docs

@@
 
-I ran lm of the data from the plots above and found that the coefficient is 10x higher in males, p value max-out low and in females the quality co-efficicent is much smaller and p.value larger.
+These results confirm that quality has a significant effect on MLH1 count per cell.
+How this affects further inferences is a little unclear since the quality score is a very rough indicator of technical error.
 
@@
 
-Each point is a single mouse
-
 ```{r G.plot, echo=FALSE, eval=FALSE}
@@
 
-The mouse 22jun15_G_m2 has strange pattern, investigate the Python SC outputs.
-
-Some of the pattern goes away when redo.crop is set to equal 'no'.
+The mouse 22jun15_G_m2 has strange pattern, investigate the Python SC outputs. Some of the pattern goes away when redo.crop is set to equal 'no'.
 
@@
 Of the final data set there are three strains had mice housed in different rooms; WSB, LEW, and MSM.
-(not balanced/ limited comparisons / )
+(This data might not balanced and there are limited comparisons.)
 
@@
 
-The brackets in the plots above reflect the output of 'compare_means()' which uses a wilcox.test.
-
-I double checked the wilcox p values for wilcox.test and compare_means, they're similar.
-
-WSB single wilcox test pvalue is  Wlcx.ROOM_WSB.female$p.value` which is the same as the raw from compare_means(). The compare_means table shows that WSB is multiplied by 2.
-
-The values look a little strange to me, I don't understand the p.adj values yet, (p values multiplied by the number of comparisons).
-To me it seems like there are 5 t.tests, .
-I don't know why there are only 3 values of adjusted p-values)
+The brackets in the plots above reflect the output of 'compare_means()' which uses a wilcox.test. I double checked the wilcox p values for wilcox.test and compare_means, they're similar.
 
 There are significant room effects in the cell level data set. This effect decreases when comparing the mouse level means.
-
-I think these results indicate that there is not a strong room effect. Also considering that
 
@@
 
- - Qualitatively, there is unlikely to be a maternal age effect, but these data are underpowered.
+ - Qualitatively, it doesn't appear that there is a maternal age effect, but these data are underpowered.
 
@@
 
-right now the only independent variable is month dissected. Should I enter info about PFA? time in humid chamber?
+Plotting euth date and mean CO number.
 

### 2020-01-09 doc/Main_Report_Results.rmd
small updates for effects docs

--- 2019-12-31 doc/Main_Report_Results.rmd
@@
 ---
-title: "MLH1 Data Report"
+title: "MLH1 Results Outline"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 #load main data file
-
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+
+#this MLH1 only goes up to batch16!
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
 #load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
@@
 #load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
+
 #load(file="data/MetaData.RData") #this was replacing most updated data file
@@
 
-load in more Data, chrm proportions, total SC,
-write / standardize the data set up for those
-
-Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
--ene endings refer to the cells/units within the cell stage? (I think)
-
 ```{r organizing.data, echo=FALSE }
 #set the order of another column, based on another variable. consistant order for ploting
+
+#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
+#-ene endings refer to the cells/units within the cell stage? (I think)
 
@@
 
-Details of dataset cleaning ect is in SetupData.rmd
+Details of data set cleaning ect is in SetupData.rmd
 Effects from different factors are reviewed in Effects.Rmd
@@
     AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-
 ```
@@
 
-figure 1, change to all white, add the subspecies lines below.
-
-```{r first.boxplots, echo=FALSE, fig.align='center'}
+```{r first.boxplots, echo=FALSE, fig.align='center', warning=FALSE}
 
@@
 #these plot currently look bad
-ss
-justMuscu
 ```
@@
 
-*cast female data came from another publication
-
-add code for the table table. may have to deal with the excess mouse issue
+*cast female data came from another publication, add this ref (Kohler?)
+
+The bottom lines and subspecies points don't exactly match up.
+
+**add in cell images like the other draft figures.**
 
@@
 
-These are the basic Mixed Models we used as a frame work to analyze all the MLH1 data.
+These are the basic Mixed Models we used as a frame work to analyze all the MLH1 count data.
 
@@
 
-Seting up the data set for using the mixed model (base data set, HQ dataset)
+Setting up the data set for using the mixed model (base data set, HQ data set)
 
@@
 
-- male ages within range
-
-Currently I ran through 3 datasets: 1) base, 2) HQ and 3) Molf (from HQ)
+- male ages within good range (5 - 12? weeks)
+
+Currently I ran through 3 data sets: 1) base, 2) HQ and 3) Molf (from HQ)
 
@@
 
-Above are the total number of mice in the data set
-
-FOR FINAL ANALYSIS, MOL Will be coded as a seperate subsp (change all of the main model reports to)
+FOR FINAL ANALYSIS, make sure that MOL is coded as a separate subsp (change all of the main model reports to).
 For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.
 
-subspecies levels: 2(3), Musc and Domesticus (Molossinus)
-
-strain levels: Dom: 3, Musc: 4  (Mol: 2)
+subspecies levels: 2(3), Musc and Domestics (Mousiness)
+
+strain levels: Don: 3, Musc: 4  (Mol: 2)
 
@@
 
-Above is the table of coefficients for MM.1, HQ dataset, lme(), random effects strain and strain specific sex effect.
+Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.
 
@@
 
-Above is the table of coefficients for MM.1.mol, molssinus coded as seperate subspecies, lme(), random effects strain and strain specific sex effect.
+Above is the table of coefficients for MM.1.mol, molssinus coded as separate subspecies, lme(), random effects strain and strain specific sex effect.
 
@@
 
-Above is the table of coefficients for MM.2, HQ dataset, lme(), random effects coded without sex specific effect.
+Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
 This is model has sex and interaction coefficient is significant.
@@
 
-Above is the table of coefficients for MM.2.2, HQ dataset, gls(), No random effects.  
+Above is the table of coefficients for MM.2.2, HQ data set, gls(), No random effects.  
 I think this gls has sex and interaction term as significant.
@@
 
-Above is the table of coefficients for MM.3, HQ dataset, lm function, sex, strain and subsp are interacting fixed effects.
+Above is the table of coefficients for MM.3, HQ data set, lm function, sex, strain and subsp are interacting fixed effects.
 
@@
 
-Above is the table of coefficients for **MM.4**, HQ dataset, lm function, just sex and strain, no subspecies.
+Above is the table of coefficients for **MM.4**, HQ data set, lm function, just sex and strain, no subspecies.
 
 Significant factors are 
-strainG (p=0.0005), strainMSM and strainPWD (very slight),
+strain G (p=0.0005), strainMSM and strainPWD (very slight),
 sex:G (p=0.001),
@@
 
-#### Summary of mean COs
+ Summary of mean COs
 
@@
 
-# Variance in CO count
-
-in addition to the mean gwRR, the variance is also an important metric to dicribe
+## Variance in CO count
+
+in addition to the mean gwRR, the variance is also an important metric to describe
 
@@
 
+### Figures
+
 ```{r variance.boxplots, echo=FALSE, eval=FALSE}
@@
 
-The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect pvalues were larger and only significant for one of the models.
-
-When Molossinus is coded as a seperate subspecies, then interaction terms are no longer significant.
+The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.
+
+When Molossinus is coded as a separate subspecies, then interaction terms are no longer significant.
 
@@
 
-list out the points whic hsuport rapid male specific evolution for MLH1 rates.
+list out the points which support rapid male specific evolution for MLH1 rates.
 
@@
 
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))
+chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
 
@@
 
-# Sexual dimorphism and heterochiasmy things
-
-total sc length / SC compaction and telomere bias in positioning are conserved / consistantly sexually dimorphism (in these mice?)  for most species
+# Sexual dimorphism and heterochiasmy
+
+total SC length / SC compaction and telomere bias in positioning are conserved / consistently sexually dimorphism (in these mice?)  for most species
 
@@
 
-# ""
+# Blank
 

### 2020-01-09 doc/MouseLevelPlots.Rmd
small updates for effects docs

@@
 
-exclude_mice <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
+exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                   "8jun15_G_f3", "8jun15_G_f4",
@@
                   "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
-                  "12sep16_MSM_f3",
+                  "12sep16_MSM_f3","31aug16_MSM_m2",
                   "27oct18_MOLF_m3",
@@
                   "3jel18_TOM_m1","3sep18_TOM_m1",
-                  "31jul17_HMI_m1",
+                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                  "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"

### 2020-01-10 doc/MouseLevelPlots.Rmd
Changed Mouse Level plots to show ALL mice (use original.DF instead of MLH1_data)

@@
 #load main data file
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.10.20.RData")#edits to OG.original.DF
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
 #load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
@@
 
+#main data frame should be original_DF then changes made to MLH1_data
+
 #remove other!
-MLH1_data <- MLH1_data[MLH1_data$strain != "other",]
+original_DF <- original_DF[original_DF$strain != "other",]
 
-MLH1_data <- droplevels(MLH1_data)
+original_DF <- droplevels(original_DF)
+```
+
+```{r clean.OG.DF, echo=FALSE}
+#because I want to display the MLH1 means from the OG list
+#Have have to do a bit of clean up
+
+#for this doc / report, remove 'X's (these changes to the DF won't be passed on)
+original_DF <- original_DF[ !grepl("X", original_DF$X) , ]
+original_DF <- original_DF[!(is.na(original_DF$nMLH1.foci) | original_DF$nMLH1.foci==""), ]
+
+original_DF <- original_DF[!(is.na(original_DF$n) | original_DF$n != 20), ]
+
+original_DF <- original_DF[original_DF$nMLH1.foci != "X",]
+original_DF <- original_DF[original_DF$nMLH1.foci != "x",]
+
+#
+original_DF$n <- as.character(original_DF$n)
+original_DF$nMLH1.foci<- as.character(original_DF$nMLH1.foci)
+original_DF$quality<- as.character(original_DF$quality)
+
+#remove non-number chrs
+original_DF$n <- as.numeric(original_DF$n)
+original_DF$nMLH1.foci<- as.numeric(original_DF$nMLH1.foci)
+original_DF$quality<- as.numeric(original_DF$quality)
+
+#when first ran there were alot of empty mouse -squares
+
+original_DF <- droplevels(original_DF)
+
+#num are nums, there are still Q5 cells
 ```
@@
 
+The main data frame for these results is original_Df, which is a step removed from the final analysis DF.
+All of the mice of the data set are shown, to give a better idea of what is take out for the analysis dataset.
+('X' labeled cells have been removed, Q5 cells remain).
+
 Yellow colored boxes are mice that will be exclude from the final (HQ) dataset for low quality cells (5).
-Consider re-runing this with the final data frame to see how the strain and mouse means change.
 
@@
 #switch this loop to ex
-for(i in 1:length(MLH1_data$mouse)){
-  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], exclude_mice), "exclude", NA)
+
+#shouldn't the DF be the OG DF?
+
+for(i in 1:length(original_DF$mouse)){
+  original_DF$quant_status[i] <- ifelse( is.element(original_DF$mouse[i], exclude_mice_list), "exclude", NA)
 }
@@
 
-df_list <- split(MLH1_data, (MLH1_data$category))
+df_list <- split(original_DF, (original_DF$category))
 #remove sections which are in category but have no data
@@
   sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                      mouse_nMLH1 = mean(adj_nMLH1.foci)
+                      mouse_nMLH1 = mean(nMLH1.foci)
   )
@@
   pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
-    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + xlim(c(0.5,5.5))+
+    geom_jitter(aes(y= nMLH1.foci, x= quality)) + xlim(c(0.5,5.5))+
 
@@
 
-      geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+      geom_hline(yintercept = c( mean(df_list[[h]]$nMLH1.foci) ), color="black" )
   #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 

### 2020-01-10 doc/totalSC.Rmd
Changed Mouse Level plots to show ALL mice (use original.DF instead of MLH1_data)

---
title: "Total_SC"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default

---

```{r setup, echo=FALSE, warning=FALSE, include = FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
#load main MLH1 files to merge MLH1 cell data with the sc data 
setwd("~./")

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent

#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)

#load BivData
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData.RData")

# Total.SC
# input: output from RW SC_skel run over Anon_Data images. The file names are random
# output: DFs of merged MLH1 and Skel data, cell and mouse levels
```

# Goals

Why include these results in the Main MLH1 results?

 - test if **greater SC area in females is a conserved trait of heterochiasmy**

 - test how average total SC correlates / contributes to mean CO

## Discribe these Data

output from RW SC_skel run over Anon_Data images. The file names are random

The Main dataframe is the Merge MLH1 + skel data.

## Running the Python Script

Python -- apply to all images in the Anon folder (so all the random named image files)

(should I update all the crops? How many images need cropping)

### ToDo

- How will I test these / and adjust for XX?

- Look up the python output

- Check for outliers in the skel data

- rerun the Python script

- check that the mice from mixed models are in these

ToDo re-run the pipeline, to get measues for SKIVE and MOLF females
molf has almost eq SC -- but this is just 1 female

merge the sc.skel outputs with the master MLH1 measures
match up random names with original names

the skel output and scripts are in the SC_skel dir. it was run on all images in the anon_data folder. 
I haven't manually confirmed that the default parameters are skeletonizing the images correctly.

```{r clean.data, echo=FALSE, warning=FALSE}
#format all factors to numbers
SC.skel$bin_size  <- as.character(SC.skel$bin_size )
SC.skel$skel_size  <- as.character(SC.skel$skel_size)

SC.skel$bin_size  <- as.numeric(SC.skel$bin_size )
SC.skel$skel_size <- as.numeric(SC.skel$skel_size)

SC.skel$rand_name <- as.character(SC.skel$rand_name)

#check for duplicatations of the random name
original_DF$Random.Name <- as.character(original_DF$Random.Name)
duplies <- original_DF[duplicated(original_DF$Random.Name),]
#3 duplicate rows in original DF -- because Random name was blank -- how will these be matched to total.SC?
```

```{r merge.DF, echo=FALSE}

#MERGEING MLH! with SKEL
MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF

```

The mixed models are mouse averages, get mouse names and ncells for each of the lists.

```{r mixed.models.setup, echo=FALSE, warning=FALSE}
#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme
#notes for testing random effects

#variance is still chr
AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
                            subsp.y,subsp.x,
                    category.y, category.x,
                        strain.x,strain.y, mouse.1  )  )

#REMOVE EXTRA STRAINS
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#mean_co.HQ  (this might remove some of the MOLF males)
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse)|DF.HetC.MixedModel.HQ$mouse==""),]

```

```{r check.lists, echo=FALSE}

MLH1.merge.Skel
OG.MLH1.merge.Skel

#checking / comparing the two lists

DF.HetC.MixedModel.HQ$mouse

skel.mice_from.MM.DF <- OG.MLH1.merge.Skel[ OG.MLH1.merge.Skel$mouse %in% DF.HetC.MixedModel.HQ$mouse, ]
#1630

#compare this list to cells in whole cell / curate (why?)
#FOR COMPARISON ACROSS APPROACHES

#curated Biv data doesn't have '.tif' -- remove .tif
skel.mice_from.MM.DF$fileName <- as.character(skel.mice_from.MM.DF$fileName)

for(j in 1:length(skel.mice_from.MM.DF$fileName)){
  print(j) 
  #MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
   templist2 = strsplit(skel.mice_from.MM.DF$fileName[j], split=".tif")#[[1]]
   #print(templist2)
   skel.mice_from.MM.DF$fileName2[j] <- templist2 #should match file name
}

DF <- skel.mice_from.MM.DF[ skel.mice_from.MM.DF$fileName2 %in% Curated_BivData$fileName, ]#299
#above is the DF which there are cell which have been curated
#merge --

DF <- add_strain(DF)
DF <- add_sex(DF)
DF <- add_category(DF)

table(DF$category) #missing MSM females

#this table now has merged bin/skel, and have been curated (but it doesn't have data on the curated bivalents)

#skel.mice_from.MM.DF$fileName %in% Curated_BivData$fileName

Skel.summary.table <- ddply(.data=skel.mice_from.MM.list, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                 var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                 mean.bin =  mean(bin_size, na.rm = TRUE)

)

#add strain, category, etc

```

```{r cleaning.up, echo=FALSE}
#clean up this DataFrame REMOVE 'X'
MLH1.merge.Skel <- MLH1.merge.Skel[ !grepl("X", MLH1.merge.Skel$Batch) , ]
MLH1.merge.Skel <- MLH1.merge.Skel[ !grepl("x", MLH1.merge.Skel$X) , ]
MLH1.merge.Skel <- MLH1.merge.Skel[!(is.na(MLH1.merge.Skel$quality) | MLH1.merge.Skel$quality==""), ]

MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)

MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)

#where does this DF come from?
#mlh1.skel.og <- add_euth_date(mlh1.skel.og)
#mlh1.skel.og <- add_age(mlh1.skel.og)
#str(mlh1.skel.og)

MLH1.merge.Skel$nMLH1.foci <- as.character(MLH1.merge.Skel$nMLH1.foci)
MLH1.merge.Skel$nMLH1.foci <- as.numeric(MLH1.merge.Skel$nMLH1.foci)

#it seems like this has duplicates
MLH1.merge.Skel.DUPs <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$Random.Name),]
#with new dataframe, there are 0 duplicates..
#make sure to seperate mice based on adult and juvinille

#remove for now
#MLH1.merge.Skel <- MLH1.merge.Skel %>% distinct()

#keep list of outliers (need to adjust threshold with lastest batch)
Total.SC.above.threshold <- MLH1.merge.Skel[MLH1.merge.Skel$skel_size > 25000,]
#remove outliers 
MLH1.merge.Skel <- MLH1.merge.Skel[MLH1.merge.Skel$skel_size < 25000,]

```

```{r write.table, echo=FALSE, eval=FALSE}
#write out these cell's info so I can double check them.
#write.table(Total.SC.above.threshold, "~./MLH1repo/data/BivData/total_SC_above_threshold.csv", sep=",", row.names = FALSE)

#write.table(MLH1.merge.Skel, "~./MLH1repo/data/BivData/total_SC_below_threshold.csv", sep=",", row.names = FALSE)
```

# More Analysis

I noticed that these cells have the same bin sizes and skel sizes
this is written within RW's code

now there are quite a few cells above the threshold
table(Total.SC.above.threshold$category)
more female cells than male

to clean up the total SC data -- use the images with no in redo crop (400 vs 2500)
could also compare to quality =1 cells

```{r ploting, echo=FALSE}
Total.SC.box <- ggplot(data = MLH1.merge.Skel, aes(y=skel_size, x=strain, color=sex) ) + geom_boxplot()+ggtitle("Total SC Measures")

Total.SC.points <- ggplot(data = MLH1.merge.Skel, aes(y=skel_size, x=strain, color=sex) ) + 
  geom_jitter()+ggtitle("Total SC Measures")
```

 test if my score of crop, causes more noise / is a cleaner cell segment
for internal checking, not need to keep

 make tables of the means ect to measure how much change there is between these subseted data sets

```{r crop.stuff, echo=FALSE}

clean.crop <- MLH1.merge.Skel[MLH1.merge.Skel$REDO.crop == 'no',]
table(clean.crop$category)
table(MLH1.merge.Skel$category)

#WTF are these?
clean.crop.table <- ddply(.data=clean.crop, 
                             .(category),
                             summarize, 
                          #mean skel, mean binary, var for the two
                          n.obs = length(unique(Original.Name)),
                          mean.skel = mean(skel_size), 
                          mean.bin = mean(bin_size),
                          var.skel = var(skel_size),
                          var.bin = var(bin_size)
                          #ncells = length(unique(fileName)),
                             #nbivs = length(unique(Obj.ID))
)

not.clean.crop.table <- ddply(.data=MLH1.merge.Skel[!MLH1.merge.Skel$REDO.crop == 'no',], 
                          .(category),
                          summarize, 
                          #mean skel, mean binary, var for the two
                          n.obs = length(unique(Original.Name)),
                          mean.skel = mean(skel_size), 
                          mean.bin = mean(bin_size),
                          var.skel = var(skel_size),
                          var.bin = var(bin_size)
                          #ncells = length(unique(fileName)),
                          #nbivs = length(unique(Obj.ID))

)

weee <- merge(clean.crop.table, not.clean.crop.table, by.x = 'category', by.y = 'category', all = FALSE)

```

table of clean crop stat
table of total (minus clean crop, stat)
 difference in means is the amount of error

for some categories the difference is ~20% between the means
usuall difference means the croped version has less compared to other

```{r more.plot, echo=FALSE}

weee$dif.skel.mean <-  ( (weee$mean.skel.x - weee$mean.skel.y) / weee$mean.skel.x ) *100
#what is this below

Total.SC.points2 <- ggplot(data = clean.crop, aes(y=skel_size, x=strain, color=sex) ) + 
  geom_jitter()+ggtitle("Total SC Measures, clean crops")

#str(MLH1.merge.Skel)
```

```{r table.making, echo=FALSE}
#mouse level
mouse.level.skel <- ddply(.data=MLH1.merge.Skel,
                          .(mouse),
                          summarize, 
                          ncells = length(unique(Original.Name)),
                          nmice = length(unique(mouse)),
                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                          mean.skel = mean(skel_size),
                          mean.bin = mean(bin_size)
)

mouse.level.skel <- add_strain(mouse.level.skel)
mouse.level.skel <- add_sex(mouse.level.skel)
mouse.level.skel <- add_category(mouse.level.skel)
mouse.level.skel <- add_subsp(mouse.level.skel)
```

```{r mouse.level.plot, echo=FALSE, warning=FALSE}

looky.mouse <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Musc",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() +facet_wrap(subsp~sex) + ggtitle("")
looky.mouse

looky.mouseDom <- ggplot(data = mouse.level.skel[mouse.level.skel$subsp == "Dom",], aes(y=mean.skel, x=mean.MLH1, color=strain) ) + geom_jitter() + ylim(c(150,1200))+facet_wrap(subsp~sex) + ggtitle("")

#theme(legend.position="none") 
```

double check a way to visually inspect the python skel output

repeat random numbers?

some of richard's analysis show negative relationship with SC length and MLH1 (his QTL word doc). Could this be sign
 of pwd like regulation?

remove MLH1 tables since they overwrite other ones

remove plots, tables for internal checking
what to keep: SC.skel, OG.MLH1.merge.Skel, MLH1.merge.Skel, MLH1.merge.Skel.DUPs, mouse.level.skel

```{r saving.data, echo=FALSE}
rm(list=setdiff(ls(), c("SC.skel", "OG.MLH1.merge.Skel", "MLH1.merge.Skel", "MLH1.merge.Skel.DUPs","mouse.level.skel")))

#output: matched random names with real names.
#save RData file
#save.image("~./MLH1repo/data/total_SC_1.6.20.RData")

```

### 2020-01-10 src/setupdata_BivData_MLH1repo.R

@@ -98,9 +98,6 @@ RawDataTable <- as.data.frame(colSums(!is.na(BivData)) )
 
 #print out list of mice that have Image folder, but are missing from the BivData
 
-
-
-
 ##############
 # CLEAN DATA #
 ##############

### 2020-01-10 src/setupdata_BivData_MLH1repo_local.R

@@ -1,103 +1,435 @@
-# set up data file for ML data output
-# input: BivData from all the runs
-# output: RData file, Table of counts, cleaned up BivData.csv
+# set up data file for algorithm data output
+# input: Currated BivDataFile
+# output: cleanBivData.csv (write out the )
 
 library(plyr)
 library(lattice)
 library(dplyr)
 library(ggplot2)
 
+
 #ToDo
-#add IFD
-#add the melted foci here
+#need to remake the large full DF -- since the makefile was leaving out rows
 
-#make a master csv file for copying measurements into
+#1. check for dups and make a list for human reference
+#2. pull all the bivalents that have good SC segmentation
+#3. make a master csv file for copying measurements into
 
-#make folder for copying the tif files (move all image output into data/BiVdata/image_output/)
-#(seperate for mat files?)
 setwd("~./MLH1repo/")
 
-#load the master curated list for cross referencing!
-master_curated_BivDatalist.csv = read.csv("data/BivData/FULL_Merged_Curated_BivData.csv")
-#already has Obj.ID and SC/foci pass cols
-#need to clean up all the random numbers in pass.sc column
-
-master_fullBivData = read.csv("data/BivData/MUS_FULL_BIVDATA_10.11.19.csv")
-#33286 obs
-#+PWD and KAZ females 
-#+PWD (male)
-#+KAZ (male)
-#+LEW (female-male)
-
-#add obj.ID
-master_fullBivData$Obj.ID <- paste(master_fullBivData$fileName, master_fullBivData$boxNumber, sep = "_")
-
-#some of the images have bad fileNames, -- so those should be removed, or Idk, checked
-#removed, PWD_f1_sp1  and 8oct14_PWDf2_sp1
-#not sure where these files/images are
-
-anyDuplicated(master_fullBivData$Obj.ID) #23393 duplicated values
-#9893 unique obj.id
-#
-
-#re-write the excel sheet for unique observations
-#mark which have been curated and which havent
-
-#making full bivData --> then requires a curate list (filtered from) Bivdata
-#that requires manual curation
-
-#merge the kaz-pwd datasets
-#MERGE THE msm and lew
-
 source("src/CommonFunc_MLH1repo.R")
[…]

### 2020-01-10 doc/totalSC.Rmd
working on organizing and analyzing all of the BivData. mice from MM have BivData

@@
 load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData.RData")
+
+#Curated list
+
+#strain specific xlsx sheets
 
@@
 
-```{r mixed.models.setup, echo=FALSE, warning=FALSE}
+```{r grab.nice.mice.setup, echo=FALSE, warning=FALSE}
 #https://rcompanion.org/handbook/G_03.html
@@
 DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#above should only apply to males
+
 DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse)|DF.HetC.MixedModel.HQ$mouse==""),]
@@
 ```{r check.lists, echo=FALSE}
-
-MLH1.merge.Skel
-OG.MLH1.merge.Skel
+#MLH1.merge.Skel
+#OG.MLH1.merge.Skel
 
@@
 
+#some WSB females missing from DF.HETC.HQ
+
 #compare this list to cells in whole cell / curate (why?)
@@
 
-#curated Biv data doesn't have '.tif' -- remove .tif
+#OG.MLH1.merge.Skel doesn't have '.tif' -- remove .tif to make matching easier
 skel.mice_from.MM.DF$fileName <- as.character(skel.mice_from.MM.DF$fileName)
@@
   #MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-   templist2 = strsplit(skel.mice_from.MM.DF$fileName[j], split=".tif")#[[1]]
+   templist2 = strsplit(skel.mice_from.MM.DF$fileName[j], split='.tif')[[1]]
    #print(templist2)
-   skel.mice_from.MM.DF$fileName2[j] <- templist2 #should match file name
+   skel.mice_from.MM.DF$fileName2[j] = templist2
 }
+#this loop puts list in fileName2
 
-DF <- skel.mice_from.MM.DF[ skel.mice_from.MM.DF$fileName2 %in% Curated_BivData$fileName, ]#299
-#above is the DF which there are cell which have been curated
-#merge --
+library(tidyr)
 
-DF <- add_strain(DF)
-DF <- add_sex(DF)
-DF <- add_category(DF)
+skel.mice_from.MM.DF$og.fileName <- skel.mice_from.MM.DF$fileName
+skel.mice_from.MM.DF <- skel.mice_from.MM.DF %>% separate(fileName, c("fileName3","blah"),'.tif')
 
-table(DF$category) #missing MSM females
+#cell level data, MLH1 and skel
+DF <- skel.mice_from.MM.DF[ skel.mice_from.MM.DF$fileName3 %in% Curated_BivData$fileName, ]#415
+#above is the DF which there are cell which have been curated (and there is total.sc data)
+#how to merge the biv level data..?
+
+#number of bivs curated?
+
+SC.pass_table <- ddply(.data=Curated_BivData, 
+                 .(fileName),
+                 summarize, 
+                 n.SC.pacc.bivs = length(unique(Obj.ID))
+
+)
+#the PWD and KAZ duplicate mice will have more than 20 -- remove them
+
+SC.pass_table <- add_mouse(SC.pass_table)
+#SC.pass_table <- add_strain(SC.pass_table)
+#SC.pass_table <- add_sex(SC.pass_table)
+#SC.pass_table <- add_category(SC.pass_table)
+
+DF$fileName <- DF$fileName3
+
+#with mouse, merge to DF
+DF.2 <- merge(DF, SC.pass_table, by = "fileName", all.y = FALSE, no.dups = TRUE )
+#not merging because of silly / weird list thing
+
+DF.2 <- DF.2[ -c(14,15,20,23,21) ]
+DF.2$mouse <- DF.2$mouse.x
+
+DF.2 <- add_strain(DF.2)
+DF.2 <- add_sex(DF.2)
+DF.2 <- add_category(DF.2)
+
+table(DF.2$category)
+
+this.plot <- ggplot(DF.2, aes(n.SC.pacc.bivs)) + geom_histogram(bins = 40)+ facet_wrap(~category)+ggtitle("Number of curated bivs per image")
+
+this.plot_WSB.female <- ggplot(DF.2[DF.2$category == "WSB female",], aes(n.SC.pacc.bivs)) + geom_histogram(bins = 40)+ facet_wrap(~mouse)+ggtitle("Number of curated bivs per image")
+
+```
+
+```{r more.blahs, echo=FALSE}
+
+#WSB -- WSB_curation, WSB_female output on big External
+#not sure why Curated bivData catching more categories than this above
 
 #this table now has merged bin/skel, and have been curated (but it doesn't have data on the curated bivalents)
+#the curated list is the current bottle neck
 
-#skel.mice_from.MM.DF$fileName %in% Curated_BivData$fileName
+#I should start inventory  how many mice are in each curated category
 

### 2020-01-10 src/SetupData.Rmd
working on organizing and analyzing all of the BivData. mice from MM have BivData

@@
 original_DF$fileName <- (original_DF$Original.Name)
+
 original_DF <- add_mouse(original_DF)
+original_DF <- add_strain(original_DF)
+original_DF <- add_category(original_DF)
+original_DF <- add_sex(original_DF)
 
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_1.10.20.RData")
 

### 2020-01-10 src/TotalSC.R

@@ -1,218 +1,0 @@
-# Total.SC
-# input: output from RW SC_skel run over Anon_Data images. The file names are random
-# output: DFs of merged MLH1 and Skel data, cell and mouse levels
-
-#merge the sc.skel outputs with the master MLH1 measures
-#match up random names with original names
-
-#the skel output and scripts are in the SC_skel dir. it was run on all images in the anon_data folder. 
-#I haven't manually confirmed that the default parameters are skeletonizing the images correctly.
-
-
-
-
-#ToDo
-#rewrite in Rmd -- to better display notes and findings
-#Check the segmentation
-#re-run python scrtip, (it runs over all anon image files)
-#is it worth croping /cleaning up these images?
-
-#load main MLH1 files to merge MLH1 cell data with the sc data 
-setwd("~./")
-
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#batch18
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent 
-
-#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
-SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
-
-#format all factors to numbers
-SC.skel$bin_size  <- as.character(SC.skel$bin_size )
-SC.skel$skel_size  <- as.character(SC.skel$skel_size)
-
-SC.skel$bin_size  <- as.numeric(SC.skel$bin_size )
-SC.skel$skel_size <- as.numeric(SC.skel$skel_size)
-
-SC.skel$rand_name <- as.character(SC.skel$rand_name)
-
-#check for duplicatations of the random name
-original_DF$Random.Name <- as.character(original_DF$Random.Name)
-duplies <- original_DF[duplicated(original_DF$Random.Name),]
-#3 duplicate rows in original DF -- because Random name was blank -- how will these be matched to total.SC?
-
-#MERGEING MLH! with SKEL
-MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
-
-
[…]

### 2020-01-10 src/makeCurateBivDataList.R

@@ -33,36 +33,51 @@ G_BivData$Obj.ID <- paste(G_BivData$fileName, G_BivData$boxNumber, sep = "_")
 CZECH_BivData = read.csv("data/BivData/algorithm\ output/SKIVE_CZECH_BBIIGG_11.15.19.csv", header=TRUE )
 CZECH_BivData$Obj.ID <- paste(CZECH_BivData$fileName, CZECH_BivData$boxNumber, sep = "_")
 
+#10.1.19
+BBIIGG_phytomorphor_4.9.19_moreFULL
 
+moreFULL_BivData = read.csv("data/BivData/curation/BBIIGG_phytomorphor_4.9.19_moreFULL.csv", header=TRUE )
+moreFULL_BivData$Obj.ID <- paste(moreFULL_BivData$fileName, moreFULL_BivData$boxNumber, sep = "_")
 
 #check for duplicate lines
-anyDuplicated(CZECH_BivData$Obj.ID)
-the.dups <- CZECH_BivData[duplicated(CZECH_BivData$Obj.ID),]
+anyDuplicated(moreFULL_BivData$Obj.ID)
+the.dups <- moreFULL_BivData[duplicated(moreFULL_BivData$Obj.ID),]
 #I think just the headers usually are duplicated
 
 #don't need to add these cols if taken from single batches
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-CZECH_BivData <- add_mouse(CZECH_BivData)
-CZECH_BivData <- add_strain(CZECH_BivData)
-CZECH_BivData <- add_subsp(CZECH_BivData)
-CZECH_BivData <- add_sex(CZECH_BivData)
-CZECH_BivData <- add_category(CZECH_BivData)
+moreFULL_BivData <- add_mouse(moreFULL_BivData)
+moreFULL_BivData <- add_strain(moreFULL_BivData)
+moreFULL_BivData <- add_subsp(moreFULL_BivData)
+moreFULL_BivData <- add_sex(moreFULL_BivData)
+moreFULL_BivData <- add_category(moreFULL_BivData)
 
-table(CZECH_BivData$strain)#only 1 strain
+table(moreFULL_BivData$strain)#only 1 strain
+#more.FULL. WSB, LEW, PWD, MSM, KAZ, SPRET
 
 #SKIVE_bivData <- FULL_BivData_REDO[FULL_BivData_REDO$strain == "SKIVE",]
-CZECH_BivData <- CZECH_BivData[CZECH_BivData$strain == "CZECH",]
[…]

### 2020-01-10 src/setupdata_BivData_MLH1repo_local.R

@@ -15,6 +15,8 @@ library(ggplot2)
 #2. pull all the bivalents that have good SC segmentation
 #3. make a master csv file for copying measurements into
 
+#make OG.curatedBivData (see SC pass =0 | 1)
+
 setwd("~./MLH1repo/")
 
 source("src/CommonFunc_MLH1repo.R")

### 2020-01-17 src/BivData_Setup.Rmd
updated BvData master sheet

---
title: "BivData_Setup"
author: "April Peterson"
date: "January 17, 2020"
output: html_document
---

```{r setup, include=FALSE}

setwd("~./MLH1repo/")
source("src/CommonFunc_MLH1repo.R")

#BIVDATA
new.Curate.FULL <- read.csv("data/BivData/curation/FULL_Curated_BivData.csv")
#80919
#61084
#60241
#60830
#57548

new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
new.Curate.FULL <- add_mouse(new.Curate.FULL)

#make OG copy
new.Curate.FULL.org <- new.Curate.FULL

new.Curate.FULL.org <- add_category(new.Curate.FULL.org)

OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)

OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)

#write the curated file
write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_17.1.20.csv", sep=",",
            row.names = TRUE)

write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_17.1.20.csv", sep=",",
            row.names = TRUE)

```

master curated bivdata file

master_curated_BivData.org = read.csv("data/BivData/FULL_Merged_Curated_BivData.csv")#6859
merge / find what's missing in old a new curate versions
'curate.csv' is the grep string!

load the big one

cleaning steps

show tables for the number of mice with curated biv

```{r CLEAN.UP, echo=FALSE, include=FALSE}

```

# basic stats

# Assemble final BivData DF

# melt position table

needed to related positions

# IFD data

make sure IFDs are applied correctly to the right DFs.

# analysis

planned analyses for later docs, but this is the list for getting everything ready

- chrom class proportions

- SC

### 2020-01-17 src/setupdata_BivData_MLH1repo_local.R

@@ -25,28 +25,47 @@ source("src/CommonFunc_MLH1repo.R")
 
 #master_curated_BivData.org = read.csv("data/BivData/FULL_Merged_Curated_BivData.csv")#6859
 #merge / find what's missing in old a new curate versions
-
+#'curate.csv' is the grep string!
 
 #BIVDATA
 new.Curate.FULL <- read.csv("data/BivData/curation/FULL_Curated_BivData.csv")
+#80919
+#61084
 #60241
 #60830
 #57548
 #this file is made with a make pipeline. cat all .csv files
-
-
 #Currently / best way is to compile this by hand -- currently have make file for compiling 
 
 #str(new.Curate.FULL) #start 27,764
-new.Curate.FULL.org <- new.Curate.FULL
+
 new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
-
-#need to fill in Obj.ID for Gough
 new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
+new.Curate.FULL <- add_mouse(new.Curate.FULL)
 
+
+#make OG copy
+new.Curate.FULL.org <- new.Curate.FULL
+
+new.Curate.FULL.org <- add_category(new.Curate.FULL.org)
+
+OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
+OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)
+
+
+OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
+
+#write the curated file
+write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_17.1.20.csv", sep=",",
+            row.names = TRUE)
+
+
+write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_17.1.20.csv", sep=",",
+            row.names = TRUE)
 
 #bc there are a bunch of dumb chrs, all the useful cols will be factors
 anyDuplicated(new.Curate.FULL$Obj.ID)#360 duplicated
+#16242
 #31186
 #13981 -- 320 rows
 
[…]

### 2020-01-17 src/BivData_Setup.Rmd
working on the BivData setup

@@
 author: "April Peterson"
-date: "January 17, 2020"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 output: html_document
@@
 ```{r setup, include=FALSE}
+library(plyr)
+library(lattice)
+library(dplyr)
+library(ggplot2)
+library(reshape2)
 
 setwd("~./MLH1repo/")
-source("src/CommonFunc_MLH1repo.R")
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
-#BIVDATA
-new.Curate.FULL <- read.csv("data/BivData/curation/FULL_Curated_BivData.csv")
+#BIVDATA MASTER SHEET
+new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
 #80919
@@
 
+#Clean away all the empty rows
 new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
 new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
+colnames(new.Curate.FULL)
 new.Curate.FULL <- add_mouse(new.Curate.FULL)
@@
 new.Curate.FULL.org <- new.Curate.FULL
-
 new.Curate.FULL.org <- add_category(new.Curate.FULL.org)
 
+#write files for basic stats to csv files
 OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
+#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
+
 OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)
-
-OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
 
@@
             row.names = TRUE)
-
+#
 write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_17.1.20.csv", sep=",",
@@
 
+#SC PASS
+new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#8933
+
 ```
+
+Goal of this doc is to make DFs that are ready for the analysis.
+
+Main files from this plot:
+
+  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures
+
+  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  3. MELT_ABS: make melt version of foci positions for abs measuers
+
+  4. MELT_PER: USE of the general foci position plots
 
@@
 
-```{r CLEAN.UP, echo=FALSE, include=FALSE}
+```{r CLEAN.UP.1, echo=FALSE, include=FALSE}
+
+#check for duplicates
+anyDuplicated(new.Curate.FULL$Obj.ID)# 3467
+
+dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
+dup.curated <- add_mouse(dup.curated)
+
+table(dup.curated$mouse)
+
+#which lines are duplicated
+#lots of PWD_female duplicated -- probably how I put together the curated files
+
+#fix these duplicated ones
+
+#remove duplicated
+Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151
+
+#not surethe value of geting these lines
+#new.Curate.FULL <- new.Curate.FULL[ !grepl("[?]", new.Curate.FULL$SC.pass) , ]
+#new.Curate.FULL <- new.Curate.FULL[ !grepl("[?]", new.Curate.FULL$hand.foci.count) , ]
+
+#new.Curate.FULL <- new.Curate.FULL.org[(new.Curate.FULL.org$SC.pass == 1 | new.Curate.FULL.org$SC.pass == 0),]
+#new.Curate.FULL <- droplevels(new.Curate.FULL)
+
+#new.Curate.FULL <- add_mouse(new.Curate.FULL)
+#new.Curate.FULL <- add_category(new.Curate.FULL)
+
+#pass.fail.table <- table(new.Curate.FULL$SC.pass, new.Curate.FULL$category)
+#this outsput shows the relative ratios of pass to fail biv segmentation
+
+Curated_BivData <- add_mouse(Curated_BivData)
+Curated_BivData <- add_category(Curated_BivData)
+Curated_BivData <- add_strain(Curated_BivData)
+Curated_BivData <- add_sex(Curated_BivData)
+Curated_BivData <- add_subsp(Curated_BivData)
+```
+
+```{r set.type, echo=FALSE}
+
+##change factor cols to numbers (factor -> chr -> number)
+Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
+
+##
+Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
+#I think the number of sig figs and rounding changes slightly from the blocks above
+
+#adding more feature calculations (IFD, ect)
+Curated_BivData <- add_IFD(Curated_BivData)
+Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+
+#IFD.2 does mulitple IFDs (just absolute)
+#Curated_BivData <- add_IFD.2(Curated_BivData)
+
+Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
+
+Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
+Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
+Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
+
+Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
+Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
+
+str(Curated_BivData)
+```
+
+IFD functions
+
+  - IFD: currently I think just does 2CO bivalents
+
+  - IFD.2: for mulitple IFDs (3CO ect.)
+
+```{r CLEAN.3}
+
+#report some of these things in the text, for the report
+
+#remove data which has paint marks
+P_fileNames <- Curated_BivData$fileName[(grep('p_rev', Curated_BivData$fileName))]
+length(P_fileNames)
+
+#what is this 
+#prev_num = length(P_fileNames)
+#fullBivData <- fullBivData[!(fullBivData$fileName %in% P_fileNames), ]
+
+#remove anything with 'DUPLICATE'
+DUP_list <- Curated_BivData$fileName[(grep('DUP', Curated_BivData$fileName))]
+DUP_num = length(DUP_list)#0
+
+#fullBivData <- fullBivData[!(fullBivData$fileName %in% DUP_list), ]
+
+#remove images that would be doubles ... 12, 12.1, 12.2
+#this isn't completely effective
+
+dup_list= c()
+for(i in Curated_BivData$fileName){  #change the file name header,  
+  # print(i)
+  ifelse(grepl("\\.[0-9]_rev", i),
+         (dup_list = c(dup_list, i)), NA)
+}
+MultiCells = c()
+for(j in dup_list){
+  stringLIST = strsplit(j, split="\\.")[[1]]
+  new_string = paste(stringLIST[[1]][1], "rev", sep="_")
+  MultiCells = c(MultiCells, new_string)
+}
+
+MultiCells = unique(MultiCells)
+multi_cell_num = length(MultiCells)#118
+
+#test this part of the function
+#I think most of these cell lines have been left out of the final DF
+fullBivData <- fullBivData[!(fullBivData$fileName %in% MultiCells), ]
+#save these number measures to report
 
@@
 
-# basic stats
+remove DUP, 'p', remove mulit cells (10, 10.1, 10.2)
+
+what colnames do all the OFD functions need?
+
+Discribe which images / Bivalents are removed
+
+- Add F4 column!
+
+# Basic stats
+
+- pass rate
+
+- number per mice?
+
+Plots
+
+1. Histograms of N.bivs per cell, facet category
+
+2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)
+
+3.
+
+Is there a way to check my assumptions for these measues?
+
+What are the sources of error fot these data?
+
+```{r cell.table1, echo=FALSE}
+
+cell.level.BivData_table <- ddply(.data=Curated_BivData, 
+                       .(fileName),
+                       summarize, 
+                       boxs.IDd = max(boxNumber),
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+
+                       ncells = length(unique(fileName)),
+                       nbivs = length(unique(Obj.ID)),
+
+#
+ #                      mean.SC_length = mean(chromosomeLength),
+                       var.SC_length = var(chromosomeLength),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3)
+)   
+                       #add the centromere and ifd here?
+                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
+#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
+#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
+#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),
+
+#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
+#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
+ #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),
+
+#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
+#                       var.IFD = var(IFD1, na.rm=TRUE),
+#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
+#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),
+
+ #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
+#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
+#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
+#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)
+
+cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
+cell.level.BivData_table <- add_strain(cell.level.BivData_table)
+cell.level.BivData_table <- add_category(cell.level.BivData_table)
+
+```
 
 # Assemble final BivData DF
+
+decide on the colnames that I want for the final DF
+
+fileName, Object_ID, box.number, centromere_ABS, centromere_PER
+chromosome length, SC.pass, Foci.pass, hand.foci.count, 
+F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD
+mouse, strain, category, subsp,
+
+merge:
+cell.MLH1.count, cell.quality, mouse.age, 
+cell.total_skel, cell.total_binary
 
@@
 
+```{r melt.table}
+
+#paste in this code
+
+coco.names <- colnames(Curated_BivData)
+
+#"fileName" ,  "chromosomeLength",        "centromere_ABS_Position", "centromere_PER_Position",
+#"boxNumber",               "Obj.ID",                  "SC.pass",                 "foci.pass",              
+#"curated.1pass.0fail",     "curation.notes",          "hand.foci.count",        "Foci1",                  
+#"Foci2",       "Foci3",                "notes",                 "mouse",                  
+#"category",                "strain",                  "sex",                    "subsp",                  
+#"IFD1_ABS",                "IFD1_PER",
+
+#all of the cols to use
+MELT_Curated_BivData <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
+              #leave out the Foci
+                "Foci1","Foci2","Foci3")],
+
+            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
+                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))
+
+#dm1$value is the number for the foci position
+#clean up dm1 (there are a bunch of rows without useful information)
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]
+
+#MELT, -- 15
+colnames(MELT_Curated_BivData) <- c(coco.names[c(1:11, 15:16)], "Foci_Type", "Position") #too many foci
+#I'm missing a number of cols -- not
+
+```
+
+```{r MELT.PER}
+
+#run the melting for the PER foci position
+
+colnames(Curated_BivData)
+
+MELT_Curated_BivData.PER <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
+              #leave out the Foci
+                "PER_Foci_1","PER_Foci_2","PER_Foci_3")],
+
+            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
+                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))
+
+MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[(!is.na(MELT_Curated_BivData.PER$value)),]
+
+#MELT, -- 15
+colnames(MELT_Curated_BivData.PER) <- c(coco.names[c(1:11, 15:16)], "Foci_Type", "PER_Position")
+
+#figure out the different hand.foci.count
+
+```
+
+```{r merge.melt}
+#merge the 2 melted DF
+FULL_MELT_Curated_BivData  <- cbind( MELT_Curated_BivData, MELT_Curated_BivData.PER[15])
+
+colnames(FULL_MELT_Curated_BivData) <- c(FULL_MELT_Curated_BivData[1:14], "Pos_ABS", "Pos_PER")
+
+#
+MELT_Curated_BivData.PER <- add_strain(MELT_Curated_BivData.PER)
+MELT_Curated_BivData.PER <- add_sex(MELT_Curated_BivData.PER)
+
+```
+
+```{r general.foci.pos, echo=FALSE}
+
+#make the normalized foci density plots
+MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
+MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]
+
+MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]
+
+#remove 0 and NA
+#remove 4foci count?
+
+#check for values over 1
+
+density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")
+
+density
+#table(the4$mouse)
+#something wrong with the 4 bivalents
+#many of the lines just have 3 lines
+
+```
+
 # IFD data
@@
 
-# analysis
+```{r save.file}
+
+#save Rdata
+save.image("~./MLH1repo/data/CleanBivData.RData")
+#save.image("data/BivData_4_PCA.RData")
+
+```
+
+# Number of whole cells
+
+- list the cells with all bivs measued or which are close to full.
+
+# Analysis
 
@@
 
+Use the Bivalent_Analysis.Rmd for analysis pieces
+BivPattern.Rmd started to analyze this but, not a very clear setup
+
 - chrom class proportions
 
-- SC
+- SC lengths, by CO number
+
+-ge

### 2020-01-20 doc/Bivalent_Analysis.Rmd
work on draft for Bivdata analysis

---
title: "Single Bivalent Level Analysis"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}

library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(reshape2)

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

#load Clean BivData
load(file="data/CleanBivData_1.20.20.RData")

#load other data files, MLH1
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18

load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#fyi loading mulitple RData files doesn't earase ther files -- but it can overwrite files named the same thing

#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

# Setup and Discription of Dataset

This plot illustrates how well the algorithm seperated indiviual bivalents per cell image. 
Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally segmented.

```{r tech.ect, echo=FALSE, eval=FALSE}

nBiv.per.cell <- ggplot(cells.BivData_table, aes(nbivs)) + ggtitle("Number of passing bivalents per cell") + geom_histogram(bins = 30) +facet_wrap( ~ category)
nBiv.per.cell

#for cells with close to 20 - they are almost complete
#they should be used to

```

consider making a table with predictions for bivalent patterns which are related to the mechanism

# Chrm Class proportions

Should I be thinking of making a MM or lm for predicting CO number (within a category)

but there is so little variance in the CO number

co ~ sc.length

I ran this model before,  I remember that these models are better predictors (more significant for males than females)

maybe -- female nuanced pattern -- falls along the lines of

simple predictions not met as strongly as in male patterns...

this reasoning might be complicated by the fact that the range in CO averages is lower in females -- so there is less power

# Applying Mix Model framework

list out pros and cons for using mouse averages of bivalent traits instead of using the single bivalent traits them selves.

**Pro Bivalent level**  
- The data are continous and closer to normal/gaussian expectations than the MLH1 counts.

- There is within cell variance that is an issue.

**Pro Mouse level averages**  
- Mouse average is a conserved framework, it would summarize general patterns which fits with the paper.

- Simple application of the same MM.

- Limited work to add to the dataset if I can justify that the collection of single bivalents are random and equivilent across mice.

# Heterochiasmy Predictions

Using the Mixed model framework which tests the effects (and interactions) of subspecies, sex and strain, I will test for evolution of the following traits.

Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);

1) SC length (cite)

2) normalized CO positions (sedell and Kirkpatrick).

Interference, measured in the physical units, has been shown to lack sexual dimorphism (petkov).

These 3 hypotheses will be tested with the mixed model framework, specifically focusing on the effect of sex.

Mouse averages for the each metric will be used.

# SC lengths

```{r show.SC.length}
#compile data
#main figures, show both mouse and bivalent level

#Curated_BivData

#Mouse_average - BivTable?

#for now remove the 
Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count <5,]

first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")
first.SC.plot
#add boxplots for mouse level means
#add second x axis, hand.foci.count

other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)) ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")

SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))

#sfsL <- ggplot(lowGEN, aes(variable, prop, fill=as.factor(Group.2))) + geom_col(position = "dodge") + #scale_fill_viridis(name="Migration Rate", discrete = TRUE) + theme(axis.text.x=element_blank()) + #xlab("Frequency in the sample") + ylab("Proportion of Total Junctions") + ggtitle("C. Generation = #2,000") + ylim(0,1)

#geom_boxplot(alpha = .2, position = "dodge")+

```

First draft of the SC length plot. I'd like to add the hand foci count to the x axis.
(also add the PWD - male data)

```{r justify.mouse.level}

#justify mouse averages
#is sampling -- bivalent dataset 
# random samples 
#equivalent sample size

```

$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$

```{r MM.SC.Length}

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

#check all the strains and male ages, should be the same for MM

DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

#mouse.level.table
#make this in datasetup
mouse.avs <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = format(round(   var(chromosomeLength),3), nsmall=3),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

mouse.avs <- add_sex(mouse.avs)
mouse.avs <- add_strain(mouse.avs)
mouse.avs <- add_subsp(mouse.avs)

#unorder the factors, but figure our
mouse.avs$strain <- unclass(mouse.avs$strain)
mouse.avs$strain <- as.factor(mouse.avs$strain)

mouse.avs$subsp <- unclass(mouse.avs$subsp)
mouse.avs$subsp <- as.factor(mouse.avs$subsp)

mouse.avs$sex <- unclass(mouse.avs$sex)
mouse.avs$sex <- as.factor(mouse.avs$sex)

BivData.MixedModel.HQ <- lme(mean_SC ~ subsp * sex, data=mouse.avs, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )

#sex is significant! with prediction

```

# IFD

```{r show.IFD, echo=FALSE}

#first draft of the IFD comparisons

#for now remove the 
Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]

first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
first.IFD.plot_ABS

first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))

#+theme(legend.position="none")
first.IFD.plot_PER

```

Above is the first plots of IFDs. Each point is a IFD observation. colors are by mouse. 
Maybe add mouse level lines for the means

Maybe I should try binned the SC lengths.

mixed model analysis for IFD (interference)

$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$

Double Check the dataset for the mixed models.

```{r MM.IFD, echo=FALSE}

#set up data set for the mixed model
#Curated_BivData

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

#check all the strains and male ages, should be the same for MM

DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

#mouse.level.table
#make this in datasetup
mouse.avs.table1 <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),

                   mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),

                   mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                        cV = cv(chromosomeLength),
                        var = format(round(   var(chromosomeLength),3), nsmall=3),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

mouse.avs.table1 <- add_sex(mouse.avs.table1)
mouse.avs.table1 <- add_strain(mouse.avs.table1)
mouse.avs.table1 <- add_subsp(mouse.avs.table1)

#unorder the factors, but figure our
mouse.avs.table1$strain <- unclass(mouse.avs.table1$strain)
mouse.avs.table1$strain <- as.factor(mouse.avs.table1$strain)

mouse.avs.table1$subsp <- unclass(mouse.avs.table1$subsp)
mouse.avs.table1$subsp <- as.factor(mouse.avs.table1$subsp)

mouse.avs.table1$sex <- unclass(mouse.avs.table1$sex)
mouse.avs.table1$sex <- as.factor(mouse.avs.table1$sex)

mouse.avs.table1 <- mouse.avs.table1[(!is.na(mouse.avs.table1$mean_IFD_ABS)),]

BivData_IFD_ABS.MixedModel.HQ <- lme(mean_IFD_ABS ~ subsp * sex, data=mouse.avs.table1, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(BivData_IFD_ABS.MixedModel.HQ)$tTable, 6 ) )
#ABD - IFD not sex specific effect

BivData_IFD_PER.MixedModel.HQ <- lme(mean_IFD_PER ~ subsp * sex, data=mouse.avs.table1, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(BivData_IFD_PER.MixedModel.HQ)$tTable, 6 ) )
#sex is a little significat

```

Preliminary results from an independant dataset indecated that PWD had longer IFDs, which goes agaisnt the simple prediction of more COs ~ denser spacing of foci on the same bivalent. This also indecated that interference distance may evolved in the house mouse complex.

```{r compare.IFD}

```

# General CO Positions

```{r CO.positions}
#main figure of density plots,
#MELT_Curated_BivData.PER

MELT_Curated_BivData.PER <- add_subsp(MELT_Curated_BivData.PER)

Musc.female_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "female") & (MELT_Curated_BivData.PER$subsp == "Musc") ), ]

CO.Pos.density_Musc_female <- ggplot(Musc.female_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc female \nNormalized Foci Positions")
CO.Pos.density_Musc_female

#Dom female
Dom.female_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "female") & (MELT_Curated_BivData.PER$subsp == "Dom") ), ]

CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
CO.Pos.density_Dom_female

#Musc male
Musc.male_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "male") & (MELT_Curated_BivData.PER$subsp == "Musc") ), ]

CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
CO.Pos.density_Musc_male

#Dom male
Dom.male_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "male") & (MELT_Curated_BivData.PER$subsp == "Dom") ), ]

CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
CO.Pos.density_Dom_male

```

Split up the plot data by subsp - sex. 
Try adding in the numbers of observations?
think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.

$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$

```{r MM_CO_pos, echo=FALSE}

#use the 1CO
#Use MELT DF
#MELT_Curated_BivData

Pos.CO.avs <- ddply(MELT_Curated_BivData.PER, c("mouse", "Foci_Type"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                    mean_PER_pos = as.numeric(format(round(  mean(PER_Position), 3 ), nsmall=3) )
)

Pos.CO.avs <- add_sex(Pos.CO.avs)
Pos.CO.avs <- add_strain(Pos.CO.avs)
Pos.CO.avs <- add_subsp(Pos.CO.avs)

#unorder the factors, but figure our
Pos.CO.avs$strain <- unclass(Pos.CO.avs$strain)
Pos.CO.avs$strain <- as.factor(Pos.CO.avs$strain)

Pos.CO.avs$subsp <- unclass(Pos.CO.avs$subsp)
Pos.CO.avs$subsp <- as.factor(Pos.CO.avs$subsp)

Pos.CO.avs$sex <- unclass(Pos.CO.avs$sex)
Pos.CO.avs$sex <- as.factor(Pos.CO.avs$sex)

#isolate the 1COPos.CO.avs
Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "PER_Foci_1",]

Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]

BivData.CO.PER.POS.MixedModel.HQ <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
#sex is significant
```

```{r pos_t.test}
#make point plots / boxplots which show differences in mean positions

#scatter + boxplot for t.tests

```

Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He sugggested doing simple t-test for the positions

```{r F1.by.F2}

#try remaking the plot Megan suggested
# for 2CO positions, Foci1, Position  on x and Foci 2 position on y

#isolate 2COs
#facet by sex and subsp

#F1.x.F2 <- ggplot(MELT_Curated_BivData.PER, aes(x=F1, fill=F2) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")

```

```{r CO.pos.ANOVA, eval=FALSE}
#what is the pattern of variance
#run analyses for each subsp*sex
#use non-melt DF

#how is the variance partioned across
#cell, mouse, strain

female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
female.Dom <- female.Dom[female.Dom$subsp == "Dom",]

female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength

#unorder strain and mouse

female.Dom$mouse <- as.factor(female.Dom$mouse)

female.Dom$strain <- unclass(female.Dom$strain)
female.Dom$strain <- as.factor(female.Dom$strain)

female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]

#1CO first
modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)

#can't get mouse and strain to have sum of square
#residual size decreases with per.F1
#residuals much larger than fileName, mouse and strain no

#model <- lm(breaks ~ wool * tension, 
#            data = warpbreaks, 
#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))

male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
male.Dom <- male.Dom[male.Dom$subsp == "Dom",]

male.Dom$mouse <- as.factor(male.Dom$mouse)

male.Dom$strain <- unclass(male.Dom$strain)
male.Dom$strain <- as.factor(male.Dom$strain)

male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]

male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength

male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
summary(aov(male.modo))

#only file name is registering as effect
#Review ANOVA frameworks
#http://www.biostathandbook.com/nestedanova.html

```

# old drafts

I went on a kick thinking of the sis-coten metric, I'm not so sure this is a good metric anymore.

```{r block1, echo=FALSE}
#new.Curate.FULL

#new.Curate.FULL <- add_strain(new.Curate.FULL)
#new.Curate.FULL <- add_sex(new.Curate.FULL)
#Curated_BivData

Curated_BivData <- subset(Curated_BivData, hand.foci.count < 5)

staggered <- ggplot(Curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("SC lengths by foci count")+theme(legend.position="none") #+ ylim(c(30,250))

fill.by.ifd <- ggplot(Curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=IFD1_PER))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("SC length Fill by IFD")#+theme(legend.position="none") #+ ylim(c(30,250))
```

 Predictions for sis-coten stabilizing selection

Because it enables tighter regulation of the spindle assembly checkpoint (SAC) the centrosomic metaphase spindle in male gametogenesis can impose selection on relative CO placement (broad scale REC landscape). (OR generally due to the stronger sensitivity of the SAC in sperm/males). Following certain assumptions, the axis length and relative positions of COs can change the amount sister-cohesin over which tension is generated on the metaphase I spindle. This can be thought of as the bivalent structure when they are in the process of bipolar orientation.

under strong purifying selection, (the idea is that bivalent structures evolve to efficienctly pass the SAC) (there might be mulitple ways to do this)

- similar amount of sister cohesion tension -related to the rate of sister cohesion degregation. So all bivalents seperate at a similar rate (enter anaphase and complete anaphase in a synchronized manner).

- tension for SAC detection

 1. In MSM and PWD males (the high recombining strains), 1COs should have less of a telomere bias compared to the other low rec mice (IFD).

The MSM and PWD males have approximately 50% 1CO, 50% 2CO compared to lower recombining strains (Dom and musc) 75% 1CO, 25% 2CO.

If the metaphase spindle / SAC evolved a new optimal amount of sister cohesion for the spindle tension to be spread over, the parameters of relative co placement and CO interference can evolve so that they are similar across .

```{r rand.1CO.pos, echo=FALSE, warning=FALSE, eval=FALSE}

PER.1CO.pos <- ggplot(Curated_BivData[(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$sex == "male"),], aes(y = PER_Foci_1, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(aes(alpha=.5))+ggtitle("Relative CO Position of 1CO across strains")

LOG.PER.1CO.pos <- ggplot(Curated_BivData[(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$sex == "male"),], aes(y = log(PER_Foci_1), x = strain, color=strain)) + geom_jitter()+ geom_boxplot(aes(alpha=.5))+ggtitle("Log of Relative CO Position of 1CO across strains")

#log(foci position, makes a bigger range), means don't really change
#but WSB, G, LEW seem to have stronger centromere supression
PER.1CO.pos
LOG.PER.1CO.pos
```

In terms of percent position, PWD, LEW and MSM have similar mean Foci1 pos. LEW is sort of the odd one out. Interestingly, LEW seems to have stronger centromere supression compared to PWD and MSM (the distribution of foci positions is larger in MSM, PWD and KAZ).

 2. Sister cohesin tension area for 1COs should trend more similar to 2COs in high recombining strains.

what's the rational for this prediction?

```{r siscoten.stabilizing, echo=FALSE, warning=FALSE, eval=FALSE}

plot.1CO <- ggplot(Curated_BivData[(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$sex == "male"),], aes(y = dis.cent, x = chromosomeLength, color=strain)) + geom_jitter()+facet_wrap(~strain)

sis.coten <- ggplot(Curated_BivData[(Curated_BivData$sex == "male"),], aes(y = SisCoTen, x = chromosomeLength, color=strain)) + geom_jitter()+facet_wrap(~strain)

sis.coten.boxplot <- ggplot(Curated_BivData[(Curated_BivData$sex == "male"),], aes(y = SisCoTen, x = strain,  color=hand.foci.count)) +geom_boxplot()+geom_jitter()

#
sfsL <- ggplot(Curated_BivData[(Curated_BivData$sex == "male"),], aes(strain, SisCoTen, fill=as.factor(hand.foci.count))) + geom_col(position = "dodge")

sfsL <- ggplot(Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$hand.foci.count < 3),], aes(strain, SisCoTen, fill=as.factor(hand.foci.count))) + geom_boxplot(aes()) + geom_jitter(aes(color=hand.foci.count))

Curated_BivData$PER.siscoten <- Curated_BivData$SisCoTen / Curated_BivData$chromosomeLength

boxplot_per.siscoten <- ggplot(Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, PER.siscoten, fill=as.factor(hand.foci.count))) + geom_boxplot(aes()) +  facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("PERCENT SIS-COTEN by mouse")+theme(axis.title.x=element_blank(), axis.text.x=element_blank())

boxplot_siscoten.mouse <- ggplot(Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, SisCoTen, fill=as.factor(hand.foci.count))) + geom_boxplot(aes())+ facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("SISCOTEN by mouse") + theme(axis.title.x=element_blank(), axis.text.x=element_blank())

boxplot_siscoten.mouse
boxplot_per.siscoten

#+ scale_fill_viridis(name="Migration Rate", discrete = TRUE) + theme(axis.text.x=element_blank()) + xlab("Frequency in the sample") + ylab("Proportion of Total Junctions") + ggtitle("C. Generation = 2,000") + ylim(0,1)
```

Mouse is shown on the x axis. It seems like PWD and MSM might have a higher variance in the range od siscoten values for 2COs, but overall the averages sis-co-tens across the strains for 2CO are very similar across all of the mice.

### 3. Lack of strain specific patterns in females

```{r female.siscoten,eval=FALSE,echo=FALSE, warning=FALSE}

boxplot_siscoten.mouse.female <- ggplot(Curated_BivData[(Curated_BivData$sex == "female") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, SisCoTen, fill=as.factor(hand.foci.count))) + geom_boxplot(aes())+ facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("SISCOTEN by mouse, female") + theme(axis.title.x=element_blank(), axis.text.x=element_blank())+ylim(c(0,150))
boxplot_siscoten.mouse.female

boxplot_PERsiscoten.mouse.female <- ggplot(Curated_BivData[(Curated_BivData$sex == "female") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, PER.siscoten, fill=as.factor(hand.foci.count))) + geom_boxplot(aes())+ facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("PER SISCOTEN by mouse, female") + theme(axis.title.x=element_blank(), axis.text.x=element_blank())
boxplot_PERsiscoten.mouse.female

```

the vls are much closer in females

# SC length traits

 Females have higher variance in more meiotic traits

```{r female.SC.plots, echo=FALSE, warning=FALSE}

SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))

SC.lengths.2CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 2,], aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("2COs male and females") +ylim(c(30,250))

SC.lengths.1CO
SC.lengths.2CO
```

The above plots seperate bivalent classes. 
(Make a table of pvalues for t.tests of male to female SC lengths across strains.)

```{r anova, echo=FALSE}
#try doing anova

ano <- aov(lm(chromosomeLength ~ hand.foci.count+sex + strain,  Curated_BivData))
summary(ano)
```

By ANOVA analysis, everything has a significant effect.

## PERMUTATIONS Accounting for sample size differences

Currently the category with the fewest observations is G female (~30). I also have may more Kaz male bivalent measures (1000)

`r kable(table(Curated_BivData$category))`

Below is code where I implement a subsampling approach/permute dataset with randomly sampled bivalent observations. 25 random bivalents are sampled from each category, and this is permuted 1000 times.

The plots shown below are histograms of p values from t.tests for **SC length** across the labeled categories from the replicated/subsampled dataset.

```{r sub.sample, echo=FALSE, eval=FALSE}

#the problem with sub-sampling is that there are so many groups I need to compare
#Easist is probably to recreate the whole data set
#G female have the lowest number of obs, 30

#sample from all the categories, the samplings are mini DFs
#-- then merge them together

#outer loop is the permutations / replications
BIG.DF.LIST <- replicate(1000,

        #filling the permuted DF is inner loop                 
         #add subsample to a list, this works!!!
list( bind_rows(list(

  Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ],
                                SubSample.WSB.male <-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB male"), Biv.sample.Number ), ],

                                Curated_BivData[ sample( which(Curated_BivData$category=="G female"), Biv.sample.Number ), ],
                                Curated_BivData[ sample( which(Curated_BivData$category=="G male"), Biv.sample.Number ), ],

                       Curated_BivData[ sample( which(Curated_BivData$category=="LEW female"), Biv.sample.Number ), ], 
                       Curated_BivData[ sample( which(Curated_BivData$category=="LEW male"), Biv.sample.Number ), ],

                       Curated_BivData[ sample( which(Curated_BivData$category=="PWD female"), Biv.sample.Number ), ], 
                       Curated_BivData[ sample( which(Curated_BivData$category=="PWD male"), Biv.sample.Number ), ],

                       Curated_BivData[ sample( which(Curated_BivData$category=="MSM female"), Biv.sample.Number ), ], 
                       Curated_BivData[ sample( which(Curated_BivData$category=="MSM male"), Biv.sample.Number ), ],

                        Curated_BivData[ sample( which(Curated_BivData$category=="KAZ male"), Biv.sample.Number ), ], 
                       Curated_BivData[ sample( which(Curated_BivData$category=="KAZ female"), Biv.sample.Number ), ],

                       Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE female"), Biv.sample.Number ), ], 
                       Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE male"), Biv.sample.Number ), ]))
)
      )
#Accessing big list / matrix
#the replicated DFs are now is a list of DFs
#head(BIG.DF.LIST[[2]]) returns head of the 2nd replicated DF

#Loop through the list of DFs and do the t.tests!
#i is this index of the dataframes in the list?

p.vals.DF <- data.frame(sex.test.pval = rep(0, length(BIG.DF.LIST) ) )
for(i in 1:length(BIG.DF.LIST)){  #current data from
  #print(i)
  #test sex
  p.vals.DF$sex.test.pval[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[BIG.DF.LIST[[i]]$sex == "female"],
         BIG.DF.LIST[[i]]$chromosomeLength[BIG.DF.LIST[[i]]$sex == "male"], paired = FALSE)$p.value

   #test female subsp
   p.vals.DF$female.subsp[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")], paired = FALSE )$p.value

   #test male subsp
     p.vals.DF$male.subsp[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")],paired = FALSE )$p.value

      p.vals.DF$male.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value

      p.vals.DF$male.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            p.vals.DF$male.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value

      #what about a null test?

      #female    
      p.vals.DF$female.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value

            #female    
      p.vals.DF$female.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            #female    
      p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value

      #test male strains (Dom)
     p.vals.DF$male.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value

      p.vals.DF$male.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value

      p.vals.DF$male.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            p.vals.DF$male.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value

       #test female strains (Dom)
     p.vals.DF$female.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value

      p.vals.DF$female.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value

      p.vals.DF$female.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            p.vals.DF$female.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value
}

#these tests might be more appropriate for averages (mouse averages)?

#melt the DF's within the BIG list for easier ggploting
Melted.pval.DF <- melt(p.vals.DF, c("sex.test.pval",    "female.subsp",     "male.subsp",       "male.PWD.MSM",     "male.PWD.KAZ",     "male.PWD.SKIVE",   "female.PWD.MSM", "female.PWD.KAZ",   "female.PWD.SKIVE", "male.WSB.G", "male.WSB.LEW", "male.WSB.KAZ",  "male.WSB.PWD","female.WSB.G", "female.WSB.LEW", "female.WSB.KAZ", "female.WSB.PWD" ), 
                         id=c() )

#clean up dm1 (there are a bunch of rows without useful information)
Melted.pval.DF <- Melted.pval.DF[(!is.na(Melted.pval.DF$value)),]
colnames(Melted.pval.DF) <- c("Test.type", "pval")

#BIG ggplot collections (skip the first type?)
biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
#I could also add the observed t.test pvalue as verticle line...
biggie
#want to try spreading the data across mulitple plots
#this is not saving mulitple pdfs
#for (i in seq(1, length(unique(Melted.pval.DF$Test.type)), 6)) {
#  pdf(paste0("tryplot_",i,".pdf"), 7, 5)
# 
#  print(ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests"))
#
# dev.off()
#
# }

#pvals.subsamp.DF.1 <- ggplot(p.vals.DF, aes(male.PWD.MSM ) )+geom_histogram(bins = 50)#PWD and MSM not sig different (with randomized subsample)
#pvals.subsamp.DF.B <- ggplot(p.vals.DF, aes(male.PWD.KAZ ) #)+geom_histogram(bins = 50)
#pvals.subsamp.DF.C <- ggplot(p.vals.DF, aes(male.PWD.SKIVE ) #)+geom_histogram(bins = 50)

#Fpvals.subsamp.DF.A <- ggplot(p.vals.DF, aes(female.PWD.MSM ) #)+geom_histogram(bins = 50)
#Fpvals.subsamp.DF.B <- ggplot(p.vals.DF, aes(female.PWD.KAZ ) #)+geom_histogram(bins = 50)
#Fpvals.subsamp.DF.C <- ggplot(p.vals.DF, aes(female.PWD.SKIVE ) #)+geom_histogram(bins = 50)
#these 3 female plots have close to uniform pvalus dist

#summaries to compare / test
#SC.by.sex.T.test <- t.test(subsampled.DF$chromosomeLength[subsampled.DF$sex == "female"],
#                           subsampled.DF$chromosomeLength[subsampled.DF$sex == "male"], paired = #FALSE)$p.value

#SubSample.WSB.female  <-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
#SubSample.WSB.male <-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB male"), Biv.sample.Number ), ]
#SubSample.G.female <- Curated_BivData[ sample( which(Curated_BivData$category=="G female"), Biv.sample.Number ), ]
#SubSample.G.male <- Curated_BivData[ sample( which(Curated_BivData$category=="G male"), Biv.sample.Number ), ]
#SubSample.LEW.female <- Curated_BivData[ sample( which(Curated_BivData$category=="LEW female"), Biv.sample.Number ), ]
#SubSample.LEW.male <- Curated_BivData[ sample( which(Curated_BivData$category=="LEW male"), Biv.sample.Number ), ]
#SubSample.PWD.female <- Curated_BivData[ sample( which(Curated_BivData$category=="PWD female"), Biv.sample.Number ), ]
#SubSample.PWD.male <- Curated_BivData[ sample( which(Curated_BivData$category=="PWD male"), Biv.sample.Number ), ]
#SubSample.MSM.female <- Curated_BivData[ sample( which(Curated_BivData$category=="MSM female"), Biv.sample.Number ), ]
#SubSample.MSM.male <- Curated_BivData[ sample( which(Curated_BivData$category=="MSM male"), Biv.sample.Number ), ]
#SubSample.KAZ.female <- Curated_BivData[ sample( which(Curated_BivData$category=="KAZ female"), Biv.sample.Number ), ]
#SubSample.KAZ.male <- Curated_BivData[ sample( which(Curated_BivData$category=="KAZ male"), Biv.sample.Number ), ]
#SubSample.SKIVE.female <- Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE female"), Biv.sample.Number ), ]
#SubSample.SKIVE.male <- Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE male"), Biv.sample.Number ), ]

#how do I make this replicate loop so that is returns the T.test pvalue in a DF?
# 1. make a big list of many subsampled DFs
# 2. make a loop for applying t.test across all of the

#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
#n use the var.equal = TRUE
#https://www.statmethods.net/stats/ttest.html
```

```{r lm.sc, echo=FALSE}
#should cell be random?
#should I compare mouse means instead?
sc.model.1CO <- lm(chromosomeLength ~ sex + strain +mouse, data = Curated_BivData[Curated_BivData$hand.foci.count == 1,])

sc.model.2CO <- lm(chromosomeLength ~ sex + strain+fileName, data = Curated_BivData[Curated_BivData$hand.foci.count == 2,])

```

Modeling SC lengths at the bivalent level is noisey. There are many levels and unequal sampling.

maybe I should compare the longest bivalent from a cell?  or the shortest 4?

# Centromere supression

Anecdotally I would say that males have stronger centromere supression. I will also add a metric for telomere distance.

```{r cent.supression.plot, echo=FALSE, warning=FALSE, eval=FALSE}
#I need to change the purple color

#simple plots for centromere - dis
cent.distance.all <- ggplot(Curated_BivData, aes(y = dis.cent, x=strain, color=sex))+geom_boxplot(aes(alpha=0.1))
#+geom_jitter(aes(alpha=0.05))+  ggtitle("1COs male and females")

cent.distance.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = dis.cent, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females")

cent.dist.2CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 2,], aes(y = dis.cent, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("2COs male and females")

cent.distance.all
cent.distance.1CO
cent.dist.2CO
```

Intesting that the distance from centromere is sexually dimorphic for 2CO bivalents but not 1CO bivalents. Females have a stronger centromere supression, strange. Did I calculate this wrong?
This indecates that female has stronger centromere supression, but only for 2CO bivalents...

```{r telomere.bias, echo=FALSE, eval=FALSE}

#I have to change the t.tests
#dis.cent

#BIG.DF.LIST
Cent.Dist.p.vals.DF <- data.frame(sex.cent.dist = rep(0, length(BIG.DF.LIST) ) )
for(i in 1:length(BIG.DF.LIST)){  #current data from
  #print(i)

      #female vs male
      Cent.Dist.p.vals.DF$sex.cent.dist[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" )],paired = FALSE )$p.value

#subsp
Cent.Dist.p.vals.DF$female.subsp[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")],paired = FALSE )$p.value

Cent.Dist.p.vals.DF$male.subsp[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")],paired = FALSE )$p.value

Cent.Dist.p.vals.DF$male.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$male.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            Cent.Dist.p.vals.DF$male.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value

      #what about a null test?

      #female    
      Cent.Dist.p.vals.DF$female.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value

            #female    
      Cent.Dist.p.vals.DF$female.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            #female    
      Cent.Dist.p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$female.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value

      #test male strains (Dom)
     Cent.Dist.p.vals.DF$male.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$male.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$male.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            Cent.Dist.p.vals.DF$male.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value

       #test female strains (Dom)
     Cent.Dist.p.vals.DF$female.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$female.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value

      Cent.Dist.p.vals.DF$female.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value

            Cent.Dist.p.vals.DF$female.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value
}

#melt the DF's within the BIG list for easier ggploting
Cent.Melted.pval.DF <- melt(Cent.Dist.p.vals.DF, c("sex.cent.dist",  "female.subsp","male.subsp","male.PWD.MSM",     "male.PWD.KAZ","male.PWD.SKIVE", "female.PWD.MSM", 
                              "female.PWD.KAZ",   "female.PWD.SKIVE", "male.WSB.G", "male.WSB.LEW", "male.WSB.KAZ",  "male.WSB.PWD","female.WSB.G", "female.WSB.LEW", "female.WSB.KAZ", "female.WSB.PWD" ), 
                         id=c() )

#clean up dm1 (there are a bunch of rows without useful information)
Cent.Melted.pval.DF <- Cent.Melted.pval.DF[(!is.na(Cent.Melted.pval.DF$value)),]
colnames(Cent.Melted.pval.DF) <- c("Test.type", "pval")

biggie2 <- ggplot(Cent.Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+ ggtitle("Cent Distance pvalues for 1000 replicated t.tests")+facet_wrap(.~Test.type) 
#sex difference is skewed
biggie2

```

Verbalizing the inference from all the p values.

Using the same permuted / subsampled data frame -- I'll look at centromere and telomere distance

How would these distributions change normalized distances?

t.tests for centromere dis and telomere distance
This result and the IFD could be the same mechanism.

## Higher recombining males have stronger interference

Do the same figure things for IFD

There might be too little 2CO observations for many dom chrm to get the IFDs

`r kable(table(Curated_BivData$category[Curated_BivData$hand.foci.count == 2]))`

```{r interference, echo=FALSE, eval=FALSE}
#for interference I might have to subsample only from 2COs

#have to seperate the hand foci number

IFD.p.vals.DF <- data.frame(sex.IFD.test = rep(0, length(BIG.DF.LIST) ) )
for(i in 1:length(BIG.DF.LIST)){  #current data from
  #print(i)

      #female vs male
      IFD.p.vals.DF$sex.IFD.test[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "female" ) & (BIG.DF.LIST[[i]]$hand.foci.count == 2) ],
         BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & (BIG.DF.LIST[[i]]$hand.foci.count == 2)], na.action=na.omit, paired = FALSE )$p.value

  #subsp
  IFD.p.vals.DF$female.subsp[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
         BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],paired = FALSE )$p.value

IFD.p.vals.DF$male.subsp[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
         BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE, na.action=na.omit)$p.value

#strain comarisons
IFD.p.vals.DF$male.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value

#IFD.p.vals.DF$male.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
#         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value

#IFD.p.vals.DF$male.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( #BIG.DF.LIST[[i]]$strain == "WSB") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
#         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain #== "G")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value

#females  (not enough PWD-MSM)
#IFD.p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == #"female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
#         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain #== "SKIVE")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value

}

#melt the DF's within the BIG list for easier ggploting
IFD.Melted.pval.DF <- melt(IFD.p.vals.DF, c( "sex.IFD.test","female.subsp", "male.subsp", "male.PWD.MSM"), 
  id=c() )

#  "sex.cent.dist",  "female.subsp","male.subsp","male.PWD.MSM",     "male.PWD.KAZ","male.PWD.SKIVE", "female.PWD.MSM", 
#                              "female.PWD.KAZ",   "female.PWD.SKIVE", "male.WSB.G", "male.WSB.LEW", "male.WSB.KAZ",  #"male.WSB.PWD","female.WSB.G", "female.WSB.LEW", "female.WSB.KAZ", "female.WSB.PWD" ),

#clean up dm1 (there are a bunch of rows without useful information)
IFD.Melted.pval.DF <- IFD.Melted.pval.DF[(!is.na(IFD.Melted.pval.DF$value)),]
colnames(IFD.Melted.pval.DF) <- c("Test.type", "pval")

biggie3 <- ggplot(IFD.Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+ ggtitle("ABS IFD pvalues for 1000 replicated t.tests")+facet_wrap(.~Test.type) 
#sex difference is skewed
biggie3

```

The

## more variation across traits for females than males

## High recombining male mice have more total SC?

I think with total SC, this is true, but it doesn't seem true for the single bivalent data. This could also be due to sampling issues.

```{r total.sc, echo=FALSE}
#FIND TOTAL sc code
```

```{r block2, echo=FALSE, eval=FALSE}
#Curated_BivData
# mmore SC based plots

#Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count <= 3,]
#Curated_BivData99 <- droplevels(Curated_BivData99)
#why does this making so many Fing NAs?

#use subset instead!!

Curated_BivData99 <- subset(Curated_BivData, hand.foci.count <= 3)

Curated_BivData <- subset(Curated_BivData, hand.foci.count <= 3)

Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$sex) ), ]

#this plot is too busy for all the data 
SC.length.points.bivs_Musc <- ggplot(Curated_BivData[Curated_BivData$subsp == "Musc",], aes(x=dis.cent, y=chromosomeLength, color= as.factor(mouse) ))+
  geom_jitter()+ xlim(c(0,120)) + ylim(c(40,120))+
  facet_grid(sex ~ hand.foci.count) + ggtitle("")+theme(legend.position="none")

SC.length.points.bivs_Dom <- ggplot(Curated_BivData[Curated_BivData$subsp == "Dom",], aes(x=dis.cent, y=chromosomeLength, color= as.factor(mouse) ))+
  geom_jitter()+ xlim(c(0,120)) + ylim(c(40,120))+
  facet_grid(sex ~ hand.foci.count) + ggtitle("")

#Making centromere rect
  #geom_vline(aes(xintercept = mean.cent.dis.PWD$mean.dist_cent ) ) +
  #annotate("rect", 
 # xmin =  ( mean.cent.dis.PWD$mean.dist_cent - (mean.cent.dis.PWD$dis.cent)*2), 
#  xmax = ( mean.cent.dis.PWD$mean.dist_cent + (mean.cent.dis.PWD$dis.cent)*2),
#              ymin=0,  ymax=Inf, alpha=0.2, fill="purple" ) +
#  labs(x="", y='Density')

```

# Interfocal comparisons

```{r IFD, echo=FALSE, eval=FALSE}

IFD.ABS.comp <- ggplot(Curated_BivData, aes(y = IFD1, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("IFD1 raw distributions")+theme(legend.position="none")

IFD.PER.comp <- ggplot(Curated_BivData, aes(y = IFD1_PER, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("IFD1 normalized distributions")+theme(legend.position="none")+ylim(c(0,1.05))

IFD.ABS.comp
IFD.PER.comp
```

The above plots shows the distribution the raw and normalized distribution of the first Inter Focal Distance.

# Sis Co ten Metric

```{r sis.co.ten, echo=FALSE, eval=FALSE}

sis.co.ten.hist.full <- ggplot(Curated_BivData, aes(SisCoTen)) + geom_histogram(bins = 50) +facet_wrap( ~ category)+ggtitle("Distribution of Sis-co-ten per bivalent")

sis.co.ten.hist <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 2,], aes(SisCoTen)) + geom_histogram(bins = 50) +facet_wrap(hand.foci.count ~ category)+ggtitle("")
#the male histograms look almost bimodal, and the female data looks more uniform
#the telomere bias skews the distribution to lower sis-co-ten measures
#PWD has sig larger

sis.co.ten.hist.full
sis.co.ten.hist

#ARE THE SIS_CO_TEN distributions different across foci.number for males and females
#males; yes?
#females

#sis.co.ten.hist.poin <- ggplot(master_curated_BivData, aes(x=SisCoTen, y=chromosomeLength, color= as.factor(hand.foci.count)  ))+
#  geom_point(aes(alpha=.5)) +facet_wrap(~ category)

#sis.co.ten.hist.poin99 <- ggplot((master_curated_BivData[1:50,]), aes(x=SisCoTen, y=chromosomeLength, color= #as.factor(hand.foci.count)  ))+
#  geom_point(aes(alpha=.5)) +facet_wrap(~ mouse)
```

The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole. Alternating IFDs. Assuming chromatid interference and no sister seperation, this should reflect the amount of sister cohesion that is pulled across the opposite pole.

need more MSM male curated measures!

### 2020-01-20 src/BivData_Setup.Rmd
work on draft for Bivdata analysis

@@
 ```
+
+ToDo:
+
+Add the new curated data.
 
@@
 #I think most of these cell lines have been left out of the final DF
-fullBivData <- fullBivData[!(fullBivData$fileName %in% MultiCells), ]
+#fullBivData <- fullBivData[!(fullBivData$fileName %in% MultiCells), ]
 #save these number measures to report
@@
 - number per mice?
+
+- number of pass.SC w/out foci position data
+
+- number of foci with 'faint' mentioned in the notes
 
@@
 F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD
+
+telomere_dist, cent_dist,
+
 mouse, strain, category, subsp,
@@
 
+```{r telo.cent_dist, echo=FALSE}
+
+#code for cal1 the telomere distance
+#space between last foci and telomere end
+
+#telo_dist <- if(hand.foci = 1){
+          #SC.Length - F1
+#}
+#elseif(hand_foci == 2){ #SC.length - F2}
+
+#code for centromere
+#space between first foci and centromere/centromere-end
+
+#cent_end_dist
+#cent_dist
+```
+
 # melt position table
@@
 ```{r melt.table}
+
+#make sure the this MELT_Curated_BivData.PER is made
+#
 
@@
 
+This is proof that the plot codes work, but move the figures in the analysis doc. remeber this is just data-setup.
+
+think about weighing the density plots by the number of bivalents per category...
+
 # IFD data
@@
 
+# Save
+
 ```{r save.file}
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData.RData")
+save.image("~./MLH1repo/data/CleanBivData_1.20.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-01-21 src/BivData_Setup.Rmd
cleaning and updates to the setup BivData file

--- 2020-01-17 src/BivData_Setup.Rmd
@@
 ---
-title: "BivData_Setup"
+title: "BivData Setup"
 author: "April Peterson"
-date: "January 17, 2020"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 output: html_document
@@
 
+# Data setup and basic cleaning
+
 ```{r setup, include=FALSE}
+library(plyr)
+library(lattice)
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+library(raster)
 
 setwd("~./MLH1repo/")
-source("src/CommonFunc_MLH1repo.R")
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
-#BIVDATA
-new.Curate.FULL <- read.csv("data/BivData/curation/FULL_Curated_BivData.csv")
+#BIVDATA MASTER SHEET
+new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
+#83976
+#83009
+#80898
 #80919
@@
 #60830
-#57548
 
+#Clean away all the empty rows
 new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
-new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
-new.Curate.FULL <- add_mouse(new.Curate.FULL)
 
+#Apply the cleaning steps below to the FULL-OG DF
+
+#remove non number charaecters for the FULL / OG DF
+#Curated_BivData <- Curated_BivData[ !grepl("[?]", Curated_BivData$SC.pass) , ]
+#Curated_BivData <- Curated_BivData[ !grepl("[?]", Curated_BivData$hand.foci.count) , ]
+
+#remove data which has paint marks
+#P_fileNames <- Curated_BivData$fileName[(grep('p_rev', Curated_BivData$fileName))]
+#length(P_fileNames)
+
+#remove anything with 'DUPLICATE'
+#DUP_list <- Curated_BivData$fileName[(grep('DUP', Curated_BivData$fileName))]
+#DUP_num = length(DUP_list)#0
+
+#dup_list= c()
+#for(i in Curated_BivData$fileName){  #change the file name header,  
+#  # print(i)
+#  ifelse(grepl("\\.[0-9]_rev", i),
+#         (dup_list = c(dup_list, i)), NA)
+#}
+#MultiCells = c()
+#for(j in dup_list){
+#  stringLIST = strsplit(j, split="\\.")[[1]]
+#  new_string = paste(stringLIST[[1]][1], "rev", sep="_")
+#  MultiCells = c(MultiCells, new_string)
+#}
+#MultiCells = unique(MultiCells)
+#multi_cell_num = length(MultiCells)#118
+
+#new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
+#colnames(new.Curate.FULL)
+
+new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
 #make OG copy
 new.Curate.FULL.org <- new.Curate.FULL
-
 new.Curate.FULL.org <- add_category(new.Curate.FULL.org)
 
-OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
+#write files for basic stats to csv files
+#OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
+
+#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
+
 OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)
 
-OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
+#write the curated file
+#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_21.1.20.csv", sep=",",
+#            row.names = TRUE)
+#
+#write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_21.1.20.csv", sep=",",
+#            row.names = TRUE)
 
-#write the curated file
-write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_17.1.20.csv", sep=",",
-            row.names = TRUE)
+#SC PASS
+new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#10480 #9819, #9769
 
-write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_17.1.20.csv", sep=",",
-            row.names = TRUE)
+#fix these duplicated ones
+#remove duplicates
+Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151
+
+Curated_BivData <- add_mouse(Curated_BivData)
+Curated_BivData <- add_category(Curated_BivData)
+Curated_BivData <- add_strain(Curated_BivData)
+Curated_BivData <- add_sex(Curated_BivData)
+Curated_BivData <- add_subsp(Curated_BivData)
+
+Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
+#can I write a quick dummy check?
+#there are 2 lines with 0 hand.foci, but with a foci position
 
@@
 
-master curated bivdata file
+edit / clean up the lines with bad hand.foci.count (61)
 
-master_curated_BivData.org = read.csv("data/BivData/FULL_Merged_Curated_BivData.csv")#6859
-merge / find what's missing in old a new curate versions
-'curate.csv' is the grep string!
+ToDo:
 
-load the big one
+1.clean up the non-numbered curated values (for the og df)
 
-cleaning steps
+2.Goal of this doc is to make DFs that are ready for the analysis.
 
-show tables for the number of mice with curated biv
+3.remove the 2 main sources of duplicate lines; 6mar16_PWD_m1, 1mar15_PWD_m1
 
-```{r CLEAN.UP, echo=FALSE, include=FALSE}
+4. decide on the right number of passing curated bivalents per category I should settle for
+Justify why the current number is enough -- or subsample -- to random equivilent sized data sets
+
+5. make sure that the saved Rdata has all the right cols and tables for the next analyses
+
+6. consider integrating manual hand / whole cell measures (check for DUPs ect.)
+
+Main files from this plot:
+
+  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
+
+  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
+
+  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  4. MELT_ABS: make melt version of foci positions for abs measuers
+
+  5. CELL and
+
+  6. mouse table
+
+```{r CLEAN.UP.1, echo=FALSE, include=FALSE}
+
+#check for duplicates, and track them down
+anyDuplicated(new.Curate.FULL$Obj.ID)  #5335, #3467
+
+dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
+dup.curated <- add_mouse(dup.curated)
+
+#table(dup.curated$mouse)
+#1mar15_PWD_m1
+
+#which lines are duplicated
+#lots of PWD_female duplicated -- probably how I put together the curated files
+
+#new.Curate.FULL <- new.Curate.FULL.org[(new.Curate.FULL.org$SC.pass == 1 | new.Curate.FULL.org$SC.pass == 0),]
+#new.Curate.FULL <- droplevels(new.Curate.FULL)
+
+#new.Curate.FULL <- add_mouse(new.Curate.FULL)
+#new.Curate.FULL <- add_category(new.Curate.FULL)
+
+#pass.fail.table <- table(new.Curate.FULL$SC.pass, new.Curate.FULL$category)
+#this outsput shows the relative ratios of pass to fail biv segmentation
 
@@
 
-# basic stats
+```{r set.type, echo=FALSE}
+
+##change factor cols to numbers (factor -> chr -> number)
+Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
+
+##Numeric
+Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
+#I think the number of sig figs and rounding changes slightly from the blocks above
+
+#adding more feature calculations (IFD, ect)
+Curated_BivData <- add_IFD(Curated_BivData)
+Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+
+#IFD.2 does mulitple IFDs (just absolute)
+#Curated_BivData <- add_IFD.2(Curated_BivData)
+
+Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
+
+Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
+Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
+Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
+Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)
+
+#not working
+Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
+Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
+
+#add telomere.dist
+#str(Curated_BivData)
+```
+
+IFD functions
+
+  - IFD: currently I think just does 2CO bivalents
+
+  - IFD.2: for mulitple IFDs (3CO ect.)
+
+# Cell Level table
+
+- pass rate (don't need the OG.DF file, because I can count the number of boxes with the max index number)
+
+- number bivs per mice?
+
+- number cells represented per mouse
+
+- number of pass.SC w/out foci position data
+
+- number of foci with 'faint' mentioned in the notes
+
+```{r cell.level.table, echo=FALSE}
+
+cell.level.BivData_table <- ddply(.data=Curated_BivData, 
+                       .(fileName),
+                       summarize, 
+                       boxs.IDd = max(boxNumber),
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                       ncells = length(unique(fileName)),
+                       nbivs = length(unique(Obj.ID)),
+
+                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       var.SC_length = var(chromosomeLength),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number
+
+                       cv.SC = cv(chromosomeLength),
+
+#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
+                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),
+
+                       mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                      var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
+                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
+                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)
+
+#cent_dis
+                      mean_cent_dist = mean(dis.cent, na.rm=TRUE)
+#telomere dist
+
+#can't think of a why to calq this by just 2CO
+
+)
+                       #add the centromere and ifd here?
+                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
+#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
+#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
+#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),
+
+#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
+#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
+ #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),
+
+#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
+#                       var.IFD = var(IFD1, na.rm=TRUE),
+#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
+#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),
+
+ #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
+#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
+#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
+#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)
+
+cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
+cell.level.BivData_table <- add_strain(cell.level.BivData_table)
+cell.level.BivData_table <- add_category(cell.level.BivData_table)
+
+```
+
+`r head(cell.level.BivData_table)`
+
+# Mouse leve table
+
+the mouse table will be the pooled stats / summary of all SC.passing +
+
+```{r mouse.table, echo=FALSE}
+
+mouse.level.BivData_table <- ddply(.data=Curated_BivData, 
+                       .(mouse),
+                       summarize,
+
+                       ncells = length(unique(fileName)),
+
+                       boxs.IDd = max(boxNumber),  #
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                       ncells = length(unique(fileName)),
+                       nbivs = length(unique(Obj.ID)),
+
+                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3)
+)
+
+```
+
+`r head(mouse.level.BivData_table)`
+
+analysis using the mouse level table -- to justify that using mouse averages is a good idea.
+
+Plots
+
+1. Histograms of N.bivs per cell, facet category
+
+2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)
+
+3.
+
+Is there a way to check my assumptions for these measues?
+
+What are the sources of error fot these data?
+
+```{r telo.cent_dist, echo=FALSE}
+
+#code for cal1 the telomere distance
+#space between last foci and telomere end
+
+#telo_dist <- if(hand.foci = 1){
+          #SC.Length - F1
+#}
+#elseif(hand_foci == 2){ #SC.length - F2}
+
+#code for centromere
+#space between first foci and centromere/centromere-end
+
+#cent_end_dist
+#cent_dist
+```
+
+# melt position table
+
+make the melt DF for the abs (raw) foci positions and then calculated the normalized (PER) foci positions from that table
+
+```{r MELT.block1, echo=FALSE}
+
+col.names.list <- colnames(Curated_BivData)
+#"fileName" ,  "chromosomeLength",        "centromere_ABS_Position", "centromere_PER_Position",
+#"boxNumber",               "Obj.ID",                  "SC.pass",                 "foci.pass",              
+#"curated.1pass.0fail",     "curation.notes",          "hand.foci.count",        "Foci1",                  
+#"Foci2",       "Foci3",                "notes",                 "mouse",                  
+#"category",                "strain",                  "sex",                    "subsp",                  
+#"IFD1_ABS",                "IFD1_PER",
+
+#all of the cols to use
+MELT_Curated_BivData <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
+              #leave out the Foci
+                "Foci1","Foci2","Foci3", "Foci4")],
+
+            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
+                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]
+
+#MELT, -- 15
+colnames(MELT_Curated_BivData) <- c(col.names.list[c(1:11, 16:17)], "Foci_Type", "Pos") #too many foci
+
+MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)
+
+#calq normalized melt pos
+MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]
+
+```
+
+```{r general.foci.pos, echo=FALSE, eval=FALSE}
+
+#make the normalized foci density plots
+#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
+#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]
+
+#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]#
+
+#remove 0 and NA
+#remove 4foci count?
+
+#check for values over 1
+
+#density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")
+
+#density
+#table(the4$mouse)
+#something wrong with the 4 bivalents
+#many of the lines just have 3 lines
+
+```
+
+This is proof that the plot codes work, but move the figures in the analysis doc. remeber this is just data-setup.
+
+think about weighing the density plots by the number of bivalents per category...
 
@@
 
-# melt position table
+Main files from this plot:
 
-needed to related positions
+  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures
 
-# IFD data
+  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
 
-make sure IFDs are applied correctly to the right DFs.
+  3. MELT_ABS: make melt version of foci positions for abs measuers
 
-# analysis
+  4. CELL and mouse table
+
+  5.
+
+MAIN BIVDATA_DF
+decide on the colnames that I want for the final DF
+
+fileName, Object_ID, box.number, centromere_ABS, centromere_PER
+chromosome length, SC.pass, Foci.pass, hand.foci.count, 
+F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD
+
+telomere_dist, cent_dist,
+
+mouse, strain, category, subsp,
+
+merge:
+cell.MLH1.count, cell.quality, mouse.age, 
+cell.total_skel, cell.total_binary
+
+# Save
+
+```{r save.file}
+
+#save Rdata
+save.image("~./MLH1repo/data/CleanBivData_1.21.20.RData")
+#save.image("data/BivData_4_PCA.RData")
+
+```
+
+# Number of whole cells
+
+- list the cells with all bivs measued or which are close to full.
+
+# Analysis
 
@@
 
+Use the Bivalent_Analysis.Rmd for analysis pieces
+BivPattern.Rmd started to analyze this but, not a very clear setup
+
 - chrom class proportions
 
-- SC
+- SC lengths, by CO number

### 2020-01-21 doc/Bivalent_Analysis.Rmd
merged in manual whole cell data to the BivData

@@
 #load Clean BivData
-load(file="data/CleanBivData_1.20.20.RData")
+load(file="data/CleanBivData_1.21.20.RData")
 
@@
 
-```{r tech.ect, echo=FALSE, eval=FALSE}
-
-nBiv.per.cell <- ggplot(cells.BivData_table, aes(nbivs)) + ggtitle("Number of passing bivalents per cell") + geom_histogram(bins = 30) +facet_wrap( ~ category)
+```{r tech.ect, echo=FALSE, eval=TRUE}
+
+nBiv.per.cell <- ggplot(cell.level.BivData_table, aes(nbivs)) + ggtitle("Number of passing bivalents per cell") + geom_histogram(bins = 30) +facet_wrap( ~ category)
 nBiv.per.cell
@@
 
+#This isn't the same as curated/passing bivalents
+
 ```
@@
 
+```{r Chrm.props, echo=FALSE}
+
+#barplots for proportions of chrm classes
+
+#consider adding this the the setup block (could have done this to category level table)
+mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$CO0 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$CO1 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$CO2 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$CO3 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$CO4 / mouse.level.BivData_table$nboxes_passed
+
+#DF needs to be melted -- so they all in 1 col
+#per code, I've done this by hand
+
+#subset DF for ease
+skinny.mouse.level <- mouse.level.BivData_table[-c(11:15 ) ]
+col.names.list.99 <- colnames(skinny.mouse.level)
+
+#"mouse"         "ncells"        "boxs.IDd"      "nboxes_passed" "Nbiv"          "CO0"           "CO1"       #[8] "CO2"           "CO3"           "CO4"           "per.0CO"       "per.1CO"       "per.2CO"       "per.3CO"      [15] "per.4CO"
+
+#all of the cols to use
+MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4",
+
+               #leave out the Foci
+                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")], 
+  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4"))
+
+#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
+
+#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
+
+MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
+
+#then this DF has to be summarized into categories
+chrm.class.category <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
+                             mean.percent.type = mean(value)
+                             #sum is the wrong!
+                             #used average percentage, might consider a different summary
+)
+
+#ggplot
+
+chrm.class.barplot <- ggplot(data = chrm.class.category, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
+chrm.class.barplot
+
+```
+
+the above plot displays the average proportion of each type of chromosome class across mice. The results mostly align with the hand measured results.
+
+Look into ordering the mouse strains to highlight the MSM and PWD differences, renameing the chrm type labels, spacing out the x axis labels or removing some of the other mice.
+
 # Applying Mix Model framework
@@
 **Pro Bivalent level**
+
 - The data are continous and closer to normal/gaussian expectations than the MLH1 counts.
@@
 **Pro Mouse level averages**
+
 - Mouse average is a conserved framework, it would summarize general patterns which fits with the paper.
@@
 
-# Heterochiasmy Predictions
+## Predictions, Heterochiasmy
 
@@
 
-1) SC length (cite)
+1) SC length (cite Lynn)
 
@@
 
-# SC lengths
-
-```{r show.SC.length}
+# SC Lengths
+
+```{r show.SC.length, echo=FALSE}
 #compile data
@@
 
-```{r justify.mouse.level}
-
-#justify mouse averages
-#is sampling -- bivalent dataset 
-# random samples 
-#equivalent sample size
-
-```
+TRY ADDING NESTED boxplots for hand.foci.count
 
@@
 
-```{r MM.SC.Length}
+Above is the model we are testing evolution of SC lengths.
+
+```{r MM.SC.Length, echo=FALSE}
 
@@
 
+The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)
+
+```{r justify.mouse.level, echo=FALSE}
+
+#start the permutation code here
+#make a randomized BivData DF
+
+#30 cells worth of bivs? 
+#30 cells * ( 19 or 20 bivs ) * 70% segment rate = decent sample size for comparing Biv-based patterns?
+# I think this would address if the sample sizes are affecting the pattern
+
+#
+
+#justify mouse averages
+#is sampling -- bivalent dataset 
+# random samples 
+#equivalent sample size
+
+```
+
 # IFD
 
-```{r show.IFD, echo=FALSE}
+**I**nter**f**ocal **D**istance, a indication of CO interference.
+
+DO I have anything to say....
+
+How can I highlight the result (I need to verify this first) for greater IFD in high rec male mice?
+
+The raw measures are pretty nosy, this might not be the best way to display these results.
+
+```{r show.IFD, echo=FALSE, warning=FALSE}
 
@@
 
-Double Check the dataset for the mixed models.
+$$mouse \ average \ normalized IFD ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
+
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure
 
@@
 
+sex may be a significant effect for the percent (normalized IFD). Is there a way to measure / account for the SC.length effect?
+
 Preliminary results from an independant dataset indecated that PWD had longer IFDs, which goes agaisnt the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
@@
 #main figure of density plots,
-#MELT_Curated_BivData.PER
-
-MELT_Curated_BivData.PER <- add_subsp(MELT_Curated_BivData.PER)
-
-Musc.female_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "female") & (MELT_Curated_BivData.PER$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_female <- ggplot(Musc.female_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc female \nNormalized Foci Positions")
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+Musc.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_female <- ggplot(Musc.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc female \nNormalized Foci Positions")
 CO.Pos.density_Musc_female
@@
 #Dom female
-Dom.female_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "female") & (MELT_Curated_BivData.PER$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
 CO.Pos.density_Dom_female
@@
 #Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "male") & (MELT_Curated_BivData.PER$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
 CO.Pos.density_Musc_male
@@
 #Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData.PER[ ((MELT_Curated_BivData.PER$sex == "male") & (MELT_Curated_BivData.PER$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER_Position, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
 CO.Pos.density_Dom_male
@@
 ```{r MM_CO_pos, echo=FALSE}
-
 #use the 1CO
@@
 
-Pos.CO.avs <- ddply(MELT_Curated_BivData.PER, c("mouse", "Foci_Type"), summarise,
+Pos.CO.avs <- ddply(MELT_Curated_BivData, c("mouse", "Foci_Type"), summarise,
                         Nmice = length(unique(mouse)),
@@
 
-                    mean_PER_pos = as.numeric(format(round(  mean(PER_Position), 3 ), nsmall=3) )
+                    mean_PER_pos = as.numeric(format(round(  mean(PER.Foci.Pos), 3 ), nsmall=3) )
 )
@@
 #isolate the 1COPos.CO.avs
-Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "PER_Foci_1",]
+Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
 
@@
 
-BivData.CO.PER.POS.MixedModel.HQ <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
+BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
 #sex is significant
@@
 
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
 #isolate 2COs
@@
 
-#F1.x.F2 <- ggplot(MELT_Curated_BivData.PER, aes(x=F1, fill=F2) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
 

### 2020-01-21 src/BivData_Setup.Rmd
merged in manual whole cell data to the BivData

@@
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+whole.cell.data_manual.OG <- read.csv("~./MLH1repo/whole_cell_measures.csv")
 
@@
 #            row.names = TRUE)
+```
+
+```{r split.up.DF, echo=FALSE}
 
@@
 ```
+
+# Manual Whole Cell
+
+```{r whole.cell.manual.clean, echo=FALSE}
+#compare and clean up whole cell measures
+
+whole.cell.data_manual <- whole.cell.data_manual.OG[whole.cell.data_manual.OG$chrm.class != "XY",]
+
+colnames(whole.cell.data_manual.OG)
+#"return5"         "fileName"        "boxNumber"       "centromere"      "SC.length"  #    
+ #[6] "rank"            "chrm.class"      "hand.foci.count" "F1"              "F2"    #         
+#[11] "F3"              "notes"           "X"               "X.1"            
+#remove, X, X.1
+whole.cell.data_manual <- whole.cell.data_manual[-c(1, 13:14)]
+whole.cell.data_manual$Foci4 <- ""
+
+#rename foci positions
+whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
+whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
+whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3
+
+#mouse, strain, sex, category
+whole.cell.data_manual <- add_mouse(whole.cell.data_manual)
+whole.cell.data_manual <- add_category(whole.cell.data_manual)
+whole.cell.data_manual <- add_strain(whole.cell.data_manual)
+whole.cell.data_manual <- add_sex(whole.cell.data_manual)
+whole.cell.data_manual <- add_subsp(whole.cell.data_manual)
+
+#whole.cell.data_manual <-   whole.cell.data_manual[(!is.na(whole.cell.data_manual$fileName)),]
+whole.cell.data_manual <- whole.cell.data_manual[!(is.na(whole.cell.data_manual$fileName) | whole.cell.data_manual$fileName==""), ]
+
+#obj.id
+whole.cell.data_manual$Obj.ID <- paste(whole.cell.data_manual$fileName, whole.cell.data_manual$boxNumber, sep = "_")
+
+#add all the extra columns, blah
+
+#percent Foci position
+
+#ADD RANK
+for(cellys in as.character(unique(whole.cell.data_manual$fileName))){
+
+ # print(cellys)
+  chrm.order = rank(whole.cell.data_manual$SC.length[whole.cell.data_manual$fileName == cellys])
+  whole.cell.data_manual$rank[whole.cell.data_manual$fileName == cellys] <- chrm.order  
+  }
+
+#IFD 
+whole.cell.data_manual <- add_IFD(whole.cell.data_manual)
+
+```
+
+ToDo
+
+- Go back and clean up the original excel data
+
+- clean the bottom, fill in missing data
+
+- check the order of foci positions for 5aug17_14jul17_LEW_f1_sp1_11_rev_blob4.
 
@@
 
+as.data.frame( head(cell.level.BivData_table) )
+
 ```
 
-`r head(cell.level.BivData_table)`
+```{r whole.cell_merging, echo=FALSE}
+
+#check if there are fileNames
+double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% Curated_BivData$fileName, ]
+
+#which cell-levels fileNames have close to 15-17 chrms?
+
+almost_whole_cell <- cell.level.BivData_table[(cell.level.BivData_table$Nbiv < 21),]
+almost_whole_cell <- almost_whole_cell[almost_whole_cell$Nbiv > 15,]
+
+#there is some discrpancy with the Nbiv and nbivs columns! ugh why, some have 20 and 10
+
+#get the Biv data 
+almost_whole_cell_BivData <- Curated_BivData[ Curated_BivData$fileName %in% almost_whole_cell$fileName, ]
+
+as.data.frame(  table( droplevels( almost_whole_cell_BivData$fileName) ) )
+length(unique(almost_whole_cell_BivData$fileName))
+
+#what next?
+
+```
+
+3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.
+
+8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev
 
@@
 
+as.data.frame( head(mouse.level.BivData_table) )
+#kable didn't work
+
 ```
-
-`r head(mouse.level.BivData_table)`
 
@@
 
-# Number of whole cells
-
-- list the cells with all bivs measued or which are close to full.
-
 # Analysis

### 2020-01-22 doc/Bivalent_Analysis.Rmd
a few small notes from home

@@
 
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 #setwd("C:/Users/April/Desktop/MLH1repo")
@@
 
+setwd("~./MLH1repo/")
 #load Clean BivData
@@
 #load other data files, MLH1
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")#added batch18
+
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
 
@@
 Look into ordering the mouse strains to highlight the MSM and PWD differences, renameing the chrm type labels, spacing out the x axis labels or removing some of the other mice.
+
+#chromosome Size effect
+
+because of chromosome size effect on CO position and rate, I'd like to try seperate bins of the 'loss' bivalents
+how does within cell variance in SC length compare across cells?
+
+```{r chrm.size.effect, echo=FALSE}
+
+show.the.cells_females <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "female",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+ facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
+show.the.cells_females
+#PWD seems to have the biggest variance across cells
+
+#the min is pretty stable
+
+show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+ facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
+show.the.cells_male
+
+#WSB has more min
+
+#almost_whole_cell
+
+#almost_whole_cell_BivData
+almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
+
+almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
+almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
+
+show.the.almost_cells_DOM_m <- ggplot(almost_whole_cell_DOM[almost_whole_cell_DOM$sex == "male",], aes(x=fileName, y= chromosomeLength, color=strain))+
+  geom_point()+ facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell BivData measures")
+
+show.the.almost_cells_DOM_m
+```
+
+given the distributions of sc lengths within cells -- do I think I can predict a random SC's length (rank)
 

### 2020-01-22 doc/Bivalent_Analysis.Rmd
first pass at analysis of Mixed models, outlined other predictions, made some notes in other files

--- 2020-01-21 doc/Bivalent_Analysis.Rmd
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-output: html_document
+output:
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '5'
+    toc_float: yes
+  pdf_document: default
 ---
@@
 ```{r setup, include=FALSE}
-
 library(knitr)
@@
 
+# Chromosome Size effect
+
+In an attempt to deal with the chromosome effect size that is know to effect total rate and CO position, we try to make a dataset of the longest 25% SC from a cell.
+
+The majority of cells from the BivData set do not have the complete completement of bivalents, (ie they are missing up to 5) -- which makes it hard to use a absolute length threshold since some cells have shorter or longer bivalents compared to the while.
+So an absolute measure might catch the lower 25% of some cells.
+
+I'd like to figure out how to best seperate out the true 4th quartille bivalents (longest 25%), even with this missing data. double check that 4th quartille is top 5/4
+
+Could I use the total SC? (I would be warry: can't ensure that they apply the same skeletonize algorithm/parameters and the XY is also included in that count and will mess everything up.)
+
+outline logic for filtering / choosing BivData set of almost whole cells
+
+- 16 to 20 bivalents in the cells
+
+```{r chrm.size.effect, echo=FALSE}
+
+#calq the mouse means
+manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
+                        median_SC = median(SC.length, na.rm = TRUE),
+                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
+#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
+
+)
+
+manual_summary_stats <- add_mouse(manual_summary_stats)
+manual_summary_stats <- add_strain(manual_summary_stats)
+manual_summary_stats <- add_sex(manual_summary_stats)
+manual_summary_stats <- add_subsp(manual_summary_stats)
+
+###
+#calq the mouse means
+Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        median_SC = median(chromosomeLength),
+                        Qar.4 = quantile(chromosomeLength)[[4]],
+
+                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
+        #add Q4
+
+)
+Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
+
+show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
+show.the.cells_females_dom
+#PWD seems to have the biggest variance across cells
+
+almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
+almost_cells_DOM_f
+
+#### Musc female
+
+show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
+show.the.cells_females_musc
+
+almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
+almost_cells_Musc_f
+
+### Dom Male
+show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
+show.the.cells_male_dom
+
+almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
+almost_cells_Dom_m
+
+#MUSC male
+show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
+show.the.cells_male_musc
+
+almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
+almost_cells_Musc_m
+
+#the min is pretty stable
+
+#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
+
+#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+ # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
+#show.the.cells_male
+
+#almost_whole_cell_BivData
+#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
+#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
+#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
+
+```
+
+All the plots above show the distrtributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
+It shows the amount of within cell variance across strains.
+There is a bit of variance across the SC length distributions in the PWD females.
+
+I'm purposing to take all the automated biv measures above the cell's 4th quartille as a dataset for 'longestest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromsome size effects.
+
+But this dataset might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females). But it's easier to just run another set of analysis.
+
+```{r long.chrm.dataset, echo=FALSE, eval=FALSE}
+
+#how many biv's are from cells with >15 bivalents
+
+al.15_bivs <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]#150 cells have al 15
+
+long_BivData_set <- Curated_BivData[ Curated_BivData$fileName %in% al.15_bivs$fileName, ] #across those 150 cells are 2654 biv measures
+
+#table(potential.data.set$category)
+long_BivData_set$fileName <- as.character(long_BivData_set$fileName)
+Biv_Data_mean_table$fileName <- as.character(Biv_Data_mean_table$fileName)
+
+#pull out the biv's which exceed the cell's Q4
+#for each line of long_BivData_set
+# check if it's SC length is greater than the Q4 for the given cell
+
+long_BivData_set99 <- long_BivData_set
+
+for(  i in 1:length( long_BivData_set99$fileName)  ){
+  long_BivData_set99$cell.Q4[i] <- Biv_Data_mean_table[ which(Biv_Data_mean_table$fileName == long_BivData_set99$fileName[i]), ]
+}
+
+for(  k in 1:length( long_BivData_set99$fileName)  ){
+long_BivData_set99$above.Q4[k] <- ifelse( (long_BivData_set99$cell.Q4 < long_BivData_set99$chromosomeLength), "yes", "no")
+}
+#This isn't working either
+
+#for(i in 1:length( long_BivData_set$fileName)  ){ 
+  # print(i)
+#  current_obj_fileName <- long_BivData_set$fileName[i]
+#  current_obj_SC_length <- long_BivData_set$chromosomeLength[i]  
+#  current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == current_obj_fileName) ]
+# current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == long_BivData_set$fileName[i]) ]
+#  long_BivData_set$above.Q4 <-ifelse( (current.Q4 < current_obj_SC_length), 1, 0)
+#  print( paste( c(current.Q4, current_obj_fileName, current_obj_SC_length), sep = "_" ) )
+#}
+#this loop isn't working
+
+```
+
+The dataset is unbalanced in some ways -- integrating manual measures might fix. MSM under rep, since most cell had high overlap.
+
 # Applying Mix Model framework
 
+I will use two different mixed models using lme() and lmer(), because the different models have limitations for reporting p values ect.
+
 list out pros and cons for using mouse averages of bivalent traits instead of using the single bivalent traits them selves.
@@
 
+```{r clean.biv.data4.MM, echo=FALSE}
+
+#remove all the non-used strains!!
+Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
+Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
+Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]
+Curated_BivData$strain <- droplevels(Curated_BivData$strain)
+#MOLF
+
+#non-ordered factors, 
+Curated_BivData$subsp <- factor(Curated_BivData$subsp, ordered = FALSE)
+Curated_BivData$sex <- factor(Curated_BivData$sex, ordered = FALSE)
+Curated_BivData$strain <- factor(Curated_BivData$strain, ordered = FALSE)
+
+#all numbers, numbers
+#all relavent factors unordered
+
+#the mouse averages will be used
+
+mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+                       #IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+)
+
+#vari
+mouse.avs_4MM <- add_sex(mouse.avs_4MM)
+mouse.avs_4MM <- add_strain(mouse.avs_4MM)
+mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
+#unordered
+mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
+mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
+mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
+
+#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
+
+#code if males are high or low?
+
+```
+
+In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .
+
 ## Predictions, Heterochiasmy
@@
 
-Interference, measured in the physical units, has been shown to lack sexual dimorphism (petkov).
+3) Interference, measured in the physical units, has been shown to lack sexual dimorphism (petkov).
 
@@
 Mouse averages for the each metric will be used.
+
+## Predictions, male polymorphism
+
+What predictions do I have for how traits will differ between the low and high Musc mouse strains?
+
+A) **SC lengths** will be longer for high Rec
+
+B.1) **Interfernce/IFD** will be shorter in high Rec
+
+or
+
+B.2) **Interference/IFD** will be longer in high Rec males, sis-co-ten metric will be maximized. (given prelim results from Beth's data)
+
+C) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+Using the same MM framework we would expect the random strain effect would be significant for these predictions.
+
+Use just the male (Musc male data)
+
+$$mouse \ average \ SC  \ length ~=~ rand(strain) + \varepsilon $$
+
+$$mouse \ average \ IFD \ trait  ~=~ rand(strain) + \varepsilon $$
+The IFD traits are:
+
+- IFD_ABS
+
+- IFD_PER
+
+$$mouse \ average \ CO \ position \ trait  ~=~ rand(strain) + \varepsilon $$
+
+The CO position traits are:
+
+- 1CO position
+
+- centromere and telomere distance
+
+- sis-co-ten
 
@@
 
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
-
-Above is the model we are testing evolution of SC lengths.
-
-```{r MM.SC.Length, echo=FALSE}
-
-#use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#check all the strains and male ages, should be the same for MM
-
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mouse.level.table
-#make this in datasetup
-mouse.avs <- ddply(Curated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = format(round(   var(chromosomeLength),3), nsmall=3),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-)
-
-mouse.avs <- add_sex(mouse.avs)
-mouse.avs <- add_strain(mouse.avs)
-mouse.avs <- add_subsp(mouse.avs)
-
-#unorder the factors, but figure our
-mouse.avs$strain <- unclass(mouse.avs$strain)
-mouse.avs$strain <- as.factor(mouse.avs$strain)
-
-mouse.avs$subsp <- unclass(mouse.avs$subsp)
-mouse.avs$subsp <- as.factor(mouse.avs$subsp)
-
-mouse.avs$sex <- unclass(mouse.avs$sex)
-mouse.avs$sex <- as.factor(mouse.avs$sex)
-
-BivData.MixedModel.HQ <- lme(mean_SC ~ subsp * sex, data=mouse.avs, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
-
-#sex is significant! with prediction
-
-```
-
-The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
+
+To estimate random effects, I'm using lmer (not lme where I could do the rand(strain*sex)). So now it's just random(strain).
+
+Above is the model we are testing evolution of SC lengths.  remember for the MM use the table **mouse.avs_4MM**
+
+```{r MM_2.SC.Length, echo=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+#p
+
+```
+
+I used the mixed model with the lme()
+
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r MM_2_SC.length, echo=FALSE}
+
+#as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
+
+#can't make the 
+SC_length.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+anova(SC_length.MM)
+
+thing1  <- summary(SC_length.MM)
+thing2 <- drop1(SC_length.MM, test="Chisq")
+
+rand_SC.Length_MM <- exactRLRT(SC_length.MM)
+rand_SC.Length_MM
+
+#this pvalue for ran is 0.0027 -- the strain effect 
+```
 
@@
 
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for SC length? (as predicted)
+
+Sex is a significant effect for SC length
+The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)
+
+> Is the random Effect of strain an effect on SC length?
+
+The exactRLRT() test indicates that the random strain effect might be significant, p= `r rand_SC.Length_MM$p.value`.
+
+### Male polymorphism Predcition
+
+Preidction, High rec males have longer SC.
+
+```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
+#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
+#MSM_SC vs MOLF_SC
+
+#2. logistic regression for sc length and chrm class?
+#if the prediction / significance
+
+#isolate data
+mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
+
+#build and test MM
+lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+
+lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
+
+anova(lmer_MM_SC_length_male.poly)
+
+rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
+
+```
+
+I can't get the above mixed models to work for testing high vs low rec groups
+
+```{r logreg, echo=FALSE}
+
+mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+#divide musc males into groups
+mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
+                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
+
+log.reg_SC.length <- glm(Rec.group ~ mean_SC, 
+                            data=mouse.avs_4MM_male, family=binomial(link="logit"))
+
+log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
+                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))
+
+#the summary for both these models is the same ... should I be worried?
+
+#mean SC slightly sig-- these are all mean SC--should I try distinguishing by chrm classes?
+```
+
+When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
+When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
+
+> Is the prediction, high rec musc male strains have long SC met?
+
+In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
+I counldn't get the Mixed models working for the male polymorphism predictions...
+
 # IFD
@@
 
-DO I have anything to say....
-
 How can I highlight the result (I need to verify this first) for greater IFD in high rec male mice?
@@
 ```{r show.IFD, echo=FALSE, warning=FALSE}
-
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blues
+
+                       'PWD'= 'red2',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
+                       'CZECH'= 'indianred4',
+
+                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
+
+                    'CAST'='green','HMI'= 'forestgreen', #greens
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
 #first draft of the IFD comparisons
-
-#for now remove the 
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
 Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
@@
 
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+boxplots_first.IFD.plot_ABS
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+boxplots_first.IFD.plot_PER
+
 first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-first.IFD.plot_ABS
 
@@
 #+theme(legend.position="none")
-first.IFD.plot_PER
 
@@
 
-mixed model analysis for IFD (interference)
-
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
-
-$$mouse \ average \ normalized IFD ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure
+mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+
+$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
+
+```{r MM_1.SC.Length, echo=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+
+#remove the SKIVE male line
+mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$mean_IFD.2CO_ABS) | mouse.avs_4MM$mean_IFD.2CO_ABS==""), ]
+
+lme_MM_IFD_ABS <- lme(mean_IFD.2CO_ABS ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_IFD_ABS)$tTable, 6 ) )
+#p
+
+lme_MM_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_IFD_PER)$tTable, 6 ) )
+```
+
+### second set of Mixed Models
+
+These seond set of models are made using the lmer()
+
+$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+$$mouse \ average \ normalized IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure. The tables below are from anova( for the lmer model )
 
@@
 
-#set up data set for the mixed model
-#Curated_BivData
-
-#use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#check all the strains and male ages, should be the same for MM
-
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mouse.level.table
-#make this in datasetup
-mouse.avs.table1 <- ddply(Curated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                   mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                   mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                        cV = cv(chromosomeLength),
-                        var = format(round(   var(chromosomeLength),3), nsmall=3),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-)
-
-mouse.avs.table1 <- add_sex(mouse.avs.table1)
-mouse.avs.table1 <- add_strain(mouse.avs.table1)
-mouse.avs.table1 <- add_subsp(mouse.avs.table1)
-
-#unorder the factors, but figure our
-mouse.avs.table1$strain <- unclass(mouse.avs.table1$strain)
-mouse.avs.table1$strain <- as.factor(mouse.avs.table1$strain)
-
-mouse.avs.table1$subsp <- unclass(mouse.avs.table1$subsp)
-mouse.avs.table1$subsp <- as.factor(mouse.avs.table1$subsp)
-
-mouse.avs.table1$sex <- unclass(mouse.avs.table1$sex)
-mouse.avs.table1$sex <- as.factor(mouse.avs.table1$sex)
-
-mouse.avs.table1 <- mouse.avs.table1[(!is.na(mouse.avs.table1$mean_IFD_ABS)),]
-
-BivData_IFD_ABS.MixedModel.HQ <- lme(mean_IFD_ABS ~ subsp * sex, data=mouse.avs.table1, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(BivData_IFD_ABS.MixedModel.HQ)$tTable, 6 ) )
-#ABD - IFD not sex specific effect
-
-BivData_IFD_PER.MixedModel.HQ <- lme(mean_IFD_PER ~ subsp * sex, data=mouse.avs.table1, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(BivData_IFD_PER.MixedModel.HQ)$tTable, 6 ) )
-#sex is a little significat
-
-```
-
-sex may be a significant effect for the percent (normalized IFD). Is there a way to measure / account for the SC.length effect?
+IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+anova(IFD_ABS.MM)
+
+#thing1  <- summary(IFD_ABS.MM)
+IFD.ABS_chi <- drop1(IFD_ABS.MM, test="Chisq") 
+IFD.ABS_chi
+
+rand_test_IFD.ABS.MM <- exactRLRT(IFD_ABS.MM)
+rand_test_IFD.ABS.MM
+
+IFD_PER.MM <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+anova(IFD_PER.MM)
+
+Chis.sp_IFD.PER.MM <- drop1(IFD_PER.MM, test="Chisq") 
+Chis.sp_IFD.PER.MM
+
+rand_test_IFD.PER.MM <- exactRLRT(IFD_PER.MM)
+rand_test_IFD.PER.MM
+
+```
+
+```{r compare.IFD}
+
+#make
+
+```
+
+### Heterochiasmy Prediction
+
+> Do the IFD measures lack significant sex effect? (as predicted)
+
+There are sex effects for the lme() models for both the ABS and PER IFD measures. There isn't evidence of Random effects from the lmer() models.
+
+> Are there signiticant strain effects?
+
+The random strain effect was tested for raw and normalized IFD, neither are signifcant.
+
+### Male Polymorphism Prediction
+
+> Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
+
+```{r IFD.t.test, echo=FALSE, eval=FALSE}
+
+#pool the whole groups
+
+ABS.t.test <- t.test( mouse.avs_4MM_M.male$mean_IFD.2CO_ABS[mouse.avs_4MM_M.male$Rec.group == 1],
+  mouse.avs_4MM_M.male$mean_IFD.2CO_ABS[mouse.avs_4MM_M.male$Rec.group == 0])
+
+PER.t.test <- t.test( mouse.avs_4MM_M.male$mean_IFD.2CO_PER[mouse.avs_4MM_M.male$Rec.group == 1],
+  mouse.avs_4MM_M.male$mean_IFD.2CO_PER[mouse.avs_4MM_M.male$Rec.group == 0])
+
+#compare strain by strain
+# p = `r ABS.t.test$p.value`  and p = `r PER.t.test$p.value`.
+```
+
+Neither t.test are sig for both the ABS and PER when I test just the Musc strains.
+The above t.tests are breaking the knitr
+
+```{r IFD.poly_log.reg, echo=FALSE}
+
+#test. for the high vs low strains and IFD
+#mouse.avs_4MM_male
+
+mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc", ]
+
+log.reg_IFD.ASB <- glm(Rec.group ~ mean_IFD.2CO_ABS, 
+                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
+#slightly sig
+
+summary(log.reg_IFD.ASB)
+
+log.reg_IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER, 
+                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
+summary(log.reg_IFD.PER)
+
+#not significant
+```
+
+None of the logistic regression models for ABS or PER IFD lengths are sigificant, even when just the Musc strains are used.
+
+Post-hoc comparisons ideas?
 
@@
 
-```{r compare.IFD}
+> Is there evidence for the alternative IFD meansure
+
+INSERT THE SISCOTEN ESTIMATES!!
+
+```{r siscoten.test, echo=FALSE}
 
@@
 
-```{r CO.positions}
+There are a few traits that fall within the CO positions
+
+- 1CO normalized position
+
+- sis.co.ten (sorta interference)
+
+- centromere and telomere distance
+
+```{r CO.positions.figs, echo=FALSE}
 #main figure of density plots,
@@
 
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count <5, ]
+# remove the na's
+MELT_Curated_BivData <- MELT_Curated_BivData[!(is.na(MELT_Curated_BivData$hand.foci.count) | MELT_Curated_BivData$hand.foci.count==""), ]
+
 Musc.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Musc") ), ]
@@
 
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+The mixed model data should only come from 1CO bivalent data.
 
@@
 
-Pos.CO.avs <- ddply(MELT_Curated_BivData, c("mouse", "Foci_Type"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                    mean_PER_pos = as.numeric(format(round(  mean(PER.Foci.Pos), 3 ), nsmall=3) )
-)
-
-Pos.CO.avs <- add_sex(Pos.CO.avs)
-Pos.CO.avs <- add_strain(Pos.CO.avs)
-Pos.CO.avs <- add_subsp(Pos.CO.avs)
-
-#unorder the factors, but figure our
-Pos.CO.avs$strain <- unclass(Pos.CO.avs$strain)
-Pos.CO.avs$strain <- as.factor(Pos.CO.avs$strain)
-
-Pos.CO.avs$subsp <- unclass(Pos.CO.avs$subsp)
-Pos.CO.avs$subsp <- as.factor(Pos.CO.avs$subsp)
-
-Pos.CO.avs$sex <- unclass(Pos.CO.avs$sex)
-Pos.CO.avs$sex <- as.factor(Pos.CO.avs$sex)
-
 #isolate the 1COPos.CO.avs
-Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
-
-Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
-
-BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
-#sex is significant
-```
+#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
+#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
+#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
+
+#can't make the 
+CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+library(lmerTest)
+#anova(CO_position.MM)
+anova(CO_position.MM, verbose = TRUE)
+
+summary(CO_position.MM) #no p values,,
+#these just give t and F stats, no pvalue
+
+#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
+#CO_pos_fix.est
+#this test, the full model is more predictive than an empty model?
+
+#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
+
+#thing2 <- drop1(SC_length.MM, test="Chisq")
+
+CO_pos_rand <- exactRLRT(CO_position.MM)
+CO_pos_rand
+
+```
+
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
+
+> is random strtain effect significant?
+
+The random strain effect seems very significant.
+
+remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
 
@@
 
+### Male Polymorphism Prediction
+
+I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
+
+```{r CO.pos.poly}
+#I
+#CO_pos
+
+```
+
 # old drafts
 
-I went on a kick thinking of the sis-coten metric, I'm not so sure this is a good metric anymore.
+I went on a kick thinking of the sis-coten metric, I'm not so sure this is a good metric anymore.  Do paired t-tests for all the male strains; test is foci position distributions are different.
+
+1CO(F1)  1CO(F1)
+
+2CO(F1) vs 2CO(F1)   2CO(F2) vs 2CO(F2)
 
@@
 
-### 3. Lack of strain specific patterns in females
+# 3. Lack of strain specific patterns in females
 
@@
 
-## PERMUTATIONS Accounting for sample size differences
+# PERMUTATIONS Accounting for sample size differences
 

### 2020-01-22 doc/Main_Report_Results.rmd
first pass at analysis of Mixed models, outlined other predictions, made some notes in other files

@@
 
+todo, add the random effect tests, consider changing all to lmer() (those have better random effects)
+
 $$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
@@
 
+#consider changing to ordered = FALSE, for the DFs used for MM
 DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
@@
 
+```{r trackcurBivData, echo=FALSE}
+#list / tally of current curated Bivdata
+Cur_biv = read.csv("~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", header = TRUE)
+
+colnames(Cur_biv) <- c('mouse', 'blank','0','1')
+
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages[ -c(9:12, 14, 18, 22:24) ]
+
+AP_mouse_table_w.Ages_HQ <- add_sex(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_strain(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_category(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_subsp(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_species(AP_mouse_table_w.Ages_HQ)
+
+#only M.musculus
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$species == "M.musculus",]
+#REMOVE EXTRA STRAINS
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$subsp != "Cast",]
+#remove musc strains
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "AST",]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "CZECH",]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "PERC",]
+#156 mice (from 193)
+
+#some mice are missing 13oct14
+
+#match the MM dataset
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$Ncells > 9,]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$age.weeks < 26,]
+#126 mice
+
+#AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[!(is.na(AP_mouse_table_w.Ages_HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mice missing from curated list
+mice.within.BivData <- AP_mouse_table_w.Ages_HQ[ AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#54 mice in (14 mice have enteries but 0 curated biv)
+
+#(14 mice need at least some curation
+
+cur.data.mice <- Cur_biv[ Cur_biv$mouse %in% AP_mouse_table_w.Ages_HQ$mouse, ]
+
+mice.missing.bivData <- AP_mouse_table_w.Ages_HQ[ ! AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#72 (includes mice not in MM)
+#72 mice that I need to upload/run the algorithm on their images
+
+#mice missing SC pass == '1'
+colnames(Cur_biv) <- c("mouse","blank","SC.fail", "SC.pass")
+NeedsCur <- Cur_biv[Cur_biv$SC.fail == 0,]
+
+NeedsCur <- add_strain(NeedsCur)
+write.table(NeedsCur, "~./MLH1repo/data/needs_cur.csv", sep=",",
+            row.names = TRUE)
+
+#44% of mice are missing curation
+
+#write this table
+write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
+            row.names = TRUE)
+
+mice.missing.bivData <- mice.missing.bivData[!(is.na(mice.missing.bivData$mouse) | mice.missing.bivData$mouse==""), ]
+write.table(mice.missing.bivData, "~./MLH1repo/data/MissingMice_13.1.20.csv", sep=",",
+            row.names = TRUE)
+
+```
+
 ```{r barlett.test, echo=FALSE}

### 2020-01-22 src/BivData_Setup.Rmd
a few small notes from home
first pass at analysis of Mixed models, outlined other predictions, made some notes in other files

@@
 # Data setup and basic cleaning
+
+test work from home ()
+
+add sis_co_ten  ref Lee 2019
+
+format manual data set so SC.length = Chrmlength (matches the bivData)
+
+add Q1 and Q4 to the cell summary stats table
+
+for final dataset, drop levels of strains not used CZECG, AST, TOM, CAST,
 
@@
 #there are 2 lines with 0 hand.foci, but with a foci position
+
+```
+
+```{r clean.data, echo=FALSE, eval=FALSE}
+
+Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
+Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
+Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]
+
+Curated_BivData <- droplevels(Curated_BivData$strain)
 

### 2020-01-22 src/MixedModels.R

@@ -9,30 +9,69 @@
 #load data
 setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
 
-load(file="data/MLH1_data_setup.RData")
+load(file="data/MLH1/MLH1_data_setup_8.29.19.RData")
 source("src/CommonFunc_MLH1repo.R")
 
 #clean data, all strains ahve bothe sexes, make sure variables set as factors and are in a consistant order
 
-MixedModel_CO_data <- AP_mouse_table[AP_mouse_table$strain != "other",]
-MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$strain != "HMI",]
+
+MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
+
+
+MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
+
+MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
+
+no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
+MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
+
+MM <- lme(mean_co ~ subsp.x * sex.x, data= MixedModel_CO_data, random=list(strain.x=pdDiag(~sex.x) ) )
+anova(MM, verbose = TRUE)
+anova(MM, verbose = TRUE, test = TRUE, type = "marginal")
+
+kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
+
+bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
+
[…]

### 2020-01-23 src/BivData_Setup.Rmd
Added sis-coten to the BivData set up

@@
 ```{r split.up.DF, echo=FALSE}
-
 #SC PASS
@@
 
-Curated_BivData <- droplevels(Curated_BivData$strain)
+Curated_BivData$strain <- droplevels(Curated_BivData$strain)
 
@@
 #add all the extra columns, blah
-
 #percent Foci position
@@
 #add telomere.dist
-#str(Curated_BivData)
+Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
+                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
+                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
+                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
+
+Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
+
+#sis.co.ten
+Curated_BivData <- add_SisCoTen(Curated_BivData)
+
 ```
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_1.21.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_1.22.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-01-23 doc/Bivalent_Analysis.Rmd
made some plots for sis-co-ten

@@
 #load Clean BivData
-load(file="data/CleanBivData_1.21.20.RData")
+load(file="data/CleanBivData_1.22.20.RData")
 
@@
 think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+
+sis.co.ten_with.plot <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
+  geom_jitter()+facet_wrap(sex~strain)
+sis.co.ten_with.plot
+
+sis.co.ten_with.plot_2 <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
+  geom_jitter()+facet_wrap(sex~strain)
+sis.co.ten_with.plot_2
+
+```
+
+the distribution of SC lengths and sis-coten seems very different across sexes
 

### 2020-01-23 doc/Bivalent_Analysis.Rmd
I think I got the code for isolating long biv data set debugged. v tedious

@@
 #then this DF has to be summarized into categories
-chrm.class.category <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
+chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
                              mean.percent.type = mean(value)
@@
 
-chrm.class.barplot <- ggplot(data = chrm.class.category, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
+chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
 chrm.class.barplot
@@
   facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
-show.the.cells_females_musc
 
@@
   facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_Musc_f
 
@@
   facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
-show.the.cells_male_musc
 
@@
   facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_Musc_m
-
-#the min is pretty stable
 
@@
 
+#adding the extra cols, messed up this part
+
 #how many biv's are from cells with >15 bivalents
@@
 
-for(  i in 1:length( long_BivData_set99$fileName)  ){
-  long_BivData_set99$cell.Q4[i] <- Biv_Data_mean_table[ which(Biv_Data_mean_table$fileName == long_BivData_set99$fileName[i]), ]
+long_BivData_set99 <- long_BivData_set99[1:50,]
+#3 unique fileNames, in abbridged DF
+
+ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
+                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
+
+newfile = strsplit(i, "_")[[1]][1:7]
+
+Biv_Data_mean_table[grepl(newfile, Biv_Data_mean_table$fileName )]
+
+#i is obj.id, this returns a table of all falses -- because it's look for the whole string
+
+o<- strsplit(i, "_")
+
+Biv_Data_mean_table$fileName[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
+#above returns the BivData, filename which the grepl matches
+
+Biv_Data_mean_table$Qar.4[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
+#this returns the Q4
+
+#this will put the cell's Q4
+#I think I need a counter
+counter = 1
+
+length(duplicated(long_BivData_set99$Obj.ID))#there are 2605 duplicated Obj.ID
+anyDuplicated(long_BivData_set99$Obj.ID)
+dup.curated <- long_BivData_set99[duplicated(long_BivData_set99$Obj.ID),]#65 duplicated obj.ID
+long_BivData_set99 <- long_BivData_set99 %>% distinct()
+
+#REDO
+#61 unique biv.id
+
+#this must be charecter for str split
+
+long_BivData_set99$Obj.ID <- as.character(long_BivData_set99$Obj.ID )
+long_BivData_set99.mini <- long_BivData_set99[905:2605,]
+
+long_BivData_set99.mini$Obj.ID <- paste(long_BivData_set99.mini$fileName, long_BivData_set99.mini$boxNumber, sep = "_")
+
+anyDuplicated(long_BivData_set99.mini$Obj.ID)#1093
+dup.mini.curated <- long_BivData_set99.mini[duplicated(long_BivData_set99.mini$Obj.ID),]
+
+#do a pipe remove
+long_BivData_set99.mini <- long_BivData_set99.mini[ ! long_BivData_set99.mini$fileName %in% dup.mini.curated$fileName, ]
+#think this removes the duplicated rows
+
+#IT COULD BE THAT THERE are filenames in the Biv_Data_mean_table which don't have the
+
+#make sure the biv.mini only have filenames which are also in the mean teabl
+long_BivData_set99.mini <- long_BivData_set99.mini[ ! long_BivData_set99.mini$fileName %in% dup.mini.curated$fileName, ]
+
+long.data.set.cells <- al.15_bivs
+
+long_BivData_set99.mini <- long_BivData_set99.mini[long_BivData_set99.mini$fileName %in% long.data.set.cells$fileName, ]
+
+long_BivData_set99.mini <- long_BivData_set99.mini[long_BivData_set99.mini$fileName %in% Biv_Data_mean_table$fileName, ]
+#removed fileNames which don't appear in Biv_Data_mean_table (which has the Q4)
+
+#long_BivData_set99.mini <- long_BivData_set99[905:2605,]
+
+for(  i in 1:length( unique(long_BivData_set99.mini$Obj.ID) ) ){
+  newfile.str = strsplit( long_BivData_set99.mini$Obj.ID[i], "_")[[1]][1:7]
+  new.egg <- paste0(newfile.str, collapse = "_")
+
+  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+  #the above can produe mulitple Q4
+
+  long_BivData_set99.mini$cell.Q4.99[i] <- the.Q4
+  #these are just using the last index,
+
+  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+
+  long_BivData_set99.mini$Q4.FN[i] <- the.FN
+
+  print(c(i, the.FN, the.Q4))
+
 }
-
-for(  k in 1:length( long_BivData_set99$fileName)  ){
-long_BivData_set99$above.Q4[k] <- ifelse( (long_BivData_set99$cell.Q4 < long_BivData_set99$chromosomeLength), "yes", "no")
-}
-#This isn't working either
-
-#for(i in 1:length( long_BivData_set$fileName)  ){ 
-  # print(i)
-#  current_obj_fileName <- long_BivData_set$fileName[i]
-#  current_obj_SC_length <- long_BivData_set$chromosomeLength[i]  
-#  current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == current_obj_fileName) ]
-# current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == long_BivData_set$fileName[i]) ]
-#  long_BivData_set$above.Q4 <-ifelse( (current.Q4 < current_obj_SC_length), 1, 0)
-#  print( paste( c(current.Q4, current_obj_fileName, current_obj_SC_length), sep = "_" ) )
+#some kind of error at 903 -- the loop should continue past it
+
+long_BivData_set99.mini$FN2 <- long_BivData_set99.mini$fileName
+#mini, looks good so far
+
+long_BivData_set99.mini$long.Biv <- ifelse( (long_BivData_set99.mini$chromosomeLength >= long_BivData_set99.mini$cell.Q4), "yes", "no" )
+
+long_BivData_set99.mini$wtf <- long_BivData_set99.mini$chromosomeLength - long_BivData_set99.mini$cell.Q4
+
+real.long.bivData <- long_BivData_set99.mini[long_BivData_set99.mini$long.Biv == "yes",]
+table(real.long.bivData$fileName)
+
+#!!! omg it finally worked!~!!
+#OMFG
+
+#OG try
+#for(  i in unique(long_BivData_set99$Obj.ID) ){
+#  newfile = strsplit(i, "_")[[1]][1:7]
+#  the.Q4 <- Biv_Data_mean_table$Qar.4[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
+#  long_BivData_set99$cell.Q4.99[counter] <- the.Q4
+#  #these are just using the last index, 
+#  the.FN <- Biv_Data_mean_table$fileName[which(grepl(newfile, Biv_Data_mean_table$fileName ) )]
+#  long_BivData_set99$Q4.FN[counter] <- the.FN
+#  print(c(i, the.FN, the.Q4))
+#  counter = counter+1
 #}
-#this loop isn't working
-
-```
+#not sure what the final error is in regards to
+#the file name matches are wronge
+
+```
+
+Ugggg, still something wrong,
+9apr18_1feb18_LEW_f1_sp1_9_rev
+
+ 0%    25%    50%    75%   100% 
+ 77.00 114.50 136.00 152.75 212.00
+
+Started compiling the long_biv data set, some cells have more than I expect:
+
+13apr15_1apr15_PWD_f2_sp1_4_rev, 5aug17_14jul17_MSM_f1_sp1_12_rev, 5aug17_14jul17_MSM_f1_sp1_26_rev, 30jun17_4jan17_LEW_f3_sp1_1_rev, 8feb17_4jan17_LEW_f1_sp1_5_rev, 8feb17_4jan17_LEW_f1_sp1_3_rev, 9apr18_1feb18_LEW_f1_sp1_9_rev, 8jun15_23apr15_PWD_f2_sp1_17_rev
+
+They have like ~10, it might have to do with not haveing = in the >=
+
+Qar.4 = quantile(chromosomeLength)[[4]],
+quantile(Curated_BivData$chromosomeLength[Curated_BivData$fileName == "8feb17_4jan17_LEW_f1_sp1_3_rev"]) #148.00
+Curated_BivData$chromosomeLength[Curated_BivData$fileName == "8feb17_4jan17_LEW_f1_sp1_3_rev"]
+
+9apr18_1feb18_LEW_f1_sp1_9_rev
+
+the correct Q4 is 148, but the value in the Q4 col is 116, and the filName listed is completely wrong
 
@@
 
-2) normalized CO positions (sedell and Kirkpatrick).
+2.A) normalized CO positions (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
 
@@
 C) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+D) Sis-co-ten predictions fall into
 
@@
 
-Sex is a significant effect for SC length
+- Sex is a significant effect for SC length
 The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)
+
+- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
 
@@
 
-There are sex effects for the lme() models for both the ABS and PER IFD measures. There isn't evidence of Random effects from the lmer() models.
+There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
 
@@
 
-> Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
+> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
 
@@
 
-> Is there evidence for the alternative IFD meansure
-
-INSERT THE SISCOTEN ESTIMATES!!
-
-```{r siscoten.test, echo=FALSE}
-
-```
+> Prediction B, Is there evidence for the alternative IFD meansure
 

### 2020-01-23 doc/Bivalent_Analysis.Rmd
updates can knit!

--- 2020-01-23 doc/Bivalent_Analysis.Rmd
@@
 
-```{r long.chrm.dataset, echo=FALSE, eval=FALSE}
-
-#how many biv's are from cells with >15 bivalents
-
-al.15_bivs <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]#150 cells have al 15
-
-long_BivData_set <- Curated_BivData[ Curated_BivData$fileName %in% al.15_bivs$fileName, ] #across those 150 cells are 2654 biv measures
-
-#table(potential.data.set$category)
-long_BivData_set$fileName <- as.character(long_BivData_set$fileName)
-Biv_Data_mean_table$fileName <- as.character(Biv_Data_mean_table$fileName)
-
-#pull out the biv's which exceed the cell's Q4
-#for each line of long_BivData_set
-# check if it's SC length is greater than the Q4 for the given cell
-
-long_BivData_set99 <- long_BivData_set
-
-for(  i in 1:length( long_BivData_set99$fileName)  ){
-  long_BivData_set99$cell.Q4[i] <- Biv_Data_mean_table[ which(Biv_Data_mean_table$fileName == long_BivData_set99$fileName[i]), ]
+Biv_Data_mean_table has the Q4 values
+
+```{r long.chrm.dataset, echo=FALSE}
+
+#1. how many biv's are from cells with >15 bivalents
+
+over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
+BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]
+
+#check for blank obj.id
+BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")
+
+#check for dup
+length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
+anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
+dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
+
+#remove duplicate obj.ids
+BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
+BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]
+
+#make sure obj.id and fileName are charcter for str stuff
+BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
+BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)
+
+#make sure 
+missing_from_cell_table <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% Biv_Data_mean_table$fileName, ]#there are 75, objs 3 cells
+#these are bivalents which don't appear the in cell level table, which can break the loop
+BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
+
+#final cleaned BivData table, 2428
+
+for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){
+
+  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
+  new.egg <- paste0(newfile.str, collapse = "_")
+
+  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+  #the above can produe mulitple Q4
+
+  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
+  #these are just using the last index,
+
+  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+
+  BivData_from_15biv_cells$Q4.FN[i] <- the.FN
+   # print(c(i, the.FN, the.Q4))
+
 }
 
-for(  k in 1:length( long_BivData_set99$fileName)  ){
-long_BivData_set99$above.Q4[k] <- ifelse( (long_BivData_set99$cell.Q4 < long_BivData_set99$chromosomeLength), "yes", "no")
-}
-#This isn't working either
-
-#for(i in 1:length( long_BivData_set$fileName)  ){ 
-  # print(i)
-#  current_obj_fileName <- long_BivData_set$fileName[i]
-#  current_obj_SC_length <- long_BivData_set$chromosomeLength[i]  
-#  current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == current_obj_fileName) ]
-# current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == long_BivData_set$fileName[i]) ]
-#  long_BivData_set$above.Q4 <-ifelse( (current.Q4 < current_obj_SC_length), 1, 0)
-#  print( paste( c(current.Q4, current_obj_fileName, current_obj_SC_length), sep = "_" ) )
-#}
-#this loop isn't working
-
-```
-
-The dataset is unbalanced in some ways -- integrating manual measures might fix. MSM under rep, since most cell had high overlap.
+BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= BivData_from_15biv_cells$cell.Q4), "yes", "no" )
+
+real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]
+table(real.long.bivData$fileName)
+
+#!!! omg it finally worked!~!!#OMFG
+
+```
+
+Well fixing those bigs took forever
+
+```{r show.long.SC, echo=FALSE}
+
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))
+female_long_biv_plot
+
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))
+male_long_biv_plot
+
+```
+
+These plots show the SC lengths for bivalents I'm calling the 'long SC data set'.
+They are supposed to be the longest 4-5 SC from cells where I could get good measures.
+These longer bivalents are useful because their patterns shouldn't be affect by chromsome size effect (which effects, CO position).
+Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
 

### 2020-01-23 doc/Bivalent_Analysis.Rmd
added a fig for sis-co-ten example

@@
 #check for dup
-length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
-anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
+#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
+#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
 dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
@@
 real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]
-table(real.long.bivData$fileName)
+#table(real.long.bivData$fileName)
 
@@
 
-Well fixing those bigs took forever
+Well fixing those bigs took forever.
+Use the DF, **real.long.bivData** for comparing the longest bivalents.
+still v unblancd across sexes and strains ~ merge with the manual whole cell biv measures later...
+
+MSM male only has data from 1 cell, and PWD 3, uh oh -- may need to merge sooner.
+
+First proof of principl will be comparisons within category (long vs all)
+
+```{r mouse.av.long.bivData, echo=FALSE}
+#real.long.bivData
+
+#calq the mouse averages for
+
+Long_biv_mouse.avs_4MM <- ddply(real.long.bivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+                       #IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+)
+
+Long_biv_mouse.avs_4MM <- add_sex(Long_biv_mouse.avs_4MM)
+Long_biv_mouse.avs_4MM <- add_strain(Long_biv_mouse.avs_4MM)
+Long_biv_mouse.avs_4MM <- add_subsp(Long_biv_mouse.avs_4MM)
+
+```
+
+in code chunk above I ran the mouse averages for the longest bivalets (680 bivalents, from 54 mice. full dataset is 10202 bivalents from 86 mice.
 
@@
 Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
+
+# Adjusting for XX
+
+The female mouse averages should have adjustments for the XX.
+working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
+Subtract this amount from the female mouse averages ...
+This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
+
+- Of all female single bivalents 5% are XXs (1 of 20).
+
+- The XX is large, thought to be within the top 25% longest bivalents of the cell (3rd largest by Mb).
+
+- What is the average % of XX for whole cell SC (sum(all bivalents)). Lets guess 12% of a cell's total SC area is XX.
+
+- do the algebra...
 
@@
 
+remeber the long biv data set in **real.long.bivData**
+
 ```{r MM_2.SC.Length, echo=FALSE}
@@
 
-```
-
-I used the mixed model with the lme()
+#model for long Bivs
+long_bivData_SC.Length_MM <- lme(mean_SC ~ subsp * sex, data= Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(long_bivData_SC.Length_MM)$tTable, 6 ) )
+
+```
+
+I used the mixed model with the lme() for the full Curated BivData and long bivdata
 
@@
 ```{r MM_2_SC.length, echo=FALSE}
+#Long_biv_mouse.avs_4MM
 
@@
 #can't make the 
-SC_length.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-anova(SC_length.MM)
-
-thing1  <- summary(SC_length.MM)
-thing2 <- drop1(SC_length.MM, test="Chisq")
-
-rand_SC.Length_MM <- exactRLRT(SC_length.MM)
+SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+anova(SC_length.lmer.MM)
+
+thing1  <- summary(SC_length.lmer.MM)
+thing2 <- drop1(SC_length.lmer.MM, test="Chisq")
+
+rand_SC.Length_MM <- exactRLRT(SC_length.lmer.MM)
 rand_SC.Length_MM
@@
 #this pvalue for ran is 0.0027 -- the strain effect
+
+##long bivalents
+long.bivs_SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long_biv_mouse.avs_4MM, REML = TRUE)
+
+anova(long.bivs_SC_length.lmer.MM) 
+#sex is larges explantion of variance
+
+thing1  <- summary(long.bivs_SC_length.lmer.MM)
+thing2 <- drop1(long.bivs_SC_length.lmer.MM, test="Chisq")
+
+rand_SC.Length_MM <- exactRLRT(long.bivs_SC_length.lmer.MM)
+rand_SC.Length_MM
+
+#strain no sig, but underpowered
+
 ```
@@
 
+- The Long Biv Data set largely agrees with the full curated dataset
+
+- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
+
 > Is the random Effect of strain an effect on SC length?
@@
 
-#the summary for both these models is the same ... should I be worried?
+summary(log.reg_SC.length_Musc)
 
 #mean SC slightly sig-- these are all mean SC--should I try distinguishing by chrm classes?
+```
+
+```{r long.sc_log.res_meanSC}
+
+long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
+long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
+
+#divide musc males into groups
+long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", long.SC.mouse.avs_4MM_male$strain), 1, 
+                                    ifelse(grepl("MSM", long.SC.mouse.avs_4MM_male$strain), 1, 0))
+
+long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
+                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
+
+summary(long.biv.log.reg)#NS effect (likelt underpowered)
 ```
@@
 # IFD
+
+ToDo fix the function to calculat IFD2, IFD3
 
@@
 
+```{r blah.again}
+#long SC Long_biv_mouse.avs_4MM
+#real.long.bivData
+
+#Only use 2CO bivalents
+real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
+real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
+
+long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+
+long.biv_IFD_look_ABS
+
+long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+
+long.biv_IFD_look_PER
+
+```
+
 Above is the first plots of IFDs. Each point is a IFD observation. colors are by mouse. 
@@
 
-#remove the SKIVE male line
+#remove the SKIVE male line which is na
 mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$mean_IFD.2CO_ABS) | mouse.avs_4MM$mean_IFD.2CO_ABS==""), ]
@@
 
-#make
+#make t.tests for pooled high and low
 
@@
 
-summary(log.reg_IFD.ASB)
+ #summary(log.reg_IFD.ASB)
 
@@
                             data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
-summary(log.reg_IFD.PER)
+ #summary(log.reg_IFD.PER)
 
@@
 - centromere and telomere distance
+
+for the long chrm, pull the object ids out of the melt table.
+
+### Normalized 1CO positions
 
@@
 #clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count <5, ]
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
 # remove the na's
-MELT_Curated_BivData <- MELT_Curated_BivData[!(is.na(MELT_Curated_BivData$hand.foci.count) | MELT_Curated_BivData$hand.foci.count==""), ]
-
-Musc.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_female <- ggplot(Musc.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc female \nNormalized Foci Positions")
-CO.Pos.density_Musc_female
-
-#Dom female
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
 Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
 CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-CO.Pos.density_Dom_female
-
 #Musc male
@@
 CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-CO.Pos.density_Musc_male
-
 #Dom male
@@
 CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-CO.Pos.density_Dom_male
-
-```
-
-Split up the plot data by subsp - sex. 
-Try adding in the numbers of observations?
+
+```
+
+Above plot focuses on the 1CO bivalent normalized positions -- since CO interference controls the general position of COs when there are multiple COs.
+This plot shows the sexual dimorphism in the density plots.
+
+consider adding annotate_text for the number of observations in each category.
 think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
 
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-
-sis.co.ten_with.plot <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
-  geom_jitter()+facet_wrap(sex~strain)
-sis.co.ten_with.plot
-
-sis.co.ten_with.plot_2 <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
-  geom_jitter()+facet_wrap(sex~strain)
-sis.co.ten_with.plot_2
-
-```
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+#real.long.bivData
+#MELT_Curated_BivData
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+
+```
+
+Tired runnin plots for the long biv data set, I think there are too few observations.
+
+add boxplots for
+
+```{r CO1_foci_pos_boxplot, echo=FALSE}
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+geom_jitter(alpha=.1)
+CO.Pos.boxplots_1CO
+
+```
+
+These bosplots show that females have a much more medial poition of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+##### Mixed models norm 1CO positions
 
@@
 
+### Siscoten
+
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+Alternating IFDs.
+Assuming chromatid interference and no sister seperation, this should reflect the amount of sister cohesion that is pulled across the opposite pole.
+
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The authors from this paper call it by something else.
+looking at this siscoten metric is a way to model different cohesin outcomes as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated from SC area -- which we are using as a proxy to the amount of cohesion at metaphase.
+
+![Logic of Sis-co-ten metric, from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+
+sis.co.ten_with.plot <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
+  geom_jitter()+facet_wrap(sex~strain)
+sis.co.ten_with.plot
+
+sis.co.ten_with.plot_2 <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
+  geom_jitter()+facet_wrap(sex~strain)
+sis.co.ten_with.plot_2
+
+```
+
+Males have much clearer seperation of siscoten across chrm classes. This is emphasized when SC length is also plotted.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistant...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r cent.dist, echo=FALSE}
+
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
+
 ### Heterochiasmy Prediction
@@
 ```{r CO.pos.poly}
-#I
+
 #CO_pos
@@
 ```
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-Alternating IFDs.
-Assuming chromatid interference and no sister seperation, this should reflect the amount of sister cohesion that is pulled across the opposite pole.
-
-need more MSM male curated measures!

### 2020-01-24 doc/Bivalent_Analysis.Rmd
git status

@@
 library(reshape2)
+#library(lmerTest) #loading this package will mess up RLRsim's exactRTRL()
+#for got the function i wanted to use
 
@@
 
-This plot illustrates how well the algorithm seperated indiviual bivalents per cell image. 
-Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally segmented.
-
-```{r tech.ect, echo=FALSE, eval=TRUE}
+Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script.
+These observations are from the automated image analysis algorithm and have been currated (threw out incorrect algorithm output).
+
+The breakdown of curated bivalents by category
+
+`r data.frame(table(droplevels(Curated_BivData$category) ) )`
+
+```{r tech.ect, echo=FALSE, eval=FALSE}
+#this fig not needed
 
@@
 #This isn't the same as curated/passing bivalents           
-
-```
-
-consider making a table with predictions for bivalent patterns which are related to the mechanism
+#This plot illustrates how well the algorithm seperated indiviual bivalents per cell image. 
+#Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally #segmented.
+
+```
 
@@
 
-Should I be thinking of making a MM or lm for predicting CO number (within a category)
-
-but there is so little variance in the CO number
-
-co ~ sc.length
-
-I ran this model before,  I remember that these models are better predictors (more significant for males than females)
-
-maybe -- female nuanced pattern -- falls along the lines of
-
-simple predictions not met as strongly as in male patterns...
-
-this reasoning might be complicated by the fact that the range in CO averages is lower in females -- so there is less power
+The figure below shows the average percentages of chromosome classes per category.
 
@@
 
-the above plot displays the average proportion of each type of chromosome class across mice. The results mostly align with the hand measured results.
-
-Look into ordering the mouse strains to highlight the MSM and PWD differences, renameing the chrm type labels, spacing out the x axis labels or removing some of the other mice.
-
-# Chromosome Size effect
-
-In an attempt to deal with the chromosome effect size that is know to effect total rate and CO position, we try to make a dataset of the longest 25% SC from a cell.
-
-The majority of cells from the BivData set do not have the complete completement of bivalents, (ie they are missing up to 5) -- which makes it hard to use a absolute length threshold since some cells have shorter or longer bivalents compared to the while.
-So an absolute measure might catch the lower 25% of some cells.
-
-I'd like to figure out how to best seperate out the true 4th quartille bivalents (longest 25%), even with this missing data. double check that 4th quartille is top 5/4
-
-Could I use the total SC? (I would be warry: can't ensure that they apply the same skeletonize algorithm/parameters and the XY is also included in that count and will mess everything up.)
-
-outline logic for filtering / choosing BivData set of almost whole cells
-
-- 16 to 20 bivalents in the cells
-
-```{r chrm.size.effect, echo=FALSE}
+These results align with the previous figure I made using percentages from hand measured results (dozens of observations, instead of hundreds).
+These results fit with the rapid male evolution of gwRR in MSM and PWD.
+
+- Look into ordering the mouse strains to highlight the MSM and PWD differences, renameing the chrm type labels, spacing out the x axis labels or removing some of the other mice.
+
+# Chromosome Size Effect
+
+In an attempt to test for and metigate noise based on chromosome identity/physical size, I tried to isolate bivalents which are in the top quartille for SC length from their cells.
+
+Since the majority of cells from the Automated BivData set do not have the complete completement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
+But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
+
+Below are examples of plots which show the SC length distributions across cells.
+The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
+There are plot pairs for each category, but they are excluded for space.
+
+Each point is a bivalent ploted by cell on the x axis.
+X's are the 4th quartile, big point is the mean and smaller black point is the median.
+I'm using these to compare the patterns of these statistics in the automated dataset which is missing some bivalent data.
+
+```{r SC.distb_manual_Auto, echo=FALSE}
 
@@
   facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
-show.the.cells_male_dom
 
@@
   facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_Dom_m
 
@@
 
-I'm purposing to take all the automated biv measures above the cell's 4th quartille as a dataset for 'longestest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromsome size effects.
+I'm purposing to take all the automated biv measures above the cell's 4th quartille as a dataset for 'longest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromsome size effects.
 
 But this dataset might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females). But it's easier to just run another set of analysis.
-
-Biv_Data_mean_table has the Q4 values
 
@@
 #table(real.long.bivData$fileName)
-
 #!!! omg it finally worked!~!!#OMFG
@@
 
-Well fixing those bigs took forever.
-Use the DF, **real.long.bivData** for comparing the longest bivalents.
-still v unblancd across sexes and strains ~ merge with the manual whole cell biv measures later...
-
-MSM male only has data from 1 cell, and PWD 3, uh oh -- may need to merge sooner.
-
-First proof of principl will be comparisons within category (long vs all)
+The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. full dataset is 
+`r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category `r data.frame(table(droplevels(real.long.bivData$category)))`.
+
+Still very unbalanced across sexes and strains ~ merge with the manual whole cell biv measures later...
+Musc males are very under-represented. MSM male only has data from 1 cell.
+
+- Try to merge this DF with the whole.cell manual measures.
+
+- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.
 
@@
 
-in code chunk above I ran the mouse averages for the longest bivalets (680 bivalents, from 54 mice. full dataset is 10202 bivalents from 86 mice.
+in code chunk above I ran the mouse averages for the longest bivalets. (680 bivalents, from 54 mice. 
+10202 bivalents from 86 mice.
 
@@
 
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")
 female_long_biv_plot
 
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")
 male_long_biv_plot
@@
 
-These plots show the SC lengths for bivalents I'm calling the 'long SC data set'.
+These plots show the SC lengths for the 'long SC data set'.
 They are supposed to be the longest 4-5 SC from cells where I could get good measures.
@@
 
-- Of all female single bivalents 5% are XXs (1 of 20).
-
-- The XX is large, thought to be within the top 25% longest bivalents of the cell (3rd largest by Mb).
-
-- What is the average % of XX for whole cell SC (sum(all bivalents)). Lets guess 12% of a cell's total SC area is XX.
-
-- do the algebra...
+- Of all female single bivalents observations, 5% are XXs (1 of 20).
+
+- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
+
+- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell dataset. Lets guess 12% of a cell's total SC area is XX.
+
+-
 
@@
 
-I will use two different mixed models using lme() and lmer(), because the different models have limitations for reporting p values ect.
-
-list out pros and cons for using mouse averages of bivalent traits instead of using the single bivalent traits them selves.
+I will use two different mixed models using lme() and lmer(), because the different models have limitations for reporting p values for fixed and random effects.
+
+Pros and con list for using mouse averages vs single bivalent traits.
 
@@
 - Limited work to add to the dataset if I can justify that the collection of single bivalents are random and equivilent across mice.
+
+The mouse averages for all the metrics is what is used in the mixed models.
 
@@
                        #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE)
 )
@@
 
-1) SC length (cite Lynn)
-
-2.A) normalized CO positions (sedell and Kirkpatrick).
+1) SC length will be sexually dimorphic (sex effect will be signicant)(cite Lynn)
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
 
@@
 
-3) Interference, measured in the physical units, has been shown to lack sexual dimorphism (petkov).
-
-These 3 hypotheses will be tested with the mixed model framework, specifically focusing on the effect of sex.
-
-Mouse averages for the each metric will be used.
-
-## Predictions, male polymorphism
-
-What predictions do I have for how traits will differ between the low and high Musc mouse strains?
-
-A) **SC lengths** will be longer for high Rec
-
-B.1) **Interfernce/IFD** will be shorter in high Rec
-
-or
-
-B.2) **Interference/IFD** will be longer in high Rec males, sis-co-ten metric will be maximized. (given prelim results from Beth's data)
-
-C) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-D) Sis-co-ten predictions fall into
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+3) Interference / IFD will not be sexually dimorphic. Previous measures has been shown to lack sexual dimorphism (cite petkov).
+
+## Predictions, Male Polymorphism (High vs Low Rec strains)
+
+A) **SC lengths** will be longer for high Rec strains.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains.
+
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
+
+C.3) telomere and centromere distances will be shorter in low rec strains
 
@@
 
-```{r MM_2_SC.length, echo=FALSE}
+```{r MM_2_SC.length, echo=FALSE, eval=FALSE}
 #Long_biv_mouse.avs_4MM
-
 #as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
 
-#can't make the 
+#did model isn't working
 SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
 
-anova(SC_length.lmer.MM)
-
-thing1  <- summary(SC_length.lmer.MM)
-thing2 <- drop1(SC_length.lmer.MM, test="Chisq")
+#anova(SC_length.lmer.MM)
+#thing1  <- summary(SC_length.lmer.MM)
+#thing2 <- drop1(SC_length.lmer.MM, test="Chisq")
 
@@
 ##long bivalents
-long.bivs_SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long_biv_mouse.avs_4MM, REML = TRUE)
-
-anova(long.bivs_SC_length.lmer.MM) 
+#long.bivs_SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long_biv_mouse.avs_4MM, REML = TRUE)
+
+#anova(long.bivs_SC_length.lmer.MM) 
 #sex is larges explantion of variance
-
-thing1  <- summary(long.bivs_SC_length.lmer.MM)
-thing2 <- drop1(long.bivs_SC_length.lmer.MM, test="Chisq")
-
-rand_SC.Length_MM <- exactRLRT(long.bivs_SC_length.lmer.MM)
-rand_SC.Length_MM
-
+#thing1  <- summary(long.bivs_SC_length.lmer.MM)
+#thing2 <- drop1(long.bivs_SC_length.lmer.MM, test="Chisq")
+
+#rand_SC.Length_MM <- exactRLRT(long.bivs_SC_length.lmer.MM)
+#rand_SC.Length_MM
 #strain no sig, but underpowered
-
-```
-
-```{r justify.mouse.level, echo=FALSE}
+```
+
+```{r justify.mouse.level, echo=FALSE, eval=FALSE}
 
@@
 #
-
 #justify mouse averages
@@
 
-The exactRLRT() test indicates that the random strain effect might be significant, p= `r rand_SC.Length_MM$p.value`.
+The exactRLRT() test indicates that the random strain effect might be significant, p= xx
 
@@
 ```{r MM.IFD, echo=FALSE}
-
+#ugg what changed, FFF kntir is broken!!
+
+library(RLRsim)
 IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
@@
 
-library(lmerTest)
+#library(lmerTest)
 #anova(CO_position.MM)
@@
 #as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
-
 #thing2 <- drop1(SC_length.MM, test="Chisq")
@@
 CO_pos_rand
-
 ```
@@
 The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-Alternating IFDs.
-Assuming chromatid interference and no sister seperation, this should reflect the amount of sister cohesion that is pulled across the opposite pole.
 
 The logic of how the sis-co-ten metric is outlined in the figure below.
-The authors from this paper call it by something else.
-looking at this siscoten metric is a way to model different cohesin outcomes as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated from SC area -- which we are using as a proxy to the amount of cohesion at metaphase.
-
-![Logic of Sis-co-ten metric, from (Lee, J.
+The goal is to use this metric to model different tension ative cohesin amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
 (2019).
@@
 Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-
-sis.co.ten_with.plot <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
   geom_jitter()+facet_wrap(sex~strain)
-sis.co.ten_with.plot
-
-sis.co.ten_with.plot_2 <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+
+jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
   geom_jitter()+facet_wrap(sex~strain)
-sis.co.ten_with.plot_2
-
-```
-
-Males have much clearer seperation of siscoten across chrm classes. This is emphasized when SC length is also plotted.
+jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+jitter_siscoten_WSBm
+```
+
+Males have much clearer seperation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males (which might)
+
+To formally test the differences in sis-co-ten I plan to write a subsampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+```{r sis.co.ten.test_HetC, echo=FALSE}
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model od predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE}
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))
+
+#test high vs low Musc strains
+mouse.avs_4MM
+
+```
+
+> is the male polymorphism
 

### 2020-01-24 src/BivData_Setup.Rmd
git status

@@
 
-add sis_co_ten  ref Lee 2019
+add mean.telomere, mean cent.dist and mean sis.coten to the mouse level table
+
+update IFD functions, to add IFD2,3
 
@@
                        sd.SC_length   = round(sd(chromosomeLength), 3),
-                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3)
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3),
+
+                       #sis.co.ten
+                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
 )

### 2020-01-24 results/XX_adjustment.R

@@ -101,15 +101,14 @@ whole.cell.data <- add_category(whole.cell.data)
 female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#384
 
 ### isolate rank 1:5 chrms
-
 #plot cos / per chrm
-
 
 #THE ranks are backwards
 pp <- ggplot(female.whole.cell[female.whole.cell$rank > 15,], aes(x=rank, y=SC.length, color=as.factor(hand.foci.count) ))+
   geom_jitter()+  facet_wrap(~strain)
+#theme(legend.position = "none"  )+ggtitle("my first plot")
 
-theme(legend.position = "none"  )+ggtitle("my first plot")
+#I think I need more whole cell measures
 
 # need more data on the ratios of 0:1:2:3 COs per biv across these ranks
 # 5 cells per mouse?
@@ -147,6 +146,10 @@ curated_cell_summary <- ddply(Curated_BivData[Curated_BivData$sex == "female",],
 )
 #there are some images with >20 Bivalents, they should be edited
 #3 female cell images with 20 bivalents, 11 cells with 19
+
+#write out curated cell summary
+write.table(curated_cell_summary, "~./MLH1repo/curate_female_cells_table.csv", sep=",", row.names = FALSE)
+
 
 
 #table summarize by rank

### 2020-01-24 doc/Bivalent_Analysis.Rmd
some work from tonight

@@
 
+```{r misc.setup, echo=FALSE}
+
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blues
+
+                       'PWD'= 'red2',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
+                       'CZECH'= 'indianred4',
+
+                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
+
+                    'CAST'='green','HMI'= 'forestgreen', #greens
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
+```
+
 # Setup and Discription of Dataset
@@
 #Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally #segmented.
-
 ```
@@
 
+#colors_of_strains
+#scale_color_manual(values=colors_of_strains)
+
 show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
@@
 
+  scale_color_manual(values=colors_of_strains)+
   facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
+
 show.the.cells_females_dom
@@
    geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
+  scale_color_manual(values=colors_of_strains)+
   facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
@@
 
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
 female_long_biv_plot
 
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
 male_long_biv_plot
@@
 #code if males are high or low?
-
 ```
@@
 
-first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")
+first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)
 first.SC.plot
@@
 
-other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)) ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")
-
-SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))
+other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
 
@@
 
-ToDo fix the function to calculat IFD2, IFD3
+3) Interference / IFD will not be sexually dimorphic.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
 
@@
 
-How can I highlight the result (I need to verify this first) for greater IFD in high rec male mice?
-
 The raw measures are pretty nosy, this might not be the best way to display these results.
 
+ Each point is a IFD observation. Maybe add mouse level lines for the means.
+
 ```{r show.IFD, echo=FALSE, warning=FALSE}
-colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
-                       #blues
-
-                       'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
-                       'CZECH'= 'indianred4',
-
-                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
-
-                    'CAST'='green','HMI'= 'forestgreen', #greens
-
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
-                    'CAROLI'='gold3',
-
-                    "F1"='grey', "other" = 'grey'
-                    )
+
 #first draft of the IFD comparisons
@@
   ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
 boxplots_first.IFD.plot_ABS
@@
   geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
 boxplots_first.IFD.plot_PER
 
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
 
@@
 #+theme(legend.position="none")
-
-```
-
-```{r blah.again}
+```
+
+```{r blah.again, echo=FALSE, eval=FALSE}
 #long SC Long_biv_mouse.avs_4MM
 #real.long.bivData
-
 #Only use 2CO bivalents
@@
   ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-
-long.biv_IFD_look_ABS
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
 
@@
   ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
-
-long.biv_IFD_look_PER
-
-```
-
-Above is the first plots of IFDs. Each point is a IFD observation. colors are by mouse. 
-Maybe add mouse level lines for the means
-
-Maybe I should try binned the SC lengths.
-
-mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+```
+
+### Fixed Effects
+
+Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
 
@@
 
-```{r MM_1.SC.Length, echo=FALSE}
+```{r MM_IFB.ABS, echo=FALSE}
 #get a list of the old mice to double check
@@
 
-### second set of Mixed Models
+Sex is a slightly significant factor for raw and normalized IFD.
+
+### Random Effects
 
@@
 
-library(RLRsim)
 IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-anova(IFD_ABS.MM)
+#anova(IFD_ABS.MM)
 
@@
 IFD.ABS_chi <- drop1(IFD_ABS.MM, test="Chisq")
-IFD.ABS_chi
 
@@
 IFD_PER.MM <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-anova(IFD_PER.MM)
+#anova(IFD_PER.MM)
 
 Chis.sp_IFD.PER.MM <- drop1(IFD_PER.MM, test="Chisq") 
-Chis.sp_IFD.PER.MM
+#Chis.sp_IFD.PER.MM
 
@@
 
-```{r compare.IFD}
-
-#make t.tests for pooled high and low
-
-```
+Random effect of strain is not significnt for ABS IFD, and only slightly significant for the IFD.PER
+
+#### IFD 3CO bivalents
+
+Run comparisons for 3CO bivalents.
 
@@
 
+```{r compare.IFD,echo=FALSE}
+#make t.tests for pooled high and low
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
+Curated_BivData <- add_IFD(Curated_BivData)
+
+Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
+     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 )]
+)#p-value = 7.955e-05
+
+Rec.group_IFD.ABS_test$p.value
+
+Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
+     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 )]
+)#p-value = 0.04231
+
+Rec.group_IFD.PER_test$p.value
+
+biv.IFDs <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,100))
+
+oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~strain, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+
+```
+
 Neither t.test are sig for both the ABS and PER when I test just the Musc strains.
 The above t.tests are breaking the knitr
+
+The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
 
@@
 
-for the long chrm, pull the object ids out of the melt table.
+#### Predictions
+
+**Heterochiasmy**
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+**Polymorphism**
+
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
+
+C.3) telomere and centromere distances will be shorter in low rec strains
 
@@
 
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
 CO.Pos.density_1CO
@@
 
-Above plot focuses on the 1CO bivalent normalized positions -- since CO interference controls the general position of COs when there are multiple COs.
+Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
 This plot shows the sexual dimorphism in the density plots.
 
-consider adding annotate_text for the number of observations in each category.
+Consider adding annotate_text for the number of observations in each category.
 think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
@@
 
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
 Dom.male_MELT_long
@@
 
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
 Musc_male_MELT_long
@@
 #there are v few observations in this dataset
-
-```
-
-Tired runnin plots for the long biv data set, I think there are too few observations.
-
-add boxplots for
+```
+
+<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
 
@@
 
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+geom_jitter(alpha=.1)
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("IFD across bivalents")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
 CO.Pos.boxplots_1CO
@@
 
+mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
+
+#<- lme()
+lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
+
 #can't make the 
@@
 #anova(CO_position.MM)
-anova(CO_position.MM, verbose = TRUE)
-
-summary(CO_position.MM) #no p values,,
+#anova(CO_position.MM, verbose = TRUE)
+
+#summary(CO_position.MM) #no p values,,
 #these just give t and F stats, no pvalue
@@
 #test high vs low Musc strains
-mouse.avs_4MM
+#mouse.avs_4MM
 
@@
 
+<!-- 
 # old drafts
@@
 ```
+-->

### 2020-01-24 src/CommonFunc_MLH1repo.R

@@ -1,4 +1,3 @@
-# Common MLH1repo functions
 #
 #
 
@@ -396,10 +395,21 @@ add_IFD <- function(DF){
   DF$IFD1_ABS <- as.numeric(DF$IFD1_ABS)
   
   
-  DF$IFD1_PER <- ifelse(  (DF$hand.foci.count >= 2), 
-                          as.numeric(DF$Foci2.PER) - as.numeric(DF$Foci1.PER),   "" )
+  DF$IFD2_ABS <- ifelse(  (DF$hand.foci.count == 3),
+                          #manual.verified.f1, 
+                          as.numeric(DF$Foci2) - as.numeric(DF$Foci3),   "" )
+  DF$IFD2_ABS <- as.numeric(DF$IFD2_ABS)
   
-  DF$IFD1_PER <- as.numeric(DF$IFD1_PER)
+  
+  DF$IFD3_ABS <- ifelse(  (DF$hand.foci.count == 4),
+                          #manual.verified.f1, 
+                          as.numeric(DF$Foci3) - as.numeric(DF$Foci4),   "" )
+  DF$IFD3_ABS <- as.numeric(DF$IFD3_ABS)
+  
+  
+ # DF$IFD1_PER <- ifelse(  (DF$hand.foci.count >= 2), 
+  #                        as.numeric(DF$Foci2.PER) - as.numeric(DF$Foci1.PER),   "" )
+  #DF$IFD1_PER <- as.numeric(DF$IFD1_PER)
   
   return(DF)
 }

### 2020-01-24 doc/Bivalent_Analysis.Rmd
spell checks

@@
 Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script.
-These observations are from the automated image analysis algorithm and have been currated (threw out incorrect algorithm output).
+These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output).
 
@@
 
-- Look into ordering the mouse strains to highlight the MSM and PWD differences, renameing the chrm type labels, spacing out the x axis labels or removing some of the other mice.
+- Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.
 
@@
 
-In an attempt to test for and metigate noise based on chromosome identity/physical size, I tried to isolate bivalents which are in the top quartille for SC length from their cells.
-
-Since the majority of cells from the Automated BivData set do not have the complete completement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
+In an attempt to test for and mitigate noise based on chromosome identity/physical size, I tried to isolate bivalents which are in the top quartille for SC length from their cells.
+
+Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
 But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
@@
 
-Each point is a bivalent ploted by cell on the x axis.
+Each point is a bivalent plotted by cell on the x axis.
 X's are the 4th quartile, big point is the mean and smaller black point is the median.
-I'm using these to compare the patterns of these statistics in the automated dataset which is missing some bivalent data.
+I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
 
@@
 
-All the plots above show the distrtributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
+All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
 It shows the amount of within cell variance across strains.
@@
 
-I'm purposing to take all the automated biv measures above the cell's 4th quartille as a dataset for 'longest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromsome size effects.
-
-But this dataset might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females). But it's easier to just run another set of analysis.
+I'm purposing to take all the automated biv measures above the cell's 4th quartille as a data set for 'longest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromosome size effects.
+
+But this data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females). But it's easier to just run another set of analysis.
 
@@
 
-The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. full dataset is 
+The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. full data set is 
 `r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category `r data.frame(table(droplevels(real.long.bivData$category)))`.
@@
 They are supposed to be the longest 4-5 SC from cells where I could get good measures.
-These longer bivalents are useful because their patterns shouldn't be affect by chromsome size effect (which effects, CO position).
+These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
 Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
@@
 
-- Of all female single bivalents observations, 5% are XXs (1 of 20).
+- Of all female single bivalents observations, 5% are XX (1 of 20).
 
@@
 
-- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell dataset. Lets guess 12% of a cell's total SC area is XX.
+- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
 
@@
 
-- The data are continous and closer to normal/gaussian expectations than the MLH1 counts.
+- The data are continuous and closer to normal/Gaussian expectations than the MLH1 counts.
 
@@
 
-- Limited work to add to the dataset if I can justify that the collection of single bivalents are random and equivilent across mice.
+- Limited work to add to the data set if I can justify that the collection of single bivalents are random and equivalent across mice.
 
@@
 
-1) SC length will be sexually dimorphic (sex effect will be signicant)(cite Lynn)
+1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
 
@@
 
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of uniform vs telomere/biased CO positioning.
 
@@
 
-TRY ADDING NESTED boxplots for hand.foci.count
+TRY ADDING NESTED box plots for hand.foci.count
 
@@
 
-remeber the long biv data set in **real.long.bivData**
+remember the long biv data set in **real.long.bivData**
 
@@
 - Sex is a significant effect for SC length
-The results seem to indicate that sex is a significant factor. Consider writing a subsampling approach (randomize / permute a data set of BivData)
+The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
 
@@
 
-Preidction, High rec males have longer SC.
+Prediction, High rec males have longer SC.
 
@@
 In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
-I counldn't get the Mixed models working for the male polymorphism predictions...
+I couldn't get the Mixed models working for the male polymorphism predictions...
 
@@
 
-These seond set of models are made using the lmer()
+These second set of models are made using the lmer()
 
@@
 
-Random effect of strain is not significnt for ABS IFD, and only slightly significant for the IFD.PER
+Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
 
@@
 
-The random strain effect was tested for raw and normalized IFD, neither are signifcant.
+The random strain effect was tested for raw and normalized IFD, neither are significant.
 
@@
 
-None of the logistic regression models for ABS or PER IFD lengths are sigificant, even when just the Musc strains are used.
+None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
 
@@
 
-Preliminary results from an independant dataset indecated that PWD had longer IFDs, which goes agaisnt the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
-This also indecated that interference distance may evolved in the house mouse complex.
+Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
+This also indicated that interference distance may evolved in the house mouse complex.
 
@@
 
-These bosplots show that females have a much more medial poition of single foci bivalents, (much closer to 50% compared to males).
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
 They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
@@
 The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension ative cohesin amounts as a consequence of different numbers and placements of chiasmata/CO.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
 This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
@@
 
-Males have much clearer seperation of siscoten across chrm classes.
+Males have much clearer separation of siscoten across chrm classes.
 This is emphasized when SC length is also plotted.
@@
 
-To formally test the differences in sis-co-ten I plan to write a subsampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
 
@@
 
-Yes, model od predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
 
@@
 In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistant...
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
 
@@
 
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He sugggested doing simple t-test for the positions
+Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
 

### 2020-01-24 doc/Bivalent_Analysis.Rmd
coffee shop work

@@
 These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output).
+The MLH1 data file is also loaded into this file.
 
@@
 
--
+I think a formula something like this can be applied to adjust for XX
+
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC araea due to XX
 
@@
               #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE)
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
 )
@@
 #MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-
 #code if males are high or low?
@@
 
-first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)
+first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
 first.SC.plot
@@
 
+#### Mixed Model Tests
+
 $$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
@@
 
-Above is the model we are testing evolution of SC lengths.  remember for the MM use the table **mouse.avs_4MM**
-
-remember the long biv data set in **real.long.bivData**
+The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.
+
+remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
 
@@
 
+explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
+
 3) Interference / IFD will not be sexually dimorphic.
@@
 
-### Normalized 1CO positions
+# Normalized 1CO positions
 
@@
 
+```{r MM.telo, echo=FALSE}
+
+lme_MM_telo.dist <- lme(mean.telo.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_telo.dist)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+```
+
 ```{r cent.dist, echo=FALSE}
-
 cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
@@
 
+```
+
+```{r MM.cent.dist, echo=FALSE}
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
 ```

### 2020-01-24 doc/Bivalent_Analysis.Rmd
first run of results

@@
                     )
+
+#add Rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
 ```
@@
 
+```{r XX.adj, echo=FALSE}
+
+#load and examine the the whole cell hand measures here
+
+```
+
 # Applying Mix Model framework
@@
 
-### Male polymorphism Predcition
+### Male Polymorphism Prediction
 
@@
 
+```{r log.reg_SC.length, echo=FALSE}
+
+mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+#divide musc males into groups
+mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
+                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
+#
+
+log.reg_SC.length <- glm(Rec.group ~ mean_SC, 
+                            data=mouse.avs_4MM_male, family=binomial(link="logit"))
+
+log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
+                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))
+mouse.av_log.reg_SC.length <- summary(log.reg_SC.length_Musc)
+
+# bivalent level log regression
+log.reg_SC.length_Musc_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+bivalent_log.reg_SC.length <- summary(log.reg_SC.length_Musc_BIV)
+
+```
+
+The mean SC logistic regresion model for mouse averages
+
+`r mouse.av_log.reg_SC.length$coefficients`
+
+and for the single bivalent levels
+
+`r bivalent_log.reg_SC.length$coefficients`
+
+#### t.test for sc length
+
+```{r t.test, echo=FALSE}
+
+T.test_Poly_SC.Len_mouse.av <- t.test(mouse.avs_4MM_male$mean_SC[mouse.avs_4MM_male$Rec.group == 1],
+
+  mouse.avs_4MM_male$mean_SC[(mouse.avs_4MM_male$subsp == "Musc")&(mouse.avs_4MM_male$sex == "male")&(mouse.avs_4MM_male$Rec.group == 0)]
+)#p-value = 0.001567
+
+T.test_Poly_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
+
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)]
+)#p-value < 2.2e-16
+
+```
+
+The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_Poly_SC.Len_mouse.av$p.value` and p=`r T.test_Poly_SC.Len_single.biv$p.value`.
+
 ```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
+#I don't think the Mixed models fit well for the high vs low polymorphism tests.
+
 #t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
@@
 rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
 ```
@@
 
-```{r logreg, echo=FALSE}
-
-mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
-                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
-
-log.reg_SC.length <- glm(Rec.group ~ mean_SC, 
-                            data=mouse.avs_4MM_male, family=binomial(link="logit"))
-
-log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
-                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))
-
-summary(log.reg_SC.length_Musc)
-
-#mean SC slightly sig-- these are all mean SC--should I try distinguishing by chrm classes?
-```
-
-```{r long.sc_log.res_meanSC}
+```{r long.sc_log.res_meanSC, eval=FALSE}
+#hold off on reporting/ working with the long SC biv resutls
 
@@
 #divide musc males into groups
-long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", long.SC.mouse.avs_4MM_male$strain), 1, 
-                                    ifelse(grepl("MSM", long.SC.mouse.avs_4MM_male$strain), 1, 0))
+long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
+                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
 
@@
 # IFD
+
+ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
 
@@
 
-```{r blah.again, echo=FALSE, eval=FALSE}
+```{r IFD.old.plots, echo=FALSE, eval=FALSE}
 #long SC Long_biv_mouse.avs_4MM
@@
 
+The polymorphism frame-work
+
 > prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
 
-```{r IFD.t.test, echo=FALSE, eval=FALSE}
-
-#pool the whole groups
-
-ABS.t.test <- t.test( mouse.avs_4MM_M.male$mean_IFD.2CO_ABS[mouse.avs_4MM_M.male$Rec.group == 1],
-  mouse.avs_4MM_M.male$mean_IFD.2CO_ABS[mouse.avs_4MM_M.male$Rec.group == 0])
-
-PER.t.test <- t.test( mouse.avs_4MM_M.male$mean_IFD.2CO_PER[mouse.avs_4MM_M.male$Rec.group == 1],
-  mouse.avs_4MM_M.male$mean_IFD.2CO_PER[mouse.avs_4MM_M.male$Rec.group == 0])
-
-#compare strain by strain
-# p = `r ABS.t.test$p.value`  and p = `r PER.t.test$p.value`.
-```
-
-```{r compare.IFD,echo=FALSE}
-#make t.tests for pooled high and low
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
-Curated_BivData <- add_IFD(Curated_BivData)
-
-Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
-     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 )]
-)#p-value = 7.955e-05
-
+logistic regression
+$$  Rec \ Group ~=~ mouse \ average \ IFD $$
+For these tests I am looking at the single IFD measure from 2CO bivalents.
+
+```{r IFD.log.reg, echo=FALSE}
+#
+log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS, 
+              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *
+
+log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+summary(log.reg_IFD_PER_mouse.av)#NS
+
+### Single Bivalent levels
+
+log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
+              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_IFD_ABS_single.biv)
+
+log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
+              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_IFD_PER_single.biv)
+```
+
+```{r t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
+
+#mouse.av - the mouse average IFD is from just 2CO
+t.test_mouse.av.IFD_ABS <- t.test(
+  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+t.test_mouse.av.IFD_PER <- t.test(
+
+  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+```
+
+```{r t.test_IFD,echo=FALSE}
+
+#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+#                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#THIS SHOULD ALREADY BE ADDED
+
+#single biv level
+
+Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 ) & (Curated_BivData$hand.foci.count == 2 )],
+
+     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 ) & (Curated_BivData$hand.foci.count == 2 )]
+)
 Rec.group_IFD.ABS_test$p.value
 
-Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
-     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 )]
+## PER
+
+Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )& (Curated_BivData$hand.foci.count == 2 )],
+
+     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 ) &  (Curated_BivData$hand.foci.count == 2 )]
 )#p-value = 0.04231
@@
 
-biv.IFDs <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+#Single Biv Leve
+single.biv.ABS.t.test <- t.test( Curated_BivData$IFD1_ABS[Curated_BivData$Rec.group == 1],
+
+         Curated_BivData$IFD1_ABS[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
+
+single.biv.IFD_PER.t.test <- t.test( Curated_BivData$IFD1_PER[Curated_BivData$Rec.group == 1],
+
+         Curated_BivData$IFD1_PER[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
+```
+
+```{r show.IFD, echo=FALSE}
+
+biv.IFDs_by_group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
   geom_jitter(alpha=.2)+
@@
   scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,100))
+biv.IFDs_by_group
+
+biv.IFDs_PER_by.group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS \n Low=0, High=1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+biv.IFDs_PER_by.group
 
@@
   ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~strain, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
-
-```
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+
+```
+
+Re-order the above figure if possible.
 
@@
 
+```{r log.reg.CO1_pos, echo=FALSE}
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+```{r t.test_1CO_pos, echo=FALSE}
+
+#mouse av
+t.test_mouse.av.1CO_pos <- t.test(
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Leve;
+single.biv.1CO_pos_t.test <- t.test(
+
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 0)]
+  )
+
+#0single biv level
+```
+
+the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
 ### Siscoten
@@
 
-> is the male polymorphism
+```{r log.reg_siscoten, echo=FALSE}
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE}
+#mouse av
+t.test_mouse.av_siscoten <- t.test(
+
+  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.
 
@@
 
-```{r cent.dist, echo=FALSE}
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
 cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
@@
 exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
 ```

### 2020-01-24 doc/Bivalent_Analysis.Rmd
too much work on analysis

@@
 #1. how many biv's are from cells with >15 bivalents
-
 over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
@@
 
+female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
+data.frame(table(droplevels(female.whole.cell$category)))
+
+#Here is the data set of female whole cell measures, apply rank to these measures ... 
+#run a ddplyr to 
+#
+
 ```
@@
                         se   = round(sd / sqrt(Ncells), 3),
-                       #IFD things
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
                 #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
@@
               mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
 
@@
 mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-#unordered
+mouse.avs_4MM <- add_category(mouse.avs_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
 mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
@@
 #code if males are high or low?
+
+#assign high and low
+#divide musc males into groups
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                    ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+
 ```
@@
 
-3) Interference / IFD will not be sexually dimorphic. Previous measures has been shown to lack sexual dimorphism (cite petkov).
+3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
 
@@
 
-```{r MM_2_SC.length, echo=FALSE, eval=FALSE}
+```{r MM_2_SC.length_2_DELETE, echo=FALSE, eval=FALSE}
 #Long_biv_mouse.avs_4MM
@@
 ```{r log.reg_SC.length, echo=FALSE}
-
-mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
-                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
-#
-
-log.reg_SC.length <- glm(Rec.group ~ mean_SC, 
-                            data=mouse.avs_4MM_male, family=binomial(link="logit"))
-
-log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
-                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))
-mouse.av_log.reg_SC.length <- summary(log.reg_SC.length_Musc)
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
+
+#1CO
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+summary(log.reg_SC.length_mouse.av_1CO)
+
+#2CO
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+
+summary(log.reg_SC.length_mouse.av_2CO)
+
+#3CO
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+summary(log.reg_SC.length_mouse.av_3CO)
 
 # bivalent level log regression
-log.reg_SC.length_Musc_BIV <- glm(Rec.group ~ chromosomeLength, 
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
                             data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
 
-bivalent_log.reg_SC.length <- summary(log.reg_SC.length_Musc_BIV)
+summary(log.reg_SC.length_BIV)
+
+#1CO
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_SC.length_BIV_1CO)
+
+#2CO
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 2),], family=binomial(link="logit"))
+summary(log.reg_SC.length_BIV_2CO)
+
+#3CO
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 3),], family=binomial(link="logit"))
+
+summary(log.reg_SC.length_BIV_3CO)
 
@@
 
-`r mouse.av_log.reg_SC.length$coefficients`
+`r log.reg_SC.length_mouse.av$coefficients`
 
@@
 
-`r bivalent_log.reg_SC.length$coefficients`
-
-#### t.test for sc length
-
-```{r t.test, echo=FALSE}
-
-T.test_Poly_SC.Len_mouse.av <- t.test(mouse.avs_4MM_male$mean_SC[mouse.avs_4MM_male$Rec.group == 1],
-
-  mouse.avs_4MM_male$mean_SC[(mouse.avs_4MM_male$subsp == "Musc")&(mouse.avs_4MM_male$sex == "male")&(mouse.avs_4MM_male$Rec.group == 0)]
+`r log.reg_SC.length_BIV$coefficients`
+
+#### t.test for SC Length
+
+```{r SC.Length_t.test, echo=FALSE}
+
+T.test_SC.Len_mouse.av <- t.test(mouse.avs_4MM$mean_SC[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean_SC[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")&(mouse.avs_4MM$Rec.group == 0)]
 )#p-value = 0.001567
 
-T.test_Poly_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
-
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)]
-)#p-value < 2.2e-16
-
-```
-
-The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_Poly_SC.Len_mouse.av$p.value` and p=`r T.test_Poly_SC.Len_single.biv$p.value`.
+#1CO
+T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#2CO
+T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#3CO
+T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#Single Biv level
+T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
+
+#1CO
+T.test_SC.Len_single.biv_1CO <- t.test(
+  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
+
+      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
+  )
+
+#2CO
+T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
+
+                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
+  )
+
+  #3CO
+  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
+
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
+  )
+```
+
+The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_SC.Len_mouse.av$p.value` and p=`r T.test_SC.Len_single.biv$p.value`.
 
@@
 #isolate data
-mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
 
@@
 rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
+
 ```
@@
 
-```{r show.IFD, echo=FALSE, warning=FALSE}
+```{r show.IFD_99, echo=FALSE, warning=FALSE}
 
@@
 
-```{r IFD.poly_log.reg, echo=FALSE}
-
+```{r IFD.poly_log.reg, echo=FALSE, eval=FALSE}
+#This might be a duplicate code chunk
 #test. for the high vs low strains and IFD
@@
 
-#Single Biv Leve;
+#Single Biv Level
 single.biv.1CO_pos_t.test <- t.test(
 
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 0)]
-  )
-
-#0single biv level
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+  )
+
 ```
@@
 
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
+
+   mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
+
+   mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
+
+   mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  )
+
+####
 #Single Biv Level
@@
   )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+#2CO
+
+#3CO
+
 ```

### 2020-01-27 doc/Bivalent_Analysis.Rmd
small work from home

@@
 
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC araea due to XX
-
-```{r XX.adj, echo=FALSE}
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC araea due to XX /*
+
+#### Whole Cell Manual Comparison
+
+```{r XX.adj_manual.whole.cell, echo=FALSE}
 
@@
 
-#Here is the data set of female whole cell measures, apply rank to these measures ... 
-#run a ddplyr to 
-#
-
-```
-
-# Applying Mix Model framework
+# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
+#cell 5aug17_14jul17_LEW_f1_sp1_11_rev is missing measures for 2 bivs
+
+data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
+#missing some strains, most have 3 cells (min). MSM has 5 full cells
+
+#remove the incomplete MSM cell
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
+
+#calq mean SC lengths by rank
+
+mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
+
+                       mean.sc =   mean(SC.length, na.rm = TRUE),
+                         nbiv = length(Obj.ID),
+                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
+                         se.SC_length   = sd.SC_length / sqrt(nbiv)
+)
+         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
+
+top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
+
+show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
+    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
+                    position=position_dodge(width=0.7), 
+                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
+          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
+
+show.SC_by.rank_top
+
+bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
+
+show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
+    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
+                    position=position_dodge(width=0.7), 
+                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
+          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
+
+#CALQ PERCENT TOTAL SC
+
+CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
+
+                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
+                         nbiv = length(Obj.ID),
+                       #rank.by.per
+
+                      rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
+                      rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
+                      rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
+                      rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
+                       rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
+                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
+                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
+)
+
+#MELT ABOVE
+MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
+              #leave out the Foci
+               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
+
+            id=c( "fileName","total.sc.cell","nbiv"))
+
+MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
+
+MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
+MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
+
+show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
+
+show_long.sc.proportions
+
+#original plot code 
+#ggplot(data = df1, aes(x, y, color = Group)) +
+#    geom_pointrange(aes(ymin = lb, ymax = ub), 
+#                    position=position_jitter(width=0.5), 
+#                    linetype='dotted') +
+#    theme_bw()
+```
+
+The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
+
+The first plot showing the mean sc lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
+
+The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
+For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
+
+(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
+
+The other figure shows of each single bivalent contributes to the total SC area.
+Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
+on average, each of the top longest bivalents make up ~10% of the cell's totl SC area.
+So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
+
+- is the difference between cell averages for males and females less that 10%?
+
+- also interesting, the pwd and msm dont have longer sc, compared to other strains.
+
+#### Automated BivData Comparison
+
+```{r XX.Automated.biv.stuff, echo=FALSE}
+
+#make automated female data to compare to
+female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
+
+female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
+
+table(female.Bivdata_comp$strain)
+
+#female.Bivdata_comp
+
+#1. calq the rate of chrm segmentation in female categories in automateted data
+
+box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
+
+                   Nmice = length(unique(mouse)),
+                   mx.box.num = max(boxNumber), # need to run this at the cell level
+                    N.biv  = length(Obj.ID),
+                   pass.biv = N.biv / mx.box.num
+)
+box.rates <- add_mouse(box.rates)
+box.rates <- add_strain(box.rates)
+box.rates <- add_category(box.rates)
+
+cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+
+cell.pass.rates
+
+category.box.rates <- ddply(box.rates, c("category"), summarise,
+
+                   mean.pass.rate = mean(pass.biv)
+)
+#category.box.rates
+
+#2. pull the 3rd rank biv's calculate the SC means
+#calq means by rank (focus on the top 5)
+
+```
+
+For the Automated dataset, I like to measure the rate of passing bivlents per cell. The mean pass rate will be multipled to the estimated XX mean_SC.
+
+The table above shows the number of bivalents from the same strains as in the manual whole cell data.
+The plot shows the bivalent passing rate across all of the individual cells from this female data set.
+For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
+
+(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be calqd on the mouse level.
+(it could even be extended to cell level -- except i dont think the XX sc length estimates wont be good.)
+
+strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
+
+# Applying Mix Model Framework
 

### 2020-01-27 src/BivData_Setup.Rmd
re-ran the setup with corrected FULL data csvs

@@
 
-update IFD functions, to add IFD2,3
+update IFD functions, to add IFD2,3 -- report the summaries
 
@@
 
+```
+
+```{r input.check,echo=FALSE}
+
+#check for mistakes that might have been made when inputing / curation process
+
+#hand foci > 4
+
+#foci pos > SC lengths
+
+#negative PER.foci.pos
+
+#it hink most (ALL) of these came from bug in my function (ugh)
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#negative.IFD <- Curated_BivData[which(Curated_BivData$IFD1_ABS < 0),]
+#caught 2
+#15mar16_16jan16_G_f1_sp1_4_rev
+#27mar16_6mar16_G_m3_sp1_12_rev
+
+#neg Nrm F1
+#negative.IFD2 <- Curated_BivData[which(Curated_BivData$PER_Foci_1 < 0),]
+#0 really?
 ```
@@
 #adding more feature calculations (IFD, ect)
-Curated_BivData <- add_IFD(Curated_BivData)
+Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
 Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_1.22.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_1.27.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-01-27 doc/Bivalent_Analysis.Rmd
last bit o work

@@
 # cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
-#cell 5aug17_14jul17_LEW_f1_sp1_11_rev is missing measures for 2 bivs
+
+#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
 
@@
 
-#remove the incomplete MSM cell
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
+#remove the incomplete LEW female cell
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
 
@@
 
-- is the difference between cell averages for males and females less that 10%?
+- Is the difference between cell averages for males and females less that 10%?
 
@@
 strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
+
+** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
+But I am not using 'whole cell' summaries to compare female and males.
+
+**What is the effect of an extra XX-autosome on single bivalent means?**
+
+use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
+Make fake datasets which sample 19 bivalents, for 'in silico' cells for males and females.
+Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
+Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
+The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total dataset.
 

### 2020-01-27 results/XX_adjust_permutations.R

@@ -40,11 +40,8 @@ Means.Here_WSB <- data.frame(
 for(i in 1:length(Means.Here_WSB$WSBmean_sc_female20)){  #current data from
   print(i)
   #filling in metrics for Means.Here
-  
   Means.Here_WSB$WSBmean_sc_female20[i] <- mean(WSBfemale.20.rand_BIG.DF.LIST[[i]]$chromosomeLength)#looks wacky, but this is right syntax
-
   Means.Here_WSB$WSBmean_sc_female19[i] <- mean(WSBfemale.19.rand_BIG.DF.LIST[[i]]$chromosomeLength)
-
   Means.Here_WSB$WSBmean_sc_male19[i] <- mean(WSBmale.19.rand_BIG.DF.LIST[[i]]$chromosomeLength)
 }
 
@@ -113,19 +110,243 @@ t.test(Means.Here_G$Gmean_sc_female20, Means.Here_G$Gmean_sc_female19) #p-value 
 
 
 
-#LEW
-
-
-
-
 #PWD
+#the script below mkes 1000 data sets 
+PWDfemale.20.rand_BIG.DF.LIST <- replicate(1000,
+                                           #add subsample to a list, this works!!!
+                                           list( bind_rows(list(Curated_BivData[ sample( which(Curated_BivData$category=="PWD female"), 20), ]))
+                                           ))
+PWDfemale.19.rand_BIG.DF.LIST <- replicate(1000,
+                                           #add subsample to a list, this works!!!
+                                           list( bind_rows(list(Curated_BivData[ sample( which(Curated_BivData$category=="PWD female"), 19), ]))
+                                           ))
+PWDmale.19.rand_BIG.DF.LIST <- replicate(1000, 
+                                         #add subsample to a list, this works!!!
+                                         list( bind_rows(list(Curated_BivData[ sample( which(Curated_BivData$category=="PWD male"), 19), ]))
+                                         ))
+#head(BIG.DF.LIST[[2]]) - the 2nd data set
+#### 
+### DFs above are dummy DF's for 1000 in silico cells
+#head(BIG.DF.LIST[[2]]) - the 2nd data set
+
+#summarize, calq the mean SC length across all the prelicates
+#I want to end up with 1000 mean.SC observations, to plot as boxplots
+
[…]

### 2020-01-28 doc/Bivalent_Analysis.Rmd
some small changes

@@
 #load Clean BivData
-load(file="data/CleanBivData_1.22.20.RData")
+load(file="data/CleanBivData_1.27.20.RData")
 
@@
 
-In an attempt to test for and mitigate noise based on chromosome identity/physical size, I tried to isolate bivalents which are in the top quartille for SC length from their cells.
-
+I tried to isolate bivalents which are in the top quartille for SC length from their cells.
+
+<!-- 
 Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
 But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
-
-Below are examples of plots which show the SC length distributions across cells.
+--> 
+Below are examples of plots of SC length distributions across cells.
 The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
-There are plot pairs for each category, but they are excluded for space.
+Most plts excluded for space.
 
@@
 I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
+(the extra stats are not correclty mapped)
 
@@
 
-I'm purposing to take all the automated biv measures above the cell's 4th quartille as a data set for 'longest' bivalents which we would predict to exclude the shortest chromsomes which would have stronger chromosome size effects.
-
-But this data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females). But it's easier to just run another set of analysis.
+This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
 
@@
 
-The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. full data set is 
-`r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category `r data.frame(table(droplevels(real.long.bivData$category)))`.
-
-Still very unbalanced across sexes and strains ~ merge with the manual whole cell biv measures later...
-Musc males are very under-represented. MSM male only has data from 1 cell.
-
-- Try to merge this DF with the whole.cell manual measures.
+The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
+The full data set is `r length(Curated_BivData$Obj.ID)`.
+This is the breakdown of bivalent observations by category for this long dataset are `r data.frame(table(droplevels(real.long.bivData$category)))`.
+
+**- Try to merge this DF with the whole.cell manual measures.**
 
@@
 
+new outline
+
+1. illustrate problem(affects mostly SC length)
+
+2. Expected impact on sex comparisons, extimated effect size of the X
+
+3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
+
+4. 19 female, 19male, 20female 20 male
+
 The female mouse averages should have adjustments for the XX.
@@
 female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
+
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
 
@@
 
+Brief lit review for SC length and MLH1 expections.
+
 ```{r show.SC.length, echo=FALSE}
-#compile data
-#main figures, show both mouse and bivalent level
-
-#Curated_BivData
-
-#Mouse_average - BivTable?
-
 #for now remove the 
-Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count <5,]
-
-first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
-first.SC.plot
+#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
+#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS
+
 #add boxplots for mouse level means
@@
 
-other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
 
@@
 
-#sfsL <- ggplot(lowGEN, aes(variable, prop, fill=as.factor(Group.2))) + geom_col(position = "dodge") + #scale_fill_viridis(name="Migration Rate", discrete = TRUE) + theme(axis.text.x=element_blank()) + #xlab("Frequency in the sample") + ylab("Proportion of Total Junctions") + ggtitle("C.
-Generation = #2,000") + ylim(0,1)
-
-#geom_boxplot(alpha = .2, position = "dodge")+
-
-```
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#nested boxplots
+SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
+SC.length_Nested_boxplots
+
+#+ +geom_jitter(aes(alpha=0.2))
+
+```
+
+In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean -- are likely due to low data number.
+
+For the 0 class chromosomes most all are around the same size of the 1CO distribution.
+
+Add in the code for the plot under 2COs males and females
 
@@
 # IFD
+
+Brief lit review for IFD / interference expectations  
+**Petkov et al 2007,  CO interference underlies sex differences in RR**
+
+"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
+However, the interference distance is the same in terms of bivalent length.
+We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
+at the pachytene stage of meiosis."
+
+Chrm1 genetic map,
+
+(other human and mice refs
+
+Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
+
+DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
+
+))
+
+*Do I believe the physical/SC results?*
 
@@
 ```
+
+```{r draft plots, eval=FALSE}
+first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
+```
+
 -->

### 2020-01-28 src/BivData_Setup.Rmd
some small changes

@@
 
-#rename foci positions
+#rename cols to match BivData
 whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
@@
 whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3
+whole.cell.data_manual$Foci4 <- whole.cell.data_manual$F4  #currently Foci4 is chr
+whole.cell.data_manual$chromosomeLength <- whole.cell.data_manual$SC.length
+
+whole.cell.data_manual$PER_Foci_1 <- whole.cell.data_manual$Foci1 / whole.cell.data_manual$chromosomeLength
+whole.cell.data_manual$PER_Foci_2 <- whole.cell.data_manual$Foci2 / whole.cell.data_manual$chromosomeLength
+whole.cell.data_manual$PER_Foci_3 <- whole.cell.data_manual$Foci3 / whole.cell.data_manual$chromosomeLength
+#whole.cell.data_manual$PER_Foci_4 <- whole.cell.data_manual$Foci4 / whole.cell.data_manual$chromosomeLength
 
@@
 whole.cell.data_manual <- add_IFD(whole.cell.data_manual)
+
+#PER Foci
+
+```
+
+```{r clean.whole.cell, echo = FALSE}
+
+#CHECKED FOR MISSING bivs -- also check for wrong directions, ect
+
+#negative IFD
+#5aug17_14jul17_LEW_f1_sp1_11_rev_blob4
+
+#PER foci positions greater th
 

### 2020-01-28 doc/Bivalent_Analysis.Rmd
in the middle of work and ideas

@@
 #load Clean BivData
-load(file="data/CleanBivData_1.27.20.RData")
+load(file="data/CleanBivData_1.28.20.RData")
 
@@
 
-Brief lit review for SC length and MLH1 expections.
-
-```{r show.SC.length, echo=FALSE}
+<!--Brief lit review for SC length and MLH1 expections. -->
+
+We expect female SC lengths to be longer (refs).
+
+In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
+
+For the 0 class chromosomes most all are around the same size of the 1CO distribution.
+
+Add in the code for the plot under 2COs males and females
+
+```{r show.SC.length, echo=FALSE, warning=FALSE}
 #for now remove the 
@@
 
-SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
+#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
 
@@
 SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
+
 SC.length_Nested_boxplots
@@
 #+ +geom_jitter(aes(alpha=0.2))
-
-```
-
-In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean -- are likely due to low data number.
-
-For the 0 class chromosomes most all are around the same size of the 1CO distribution.
-
-Add in the code for the plot under 2COs males and females
-
-First draft of the SC length plot. I'd like to add the hand foci count to the x axis.
-(also add the PWD - male data)
-
-TRY ADDING NESTED box plots for hand.foci.count
-
-#### Mixed Model Tests
+```
+
+The dependant variables I'll be testing in the mixed model framework are:
+
+- Pooled SC lengths
+
+- 1CO
+
+- 2CO
+
+- 3CO
+
+- long.biv
+
+### Mixed Model Tests
 
@@
 remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
+
+#### Pooled SC Lengths
 
@@
 ```{r MM_2_SC.length_2_DELETE, echo=FALSE, eval=FALSE}
+#review this block!!
+
 #Long_biv_mouse.avs_4MM
@@
 
-```{r justify.mouse.level, echo=FALSE, eval=FALSE}
-
-#start the permutation code here
-#make a randomized BivData DF
-
-#30 cells worth of bivs? 
-#30 cells * ( 19 or 20 bivs ) * 70% segment rate = decent sample size for comparing Biv-based patterns?
-# I think this would address if the sample sizes are affecting the pattern
-
-#
-#justify mouse averages
-#is sampling -- bivalent dataset 
-# random samples 
-#equivalent sample size
-
-```
-
-### Heterochiasmy Prediction
+#### 1CO
+
+Insert the results for the MMs using the dependant variable 1CO
+
+#### 2CO
+
+#### 3CO
+
+#### long.biv
+
+## Strain Comparisons
+
+Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
+
+```{r WSB, echo=FALSE}
+#make the dataset
+WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
+#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
+table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
+wsb.ploty
+
+#make a table of the means and n ect
+wsb.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+```
+
+```{r LEW, echo=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+LEW.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+LEW.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+
+```
+
+```{r G, echo=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+G.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+G.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+```
+
+```{r PWD, echo=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+PWD.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+PWD.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+```
+
+```{r KAZ, echo=FALSE}
+#make the dataset
+KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+KAZ.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+KAZ.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+
+```
+
+```{r SKIVE, echo=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+SKIVE.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+SKIVE.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+```
+
+```{r MSM, echo=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MSM.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+```
+
+```{r MOLF, eval=FALSE}
+#make the dataset
+MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MOLF.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MOLF.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+### Summary
 
@@
 
+<!--
 ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
@@
 explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
+-->
 
@@
 
+## Strain Comparisons
+
+Dive deeper into the sex specific pattern for each strain.
+
+```{r WSB.IFD, echo=FALSE}
+#make the dataset
+#WSB.Bivalents.DF
+
+table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
+
+#plot the differences
+wsb.ploty_IFD <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex, color= as.factor(mouse)))+geom_jitter()+ggtitle("")
+wsb.ploty_IFD
+
+#try plotting the other IFD
+
+#make a table of the means and n ect
+wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+
+                       mean_SC = mean(chromosomeLength),
+                       n_SC = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats?
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+```
+
+```{r LEW.IFD, echo=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+LEW.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+LEW.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+
+```
+
+```{r G.IFD, echo=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+G.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+G.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+```
+
+```{r PWD.IFD, echo=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+PWD.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+PWD.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+```
+
+```{r KAZ.IFD, echo=FALSE}
+#make the dataset
+KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+KAZ.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+KAZ.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+
+```
+
+```{r SKIVE.IFD, echo=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+SKIVE.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+SKIVE.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+```
+
+```{r MSM.IFD, echo=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MSM.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+```
+
+```{r MOLF.IFD, eval=FALSE}
+#make the dataset
+MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MOLF.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MOLF.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
 #### IFD 3CO bivalents
@@
 <!-- 
-# old drafts
+# old drafts / trash pile
+
+```{r justify.mouse.level, echo=FALSE, eval=FALSE}
+
+#start the permutation code here
+#make a randomized BivData DF
+
+#30 cells worth of bivs? 
+#30 cells * ( 19 or 20 bivs ) * 70% segment rate = decent sample size for comparing Biv-based patterns?
+# I think this would address if the sample sizes are affecting the pattern
+
+#
+#justify mouse averages
+#is sampling -- bivalent dataset 
+# random samples 
+#equivalent sample size
+
+```
 

### 2020-01-28 results/Chapter3_ideas99.Rmd
in the middle of work and ideas

@@
 whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
+
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
 #load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
@@
 
+whole.cell.data <- whole.cell.data[!(is.na(whole.cell.data$SC.length) | whole.cell.data$SC.length==""), ]
+
 #ADD THE RANK -- this stoped working for a bit, 
@@
 female.biv.count_table$ratio.now <- var(refMus.geome_Mb_XX) / female.biv.count_table$var_sc
+
 female.biv.count_table$ratio.invert <-  female.biv.count_table$var_sc / var(refMus.geome_Mb_XX)
 
-```
+#plot these variances, do them for male and female
+
+```
+
+```{r var.ratios.plot, echo=FALSE}
+
+#var(refMus.geome_Mb_XX) #1146.608
+#var(refMus.geome_Mb_male) #1109.412
+
+#female.biv.count_table, table of 18 cells with their summary stats
+
+female.cell.var.ratios <- ggplot(female.biv.count_table, aes(ratio.now, fileName))+geom_point()
+female.cell.var.ratios
+
+male.cell.var.ratios <- ggplot(male.biv.count_table, aes(ratio.now, fileName))+geom_point()
+male.cell.var.ratios
+
+female.biv.count_table$sex <- "female"
+male.biv.count_table$sex <- "male"
+
+biv.ratio.table <- rbind(female.biv.count_table, male.biv.count_table)
+
+biv.ratio.table <- add_mouse(biv.ratio.table)
+biv.ratio.table <- add_strain(biv.ratio.table)
+
+all.ratios <- ggplot(biv.ratio.table, aes(y=ratio.now, x=strain, color=strain))+geom_jitter()+facet_wrap(~sex)+ggtitle("")+ylab(" var(chrm size Mb) : var(SC lengths)")+theme_bw()
+
+```
+
+A question which is fundamental to these Chap3 analyses:
+
+ - Is chrom axis compaction an accurate reflection on the actual physical chrm size? (is ranked SC length a decent predictor for chromosome identity?)
+
+Trying to address this by calculating the ratio between the within cell variance of SC lengths and the variance of chromsomes, ohysical sizes (mb) (haploid sizes from ref genome)
+
+The variance across SC is always smaller than variance across chrm sizes. could be due to different resolution of microscope (SC).
+
+what is a better name for this ratio? Chrm size variation (mB), and bivalent length variation (uM)
+
+true compaction ratio, compaction fedelity ratio, ... it's a ratio of variances, so those might be not be the best choices
+
+there is a negative relatioship with total SC and this ratio
+
+- is total SC a good metric for comparing with this?
+
+-is it strange that males have a clear min/higher min?
+
+-do the strains cluster with their ratios? ... still unbalanced data
+
+Male cell ratios are mostly
+
+above are ratios for indiviual cells, ratio of the variance of the physical chrm lengths (predicted, no centromere or telomere) to the variance of bivalents in SC.
 

### 2020-01-28 src/BivData_Setup.Rmd
in the middle of work and ideas

@@
 test work from home ()
+
+check the numbers of bivalents with a SC length, but no hand.foci.count
 
@@
 #PER foci positions greater th
+
+#check the numbers of bivalents with a SC length, but no hand.foci.count
 
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_1.27.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_1.28.20.RData")
 #save.image("data/BivData_4_PCA.RData")
@@
 
-- chrom class proportions
+-chrom class proportions
 
-- SC lengths, by CO number
+-SC lengths, by CO number

### 2020-01-29 doc/Bivalent_Analysis.Rmd
new pattern for sex diffs in IFDs

@@
 #plot the differences
-wsb.ploty_IFD <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex, color= as.factor(mouse)))+geom_jitter()+ggtitle("")
-wsb.ploty_IFD
+wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_raw
+
+wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_PER
 
@@
 LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
 LEW.ploty
 #remove the NAs
+
+wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_raw
+
+wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_PER
 
@@
 
-#plot the differences
-G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-G.ploty
+#plot the metric
 #remove the NAs
+
+G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
+G.ploty_IFD_raw
+
+#
+G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
+G.ploty_IFD_PER
 
@@
 #plot the differences
-PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-PWD.ploty
 #remove the NAs
+
+PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
+PWD.ploty_IFD_raw
+
+PWD.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
+PWD.ploty_IFD_PER
 
@@
 #plot the differences
-KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-KAZ.ploty
+
 #remove the NAs
+KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_raw
+
+KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_PER
 
@@
 #plot the differences
-SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-SKIVE.ploty
+
 #remove the NAs
+SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_raw
+
+SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_PER
 
@@
 
+MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_raw
+
+MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_PER
+
+#another plot
+MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
+
+MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
+
+#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
+
 #make a table of the means and n ect
@@
 #remove the NAs
+molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
+molf.ploty_IFD_raw
+
+molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
+molf.ploty_IFD_PER
 

### 2020-01-29 doc/Bivalent_Analysis.Rmd
notes on IFD and sis-co-ten patterns

@@
 
-#add Rec group
+#add Rec group,
+#reclassified SKIVE as intermediate
+Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
 Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
@@
 #scale_color_manual(values=colors_of_strains)
+
+#un-order these factors
+
+```
+
+```{r long.biv.plots, echo=FALSE}
 
@@
 
+Long_biv_mouse.avs_4MM$strain <- unclass(Long_biv_mouse.avs_4MM$strain)
+Long_biv_mouse.avs_4MM$strain <- as.factor(Long_biv_mouse.avs_4MM$strain)
+
+Long_biv_mouse.avs_4MM$sex <- unclass(Long_biv_mouse.avs_4MM$sex)
+Long_biv_mouse.avs_4MM$sex <- as.factor(Long_biv_mouse.avs_4MM$sex)
+
+Long_biv_mouse.avs_4MM$subsp <- unclass(Long_biv_mouse.avs_4MM$v)
+Long_biv_mouse.avs_4MM$subsp <- as.factor(Long_biv_mouse.avs_4MM$subsp)
+
 ```
@@
 2. Expected impact on sex comparisons, extimated effect size of the X  
-
 3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
@@
 #divide musc males into groups
+mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
+
 mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
@@
 
+```
+
+I used the mixed model with the lme() for the full Curated BivData and long bivdata
+
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+#### 1CO
+
+Insert the results for the MMs using the dependant variable 1CO
+
+```{r 1CO.MM, echo=FALSE}
+#I think I need to re-calq the mouse averages for just 1CO
+#mean.SC_1CO is mean 1CO
+
+SC.Length_1CO_MM <- lme(mean.SC_1CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(SC.Length_1CO_MM)$tTable, 6 ) )
+```
+
+#### 2CO
+
+```{r 2CO.MM, echo=FALSE, eval=FALSE}
+
+#remove MOLF
+mouse.avs_4MM.99 <- mouse.avs_4MM[mouse.avs_4MM$strain != "MOLF",]
+
+SC.Length_2CO_MM <- lme(mean.SC_2CO ~ subsp * sex, 
+                  random=list(strain=pdDiag(~sex),
+                        data= mouse.avs_4MM.99, 
+                                                                                       na.action=na.omit ) )
+#keep getting errors for building this model
+
+as.data.frame( round(  summary(SC.Length_2CO_MM)$tTable, 6 ) )
+
+table(mouse.avs_4MM$strain, mouse.avs_4MM$sex)
+
+```
+
+#### 3CO
+
+```{r 3CO.MM, echo=FALSE}
+
+SC.Length_3CO_MM <- lme(mean.SC_3CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),  na.action=na.omit  )
+#this model is also generating error...
+#don't know how to fix
+
+as.data.frame( round(  summary(SC.Length_3CO_MM)$tTable, 6 ) )
+```
+
+#### long.biv
+
+```{r long.biv.MM, echo=FALSE, eval=FALSE}
+
+SC.Length_long.biv_MM <- lme(mean_SC ~ subsp * sex, data= Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex),  na.action=na.omit  ) )
+
+#now there is an error after un-ordering them
+
+as.data.frame( round(  summary(SC.Length_long.biv_MM)$tTable, 6 ) )
+
 #model for long Bivs
-long_bivData_SC.Length_MM <- lme(mean_SC ~ subsp * sex, data= Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(long_bivData_SC.Length_MM)$tTable, 6 ) )
-
-```
-
-I used the mixed model with the lme() for the full Curated BivData and long bivdata
-
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r MM_2_SC.length_2_DELETE, echo=FALSE, eval=FALSE}
-#review this block!!
-
-#Long_biv_mouse.avs_4MM
-#as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )
-
-#did model isn't working
-SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-#anova(SC_length.lmer.MM)
-#thing1  <- summary(SC_length.lmer.MM)
-#thing2 <- drop1(SC_length.lmer.MM, test="Chisq")
-
-rand_SC.Length_MM <- exactRLRT(SC_length.lmer.MM)
-rand_SC.Length_MM
-
-#this pvalue for ran is 0.0027 -- the strain effect
-
-##long bivalents
-#long.bivs_SC_length.lmer.MM <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long_biv_mouse.avs_4MM, REML = TRUE)
-
-#anova(long.bivs_SC_length.lmer.MM) 
-#sex is larges explantion of variance
-#thing1  <- summary(long.bivs_SC_length.lmer.MM)
-#thing2 <- drop1(long.bivs_SC_length.lmer.MM, test="Chisq")
-
-#rand_SC.Length_MM <- exactRLRT(long.bivs_SC_length.lmer.MM)
-#rand_SC.Length_MM
-#strain no sig, but underpowered
-```
-
-#### 1CO
-
-Insert the results for the MMs using the dependant variable 1CO
-
-#### 2CO
-
-#### 3CO
-
-#### long.biv
+#long_bivData_SC.Length_MM <- lme(mean_SC ~ subsp * sex, data= #Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(long_bivData_SC.Length_MM)$tTable, 6 ) )
+```
 
@@
 
-```{r WSB, echo=FALSE}
+```{r WSB.SC, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-wsb.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                  # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
@@
                    mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
 
                    mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
 
                   mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
 
@@
 
-```{r LEW, echo=FALSE}
+```{r LEW.SC, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-LEW.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+LEW.SC.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 
-```{r G, echo=FALSE}
+```{r G.SC, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-G.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+G.SC.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 
-```{r PWD, echo=FALSE}
+```{r PWD.SC, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-PWD.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
+PWD.SC.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 
-```{r KAZ, echo=FALSE}
+```{r KAZ.SC, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-KAZ.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
+KAZ.SC.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 
-```{r SKIVE, echo=FALSE}
+```{r SKIVE.Sc, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-SKIVE.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+SKIVE.SC.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 
-```{r MSM, echo=FALSE}
+```{r MSM.SC, echo=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-MSM.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
+MSM.SC.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 
-```{r MOLF, eval=FALSE}
+```{r MOLF.SC, eval=FALSE}
 #make the dataset
@@
 #make a table of the means and n ect
-MOLF.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+MOLF.SC.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
                              mean_SC = mean(chromosomeLength),
@@
 summary(log.reg_SC.length_BIV_3CO)
-
 ```
@@
   ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-boxplots_first.IFD.plot_ABS
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
 #order is out of wack, but -- good for now
@@
   ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-boxplots_first.IFD.plot_PER
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
 
@@
 
-### Fixed Effects
+Still working on the best way to display the general IFD patterns.
+
+### Mixed Model Tests, Fixed Effects
 
@@
 
-lme_MM_IFD_ABS <- lme(mean_IFD.2CO_ABS ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+lme_MM_IFD_ABS <- lme(mean_IFD.2CO_ABS ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),na.action=na.omit  )
+
 as.data.frame( round(  summary(lme_MM_IFD_ABS)$tTable, 6 ) )
 #p
-
-lme_MM_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+#same missing values in obc
+
+lme_MM_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ), na.action=na.omit )
+
 as.data.frame( round(  summary(lme_MM_IFD_PER)$tTable, 6 ) )
+
+#remove PWD, SKIVE and MSM
+another.DF <- mouse.avs_4MM[mouse.avs_4MM$strain != "PWD",]
+another.DF <- another.DF[another.DF$strain != "MSM",]
+another.DF <- another.DF[another.DF$strain != "SKIVE",]
+
+lme_MM_another_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= another.DF, random=list(strain=pdDiag(~sex) ), na.action=na.omit )
+
+as.data.frame( round(  summary(lme_MM_another_IFD_PER)$tTable, 6 ) )
+
 ```
@@
 
-### Random Effects
+The table above should display the slightly unsual pattern, where the coefficients for the significant sex fixed effect are positive and neagtive in the raw and normalized values respectively.
+That is for the raw IFD values, females are signiciantly longer but the normalized IFD values, males are significantly longer.
+
+show a raw plot, and norm plot and a scatter plot
+
+#### Random Effects
 
@@
 rand_test_IFD.PER.MM
-
 ```
@@
 
-```{r WSB.IFD, echo=FALSE}
+these are example strains that show the 'unusual' results.
+show a raw plot, and norm plot and a scatter plot.
+Could the female bivalents with IFD under 30% be a different class of interference?
+(What does that mean?).
+I think the General pattern is that this rec pattern is not observed in males.
+That is there are rarely instances where 2 foci are positioned closer than 30% of the total SC length.
+
+```{r WSB.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
@@
 
-table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
+#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
 
@@
 
+WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+WSB.IFDnorm_scatter.plot
+#this plot is much messier than the MSM (PWD?)
+
 #try plotting the other IFD
@@
 
-                       mean_SC = mean(chromosomeLength),
-                       n_SC = length(Obj.ID),
+                       total.biv = length(Obj.ID),
 
@@
 
-#can I plot these stats?
-
 #run a t.test
@@
 
-```{r LEW.IFD, echo=FALSE}
+The range of normalized IFDs overlap closer in males and females in the WSB data.
+
+```{r LEW.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
 LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
 
 #plot the differences
-LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-LEW.ploty
-#remove the NAs
-
-wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_raw
-
-wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_PER
+#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+lew.ploty_IFD_raw <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_raw
+
+lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_PER
+
+LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+LEW.IFDnorm_scatter.plot
 
 #make a table of the means and n ect
-LEW.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
+lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+
+                  #     mean_SC = mean(chromosomeLength),
                        n_SC = length(Obj.ID),
 
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
-)
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
 #can I plot these stats
@@
 
-```{r G.IFD, echo=FALSE}
+The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
+
+```{r G.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
 G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
 
@@
                           geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
+
 G.ploty_IFD_raw
-
 #
@@
                           geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
 G.ploty_IFD_PER
 
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.IFDnorm_scatter.plot
+
 #make a table of the means and n ect
-G.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
+
+G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
 )
@@
 
-```{r PWD.IFD, echo=FALSE}
+```{r PWD.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
 PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
 
@@
                           geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
-PWD.ploty_IFD_raw
-
-PWD.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+
+PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
                           geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
-PWD.ploty_IFD_PER
+
+PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+PWD.IFDnorm_scatter.plot
 
@@
 PWD.table <-  ddply(PWD.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
 
@@
 
-```{r KAZ.IFD, echo=FALSE}
+For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
+
+```{r KAZ.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
 KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
 
@@
 
+KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+KAZ.IFDnorm_scatter.plot
+
 #make a table of the means and n ect
 KAZ.table <-  ddply(KAZ.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
 
@@
 
-```{r SKIVE.IFD, echo=FALSE}
+For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
+
+```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
 SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
 
@@
 
-#make a table of the means and n ect
+SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+SKIVE.IFDnorm_scatter.plot
+
 SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
 
@@
 
-```{r MSM.IFD, echo=FALSE}
+In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
+
+```{r MSM.IFD, echo=FALSE, warning=FALSE}
 #make the dataset
 MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
 
@@
 MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty
-#remove the NAs
 
@@
 
+MSM.IFDnorm_another.plot
+
 #this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
@@
 MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
-)
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
 #can I plot these stats
@@
 
-```{r MOLF.IFD, eval=FALSE}
+The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
+
+```{r MOLF.IFD, eval=FALSE, echo=FALSE}
 #make the dataset
@@
 MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum(which(hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum(which(hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum(which(hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum(which(hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
 
@@
 
+The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
+The other strains which have more overlap between males and females are the Dom strains and KAZ.
+
 #### IFD 3CO bivalents
@@
 
-### Heterochiasmy Prediction
+```{r CO3.comparisons, echo=FALSE}
+
+#take a look at the 3CO things
+
+```
+
+<!--
+ Heterochiasmy Prediction
 
@@
 The random strain effect was tested for raw and normalized IFD, neither are significant.
-
-### Male Polymorphism Prediction
+-->
+
+## Male Polymorphism Prediction
 
@@
 For these tests I am looking at the single IFD measure from 2CO bivalents.
+
+For the logistic regression, comparing High and Low groups (there needs to be 2 groups)
+I think I should compare 2 sets of logistic models, where SKIVE (intermediate gwRR), is in high and low.
 
@@
   ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,100))
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,100))
 biv.IFDs_by_group
+#+theme(legend.position="none")+
 
@@
 
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("IFD across bivalents")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
 CO.Pos.boxplots_1CO
@@
   geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_LEW_female
+#sis.co.ten.plot_LEW_female
 
@@
   geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_LEW_male
+#sis.co.ten.plot_LEW_male
 
@@
   geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_PWD_female
+#sis.co.ten.plot_PWD_female
 
@@
   geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_PWD_male
+#sis.co.ten.plot_PWD_male
 
@@
   geom_jitter()+facet_wrap(sex~strain)
-jitter_siscoten_PWDf
+#jitter_siscoten_PWDf
 
@@
   geom_jitter()+facet_wrap(sex~strain)
-jitter_siscoten_PWDm
+#jitter_siscoten_PWDm
 
@@
   geom_jitter()+facet_wrap(sex~strain)
-jitter_siscoten_WSBf
+#jitter_siscoten_WSBf
 
@@
   geom_jitter()+facet_wrap(sex~strain)
-jitter_siscoten_WSBm
+#jitter_siscoten_WSBm
 ```
@@
 To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
 

### 2020-01-30 doc/Bivalent_Analysis.Rmd
updating the MLH1 Mixed model results

@@
 
+#apply this before ploting, to display strains in desired order
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, category, mouse) %>%
+#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+# sort your dataframe, by the focal categories
+#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
 ```

### 2020-01-30 doc/Main_Report_Results.rmd
updating the MLH1 Mixed model results

@@
 
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
+
 #this MLH1 only goes up to batch16!
@@
 
+#apply this before ploting
 MLH1_data <- MLH1_data %>%
@@
 
-```{r rand, echo=FALSE, include=FALSE, eval=FALSE}
+```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
 #str(AP_mouse_table_w.Ages)
@@
     AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-```
-
-# Main Figures
+
+#why did I male just a female table?
+female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
+
+female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)
+
+```
+
+# A. Models for av MLH1 counts foci per cell
+
+ToDo: add the random effect tests!!
+
+These are the basic Mixed Models we used as a frame work to analyze all the MLH1 count data.
+
+$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)  + \varepsilon$$
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
+
+The mixed model lets us separate effects to see how they effect a mouse mean CO count. The subspecies effect is a proxy for **Divergence** and the random strain effect is a proxy for **Polymorphism**.
+
+Setting up the data set for using the mixed model (base data set, HQ data set)
+
+- only Mus musculus strains with sex matched observations
+
+- Unorder factors
+
+- male ages within good range (5 - 12? weeks)
+
+- Mol as a different subsp
+
+- ( Nested strain in subsp?)
+
+*The mouse average table used in this part excluded Q5 cells*
+
+*the MLH1 mean is from mean - adj_MLH1 (males have +1 CO to count for PAR)*
+
+```{r mixed.model.data.setup, echo=FALSE, warning=FALSE}
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+
+#What are the main DFs for this code chunk
+#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
+#DF.HetC.MixedModel   <- only Mmusculus
+#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
+
+#variance is still chr
+AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+
+#use the merged DF to filter by age 
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+#remove the extra merged cols
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+#DONT NEED CATEGORY
+
+#REMOVE EXTRA STRAINS
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+
+#SWITCH TO HQ.DF
+#mean_co.HQ  (this might remove some of the MOLF males)
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]  #this removes the MOLF female
+
+table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#afriad this is removing NA's
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
+
+table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#reformate Molosinus subsp
+#mean_co.HQ.mol
+DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
+        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))
+
+# For some reason I changed subsp to nums (wtf)
+DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
+        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))
+
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#remove blank rows (which are these?)
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
+#151 mice, no rows with nas for values
+
+#main df with variables coded as chr
+DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+
+#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+```
+
+This is the size of mice used within the Mixed model
+
+`r kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
+
+Can I indedate the average number of cells with go into each mouse.?
+
+FOR FINAL ANALYSIS, make sure that MOL is coded as a separate subsp (change all of the main model reports to).
+For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.
+
+subspecies levels: 2(3), Musc and Domestics and Molossinus
+
+strain levels: Don: 3, Musc: 4  Mol: 2
+
+sex: 2 female male
+
+## Mean CO Count
+
+todo, add the random effect tests, consider changing all to lmer() (those have better random effects)
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
+
+```{r HQ.MixedModel.1, echo=FALSE, digits = 2}
+#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+#mean_co.HQ  (this might remove some of the MOLF males)
+
+HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )
+
+ranef(HetC.MixedModel.HQ)
+#which is the test for random effects ?!
+
+```
+
+Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
+
+Above is the table of coefficients for MM.1.mol, molssinus coded as separate subspecies, lme(), random effects strain and strain specific sex effect.
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+```{r MixedModel.2, echo=FALSE}
+#reduced models
+Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+#summary(Reduced.strain.HQ)
+
+as.data.frame(round(summary(Reduced.strain.HQ)$tTable, 6))
+#add a aption to this code chunk
+#ranef(Reduced.strain.HQ)
+
+```
+
+Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
+This is model has sex and interaction coefficient is significant.
+
+Can the above model be re-written with lmer so that random effect can be tested?
+
+> MM3
+
+$$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
+
+```{r MixedModel.3, echo=FALSE}
+#strain = fixed
+lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+#library(xtable)
+
+#this table of results is broken -- might be too long -- lm
+#lm has different names
+as.data.frame(round( summary(lm.strain.fixed.full)$coefficients, 6))
+
+#summary(strain.fixed.full)
+#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
+
+#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are 
+```
+
+Above is the table of coefficients for MM.3, HQ data set, lm function, sex, strain and subsp are interacting fixed effects.
+
+Strain G is significant (p=0.0005),
+
+interaction terms of sex and strain for (G, MSM, PWD).
+
+How are these models different from the last one -- MM4 is missing subsp, so only sex * strain effects.
+
+> MM.4
+
+$$mouse .mean CO ~=~ sex * strain + \varepsilon$$
+
+```{r MixedModel.4, echo=FALSE}
+
+strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+#below table is broken
+as.data.frame( round( summary(strain.fixed)$coefficients, 6))
+
+#summary(strain.fixed)
+#maybe this WSB effec is just in the HQ data set?
+```
+
+Above is the table of coefficients for **MM.4**, HQ data set, lm function, just sex and strain, no subspecies.
+
+Significant factors are 
+strain G (p=0.0005), strainMSM and strainPWD (very slight),
+sex:G (p=0.001),
+sex:MSM, Sex:PWD (very slight p=.05, p=0.3),
+
+#### ANOVA, mean CO count full vs reduced
+
+Use Anova to compare full vs reduced models
+the reduced models will be those missing either, strain, sex, or subsp.
+
+```{r anova.full.v.reduced, echo=FALSE, eval=FALSE}
+
+#strain as fixed
+#CONTRAST models
+#newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
+
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
+
+full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
+
+#base vs strain
+fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co) #<.0001
+
+#show scientific vs normal
+#vv <- fullVstrain$`p-value`
+#round(vv[2], 8)
+#format(vv[2], scientific=F)
+
+#strain vs no random full.reduce
+fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+#fullreduce vs strain
+#can't distinguish which random effects are more significant -- both tests are significant
+
+fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
+
+#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
+full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
+
+full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
+#this comparison not useful when there are different fixed effects
+
+#full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ) # < .0001
+
+#strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ) #NS
+
+#anova tests
+#full vs strain
+full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
+#full vs non
+full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
+#strain vs non
+strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
+
+anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
+
+#base vs mol model (this only makes sense if these are nested models)
+anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
+```
+
+Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
+Models with random effects more significant than nulls (both just strain and no random effects).
+The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
+
+```{r Base.strain.fixed, echo=FALSE}
+#make mixed model with strain as fixed, since there are interesting results
+#Consider including these results from all FIXED MODEL
+
+HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
+#this produces several significant factors
+
+HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
+
+#summary(HetC.MM.strain.fixed2)
+#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
+
+#summary(HetC.MM.strain.fixed)
+#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
+#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
+#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
+#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
+#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
+#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
+#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
+#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
+
+#male effect lowers the mean from intercept
+#while strain effects 
+```
+
+ Summary of mean COs
+
+MM.2 has significant coefficients, sex and sex*subsp interaction.
+
+MM.2.Mol, the sex and interaction term (subsp*sex) are significant.
+
+MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.
+
+## Variance in CO count
+
+DECIDE WHICH RESULTS TO DISPLAY.
+
+in addition to the mean gwRR, the variance is also an important metric to describe
+
+Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
+
+RW's main qtl paper used the within animal variance, 
+"We mapped QTL for
+MLH1 count variance within individuals, applying an approach
+to deconvolve the mean-variance relationship (see
+Materials and Methods)."
+
+above method remove / accounts for mean-variance correlation -
+
+The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
+
+QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
+
+@article{wang2017genetics,
+  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
+  author={Wang, Richard J and Payseur, Bret A},
+  journal={Genetics},
+  volume={206},
+  number={4},
+  pages={1841--1852},
+  year={2017},
+  publisher={Genetics Soc America}
+}
+
+references for phenotypic variance connected to robustness to environment, increase in developmental consistancy
+Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
+
+phenotypic variance leading to adaptation - 
+(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
+
+Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
+
+meiotic Checkpoints for ensuring consistancy (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
+
+#### MLH1 variance Figures
+
+Not sure why I have this section -- think this should be into a supplemental doc or something.
+
+```{r variance.boxplots, echo=FALSE, eval=FALSE}
+#how did I make this 
+#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
+#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
+#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
+#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
+
+#plot of variance
+var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
+#this plot shows variance as points/single values
+
+#plot by sex / seperately
+# HQ data DF.HetC.MixedModel.HQ
+
+var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+var.male <- var.male   + facet_wrap( ~ strain, scales="free")
+
+cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
+ ggtitle("Male cV levels") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
+
+cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
+
+```
+
+```{r variance.hist, echo=FALSE, eval=FALSE}
+
+#make series for total dataset?
+Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("Histograms of within mouse variance, all mice")
+Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+Var.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("Histograms of within mouse variance, 10 cell min")
+Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
+Var.Hist
+
+Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
+Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
+Var.Hist.15cells
+
+```
+
+A general pattern is that the female data has higher within mouse variance (both var and cV).
+Some strains more than others.
+I think the dissection date needs to be tested as a predictor factor.
+It might be tricky to add the dissection date metric.
+Test if dissection date is more significant predictor than strain (for females)
+Also test how the number of 0CO bivalents per cell affects mouse with variance.
+
+(LEW females with top 3 variance are all from different dissection dates)
+
+I don't have enough mice with >15 cells for a complete idea of the distributions, but comparing single mice with enough good quality cells show be
+
+```{r cv.plots, echo=FALSE, eval=FALSE}
+
+cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("cV of within mouse variance, all mice")
+cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+cV.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("cV of within mouse variance, 10 cell min")
+cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
+cV.Hist
+
+```
+
+### Modeling Variance
+
+(use the same models are mean_co, )
+
+1.	HetC.MixedModel.HQ <- lme(mean_co ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+2.	Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+3.	lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+
+4.	strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+$$mouse .cV CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$
+
+```{r cV.model.setup, echo=FALSE}
+#modeling within mouse variance
+
+DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
+
+#HQ var.CO
+HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+#summary(HetC.MixedModel.HQ.var_co)
+
+as.data.frame( round(summary(HetC.MixedModel.HQ.var_co)$tTable, 6))
+
+#ranef(HetC.MixedModel.HQ.var_co)
+```
+
+$$mouse .cV CO ~=~ subsp * sex + rand(1|strain)+ \varepsilon$$
+
+```{r cV.models, echo=FALSE}
+HQjust.strain.var <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+as.data.frame( round(summary(HQjust.strain.var)$tTable, 6))
+
+#summary(HQjust.strain.var_co)
+```
+
+$$mouse .var CO ~=~ subsp * sex  + \varepsilon$$
+```{r base.var.mol, echo=FALSE, eval=FALSE}
+HetC.MixedModel.mol.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(HetC.MixedModel.mol.var_co)$tTable, 6))
+
+```
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain * sex) + \varepsilon$$
+```{r var.model.1, echo=FALSE, eval=FALSE}
+#main cV.CO
+HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(HetC.MixedModel.varco)$tTable, 6))
+
+#summary(HetC.MixedModel.cVco)
+#ranef(HetC.MixedModel.cVco)
+```
+
+```{r var.contrast, echo=FALSE, eval=FALSE}
+#contrast models
+# review this logic again
+
+full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
+
+full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
+
+strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
+#for var, p = 0.0219
+```
+
+#### Summary variance
+
+The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.
+
+When Molossinus is coded as a separate subspecies, then interaction terms are no longer significant.  (subsp *sex )  interaction?
+
+<!--
+For later analysis.  Not sure what these are good for.
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
+
+# check if the main mixed model results hold up when all cells are pooled
+
+#pooled.cells
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+#consider changing to ordered = FALSE, for the DFs used for MM
+DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
+
+DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
+
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+
+```
+
+- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
+This means there is 0 variance added to mean_CO to mean due to being MSM.
+Does this mean the model is very good at predicting MSM mean_CO.
+The fixed strain - sex effects are for musc are large.
+This will likely change with MOLF (male molf are low, so smaller fixed effect and )
+
+- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
+In the first model MSM has largest random effect, since it's so far from WSB female.
+
+-adding MOLF also doesn't have significant factors
+
+changed all to unordered factors.
+all the variables are coded as ordered factors.
+When the variables are kept factors, and unordered the names appear! yay!
+
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
+-->
+
+## Summary of Mixed Model Analyses
+
+**mouse average MLH1 count**
+
+Multiple models were tested to test for potential evo models -- for the mouse average CO counts.  (there were a mix of results -- and nuanced)
+
+First full Mixed model -- no significant fixed factors --(the random effect might be significant).
+
+The second model were strain was (nested and random), the fixed effects, sex and interaction of (subsp and sex), are much more significant.
+The coefficients indicate, males in general have 1 less in average, and the musc and molf subsp have ~3 more on average.
+
+**Variance effect**
+
+** Do these patterns hold across Quality bins? **
+
+# Figure 1
+
+Main figure, MLH1 strain avs on top, chrm proportion bar plot on bottom
 
@@
 ```{r table.for.plot, echo=FALSE}
+#why do I add in a different table?
 
@@
 
-*cast female data came from another publication, add this ref (Kohler?)
-
-The bottom lines and subspecies points don't exactly match up.
-
-**add in cell images like the other draft figures.**
-
-# A. Models for mean and variance of MLH1 foci per cell
-
-These are the basic Mixed Models we used as a frame work to analyze all the MLH1 count data.
-
-$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)  + \varepsilon$$
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
-
-The mixed model lets us separate effects to see how they effect a mouse mean CO count. The subspecies effect is a proxy for **Divergence** and the random strain effect is a proxy for **Polymorphism**.
-
-Initial results indicate that male MLH1 counts have less variance within subspecies but more between subspecies compared to female data.
-
-```{r mixed.models.setup, echo=FALSE, warning=FALSE}
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-
-#notes for testing random effects
-
-#variance is still chr
-AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
-
-#use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
-                            subsp.y,subsp.x,
-                    category.y, category.x,
-                        strain.x,strain.y, mouse.1  )  )
-
-#why did I male just a female table?
-female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-
-female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-                        #quality?
-)
-
-#REMOVE EXTRA STRAINS
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
-
-#remove blank rows (which are these?)
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
-#151 mice, no rows with nas for values
-
-#main df with variables coded as chr
-DF.HetC.MixedModel <- DF.HetC.MixedModel
-DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
-DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
-DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
-```
-
-```{r HQ.dataset, echo=FALSE}
-
-#also make a HQ dataset, tighter window for ages,
-
-#mean_co.HQ  (this might remove some of the MOLF males)
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-```
-
-```{r Mol.dataset, echo=FALSE}
-#mean_co.HQ.mol
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
-        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))
-
-DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
-        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
-
-DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
-
-```
-
-Setting up the data set for using the mixed model (base data set, HQ data set)
-
-- only Mus musculus strains with sex matched observations
-
-- Unorder factors
-
-- male ages within good range (5 - 12? weeks)
-
-Currently I ran through 3 data sets: 1) base, 2) HQ and 3) Molf (from HQ)
-
-`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`
-
-FOR FINAL ANALYSIS, make sure that MOL is coded as a separate subsp (change all of the main model reports to).
-For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.
-
-subspecies levels: 2(3), Musc and Domestics (Mousiness)
-
-strain levels: Don: 3, Musc: 4  (Mol: 2)
-
-sex: 2 female male
-
-## Mean CO Count
-
-todo, add the random effect tests, consider changing all to lmer() (those have better random effects)
-
-$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
-
-```{r HQ.MixedModel.1, echo=FALSE, digits = 2}
-#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-#mean_co.HQ  (this might remove some of the MOLF males)
-
-HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )
-
-#ranef(HetC.MixedModel.HQ)
-#which is the test for random effects
-
-```
-
-Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.
-
-$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
-
-```{r MM.1.Mol, echo=FALSE}
-
-HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(   summary(HetC.MixedModel.HQ.mol)$tTable, 6) )
-
-```
-
-Above is the table of coefficients for MM.1.mol, molssinus coded as separate subspecies, lme(), random effects strain and strain specific sex effect.
-
-$$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-```{r MixedModel.2, echo=FALSE}
-#reduced models
-Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-#summary(Reduced.strain.HQ)
-
-as.data.frame(round(summary(Reduced.strain.HQ)$tTable, 6))
-#add a aption to this code chunk
-#ranef(Reduced.strain.HQ)
-
-```
-
-Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
-This is model has sex and interaction coefficient is significant.
-
-> MM.2.Mol
-
-```{r MixedModel.2.mol, echo=FALSE}
-#reduced models
-Reduced.strain.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random = ~1|strain)
-#summary(Reduced.strain.HQ)
-
-as.data.frame(round(summary(Reduced.strain.HQ.mol)$tTable, 6))
-#add a aption to this code chunk
-#ranef(Reduced.strain.HQ.mol)
-
-```
-
-The model for MM.2.Mol has similar results as the above model, with just 2 subsps
-This model has significant coefficients for sex and the 2 interaction effects (sex:musc, sex:mol)
-
-> MM2.2
-
-$$mouse .mean CO ~=~ subsp * sex  + \varepsilon$$
-
-```{r MixedModel.2.2, echo=FALSE}
-Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
-
-as.data.frame( round(summary(Reduced.all.HQ)$tTable, 6))
-#summary(Reduced.all.HQ)
-#ranef(Reduced.all.HQ)
-```
-
-Above is the table of coefficients for MM.2.2, HQ data set, gls(), No random effects.  
-I think this gls has sex and interaction term as significant.
-
-> MM3
-
-$$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
-
-```{r MixedModel.3, echo=FALSE}
-#strain = fixed
-strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-#library(xtable)
-
-#this table of results is broken -- might be too long -- lm
-#lm has different names
-as.data.frame(round( summary(strain.fixed.full)$coefficients, 6))
-
-#summary(strain.fixed.full)
-#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
-
-#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are 
-```
-
-Above is the table of coefficients for MM.3, HQ data set, lm function, sex, strain and subsp are interacting fixed effects.
-
-Strain G is significant (p=0.0005),
-
-interaction terms of sex and strain for (G, MSM, PWD).
-
-How are these models different from the last one -- MM4 is missing subsp, so only sex * strain effects.
-
-> MM.4
-
-$$mouse .mean CO ~=~ sex * strain + \varepsilon$$
-
-```{r MixedModel.4, echo=FALSE}
-
-strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-#below table is broken
-as.data.frame( round( summary(strain.fixed)$coefficients, 6))
-
-#summary(strain.fixed)
-#maybe this WSB effec is just in the HQ data set?
-```
-
-Above is the table of coefficients for **MM.4**, HQ data set, lm function, just sex and strain, no subspecies.
-
-Significant factors are 
-strain G (p=0.0005), strainMSM and strainPWD (very slight),
-sex:G (p=0.001),
-sex:MSM, Sex:PWD (very slight p=.05, p=0.3),
-
-#### ANOVA, mean CO count full vs reduced
-
-Use Anova to compare full vs reduced models
-the reduced models will be those missing either, strain, sex, or subsp.
-
-```{r anova.full.v.reduced, echo=FALSE, eval=FALSE}
-
-#strain as fixed
-#CONTRAST models
-#newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
-
-HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
-
-full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
-
-#base vs strain
-fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co) #<.0001
-
-#show scientific vs normal
-#vv <- fullVstrain$`p-value`
-#round(vv[2], 8)
-#format(vv[2], scientific=F)
-
-#strain vs no random full.reduce
-fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
-
-strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
-
-#fullreduce vs strain
-#can't distinguish which random effects are more significant -- both tests are significant
-
-fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
-
-#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
-full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
-
-full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
-#this comparison not useful when there are different fixed effects
-
-full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ) # < .0001
-
-strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ) #NS
-```
-
-```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
-#base model with molossinus level ... why am I making molossinus model?
-#mean.co
-HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )
-
-#mreducing the model (reducing the random effects)
-Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
-
-Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
-
-#anova tests
-#full vs strain
-full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
-#full vs non
-full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
-#strain vs non
-strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
-
-anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
-
-#base vs mol model (this only makes sense if these are nested models)
-anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
-```
-
-Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
-Models with random effects more significant than nulls (both just strain and no random effects).
-The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
-
-```{r Base.strain.fixed, echo=FALSE}
-
-#make mixed model with strain as fixed, since there are interesting results
-
-HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
-#this produces several significant factors
-
-HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
-
-#summary(HetC.MM.strain.fixed2)
-#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
-
-#summary(HetC.MM.strain.fixed)
-#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
-#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
-#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
-#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
-#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
-#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
-#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
-#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
-
-#male effect lowers the mean from intercept
-#while strain effects 
-```
-
- Summary of mean COs
-
-MM.2 has significant coefficients, sex and sex*subsp interaction.
-
-MM.2.Mol, the sex and interaction term (subsp*sex) are significant.
-
-MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.
-
-## Variance in CO count
-
-in addition to the mean gwRR, the variance is also an important metric to describe
-
-Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
-
-RW's main qtl paper used the within animal variance, 
-"We mapped QTL for
-MLH1 count variance within individuals, applying an approach
-to deconvolve the mean-variance relationship (see
-Materials and Methods)."
-
-above method remove / accounts for mean-variance correlation -
-
-The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
-
-QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
-
-@article{wang2017genetics,
-  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
-  author={Wang, Richard J and Payseur, Bret A},
-  journal={Genetics},
-  volume={206},
-  number={4},
-  pages={1841--1852},
-  year={2017},
-  publisher={Genetics Soc America}
-}
-
-references for phenotypic variance connected to robustness to environment, increase in developmental consistancy
-Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
-
-phenotypic variance leading to adaptation - 
-(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
-
-Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
-
-meiotic Checkpoints for ensuring consistancy (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-
-### Figures
-
-```{r variance.boxplots, echo=FALSE, eval=FALSE}
-#how did I make this 
-#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
-#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
-#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
-#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
-
-#plot of variance
-var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
-var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
-#this plot shows variance as points/single values
-
-#plot by sex / seperately
-# HQ data DF.HetC.MixedModel.HQ
-
-var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-var.male <- var.male   + facet_wrap( ~ strain, scales="free")
-
-cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male cV levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
-
-```
-
-```{r variance.hist, echo=FALSE}
-
-#make series for total dataset?
-Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("Histograms of within mouse variance, all mice")
-Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-Var.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("Histograms of within mouse variance, 10 cell min")
-Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
-Var.Hist
-
-Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
-Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
-Var.Hist.15cells
-
-```
-
-A general pattern is that the female data has higher within mouse variance (both var and cV).
-Some strains more than others.
-I think the dissection date needs to be tested as a predictor factor.
-It might be tricky to add the dissection date metric.
-Test if dissection date is more significant predictor than strain (for females)
-Also test how the number of 0CO bivalents per cell affects mouse with variance.
-
-(LEW females with top 3 variance are all from different dissection dates)
-
-I don't have enough mice with >15 cells for a complete idea of the distributions, but comparing single mice with enough good quality cells show be
-
-```{r cv.plots, echo=FALSE}
-
-cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("cV of within mouse variance, all mice")
-cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-cV.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("cV of within mouse variance, 10 cell min")
-cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
-cV.Hist
-
-```
-
-### modeling variance
-
-$$mouse .cV CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$
-
-```{r cV.model.setup, echo=FALSE}
-#modeling within mouse variance
-
-DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
-
-#HQ var.CO
-HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-#summary(HetC.MixedModel.HQ.var_co)
-
-as.data.frame( round(summary(HetC.MixedModel.HQ.var_co)$tTable, 6))
-
-#ranef(HetC.MixedModel.HQ.var_co)
-```
-
-lme function, Cv modeled. Only sex is significant.
-
-```{r cv.mod1.mol, echo=FALSE}
-
-HetC.MixedModel.HQ_mol.cV_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(HetC.MixedModel.HQ_mol.cV_co)$tTable, 6))
-
-```
-
-$$mouse .cV CO ~=~ subsp * sex + rand(1|strain)+ \varepsilon$$
-
-```{r cV.models, echo=FALSE}
-HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-
-as.data.frame( round(summary(Reduced.all.HQ)$tTable, 6))
-
-#summary(HQjust.strain.var_co)
-```
-
-Sex and sex*subsp interaction effect are significant.
-
-```{r cV.mod.mol, echo=FALSE}
-
-HQ.mol.cV_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random = ~1|strain)
-
-as.data.frame( round(summary(HQ.mol.cV_co)$tTable, 6))
-
-```
-
-Sex is significant, when mol is a seperate subspecies interaction is no longer significant.
-
-$$mouse .cV CO ~=~ subsp * sex + \varepsilon$$
-```{r more.cV.CO, echo=FALSE}
-HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
-as.data.frame( round(summary(HQfull.Reduced.cv_co)$tTable, 6))
-#summary(HQfull.Reduced.cv_co)
-
-#contrast models
-full2strain.CV <- anova(HetC.MixedModel.HQ.var_co, HQjust.strain.var_co)#
-
-full2noRandom.CV <- anova(HetC.MixedModel.HQ.var_co, HQfull.Reduced.cv_co)#
-
-strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
-```
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain*sex) + \varepsilon$$
-
-```{r Base.Variance.MM, echo=FALSE, eval=FALSE}
-
-#there's 1 mouse with just 1 cell remove
-DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
-DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
-
-#main var.CO
-HetC.MixedModel.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(HetC.MixedModel.var_co)$tTable, 6))
-
-#summary(HetC.MixedModel.var_co)
-#ranef(HetC.MixedModel.var_co)
-```
-Above error says missing values in object -- I don't understand. Mol version of data frame works...
-
-Base DF, lme function, random sex specific strain effects
-
-$$mouse .var CO ~=~ subsp * sex  + \varepsilon$$
-```{r base.var.mol, echo=FALSE}
-HetC.MixedModel.mol.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(HetC.MixedModel.mol.var_co)$tTable, 6))
-
-```
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain * sex) + \varepsilon$$
-```{r var.model.1, echo=FALSE, eval=FALSE}
-#main cV.CO
-HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(HetC.MixedModel.varco)$tTable, 6))
-
-#summary(HetC.MixedModel.cVco)
-#ranef(HetC.MixedModel.cVco)
-```
-
-$$mouse .var CO ~=~ subsp * sex + rand(1|strain) + \varepsilon$$
-
-```{r var.model.2, echo=FALSE, eval=FALSE}
-just.strain.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)
-
-as.data.frame( round(summary(just.strain.var_co)$tTable, 6))
-
-#summary(just.strain.cV_co)
-```
-
-lme model above
-
-$$mouse .var CO ~=~ subsp * sex + \varepsilon$$
-
-```{r var.mod.3, echo=FALSE, eval=FALSE}
-
-full.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel)
-#summary(full.Reduced.cv_co)
-
-as.data.frame( (summary(full.Reduced.cv_co)$tTable))
-
-```
-
-gls model model above
-
-```{r var.contrast, echo=FALSE, eval=FALSE}
-#contrast models
-full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
-
-full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
-
-strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
-#for var, p = 0.0219
-```
-
-#### Summary variance
-
-The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.
-
-When Molossinus is coded as a separate subspecies, then interaction terms are no longer significant.
-
-#### Pooled cells
-
-For later analysis.
-
-```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
-
-# check if the main mixed model results hold up when all cells are pooled
-
-#pooled.cells
-DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
-
-#remove PERC, MOLF, CZECH,
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
-
-#sex only has 2 levels and this breaks...
-model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.pooled)# indeces for this object: [1]-structure
-ranef(model.pooled)
-
-model.pooled$coefficients
-
-#pooled cells,  DF.HetC.MixedModel.pooled
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
-
-#consider changing to ordered = FALSE, for the DFs used for MM
-DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
-
-DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
-
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-
-```
-
-- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
-This means there is 0 variance added to mean_CO to mean due to being MSM.
-Does this mean the model is very good at predicting MSM mean_CO.
-The fixed strain - sex effects are for musc are large.
-This will likely change with MOLF (male molf are low, so smaller fixed effect and )
-
-- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
-In the first model MSM has largest random effect, since it's so far from WSB female.
-
--adding MOLF also doesn't have significant factors
-
-changed all to unordered factors.
-all the variables are coded as ordered factors.
-When the variables are kept factors, and unordered the names appear! yay!
-
-Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
-
-# Rapid male specific evolution
-
-list out the points which support rapid male specific evolution for MLH1 rates.
-
-## Proportion of chrm classes
+#### Proportion of chrm classes
+
+make a smaller propoertion chrm block --
 
@@
 
-```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}
-
-#this is a second mode of
-
-#bivalent data might be too big..
-#load(file="~./MLH1repo/data/CleanBivData.RData")
-#Curates_BivData -- hand.foci.count!
-
-#For each mouse I need to count up
-#the number of bivalent classes are not counted up
-
-chrm.propotion.category <- ddply(.data=Curated_BivData, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(fileName)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = length(Obj.ID),
-
-                 n0CO = length(which(hand.foci.count == 0)),
-                  n1CO = length(which(hand.foci.count == 1)),
-                  n2CO = length(which(hand.foci.count == 2)),
-                  n3CO = length(which(hand.foci.count == 3)),
-                  #n4CO = sum(n4CO),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv )
-                # per4CO= sum(n4CO)/ tot_biv  )
-
-Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
-  "per0CO", "per1CO", "per2CO", "per3CO"), 
-  id=c("category", "tot_biv") )
-colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
-
-# I think I need to make 2 melt tables -- for count and proportions
-Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
-  id=c("category", "tot_biv") )
-colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
-
-Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
-colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
-
-Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
-
-Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
-
-Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
-ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
-      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
-        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
-        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
-        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
-        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
-
-Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
-                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
-#no mol yet
-
-Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
-
-#need to make list of positions
-
-chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
-
-chrm_prop.BivData_plot
-
-```
+*cast female data came from another publication, add this ref (Kohler?)
+
+The bottom lines and subspecies points don't exactly match up.
+
+**add in cell images like the other draft figures.**
+
+# Rapid male specific evolution
+
+list out the points which support rapid male specific evolution for MLH1 rates.
 
@@
 #write this table
-write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
-            row.names = TRUE)
+#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
+#            row.names = TRUE)
 
@@
 
+# Trash / Extra
+
+<!--
+
+```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}
+
+#this chunk migth need to be deleted // I think everything is already done within the BivData
+#this is a second mode of
+
+#bivalent data might be too big..
+#load(file="~./MLH1repo/data/CleanBivData.RData")
+#Curates_BivData -- hand.foci.count!
+
+#For each mouse I need to count up
+#the number of bivalent classes are not counted up
+
+chrm.propotion.category <- ddply(.data=Curated_BivData, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(fileName)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = length(Obj.ID),
+
+                 n0CO = length(which(hand.foci.count == 0)),
+                  n1CO = length(which(hand.foci.count == 1)),
+                  n2CO = length(which(hand.foci.count == 2)),
+                  n3CO = length(which(hand.foci.count == 3)),
+                  #n4CO = sum(n4CO),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv )
+                # per4CO= sum(n4CO)/ tot_biv  )
+
+Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
+  "per0CO", "per1CO", "per2CO", "per3CO"), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
+
+# I think I need to make 2 melt tables -- for count and proportions
+Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
+
+Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
+colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
+
+Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
+
+Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
+
+Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
+ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
+      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
+        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
+        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
+        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
+        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
+
+Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
+                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
+#no mol yet
+
+Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
+
+#need to make list of positions
+
+chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
+
+chrm_prop.BivData_plot
+
+```
+
+-->
+
 # References

### 2020-01-30 doc/Main_Report_Results.rmd
organizing Mixed Model results

@@
 
-Details of data set cleaning ect is in SetupData.rmd
+Details of data set cleaning etc is in SetupData.rmd
 Effects from different factors are reviewed in Effects.Rmd
@@
 
-table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
 
@@
 
-table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
 
@@
 
-This is the size of mice used within the Mixed model
+This is the break down of mice used within the Mixed model
 
@@
 
-Can I indedate the average number of cells with go into each mouse.?
-
-FOR FINAL ANALYSIS, make sure that MOL is coded as a separate subsp (change all of the main model reports to).
-For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.
+<!--Can I indicate the average number of cells with go into each mouse.? -->
 
@@
 
+*mouse averages exclude Q5 and male counts are adjusted +1 for PAR)*
+
 ## Mean CO Count
@@
 
+For now I will focus on 4 different models run mouse averages on MLH1 counts.
+
+1. $$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
+
+2. $$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+3. $$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
+
+4. $$mouse .mean CO ~=~ sex * strain + \varepsilon$$
+
+> Model 1
+
 $$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
+
+1.	HetC.MixedModel.HQ <- lme(mean_co ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
 
@@
 
-ranef(HetC.MixedModel.HQ)
+#ranef(HetC.MixedModel.HQ)
 #which is the test for random effects ?!
-
 ```
@@
 
-$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
-
-Above is the table of coefficients for MM.1.mol, molssinus coded as separate subspecies, lme(), random effects strain and strain specific sex effect.
+> Model 2.
 
@@
 #ranef(Reduced.strain.HQ)
-
 ```
@@
 
-> MM3
+> Model 3
 
@@
 
-> MM.4
+> Model 4
 
@@
 
-references for phenotypic variance connected to robustness to environment, increase in developmental consistancy
+references for phenotypic variance connected to robustness to environment, increase in developmental consistency
 Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
@@
 
-meiotic Checkpoints for ensuring consistancy (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
+meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
 
@@
 
-make a smaller propoertion chrm block --
+make a smaller proportion chrm block --
 
@@
 
-insert male total sc results here
+insert male total SC results here
 

### 2020-01-30 doc/TotalSC.R

@@ -1,0 +1,208 @@
+
+
+
+#ToDo
+#rewrite in Rmd -- to better display notes and findings
+#Check the segmentation
+#re-run python scrtip, (it runs over all anon image files)
+#is it worth croping /cleaning up these images?
+
+#load main MLH1 files to merge MLH1 cell data with the sc data 
+setwd("~./")
+
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#batch18
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent 
+
+#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
+SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
+
+#format all factors to numbers
+SC.skel$bin_size  <- as.character(SC.skel$bin_size )
+SC.skel$skel_size  <- as.character(SC.skel$skel_size)
+
+SC.skel$bin_size  <- as.numeric(SC.skel$bin_size )
+SC.skel$skel_size <- as.numeric(SC.skel$skel_size)
+
+SC.skel$rand_name <- as.character(SC.skel$rand_name)
+
+#check for duplicatations of the random name
+original_DF$Random.Name <- as.character(original_DF$Random.Name)
+duplies <- original_DF[duplicated(original_DF$Random.Name),]
+#3 duplicate rows in original DF -- because Random name was blank -- how will these be matched to total.SC?
+
+#MERGEING MLH! with SKEL
+MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
+OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
+
+
+#clean up this DataFrame REMOVE 'X'
+MLH1.merge.Skel <- MLH1.merge.Skel[ !grepl("X", MLH1.merge.Skel$Batch) , ]
+MLH1.merge.Skel <- MLH1.merge.Skel[ !grepl("x", MLH1.merge.Skel$X) , ]
[…]

### 2020-01-31 doc/Bivalent_Analysis.Rmd
small bit of work

@@
 
-load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
+#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
 #fyi loading mulitple RData files doesn't earase ther files -- but it can overwrite files named the same thing

### 2020-01-31 doc/Main_Report_Results.rmd
small bit of work

@@
 
+setwd("~./MLH1repo/")
+
+load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
+load(file="/data/CleanBivData_1.28.20.RData")
+
 #setwd("~./MLH1repo/doc/")
-setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
-#setwd("C:/Users/April/Desktop/MLH1repo")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
+#setwd("C:/Users/April/Documents/MLH1repo")
 #load(file="data/MLH1/MLH1_data_setup.RData")
@@
 #load main data file
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
-
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
+#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
 
@@
 
-$$mouse .cV CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$
-
-```{r cV.model.setup, echo=FALSE}
+$$mouse .var CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$
+
+```{r M1.var, echo=FALSE}
 #modeling within mouse variance
-
-DF.HetC.MixedModel.HQ$var <-  as.numeric(DF.HetC.MixedModel.HQ$var)
+#use  DF.HetC.MixedModel.HQ
 
 #HQ var.CO
-HetC.MixedModel.HQ.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+M1.var_co <- lme( var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
 
@@
 
-as.data.frame( round(summary(HetC.MixedModel.HQ.var_co)$tTable, 6))
+as.data.frame( round(summary(M1.var_co)$tTable, 6))
 
 #ranef(HetC.MixedModel.HQ.var_co)
+```
+
+```{r M1.Q12_table, echo=FALSE}
+#use the Q1Q2 calq'd table
+
+#Q12_mouse_table_w.Ages$subsp <- ifelse(grepl(1, Q12_mouse_table_w.Ages$subsp), "Dom", 
+#        ifelse(grepl(2, Q12_mouse_table_w.Ages$subsp), "Musc", Q12_mouse_table_w.Ages$subsp))
+
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | #DF.HetC.MixedModel.HQ$mouse==""), ]
+
+Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
+
+Q12_mouse_table_w.Ages$subsp <- ifelse(grepl("MSM", Q12_mouse_table_w.Ages$strain), "Mol", 
+        ifelse(grepl("MOLF", Q12_mouse_table_w.Ages$strain), "Mol", NA))
+
+#Q12_mouse_table_w.Ages
+
+M1.Q12_var_co <- lme( var ~ subsp * sex, data=Q12_mouse_table_w.Ages, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(M1.Q12_var_co)$tTable, 6))
+
 ```

### 2020-01-31 src/SetupData.Rmd
small bit of work

@@
 
+```{r Q1Q2.mouse.table, echo=FALSE}
+
+#Make a mouse level table
+Q12_mouse_table <- ddply(MLH1_data[MLH1_data$quality < 3,], c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)
+
+#REMOVE mice with fewer than 3 cells
+Q12_mouse_table <- Q12_mouse_table[Q12_mouse_table$Ncells > 3,]
+
+Q12_mouse_table <- add_strain(Q12_mouse_table)
+Q12_mouse_table <- add_subsp(Q12_mouse_table)
+Q12_mouse_table <- add_sex(Q12_mouse_table)
+Q12_mouse_table <- add_category(Q12_mouse_table)
+
+#use things from prv block
+#mice.frm.meta 
+#meta.data
+#mouse.table_w.Ages
+
+#MERGE HERE
+Q12_mouse_table_w.Ages <- merge(Q12_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+#what are the differences between these two
+
+#remove the NAs
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[!(is.na(Q12_mouse_table_w.Ages$category.x) | Q12_mouse_table_w.Ages$category.x==""), ]
+
+#CLEAN up and format for mixed models
+Q12_mouse_table_w.Ages <- subset(Q12_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+Q12_mouse_table_w.Ages <- add_strain(Q12_mouse_table_w.Ages)
+Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
+Q12_mouse_table_w.Ages <- add_sex(Q12_mouse_table_w.Ages)
+Q12_mouse_table_w.Ages <- add_category(Q12_mouse_table_w.Ages)
+
+Q12_mouse_table_w.Ages <- add_species(Q12_mouse_table_w.Ages)
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$species == "M.musculus",]
+
+#REMOVE EXTRA STRAINS
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
+#remove musc strains
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]
+
+#REMOVE WRONG AGES
+#afriad this is removing NA's
+Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[( (Q12_mouse_table_w.Ages$age.weeks < 25) &  (Q12_mouse_table_w.Ages$age.days != 13) ) | is.na(Q12_mouse_table_w.Ages$age.weeks), ] #keeps mice without ages, and removes juvinilles
+
+#MOL = subsp
+Q12_mouse_table_w.Ages$subsp <- ifelse(grepl("MSM", Q12_mouse_table_w.Ages$strain), "Mol", 
+        ifelse(grepl("MOLF", Q12_mouse_table_w.Ages$strain), "Mol", Q12_mouse_table_w.Ages$subsp))
+
+#unclass order factors
+#main df with variables coded as chr
+Q12_mouse_table_w.Ages$subsp <- factor(Q12_mouse_table_w.Ages$subsp, ordered = FALSE)
+Q12_mouse_table_w.Ages$sex <- factor(Q12_mouse_table_w.Ages$sex, ordered = FALSE)
+Q12_mouse_table_w.Ages$strain <- factor(Q12_mouse_table_w.Ages$strain, ordered = FALSE)
+Q12_mouse_table_w.Ages$species <- factor(Q12_mouse_table_w.Ages$species, ordered = FALSE)
+
+```
+
 ```{r strain.table, echo=FALSE}

### 2020-02-03 src/SetupData.Rmd
removed juvi male mouse observations, re-ran MLH1

@@
 ---
+
+to Do
+
+- make sure to remove some of the juvi mice!! from the AP_mouse table
 
@@
                   )
+MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]
 
-MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% exclude_mice_list, ]
 ```
@@
 #merge with DOB
-mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = "mouse", by.x = "Var1", all = TRUE)
+
+#why are there 2
+#mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = #"mouse", by.x = "Var1", all = TRUE)
+
 AP_mouse_table_w.Ages <- merge(AP_mouse_table.HQ, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
@@
 
+juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
+
+juvi.mice_DF <- juvi.mice[!(is.na(juvi.mice$mouse) | juvi.mice$mouse==""), ]
+
+#remove bad ages
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[( (AP_mouse_table_w.Ages$age.weeks < 25) &  (AP_mouse_table_w.Ages$age.days != 13) ) | is.na(AP_mouse_table_w.Ages$age.weeks), ]
+
+#remove juvi mice
+MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi.mice_DF$mouse, ]
+
 #remove the NAs
 AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$category.x) | AP_mouse_table_w.Ages$category.x==""), ]
+
 ```
@@
 ```{r Q1Q2.mouse.table, echo=FALSE}
-
 #Make a mouse level table
@@
 
+```{r Q123_mouse.table, echo=FALSE}
+
+Q123_mouse_table <- ddply(MLH1_data[MLH1_data$quality < 4,], c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)
+
+#REMOVE mice with fewer than 3 cells
+Q123_mouse_table <- Q123_mouse_table[Q123_mouse_table$Ncells > 3,]
+
+Q123_mouse_table <- add_strain(Q123_mouse_table)
+Q123_mouse_table <- add_subsp(Q123_mouse_table)
+Q123_mouse_table <- add_sex(Q123_mouse_table)
+Q123_mouse_table <- add_category(Q123_mouse_table)
+
+```
+
 ```{r strain.table, echo=FALSE}
@@
 
-male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[male_mouse_table_w.Ages$age.weeks < 20,]
+#male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[male_mouse_table_w.Ages$age.weeks < 20,]
+
+male_mouse_table_w.good.ages <- male_mouse_table_w.Ages[( (male_mouse_table_w.Ages$age.weeks < 25) &  (male_mouse_table_w.Ages$age.days != 13) ) | is.na(male_mouse_table_w.Ages$age.weeks), ]
+
+#remove juvi mice!!
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ]
+
 #droplevels on DF above
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_1.10.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_2.2.20.RData")
 
@@
 
--age (for males)
--enough cells
--both male and female
+- age (for males)
+- enough cells
+- both male and female

### 2020-02-03 doc/Bivalent_Analysis.Rmd
too much work from home

@@
 
-```{r clean.biv.data4.MM, echo=FALSE}
+```{r biv.data_table.4MM.HetC, echo=FALSE}
 
@@
 
+#I'd like to merge the additional strains used for the male tests
+
 #the mouse averages will be used
@@
                         Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
                         mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
@@
 
+#code intp chr
+
 #assign high and low
@@
 
+```{r biv.data_table.4MM.Male, echo=FALSE}
+#remove female, 
+#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
+
+#Curated_BivData_male
+
+male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+male.av_4MM <- add_sex(male.av_4MM)
+male.av_4MM <- add_strain(male.av_4MM)
+male.av_4MM <- add_subsp(male.av_4MM)
+male.av_4MM <- add_category(male.av_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
+male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
+male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
+
+```
+
 ## Predictions, Heterochiasmy
@@
 ## Predictions, Male Polymorphism (High vs Low Rec strains)
+
+Rerun all these predictions for MSM > MOLF as well
 
@@
 
+APRIL, consider moving these strain chunks to the bivData set up
+
 ```{r WSB.SC, echo=FALSE}
@@
 
+#### male strain comparisons
+
+```{r male.sc.length.strain, echo=FALSE}
+
+Curated_BivData <- add_species(Curated_BivData)
+#isolate all musc
+Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+#make table
+SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+
+```
+
+```{r SC.length.male_tests, echo=FALSE}
+
+## run the tests to check following expectations
+
+#1.Dom strains not different from ech other
+#2. PWD and SKIVE are different from the other strains
+#3.MSM and MOLF are different from each other
+
+#decide between lm or glm..?
+#mouse avs, males
+
+male.Dom.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Dom"), ]
+#the subsp labels get messed up, unless they are designated chr first
+male.Musc.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == 2), ]
+
+mouse.avs_4MM$subsp <- as.character(mouse.avs_4MM$subsp)
+mouse.avs_4MM$subsp <- ifelse(grepl("MSM", mouse.avs_4MM$strain), "Mol", 
+        ifelse(grepl("MOLF", mouse.avs_4MM$strain), "Mol", mouse.avs_4MM$subsp))
+male.Mol.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Mol"), ]
+
+Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
+
+Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  )
+Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
+#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
+#total sc sig, 1CO sig, 2CO sig
+
+Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
+#0.000369 ***
+
+Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
+#0.000816 ***
+
+Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
+#0.0148 *
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
+#making plots above show that mouse average SC lengths are lower could there be microscope issue?
+
+#make sure to remove some of the juvi mice!! from the AP_mouse table
+
+show.Dom.male.CO <- ggplot(
+AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") & (AP_mouse_table_w.Ages$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
+
+Dom.male_CO.aov <- aov(mean_co ~ strain, data = AP_mouse_table.HQ[(AP_mouse_table.HQ$sex == "male") & (AP_mouse_table.HQ$subsp == "Dom"),])
+#P not sig for CO mean
+
+#seems like there are more mice in the MLH1 data set than bivData
+#
+
+#WSB all dissected and imaged in '15
+#G imaged 15 - 16
+#LEW imaged 15 to 17
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "1CO SC Length", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "2CO SC Length", xlab = "")
+```
+
+```{r SC.length.pairwise.T, echo=FALSE}
+#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
+
+male.Musc.mouse.av <- male.av_4MM[(male.av_4MM$subsp == "Musc")&(male.av_4MM$sex == "male"),]
+
+Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
+#PWD and MSM are most diff from SKIVE?
+
+Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
+#the pooled SC value tests are sig
+
+#the mol comparisons are different due to adj p values
+
+Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
+#no sig (only 3) -- does include CZECCH
+
+#non-sig, with mol pooled with musc
+
+Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
+
+Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
+
+Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
+#non of these sig with the mulitple adjust
+
+Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
+
+pairwise.t.test(Ozone, Month, p.adj = "bonf")
+
+```
+
+If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
+
+Ideas for the above 3 tests,
+
+1. no difference in DOM
+- no sig logistic regression test
+- glm, strain effect?
+
 ### Summary
@@
 
-### Male Polymorphism Prediction
-
+(merge this with the code above)
 Prediction, High rec males have longer SC.
@@
 
+```{r male.IFD_strain, echo=FALSE}
+
+#make table
+Musc_male_IFD_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 2,], c("strain"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_IFD = mean(IFD1_ABS, na.rm=TRUE),
+                        mean_IFD.nrm = mean(IFD1_PER,na.rm=TRUE)
+                         )
+
+        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+```
+
 ```{r IFD.log.reg, echo=FALSE}
@@
 I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
+
+```{r male.Nrm1CO_strain, echo=FALSE}
+
+#isolate all musc
+Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+#make table
+Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
+
+                         )
+
+        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+```
 

### 2020-02-03 doc/Main_Report_Results.rmd
too much work from home

@@
 
-setwd("~./MLH1repo/")
-
-load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
-load(file="/data/CleanBivData_1.28.20.RData")
+#setwd("~./MLH1repo/")
+
+load(file = "C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
+
+#2.2.20
+
+load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
+
+#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
 
@@
 
-```{r HQ.MixedModel.1, echo=FALSE, digits = 2}
+```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2}
 #table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
@@
 MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.
+
+## Musc Males
+
+test a model for fixed effects of strains within musc. The purpose is to distinguihsi musc strains from low 'Dom' level strains.
+
+```{r Musc.male.Model.data.setup, echo=FALSE}
+#nested mouse in strain
+
+#remove the extra merged cols
+Musc.male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+Musc.male_mouse.av <- add_strain(Musc.male_mouse.av)
+Musc.male_mouse.av <- add_sex(Musc.male_mouse.av)
+Musc.male_mouse.av <- add_subsp(Musc.male_mouse.av)
+Musc.male_mouse.av <- add_species(Musc.male_mouse.av)
+
+Musc.male_mouse.av$subsp <- as.character(Musc.male_mouse.av$subsp)
+
+Musc.male_mouse.av$subsp[Musc.male_mouse.av$strain== "MSM"] <- 'Mol'
+Musc.male_mouse.av$subsp[Musc.male_mouse.av$strain== "MOLF"] <- "Mol"
+
+#use original for getting the other strains
+
+Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$species == "M.musculus",]
+Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$subsp == "Musc",]
+Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$sex == "male",]
+
+#set the factors
+Musc.male_mouse.av$strain <- factor(Musc.male_mouse.av$strain, ordered = FALSE)
+Musc.male_mouse.av$mouse <- factor(Musc.male_mouse.av$mouse, ordered = FALSE)
+
+#compare the the higher quality plots
+#Q1Q2
+
+#Q12_mouse_table_w.Ages
+
+Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
+Q12_mouse_table_w.Ages <- add_species(Q12_mouse_table_w.Ages)
+
+Q12_mouse_table_w.Ages$subsp <- as.character(Q12_mouse_table_w.Ages$subsp)
+
+Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MSM"] <- 'Mol'
+Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MOLF"] <- "Mol"
+
+Musc.male.Q12 <- Q12_mouse_table_w.Ages[(Q12_mouse_table_w.Ages$subsp == "Musc")&(Q12_mouse_table_w.Ages$sex == "male"),]
+
+Musc.male.Q12$strain <- factor(Musc.male.Q12$strain, ordered = FALSE)
+Musc.male.Q12$mouse <- factor(Musc.male.Q12$mouse, ordered = FALSE)
+
+#Q123  Q123_mouse_table
+
+Q123_mouse_table <- add_subsp(Q123_mouse_table)
+Q123_mouse_table <- add_species(Q123_mouse_table)
+
+Q123_mouse_table$subsp <- as.character(Q123_mouse_table$subsp)
+
+Q123_mouse_table$subsp[Q123_mouse_table$strain== "MSM"] <- 'Mol'
+Q123_mouse_table$subsp[Q123_mouse_table$strain== "MOLF"] <- "Mol"
+
+Musc.male.Q123 <- Q123_mouse_table[(Q123_mouse_table$subsp == "Musc")&(Q123_mouse_table$sex == "male"),]
+
+Musc.male.Q123$strain <- factor(Musc.male.Q123$strain, ordered = FALSE)
+Musc.male.Q123$mouse <- factor(Musc.male.Q123$mouse, ordered = FALSE)
+
+```
+
+```{r Musc.male.show.plots, echo=FALSE}
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +  
+musc.male.plot
+
+Q123.musc.male.plot <- ggplot(Musc.male.Q123,
+                             aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q123 Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Q123.musc.male.plot
+#THE q12,data set only 
+#Q123 has 5 strain rep (no TOM), there are still some mouse outliers
+
+```
+
+This is the table and plot of new Musc strains
+
+`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
+
+Given the plot, make note of the potential outlier mice.  The strains have unequal variance
+
+KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
+
+PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
+
+```{r M1_Musc.male, echo=FALSE}
+
+#mouse.av ~ mouse | strain
+
+#not sure the right syntax for nested factors
+#Musc.male.M1 <- lm(mean_co ~ (mouse|strain), data=Musc.male_mouse.av )
+
+Musc.male.M1 <- lm(mean_co ~ strain, data=Musc.male_mouse.av )
+
+as.data.frame(  round(summary(Musc.male.M1)$coefficients, 4) )
+
+#all of the strain have a significant effect agaisnt the group mean / intercept
+#looking at the intercept -- the PWD observations are pushing up the average
+
+#musc male Q123  Q123_mouse_table
+
+Musc.male.Q123.M1 <- lm(mean_co ~ strain, data=Musc.male.Q123 )
+
+as.data.frame(  round(summary(Musc.male.Q123.M1)$coefficients, 4) )
+
+#mostly all strains, (but SKIVE) are significant -- similar to the top model
+
+#shoul I compare with subsampling
+
+```
+
+```{r Musc.male.M2}
+
+#mouse.av ~ strain
+
+```
 
@@
 #ranef(HetC.MixedModel.HQ.var_co)
+
+#CV
+M1.cv_co <- lme( cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+#summary(HetC.MixedModel.HQ.var_co)
+as.data.frame( round(summary(M1.cv_co)$tTable, 6))
 ```
@@
 
-Q12_mouse_table_w.Ages$subsp <- ifelse(grepl("MSM", Q12_mouse_table_w.Ages$strain), "Mol", 
-        ifelse(grepl("MOLF", Q12_mouse_table_w.Ages$strain), "Mol", NA))
+#need to change to chr first
+
+Q12_mouse_table_w.Ages$subsp <- as.character(Q12_mouse_table_w.Ages$subsp)
+
+Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MSM"] <- 'Mol'
+
+Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MOLF"] <- "Mol"
 
@@
 
-M1.Q12_var_co <- lme( var ~ subsp * sex, data=Q12_mouse_table_w.Ages, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(M1.Q12_var_co)$tTable, 6))
+M1.Q12_var <- lme( var ~ subsp * sex, data=Q12_mouse_table_w.Ages, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(M1.Q12_var)$tTable, 6))
+
+M1.Q12_CV <- lme( cV ~ subsp * sex, data=Q12_mouse_table_w.Ages, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(summary(M1.Q12_CV)$tTable, 6))
 
@@
 
-```{r cV.models, echo=FALSE}
-HQjust.strain.var <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-
-as.data.frame( round(summary(HQjust.strain.var)$tTable, 6))
+```{r M2.var, echo=FALSE}
+#this model must be the same as M! -- since all the results are the same
+M2.var <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+as.data.frame( round(summary(M2.var)$tTable, 6))
 
 #summary(HQjust.strain.var_co)
+
+M2.CV <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+
+as.data.frame( round(summary(M2.CV)$tTable, 6))
+
+```
+
+```{r M2.var.Q12, echo=FALSE}
+M2.Q12_var_co <- lme(var ~ subsp * sex, data=Q12_mouse_table_w.Ages, random = ~1|strain)
+
+as.data.frame( round(summary(M2.Q12_var_co)$tTable, 6))
+
+M2.Q12_CV <- lme(cV ~ subsp * sex, data=Q12_mouse_table_w.Ages, random = ~1|strain)
+
+as.data.frame( round(summary(M2.Q12_CV)$tTable, 6))
 ```
@@
 $$mouse .var CO ~=~ subsp * sex  + \varepsilon$$
-```{r base.var.mol, echo=FALSE, eval=FALSE}
-HetC.MixedModel.mol.var_co <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(HetC.MixedModel.mol.var_co)$tTable, 6))
-
+```{r M3.var, echo=FALSE, eval=FALSE}
+M3.var <- lm(var ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+as.data.frame( round(summary(M3.var)$tTable, 6))
+
+M3.cv <- lm(cV ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+as.data.frame( round(summary(M3.cv)$tTable, 6))
+#these models won't work
+```
+
+```{r M3VAR.Q12, echo=FALSE, eval=FALSE}
+#ugh, don't know how to get these to work 
+M3.Q12_var <- glm(var ~ subsp * sex * strain, data=Q12_mouse_table_w.Ages)
+
+as.data.frame(summary(M3.Q12_var)$tTable, 6)
+#results don't show
+
+M3.Q12_cv <- lm(cv ~ subsp * sex * strain, data=Q12_mouse_table_w.Ages)
+
+as.data.frame( round(summary(M3.Q12_cv)$tTable, 6))
+```
+
+```{r M4.var, echo=FALSE, eval=FALSE}
+M4.var <- lm(var ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+as.data.frame( round(summary(M4.var)$tTable, 6))
+
+M4.cv <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+as.data.frame( round(summary(M4.cv)$tTable, 6))
+```
+
+```{r M4.var_Q12, echo=FALSE, eval=FALSE}
+
+M4.var_Q12 <- lm(var ~ sex * strain, data=Q12_mouse_table_w.Ages)
+as.data.frame( round(summary(M4.var_Q12)$tTable, 6))
+
+#these results not print / working
+M4.CV_Q12 <- lm(cV ~ sex * strain, data=Q12_mouse_table_w.Ages)
+as.data.frame( round(summary(M4.CV_Q12)$tTable, 6))
 ```
@@
 $$mouse .var CO ~=~ subsp * sex + rand(strain * sex) + \varepsilon$$
-```{r var.model.1, echo=FALSE, eval=FALSE}
-#main cV.CO
-HetC.MixedModel.varco <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(HetC.MixedModel.varco)$tTable, 6))
-
-#summary(HetC.MixedModel.cVco)
-#ranef(HetC.MixedModel.cVco)
-```
 
 ```{r var.contrast, echo=FALSE, eval=FALSE}
-#contrast models
+#contrast models -- what's the points for this block
 # review this logic again
-
 full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
@@
 #### Summary variance
+
+I couldn't get many of the models to return results.
 
@@
 
-chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
 
@@
 
-```{r DMC1.setup, echo=FALSE}
+```{r DMC1.setup, echo=FALSE, eval=FALSE}
 #load DMC1 csv
 DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+
 DMC1.DF$fileName <- DMC1.DF$Original.Name
@@
 
-```{r normality, echo=FALSE, warning=FALSE}
+```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
 #histograms for each
@@
 
-```{r DMC1.plot, echo=FALSE, warning=FALSE}
+```{r DMC1.plot, echo=FALSE, warning=FALSE, eval=FALSE}
 #think of better color scheme
@@
 
-```{r DMC1_MLH1.cor, echo=FALSE}
+```{r DMC1_MLH1.cor, echo=FALSE, eval=FALSE}
 #make category averages from AP_mouse_table.HQ
@@
 
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+The correlation with MLH1 and leptotene cells is r MLH1.DMC1.L_cor.
+
+The correlation with MLH1 and zygotene cells is r MLH1.DMC1.Z_cor.
+
+# Single Bivalent Patterns
+
+<!--
 
@@
 
-# Blank
-
-```{r trackcurBivData, echo=FALSE}
-#list / tally of current curated Bivdata
-Cur_biv = read.csv("~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", header = TRUE)
-
-colnames(Cur_biv) <- c('mouse', 'blank','0','1')
-
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages[ -c(9:12, 14, 18, 22:24) ]
-
-AP_mouse_table_w.Ages_HQ <- add_sex(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_strain(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_category(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_subsp(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_species(AP_mouse_table_w.Ages_HQ)
-
-#only M.musculus
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$species == "M.musculus",]
-#REMOVE EXTRA STRAINS
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$subsp != "Cast",]
-#remove musc strains
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "AST",]
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "CZECH",]
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "PERC",]
-#156 mice (from 193)
-
-#some mice are missing 13oct14
-
-#match the MM dataset
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$Ncells > 9,]
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$age.weeks < 26,]
-#126 mice
-
-#AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[!(is.na(AP_mouse_table_w.Ages_HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mice missing from curated list
-mice.within.BivData <- AP_mouse_table_w.Ages_HQ[ AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#54 mice in (14 mice have enteries but 0 curated biv)
-
-#(14 mice need at least some curation
-
-cur.data.mice <- Cur_biv[ Cur_biv$mouse %in% AP_mouse_table_w.Ages_HQ$mouse, ]
-
-mice.missing.bivData <- AP_mouse_table_w.Ages_HQ[ ! AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#72 (includes mice not in MM)
-#72 mice that I need to upload/run the algorithm on their images
-
-#mice missing SC pass == '1'
-colnames(Cur_biv) <- c("mouse","blank","SC.fail", "SC.pass")
-NeedsCur <- Cur_biv[Cur_biv$SC.fail == 0,]
-
-NeedsCur <- add_strain(NeedsCur)
-write.table(NeedsCur, "~./MLH1repo/data/needs_cur.csv", sep=",",
-            row.names = TRUE)
-
-#44% of mice are missing curation
-
-#write this table
-#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
-#            row.names = TRUE)
-
-mice.missing.bivData <- mice.missing.bivData[!(is.na(mice.missing.bivData$mouse) | mice.missing.bivData$mouse==""), ]
-write.table(mice.missing.bivData, "~./MLH1repo/data/MissingMice_13.1.20.csv", sep=",",
-            row.names = TRUE)
-
-```
-
-```{r barlett.test, echo=FALSE}
-
-#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
-bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
-
-```
-
-Evidence for non-equal variance across strains for zygotene cells.
-
-Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
-
-From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova <- anova(mod.L) #0.0002701 ***
-z.anova <- anova(mod.Z) #NS
-
-l.anova
-z.anova
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#L vs Z (all species)
-#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
-#p-value = 1.032e-05
-#not a valid interpertation
-
-#high vs low group (L)
-ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
+-->
 
@@
 ```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
+
+#REMOVE This BLOCK?
+
 #### Distributions by mouse
@@
 
+# Blank
+
+```{r trackcurBivData, echo=FALSE, eval=FALSE}
+#REMOVE THIS BLOCK??
+
+#list / tally of current curated Bivdata
+
+Cur_biv = read.csv("~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", header = TRUE)
+
+colnames(Cur_biv) <- c('mouse', 'blank','0','1')
+
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages[ -c(9:12, 14, 18, 22:24) ]
+
+AP_mouse_table_w.Ages_HQ <- add_sex(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_strain(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_category(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_subsp(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_species(AP_mouse_table_w.Ages_HQ)
+
+#only M.musculus
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$species == "M.musculus",]
+#REMOVE EXTRA STRAINS
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$subsp != "Cast",]
+#remove musc strains
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "AST",]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "CZECH",]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "PERC",]
+#156 mice (from 193)
+
+#some mice are missing 13oct14
+
+#match the MM dataset
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$Ncells > 9,]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$age.weeks < 26,]
+#126 mice
+
+#AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[!(is.na(AP_mouse_table_w.Ages_HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mice missing from curated list
+mice.within.BivData <- AP_mouse_table_w.Ages_HQ[ AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#54 mice in (14 mice have enteries but 0 curated biv)
+
+#(14 mice need at least some curation
+
+cur.data.mice <- Cur_biv[ Cur_biv$mouse %in% AP_mouse_table_w.Ages_HQ$mouse, ]
+
+mice.missing.bivData <- AP_mouse_table_w.Ages_HQ[ ! AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#72 (includes mice not in MM)
+#72 mice that I need to upload/run the algorithm on their images
+
+#mice missing SC pass == '1'
+colnames(Cur_biv) <- c("mouse","blank","SC.fail", "SC.pass")
+NeedsCur <- Cur_biv[Cur_biv$SC.fail == 0,]
+
+NeedsCur <- add_strain(NeedsCur)
+write.table(NeedsCur, "~./MLH1repo/data/needs_cur.csv", sep=",",
+            row.names = TRUE)
+
+#44% of mice are missing curation
+
+#write this table
+#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
+#            row.names = TRUE)
+
+mice.missing.bivData <- mice.missing.bivData[!(is.na(mice.missing.bivData$mouse) | mice.missing.bivData$mouse==""), ]
+write.table(mice.missing.bivData, "~./MLH1repo/data/MissingMice_13.1.20.csv", sep=",",
+            row.names = TRUE)
+
+```
+
+```{r barlett.test, echo=FALSE, eval=FALSE}
+
+#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
+bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
+
+```
+
+Evidence for non-equal variance across strains for zygotene cells.
+
+Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
+
+From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE, eval=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#L vs Z (all species)
+#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
+#p-value = 1.032e-05
+#not a valid interpertation
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = r ttest.HighLow.L$p.value) while the Z cells, there is not a significant difference across the high and low groups (value = r ttest.HighLow.Z$p.value).
+
 # Trash / Extra

### 2020-02-03 src/BivData_Setup.Rmd
added todo note
notes on IFD and sis-co-ten patterns
small bit of work
too much work from home

@@
 
-test work from home ()
+test work from home (),
+
+- remove BivData! (these are all Peromyscus!!)
+
+calq the nromalized.siscoten, double check the negative values; 13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14, 27aug15_12jun15_WSB_m2_sp1_18_rev_18, 21nov17_16nov17_MSM_f3_sp1_3_rev_1
+
+foci.pass = 2  5aug17_14jul17_MSM_f1_sp1_13_rev_3
+
+hand foci =0 but there is a foci position
+
+too high hand foci count
+
+declare the Strain specific BivData_DFs  (for convinice)
 
@@
 Curated_BivData$strain <- droplevels(Curated_BivData$strain)
+
+```
+
+```{r no.Foci.Pos, echo=FALSE}
+
+just.SC.length <- Curated_BivData[is.na(Curated_BivData$hand.foci.count),]#650 lines
+
+#table(just.SC.length$category)
+#decide what to do with these lines
+#how will they be represented in Mat Methods?
 
@@
 
+```{r Q1Q2.mouse.table, echo=FALSE}
+
+```
+
 analysis using the mouse level table -- to justify that using mouse averages is a good idea.

### 2020-02-03 doc/Bivalent_Analysis.Rmd
moved some DF setup from analysis to set up script

@@
 
-```{r WSB.SC, echo=FALSE}
+(adding all of these blocks to -- the setup script); strain.Bivalents.DF
+
+```{r WSB.SC, echo=FALSE, eval=FALSE}
 #make the dataset
@@
 wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
-
-```{r LEW.SC, echo=FALSE}
-#make the dataset
-LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-LEW.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-LEW.SC.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-
-```
-
-```{r G.SC, echo=FALSE}
-#make the dataset
-G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-G.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-G.SC.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-
-```
-
-```{r PWD.SC, echo=FALSE}
-#make the dataset
-PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-PWD.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-PWD.SC.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-```
-
-```{r KAZ.SC, echo=FALSE}
-#make the dataset
-KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-KAZ.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-KAZ.SC.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-
-```
-
-```{r SKIVE.Sc, echo=FALSE}
-#make the dataset
-SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-SKIVE.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-SKIVE.SC.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-```
-
-```{r MSM.SC, echo=FALSE}
-#make the dataset
-MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-MSM.SC.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-```
-
-```{r MOLF.SC, eval=FALSE}
-#make the dataset
-MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MOLF.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-MOLF.SC.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
 

### 2020-02-03 src/BivData_Setup.Rmd
moved some DF setup from analysis to set up script

@@
 
+## Strain specific data sets
+
+Compile dataframes for both sexes of each strain.
+
+```{r WSB.SC, echo=FALSE}
+#make the dataset
+WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
+#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
+table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
+wsb.ploty
+
+#make a table of the means and n ect
+wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+```
+
+```{r LEW.SC, echo=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+LEW.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+LEW.SC.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+
+```
+
+```{r G.SC, echo=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+G.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+G.SC.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+```
+
+```{r PWD.SC, echo=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+PWD.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+PWD.SC.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+```
+
+```{r KAZ.SC, echo=FALSE}
+#make the dataset
+KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+KAZ.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+KAZ.SC.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+
+```
+
+```{r SKIVE.Sc, echo=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+SKIVE.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+SKIVE.SC.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+```
+
+```{r MSM.SC, echo=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MSM.SC.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+```
+
+```{r MOLF.SC, eval=FALSE}
+#make the dataset
+MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MOLF.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MOLF.SC.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
 # Cell Level table

### 2020-02-03 doc/Bivalent_Analysis.Rmd
starting to shift initializing files to setup script

@@
 #un-order these factors
-
 ```
@@
 
-```{r long.chrm.dataset, echo=FALSE}
-
-#1. how many biv's are from cells with >15 bivalents
-over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
-BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]
-
-#check for blank obj.id
-BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")
-
-#check for dup
-#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
-#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
-dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
-
-#remove duplicate obj.ids
-BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
-BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]
-
-#make sure obj.id and fileName are charcter for str stuff
-BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
-BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)
-
-#make sure 
-missing_from_cell_table <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% Biv_Data_mean_table$fileName, ]#there are 75, objs 3 cells
-#these are bivalents which don't appear the in cell level table, which can break the loop
-BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
-
-#final cleaned BivData table, 2428
-
-for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){
-
-  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
-  new.egg <- paste0(newfile.str, collapse = "_")
-
-  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
-  #the above can produe mulitple Q4
-
-  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
-  #these are just using the last index,
-
-  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]
-
-  BivData_from_15biv_cells$Q4.FN[i] <- the.FN
-   # print(c(i, the.FN, the.Q4))
-
-}
-
-BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= BivData_from_15biv_cells$cell.Q4), "yes", "no" )
-
-real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]
-#table(real.long.bivData$fileName)
-#!!! omg it finally worked!~!!#OMFG
-
-```
-
 The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
@@
 - Try estimating which 'loose' bivalent observations might be within the long class of bivalents.
-
-```{r mouse.av.long.bivData, echo=FALSE}
-#real.long.bivData
-
-#calq the mouse averages for
-
-Long_biv_mouse.avs_4MM <- ddply(real.long.bivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-                       #IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-
-)
-
-Long_biv_mouse.avs_4MM <- add_sex(Long_biv_mouse.avs_4MM)
-Long_biv_mouse.avs_4MM <- add_strain(Long_biv_mouse.avs_4MM)
-Long_biv_mouse.avs_4MM <- add_subsp(Long_biv_mouse.avs_4MM)
-
-Long_biv_mouse.avs_4MM$strain <- unclass(Long_biv_mouse.avs_4MM$strain)
-Long_biv_mouse.avs_4MM$strain <- as.factor(Long_biv_mouse.avs_4MM$strain)
-
-Long_biv_mouse.avs_4MM$sex <- unclass(Long_biv_mouse.avs_4MM$sex)
-Long_biv_mouse.avs_4MM$sex <- as.factor(Long_biv_mouse.avs_4MM$sex)
-
-Long_biv_mouse.avs_4MM$subsp <- unclass(Long_biv_mouse.avs_4MM$v)
-Long_biv_mouse.avs_4MM$subsp <- as.factor(Long_biv_mouse.avs_4MM$subsp)
-
-```
 
@@
 
+<!--
 $$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
@@
 
+-->
+
 $$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
-#### 1CO
-
-Insert the results for the MMs using the dependant variable 1CO
-
-```{r 1CO.MM, echo=FALSE}
+#### Mixed Model 1
+
+Insert the results for the MMs using the dependant variable 1CO - use the Mixed model (lmer), for all of the meterics (pooled SC, 1CO 2CO ect)
+
+```{r M1_1CO, echo=FALSE}
+
+#run Mixed models
+
+#all pooled SC lengths
+SCLength_all_.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=mouse.avs_4MM) 
+#summary(Reduced.strain.HQ)
+
+drop1(SCLength_all_.MM.M1, test="Chisq") # p = 0.0002045 ***
+
+exactRLRT(SCLength_all_.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0
+as.data.frame(round(summary(SCLength_all_.MM.M1)$tTable, 6))
+
+#1CO
+
+#2CO
+
 #I think I need to re-calq the mouse averages for just 1CO
@@
 
+```{r strain.SClength.t.test, echo=FALSE}
+#t.tests for all the strain specific data frames
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
 #### male strain comparisons
-
-```{r male.sc.length.strain, echo=FALSE}
-
-Curated_BivData <- add_species(Curated_BivData)
-#isolate all musc
-Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-#make table
-SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-
-```
 

### 2020-02-03 src/BivData_Setup.Rmd
starting to shift initializing files to setup script

@@
 
-test work from home (),
-
-- remove BivData! (these are all Peromyscus!!)
+- remove BivData! (these are all Peromyscus!!) (not sure where this DF comes from)
 
@@
 
-for final dataset, drop levels of strains not used CZECG, AST, TOM, CAST,
+ToDo
+
+- Go back and clean up the original excel data
+
+- clean the bottom, fill in missing data
+
+- check the order of foci positions for 5aug17_14jul17_LEW_f1_sp1_11_rev_blob4.
+
+edit / clean up the lines with bad hand.foci.count (61)
+
+ToDo:
+
+1.clean up the non-numbered curated values (for the og df)
+
+2.Goal of this doc is to make DFs that are ready for the analysis.
+
+3.remove the 2 main sources of duplicate lines; 6mar16_PWD_m1, 1mar15_PWD_m1
+
+4. decide on the right number of passing curated bivalents per category I should settle for
+Justify why the current number is enough -- or subsample -- to random equivilent sized data sets
+
+5. make sure that the saved Rdata has all the right cols and tables for the next analyses
+
+6. consider integrating manual hand / whole cell measures (check for DUPs ect.)
 
@@
 
+new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
+#make OG copy
+new.Curate.FULL.org <- new.Curate.FULL
+new.Curate.FULL.org <- add_category(new.Curate.FULL.org)
+
+#write files for basic stats to csv files
+#OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
+#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
+
+OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)
+
+#write the curated file
+#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_21.1.20.csv", sep=",",
+#            row.names = TRUE)
+#write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_21.1.20.csv", sep=",",
+#            row.names = TRUE)
+```
+
+```{r split.up.DF, echo=FALSE}
+
+#new.Curate.FULL to Curated_BivData
+
+#SC PASS
+new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#10480 #9819, #9769
+
+#fix these duplicated ones
+#remove duplicates
+Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151
+
+Curated_BivData <- add_mouse(Curated_BivData)
+Curated_BivData <- add_category(Curated_BivData)
+Curated_BivData <- add_strain(Curated_BivData)
+Curated_BivData <- add_sex(Curated_BivData)
+Curated_BivData <- add_subsp(Curated_BivData)
+Curated_BivData <- add_species(Curated_BivData)
+
+Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
+```
+
+```{r no.Foci.Pos, echo=FALSE}
+
+just.SC.length <- Curated_BivData[is.na(Curated_BivData$hand.foci.count),]#650 lines
+
+#table(just.SC.length$category)
+#decide what to do with these lines
+#how will they be represented in Mat Methods?
+```
+
+```{r input.check,echo=FALSE}
+
+#check for mistakes that might have been made when inputing / curation process
+
+#hand foci > 4
+#foci pos > SC lengths
+#negative PER.foci.pos
+
+#it hink most (ALL) of these came from bug in my function (ugh)
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#negative.IFD <- Curated_BivData[which(Curated_BivData$IFD1_ABS < 0),]
+#caught 2
+#15mar16_16jan16_G_f1_sp1_4_rev
+#27mar16_6mar16_G_m3_sp1_12_rev
+
+#neg Nrm F1
+#negative.IFD2 <- Curated_BivData[which(Curated_BivData$PER_Foci_1 < 0),]
+#0 really?
+```
+
+Main files from this plot:
+
+  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
+
+  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
+
+  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  4. MELT_ABS: make melt version of foci positions for abs measuers
+
+  5. CELL and
+
+  6. mouse table
+
+  7. **Strain DFs** for Het models
+
+  8. **male Subsp DFs** for Polymorphism models
+
+```{r CLEAN.UP.1, echo=FALSE, include=FALSE}
+
+#check for duplicates, and track them down
+anyDuplicated(new.Curate.FULL$Obj.ID)  #5335, #3467
+
+dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
+dup.curated <- add_mouse(dup.curated)
+
+#table(dup.curated$mouse)
+#1mar15_PWD_m1
+
+#which lines are duplicated
+#lots of PWD_female duplicated -- probably how I put together the curated files
+
+#new.Curate.FULL <- new.Curate.FULL.org[(new.Curate.FULL.org$SC.pass == 1 | new.Curate.FULL.org$SC.pass == 0),]
+#new.Curate.FULL <- droplevels(new.Curate.FULL)
+
+#new.Curate.FULL <- add_mouse(new.Curate.FULL)
+#new.Curate.FULL <- add_category(new.Curate.FULL)
+
+#pass.fail.table <- table(new.Curate.FULL$SC.pass, new.Curate.FULL$category)
+#this outsput shows the relative ratios of pass to fail biv segmentation
+
+```
+
+```{r remove.mice, echo=FALSE, eval=FALSE}
+
+#mice to remove from the Curated_BivData -- should be similar to the mice exclude from the MLH1
+
+#poor quality?
+
+#reaxamine these mice for if they should be included in 
+exclude_mice_list_MLH1 <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
+                  "8jun15_G_f3", "8jun15_G_f4",
+                  "4jan17_LEW_f2",
+                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
+                  "12sep16_MSM_f3","31aug16_MSM_m2",
+                  "27oct18_MOLF_m3",
+                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
+
+#    Musc males
+    "27aug18_CZECH_m1","8jun18_CZECH_m1",
+
+   "3sep18_AST_m1",
+
+    "3jul18_TOM_m1","3sep18_TOM_m1",
+
+                "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
+                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
+                  )
+#REaxamine and rename to exclude mice_BivData
+
+#Curated_BivData <- Curated_BivData[ ! Curated_BivData$mouse %in% exclude_mice_list, ]
+
+exmice.Curated_BivData <- Curated_BivData[ Curated_BivData$mouse %in% exclude_mice_list, ]
+
+#juvi males
+#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
+#juvi.mice_DF <- juvi.mice[!(is.na(juvi.mice$mouse) | juvi.mice$mouse==""), ]
+
+#old males
+
+```
+
+I'm considering removing the same mice from the MLH1 counts, from the BivData, but this might -- limit the comparisons -- might be too stict.
+IF foci are bright and clear, SC/bivalents intact -- include.
+
+IF it seems like primary reason for excluding is low cell number include bivalents
+
+Pro remove:
+
+- more similarity across data sets
+
+Con remove:
+
+- the reasons for poor quality of the exclude mice in above list, might be from 2-3 bad bivalents -- which are removed from the biv data during the curation process.
+
+```{r set.type, echo=FALSE}
+##change factor cols to numbers (factor -> chr -> number)
+Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
+
+##Numeric
+Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
+#I think the number of sig figs and rounding changes slightly from the blocks above
+
+#adding more feature calculations (IFD, ect)
+Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
+Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+
+#IFD.2 does mulitple IFDs (just absolute)
+#Curated_BivData <- add_IFD.2(Curated_BivData)
+
+Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
+
+Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
+Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
+Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
+Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)
+
+#not working
+Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
+Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
+
+#add telomere.dist
+Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
+                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
+                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
+                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
+
+Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
+
+#sis.co.ten
+Curated_BivData <- add_SisCoTen(Curated_BivData)
+
+```
+
+Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.
+
+- HetC_MM  : subset only mice
+
+IFD functions
+
+  - IFD: currently I think just does 2CO bivalents
+
+  - IFD.2: for mulitple IFDs (3CO ect.)
+
+## Strain specific data sets
+
+Compile dataframes for both sexes of each strain.
+
+```{r WSB.SC, echo=FALSE}
+#make the dataset
+WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
+#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
+table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
+wsb.ploty
+
+#make a table of the means and n ect
+wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r LEW.SC, echo=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+LEW.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+LEW.SC.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r G.SC, echo=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+G.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+G.SC.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r PWD.SC, echo=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+PWD.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+PWD.SC.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r KAZ.SC, echo=FALSE}
+#make the dataset
+KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+KAZ.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+KAZ.SC.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r SKIVE.Sc, echo=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+SKIVE.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+SKIVE.SC.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r MSM.SC, echo=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MSM.SC.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+```{r MOLF.SC, eval=FALSE}
+#make the dataset
+MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MOLF.ploty
+#remove the NAs
+
+#make a table of the means and n ect
+MOLF.SC.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+#can I plot these stats
+
+```
+
+## Male specific data set
+
+```{r male.sc.length.strain, echo=FALSE}
+
+#isolate all musc and male
+
+Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == ),]
+
+#make table
+SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+
+```
+
+# Manual Whole Cell
+
+Cleaning and integrating the dataframes for manual measured data.
+
+```{r whole.cell.manual.clean, echo=FALSE}
+#compare and clean up whole cell measures
+
+whole.cell.data_manual <- whole.cell.data_manual.OG[whole.cell.data_manual.OG$chrm.class != "XY",]
+
+colnames(whole.cell.data_manual.OG)
+#"return5"         "fileName"        "boxNumber"       "centromere"      "SC.length"  #    
+ #[6] "rank"            "chrm.class"      "hand.foci.count" "F1"              "F2"    #         
+#[11] "F3"              "notes"           "X"               "X.1"            
+#remove, X, X.1
+whole.cell.data_manual <- whole.cell.data_manual[-c(1, 13:14)]
+whole.cell.data_manual$Foci4 <- ""
+
+#rename cols to match BivData
+whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
+whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
+whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3
+whole.cell.data_manual$Foci4 <- whole.cell.data_manual$F4  #currently Foci4 is chr
+whole.cell.data_manual$chromosomeLength <- whole.cell.data_manual$SC.length
+
+whole.cell.data_manual$PER_Foci_1 <- whole.cell.data_manual$Foci1 / whole.cell.data_manual$chromosomeLength
+whole.cell.data_manual$PER_Foci_2 <- whole.cell.data_manual$Foci2 / whole.cell.data_manual$chromosomeLength
+whole.cell.data_manual$PER_Foci_3 <- whole.cell.data_manual$Foci3 / whole.cell.data_manual$chromosomeLength
+#whole.cell.data_manual$PER_Foci_4 <- whole.cell.data_manual$Foci4 / whole.cell.data_manual$chromosomeLength
+
+#mouse, strain, sex, category
+whole.cell.data_manual <- add_mouse(whole.cell.data_manual)
+whole.cell.data_manual <- add_category(whole.cell.data_manual)
+whole.cell.data_manual <- add_strain(whole.cell.data_manual)
+whole.cell.data_manual <- add_sex(whole.cell.data_manual)
+whole.cell.data_manual <- add_subsp(whole.cell.data_manual)
+
+#whole.cell.data_manual <-   whole.cell.data_manual[(!is.na(whole.cell.data_manual$fileName)),]
+whole.cell.data_manual <- whole.cell.data_manual[!(is.na(whole.cell.data_manual$fileName) | whole.cell.data_manual$fileName==""), ]
+
+#obj.id
+whole.cell.data_manual$Obj.ID <- paste(whole.cell.data_manual$fileName, whole.cell.data_manual$boxNumber, sep = "_")
+
+#add all the extra columns, blah
+#percent Foci position
+
+#ADD RANK
+for(cellys in as.character(unique(whole.cell.data_manual$fileName))){
+
+ # print(cellys)
+  chrm.order = rank(whole.cell.data_manual$SC.length[whole.cell.data_manual$fileName == cellys])
+  whole.cell.data_manual$rank[whole.cell.data_manual$fileName == cellys] <- chrm.order  
+  }
+
+#IFD 
+whole.cell.data_manual <- add_IFD(whole.cell.data_manual)
+
+#PER Foci
+
+```
+
+```{r clean.whole.cell, echo = FALSE}
+
+#CHECKED FOR MISSING bivs -- also check for wrong directions, ect
+
+#negative IFD
+#5aug17_14jul17_LEW_f1_sp1_11_rev_blob4
+
+#PER foci positions greater th
+
+#check the numbers of bivalents with a SC length, but no hand.foci.count
+
+```
+
+# Cell Level table
+
+- pass rate (don't need the OG.DF file, because I can count the number of boxes with the max index number)
+
+- number bivs per mice?
+
+- number cells represented per mouse
+
+- number of pass.SC w/out foci position data
+
+- number of foci with 'faint' mentioned in the notes
+
+The below cell level table is used for making the **long.biv.dataset**, (any others?)
+
+```{r cell.level.table, echo=FALSE}
+
+cell.level.BivData_table <- ddply(.data=Curated_BivData, 
+                       .(fileName),
+                       summarize, 
+                       boxs.IDd = max(boxNumber),
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                       ncells = length(unique(fileName)),
+                       nbivs = length(unique(Obj.ID)),
+
+                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       var.SC_length = var(chromosomeLength),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number
+
+                       cv.SC = cv(chromosomeLength),
+
+#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
+                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),
+
+                       mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                      var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
+                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
+                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)
+
+#cent_dis
+                      mean_cent_dist = mean(dis.cent, na.rm=TRUE)
+#telomere dist
+
+#can't think of a why to calq this by just 2CO
+
+)
+                       #add the centromere and ifd here?
+                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
+#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
+#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
+#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),
+
+#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
+#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
+ #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),
+
+#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
+#                       var.IFD = var(IFD1, na.rm=TRUE),
+#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
+#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),
+
+ #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
+#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
+#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
+#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)
+
+cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
+cell.level.BivData_table <- add_strain(cell.level.BivData_table)
+cell.level.BivData_table <- add_category(cell.level.BivData_table)
+
+as.data.frame( head(cell.level.BivData_table) )
+
+```
+
+```{r whole.cell_merging, echo=FALSE, eval=FALSE}
+#I think this block is redundant, could delete
+
+#check if there are fileNames
+double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% Curated_BivData$fileName, ]
+
+#which cell-levels fileNames have close to 15-17 chrms?
+almost_whole_cell <- cell.level.BivData_table[(cell.level.BivData_table$Nbiv < 21),] #cut off cells with too many
+almost_whole_cell <- almost_whole_cell[almost_whole_cell$Nbiv > 15,]
+
+#there is some discrpancy with the Nbiv and nbivs columns! ugh why, some have 20 and 10
+
+#get the Biv data 
+almost_whole_cell_BivData <- Curated_BivData[ Curated_BivData$fileName %in% almost_whole_cell$fileName, ]
+
+as.data.frame(  table( droplevels( almost_whole_cell_BivData$fileName) ) )
+#length(unique(almost_whole_cell_BivData$fileName))
+
+#what is almost whole cell BivData used for...
+```
+
+3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.  
+8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev
+
+## Long Biv Data Set
+
+I want to make a dataset of 'longest' 25% chromosomes from cells with at least 16 segmented bivalents from a single cell.
+
+```{r long.chrm.dataset, echo=FALSE}
+
+#1. how many biv's are from cells with >15 bivalents
+over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
+BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]
+
+#check for blank obj.id
+BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")
+
+#check for dup
+#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
+#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
+dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
+
+#remove duplicate obj.ids
+BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
+BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]
+
+#make sure obj.id and fileName are charcter for str stuff
+BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
+BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)
+
+#make sure 
+missing_from_cell_table <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% Biv_Data_mean_table$fileName, ]#there are 75, objs 3 cells
+#these are bivalents which don't appear the in cell level table, which can break the loop
+BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
+#final cleaned BivData table, 2428
+
+#loop for greping the mouse from file name ect
+for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){
+
+  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
+  new.egg <- paste0(newfile.str, collapse = "_")
+
+  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+  #the above can produe mulitple Q4
+
+  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
+  #these are just using the last index,
+
+  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+
+  BivData_from_15biv_cells$Q4.FN[i] <- the.FN
+   # print(c(i, the.FN, the.Q4))
+
+}
+
+BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= BivData_from_15biv_cells$cell.Q4), "yes", "no" )
+
+real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]
+
+```
+
+```{r table.long.biv, echo=FALSE}
+
+#real.long.bivData
+#calq the mouse averages for
+
+Long_biv_mouse.avs_4MM <- ddply(real.long.bivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+                       #IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+)
+
+Long_biv_mouse.avs_4MM <- add_sex(Long_biv_mouse.avs_4MM)
+Long_biv_mouse.avs_4MM <- add_strain(Long_biv_mouse.avs_4MM)
+Long_biv_mouse.avs_4MM <- add_subsp(Long_biv_mouse.avs_4MM)
+
+Long_biv_mouse.avs_4MM$strain <- unclass(Long_biv_mouse.avs_4MM$strain)
+Long_biv_mouse.avs_4MM$strain <- as.factor(Long_biv_mouse.avs_4MM$strain)
+
+Long_biv_mouse.avs_4MM$sex <- unclass(Long_biv_mouse.avs_4MM$sex)
+Long_biv_mouse.avs_4MM$sex <- as.factor(Long_biv_mouse.avs_4MM$sex)
+
+Long_biv_mouse.avs_4MM$subsp <- unclass(Long_biv_mouse.avs_4MM$v)
+Long_biv_mouse.avs_4MM$subsp <- as.factor(Long_biv_mouse.avs_4MM$subsp)
+
+```
+
+# Mouse leve table
+
+the mouse table will be the pooled stats / summary of all SC.passing +
+
+```{r mouse.table, echo=FALSE}
+
+mouse.level.BivData_table <- ddply(.data=Curated_BivData, 
+                       .(mouse),
+                       summarize,
+
+                       ncells = length(unique(fileName)),
+
+                       boxs.IDd = max(boxNumber),  #
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                       ncells = length(unique(fileName)),
+                       nbivs = length(unique(Obj.ID)),
+
+                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3),
+
+                       #sis.co.ten
+                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
+)
+
+as.data.frame( head(mouse.level.BivData_table) )
+#kable didn't work
+
+```
+
+```{r table.HetC.MM, echo=FALSE}
+
+```
+
+```{r Q1Q2.mouse.table, echo=FALSE}
+
+```
+
+analysis using the mouse level table -- to justify that using mouse averages is a good idea.
+
+Plots
+
+1. Histograms of N.bivs per cell, facet category
+
+2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)
+
+3.
+
+Is there a way to check my assumptions for these measues?
+
+What are the sources of error fot these data?
+
+```{r telo.cent_dist, echo=FALSE}
+
+#code for cal1 the telomere distance
+#space between last foci and telomere end
+
+#telo_dist <- if(hand.foci = 1){
+          #SC.Length - F1
+#}
+#elseif(hand_foci == 2){ #SC.length - F2}
+
+#code for centromere
+#space between first foci and centromere/centromere-end
+
+#cent_end_dist
+#cent_dist
+```
+
+# melt position table
+
+make the melt DF for the abs (raw) foci positions and then calculated the normalized (PER) foci positions from that table
+
+```{r MELT.block1, echo=FALSE}
+
+col.names.list <- colnames(Curated_BivData)
+#"fileName" ,  "chromosomeLength",        "centromere_ABS_Position", "centromere_PER_Position",
+#"boxNumber",               "Obj.ID",                  "SC.pass",                 "foci.pass",              
+#"curated.1pass.0fail",     "curation.notes",          "hand.foci.count",        "Foci1",                  
+#"Foci2",       "Foci3",                "notes",                 "mouse",                  
+#"category",                "strain",                  "sex",                    "subsp",                  
+#"IFD1_ABS",                "IFD1_PER",
+
+#all of the cols to use
+MELT_Curated_BivData <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
+              #leave out the Foci
+                "Foci1","Foci2","Foci3", "Foci4")],
+
+            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
+                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]
+
+#MELT, -- 15
+colnames(MELT_Curated_BivData) <- c(col.names.list[c(1:11, 16:17)], "Foci_Type", "Pos") #too many foci
+
+MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)
+
+#calq normalized melt pos
+MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]
+
+```
+
+```{r general.foci.pos, echo=FALSE, eval=FALSE}
+
+#make the normalized foci density plots
+#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
+#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]
+
+#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]#
+
+#remove 0 and NA
+#remove 4foci count?
+
+#check for values over 1
+
+#density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")
+
+#density
+#table(the4$mouse)
+#something wrong with the 4 bivalents
+#many of the lines just have 3 lines
+
+```
+
+This is proof that the plot codes work, but move the figures in the analysis doc. remeber this is just data-setup.
+
+think about weighing the density plots by the number of bivalents per category...
+
+# Assemble final BivData DF
+
+Main files from this plot:
+
+  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures
+
+  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  3. MELT_ABS: make melt version of foci positions for abs measuers
+
+  4. CELL and mouse table
+
+  5.
+
+MAIN BIVDATA_DF
+decide on the colnames that I want for the final DF
+
+fileName, Object_ID, box.number, centromere_ABS, centromere_PER
+chromosome length, SC.pass, Foci.pass, hand.foci.count, 
+F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD
+
+telomere_dist, cent_dist,
+
+mouse, strain, category, subsp,
+
+merge:
+cell.MLH1.count, cell.quality, mouse.age, 
+cell.total_skel, cell.total_binary
+
+# Save
+
+```{r save.file}
+
+#save Rdata
+save.image("~./MLH1repo/data/CleanBivData_1.28.20.RData")
+#save.image("data/BivData_4_PCA.RData")
+
+```
+
+# Analysis
+
+planned analyses for later docs, but this is the list for getting everything ready
+
+Use the Bivalent_Analysis.Rmd for analysis pieces
+BivPattern.Rmd started to analyze this but, not a very clear setup
+
+-chrom class proportions
+
+-SC lengths, by CO number
+
+<!--
+trash / junk
 #Apply the cleaning steps below to the FULL-OG DF
@@
 
-new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
-#make OG copy
-new.Curate.FULL.org <- new.Curate.FULL
-new.Curate.FULL.org <- add_category(new.Curate.FULL.org)
-
-#write files for basic stats to csv files
-#OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
-
-#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)
-
-OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)
-
-#write the curated file
-#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_21.1.20.csv", sep=",",
-#            row.names = TRUE)
-#
-#write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_21.1.20.csv", sep=",",
-#            row.names = TRUE)
-```
-
-```{r split.up.DF, echo=FALSE}
-#SC PASS
-new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#10480 #9819, #9769
-
-#fix these duplicated ones
-#remove duplicates
-Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151
-
-Curated_BivData <- add_mouse(Curated_BivData)
-Curated_BivData <- add_category(Curated_BivData)
-Curated_BivData <- add_strain(Curated_BivData)
-Curated_BivData <- add_sex(Curated_BivData)
-Curated_BivData <- add_subsp(Curated_BivData)
-
-Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
-#can I write a quick dummy check?
-#there are 2 lines with 0 hand.foci, but with a foci position
-
-```
-
 ```{r clean.data, echo=FALSE, eval=FALSE}
-
-Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
-Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
-Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]
-
-Curated_BivData$strain <- droplevels(Curated_BivData$strain)
-
-```
-
-```{r no.Foci.Pos, echo=FALSE}
-
-just.SC.length <- Curated_BivData[is.na(Curated_BivData$hand.foci.count),]#650 lines
-
-#table(just.SC.length$category)
-#decide what to do with these lines
-#how will they be represented in Mat Methods?
-
-```
-
-```{r input.check,echo=FALSE}
-
-#check for mistakes that might have been made when inputing / curation process
-
-#hand foci > 4
-
-#foci pos > SC lengths
-
-#negative PER.foci.pos
-
-#it hink most (ALL) of these came from bug in my function (ugh)
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#negative.IFD <- Curated_BivData[which(Curated_BivData$IFD1_ABS < 0),]
-#caught 2
-#15mar16_16jan16_G_f1_sp1_4_rev
-#27mar16_6mar16_G_m3_sp1_12_rev
-
-#neg Nrm F1
-#negative.IFD2 <- Curated_BivData[which(Curated_BivData$PER_Foci_1 < 0),]
-#0 really?
-```
-
-# Manual Whole Cell
-
-```{r whole.cell.manual.clean, echo=FALSE}
-#compare and clean up whole cell measures
-
-whole.cell.data_manual <- whole.cell.data_manual.OG[whole.cell.data_manual.OG$chrm.class != "XY",]
-
-colnames(whole.cell.data_manual.OG)
-#"return5"         "fileName"        "boxNumber"       "centromere"      "SC.length"  #    
- #[6] "rank"            "chrm.class"      "hand.foci.count" "F1"              "F2"    #         
-#[11] "F3"              "notes"           "X"               "X.1"            
-#remove, X, X.1
-whole.cell.data_manual <- whole.cell.data_manual[-c(1, 13:14)]
-whole.cell.data_manual$Foci4 <- ""
-
-#rename cols to match BivData
-whole.cell.data_manual$Foci1 <- whole.cell.data_manual$F1
-whole.cell.data_manual$Foci2 <- whole.cell.data_manual$F2
-whole.cell.data_manual$Foci3 <- whole.cell.data_manual$F3
-whole.cell.data_manual$Foci4 <- whole.cell.data_manual$F4  #currently Foci4 is chr
-whole.cell.data_manual$chromosomeLength <- whole.cell.data_manual$SC.length
-
-whole.cell.data_manual$PER_Foci_1 <- whole.cell.data_manual$Foci1 / whole.cell.data_manual$chromosomeLength
-whole.cell.data_manual$PER_Foci_2 <- whole.cell.data_manual$Foci2 / whole.cell.data_manual$chromosomeLength
-whole.cell.data_manual$PER_Foci_3 <- whole.cell.data_manual$Foci3 / whole.cell.data_manual$chromosomeLength
-#whole.cell.data_manual$PER_Foci_4 <- whole.cell.data_manual$Foci4 / whole.cell.data_manual$chromosomeLength
-
-#mouse, strain, sex, category
-whole.cell.data_manual <- add_mouse(whole.cell.data_manual)
-whole.cell.data_manual <- add_category(whole.cell.data_manual)
-whole.cell.data_manual <- add_strain(whole.cell.data_manual)
-whole.cell.data_manual <- add_sex(whole.cell.data_manual)
-whole.cell.data_manual <- add_subsp(whole.cell.data_manual)
-
-#whole.cell.data_manual <-   whole.cell.data_manual[(!is.na(whole.cell.data_manual$fileName)),]
-whole.cell.data_manual <- whole.cell.data_manual[!(is.na(whole.cell.data_manual$fileName) | whole.cell.data_manual$fileName==""), ]
-
-#obj.id
-whole.cell.data_manual$Obj.ID <- paste(whole.cell.data_manual$fileName, whole.cell.data_manual$boxNumber, sep = "_")
-
-#add all the extra columns, blah
-#percent Foci position
-
-#ADD RANK
-for(cellys in as.character(unique(whole.cell.data_manual$fileName))){
-
- # print(cellys)
-  chrm.order = rank(whole.cell.data_manual$SC.length[whole.cell.data_manual$fileName == cellys])
-  whole.cell.data_manual$rank[whole.cell.data_manual$fileName == cellys] <- chrm.order  
-  }
-
-#IFD 
-whole.cell.data_manual <- add_IFD(whole.cell.data_manual)
-
-#PER Foci
-
-```
-
-```{r clean.whole.cell, echo = FALSE}
-
-#CHECKED FOR MISSING bivs -- also check for wrong directions, ect
-
-#negative IFD
-#5aug17_14jul17_LEW_f1_sp1_11_rev_blob4
-
-#PER foci positions greater th
-
-#check the numbers of bivalents with a SC length, but no hand.foci.count
-
-```
-
-ToDo
-
-- Go back and clean up the original excel data
-
-- clean the bottom, fill in missing data
-
-- check the order of foci positions for 5aug17_14jul17_LEW_f1_sp1_11_rev_blob4.
-
-edit / clean up the lines with bad hand.foci.count (61)
-
-ToDo:
-
-1.clean up the non-numbered curated values (for the og df)
-
-2.Goal of this doc is to make DFs that are ready for the analysis.
-
-3.remove the 2 main sources of duplicate lines; 6mar16_PWD_m1, 1mar15_PWD_m1
-
-4. decide on the right number of passing curated bivalents per category I should settle for
-Justify why the current number is enough -- or subsample -- to random equivilent sized data sets
-
-5. make sure that the saved Rdata has all the right cols and tables for the next analyses
-
-6. consider integrating manual hand / whole cell measures (check for DUPs ect.)
-
-Main files from this plot:
-
-  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
-
-  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
-
-  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
-
-  4. MELT_ABS: make melt version of foci positions for abs measuers
-
-  5. CELL and
-
-  6. mouse table
-
-```{r CLEAN.UP.1, echo=FALSE, include=FALSE}
-
-#check for duplicates, and track them down
-anyDuplicated(new.Curate.FULL$Obj.ID)  #5335, #3467
-
-dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
-dup.curated <- add_mouse(dup.curated)
-
-#table(dup.curated$mouse)
-#1mar15_PWD_m1
-
-#which lines are duplicated
-#lots of PWD_female duplicated -- probably how I put together the curated files
-
-#new.Curate.FULL <- new.Curate.FULL.org[(new.Curate.FULL.org$SC.pass == 1 | new.Curate.FULL.org$SC.pass == 0),]
-#new.Curate.FULL <- droplevels(new.Curate.FULL)
-
-#new.Curate.FULL <- add_mouse(new.Curate.FULL)
-#new.Curate.FULL <- add_category(new.Curate.FULL)
-
-#pass.fail.table <- table(new.Curate.FULL$SC.pass, new.Curate.FULL$category)
-#this outsput shows the relative ratios of pass to fail biv segmentation
-
-```
-
-```{r set.type, echo=FALSE}
-
-##change factor cols to numbers (factor -> chr -> number)
-Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
-Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
-Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
-
-Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
-
-Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
-Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
-Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
-Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
-
-##Numeric
-Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
-Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
-Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
-
-Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
-
-Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
-Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
-Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
-Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
-#I think the number of sig figs and rounding changes slightly from the blocks above
-
-#adding more feature calculations (IFD, ect)
-Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
-Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
-
-#IFD.2 does mulitple IFDs (just absolute)
-#Curated_BivData <- add_IFD.2(Curated_BivData)
-
-Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
-
-Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
-Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
-Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
-Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)
-
-#not working
-Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
-Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
-
-#add telomere.dist
-Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
-                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
-                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
-                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
-
-Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
-
-#sis.co.ten
-Curated_BivData <- add_SisCoTen(Curated_BivData)
-
-```
-
-IFD functions
-
-  - IFD: currently I think just does 2CO bivalents
-
-  - IFD.2: for mulitple IFDs (3CO ect.)
-
-## Strain specific data sets
-
-Compile dataframes for both sexes of each strain.
-
-```{r WSB.SC, echo=FALSE}
-#make the dataset
-WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
-#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
-table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
-wsb.ploty
-
-#make a table of the means and n ect
-wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
-
-```{r LEW.SC, echo=FALSE}
-#make the dataset
-LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-LEW.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-LEW.SC.table <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-
-```
-
-```{r G.SC, echo=FALSE}
-#make the dataset
-G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-G.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-G.SC.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-
-```
-
-```{r PWD.SC, echo=FALSE}
-#make the dataset
-PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-PWD.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-PWD.SC.table <- ddply(PWD.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-```
-
-```{r KAZ.SC, echo=FALSE}
-#make the dataset
-KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-KAZ.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-KAZ.SC.table <- ddply(KAZ.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-
-```
-
-```{r SKIVE.Sc, echo=FALSE}
-#make the dataset
-SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-SKIVE.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-SKIVE.SC.table <- ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-```
-
-```{r MSM.SC, echo=FALSE}
-#make the dataset
-MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-MSM.SC.table <- ddply(MSM.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-```
-
-```{r MOLF.SC, eval=FALSE}
-#make the dataset
-MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MOLF.ploty
-#remove the NAs
-
-#make a table of the means and n ect
-MOLF.SC.table <- ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-# Cell Level table
-
-- pass rate (don't need the OG.DF file, because I can count the number of boxes with the max index number)
-
-- number bivs per mice?
-
-- number cells represented per mouse
-
-- number of pass.SC w/out foci position data
-
-- number of foci with 'faint' mentioned in the notes
-
-```{r cell.level.table, echo=FALSE}
-
-cell.level.BivData_table <- ddply(.data=Curated_BivData, 
-                       .(fileName),
-                       summarize, 
-                       boxs.IDd = max(boxNumber),
-                       nboxes_passed = length(boxNumber),
-
-                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
-                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
-                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
-                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
-                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
-
-                       ncells = length(unique(fileName)),
-                       nbivs = length(unique(Obj.ID)),
-
-                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
-                       var.SC_length = var(chromosomeLength),
-                       sd.SC_length   = round(sd(chromosomeLength), 3),
-                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number
-
-                       cv.SC = cv(chromosomeLength),
-
-#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
-                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),
-
-                       mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                      var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
-                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
-                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)
-
-#cent_dis
-                      mean_cent_dist = mean(dis.cent, na.rm=TRUE)
-#telomere dist
-
-#can't think of a why to calq this by just 2CO
-
-)
-                       #add the centromere and ifd here?
-                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
-#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
-#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
-#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),
-
-#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
-#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
- #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),
-
-#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
-#                       var.IFD = var(IFD1, na.rm=TRUE),
-#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
-#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),
-
- #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
-#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
-#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
-#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)
-
-cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
-cell.level.BivData_table <- add_strain(cell.level.BivData_table)
-cell.level.BivData_table <- add_category(cell.level.BivData_table)
-
-as.data.frame( head(cell.level.BivData_table) )
-
-```
-
-```{r whole.cell_merging, echo=FALSE}
-
-#check if there are fileNames
-double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% Curated_BivData$fileName, ]
-
-#which cell-levels fileNames have close to 15-17 chrms?
-
-almost_whole_cell <- cell.level.BivData_table[(cell.level.BivData_table$Nbiv < 21),]
-almost_whole_cell <- almost_whole_cell[almost_whole_cell$Nbiv > 15,]
-
-#there is some discrpancy with the Nbiv and nbivs columns! ugh why, some have 20 and 10
-
-#get the Biv data 
-almost_whole_cell_BivData <- Curated_BivData[ Curated_BivData$fileName %in% almost_whole_cell$fileName, ]
-
-as.data.frame(  table( droplevels( almost_whole_cell_BivData$fileName) ) )
-length(unique(almost_whole_cell_BivData$fileName))
-
-#what next?
-
-```
-
-3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.
-
-8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev
-
-# Mouse leve table
-
-the mouse table will be the pooled stats / summary of all SC.passing +
-
-```{r mouse.table, echo=FALSE}
-
-mouse.level.BivData_table <- ddply(.data=Curated_BivData, 
-                       .(mouse),
-                       summarize,
-
-                       ncells = length(unique(fileName)),
-
-                       boxs.IDd = max(boxNumber),  #
-                       nboxes_passed = length(boxNumber),
-
-                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
-                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
-                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
-                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
-                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
-
-                       ncells = length(unique(fileName)),
-                       nbivs = length(unique(Obj.ID)),
-
-                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
-                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
-                       sd.SC_length   = round(sd(chromosomeLength), 3),
-                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3),
-
-                       #sis.co.ten
-                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
-)
-
-as.data.frame( head(mouse.level.BivData_table) )
-#kable didn't work
-
-```
-
-```{r Q1Q2.mouse.table, echo=FALSE}
-
-```
-
-analysis using the mouse level table -- to justify that using mouse averages is a good idea.
-
-Plots
-
-1. Histograms of N.bivs per cell, facet category
-
-2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)
-
-3.
-
-Is there a way to check my assumptions for these measues?
-
-What are the sources of error fot these data?
-
-```{r telo.cent_dist, echo=FALSE}
-
-#code for cal1 the telomere distance
-#space between last foci and telomere end
-
-#telo_dist <- if(hand.foci = 1){
-          #SC.Length - F1
-#}
-#elseif(hand_foci == 2){ #SC.length - F2}
-
-#code for centromere
-#space between first foci and centromere/centromere-end
-
-#cent_end_dist
-#cent_dist
-```
-
-# melt position table
-
-make the melt DF for the abs (raw) foci positions and then calculated the normalized (PER) foci positions from that table
-
-```{r MELT.block1, echo=FALSE}
-
-col.names.list <- colnames(Curated_BivData)
-#"fileName" ,  "chromosomeLength",        "centromere_ABS_Position", "centromere_PER_Position",
-#"boxNumber",               "Obj.ID",                  "SC.pass",                 "foci.pass",              
-#"curated.1pass.0fail",     "curation.notes",          "hand.foci.count",        "Foci1",                  
-#"Foci2",       "Foci3",                "notes",                 "mouse",                  
-#"category",                "strain",                  "sex",                    "subsp",                  
-#"IFD1_ABS",                "IFD1_PER",
-
-#all of the cols to use
-MELT_Curated_BivData <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
-              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
-              #leave out the Foci
-                "Foci1","Foci2","Foci3", "Foci4")],
-
-            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
-                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
-              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))
-
-MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]
-
-#MELT, -- 15
-colnames(MELT_Curated_BivData) <- c(col.names.list[c(1:11, 16:17)], "Foci_Type", "Pos") #too many foci
-
-MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)
-
-#calq normalized melt pos
-MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength
-
-MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]
-
-```
-
-```{r general.foci.pos, echo=FALSE, eval=FALSE}
-
-#make the normalized foci density plots
-#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
-#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]
-
-#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]#
-
-#remove 0 and NA
-#remove 4foci count?
-
-#check for values over 1
-
-#density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")
-
-#density
-#table(the4$mouse)
-#something wrong with the 4 bivalents
-#many of the lines just have 3 lines
-
-```
-
-This is proof that the plot codes work, but move the figures in the analysis doc. remeber this is just data-setup.
-
-think about weighing the density plots by the number of bivalents per category...
-
-# Assemble final BivData DF
-
-Main files from this plot:
-
-  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures
-
-  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
-
-  3. MELT_ABS: make melt version of foci positions for abs measuers
-
-  4. CELL and mouse table
-
-  5.
-
-MAIN BIVDATA_DF
-decide on the colnames that I want for the final DF
-
-fileName, Object_ID, box.number, centromere_ABS, centromere_PER
-chromosome length, SC.pass, Foci.pass, hand.foci.count, 
-F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD
-
-telomere_dist, cent_dist,
-
-mouse, strain, category, subsp,
-
-merge:
-cell.MLH1.count, cell.quality, mouse.age, 
-cell.total_skel, cell.total_binary
-
-# Save
-
-```{r save.file}
-
-#save Rdata
-save.image("~./MLH1repo/data/CleanBivData_1.28.20.RData")
-#save.image("data/BivData_4_PCA.RData")
-
-```
-
-# Analysis
-
-planned analyses for later docs, but this is the list for getting everything ready
-
-Use the Bivalent_Analysis.Rmd for analysis pieces
-BivPattern.Rmd started to analyze this but, not a very clear setup
-
--chrom class proportions
-
--SC lengths, by CO number
+#Keep these biv data for the male polymorphism version
+
+#Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
+#Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
+#Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]
+
+#Curated_BivData$strain <- droplevels(Curated_BivData$strain)
+```
+
+-->

### 2020-02-03 src/BivData_Setup.Rmd
working on cleaning up-organizing script

@@
 ---
+
+# Main Files
+
+Main files from this Script:
+
+  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
+
+  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
+
+  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  4. **MELT_ABS**: make melt version of foci positions for abs measuers
+
+  5. CELL level table (full all mice) (use this to make long biv dataset)
+
+  6. Long Biv Dataset
+
+  7. mouse table (full all mice)
+
+  8. **Strain DFs** for Het models **Mouse.Table_BivData_4MM**
+
+  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
 
@@
 
-Main files from this plot:
+Main files from this Script:
 
@@
 
-  4. MELT_ABS: make melt version of foci positions for abs measuers
-
-  5. CELL and
-
-  6. mouse table
-
-  7. **Strain DFs** for Het models
-
-  8. **male Subsp DFs** for Polymorphism models
+  4. **MELT_ABS**: make melt version of foci positions for abs measuers
+
+  5. CELL level table (full all mice) (use this to make long biv dataset)
+
+  6. Long Biv Dataset
+
+  7. mouse table (full all mice)
+
+  8. **Strain DFs** for Het models (summary strain DFs and tables?)
+
+  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models
+
+  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
 
@@
 
+#these measure ect should be in the male - mouse 
 #isolate all musc and male
-
-Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == ),]
+Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "Musculus"),]
 
@@
 
-## Long Biv Data Set
+# Long Biv Data Set
 
@@
 
-# Mouse leve table
+# Mouse level table
+
+Mouse level tables are the main meat of the models. There will be several --- which account for different combinations of quality scores.
+
+BC their for MM, format the tables for run the mixed models (unordered factors, ect)
 
@@
 
-```{r table.HetC.MM, echo=FALSE}
+```{r mouse.table.HetC.MM, echo=FALSE}
+
+#this mouse level table -- should not have strains without female obs
+#only Mmusculus
+
+#is there a way that the above table would be different than the 
+Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+Mouse.Table_BivData_4MM <- add_sex(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_strain(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_subsp(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_category(Mouse.Table_BivData_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Mouse.Table_BivData_4MM$subsp <- factor(Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
+Mouse.Table_BivData_4MM$sex <- factor(Mouse.Table_BivData_4MM$sex, ordered = FALSE)
+Mouse.Table_BivData_4MM$strain <- factor(Mouse.Table_BivData_4MM$strain, ordered = FALSE)
+
+#remove all the non-used strains!!
+Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$strain != "AST",]
+Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[Curated_BivData$strain != "TOM",]
+Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$strain != "CZECH",]
+Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
+#MOLF
+
+#These shouldn't be needed for now.
+
+#assign high and low
+#divide musc males into groups
+#mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+#                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+#                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))#
+#
+#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+#                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+```
+
+```{r mouse.table.Male.Poly, echo=FALSE}
+
+#this table will have male BivData
+#it will include Musc strains
+
+Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+Male.poly.Mouse.Table_BivData_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+Male.poly.Mouse.Table_BivData_4MM <- add_sex(Male.poly.Mouse.Table_BivData_4MM)
+Male.poly.Mouse.Table_BivData_4MM <- add_strain(Male.poly.Mouse.Table_BivData_4MM)
+Male.poly.Mouse.Table_BivData_4MM <- add_subsp(Male.poly.Mouse.Table_BivData_4MM)
+Male.poly.Mouse.Table_BivData_4MM <- add_category(Male.poly.Mouse.Table_BivData_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Male.poly.Mouse.Table_BivData_4MM$subsp <- factor(Male.poly.Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
+Male.poly.Mouse.Table_BivData_4MM$sex <- factor(Male.poly.Mouse.Table_BivData_4MM$sex, ordered = FALSE)
+Male.poly.Mouse.Table_BivData_4MM$strain <- factor(Male.poly.Mouse.Table_BivData_4MM$strain, ordered = FALSE)
+
+#remove females
+Male.poly.Mouse.Table_BivData_4MM$sex <- droplevels(Male.poly.Mouse.Table_BivData_4MM$sex)
+#MOLF
 
@@
 
+#How Do I apply quality score things to BivData?
+#make the same table but
+
 ```
@@
 
-```{r telo.cent_dist, echo=FALSE}
-
-#code for cal1 the telomere distance
-#space between last foci and telomere end
-
-#telo_dist <- if(hand.foci = 1){
-          #SC.Length - F1
-#}
-#elseif(hand_foci == 2){ #SC.length - F2}
-
-#code for centromere
-#space between first foci and centromere/centromere-end
-
-#cent_end_dist
-#cent_dist
-```
-
-# melt position table
+# Melt Position Table
 
@@
 
-# Assemble final BivData DF
-
-Main files from this plot:
+# Assemble Final BivData DF
+
+Main files from this script:
 
@@
 
+```{r telo.cent_dist, echo=FALSE}
+
+#code for cal1 the telomere distance
+#space between last foci and telomere end
+
+#telo_dist <- if(hand.foci = 1){
+          #SC.Length - F1
+#}
+#elseif(hand_foci == 2){ #SC.length - F2}
+
+#code for centromere
+#space between first foci and centromere/centromere-end
+
+#cent_end_dist
+#cent_dist
+
 -->

### 2020-02-03 src/BivData_Setup.Rmd
lots of cleaning up and de-bugging

@@
 - remove BivData! (these are all Peromyscus!!) (not sure where this DF comes from)
-
-calq the nromalized.siscoten, double check the negative values; 13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14, 27aug15_12jun15_WSB_m2_sp1_18_rev_18, 21nov17_16nov17_MSM_f3_sp1_3_rev_1
-
-foci.pass = 2  5aug17_14jul17_MSM_f1_sp1_13_rev_3
-
-hand foci =0 but there is a foci position
-
-too high hand foci count
 
@@
 new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
+#83975
 #83976
@@
 new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
+#this takes a while to run
 
@@
 ```{r split.up.DF, echo=FALSE}
-
 #new.Curate.FULL to Curated_BivData
@@
 #SC PASS
-new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#10480 #9819, #9769
+new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#
+#10480 #9819, #9769
 
@@
 
-Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
-```
+#don't think I need this one
+#Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
+```
+
+```{r set.type_Curated_BivData, echo=FALSE}
+##change factor cols to numbers (factor -> chr -> number)
+Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
+
+##Numeric
+Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
+#I think the number of sig figs and rounding changes slightly from the blocks above
+
+#adding more feature calculations (IFD, ect)
+Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
+Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+
+#IFD functions
+#  - IFD: currently I think just does 2CO bivalents 
+#  - IFD.2: for mulitple IFDs (3CO ect.)
+
+#IFD.2 does mulitple IFDs (just absolute)
+#Curated_BivData <- add_IFD.2(Curated_BivData)
+
+Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
+
+Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
+Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
+Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
+Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)
+
+#not working
+Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
+Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
+
+#add telomere.dist
+Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
+                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
+                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
+                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
+
+Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
+
+#sis.co.ten
+Curated_BivData <- add_SisCoTen(Curated_BivData)
+
+```
+
+Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.
 
@@
 
-just.SC.length <- Curated_BivData[is.na(Curated_BivData$hand.foci.count),]#650 lines
+#no posittion
+#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
+just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]
 
 #table(just.SC.length$category)
+
 #decide what to do with these lines
@@
 ```{r input.check,echo=FALSE}
+
+#10jan18_13jan17_LEW_f1_sp1_14_rev_12  '?'
+
+##calq the nromalized.siscoten, double check the negative values;  #13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14, #27aug15_12jun15_WSB_m2_sp1_18_rev_18, #21nov17_16nov17_MSM_f3_sp1_3_rev_1
+
+#foci.pass = 2  #5aug17_14jul17_MSM_f1_sp1_13_rev_3
+#hand foci =0 but there is a foci position
+#too high hand foci count
 
@@
 #0 really?
+
 ```
@@
 
+<!--
 I'm considering removing the same mice from the MLH1 counts, from the BivData, but this might -- limit the comparisons -- might be too stict.
@@
 
-IF it seems like primary reason for excluding is low cell number include bivalents
+If it seems like primary reason for excluding is low cell number include bivalents
 
@@
 - the reasons for poor quality of the exclude mice in above list, might be from 2-3 bad bivalents -- which are removed from the biv data during the curation process. 
-
-```{r set.type, echo=FALSE}
-##change factor cols to numbers (factor -> chr -> number)
-Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
-Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
-Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
-
-Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
-
-Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
-Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
-Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
-Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
-
-##Numeric
-Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
-Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
-Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
-
-Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
-
-Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
-Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
-Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
-Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
-#I think the number of sig figs and rounding changes slightly from the blocks above
-
-#adding more feature calculations (IFD, ect)
-Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
-Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
-
-#IFD.2 does mulitple IFDs (just absolute)
-#Curated_BivData <- add_IFD.2(Curated_BivData)
-
-Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
-
-Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
-Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
-Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
-Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)
-
-#not working
-Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
-Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
-
-#add telomere.dist
-Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
-                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
-                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
-                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
-
-Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
-
-#sis.co.ten
-Curated_BivData <- add_SisCoTen(Curated_BivData)
-
-```
-
-Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.
-
-- HetC_MM  : subset only mice
-
-IFD functions
-
-  - IFD: currently I think just does 2CO bivalents
-
-  - IFD.2: for mulitple IFDs (3CO ect.)
-
-## Strain specific data sets
+-->
+
+## Strain specific BivData sets
 
@@
 
-## Male specific data set
+## Male specific BivData set
 
@@
 #isolate all musc and male
-Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "Musculus"),]
-
-#make table
+Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "M.musculus"),]
+
+#make strain average table
 SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
@@
 )
-
-```
-
-# Manual Whole Cell
+```
+
+# Manual Whole Cell BivData
 
@@
 #PER Foci
-
-```
-
-```{r clean.whole.cell, echo = FALSE}
+```
+
+```{r clean.whole.cell_table, echo = FALSE, eval=FALSE}
+
+#calq the mouse means
+manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
+                        median_SC = median(SC.length, na.rm = TRUE),
+                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
+#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
+
+)
+
+manual_summary_stats <- add_mouse(manual_summary_stats)
+manual_summary_stats <- add_strain(manual_summary_stats)
+manual_summary_stats <- add_sex(manual_summary_stats)
+manual_summary_stats <- add_subsp(manual_summary_stats)
 
@@
 
-# Cell Level table
+# Cell Level Table
 
@@
 
+                #this below might not be correct
                      mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+
                        var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
@@
 #cent_dis
-                      mean_cent_dist = mean(dis.cent, na.rm=TRUE)
-#telomere dist
-
-#can't think of a why to calq this by just 2CO
-
-)
+                      mean_cent_dist = mean(dis.cent, na.rm=TRUE),
+                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]] #calq threshold for all cells, pull out 'almost.whole.cells' later
+
+                        )
+
 #add the centromere and ifd here?
@@
 
-as.data.frame( head(cell.level.BivData_table) )
+#as.data.frame( head(cell.level.BivData_table) )
 
@@
 
-# Long Biv Data Set
+# Long BivData Set
 
@@
 
-#1. how many biv's are from cells with >15 bivalents
+#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
 over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
@@
 dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
+#Dups -- 50, all from 1mar15_PWD_m1
 
@@
 
-#make sure obj.id and fileName are charcter for str stuff
+#make sure obj.id and fileName are charcter for studd below
 BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
@@
 
-#make sure 
-missing_from_cell_table <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% Biv_Data_mean_table$fileName, ]#there are 75, objs 3 cells
+#what are the Biv_Data_mean_table ... ? old version of cell table..
+#replacing Biv_Data_mea_table with... cell.level.BivData_table
+
+missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ]#0, 
+#all these files stem from
+
 #these are bivalents which don't appear the in cell level table, which can break the loop
@@
 
+#2. calq the Q4 threshold
 #loop for greping the mouse from file name ect
+
+#mini.now  <- BivData_from_15biv_cells[1:300,]
+#broken again -- confusion between the number Q4 and the fileName
+
 for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){
+
+  #print(c(i, " the i") )
 
@@
 
-  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
+  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+  #there is no Qar4 in this DF
+
+  #print(c("the Q4  ",the.Q4)  )
+
   #the above can produe mulitple Q4
@@
   #these are just using the last index,
-
-  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]
-
-  BivData_from_15biv_cells$Q4.FN[i] <- the.FN
+  #this might be wrong....
+
+  the.FN <- as.character(cell.level.BivData_table$fileName[which(grepl(new.egg, cell.level.BivData_table$fileName ) )] )
+
+  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)  #this FN is NOT RIGHT
    # print(c(i, the.FN, the.Q4))
-
 }
 
-BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= BivData_from_15biv_cells$cell.Q4), "yes", "no" )
-
-real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]
-
-```
+#3.mark which bivalents are above the threshold
+#WHY DOESN"T THIS WORK?
+
+BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= as.numeric(BivData_from_15biv_cells$cell.Q4)  ), "yes", "no" )
+
+real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]#713 observations
+```
+
+The above blocks of code 1) isolates bivalents measures from cells with >15 segmented bivalents 2) for the cells in these data calculates the Q4 threshold and 3) marks which bivalents are longer than the threshold.
 
@@
 
-#real.long.bivData
-#calq the mouse averages for
-
-Long_biv_mouse.avs_4MM <- ddply(real.long.bivData, c("mouse"), summarise,
+# make a table for the Mixed Models -- from the long.BivData
+# Calq mouse averages
+
+#add in all the other stats there are in the other tables
+
+Long.biv_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+
                               Nmice = length(unique(mouse)),
                         Ncells  = length(Obj.ID),
+
                           mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+
                         cV = cv(chromosomeLength),
@@
                         se   = round(sd / sqrt(Ncells), 3),
-                       #IFD things
+
+                        #SC lengths by class
+                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
+                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
                 #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
@@
 
-)
-
-Long_biv_mouse.avs_4MM <- add_sex(Long_biv_mouse.avs_4MM)
-Long_biv_mouse.avs_4MM <- add_strain(Long_biv_mouse.avs_4MM)
-Long_biv_mouse.avs_4MM <- add_subsp(Long_biv_mouse.avs_4MM)
-
-Long_biv_mouse.avs_4MM$strain <- unclass(Long_biv_mouse.avs_4MM$strain)
-Long_biv_mouse.avs_4MM$strain <- as.factor(Long_biv_mouse.avs_4MM$strain)
-
-Long_biv_mouse.avs_4MM$sex <- unclass(Long_biv_mouse.avs_4MM$sex)
-Long_biv_mouse.avs_4MM$sex <- as.factor(Long_biv_mouse.avs_4MM$sex)
-
-Long_biv_mouse.avs_4MM$subsp <- unclass(Long_biv_mouse.avs_4MM$v)
-Long_biv_mouse.avs_4MM$subsp <- as.factor(Long_biv_mouse.avs_4MM$subsp)
-
-```
-
-# Mouse level table
-
-Mouse level tables are the main meat of the models. There will be several --- which account for different combinations of quality scores.
-
-BC their for MM, format the tables for run the mixed models (unordered factors, ect)
-
-the mouse table will be the pooled stats / summary of all SC.passing +
-
-```{r mouse.table, echo=FALSE}
+        #          mean_siscoten = mean(siscoten)
+        #
+)
+#mouse averages for long.biv
+
+Long.biv_mouse_table <- add_sex(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_strain(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_subsp(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_category(Long.biv_mouse_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Long.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
+Long.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
+Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)
+
+#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)
+
+#1) Long Biv for 
+#remove all the non-used strains!!
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table
+
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
+Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
+
+#table(droplevels( Long.biv_mouse_table_HetC$category) )  #for double checking
+
+#2) Long Biv for male.poly
+#just males, just Musculus
+Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
+
+Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
+
+#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
+```
+
+The above code chunk makes 2 mouse level tables; **Long.biv_mouse_table_HetC** and **Long.biv_mouse_table_male.poly**. Run these through the model systems.
+
+# Mouse Level Tables
+
+Mouse level tables are the main meat of the models. There will be several for use in different models. Format these tables for run the mixed models (unordered factors, ect).
+
+```{r FULL.mouse.table, echo=FALSE}
 
@@
 
-#vari
+#
 Mouse.Table_BivData_4MM <- add_sex(Mouse.Table_BivData_4MM)
@@
 Male.poly.Mouse.Table_BivData_4MM$sex <- droplevels(Male.poly.Mouse.Table_BivData_4MM$sex)
-#MOLF
 
@@
 
-Plots
-
-1. Histograms of N.bivs per cell, facet category
-
-2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)
-
-3.
-
-Is there a way to check my assumptions for these measues?
-
-What are the sources of error fot these data?
-
 # Melt Position Table
 
-make the melt DF for the abs (raw) foci positions and then calculated the normalized (PER) foci positions from that table
+The BivData MELT DF is primarily for the nrmF1.pos plots/analysis. It is a Melted DF for raw and normalized foci positions from the Curated_BivData.
 
@@
 MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]
-
-```
-
-```{r general.foci.pos, echo=FALSE, eval=FALSE}
-
-#make the normalized foci density plots
-#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
-#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]
-
-#MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]#
-
-#remove 0 and NA
-#remove 4foci count?
-
-#check for values over 1
-
-#density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")
-
-#density
-#table(the4$mouse)
-#something wrong with the 4 bivalents
-#many of the lines just have 3 lines
-
-```
-
-This is proof that the plot codes work, but move the figures in the analysis doc. remeber this is just data-setup.
-
-think about weighing the density plots by the number of bivalents per category...
+```
 
@@
 
-Main files from this script:
-
-  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures
-
-  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
-
-  3. MELT_ABS: make melt version of foci positions for abs measuers
-
-  4. CELL and mouse table
-
-  5.
+Main files from this Script:
+
+  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
+
+  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
+
+  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  4. **MELT_ABS**: make melt version of foci positions for abs measuers
+
+  5. CELL level table (full all mice) (use this to make long biv dataset)
+
+  6. Long Biv Dataset
+
+  7. mouse table (full/all mice)
+
+  8. **Strain DFs** for Het models **Mouse.Table_BivData_4MM**
+
+  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
 
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_1.28.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_3.2.20.RData")
 #save.image("data/BivData_4_PCA.RData")
@@
 trash / junk
+
+Plots
+
+1. Histograms of N.bivs per cell, facet category
+
+2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)
+
+3.
+
+Is there a way to check my assumptions for these measues?
+
+What are the sources of error fot these data?
+
 #Apply the cleaning steps below to the FULL-OG DF
@@
 
-```{r telo.cent_dist, echo=FALSE}
+```{r telo.cent_dist, echo=FALSE, eval=FALSE}
 
@@
 #cent_dist
+```
 

### 2020-02-04 doc/Bivalent_Analysis.Rmd
work and notes

@@
 #load Clean BivData
-load(file="data/CleanBivData_1.28.20.RData")
+load(file="data/CleanBivData_3.2.20.RData")
+#load(file="data/CleanBivData_1.28.20.RData")
 
 #load other data files, MLH1
-load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")#added batch18
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_2.2.20.RData")#added batch18
 
@@
 
-#remove all the non-used strains!!
-Curated_BivData <- Curated_BivData[Curated_BivData$strain != "AST",]
-Curated_BivData <- Curated_BivData[Curated_BivData$strain != "TOM",]
-Curated_BivData <- Curated_BivData[Curated_BivData$strain != "CZECH",]
-Curated_BivData$strain <- droplevels(Curated_BivData$strain)
-#MOLF
-
-#non-ordered factors, 
-Curated_BivData$subsp <- factor(Curated_BivData$subsp, ordered = FALSE)
-Curated_BivData$sex <- factor(Curated_BivData$sex, ordered = FALSE)
-Curated_BivData$strain <- factor(Curated_BivData$strain, ordered = FALSE)
-
-#all numbers, numbers
-#all relavent factors unordered
-
-#I'd like to merge the additional strains used for the male tests
-
 #the mouse averages will be used
@@
 - long.biv
-
-### Mixed Model Tests
 
@@
 
-#### Mixed Model 1
+#### HetC_M1
 
@@
 
-```{r M1_1CO, echo=FALSE}
-
-#run Mixed models
+Below's the code for Mixed model results -- try to organize them into a table or somethinf
+
+```{r M1_many.metrics, echo=FALSE}
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
 
 #all pooled SC lengths
-SCLength_all_.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=mouse.avs_4MM) 
-#summary(Reduced.strain.HQ)
-
-drop1(SCLength_all_.MM.M1, test="Chisq") # p = 0.0002045 ***
-
-exactRLRT(SCLength_all_.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0
-as.data.frame(round(summary(SCLength_all_.MM.M1)$tTable, 6))
+lmer.SCLength_all_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
+anova(lmer.SCLength_all_MM.M1) # also F values
+
+exactRLRT(lmer.SCLength_all_MM.M1)
+as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
+#can't get p values from this summary
 
 #1CO
+SCLength_1CO_MM.M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+drop1(SCLength_1CO_MM.M1, test="Chisq")
+exactRLRT(SCLength_1CO_MM.M1)
+as.data.frame(round(summary(SCLength_1CO_MM.M1)$coefficients, 6))
 
 #2CO
-
-#I think I need to re-calq the mouse averages for just 1CO
-#mean.SC_1CO is mean 1CO
-
-SC.Length_1CO_MM <- lme(mean.SC_1CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(SC.Length_1CO_MM)$tTable, 6 ) )
-```
-
-#### 2CO
-
-```{r 2CO.MM, echo=FALSE, eval=FALSE}
-
-#remove MOLF
-mouse.avs_4MM.99 <- mouse.avs_4MM[mouse.avs_4MM$strain != "MOLF",]
-
-SC.Length_2CO_MM <- lme(mean.SC_2CO ~ subsp * sex, 
-                  random=list(strain=pdDiag(~sex),
-                        data= mouse.avs_4MM.99, 
-                                                                                       na.action=na.omit ) )
-#keep getting errors for building this model
-
-as.data.frame( round(  summary(SC.Length_2CO_MM)$tTable, 6 ) )
-
-table(mouse.avs_4MM$strain, mouse.avs_4MM$sex)
-
-```
-
-#### 3CO
-
-```{r 3CO.MM, echo=FALSE}
-
-SC.Length_3CO_MM <- lme(mean.SC_3CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),  na.action=na.omit  )
-#this model is also generating error...
-#don't know how to fix
-
-as.data.frame( round(  summary(SC.Length_3CO_MM)$tTable, 6 ) )
-```
-
-#### long.biv
-
-```{r long.biv.MM, echo=FALSE, eval=FALSE}
-
-SC.Length_long.biv_MM <- lme(mean_SC ~ subsp * sex, data= Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex),  na.action=na.omit  ) )
-
-#now there is an error after un-ordering them
-
-as.data.frame( round(  summary(SC.Length_long.biv_MM)$tTable, 6 ) )
-
-#model for long Bivs
-#long_bivData_SC.Length_MM <- lme(mean_SC ~ subsp * sex, data= #Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(long_bivData_SC.Length_MM)$tTable, 6 ) )
+SCLength_2CO_MM.M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+drop1(SCLength_2CO_MM.M1, test="Chisq")
+exactRLRT(SCLength_2CO_MM.M1)
+as.data.frame(round(summary(SCLength_2CO_MM.M1)$coefficients, 6))
+
+#Long.Biv
+SCLength_long.biv_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
+
+drop1(SCLength_long.biv_MM.M1, test="Chisq")
+exactRLRT(SCLength_long.biv_MM.M1)
+as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+
+```
+
+#### HetC_M2
+
+All fixed and all interactions
+
+```{r M2_many.metrics, echo=FALSE, eval=FALSE}
+
+#feed these 2 DFs into M2
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+```
+
+#### HetC_M3
+
+```{r HetC_M3, echo=FALSE}
+
 ```
@@
 
-APRIL, consider moving these strain chunks to the bivData set up
-
 (adding all of these blocks to -- the setup script); strain.Bivalents.DF
-
-```{r WSB.SC, echo=FALSE, eval=FALSE}
-#make the dataset
-WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
-#correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
-table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
-wsb.ploty
-
-#make a table of the means and n ect
-wsb.SC.table <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-#can I plot these stats
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
 
@@
 
-#### male strain comparisons
+#### Male Poly
+
+These are the models for male polymorphism, (I am re-thinking the tests I think)
+
+Use these DFs for the models:  **Male.poly.Mouse.Table_BivData_4MM**  and **Long.biv_mouse_table_male.poly**
 
 ```{r SC.length.male_tests, echo=FALSE}
-
 ## run the tests to check following expectations
@@
 #3.MSM and MOLF are different from each other
+
+#Male.poly.Mouse.Table_BivData_4MM
+#Long.biv_mouse_table_male.poly
 
@@
 
-pairwise.t.test(Ozone, Month, p.adj = "bonf")
+#pairwise.t.test(Ozone, Month, p.adj = "bonf")
 
@@
 #### t.test for SC Length
+
+what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
 
@@
 
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
-
-```{r MM_IFB.ABS, echo=FALSE}
+$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r HetC_M1_IFD_many.metrics, echo=FALSE}
 #get a list of the old mice to double check
@@
 
-#remove the SKIVE male line which is na
-mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$mean_IFD.2CO_ABS) | mouse.avs_4MM$mean_IFD.2CO_ABS==""), ]
-
-lme_MM_IFD_ABS <- lme(mean_IFD.2CO_ABS ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),na.action=na.omit  )
-
-as.data.frame( round(  summary(lme_MM_IFD_ABS)$tTable, 6 ) )
-#p
-#same missing values in obc
-
-lme_MM_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ), na.action=na.omit )
-
-as.data.frame( round(  summary(lme_MM_IFD_PER)$tTable, 6 ) )
-
-#remove PWD, SKIVE and MSM
-another.DF <- mouse.avs_4MM[mouse.avs_4MM$strain != "PWD",]
-another.DF <- another.DF[another.DF$strain != "MSM",]
-another.DF <- another.DF[another.DF$strain != "SKIVE",]
-
-lme_MM_another_IFD_PER <- lme(mean_IFD.2CO_PER ~ subsp * sex, data= another.DF, random=list(strain=pdDiag(~sex) ), na.action=na.omit )
-
-as.data.frame( round(  summary(lme_MM_another_IFD_PER)$tTable, 6 ) )
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+anova(lmer.IFD.PER_MM.M1)
+
+exactRLRT(lmer.IFD.PER_MM.M1)
+as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
+
+#M1_IFD.RAW_2CO
+lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
+anova(lmer.IFD.RAW_MM.M1) # also F values
+
+exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
+as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
 
@@
 
-#### Random Effects
-
-These second set of models are made using the lmer()
-
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-$$mouse \ average \ normalized IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure. The tables below are from anova( for the lmer model )
-
-```{r MM.IFD, echo=FALSE}
-#ugg what changed, FFF kntir is broken!!
-
-IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(IFD_ABS.MM)
-
-#thing1  <- summary(IFD_ABS.MM)
-IFD.ABS_chi <- drop1(IFD_ABS.MM, test="Chisq")
-
-rand_test_IFD.ABS.MM <- exactRLRT(IFD_ABS.MM)
-rand_test_IFD.ABS.MM
-
-IFD_PER.MM <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(IFD_PER.MM)
-
-Chis.sp_IFD.PER.MM <- drop1(IFD_PER.MM, test="Chisq") 
-#Chis.sp_IFD.PER.MM
-
-rand_test_IFD.PER.MM <- exactRLRT(IFD_PER.MM)
-rand_test_IFD.PER.MM
-```
-
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
+The tables below are from anova( for the lmer model ).
 Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
@@
 Dive deeper into the sex specific pattern for each strain.
-
-these are example strains that show the 'unusual' results.
-show a raw plot, and norm plot and a scatter plot.
-Could the female bivalents with IFD under 30% be a different class of interference?
-(What does that mean?).
-I think the General pattern is that this rec pattern is not observed in males.
-That is there are rarely instances where 2 foci are positioned closer than 30% of the total SC length.
+Below are code chunks which show the unusal sex specific results for IFD measures.
+The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
+The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
+
+For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
 
@@
 
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
 #plot the metric
@@
 #make a table of the means and n ect
-
 G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
@@
 G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+#calQ percent < 30%
+
+oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
+oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
+
+#the na's are being counted (ughhhhhh)
+#use the file thing instead
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>% filter(IFD1_PER <= .3) #this makes DF
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male")
+ratio.male <- 18/1031 #2%
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>% filter(IFD1_PER <= .3)
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")
+ratio.female <- 36/766  #5%
+
+LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
+  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
+
+#uggh how does this not work ...
+
+#more males have lower than 30% IFD?
+
+LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
+#0.7062663
+
+#would Chi-sq be a good test to compare the ratios ect?
+
+#the sum of these doesn't add
+
+G.Bivalents.DF
 
@@
 
+# old junk
+
 <!-- 
 # old drafts / trash pile
+
+```{r long.biv.MM, echo=FALSE, eval=FALSE}
+
+SC.Length_long.biv_MM <- lme(mean_SC ~ subsp * sex, data= Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex),  na.action=na.omit  ) )
+
+#now there is an error after un-ordering them
+
+as.data.frame( round(  summary(SC.Length_long.biv_MM)$tTable, 6 ) )
+
+#model for long Bivs
+#long_bivData_SC.Length_MM <- lme(mean_SC ~ subsp * sex, data= #Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(long_bivData_SC.Length_MM)$tTable, 6 ) )
+```
 
@@
 
+```{r 2CO.MM, echo=FALSE, eval=FALSE}
+
+#remove MOLF
+mouse.avs_4MM.99 <- mouse.avs_4MM[mouse.avs_4MM$strain != "MOLF",]
+
+SC.Length_2CO_MM <- lme(mean.SC_2CO ~ subsp * sex, 
+                  random=list(strain=pdDiag(~sex),
+                        data= mouse.avs_4MM.99, 
+                                                                                       na.action=na.omit ) )
+#keep getting errors for building this model
+
+as.data.frame( round(  summary(SC.Length_2CO_MM)$tTable, 6 ) )
+
+table(mouse.avs_4MM$strain, mouse.avs_4MM$sex)
+
+```
+
+```{r 3CO.MM, echo=FALSE}
+
+SC.Length_3CO_MM <- lme(mean.SC_3CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),  na.action=na.omit  )
+#this model is also generating error...
+#don't know how to fix
+
+as.data.frame( round(  summary(SC.Length_3CO_MM)$tTable, 6 ) )
+```
+
 -->

### 2020-02-04 doc/Main_Report_Results.rmd
work and notes

@@
 
-load(file = "C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
+#load(file = "C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
+
+load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
 
 #2.2.20
-
-load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
+#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
 
@@
 
-ToDo: add the random effect tests!!
+ToDo: change the Mixed models to lmer() so that I can confidantly test random effect (variance)
 
@@
 
-$$mouse .var CO ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$
+I won't run the model above -- even tho it is logical and makes sense -- I can't make good estimates for random effects.
+
+$$mouse .var CO ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
@@
 
-- ( Nested strain in subsp?)
+(Move this into the Methods)
 
@@
 ```{r mixed.model.data.setup, echo=FALSE, warning=FALSE}
+# MOVE this into BIVDATA SETUP!!!
+
 #https://rcompanion.org/handbook/G_03.html
@@
 
-#reformate Molosinus subsp
+#add_subsp function re-writen
+
 #mean_co.HQ.mol
-DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
-        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))
+#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
+#        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))
 
 # For some reason I changed subsp to nums (wtf)
-DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
-        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))
-
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
+#        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))
+
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | #DF.HetC.MixedModel.HQ$mouse==""), ]
 
@@
 
+USE the bottom one
+
 2. $$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
@@
 
+<!--
+
 > Model 1
@@
 
-```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2}
+```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2, eval=FALSE}
 #table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
@@
 Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.  
-
-> Model 2.
+-->
+
+> Model 2. THIS SHOULD BE THE MAIN MIXED MODEL --lmer
 
@@
 
-```{r MixedModel.2, echo=FALSE}
+```{r MixedModel.1, echo=FALSE, digits = 3}
+#THIS IS THE NEW MIXED MODEL
+
 #reduced models
-Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
 #summary(Reduced.strain.HQ)
 
-as.data.frame(round(summary(Reduced.strain.HQ)$tTable, 6))
+drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***
+
+exactRLRT(MLH1.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0
+
+as.data.frame(round(summary(MLH1.MM.M1)$tTable, 6))
 #add a aption to this code chunk
@@
 
+Above is the mixed model M1,
+
 Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
@@
 
+All fixed effects with all interactions.
+
 $$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
@@
 ```{r MixedModel.3, echo=FALSE}
+
 #strain = fixed

### 2020-02-04 src/BivData_Setup.Rmd
work and notes

@@
   #print(c(i, " the i") )
-
   newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
@@
 Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
+
+#remove na's
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]
 

### 2020-02-04 src/SetupData.Rmd
too much work from home
work and notes

@@
 
+- code Mol as discint subsp
+
 - make sure to remove some of the juvi mice!! from the AP_mouse table
+
+- format the DF for applying to Mixed models
 
@@
                   "3sep18_AST_m1",
-                  "3jel18_TOM_m1","3sep18_TOM_m1",
+                  "3jul18_TOM_m1","3sep18_TOM_m1",
                   "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
@@
 
+#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
+#juvi.mice_DF <- juvi.mice_DF[!(is.na(juvi.mice_DF$mouse) | juvi.mice_DF$mouse==""), ]
+
+#juvi mice ~ 2 weeks old
+juvi_mice <- c(
+  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
+  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
+  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
+
+  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
+   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
+  )
+
+#make juvi.mice DF
+juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice),]
+
+#remove juvi mice from big MLH1 -- but Then i need to remake the mouse_table
+MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice, ]
+
+# > 25 weeks
+# I might need to keep some -- espcially outgroups
+#old mice -- This list includes mice which don't have MLH1 observations, but I have records of dissecting them
+
+old.mice <- c(
+  "12oct15_WSB_m1","9oct17_WSB_m1","1mar18_WSB_m2","21dec14_WSB_m2","21dec14_WSB_m3", "1mar18_WSB_m1",
+  "14mar15_G_m1","14mar15_G_m2","24mar15_G_m1","19oct15_G_m1",
+  "8may17_LEW_m1","18sep17_LEW_m1","8may17_LEW_m2","20dec16_LEW_m2",
+  "5jul17_PERC_m1",
+
+  "27oct18_SKIVE_m2","4feb19_SKIVE_m1",
+  "1feb18_KAZ_m1","11jan18_KAZ_m1","19may18_KAZ_m2","9apr18_KAZ_m1","9apr18_KAZ_m2",
+  "20dec18_CZECH_m1","3dec18_CZECH_m1",
+  "28jun18_TOM_m1","3jul18_TOM_m1",
+  "27aug18_AST_m1","7aug18_AST_m1",
+
+  "13nov16_MSM_m1","13nov16_MSM_m2","4feb19_MSM_m3","4feb19_MSM_m4","9oct17_MSM_m2","4feb19_MSM_m2","30may17_MSM_m1","5jul17_MSM_m1",
+  "4feb19_MOLF_m1",
+
+  "24aug17_CAST_m1","23sep17_CAST_m1","23sep17_CAST_m2","30jun16_CAST_m3","5jul17_CAST_m1",
+  "27aug18_HMI_m2","28jun18_HMI_m1","27aug18_HMI_m1","4dec17_HMI_m1",
+
+  "11jan18_SPRET_m1","27aug18_SPRET_m1","7aug18_SPRET_m2","7aug18_SPRET_m1","13jul17_SPRET_m1","2aug18_SPRET_m1",
+  "19may18_SPIC_m1","19may18_SPIC_m2","30oct17_SPIC_m1","11jan18_SPIC_m1",
+  "1mar18_CAROLI_m1","9apr18_CAROLI_m1"
+)
+
 ```
+
+I have made a list of the old mice, but from the effect document the older mice might not be that different.
 
@@
 ```{r reorder.DF, echo=FALSE}
+#this is most important for plotting
+
 #reorder dataframe
@@
 
-Make mouse level tables
+Make mouse level tables  (re think this name, AP to distinguish BD -- but this isn't really an issue)
+
+```{r all.data.table, echo=FALSE}
+
+#make the table with ALL MLH1 obs
+
+FULL_mouse_table_MLH1 <- ddply(MLH1_data, c("mouse"), summarise,
+
+                             Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+
+)
+
+```
 
@@
 
-#Make a mouse level table
-AP_mouse_table.HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
+#remove juvi mice
+
+#Make a mouse level table -- why are there many old mice without any MLH1 (bc these are mice which were dissected, but not imaged/quantified)
+
+AP_mouse_table.MLH1_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
 
-AP_mouse_table.HQ <- add_strain(AP_mouse_table.HQ)
-AP_mouse_table.HQ <- add_subsp(AP_mouse_table.HQ)
-AP_mouse_table.HQ <- add_sex(AP_mouse_table.HQ)
-AP_mouse_table.HQ <- add_category(AP_mouse_table.HQ)
-
-#make sure they are all factors
-AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
-AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
-AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
-AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)
+AP_mouse_table.MLH1_HQ <- add_strain(AP_mouse_table.MLH1_HQ)
+AP_mouse_table.MLH1_HQ <- add_subsp(AP_mouse_table.MLH1_HQ)
+AP_mouse_table.MLH1_HQ <- add_sex(AP_mouse_table.MLH1_HQ)
+AP_mouse_table.MLH1_HQ <- add_category(AP_mouse_table.MLH1_HQ)
 
 #add DOB to this table?
-mice.frm.meta <- AP_mouse_table.HQ[(AP_mouse_table.HQ$mouse %in% meta.data$mouse),]
+mice.frm.meta <- AP_mouse_table.MLH1_HQ[(AP_mouse_table.MLH1_HQ$mouse %in% meta.data$mouse),]
 
@@
 
-#merge with DOB
-
-#why are there 2
-#mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = #"mouse", by.x = "Var1", all = TRUE)
-
-AP_mouse_table_w.Ages <- merge(AP_mouse_table.HQ, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+#during merging -- try to just merge a few cols: 
+AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
 #what are the differences between these two
 
-juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
+#remove mice without MLH1 obs
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice == ""), ]
 
-juvi.mice_DF <- juvi.mice[!(is.na(juvi.mice$mouse) | juvi.mice$mouse==""), ]
+#remove cols with bad suffix
+AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x, strain.y,
+                    mouse.1  )  )
+
+#add all the cols back
+AP_mouse_table_w.Ages <- add_strain(AP_mouse_table_w.Ages)
+AP_mouse_table_w.Ages <- add_subsp(AP_mouse_table_w.Ages)
+AP_mouse_table_w.Ages <- add_sex(AP_mouse_table_w.Ages)
+AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+
+#make sure they are all factors
+AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
+AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
+AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
+AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)
+
+old.mice.df <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.weeks > 25),]
+#remove the NAs
+old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]
 
 #remove bad ages
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[( (AP_mouse_table_w.Ages$age.weeks < 25) &  (AP_mouse_table_w.Ages$age.days != 13) ) | is.na(AP_mouse_table_w.Ages$age.weeks), ]
-
-#remove juvi mice
-MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi.mice_DF$mouse, ]
-
-#remove the NAs
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$category.x) | AP_mouse_table_w.Ages$category.x==""), ]
-
+#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[( (AP_mouse_table_w.Ages$age.weeks < 25) &  (AP_mouse_table_w.Ages$age.days != 13) ) | #is.na(AP_mouse_table_w.Ages$age.weeks), ]
 ```
 
-```{r Q1Q2.mouse.table, echo=FALSE}
+Made up the main mouse level table:
+
+- (Q1-Q4)
+
+- no juvi mice
+
+Subset DFs from here of other analysis.
+
+```{r Q12_mouse.table, echo=FALSE}
+
+#This DF is used for some MM, apply the HetC MM certia...
+#this is a DF with a different set of quality certeria
+
 #Make a mouse level table
-Q12_mouse_table <- ddply(MLH1_data[MLH1_data$quality < 3,], c("mouse"), summarise,
+Q12_mouse_table_MLH1 <- ddply(MLH1_data[MLH1_data$quality < 3,], c("mouse"), summarise,
 
@@
 #REMOVE mice with fewer than 3 cells
-Q12_mouse_table <- Q12_mouse_table[Q12_mouse_table$Ncells > 3,]
+Q12_mouse_table_MLH1 <- Q12_mouse_table_MLH1[Q12_mouse_table_MLH1$Ncells > 3,]
 
-Q12_mouse_table <- add_strain(Q12_mouse_table)
-Q12_mouse_table <- add_subsp(Q12_mouse_table)
-Q12_mouse_table <- add_sex(Q12_mouse_table)
-Q12_mouse_table <- add_category(Q12_mouse_table)
+Q12_mouse_table_MLH1 <- add_strain(Q12_mouse_table_MLH1)
+Q12_mouse_table_MLH1 <- add_subsp(Q12_mouse_table_MLH1)
+Q12_mouse_table_MLH1 <- add_sex(Q12_mouse_table_MLH1)
+Q12_mouse_table_MLH1 <- add_category(Q12_mouse_table_MLH1)
 
@@
 #MERGE HERE
-Q12_mouse_table_w.Ages <- merge(Q12_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+Q12_mouse_table_w.Ages <- merge(Q12_mouse_table_MLH1, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
 #what are the differences between these two
@@
 
-#MOL = subsp
-Q12_mouse_table_w.Ages$subsp <- ifelse(grepl("MSM", Q12_mouse_table_w.Ages$strain), "Mol", 
-        ifelse(grepl("MOLF", Q12_mouse_table_w.Ages$strain), "Mol", Q12_mouse_table_w.Ages$subsp))
-
-#unclass order factors
+#unclass order factors --(applies to the models)
 #main df with variables coded as chr
@@
 
-```{r strain.table, echo=FALSE}
+### DataFrame for Mixed Models
+
+Below I format the DF's for the models;
+
+i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus. start with **AP_mouse_table**
+
+```{r HetC.DF.4MM, echo=FALSE}
+
+#for the DF for use in the HetC Mixed Model
+#Which Start with the AP_mouse_table
+
+#variance is still chr
+AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+
+#use the merged DF to filter by age 
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+#remove the extra merged cols
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+#REMOVE STRAINS WITHOUT FEMALES
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+
+#Only us the righ age window
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
+
+#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#remove blank rows (which are these?)
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
+#151 mice, no rows with nas for values
+
+#main df with variables coded as chr
+DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+
+#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+
+#only female male reps
 

### 2020-02-04 doc/Main_Report_Results.rmd
filling in pvalues and making notes for the mouse mlh1 models

@@
 
-```{r mixed.model.data.setup, echo=FALSE, warning=FALSE}
+```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=FALSE}
 # MOVE this into BIVDATA SETUP!!!
+
+#should be able to use the loaded 
+#DF.HetC.MixedModel.HQ
 
@@
 #afriad this is removing NA's
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
 
@@
 #THIS IS THE NEW MIXED MODEL
-
-#reduced models
+#should be able to use the loaded 
+#DF.HetC.MixedModel.HQ
+
 MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
@@
 
+summary(MLH1.MM.M1)
 drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***
+#only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
+anova(MLH1.MM.M1)
 
@@
 
-as.data.frame(round(summary(MLH1.MM.M1)$tTable, 6))
+as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
 #add a aption to this code chunk
@@
 
-## Variance in CO count
-
-DECIDE WHICH RESULTS TO DISPLAY.
-
+## Within Mouse Variance in CO Count per Cell
+
+<!-- intro
 in addition to the mean gwRR, the variance is also an important metric to describe
@@
 meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-
-#### MLH1 variance Figures
-
-Not sure why I have this section -- think this should be into a supplemental doc or something.
-
-```{r variance.boxplots, echo=FALSE, eval=FALSE}
-#how did I make this 
-#Species.COcount.mixed <- lmer(nMLH1.foci ~ species + quality + (1|mouse), data=MLH1.hq.lab, REML = TRUE)
-#Species.COcount.mixed.sum <- summary(Species.COcount.mixed)
-#SpeciesCOcount.fix.est <- drop1(Species.COcount.mixed, test="Chisq") 
-#SpeciesCOcount.Rand.est <- exactRLRT(Species.COcount.mixed)
-
-#plot of variance
-var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
-var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
-#this plot shows variance as points/single values
-
-#plot by sex / seperately
-# HQ data DF.HetC.MixedModel.HQ
-
-var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-var.male <- var.male   + facet_wrap( ~ strain, scales="free")
-
-cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
- ggtitle("Male cV levels") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
-  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )
-
-cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")
-
-```
+-->
+
+Histograms might not be the best way to represent, but they show the general pattern than female mice have higher within mouse variation for CO counts.
 
 ```{r variance.hist, echo=FALSE, eval=FALSE}
+#figures of histograms,
 
@@
 Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
-Var.Hist
 
@@
 Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
-Var.Hist.15cells
+
+Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
+
+Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
+Q12.var.hist.plot
+
+#some LEW female still have v high within mouse variance (2 mice, )
 
@@
 A general pattern is that the female data has higher within mouse variance (both var and cV).
-Some strains more than others.
-I think the dissection date needs to be tested as a predictor factor.
-It might be tricky to add the dissection date metric.
-Test if dissection date is more significant predictor than strain (for females)
-Also test how the number of 0CO bivalents per cell affects mouse with variance.
-
+Some strains more than others (invetigate potential outliers).
 (LEW females with top 3 variance are all from different dissection dates)
-
-I don't have enough mice with >15 cells for a complete idea of the distributions, but comparing single mice with enough good quality cells show be
+This pattern also holds for within mouse cv for MLH1 count (not shown).
+
+**todo**
+
+Test if dissection date is more significant predictor than strain (for females) (Effect doc)
+Also test how the number of 0CO bivalents per cell affects mouse with variance. (lm)
 
@@
 cV.Hist
-
-```
-
-### Modeling Variance
+```
+
+### Modeling Within Mouse Variance
 
@@
 
-1.	HetC.MixedModel.HQ <- lme(mean_co ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-2.	Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-
-3.	lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-
-4.	strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$
-
-```{r M1.var, echo=FALSE}
+XXX (old 1.)	HetC.MixedModel.HQ <- lme(mouse.metric ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+1.(new 1)	Reduced.strain.HQ <- lmer(mouse.metric ~ subsp * sex+ (1|strain), data=DF.HetC.MixedModel.HQ)
+
+3.	lm(mouse.metric ~ subsp \* sex \* strain, data=DF.HetC.MixedModel.HQ)
+
+4.	lm(mouse.metric ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+$$mouse metric  ~=~ subsp * sex + rand(strain)+ \varepsilon$$
+
+```{r M1.MLH1.var_cv, echo=FALSE}
 #modeling within mouse variance
@@
 #HQ var.CO
-M1.var_co <- lme( var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-#summary(HetC.MixedModel.HQ.var_co)
-
-as.data.frame( round(summary(M1.var_co)$tTable, 6))
-
-#ranef(HetC.MixedModel.HQ.var_co)
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+summary(MLH1var_M1)
+
+drop1(MLH1var_M1, test="Chisq") #interaction term NS
+#when switched to additive fixed, sex is very sig
+
+anova(MLH1var_M1)#sex explains most variance...
+
+exactRLRT(MLH1var_M1) #strain
+as.data.frame( round(summary(MLH1var_M1)$tTable, 6))
 
 #CV
-M1.cv_co <- lme( cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-#summary(HetC.MixedModel.HQ.var_co)
+MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+summary(MLH1cv_M1)
+
+drop1(MLH1cv_M1, test="Chisq") #interaction term NS
+#when switched to additive fixed, sex is very sig
+anova(MLH1cv_M1)  #sex F value is large
+exactRLRT(MLH1cv_M1) #strain
+
 as.data.frame( round(summary(M1.cv_co)$tTable, 6))
-```
-
-```{r M1.Q12_table, echo=FALSE}
-#use the Q1Q2 calq'd table
-
-#Q12_mouse_table_w.Ages$subsp <- ifelse(grepl(1, Q12_mouse_table_w.Ages$subsp), "Dom", 
-#        ifelse(grepl(2, Q12_mouse_table_w.Ages$subsp), "Musc", Q12_mouse_table_w.Ages$subsp))
-
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | #DF.HetC.MixedModel.HQ$mouse==""), ]
-
-Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
-
-#need to change to chr first
-
-Q12_mouse_table_w.Ages$subsp <- as.character(Q12_mouse_table_w.Ages$subsp)
-
-Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MSM"] <- 'Mol'
-
-Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MOLF"] <- "Mol"
-
+
+#USE Q12 (Q123) table
 #Q12_mouse_table_w.Ages
-
-M1.Q12_var <- lme( var ~ subsp * sex, data=Q12_mouse_table_w.Ages, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(M1.Q12_var)$tTable, 6))
-
-M1.Q12_CV <- lme( cV ~ subsp * sex, data=Q12_mouse_table_w.Ages, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(summary(M1.Q12_CV)$tTable, 6))
-
-```
-
-$$mouse .cV CO ~=~ subsp * sex + rand(1|strain)+ \varepsilon$$
-
-```{r M2.var, echo=FALSE}
-#this model must be the same as M! -- since all the results are the same
-M2.var <- lme(var ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-
-as.data.frame( round(summary(M2.var)$tTable, 6))
-
-#summary(HQjust.strain.var_co)
-
-M2.CV <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
-
-as.data.frame( round(summary(M2.CV)$tTable, 6))
-
-```
-
-```{r M2.var.Q12, echo=FALSE}
-M2.Q12_var_co <- lme(var ~ subsp * sex, data=Q12_mouse_table_w.Ages, random = ~1|strain)
-
-as.data.frame( round(summary(M2.Q12_var_co)$tTable, 6))
-
-M2.Q12_CV <- lme(cV ~ subsp * sex, data=Q12_mouse_table_w.Ages, random = ~1|strain)
-
-as.data.frame( round(summary(M2.Q12_CV)$tTable, 6))
-```
-
-$$mouse .var CO ~=~ subsp * sex  + \varepsilon$$
-```{r M3.var, echo=FALSE, eval=FALSE}
-M3.var <- lm(var ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-as.data.frame( round(summary(M3.var)$tTable, 6))
-
-M3.cv <- lm(cV ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-as.data.frame( round(summary(M3.cv)$tTable, 6))
-#these models won't work
-```
-
-```{r M3VAR.Q12, echo=FALSE, eval=FALSE}
-#ugh, don't know how to get these to work 
-M3.Q12_var <- glm(var ~ subsp * sex * strain, data=Q12_mouse_table_w.Ages)
-
-as.data.frame(summary(M3.Q12_var)$tTable, 6)
-#results don't show
-
-M3.Q12_cv <- lm(cv ~ subsp * sex * strain, data=Q12_mouse_table_w.Ages)
-
-as.data.frame( round(summary(M3.Q12_cv)$tTable, 6))
-```
-
-```{r M4.var, echo=FALSE, eval=FALSE}
-M4.var <- lm(var ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-as.data.frame( round(summary(M4.var)$tTable, 6))
-
-M4.cv <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-as.data.frame( round(summary(M4.cv)$tTable, 6))
-```
-
-```{r M4.var_Q12, echo=FALSE, eval=FALSE}
-
-M4.var_Q12 <- lm(var ~ sex * strain, data=Q12_mouse_table_w.Ages)
-as.data.frame( round(summary(M4.var_Q12)$tTable, 6))
-
-#these results not print / working
-M4.CV_Q12 <- lm(cV ~ sex * strain, data=Q12_mouse_table_w.Ages)
-as.data.frame( round(summary(M4.CV_Q12)$tTable, 6))
-```
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain * sex) + \varepsilon$$
-
-```{r var.contrast, echo=FALSE, eval=FALSE}
-#contrast models -- what's the points for this block
-# review this logic again
-full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
-
-full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
-
-strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
-#for var, p = 0.0219
-```
-
-#### Summary variance
-
-I couldn't get many of the models to return results.
+#Q12 var
+MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+
+drop1(MLH1var_M1_Q12, test="Chisq")#sig
+anova(MLH1var_M1_Q12)
+exactRLRT(MLH1var_M1_Q12)#NS
+
+as.data.frame( round(summary(MLH1var_M1_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+
+drop1(MLH1cv_M1_Q12, test="Chisq")#NS
+anova(MLH1cv_M1_Q12)#sex large F
+exactRLRT(MLH1cv_M1_Q12)#NS
+
+as.data.frame( round(summary(MLH1cv_M1_Q12)$coefficients, 6))
+
+```
+
+The fixed sex effect has the largest effect sizes.
+All are negative (for male).
+(I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
+The strain effects are only slightly significant for within mouse variance measures in the full dataset (includes low quality cells)
+
+$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+M2 slash M3
+
+```{r M3.MLH1.var, echo=FALSE}
+
+MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+summary(MLH1var_M3)
+anova(MLH1var_M3)  
+as.data.frame( round(summary(MLH1var_M3)$tTable, 6))
+
+MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+summary(MLH1cv_M3)
+
+anova(MLH1cv_M3)
+as.data.frame( round(summary(M1.cv_co)$tTable, 6))
+
+#Q12 var
+MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+summary(MLH1var_M3_Q12)
+anova(MLH1var_M3_Q12)  
+as.data.frame( round(summary(MLH1var_M3_Q12)$tTable, 6))
+
+#Q12 cv
+MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+summary(MLH1cv_M3_Q12)
+anova(MLH1cv_M3_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
+
+```
+
+For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
+Most of the strain specific differences in within mouse variance go away in the Q12 dataset.
+
+$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+```{r M4.MLH1.var, echo=FALSE}
+
+MLH1var_M4 <- lm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+summary(MLH1var_M4)
+
+drop1(MLH1var_M4, test="Chisq")
+anova(MLH1var_M4)
+as.data.frame( round(summary(MLH1var_M4)$tTable, 6))
+
+MLH1cv_M4 <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summary(MLH1cv_M4)
+
+drop1(MLH1cv_M4, test="Chisq")
+anova(MLH1cv_M4)  
+exactRLRT(MLH1cv_M4)
+as.data.frame( round(summary(MLH1cv_M4)$tTable, 6))
+
+### Q12 var
+MLH1var_M4_Q12 <- lm( var ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+anova(MLH1var_M4_Q12)
+as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M4_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+summary(MLH1cv_M4_Q12)
+anova(MLH1cv_M4_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
+
+```
+
+#### Summary witin mouse Variance
 
@@
 
-When Molossinus is coded as a separate subspecies, then interaction terms are no longer significant.  (subsp *sex )  interaction?
+(some strains have significant effect -- some LEW females had large leverage).
 
@@
 
+```{r var.contrast, echo=FALSE, eval=FALSE}
+#contrast models -- what's the points for this block
+# review this logic again
+full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
+
+full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
+
+strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
+#for var, p = 0.0219
+```
+
 ```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}

### 2020-02-04 doc/Main_Report_Results.rmd
debugging and Cleaning up Main Results file

@@
 exactRLRT(MLH1var_M1) #strain
-as.data.frame( round(summary(MLH1var_M1)$tTable, 6))
+as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))
 
@@
 
-as.data.frame( round(summary(M1.cv_co)$tTable, 6))
+as.data.frame( round(summary(MLH1cv_M1)$coefficients, 6))
 
@@
 anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$tTable, 6))
+as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
 
@@
 anova(MLH1cv_M3)
-as.data.frame( round(summary(M1.cv_co)$tTable, 6))
+as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
 
@@
 anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$tTable, 6))
+as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
 
@@
 anova(MLH1var_M4)
-as.data.frame( round(summary(MLH1var_M4)$tTable, 6))
+as.data.frame( round(summary(MLH1var_M4)$coefficients, 6))
 
@@
 summary(MLH1cv_M4)
-
-drop1(MLH1cv_M4, test="Chisq")
 anova(MLH1cv_M4)  
-exactRLRT(MLH1cv_M4)
-as.data.frame( round(summary(MLH1cv_M4)$tTable, 6))
+as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))
 
 ### Q12 var
-MLH1var_M4_Q12 <- lm( var ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1var_M4_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+
+summary(MLH1var_M4_Q12)
+
 anova(MLH1var_M4_Q12)
@@
 #Q12 cv
-MLH1cv_M4_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1cv_M4_Q12 <- lm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
 
 summary(MLH1cv_M4_Q12)
-anova(MLH1cv_M4_Q12)#sex large F
+anova(MLH1cv_M4_Q12)
 
@@
 
-Main figure, MLH1 strain avs on top, chrm proportion bar plot on bottom
+work more on figure 1. Main figure, MLH1 strain avs on top, chrm proportion bar plot on bottom
 
@@
 
+What are the steps the each chunks walk through.
+
 #### Proportion of chrm classes
@@
 chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#try adding the BivData version of the proportions
 
@@
 
-# Rapid male specific evolution
-
-list out the points which support rapid male specific evolution for MLH1 rates.
-
-number above bars are the raw counts for bivalent classes.
-
-Potential figure 2 (eventually). Try to make the figure taller.
--- add more data to make sample sizes more equal.
--- add space to indicate where female data is missing
-
-SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
-What's up with Lew females? there's lots of 2COs, also WSB female.
-
-This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
-If these data are random samples of bivalents, the X
-
-ToDo: chisq tests for difference in proportions, 
-- within strain across sex,
-- males, across high and low groups
-- look at mouse level proportions
-
-## DMC1 Results Spermatocytes
+# DMC1 Results Spermatocytes
 
@@
 
-```{r DMC1.setup, echo=FALSE, eval=FALSE}
+```{r DMC1.setup, echo=FALSE}
 #load DMC1 csv
-DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
 
@@
 DMC1.mean.obs.Z$stage <- "Z"
-
-#these tables are merged latter
-
-```
-
-### Testing Normality Assumptions
+```
 
 ```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-#histograms for each
+
+#these are histogram plots for testing Normality shape -- move to supplemenatry section
 
@@
 
-```{r DMC1.plot, echo=FALSE, warning=FALSE, eval=FALSE}
+```{r DMC1.plot, echo=FALSE, warning=FALSE}
 #think of better color scheme
@@
   geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")
+
 DMC1.boxplot
@@
 
+Summary:
+
+The distributions of counts is close to normal (also the higher count is better than the MLH1 counts)
+
 Above figure shows p-values for comparison of means across the cell stages.
@@
 
-```{r DMC1.subsampling, echo=FALSE, eval=FALSE}
+```{r DMC1.subsampling, echo=FALSE}
 
@@
 
-plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-plot.pvals.L
-plot.pvals.Z
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
+#plot.pvals.L
+#plot.pvals.Z
 
@@
 
-```{r DMC1_MLH1.cor, echo=FALSE, eval=FALSE}
+```{r DMC1_MLH1.cor, echo=FALSE}
 #make category averages from AP_mouse_table.HQ
@@
 
-The correlation with MLH1 and leptotene cells is r MLH1.DMC1.L_cor.
-
-The correlation with MLH1 and zygotene cells is r MLH1.DMC1.Z_cor.
-
-# Single Bivalent Patterns
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+
+```{r barlett.test, echo=FALSE, eval=FALSE}
+
+#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
+bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
+
+```
+
+Evidence for non-equal variance across strains for zygotene cells.
+
+Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
+
+From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#L vs Z (all species)
+#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
+#p-value = 1.032e-05
+#not a valid interpertation
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = r ttest.HighLow.L$p.value) while the Z cells, there is not a significant difference across the high and low groups (value = r ttest.HighLow.Z$p.value).
 
 <!--
+ Single Bivalent Patterns
 
@@
 -->
+
+# Rapid male specific evolution
+
+list out the points which support rapid male specific evolution for MLH1 rates.
+
+number above bars are the raw counts for bivalent classes.
+
+Potential figure 2 (eventually). Try to make the figure taller.
+-- add more data to make sample sizes more equal.
+-- add space to indicate where female data is missing
+
+SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
+What's up with Lew females? there's lots of 2COs, also WSB female.
+
+This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
+If these data are random samples of bivalents, the X
+
+ToDo: chisq tests for difference in proportions, 
+- within strain across sex,
+- males, across high and low groups
+- look at mouse level proportions
 
@@
 
-# Blank
+# Trash / Extra
+
+<!--
 
@@
 
-```{r barlett.test, echo=FALSE, eval=FALSE}
-
-#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
-bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
-
-```
-
-Evidence for non-equal variance across strains for zygotene cells.
-
-Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
-
-From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova <- anova(mod.L) #0.0002701 ***
-z.anova <- anova(mod.Z) #NS
-
-l.anova
-z.anova
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE, eval=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#L vs Z (all species)
-#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
-#p-value = 1.032e-05
-#not a valid interpertation
-
-#high vs low group (L)
-ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = r ttest.HighLow.L$p.value) while the Z cells, there is not a significant difference across the high and low groups (value = r ttest.HighLow.Z$p.value).
-
-# Trash / Extra
-
-<!--
-
 ```{r var.contrast, echo=FALSE, eval=FALSE}

### 2020-02-04 doc/BivPattern.Rmd
summarizing all results, updating saving files

---
title: "BivData Patterns"
author: "April Peterson"
date: "October 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup.data, echo=FALSE}

library(ggplot2)

setwd("~./MLH1repo/")
#load curated data.frame
load(file="data/BivData_4_PCA.RData")
```

first discribe data (table functions)

Test everything from chrms of the same class across strains
How similar are 1COs bivalents in MSM,PWD, and KAZ, LEW ect?
(male and female?)

1CO bivalents;

> position
> centromere position
> SC length

```{r blah}

master_curated_BivData <- curated.oneCO.male[!(is.na(curated.oneCO.male$fileName) | curated.oneCO.male$fileName ==""), ]
master_curated_BivData <- master_curated_BivData[master_curated_BivData$hand.foci.count != 8,]

all1CO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 1,]
all1CO <- all1CO[!(is.na(all1CO$fileName) | all1CO$fileName ==""), ]

blahblah <- ggplot(all1CO, aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))

all2CO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 2,]
all2CO <- all2CO[!(is.na(all2CO$fileName) | all2CO$fileName ==""), ]

blahblah2 <- ggplot(all2CO, aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("2COs male and females")+ ylim(c(30,250))

## (try IFD)
staggered <- ggplot(master_curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("")+theme(legend.position="none") #+ ylim(c(30,250))

fill.by.ifd <- ggplot(master_curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=IFD1_PER))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("SC length Fill by IFD")#+theme(legend.position="none") #+ ylim(c(30,250))

ifd.rand <- ggplot(master_curated_BivData, aes(y = IFD1_PER, x=hand.foci.count, color=strain))+geom_jitter(aes(alpha=0.2))+ facet_wrap(~sex)+ ggtitle("SC length Fill by IFD")#+theme(legend.position="none") #+ ylim(c(30,250))

IFD.comp <- ggplot(master_curated_BivData, aes(y = IFD1_PER, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("IFD1 distributions")+theme(legend.position="none")

```

```{r contrast.chrm.class, echo=FALSE}

curated.oneCO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 1,]
curated.oneCO.male <- curated.oneCO[curated.oneCO$sex == "male",]

curated.oneCO.male$Foci1.PER <- curated.oneCO.male$Foci1 / curated.oneCO.male$chromosomeLength

curated.oneCO.male <- curated.oneCO.male[!(is.na(curated.oneCO.male$fileName) | curated.oneCO.male$fileName ==""), ]

#lm - but what trait do i want to predict
#one.co.lm <- lm(data = curated.oneCO.male)

#ggplots
oocc <- ggplot(curated.oneCO.male, aes(y = dis.cent, x = Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)+ylim(c(0,120))

mabeh <- ggplot(curated.oneCO.male, aes(x= dis.cent, y =chromosomeLength ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)#+ ylim(c(0,125))+xlim(c(30,150))

ghost <- ggplot(curated.oneCO.male, aes(y = chromosomeLength, x =Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)+ylim(c(30,150))

ghost.ssdgm <- ggplot(curated.oneCO.male, aes(y = chromosomeLength, x = Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ylim(c(30,150))

Length.1CO <- ggplot(curated.oneCO.male, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter()+
  ggtitle("1CO chromosome Length") #+facet_wrap(~strain) #+ylim(c(0,150))
#check the outlier

oocc
mabeh
ghost
```

KAZ has more terminal 1CO. 
are 1COs longer?

MSM has higher mean, the PWD and MSM have stronger floor. LEW has stronger ceiling?

MSM and PWD have long chromosome length. (and stronger floor - what is the biological significance of the floor?)

predict that rank would have strata. (there is true bivalent rank and realized rank)

```{r two.co.biv, echo=FALSE}

curated.twoCO <- master_curated_BivData[master_curated_BivData$hand.foci.count == 2,]
curated.twoCO.male <- curated.twoCO[curated.twoCO$sex == "male",]

curated.twoCO.male$Foci1.PER <- curated.twoCO.male$Foci1 / curated.twoCO.male$chromosomeLength
curated.twoCO.male$Foci2.PER <- curated.twoCO.male$Foci2 / curated.twoCO.male$chromosomeLength

curated.twoCO.male <- curated.twoCO.male[!(is.na(curated.twoCO.male$fileName) | curated.twoCO.male$fileName ==""), ]

#ggplots
ttoo.oo <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x = IFD1 ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ylim(c(0,150))

per.ifd.oo <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter()+
  ggtitle("2CO chromosome Length") #+facet_wrap(~strain) #+ylim(c(0,150))

mabeh.two <- ggplot(curated.twoCO.male, aes(x = dis.cent, y =chromosomeLength ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ ylim(c(0,125))+xlim(c(30,120))

ghost.two <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x = Foci1.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain)+ylim(c(30,150))

ghost.two2 <- ggplot(curated.twoCO.male, aes(y = chromosomeLength, x = Foci2.PER ))+geom_jitter()+
  ggtitle("")+facet_wrap(~strain) #+ylim(c(30,150))

```

there seems to be less variance in the MSM and PWD IFD percent. the lengths are similar, but the PER.IFDs are tighter.
Also, in PWD and MSM there are stronger 'floors' for the 2CO bivalents

```{r female.patterns, echo=FALSE}

curated.female.bivs <- master_curated_BivData[master_curated_BivData$sex == "female",]
curated.female.1CO <- curated.female.bivs[curated.female.bivs$hand.foci.count == 1,]
curated.female.2CO <- curated.female.bivs[curated.female.bivs$hand.foci.count == 2,]

curated.female.1CO <- curated.female.1CO[!(is.na(curated.female.1CO$fileName) | curated.female.1CO$fileName ==""), ]
curated.female.2CO <- curated.female.2CO[!(is.na(curated.female.2CO$fileName) | curated.female.2CO$fileName ==""), ]

curated.female.1CO$dist.cent.per <- curated.female.1CO$dis.cent / curated.female.1CO$chromosomeLength
curated.female.2CO$dist.cent.per <- curated.female.2CO$dis.cent / curated.female.2CO$chromosomeLength

Length.1CO.female <- ggplot(curated.female.1CO, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(color=dis.cent))+
  ggtitle("1CO chromosome Length Female")

Length.2CO.female <- ggplot(curated.female.2CO, aes(y = chromosomeLength, x=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(color=dist.cent.per))+
  ggtitle("2CO chromosome Length Female")

```

weaker basements? (how can I test) violin plots?
KAZ and LEW have short 1CO bivalents.

Lew shorter 2CO chrms, 
weaker basments in 2CO chrms

within category variation

```{r kaz.male, echo=FALSE}
#master_curated_BivData
#KAZ males
kaz.male <- master_curated_BivData[master_curated_BivData$category == "KAZ male",]
#3 mice, 73 cells

kaz.male.1CO <- kaz.male[kaz.male$hand.foci.count == 1,]

first.model <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=kaz.male.1CO)
summary(first.model)
#1CO foci position is driven by SC length

kaz.male.2CO <- kaz.male[kaz.male$hand.foci.count == 2,]
kaz.male.2CO <- kaz.male.2CO[!(is.na(kaz.male.2CO$fileName) | kaz.male.2CO$fileName==""), ]

two.model <- lm(Foci2 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+Foci1,  data=kaz.male.2CO)
summary(two.model)
#Foci1 not predictable, but foci2 position (raw), predicted by chrm length and mouse

SC.length.lm <- lm(chromosomeLength ~ mouse +centromere_ABS_Position+centromere_PER_Position+avergeRED+Foci1,  data=kaz.male.1CO)
summary(SC.length.lm)

SC.length.lm2 <- lm(chromosomeLength ~  IFD1_ABS+centromere_ABS_Position+Foci1+Foci2+dis.cent+IFD1,  data=kaz.male.2CO)
summary(SC.length.lm2)
#the significance depends on the order, 
#centromere position is significant for predicting SC length

length(unique(kaz.male$fileName))

```

```{r KAZ.female}

kaz.female <- master_curated_BivData[master_curated_BivData$category == "KAZ female",]
kaz.female.1CO <- kaz.female[kaz.female$hand.foci.count == 1,]
kaz.female.2CO <- kaz.female[kaz.female$hand.foci.count == 2,]

kaz.female.1CO <- kaz.female.1CO[!(is.na(kaz.female.1CO$fileName) | kaz.female.1CO$fileName==""), ]
kaz.female.2CO <- kaz.female.2CO[!(is.na(kaz.female.2CO$fileName) | kaz.female.2CO$fileName==""), ]

first.model <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=kaz.female.1CO)
summary(first.model)
#chrm length is strongest predictor

two.model <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+IFD1+Foci2,  data=kaz.female.2CO)
summary(two.model)
#chrm length and foci1, predict foci2, and IFD1
#to predict foci1, IFD, chrm length, and foci 2 are strong predictors

length.model <- lm(chromosomeLength ~ mouse +Foci1 +centromere_ABS_Position+centromere_PER_Position+avergeRED+hand.foci.count,  data=kaz.female)
summary(length.model)
#centromere position is sig, IFD1, FOci1 is sig -- this is true for 1 and 2CO
#for all chrms, hand.foci.count is significant

```

```{r MSM.male, echo=FALSE}
#MSM male
MSM.male <- master_curated_BivData[master_curated_BivData$category == "MSM male",]
droplevels(MSM.male)
MSM.male.1CO <- MSM.male[MSM.male$hand.foci.count == 1,]
MSM.male.2CO <- MSM.male[MSM.male$hand.foci.count == 2,]

MSM.male.1CO <- MSM.male.1CO[!(is.na(MSM.male.1CO$fileName) | MSM.male.1CO$fileName==""), ]
MSM.male.2CO <- MSM.male.2CO[!(is.na(MSM.male.2CO$fileName) | MSM.male.2CO$fileName==""), ]

first.mo <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=MSM.male.1CO)
summary(first.mo)
#chrm length is most sig

two.mo <- lm(Foci2 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+IFD1+Foci1+SisCoTen,  data=MSM.male.2CO)
summary(two.mo)
#one of the males is sig, chrm length and IFD1, foci2 are signicant

length.mo <- lm(chromosomeLength ~ centromere_ABS_Position+centromere_PER_Position+avergeRED+hand.foci.count+mouse,  data=MSM.male)
summary(length.mo)

#foci1, and centromere positions are sig for 2CO
#centromeres sig
#for all bivs, centromere and hand.foci are the sig factors

```

If all bivalents included for a strain, and sex included, sex will be significant for chrm (for foci as well?) -- what does that add -- 
models don't predict foci, centromere doesn't predict.
I think the chromosome length predicted by centromere position might be related to resolution.

```{r MSM.female}

MSM.female <- master_curated_BivData[master_curated_BivData$category == "MSM female",]
droplevels(MSM.female)
MSM.female.1CO <- MSM.female[MSM.female$hand.foci.count == 1,]
MSM.female.2CO <- MSM.female[MSM.female$hand.foci.count == 2,]

MSM.female.1CO <- MSM.female.1CO[!(is.na(MSM.female.1CO$fileName) | MSM.female.1CO$fileName==""), ]
MSM.female.2CO <- MSM.female.2CO[!(is.na(MSM.female.2CO$fileName) | MSM.female.2CO$fileName==""), ]

first.mo <- lm(Foci1 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED,  data=MSM.female.1CO)
summary(first.mo)
#chrm length is most sig

two.mo <- lm(Foci2 ~ mouse  + chromosomeLength+centromere_ABS_Position+centromere_PER_Position+avergeRED+IFD1+Foci1+SisCoTen,  data=MSM.female.2CO)
summary(two.mo)
#one of the males is sig, chrm length and IFD1, foci2 are signicant

length.mo <- lm(chromosomeLength ~ centromere_ABS_Position+centromere_PER_Position+avergeRED+hand.foci.count+mouse,  data=MSM.female)
summary(length.mo)
#for chrm length in 2COs, intercept, and centromere postions are sig (hand foci count is not significant if i only use 1 chrm class), it is sig for all bivs
#1 of the mice is significant

```

what does a significant intercept for foci1 mean? (it's different than 0?)

```{r melt.foci.positions, echo=FALSE}

```

### 2020-02-04 doc/Main_Report_Results.rmd
summarizing all results, updating saving files

@@
 ---
-
-# Results
 
@@
 
-Details of data set cleaning etc is in SetupData.rmd
-Effects from different factors are reviewed in Effects.Rmd
+<!--  insert Details of data set cleaning etc is in SetupData.rmd
+Effects from different factors are reviewed in Effects.Rmd -->
 
@@
 
-# A. Models for av MLH1 counts foci per cell
-
-ToDo: change the Mixed models to lmer() so that I can confidantly test random effect (variance)
+# Models Framework
 
@@
 
-$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)  + \varepsilon$$
-
-I won't run the model above -- even tho it is logical and makes sense -- I can't make good estimates for random effects.
-
-$$mouse .var CO ~=~ subsp * sex + rand(strain) + \varepsilon $$
+$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
@@
 
-- Unorder factors
-
-- male ages within good range (5 - 12? weeks)
-
-- Mol as a different subsp
-
-(Move this into the Methods)
-
-*The mouse average table used in this part excluded Q5 cells*
-
-*the MLH1 mean is from mean - adj_MLH1 (males have +1 CO to count for PAR)*
+- Male gwRR have been adjusted for the PAR
+
+- Q5 excluded
+
+- All the effects are unorder factors
+
+- male ages within 5 - 20 weeks
+
+- Molossinus is coded as a subsp level
 
@@
 
-subspecies levels: 2(3), Musc and Domestics and Molossinus
-
-strain levels: Don: 3, Musc: 4  Mol: 2
+subspecies levels: 3, Musc, Dom and Molossinus
+
+strain levels: Dom: 3, Musc: 3  Mol: 2
 
@@
 
-*mouse averages exclude Q5 and male counts are adjusted +1 for PAR)*
-
-## Mean CO Count
-
-todo, add the random effect tests, consider changing all to lmer() (those have better random effects)
-
-For now I will focus on 4 different models run mouse averages on MLH1 counts.
-
-1. $$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
-
-USE the bottom one
-
-2. $$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-3. $$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
-
-4. $$mouse .mean CO ~=~ sex * strain + \varepsilon$$
-
-<!--
-
-> Model 1
-
-$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$
-
-1.	HetC.MixedModel.HQ <- lme(mean_co ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-```{r M1_HQ.MixedModel.1, echo=FALSE, digits = 2, eval=FALSE}
-#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-#mean_co.HQ  (this might remove some of the MOLF males)
-
-HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )
-
-#ranef(HetC.MixedModel.HQ)
-#which is the test for random effects ?!
-```
-
-Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.  
--->
-
-> Model 2. THIS SHOULD BE THE MAIN MIXED MODEL --lmer
+For now I will focus on 3 different models for comparing the subsp, sex and strain effects across relevant dependant variables.
+
+1. $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+Allows random effect from strain. Considers the wild derived inbred strains as random samples from each subspecies territory.
+
+2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
+
+Post hoc model for investigating strain as a fixed effect.
+
+3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$
+
+Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.
+
+## MLH1 count M1
 
@@
 
-```{r MixedModel.1, echo=FALSE, digits = 3}
+```{r MLH1co.av.M1, echo=FALSE, digits = 3}
 #THIS IS THE NEW MIXED MODEL
@@
 
-summary(MLH1.MM.M1)
+#summary(MLH1.MM.M1)
 drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***
+
+as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
+
 #only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
@@
 
-as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
-#add a aption to this code chunk
-#ranef(Reduced.strain.HQ)
 ```
@@
 
-Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
-This is model has sex and interaction coefficient is significant.
-
-Can the above model be re-written with lmer so that random effect can be tested?
-
-> Model 3
-
-All fixed effects with all interactions.
-
-$$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$
-
-```{r MixedModel.3, echo=FALSE}
-
+## MLH1 count M2
+
+$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$
+
+```{r MixedModel.2, echo=FALSE,  digits = 3}
 #strain = fixed
 lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-#library(xtable)
+
+anova(lm.strain.fixed.full)
 
@@
 
-> Model 4
-
-$$mouse .mean CO ~=~ sex * strain + \varepsilon$$
-
-```{r MixedModel.4, echo=FALSE}
+## MLH1 count M3
+
+$$mouse \ av \ metric ~=~ sex * strain + \varepsilon$$
+
+```{r MixedModel.3, echo=FALSE, digits = 3}
 
@@
 
-Above is the table of coefficients for **MM.4**, HQ data set, lm function, just sex and strain, no subspecies.
-
-Significant factors are 
-strain G (p=0.0005), strainMSM and strainPWD (very slight),
-sex:G (p=0.001),
-sex:MSM, Sex:PWD (very slight p=.05, p=0.3),
-
-#### ANOVA, mean CO count full vs reduced
+<!--
+ Contrasts and ANOVA, mean CO count full vs reduced
 
@@
 
-Full model, so significant fixed effects. (For the full model, The proportions of the random effects make sense looking at the average CO data
-Models with random effects more significant than nulls (both just strain and no random effects).
-The models with fewer random effects, the fixed effects, sex and sex*subspec (interaction) become significant.
-
-```{r Base.strain.fixed, echo=FALSE}
-#make mixed model with strain as fixed, since there are interesting results
-#Consider including these results from all FIXED MODEL
-
-HetC.MM.strain.fixed <- lm(mean_co ~ strain*sex, data= DF.HetC.MixedModel )
-#this produces several significant factors
-
-HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
-
-#summary(HetC.MM.strain.fixed2)
-#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))
-
-#summary(HetC.MM.strain.fixed)
-#strainKAZ           -1.57392    0.86769  -1.814 0.071773 . 
-#strainPWD           -1.98205    0.74191  -2.672 0.008420 ** 
-#strainWSB           -3.09482    0.74191  -4.171 5.21e-05 *** 
-#sexmale             -2.89186    0.73333  -3.943 0.000125 ***
-#strainMSM:sexmale    6.83426    1.16149   5.884 2.69e-08 *** 
-#strainPWD:sexmale    7.04433    1.11783   6.302 3.38e-09 ***  
-#strainSKIVE:sexmale  4.04919    2.38756   1.696 0.092056 .   
-#strainWSB:sexmale    2.85301    1.04317   2.735 0.007024 **
-
-#male effect lowers the mean from intercept
-#while strain effects 
-```
-
- Summary of mean COs
-
-MM.2 has significant coefficients, sex and sex*subsp interaction.
-
-MM.2.Mol, the sex and interaction term (subsp*sex) are significant.
-
-MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.
-
-## Musc Males
-
-test a model for fixed effects of strains within musc. The purpose is to distinguihsi musc strains from low 'Dom' level strains.
-
-```{r Musc.male.Model.data.setup, echo=FALSE}
-#nested mouse in strain
-
-#remove the extra merged cols
-Musc.male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-Musc.male_mouse.av <- add_strain(Musc.male_mouse.av)
-Musc.male_mouse.av <- add_sex(Musc.male_mouse.av)
-Musc.male_mouse.av <- add_subsp(Musc.male_mouse.av)
-Musc.male_mouse.av <- add_species(Musc.male_mouse.av)
-
-Musc.male_mouse.av$subsp <- as.character(Musc.male_mouse.av$subsp)
-
-Musc.male_mouse.av$subsp[Musc.male_mouse.av$strain== "MSM"] <- 'Mol'
-Musc.male_mouse.av$subsp[Musc.male_mouse.av$strain== "MOLF"] <- "Mol"
-
-#use original for getting the other strains
-
-Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$species == "M.musculus",]
-Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$subsp == "Musc",]
-Musc.male_mouse.av <- Musc.male_mouse.av[Musc.male_mouse.av$sex == "male",]
-
-#set the factors
-Musc.male_mouse.av$strain <- factor(Musc.male_mouse.av$strain, ordered = FALSE)
-Musc.male_mouse.av$mouse <- factor(Musc.male_mouse.av$mouse, ordered = FALSE)
-
-#compare the the higher quality plots
-#Q1Q2
-
-#Q12_mouse_table_w.Ages
-
-Q12_mouse_table_w.Ages <- add_subsp(Q12_mouse_table_w.Ages)
-Q12_mouse_table_w.Ages <- add_species(Q12_mouse_table_w.Ages)
-
-Q12_mouse_table_w.Ages$subsp <- as.character(Q12_mouse_table_w.Ages$subsp)
-
-Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MSM"] <- 'Mol'
-Q12_mouse_table_w.Ages$subsp[Q12_mouse_table_w.Ages$strain== "MOLF"] <- "Mol"
-
-Musc.male.Q12 <- Q12_mouse_table_w.Ages[(Q12_mouse_table_w.Ages$subsp == "Musc")&(Q12_mouse_table_w.Ages$sex == "male"),]
-
-Musc.male.Q12$strain <- factor(Musc.male.Q12$strain, ordered = FALSE)
-Musc.male.Q12$mouse <- factor(Musc.male.Q12$mouse, ordered = FALSE)
-
-#Q123  Q123_mouse_table
-
-Q123_mouse_table <- add_subsp(Q123_mouse_table)
-Q123_mouse_table <- add_species(Q123_mouse_table)
-
-Q123_mouse_table$subsp <- as.character(Q123_mouse_table$subsp)
-
-Q123_mouse_table$subsp[Q123_mouse_table$strain== "MSM"] <- 'Mol'
-Q123_mouse_table$subsp[Q123_mouse_table$strain== "MOLF"] <- "Mol"
-
-Musc.male.Q123 <- Q123_mouse_table[(Q123_mouse_table$subsp == "Musc")&(Q123_mouse_table$sex == "male"),]
-
-Musc.male.Q123$strain <- factor(Musc.male.Q123$strain, ordered = FALSE)
-Musc.male.Q123$mouse <- factor(Musc.male.Q123$mouse, ordered = FALSE)
-
-```
-
-```{r Musc.male.show.plots, echo=FALSE}
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +  
-musc.male.plot
-
-Q123.musc.male.plot <- ggplot(Musc.male.Q123,
-                             aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q123 Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Q123.musc.male.plot
-#THE q12,data set only 
-#Q123 has 5 strain rep (no TOM), there are still some mouse outliers
-
-```
-
-This is the table and plot of new Musc strains
-
-`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
-
-Given the plot, make note of the potential outlier mice.  The strains have unequal variance
-
-KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
-
-PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
-
-```{r M1_Musc.male, echo=FALSE}
-
-#mouse.av ~ mouse | strain
-
-#not sure the right syntax for nested factors
-#Musc.male.M1 <- lm(mean_co ~ (mouse|strain), data=Musc.male_mouse.av )
-
-Musc.male.M1 <- lm(mean_co ~ strain, data=Musc.male_mouse.av )
-
-as.data.frame(  round(summary(Musc.male.M1)$coefficients, 4) )
-
-#all of the strain have a significant effect agaisnt the group mean / intercept
-#looking at the intercept -- the PWD observations are pushing up the average
-
-#musc male Q123  Q123_mouse_table
-
-Musc.male.Q123.M1 <- lm(mean_co ~ strain, data=Musc.male.Q123 )
-
-as.data.frame(  round(summary(Musc.male.Q123.M1)$coefficients, 4) )
-
-#mostly all strains, (but SKIVE) are significant -- similar to the top model
-
-#shoul I compare with subsampling
-
-```
-
-```{r Musc.male.M2}
-
-#mouse.av ~ strain
-
-```
+-->
 
@@
 
+## Male Specific Polymorphism
+
+Follow up on this finding -- included more strains which were excluded from the HetC analysis.
+
+-- what's the driving question for this section --
+
+> is there a strain effect in Dom?
+
+> How different are the Musc strains?
+
+test a model for fixed effects of strains within musc. The purpose is to distinguihsi musc strains from low 'Dom' level strains.
+
+```{r male.show.plots, echo=FALSE}
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
+fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
+fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
+fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
+
+table( droplevels(fomo.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.male.plot
+
+#Dom version 
+Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.male.plot
+
+```
+
+This is the table and plot of new Musc strains
+
+`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
+
+Given the plot, make note of the potential outlier mice.  The strains have unequal variance
+
+KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
+
+PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
+
+```{r Male.specific.test, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+
+#1. determine which strain effects are sig
+
+bhh11 <- lm(mean_co ~ subsp * strain, data= fomo.male_mouse.av)
+#bhh11$coefficients
+
+booo <- summary(bhh11)
+booo$coefficients
+
+lm.male.strain <- lm(mean_co ~ strain, data= fomo.male_mouse.av)
+ll <- summary(lm.male.strain)
+ll$coefficients
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+```
+
+msm, pwd skive are when all other male strains are added. skive is less sig than the others -- but it's still sig.  Given these results I can group SKIVE, PWD and MSM into the 'high rec' group.
+
+*PWD, MSM, SKIVE are in 'high rec group'*
+
 #### Proportion of chrm classes

### 2020-02-04 src/SetupData.Rmd
summarizing all results, updating saving files

@@
 
+-rm Peromycus data BivData!!
+
 - code Mol as discint subsp
@@
 - format the DF for applying to Mixed models
+
+- rm all of BD data?
 
@@
                  "3sep18_AST_m1",
-                  "3jul18_TOM_m1","3sep18_TOM_m1",
+                  "3sep18_TOM_m1",
                   "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
@@
 I have made a list of the old mice, but from the effect document the older mice might not be that different.
+
+the excluded TOM, had low quality cells, only 4.
+The exclude AST, had low poor quality cells, there are 3 other mice.
+The excluded CZECH males are poor quality (but near the mean).
+There are 3 other males.
 
@@
                         se   = round(sd / sqrt(Ncells), 3)
-
 )
@@
 
-```{r Q123_mouse.table, echo=FALSE}
+```{r Q123_mouse.table, echo=FALSE, eval=FALSE}
 
@@
 #only female male reps
+
+```
+
+```{r male.specific.table,echo=FALSE}
+
+#AP_mouse_table_w.Ages
+#use the above table to pull the male mouse levels
+
+male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")
+
+#formate the table for pluggin into models
 

### 2020-02-05 doc/Main_Report_Results.rmd
new file MERGing MLH1 and bivalent based results

@@
 
-# Models Framework
+# Average MLH1 Count Model
 
@@
 
-For now I will focus on 3 different models for comparing the subsp, sex and strain effects across relevant dependant variables.
+For now I will focus on 3 different models for comparing the subsp, sex and strain effects across relevant dependent variables.
 
@@
 -->
+
+#### Summary of Mixed models
 
@@
 A general pattern is that the female data has higher within mouse variance (both var and cV).
-Some strains more than others (invetigate potential outliers).
+Some strains more than others (investigate potential outliers).
 (LEW females with top 3 variance are all from different dissection dates)
@@
 
-### Modeling Within Mouse Variance
+> Modeling Within Mouse Variance
 
@@
 4.	lm(mouse.metric ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+## MLH1 variance M1
 
@@
 
+## MLH1 variance M2
+
 $$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
@@
 
-```{r M3.MLH1.var, echo=FALSE}
+```{r M2.MLH1.var, echo=FALSE}
 
@@
 
-$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-```{r M4.MLH1.var, echo=FALSE}
+## MLH1 variance M3
+
+$$mouse \ CO \ metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+```{r M3.MLH1.var, echo=FALSE}
 
@@
 
-test a model for fixed effects of strains within musc. The purpose is to distinguihsi musc strains from low 'Dom' level strains.
+test a model for fixed effects of strains within musc. The purpose is to distinguish musc strains from low 'Dom' level strains.
 

### 2020-02-05 doc/Results.rmd
new file MERGing MLH1 and bivalent based results

--- 2020-02-04 doc/Bivalent_Analysis.Rmd
@@
 ---
-title: "Single Bivalent Level Analysis"
-author: "April Peterson"
+title: "MLH1 Results Outline"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
   pdf_document: default
+bibliography: refs_Main_Report.bib
 ---
 
-```{r setup, include=FALSE}
+```{r setup, echo=FALSE, include = FALSE}
 library(knitr)
@@
 library(RLRsim)
+library(ggpubr)
 library(reshape2)
-#library(lmerTest) #loading this package will mess up RLRsim's exactRTRL()
-#for got the function i wanted to use
-
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-#setwd("C:/Users/April/Desktop/MLH1repo")
+
+setwd("~.")
+
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
+
+#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
+
+#2.2.20
+#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
+
+load(file="~./MLH1repo/data/CleanBivData_3.2.20.RData")
+
+#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
+
+#setwd("~./MLH1repo/doc/")
+#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
+#setwd("C:/Users/April/Documents/MLH1repo")
 #load(file="data/MLH1/MLH1_data_setup.RData")
 
-setwd("~./MLH1repo/")
-#load Clean BivData
-load(file="data/CleanBivData_3.2.20.RData")
-#load(file="data/CleanBivData_1.28.20.RData")
-
-#load other data files, MLH1
-load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_2.2.20.RData")#added batch18
-
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
-
+#load main data file
+#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
+#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
+
+#this MLH1 only goes up to batch16!
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
+#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
 #load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
-#fyi loading mulitple RData files doesn't earase ther files -- but it can overwrite files named the same thing
+#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
+#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
+#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
+
+#load(file="data/MetaData.RData") #this was replacing most updated data file
 
@@
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+```
+
+```{r organizing.data, echo=FALSE }
+#set the order of another column, based on another variable. consistant order for ploting
+
+#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
+#-ene endings refer to the cells/units within the cell stage? (I think)
+
+#apply this before ploting
+MLH1_data <- MLH1_data %>%
+  arrange(strain, sex, category, mouse) %>%
+  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+# sort your dataframe, by the focal categories
+MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+#use the above line for when order is out of wak
+
+#count the non-quality measures,  #remove non qualit
+#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
+#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
+
+#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
+#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
+
+#MLH1_data <- MLH1_data %>%
+#  arrange(category) %>%
+#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+
+#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
+AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+
+AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
+                 .(strain,sex),
+                 summarize,
+                 sum.ncells = sum(Ncells),
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co),
+                 AV.var_MLH1 = mean(var, na.rm=TRUE)
+
+)
+
+#there might be an error for changing wd
+#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
+
+AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
+AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+
+var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
+  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
+
+```
+
+<!--  insert Details of data set cleaning etc is in SetupData.rmd
+Effects from different factors are reviewed in Effects.Rmd -->
+
+```{r basic.numbers, echo=FALSE, eval=FALSE}
+
+kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+
+kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+
+```
+
+```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
+#str(AP_mouse_table_w.Ages)
+
+AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+
+#remove sex.x == na, and strain.x=na or F1
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
+#plot variance and mean (facet sex)
+
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
+
+VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
+
+cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
+
+str(AP_mouse_table.HQ)
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
+#[1] 0.5750399
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
+#0.321601
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
+# 0.3515544
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
+#-0.03713536
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
+# p-value = 8.623e-09
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
+
+#remove mice with fewer than 10 cells
+
+AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
+
+VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
+
+cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
+
+cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
+    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
+
+#why did I male just a female table?
+female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
+
+female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)
+
+```
+
+# Average MLH1 Count Model
+
+These are the basic Mixed Models we used as a frame work to analyze all the MLH1 count data.
+
+$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+The mixed model lets us separate effects to see how they effect a mouse mean CO count. The subspecies effect is a proxy for **Divergence** and the random strain effect is a proxy for **Polymorphism**.
+
+Setting up the data set for using the mixed model (base data set, HQ data set)
+
+- only Mus musculus strains with sex matched observations
+
+- Male gwRR have been adjusted for the PAR
+
+- Q5 excluded
+
+- All the effects are unorder factors
+
+- male ages within 5 - 20 weeks
+
+- Molossinus is coded as a subsp level
+
+```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=FALSE}
+# MOVE this into BIVDATA SETUP!!!
+
+#should be able to use the loaded 
+#DF.HetC.MixedModel.HQ
+
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+
+#What are the main DFs for this code chunk
+#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
+#DF.HetC.MixedModel   <- only Mmusculus
+#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
+
+#variance is still chr
+AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+
+#use the merged DF to filter by age 
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+#remove the extra merged cols
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+#DONT NEED CATEGORY
+
+#REMOVE EXTRA STRAINS
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+
+#SWITCH TO HQ.DF
+#mean_co.HQ  (this might remove some of the MOLF males)
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]  #this removes the MOLF female
+
+#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
+
+#afriad this is removing NA's
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
+
+#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
+
+#add_subsp function re-writen
+
+#mean_co.HQ.mol
+#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
+#        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))
+
+# For some reason I changed subsp to nums (wtf)
+#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
+#        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))
+
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | #DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#remove blank rows (which are these?)
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
+#151 mice, no rows with nas for values
+
+#main df with variables coded as chr
+DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+
+#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+```
+
+This is the break down of mice used within the Mixed model
+
+`r kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
+
+<!--Can I indicate the average number of cells with go into each mouse.? -->
+
+subspecies levels: 3, Musc, Dom and Molossinus
+
+strain levels: Dom: 3, Musc: 3  Mol: 2
+
+sex: 2 female male
+
+For now I will focus on 3 different models for comparing the subsp, sex and strain effects across relevant dependent variables.
+
+1. $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+Allows random effect from strain. Considers the wild derived inbred strains as random samples from each subspecies territory.
+
+2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
+
+Post hoc model for investigating strain as a fixed effect.
+
+3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$
+
+Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.
+
+## MLH1 count M1
+
+$$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+```{r MLH1co.av.M1, echo=FALSE, digits = 3}
+#THIS IS THE NEW MIXED MODEL
+#should be able to use the loaded 
+#DF.HetC.MixedModel.HQ
+
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+#summary(Reduced.strain.HQ)
+
+#summary(MLH1.MM.M1)
+drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***
+
+as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
+
+#only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
+anova(MLH1.MM.M1)
+
+exactRLRT(MLH1.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0
+
+```
+
+Above is the mixed model M1,
+
+## MLH1 count M2
+
+$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$
+
+```{r MixedModel.2, echo=FALSE,  digits = 3}
+#strain = fixed
+lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+
+anova(lm.strain.fixed.full)
+
+#this table of results is broken -- might be too long -- lm
+#lm has different names
+as.data.frame(round( summary(lm.strain.fixed.full)$coefficients, 6))
+
+#summary(strain.fixed.full)
+#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
+
+#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are 
+```
+
+Above is the table of coefficients for MM.3, HQ data set, lm function, sex, strain and subsp are interacting fixed effects.
+
+Strain G is significant (p=0.0005),
+
+interaction terms of sex and strain for (G, MSM, PWD).
+
+How are these models different from the last one -- MM4 is missing subsp, so only sex * strain effects.
+
+## MLH1 count M3
+
+$$mouse \ av \ metric ~=~ sex * strain + \varepsilon$$
+
+```{r MixedModel.3, echo=FALSE, digits = 3}
+
+strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+#below table is broken
+as.data.frame( round( summary(strain.fixed)$coefficients, 6))
+
+#summary(strain.fixed)
+#maybe this WSB effec is just in the HQ data set?
+```
+
+<!--
+ Contrasts and ANOVA, mean CO count full vs reduced
+
+Use Anova to compare full vs reduced models
+the reduced models will be those missing either, strain, sex, or subsp.
+
+```{r anova.full.v.reduced, echo=FALSE, eval=FALSE}
+
+#strain as fixed
+#CONTRAST models
+#newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
+
+HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
+R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
+
+full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
+
+#base vs strain
+fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co) #<.0001
+
+#show scientific vs normal
+#vv <- fullVstrain$`p-value`
+#round(vv[2], 8)
+#format(vv[2], scientific=F)
+
+#strain vs no random full.reduce
+fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
+
+#fullreduce vs strain
+#can't distinguish which random effects are more significant -- both tests are significant
+
+fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
+
+#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
+full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
+
+full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
+#this comparison not useful when there are different fixed effects
+
+#full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ) # < .0001
+
+#strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ) #NS
+
+#anova tests
+#full vs strain
+full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
+#full vs non
+full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
+#strain vs non
+strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
+
+anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
+
+#base vs mol model (this only makes sense if these are nested models)
+anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
+```
+
+-->
+
+#### Summary of Mixed models
+
+## Within Mouse Variance in CO Count per Cell
+
+<!-- intro
+in addition to the mean gwRR, the variance is also an important metric to describe
+
+Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
+
+RW's main qtl paper used the within animal variance, 
+"We mapped QTL for
+MLH1 count variance within individuals, applying an approach
+to deconvolve the mean-variance relationship (see
+Materials and Methods)."
+
+above method remove / accounts for mean-variance correlation -
+
+The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
+
+QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
+
+@article{wang2017genetics,
+  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
+  author={Wang, Richard J and Payseur, Bret A},
+  journal={Genetics},
+  volume={206},
+  number={4},
+  pages={1841--1852},
+  year={2017},
+  publisher={Genetics Soc America}
+}
+
+references for phenotypic variance connected to robustness to environment, increase in developmental consistency
+Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
+
+phenotypic variance leading to adaptation - 
+(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
+
+Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
+
+meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
+-->
+
+Histograms might not be the best way to represent, but they show the general pattern than female mice have higher within mouse variation for CO counts.
+
+```{r variance.hist, echo=FALSE, eval=FALSE}
+#figures of histograms,
+
+#make series for total dataset?
+Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("Histograms of within mouse variance, all mice")
+Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+Var.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("Histograms of within mouse variance, 10 cell min")
+Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
+
+Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
+Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
+
+Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
+
+Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
+Q12.var.hist.plot
+
+#some LEW female still have v high within mouse variance (2 mice, )
+
+```
+
+A general pattern is that the female data has higher within mouse variance (both var and cV).
+Some strains more than others (investigate potential outliers).
+(LEW females with top 3 variance are all from different dissection dates)
+This pattern also holds for within mouse cv for MLH1 count (not shown).
+
+**todo**
+
+Test if dissection date is more significant predictor than strain (for females) (Effect doc)
+Also test how the number of 0CO bivalents per cell affects mouse with variance. (lm)
+
+```{r cv.plots, echo=FALSE, eval=FALSE}
+
+cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("cV of within mouse variance, all mice")
+cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+cV.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("cV of within mouse variance, 10 cell min")
+cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
+cV.Hist
+```
+
+> Modeling Within Mouse Variance
+
+(use the same models are mean_co, )
+
+XXX (old 1.)	HetC.MixedModel.HQ <- lme(mouse.metric ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
+
+1.(new 1)	Reduced.strain.HQ <- lmer(mouse.metric ~ subsp * sex+ (1|strain), data=DF.HetC.MixedModel.HQ)
+
+3.	lm(mouse.metric ~ subsp \* sex \* strain, data=DF.HetC.MixedModel.HQ)
+
+4.	lm(mouse.metric ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+## MLH1 variance M1
+
+$$mouse metric  ~=~ subsp * sex + rand(strain)+ \varepsilon$$
+
+```{r M1.MLH1.var_cv, echo=FALSE}
+#modeling within mouse variance
+#use  DF.HetC.MixedModel.HQ
+
+#HQ var.CO
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+summary(MLH1var_M1)
+
+drop1(MLH1var_M1, test="Chisq") #interaction term NS
+#when switched to additive fixed, sex is very sig
+
+anova(MLH1var_M1)#sex explains most variance...
+
+exactRLRT(MLH1var_M1) #strain
+as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))
+
+#CV
+MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+summary(MLH1cv_M1)
+
+drop1(MLH1cv_M1, test="Chisq") #interaction term NS
+#when switched to additive fixed, sex is very sig
+anova(MLH1cv_M1)  #sex F value is large
+exactRLRT(MLH1cv_M1) #strain
+
+as.data.frame( round(summary(MLH1cv_M1)$coefficients, 6))
+
+#USE Q12 (Q123) table
+#Q12_mouse_table_w.Ages
+#Q12 var
+MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+
+drop1(MLH1var_M1_Q12, test="Chisq")#sig
+anova(MLH1var_M1_Q12)
+exactRLRT(MLH1var_M1_Q12)#NS
+
+as.data.frame( round(summary(MLH1var_M1_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+
+drop1(MLH1cv_M1_Q12, test="Chisq")#NS
+anova(MLH1cv_M1_Q12)#sex large F
+exactRLRT(MLH1cv_M1_Q12)#NS
+
+as.data.frame( round(summary(MLH1cv_M1_Q12)$coefficients, 6))
+
+```
+
+The fixed sex effect has the largest effect sizes.
+All are negative (for male).
+(I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
+The strain effects are only slightly significant for within mouse variance measures in the full dataset (includes low quality cells)
+
+## MLH1 variance M2
+
+$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+M2 slash M3
+
+```{r M2.MLH1.var, echo=FALSE}
+
+MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+summary(MLH1var_M3)
+anova(MLH1var_M3)  
+as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
+
+MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+summary(MLH1cv_M3)
+
+anova(MLH1cv_M3)
+as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
+
+#Q12 var
+MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+summary(MLH1var_M3_Q12)
+anova(MLH1var_M3_Q12)  
+as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+summary(MLH1cv_M3_Q12)
+anova(MLH1cv_M3_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
+
+```
+
+For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
+Most of the strain specific differences in within mouse variance go away in the Q12 dataset.
+
+## MLH1 variance M3
+
+$$mouse \ CO \ metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+```{r M3.MLH1.var, echo=FALSE}
+
+MLH1var_M4 <- lm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+summary(MLH1var_M4)
+
+drop1(MLH1var_M4, test="Chisq")
+anova(MLH1var_M4)
+as.data.frame( round(summary(MLH1var_M4)$coefficients, 6))
+
+MLH1cv_M4 <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summary(MLH1cv_M4)
+anova(MLH1cv_M4)  
+as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))
+
+### Q12 var
+MLH1var_M4_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+
+summary(MLH1var_M4_Q12)
+
+anova(MLH1var_M4_Q12)
+as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M4_Q12 <- lm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+
+summary(MLH1cv_M4_Q12)
+anova(MLH1cv_M4_Q12)
+
+as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
+
+```
+
+#### Summary witin mouse Variance
+
+The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.
+
+(some strains have significant effect -- some LEW females had large leverage).
+
+<!--
+For later analysis.  Not sure what these are good for.
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
+
+# check if the main mixed model results hold up when all cells are pooled
+
+#pooled.cells
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+#consider changing to ordered = FALSE, for the DFs used for MM
+DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
+
+DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
+
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+
+```
+
+- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
+This means there is 0 variance added to mean_CO to mean due to being MSM.
+Does this mean the model is very good at predicting MSM mean_CO.
+The fixed strain - sex effects are for musc are large.
+This will likely change with MOLF (male molf are low, so smaller fixed effect and )
+
+- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
+In the first model MSM has largest random effect, since it's so far from WSB female.
+
+-adding MOLF also doesn't have significant factors
+
+changed all to unordered factors.
+all the variables are coded as ordered factors.
+When the variables are kept factors, and unordered the names appear! yay!
+
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
+-->
+
+## Summary of Mixed Model Analyses
+
+**mouse average MLH1 count**
+
+Multiple models were tested to test for potential evo models -- for the mouse average CO counts.  (there were a mix of results -- and nuanced)
+
+First full Mixed model -- no significant fixed factors --(the random effect might be significant).
+
+The second model were strain was (nested and random), the fixed effects, sex and interaction of (subsp and sex), are much more significant.
+The coefficients indicate, males in general have 1 less in average, and the musc and molf subsp have ~3 more on average.
+
+**Variance effect**
+
+** Do these patterns hold across Quality bins? **
+
+# Figure 1
+
+work more on figure 1. Main figure, MLH1 strain avs on top, chrm proportion bar plot on bottom
+
+```{r first.boxplots, echo=FALSE, fig.align='center', warning=FALSE}
+
+#the below ordering works!! keep it
+MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
+              "PWD","CZECH","SKIVE",
+              "KAZ","TOM","AST",
+              "MSM","MOLF",
+              "CAST","HMI","SPRET","SPIC", "other"), order=T )
+#order the data frame
+MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
+#can't re-order the miceagg
+#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
+#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
+#        "grey","grey","grey",  "grey","grey","grey" )) +
+# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
+#don't show M.musculus by mouse
+#scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
+ #       "green","green4","grey",  "grey","grey","grey" )) +
+
+#order MLH1_data by strain!!
+#colors_of_strains, colors .. 
+#there is no order to 'mouse' so it looks shitty
+
+justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
+  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+
+
+justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")
+
+#scales="free_y"
+#this is ok except for the set placement of mice on x axis
+#if the full dataset was ordered correctly, this would kinda fix it
+
+#boxplots by strain
+ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())+
+  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )+scale_fill_manual(values=colors_of_strains)
+
+ss <- ss   + facet_wrap(~ sex, scales="free")
+
+#these plot currently look bad
+```
+
+```{r table.for.plot, echo=FALSE}
+#why do I add in a different table?
+
+AP_strain.table.HQ <- ddply(MLH1_data, c("strain", "sex"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  =length(nMLH1.foci),
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci, na.rm = TRUE),
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                       #Standard error (mice vs cells)
+                        se  = round(sd / sqrt(Ncells), 3)
+    )
+
+```
+
+```{r Fig1, echo=FALSE, warning=FALSE}
+
+Fig1_v1_strain <- ggplot(AP_strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+  geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+ 
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
+#remove y labfor male 
+  #segments
+  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
+  #cast
+  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
+  #outgroups
+  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
+           y = c(18.5,18.5,33,18.5, 18.5),
+
+   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
+
+Fig1_v1_strain
+```
+
+What are the steps the each chunks walk through.
+
+## Male Specific Polymorphism
+
+Follow up on this finding -- included more strains which were excluded from the HetC analysis.
+
+-- what's the driving question for this section --
+
+> is there a strain effect in Dom?
+
+> How different are the Musc strains?
+
+test a model for fixed effects of strains within musc. The purpose is to distinguish musc strains from low 'Dom' level strains.
+
+```{r male.show.plots, echo=FALSE}
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
+fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
+fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
+fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
+
+table( droplevels(fomo.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.male.plot
+
+#Dom version 
+Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.male.plot
+
+```
+
+This is the table and plot of new Musc strains
+
+`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
+
+Given the plot, make note of the potential outlier mice.  The strains have unequal variance
+
+KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
+
+PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
+
+```{r Male.specific.test, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+
+#1. determine which strain effects are sig
+
+bhh11 <- lm(mean_co ~ subsp * strain, data= fomo.male_mouse.av)
+#bhh11$coefficients
+
+booo <- summary(bhh11)
+booo$coefficients
+
+lm.male.strain <- lm(mean_co ~ strain, data= fomo.male_mouse.av)
+ll <- summary(lm.male.strain)
+ll$coefficients
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+```
+
+msm, pwd skive are when all other male strains are added. skive is less sig than the others -- but it's still sig.  Given these results I can group SKIVE, PWD and MSM into the 'high rec' group.
+
+*PWD, MSM, SKIVE are in 'high rec group'*
+
+# Proportion of chrm classes
+
+make a smaller proportion chrm block --
+
+These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes.
+Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs. (the gwRR for most mice is low/close to the minimum)
+
+This is also another way to score heterochiasmy (that gets around the sex chrms, i think)
+
+The data below are from hand measured cells (not from bivalent data)
+generate smaller data set of chrm proportions (0CO, 1CO,2CO etc)
+10 cells from each musc category. **Add Molf female**
+
+```{r chrm_prop_data, echo=FALSE}
+#shut off eval for now because of wd / path error
+#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
+chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#try adding the BivData version of the proportions
+
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
+
+chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
+chrm.class.DF <- add_mouse(chrm.class.DF)
+chrm.class.DF <- add_strain(chrm.class.DF)
+chrm.class.DF <- add_sex(chrm.class.DF)
+chrm.class.DF <- add_category(chrm.class.DF)
+#chrm.class.DF <- add_euth_date(meta.data.file2)
+#chrm.class.DF <- add_age(meta.data.file2)
+
+#table(chrm.class.DF$category)
+#str(chrm.class.DF)
+
+#summarize by category
+proportion_stats_category  <- ddply(.data=chrm.class.DF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
+
+                 n0CO = sum(n0CO, na.rm = TRUE),
+                  n1CO = sum(n1CO,na.rm = TRUE),
+                  n2CO = sum(n2CO,na.rm = TRUE),
+                  n3CO = sum(n3CO,na.rm = TRUE),
+                  n4CO = sum(n4CO,na.rm = TRUE),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv,
+                 per4CO= sum(n4CO)/ tot_biv,
+
+                 mean.MLH1.foci = mean(MLH1.sum.check),
+                 mean.MLH1.foci2 = mean(nMLH1.foci),
+                 meanQ = mean(quality)
+
+)
+#this is the table that needs melting
+#but is also easy to redo by hand
+
+#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
+```
+
+```{r plot.chrm.prop.melt, echo=FALSE}
+
+#add subspecies to these values
+melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
+
+#add the subsps
+
+melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
+                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
+
+melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
+
+chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
+
+#need a second DF / the first df
+#+ annotate() text 
+#annotate()text
+
+#proportion_stats_category,  tot_biv, ncells
+
+#add number of cells and bivalents in the corner
+
+chrm.class.barplot
+```
+
+```{r test.Chrm.Prop, echo=FALSE}
+
+#this first one is the 'first' MLH1 count
+#which is this variable
+
+#oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
+#summary(oopp2)
+#none of the classes are significant
+
+#mean is the new calculated-- mean 0CO is significant and intercept
+#oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = #proportion_stats_category)
+
+#summary(oopp)
+#not sure if these are inter-correlated
+#I don't know why 0COs and the intercept are significant..
+#0CO and 1CO are significant when only chrm classes are included (and intercept)
+
+#no longer significant when used
+#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
+#after added more MSM data, tot_biv is not significant
+#0CO is now more significant, and 2CO is sightly significant
+
+#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
+
+#MSM needs more bialent observations
+#use melt to put all propotions 
+#display / visualize / compare proportions
+#the table needs to be melted, because it's split by category
+
+#Need to run the chrm proportions thru melt code.
+
+chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
+
+#chrm.class.DF$biv.sum.check
+```
+
+High musc males have half 1CO and 2CO bivalents.
+For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
+
+Add p.values of tests for the how the proportions differ.
+
+*cast female data came from another publication, add this ref (Kohler?)
+
+The bottom lines and subspecies points don't exactly match up.
+
+**add in cell images like the other draft figures.**
+
+# Chrm Class proportions
+
+The figure below shows the average percentages of chromosome classes per category.
+
+```{r Chrm.props, echo=FALSE}
+
+#barplots for proportions of chrm classes
+
+#consider adding this the the setup block (could have done this to category level table)
+mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$CO0 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$CO1 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$CO2 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$CO3 / mouse.level.BivData_table$nboxes_passed
+
+mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$CO4 / mouse.level.BivData_table$nboxes_passed
+
+#DF needs to be melted -- so they all in 1 col
+#per code, I've done this by hand
+
+#subset DF for ease
+skinny.mouse.level <- mouse.level.BivData_table[-c(11:15 ) ]
+col.names.list.99 <- colnames(skinny.mouse.level)
+
+#"mouse"         "ncells"        "boxs.IDd"      "nboxes_passed" "Nbiv"          "CO0"           "CO1"       #[8] "CO2"           "CO3"           "CO4"           "per.0CO"       "per.1CO"       "per.2CO"       "per.3CO"      [15] "per.4CO"
+
+#all of the cols to use
+MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4",
+
+               #leave out the Foci
+                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")], 
+  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4"))
+
+#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
+
+#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
+
+MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
+
+#then this DF has to be summarized into categories
+chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
+                             mean.percent.type = mean(value)
+                             #sum is the wrong!
+                             #used average percentage, might consider a different summary
+)
+
+#ggplot
+
+chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
+chrm.class.barplot
+
+```
+
+These results align with the previous figure I made using percentages from hand measured results (dozens of observations, instead of hundreds).
+These results fit with the rapid male evolution of gwRR in MSM and PWD.
+
+- Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.
+
+# DMC1 Results Spermatocytes
+
+These data are meant to test if CO precursors, double strand breaks (DSBs), are significant predictors for MLH1 count variation. Simply, do mice with more DSBs also have more COs?
+
+ToDo:
+  - attempt to supplement data with more KAZ DMC1 cell observations.
+  - change all the bad variable names
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize, 
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+```
+
+```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#these are histogram plots for testing Normality shape -- move to supplemenatry section
+
+#- check that distributions are close to normal
+#- check for equal variance.
+
+DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
+
+DMC1.histogram
+#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
+#bartlet's test, Levine's test
+```
+
+The counts look close to normal.
+
+Potentially figure 2. Using a function to display the p-vales for mean comparison
+
+```{r DMC1.plot, echo=FALSE, warning=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")
+
+DMC1.boxplot
+  #this will add neat bars, but still need to figure out what groups ect it's actually using
+#THE stat_compare function adds a arm thing and pvalue for 
+#  stat_compare_means(comparisons = my_comparisons)
+
+#status - legend, expression y-axis, facet, facet_wrap SYMBOL
+#legend - stage. y=total count, FACet strain
+
+#add column to distinguish high vs low rec strains
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
+
+# annotation table with adjusted pvals and y-position of the labels
+anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",]) %>% mutate(y_pos = 300)
+
+DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=stage), position=position_jitterdodge()) + 
+  facet_wrap(~stage, nrow=1) + 
+  ggsignif::geom_signif(
+    data=anno_df, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")
+#this is kinda the plot I'm enivioning 
+DMC1.bxlt_High.to.Low
+
+#I want a test between 1) high vs low cell stages, 
+#I think a better way to display would be
+#facet by stage, x is the strain
+#compare means across high vs low  (this level isn't defined)
+
+#example from here
+#https://github.com/kassambara/ggpubr/issues/65
+
+#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
+
+```
+
+Summary:
+
+The distributions of counts is close to normal (also the higher count is better than the MLH1 counts)
+
+Above figure shows p-values for comparison of means across the cell stages.
+
+Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days).
+One mouse represents each strain.
+The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
+
+ToDo think of ways to add in the compared means for the high and low groups (across cell stage)
+how similar are G-WSB or PWD-MSM to each other?
+
+## DMC1 Permutations with equal sampling
+
+The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
+Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
+Permutations were done for both cell stages.
+The actual p value is in red.
+
+The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
+
+```{r DMC1.subsampling, echo=FALSE}
+
+#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
+# the 2 groups
+
+#make the 2 pools
+High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+
+Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+
+#permute drawing / subsampling and performing the t-test
+#put the p.value results into a DF!
+
+#permut.SC.cv.diff  <- as.data.frame(
+#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
+
+#permuting samples -- and re-licating t-tests
+
+#nrep <- 1000
+
+#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
+bunchOpvals.L <- replicate(1000, 
+    # t.test
+    t.test( sample(High.L.pool, 25, replace = FALSE),
+            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
+          )
+bunchOpvals.L <- as.data.frame(bunchOpvals.L)
+bunchOpvals.L$stage <- "L"
+
+bunchOpvals.Z <- replicate(1000, 
+    # t.test
+    t.test( sample(High.Z.pool, 25, replace = FALSE),
+            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
+          )
+
+bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
+bunchOpvals.Z$stage <- "Z"
+
+#could merge the lists of p-values
+replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
+#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
+
+#ASSUMPTIONS; normality, equal variance
+
+#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
+
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
+#plot.pvals.L
+#plot.pvals.Z
+
+#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
+#n use the var.equal = TRUE
+#https://www.statmethods.net/stats/ttest.html
+```
+
+```{r DMC1_MLH1.cor, echo=FALSE}
+#make category averages from AP_mouse_table.HQ
+mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+#attach this mini DF to the DMC1 data
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+```
+
+> At what stage are DMC1 foci numbers most correlated with MLH1?
+
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+
+```{r barlett.test, echo=FALSE, eval=FALSE}
+
+#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
+bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
+
+```
+
+Evidence for non-equal variance across strains for zygotene cells.
+
+Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
+
+From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#L vs Z (all species)
+#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
+#p-value = 1.032e-05
+#not a valid interpertation
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = r ttest.HighLow.L$p.value) while the Z cells, there is not a significant difference across the high and low groups (value = r ttest.HighLow.Z$p.value).
+
+<!--
+ Single Bivalent Patterns
+
+## Total SC Results
+
+insert male total SC results here
+
+# Female patterns
+
+Nuanced
+there is strain specific evolution for females ...
+
+# Sexual dimorphism and heterochiasmy
+
+total SC length / SC compaction and telomere bias in positioning are conserved / consistently sexually dimorphism (in these mice?)  for most species
+
+make the heterochiasmy plot, with female strain averages adjusted for X
+(measure variance in CO counts for the top 5 female whole cell bivalents)
+(variance in )
+how does this compare to variance of mid 5, bottom 5?
+
+# Predictions / tests for selection on Bivalent structure
+
+-->
+
+<!--
+# Rapid male specific evolution
+
+list out the points which support rapid male specific evolution for MLH1 rates.
+
+number above bars are the raw counts for bivalent classes.
+
+Potential figure 2 (eventually). Try to make the figure taller.
+-- add more data to make sample sizes more equal.
+-- add space to indicate where female data is missing
+
+SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
+What's up with Lew females? there's lots of 2COs, also WSB female.
+
+This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
+If these data are random samples of bivalents, the X
+
+ToDo: chisq tests for difference in proportions, 
+- within strain across sex,
+- males, across high and low groups
+- look at mouse level proportions
+-->
+
+# Note on Heterochiasmy Definition
+
+I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
+In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
+
+Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
+(I note if the and Y are paired, which they are at a high rate).
+The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
+
+To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
+
+1. Compile full-cell data from females (all 20 bivalents measured)
+2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
+3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
+
+According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
+
+```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
+#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
+#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
+#remove quality 5 cells
+
+#outgroups
+#SPIC
+HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
+#unique(HetC.SPIC$mouse)
+#non-passing mice
+SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
+HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
+#calq sex mean
+SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
+
+#SPRET
+HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
+#unique(HetC.SPRET$mouse)
+#non-passing mice
+#SPRET.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
+
+#CAROLI
+HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
+#unique(HetC.CAROLI$mouse)
+#non-passing mice
+#SPRET.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+#cast
+HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
+#unique(HetC.CAST$mouse)
+#non-passing mice
+#X.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
+
+#HMI
+HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
+#unique(HetC.HMI$mouse)
+#non-passing mice
+#X.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+#dom
+
+#WSB
+HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
+#unique(HetC.WSB$mouse)
+#non-passing mice
+WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
+HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
+#calq sex mean
+WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
+
+#LEW
+HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
+unique(HetC.LEW$mouse)
+#non-passing mice
+LEW.remove.mice <- c("4jan17_LEW_f2")
+HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
+
+#calq sex mean
+LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
+
+#G
+HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
+#unique(HetC.G$mouse)
+#non-passing mice
+G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
+HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
+
+#calq sex mean
+G.mean.MLH1.obs <- ddply(.data=HetC.G, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
+
+#PERC
+HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
+#unique(HetC.PERC$mouse)
+#non-passing mice
+#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
+#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
+
+#calq sex mean
+PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
+
+#MUSC
+
+#PWD
+HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
+#unique(HetC.PWD$mouse)
+#non-passing mice
+PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
+
+# CZECH
+HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
+#unique(HetC.CZECH$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
+
+# SKIVE
+HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
+#unique(HetC.SKIVE$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
+
+# KAZ
+HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
+#unique(HetC.KAZ$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
+
+#TOM
+HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
+#unique(HetC.TOM$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
+
+#AST
+HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
+#unique(HetC.AST$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
+
+#Molosinus
+
+#MSM
+HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
+unique(HetC.MSM$mouse)
+
+#non-passing mice
+MSM.remove.mice <- c("12sep16_MSM_f3")
+HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
+
+#calq sex mean
+MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
+
+#MOLF
+HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
+unique(HetC.MOLF$mouse)
+
+#non-passing mice
+MOLF.remove.mice <- c("27oct18_MOLF_m3")
+HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
+
+#calq sex mean
+MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
+
+```
+
+```{r HetC.plot, echo=FALSE, warning=FALSE}
+#HETC plot
+
+#set up data
+
+HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#
+#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
+
+HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
+
+#remove the 'other'
+
+HetC_strain_table.M$sex <- "male"
+HetC_strain_table.F$sex <- "female"
+
+HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
+
+HetC_strain_table <- add_subsp(HetC_strain_table)
+
+#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
+
+HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
+HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
+#str(HetC_strain_table)
+
+Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
+  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
+
+Het.C.plot
+#this plot is missing several
+
+```
+
+There is now MOLF, which has female biased hetC
+3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
+1 of the musc strains has female biased heterochiasmy, KAZ.
+
+```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
+
+#REMOVE This BLOCK?
+
+#### Distributions by mouse
+#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
+
+#remove this block since it has it's own src/script
+
+### to integrate 'pass'/fail
+### this should be in setup data
+
+#mice that don't have female data will break the chain-loop
+#MOLF female, HMI, CZECH female, CAROLI female, Tom male
+
+#instead of adding fake data, remove the categories from list
+
+#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
+#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
+#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
+#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
+#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
+#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
+#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
+#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
+#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
+
+pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
+
+#this loop makes a quant_status col for a list of mice
+for(i in 1:length(MLH1_data$mouse)){
+  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+}
+
+df_list <- split(MLH1_data, (MLH1_data$category))
+#remove sections which are in category but have no data
+#this is where you remove the non-data data!
+df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
+
+#7 PERC, 12 MOLF female, 
+#
+
+#df list, is a list of df's which were split() from BIG df by category
+df_list 
+o=0
+i=0
+
+MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
+
+#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
+for(i in 1:(length(df_list))){
+ # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
+  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
+  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
+                      mouse_nMLH1 = mean(adj_nMLH1.foci)
+  )
+# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
+  for(o in 1:length(df_list[[i]]$Original.Name)){
+    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
+        }
+}
+
+plot_list = list()
+for (h in 1:length(df_list)) {
+  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
+    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
+    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
+    facet_wrap(~ mouse, drop = TRUE) + 
+    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
+  #make an empty ggplot? for missing data
+
+  plot_list[[h]] =   pf
+}
+
+#orig (I think this is the same as above)
+#plot_list = list()
+#for (h in 1:length(df_list)) {
+#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
+#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
+
+#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
+#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
+
+ #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
+
+    #theme for indiviual mice, 
+#theme(plot.background = element_rect(fill = "green"))
+#  plot_list[[h]] =   pf
+#}
+#can't change the alpha/transparency of the red rec
+
+#aes must be same length ---   #geom_point(stat = "identity") +
+# the last one is correct....?
+
+#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
+```
+
+```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
+geom_hline(aes(yintercept = med, group = gr), colour = 'red')
+#no plot is made that goes in the list -- so it can't be shown
+for (gg in 1:length(plot_list) ){
+  show(plot_list[gg])
+ # invisible(lapply(obj, function(x) plot(x,main="some plot")))
+}
+
+#this stops at MSM
+```
+
+The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
+
+Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
+The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
+
+```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
+
+no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
+
+MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
+
+MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
+
+MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
+
+MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
+
+kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
+
+bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
+
+#add titles
+#use multiplot
+#remove X axis
+
+multiplot(kuku,bubu, cols=2)
 ```
@@
 
-# Setup and Discription of Dataset
+<!-- merging the 2 data sets-->
+
+# Single Bivalent Level Results
 
@@
 
-```{r tech.ect, echo=FALSE, eval=FALSE}
-#this fig not needed
-
-nBiv.per.cell <- ggplot(cell.level.BivData_table, aes(nbivs)) + ggtitle("Number of passing bivalents per cell") + geom_histogram(bins = 30) +facet_wrap( ~ category)
-nBiv.per.cell
-
-#for cells with close to 20 - they are almost complete
-#they should be used to
-
-#This isn't the same as curated/passing bivalents           
-#This plot illustrates how well the algorithm seperated indiviual bivalents per cell image. 
-#Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally #segmented.
-```
-
-# Chrm Class proportions
-
-The figure below shows the average percentages of chromosome classes per category.
-
-```{r Chrm.props, echo=FALSE}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$CO0 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$CO1 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$CO2 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$CO3 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$CO4 / mouse.level.BivData_table$nboxes_passed
-
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(11:15 ) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#"mouse"         "ncells"        "boxs.IDd"      "nboxes_passed" "Nbiv"          "CO0"           "CO1"       #[8] "CO2"           "CO3"           "CO4"           "per.0CO"       "per.1CO"       "per.2CO"       "per.3CO"      [15] "per.4CO"
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")], 
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4"))
-
-#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-
-#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#ggplot
-
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
-chrm.class.barplot
-
-```
-
-These results align with the previous figure I made using percentages from hand measured results (dozens of observations, instead of hundreds).
-These results fit with the rapid male evolution of gwRR in MSM and PWD.
-
-- Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.
-
-# Chromosome Size Effect
-
-I tried to isolate bivalents which are in the top quartille for SC length from their cells.
+> Q1. driving questions, which traits are sexually dimorphic?
+
+> Q2. which traits fit male polymorphism predictions?
+
+# Model Frameworks
+
+> Q1. same models for MLH1 based sex differences (and evolution)
+
+> Q2 the same models for the male polymorphism and logistic regression for testing if bivalent traits can predcit if they are from high or low rec group
+
+I decided to use the mouse averages for most of the bivalent metric tests. some of my logic is below.
+
+**Pro Bivalent level**
+
+- The data are continuous and closer to normal/Gaussian expectations than the MLH1 counts.
+
+- There is within cell variance that is an issue (would affect single bivalent observations).
+
+**Pro Mouse averages**
+
+- Mouse average is a conservative choice, it would summarize general patterns which fits with the paper.
+
+- Simple application of the same models as MLH1 based reseults.
+
+- Limited work to add to the data set if I can justify that the collection of single bivalents are random and equivalent across mice.
+
+So I'll use the mouse averages for all the metrics is what is used in the mixed models.
+
+```{r biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}
+
+#this block should be in the data set up
+
+#the mouse averages will be used
+
+mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+mouse.avs_4MM <- add_sex(mouse.avs_4MM)
+mouse.avs_4MM <- add_strain(mouse.avs_4MM)
+mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
+mouse.avs_4MM <- add_category(mouse.avs_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
+mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
+mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
+
+#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
+#code if males are high or low?
+
+#code intp chr
+
+#assign high and low
+#divide musc males into groups
+mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+
+```
+
+In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .
+
+```{r biv.data_table.4MM.Male, echo=FALSE}
+#remove female, 
+#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
+
+#Curated_BivData_male
+
+male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+male.av_4MM <- add_sex(male.av_4MM)
+male.av_4MM <- add_strain(male.av_4MM)
+male.av_4MM <- add_subsp(male.av_4MM)
+male.av_4MM <- add_category(male.av_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
+male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
+male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
+
+```
+
+## Q1 Analysis, Predictions for Heterochiasmy Q1
+
+Using the Mixed model framework which tests the effects (and interactions) of subspecies, sex and strain, I will test for evolution of the following traits.
+
+Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
+
+1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of uniform vs telomere/biased CO positioning.
+
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
+
+# SC Lengths
+
+<!--Brief lit review for SC length and MLH1 expections. -->
+
+We expect female SC lengths to be longer (refs).
+
+In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
+
+For the 0 class chromosomes most all are around the same size of the 1CO distribution.
+
+Add in the code for the plot under 2COs males and females
+
+```{r show.SC.length, echo=FALSE, warning=FALSE}
+#for now remove the 
+#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
+#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS
+
+#add boxplots for mouse level means
+#add second x axis, hand.foci.count
+
+#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#nested boxplots
+SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
+
+SC.length_Nested_boxplots
+
+#+ +geom_jitter(aes(alpha=0.2))
+```
+
+The dependant variables I'll be testing in the mixed model framework are:
+
+- Pooled SC lengths
+
+- 1CO
+
+- 2CO
+
+- 3CO
+
+- long.biv
+
+<!--
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
+
+To estimate random effects, I'm using lmer (not lme where I could do the rand(strain*sex)). So now it's just random(strain).
+
+The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.
+
+remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
+
+#### Pooled SC Lengths
+
+```{r MM_2.SC.Length, echo=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+#p
+
+```
+
+I used the mixed model with the lme() for the full Curated BivData and long bivdata
+
+-->
+
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+#### HetC_M1
+
+Insert the results for the MMs using the dependant variable 1CO - use the Mixed model (lmer), for all of the meterics (pooled SC, 1CO 2CO ect)
+
+Below's the code for Mixed model results -- try to organize them into a table or somethinf
+
+```{r M1_many.metrics, echo=FALSE}
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#all pooled SC lengths
+lmer.SCLength_all_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
+anova(lmer.SCLength_all_MM.M1) # also F values
+
+exactRLRT(lmer.SCLength_all_MM.M1)
+as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
+#can't get p values from this summary
+
+#1CO
+SCLength_1CO_MM.M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+drop1(SCLength_1CO_MM.M1, test="Chisq")
+exactRLRT(SCLength_1CO_MM.M1)
+as.data.frame(round(summary(SCLength_1CO_MM.M1)$coefficients, 6))
+
+#2CO
+SCLength_2CO_MM.M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+drop1(SCLength_2CO_MM.M1, test="Chisq")
+exactRLRT(SCLength_2CO_MM.M1)
+as.data.frame(round(summary(SCLength_2CO_MM.M1)$coefficients, 6))
+
+#Long.Biv
+SCLength_long.biv_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
+
+drop1(SCLength_long.biv_MM.M1, test="Chisq")
+exactRLRT(SCLength_long.biv_MM.M1)
+as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+
+```
+
+#### HetC_M2
+
+All fixed and all interactions
+
+```{r M2_many.metrics, echo=FALSE, eval=FALSE}
+
+#feed these 2 DFs into M2
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+```
+
+#### HetC_M3
+
+```{r HetC_M3, echo=FALSE}
+
+```
+
+#### Strain Comparisons of SC Length
+
+Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
+
+(adding all of these blocks to -- the setup script); strain.Bivalents.DF
+
+```{r strain.SClength.t.test, echo=FALSE}
+#t.tests for all the strain specific data frames
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+# Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
+
+The general predictions across the males and subspecies based on th above MLH1 results.
+
+For positive correlation traits/metrics
+
+i) in DOM strains, low to no difference across strains
+
+ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
+
+ii) in Mol, MSM > MOLF
+
+A) **SC lengths** will be longer for high Rec strains.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
+
+C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
+
+The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
+When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
+
+C.2) sis-co-ten metric ... (what about the clustering?)
+
+C.3) telomere and centromere distances ...
+
+$$mouse \ average \ SC  \ length ~=~ rand(strain) + \varepsilon $$
+
+$$mouse \ average \ IFD \ trait  ~=~ rand(strain) + \varepsilon $$
+The IFD traits are:
+
+- IFD_ABS
+
+- IFD_PER
+
+$$mouse \ average \ CO \ position \ trait  ~=~ rand(strain) + \varepsilon $$
+
+These are the models for male polymorphism, (I am re-thinking the tests I think)
+
+Use these DFs for the models:  **Male.poly.Mouse.Table_BivData_4MM**  and **Long.biv_mouse_table_male.poly**
+
+** add chunks for glm()s of male polymorphism**
+
+```{r male.show.plots, echo=FALSE, eval=FALSE}
+
+#REWRITE THIS FOR SC LENGTH COMPARISONS
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
+fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
+fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
+fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
+
+table( droplevels(fomo.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.male.plot
+
+#Dom version 
+Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.male.plot
+```
+
+```{r SC.length.male_tests, echo=FALSE}
+## run the tests to check following expectations
+
+#1.Dom strains not different from ech other
+#2. PWD and SKIVE are different from the other strains
+#3.MSM and MOLF are different from each other
+
+#Male.poly.Mouse.Table_BivData_4MM
+#Long.biv_mouse_table_male.poly
+
+#decide between lm or glm..?
+#mouse avs, males
+
+male.Dom.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Dom"), ]
+#the subsp labels get messed up, unless they are designated chr first
+male.Musc.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == 2), ]
+
+mouse.avs_4MM$subsp <- as.character(mouse.avs_4MM$subsp)
+mouse.avs_4MM$subsp <- ifelse(grepl("MSM", mouse.avs_4MM$strain), "Mol", 
+        ifelse(grepl("MOLF", mouse.avs_4MM$strain), "Mol", mouse.avs_4MM$subsp))
+male.Mol.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Mol"), ]
+
+Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
+
+Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  )
+Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
+#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
+#total sc sig, 1CO sig, 2CO sig
+
+Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
+#0.000369 ***
+
+Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
+#0.000816 ***
+
+Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
+#0.0148 *
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
+#making plots above show that mouse average SC lengths are lower could there be microscope issue?
+
+#make sure to remove some of the juvi mice!! from the AP_mouse table
+
+show.Dom.male.CO <- ggplot(
+AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") & (AP_mouse_table_w.Ages$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
+
+Dom.male_CO.aov <- aov(mean_co ~ strain, data = AP_mouse_table.HQ[(AP_mouse_table.HQ$sex == "male") & (AP_mouse_table.HQ$subsp == "Dom"),])
+#P not sig for CO mean
+
+#seems like there are more mice in the MLH1 data set than bivData
+#
+
+#WSB all dissected and imaged in '15
+#G imaged 15 - 16
+#LEW imaged 15 to 17
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "1CO SC Length", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "2CO SC Length", xlab = "")
+```
+
+```{r SC.length.pairwise.T, echo=FALSE}
+#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
+
+male.Musc.mouse.av <- male.av_4MM[(male.av_4MM$subsp == "Musc")&(male.av_4MM$sex == "male"),]
+
+Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
+#PWD and MSM are most diff from SKIVE?
+
+Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
+#the pooled SC value tests are sig
+
+#the mol comparisons are different due to adj p values
+
+Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
+#no sig (only 3) -- does include CZECCH
+
+#non-sig, with mol pooled with musc
+
+Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
+
+Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
+
+Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
+#non of these sig with the mulitple adjust
+
+Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
+
+#pairwise.t.test(Ozone, Month, p.adj = "bonf")
+
+```
+
+If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
+
+Ideas for the above 3 tests,
+
+1. no difference in DOM
+- no sig logistic regression test
+- glm, strain effect?
+
+### Summary
+
+> Is sex a significant effect for SC length? (as predicted)
+
+- Sex is a significant effect for SC length
+The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
+
+- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
+
+- The Long Biv Data set largely agrees with the full curated dataset
+
+- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
+
+> Is the random Effect of strain an effect on SC length?
+
+The exactRLRT() test indicates that the random strain effect might be significant, p= xx
+
+(merge this with the code above)
+Prediction, High rec males have longer SC.
+
+```{r log.reg_SC.length, echo=FALSE}
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
+
+#1CO
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+summary(log.reg_SC.length_mouse.av_1CO)
+
+#2CO
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+
+summary(log.reg_SC.length_mouse.av_2CO)
+
+#3CO
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
+
+                         , family=binomial(link="logit"))
+summary(log.reg_SC.length_mouse.av_3CO)
+
+# bivalent level log regression
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_SC.length_BIV)
+
+#1CO
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_SC.length_BIV_1CO)
+
+#2CO
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 2),], family=binomial(link="logit"))
+summary(log.reg_SC.length_BIV_2CO)
+
+#3CO
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 3),], family=binomial(link="logit"))
+
+summary(log.reg_SC.length_BIV_3CO)
+```
+
+The mean SC logistic regresion model for mouse averages
+
+`r log.reg_SC.length_mouse.av$coefficients`
+
+and for the single bivalent levels
+
+`r log.reg_SC.length_BIV$coefficients`
+
+#### t.test for SC Length
+
+what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
+
+```{r SC.Length_t.test, echo=FALSE}
+
+T.test_SC.Len_mouse.av <- t.test(mouse.avs_4MM$mean_SC[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean_SC[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")&(mouse.avs_4MM$Rec.group == 0)]
+)#p-value = 0.001567
+
+#1CO
+T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#2CO
+T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#3CO
+T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#Single Biv level
+T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
+
+#1CO
+T.test_SC.Len_single.biv_1CO <- t.test(
+  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
+
+      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
+  )
+
+#2CO
+T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
+
+                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
+  )
+
+  #3CO
+  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
+
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
+  )
+```
+
+The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_SC.Len_mouse.av$p.value` and p=`r T.test_SC.Len_single.biv$p.value`.
+
+```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
+#I don't think the Mixed models fit well for the high vs low polymorphism tests.
+
+#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
+#MSM_SC vs MOLF_SC
+
+#2. logistic regression for sc length and chrm class?
+#if the prediction / significance
+
+#isolate data
+#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
+
+#build and test MM
+lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+
+lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
+
+anova(lmer_MM_SC_length_male.poly)
+
+rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
+
+```
+
+I can't get the above mixed models to work for testing high vs low rec groups
+
+```{r long.sc_log.res_meanSC, eval=FALSE}
+#hold off on reporting/ working with the long SC biv resutls
+
+long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
+long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
+
+#divide musc males into groups
+long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
+                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
+
+long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
+                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
+
+summary(long.biv.log.reg)#NS effect (likelt underpowered)
+```
+
+When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
+When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
+
+> Is the prediction, high rec musc male strains have long SC met?
+
+In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
+I couldn't get the Mixed models working for the male polymorphism predictions...
+
+# IFD
+
+Brief lit review for IFD / interference expectations  
+**Petkov et al 2007,  CO interference underlies sex differences in RR**
+
+"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
+However, the interference distance is the same in terms of bivalent length.
+We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
+at the pachytene stage of meiosis."
+
+Chrm1 genetic map,
+
+(other human and mice refs
+
+Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
+
+DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
+
+))
+
+*Do I believe the physical/SC results?*
+
+<!--
+ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
+
+explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
+-->
+
+3) Interference / IFD will not be sexually dimorphic.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
+
+**I**nter**f**ocal **D**istance, a indication of CO interference.
+
+The raw measures are pretty nosy, this might not be the best way to display these results.
+
+ Each point is a IFD observation. Maybe add mouse level lines for the means.
+
+```{r show.IFD_99, echo=FALSE, warning=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+
+#+theme(legend.position="none")
+```
+
+```{r IFD.old.plots, echo=FALSE, eval=FALSE}
+#long SC Long_biv_mouse.avs_4MM
+#real.long.bivData
+#Only use 2CO bivalents
+real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
+real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
+
+long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+```
+
+Still working on the best way to display the general IFD patterns.
+
+### Mixed Model Tests, Fixed Effects
+
+Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+
+$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r HetC_M1_IFD_many.metrics, echo=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+anova(lmer.IFD.PER_MM.M1)
+
+exactRLRT(lmer.IFD.PER_MM.M1)
+as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
+
+#M1_IFD.RAW_2CO
+lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
+anova(lmer.IFD.RAW_MM.M1) # also F values
+
+exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
+as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+```
+
+Sex is a slightly significant factor for raw and normalized IFD.
+
+The table above should display the slightly unsual pattern, where the coefficients for the significant sex fixed effect are positive and neagtive in the raw and normalized values respectively.
+That is for the raw IFD values, females are signiciantly longer but the normalized IFD values, males are significantly longer.
+
+show a raw plot, and norm plot and a scatter plot
+
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
+The tables below are from anova( for the lmer model ).
+Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
+
+## Strain Comparisons
+
+Dive deeper into the sex specific pattern for each strain.
+Below are code chunks which show the unusal sex specific results for IFD measures.
+The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
+The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
+
+For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
+
+```{r WSB.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+#WSB.Bivalents.DF
+
+#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
+
+#plot the differences
+wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_raw
+
+wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_PER
+
+WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+WSB.IFDnorm_scatter.plot
+#this plot is much messier than the MSM (PWD?)
+
+#try plotting the other IFD
+
+#make a table of the means and n ect
+wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+
+                       total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+```
+
+The range of normalized IFDs overlap closer in males and females in the WSB data.
+
+```{r LEW.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+lew.ploty_IFD_raw <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_raw
+
+lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_PER
+
+LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+LEW.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+
+                  #     mean_SC = mean(chromosomeLength),
+                       n_SC = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+
+```
+
+The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
+
+```{r G.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+#plot the metric
+#remove the NAs
+
+G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
+
+G.ploty_IFD_raw
+#
+G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.ploty_IFD_PER
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+#calQ percent < 30%
+
+oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
+oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
+
+#the na's are being counted (ughhhhhh)
+#use the file thing instead
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>% filter(IFD1_PER <= .3) #this makes DF
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male")
+ratio.male <- 18/1031 #2%
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>% filter(IFD1_PER <= .3)
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")
+ratio.female <- 36/766  #5%
+
+LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
+  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
+
+#uggh how does this not work ...
+
+#more males have lower than 30% IFD?
+
+LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
+#0.7062663
+
+#would Chi-sq be a good test to compare the ratios ect?
+
+#the sum of these doesn't add
+
+G.Bivalents.DF
+
+```
+
+```{r PWD.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#remove the NAs
+
+PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
+
+PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
+
+PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+PWD.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+PWD.table <-  ddply(PWD.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+```
+
+For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
+
+```{r KAZ.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+
+#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+
+#remove the NAs
+KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_raw
+
+KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_PER
+
+KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+KAZ.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+KAZ.table <-  ddply(KAZ.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+
+```
+
+For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
+
+```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+
+#remove the NAs
+SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_raw
+
+SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_PER
+
+SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+SKIVE.IFDnorm_scatter.plot
+
+SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+```
+
+In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
+
+```{r MSM.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_raw
+
+MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_PER
+
+#another plot
+MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
+
+MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
+
+MSM.IFDnorm_another.plot
+
+#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
+
+#make a table of the means and n ect
+MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+```
+
+The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
+
+```{r MOLF.IFD, eval=FALSE, echo=FALSE}
+#make the dataset
+MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MOLF.ploty
+#remove the NAs
+molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
+molf.ploty_IFD_raw
+
+molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
+molf.ploty_IFD_PER
+
+#make a table of the means and n ect
+MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
+The other strains which have more overlap between males and females are the Dom strains and KAZ.
+
+#### IFD 3CO bivalents
+
+Run comparisons for 3CO bivalents.
+
+```{r CO3.comparisons, echo=FALSE}
+
+#take a look at the 3CO things
+
+```
+
+<!--
+ Heterochiasmy Prediction
+
+> Do the IFD measures lack significant sex effect? (as predicted)
+
+There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
+
+> Are there signiticant strain effects?
+
+The random strain effect was tested for raw and normalized IFD, neither are significant.
+-->
+
+## Male Polymorphism Prediction
+
+The polymorphism frame-work
+
+> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
+
+logistic regression
+$$  Rec \ Group ~=~ mouse \ average \ IFD $$
+For these tests I am looking at the single IFD measure from 2CO bivalents.
+
+For the logistic regression, comparing High and Low groups (there needs to be 2 groups)
+I think I should compare 2 sets of logistic models, where SKIVE (intermediate gwRR), is in high and low.
+
+```{r male.IFD_strain, echo=FALSE}
+
+#make table
+Musc_male_IFD_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 2,], c("strain"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_IFD = mean(IFD1_ABS, na.rm=TRUE),
+                        mean_IFD.nrm = mean(IFD1_PER,na.rm=TRUE)
+                         )
+
+        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+```
+
+```{r IFD.log.reg, echo=FALSE}
+#
+log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *
+
+log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+summary(log.reg_IFD_PER_mouse.av)#NS
+
+### Single Bivalent levels
+
+log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
+              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_IFD_ABS_single.biv)
+
+log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
+              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_IFD_PER_single.biv)
+```
+
+```{r t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
+
+#mouse.av - the mouse average IFD is from just 2CO
+t.test_mouse.av.IFD_ABS <- t.test(
+  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+t.test_mouse.av.IFD_PER <- t.test(
+
+  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+```
+
+```{r t.test_IFD,echo=FALSE}
+
+#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+#                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#THIS SHOULD ALREADY BE ADDED
+
+#single biv level
+
+Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 ) & (Curated_BivData$hand.foci.count == 2 )],
+
+     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 ) & (Curated_BivData$hand.foci.count == 2 )]
+)
+Rec.group_IFD.ABS_test$p.value
+
+## PER
+
+Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )& (Curated_BivData$hand.foci.count == 2 )],
+
+     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 ) &  (Curated_BivData$hand.foci.count == 2 )]
+)#p-value = 0.04231
+
+Rec.group_IFD.PER_test$p.value
+
+#Single Biv Leve
+single.biv.ABS.t.test <- t.test( Curated_BivData$IFD1_ABS[Curated_BivData$Rec.group == 1],
+
+         Curated_BivData$IFD1_ABS[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
+
+single.biv.IFD_PER.t.test <- t.test( Curated_BivData$IFD1_PER[Curated_BivData$Rec.group == 1],
+
+         Curated_BivData$IFD1_PER[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
+```
+
+```{r show.IFD, echo=FALSE}
+
+biv.IFDs_by_group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,100))
+biv.IFDs_by_group
+#+theme(legend.position="none")+
+
+biv.IFDs_PER_by.group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS \n Low=0, High=1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+biv.IFDs_PER_by.group
+
+oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+
+```
+
+Re-order the above figure if possible.
+
+Neither t.test are sig for both the ABS and PER when I test just the Musc strains.
+The above t.tests are breaking the knitr
+
+The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
+
+```{r IFD.poly_log.reg, echo=FALSE, eval=FALSE}
+#This might be a duplicate code chunk
+#test. for the high vs low strains and IFD
+#mouse.avs_4MM_male
+
+mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc", ]
+
+log.reg_IFD.ASB <- glm(Rec.group ~ mean_IFD.2CO_ABS, 
+                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
+#slightly sig
+
+ #summary(log.reg_IFD.ASB)
+
+log.reg_IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER, 
+                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
+ #summary(log.reg_IFD.PER)
+
+#not significant
+```
+
+None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
+
+Post-hoc comparisons ideas?
+
+Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
+This also indicated that interference distance may evolved in the house mouse complex.
+
+> Prediction B, Is there evidence for the alternative IFD meansure
+
+# General CO Positions
+
+There are a few traits that fall within the CO positions
+
+- 1CO normalized position
+
+- sis.co.ten (sorta interference)
+
+- centromere and telomere distance
+
+#### Predictions
+
+**Heterochiasmy**
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+**Polymorphism**
+
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
+
+C.3) telomere and centromere distances will be shorter in low rec strains
+
+# Normalized 1CO positions
+
+```{r CO.positions.figs, echo=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
+This plot shows the sexual dimorphism in the density plots.
+
+Consider adding annotate_text for the number of observations in each category.
+think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
+
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+#real.long.bivData
+#MELT_Curated_BivData
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+```
+
+<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
+
+```{r CO1_foci_pos_boxplot, echo=FALSE}
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+##### Mixed models norm 1CO positions
+
+the distribution of SC lengths and sis-coten seems very different across sexes
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+The mixed model data should only come from 1CO bivalent data.
+
+```{r MM_CO_pos, echo=FALSE}
+#use the 1CO
+#Use MELT DF
+#MELT_Curated_BivData
+
+#isolate the 1COPos.CO.avs
+#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
+#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
+#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
+
+mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
+
+#<- lme()
+lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
+
+#can't make the 
+CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+#library(lmerTest)
+#anova(CO_position.MM)
+#anova(CO_position.MM, verbose = TRUE)
+
+#summary(CO_position.MM) #no p values,,
+#these just give t and F stats, no pvalue
+
+#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
+#CO_pos_fix.est
+#this test, the full model is more predictive than an empty model?
+
+#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
+#thing2 <- drop1(SC_length.MM, test="Chisq")
+
+CO_pos_rand <- exactRLRT(CO_position.MM)
+CO_pos_rand
+```
+
+```{r log.reg.CO1_pos, echo=FALSE}
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+```{r t.test_1CO_pos, echo=FALSE}
+
+#mouse av
+t.test_mouse.av.1CO_pos <- t.test(
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv.1CO_pos_t.test <- t.test(
+
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+  )
+
+```
+
+the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
+### Siscoten
+
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males (which might)
+
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE}
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE}
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE}
+#mouse av
+t.test_mouse.av_siscoten <- t.test(
+
+  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
+
+   mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
+
+   mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
+
+   mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+#2CO
+
+#3CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+
+lme_MM_telo.dist <- lme(mean.telo.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_telo.dist)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE}
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
+
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
+
+> is random strtain effect significant?
+
+The random strain effect seems very significant.
+
+remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
+
+```{r pos_t.test}
+#make point plots / boxplots which show differences in mean positions
+
+#scatter + boxplot for t.tests
+
+```
+
+Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
+
+```{r F1.by.F2}
+
+#try remaking the plot Megan suggested
+# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
+
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
+#isolate 2COs
+#facet by sex and subsp
+
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
+
+```
+
+```{r CO.pos.ANOVA, eval=FALSE}
+#what is the pattern of variance
+#run analyses for each subsp*sex
+#use non-melt DF
+
+#how is the variance partioned across
+#cell, mouse, strain
+
+female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
+female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
+
+female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
+
+#unorder strain and mouse
+
+female.Dom$mouse <- as.factor(female.Dom$mouse)
+
+female.Dom$strain <- unclass(female.Dom$strain)
+female.Dom$strain <- as.factor(female.Dom$strain)
+
+female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
+female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
+
+#1CO first
+modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
+
+#can't get mouse and strain to have sum of square
+#residual size decreases with per.F1
+#residuals much larger than fileName, mouse and strain no
+
+#model <- lm(breaks ~ wool * tension, 
+#            data = warpbreaks, 
+#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
+
+male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
+male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
+
+male.Dom$mouse <- as.factor(male.Dom$mouse)
+
+male.Dom$strain <- unclass(male.Dom$strain)
+male.Dom$strain <- as.factor(male.Dom$strain)
+
+male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
+male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
+
+male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
+
+male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
+summary(aov(male.modo))
+
+#only file name is registering as effect
+#Review ANOVA frameworks
+#http://www.biostathandbook.com/nestedanova.html
+
+```
+
+### Male Polymorphism Prediction
+
+I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
+
+```{r male.Nrm1CO_strain, echo=FALSE}
+
+#isolate all musc
+Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+#make table
+Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
+
+                         )
+
+        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+```
+
+#Caveats
+
+put all of the code chuncks/analysis for caveats here
+
+## Chromosome Size Effect
+
+I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)
 
@@
 
-# Adjusting for XX
-
+## Adjusting for XX
+
+(rethink where this section should go)
 new outline
@@
 
-# Applying Mix Model Framework
-
-I will use two different mixed models using lme() and lmer(), because the different models have limitations for reporting p values for fixed and random effects.
-
-Pros and con list for using mouse averages vs single bivalent traits.
-
-**Pro Bivalent level**
-
-- The data are continuous and closer to normal/Gaussian expectations than the MLH1 counts.
-
-- There is within cell variance that is an issue.
-
-**Pro Mouse level averages**
-
-- Mouse average is a conserved framework, it would summarize general patterns which fits with the paper.
-
-- Simple application of the same MM.
-
-- Limited work to add to the data set if I can justify that the collection of single bivalents are random and equivalent across mice.
-
-The mouse averages for all the metrics is what is used in the mixed models.
-
-```{r biv.data_table.4MM.HetC, echo=FALSE}
-
-#the mouse averages will be used
-
-mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-mouse.avs_4MM <- add_sex(mouse.avs_4MM)
-mouse.avs_4MM <- add_strain(mouse.avs_4MM)
-mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-mouse.avs_4MM <- add_category(mouse.avs_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
-mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
-mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
-
-#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-#code if males are high or low?
-
-#code intp chr
-
-#assign high and low
-#divide musc males into groups
-mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
-
-```
-
-In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .
-
-```{r biv.data_table.4MM.Male, echo=FALSE}
-#remove female, 
-#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
-
-#Curated_BivData_male
-
-male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-male.av_4MM <- add_sex(male.av_4MM)
-male.av_4MM <- add_strain(male.av_4MM)
-male.av_4MM <- add_subsp(male.av_4MM)
-male.av_4MM <- add_category(male.av_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
-male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
-male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-
-```
-
-## Predictions, Heterochiasmy
-
-Using the Mixed model framework which tests the effects (and interactions) of subspecies, sex and strain, I will test for evolution of the following traits.
-
-Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
-
-1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of uniform vs telomere/biased CO positioning.
-
-  C) Centromere and telomere distances will be sexually dimorphic.
-
-3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
-
-## Predictions, Male Polymorphism (High vs Low Rec strains)
-
-Rerun all these predictions for MSM > MOLF as well
-
-A) **SC lengths** will be longer for high Rec strains.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains.
-
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
-
-C.3) telomere and centromere distances will be shorter in low rec strains
-
-Using the same MM framework we would expect the random strain effect would be significant for these predictions.
-
-Use just the male (Musc male data)
-
-$$mouse \ average \ SC  \ length ~=~ rand(strain) + \varepsilon $$
-
-$$mouse \ average \ IFD \ trait  ~=~ rand(strain) + \varepsilon $$
-The IFD traits are:
-
-- IFD_ABS
-
-- IFD_PER
-
-$$mouse \ average \ CO \ position \ trait  ~=~ rand(strain) + \varepsilon $$
-
-The CO position traits are:
-
-- 1CO position
-
-- centromere and telomere distance
-
-- sis-co-ten
-
-# SC Lengths
-
-<!--Brief lit review for SC length and MLH1 expections. -->
-
-We expect female SC lengths to be longer (refs).
-
-In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
-
-For the 0 class chromosomes most all are around the same size of the 1CO distribution.
-
-Add in the code for the plot under 2COs males and females
-
-```{r show.SC.length, echo=FALSE, warning=FALSE}
-#for now remove the 
-#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
-#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS
-
-#add boxplots for mouse level means
-#add second x axis, hand.foci.count
-
-#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#nested boxplots
-SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
-
-SC.length_Nested_boxplots
-
-#+ +geom_jitter(aes(alpha=0.2))
-```
-
-The dependant variables I'll be testing in the mixed model framework are:
-
-- Pooled SC lengths
-
-- 1CO
-
-- 2CO
-
-- 3CO
-
-- long.biv
+# Trash / Extra
 
 <!--
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
-
-To estimate random effects, I'm using lmer (not lme where I could do the rand(strain*sex)). So now it's just random(strain).
-
-The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.
-
-remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
-
-#### Pooled SC Lengths
-
-```{r MM_2.SC.Length, echo=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-#p
-
-```
-
-I used the mixed model with the lme() for the full Curated BivData and long bivdata
+
+```{r trackcurBivData, echo=FALSE, eval=FALSE}
+#REMOVE THIS BLOCK??
+
+#list / tally of current curated Bivdata
+
+Cur_biv = read.csv("~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", header = TRUE)
+
+colnames(Cur_biv) <- c('mouse', 'blank','0','1')
+
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages[ -c(9:12, 14, 18, 22:24) ]
+
+AP_mouse_table_w.Ages_HQ <- add_sex(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_strain(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_category(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_subsp(AP_mouse_table_w.Ages_HQ)
+AP_mouse_table_w.Ages_HQ <- add_species(AP_mouse_table_w.Ages_HQ)
+
+#only M.musculus
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$species == "M.musculus",]
+#REMOVE EXTRA STRAINS
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$subsp != "Cast",]
+#remove musc strains
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "AST",]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "CZECH",]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "PERC",]
+#156 mice (from 193)
+
+#some mice are missing 13oct14
+
+#match the MM dataset
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$Ncells > 9,]
+AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$age.weeks < 26,]
+#126 mice
+
+#AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[!(is.na(AP_mouse_table_w.Ages_HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
+
+#mice missing from curated list
+mice.within.BivData <- AP_mouse_table_w.Ages_HQ[ AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#54 mice in (14 mice have enteries but 0 curated biv)
+
+#(14 mice need at least some curation
+
+cur.data.mice <- Cur_biv[ Cur_biv$mouse %in% AP_mouse_table_w.Ages_HQ$mouse, ]
+
+mice.missing.bivData <- AP_mouse_table_w.Ages_HQ[ ! AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#72 (includes mice not in MM)
+#72 mice that I need to upload/run the algorithm on their images
+
+#mice missing SC pass == '1'
+colnames(Cur_biv) <- c("mouse","blank","SC.fail", "SC.pass")
+NeedsCur <- Cur_biv[Cur_biv$SC.fail == 0,]
+
+NeedsCur <- add_strain(NeedsCur)
+write.table(NeedsCur, "~./MLH1repo/data/needs_cur.csv", sep=",",
+            row.names = TRUE)
+
+#44% of mice are missing curation
+
+#write this table
+#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
+#            row.names = TRUE)
+
+mice.missing.bivData <- mice.missing.bivData[!(is.na(mice.missing.bivData$mouse) | mice.missing.bivData$mouse==""), ]
+write.table(mice.missing.bivData, "~./MLH1repo/data/MissingMice_13.1.20.csv", sep=",",
+            row.names = TRUE)
+
+```
+
+```{r var.contrast, echo=FALSE, eval=FALSE}
+#contrast models -- what's the points for this block
+# review this logic again
+full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
+
+full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
+
+strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
+#for var, p = 0.0219
+```
+
+```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}
+
+#this chunk migth need to be deleted // I think everything is already done within the BivData
+#this is a second mode of
+
+#bivalent data might be too big..
+#load(file="~./MLH1repo/data/CleanBivData.RData")
+#Curates_BivData -- hand.foci.count!
+
+#For each mouse I need to count up
+#the number of bivalent classes are not counted up
+
+chrm.propotion.category <- ddply(.data=Curated_BivData, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(fileName)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = length(Obj.ID),
+
+                 n0CO = length(which(hand.foci.count == 0)),
+                  n1CO = length(which(hand.foci.count == 1)),
+                  n2CO = length(which(hand.foci.count == 2)),
+                  n3CO = length(which(hand.foci.count == 3)),
+                  #n4CO = sum(n4CO),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv )
+                # per4CO= sum(n4CO)/ tot_biv  )
+
+Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
+  "per0CO", "per1CO", "per2CO", "per3CO"), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
+
+# I think I need to make 2 melt tables -- for count and proportions
+Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
+  id=c("category", "tot_biv") )
+colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
+
+Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
+colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
+
+Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
+
+Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
+
+Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
+ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
+      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
+        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
+        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
+        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
+        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
+
+Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
+                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
+#no mol yet
+
+Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
+
+#need to make list of positions
+
+chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
+
+chrm_prop.BivData_plot
+
+```
 
@@
 
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-#### HetC_M1
-
-Insert the results for the MMs using the dependant variable 1CO - use the Mixed model (lmer), for all of the meterics (pooled SC, 1CO 2CO ect)
-
-Below's the code for Mixed model results -- try to organize them into a table or somethinf
-
-```{r M1_many.metrics, echo=FALSE}
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#all pooled SC lengths
-lmer.SCLength_all_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
-anova(lmer.SCLength_all_MM.M1) # also F values
-
-exactRLRT(lmer.SCLength_all_MM.M1)
-as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
-#can't get p values from this summary
-
-#1CO
-SCLength_1CO_MM.M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-drop1(SCLength_1CO_MM.M1, test="Chisq")
-exactRLRT(SCLength_1CO_MM.M1)
-as.data.frame(round(summary(SCLength_1CO_MM.M1)$coefficients, 6))
-
-#2CO
-SCLength_2CO_MM.M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-drop1(SCLength_2CO_MM.M1, test="Chisq")
-exactRLRT(SCLength_2CO_MM.M1)
-as.data.frame(round(summary(SCLength_2CO_MM.M1)$coefficients, 6))
-
-#Long.Biv
-SCLength_long.biv_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
-
-drop1(SCLength_long.biv_MM.M1, test="Chisq")
-exactRLRT(SCLength_long.biv_MM.M1)
-as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-
-```
-
-#### HetC_M2
-
-All fixed and all interactions
-
-```{r M2_many.metrics, echo=FALSE, eval=FALSE}
-
-#feed these 2 DFs into M2
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-```
-
-#### HetC_M3
-
-```{r HetC_M3, echo=FALSE}
-
-```
-
-## Strain Comparisons
-
-Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
-
-(adding all of these blocks to -- the setup script); strain.Bivalents.DF
-
-```{r strain.SClength.t.test, echo=FALSE}
-#t.tests for all the strain specific data frames
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-#### Male Poly
-
-These are the models for male polymorphism, (I am re-thinking the tests I think)
-
-Use these DFs for the models:  **Male.poly.Mouse.Table_BivData_4MM**  and **Long.biv_mouse_table_male.poly**
-
-```{r SC.length.male_tests, echo=FALSE}
-## run the tests to check following expectations
-
-#1.Dom strains not different from ech other
-#2. PWD and SKIVE are different from the other strains
-#3.MSM and MOLF are different from each other
-
-#Male.poly.Mouse.Table_BivData_4MM
-#Long.biv_mouse_table_male.poly
-
-#decide between lm or glm..?
-#mouse avs, males
-
-male.Dom.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Dom"), ]
-#the subsp labels get messed up, unless they are designated chr first
-male.Musc.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == 2), ]
-
-mouse.avs_4MM$subsp <- as.character(mouse.avs_4MM$subsp)
-mouse.avs_4MM$subsp <- ifelse(grepl("MSM", mouse.avs_4MM$strain), "Mol", 
-        ifelse(grepl("MOLF", mouse.avs_4MM$strain), "Mol", mouse.avs_4MM$subsp))
-male.Mol.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Mol"), ]
-
-Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
-
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  )
-Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
-#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
-#total sc sig, 1CO sig, 2CO sig
-
-Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
-#0.000369 ***
-
-Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
-#0.000816 ***
-
-Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
-#0.0148 *
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
-#making plots above show that mouse average SC lengths are lower could there be microscope issue?
-
-#make sure to remove some of the juvi mice!! from the AP_mouse table
-
-show.Dom.male.CO <- ggplot(
-AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") & (AP_mouse_table_w.Ages$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = AP_mouse_table.HQ[(AP_mouse_table.HQ$sex == "male") & (AP_mouse_table.HQ$subsp == "Dom"),])
-#P not sig for CO mean
-
-#seems like there are more mice in the MLH1 data set than bivData
-#
-
-#WSB all dissected and imaged in '15
-#G imaged 15 - 16
-#LEW imaged 15 to 17
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "1CO SC Length", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "2CO SC Length", xlab = "")
-```
-
-```{r SC.length.pairwise.T, echo=FALSE}
-#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
-
-male.Musc.mouse.av <- male.av_4MM[(male.av_4MM$subsp == "Musc")&(male.av_4MM$sex == "male"),]
-
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
-#PWD and MSM are most diff from SKIVE?
-
-Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
-#the pooled SC value tests are sig
-
-#the mol comparisons are different due to adj p values
-
-Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
-#no sig (only 3) -- does include CZECCH
-
-#non-sig, with mol pooled with musc
-
-Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
-
-Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
-
-Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
-#non of these sig with the mulitple adjust
-
-Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
-
-#pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```
-
-If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
-
-Ideas for the above 3 tests,
-
-1. no difference in DOM
-- no sig logistic regression test
-- glm, strain effect?
-
-### Summary
-
-> Is sex a significant effect for SC length? (as predicted)
-
-- Sex is a significant effect for SC length
-The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
-
-- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
-
-- The Long Biv Data set largely agrees with the full curated dataset
-
-- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
-
-> Is the random Effect of strain an effect on SC length?
-
-The exactRLRT() test indicates that the random strain effect might be significant, p= xx
-
-(merge this with the code above)
-Prediction, High rec males have longer SC.
-
-```{r log.reg_SC.length, echo=FALSE}
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
-summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
-summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
-
-summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
-summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_SC.length_BIV)
-
-#1CO
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 2),], family=binomial(link="logit"))
-summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 3),], family=binomial(link="logit"))
-
-summary(log.reg_SC.length_BIV_3CO)
-```
-
-The mean SC logistic regresion model for mouse averages
-
-`r log.reg_SC.length_mouse.av$coefficients`
-
-and for the single bivalent levels
-
-`r log.reg_SC.length_BIV$coefficients`
-
-#### t.test for SC Length
-
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```{r SC.Length_t.test, echo=FALSE}
-
-T.test_SC.Len_mouse.av <- t.test(mouse.avs_4MM$mean_SC[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean_SC[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")&(mouse.avs_4MM$Rec.group == 0)]
-)#p-value = 0.001567
-
-#1CO
-T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#2CO
-T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#3CO
-T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#Single Biv level
-T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
-
-#1CO
-T.test_SC.Len_single.biv_1CO <- t.test(
-  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
-
-      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
-  )
-
-#2CO
-T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
-
-                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
-  )
-
-  #3CO
-  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
-
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
-  )
-```
-
-The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_SC.Len_mouse.av$p.value` and p=`r T.test_SC.Len_single.biv$p.value`.
-
-```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
-#I don't think the Mixed models fit well for the high vs low polymorphism tests.
-
-#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
-#MSM_SC vs MOLF_SC
-
-#2. logistic regression for sc length and chrm class?
-#if the prediction / significance
-
-#isolate data
-#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
-
-#build and test MM
-lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-
-lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
-
-anova(lmer_MM_SC_length_male.poly)
-
-rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
-```
-
-I can't get the above mixed models to work for testing high vs low rec groups
-
-```{r long.sc_log.res_meanSC, eval=FALSE}
-#hold off on reporting/ working with the long SC biv resutls
-
-long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
-long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
-                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
-
-long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
-                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
-
-summary(long.biv.log.reg)#NS effect (likelt underpowered)
-```
-
-When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
-When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
-
-> Is the prediction, high rec musc male strains have long SC met?
-
-In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
-I couldn't get the Mixed models working for the male polymorphism predictions...
-
-# IFD
-
-Brief lit review for IFD / interference expectations  
-**Petkov et al 2007,  CO interference underlies sex differences in RR**
-
-"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
-However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
-at the pachytene stage of meiosis."
-
-Chrm1 genetic map,
-
-(other human and mice refs
-
-Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
-
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-
-))
-
-*Do I believe the physical/SC results?*
-
-<!--
-ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
-
-explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
--->
-
-3) Interference / IFD will not be sexually dimorphic.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
-
-**I**nter**f**ocal **D**istance, a indication of CO interference.
-
-The raw measures are pretty nosy, this might not be the best way to display these results.
-
- Each point is a IFD observation. Maybe add mouse level lines for the means.
-
-```{r show.IFD_99, echo=FALSE, warning=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r IFD.old.plots, echo=FALSE, eval=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
-Still working on the best way to display the general IFD patterns.
-
-### Mixed Model Tests, Fixed Effects
-
-Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
-
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r HetC_M1_IFD_many.metrics, echo=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-anova(lmer.IFD.PER_MM.M1)
-
-exactRLRT(lmer.IFD.PER_MM.M1)
-as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
-
-#M1_IFD.RAW_2CO
-lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
-anova(lmer.IFD.RAW_MM.M1) # also F values
-
-exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
-as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-```
-
-Sex is a slightly significant factor for raw and normalized IFD.
-
-The table above should display the slightly unsual pattern, where the coefficients for the significant sex fixed effect are positive and neagtive in the raw and normalized values respectively.
-That is for the raw IFD values, females are signiciantly longer but the normalized IFD values, males are significantly longer.
-
-show a raw plot, and norm plot and a scatter plot
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
-The tables below are from anova( for the lmer model ).
-Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
-
-## Strain Comparisons
-
-Dive deeper into the sex specific pattern for each strain.
-Below are code chunks which show the unusal sex specific results for IFD measures.
-The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
-The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
-
-For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
-
-```{r WSB.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#WSB.Bivalents.DF
-
-#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
-
-#plot the differences
-wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_raw
-
-wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_PER
-
-WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-WSB.IFDnorm_scatter.plot
-#this plot is much messier than the MSM (PWD?)
-
-#try plotting the other IFD
-
-#make a table of the means and n ect
-wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
-
-                       total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
-
-The range of normalized IFDs overlap closer in males and females in the WSB data.
-
-```{r LEW.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-lew.ploty_IFD_raw <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex))
-lew.ploty_IFD_raw
-
-lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
-lew.ploty_IFD_PER
-
-LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-LEW.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-
-                  #     mean_SC = mean(chromosomeLength),
-                       n_SC = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-
-```
-
-The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
-
-```{r G.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-#plot the metric
-#remove the NAs
-
-G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
-
-G.ploty_IFD_raw
-#
-G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.ploty_IFD_PER
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-
-#calQ percent < 30%
-
-oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
-oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
-
-#the na's are being counted (ughhhhhh)
-#use the file thing instead
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>% filter(IFD1_PER <= .3) #this makes DF
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male")
-ratio.male <- 18/1031 #2%
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>% filter(IFD1_PER <= .3)
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")
-ratio.female <- 36/766  #5%
-
-LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
-  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
-
-#uggh how does this not work ...
-
-#more males have lower than 30% IFD?
-
-LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
-#0.7062663
-
-#would Chi-sq be a good test to compare the ratios ect?
-
-#the sum of these doesn't add
-
-G.Bivalents.DF
-
-```
-
-```{r PWD.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#remove the NAs
-
-PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
-
-PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
-
-PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-PWD.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-PWD.table <-  ddply(PWD.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-```
-
-For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
-
-```{r KAZ.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-
-#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_raw
-
-KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_PER
-
-KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-KAZ.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-KAZ.table <-  ddply(KAZ.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-
-```
-
-For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
-
-```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_raw
-
-SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_PER
-
-SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-SKIVE.IFDnorm_scatter.plot
-
-SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-```
-
-In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
-
-```{r MSM.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_raw
-
-MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_PER
-
-#another plot
-MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
-
-MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
-
-MSM.IFDnorm_another.plot
-
-#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
-
-#make a table of the means and n ect
-MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-```
-
-The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
-
-```{r MOLF.IFD, eval=FALSE, echo=FALSE}
-#make the dataset
-MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MOLF.ploty
-#remove the NAs
-molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
-molf.ploty_IFD_raw
-
-molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
-molf.ploty_IFD_PER
-
-#make a table of the means and n ect
-MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
-The other strains which have more overlap between males and females are the Dom strains and KAZ.
-
-#### IFD 3CO bivalents
-
-Run comparisons for 3CO bivalents.
-
-```{r CO3.comparisons, echo=FALSE}
-
-#take a look at the 3CO things
-
-```
-
-<!--
- Heterochiasmy Prediction
-
-> Do the IFD measures lack significant sex effect? (as predicted)
-
-There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
-
-> Are there signiticant strain effects?
-
-The random strain effect was tested for raw and normalized IFD, neither are significant.
--->
-
-## Male Polymorphism Prediction
-
-The polymorphism frame-work
-
-> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
-
-logistic regression
-$$  Rec \ Group ~=~ mouse \ average \ IFD $$
-For these tests I am looking at the single IFD measure from 2CO bivalents.
-
-For the logistic regression, comparing High and Low groups (there needs to be 2 groups)
-I think I should compare 2 sets of logistic models, where SKIVE (intermediate gwRR), is in high and low.
-
-```{r male.IFD_strain, echo=FALSE}
-
-#make table
-Musc_male_IFD_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 2,], c("strain"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_IFD = mean(IFD1_ABS, na.rm=TRUE),
-                        mean_IFD.nrm = mean(IFD1_PER,na.rm=TRUE)
-                         )
-
-        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-```
-
-```{r IFD.log.reg, echo=FALSE}
-#
-log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *
-
-log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-summary(log.reg_IFD_PER_mouse.av)#NS
-
-### Single Bivalent levels
-
-log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
-              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_IFD_ABS_single.biv)
-
-log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
-              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_IFD_PER_single.biv)
-```
-
-```{r t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
-
-#mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-t.test_mouse.av.IFD_PER <- t.test(
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-```
-
-```{r t.test_IFD,echo=FALSE}
-
-#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-#                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-
-Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 ) & (Curated_BivData$hand.foci.count == 2 )],
-
-     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 ) & (Curated_BivData$hand.foci.count == 2 )]
-)
-Rec.group_IFD.ABS_test$p.value
-
-## PER
-
-Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )& (Curated_BivData$hand.foci.count == 2 )],
-
-     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 ) &  (Curated_BivData$hand.foci.count == 2 )]
-)#p-value = 0.04231
-
-Rec.group_IFD.PER_test$p.value
-
-#Single Biv Leve
-single.biv.ABS.t.test <- t.test( Curated_BivData$IFD1_ABS[Curated_BivData$Rec.group == 1],
-
-         Curated_BivData$IFD1_ABS[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
-
-single.biv.IFD_PER.t.test <- t.test( Curated_BivData$IFD1_PER[Curated_BivData$Rec.group == 1],
-
-         Curated_BivData$IFD1_PER[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
-```
-
-```{r show.IFD, echo=FALSE}
-
-biv.IFDs_by_group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,100))
-biv.IFDs_by_group
-#+theme(legend.position="none")+
-
-biv.IFDs_PER_by.group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \n Low=0, High=1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-biv.IFDs_PER_by.group
-
-oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
-
-```
-
-Re-order the above figure if possible.
-
-Neither t.test are sig for both the ABS and PER when I test just the Musc strains.
-The above t.tests are breaking the knitr
-
-The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
-
-```{r IFD.poly_log.reg, echo=FALSE, eval=FALSE}
-#This might be a duplicate code chunk
-#test. for the high vs low strains and IFD
-#mouse.avs_4MM_male
-
-mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc", ]
-
-log.reg_IFD.ASB <- glm(Rec.group ~ mean_IFD.2CO_ABS, 
-                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
-#slightly sig
-
- #summary(log.reg_IFD.ASB)
-
-log.reg_IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER, 
-                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
- #summary(log.reg_IFD.PER)
-
-#not significant
-```
-
-None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
-
-Post-hoc comparisons ideas?
-
-Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
-This also indicated that interference distance may evolved in the house mouse complex.
-
-> Prediction B, Is there evidence for the alternative IFD meansure
-
-# General CO Positions
-
-There are a few traits that fall within the CO positions
-
-- 1CO normalized position
-
-- sis.co.ten (sorta interference)
-
-- centromere and telomere distance
-
-#### Predictions
-
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-
-  C) Centromere and telomere distances will be sexually dimorphic.
-
-**Polymorphism**
-
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
-
-C.3) telomere and centromere distances will be shorter in low rec strains
-
-# Normalized 1CO positions
-
-```{r CO.positions.figs, echo=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-
-```
-
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-#real.long.bivData
-#MELT_Curated_BivData
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-```{r CO1_foci_pos_boxplot, echo=FALSE}
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-##### Mixed models norm 1CO positions
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-The mixed model data should only come from 1CO bivalent data.
-
-```{r MM_CO_pos, echo=FALSE}
-#use the 1CO
-#Use MELT DF
-#MELT_Curated_BivData
-
-#isolate the 1COPos.CO.avs
-#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
-#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
-#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
-
-mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
-
-#<- lme()
-lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
-
-#can't make the 
-CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-#library(lmerTest)
-#anova(CO_position.MM)
-#anova(CO_position.MM, verbose = TRUE)
-
-#summary(CO_position.MM) #no p values,,
-#these just give t and F stats, no pvalue
-
-#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
-#CO_pos_fix.est
-#this test, the full model is more predictive than an empty model?
-
-#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
-#thing2 <- drop1(SC_length.MM, test="Chisq")
-
-CO_pos_rand <- exactRLRT(CO_position.MM)
-CO_pos_rand
-```
-
-```{r log.reg.CO1_pos, echo=FALSE}
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-```{r t.test_1CO_pos, echo=FALSE}
-
-#mouse av
-t.test_mouse.av.1CO_pos <- t.test(
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv.1CO_pos_t.test <- t.test(
-
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-  )
-
-```
-
-the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males (which might)
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE}
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE}
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE}
-#mouse av
-t.test_mouse.av_siscoten <- t.test(
-
-  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
-
-   mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
-
-   mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
-
-   mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-
-lme_MM_telo.dist <- lme(mean.telo.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_telo.dist)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE}
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
-### Heterochiasmy Prediction
-
-> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
-
-> is random strtain effect significant?
-
-The random strain effect seems very significant.
-
-remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
-
-```{r pos_t.test}
-#make point plots / boxplots which show differences in mean positions
-
-#scatter + boxplot for t.tests
-
-```
-
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
-
-```{r F1.by.F2}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-```{r CO.pos.ANOVA, eval=FALSE}
-#what is the pattern of variance
-#run analyses for each subsp*sex
-#use non-melt DF
-
-#how is the variance partioned across
-#cell, mouse, strain
-
-female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
-female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
-
-female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
-
-#unorder strain and mouse
-
-female.Dom$mouse <- as.factor(female.Dom$mouse)
-
-female.Dom$strain <- unclass(female.Dom$strain)
-female.Dom$strain <- as.factor(female.Dom$strain)
-
-female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
-female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
-
-#1CO first
-modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
-
-#can't get mouse and strain to have sum of square
-#residual size decreases with per.F1
-#residuals much larger than fileName, mouse and strain no
-
-#model <- lm(breaks ~ wool * tension, 
-#            data = warpbreaks, 
-#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
-
-male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
-male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
-
-male.Dom$mouse <- as.factor(male.Dom$mouse)
-
-male.Dom$strain <- unclass(male.Dom$strain)
-male.Dom$strain <- as.factor(male.Dom$strain)
-
-male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
-male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
-
-male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
-
-male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
-summary(aov(male.modo))
-
-#only file name is registering as effect
-#Review ANOVA frameworks
-#http://www.biostathandbook.com/nestedanova.html
-
-```
-
-### Male Polymorphism Prediction
-
-I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
-
-```{r male.Nrm1CO_strain, echo=FALSE}
-
-#isolate all musc
-Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-#make table
-Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
-
-                         )
-
-        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-```
-
-```{r CO.pos.poly}
-
-#CO_pos
-
-```
-
-# old junk
-
-<!-- 
-# old drafts / trash pile
-
-```{r long.biv.MM, echo=FALSE, eval=FALSE}
-
-SC.Length_long.biv_MM <- lme(mean_SC ~ subsp * sex, data= Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex),  na.action=na.omit  ) )
-
-#now there is an error after un-ordering them
-
-as.data.frame( round(  summary(SC.Length_long.biv_MM)$tTable, 6 ) )
-
-#model for long Bivs
-#long_bivData_SC.Length_MM <- lme(mean_SC ~ subsp * sex, data= #Long_biv_mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(long_bivData_SC.Length_MM)$tTable, 6 ) )
-```
-
-```{r justify.mouse.level, echo=FALSE, eval=FALSE}
-
-#start the permutation code here
-#make a randomized BivData DF
-
-#30 cells worth of bivs? 
-#30 cells * ( 19 or 20 bivs ) * 70% segment rate = decent sample size for comparing Biv-based patterns?
-# I think this would address if the sample sizes are affecting the pattern
-
-#
-#justify mouse averages
-#is sampling -- bivalent dataset 
-# random samples 
-#equivalent sample size
-
-```
-
-I went on a kick thinking of the sis-coten metric, I'm not so sure this is a good metric anymore.  Do paired t-tests for all the male strains; test is foci position distributions are different.
-
-1CO(F1)  1CO(F1)
-
-2CO(F1) vs 2CO(F1)   2CO(F2) vs 2CO(F2)
-
-```{r block1, echo=FALSE}
-#new.Curate.FULL
-
-#new.Curate.FULL <- add_strain(new.Curate.FULL)
-#new.Curate.FULL <- add_sex(new.Curate.FULL)
-#Curated_BivData
-
-Curated_BivData <- subset(Curated_BivData, hand.foci.count < 5)
-
-staggered <- ggplot(Curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("SC lengths by foci count")+theme(legend.position="none") #+ ylim(c(30,250))
-
-fill.by.ifd <- ggplot(Curated_BivData, aes(y = chromosomeLength, x=hand.foci.count, color=IFD1_PER))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("SC length Fill by IFD")#+theme(legend.position="none") #+ ylim(c(30,250))
-```
-
- Predictions for sis-coten stabilizing selection
-
-Because it enables tighter regulation of the spindle assembly checkpoint (SAC) the centrosomic metaphase spindle in male gametogenesis can impose selection on relative CO placement (broad scale REC landscape).
-(OR generally due to the stronger sensitivity of the SAC in sperm/males).
-Following certain assumptions, the axis length and relative positions of COs can change the amount sister-cohesin over which tension is generated on the metaphase I spindle.
-This can be thought of as the bivalent structure when they are in the process of bipolar orientation.
-
-under strong purifying selection, (the idea is that bivalent structures evolve to efficienctly pass the SAC) (there might be mulitple ways to do this)
-
-- similar amount of sister cohesion tension -related to the rate of sister cohesion degregation.
-So all bivalents seperate at a similar rate (enter anaphase and complete anaphase in a synchronized manner).
-
-- tension for SAC detection
-
- 1. In MSM and PWD males (the high recombining strains), 1COs should have less of a telomere bias compared to the other low rec mice (IFD).
-
-The MSM and PWD males have approximately 50% 1CO, 50% 2CO compared to lower recombining strains (Dom and musc) 75% 1CO, 25% 2CO.
-
-If the metaphase spindle / SAC evolved a new optimal amount of sister cohesion for the spindle tension to be spread over, the parameters of relative co placement and CO interference can evolve so that they are similar across .
-
-```{r rand.1CO.pos, echo=FALSE, warning=FALSE, eval=FALSE}
-
-PER.1CO.pos <- ggplot(Curated_BivData[(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$sex == "male"),], aes(y = PER_Foci_1, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(aes(alpha=.5))+ggtitle("Relative CO Position of 1CO across strains")
-
-LOG.PER.1CO.pos <- ggplot(Curated_BivData[(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$sex == "male"),], aes(y = log(PER_Foci_1), x = strain, color=strain)) + geom_jitter()+ geom_boxplot(aes(alpha=.5))+ggtitle("Log of Relative CO Position of 1CO across strains")
-
-#log(foci position, makes a bigger range), means don't really change
-#but WSB, G, LEW seem to have stronger centromere supression
-PER.1CO.pos
-LOG.PER.1CO.pos
-```
-
-In terms of percent position, PWD, LEW and MSM have similar mean Foci1 pos.
-LEW is sort of the odd one out.
-Interestingly, LEW seems to have stronger centromere supression compared to PWD and MSM (the distribution of foci positions is larger in MSM, PWD and KAZ).
-
- 2. Sister cohesin tension area for 1COs should trend more similar to 2COs in high recombining strains.
-
-what's the rational for this prediction?
-
-```{r siscoten.stabilizing, echo=FALSE, warning=FALSE, eval=FALSE}
-
-plot.1CO <- ggplot(Curated_BivData[(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$sex == "male"),], aes(y = dis.cent, x = chromosomeLength, color=strain)) + geom_jitter()+facet_wrap(~strain)
-
-sis.coten <- ggplot(Curated_BivData[(Curated_BivData$sex == "male"),], aes(y = SisCoTen, x = chromosomeLength, color=strain)) + geom_jitter()+facet_wrap(~strain)
-
-sis.coten.boxplot <- ggplot(Curated_BivData[(Curated_BivData$sex == "male"),], aes(y = SisCoTen, x = strain,  color=hand.foci.count)) +geom_boxplot()+geom_jitter()
-
-#
-sfsL <- ggplot(Curated_BivData[(Curated_BivData$sex == "male"),], aes(strain, SisCoTen, fill=as.factor(hand.foci.count))) + geom_col(position = "dodge")
-
-sfsL <- ggplot(Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$hand.foci.count < 3),], aes(strain, SisCoTen, fill=as.factor(hand.foci.count))) + geom_boxplot(aes()) + geom_jitter(aes(color=hand.foci.count))
-
-Curated_BivData$PER.siscoten <- Curated_BivData$SisCoTen / Curated_BivData$chromosomeLength
-
-boxplot_per.siscoten <- ggplot(Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, PER.siscoten, fill=as.factor(hand.foci.count))) + geom_boxplot(aes()) +  facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("PERCENT SIS-COTEN by mouse")+theme(axis.title.x=element_blank(), axis.text.x=element_blank())
-
-boxplot_siscoten.mouse <- ggplot(Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, SisCoTen, fill=as.factor(hand.foci.count))) + geom_boxplot(aes())+ facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("SISCOTEN by mouse") + theme(axis.title.x=element_blank(), axis.text.x=element_blank())
-
-boxplot_siscoten.mouse
-boxplot_per.siscoten
-
-#+ scale_fill_viridis(name="Migration Rate", discrete = TRUE) + theme(axis.text.x=element_blank()) + xlab("Frequency in the sample") + ylab("Proportion of Total Junctions") + ggtitle("C.
-Generation = 2,000") + ylim(0,1)
-```
-
-Mouse is shown on the x axis.
-It seems like PWD and MSM might have a higher variance in the range od siscoten values for 2COs, but overall the averages sis-co-tens across the strains for 2CO are very similar across all of the mice.
-
-# 3. Lack of strain specific patterns in females
-
-```{r female.siscoten,eval=FALSE,echo=FALSE, warning=FALSE}
-
-boxplot_siscoten.mouse.female <- ggplot(Curated_BivData[(Curated_BivData$sex == "female") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, SisCoTen, fill=as.factor(hand.foci.count))) + geom_boxplot(aes())+ facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("SISCOTEN by mouse, female") + theme(axis.title.x=element_blank(), axis.text.x=element_blank())+ylim(c(0,150))
-boxplot_siscoten.mouse.female
-
-boxplot_PERsiscoten.mouse.female <- ggplot(Curated_BivData[(Curated_BivData$sex == "female") & (Curated_BivData$hand.foci.count < 4),], aes(mouse, PER.siscoten, fill=as.factor(hand.foci.count))) + geom_boxplot(aes())+ facet_wrap(~ strain, nrow = 3, scales="free")  + ggtitle("PER SISCOTEN by mouse, female") + theme(axis.title.x=element_blank(), axis.text.x=element_blank())
-boxplot_PERsiscoten.mouse.female
-
-```
-
-the vls are much closer in females
-
-# SC length traits
-
- Females have higher variance in more meiotic traits
-
-```{r female.SC.plots, echo=FALSE, warning=FALSE}
-
-SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))
-
-SC.lengths.2CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 2,], aes(y = chromosomeLength, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("2COs male and females") +ylim(c(30,250))
-
-SC.lengths.1CO
-SC.lengths.2CO
-```
-
-The above plots seperate bivalent classes. 
-(Make a table of pvalues for t.tests of male to female SC lengths across strains.)
-
-```{r anova, echo=FALSE}
-#try doing anova
-
-ano <- aov(lm(chromosomeLength ~ hand.foci.count+sex + strain,  Curated_BivData))
-summary(ano)
-```
-
-By ANOVA analysis, everything has a significant effect.
-
-# PERMUTATIONS Accounting for sample size differences
-
-Currently the category with the fewest observations is G female (~30). I also have may more Kaz male bivalent measures (1000)
-
-`r kable(table(Curated_BivData$category))`
-
-Below is code where I implement a subsampling approach/permute dataset with randomly sampled bivalent observations. 25 random bivalents are sampled from each category, and this is permuted 1000 times.
-
-The plots shown below are histograms of p values from t.tests for **SC length** across the labeled categories from the replicated/subsampled dataset.
-
-```{r sub.sample, echo=FALSE, eval=FALSE}
-
-#the problem with sub-sampling is that there are so many groups I need to compare
-#Easist is probably to recreate the whole data set
-#G female have the lowest number of obs, 30
-
-#sample from all the categories, the samplings are mini DFs
-#-- then merge them together
-
-#outer loop is the permutations / replications
-BIG.DF.LIST <- replicate(1000,
-
-        #filling the permuted DF is inner loop                 
-         #add subsample to a list, this works!!!
-list( bind_rows(list(
-
-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ],
-                                SubSample.WSB.male <-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB male"), Biv.sample.Number ), ],
-
-                                Curated_BivData[ sample( which(Curated_BivData$category=="G female"), Biv.sample.Number ), ],
-                                Curated_BivData[ sample( which(Curated_BivData$category=="G male"), Biv.sample.Number ), ],
-
-                       Curated_BivData[ sample( which(Curated_BivData$category=="LEW female"), Biv.sample.Number ), ], 
-                       Curated_BivData[ sample( which(Curated_BivData$category=="LEW male"), Biv.sample.Number ), ],
-
-                       Curated_BivData[ sample( which(Curated_BivData$category=="PWD female"), Biv.sample.Number ), ], 
-                       Curated_BivData[ sample( which(Curated_BivData$category=="PWD male"), Biv.sample.Number ), ],
-
-                       Curated_BivData[ sample( which(Curated_BivData$category=="MSM female"), Biv.sample.Number ), ], 
-                       Curated_BivData[ sample( which(Curated_BivData$category=="MSM male"), Biv.sample.Number ), ],
-
-                        Curated_BivData[ sample( which(Curated_BivData$category=="KAZ male"), Biv.sample.Number ), ], 
-                       Curated_BivData[ sample( which(Curated_BivData$category=="KAZ female"), Biv.sample.Number ), ],
-
-                       Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE female"), Biv.sample.Number ), ], 
-                       Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE male"), Biv.sample.Number ), ]))
-)
-      )
-#Accessing big list / matrix
-#the replicated DFs are now is a list of DFs
-#head(BIG.DF.LIST[[2]]) returns head of the 2nd replicated DF
-
-#Loop through the list of DFs and do the t.tests!
-#i is this index of the dataframes in the list?
-
-p.vals.DF <- data.frame(sex.test.pval = rep(0, length(BIG.DF.LIST) ) )
-for(i in 1:length(BIG.DF.LIST)){  #current data from
-  #print(i)
-  #test sex
-  p.vals.DF$sex.test.pval[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[BIG.DF.LIST[[i]]$sex == "female"],
-         BIG.DF.LIST[[i]]$chromosomeLength[BIG.DF.LIST[[i]]$sex == "male"], paired = FALSE)$p.value
-
-   #test female subsp
-   p.vals.DF$female.subsp[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")], paired = FALSE )$p.value
-
-   #test male subsp
-     p.vals.DF$male.subsp[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")],paired = FALSE )$p.value
-
-      p.vals.DF$male.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value
-
-      p.vals.DF$male.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            p.vals.DF$male.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value
-
-      #what about a null test?
-
-      #female    
-      p.vals.DF$female.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value
-
-            #female    
-      p.vals.DF$female.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            #female    
-      p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value
-
-      #test male strains (Dom)
-     p.vals.DF$male.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value
-
-      p.vals.DF$male.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value
-
-      p.vals.DF$male.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            p.vals.DF$male.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value
-
-       #test female strains (Dom)
-     p.vals.DF$female.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value
-
-      p.vals.DF$female.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value
-
-      p.vals.DF$female.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            p.vals.DF$female.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$chromosomeLength[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value
-}
-
-#these tests might be more appropriate for averages (mouse averages)?
-
-#melt the DF's within the BIG list for easier ggploting
-Melted.pval.DF <- melt(p.vals.DF, c("sex.test.pval",    "female.subsp",     "male.subsp",       "male.PWD.MSM",     "male.PWD.KAZ",     "male.PWD.SKIVE",   "female.PWD.MSM", "female.PWD.KAZ",   "female.PWD.SKIVE", "male.WSB.G", "male.WSB.LEW", "male.WSB.KAZ",  "male.WSB.PWD","female.WSB.G", "female.WSB.LEW", "female.WSB.KAZ", "female.WSB.PWD" ),
-                         id=c() )
-
-#clean up dm1 (there are a bunch of rows without useful information)
-Melted.pval.DF <- Melted.pval.DF[(!is.na(Melted.pval.DF$value)),]
-colnames(Melted.pval.DF) <- c("Test.type", "pval")
-
-#BIG ggplot collections (skip the first type?)
-biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
-#I could also add the observed t.test pvalue as verticle line...
-biggie
-#want to try spreading the data across mulitple plots
-#this is not saving mulitple pdfs
-#for (i in seq(1, length(unique(Melted.pval.DF$Test.type)), 6)) {
-#  pdf(paste0("tryplot_",i,".pdf"), 7, 5)
-# 
-#  print(ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests"))
-#
-# dev.off()
-#
-# }
-
-#pvals.subsamp.DF.1 <- ggplot(p.vals.DF, aes(male.PWD.MSM ) )+geom_histogram(bins = 50)#PWD and MSM not sig different (with randomized subsample)
-#pvals.subsamp.DF.B <- ggplot(p.vals.DF, aes(male.PWD.KAZ ) #)+geom_histogram(bins = 50)
-#pvals.subsamp.DF.C <- ggplot(p.vals.DF, aes(male.PWD.SKIVE ) #)+geom_histogram(bins = 50)
-
-#Fpvals.subsamp.DF.A <- ggplot(p.vals.DF, aes(female.PWD.MSM ) #)+geom_histogram(bins = 50)
-#Fpvals.subsamp.DF.B <- ggplot(p.vals.DF, aes(female.PWD.KAZ ) #)+geom_histogram(bins = 50)
-#Fpvals.subsamp.DF.C <- ggplot(p.vals.DF, aes(female.PWD.SKIVE ) #)+geom_histogram(bins = 50)
-#these 3 female plots have close to uniform pvalus dist
-
-#summaries to compare / test
-#SC.by.sex.T.test <- t.test(subsampled.DF$chromosomeLength[subsampled.DF$sex == "female"],
-#                           subsampled.DF$chromosomeLength[subsampled.DF$sex == "male"], paired = #FALSE)$p.value
-
-#SubSample.WSB.female  <-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB female"), Biv.sample.Number ), ]
-#SubSample.WSB.male <-  Curated_BivData[ sample( which(Curated_BivData$category=="WSB male"), Biv.sample.Number ), ]
-#SubSample.G.female <- Curated_BivData[ sample( which(Curated_BivData$category=="G female"), Biv.sample.Number ), ]
-#SubSample.G.male <- Curated_BivData[ sample( which(Curated_BivData$category=="G male"), Biv.sample.Number ), ]
-#SubSample.LEW.female <- Curated_BivData[ sample( which(Curated_BivData$category=="LEW female"), Biv.sample.Number ), ]
-#SubSample.LEW.male <- Curated_BivData[ sample( which(Curated_BivData$category=="LEW male"), Biv.sample.Number ), ]
-#SubSample.PWD.female <- Curated_BivData[ sample( which(Curated_BivData$category=="PWD female"), Biv.sample.Number ), ]
-#SubSample.PWD.male <- Curated_BivData[ sample( which(Curated_BivData$category=="PWD male"), Biv.sample.Number ), ]
-#SubSample.MSM.female <- Curated_BivData[ sample( which(Curated_BivData$category=="MSM female"), Biv.sample.Number ), ]
-#SubSample.MSM.male <- Curated_BivData[ sample( which(Curated_BivData$category=="MSM male"), Biv.sample.Number ), ]
-#SubSample.KAZ.female <- Curated_BivData[ sample( which(Curated_BivData$category=="KAZ female"), Biv.sample.Number ), ]
-#SubSample.KAZ.male <- Curated_BivData[ sample( which(Curated_BivData$category=="KAZ male"), Biv.sample.Number ), ]
-#SubSample.SKIVE.female <- Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE female"), Biv.sample.Number ), ]
-#SubSample.SKIVE.male <- Curated_BivData[ sample( which(Curated_BivData$category=="SKIVE male"), Biv.sample.Number ), ]
-
-#how do I make this replicate loop so that is returns the T.test pvalue in a DF?
-# 1. make a big list of many subsampled DFs
-# 2. make a loop for applying t.test across all of the
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
-```{r lm.sc, echo=FALSE}
-#should cell be random?
-#should I compare mouse means instead?
-sc.model.1CO <- lm(chromosomeLength ~ sex + strain +mouse, data = Curated_BivData[Curated_BivData$hand.foci.count == 1,])
-
-sc.model.2CO <- lm(chromosomeLength ~ sex + strain+fileName, data = Curated_BivData[Curated_BivData$hand.foci.count == 2,])
-
-```
-
-Modeling SC lengths at the bivalent level is noisey. There are many levels and unequal sampling.
-
-maybe I should compare the longest bivalent from a cell?  or the shortest 4?
-
-# Centromere supression
-
-Anecdotally I would say that males have stronger centromere supression. I will also add a metric for telomere distance.
-
-```{r cent.supression.plot, echo=FALSE, warning=FALSE, eval=FALSE}
-#I need to change the purple color
-
-#simple plots for centromere - dis
-cent.distance.all <- ggplot(Curated_BivData, aes(y = dis.cent, x=strain, color=sex))+geom_boxplot(aes(alpha=0.1))
-#+geom_jitter(aes(alpha=0.05))+  ggtitle("1COs male and females")
-
-cent.distance.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = dis.cent, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females")
-
-cent.dist.2CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 2,], aes(y = dis.cent, x=strain, fill=sex))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("2COs male and females")
-
-cent.distance.all
-cent.distance.1CO
-cent.dist.2CO
-```
-
-Intesting that the distance from centromere is sexually dimorphic for 2CO bivalents but not 1CO bivalents. Females have a stronger centromere supression, strange. Did I calculate this wrong?
-This indecates that female has stronger centromere supression, but only for 2CO bivalents...
-
-```{r telomere.bias, echo=FALSE, eval=FALSE}
-
-#I have to change the t.tests
-#dis.cent
-
-#BIG.DF.LIST
-Cent.Dist.p.vals.DF <- data.frame(sex.cent.dist = rep(0, length(BIG.DF.LIST) ) )
-for(i in 1:length(BIG.DF.LIST)){  #current data from
-  #print(i)
-
-      #female vs male
-      Cent.Dist.p.vals.DF$sex.cent.dist[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" )],paired = FALSE )$p.value
-
-#subsp
-Cent.Dist.p.vals.DF$female.subsp[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")],paired = FALSE )$p.value
-
-Cent.Dist.p.vals.DF$male.subsp[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")],paired = FALSE )$p.value
-
-Cent.Dist.p.vals.DF$male.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$male.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            Cent.Dist.p.vals.DF$male.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value
-
-      #what about a null test?
-
-      #female    
-      Cent.Dist.p.vals.DF$female.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")],paired = FALSE )$p.value
-
-            #female    
-      Cent.Dist.p.vals.DF$female.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            #female    
-      Cent.Dist.p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$female.PWD.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")],paired = FALSE )$p.value
-
-      #test male strains (Dom)
-     Cent.Dist.p.vals.DF$male.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$male.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$male.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            Cent.Dist.p.vals.DF$male.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value
-
-       #test female strains (Dom)
-     Cent.Dist.p.vals.DF$female.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "G")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$female.WSB.LEW[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "LEW")],paired = FALSE )$p.value
-
-      Cent.Dist.p.vals.DF$female.WSB.KAZ[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "KAZ")],paired = FALSE )$p.value
-
-            Cent.Dist.p.vals.DF$female.WSB.PWD[i] <- t.test(BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "WSB") ],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD")],paired = FALSE )$p.value
-}
-
-#melt the DF's within the BIG list for easier ggploting
-Cent.Melted.pval.DF <- melt(Cent.Dist.p.vals.DF, c("sex.cent.dist",  "female.subsp","male.subsp","male.PWD.MSM",     "male.PWD.KAZ","male.PWD.SKIVE", "female.PWD.MSM", 
-                              "female.PWD.KAZ",   "female.PWD.SKIVE", "male.WSB.G", "male.WSB.LEW", "male.WSB.KAZ",  "male.WSB.PWD","female.WSB.G", "female.WSB.LEW", "female.WSB.KAZ", "female.WSB.PWD" ),
-                         id=c() )
-
-#clean up dm1 (there are a bunch of rows without useful information)
-Cent.Melted.pval.DF <- Cent.Melted.pval.DF[(!is.na(Cent.Melted.pval.DF$value)),]
-colnames(Cent.Melted.pval.DF) <- c("Test.type", "pval")
-
-biggie2 <- ggplot(Cent.Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+ ggtitle("Cent Distance pvalues for 1000 replicated t.tests")+facet_wrap(.~Test.type) 
-#sex difference is skewed
-biggie2
-
-```
-
-Verbalizing the inference from all the p values.
-
-Using the same permuted / subsampled data frame -- I'll look at centromere and telomere distance
-
-How would these distributions change normalized distances?
-
-t.tests for centromere dis and telomere distance
-This result and the IFD could be the same mechanism.
-
-## Higher recombining males have stronger interference
-
-Do the same figure things for IFD
-
-There might be too little 2CO observations for many dom chrm to get the IFDs
-
-`r kable(table(Curated_BivData$category[Curated_BivData$hand.foci.count == 2]))`
-
-```{r interference, echo=FALSE, eval=FALSE}
-#for interference I might have to subsample only from 2COs
-
-#have to seperate the hand foci number
-
-IFD.p.vals.DF <- data.frame(sex.IFD.test = rep(0, length(BIG.DF.LIST) ) )
-for(i in 1:length(BIG.DF.LIST)){  #current data from
-  #print(i)
-
-      #female vs male
-      IFD.p.vals.DF$sex.IFD.test[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "female" ) & (BIG.DF.LIST[[i]]$hand.foci.count == 2) ],
-         BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & (BIG.DF.LIST[[i]]$hand.foci.count == 2)], na.action=na.omit, paired = FALSE )$p.value
-
-  #subsp
-  IFD.p.vals.DF$female.subsp[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
-         BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],paired = FALSE )$p.value
-
-IFD.p.vals.DF$male.subsp[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Dom") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
-         BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$subsp == "Musc")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE, na.action=na.omit)$p.value
-
-#strain comarisons
-IFD.p.vals.DF$male.PWD.MSM[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
-         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "MSM")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value
-
-#IFD.p.vals.DF$male.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
-#         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain == "SKIVE")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value
-
-#IFD.p.vals.DF$male.WSB.G[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( #BIG.DF.LIST[[i]]$strain == "WSB") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
-#         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "male" ) & ( BIG.DF.LIST[[i]]$strain #== "G")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value
-
-#females  (not enough PWD-MSM)
-#IFD.p.vals.DF$female.PWD.SKIVE[i] <- t.test(BIG.DF.LIST[[i]]$IFD1[ ( BIG.DF.LIST[[i]]$sex == #"female" ) & ( BIG.DF.LIST[[i]]$strain == "PWD") & (BIG.DF.LIST[[i]]$hand.foci.count == 2)],
-#         BIG.DF.LIST[[i]]$dis.cent[ ( BIG.DF.LIST[[i]]$sex == "female" ) & ( BIG.DF.LIST[[i]]$strain #== "SKIVE")& (BIG.DF.LIST[[i]]$hand.foci.count == 2)],na.action=na.omit, paired = FALSE )$p.value
-
-}
-
-#melt the DF's within the BIG list for easier ggploting
-IFD.Melted.pval.DF <- melt(IFD.p.vals.DF, c( "sex.IFD.test","female.subsp", "male.subsp", "male.PWD.MSM"), 
-  id=c() )
-
-#  "sex.cent.dist",  "female.subsp","male.subsp","male.PWD.MSM",     "male.PWD.KAZ","male.PWD.SKIVE", "female.PWD.MSM", 
-#                              "female.PWD.KAZ",   "female.PWD.SKIVE", "male.WSB.G", "male.WSB.LEW", "male.WSB.KAZ",  #"male.WSB.PWD","female.WSB.G", "female.WSB.LEW", "female.WSB.KAZ", "female.WSB.PWD" ),
-
-#clean up dm1 (there are a bunch of rows without useful information)
-IFD.Melted.pval.DF <- IFD.Melted.pval.DF[(!is.na(IFD.Melted.pval.DF$value)),]
-colnames(IFD.Melted.pval.DF) <- c("Test.type", "pval")
-
-biggie3 <- ggplot(IFD.Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+ ggtitle("ABS IFD pvalues for 1000 replicated t.tests")+facet_wrap(.~Test.type) 
-#sex difference is skewed
-biggie3
-
-```
-
-The
-
-## more variation across traits for females than males
-
-## High recombining male mice have more total SC?
-
-I think with total SC, this is true, but it doesn't seem true for the single bivalent data. This could also be due to sampling issues.
-
-```{r total.sc, echo=FALSE}
-#FIND TOTAL sc code
-```
-
-```{r block2, echo=FALSE, eval=FALSE}
-#Curated_BivData
-# mmore SC based plots
-
-#Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count <= 3,]
-#Curated_BivData99 <- droplevels(Curated_BivData99)
-#why does this making so many Fing NAs?
-
-#use subset instead!!
-
-Curated_BivData99 <- subset(Curated_BivData, hand.foci.count <= 3)
-
-Curated_BivData <- subset(Curated_BivData, hand.foci.count <= 3)
-
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$sex) ), ]
-
-#this plot is too busy for all the data 
-SC.length.points.bivs_Musc <- ggplot(Curated_BivData[Curated_BivData$subsp == "Musc",], aes(x=dis.cent, y=chromosomeLength, color= as.factor(mouse) ))+
-  geom_jitter()+ xlim(c(0,120)) + ylim(c(40,120))+
-  facet_grid(sex ~ hand.foci.count) + ggtitle("")+theme(legend.position="none")
-
-SC.length.points.bivs_Dom <- ggplot(Curated_BivData[Curated_BivData$subsp == "Dom",], aes(x=dis.cent, y=chromosomeLength, color= as.factor(mouse) ))+
-  geom_jitter()+ xlim(c(0,120)) + ylim(c(40,120))+
-  facet_grid(sex ~ hand.foci.count) + ggtitle("")
-
-#Making centromere rect
-  #geom_vline(aes(xintercept = mean.cent.dis.PWD$mean.dist_cent ) ) +
-  #annotate("rect", 
- # xmin =  ( mean.cent.dis.PWD$mean.dist_cent - (mean.cent.dis.PWD$dis.cent)*2), 
-#  xmax = ( mean.cent.dis.PWD$mean.dist_cent + (mean.cent.dis.PWD$dis.cent)*2),
-#              ymin=0,  ymax=Inf, alpha=0.2, fill="purple" ) +
-#  labs(x="", y='Density')
-
-```
-
-# Interfocal comparisons
-
-```{r IFD, echo=FALSE, eval=FALSE}
-
-IFD.ABS.comp <- ggplot(Curated_BivData, aes(y = IFD1, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("IFD1 raw distributions")+theme(legend.position="none")
-
-IFD.PER.comp <- ggplot(Curated_BivData, aes(y = IFD1_PER, x=hand.foci.count, color=mouse))+geom_jitter(aes(alpha=0.2))+ facet_wrap(strain~sex)+ ggtitle("IFD1 normalized distributions")+theme(legend.position="none")+ylim(c(0,1.05))
-
-IFD.ABS.comp
-IFD.PER.comp
-```
-
-The above plots shows the distribution the raw and normalized distribution of the first Inter Focal Distance.
-
-# Sis Co ten Metric
-
-```{r sis.co.ten, echo=FALSE, eval=FALSE}
-
-sis.co.ten.hist.full <- ggplot(Curated_BivData, aes(SisCoTen)) + geom_histogram(bins = 50) +facet_wrap( ~ category)+ggtitle("Distribution of Sis-co-ten per bivalent")
-
-sis.co.ten.hist <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 2,], aes(SisCoTen)) + geom_histogram(bins = 50) +facet_wrap(hand.foci.count ~ category)+ggtitle("")
-#the male histograms look almost bimodal, and the female data looks more uniform
-#the telomere bias skews the distribution to lower sis-co-ten measures
-#PWD has sig larger
-
-sis.co.ten.hist.full
-sis.co.ten.hist
-
-#ARE THE SIS_CO_TEN distributions different across foci.number for males and females
-#males; yes?
-#females
-
-#sis.co.ten.hist.poin <- ggplot(master_curated_BivData, aes(x=SisCoTen, y=chromosomeLength, color= as.factor(hand.foci.count)  ))+
-#  geom_point(aes(alpha=.5)) +facet_wrap(~ category)
-
-#sis.co.ten.hist.poin99 <- ggplot((master_curated_BivData[1:50,]), aes(x=SisCoTen, y=chromosomeLength, color= #as.factor(hand.foci.count)  ))+
-#  geom_point(aes(alpha=.5)) +facet_wrap(~ mouse)
-```
-
-```{r draft plots, eval=FALSE}
-first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
-```
-
-```{r 2CO.MM, echo=FALSE, eval=FALSE}
-
-#remove MOLF
-mouse.avs_4MM.99 <- mouse.avs_4MM[mouse.avs_4MM$strain != "MOLF",]
-
-SC.Length_2CO_MM <- lme(mean.SC_2CO ~ subsp * sex, 
-                  random=list(strain=pdDiag(~sex),
-                        data= mouse.avs_4MM.99, 
-                                                                                       na.action=na.omit ) )
-#keep getting errors for building this model
-
-as.data.frame( round(  summary(SC.Length_2CO_MM)$tTable, 6 ) )
-
-table(mouse.avs_4MM$strain, mouse.avs_4MM$sex)
-
-```
-
-```{r 3CO.MM, echo=FALSE}
-
-SC.Length_3CO_MM <- lme(mean.SC_3CO ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ),  na.action=na.omit  )
-#this model is also generating error...
-#don't know how to fix
-
-as.data.frame( round(  summary(SC.Length_3CO_MM)$tTable, 6 ) )
-```
-
--->
+# References

### 2020-02-05 doc/Results.rmd
debugging and standarizing the DFs used

@@
 
-```{r biv.data_table.4MM.Male, echo=FALSE}
+```{r biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+
+#add this to setup
+
 #remove female, 
@@
 male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-
 ```
@@
 
-```{r MM_2.SC.Length, echo=FALSE}
+```{r MM_2.SC.Length, echo=FALSE, eval=FALSE}
+
 #get a list of the old mice to double check
@@
 #MIXED MODEL
-lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
 as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
@@
 #run a t.test seperate out by the chrm classes
-MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
 
@@
 
-```{r male.show.plots, echo=FALSE, eval=FALSE}
+```{r male.show.plotsEXXXample, echo=FALSE, eval=FALSE}
 
@@
 ```{r SC.length.male_tests, echo=FALSE}
+
 ## run the tests to check following expectations
@@
 
-male.Dom.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Dom"), ]
+#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
+
+male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
 #the subsp labels get messed up, unless they are designated chr first
-male.Musc.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == 2), ]
-
-mouse.avs_4MM$subsp <- as.character(mouse.avs_4MM$subsp)
-mouse.avs_4MM$subsp <- ifelse(grepl("MSM", mouse.avs_4MM$strain), "Mol", 
-        ifelse(grepl("MOLF", mouse.avs_4MM$strain), "Mol", mouse.avs_4MM$subsp))
-male.Mol.mouse.av <- mouse.avs_4MM[ (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Mol"), ]
-
 Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
 
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  )
+Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
 Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
@@
 #0.000369 ***
-
 Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
 #0.000816 ***
-
 Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
@@
 #make sure to remove some of the juvi mice!! from the AP_mouse table
-
 show.Dom.male.CO <- ggplot(
-AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") & (AP_mouse_table_w.Ages$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = AP_mouse_table.HQ[(AP_mouse_table.HQ$sex == "male") & (AP_mouse_table.HQ$subsp == "Dom"),])
+male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
+
+Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
 #P not sig for CO mean
@@
 
-male.Musc.mouse.av <- male.av_4MM[(male.av_4MM$subsp == "Musc")&(male.av_4MM$sex == "male"),]
+male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
+male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
+male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
 
@@
 ```{r log.reg_SC.length, echo=FALSE}
+#male_mouse.av
+#Male.poly.Mouse.Table_BivData_4MM
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
+                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
+                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
+
 #All chrm
 log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
 summary(log.reg_SC.length_mouse.av)
@@
 log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
 summary(log.reg_SC.length_mouse.av_1CO)
@@
 log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
 
@@
 log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]
-
-                         , family=binomial(link="logit"))
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
 summary(log.reg_SC.length_mouse.av_3CO)
@@
 # bivalent level log regression
+
 log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
 
@@
 
-#1CO
+#1CO  (BIVDATA)
 log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
 
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
 summary(log.reg_SC.length_BIV_1CO)
@@
 log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 2),], family=binomial(link="logit"))
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
 summary(log.reg_SC.length_BIV_2CO)
@@
 log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 3),], family=binomial(link="logit"))
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
 
@@
 
-```{r SC.Length_t.test, echo=FALSE}
-
-T.test_SC.Len_mouse.av <- t.test(mouse.avs_4MM$mean_SC[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean_SC[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")&(mouse.avs_4MM$Rec.group == 0)]
+```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
+
+#comparing all the strains
+
+T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
 )#p-value = 0.001567
@@
 
-The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_SC.Len_mouse.av$p.value` and p=`r T.test_SC.Len_single.biv$p.value`.
+The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
 
@@
 
-I can't get the above mixed models to work for testing high vs low rec groups
-
 ```{r long.sc_log.res_meanSC, eval=FALSE}
+#Long Biv comparison
+
 #hold off on reporting/ working with the long SC biv resutls
@@
 
-```{r IFD.log.reg, echo=FALSE}
-#
+```{r IFD.log.reg, echo=FALSE, eval=FALSE}
+
+# LOGISTIC REGRESSION FOR IFD
+
 log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS,
@@
 
+#t.test for IFD
 #mouse.av - the mouse average IFD is from just 2CO
@@
 
-oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+#oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+#  geom_jitter(alpha=.2)+
+#  ggtitle("Interfocal Distances, ABS")+
+#  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = #"free")+ylim(c(0,1))+theme(legend.position="none")
 
@@
 
-```{r MM_CO_pos, echo=FALSE}
+```{r MM_CO_pos, echo=FALSE, eval=FALSE}
 #use the 1CO
@@
 
-```{r log.reg.CO1_pos, echo=FALSE}
+```{r log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+
+#LOGISTIC REGRESSION for FOCI POSITION
+
 #mouse av, norm.first foci is from 1CO bivalents
@@
 
-```{r t.test_1CO_pos, echo=FALSE}
+```{r t.test_1CO_pos, echo=FALSE, eval=FALSE}
+
+#t.test for high and low group  -- logistic regression still need to format the high and low group
 
@@
 
-```{r sis.co.ten.test_HetC, echo=FALSE}
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+
 #mouse.avs_4MM  - mouse averages with mean sis.co.ten
@@
 
-```{r siscoten.test.poly, echo=FALSE}
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
 
 mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
 
@@
 ```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
 #SISCOTEN
 #mouse av, norm.first foci is from 1CO bivalents
+
 log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
 
-                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
 summary(log.reg_SIS.co.ten_mouse.av)
@@
 
-```{r t.test_sis.co.ten,echo=FALSE}
-#mouse av
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
 t.test_mouse.av_siscoten <- t.test(
 
-  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean.siscoten[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
   )
@@
 t.test_mouse.av_siscoten_1CO <- t.test( 
-  mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
-
-   mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
   )
@@
 t.test_mouse.av_siscoten_2CO <- t.test( 
-  mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
-
-   mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
   )
@@
 t.test_mouse.av_siscoten_3CO <- t.test( 
-  mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],
-
-   mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
   )
@@
 ####
-#Single Biv Level
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
 t.test_single.biv_siscoten <- t.test(
@@
 ```{r MM.telo, echo=FALSE}
-
-lme_MM_telo.dist <- lme(mean.telo.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_telo.dist)$tTable, 6 ) )
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+
+drop1(lme_MM_telo.dist, test="Chisq")
+
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
 
@@
 
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
 #anova(lmer_MM_siscoten)
@@
 #thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
+drop1(lmer_MM_telo.dist, test="Chisq") 
 rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
 rand_telo.dist_MM
+
 ```
@@
 ```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
 #TELO.DIST
@@
 
-                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
 summary(log.reg_Telo.Dist_mouse.av)
@@
 
-  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean.telo.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
   )
@@
 
-```{r MM.cent.dist, echo=FALSE}
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
 as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
 
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
 
@@
 
-                                 data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
 summary(log.reg_CENT.Dist_mouse.av)
@@
 
-  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean.cent.dist[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
   )
@@
 #scatter + boxplot for t.tests
-
 ```

### 2020-02-05 doc/MatMeth.Rmd
working on the chrm class proportions sections, other small edits
start working on MatMethods again
reviewing MatMethods doc adding notes based on new analysis

@@
 To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
-The Mus musculus PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp).
-The related species of *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jax labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
+The Mus musculus strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
+The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
+(https://en.brc.riken.jp) and rederived by the UW-Biotech centre (Madison UW).
+The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
 
- All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.  A breeding colony of wild derived 
-*Mus musculus* mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
- Mice were feed on dry standard breeder chow.
+ All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.
+A breeding colony of wild derived *Mus musculus* mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
+ Mice were fed on dry standard breeder chow.
 Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
@@
 (evidence for environmental (room) effect on (MLH1 / SC lengths / a variety of meiotic traits was tested for.
-We observed no effect, Supplemental figure))
+Effects on average MLH1 counts were compared -- for strains which were kept at mulitple institutions.
+((We observed no effect, Supplemental figure))
 
@@
 
-The vast majority of oocyte data was collected from neonate litters/mice between 5 to 48 hours.
+The vast majority of oocyte data was collected from neonate litters/mice between 5 to 48 hours old.
 This approach maintain breeding pairing and still result in prophase oocytes (cite timeline of oocytes in neonates).
 (differences were assessed between embryo samples and neonates?).
-Precise staging of embryos by checking copulary plugs was difficult/not feasible in many of these wild derived strains due to their behavior.
+
+Precise staging of embryos via copulary plugs was not feasible in many of these wild derived strains due to their behavior.
 Embryonic samples were collected when pregnancy noted in females, embryos were staged based on (X table markers).
@@
 
+SYCP3 was used to mark  visualize the Axial element of the synaptonemal complex.
+Sicne all mouse chromosomes are telocentric/acrocentric, CREST was used to indicate the centromere -- indicate the side of the bivalent with the centromere.
+
+the genome wide recombination rate was charaecterized / estimated with staining of MLH1 antibody, 
 The same staining protocol was applied to spermatocyte and oocyte spreads.
-The staining protocol was based on X with some adaptations, and same as previously described in Peterson et al 2019.
+The staining protocol was based on X with some adaptations, and same as previously described in [@peterson_2019].
 
-Explain list recipes for DMC1 since these haven't been listed before
+Precursurs to COs, DSBs were quantified with the DMC1 staning antibody.
+Due to resource constraints -- only a subset of strains were stained for this marker.
+Only spermatocyte spreads produced enough good cells for DMC1 quantification.
 
@@
 
-For MLH1 pachytene characterization, Cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
-Reprocessing in Photoshop (cite). Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
+For MLH1 pachytene characterization, cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
+Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
 
@@
 
+Foci counts/numbers were manually scored.
+
+Single bivalent features were measured both manual and with an automated image analysis software [@Peterson_2019]
+
+DMC1 -- leptotene and sygotene cells were identified by SC morphology (by degree of synapsis) (cite another paper?).
+foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
+
+total number of DMC1 foci per cell were reported.
+
 # Statisical Analysis
+
+DMC1 -- Ttest
 
@@
 
-We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
+The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R). these mouse level summary statistics were analyzed -- within our model frameworks.
+
+We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio]. (lme)
 
@@
 
+### Single bivalet quant
+
+Single bivalent chromosomes were isolated from cytology cell images as discribed in (X ). the single bivalents were 'curated' (removed errors in the segmentation algorithm)
+
+to account for effects of CO number and physical chromosome size -- the data was also analyzed after spliting up by chromosome class and in long and short chromosome bins.
+
+- Split up 'classes' of bivalents depending on the number of foci -- to tests/ comparisons more comprable.
+
+- The performance of the segmentaion algorithm was assessed by counting the number of objects segmented per single cell image (expected 19 or 20 in males and females respectively).
+All automated bivalent data was manually curated before analysis.
+
+- whole cell bivalent measures (measures of all bivalents ), were compiled by mixed automated bivalent measures with hand measurements. (Since the error in automated measures was so low)
+
+**SC length**
+
+The axial element (AE) of the SC -- for single bivalents (synapsesd homolog pairs), was quantified as the skeletonized (single pixel width line) --
+
+-comparison to human measures previously in permyscus species -- and in musculuc (shown in the supplement)
+
+- The long bivalent and short bivalent data sets were compiled by calculating the 1st and 4th quartille of SC lengths for each cell with at least 16 bivalent measures.
+Bivalents from each cell were assigned into the long and short bins based on these SC Length cut offs.
+
+**CO interference**
+
+- how we choose to measure interference, the average distance between2 foci on same bivalent.
+
+- downside to using the gamma distribution -- not the primary meteric for CO interference.
+
+- Also pooling intervals for categories and fitting a gamma distribution (in supplement)
+
+**rec landscape** **1CO normalized position**
+
+- how why measure the recombination landscape -- coohse to charaecterize the rec landscaoe bby isloating the 1CO bivalents; since there are known sexual dimorphic features of this and since bivalents with more than 1 foci will be primarily influenced by CO interference.
+We look at both the raw and
+
+the distance to the telomere and centromere end of the SC shape were measured -- for all / pooled bivalents
+
 -specific R packages
@@
 -hypothesis tested (size, DSB number)
+
+-Mixed models and glm models were built
+
+-tests for fixed effects
+-tests of random effects in mixed models was LTR of exactRLRT function in X package.
+
+single bivalent analysis
+
+# Numbers of mice cells ect for each method?
+
+above belongs more in results
 

### 2020-02-05 doc/Results.rmd
updated a chrm proportion figure

@@
 
-## MLH1 variance M1
+### MLH1 variance M1
 
@@
 
-```{r M1.MLH1.var_cv, echo=FALSE}
+```{r M1.MLH1.var_cv, echo=FALSE, include=FALSE}
 #modeling within mouse variance
@@
 MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-summary(MLH1var_M1)
+#summary(MLH1var_M1)
 
@@
 MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-summary(MLH1cv_M1)
+#summary(MLH1cv_M1)
 
@@
 
-## MLH1 variance M2
+### MLH1 variance M2
 
 $$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-M2 slash M3
-
-```{r M2.MLH1.var, echo=FALSE}
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE}
 
@@
 
-summary(MLH1var_M3)
+#summary(MLH1var_M3)
 anova(MLH1var_M3)  
@@
 
-summary(MLH1cv_M3)
+#summary(MLH1cv_M3)
 
@@
 
-summary(MLH1var_M3_Q12)
+#summary(MLH1var_M3_Q12)
 anova(MLH1var_M3_Q12)  
@@
 
-summary(MLH1cv_M3_Q12)
+#summary(MLH1cv_M3_Q12)
 anova(MLH1cv_M3_Q12)#sex large F
@@
 
-## MLH1 variance M3
+### MLH1 variance M3
 
@@
 
-```{r M3.MLH1.var, echo=FALSE}
+```{r M3.MLH1.var, echo=FALSE, include=FALSE}
 
@@
 
-summary(MLH1var_M4)
+#summary(MLH1var_M4)
 
@@
 
-summary(MLH1cv_M4)
+#summary(MLH1cv_M4)
+
 anova(MLH1cv_M4)  
@@
 
-summary(MLH1var_M4_Q12)
+#summary(MLH1var_M4_Q12)
 
@@
 
-summary(MLH1cv_M4_Q12)
+#summary(MLH1cv_M4_Q12)
+
 anova(MLH1cv_M4_Q12)
@@
 
-#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
-#can't re-order the miceagg
-#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
-#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
-#        "grey","grey","grey",  "grey","grey","grey" )) +
-# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
-#don't show M.musculus by mouse
-#scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
- #       "green","green4","grey",  "grey","grey","grey" )) +
-
-#order MLH1_data by strain!!
-#colors_of_strains, colors .. 
-#there is no order to 'mouse' so it looks shitty
-
 justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
@@
 
-AP_strain.table.HQ <- ddply(MLH1_data, c("strain", "sex"), summarise,
+#this move this to the setup file
+
+strain.table.HQ_Fig <- ddply(MLH1_data, c("strain", "sex"), summarise,
                         Nmice =length(unique(mouse)),
@@
 
-Fig1_v1_strain <- ggplot(AP_strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+#might need to nudge the X values
+
+Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
   ylab("Average MLH1 Foci per Cell")+
@@
   geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+
     geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
@@
  scale_color_manual(values=colors_of_strains)+
+
     theme(legend.position="none",
@@
 #remove y labfor male
+
    #segments
@@
 
-> is there a strain effect in Dom?
+> Is there a strain effect in Dom?
 
@@
 
-```{r male.show.plots, echo=FALSE}
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+#try customize the plot size
 
@@
 
+(Try removing the alpha legend)
+
 This is the table and plot of new Musc strains
@@
 booo <- summary(bhh11)
-booo$coefficients
 
@@
 ll <- summary(lm.male.strain)
-ll$coefficients
+
 #even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
@@
 
+#no filename for add mouse, strain subsp
 MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
@@
 
+#can't apply function use this
+chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
+             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
+                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
+                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
+
+          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
+              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
+
+                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
+                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
+            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
+                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
+                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
+                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
+                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
+                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
+
+             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
+               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
+
+chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",
+                                                                 "Cast",
+                                                                 "Musc",
+                                                                 "Mol",
+                                                                 "Spretus","Spic",
+                                                                 "other") )
 #ggplot
 
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
+chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
 chrm.class.barplot
@@
 
-#### HetC_M1
+#### SC_HetC_M1
 
@@
 
-#### HetC_M2
+#### SC_HetC_M2
 
@@
 
-#### HetC_M3
+#### SC_HetC_M3
 
@@
 
-$$mouse \ average \ SC  \ length ~=~ rand(strain) + \varepsilon $$
-
-$$mouse \ average \ IFD \ trait  ~=~ rand(strain) + \varepsilon $$
+M1 glm model for fixed strain effect across male averages.
+
+$$mouse \ average \ SC  \ length ~=~ strain + \varepsilon $$
+M2 anova for glm, of each subset, does at least one of the strains differ from others?
+
+M3 logistic regression, can a bivalent metric predict the high or low group?
+
+$$ log( strain ~=~ \ mouse \ average \ SC  \ length ) $$
+
 The IFD traits are:
@@
 
-# IFD
+# Q1. IFD
 
@@
 
+### IFD_HetC_M1
+
 ```{r HetC_M1_IFD_many.metrics, echo=FALSE}
@@
 as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
 ```
@@
 Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
+
+### IFD_HetC_M2
+
+```{r IFD_HetC_M2, echo=FALSE}
+
+```
+
+### IFD_HetC_M3
+
+```{r IFD_HetC_M3, echo=FALSE}
+
+```
 
@@
 
-## Male Polymorphism Prediction
+# Q1. General CO Positions
+
+There are a few traits that fall within the CO positions
+
+- 1CO normalized position
+
+- sis.co.ten (sorta interference)
+
+- centromere and telomere distance
+
+#### Predictions
+
+**Heterochiasmy**
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+**Polymorphism**
+
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
+
+C.3) telomere and centromere distances will be shorter in low rec strains
+
+# Q1. Normalized 1CO positions
+
+**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
+
+```{r CO.positions.figs, echo=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
+This plot shows the sexual dimorphism in the density plots.
+
+Consider adding annotate_text for the number of observations in each category.
+think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
+
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+#real.long.bivData
+#MELT_Curated_BivData
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+```
+
+<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
+
+```{r CO1_foci_pos_boxplot, echo=FALSE}
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+##### Mixed models norm 1CO positions
+
+the distribution of SC lengths and sis-coten seems very different across sexes
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+The mixed model data should only come from 1CO bivalent data.
+
+```{r MM_CO_pos, echo=FALSE, eval=FALSE}
+#use the 1CO
+#Use MELT DF
+#MELT_Curated_BivData
+
+#isolate the 1COPos.CO.avs
+#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
+#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
+#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
+
+mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
+
+#<- lme()
+lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
+
+#can't make the 
+CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+#library(lmerTest)
+#anova(CO_position.MM)
+#anova(CO_position.MM, verbose = TRUE)
+
+#summary(CO_position.MM) #no p values,,
+#these just give t and F stats, no pvalue
+
+#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
+#CO_pos_fix.est
+#this test, the full model is more predictive than an empty model?
+
+#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
+#thing2 <- drop1(SC_length.MM, test="Chisq")
+
+CO_pos_rand <- exactRLRT(CO_position.MM)
+CO_pos_rand
+```
+
+```{r log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+
+#LOGISTIC REGRESSION for FOCI POSITION
+
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+```{r t.test_1CO_pos, echo=FALSE, eval=FALSE}
+
+#t.test for high and low group  -- logistic regression still need to format the high and low group
+
+#mouse av
+t.test_mouse.av.1CO_pos <- t.test(
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv.1CO_pos_t.test <- t.test(
+
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+  )
+
+```
+
+the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
+#### Nrm_1COpos_HetC_M1
+
+#### Nrm_1COpos_HetC_M2
+
+#### Nrm_1COpos_HetC_M3
+
+### Siscoten
+
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males.
+
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
+t.test_mouse.av_siscoten <- t.test(
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+#2CO
+
+#3CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+
+drop1(lme_MM_telo.dist, test="Chisq")
+
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+drop1(lmer_MM_telo.dist, test="Chisq") 
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
+
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
+
+> is random strtain effect significant?
+
+The random strain effect seems very significant.
+
+remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
+
+```{r pos_t.test}
+#make point plots / boxplots which show differences in mean positions
+
+#scatter + boxplot for t.tests
+```
+
+Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
+
+```{r F1.by.F2}
+
+#try remaking the plot Megan suggested
+# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
+
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
+#isolate 2COs
+#facet by sex and subsp
+
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
+
+```
+
+```{r CO.pos.ANOVA, eval=FALSE}
+#what is the pattern of variance
+#run analyses for each subsp*sex
+#use non-melt DF
+
+#how is the variance partioned across
+#cell, mouse, strain
+
+female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
+female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
+
+female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
+
+#unorder strain and mouse
+
+female.Dom$mouse <- as.factor(female.Dom$mouse)
+
+female.Dom$strain <- unclass(female.Dom$strain)
+female.Dom$strain <- as.factor(female.Dom$strain)
+
+female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
+female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
+
+#1CO first
+modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
+
+#can't get mouse and strain to have sum of square
+#residual size decreases with per.F1
+#residuals much larger than fileName, mouse and strain no
+
+#model <- lm(breaks ~ wool * tension, 
+#            data = warpbreaks, 
+#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
+
+male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
+male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
+
+male.Dom$mouse <- as.factor(male.Dom$mouse)
+
+male.Dom$strain <- unclass(male.Dom$strain)
+male.Dom$strain <- as.factor(male.Dom$strain)
+
+male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
+male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
+
+male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
+
+male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
+summary(aov(male.modo))
+
+#only file name is registering as effect
+#Review ANOVA frameworks
+#http://www.biostathandbook.com/nestedanova.html
+
+```
+
+# Q2 SC
+
+I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
+
+```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
+
+#isolate all musc
+Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+#make table
+Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
+
+                         )
+
+        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+```
+
+## SC_male.poly_glm
+
+remeber the 3 predictions for subsp;
+
+i) Dom, no difference
+
+ii) Musc, PWD and SKIVE greater than others
+
+iii) Mol, MSM > MOLF
+
+```{r Q2.SC.ploy.mal, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+
+#1. determine which strain effects are sig
+#use mouse biv average
+#Male.poly.Mouse.Table_BivData_4MM
+
+glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm1.SC.male.poly) 
+#SC many strains are different (more than the gwRR)
+#subsps are sig
+
+glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm2.SC.male.poly)
+
+#Dom
+glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+
+#summary(glm_SC.male_Dom)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+
+#summary(glm_SC.male_Musc)
+
+#Mol
+glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+
+#summary(glm_SC.male_Mol)
+
+#Logistic regressions
+
+log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+
+                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
+
+#summary(log.SC.male.poly)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_musc)
+#not sig -- when just looking at Musc
+
+#significant in predicting group
+log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_mol)
+#not sig -- when just looking at Mol
+```
+
+#### Q2 SC by class
+
+I think these comparisons should be done with the **Bivalent observations**
+
+The logistic regression showed rec groups could be predicted by SC length - re-run them while seperating out by chrm class
+
+```{r Q2.SC.Chrm.Class, echo=FALSE}
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+```
+
+These are the plots of the logistic regression.
+
+```{r Q2.SC.log.regression.show.plots, echo=FALSE}
+
+plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.blah.mouse.av
+
+plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.logReg.mouse.av_1CO
+
+plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.logReg.mouse.av_2CO
+
+```
+
+the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
+
+> **Summary**
+
+When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
+I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
+(IE the scatter plots of SC length, binned by chrm class on the X).
+
+Each point is a mouse average SC.
+When all bivalents are pooled, the logistic regression SC lengths predict the groups.
+When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
+For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
+
+ and the mean 2COs SC are higher in the high strains
+(this must have something to do with the physical size effect)
+
+# Q2 IFD
+
+this is the Q2 prediction for IFD / interference
+
+> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
+
+1. glm across subsp
+
+2. anova within subsp
+
+3. logistic regression for musc (or all males)
+
+```{r Q2.IFD, echo=FALSE}
+
+#glm
+
+#anova for glm
+
+#logistic regression for high and low group
+
+```
+
+> Male Polymorphism Prediction
 
@@
 
-> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
-
-logistic regression
+test 3. logistic regression
 $$  Rec \ Group ~=~ mouse \ average \ IFD $$
@@
 
-```{r male.IFD_strain, echo=FALSE}
+```{r male.IFD_strain, echo=FALSE, eval=FALSE}
+#integrate this into the setup scripts
 
@@
                          )
-
-        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
 
@@
 
-# General CO Positions
-
-There are a few traits that fall within the CO positions
-
-- 1CO normalized position
-
-- sis.co.ten (sorta interference)
-
-- centromere and telomere distance
-
-#### Predictions
-
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-
-  C) Centromere and telomere distances will be sexually dimorphic.
-
-**Polymorphism**
-
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
-
-C.3) telomere and centromere distances will be shorter in low rec strains
-
-# Normalized 1CO positions
-
-```{r CO.positions.figs, echo=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-
-```
-
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-#real.long.bivData
-#MELT_Curated_BivData
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-```{r CO1_foci_pos_boxplot, echo=FALSE}
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-##### Mixed models norm 1CO positions
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-The mixed model data should only come from 1CO bivalent data.
-
-```{r MM_CO_pos, echo=FALSE, eval=FALSE}
-#use the 1CO
-#Use MELT DF
-#MELT_Curated_BivData
-
-#isolate the 1COPos.CO.avs
-#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
-#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
-#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
-
-mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
-
-#<- lme()
-lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
-
-#can't make the 
-CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-#library(lmerTest)
-#anova(CO_position.MM)
-#anova(CO_position.MM, verbose = TRUE)
-
-#summary(CO_position.MM) #no p values,,
-#these just give t and F stats, no pvalue
-
-#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
-#CO_pos_fix.est
-#this test, the full model is more predictive than an empty model?
-
-#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
-#thing2 <- drop1(SC_length.MM, test="Chisq")
-
-CO_pos_rand <- exactRLRT(CO_position.MM)
-CO_pos_rand
-```
-
-```{r log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-```{r t.test_1CO_pos, echo=FALSE, eval=FALSE}
-
-#t.test for high and low group  -- logistic regression still need to format the high and low group
-
-#mouse av
-t.test_mouse.av.1CO_pos <- t.test(
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv.1CO_pos_t.test <- t.test(
-
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-  )
-
-```
-
-the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males (which might)
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
-drop1(lme_MM_telo.dist, test="Chisq")
-
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
-### Heterochiasmy Prediction
-
-> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
-
-> is random strtain effect significant?
-
-The random strain effect seems very significant.
-
-remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
-
-```{r pos_t.test}
-#make point plots / boxplots which show differences in mean positions
-
-#scatter + boxplot for t.tests
-```
-
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
-
-```{r F1.by.F2}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-```{r CO.pos.ANOVA, eval=FALSE}
-#what is the pattern of variance
-#run analyses for each subsp*sex
-#use non-melt DF
-
-#how is the variance partioned across
-#cell, mouse, strain
-
-female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
-female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
-
-female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
-
-#unorder strain and mouse
-
-female.Dom$mouse <- as.factor(female.Dom$mouse)
-
-female.Dom$strain <- unclass(female.Dom$strain)
-female.Dom$strain <- as.factor(female.Dom$strain)
-
-female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
-female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
-
-#1CO first
-modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
-
-#can't get mouse and strain to have sum of square
-#residual size decreases with per.F1
-#residuals much larger than fileName, mouse and strain no
-
-#model <- lm(breaks ~ wool * tension, 
-#            data = warpbreaks, 
-#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
-
-male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
-male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
-
-male.Dom$mouse <- as.factor(male.Dom$mouse)
-
-male.Dom$strain <- unclass(male.Dom$strain)
-male.Dom$strain <- as.factor(male.Dom$strain)
-
-male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
-male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
-
-male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
-
-male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
-summary(aov(male.modo))
-
-#only file name is registering as effect
-#Review ANOVA frameworks
-#http://www.biostathandbook.com/nestedanova.html
-
-```
-
-### Male Polymorphism Prediction
-
-I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
-
-```{r male.Nrm1CO_strain, echo=FALSE}
-
-#isolate all musc
-Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-#make table
-Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
-
-                         )
-
-        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-```
+## Normalized CO positions
+
+C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
+
+The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
+When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
+
+C.2) sis-co-ten metric ... (what about the clustering?)
+
+C.3) telomere and centromere distances ...
 

### 2020-02-07 doc/Results.rmd
female specific analysis patterns

@@
 
-## Male Specific Polymorphism
+# Female Specific Polymorphism
+
+I should run the same questions for the female data --
+
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+table( droplevels(female_mouse.avs_for.analysis$strain) )
+#there are cast strains too
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#Mol
+Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.female.plot
+
+#Dom version 
+Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.female.plot
+#Gough female is higher
+
+#variance
+musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.female.plot
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.female.plot
+
+```
+
+```{r Female.specific.test, echo=FALSE}
+
+lm.female.strain.subsp <- lm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.effects <- summary(lm.female.strain.subsp)
+#strainG, 1.06e-05 ***
+#strainLEW  0.0273 *
+
+lm.female.strain <- lm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+lm.female.strain_sum <- summary(lm.female.strain)
+#G  1.06e-05 ***
+#LEW   0.0273 *
+#PWD  0.0810 .
+#MSM  3.02e-05 ***
+
+#how much greater is G female
+G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
+mean.G.female <- mean(G_female$mean_co)
+
+sans_G_female <- female_mouse.avs_for.analysis %>% filter(strain != "G")
+mean.all.other <- mean(sans_G_female$mean_co)
+
+## within mouse variance models 
+lmVAR.female.strain.subsp <- lm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- lm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- lm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+#LEW not sig
+
+lm.m2.Q12_female <- lm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
+#LEW sig
+```
+
+# Male Specific Polymorphism
 
 Follow up on this finding -- included more strains which were excluded from the HetC analysis.
-
--- what's the driving question for this section --
 
@@
 
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-table( droplevels(fomo.male_mouse.av$strain) )
+just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#table( droplevels(just.male_mouse.av$strain) )
 #there are cast strains too
 
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
 
@@
 #make the Mol and
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
 
@@
 #Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
 
@@
 
-bhh11 <- lm(mean_co ~ subsp * strain, data= fomo.male_mouse.av)
-#bhh11$coefficients
-
-booo <- summary(bhh11)
-
-lm.male.strain <- lm(mean_co ~ strain, data= fomo.male_mouse.av)
-ll <- summary(lm.male.strain)
+lm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.effects <- summary(lm.male.strain.subsp)
+
+lm.male.strain <- lm(mean_co ~ strain, data= just.male_mouse.av)
+lm.male.strain_sum <- summary(lm.male.strain)
 
 #even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+
+#how much greater are the High Rec gwRRs
+just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
+                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
+                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
+
+sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+#mean(sans_LewRec_male$mean_co)
+
+PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+#mean(SKIVE_male$mean_co) /  
+  mean(sans_LewRec_male$mean_co)
+
+## within mouse variance models 
+lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+#all p values kinda high, MSM and TOM are slightly sig
+
+lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+#non sig
+
+lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+#NS
+
+lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+#NS
+
 ```
@@
 # Proportion of chrm classes
+
+TODO -- try running a chi square test on the proportions.
 
@@
 
+As alex said, be careful about mixed models and interperting the random effect (radnom variance) -- assigning strain as random effect is slightly problamatic as i can give it a biological interpertation (it's due to strain -- genetics of each of the strains).
+
 > Q1. same models for MLH1 based sex differences (and evolution)
@@
 3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
+
+## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
+
+The general predictions across the males and subspecies based on th above MLH1 results.
+
+For positive correlation traits/metrics
+
+i) in DOM strains, low to no difference across strains
+
+ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
+
+ii) in Mol, MSM > MOLF
+
+A) **SC lengths** will be longer for high Rec strains.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
+
+C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
+
+The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
+When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
+
+C.2) sis-co-ten metric ... (what about the clustering?)
+
+C.3) telomere and centromere distances ...
+
+M1 glm model for fixed strain effect across male averages.
 
@@
 
-#### SC_HetC_M1
+#### Q1.SC_HetC_M1
 
@@
 
-# Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
-
-The general predictions across the males and subspecies based on th above MLH1 results.
-
-For positive correlation traits/metrics
-
-i) in DOM strains, low to no difference across strains
-
-ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
-
-ii) in Mol, MSM > MOLF
-
-A) **SC lengths** will be longer for high Rec strains.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
-
-The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
-When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
-
-C.2) sis-co-ten metric ... (what about the clustering?)
-
-C.3) telomere and centromere distances ...
-
-M1 glm model for fixed strain effect across male averages.
-
 $$mouse \ average \ SC  \ length ~=~ strain + \varepsilon $$
@@
 #these 2 below are LR tests
-drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-anova(lmer.IFD.PER_MM.M1)
-
-exactRLRT(lmer.IFD.PER_MM.M1)
-as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
+#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+#anova(lmer.IFD.PER_MM.M1)
+
+#exactRLRT(lmer.IFD.PER_MM.M1)
+#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
 
@@
 #these 2 below are LR tests
-drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
-anova(lmer.IFD.RAW_MM.M1) # also F values
-
-exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
-as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
+#anova(lmer.IFD.RAW_MM.M1) # also F values
+
+#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
+#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
 ```
@@
 
+```{r Comp_Ratios, echo=FALSE}
+
+#calq ratios of belo 30
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
+
+#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
+
+DF.strain.list <- list(
+  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
+  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
+  MSM.Bivalents.DF, MOLF.Bivalents.DF
+  )
+
+entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
+        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
+          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
+
+i = 1
+for(x in (DF.strain.list) ){
+
+  #print(head())
+
+  #these filters won't apply to the indexed DF
+  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
+  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
+  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
+  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
+  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+  i = i+1
+
+  }
+entr_DF
+
+entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
+
+```
+
+Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
+
 ```{r WSB.IFD, echo=FALSE, warning=FALSE}
@@
 #table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
+WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
 
@@
 LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+
+LEW.Bivalents.DF_IFD <- LEW.Bivalents.DF[(!is.na(LEW.Bivalents.DF$sex)) & (LEW.Bivalents.DF$hand.foci.count == 2),]
+
+LEW.Bivalents.DF_IFD$sex <- droplevels(LEW.Bivalents.DF_IFD$sex)
+
 #table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
@@
 
-lew.ploty_IFD_raw <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_raw <- ggplot( LEW.Bivalents.DF_IFD,  aes(y=IFD1_ABS, x = sex, na.rm = TRUE))+                     geom_jitter( na.rm = TRUE)+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex, na.rm = TRUE))
 lew.ploty_IFD_raw
+#arg, little annoyed that the x NA won't go away
 
@@
 
-LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  
+  scale_color_manual(values=c('#56B4E9','#E69F00'))+
+  geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
 
@@
                   total.biv = length(Obj.ID),
-
                   mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                   mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
                   n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
 )
-#can I plot these stats
 
@@
 G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>% filter(IFD1_PER <= .3) #this makes DF
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male")
-ratio.male <- 18/1031 #2%
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>% filter(IFD1_PER <= .3)
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")
-ratio.female <- 36/766  #5%
 
@@
 
-PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
+PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
+
+PWD.ploty_IFD_raw
 
@@
                           geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
-
-PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+PWD.ploty_IFD_PER
+
+PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
 
 PWD.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-PWD.table <-  ddply(PWD.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
 
@@
 #make the dataset
-KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+#KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
 
@@
 
-KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
 
 KAZ.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-KAZ.table <-  ddply(KAZ.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
 
@@
 
-```{r MOLF.IFD, eval=FALSE, echo=FALSE}
+```{r MOLF.IFD, echo=FALSE}
 #make the dataset
@@
 
+**What is the overlap for cell images in the MLH1 comparisons and these bivalent comparisons?**
+
+should all mice be represented.
+
 ```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
@@
 
-        #           mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-         #         n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-      #             mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-      #              n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-      #             mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-      #             n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-      #            mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-      #            n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
 ```
@@
 
-```{r Q2.SC.ploy.mal, echo=FALSE}
+```{r Q2.SC.ploy.male, echo=FALSE}
 # 1 way anova, is at least 1 strain different from the rest?
@@
 ```
+
+The glm for SC lengths -- sub-grouped by subsp -- indicates that all strain effects are significant for the average mouseisn't met.  SC length.
+My prediction for Dom, isn't met.
+
+(indecation that there is more variation for the SC lengths than for gwRR / CO counts)
 

### 2020-02-07 src/SetupData.Rmd
can't remeber changes

@@
 #use the above table to pull the male mouse levels
-
 male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")
 
+#add the Rec.group markers
+
+#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+#                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+#                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
 #formate the table for pluggin into models
-
 ```

### 2020-02-07 doc/Results.rmd
added some results for Nrm Foci Pos

@@
 ```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
-
 #isolate all musc
@@
 Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
-
                         n_obs = length(Obj.ID),
@@
                         mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
-
-                         )
+                         )
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
 
@@
 glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-
 #summary(glm_SC.male_Dom)  #ALL strain are sig
@@
 glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-
 #summary(glm_SC.male_Musc)
@@
 glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-
 #summary(glm_SC.male_Mol)
@@
 #not sig -- when just looking at Mol
+```
+
+```{r Q2.SC.long.biv, echo=FALSE}
+
+#mouse.av.long.biv.table
+
+glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
+summary(glm1.SC.male_long.biv)
+
+glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
+summary(glm2.SC.male_long.biv)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
+summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
+summary(glm_SC.male_Musc_long.biv)
+
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
+summary(glm_SC.male_Mol_long.biv)
+
+## logistic regression for Rec group
+
+log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
+              data=mouse.av.long.biv.table,  family=binomial(link="logit"))
+
+summary(log.SC.male_long.SC)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
 ```
@@
 
+#there's no rec group for this table
 log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
@@
 
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
+#there might not be enough data 
+#
+ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
 ```
@@
 (this must have something to do with the physical size effect)
+
+# Q2 Normalized CO positions
+
+<!-- 
+C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
+
+The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
+When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
+
+C.2) sis-co-ten metric ... (what about the clustering?)
+
+C.3) telomere and centromere distances ... 
+-->
+
+Since there aren't many good predictions for how the 1CO normalized landscape relates to gwRR variation, these will be the main questions
+
+- Has there been evolution in the 1CO pattern?
+
+- Is there a consistant pattern with the high rec strains?
+
+Normalized 1CO foci positions will be used. (F1_PER). Check fixed effects of subsp and strains.
+
+sis-co-ten, telomere and centromere distance of foci are metrics which draw from a wider pool of samples. (So maybe I will use them... but avoiding focus on them)
+
+```{r Q2.show.nrmF1, echo=FALSE}
+#Curated_BivData_male
+
+#boxplots of nrm F1
+male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
+
+Q2.show.nrmF1
+
+nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= norm.first.foci.pos, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
+
+nrmF1_mouse_level
+
+```
+
+Remade the Position table, -- and MSM has a more telomeric distribution. MOLF has the most central distribution.
+
+```{r Q2.F1.pos.long.biv, echo=FALSE}
+
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
+
+Q2.show.nrmF1_long.biv
+
+#2 version of the boxplot data -- not sure why th pattern differd
+
+#more bivData
+#nrmF1_with.SC.length <- ggplot(Curated_BivData_male, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
+
+```
+
+I think the plot shows that there are fewer 1CO for the High Rec group.
+
+Maybe showing showin
+
+```{r Q2.1CO.landscape, echo=FALSE}
+#Male.poly.Mouse.Table_BivData_4MM
+#Curated_BivData_male
+
+#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm1.SC.male.poly) 
+#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm2.SC.male.poly)
+
+#M1
+glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1_Q2.1COnrm)
+
+#M2
+glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+
+summary(glm.M2_Q2.1COnrm)
+
+#Dom
+glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(glmQ2_M2.1COnrm_Dom)
+
+#Musc
+glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(glmQ2_M2.1COnrm_Musc)
+
+#Mol
+glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(glmQ2_M2.1COnrm_Mol)
+
+#m3 (log)
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm)
+
+#Musc
+Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
+              family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm_Musc)
+
+#Mol
+Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
+              family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm_Mol)
+
+#biv level logistic regression
+#Curated_BivData_male
+#male.bivdata.1CO
+
+male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
+
+Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO,  family=binomial(link="logit"))
+summary(Log.glmQ2_M2.1COnrm)
+
+#Musc
+Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
+              family=binomial(link="logit"))
+summary(Log.Q2.F1_PER_Musc)
+
+#Mol
+Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
+              family=binomial(link="logit"))
+
+summary(Log.glmQ2_M2.1COnrm_Mol)
+
+```
+
+The lm models
+
+#### Summary
+
+Not really a clear summary for the normalized 1CO position and rec groups.
 
@@
 > Prediction B, Is there evidence for the alternative IFD meansure
-
-## Normalized CO positions
-
-C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
-
-The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
-When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
-
-C.2) sis-co-ten metric ... (what about the clustering?)
-
-C.3) telomere and centromere distances ...
 

### 2020-02-09 doc/Results.rmd
finalizing drafts of Q2 section. debugging the knitr

@@
 )
-
 ```
@@
 #put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
+MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
 
@@
 ```{r MOLF.IFD, echo=FALSE}
+
+#SHUT DOWN THIS BLOCK UNTIL MOLF FEMALE DATA
+
 #make the dataset
-MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+#MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+
+MOLF.Bivalents.DF <- Curated_BivData %>%  filter(strain == "MOLF")
+
 table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
@@
 #plot the differences
-MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MOLF.ploty
+#MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
 #remove the NAs
-molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
-molf.ploty_IFD_raw
-
-molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
-molf.ploty_IFD_PER
+#molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
+
+#molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+  geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
 
@@
 
-```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
+```{r male.Nrm1CO_strain, echo=FALSE}
 #isolate all musc
@@
 
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
 glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
@@
 
+<!--  main DFs for this section
+
+Male.poly.Mouse.Table_BivData_4MM
+
+Bivalent_1o2  (for logistic regression, 1CO and 2COs)
+male.long.biv
+
+-->
 I think these comparisons should be done with the **Bivalent observations**
-
 The logistic regression showed rec groups could be predicted by SC length - re-run them while seperating out by chrm class
@@
 
+```{r Q2.SClength.t.test, echo=FALSE}
+#MOUSE AVE, all
+Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
+
+#mouse.av 1CO
+Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
+
+#MOUSE.av 2CO
+Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
+
+#single biv level
+#male.biv?
+
+#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value
+
+```
+
+The top t.tests indicate that when all the mice averages are pooled, there's a sig difference in SC lengths.
+But When the means are compared across chrm class, the mouse averages are no longer significant.
+
+### Q2 SC length Chrm class prediction
+
+How well does SC length predict chromosome class?
+Prediction, High rec strains will have more sig pvalue, given the lower overlap in SC lengths across chrm class.
+(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
 > **Summary**
@@
 <!-- 
-C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
-
-The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
-When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
-
-C.2) sis-co-ten metric ... (what about the clustering?)
-
-C.3) telomere and centromere distances ... 
+Main DFs for this section:
+
+Male.poly.Mouse.Table_BivData_4MM
+male.bivdata.1CO
+
 -->
@@
 nrmF1_mouse_level
-
-```
-
-Remade the Position table, -- and MSM has a more telomeric distribution. MOLF has the most central distribution.
+```
+
+Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.
 
@@
 #nrmF1_with.SC.length <- ggplot(Curated_BivData_male, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
-
 ```
@@
 I think the plot shows that there are fewer 1CO for the High Rec group.
-
-Maybe showing showin
 
@@
 glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1_Q2.1COnrm)
+#summary(glm.M1_Q2.1COnrm)
 
@@
 glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-
-summary(glm.M2_Q2.1COnrm)
+#summary(glm.M2_Q2.1COnrm)
 
@@
               data=male.bivdata.1CO,  family=binomial(link="logit"))
-summary(Log.glmQ2_M2.1COnrm)
+#summary(Log.glmQ2_M2.1COnrm)
 
@@
               family=binomial(link="logit"))
-summary(Log.Q2.F1_PER_Musc)
+#summary(Log.Q2.F1_PER_Musc)
 
@@
 
-summary(Log.glmQ2_M2.1COnrm_Mol)
-
-```
-
-The lm models
+#summary(Log.glmQ2_M2.1COnrm_Mol)
+
+```
+
+```{r Q2.t.test_COlandscape,echo=FALSE}
+
+#MOUSE AVE
+Rec.group.mouse.F1pos_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.F1pos_Ttest$p.value
+#p-value = 0.9098
+
+#single biv level
+Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+Rec.group_BivData_F1post_Ttest$p.value 
+#p-value = 0.238
+
+```
+
+The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
 
@@
 
-```{r Q2.IFD, echo=FALSE}
-
-#glm
-
-#anova for glm
-
-#logistic regression for high and low group
-
-```
-
-> Male Polymorphism Prediction
-
-The polymorphism frame-work
-
-test 3. logistic regression
-$$  Rec \ Group ~=~ mouse \ average \ IFD $$
-For these tests I am looking at the single IFD measure from 2CO bivalents.
-
-For the logistic regression, comparing High and Low groups (there needs to be 2 groups)
-I think I should compare 2 sets of logistic models, where SKIVE (intermediate gwRR), is in high and low.
-
-```{r male.IFD_strain, echo=FALSE, eval=FALSE}
-#integrate this into the setup scripts
-
-#make table
-Musc_male_IFD_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 2,], c("strain"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_IFD = mean(IFD1_ABS, na.rm=TRUE),
-                        mean_IFD.nrm = mean(IFD1_PER,na.rm=TRUE)
-                         )
-
-```
-
-```{r IFD.log.reg, echo=FALSE, eval=FALSE}
-
-# LOGISTIC REGRESSION FOR IFD
-
-log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *
-
-log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-summary(log.reg_IFD_PER_mouse.av)#NS
-
-### Single Bivalent levels
-
-log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
-              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_IFD_ABS_single.biv)
-
-log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
-              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_IFD_PER_single.biv)
-```
-
-```{r t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
+```{r Q2.show.IFD, echo=FALSE}
+
+#subset
+male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
+
+biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
+
+biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+biv.IFDs_PER_by.group
+```
+
+The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
+ The normalized plot displays the pattern of higher rec group having slightly higher longer IFDs.
+
+```{r Q2.IFD.glm.test, echo=FALSE}
+#glm --
+#mouse aver
+
+#M1.ABS
+glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1.Q2_IFD.raw) 
+#M1.PER
+glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1.Q2_IFD.PER)
+
+#M2.raw
+glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M2_Q2.1COnrm)
+
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+summary(Q2.glm.M2_IFD_Mol)
+
+#M2.PER
+glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M2_Q2.IFD.PER)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+summary(Q2.glm.M2_IFD.PER_Mol)
+```
+
+```{r Q2IFD_logreg, echo=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.PER, echo=FALSE}
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
 
@@
 
-```{r t.test_IFD,echo=FALSE}
-
-#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-#                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+```{r Q2.t.test_IFD,echo=FALSE}
+
+#already made
+#subset
+male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+#t.tests for the Rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
 
@@
 #single biv level
-
-Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 ) & (Curated_BivData$hand.foci.count == 2 )],
-
-     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 ) & (Curated_BivData$hand.foci.count == 2 )]
-)
-Rec.group_IFD.ABS_test$p.value
-
-## PER
-
-Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )& (Curated_BivData$hand.foci.count == 2 )],
-
-     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 ) &  (Curated_BivData$hand.foci.count == 2 )]
-)#p-value = 0.04231
-
-Rec.group_IFD.PER_test$p.value
-
-#Single Biv Leve
-single.biv.ABS.t.test <- t.test( Curated_BivData$IFD1_ABS[Curated_BivData$Rec.group == 1],
-
-         Curated_BivData$IFD1_ABS[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
-
-single.biv.IFD_PER.t.test <- t.test( Curated_BivData$IFD1_PER[Curated_BivData$Rec.group == 1],
-
-         Curated_BivData$IFD1_PER[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc") &(Curated_BivData$Rec.group == 0)] )
-```
-
-```{r show.IFD, echo=FALSE}
-
-biv.IFDs_by_group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
+                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
+
+Rec.group.BivData_IFD.PER_Ttest$p.value
+#p-value < 2.2e-16
+
+#MOUSE AVE
+Rec.group_IFD.PER_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group_IFD.PER_Ttest$p.value 
+#p-value = 6.872e-07
+
+```
+
+The t.tests return very significant pvalues for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
+
+### Q2 Long Bivalent Dataset
+
+Examine the pattern in the Long bivalent dataset to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalant data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+The section on IFD only looks at the 2CO bivalents.
+
+```{r Q2.long_biv.IFD, echo=FALSE}
+#long BivData
+male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
+
+#show
+Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
   geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+ylim(c(0,100))
-biv.IFDs_by_group
-#+theme(legend.position="none")+
-
-biv.IFDs_PER_by.group <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \n Low=0, High=1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-biv.IFDs_PER_by.group
-
-#oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-#  geom_jitter(alpha=.2)+
-#  ggtitle("Interfocal Distances, ABS")+
-#  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = #"free")+ylim(c(0,1))+theme(legend.position="none")
-
-```
-
-Re-order the above figure if possible.
+  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+
+Long.biv.IFDs_PER_by.group
+
+```
+
+table(droplevels(male.long.biv.2CO_IFD$strain) )
+
+The plot above shows the normalized IFD measures from 2CO bivalents
+
+There are alot fewer long bivalent data observations, but the same positive correlation with
+
+-table of long bivalents
+
+- glms
+
+- logistic regression
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE}
+
+#Full dataset
+
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+
+```
+
+There are not enough observations for subsetting the data into strains or use the mouse averages.
 
@@
 
-```{r IFD.poly_log.reg, echo=FALSE, eval=FALSE}
-#This might be a duplicate code chunk
-#test. for the high vs low strains and IFD
-#mouse.avs_4MM_male
-
-mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc", ]
-
-log.reg_IFD.ASB <- glm(Rec.group ~ mean_IFD.2CO_ABS, 
-                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
-#slightly sig
-
- #summary(log.reg_IFD.ASB)
-
-log.reg_IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER, 
-                            data=mouse.avs_4MM_M.male, family=binomial(link="logit"))
- #summary(log.reg_IFD.PER)
-
-#not significant
-```
-
 None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
-
-Post-hoc comparisons ideas?
 
@@
 This also indicated that interference distance may evolved in the house mouse complex.
-
-> Prediction B, Is there evidence for the alternative IFD meansure
 
@@
 <!--
+
+```{r Q2.IFD.log.reg, echo=FALSE, eval=FALSE}
+
+# LOGISTIC REGRESSION FOR IFD
+
+log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *
+
+log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+summary(log.reg_IFD_PER_mouse.av)#NS
+
+### Single Bivalent levels
+
+log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
+              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_IFD_ABS_single.biv)
+
+log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
+              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_IFD_PER_single.biv)
+```
 

### 2020-02-10 doc/Results.rmd
Cleaning up notes and better organization of this doc

@@
 
-- All the effects are unorder factors
+- All the effects are un-ordered factors
 
@@
 (I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
-The strain effects are only slightly significant for within mouse variance measures in the full dataset (includes low quality cells)
+The strain effects are only slightly significant for within mouse variance measures in the full data set (includes low quality cells)
 
@@
 For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
-Most of the strain specific differences in within mouse variance go away in the Q12 dataset.
+Most of the strain specific differences in within mouse variance go away in the Q12 data set.
 
@@
 
+<!--  come back to Fig1
 # Figure 1
@@
 
+-->
+
 # Female Specific Polymorphism
@@
 
-# Proportion of chrm classes
-
+# Proportion of chromosome classes
+
+<!--
 TODO -- try running a chi square test on the proportions.
@@
 make a smaller proportion chrm block --
+Add p.values of tests for the how the proportions differ.
+*cast female data came from another publication, add this ref (Kohler?)
+The bottom lines and subspecies points don't exactly match up.
+**add in cell images like the other draft figures.**
+-->
 
@@
 Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs. (the gwRR for most mice is low/close to the minimum)
-
-This is also another way to score heterochiasmy (that gets around the sex chrms, i think)
 
@@
 
-High musc males have half 1CO and 2CO bivalents.
 For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
 
-Add p.values of tests for the how the proportions differ.
-
-*cast female data came from another publication, add this ref (Kohler?)
-
-The bottom lines and subspecies points don't exactly match up.
-
-**add in cell images like the other draft figures.**
-
-# Chrm Class proportions
-
-The figure below shows the average percentages of chromosome classes per category.
+The figure below shows the average percentages of chromosome classes per category gathered from BivData
 
@@
 
-- Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.
-
-# DMC1 Results Spermatocytes
+Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.
+
+# DMC1 Results, Spermatocytes
 
@@
 
-# Note on Heterochiasmy Definition
+## Note on Heterochiasmy Definition
+
+<!-- add outline for new XX adjust section 
+adjust the colors for the plot below (standardize)
+-->
 
 I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
-In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
-
+In females the XX bivalent is indistinguishable from the autosomes.
+To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
 Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
@@
 The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
-
 To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
@@
 According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
+
+(Put the XX adjustment section here)
 
@@
 
-The breakdown of curated bivalents by category
+The breakdown of single bivalent observations by category
 
@@
 
+Two Main questions for these analyses.
+
 > Q1. driving questions, which traits are sexually dimorphic?
@@
 
+These questions use different datasets, q1 sex matched strains
+
+Q2 all male observations with including strains not in the q1 data set
+
 # Model Frameworks
 
-As alex said, be careful about mixed models and interperting the random effect (radnom variance) -- assigning strain as random effect is slightly problamatic as i can give it a biological interpertation (it's due to strain -- genetics of each of the strains).
-
-> Q1. same models for MLH1 based sex differences (and evolution)
-
-> Q2 the same models for the male polymorphism and logistic regression for testing if bivalent traits can predcit if they are from high or low rec group
+<!--
+As alex said, be careful about mixed models and interpreting the random effect (random variance) -- assigning strain as random effect is slightly problematic as i can give it a biological interpretation (it's due to strain -- genetics of each of the strains).
+-->
+
+Q1. Model 1, mixed model lmer()  
+$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+Q1. Model 2, glm()
+$$mouse \ av.\  metric ~=~ sex * strain + \varepsilon $$
+
+Q2 Model 1, mixed model lmer()
+
+$$mouse \ av.\  metric ~=~ subsp *  + rand(strain) + \varepsilon $$
+
+Q2 Model 2,  glm()
+
+$$mouse \ av.\  metric ~=~ strain + \varepsilon $$
+
+Q2 Model 3, logistic regression
+
+$$Rec \ group  ~=~ mouse \ av.\  metric  $$
 
@@
 
-- Simple application of the same models as MLH1 based reseults.
+- Simple application of the same models as MLH1 based results.
 
@@
 
-```{r biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}
+```{r Q1.biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}
 
@@
 
-```{r biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
 
@@
 
-## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
-
-The general predictions across the males and subspecies based on th above MLH1 results.
-
-For positive correlation traits/metrics
-
-i) in DOM strains, low to no difference across strains
-
-ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
-
-ii) in Mol, MSM > MOLF
-
-A) **SC lengths** will be longer for high Rec strains.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
-
-The mouse averages for the other position meterics will be highly influenced by proportions of the 1CO and 2CO bivalents.
-When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
-
-C.2) sis-co-ten metric ... (what about the clustering?)
-
-C.3) telomere and centromere distances ...
-
-M1 glm model for fixed strain effect across male averages.
-
-# SC Lengths
+# Q1 SC Lengths
 
@@
 We expect female SC lengths to be longer (refs).
-
-In the plot above the SC length ~ higher hand foci cells. Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
-
+In the plot above the SC length ~ higher hand foci cells.
+Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
 For the 0 class chromosomes most all are around the same size of the 1CO distribution.
-
 Add in the code for the plot under 2COs males and females
@@
 #SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
-
 Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
@@
 
-The dependant variables I'll be testing in the mixed model framework are:
+> M1. subsp * sex + (strain)
+
+> M2 subsp \* sex \* strain  (don't rally use)
+
+> M3 sex * strain
+
+The dependent variables I'll be testing in the model framework are:
 
@@
 
-- long.biv
+- Xlong.biv  *Don't do long bivData for testing sex effect, since this would have the XX
+
+- Short.biv dataset *this would be a chrm size effect, without the XX
 
@@
 
-Insert the results for the MMs using the dependant variable 1CO - use the Mixed model (lmer), for all of the meterics (pooled SC, 1CO 2CO ect)
-
-Below's the code for Mixed model results -- try to organize them into a table or somethinf
-
-```{r M1_many.metrics, echo=FALSE}
+Insert the results for the MMs using the dependent variable 1CO - use the Mixed model (lmer), for all of the metrics (pooled SC, 1CO 2CO ect)
+
+Below's the code for Mixed model results -- try to organize them into a table or something
+
+```{r 1.SC.Length_M1, echo=FALSE}
 #https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
@@
 #all pooled SC lengths
-lmer.SCLength_all_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+lmer.Q1_SCLength_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
 
 #these 2 below are LR tests
-drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
-anova(lmer.SCLength_all_MM.M1) # also F values
-
-exactRLRT(lmer.SCLength_all_MM.M1)
-as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+#drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
+#anova(lmer.SCLength_all_MM.M1) # also F values
+exactRLRT(lmer.Q1_SCLength_M1)
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
 
@@
 
-#1CO
-SCLength_1CO_MM.M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-drop1(SCLength_1CO_MM.M1, test="Chisq")
-exactRLRT(SCLength_1CO_MM.M1)
-as.data.frame(round(summary(SCLength_1CO_MM.M1)$coefficients, 6))
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+#drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")
+exactRLRT(Q1.lmer_SCLength_1CO_M1)
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
 
 #2CO
-SCLength_2CO_MM.M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-drop1(SCLength_2CO_MM.M1, test="Chisq")
-exactRLRT(SCLength_2CO_MM.M1)
-as.data.frame(round(summary(SCLength_2CO_MM.M1)$coefficients, 6))
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+#drop1(SCLength_2CO_MM.M1, test="Chisq")
+exactRLRT(Q1.lmer_SCLength_2CO_M1)
+as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
 
 #Long.Biv
-SCLength_long.biv_MM.M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
-
-drop1(SCLength_long.biv_MM.M1, test="Chisq")
-exactRLRT(SCLength_long.biv_MM.M1)
-as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-
-```
-
-#### SC_HetC_M2
+Q1.lmer_SCLength_long.biv_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
+
+#drop1(Q1.lmer_SCLength_long.biv_M1, test="Chisq")
+#exactRLRT(Q1.lmer_SCLength_long.biv_M1)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
+
+#### Q1. SC_HetC_M2
 
@@
 
-```{r M2_many.metrics, echo=FALSE, eval=FALSE}
-
+```{r Q1.SC.Length_M2, echo=FALSE}
 #feed these 2 DFs into M2
@@
 #Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-```
-
-#### SC_HetC_M3
-
-```{r HetC_M3, echo=FALSE}
-
-```
+#glm(metric ~ strain)  ---mean_SC
+
+Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_SCLength_M2)
+
+#Dom.all
+Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.Dom)
+
+#Dom.1CO
+Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.1CO.Dom)
+
+#Dom.2CO
+Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.2CO.Dom)
+
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.Musc)
+
+#Musc.1CO
+Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.1CO.Musc)
+
+#Musc.2CO
+Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.2CO.Musc)
+
+#Mol.all
+Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.Mol)
+
+#Mol.1CO
+Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.1CO.Mol)
+
+#Mol.2CO
+Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.2CO.Mol)
+```
+
+```{r Q1.SC.M2plot, echo=FALSE}
+
+#Q1 plots for SC Lengths across Musc strains
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#mouse av
+Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
+Q1.SClength_Musc_plot
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+Q1.SClength_Musc.BivData_plot
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+Q1.SClength_Dom.BivData_plot
+
+```
+
+I think the strange thing about the Musc is that Skive females have shorter SC lengths.
 
@@
 
-$$mouse \ average \ SC  \ length ~=~ strain + \varepsilon $$
-M2 anova for glm, of each subset, does at least one of the strains differ from others?
-
-M3 logistic regression, can a bivalent metric predict the high or low group?
-
-$$ log( strain ~=~ \ mouse \ average \ SC  \ length ) $$
-
-The IFD traits are:
-
-- IFD_ABS
-
-- IFD_PER
-
-$$mouse \ average \ CO \ position \ trait  ~=~ rand(strain) + \varepsilon $$
-
-These are the models for male polymorphism, (I am re-thinking the tests I think)
-
-Use these DFs for the models:  **Male.poly.Mouse.Table_BivData_4MM**  and **Long.biv_mouse_table_male.poly**
-
-** add chunks for glm()s of male polymorphism**
+# Q1. IFD
+
+Brief lit review for IFD / interference expectations  
+**Petkov et al 2007,  CO interference underlies sex differences in RR**
+
+"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
+However, the interference distance is the same in terms of bivalent length.
+We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
+at the pachytene stage of meiosis."
+
+Chrm1 genetic map,
+
+(other human and mice refs
+
+Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
+
+DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
+
+))
+
+*Do I believe the physical/SC results?*
+
+<!--
+ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
+
+explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
+-->
+
+3) Interference / IFD will not be sexually dimorphic.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
+
+**I**nter**f**ocal **D**istance, a indication of CO interference.
+
+The raw measures are pretty nosy, this might not be the best way to display these results.
+
+ Each point is a IFD observation. Maybe add mouse level lines for the means.
+
+```{r show.IFD_99, echo=FALSE, warning=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+
+#+theme(legend.position="none")
+```
+
+```{r IFD.old.plots, echo=FALSE, eval=FALSE}
+#long SC Long_biv_mouse.avs_4MM
+#real.long.bivData
+#Only use 2CO bivalents
+real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
+real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
+
+long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+```
+
+Still working on the best way to display the general IFD patterns.
+
+### Mixed Model Tests, Fixed Effects
+
+Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+
+$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+### IFD_HetC_M1
+
+```{r HetC_M1_IFD_many.metrics, echo=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+#anova(lmer.IFD.PER_MM.M1)
+
+#exactRLRT(lmer.IFD.PER_MM.M1)
+#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
+
+#M1_IFD.RAW_2CO
+lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
+#anova(lmer.IFD.RAW_MM.M1) # also F values
+
+#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
+#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+```
+
+Sex is a slightly significant factor for raw and normalized IFD.
+
+The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
+That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
+
+show a raw plot, and norm plot and a scatter plot
+
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
+The tables below are from anova( for the lmer model ).
+Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
+
+### IFD_HetC_M2
+
+```{r IFD_HetC_M2, echo=FALSE}
+
+```
+
+### IFD_HetC_M3
+
+```{r IFD_HetC_M3, echo=FALSE}
+
+```
+
+## Strain Comparisons
+
+Dive deeper into the sex specific pattern for each strain.
+Below are code chunks which show the unusual sex specific results for IFD measures.
+The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
+The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
+
+For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
+
+```{r Comp_Ratios, echo=FALSE}
+
+#calq ratios of belo 30
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
+
+#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
+MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
+
+DF.strain.list <- list(
+  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
+  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
+  MSM.Bivalents.DF, MOLF.Bivalents.DF
+  )
+
+entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
+        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
+          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
+
+i = 1
+for(x in (DF.strain.list) ){
+
+  #print(head())
+
+  #these filters won't apply to the indexed DF
+  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
+  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
+  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
+  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
+  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+  i = i+1
+
+  }
+#entr_DF
+
+entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
+
+```
+
+Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
+
+```{r WSB.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+#WSB.Bivalents.DF
+
+#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
+WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
+
+#plot the differences
+wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_raw
+
+wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_PER
+
+WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+WSB.IFDnorm_scatter.plot
+#this plot is much messier than the MSM (PWD?)
+
+#try plotting the other IFD
+
+#make a table of the means and n ect
+wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+
+                       total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+```
+
+The range of normalized IFDs overlap closer in males and females in the WSB data.
+
+```{r LEW.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+
+LEW.Bivalents.DF_IFD <- LEW.Bivalents.DF[(!is.na(LEW.Bivalents.DF$sex)) & (LEW.Bivalents.DF$hand.foci.count == 2),]
+
+LEW.Bivalents.DF_IFD$sex <- droplevels(LEW.Bivalents.DF_IFD$sex)
+
+#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+lew.ploty_IFD_raw <- ggplot( LEW.Bivalents.DF_IFD,  aes(y=IFD1_ABS, x = sex, na.rm = TRUE))+                     geom_jitter( na.rm = TRUE)+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex, na.rm = TRUE))
+lew.ploty_IFD_raw
+#arg, little annoyed that the x NA won't go away
+
+lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_PER
+
+LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  
+  scale_color_manual(values=c('#56B4E9','#E69F00'))+
+  geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+LEW.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+
+                  #     mean_SC = mean(chromosomeLength),
+                       n_SC = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+
+```
+
+The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
+
+```{r G.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+#plot the metric
+#remove the NAs
+
+G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
+
+G.ploty_IFD_raw
+#
+G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.ploty_IFD_PER
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+#calQ percent < 30%
+
+oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
+oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
+
+#the na's are being counted (ughhhhhh)
+#use the file thing instead
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
+  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
+
+LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
+#0.7062663
+
+#would Chi-sq be a good test to compare the ratios ect?
+#the sum of these doesn't add 
+#G.Bivalents.DF
+
+```
+
+```{r PWD.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#remove the NAs
+
+PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
+
+PWD.ploty_IFD_raw
+
+PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
+PWD.ploty_IFD_PER
+
+PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
+
+PWD.IFDnorm_scatter.plot
+
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+```
+
+For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
+
+```{r KAZ.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+#KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+
+#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+
+#remove the NAs
+KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_raw
+
+KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_PER
+
+KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
+
+KAZ.IFDnorm_scatter.plot
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+
+```
+
+For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
+
+```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+
+#remove the NAs
+SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_raw
+
+SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_PER
+
+SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+SKIVE.IFDnorm_scatter.plot
+
+SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+```
+
+In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
+
+```{r MSM.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_raw
+
+MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_PER
+
+#another plot
+MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
+
+MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
+
+MSM.IFDnorm_another.plot
+
+#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
+
+#make a table of the means and n ect
+MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+```
+
+The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
+
+```{r MOLF.IFD, echo=FALSE}
+
+#SHUT DOWN THIS BLOCK UNTIL MOLF FEMALE DATA
+
+#make the dataset
+#MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+
+MOLF.Bivalents.DF <- Curated_BivData %>%  filter(strain == "MOLF")
+
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+#remove the NAs
+#molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
+
+#molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+  geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
+
+#make a table of the means and n ect
+MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
+The other strains which have more overlap between males and females are the Dom strains and KAZ.
+
+#### IFD 3CO bivalents
+
+Run comparisons for 3CO bivalents.
+
+```{r CO3.comparisons, echo=FALSE}
+
+#take a look at the 3CO things
+
+```
+
+<!--
+ Heterochiasmy Prediction
+
+> Do the IFD measures lack significant sex effect? (as predicted)
+
+There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
+
+> Are there signiticant strain effects?
+
+The random strain effect was tested for raw and normalized IFD, neither are significant.
+-->
+
+# Q1. General CO Positions
+
+There are a few traits that fall within the CO positions
+
+- 1CO normalized position
+
+- sis.co.ten (sorta interference)
+
+- centromere and telomere distance
+
+#### Predictions
+
+**Heterochiasmy**
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+**Polymorphism**
+
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
+
+C.3) telomere and centromere distances will be shorter in low rec strains
+
+# Q1. Normalized 1CO positions
+
+**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
+
+```{r CO.positions.figs, echo=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+```{r Q1.Nrm.F1pos, echo=FALSE}
+
+#calq the means for normalized F1
+
+```
+
+Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
+This plot shows the sexual dimorphism in the density plots.
+
+Consider adding annotate_text for the number of observations in each category.
+think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
+
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#MELT_Curated_BivData
+
+#these isn't long.biv data
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+real.long.bivData
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+```
+
+<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
+
+```{r CO1_foci_pos_boxplot, echo=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+##### Mixed models norm 1CO positions
+
+the distribution of SC lengths and sis-coten seems very different across sexes
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+The mixed model data should only come from 1CO bivalent data.
+
+```{r MM_CO_pos, echo=FALSE, eval=FALSE}
+#use the 1CO
+#Use MELT DF
+#MELT_Curated_BivData
+
+#isolate the 1COPos.CO.avs
+#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
+#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
+#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
+
+mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
+
+#<- lme()
+lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
+
+#can't make the 
+CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+#library(lmerTest)
+#anova(CO_position.MM)
+#anova(CO_position.MM, verbose = TRUE)
+
+#summary(CO_position.MM) #no p values,,
+#these just give t and F stats, no pvalue
+
+#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
+#CO_pos_fix.est
+#this test, the full model is more predictive than an empty model?
+
+#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
+#thing2 <- drop1(SC_length.MM, test="Chisq")
+
+CO_pos_rand <- exactRLRT(CO_position.MM)
+CO_pos_rand
+```
+
+```{r log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+
+#LOGISTIC REGRESSION for FOCI POSITION
+
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+```{r t.test_1CO_pos, echo=FALSE, eval=FALSE}
+
+#t.test for high and low group  -- logistic regression still need to format the high and low group
+
+#mouse av
+t.test_mouse.av.1CO_pos <- t.test(
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv.1CO_pos_t.test <- t.test(
+
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+  )
+
+```
+
+the mouse average foci1 pos is more significant in t.test, but not log regression...
+(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
+#### Nrm_1COpos_HetC_M1
+
+#### Nrm_1COpos_HetC_M2
+
+#### Nrm_1COpos_HetC_M3
+
+### Siscoten
+
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males.
+
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
+t.test_mouse.av_siscoten <- t.test(
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+#2CO
+
+#3CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+
+drop1(lme_MM_telo.dist, test="Chisq")
+
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+drop1(lmer_MM_telo.dist, test="Chisq") 
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
+
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
+
+> is random strtain effect significant?
+
+The random strain effect seems very significant.
+
+remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
+
+```{r pos_t.test}
+#make point plots / boxplots which show differences in mean positions
+
+#scatter + boxplot for t.tests
+```
+
+Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
+
+```{r F1.by.F2}
+
+#try remaking the plot Megan suggested
+# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
+
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
+#isolate 2COs
+#facet by sex and subsp
+
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
+
+```
+
+```{r CO.pos.ANOVA, eval=FALSE}
+#what is the pattern of variance
+#run analyses for each subsp*sex
+#use non-melt DF
+
+#how is the variance partioned across
+#cell, mouse, strain
+
+female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
+female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
+
+female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
+
+#unorder strain and mouse
+
+female.Dom$mouse <- as.factor(female.Dom$mouse)
+
+female.Dom$strain <- unclass(female.Dom$strain)
+female.Dom$strain <- as.factor(female.Dom$strain)
+
+female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
+female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
+
+#1CO first
+modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
+
+#can't get mouse and strain to have sum of square
+#residual size decreases with per.F1
+#residuals much larger than fileName, mouse and strain no
+
+#model <- lm(breaks ~ wool * tension, 
+#            data = warpbreaks, 
+#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
+
+male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
+male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
+
+male.Dom$mouse <- as.factor(male.Dom$mouse)
+
+male.Dom$strain <- unclass(male.Dom$strain)
+male.Dom$strain <- as.factor(male.Dom$strain)
+
+male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
+male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
+
+male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
+
+male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
+summary(aov(male.modo))
+
+#only file name is registering as effect
+#Review ANOVA frameworks
+#http://www.biostathandbook.com/nestedanova.html
+
+```
+
+## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
+
+The general predictions across the males and subspecies based on th above MLH1 results.
+
+For positive correlation traits/metrics
+
+i) in DOM strains, low to no difference across strains
+
+ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
+
+ii) in Mol, MSM > MOLF
+
+A) **SC lengths** will be longer for high Rec strains.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
+
+C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
+
+The mouse averages for the other position metrics will be highly influenced by proportions of the 1CO and 2CO bivalents.
+When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
+
+C.2) sis-co-ten metric ... (what about the clustering?)
+
+C.3) telomere and centromere distances ...
+
+M1 glm model for fixed strain effect across male averages.
+
+# Q2 SC
+
+<!--  main DFs for this section
+
+Male.poly.Mouse.Table_BivData_4MM
+
+Bivalent_1o2  (for logistic regression, 1CO and 2COs)
+male.long.biv
+
+-->
+
+remember the 3 predictions for subsp;
+
+i) Dom, no difference
+
+ii) Musc, PWD and SKIVE greater than others
+
+iii) Mol, MSM > MOLF
+
+```{r Q2.male.Nrm1CO_strain, echo=FALSE}
+#isolate all musc
+#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
+```
+
+```{r Q2.SC.ploy.male, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+
+#1. determine which strain effects are sig
+#use mouse biv average
+#Male.poly.Mouse.Table_BivData_4MM
+
+glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm1.SC.male.poly) 
+#SC many strains are different (more than the gwRR)
+#subsps are sig
+
+glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm2.SC.male.poly)
+
+#Dom
+glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(glm_SC.male_Dom)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(glm_SC.male_Musc)
+
+#Mol
+glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(glm_SC.male_Mol)
+
+#Logistic regressions
+
+log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+
+                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
+
+#summary(log.SC.male.poly)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_musc)
+#not sig -- when just looking at Musc
+
+#significant in predicting group
+log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_mol)
+#not sig -- when just looking at Mol
+```
+
+```{r Q2.SC.long.biv, echo=FALSE}
+
+#mouse.av.long.biv.table
+
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
+#summary(glm1.SC.male_long.biv)
+
+glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
+#summary(glm2.SC.male_long.biv)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
+#summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
+#summary(glm_SC.male_Mol_long.biv)
+
+## logistic regression for Rec group
+
+log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
+              data=mouse.av.long.biv.table,  family=binomial(link="logit"))
+
+#summary(log.SC.male_long.SC)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+```
+
+Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
+
+The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).
+
+#### Q2 SC by class
+
+The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
+Mainly that 1CO are shorter in the high rec strains then low rec strains.
+This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
+Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
+This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
+They have a lower probability of 2COs below a certain SC length threshold.
+
+I think these comparisons should be done with the **Bivalent observations**
+The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class
+
+```{r Q2.SC.Chrm.Class, echo=FALSE}
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+#there's no rec group for this table
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
+#there might not be enough data 
+#
+ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
+```
+
+These are the plots of the logistic regression.
+
+```{r Q2.SC.log.regression.show.plots, echo=FALSE}
+
+plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.blah.mouse.av
+
+plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.logReg.mouse.av_1CO
+
+plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.logReg.mouse.av_2CO
+
+```
+
+the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
+
+```{r Q2.SClength.t.test, echo=FALSE}
+#MOUSE AVE, all
+Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
+
+#mouse.av 1CO
+Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
+
+#MOUSE.av 2CO
+Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
+
+#single biv level
+#male.biv?
+
+#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value
+
+```
+
+The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths.
+But When the means are compared across chrm class, the mouse averages are no longer significant.
+
+### Q2 SC length Chrm class prediction
+
+How well does SC length predict chromosome class?
+Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class.
+(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+<!--
+blocks of other male Q2 - SC lengths
+
+what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
+
+```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
+
+#comparing all the strains
+
+T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+)#p-value = 0.001567
+
+#1CO
+T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#2CO
+T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#3CO
+T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#Single Biv level
+T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
+
+#1CO
+T.test_SC.Len_single.biv_1CO <- t.test(
+  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
+
+      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
+  )
+
+#2CO
+T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
+
+                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
+  )
+
+  #3CO
+  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
+
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
+  )
+```
+
+The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
+
+```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
+#I don't think the Mixed models fit well for the high vs low polymorphism tests.
+
+#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
+#MSM_SC vs MOLF_SC
+
+#2. logistic regression for sc length and chrm class?
+#if the prediction / significance
+
+#isolate data
+#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
+
+#build and test MM
+lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+
+lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
+
+anova(lmer_MM_SC_length_male.poly)
+
+rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
+
+```
+
+```{r long.sc_log.res_meanSC, eval=FALSE}
+#Long Biv comparison
+
+#hold off on reporting/ working with the long SC biv resutls
+
+long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
+long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
+
+#divide musc males into groups
+long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
+                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
+
+long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
+                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
+
+summary(long.biv.log.reg)#NS effect (likelt underpowered)
+```
+
+-->
+
+> **Summary**
+
+When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
+I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
+(IE the scatter plots of SC length, binned by chrm class on the X).
+
+Each point is a mouse average SC.
+When all bivalents are pooled, the logistic regression SC lengths predict the groups.
+When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
+For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
+
+ and the mean 2COs SC are higher in the high strains
+(this must have something to do with the physical size effect)
 
@@
 
-### Summary
-
 > Is sex a significant effect for SC length? (as predicted)
@@
 - I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
-
-> Is the random Effect of strain an effect on SC length?
-
-The exactRLRT() test indicates that the random strain effect might be significant, p= xx
-
-(merge this with the code above)
-Prediction, High rec males have longer SC.
 
@@
 
+#male poly shouldn't be here
 Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
@@
                             data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SC.length_mouse.av)
+#summary(log.reg_SC.length_mouse.av)
 
@@
                             data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SC.length_mouse.av_1CO)
+#summary(log.reg_SC.length_mouse.av_1CO)
 
@@
 
-summary(log.reg_SC.length_mouse.av_2CO)
+#summary(log.reg_SC.length_mouse.av_2CO)
 
@@
                             data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SC.length_mouse.av_3CO)
+#summary(log.reg_SC.length_mouse.av_3CO)
 
@@
 
-summary(log.reg_SC.length_BIV)
+#summary(log.reg_SC.length_BIV)
 
@@
                 data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_SC.length_BIV_1CO)
+#summary(log.reg_SC.length_BIV_1CO)
 
@@
                             data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_SC.length_BIV_2CO)
+#summary(log.reg_SC.length_BIV_2CO)
 
@@
 
-summary(log.reg_SC.length_BIV_3CO)
-```
-
-The mean SC logistic regresion model for mouse averages
+#summary(log.reg_SC.length_BIV_3CO)
+```
+
+The mean SC logistic regression model for mouse averages
 
@@
 `r log.reg_SC.length_BIV$coefficients`
-
-#### t.test for SC Length
-
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
-
-#comparing all the strains
-
-T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-)#p-value = 0.001567
-
-#1CO
-T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#2CO
-T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#3CO
-T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#Single Biv level
-T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
-
-#1CO
-T.test_SC.Len_single.biv_1CO <- t.test(
-  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
-
-      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
-  )
-
-#2CO
-T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
-
-                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
-  )
-
-  #3CO
-  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
-
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
-  )
-```
-
-The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-
-```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
-#I don't think the Mixed models fit well for the high vs low polymorphism tests.
-
-#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
-#MSM_SC vs MOLF_SC
-
-#2. logistic regression for sc length and chrm class?
-#if the prediction / significance
-
-#isolate data
-#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
-
-#build and test MM
-lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-
-lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
-
-anova(lmer_MM_SC_length_male.poly)
-
-rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
-```
-
-```{r long.sc_log.res_meanSC, eval=FALSE}
-#Long Biv comparison
-
-#hold off on reporting/ working with the long SC biv resutls
-
-long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
-long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
-                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
-
-long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
-                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
-
-summary(long.biv.log.reg)#NS effect (likelt underpowered)
-```
 
@@
 
-# Q1. IFD
-
-Brief lit review for IFD / interference expectations  
-**Petkov et al 2007,  CO interference underlies sex differences in RR**
-
-"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
-However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
-at the pachytene stage of meiosis."
-
-Chrm1 genetic map,
-
-(other human and mice refs
-
-Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
-
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-
-))
-
-*Do I believe the physical/SC results?*
-
-<!--
-ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
-
-explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
--->
-
-3) Interference / IFD will not be sexually dimorphic.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
-
-**I**nter**f**ocal **D**istance, a indication of CO interference.
-
-The raw measures are pretty nosy, this might not be the best way to display these results.
-
- Each point is a IFD observation. Maybe add mouse level lines for the means.
-
-```{r show.IFD_99, echo=FALSE, warning=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r IFD.old.plots, echo=FALSE, eval=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
-Still working on the best way to display the general IFD patterns.
-
-### Mixed Model Tests, Fixed Effects
-
-Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
-
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-### IFD_HetC_M1
-
-```{r HetC_M1_IFD_many.metrics, echo=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-#anova(lmer.IFD.PER_MM.M1)
-
-#exactRLRT(lmer.IFD.PER_MM.M1)
-#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
-
-#M1_IFD.RAW_2CO
-lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
-#anova(lmer.IFD.RAW_MM.M1) # also F values
-
-#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
-#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-```
-
-Sex is a slightly significant factor for raw and normalized IFD.
-
-The table above should display the slightly unsual pattern, where the coefficients for the significant sex fixed effect are positive and neagtive in the raw and normalized values respectively.
-That is for the raw IFD values, females are signiciantly longer but the normalized IFD values, males are significantly longer.
-
-show a raw plot, and norm plot and a scatter plot
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
-The tables below are from anova( for the lmer model ).
-Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
-
-### IFD_HetC_M2
-
-```{r IFD_HetC_M2, echo=FALSE}
-
-```
-
-### IFD_HetC_M3
-
-```{r IFD_HetC_M3, echo=FALSE}
-
-```
-
-## Strain Comparisons
-
-Dive deeper into the sex specific pattern for each strain.
-Below are code chunks which show the unusal sex specific results for IFD measures.
-The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
-The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
-
-For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
-
-```{r Comp_Ratios, echo=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
-
-```{r WSB.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#WSB.Bivalents.DF
-
-#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
-WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
-
-#plot the differences
-wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_raw
-
-wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_PER
-
-WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-WSB.IFDnorm_scatter.plot
-#this plot is much messier than the MSM (PWD?)
-
-#try plotting the other IFD
-
-#make a table of the means and n ect
-wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
-
-                       total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
-
-The range of normalized IFDs overlap closer in males and females in the WSB data.
-
-```{r LEW.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-
-LEW.Bivalents.DF_IFD <- LEW.Bivalents.DF[(!is.na(LEW.Bivalents.DF$sex)) & (LEW.Bivalents.DF$hand.foci.count == 2),]
-
-LEW.Bivalents.DF_IFD$sex <- droplevels(LEW.Bivalents.DF_IFD$sex)
-
-#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-lew.ploty_IFD_raw <- ggplot( LEW.Bivalents.DF_IFD,  aes(y=IFD1_ABS, x = sex, na.rm = TRUE))+                     geom_jitter( na.rm = TRUE)+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex, na.rm = TRUE))
-lew.ploty_IFD_raw
-#arg, little annoyed that the x NA won't go away
-
-lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
-lew.ploty_IFD_PER
-
-LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  
-  scale_color_manual(values=c('#56B4E9','#E69F00'))+
-  geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-LEW.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-
-                  #     mean_SC = mean(chromosomeLength),
-                       n_SC = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-
-```
-
-The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
-
-```{r G.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-#plot the metric
-#remove the NAs
-
-G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
-
-G.ploty_IFD_raw
-#
-G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.ploty_IFD_PER
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-
-#calQ percent < 30%
-
-oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
-oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
-
-#the na's are being counted (ughhhhhh)
-#use the file thing instead
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
-  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
-
-#uggh how does this not work ...
-
-#more males have lower than 30% IFD?
-
-LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
-#0.7062663
-
-#would Chi-sq be a good test to compare the ratios ect?
-
-#the sum of these doesn't add
-
-G.Bivalents.DF
-
-```
-
-```{r PWD.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#remove the NAs
-
-PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
-
-PWD.ploty_IFD_raw
-
-PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
-PWD.ploty_IFD_PER
-
-PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
-
-PWD.IFDnorm_scatter.plot
-
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-```
-
-For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
-
-```{r KAZ.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-
-#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_raw
-
-KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_PER
-
-KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
-
-KAZ.IFDnorm_scatter.plot
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-
-```
-
-For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
-
-```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_raw
-
-SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_PER
-
-SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-SKIVE.IFDnorm_scatter.plot
-
-SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-```
-
-In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
-
-```{r MSM.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_raw
-
-MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_PER
-
-#another plot
-MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
-
-MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
-
-MSM.IFDnorm_another.plot
-
-#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
-
-#make a table of the means and n ect
-MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-```
-
-The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
-
-```{r MOLF.IFD, echo=FALSE}
-
-#SHUT DOWN THIS BLOCK UNTIL MOLF FEMALE DATA
-
-#make the dataset
-#MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-
-MOLF.Bivalents.DF <- Curated_BivData %>%  filter(strain == "MOLF")
-
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-#remove the NAs
-#molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
-
-#molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+  geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
-
-#make a table of the means and n ect
-MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
-The other strains which have more overlap between males and females are the Dom strains and KAZ.
-
-#### IFD 3CO bivalents
-
-Run comparisons for 3CO bivalents.
-
-```{r CO3.comparisons, echo=FALSE}
-
-#take a look at the 3CO things
-
-```
-
-<!--
- Heterochiasmy Prediction
-
-> Do the IFD measures lack significant sex effect? (as predicted)
-
-There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
-
-> Are there signiticant strain effects?
-
-The random strain effect was tested for raw and normalized IFD, neither are significant.
--->
-
-# Q1. General CO Positions
-
-There are a few traits that fall within the CO positions
-
-- 1CO normalized position
-
-- sis.co.ten (sorta interference)
-
-- centromere and telomere distance
-
-#### Predictions
-
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-
-  C) Centromere and telomere distances will be sexually dimorphic.
-
-**Polymorphism**
-
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
-
-C.3) telomere and centromere distances will be shorter in low rec strains
-
-# Q1. Normalized 1CO positions
-
-**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
-
-```{r CO.positions.figs, echo=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-
-```
-
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-#real.long.bivData
-#MELT_Curated_BivData
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-```{r CO1_foci_pos_boxplot, echo=FALSE}
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-##### Mixed models norm 1CO positions
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-The mixed model data should only come from 1CO bivalent data.
-
-```{r MM_CO_pos, echo=FALSE, eval=FALSE}
-#use the 1CO
-#Use MELT DF
-#MELT_Curated_BivData
-
-#isolate the 1COPos.CO.avs
-#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
-#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
-#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
-
-mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
-
-#<- lme()
-lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
-
-#can't make the 
-CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-#library(lmerTest)
-#anova(CO_position.MM)
-#anova(CO_position.MM, verbose = TRUE)
-
-#summary(CO_position.MM) #no p values,,
-#these just give t and F stats, no pvalue
-
-#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
-#CO_pos_fix.est
-#this test, the full model is more predictive than an empty model?
-
-#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
-#thing2 <- drop1(SC_length.MM, test="Chisq")
-
-CO_pos_rand <- exactRLRT(CO_position.MM)
-CO_pos_rand
-```
-
-```{r log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-```{r t.test_1CO_pos, echo=FALSE, eval=FALSE}
-
-#t.test for high and low group  -- logistic regression still need to format the high and low group
-
-#mouse av
-t.test_mouse.av.1CO_pos <- t.test(
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv.1CO_pos_t.test <- t.test(
-
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-  )
-
-```
-
-the mouse average foci1 pos is more sig in t.test, but not log regression... (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-#### Nrm_1COpos_HetC_M1
-
-#### Nrm_1COpos_HetC_M2
-
-#### Nrm_1COpos_HetC_M3
-
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion struction/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
-drop1(lme_MM_telo.dist, test="Chisq")
-
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
-### Heterochiasmy Prediction
-
-> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
-
-> is random strtain effect significant?
-
-The random strain effect seems very significant.
-
-remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
-
-```{r pos_t.test}
-#make point plots / boxplots which show differences in mean positions
-
-#scatter + boxplot for t.tests
-```
-
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
-
-```{r F1.by.F2}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-```{r CO.pos.ANOVA, eval=FALSE}
-#what is the pattern of variance
-#run analyses for each subsp*sex
-#use non-melt DF
-
-#how is the variance partioned across
-#cell, mouse, strain
-
-female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
-female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
-
-female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
-
-#unorder strain and mouse
-
-female.Dom$mouse <- as.factor(female.Dom$mouse)
-
-female.Dom$strain <- unclass(female.Dom$strain)
-female.Dom$strain <- as.factor(female.Dom$strain)
-
-female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
-female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
-
-#1CO first
-modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
-
-#can't get mouse and strain to have sum of square
-#residual size decreases with per.F1
-#residuals much larger than fileName, mouse and strain no
-
-#model <- lm(breaks ~ wool * tension, 
-#            data = warpbreaks, 
-#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
-
-male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
-male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
-
-male.Dom$mouse <- as.factor(male.Dom$mouse)
-
-male.Dom$strain <- unclass(male.Dom$strain)
-male.Dom$strain <- as.factor(male.Dom$strain)
-
-male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
-male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
-
-male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
-
-male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
-summary(aov(male.modo))
-
-#only file name is registering as effect
-#Review ANOVA frameworks
-#http://www.biostathandbook.com/nestedanova.html
-
-```
-
-# Q2 SC
-
-I didn't have a good prediction for the male polymorphism... So I'll just do post-hoc comparisons for the groups to test if they are different.
-
-**What is the overlap for cell images in the MLH1 comparisons and these bivalent comparisons?**
-
-should all mice be represented.
-
-```{r male.Nrm1CO_strain, echo=FALSE}
-#isolate all musc
-Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-#make table
-Musc_male_1COpos_table <- ddply(Curated_BivData_male[Curated_BivData_male$hand.foci.count == 1,], c("strain"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE)
-                         )
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-
-```
-
-## SC_male.poly_glm
-
-remeber the 3 predictions for subsp;
-
-i) Dom, no difference
-
-ii) Musc, PWD and SKIVE greater than others
-
-iii) Mol, MSM > MOLF
-
-```{r Q2.SC.ploy.male, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-
-#1. determine which strain effects are sig
-#use mouse biv average
-#Male.poly.Mouse.Table_BivData_4MM
-
-glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#SC many strains are different (more than the gwRR)
-#subsps are sig
-
-glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#Dom
-glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glm_SC.male_Dom)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glm_SC.male_Musc)
-
-#Mol
-glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glm_SC.male_Mol)
-
-#Logistic regressions
-
-log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-
-                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
-
-#summary(log.SC.male.poly)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_musc)
-#not sig -- when just looking at Musc
-
-#significant in predicting group
-log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_mol)
-#not sig -- when just looking at Mol
-```
-
-```{r Q2.SC.long.biv, echo=FALSE}
-
-#mouse.av.long.biv.table
-
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
-summary(glm1.SC.male_long.biv)
-
-glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
-summary(glm2.SC.male_long.biv)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
-summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
-summary(glm_SC.male_Musc_long.biv)
-
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
-summary(glm_SC.male_Mol_long.biv)
-
-## logistic regression for Rec group
-
-log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-              data=mouse.av.long.biv.table,  family=binomial(link="logit"))
-
-summary(log.SC.male_long.SC)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-```
-
-The glm for SC lengths -- sub-grouped by subsp -- indicates that all strain effects are significant for the average mouseisn't met.  SC length.
-My prediction for Dom, isn't met.
-
-(indecation that there is more variation for the SC lengths than for gwRR / CO counts)
-
-#### Q2 SC by class
-
-<!--  main DFs for this section
-
-Male.poly.Mouse.Table_BivData_4MM
-
-Bivalent_1o2  (for logistic regression, 1CO and 2COs)
-male.long.biv
-
--->
-I think these comparisons should be done with the **Bivalent observations**
-The logistic regression showed rec groups could be predicted by SC length - re-run them while seperating out by chrm class
-
-```{r Q2.SC.Chrm.Class, echo=FALSE}
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
-#there might not be enough data 
-#
-ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
-```
-
-These are the plots of the logistic regression.
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.blah.mouse.av
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.logReg.mouse.av_1CO
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.logReg.mouse.av_2CO
-
-```
-
-the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
-
-```{r Q2.SClength.t.test, echo=FALSE}
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
-
-```
-
-The top t.tests indicate that when all the mice averages are pooled, there's a sig difference in SC lengths.
-But When the means are compared across chrm class, the mouse averages are no longer significant.
-
-### Q2 SC length Chrm class prediction
-
-How well does SC length predict chromosome class?
-Prediction, High rec strains will have more sig pvalue, given the lower overlap in SC lengths across chrm class.
-(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-> **Summary**
-
-When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
-I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
-(IE the scatter plots of SC length, binned by chrm class on the X).
-
-Each point is a mouse average SC.
-When all bivalents are pooled, the logistic regression SC lengths predict the groups.
-When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
-
- and the mean 2COs SC are higher in the high strains
-(this must have something to do with the physical size effect)
-
 # Q2 Normalized CO positions
@@
 
-- Is there a consistant pattern with the high rec strains?
+- Is there a consistent pattern with the high rec strains?
 
@@
 
-The t.tests return very significant pvalues for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
+The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
 
@@
 
-Examine the pattern in the Long bivalent dataset to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalant data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
 The section on IFD only looks at the 2CO bivalents.
@@
 
-There are alot fewer long bivalent data observations, but the same positive correlation with
+There are many fewer long bivalent data observations, but the same positive correlation with
 
@@
 
-There are not enough observations for subsetting the data into strains or use the mouse averages.
-
-Neither t.test are sig for both the ABS and PER when I test just the Musc strains.
+There are not enough observations for sub setting the data into strains or use the mouse averages.
+
+Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
 The above t.tests are breaking the knitr
@@
 
-put all of the code chuncks/analysis for caveats here
+put all of the code chunks/analysis for caveats here
 
@@
 The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
-Most plts excluded for space.
+Most plots excluded for space.
 
@@
 I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
-(the extra stats are not correclty mapped)
+(the extra stats are not correctly mapped)
 
@@
 
-in code chunk above I ran the mouse averages for the longest bivalets. (680 bivalents, from 54 mice. 
+in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
 10202 bivalents from 86 mice.
@@
 
-2. Expected impact on sex comparisons, extimated effect size of the X  
+2. Expected impact on sex comparisons, estimated effect size of the X  
 3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
@@
 
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC araea due to XX /*
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
 
@@
 
-The first plot showing the mean sc lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
+The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
 
@@
 Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
-on average, each of the top longest bivalents make up ~10% of the cell's totl SC area.
+on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
 So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
@@
 
-- also interesting, the pwd and msm dont have longer sc, compared to other strains.
+- also interesting, the pwd and msm don't have longer SC, compared to other strains.
 
@@
 
-For the Automated dataset, I like to measure the rate of passing bivlents per cell. The mean pass rate will be multipled to the estimated XX mean_SC.
+For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
 
@@
 
-(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be calqd on the mouse level.
-(it could even be extended to cell level -- except i dont think the XX sc length estimates wont be good.)
+(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
+(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
 
@@
 use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
-Make fake datasets which sample 19 bivalents, for 'in silico' cells for males and females.
+Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
 Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
 Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
-The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total dataset.
+The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
 

### 2020-02-10 doc/Results.rmd
work on 1CO landscape section

@@
 
-#### Predictions
-
 **Heterochiasmy**
@@
   C) Centromere and telomere distances will be sexually dimorphic.
-
-**Polymorphism**
-
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
-
-C.3) telomere and centromere distances will be shorter in low rec strains
 
@@
 
-```{r Q1.Nrm.F1pos, echo=FALSE}
+```{r Q1.Nrm.F1pos_M2, echo=FALSE}
 
 #calq the means for normalized F1
+#M2
+
+Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_nrmF1pos_M2)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+```{r Q1.Nrm.F1pos_M2_long.biv, echo=FALSE}
+
+#table for the 1CO long biv across strain and sexes?
 
@@
 
-```{r MM_CO_pos, echo=FALSE, eval=FALSE}
+#### Nrm_1COpos_HetC_M1
+
+```{r Q1.MM_CO_pos, echo=FALSE, eval=FALSE}
 #use the 1CO
@@
 
-```{r log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+```{r Q1.nrmFoci1_M1, echo=FALSE}
+
+#M1
+
+```
+
+the mouse average foci1 pos is more significant in t.test, but not log regression...
+(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
+#### Nrm_1COpos_HetC_M2
+
+```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
 
@@
 
-```{r t.test_1CO_pos, echo=FALSE, eval=FALSE}
+```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}
 
@@
 ```
-
-the mouse average foci1 pos is more significant in t.test, but not log regression...
-(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-#### Nrm_1COpos_HetC_M1
-
-#### Nrm_1COpos_HetC_M2
-
-#### Nrm_1COpos_HetC_M3
 
@@
 male.bivdata.1CO
+
+**Polymorphism**
+
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
+
+C.3) telomere and centromere distances will be shorter in low rec strains
 

### 2020-02-10 doc/Results.rmd
add motivating sents to v2 outline

@@
 
+#2 sample prop test
+prop.test(x, y)
+
+x - list of the number of 'successes' across groups  (1COs)
+y - list of the number of 'failures' across groups  (2COs)
+
+(does a chi square test)
+
 make a smaller proportion chrm block --
@@
 
-#this first one is the 'first' MLH1 count
-#which is this variable
+#prop.test(x, y)  x number of successes, y total  
+#strain sums of n
+#look up continuity correction
+
+Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),]
+
+female.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(sex == "female")
+male.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(sex == "male")
+
+n1CO_list_F <- c( female.mouse.av$n1CO)
+n1CO_list_M <- c( male.mouse.av$n1CO)
+
+total_biv_F_list <- c( female.mouse.av$tot.biv)
+total_biv_M_list <- c( male.mouse.av$tot.biv)
+
+nbiv_total_list <- c(Mouse.Table_BivData_4MM$tot.biv)
+
+prop.test(n1CO_list_F, total_biv_F_list )
+
+#perform test across strain
+
+Chrm.prop_strain_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
+
+                  nbiv = sum(tot.biv ),              
+                 n0CO = sum(n0CO ),
+                  n1CO = sum(n1CO ),
+                 n2CO = sum(n2CO ),
+                  n3CO = sum(n3CO )
+
+                          )
+
+chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
+
+female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$nbiv )
+female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$nbiv )
+female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$nbiv )
+
+chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
+#chrm.prop.strain.male.high.rec <-  Chrm.prop_strain_table %>%  filter(sex == "male")
+#chrm.prop.strain.male.low.rec <-  Chrm.prop_strain_table %>%  filter(sex == "male")
+
+male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$nbiv )
+male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$nbiv )
+male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$nbiv )
 
@@
 #oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = #proportion_stats_category)
-
 #summary(oopp)
@@
 #0CO and 1CO are significant when only chrm classes are included (and intercept)
-
 #no longer significant when used
@@
 #0CO is now more significant, and 2CO is sightly significant
-
 #2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
-
 #MSM needs more bialent observations
@@
 
-<!--Brief lit review for SC length and MLH1 expections. -->
+<!--Brief lit review for SC length and MLH1 expections.
+add XX adjusment / arguments
+-->
 
@@
 
-#### Strain Comparisons of SC Length
-
-Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
-
-(adding all of these blocks to -- the setup script); strain.Bivalents.DF
+#### Q1. SC Length t.test
 
@@
 
+#### Strain Comparisons of SC Length
+
+Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
+
+(adding all of these blocks to -- the setup script); strain.Bivalents.DF
+
+#### XX adjustment
+
+Generate inteligible summaries for the permutation / subsampling results.
+
+```{r Q1_XX.adjustment, echo=FALSE}
+
+```
+
 # Q1. IFD
@@
 
+<!--
+
 There are a few traits that fall within the CO positions  
-
 - 1CO normalized position  
-
 - sis.co.ten (sorta interference)  
-
 - centromere and telomere distance  
-
 **Heterochiasmy**
@@
 2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
-
   B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-
   C) Centromere and telomere distances will be sexually dimorphic.
-
-# Q1. Normalized 1CO positions
-
 **for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
+-->
 
@@
 CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
+
+#make a table for the sex specific values of the normalized F1 positions
+
+#for strain, sex -- Curated_BivData
+#or (Mouse.Table_BivData_4MM)
+#norm.first.foci.pos
+
+Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
+
+                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
+                  n1CO = sum(n1CO )
+
+                          )
 

### 2020-02-10 src/BivData_Setup.Rmd
add motivating sents to v2 outline

@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
+                        Ncells  = length(  unique(fileName) ),
+                      tot.biv = length(Obj.ID),
+
+                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
 
@@
 )
+Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),]
 
@@
 Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
-#MOLF
-
-#These shouldn't be needed for now.
 
@@
                         se   = round(sd / sqrt(Ncells), 3),
+
+                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
 

### 2020-02-11 doc/Results.rmd
reordering sections to match the outline, cutting/summarizing and simplifying

@@
 
-# Average MLH1 Count Model
+# gwRR estimates
+
+make table of the strain averages
+
+```{r MLH1.summ.stats, echo=FALSE}
+
+#table for the means by strain
+
+```
+
+## Female Specific Analysis
+
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+table( droplevels(female_mouse.avs_for.analysis$strain) )
+#there are cast strains too
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.female.plot
+
+#Mol
+Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.female.plot
+
+#Dom version 
+Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.female.plot
+#Gough female is higher
+
+#variance plots
+musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+```
+
+```{r female.strain.table, echo=FALSE}
+
+#female strain table
+#female_mouse.avs_for.analysis
+F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  = sum(Ncells),
+                        mean_co =as.numeric(format(round(mean(mean_co), 3 ), nsmall=3) ),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se)
+    )
+
+```
+
+```{r Female.specific.glm, echo=FALSE}
+
+glm.female.strain.subsp <- lm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+
+#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+lm.female.strain_sum <- summary(lm.female.strain)
+
+#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+
+#how much greater is G female
+G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
+mean.G.female <- mean(G_female$mean_co)
+
+sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM")
+mean.all.other <- mean(sans_high_female$mean_co)
+
+#how much greater is MSM female
+MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
+mean.MSN.female <- mean(MSM_female$mean_co)
+
+sans_MSM_female <- female_mouse.avs_for.analysis %>% filter(strain != "MSM")
+mean.all.other <- mean(sans_MSM_female$mean_co)
+
+mean.MSN.female / mean.all.other
+
+## within mouse variance models 
+lmVAR.female.strain.subsp <- lm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- lm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- lm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+#LEW not sig
+
+lm.m2.Q12_female <- lm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
+#LEW sig
+```
+
+The range of strain means is `r min()`
+
+the ranges of mouse means for each strain is X
+
+MSM, G, LEW have significant strain effects.
+
+## Male Specific Analysis
+
+Follow up on this finding -- included more strains which were excluded from the HetC analysis.
+
+> Is there a strain effect in Dom?
+
+> How different are the Musc strains?
+
+test a model for fixed effects of strains within musc. The purpose is to distinguish musc strains from low 'Dom' level strains.
+
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+#try customize the plot size
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#table( droplevels(just.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.male.plot
+
+#Dom version 
+Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.male.plot
+
+```
+
+(Try removing the alpha legend)
+
+This is the table and plot of new Musc strains
+
+`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
+
+Given the plot, make note of the potential outlier mice.  The strains have unequal variance
+
+KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
+
+PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
+
+```{r male.strain.table, echo=FALSE}
+
+#female strain table
+#female_mouse.avs_for.analysis
+M.strain.table <- ddply(just.male_mouse.av, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  = sum(Ncells),
+                        mean_co =as.numeric(format(round(mean(mean_co), 3 ), nsmall=3) ),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se)
+    )
+
+```
+
+```{r Male.specific.glm, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+
+summry.male.strain.effects <- summary(glm.male.strain.subsp)
+
+#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+
+#1. determine which strain effects are sig
+#M2
+lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
+lm.male.strain_sum<- summary(lm.male.strain)
+
+#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+
+#how much greater are the High Rec gwRRs
+just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
+                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
+                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
+
+#how much bigger are the high rec male means?
+sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+#mean(sans_LewRec_male$mean_co)
+
+PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+#mean(SKIVE_male$mean_co) /  
+  mean(sans_LewRec_male$mean_co)
+
+## within mouse variance models 
+lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+#all p values kinda high, MSM and TOM are slightly sig
+
+lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+#non sig
+
+lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+#NS
+
+lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+#NS
+
+```
+
+the ranges of means are X
+
+PWD, MSM and SKIVE have strain significant strain effects so they will be designedated into the high rec group.
+
+# Analysis for Evolutionary Patterns
+
+In order to understand the variance within an evolutionary framework we fit the mouse gwRRs to a mixed model, which specified subspecies, sex and the interaction as fixed effects.
+The subspecies effect - is a proxy for measuring the degree of divergence across subspecies.
+The strains were coded as a random effect to -simulate / approximate / the standing genetic variation across subspecies.
 
@@
 
+<!--
 Setting up the data set for using the mixed model (base data set, HQ data set)
@@
 - only Mus musculus strains with sex matched observations  
-
 - Male gwRR have been adjusted for the PAR
-
 - Q5 excluded
-
 - All the effects are un-ordered factors  
-
 - male ages within 5 - 20 weeks
-
 - Molossinus is coded as a subsp level
 
+subspecies levels: 3, Musc, Dom and Molossinus
+
+strain levels: Dom: 3, Musc: 3  Mol: 2
+
+sex: 2 female male  
+-->
+
 ```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=FALSE}
-# MOVE this into BIVDATA SETUP!!!
-
-#should be able to use the loaded 
 #DF.HetC.MixedModel.HQ
@@
 
-<!--Can I indicate the average number of cells with go into each mouse.? -->
-
-subspecies levels: 3, Musc, Dom and Molossinus
-
-strain levels: Dom: 3, Musc: 3  Mol: 2
-
-sex: 2 female male
-
-For now I will focus on 3 different models for comparing the subsp, sex and strain effects across relevant dependent variables.
-
 1. $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
@@
 
-2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
-
-Post hoc model for investigating strain as a fixed effect.
-
-3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$
-
-Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.
-
-## MLH1 count M1
-
-$$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-```{r MLH1co.av.M1, echo=FALSE, digits = 3}
+```{r MLH1co.MM, echo=FALSE, digits = 3}
 #THIS IS THE NEW MIXED MODEL
@@
 MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-#summary(Reduced.strain.HQ)
-
 #summary(MLH1.MM.M1)
-drop1(MLH1.MM.M1, test="Chisq") # p = 0.0002045 ***
-
-as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
+
+MM.Fixed.pval <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+
+#as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
 
 #only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
-anova(MLH1.MM.M1)
-
-exactRLRT(MLH1.MM.M1) #< 2.2e-16  Random strain effect, the variance due to random strain effect is NOT 0
-
-```
-
-Above is the mixed model M1,
-
-## MLH1 count M2
-
-$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$
+
+```
+
+The significance of the interaction effect for the 2 fixed effects is `r MM.Fixed.pval`. The random effect for strain effect is `r MM.rand.pval`.
+
+### glm followup
+
+To follow up on the mixed model results, we also ran glm / linear models to get a better idea of the strain by sex interaction effects.
+
+2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
+
+Post hoc model for investigating strain as a fixed effect.
 
@@
 #strain = fixed
-lm.strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-
-anova(lm.strain.fixed.full)
-
-#this table of results is broken -- might be too long -- lm
-#lm has different names
-as.data.frame(round( summary(lm.strain.fixed.full)$coefficients, 6))
-
-#summary(strain.fixed.full)
-#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
-
-#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are 
+glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summ <- summary(glm.strain.fixed.full)
+
+#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
+
 ```
@@
 
-## MLH1 count M3
+3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$
+
+Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.
+
+$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$
 
@@
 -->
-
-#### Summary of Mixed models
 
@@
 
-### MLH1 variance M1
-
 $$mouse metric  ~=~ subsp * sex + rand(strain)+ \varepsilon$$
@@
 
-### MLH1 variance M2
-
 $$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
@@
 Most of the strain specific differences in within mouse variance go away in the Q12 data set.
-
-### MLH1 variance M3
 
@@
 
-# Female Specific Polymorphism
-
-I should run the same questions for the female data --
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#Mol
-Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
-
-Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.female.plot
-
-#Dom version 
-Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
-
-Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.female.plot
-#Gough female is higher
-
-#variance
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.female.plot
-
-Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.female.plot
-
-```
-
-```{r Female.specific.test, echo=FALSE}
-
-lm.female.strain.subsp <- lm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.effects <- summary(lm.female.strain.subsp)
-#strainG, 1.06e-05 ***
-#strainLEW  0.0273 *
-
-lm.female.strain <- lm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-lm.female.strain_sum <- summary(lm.female.strain)
-#G  1.06e-05 ***
-#LEW   0.0273 *
-#PWD  0.0810 .
-#MSM  3.02e-05 ***
-
-#how much greater is G female
-G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co)
-
-sans_G_female <- female_mouse.avs_for.analysis %>% filter(strain != "G")
-mean.all.other <- mean(sans_G_female$mean_co)
-
-## within mouse variance models 
-lmVAR.female.strain.subsp <- lm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- lm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- lm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-#LEW not sig
-
-lm.m2.Q12_female <- lm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
-#LEW sig
-```
-
-# Male Specific Polymorphism
-
-Follow up on this finding -- included more strains which were excluded from the HetC analysis.
-
-> Is there a strain effect in Dom?
-
-> How different are the Musc strains?
-
-test a model for fixed effects of strains within musc. The purpose is to distinguish musc strains from low 'Dom' level strains.
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
-
-#Dom version 
-Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-
-```
-
-(Try removing the alpha legend)
-
-This is the table and plot of new Musc strains
-
-`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
-
-Given the plot, make note of the potential outlier mice.  The strains have unequal variance
-
-KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
-
-PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
-
-```{r Male.specific.test, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-
-#1. determine which strain effects are sig
-
-lm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.effects <- summary(lm.male.strain.subsp)
-
-lm.male.strain <- lm(mean_co ~ strain, data= just.male_mouse.av)
-lm.male.strain_sum <- summary(lm.male.strain)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /  
-  mean(sans_LewRec_male$mean_co)
-
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#all p values kinda high, MSM and TOM are slightly sig
-
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#NS
-
-```
-
-msm, pwd skive are when all other male strains are added. skive is less sig than the others -- but it's still sig.  Given these results I can group SKIVE, PWD and MSM into the 'high rec' group.
-
-*PWD, MSM, SKIVE are in 'high rec group'*
-
-# Proportion of chromosome classes
+# Proportion of Chromosome Classes
+
+In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes with different numbers of COs)
 
@@
 
-```{r test.Chrm.Prop, echo=FALSE}
+```{r test.Chrm.Prop, echo=FALSE, eval=FALSE}
 
@@
                   n3CO = sum(n3CO )
-
                           )

### 2020-02-11 src/BivData_Setup.Rmd
fixed a bug for a strain level table

@@
                        Nbiv = length(Obj.ID), #if there are na's below -- it breaks
-                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
-                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
-                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
-                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
-
-                       ncells = length(unique(fileName)),
-                       nbivs = length(unique(Obj.ID)),
+                       tot.biv = length(Obj.ID),
+
+                       n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
+                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),
 
@@
                        sd.SC_length   = round(sd(chromosomeLength), 3),
-                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(Nbiv), 3),
 
@@
 
-as.data.frame( head(mouse.level.BivData_table) )
-#kable didn't work
+mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_category(mouse.level.BivData_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
 
@@
 
-#is there a way that the above table would be different than the 
 Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
@@
 )
-Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),]
 
@@
 #remove all the non-used strains!!
-Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$strain != "AST",]
-Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[Curated_BivData$strain != "TOM",]
-Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$strain != "CZECH",]
-Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
-
-#assign high and low
-#divide musc males into groups
-#mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-#                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-#                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))#
-#
-#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-#                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+
+Mouse.Table_BivData_4MM <-  Mouse.Table_BivData_4MM %>%  filter(strain != "AST")  %>% filter(strain != "TOM")  %>% filter(strain != "CZECH")
+
+#Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
+#Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),] #so many NAs still
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                              ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)))
+
 ```
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_3.2.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_11.2.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-02-11 src/BivData_Setup.Rmd
removed the empty hand foci lines from a mouse table loop

@@
 ToDo
+
+- Fix the hand.foci.count ==2, 5jan19_27oct18_SKIVE_m2_sp1_3.2_rev_13
+
+- Fix all the hand.foci == 4, which don't have the right number of positions
 
@@
 
-mouse.level.BivData_table <- ddply(.data=Curated_BivData, 
+#remove bivalent entries without hand foci count values
+
+cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
+                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
+
+mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
                        .(mouse),
@@
                        Nbiv = length(Obj.ID), #if there are na's below -- it breaks
-                       tot.biv = length(Obj.ID),
+
+                      tot.biv = length( !is.na(hand.foci.count) ),
 
@@
 
+mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
+                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
 ```

### 2020-02-12 doc/Results.rmd
end o day o

@@
 
-load(file="~./MLH1repo/data/CleanBivData_3.2.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_11.2.20.RData")
 
@@
 
+# Main Figure
+
+```{r first.boxplots, echo=FALSE, fig.align='center', warning=FALSE}
+
+#the below ordering works!! keep it
+MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
+              "PWD","CZECH","SKIVE",
+              "KAZ","TOM","AST",
+              "MSM","MOLF",
+              "CAST","HMI","SPRET","SPIC", "other"), order=T )
+#order the data frame
+MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
+  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+ ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
+#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+
+
+justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")
+
+#scales="free_y"
+#this is ok except for the set placement of mice on x axis
+#if the full dataset was ordered correctly, this would kinda fix it
+
+#boxplots by strain
+ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())+
+  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )+scale_fill_manual(values=colors_of_strains)
+
+ss <- ss   + facet_wrap(~ sex, scales="free")
+
+#these plot currently look bad
+```
+
+```{r table.for.plot, echo=FALSE}
+#why do I add in a different table?
+
+#this move this to the setup file
+
+strain.table.HQ_Fig.og <- ddply(MLH1_data, c("strain", "sex"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  =length(nMLH1.foci),
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci, na.rm = TRUE),
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                       #Standard error (mice vs cells)
+                        se  = round(sd / sqrt(Ncells), 3)
+    )
+
+strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
+                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")
+
+#adjust order of plot here
+strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
+     "MSM", "MOLF",   
+  "PWD", "SKIVE",  "KAZ",
+  "WSB","G","LEW","PERC",
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+```
+
+```{r Fig1, echo=FALSE, warning=FALSE}
+
+#might need to nudge the X values
+
+Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+
+
+    theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+#remove y labfor male
+
+   #segments
+#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
+  #cast
+#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
+  #outgroups
+ # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+
+ #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
+   #        y = c(18.5,18.5,33,18.5, 18.5),
+
+  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
+    #       fontface = 'italic', hjust = 0.5, size=3.5)
+
+Fig1_v1_strain
+```
+
+This is a draft of the main figure. Outgroups are not included in the analysis below.
+
 <!--  insert Details of data set cleaning etc is in SetupData.rmd
 Effects from different factors are reviewed in Effects.Rmd -->
+
+```{r maw.fig1s, echo=FALSE, include=FALSE}
+
+strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
+
+Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point( size=3, position=position_dodge(.5))+
+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+
+
+    theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+
+#strain.table.HQ_Fig.og
+
+strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")
+
+strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
+     "MSM", "MOLF",   
+  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
+  "WSB","G","LEW"), order=T )
+
+Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(size=3, position=position_dodge(.5))+
+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+
+
+    theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+
+```
 
@@
 
-# gwRR estimates
-
-make table of the strain averages
+# Genome wide recombiantion rate estimates
 
@@
 #Gough female is higher
-
-#variance plots
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
 
@@
 
+G and LEW have a higher strain means compared to WSB. MSM (and PWD) also has an elevated mean.
+
+The variance across mouse means for each strain is also variable across strains.
+
+In the dom  subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in WSB, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in G, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in LEW.
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in PWD, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in KAZ.
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in MSM.
+
+Two glm models were run to test the effects of strain and subspecies.
+
 ```{r Female.specific.glm, echo=FALSE}
@@
 summry.female.strain.effects <- summary(glm.female.strain.subsp)
+
+round(summry.female.strain.effects$coefficients[,c(1,4)],5)
 
@@
 lm.female.strain_sum <- summary(lm.female.strain)
+
+round(lm.female.strain_sum$coefficients[,c(1,4)],5)
 
@@
 
-sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM")
-mean.all.other <- mean(sans_high_female$mean_co)
+LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
+mean.LEW.female <- mean(LEW_female$mean_co)
 
@@
 MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSN.female <- mean(MSM_female$mean_co)
-
-sans_MSM_female <- female_mouse.avs_for.analysis %>% filter(strain != "MSM")
-mean.all.other <- mean(sans_MSM_female$mean_co)
-
-mean.MSN.female / mean.all.other
+mean.MSM.female <- mean(MSM_female$mean_co)
+
+sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
+mean.all.other <- mean(sans_high_female$mean_co)
+
+#sans_MSM_female / mean.all.other
 
@@
 
-The range of strain means is `r min()`
-
-the ranges of mouse means for each strain is X
-
-MSM, G, LEW have significant strain effects.
+Above are the coefficents for the two glm's of the female specific mouse MLH1 averages, which include subspecies and strain as fixed effects.
+G has the most consistant significant strain effects in both models.
+MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.
+
+G is `r mean.G.female / mean.all.other` higher the the other means.
+LEW is `r mean.LEW.female / mean.all.other` higher  and MSM is `r mean.MSM.female / mean.all.other`.
+These three will be designated as 'moderate high rec' strains for later sex specific analysis.
 
@@
 
-Follow up on this finding -- included more strains which were excluded from the HetC analysis.
-
-> Is there a strain effect in Dom?
-
-> How different are the Musc strains?
-
-test a model for fixed effects of strains within musc. The purpose is to distinguish musc strains from low 'Dom' level strains.
+The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
 
@@
 
-(Try removing the alpha legend)
-
-This is the table and plot of new Musc strains
-
-`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`
-
-Given the plot, make note of the potential outlier mice.  The strains have unequal variance
-
-KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2
-
-PWD: 24sep14_PWD_m1, 18may15_PWD_m1?
+- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
+
+- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
+
+- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
 
@@
 
+Two models were fitted to the male specific data to test the effects of strain and subspecies.
+
 ```{r Male.specific.glm, echo=FALSE}
@@
 #1. determine which strain effects are sig
+
 #M2
@@
 lm.male.strain_sum<- summary(lm.male.strain)
+
+round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
 
@@
 #mean(SKIVE_male$mean_co) /
-  mean(sans_LewRec_male$mean_co)
 
@@
 
-the ranges of means are X
-
-PWD, MSM and SKIVE have strain significant strain effects so they will be designedated into the high rec group.
+The strain average for PWD, MSM and SKIVE are  ` r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
+Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
 
@@
 The subspecies effect - is a proxy for measuring the degree of divergence across subspecies.
-The strains were coded as a random effect to -simulate / approximate / the standing genetic variation across subspecies.
+The strains were coded as a random effect to approximate standing genetic variation across subspecies ranges.
 
@@
 #kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+
+all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
+                 .(sex, strain),
+                 summarize,
+                 all.cells = sum(Ncells)
+)
+
 ```
@@
 
-`r kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
-
-1. $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
+
+add number of cells
+
+`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
+
+ $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
 
@@
 MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
 #summary(MLH1.MM.M1)
@@
 MM.Fixed.pval <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+
+anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)
+anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)
+
 MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
@@
 
-2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
+ $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
 
@@
 summ <- summary(glm.strain.fixed.full)
+#table of just coeff anf pvalue
+round(summ$coefficients[,c(1,4)],5 )
 
@@
 
-Strain G is significant (p=0.0005),
+Strain G is significant (p=0.0005), (remember female result)
 
@@
 
-3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$
-
-Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.
-
-$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$
-
 $$mouse \ av \ metric ~=~ sex * strain + \varepsilon$$
@@
 
-#below table is broken
-as.data.frame( round( summary(strain.fixed)$coefficients, 6))
+summ.strain.fixed <- summary(strain.fixed)
+
+round(summ.strain.fixed$coefficients[,c(1,4)],5)
+
+#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+
+#as.data.frame( round( summary(strain.fixed)$coefficients, 6))
 
@@
 
-## Within Mouse Variance in CO Count per Cell
+## Summary of Mixed Model Analyses
+
+**mouse average MLH1 count**
+
+Multiple models were tested to test for potential evo models -- for the mouse average CO counts.  (there were a mix of results -- and nuanced)
+
+First full Mixed model -- no significant fixed factors --(the random effect might be significant).
+
+The second model were strain was (nested and random), the fixed effects, sex and interaction of (subsp and sex), are much more significant.
+The coefficients indicate, males in general have 1 less in average, and the musc and molf subsp have ~3 more on average.
+
+**Variance effect**
+
+** Do these patterns hold across Quality bins? **
+
+# Within Mouse Variance in CO Count per Cell
 
@@
 
-**todo**
-
-Test if dissection date is more significant predictor than strain (for females) (Effect doc)
-Also test how the number of 0CO bivalents per cell affects mouse with variance. (lm)
-
 ```{r cv.plots, echo=FALSE, eval=FALSE}
@@
 
-> Modeling Within Mouse Variance
-
-(use the same models are mean_co, )
-
-XXX (old 1.)	HetC.MixedModel.HQ <- lme(mouse.metric ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )
-
-1.(new 1)	Reduced.strain.HQ <- lmer(mouse.metric ~ subsp * sex+ (1|strain), data=DF.HetC.MixedModel.HQ)
-
-3.	lm(mouse.metric ~ subsp \* sex \* strain, data=DF.HetC.MixedModel.HQ)
-
-4.	lm(mouse.metric ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5}
+
+#variance plots
+musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot
+
+Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.female.VAR.plot
+
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.female.VAR.plot
+
+```
+
+Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
+
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5}
+
+#variance plots
+musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.male.VAR.plot
+
+Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.male.VAR.plot
+
+Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.male.VAR.plot
+
+```
+
+from the Mixed models, there should be 4, 1-2 models from the full dataset for var and cv, and 3-4 from a higher quality data set for var and cv. Only a subset of the model outputs will be shown.
+
+For the mixed models the indiviual fixed effects are tested by a anova of the full and reduced models. The interaction fixed effect pvalue comes from drop1() a LTR test.
 
@@
 
-```{r M1.MLH1.var_cv, echo=FALSE, include=FALSE}
+```{r M1.MLH1.var_cv, echo=FALSE}
 #modeling within mouse variance
@@
 MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
 #summary(MLH1var_M1)
-
-drop1(MLH1var_M1, test="Chisq") #interaction term NS
+#drop1(MLH1var_M1, test="Chisq") #interaction term NS
 #when switched to additive fixed, sex is very sig
 
-anova(MLH1var_M1)#sex explains most variance...
-
-exactRLRT(MLH1var_M1) #strain
-as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))
+#anova(MLH1var_M1, MLH1var_M1.reduce.sex)
+#anova(MLH1var_M1, MLH1var_M1.reduce.subp)
+
+#exactRLRT(MLH1var_M1) #strain
+#as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))
 
 #CV
-MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-#summary(MLH1cv_M1)
-
-drop1(MLH1cv_M1, test="Chisq") #interaction term NS
-#when switched to additive fixed, sex is very sig
-anova(MLH1cv_M1)  #sex F value is large
-exactRLRT(MLH1cv_M1) #strain
-
-as.data.frame( round(summary(MLH1cv_M1)$coefficients, 6))
+MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#drop1(MLH1cV_M1, test="Chisq") #interaction term NS #when switched to additive fixed, sex is very sig
+#anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)
+#anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)
+#exactRLRT(MLH1cV_M1) #strain
 
@@
 MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-
-drop1(MLH1var_M1_Q12, test="Chisq")#sig
-anova(MLH1var_M1_Q12)
-exactRLRT(MLH1var_M1_Q12)#NS
-
-as.data.frame( round(summary(MLH1var_M1_Q12)$coefficients, 6))
+MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+#drop1(MLH1var_M1_Q12, test="Chisq")#sig
+
+#anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)
+#anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)
+#exactRLRT(MLH1var_M1_Q12)#NS
 
@@
 MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-
-drop1(MLH1cv_M1_Q12, test="Chisq")#NS
-anova(MLH1cv_M1_Q12)#sex large F
-exactRLRT(MLH1cv_M1_Q12)#NS
-
-as.data.frame( round(summary(MLH1cv_M1_Q12)$coefficients, 6))
-
-```
-
-The fixed sex effect has the largest effect sizes.
-All are negative (for male).
-(I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
-The strain effects are only slightly significant for within mouse variance measures in the full data set (includes low quality cells)
-
-$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE}
+MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+#drop1(MLH1cv_M1_Q12, test="Chisq")#NS
+
+pval.sig <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)
+CV.anova.p <- pval.sig$`Pr(>Chisq)`[2]
+
+# <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)
+
+#exactRLRT(MLH1cv_M1_Q12)#NS
+
+#make a table of the Mixed model pvalues, which are caluclated for subp, sex, sex:subsp and strain across 4 models ()
+```
+
+**General pattern is that sex has a very significant effect.** ( on the order of `r round(CV.anova.p, 6) ` )  this applies to different measures of within animal (across cell) variance and also to subsetted data with higher quality cells.
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
 
@@
 
+<!-- insert the sex specific variance effect here / maybe also the plots of variance -->
+
 For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
-Most of the strain specific differences in within mouse variance go away in the Q12 data set.
+Most of the strain specific differences in within mouse variance go away in the Q12 data set.  (These strain effects could indicate that the distribution of across cell variance in CO number evolves.)
+
+The second set of models analyzed have this form.
 
@@
 
-```{r M3.MLH1.var, echo=FALSE, include=FALSE}
+```{r M3.MLH1.var, echo=FALSE}
 
@@
 
-#summary(MLH1var_M4)
-
-drop1(MLH1var_M4, test="Chisq")
-anova(MLH1var_M4)
-as.data.frame( round(summary(MLH1var_M4)$coefficients, 6))
+#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+#as.data.frame( round(summary(MLH1var_M4)$coefficients[,c(1,4)], 6))
 
@@
 #summary(MLH1cv_M4)
-
-anova(MLH1cv_M4)  
-as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))
+#as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))
 
@@
 MLH1var_M4_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-
 #summary(MLH1var_M4_Q12)
-
-anova(MLH1var_M4_Q12)
-as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))
+#as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))
+#as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients[,c(1,4)], 6))
 
@@
 
-anova(MLH1cv_M4_Q12)
-
-as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
-
-```
-
-#### Summary witin mouse Variance
-
-The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.
+#anova(MLH1cv_M4_Q12)
+
+#as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
+
+```
+
+The glm's support the mixed models with the sex effect being the consistant significant effect
+
+#### Summary for witin mouse Variance
+
+The general pattern across all the models, is that sex is a significant effect for all models.
+**General pattern is that sex has a very significant effect.** The sex-subsp interaction effect p values were larger and only significant for one of the models.
+
+The models run over higher quality cells, suggest that staining or technical error in female data is unlikely to be the only cause for increased variance of number of MLH1 foci per oocyte.
 
@@
 
+# DMC1 Results, Spermatocytes
+
+These data are meant to test if CO precursors, double strand breaks (DSBs), are significant predictors for MLH1 count variation. Simply, do mice with more DSBs also have more COs?
+
+Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days).
+One mouse represents each strain.
+The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize,
+
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
+
+#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
+#            row.names = FALSE)
+
+DMC1.mean.table
+
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+
+```
+
+```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#these are histogram plots for testing Normality shape -- move to supplemenatry section
+
+#- check that distributions are close to normal
+#- check for equal variance.
+
+DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
+
+DMC1.histogram
+#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
+#bartlet's test, Levine's test
+
+#The counts look close to normal.
+#Potentially figure 2. Using a function to #display the p-vales for mean comparison
+
+```
+
+```{r DMC1.plot, echo=FALSE, warning=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+#change order
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
+                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.boxplot
+```
+
+```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
+
+anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#
+
+DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
+
+  #facet_wrap(~stage, nrow=1) +
+
+  ggsignif::geom_signif(
+    data=anno_df_stage, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.bxlt_Z_to_L
+
+#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
+
+ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
+               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
+
+DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
+
+ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
+               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
+
+DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
+
+ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
+               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
+```
+
+the pvalues for the differences between time points are  `r ttest.ZtoL$p.value`,  `r ttest.ZtoL.high$p.value` , and  `r ttest.ZtoL.Low$p.value` for all observations, the high rec group and the low group respectively.
+
+```{r plot.t.test, echo=FALSE}
+
+  #this will add neat bars, but still need to figure out what groups ect it's actually using
+#THE stat_compare function adds a arm thing and pvalue for 
+#  stat_compare_means(comparisons = my_comparisons)
+
+#status - legend, expression y-axis, facet, facet_wrap SYMBOL
+#legend - stage. y=total count, FACet strain
+
+#add column to distinguish high vs low rec strains
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+
+DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
+
+# annotation table with adjusted pvals and y-position of the labels
+
+anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#
+
+DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=stage), position=position_jitterdodge()) + 
+  facet_wrap(~stage, nrow=1) + 
+  ggsignif::geom_signif(
+    data=anno_df, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.bxlt_High.to.Low
+
+#example for ggplot+ test from here
+#https://github.com/kassambara/ggpubr/issues/65
+
+#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
+
+```
+
+The category difference is only observed in the earlier Leptotene stage (p values report a t.test)
+
 <!--
-For later analysis.  Not sure what these are good for.
-
-```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
-
-# check if the main mixed model results hold up when all cells are pooled
-
-#pooled.cells
-DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
-
-#remove PERC, MOLF, CZECH,
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
-
-#sex only has 2 levels and this breaks...
-model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.pooled)# indeces for this object: [1]-structure
-ranef(model.pooled)
-
-model.pooled$coefficients
-
-#pooled cells,  DF.HetC.MixedModel.pooled
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
-
-#consider changing to ordered = FALSE, for the DFs used for MM
-DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
-
-DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
-
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-
-```
-
-- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
-This means there is 0 variance added to mean_CO to mean due to being MSM.
-Does this mean the model is very good at predicting MSM mean_CO.
-The fixed strain - sex effects are for musc are large.
-This will likely change with MOLF (male molf are low, so smaller fixed effect and )
-
-- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
-In the first model MSM has largest random effect, since it's so far from WSB female.
-
--adding MOLF also doesn't have significant factors
-
-changed all to unordered factors.
-all the variables are coded as ordered factors.
-When the variables are kept factors, and unordered the names appear! yay!
-
-Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
+Subsampling code -- don't report
+
+The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
+Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
+Permutations were done for both cell stages.
+The actual p value is in red.
+
+The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
+
+```{r DMC1.subsampling, echo=FALSE}
+
+#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
+# the 2 groups
+
+#make the 2 pools
+High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+
+Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+
+#permute drawing / subsampling and performing the t-test
+#put the p.value results into a DF!
+
+#permut.SC.cv.diff  <- as.data.frame(
+#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
+
+#permuting samples -- and re-licating t-tests
+
+#nrep <- 1000
+
+#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
+bunchOpvals.L <- replicate(1000, 
+    # t.test
+    t.test( sample(High.L.pool, 25, replace = FALSE),
+            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
+          )
+bunchOpvals.L <- as.data.frame(bunchOpvals.L)
+bunchOpvals.L$stage <- "L"
+
+bunchOpvals.Z <- replicate(1000, 
+    # t.test
+    t.test( sample(High.Z.pool, 25, replace = FALSE),
+            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
+          )
+
+bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
+bunchOpvals.Z$stage <- "Z"
+
+#could merge the lists of p-values
+replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
+#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
+
+#ASSUMPTIONS; normality, equal variance
+
+#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
+
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
+#plot.pvals.L
+#plot.pvals.Z
+
+#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
+#n use the var.equal = TRUE
+#https://www.statmethods.net/stats/ttest.html
+```
+
 -->
 
-## Summary of Mixed Model Analyses
-
-**mouse average MLH1 count**
-
-Multiple models were tested to test for potential evo models -- for the mouse average CO counts.  (there were a mix of results -- and nuanced)
-
-First full Mixed model -- no significant fixed factors --(the random effect might be significant).
-
-The second model were strain was (nested and random), the fixed effects, sex and interaction of (subsp and sex), are much more significant.
-The coefficients indicate, males in general have 1 less in average, and the musc and molf subsp have ~3 more on average.
-
-**Variance effect**
-
-** Do these patterns hold across Quality bins? **
-
-<!--  come back to Fig1
-# Figure 1
-
-work more on figure 1. Main figure, MLH1 strain avs on top, chrm proportion bar plot on bottom
-
-```{r first.boxplots, echo=FALSE, fig.align='center', warning=FALSE}
-
-#the below ordering works!! keep it
-MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
-              "PWD","CZECH","SKIVE",
-              "KAZ","TOM","AST",
-              "MSM","MOLF",
-              "CAST","HMI","SPRET","SPIC", "other"), order=T )
-#order the data frame
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
-  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
-#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+
-
-justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")
-
-#scales="free_y"
-#this is ok except for the set placement of mice on x axis
-#if the full dataset was ordered correctly, this would kinda fix it
-
-#boxplots by strain
-ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+
-  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )+scale_fill_manual(values=colors_of_strains)
-
-ss <- ss   + facet_wrap(~ sex, scales="free")
-
-#these plot currently look bad
-```
-
-```{r table.for.plot, echo=FALSE}
-#why do I add in a different table?
-
-#this move this to the setup file
-
-strain.table.HQ_Fig <- ddply(MLH1_data, c("strain", "sex"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  =length(nMLH1.foci),
-                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(nMLH1.foci, na.rm = TRUE),
-                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
-                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
-                       #Standard error (mice vs cells)
-                        se  = round(sd / sqrt(Ncells), 3)
-    )
-
-```
-
-```{r Fig1, echo=FALSE, warning=FALSE}
-
-#might need to nudge the X values
-
-Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
-#remove y labfor male
-
-   #segments
-  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
-  #Musc
-  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
-  #mol
-  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
-  #cast
-  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
-  #outgroups
-  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
-  #label
-  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
-           y = c(18.5,18.5,33,18.5, 18.5),
-
-   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
-           fontface = 'italic', hjust = 0.5, size=3.5)
-
-Fig1_v1_strain
-```
-
-What are the steps the each chunks walk through.
+```{r DMC1_MLH1.cor, echo=FALSE}
+#make category averages from AP_mouse_table.HQ
+mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+#attach this mini DF to the DMC1 data
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+```
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova <- anova(mod.L) #0.0002701 ***
+z.anova <- anova(mod.Z) #NS
+
+l.anova
+z.anova
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+
+```{r barlett.test, echo=FALSE, eval=FALSE}
+
+#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
+bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
+
+```
+
+Since there is evidence for non-equal variance across strains for zygotene cells, so don't rely on lm()s that estimate the effect of strains.
+
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
+
+<!--
+ Single Bivalent Patterns
+
+## Total SC Results
+
+insert male total SC results here
+
+# Female patterns
+
+Nuanced
+there is strain specific evolution for females ...
+
+# Sexual dimorphism and heterochiasmy
+
+total SC length / SC compaction and telomere bias in positioning are conserved / consistently sexually dimorphism (in these mice?)  for most species
+
+make the heterochiasmy plot, with female strain averages adjusted for X
+(measure variance in CO counts for the top 5 female whole cell bivalents)
+(variance in )
+how does this compare to variance of mid 5, bottom 5?
+
+# Predictions / tests for selection on Bivalent structure
 
@@
 
-# Proportion of Chromosome Classes
-
-In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes with different numbers of COs)
-
 <!--
-TODO -- try running a chi square test on the proportions.
-
-#2 sample prop test
-prop.test(x, y)
-
-x - list of the number of 'successes' across groups  (1COs)
-y - list of the number of 'failures' across groups  (2COs)
-
-(does a chi square test)
-
-make a smaller proportion chrm block --
-Add p.values of tests for the how the proportions differ.
-*cast female data came from another publication, add this ref (Kohler?)
-The bottom lines and subspecies points don't exactly match up.
-**add in cell images like the other draft figures.**
+# Rapid male specific evolution
+
+list out the points which support rapid male specific evolution for MLH1 rates.
+
+number above bars are the raw counts for bivalent classes.
+
+Potential figure 2 (eventually). Try to make the figure taller.
+-- add more data to make sample sizes more equal.
+-- add space to indicate where female data is missing
+
+SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
+What's up with Lew females? there's lots of 2COs, also WSB female.
+
+This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
+If these data are random samples of bivalents, the X
+
+ToDo: chisq tests for difference in proportions, 
+- within strain across sex,
+- males, across high and low groups
+- look at mouse level proportions
 -->
 
-These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes.
-Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs. (the gwRR for most mice is low/close to the minimum)
-
-The data below are from hand measured cells (not from bivalent data)
-generate smaller data set of chrm proportions (0CO, 1CO,2CO etc)
-10 cells from each musc category. **Add Molf female**
+```{r misc.setup, echo=FALSE}
+
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blues
+
+                       'PWD'= 'red2',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
+                       'CZECH'= 'indianred4',
+
+                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
+
+                    'CAST'='green','HMI'= 'forestgreen', #greens
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
+
+#add Rec group,
+#reclassified SKIVE as intermediate
+Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
+#apply this before ploting, to display strains in desired order
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, category, mouse) %>%
+#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+# sort your dataframe, by the focal categories
+#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
+```
+
+<!-- merging the 2 data sets-->
+
+# Chromosome Class Proportions
+
+In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes with different numbers of COs.
+The two plots show the chromosome class proportions from hand measured and the curated BivData.
 
@@
 
-Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),]
-
-female.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(sex == "female")
-male.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(sex == "male")
-
-n1CO_list_F <- c( female.mouse.av$n1CO)
-n1CO_list_M <- c( male.mouse.av$n1CO)
-
-total_biv_F_list <- c( female.mouse.av$tot.biv)
-total_biv_M_list <- c( male.mouse.av$tot.biv)
-
-nbiv_total_list <- c(Mouse.Table_BivData_4MM$tot.biv)
-
-prop.test(n1CO_list_F, total_biv_F_list )
+#mouse.level.BivData_table <-- has more male strains
+#Mouse.Table_BivData_4MM <- only sex matched strains
 
 #perform test across strain 
-
-Chrm.prop_strain_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
+#removed all lines with no hand.foci.count entry
+Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
 
@@
                  n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO )
-                          )
+                  n3CO = sum(n3CO ),
+                 n4CO = sum(n4CO )
+                          )
+
+Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
+
+for(g in 1:length(Chrm.prop_strain_table$nbiv)){
+
+  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
+                                     Chrm.prop_strain_table$n1CO[g],
+                                     Chrm.prop_strain_table$n2CO[g],
+                                     Chrm.prop_strain_table$n3CO[g] ))  
+}
+
+Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
+#there are still some wacky entries -- but much fewer than before
 
@@
 
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$nbiv )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$nbiv )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$nbiv )
+female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
+female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
+female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
+#the pvalues for 0 are not sig dif
+#the p values for 1 and 2CO are sig diff
+
+#seperate by subsp
+chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
+chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
+chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
+
+Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
+Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
+Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
+Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
+#WSB has fewer 2COs than other 3 strains
+
+Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
+Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
+Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
+#KAZ has slightly more 1CO than the other two
+
+#high.female_G.LEW.
+chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
+
+high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
+high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
+high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
+
+#Low.female.G.LEW.
+chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
+
+low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
+low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
+low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
 
 chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-#chrm.prop.strain.male.high.rec <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-#chrm.prop.strain.male.low.rec <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$nbiv )
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$nbiv )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$nbiv )
+male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
+male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
+male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
+#the pvals for these 3 class are sig dif
+
+#seperate by subsp
+chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
+chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
+chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
+
+Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
+Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
+Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
+#non of the Dom proportions are different
+
+Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
+Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
+Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
+#all v different
+
+Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
+Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
+Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
+
+#high vs low
+chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
+                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
+                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
+
+chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
+chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
+
+chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
+chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
+
+Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
+Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
+Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
+#the low strains do vary slightly in their 1CO and 2CO proportions
+
+Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
+Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
+Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
+#all very sig p values
+
+High2_male.0CO.prop.test <- prop.test(chrm.prop.male.high2$n0CO, chrm.prop.male.high2$nbiv )
+High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO, chrm.prop.male.high2$nbiv )
+High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO, chrm.prop.male.high2$nbiv )
+#these proportions are v different
+
+High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
+High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
+High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
+#MSM and PWD only slightly sig dif 2CO
+
+#Need to run the chrm proportions thru melt code.
+
+chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
+chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
+
+#chrm.class.DF$biv.sum.check
+```
+
+These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes. Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs.
+
+Most all the the p values for the proportion tests are significant, indicating there are slight but significant shifts across the classes of chromosomes.
+However the most striking male pattern is the propotion of 2COs
+
+```{r Chrm.props.plot2, echo=FALSE}
+
+#barplots for proportions of chrm classes
+
+#consider adding this the the setup block (could have done this to category level table)
+mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
+
+#DF needs to be melted -- so they all in 1 col
+#per code, I've done this by hand
+
+#subset DF for ease
+skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
+col.names.list.99 <- colnames(skinny.mouse.level)
+
+#all of the cols to use
+MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
+
+               #leave out the Foci
+                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
+
+  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
+
+MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
+MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
+
+#no filename for add mouse, strain subsp
+MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
+
+#then this DF has to be summarized into categories
+chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
+                             mean.percent.type = mean(value)
+                             #sum is the wrong!
+                             #used average percentage, might consider a different summary
+)
+
+#can't apply function use this
+chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
+             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
+                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
+                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
+
+          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
+              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
+
+                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
+                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
+            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
+                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
+                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
+                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
+                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
+                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
+
+             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
+               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
+
+chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",
+                                                                 "Cast",
+                                                                 "Musc",
+                                                                 "Mol",
+                                                                 "Spretus","Spic",
+                                                                 "other") )
+#ggplot
+
+chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
+chrm.class.barplot
+
+```
+
+(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class.
+The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.
+
+- High female strains, G and Lew have significantly more 2CO bivalents.
+
+- The overall male pattern that's most striking is the gradient of 2CO proportions
+
+**MSM 60% > PWD 50% > SKIVE 30% >  20 - 10% remaining (low) strains**
+
+# Single Bivalent Level Results
+
+Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script.
+These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output).
+The MLH1 data file is also loaded into this file.
+
+The breakdown of single bivalent observations by category
+
+`r data.frame(table(droplevels(Curated_BivData$category) ) )`
+
+And these are the number of bivalent observations with filled hand.foci.count values.
+
+`r data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )`
+
+Validity of comparing bivalent observations
+While the automated software doesn't isolate all bivalents/chromosomes from each cell (on average 17), we assume that the isolation process is not biased.
+Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.
+
+Two Main questions for these analyses.
+
+> Q1. driving questions, which traits are sexually dimorphic?
+
+> Q2. which traits fit male polymorphism predictions?
+
+These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.
+
+The mouse averages for 3 bivalent level metrics will be analyzed across these questions:
+
+- SC lengths (or total.SC)
+
+- Normalized Rec landscape of 1CO
+
+- IFDs of 2CO
+
+The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.
+
+Q1. Model 1, mixed model lmer()  
+$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+Q1. Model 2, glm()
+$$mouse \ av.\  metric ~=~ sex * strain + \varepsilon $$
+
+Q2 Model 3, logistic regression
+
+$$Rec \ group  ~=~ mouse \ av.\  metric  $$
+
+<!--
+I decided to use the mouse averages for most of the bivalent metric tests. some of my logic is below.
+
+**Pro Bivalent level**
+
+- The data are continuous and closer to normal/Gaussian expectations than the MLH1 counts.
+
+- There is within cell variance that is an issue (would affect single bivalent observations).
+
+**Pro Mouse averages**
+
+- Mouse average is a conservative choice, it would summarize general patterns which fits with the paper.
+
+- Simple application of the same models as MLH1 based results.
+
+- Limited work to add to the data set if I can justify that the collection of single bivalents are random and equivalent across mice.
+
+So I'll use the mouse averages for all the metrics is what is used in the mixed models.
+
+-->
+
+```{r Q1.biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}
+
+#this block should be in the data set up
+
+#the mouse averages will be used
+
+mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+mouse.avs_4MM <- add_sex(mouse.avs_4MM)
+mouse.avs_4MM <- add_strain(mouse.avs_4MM)
+mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
+mouse.avs_4MM <- add_category(mouse.avs_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
+mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
+mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
+
+#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
+#code if males are high or low?
+
+#code intp chr
+
+#assign high and low
+#divide musc males into groups
+mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+
+```
+
+In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .
+
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+
+#add this to setup
+
+#remove female, 
+#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
+
+#Curated_BivData_male
+
+male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+male.av_4MM <- add_sex(male.av_4MM)
+male.av_4MM <- add_strain(male.av_4MM)
+male.av_4MM <- add_subsp(male.av_4MM)
+male.av_4MM <- add_category(male.av_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
+male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
+male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
+```
+
+# Q1 Analysis, Predictions for Heterochiasmy Q1
+
+Using the Mixed model framework which tests the effects (and interactions) of subspecies, sex and strain, I will test for evolution of the following traits.
+
+Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
+
+1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of uniform vs telomere/biased CO positioning.
+
+  C) Centromere and telomere distances will be sexually dimorphic.
+
+3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
+
+# Q1 SC Lengths
+
+<!--Brief lit review for SC length and MLH1 expections.
+add XX adjusment / arguments
+-->
+
+We expect female SC lengths to be longer (refs).
+In the plot above the SC length ~ higher hand foci cells.
+Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
+For the 0 class chromosomes most all are around the same size of the 1CO distribution.
+Add in the code for the plot under 2COs males and females
+
+```{r show.SC.length, echo=FALSE, warning=FALSE}
+#for now remove the 
+#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
+#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS
+
+#add boxplots for mouse level means
+#add second x axis, hand.foci.count
+
+#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#nested boxplots
+SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
+
+SC.length_Nested_boxplots
+
+#+ +geom_jitter(aes(alpha=0.2))
+```
+
+> M1. subsp * sex + (strain)
+
+> M2. subsp \* sex \* strain  (slightly redundant model)
+
+> M3. sex * strain
+
+The dependent variables I'll be testing in the model framework are:
+
+- Pooled SC lengths
+
+- 1CO
+
+- 2CO
+
+- 3CO
+
+- Xlong.biv  *Don't do long bivData for testing sex effect, since this would have the XX
+
+- Short.biv dataset *this would be a chrm size effect, without the XX
+
+<!--
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
+
+To estimate random effects, I'm using lmer (not lme where I could do the rand(strain*sex)). So now it's just random(strain).
+
+The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.
+
+remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
+
+#### Pooled SC Lengths
+
+```{r MM_2.SC.Length, echo=FALSE, eval=FALSE}
+
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+#p
+
+```
+
+I used the mixed model with the lme() for the full Curated BivData and long bivdata
+
+-->
+
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+#### Q1.SC_HetC_M1
+
+Insert the results for the MMs using the dependent variable 1CO - use the Mixed model (lmer), for all of the metrics (pooled SC, 1CO 2CO ect)
+
+Below's the code for Mixed model results -- try to organize them into a table or something
+
+```{r 1.SC.Length_M1, echo=FALSE}
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#all pooled SC lengths
+lmer.Q1_SCLength_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.Q1_SCLength_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.Q1_SCLength_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+drop1(lmer.Q1_SCLength_M1, test="Chisq" ) #other options
+
+anova(lmer.Q1_SCLength_M1, lmer.Q1_SCLength_M1_R.sex) # also F 
+anova(lmer.Q1_SCLength_M1, lmer.Q1_SCLength_M1_R.subsp)
+
+exactRLRT(lmer.Q1_SCLength_M1)
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
+#can't get p values from this summary
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+#drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")
+exactRLRT(Q1.lmer_SCLength_1CO_M1)
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+#drop1(SCLength_2CO_MM.M1, test="Chisq")
+exactRLRT(Q1.lmer_SCLength_2CO_M1)
+as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+#Long.Biv
+Q1.lmer_SCLength_long.biv_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
+Q1.lmer_SCLength_long.biv_M1.R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Long.biv_mouse_table_HetC, REML = TRUE) 
+Q1.lmer_SCLength_long.biv_M1.R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Long.biv_mouse_table_HetC, REML = TRUE)
+
+anova(Q1.lmer_SCLength_long.biv_M1, Q1.lmer_SCLength_long.biv_M1.R.sex)
+anova(Q1.lmer_SCLength_long.biv_M1, Q1.lmer_SCLength_long.biv_M1.R.subsp)
+
+drop1(Q1.lmer_SCLength_long.biv_M1, test="Chisq")
+exactRLRT(Q1.lmer_SCLength_long.biv_M1)
+
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
+
+#### Q1. SC_HetC_M2
+
+All fixed and all interactions
+
+```{r Q1.SC.Length_M2, echo=FALSE}
+#feed these 2 DFs into M2
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+#glm(metric ~ strain)  ---mean_SC
+
+Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_SCLength_M2)
+
+#Dom.all
+Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.Dom)
+
+#Dom.1CO
+Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.1CO.Dom)
+
+#Dom.2CO
+Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.2CO.Dom)
+
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.Musc)
+
+#Musc.1CO
+Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.1CO.Musc)
+
+#Musc.2CO
+Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.2CO.Musc)
+
+#Mol.all
+Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.Mol)
+
+#Mol.1CO
+Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.1CO.Mol)
+
+#Mol.2CO
+Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.2CO.Mol)
+```
+
+```{r Q1.SC.M2plot, echo=FALSE}
+
+#Q1 plots for SC Lengths across Musc strains
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#mouse av
+Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
+Q1.SClength_Musc_plot
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+Q1.SClength_Musc.BivData_plot
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+Q1.SClength_Dom.BivData_plot
+
+```
+
+I think the strange thing about the Musc is that Skive females have shorter SC lengths.
+
+#### Q1. SC Length t.test
+
+```{r strain.SClength.t.test, echo=FALSE}
+#t.tests for all the strain specific data frames
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+#run a t.test seperate out by the chrm classes
+#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+#### Strain Comparisons of SC Length
+
+Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
+
+(adding all of these blocks to -- the setup script); strain.Bivalents.DF
+
+#### XX adjustment
+
+Generate inteligible summaries for the permutation / subsampling results.
+
+```{r Q1_XX.adjustment, echo=FALSE}
+
+```
+
+# Q1. IFD
+
+Brief lit review for IFD / interference expectations  
+**Petkov et al 2007,  CO interference underlies sex differences in RR**
+
+"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
+However, the interference distance is the same in terms of bivalent length.
+We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
+at the pachytene stage of meiosis."
+
+Chrm1 genetic map,
+
+(other human and mice refs
+
+Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
+
+DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
+
+))
+
+*Do I believe the physical/SC results?*
+
+<!--
+ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
+
+explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
+-->
+
+3) Interference / IFD will not be sexually dimorphic.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
+
+**I**nter**f**ocal **D**istance, a indication of CO interference.
+
+The raw measures are pretty nosy, this might not be the best way to display these results.
+
+ Each point is a IFD observation. Maybe add mouse level lines for the means.
+
+```{r show.IFD_99, echo=FALSE, warning=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+
+#+theme(legend.position="none")
+```
+
+```{r IFD.old.plots, echo=FALSE, eval=FALSE}
+#long SC Long_biv_mouse.avs_4MM
+#real.long.bivData
+#Only use 2CO bivalents
+real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
+real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
+
+long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+```
+
+Still working on the best way to display the general IFD patterns.
+
+### Mixed Model Tests, Fixed Effects
+
+Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+
+$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+### IFD_HetC_M1
+
+```{r HetC_M1_IFD_many.metrics, echo=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+#anova(lmer.IFD.PER_MM.M1)
+
+#exactRLRT(lmer.IFD.PER_MM.M1)
+#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
+
+#M1_IFD.RAW_2CO
+lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#these 2 below are LR tests
+#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
+#anova(lmer.IFD.RAW_MM.M1) # also F values
+
+#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
+#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+```
+
+Sex is a slightly significant factor for raw and normalized IFD.
+
+The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
+That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
+
+show a raw plot, and norm plot and a scatter plot
+
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
+The tables below are from anova( for the lmer model ).
+Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
+
+### IFD_HetC_M2
+
+```{r IFD_HetC_M2, echo=FALSE}
+
+```
+
+### IFD_HetC_M3
+
+```{r IFD_HetC_M3, echo=FALSE}
+
+```
+
+## Strain Comparisons
+
+Dive deeper into the sex specific pattern for each strain.
+Below are code chunks which show the unusual sex specific results for IFD measures.
+The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
+The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
+
+For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
+
+```{r Comp_Ratios, echo=FALSE}
+
+#calq ratios of belo 30
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
+
+#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
+MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
+
+DF.strain.list <- list(
+  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
+  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
+  MSM.Bivalents.DF, MOLF.Bivalents.DF
+  )
+
+entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
+        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
+          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
+
+i = 1
+for(x in (DF.strain.list) ){
+
+  #print(head())
+
+  #these filters won't apply to the indexed DF
+  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
+  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
+  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
+  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
+  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+  i = i+1
+
+  }
+#entr_DF
+
+entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
+
+```
+
+Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
+
+```{r WSB.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+#WSB.Bivalents.DF
+
+#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
+WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
+
+#plot the differences
+wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_raw
+
+wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
+wsb.ploty_IFD_PER
+
+WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+WSB.IFDnorm_scatter.plot
+#this plot is much messier than the MSM (PWD?)
+
+#try plotting the other IFD
+
+#make a table of the means and n ect
+wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
+
+                       total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test
+wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
+
+```
+
+The range of normalized IFDs overlap closer in males and females in the WSB data.
+
+```{r LEW.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
+
+LEW.Bivalents.DF_IFD <- LEW.Bivalents.DF[(!is.na(LEW.Bivalents.DF$sex)) & (LEW.Bivalents.DF$hand.foci.count == 2),]
+
+LEW.Bivalents.DF_IFD$sex <- droplevels(LEW.Bivalents.DF_IFD$sex)
+
+#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+lew.ploty_IFD_raw <- ggplot( LEW.Bivalents.DF_IFD,  aes(y=IFD1_ABS, x = sex, na.rm = TRUE))+                     geom_jitter( na.rm = TRUE)+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex, na.rm = TRUE))
+lew.ploty_IFD_raw
+#arg, little annoyed that the x NA won't go away
+
+lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
+lew.ploty_IFD_PER
+
+LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  
+  scale_color_manual(values=c('#56B4E9','#E69F00'))+
+  geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+LEW.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
+
+                  #     mean_SC = mean(chromosomeLength),
+                       n_SC = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
+
+```
+
+The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
+
+```{r G.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
+#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+#plot the metric
+#remove the NAs
+
+G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
+
+G.ploty_IFD_raw
+#
+G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.ploty_IFD_PER
+
+G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+G.IFDnorm_scatter.plot
+
+#make a table of the means and n ect
+G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
+
+#calQ percent < 30%
+
+oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
+oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
+
+#the na's are being counted (ughhhhhh)
+#use the file thing instead
+
+G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
+
+LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
+  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
+
+LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
+#0.7062663
+
+#would Chi-sq be a good test to compare the ratios ect?
+#the sum of these doesn't add 
+#G.Bivalents.DF
+
+```
+
+```{r PWD.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
+#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#remove the NAs
+
+PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
+
+PWD.ploty_IFD_raw
+
+PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
+PWD.ploty_IFD_PER
+
+PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
+
+PWD.IFDnorm_scatter.plot
+
+#run a t.test seperate out by the chrm classes
+PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
+
+```
+
+For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
+
+```{r KAZ.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+#KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
+
+#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+
+#remove the NAs
+KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_raw
+
+KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
+KAZ.ploty_IFD_PER
+
+KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
+
+KAZ.IFDnorm_scatter.plot
+
+#run a t.test seperate out by the chrm classes
+KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
+
+```
+
+For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
+
+```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
+#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+
+#remove the NAs
+SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_raw
+
+SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
+SKIVE.ploty_IFD_PER
+
+SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
+
+SKIVE.IFDnorm_scatter.plot
+
+SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
+
+```
+
+In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
+
+```{r MSM.IFD, echo=FALSE, warning=FALSE}
+#make the dataset
+MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
+#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
+                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_raw
+
+MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
+                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
+MSM.ploty_IFD_PER
+
+#another plot
+MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
+
+MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
+
+MSM.IFDnorm_another.plot
+
+#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
+
+#make a table of the means and n ect
+MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#can I plot these stats
+
+#run a t.test seperate out by the chrm classes
+MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
+
+```
+
+The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
+
+```{r MOLF.IFD, echo=FALSE}
+
+#SHUT DOWN THIS BLOCK UNTIL MOLF FEMALE DATA
+
+#make the dataset
+#MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
+
+MOLF.Bivalents.DF <- Curated_BivData %>%  filter(strain == "MOLF")
+
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+
+#plot the differences
+#MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+
+#remove the NAs
+#molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
+
+#molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+  geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
+
+#make a table of the means and n ect
+MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
+
+                  total.biv = length(Obj.ID),
+
+                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
+)
+
+#run a t.test seperate out by the chrm classes
+#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
+The other strains which have more overlap between males and females are the Dom strains and KAZ.
+
+#### IFD 3CO bivalents
+
+Run comparisons for 3CO bivalents.
+
+```{r CO3.comparisons, echo=FALSE}
+
+#take a look at the 3CO things
+
+```
+
+<!--
+ Heterochiasmy Prediction
+
+> Do the IFD measures lack significant sex effect? (as predicted)
+
+There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
+
+> Are there signiticant strain effects?
+
+The random strain effect was tested for raw and normalized IFD, neither are significant.
+-->
+
+# Q1. General CO Positions
+
+Try accounting for chrm size effects. **Also i didn't notice that I'm using density plots and scatter+boxplots across sections**
+
+<!--
+There are a few traits that fall within the CO positions  
+- 1CO normalized position  
+- sis.co.ten (sorta interference)  
+- centromere and telomere distance  
+**Heterochiasmy**
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+  C) Centromere and telomere distances will be sexually dimorphic.
+**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
+-->
+
+```{r CO.positions.figs, echo=FALSE, include=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+```{r Q1.F1.pos.LONG.BIV_scatter, echo=FALSE}
+
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+real.long.bivData_1CO <- real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+Q1.show.nrmF1_long.biv <- ggplot(real.long.bivData_1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
+
+Q1.show.nrmF1_long.biv
+
+Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Nbiv  = length(Obj.ID),
+                       mean.1CO_PER = mean(PER_Foci_1)
+                      # mean_CO = mean()
+                       )
+
+Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
+Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
+
+Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
+
+Q2.shoooo
+
+```
+
+```{r show.long.biv, echo=FALSE}
+#real.long.bivData
+
+```
+
+```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
+
+#make a table for the sex specific values of the normalized F1 positions
+
+#for strain, sex -- Curated_BivData
+#or (Mouse.Table_BivData_4MM)
+#norm.first.foci.pos
+
+Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
+
+                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
+                  n1CO = sum(n1CO )
+
+                          )
+
+```
+
+```{r Q1.Nrm.F1pos_M2, echo=FALSE}
+
+#calq the means for normalized F1
+#M2
+
+Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_nrmF1pos_M2)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+```{r Q1.Nrm.F1pos_M2_long.biv, echo=FALSE}
+
+#table for the 1CO long biv across strain and sexes?
+
+```
+
+Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
+This plot shows the sexual dimorphism in the density plots.
+
+Consider adding annotate_text for the number of observations in each category.
+think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
+
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#MELT_Curated_BivData
+
+#these isn't long.biv data
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+real.long.bivData
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+```
+
+<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
+
+```{r CO1_foci_pos_boxplot, echo=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+##### Mixed models norm 1CO positions
+
+the distribution of SC lengths and sis-coten seems very different across sexes
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+The mixed model data should only come from 1CO bivalent data.
+
+#### Nrm_1COpos_HetC_M1
+
+```{r Q1.MM_CO_pos, echo=FALSE, eval=FALSE}
+#use the 1CO
+#Use MELT DF
+#MELT_Curated_BivData
+
+#isolate the 1COPos.CO.avs
+#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
+#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
+#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
+#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
+
+mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
+
+#<- lme()
+lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+
+as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
+
+#can't make the 
+CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+
+#library(lmerTest)
+#anova(CO_position.MM)
+#anova(CO_position.MM, verbose = TRUE)
+
+#summary(CO_position.MM) #no p values,,
+#these just give t and F stats, no pvalue
+
+#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
+#CO_pos_fix.est
+#this test, the full model is more predictive than an empty model?
+
+#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
+#thing2 <- drop1(SC_length.MM, test="Chisq")
+
+CO_pos_rand <- exactRLRT(CO_position.MM)
+CO_pos_rand
+```
+
+```{r Q1.nrmFoci1_M1, echo=FALSE}
+
+#M1
+
+```
+
+the mouse average foci1 pos is more significant in t.test, but not log regression...
+(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
+#### Nrm_1COpos_HetC_M2
+
+```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+
+#LOGISTIC REGRESSION for FOCI POSITION
+
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}
+
+#t.test for high and low group  -- logistic regression still need to format the high and low group
+
+#mouse av
+t.test_mouse.av.1CO_pos <- t.test(
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv.1CO_pos_t.test <- t.test(
+
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+  )
+
+```
+
+### Siscoten
+
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males.
+
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
+t.test_mouse.av_siscoten <- t.test(
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+#2CO
+
+#3CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+
+drop1(lme_MM_telo.dist, test="Chisq")
+
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+drop1(lmer_MM_telo.dist, test="Chisq") 
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
+
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
+
+> is random strtain effect significant?
+
+The random strain effect seems very significant.
+
+remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
+
+```{r pos_t.test}
+#make point plots / boxplots which show differences in mean positions
+
+#scatter + boxplot for t.tests
+```
+
+Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
+
+```{r F1.by.F2}
+
+#try remaking the plot Megan suggested
+# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
+
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
+#isolate 2COs
+#facet by sex and subsp
+
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
+
+```
+
+```{r CO.pos.ANOVA, eval=FALSE}
+#what is the pattern of variance
+#run analyses for each subsp*sex
+#use non-melt DF
+
+#how is the variance partioned across
+#cell, mouse, strain
+
+female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
+female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
+
+female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
+
+#unorder strain and mouse
+
+female.Dom$mouse <- as.factor(female.Dom$mouse)
+
+female.Dom$strain <- unclass(female.Dom$strain)
+female.Dom$strain <- as.factor(female.Dom$strain)
+
+female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
+female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
+
+#1CO first
+modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
+
+#can't get mouse and strain to have sum of square
+#residual size decreases with per.F1
+#residuals much larger than fileName, mouse and strain no
+
+#model <- lm(breaks ~ wool * tension, 
+#            data = warpbreaks, 
+#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
+
+male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
+male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
+
+male.Dom$mouse <- as.factor(male.Dom$mouse)
+
+male.Dom$strain <- unclass(male.Dom$strain)
+male.Dom$strain <- as.factor(male.Dom$strain)
+
+male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
+male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
+
+male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
+
+male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
+summary(aov(male.modo))
+
+#only file name is registering as effect
+#Review ANOVA frameworks
+#http://www.biostathandbook.com/nestedanova.html
+
+```
+
+## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
+
+The general predictions across the males and subspecies based on th above MLH1 results.
+
+For positive correlation traits/metrics
+
+i) in DOM strains, low to no difference across strains
+
+ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
+
+ii) in Mol, MSM > MOLF
+
+A) **SC lengths** will be longer for high Rec strains.
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
+
+C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
+
+The mouse averages for the other position metrics will be highly influenced by proportions of the 1CO and 2CO bivalents.
+When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
+
+C.2) sis-co-ten metric ... (what about the clustering?)
+
+C.3) telomere and centromere distances ...
+
+M1 glm model for fixed strain effect across male averages.
+
+# Q2 SC
+
+<!--  main DFs for this section
+
+Male.poly.Mouse.Table_BivData_4MM
+
+Bivalent_1o2  (for logistic regression, 1CO and 2COs)
+male.long.biv
+
+-->
+
+remember the 3 predictions for subsp;
+
+i) Dom, no difference
+
+ii) Musc, PWD and SKIVE greater than others
+
+iii) Mol, MSM > MOLF
+
+```{r Q2.male.Nrm1CO_strain, echo=FALSE}
+#isolate all musc
+#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
+```
+
+```{r Q2.SC.ploy.male, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+
+#1. determine which strain effects are sig
+#use mouse biv average
+#Male.poly.Mouse.Table_BivData_4MM
+
+glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm1.SC.male.poly) 
+#SC many strains are different (more than the gwRR)
+#subsps are sig
+
+glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm2.SC.male.poly)
+
+#Dom
+glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(glm_SC.male_Dom)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(glm_SC.male_Musc)
+
+#Mol
+glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(glm_SC.male_Mol)
+
+#Logistic regressions
+
+log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+
+                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
+
+#summary(log.SC.male.poly)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_musc)
+#not sig -- when just looking at Musc
+
+#significant in predicting group
+log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_mol)
+#not sig -- when just looking at Mol
+```
+
+```{r Q2.SC.long.biv, echo=FALSE}
+
+#mouse.av.long.biv.table
+
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
+#summary(glm1.SC.male_long.biv)
+
+glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
+#summary(glm2.SC.male_long.biv)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
+#summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
+#summary(glm_SC.male_Mol_long.biv)
+
+## logistic regression for Rec group
+
+log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
+              data=mouse.av.long.biv.table,  family=binomial(link="logit"))
+
+#summary(log.SC.male_long.SC)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+```
+
+Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
+
+The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).
+
+#### Q2 SC by class
+
+The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
+Mainly that 1CO are shorter in the high rec strains then low rec strains.
+This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
+Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
+This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
+They have a lower probability of 2COs below a certain SC length threshold.
+
+I think these comparisons should be done with the **Bivalent observations**
+The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class
+
+```{r Q2.SC.Chrm.Class, echo=FALSE}
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+#there's no rec group for this table
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
+#there might not be enough data 
+#
+ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
+```
+
+These are the plots of the logistic regression.
+
+```{r Q2.SC.log.regression.show.plots, echo=FALSE}
+
+plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.blah.mouse.av
+
+plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.logReg.mouse.av_1CO
+
+plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+plot.logReg.mouse.av_2CO
+
+```
+
+the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
+
+```{r Q2.SClength.t.test, echo=FALSE}
+#MOUSE AVE, all
+Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
+
+#mouse.av 1CO
+Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
+
+#MOUSE.av 2CO
+Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
+
+#single biv level
+#male.biv?
+
+#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value
+
+```
+
+The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths.
+But When the means are compared across chrm class, the mouse averages are no longer significant.
+
+### Q2 SC length Chrm class prediction
+
+How well does SC length predict chromosome class?
+Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class.
+(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+<!--
+blocks of other male Q2 - SC lengths
+
+what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
+
+```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
+
+#comparing all the strains
+
+T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+)#p-value = 0.001567
+
+#1CO
+T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#2CO
+T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#3CO
+T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
+
+  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
+
+  )
+
+#Single Biv level
+T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
+
+#1CO
+T.test_SC.Len_single.biv_1CO <- t.test(
+  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
+
+      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
+  )
+
+#2CO
+T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
+
+                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
+  )
+
+  #3CO
+  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
+
+  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
+  )
+```
+
+The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
+
+```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
+#I don't think the Mixed models fit well for the high vs low polymorphism tests.
+
+#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
+#MSM_SC vs MOLF_SC
+
+#2. logistic regression for sc length and chrm class?
+#if the prediction / significance
+
+#isolate data
+#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
+
+#build and test MM
+lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+
+lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
+
+anova(lmer_MM_SC_length_male.poly)
+
+rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
+
+```
+
+```{r long.sc_log.res_meanSC, eval=FALSE}
+#Long Biv comparison
+
+#hold off on reporting/ working with the long SC biv resutls
+
+long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
+long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
+
+#divide musc males into groups
+long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
+                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
+
+long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
+                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
+
+summary(long.biv.log.reg)#NS effect (likelt underpowered)
+```
+
+-->
+
+> **Summary**
+
+When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
+I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
+(IE the scatter plots of SC length, binned by chrm class on the X).
+
+Each point is a mouse average SC.
+When all bivalents are pooled, the logistic regression SC lengths predict the groups.
+When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
+For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
+
+ and the mean 2COs SC are higher in the high strains
+(this must have something to do with the physical size effect)
+
+```{r male.show.plotsEXXXample, echo=FALSE, eval=FALSE}
+
+#REWRITE THIS FOR SC LENGTH COMPARISONS
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
+fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
+fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
+fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
+
+table( droplevels(fomo.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.male.plot
+
+#Dom version 
+Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.male.plot
+```
+
+```{r SC.length.male_tests, echo=FALSE}
+
+## run the tests to check following expectations
+
+#1.Dom strains not different from ech other
+#2. PWD and SKIVE are different from the other strains
+#3.MSM and MOLF are different from each other
+
+#Male.poly.Mouse.Table_BivData_4MM
+#Long.biv_mouse_table_male.poly
+
+#decide between lm or glm..?
+#mouse avs, males
+
+#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
+
+male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
+#the subsp labels get messed up, unless they are designated chr first
+Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
+
+Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
+Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
+#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
+#total sc sig, 1CO sig, 2CO sig
+
+Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
+#0.000369 ***
+Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
+#0.000816 ***
+Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
+#0.0148 *
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
+#making plots above show that mouse average SC lengths are lower could there be microscope issue?
+
+#make sure to remove some of the juvi mice!! from the AP_mouse table
+show.Dom.male.CO <- ggplot(
+male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
+
+Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
+#P not sig for CO mean
+
+#seems like there are more mice in the MLH1 data set than bivData
+#
+
+#WSB all dissected and imaged in '15
+#G imaged 15 - 16
+#LEW imaged 15 to 17
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "1CO SC Length", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "2CO SC Length", xlab = "")
+```
+
+```{r SC.length.pairwise.T, echo=FALSE}
+#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
+
+male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
+male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
+male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
+
+Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
+#PWD and MSM are most diff from SKIVE?
+
+Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
+#the pooled SC value tests are sig
+
+#the mol comparisons are different due to adj p values
+
+Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
+#no sig (only 3) -- does include CZECCH
+
+#non-sig, with mol pooled with musc
+
+Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
+
+Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
+
+Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
+#non of these sig with the mulitple adjust
+
+Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
+
+#pairwise.t.test(Ozone, Month, p.adj = "bonf")
+
+```
+
+If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
+
+Ideas for the above 3 tests,
+
+1. no difference in DOM
+- no sig logistic regression test
+- glm, strain effect?
+
+> Is sex a significant effect for SC length? (as predicted)
+
+- Sex is a significant effect for SC length
+The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
+
+- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
+
+- The Long Biv Data set largely agrees with the full curated dataset
+
+- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
+
+```{r log.reg_SC.length, echo=FALSE}
+#male_mouse.av
+#Male.poly.Mouse.Table_BivData_4MM
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+#male poly shouldn't be here
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
+                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
+                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
+
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
+
+#1CO
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_1CO)
+
+#2CO
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_mouse.av_2CO)
+
+#3CO
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_3CO)
+
+# bivalent level log regression
+
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV)
+
+#1CO  (BIVDATA)
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_1CO)
+
+#2CO
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_2CO)
+
+#3CO
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV_3CO)
+```
+
+The mean SC logistic regression model for mouse averages
+
+`r log.reg_SC.length_mouse.av$coefficients`
+
+and for the single bivalent levels
+
+`r log.reg_SC.length_BIV$coefficients`
+
+When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
+When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
+
+> Is the prediction, high rec musc male strains have long SC met?
+
+In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
+I couldn't get the Mixed models working for the male polymorphism predictions...
+
+# Q2 Normalized CO positions
+
+A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.
+
+or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.
+
+<!-- 
+Main DFs for this section:
+Male.poly.Mouse.Table_BivData_4MM
+male.bivdata.1CO
+
+**Polymorphism**
+C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
+C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.  
+C.3) telomere and centromere distances will be shorter in low rec strains
+
+-->
+
+Since there aren't many good predictions for how the 1CO normalized landscape relates to gwRR variation, these will be the main questions
+
+- Has there been evolution in the 1CO pattern?
+
+- Is there a consistent pattern with the high rec strains?
+
+Normalized 1CO foci positions will be used. (F1_PER). Check fixed effects of subsp and strains.
+
+sis-co-ten, telomere and centromere distance of foci are metrics which draw from a wider pool of samples. (So maybe I will use them... but avoiding focus on them)
+
+```{r Q2.show.nrmF1, echo=FALSE}
+#Curated_BivData_male
+
+#boxplots of nrm F1
+male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
+
+Q2.show.nrmF1
+
+nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
+
+nrmF1_mouse_level
+```
+
+Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.
+
+```{r Q2.F1.pos.LONG.BIV, echo=FALSE}
+
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
+
+Q2.show.nrmF1_long.biv
+
+Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Nbiv  = length(Obj.ID),
+                       mean.1CO_PER = mean(PER_Foci_1)
+                      # mean_CO = mean()
+                       )
+
+Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
+Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
+
+Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
+
+Q2.shoooo
+
+```
+
+So the plot for the long bivs, shows that the PWD, SKIVE and MSM have on average more centrally placed F1 positions.
+It's kinda werid that the glms/models suggest that WSB has the most terminal F1 -- but, out of all the strains, for the long bivs, WSB has the lowest mean.
+
+But there are v few observations. I think the plot shows that there are fewer 1CO for the High Rec group. not sure there would be enough for feeding into models/
+
+```{r Q2.1CO.landscape, echo=FALSE}
+#Male.poly.Mouse.Table_BivData_4MM
+#Curated_BivData_male
+
+#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm1.SC.male.poly) 
+#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm2.SC.male.poly)
+
+#M1
+glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm.M1_Q2.1COnrm)
+
+#M2
+glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm.M2_Q2.1COnrm)
+
+#Dom
+glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(glmQ2_M2.1COnrm_Dom)
+
+#Musc
+glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(glmQ2_M2.1COnrm_Musc)
+
+#Mol
+glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(glmQ2_M2.1COnrm_Mol)
+
+#m3 (log)
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm)
+
+#Musc
+Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
+              family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm_Musc)
+
+#Mol
+Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
+              family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm_Mol)
+
+#biv level logistic regression
+#Curated_BivData_male
+#male.bivdata.1CO
+
+male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
+
+Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO,  family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm)
+
+#Musc
+Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
+              family=binomial(link="logit"))
+#summary(Log.Q2.F1_PER_Musc)
+
+#Mol
+Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
+              family=binomial(link="logit"))
+
+#summary(Log.glmQ2_M2.1COnrm_Mol)
+
+```
+
+```{r Q2.t.test_COlandscape,echo=FALSE}
+
+#MOUSE AVE
+Rec.group.mouse.F1pos_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.F1pos_Ttest$p.value
+#p-value = 0.9098
+
+#single biv level
+Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+Rec.group_BivData_F1post_Ttest$p.value 
+#p-value = 0.238
+
+```
+
+The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
+
+#### Summary
+
+Not really a clear summary for the normalized 1CO position and rec groups.
+
+# Q2 IFD
+
+this is the Q2 prediction for IFD / interference
+
+> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
+
+B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
+
+1. glm across subsp
+
+2. anova within subsp
+
+3. logistic regression for musc (or all males)
+
+```{r Q2.show.IFD, echo=FALSE}
+
+#subset
+male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
+
+biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
+
+biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+biv.IFDs_PER_by.group
+```
+
+The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
+ The normalized plot displays the pattern of higher rec group having slightly higher longer IFDs.
+
+```{r Q2.IFD.glm.test, echo=FALSE}
+#glm --
+#mouse aver
+
+#M1.ABS
+glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1.Q2_IFD.raw) 
+#M1.PER
+glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1.Q2_IFD.PER)
+
+#M2.raw
+glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M2_Q2.1COnrm)
+
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+summary(Q2.glm.M2_IFD_Mol)
+
+#M2.PER
+glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M2_Q2.IFD.PER)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+summary(Q2.glm.M2_IFD.PER_Mol)
+```
+
+```{r Q2IFD_logreg, echo=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.PER, echo=FALSE}
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
+
+#t.test for IFD
+#mouse.av - the mouse average IFD is from just 2CO
+t.test_mouse.av.IFD_ABS <- t.test(
+  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+t.test_mouse.av.IFD_PER <- t.test(
+
+  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+```
+
+```{r Q2.t.test_IFD,echo=FALSE}
+
+#already made
+#subset
+male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+#t.tests for the Rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#THIS SHOULD ALREADY BE ADDED
+
+#single biv level
+Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
+                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
+
+Rec.group.BivData_IFD.PER_Ttest$p.value
+#p-value < 2.2e-16
+
+#MOUSE AVE
+Rec.group_IFD.PER_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group_IFD.PER_Ttest$p.value 
+#p-value = 6.872e-07
+
+```
+
+The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
+
+### Q2 Long Bivalent Dataset
+
+Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+The section on IFD only looks at the 2CO bivalents.
+
+```{r Q2.long_biv.IFD, echo=FALSE}
+#long BivData
+male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
+
+#show
+Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+
+Long.biv.IFDs_PER_by.group
+
+```
+
+table(droplevels(male.long.biv.2CO_IFD$strain) )
+
+The plot above shows the normalized IFD measures from 2CO bivalents
+
+There are many fewer long bivalent data observations, but the same positive correlation with
+
+-table of long bivalents
+
+- glms
+
+- logistic regression
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE}
+
+#Full dataset
+
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+
+```
+
+There are not enough observations for sub setting the data into strains or use the mouse averages.
+
+Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
+The above t.tests are breaking the knitr
+
+The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
+
+None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
+
+Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
+This also indicated that interference distance may evolved in the house mouse complex.
+
+#Caveats
+
+put all of the code chunks/analysis for caveats here
+
+## Chromosome Size Effect
+
+I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)
+
+<!-- 
+Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
+But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
+--> 
+Below are examples of plots of SC length distributions across cells.
+The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
+Most plots excluded for space.
+
+Each point is a bivalent plotted by cell on the x axis.
+X's are the 4th quartile, big point is the mean and smaller black point is the median.
+I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
+(the extra stats are not correctly mapped)
+
+```{r SC.distb_manual_Auto, echo=FALSE}
+
+#calq the mouse means
+manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
+                        median_SC = median(SC.length, na.rm = TRUE),
+                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
+#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
+
+)
+
+manual_summary_stats <- add_mouse(manual_summary_stats)
+manual_summary_stats <- add_strain(manual_summary_stats)
+manual_summary_stats <- add_sex(manual_summary_stats)
+manual_summary_stats <- add_subsp(manual_summary_stats)
+
+###
+#calq the mouse means
+Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        median_SC = median(chromosomeLength),
+                        Qar.4 = quantile(chromosomeLength)[[4]],
+
+                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
+        #add Q4
+
+)
+Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
+
+#colors_of_strains
+#scale_color_manual(values=colors_of_strains)
+
+#un-order these factors
+```
+
+```{r long.biv.plots, echo=FALSE}
+
+show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
+
+show.the.cells_females_dom
+#PWD seems to have the biggest variance across cells
+
+almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+  scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
+almost_cells_DOM_f
+
+#### Musc female
+
+show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
+
+almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
+
+### Dom Male
+show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
+
+almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
+
+#MUSC male
+show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
+
+almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
+
+#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
+
+#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+ # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
+#show.the.cells_male
+
+#almost_whole_cell_BivData
+#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
+#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
+#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
+
+```
+
+All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
+It shows the amount of within cell variance across strains.
+There is a bit of variance across the SC length distributions in the PWD females.
+
+This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
+
+The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
+The full data set is `r length(Curated_BivData$Obj.ID)`.
+This is the breakdown of bivalent observations by category for this long dataset are `r data.frame(table(droplevels(real.long.bivData$category)))`.
+
+**- Try to merge this DF with the whole.cell manual measures.**
+
+- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.
+
+in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
+10202 bivalents from 86 mice.
+
+```{r show.long.SC, echo=FALSE}
+
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
+female_long_biv_plot
+
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
+male_long_biv_plot
+
+```
+
+These plots show the SC lengths for the 'long SC data set'.
+They are supposed to be the longest 4-5 SC from cells where I could get good measures.
+These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
+Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
+
+## Adjusting for XX
+
+(rethink where this section should go)
+new outline
+
+1. illustrate problem(affects mostly SC length)
+
+2. Expected impact on sex comparisons, estimated effect size of the X  
+3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
+
+4. 19 female, 19male, 20female 20 male
+
+The female mouse averages should have adjustments for the XX.
+working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
+Subtract this amount from the female mouse averages ...
+This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
+
+- Of all female single bivalents observations, 5% are XX (1 of 20).
+
+- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
+
+- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
+
+I think a formula something like this can be applied to adjust for XX
+
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
+
+#### Whole Cell Manual Comparison
+
+```{r XX.adj_manual.whole.cell, echo=FALSE}
+
+#load and examine the the whole cell hand measures here
+
+female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
+data.frame(table(droplevels(female.whole.cell$category)))
+
+# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
+
+#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
+
+data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
+#missing some strains, most have 3 cells (min). MSM has 5 full cells
+
+#remove the incomplete LEW female cell
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
+
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
+
+#calq mean SC lengths by rank
+
+mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
+
+                       mean.sc =   mean(SC.length, na.rm = TRUE),
+                         nbiv = length(Obj.ID),
+                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
+                         se.SC_length   = sd.SC_length / sqrt(nbiv)
+)
+         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
+
+top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
+
+show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
+    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
+                    position=position_dodge(width=0.7), 
+                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
+          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
+
+show.SC_by.rank_top
+
+bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
+
+show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
+    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
+                    position=position_dodge(width=0.7), 
+                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
+          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
+
+#CALQ PERCENT TOTAL SC
+
+CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
+
+                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
+                         nbiv = length(Obj.ID),
+                       #rank.by.per
+
+                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
+                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
+                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
+                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
+                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
+                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
+                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
+)
+
+#MELT ABOVE
+MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
+              #leave out the Foci
+               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
+
+            id=c( "fileName","total.sc.cell","nbiv"))
+
+MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
+
+MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
+MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
+
+show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
+
+show_long.sc.proportions
+
+#original plot code 
+#ggplot(data = df1, aes(x, y, color = Group)) +
+#    geom_pointrange(aes(ymin = lb, ymax = ub), 
+#                    position=position_jitter(width=0.5), 
+#                    linetype='dotted') +
+#    theme_bw()
+```
+
+The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
+
+The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
+
+The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
+For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
+
+(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
+
+The other figure shows of each single bivalent contributes to the total SC area.
+Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
+on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
+So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
+
+- Is the difference between cell averages for males and females less that 10%?
+
+- also interesting, the pwd and msm don't have longer SC, compared to other strains.
+
+#### Automated BivData Comparison
+
+```{r XX.Automated.biv.stuff, echo=FALSE}
+
+#make automated female data to compare to
+female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
+
+female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
+
+table(female.Bivdata_comp$strain)
+
+#female.Bivdata_comp
+
+#1. calq the rate of chrm segmentation in female categories in automateted data
+
+box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
+
+                   Nmice = length(unique(mouse)),
+                   mx.box.num = max(boxNumber), # need to run this at the cell level
+                    N.biv  = length(Obj.ID),
+                   pass.biv = N.biv / mx.box.num
+)
+box.rates <- add_mouse(box.rates)
+box.rates <- add_strain(box.rates)
+box.rates <- add_category(box.rates)
+
+cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+
+cell.pass.rates
+
+category.box.rates <- ddply(box.rates, c("category"), summarise,
+
+                   mean.pass.rate = mean(pass.biv)
+)
+#category.box.rates
+
+#2. pull the 3rd rank biv's calculate the SC means
+#calq means by rank (focus on the top 5)
+
+```
+
+For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
+
+The table above shows the number of bivalents from the same strains as in the manual whole cell data.
+The plot shows the bivalent passing rate across all of the individual cells from this female data set.
+For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
+
+(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
+(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
+
+strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
+
+** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
+But I am not using 'whole cell' summaries to compare female and males.
+
+**What is the effect of an extra XX-autosome on single bivalent means?**
+
+use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
+Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
+Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
+Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
+The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
+
+# Note on Heterochiasmy Definition
+
+<!-- add outline for new XX adjust section 
+adjust the colors for the plot below (standardize)
+-->
+
+I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
+In females the XX bivalent is indistinguishable from the autosomes.
+To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
+Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
+(I note if the and Y are paired, which they are at a high rate).
+The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
+To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
+
+1. Compile full-cell data from females (all 20 bivalents measured)
+2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
+3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
+
+According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
+
+(Put the XX adjustment section here)
+
+```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
+#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
+#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
+#remove quality 5 cells
+
+#outgroups
+#SPIC
+HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
+#unique(HetC.SPIC$mouse)
+#non-passing mice
+SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
+HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
+#calq sex mean
+SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
+
+#SPRET
+HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
+#unique(HetC.SPRET$mouse)
+#non-passing mice
+#SPRET.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
+
+#CAROLI
+HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
+#unique(HetC.CAROLI$mouse)
+#non-passing mice
+#SPRET.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+#cast
+HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
+#unique(HetC.CAST$mouse)
+#non-passing mice
+#X.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
+
+#HMI
+HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
+#unique(HetC.HMI$mouse)
+#non-passing mice
+#X.remove.mice <- c("", "")
+#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
+#calq sex mean
+HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+#dom
+
+#WSB
+HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
+#unique(HetC.WSB$mouse)
+#non-passing mice
+WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
+HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
+#calq sex mean
+WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
+
+#LEW
+HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
+unique(HetC.LEW$mouse)
+#non-passing mice
+LEW.remove.mice <- c("4jan17_LEW_f2")
+HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
+
+#calq sex mean
+LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
+
+#G
+HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
+#unique(HetC.G$mouse)
+#non-passing mice
+G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
+HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
+
+#calq sex mean
+G.mean.MLH1.obs <- ddply(.data=HetC.G, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
+
+#PERC
+HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
+#unique(HetC.PERC$mouse)
+#non-passing mice
+#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
+#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
+
+#calq sex mean
+PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
+
+#MUSC
+
+#PWD
+HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
+#unique(HetC.PWD$mouse)
+#non-passing mice
+PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
+
+# CZECH
+HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
+#unique(HetC.CZECH$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
+
+# SKIVE
+HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
+#unique(HetC.SKIVE$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
+
+# KAZ
+HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
+#unique(HetC.KAZ$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
+
+#TOM
+HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
+#unique(HetC.TOM$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
+
+#AST
+HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
+#unique(HetC.AST$mouse)
+
+#non-passing mice
+#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
+#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
+
+#calq sex mean
+AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
+
+#Molosinus
+
+#MSM
+HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
+unique(HetC.MSM$mouse)
+
+#non-passing mice
+MSM.remove.mice <- c("12sep16_MSM_f3")
+HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
+
+#calq sex mean
+MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
+
+#MOLF
+HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
+unique(HetC.MOLF$mouse)
+
+#non-passing mice
+MOLF.remove.mice <- c("27oct18_MOLF_m3")
+HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
+
+#calq sex mean
+MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci),
+                 var.MLH1 = var(adj_nMLH1.foci)
+)
+
+MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
+
+```
+
+```{r HetC.plot, echo=FALSE, warning=FALSE}
+#HETC plot
+
+#set up data
+
+HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#
+#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
+
+HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
+                         Nmice = length(unique(mouse)),
+                         Ncells  = length(adj_nMLH1.foci),
+                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
+                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         se   = round(sd / sqrt(Ncells), 3),
+                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
+)
+#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
+
+#remove the 'other'
+
+HetC_strain_table.M$sex <- "male"
+HetC_strain_table.F$sex <- "female"
+
+HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
+
+HetC_strain_table <- add_subsp(HetC_strain_table)
+
+#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
+
+HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
+HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
+#str(HetC_strain_table)
+
+Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
+  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
+
+Het.C.plot
+#this plot is missing several
+
+```
+
+There is now MOLF, which has female biased hetC
+3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
+1 of the musc strains has female biased heterochiasmy, KAZ.
+
+```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
+
+#REMOVE This BLOCK?
+
+#### Distributions by mouse
+#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
+
+#remove this block since it has it's own src/script
+
+### to integrate 'pass'/fail
+### this should be in setup data
+
+#mice that don't have female data will break the chain-loop
+#MOLF female, HMI, CZECH female, CAROLI female, Tom male
+
+#instead of adding fake data, remove the categories from list
+
+#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
+#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
+#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
+#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
+#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
+#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
+#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
+#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
+#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
+
+pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
+
+#this loop makes a quant_status col for a list of mice
+for(i in 1:length(MLH1_data$mouse)){
+  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
+}
+
+df_list <- split(MLH1_data, (MLH1_data$category))
+#remove sections which are in category but have no data
+#this is where you remove the non-data data!
+df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
+
+#7 PERC, 12 MOLF female, 
+#
+
+#df list, is a list of df's which were split() from BIG df by category
+df_list 
+o=0
+i=0
+
+MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
+
+#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
+for(i in 1:(length(df_list))){
+ # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
+  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
+  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
+                      mouse_nMLH1 = mean(adj_nMLH1.foci)
+  )
+# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
+  for(o in 1:length(df_list[[i]]$Original.Name)){
+    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
+        }
+}
+
+plot_list = list()
+for (h in 1:length(df_list)) {
+  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
+    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
+    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
+    facet_wrap(~ mouse, drop = TRUE) + 
+    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
+  #make an empty ggplot? for missing data
+
+  plot_list[[h]] =   pf
+}
+
+#orig (I think this is the same as above)
+#plot_list = list()
+#for (h in 1:length(df_list)) {
+#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
+#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
+
+#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
+#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
+#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
+
+ #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
+
+    #theme for indiviual mice, 
+#theme(plot.background = element_rect(fill = "green"))
+#  plot_list[[h]] =   pf
+#}
+#can't change the alpha/transparency of the red rec
+
+#aes must be same length ---   #geom_point(stat = "identity") +
+# the last one is correct....?
+
+#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
+```
+
+```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
+geom_hline(aes(yintercept = med, group = gr), colour = 'red')
+#no plot is made that goes in the list -- so it can't be shown
+for (gg in 1:length(plot_list) ){
+  show(plot_list[gg])
+ # invisible(lapply(obj, function(x) plot(x,main="some plot")))
+}
+
+#this stops at MSM
+```
+
+The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
+
+Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
+The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
+
+```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
+
+no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
+
+MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
+CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
+
+MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
+
+MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
+
+MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
+
+kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
+
+bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
+
+#add titles
+#use multiplot
+#remove X axis
+
+multiplot(kuku,bubu, cols=2)
+```
+
+# Deleted notes
+<!--
+For later analysis.  Not sure what these are good for.
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
+
+# check if the main mixed model results hold up when all cells are pooled
+
+#pooled.cells
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+#consider changing to ordered = FALSE, for the DFs used for MM
+DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
+
+DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
+
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+
+```
+
+- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
+This means there is 0 variance added to mean_CO to mean due to being MSM.
+Does this mean the model is very good at predicting MSM mean_CO.
+The fixed strain - sex effects are for musc are large.
+This will likely change with MOLF (male molf are low, so smaller fixed effect and )
+
+- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
+In the first model MSM has largest random effect, since it's so far from WSB female.
+
+-adding MOLF also doesn't have significant factors
+
+changed all to unordered factors.
+all the variables are coded as ordered factors.
+When the variables are kept factors, and unordered the names appear! yay!
+
+Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
+-->
+
+<!--
+
+#from chrm proportion black
 
@@
 
-#Need to run the chrm proportions thru melt code.
-
-chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
-
-#chrm.class.DF$biv.sum.check
-```
-
-For following up on the biv stuff, find ways to compare the 1CO to 2COs in Dom and Musc. (sis-coten, length, distance from centromere).
-
-The figure below shows the average percentages of chromosome classes per category gathered from BivData
-
-```{r Chrm.props, echo=FALSE}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$CO0 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$CO1 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$CO2 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$CO3 / mouse.level.BivData_table$nboxes_passed
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$CO4 / mouse.level.BivData_table$nboxes_passed
-
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(11:15 ) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#"mouse"         "ncells"        "boxs.IDd"      "nboxes_passed" "Nbiv"          "CO0"           "CO1"       #[8] "CO2"           "CO3"           "CO4"           "per.0CO"       "per.1CO"       "per.2CO"       "per.3CO"      [15] "per.4CO"
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")], 
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","CO0","CO1","CO2", "CO3","CO4"))
-
-#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-
-#MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-#no filename for add mouse, strain subsp
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#can't apply function use this
-chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
-             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
-                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
-                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
-
-          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
-              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
-
-                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
-                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
-            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
-                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
-                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
-                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
-                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
-                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
-
-             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
-               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
-
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",
-                                                                 "Cast",
-                                                                 "Musc",
-                                                                 "Mol",
-                                                                 "Spretus","Spic",
-                                                                 "other") )
-#ggplot
-
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
-chrm.class.barplot
-
-```
-
-These results align with the previous figure I made using percentages from hand measured results (dozens of observations, instead of hundreds).
-These results fit with the rapid male evolution of gwRR in MSM and PWD.
-
-Look into ordering the mouse strains to highlight the MSM and PWD differences, renaming the chrm type labels, spacing out the x axis labels or removing some of the other mice.
-
-# DMC1 Results, Spermatocytes
-
-These data are meant to test if CO precursors, double strand breaks (DSBs), are significant predictors for MLH1 count variation. Simply, do mice with more DSBs also have more COs?
-
-ToDo:
-  - attempt to supplement data with more KAZ DMC1 cell observations.
-  - change all the bad variable names
-
-```{r DMC1.setup, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
-                 .(category),
-                 summarize, 
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "L"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
-                 .(category),
-                 summarize, 
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "Z"
-```
-
-```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#these are histogram plots for testing Normality shape -- move to supplemenatry section
-
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-DMC1.histogram
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-```
-
-The counts look close to normal.
-
-Potentially figure 2. Using a function to display the p-vales for mean comparison
-
-```{r DMC1.plot, echo=FALSE, warning=FALSE}
-#think of better color scheme
-#make the jitter col thinner or make the stacked dots.
-#remember that box-plots add outlier dots.
-#change y labels
-
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")
-
-DMC1.boxplot
-  #this will add neat bars, but still need to figure out what groups ect it's actually using
-#THE stat_compare function adds a arm thing and pvalue for 
-#  stat_compare_means(comparisons = my_comparisons)
-
-#status - legend, expression y-axis, facet, facet_wrap SYMBOL
-#legend - stage. y=total count, FACet strain
-
-#add column to distinguish high vs low rec strains
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
-
-# annotation table with adjusted pvals and y-position of the labels
-anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",]) %>% mutate(y_pos = 300)
-
-DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=stage), position=position_jitterdodge()) + 
-  facet_wrap(~stage, nrow=1) + 
-  ggsignif::geom_signif(
-    data=anno_df, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")
-#this is kinda the plot I'm enivioning 
-DMC1.bxlt_High.to.Low
-
-#I want a test between 1) high vs low cell stages, 
-#I think a better way to display would be
-#facet by stage, x is the strain
-#compare means across high vs low  (this level isn't defined)
-
-#example from here
-#https://github.com/kassambara/ggpubr/issues/65
-
-#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
-
-```
-
-Summary:
-
-The distributions of counts is close to normal (also the higher count is better than the MLH1 counts)
-
-Above figure shows p-values for comparison of means across the cell stages.
-
-Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days).
-One mouse represents each strain.
-The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
-
-ToDo think of ways to add in the compared means for the high and low groups (across cell stage)
-how similar are G-WSB or PWD-MSM to each other?
-
-## DMC1 Permutations with equal sampling
-
-The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
-Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
-Permutations were done for both cell stages.
-The actual p value is in red.
-
-The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
-
-```{r DMC1.subsampling, echo=FALSE}
-
-#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
-# the 2 groups
-
-#make the 2 pools
-High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-
-Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-
-#permute drawing / subsampling and performing the t-test
-#put the p.value results into a DF!
-
-#permut.SC.cv.diff  <- as.data.frame(
-#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
-
-#permuting samples -- and re-licating t-tests
-
-#nrep <- 1000
-
-#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
-bunchOpvals.L <- replicate(1000, 
-    # t.test
-    t.test( sample(High.L.pool, 25, replace = FALSE),
-            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
-          )
-bunchOpvals.L <- as.data.frame(bunchOpvals.L)
-bunchOpvals.L$stage <- "L"
-
-bunchOpvals.Z <- replicate(1000, 
-    # t.test
-    t.test( sample(High.Z.pool, 25, replace = FALSE),
-            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
-          )
-
-bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
-bunchOpvals.Z$stage <- "Z"
-
-#could merge the lists of p-values
-replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
-#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
-
-#ASSUMPTIONS; normality, equal variance
-
-#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-#plot.pvals.L
-#plot.pvals.Z
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
-```{r DMC1_MLH1.cor, echo=FALSE}
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-#attach this mini DF to the DMC1 data
-
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-```
-
-> At what stage are DMC1 foci numbers most correlated with MLH1?
-
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-
-```{r barlett.test, echo=FALSE, eval=FALSE}
-
-#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
-bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
-
-```
-
-Evidence for non-equal variance across strains for zygotene cells.
-
-Ideas for post-hoc tests, remove strains 1 at a time and test variance. Test variance for L cells without KAz.
-
-From ANOVA analysis strain is significant for Leptotene cells but not Zygotene cells. (but take these results with a grain of salt before assumptions are tested.)
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova <- anova(mod.L) #0.0002701 ***
-z.anova <- anova(mod.Z) #NS
-
-l.anova
-z.anova
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#L vs Z (all species)
-#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
-#p-value = 1.032e-05
-#not a valid interpertation
-
-#high vs low group (L)
-ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-Divide the Musculus strains into high and low recombining groups to characterize the difference in distributions.
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = r ttest.HighLow.L$p.value) while the Z cells, there is not a significant difference across the high and low groups (value = r ttest.HighLow.Z$p.value).
-
-<!--
- Single Bivalent Patterns
-
-## Total SC Results
-
-insert male total SC results here
-
-# Female patterns
-
-Nuanced
-there is strain specific evolution for females ...
-
-# Sexual dimorphism and heterochiasmy
-
-total SC length / SC compaction and telomere bias in positioning are conserved / consistently sexually dimorphism (in these mice?)  for most species
-
-make the heterochiasmy plot, with female strain averages adjusted for X
-(measure variance in CO counts for the top 5 female whole cell bivalents)
-(variance in )
-how does this compare to variance of mid 5, bottom 5?
-
-# Predictions / tests for selection on Bivalent structure
-
--->
-
-<!--
-# Rapid male specific evolution
-
-list out the points which support rapid male specific evolution for MLH1 rates.
-
-number above bars are the raw counts for bivalent classes.
-
-Potential figure 2 (eventually). Try to make the figure taller.
--- add more data to make sample sizes more equal.
--- add space to indicate where female data is missing
-
-SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
-What's up with Lew females? there's lots of 2COs, also WSB female.
-
-This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
-If these data are random samples of bivalents, the X
-
-ToDo: chisq tests for difference in proportions, 
-- within strain across sex,
-- males, across high and low groups
-- look at mouse level proportions
--->
-
-## Note on Heterochiasmy Definition
-
-<!-- add outline for new XX adjust section 
-adjust the colors for the plot below (standardize)
--->
-
-I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
-In females the XX bivalent is indistinguishable from the autosomes.
-To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
-Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
-(I note if the and Y are paired, which they are at a high rate).
-The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
-To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
-
-1. Compile full-cell data from females (all 20 bivalents measured)
-2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
-3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
-
-According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
-
-(Put the XX adjustment section here)
-
-```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
-#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
-#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
-#remove quality 5 cells
-
-#outgroups
-#SPIC
-HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
-#unique(HetC.SPIC$mouse)
-#non-passing mice
-SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
-HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
-#calq sex mean
-SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
-
-#SPRET
-HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
-#unique(HetC.SPRET$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
-
-#CAROLI
-HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
-#unique(HetC.CAROLI$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-#cast
-HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
-#unique(HetC.CAST$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
-
-#HMI
-HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
-#unique(HetC.HMI$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-#dom
-
-#WSB
-HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
-#unique(HetC.WSB$mouse)
-#non-passing mice
-WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
-HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
-#calq sex mean
-WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
-
-#LEW
-HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
-unique(HetC.LEW$mouse)
-#non-passing mice
-LEW.remove.mice <- c("4jan17_LEW_f2")
-HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
-
-#calq sex mean
-LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
-
-#G
-HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
-#unique(HetC.G$mouse)
-#non-passing mice
-G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
-
-#calq sex mean
-G.mean.MLH1.obs <- ddply(.data=HetC.G, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#PERC
-HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
-#unique(HetC.PERC$mouse)
-#non-passing mice
-#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
-
-#calq sex mean
-PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#MUSC
-
-#PWD
-HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
-#unique(HetC.PWD$mouse)
-#non-passing mice
-PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
-
-# CZECH
-HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
-#unique(HetC.CZECH$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
-
-# SKIVE
-HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
-#unique(HetC.SKIVE$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
-
-# KAZ
-HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
-#unique(HetC.KAZ$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
-
-#TOM
-HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
-#unique(HetC.TOM$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
-
-#AST
-HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
-#unique(HetC.AST$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
-
-#Molosinus
-
-#MSM
-HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
-unique(HetC.MSM$mouse)
-
-#non-passing mice
-MSM.remove.mice <- c("12sep16_MSM_f3")
-HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
-
-#calq sex mean
-MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
-
-#MOLF
-HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
-unique(HetC.MOLF$mouse)
-
-#non-passing mice
-MOLF.remove.mice <- c("27oct18_MOLF_m3")
-HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
-
-#calq sex mean
-MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
-
-```
-
-```{r HetC.plot, echo=FALSE, warning=FALSE}
-#HETC plot
-
-#set up data
-
-HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#
-#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
-
-HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
-
-#remove the 'other'
-
-HetC_strain_table.M$sex <- "male"
-HetC_strain_table.F$sex <- "female"
-
-HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
-
-HetC_strain_table <- add_subsp(HetC_strain_table)
-
-#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
-
-HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
-HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
-#str(HetC_strain_table)
-
-Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
-  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
-
-Het.C.plot
-#this plot is missing several
-
-```
-
-There is now MOLF, which has female biased hetC
-3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
-1 of the musc strains has female biased heterochiasmy, KAZ.
-
-```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
-
-#REMOVE This BLOCK?
-
-#### Distributions by mouse
-#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
-
-#remove this block since it has it's own src/script
-
-### to integrate 'pass'/fail
-### this should be in setup data
-
-#mice that don't have female data will break the chain-loop
-#MOLF female, HMI, CZECH female, CAROLI female, Tom male
-
-#instead of adding fake data, remove the categories from list
-
-#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
-#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
-#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
-#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
-#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
-#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
-#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
-#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
-#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
-
-pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
-
-#this loop makes a quant_status col for a list of mice
-for(i in 1:length(MLH1_data$mouse)){
-  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-}
-
-df_list <- split(MLH1_data, (MLH1_data$category))
-#remove sections which are in category but have no data
-#this is where you remove the non-data data!
-df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
-
-#7 PERC, 12 MOLF female, 
-#
-
-#df list, is a list of df's which were split() from BIG df by category
-df_list 
-o=0
-i=0
-
-MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
-
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-for(i in 1:(length(df_list))){
- # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                      mouse_nMLH1 = mean(adj_nMLH1.foci)
-  )
-# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-  for(o in 1:length(df_list[[i]]$Original.Name)){
-    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
-        }
-}
-
-plot_list = list()
-for (h in 1:length(df_list)) {
-  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
-    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
-    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
-    facet_wrap(~ mouse, drop = TRUE) + 
-    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
-  #make an empty ggplot? for missing data
-
-  plot_list[[h]] =   pf
-}
-
-#orig (I think this is the same as above)
-#plot_list = list()
-#for (h in 1:length(df_list)) {
-#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
-
-#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
-#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
-
- #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
-
-    #theme for indiviual mice, 
-#theme(plot.background = element_rect(fill = "green"))
-#  plot_list[[h]] =   pf
-#}
-#can't change the alpha/transparency of the red rec
-
-#aes must be same length ---   #geom_point(stat = "identity") +
-# the last one is correct....?
-
-#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
-```
-
-```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
-geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-#no plot is made that goes in the list -- so it can't be shown
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
- # invisible(lapply(obj, function(x) plot(x,main="some plot")))
-}
-
-#this stops at MSM
-```
-
-The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
-
-Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
-The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
-
-```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
-
-no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
-
-MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
-
-MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
-
-kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
-
-bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
-
-#add titles
-#use multiplot
-#remove X axis
-
-multiplot(kuku,bubu, cols=2)
-```
-
-```{r misc.setup, echo=FALSE}
-
-colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
-                       #blues
-
-                       'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
-                       'CZECH'= 'indianred4',
-
-                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
-
-                    'CAST'='green','HMI'= 'forestgreen', #greens
-
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
-                    'CAROLI'='gold3',
-
-                    "F1"='grey', "other" = 'grey'
-                    )
-
-#add Rec group,
-#reclassified SKIVE as intermediate
-Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
-#apply this before ploting, to display strains in desired order
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, category, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-```
-
-<!-- merging the 2 data sets-->
-
-# Single Bivalent Level Results
-
-Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script.
-These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output).
-The MLH1 data file is also loaded into this file.
-
-The breakdown of single bivalent observations by category
-
-`r data.frame(table(droplevels(Curated_BivData$category) ) )`
-
-Two Main questions for these analyses.
-
-> Q1. driving questions, which traits are sexually dimorphic?
-
-> Q2. which traits fit male polymorphism predictions?
-
-These questions use different datasets, q1 sex matched strains
-
-Q2 all male observations with including strains not in the q1 data set
-
-# Model Frameworks
-
-<!--
-As alex said, be careful about mixed models and interpreting the random effect (random variance) -- assigning strain as random effect is slightly problematic as i can give it a biological interpretation (it's due to strain -- genetics of each of the strains).
--->
-
-Q1. Model 1, mixed model lmer()  
-$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-Q1. Model 2, glm()
-$$mouse \ av.\  metric ~=~ sex * strain + \varepsilon $$
-
-Q2 Model 1, mixed model lmer()
-
-$$mouse \ av.\  metric ~=~ subsp *  + rand(strain) + \varepsilon $$
-
-Q2 Model 2,  glm()
-
-$$mouse \ av.\  metric ~=~ strain + \varepsilon $$
-
-Q2 Model 3, logistic regression
-
-$$Rec \ group  ~=~ mouse \ av.\  metric  $$
-
-I decided to use the mouse averages for most of the bivalent metric tests. some of my logic is below.
-
-**Pro Bivalent level**
-
-- The data are continuous and closer to normal/Gaussian expectations than the MLH1 counts.
-
-- There is within cell variance that is an issue (would affect single bivalent observations).
-
-**Pro Mouse averages**
-
-- Mouse average is a conservative choice, it would summarize general patterns which fits with the paper.
-
-- Simple application of the same models as MLH1 based results.
-
-- Limited work to add to the data set if I can justify that the collection of single bivalents are random and equivalent across mice.
-
-So I'll use the mouse averages for all the metrics is what is used in the mixed models.
-
-```{r Q1.biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}
-
-#this block should be in the data set up
-
-#the mouse averages will be used
-
-mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-mouse.avs_4MM <- add_sex(mouse.avs_4MM)
-mouse.avs_4MM <- add_strain(mouse.avs_4MM)
-mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-mouse.avs_4MM <- add_category(mouse.avs_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
-mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
-mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
-
-#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-#code if males are high or low?
-
-#code intp chr
-
-#assign high and low
-#divide musc males into groups
-mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
-
-```
-
-In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#add this to setup
-
-#remove female, 
-#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
-
-#Curated_BivData_male
-
-male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-male.av_4MM <- add_sex(male.av_4MM)
-male.av_4MM <- add_strain(male.av_4MM)
-male.av_4MM <- add_subsp(male.av_4MM)
-male.av_4MM <- add_category(male.av_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
-male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
-male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-```
-
-## Q1 Analysis, Predictions for Heterochiasmy Q1
-
-Using the Mixed model framework which tests the effects (and interactions) of subspecies, sex and strain, I will test for evolution of the following traits.
-
-Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
-
-1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of uniform vs telomere/biased CO positioning.
-
-  C) Centromere and telomere distances will be sexually dimorphic.
-
-3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
-
-# Q1 SC Lengths
-
-<!--Brief lit review for SC length and MLH1 expections.
-add XX adjusment / arguments
--->
-
-We expect female SC lengths to be longer (refs).
-In the plot above the SC length ~ higher hand foci cells.
-Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
-For the 0 class chromosomes most all are around the same size of the 1CO distribution.
-Add in the code for the plot under 2COs males and females
-
-```{r show.SC.length, echo=FALSE, warning=FALSE}
-#for now remove the 
-#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
-#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS
-
-#add boxplots for mouse level means
-#add second x axis, hand.foci.count
-
-#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#nested boxplots
-SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
-
-SC.length_Nested_boxplots
-
-#+ +geom_jitter(aes(alpha=0.2))
-```
-
-> M1. subsp * sex + (strain)
-
-> M2 subsp \* sex \* strain  (don't rally use)
-
-> M3 sex * strain
-
-The dependent variables I'll be testing in the model framework are:
-
-- Pooled SC lengths
-
-- 1CO
-
-- 2CO
-
-- 3CO
-
-- Xlong.biv  *Don't do long bivData for testing sex effect, since this would have the XX
-
-- Short.biv dataset *this would be a chrm size effect, without the XX
-
-<!--
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
-
-To estimate random effects, I'm using lmer (not lme where I could do the rand(strain*sex)). So now it's just random(strain).
-
-The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.
-
-remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
-
-#### Pooled SC Lengths
-
-```{r MM_2.SC.Length, echo=FALSE, eval=FALSE}
-
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-#p
-
-```
-
-I used the mixed model with the lme() for the full Curated BivData and long bivdata
-
--->
-
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-#### Q1.SC_HetC_M1
-
-Insert the results for the MMs using the dependent variable 1CO - use the Mixed model (lmer), for all of the metrics (pooled SC, 1CO 2CO ect)
-
-Below's the code for Mixed model results -- try to organize them into a table or something
-
-```{r 1.SC.Length_M1, echo=FALSE}
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#all pooled SC lengths
-lmer.Q1_SCLength_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-#drop1(lmer.SCLength_all_MM.M1, test="Chisq" ) #other options for test don't work
-#anova(lmer.SCLength_all_MM.M1) # also F values
-exactRLRT(lmer.Q1_SCLength_M1)
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
-#can't get p values from this summary
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-#drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")
-exactRLRT(Q1.lmer_SCLength_1CO_M1)
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-#drop1(SCLength_2CO_MM.M1, test="Chisq")
-exactRLRT(Q1.lmer_SCLength_2CO_M1)
-as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-#Long.Biv
-Q1.lmer_SCLength_long.biv_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
-
-#drop1(Q1.lmer_SCLength_long.biv_M1, test="Chisq")
-#exactRLRT(Q1.lmer_SCLength_long.biv_M1)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
-
-#### Q1. SC_HetC_M2
-
-All fixed and all interactions
-
-```{r Q1.SC.Length_M2, echo=FALSE}
-#feed these 2 DFs into M2
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-#glm(metric ~ strain)  ---mean_SC
-
-Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_SCLength_M2)
-
-#Dom.all
-Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.Dom)
-
-#Dom.1CO
-Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.1CO.Dom)
-
-#Dom.2CO
-Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.2CO.Dom)
-
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.Musc)
-
-#Musc.1CO
-Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.1CO.Musc)
-
-#Musc.2CO
-Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.2CO.Musc)
-
-#Mol.all
-Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.Mol)
-
-#Mol.1CO
-Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.1CO.Mol)
-
-#Mol.2CO
-Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.2CO.Mol)
-```
-
-```{r Q1.SC.M2plot, echo=FALSE}
-
-#Q1 plots for SC Lengths across Musc strains
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#mouse av
-Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
-Q1.SClength_Musc_plot
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-Q1.SClength_Musc.BivData_plot
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-Q1.SClength_Dom.BivData_plot
-
-```
-
-I think the strange thing about the Musc is that Skive females have shorter SC lengths.
-
-#### Q1. SC Length t.test
-
-```{r strain.SClength.t.test, echo=FALSE}
-#t.tests for all the strain specific data frames
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-#### Strain Comparisons of SC Length
-
-Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
-
-(adding all of these blocks to -- the setup script); strain.Bivalents.DF
-
-#### XX adjustment
-
-Generate inteligible summaries for the permutation / subsampling results.
-
-```{r Q1_XX.adjustment, echo=FALSE}
-
-```
-
-# Q1. IFD
-
-Brief lit review for IFD / interference expectations  
-**Petkov et al 2007,  CO interference underlies sex differences in RR**
-
-"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
-However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
-at the pachytene stage of meiosis."
-
-Chrm1 genetic map,
-
-(other human and mice refs
-
-Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
-
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-
-))
-
-*Do I believe the physical/SC results?*
-
-<!--
-ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
-
-explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
--->
-
-3) Interference / IFD will not be sexually dimorphic.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
-
-**I**nter**f**ocal **D**istance, a indication of CO interference.
-
-The raw measures are pretty nosy, this might not be the best way to display these results.
-
- Each point is a IFD observation. Maybe add mouse level lines for the means.
-
-```{r show.IFD_99, echo=FALSE, warning=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r IFD.old.plots, echo=FALSE, eval=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
-Still working on the best way to display the general IFD patterns.
-
-### Mixed Model Tests, Fixed Effects
-
-Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
-
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-### IFD_HetC_M1
-
-```{r HetC_M1_IFD_many.metrics, echo=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-#anova(lmer.IFD.PER_MM.M1)
-
-#exactRLRT(lmer.IFD.PER_MM.M1)
-#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
-
-#M1_IFD.RAW_2CO
-lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
-#anova(lmer.IFD.RAW_MM.M1) # also F values
-
-#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
-#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-```
-
-Sex is a slightly significant factor for raw and normalized IFD.
-
-The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
-That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
-
-show a raw plot, and norm plot and a scatter plot
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
-The tables below are from anova( for the lmer model ).
-Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
-
-### IFD_HetC_M2
-
-```{r IFD_HetC_M2, echo=FALSE}
-
-```
-
-### IFD_HetC_M3
-
-```{r IFD_HetC_M3, echo=FALSE}
-
-```
-
-## Strain Comparisons
-
-Dive deeper into the sex specific pattern for each strain.
-Below are code chunks which show the unusual sex specific results for IFD measures.
-The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
-The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
-
-For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
-
-```{r Comp_Ratios, echo=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
-
-```{r WSB.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#WSB.Bivalents.DF
-
-#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
-WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
-
-#plot the differences
-wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_raw
-
-wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_PER
-
-WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-WSB.IFDnorm_scatter.plot
-#this plot is much messier than the MSM (PWD?)
-
-#try plotting the other IFD
-
-#make a table of the means and n ect
-wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
-
-                       total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
-
-The range of normalized IFDs overlap closer in males and females in the WSB data.
-
-```{r LEW.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-
-LEW.Bivalents.DF_IFD <- LEW.Bivalents.DF[(!is.na(LEW.Bivalents.DF$sex)) & (LEW.Bivalents.DF$hand.foci.count == 2),]
-
-LEW.Bivalents.DF_IFD$sex <- droplevels(LEW.Bivalents.DF_IFD$sex)
-
-#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-lew.ploty_IFD_raw <- ggplot( LEW.Bivalents.DF_IFD,  aes(y=IFD1_ABS, x = sex, na.rm = TRUE))+                     geom_jitter( na.rm = TRUE)+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex, na.rm = TRUE))
-lew.ploty_IFD_raw
-#arg, little annoyed that the x NA won't go away
-
-lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
-lew.ploty_IFD_PER
-
-LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  
-  scale_color_manual(values=c('#56B4E9','#E69F00'))+
-  geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-LEW.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-
-                  #     mean_SC = mean(chromosomeLength),
-                       n_SC = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-
-```
-
-The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
-
-```{r G.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-#plot the metric
-#remove the NAs
-
-G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
-
-G.ploty_IFD_raw
-#
-G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.ploty_IFD_PER
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-
-#calQ percent < 30%
-
-oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
-oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
-
-#the na's are being counted (ughhhhhh)
-#use the file thing instead
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
-  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
-
-LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
-#0.7062663
-
-#would Chi-sq be a good test to compare the ratios ect?
-#the sum of these doesn't add 
-#G.Bivalents.DF
-
-```
-
-```{r PWD.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#remove the NAs
-
-PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
-
-PWD.ploty_IFD_raw
-
-PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
-PWD.ploty_IFD_PER
-
-PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
-
-PWD.IFDnorm_scatter.plot
-
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-```
-
-For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
-
-```{r KAZ.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-
-#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_raw
-
-KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_PER
-
-KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
-
-KAZ.IFDnorm_scatter.plot
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-
-```
-
-For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
-
-```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_raw
-
-SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_PER
-
-SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-SKIVE.IFDnorm_scatter.plot
-
-SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-```
-
-In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
-
-```{r MSM.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_raw
-
-MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_PER
-
-#another plot
-MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
-
-MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
-
-MSM.IFDnorm_another.plot
-
-#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
-
-#make a table of the means and n ect
-MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-```
-
-The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
-
-```{r MOLF.IFD, echo=FALSE}
-
-#SHUT DOWN THIS BLOCK UNTIL MOLF FEMALE DATA
-
-#make the dataset
-#MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-
-MOLF.Bivalents.DF <- Curated_BivData %>%  filter(strain == "MOLF")
-
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-#remove the NAs
-#molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
-
-#molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+  geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
-
-#make a table of the means and n ect
-MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
-The other strains which have more overlap between males and females are the Dom strains and KAZ.
-
-#### IFD 3CO bivalents
-
-Run comparisons for 3CO bivalents.
-
-```{r CO3.comparisons, echo=FALSE}
-
-#take a look at the 3CO things
-
-```
-
-<!--
- Heterochiasmy Prediction
-
-> Do the IFD measures lack significant sex effect? (as predicted)
-
-There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
-
-> Are there signiticant strain effects?
-
-The random strain effect was tested for raw and normalized IFD, neither are significant.
--->
-
-# Q1. General CO Positions
-
-<!--
-
-There are a few traits that fall within the CO positions  
-- 1CO normalized position  
-- sis.co.ten (sorta interference)  
-- centromere and telomere distance  
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-  C) Centromere and telomere distances will be sexually dimorphic.
-**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
--->
-
-```{r CO.positions.figs, echo=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-
-```
-
-```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
-
-#make a table for the sex specific values of the normalized F1 positions
-
-#for strain, sex -- Curated_BivData
-#or (Mouse.Table_BivData_4MM)
-#norm.first.foci.pos
-
-Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
-
-                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
-                  n1CO = sum(n1CO )
-
-                          )
-
-```
-
-```{r Q1.Nrm.F1pos_M2, echo=FALSE}
-
-#calq the means for normalized F1
-#M2
-
-Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_nrmF1pos_M2)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-```{r Q1.Nrm.F1pos_M2_long.biv, echo=FALSE}
-
-#table for the 1CO long biv across strain and sexes?
-
-```
-
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#MELT_Curated_BivData
-
-#these isn't long.biv data
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-real.long.bivData
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-```{r CO1_foci_pos_boxplot, echo=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-##### Mixed models norm 1CO positions
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-The mixed model data should only come from 1CO bivalent data.
-
-#### Nrm_1COpos_HetC_M1
-
-```{r Q1.MM_CO_pos, echo=FALSE, eval=FALSE}
-#use the 1CO
-#Use MELT DF
-#MELT_Curated_BivData
-
-#isolate the 1COPos.CO.avs
-#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
-#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
-#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
-
-mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
-
-#<- lme()
-lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
-
-#can't make the 
-CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-#library(lmerTest)
-#anova(CO_position.MM)
-#anova(CO_position.MM, verbose = TRUE)
-
-#summary(CO_position.MM) #no p values,,
-#these just give t and F stats, no pvalue
-
-#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
-#CO_pos_fix.est
-#this test, the full model is more predictive than an empty model?
-
-#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
-#thing2 <- drop1(SC_length.MM, test="Chisq")
-
-CO_pos_rand <- exactRLRT(CO_position.MM)
-CO_pos_rand
-```
-
-```{r Q1.nrmFoci1_M1, echo=FALSE}
-
-#M1
-
-```
-
-the mouse average foci1 pos is more significant in t.test, but not log regression...
-(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-#### Nrm_1COpos_HetC_M2
-
-```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}
-
-#t.test for high and low group  -- logistic regression still need to format the high and low group
-
-#mouse av
-t.test_mouse.av.1CO_pos <- t.test(
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv.1CO_pos_t.test <- t.test(
-
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-  )
-
-```
-
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
-drop1(lme_MM_telo.dist, test="Chisq")
-
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
-### Heterochiasmy Prediction
-
-> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
-
-> is random strtain effect significant?
-
-The random strain effect seems very significant.
-
-remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
-
-```{r pos_t.test}
-#make point plots / boxplots which show differences in mean positions
-
-#scatter + boxplot for t.tests
-```
-
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
-
-```{r F1.by.F2}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-```{r CO.pos.ANOVA, eval=FALSE}
-#what is the pattern of variance
-#run analyses for each subsp*sex
-#use non-melt DF
-
-#how is the variance partioned across
-#cell, mouse, strain
-
-female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
-female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
-
-female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
-
-#unorder strain and mouse
-
-female.Dom$mouse <- as.factor(female.Dom$mouse)
-
-female.Dom$strain <- unclass(female.Dom$strain)
-female.Dom$strain <- as.factor(female.Dom$strain)
-
-female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
-female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
-
-#1CO first
-modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
-
-#can't get mouse and strain to have sum of square
-#residual size decreases with per.F1
-#residuals much larger than fileName, mouse and strain no
-
-#model <- lm(breaks ~ wool * tension, 
-#            data = warpbreaks, 
-#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
-
-male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
-male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
-
-male.Dom$mouse <- as.factor(male.Dom$mouse)
-
-male.Dom$strain <- unclass(male.Dom$strain)
-male.Dom$strain <- as.factor(male.Dom$strain)
-
-male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
-male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
-
-male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
-
-male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
-summary(aov(male.modo))
-
-#only file name is registering as effect
-#Review ANOVA frameworks
-#http://www.biostathandbook.com/nestedanova.html
-
-```
-
-## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
-
-The general predictions across the males and subspecies based on th above MLH1 results.
-
-For positive correlation traits/metrics
-
-i) in DOM strains, low to no difference across strains
-
-ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
-
-ii) in Mol, MSM > MOLF
-
-A) **SC lengths** will be longer for high Rec strains.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
-
-The mouse averages for the other position metrics will be highly influenced by proportions of the 1CO and 2CO bivalents.
-When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
-
-C.2) sis-co-ten metric ... (what about the clustering?)
-
-C.3) telomere and centromere distances ...
-
-M1 glm model for fixed strain effect across male averages.
-
-# Q2 SC
-
-<!--  main DFs for this section
-
-Male.poly.Mouse.Table_BivData_4MM
-
-Bivalent_1o2  (for logistic regression, 1CO and 2COs)
-male.long.biv
-
--->
-
-remember the 3 predictions for subsp;
-
-i) Dom, no difference
-
-ii) Musc, PWD and SKIVE greater than others
-
-iii) Mol, MSM > MOLF
-
-```{r Q2.male.Nrm1CO_strain, echo=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
-```{r Q2.SC.ploy.male, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-
-#1. determine which strain effects are sig
-#use mouse biv average
-#Male.poly.Mouse.Table_BivData_4MM
-
-glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#SC many strains are different (more than the gwRR)
-#subsps are sig
-
-glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#Dom
-glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glm_SC.male_Dom)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glm_SC.male_Musc)
-
-#Mol
-glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glm_SC.male_Mol)
-
-#Logistic regressions
-
-log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-
-                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
-
-#summary(log.SC.male.poly)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_musc)
-#not sig -- when just looking at Musc
-
-#significant in predicting group
-log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_mol)
-#not sig -- when just looking at Mol
-```
-
-```{r Q2.SC.long.biv, echo=FALSE}
-
-#mouse.av.long.biv.table
-
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
-#summary(glm1.SC.male_long.biv)
-
-glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
-#summary(glm2.SC.male_long.biv)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
-#summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
-#summary(glm_SC.male_Mol_long.biv)
-
-## logistic regression for Rec group
-
-log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-              data=mouse.av.long.biv.table,  family=binomial(link="logit"))
-
-#summary(log.SC.male_long.SC)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-```
-
-Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
-
-The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).
-
-#### Q2 SC by class
-
-The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high rec strains then low rec strains.
-This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
-Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
-This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
-They have a lower probability of 2COs below a certain SC length threshold.
-
-I think these comparisons should be done with the **Bivalent observations**
-The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class
-
-```{r Q2.SC.Chrm.Class, echo=FALSE}
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
-#there might not be enough data 
-#
-ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
-```
-
-These are the plots of the logistic regression.
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.blah.mouse.av
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.logReg.mouse.av_1CO
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.logReg.mouse.av_2CO
-
-```
-
-the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
-
-```{r Q2.SClength.t.test, echo=FALSE}
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
-
-```
-
-The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths.
-But When the means are compared across chrm class, the mouse averages are no longer significant.
-
-### Q2 SC length Chrm class prediction
-
-How well does SC length predict chromosome class?
-Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class.
-(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-<!--
-blocks of other male Q2 - SC lengths
-
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
-
-#comparing all the strains
-
-T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-)#p-value = 0.001567
-
-#1CO
-T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#2CO
-T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#3CO
-T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#Single Biv level
-T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
-
-#1CO
-T.test_SC.Len_single.biv_1CO <- t.test(
-  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
-
-      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
-  )
-
-#2CO
-T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
-
-                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
-  )
-
-  #3CO
-  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
-
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
-  )
-```
-
-The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-
-```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
-#I don't think the Mixed models fit well for the high vs low polymorphism tests.
-
-#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
-#MSM_SC vs MOLF_SC
-
-#2. logistic regression for sc length and chrm class?
-#if the prediction / significance
-
-#isolate data
-#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
-
-#build and test MM
-lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-
-lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
-
-anova(lmer_MM_SC_length_male.poly)
-
-rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
-```
-
-```{r long.sc_log.res_meanSC, eval=FALSE}
-#Long Biv comparison
-
-#hold off on reporting/ working with the long SC biv resutls
-
-long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
-long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
-                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
-
-long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
-                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
-
-summary(long.biv.log.reg)#NS effect (likelt underpowered)
-```
-
--->
-
-> **Summary**
-
-When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
-I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
-(IE the scatter plots of SC length, binned by chrm class on the X).
-
-Each point is a mouse average SC.
-When all bivalents are pooled, the logistic regression SC lengths predict the groups.
-When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
-
- and the mean 2COs SC are higher in the high strains
-(this must have something to do with the physical size effect)
-
-```{r male.show.plotsEXXXample, echo=FALSE, eval=FALSE}
-
-#REWRITE THIS FOR SC LENGTH COMPARISONS
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-table( droplevels(fomo.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
-
-#Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-```
-
-```{r SC.length.male_tests, echo=FALSE}
-
-## run the tests to check following expectations
-
-#1.Dom strains not different from ech other
-#2. PWD and SKIVE are different from the other strains
-#3.MSM and MOLF are different from each other
-
-#Male.poly.Mouse.Table_BivData_4MM
-#Long.biv_mouse_table_male.poly
-
-#decide between lm or glm..?
-#mouse avs, males
-
-#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
-
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
-#the subsp labels get messed up, unless they are designated chr first
-Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
-
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
-Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
-#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
-#total sc sig, 1CO sig, 2CO sig
-
-Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
-#0.000369 ***
-Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
-#0.000816 ***
-Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
-#0.0148 *
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
-#making plots above show that mouse average SC lengths are lower could there be microscope issue?
-
-#make sure to remove some of the juvi mice!! from the AP_mouse table
-show.Dom.male.CO <- ggplot(
-male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
-#P not sig for CO mean
-
-#seems like there are more mice in the MLH1 data set than bivData
-#
-
-#WSB all dissected and imaged in '15
-#G imaged 15 - 16
-#LEW imaged 15 to 17
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "1CO SC Length", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "2CO SC Length", xlab = "")
-```
-
-```{r SC.length.pairwise.T, echo=FALSE}
-#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
-
-male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
-male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
-
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
-#PWD and MSM are most diff from SKIVE?
-
-Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
-#the pooled SC value tests are sig
-
-#the mol comparisons are different due to adj p values
-
-Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
-#no sig (only 3) -- does include CZECCH
-
-#non-sig, with mol pooled with musc
-
-Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
-
-Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
-
-Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
-#non of these sig with the mulitple adjust
-
-Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
-
-#pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```
-
-If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
-
-Ideas for the above 3 tests,
-
-1. no difference in DOM
-- no sig logistic regression test
-- glm, strain effect?
-
-> Is sex a significant effect for SC length? (as predicted)
-
-- Sex is a significant effect for SC length
-The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
-
-- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
-
-- The Long Biv Data set largely agrees with the full curated dataset
-
-- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
-
-```{r log.reg_SC.length, echo=FALSE}
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-The mean SC logistic regression model for mouse averages
-
-`r log.reg_SC.length_mouse.av$coefficients`
-
-and for the single bivalent levels
-
-`r log.reg_SC.length_BIV$coefficients`
-
-When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
-When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
-
-> Is the prediction, high rec musc male strains have long SC met?
-
-In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
-I couldn't get the Mixed models working for the male polymorphism predictions...
-
-# Q2 Normalized CO positions
-
-<!-- 
-Main DFs for this section:
-
-Male.poly.Mouse.Table_BivData_4MM
-male.bivdata.1CO
-
-**Polymorphism**
-
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.
-
-C.3) telomere and centromere distances will be shorter in low rec strains
-
--->
-
-Since there aren't many good predictions for how the 1CO normalized landscape relates to gwRR variation, these will be the main questions
-
-- Has there been evolution in the 1CO pattern?
-
-- Is there a consistent pattern with the high rec strains?
-
-Normalized 1CO foci positions will be used. (F1_PER). Check fixed effects of subsp and strains.
-
-sis-co-ten, telomere and centromere distance of foci are metrics which draw from a wider pool of samples. (So maybe I will use them... but avoiding focus on them)
-
-```{r Q2.show.nrmF1, echo=FALSE}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
-
-Q2.show.nrmF1
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= norm.first.foci.pos, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
-
-nrmF1_mouse_level
-```
-
-Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.
-
-```{r Q2.F1.pos.long.biv, echo=FALSE}
-
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Q2.show.nrmF1_long.biv
-
-#2 version of the boxplot data -- not sure why th pattern differd
-
-#more bivData
-#nrmF1_with.SC.length <- ggplot(Curated_BivData_male, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(aes(alpha=.5) )+ geom_boxplot(aes(y= PER_Foci_1, alpha=.5) ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
-```
-
-I think the plot shows that there are fewer 1CO for the High Rec group.
-
-```{r Q2.1CO.landscape, echo=FALSE}
-#Male.poly.Mouse.Table_BivData_4MM
-#Curated_BivData_male
-
-#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#M1
-glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm.M1_Q2.1COnrm)
-
-#M2
-glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm.M2_Q2.1COnrm)
-
-#Dom
-glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glmQ2_M2.1COnrm_Dom)
-
-#Musc
-glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glmQ2_M2.1COnrm_Musc)
-
-#Mol
-glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glmQ2_M2.1COnrm_Mol)
-
-#m3 (log)
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm_Musc)
-
-#Mol
-Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
-              family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-#biv level logistic regression
-#Curated_BivData_male
-#male.bivdata.1CO
-
-male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
-
-Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.Q2.F1_PER_Musc)
-
-#Mol
-Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
-              family=binomial(link="logit"))
-
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-```
-
-```{r Q2.t.test_COlandscape,echo=FALSE}
-
-#MOUSE AVE
-Rec.group.mouse.F1pos_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.F1pos_Ttest$p.value
-#p-value = 0.9098
-
-#single biv level
-Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-Rec.group_BivData_F1post_Ttest$p.value 
-#p-value = 0.238
-
-```
-
-The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
-
-#### Summary
-
-Not really a clear summary for the normalized 1CO position and rec groups.
-
-# Q2 IFD
-
-this is the Q2 prediction for IFD / interference
-
-> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-1. glm across subsp
-
-2. anova within subsp
-
-3. logistic regression for musc (or all males)
-
-```{r Q2.show.IFD, echo=FALSE}
-
-#subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
-
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-biv.IFDs_PER_by.group
-```
-
-The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
- The normalized plot displays the pattern of higher rec group having slightly higher longer IFDs.
-
-```{r Q2.IFD.glm.test, echo=FALSE}
-#glm --
-#mouse aver
-
-#M1.ABS
-glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.raw) 
-#M1.PER
-glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.PER)
-
-#M2.raw
-glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.1COnrm)
-
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD_Mol)
-
-#M2.PER
-glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.IFD.PER)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD.PER_Mol)
-```
-
-```{r Q2IFD_logreg, echo=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.PER, echo=FALSE}
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
-
-#t.test for IFD
-#mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-t.test_mouse.av.IFD_PER <- t.test(
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-```
-
-```{r Q2.t.test_IFD,echo=FALSE}
-
-#already made
-#subset
-male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-#t.tests for the Rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
-                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
-
-Rec.group.BivData_IFD.PER_Ttest$p.value
-#p-value < 2.2e-16
-
-#MOUSE AVE
-Rec.group_IFD.PER_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group_IFD.PER_Ttest$p.value 
-#p-value = 6.872e-07
-
-```
-
-The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
-
-### Q2 Long Bivalent Dataset
-
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.
-
-```{r Q2.long_biv.IFD, echo=FALSE}
-#long BivData
-male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
-
-#show
-Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-
-Long.biv.IFDs_PER_by.group
-
-```
-
-table(droplevels(male.long.biv.2CO_IFD$strain) )
-
-The plot above shows the normalized IFD measures from 2CO bivalents
-
-There are many fewer long bivalent data observations, but the same positive correlation with
-
--table of long bivalents
-
-- glms
-
-- logistic regression
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-
-```
-
-There are not enough observations for sub setting the data into strains or use the mouse averages.
-
-Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
-The above t.tests are breaking the knitr
-
-The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
-
-None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
-
-Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
-This also indicated that interference distance may evolved in the house mouse complex.
-
-#Caveats
-
-put all of the code chunks/analysis for caveats here
-
-## Chromosome Size Effect
-
-I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)
-
-<!-- 
-Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
-But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
---> 
-Below are examples of plots of SC length distributions across cells.
-The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
-Most plots excluded for space.
-
-Each point is a bivalent plotted by cell on the x axis.
-X's are the 4th quartile, big point is the mean and smaller black point is the median.
-I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
-(the extra stats are not correctly mapped)
-
-```{r SC.distb_manual_Auto, echo=FALSE}
-
-#calq the mouse means
-manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
-                        median_SC = median(SC.length, na.rm = TRUE),
-                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
-#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
-
-)
-
-manual_summary_stats <- add_mouse(manual_summary_stats)
-manual_summary_stats <- add_strain(manual_summary_stats)
-manual_summary_stats <- add_sex(manual_summary_stats)
-manual_summary_stats <- add_subsp(manual_summary_stats)
-
-###
-#calq the mouse means
-Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        median_SC = median(chromosomeLength),
-                        Qar.4 = quantile(chromosomeLength)[[4]],
-
-                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
-        #add Q4
-
-)
-Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
-
-#colors_of_strains
-#scale_color_manual(values=colors_of_strains)
-
-#un-order these factors
-```
-
-```{r long.biv.plots, echo=FALSE}
-
-show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
-
-show.the.cells_females_dom
-#PWD seems to have the biggest variance across cells
-
-almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_DOM_f
-
-#### Musc female
-
-show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
-
-almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
-
-### Dom Male
-show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
-
-almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#MUSC male
-show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
-
-almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
-
-#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
- # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
-#show.the.cells_male
-
-#almost_whole_cell_BivData
-#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
-#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
-#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
-
-```
-
-All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
-It shows the amount of within cell variance across strains.
-There is a bit of variance across the SC length distributions in the PWD females.
-
-This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
-
-The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
-The full data set is `r length(Curated_BivData$Obj.ID)`.
-This is the breakdown of bivalent observations by category for this long dataset are `r data.frame(table(droplevels(real.long.bivData$category)))`.
-
-**- Try to merge this DF with the whole.cell manual measures.**
-
-- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.
-
-in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
-10202 bivalents from 86 mice.
-
-```{r show.long.SC, echo=FALSE}
-
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
-female_long_biv_plot
-
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
-male_long_biv_plot
-
-```
-
-These plots show the SC lengths for the 'long SC data set'.
-They are supposed to be the longest 4-5 SC from cells where I could get good measures.
-These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
-Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
-
-## Adjusting for XX
-
-(rethink where this section should go)
-new outline
-
-1. illustrate problem(affects mostly SC length)
-
-2. Expected impact on sex comparisons, estimated effect size of the X  
-3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
-
-4. 19 female, 19male, 20female 20 male
-
-The female mouse averages should have adjustments for the XX.
-working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
-Subtract this amount from the female mouse averages ...
-This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
-
-- Of all female single bivalents observations, 5% are XX (1 of 20).
-
-- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
-
-- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
-
-I think a formula something like this can be applied to adjust for XX
-
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
-
-#### Whole Cell Manual Comparison
-
-```{r XX.adj_manual.whole.cell, echo=FALSE}
-
-#load and examine the the whole cell hand measures here
-
-female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
-data.frame(table(droplevels(female.whole.cell$category)))
-
-# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
-
-#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
-
-data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
-#missing some strains, most have 3 cells (min). MSM has 5 full cells
-
-#remove the incomplete LEW female cell
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
-
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
-
-#calq mean SC lengths by rank
-
-mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
-
-                       mean.sc =   mean(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                         se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
-
-top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
-
-show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
-
-show.SC_by.rank_top
-
-bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
-
-show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
-
-#CALQ PERCENT TOTAL SC
-
-CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
-
-                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       #rank.by.per
-
-                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
-                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
-                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
-                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
-                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
-                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-
-#MELT ABOVE
-MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
-              #leave out the Foci
-               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
-
-            id=c( "fileName","total.sc.cell","nbiv"))
-
-MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
-
-MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
-MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
-
-show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
-
-show_long.sc.proportions
-
-#original plot code 
-#ggplot(data = df1, aes(x, y, color = Group)) +
-#    geom_pointrange(aes(ymin = lb, ymax = ub), 
-#                    position=position_jitter(width=0.5), 
-#                    linetype='dotted') +
-#    theme_bw()
-```
-
-The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
-
-The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
-
-The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
-For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
-
-(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
-
-The other figure shows of each single bivalent contributes to the total SC area.
-Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
-on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
-So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
-
-- Is the difference between cell averages for males and females less that 10%?
-
-- also interesting, the pwd and msm don't have longer SC, compared to other strains.
-
-#### Automated BivData Comparison
-
-```{r XX.Automated.biv.stuff, echo=FALSE}
-
-#make automated female data to compare to
-female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
-
-female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
-
-table(female.Bivdata_comp$strain)
-
-#female.Bivdata_comp
-
-#1. calq the rate of chrm segmentation in female categories in automateted data
-
-box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
-
-                   Nmice = length(unique(mouse)),
-                   mx.box.num = max(boxNumber), # need to run this at the cell level
-                    N.biv  = length(Obj.ID),
-                   pass.biv = N.biv / mx.box.num
-)
-box.rates <- add_mouse(box.rates)
-box.rates <- add_strain(box.rates)
-box.rates <- add_category(box.rates)
-
-cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-
-cell.pass.rates
-
-category.box.rates <- ddply(box.rates, c("category"), summarise,
-
-                   mean.pass.rate = mean(pass.biv)
-)
-#category.box.rates
-
-#2. pull the 3rd rank biv's calculate the SC means
-#calq means by rank (focus on the top 5)
-
-```
-
-For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
-
-The table above shows the number of bivalents from the same strains as in the manual whole cell data.
-The plot shows the bivalent passing rate across all of the individual cells from this female data set.
-For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
-
-(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
-(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
-
-strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
-
-** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
-But I am not using 'whole cell' summaries to compare female and males.
-
-**What is the effect of an extra XX-autosome on single bivalent means?**
-
-use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
-Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
-Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
-Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
-The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
-
-# Trash / Extra
-
-<!--
-
 ```{r Q2.IFD.log.reg, echo=FALSE, eval=FALSE}

### 2020-02-12 doc/Results.rmd
more updates and knit

@@
 
-```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=FALSE}
+```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=TRUE}
 #DF.HetC.MixedModel.HQ
@@
 
-*Do I believe the physical/SC results?*
-
 <!--
@@
 -->
-
-3) Interference / IFD will not be sexually dimorphic.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.
-
-**I**nter**f**ocal **D**istance, a indication of CO interference.
-
-The raw measures are pretty nosy, this might not be the best way to display these results.
-
- Each point is a IFD observation. Maybe add mouse level lines for the means.
 
@@
 
-### IFD_HetC_M1
-
-```{r HetC_M1_IFD_many.metrics, echo=FALSE}
+```{r Q1.IFD.M1, echo=FALSE}
 #get a list of the old mice to double check
@@
 #M1_IFD_2CO
+lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)
+anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)
+
+drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )
+
+exactRLRT(lmer.IFD.ABS_MM.M1)
+
+#M1_IFD_2CO
 lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
+anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
+
 #these 2 below are LR tests
-#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-#anova(lmer.IFD.PER_MM.M1)
-
-#exactRLRT(lmer.IFD.PER_MM.M1)
-#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))
-
-#M1_IFD.RAW_2CO
-lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
-#anova(lmer.IFD.RAW_MM.M1) # also F values
-
-#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
-#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-```
-
-Sex is a slightly significant factor for raw and normalized IFD.
+drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+exactRLRT(lmer.IFD.PER_MM.M1)
+
+```
 
@@
 That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
-
-show a raw plot, and norm plot and a scatter plot
 
@@
 
-### IFD_HetC_M2
+$$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
 
@@
 
-```
-
-### IFD_HetC_M3
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+
+summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
+#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+
+summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
+#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
 
@@
 
-```
+glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
+
+summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
+round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
+
+write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
+round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
+
+write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+
+```
+
+For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.
+
+the interaction effects were slightly significant for both raw and nrm.IFD.
+
+The most significant value was the sex effect for nrm.IFDs.
+
+The the random strain effect was not significant for either model.
+
+For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
+The other effects are for PWD ansD SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
+For the M3, only SKIVE*male effect is close to significant
+
+For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
+SKIVE*male is the only other consistantly significant effect, which also increases the nrm.IFD measure.
+
+Overall There's a low amount of significant effects across the 2CO IFD measures.
+This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
 
@@
 
+density plots
 ```{r CO.positions.figs, echo=FALSE, include=FALSE}
@@
 
-```{r show.long.biv, echo=FALSE}
-#real.long.bivData
-
-```
-
 ```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
@@
 
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.nrmF1CO.pos_M1, echo=FALSE}
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
+anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)
+
+drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )
+
+exactRLRT(lmer.Q1.nrm1CO_M1)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
+
 ```{r Q1.Nrm.F1pos_M2, echo=FALSE}
@@
 
-Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
+Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Mouse.Table_BivData_4MM)
 #summary(Q1.glm_nrmF1pos_M2)
+
+summry.Q1.glm_nrmF1pos_M2 <- summary(Q1.glm_nrmF1pos_M2)
+round(summry.Q1.glm_nrmF1pos_M2$coefficients[,c(1,4)],5)
+
+write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", sep=",",row.names = TRUE)
 
@@
 
-```{r Q1.Nrm.F1pos_M2_long.biv, echo=FALSE}
-
-#table for the 1CO long biv across strain and sexes?
-
-```
+$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.nrmCO.pos_M3, echo=FALSE, eval=FALSE}
+
+Q1.glm_nrmF1pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_nrmF1pos_M2)
+
+summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
+round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)
+
+write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+```
+
+The model results aren't as clear as I'd like.
+For the Mol, both the sex effect and MOLF strain effect are significant (this means both female and male rec landscapes are effected - most of the time towards the middle
+
+GENERALLY - Sex is the biggest effect for the 1CO landscape (female middle male telomereic)
+molf
+Not sure I understand how MSM is supper significant for 
+There must be some strange effect due to the Mol subsp - MOLF isn't even in M2 (
+Male is consistently significant across the two models.
+
+Male and MOLF are the most significant effects for M3
 
@@
 
-##### Mixed models norm 1CO positions
-
 the distribution of SC lengths and sis-coten seems very different across sexes
 
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
 The mixed model data should only come from 1CO bivalent data.
-
-#### Nrm_1COpos_HetC_M1
-
-```{r Q1.MM_CO_pos, echo=FALSE, eval=FALSE}
-#use the 1CO
-#Use MELT DF
-#MELT_Curated_BivData
-
-#isolate the 1COPos.CO.avs
-#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
-#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
-#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
-#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )
-
-mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]
-
-#<- lme()
-lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-
-as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )
-
-#can't make the 
-CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-
-#library(lmerTest)
-#anova(CO_position.MM)
-#anova(CO_position.MM, verbose = TRUE)
-
-#summary(CO_position.MM) #no p values,,
-#these just give t and F stats, no pvalue
-
-#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
-#CO_pos_fix.est
-#this test, the full model is more predictive than an empty model?
-
-#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
-#thing2 <- drop1(SC_length.MM, test="Chisq")
-
-CO_pos_rand <- exactRLRT(CO_position.MM)
-CO_pos_rand
-```
-
-```{r Q1.nrmFoci1_M1, echo=FALSE}
-
-#M1
-
-```
 
@@
 (is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-#### Nrm_1COpos_HetC_M2
 
@@
 
-```{r var.contrast, echo=FALSE, eval=FALSE}
-#contrast models -- what's the points for this block
-# review this logic again
-full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997
-
-full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012
-
-strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
-#for var, p = 0.0219
-```
-
 ```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}

### 2020-02-12 doc/Results.rmd
mostly finished with pp

@@
 
-#### Q1. SC_HetC_M2
-
 All fixed and all interactions
@@
 
+```{r Q1.SC.show.long.biv, echo=FALSE}
+
+mouse.shoo.long.biv <- ggplot(real.long.bivData, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+mouse.shoo.long.biv
+
+```
+
+```{r Q2.SC.long.biv, echo=FALSE}
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+Q1.long.biv_cell_table <- ddply(real.long.bivData, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+Q1.long.biv_cell_table <- add_mouse(Q1.long.biv_cell_table)
+Q1.long.biv_cell_table <- add_strain(Q1.long.biv_cell_table)
+Q1.long.biv_cell_table <- add_sex(Q1.long.biv_cell_table)
+Q1.long.biv_cell_table <- add_subsp(Q1.long.biv_cell_table)
+
+Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
+Q1_cell_long.biv
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+cell.avlong.biv
+
+#mouse av for long biv
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+mouse.shoo.long.biv
+
+```
+
 #### XX adjustment
@@
 
-<!--  main DFs for this section
-
-Male.poly.Mouse.Table_BivData_4MM
-
-Bivalent_1o2  (for logistic regression, 1CO and 2COs)
-male.long.biv
-
--->
-
-remember the 3 predictions for subsp;
-
-i) Dom, no difference
-
-ii) Musc, PWD and SKIVE greater than others
-
-iii) Mol, MSM > MOLF
+Focus on comparing the differences between long.biv (and short biv once I make that dataset)
 
@@
 ```
+
+```{r Q2.SC.long.biv, echo=FALSE}
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+shoo.long.biv
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+cell.avlong.biv
+
+#mouse av for long biv
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+mouse.shoo.long.biv
+
+```
+
+Not the pattern I expected, there might be more noise than I thought in this dataset.
 

### 2020-02-12 src/BivData_Setup.Rmd
coded in Quartille2 and short BivData set

@@
                       mean_cent_dist = mean(dis.cent, na.rm=TRUE),
-                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]] #calq threshold for all cells, pull out 'almost.whole.cells' later
-
-                        )
-
-#add the centromere and ifd here?
-                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
-#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
-#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
-#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),
-
-#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
-#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
- #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),
-
-#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
-#                       var.IFD = var(IFD1, na.rm=TRUE),
-#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
-#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),
-
- #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
-#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
-#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
-#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)
+                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]], #calq threshold for all cells, pull out 'almost.whole.cells' later
+                      Qar.3 = quantile(chromosomeLength, na.rm = TRUE)[[3]],
+                      Qar.2 = quantile(chromosomeLength, na.rm = TRUE)[[2]],
+                      Qar.1 = quantile(chromosomeLength, na.rm = TRUE)[[1]]
+
+                        )
 
@@
 
+#use quartille 2 as a min, on average it grabs 4
+
 ```
@@
 #check if there are fileNames
-double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% Curated_BivData$fileName, ]
+#double.filename <- whole.cell.data_manual[ whole.cell.data_manual$fileName %in% #Curated_BivData$fileName, ]
 
@@
 
-missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ]#0, 
+missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ] 
 #all these files stem from
@@
 #mini.now  <- BivData_from_15biv_cells[1:300,]
-#broken again -- confusion between the number Q4 and the fileName
 
@@
 
-  #print(c(i, " the i") )
-  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
+  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_" )[[1]][1:7]
   new.egg <- paste0(newfile.str, collapse = "_")
 
-  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
-  #there is no Qar4 in this DF
-
-  #print(c("the Q4  ",the.Q4)  )
-
-  #the above can produe mulitple Q4
+  the.Q4 <-cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+   the.Q3 <-cell.level.BivData_table$Qar.3[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+  the.Q2 <-cell.level.BivData_table$Qar.2[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
+  the.Q1 <-cell.level.BivData_table$Qar.1[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
 
   BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
-  #these are just using the last index,
-  #this might be wrong....
+    BivData_from_15biv_cells$cell.Q3[i] <- the.Q3
+  BivData_from_15biv_cells$cell.Q2[i] <- the.Q2
+  BivData_from_15biv_cells$cell.Q1[i] <- the.Q1
 
@@
 
-  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)  #this FN is NOT RIGHT
-   # print(c(i, the.FN, the.Q4))
+  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)
+  BivData_from_15biv_cells$Q1.FN[i] <- as.character(the.FN)
 }
@@
 real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]#713 observations
+
+#isolate the short bivdata
+BivData_from_15biv_cells$short.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength <= as.numeric(BivData_from_15biv_cells$cell.Q2)  ), "yes", "no" )
+
+real.short.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$short.Biv == "yes",]# 708 much few than
+
+```
+
+```{r short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}
+
+#notes for short biv
+
+mini_G.15_15biv_FN <- BivData_from_15biv_cells %>% filter(category == "G male")
+#ploting short biv data
+G.15_15biv_male <- BivData_from_15biv_cells %>% filter(category == "G male")
+mini_G.15_15biv_FN <- sample( G.15_15biv_male$fileName, 10)
+mini_G.15_15biv  <- BivData_from_15biv_cells[BivData_from_15biv_cells$fileName %in% mini_G.15_15biv_FN, ]
+
+ppll <- ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
+  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
+  geom_hline(aes(yintercept= cell.Q2), color='red') +
+ geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
+ geom_hline(aes(yintercept= cell.Q4), color='black')
+
+  #this plot shows that the Q1 is capturing only single values -- I should try Q2
+#use Q2 as a 
 ```
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_11.2.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_12.2.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-02-12 doc/Results.rmd
encountering bugs when knitting

@@
 
-load(file="~./MLH1repo/data/CleanBivData_11.2.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_12.2.20.RData")
 
@@
 
-```{r Q1.SC.show.long.biv, echo=FALSE}
-
-mouse.shoo.long.biv <- ggplot(real.long.bivData, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+```{r Q1.SC.show.long.biv, echo=FALSE, eval=FALSE}
+
+mouse.shoo.long.biv <- ggplot(real.long.bivData, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
 \nMale Long Biv SC Lengths")
@@
 
-```{r Q2.SC.long.biv, echo=FALSE}
+```{r Q2.SC.long.biv_wacko, echo=FALSE}
 male.long.biv <- real.long.bivData %>% filter(sex== "male")
@@
 
-```{r Q2.SC.long.biv, echo=FALSE}
+```{r Q2.SC.long.biv_2, echo=FALSE}
 male.long.biv <- real.long.bivData %>% filter(sex== "male")
@@
 
-```{r Q2.SC.long.biv, echo=FALSE}
+```{r Q2.SC.long.biv_3, echo=FALSE}
 
@@
 
-#mouse.av.long.biv.table
+#mouse.av.long.biv.table <- make this is set up
 

### 2020-02-13 doc/Results.rmd
cleaning up more

@@
   pdf_document: default
+  word_document:
+    toc: yes
+    toc_depth: '5'
 bibliography: refs_Main_Report.bib
@@
 
+options(digits=2)
+
 setwd("~.")
@@
 
+```{r misc.setup, echo=FALSE}
+
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blues
+
+                       'PWD'= 'red2',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
+                       'CZECH'= 'indianred4',
+
+                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
+
+                    'CAST'='green','HMI'= 'forestgreen', #greens
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
+
+#add Rec group,
+#reclassified SKIVE as intermediate
+Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
+#apply this before ploting, to display strains in desired order
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, category, mouse) %>%
+#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
+#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
+
+# sort your dataframe, by the focal categories
+#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
+```
+
+```{r FEMALE.TABLE, echo=FALSE}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+#Mol
+#Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
+
+#Dom 
+Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
+
+```
+
+```{r MALE.TABLE, echo=FALSE}
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
+fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
+fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
+fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
+
+Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+#Dom version 
+Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+```
+
+```{r MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
+#DF.HetC.MixedModel.HQ
+
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+
+#What are the main DFs for this code chunk
+#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
+#DF.HetC.MixedModel   <- only Mmusculus
+#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
+
+#variance is still chr
+AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+
+#use the merged DF to filter by age 
+AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+#main file, make everything unordered factors
+AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+#remove the extra merged cols
+DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+#DONT NEED CATEGORY
+
+#REMOVE EXTRA STRAINS
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#remove PERC, MOLF, CZECH, (no females)
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+
+#main df with variables coded as chr
+DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+
+#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+
+all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
+                 .(sex, strain),
+                 summarize,
+                 all.cells = sum(Ncells)
+)
+
+```
+
 # Main Figure
-
-```{r first.boxplots, echo=FALSE, fig.align='center', warning=FALSE}
-
-#the below ordering works!! keep it
-MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
-              "PWD","CZECH","SKIVE",
-              "KAZ","TOM","AST",
-              "MSM","MOLF",
-              "CAST","HMI","SPRET","SPIC", "other"), order=T )
-#order the data frame
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
-  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank()) +
- ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
-#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+
-
-justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")
-
-#scales="free_y"
-#this is ok except for the set placement of mice on x axis
-#if the full dataset was ordered correctly, this would kinda fix it
-
-#boxplots by strain
-ss <- ggplot(MLH1_data, aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain )) + geom_boxplot(aes(alpha=0.5))  + ggtitle("Boxplots of MLH1 distributions by strain") + theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+
-  ylim( c( (min(MLH1_data$adj_nMLH1.foci)-3), (max(MLH1_data$adj_nMLH1.foci)+5)) )+scale_fill_manual(values=colors_of_strains)
-
-ss <- ss   + facet_wrap(~ sex, scales="free")
-
-#these plot currently look bad
-```
 
@@
 #this move this to the setup file
-
-strain.table.HQ_Fig.og <- ddply(MLH1_data, c("strain", "sex"), summarise,
+#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed
+
+#pooled
+strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                         Nmice =length(unique(mouse)),
@@
                         se  = round(sd / sqrt(Ncells), 3)
+
+                       #range_mouse_means = range()
+                   #consider adding crhm proportions?
     )
@@
 
-```{r maw.fig1s, echo=FALSE, include=FALSE}
+```{r main.fig1s, echo=FALSE, include=FALSE}
 
@@
 
-#table for the means by strain
-
-```
-
-## Female Specific Analysis
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.female.plot
-
-#Mol
-Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
-
-Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.female.plot
-
-#Dom version 
-Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
-
-Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.female.plot
-#Gough female is higher
-
-```
-
-```{r female.strain.table, echo=FALSE}
-
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  = sum(Ncells),
-                        mean_co =as.numeric(format(round(mean(mean_co), 3 ), nsmall=3) ),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se)
-    )
-
-```
-
-G and LEW have a higher strain means compared to WSB. MSM (and PWD) also has an elevated mean.
-
-The variance across mouse means for each strain is also variable across strains.
-
-In the dom  subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in WSB, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in G, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in LEW.
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in PWD, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in KAZ.
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in MSM.
-
-Two glm models were run to test the effects of strain and subspecies.
-
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- lm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-
-round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-lm.female.strain_sum <- summary(lm.female.strain)
-
-round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-
-#how much greater is G female
-G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co)
-
-LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
-mean.LEW.female <- mean(LEW_female$mean_co)
-
-#how much greater is MSM female
-MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSM.female <- mean(MSM_female$mean_co)
-
-sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
-mean.all.other <- mean(sans_high_female$mean_co)
-
-#sans_MSM_female / mean.all.other
-
-## within mouse variance models 
-lmVAR.female.strain.subsp <- lm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- lm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- lm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-#LEW not sig
-
-lm.m2.Q12_female <- lm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
-#LEW sig
-```
-
-Above are the coefficents for the two glm's of the female specific mouse MLH1 averages, which include subspecies and strain as fixed effects.
-G has the most consistant significant strain effects in both models.
-MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.
-
-G is `r mean.G.female / mean.all.other` higher the the other means.
-LEW is `r mean.LEW.female / mean.all.other` higher  and MSM is `r mean.MSM.female / mean.all.other`.
-These three will be designated as 'moderate high rec' strains for later sex specific analysis.
-
-## Male Specific Analysis
-
-The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
-
-#Dom version 
-Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-
-```
-
-- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
-
-- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
-
-- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
-
-```{r male.strain.table, echo=FALSE}
-
-#female strain table
-#female_mouse.avs_for.analysis
-M.strain.table <- ddply(just.male_mouse.av, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  = sum(Ncells),
-                        mean_co =as.numeric(format(round(mean(mean_co), 3 ), nsmall=3) ),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se)
-    )
-
-```
-
-Two models were fitted to the male specific data to test the effects of strain and subspecies.
-
-```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-summry.male.strain.effects <- summary(glm.male.strain.subsp)
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#1. determine which strain effects are sig
-
-#M2
-lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
-lm.male.strain_sum<- summary(lm.male.strain)
-
-round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /
-
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#all p values kinda high, MSM and TOM are slightly sig
-
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#NS
-
-```
-
-The strain average for PWD, MSM and SKIVE are  ` r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
-Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
+#strain.table.HQ_Fig.og  all strains
+#strain.table.HQ_Fig  extra strains removed
+
+main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
+
+main.table <- add_subsp(main.table)
+
+kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
+
+#strain.table.HQ_Fig
+
+#sum( (main.table  %>% filter(sex == "male"))$Ncells)
+#the kable version is bulkier than the other versaion
+```
+
+In order to estimate the genome wide recombination rates across our panel of `r length(unique(main.table$strain))` strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)`.
+
+The strain means for females were pretty similar.
+For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _M.
+m.
+domesticus^WSB^_, _M.
+m.
+domesticus^G^_, and _M.
+m.
+domesticus$^LEW^_ respectively.
+In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _M.
+m.
+musculus^ PWD}^_, _M.
+m.
+musculus^ SKIVE^_, and _M.
+m.
+musculus^KAZ^_ respectively.
+In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _M.
+m.
+molossinus^ MSM^_ and _M.
+m.
+molossinus^MOLF^_ respectively.
+
+While in two of the subspecies, the means for males show a greater range across means, notable in the mmmolossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _M.
+m.
+molossinus^MSM^_ and  _M.
+m.
+molossinus^MOLF^_ respectively.
+In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _M.
+m.
+musculus^PWD^_, _M.
+m.
+musculus^SKIVE^_, _M.
+m.
+musculus^KAZ^_, _M.
+m.
+musculus^TOM^_, _M.
+m.
+musculus^AST^_, and _M.
+m.
+musculus^CZECH^_ respectively.
+In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _M.
+m.
+domesticus^WSB^_, _M.
+m.
+domesticus^G^_, _M.
+m.
+domesticus^LEW^_ and _M.
+m.
+domesticus^PERC^_ respectively.
+These
+
+<!-- mouse level means-->
+The ranges of mean MLH1 counts per cell are
+
+<!--(the figure is made from pooling MLH1 cells)
+the male
+
+- the ranges of mouse means are displayed in Sup table X
+
+- How the means were calculated
+
+-The male gwRR is adjusted for the PAR
+-The means for each strains are report in Table 1
+
+-Normality was assessed for the MLH1 counts (include in Supplement histograms)
+
+, -->
 
@@
 
-In order to understand the variance within an evolutionary framework we fit the mouse gwRRs to a mixed model, which specified subspecies, sex and the interaction as fixed effects.
-The subspecies effect - is a proxy for measuring the degree of divergence across subspecies.
-The strains were coded as a random effect to approximate standing genetic variation across subspecies ranges.
-
-These are the basic Mixed Models we used as a frame work to analyze all the MLH1 count data.
-
-$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-The mixed model lets us separate effects to see how they effect a mouse mean CO count. The subspecies effect is a proxy for **Divergence** and the random strain effect is a proxy for **Polymorphism**.
+In order to test the effects of subspecies, sex and strain on the mean and variance of MLH1 counts per cell, we fit a data set of `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages for MLH1 foci per cell.
+We constructed the model to set subspecies, sex and their interaction as fixed effects, which strain was coded as a random effect.
+The predicted coeffecients and pvalues are reporter below in table X.
+
+$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
 <!--
+This is the break down of mice used within the Mixed model
+
+`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
+
+add number of cells
+
+`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
+
+ $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+In order to understand the variance within an evolutionary framework, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse level MLH1 foci per cell means as the dependant variable with a mixed model.
+
 Setting up the data set for using the mixed model (base data set, HQ data set)
@@
 subspecies levels: 3, Musc, Dom and Molossinus
-
 strain levels: Dom: 3, Musc: 3  Mol: 2
-
 sex: 2 female male  
 -->
-
-```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=TRUE}
-#DF.HetC.MixedModel.HQ
-
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-
-#What are the main DFs for this code chunk
-#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
-#DF.HetC.MixedModel   <- only Mmusculus
-#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
-
-#variance is still chr
-AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
-
-#use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#remove the extra merged cols
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-#DONT NEED CATEGORY
-
-#REMOVE EXTRA STRAINS
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
-
-#SWITCH TO HQ.DF
-#mean_co.HQ  (this might remove some of the MOLF males)
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]  #this removes the MOLF female
-
-#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
-
-#afriad this is removing NA's
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
-
-#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
-
-#add_subsp function re-writen
-
-#mean_co.HQ.mol
-#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ$strain), "Mol", 
-#        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ$strain), "Mol", DF.HetC.MixedModel.HQ$subsp))
-
-# For some reason I changed subsp to nums (wtf)
-#DF.HetC.MixedModel.HQ$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ$subsp), "Dom", 
-#        ifelse(grepl(2, DF.HetC.MixedModel.HQ$subsp), "Musc", DF.HetC.MixedModel.HQ$subsp))
-
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | #DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#remove blank rows (which are these?)
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
-#151 mice, no rows with nas for values
-
-#main df with variables coded as chr
-DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
-
-all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
-                 .(sex, strain),
-                 summarize,
-                 all.cells = sum(Ncells)
-)
-
-```
-
-This is the break down of mice used within the Mixed model
-
-`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
-
-add number of cells
-
-`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
-
- $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-Allows random effect from strain. Considers the wild derived inbred strains as random samples from each subspecies territory.
 
@@
 
-#summary(MLH1.MM.M1)
-
-MM.Fixed.pval <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-
-anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)
-anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)
-
-MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-#as.data.frame(round(summary(MLH1.MM.M1)$coefficients, 6))
-
-#only the interaction term shows up, when the 2 terms are additive, they both show up in the drop1 results
-
-```
-
-The significance of the interaction effect for the 2 fixed effects is `r MM.Fixed.pval`. The random effect for strain effect is `r MM.rand.pval`.
-
-### glm followup
-
-To follow up on the mixed model results, we also ran glm / linear models to get a better idea of the strain by sex interaction effects.
-
- $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$
-
-Post hoc model for investigating strain as a fixed effect.
+summary(MLH1.MM.M1)$coefficients
+
+MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+
+MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
+sex.results =  MLH1.MM_sex, 
+               interaction.pval = MLH1.MM_interaction,  rand.pval = 
+MLH1.MM_strain)
+
+```
+
+Because all of the mixed model effects had significant effects, (p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r  round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect), we followed up with two general linear models to investigate the specific strain effects.
+
+ $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
+
+and
+
+$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
 
@@
 #table of just coeff anf pvalue
-round(summ$coefficients[,c(1,4)],5 )
+#round(summ$coefficients[,c(1,4)],5 )
 
 #write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-
-```
-
-Above is the table of coefficients for MM.3, HQ data set, lm function, sex, strain and subsp are interacting fixed effects.
-
-Strain G is significant (p=0.0005), (remember female result)
-
-interaction terms of sex and strain for (G, MSM, PWD).
-
-How are these models different from the last one -- MM4 is missing subsp, so only sex * strain effects.
-
-$$mouse \ av \ metric ~=~ sex * strain + \varepsilon$$
+```
+
+In the first glm, the fixed subspecies effects were not significant, so we only report for values for the second model which is limited to sex and strain fixed effects.
 
@@
 #write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-
-#as.data.frame( round( summary(strain.fixed)$coefficients, 6))
-
-#summary(strain.fixed)
-#maybe this WSB effec is just in the HQ data set?
-```
-
-<!--
- Contrasts and ANOVA, mean CO count full vs reduced
-
-Use Anova to compare full vs reduced models
-the reduced models will be those missing either, strain, sex, or subsp.
-
-```{r anova.full.v.reduced, echo=FALSE, eval=FALSE}
-
-#strain as fixed
-#CONTRAST models
-#newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )
-
-HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
-R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male
-
-full.Reduced.HetC.MixedModel.mean_co <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel)
-
-#base vs strain
-fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co) #<.0001
-
-#show scientific vs normal
-#vv <- fullVstrain$`p-value`
-#round(vv[2], 8)
-#format(vv[2], scientific=F)
-
-#strain vs no random full.reduce
-fullVnon <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
-
-strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001
-
-#fullreduce vs strain
-#can't distinguish which random effects are more significant -- both tests are significant
-
-fullVreduce22 <- anova(HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)
-
-#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
-full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant
-
-full.v.fixed <- anova(HetC.MixedModel.HQ, strain.fixed)
-#this comparison not useful when there are different fixed effects
-
-#full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ) # < .0001
-
-#strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ) #NS
-
-#anova tests
-#full vs strain
-full2strain.mol <- anova(HetC.MixedModel.mol, Reduced.strain.mol)# < .0001
-#full vs non
-full2noRandom.mol <- anova(HetC.MixedModel.mol, Reduced.all.mol)# < .0001
-#strain vs non
-strain2noRandom.mol <- anova(Reduced.strain.mol, Reduced.all.mol)# < .0001
-
-anova(Reduced.strain.mol, Reduced.all.mol, HetC.MixedModel.mol)
-
-#base vs mol model (this only makes sense if these are nested models)
-anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
-```
-
--->
-
-## Summary of Mixed Model Analyses
-
-**mouse average MLH1 count**
-
-Multiple models were tested to test for potential evo models -- for the mouse average CO counts.  (there were a mix of results -- and nuanced)
-
-First full Mixed model -- no significant fixed factors --(the random effect might be significant).
+```
+
+In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _M.
+m.
+musculus^PWD^_ and _M.
+m.
+molossinus^MSM^_  respectively.
+These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.
+
+Additionally, _M.
+m.
+domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
+These results confirm the qualtative pattern of the _M.
+m.
+domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
 
@@
 
-**Variance effect**
-
-** Do these patterns hold across Quality bins? **
-
 # Within Mouse Variance in CO Count per Cell
 
-<!-- intro
-in addition to the mean gwRR, the variance is also an important metric to describe
+In order to more fully discribe the distributions we examined the variance in addition to the mean applying the same mixed models and linear models, with _within mouse variance_ of MLH1 counts per cell as the dependant variable.
+
+<!--
+motivation / background: maybe checkpoint / regulation
+
+limitations: observations at the mouse level, decreased power to estiamate the variance --
+
+setup
+
+-quality / technical error
+
+- variance and cv
+
+The mixed models have most significant variables,
+
+M3.glm for variance was less significant (0.044181 ),
+
+the Q12 models in general were more significant
+
+cV's were more significant than var
+
+var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
+
+cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
+-->
+
+$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$
+
+```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
+
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M1)
+LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
+
+LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
+sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
+LmerMLH1_VAR_results
+
+#CV
+MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
+
+LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
+sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
+LmerMLH1_CV_results
+
+#USE Q12 (Q123) table
+#Q12_mouse_table_w.Ages
+#Q12 var
+MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
+
+LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
+sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
+LmerMLH1_M1_Q12_VAR_results
+
+#Q12 cv
+MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
+
+LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
+sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
+
+LmerMLH1_M1_Q12_VAR_results
+```
+
+These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
+
+```{r cv.plots, echo=FALSE, eval=FALSE}
+
+cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("cV of within mouse variance, all mice")
+cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+cV.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("cV of within mouse variance, 10 cell min")
+cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
+cV.Hist
+```
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M3)
+anova(MLH1var_M3)  
+as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
+
+MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+#summary(MLH1cv_M3)
+
+anova(MLH1cv_M3)
+as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
+
+#Q12 var
+MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1var_M3_Q12)
+anova(MLH1var_M3_Q12)  
+as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M3_Q12)
+anova(MLH1cv_M3_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
+
+```
+
+We followed up the mixed models with glms. The coefficients are listed below for the following glms: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
+
+$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
+
+```{r M3.MLH1.var, echo=FALSE}
+
+glmMLH1Var_M3 <- lm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
+round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
+
+MLH1cv_M3 <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+#summary(MLH1cv_M4)
+round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
+
+### Q12 var
+MLH1var_M3_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+
+MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
+#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
+round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
+
+#Q12 cv
+MLH1cv_M4_Q12 <- lm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
+
+#summary(MLH1cv_M4_Q12)
+#anova(MLH1cv_M4_Q12)
+#as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
+
+```
+
+Across all four data sets, the sex effect has the most significant p value.
+The highest p value is for the sex effect for VAR and Cv in the full models.
+This might be due to extra noise from technical sources.
+
+<!--
+var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
+
+cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
+-->
+
+For both the mixed model and linear model, sex was the most signifcant effect (p= `r LmerMLH1_VAR_results$sex.results` p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` (for the high quality dataset)  ) respectively.
+
+The measures for variance are more supciptibal to thechnical error effects (noise).
+For this reasons, we analysze the models by restricting the data to a subset cells which were scored as highest quality.
+Sex was also the most significant effect (p)
+
+<!--
+Unequal variance across strains might indicate that the within animal variance for MLH1 foci per cell may have evolved, yet --our low power and number of mice is not equal, -- number of mice across
+
+Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
+
+<!--
+
+even tho it's an important trait for XX reasons, -- there are limitations for our analysis
+
+Unlike the mean MLH1 counts -- the mean per mouse is calculated across a sample of cells (which serve as replicates)
+
+the within mouse variance for MLH1 count per cell, on the other hand -- is calculated at the mouse level, var()
+
+-- the replicates are at the level of mouse, unlike MLH1 counts per cell
 
@@
 
-above method remove / accounts for mean-variance correlation -
-
 The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
-
 QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
@@
 Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
-
 meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
--->
-
-Histograms might not be the best way to represent, but they show the general pattern than female mice have higher within mouse variation for CO counts.
-
-```{r variance.hist, echo=FALSE, eval=FALSE}
-#figures of histograms,
-
-#make series for total dataset?
-Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("Histograms of within mouse variance, all mice")
-Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-Var.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("Histograms of within mouse variance, 10 cell min")
-Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
-
-Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
-Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
-
-Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
-
-Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
-Q12.var.hist.plot
-
-#some LEW female still have v high within mouse variance (2 mice, )
-
-```
-
-A general pattern is that the female data has higher within mouse variance (both var and cV).
-Some strains more than others (investigate potential outliers).
-(LEW females with top 3 variance are all from different dissection dates)
-This pattern also holds for within mouse cv for MLH1 count (not shown).
-
-```{r cv.plots, echo=FALSE, eval=FALSE}
-
-cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("cV of within mouse variance, all mice")
-cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-cV.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("cV of within mouse variance, 10 cell min")
-cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
-cV.Hist
-```
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5}
+-->s
+
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
 
@@
 
-Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.female.VAR.plot
+#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+#Mol.female.VAR.plot
 
@@
 
-Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5}
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
 
@@
 
-from the Mixed models, there should be 4, 1-2 models from the full dataset for var and cv, and 3-4 from a higher quality data set for var and cv. Only a subset of the model outputs will be shown.
-
-For the mixed models the indiviual fixed effects are tested by a anova of the full and reduced models. The interaction fixed effect pvalue comes from drop1() a LTR test.
-
-$$mouse metric  ~=~ subsp * sex + rand(strain)+ \varepsilon$$
-
-```{r M1.MLH1.var_cv, echo=FALSE}
-#modeling within mouse variance
-#use  DF.HetC.MixedModel.HQ
-
-#HQ var.CO
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-#drop1(MLH1var_M1, test="Chisq") #interaction term NS
-#when switched to additive fixed, sex is very sig
-
-#anova(MLH1var_M1, MLH1var_M1.reduce.sex)
-#anova(MLH1var_M1, MLH1var_M1.reduce.subp)
-
-#exactRLRT(MLH1var_M1) #strain
-#as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#drop1(MLH1cV_M1, test="Chisq") #interaction term NS #when switched to additive fixed, sex is very sig
-#anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)
-#anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)
-#exactRLRT(MLH1cV_M1) #strain
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-#drop1(MLH1var_M1_Q12, test="Chisq")#sig
-
-#anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)
-#anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)
-#exactRLRT(MLH1var_M1_Q12)#NS
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-#drop1(MLH1cv_M1_Q12, test="Chisq")#NS
-
-pval.sig <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)
-CV.anova.p <- pval.sig$`Pr(>Chisq)`[2]
-
-# <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)
-
-#exactRLRT(MLH1cv_M1_Q12)#NS
-
-#make a table of the Mixed model pvalues, which are caluclated for subp, sex, sex:subsp and strain across 4 models ()
-```
-
-**General pattern is that sex has a very significant effect.** ( on the order of `r round(CV.anova.p, 6) ` )  this applies to different measures of within animal (across cell) variance and also to subsetted data with higher quality cells.
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
-
-#Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
-```
-
-<!-- insert the sex specific variance effect here / maybe also the plots of variance -->
-
-For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
-Most of the strain specific differences in within mouse variance go away in the Q12 data set.  (These strain effects could indicate that the distribution of across cell variance in CO number evolves.)
-
-The second set of models analyzed have this form.
-
-$$mouse \ CO \ metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-```{r M3.MLH1.var, echo=FALSE}
-
-MLH1var_M4 <- lm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-#as.data.frame( round(summary(MLH1var_M4)$coefficients[,c(1,4)], 6))
-
-MLH1cv_M4 <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M4)
-#as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))
-
-### Q12 var
-MLH1var_M4_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#summary(MLH1var_M4_Q12)
-#as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))
-#as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients[,c(1,4)], 6))
-
-#Q12 cv
-MLH1cv_M4_Q12 <- lm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M4_Q12)
-
-#anova(MLH1cv_M4_Q12)
-
-#as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
-
-```
-
-The glm's support the mixed models with the sex effect being the consistant significant effect
-
-#### Summary for witin mouse Variance
-
-The general pattern across all the models, is that sex is a significant effect for all models.
-**General pattern is that sex has a very significant effect.** The sex-subsp interaction effect p values were larger and only significant for one of the models.
-
-The models run over higher quality cells, suggest that staining or technical error in female data is unlikely to be the only cause for increased variance of number of MLH1 foci per oocyte.
-
-(some strains have significant effect -- some LEW females had large leverage).
-
-# DMC1 Results, Spermatocytes
-
-These data are meant to test if CO precursors, double strand breaks (DSBs), are significant predictors for MLH1 count variation. Simply, do mice with more DSBs also have more COs?
-
-Foci of DMC1 were scored from Leptotene and Zygotene spermatocytes of juvenile mice (12-14-18 days).
-One mouse represents each strain.
-The main comparison to examine is that between the high (PWD and MSM) and low recombining (WSB, G, KAZ) strains.
+## Female Specific Analysis
+
+In order to more finely discirbe the sex specific patterns, we foucsed on data sets including single sexes.
+
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+#table( droplevels(female_mouse.avs_for.analysis$strain) )
+#there are cast strains too
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Female mouse av.
+MLH1 count per cell ")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))
+
+female.specific.plot
+
+musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+#Mol
+Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+#Dom version 
+Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+```
+
+```{r female.strain.table, echo=FALSE}
+
+#female strain table
+#female_mouse.avs_for.analysis
+F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  = sum(Ncells),
+                        mean_co =as.numeric(format(round(mean(mean_co), 5 ), nsmall=3) ),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se)
+    )
+F.strain.table
+```
+
+For some strains, we were only able to quantified or include one mouse and the Cast observation comes from 1 1 cell.
+To analyze the female specific dataset,
+
+In the dom  subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M.
+m.
+domesticus$^WSB^_
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_
+
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _M. m. domesticus$^LEW^_
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _M.
+m.
+musculus$^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _M.
+m.
+musculus$^KAZ^_
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _M.
+m.
+molossinus$^MSM^_
+
+Two glm models were run to test the effects of strain and subspecies on mean MLH1 coount per cell per mouse. The coefficients for the following model are below.
+
+$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
+
+```{r Female.specific.glm, echo=FALSE}
+
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+lm.female.strain_sum <- summary(lm.female.strain)
+
+round(lm.female.strain_sum$coefficients[,c(1,4)],5)
+
+#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+
+#how much greater are the high strains?
+G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
+mean.G.female <- mean(G_female$mean_co)
+
+LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
+mean.LEW.female <- mean(LEW_female$mean_co)
+
+#how much greater is MSM female
+MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
+mean.MSM.female <- mean(MSM_female$mean_co)
+
+sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
+mean.all.other <- mean(sans_high_female$mean_co)
+```
+
+For mean CO count, _M. m. domesticus^G^_  strain effects in both models (p = )
+_M. m. musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).
+
+Above are the coefficents for the two glm's of the female specific mouse MLH1 averages, which include subspecies and strain as fixed effects.
+G has the most consistant significant strain effects in both models.
+MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.
+
+_M.
+m.
+domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means.
+_M.
+m.
+domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _M.
+m.
+musculus^MSM^_ is `r mean.MSM.female / mean.all.other`.
+These three will be designated as 'moderate high rec' strains for later sex specific analysis.
+
+Within mouse variance was assess to follow up on the results from the full dataset.
+
+```{r FEMALE.VAR.MODELS, echo=FALSE}
+## within mouse variance models 
+lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#round( #round isn't working for this sumary
+lmVAR.female.strain_sum$coefficients[,c(1,4)]
+
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+
+#round(
+lm.m1.Q12_female_sum$coefficients[,c(1,4)]
+#LEW not sig
+
+lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
+
+lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
+#LEW sig
+
+```
+
+The variance across mouse means for each strain is also variable across strains. With _M. m. domesticus$^LEW^_ having the most significant strain effect in the second model (p= ).
+
+## Male Specific Analysis
+
+The male specific patterns were also investigated.
+
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+#try customize the plot size
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#table( droplevels(just.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+male.plot
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+#Dom version 
+Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+```
+
+```{r male.strain.table, echo=FALSE}
+
+#female strain table
+#female_mouse.avs_for.analysis
+M.strain.table <- ddply(just.male_mouse.av, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  = sum(Ncells),
+                        mean_co =as.numeric(format(round(mean(mean_co), 4 ), nsmall=3) ),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se)
+    )
+M.strain.table
+```
+
+The same model framwrok was applied to the male specific data.
+The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
+
+ _M. m. domesticus^LEW^_
+
+- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
+
+- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
+
+- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
+
+Two models were fitted to the male specific data to test the effects of strain and subspecies.
+
+```{r Male.specific.glm, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+
+summry.male.strain.effects <- summary(glm.male.strain.subsp)
+round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
+
+#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+
+#1. determine which strain effects are sig
+
+#M2
+lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
+lm.male.strain_sum<- summary(lm.male.strain)
+
+round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+
+#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+
+#how much greater are the High Rec gwRRs
+just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
+                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
+                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
+
+#how much bigger are the high rec male means?
+sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+#mean(sans_LewRec_male$mean_co)
+
+PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+#mean(SKIVE_male$mean_co) /  
+```
+
+the coefficients for glms of the CO means are listed above.
+
+```{r MALE.SPECIFIC.variance.model, echo=FALSE}
+## within mouse variance models 
+lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
+
+#all p values kinda high, MSM and TOM are slightly sig
+lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
+#non sig
+
+lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+```
+
+The coefficients for the variance models are listed above
+
+The strain average for  _M.
+m.
+musculus^PWD^_ , _M.
+m.
+molossinus^MSM^_ and  _M.
+m.
+musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
+Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
+
+# Variation in DSB number per cell
 
@@
 
+In order to get a good idea on wether variation with mean CO number per cell is associated with the number of percursors upstream in prophase, we quantified DMC1 foci, a marker for DSBs.
+DMC1 foci were scored from X `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of juvenile mice (12-14-18 days).
+A subset of strains were quantified  _M.
+m.
+musculus^PWD^_, _M.
+m.
+molossinus^MSM^_, _M.
+m.
+musculus^KAZ^_ ,  _M.
+m.
+domesticus^WSB^_ , and _M.
+m.
+domesticus^G^_.
+Leptotene and zygotene cells were staged based on SC-AE and centromere morphology.
+
 ```{r DMC1.plot, echo=FALSE, warning=FALSE}
@@
 ```
+
+(discription of the nice boxplot figure)
+
+- overall the range is very high (CO homeostasis, technical noise, process is noiser continous repair and new DSBs)
+
+- exclude KAZ?
+
+- expected difference between cell stages for all strains
+
+- pattern of musc strains being higher than the dom strains (subspe pooled t-test)
+
+- MSM is much higher
+
+- the ratios of DMC1 : MLH1 counts!!
 
@@
         panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.bxlt_Z_to_L
 
@@
 
-DMC1.bxlt_High.to.Low
-
 #example for ggplot+ test from here
@@
 
-The category difference is only observed in the earlier Leptotene stage (p values report a t.test)
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
 
@@
 
-```
+##RATIOS
+DMC1table.main$MLH1.L_ratio <- (DMC1table.main$mean.count.L / DMC1table.main$mean.MLH1)
+DMC1table.main$MLH1.Z_ratio <- (DMC1table.main$mean.count.Z / DMC1table.main$mean.MLH1)
+
+#DMC1.mean.table
+
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+```
+
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+
+Why calculate ratios using the mean Leptotene cells -- they represent the a metric for the CO : NCO decision ( DSB / COs = NCO)
+
+ratios are .. not that different (slightly suprisingly)
+
+`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M.
+m.
+domesticus^WSB^_ , _M.
+m.
+domesticus^G^_, , _M.
+m.
+molossinus^MSM^_ and  _M.
+m.
+musculus^PWD^_ respecitively.
+
+These are the DSB:CO ratios for the L means `r DMC1table.main$MLH1.L_ratio` and Z means `r DMC1table.main$MLH1.Z_ratio`.
+
+(maybe ignore KAW.L? 1 cell)
 
@@
 
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-
-```{r barlett.test, echo=FALSE, eval=FALSE}
-
-#bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "L",])#L (don't work with single KAZ L cell)
-bartlett.test(total.count~strain, data=DMC1.DF[DMC1.DF$stage == "Z",])#Z
-
-```
-
 Since there is evidence for non-equal variance across strains for zygotene cells, so don't rely on lm()s that estimate the effect of strains.
 
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the Z cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
-
-<!--
- Single Bivalent Patterns
-
-## Total SC Results
-
-insert male total SC results here
-
-# Female patterns
-
-Nuanced
-there is strain specific evolution for females ...
-
-# Sexual dimorphism and heterochiasmy
-
-total SC length / SC compaction and telomere bias in positioning are conserved / consistently sexually dimorphism (in these mice?)  for most species
-
-make the heterochiasmy plot, with female strain averages adjusted for X
-(measure variance in CO counts for the top 5 female whole cell bivalents)
-(variance in )
-how does this compare to variance of mid 5, bottom 5?
-
-# Predictions / tests for selection on Bivalent structure
-
--->
+The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the zygotene cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
 
@@
 
-```{r misc.setup, echo=FALSE}
-
-colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
-                       #blues
-
-                       'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
-                       'CZECH'= 'indianred4',
-
-                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
-
-                    'CAST'='green','HMI'= 'forestgreen', #greens
-
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
-                    'CAROLI'='gold3',
-
-                    "F1"='grey', "other" = 'grey'
-                    )
-
-#add Rec group,
-#reclassified SKIVE as intermediate
-Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
-#apply this before ploting, to display strains in desired order
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, category, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-```
-
 <!-- merging the 2 data sets-->
@@
 
-```{r chrm_prop_data, echo=FALSE}
+```{r chrm_prop_data, echo=FALSE, warning=FALSE}
 #shut off eval for now because of wd / path error
@@
 
-```{r plot.chrm.prop.melt, echo=FALSE}
+```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
 
@@
 melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+
+#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+
 melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
@@
 chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
-chrm.class.barplot
+#chrm.class.barplot
 
@@
 
-**MSM 60% > PWD 50% > SKIVE 30% >  20 - 10% remaining (low) strains**
+**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
 
@@
 
-Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script.
-These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output).
-The MLH1 data file is also loaded into this file.
-
-The breakdown of single bivalent observations by category
-
-`r data.frame(table(droplevels(Curated_BivData$category) ) )`
-
-And these are the number of bivalent observations with filled hand.foci.count values.
-
-`r data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )`
-
-Validity of comparing bivalent observations
+In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining single chromosome patterns.
+
+From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software.
+After the human curation step, `r length(new.Curate.FULL.org$Obj.ID)` single bivalent measures were left.
+
+```{r algorithm_performance_stats, echo=FALSE}
+
+#sort by category,
+
+#number of mice, number of cells
+#mean number of isolate bivs per cell,
+
+data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
+```
+
+Table X illustrates the break down of single bivalent observations by category
+
+Error compared to human measures and other details have been discribed in (@Peterson_2019).
+
 While the automated software doesn't isolate all bivalents/chromosomes from each cell (on average 17), we assume that the isolation process is not biased.
@@
 
-Two Main questions for these analyses.
-
-> Q1. driving questions, which traits are sexually dimorphic?
-
-> Q2. which traits fit male polymorphism predictions?
+The mouse averages for 3 bivalent level metrics will be analyzed across these questions:  
+1)SC lengths (or total.SC)/chromatin compaction (using all cells regaurdless of CO number) 2) Interfocal Distance of 2CO bivalents and 3) the normalized foci position from 1CO bivalents.
+
+We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
 
@@
 
-The mouse averages for 3 bivalent level metrics will be analyzed across these questions:
-
-- SC lengths (or total.SC)
-
-- Normalized Rec landscape of 1CO
-
-- IFDs of 2CO
+(We will first list the results for the Q1)
+
+These are our predictions for the sex specifici patterns on the single bivalent lanscape
+
+ Q1 Analysis, Predictions for Heterochiasmy Q1
+
+Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
+
+1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
+
+2. Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
+
+3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (deBoer et al 1996, petkov 2001).
 
 The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.
-
-Q1. Model 1, mixed model lmer()  
-$$mouse \ av.\  metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-Q1. Model 2, glm()
-$$mouse \ av.\  metric ~=~ sex * strain + \varepsilon $$
-
-Q2 Model 3, logistic regression
-
-$$Rec \ group  ~=~ mouse \ av.\  metric  $$
-
-<!--
-I decided to use the mouse averages for most of the bivalent metric tests. some of my logic is below.
-
-**Pro Bivalent level**
-
-- The data are continuous and closer to normal/Gaussian expectations than the MLH1 counts.
-
-- There is within cell variance that is an issue (would affect single bivalent observations).
-
-**Pro Mouse averages**
-
-- Mouse average is a conservative choice, it would summarize general patterns which fits with the paper.
-
-- Simple application of the same models as MLH1 based results.
-
-- Limited work to add to the data set if I can justify that the collection of single bivalents are random and equivalent across mice.
-
-So I'll use the mouse averages for all the metrics is what is used in the mixed models.
-
--->
 
@@
 
-# Q1 Analysis, Predictions for Heterochiasmy Q1
-
-Using the Mixed model framework which tests the effects (and interactions) of subspecies, sex and strain, I will test for evolution of the following traits.
-
-Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
-
-1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of uniform vs telomere/biased CO positioning.
-
-  C) Centromere and telomere distances will be sexually dimorphic.
-
-3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).
-
 # Q1 SC Lengths
@@
 
+- Caveats XX
+
+The most convincing should be the short biv data set
+
 ```{r show.SC.length, echo=FALSE, warning=FALSE}
@@
 
-SC.length_Nested_boxplots
-
-#+ +geom_jitter(aes(alpha=0.2))
-```
-
-> M1. subsp * sex + (strain)
-
-> M2. subsp \* sex \* strain  (slightly redundant model)
-
-> M3. sex * strain
-
-The dependent variables I'll be testing in the model framework are:
-
-- Pooled SC lengths
-
-- 1CO
-
-- 2CO
-
-- 3CO
-
-- Xlong.biv  *Don't do long bivData for testing sex effect, since this would have the XX
-
-- Short.biv dataset *this would be a chrm size effect, without the XX
-
-<!--
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain*sex) + \varepsilon $$
-
-To estimate random effects, I'm using lmer (not lme where I could do the rand(strain*sex)). So now it's just random(strain).
-
-The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.
-
-remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**
-
-#### Pooled SC Lengths
-
-```{r MM_2.SC.Length, echo=FALSE, eval=FALSE}
-
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-lme_MM_SC.length <- lme(mean_SC ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-#p
-
-```
-
-I used the mixed model with the lme() for the full Curated BivData and long bivdata
-
--->
+```
 
@@
 
-#### Q1.SC_HetC_M1
-
-Insert the results for the MMs using the dependent variable 1CO - use the Mixed model (lmer), for all of the metrics (pooled SC, 1CO 2CO ect)
-
 Below's the code for Mixed model results -- try to organize them into a table or something
 
-```{r 1.SC.Length_M1, echo=FALSE}
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+Because of confonding effects across chromsomes of different physical lenghts, we attempt to ...
+reduce the error from mixing chromosomes by constructing reduced bivalent sets, for the longest 5 and shortest 5 bivalents from the same cell.
+The mouse means for these two data sets were then calculated.
+
+```{r Q1.Short.Long_biv.setup, echo=FALSE}
+Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength)
+                         )
+
+Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
+Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
+Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
+
+Q1.real.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength)
+                         )
+
+Q1.real.long_mouse_table <- add_strain(Q1.real.long_mouse_table)
+Q1.real.long_mouse_table <- add_sex(Q1.real.long_mouse_table)
+Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)
+
+```
+
+```{r show.short.biv, echo=FALSE}
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+
+Q1.SC.short.mouse_scatter_boxplot
+```
+
+```{r show.long.biv, echo=FALSE}
+#mouse av
+Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
+
+Q1.SC.long.mouse_scatter_boxplot
+```
+
+```{r Q1_short_biv_lmer, echo=FALSE}
 
@@
 #all pooled SC lengths
-lmer.Q1_SCLength_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.Q1_SCLength_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.Q1_SCLength_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-drop1(lmer.Q1_SCLength_M1, test="Chisq" ) #other options
-
-anova(lmer.Q1_SCLength_M1, lmer.Q1_SCLength_M1_R.sex) # also F 
-anova(lmer.Q1_SCLength_M1, lmer.Q1_SCLength_M1_R.subsp)
-
-exactRLRT(lmer.Q1_SCLength_M1)
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
 #as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
 
-#glmer.SCLength_all_MM.M1 <- glmer(mean_SC ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM) 
-#can't get p values from this summary
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+Q1lmer.MM_SC_short_results
+
+```
+
+These plot above show that qualitatively, the average female bivalent SC length is significantly longer than the mean length for males from the same strain.
+
+```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE}
+###
+lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value
+
+Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
+sex.results =  Q1lmer.MM_SC_long_sex, 
+               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
+Q1lmer.MM_SC_long_strain)
+
+Q1lmer.MM_SC_long_results
+
+```
+
+Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
+
+```{r Q1other_SC_lmer, echo=FALSE}
 
@@
 Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-#drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")
-exactRLRT(Q1.lmer_SCLength_1CO_M1)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
 #as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+Q1.lmer_SCLength.1CO_results
 
@@
 Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-#drop1(SCLength_2CO_MM.M1, test="Chisq")
-exactRLRT(Q1.lmer_SCLength_2CO_M1)
-as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-#Long.Biv
-Q1.lmer_SCLength_long.biv_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Long.biv_mouse_table_HetC)
-Q1.lmer_SCLength_long.biv_M1.R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Long.biv_mouse_table_HetC, REML = TRUE) 
-Q1.lmer_SCLength_long.biv_M1.R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Long.biv_mouse_table_HetC, REML = TRUE)
-
-anova(Q1.lmer_SCLength_long.biv_M1, Q1.lmer_SCLength_long.biv_M1.R.sex)
-anova(Q1.lmer_SCLength_long.biv_M1, Q1.lmer_SCLength_long.biv_M1.R.subsp)
-
-drop1(Q1.lmer_SCLength_long.biv_M1, test="Chisq")
-exactRLRT(Q1.lmer_SCLength_long.biv_M1)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+Q1.lmer_SCLength.2CO_results
 
@@
 
+Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
+
 The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
 
-All fixed and all interactions
+$$mouse \ average \ SC  \ length ~=~ sex * strain + \varepsilon $$
 
@@
 
+```{r Q1.SC.length.show_long.short, echo=FALSE}
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+Q1.SClength_Musc.BivData_plot
+
+```
+
 ```{r Q1.SC.M2plot, echo=FALSE}
 
-#Q1 plots for SC Lengths across Musc strains
+#removing these plots for space
 
@@
 SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
-Q1.SClength_Musc_plot
 
@@
 
-Q1.SClength_Musc.BivData_plot
-
 #BivData Level
@@
 
-Q1.SClength_Dom.BivData_plot
-
-```
-
-I think the strange thing about the Musc is that Skive females have shorter SC lengths.
-
-#### Q1. SC Length t.test
+```
 
@@
 
-#### Strain Comparisons of SC Length
-
-Dive deeper into the sex specific pattern for each strain.  (I think for all the within strain comparisons across sexes the general pattern is met of females being longer)
-
-(adding all of these blocks to -- the setup script); strain.Bivalents.DF
-
-```{r Q1.SC.show.long.biv, echo=FALSE, eval=FALSE}
-
-mouse.shoo.long.biv <- ggplot(real.long.bivData, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-mouse.shoo.long.biv
-
-```
-
-```{r Q2.SC.long.biv_wacko, echo=FALSE}
+```{r Q1.SC.long.biv_wacko, echo=FALSE}
+#removed these plots for space
+
 male.long.biv <- real.long.bivData %>% filter(sex== "male")
@@
 Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
-Q1_cell_long.biv
 
@@
 \nMale Long Biv SC Lengths")
-cell.avlong.biv
 
@@
 \nMale Long Biv SC Lengths")
-mouse.shoo.long.biv
-
-```
-
-#### XX adjustment
-
-Generate inteligible summaries for the permutation / subsampling results.
-
-```{r Q1_XX.adjustment, echo=FALSE}
-
-```
-
-# Q1. IFD
+
+```
+
+<!-- 
+lmr - sex is most significant effect
+
+plots,
+
+- ranges overlap,
+
+- try the short and long mouse averages
+
+G and MSM female?
+
+G male vs female
+
+-->
+
+# Q1. General CO Positions
+
+To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
+
+In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
+
+we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
+
+Chromosome size effects are a confonding factor for CO position, with shorter chromosomes having a more uniform landscape compared to larger chromosomes (cite).
+To account for this we again use the reduct bivalent sets (long.biv and short.biv)
+
+<!--
+-full 1CO tests
+
+long 1CO tests
+
+short 1CO tests
+
+There are a few traits that fall within the CO positions  
+- 1CO normalized position  
+- sis.co.ten (sorta interference)  
+- centromere and telomere distance  
+**Heterochiasmy**
+
+2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
+  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
+  C) Centromere and telomere distances will be sexually dimorphic.
+**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
+-->
+
+```{r Q1show.F1pos_Full, echo=FALSE}
+
+Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+Q1.short.biv_scatr_box
+
+```
+
+```{r CO1_foci_pos_boxplot, echo=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+```{r Q1.F1pos_long.short_setup, echo=FALSE}
+
+Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+Q1.long.1COpos_mouse.av <-  ddply(Q1.long.biv_1CO, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength), 
+                        mean_F1pos_PER =  mean(PER_Foci_1),
+                         mean_F1pos_ABS =  mean(Foci1)
+                         )
+Q1.long.1COpos_mouse.av <- add_strain(Q1.long.1COpos_mouse.av)
+Q1.long.1COpos_mouse.av <- add_sex(Q1.long.1COpos_mouse.av)
+Q1.long.1COpos_mouse.av <- add_subsp(Q1.long.1COpos_mouse.av)
+
+Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+Q1.short.1COpos_mouse.av <-  ddply(Q1.short.biv_1CO, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength),
+                        mean_F1pos_PER =  mean(PER_Foci_1),
+                         mean_F1pos_ABS =  mean(Foci1)
+                         )
+
+Q1.short.1COpos_mouse.av <- add_strain(Q1.short.1COpos_mouse.av)
+Q1.short.1COpos_mouse.av <- add_sex(Q1.short.1COpos_mouse.av)
+Q1.short.1COpos_mouse.av <- add_subsp(Q1.short.1COpos_mouse.av)
+
+```
+
+```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
+
+Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+Q1.short.biv_scatr_box
+
+Q1.long.biv_scatr_box   <- ggplot(Q1.long.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+Q1.long.biv_scatr_box
+```
+
+```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
+
+#make the same plots as above but with the long biv data set
+
+#mouse.av.long.biv.table
+Q1.av.long.biv_scatr_box   <- ggplot(Q1.long.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+Q1.av.long.biv_scatr_box
+
+Q1.av.short.biv_scatr_box   <- ggplot(Q1.short.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+Q1.av.short.biv_scatr_box
+
+```
+
+As expected -- confirming previous results, the normalized position of single crossovers varies between males and females.
+The strain effects indicate that within the musculus strains the difference is less compared to
+This is an indication of evolution of the sexual dimorphism for the pattern of single foci positions.
+
+The long bivalent pattern -- shows a different effect, both female and males have positions close to the center. This is a lack of telomeric pattern efpected for males
+
+For short bivalents, we predict that the normalized position will be more medial
+
+The short bivalent landscape pattern is more medial compared to males in all strains. However some strains have larger sex differences, WSB
+
+```{r CO.positions.figs, echo=FALSE, include=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
+
+#make a table for the sex specific values of the normalized F1 positions
+
+#for strain, sex -- Curated_BivData
+#or (Mouse.Table_BivData_4MM)
+#norm.first.foci.pos
+
+Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
+
+                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
+                  n1CO = sum(n1CO )
+
+                          )
+
+```
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.nrmF1CO.pos_M1, echo=FALSE}
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
+anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
+anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)
+
+drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )
+
+exactRLRT(lmer.Q1.nrm1CO_M1)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
+
+```{r Q1.Nrm.F1pos_M2, echo=FALSE}
+
+#calq the means for normalized F1
+#M2
+
+Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_nrmF1pos_M2)
+
+summry.Q1.glm_nrmF1pos_M2 <- summary(Q1.glm_nrmF1pos_M2)
+round(summry.Q1.glm_nrmF1pos_M2$coefficients[,c(1,4)],5)
+
+write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", sep=",",row.names = TRUE)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.nrmCO.pos_M3, echo=FALSE, eval=FALSE}
+
+Q1.glm_nrmF1pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_nrmF1pos_M2)
+
+summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
+round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)
+
+summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
+round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)
+
+write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+```
+
+The model results aren't as clear as I'd like.
+For the Mol, both the sex effect and MOLF strain effect are significant (this means both female and male rec landscapes are effected - most of the time towards the middle
+
+GENERALLY - Sex is the biggest effect for the 1CO landscape (female middle male telomereic)
+molf
+Not sure I understand how MSM is supper significant for 
+There must be some strange effect due to the Mol subsp - MOLF isn't even in M2 (
+Male is consistently significant across the two models.
+
+Male and MOLF are the most significant effects for M3
+
+Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
+This plot shows the sexual dimorphism in the density plots.
+
+Consider adding annotate_text for the number of observations in each category.
+think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
+
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#MELT_Curated_BivData
+
+#these isn't long.biv data
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+real.long.bivData
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+```
+
+<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+the distribution of SC lengths and sis-coten seems very different across sexes
+
+The mixed model data should only come from 1CO bivalent data.
+
+the mouse average foci1 pos is more significant in t.test, but not log regression...
+(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
+
+```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+
+#LOGISTIC REGRESSION for FOCI POSITION
+
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}
+
+#t.test for high and low group  -- logistic regression still need to format the high and low group
+
+#mouse av
+t.test_mouse.av.1CO_pos <- t.test(
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
+
+  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv.1CO_pos_t.test <- t.test(
+
+  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+  )
+
+```
+
+### Siscoten
+
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males.
+
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
+t.test_mouse.av_siscoten <- t.test(
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+#2CO
+
+#3CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+
+drop1(lme_MM_telo.dist, test="Chisq")
+
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+drop1(lmer_MM_telo.dist, test="Chisq") 
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
+
+### Heterochiasmy Prediction
+
+> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
+
+> is random strtain effect significant?
+
+The random strain effect seems very significant.
+
+remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
+
+```{r pos_t.test}
+#make point plots / boxplots which show differences in mean positions
+
+#scatter + boxplot for t.tests
+```
+
+Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
+
+```{r F1.by.F2}
+
+#try remaking the plot Megan suggested
+# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
+
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
+#isolate 2COs
+#facet by sex and subsp
+
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
+
+```
+
+```{r CO.pos.ANOVA, eval=FALSE}
+#what is the pattern of variance
+#run analyses for each subsp*sex
+#use non-melt DF
+
+#how is the variance partioned across
+#cell, mouse, strain
+
+female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
+female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
+
+female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
+
+#unorder strain and mouse
+
+female.Dom$mouse <- as.factor(female.Dom$mouse)
+
+female.Dom$strain <- unclass(female.Dom$strain)
+female.Dom$strain <- as.factor(female.Dom$strain)
+
+female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
+female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
+
+#1CO first
+modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
+
+#can't get mouse and strain to have sum of square
+#residual size decreases with per.F1
+#residuals much larger than fileName, mouse and strain no
+
+#model <- lm(breaks ~ wool * tension, 
+#            data = warpbreaks, 
+#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
+
+male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
+male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
+
+male.Dom$mouse <- as.factor(male.Dom$mouse)
+
+male.Dom$strain <- unclass(male.Dom$strain)
+male.Dom$strain <- as.factor(male.Dom$strain)
+
+male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
+male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
+
+male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
+
+male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
+summary(aov(male.modo))
+
+#only file name is registering as effect
+#Review ANOVA frameworks
+#http://www.biostathandbook.com/nestedanova.html
+
+```
+
+#Q1. IFD
+
+In order to test if there are sex differences between crossover interfernce, a major determinant of the positioning of crossovers along chromosomes, we examined ...
+ran models with the interfocal distance (IFD) of two foci on the same bivalent.
+
+-We focus on observations from two-crossover bivalents for more comparable observations.
+
+-comparisons of all IFDs of mulit crossover bivalents is also included
+
+Interference is a major determinant of the positioning of chromosomes
+
+<!--
 
@@
 DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-
-))
-
 <!--
@@
 lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-
 lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-
 lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
 
-anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)
-anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)
-
-drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )
-
-exactRLRT(lmer.IFD.ABS_MM.M1)
+ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
+               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
 
@@
 lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-
 lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-
 lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
 
-anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
-anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
+sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
+               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
+#don't understand why 2 rows?
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
+#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
 
 #these 2 below are LR tests
-drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-exactRLRT(lmer.IFD.PER_MM.M1)
+#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+#exactRLRT(lmer.IFD.PER_MM.M1)
 
@@
 
-# Q1. General CO Positions
-
-Try accounting for chrm size effects. **Also i didn't notice that I'm using density plots and scatter+boxplots across sections**
-
-<!--
-There are a few traits that fall within the CO positions  
-- 1CO normalized position  
-- sis.co.ten (sorta interference)  
-- centromere and telomere distance  
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-  C) Centromere and telomere distances will be sexually dimorphic.
-**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
--->
-
-density plots
-```{r CO.positions.figs, echo=FALSE, include=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-
-```
-
-```{r Q1.F1.pos.LONG.BIV_scatter, echo=FALSE}
-
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-real.long.bivData_1CO <- real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q1.show.nrmF1_long.biv <- ggplot(real.long.bivData_1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Q1.show.nrmF1_long.biv
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-Q2.shoooo
-
-```
-
-```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
-
-#make a table for the sex specific values of the normalized F1 positions
-
-#for strain, sex -- Curated_BivData
-#or (Mouse.Table_BivData_4MM)
-#norm.first.foci.pos
-
-Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
-
-                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
-                  n1CO = sum(n1CO )
-
-                          )
-
-```
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.nrmF1CO.pos_M1, echo=FALSE}
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
-anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)
-
-drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )
-
-exactRLRT(lmer.Q1.nrm1CO_M1)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.Nrm.F1pos_M2, echo=FALSE}
-
-#calq the means for normalized F1
-#M2
-
-Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_nrmF1pos_M2)
-
-summry.Q1.glm_nrmF1pos_M2 <- summary(Q1.glm_nrmF1pos_M2)
-round(summry.Q1.glm_nrmF1pos_M2$coefficients[,c(1,4)],5)
-
-write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.nrmCO.pos_M3, echo=FALSE, eval=FALSE}
-
-Q1.glm_nrmF1pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_nrmF1pos_M2)
-
-summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
-round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)
-
-write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-```
-
-The model results aren't as clear as I'd like.
-For the Mol, both the sex effect and MOLF strain effect are significant (this means both female and male rec landscapes are effected - most of the time towards the middle
-
-GENERALLY - Sex is the biggest effect for the 1CO landscape (female middle male telomereic)
-molf
-Not sure I understand how MSM is supper significant for 
-There must be some strange effect due to the Mol subsp - MOLF isn't even in M2 (
-Male is consistently significant across the two models.
-
-Male and MOLF are the most significant effects for M3
-
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#MELT_Curated_BivData
-
-#these isn't long.biv data
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-real.long.bivData
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-```{r CO1_foci_pos_boxplot, echo=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-The mixed model data should only come from 1CO bivalent data.
-
-the mouse average foci1 pos is more significant in t.test, but not log regression...
-(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}
-
-#t.test for high and low group  -- logistic regression still need to format the high and low group
-
-#mouse av
-t.test_mouse.av.1CO_pos <- t.test(
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv.1CO_pos_t.test <- t.test(
-
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-  )
-
-```
-
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
-drop1(lme_MM_telo.dist, test="Chisq")
-
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
-### Heterochiasmy Prediction
-
-> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
-
-> is random strtain effect significant?
-
-The random strain effect seems very significant.
-
-remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
-
-```{r pos_t.test}
-#make point plots / boxplots which show differences in mean positions
-
-#scatter + boxplot for t.tests
-```
-
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
-
-```{r F1.by.F2}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-```{r CO.pos.ANOVA, eval=FALSE}
-#what is the pattern of variance
-#run analyses for each subsp*sex
-#use non-melt DF
-
-#how is the variance partioned across
-#cell, mouse, strain
-
-female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
-female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
-
-female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
-
-#unorder strain and mouse
-
-female.Dom$mouse <- as.factor(female.Dom$mouse)
-
-female.Dom$strain <- unclass(female.Dom$strain)
-female.Dom$strain <- as.factor(female.Dom$strain)
-
-female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
-female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
-
-#1CO first
-modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
-
-#can't get mouse and strain to have sum of square
-#residual size decreases with per.F1
-#residuals much larger than fileName, mouse and strain no
-
-#model <- lm(breaks ~ wool * tension, 
-#            data = warpbreaks, 
-#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
-
-male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
-male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
-
-male.Dom$mouse <- as.factor(male.Dom$mouse)
-
-male.Dom$strain <- unclass(male.Dom$strain)
-male.Dom$strain <- as.factor(male.Dom$strain)
-
-male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
-male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
-
-male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
-
-male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
-summary(aov(male.modo))
-
-#only file name is registering as effect
-#Review ANOVA frameworks
-#http://www.biostathandbook.com/nestedanova.html
-
-```
-
 ## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
@@
 ```{r Q2.SC.long.biv_2, echo=FALSE}
+
+#>>>>>>> 7ebba2e601079e1783e92c05ba74529ea765a5ed
+
 male.long.biv <- real.long.bivData %>% filter(sex== "male")
@@
 
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+cell.avlong.biv
+
 ```
@@
 
-glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= mouse.av.long.biv.table)
+glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= Long.biv_mouse_table)
 #summary(glm1.SC.male_long.biv)
 
-glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table)
+glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table)
 #summary(glm2.SC.male_long.biv)
@@
 #Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Dom",])
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
 #summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
@@
 #Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Musc",])
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
 #summary(glm_SC.male_Musc_long.biv)
@@
 #Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= mouse.av.long.biv.table[mouse.av.long.biv.table$subsp == "Mol",])
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
 #summary(glm_SC.male_Mol_long.biv)
@@
 ## logistic regression for Rec group
-
-log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-              data=mouse.av.long.biv.table,  family=binomial(link="logit"))
+#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
+#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
 
@@
 
-```{r male.show.plotsEXXXample, echo=FALSE, eval=FALSE}
-
-#REWRITE THIS FOR SC LENGTH COMPARISONS
+```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
 
@@
 
+#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
 ```
@@
 The full data set is `r length(Curated_BivData$Obj.ID)`.
-This is the breakdown of bivalent observations by category for this long dataset are `r data.frame(table(droplevels(real.long.bivData$category)))`.
+This is the breakdown of bivalent observations by category for this long dataset are
 
@@
 # Deleted notes
+
+```{r variance.hist, echo=FALSE, eval=FALSE}
+#figures of histograms,
+
+#make series for total dataset?
+Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +
+  ggtitle("Histograms of within mouse variance, all mice")
+Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
+Var.Hist.total.mice
+
+#DF.HetC.MixedModel.HQ
+Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  + 
+  ggtitle("Histograms of within mouse variance, 10 cell min")
+Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
+
+Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
+Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
+
+Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
+  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
+
+Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
+Q12.var.hist.plot
+
+#some LEW female still have v high within mouse variance (2 mice, )
+
+```
+
 <!--

### 2020-02-13 src/BivData_Setup.Rmd
cleaning up more

@@
 
-  6. Long Biv Dataset
+  6. Long Biv Dataset and short biv dataset
 
@@
 
+new.Curate.FULL.org$Obj.ID <- paste(new.Curate.FULL.org$fileName, new.Curate.FULL.org$boxNumber, sep = "_")
+
 #write files for basic stats to csv files
@@
 
-The below cell level table is used for making the **long.biv.dataset**, (any others?)
+The below cell level table is used for making the **long.biv.dataset** and short, (any others?)
 
@@
 
-# Long BivData Set
+# Long and Short BivData Set
 
@@
 
-  the.Q4 <-cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
   the.Q3 <-cell.level.BivData_table$Qar.3[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
   the.Q2 <-cell.level.BivData_table$Qar.2[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
@@
 #3.mark which bivalents are above the threshold
-#WHY DOESN"T THIS WORK?
 
@@
 Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)
+
+#rec group
+Long.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Long.biv_mouse_table$category), 1, 
+                            ifelse(grepl("MSM", Long.biv_mouse_table$category), 1,
+                             ifelse(grepl("SKIVE", Long.biv_mouse_table$category), 1, 0)) )
 

### 2020-02-14 src/BivData_Setup.Rmd
re-ran setup, rdata wasn't loading, clean up todos

@@
 - remove BivData! (these are all Peromyscus!!) (not sure where this DF comes from)
-
-declare the Strain specific BivData_DFs  (for convinice)
-
-check the numbers of bivalents with a SC length, but no hand.foci.count
 
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_12.2.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_13.2.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-02-14 doc/Results.rmd
trying to fix bug for kniting
knit successful, adding more results

@@
 
-load(file="~./MLH1repo/data/CleanBivData_12.2.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_13.2.20.RData")
 
@@
 
-glmMLH1Var_M3 <- lm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
 glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
@@
 
-MLH1cv_M3 <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
 #summary(MLH1cv_M4)
@@
 ### Q12 var
-MLH1var_M3_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
 
@@
 #Q12 cv
-MLH1cv_M4_Q12 <- lm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
 round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-
-#summary(MLH1cv_M4_Q12)
-#anova(MLH1cv_M4_Q12)
-#as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))
 
@@
 ```{r Q1_short_biv_lmer, echo=FALSE}
-
 #Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
@@
 
-```{r Q1.SC.length.show_long.short, echo=FALSE}
+```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
 
@@
 SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-Q1.SClength_Musc.BivData_plot
 
@@
 
-```{r Q1show.F1pos_Full, echo=FALSE}
+```{r Q1show.F1pos_Full, echo=FALSE, eval=FALSE}
 
@@
 
-```{r CO1_foci_pos_boxplot, echo=FALSE}
+```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
 
@@
 
+<!--
 ### Siscoten
@@
 
-### Heterochiasmy Prediction
-
-> Is sex a significant effect for the 1CO normalized CO position? (as predicted)
-
-> is random strtain effect significant?
-
-The random strain effect seems very significant.
-
-remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)
-
-```{r pos_t.test}
-#make point plots / boxplots which show differences in mean positions
-
-#scatter + boxplot for t.tests
-```
-
-Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions
-
-```{r F1.by.F2}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-```{r CO.pos.ANOVA, eval=FALSE}
-#what is the pattern of variance
-#run analyses for each subsp*sex
-#use non-melt DF
-
-#how is the variance partioned across
-#cell, mouse, strain
-
-female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
-female.Dom <- female.Dom[female.Dom$subsp == "Dom",]
-
-female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength
-
-#unorder strain and mouse
-
-female.Dom$mouse <- as.factor(female.Dom$mouse)
-
-female.Dom$strain <- unclass(female.Dom$strain)
-female.Dom$strain <- as.factor(female.Dom$strain)
-
-female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
-female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]
-
-#1CO first
-modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)
-
-#can't get mouse and strain to have sum of square
-#residual size decreases with per.F1
-#residuals much larger than fileName, mouse and strain no
-
-#model <- lm(breaks ~ wool * tension, 
-#            data = warpbreaks, 
-#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))
-
-male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
-male.Dom <- male.Dom[male.Dom$subsp == "Dom",]
-
-male.Dom$mouse <- as.factor(male.Dom$mouse)
-
-male.Dom$strain <- unclass(male.Dom$strain)
-male.Dom$strain <- as.factor(male.Dom$strain)
-
-male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
-male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]
-
-male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength
-
-male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
-summary(aov(male.modo))
-
-#only file name is registering as effect
-#Review ANOVA frameworks
-#http://www.biostathandbook.com/nestedanova.html
-
-```
+-->
 
@@
 <!-- 
-
 Brief lit review for IFD / interference expectations  
@@
 Chrm1 genetic map, 
-
 (other human and mice refs 
-
 Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
@@
 
+total.IFDs <- Curated_BivData %>% filter(hand.foci.count > 1)
+
 #take a look at the 3CO things
@@
 
-C.2) sis-co-ten metric ... (what about the clustering?)
-
-C.3) telomere and centromere distances ...
-
-M1 glm model for fixed strain effect across male averages.
-
 # Q2 SC
 
+In order to test if sc lengths (sc area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
+
+The average sc lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confonded by the different proportions across strains.
+
+long.biv data annd short bivdata
+
 Focus on comparing the differences between long.biv (and short biv once I make that dataset)
 
-```{r Q2.male.Nrm1CO_strain, echo=FALSE}
+```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
 #isolate all musc
@@
 
+```{r wacky, echo=FALSE}
+
+real.long.bivData$bin <- "long"
+real.short.bivData$bin <- "short"
+
+lng.shrt <- rbind(real.long.bivData[,c(1:2, 17,19)], real.short.bivData[,c(1:2, 17,19)])
+
+uu <- ggplot(lng.shrt, aes(x=fileName, y=chromosomeLength) )  + geom_point()+facet_wrap(~strain, scales="free")
+
+```
+
+```{r Q2.SC.short.long.setup, echo=FALSE}
+table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+mouse.av.long.biv.table <- add_mouse(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+
+mouse.av.short.biv.table <- add_mouse(mouse.av.short.biv.table)
+mouse.av.short.biv.table <- add_strain(mouse.av.short.biv.table)
+mouse.av.short.biv.table <- add_subsp(mouse.av.short.biv.table)
+mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
+mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
+mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
+```
+
+```{r Q2.SC.long.biv_2, echo=FALSE}
+shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+```
+
 ```{r Q2.SC.ploy.male, echo=FALSE}
@@
 
-```{r Q2.SC.long.biv_2, echo=FALSE}
-
-#>>>>>>> 7ebba2e601079e1783e92c05ba74529ea765a5ed
-
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-shoo.long.biv
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-cell.avlong.biv
-
-#mouse av for long biv
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-mouse.shoo.long.biv
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-cell.avlong.biv
-
-```
-
-Not the pattern I expected, there might be more noise than I thought in this dataset.
-
-```{r Q2.SC.long.biv_3, echo=FALSE}
-
-#mouse.av.long.biv.table
-
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table <- make this is set up
-
-glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= Long.biv_mouse_table)
-#summary(glm1.SC.male_long.biv)
-
-glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table)
-#summary(glm2.SC.male_long.biv)
+```{r Q2.glm.SC.long, echo=FALSE}
+glm.Q2.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
+round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
+
+table(mouse.av.long.biv.table$mouse)
+
+glmQ2.M2.SC.male_long.biv <- glm(mean.SC ~ strain, data= mouse.av.long.biv.table)
+round(summary(glmQ2.M2.SC.male_long.biv)$coefficients[,c(1,4)], 6)
+```
+In the cell average Long biv data set, MSM G and CZECH strains have sig strain effects
+
+very few strains have significant values for the mouse
+
+```{r Q2.glm.SC.short, echo=FALSE}
+glmQ2_SC_mouse.av.short_biv <- glm(mean.SC ~ strain, data= mouse.av.short.biv.table)
+
+round(summary(glmQ2_SC_mouse.av.short_biv)$coefficients[,c(1,4)], 6)
+```
+Basically all strains have a significant effect for the mouse.av of long biv
+
+```{r Q2.glm.SC.all.strain, echo=FALSE}
 
@@
 glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
+round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
 
@@
 glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
+
 #summary(glm_SC.male_Musc_long.biv)
@@
 glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-#summary(glm_SC.male_Mol_long.biv)
+round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
 
@@
 
+abova are glm tables for mouse.av SC lengths for each strain.
+
 Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
@@
 The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).
-
-#### Q2 SC by class
 
@@
 SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.blah.mouse.av
 
@@
 SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.logReg.mouse.av_1CO
 
@@
 SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-plot.logReg.mouse.av_2CO
 
@@
 
-```{r Q2.SClength.t.test, echo=FALSE}
+```{r Q2.SClength.ttest, echo=FALSE}
 #MOUSE AVE, all
@@
 But When the means are compared across chrm class, the mouse averages are no longer significant.
-
-### Q2 SC length Chrm class prediction
 
@@
 
+```{r F1.by.F2, eval=FALSE}
+
+#try remaking the plot Megan suggested
+# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
+
+CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
+
+#isolate 2COs
+#facet by sex and subsp
+
+F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
+F1.x.F2
+
+```
+
 # Deleted notes

### 2020-02-14 doc/Results.rmd
starting total SC analysis

@@
 
+load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
+
+#skeletonize
+SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
+
 #load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
@@
                        'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
-                       'CZECH'= 'indianred4',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4',
+
+                       'TOM'= 'indianred1','AST' = 'tan2',
+                       'CZECH'= 'brown1',
 
@@
                     )
+
+color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF")
 
@@
 ```
+
+```{r totSC_setup, echo=FALSE}
+#SC.skel
+#"anon.batch" "rand_name"  "bin_size"   "skel_size"
+
+#MERGEING MLH1 with SKEL
+MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
+OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
+
+MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
+
+MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
+
+MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
+MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
+
+#table(MLH1.merge.Skel$category)
+
+```
+
+!Note! the most recent tot.sc csv was missing ~2 batches... rerun
+
+```{r tot.sc.viz4outliers, echo=FALSE}
+
+tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=mouse))+ geom_point()
+tot.sc2 <- ggplot(data= MLH1.merge.Skel, aes(y = skel_size, x=mouse))+ geom_point()
+
+#seems like the outliers have been removed
+```
+
+```{r tot.skel.table, echo=FALSE}
+
+totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+
+#unorder the factors
+totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
+totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
+totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
+
+#add rec group
+totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
+
+OG.totSC.mouse.av.table <- totSC.mouse.av.table
+
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(sex~subsp, scales="free")
+
+tot.sc77 <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(0, 800))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+
+tot.sc_bin <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())  +ylim(c(0, 3000))
+
+#there should be tom
+
+```
+From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
+
+the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
+
+INSERT TABLE FOR THE TOTAL SC means ect
+
+The difference between
+
+bin is short for binary, it reports the binarized area of the SC
+
+skel -- takes the skeletonized midline of the bin objects
 
@@
 
+In order to test one of the major/known correlates with genome wide recombiation rate variation, we compared the average
+```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+#nested boxplots, pooled Bivalents
+
+Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+Q1.SC.length_Nested_boxplots_Dom
+
+Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+Q1.SC.length_Nested_boxplots_Musc
+
+Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
+Q1.SC.length_Nested_boxplots_Mol
+
+```
+
+The full mFull data set of SC lengths shows that
+
+- there is a positive correlation between mean SC length and number of COs per chrm
+
 We expect female SC lengths to be longer (refs).
@@
 
-```{r show.SC.length, echo=FALSE, warning=FALSE}
-#for now remove the 
-#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
-#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS
-
-#add boxplots for mouse level means
-#add second x axis, hand.foci.count
-
-#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#nested boxplots
-SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))
-
-```
-
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-Below's the code for Mixed model results -- try to organize them into a table or something
-
-Because of confonding effects across chromsomes of different physical lenghts, we attempt to ...
-reduce the error from mixing chromosomes by constructing reduced bivalent sets, for the longest 5 and shortest 5 bivalents from the same cell.
-The mouse means for these two data sets were then calculated.
-
 ```{r Q1.Short.Long_biv.setup, echo=FALSE}
@@
 ```
+
+Chromosome size effect.
+
+$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+Below's the code for Mixed model results -- try to organize them into a table or something
+
+Because of confonding effects across chromsomes of different physical lenghts, we attempt to ...
+reduce the error from mixing chromosomes by constructing reduced bivalent sets, for the longest 5 and shortest 5 bivalents from the same cell.
+The mouse means for these two data sets were then calculated.
 
@@
 
+```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
 ```{r Q1show.F1pos_Full, echo=FALSE, eval=FALSE}
@@
 
-write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
+#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
 
@@
 
-write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
 
@@
 
-long.biv data annd short bivdata
-
-Focus on comparing the differences between long.biv (and short biv once I make that dataset)
+First we examine the pattern of SC length variation across chromosome classes.
+
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
+
+#nested boxplots, pooled Bivalents
+#x axis for hand.focicount -- good for seperating the points --- but it pools
+
+#prediction for Dom strains: low variation 
+Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+Q2.SC.length_Nested_boxplots_Dom2
+
+Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+Q2.SC.length_Nested_boxplots_Musc2
+
+Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
+Q2.SC.length_Nested_boxplots_Mol
+
+#+theme(legend.position="none") 
+```
+
+**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.
+
+The positive correlation of sc lengths has a positive correlation with the number of COs on single chromosomes ,
+
+in the musculus and molossinus strains, the positive correlation is more
+
+In Domisticus the -- postive correlation only holds for the first 3 classes of bivalents (0CO, 1CO, 2CO) -- this illustrates how the differences between proportions of chromosome classes across strains confonds comparisons of SC length across strains.
+
+The glm's for the full dataset across -- the coefficients are listed below
 
@@
 uu <- ggplot(lng.shrt, aes(x=fileName, y=chromosomeLength) )  + geom_point()+facet_wrap(~strain, scales="free")
+
+```
+
+We purpose -- we think that running the models for comparing SC means by chromosome class and devided by rec group.
+(since the chromsome porportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )
+
+The null hypothesis is that these won't be highly significant strain effects.
+If there are still strain effects fo these models, that would be evidence that there has been eevolution for SC length compaction across these strains
+
+```{r show.tot.sc, echo=FALSE}
+
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(sex~subsp, scales="free")
+
+tot.sc77 <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(150, 650))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+tot.sc77
+
+tot.sc_bin <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())  +ylim(c(0, 3000))
+```
+
+The skeletonized measures show that the general pattern holds for total SC and the high rec strains.
+However there are a few exceptions -- However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel
+
+--
+
+-the SKIVE are within dom and kaz level -- so why does SKIVE have inntermediate CO proportions and gwRR
+
+24sep14_PWD_m1 is lower than the others
+5jul17_MSM_m1 is the low MSM
+
+The CZECH and AST are higher than KAZ
+
+```{r Q2.SC_class.rec.group_setup, echo=FALSE}
+
+#rerun the models for datasets -- deviding by chrm class and rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
+
+Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
+Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
+
+#calq_mouse av
+mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        n_cells = length(unique(fileName) ),
+                        mean.SC.total = mean(chromosomeLength),
+
+                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+                         )
+
+mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
+mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
+
+mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
+mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
+
+#calq_mouse av
+mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
+
+mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
+mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
+```
+
+The number of cells and mice is difference across the low group strains, 
+-- permut the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)
+
+```{r Q2.SC.Length.stuff, echo=FALSE}
+#all pooled SC lengths
+#WSB     G   LEW  MOLF   KAZ CZECH 
+#    5     9     5     4     3     3
+
+#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14
+
+#G sample 5 mice
+
+#break down of mice in the low rec dataset
+
+#glm.1CO
+Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
+summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
+round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)
+
+#glm.2CO
+Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
+summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
+round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
+#fewer significant differences
+
+Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
+sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
+round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
+#all but KAZ are slightly significant, G and CZECH more so
+
+#glm.3CO
+Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
+#not sig differences
+summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
+round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)
+
+#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)
+
+#all pooled SC lengths
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+Q1lmer.MM_SC_short_results
 

### 2020-02-14 doc/totalSC.Rmd
starting total SC analysis

@@
 #load BivData
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData.RData")
+load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")
 
@@
 ```
-
-The mixed models are mouse averages, get mouse names and ncells for each of the lists.
 
@@
 ```{r more.blahs, echo=FALSE}
-
 #WSB -- WSB_curation, WSB_female output on big External
@@
 #save RData file
-#save.image("~./MLH1repo/data/total_SC_1.6.20.RData")
+#save.image("~./MLH1repo/data/total_SC_14.2.20.RData")
+
+#changing to just loading the csv file
 

### 2020-02-15 doc/Results.rmd
starting checking the SC_skel algorithim output

@@
 #skeletonize
-SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
+SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
 
@@
 
-color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF")
+color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
+#4!!
+
+show_col(hue_pal()(5))
 
@@
 #SC.skel
-#"anon.batch" "rand_name"  "bin_size"   "skel_size"
+#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included
+
+SC.skel.OG <- SC.skel #5683
 
@@
 MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
 
@@
 MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
+#5110,
+
+OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
+
+#check for DUPs, #check for duplicatations of the random name
+MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
+tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
 
@@
 MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
-
 #table(MLH1.merge.Skel$category)
 
-```
-
-!Note! the most recent tot.sc csv was missing ~2 batches... rerun
+#remove outliers
+MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
+
+```
 
@@
 
-tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=mouse))+ geom_point()
-tot.sc2 <- ggplot(data= MLH1.merge.Skel, aes(y = skel_size, x=mouse))+ geom_point()
+tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
 
 #seems like the outliers have been removed
-```
+
+MLH1.merge.Skel.wsb.male <- MLH1.merge.Skel  %>% filter(category < "WSB male" )
+
+tot.sc.wsb.male <- ggplot(data= MLH1.merge.Skel.wsb.male, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
+
+tot.sc.single.male <- ggplot(data= MLH1.merge.Skel  %>% filter(category == "G female" ) , aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
+
+```
+
+Checking the outliers at category level,
+
+WSB male,
+10apr15_31mar15_WSB_m3_sp1_4_rev.tif
+
+G outliers, 
+G 19oct15_12oct15_G_m1_sp1_17.1_rev.tif
+
+LEW, low outlier, 
+6mar18_1feb18_LEW_m1_sp1_21.5_rev
+
+PWD outliers, 6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif, 7mar15_1mar15_PWD_m1_sp4_2_rev.tif, and 20jul15_18may15_PWD_m2_sp1_10.2_rev.tif
+
+G female; 1mar16_20feb16_G_f3_sp1_8_rev.tif and 21nov16_16jan16_G_f2_sp1_15_rev.tif
 
@@
 
+Total SC
+
 ```{r show.tot.sc, echo=FALSE}
-
 totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

### 2020-02-15 doc/Results.rmd
de-bugged for knitting, cleaned up outliers for total SC

@@
 
+# Data set ups
+
 ```{r organizing.data, echo=FALSE }
@@
 
-show_col(hue_pal()(5))
+#show_col(hue_pal()(5))
 
@@
 
+# -- EXAMINE the ggplot scatter plots for outliers
 tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
 
-#seems like the outliers have been removed
-
-MLH1.merge.Skel.wsb.male <- MLH1.merge.Skel  %>% filter(category < "WSB male" )
-
-tot.sc.wsb.male <- ggplot(data= MLH1.merge.Skel.wsb.male, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
-
-tot.sc.single.male <- ggplot(data= MLH1.merge.Skel  %>% filter(category == "G female" ) , aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
-
-```
-
-Checking the outliers at category level,
-
-WSB male,
-10apr15_31mar15_WSB_m3_sp1_4_rev.tif
-
-G outliers, 
-G 19oct15_12oct15_G_m1_sp1_17.1_rev.tif
-
-LEW, low outlier, 
-6mar18_1feb18_LEW_m1_sp1_21.5_rev
-
-PWD outliers, 6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif, 7mar15_1mar15_PWD_m1_sp4_2_rev.tif, and 20jul15_18may15_PWD_m2_sp1_10.2_rev.tif
-
-G female; 1mar16_20feb16_G_f3_sp1_8_rev.tif and 21nov16_16jan16_G_f2_sp1_15_rev.tif
+#round 1
+SC.skel.outliers_round1 <- c(
+  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
+  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
+            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
+
+    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
+                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
+              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
+          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",
+
+  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
+          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",
+
+       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
+                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",
+
+   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
+   "6may15_17apr15_PWD_f1_sp1_7_rev",
+
+      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
+            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",
+
+    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
+    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
+  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
+ "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
+ "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
+ "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
+"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,
+
+  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
+    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
+
+# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
+"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
+
+#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
+"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
+
+#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
+"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
+"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
+"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
+
+#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
+"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
+ "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
+ "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
+ "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
+ "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
+ "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
+"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
+"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
+ "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
+ "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
+ "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
+ "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
+
+MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
+
+remove.mice_round2 <- c(
+    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
+    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
+
+    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
+  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
+"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
+  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",
+
+  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
+  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
+"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
+"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
+
+# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
+"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
+"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
+
+#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
+"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
+#SPIC male  x
+#SPIC female  x
+#TOM x
+#AST x
+#CZECH male x
+
+#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
+ "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
+#SPRET female X
+#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
+"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
+)
+#3remove.mice_round2
+
+MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
+MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
+
+all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
+
+```
+
+<!-- 
+two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
+-->
 
@@
 totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
 
@@
 totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
+totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
 
@@
 
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(sex~subsp, scales="free")
-
-tot.sc77 <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(0, 800))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-
-tot.sc_bin <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())  +ylim(c(0, 3000))
-
-#there should be tom
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
+total SC area")
+
+mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
 
@@
 
+```{r Q1.tot.SC_show, echo=FALSE}
+
+Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
+
+Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
+   "WSB","G","LEW","PERC", 
+  "PWD", "SKIVE",  "KAZ",
+  "MSM", "MOLF"), order=T )
+
+Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
+
+#order the strains
+Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
+  arrange(strain) %>%
+  mutate(mouse = factor(mouse))
+
+Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
+
+#add category mean
+
+mouse.av_Tot.SC <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(sex~subsp, scales="free")+ 
+  theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(0,3000 ))+ggtitle("Mouse av. total SC area")
+
+mouse.av_Tot.SC
+```
+
+<!-- can't figure out the ordering -->
+
 The full mFull data set of SC lengths shows that
@@
 
-Total SC
+## Total SC
+
+`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
+X observations removed for having extra signal from other cells.
+`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
 
 ```{r show.tot.sc, echo=FALSE}
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(sex~subsp, scales="free")
-
-tot.sc77 <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(150, 650))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-tot.sc77
-
-tot.sc_bin <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())  +ylim(c(0, 3000))
+
+#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, mouse) %>%
+#  mutate(file.name = factor(file.name))
+
+mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
+mouse.av_sc_male
+
 ```

### 2020-02-16 doc/Results.rmd
Q1 SC section. good motivating sentence

@@
 #skeletonize
-SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
 
@@
 all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
-
 ```
@@
 ```
+
+```{r, tot.SC.setup, echo = FALSE}
+
+totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
+                 .(sex,strain),
+                 summarize, 
+                 ncells = sum(ncells),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(mean.skel, na.rm = TRUE),
+                 mean.bin =  mean(mean.bin, na.rm = TRUE)
+
+)
+
+totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
+
+```
+
 From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
@@
 
+Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of PWD and MSM.
+A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
+
+```{r Q1.Short.Long_biv.setup, echo=FALSE}
+Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength)
+                         )
+
+Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
+Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
+Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
+
+Q1.real.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength)
+                         )
+
+Q1.real.long_mouse_table <- add_strain(Q1.real.long_mouse_table)
+Q1.real.long_mouse_table <- add_sex(Q1.real.long_mouse_table)
+Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)
+
+```
+
 <!--Brief lit review for SC length and MLH1 expections.
 add XX adjusment / arguments
+
+In order to test one of the major/known correlates with genome wide recombiation rate variation, we compared the average
+
 -->
 
-In order to test one of the major/known correlates with genome wide recombiation rate variation, we compared the average
+The plots below show the distributions of mouse average total SC area.
+
 ```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
@@
 Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-Q1.SC.length_Nested_boxplots_Dom
 
 Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-Q1.SC.length_Nested_boxplots_Musc
 
@@
   ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-Q1.SC.length_Nested_boxplots_Mol
 
@@
 #add category mean
-
-mouse.av_Tot.SC <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(sex~subsp, scales="free")+ 
-  theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(0,3000 ))+ggtitle("Mouse av. total SC area")
-
-mouse.av_Tot.SC
-```
-
-<!-- can't figure out the ordering -->
-
-The full mFull data set of SC lengths shows that
-
-- there is a positive correlation between mean SC length and number of COs per chrm
-
-We expect female SC lengths to be longer (refs).
-In the plot above the SC length ~ higher hand foci cells.
-Any of the chromosome classes above 3, that don't have a higher mean are likely due to low data number.
-For the 0 class chromosomes most all are around the same size of the 1CO distribution.
-Add in the code for the plot under 2COs males and females
-
-- Caveats XX
-
-The most convincing should be the short biv data set
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength)
-                         )
-
-Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
-
-Q1.real.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength)
-                         )
-
-Q1.real.long_mouse_table <- add_strain(Q1.real.long_mouse_table)
-Q1.real.long_mouse_table <- add_sex(Q1.real.long_mouse_table)
-Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)
+#totSC.strain.av.table %>% filter
+totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
+
+totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
+
+Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")
+
+Tot.SC_mouse.av_Dom
+
+totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
+totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
+
+Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
+
+Tot.SC_mouse.av_Musc
+```
+
+these are the number of mice for each of the categories
+
+`r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )`
+
+```{r Q1.total.SC_t.test, echo=FALSE}
+
+#run t-test for male female avergage for each strain
+
+#WSB
+Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
+#G
+Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
+#LEW
+Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
+
+#PWD
+Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
+
+#KAZ
+Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
+
+#SKIVE
+#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
+#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")
+
+#error not enought female observations
+
+#MSM
+Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
+
+```
+
+The full data set of SC lengths illustrates the general pattern than females have longer SC-AE.
+
+The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
+`r Q1_tot.sc_t.test.WSB$p.value`, `r Q1_tot.sc_t.test.G$p.value`, and `r Q1_tot.sc_t.test.LEW$p.value`
+for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _ respectively).
+
+Some of the p values for PWD and MSM are much closer (less significant), ( p = `r Q1_tot.sc_t.test.PWD$p.value`, `r Q1_tot.sc_t.test.KAZ$p.value`, and `r Q1_tot.sc_t.test.MSM$p.value` for the _M.
+m.
+musculus^PWD^ _, _M.
+m.
+musculus^KAZ^ _, and _M.
+m.
+molossinus^MSM^_ respectively)
+
+```{r Q1_SC.biv.short_ttest, echo=FALSE}
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+#WSB
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_shortSC_t.test.MSM <- t.test((Q1.real.short_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Q1.real.short_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
+
+```
+
+the pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data
+
+with the mouse average for 'short' bivalents
+
+(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r Q1_mus.av_shortSC_t.test.G$p.value`, and `r Q1_mus.av_shortSC_t.test.LEW$p.value`
+for WSB, G and LEW respectively.
+
+(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`
+
+for PWD, KAZ, SKIVE and MSM.
+
+<!--
+for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
+-->
+
+```{r Q1_SC.biv.long_ttest, echo=FALSE}
+
+#Q1.real.long_mouse_table
+
+Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
 
@@
 
-Chromosome size effect.
-
-$$mouse \ average \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-Below's the code for Mixed model results -- try to organize them into a table or something
-
+mixed models and glms were donw to test for evolution
+
+$$mouse \ average \ total \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.lmer_totalSC, echo=FALSE}
+
+Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
+str(Q1.lmer_totSC.mouse.av.table)
+
+#unorder factors
+Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
+
+#TOTAL SC AREA!!
+lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
+
+lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
+
+lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
+
+lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
+sex.results =  lmer.Q1_mus.av_totSC_sex, 
+               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
+lmer.Q1_mus.av_totSC_strain)
+
+Q1.lmer.totSC_results
+
+#G
+```
+there are more significant effects for strain and subsp for the mixed model of mouse.av total sc
+
+$$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.glm_totalSC, echo=FALSE}
+#total-SC
+
+#Q1.lmer_totSC.mouse.av.table
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+summary(Q1.glm_totSC_M2)
+#subsp effect, many strain, #male
+
+#G * malev.sig and G*LEW
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+summary(Q1.glm_totSC_M3)
+#LOT OF SIGNIFICant strain effects in this model
+
+#Dom.all
+#Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.Dom)
+
+#Dom.1CO
+#Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.1CO.Dom)
+
+#Dom.2CO
+#Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.2CO.Dom)
+
+#Musc.all
+#Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.Musc)
+
+#Musc.1CO
+#Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.1CO.Musc)
+
+#Musc.2CO
+#Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.2CO.Musc)
+
+#Mol.all
+#Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.Mol)
+
+#Mol.1CO
+#Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.1CO.Mol)
+
+#Mol.2CO
+#Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.2CO.Mol)
+```
+
+<!--
 Because of confonding effects across chromsomes of different physical lenghts, we attempt to ...
@@
 The mouse means for these two data sets were then calculated.
-
-```{r Q1_short_biv_lmer, echo=FALSE}
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#all pooled SC lengths
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-
-```
-
-These plot above show that qualitatively, the average female bivalent SC length is significantly longer than the mean length for males from the same strain.
-
-```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE}
-###
-lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value
-
-Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
-sex.results =  Q1lmer.MM_SC_long_sex, 
-               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
-Q1lmer.MM_SC_long_strain)
-
-Q1lmer.MM_SC_long_results
-
-```
-
-Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
-
-```{r Q1other_SC_lmer, echo=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-Q1.lmer_SCLength.1CO_results
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-Q1.lmer_SCLength.2CO_results
-
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
-
-The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
-
-$$mouse \ average \ SC  \ length ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.SC.Length_M2, echo=FALSE}
-#feed these 2 DFs into M2
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-#glm(metric ~ strain)  ---mean_SC
-
-Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_SCLength_M2)
-
-#Dom.all
-Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.Dom)
-
-#Dom.1CO
-Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.1CO.Dom)
-
-#Dom.2CO
-Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.2CO.Dom)
-
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.Musc)
-
-#Musc.1CO
-Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.1CO.Musc)
-
-#Musc.2CO
-Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.2CO.Musc)
-
-#Mol.all
-Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.Mol)
-
-#Mol.1CO
-Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.1CO.Mol)
-
-#Mol.2CO
-Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.2CO.Mol)
-```
-
-```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-```{r Q1.SC.M2plot, echo=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#mouse av
-Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-```{r strain.SClength.t.test, echo=FALSE}
-#t.tests for all the strain specific data frames
-
+-->
+
+```{r Q1.SC_short_t.test, echo=FALSE}
+
+#t.tests for a
+Q1_short.sc_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+#sig
+
+#Using the full data set has strong signle but -- the interpertation is complicated
+#FULL SC t.test
 #run a t.test
@@
 #MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
+
+```
+
+```{r Q1_short_biv_lmer, echo=FALSE}
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#all pooled SC lengths
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+Q1lmer.MM_SC_short_results
+
+```
+
+These plot above show that qualitatively, the average female bivalent SC length is significantly longer than the mean length for males from the same strain.
+
+```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE, include=FALSE}
+###
+lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value
+
+Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
+sex.results =  Q1lmer.MM_SC_long_sex, 
+               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
+Q1lmer.MM_SC_long_strain)
+
+```
+
+Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
+
+```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
+
+The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
+
+$$mouse \ average \ SC  \ length ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.SC.Length_M2, echo=FALSE}
+#feed these 2 DFs into M2
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+#glm(metric ~ strain)  ---mean_SC
+
+Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
+#summary(Q1.glm_SCLength_M2)
+
+#Dom.all
+Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.Dom)
+
+#Dom.1CO
+Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.1CO.Dom)
+
+#Dom.2CO
+Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.2CO.Dom)
+
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.Musc)
+
+#Musc.1CO
+Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.1CO.Musc)
+
+#Musc.2CO
+Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.2CO.Musc)
+
+#Mol.all
+Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.Mol)
+
+#Mol.1CO
+Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.1CO.Mol)
+
+#Mol.2CO
+Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.2CO.Mol)
+```
+
+```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+```{r Q1.SC.M2plot, echo=FALSE}
+
+#removing these plots for space
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#mouse av
+Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
 

### 2020-02-16 doc/Results.rmd
adding notes and cutting fat for Q1-SC results

@@
 
-```{r MixedModel.2, echo=FALSE,  digits = 3}
+```{r MLH1.glmM2, echo=FALSE,  digits = 3}
 #strain = fixed
@@
 
-```{r MixedModel.3, echo=FALSE, digits = 3}
+```{r MLH1.glm.M3, echo=FALSE, digits = 3}
 
@@
 ```{r Q1.Short.Long_biv.setup, echo=FALSE}
+
 Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
@@
 Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
+Q1.real.short_mouse_table <- add_species(Q1.real.short_mouse_table)
+
+Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+Q1.real.short_mouse_table$species <- factor(Q1.real.short_mouse_table$species, ordered = FALSE)
 
@@
 
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_shortSC_t.test.MSM <- t.test((Q1.real.short_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.short_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-```
-
-the pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data
-
-with the mouse average for 'short' bivalents
-
-(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r Q1_mus.av_shortSC_t.test.G$p.value`, and `r Q1_mus.av_shortSC_t.test.LEW$p.value`
-for WSB, G and LEW respectively.
-
-(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`
-
-for PWD, KAZ, SKIVE and MSM.
-
-<!--
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
--->
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-
-```
-
-```{r show.short.biv, echo=FALSE}
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
-Q1.SC.short.mouse_scatter_boxplot
-```
-
-```{r show.long.biv, echo=FALSE}
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-Q1.SC.long.mouse_scatter_boxplot
-```
-
-mixed models and glms were donw to test for evolution
-
 $$mouse \ average \ total \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
@@
 Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-str(Q1.lmer_totSC.mouse.av.table)
+
+#str(Q1.lmer_totSC.mouse.av.table)
 
@@
 
-```{r Q1.glm_totalSC, echo=FALSE}
+```{r Q1.glm_totalSC, echo=FALSE, }
 #total-SC
@@
 
-#Dom.all
-#Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.Dom)
-
-#Dom.1CO
-#Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.1CO.Dom)
-
-#Dom.2CO
-#Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.2CO.Dom)
-
-#Musc.all
-#Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.Musc)
-
-#Musc.1CO
-#Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.1CO.Musc)
-
-#Musc.2CO
-#Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.2CO.Musc)
-
-#Mol.all
-#Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.Mol)
-
-#Mol.1CO
-#Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.1CO.Mol)
-
-#Mol.2CO
-#Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.2CO.Mol)
-```
+```
+
+```{r show.short.biv, echo=FALSE}
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+
+Q1.SC.short.mouse_scatter_boxplot
+
+#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
+skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+#skiveQ1.SC.short.mouse_scatter_boxplot
+#there are 2 lowish female means,
+
+```
+
+When skive is plotted alone it does have overlapping ranges.
+
+```{r Q1_SC.biv.short_ttest, echo=FALSE}
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+#WSB
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
+#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
+#1 male, 3 females measures
+#try to merge the manual - whole cell observations (at least 2 males)
+
+```
+
+The pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data
+
+with the mouse average for 'short' bivalents
+
+(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r Q1_mus.av_shortSC_t.test.G$p.value`, and `r Q1_mus.av_shortSC_t.test.LEW$p.value`
+for WSB, G and LEW respectively.
+
+(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value` for PWD, KAZ, SKIVE
+
+for the short bivalents are not significantly differen, for this strain there are 4 female and 4 male
+
+** there are more SKIVE female mouse observations -- but they are filtered out because I excluded them from the MLH1 dataset.  -- for low quality -- they might have good total SC spreads -- // but
+
+* it think total sc only has 1 female observation but the short bivalent has 4 skive females
+
+<!--
+for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
+-->
+
+Mixed models for the sex differences in mouse averages for short bivalents (reduced dataset)
+
+$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1_short_biv_lmer, echo=FALSE}
+
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+Q1lmer.MM_SC_short_results
+
+```
+
+Sex is the most significant factor for the mixed model of mean short bivalents.
+
+$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$
+
+```{r Q1.SC.Length_M2, echo=FALSE}
+
+Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+
+round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
+
+#skive strin very slight sign
+#male highly sig
+#skive * male slightly sig
+
+```
+
+skive strin very slight sign, male highly sig,  skive * male slightly sig
+
+$$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.glm_short.SC_M3, echo=FALSE}
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
+
+summary(Q1.glm_short.SC_M3)
+
+summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
+
+round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#sex male very sig -- only sig
+```
+
+male is the only significant p value from this model.
 
@@
 
-```{r Q1.SC_short_t.test, echo=FALSE}
+Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
+
+Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
+
+The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
+
+```{r show.long.biv, echo=FALSE, include=FALSE}
+#mouse av
+Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
+
+```
+
+```{r JUNK.Q1.SC_short_t.test, echo=FALSE, include=FALSE}
 
@@
 
-```{r Q1_short_biv_lmer, echo=FALSE}
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#all pooled SC lengths
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-
-```
-
-These plot above show that qualitatively, the average female bivalent SC length is significantly longer than the mean length for males from the same strain.
+```{r Q1.strain.specific.junk, include=FALSE}
+
+#Dom.all
+#Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.Dom)
+
+#Dom.1CO
+#Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.1CO.Dom)
+
+#Dom.2CO
+#Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1glm_SCLength_M2.2CO.Dom)
+
+#Musc.all
+#Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.Musc)
+
+#Musc.1CO
+#Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.1CO.Musc)
+
+#Musc.2CO
+#Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1glm_SCLength_M2.2CO.Musc)
+
+#Mol.all
+#Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.Mol)
+
+#Mol.1CO
+#Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.1CO.Mol)
+
+#Mol.2CO
+#Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1glm_SCLength_M2.2CO.Mol)
+```
+
+```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
+
+#Q1.real.long_mouse_table
+
+Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+
+```
+
+```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
+
+#removing these plots for space
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#mouse av
+Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE}
+#removed these plots for space
+
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+Q1.long.biv_cell_table <- ddply(real.long.bivData, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+Q1.long.biv_cell_table <- add_mouse(Q1.long.biv_cell_table)
+Q1.long.biv_cell_table <- add_strain(Q1.long.biv_cell_table)
+Q1.long.biv_cell_table <- add_sex(Q1.long.biv_cell_table)
+Q1.long.biv_cell_table <- add_subsp(Q1.long.biv_cell_table)
+
+Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+#mouse av for long biv
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.sc = mean(chromosomeLength)
+                         )
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+```
 
@@
 ```
-
-Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
-
-The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
-
-$$mouse \ average \ SC  \ length ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.SC.Length_M2, echo=FALSE}
-#feed these 2 DFs into M2
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-#glm(metric ~ strain)  ---mean_SC
-
-Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_SCLength_M2)
-
-#Dom.all
-Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.Dom)
-
-#Dom.1CO
-Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.1CO.Dom)
-
-#Dom.2CO
-Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.2CO.Dom)
-
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.Musc)
-
-#Musc.1CO
-Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.1CO.Musc)
-
-#Musc.2CO
-Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.2CO.Musc)
-
-#Mol.all
-Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.Mol)
-
-#Mol.1CO
-Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.1CO.Mol)
-
-#Mol.2CO
-Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.2CO.Mol)
-```
-
-```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-```{r Q1.SC.M2plot, echo=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#mouse av
-Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-```{r Q1.SC.long.biv_wacko, echo=FALSE}
-#removed these plots for space
-
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-Q1.long.biv_cell_table <- ddply(real.long.bivData, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-Q1.long.biv_cell_table <- add_mouse(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_strain(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_sex(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_subsp(Q1.long.biv_cell_table)
-
-Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-#mouse av for long biv
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-```
-
-<!-- 
-lmr - sex is most significant effect
-
-plots,
-
-- ranges overlap,
-
-- try the short and long mouse averages
-
-G and MSM female?
-
-G male vs female
-
--->
 

### 2020-02-17 doc/Results.rmd
trying to finalize the Q1.1CO pos section

@@
 
-```{r FEMALE.TABLE, echo=FALSE}
+```{r MLH1.FEMALE.TABLE, echo=FALSE}
 
@@
 
-```{r MALE.TABLE, echo=FALSE}
-
+```{r MLH1.MALE.TABLE, echo=FALSE}
 male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
@@
 
-```{r MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
+```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
 #DF.HetC.MixedModel.HQ
@@
 
+```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
+
+#this block should be in the data set up
+#the mouse averages will be used
+
+mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+mouse.avs_4MM <- add_sex(mouse.avs_4MM)
+mouse.avs_4MM <- add_strain(mouse.avs_4MM)
+mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
+mouse.avs_4MM <- add_category(mouse.avs_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
+mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
+mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
+
+#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
+#code if males are high or low?
+
+#code intp chr
+
+#assign high and low
+#divide musc males into groups
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
+
+```
+
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+
+#add this to setup
+
+#remove female, 
+#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
+
+#Curated_BivData_male
+
+male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+male.av_4MM <- add_sex(male.av_4MM)
+male.av_4MM <- add_strain(male.av_4MM)
+male.av_4MM <- add_subsp(male.av_4MM)
+male.av_4MM <- add_category(male.av_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
+male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
+male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
+```
+
 ```{r totSC_setup, echo=FALSE}
@@
 
-```{r tot.sc.viz4outliers, echo=FALSE}
+```{r totSC.viz4outliers, echo=FALSE}
 
@@
 
+<!--
 From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
@@
 skel -- takes the skeletonized midline of the bin objects
-
+-->
 # Main Figure
@@
 m.
-musculus^ PWD}^_, _M.
+musculus^PWD^_, _M.
 m.
-musculus^ SKIVE^_, and _M.
+musculus^SKIVE^_, and _M.
 m.
@@
 domesticus^PERC^_ respectively.
-These
 
 <!-- mouse level means-->
-The ranges of mean MLH1 counts per cell are
 
@@
 
-# Analysis for Evolutionary Patterns
+# Analysis of Evolutionary Patterns
 
@@
 meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
--->s
+-->
 
@@
 
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
 
@@
 F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  = sum(Ncells),
-                        mean_co =as.numeric(format(round(mean(mean_co), 5 ), nsmall=3) ),
+
+                        Nmice = length(unique(mouse)),
+
+                        total.cells  = sum(Ncells),
+
+                        mouse.mean_co = mean(mean_co),
 
@@
 
-                        mean.mouse_sd   = mean(sd),
+                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
 
@@
 F.strain.table
+
+# I don't know why most of the table is blank
 ```
@@
 
-In the dom  subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M.
-m.
-domesticus$^WSB^_
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_
+In the dom  subspecies the ranges of mouse means are 
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to
+
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M. m. domesticus$^WSB^_ ,
+
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to
+
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_
 
@@
 
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
 #try customize the plot size
@@
 M.strain.table
+#same issue as the female table .. ugh
 ```
@@
 
+<!--
 (discription of the nice boxplot figure)
@@
 - the ratios of DMC1 : MLH1 counts!!
-
+-->
 ```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
@@
 
-```{r DMC1_MLH1.cor, echo=FALSE}
+```{r DMC1_MLH1.cor, echo=FALSE, warning = FALSE}
 #make category averages from AP_mouse_table.HQ
@@
 PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-```
-
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-
-Why calculate ratios using the mean Leptotene cells -- they represent the a metric for the CO : NCO decision ( DSB / COs = NCO)
-
-ratios are .. not that different (slightly suprisingly)
+
+as.data.frame(DMC1table.main)
+
+```
+
+The table above indicates, there is not much variation across the MLH1 : DMC1 ratio (a proxy for the proportion of NCO).
+
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`. The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
 
@@
 m.
-domesticus^G^_, , _M.
+domesticus^G^_, _M.
 m.
-molossinus^MSM^_ and  _M.
+molossinus^MSM^_, and  _M.
 m.
@@
 These are the DSB:CO ratios for the L means `r DMC1table.main$MLH1.L_ratio` and Z means `r DMC1table.main$MLH1.Z_ratio`.
-
-(maybe ignore KAW.L? 1 cell)
 
@@
 
+<!--
 Since there is evidence for non-equal variance across strains for zygotene cells, so don't rely on lm()s that estimate the effect of strains.
@@
 The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the zygotene cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
+-->
 
@@
 
-ToDo: chisq tests for difference in proportions, 
+ToDo:
 - within strain across sex,
@@
 -->
-
-<!-- merging the 2 data sets-->
 
@@
 The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.
-
-```{r Q1.biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}
-
-#this block should be in the data set up
-
-#the mouse averages will be used
-
-mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-mouse.avs_4MM <- add_sex(mouse.avs_4MM)
-mouse.avs_4MM <- add_strain(mouse.avs_4MM)
-mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-mouse.avs_4MM <- add_category(mouse.avs_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
-mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
-mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
-
-#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-#code if males are high or low?
-
-#code intp chr
-
-#assign high and low
-#divide musc males into groups
-mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
-
-```
-
-In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#add this to setup
-
-#remove female, 
-#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
-
-#Curated_BivData_male
-
-male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-male.av_4MM <- add_sex(male.av_4MM)
-male.av_4MM <- add_strain(male.av_4MM)
-male.av_4MM <- add_subsp(male.av_4MM)
-male.av_4MM <- add_category(male.av_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
-male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
-male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-```
 
@@
 Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-summary(Q1.glm_totSC_M2)
+#summary(Q1.glm_totSC_M2)
 #subsp effect, many strain, #male
@@
 Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-summary(Q1.glm_totSC_M3)
+#summary(Q1.glm_totSC_M3)
 #LOT OF SIGNIFICant strain effects in this model
@@
 ```{r Q1.SC.Length_M2, echo=FALSE}
-
 Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
@@
 
-```{r JUNK.Q1.SC_short_t.test, echo=FALSE, include=FALSE}
-
-#t.tests for a
-Q1_short.sc_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-#sig
-
-#Using the full data set has strong signle but -- the interpertation is complicated
-#FULL SC t.test
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-#run a t.test seperate out by the chrm classes
-#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
 ```{r Q1.strain.specific.junk, include=FALSE}
@@
 
+```{r Q1.F1pos_long.short_setup, echo=FALSE}
+
+Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+Q1.long.1COpos_mouse.av <-  ddply(Q1.long.biv_1CO, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength), 
+                        mean_F1pos_PER =  mean(PER_Foci_1),
+                         mean_F1pos_ABS =  mean(Foci1)
+                         )
+Q1.long.1COpos_mouse.av <- add_strain(Q1.long.1COpos_mouse.av)
+Q1.long.1COpos_mouse.av <- add_sex(Q1.long.1COpos_mouse.av)
+Q1.long.1COpos_mouse.av <- add_subsp(Q1.long.1COpos_mouse.av)
+
+Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+Q1.short.1COpos_mouse.av <-  ddply(Q1.short.biv_1CO, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean_SC = mean(chromosomeLength),
+                        mean_F1pos_PER =  mean(PER_Foci_1),
+                         mean_F1pos_ABS =  mean(Foci1)
+                         )
+
+Q1.short.1COpos_mouse.av <- add_strain(Q1.short.1COpos_mouse.av)
+Q1.short.1COpos_mouse.av <- add_sex(Q1.short.1COpos_mouse.av)
+Q1.short.1COpos_mouse.av <- add_subsp(Q1.short.1COpos_mouse.av)
+
+```
+
+<!-- motivating
+
 To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
-
 In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
+-->
 
 we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
+We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ).
+(all chromosomes in house mouse are acrocentric)
 
@@
 To account for this we again use the reduct bivalent sets (long.biv and short.biv)
+
+Below are the general patterns / plots -- short and long bivalent groups are shown to try and account for the chromsomes size effect.
+
+```{r CO.positions.figs, echo=FALSE, include=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+
+```
+
+```{r Q1show.F1pos_Full, echo=FALSE}
+
+Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \n all bivalents")
+Q1.mouse.av.biv_scatr_box
+
+```
+
+```{r Q1_1COpos.t.test, echo=FALSE}
+#testing,
+#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
+
+#WSB
+Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
+#G
+Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
+#LEW
+Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
+
+#PWD
+Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
+
+#KAZ
+Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
+
+#SKIVE
+Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
+           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
+
+#MSM
+Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
+
+#Single Biv Level
+#single.biv.1CO_pos_t.test <- t.test(
+
+ # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+#  )
+
+#t.tests for mous.av_nrm.1CO.pos 
+```
+
+In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (with female)
+are highly significantly more medially placed (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
+With the musclus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
+Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).
+
+<!-- lmer and glm here -->
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.CO1pos_lmer.M1, echo=FALSE}
+
+#make sure this dataset has all the right strains
+Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
+lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
+
+lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
+lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
+
+lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
+sex.results = lmer.Q1.nrm1CO_M1.sex , 
+               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
+
+lmer.Q1.nrm1CO_M1_results
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+For the Mixed model, sex and strain are significant effects for the normalized single foci position.
+
+$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
+
+```{r Q1.CO1pos_glm.M2, echo=FALSE}
+
+#in case factors aren't ordered
+#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
+summary(Q1.glm_F1pos_M2)
+
+summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
+round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=FALSE}
+
+summry.Q1.glm_F1pos_M2
+
+Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
+summary(Q1.glm_1CO.pos_M3)
+
+summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
+round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+```
+
+Lew and KAZ had slightly significant strain speicific effects across the two glms (these results will be explored more fully in the Q2 section).
+The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).
 
@@
 
-```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, eval=FALSE}
-
-Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-Q1.short.biv_scatr_box
-
-```
-
-```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE}
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-Q1.long.1COpos_mouse.av <-  ddply(Q1.long.biv_1CO, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength), 
-                        mean_F1pos_PER =  mean(PER_Foci_1),
-                         mean_F1pos_ABS =  mean(Foci1)
-                         )
-Q1.long.1COpos_mouse.av <- add_strain(Q1.long.1COpos_mouse.av)
-Q1.long.1COpos_mouse.av <- add_sex(Q1.long.1COpos_mouse.av)
-Q1.long.1COpos_mouse.av <- add_subsp(Q1.long.1COpos_mouse.av)
-
-Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-Q1.short.1COpos_mouse.av <-  ddply(Q1.short.biv_1CO, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength),
-                        mean_F1pos_PER =  mean(PER_Foci_1),
-                         mean_F1pos_ABS =  mean(Foci1)
-                         )
-
-Q1.short.1COpos_mouse.av <- add_strain(Q1.short.1COpos_mouse.av)
-Q1.short.1COpos_mouse.av <- add_sex(Q1.short.1COpos_mouse.av)
-Q1.short.1COpos_mouse.av <- add_subsp(Q1.short.1COpos_mouse.av)
-
-```
-
 ```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
@@
 
-As expected -- confirming previous results, the normalized position of single crossovers varies between males and females.
-The strain effects indicate that within the musculus strains the difference is less compared to
-This is an indication of evolution of the sexual dimorphism for the pattern of single foci positions.
-
-The long bivalent pattern -- shows a different effect, both female and males have positions close to the center. This is a lack of telomeric pattern efpected for males
-
-For short bivalents, we predict that the normalized position will be more medial
-
-The short bivalent landscape pattern is more medial compared to males in all strains. However some strains have larger sex differences, WSB
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
+#IGNORE THIS FOR KNOW, THIS
+
+#filter all Curated biv data for 1 strain
+#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
+
+# the stardell paper sums the number of COs for bins (10)
+
+#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
+#but then the model would be close to logistic
+
+#summing the number of COs per bin, losses the chrom specific data....
+#!! need to use a melted Data frame to capture all nrm.foci obs
+#wsb.ttest<- mouse.avs_4MM %>% filter()
+#start with same chrm classes
+#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
+
+WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
+
+Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
+
+PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
+
+Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
+Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
+Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
+
+anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
+
+anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
+
+anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
+
+#or AIC(model1,model2)
 
@@
 ```
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.nrmF1CO.pos_M1, echo=FALSE}
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
-anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
-anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)
-
-drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )
-
-exactRLRT(lmer.Q1.nrm1CO_M1)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.Nrm.F1pos_M2, echo=FALSE}
-
-#calq the means for normalized F1
-#M2
-
-Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_nrmF1pos_M2)
-
-summry.Q1.glm_nrmF1pos_M2 <- summary(Q1.glm_nrmF1pos_M2)
-round(summry.Q1.glm_nrmF1pos_M2$coefficients[,c(1,4)],5)
-
-write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.nrmCO.pos_M3, echo=FALSE, eval=FALSE}
-
-Q1.glm_nrmF1pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
-#summary(Q1.glm_nrmF1pos_M2)
-
-summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
-round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)
-
-summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
-round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)
-
-write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-```
-
-The model results aren't as clear as I'd like.
-For the Mol, both the sex effect and MOLF strain effect are significant (this means both female and male rec landscapes are effected - most of the time towards the middle
-
-GENERALLY - Sex is the biggest effect for the 1CO landscape (female middle male telomereic)
-molf
-Not sure I understand how MSM is supper significant for 
-There must be some strange effect due to the Mol subsp - MOLF isn't even in M2 (
-Male is consistently significant across the two models.
-
-Male and MOLF are the most significant effects for M3
 
@@
 summary(log.reg_1CO_pos_biv)
-
-```
-
-```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}
-
-#t.test for high and low group  -- logistic regression still need to format the high and low group
-
-#mouse av
-t.test_mouse.av.1CO_pos <- t.test(
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv.1CO_pos_t.test <- t.test(
-
-  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-  )
 

### 2020-02-17 doc/Results.rmd
added interpertation notes for DSB section

@@
 
+<!--
+interpertation notes,
+
+the number of DMC1 across stages (L and Z stages) are driven by different equations
+
+Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
+
+Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
+
+*the repair decision can happen at both stages and remove DMC1.
+formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
+
+back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
+
+expectations, predictions if this is the case
+
+-->
+
 ```{r DMC1.setup, echo=FALSE}

### 2020-02-17 doc/Results.rmd
Finalizing the DMC1 section, hopefully

@@
 
-The same model framwrok was applied to the male specific data.
+The same model framework was applied to the male specific data.
 The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
@@
 
-<!--
-interpertation notes,
-
-the number of DMC1 across stages (L and Z stages) are driven by different equations
-
-Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
-
-Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
-
-*the repair decision can happen at both stages and remove DMC1.
-formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
-
-back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
-
-expectations, predictions if this is the case
-
--->
-
 ```{r DMC1.setup, echo=FALSE}
@@
 
-DMC1.mean.table
-
 DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
@@
 
-```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#these are histogram plots for testing Normality shape -- move to supplemenatry section
-
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-DMC1.histogram
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-
-#The counts look close to normal.
-#Potentially figure 2. Using a function to #display the p-vales for mean comparison
-
-```
-
-In order to get a good idea on wether variation with mean CO number per cell is associated with the number of percursors upstream in prophase, we quantified DMC1 foci, a marker for DSBs.
-DMC1 foci were scored from X `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of juvenile mice (12-14-18 days).
-A subset of strains were quantified  _M.
-m.
-musculus^PWD^_, _M.
-m.
-molossinus^MSM^_, _M.
-m.
-musculus^KAZ^_ ,  _M.
-m.
-domesticus^WSB^_ , and _M.
-m.
-domesticus^G^_.
-Leptotene and zygotene cells were staged based on SC-AE and centromere morphology.
-
-```{r DMC1.plot, echo=FALSE, warning=FALSE}
+```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
 #think of better color scheme
@@
 DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black"))
@@
 
-<!--
-(discription of the nice boxplot figure)
-
-- overall the range is very high (CO homeostasis, technical noise, process is noiser continous repair and new DSBs)
-
-- exclude KAZ?
-
-- expected difference between cell stages for all strains
-
-- pattern of musc strains being higher than the dom strains (subspe pooled t-test)
-
-- MSM is much higher
-
-- the ratios of DMC1 : MLH1 counts!!
--->
 ```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
@@
 
-the pvalues for the differences between time points are  `r ttest.ZtoL$p.value`,  `r ttest.ZtoL.high$p.value` , and  `r ttest.ZtoL.Low$p.value` for all observations, the high rec group and the low group respectively.
-
-```{r plot.t.test, echo=FALSE}
+```{r plot.t.test, echo=FALSE, include=FALSE}
 
@@
 #high vs low group (L)
-ttest.HighLow.L <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
 # p-value =  0.002145
@@
 #high vs low group (Z)
-ttest.HighLow.Z <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
 # p-value = 0.6628
@@
 
-<!--
-Subsampling code -- don't report
-
-The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
-Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
-Permutations were done for both cell stages.
-The actual p value is in red.
-
-The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
-
-```{r DMC1.subsampling, echo=FALSE}
-
-#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
-# the 2 groups
-
-#make the 2 pools
-High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-
-Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-
-#permute drawing / subsampling and performing the t-test
-#put the p.value results into a DF!
-
-#permut.SC.cv.diff  <- as.data.frame(
-#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
-
-#permuting samples -- and re-licating t-tests
-
-#nrep <- 1000
-
-#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
-bunchOpvals.L <- replicate(1000, 
-    # t.test
-    t.test( sample(High.L.pool, 25, replace = FALSE),
-            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
-          )
-bunchOpvals.L <- as.data.frame(bunchOpvals.L)
-bunchOpvals.L$stage <- "L"
-
-bunchOpvals.Z <- replicate(1000, 
-    # t.test
-    t.test( sample(High.Z.pool, 25, replace = FALSE),
-            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
-          )
-
-bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
-bunchOpvals.Z$stage <- "Z"
-
-#could merge the lists of p-values
-replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
-#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
-
-#ASSUMPTIONS; normality, equal variance
-
-#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-#plot.pvals.L
-#plot.pvals.Z
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
--->
-
-```{r DMC1_MLH1.cor, echo=FALSE, warning = FALSE}
+```{r DMC1_anova, echo=FALSE, warning=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+
+#unorder strain
+DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
+
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
+z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
+```
+
+```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
 #make category averages from AP_mouse_table.HQ
@@
 ##RATIOS
-DMC1table.main$MLH1.L_ratio <- (DMC1table.main$mean.count.L / DMC1table.main$mean.MLH1)
-DMC1table.main$MLH1.Z_ratio <- (DMC1table.main$mean.count.Z / DMC1table.main$mean.MLH1)
+DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
+DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
 
 #DMC1.mean.table
+DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
+
+colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
+                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
+
+DMC1.mean.table.99
 
@@
 
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
 as.data.frame(DMC1table.main)
 
-```
-
-The table above indicates, there is not much variation across the MLH1 : DMC1 ratio (a proxy for the proportion of NCO).
-
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`. The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
+In an attempt to locatize the above male specific crossoever number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase meiocytes.
+
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of (X for each strain) juvenile mice (12 to 18 days).
+
+A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _ musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
+One mouse was sampled for each strain.
+
+As can be seen in the figure, the high recombining groups have significantly more DMC1 for for leptotene (t.test; `r ttest.HighLow.L.pval` , one-way-anova; `r round(l.anova.pval, 5)`), but not the zygotene cells (t.test; `r ttest.HighLow.Z.pval`, one-way-anova; `r z.anova.pval`).
+
+The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio).
+Notably there is not much variation and do not show a significant difference between the high and the low recombining groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+These results indicate that the best predictor for CO number may be established before DSB formation (cite Baier et al 2014).
+
+<!--
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
 `r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M.
@@
 musculus^PWD^_ respecitively.
-
-These are the DSB:CO ratios for the L means `r DMC1table.main$MLH1.L_ratio` and Z means `r DMC1table.main$MLH1.Z_ratio`.
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE, eval=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova <- anova(mod.L) #0.0002701 ***
-z.anova <- anova(mod.Z) #NS
-
-l.anova
-z.anova
+-->
+
+```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
+
+#Subsampling code -- don't report
+
+#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
+Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
+Permutations were done for both cell stages.
+The actual p value is in red.
+
+#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
+
+#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
+# the 2 groups
+
+#make the 2 pools
+High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+
+Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+
+#permute drawing / subsampling and performing the t-test
+#put the p.value results into a DF!
+
+#permut.SC.cv.diff  <- as.data.frame(
+#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
+
+#permuting samples -- and re-licating t-tests
+
+#nrep <- 1000
+
+#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
+bunchOpvals.L <- replicate(1000, 
+    # t.test
+    t.test( sample(High.L.pool, 25, replace = FALSE),
+            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
+          )
+bunchOpvals.L <- as.data.frame(bunchOpvals.L)
+bunchOpvals.L$stage <- "L"
+
+bunchOpvals.Z <- replicate(1000, 
+    # t.test
+    t.test( sample(High.Z.pool, 25, replace = FALSE),
+            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
+          )
+
+bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
+bunchOpvals.Z$stage <- "Z"
+
+#could merge the lists of p-values
+replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
+#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
+
+#ASSUMPTIONS; normality, equal variance
+
+#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
+
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
+#plot.pvals.L
+#plot.pvals.Z
+
+#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
+#n use the var.equal = TRUE
+#https://www.statmethods.net/stats/ttest.html
+```
+
+```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#these are histogram plots for testing Normality shape -- move to supplemenatry section
+
+#- check that distributions are close to normal
+#- check for equal variance.
+
+DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
+
+DMC1.histogram
+#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
+#bartlet's test, Levine's test
+
+#The counts look close to normal.
+#Potentially figure 2. Using a function to #display the p-vales for mean comparison
+
 ```
@@
 <!--
-Since there is evidence for non-equal variance across strains for zygotene cells, so don't rely on lm()s that estimate the effect of strains.
-
-The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the zygotene cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
+interpertation notes, 
+the number of DMC1 across stages (L and Z stages) are driven by different equations
+
+Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
+
+Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
+
+*the repair decision can happen at both stages and remove DMC1.
+formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
+
+back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
+
+expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
+OR
+
+variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
+-Baier et al 2014 supports this model (Chmr axis / DNA loop model)
 -->
@@
 ```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
-
 male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,

### 2020-02-18 doc/Results.rmd
finer editing for the begining MLH1 section

@@
 
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")
 
@@
 
-```{r table.for.plot, echo=FALSE}
+```{r table.for.plot, echo=FALSE, warning=FALSE}
 #why do I add in a different table?
@@
 ```{r main.fig1s, echo=FALSE, include=FALSE}
-
 strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
@@
 main.table <- add_subsp(main.table)
+main.table <- add_species(main.table)
 
@@
 
-In order to estimate the genome wide recombination rates across our panel of `r length(unique(main.table$strain))` strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)`.
-
+In order to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` _M.
+m.
+musculus_ strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)` oocytes.
+
+(mouse means were calculated and statisics in Table 2)
+
+Figure 1(A) shows the --
+
+--notes on error /
+
+-- connections to previous measures in the literature
+
+<!--
 The strain means for females were pretty similar.
-For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _M.
-m.
-domesticus^WSB^_, _M.
-m.
-domesticus^G^_, and _M.
-m.
-domesticus$^LEW^_ respectively.
-In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _M.
-m.
-musculus^PWD^_, _M.
-m.
-musculus^SKIVE^_, and _M.
-m.
-musculus^KAZ^_ respectively.
-In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _M.
-m.
-molossinus^ MSM^_ and _M.
-m.
-molossinus^MOLF^_ respectively.
-
-While in two of the subspecies, the means for males show a greater range across means, notable in the mmmolossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _M.
-m.
-molossinus^MSM^_ and  _M.
-m.
-molossinus^MOLF^_ respectively.
-In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _M.
-m.
-musculus^PWD^_, _M.
-m.
-musculus^SKIVE^_, _M.
-m.
-musculus^KAZ^_, _M.
-m.
-musculus^TOM^_, _M.
-m.
-musculus^AST^_, and _M.
-m.
-musculus^CZECH^_ respectively.
-In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _M.
-m.
-domesticus^WSB^_, _M.
-m.
-domesticus^G^_, _M.
-m.
-domesticus^LEW^_ and _M.
-m.
-domesticus^PERC^_ respectively.
-
-<!-- mouse level means-->
+For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively.
+In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively.
+In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively.
+-->
+
+While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively.
+In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively.
+In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.
 
@@
 the male 
-
 - the ranges of mouse means are displayed in Sup table X
-
 - How the means were calculated
-
 -The male gwRR is adjusted for the PAR
 -The means for each strains are report in Table 1
-
 -Normality was assessed for the MLH1 counts (include in Supplement histograms)
-
-, -->
+-->
 
@@
 
-In order to test the effects of subspecies, sex and strain on the mean and variance of MLH1 counts per cell, we fit a data set of `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages for MLH1 foci per cell.
-We constructed the model to set subspecies, sex and their interaction as fixed effects, which strain was coded as a random effect.
-The predicted coeffecients and pvalues are reporter below in table X.
-
+In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
+(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
+
+<!--
 $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
+-->
 <!--
@@
 add number of cells
-
 `r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
-
  $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-In order to understand the variance within an evolutionary framework, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse level MLH1 foci per cell means as the dependant variable with a mixed model.
 
@@
 
-summary(MLH1.MM.M1)$coefficients
+#summary(MLH1.MM.M1)$coefficients
 
@@
 MLH1.MM_strain)
-
-```
-
-Because all of the mixed model effects had significant effects, (p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r  round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect), we followed up with two general linear models to investigate the specific strain effects.
-
+```
+
+Our models were designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
+
+All of the effects from the mixed model significant, (LRT; p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect).
+
+<!--
  $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-
 and 
-
-$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$  -->
 
@@
 
-In the first glm, the fixed subspecies effects were not significant, so we only report for values for the second model which is limited to sex and strain fixed effects.
+We followed up the mixed model effects with general linear models.
+The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions.
+(*what are the points and differences?
+)*
 
@@
 
-round(summ.strain.fixed$coefficients[,c(1,4)],5)
+#round(summ.strain.fixed$coefficients[,c(1,4)],5)
 
@@
 
-In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _M.
-m.
-musculus^PWD^_ and _M.
-m.
-molossinus^MSM^_  respectively.
+In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
 These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.
 
-Additionally, _M.
-m.
-domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
-These results confirm the qualtative pattern of the _M.
-m.
-domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
-
-The second model were strain was (nested and random), the fixed effects, sex and interaction of (subsp and sex), are much more significant.
-The coefficients indicate, males in general have 1 less in average, and the musc and molf subsp have ~3 more on average.
-
-# Within Mouse Variance in CO Count per Cell
-
-In order to more fully discribe the distributions we examined the variance in addition to the mean applying the same mixed models and linear models, with _within mouse variance_ of MLH1 counts per cell as the dependant variable.
-
-<!--
-motivation / background: maybe checkpoint / regulation
-
-limitations: observations at the mouse level, decreased power to estiamate the variance --
-
-setup
-
--quality / technical error
-
-- variance and cv
-
-The mixed models have most significant variables,
-
-M3.glm for variance was less significant (0.044181 ),
-
-the Q12 models in general were more significant
-
-cV's were more significant than var
-
-var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-
-cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
--->
-
-$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$
+Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
+These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
+
+## Within Mouse Variance in CO Count per Cell
+
+In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).
 
@@
 sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-LmerMLH1_VAR_results
 
@@
 sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-LmerMLH1_CV_results
 
@@
 sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-LmerMLH1_M1_Q12_VAR_results
 
@@
 sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-
-LmerMLH1_M1_Q12_VAR_results
-```
-
-These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
-
-```{r cv.plots, echo=FALSE, eval=FALSE}
-
-cV.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x =cV , fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("cV of within mouse variance, all mice")
-cV.Hist.total.mice <- cV.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-cV.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-cV.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = cV, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("cV of within mouse variance, 10 cell min")
-cV.Hist <- cV.Hist   + facet_wrap(~ strain, scales="free")
-cV.Hist
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
-
-#Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
-```
-
-We followed up the mixed models with glms. The coefficients are listed below for the following glms: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
-
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
-
+```
+<!--
+$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$-->
 ```{r M3.MLH1.var, echo=FALSE}
@@
 glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
+#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
 
@@
 #summary(MLH1cv_M4)
-round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
+#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
 
@@
 MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-
 MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
+
 #as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
+#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
 
@@
 MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-
-```
-
-Across all four data sets, the sex effect has the most significant p value.
-The highest p value is for the sex effect for VAR and Cv in the full models.
-This might be due to extra noise from technical sources.
-
+#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
+```
 <!--
+These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
+
 var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
@@
 
-For both the mixed model and linear model, sex was the most signifcant effect (p= `r LmerMLH1_VAR_results$sex.results` p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` (for the high quality dataset)  ) respectively.
-
-The measures for variance are more supciptibal to thechnical error effects (noise).
-For this reasons, we analysze the models by restricting the data to a subset cells which were scored as highest quality.
-Sex was also the most significant effect (p)
-
+For both the mixed model and linear model, sex was the most significant effect while the subspecies, interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively.
+
+*notes on caveats for calculating variance -- lower power
+
+Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
+background noise).
+For this reasons, we replicated the model analysis from data sets in the subset of cells with higher quality scores compared to the full data set.
+(these analysis/ models replicated the results of the full data set (sex being the most significant
+(LTR; p = `r LmerMLH1_M1_Q12_VAR_results$sex.results`, and glm; p = `r  MLH1var_M3_Q12_sum$coefficients[2,4]` )
+
+**female strongest effect, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.
+
+<!--
+motivation / background: maybe checkpoint / regulation
+
+limitations: observations at the mouse level, decreased power to estiamate the variance --
+setup
+-quality / technical error
+- variance and cv
+The mixed models have most significant variables, 
+M3.glm for variance was less significant (0.044181 ), 
+the Q12 models in general were more significant
+cV's were more significant than var
+
+var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
+
+cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
+
+$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
 <!--
@@
 Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
-
 <!--
@@
 
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M3)
+anova(MLH1var_M3)  
+as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
+
+MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+#summary(MLH1cv_M3)
+
+anova(MLH1cv_M3)
+as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
+
+#Q12 var
+MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1var_M3_Q12)
+anova(MLH1var_M3_Q12)  
+as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M3_Q12)
+anova(MLH1cv_M3_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
+
+```
+
 ```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
@@
 
-## Female Specific Analysis
-
-In order to more finely discirbe the sex specific patterns, we foucsed on data sets including single sexes.
+# Female Specific Analysis
+
+In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.
 
@@
 
-                        total.cells  = sum(Ncells),
-
-                        mouse.mean_co = mean(mean_co),
+                        total.cells  = sum(Ncells, na.rm = TRUE),
+
+                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
 
@@
                         #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
     )
@@
 For some strains, we were only able to quantified or include one mouse and the Cast observation comes from 1 1 cell.
-To analyze the female specific dataset,
-
-In the dom  subspecies the ranges of mouse means are 
+To analyze the female specific data set,
+
+In the _domesticus_  subspecies the ranges of mouse means are 
 `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to
 
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M. m. domesticus$^WSB^_ ,
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,
 
@@
 
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _M. m. domesticus$^LEW^_
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _M.
-m.
-musculus$^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _M.
-m.
-musculus$^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _M.
-m.
-molossinus$^MSM^_
-
-Two glm models were run to test the effects of strain and subspecies on mean MLH1 coount per cell per mouse. The coefficients for the following model are below.
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _domesticus^G^_
+
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _ musculus^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_
+
+Two glm models were run to test the effects of strain and subspecies on mean MLH1 count per cell per mouse. The coefficients for the following model are below.
 
@@
 summry.female.strain.effects <- summary(glm.female.strain.subsp)
-round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
 #write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
@@
 
-round(lm.female.strain_sum$coefficients[,c(1,4)],5)
+#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
 
@@
 
-For mean CO count, _M. m. domesticus^G^_  strain effects in both models (p = )
-_M. m. musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).
+For mean CO count, _domesticus^G^_  strain effects in both models (p = )
+_musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).
 
@@
 
-_M.
-m.
-domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means.
-_M.
-m.
-domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _M.
-m.
-musculus^MSM^_ is `r mean.MSM.female / mean.all.other`.
+_domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means.
+_ domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _musculus^MSM^_ is `r mean.MSM.female / mean.all.other`.
 These three will be designated as 'moderate high rec' strains for later sex specific analysis.
@@
 #round( #round isn't working for this sumary
-lmVAR.female.strain_sum$coefficients[,c(1,4)]
+#lmVAR.female.strain_sum$coefficients[,c(1,4)]
 
@@
 #round(
-lm.m1.Q12_female_sum$coefficients[,c(1,4)]
+#lm.m1.Q12_female_sum$coefficients[,c(1,4)]
 #LEW not sig
@@
 
-lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
+#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
 #LEW sig
@@
 
-The variance across mouse means for each strain is also variable across strains. With _M. m. domesticus$^LEW^_ having the most significant strain effect in the second model (p= ).
-
-## Male Specific Analysis
-
-The male specific patterns were also investigated.
+The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (p= ).
+
+# Male Specific Analysis
+
+The male specific patterns were also investigated. (what's a better motivation for this section)
 
@@
 
+#fomo.male_mouse.av
+
 male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
@@
 Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
 ```
@@
 
-#female strain table
-#female_mouse.avs_for.analysis
-M.strain.table <- ddply(just.male_mouse.av, c("strain"), summarise,
+M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                         Nmice =length(unique(mouse)),
-                        Ncells  = sum(Ncells),
-                        mean_co =as.numeric(format(round(mean(mean_co), 4 ), nsmall=3) ),
+
+                        totNcells  = sum(Ncells, na.rm=TRUE),
+
+                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
 
@@
 
-                        mean.mouse_sd   = mean(sd),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se)
+                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
+
+                        #Standard error (mice . cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
     )
@@
 
- _M. m. domesticus^LEW^_
+ _domesticus^LEW^_
 
@@
 
-round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
 
@@
 summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
+#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
 
@@
 lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
+#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
 #non sig
@@
 lmcv.male.strain_sum <- summary(lmcv.male.strain)
-round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
+#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
 #HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
@@
 lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
+#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
 #NS
@@
 lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
+#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
 #NS
-
-```
-
-The coefficients for the variance models are listed above
-
-The strain average for  _M.
-m.
-musculus^PWD^_ , _M.
-m.
-molossinus^MSM^_ and  _M.
-m.
-musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
+```
+
+The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
 Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
@@
 ```{r Q1_short_biv_lmer, echo=FALSE}
-
 lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
@@
 #skive * male slightly sig
-
 ```
@@
 
-```{r Q1.strain.specific.junk, include=FALSE}
-
-#Dom.all
-#Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.Dom)
-
-#Dom.1CO
-#Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table[Q1.lmer_totSC.mouse.av.table$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.1CO.Dom)
-
-#Dom.2CO
-#Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1glm_SCLength_M2.2CO.Dom)
-
-#Musc.all
-#Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.Musc)
-
-#Musc.1CO
-#Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.1CO.Musc)
-
-#Musc.2CO
-#Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1glm_SCLength_M2.2CO.Musc)
-
-#Mol.all
-#Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.Mol)
-
-#Mol.1CO
-#Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.1CO.Mol)
-
-#Mol.2CO
-#Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1glm_SCLength_M2.2CO.Mol)
-```
-
 ```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
@@
 #run t.tests for the mouse av. for reduced single bivalent datasets
-
 ```
@@
 SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
 ```
@@
 Q1lmer.MM_SC_long_strain)
-
 ```
@@
 CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-
 ```
@@
 
-```{r WSB.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#WSB.Bivalents.DF
-
-#table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)#the 2CO's will be used
-WSB.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "WSB",]
-
-#plot the differences
-wsb.ploty_IFD_raw <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_raw
-
-wsb.ploty_IFD_PER <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, WSB")+geom_boxplot(aes(x=sex))
-wsb.ploty_IFD_PER
-
-WSB.IFDnorm_scatter.plot <- ggplot(WSB.Bivalents.DF[WSB.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+        geom_jitter()+ggtitle("WSB \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-WSB.IFDnorm_scatter.plot
-#this plot is much messier than the MSM (PWD?)
-
-#try plotting the other IFD
-
-#make a table of the means and n ect
-wsb.table_IFD <- ddply(WSB.Bivalents.DF, c("sex"), summarise,
-
-                       total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test
-wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])
-
-```
-
-The range of normalized IFDs overlap closer in males and females in the WSB data.
-
-```{r LEW.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-
-LEW.Bivalents.DF_IFD <- LEW.Bivalents.DF[(!is.na(LEW.Bivalents.DF$sex)) & (LEW.Bivalents.DF$hand.foci.count == 2),]
-
-LEW.Bivalents.DF_IFD$sex <- droplevels(LEW.Bivalents.DF_IFD$sex)
-
-#table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= #as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-lew.ploty_IFD_raw <- ggplot( LEW.Bivalents.DF_IFD,  aes(y=IFD1_ABS, x = sex, na.rm = TRUE))+                     geom_jitter( na.rm = TRUE)+ggtitle("Raw IFD, LEW")+geom_boxplot(aes(x=sex, na.rm = TRUE))
-lew.ploty_IFD_raw
-#arg, little annoyed that the x NA won't go away
-
-lew.ploty_IFD_PER <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, LEW")+geom_boxplot(aes(x=sex))
-lew.ploty_IFD_PER
-
-LEW.IFDnorm_scatter.plot <- ggplot(LEW.Bivalents.DF[LEW.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  
-  scale_color_manual(values=c('#56B4E9','#E69F00'))+
-  geom_jitter()+ggtitle("LEW \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-LEW.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-lew.table_IFD <- ddply(LEW.Bivalents.DF, c("sex"), summarise,
-
-                  #     mean_SC = mean(chromosomeLength),
-                       n_SC = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
-
-```
-
-The Lew pattern doesn't have a clean cut off of nrm.IFD. the range of male and females overlap, but there are more female observations below.
-
-```{r G.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-#table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-#plot the metric
-#remove the NAs
-
-G.ploty_IFD_raw <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, G")+geom_boxplot(aes(x=sex))
-
-G.ploty_IFD_raw
-#
-G.ploty_IFD_PER <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, G")+geom_boxplot(aes(x=sex))
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.ploty_IFD_PER
-
-G.IFDnorm_scatter.plot <- ggplot(G.Bivalents.DF[G.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+  geom_jitter()+ggtitle("G \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-G.IFDnorm_scatter.plot
-
-#make a table of the means and n ect
-G.IFD.table <- ddply(G.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
-
-#calQ percent < 30%
-
-oo <- G.Bivalents.DF[ (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER <= .3 ), ]
-oo <- oo[!(is.na(oo$fileName) | oo$fileName == ""), ]
-
-#the na's are being counted (ughhhhhh)
-#use the file thing instead
-
-G.Bivalents.DF <- G.Bivalents.DF[!(is.na(G.Bivalents.DF$fileName) | G.Bivalents.DF$fileName == ""), ]
-
-LT30_PER_male <-  length(G.Bivalents.DF$Obj.ID[ (  (G.Bivalents.DF$sex == "male") & (G.Bivalents.DF$IFD1_PER >= .3 ) ) ]) / 
-  length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "male") ]) #0.8302619
-
-LT30_PER_female <-  length(G.Bivalents.DF$Obj.ID[(G.Bivalents.DF$sex == "female") & (G.Bivalents.DF$IFD1_PER <= .3 ) ]) / length(G.Bivalents.DF$Obj.ID[ (G.Bivalents.DF$sex == "female") ])
-#0.7062663
-
-#would Chi-sq be a good test to compare the ratios ect?
-#the sum of these doesn't add 
-#G.Bivalents.DF
-
-```
-
-```{r PWD.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-#table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#remove the NAs
-
-PWD.ploty_IFD_raw <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+                          geom_jitter()+ggtitle("Raw IFD, PWD")+geom_boxplot(aes(x=sex))
-
-PWD.ploty_IFD_raw
-
-PWD.ploty_IFD_PER <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, PWD")+geom_boxplot(aes(x=sex))
-PWD.ploty_IFD_PER
-
-PWD.IFDnorm_scatter.plot <- ggplot(PWD.Bivalents.DF[PWD.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("PWD \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
-
-PWD.IFDnorm_scatter.plot
-
-#run a t.test seperate out by the chrm classes
-PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])
-
-```
-
-For PWD, there are a few observations of the short IFDs for males, but there seems to be a cut-off / threshold at .3
-
-```{r KAZ.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-#KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-
-#table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-KAZ.ploty_IFD_raw <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_raw
-
-KAZ.ploty_IFD_PER <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, KAZ")+geom_boxplot(aes(x=sex))
-KAZ.ploty_IFD_PER
-
-KAZ.IFDnorm_scatter.plot <- ggplot(KAZ.Bivalents.DF[KAZ.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("KAZ \n Normalized IFD and SC Length")+ylim(c(0,1.1))+scale_color_manual(values=c('#56B4E9','#E69F00'))
-
-KAZ.IFDnorm_scatter.plot
-
-#run a t.test seperate out by the chrm classes
-KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
-
-```
-
-For the KAZ, pattern the distinction between the male and female pattern is less distinct. There are fewer instances of females with v close IFD distances.
-
-```{r SKIVE.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-#table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-
-#remove the NAs
-SKIVE.ploty_IFD_raw <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_raw
-
-SKIVE.ploty_IFD_PER <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, SKIVE")+geom_boxplot(aes(x=sex))
-SKIVE.ploty_IFD_PER
-
-SKIVE.IFDnorm_scatter.plot <- ggplot(SKIVE.Bivalents.DF[SKIVE.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+ geom_jitter()+ggtitle("SKIVE \n Normalized IFD and SC Length")+ylim(c(0,1.1))
-
-SKIVE.IFDnorm_scatter.plot
-
-SKIVE.table <-  ddply(SKIVE.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])
-
-```
-
-In the Skive data, it could be the case that the v. short IFD measures in females are rare / another class of observations.
-
-```{r MSM.IFD, echo=FALSE, warning=FALSE}
-#make the dataset
-MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-#table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-MSM.ploty_IFD_raw <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+
-                          geom_jitter()+ggtitle("Raw IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_raw
-
-MSM.ploty_IFD_PER <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+
-                          geom_jitter()+ggtitle("Percent IFD, MSM")+geom_boxplot(aes(x=sex))
-MSM.ploty_IFD_PER
-
-#another plot
-MSM.IFD_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")
-
-MSM.IFDnorm_another.plot <- ggplot(MSM.Bivalents.DF[MSM.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = chromosomeLength, color = sex))+                          geom_jitter()+ggtitle("MSM")+ylim(c(0,1.1))
-
-MSM.IFDnorm_another.plot
-
-#this plot helps illustrate sex-specific differences for normalized IFDs -- both sexes have a large class / proportion of bivalents with similar IFD_PER values -- but the mean in females is lower due to them having MORE low IFD_PER values
-
-#make a table of the means and n ect
-MSM.table  <-  ddply(MSM.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#can I plot these stats
-
-#run a t.test seperate out by the chrm classes
-MSM.ttest  <- t.test(MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "female"], MSM.Bivalents.DF$chromosomeLength[MSM.Bivalents.DF$sex == "male"])
-
-```
-
-The MSM pattern has a short range and longer range of nrm.IFD in males and females respectively.
-
-```{r MOLF.IFD, echo=FALSE}
-
-#SHUT DOWN THIS BLOCK UNTIL MOLF FEMALE DATA
-
-#make the dataset
-#MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-
-MOLF.Bivalents.DF <- Curated_BivData %>%  filter(strain == "MOLF")
-
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
-
-#plot the differences
-#MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
-
-#remove the NAs
-#molf.ploty_IFD_raw <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_ABS, x = sex))+geom_jitter()+ggtitle("Raw IFD, MOLF")+geom_boxplot(aes(x=sex))
-
-#molf.ploty_IFD_PER <- ggplot(MOLF.Bivalents.DF[MOLF.Bivalents.DF$hand.foci.count == 2,],  aes(y=IFD1_PER, x = sex))+  geom_jitter()+ggtitle("Percent IFD, MOLF")+geom_boxplot(aes(x=sex))
-
-#make a table of the means and n ect
-MOLF.table  <-  ddply(MOLF.Bivalents.DF, c("sex"), summarise,
-
-                  total.biv = length(Obj.ID),
-
-                  mean.IFD_2CO = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                  mean.IFD.PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                  n_2CO =  sum(hand.foci.count == 2, na.rm = TRUE) #this is the right syntax
-)
-
-#run a t.test seperate out by the chrm classes
-#MOLF.ttest  <- t.test(MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "female"], #MOLF.Bivalents.DF$chromosomeLength[MOLF.Bivalents.DF$sex == "male"])
-
-```
-
-The strains which show a clean "30% threshold" for normalized IFD in males are: PWD, SKIVE, and MSM (which are the 2 high Rec and a intermediate strain).
-The other strains which have more overlap between males and females are the Dom strains and KAZ.
-
-#### IFD 3CO bivalents
-
-Run comparisons for 3CO bivalents.
-
-```{r CO3.comparisons, echo=FALSE}
-
-total.IFDs <- Curated_BivData %>% filter(hand.foci.count > 1)
-
-#take a look at the 3CO things
-
-```
-
-<!--
- Heterochiasmy Prediction
-
-> Do the IFD measures lack significant sex effect? (as predicted)
-
-There are sex effects for the lme() models for both the ABS and PER IFD measures. There is slight evidence for significant strain effect.
-
-> Are there signiticant strain effects?
-
-The random strain effect was tested for raw and normalized IFD, neither are significant.
+# Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
+
+<!-- keep blank for now 
+-basic stats for the new dataset
+
 -->
-
-## Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
-
-The general predictions across the males and subspecies based on th above MLH1 results.
-
-For positive correlation traits/metrics
-
-i) in DOM strains, low to no difference across strains
-
-ii) in Musc, PWD > SKIVE > KAZ, CZECH all the others
-
-ii) in Mol, MSM > MOLF
-
-A) **SC lengths** will be longer for high Rec strains.
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-C.1) not enough is known about variation within species for the  **1CO normalized positions**. Null prediction, no difference in the 'telomeric pattern'.
-
-The mouse averages for the other position metrics will be highly influenced by proportions of the 1CO and 2CO bivalents.
-When class of chromosome and SC length is account for, there won't be a difference, however, not enough is known about these patterns.
 
@@
 However there are a few exceptions -- However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel
-
---
 
@@
 
-<!--
-For later analysis.  Not sure what these are good for.
-
-```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
-
-# check if the main mixed model results hold up when all cells are pooled
-
-#pooled.cells
-DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
-
-#remove PERC, MOLF, CZECH,
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
-
-#sex only has 2 levels and this breaks...
-model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
-
-#pull out values / other things
-summary(model.pooled)# indeces for this object: [1]-structure
-ranef(model.pooled)
-
-model.pooled$coefficients
-
-#pooled cells,  DF.HetC.MixedModel.pooled
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
-
-#consider changing to ordered = FALSE, for the DFs used for MM
-DF.HetC.MixedModel.pooled$sex <- factor(DF.HetC.MixedModel.pooled$sex, ordered = TRUE, levels =c("male", "female") )
-
-DF.HetC.MixedModel.pooled$strain<- factor(DF.HetC.MixedModel.pooled$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PWD", "MSM", "SKIVE", "KAZ") )
-
-DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
-
-```
-
-- The MSM strain effect is 0, but the sex*strain for molossinus is 7 (this seems very high).
-This means there is 0 variance added to mean_CO to mean due to being MSM.
-Does this mean the model is very good at predicting MSM mean_CO.
-The fixed strain - sex effects are for musc are large.
-This will likely change with MOLF (male molf are low, so smaller fixed effect and )
-
-- I think adding Mol, moves the highest strain into it's own category, so the coefficient is large.
-In the first model MSM has largest random effect, since it's so far from WSB female.
-
--adding MOLF also doesn't have significant factors
-
-changed all to unordered factors.
-all the variables are coded as ordered factors.
-When the variables are kept factors, and unordered the names appear! yay!
-
-Character assigned variables.  The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independent of subspecies or strain.
--->
-
-<!--
-
-#from chrm proportion black
-
-#oopp2 <- lm( mean.MLH1.foci2 ~ per0CO+ per1CO +per2CO +per3CO, data = proportion_stats_category)
-#summary(oopp2)
-#none of the classes are significant
-
-#mean is the new calculated-- mean 0CO is significant and intercept
-#oopp <-  lm( mean.MLH1.foci ~ per0CO+ per1CO +per2CO +per3CO, data = #proportion_stats_category)
-#summary(oopp)
-#not sure if these are inter-correlated
-#I don't know why 0COs and the intercept are significant..
-#0CO and 1CO are significant when only chrm classes are included (and intercept)
-#no longer significant when used
-#lm regression indicates that the proportions aren't significant contributers to meanMLH1 -- but this could be due to error
-#after added more MSM data, tot_biv is not significant
-#0CO is now more significant, and 2CO is sightly significant
-#2CO proportion is the most significant, 1CO is second most significat, 3CO doesn't have enough power - total bivalents is also significant, there might be a bias
-#MSM needs more bialent observations
-#use melt to put all propotions 
-#display / visualize / compare proportions
-#the table needs to be melted, because it's split by category
-
-```{r Q2.IFD.log.reg, echo=FALSE, eval=FALSE}
-
-# LOGISTIC REGRESSION FOR IFD
-
-log.reg_IFD_ABS_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-summary(log.reg_IFD_ABS_mouse.av)# 0.0366 *
-
-log.reg_IFD_PER_mouse.av <- glm(Rec.group ~ mean_IFD.2CO_PER, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-summary(log.reg_IFD_PER_mouse.av)#NS
-
-### Single Bivalent levels
-
-log.reg_IFD_ABS_single.biv <- glm(Rec.group ~ IFD1_ABS, 
-              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_IFD_ABS_single.biv)
-
-log.reg_IFD_PER_single.biv <- glm(Rec.group ~ IFD1_PER, 
-              data=Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_IFD_PER_single.biv)
-```
-
-```{r trackcurBivData, echo=FALSE, eval=FALSE}
-#REMOVE THIS BLOCK??
-
-#list / tally of current curated Bivdata
-
-Cur_biv = read.csv("~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", header = TRUE)
-
-colnames(Cur_biv) <- c('mouse', 'blank','0','1')
-
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages[ -c(9:12, 14, 18, 22:24) ]
-
-AP_mouse_table_w.Ages_HQ <- add_sex(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_strain(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_category(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_subsp(AP_mouse_table_w.Ages_HQ)
-AP_mouse_table_w.Ages_HQ <- add_species(AP_mouse_table_w.Ages_HQ)
-
-#only M.musculus
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$species == "M.musculus",]
-#REMOVE EXTRA STRAINS
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$subsp != "Cast",]
-#remove musc strains
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "AST",]
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "CZECH",]
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$strain != "PERC",]
-#156 mice (from 193)
-
-#some mice are missing 13oct14
-
-#match the MM dataset
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$Ncells > 9,]
-AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[AP_mouse_table_w.Ages_HQ$age.weeks < 26,]
-#126 mice
-
-#AP_mouse_table_w.Ages_HQ <- AP_mouse_table_w.Ages_HQ[!(is.na(AP_mouse_table_w.Ages_HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
-
-#mice missing from curated list
-mice.within.BivData <- AP_mouse_table_w.Ages_HQ[ AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#54 mice in (14 mice have enteries but 0 curated biv)
-
-#(14 mice need at least some curation
-
-cur.data.mice <- Cur_biv[ Cur_biv$mouse %in% AP_mouse_table_w.Ages_HQ$mouse, ]
-
-mice.missing.bivData <- AP_mouse_table_w.Ages_HQ[ ! AP_mouse_table_w.Ages_HQ$mouse %in% Cur_biv$mouse, ]#72 (includes mice not in MM)
-#72 mice that I need to upload/run the algorithm on their images
-
-#mice missing SC pass == '1'
-colnames(Cur_biv) <- c("mouse","blank","SC.fail", "SC.pass")
-NeedsCur <- Cur_biv[Cur_biv$SC.fail == 0,]
-
-NeedsCur <- add_strain(NeedsCur)
-write.table(NeedsCur, "~./MLH1repo/data/needs_cur.csv", sep=",",
-            row.names = TRUE)
-
-#44% of mice are missing curation
-
-#write this table
-#write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_13.1.20.csv", sep=",",
-#            row.names = TRUE)
-
-mice.missing.bivData <- mice.missing.bivData[!(is.na(mice.missing.bivData$mouse) | mice.missing.bivData$mouse==""), ]
-write.table(mice.missing.bivData, "~./MLH1repo/data/MissingMice_13.1.20.csv", sep=",",
-            row.names = TRUE)
-
-```
-
-```{r chrm.pro.Curated.BivData, echo=FALSE, eval=FALSE}
-
-#this chunk migth need to be deleted // I think everything is already done within the BivData
-#this is a second mode of
-
-#bivalent data might be too big..
-#load(file="~./MLH1repo/data/CleanBivData.RData")
-#Curates_BivData -- hand.foci.count!
-
-#For each mouse I need to count up
-#the number of bivalent classes are not counted up
-
-chrm.propotion.category <- ddply(.data=Curated_BivData, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(fileName)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = length(Obj.ID),
-
-                 n0CO = length(which(hand.foci.count == 0)),
-                  n1CO = length(which(hand.foci.count == 1)),
-                  n2CO = length(which(hand.foci.count == 2)),
-                  n3CO = length(which(hand.foci.count == 3)),
-                  #n4CO = sum(n4CO),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv )
-                # per4CO= sum(n4CO)/ tot_biv  )
-
-Chrm.proportions_Biv.Data <- melt(chrm.propotion.category, c(
-  "per0CO", "per1CO", "per2CO", "per3CO"), 
-  id=c("category", "tot_biv") )
-colnames(Chrm.proportions_Biv.Data) <- c("category","total_biv","Chrm.Type","value")
-
-# I think I need to make 2 melt tables -- for count and proportions
-Chrm.numbers_Biv.Data <- melt(chrm.propotion.category, c("n0CO", "n1CO", "n2CO", "n3CO" ), 
-  id=c("category", "tot_biv") )
-colnames(Chrm.numbers_Biv.Data) <- c("category","total_biv", "Chrm.Type", "biv_count")
-
-Chrm.prop_Biv.Data <- cbind(Chrm.proportions_Biv.Data, Chrm.numbers_Biv.Data[,4])
-colnames(Chrm.prop_Biv.Data) <- c("category","total_biv", "Chrm.Type","value", "biv_count")
-
-Chrm.prop_Biv.Data <-Chrm.prop_Biv.Data[(!is.na(Chrm.prop_Biv.Data$value)),]
-
-Chrm.prop_Biv.Data$sex <- ifelse(grepl("female", Chrm.prop_Biv.Data$category), "female", ifelse(grepl("male", Chrm.prop_Biv.Data$category), "male", "" ))
-
-Chrm.prop_Biv.Data$strain <- ifelse(grepl("WSB", Chrm.prop_Biv.Data$category), "WSB", ifelse(grepl("G", Chrm.prop_Biv.Data$category), "G",
-ifelse(grepl("LEW", Chrm.prop_Biv.Data$category), "LEW",                                                                                                           
-      ifelse(grepl("KAZ", Chrm.prop_Biv.Data$category), "KAZ",           
-        ifelse(grepl("CZECH", Chrm.prop_Biv.Data$category), "CZECH", 
-        ifelse(grepl("SKIVE", Chrm.prop_Biv.Data$category), "SKIVE",
-        ifelse(grepl("PWD", Chrm.prop_Biv.Data$category), "PWD",  
-        ifelse(grepl("MSM", Chrm.prop_Biv.Data$category), "MSM", "" ))))))))
-
-Chrm.prop_Biv.Data$strain<- factor(Chrm.prop_Biv.Data$strain,levels =c("WSB","G","LEW",
-                       "KAZ","CZECH", "SKIVE","PWD","MSM"), order=T )
-#no mol yet
-
-Chrm.prop_Biv.Data <- with(Chrm.prop_Biv.Data, Chrm.prop_Biv.Data[order(strain),])
-
-#need to make list of positions
-
-chrm_prop.BivData_plot <- ggplot(data = Chrm.prop_Biv.Data, aes(y = value, x = strain, fill=as.factor(Chrm.Type)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData")+ facet_wrap(~sex, scales = "free", ncol = 1 )+ylim(c(0,1)) + labs(x="", y='Proportions') +   geom_text(aes(label = biv_count), position = position_dodge(0.99))
-
-chrm_prop.BivData_plot
-
-```
-
--->
-
 # References

### 2020-02-18 junk_text.R

@@ -1,0 +1,673 @@
+#
+# junk text -- dumping the junk text
+
+
+
+
+
+
+
+<!--
+  For later analysis.  Not sure what these are good for.
+
+```{r Mixed.models.pooled.cells, echo=FALSE, eval=FALSE, include=FALSE}
+
+# check if the main mixed model results hold up when all cells are pooled
+
+#pooled.cells
+DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
+#remove musc strains
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]
+
+#remove PERC, MOLF, CZECH,
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
+
+
+#sex only has 2 levels and this breaks...
+model.pooled <- lme(nMLH1.foci ~ subsp * sex, data=DF.HetC.MixedModel.pooled, random=list(strain=pdDiag(~sex) ) )
+
+#pull out values / other things
+summary(model.pooled)# indeces for this object: [1]-structure
+ranef(model.pooled)
+
+model.pooled$coefficients
+
+#pooled cells,  DF.HetC.MixedModel.pooled
+DF.HetC.MixedModel.pooled <- droplevels(DF.HetC.MixedModel.pooled)
+DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[!(is.na(DF.HetC.MixedModel.pooled$mouse) | DF.HetC.MixedModel.pooled$mouse==""), ]
+
+#consider changing to ordered = FALSE, for the DFs used for MM
[…]

### 2020-02-18 doc/Results.rmd
small changes to ordering and geting numbers to appear

@@
 
-In order to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` _M.
-m.
-musculus_ strains, MLH1 foci were quantified from a total of `r sum( (main.table  %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table  %>% filter(sex == "female"))$Ncells)` oocytes.
-
-(mouse means were calculated and statisics in Table 2)
-
-Figure 1(A) shows the --
+In order to estimate the genome wide recombination rates across our panel of three subspecies from the house mouse species complex, _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. 
+`r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains were used with reflect samples of natural genetic variation across the terretories of these subspecies.
+
+MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
+The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.
+
+1. -female strain means are v close  -male strain means are more varied!!
+-Comparisons to references / literatures
+-house mouse close to minimum of 1  obligate CO per chrm
+
+2. The largest female difference is G, 1.07X  the largest male difference is PWD and MSM which are 1.3 and 1.2 respectively.
+
+<!-- female -->
+In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.
+
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+#table( droplevels(female_mouse.avs_for.analysis$strain) )
+#there are cast strains too
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Female mouse av.
+MLH1 count per cell ")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))
+
+#female.specific.plot
+
+musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+#Mol
+Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+#Dom version 
+Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+```
+
+## female section
+
+```{r female.strain.table, echo=FALSE}
+
+#female strain table
+#female_mouse.avs_for.analysis
+F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
+
+                        Nmice = length(unique(mouse)),
+
+                        total.cells  = sum(Ncells, na.rm = TRUE),
+
+                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#F.strain.table
+
+# I don't know why most of the table is blank
+```
+
+```{r female.dif, echo=FALSE}
+#how much greater are the high strains?
+G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
+mean.G.female <- mean(G_female$mean_co)
+
+LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
+mean.LEW.female <- mean(LEW_female$mean_co)
+
+#how much greater is MSM female
+MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
+mean.MSM.female <- mean(MSM_female$mean_co)
+
+sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
+mean.all.other <- mean(sans_high_female$mean_co)
+
+G.female.diff <- (mean.G.female / mean.all.other)
+LEW.female.diff <- (mean.LEW.female / mean.all.other)
+MSM.female.diff <- (mean.MSM.female / mean.all.other)
+
+```
+
+The overall pattern for the female mean MLH1 counts per cell is low variation. ( in the most different strains
+In females specifically _domesticus^G^_ is `r G.female.diff` higher the the other means. 
+while _ domesticu^LEW^_  is `r LEW.female.diff` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.
+
+In the _domesticus_ subspecies the ranges of mouse means are 
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to
+
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,
+
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to
+
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )`
+
+in _domesticus^G^_
+
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _musculus^PWD^_,
+
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to
+
+`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_
+
+### male section
+
+While the male pattern shows more variation across the strain means.
+
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
+#try customize the plot size
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#table( droplevels(just.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+#fomo.male_mouse.av
+
+male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+male.plot
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+#Dom version 
+Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+```
+
+```{r male.strain.table, echo=FALSE}
+M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+
+                        totNcells  = sum(Ncells, na.rm=TRUE),
+
+                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
+
+                        #Standard error (mice . cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#M.strain.table
+#same issue as the female table .. ugh
+```
+
+```{r male.diff, echo=FALSE}
+
+#how much greater are the High Rec gwRRs
+just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
+                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
+                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
+
+#how much bigger are the high rec male means?
+sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+#mean(sans_LewRec_male$mean_co)
+
+PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+#mean(SKIVE_male$mean_co) /  
+```
+
+- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
+
+- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
+
+The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r (mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co) )`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
+Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
+
+- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
 
 --notes on error / 
-
 -- connections to previous measures in the literature
 
+The same model framework was applied to the male specific data.
+The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
 <!--
@@
 
-# Female Specific Analysis
-
-In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-#table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Female mouse av.
-MLH1 count per cell ")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
-  theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))
-
-female.specific.plot
-
-musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-#Mol
-Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
-
-Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-#Dom version 
-Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
-
-Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-```
-
-```{r female.strain.table, echo=FALSE}
-
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-
-                        Nmice = length(unique(mouse)),
-
-                        total.cells  = sum(Ncells, na.rm = TRUE),
-
-                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-F.strain.table
-
-# I don't know why most of the table is blank
-```
-
-For some strains, we were only able to quantified or include one mouse and the Cast observation comes from 1 1 cell.
-To analyze the female specific data set,
-
-In the _domesticus_  subspecies the ranges of mouse means are 
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _domesticus^G^_
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _ musculus^PWD^_,`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_
+MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
 
@@
 
-#how much greater are the high strains?
-G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co)
-
-LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
-mean.LEW.female <- mean(LEW_female$mean_co)
-
-#how much greater is MSM female
-MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSM.female <- mean(MSM_female$mean_co)
-
-sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
-mean.all.other <- mean(sans_high_female$mean_co)
 ```
@@
 MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.
-
-_domesticus^G^_ is `r mean.G.female / mean.all.other` higher the the other means.
-_ domesticu^LEW^_  is `r mean.LEW.female / mean.all.other` higher  and _musculus^MSM^_ is `r mean.MSM.female / mean.all.other`.
-These three will be designated as 'moderate high rec' strains for later sex specific analysis.
 
@@
 
-# Male Specific Analysis
-
-The male specific patterns were also investigated. (what's a better motivation for this section)
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-#fomo.male_mouse.av
-
-male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-male.plot
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-#Dom version 
-Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-```
-
-```{r male.strain.table, echo=FALSE}
-
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-M.strain.table
-#same issue as the female table .. ugh
-```
-
-The same model framework was applied to the male specific data.
-The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
-
- _domesticus^LEW^_
-
-- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
-
-- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
-
-- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
+MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
 
@@
 
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /  
 ```
@@
 ```
-
-The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
-Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
 

### 2020-02-19 doc/Debug_Rmd.Rmd
messy work from home

--- 2020-02-18 doc/Results.rmd
@@
 title: "MLH1 Results Outline"
-date: "`r format(Sys.time(), '%d %B %Y')`"
 output:
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-  pdf_document: default
-  word_document:
-    toc: yes
-    toc_depth: '5'
-bibliography: refs_Main_Report.bib
+  html_notebook: default
+  html_document: default
 ---
@@
     )
-#F.strain.table
+F.strain.table
 
@@
 ```{r female.dif, echo=FALSE}
+
 #how much greater are the high strains?
@@
 
+##showing values
+
+'tostring()  didn't work
+
 The overall pattern for the female mean MLH1 counts per cell is low variation. ( in the most different strains
-In females specifically _domesticus^G^_ is `r G.female.diff` higher the the other means. 
+In females specifically _domesticus^G^_ is `r toString(G.female.diff)` higher the the other means. 
 while _ domesticu^LEW^_  is `r LEW.female.diff` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.
@@
 `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )`
-
-in _domesticus^G^_
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _musculus^PWD^_,
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_
-
-### male section
-
-While the male pattern shows more variation across the strain means.
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-#fomo.male_mouse.av
-
-male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-male.plot
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-#Dom version 
-Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-```
-
-```{r male.strain.table, echo=FALSE}
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#M.strain.table
-#same issue as the female table .. ugh
-```
-
-```{r male.diff, echo=FALSE}
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /  
-```
-
-- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
-
-- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
-
-The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r (mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co) )`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
-Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
-
-- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
-
---notes on error / 
--- connections to previous measures in the literature
-
-The same model framework was applied to the male specific data.
-The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
-<!--
-The strain means for females were pretty similar.
-For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively.
-In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively.
-In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively.
--->
-
-While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively.
-In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively.
-In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.
-
-<!--(the figure is made from pooling MLH1 cells)
-the male 
-- the ranges of mouse means are displayed in Sup table X
-- How the means were calculated
--The male gwRR is adjusted for the PAR
--The means for each strains are report in Table 1
--Normality was assessed for the MLH1 counts (include in Supplement histograms)
--->
-
-# Analysis of Evolutionary Patterns
-
-In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
-(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
-
-<!--
-$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
--->
-<!--
-This is the break down of mice used within the Mixed model
-
-`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
-
-add number of cells
-`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
- $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-Setting up the data set for using the mixed model (base data set, HQ data set)
-
-- only Mus musculus strains with sex matched observations  
-- Male gwRR have been adjusted for the PAR
-- Q5 excluded
-- All the effects are un-ordered factors  
-- male ages within 5 - 20 weeks
-- Molossinus is coded as a subsp level
-
-subspecies levels: 3, Musc, Dom and Molossinus
-strain levels: Dom: 3, Musc: 3  Mol: 2
-sex: 2 female male  
--->
-
-```{r MLH1co.MM, echo=FALSE, digits = 3}
-#THIS IS THE NEW MIXED MODEL
-#should be able to use the loaded 
-#DF.HetC.MixedModel.HQ
-
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
-
-MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
-sex.results =  MLH1.MM_sex, 
-               interaction.pval = MLH1.MM_interaction,  rand.pval = 
-MLH1.MM_strain)
-```
-
-Our models were designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
-
-All of the effects from the mixed model significant, (LRT; p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect).
-
-<!--
- $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-and 
-$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$  -->
-
-```{r MLH1.glmM2, echo=FALSE,  digits = 3}
-#strain = fixed
-glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-
-summ <- summary(glm.strain.fixed.full)
-#table of just coeff anf pvalue
-#round(summ$coefficients[,c(1,4)],5 )
-
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-```
-
-We followed up the mixed model effects with general linear models.
-The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions.
-(*what are the points and differences?
-)*
-
-```{r MLH1.glm.M3, echo=FALSE, digits = 3}
-
-strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-summ.strain.fixed <- summary(strain.fixed)
-
-#round(summ.strain.fixed$coefficients[,c(1,4)],5)
-
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-```
-
-In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
-These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.
-
-Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
-These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
-
-## Within Mouse Variance in CO Count per Cell
-
-In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-<!--
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$-->
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-<!--
-These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
-
-var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-
-cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
--->
-
-For both the mixed model and linear model, sex was the most significant effect while the subspecies, interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively.
-
-*notes on caveats for calculating variance -- lower power
-
-Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
-background noise).
-For this reasons, we replicated the model analysis from data sets in the subset of cells with higher quality scores compared to the full data set.
-(these analysis/ models replicated the results of the full data set (sex being the most significant
-(LTR; p = `r LmerMLH1_M1_Q12_VAR_results$sex.results`, and glm; p = `r  MLH1var_M3_Q12_sum$coefficients[2,4]` )
-
-**female strongest effect, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.
-
-<!--
-motivation / background: maybe checkpoint / regulation
-
-limitations: observations at the mouse level, decreased power to estiamate the variance --
-setup
--quality / technical error
-- variance and cv
-The mixed models have most significant variables, 
-M3.glm for variance was less significant (0.044181 ), 
-the Q12 models in general were more significant
-cV's were more significant than var
-
-var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-
-cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
-
-$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
-<!--
-Unequal variance across strains might indicate that the within animal variance for MLH1 foci per cell may have evolved, yet --our low power and number of mice is not equal, -- number of mice across
-
-Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
-<!--
-
-even tho it's an important trait for XX reasons, -- there are limitations for our analysis
-
-Unlike the mean MLH1 counts -- the mean per mouse is calculated across a sample of cells (which serve as replicates)
-
-the within mouse variance for MLH1 count per cell, on the other hand -- is calculated at the mouse level, var()
-
--- the replicates are at the level of mouse, unlike MLH1 counts per cell
-
-Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
-
-RW's main qtl paper used the within animal variance, 
-"We mapped QTL for
-MLH1 count variance within individuals, applying an approach
-to deconvolve the mean-variance relationship (see
-Materials and Methods)."
-
-The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
-QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
-
-@article{wang2017genetics,
-  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
-  author={Wang, Richard J and Payseur, Bret A},
-  journal={Genetics},
-  volume={206},
-  number={4},
-  pages={1841--1852},
-  year={2017},
-  publisher={Genetics Soc America}
-}
-
-references for phenotypic variance connected to robustness to environment, increase in developmental consistency
-Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
-
-phenotypic variance leading to adaptation - 
-(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
-
-Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
-meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
--->
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
-
-#Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
-```
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-#variance plots
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-#Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.male.VAR.plot
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.male.VAR.plot
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.male.VAR.plot
-
-```
-
-MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
-
-Two glm models were run to test the effects of strain and subspecies on mean MLH1 count per cell per mouse. The coefficients for the following model are below.
-
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
-
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-lm.female.strain_sum <- summary(lm.female.strain)
-
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-
-```
-
-For mean CO count, _domesticus^G^_  strain effects in both models (p = )
-_musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).
-
-Above are the coefficents for the two glm's of the female specific mouse MLH1 averages, which include subspecies and strain as fixed effects.
-G has the most consistant significant strain effects in both models.
-MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.
-
-Within mouse variance was assess to follow up on the results from the full dataset.
-
-```{r FEMALE.VAR.MODELS, echo=FALSE}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-#round(
-#lm.m1.Q12_female_sum$coefficients[,c(1,4)]
-#LEW not sig
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
-
-The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (p= ).
-
-MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
-
-Two models were fitted to the male specific data to test the effects of strain and subspecies.
-
-```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-summry.male.strain.effects <- summary(glm.male.strain.subsp)
-round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#1. determine which strain effects are sig
-
-#M2
-lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
-lm.male.strain_sum<- summary(lm.male.strain)
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-
-```
-
-the coefficients for glms of the CO means are listed above.
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-```
-
-# Variation in DSB number per cell
-
-```{r DMC1.setup, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "L"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "Z"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
-#            row.names = FALSE)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
-```
-
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
-#think of better color scheme
-#make the jitter col thinner or make the stacked dots.
-#remember that box-plots add outlier dots.
-#change y labels
-
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
-#change order
-
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.boxplot
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-  ggsignif::geom_signif(
-    data=anno_df_stage, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
-```
-
-```{r plot.t.test, echo=FALSE, include=FALSE}
-
-  #this will add neat bars, but still need to figure out what groups ect it's actually using
-#THE stat_compare function adds a arm thing and pvalue for 
-#  stat_compare_means(comparisons = my_comparisons)
-
-#status - legend, expression y-axis, facet, facet_wrap SYMBOL
-#legend - stage. y=total count, FACet strain
-
-#add column to distinguish high vs low rec strains
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
-DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
-
-# annotation table with adjusted pvals and y-position of the labels
-
-anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=stage), position=position_jitterdodge()) + 
-  facet_wrap(~stage, nrow=1) + 
-  ggsignif::geom_signif(
-    data=anno_df, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#example for ggplot+ test from here
-#https://github.com/kassambara/ggpubr/issues/65
-
-#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
-
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-
-#unorder strain
-DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
-
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
-z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
-```
-
-```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-#attach this mini DF to the DMC1 data
-
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-
-#DMC1.mean.table
-DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
-
-colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
-                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
-
-DMC1.mean.table.99
-
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-as.data.frame(DMC1table.main)
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
-In an attempt to locatize the above male specific crossoever number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase meiocytes.
-
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of (X for each strain) juvenile mice (12 to 18 days).
-
-A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _ musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
-One mouse was sampled for each strain.
-
-As can be seen in the figure, the high recombining groups have significantly more DMC1 for for leptotene (t.test; `r ttest.HighLow.L.pval` , one-way-anova; `r round(l.anova.pval, 5)`), but not the zygotene cells (t.test; `r ttest.HighLow.Z.pval`, one-way-anova; `r z.anova.pval`).
-
-The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio).
-Notably there is not much variation and do not show a significant difference between the high and the low recombining groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
-These results indicate that the best predictor for CO number may be established before DSB formation (cite Baier et al 2014).
-
-<!--
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M.
-m.
-domesticus^WSB^_ , _M.
-m.
-domesticus^G^_, _M.
-m.
-molossinus^MSM^_, and  _M.
-m.
-musculus^PWD^_ respecitively.
--->
-
-```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
-
-#Subsampling code -- don't report
-
-#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
-Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
-Permutations were done for both cell stages.
-The actual p value is in red.
-
-#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
-
-#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
-# the 2 groups
-
-#make the 2 pools
-High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-
-Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-
-#permute drawing / subsampling and performing the t-test
-#put the p.value results into a DF!
-
-#permut.SC.cv.diff  <- as.data.frame(
-#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
-
-#permuting samples -- and re-licating t-tests
-
-#nrep <- 1000
-
-#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
-bunchOpvals.L <- replicate(1000, 
-    # t.test
-    t.test( sample(High.L.pool, 25, replace = FALSE),
-            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
-          )
-bunchOpvals.L <- as.data.frame(bunchOpvals.L)
-bunchOpvals.L$stage <- "L"
-
-bunchOpvals.Z <- replicate(1000, 
-    # t.test
-    t.test( sample(High.Z.pool, 25, replace = FALSE),
-            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
-          )
-
-bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
-bunchOpvals.Z$stage <- "Z"
-
-#could merge the lists of p-values
-replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
-#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
-
-#ASSUMPTIONS; normality, equal variance
-
-#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-#plot.pvals.L
-#plot.pvals.Z
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
-```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#these are histogram plots for testing Normality shape -- move to supplemenatry section
-
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-DMC1.histogram
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-
-#The counts look close to normal.
-#Potentially figure 2. Using a function to #display the p-vales for mean comparison
-
-```
-
-<!--
-interpertation notes, 
-the number of DMC1 across stages (L and Z stages) are driven by different equations
-
-Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
-
-Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
-
-*the repair decision can happen at both stages and remove DMC1.
-formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
-
-back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
-
-expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
-OR
-
-variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
--Baier et al 2014 supports this model (Chmr axis / DNA loop model)
--->
-
-<!--
-# Rapid male specific evolution
-
-list out the points which support rapid male specific evolution for MLH1 rates.
-
-number above bars are the raw counts for bivalent classes.
-
-Potential figure 2 (eventually). Try to make the figure taller.
--- add more data to make sample sizes more equal.
--- add space to indicate where female data is missing
-
-SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
-What's up with Lew females? there's lots of 2COs, also WSB female.
-
-This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
-If these data are random samples of bivalents, the X
-
-ToDo:
-- within strain across sex,
-- males, across high and low groups
-- look at mouse level proportions
--->
-
-# Chromosome Class Proportions
-
-In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes with different numbers of COs.
-The two plots show the chromosome class proportions from hand measured and the curated BivData.
-
-```{r chrm_prop_data, echo=FALSE, warning=FALSE}
-#shut off eval for now because of wd / path error
-#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#try adding the BivData version of the proportions
-
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
-
-                 n0CO = sum(n0CO, na.rm = TRUE),
-                  n1CO = sum(n1CO,na.rm = TRUE),
-                  n2CO = sum(n2CO,na.rm = TRUE),
-                  n3CO = sum(n3CO,na.rm = TRUE),
-                  n4CO = sum(n4CO,na.rm = TRUE),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#this is the table that needs melting
-#but is also easy to redo by hand
-
-#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
-```
-
-```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
-
-#add subspecies to these values
-melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
-
-#add the subsps
-
-melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
-
-melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
-
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
-
-#need a second DF / the first df
-#+ annotate() text 
-#annotate()text
-
-#proportion_stats_category,  tot_biv, ncells
-
-#add number of cells and bivalents in the corner
-
-chrm.class.barplot
-```
-
-```{r test.Chrm.Prop, echo=FALSE, eval=FALSE}
-
-#prop.test(x, y)  x number of successes, y total  
-#strain sums of n
-#look up continuity correction
-
-#mouse.level.BivData_table <-- has more male strains
-#Mouse.Table_BivData_4MM <- only sex matched strains
-
-#perform test across strain 
-#removed all lines with no hand.foci.count entry
-Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
-
-                  nbiv = sum(tot.biv ),              
-                 n0CO = sum(n0CO ),
-                  n1CO = sum(n1CO ),
-                 n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO ),
-                 n4CO = sum(n4CO )
-                          )
-
-Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
-
-for(g in 1:length(Chrm.prop_strain_table$nbiv)){
-
-  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
-                                     Chrm.prop_strain_table$n1CO[g],
-                                     Chrm.prop_strain_table$n2CO[g],
-                                     Chrm.prop_strain_table$n3CO[g] ))  
-}
-
-Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
-#there are still some wacky entries -- but much fewer than before
-
-chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
-
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
-#the pvalues for 0 are not sig dif
-#the p values for 1 and 2CO are sig diff
-
-#seperate by subsp
-chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
-chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
-chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
-
-Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
-Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
-Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
-Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
-#WSB has fewer 2COs than other 3 strains
-
-Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
-Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
-Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
-#KAZ has slightly more 1CO than the other two
-
-#high.female_G.LEW.
-chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
-
-high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
-high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
-high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
-
-#Low.female.G.LEW.
-chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
-
-low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
-low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
-low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
-
-chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
-#the pvals for these 3 class are sig dif
-
-#seperate by subsp
-chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
-chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
-chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
-
-Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
-Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
-Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
-#non of the Dom proportions are different
-
-Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
-Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
-Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
-#all v different
-
-Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
-Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
-Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
-
-#high vs low
-chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
-                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
-                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
-
-chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
-chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
-
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
-
-Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
-Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
-Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
-#the low strains do vary slightly in their 1CO and 2CO proportions
-
-Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
-Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
-Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
-#all very sig p values
-
-High2_male.0CO.prop.test <- prop.test(chrm.prop.male.high2$n0CO, chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO, chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO, chrm.prop.male.high2$nbiv )
-#these proportions are v different
-
-High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
-High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
-High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
-#MSM and PWD only slightly sig dif 2CO
-
-#Need to run the chrm proportions thru melt code.
-
-chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
-
-#chrm.class.DF$biv.sum.check
-```
-
-These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes. Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs.
-
-Most all the the p values for the proportion tests are significant, indicating there are slight but significant shifts across the classes of chromosomes.
-However the most striking male pattern is the propotion of 2COs
-
-```{r Chrm.props.plot2, echo=FALSE}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
-
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
-
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
-
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-#no filename for add mouse, strain subsp
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#can't apply function use this
-chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
-             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
-                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
-                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
-
-          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
-              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
-
-                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
-                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
-            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
-                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
-                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
-                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
-                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
-                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
-
-             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
-               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
-
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",
-                                                                 "Cast",
-                                                                 "Musc",
-                                                                 "Mol",
-                                                                 "Spretus","Spic",
-                                                                 "other") )
-#ggplot
-
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
-#chrm.class.barplot
-
-```
-
-(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class.
-The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.
-
-- High female strains, G and Lew have significantly more 2CO bivalents.
-
-- The overall male pattern that's most striking is the gradient of 2CO proportions
-
-**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
-
-# Single Bivalent Level Results
-
-In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining single chromosome patterns.
-
-From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(new.Curate.FULL.org$Obj.ID)` single bivalent measures were left.
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#sort by category,
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-
-data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-Table X illustrates the break down of single bivalent observations by category
-
-Error compared to human measures and other details have been discribed in (@Peterson_2019).
-
-While the automated software doesn't isolate all bivalents/chromosomes from each cell (on average 17), we assume that the isolation process is not biased.
-Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.
-
-The mouse averages for 3 bivalent level metrics will be analyzed across these questions:  
-1)SC lengths (or total.SC)/chromatin compaction (using all cells regaurdless of CO number) 2) Interfocal Distance of 2CO bivalents and 3) the normalized foci position from 1CO bivalents.
-
-We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
-
-These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.
-
-(We will first list the results for the Q1)
-
-These are our predictions for the sex specifici patterns on the single bivalent lanscape
-
- Q1 Analysis, Predictions for Heterochiasmy Q1
-
-Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
-
-1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
-
-2. Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (deBoer et al 1996, petkov 2001).
-
-The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.
-
-# Q1 SC Lengths
-
-Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of PWD and MSM.
-A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-
-Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength)
-                         )
-
-Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_species(Q1.real.short_mouse_table)
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-Q1.real.short_mouse_table$species <- factor(Q1.real.short_mouse_table$species, ordered = FALSE)
-
-Q1.real.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength)
-                         )
-
-Q1.real.long_mouse_table <- add_strain(Q1.real.long_mouse_table)
-Q1.real.long_mouse_table <- add_sex(Q1.real.long_mouse_table)
-Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)
-
-```
-
-<!--Brief lit review for SC length and MLH1 expections.
-add XX adjusment / arguments
-
-In order to test one of the major/known correlates with genome wide recombiation rate variation, we compared the average
-
--->
-
-The plots below show the distributions of mouse average total SC area.
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-#nested boxplots, pooled Bivalents
-
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
-  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
-```{r Q1.tot.SC_show, echo=FALSE}
-
-Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
-
-Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
-   "WSB","G","LEW","PERC", 
-  "PWD", "SKIVE",  "KAZ",
-  "MSM", "MOLF"), order=T )
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#order the strains
-Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
-  arrange(strain) %>%
-  mutate(mouse = factor(mouse))
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#add category mean
-#totSC.strain.av.table %>% filter
-totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
-
-totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
-
-Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")
-
-Tot.SC_mouse.av_Dom
-
-totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
-totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
-
-Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
-
-Tot.SC_mouse.av_Musc
-```
-
-these are the number of mice for each of the categories
-
-`r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )`
-
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")
-
-#error not enought female observations
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-The full data set of SC lengths illustrates the general pattern than females have longer SC-AE.
-
-The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
-`r Q1_tot.sc_t.test.WSB$p.value`, `r Q1_tot.sc_t.test.G$p.value`, and `r Q1_tot.sc_t.test.LEW$p.value`
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _ respectively).
-
-Some of the p values for PWD and MSM are much closer (less significant), ( p = `r Q1_tot.sc_t.test.PWD$p.value`, `r Q1_tot.sc_t.test.KAZ$p.value`, and `r Q1_tot.sc_t.test.MSM$p.value` for the _M.
-m.
-musculus^PWD^ _, _M.
-m.
-musculus^KAZ^ _, and _M.
-m.
-molossinus^MSM^_ respectively)
-
-$$mouse \ average \ total \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.lmer_totalSC, echo=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA!!
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-Q1.lmer.totSC_results
-
-#G
-```
-there are more significant effects for strain and subsp for the mixed model of mouse.av total sc
-
-$$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.glm_totalSC, echo=FALSE, }
-#total-SC
-
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-#summary(Q1.glm_totSC_M2)
-#subsp effect, many strain, #male
-
-#G * malev.sig and G*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-#summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-
-```
-
-```{r show.short.biv, echo=FALSE}
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
-Q1.SC.short.mouse_scatter_boxplot
-
-#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
-skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-#skiveQ1.SC.short.mouse_scatter_boxplot
-#there are 2 lowish female means,
-
-```
-
-When skive is plotted alone it does have overlapping ranges.
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
-#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
-#1 male, 3 females measures
-#try to merge the manual - whole cell observations (at least 2 males)
-
-```
-
-The pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data
-
-with the mouse average for 'short' bivalents
-
-(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r Q1_mus.av_shortSC_t.test.G$p.value`, and `r Q1_mus.av_shortSC_t.test.LEW$p.value`
-for WSB, G and LEW respectively.
-
-(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value` for PWD, KAZ, SKIVE
-
-for the short bivalents are not significantly differen, for this strain there are 4 female and 4 male
-
-** there are more SKIVE female mouse observations -- but they are filtered out because I excluded them from the MLH1 dataset.  -- for low quality -- they might have good total SC spreads -- // but
-
-* it think total sc only has 1 female observation but the short bivalent has 4 skive females
-
-<!--
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
--->
-
-Mixed models for the sex differences in mouse averages for short bivalents (reduced dataset)
-
-$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1_short_biv_lmer, echo=FALSE}
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-
-```
-
-Sex is the most significant factor for the mixed model of mean short bivalents.
-
-$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.SC.Length_M2, echo=FALSE}
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-#skive strin very slight sign
-#male highly sig
-#skive * male slightly sig
-```
-
-skive strin very slight sign, male highly sig,  skive * male slightly sig
-
-$$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.glm_short.SC_M3, echo=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
-
-summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-male is the only significant p value from this model.
-
-<!--
-Because of confonding effects across chromsomes of different physical lenghts, we attempt to ...
-reduce the error from mixing chromosomes by constructing reduced bivalent sets, for the longest 5 and shortest 5 bivalents from the same cell.
-The mouse means for these two data sets were then calculated.
--->
-
-Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
-
-Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
-
-The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#mouse av
-Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE}
-#removed these plots for space
-
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-Q1.long.biv_cell_table <- ddply(real.long.bivData, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-Q1.long.biv_cell_table <- add_mouse(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_strain(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_sex(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_subsp(Q1.long.biv_cell_table)
-
-Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-#mouse av for long biv
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE, include=FALSE}
-###
-lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value
-
-Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
-sex.results =  Q1lmer.MM_SC_long_sex, 
-               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
-Q1lmer.MM_SC_long_strain)
-```
-
-# Q1. General CO Positions
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE}
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-Q1.long.1COpos_mouse.av <-  ddply(Q1.long.biv_1CO, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength), 
-                        mean_F1pos_PER =  mean(PER_Foci_1),
-                         mean_F1pos_ABS =  mean(Foci1)
-                         )
-Q1.long.1COpos_mouse.av <- add_strain(Q1.long.1COpos_mouse.av)
-Q1.long.1COpos_mouse.av <- add_sex(Q1.long.1COpos_mouse.av)
-Q1.long.1COpos_mouse.av <- add_subsp(Q1.long.1COpos_mouse.av)
-
-Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-Q1.short.1COpos_mouse.av <-  ddply(Q1.short.biv_1CO, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength),
-                        mean_F1pos_PER =  mean(PER_Foci_1),
-                         mean_F1pos_ABS =  mean(Foci1)
-                         )
-
-Q1.short.1COpos_mouse.av <- add_strain(Q1.short.1COpos_mouse.av)
-Q1.short.1COpos_mouse.av <- add_sex(Q1.short.1COpos_mouse.av)
-Q1.short.1COpos_mouse.av <- add_subsp(Q1.short.1COpos_mouse.av)
-
-```
-
-<!-- motivating
-
-To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
-In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
--->
-
-we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
-We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ).
-(all chromosomes in house mouse are acrocentric)
-
-Chromosome size effects are a confonding factor for CO position, with shorter chromosomes having a more uniform landscape compared to larger chromosomes (cite).
-To account for this we again use the reduct bivalent sets (long.biv and short.biv)
-
-Below are the general patterns / plots -- short and long bivalent groups are shown to try and account for the chromsomes size effect.
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-Q1.mouse.av.biv_scatr_box
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
-           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test(
-
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (with female)
-are highly significantly more medially placed (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
-With the musclus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
-Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).
-
-<!-- lmer and glm here -->
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-For the Mixed model, sex and strain are significant effects for the normalized single foci position.
-
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=FALSE}
-
-summry.Q1.glm_F1pos_M2
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-```
-
-Lew and KAZ had slightly significant strain speicific effects across the two glms (these results will be explored more fully in the Q2 section).
-The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).
-
-<!--
--full 1CO tests
-
-long 1CO tests
-
-short 1CO tests
-
-There are a few traits that fall within the CO positions  
-- 1CO normalized position  
-- sis.co.ten (sorta interference)  
-- centromere and telomere distance  
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-  C) Centromere and telomere distances will be sexually dimorphic.
-**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
--->
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-
-Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-Q1.short.biv_scatr_box
-
-Q1.long.biv_scatr_box   <- ggplot(Q1.long.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-Q1.long.biv_scatr_box
-```
-
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
-
-#make the same plots as above but with the long biv data set
-
-#mouse.av.long.biv.table
-Q1.av.long.biv_scatr_box   <- ggplot(Q1.long.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-Q1.av.long.biv_scatr_box
-
-Q1.av.short.biv_scatr_box   <- ggplot(Q1.short.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-Q1.av.short.biv_scatr_box
-
-```
-
-```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
-#IGNORE THIS FOR KNOW, THIS
-
-#filter all Curated biv data for 1 strain
-#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
-
-# the stardell paper sums the number of COs for bins (10)
-
-#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
-#but then the model would be close to logistic
-
-#summing the number of COs per bin, losses the chrom specific data....
-#!! need to use a melted Data frame to capture all nrm.foci obs
-#wsb.ttest<- mouse.avs_4MM %>% filter()
-#start with same chrm classes
-#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
-
-WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
-
-PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
-
-anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
-
-#or AIC(model1,model2)
-
-```
-
-```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
-
-#make a table for the sex specific values of the normalized F1 positions
-
-#for strain, sex -- Curated_BivData
-#or (Mouse.Table_BivData_4MM)
-#norm.first.foci.pos
-
-Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
-
-                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
-                  n1CO = sum(n1CO )
-
-                          )
-
-```
-
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#MELT_Curated_BivData
-
-#these isn't long.biv data
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-real.long.bivData
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-The mixed model data should only come from 1CO bivalent data.
-
-the mouse average foci1 pos is more significant in t.test, but not log regression...
-(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
-```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-<!--
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
-drop1(lme_MM_telo.dist, test="Chisq")
-
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
--->
-
-#Q1. IFD
-
-In order to test if there are sex differences between crossover interfernce, a major determinant of the positioning of crossovers along chromosomes, we examined ...
-ran models with the interfocal distance (IFD) of two foci on the same bivalent.
-
--We focus on observations from two-crossover bivalents for more comparable observations.
-
--comparisons of all IFDs of mulit crossover bivalents is also included
-
-Interference is a major determinant of the positioning of chromosomes
-
-<!-- 
-Brief lit review for IFD / interference expectations  
-**Petkov et al 2007,  CO interference underlies sex differences in RR**
-
-"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
-However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
-at the pachytene stage of meiosis."
-
-Chrm1 genetic map, 
-(other human and mice refs 
-Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
-
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-<!--
-ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
-
-explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
--->
-
-```{r show.IFD_99, echo=FALSE, warning=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r IFD.old.plots, echo=FALSE, eval=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
-Still working on the best way to display the general IFD patterns.
-
-### Mixed Model Tests, Fixed Effects
-
-Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
-
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.IFD.M1, echo=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#M1_IFD_2CO
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-#don't understand why 2 rows?
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
-#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-#exactRLRT(lmer.IFD.PER_MM.M1)
-
-```
-
-The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
-That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
-The tables below are from anova( for the lmer model ).
-Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
-
-$$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r IFD_HetC_M2, echo=FALSE}
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
-
-```{r IFD_HetC_M3, echo=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-
-```
-
-For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.
-
-the interaction effects were slightly significant for both raw and nrm.IFD.
-
-The most significant value was the sex effect for nrm.IFDs.
-
-The the random strain effect was not significant for either model.
-
-For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
-The other effects are for PWD ansD SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
-For the M3, only SKIVE*male effect is close to significant
-
-For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
-SKIVE*male is the only other consistantly significant effect, which also increases the nrm.IFD measure.
-
-Overall There's a low amount of significant effects across the 2CO IFD measures.
-This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
-
-## Strain Comparisons
-
-Dive deeper into the sex specific pattern for each strain.
-Below are code chunks which show the unusual sex specific results for IFD measures.
-The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
-The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
-
-For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
-
-```{r Comp_Ratios, echo=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
-
-# Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
-
-<!-- keep blank for now 
--basic stats for the new dataset
-
--->
-
-# Q2 SC
-
-In order to test if sc lengths (sc area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
-
-The average sc lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confonded by the different proportions across strains.
-
-First we examine the pattern of SC length variation across chromosome classes.
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-Q2.SC.length_Nested_boxplots_Dom2
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-Q2.SC.length_Nested_boxplots_Musc2
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-Q2.SC.length_Nested_boxplots_Mol
-
-#+theme(legend.position="none") 
-```
-
-**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.
-
-The positive correlation of sc lengths has a positive correlation with the number of COs on single chromosomes ,
-
-in the musculus and molossinus strains, the positive correlation is more
-
-In Domisticus the -- postive correlation only holds for the first 3 classes of bivalents (0CO, 1CO, 2CO) -- this illustrates how the differences between proportions of chromosome classes across strains confonds comparisons of SC length across strains.
-
-The glm's for the full dataset across -- the coefficients are listed below
-
-```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
-```{r wacky, echo=FALSE}
-
-real.long.bivData$bin <- "long"
-real.short.bivData$bin <- "short"
-
-lng.shrt <- rbind(real.long.bivData[,c(1:2, 17,19)], real.short.bivData[,c(1:2, 17,19)])
-
-uu <- ggplot(lng.shrt, aes(x=fileName, y=chromosomeLength) )  + geom_point()+facet_wrap(~strain, scales="free")
-
-```
-
-We purpose -- we think that running the models for comparing SC means by chromosome class and devided by rec group.
-(since the chromsome porportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )
-
-The null hypothesis is that these won't be highly significant strain effects.
-If there are still strain effects fo these models, that would be evidence that there has been eevolution for SC length compaction across these strains
-
-## Total SC
-
-`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
-X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
-
-```{r show.tot.sc, echo=FALSE}
-
-#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, mouse) %>%
-#  mutate(file.name = factor(file.name))
-
-mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-mouse.av_sc_male
-
-```
-
-The skeletonized measures show that the general pattern holds for total SC and the high rec strains.
-However there are a few exceptions -- However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel
-
--the SKIVE are within dom and kaz level -- so why does SKIVE have inntermediate CO proportions and gwRR
-
-24sep14_PWD_m1 is lower than the others
-5jul17_MSM_m1 is the low MSM
-
-The CZECH and AST are higher than KAZ
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE}
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
-The number of cells and mice is difference across the low group strains, 
--- permut the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)
-
-```{r Q2.SC.Length.stuff, echo=FALSE}
-#all pooled SC lengths
-#WSB     G   LEW  MOLF   KAZ CZECH 
-#    5     9     5     4     3     3
-
-#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14
-
-#G sample 5 mice
-
-#break down of mice in the low rec dataset
-
-#glm.1CO
-Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
-summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
-round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)
-
-#glm.2CO
-Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
-summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
-round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
-#fewer significant differences
-
-Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
-sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
-round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
-#all but KAZ are slightly significant, G and CZECH more so
-
-#glm.3CO
-Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
-#not sig differences
-summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
-round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)
-
-#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)
-
-#all pooled SC lengths
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-
-```
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.long.biv.table <- add_mouse(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-
-mouse.av.short.biv.table <- add_mouse(mouse.av.short.biv.table)
-mouse.av.short.biv.table <- add_strain(mouse.av.short.biv.table)
-mouse.av.short.biv.table <- add_subsp(mouse.av.short.biv.table)
-mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
-mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
-mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
-```
-
-```{r Q2.SC.long.biv_2, echo=FALSE}
-shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2.SC.ploy.male, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-
-#1. determine which strain effects are sig
-#use mouse biv average
-#Male.poly.Mouse.Table_BivData_4MM
-
-glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#SC many strains are different (more than the gwRR)
-#subsps are sig
-
-glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#Dom
-glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glm_SC.male_Dom)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glm_SC.male_Musc)
-
-#Mol
-glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glm_SC.male_Mol)
-
-#Logistic regressions
-
-log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-
-                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
-
-#summary(log.SC.male.poly)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_musc)
-#not sig -- when just looking at Musc
-
-#significant in predicting group
-log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_mol)
-#not sig -- when just looking at Mol
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-glm.Q2.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
-round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
-
-table(mouse.av.long.biv.table$mouse)
-
-glmQ2.M2.SC.male_long.biv <- glm(mean.SC ~ strain, data= mouse.av.long.biv.table)
-round(summary(glmQ2.M2.SC.male_long.biv)$coefficients[,c(1,4)], 6)
-```
-In the cell average Long biv data set, MSM G and CZECH strains have sig strain effects
-
-very few strains have significant values for the mouse
-
-```{r Q2.glm.SC.short, echo=FALSE}
-glmQ2_SC_mouse.av.short_biv <- glm(mean.SC ~ strain, data= mouse.av.short.biv.table)
-
-round(summary(glmQ2_SC_mouse.av.short_biv)$coefficients[,c(1,4)], 6)
-```
-Basically all strains have a significant effect for the mouse.av of long biv
-
-```{r Q2.glm.SC.all.strain, echo=FALSE}
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
-
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
-
-#summary(glm_SC.male_Musc_long.biv)
-
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
-
-## logistic regression for Rec group
-#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
-
-#summary(log.SC.male_long.SC)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-```
-
-abova are glm tables for mouse.av SC lengths for each strain.
-
-Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
-
-The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).
-
-The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high rec strains then low rec strains.
-This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
-Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
-This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
-They have a lower probability of 2COs below a certain SC length threshold.
-
-I think these comparisons should be done with the **Bivalent observations**
-The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class
-
-```{r Q2.SC.Chrm.Class, echo=FALSE}
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
-#there might not be enough data 
-#
-ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
-```
-
-These are the plots of the logistic regression.
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
-
-```{r Q2.SClength.ttest, echo=FALSE}
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
-
-```
-
-The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths.
-But When the means are compared across chrm class, the mouse averages are no longer significant.
-
-How well does SC length predict chromosome class?
-Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class.
-(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-<!--
-blocks of other male Q2 - SC lengths
-
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
-
-#comparing all the strains
-
-T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-)#p-value = 0.001567
-
-#1CO
-T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#2CO
-T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#3CO
-T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#Single Biv level
-T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
-
-#1CO
-T.test_SC.Len_single.biv_1CO <- t.test(
-  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
-
-      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
-  )
-
-#2CO
-T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
-
-                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
-  )
-
-  #3CO
-  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
-
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
-  )
-```
-
-The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-
-```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
-#I don't think the Mixed models fit well for the high vs low polymorphism tests.
-
-#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
-#MSM_SC vs MOLF_SC
-
-#2. logistic regression for sc length and chrm class?
-#if the prediction / significance
-
-#isolate data
-#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
-
-#build and test MM
-lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-
-lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
-
-anova(lmer_MM_SC_length_male.poly)
-
-rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
-```
-
-```{r long.sc_log.res_meanSC, eval=FALSE}
-#Long Biv comparison
-
-#hold off on reporting/ working with the long SC biv resutls
-
-long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
-long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
-                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
-
-long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
-                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
-
-summary(long.biv.log.reg)#NS effect (likelt underpowered)
-```
-
--->
-
-> **Summary**
-
-When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
-I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
-(IE the scatter plots of SC length, binned by chrm class on the X).
-
-Each point is a mouse average SC.
-When all bivalents are pooled, the logistic regression SC lengths predict the groups.
-When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
-
- and the mean 2COs SC are higher in the high strains
-(this must have something to do with the physical size effect)
-
-```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-table( droplevels(fomo.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
-
-#Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-```
-
-```{r SC.length.male_tests, echo=FALSE}
-
-## run the tests to check following expectations
-
-#1.Dom strains not different from ech other
-#2. PWD and SKIVE are different from the other strains
-#3.MSM and MOLF are different from each other
-
-#Male.poly.Mouse.Table_BivData_4MM
-#Long.biv_mouse_table_male.poly
-
-#decide between lm or glm..?
-#mouse avs, males
-
-#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
-
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
-#the subsp labels get messed up, unless they are designated chr first
-Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
-
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
-Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
-#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
-#total sc sig, 1CO sig, 2CO sig
-
-Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
-#0.000369 ***
-Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
-#0.000816 ***
-Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
-#0.0148 *
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
-#making plots above show that mouse average SC lengths are lower could there be microscope issue?
-
-#make sure to remove some of the juvi mice!! from the AP_mouse table
-show.Dom.male.CO <- ggplot(
-male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
-#P not sig for CO mean
-
-#seems like there are more mice in the MLH1 data set than bivData
-#
-
-#WSB all dissected and imaged in '15
-#G imaged 15 - 16
-#LEW imaged 15 to 17
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "1CO SC Length", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "2CO SC Length", xlab = "")
-```
-
-```{r SC.length.pairwise.T, echo=FALSE}
-#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
-
-male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
-male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
-
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
-#PWD and MSM are most diff from SKIVE?
-
-Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
-#the pooled SC value tests are sig
-
-#the mol comparisons are different due to adj p values
-
-Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
-#no sig (only 3) -- does include CZECCH
-
-#non-sig, with mol pooled with musc
-
-Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
-
-Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
-
-Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
-#non of these sig with the mulitple adjust
-
-Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
-
-#pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```
-
-If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
-
-Ideas for the above 3 tests,
-
-1. no difference in DOM
-- no sig logistic regression test
-- glm, strain effect?
-
-> Is sex a significant effect for SC length? (as predicted)
-
-- Sex is a significant effect for SC length
-The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
-
-- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
-
-- The Long Biv Data set largely agrees with the full curated dataset
-
-- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
-
-```{r log.reg_SC.length, echo=FALSE}
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-The mean SC logistic regression model for mouse averages
-
-`r log.reg_SC.length_mouse.av$coefficients`
-
-and for the single bivalent levels
-
-`r log.reg_SC.length_BIV$coefficients`
-
-When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
-When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
-
-> Is the prediction, high rec musc male strains have long SC met?
-
-In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
-I couldn't get the Mixed models working for the male polymorphism predictions...
-
-# Q2 Normalized CO positions
-
-A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.
-
-or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.
-
-<!-- 
-Main DFs for this section:
-Male.poly.Mouse.Table_BivData_4MM
-male.bivdata.1CO
-
-**Polymorphism**
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.  
-C.3) telomere and centromere distances will be shorter in low rec strains
-
--->
-
-Since there aren't many good predictions for how the 1CO normalized landscape relates to gwRR variation, these will be the main questions
-
-- Has there been evolution in the 1CO pattern?
-
-- Is there a consistent pattern with the high rec strains?
-
-Normalized 1CO foci positions will be used. (F1_PER). Check fixed effects of subsp and strains.
-
-sis-co-ten, telomere and centromere distance of foci are metrics which draw from a wider pool of samples. (So maybe I will use them... but avoiding focus on them)
-
-```{r Q2.show.nrmF1, echo=FALSE}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
-
-Q2.show.nrmF1
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
-
-nrmF1_mouse_level
-```
-
-Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.
-
-```{r Q2.F1.pos.LONG.BIV, echo=FALSE}
-
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Q2.show.nrmF1_long.biv
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-Q2.shoooo
-
-```
-
-So the plot for the long bivs, shows that the PWD, SKIVE and MSM have on average more centrally placed F1 positions.
-It's kinda werid that the glms/models suggest that WSB has the most terminal F1 -- but, out of all the strains, for the long bivs, WSB has the lowest mean.
-
-But there are v few observations. I think the plot shows that there are fewer 1CO for the High Rec group. not sure there would be enough for feeding into models/
-
-```{r Q2.1CO.landscape, echo=FALSE}
-#Male.poly.Mouse.Table_BivData_4MM
-#Curated_BivData_male
-
-#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#M1
-glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm.M1_Q2.1COnrm)
-
-#M2
-glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm.M2_Q2.1COnrm)
-
-#Dom
-glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glmQ2_M2.1COnrm_Dom)
-
-#Musc
-glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glmQ2_M2.1COnrm_Musc)
-
-#Mol
-glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glmQ2_M2.1COnrm_Mol)
-
-#m3 (log)
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm_Musc)
-
-#Mol
-Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
-              family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-#biv level logistic regression
-#Curated_BivData_male
-#male.bivdata.1CO
-
-male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
-
-Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.Q2.F1_PER_Musc)
-
-#Mol
-Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
-              family=binomial(link="logit"))
-
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-```
-
-```{r Q2.t.test_COlandscape,echo=FALSE}
-
-#MOUSE AVE
-Rec.group.mouse.F1pos_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.F1pos_Ttest$p.value
-#p-value = 0.9098
-
-#single biv level
-Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-Rec.group_BivData_F1post_Ttest$p.value 
-#p-value = 0.238
-
-```
-
-The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
-
-#### Summary
-
-Not really a clear summary for the normalized 1CO position and rec groups.
-
-# Q2 IFD
-
-this is the Q2 prediction for IFD / interference
-
-> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-1. glm across subsp
-
-2. anova within subsp
-
-3. logistic regression for musc (or all males)
-
-```{r Q2.show.IFD, echo=FALSE}
-
-#subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
-
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-biv.IFDs_PER_by.group
-```
-
-The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
- The normalized plot displays the pattern of higher rec group having slightly higher longer IFDs.
-
-```{r Q2.IFD.glm.test, echo=FALSE}
-#glm --
-#mouse aver
-
-#M1.ABS
-glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.raw) 
-#M1.PER
-glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.PER)
-
-#M2.raw
-glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.1COnrm)
-
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD_Mol)
-
-#M2.PER
-glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.IFD.PER)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD.PER_Mol)
-```
-
-```{r Q2IFD_logreg, echo=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.PER, echo=FALSE}
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
-
-#t.test for IFD
-#mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-t.test_mouse.av.IFD_PER <- t.test(
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-```
-
-```{r Q2.t.test_IFD,echo=FALSE}
-
-#already made
-#subset
-male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-#t.tests for the Rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
-                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
-
-Rec.group.BivData_IFD.PER_Ttest$p.value
-#p-value < 2.2e-16
-
-#MOUSE AVE
-Rec.group_IFD.PER_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group_IFD.PER_Ttest$p.value 
-#p-value = 6.872e-07
-
-```
-
-The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
-
-### Q2 Long Bivalent Dataset
-
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.
-
-```{r Q2.long_biv.IFD, echo=FALSE}
-#long BivData
-male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
-
-#show
-Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-
-Long.biv.IFDs_PER_by.group
-
-```
-
-table(droplevels(male.long.biv.2CO_IFD$strain) )
-
-The plot above shows the normalized IFD measures from 2CO bivalents
-
-There are many fewer long bivalent data observations, but the same positive correlation with
-
--table of long bivalents
-
-- glms
-
-- logistic regression
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-
-```
-
-There are not enough observations for sub setting the data into strains or use the mouse averages.
-
-Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
-The above t.tests are breaking the knitr
-
-The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
-
-None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
-
-Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
-This also indicated that interference distance may evolved in the house mouse complex.
-
-#Caveats
-
-put all of the code chunks/analysis for caveats here
-
-## Chromosome Size Effect
-
-I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)
-
-<!-- 
-Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
-But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
---> 
-Below are examples of plots of SC length distributions across cells.
-The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
-Most plots excluded for space.
-
-Each point is a bivalent plotted by cell on the x axis.
-X's are the 4th quartile, big point is the mean and smaller black point is the median.
-I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
-(the extra stats are not correctly mapped)
-
-```{r SC.distb_manual_Auto, echo=FALSE}
-
-#calq the mouse means
-manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
-                        median_SC = median(SC.length, na.rm = TRUE),
-                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
-#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
-
-)
-
-manual_summary_stats <- add_mouse(manual_summary_stats)
-manual_summary_stats <- add_strain(manual_summary_stats)
-manual_summary_stats <- add_sex(manual_summary_stats)
-manual_summary_stats <- add_subsp(manual_summary_stats)
-
-###
-#calq the mouse means
-Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        median_SC = median(chromosomeLength),
-                        Qar.4 = quantile(chromosomeLength)[[4]],
-
-                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
-        #add Q4
-
-)
-Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
-
-#colors_of_strains
-#scale_color_manual(values=colors_of_strains)
-
-#un-order these factors
-```
-
-```{r long.biv.plots, echo=FALSE}
-
-show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
-
-show.the.cells_females_dom
-#PWD seems to have the biggest variance across cells
-
-almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_DOM_f
-
-#### Musc female
-
-show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
-
-almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
-
-### Dom Male
-show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
-
-almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#MUSC male
-show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
-
-almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
-
-#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
- # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
-#show.the.cells_male
-
-#almost_whole_cell_BivData
-#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
-#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
-#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
-
-#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
-```
-
-All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
-It shows the amount of within cell variance across strains.
-There is a bit of variance across the SC length distributions in the PWD females.
-
-This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
-
-The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
-The full data set is `r length(Curated_BivData$Obj.ID)`.
-This is the breakdown of bivalent observations by category for this long dataset are
-
-**- Try to merge this DF with the whole.cell manual measures.**
-
-- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.
-
-in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
-10202 bivalents from 86 mice.
-
-```{r show.long.SC, echo=FALSE}
-
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
-female_long_biv_plot
-
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
-male_long_biv_plot
-
-```
-
-These plots show the SC lengths for the 'long SC data set'.
-They are supposed to be the longest 4-5 SC from cells where I could get good measures.
-These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
-Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
-
-## Adjusting for XX
-
-(rethink where this section should go)
-new outline
-
-1. illustrate problem(affects mostly SC length)
-
-2. Expected impact on sex comparisons, estimated effect size of the X  
-3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
-
-4. 19 female, 19male, 20female 20 male
-
-The female mouse averages should have adjustments for the XX.
-working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
-Subtract this amount from the female mouse averages ...
-This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
-
-- Of all female single bivalents observations, 5% are XX (1 of 20).
-
-- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
-
-- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
-
-I think a formula something like this can be applied to adjust for XX
-
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
-
-#### Whole Cell Manual Comparison
-
-```{r XX.adj_manual.whole.cell, echo=FALSE}
-
-#load and examine the the whole cell hand measures here
-
-female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
-data.frame(table(droplevels(female.whole.cell$category)))
-
-# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
-
-#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
-
-data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
-#missing some strains, most have 3 cells (min). MSM has 5 full cells
-
-#remove the incomplete LEW female cell
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
-
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
-
-#calq mean SC lengths by rank
-
-mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
-
-                       mean.sc =   mean(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                         se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
-
-top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
-
-show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
-
-show.SC_by.rank_top
-
-bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
-
-show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
-
-#CALQ PERCENT TOTAL SC
-
-CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
-
-                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       #rank.by.per
-
-                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
-                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
-                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
-                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
-                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
-                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-
-#MELT ABOVE
-MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
-              #leave out the Foci
-               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
-
-            id=c( "fileName","total.sc.cell","nbiv"))
-
-MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
-
-MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
-MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
-
-show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
-
-show_long.sc.proportions
-
-#original plot code 
-#ggplot(data = df1, aes(x, y, color = Group)) +
-#    geom_pointrange(aes(ymin = lb, ymax = ub), 
-#                    position=position_jitter(width=0.5), 
-#                    linetype='dotted') +
-#    theme_bw()
-```
-
-The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
-
-The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
-
-The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
-For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
-
-(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
-
-The other figure shows of each single bivalent contributes to the total SC area.
-Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
-on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
-So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
-
-- Is the difference between cell averages for males and females less that 10%?
-
-- also interesting, the pwd and msm don't have longer SC, compared to other strains.
-
-#### Automated BivData Comparison
-
-```{r XX.Automated.biv.stuff, echo=FALSE}
-
-#make automated female data to compare to
-female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
-
-female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
-
-table(female.Bivdata_comp$strain)
-
-#female.Bivdata_comp
-
-#1. calq the rate of chrm segmentation in female categories in automateted data
-
-box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
-
-                   Nmice = length(unique(mouse)),
-                   mx.box.num = max(boxNumber), # need to run this at the cell level
-                    N.biv  = length(Obj.ID),
-                   pass.biv = N.biv / mx.box.num
-)
-box.rates <- add_mouse(box.rates)
-box.rates <- add_strain(box.rates)
-box.rates <- add_category(box.rates)
-
-cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-
-cell.pass.rates
-
-category.box.rates <- ddply(box.rates, c("category"), summarise,
-
-                   mean.pass.rate = mean(pass.biv)
-)
-#category.box.rates
-
-#2. pull the 3rd rank biv's calculate the SC means
-#calq means by rank (focus on the top 5)
-
-```
-
-For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
-
-The table above shows the number of bivalents from the same strains as in the manual whole cell data.
-The plot shows the bivalent passing rate across all of the individual cells from this female data set.
-For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
-
-(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
-(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
-
-strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
-
-** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
-But I am not using 'whole cell' summaries to compare female and males.
-
-**What is the effect of an extra XX-autosome on single bivalent means?**
-
-use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
-Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
-Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
-Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
-The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
-
-# Note on Heterochiasmy Definition
-
-<!-- add outline for new XX adjust section 
-adjust the colors for the plot below (standardize)
--->
-
-I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
-In females the XX bivalent is indistinguishable from the autosomes.
-To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
-Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
-(I note if the and Y are paired, which they are at a high rate).
-The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
-To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
-
-1. Compile full-cell data from females (all 20 bivalents measured)
-2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
-3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
-
-According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
-
-(Put the XX adjustment section here)
-
-```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
-#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
-#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
-#remove quality 5 cells
-
-#outgroups
-#SPIC
-HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
-#unique(HetC.SPIC$mouse)
-#non-passing mice
-SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
-HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
-#calq sex mean
-SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
-
-#SPRET
-HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
-#unique(HetC.SPRET$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
-
-#CAROLI
-HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
-#unique(HetC.CAROLI$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-#cast
-HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
-#unique(HetC.CAST$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
-
-#HMI
-HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
-#unique(HetC.HMI$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-#dom
-
-#WSB
-HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
-#unique(HetC.WSB$mouse)
-#non-passing mice
-WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
-HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
-#calq sex mean
-WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
-
-#LEW
-HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
-unique(HetC.LEW$mouse)
-#non-passing mice
-LEW.remove.mice <- c("4jan17_LEW_f2")
-HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
-
-#calq sex mean
-LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
-
-#G
-HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
-#unique(HetC.G$mouse)
-#non-passing mice
-G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
-
-#calq sex mean
-G.mean.MLH1.obs <- ddply(.data=HetC.G, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#PERC
-HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
-#unique(HetC.PERC$mouse)
-#non-passing mice
-#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
-
-#calq sex mean
-PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#MUSC
-
-#PWD
-HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
-#unique(HetC.PWD$mouse)
-#non-passing mice
-PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
-
-# CZECH
-HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
-#unique(HetC.CZECH$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
-
-# SKIVE
-HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
-#unique(HetC.SKIVE$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
-
-# KAZ
-HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
-#unique(HetC.KAZ$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
-
-#TOM
-HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
-#unique(HetC.TOM$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
-
-#AST
-HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
-#unique(HetC.AST$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
-
-#Molosinus
-
-#MSM
-HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
-unique(HetC.MSM$mouse)
-
-#non-passing mice
-MSM.remove.mice <- c("12sep16_MSM_f3")
-HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
-
-#calq sex mean
-MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
-
-#MOLF
-HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
-unique(HetC.MOLF$mouse)
-
-#non-passing mice
-MOLF.remove.mice <- c("27oct18_MOLF_m3")
-HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
-
-#calq sex mean
-MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
-)
-
-MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
-
-```
-
-```{r HetC.plot, echo=FALSE, warning=FALSE}
-#HETC plot
-
-#set up data
-
-HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#
-#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
-
-HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
-
-#remove the 'other'
-
-HetC_strain_table.M$sex <- "male"
-HetC_strain_table.F$sex <- "female"
-
-HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
-
-HetC_strain_table <- add_subsp(HetC_strain_table)
-
-#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
-
-HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
-HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
-#str(HetC_strain_table)
-
-Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
-  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
-
-Het.C.plot
-#this plot is missing several
-
-```
-
-There is now MOLF, which has female biased hetC
-3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
-1 of the musc strains has female biased heterochiasmy, KAZ.
-
-```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
-
-#REMOVE This BLOCK?
-
-#### Distributions by mouse
-#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
-
-#remove this block since it has it's own src/script
-
-### to integrate 'pass'/fail
-### this should be in setup data
-
-#mice that don't have female data will break the chain-loop
-#MOLF female, HMI, CZECH female, CAROLI female, Tom male
-
-#instead of adding fake data, remove the categories from list
-
-#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
-#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
-#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
-#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
-#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
-#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
-#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
-#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
-#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
-
-pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
-
-#this loop makes a quant_status col for a list of mice
-for(i in 1:length(MLH1_data$mouse)){
-  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-}
-
-df_list <- split(MLH1_data, (MLH1_data$category))
-#remove sections which are in category but have no data
-#this is where you remove the non-data data!
-df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
-
-#7 PERC, 12 MOLF female, 
-#
-
-#df list, is a list of df's which were split() from BIG df by category
-df_list 
-o=0
-i=0
-
-MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
-
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-for(i in 1:(length(df_list))){
- # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                      mouse_nMLH1 = mean(adj_nMLH1.foci)
-  )
-# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-  for(o in 1:length(df_list[[i]]$Original.Name)){
-    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
-        }
-}
-
-plot_list = list()
-for (h in 1:length(df_list)) {
-  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
-    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
-    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
-    facet_wrap(~ mouse, drop = TRUE) + 
-    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
-  #make an empty ggplot? for missing data
-
-  plot_list[[h]] =   pf
-}
-
-#orig (I think this is the same as above)
-#plot_list = list()
-#for (h in 1:length(df_list)) {
-#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
-
-#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
-#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
-
- #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
-
-    #theme for indiviual mice, 
-#theme(plot.background = element_rect(fill = "green"))
-#  plot_list[[h]] =   pf
-#}
-#can't change the alpha/transparency of the red rec
-
-#aes must be same length ---   #geom_point(stat = "identity") +
-# the last one is correct....?
-
-#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
-```
-
-```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
-geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-#no plot is made that goes in the list -- so it can't be shown
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
- # invisible(lapply(obj, function(x) plot(x,main="some plot")))
-}
-
-#this stops at MSM
-```
-
-The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
-
-Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
-The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
-
-```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
-
-no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
-
-MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
-
-MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
-
-kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
-
-bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
-
-#add titles
-#use multiplot
-#remove X axis
-
-multiplot(kuku,bubu, cols=2)
-```
-
-```{r F1.by.F2, eval=FALSE}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
-
-```
-
-# Deleted notes
-
-```{r variance.hist, echo=FALSE, eval=FALSE}
-#figures of histograms,
-
-#make series for total dataset?
-Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("Histograms of within mouse variance, all mice")
-Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-Var.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("Histograms of within mouse variance, 10 cell min")
-Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
-
-Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
-Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
-
-Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
-
-Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
-Q12.var.hist.plot
-
-#some LEW female still have v high within mouse variance (2 mice, )
-
-```
-
-# References

### 2020-02-19 doc/Results.rmd
messy work from home

@@
 title: "MLH1 Results Outline"
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: 
 output:
@@
     toc_float: yes
+  html_notebook: default
   pdf_document: default
@@
 ```{r female.dif, echo=FALSE}
+
 #how much greater are the high strains?
 G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co)
+mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
 
 LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
-mean.LEW.female <- mean(LEW_female$mean_co)
+mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
 
@@
 MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSM.female <- mean(MSM_female$mean_co)
+mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
 
 sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
-mean.all.other <- mean(sans_high_female$mean_co)
+mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
 
@@
 MSM.female.diff <- (mean.MSM.female / mean.all.other)
-
-```
-
-The overall pattern for the female mean MLH1 counts per cell is low variation. ( in the most different strains
-In females specifically _domesticus^G^_ is `r G.female.diff` higher the the other means. 
-while _ domesticu^LEW^_  is `r LEW.female.diff` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.
-
+```
+
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+#variance plots
+musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot
+
+#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+#Mol.female.VAR.plot
+
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.female.VAR.plot
+
+```
+
+```{r Female.specific.glm, echo=FALSE}
+
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+lm.female.strain_sum <- summary(lm.female.strain)
+
+#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
+
+#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+
+```
+
+Within mouse variance was assess to follow up on the results from the full dataset.
+
+```{r FEMALE.VAR.MODELS, echo=FALSE}
+## within mouse variance models 
+lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#round( #round isn't working for this sumary
+#lmVAR.female.strain_sum$coefficients[,c(1,4)]
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+
+lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
+
+lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
+lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
+
+#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
+#LEW sig
+
+```
+
+The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (p= ).
+
+The overall pattern for the female mean MLH1 counts per cell is low variation. 
+(the most significantly differnt strains in the female data), For _domesticus^G^_ the G strain is `r G.female.diff` higher the the other means.
+While _domesticu^LEW^_  is `r round(LEW.female.diff,4)` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.
+
+ mouse mean ranges
 In the _domesticus_ subspecies the ranges of mouse means are 
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )`
-
-in _domesticus^G^_
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _musculus^PWD^_,
-
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to
-
-`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )`
+in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
+
+-These results are above the minimum of 20 which is the expected minimum following a obligate crossover per chromosomes.
 
@@
 
-just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")
+
+just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
 
@@
 ```{r male.diff, echo=FALSE}
+just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
 
@@
 
-- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`
-
-- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.
-
-The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r (mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co) )`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively.
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
+
+#variance plots
+musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.male.VAR.plot
+
+Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.male.VAR.plot
+
+Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.male.VAR.plot
+```
+
+```{r Male.specific.glm, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+
+summry.male.strain.effects <- summary(glm.male.strain.subsp)
+round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
+
+#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+
+#1. determine which strain effects are sig
+
+#M2
+lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
+lm.male.strain_sum<- summary(lm.male.strain)
+
+#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+
+#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+```
+
+the coefficients for glms of the CO means are listed above.
+
+```{r MALE.SPECIFIC.variance.model, echo=FALSE}
+## within mouse variance models 
+lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
+
+#all p values kinda high, MSM and TOM are slightly sig
+lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
+#non sig
+
+lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
+#NS
+```
+
+MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
+
+Two glm models were run to test the effects of strain and subspecies on mean MLH1 count per cell per mouse. The coefficients for the following model are below.
+
+<!--
+$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$ -->
+
+- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
+
+- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
+
+The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,
+
+`r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means respectively.
+
+**general linear models** -- indicate the 3 strains have significant strains effects
+
 Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
@@
 
-# Analysis of Evolutionary Patterns
+# Evolutionary framework
 
@@
 $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
--->
-<!--
+
 This is the break down of mice used within the Mixed model
@@
 
-```{r MLH1co.MM, echo=FALSE, digits = 3}
+```{r MLH1co.MM, echo=FALSE, warning = FALSE, digits = 3}
 #THIS IS THE NEW MIXED MODEL
@@
 
-Our models were designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
-
-All of the effects from the mixed model significant, (LRT; p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect).
-
 <!--
@@
 
-We followed up the mixed model effects with general linear models.
+Our models were designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
+
+All of the effects from the mixed model significant (Table X) so we followed up with general linear models.
 The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions.
@@
 In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
-These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.
-
+These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.
 Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
@@
 In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).
+
+Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
 
@@
 These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
-
 var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
 1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-
 cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
@@
 
-For both the mixed model and linear model, sex was the most significant effect while the subspecies, interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively.
-
-*notes on caveats for calculating variance -- lower power
+For both the mixed model and linear model, sex was the only significant effect while the subspecies, while interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively).
 
@@
 
-**female strongest effect, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.
+<!--*notes on caveats for calculating variance -- lower power-->
+For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell.
+This limits the power for detecting significance differences.
+
+**female strongest effect**, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.
+
+in the sex specific data was analyzed in glm's to test for significance differences in the magnitude of within mouse variance across strains.
+The male specific models, showed no significant strain effects
+
+For male specific glms for within mouse variance -- had no significant strain effects
+
+the female data models  --- inconsistant and moderate significance p values
 
@@
 motivation / background: maybe checkpoint / regulation
-
 limitations: observations at the mouse level, decreased power to estiamate the variance --
@@
 <!--
-Unequal variance across strains might indicate that the within animal variance for MLH1 foci per cell may have evolved, yet --our low power and number of mice is not equal, -- number of mice across
-
-Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
-<!--
-
-even tho it's an important trait for XX reasons, -- there are limitations for our analysis
-
-Unlike the mean MLH1 counts -- the mean per mouse is calculated across a sample of cells (which serve as replicates)
-
-the within mouse variance for MLH1 count per cell, on the other hand -- is calculated at the mouse level, var()
-
--- the replicates are at the level of mouse, unlike MLH1 counts per cell
-
-Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
 
@@
 }
-
 references for phenotypic variance connected to robustness to environment, increase in developmental consistency
 Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
-
 phenotypic variance leading to adaptation - 
 (Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
-
 Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
@@
 
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-#variance plots
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-#Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.male.VAR.plot
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.male.VAR.plot
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.male.VAR.plot
-
-```
-
-MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
-
-Two glm models were run to test the effects of strain and subspecies on mean MLH1 count per cell per mouse. The coefficients for the following model are below.
-
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
-
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-lm.female.strain_sum <- summary(lm.female.strain)
-
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-
-```
-
 For mean CO count, _domesticus^G^_  strain effects in both models (p = )
@@
 
-Within mouse variance was assess to follow up on the results from the full dataset.
-
-```{r FEMALE.VAR.MODELS, echo=FALSE}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-#round(
-#lm.m1.Q12_female_sum$coefficients[,c(1,4)]
-#LEW not sig
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
-
-The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (p= ).
-
 MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
@@
 Two models were fitted to the male specific data to test the effects of strain and subspecies.
-
-```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-summry.male.strain.effects <- summary(glm.male.strain.subsp)
-round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#1. determine which strain effects are sig
-
-#M2
-lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
-lm.male.strain_sum<- summary(lm.male.strain)
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-
-```
-
-the coefficients for glms of the CO means are listed above.
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-```
 
@@
                ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
 ```
@@
 
-DMC1.mean.table.99
-
 WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
@@
 DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-as.data.frame(DMC1table.main)
+
+kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
 
@@
 
-In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes with different numbers of COs.
-The two plots show the chromosome class proportions from hand measured and the curated BivData.
-
 ```{r chrm_prop_data, echo=FALSE, warning=FALSE}
@@
 
-```{r test.Chrm.Prop, echo=FALSE, eval=FALSE}
+```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
 
@@
 
+#all females
+
 chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
 male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
+
 male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
@@
 
+#includes SKIVE
 Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
@@
 
-High2_male.0CO.prop.test <- prop.test(chrm.prop.male.high2$n0CO, chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO, chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO, chrm.prop.male.high2$nbiv )
+High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
+High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
+High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
 #these proportions are v different
 
+#no SKIVE
 High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
@@
 
-#Need to run the chrm proportions thru melt code.
-
-chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
-chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check
-
-#chrm.class.DF$biv.sum.check
-```
-
-These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes. Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs.
-
-Most all the the p values for the proportion tests are significant, indicating there are slight but significant shifts across the classes of chromosomes.
-However the most striking male pattern is the propotion of 2COs
+#MSM-MOLF
+MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+#PWD-KAZ
+PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+```
 
@@
 
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",
-                                                                 "Cast",
+chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
                                                                  "Musc",
@@
 
+```{r Chrm.Prop.misc.calulations, echo=FALSE}
+
+total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
+
+sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
+sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
+```
+
+<!--
+- This is motivation to investigate more traits at the chromosome level
+1.	Basic stats - bivalent data numbers 
+2.	Rang of co per chrm, mostly 1co 2co
+3.	High rec strains have evolved a higher proportion of 2CO
+For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
+MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
+4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
+-->
+
+In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes grouped by the number crossovers on the same bivalent.
+A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).
+
+**most of the male proportions are 2/3's 1CO and 1/3 2COs. **
+
+Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
+They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
+This significant p value indicates that at least one of the six strains had significant differences).
+
+**The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
+
+PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
+
+While the proportions are significantly differennt, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
+ and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+
+The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`).
+This relpicates what has been notice in the literature (cite stapley).
+
+Most of the bivalents are 1CO and 2CO class bivalents -- and the proportion of these two classes distinguish high and low rec male strains. (distinguish the high rec males from minimal rec groups)
+
 (A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class.
 The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.
-
-- High female strains, G and Lew have significantly more 2CO bivalents.
+*(much smaller proportions and variation in the fringe classes*
+
+- High female strains, G and Lew have significantly more 2CO bivalents -- there are some differences in the female strains.
 
@@
 
-Error compared to human measures and other details have been discribed in (@Peterson_2019).
+Error compared to human measures and other details have been discribed in (Peterson_2019).
 

### 2020-02-19 doc/Results.rmd
added rec.group col to main bivData DF

@@
 
-load(file="~./MLH1repo/data/CleanBivData_13.2.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_19.2.20.RData")
 
@@
 
+```{r juvimice.list, echo=FALSE}
+juvi_mice <- c(
+  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
+  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
+  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
+
+  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
+   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
+  )
+
+#REMOVE JUVI mice
+#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
+#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
+
+```
+
 ```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
@@
 #the mouse averages will be used
+
+#has 
+#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
+Curated_BivData <- Curated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
 
@@
 
-#code intp chr
+#REMOVE JUVI mice
+juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
+mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
 
@@
 
+mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+
 ```
@@
 male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
+
+#REMOVE JUVI mice
+juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
+mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
+
 ```
@@
 OG.totSC.mouse.av.table <- totSC.mouse.av.table
+
+#REMOVE JUVI mice
+tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
 
@@
 
-# Q2 SC
+# Q2 SC Length
+
+<!-- place set ups for total SC, short biv and long biv here -->
+
+I made a sereies of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromsomes (possitive correlation of sc length and CO number).
+Yet, given the complications from the chromsomes proportions outlined above, comparing the bivalents across chromosome classes is complecated
+
+```{r juvimice.list.copy, echo=FALSE}
+juvi_mice <- c(
+  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
+  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
+  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
+
+  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
+   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
+  )
+
+ppmm <- OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$mouse  %in% juvi_mice, ]   
+```
+
+```{r Q2.SC.short.long.setup, echo=FALSE}
+table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+
+Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                         #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+                         )
+
+Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
+Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
+
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                        #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+                         )
+
+Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
+
+Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
+Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
+
+Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
+```
 
@@
   ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-Q2.SC.length_Nested_boxplots_Dom2
 
@@
   ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-Q2.SC.length_Nested_boxplots_Musc2
 
 Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-Q2.SC.length_Nested_boxplots_Mol
 
@@
 
-**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.
-
-The positive correlation of sc lengths has a positive correlation with the number of COs on single chromosomes ,
-
-in the musculus and molossinus strains, the positive correlation is more
-
-In Domisticus the -- postive correlation only holds for the first 3 classes of bivalents (0CO, 1CO, 2CO) -- this illustrates how the differences between proportions of chromosome classes across strains confonds comparisons of SC length across strains.
-
-The glm's for the full dataset across -- the coefficients are listed below
-
-```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
-```{r wacky, echo=FALSE}
-
-real.long.bivData$bin <- "long"
-real.short.bivData$bin <- "short"
-
-lng.shrt <- rbind(real.long.bivData[,c(1:2, 17,19)], real.short.bivData[,c(1:2, 17,19)])
-
-uu <- ggplot(lng.shrt, aes(x=fileName, y=chromosomeLength) )  + geom_point()+facet_wrap(~strain, scales="free")
-
-```
+```{r Q2.SC.long.biv_plots, echo=FALSE}
+
+shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+mouse.shoo.long.biv
+```
+
+```{r Q2.SCshort_plot, echo=FALSE}
+
+```
+
+```{r Q2_tot.sc_plot, echo=FALSE}
+
+#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, mouse) %>%
+#  mutate(file.name = factor(file.name))
+
+mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
+
+mouse.av_totsc_male
+
+```
+
+```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
+#there might not be enough data 
+#
+ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
+```
+
+<!-- basic stats -->
+
+total SC for males (which ar)
+X additional cells were added to the male specifi total SC data set discirbed above
+
+-reduced bivalent dataset (long and short) is made up of bined bivalents from X cells and X mice
+
+<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
+remember the pairwise t.test
+Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
+G-WSB is sig, WSB-LEW are slightly sig
+
+-->
+
+```{r Q2.totSC.ttest, echo=FALSE}
+
+#totalSC high and low rec group
+
+#
+Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
+
+                               (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+#p-value = 0.007
+
+```
+
+```{r Q2.short.long.ttest, echo=FALSE}
+
+#long and short 
+#mouse.av.short.biv.table
+#Q2mouse.av.short.biv.table
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
+
+#p-value = 0.9
+
+#mouse.av.long.biv.table
+Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
+#p-value = 0.2
+
+```
+
+**we ran t.tests on the SC based metrics to test is the high and low groups are significanly different**
+
+**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
+`r Q2tot.sc_t.test_High.Low$p.value`).**
+
+**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
+(the summary of mouse averages for this subset of ) **
+
+this might indicate -there's error in the total SC algorithm
+
+- mouse effects
+
+- low power / sampling bias in reduced data set
+
+- total SC is capturing differences in the XY signal which is excluded from the bivalent dataset
+
+glms
+
+```{r Q2.totSC.glm, echo=FALSE}
+
+#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel
+
+Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+
+sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
+
+#sum.Q2.glm_mus.av_totSC_M1.coef <-
+
+#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
+#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
+#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)
+
+Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+
+sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
+#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
+```
+
+(the subspecis effect is tested in M1)
+For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
+
+strain effects from M1; G, LEW, MSM, SKIVE and TOM
+
+strain effects  (basicall all strain effects are significant) (tech effects / error?)
+All strain effects were significant when tested in M2 (Supplemental table X)
+
+For M2 all
+
+```{r Q2.glm.SC.long, echo=FALSE}
+
+#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
+#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
+#table(mouse.av.long.biv.table$mouse)
+#Q2mouse.av.long.biv.table
+
+glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+
+sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
+
+glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+
+sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
+
+#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+```
+
+**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
+The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
+
+**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
+
+```{r Q2.glm.SC.short, echo=FALSE}
+#Q2mouse.av.short.biv.table
+#mouse.av.short.biv.table
+
+glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
+sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
+
+#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
+
+glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
+glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
+
+```
+
+In the short bivalent data set -- analyzed by M1, both subspeces are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+And the strain effec for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).
+
+For the M2 modeanalyzing the mouse means short bivalents, all strains have significant effect but LEW
+G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.
+
+```{r Q2.SC.long.glm.bivglm, echo=FALSE}
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
+
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
+
+#summary(glm_SC.male_Musc_long.biv)
+
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
+
+## logistic regression for Rec group
+#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
+#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
+
+#summary(log.SC.male_long.SC)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+```
+
+**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.**
 
@@
 If there are still strain effects fo these models, that would be evidence that there has been eevolution for SC length compaction across these strains
-
-## Total SC
 
@@
 
-```{r show.tot.sc, echo=FALSE}
-
-#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, mouse) %>%
-#  mutate(file.name = factor(file.name))
-
-mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-mouse.av_sc_male
-
-```
-
 The skeletonized measures show that the general pattern holds for total SC and the high rec strains.
@@
 
-```{r Q2.SC_class.rec.group_setup, echo=FALSE}
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
 The number of cells and mice is difference across the low group strains, 
 -- permut the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)
-
-```{r Q2.SC.Length.stuff, echo=FALSE}
-#all pooled SC lengths
-#WSB     G   LEW  MOLF   KAZ CZECH 
-#    5     9     5     4     3     3
-
-#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14
-
-#G sample 5 mice
-
-#break down of mice in the low rec dataset
-
-#glm.1CO
-Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
-summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
-round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)
-
-#glm.2CO
-Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
-summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
-round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
-#fewer significant differences
-
-Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
-sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
-round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
-#all but KAZ are slightly significant, G and CZECH more so
-
-#glm.3CO
-Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
-#not sig differences
-summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
-round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)
-
-#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)
-
-#all pooled SC lengths
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-
-```
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.long.biv.table <- add_mouse(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-
-mouse.av.short.biv.table <- add_mouse(mouse.av.short.biv.table)
-mouse.av.short.biv.table <- add_strain(mouse.av.short.biv.table)
-mouse.av.short.biv.table <- add_subsp(mouse.av.short.biv.table)
-mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
-mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
-mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
-```
-
-```{r Q2.SC.long.biv_2, echo=FALSE}
-shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2.SC.ploy.male, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-
-#1. determine which strain effects are sig
-#use mouse biv average
-#Male.poly.Mouse.Table_BivData_4MM
-
-glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#SC many strains are different (more than the gwRR)
-#subsps are sig
-
-glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#Dom
-glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glm_SC.male_Dom)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glm_SC.male_Musc)
-
-#Mol
-glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glm_SC.male_Mol)
-
-#Logistic regressions
-
-log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-
-                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
-
-#summary(log.SC.male.poly)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_musc)
-#not sig -- when just looking at Musc
-
-#significant in predicting group
-log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_mol)
-#not sig -- when just looking at Mol
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-glm.Q2.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
-round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
-
-table(mouse.av.long.biv.table$mouse)
-
-glmQ2.M2.SC.male_long.biv <- glm(mean.SC ~ strain, data= mouse.av.long.biv.table)
-round(summary(glmQ2.M2.SC.male_long.biv)$coefficients[,c(1,4)], 6)
-```
-In the cell average Long biv data set, MSM G and CZECH strains have sig strain effects
-
-very few strains have significant values for the mouse
-
-```{r Q2.glm.SC.short, echo=FALSE}
-glmQ2_SC_mouse.av.short_biv <- glm(mean.SC ~ strain, data= mouse.av.short.biv.table)
-
-round(summary(glmQ2_SC_mouse.av.short_biv)$coefficients[,c(1,4)], 6)
-```
-Basically all strains have a significant effect for the mouse.av of long biv
-
-```{r Q2.glm.SC.all.strain, echo=FALSE}
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
-
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
-
-#summary(glm_SC.male_Musc_long.biv)
-
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
-
-## logistic regression for Rec group
-#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
-
-#summary(log.SC.male_long.SC)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-```
 
@@
 
-I think these comparisons should be done with the **Bivalent observations**
-The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class
-
-```{r Q2.SC.Chrm.Class, echo=FALSE}
+<!--
+blocks of other male Q2 - SC lengths
+
+what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
+The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
+
+```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
+#I don't think the Mixed models fit well for the high vs low polymorphism tests.
+
+#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
+#MSM_SC vs MOLF_SC
+
+#2. logistic regression for sc length and chrm class?
+#if the prediction / significance
+
+#isolate data
+#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
+#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
+
+mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
+
+#build and test MM
+lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
+as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
+
+lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
+
+anova(lmer_MM_SC_length_male.poly)
+
+rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
+
+```
+
+-->
+
+> **Summary**
+
+When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
+I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
+(IE the scatter plots of SC length, binned by chrm class on the X).
+
+Each point is a mouse average SC.
+When all bivalents are pooled, the logistic regression SC lengths predict the groups.
+When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
+For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
+
+ and the mean 2COs SC are higher in the high strains
+(this must have something to do with the physical size effect)
+
+If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
+
+Ideas for the above 3 tests,
+
+1. no difference in DOM
+- no sig logistic regression test
+- glm, strain effect?
+
+When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
+When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
+
+> Is the prediction, high rec musc male strains have long SC met?
+
+In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
+I couldn't get the Mixed models working for the male polymorphism predictions...
+
+<!-- junk chunks below -->
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
+#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
 
@@
 
-```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
-#there might not be enough data 
-#
-ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
-```
-
 These are the plots of the logistic regression.
 
-```{r Q2.SC.log.regression.show.plots, echo=FALSE}
+```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
 
@@
 
-```{r Q2.SClength.ttest, echo=FALSE}
+```{r Q2.SClength.ttest, echo=FALSE, include=FALSE}
+#COMPARISONS ACROSS ALL CLASSES
+
 #MOUSE AVE, all
@@
 
-The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths.
-But When the means are compared across chrm class, the mouse averages are no longer significant.
-
-How well does SC length predict chromosome class?
-Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class.
-(Note this test can only be done with bivalent level observations, because 1CO or 2CO)
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-<!--
-blocks of other male Q2 - SC lengths
-
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```{r SC.Length_t.test, echo=FALSE, eval=FALSE}
-
-#comparing all the strains
-
-T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-)#p-value = 0.001567
+```{r log.reg_SC.length, echo=FALSE, include=FALSE}
+#REMOVE THIS COMPARES FROM FULL BIV DATA
+
+#male_mouse.av
+#Male.poly.Mouse.Table_BivData_4MM
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+#male poly shouldn't be here
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
+                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
+                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
+
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
 
 #1CO
-T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_1CO)
 
 #2CO
-T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_mouse.av_2CO)
 
 #3CO
-T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],
-
-  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]
-
-  )
-
-#Single Biv level
-T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )
-
-#1CO
-T.test_SC.Len_single.biv_1CO <- t.test(
-  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],
-
-      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
-  )
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_3CO)
+
+# bivalent level log regression
+
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV)
+
+#1CO  (BIVDATA)
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_1CO)
 
 #2CO
-T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],
-
-                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ]
-  )
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_2CO)
 
 #3CO
-  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],
-
-  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
-  )
-```
-
-The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-
-```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
-#I don't think the Mixed models fit well for the high vs low polymorphism tests.
-
-#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
-#MSM_SC vs MOLF_SC
-
-#2. logistic regression for sc length and chrm class?
-#if the prediction / significance
-
-#isolate data
-#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
-
-#build and test MM
-lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-
-lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
-
-anova(lmer_MM_SC_length_male.poly)
-
-rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
-```
-
-```{r long.sc_log.res_meanSC, eval=FALSE}
-#Long Biv comparison
-
-#hold off on reporting/ working with the long SC biv resutls
-
-long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
-long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]
-
-#divide musc males into groups
-long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
-                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))
-
-long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
-                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))
-
-summary(long.biv.log.reg)#NS effect (likelt underpowered)
-```
-
--->
-
-> **Summary**
-
-When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
-I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
-(IE the scatter plots of SC length, binned by chrm class on the X).
-
-Each point is a mouse average SC.
-When all bivalents are pooled, the logistic regression SC lengths predict the groups.
-When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
-
- and the mean 2COs SC are higher in the high strains
-(this must have something to do with the physical size effect)
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV_3CO)
+```
+
+```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
+#REMOVE THIS, IT uses total bivalent data
+
+#rerun the models for datasets -- deviding by chrm class and rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
+
+Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
+Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
+
+#calq_mouse av
+mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        n_cells = length(unique(fileName) ),
+                        mean.SC.total = mean(chromosomeLength),
+
+                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+                         )
+
+mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
+mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
+
+mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
+mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
+
+#calq_mouse av
+mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
+
+mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
+mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
+```
+
+```{r Q2.SC.Length.stuff, echo=FALSE, include=FALSE}
+#REMOVE THIS, COMPARES FULL BIVDATA ACROSS CHRM CLASS
+
+#all pooled SC lengths
+#WSB     G   LEW  MOLF   KAZ CZECH 
+#    5     9     5     4     3     3
+
+#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14
+
+#G sample 5 mice
+
+#break down of mice in the low rec dataset
+
+#glm.1CO
+Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
+summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
+round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)
+
+#glm.2CO
+Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
+summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
+round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
+#fewer significant differences
+
+Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
+sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
+round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
+#all but KAZ are slightly significant, G and CZECH more so
+
+#glm.3CO
+Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
+#not sig differences
+summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
+round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)
+
+#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)
+
+#all pooled SC lengths
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+Q1lmer.MM_SC_short_results
+```
+
+```{r Q2.SC.ploy.male, echo=FALSE, include=FALSE}
+#REMOVE THIS< COMARES FROM FULL BIVDATA set
+
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+
+#1. determine which strain effects are sig
+#use mouse biv average
+#Male.poly.Mouse.Table_BivData_4MM
+
+glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm1.SC.male.poly) 
+#SC many strains are different (more than the gwRR)
+#subsps are sig
+
+glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+#summary(glm2.SC.male.poly)
+
+#Dom
+glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(glm_SC.male_Dom)  #ALL strain are sig
+#both G and LEW have significant mean.SC, doesn't meet predictions
+
+#Musc
+glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(glm_SC.male_Musc)
+
+#Mol
+glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(glm_SC.male_Mol)
+
+#Logistic regressions
+
+log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+
+                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
+
+#summary(log.SC.male.poly)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_musc)
+#not sig -- when just looking at Musc
+
+#significant in predicting group
+log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC.male.poly_mol)
+#not sig -- when just looking at Mol
+```
 
@@
 
-```{r SC.length.male_tests, echo=FALSE}
+```{r SC.length.male_tests, echo=FALSE, include=FALSE}
 
@@
 
-```{r SC.length.pairwise.T, echo=FALSE}
+```{r SC.length.pairwise.T, echo=FALSE, include=FALSE}
+
+#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
 #male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
@@
 
-If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
-
-Ideas for the above 3 tests,
-
-1. no difference in DOM
-- no sig logistic regression test
-- glm, strain effect?
-
-> Is sex a significant effect for SC length? (as predicted)
-
-- Sex is a significant effect for SC length
-The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)
-
-- According to anova, sex effect explains most of the variance in single bivalent SC lengths.
-
-- The Long Biv Data set largely agrees with the full curated dataset
-
-- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---
-
-```{r log.reg_SC.length, echo=FALSE}
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-The mean SC logistic regression model for mouse averages
-
-`r log.reg_SC.length_mouse.av$coefficients`
-
-and for the single bivalent levels
-
-`r log.reg_SC.length_BIV$coefficients`
-
-When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
-When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
-
-> Is the prediction, high rec musc male strains have long SC met?
-
-In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
-I couldn't get the Mixed models working for the male polymorphism predictions...
-
 # Q2 Normalized CO positions
 
-A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.
-
-or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.
-
-<!-- 
-Main DFs for this section:
-Male.poly.Mouse.Table_BivData_4MM
-male.bivdata.1CO
-
-**Polymorphism**
-C.1) It's unknown if **1CO normalized positions** between high and low musc strains, will differ. (no good predictions come to mind)
-C.2) sis-co-ten metric will be maximized in higher recombining strains .. because.  
-C.3) telomere and centromere distances will be shorter in low rec strains
+In order to test /
+
+We tested another aspect of the rec landscape -- focuing on differences between the high and the low rec strains.
+
+To detect if there are significant differences between the high and the low rec strains we performed t.tests across the two groups
+
+(and logistic regressions?)
+
+<!-- set up chunks here
 
@@
 
-Since there aren't many good predictions for how the 1CO normalized landscape relates to gwRR variation, these will be the main questions
-
-- Has there been evolution in the 1CO pattern?
-
-- Is there a consistent pattern with the high rec strains?
-
-Normalized 1CO foci positions will be used. (F1_PER). Check fixed effects of subsp and strains.
-
-sis-co-ten, telomere and centromere distance of foci are metrics which draw from a wider pool of samples. (So maybe I will use them... but avoiding focus on them)
+```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
+#isolate all musc
+#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
+```
+
+<!-- plots /figures below-->
 
@@
 Q2.shoooo
-
-```
-
-So the plot for the long bivs, shows that the PWD, SKIVE and MSM have on average more centrally placed F1 positions.
-It's kinda werid that the glms/models suggest that WSB has the most terminal F1 -- but, out of all the strains, for the long bivs, WSB has the lowest mean.
-
-But there are v few observations. I think the plot shows that there are fewer 1CO for the High Rec group. not sure there would be enough for feeding into models/
+```
+
+A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.
+
+or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.
+
+<!-- 
+Main DFs for this section:
+Male.poly.Mouse.Table_BivData_4MM
+male.bivdata.1CO
+-->
+
+```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
+
+#mouse.avs_4MM
+#31may19_SKIVE_m2  has NA for norm.first.foci.pos
+
+#High = MSM, PWD, SKIVE
+Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+#p-value = 1
+
+#high = PWD and MSM
+Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
+#p-value = 1
+
+#single biv level
+Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value 
+#p-value = 0.238
+
+```
+
+The t.tests are not significant for normalized foci position comparing all bivalets
+
+The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
+
+```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
+#these don't have mean.first.foci
+
+#short
+Q2.1COpos_short_ttest_HL <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
+            (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+#p-value = 0.06
+
+#remove SKIVE
+Q2.1COpos_short_ttest_HL.ex <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+            (Q2mouse.av.short.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
+#p-value = 0.4
+
+#long
+Q2mouse.av.long.biv.table_HL  <-  t.test( (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
+            (Q2mouse.av.long.biv.table %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+#p-value = 0.5
+
+#long exclude skive  Q2mouse.av.long.biv.table
+Q2mouse.av.long.biv.table_HL.ex  <-  t.test( (Q2mouse.av.long.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+            (Q2mouse.av.long.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
+#p-value = 0.08
+
+#they kinda have ooposite effects
+
+```
+
+The t-tests for the high and low rec strains for the reduced bivalent datasets are mostly NS
+
+mouse mean
+short bivalent
+
+So try using the bivalent sample levels
+
+```{r Q2.1COpos_BIV_ttest, echo=FALSE}
+
+whoo.mama <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1,
+
+            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group == "1"))$PER_Foci_1   )
+
+```
 
@@
 
-```{r Q2.t.test_COlandscape,echo=FALSE}
-
-#MOUSE AVE
-Rec.group.mouse.F1pos_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$norm.first.foci.pos[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.F1pos_Ttest$p.value
-#p-value = 0.9098
-
-#single biv level
-Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-Rec.group_BivData_F1post_Ttest$p.value 
-#p-value = 0.238
-
-```
-
-The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
-
-#### Summary
-
-Not really a clear summary for the normalized 1CO position and rec groups.
+## JUNK Section
 
@@
 
-this is the Q2 prediction for IFD / interference
-
-> prediction A, Is the Male polymorphism Prediction met? High rec strains have shorter IFDs?
+<!-- set up chunks here-->
 

### 2020-02-19 src/BivData_Setup.Rmd
smllchanges
added rec.group col to main bivData DF

@@
 format manual data set so SC.length = Chrmlength (matches the bivData)
-
-add Q1 and Q4 to the cell summary stats table
 
@@
 
+#add rec.group
+#divide musc males into groups
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                  ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                  ifelse(grepl("SKIVE male", Curated_BivData$category), 1,
+                                           0)))
+
+Curated_BivData$Rec.group.exclu <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+
 ```
@@
 
-<!--
 I'm considering removing the same mice from the MLH1 counts, from the BivData, but this might -- limit the comparisons -- might be too stict.
@@
 - the reasons for poor quality of the exclude mice in above list, might be from 2-3 bad bivalents -- which are removed from the biv data during the curation process.
--->
 
@@
 )
+
+#add all the other calculations
+
 ```
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_13.2.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_19.2.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-02-19 doc/Results.rmd
knited a new version

@@
 -We focus on observations from two-crossover bivalents for more comparable observations.
-
 -comparisons of all IFDs of mulit crossover bivalents is also included
-
 Interference is a major determinant of the positioning of chromosomes
@@
 
-```{r show.IFD_99, echo=FALSE, warning=FALSE}
+<!--  set up -->
+
+<!--  plots-->
+
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
 
@@
 
-```{r IFD.old.plots, echo=FALSE, eval=FALSE}
+```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
 #long SC Long_biv_mouse.avs_4MM
@@
 
-Still working on the best way to display the general IFD patterns.
+<!--t.tests -->
+
+```{r Q1.IFD_Ttests, echo=FALSE}
+
+#below are calq just from 2COs, have no fear
+#mean_IFD.2CO_ABS
+#mean_IFD.2CO_PER
+
+#FULL all strains
+Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
+#p-value = 0.07
+
+Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
+#p-value = 1e-12
+
+```
+
+The raw IFD measures for 2CO bivalents are slightly significant (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
+While the normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`)
+
+(interpertation statement)  Males have stronger interference
+
+<!-- lmer/ glms-->
 
@@
 
-## Strain Comparisons
-
 Dive deeper into the sex specific pattern for each strain.
@@
 Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
-
-# Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)
-
-<!-- keep blank for now 
--basic stats for the new dataset
-
--->
 
@@
 whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
-
             (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group == "1"))$PER_Foci_1   )
-
-```
+#p-value = 0.2
+
+whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
+            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
+
+#p-value = 0.4
+
+```
+
+The t.test at the bivalent level is also NS (test; p = )
+
+```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
+
+#compare within subsp
+
+#Musc
+Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+
+# p-value = 0.4
+# exclusive rec group not sig
+```
+
+**Most all of the t.tests I've done indecate that there are no sig diff between the High and the low rec groups**
+
+The normalized 1CO position is not a good metric to distinguish the high and low rec groups.
+
+GLMs
+
+glms are constructed to test if there has been evolution between strains ... ect.
+I will use the full data set the reduced datasets have such low power
+
+```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
+
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
+
+Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+
+sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
+
+#M2
+Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+
+sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
+
+```
+
+for M1
+Mol subsp is significant, strain G (LEW slight sig), strain MSM -- strain MOLF is NA
+
+for M2,
+
+G, LEW, PWD, MSM, MOLF, SKIVE, KAZ have significant effects
+
+put together, these glm's indeceate that MOLF is the most different in it's 1CO landscape (compared to other males). Is this a sampling issue
+
+For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
+
+there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
+(motivation for MOLF female observations and more male MSM biv measures)
 

### 2020-02-19 doc/Results.rmd
notes before knitting again

@@
 
-# Data set ups
+# Data Set Up Notes
 
@@
 
-In order to estimate the genome wide recombination rates across our panel of three subspecies from the house mouse species complex, _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. 
-`r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains were used with reflect samples of natural genetic variation across the terretories of these subspecies.
-
+In order to estimate the genome wide recombination rates across our panel of three subspecies from the house mouse species complex, _M.
+m.
+domesticus,_ _M.
+m.
+musculus_.
+and _M.
+m.
+molossinus_ using `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains.
 MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
@@
 
+<!--
+transition first aragraph to strain specific
+Main points
 1. -female strain means are v close  -male strain means are more varied!!
@@
 2. The largest female difference is G, 1.07X  the largest male difference is PWD and MSM which are 1.3 and 1.2 respectively.  
-
-<!-- female -->
-In order to more finely describe the sex specific patterns, we focused on data sets including single sexes.
+-->
+
+<!-- transition from first part into female -->
 
@@
 
-Within mouse variance was assess to follow up on the results from the full dataset.
+<!--
+The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (glm; p= ).
+ -->
 
@@
 
-The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (p= ).
-
-The overall pattern for the female mean MLH1 counts per cell is low variation. 
-(the most significantly differnt strains in the female data), For _domesticus^G^_ the G strain is `r G.female.diff` higher the the other means.
-While _domesticu^LEW^_  is `r round(LEW.female.diff,4)` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.
-
- mouse mean ranges
-In the _domesticus_ subspecies the ranges of mouse means are 
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )`
-in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
-
--These results are above the minimum of 20 which is the expected minimum following a obligate crossover per chromosomes.
-
 ### male section
-
-While the male pattern shows more variation across the strain means.
 
@@
 
+Intial examinations of the sex specific data revealed that.. there is more variety in across the strain means in meals compared to females.
+
+The overall pattern for the female mean MLH1 counts per cell is low variation. 
+(the most significantly differnt strains in the female data), For _domesticus^G^_ the G strain is `r G.female.diff` higher the the other means.
+While _domesticu^LEW^_  is `r round(LEW.female.diff,4)` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means in females.
+
+while in the male data the strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,
+
+`r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means respectively.
+
+In the _domesticus_ subspecies the ranges of mouse means are 
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )`
+in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
+
+-These results are above the minimum of 20 which is the expected minimum following a obligate crossover per chromosomes.
+
 MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
@@
 - Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
-
-The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,
-
-`r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means respectively.
 
@@
 
+Check the DF of the number cited above
+
 <!--
+
 $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
@@
 
-```{r MLH1co.MM, echo=FALSE, warning = FALSE, digits = 3}
+```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3}
 #THIS IS THE NEW MIXED MODEL
@@
 
-All of the effects from the mixed model significant (Table X) so we followed up with general linear models.
-The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions.
-(*what are the points and differences?
-)*
+All of the effects from the mixed model significant (Table X) 
+All of the effects within the mixed model - came out as significant -- (LRT: subsp p= `r MLH1.MMlmer_results$blah`
+sex p = `r MLH1.MMlmer_results$blah`, subspecies * sex p = `r MLH1.MMlmer_results$blah`).
+(random LTR; strain p= `r MLH1.MMlmer_results$blah`).
 
@@
 ```
+
+Following up with linear models which tested for (subsecies M1), and strain effects alone (M3)
+indecated ... identified several signifcant sex by strain interaction effects, (MSM and PWD) (glm; PWD*male p= and MSM-male p = )
 
@@
 
+<!-- transition gwRR to within mouse variance -->
+
 ## Within Mouse Variance in CO Count per Cell
@@
 Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
+
+RENAME THE MODELS M2 subsp * strain \* sex
 
@@
 $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$-->
+
 ```{r M3.MLH1.var, echo=FALSE}
@@
 
-For both the mixed model and linear model, sex was the only significant effect while the subspecies, while interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively).
+For both the mixed model and linear model, sex was the only significant effect while the subspecies, while interaction and random strain effects were not significant (Table X) (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
 
@@
 (these analysis/ models replicated the results of the full data set (sex being the most significant
-(LTR; p = `r LmerMLH1_M1_Q12_VAR_results$sex.results`, and glm; p = `r  MLH1var_M3_Q12_sum$coefficients[2,4]` )
+(LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` )
 
@@
 For male specific glms for within mouse variance -- had no significant strain effects
-
 the female data models  --- inconsistant and moderate significance p values
@@
 <!--
-
 RW's main qtl paper used the within animal variance,
-"We mapped QTL for
-MLH1 count variance within individuals, applying an approach
-to deconvolve the mean-variance relationship (see
-Materials and Methods)."
-
-The fig5 also has a histogram of within animal variances in MLH1 counts (For F2s, with the parents as reference) (I could do this for males and females)
-QTL on chrm 14, decreased variance in GG genotype, compared to hetZ and WSB (I think this also raised the mean)
 
@@
 references for phenotypic variance connected to robustness to environment, increase in developmental consistency
+
 Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
@@
 
-In an attempt to locatize the above male specific crossoever number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase meiocytes.
-
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of (X for each strain) juvenile mice (12 to 18 days).
-
+explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
+
+In an attempt to locatize the above male specific crossoever number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of (X for each strain) juvenile mice (12 to 18 days) when the proportion of early prophase cells is higher (cite).
 A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _ musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
@@
 
-As can be seen in the figure, the high recombining groups have significantly more DMC1 for for leptotene (t.test; `r ttest.HighLow.L.pval` , one-way-anova; `r round(l.anova.pval, 5)`), but not the zygotene cells (t.test; `r ttest.HighLow.Z.pval`, one-way-anova; `r z.anova.pval`).
+(the differences in mean DMC1 foci across the 2 time points confirms expected results. (t.test for pooled DMC1 foci) )
+
+The high recombining groups have significantly more DMC1 for for leptotene (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r round(l.anova.pval, 5)`), but not the zygotene cells (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
 
 The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio).
-Notably there is not much variation and do not show a significant difference between the high and the low recombining groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+Notably these ratios do not show a significant difference between the high and the low recombining groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
 These results indicate that the best predictor for CO number may be established before DSB formation (cite Baier et al 2014).
@@
 #proportion_stats_category,  tot_biv, ncells
-
 #add number of cells and bivalents in the corner
@@
 
-sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
-sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
+#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
+#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
 ```
@@
 
-**most of the male proportions are 2/3's 1CO and 1/3 2COs. **
+**The vast majority of all bivalents in house mouse have one or two crossoveres (1CO an 2CO), even across sexes**
+
+- fits with low rec rates observed in house mouse.
+
+(rates of 0CO, 3CO and 4CO, these are marginal chromsomes classes)
+
+**most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
+
+**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
+
+**pvales for chi square test, are larger for PSD-MSM compared to low rec strains within subsp, PWD-KAZ and MSM-MOLF**
+
+**these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**
 
@@
 
-**The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
-
 PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
@@
 
-**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
-
 # Single Bivalent Level Results
 
-In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining single chromosome patterns.
+In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining features of the recombiantion landscape at the single bivalent level.
+recombination landscape refers to the i) total length of the SC-AE (potential area for COs), ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.
 
@@
 After the human curation step, `r length(new.Curate.FULL.org$Obj.ID)` single bivalent measures were left.
+(make sure these are the hand.foci bivalents).
+
+how do these agree with other patterns of MLH1
 
@@
 #sort by category,
+
+#compare total boxes and rate passed
+#Curated_BivData  -- pass rate?
 
@@
 Table X illustrates the break down of single bivalent observations by category 
-
 Error compared to human measures and other details have been discribed in (Peterson_2019).
@@
 
-The mouse averages for 3 bivalent level metrics will be analyzed across these questions:  
-1)SC lengths (or total.SC)/chromatin compaction (using all cells regaurdless of CO number) 2) Interfocal Distance of 2CO bivalents and 3) the normalized foci position from 1CO bivalents.
-
 We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
@@
 
-(We will first list the results for the Q1)
-
-These are our predictions for the sex specifici patterns on the single bivalent lanscape
-
- Q1 Analysis, Predictions for Heterochiasmy Q1
-
-Two bivalent level traits are predicted to display heterochiasmy (ie significant effects of sex);
-
-1) SC length will be sexually dimorphic (sex effect will be significant)(cite Lynn)
-
-2. Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick).
-
-3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (deBoer et al 1996, petkov 2001).
-
 The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.
@@
 # Q1 SC Lengths
+
+<!-- breif background on longer female SC  (mouse, human) Gruhn-->
 
@@
 A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
+
+EXPLAIN WHY USING TOTAL SC AREA --
+
+turn out comparing the means of single bivalents is complicated by the fact we can't identify any of the bivalents. -- so we use total sc
 
@@
 
-<!--Brief lit review for SC length and MLH1 expections.
-add XX adjusment / arguments
-
-In order to test one of the major/known correlates with genome wide recombiation rate variation, we compared the average
-
--->
-
 The plots below show the distributions of mouse average total SC area.
@@
 
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+#Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+#Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#CLEAN UP THE FULL DATA SET
 
@@
 The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
-`r Q1_tot.sc_t.test.WSB$p.value`, `r Q1_tot.sc_t.test.G$p.value`, and `r Q1_tot.sc_t.test.LEW$p.value`
+`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific=F )`, and `r Q1_tot.sc_t.test.LEW$p.value`
 for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _ respectively).

### 2020-02-20 doc/Results.rmd
small formating edits

@@
 
-#Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-#Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
 
@@
 The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
-`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific=F )`, and `r Q1_tot.sc_t.test.LEW$p.value`
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _ respectively).
-
-Some of the p values for PWD and MSM are much closer (less significant), ( p = `r Q1_tot.sc_t.test.PWD$p.value`, `r Q1_tot.sc_t.test.KAZ$p.value`, and `r Q1_tot.sc_t.test.MSM$p.value` for the _M.
-m.
-musculus^PWD^ _, _M.
-m.
-musculus^KAZ^ _, and _M.
-m.
-molossinus^MSM^_ respectively)
+`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
+for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).
+
+Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)
 
@@
 
-```{r Q1.lmer_totalSC, echo=FALSE}
+```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE}
 
@@
 
-Q1.lmer.totSC_results
+#Q1.lmer.totSC_results
 
@@
 ```
+
 there are more significant effects for strain and subsp for the mixed model of mouse.av total sc
 
-$$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
+the 3 fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
+sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`)
+
+These effects were followed up the glms
+
+**M2.** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
+
+**M3.** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
 
@@
 
-#G * malev.sig and G*LEW
+#G * male v.sig and male*LEW
 
@@
 ```
+
+The results from those models was that there were 
+...lots of significant effects.  (G male, )
+
+The summary for the total SC area comparison is...
+
+<!-- transition to short-->
+
+The short bivalent dataset was used -- because there are no confounding factors of the XX -- so makes hapier comparisons across male and females
 
@@
 ```
-
-When skive is plotted alone it does have overlapping ranges.
 
@@
 
-(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r Q1_mus.av_shortSC_t.test.G$p.value`, and `r Q1_mus.av_shortSC_t.test.LEW$p.value`
+(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and `r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
 for WSB, G and LEW respectively.
 
-(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value` for PWD, KAZ, SKIVE
+(t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for PWD, KAZ, SKIVE
 
@@
 
-$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1_short_biv_lmer, echo=FALSE}
+**M2.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$
+
+```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE}
+
+Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+
+#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
+
+#skive strin very slight sign
+#male highly sig
+#skive * male slightly sig
+```
+
+skive strin very slight sign, male highly sig,  skive * male slightly sig
+
+**M3.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE}
 lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
@@
 Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-
-```
-
-Sex is the most significant factor for the mixed model of mean short bivalents.
-
-$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.SC.Length_M2, echo=FALSE}
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-#skive strin very slight sign
-#male highly sig
-#skive * male slightly sig
-```
-
-skive strin very slight sign, male highly sig,  skive * male slightly sig
-
-$$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$
+
+#Q1lmer.MM_SC_short_results
+
+```
+
+Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+interaction effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`)
+
+**M3.** $$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$
 
@@
 
+In summary .... females (oocytes) have different chromatin compaction patterns compared to males.
+
 ```{r show.long.biv, echo=FALSE, include=FALSE}
@@
 
-we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
+We focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
 We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ).
@@
 
-```{r CO.positions.figs, echo=FALSE, include=FALSE}
+```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
 #main figure of density plots,

### 2020-02-20 doc/Results.rmd
more slow progress

@@
 
-In order to estimate the genome wide recombination rates across our panel of three subspecies from the house mouse species complex, _M.
+We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from thress subspecies within the house mouse species complex;  _M.
 m.
@@
 m.
-molossinus_ using `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains.
+molossinus_.
 MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
@@
 <!--
+more basic stats, error, comparison to other measures
+
 transition first aragraph to strain specific
@@
 
-## female section
+<!-- female section -->
 
@@
 
+female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
 #female strain table
@@
 
+#F.strain.table
+
 # I don't know why most of the table is blank
@@
 
-lm.female.strain <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-lm.female.strain_sum <- summary(lm.female.strain)
+MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
 
@@
 #write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-
 ```
@@
 lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lmVAR.female.strain)
+lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
 
@@
 
-### male section
+<!-- male section -->
 
@@
 # given a lm, how sig is the strain effect
-glm.male.strain.subsp <- lm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-summry.male.strain.effects <- summary(glm.male.strain.subsp)
-round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
+glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+
+#summry.male.strain.effects <- summary(glm.male.strain.subsp)
+#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
 
@@
 #M2
-lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
-lm.male.strain_sum<- summary(lm.male.strain)
+MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
+sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
 
@@
 ```
-
-the coefficients for glms of the CO means are listed above.
 
@@
 
-Intial examinations of the sex specific data revealed that.. there is more variety in across the strain means in meals compared to females.
-
-The overall pattern for the female mean MLH1 counts per cell is low variation. 
-(the most significantly differnt strains in the female data), For _domesticus^G^_ the G strain is `r G.female.diff` higher the the other means.
-While _domesticu^LEW^_  is `r round(LEW.female.diff,4)` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means in females.
-
-while in the male data the strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,
-
-`r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means respectively.
-
+Intial examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.
+In females, _domesticus^G^_, _musculus^MSM^_, and _domesticu^LEW^_ had the greatest differences in means compared to other strains, `r G.female.diff`, `r round(LEW.female.diff,4)`, and
+`r MSM.female.diff` respectively.
+While in the male data the strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`, `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.
+
+<notabley these strains come from Musc and Mol -- Mol natural hybrid of musculus and castaneus. this comments out text?>
+
+The musculus and molossinus subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
+
+<!-- dom, low variation and close to the expected in
+within the male means, there is a low degree of strain varaince in _Domesticus_ subspecies with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
+-->
+
+<!-- ranges of mouse means
 In the _domesticus_ subspecies the ranges of mouse means are 
@@
 The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
-
--These results are above the minimum of 20 which is the expected minimum following a obligate crossover per chromosomes.
-
-MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
-
-Two glm models were run to test the effects of strain and subspecies on mean MLH1 count per cell per mouse. The coefficients for the following model are below.
+-->
 
@@
 
-- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
-
-- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
-
-**general linear models** -- indicate the 3 strains have significant strains effects
-
-Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.
-
-- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are
-
---notes on error / 
--- connections to previous measures in the literature
-
-The same model framework was applied to the male specific data.
-The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.
 <!--
@@
 
-# Evolutionary framework
+# Evolutionary Framework
+
+<to better understand this varation in RR within an evolutionary framework>
 
 In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
+
+<!--
 (*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
@@
 Check the DF of the number cited above
-
-<!--
-
-$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
 This is the break down of mice used within the Mixed model
-
 `r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
@@
 
-```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3}
+**M1.Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+Our model, M1., was designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
+
+```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3, message=FALSE}
 #THIS IS THE NEW MIXED MODEL
@@
 
-<!--
- $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-and 
-$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$  -->
-
-```{r MLH1.glmM2, echo=FALSE,  digits = 3}
+All of the effects within the mixed model were significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
+
+```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
 #strain = fixed
-glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-
-summ <- summary(glm.strain.fixed.full)
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
+
 #table of just coeff anf pvalue
@@
 
-Our models were designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
-
-All of the effects from the mixed model significant (Table X) 
-All of the effects within the mixed model - came out as significant -- (LRT: subsp p= `r MLH1.MMlmer_results$blah`
-sex p = `r MLH1.MMlmer_results$blah`, subspecies * sex p = `r MLH1.MMlmer_results$blah`).
-(random LTR; strain p= `r MLH1.MMlmer_results$blah`).
-
 ```{r MLH1.glm.M3, echo=FALSE, digits = 3}
 
-strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-summ.strain.fixed <- summary(strain.fixed)
+MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
 
@@
 
-Following up with linear models which tested for (subsecies M1), and strain effects alone (M3)
-indecated ... identified several signifcant sex by strain interaction effects, (MSM and PWD) (glm; PWD*male p= and MSM-male p = )
-
-In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
+ **M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
+
+ **M3. LInear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+
+Following up with linear models which tested for (subsecies M2), and strain effects alone (M3)
+both showed similar -- results -- **so we just report those from M3. also just report effects which were consisantly significant across both M2 and M3**
+
+<they also have sig strain effects .. PWD strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`>
+
+which identified several signifcant sex by strain interaction effects, ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
+and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`
+)
+
+<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
 These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.
-Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
+
+<M3 also has a few less significant effects>
+<not the reverse coefficient indexing -- this will give the name >
+
+Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
 These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
 
+_domesticus^G^_ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
+and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`
+
+<!--
+also g strain effect/ dom^G^ strain effect 
+`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[12,][4]` 
+In both models strain by sex interaction effects were significant for (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
+-->
+<!-- glms for sex specific data also run -->
+
+A similar glm models (M4) was run on the sex specific data set, (just containing subspecies, strain and the interaction).
+
+ **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$
+
+For the male specific data, three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
+(with effect sizes ranging from 5, 7, and 2 foci respectively).
+Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
+
+While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
+
 <!-- transition gwRR to within mouse variance -->
@@
 
-RENAME THE MODELS M2 subsp * strain \* sex
+Following the same basic framework we used for analyzing mean number of MLH1 foci per cell, we applied a set of mixed and linear models -- to test effects (sex differences for the within mouse variance in MLH1 foci per cell).
+
+ **M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+ **M6. Linear Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex * strain + \varepsilon$$
+
+ **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
 
@@
 ```
-<!--
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$-->
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M3)
+anova(MLH1var_M3)  
+as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
+
+MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+#summary(MLH1cv_M3)
+
+anova(MLH1cv_M3)
+as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
+
+#Q12 var
+MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1var_M3_Q12)
+anova(MLH1var_M3_Q12)  
+as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M3_Q12)
+anova(MLH1cv_M3_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
+
+```
 
@@
 
-For both the mixed model and linear model, sex was the only significant effect while the subspecies, while interaction and random strain effects were not significant (Table X) (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
-
+For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
 Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
 background noise).
-For this reasons, we replicated the model analysis from data sets in the subset of cells with higher quality scores compared to the full data set.
-(these analysis/ models replicated the results of the full data set (sex being the most significant
-(LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` )
-
-<!--*notes on caveats for calculating variance -- lower power-->
+For this reasons, we replicated the model analysis using a subset of cells with higher quality scores.
+These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).
+Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
+While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`) for the model with the full dataset.
+While in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
 For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell.
-This limits the power for detecting significance differences.
-
-**female strongest effect**, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.
-
-in the sex specific data was analyzed in glm's to test for significance differences in the magnitude of within mouse variance across strains.
-The male specific models, showed no significant strain effects
-
-For male specific glms for within mouse variance -- had no significant strain effects
-the female data models  --- inconsistant and moderate significance p values
+This limits the power for detecting significance differences across groups, <but the main/general pattern, is that females have significantly higher within mouse variance compared to males.
+This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.
 
@@
 limitations: observations at the mouse level, decreased power to estiamate the variance --
-setup
--quality / technical error
-- variance and cv
-The mixed models have most significant variables, 
-M3.glm for variance was less significant (0.044181 ), 
-the Q12 models in general were more significant
-cV's were more significant than var
-
-var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-
-cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
 
@@
 
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
-
-#Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
-```
-
-For mean CO count, _domesticus^G^_  strain effects in both models (p = )
-_musculus^MSM^_ was a significant strain effect on female mean MLH1 counts per cell for only the second model (p= ).
-
-Above are the coefficents for the two glm's of the female specific mouse MLH1 averages, which include subspecies and strain as fixed effects.
-G has the most consistant significant strain effects in both models.
-MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.
-
-MOVE THESE MODEL RESULTS WITH THE OTHER MODEL RESULTS
-
-Two models were fitted to the male specific data to test the effects of strain and subspecies.
-
-# Variation in DSB number per cell
+< SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE >
+
+# Variation in DSB Number Per Cell
 
@@
 
+<!--
 explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
-
-In an attempt to locatize the above male specific crossoever number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes of (X for each strain) juvenile mice (12 to 18 days) when the proportion of early prophase cells is higher (cite).
-A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _ musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
+-->
+
+In an attempt to locatize the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days).
+A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
 One mouse was sampled for each strain.
 
-(the differences in mean DMC1 foci across the 2 time points confirms expected results. (t.test for pooled DMC1 foci) )
-
-The high recombining groups have significantly more DMC1 for for leptotene (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r round(l.anova.pval, 5)`), but not the zygotene cells (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
-
-The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio).
-Notably these ratios do not show a significant difference between the high and the low recombining groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
-These results indicate that the best predictor for CO number may be established before DSB formation (cite Baier et al 2014).
+For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = ), which reflects the expected patterns across the timing of meiosis.
+
+Confirming the prediction that (more crossovers are associated with more initial DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
+
+The differences in DMC1 foci were not significant for zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
+<!-- which maybe be attributed to noise effects
+
+The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->
+
+The ratios of average number of MLH1 and DMC1 foci per cell -- have  .... biological significance
+
+The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+
+These results indicate that the changes -- which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
 
@@
 
-chrm.class.barplot
+#chrm.class.barplot
 ```
@@
 #chrm.class.barplot
-
+chrm.class.barplot
 ```
@@
 
-In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes grouped by the number crossovers on the same bivalent.
+In order to decompose the cell wide rate, we decided to look at the proportion of chromosome classes (ie by the number crossovers on the same bivalent).
 A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).
 
-**The vast majority of all bivalents in house mouse have one or two crossoveres (1CO an 2CO), even across sexes**
-
-- fits with low rec rates observed in house mouse.
-
-(rates of 0CO, 3CO and 4CO, these are marginal chromsomes classes)
-
-**most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
-
-**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
-
-**pvales for chi square test, are larger for PSD-MSM compared to low rec strains within subsp, PWD-KAZ and MSM-MOLF**
-
-**these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**
-
-Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
-They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
-This significant p value indicates that at least one of the six strains had significant differences).
-
-PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
-
-While the proportions are significantly differennt, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
- and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+The first attern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes.
+<these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>
+
+< note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?) >
 
@@
 
-- High female strains, G and Lew have significantly more 2CO bivalents -- there are some differences in the female strains.
-
-- The overall male pattern that's most striking is the gradient of 2CO proportions
+**2. most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
+
+**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
+
+PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
+
+While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
+ and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+
+**3.
+these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**
+
+<Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
+They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
+This significant p value indicates that at least one of the six strains had significant differences).>
+
+<SUMMARY / TRANSITION to single bivalen measures>
 
@@
 
-**M2.** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
-
-**M3.** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
+**M2.Linear Model** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
+
+**M3. Linear Model** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
 

### 2020-02-21 doc/Results.rmd
wacky idea about nrm IFDs

@@
     toc_float: yes
-  html_notebook: default
   pdf_document: default
@@
 
-< note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?) >
+<note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?)>
 
@@
 
-<SUMMARY / TRANSITION to single bivalen measures>
+Breif conclusion: the evolution of genome wide recombiantion rates proceeds through the chromosome level.
+
+<something about the female results>
+
+<SUMMARY the / TRANSITION to single bivalen measures>
 
@@
 From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(new.Curate.FULL.org$Obj.ID)` single bivalent measures were left.
+After the human curation step, `r length(Curated_BivData$Obj.ID)` single bivalent measures were left.
 (make sure these are the hand.foci bivalents).
 
-how do these agree with other patterns of MLH1
+<how do these agree with other patterns of MLH1>
 
@@
 #sort by category, 
-
 #compare total boxes and rate passed
@@
 #mean number of isolate bivs per cell,
+#use these two metrics
+#cell.level.BivData_table$boxs.IDd 
+#cell.level.BivData_table$nboxes_passed
+
+#calq the mean rate by category
+
+cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
+
+algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
+                        n_obs = length( unique(fileName) ),
+                         mean_pass.rate = mean(pass.rate)
+                         )
+
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
+
+algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
+                        n_obs = length(unique(fileName)),
+                        mean_pass.rate = mean(pass.rate)
+                         )
 
@@
 
-Table X illustrates the break down of single bivalent observations by category 
+Table X illustrates the break down of single bivalent observations by category. 
 Error compared to human measures and other details have been discribed in (Peterson_2019).
 
-While the automated software doesn't isolate all bivalents/chromosomes from each cell (on average 17), we assume that the isolation process is not biased.
+The ranges of of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in MSM male to 
+`r max(algorithm.stat.category$mean_pass.rate)` ) KAZ female, we assume that the isolation process is not biased.
 Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.
+We assume the bivalent datasets across all categories are equally eqivilent for representing general patterns.
+
+<curation, x from curation, XY excluded, replication to hand measures>
 
@@
 
-These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.
-
-The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.
-
-# Q1 SC Lengths
+<These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.>
+<The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.>
+
+# Q1 Sex Differences in SC-AE Lengths
 
@@
 
-EXPLAIN WHY USING TOTAL SC AREA --
-
-turn out comparing the means of single bivalents is complicated by the fact we can't identify any of the bivalents. -- so we use total sc
+Comparing the means of single bivalents is complicated by the fact we can't identify the chromosome identitied the physical size (in terms of SC-AE) is the closet approximation we can make.
 
@@
                         n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
                          mean_SC = mean(chromosomeLength)
@@
 Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)
-
-```
-
-The plots below show the distributions of mouse average total SC area.
+```
+
+For this reason we use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells).
+A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes.
+
+Even though this smaller data set has decreased power due to redcued sample size, ...
+
+This 'short biv' data set has advantages in that it is closer to sampling across a similar set of chromosomes (ideally Chromosomes 15 to 19) and it excludes the XX in females, which is predicted to be the 3rd longest based on the physical (Mb) length.
 
@@
 
-```{r Q1.tot.SC_show, echo=FALSE}
+```{r Q1.SC.show.short.biv, echo=FALSE}
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+
+Q1.SC.short.mouse_scatter_boxplot
+
+#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
+skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+#skiveQ1.SC.short.mouse_scatter_boxplot
+#there are 2 lowish female means,
+
+```
+
+```{r Q1_SC.biv.short_ttest, echo=FALSE}
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+#WSB
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
+#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
+#1 male, 3 females measures
+#try to merge the manual - whole cell observations (at least 2 males)
+
+```
+
+<in the above boxplots of mouse means for shortbivalents>
+
+For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer, the largest difference across sexes by strain are>  (t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and `r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
+for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+and (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
+
+<EXPLAIN WHY USING TOTAL SC AREA>
+
+An additional measure of the SC-AE we tested is the total SC area per cell (the whole genome).**to get a comparison of the SC-AE area for the whole genome (not just a subset of chromosomes)** We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
+A total of X images were run trhought the image analysis pipeline, and after outliers removed, X remained.
+/ cell based measures were left.
+Mouse means were calculated for their total SC-AE area and are displayed in figure X with the strain means shown as horizontal lines (below).
+
+```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
 
@@
 
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
+
+Tot.SC_mouse.av_Musc
+
+totsc.mol <- Q1.totSC.mouse.av.table %>% filter( subsp == "Mol" )
+totsc_strain.av.mol <- totSC.strain.av.table %>% filter( subsp == "Mol" )
+
+Tot.SC_mouse.av_Mol <- ggplot(data= totsc.mol, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totsc_strain.av.mol, aes(yintercept=mean.skel) )+
+
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
 
-Tot.SC_mouse.av_Musc
-```
-
-these are the number of mice for each of the categories
-
-`r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )`
+Tot.SC_mouse.av_Mol
+```
+
+<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )
 
@@
 
-The full data set of SC lengths illustrates the general pattern than females have longer SC-AE.
-
-The most averages for total SC (skeletonizes) are significantly longer in females for all strains (p=
+```{r Q1.totSC_male.first.comp, echo=FALSE}
+
+Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
+
+Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
+```
+
+The most averages for total SC-AE area (skeletonizes) are significantly longer in females for all strains (p=
 `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
@@
 
-$$mouse \ average \ total \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE}
+The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
+(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
+<But we use this metric anyways>
+
+<We estimate the effect of the sex chromosomes from hand measured data...the XX (3rd longest chromsomes) in females is ~10% of total the summed SC area, the average propotion of the XY in spermatocytes is XX%?
+however the morphology of the XY is highly variable across cells>
+
+The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>
+-- we follow up with models (which focus on sex differences).
+
+<explain why applying models.  We then use models to examine/ think about the evolution>
+
+The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mean mouse total sc and mouse mean short bivalent.
+
+**M3. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-there are more significant effects for strain and subsp for the mixed model of mouse.av total sc
-
-the 3 fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
-sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`)
-
-These effects were followed up the glms
-
-**M2.Linear Model** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
-
-**M3. Linear Model** $$mouse \ average \ total SC  \ area ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.glm_totalSC, echo=FALSE, }
-#total-SC
-
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-#summary(Q1.glm_totSC_M2)
-#subsp effect, many strain, #male
-
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-#summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-
-```
-
-The results from those models was that there were 
-...lots of significant effects.  (G male, )
-
-The summary for the total SC area comparison is...
-
-<!-- transition to short-->
-
-The short bivalent dataset was used -- because there are no confounding factors of the XX -- so makes hapier comparisons across male and females
-
-```{r show.short.biv, echo=FALSE}
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
-Q1.SC.short.mouse_scatter_boxplot
-
-#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
-skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-#skiveQ1.SC.short.mouse_scatter_boxplot
-#there are 2 lowish female means,
-
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
-#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
-#1 male, 3 females measures
-#try to merge the manual - whole cell observations (at least 2 males)
-
-```
-
-The pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data
-
-with the mouse average for 'short' bivalents
-
-(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and `r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
-for WSB, G and LEW respectively.
-
-(t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for PWD, KAZ, SKIVE
-
-for the short bivalents are not significantly differen, for this strain there are 4 female and 4 male
-
-** there are more SKIVE female mouse observations -- but they are filtered out because I excluded them from the MLH1 dataset.  -- for low quality -- they might have good total SC spreads -- // but
-
-* it think total sc only has 1 female observation but the short bivalent has 4 skive females
-
-<!--
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
--->
-
-Mixed models for the sex differences in mouse averages for short bivalents (reduced dataset)
-
-**M2.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE}
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-#skive strin very slight sign
-#male highly sig
-#skive * male slightly sig
-```
-
-skive strin very slight sign, male highly sig,  skive * male slightly sig
-
-**M3.** $$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE}
+For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed model of mouse.av total sc.
+
+The three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
+sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
+
+```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
 lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
@@
 
-Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
-interaction effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`)
-
-**M3.** $$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.glm_short.SC_M3, echo=FALSE}
+<and the short biv>
+Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
+
+These effects were followed up the glms, again applied to both mouse means for short biv SC lengths and total sc area per cell.
+
+**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.glm_totalSC, echo=FALSE, message=FALSE}
+#total-SC
+#Q1.lmer_totSC.mouse.av.table
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
+
+#subsp effect, many strain, #male
+#G * male v.sig and male*LEW
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
+#LOT OF SIGNIFICant strain effects in this model
+```
+
+The following effects had significant p values given the above model
+
+subspecies musc p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`, subspecies mol p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`, sex male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`, strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
+<this is the smallest p vale, elevated female total, driven by the high means in females>
+
+(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
+
+These results indicate that there are strain specific effects for the SC-AE area --
+
+the large G strain effect is likely driven by the high female mean total SC area
+
+sex is the next strongest (most significant) effect.
+
+<!--
+for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
+-->
+
+For the short bivalent data -- run through the same models...
+
+```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE, message=FALSE}
+
+Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+
+#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
+
+```
+
+```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
 
@@
 
-summary(Q1.glm_short.SC_M3)
+#summary(Q1.glm_short.SC_M3)
 
@@
 
-round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
 #sex male very sig -- only sig
@@
 
-male is the only significant p value from this model.
-
-<!--
-Because of confonding effects across chromsomes of different physical lenghts, we attempt to ...
-reduce the error from mixing chromosomes by constructing reduced bivalent sets, for the longest 5 and shortest 5 bivalents from the same cell.
-The mouse means for these two data sets were then calculated.
--->
-
-Above are the p values for the components of the mixed models for the short and long biv mouse means. In both sets of models the sex effect is the most significant p value.
-
-Above are the mixed model results seperated by mouse averages for 1 crossover and 2 crossover bivalents respectively.
-
-The fixed sex effect is highly significant for the mouse means of SC length. The data sets tested were all pooled SC lengths, the long bivalents, 1CO and 2CO seperately.
-
-In summary .... females (oocytes) have different chromatin compaction patterns compared to males.
+From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][10]`).
+while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.
+
+In the M2 model, the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][10]`).
+<should apply caution to the short bivalent dataset> The significant SKIVE effect is driven by  2 of the 4 female skive mice -- having much shorter bivalents comapred to the other female musc strain means.
+
+<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-AE.  >
+
+In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.
 
@@
 
-# Q1. General CO Positions
+# short transition
+
+<transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents
+1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
+-- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).
+
+2.
+We focus on two metrics/ aspects i) the placement of single foci along a bivalent (just 1CO) and ii) the placement of two foci on the same bivalent / the interfocal distance of 2CO bivalents (reflects crossover interference)
+
+# Q1. 1CO Position Sex Differences
 
@@
 
-We focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
-We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ).
-(all chromosomes in house mouse are acrocentric)
-
-Chromosome size effects are a confonding factor for CO position, with shorter chromosomes having a more uniform landscape compared to larger chromosomes (cite).
-To account for this we again use the reduct bivalent sets (long.biv and short.biv)
-
-Below are the general patterns / plots -- short and long bivalent groups are shown to try and account for the chromsomes size effect.
+We focus on the foci positions from one crossover (1CO n= 
+) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
+
+We anchor the position measures at the centromere (and normalize by deviding the position by the length of the chromosome.
+The foci positions then have a range of 0-1, with smaller values representing positions closer to the centromere end of the bivalent, and larger values meaning foci positions closer to the opposite end (refered here as telomeric position).
+
+We plot the mouse averages across all of the single bivalent measures, with the strain averages shown as boxplots...
+
+in order to get a better picture of the sex differences in this foci-position metric
 
@@
 
-```{r Q1show.F1pos_Full, echo=FALSE}
+```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
 
@@
 are highly significantly more medially placed (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
-With the musclus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
+With the musculus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
 Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).
 
-<!-- lmer and glm here -->
+After confirming the significan sex differences for 1CO position, we apply the familiar models
+
+**M1. Mixed Model**
 
@@
 
-```{r Q1.CO1pos_lmer.M1, echo=FALSE}
+```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
 
@@
 
-lmer.Q1.nrm1CO_M1_results
+#lmer.Q1.nrm1CO_M1_results
 
@@
 ```
-For the Mixed model, sex and strain are significant effects for the normalized single foci position.
+
+For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
+(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
+These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).
+
+**M2. Linear model**
 
@@
 Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-summary(Q1.glm_F1pos_M2)
+
+#summary(Q1.glm_F1pos_M2)
 
 summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
+#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
 
@@
 
+**M3. Linear model**
+
 $$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
 
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=FALSE}
-
-summry.Q1.glm_F1pos_M2
+```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
 
 Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-summary(Q1.glm_1CO.pos_M3)
+
+#summary(Q1.glm_1CO.pos_M3)
 
 summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-
+#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
 #write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-```
-
-Lew and KAZ had slightly significant strain speicific effects across the two glms (these results will be explored more fully in the Q2 section).
+
+```
+
+The male effects were the most significant (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+While the musc subspecies, LEW and KAZ strains had slightly signifcant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
+
+(these results will be explored more fully in the Q2 section).
 The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).
 
-<!--
--full 1CO tests
-
-long 1CO tests
-
-short 1CO tests
-
-There are a few traits that fall within the CO positions  
-- 1CO normalized position  
-- sis.co.ten (sorta interference)  
-- centromere and telomere distance  
-**Heterochiasmy**
-
-2.A) Normalized 1CO positions will be sexually dimorphic (sedell and Kirkpatrick). 
-  B) **sis**ter **co**hesion **ten**sion (sis-co-ten) will be sexually dimorphic as it reflects the general property of unifrom vs telomere/biased CO positioning.
-  C) Centromere and telomere distances will be sexually dimorphic.
-**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.
--->
+<we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>
+
+Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite).
+To account for this we again use the reduced bivalent sets (long.biv and short.biv)
 
 ```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-
 Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nShort Biv")
-Q1.short.biv_scatr_box
 
@@
 \nNormalized 1CO Foci pos \nLong Biv")
-Q1.long.biv_scatr_box
+
 ```
@@
 \nNormalized 1CO Foci pos \nLong Biv")
-Q1.av.long.biv_scatr_box
 
@@
 \nNormalized 1CO Foci pos \nShort Biv")
-Q1.av.short.biv_scatr_box
-
-```
+
+```
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+<!--
+### Siscoten
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males.
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
+
+> Is the heterochiasmy prediction met?
+
+Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
+
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
+t.test_mouse.av_siscoten <- t.test(
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+
+drop1(lme_MM_telo.dist, test="Chisq")
+
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+drop1(lmer_MM_telo.dist, test="Chisq") 
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.-->
 
@@
 
-Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs.
-This plot shows the sexual dimorphism in the density plots.
-
-Consider adding annotate_text for the number of observations in each category.
-think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.
-
 ```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
@@
 
-<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-the distribution of SC lengths and sis-coten seems very different across sexes
-
-The mixed model data should only come from 1CO bivalent data.
-
-the mouse average foci1 pos is more significant in t.test, but not log regression...
-(is something wrong?) Check the mouse averages for the F1_pos, there might be an outlier or mouse with v.few observations.
-
 ```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
@@
 
-<!--
-### Siscoten
-
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-#2CO
-
-#3CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
-drop1(lme_MM_telo.dist, test="Chisq")
-
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.
-
--->
-
-#Q1. IFD
-
-In order to test if there are sex differences between crossover interfernce, a major determinant of the positioning of crossovers along chromosomes, we examined ...
+#Q1. Sex Differences in CO Interference (IFD)
+
+In order to test if there are sex differences between crossover interference, a major determinant of the positioning of crossovers along chromosomes, we examined ...
 ran models with the interfocal distance (IFD) of two foci on the same bivalent.
@@
 However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes
-at the pachytene stage of meiosis."
+We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes at the pachytene stage of meiosis."
 
@@
 first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
+first.IFD.plot_ABS
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+first.IFD.plot_PER
 #+theme(legend.position="none")
@@
 
+<!-- EXPLAIN WHY WE SHOULD CARE ABOUT RAW AND PER IFD values-->
+
+The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical - force mediated through the SC -- In some models this metric should be thought of as the best metric for charaecterizing crossover interference -- bc it reflect the physcial --
+
+The above plots display the raw and normalized area which seperate two foci on the same chromosome.
+In the normalized plot, 9 observations fall above the length threshold of 150.
+The female measures fall more uniformly across the range of (0 to 1).
+
+In the abosolute plots, you can start to see clustering/ enrichment of IFD values between 40 and 80 pixels for PWD and SKIVE, 
+(and maybe slight clustering for MSM).
+
 <!--t.tests -->
 
+t.tests are performed on the normalized metrics for establish if there are sex differences in the IFD meterics (raw and normalized).
+
 ```{r Q1.IFD_Ttests, echo=FALSE}
-
 #below are calq just from 2COs, have no fear
@@
 Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-
                                 ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
@@
 
-```
-
-The raw IFD measures for 2CO bivalents are slightly significant (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
-While the normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`)
-
-(interpertation statement)  Males have stronger interference
+#WSB
+Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
+                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
+
+#LEW
+Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
+
+#G
+Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
+
+#PWD
+Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
+
+#KAZ
+Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
+
+#SKIVE
+Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+
+#MSM
+Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
+
+#FULL all strains
+Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
+#p-value = 0.3
+
+```
+
+The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
+BUT WHEN G IS REMOVE, the pvalue is no longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
+<look closer at G>
+
+The normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`) between sexes, indicating there a
+
+When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
+MSM
+SKIVE
+PWD
+MSM
+LEW
+G
+WSB
+
+This indicates a general pattern of two foci are seperated by more area (hance have stronger interference) in males, this pattern is also for genetic maps.
+
+<!-- short and long -->
+
+```{r Q1.IFD.long.short.setup, echo=FALSE}
+
+#real.long.bivData
+#real.short.bivData
+
+Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.long <- add_sex(Q1.IFD.long)
+Q1.IFD.long <- add_strain(Q1.IFD.long)
+Q1.IFD.long <- add_subsp(Q1.IFD.long)
+
+Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
+Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
+
+Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.short <- add_sex(Q1.IFD.short)
+Q1.IFD.short <- add_strain(Q1.IFD.short)
+Q1.IFD.short <- add_subsp(Q1.IFD.short)
+
+Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
+Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
+
+#t.test. mouse av
+Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+
+Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+#both the short and long norm IFD are NS
+
+#t.test Biv
+Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+
+Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+
+#THIS BIV OBS are also SIG
+
+nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
+#looked up image, measures confirmed
+
+####
+nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#21may18_KAZ_m2 -- seems like it might be an outlier
+#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
+#second foci slightly smaller, kinda fainter -- but also much brighter than background
+
+#check all KAZ males
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+###look at why KAZ males are weird
+
+nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ.female_BIV.data
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ_BIV.data
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#the difference the the Nrm IFD cut off is striking with these plots
+
+nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD_BIV.data
+#this might
+
+nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_MSM_BIV.data
+
+nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_SKIVE_BIV.data
+
+nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G_BIV.data
+```
+
+The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes?
+they were supposed to be 1CO -- but somehow a second CO formed anyways ..
+and very close by
+
+Where as in PWD, SKIVE and MSM there ARE NO 2CO bivalents which have such short IFD
+WHAT espicially werid about this --
+
+whats the expectations following the confonding proportions? 
+high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.
+
+to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
+Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)
+
+-- run these comparisons to see if the small NRM IFD are due to shorter chromosomes, which are predicted to have weaker interference.
+
+Looked at the KAZ pattern
+
+Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
+These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).
+
+Considering that the short-long bivData is much smaller than the full data set
 
@@
 
-$$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.IFD.M1, echo=FALSE}
+**M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.IFD.M1, echo=FALSE, message=FALSE}
 #get a list of the old mice to double check
@@
 
-$$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
+**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
 
@@
 
+**M3. Linear model**
+
 $$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
@@
 summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
+#round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
 
@@
 For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
-The other effects are for PWD ansD SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
+The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
 For the M3, only SKIVE*male effect is close to significant
@@
 For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
-SKIVE*male is the only other consistantly significant effect, which also increases the nrm.IFD measure.
+SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure.
 
@@
 
-I made a sereies of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromsomes (possitive correlation of sc length and CO number).
-Yet, given the complications from the chromsomes proportions outlined above, comparing the bivalents across chromosome classes is complecated
+I made a series of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromosome (positive correlation of SC length and CO number).
+Yet, given the complications from the chromosome proportions outlined above, comparing the bivalents across chromosome classes is complicated
 
@@
 
-In order to test if sc lengths (sc area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
-
-The average sc lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confonded by the different proportions across strains.
+In order to test if SC lengths (SC area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
+
+The average SC lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confounded by the different proportions across strains.
 
@@
 total SC for males (which ar)
-X additional cells were added to the male specifi total SC data set discirbed above
-
--reduced bivalent dataset (long and short) is made up of bined bivalents from X cells and X mice
+X additional cells were added to the male specific total SC data set described above
+
+-reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice
 
@@
 
-- total SC is capturing differences in the XY signal which is excluded from the bivalent dataset
+- total SC is capturing differences in the XY signal which is excluded from the bivalent data set
 
@@
 
-(the subspecis effect is tested in M1)
+(the subspecies effect is tested in M1)
 For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
@@
 
-strain effects  (basicall all strain effects are significant) (tech effects / error?)
+strain effects  (basically all strain effects are significant) (tech effects / error?)
 All strain effects were significant when tested in M2 (Supplemental table X)
@@
 
-In the short bivalent data set -- analyzed by M1, both subspeces are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-And the strain effec for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).
-
-For the M2 modeanalyzing the mouse means short bivalents, all strains have significant effect but LEW
+In the short bivalent data set -- analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).
+
+For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW
 G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.
@@
 
-We purpose -- we think that running the models for comparing SC means by chromosome class and devided by rec group.
-(since the chromsome porportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )
+We purpose -- we think that running the models for comparing SC means by chromosome class and divided by rec group.
+(since the chromosome proportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )
 
 The null hypothesis is that these won't be highly significant strain effects.
-If there are still strain effects fo these models, that would be evidence that there has been eevolution for SC length compaction across these strains
+If there are still strain effects for these models, that would be evidence that there has been evolution for SC length compaction across these strains
 
@@
 
--the SKIVE are within dom and kaz level -- so why does SKIVE have inntermediate CO proportions and gwRR
+-the SKIVE are within dom and kaz level -- so why does SKIVE have intermediate CO proportions and gwRR
 
@@
 The number of cells and mice is difference across the low group strains, 
--- permut the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)
-
-abova are glm tables for mouse.av SC lengths for each strain.
+-- permute the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)
+
+above are glm tables for mouse.av SC lengths for each strain.
 
@@
 
-We tested another aspect of the rec landscape -- focuing on differences between the high and the low rec strains.
+We tested another aspect of the rec landscape -- focusing on differences between the high and the low rec strains.
 

### 2020-02-21 doc/Results.rmd
small changes

@@
 
+DO THE HIGH QUALITY
+
 <!-- breif background on longer female SC  (mouse, human) Gruhn-->
@@
 When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
-MSM
-SKIVE
-PWD
-MSM
-LEW
-G
-WSB
+MSM   `r `
+SKIVE  `r `
+PWD  `r `
+MSM  `r `
+LEW  `r `
+G  `r `
+WSB  `r `
 
@@
 high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.
+
+**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference
 

### 2020-02-21 src/BivData_Setup.Rmd
small changes

@@
 cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
 cell.level.BivData_table <- add_strain(cell.level.BivData_table)

### 2020-02-21 doc/Results.rmd
short notes on the IFD part

@@
 
-DO THE HIGH QUALITY
+DO THE HIGH QUALITY dataset tests
 
@@
                 ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-#THIS BIV OBS are also SIG
+```
+
+```{r Q1.show.IFD.things, echo=FALSE}
 
@@
 
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-###look at why KAZ males are weird
-
 nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
@@
 
+nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD.female_BIV.data
+
+nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_LEW.female_BIV.data
+
+nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G.female_BIV.data
+
+nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+###look at why KAZ males are weird
+
 nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ_BIV.data
 
@@
 nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD_BIV.data
+
 #this might
@@
 nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_MSM_BIV.data
 
 nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_SKIVE_BIV.data
 
 nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G_BIV.data
-```
-
-The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes?
+
+```
+
+see the reference plots within this dir, --
+
+this section -- all strains but KAZ, males have a higher mean normalized IFDs
+
+the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs
+
+The KAZ strain is the only strain without significant p value for the normalized IFDs, due to KAZ males having a higher rate (enrichment of short normalized IFDs.
+
+There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
+off.
+Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?
+
+<The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes?
 they were supposed to be 1CO -- but somehow a second CO formed anyways ..
@@
 whats the expectations following the confonding proportions? 
-high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.
+high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.>
 

### 2020-02-21 doc/Results.rmd
clean up and org for Q2 SC length

@@
 
+# transition Q1 to Q2
+
+we tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.
+
 # Q2 SC Length
 
+motivation:
+the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)
+
+<pattern for the single bivalent, pos correlation>
+The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
+Mainly that 1CO are shorter in the high rec strains then low rec strains.
+This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
+Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
+This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
+They have a lower probability of 2COs below a certain SC length threshold.
+
+<!-- basic stats -->
+
+`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
+X observations removed for having extra signal from other cells.
+`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
+
+<NEW STRAINS INTRODUCED NOT IN PREV DATA>
+
+this section / these tests follow up with those made in Q1 which found that the male musc and mol total-SC were signifcantly higher compared to the male Dom total SC measures.
+These test include additional measues from strains which only had male observations.
+
+In order to test if SC lengths (SC area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
+
+total SC for males (which are), X additional cells were added to the male specific total SC data set described above
+
+-reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice
+
 <!-- place set ups for total SC, short biv and long biv here -->
 
-I made a series of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromosome (positive correlation of SC length and CO number).
-Yet, given the complications from the chromosome proportions outlined above, comparing the bivalents across chromosome classes is complicated
-
-```{r juvimice.list.copy, echo=FALSE}
+```{r Q2.SC.juvimice.list.copy, echo=FALSE}
 juvi_mice <- c(
@@
                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
                          )
@@
 
-In order to test if SC lengths (SC area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
-
-The average SC lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confounded by the different proportions across strains.
-
-First we examine the pattern of SC length variation across chromosome classes.
+<!-- plots -->
 
@@
 \nMale Long Biv SC Lengths")
-mouse.shoo.long.biv
-```
-
-```{r Q2.SCshort_plot, echo=FALSE}
-
 ```
@@
 mouse.av_totsc_male
-
-```
-
-```{r Q2_SC_ChrmClass_long.biv, echo=FALSE}
-#there might not be enough data 
-#
-ii <- as.data.frame( table(male.long.biv$hand.foci.count, male.long.biv$strain) )
-```
-
-<!-- basic stats -->
-
-total SC for males (which ar)
-X additional cells were added to the male specific total SC data set described above
-
--reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice
-
+```
+
+I made a series of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromosome (positive correlation of SC length and CO number).
+Yet, given the complications from the chromosome proportions outlined above, comparing the bivalents across chromosome classes is complicated
+
+<!-- t tests-->
 <!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
@@
 G-WSB is sig, WSB-LEW are slightly sig
-
 -->
@@
 ```{r Q2.totSC.ttest, echo=FALSE}
-
 #totalSC high and low rec group
 
-#
 Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-
                 (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
 #p-value = 0.007
+
+#mol
+Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+
+#musc
+Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
 
@@
 
-**we ran t.tests on the SC based metrics to test is the high and low groups are significanly different**
-
+The average SC lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confounded by the different proportions across strains.
+
+total SC t test
+
+ p = `r Q2tot.sc_t.test_High.Low$p.value`
 **mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
@@
 
+musc and mol ttests
+
+p = `r`
+
+p= `r`
+
+Additionally we ran similar tests using the short and long bivalent data sets which help to alimeorate potential chromosome size effects.
+
 **The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
@@
 
-this might indicate -there's error in the total SC algorithm
-
-- mouse effects
-
+<this might indicate -there's error in the total SC algorithm
+- mouse effect
 - low power / sampling bias in reduced data set
-
-- total SC is capturing differences in the XY signal which is excluded from the bivalent data set
-
-glms
+- total SC is capturing differences in the XY signal which is excluded from the bivalent data set>
+
+<!-- models -->
 
@@
 ```
-
-(the subspecies effect is tested in M1)
-For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
-
-strain effects from M1; G, LEW, MSM, SKIVE and TOM
-
-strain effects  (basically all strain effects are significant) (tech effects / error?)
-All strain effects were significant when tested in M2 (Supplemental table X)
-
-For M2 all
 
@@
 
-**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
-The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
-
-**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
-
 ```{r Q2.glm.SC.short, echo=FALSE}
@@
 
-In the short bivalent data set -- analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).
-
-For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW
-G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.
-
 ```{r Q2.SC.long.glm.bivglm, echo=FALSE}
@@
 
-**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.**
-
-We purpose -- we think that running the models for comparing SC means by chromosome class and divided by rec group.
+<we ran models to start examine strain specific effects ... modeling the fixed effects of subspeices and strain.>
+
+Thse are our prediction for the glm models -- based on the variation of the gwRR's
+
+1. no difference in DOM strains (for SC lengths (CONFIRM IF THERE ARE DIFFERENECES FOR GWRR))
+
+2. PWD > SKIVE > the rest
+
+3. Msm > mol
+
+(the subspecies effect is tested in M1)
+For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
+
+strain effects from M1; G, LEW, MSM, SKIVE and TOM
+
+strain effects  (basically all strain effects are significant) (tech effects / error?)
+All strain effects were significant when tested in M2 (Supplemental table X)
+
+**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
+The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
+
+**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
+
+In the short bivalent data set -- analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).
+
+For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW
+G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.
+
+We purpose we think that running the models for comparing SC means by chromosome class and divided by rec group.
 (since the chromosome proportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )
@@
 
-`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
-X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
-
 The skeletonized measures show that the general pattern holds for total SC and the high rec strains.
-However there are a few exceptions -- However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel
+However there are a few exceptions -However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel
 
@@
 
-The general pattern is that the high rec strain have a greater mean SC length for all pooled bivalents (`r length(Curated_BivData$Obj.ID[Curated_BivData$sex == "male"])` males bivalent observations).
-
-The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high rec strains then low rec strains.
-This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
-Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
-This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
-They have a lower probability of 2COs below a certain SC length threshold.
-
 <!--
@@
 The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-
-```{r male.poly.test.MM, echo=FALSE, eval=FALSE}
-#I don't think the Mixed models fit well for the high vs low polymorphism tests.
-
-#t.test comparisons of PWD_SC vs (KAZ, SKIVE, )
-#MSM_SC vs MOLF_SC
-
-#2. logistic regression for sc length and chrm class?
-#if the prediction / significance
-
-#isolate data
-#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
-#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]
-
-mouse.avs_4MM_M.male$mouse <- factor(mouse.avs_4MM_M.male$mouse, ordered = FALSE)
-
-#build and test MM
-lme_MM_SC_length_male.poly <- lme(mean_SC ~ mouse + 1|strain, data= mouse.avs_4MM_M.male)#this syntax not working
-as.data.frame( round(  summary(lme_MM_SC.length)$tTable, 6 ) )
-
-lmer_MM_SC_length_male.poly <- lmer(mean_SC ~ (1|strain), data=mouse.avs_4MM_M.male, REML = TRUE)
-
-anova(lmer_MM_SC_length_male.poly)
-
-rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
-
-```
 
@@
 
-Each point is a mouse average SC.
+<Each point is a mouse average SC.
 When all bivalents are pooled, the logistic regression SC lengths predict the groups.
 When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.
-
- and the mean 2COs SC are higher in the high strains
-(this must have something to do with the physical size effect)
-
-If I want to try to use mouse nested in strain -- I should use cell level metrics (but those have there own flaws)
-
-Ideas for the above 3 tests,
-
-1. no difference in DOM
-- no sig logistic regression test
-- glm, strain effect?
+For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.>
 

### 2020-02-21 doc/Results.rmd
organizing the Q2 IFD section

@@
 total SC t test
-
  p = `r Q2tot.sc_t.test_High.Low$p.value`
@@
 musc and mol ttests
-
-p = `r`
-
-p= `r`
+p = 
+p=
 
@@
 
+<Compare by chrm class, but only for strains with similar proportions>
+
 We purpose we think that running the models for comparing SC means by chromosome class and divided by rec group.
@@
 Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
+
+<lots of variation across SC lengths -- doesn't fit clean expectation or signal>
+<balancing the interpertation of general positive correlation of CO number and SC length (SC length drives CO number) --- yet the pattern breaks down -it's not completely predictive -- /at finer levels?>  <noise/mouse effects/chromosome identity effects>
+
+<just like the sex differences in SC area -- this points to differential usage of the SC for placing CO -- investigate the recombination landscape>
 
@@
 
+# breief transition, rec landscape
+
+the different lack of predictive power of SC length to predict CO number -- indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.
+
 # Q2 Normalized CO positions
 
-In order to test /
+<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->
 
@@
 
-<!-- set up chunks here
-
--->
-
-```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
+<!-- set up chunks here -->
+
+```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
 #isolate all musc
@@
 
-```{r Q2.show.nrmF1, echo=FALSE}
+```{r Q2.1CO.show_plots, echo=FALSE}
 #Curated_BivData_male
@@
 
-Remade the Position table. and MSM has a more telomeric distribution. MOLF has the most central distribution.
-
-```{r Q2.F1.pos.LONG.BIV, echo=FALSE}
+```{r Q2.1COpos.LONG.BIV, echo=FALSE}
 
@@
 
-A main / biggish cavest to address for this section is the chromosome size effect, use real.long.bivData.
-
-or have chromosome lenght as a effect? would only work for models using the single bivalents, not mouse averages.
-
-<!-- 
-Main DFs for this section:
-Male.poly.Mouse.Table_BivData_4MM
-male.bivdata.1CO
--->
+<!-- t test -->
 
@@
 
-The t.tests are not significant for normalized foci position comparing all bivalets
-
-The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
-
 ```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
@@
 
-The t-tests for the high and low rec strains for the reduced bivalent datasets are mostly NS
-
-mouse mean
-short bivalent
-
-So try using the bivalent sample levels
-
 ```{r Q2.1COpos_BIV_ttest, echo=FALSE}
@@
 
-The t.test at the bivalent level is also NS (test; p = )
-
 ```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
@@
 
+The t.tests are not significant for normalized foci position comparing all bivalets
+The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
+
+The t-tests for the high and low rec strains for the reduced bivalent datasets are mostly NS
+
 **Most all of the t.tests I've done indecate that there are no sig diff between the High and the low rec groups**
 
+For normalized position of 1CO. 
 The normalized 1CO position is not a good metric to distinguish the high and low rec groups.
 
-GLMs
-
-glms are constructed to test if there has been evolution between strains ... ect.
-I will use the full data set the reduced datasets have such low power
+<!-- GLMs -->
+
+<toal data set, short and long biv>
 
@@
 
-for M1
-Mol subsp is significant, strain G (LEW slight sig), strain MSM -- strain MOLF is NA
-
-for M2,
-
-G, LEW, PWD, MSM, MOLF, SKIVE, KAZ have significant effects
-
-put together, these glm's indeceate that MOLF is the most different in it's 1CO landscape (compared to other males). Is this a sampling issue
-
-For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
-
-there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
-(motivation for MOLF female observations and more male MSM biv measures)
-
-```{r Q2.1CO.landscape, echo=FALSE}
+```{r Q2.1COpos.glm, echo=FALSE}
 #Male.poly.Mouse.Table_BivData_4MM
@@
 
-## JUNK Section
+<no prediction for the strains within subspecies>
+
+glms are constructed to test if there has been evolution between strains ... ect.
+I will use the full data set the reduced datasets have such low power
+
+for M1
+Mol subsp is significant, strain G (LEW slight sig), strain MSM -- strain MOLF is NA
+
+for M2,
+
+G, LEW, PWD, MSM, MOLF, SKIVE, KAZ have significant effects
+
+put together, these glm's indeceate that MOLF is the most different in it's 1CO landscape (compared to other males). Is this a sampling issue
+
+For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
+
+there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
+(motivation for MOLF female observations and more male MSM biv measures)
 
@@
 
+```{r Q2.IFD.long.short.setup, echo=FALSE}
+#long BivData
+male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
+
+#show
+Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+
+Long.biv.IFDs_PER_by.group
+
+```
+
 <!-- set up chunks here-->
-
-B.1) **Interfernce/IFD** will be shorter in high Rec strains. Use IFD_PER to account for SC length differences.
-
-1. glm across subsp
-
-2. anova within subsp
-
-3. logistic regression for musc (or all males)
 
@@
 
+<motivation background / fit more COs along a chromsome>
+
+t.test, are the high and low rec strains different?
+
+-glms -- are predictions based on the gwRR met for interference
+
+1. glm across subsp
+
+2. anova within subsp
+
+3. logistic regression for musc (or all males)
+
 The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
@@
 
-```{r Q2.IFD.glm.test, echo=FALSE}
-#glm --
-#mouse aver
-
-#M1.ABS
-glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.raw) 
-#M1.PER
-glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.PER)
-
-#M2.raw
-glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.1COnrm)
-
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD_Mol)
-
-#M2.PER
-glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.IFD.PER)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD.PER_Mol)
-```
-
-```{r Q2IFD_logreg, echo=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.PER, echo=FALSE}
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
+Full_2CO dataset
+- short biv data
+- long bivdata
+
+Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+The section on IFD only looks at the 2CO bivalents.
 
@@
 
-### Q2 Long Bivalent Dataset
-
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.
-
-```{r Q2.long_biv.IFD, echo=FALSE}
-#long BivData
-male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
-
-#show
-Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-
-Long.biv.IFDs_PER_by.group
-
-```
-
-table(droplevels(male.long.biv.2CO_IFD$strain) )
-
-The plot above shows the normalized IFD measures from 2CO bivalents
-
-There are many fewer long bivalent data observations, but the same positive correlation with
-
--table of long bivalents
-
-- glms
-
-- logistic regression
+The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
+
+<glms>
+
+```{r Q2.IFD.glm.test, echo=FALSE}
+#glm --
+#mouse aver
+
+#M1.ABS
+glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1.Q2_IFD.raw) 
+#M1.PER
+glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M1.Q2_IFD.PER)
+
+#M2.raw
+glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M2_Q2.1COnrm)
+
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+summary(Q2.glm.M2_IFD_Mol)
+
+#M2.PER
+glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
+summary(glm.M2_Q2.IFD.PER)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+summary(Q2.glm.M2_IFD.PER_Mol)
+```
+
+```{r Q2.IFD.PER.glm, echo=FALSE}
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
 
@@
 
-There are not enough observations for sub setting the data into strains or use the mouse averages.
-
-Neither t.test are significant for both the ABS and PER when I test just the Musc strains.
-The above t.tests are breaking the knitr
-
-The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
-
-None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.
+<None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.>
 
@@
 This also indicated that interference distance may evolved in the house mouse complex.
-
-#Caveats
-
-put all of the code chunks/analysis for caveats here
-
-## Chromosome Size Effect
-
-I tried to isolate bivalents which are in the top quartille for SC length from their cells. (re think where this section should go)
 
@@
 -->
-Below are examples of plots of SC length distributions across cells.
+
+<Below are examples of plots of SC length distributions across cells.
 The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
@@
 I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
-(the extra stats are not correctly mapped)
+(the extra stats are not correctly mapped)>
 
@@
                         var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
-        #add Q4
 
@@
 
+<below chunk goes with the above chunk>
 ```{r long.biv.plots, echo=FALSE}
@@
 
-in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice. 
-10202 bivalents from 86 mice.
+in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.
 
@@
 
-## Adjusting for XX
+# Adjusting for XX
 
@@
 
-#### Whole Cell Manual Comparison
+# Whole Cell Manual Comparison
 

### 2020-02-21 doc/Results.rmd
centralized short and long mouse.av tables

@@
 
+```{r LONG.SHORT.biv_setup, echo=FALSE}
+
+#FROM q1
+#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+#from Q2 IFD
+#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+##long biv
+#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+#FROM q1 
+REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
+                         mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                      #make sure 1CO  
+                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
+                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
+                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                         )
+
+REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
+
+REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
+REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
+REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
+                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
+                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
+                                           0)))
+
+REAL.short_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
+                                ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1, 0))
+
+###LONG BIV
+
+REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+                       n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
+                         mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
+                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
+                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+                         )
+
+REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
+
+REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
+REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
+REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
+                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
+                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
+                                           0)))
+
+REAL.long_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
+                                ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1, 0))
+
+#find which one 
+```
+
 <!-- 
@@
 
-Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength)
-                         )
-
-Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
-Q1.real.short_mouse_table <- add_species(Q1.real.short_mouse_table)
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-Q1.real.short_mouse_table$species <- factor(Q1.real.short_mouse_table$species, ordered = FALSE)
-
-Q1.real.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength)
-                         )
-
-Q1.real.long_mouse_table <- add_strain(Q1.real.long_mouse_table)
-Q1.real.long_mouse_table <- add_sex(Q1.real.long_mouse_table)
-Q1.real.long_mouse_table <- add_subsp(Q1.real.long_mouse_table)
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#make the Q1 strain list
+Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
 ```
@@
 
-```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE}
+```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE, eval=FALSE}
+
+#REAL.long_mouse_table
+
 #removed these plots for space
-
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-Q1.long.biv_cell_table <- ddply(real.long.bivData, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-Q1.long.biv_cell_table <- add_mouse(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_strain(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_sex(Q1.long.biv_cell_table)
-Q1.long.biv_cell_table <- add_subsp(Q1.long.biv_cell_table)
-
-Q1_cell_long.biv <- ggplot(Q1.long.biv_cell_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
+Q1_cell_long.biv <- ggplot(REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
 
@@
 
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+#cell.avlong.biv <- ggplot( (REAL.long_mouse_table %>% filter(sex == "male")), aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
 \nMale Long Biv SC Lengths")
@@
 #mouse av for long biv
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+
+mouse.show.long.biv <- ggplot((REAL.long_mouse_table %>% filter(sex == "male")), aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
 \nMale Long Biv SC Lengths")
 
+mouse.show.long.biv
+
 ```
@@
 ```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE, include=FALSE}
+
+Q1.real.long_mouse_table <- REAL.long_mouse_table
+
 ###
@@
 
+#REAL.short_mouse_table
+#REAL.long_mouse_table
+
 Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-Q1.long.1COpos_mouse.av <-  ddply(Q1.long.biv_1CO, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength), 
-                        mean_F1pos_PER =  mean(PER_Foci_1),
-                         mean_F1pos_ABS =  mean(Foci1)
-                         )
-Q1.long.1COpos_mouse.av <- add_strain(Q1.long.1COpos_mouse.av)
-Q1.long.1COpos_mouse.av <- add_sex(Q1.long.1COpos_mouse.av)
-Q1.long.1COpos_mouse.av <- add_subsp(Q1.long.1COpos_mouse.av)
-
-Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-Q1.short.1COpos_mouse.av <-  ddply(Q1.short.biv_1CO, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean_SC = mean(chromosomeLength),
-                        mean_F1pos_PER =  mean(PER_Foci_1),
-                         mean_F1pos_ABS =  mean(Foci1)
-                         )
-
-Q1.short.1COpos_mouse.av <- add_strain(Q1.short.1COpos_mouse.av)
-Q1.short.1COpos_mouse.av <- add_sex(Q1.short.1COpos_mouse.av)
-Q1.short.1COpos_mouse.av <- add_subsp(Q1.short.1COpos_mouse.av)
 
@@
 ```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-Q1.short.biv_scatr_box   <- ggplot(Q1.short.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+
+#real.long.bivData
+
+Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nShort Biv")
-
-Q1.long.biv_scatr_box   <- ggplot(Q1.long.biv_1CO, aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+Q1.short.biv_scatr_box
+
+Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nLong Biv")
-
+Q1.long.biv_scatr_box
 ```
@@
 
-#make the same plots as above but with the long biv data set
-
+#REAL.short_mouse_table
+#REAL.long_mouse_table
 #mouse.av.long.biv.table
-Q1.av.long.biv_scatr_box   <- ggplot(Q1.long.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+
+Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nLong Biv")
 
-Q1.av.short.biv_scatr_box   <- ggplot(Q1.short.1COpos_mouse.av, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nShort Biv")
-
 ```
@@
 ```{r show.norm.siscoten, echo=FALSE}
-
 Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
@@
 #glm for siscoten
-
 #mouse.avs_4MM  - mouse averages with mean sis.co.ten
-
 #MM
@@
 #sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
 lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
@@
 
-The fixed effects, sex and sex*subsp are significant. The random strain effect is also significant.
-
-> Is the heterochiasmy prediction met?
-
-Yes, model  predicting the mouse average siscoten, sex and sex-subp interaction are significant factors. The Random strain effect is also significant.
-
 ```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
@@
 #mouse.avs_4MM
-
 ```
@@
 #why do a different number of rows removed each time??
-
 ```
@@
 lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-
 drop1(lme_MM_telo.dist, test="Chisq")
-
 as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
@@
 rand_telo.dist_MM
-
 ```
@@
 #Only use 2CO bivalents
+
 real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
@@
 ```{r Q1.show.IFD.things, echo=FALSE}
-
 nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
@@
 ```{r Q2.SC.short.long.setup, echo=FALSE}
-table(male.long.biv$strain)
+
+#table(male.long.biv$strain)
 #the Long biv dataset is quite unbalance (esp MSM)
@@
                         norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         )
+                         ) )
 
@@
 
-shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#real.long.bivData
+#real.short.bivData
+
+shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
 \nMale Long Biv SC Lengths")
 
-mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
 \nMale Long Biv SC Lengths")
 
-cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
 \nMale Long Biv SC Lengths")
+
 ```
@@
 
-<!-- t test -->
+<!-- t tests -->
 
@@
 #remove SKIVE
-Q2.1COpos_short_ttest_HL.ex <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+Q2.1COpos_short_ttest_HL.sansSKVIE <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
             (Q2mouse.av.short.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
@@
 ```
+
+t.tests for high and Low rec strains
+
+long 
+Q2.1COpos_short_ttest_HL
+
+short
 

### 2020-02-22 doc/Results.rmd
reaching dead end for the day, plus commit some old files

@@
 
-#has 
 #18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
-Curated_BivData <- Curated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
-
-mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
+
+#cleanCurated_BivData
+#removed the no hand foci count
+
+mouse.avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
 
-male.av_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
 
-We anchor the position measures at the centromere (and normalize by deviding the position by the length of the chromosome.
-The foci positions then have a range of 0-1, with smaller values representing positions closer to the centromere end of the bivalent, and larger values meaning foci positions closer to the opposite end (refered here as telomeric position).
+We anchor the position measures at the centromere (and normalize by deviding the position by the length of the chromosome, so the value range from 0 to 1, with smaller values representing positions closer to the centromere end of the bivalent, and larger values meaning foci positions closer to the opposite end (refered here as terminal or telomeric position).
 
@@
 
-In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (with female)
-are highly significantly more medially placed (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
+In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (fitting the predictions/pattern of the 'typical recombination landscape' (Sardell)).
+The single CO positions are medially placed in females and terminall placed in males.
+
+The pavlues across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
+
 With the musculus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
@@
 
-# Q2 Normalized CO positions
+# Q2 Normalized Single CO positions
 
 <!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->
-
-We tested another aspect of the rec landscape -- focusing on differences between the high and the low rec strains.
-
-To detect if there are significant differences between the high and the low rec strains we performed t.tests across the two groups
-
-(and logistic regressions?)
-
 <!-- set up chunks here -->
-
-```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
 <!-- plots /figures below-->
 
-```{r Q2.1CO.show_plots, echo=FALSE}
+```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
 #Curated_BivData_male
@@
 
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos")
-
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos All bivalents")
 Q2.show.nrmF1
@@
 
-nrmF1_mouse_level
-```
-
-```{r Q2.1COpos.LONG.BIV, echo=FALSE}
-
+```
+
+<not sure what the beenfit is to test for chrm effectss>
+
+```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
 #make the same plots as above but with the long biv data set
@@
 Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Q2.show.nrmF1_long.biv
 
@@
 
-Q2.shoooo
-```
-
-<!-- t tests -->
+```
+
+The previous section found that the single crossover landscape is significantly different between males and females.
+In this section -- we focus on differences between the high and the low rec male strains
+In order to test if if there are distinguishing meterics for these two groups.
+
+<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->
 
 ```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
-
-#mouse.avs_4MM
-#31may19_SKIVE_m2  has NA for norm.first.foci.pos
+#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
 
@@
 
-#single biv level
-Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value 
-#p-value = 0.238
-
+```
+
+```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
+
+#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
+
+#Musc
+Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+# p-value = 0.4
+
+#Mol
+Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+# p-value = 0.07
+# exclusive rec group not sig
 ```
@@
 
-#short
-Q2.1COpos_short_ttest_HL <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
-#p-value = 0.06
-
-#remove SKIVE
-Q2.1COpos_short_ttest_HL.sansSKVIE <- t.test( (Q2mouse.av.short.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
-            (Q2mouse.av.short.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
-#p-value = 0.4
-
-#long
-Q2mouse.av.long.biv.table_HL  <-  t.test( (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (Q2mouse.av.long.biv.table %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+#short -- real.long.bivData -- REAL.short_mouse_table
+Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
+
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
+#p-value = 0.3
+
+Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
+# p-value = 0.1
+
+#remove SKIVE??
+```
+
+```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
+
+#real.long.bivData
+#single biv level
+Rec.group_LONG.BivData_F1post_Ttest <- t.test(
+  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
+(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
 #p-value = 0.5
 
-#long exclude skive  Q2mouse.av.long.biv.table
-Q2mouse.av.long.biv.table_HL.ex  <-  t.test( (Q2mouse.av.long.biv.table  %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
-            (Q2mouse.av.long.biv.table %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos   )
-#p-value = 0.08
-
-#they kinda have ooposite effects
-
-```
-
-```{r Q2.1COpos_BIV_ttest, echo=FALSE}
-
-whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
+Rec.group_BivData_F1post_Ttest <- t.test(
+  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
 (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
@@
 
-whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
-            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
-
+#
+#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
+#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
+#p-value = 0.2
+
+#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
+#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
 #p-value = 0.4
@@
 
-```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
-
-#compare within subsp
-
-#Musc
-Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
-
-# p-value = 0.4
-# exclusive rec group not sig
-```
-
-t.tests for high and Low rec strains
-
-long 
-Q2.1COpos_short_ttest_HL
-
-short
-
-The t.tests are not significant for normalized foci position comparing all bivalets
-The normalized Foci Positions are not significantly different between the high and low rec groups (full data set, mouse av and single bivalent level).
-
-The t-tests for the high and low rec strains for the reduced bivalent datasets are mostly NS
-
-**Most all of the t.tests I've done indecate that there are no sig diff between the High and the low rec groups**
-
-For normalized position of 1CO. 
-The normalized 1CO position is not a good metric to distinguish the high and low rec groups.
-
-<!-- GLMs -->
-
-<toal data set, short and long biv>
+**The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**
+
+The first use t.tests to test if there are significant differences in the single CO landscapes between the high and low rec groups.
+
+T tests of the single crossover positions reveal that the normalized 1 foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
+
+further tests -- within the subspecies, musculus and molossinus,
+
+The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
+
+While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`).
+This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.
+
+ -- BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF,
+
+(MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.
+
+Where as something is
+
+<long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`> 
+<!-- GLMs because there's no difference 
+toal data set, short and long biv. There's only a difference between MSM and MOKF
+test ppredictions, Dom = NULL,  -->
 
@@
 Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-
 sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
+#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
+
+#with.chrm.legth  ?
+Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
 
@@
 Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-
 sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
@@
 
-```{r Q2.1COpos.glm, echo=FALSE}
-#Male.poly.Mouse.Table_BivData_4MM
-#Curated_BivData_male
-
-#glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#M1
-glm.M1_Q2.1COnrm <- glm(norm.first.foci.pos ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm.M1_Q2.1COnrm)
-
-#M2
-glm.M2_Q2.1COnrm <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm.M2_Q2.1COnrm)
-
-#Dom
-glmQ2_M2.1COnrm_Dom <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glmQ2_M2.1COnrm_Dom)
-
-#Musc
-glmQ2_M2.1COnrm_Musc <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glmQ2_M2.1COnrm_Musc)
-
-#Mol
-glmQ2_M2.1COnrm_Mol <- glm(norm.first.foci.pos ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glmQ2_M2.1COnrm_Mol)
-
-#m3 (log)
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                            ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-Log.glmQ2_M2.1COnrm <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.glmQ2_M2.1COnrm_Musc <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm_Musc)
-
-#Mol
-Log.glmQ2_M2.1COnrm_Mol <- glm(Rec.group ~ norm.first.foci.pos,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],
-              family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-#biv level logistic regression
-#Curated_BivData_male
-#male.bivdata.1CO
-
+in the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
+..
+the chromosome length has the most significant effect, LEW by chrm effect are significant
+
+In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects
+
+**For the M2 (subsp)**
+There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).
+
+LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).
+
+But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).
+
+and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)
+
+**For M3 (all strains)**
+
+strain G (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][2]`).
+
+strain Lew (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][3]`).
+
+strain PWD ( p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][4]`).
+
+slight strain MSM  (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][5]`).
+
+very sig strain MOLF (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][6]`).
+
+strain SKIVE (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][7]`).
+
+strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
+
+from the main figure above, you can see that G and MOLF 1CO foci positions are the most different
+
+<no prediction for the strains within subspecies>
+
+For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
+
+there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
+(motivation for MOLF female observations and more male MSM biv measures)
+
+```{r Q2.1COpos.show.MOL, echo=FALSE}
+
+(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos
+
+mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
+                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
+
+#+facet_wrap(~mouse, scales = "free")
+
+single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
+                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
+
+#+scale_color_manual(values=colors_of_strains)
+
+single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
+                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
+
+#CHRM.length is very significant-- some of the same subsp and strains are sig
+
+```
+
+I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.
+
+**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$
+
+```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
+#remove this cell,
+
+Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
+
+Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
+
+#making new models
+
+Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
+Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
+
+chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+
+sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
+#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
+
+chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+
+#summary(chrm.1CO.pos.chrmLength_strain)
+
+```
+
+for the first model,
+
+chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`
+
+subsp MOL `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][4]`  
+the mol obj above isn't showing the reight value.
+
+strain G   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][5]`
+
+strain LEW   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][6]`
+
+Strain MSM `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][8]`
+
+chrm * LEW `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][14]`
+
+chrm *MSM  `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][16]`
+
+M2 / for the second model ...
+
+-chromosome length, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][2]`
+
+-strain G, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][3]`
+
+-strain LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][4]`
+
+-strain MOLF, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][7]`
+
+-strain SKIVE, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][8]`
+
+-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`
+
+**males from WSB and MOLF strains have 'strang'/unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation**
+
+**There are significant difference across the 1CO positions, yet they don't match /fit any of the predictions. there are also complications due to the chromsome class proportions that **
+
+```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
+#isolate all musc
+#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
+```
+
+```{r Q2.1COpos.glm, echo=FALSE, eval=FALSE}
 male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
@@
 #summary(Log.glmQ2_M2.1COnrm_Mol)
-
-```
-
-<no prediction for the strains within subspecies>
-
-glms are constructed to test if there has been evolution between strains ... ect.
-I will use the full data set the reduced datasets have such low power
-
-for M1
-Mol subsp is significant, strain G (LEW slight sig), strain MSM -- strain MOLF is NA
-
-for M2,
-
-G, LEW, PWD, MSM, MOLF, SKIVE, KAZ have significant effects
-
-put together, these glm's indeceate that MOLF is the most different in it's 1CO landscape (compared to other males). Is this a sampling issue
-
-For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
-
-there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
-(motivation for MOLF female observations and more male MSM biv measures)
+```
 

### 2020-02-22 doc/Results.rmd
fixed some erroneus p values in Q1 SC section

@@
 
-Comparing the means of single bivalents is complicated by the fact we can't identify the chromosome identitied the physical size (in terms of SC-AE) is the closet approximation we can make.
+<Comparing the means of single bivalents is complicated by the fact we can't identify the chromosome identitied the physical size (in terms of SC-AE) is the closet approximation we can make.>
 
@@
 
-For this reason we use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells).
+We use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data.
 A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes.
 
-Even though this smaller data set has decreased power due to redcued sample size, ...
+<Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes>
 
@@
 
-For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer, the largest difference across sexes by strain are>  (t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, `r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and `r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
+For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer, the largest difference across sexes by strain are>  (t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
 for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
-and (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
+This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
 
@@
 
+```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE, message=FALSE}
+
+Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+
+#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
+
+```
+
+```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M3)
+
+summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
+
+#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#sex male very sig -- only sig
+```
+
 The following effects had significant p values given the above model
@@
 (subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
-
-These results indicate that there are strain specific effects for the SC-AE area --
-
-the large G strain effect is likely driven by the high female mean total SC area
-
-sex is the next strongest (most significant) effect.
 
@@
 
-For the short bivalent data -- run through the same models...
-
-```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][10]`).
-while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.
-
-In the M2 model, the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][10]`).
-<should apply caution to the short bivalent dataset> The significant SKIVE effect is driven by  2 of the 4 female skive mice -- having much shorter bivalents comapred to the other female musc strain means.
+From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).
+
+while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.
+
+In the M2 model, the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).
 

### 2020-02-23 doc/Results.rmd
more calqs from Q2 sc length

@@
 
-<and the short biv>
 Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
-interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
+The interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
 
@@
 
-```{r Q1.glm_totalSC, echo=FALSE, message=FALSE}
+```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
 #total-SC
@@
 
-```{r Q1.SC.Length_M2, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-The following effects had significant p values given the above model
+The following effects had significant p values for the **total SC** from the above model
 
@@
 
-From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).
+Fr the **short bivalent** sc the model results are listed below. From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).
 
@@
 
-In the M2 model, the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).
-
-<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-AE.  >
-
-In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.
+In the M3 model (strains and sex as fixed effects), the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).
+
+<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >
+
+<In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
+
+<1. the longer female sc's is consistant and seen across all strains,
+even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
+2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>
 
@@
 
-```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Q1.real.long_mouse_table, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-```
-
 ```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
@@
 
-#mouse av
-Q1.SClength_Musc_plot <- ggplot( Musc.mouse.av, aes(y = mean_SC, x = strain, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length")+facet_wrap(~sex)+ylim(c(75,125) )
-
 #BivData Level
@@
 
-```
-
-```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE, eval=FALSE}
-
-#REAL.long_mouse_table
-
-#removed these plots for space
-Q1_cell_long.biv <- ggplot(REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,200))+geom_boxplot(alpha=.2) + ggtitle("Cell averages for Long Bivs")+theme(legend.position="none")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.sc = mean(chromosomeLength)
-                         )
-
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-#cell.avlong.biv <- ggplot( (REAL.long_mouse_table %>% filter(sex == "male")), aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-#mouse av for long biv
-
-mouse.show.long.biv <- ggplot((REAL.long_mouse_table %>% filter(sex == "male")), aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-mouse.show.long.biv
-
-```
-
-```{r Q1_long_biv_lmer, echo=FALSE, message=FALSE, include=FALSE}
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table
-
-###
-lmer.Q1_SCLength_long_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_long_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_long_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.long_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_long_interaction  <-drop1(lmer.Q1_SCLength_long_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_long_sex  <-anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_long_subp  <- anova(lmer.Q1_SCLength_long_M1, lmer.Q1_SCLength_long_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_long_strain  <- exactRLRT(lmer.Q1_SCLength_long_M1)$p.value
-
-Q1lmer.MM_SC_long_results <- data.frame(sub.sp = Q1lmer.MM_SC_long_subp, 
-sex.results =  Q1lmer.MM_SC_long_sex, 
-               interaction.pval = Q1lmer.MM_SC_long_interaction,  rand.pval = 
-Q1lmer.MM_SC_long_strain)
 ```
@@
 
-motivation:
-the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)
+<motivation:
+the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
 
@@
 
+while testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR
+
 <!-- basic stats -->
@@
 
-In order to test if SC lengths (SC area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.
-
-total SC for males (which are), X additional cells were added to the male specific total SC data set described above
-
--reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice
+-reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice.
+(Since comparing within sex, there are no sex chromosomes differences and the long bivalent dataset can be used.
 
 <!-- place set ups for total SC, short biv and long biv here -->
-
-```{r Q2.SC.juvimice.list.copy, echo=FALSE}
-juvi_mice <- c(
-  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
-  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
-  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
-
-  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
-   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
-  )
-
-ppmm <- OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$mouse  %in% juvi_mice, ]   
-```
 
@@
 
+#mUSC-short
+Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-short
+#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
+
+#musc-long
+Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-long
+#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
+
+#not enought observations
+
+```
+
+```{r Q2.SC_HvL_allBiv, echo=FALSE}
+
+#High-Low_All-BivData
+
+Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#short
+Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#long
+Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
 ```
@@
 musc and mol ttests
-p = 
-p=
+mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` Mole have significant difference between mouse means
+
+musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`
+the total sc mouse means are not significant for the high and low strains in musc
 
@@
 
-**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
+**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
 (the summary of mouse averages for this subset of ) **
+
+within subspecies
+
+musc-short
+
+musc-long
+
+mol-short
+
+mol-long
+
+```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
+
+#Dom-Musc
+Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
+                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
+#p-value = 0.5
+
+#Dom-Mol
+Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
+                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
+#p-value = 0.1
+```
+
+```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
+
+#DOM-musc-long
+Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.08
+
+#DOM-musc-short
+Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.09
+
+```
+
+```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
+
+#DOM-MUSC-ALL
+Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 4e-13
+
+#DOM-MOL-ALL
+Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 6e-10
+
+#DOM-musc-long
+Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.2
+
+#DOM-musc-short
+Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.06
+
+```
+
+across subspecies:
+
+dom - musc
+
+dom-mol
 

### 2020-02-23 doc/Results.rmd
must make progress

@@
 
-total SC t test
+total SC t tests
  p = `r Q2tot.sc_t.test_High.Low$p.value`
@@
 
-within subspecies
-
+<within subspecies
 musc-short
@@
 mol-long
+very few significant comparisons>
 
@@
 
-across subspecies:
-
-dom - musc
-
-dom-mol
-
-<this might indicate -there's error in the total SC algorithm
-- mouse effect
-- low power / sampling bias in reduced data set
-- total SC is capturing differences in the XY signal which is excluded from the bivalent data set>
+across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).
 
@@
 ```{r Q2.totSC.glm, echo=FALSE}
-
 #(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel
@@
 sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
-
-#sum.Q2.glm_mus.av_totSC_M1.coef <-
 
@@
 glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
 ```
@@
 #round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
-
 #summary(glm_SC.male_Musc_long.biv)
@@
 
-<we ran models to start examine strain specific effects ... modeling the fixed effects of subspeices and strain.>
-
-Thse are our prediction for the glm models -- based on the variation of the gwRR's
+**For the total SC glm**
+
+Thse are our prediction for the glm models based on the variation of the gwRR's
 
@@
 (the subspecies effect is tested in M1)
-For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
-
-strain effects from M1; G, LEW, MSM, SKIVE and TOM
+
+For total SC in M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
+
+strain effects from M3; G, LEW, MSM, SKIVE and TOM are significant
 
@@
 
-<Compare by chrm class, but only for strains with similar proportions>
-
-We purpose we think that running the models for comparing SC means by chromosome class and divided by rec group.
-(since the chromosome proportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )
-
-The null hypothesis is that these won't be highly significant strain effects.
-If there are still strain effects for these models, that would be evidence that there has been evolution for SC length compaction across these strains
-
 The skeletonized measures show that the general pattern holds for total SC and the high rec strains.
@@
 
-The CZECH and AST are higher than KAZ
-
-The number of cells and mice is difference across the low group strains, 
--- permute the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)
-
-above are glm tables for mouse.av SC lengths for each strain.
-
-Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
+The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)
 
@@
 blocks of other male Q2 - SC lengths
-
 what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
 The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-
 -->
 
-> **Summary**
-
-When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression.
-I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length.
-(IE the scatter plots of SC length, binned by chrm class on the X).
+> **Summary Points**
+
+- confirm that the positive correlation of SC and CO number is maintained
+
+- Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.
+
+- (the glms, indecate there are many strain effects), Using the mouse average SC lengths, in the full data set all strain effects are significant.
+This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
+
+- For the M3 models of mouse av.s across SC meterics, for many of the SC metrics, most all of the strain effects are significant
+
+- for the M2 models of the mouse av, musc - mol are significant fixed effects
+
+- many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging
 
@@
 
-When all male mice are used, the predictive power is greater, than when just the Musc strains are used.
-When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)
-
-> Is the prediction, high rec musc male strains have long SC met?
-
-In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.
-I couldn't get the Mixed models working for the male polymorphism predictions...
-
 <!-- junk chunks below -->
@@
 
-These are the plots of the logistic regression.
+<These are the plots of the logistic regression.>
 
@@
 ```
-
-the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group
 
@@
 
-```{r Q2.SC.Length.stuff, echo=FALSE, include=FALSE}
-#REMOVE THIS, COMPARES FULL BIVDATA ACROSS CHRM CLASS
-
-#all pooled SC lengths
-#WSB     G   LEW  MOLF   KAZ CZECH 
-#    5     9     5     4     3     3
-
-#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14
-
-#G sample 5 mice
-
-#break down of mice in the low rec dataset
-
-#glm.1CO
-Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
-summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
-round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)
-
-#glm.2CO
-Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
-summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
-round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
-#fewer significant differences
-
-Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
-sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
-round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
-#all but KAZ are slightly significant, G and CZECH more so
-
-#glm.3CO
-Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
-#not sig differences
-summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
-round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)
-
-#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)
-
-#all pooled SC lengths
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-Q1lmer.MM_SC_short_results
-```
-
-```{r Q2.SC.ploy.male, echo=FALSE, include=FALSE}
-#REMOVE THIS< COMARES FROM FULL BIVDATA set
-
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-
-#1. determine which strain effects are sig
-#use mouse biv average
-#Male.poly.Mouse.Table_BivData_4MM
-
-glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm1.SC.male.poly) 
-#SC many strains are different (more than the gwRR)
-#subsps are sig
-
-glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-#summary(glm2.SC.male.poly)
-
-#Dom
-glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(glm_SC.male_Dom)  #ALL strain are sig
-#both G and LEW have significant mean.SC, doesn't meet predictions
-
-#Musc
-glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(glm_SC.male_Musc)
-
-#Mol
-glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(glm_SC.male_Mol)
-
-#Logistic regressions
-
-log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-
-                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],
-
-#summary(log.SC.male.poly)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_musc)
-#not sig -- when just looking at Musc
-
-#significant in predicting group
-log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC.male.poly_mol)
-#not sig -- when just looking at Mol
-```
-
 ```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
@@
 
-```{r SC.length.male_tests, echo=FALSE, include=FALSE}
-
+```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
 ## run the tests to check following expectations
@@
 #3.MSM and MOLF are different from each other
-
-#Male.poly.Mouse.Table_BivData_4MM
-#Long.biv_mouse_table_male.poly
-
-#decide between lm or glm..?
-#mouse avs, males
 
@@
 
-```{r SC.length.pairwise.T, echo=FALSE, include=FALSE}
+```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}
 
@@
 #pairwise.t.test(Ozone, Month, p.adj = "bonf")
-
-```
-
-# breief transition, rec landscape
+```
+
+# brief transition, rec landscape
 

### 2020-02-23 doc/Results.rmd
small changes

@@
 musc and mol ttests
-mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` Mole have significant difference between mouse means
+mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` Mole have significant difference between mouse means for total sc, between the high and low srains
 
@@
 
-Thse are our prediction for the glm models based on the variation of the gwRR's
+Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
 
@@
 - many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging
+
+- th predictions for M3 (strain) differencs (motivated by the gwrr variation are not met,  (correlation of SC means and CO means?))
 
@@
 
-the different lack of predictive power of SC length to predict CO number -- indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.
+the lack of predictive power of SC length to predict CO number indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.
 
@@
 The previous section found that the single crossover landscape is significantly different between males and females.
-In this section -- we focus on differences between the high and the low rec male strains
+In this section we focus on differences between the high and the low rec male strains
 In order to test if if there are distinguishing meterics for these two groups.
@@
 
-The first use t.tests to test if there are significant differences in the single CO landscapes between the high and low rec groups.
-
-T tests of the single crossover positions reveal that the normalized 1 foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
-
-further tests -- within the subspecies, musculus and molossinus,
+T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
+
+<further tests -- within the subspecies, musculus and molossinus?
 
@@
 
- -- BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF,
-
-(MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.
-
-Where as something is
+ <BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF, 
+ (MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.>
 
@@
 ```
+
+in addition models which use mouse average of foci position as a dependat variable, we decided to run models for the single bivalent level, and which include chromosome length as a fixed effect.
 

### 2020-02-24 doc/Results.rmd
cleaning up Q2 IFD section

@@
 #CHRM.length is very significant-- some of the same subsp and strains are sig     
-
 ```
@@
 
-for the first model,
+<for the first model,
 
@@
 
--chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`
+-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`>
 
@@
 
-Long.biv.IFDs_PER_by.group
-
 ```
@@
 
-```{r Q2.show.IFD, echo=FALSE}
+```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
 
@@
 
-<motivation background / fit more COs along a chromsome>
-
-t.test, are the high and low rec strains different?
-
--glms -- are predictions based on the gwRR met for interference
-
-1. glm across subsp
-
-2. anova within subsp
-
-3. logistic regression for musc (or all males)
-
-The above plots are not the best at showing the pattern of higher rec group having slightly higher raw IFD - because their SC are longer on average. 
- The normalized plot displays the pattern of higher rec group having slightly higher longer IFDs.
-
-Full_2CO dataset
-- short biv data
-- long bivdata
-
+<Instead of the long and short biva data, the total 2CO bivalent dataset is used to make models.>
+
+<Previous prediction for differences in the long and short bivalent dataset.
 Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
 In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.
-
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=FALSE}
+The section on IFD only looks at the 2CO bivalents.>
+
+```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
 
@@
 #mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_ABS[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-t.test_mouse.av.IFD_PER <- t.test(
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$Rec.group == 1)],
-
-  mouse.avs_4MM$mean_IFD.2CO_PER[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
-  )
-
-```
-
-```{r Q2.t.test_IFD,echo=FALSE}
+t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
+
+```
+
+```{r Q2.t.test_IFD,echo=FALSE, eval=FALSE}
 
@@
 
-#MOUSE AVE
-Rec.group_IFD.PER_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_IFD.2CO_PER[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
+Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
+                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
 
@@
 
-The t.tests return very significant p values for the normalized IFD between the high and low rec groups (for both mouse av and single bivalent observations levels).
-
-The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.
+The mouse averages for IFDs were significantly longer in the high rec groups ( p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The pattern is confirmed with t.test comparing groups within the same subepcies.
+**IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.**
 
@@
 
+The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
+
+**model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
+
+**model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
+
 ```{r Q2.IFD.glm.test, echo=FALSE}
-#glm --
-#mouse aver
+#REPLACE Male.poly with ..
+#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
 
 #M1.ABS
-glm.M1.Q2_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.raw) 
+Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
+#summary(Q2.M1glm_IFD.raw)
+
 #M1.PER
-glm.M1.Q2_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M1.Q2_IFD.PER)
+Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
+
+sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
+#summary(Q2.M1glm_IFD.PER)
 
 #M2.raw
-glm.M2_Q2.1COnrm <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.1COnrm)
-
+Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+#summary(Q2.M2glm_IFD.raw)
+
+Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+#within subspecies
 #Dom
 Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD_Dom)
+#summary(Q2.glm.M2_IFD_Dom)
 
@@
 Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD_Musc)
+#summary(Q2.glm.M2_IFD_Musc)
 
@@
 Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD_Mol)
-
-#M2.PER
-glm.M2_Q2.IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
-summary(glm.M2_Q2.IFD.PER)
+#summary(Q2.glm.M2_IFD_Mol)
 
@@
 Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-summary(Q2.glm.M2_IFD.PER_Dom)
+#summary(Q2.glm.M2_IFD.PER_Dom)
 
@@
 Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-summary(Q2.glm.M2_IFD.PER_Musc)
+#summary(Q2.glm.M2_IFD.PER_Musc)
 
@@
 Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-summary(Q2.glm.M2_IFD.PER_Mol)
-```
-
-```{r Q2.IFD.PER.glm, echo=FALSE}
+#summary(Q2.glm.M2_IFD.PER_Mol)
+```
+
+In M1, the model build on the mouse averages of IFD metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]`  for PWD, MSM and SKIVE respectively).
+In M2, the mouse average for normalzied IFD is modeled with only strain effects.
+Only PWD and SKIVE IFD^PER^ had significant strain effects (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]` for PWD and SKIVE respectively).
+
+<when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF>
+
+We ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length.
+
+**model BivLen.M1 ** $$Single bivalent\ IFD ~=~ subsp * strain * chromosomeLength + \varepsilon $$
+
+**model BivLen.M2 ** $$Single bivalent\ IFD ~=~ strain * chromosomeLength + \varepsilon $$
+
+```{r Q2.IFD.glm.chrmLength, echo=FALSE}
+
+#repeating chrm models with chromsome length
+#FOR BIVALENT LEVEL MODELS
+
+Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
+
+#M.Len.ABS
+Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.raw)
+
+#M.Len.PER
+Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.PER)
+
+#M.Len.raw
+Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
+
+##M.Len PER
+Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+```
+
+For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs).
+Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^.
+The Mol\*chrmLength is negative (IN MOL as chrm length got longer - IFD gets smaller -- also in SKIVE\*chrm
+
+For the strain only model,
+
+- PWD, MSM, MOLF, SKIVE are the only significant strain effects.
+
+- all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
+
+The general chromsome Length effect is that
+
+The single fixed effects can be interperted as, with each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
+All of the high rec strains and MOLF have significantly longer IFD^PER^.
+
+There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
+
+```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
+
+#Log regression all show significant prediction
 
@@
               data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER)
+#summary(log.reg.Q2.IFD.PER)
 
@@
               data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Musc)
+#summary(log.reg.Q2.IFD.PER.Musc)
 
@@
               data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-summary(log.reg.Q2.IFD.PER.Mol)
+#summary(log.reg.Q2.IFD.PER.Mol)
 
@@
               data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-summary(Log.Q2.IFD.per_bivdata)
+#summary(Log.Q2.IFD.per_bivdata)
 
@@
 
-summary(Log.Q2.IFD.per_bivdata.Musc)
+#summary(Log.Q2.IFD.per_bivdata.Musc)
 
@@
 
-summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE}
+#summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
 
@@
 #The PER.IFDs don't have significant strain effects
-
 ```
@@
 
-```{r SC.distb_manual_Auto, echo=FALSE}
-
-#calq the mouse means
-manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
-                        median_SC = median(SC.length, na.rm = TRUE),
-                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
-#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
-
-)
-
-manual_summary_stats <- add_mouse(manual_summary_stats)
-manual_summary_stats <- add_strain(manual_summary_stats)
-manual_summary_stats <- add_sex(manual_summary_stats)
-manual_summary_stats <- add_subsp(manual_summary_stats)
-
-###
-#calq the mouse means
-Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        median_SC = median(chromosomeLength),
-                        Qar.4 = quantile(chromosomeLength)[[4]],
-
-                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
-
-)
-Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
-
-#colors_of_strains
-#scale_color_manual(values=colors_of_strains)
-
-#un-order these factors
-```
-
 <below chunk goes with the above chunk>
-```{r long.biv.plots, echo=FALSE}
+```{r long.biv.plots, echo=FALSE, eval=FALSE}
 
@@
 
-All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
+<All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
 It shows the amount of within cell variance across strains.
 There is a bit of variance across the SC length distributions in the PWD females.
-
 This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
-
 The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
@@
 **- Try to merge this DF with the whole.cell manual measures.**  
-
 - Try estimating which 'loose' bivalent observations might be within the long class of bivalents.  
-
-in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.
-
-```{r show.long.SC, echo=FALSE}
+in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.>
+
+```{r show.long.SC, echo=FALSE, eval=FALSE}
 
 female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
-female_long_biv_plot
 
 male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
-male_long_biv_plot
-
-```
-
-These plots show the SC lengths for the 'long SC data set'.
+
+```
+
+<ABOVE plots show the SC lengths for the 'long SC data set'.
 They are supposed to be the longest 4-5 SC from cells where I could get good measures.
 These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
-Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
+Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
+
+<!-- unused chunkcs -->
+
+```{r SC.distb_manual_Auto, echo=FALSE, eval=FALSE}
+
+#calq the mouse means
+manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
+                        median_SC = median(SC.length, na.rm = TRUE),
+                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
+#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
+
+)
+
+manual_summary_stats <- add_mouse(manual_summary_stats)
+manual_summary_stats <- add_strain(manual_summary_stats)
+manual_summary_stats <- add_sex(manual_summary_stats)
+manual_summary_stats <- add_subsp(manual_summary_stats)
+
+###
+#calq the mouse means
+Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        median_SC = median(chromosomeLength),
+                        Qar.4 = quantile(chromosomeLength)[[4]],
+
+                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
+
+)
+Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
+
+#colors_of_strains
+#scale_color_manual(values=colors_of_strains)
+
+#un-order these factors
+```
 
@@
 F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-F1.x.F2
 

### 2020-02-24 doc/Results.rmd
small editing changes

@@
 
-<motivation:
-the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
+<motivation: the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
 
@@
 
-total SC t tests
- p = `r Q2tot.sc_t.test_High.Low$p.value`
 **mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
@@
 
-musc and mol ttests
-mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` Mole have significant difference between mouse means for total sc, between the high and low srains
-
-musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`
-the total sc mouse means are not significant for the high and low strains in musc
-
-Additionally we ran similar tests using the short and long bivalent data sets which help to alimeorate potential chromosome size effects.
-
-**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
-(the summary of mouse averages for this subset of ) **
+Molossinus have significant difference between mouse means for total sc, between the high and low strains (mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
+The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+<Additionally we ran similar tests using the short and long bivalent data sets which help to alimeorate potential chromosome size effects.>
 
@@
 mol-short
-
 mol-long
-very few significant comparisons>
+very few significant effects>
 
@@
 
-across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).
+<across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).>
 
@@
 
-(the subspecies effect is tested in M1)
-
 For total SC in M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
 
-strain effects from M3; G, LEW, MSM, SKIVE and TOM are significant
-
-strain effects  (basically all strain effects are significant) (tech effects / error?)
-All strain effects were significant when tested in M2 (Supplemental table X)
-
-**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
-The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
-
-**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
-
-In the short bivalent data set -- analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).
-
-For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW
-G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.
+In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
+
+<For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW, G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.>
 
@@
 
-24sep14_PWD_m1 is lower than the others
-5jul17_MSM_m1 is the low MSM
-
-The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)
+<24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM>
+
+<The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)>
 
@@
 -->
+
+**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
+(the summary of mouse averages for this subset of ) **
+
+**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
+The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
+
+**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
 
@@
 The pattern is confirmed with t.test comparing groups within the same subepcies.
-**IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.**
 
@@
 Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^.
-The Mol\*chrmLength is negative (IN MOL as chrm length got longer - IFD gets smaller -- also in SKIVE\*chrm
-
-For the strain only model,
+The Mol\*chrmLength is negative
+
+<IN MOL as chrm length got longer - IFD gets smaller -- also in SKIVE x chrm>
+
+For the strain only model (M2),
 
@@
 
-The general chromsome Length effect is that
-
-The single fixed effects can be interperted as, with each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
+With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
 All of the high rec strains and MOLF have significantly longer IFD^PER^.
-
 There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
@@
 
-<None of the logistic regression models for ABS or PER IFD lengths are significant, even when just the Musc strains are used.>
-
-Preliminary results from an independent data set indicated that PWD had longer IFDs, which goes against the simple prediction of more COs ~ denser spacing of foci on the same bivalent.
-This also indicated that interference distance may evolved in the house mouse complex.
+ **IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.**
 
@@
 
-<below chunk goes with the above chunk>
+<HAND measure METRICS, below chunk goes with the above chunk>
 ```{r long.biv.plots, echo=FALSE, eval=FALSE}

### 2020-02-24 doc/Results.rmd
slight changes in formating

@@
 
-Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
-
-1. no difference in DOM strains (for SC lengths (CONFIRM IF THERE ARE DIFFERENECES FOR GWRR))
-
+<Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
+1. no difference in DOM strains (for SC lengths )
 2. PWD > SKIVE > the rest
-
-3. Msm > mol
+3. Msm > mol>
 
@@
 
-The skeletonized measures show that the general pattern holds for total SC and the high rec strains.
-However there are a few exceptions -However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel
-
--the SKIVE are within dom and kaz level -- so why does SKIVE have intermediate CO proportions and gwRR
-
-<24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM>
+The skeletonized measures show that the general pattern holds for total SC and the high rec strains. However there are a few exceptions.
+
+<-the SKIVE are within dom and kaz level -- so why does SKIVE have intermediate CO proportions and gwRR>
+
+<!--24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM-->
 
@@
 
-**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
-(the summary of mouse averages for this subset of ) **
+**The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).**
 
@@
 
-I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.
-
-**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$
+<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.
+
+**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$>
 

### 2020-02-25 src/SetupData.Rmd
Finally debugged the kniting part, used here()

@@
 
--rm Peromycus data BivData!!
+- rm Peromycus data BivData!!
 
@@
 
+- try coding i) date image for each cell image, ii) most recent image date for each mouse
+
 ```{r setup, include=FALSE}
@@
 library(kableExtra)
-setwd("~./MLH1repo/")
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+library(here)
 
-#forgot why I'm reading in this metadata file
-meta.data = read.csv("data/clean.Meta.Data.txt",sep = "\t", header=TRUE)
+#knitr::opts_chunk$set(echo = FALSE)
+
+#opts_knit$set(root.dir = "~./MLH1repo/")
+#setwd("~./MLH1repo/")
+
+meta.data = read.csv(here("data", "/clean.Meta.Data.txt"),sep = "\t", header=TRUE)
+
 #remove all MAD (peromyscus) from meta.data
@@
 
-MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE)
+MLH1_data = read.csv(here("data", "/MLH1/AnonData.csv"), header=TRUE)
 
-knitr::opts_chunk$set(echo = FALSE)
+source(here("src", "/CommonFunc_MLH1repo.R") )
+
 ```
@@
 
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 original_DF <- add_mouse(original_DF)
@@
 MLH1_data$fileName <- (MLH1_data$Original.Name)
+
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
@@
 
+```{r Date.imaged, echo=FALSE}
+
+#for issues with the total SC metrics, code in date imaged
+#use the add_mouse function
+
+#mini <- MLH1_data[1:500,]
+
+count=1
+for(i in MLH1_data$Original.Name){
+   # print(i)
+    templist= strsplit(i, split="_")[[1]]
+    c = (templist[1])
+    MLH1_data$image.string[count] <-  c
+    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
+    count= count +1
+}
+#works better if dat is converted out of image
+MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')
+
+#MLH1_data <- add_mouse2(MLH1_data)
+
+mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Random.Name),
+                        #most recent date
+                        oldest = min(date.imaged),
+                        newest = max(date.imaged)
+
+)
+
+```
+
 # Mouse Data
@@
 
-```{r mouse.table, echo=FALSE, warning=FALSE}
-
-#remove juvi mice
-
-#Make a mouse level table -- why are there many old mice without any MLH1 (bc these are mice which were dissected, but not imaged/quantified)
+```{r mouse.table, echo=FALSE}
+#Make a mouse level table
 
@@
                         Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci, na.rm=TRUE),
+                        var = as.numeric(format(round(   var(adj_nMLH1.foci, na.rm=TRUE),3), nsmall=3)),
                         sd   = round(sd(adj_nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                        oldest.image.date = min(date.imaged),
+                        newest.image.date = max(date.imaged)
                         #quality?
 )
+
+#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
@@
 #add DOB to this table?
-mice.frm.meta <- AP_mouse_table.MLH1_HQ[(AP_mouse_table.MLH1_HQ$mouse %in% meta.data$mouse),]
+mice.frm.meta <- AP_mouse_table.MLH1_HQ[as.character(AP_mouse_table.MLH1_HQ$mouse) %in% as.character(meta.data$mouse),]
 
@@
 #during merging -- try to just merge a few cols: 
-AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
+AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data)
+#don't use by, that makes duplicate cols
+# by.y = "mouse", by.x = "mouse", all = TRUE)
 #what are the differences between these two
@@
 #remove cols with bad suffix
-AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x, strain.y,
-                    mouse.1  )  )
+#AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+#                            subsp.y,subsp.x,
+#                    category.y,category.x, strain.x, strain.y,
+#                    mouse.1  )  )
 
@@
 
-#make sure they are all factors
-AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
-AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
-AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
-AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)
+#make sure they are all factors - this was breaking the kniting
+#AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
+#AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
+#AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
+#AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)
 
@@
 old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]
+
+#remove juvi mice via the list
 
@@
 ```{r HetC.DF.4MM, echo=FALSE}
+#this block has so many fing errors
 
@@
 #use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
 #main file, make everything unordered factors
+
+#setwd("~./MLH1repo/src/")
+
+source(here("src","CommonFunc_MLH1repo.R") )
 AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages %>% filter(species == "M.musculus")
 
 #remove the extra merged cols
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
+#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+#                            subsp.y,subsp.x,
+#                    category.y,category.x, strain.x,
+#                    strain.y, mouse.1  )  )
 
 #REMOVE STRAINS WITHOUT FEMALES
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel %>% filter( subsp != "Cast") %>% filter(strain != "AST")
+
+# These are removed in the analysis
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
 #remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
 #remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+#F.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
 
 #Only us the righ age window
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
+#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
 
@@
 #main df with variables coded as chr
-DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+#DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+#DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+#DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
 
 #kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
-
 #only female male reps
@@
 #use the above table to pull the male mouse levels
-male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")
+mlh1male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")
 
@@
 
-male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") &(AP_mouse_table_w.Ages$category.x != "other"),]
+male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex == "male") &(AP_mouse_table_w.Ages$category != "other"),]
 
-all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category.x)+ggtitle("Male Age Distributions")
+all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
 all.ages_plot
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_2.2.20.RData")
+#remove non relevant files -- not sure where they load from
+#BD data and Peromyscus data
+#rm(c(BD_MLH1_data, BDmouse_level_table, MLH1_data.hq) )
+
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
 

### 2020-02-25 src/SetupData.Rmd
small changes

@@
 
-mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Random.Name),
-                        #most recent date
-                        oldest = min(date.imaged),
-                        newest = max(date.imaged)
-
-)
+#mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
+#                        Nmice = length(unique(mouse)),
+#                        Ncells  = length(Random.Name),
+#                        #most recent date
+#                        oldest = min(date.imaged),
+#                        newest = max(date.imaged)
+#)
 

### 2020-02-25 doc/Results.rmd
learn to recognz the this column dont exsist errors.

@@
 
-setwd("~.")
-
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
+setwd("~.")#switch over to here()
+
+#MLH1_data_setup_2.24.20.RData
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
 
@@
 #add Rec group,
-#reclassified SKIVE as intermediate
-Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                               ifelse(grepl("MSM male", Curated_BivData$category), 1,
@@
 
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                               ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
@@
 ```{r MLH1.MALE.TABLE, echo=FALSE}
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+male_MLH1mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                             subsp.y,subsp.x,
@@
 
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+male_MLH1mouse.av <- add_sex(male_MLH1mouse.av)
+male_MLH1mouse.av <- add_strain(male_MLH1mouse.av)
+male_MLH1mouse.av <- add_subsp(male_MLH1mouse.av)
+male_MLH1mouse.av <- add_species(male_MLH1mouse.av)
+
+fomo.male_mouse.av <-  male_MLH1mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
 
@@
 
+#this blck is duplicated in setupData.rm
+
 #https://rcompanion.org/handbook/G_03.html
@@
 #variance is still chr
-AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+#AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
 
 #use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
+#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
+#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
+#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
 #main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
+#AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
 
@@
 #remove the extra merged cols
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-#DONT NEED CATEGORY
+#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
+#                            subsp.y,subsp.x,
+#                    category.y,category.x, strain.x,
+#                    strain.y, mouse.1  )  )
 
 #REMOVE EXTRA STRAINS
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
 #remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
 #remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
+#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
+
+#DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
+
+#AP_mouse_table excluide Qual5 cells
+DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
+
+DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
 
@@
 DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
 
@@
 #norm.first.foci.pos
-
 Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
@@
 
+<put in plot for total SC and mean gwRR count>
+
 <motivation: the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
 
-<pattern for the single bivalent, pos correlation>
-The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
+<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
 Mainly that 1CO are shorter in the high rec strains then low rec strains.
@@
 This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
-They have a lower probability of 2COs below a certain SC length threshold.
+They have a lower probability of 2COs below a certain SC length threshold.>
 
@@
 
-<!-- basic stats -->
-
+<!-- basic stats
 `r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
@@
 
-<NEW STRAINS INTRODUCED NOT IN PREV DATA>
-
-this section / these tests follow up with those made in Q1 which found that the male musc and mol total-SC were signifcantly higher compared to the male Dom total SC measures.
-These test include additional measues from strains which only had male observations.
-
--reduced bivalent data set (long and short) is made up of binned bivalents from X cells and X mice.
-(Since comparing within sex, there are no sex chromosomes differences and the long bivalent dataset can be used.
-
-<!-- place set ups for total SC, short biv and long biv here -->
+ place set ups for total SC, short biv and long biv here -->
 
@@
 
-```{r Q2_tot.sc_plot, echo=FALSE}
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
 
@@
 
-I made a series of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromosome (positive correlation of SC length and CO number).
-Yet, given the complications from the chromosome proportions outlined above, comparing the bivalents across chromosome classes is complicated
-
-<!-- t tests-->
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE}
+
+#make plot of mouse averages for MLH1 and total SC
+
+bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
+
+X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+
+X.by.texas
+
+Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+#not sure what my expectations for this plot are,
+
+#there might be an issue with the algorithm performance 
+#given the parrelle diagonal lines -- 
+#this might be due to the dates imaged
+
+```
+
 <!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
@@
 
-The average SC lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confounded by the different proportions across strains.
-
 **mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
@@
 The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-<Additionally we ran similar tests using the short and long bivalent data sets which help to alimeorate potential chromosome size effects.>
-
-<within subspecies
-musc-short
-
-musc-long
-
-mol-short
-mol-long
-very few significant effects>
 
@@
 
-For total SC in M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
+For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
 
 In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+
 <And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
@@
 
-The skeletonized measures show that the general pattern holds for total SC and the high rec strains. However there are a few exceptions.
-
-<-the SKIVE are within dom and kaz level -- so why does SKIVE have intermediate CO proportions and gwRR>
-
 <!--24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM-->
-
-<The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)>
-
-<lots of variation across SC lengths -- doesn't fit clean expectation or signal>
-<balancing the interpertation of general positive correlation of CO number and SC length (SC length drives CO number) --- yet the pattern breaks down -it's not completely predictive -- /at finer levels?>  <noise/mouse effects/chromosome identity effects>
 
@@
 
-(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
 
@@
 
-M2 / for the second model ...
+M2
 
@@
 
-**males from WSB and MOLF strains have 'strang'/unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation**
-
-**There are significant difference across the 1CO positions, yet they don't match /fit any of the predictions. there are also complications due to the chromsome class proportions that **
+males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
+
+There are significant difference across the 1CO positions, yet they don't match fit any of the predictions. there are also complications due to the chromsome class proportions that
 
@@
 #summary(Log.glmQ2_M2.1COnrm_Mol)
-```
-
-# Q2 IFD
+
+```
+
+# Q2 Evolution of interference is associated with genome wide recombination rate evolution
 
@@
   ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none", axis.text.x = element_blank())+ylim(c(0,1.1))
+
 biv.IFDs_PER_by.group
-```
-
-<Instead of the long and short biva data, the total 2CO bivalent dataset is used to make models.>
+
+```
 
@@
 
-The mouse averages for IFDs were significantly longer in the high rec groups ( p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
 The pattern is confirmed with t.test comparing groups within the same subepcies.
@@
 
-**model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
-
-**model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
+**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
+
+**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
 
@@
 
-We ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length.
+<!-- we ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length.
 
@@
 The Mol\*chrmLength is negative
-
-<IN MOL as chrm length got longer - IFD gets smaller -- also in SKIVE x chrm>
-
-For the strain only model (M2),
-
-- PWD, MSM, MOLF, SKIVE are the only significant strain effects.
-
+<PWD, MSM, MOLF, SKIVE are the only significant strain effects.
 - all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
-
 With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
@@
 There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
+-->
+
+```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
+
+#calculate the rates of 'close foci' (<30%)
+#examine the changes in the distribution closer.
+
+#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
+
+close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                      n.close.foci = length(which(IFD1_PER <= .3)),
+
+                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                      rate.close.foci = n.close.foci / N.obs
+)
+
+close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
+
+close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("strain"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER),
+
+                      inverse.close.foci = length(which(IFD1_PER > .3)),  
+                      n.close.foci = length(which(IFD1_PER < .3)),
+                      rate.close.foci = n.close.foci / N.obs,
+
+                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
+
+)
+#the
+
+show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
+
+#add rec.group
+close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
+
+#POOL all short and 
+#chi square test
+high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
+                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
+#p-value = 0.03, slight dif tween 3 high rec strains
+
+#don't need to run chi square for this pattern, since it's enough to
+
+```
+
+We determined that the main difference in IFD disbribution across these groups is an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC length in low rec strains.
+The rate of these close foci range from 8.2% (G) to 16% (KAZ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for skive, MSM and PWD respectively).
+<no chi sqaure test see above>
 
@@
 
-```{r Q2.IFD.longbiv_glm, echo=FALSE}
+```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
 
@@
 
- **IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.**
+ **(unlike total SC length and 1CO pattern, IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.
+However, they go agaisnt a standard perdiction for the relationship between the chromosome level recombination landscape and genome wide recombiantion rate.**
 

### 2020-02-25 doc/Results.rmd
small change to setup, more cleaning ect.

@@
 ```{r MLH1.FEMALE.TABLE, echo=FALSE}
-
 female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
@@
 ```{r MLH1.MALE.TABLE, echo=FALSE}
-male_MLH1mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
+#AP_mouse_table.MLH1_HQ
+
+#find the table that includes the las image data
+male_MLH1mouse.av <- AP_mouse_table_w.Ages
 
@@
 
+#investigating potential skew / technical effects in the algorithm
+
 Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
@@
 
+median(bep$oldest.image.date) #"2017-08-03"
 #there might be an issue with the algorithm performance 
 #given the parrelle diagonal lines -- 
-#this might be due to the dates imaged
+#this might be due to the dates imaged  2017-02-07
+
+bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
+
+Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
+
+#there's evidence that the algorithm might have some biased effects -- 
+#that will be investigated later.
 

### 2020-02-25 src/SetupData.Rmd
small change to setup, more cleaning ect.

@@
 AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data)
+
+#this should be transfering the image dataes!
+
 #don't use by, that makes duplicate cols

### 2020-02-25 doc/Results.rmd
refining order of sections
improving points for main outline of begining

@@
 library(reshape2)
+library(patchwork)
 
@@
 
-# Data Set Up Notes
+<!-- Data Set Up Notes -->
 
@@
 
-Fig1_v1_strain
+```
+
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+#table( droplevels(female_mouse.avs_for.analysis$strain) )
+#there are cast strains too
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+#female.specific.plot
+
+```
+
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
+#try customize the plot size
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")
+
+just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#table( droplevels(just.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+#fomo.male_mouse.av
+
+male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
+male.plot
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+```
+
+```{r three.plots, echo=FALSE, warning=FALSE}
+#merge the 3 plots
+#try patchwork
+
+#Fig1_v1_strain + female.specific.plot + male.plot
+
+#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
+Fig1_v1_strain / (female.specific.plot + male.plot )
+
+#(p1 | p2 | p3)/p4
+
 ```
@@
 
-kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
+main.table %>% filter(sex == "female")
+
+main.table %>% filter(sex == "male")
+
+G.HETC.val #1.2
+LEW.HETC.val  #1
+WSB.HETC.val  #1
+KAZ.HETC.val  #1.1
+
+SKIVE.HETC.val # 0.96
+PWD.HETC.val  #0.88
+MSM.HETC.val #0.9
+#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
 
@@
 
-We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from thress subspecies within the house mouse species complex;  _M.
+CALCULATE THE MAGNITUDES of HETC
+
+We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M.
 m.
@@
 MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
+<!-- state the average number of mice-->
 The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.
@@
 <!--
-more basic stats, error, comparison to other measures
-
-transition first aragraph to strain specific
-Main points
-1. -female strain means are v close  -male strain means are more varied!!
--Comparisons to references / literatures
--house mouse close to minimum of 1  obligate CO per chrm
-
-2. The largest female difference is G, 1.07X  the largest male difference is PWD and MSM which are 1.3 and 1.2 respectively.  
+1. Fig1A  magnitude and direction of heterochiasmy;
+-general patterns i) female higher and ii) low degree of hetC
+(all strain means are above the expected min)
+
+2. Fig1B and C mouse means and sex specific patterns; 
+female distributions more random / uniform
+male distributions are more clustered, low and high
+
 -->
@@
 <!-- transition from first part into female -->
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-#table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Female mouse av.
-MLH1 count per cell ")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
-  theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))
-
-#female.specific.plot
-
-musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-#Mol
-Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
-
-Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-#Dom version 
-Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
-
-Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-```
 
@@
 <!-- male section -->
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")
-
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-#fomo.male_mouse.av
-
-male.plot <- ggplot(just.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Male mouse mean MLH1 per cell")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-male.plot
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-#Dom version 
-Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-```
 
@@
 
-Intial examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.
-In females, _domesticus^G^_, _musculus^MSM^_, and _domesticu^LEW^_ had the greatest differences in means compared to other strains, `r G.female.diff`, `r round(LEW.female.diff,4)`, and
-`r MSM.female.diff` respectively.
-While in the male data the strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`, `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.
+Examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.
+
+<!-- female strain means are have more variation in a random way. The range of genome wide rec strains is X to X, -- no clustering of strain means
+
+-->
+
+In females the strains with the greatest difference
+in mean gwnome wide recombiantion rate was `r G.female.diff` ( _domesticus^G^_) more compared to other strains. 
+<!-- _domesticus^LEW^_  (`r round(LEW.female.diff,4)`),
+and _musculus^MSM^_  (`r MSM.female.diff` ),-->
+
+<!-- male variation is a greater magnitude, but divides into 2 groups, magjority of strain means are around 23, while PWD and MSM are near 30, skive is intermediate-->
+
+While in the male data (most strain averages are close to 23 (Figure1).
+
+the genome wide recombination rate for had a much larger range, up to `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )` ( _molossinus^MSM^_ ).
+
+_musculus^PWD^_ ,and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,  and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.
 
@@
 
-The musculus and molossinus subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
+The _musculus_ and _molossinus_ subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
 
@@
 
+<from the means the expected number of COs per chroosomes ranges from X to X we confirm these numbers Can't have fractions of COs>
+
 In order to decompose the cell wide rate, we decided to look at the proportion of chromosome classes (ie by the number crossovers on the same bivalent).
@@
 
+## Adjusting for XX
+
+(rethink where this section should go)
+new outline
+
+1. illustrate problem(affects mostly SC length)
+
+2. Expected impact on sex comparisons, estimated effect size of the X  
+3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
+
+4. 19 female, 19male, 20female 20 male
+
+The female mouse averages should have adjustments for the XX.
+working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
+Subtract this amount from the female mouse averages ...
+This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
+
+- Of all female single bivalents observations, 5% are XX (1 of 20).
+
+- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
+
+- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
+
+I think a formula something like this can be applied to adjust for XX
+
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
+
 # short transition
@@
 X.by.texas
+```
+
+```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
 
@@
 
-# Adjusting for XX
-
-(rethink where this section should go)
-new outline
-
-1. illustrate problem(affects mostly SC length)
-
-2. Expected impact on sex comparisons, estimated effect size of the X  
-3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
-
-4. 19 female, 19male, 20female 20 male
-
-The female mouse averages should have adjustments for the XX.
-working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
-Subtract this amount from the female mouse averages ...
-This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
-
-- Of all female single bivalents observations, 5% are XX (1 of 20).
-
-- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
-
-- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
-
-I think a formula something like this can be applied to adjust for XX
-
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
-
-# Whole Cell Manual Comparison
-
-```{r XX.adj_manual.whole.cell, echo=FALSE}
+<!-- Whole Cell Manual Comparison -->
+
+```{r XX.adj_manual.whole.cell, echo=FALSE, eval=FALSE}
 
@@
 
-#### Automated BivData Comparison
+<!-- Automated BivData Comparison -->
 
@@
 
-table(female.Bivdata_comp$strain)
+#table(female.Bivdata_comp$strain)
 
@@
 
-# Deleted notes
+<!-- Deleted notes -->
 

### 2020-02-25 junk_text.R

@@ -3,7 +3,34 @@
 
 
 
+#from sex specific plots
 
+musc.female.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+#Mol
+Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
+
+Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+
+#Dom version 
+Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
+
+Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom female of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+
+
+#make the Mol and
+Mol.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+
+
+
+#Dom version 
+Dom.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
 
 
 

### 2020-02-25 doc/Results.rmd
starting on evo framework section

@@
 
-```{r juvimice.list, echo=FALSE}
-juvi_mice <- c(
-  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
-  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
-  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
-
-  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
-   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
-  )
-
-#REMOVE JUVI mice
-#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
-#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
-
-```
-
-```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
-
-#this block should be in the data set up
-#the mouse averages will be used
-
-#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
-cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
-
-#cleanCurated_BivData
-#removed the no hand foci count
-
-mouse.avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-mouse.avs_4MM <- add_sex(mouse.avs_4MM)
-mouse.avs_4MM <- add_strain(mouse.avs_4MM)
-mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-mouse.avs_4MM <- add_category(mouse.avs_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
-mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
-mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
-
-#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-#code if males are high or low?
-
-#REMOVE JUVI mice
-juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
-mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
-
-#assign high and low
-#divide musc males into groups
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
-
-mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
-
-```
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#add this to setup
-
-#remove female, 
-#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
-
-#Curated_BivData_male
-
-male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-male.av_4MM <- add_sex(male.av_4MM)
-male.av_4MM <- add_strain(male.av_4MM)
-male.av_4MM <- add_subsp(male.av_4MM)
-male.av_4MM <- add_category(male.av_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
-male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
-male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-
-#REMOVE JUVI mice
-juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
-mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
-
-```
-
-```{r totSC_setup, echo=FALSE}
-#SC.skel
-#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included
-
-SC.skel.OG <- SC.skel #5683
-
-#MERGEING MLH1 with SKEL
-MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-
-MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
-#5110,
-
-OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
-
-#check for DUPs, #check for duplicatations of the random name
-MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
-tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
-
-MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
-
-MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
-MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
-#table(MLH1.merge.Skel$category)
-
-#remove outliers
-MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
-
-```
-
-```{r totSC.viz4outliers, echo=FALSE}
-
-# -- EXAMINE the ggplot scatter plots for outliers
-tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
-
-#round 1
-SC.skel.outliers_round1 <- c(
-  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
-  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
-            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
-
-    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
-                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
-              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
-          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",
-
-  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
-          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",
-
-       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
-                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",
-
-   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
-   "6may15_17apr15_PWD_f1_sp1_7_rev",
-
-      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
-            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",
-
-    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
-    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
-  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
- "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
- "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
- "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
-"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,
-
-  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
-    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
-
-# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
-"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
-
-#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
-"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
-
-#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
-"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
-"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
-"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
-
-#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
-"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
- "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
- "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
- "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
- "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
- "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
-"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
-"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
- "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
- "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
- "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
- "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
-
-MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
-
-remove.mice_round2 <- c(
-    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
-    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
-
-    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
-  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
-"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
-  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",
-
-  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
-  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
-"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
-"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
-
-# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
-"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
-"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
-
-#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
-"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
-#SPIC male  x
-#SPIC female  x
-#TOM x
-#AST x
-#CZECH male x
-
-#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
- "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
-#SPRET female X
-#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
-"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
-)
-#3remove.mice_round2
-
-MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
-MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
-
-all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
-```
-
-```{r LONG.SHORT.biv_setup, echo=FALSE}
-
-#FROM q1
-#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#from Q2 IFD
-#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-##long biv
-#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-#FROM q1 
-REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO  
-                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                         )
-
-REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
-
-REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
-REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
-REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
-                                           0)))
-
-REAL.short_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1, 0))
-
-###LONG BIV
-
-REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                       n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-                         )
-
-REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
-
-REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
-REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
-REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
-                                           0)))
-
-REAL.long_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1, 0))
-
-#find which one 
-```
-
-<!-- 
-two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
--->
-
-```{r tot.skel.table, echo=FALSE}
-
-totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
-
-#unorder the factors
-totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
-totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
-totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
-totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
-
-OG.totSC.mouse.av.table <- totSC.mouse.av.table
-
-#REMOVE JUVI mice
-tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
-total SC area")
-
-mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-```
-
-```{r, tot.SC.setup, echo = FALSE}
-
-totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
-                 .(sex,strain),
-                 summarize, 
-                 ncells = sum(ncells),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(mean.skel, na.rm = TRUE),
-                 mean.bin =  mean(mean.bin, na.rm = TRUE)
-
-)
-
-totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
-
-```
-
-<!--
-From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
-
-the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
-
-INSERT TABLE FOR THE TOTAL SC means ect
-
-The difference between
-
-bin is short for binary, it reports the binarized area of the SC
-
-skel -- takes the skeletonized midline of the bin objects
--->
-# Main Figure
-
-```{r table.for.plot, echo=FALSE, warning=FALSE}
-#why do I add in a different table?
-
-#this move this to the setup file
-#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed
-
-#pooled
-strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  =length(nMLH1.foci),
-                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(nMLH1.foci, na.rm = TRUE),
-                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
-                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
-                       #Standard error (mice vs cells)
-                        se  = round(sd / sqrt(Ncells), 3)
-
-                       #range_mouse_means = range()
-                   #consider adding crhm proportions?
-    )
-
-strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
-                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")
-
-#adjust order of plot here
-strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
-     "MSM", "MOLF",   
-  "PWD", "SKIVE",  "KAZ",
-  "WSB","G","LEW","PERC",
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-```
-
-```{r Fig1, echo=FALSE, warning=FALSE}
-
-#might need to nudge the X values
-
-Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-#remove y labfor male
-
-   #segments
-#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
-  #Musc
-#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
-  #mol
-#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
-  #cast
-#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
-  #outgroups
- # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
-  #label
-
- #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
-   #        y = c(18.5,18.5,33,18.5, 18.5),
-
-  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
-    #       fontface = 'italic', hjust = 0.5, size=3.5)
-
-```
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-#table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
-  theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-#female.specific.plot
-
-```
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")
-
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-#fomo.male_mouse.av
-
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-male.plot
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-```
-
-```{r three.plots, echo=FALSE, warning=FALSE}
-#merge the 3 plots
-#try patchwork
-
-#Fig1_v1_strain + female.specific.plot + male.plot
-
-#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
-Fig1_v1_strain / (female.specific.plot + male.plot )
-
-#(p1 | p2 | p3)/p4
-
-```
-
-This is a draft of the main figure. Outgroups are not included in the analysis below.
-
-<!--  insert Details of data set cleaning etc is in SetupData.rmd
-Effects from different factors are reviewed in Effects.Rmd -->
-
-```{r main.fig1s, echo=FALSE, include=FALSE}
-strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
-
-Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point( size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-
-#strain.table.HQ_Fig.og
-
-strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")
-
-strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
-     "MSM", "MOLF",   
-  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
-  "WSB","G","LEW"), order=T )
-
-Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-
-```
-
-```{r basic.numbers, echo=FALSE, eval=FALSE}
-
-kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-
-kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-```
-
-```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
-#str(AP_mouse_table_w.Ages)
-
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-#remove sex.x == na, and strain.x=na or F1
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
-
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
-#plot variance and mean (facet sex)
-
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
-
-VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
-
-cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
-
-str(AP_mouse_table.HQ)
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
-#[1] 0.5750399
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
-#0.321601
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
-# 0.3515544
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
-#-0.03713536
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
-# p-value = 8.623e-09
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
-
-#remove mice with fewer than 10 cells
-
-AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
-
-VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
-
-cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
-
-cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
-    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-
-#why did I male just a female table?
-female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-
-female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-                        #quality?
-)
-```
-
-# Genome wide recombiantion rate estimates
-
-```{r MLH1.summ.stats, echo=FALSE}
-
-#strain.table.HQ_Fig.og  all strains
-#strain.table.HQ_Fig  extra strains removed
-
-main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
-
-main.table <- add_subsp(main.table)
-main.table <- add_species(main.table)
-
-main.table %>% filter(sex == "female")
-
-main.table %>% filter(sex == "male")
-
-G.HETC.val #1.2
-LEW.HETC.val  #1
-WSB.HETC.val  #1
-KAZ.HETC.val  #1.1
-
-SKIVE.HETC.val # 0.96
-PWD.HETC.val  #0.88
-MSM.HETC.val #0.9
-#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
-
-#strain.table.HQ_Fig
-
-#sum( (main.table  %>% filter(sex == "male"))$Ncells)
-#the kable version is bulkier than the other versaion
-```
-
-CALCULATE THE MAGNITUDES of HETC
-
-We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M.
-m.
-domesticus,_ _M.
-m.
-musculus_.
-and _M.
-m.
-molossinus_.
-MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
-<!-- state the average number of mice-->
-The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.
-
-<!--
-1. Fig1A  magnitude and direction of heterochiasmy;
--general patterns i) female higher and ii) low degree of hetC
-(all strain means are above the expected min)
-
-2. Fig1B and C mouse means and sex specific patterns; 
-female distributions more random / uniform
-male distributions are more clustered, low and high
-
--->
-
-<!-- transition from first part into female -->
-
-<!-- female section -->
-
-```{r female.strain.table, echo=FALSE}
-
-female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-
-                        Nmice = length(unique(mouse)),
-
-                        total.cells  = sum(Ncells, na.rm = TRUE),
-
-                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#F.strain.table
-
-#F.strain.table
-
-# I don't know why most of the table is blank
-```
-
-```{r female.dif, echo=FALSE}
-
-#how much greater are the high strains?
-G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
-
-LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
-mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
-
-#how much greater is MSM female
-MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
-
-sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
-mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
-
-G.female.diff <- (mean.G.female / mean.all.other)
-LEW.female.diff <- (mean.LEW.female / mean.all.other)
-MSM.female.diff <- (mean.MSM.female / mean.all.other)
-```
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-#variance plots
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-#Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
-
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-```
-
-<!--
-The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (glm; p= ).
- -->
-
-```{r FEMALE.VAR.MODELS, echo=FALSE}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
-
-lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
-lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
-
-<!-- male section -->
-
-```{r male.strain.table, echo=FALSE}
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#M.strain.table
-#same issue as the female table .. ugh
-```
-
-```{r male.diff, echo=FALSE}
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /  
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.male.VAR.plot
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.male.VAR.plot
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.male.VAR.plot
-```
-
-```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-#summry.male.strain.effects <- summary(glm.male.strain.subsp)
-#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#1. determine which strain effects are sig
-
-#M2
-MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
-sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-```
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-```
-
-Examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.
-
-<!-- female strain means are have more variation in a random way. The range of genome wide rec strains is X to X, -- no clustering of strain means
-
--->
-
-In females the strains with the greatest difference
-in mean gwnome wide recombiantion rate was `r G.female.diff` ( _domesticus^G^_) more compared to other strains. 
-<!-- _domesticus^LEW^_  (`r round(LEW.female.diff,4)`),
-and _musculus^MSM^_  (`r MSM.female.diff` ),-->
-
-<!-- male variation is a greater magnitude, but divides into 2 groups, magjority of strain means are around 23, while PWD and MSM are near 30, skive is intermediate-->
-
-While in the male data (most strain averages are close to 23 (Figure1).
-
-the genome wide recombination rate for had a much larger range, up to `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )` ( _molossinus^MSM^_ ).
-
-_musculus^PWD^_ ,and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,  and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.
-
-<notabley these strains come from Musc and Mol -- Mol natural hybrid of musculus and castaneus. this comments out text?>
-
-The _musculus_ and _molossinus_ subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
-
-<!-- dom, low variation and close to the expected in
-within the male means, there is a low degree of strain varaince in _Domesticus_ subspecies with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
--->
-
-<!-- ranges of mouse means
-In the _domesticus_ subspecies the ranges of mouse means are 
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )`
-in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
--->
-
-<!--
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$ -->
-
-<!--
-The strain means for females were pretty similar.
-For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively.
-In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively.
-In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively.
--->
-
-While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively.
-In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively.
-In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.
-
-<!--(the figure is made from pooling MLH1 cells)
-the male 
-- the ranges of mouse means are displayed in Sup table X
-- How the means were calculated
--The male gwRR is adjusted for the PAR
--The means for each strains are report in Table 1
--Normality was assessed for the MLH1 counts (include in Supplement histograms)
--->
-
-# Evolutionary Framework
-
-<to better understand this varation in RR within an evolutionary framework>
-
-In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
-
-<!--
-(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
-
-Check the DF of the number cited above
-This is the break down of mice used within the Mixed model
-`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
-
-add number of cells
-`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
- $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-Setting up the data set for using the mixed model (base data set, HQ data set)
-
-- only Mus musculus strains with sex matched observations  
-- Male gwRR have been adjusted for the PAR
-- Q5 excluded
-- All the effects are un-ordered factors  
-- male ages within 5 - 20 weeks
-- Molossinus is coded as a subsp level
-
-subspecies levels: 3, Musc, Dom and Molossinus
-strain levels: Dom: 3, Musc: 3  Mol: 2
-sex: 2 female male  
--->
-
-**M1.Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-Our model, M1., was designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.
-
-```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3, message=FALSE}
-#THIS IS THE NEW MIXED MODEL
-#should be able to use the loaded 
-#DF.HetC.MixedModel.HQ
-
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
-
-MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
-sex.results =  MLH1.MM_sex, 
-               interaction.pval = MLH1.MM_interaction,  rand.pval = 
-MLH1.MM_strain)
-```
-
-All of the effects within the mixed model were significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
-
-```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
-#strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
-
-summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
-
-#table of just coeff anf pvalue
-#round(summ$coefficients[,c(1,4)],5 )
-
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-```
-
-```{r MLH1.glm.M3, echo=FALSE, digits = 3}
-
-MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-
-summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
-
-#round(summ.strain.fixed$coefficients[,c(1,4)],5)
-
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-```
-
- **M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-
- **M3. LInear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
-
-Following up with linear models which tested for (subsecies M2), and strain effects alone (M3)
-both showed similar -- results -- **so we just report those from M3. also just report effects which were consisantly significant across both M2 and M3**
-
-<they also have sig strain effects .. PWD strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`>
-
-which identified several signifcant sex by strain interaction effects, ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
-and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`
-)
-
-<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
-These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.
-
-<M3 also has a few less significant effects>
-<not the reverse coefficient indexing -- this will give the name >
-
-Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
-These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
-
-_domesticus^G^_ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
-and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`
-
-<!--
-also g strain effect/ dom^G^ strain effect 
-`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[12,][4]` 
-In both models strain by sex interaction effects were significant for (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
--->
-<!-- glms for sex specific data also run -->
-
-A similar glm models (M4) was run on the sex specific data set, (just containing subspecies, strain and the interaction).
-
- **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$
-
-For the male specific data, three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
-(with effect sizes ranging from 5, 7, and 2 foci respectively).
-Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
-
-While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
-
-<!-- transition gwRR to within mouse variance -->
-
-## Within Mouse Variance in CO Count per Cell
-
-In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).
-
-Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
-
-Following the same basic framework we used for analyzing mean number of MLH1 foci per cell, we applied a set of mixed and linear models -- to test effects (sex differences for the within mouse variance in MLH1 foci per cell).
-
- **M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
- **M6. Linear Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex * strain + \varepsilon$$
-
- **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
-
-#Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-<!--
-These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
-var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
--->
-
-For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
-Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
-background noise).
-For this reasons, we replicated the model analysis using a subset of cells with higher quality scores.
-These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).
-Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
-While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`) for the model with the full dataset.
-While in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
-For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell.
-This limits the power for detecting significance differences across groups, <but the main/general pattern, is that females have significantly higher within mouse variance compared to males.
-This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.
-
-<!--
-motivation / background: maybe checkpoint / regulation
-limitations: observations at the mouse level, decreased power to estiamate the variance --
-
-$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
-<!--
-RW's main qtl paper used the within animal variance,
-
-@article{wang2017genetics,
-  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
-  author={Wang, Richard J and Payseur, Bret A},
-  journal={Genetics},
-  volume={206},
-  number={4},
-  pages={1841--1852},
-  year={2017},
-  publisher={Genetics Soc America}
-}
-references for phenotypic variance connected to robustness to environment, increase in developmental consistency
-
-Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
-phenotypic variance leading to adaptation - 
-(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
-Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
-meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
--->
-
-< SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE >
-
-# Variation in DSB Number Per Cell
-
-```{r DMC1.setup, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "L"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "Z"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
-#            row.names = FALSE)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-```
-
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
-#think of better color scheme
-#make the jitter col thinner or make the stacked dots.
-#remember that box-plots add outlier dots.
-#change y labels
-
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
-#change order
-
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.boxplot
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-  ggsignif::geom_signif(
-    data=anno_df_stage, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
-```
-
-```{r plot.t.test, echo=FALSE, include=FALSE}
-
-  #this will add neat bars, but still need to figure out what groups ect it's actually using
-#THE stat_compare function adds a arm thing and pvalue for 
-#  stat_compare_means(comparisons = my_comparisons)
-
-#status - legend, expression y-axis, facet, facet_wrap SYMBOL
-#legend - stage. y=total count, FACet strain
-
-#add column to distinguish high vs low rec strains
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
-DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
-
-# annotation table with adjusted pvals and y-position of the labels
-
-anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=stage), position=position_jitterdodge()) + 
-  facet_wrap(~stage, nrow=1) + 
-  ggsignif::geom_signif(
-    data=anno_df, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#example for ggplot+ test from here
-#https://github.com/kassambara/ggpubr/issues/65
-
-#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
-
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-
-#unorder strain
-DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
-
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
-z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
-```
-
-```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-#attach this mini DF to the DMC1 data
-
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-
-#DMC1.mean.table
-DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
-
-colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
-                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
-
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
-<!--
-explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
--->
-
-In an attempt to locatize the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days).
-A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
-One mouse was sampled for each strain.
-
-For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = ), which reflects the expected patterns across the timing of meiosis.
-
-Confirming the prediction that (more crossovers are associated with more initial DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
-
-The differences in DMC1 foci were not significant for zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
-<!-- which maybe be attributed to noise effects
-
-The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->
-
-The ratios of average number of MLH1 and DMC1 foci per cell -- have  .... biological significance
-
-The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
-
-These results indicate that the changes -- which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
-
-<!--
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M.
-m.
-domesticus^WSB^_ , _M.
-m.
-domesticus^G^_, _M.
-m.
-molossinus^MSM^_, and  _M.
-m.
-musculus^PWD^_ respecitively.
--->
-
-```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
-
-#Subsampling code -- don't report
-
-#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
-Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
-Permutations were done for both cell stages.
-The actual p value is in red.
-
-#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
-
-#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
-# the 2 groups
-
-#make the 2 pools
-High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-
-Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-
-#permute drawing / subsampling and performing the t-test
-#put the p.value results into a DF!
-
-#permut.SC.cv.diff  <- as.data.frame(
-#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
-
-#permuting samples -- and re-licating t-tests
-
-#nrep <- 1000
-
-#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
-bunchOpvals.L <- replicate(1000, 
-    # t.test
-    t.test( sample(High.L.pool, 25, replace = FALSE),
-            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
-          )
-bunchOpvals.L <- as.data.frame(bunchOpvals.L)
-bunchOpvals.L$stage <- "L"
-
-bunchOpvals.Z <- replicate(1000, 
-    # t.test
-    t.test( sample(High.Z.pool, 25, replace = FALSE),
-            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
-          )
-
-bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
-bunchOpvals.Z$stage <- "Z"
-
-#could merge the lists of p-values
-replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
-#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
-
-#ASSUMPTIONS; normality, equal variance
-
-#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-#plot.pvals.L
-#plot.pvals.Z
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
-```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#these are histogram plots for testing Normality shape -- move to supplemenatry section
-
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-DMC1.histogram
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-
-#The counts look close to normal.
-#Potentially figure 2. Using a function to #display the p-vales for mean comparison
-
-```
-
-<!--
-interpertation notes, 
-the number of DMC1 across stages (L and Z stages) are driven by different equations
-
-Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
-
-Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
-
-*the repair decision can happen at both stages and remove DMC1.
-formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
-
-back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
-
-expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
-OR
-
-variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
--Baier et al 2014 supports this model (Chmr axis / DNA loop model)
--->
-
-<!--
-# Rapid male specific evolution
-
-list out the points which support rapid male specific evolution for MLH1 rates.
-
-number above bars are the raw counts for bivalent classes.
-
-Potential figure 2 (eventually). Try to make the figure taller.
--- add more data to make sample sizes more equal.
--- add space to indicate where female data is missing
-
-SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
-What's up with Lew females? there's lots of 2COs, also WSB female.
-
-This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
-If these data are random samples of bivalents, the X
-
-ToDo:
-- within strain across sex,
-- males, across high and low groups
-- look at mouse level proportions
--->
-
-# Chromosome Class Proportions
-
-```{r chrm_prop_data, echo=FALSE, warning=FALSE}
-#shut off eval for now because of wd / path error
-#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#try adding the BivData version of the proportions
-
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
-
-                 n0CO = sum(n0CO, na.rm = TRUE),
-                  n1CO = sum(n1CO,na.rm = TRUE),
-                  n2CO = sum(n2CO,na.rm = TRUE),
-                  n3CO = sum(n3CO,na.rm = TRUE),
-                  n4CO = sum(n4CO,na.rm = TRUE),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#this is the table that needs melting
-#but is also easy to redo by hand
-
-#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
-```
-
-```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
-
-#add subspecies to these values
-melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
-
-#add the subsps
-
-melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
-
-melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
-
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
-
-#need a second DF / the first df
-#+ annotate() text 
-#annotate()text
-
-#proportion_stats_category,  tot_biv, ncells
-#add number of cells and bivalents in the corner
-
-#chrm.class.barplot
-```
-
-```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
-
-#prop.test(x, y)  x number of successes, y total  
-#strain sums of n
-#look up continuity correction
-
-#mouse.level.BivData_table <-- has more male strains
-#Mouse.Table_BivData_4MM <- only sex matched strains
-
-#perform test across strain 
-#removed all lines with no hand.foci.count entry
-Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
-
-                  nbiv = sum(tot.biv ),              
-                 n0CO = sum(n0CO ),
-                  n1CO = sum(n1CO ),
-                 n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO ),
-                 n4CO = sum(n4CO )
-                          )
-
-Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
-
-for(g in 1:length(Chrm.prop_strain_table$nbiv)){
-
-  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
-                                     Chrm.prop_strain_table$n1CO[g],
-                                     Chrm.prop_strain_table$n2CO[g],
-                                     Chrm.prop_strain_table$n3CO[g] ))  
-}
-
-Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
-#there are still some wacky entries -- but much fewer than before
-
-chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
-
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
-#the pvalues for 0 are not sig dif
-#the p values for 1 and 2CO are sig diff
-
-#seperate by subsp
-chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
-chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
-chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
-
-Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
-Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
-Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
-Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
-#WSB has fewer 2COs than other 3 strains
-
-Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
-Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
-Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
-#KAZ has slightly more 1CO than the other two
-
-#high.female_G.LEW.
-chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
-
-high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
-high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
-high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
-
-#Low.female.G.LEW.
-chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
-
-low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
-low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
-low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
-
-#all females
-
-chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
-
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
-#the pvals for these 3 class are sig dif
-
-#seperate by subsp
-chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
-chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
-chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
-
-Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
-Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
-Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
-#non of the Dom proportions are different
-
-Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
-Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
-Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
-#all v different
-
-Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
-Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
-Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
-
-#high vs low
-chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
-                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
-                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
-
-chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
-chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
-
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
-
-Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
-Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
-Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
-#the low strains do vary slightly in their 1CO and 2CO proportions
-
-#includes SKIVE
-Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
-Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
-Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
-#all very sig p values
-
-High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
-#these proportions are v different
-
-#no SKIVE
-High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
-High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
-High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
-#MSM and PWD only slightly sig dif 2CO
-
-#MSM-MOLF
-MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-#PWD-KAZ
-PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-```
-
-```{r Chrm.props.plot2, echo=FALSE}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
-
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
-
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
-
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-#no filename for add mouse, strain subsp
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#can't apply function use this
-chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
-             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
-                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
-                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
-
-          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
-              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
-
-                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
-                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
-            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
-                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
-                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
-                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
-                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
-                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
-
-             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
-               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
-
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
-                                                                 "Musc",
-                                                                 "Mol",
-                                                                 "Spretus","Spic",
-                                                                 "other") )
-#ggplot
-
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
-#chrm.class.barplot
-chrm.class.barplot
-```
-
-```{r Chrm.Prop.misc.calulations, echo=FALSE}
-
-total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-
-#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
-#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
-```
-
-<!--
-- This is motivation to investigate more traits at the chromosome level
-1.	Basic stats - bivalent data numbers 
-2.	Rang of co per chrm, mostly 1co 2co
-3.	High rec strains have evolved a higher proportion of 2CO
-For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
-MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
-4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
--->
-
-<from the means the expected number of COs per chroosomes ranges from X to X we confirm these numbers Can't have fractions of COs>
-
-In order to decompose the cell wide rate, we decided to look at the proportion of chromosome classes (ie by the number crossovers on the same bivalent).
-A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).
-
-The first attern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes.
-<these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>
-
-<note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?)>
-
-The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`).
-This relpicates what has been notice in the literature (cite stapley).
-
-Most of the bivalents are 1CO and 2CO class bivalents -- and the proportion of these two classes distinguish high and low rec male strains. (distinguish the high rec males from minimal rec groups)
-
-(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class.
-The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.
-*(much smaller proportions and variation in the fringe classes*
-
-**2. most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
-
-**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
-
-PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
-
-While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
- and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-
-**3.
-these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**
-
-<Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
-They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
-This significant p value indicates that at least one of the six strains had significant differences).>
-
-Breif conclusion: the evolution of genome wide recombiantion rates proceeds through the chromosome level.
-
-<something about the female results>
-
-<SUMMARY the / TRANSITION to single bivalen measures>
-
-# Single Bivalent Level Results
-
-In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining features of the recombiantion landscape at the single bivalent level.
-recombination landscape refers to the i) total length of the SC-AE (potential area for COs), ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.
-
-From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(Curated_BivData$Obj.ID)` single bivalent measures were left.
-(make sure these are the hand.foci bivalents).
-
-<how do these agree with other patterns of MLH1>
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#sort by category, 
-#compare total boxes and rate passed
-#Curated_BivData  -- pass rate?
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-#use these two metrics
-#cell.level.BivData_table$boxs.IDd 
-#cell.level.BivData_table$nboxes_passed
-
-#calq the mean rate by category
-
-cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
-
-algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
-                        n_obs = length( unique(fileName) ),
-                         mean_pass.rate = mean(pass.rate)
-                         )
-
-cell.level.BivData_table <- add_sex(cell.level.BivData_table)
-
-algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
-                        n_obs = length(unique(fileName)),
-                        mean_pass.rate = mean(pass.rate)
-                         )
-
-data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-Table X illustrates the break down of single bivalent observations by category. 
-Error compared to human measures and other details have been discribed in (Peterson_2019).
-
-The ranges of of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in MSM male to 
-`r max(algorithm.stat.category$mean_pass.rate)` ) KAZ female, we assume that the isolation process is not biased.
-Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.
-We assume the bivalent datasets across all categories are equally eqivilent for representing general patterns.
-
-<curation, x from curation, XY excluded, replication to hand measures>
-
-We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
-
-<These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.>
-<The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.>
-
-# Q1 Sex Differences in SC-AE Lengths
-
-DO THE HIGH QUALITY dataset tests
-
-<!-- breif background on longer female SC  (mouse, human) Gruhn-->
-
-Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of PWD and MSM.
-A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
-
-<Comparing the means of single bivalents is complicated by the fact we can't identify the chromosome identitied the physical size (in terms of SC-AE) is the closet approximation we can make.>
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-We use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data.
-A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes.
-
-<Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes>
-
-This 'short biv' data set has advantages in that it is closer to sampling across a similar set of chromosomes (ideally Chromosomes 15 to 19) and it excludes the XX in females, which is predicted to be the 3rd longest based on the physical (Mb) length.
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#CLEAN UP THE FULL DATA SET
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-#nested boxplots, pooled Bivalents
-
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
-  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
-```{r Q1.SC.show.short.biv, echo=FALSE}
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
-Q1.SC.short.mouse_scatter_boxplot
-
-#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
-skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-#skiveQ1.SC.short.mouse_scatter_boxplot
-#there are 2 lowish female means,
-
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
-#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
-#1 male, 3 females measures
-#try to merge the manual - whole cell observations (at least 2 males)
-
-```
-
-<in the above boxplots of mouse means for shortbivalents>
-
-For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer, the largest difference across sexes by strain are>  (t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
-for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
-This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
-
-<EXPLAIN WHY USING TOTAL SC AREA>
-
-An additional measure of the SC-AE we tested is the total SC area per cell (the whole genome).**to get a comparison of the SC-AE area for the whole genome (not just a subset of chromosomes)** We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
-A total of X images were run trhought the image analysis pipeline, and after outliers removed, X remained.
-/ cell based measures were left.
-Mouse means were calculated for their total SC-AE area and are displayed in figure X with the strain means shown as horizontal lines (below).
-
-```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
-
-Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
-
-Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
-   "WSB","G","LEW","PERC", 
-  "PWD", "SKIVE",  "KAZ",
-  "MSM", "MOLF"), order=T )
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#order the strains
-Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
-  arrange(strain) %>%
-  mutate(mouse = factor(mouse))
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#add category mean
-#totSC.strain.av.table %>% filter
-totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
-
-totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
-
-Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")
-
-Tot.SC_mouse.av_Dom
-
-totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
-totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
-
-Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
-
-Tot.SC_mouse.av_Musc
-
-totsc.mol <- Q1.totSC.mouse.av.table %>% filter( subsp == "Mol" )
-totsc_strain.av.mol <- totSC.strain.av.table %>% filter( subsp == "Mol" )
-
-Tot.SC_mouse.av_Mol <- ggplot(data= totsc.mol, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.mol, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
-
-Tot.SC_mouse.av_Mol
-```
-
-<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )
-
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")
-
-#error not enought female observations
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-```{r Q1.totSC_male.first.comp, echo=FALSE}
-
-Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
-```
-
-The most averages for total SC-AE area (skeletonizes) are significantly longer in females for all strains (p=
-`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
-for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).
-
-Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)
-
-The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
-(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
-<But we use this metric anyways>
-
-<We estimate the effect of the sex chromosomes from hand measured data...the XX (3rd longest chromsomes) in females is ~10% of total the summed SC area, the average propotion of the XY in spermatocytes is XX%?
-however the morphology of the XY is highly variable across cells>
-
-The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>
--- we follow up with models (which focus on sex differences).
-
-<explain why applying models.  We then use models to examine/ think about the evolution>
-
-The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mean mouse total sc and mouse mean short bivalent.
-
-**M3. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA!!
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-#Q1.lmer.totSC_results
-
-#G
-```
-
-For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed model of mouse.av total sc.
-
-The three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
-sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
-
-```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-
-```
-
-Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
-The interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
-
-These effects were followed up the glms, again applied to both mouse means for short biv SC lengths and total sc area per cell.
-
-**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
-
-**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-#subsp effect, many strain, #male
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-The following effects had significant p values for the **total SC** from the above model
-
-subspecies musc p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`, subspecies mol p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`, sex male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`, strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
-<this is the smallest p vale, elevated female total, driven by the high means in females>
-
-(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
-
-<!--
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
--->
-
-Fr the **short bivalent** sc the model results are listed below. From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).
-
-while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.
-
-In the M3 model (strains and sex as fixed effects), the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).
-
-<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >
-
-<In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
-
-<1. the longer female sc's is consistant and seen across all strains,
-even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
-2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-## Adjusting for XX
-
-(rethink where this section should go)
-new outline
-
-1. illustrate problem(affects mostly SC length)
-
-2. Expected impact on sex comparisons, estimated effect size of the X  
-3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
-
-4. 19 female, 19male, 20female 20 male
-
-The female mouse averages should have adjustments for the XX.
-working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
-Subtract this amount from the female mouse averages ...
-This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
-
-- Of all female single bivalents observations, 5% are XX (1 of 20).
-
-- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
-
-- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
-
-I think a formula something like this can be applied to adjust for XX
-
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
-
-# short transition
-
-<transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents
-1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
--- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).
-
-2.
-We focus on two metrics/ aspects i) the placement of single foci along a bivalent (just 1CO) and ii) the placement of two foci on the same bivalent / the interfocal distance of 2CO bivalents (reflects crossover interference)
-
-# Q1. 1CO Position Sex Differences
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE}
-
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
-<!-- motivating
-
-To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
-In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
--->
-
-We focus on the foci positions from one crossover (1CO n= 
-) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
-
-We anchor the position measures at the centromere (and normalize by deviding the position by the length of the chromosome, so the value range from 0 to 1, with smaller values representing positions closer to the centromere end of the bivalent, and larger values meaning foci positions closer to the opposite end (refered here as terminal or telomeric position).
-
-We plot the mouse averages across all of the single bivalent measures, with the strain averages shown as boxplots...
-
-in order to get a better picture of the sex differences in this foci-position metric
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
-#main figure of density plots,
-
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-Q1.mouse.av.biv_scatr_box
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
-           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test(
-
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
-
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (fitting the predictions/pattern of the 'typical recombination landscape' (Sardell)).
-The single CO positions are medially placed in females and terminall placed in males.
-
-The pavlues across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
-
-With the musculus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
-Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).
-
-After confirming the significan sex differences for 1CO position, we apply the familiar models
-
-**M1. Mixed Model**
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-#lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
-(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
-These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).
-
-**M2. Linear model**
-
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-**M3. Linear model**
-
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-
-```
-
-The male effects were the most significant (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-While the musc subspecies, LEW and KAZ strains had slightly signifcant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
-
-(these results will be explored more fully in the Q2 section).
-The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).
-
-<we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>
-
-Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite).
-To account for this we again use the reduced bivalent sets (long.biv and short.biv)
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-
-#real.long.bivData
-
-Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-Q1.short.biv_scatr_box
-
-Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-Q1.long.biv_scatr_box
-```
-
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
-
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-#mouse.av.long.biv.table
-
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
-
-<!--
-### Siscoten
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
-```{r show.siscoten, echo=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-BUT females have a greater range -- so maybe it's just a scale issue.
-
-```{r show.norm.siscoten, echo=FALSE}
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-```
-
-```{r log.reg_siscoten, echo=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-```
-
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
-```{r telo.dist, echo=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
-
-```{r MM.telo, echo=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-drop1(lme_MM_telo.dist, test="Chisq")
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-```
-
-Mixed model result summary:
-
-```{r log.reg_telo.dist, echo=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.-->
-
-```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
-#IGNORE THIS FOR KNOW, THIS
-
-#filter all Curated biv data for 1 strain
-#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
-
-# the stardell paper sums the number of COs for bins (10)
-
-#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
-#but then the model would be close to logistic
-
-#summing the number of COs per bin, losses the chrom specific data....
-#!! need to use a melted Data frame to capture all nrm.foci obs
-#wsb.ttest<- mouse.avs_4MM %>% filter()
-#start with same chrm classes
-#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
-
-WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
-
-PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
-
-anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
-
-#or AIC(model1,model2)
-
-```
-
-```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
-
-#make a table for the sex specific values of the normalized F1 positions
-
-#for strain, sex -- Curated_BivData
-#or (Mouse.Table_BivData_4MM)
-#norm.first.foci.pos
-Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
-
-                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
-                  n1CO = sum(n1CO )
-
-)
-
-```
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#MELT_Curated_BivData
-
-#these isn't long.biv data
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-real.long.bivData
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-#Q1. Sex Differences in CO Interference (IFD)
-
-In order to test if there are sex differences between crossover interference, a major determinant of the positioning of crossovers along chromosomes, we examined ...
-ran models with the interfocal distance (IFD) of two foci on the same bivalent.
-
--We focus on observations from two-crossover bivalents for more comparable observations.
--comparisons of all IFDs of mulit crossover bivalents is also included
-Interference is a major determinant of the positioning of chromosomes
-
-<!-- 
-Brief lit review for IFD / interference expectations  
-**Petkov et al 2007,  CO interference underlies sex differences in RR**
-
-"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
-However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes at the pachytene stage of meiosis."
-
-Chrm1 genetic map, 
-(other human and mice refs 
-Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
-
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-<!--
-ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
-
-explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
--->
-
-<!--  set up -->
-
-<!--  plots-->
-
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-first.IFD.plot_ABS
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-first.IFD.plot_PER
-#+theme(legend.position="none")
-```
-
-```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
-<!-- EXPLAIN WHY WE SHOULD CARE ABOUT RAW AND PER IFD values-->
-
-The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical - force mediated through the SC -- In some models this metric should be thought of as the best metric for charaecterizing crossover interference -- bc it reflect the physcial --
-
-The above plots display the raw and normalized area which seperate two foci on the same chromosome.
-In the normalized plot, 9 observations fall above the length threshold of 150.
-The female measures fall more uniformly across the range of (0 to 1).
-
-In the abosolute plots, you can start to see clustering/ enrichment of IFD values between 40 and 80 pixels for PWD and SKIVE, 
-(and maybe slight clustering for MSM).
-
-<!--t.tests -->
-
-t.tests are performed on the normalized metrics for establish if there are sex differences in the IFD meterics (raw and normalized).
-
-```{r Q1.IFD_Ttests, echo=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
-#SKIVE
-Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-
-```
-
-The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
-BUT WHEN G IS REMOVE, the pvalue is no longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
-<look closer at G>
-
-The normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`) between sexes, indicating there a
-
-When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
-MSM   `r `
-SKIVE  `r `
-PWD  `r `
-MSM  `r `
-LEW  `r `
-G  `r `
-WSB  `r `
-
-This indicates a general pattern of two foci are seperated by more area (hance have stronger interference) in males, this pattern is also for genetic maps.
-
-<!-- short and long -->
-
-```{r Q1.IFD.long.short.setup, echo=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-```{r Q1.show.IFD.things, echo=FALSE}
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-###look at why KAZ males are weird
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#this might
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-```
-
-see the reference plots within this dir, --
-
-this section -- all strains but KAZ, males have a higher mean normalized IFDs
-
-the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs
-
-The KAZ strain is the only strain without significant p value for the normalized IFDs, due to KAZ males having a higher rate (enrichment of short normalized IFDs.
-
-There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
-off.
-Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?
-
-<The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes?
-they were supposed to be 1CO -- but somehow a second CO formed anyways ..
-and very close by
-
-Where as in PWD, SKIVE and MSM there ARE NO 2CO bivalents which have such short IFD
-WHAT espicially werid about this --
-
-whats the expectations following the confonding proportions? 
-high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.>
-
-**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference
-
-to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
-Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)
-
--- run these comparisons to see if the small NRM IFD are due to shorter chromosomes, which are predicted to have weaker interference.
-
-Looked at the KAZ pattern
-
-Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
-These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).
-
-Considering that the short-long bivData is much smaller than the full data set
-
-<!-- lmer/ glms-->
-
-### Mixed Model Tests, Fixed Effects
-
-Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
-
-**M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.IFD.M1, echo=FALSE, message=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#M1_IFD_2CO
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-#don't understand why 2 rows?
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
-#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-#exactRLRT(lmer.IFD.PER_MM.M1)
-
-```
-
-The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
-That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
-The tables below are from anova( for the lmer model ).
-Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
-
-**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
-
-```{r IFD_HetC_M2, echo=FALSE}
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-**M3. Linear model**
-
-$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
-
-```{r IFD_HetC_M3, echo=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-#round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-
-```
-
-For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.
-
-the interaction effects were slightly significant for both raw and nrm.IFD.
-
-The most significant value was the sex effect for nrm.IFDs.
-
-The the random strain effect was not significant for either model.
-
-For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
-The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
-For the M3, only SKIVE*male effect is close to significant
-
-For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
-SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure.
-
-Overall There's a low amount of significant effects across the 2CO IFD measures.
-This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
-
-Dive deeper into the sex specific pattern for each strain.
-Below are code chunks which show the unusual sex specific results for IFD measures.
-The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
-The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
-
-For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
-
-```{r Comp_Ratios, echo=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
-
-# transition Q1 to Q2
-
-we tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.
-
-# Q2 SC Length
-
-<put in plot for total SC and mean gwRR count>
-
-<motivation: the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
-
-<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high rec strains then low rec strains.
-This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
-Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
-This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
-They have a lower probability of 2COs below a certain SC length threshold.>
-
-while testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR
-
-<!-- basic stats
-`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
-X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
-
- place set ups for total SC, short biv and long biv here -->
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
-```
-
-<!-- plots -->
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-
-#+theme(legend.position="none") 
-```
-
-```{r Q2.SC.long.biv_plots, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, mouse) %>%
-#  mutate(file.name = factor(file.name))
-
-mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-mouse.av_totsc_male
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE}
-
-#make plot of mouse averages for MLH1 and total SC
-
-bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
-
-X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-
-X.by.texas
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
-remember the pairwise t.test
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-G-WSB is sig, WSB-LEW are slightly sig
--->
-
-```{r Q2.totSC.ttest, echo=FALSE}
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol
-Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
-`r Q2tot.sc_t.test_High.Low$p.value`).**
-
-Molossinus have significant difference between mouse means for total sc, between the high and low strains (mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
-The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.09
-
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-<across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).>
-
-<!-- models -->
-
-```{r Q2.totSC.glm, echo=FALSE}
-#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel
-
-Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-
-sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
-
-#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
-#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
-#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)
-
-Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-
-sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
-#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-
-#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
-#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
-#table(mouse.av.long.biv.table$mouse)
-#Q2mouse.av.long.biv.table
-
-glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
-
-sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
-
-glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
-
-sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
-
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-```
-
-```{r Q2.SC.long.glm.bivglm, echo=FALSE}
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
-
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
-#summary(glm_SC.male_Musc_long.biv)
-
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
-
-## logistic regression for Rec group
-#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
-
-#summary(log.SC.male_long.SC)
-
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-```
-
-**For the total SC glm**
-
-<Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
-1. no difference in DOM strains (for SC lengths )
-2. PWD > SKIVE > the rest
-3. Msm > mol>
-
-For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
-
-In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-
-<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
-
-<For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW, G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.>
-
-<!--24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM-->
-
-<just like the sex differences in SC area -- this points to differential usage of the SC for placing CO -- investigate the recombination landscape>
-
-<!--
-blocks of other male Q2 - SC lengths
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
--->
-
-**The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).**
-
-**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
-The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
-
-**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
-
-> **Summary Points**
-
-- confirm that the positive correlation of SC and CO number is maintained
-
-- Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.
-
-- (the glms, indecate there are many strain effects), Using the mouse average SC lengths, in the full data set all strain effects are significant.
-This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
-
-- For the M3 models of mouse av.s across SC meterics, for many of the SC metrics, most all of the strain effects are significant
-
-- for the M2 models of the mouse av, musc - mol are significant fixed effects
-
-- many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging
-
-- th predictions for M3 (strain) differencs (motivated by the gwrr variation are not met,  (correlation of SC means and CO means?))
-
-<Each point is a mouse average SC.
-When all bivalents are pooled, the logistic regression SC lengths predict the groups.
-When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.>
-
-<!-- junk chunks below -->
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-<These are the plots of the logistic regression.>
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SClength.ttest, echo=FALSE, include=FALSE}
-#COMPARISONS ACROSS ALL CLASSES
-
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
-
-```
-
-```{r log.reg_SC.length, echo=FALSE, include=FALSE}
-#REMOVE THIS COMPARES FROM FULL BIV DATA
-
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
-#REMOVE THIS, IT uses total bivalent data
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
-```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-table( droplevels(fomo.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
-
-#Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-```
-
-```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
-## run the tests to check following expectations
-
-#1.Dom strains not different from ech other
-#2. PWD and SKIVE are different from the other strains
-#3.MSM and MOLF are different from each other
-
-#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
-
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
-#the subsp labels get messed up, unless they are designated chr first
-Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
-
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
-Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
-#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
-#total sc sig, 1CO sig, 2CO sig
-
-Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
-#0.000369 ***
-Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
-#0.000816 ***
-Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
-#0.0148 *
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
-#making plots above show that mouse average SC lengths are lower could there be microscope issue?
-
-#make sure to remove some of the juvi mice!! from the AP_mouse table
-show.Dom.male.CO <- ggplot(
-male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
-#P not sig for CO mean
-
-#seems like there are more mice in the MLH1 data set than bivData
-#
-
-#WSB all dissected and imaged in '15
-#G imaged 15 - 16
-#LEW imaged 15 to 17
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "1CO SC Length", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "2CO SC Length", xlab = "")
-```
-
-```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}
-
-#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
-#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
-
-male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
-male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
-
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
-#PWD and MSM are most diff from SKIVE?
-
-Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
-#the pooled SC value tests are sig
-
-#the mol comparisons are different due to adj p values
-
-Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
-#no sig (only 3) -- does include CZECCH
-
-#non-sig, with mol pooled with musc
-
-Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
-
-Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
-
-Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
-#non of these sig with the mulitple adjust
-
-Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
-
-#pairwise.t.test(Ozone, Month, p.adj = "bonf")
-```
-
-# brief transition, rec landscape
-
-the lack of predictive power of SC length to predict CO number indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.
-
-# Q2 Normalized Single CO positions
-
-<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->
-<!-- set up chunks here -->
-<!-- plots /figures below-->
-
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos All bivalents")
-Q2.show.nrmF1
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
-
-```
-
-<not sure what the beenfit is to test for chrm effectss>
-
-```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-The previous section found that the single crossover landscape is significantly different between males and females.
-In this section we focus on differences between the high and the low rec male strains
-In order to test if if there are distinguishing meterics for these two groups.
-
-<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->
-
-```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
-#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
-
-#High = MSM, PWD, SKIVE
-Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-#high = PWD and MSM
-Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-```
-
-```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
-
-#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
-
-#Musc
-Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
-# p-value = 0.4
-
-#Mol
-Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
-# p-value = 0.07
-# exclusive rec group not sig
-```
-
-```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
-#these don't have mean.first.foci
-
-#short -- real.long.bivData -- REAL.short_mouse_table
-Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
-
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
-#p-value = 0.3
-
-Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
-# p-value = 0.1
-
-#remove SKIVE??
-```
-
-```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
-
-#real.long.bivData
-#single biv level
-Rec.group_LONG.BivData_F1post_Ttest <- t.test(
-  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.5
-
-Rec.group_BivData_F1post_Ttest <- t.test(
-  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.2
-
-#
-#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
-#p-value = 0.2
-
-#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
-#p-value = 0.4
-
-```
-
-**The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**
-
-T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
-
-<further tests -- within the subspecies, musculus and molossinus?
-
-The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
-
-While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`).
-This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.
-
- <BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF, 
- (MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.>
-
-<long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`> 
-<!-- GLMs because there's no difference 
-toal data set, short and long biv. There's only a difference between MSM and MOKF
-test ppredictions, Dom = NULL,  -->
-
-```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
-
-Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
-#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-#with.chrm.legth  ?
-Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-
-#M2
-Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
-
-```
-
-in addition models which use mouse average of foci position as a dependat variable, we decided to run models for the single bivalent level, and which include chromosome length as a fixed effect.
-
-in the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
-..
-the chromosome length has the most significant effect, LEW by chrm effect are significant
-
-In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects
-
-**For the M2 (subsp)**
-There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).
-
-LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).
-
-But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).
-
-and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)
-
-**For M3 (all strains)**
-
-strain G (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][2]`).
-
-strain Lew (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][3]`).
-
-strain PWD ( p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][4]`).
-
-slight strain MSM  (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][5]`).
-
-very sig strain MOLF (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][6]`).
-
-strain SKIVE (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][7]`).
-
-strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
-
-from the main figure above, you can see that G and MOLF 1CO foci positions are the most different
-
-<no prediction for the strains within subspecies>
-
-For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
-
-there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
-(motivation for MOLF female observations and more male MSM biv measures)
-
-```{r Q2.1COpos.show.MOL, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
-
-mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
-                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-
-#+facet_wrap(~mouse, scales = "free")
-
-single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
-                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
-
-#+scale_color_manual(values=colors_of_strains)
-
-single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
-                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-
-#CHRM.length is very significant-- some of the same subsp and strains are sig     
-```
-
-<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.
-
-**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$>
-
-```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
-#remove this cell,
-
-Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
-
-Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
-
-#making new models
-
-Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
-Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
-
-chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
-
-sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
-#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
-
-#summary(chrm.1CO.pos.chrmLength_strain)
-
-```
-
-<for the first model,
-
-chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`
-
-subsp MOL `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][4]`  
-the mol obj above isn't showing the reight value.
-
-strain G   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][5]`
-
-strain LEW   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][6]`
-
-Strain MSM `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][8]`
-
-chrm * LEW `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][14]`
-
-chrm *MSM  `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][16]`
-
-M2
-
--chromosome length, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][2]`
-
--strain G, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][3]`
-
--strain LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][4]`
-
--strain MOLF, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][7]`
-
--strain SKIVE, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][8]`
-
--chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`>
-
-males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
-
-There are significant difference across the 1CO positions, yet they don't match fit any of the predictions. there are also complications due to the chromsome class proportions that
-
-```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
-```{r Q2.1COpos.glm, echo=FALSE, eval=FALSE}
-male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
-
-Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.Q2.F1_PER_Musc)
-
-#Mol
-Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
-              family=binomial(link="logit"))
-
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-```
-
-# Q2 Evolution of interference is associated with genome wide recombination rate evolution
-
-```{r Q2.IFD.long.short.setup, echo=FALSE}
-#long BivData
-male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
-
-#show
-Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-
-```
-
-<!-- set up chunks here-->
-
-```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
-
-#subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
-
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none", axis.text.x = element_blank())+ylim(c(0,1.1))
-
-biv.IFDs_PER_by.group
-
-```
-
-<Previous prediction for differences in the long and short bivalent dataset.
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.>
-
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
-
-#t.test for IFD
-#mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
-
-```
-
-```{r Q2.t.test_IFD,echo=FALSE, eval=FALSE}
-
-#already made
-#subset
-male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-#t.tests for the Rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
-                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
-
-Rec.group.BivData_IFD.PER_Ttest$p.value
-#p-value < 2.2e-16
-
-#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
-Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
-                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
-
-Rec.group_IFD.PER_Ttest$p.value 
-#p-value = 6.872e-07
-
-```
-
-The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
-The pattern is confirmed with t.test comparing groups within the same subepcies.
-
-<glms>
-
-The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
-
-**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
-
-**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
-
-```{r Q2.IFD.glm.test, echo=FALSE}
-#REPLACE Male.poly with ..
-#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-#summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-#summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD.PER_Mol)
-```
-
-In M1, the model build on the mouse averages of IFD metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]`  for PWD, MSM and SKIVE respectively).
-In M2, the mouse average for normalzied IFD is modeled with only strain effects.
-Only PWD and SKIVE IFD^PER^ had significant strain effects (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]` for PWD and SKIVE respectively).
-
-<when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF>
-
-<!-- we ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length.
-
-**model BivLen.M1 ** $$Single bivalent\ IFD ~=~ subsp * strain * chromosomeLength + \varepsilon $$
-
-**model BivLen.M2 ** $$Single bivalent\ IFD ~=~ strain * chromosomeLength + \varepsilon $$
-
-```{r Q2.IFD.glm.chrmLength, echo=FALSE}
-
-#repeating chrm models with chromsome length
-#FOR BIVALENT LEVEL MODELS
-
-Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
-
-#M.Len.ABS
-Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.raw)
-
-#M.Len.PER
-Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.PER)
-
-#M.Len.raw
-Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
-
-##M.Len PER
-Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-```
-
-For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs).
-Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^.
-The Mol\*chrmLength is negative
-<PWD, MSM, MOLF, SKIVE are the only significant strain effects.
-- all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
-With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
-All of the high rec strains and MOLF have significantly longer IFD^PER^.
-There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
--->
-
-```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
-
-#calculate the rates of 'close foci' (<30%)
-#examine the changes in the distribution closer.
-
-#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
-
-close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                      n.close.foci = length(which(IFD1_PER <= .3)),
-
-                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                      rate.close.foci = n.close.foci / N.obs
-)
-
-close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
-
-close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("strain"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER),
-
-                      inverse.close.foci = length(which(IFD1_PER > .3)),  
-                      n.close.foci = length(which(IFD1_PER < .3)),
-                      rate.close.foci = n.close.foci / N.obs,
-
-                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
-
-)
-#the
-
-show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
-
-#add rec.group
-close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
-
-#POOL all short and 
-#chi square test
-high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
-                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
-#p-value = 0.03, slight dif tween 3 high rec strains
-
-#don't need to run chi square for this pattern, since it's enough to
-
-```
-
-We determined that the main difference in IFD disbribution across these groups is an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC length in low rec strains.
-The rate of these close foci range from 8.2% (G) to 16% (KAZ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for skive, MSM and PWD respectively).
-<no chi sqaure test see above>
-
-```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
-
-#Log regression all show significant prediction
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-```
-
- **(unlike total SC length and 1CO pattern, IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.
-However, they go agaisnt a standard perdiction for the relationship between the chromosome level recombination landscape and genome wide recombiantion rate.**
-
-<!-- 
-Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
-But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
--->
-
-<Below are examples of plots of SC length distributions across cells.
-The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
-Most plots excluded for space.
-
-Each point is a bivalent plotted by cell on the x axis.
-X's are the 4th quartile, big point is the mean and smaller black point is the median.
-I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
-(the extra stats are not correctly mapped)>
-
-<HAND measure METRICS, below chunk goes with the above chunk>
-```{r long.biv.plots, echo=FALSE, eval=FALSE}
-
-show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
-
-show.the.cells_females_dom
-#PWD seems to have the biggest variance across cells
-
-almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_DOM_f
-
-#### Musc female
-
-show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
-
-almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
-
-### Dom Male
-show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
-
-almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#MUSC male
-show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
-
-almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
-
-#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
- # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
-#show.the.cells_male
-
-#almost_whole_cell_BivData
-#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
-#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
-#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
-
-#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
-```
-
-<All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
-It shows the amount of within cell variance across strains.
-There is a bit of variance across the SC length distributions in the PWD females.
-This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
-The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
-The full data set is `r length(Curated_BivData$Obj.ID)`.
-This is the breakdown of bivalent observations by category for this long dataset are
-
-**- Try to merge this DF with the whole.cell manual measures.**  
-- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.  
-in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.>
-
-```{r show.long.SC, echo=FALSE, eval=FALSE}
-
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
-
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
-
-```
-
-<ABOVE plots show the SC lengths for the 'long SC data set'.
-They are supposed to be the longest 4-5 SC from cells where I could get good measures.
-These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
-Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
-
-<!-- unused chunkcs -->
-
-```{r SC.distb_manual_Auto, echo=FALSE, eval=FALSE}
-
-#calq the mouse means
-manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
-                        median_SC = median(SC.length, na.rm = TRUE),
-                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
-#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
-
-)
-
-manual_summary_stats <- add_mouse(manual_summary_stats)
-manual_summary_stats <- add_strain(manual_summary_stats)
-manual_summary_stats <- add_sex(manual_summary_stats)
-manual_summary_stats <- add_subsp(manual_summary_stats)
-
-###
-#calq the mouse means
-Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        median_SC = median(chromosomeLength),
-                        Qar.4 = quantile(chromosomeLength)[[4]],
-
-                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
-
-)
-Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
-
-#colors_of_strains
-#scale_color_manual(values=colors_of_strains)
-
-#un-order these factors
-```
-
-<!-- Whole Cell Manual Comparison -->
-
-```{r XX.adj_manual.whole.cell, echo=FALSE, eval=FALSE}
-
-#load and examine the the whole cell hand measures here
-
-female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
-data.frame(table(droplevels(female.whole.cell$category)))
-
-# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
-
-#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
-
-data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
-#missing some strains, most have 3 cells (min). MSM has 5 full cells
-
-#remove the incomplete LEW female cell
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
-
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
-
-#calq mean SC lengths by rank
-
-mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
-
-                       mean.sc =   mean(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                         se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
-
-top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
-
-show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
-
-show.SC_by.rank_top
-
-bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
-
-show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
-
-#CALQ PERCENT TOTAL SC
-
-CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
-
-                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       #rank.by.per
-
-                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
-                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
-                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
-                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
-                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
-                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-
-#MELT ABOVE
-MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
-              #leave out the Foci
-               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
-
-            id=c( "fileName","total.sc.cell","nbiv"))
-
-MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
-
-MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
-MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
-
-show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
-
-show_long.sc.proportions
-
-#original plot code 
-#ggplot(data = df1, aes(x, y, color = Group)) +
-#    geom_pointrange(aes(ymin = lb, ymax = ub), 
-#                    position=position_jitter(width=0.5), 
-#                    linetype='dotted') +
-#    theme_bw()
-```
-
-The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
-
-The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
-
-The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
-For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
-
-(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
-
-The other figure shows of each single bivalent contributes to the total SC area.
-Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
-on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
-So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
-
-- Is the difference between cell averages for males and females less that 10%?
-
-- also interesting, the pwd and msm don't have longer SC, compared to other strains.
-
-<!-- Automated BivData Comparison -->
-
-```{r XX.Automated.biv.stuff, echo=FALSE}
-
-#make automated female data to compare to
-female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
-
-female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
-
-#table(female.Bivdata_comp$strain)
-
-#female.Bivdata_comp
-
-#1. calq the rate of chrm segmentation in female categories in automateted data
-
-box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
-
-                   Nmice = length(unique(mouse)),
-                   mx.box.num = max(boxNumber), # need to run this at the cell level
-                    N.biv  = length(Obj.ID),
-                   pass.biv = N.biv / mx.box.num
-)
-box.rates <- add_mouse(box.rates)
-box.rates <- add_strain(box.rates)
-box.rates <- add_category(box.rates)
-
-cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-
-cell.pass.rates
-
-category.box.rates <- ddply(box.rates, c("category"), summarise,
-
-                   mean.pass.rate = mean(pass.biv)
-)
-#category.box.rates
-
-#2. pull the 3rd rank biv's calculate the SC means
-#calq means by rank (focus on the top 5)
-
-```
-
-For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
-
-The table above shows the number of bivalents from the same strains as in the manual whole cell data.
-The plot shows the bivalent passing rate across all of the individual cells from this female data set.
-For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
-
-(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
-(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
-
-strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
-
-** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
-But I am not using 'whole cell' summaries to compare female and males.
-
-**What is the effect of an extra XX-autosome on single bivalent means?**
-
-use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
-Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
-Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
-Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
-The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
-
-# Note on Heterochiasmy Definition
-
-<!-- add outline for new XX adjust section 
-adjust the colors for the plot below (standardize)
--->
-
-I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
-In females the XX bivalent is indistinguishable from the autosomes.
-To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
-Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
-(I note if the and Y are paired, which they are at a high rate).
-The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
-To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
-
-1. Compile full-cell data from females (all 20 bivalents measured)
-2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
-3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
-
-According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
-
-(Put the XX adjustment section here)
-
-```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
+```{r MLH1.HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
 #calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
@@
 
-```{r HetC.plot, echo=FALSE, warning=FALSE}
+```{r MLH1.HetC.plot, echo=FALSE, warning=FALSE}
 #HETC plot
-
-#set up data
+#change the main data?
 
@@
 
-Het.C.plot
-#this plot is missing several
-
-```
+```
+
+```{r juvimice.list, echo=FALSE}
+juvi_mice <- c(
+  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
+  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
+  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
+
+  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
+   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
+  )
+
+#REMOVE JUVI mice
+#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
+#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
+
+```
+
+```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
+
+#this block should be in the data set up
+#the mouse averages will be used
+
+#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
+cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
+
+#cleanCurated_BivData
+#removed the no hand foci count
+
+mouse.avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+mouse.avs_4MM <- add_sex(mouse.avs_4MM)
+mouse.avs_4MM <- add_strain(mouse.avs_4MM)
+mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
+mouse.avs_4MM <- add_category(mouse.avs_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
+mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
+mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
+
+#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
+#code if males are high or low?
+
+#REMOVE JUVI mice
+juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
+mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
+
+#assign high and low
+#divide musc males into groups
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
+
+mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
+
+```
+
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+
+#add this to setup
+
+#remove female, 
+#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
+
+#Curated_BivData_male
+
+male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+male.av_4MM <- add_sex(male.av_4MM)
+male.av_4MM <- add_strain(male.av_4MM)
+male.av_4MM <- add_subsp(male.av_4MM)
+male.av_4MM <- add_category(male.av_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
+male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
+male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
+
+#REMOVE JUVI mice
+juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
+mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
+
+```
+
+```{r totSC_setup, echo=FALSE}
+#SC.skel
+#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included
+
+SC.skel.OG <- SC.skel #5683
+
+#MERGEING MLH1 with SKEL
+MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
+
+MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
+MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
+#5110,
+
+OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
+
+#check for DUPs, #check for duplicatations of the random name
+MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
+tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
+
+MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
+
+MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
+MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
+#table(MLH1.merge.Skel$category)
+
+#remove outliers
+MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
+
+```
+
+```{r totSC.viz4outliers, echo=FALSE}
+
+# -- EXAMINE the ggplot scatter plots for outliers
+tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
+
+#round 1
+SC.skel.outliers_round1 <- c(
+  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
+  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
+            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
+
+    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
+                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
+              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
+          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",
+
+  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
+          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",
+
+       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
+                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",
+
+   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
+   "6may15_17apr15_PWD_f1_sp1_7_rev",
+
+      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
+            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",
+
+    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
+    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
+  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
+ "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
+ "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
+ "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
+"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,
+
+  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
+    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
+
+# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
+"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
+
+#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
+"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
+
+#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
+"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
+"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
+"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
+
+#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
+"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
+ "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
+ "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
+ "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
+ "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
+ "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
+"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
+"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
+ "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
+ "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
+ "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
+ "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
+
+MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
+
+remove.mice_round2 <- c(
+    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
+    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
+
+    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
+  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
+"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
+  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",
+
+  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
+  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
+"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
+"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
+
+# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
+"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
+"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
+
+#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
+"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
+#SPIC male  x
+#SPIC female  x
+#TOM x
+#AST x
+#CZECH male x
+
+#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
+ "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
+#SPRET female X
+#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
+"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
+)
+#3remove.mice_round2
+
+MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
+MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
+
+all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
+```
+
+```{r LONG.SHORT.biv_setup, echo=FALSE}
+
+#FROM q1
+#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+#from Q2 IFD
+#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+##long biv
+#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+#FROM q1 
+REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
+                         mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                      #make sure 1CO  
+                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
+                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
+                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                         )
+
+REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
+
+REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
+REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
+REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
+                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
+                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
+                                           0)))
+
+REAL.short_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
+                                ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1, 0))
+
+###LONG BIV
+
+REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+                       n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
+                         mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
+                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
+                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+                         )
+
+REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
+
+REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
+REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
+REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
+                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
+                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
+                                           0)))
+
+REAL.long_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
+                                ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1, 0))
+
+#find which one 
+```
+
+<!-- 
+two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
+-->
+
+```{r tot.skel.table, echo=FALSE}
+
+totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
+
+#unorder the factors
+totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
+totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
+totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
+totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
+
+OG.totSC.mouse.av.table <- totSC.mouse.av.table
+
+#REMOVE JUVI mice
+tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
+total SC area")
+
+mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
+
+```
+
+```{r, tot.SC.setup, echo = FALSE}
+
+totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
+                 .(sex,strain),
+                 summarize, 
+                 ncells = sum(ncells),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(mean.skel, na.rm = TRUE),
+                 mean.bin =  mean(mean.bin, na.rm = TRUE)
+
+)
+
+totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
+
+```
+
+<!--
+From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
+
+the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
+
+INSERT TABLE FOR THE TOTAL SC means ect
+
+The difference between
+
+bin is short for binary, it reports the binarized area of the SC
+
+skel -- takes the skeletonized midline of the bin objects
+-->
+# Main Figure
+
+```{r table.for.plot, echo=FALSE, warning=FALSE}
+#why do I add in a different table?
+
+#this move this to the setup file
+#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed
+
+#pooled
+strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
+                        Nmice =length(unique(mouse)),
+                        Ncells  =length(nMLH1.foci),
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci, na.rm = TRUE),
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                       #Standard error (mice vs cells)
+                        se  = round(sd / sqrt(Ncells), 3)
+
+                       #range_mouse_means = range()
+                   #consider adding crhm proportions?
+    )
+
+strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
+                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")
+
+#adjust order of plot here
+strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
+     "MSM", "MOLF",   
+  "PWD", "SKIVE",  "KAZ",
+  "WSB","G","LEW","PERC",
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+```
+
+```{r Fig1, echo=FALSE, warning=FALSE}
+
+#might need to nudge the X values
+
+Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+
+
+    theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+#remove y labfor male
+
+   #segments
+#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc
+#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #mol
+#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
+  #cast
+#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
+  #outgroups
+ # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
+  #label
+
+ #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
+   #        y = c(18.5,18.5,33,18.5, 18.5),
+
+  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
+    #       fontface = 'italic', hjust = 0.5, size=3.5)
+
+```
+
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+
+female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+female_mouse.av <- add_sex(female_mouse.av)
+female_mouse.av <- add_strain(female_mouse.av)
+female_mouse.av <- add_subsp(female_mouse.av)
+female_mouse.av <- add_species(female_mouse.av)
+
+female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
+
+female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
+female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
+female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
+female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
+
+#table( droplevels(female_mouse.avs_for.analysis$strain) )
+#there are cast strains too
+
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+  theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+#female.specific.plot
+
+```
+
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
+#try customize the plot size
+
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")
+
+just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#table( droplevels(just.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+#fomo.male_mouse.av
+
+male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
+male.plot
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+
+```
+
+```{r three.plots, echo=FALSE, warning=FALSE}
+#merge the 3 plots
+#try patchwork
+
+#Fig1_v1_strain + female.specific.plot + male.plot
+
+#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
+Fig1_v1_strain / (female.specific.plot + male.plot )
+
+#(p1 | p2 | p3)/p4
+
+```
+
+This is a draft of the main figure. Outgroups are not included in the analysis below.
+
+<!--  insert Details of data set cleaning etc is in SetupData.rmd
+Effects from different factors are reviewed in Effects.Rmd -->
+
+```{r main.fig1s, echo=FALSE, include=FALSE}
+strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
+
+Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point( size=3, position=position_dodge(.5))+
+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+
+
+    theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+
+#strain.table.HQ_Fig.og
+
+strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")
+
+strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
+     "MSM", "MOLF",   
+  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
+  "WSB","G","LEW"), order=T )
+
+Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
+  ylab("Average MLH1 Foci per Cell")+
+
+  geom_point(size=3, position=position_dodge(.5))+
+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+
+  ylim(c(18, 35))+
+ scale_color_manual(values=colors_of_strains)+
+
+    theme(legend.position="none",
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+
+```
+
+```{r basic.numbers, echo=FALSE, eval=FALSE}
+
+kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+
+kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+
+```
+
+```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
+#str(AP_mouse_table_w.Ages)
+
+AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+
+#remove sex.x == na, and strain.x=na or F1
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+
+AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
+#plot variance and mean (facet sex)
+
+AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
+
+VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
+
+cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
+
+str(AP_mouse_table.HQ)
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
+#[1] 0.5750399
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
+#0.321601
+
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
+# 0.3515544
+cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
+#-0.03713536
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
+# p-value = 8.623e-09
+
+cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
+    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
+
+#remove mice with fewer than 10 cells
+
+AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
+
+VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
+
+cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
+
+cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
+    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
+
+#why did I male just a female table?
+female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
+
+female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(adj_nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(adj_nMLH1.foci),
+                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+                        #quality?
+)
+```
+
+# Genome wide recombiantion rate estimates
+
+```{r MLH1.summ.stats, echo=FALSE}
+
+#strain.table.HQ_Fig.og  all strains
+#strain.table.HQ_Fig  extra strains removed
+
+main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
+
+main.table <- add_subsp(main.table)
+main.table <- add_species(main.table)
+
+main.table %>% filter(sex == "female")
+
+main.table %>% filter(sex == "male")
+
+#G.HETC.val #1.2
+#LEW.HETC.val  #1
+#WSB.HETC.val  #1
+#KAZ.HETC.val  #1.1
+
+#SKIVE.HETC.val # 0.96
+#PWD.HETC.val  #0.88
+#MSM.HETC.val #0.9
+
+#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
+
+#strain.table.HQ_Fig
+
+#sum( (main.table  %>% filter(sex == "male"))$Ncells)
+#the kable version is bulkier than the other versaion
+
+```
+
+We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M.
+m.
+domesticus,_ _M.
+m.
+musculus_.
+and _M.
+m.
+molossinus_.
+MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
+<!-- state the average number of mice
+The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises. -->
+
+<!--
+1. Fig1A  magnitude and direction of heterochiasmy;
+-general patterns i) female higher and ii) low degree of hetC  (range of 1 to 1.2) (human is 1.6)
+-All strain means are above the expected minimum of 20 (1CO per chromosome)
+-there are exceptions they are followed up in the sex-specific plots  (male higher)
+
+2. Fig1B and C mouse means and sex specific patterns; 
+-female mouse mean distributions are have more spread across the range of means
+-male distributions are more clustered, low and high (~30% evolution largest-rapid evolution of gwRR)
+-sex-specific evolution, both sexes don't show equivalent elevation of gwRR
+-Dom vs Molossinus and Musculus
+
+-->
+
+Figure 1A shows the strain averages for each sex together - this is a way to display the general patterns of sexual dimorphism (or heterochiasmy) across strains. 
+Taking note of the direction and the magnitude our results confirm two general patterns that i) the genome wide recombination rates averages are greater females compared to males and ii) the degree of heterochiasmy (F:M rates) is generally low, rangeing from `r G.HETC.val` in _G_ to `r LEW.HETC.val` in _LEW_.
+
+The notable exceptions to this pattern are from strains with male averages that exceed the female averages, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` for PWD, MSM and SKIVE respectively.
+
+Examining the mouse means for the genome wide recombinaiton rate seperately for each sex reveals disctinct patterns of variation (Figure1 B and C).
+
+The female means can be discribed as more evenly distributed around the sex-wide mean of ~25
+
+While in males, the strain specific means -- can be seperated into groups of high genome wide recombination rates (near 30 COs per cell) and low genome wide recombiantion rates, near 23 COs per cell.
+
+<!-- transition from first part into female -->
+<!-- female section -->
+
+```{r female.strain.table, echo=FALSE}
+
+female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
+#female strain table
+#female_mouse.avs_for.analysis
+F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
+
+                        Nmice = length(unique(mouse)),
+
+                        total.cells  = sum(Ncells, na.rm = TRUE),
+
+                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#F.strain.table
+
+#F.strain.table
+
+# I don't know why most of the table is blank
+```
+
+```{r female.dif, echo=FALSE}
+
+#how much greater are the high strains?
+G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
+mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
+
+LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
+mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
+
+#how much greater is MSM female
+MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
+mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
+
+sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
+mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
+
+G.female.diff <- (mean.G.female / mean.all.other)
+LEW.female.diff <- (mean.LEW.female / mean.all.other)
+MSM.female.diff <- (mean.MSM.female / mean.all.other)
+```
+
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
+Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
+
+#variance plots
+musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot
+
+#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+#Mol.female.VAR.plot
+
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.female.VAR.plot
+
+```
+
+```{r Female.specific.glm, echo=FALSE}
+
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
+
+#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
+
+#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+```
+
+<!--
+The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (glm; p= ).
+ -->
+
+```{r FEMALE.VAR.MODELS, echo=FALSE}
+## within mouse variance models 
+lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#round( #round isn't working for this sumary
+#lmVAR.female.strain_sum$coefficients[,c(1,4)]
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+
+lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
+
+lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
+lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
+
+#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
+#LEW sig
+
+```
+
+<!-- male section -->
+
+```{r male.strain.table, echo=FALSE}
+M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+
+                        totNcells  = sum(Ncells, na.rm=TRUE),
+
+                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
+
+                        #Standard error (mice . cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#M.strain.table
+#same issue as the female table .. ugh
+```
+
+```{r male.diff, echo=FALSE}
+just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#how much greater are the High Rec gwRRs
+just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
+                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
+                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
+
+#how much bigger are the high rec male means?
+sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+#mean(sans_LewRec_male$mean_co)
+
+PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+#mean(SKIVE_male$mean_co) /  
+```
+
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
+
+#variance plots
+musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.male.VAR.plot
+
+Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.male.VAR.plot
+
+Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.male.VAR.plot
+```
+
+```{r Male.specific.glm, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+
+#summry.male.strain.effects <- summary(glm.male.strain.subsp)
+#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
+
+#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+
+#1. determine which strain effects are sig
+
+#M2
+MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
+sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
+
+#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+
+#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+```
+
+```{r MALE.SPECIFIC.variance.model, echo=FALSE}
+## within mouse variance models 
+lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
+
+#all p values kinda high, MSM and TOM are slightly sig
+lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
+#non sig
+
+lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
+#NS
+```
+
+Examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.
+
+<!-- female strain means are have more variation in a random way. The range of genome wide rec strains is X to X, -- no clustering of strain means
+
+-->
+
+In females the strains with the greatest difference
+in mean gwnome wide recombiantion rate was `r G.female.diff` ( _domesticus^G^_) more compared to other strains. 
+<!-- _domesticus^LEW^_  (`r round(LEW.female.diff,4)`),
+and _musculus^MSM^_  (`r MSM.female.diff` ),-->
+
+<!-- male variation is a greater magnitude, but divides into 2 groups, magjority of strain means are around 23, while PWD and MSM are near 30, skive is intermediate-->
+
+While in the male data (most strain averages are close to 23 (Figure1).
+
+the genome wide recombination rate for had a much larger range, up to `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )` ( _molossinus^MSM^_ ).
+
+_musculus^PWD^_ ,and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,  and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.
+
+<notabley these strains come from Musc and Mol -- Mol natural hybrid of musculus and castaneus. this comments out text?>
+
+The _musculus_ and _molossinus_ subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
+
+<!-- dom, low variation and close to the expected in
+within the male means, there is a low degree of strain varaince in _Domesticus_ subspecies with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
+-->
+
+<!-- ranges of mouse means
+In the _domesticus_ subspecies the ranges of mouse means are 
+`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )`
+in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.
+
+In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_
+
+The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
+-->
+
+<!--
+$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$ -->
+
+<!--
+The strain means for females were pretty similar.
+For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively.
+In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively.
+In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively.
+-->
+
+While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively.
+In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively.
+In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.
+
+<!--(the figure is made from pooling MLH1 cells)
+the male 
+- the ranges of mouse means are displayed in Sup table X
+- How the means were calculated
+-The male gwRR is adjusted for the PAR
+-The means for each strains are report in Table 1
+-Normality was assessed for the MLH1 counts (include in Supplement histograms)
+-->
+
+# Evolutionary Framework
+
+<to better understand this varation in RR within an evolutionary framework. Make table of models>
+
+**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
+
+**M3. LInear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+
+We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework.
+
+```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3, message=FALSE}
+#THIS IS THE NEW MIXED MODEL
+#should be able to use the loaded 
+#DF.HetC.MixedModel.HQ
+
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1.MM.M1)$coefficients
+
+MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+
+MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
+sex.results =  MLH1.MM_sex, 
+               interaction.pval = MLH1.MM_interaction,  rand.pval = 
+MLH1.MM_strain)
+```
+
+All of the effects within the mixed model M1.
+were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
+Confirming the qualitative observations of wide spread variance.
+<!-- indicating XXX -->
+
+```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
+#strain = fixed
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
+
+#table of just coeff anf pvalue
+#round(summ$coefficients[,c(1,4)],5 )
+
+#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
+```
+
+```{r MLH1.glm.M3, echo=FALSE, digits = 3}
+
+MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+
+summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
+
+#round(summ.strain.fixed$coefficients[,c(1,4)],5)
+
+#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+```
+
+The linear models (M2. and M3.) allow more == specific comparisons
+
+Following up with linear models which tested for (subsecies M2), and strain effects alone (M3)
+both showed similar -- results -- **so we just report those from M3. also just report effects which were consisantly significant across both M2 and M3**
+
+<they also have sig strain effects .. PWD strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`>
+
+which identified several signifcant sex by strain interaction effects, ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
+and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`
+)
+
+<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
+These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.
+
+<M3 also has a few less significant effects>
+<not the reverse coefficient indexing -- this will give the name >
+
+Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
+These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
+
+_domesticus^G^_ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
+and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`
+
+<!--
+In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
+
+(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
+
+Check the DF of the number cited above
+This is the break down of mice used within the Mixed model
+`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
+
+add number of cells
+`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
+ $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
+
+Setting up the data set for using the mixed model (base data set, HQ data set)
+
+- only Mus musculus strains with sex matched observations  
+- Male gwRR have been adjusted for the PAR
+- Q5 excluded
+- All the effects are un-ordered factors  
+- male ages within 5 - 20 weeks
+- Molossinus is coded as a subsp level
+
+subspecies levels: 3, Musc, Dom and Molossinus
+strain levels: Dom: 3, Musc: 3  Mol: 2
+sex: 2 female male  
+-->
+
+<!--
+also g strain effect/ dom^G^ strain effect 
+`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[12,][4]` 
+In both models strain by sex interaction effects were significant for (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
+-->
+<!-- glms for sex specific data also run -->
+
+A similar glm models (M4) was run on the sex specific data set, (just containing subspecies, strain and the interaction).
+
+ **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$
+
+For the male specific data, three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
+(with effect sizes ranging from 5, 7, and 2 foci respectively).
+Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
+
+While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
+
+<!-- transition gwRR to within mouse variance -->
+
+## Within Mouse Variance in CO Count per Cell
+
+In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).
+
+Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
+
+Following the same basic framework we used for analyzing mean number of MLH1 foci per cell, we applied a set of mixed and linear models -- to test effects (sex differences for the within mouse variance in MLH1 foci per cell).
+
+ **M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+ **M6. Linear Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex * strain + \varepsilon$$
+
+ **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
+
+```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
+
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M1)
+LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
+
+LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
+sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
+
+#CV
+MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
+
+LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
+sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
+
+#USE Q12 (Q123) table
+#Q12_mouse_table_w.Ages
+#Q12 var
+MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
+
+LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
+sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
+
+#Q12 cv
+MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
+
+LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
+sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
+```
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M3)
+anova(MLH1var_M3)  
+as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
+
+MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+#summary(MLH1cv_M3)
+
+anova(MLH1cv_M3)
+as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
+
+#Q12 var
+MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1var_M3_Q12)
+anova(MLH1var_M3_Q12)  
+as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M3_Q12)
+anova(MLH1cv_M3_Q12)#sex large F
+
+as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
+
+```
+
+```{r M3.MLH1.var, echo=FALSE}
+
+glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
+#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
+
+MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+#summary(MLH1cv_M4)
+#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
+
+### Q12 var
+MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
+
+#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
+#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
+
+#Q12 cv
+MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
+```
+<!--
+These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
+var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
+cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
+5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
+-->
+
+For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
+Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
+background noise).
+For this reasons, we replicated the model analysis using a subset of cells with higher quality scores.
+These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).
+Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
+While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`) for the model with the full dataset.
+While in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
+For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell.
+This limits the power for detecting significance differences across groups, <but the main/general pattern, is that females have significantly higher within mouse variance compared to males.
+This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.
+
+<!--
+motivation / background: maybe checkpoint / regulation
+limitations: observations at the mouse level, decreased power to estiamate the variance --
+
+$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
+<!--
+RW's main qtl paper used the within animal variance,
+
+@article{wang2017genetics,
+  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
+  author={Wang, Richard J and Payseur, Bret A},
+  journal={Genetics},
+  volume={206},
+  number={4},
+  pages={1841--1852},
+  year={2017},
+  publisher={Genetics Soc America}
+}
+references for phenotypic variance connected to robustness to environment, increase in developmental consistency
+
+Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
+phenotypic variance leading to adaptation - 
+(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
+Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
+meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
+-->
+
+< SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE >
+
+# Variation in DSB Number Per Cell
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize,
+
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
+
+#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
+#            row.names = FALSE)
+
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+```
+
+```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+#change order
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
+                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.boxplot
+```
+
+```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
+
+anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#
+
+DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
+
+  #facet_wrap(~stage, nrow=1) +
+
+  ggsignif::geom_signif(
+    data=anno_df_stage, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
+
+ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
+               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
+
+DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
+
+ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
+               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
+
+DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
+
+ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
+               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
+```
+
+```{r plot.t.test, echo=FALSE, include=FALSE}
+
+  #this will add neat bars, but still need to figure out what groups ect it's actually using
+#THE stat_compare function adds a arm thing and pvalue for 
+#  stat_compare_means(comparisons = my_comparisons)
+
+#status - legend, expression y-axis, facet, facet_wrap SYMBOL
+#legend - stage. y=total count, FACet strain
+
+#add column to distinguish high vs low rec strains
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+
+DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
+
+# annotation table with adjusted pvals and y-position of the labels
+
+anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#
+
+DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=stage), position=position_jitterdodge()) + 
+  facet_wrap(~stage, nrow=1) + 
+  ggsignif::geom_signif(
+    data=anno_df, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#example for ggplot+ test from here
+#https://github.com/kassambara/ggpubr/issues/65
+
+#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
+
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+```{r DMC1_anova, echo=FALSE, warning=FALSE}
+#1. ANOVA, is there a difference in DMC1 counts across strains?
+
+#unorder strain
+DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
+
+mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
+mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
+
+l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
+z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
+```
+
+```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
+#make category averages from AP_mouse_table.HQ
+mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+#attach this mini DF to the DMC1 data
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+##RATIOS
+DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
+DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+
+#DMC1.mean.table
+DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
+
+colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
+                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
+
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
+
+kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
+<!--
+explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
+-->
+
+In an attempt to locatize the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days).
+A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
+One mouse was sampled for each strain.
+
+For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = ), which reflects the expected patterns across the timing of meiosis.
+
+Confirming the prediction that (more crossovers are associated with more initial DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
+
+The differences in DMC1 foci were not significant for zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
+<!-- which maybe be attributed to noise effects
+
+The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->
+
+The ratios of average number of MLH1 and DMC1 foci per cell -- have  .... biological significance
+
+The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+
+These results indicate that the changes -- which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
+
+<!--
+The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
+The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
+`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M.
+m.
+domesticus^WSB^_ , _M.
+m.
+domesticus^G^_, _M.
+m.
+molossinus^MSM^_, and  _M.
+m.
+musculus^PWD^_ respecitively.
+-->
+
+```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
+
+#Subsampling code -- don't report
+
+#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
+Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
+Permutations were done for both cell stages.
+The actual p value is in red.
+
+#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
+
+#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
+# the 2 groups
+
+#make the 2 pools
+High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
+
+Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
+
+#permute drawing / subsampling and performing the t-test
+#put the p.value results into a DF!
+
+#permut.SC.cv.diff  <- as.data.frame(
+#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
+
+#permuting samples -- and re-licating t-tests
+
+#nrep <- 1000
+
+#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
+bunchOpvals.L <- replicate(1000, 
+    # t.test
+    t.test( sample(High.L.pool, 25, replace = FALSE),
+            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
+          )
+bunchOpvals.L <- as.data.frame(bunchOpvals.L)
+bunchOpvals.L$stage <- "L"
+
+bunchOpvals.Z <- replicate(1000, 
+    # t.test
+    t.test( sample(High.Z.pool, 25, replace = FALSE),
+            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
+          )
+
+bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
+bunchOpvals.Z$stage <- "Z"
+
+#could merge the lists of p-values
+replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
+#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
+
+#ASSUMPTIONS; normality, equal variance
+
+#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
+
+#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
+
+#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
+#plot.pvals.L
+#plot.pvals.Z
+
+#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
+#n use the var.equal = TRUE
+#https://www.statmethods.net/stats/ttest.html
+```
+
+```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#these are histogram plots for testing Normality shape -- move to supplemenatry section
+
+#- check that distributions are close to normal
+#- check for equal variance.
+
+DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
+
+DMC1.histogram
+#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
+#bartlet's test, Levine's test
+
+#The counts look close to normal.
+#Potentially figure 2. Using a function to #display the p-vales for mean comparison
+
+```
+
+<!--
+interpertation notes, 
+the number of DMC1 across stages (L and Z stages) are driven by different equations
+
+Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
+
+Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
+
+*the repair decision can happen at both stages and remove DMC1.
+formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
+
+back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
+
+expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
+OR
+
+variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
+-Baier et al 2014 supports this model (Chmr axis / DNA loop model)
+-->
+
+<!--
+# Rapid male specific evolution
+
+list out the points which support rapid male specific evolution for MLH1 rates.
+
+number above bars are the raw counts for bivalent classes.
+
+Potential figure 2 (eventually). Try to make the figure taller.
+-- add more data to make sample sizes more equal.
+-- add space to indicate where female data is missing
+
+SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
+What's up with Lew females? there's lots of 2COs, also WSB female.
+
+This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
+If these data are random samples of bivalents, the X
+
+ToDo:
+- within strain across sex,
+- males, across high and low groups
+- look at mouse level proportions
+-->
+
+# Chromosome Class Proportions
+
+```{r chrm_prop_data, echo=FALSE, warning=FALSE}
+#shut off eval for now because of wd / path error
+#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
+chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#try adding the BivData version of the proportions
+
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
+
+chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
+chrm.class.DF <- add_mouse(chrm.class.DF)
+chrm.class.DF <- add_strain(chrm.class.DF)
+chrm.class.DF <- add_sex(chrm.class.DF)
+chrm.class.DF <- add_category(chrm.class.DF)
+#chrm.class.DF <- add_euth_date(meta.data.file2)
+#chrm.class.DF <- add_age(meta.data.file2)
+
+#table(chrm.class.DF$category)
+#str(chrm.class.DF)
+
+#summarize by category
+proportion_stats_category  <- ddply(.data=chrm.class.DF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
+
+                 n0CO = sum(n0CO, na.rm = TRUE),
+                  n1CO = sum(n1CO,na.rm = TRUE),
+                  n2CO = sum(n2CO,na.rm = TRUE),
+                  n3CO = sum(n3CO,na.rm = TRUE),
+                  n4CO = sum(n4CO,na.rm = TRUE),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv,
+                 per4CO= sum(n4CO)/ tot_biv,
+
+                 mean.MLH1.foci = mean(MLH1.sum.check),
+                 mean.MLH1.foci2 = mean(nMLH1.foci),
+                 meanQ = mean(quality)
+
+)
+#this is the table that needs melting
+#but is also easy to redo by hand
+
+#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
+```
+
+```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
+
+#add subspecies to these values
+melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+
+#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+
+melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
+
+#add the subsps
+
+melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
+                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
+
+melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
+
+chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
+
+#need a second DF / the first df
+#+ annotate() text 
+#annotate()text
+
+#proportion_stats_category,  tot_biv, ncells
+#add number of cells and bivalents in the corner
+
+#chrm.class.barplot
+```
+
+```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
+
+#prop.test(x, y)  x number of successes, y total  
+#strain sums of n
+#look up continuity correction
+
+#mouse.level.BivData_table <-- has more male strains
+#Mouse.Table_BivData_4MM <- only sex matched strains
+
+#perform test across strain 
+#removed all lines with no hand.foci.count entry
+Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
+
+                  nbiv = sum(tot.biv ),              
+                 n0CO = sum(n0CO ),
+                  n1CO = sum(n1CO ),
+                 n2CO = sum(n2CO ),
+                  n3CO = sum(n3CO ),
+                 n4CO = sum(n4CO )
+                          )
+
+Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
+
+for(g in 1:length(Chrm.prop_strain_table$nbiv)){
+
+  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
+                                     Chrm.prop_strain_table$n1CO[g],
+                                     Chrm.prop_strain_table$n2CO[g],
+                                     Chrm.prop_strain_table$n3CO[g] ))  
+}
+
+Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
+#there are still some wacky entries -- but much fewer than before
+
+chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
+
+female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
+female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
+female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
+#the pvalues for 0 are not sig dif
+#the p values for 1 and 2CO are sig diff
+
+#seperate by subsp
+chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
+chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
+chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
+
+Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
+Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
+Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
+Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
+#WSB has fewer 2COs than other 3 strains
+
+Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
+Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
+Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
+#KAZ has slightly more 1CO than the other two
+
+#high.female_G.LEW.
+chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
+
+high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
+high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
+high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
+
+#Low.female.G.LEW.
+chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
+
+low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
+low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
+low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
+
+#all females
+
+chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
+male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
+
+male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
+male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
+#the pvals for these 3 class are sig dif
+
+#seperate by subsp
+chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
+chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
+chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
+
+Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
+Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
+Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
+#non of the Dom proportions are different
+
+Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
+Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
+Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
+#all v different
+
+Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
+Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
+Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
+
+#high vs low
+chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
+                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
+                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
+
+chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
+chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
+
+chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
+chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
+
+Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
+Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
+Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
+#the low strains do vary slightly in their 1CO and 2CO proportions
+
+#includes SKIVE
+Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
+Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
+Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
+#all very sig p values
+
+High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
+High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
+High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
+#these proportions are v different
+
+#no SKIVE
+High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
+High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
+High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
+#MSM and PWD only slightly sig dif 2CO
+
+#MSM-MOLF
+MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+#PWD-KAZ
+PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+```
+
+```{r Chrm.props.plot2, echo=FALSE}
+
+#barplots for proportions of chrm classes
+
+#consider adding this the the setup block (could have done this to category level table)
+mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
+
+mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
+
+#DF needs to be melted -- so they all in 1 col
+#per code, I've done this by hand
+
+#subset DF for ease
+skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
+col.names.list.99 <- colnames(skinny.mouse.level)
+
+#all of the cols to use
+MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
+
+               #leave out the Foci
+                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
+
+  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
+
+MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
+MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
+
+#no filename for add mouse, strain subsp
+MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
+
+#then this DF has to be summarized into categories
+chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
+                             mean.percent.type = mean(value)
+                             #sum is the wrong!
+                             #used average percentage, might consider a different summary
+)
+
+#can't apply function use this
+chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
+             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
+                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
+                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
+
+          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
+              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
+
+                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
+                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
+            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
+                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
+                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
+                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
+                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
+                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
+
+             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
+               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
+
+chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
+                                                                 "Musc",
+                                                                 "Mol",
+                                                                 "Spretus","Spic",
+                                                                 "other") )
+#ggplot
+
+chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
+#chrm.class.barplot
+chrm.class.barplot
+```
+
+```{r Chrm.Prop.misc.calulations, echo=FALSE}
+
+total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
+
+#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
+#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
+```
+
+<!--
+- This is motivation to investigate more traits at the chromosome level
+1.	Basic stats - bivalent data numbers 
+2.	Rang of co per chrm, mostly 1co 2co
+3.	High rec strains have evolved a higher proportion of 2CO
+For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
+MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
+4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
+-->
+
+<from the means the expected number of COs per chroosomes ranges from X to X we confirm these numbers Can't have fractions of COs>
+
+In order to decompose the cell wide rate, we decided to look at the proportion of chromosome classes (ie by the number crossovers on the same bivalent).
+A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).
+
+The first attern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes.
+<these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>
+
+<note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?)>
+
+The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`).
+This relpicates what has been notice in the literature (cite stapley).
+
+Most of the bivalents are 1CO and 2CO class bivalents -- and the proportion of these two classes distinguish high and low rec male strains. (distinguish the high rec males from minimal rec groups)
+
+(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class.
+The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.
+*(much smaller proportions and variation in the fringe classes*
+
+**2. most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
+
+**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
+
+PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
+
+While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
+ and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+
+**3.
+these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**
+
+<Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
+They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
+This significant p value indicates that at least one of the six strains had significant differences).>
+
+Breif conclusion: the evolution of genome wide recombiantion rates proceeds through the chromosome level.
+
+<something about the female results>
+
+<SUMMARY the / TRANSITION to single bivalen measures>
+
+# Single Bivalent Level Results
+
+In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining features of the recombiantion landscape at the single bivalent level.
+recombination landscape refers to the i) total length of the SC-AE (potential area for COs), ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.
+
+From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software.
+After the human curation step, `r length(Curated_BivData$Obj.ID)` single bivalent measures were left.
+(make sure these are the hand.foci bivalents).
+
+<how do these agree with other patterns of MLH1>
+
+```{r algorithm_performance_stats, echo=FALSE}
+
+#sort by category, 
+#compare total boxes and rate passed
+#Curated_BivData  -- pass rate?
+
+#number of mice, number of cells
+#mean number of isolate bivs per cell,
+#use these two metrics
+#cell.level.BivData_table$boxs.IDd 
+#cell.level.BivData_table$nboxes_passed
+
+#calq the mean rate by category
+
+cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
+
+algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
+                        n_obs = length( unique(fileName) ),
+                         mean_pass.rate = mean(pass.rate)
+                         )
+
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
+
+algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
+                        n_obs = length(unique(fileName)),
+                        mean_pass.rate = mean(pass.rate)
+                         )
+
+data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
+```
+
+Table X illustrates the break down of single bivalent observations by category. 
+Error compared to human measures and other details have been discribed in (Peterson_2019).
+
+The ranges of of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in MSM male to 
+`r max(algorithm.stat.category$mean_pass.rate)` ) KAZ female, we assume that the isolation process is not biased.
+Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.
+We assume the bivalent datasets across all categories are equally eqivilent for representing general patterns.
+
+<curation, x from curation, XY excluded, replication to hand measures>
+
+We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
+
+<These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.>
+<The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.>
+
+# Q1 Sex Differences in SC-AE Lengths
+
+DO THE HIGH QUALITY dataset tests
+
+<!-- breif background on longer female SC  (mouse, human) Gruhn-->
+
+Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of PWD and MSM.
+A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
+
+<Comparing the means of single bivalents is complicated by the fact we can't identify the chromosome identitied the physical size (in terms of SC-AE) is the closet approximation we can make.>
+
+```{r Q1.Short.Long_biv.setup, echo=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#make the Q1 strain list
+Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+```
+
+We use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data.
+A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes.
+
+<Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes>
+
+This 'short biv' data set has advantages in that it is closer to sampling across a similar set of chromosomes (ideally Chromosomes 15 to 19) and it excludes the XX in females, which is predicted to be the 3rd longest based on the physical (Mb) length.
+
+```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#CLEAN UP THE FULL DATA SET
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+#nested boxplots, pooled Bivalents
+
+Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+
+Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+
+Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+
+```
+
+```{r Q1.SC.show.short.biv, echo=FALSE}
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+
+Q1.SC.short.mouse_scatter_boxplot
+
+#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
+skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+#skiveQ1.SC.short.mouse_scatter_boxplot
+#there are 2 lowish female means,
+
+```
+
+```{r Q1_SC.biv.short_ttest, echo=FALSE}
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+#WSB
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
+#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
+#1 male, 3 females measures
+#try to merge the manual - whole cell observations (at least 2 males)
+
+```
+
+<in the above boxplots of mouse means for shortbivalents>
+
+For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer, the largest difference across sexes by strain are>  (t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
+for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
+
+<EXPLAIN WHY USING TOTAL SC AREA>
+
+An additional measure of the SC-AE we tested is the total SC area per cell (the whole genome).**to get a comparison of the SC-AE area for the whole genome (not just a subset of chromosomes)** We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
+A total of X images were run trhought the image analysis pipeline, and after outliers removed, X remained.
+/ cell based measures were left.
+Mouse means were calculated for their total SC-AE area and are displayed in figure X with the strain means shown as horizontal lines (below).
+
+```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
+
+Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
+
+Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
+   "WSB","G","LEW","PERC", 
+  "PWD", "SKIVE",  "KAZ",
+  "MSM", "MOLF"), order=T )
+
+Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
+
+#order the strains
+Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
+  arrange(strain) %>%
+  mutate(mouse = factor(mouse))
+
+Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
+
+#add category mean
+#totSC.strain.av.table %>% filter
+totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
+
+totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
+
+Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")
+
+Tot.SC_mouse.av_Dom
+
+totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
+totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
+
+Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
+
+Tot.SC_mouse.av_Musc
+
+totsc.mol <- Q1.totSC.mouse.av.table %>% filter( subsp == "Mol" )
+totsc_strain.av.mol <- totSC.strain.av.table %>% filter( subsp == "Mol" )
+
+Tot.SC_mouse.av_Mol <- ggplot(data= totsc.mol, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totsc_strain.av.mol, aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
+
+Tot.SC_mouse.av_Mol
+```
+
+<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )
+
+```{r Q1.total.SC_t.test, echo=FALSE}
+
+#run t-test for male female avergage for each strain
+
+#WSB
+Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
+#G
+Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
+#LEW
+Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
+
+#PWD
+Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
+
+#KAZ
+Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
+
+#SKIVE
+#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
+#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")
+
+#error not enought female observations
+
+#MSM
+Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
+
+```
+
+```{r Q1.totSC_male.first.comp, echo=FALSE}
+
+Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
+
+Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
+```
+
+The most averages for total SC-AE area (skeletonizes) are significantly longer in females for all strains (p=
+`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
+for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).
+
+Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)
+
+The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
+(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
+<But we use this metric anyways>
+
+<We estimate the effect of the sex chromosomes from hand measured data...the XX (3rd longest chromsomes) in females is ~10% of total the summed SC area, the average propotion of the XY in spermatocytes is XX%?
+however the morphology of the XY is highly variable across cells>
+
+The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>
+-- we follow up with models (which focus on sex differences).
+
+<explain why applying models.  We then use models to examine/ think about the evolution>
+
+The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mean mouse total sc and mouse mean short bivalent.
+
+**M3. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
+
+Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
+
+#str(Q1.lmer_totSC.mouse.av.table)
+
+#unorder factors
+Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
+
+#TOTAL SC AREA!!
+lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
+
+lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
+
+lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
+
+lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
+sex.results =  lmer.Q1_mus.av_totSC_sex, 
+               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
+lmer.Q1_mus.av_totSC_strain)
+
+#Q1.lmer.totSC_results
+
+#G
+```
+
+For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed model of mouse.av total sc.
+
+The three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
+sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
+
+```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+
+#Q1lmer.MM_SC_short_results
+
+```
+
+Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+The interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
+
+These effects were followed up the glms, again applied to both mouse means for short biv SC lengths and total sc area per cell.
+
+**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
+#total-SC
+#Q1.lmer_totSC.mouse.av.table
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
+
+#subsp effect, many strain, #male
+#G * male v.sig and male*LEW
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
+#LOT OF SIGNIFICant strain effects in this model
+```
+
+```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
+
+Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+
+#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
+
+```
+
+```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
+
+#summary(Q1.glm_short.SC_M3)
+
+summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
+
+#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#sex male very sig -- only sig
+```
+
+The following effects had significant p values for the **total SC** from the above model
+
+subspecies musc p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`, subspecies mol p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`, sex male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`, strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
+<this is the smallest p vale, elevated female total, driven by the high means in females>
+
+(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
+
+<!--
+for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
+-->
+
+Fr the **short bivalent** sc the model results are listed below. From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).
+
+while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.
+
+In the M3 model (strains and sex as fixed effects), the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).
+
+<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >
+
+<In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
+
+<1. the longer female sc's is consistant and seen across all strains,
+even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
+2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>
+
+```{r show.long.biv, echo=FALSE, include=FALSE}
+#mouse av
+Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
+
+```
+
+```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
+
+#Q1.real.long_mouse_table
+
+Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE
+Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+```
+
+```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
+
+#removing these plots for space
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+## Adjusting for XX
+
+(rethink where this section should go)
+new outline
+
+1. illustrate problem(affects mostly SC length)
+
+2. Expected impact on sex comparisons, estimated effect size of the X  
+3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
+
+4. 19 female, 19male, 20female 20 male
+
+The female mouse averages should have adjustments for the XX.
+working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
+Subtract this amount from the female mouse averages ...
+This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
+
+- Of all female single bivalents observations, 5% are XX (1 of 20).
+
+- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
+
+- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
+
+I think a formula something like this can be applied to adjust for XX
+
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
+
+# short transition
+
+<transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents
+1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
+-- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).
+
+2.
+We focus on two metrics/ aspects i) the placement of single foci along a bivalent (just 1CO) and ii) the placement of two foci on the same bivalent / the interfocal distance of 2CO bivalents (reflects crossover interference)
+
+# Q1. 1CO Position Sex Differences
+
+```{r Q1.F1pos_long.short_setup, echo=FALSE}
+
+#REAL.short_mouse_table
+#REAL.long_mouse_table
+
+Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+```
+
+<!-- motivating
+
+To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
+In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
+-->
+
+We focus on the foci positions from one crossover (1CO n= 
+) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
+
+We anchor the position measures at the centromere (and normalize by deviding the position by the length of the chromosome, so the value range from 0 to 1, with smaller values representing positions closer to the centromere end of the bivalent, and larger values meaning foci positions closer to the opposite end (refered here as terminal or telomeric position).
+
+We plot the mouse averages across all of the single bivalent measures, with the strain averages shown as boxplots...
+
+in order to get a better picture of the sex differences in this foci-position metric
+
+```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
+#main figure of density plots,
+
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+CO.Pos.density_1CO
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+```
+
+```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
+
+Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \n all bivalents")
+Q1.mouse.av.biv_scatr_box
+
+```
+
+```{r Q1_1COpos.t.test, echo=FALSE}
+#testing,
+#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
+
+#WSB
+Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
+#G
+Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
+#LEW
+Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
+
+#PWD
+Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
+
+#KAZ
+Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
+
+#SKIVE
+Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
+           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
+
+#MSM
+Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
+
+#Single Biv Level
+#single.biv.1CO_pos_t.test <- t.test(
+
+ # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],
+
+#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+#  )
+
+#t.tests for mous.av_nrm.1CO.pos 
+```
+
+In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (fitting the predictions/pattern of the 'typical recombination landscape' (Sardell)).
+The single CO positions are medially placed in females and terminall placed in males.
+
+The pavlues across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
+
+With the musculus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
+Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).
+
+After confirming the significan sex differences for 1CO position, we apply the familiar models
+
+**M1. Mixed Model**
+
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
+
+#make sure this dataset has all the right strains
+Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
+lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
+
+lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
+lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
+
+lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
+sex.results = lmer.Q1.nrm1CO_M1.sex , 
+               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
+
+#lmer.Q1.nrm1CO_M1_results
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
+(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
+These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).
+
+**M2. Linear model**
+
+$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
+
+```{r Q1.CO1pos_glm.M2, echo=FALSE}
+
+#in case factors aren't ordered
+#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
+
+#summary(Q1.glm_F1pos_M2)
+
+summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
+#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+**M3. Linear model**
+
+$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
+
+Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
+
+#summary(Q1.glm_1CO.pos_M3)
+
+summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
+#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
+#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+
+```
+
+The male effects were the most significant (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+While the musc subspecies, LEW and KAZ strains had slightly signifcant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
+
+(these results will be explored more fully in the Q2 section).
+The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).
+
+<we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>
+
+Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite).
+To account for this we again use the reduced bivalent sets (long.biv and short.biv)
+
+```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
+
+#real.long.bivData
+
+Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+Q1.short.biv_scatr_box
+
+Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+Q1.long.biv_scatr_box
+```
+
+```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
+
+#REAL.short_mouse_table
+#REAL.long_mouse_table
+#mouse.av.long.biv.table
+
+Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+
+Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+```
+
+These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
+They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
+MOLF males have much more medial positions than other strains.
+
+<!--
+### Siscoten
+The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
+The logic of how the sis-co-ten metric is outlined in the figure below.
+The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
+This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
+
+![from (Lee, J.
+(2019).
+Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
+Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
+
+```{r show.siscoten, echo=FALSE}
+#think of a good way to display this
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_female
+
+sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+sis.co.ten.plot_WSB_male
+
+sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_female
+
+sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_LEW_male
+
+sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_female
+
+sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+#sis.co.ten.plot_PWD_male
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDf
+
+jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_PWDm
+
+jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBf
+
+jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
+  geom_jitter()+facet_wrap(sex~strain)
+#jitter_siscoten_WSBm
+```
+
+Males have much clearer separation of siscoten across chrm classes.
+This is emphasized when SC length is also plotted.
+It seems like musc males have higher amounts of this metric compared to Dom males.
+To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
+BUT females have a greater range -- so maybe it's just a scale issue.
+
+```{r show.norm.siscoten, echo=FALSE}
+Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
+
+Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_WSB
+
+Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_G
+
+Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_LEW
+
+Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_PWD
+
+Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_KAZ
+
+Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_SKIVE
+
+Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
+Nrm.sis.co.ten.plot_MSM
+
+```
+
+I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
+
+```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
+
+#glm for siscoten
+#mouse.avs_4MM  - mouse averages with mean sis.co.ten
+#MM
+lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+#drop1(lmer_MM_siscoten, test="Chisq")
+
+rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
+rand_siscoten_MM
+#slightly sig
+```
+
+```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
+
+#move this to the set up blocks
+
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
+                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
+
+#test high vs low Musc strains
+#mouse.avs_4MM
+```
+
+```{r log.reg_siscoten, echo=FALSE}
+
+#replace the table
+
+male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
+                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
+                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
+                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
+
+#SISCOTEN
+#mouse av, norm.first foci is from 1CO bivalents
+
+log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
+
+                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+summary(log.reg_SIS.co.ten_mouse.av)
+
+#single biv level
+log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Sis.co.ten_biv)
+
+```
+
+```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
+
+#bivData --
+
+t.test_mouse.av_siscoten <- t.test(
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
+
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#1CO
+t.test_mouse.av_siscoten_1CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#2CO
+t.test_mouse.av_siscoten_2CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+#3CO
+t.test_mouse.av_siscoten_3CO <- t.test( 
+  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
+  )
+
+####
+#Single Biv Level -- from the main 
+#pull the single BivData -- MUSC
+
+#check that these have the extra male strains
+
+t.test_single.biv_siscoten <- t.test(
+
+  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+#1CO
+mouse.avs_4MM$mean.siscoten_2CO
+
+```
+
+All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
+
+### Telomere and centromere Distance
+
+My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
+In the plots below each point is a single bivalent.
+I choose not to use the mark for centromere because it seems noisy and inconsistent...
+
+```{r telo.dist, echo=FALSE}
+
+sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
+  geom_point()+facet_wrap(sex~strain)
+
+Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
+
+telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
+
+telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
+telo_jitter
+
+nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
+nrm.telo_jitter
+
+#why do a different number of rows removed each time??
+```
+
+Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
+In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
+In females the ranges of telomere distances have much more overlap.
+
+```{r MM.telo, echo=FALSE}
+#HetC
+
+lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
+drop1(lme_MM_telo.dist, test="Chisq")
+as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
+
+#sex and sex*subsp interaction effect are significant for mean sis.co.ten
+
+lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+#anova(lmer_MM_siscoten)
+
+#thing1  <- summary(lmer_MM_siscoten)
+drop1(lmer_MM_telo.dist, test="Chisq") 
+rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
+rand_telo.dist_MM
+```
+
+Mixed model result summary:
+
+```{r log.reg_telo.dist, echo=FALSE}
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
+
+#TELO.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_Telo.Dist_mouse.av)
+
+#single biv level
+log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist_biv)
+
+#PER versions
+log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_Telo.Dist.PER_biv)
+```
+
+```{r t.test_Telo.Dist, echo=FALSE}
+
+#mouse av
+t.test_mouse.av_telo.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Telo.Dist <- t.test(
+
+  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Telo Dist PER
+single.biv_t.test_Telo.Dist_PER <- t.test(
+
+  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+```{r cent.dist_show, echo=FALSE}
+cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
+
+cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
+cent.dist_jitter
+
+norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
+norm.cent.dist_jitter
+
+```
+
+```{r MM.cent.dist, echo=FALSE, eval=FALSE}
+
+lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
+as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
+
+lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+exactRLRT(lmer_MM_cent.dist)
+```
+
+```{r log.reg_CENT.dist, echo=FALSE}
+
+#CENT.DIST
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
+
+                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_CENT.Dist_mouse.av)
+
+#single biv level
+log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist_biv)
+
+#PER versions
+log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+summary(log.reg_CENT.Dist.PER_biv)
+```
+
+```{r t.test_CENT.Dist, echo=FALSE}
+#CENT.DIST
+#mouse av
+t.test_mouse.av_Cent.dist <- t.test(
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
+
+  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
+  )
+
+#Single Biv Level
+single.biv_t.test_Cent.Dist <- t.test(
+
+  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
+  )
+
+#Cent Dist PER
+single.biv_t.test_Cent.Dist_PER <- t.test(
+
+  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
+
+   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
+  )
+
+```
+
+The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
+Females have more overlap in the distributions of centromere distances across chromosome class compared to males.-->
+
+```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
+CO.Pos.boxplots_1CO
+
+```
+
+```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
+#IGNORE THIS FOR KNOW, THIS
+
+#filter all Curated biv data for 1 strain
+#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
+
+# the stardell paper sums the number of COs for bins (10)
+
+#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
+#but then the model would be close to logistic
+
+#summing the number of COs per bin, losses the chrom specific data....
+#!! need to use a melted Data frame to capture all nrm.foci obs
+#wsb.ttest<- mouse.avs_4MM %>% filter()
+#start with same chrm classes
+#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
+
+WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
+
+Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
+
+PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
+
+Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
+Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
+Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
+
+anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
+
+anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
+
+anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
+
+#or AIC(model1,model2)
+
+```
+
+```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
+
+#make a table for the sex specific values of the normalized F1 positions
+
+#for strain, sex -- Curated_BivData
+#or (Mouse.Table_BivData_4MM)
+#norm.first.foci.pos
+Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
+
+                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
+                  n1CO = sum(n1CO )
+
+)
+
+```
+
+```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#MELT_Curated_BivData
+
+#these isn't long.biv data
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+real.long.bivData
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
+
+MELT_long_biv <- add_sex(MELT_long_biv)
+MELT_long_biv <- add_strain(MELT_long_biv)
+MELT_long_biv <- add_subsp(MELT_long_biv)
+MELT_long_biv <- add_category(MELT_long_biv)
+
+long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
+
+Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Dom.male_MELT_long
+
+long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
+
+Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
+Musc_male_MELT_long
+
+#there are v few observations in this dataset
+```
+
+```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
+
+#LOGISTIC REGRESSION for FOCI POSITION
+
+#mouse av, norm.first foci is from 1CO bivalents
+log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
+                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_mouse.av)
+
+#single biv level
+log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
+summary(log.reg_1CO_pos_biv)
+
+```
+
+#Q1. Sex Differences in CO Interference (IFD)
+
+In order to test if there are sex differences between crossover interference, a major determinant of the positioning of crossovers along chromosomes, we examined ...
+ran models with the interfocal distance (IFD) of two foci on the same bivalent.
+
+-We focus on observations from two-crossover bivalents for more comparable observations.
+-comparisons of all IFDs of mulit crossover bivalents is also included
+Interference is a major determinant of the positioning of chromosomes
+
+<!-- 
+Brief lit review for IFD / interference expectations  
+**Petkov et al 2007,  CO interference underlies sex differences in RR**
+
+"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
+However, the interference distance is the same in terms of bivalent length.
+We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes at the pachytene stage of meiosis."
+
+Chrm1 genetic map, 
+(other human and mice refs 
+Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
+
+DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
+<!--
+ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
+
+explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
+-->
+
+<!--  set up -->
+
+<!--  plots-->
+
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+first.IFD.plot_ABS
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+first.IFD.plot_PER
+#+theme(legend.position="none")
+```
+
+```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
+#long SC Long_biv_mouse.avs_4MM
+#real.long.bivData
+#Only use 2CO bivalents
+
+real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
+real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
+
+long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+
+long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
+```
+
+<!-- EXPLAIN WHY WE SHOULD CARE ABOUT RAW AND PER IFD values-->
+
+The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical - force mediated through the SC -- In some models this metric should be thought of as the best metric for charaecterizing crossover interference -- bc it reflect the physcial --
+
+The above plots display the raw and normalized area which seperate two foci on the same chromosome.
+In the normalized plot, 9 observations fall above the length threshold of 150.
+The female measures fall more uniformly across the range of (0 to 1).
+
+In the abosolute plots, you can start to see clustering/ enrichment of IFD values between 40 and 80 pixels for PWD and SKIVE, 
+(and maybe slight clustering for MSM).
+
+<!--t.tests -->
+
+t.tests are performed on the normalized metrics for establish if there are sex differences in the IFD meterics (raw and normalized).
+
+```{r Q1.IFD_Ttests, echo=FALSE}
+#below are calq just from 2COs, have no fear
+#mean_IFD.2CO_ABS
+#mean_IFD.2CO_PER
+
+#FULL all strains
+Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
+#p-value = 0.07
+
+Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
+#p-value = 1e-12
+
+#WSB
+Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
+                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
+
+#LEW
+Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
+
+#G
+Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
+
+#PWD
+Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
+
+#KAZ
+Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
+
+#SKIVE
+Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+
+#MSM
+Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
+
+#FULL all strains
+Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
+#p-value = 0.3
+
+```
+
+The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
+BUT WHEN G IS REMOVE, the pvalue is no longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
+<look closer at G>
+
+The normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`) between sexes, indicating there a
+
+When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
+MSM   `r `
+SKIVE  `r `
+PWD  `r `
+MSM  `r `
+LEW  `r `
+G  `r `
+WSB  `r `
+
+This indicates a general pattern of two foci are seperated by more area (hance have stronger interference) in males, this pattern is also for genetic maps.
+
+<!-- short and long -->
+
+```{r Q1.IFD.long.short.setup, echo=FALSE}
+
+#real.long.bivData
+#real.short.bivData
+
+Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.long <- add_sex(Q1.IFD.long)
+Q1.IFD.long <- add_strain(Q1.IFD.long)
+Q1.IFD.long <- add_subsp(Q1.IFD.long)
+
+Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
+Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
+
+Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.short <- add_sex(Q1.IFD.short)
+Q1.IFD.short <- add_strain(Q1.IFD.short)
+Q1.IFD.short <- add_subsp(Q1.IFD.short)
+
+Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
+Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
+
+#t.test. mouse av
+Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+
+Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+#both the short and long norm IFD are NS
+
+#t.test Biv
+Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+
+Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+```
+
+```{r Q1.show.IFD.things, echo=FALSE}
+nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
+#looked up image, measures confirmed
+
+####
+nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#21may18_KAZ_m2 -- seems like it might be an outlier
+#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
+#second foci slightly smaller, kinda fainter -- but also much brighter than background
+
+#check all KAZ males
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ.female_BIV.data
+
+nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD.female_BIV.data
+
+nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_LEW.female_BIV.data
+
+nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G.female_BIV.data
+
+nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+###look at why KAZ males are weird
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#the difference the the Nrm IFD cut off is striking with these plots
+
+nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#this might
+
+nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+```
+
+see the reference plots within this dir, --
+
+this section -- all strains but KAZ, males have a higher mean normalized IFDs
+
+the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs
+
+The KAZ strain is the only strain without significant p value for the normalized IFDs, due to KAZ males having a higher rate (enrichment of short normalized IFDs.
+
+There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
+off.
+Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?
+
+<The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes?
+they were supposed to be 1CO -- but somehow a second CO formed anyways ..
+and very close by
+
+Where as in PWD, SKIVE and MSM there ARE NO 2CO bivalents which have such short IFD
+WHAT espicially werid about this --
+
+whats the expectations following the confonding proportions? 
+high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.>
+
+**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference
+
+to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
+Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)
+
+-- run these comparisons to see if the small NRM IFD are due to shorter chromosomes, which are predicted to have weaker interference.
+
+Looked at the KAZ pattern
+
+Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
+These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).
+
+Considering that the short-long bivData is much smaller than the full data set
+
+<!-- lmer/ glms-->
+
+### Mixed Model Tests, Fixed Effects
+
+Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+
+**M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+```{r Q1.IFD.M1, echo=FALSE, message=FALSE}
+#get a list of the old mice to double check
+#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
+#MIXED MODEL
+
+#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
+
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
+
+#M1_IFD_2CO
+lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
+               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
+               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
+#don't understand why 2 rows?
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
+#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
+
+#these 2 below are LR tests
+#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
+#exactRLRT(lmer.IFD.PER_MM.M1)
+
+```
+
+The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
+That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
+
+I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
+The tables below are from anova( for the lmer model ).
+Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
+
+**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
+
+```{r IFD_HetC_M2, echo=FALSE}
+
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+
+summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
+#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+
+summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
+#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+**M3. Linear model**
+
+$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
+
+```{r IFD_HetC_M3, echo=FALSE}
+
+glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
+
+summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
+#round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
+round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+
+```
+
+For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.
+
+the interaction effects were slightly significant for both raw and nrm.IFD.
+
+The most significant value was the sex effect for nrm.IFDs.
+
+The the random strain effect was not significant for either model.
+
+For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
+The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
+For the M3, only SKIVE*male effect is close to significant
+
+For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
+SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure.
+
+Overall There's a low amount of significant effects across the 2CO IFD measures.
+This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
+
+Dive deeper into the sex specific pattern for each strain.
+Below are code chunks which show the unusual sex specific results for IFD measures.
+The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
+The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
+
+For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
+
+```{r Comp_Ratios, echo=FALSE}
+
+#calq ratios of belo 30
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
+
+#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
+MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
+
+DF.strain.list <- list(
+  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
+  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
+  MSM.Bivalents.DF, MOLF.Bivalents.DF
+  )
+
+entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
+        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
+          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
+
+i = 1
+for(x in (DF.strain.list) ){
+
+  #print(head())
+
+  #these filters won't apply to the indexed DF
+  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
+  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
+  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
+  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
+  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+  i = i+1
+
+  }
+#entr_DF
+
+entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
+
+```
+
+Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
+
+# transition Q1 to Q2
+
+we tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.
+
+# Q2 SC Length
+
+<put in plot for total SC and mean gwRR count>
+
+<motivation: the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
+
+<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
+Mainly that 1CO are shorter in the high rec strains then low rec strains.
+This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
+Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
+This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
+They have a lower probability of 2COs below a certain SC length threshold.>
+
+while testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR
+
+<!-- basic stats
+`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
+X observations removed for having extra signal from other cells.
+`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
+
+ place set ups for total SC, short biv and long biv here -->
+
+```{r Q2.SC.short.long.setup, echo=FALSE}
+
+#table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+
+Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                         #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+                         )
+
+Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
+Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
+
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                        #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
+                         ) )
+
+Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
+
+Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
+Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
+
+Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
+```
+
+<!-- plots -->
+
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
+
+#nested boxplots, pooled Bivalents
+#x axis for hand.focicount -- good for seperating the points --- but it pools
+
+#prediction for Dom strains: low variation 
+Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
+
+#+theme(legend.position="none") 
+```
+
+```{r Q2.SC.long.biv_plots, echo=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#real.long.bivData
+#real.short.bivData
+
+shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+```
+
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
+
+#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, mouse) %>%
+#  mutate(file.name = factor(file.name))
+
+mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
+
+mouse.av_totsc_male
+```
+
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE}
+
+#make plot of mouse averages for MLH1 and total SC
+
+bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
+
+X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+
+X.by.texas
+```
+
+```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
+
+#investigating potential skew / technical effects in the algorithm
+
+Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+#not sure what my expectations for this plot are,
+
+median(bep$oldest.image.date) #"2017-08-03"
+#there might be an issue with the algorithm performance 
+#given the parrelle diagonal lines -- 
+#this might be due to the dates imaged  2017-02-07
+
+bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
+
+Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
+
+#there's evidence that the algorithm might have some biased effects -- 
+#that will be investigated later.
+
+```
+
+<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
+remember the pairwise t.test
+Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
+G-WSB is sig, WSB-LEW are slightly sig
+-->
+
+```{r Q2.totSC.ttest, echo=FALSE}
+#totalSC high and low rec group
+
+Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
+                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+#p-value = 0.007
+
+#mol
+Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+
+#musc
+Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+
+```
+
+```{r Q2.short.long.ttest, echo=FALSE}
+
+#long and short 
+#mouse.av.short.biv.table
+#Q2mouse.av.short.biv.table
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
+
+#p-value = 0.9
+
+#mouse.av.long.biv.table
+Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
+#p-value = 0.2
+
+#mUSC-short
+Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-short
+#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
+
+#musc-long
+Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-long
+#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
+
+#not enought observations
+
+```
+
+```{r Q2.SC_HvL_allBiv, echo=FALSE}
+
+#High-Low_All-BivData
+
+Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#short
+Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#long
+Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+```
+
+**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
+`r Q2tot.sc_t.test_High.Low$p.value`).**
+
+Molossinus have significant difference between mouse means for total sc, between the high and low strains (mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
+The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
+
+#Dom-Musc
+Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
+                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
+#p-value = 0.5
+
+#Dom-Mol
+Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
+                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
+#p-value = 0.1
+```
+
+```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
+
+#DOM-musc-long
+Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.08
+
+#DOM-musc-short
+Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.09
+
+```
+
+```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
+
+#DOM-MUSC-ALL
+Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 4e-13
+
+#DOM-MOL-ALL
+Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 6e-10
+
+#DOM-musc-long
+Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.2
+
+#DOM-musc-short
+Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.06
+
+```
+
+<across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).>
+
+<!-- models -->
+
+```{r Q2.totSC.glm, echo=FALSE}
+#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel
+
+Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+
+sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
+
+#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
+#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
+#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)
+
+Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+
+sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
+#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
+```
+
+```{r Q2.glm.SC.long, echo=FALSE}
+
+#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
+#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
+#table(mouse.av.long.biv.table$mouse)
+#Q2mouse.av.long.biv.table
+
+glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+
+sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
+
+glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+
+sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
+
+#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+```
+
+```{r Q2.glm.SC.short, echo=FALSE}
+#Q2mouse.av.short.biv.table
+#mouse.av.short.biv.table
+
+glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
+sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
+
+#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
+
+glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
+glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
+```
+
+```{r Q2.SC.long.glm.bivglm, echo=FALSE}
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
+
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
+#summary(glm_SC.male_Musc_long.biv)
+
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
+
+## logistic regression for Rec group
+#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
+#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
+
+#summary(log.SC.male_long.SC)
+
+log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+
+```
+
+**For the total SC glm**
+
+<Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
+1. no difference in DOM strains (for SC lengths )
+2. PWD > SKIVE > the rest
+3. Msm > mol>
+
+For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
+
+In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+
+<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
+
+<For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW, G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.>
+
+<!--24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM-->
+
+<just like the sex differences in SC area -- this points to differential usage of the SC for placing CO -- investigate the recombination landscape>
+
+<!--
+blocks of other male Q2 - SC lengths
+what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
+The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
+-->
+
+**The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).**
+
+**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
+The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
+
+**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
+
+> **Summary Points**
+
+- confirm that the positive correlation of SC and CO number is maintained
+
+- Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.
+
+- (the glms, indecate there are many strain effects), Using the mouse average SC lengths, in the full data set all strain effects are significant.
+This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
+
+- For the M3 models of mouse av.s across SC meterics, for many of the SC metrics, most all of the strain effects are significant
+
+- for the M2 models of the mouse av, musc - mol are significant fixed effects
+
+- many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging
+
+- th predictions for M3 (strain) differencs (motivated by the gwrr variation are not met,  (correlation of SC means and CO means?))
+
+<Each point is a mouse average SC.
+When all bivalents are pooled, the logistic regression SC lengths predict the groups.
+When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
+For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.>
+
+<!-- junk chunks below -->
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
+#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+#there's no rec group for this table
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+<These are the plots of the logistic regression.>
+
+```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
+
+plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SClength.ttest, echo=FALSE, include=FALSE}
+#COMPARISONS ACROSS ALL CLASSES
+
+#MOUSE AVE, all
+Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
+
+#mouse.av 1CO
+Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
+
+#MOUSE.av 2CO
+Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
+
+#single biv level
+#male.biv?
+
+#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value
+
+```
+
+```{r log.reg_SC.length, echo=FALSE, include=FALSE}
+#REMOVE THIS COMPARES FROM FULL BIV DATA
+
+#male_mouse.av
+#Male.poly.Mouse.Table_BivData_4MM
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+#male poly shouldn't be here
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
+                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
+                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
+
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
+
+#1CO
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_1CO)
+
+#2CO
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_mouse.av_2CO)
+
+#3CO
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_3CO)
+
+# bivalent level log regression
+
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV)
+
+#1CO  (BIVDATA)
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_1CO)
+
+#2CO
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_2CO)
+
+#3CO
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV_3CO)
+```
+
+```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
+#REMOVE THIS, IT uses total bivalent data
+
+#rerun the models for datasets -- deviding by chrm class and rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
+
+Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
+Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
+
+#calq_mouse av
+mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        n_cells = length(unique(fileName) ),
+                        mean.SC.total = mean(chromosomeLength),
+
+                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+                         )
+
+mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
+mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
+
+mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
+mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
+
+#calq_mouse av
+mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
+
+mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
+mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
+```
+
+```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
+male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
+                            subsp.y,subsp.x,
+                    category.y,category.x, strain.x,
+                    strain.y, mouse.1  )  )
+
+male_mouse.av <- add_sex(male_mouse.av)
+male_mouse.av <- add_strain(male_mouse.av)
+male_mouse.av <- add_subsp(male_mouse.av)
+male_mouse.av <- add_species(male_mouse.av)
+
+fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
+fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
+fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
+fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
+
+table( droplevels(fomo.male_mouse.av$strain) )
+#there are cast strains too
+
+Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+musc.male.plot
+
+# + theme(axis.title.x=element_blank(),
+#        axis.text.x=element_blank(),
+#        axis.ticks.x=element_blank()) +
+#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+
+#make the Mol and
+Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+
+Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Mol.male.plot
+
+#Dom version 
+Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
+Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
+Dom.male.plot
+```
+
+```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
+## run the tests to check following expectations
+
+#1.Dom strains not different from ech other
+#2. PWD and SKIVE are different from the other strains
+#3.MSM and MOLF are different from each other
+
+#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
+
+male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
+#the subsp labels get messed up, unless they are designated chr first
+Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
+
+Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
+Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
+#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
+#total sc sig, 1CO sig, 2CO sig
+
+Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
+#0.000369 ***
+Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
+#0.000816 ***
+Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
+#0.0148 *
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
+#making plots above show that mouse average SC lengths are lower could there be microscope issue?
+
+#make sure to remove some of the juvi mice!! from the AP_mouse table
+show.Dom.male.CO <- ggplot(
+male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
+
+Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
+#P not sig for CO mean
+
+#seems like there are more mice in the MLH1 data set than bivData
+#
+
+#WSB all dissected and imaged in '15
+#G imaged 15 - 16
+#LEW imaged 15 to 17
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "1CO SC Length", xlab = "")
+
+ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
+          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
+          ylab = "2CO SC Length", xlab = "")
+```
+
+```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}
+
+#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
+#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
+
+male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
+male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
+male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
+
+Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
+#PWD and MSM are most diff from SKIVE?
+
+Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
+#the pooled SC value tests are sig
+
+#the mol comparisons are different due to adj p values
+
+Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
+#G-WSB is sig, WSB-LEW are slightly sig
+
+Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
+#no sig (only 3) -- does include CZECCH
+
+#non-sig, with mol pooled with musc
+
+Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
+
+Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
+
+Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
+#non of these sig with the mulitple adjust
+
+Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
+
+#pairwise.t.test(Ozone, Month, p.adj = "bonf")
+```
+
+# brief transition, rec landscape
+
+the lack of predictive power of SC length to predict CO number indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.
+
+# Q2 Normalized Single CO positions
+
+<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->
+<!-- set up chunks here -->
+<!-- plots /figures below-->
+
+```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
+#Curated_BivData_male
+
+#boxplots of nrm F1
+male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos All bivalents")
+Q2.show.nrmF1
+
+nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
+
+```
+
+<not sure what the beenfit is to test for chrm effectss>
+
+```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
+
+Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Nbiv  = length(Obj.ID),
+                       mean.1CO_PER = mean(PER_Foci_1)
+                      # mean_CO = mean()
+                       )
+
+Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
+Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
+
+Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
+
+```
+
+The previous section found that the single crossover landscape is significantly different between males and females.
+In this section we focus on differences between the high and the low rec male strains
+In order to test if if there are distinguishing meterics for these two groups.
+
+<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->
+
+```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
+#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
+
+#High = MSM, PWD, SKIVE
+Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+#p-value = 1
+
+#high = PWD and MSM
+Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
+#p-value = 1
+
+```
+
+```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
+
+#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
+
+#Musc
+Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+# p-value = 0.4
+
+#Mol
+Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+# p-value = 0.07
+# exclusive rec group not sig
+```
+
+```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
+#these don't have mean.first.foci
+
+#short -- real.long.bivData -- REAL.short_mouse_table
+Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
+
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
+#p-value = 0.3
+
+Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
+# p-value = 0.1
+
+#remove SKIVE??
+```
+
+```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
+
+#real.long.bivData
+#single biv level
+Rec.group_LONG.BivData_F1post_Ttest <- t.test(
+  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
+(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
+#p-value = 0.5
+
+Rec.group_BivData_F1post_Ttest <- t.test(
+  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
+(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
+#p-value = 0.2
+
+#
+#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
+#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
+#p-value = 0.2
+
+#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
+#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
+#p-value = 0.4
+
+```
+
+**The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**
+
+T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
+
+<further tests -- within the subspecies, musculus and molossinus?
+
+The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
+
+While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`).
+This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.
+
+ <BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF, 
+ (MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.>
+
+<long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`> 
+<!-- GLMs because there's no difference 
+toal data set, short and long biv. There's only a difference between MSM and MOKF
+test ppredictions, Dom = NULL,  -->
+
+```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
+
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
+
+Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
+#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
+
+#with.chrm.legth  ?
+Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+
+#M2
+Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
+
+```
+
+in addition models which use mouse average of foci position as a dependat variable, we decided to run models for the single bivalent level, and which include chromosome length as a fixed effect.
+
+in the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
+..
+the chromosome length has the most significant effect, LEW by chrm effect are significant
+
+In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects
+
+**For the M2 (subsp)**
+There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).
+
+LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).
+
+But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).
+
+and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)
+
+**For M3 (all strains)**
+
+strain G (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][2]`).
+
+strain Lew (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][3]`).
+
+strain PWD ( p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][4]`).
+
+slight strain MSM  (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][5]`).
+
+very sig strain MOLF (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][6]`).
+
+strain SKIVE (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][7]`).
+
+strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
+
+from the main figure above, you can see that G and MOLF 1CO foci positions are the most different
+
+<no prediction for the strains within subspecies>
+
+For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
+
+there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
+(motivation for MOLF female observations and more male MSM biv measures)
+
+```{r Q2.1COpos.show.MOL, echo=FALSE}
+
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
+
+mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
+                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
+
+#+facet_wrap(~mouse, scales = "free")
+
+single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
+                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
+
+#+scale_color_manual(values=colors_of_strains)
+
+single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
+                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
+
+#CHRM.length is very significant-- some of the same subsp and strains are sig     
+```
+
+<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.
+
+**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$>
+
+```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
+#remove this cell,
+
+Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
+
+Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
+
+#making new models
+
+Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
+Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
+
+chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+
+sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
+#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
+
+chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+
+#summary(chrm.1CO.pos.chrmLength_strain)
+
+```
+
+<for the first model,
+
+chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`
+
+subsp MOL `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][4]`  
+the mol obj above isn't showing the reight value.
+
+strain G   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][5]`
+
+strain LEW   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][6]`
+
+Strain MSM `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][8]`
+
+chrm * LEW `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][14]`
+
+chrm *MSM  `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][16]`
+
+M2
+
+-chromosome length, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][2]`
+
+-strain G, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][3]`
+
+-strain LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][4]`
+
+-strain MOLF, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][7]`
+
+-strain SKIVE, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][8]`
+
+-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`>
+
+males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
+
+There are significant difference across the 1CO positions, yet they don't match fit any of the predictions. there are also complications due to the chromsome class proportions that
+
+```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
+#isolate all musc
+#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
+```
+
+```{r Q2.1COpos.glm, echo=FALSE, eval=FALSE}
+male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
+
+Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO,  family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm)
+
+#Musc
+Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
+              family=binomial(link="logit"))
+#summary(Log.Q2.F1_PER_Musc)
+
+#Mol
+Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
+              family=binomial(link="logit"))
+
+#summary(Log.glmQ2_M2.1COnrm_Mol)
+
+```
+
+# Q2 Evolution of interference is associated with genome wide recombination rate evolution
+
+```{r Q2.IFD.long.short.setup, echo=FALSE}
+#long BivData
+male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
+
+#show
+Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+
+```
+
+<!-- set up chunks here-->
+
+```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
+
+#subset
+male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
+
+biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
+
+biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none", axis.text.x = element_blank())+ylim(c(0,1.1))
+
+biv.IFDs_PER_by.group
+
+```
+
+<Previous prediction for differences in the long and short bivalent dataset.
+Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+The section on IFD only looks at the 2CO bivalents.>
+
+```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
+
+#t.test for IFD
+#mouse.av - the mouse average IFD is from just 2CO
+t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
+
+```
+
+```{r Q2.t.test_IFD,echo=FALSE, eval=FALSE}
+
+#already made
+#subset
+male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+#t.tests for the Rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#THIS SHOULD ALREADY BE ADDED
+
+#single biv level
+Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
+                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
+
+Rec.group.BivData_IFD.PER_Ttest$p.value
+#p-value < 2.2e-16
+
+#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
+Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
+                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
+
+Rec.group_IFD.PER_Ttest$p.value 
+#p-value = 6.872e-07
+
+```
+
+The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The pattern is confirmed with t.test comparing groups within the same subepcies.
+
+<glms>
+
+The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
+
+**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
+
+**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
+
+```{r Q2.IFD.glm.test, echo=FALSE}
+#REPLACE Male.poly with ..
+#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
+
+#M1.ABS
+Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
+#summary(Q2.M1glm_IFD.raw)
+
+#M1.PER
+Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
+
+sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
+#summary(Q2.M1glm_IFD.PER)
+
+#M2.raw
+Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+#summary(Q2.M2glm_IFD.raw)
+
+Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+#within subspecies
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M2_IFD_Mol)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M2_IFD.PER_Mol)
+```
+
+In M1, the model build on the mouse averages of IFD metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]`  for PWD, MSM and SKIVE respectively).
+In M2, the mouse average for normalzied IFD is modeled with only strain effects.
+Only PWD and SKIVE IFD^PER^ had significant strain effects (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]` for PWD and SKIVE respectively).
+
+<when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF>
+
+<!-- we ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length.
+
+**model BivLen.M1 ** $$Single bivalent\ IFD ~=~ subsp * strain * chromosomeLength + \varepsilon $$
+
+**model BivLen.M2 ** $$Single bivalent\ IFD ~=~ strain * chromosomeLength + \varepsilon $$
+
+```{r Q2.IFD.glm.chrmLength, echo=FALSE}
+
+#repeating chrm models with chromsome length
+#FOR BIVALENT LEVEL MODELS
+
+Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
+
+#M.Len.ABS
+Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.raw)
+
+#M.Len.PER
+Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.PER)
+
+#M.Len.raw
+Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
+
+##M.Len PER
+Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+```
+
+For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs).
+Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^.
+The Mol\*chrmLength is negative
+<PWD, MSM, MOLF, SKIVE are the only significant strain effects.
+- all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
+With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
+All of the high rec strains and MOLF have significantly longer IFD^PER^.
+There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
+-->
+
+```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
+
+#calculate the rates of 'close foci' (<30%)
+#examine the changes in the distribution closer.
+
+#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
+
+close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                      n.close.foci = length(which(IFD1_PER <= .3)),
+
+                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                      rate.close.foci = n.close.foci / N.obs
+)
+
+close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
+
+close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("strain"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER),
+
+                      inverse.close.foci = length(which(IFD1_PER > .3)),  
+                      n.close.foci = length(which(IFD1_PER < .3)),
+                      rate.close.foci = n.close.foci / N.obs,
+
+                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
+
+)
+#the
+
+show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
+
+#add rec.group
+close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
+
+#POOL all short and 
+#chi square test
+high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
+                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
+#p-value = 0.03, slight dif tween 3 high rec strains
+
+#don't need to run chi square for this pattern, since it's enough to
+
+```
+
+We determined that the main difference in IFD disbribution across these groups is an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC length in low rec strains.
+The rate of these close foci range from 8.2% (G) to 16% (KAZ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for skive, MSM and PWD respectively).
+<no chi sqaure test see above>
+
+```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
+
+#Log regression all show significant prediction
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
+
+#Full dataset
+
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+```
+
+ **(unlike total SC length and 1CO pattern, IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.
+However, they go agaisnt a standard perdiction for the relationship between the chromosome level recombination landscape and genome wide recombiantion rate.**
+
+<!-- 
+Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
+But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
+-->
+
+<Below are examples of plots of SC length distributions across cells.
+The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
+Most plots excluded for space.
+
+Each point is a bivalent plotted by cell on the x axis.
+X's are the 4th quartile, big point is the mean and smaller black point is the median.
+I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
+(the extra stats are not correctly mapped)>
+
+<HAND measure METRICS, below chunk goes with the above chunk>
+```{r long.biv.plots, echo=FALSE, eval=FALSE}
+
+show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
+
+show.the.cells_females_dom
+#PWD seems to have the biggest variance across cells
+
+almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+  scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
+almost_cells_DOM_f
+
+#### Musc female
+
+show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
+
+almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
+
+### Dom Male
+show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
+
+almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
+
+#MUSC male
+show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
+
+almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
+
+  geom_point()+ 
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
+  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
+
+#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+
+#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
+
+#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
+
+#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
+
+ # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
+#show.the.cells_male
+
+#almost_whole_cell_BivData
+#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
+#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
+#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
+
+#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
+```
+
+<All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
+It shows the amount of within cell variance across strains.
+There is a bit of variance across the SC length distributions in the PWD females.
+This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
+The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
+The full data set is `r length(Curated_BivData$Obj.ID)`.
+This is the breakdown of bivalent observations by category for this long dataset are
+
+**- Try to merge this DF with the whole.cell manual measures.**  
+- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.  
+in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.>
+
+```{r show.long.SC, echo=FALSE, eval=FALSE}
+
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
+
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
+
+```
+
+<ABOVE plots show the SC lengths for the 'long SC data set'.
+They are supposed to be the longest 4-5 SC from cells where I could get good measures.
+These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
+Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
+
+<!-- unused chunkcs -->
+
+```{r SC.distb_manual_Auto, echo=FALSE, eval=FALSE}
+
+#calq the mouse means
+manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
+                        median_SC = median(SC.length, na.rm = TRUE),
+                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
+#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
+
+)
+
+manual_summary_stats <- add_mouse(manual_summary_stats)
+manual_summary_stats <- add_strain(manual_summary_stats)
+manual_summary_stats <- add_sex(manual_summary_stats)
+manual_summary_stats <- add_subsp(manual_summary_stats)
+
+###
+#calq the mouse means
+Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        median_SC = median(chromosomeLength),
+                        Qar.4 = quantile(chromosomeLength)[[4]],
+
+                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
+
+)
+Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
+Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
+
+#colors_of_strains
+#scale_color_manual(values=colors_of_strains)
+
+#un-order these factors
+```
+
+<!-- Whole Cell Manual Comparison -->
+
+```{r XX.adj_manual.whole.cell, echo=FALSE, eval=FALSE}
+
+#load and examine the the whole cell hand measures here
+
+female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
+data.frame(table(droplevels(female.whole.cell$category)))
+
+# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
+
+#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
+
+data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
+#missing some strains, most have 3 cells (min). MSM has 5 full cells
+
+#remove the incomplete LEW female cell
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
+
+female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
+
+#calq mean SC lengths by rank
+
+mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
+
+                       mean.sc =   mean(SC.length, na.rm = TRUE),
+                         nbiv = length(Obj.ID),
+                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
+                         se.SC_length   = sd.SC_length / sqrt(nbiv)
+)
+         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
+
+top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
+
+show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
+    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
+                    position=position_dodge(width=0.7), 
+                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
+          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
+
+show.SC_by.rank_top
+
+bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
+
+show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
+    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
+                    position=position_dodge(width=0.7), 
+                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
+          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
+
+#CALQ PERCENT TOTAL SC
+
+CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
+
+                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
+                         nbiv = length(Obj.ID),
+                       #rank.by.per
+
+                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
+                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
+                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
+                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
+                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
+                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
+                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
+)
+
+#MELT ABOVE
+MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
+              #leave out the Foci
+               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
+
+            id=c( "fileName","total.sc.cell","nbiv"))
+
+MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
+
+MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
+MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
+
+show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
+
+show_long.sc.proportions
+
+#original plot code 
+#ggplot(data = df1, aes(x, y, color = Group)) +
+#    geom_pointrange(aes(ymin = lb, ymax = ub), 
+#                    position=position_jitter(width=0.5), 
+#                    linetype='dotted') +
+#    theme_bw()
+```
+
+The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
+
+The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
+
+The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
+For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
+
+(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
+
+The other figure shows of each single bivalent contributes to the total SC area.
+Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
+on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
+So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
+
+- Is the difference between cell averages for males and females less that 10%?
+
+- also interesting, the pwd and msm don't have longer SC, compared to other strains.
+
+<!-- Automated BivData Comparison -->
+
+```{r XX.Automated.biv.stuff, echo=FALSE}
+
+#make automated female data to compare to
+female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
+
+female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
+
+#table(female.Bivdata_comp$strain)
+
+#female.Bivdata_comp
+
+#1. calq the rate of chrm segmentation in female categories in automateted data
+
+box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
+
+                   Nmice = length(unique(mouse)),
+                   mx.box.num = max(boxNumber), # need to run this at the cell level
+                    N.biv  = length(Obj.ID),
+                   pass.biv = N.biv / mx.box.num
+)
+box.rates <- add_mouse(box.rates)
+box.rates <- add_strain(box.rates)
+box.rates <- add_category(box.rates)
+
+cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+
+cell.pass.rates
+
+category.box.rates <- ddply(box.rates, c("category"), summarise,
+
+                   mean.pass.rate = mean(pass.biv)
+)
+#category.box.rates
+
+#2. pull the 3rd rank biv's calculate the SC means
+#calq means by rank (focus on the top 5)
+
+```
+
+For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
+
+The table above shows the number of bivalents from the same strains as in the manual whole cell data.
+The plot shows the bivalent passing rate across all of the individual cells from this female data set.
+For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
+
+(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
+(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
+
+strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
+
+** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
+But I am not using 'whole cell' summaries to compare female and males.
+
+**What is the effect of an extra XX-autosome on single bivalent means?**
+
+use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
+Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
+Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
+Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
+The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
+
+# Note on Heterochiasmy Definition
+
+<!-- add outline for new XX adjust section 
+adjust the colors for the plot below (standardize)
+-->
+
+I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
+In females the XX bivalent is indistinguishable from the autosomes.
+To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
+Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
+(I note if the and Y are paired, which they are at a high rate).
+The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
+To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
+
+1. Compile full-cell data from females (all 20 bivalents measured)
+2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
+3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
+
+According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
+
+(Put the XX adjustment section here)
 

### 2020-02-25 doc/Results.rmd
more careful editing and reading

@@
 
+ **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$
+
 We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework.
@@
 ```
-
-All of the effects within the mixed model M1.
+(not suprisingly) All of the effects within the mixed model M1.
 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
@@
 
-The linear models (M2. and M3.) allow more == specific comparisons
-
-Following up with linear models which tested for (subsecies M2), and strain effects alone (M3)
-both showed similar -- results -- **so we just report those from M3. also just report effects which were consisantly significant across both M2 and M3**
-
-<they also have sig strain effects .. PWD strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`>
-
-which identified several signifcant sex by strain interaction effects, ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
-and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`
-)
-
-<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
+allow us to test / quantifying the significance and focus on how specific levels and interactions create these differences.
+
+The linear models (M2.
+and M3.) had similar signifncat effects, so we just show the results from M3.
+In addtion to significant strain effects, ( p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both PWD and MSM had significant strain by sex effects ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
+and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`)
+
 These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.
-
-<M3 also has a few less significant effects>
-<not the reverse coefficient indexing -- this will give the name >
 
@@
 
+A linear model was run on the sex specific data set (M4), (just containing subspecies, strain and the interaction). X number of male and female mean MLH1 foci counts were used in each of these moels.
+
+For the male specific data, (the same PWD, MSM strain effects were signifciant -in addition to SKIVE (glm; p = `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`)(which is intrmediate)) .
+Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
+
+<three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
+(with effect sizes ranging from 5, 7, and 2 foci respectively).>
+
+While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
+
+<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
 <!--
 In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
-
 (*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
-
 Check the DF of the number cited above
@@
 `r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
-
 add number of cells
@@
  $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-Setting up the data set for using the mixed model (base data set, HQ data set)
 
@@
 - Molossinus is coded as a subsp level
-
-subspecies levels: 3, Musc, Dom and Molossinus
-strain levels: Dom: 3, Musc: 3  Mol: 2
-sex: 2 female male  
--->
-
-<!--
 also g strain effect/ dom^G^ strain effect 
@@
 -->
-<!-- glms for sex specific data also run -->
-
-A similar glm models (M4) was run on the sex specific data set, (just containing subspecies, strain and the interaction).
-
- **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$
-
-For the male specific data, three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
-(with effect sizes ranging from 5, 7, and 2 foci respectively).
-Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
-
-While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
 
@@
 
-In order to more fully discribe the MLH1 count per cell distributions we examined the variance applying the same mixed models and linear models, using the within mouse variance of MLH1 counts per cell as the dependant variable (Table 1).
-
-Within animal variance in CO count / genome wide RR is important because it could shape constraints (and determine how effective stabilizing selection can act (ref))
-
-Following the same basic framework we used for analyzing mean number of MLH1 foci per cell, we applied a set of mixed and linear models -- to test effects (sex differences for the within mouse variance in MLH1 foci per cell).
-
- **M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+MLH1 mean -- is quantitative trait -- variance of a trait can contain information pertinant to it's mode of evolution -- and mechanistically could reflect the strength of constraints acting on the trait.
+<in some traits, like gene expression patterns, evolution of the mean -- also has an effect on the variance>
+
+We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependant variable).
+<note about reduced power>
+
+<**M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
@@
 
- **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$
+ **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$>
 
@@
 
+**the general pattern is that females have more variance (alsmot twice as much (Figure 1))**
+
 For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
 Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
-background noise).
-For this reasons, we replicated the model analysis using a subset of cells with higher quality scores.
+background noise), we replicated the model analysis using a subset of cells with higher quality scores.
 These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).
-Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
-While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`) for the model with the full dataset.
-While in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
+The Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
+While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`).
+However, in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
+
+<there are some caveats, but the general pattern is that females have more variance ...
 For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell.
-This limits the power for detecting significance differences across groups, <but the main/general pattern, is that females have significantly higher within mouse variance compared to males.
-This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.
+This limits the power for detecting significance differences across groups,
+
+the main/general pattern, is that females have significantly higher within mouse variance compared to males.
+This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.>
 
 <!--
-motivation / background: maybe checkpoint / regulation
-limitations: observations at the mouse level, decreased power to estiamate the variance --
-
 $$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
@@
 
-< SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE >
+<SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE>
+
+<The greater amount of within mouse variance in females (almost and 
+(males have no detecable difference across strains for within mouse variance)
+
+-- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes. >
 
@@
 
-In an attempt to locatize the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
+In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
 DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days).
-A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_.
+A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_ (Figure2).
 One mouse was sampled for each strain.
@@
 
-Confirming the prediction that (more crossovers are associated with more initial DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval` , one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
-
-The differences in DMC1 foci were not significant for zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
+Confirming the prediction that evolution of the mean number of crossovers per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
+
+However the the differences in DMC1 foci were not significant for (later prophase) zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
+Indicating that this marker at early prophase is more precitive of the downstream crossover number differences.)
+
+The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+These results indicate that the targets of evolution which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
+
 <!-- which maybe be attributed to noise effects 
-
 The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->
-
-The ratios of average number of MLH1 and DMC1 foci per cell -- have  .... biological significance
-
-The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
-
-These results indicate that the changes -- which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
 
@@
 the number of DMC1 across stages (L and Z stages) are driven by different equations
-
 Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
-
 Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
@@
 -Baier et al 2014 supports this model (Chmr axis / DNA loop model)
--->
-
-<!--
-# Rapid male specific evolution
-
-list out the points which support rapid male specific evolution for MLH1 rates.
-
-number above bars are the raw counts for bivalent classes.
-
-Potential figure 2 (eventually). Try to make the figure taller.
--- add more data to make sample sizes more equal.
--- add space to indicate where female data is missing
-
-SKIVE male has lower proportion of 2CO compared to the hand measured, but it still looks like it's more than other strains.
-What's up with Lew females? there's lots of 2COs, also WSB female.
-
-This type of plot is another way to score heterochiasmy that gets around the sex chromosomes, because the data come from samples of bivalents and not averages of cells.
-If these data are random samples of bivalents, the X
-
-ToDo:
-- within strain across sex,
-- males, across high and low groups
-- look at mouse level proportions
 -->
@@
 
-The first attern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes.
+The first pattern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes.
 <these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>

### 2020-02-25 doc/Results.rmd
more slow maybe steady edits

@@
 
-This is a draft of the main figure. Outgroups are not included in the analysis below.
+This is a draft of the main figure. Outgroups are not included in the analysis below. (add horizontal lines for denoting subspecies, group means and expected minimum)
 
@@
 
-kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-
-kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+
+#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
 
@@
 
-# Genome wide recombiantion rate estimates
+# Genome wide recombiantion rate estimates for both sexes
 
@@
 
-main.table %>% filter(sex == "female")
-
-main.table %>% filter(sex == "male")
+#main.table %>% filter(sex == "female")
+#main.table %>% filter(sex == "male")
 
@@
 molossinus_.
-MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
+Mean MLH1 foci counts were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
 <!-- state the average number of mice
@@
 -sex-specific evolution, both sexes don't show equivalent elevation of gwRR
--Dom vs Molossinus and Musculus
-
--->
-
-Figure 1A shows the strain averages for each sex together - this is a way to display the general patterns of sexual dimorphism (or heterochiasmy) across strains. 
-Taking note of the direction and the magnitude our results confirm two general patterns that i) the genome wide recombination rates averages are greater females compared to males and ii) the degree of heterochiasmy (F:M rates) is generally low, rangeing from `r G.HETC.val` in _G_ to `r LEW.HETC.val` in _LEW_.
-
-The notable exceptions to this pattern are from strains with male averages that exceed the female averages, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` for PWD, MSM and SKIVE respectively.
-
-Examining the mouse means for the genome wide recombinaiton rate seperately for each sex reveals disctinct patterns of variation (Figure1 B and C).
-
-The female means can be discribed as more evenly distributed around the sex-wide mean of ~25
-
-While in males, the strain specific means -- can be seperated into groups of high genome wide recombination rates (near 30 COs per cell) and low genome wide recombiantion rates, near 23 COs per cell.
-
-<!-- transition from first part into female -->
-<!-- female section -->
+-Dom vs Molossinus and Musculus -->
+
+The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
+Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) that the genome wide recombination rates averages are greater females compared to males (female biaded heterochiasmy) and ii) the degree of heterochiasmy (F:M rates) is generally low in these house mouse strains, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
+Three notable exceptions of male biased heterochiasmy are from the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
+
+Examining the mouse means of MLH1 foci per cell seperately for each sex reveal disctinct patterns of variation (Figure1 B and C).
+The female means can be discribed as more evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B).
+While in males, the strain specific means seperate more clearly into two groups of high genome wide recombination rates near 30 MLH1 foci per cell and low genome wide recombiantion rates, near 23 MLH1 foci per cell (Figure1 C).
+
+<illustrate sex differences female mean `r mean((main.table %>% filter(sex=="female") )$mean_co )`>
+
+<This striking differences across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
+The pattern of strain averages distributed around the total sample mean in females is supportive of stabilizing selection or neutral evolution.
+In contrast to the pattern from male means which have rapid evolution in a subset of genetic backgrounds this fits a model of directional selection.
+(We attempt to make better evolutionary models by applying an evolutionary framework)>
 
 ```{r female.strain.table, echo=FALSE}
-
 female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
@@
 ```
-
 <!-- male section -->
@@
 ```
-
-Examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.
-
-<!-- female strain means are have more variation in a random way. The range of genome wide rec strains is X to X, -- no clustering of strain means
-
--->
-
-In females the strains with the greatest difference
-in mean gwnome wide recombiantion rate was `r G.female.diff` ( _domesticus^G^_) more compared to other strains. 
-<!-- _domesticus^LEW^_  (`r round(LEW.female.diff,4)`),
-and _musculus^MSM^_  (`r MSM.female.diff` ),-->
-
-<!-- male variation is a greater magnitude, but divides into 2 groups, magjority of strain means are around 23, while PWD and MSM are near 30, skive is intermediate-->
-
-While in the male data (most strain averages are close to 23 (Figure1).
-
-the genome wide recombination rate for had a much larger range, up to `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )` ( _molossinus^MSM^_ ).
-
-_musculus^PWD^_ ,and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,  and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means.
-
-<notabley these strains come from Musc and Mol -- Mol natural hybrid of musculus and castaneus. this comments out text?>
-
-The _musculus_ and _molossinus_ subspecies have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.
-
-<!-- dom, low variation and close to the expected in
-within the male means, there is a low degree of strain varaince in _Domesticus_ subspecies with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`
--->
-
-<!-- ranges of mouse means
-In the _domesticus_ subspecies the ranges of mouse means are 
-`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )`
-in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.
-
-In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_
-
-The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.
--->
-
-<!--
-$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$ -->
-
-<!--
-The strain means for females were pretty similar.
-For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _domesticus^WSB^_, _domesticus^G^_, and _domesticus^LEW^_ respectively.
-In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, and _musculus^KAZ^_ respectively.
-In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and _molossinus^MOLF^_ respectively.
--->
-
-While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively.
-In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively.
-In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.
 
@@
 
- **M4.** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subspecies * strain + \varepsilon$$
-
-We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework.
+**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
 
@@
 ```
-(not suprisingly) All of the effects within the mixed model M1.
+
+We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework.
+All of the effects within the mixed model M1.
 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
-Confirming the qualitative observations of wide spread variance.
-<!-- indicating XXX -->
+
+<These results allow use to reject the hypothesis of genome wide recombination rates (varying consistantly by sex across subepcies -- and having variance within subspecies due to strains)>
+
+ <!-- (random strain effect was meant to reflect sampling of gentic backgrounds/standing genetic variation for a subspecies) -- (significant pvalue indicates that the variance is not 0)>
+ indicating, fixed effects -- random effect variance due to each strain nested within subspecies -->
 
 ```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
 #strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
 
@@
 
-MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
 
@@
 
-allow us to test / quantifying the significance and focus on how specific levels and interactions create these differences.
-
-The linear models (M2.
-and M3.) had similar signifncat effects, so we just show the results from M3.
-In addtion to significant strain effects, ( p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and MSM strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both PWD and MSM had significant strain by sex effects ( _musculus^PWD^_*male and  _musculus^MSM^_*male ) (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
-and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`)
-
-These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell.
-
-Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
+<The linear models, allow us to quantifying the significance in sepecific fixed effects and focus on how specific levels and interactions create these differences.>
+
+The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombiantion rates.
+The subspecies fixed effects tested in M2, had weak to no significance (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
+m.
+musculus_ and _M.
+m.
+molossinus_ respectively), indicating that most of the variance is due to strain effects.
+As such, we focus on results from M3.
+The two most significant strain effects were _domesticus^G^_ and _musculus^MSM^_ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]` for _domesticus^G^_ and _molossinus^MSM^_ respectively).
+
+In addition to significant strain effects, ( p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and _molossinus^MSM^_ strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
+and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.
+
+Combined these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of sexual dimorphism /heterochiasmy in _domesticus^G^_.
+No significant sex effect showing that the sex effect on mean CO number is not uniform across the sample but has synergy with speciefic strains.
+
+<domesticus^G^ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
+and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`>
+
+A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` number of male and `r length(female_mouse.avs_for.analysis$mouse)` female mean MLH1 foci counts were used in each of these models.
+For the male specific data,
+
+_musculus^PWD^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
+(with effect sizes ranging from 5, 7, and 2 foci respectively).
+Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
+
+three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, _musculus^PWD^_  (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
+(with effect sizes ranging from 5, 7, and 2 foci respectively).
+
+While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
+
+<!--
+M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
+<Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
 These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
 
-_domesticus^G^_ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
-and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`
-
-A linear model was run on the sex specific data set (M4), (just containing subspecies, strain and the interaction). X number of male and female mean MLH1 foci counts were used in each of these moels.
-
-For the male specific data, (the same PWD, MSM strain effects were signifciant -in addition to SKIVE (glm; p = `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`)(which is intrmediate)) .
-Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
-
-<three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, PWD (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), SKIVE (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and MSM (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
-(with effect sizes ranging from 5, 7, and 2 foci respectively).>
-
-While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
-
-<M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
-<!--
 In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
@@
 
-MLH1 mean -- is quantitative trait -- variance of a trait can contain information pertinant to it's mode of evolution -- and mechanistically could reflect the strength of constraints acting on the trait.
-<in some traits, like gene expression patterns, evolution of the mean -- also has an effect on the variance>
-
+Variance of a trait can contain information pertinant to it's mode of evolution and mechanistically could reflect the strength of constraints acting on the trait.
 We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependant variable).
 <note about reduced power>
-
-<**M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+<in some traits, like gene expression patterns, evolution of the mean -- also has an effect on the variance>
+
+<!-- 
+**M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
@@
  **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$>
+-->
 
@@
 
-For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` mixed and glm models respectively).
+For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`).
 Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
@@
 These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).
-The Sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
+
+The sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
 While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`).
-However, in the model using the high quality cells, PWD has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
+However, in the model using the high quality cells, _musculus^PWD^_ has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
+
+**the general pattern is that females have more variance (alsmot twice as much (Figure 1))**
 
@@
 
-<The greater amount of within mouse variance in females (almost and 
-(males have no detecable difference across strains for within mouse variance)
-
--- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes. >
-
-# Variation in DSB Number Per Cell
+(combined the summary of the MLH1 distributions evolution in the mean and consistant pattern of sex differences)
+
+<The greater amount of within mouse variance in females (almost and, (males have no detecable difference across strains for within mouse variance)
+-- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes.>
+
+3. -- the striking differences / patterns of variation across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories for male and female rates. 
+For female the pattern of Strain averages distributed around a species wide average - fits a model of stabilizing / relaxed / neutral evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic backgrounds / strains which fits a model of directional selection on genome wide recombination rate.
+
+# Evolution of Genome wide recombination rate associated with evolution of mean precursor number
+
+<transition, >
 
@@
 
-For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = ), which reflects the expected patterns across the timing of meiosis.
-
+For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the timing of meiosis.
 Confirming the prediction that evolution of the mean number of crossovers per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
@@
 However the the differences in DMC1 foci were not significant for (later prophase) zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
-Indicating that this marker at early prophase is more precitive of the downstream crossover number differences.)
+Indicating that this marker at early prophase is more precitive of the downstream crossover number differences.
+
+<explain the biological significance of the ratios!>
 
@@
 
-# Chromosome Class Proportions
+# Evolution of genome wide recombination rate is reflected at the single chromosome level
+
+<come up with a better section title>
 
@@
 chrm.class.barplot
+
 ```
@@
 #sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
+
 ```
@@
 
-In order to decompose the cell wide rate, we decided to look at the proportion of chromosome classes (ie by the number crossovers on the same bivalent).
-A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).
-
-The first pattern we notice from looking at theis large number of samples is that The vast majority of all bivalents in house mouse have one or two crossoverrs (1CO an 2CO), even across sexes.
+We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per chromosomes (bivalent).
+
+Because our observations of MLH1 foci are made in pachtyene cells, our chromosome opbservations are fully synapsesed paired homologous chromosomes (i.e.
+4n copies of each chromsomes) -- additionally we can distinguish the idenetiy of indiviual chromsomes (Chromsome 1 or Chromsome 2)
+for these reasons we use the term bivalent.
+
+Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers.
 <these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>
 
-<note about the rates of 0CO, 3CO and 4CO? these are marginal chromsomes classes / we'll discuss later?)>
-
-The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`).
-This relpicates what has been notice in the literature (cite stapley).
+The proportion of 1CO:2CO (are displayed in figure X) --
+
+-- distinguishes -- the high and low rec males
 
@@
 
-<Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
+The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`).
+<This relpicates what has been notice in the literature (cite stapley).>
+
+Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
 They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
@@
 
-From our total set of cell images `r length(new.Curate.FULL.org$Obj.ID)` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(Curated_BivData$Obj.ID)` single bivalent measures were left.
+From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
+After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent measures were left.
 (make sure these are the hand.foci bivalents).
-
-<how do these agree with other patterns of MLH1>
 
@@
 
-data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-Table X illustrates the break down of single bivalent observations by category. 
-Error compared to human measures and other details have been discribed in (Peterson_2019).
-
-The ranges of of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in MSM male to 
-`r max(algorithm.stat.category$mean_pass.rate)` ) KAZ female, we assume that the isolation process is not biased.
-Because there are hundreds of observations per category, we assume that each of the 19 autosomes (chromosomes) is equally represented in the dataset of single bivalents.
+#showing the table..
+#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
+```
+
+<Table X illustrates the break down of single bivalent observations by category. Error compared to human measures and other details have been discribed in (Peterson_2019).>
+
+The ranges of of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to 
+`r max(algorithm.stat.category$mean_pass.rate)` ) _musculus^KAZ^_ female, we assume that the isolation process is not biased.
+Because there are hundreds of observations per category, we assume that each of the 19 or 20 chromosomes are equally represented in male and female single bivalent data datasets.
 We assume the bivalent datasets across all categories are equally eqivilent for representing general patterns.
@@
 
-We approach two main questions using this single bivalent data set, 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
+Using this data set, We address two main questions 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
 
@@
 
-Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of PWD and MSM.
-A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
+Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of _musculus^PWD^_  and _molossinus^MSM^_  A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
 

### 2020-02-25 junk_text.R

@@ -1,6 +1,45 @@
 #
 # junk text -- dumping the junk text
 
+<!--While in two of the subspecies, the means for males show a greater range across means, notable in the molossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _molossinus^MSM^_ and  _molossinus^MOLF^_ respectively. In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _musculus^PWD^_, _musculus^SKIVE^_, _musculus^KAZ^_, _musculus^TOM^_, _musculus^AST^_, and _musculus^CZECH^_ respectively. In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _domesticus^WSB^_, _domesticus^G^_, _domesticus^LEW^_ and _domesticus^PERC^_ respectively.-->
+
+
+from the intro / first part
+
+<!-- _domesticus^LEW^_  (`r round(LEW.female.diff,4)`),
+and _musculus^MSM^_  (`r MSM.female.diff` ),-->
+  
+  <!-- male variation is a greater magnitude, but divides into 2 groups, magjority of strain means are around 23, while PWD and MSM are near 30, skive is intermediate
+
+Examinations of the sex specific data revealed that greater variety in across the strain means in males compared to females.  (exceptions to the general patterns of heterochiasmy)
+
+In females the strains with the greatest difference
+in mean gwnome wide recombiantion rate was `r G.female.diff` ( _domesticus^G^_) more compared to other strains. The genome wide recombination rate for had a much larger range, up to `r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )` ( _molossinus^MSM^_ ). _musculus^PWD^_ ,and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,  and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means. 
+
[…]

### 2020-02-25 doc/Results.rmd
end O day

@@
 
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
+chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
+chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
 
@@
 
+#MELT_Chrm.Proportion$Rec.group <- ifelse(grepl("PWD male", MELT_Chrm.Proportion$category), 1, 
+#                              ifelse(grepl("MSM male", MELT_Chrm.Proportion$category), 1,
+#                                ifelse(grepl("SKIVE male", MELT_Chrm.Proportion$category), 1, 0)))
+
+#Chrm.prop.male.low <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 0)
+#Chrm.prop.male.HIGH <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 1)
+
+#strain av tables
+#chrm.prop.male.high1
+#chrm.prop.male.low
 ```
@@
 We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per chromosomes (bivalent).
-
 Because our observations of MLH1 foci are made in pachtyene cells, our chromosome opbservations are fully synapsesed paired homologous chromosomes (i.e.
-4n copies of each chromsomes) -- additionally we can distinguish the idenetiy of indiviual chromsomes (Chromsome 1 or Chromsome 2)
+4n copies of each chromsomes).
+Additionally we can distinguish the idenetiy of indiviual chromsomes (Chromsome 1 or Chromsome 2)
 for these reasons we use the term bivalent.
 
-Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers.
-<these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>
-
-The proportion of 1CO:2CO (are displayed in figure X) --
-
--- distinguishes -- the high and low rec males
-
-Most of the bivalents are 1CO and 2CO class bivalents -- and the proportion of these two classes distinguish high and low rec male strains. (distinguish the high rec males from minimal rec groups)
-
-(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class.
-The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.
-*(much smaller proportions and variation in the fringe classes*
-
-**2. most of the male proportions are 2/3's 1CO and 1/3 2COs. but The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**
-
-**MSM 60%  PWD 50%  SKIVE 30%   20 - 10% remaining (low) strains**
-
-PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).
-
-While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
- and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-
-**3.
-these results confirm the intuitive conclusion/result that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents**
-
-The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`).
-<This relpicates what has been notice in the literature (cite stapley).>
-
-Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
+Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers (Figure X).
+<these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>.
+The proportion of 1CO:2CO distinguishes the high and low rec males (Figure X).
+Confirming the intuitive interpertation that high recombination strains have more 2CO bivalents at the expense of 1CO bivalents.
+Across the female observations the propotions of chromsome classes are not notably different.
+Most of the low rec strains range from `r min(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` _musculus^KAZ^_ to `r max(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` ( _molossinus^MOLF^_ ).
+In the high rec strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
+
+While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+
+These results confirm the intuitive conclusion that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents.
+
+<(Also comparing the low group together ... Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
 They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
@@
 
-Breif conclusion: the evolution of genome wide recombiantion rates proceeds through the chromosome level.
-
-<something about the female results>
-
 <SUMMARY the / TRANSITION to single bivalen measures>
@@
 # Single Bivalent Level Results
+
+-- Confirmation that it's the increase in 2CO bivalents -- that comprises a higher cell wide number of COs (total number)
+
+Let's look at the Rec landscape --
 

### 2020-02-26 doc/Results.rmd
small changes and notes

@@
 
+after DSB formation, DSBs are repaired as either NCO or CO (CO designation step)
+The vast majority of DSBs are repaired as NCOs, with a minority (in most species) of approximately 10% are repaired as crossovers.
+
+-DSBs are either repaired as crossovers or noncrossovers (NCO), with conversion tract lengths (30 to 100bp long). (this is also known as the CO repair deicison )
+
+thus the ratio of CO:DSB is in part an indicator of the proportion of DSBs which are designated into COs -- (and the others are repaired as NCO).
+
 The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
 These results indicate that the targets of evolution which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
-
-<!-- which maybe be attributed to noise effects 
-The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->
 
@@
 
-<come up with a better section title>
-
 ```{r chrm_prop_data, echo=FALSE, warning=FALSE}
@@
 
-**M3. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
+**M1. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$
+
+**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
+```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
 lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
@@
 These effects were followed up the glms, again applied to both mouse means for short biv SC lengths and total sc area per cell.
-
-**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
-
-**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
 
@@
 
-(rethink where this section should go)
-new outline
-
 1. illustrate problem(affects mostly SC length)  
-
 2. Expected impact on sex comparisons, estimated effect size of the X  
 3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
-
-4. 19 female, 19male, 20female 20 male
-
-The female mouse averages should have adjustments for the XX.
-working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains.
-Subtract this amount from the female mouse averages ...
-This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)
-
+4. permutations of 19 female, 19male, 20female 20 male
+
+<
 - Of all female single bivalents observations, 5% are XX (1 of 20). 
-
 - The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
-
 - The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
-
-I think a formula something like this can be applied to adjust for XX
-
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*
+- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX>
 
@@
 <transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents
+
 1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs

### 2020-02-26 doc/Results.rmd
blah

@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
+  pdf_document: default
   html_document:
@@
     toc_float: yes
-  pdf_document: default
-  word_document:
-    toc: yes
-    toc_depth: '5'
 bibliography: refs_Main_Report.bib

### 2020-02-26 doc/Results.rmd
working from bottom up

--- 2020-02-26 doc/Results.rmd
@@
 
-DO THE HIGH QUALITY dataset tests
-
 <!-- breif background on longer female SC  (mouse, human) Gruhn-->
@@
 
+The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
+(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
+<But we use this metric anyways>
+
 An additional measure of the SC-AE we tested is the total SC area per cell (the whole genome).**to get a comparison of the SC-AE area for the whole genome (not just a subset of chromosomes)** We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
@@
 #run t-test for male female avergage for each strain
+
+#ALL
+Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)
+
+#dom
+Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
 
@@
 
+#Musc
+Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
+                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
+
 #PWD
@@
 #SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")
-
-#error not enought female observations
+#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
+#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
+#cell level total SC
+#MLH1.merge.Skel   $skel_size
+#error not enought female observations, ONLY 1 female SKIVE mouse
+
+Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
+           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
 
@@
 
-The most averages for total SC-AE area (skeletonizes) are significantly longer in females for all strains (p=
-`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
-for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).
-
-Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)
-
-The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
-(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
-<But we use this metric anyways>
+In all of the strains, the female mouse mean total SC per cell was significantly greater in females than males (ttest; WSB p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, G p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, LEW p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`,  PWD p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, KAZ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, MSM `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`
+
+for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively), for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)
 
@@
 
-The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>
--- we follow up with models (which focus on sex differences).
+<!-- Dom males are shorter than this set of high rec mice
+The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>-- we follow up with models (which focus on sex differences).
+-->
+
+<**t.test results for total.SC**
+all values (F vs M) (p= `r Q1_tot.sc_t.test.all$p.value`)
+Dom (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.DOM$p.value`)
+Musc (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.MUSC$p.value`)  (note, missing SKIVE, not enough female measures)  
+MSM (F vs M) -- (ttest; p= `r Q1_tot.sc_t.test.MSM$p.value`)>
 
 <explain why applying models.  We then use models to examine/ think about the evolution>
-
-The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mean mouse total sc and mouse mean short bivalent.
+The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mouse mean short bivalent length and mean mouse total SC.
 
@@
 **M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$
-
-**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
-
-**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
 
@@
 
-For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed model of mouse.av total sc.
-
-The three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
-sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
-
 ```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
@@
 
-Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
-The interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
-
-These effects were followed up the glms, again applied to both mouse means for short biv SC lengths and total sc area per cell.
+Mixed models for the sex differences in mouse in the mixed models using mean **short bivalents** length as the dependant variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
+
+Subsp and sex had the most significant p values, --- For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed modelusing mean **total sc** per cell as the dependant variable, the three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
+
+(these results from the mixed models -- are similar to the mean MLH1 per cell results -- so we follow up -- with tests of significant strain effects).
 
@@
 
-The following effects had significant p values for the **total SC** from the above model
-
-subspecies musc p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`, subspecies mol p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`, sex male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`, strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
-<this is the smallest p vale, elevated female total, driven by the high means in females>
-
+In linear models (M2 and M3) for **mean short bivalent length**, male is most significant (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`).
+In the linear model M2 the skive strain and strain by sex effects are slightly significant (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively), however these could be due to low sampling of skive females.
+
+The following effects had significant p values for the **total SC** from the above model (for the linear models M2 and M3), 
+subspecies _musculus_ (glm M2; p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`), subspecies _molossinus_ (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`), sex male (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`), strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
+
+For M3 -- many strain effects are significant, but the largest are sex (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`)
+but also some strain effects, G strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][2]`), LEW strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][3]`), PWD strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][4]`).
+
+Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel.
+when chromatin condensation is summarized / quantified at this level.
+(can't exclude the possibility that the XY morphology is different between these strains)
+
+<!--
 (subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
-
-<!--
 for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
 -->
-
-Fr the **short bivalent** sc the model results are listed below. From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).
-
-while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.
-
-In the M3 model (strains and sex as fixed effects), the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).
 
@@
 2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>
+
+```{r Q1.SC.Q12, echo=FALSE}
+
+#make comparisons for the high quality cells
+
+#get Q12 cells from MLH1.merge.Skel
+#calq the mouse averages
+
+MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
+
+totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
+
+#unorder the factors
+totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
+totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
+totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
+totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
+
+#REMOVE JUVI mice
+tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
+totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
+                 .(strain, sex),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
+
+#unorder the factors
+totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
+totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
+totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
+totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
+
+```
+
+```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
+
+Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Dom
+#Tot.SC_mouse.av_Dom
+
+Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Musc
+#Tot.SC_mouse.av_Musc
+
+Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+
+Q12.Tot.SC_mouse.av_Mol
+Tot.SC_mouse.av_Mol
+
+```
+
+<Q12 total SC mouse means!!
+In Dom when running the Q12 SC lengths, the sex difference seems to get larger (male mean total SC gets lower).
+In Musc the female mean total sc gets lower (maybe fewer cells (SKIVE is removed)).
+For Mol / MSM female mean stays about the same, MSM and MOLF means converge around 1600.>
 
@@
 
-# Q2 Normalized Single CO positions
-
-<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->
+# Q2.1CO rec landscape evolution is decoupled from gwRR evolution
+
+<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents
+
+Main results:  
+1. WSB has the most terminal 1CO rec landscape and MOLF has the most central.
+2. Evidence for evolution of the 1CO rec landscape, but it is decoupled from the evolution in gwRR we observe in figure 1
+2.5  The normalized 1CO position is not a good metric for distinguishing strains with rapid gwRR evolution
+
+-->
 <!-- set up chunks here -->
@@
 ```
-
-<not sure what the beenfit is to test for chrm effectss>
 
@@
 ```
-
-The previous section found that the single crossover landscape is significantly different between males and females.
-In this section we focus on differences between the high and the low rec male strains
-In order to test if if there are distinguishing meterics for these two groups.
 
@@
 
+The previous section found that the single crossover landscape is significantly different between males and females.
+In this section we focus on differences between the high and the low rec male strains
+In order to test if if there are distinguishing meterics for these two groups.
 **The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**
@@
 T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
-
-<further tests -- within the subspecies, musculus and molossinus?
-
 The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
-
 While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`).
 This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.
-
- <BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF, 
- (MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.>
 
@@
 
-in addition models which use mouse average of foci position as a dependat variable, we decided to run models for the single bivalent level, and which include chromosome length as a fixed effect.
-
-in the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
+<!--
+In the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
 ..
@@
 
-In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects
-
+In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects -->
+
+<!-- Just hold these for reference
 **For the M2 (subsp)**
@@
 LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).  
-
 But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).  
-
 and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)  
-
 **For M3 (all strains)**
@@
 strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
-
-from the main figure above, you can see that G and MOLF 1CO foci positions are the most different
-
-<no prediction for the strains within subspecies>
-
-For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice
-
-there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
-(motivation for MOLF female observations and more male MSM biv measures)
+-->
 
@@
 
-<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.
-
-**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$>
+given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models.
+Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
+
+<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>
+
+<!--
+**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$ -->
 
@@
 
-<for the first model,
-
+<!-- hide these linear model results for now
+for the first model, 
 chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`
@@
 
--chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`>
-
-males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
-
-There are significant difference across the 1CO positions, yet they don't match fit any of the predictions. there are also complications due to the chromsome class proportions that
+-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]` -->
 
@@
 
-<glms>
-
-The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
+<glms> The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
 
@@
 
-<when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF>
+<!--when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF-->
 
@@
 
-We determined that the main difference in IFD disbribution across these groups is an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC length in low rec strains.
-The rate of these close foci range from 8.2% (G) to 16% (KAZ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for skive, MSM and PWD respectively).
-<no chi sqaure test see above>
-
 ```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
@@
 
+We determined that the main difference in IFD disbribution across these groups is an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC length in low rec strains.
+The rate of these close foci range from 8.2% (G) to 16% (KAZ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for skive, MSM and PWD respectively).
+<no chi sqaure test see above>
+
  **(unlike total SC length and 1CO pattern, IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.
@@
 (the extra stats are not correctly mapped)>
+
+<!-- final summary, BE brief-->
+
+BE brief,
+
+final summary, highlight the themes of 1. more variance in female traits (MLH1 foci per cell, rec landscape / interference)
+
+while the male traits -- more focused (less variance in MLH1 counts per cell, shorter chromatin comparction?  bias position, stronger interference)
+
+(should I measure the var of all traits in Q1)
+
+2. rapid evolution of genome wide rec rates can fit into this -
+
+- evolution of more COs requires more DSBs (evidence points to a earlier step which is likely the chromatin compaction (amount of axis))
+
+- the strongest single bivalent predictor/correlate of rapid gwRR for males is interferece strength in 2COs
+
+AND somewhat puzzeling result of (having stronger interference), meaning there is a greater proportion of a bivalent that seperates two foci  (2 foci push to the ends of chrms (maybe decreased centromere supression))
+
+# Blank1
+
+<!-- unused chunks -->
+
+```{r show.long.SC, echo=FALSE, eval=FALSE}
+
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
+
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
+
+```
+
+<ABOVE plots show the SC lengths for the 'long SC data set'.
+They are supposed to be the longest 4-5 SC from cells where I could get good measures.
+These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
+Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
 
@@
 
-```{r show.long.SC, echo=FALSE, eval=FALSE}
-
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
-
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
-
-```
-
-<ABOVE plots show the SC lengths for the 'long SC data set'.
-They are supposed to be the longest 4-5 SC from cells where I could get good measures.
-These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
-Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
-
-<!-- unused chunkcs -->
-
 ```{r SC.distb_manual_Auto, echo=FALSE, eval=FALSE}
@@
 
-<!-- Whole Cell Manual Comparison -->
+<!-- Whole Cell Manual Comparison
 
@@
 
-# Note on Heterochiasmy Definition
+# Blank2
+<!-- Note on Heterochiasmy Definition
 
@@
 
+  -->
+
 # References

### 2020-02-26 doc/Results.rmd
cleaning up english

@@
 
-Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers (Figure X).
-<these results fit with previous observations on the number of crossovers per chromosome (Stapley 2019)>.
-The proportion of 1CO:2CO distinguishes the high and low rec males (Figure X).
-Confirming the intuitive interpertation that high recombination strains have more 2CO bivalents at the expense of 1CO bivalents.
-Across the female observations the propotions of chromsome classes are not notably different.
-Most of the low rec strains range from `r min(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` _musculus^KAZ^_ to `r max(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` ( _molossinus^MOLF^_ ).
+Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers (Figure X) with proportion of 1CO:2CO distinguishes the high and low rec males (Figure X).
+THis confirms the intuitive interpertation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents.
+Across the female observations the proportions of chromosome classes are not notably different.
+
+<Most of the low rec strains range from `r min(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` _musculus^KAZ^_ to `r max(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` ( _molossinus^MOLF^_ ).>
+
 In the high rec strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
 
-While the proportions are significantly different, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-
-These results confirm the intuitive conclusion that the rapid male specific evolution in gwRR proceeds though the chromosome level/though increasing the number of 2CO bivalents at the expense of 1CO bivalents.
-
-<(Also comparing the low group together ... Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
-They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`).
-This significant p value indicates that at least one of the six strains had significant differences).>
+While the proportions are significantly different
+(p = `r`), PWD male and MSM male proportions are more similar to each other than strains from the same subspcies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
 
@@
 
--- Confirmation that it's the increase in 2CO bivalents -- that comprises a higher cell wide number of COs (total number)
-
-Let's look at the Rec landscape --
-
-In order to better deconstruct the genome wide recombiation rate, we decided to deconstruct cell wide average by examining features of the recombiantion landscape at the single bivalent level.
-recombination landscape refers to the i) total length of the SC-AE (potential area for COs), ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.
+In order to deconstruct the cell wide MLH1 counts
+deconstruct we 
+look at the recombination landscape at the single bivalent level.
 
@@
 After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent measures were left.
-(make sure these are the hand.foci bivalents).
+
+We use these observations to look at the recombination landscapes, here refereing to the i) total length of the SC ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.
+<the position is general, not genomic corrdinates>
 
@@
 
-The ranges of of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to 
+A potential downside of the image analysis algorithm we used is that not all bivalents per cell are isolated.
+The ranges of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to
 `r max(algorithm.stat.category$mean_pass.rate)` ) _musculus^KAZ^_ female, we assume that the isolation process is not biased.
-Because there are hundreds of observations per category, we assume that each of the 19 or 20 chromosomes are equally represented in male and female single bivalent data datasets.
-We assume the bivalent datasets across all categories are equally eqivilent for representing general patterns.
-
-<curation, x from curation, XY excluded, replication to hand measures>
-
-Using this data set, We address two main questions 1) Which traits are sexually dimorphic? and 2) which traits fit distinsutinguish the high and low recombining strains for males?
-
-<These questions use different datasets: Q1 uses a dataset with only sex matched strains  and Q2 uses the data set with all male observations with including strains including those not in the Q1 set.>
-<The same basic models from the MLH1 counts per cell will be used. In addition to basic t.tests and logistic regression models for Q2 to distinguish betwee high and low recombining strains.>
+However, because there are hundreds of observations per category, we assume the set of single bivalents are equally representative of general patterns.
+
+Using this data set, we address two main questions 1) Which traits are sexually dimorphic? and 2) Which traits fit distinguish the high and low recombining strains for males?
 
 # Q1 Sex Differences in SC-AE Lengths
-
 <!-- breif background on longer female SC  (mouse, human) Gruhn-->
 
-Our data set provides an opportunity to test if the canonical pattern of females having higher gwRR and more SC area) have been uncoupled in the instance of rapid male specific evolution in the gwRR of _musculus^PWD^_  and _molossinus^MSM^_  A simple model might predict if SC area is the strongest predictor for gwRR / CO number, the high recombining males would have greater SC area proportional to the difference in CO number per cell.
-
-<Comparing the means of single bivalents is complicated by the fact we can't identify the chromosome identitied the physical size (in terms of SC-AE) is the closet approximation we can make.>
+(in many mammal species -- SC lengths are longer in females - yet most all of these observations come from female biased heterochiasmy species) -- Our data set provides an opportunity to test (how SC length and chromatin compaction relates to heterochiasmy)
+
+if the canonical female biased heterochiasmy pattern and long SC lengths are uncoupled in instances / our strain with male-biased heterochiasmy ( _musculus^PWD^_  and _molossinus^MSM^_).
+A simple model might predict if SC length is the strongest predictor for number of MLH1 foci er cell the high recombining males would have longer SC length proportional to differences in the number of MLH1 foci per cell.
 
@@
 
-We use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data.
+To account of the obvious confounding effect of sex chromosome differences in females and males, we use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data.
 A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes.
-
-<Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes>
+Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes.
 
@@
 
-<in the above boxplots of mouse means for shortbivalents>
-
-For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer, the largest difference across sexes by strain are>  (t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)`
-for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer the largest difference across sexes by strain are>  
+**`r` the magnitude differences between the mouse mean short bivalent SC range from X to X**
+
+(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
 This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
-
-<EXPLAIN WHY USING TOTAL SC AREA>
-
-The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
-(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
-<But we use this metric anyways>
 
@@
 Mouse means were calculated for their total SC-AE area and are displayed in figure X with the strain means shown as horizontal lines (below).
+
+The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
+(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
+<But we use this metric anyways>
 
@@
 
-# Q1. 1CO Position Sex Differences
+#Q1. 1CO Position Sex Differences
 

### 2020-02-26 doc/Results.rmd
more spell check

@@
 
-An additional measure of the SC-AE we tested is the total SC area per cell (the whole genome).**to get a comparison of the SC-AE area for the whole genome (not just a subset of chromosomes)** We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
+An additional measure of a smaller subset of single bivalents (which excludes the sex chromsoomes), we tested/compared a measure of the SC lengths for the whole genome (whole cells) using a image nalysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang)
+
+We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
 A total of X images were run trhought the image analysis pipeline, and after outliers removed, X remained.
-/ cell based measures were left.
-Mouse means were calculated for their total SC-AE area and are displayed in figure X with the strain means shown as horizontal lines (below).
+Mouse means were calculated for their total SC area per cell and are displayed in figure X.
 
 The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
-(simply put, the oocytes have an additional autosome, and the spermatocytes have mroe mostly non-recombinationly active SC-AE area, the XY).
-<But we use this metric anyways>
+Breifly/in effect the oocytes have an additional autosome, and the spermatocytes have a mostly recombinationly inactive unit of SC (the XY).
 
@@
 
-<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )
+<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )>
 
@@
 
-In all of the strains, the female mouse mean total SC per cell was significantly greater in females than males (ttest; WSB p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, G p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, LEW p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`,  PWD p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, KAZ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, MSM `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`
-
-for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively), for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)
+In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; _domesticues^WSB^ _ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^ _  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^ _ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^ _ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, KAZ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^ _ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`
 
@@
 
-<explain why applying models.  We then use models to examine/ think about the evolution>
-The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mouse mean short bivalent length and mean mouse total SC.
+<explain why I'm applying models>
+
+We follow up the general result of females have longer SC lengths -- by applying the same model frameworks which we used to analyze the mean MLH1 foci per cell -- to understsand the evolutionary framework.
 
@@
 
-Mixed models for the sex differences in mouse in the mixed models using mean **short bivalents** length as the dependant variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+In the mixed models for the sex differences mean **short bivalents** length as the dependant variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
 The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
 
-Subsp and sex had the most significant p values, --- For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed modelusing mean **total sc** per cell as the dependant variable, the three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
-
-(these results from the mixed models -- are similar to the mean MLH1 per cell results -- so we follow up -- with tests of significant strain effects).
+For the mixed model run on the mean total SC metric, there are more significant effects for strain and subsp for the mixed model using mean **total sc** per cell as the dependant variable, the three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
 
@@
 
-The following effects had significant p values for the **total SC** from the above model (for the linear models M2 and M3), 
-subspecies _musculus_ (glm M2; p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`), subspecies _molossinus_ (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`), sex male (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`), strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
-
-For M3 -- many strain effects are significant, but the largest are sex (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`)
+The following effects had significant p values for the **total SC** from the above model (for the linear models M2 and M3), subspecies _musculus_ (glm M2; p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`), subspecies _molossinus_ (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`), sex male (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`), strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
+
+For M3 in the mean total SC lengths there were many strain effects are significant, but the largest are sex (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`)
 but also some strain effects, G strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][2]`), LEW strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][3]`), PWD strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][4]`).
-
-Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel.
-when chromatin condensation is summarized / quantified at this level.
-(can't exclude the possibility that the XY morphology is different between these strains)
+(this means...).
+Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized / quantified at this level.
 
@@
 
-<In summary, females (oocytes) have loser different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
-
-<1. the longer female sc's is consistant and seen across all strains,
-even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
+<In summary, females (oocytes) have different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
+
+<1.
+the longer female sc's is consistant and seen across all strains, even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
 2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>
@@
 
-<Q12 total SC mouse means!!
+<Q12 total SC mouse means!
 In Dom when running the Q12 SC lengths, the sex difference seems to get larger (male mean total SC gets lower).
@@
 
+Summarize for now, refine this section later.
+
 1. illustrate problem(affects mostly SC length)  
@@
 
-<
-- Of all female single bivalents observations, 5% are XX (1 of 20). 
+<Of all female single bivalents observations, 5% are XX (1 of 20). 
 - The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
@@
 
-<transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents
+<transition>  
+-following up on SC area - with how COs are place - placement and number of COs on single bivalents
 

### 2020-02-26 doc/Results.rmd
cleaning up Q1.1CO and Q1.IFD

@@
 
-We focus on the foci positions from one crossover (1CO n= 
-) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.
-
-We anchor the position measures at the centromere (and normalize by deviding the position by the length of the chromosome, so the value range from 0 to 1, with smaller values representing positions closer to the centromere end of the bivalent, and larger values meaning foci positions closer to the opposite end (refered here as terminal or telomeric position).
-
-We plot the mouse averages across all of the single bivalent measures, with the strain averages shown as boxplots...
-
-in order to get a better picture of the sex differences in this foci-position metric
+We focus on the foci positions from one crossover (1CO n= `r `) bivalents since the landscape patterns for multi crossover bivalents (2COs) will be highly influenced by crossover interference.
+
+The CO positions are normalized by the SC length and anchored at the centromere with the normalized position values ranging from 0 to 1, (refered here as terminal or telomeric position).
 
@@
 
-In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (fitting the predictions/pattern of the 'typical recombination landscape' (Sardell)).
-The single CO positions are medially placed in females and terminall placed in males.
-
-The pavlues across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
+In the domesticus strains the mean normalized foci positions were significantly different between sexes and fit the predictions of the 'typical recombination landscape' (with female positions being medially placed and terminally placed in males).
+The p.values across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
 
@@
 
-After confirming the significan sex differences for 1CO position, we apply the familiar models
+After confirming the significan sex differences for 1CO position, we apply the model frameworks
 
@@
 $$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear model**
+
+$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
+
+**M3. Linear model**
+
+$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
 
@@
 ```
-
-For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
-(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
-These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).
-
-**M2. Linear model**
-
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
 
@@
 
-**M3. Linear model**
-
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
 ```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
@@
 
+For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
+(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
+These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).
+
 The male effects were the most significant (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
@@
 
-<we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>
-
+REMOVE THE SISCOTEN PLOTs...
+
+<ONLY SHOW 1 PLOT, we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>
+
+<!-- Ignore this section / kind of results for know
 Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite).
-To account for this we again use the reduced bivalent sets (long.biv and short.biv)
+To account for this we again use the reduced bivalent sets (long.biv and short.biv) -->
 
@@
 ```
-
-These box plot show that females have a much more medial position of single foci bivalents, (much closer to 50% compared to males).
-They also show that Musc males' Foci1 position is slightly more central / medial compared to the same type of positions in the Dom male strains.
-MOLF males have much more medial positions than other strains.
 
@@
 
-```{r show.siscoten, echo=FALSE}
+```{r show.siscoten, echo=FALSE, eval=FALSE}
 #think of a good way to display this
@@
 
-```{r show.norm.siscoten, echo=FALSE}
+```{r show.norm.siscoten, echo=FALSE, eval=FALSE}
 Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
@@
 
-```{r log.reg_siscoten, echo=FALSE}
+```{r log.reg_siscoten, echo=FALSE, eval=FALSE}
 
@@
 
-```{r telo.dist, echo=FALSE}
+```{r telo.dist, echo=FALSE, eval=FALSE}
 
@@
 
-```{r MM.telo, echo=FALSE}
+```{r MM.telo, echo=FALSE, eval=FALSE}
 #HetC
@@
 
-```{r log.reg_telo.dist, echo=FALSE}
+```{r log.reg_telo.dist, echo=FALSE, eval=FALSE}
 
@@
 
-```{r t.test_Telo.Dist, echo=FALSE}
+```{r t.test_Telo.Dist, echo=FALSE, eval=FALSE}
 
@@
 
-```{r cent.dist_show, echo=FALSE}
+```{r cent.dist_show, echo=FALSE, eval=FALSE}
 cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
@@
 
-```{r log.reg_CENT.dist, echo=FALSE}
+```{r log.reg_CENT.dist, echo=FALSE, eval=FALSE}
 
@@
 
-```{r t.test_CENT.Dist, echo=FALSE}
+```{r t.test_CENT.Dist, echo=FALSE, eval=FALSE}
 #CENT.DIST
@@
 
-```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
+```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, eval=FALSE, include=FALSE}
 #IGNORE THIS FOR KNOW, THIS
@@
 
-```{r Q1.mean.nrmF1_table, echo=FALSE,eval=FALSE}
+```{r Q1.mean.nrmF1_table, echo=FALSE, eval=FALSE}
 
@@
 
-In order to test if there are sex differences between crossover interference, a major determinant of the positioning of crossovers along chromosomes, we examined ...
-ran models with the interfocal distance (IFD) of two foci on the same bivalent.
-
--We focus on observations from two-crossover bivalents for more comparable observations.
--comparisons of all IFDs of mulit crossover bivalents is also included
-Interference is a major determinant of the positioning of chromosomes
-
 <!-- 
@@
 
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns
-<!--
-ToDo: make a melt.DF with the IFD1, IFD2, IFD3 in a single cols
-
-explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?
--->
+DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns -->
 
@@
 
-The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical - force mediated through the SC -- In some models this metric should be thought of as the best metric for charaecterizing crossover interference -- bc it reflect the physcial --
-
-The above plots display the raw and normalized area which seperate two foci on the same chromosome.
-In the normalized plot, 9 observations fall above the length threshold of 150.
-The female measures fall more uniformly across the range of (0 to 1).
+Crossover interference is one of the major determinates -- of the CO landscape. We use X number of 2CO bivalents, to test if there are diffferences between males and females in interference. 
+The mean interfocal distance (IFD) for raw and normalized values was calculated from X female and X male mice.
+
+The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical force mediated through the SC. Previous measures found no sex difference in IFD in B6 mice (de Boer et al )
 
@@
 
+<remove the chrm size effects -- or check out model with chromosome length>
+
 to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
@@
 
--- run these comparisons to see if the small NRM IFD are due to shorter chromosomes, which are predicted to have weaker interference.
-
-Looked at the KAZ pattern
-
 Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
@@
 
-Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+Since the most significant t.test were significantly different across sexes for the NORMALIZED effects -- we follow up with the model framework to --check out the potential evolutionary patterns--
+
+We applied the same frame work of models, Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
 
 **M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
+
+**M3. Linear model**
+
+$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
 
@@
 
-The table above should display the slightly unusual pattern, where the coefficients for the significant sex fixed effect are positive and negative in the raw and normalized values respectively.
-That is for the raw IFD values, females are significantly longer but the normalized IFD values, males are significantly longer.
-
-I tested 2 versions of the mixed model for this flavor of trait, raw IFD and normalized IFD measure.
-The tables below are from anova( for the lmer model ).
-Random effect of strain is not significant for ABS IFD, and only slightly significant for the IFD.PER
-
-**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
-
 ```{r IFD_HetC_M2, echo=FALSE}
@@
 ```
-
-**M3. Linear model**
-
-$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
 
@@
 
+ (IFD.PER)
+sex
+MM p = `r`
+
+subsp
+p = `r`
+
 the interaction effects were slightly significant for both raw and nrm.IFD.
-
 The most significant value was the sex effect for nrm.IFDs. 
-
 The the random strain effect was not significant for either model.
 
-For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
+<For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
 The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
-For the M3, only SKIVE*male effect is close to significant
+For the M3, only SKIVE*male effect is close to significant>
 
@@
 This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
-
-Dive deeper into the sex specific pattern for each strain.
-Below are code chunks which show the unusual sex specific results for IFD measures.
-The general pattern is that, female raw IFD > male IFD and  female PER IFD < male PER IFD.
-The scatter plots show that female raw measures are longer than male and for the PER values, the female mean is brought down by an enrichment of short IFDs.
-
-For some strains, PWD, MSM and SKIVE there's a 30% threshold in the male PER IFD distributions. (What does that mean?). How do I test / quantify this pattern? Cluster metric?
 
@@
 
-Above is a table of the proportion of 2Co bivalents which have a norm IFD below 30%, For all strains but KAZ, the females have a greater proportion of these shorted IFD values.
+<!-- junk for Q1.IFD
+
+The above plots display the raw and normalized area which seperate two foci on the same chromosome.
+In the normalized plot, 9 observations fall above the length threshold of 150.
+The female measures fall more uniformly across the range of (0 to 1).
+-->
 

### 2020-02-27 doc/Results.rmd
missed the bus

@@
 
-log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
+#              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
 
@@
 
-```{r Q2.SClength.ttest, echo=FALSE, include=FALSE}
+```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
 #COMPARISONS ACROSS ALL CLASSES

### 2020-02-27 Results_library.Rmd
mostly notes and changes to Q1 and Q2 IFD

---
title: "MLH1 Results Outline"
date: 
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

```{r setup, echo=FALSE, include = FALSE}
library(knitr)
library(ggplot2)
#library(pwr)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)
library(patchwork)
library(here)
options(digits=2)
#setwd("~.")#switch over to here()

#MLH1_data_setup_2.24.20.RData
load(file = here( "MLH1/MLH1_data_setup_2.24.20.RData") )
#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")

#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
#2.2.20
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
load(file="~./MLH1repo/data/CleanBivData_19.2.20.RData")
load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
#skeletonize
SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
#setwd("~./MLH1repo/doc/")
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
#setwd("C:/Users/April/Documents/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")
#load main data file
#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
#load(file="data/MetaData.RData") #this was replacing most updated data file
#load functions
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
```

<!-- Data Set Up Notes -->

```{r organizing.data, echo=FALSE }
#set the order of another column, based on another variable. consistant order for ploting
#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
#-ene endings refer to the cells/units within the cell stage? (I think)
#apply this before ploting
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, category, mouse) %>%
  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
# sort your dataframe, by the focal categories
MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
#use the above line for when order is out of wak
#count the non-quality measures,  #remove non qualit
#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
#MLH1_data <- MLH1_data %>%
#  arrange(category) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
                 .(strain,sex),
                 summarize,
                 sum.ncells = sum(Ncells),
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co),
                 AV.var_MLH1 = mean(var, na.rm=TRUE)

)
#there might be an error for changing wd
#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]

var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
```

```{r misc.setup, echo=FALSE}
colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4',

                       'TOM'= 'indianred1','AST' = 'tan2',
                       'CZECH'= 'brown1',

                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice

                    'CAST'='green','HMI'= 'forestgreen', #greens

                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )
color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
#4!!
#show_col(hue_pal()(5))
#add Rec group,
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
#apply this before ploting, to display strains in desired order
#MLH1_data <- MLH1_data %>%
#  arrange(strain, sex, category, mouse) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
# sort your dataframe, by the focal categories
#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
```

```{r MLH1.FEMALE.TABLE, echo=FALSE}
female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )
female_mouse.av <- add_sex(female_mouse.av)
female_mouse.av <- add_strain(female_mouse.av)
female_mouse.av <- add_subsp(female_mouse.av)
female_mouse.av <- add_species(female_mouse.av)
female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")
female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
#Mol
#Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
#Dom 
Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
```

```{r MLH1.MALE.TABLE, echo=FALSE}
#AP_mouse_table.MLH1_HQ
#find the table that includes the las image data
male_MLH1mouse.av <- AP_mouse_table_w.Ages
male_MLH1mouse.av <- add_sex(male_MLH1mouse.av)
male_MLH1mouse.av <- add_strain(male_MLH1mouse.av)
male_MLH1mouse.av <- add_subsp(male_MLH1mouse.av)
male_MLH1mouse.av <- add_species(male_MLH1mouse.av)
fomo.male_mouse.av <-  male_MLH1mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus") 
fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
#Dom version 
Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
```

```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ
#this blck is duplicated in setupData.rm
#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme
#What are the main DFs for this code chunk
#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
#DF.HetC.MixedModel   <- only Mmusculus
#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
#variance is still chr
#AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
#use the merged DF to filter by age 
#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
#AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
#remove the extra merged cols
#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
#                            subsp.y,subsp.x,
#                    category.y,category.x, strain.x,
#                    strain.y, mouse.1  )  )
#REMOVE EXTRA STRAINS
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
#DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
#AP_mouse_table excluide Qual5 cells
DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
#main df with variables coded as chr
DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
                 .(sex, strain),
                 summarize,
                 all.cells = sum(Ncells)
)
```

```{r MLH1.HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells
#outgroups
#SPIC
HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
#unique(HetC.SPIC$mouse)
#non-passing mice
SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
#calq sex mean
SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
#SPRET
HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
#unique(HetC.SPRET$mouse)
#non-passing mice
#SPRET.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
#CAROLI
HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
#unique(HetC.CAROLI$mouse)
#non-passing mice
#SPRET.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
#cast
HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
#unique(HetC.CAST$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
#HMI
HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
#unique(HetC.HMI$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
#dom
#WSB
HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
#unique(HetC.WSB$mouse)
#non-passing mice
WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
#calq sex mean
WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
#LEW
HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
unique(HetC.LEW$mouse)
#non-passing mice
LEW.remove.mice <- c("4jan17_LEW_f2")
HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
#calq sex mean
LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
#G
HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
#unique(HetC.G$mouse)
#non-passing mice
G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
#calq sex mean
G.mean.MLH1.obs <- ddply(.data=HetC.G, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
#PERC
HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
#unique(HetC.PERC$mouse)
#non-passing mice
#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
#calq sex mean
PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
#MUSC
#PWD
HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
#unique(HetC.PWD$mouse)
#non-passing mice
PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
#calq sex mean
PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
# CZECH
HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
#unique(HetC.CZECH$mouse)
#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
#calq sex mean
CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
# SKIVE
HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
#unique(HetC.SKIVE$mouse)
#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
#calq sex mean
SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
# KAZ
HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
#unique(HetC.KAZ$mouse)
#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
#calq sex mean
KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
#TOM
HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
#unique(HetC.TOM$mouse)
#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
#calq sex mean
TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
#AST
HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
#unique(HetC.AST$mouse)
#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
#calq sex mean
AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
#Molosinus
#MSM
HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
unique(HetC.MSM$mouse)
#non-passing mice
MSM.remove.mice <- c("12sep16_MSM_f3")
HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
#calq sex mean
MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
#MOLF
HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
unique(HetC.MOLF$mouse)
#non-passing mice
MOLF.remove.mice <- c("27oct18_MOLF_m3")
HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
#calq sex mean
MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
```

```{r MLH1.HetC.plot, echo=FALSE, warning=FALSE}
#HETC plot
#change the main data? 
HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#
#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length(adj_nMLH1.foci),
                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd(adj_nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
#remove the 'other'
HetC_strain_table.M$sex <- "male"
HetC_strain_table.F$sex <- "female"
HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
HetC_strain_table <- add_subsp(HetC_strain_table)
#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
#str(HetC_strain_table)
Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
```

```{r juvimice.list, echo=FALSE}
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )
#REMOVE JUVI mice
#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ] 
```

```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
#this block should be in the data set up
#the mouse averages will be used
#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
#cleanCurated_BivData
#removed the no hand foci count
mouse.avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)
#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
mouse.avs_4MM <- add_category(mouse.avs_4MM)
#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
#code if males are high or low?
#REMOVE JUVI mice
juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]  
#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
```

```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
#add this to setup
#remove female, 
#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
#Curated_BivData_male
male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)
#vari
male.av_4MM <- add_sex(male.av_4MM)
male.av_4MM <- add_strain(male.av_4MM)
male.av_4MM <- add_subsp(male.av_4MM)
male.av_4MM <- add_category(male.av_4MM)
#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
#REMOVE JUVI mice
juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ] 
```

```{r totSC_setup, echo=FALSE}
#SC.skel
#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included
SC.skel.OG <- SC.skel #5683
#MERGEING MLH1 with SKEL
MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
#5110, 
OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
#check for DUPs, #check for duplicatations of the random name
MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
#table(MLH1.merge.Skel$category)
#remove outliers
MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
```

```{r totSC.viz4outliers, echo=FALSE}
# -- EXAMINE the ggplot scatter plots for outliers
tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
#round 1
SC.skel.outliers_round1 <- c(
  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",

  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",

  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",

       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",

   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
   "6may15_17apr15_PWD_f1_sp1_7_rev",

      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",

    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
 "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
 "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
 "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,

  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
 "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
 "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
 "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
 "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
 "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
 "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
 "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
 "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
remove.mice_round2 <- c(
    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",

    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",

#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",  
#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",

  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
#SPIC male  x
#SPIC female  x
#TOM x
#AST x
#CZECH male x 
#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
 "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
#SPRET female X
#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
)
#3remove.mice_round2
MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
```

```{r LONG.SHORT.biv_setup, echo=FALSE}
#FROM q1
#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
#from Q2 IFD
#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
##long biv
#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
#table( droplevels( male.long.biv$strain)  )
#FROM q1 
REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                        ncells = length(unique(fileName)),
                         mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                      #make sure 1CO  
                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)

                         )
REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
#divide musc males into groups
REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
                                           0)))
REAL.short_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
                                ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1, 0))
###LONG BIV
REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
                       n_obs = length(Obj.ID),
                        ncells = length(unique(fileName)),
                         mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
                         )
REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
#divide musc males into groups
REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
                                           0)))
REAL.long_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
                                ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1, 0))
#find which one 
```

<!-- 
two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
-->

```{r tot.skel.table, echo=FALSE}
totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                #se for skel

                 mean.bin =  mean(bin_size, na.rm = TRUE)

)
totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
#unorder the factors
totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
#add rec group
totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
OG.totSC.mouse.av.table <- totSC.mouse.av.table
#REMOVE JUVI mice
tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ] 
totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av. total SC area")
mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")
```

```{r, tot.SC.setup, echo = FALSE}
totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
                 .(sex,strain),
                 summarize, 
                 ncells = sum(ncells),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(mean.skel, na.rm = TRUE),
                 mean.bin =  mean(mean.bin, na.rm = TRUE)

)
totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
```

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
INSERT TABLE FOR THE TOTAL SC means ect
The difference between 
bin is short for binary, it reports the binarized area of the SC
skel -- takes the skeletonized midline of the bin objects
-->
# Main Figure

```{r table.for.plot, echo=FALSE, warning=FALSE}
#why do I add in a different table?
#this move this to the setup file
#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed
#pooled
strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)

                       #range_mouse_means = range()
                   #consider adding crhm proportions?
    )
strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")
#adjust order of plot here
strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ",
  "WSB","G","LEW","PERC",
                     "CAST",
                         "SPRET","SPIC" ), order=T )
```

```{r Fig1, echo=FALSE, warning=FALSE}
#might need to nudge the X values
Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+
  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#remove y labfor male

   #segments
#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
 # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label
 #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
   #        y = c(18.5,18.5,33,18.5, 18.5),

  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
    #       fontface = 'italic', hjust = 0.5, size=3.5)
```

```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )
female_mouse.av <- add_sex(female_mouse.av)
female_mouse.av <- add_strain(female_mouse.av)
female_mouse.av <- add_subsp(female_mouse.av)
female_mouse.av <- add_species(female_mouse.av)
female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") 
female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
#table( droplevels(female_mouse.avs_for.analysis$strain) )
#there are cast strains too
Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
#female.specific.plot
```

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
#try customize the plot size
male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )
male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)
#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus") 
just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus") 
just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
#table( droplevels(just.male_mouse.av$strain) )
#there are cast strains too
Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
#fomo.male_mouse.av
male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
male.plot
musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
```

```{r three.plots, echo=FALSE, warning=FALSE}
#merge the 3 plots
#try patchwork
#Fig1_v1_strain + female.specific.plot + male.plot
#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
Fig1_v1_strain / (female.specific.plot + male.plot )
#(p1 | p2 | p3)/p4
```

This is a draft of the main figure. Outgroups are not included in the analysis below. (add horizontal lines for denoting subspecies, group means and expected minimum)

<!--  insert Details of data set cleaning etc is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd -->

```{r main.fig1s, echo=FALSE, include=FALSE}
strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+
  geom_point( size=3, position=position_dodge(.5))+
    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#strain.table.HQ_Fig.og
strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")
strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
  "WSB","G","LEW"), order=T )
Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3, position=position_dodge(.5))+
    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
```

```{r basic.numbers, echo=FALSE, eval=FALSE}
#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
```

```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
#str(AP_mouse_table_w.Ages)
AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
#remove sex.x == na, and strain.x=na or F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
#plot variance and mean (facet sex)
AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
str(AP_mouse_table.HQ)
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
#[1] 0.5750399
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
#0.321601
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
# 0.3515544
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
#-0.03713536
cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
# p-value = 8.623e-09
cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
#remove mice with fewer than 10 cells
AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
#why did I male just a female table?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)
```

# Genome wide recombiantion rate estimates for both sexes

```{r MLH1.summ.stats, echo=FALSE}
#strain.table.HQ_Fig.og  all strains
#strain.table.HQ_Fig  extra strains removed
main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
main.table <- add_subsp(main.table)
main.table <- add_species(main.table)
#main.table %>% filter(sex == "female")
#main.table %>% filter(sex == "male")
#G.HETC.val #1.2
#LEW.HETC.val  #1
#WSB.HETC.val  #1
#KAZ.HETC.val  #1.1
#SKIVE.HETC.val # 0.96
#PWD.HETC.val  #0.88
#MSM.HETC.val #0.9
#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
#strain.table.HQ_Fig
#sum( (main.table  %>% filter(sex == "male"))$Ncells)
#the kable version is bulkier than the other versaion
```

We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. Mean MLH1 foci counts were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes. <!-- state the average number of mice
The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises. -->

<!--
1. Fig1A  magnitude and direction of heterochiasmy;
-general patterns i) female higher and ii) low degree of hetC  (range of 1 to 1.2) (human is 1.6)
-All strain means are above the expected minimum of 20 (1CO per chromosome)
-there are exceptions they are followed up in the sex-specific plots  (male higher)
2. Fig1B and C mouse means and sex specific patterns; 
-female mouse mean distributions are have more spread across the range of means
-male distributions are more clustered, low and high (~30% evolution largest-rapid evolution of gwRR)
-sex-specific evolution, both sexes don't show equivalent elevation of gwRR
-Dom vs Molossinus and Musculus -->

The general patterns of heterochiasmy in house mouse are displayed in Figure 1A. Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) that the genome wide recombination rates averages are greater females compared to males (female biaded heterochiasmy) and ii) the degree of heterochiasmy (F:M rates) is generally low in these house mouse strains, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_. Three notable exceptions of male biased heterochiasmy are from the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.

Examining the mouse means of MLH1 foci per cell seperately for each sex reveal disctinct patterns of variation (Figure1 B and C). The female means can be discribed as more evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B). While in males, the strain specific means seperate more clearly into two groups of high genome wide recombination rates near 30 MLH1 foci per cell and low genome wide recombiantion rates, near 23 MLH1 foci per cell (Figure1 C).

<illustrate sex differences female mean `r mean((main.table %>% filter(sex=="female") )$mean_co )`>

<This striking differences across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories in males and females of the same species. The pattern of strain averages distributed around the total sample mean in females is supportive of stabilizing selection or neutral evolution. In contrast to the pattern from male means which have rapid evolution in a subset of genetic backgrounds this fits a model of directional selection. (We attempt to make better evolutionary models by applying an evolutionary framework)>

```{r female.strain.table, echo=FALSE}
female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells, na.rm = TRUE),

                        mouse.mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#F.strain.table
#F.strain.table
# I don't know why most of the table is blank
```

```{r female.dif, echo=FALSE}
#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)
```

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
#variance plots
musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.female.VAR.plot
#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
#Mol.female.VAR.plot
Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.female.VAR.plot
```

```{r Female.specific.glm, echo=FALSE}
glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
summry.female.strain.effects <- summary(glm.female.strain.subsp)
#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
```

<!--
The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (glm; p= ).
 -->

```{r FEMALE.VAR.MODELS, echo=FALSE}
## within mouse variance models 
lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]
#Q12
Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
#LEW sig
```
<!-- male section -->

```{r male.strain.table, echo=FALSE}
M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

                        totNcells  = sum(Ncells, na.rm=TRUE),

                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),

                        mean.mouse_cV = mean(cV, na.rm = TRUE),

                        mean.mouse_var = mean(var, na.rm = TRUE ),

                        mean.mouse_sd   = mean(sd, na.rm=TRUE),

                        #Standard error (mice . cells)
                        mean.mouse_se  = mean(se, na.rm=TRUE)
    )
#M.strain.table
#same issue as the female table .. ugh
```

```{r male.diff, echo=FALSE}
just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus") 
just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
#how much greater are the High Rec gwRRs
just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
#how much bigger are the high rec male means?
sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
#mean(sans_LewRec_male$mean_co)
PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
#mean(SKIVE_male$mean_co) /  
```

```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
#variance plots
musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.male.VAR.plot
Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Mol.male.VAR.plot
Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.male.VAR.plot
```

```{r Male.specific.glm, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect
glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
#summry.male.strain.effects <- summary(glm.male.strain.subsp)
#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
#1. determine which strain effects are sig
#M2
MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

```{r MALE.SPECIFIC.variance.model, echo=FALSE}
## within mouse variance models 
lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
#all p values kinda high, MSM and TOM are slightly sig
lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
#non sig
lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
lmcv.male.strain_sum <- summary(lmcv.male.strain)
#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
#Q12
Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
#NS
lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
#NS
```

<!--(the figure is made from pooling MLH1 cells)
the male 
- the ranges of mouse means are displayed in Sup table X
- How the means were calculated
-The male gwRR is adjusted for the PAR
-The means for each strains are report in Table 1
-Normality was assessed for the MLH1 counts (include in Supplement histograms)
-->

# Evolutionary Framework

<to better understand this varation in RR within an evolutionary framework. Make table of models>

**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$

**M3. LInear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$

**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$

```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3, message=FALSE}
#THIS IS THE NEW MIXED MODEL
#should be able to use the loaded 
#DF.HetC.MixedModel.HQ
MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
#summary(MLH1.MM.M1)$coefficients
MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value 
MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)
```

We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework. All of the effects within the mixed model M1. were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).

<These results allow use to reject the hypothesis of genome wide recombination rates (varying consistantly by sex across subepcies -- and having variance within subspecies due to strains)>

 <!-- (random strain effect was meant to reflect sampling of gentic backgrounds/standing genetic variation for a subspecies) -- (significant pvalue indicates that the variance is not 0)>
 indicating, fixed effects -- random effect variance due to each strain nested within subspecies -->

```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
#strain = fixed
MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
#table of just coeff anf pvalue
#round(summ$coefficients[,c(1,4)],5 )
#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
```

```{r MLH1.glm.M3, echo=FALSE, digits = 3}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
#round(summ.strain.fixed$coefficients[,c(1,4)],5)
#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
```

<The linear models, allow us to quantifying the significance in sepecific fixed effects and focus on how specific levels and interactions create these differences.>

The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombiantion rates. The subspecies fixed effects tested in M2, had weak to no significance (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M. m. musculus_ and _M. m. molossinus_ respectively), indicating that most of the variance is due to strain effects. As such, we focus on results from M3. The two most significant strain effects were _domesticus^G^_ and _musculus^MSM^_ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]` for _domesticus^G^_ and _molossinus^MSM^_ respectively).

In addition to significant strain effects, ( p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and _molossinus^MSM^_ strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.

Combined these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of sexual dimorphism /heterochiasmy in _domesticus^G^_. No significant sex effect showing that the sex effect on mean CO number is not uniform across the sample but has synergy with speciefic strains.

<domesticus^G^ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`>

A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` number of male and `r length(female_mouse.avs_for.analysis$mouse)` female mean MLH1 foci counts were used in each of these models. For the male specific data,

_musculus^PWD^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ). (with effect sizes ranging from 5, 7, and 2 foci respectively). Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.

three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, _musculus^PWD^_  (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ). (with effect sizes ranging from 5, 7, and 2 foci respectively).

While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.

<!--
M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
<Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect). These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models. 
(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
Check the DF of the number cited above
This is the break down of mice used within the Mixed model
`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
add number of cells
`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
 $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
- only Mus musculus strains with sex matched observations  
- Male gwRR have been adjusted for the PAR
- Q5 excluded
- All the effects are un-ordered factors  
- male ages within 5 - 20 weeks
- Molossinus is coded as a subsp level
also g strain effect/ dom^G^ strain effect 
`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[12,][4]` 
In both models strain by sex interaction effects were significant for (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively. 
-->

<!-- transition gwRR to within mouse variance -->

## Within Mouse Variance in CO Count per Cell

Variance of a trait can contain information pertinant to it's mode of evolution and mechanistically could reflect the strength of constraints acting on the trait. We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependant variable). <note about reduced power>
<in some traits, like gene expression patterns, evolution of the mean -- also has an effect on the variance>

<!-- 
**M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

 **M6. Linear Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex * strain + \varepsilon$$

 **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$>
-->

```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
#summary(MLH1var_M1)
LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
#CV
MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
#USE Q12 (Q123) table
#Q12_mouse_table_w.Ages
#Q12 var
MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
#Q12 cv
MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
```

```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
#summary(MLH1var_M3)
anova(MLH1var_M3)  
as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M3)
anova(MLH1cv_M3)
as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
#Q12 var
MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
#summary(MLH1var_M3_Q12)
anova(MLH1var_M3_Q12)  
as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
#Q12 cv
MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
#summary(MLH1cv_M3_Q12)
anova(MLH1cv_M3_Q12)#sex large F
as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
```

```{r M3.MLH1.var, echo=FALSE}
glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
### Q12 var
MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```
<!--
These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
-->

**the general pattern is that females have more variance (alsmot twice as much (Figure 1))**

For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`). Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e. background noise), we replicated the model analysis using a subset of cells with higher quality scores. These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).

The sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males. While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`). However, in the model using the high quality cells, _musculus^PWD^_ has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.

**the general pattern is that females have more variance (alsmot twice as much (Figure 1))**

<there are some caveats, but the general pattern is that females have more variance ... For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell. This limits the power for detecting significance differences across groups,

the main/general pattern, is that females have significantly higher within mouse variance compared to males. This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.>

<!--
$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
<!--
RW's main qtl paper used the within animal variance, 
@article{wang2017genetics,
  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
  author={Wang, Richard J and Payseur, Bret A},
  journal={Genetics},
  volume={206},
  number={4},
  pages={1841--1852},
  year={2017},
  publisher={Genetics Soc America}
}
references for phenotypic variance connected to robustness to environment, increase in developmental consistency
Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
phenotypic variance leading to adaptation - 
(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-->

<SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE>

(combined the summary of the MLH1 distributions evolution in the mean and consistant pattern of sex differences)

<The greater amount of within mouse variance in females (almost and, (males have no detecable difference across strains for within mouse variance)
-- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes.>

3. -- the striking differences / patterns of variation across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories for male and female rates. 
For female the pattern of Strain averages distributed around a species wide average - fits a model of stabilizing / relaxed / neutral evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic backgrounds / strains which fits a model of directional selection on genome wide recombination rate.

# Evolution of Genome wide recombination rate associated with evolution of mean precursor number

<transition, >

```{r DMC1.setup, echo=FALSE}
#load DMC1 csv
#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_category(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
                 .(category),
                 summarize,

                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.L$stage <- "L"
DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "Z"
DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
#            row.names = FALSE)
DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
```

```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
#think of better color scheme
#make the jitter col thinner or make the stacked dots.
#remember that box-plots add outlier dots.
#change y labels
#experimenting with ggboplot
everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
          color = "stage", palette = "jco",
          add = "jitter", short.panel.labs = FALSE)
  #library gpubr has functions to add in sig/test bars.
#change order
DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
DMC1.boxplot
```

```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
#
DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=rec_group), position=position_jitterdodge()) +

  #facet_wrap(~stage, nrow=1) +

  ggsignif::geom_signif(
    data=anno_df_stage, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High") 
ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low") 
ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
```

```{r plot.t.test, echo=FALSE, include=FALSE}
  #this will add neat bars, but still need to figure out what groups ect it's actually using
#THE stat_compare function adds a arm thing and pvalue for 
#  stat_compare_means(comparisons = my_comparisons)
#status - legend, expression y-axis, facet, facet_wrap SYMBOL
#legend - stage. y=total count, FACet strain
#add column to distinguish high vs low rec strains
DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)

# annotation table with adjusted pvals and y-position of the labels
anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
#
DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=stage), position=position_jitterdodge()) + 
  facet_wrap(~stage, nrow=1) + 
  ggsignif::geom_signif(
    data=anno_df, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
#example for ggplot+ test from here
#https://github.com/kassambara/ggpubr/issues/65
#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
```

```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
#2. t-tests for the DMC1 counts 
#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
#high vs low group (L)
ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value =  0.002145
#high vs low group (Z)
ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value = 0.6628
#dom vs musc (L) (2 subsp vs)
ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
# p-value = 0.003016
#dom vs musc (Z)
ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
#p-value = 0.09588
#KAZ vs Dom (Z) (only 1 L cell)
ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
#p-value = 0.004148
#Kaz Z are higher than Dom
```

```{r DMC1_anova, echo=FALSE, warning=FALSE}
#1. ANOVA, is there a difference in DMC1 counts across strains?
#unorder strain
DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
```

```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
#make category averages from AP_mouse_table.HQ
mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co)
)
DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
                   category == "WSB male" |
                     category == "KAZ male" |
                  category == "MSM male" |
                    category == "PWD male")
#attach this mini DF to the DMC1 data
DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
##RATIOS
DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z

#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
#DMC1.mean.table
DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1

G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )

DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
```

<!--
explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
-->

In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes. DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days). A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_ (Figure2). One mouse was sampled for each strain.

For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the timing of meiosis. Confirming the prediction that evolution of the mean number of crossovers per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).

However the the differences in DMC1 foci were not significant for (later prophase) zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`). Indicating that this marker at early prophase is more precitive of the downstream crossover number differences.

<explain the biological significance of the ratios!>

after DSB formation, DSBs are repaired as either NCO or CO (CO designation step)
The vast majority of DSBs are repaired as NCOs, with a minority (in most species) of approximately 10% are repaired as crossovers.

-DSBs are either repaired as crossovers or noncrossovers (NCO), with conversion tract lengths (30 to 100bp long). (this is also known as the CO repair deicison )

thus the ratio of CO:DSB is in part an indicator of the proportion of DSBs which are designated into COs -- (and the others are repaired as NCO).

The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively). These results indicate that the targets of evolution which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).

<!--
The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`. The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`. `r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M. m. domesticus^WSB^_ , _M. m. domesticus^G^_, _M. m. molossinus^MSM^_, and  _M. m. musculus^PWD^_ respecitively. 
-->

```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
#Subsampling code -- don't report

#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells. Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups. Permutations were done for both cell stages. The actual p value is in red. 
#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
# the 2 groups
#make the 2 pools
High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
#permute drawing / subsampling and performing the t-test
#put the p.value results into a DF!
#permut.SC.cv.diff  <- as.data.frame(
#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
#permuting samples -- and re-licating t-tests
#nrep <- 1000
#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
bunchOpvals.L <- replicate(1000, 
    # t.test
    t.test( sample(High.L.pool, 25, replace = FALSE),
            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
          )
bunchOpvals.L <- as.data.frame(bunchOpvals.L)
bunchOpvals.L$stage <- "L"
bunchOpvals.Z <- replicate(1000, 
    # t.test
    t.test( sample(High.Z.pool, 25, replace = FALSE),
            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
          )
bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
bunchOpvals.Z$stage <- "Z"
#could merge the lists of p-values
replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
#ASSUMPTIONS; normality, equal variance
#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
#plot.pvals.L
#plot.pvals.Z
#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
#n use the var.equal = TRUE
#https://www.statmethods.net/stats/ttest.html
```

```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
#these are histogram plots for testing Normality shape -- move to supplemenatry section
#- check that distributions are close to normal
#- check for equal variance.
DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
DMC1.histogram
#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
#bartlet's test, Levine's test
#The counts look close to normal.
#Potentially figure 2. Using a function to #display the p-vales for mean comparison
```

<!--
interpertation notes, 
the number of DMC1 across stages (L and Z stages) are driven by different equations
Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
*the repair decision can happen at both stages and remove DMC1. formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
OR 
variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
-Baier et al 2014 supports this model (Chmr axis / DNA loop model)
-->

# Evolution of genome wide recombination rate is reflected at the single chromosome level

```{r chrm_prop_data, echo=FALSE, warning=FALSE}
#shut off eval for now because of wd / path error
#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
#try adding the BivData version of the proportions
#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
chrm.class.DF <- add_mouse(chrm.class.DF)
chrm.class.DF <- add_strain(chrm.class.DF)
chrm.class.DF <- add_sex(chrm.class.DF)
chrm.class.DF <- add_category(chrm.class.DF)
#chrm.class.DF <- add_euth_date(meta.data.file2)
#chrm.class.DF <- add_age(meta.data.file2)
#table(chrm.class.DF$category)
#str(chrm.class.DF)
#summarize by category
proportion_stats_category  <- ddply(.data=chrm.class.DF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
                 n0CO = sum(n0CO, na.rm = TRUE),
                  n1CO = sum(n1CO,na.rm = TRUE),
                  n2CO = sum(n2CO,na.rm = TRUE),
                  n3CO = sum(n3CO,na.rm = TRUE),
                  n4CO = sum(n4CO,na.rm = TRUE),

                 per0CO = sum(n0CO) / tot_biv,                 
                 per1CO= sum(n1CO)/ tot_biv,
                 per2CO= sum(n2CO)/ tot_biv,
                 per3CO= sum(n3CO)/ tot_biv,
                 per4CO= sum(n4CO)/ tot_biv,

                 mean.MLH1.foci = mean(MLH1.sum.check),
                 mean.MLH1.foci2 = mean(nMLH1.foci),
                 meanQ = mean(quality)
)
#this is the table that needs melting
#but is also easy to redo by hand
#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
```

```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
#add subspecies to these values
melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
#add the subsps
melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
#need a second DF / the first df
#+ annotate() text 
#annotate()text
#proportion_stats_category,  tot_biv, ncells
#add number of cells and bivalents in the corner
#chrm.class.barplot
```

```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
#prop.test(x, y)  x number of successes, y total  
#strain sums of n
#look up continuity correction
#mouse.level.BivData_table <-- has more male strains
#Mouse.Table_BivData_4MM <- only sex matched strains
#perform test across strain 
#removed all lines with no hand.foci.count entry
Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,

                  nbiv = sum(tot.biv ),              
                 n0CO = sum(n0CO ),
                  n1CO = sum(n1CO ),
                 n2CO = sum(n2CO ),
                  n3CO = sum(n3CO ),
                 n4CO = sum(n4CO )
                          )
Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
for(g in 1:length(Chrm.prop_strain_table$nbiv)){

  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
                                     Chrm.prop_strain_table$n1CO[g],
                                     Chrm.prop_strain_table$n2CO[g],
                                     Chrm.prop_strain_table$n3CO[g] ))  
}
Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
#there are still some wacky entries -- but much fewer than before
chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
#the pvalues for 0 are not sig dif
#the p values for 1 and 2CO are sig diff
#seperate by subsp
chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
#WSB has fewer 2COs than other 3 strains
Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
#KAZ has slightly more 1CO than the other two
#high.female_G.LEW.
chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
#Low.female.G.LEW.
chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
#all females
chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
#the pvals for these 3 class are sig dif
#seperate by subsp
chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
#non of the Dom proportions are different
Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
#all v different
Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
#high vs low
chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
#the low strains do vary slightly in their 1CO and 2CO proportions
#includes SKIVE
Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
#all very sig p values
High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
#these proportions are v different
#no SKIVE
High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
#MSM and PWD only slightly sig dif 2CO
#MSM-MOLF
MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

#PWD-KAZ
PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
```

```{r Chrm.props.plot2, echo=FALSE}
#barplots for proportions of chrm classes
#consider adding this the the setup block (could have done this to category level table)
mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
#DF needs to be melted -- so they all in 1 col
#per code, I've done this by hand
#subset DF for ease
skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
col.names.list.99 <- colnames(skinny.mouse.level)
#all of the cols to use
MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",

               #leave out the Foci
                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],

  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
#no filename for add mouse, strain subsp
MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
#then this DF has to be summarized into categories
chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
                             mean.percent.type = mean(value)
                             #sum is the wrong!
                             #used average percentage, might consider a different summary
)
#can't apply function use this
chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",

          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",

                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",

             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))

chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
                                                                 "Musc",
                                                                 "Mol",
                                                                 "Spretus","Spic",
                                                                 "other") )
#ggplot
chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
#chrm.class.barplot
chrm.class.barplot
```

```{r Chrm.Prop.misc.calulations, echo=FALSE}
total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
#MELT_Chrm.Proportion$Rec.group <- ifelse(grepl("PWD male", MELT_Chrm.Proportion$category), 1, 
#                              ifelse(grepl("MSM male", MELT_Chrm.Proportion$category), 1,
#                                ifelse(grepl("SKIVE male", MELT_Chrm.Proportion$category), 1, 0)))

#Chrm.prop.male.low <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 0)
#Chrm.prop.male.HIGH <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 1)
#strain av tables
#chrm.prop.male.high1
#chrm.prop.male.low
```

<!--
- This is motivation to investigate more traits at the chromosome level
1.	Basic stats - bivalent data numbers 
2.	Rang of co per chrm, mostly 1co 2co
3.	High rec strains have evolved a higher proportion of 2CO
For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
-->

<from the means the expected number of COs per chroosomes ranges from X to X we confirm these numbers Can't have fractions of COs>

We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per chromosomes (bivalent).  Because our observations of MLH1 foci are made in pachtyene cells, our chromosome opbservations are fully synapsesed paired homologous chromosomes (i.e. 4n copies of each chromsomes). Additionally we can distinguish the idenetiy of indiviual chromsomes (Chromsome 1 or Chromsome 2)
for these reasons we use the term bivalent.

Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers (Figure X) with proportion of 1CO:2CO distinguishes the high and low rec males (Figure X). THis confirms the intuitive interpertation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents. Across the female observations the proportions of chromosome classes are not notably different.

<Most of the low rec strains range from `r min(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` _musculus^KAZ^_ to `r max(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` ( _molossinus^MOLF^_ ).>

In the high rec strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).

While the proportions are significantly different
(p = `r`), PWD male and MSM male proportions are more similar to each other than strains from the same subspcies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).

<SUMMARY the / TRANSITION to single bivalen measures>

# Single Bivalent Level Results

In order to deconstruct the cell wide MLH1 counts
deconstruct we 
look at the recombination landscape at the single bivalent level.

From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software. After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent measures were left.

We use these observations to look at the recombination landscapes, here refereing to the i) total length of the SC ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent. <the position is general, not genomic corrdinates>

```{r algorithm_performance_stats, echo=FALSE}
#sort by category, 
#compare total boxes and rate passed
#Curated_BivData  -- pass rate?
#number of mice, number of cells
#mean number of isolate bivs per cell,
#use these two metrics
#cell.level.BivData_table$boxs.IDd 
#cell.level.BivData_table$nboxes_passed
#calq the mean rate by category
cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd 
algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
                        n_obs = length( unique(fileName) ),
                         mean_pass.rate = mean(pass.rate)
                         )
cell.level.BivData_table <- add_sex(cell.level.BivData_table)
algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
                        n_obs = length(unique(fileName)),
                        mean_pass.rate = mean(pass.rate)
                         )
#showing the table..
#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
```

<Table X illustrates the break down of single bivalent observations by category. Error compared to human measures and other details have been discribed in (Peterson_2019).>

A potential downside of the image analysis algorithm we used is that not all bivalents per cell are isolated. The ranges of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to 
`r max(algorithm.stat.category$mean_pass.rate)` ) _musculus^KAZ^_ female, we assume that the isolation process is not biased. However, because there are hundreds of observations per category, we assume the set of single bivalents are equally representative of general patterns.

Using this data set, we address two main questions 1) Which traits are sexually dimorphic? and 2) Which traits fit distinguish the high and low recombining strains for males?

# Q1 Sex Differences in SC-AE Lengths
<!-- breif background on longer female SC  (mouse, human) Gruhn-->

(in many mammal species -- SC lengths are longer in females - yet most all of these observations come from female biased heterochiasmy species) -- Our data set provides an opportunity to test (how SC length and chromatin compaction relates to heterochiasmy)

if the canonical female biased heterochiasmy pattern and long SC lengths are uncoupled in instances / our strain with male-biased heterochiasmy ( _musculus^PWD^_  and _molossinus^MSM^_). A simple model might predict if SC length is the strongest predictor for number of MLH1 foci er cell the high recombining males would have longer SC length proportional to differences in the number of MLH1 foci per cell.

```{r Q1.Short.Long_biv.setup, echo=FALSE}
#REAL.long_mouse_table
#REAL.short_mouse_table
#make the Q1 strain list
Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
```

To account of the obvious confounding effect of sex chromosome differences in females and males, we use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data. A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes. Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes.

This 'short biv' data set has advantages in that it is closer to sampling across a similar set of chromosomes (ideally Chromosomes 15 to 19) and it excludes the XX in females, which is predicted to be the 3rd longest based on the physical (Mb) length.

```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
#CLEAN UP THE FULL DATA SET
#adjust order of plot here
Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
  "WSB","G","LEW","PERC",  
"PWD", "SKIVE",  "KAZ",
     "MSM", "MOLF",   
                     "CAST",
                         "SPRET","SPIC" ), order=T )
#nested boxplots, pooled Bivalents
Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
```

```{r Q1.SC.show.short.biv, echo=FALSE}
Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
Q1.SC.short.mouse_scatter_boxplot
#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
#skiveQ1.SC.short.mouse_scatter_boxplot
#there are 2 lowish female means, 
```

```{r Q1_SC.biv.short_ttest, echo=FALSE}
#run t.tests for the mouse av. for reduced single bivalent datasets
#WSB
Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
#G
Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
#LEW
Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
#PWD
Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
#KAZ
Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
#SKIVE
Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
#MSM
#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
#1 male, 3 females measures
#try to merge the manual - whole cell observations (at least 2 males)
```

For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer the largest difference across sexes by strain are>  
**`r` the magnitude differences between the mouse mean short bivalent SC range from X to X**

(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.

An additional measure of a smaller subset of single bivalents (which excludes the sex chromsoomes), we tested/compared a measure of the SC lengths for the whole genome (whole cells) using a image nalysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang)

We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads. A total of X images were run trhought the image analysis pipeline, and after outliers removed, X remained. Mouse means were calculated for their total SC area per cell and are displayed in figure X.

The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes. Breifly/in effect the oocytes have an additional autosome, and the spermatocytes have a mostly recombinationly inactive unit of SC (the XY).

```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")

Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
   "WSB","G","LEW","PERC", 
  "PWD", "SKIVE",  "KAZ",
  "MSM", "MOLF"), order=T )
Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
#order the strains
Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
  arrange(strain) %>%
  mutate(mouse = factor(mouse))
Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
#add category mean
#totSC.strain.av.table %>% filter
totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")
Tot.SC_mouse.av_Dom
totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
Tot.SC_mouse.av_Musc
totsc.mol <- Q1.totSC.mouse.av.table %>% filter( subsp == "Mol" )
totsc_strain.av.mol <- totSC.strain.av.table %>% filter( subsp == "Mol" )
Tot.SC_mouse.av_Mol <- ggplot(data= totsc.mol, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totsc_strain.av.mol, aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
Tot.SC_mouse.av_Mol
```

<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )>

```{r Q1.total.SC_t.test, echo=FALSE}
#run t-test for male female avergage for each strain
#ALL
Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)
#dom
Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
#WSB
Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
#G
Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
#LEW
Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
#Musc
Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
#PWD
Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
#KAZ
Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
#SKIVE
#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
#cell level total SC
#MLH1.merge.Skel   $skel_size
#error not enought female observations, ONLY 1 female SKIVE mouse
Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
#MSM
Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
```

```{r Q1.totSC_male.first.comp, echo=FALSE}
Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
```

In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; _domesticues^WSB^ _ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^ _  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^ _ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^ _ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, KAZ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^ _ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` 
<We estimate the effect of the sex chromosomes from hand measured data...the XX (3rd longest chromsomes) in females is ~10% of total the summed SC area, the average propotion of the XY in spermatocytes is XX%?  however the morphology of the XY is highly variable across cells>
<!-- Dom males are shorter than this set of high rec mice
The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>-- we follow up with models (which focus on sex differences).  -->
<**t.test results for total.SC**
all values (F vs M) (p= `r Q1_tot.sc_t.test.all$p.value`)
Dom (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.DOM$p.value`)
Musc (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.MUSC$p.value`)  (note, missing SKIVE, not enough female measures)  
MSM (F vs M) -- (ttest; p= `r Q1_tot.sc_t.test.MSM$p.value`)>
<explain why I'm applying models>
We follow up the general result of females have longer SC lengths -- by applying the same model frameworks which we used to analyze the mean MLH1 foci per cell -- to understsand the evolutionary framework.
**M1. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$
**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
#str(Q1.lmer_totSC.mouse.av.table)
#unorder factors
Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
#TOTAL SC AREA!!
lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE) 
lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE) 
lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE) 
lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value 
#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
sex.results =  lmer.Q1_mus.av_totSC_sex, 
               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
lmer.Q1_mus.av_totSC_strain)
#Q1.lmer.totSC_results
#G
```
```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value 
#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)
#Q1lmer.MM_SC_short_results
```
In the mixed models for the sex differences mean **short bivalents** length as the dependant variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).  
For the mixed model run on the mean total SC metric, there are more significant effects for strain and subsp for the mixed model using mean **total sc** per cell as the dependant variable, the three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
#total-SC
#Q1.lmer_totSC.mouse.av.table
Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
#subsp effect, many strain, #male
#G * male v.sig and male*LEW
Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
#LOT OF SIGNIFICant strain effects in this model
```
```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
#summary(Q1.glm_short.SC_M2)
summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
```
```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
#summary(Q1.glm_short.SC_M3)
summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
#sex male very sig -- only sig
```
In linear models (M2 and M3) for **mean short bivalent length**, male is most significant (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`). In the linear model M2 the skive strain and strain by sex effects are slightly significant (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively), however these could be due to low sampling of skive females.
The following effects had significant p values for the **total SC** from the above model (for the linear models M2 and M3), subspecies _musculus_ (glm M2; p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`), subspecies _molossinus_ (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`), sex male (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`), strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,

For M3 in the mean total SC lengths there were many strain effects are significant, but the largest are sex (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`)
but also some strain effects, G strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][2]`), LEW strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][3]`), PWD strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][4]`). (this means...). Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized / quantified at this level.

<!--
(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
-->

<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >

<In summary, females (oocytes) have different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>

<1. the longer female sc's is consistant and seen across all strains, even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>

```{r Q1.SC.Q12, echo=FALSE}
#make comparisons for the high quality cells
#get Q12 cells from MLH1.merge.Skel
#calq the mouse averages
MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                #se for skel

                 mean.bin =  mean(bin_size, na.rm = TRUE)

)
totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
#unorder the factors
totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
#add rec group
totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
#REMOVE JUVI mice
tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ] 
totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
                 .(strain, sex),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                #se for skel

                 mean.bin =  mean(bin_size, na.rm = TRUE)

)
#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
#unorder the factors
totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
```

```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
#Q12.Tot.SC_mouse.av_Dom
#Tot.SC_mouse.av_Dom
Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
#Q12.Tot.SC_mouse.av_Musc
#Tot.SC_mouse.av_Musc
Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
Q12.Tot.SC_mouse.av_Mol
Tot.SC_mouse.av_Mol
```

<Q12 total SC mouse means! In Dom when running the Q12 SC lengths, the sex difference seems to get larger (male mean total SC gets lower).  In Musc the female mean total sc gets lower (maybe fewer cells (SKIVE is removed)). For Mol / MSM female mean stays about the same, MSM and MOLF means converge around 1600.>

```{r show.long.biv, echo=FALSE, include=FALSE}
#mouse av
Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
```

```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
#Q1.real.long_mouse_table
Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
#G
Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
#LEW
Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
#PWD
Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
#KAZ
Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
#SKIVE
Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
#MSM
#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
#run t.tests for the mouse av. for reduced single bivalent datasets
```

```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
#1CO  lmer.Q1_SCLength_M1
Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
sex.results =  Q1.lmer_SCLength.1CO_sex, 
               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
Q1.lmer_SCLength.1CO_strain)
#2CO
Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
sex.results =  Q1.lmer_SCLength.2CO_sex, 
               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
Q1.lmer_SCLength.2CO_strain)
#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
```

```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
#removing these plots for space
Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
#Musc.all
Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
#BivData Level
Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc") 
Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av. SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
#BivData Level
Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom") 
Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
```

## Adjusting for XX

Summarize for now, refine this section later.

1. illustrate problem(affects mostly SC length)  
2. Expected impact on sex comparisons, estimated effect size of the X  
3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
4. permutations of 19 female, 19male, 20female 20 male

<Of all female single bivalents observations, 5% are XX (1 of 20). 
- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX>

# short transition

<transition>  
-following up on SC area - with how COs are place - placement and number of COs on single bivalents

1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
-- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).

2. We focus on two metrics/ aspects i) the placement of single foci along a bivalent (just 1CO) and ii) the placement of two foci on the same bivalent / the interfocal distance of 2CO bivalents (reflects crossover interference)

#Q1. 1CO Position Sex Differences

```{r Q1.F1pos_long.short_setup, echo=FALSE}
#REAL.short_mouse_table
#REAL.long_mouse_table
Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
```

<!-- motivating
To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
-->

We focus on the foci positions from one crossover (1CO n= `r `) bivalents since the landscape patterns for multi crossover bivalents (2COs) will be highly influenced by crossover interference.

The CO positions are normalized by the SC length and anchored at the centromere with the normalized position values ranging from 0 to 1, (refered here as terminal or telomeric position).

```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
#main figure of density plots,
MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
#clean up all the non 1 to 4 hand.foci counts
MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
# remove the na's
MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
CO.Pos.density_1CO
#category specific plots, don't run for now.
Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
#Musc male
Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
#Dom male
Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
```

```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \n all bivalents")
Q1.mouse.av.biv_scatr_box
```

```{r Q1_1COpos.t.test, echo=FALSE}
#testing,
#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
#WSB
Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
#G
Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
#LEW
Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
#PWD
Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
#KAZ
Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
#SKIVE
Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
#MSM
Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
#Single Biv Level
#single.biv.1CO_pos_t.test <- t.test(

 # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],

#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
#  )
#t.tests for mous.av_nrm.1CO.pos 
```

In the domesticus strains the mean normalized foci positions were significantly different between sexes and fit the predictions of the 'typical recombination landscape' (with female positions being medially placed and terminally placed in males). The p.values across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).

With the musculus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM. Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).

After confirming the significan sex differences for 1CO position, we apply the model frameworks

**M1. Mixed Model**

$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$

**M2. Linear model**

$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$

**M3. Linear model**

$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$

```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
#make sure this dataset has all the right strains
Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE) 
lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value 
lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
sex.results = lmer.Q1.nrm1CO_M1.sex , 
               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
#lmer.Q1.nrm1CO_M1_results
#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
```

```{r Q1.CO1pos_glm.M2, echo=FALSE}
#in case factors aren't ordered
#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
#summary(Q1.glm_F1pos_M2)
summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
#DOM
Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1.glm_nrmF1pos_M2_Dom)
#MUSC
Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1.glm_nrmF1pos_M2_Musc)
#MOL
Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1.glm_nrmF1pos_M2_Mol)
```

```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
#summary(Q1.glm_1CO.pos_M3)
summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
```

For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`). These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).

The male effects were the most significant (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`). While the musc subspecies, LEW and KAZ strains had slightly signifcant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).

(these results will be explored more fully in the Q2 section). The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).

REMOVE THE SISCOTEN PLOTs...

<ONLY SHOW 1 PLOT, we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>

<!-- Ignore this section / kind of results for know
Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite). To account for this we again use the reduced bivalent sets (long.biv and short.biv) -->

```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
#real.long.bivData
Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nShort Biv")
Q1.short.biv_scatr_box
Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nLong Biv")
Q1.long.biv_scatr_box
```

```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
#REAL.short_mouse_table
#REAL.long_mouse_table
#mouse.av.long.biv.table
Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nLong Biv")
Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nShort Biv")
```

<!--
### Siscoten
The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
The logic of how the sis-co-ten metric is outlined in the figure below. The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO. This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
![from (Lee, J. (2019). Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?. Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
```{r show.siscoten, echo=FALSE, eval=FALSE}
#think of a good way to display this
sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
sis.co.ten.plot_WSB_female
sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
sis.co.ten.plot_WSB_male
sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_LEW_female
sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_LEW_male
sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_PWD_female
sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
#sis.co.ten.plot_PWD_male
Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_PWDf
jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_PWDm
jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_WSBf
jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
  geom_jitter()+facet_wrap(sex~strain)
#jitter_siscoten_WSBm
```
Males have much clearer separation of siscoten across chrm classes. This is emphasized when SC length is also plotted. It seems like musc males have higher amounts of this metric compared to Dom males.
To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
BUT females have a greater range -- so maybe it's just a scale issue.  
```{r show.norm.siscoten, echo=FALSE, eval=FALSE}
Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
Nrm.sis.co.ten.plot_WSB
Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
Nrm.sis.co.ten.plot_G
Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
Nrm.sis.co.ten.plot_LEW
Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
Nrm.sis.co.ten.plot_PWD
Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
Nrm.sis.co.ten.plot_KAZ
Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
Nrm.sis.co.ten.plot_SKIVE
Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
Nrm.sis.co.ten.plot_MSM
```
I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
#glm for siscoten
#mouse.avs_4MM  - mouse averages with mean sis.co.ten
#MM
lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
#sex and sex*subsp interaction effect are significant for mean sis.co.ten
lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(lmer_MM_siscoten) 
#thing1  <- summary(lmer_MM_siscoten)
#drop1(lmer_MM_siscoten, test="Chisq") 
rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
rand_siscoten_MM
#slightly sig
```
```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
#move this to the set up blocks
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
#test high vs low Musc strains
#mouse.avs_4MM
```
```{r log.reg_siscoten, echo=FALSE, eval=FALSE}
#replace the table
male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
#SISCOTEN
#mouse av, norm.first foci is from 1CO bivalents
log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,

                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
summary(log.reg_SIS.co.ten_mouse.av)
#single biv level
log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_Sis.co.ten_biv)
```
```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
#bivData -- 
t.test_mouse.av_siscoten <- t.test(

  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],

  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
  )
#1CO
t.test_mouse.av_siscoten_1CO <- t.test( 
  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],

   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
  )
#2CO
t.test_mouse.av_siscoten_2CO <- t.test( 
  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],

   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
  )
#3CO
t.test_mouse.av_siscoten_3CO <- t.test( 
  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],

   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
  )
####
#Single Biv Level -- from the main 
#pull the single BivData -- MUSC
#check that these have the extra male strains
t.test_single.biv_siscoten <- t.test(

  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],

   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )
#1CO
mouse.avs_4MM$mean.siscoten_2CO
```
All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
### Telomere and centromere Distance
My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC). In the plots below each point is a single bivalent. I choose not to use the mark for centromere because it seems noisy and inconsistent... 
```{r telo.dist, echo=FALSE, eval=FALSE}
sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
  geom_point()+facet_wrap(sex~strain)
Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
telo_jitter
nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
nrm.telo_jitter
#why do a different number of rows removed each time??
```
Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females. In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range. In females the ranges of telomere distances have much more overlap.   
```{r MM.telo, echo=FALSE, eval=FALSE}
#HetC
lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
drop1(lme_MM_telo.dist, test="Chisq")
as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
#sex and sex*subsp interaction effect are significant for mean sis.co.ten
lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
#anova(lmer_MM_siscoten) 
#thing1  <- summary(lmer_MM_siscoten)
drop1(lmer_MM_telo.dist, test="Chisq") 
rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
rand_telo.dist_MM
```
Mixed model result summary:
```{r log.reg_telo.dist, echo=FALSE, eval=FALSE}
Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
#TELO.DIST
#mouse av, norm.first foci is from 1CO bivalents
log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,

                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_Telo.Dist_mouse.av)
#single biv level
log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_Telo.Dist_biv)
#PER versions
log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_Telo.Dist.PER_biv)
```
```{r t.test_Telo.Dist, echo=FALSE, eval=FALSE}
#mouse av
t.test_mouse.av_telo.dist <- t.test(

  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],

  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
  )
#Single Biv Level
single.biv_t.test_Telo.Dist <- t.test(

  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

#Telo Dist PER
single.biv_t.test_Telo.Dist_PER <- t.test(

  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],

   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

```
```{r cent.dist_show, echo=FALSE, eval=FALSE}
cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
cent.dist_jitter
norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
norm.cent.dist_jitter
```
```{r MM.cent.dist, echo=FALSE, eval=FALSE}
lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
exactRLRT(lmer_MM_cent.dist)
```
```{r log.reg_CENT.dist, echo=FALSE, eval=FALSE}
#CENT.DIST
#mouse av, norm.first foci is from 1CO bivalents
log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,

                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_CENT.Dist_mouse.av)
#single biv level
log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_CENT.Dist_biv)
#PER versions
log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_CENT.Dist.PER_biv)
```
```{r t.test_CENT.Dist, echo=FALSE, eval=FALSE}
#CENT.DIST
#mouse av
t.test_mouse.av_Cent.dist <- t.test(

  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],

  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
  )
#Single Biv Level
single.biv_t.test_Cent.Dist <- t.test(

  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
  )

#Cent Dist PER
single.biv_t.test_Cent.Dist_PER <- t.test(

  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )

```
The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males. Females have more overlap in the distributions of centromere distances across chromosome class compared to males.-->

```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
CO.Pos.boxplots_1CO
```

```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
CO.Pos.boxplots_1CO
```

```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, eval=FALSE, include=FALSE}
#IGNORE THIS FOR KNOW, THIS 
#filter all Curated biv data for 1 strain
#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
# the stardell paper sums the number of COs for bins (10)
#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
#but then the model would be close to logistic
#summing the number of COs per bin, losses the chrom specific data....
#!! need to use a melted Data frame to capture all nrm.foci obs
#wsb.ttest<- mouse.avs_4MM %>% filter()
#start with same chrm classes
#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
#or AIC(model1,model2)
```

```{r Q1.mean.nrmF1_table, echo=FALSE, eval=FALSE}
#make a table for the sex specific values of the normalized F1 positions
#for strain, sex -- Curated_BivData
#or (Mouse.Table_BivData_4MM)
#norm.first.foci.pos
Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,

                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
                  n1CO = sum(n1CO )
)
```

```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
#real.long.bivData
#MELT_Curated_BivData
#these isn't long.biv data
MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
real.long.bivData
MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
MELT_long_biv <- add_sex(MELT_long_biv)
MELT_long_biv <- add_strain(MELT_long_biv)
MELT_long_biv <- add_subsp(MELT_long_biv)
MELT_long_biv <- add_category(MELT_long_biv)
long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
Dom.male_MELT_long
long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
Musc_male_MELT_long
#there are v few observations in this dataset
```

```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
#LOGISTIC REGRESSION for FOCI POSITION
#mouse av, norm.first foci is from 1CO bivalents
log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
summary(log.reg_1CO_pos_mouse.av)
#single biv level
log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
summary(log.reg_1CO_pos_biv)
```

#Q1. Sex Differences in CO Interference (IFD)

<!-- 
Brief lit review for IFD / interference expectations  
**Petkov et al 2007,  CO interference underlies sex differences in RR**  
"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb. However, the interference distance is the same in terms of bivalent length. We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes at the pachytene stage of meiosis."
Chrm1 genetic map, 
(other human and mice refs 
Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns -->

<!--  set up -->

<!--  plots-->

```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
#first draft of the IFD comparisons
#make the mouse boxplots -- instead of geom_points
#Only use 2CO bivalents
Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
#boxplots
boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
#order is out of wack, but -- good for now
boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, PER")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)

first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))

#+theme(legend.position="none")
```

```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
```

<!-- EXPLAIN WHY WE SHOULD CARE ABOUT RAW AND PER IFD values-->

Crossover interference is one of the major determinates of the CO landscape. We use X number of 2CO bivalents, to test if there are diffferences between males and females in interference. 
The mean interfocal distance (IFD) for raw and normalized values was calculated from X `r ` female and X`r ` male mice.
The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical force mediated through the SC. Previous measures found no sex difference in IFD in B6 mice (de Boer et al )

<!--t.tests -->

```{r Q1.IFD_Ttests, echo=FALSE}
#below are calq just from 2COs, have no fear
#mean_IFD.2CO_ABS
#mean_IFD.2CO_PER
#FULL all strains
Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,

                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
#p-value = 0.07
Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
#p-value = 1e-12
#WSB
Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
#LEW
Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
#G
Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
#PWD
Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
#KAZ
Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
#SKIVE
Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
#MSM
Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
#FULL all strains
Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,

                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
#p-value = 0.3
```

t.tests are performed on the normalized metrics for establish if there are sex differences in the IFD meterics (raw and normalized).

The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`). BUT WHEN G IS REMOVE, the pvalue is no longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`). <look closer at G>

Mean IFD^PER^ measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`) between sexes, with males have stronger crossover interference (in 2CO bivalents)

When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
MSM   `r Q1.IFD.PER_t.test_MSM$p.value`
SKIVE  `r Q1.IFD.PER_t.test_SKIVE$p.value`
PWD  `r Q1.IFD.PER_t.test_PWD$p.value`
MSM  `r Q1.IFD.PER_t.test_MSM$p.value`
LEW  `r Q1.IFD.PER_t.test_LEW$p.value`
G  `r Q1.IFD.PER_t.test_G$p.value`
WSB  `r Q1.IFD.PER_t.test_WSB$p.value`

<The KAZ strain is the only strain without significant p value for the normalized IFDs, due to KAZ males having a higher rate/enrichment of short normalized IFDs.>

(raw measues of IFD, female lengths are weakly significantly longer reflecting the longer axis lengths). When the underlying differences in chromatin compaction are controlled for (IFD^PER^), males have a greater propotion of single bivalents seperating two foci on the same bivalent, indicative of stronger interference.

<!-- short and long -->

```{r Q1.IFD.long.short.setup, echo=FALSE}
#real.long.bivData
#real.short.bivData
Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)

                     )
Q1.IFD.long <- add_sex(Q1.IFD.long)
Q1.IFD.long <- add_strain(Q1.IFD.long)
Q1.IFD.long <- add_subsp(Q1.IFD.long)
Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)

                     )
Q1.IFD.short <- add_sex(Q1.IFD.short)
Q1.IFD.short <- add_strain(Q1.IFD.short)
Q1.IFD.short <- add_subsp(Q1.IFD.short)
Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
#t.test. mouse av
Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)

Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
#both the short and long norm IFD are NS

#t.test Biv
Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
```

```{r Q1.show.IFD.things, echo=FALSE, eval=FALSE}
nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
#looked up image, measures confirmed 
####
nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
#21may18_KAZ_m2 -- seems like it might be an outlier
#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
#second foci slightly smaller, kinda fainter -- but also much brighter than background
#check all KAZ males
nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_KAZ.female_BIV.data
nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_PWD.female_BIV.data
nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_LEW.female_BIV.data
nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_G.female_BIV.data
nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
###look at why KAZ males are weird
nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
#the difference the the Nrm IFD cut off is striking with these plots
nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
#this might 
nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
```

(random results below)

There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
off. Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?  (uniform landscape in females -- predicts a mean 1CO position of .5)

Female landscape / distribution of interference – seems like there might be constraints -- a lower cut off og 25% and an upper cutt off of 75%

there is a cut off of 75% -- it is rare to oberserve bivalents with very strong interference (more than 75% or less than 25% of the chrm separating 2 foci on the same chromsomes) (is this differen in males  -- do their distributions exceed 75% IFD^PER^?)

**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference

<remove the chrm size effects -- or check out model with chromosome length>

to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets. Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)

Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female). These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).

Considering that the short-long bivData is much smaller than the full data set

<!-- lmer/ glms-->

### Mixed Model Tests, Fixed Effects

Since the most significant t.test were significantly different across sexes for the NORMALIZED effects -- we follow up with the model framework to --check out the potential evolutionary patterns--

We applied the same frame work of models, Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.

**M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$

**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$

**M3. Linear model**

$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$

```{r Q1.IFD.M1, echo=FALSE, message=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL
#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
#M1_IFD_2CO
lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
#M1_IFD_2CO
lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
#don't understand why 2 rows?
#M1_IFD_2CO
lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
#these 2 below are LR tests
#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
#exactRLRT(lmer.IFD.PER_MM.M1)
```

```{r IFD_HetC_M2, echo=FALSE}
glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)
#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
#PER
glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)
#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
```

```{r IFD_HetC_M3, echo=FALSE}
glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
#round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
#PER
glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
```

For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.

 (IFD.PER)
sex
MM p = `r`

subsp
p = `r`

the interaction effects were slightly significant for both raw and nrm.IFD.
The most significant value was the sex effect for nrm.IFDs. 
The the random strain effect was not significant for either model.

<For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw. The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance. For the M3, only SKIVE*male effect is close to significant>

For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males. SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure.

Overall There's a low amount of significant effects across the 2CO IFD measures. This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.

```{r Comp_Ratios, echo=FALSE}
#calq ratios of belo 30
top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3) 
bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) 
ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
DF.strain.list <- list(
  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
  MSM.Bivalents.DF, MOLF.Bivalents.DF
  )
entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))

i = 1
for(x in (DF.strain.list) ){

  #print(head())

  #these filters won't apply to the indexed DF
  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)

  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)

  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)

  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)

  i = i+1

  }
#entr_DF
entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
```

<!-- junk for Q1.IFD
The above plots display the raw and normalized area which seperate two foci on the same chromosome. In the normalized plot, 9 observations fall above the length threshold of 150.  The female measures fall more uniformly across the range of (0 to 1).
-->

# transition Q1 to Q2

we tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.

# Q2 SC Length

<put in plot for total SC and mean gwRR count>

<motivation: the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>

<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class. Mainly that 1CO are shorter in the high rec strains then low rec strains. This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up. Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.  This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group. They have a lower probability of 2COs below a certain SC length threshold.>

while testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR

<!-- basic stats
`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images. X observations removed for having extra signal from other cells. `r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
 place set ups for total SC, short biv and long biv here -->

```{r Q2.SC.short.long.setup, echo=FALSE}
#table(male.long.biv$strain)
#the Long biv dataset is quite unbalance (esp MSM)
male.long.biv <- real.long.bivData %>% filter(sex== "male")
#mouse av for long biv
cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                         #mean.first.1CO
                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
                         )
Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
#divide musc males into groups
Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
                                           0)))
Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
male.short.biv <- real.short.bivData %>% filter(sex== "male")
Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                        #mean.first.1CO
                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
                         ) ) 
Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
                                           0)))
Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
```

<!-- plots -->

```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
#adjust order of plot here
Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
  "WSB","G","LEW","PERC",  
"PWD", "SKIVE",  "KAZ",
     "MSM", "MOLF",   
                     "CAST",
                         "SPRET","SPIC" ), order=T )
Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
#nested boxplots, pooled Bivalents
#x axis for hand.focicount -- good for seperating the points --- but it pools 
#prediction for Dom strains: low variation 
Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
#+theme(legend.position="none") 
```

```{r Q2.SC.long.biv_plots, echo=FALSE}
#REAL.long_mouse_table
#REAL.short_mouse_table
#real.long.bivData
#real.short.bivData
shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")
mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av. \nMale Long Biv SC Lengths")

#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av. \nMale Long Biv SC Lengths")
```

```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
#MLH1_data <- MLH1_data %>%
#  arrange(strain, sex, mouse) %>%
#  mutate(file.name = factor(file.name))
mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")
mouse.av_totsc_male
```

```{r Q2.SC_MLH1.by.totalSC, echo=FALSE}
#make plot of mouse averages for MLH1 and total SC
bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
X.by.texas
```

```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
#investigating potential skew / technical effects in the algorithm
Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
#not sure what my expectations for this plot are, 
median(bep$oldest.image.date) #"2017-08-03"
#there might be an issue with the algorithm performance 
#given the parrelle diagonal lines -- 
#this might be due to the dates imaged  2017-02-07
bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
#there's evidence that the algorithm might have some biased effects -- 
#that will be investigated later.
```

<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
remember the pairwise t.test
Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
G-WSB is sig, WSB-LEW are slightly sig
-->

```{r Q2.totSC.ttest, echo=FALSE}
#totalSC high and low rec group
Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
#p-value = 0.007
#mol
Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
#musc
Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
```

```{r Q2.short.long.ttest, echo=FALSE}
#long and short 
#mouse.av.short.biv.table
#Q2mouse.av.short.biv.table
Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
#p-value = 0.9
#mouse.av.long.biv.table
Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
#p-value = 0.2
#mUSC-short
Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
#mol-short
#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
#musc-long
Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
#mol-long
#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
#not enought observations 
```

```{r Q2.SC_HvL_allBiv, echo=FALSE}
#High-Low_All-BivData
Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
#short
Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
#long
Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
```

**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

Molossinus have significant difference between mouse means for total sc, between the high and low strains (mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`). The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).

```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
#Dom-Musc
Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
#p-value = 0.5
#Dom-Mol
Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
#p-value = 0.1
```

```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
#DOM-musc-long
Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
#p-value = 0.08
#DOM-musc-short
Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
#p-value = 0.8
#DOM-mol-long
Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
#p-value = 0.4
#DOM-mol-short
Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
#p-value = 0.09
```

```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
#DOM-MUSC-ALL
Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 4e-13
#DOM-MOL-ALL
Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 6e-10
#DOM-musc-long
Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.2
#DOM-musc-short
Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.8
#DOM-mol-long
Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.4
#DOM-mol-short
Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
#p-value = 0.06
```

<across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).>

<!-- models -->

```{r Q2.totSC.glm, echo=FALSE}
#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel
Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)
Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
```

```{r Q2.glm.SC.long, echo=FALSE}
#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
#table(mouse.av.long.biv.table$mouse)
#Q2mouse.av.long.biv.table
glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
```

```{r Q2.glm.SC.short, echo=FALSE}
#Q2mouse.av.short.biv.table
#mouse.av.short.biv.table
glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
```

```{r Q2.SC.long.glm.bivglm, echo=FALSE}
#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
#summary(glm_SC.male_Musc_long.biv)
#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
## logistic regression for Rec group
#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
#summary(log.SC.male_long.SC)
#log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
#              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
```

**For the total SC glm**

<Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
1. no difference in DOM strains (for SC lengths )
2. PWD > SKIVE > the rest
3. Msm > mol>

For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)

In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).

<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>

<For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW, G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.>

<!--24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM-->

<just like the sex differences in SC area -- this points to differential usage of the SC for placing CO -- investigate the recombination landscape>

<!--
blocks of other male Q2 - SC lengths
what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-->

**The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).**

**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`). The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**

**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**

> **Summary Points**
- confirm that the positive correlation of SC and CO number is maintained

- Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.

- (the glms, indecate there are many strain effects), Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.

- For the M3 models of mouse av.s across SC meterics, for many of the SC metrics, most all of the strain effects are significant

- for the M2 models of the mouse av, musc - mol are significant fixed effects

- many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging

- th predictions for M3 (strain) differencs (motivated by the gwrr variation are not met,  (correlation of SC means and CO means?))

<Each point is a mouse average SC. When all bivalents are pooled, the logistic regression SC lengths predict the groups. When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class. For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.>

<!-- junk chunks below -->

```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2,  family=binomial(link="logit"))
summary(log.SC_male_chrm.class_all)
#use the 1CO or 2CO DF
Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
summary(log.SC_male_chrm.class.WSB)
Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_G,  family=binomial(link="logit"))
summary(log.SC_chrm.class.G)
Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
summary(log.SC_chrm.class.LEW)
Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
summary(log.SC_chrm.class.PWD)
Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
summary(log.SC_chrm.class.SKIVE)
Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
summary(log.SC_chrm.class.KAZ)
Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
summary(log.SC_chrm.class.CZECH)
Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
summary(log.SC_chrm.class.MSM)
Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
summary(log.SC_chrm.class.MOLF)
```

```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
log.reg.chrm.class_all.strains
WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
```

```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
#1CO
log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_1CO.poly)
log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_1CO.Musc)
#double check these directions
log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_1CO.Mol)
#2CO
log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_2CO.poly)
log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_2CO.Musc)
log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_2CO.Mol)
#3CO
log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_3CO.poly)
log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Musc)
#there's no rec group for this table
log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Mol)
#reverse logistic regression (single bivalent level)
#how well does SC length predict 1 or 2 COs?
Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3) 
#need to switch the co counts to 1 - 0
Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
              data=Bivalent_1o2,  family=binomial(link="logit"))
summary(log.SC_male_chrm.class)
#not sure if this is the equivilent of running each strain seperatly
#the strain effect is probably also pulling in the different proportion of chrm class
#useing the chrm.length*strain gives fewer significant factors. 
#i'd be more convinced by running all those log regression plots for each strain
#come back later
```

<These are the plots of the logistic regression.>

```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
```

```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
#COMPARISONS ACROSS ALL CLASSES
#MOUSE AVE, all
Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
#mouse.av 1CO
Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
#MOUSE.av 2CO
Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
#single biv level
#male.biv?
#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])

#Rec.group_BivData_F1post_Ttest$p.value 
```

```{r log.reg_SC.length, echo=FALSE, include=FALSE}
#REMOVE THIS COMPARES FROM FULL BIV DATA
#male_mouse.av
#Male.poly.Mouse.Table_BivData_4MM
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
#male poly shouldn't be here
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
#All chrm
log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av)
#considering seperating out by hand foci counts
#1CO
log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_1CO)
#2CO
log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_2CO)
#3CO
log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_mouse.av_3CO)
# bivalent level log regression
log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV)
#1CO  (BIVDATA)
log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,

                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_1CO)
#2CO
log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_2CO)
#3CO
log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
#summary(log.reg_SC.length_BIV_3CO)
```

```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
#REMOVE THIS, IT uses total bivalent data
#rerun the models for datasets -- deviding by chrm class and rec group
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
#calq_mouse av
mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                        n_cells = length(unique(fileName) ),
                        mean.SC.total = mean(chromosomeLength),
                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

                         )
mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
#calq_mouse av
mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
```

```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )
male_mouse.av <- add_sex(male_mouse.av)
male_mouse.av <- add_strain(male_mouse.av)
male_mouse.av <- add_subsp(male_mouse.av)
male_mouse.av <- add_species(male_mouse.av)
fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus") 
fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
table( droplevels(fomo.male_mouse.av$strain) )
#there are cast strains too
Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
musc.male.plot
# + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +
#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +  
#make the Mol and
Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
Mol.male.plot
#Dom version 
Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
Dom.male.plot
```

```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
## run the tests to check following expectations
#1.Dom strains not different from ech other
#2. PWD and SKIVE are different from the other strains
#3.MSM and MOLF are different from each other
#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]

male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
#the subsp labels get messed up, unless they are designated chr first
Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
#total sc sig, 1CO sig, 2CO sig
Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
#0.000369 ***
Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
#0.000816 ***
Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
#0.0148 *
show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
#making plots above show that mouse average SC lengths are lower could there be microscope issue?
#make sure to remove some of the juvi mice!! from the AP_mouse table
show.Dom.male.CO <- ggplot(
male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
#P not sig for CO mean
#seems like there are more mice in the MLH1 data set than bivData
#
#WSB all dissected and imaged in '15
#G imaged 15 - 16
#LEW imaged 15 to 17
ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "", xlab = "")
show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "", xlab = "")
ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "1CO SC Length", xlab = "")
ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "2CO SC Length", xlab = "")
```

```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}
#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
#G-WSB is sig, WSB-LEW are slightly sig
Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
#PWD and MSM are most diff from SKIVE?
Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
#the pooled SC value tests are sig
#the mol comparisons are different due to adj p values
Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
#G-WSB is sig, WSB-LEW are slightly sig
Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
#no sig (only 3) -- does include CZECCH
#non-sig, with mol pooled with musc
Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
#non of these sig with the mulitple adjust
Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
#pairwise.t.test(Ozone, Month, p.adj = "bonf")
```

# brief transition, rec landscape

the lack of predictive power of SC length to predict CO number indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.

# Q2.1CO rec landscape evolution is decoupled from gwRR evolution

<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents 
Main results:  
1. WSB has the most terminal 1CO rec landscape and MOLF has the most central.
2. Evidence for evolution of the 1CO rec landscape, but it is decoupled from the evolution in gwRR we observe in figure 1
2.5  The normalized 1CO position is not a good metric for distinguishing strains with rapid gwRR evolution
-->
<!-- set up chunks here -->
<!-- plots /figures below-->

```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
#Curated_BivData_male
#boxplots of nrm F1
male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)

Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos All bivalents")
Q2.show.nrmF1
nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
```

```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
#make the same plots as above but with the long biv data set
male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
#mouse.av.long.biv.table
Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Nbiv  = length(Obj.ID),
                       mean.1CO_PER = mean(PER_Foci_1)
                      # mean_CO = mean()
                       )
Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
```

<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->

```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
#High = MSM, PWD, SKIVE
Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
#p-value = 1
#high = PWD and MSM
Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
#p-value = 1
```

```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
#Musc
Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,

            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
# p-value = 0.4
#Mol
Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,

              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
# p-value = 0.07
# exclusive rec group not sig
```

```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
#these don't have mean.first.foci
#short -- real.long.bivData -- REAL.short_mouse_table
Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,

      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
#p-value = 0.3
Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
# p-value = 0.1
#remove SKIVE??
```

```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
#real.long.bivData
#single biv level
Rec.group_LONG.BivData_F1post_Ttest <- t.test(
  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
#p-value = 0.5 
Rec.group_BivData_F1post_Ttest <- t.test(
  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
#p-value = 0.2
#
#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
#p-value = 0.2
#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
#p-value = 0.4
```

The previous section found that the single crossover landscape is significantly different between males and females. In this section we focus on differences between the high and the low rec male strains
In order to test if if there are distinguishing meterics for these two groups. **The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**

T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`). The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`). While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`). This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.

<long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`> 
<!-- GLMs because there's no difference 
toal data set, short and long biv. There's only a difference between MSM and MOKF
test ppredictions, Dom = NULL,  -->

```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
#with.chrm.legth  ?
Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
#M2
Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
```

<!--
In the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower. .. the chromosome length has the most significant effect, LEW by chrm effect are significant
In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects -->

<!-- Just hold these for reference
**For the M2 (subsp)**
There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).
LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).  
But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).  
and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)  
**For M3 (all strains)**
strain G (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][2]`).
strain Lew (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][3]`).
strain PWD ( p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][4]`).
slight strain MSM  (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][5]`).
very sig strain MOLF (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][6]`).
strain SKIVE (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][7]`).
strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
-->

```{r Q2.1COpos.show.MOL, echo=FALSE}
#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
#+facet_wrap(~mouse, scales = "free")
single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
#+scale_color_manual(values=colors_of_strains)
single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
#CHRM.length is very significant-- some of the same subsp and strains are sig     
```

given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models. Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation

<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>

<!--
**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$ -->

```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
#remove this cell, 
Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
#making new models 
Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
#summary(chrm.1CO.pos.chrmLength_strain)
```

<!-- hide these linear model results for now
for the first model, 
chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`  
subsp MOL `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][4]`  
the mol obj above isn't showing the reight value. 
strain G   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][5]`  
strain LEW   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][6]`  
Strain MSM `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][8]`  
chrm * LEW `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][14]`  
chrm *MSM  `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][16]`
M2
-chromosome length, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][2]`  
-strain G, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][3]`  
-strain LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][4]`  
-strain MOLF, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][7]`  
-strain SKIVE, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][8]`  
-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]` -->

```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
##long biv
male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
#table( droplevels( male.long.biv$strain)  )
mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,

                        n_obs = length(Obj.ID),
                         mean_F1 = mean(Foci1, na.rm=TRUE),
                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),

                        mean_SC = mean(chromosomeLength),
                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
                                               )
#add subsp
mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
#add rec group
mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
```

```{r Q2.1COpos.glm, echo=FALSE, eval=FALSE}
male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO,  family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm)
#Musc
Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
              family=binomial(link="logit"))
#summary(Log.Q2.F1_PER_Musc)
#Mol
Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
              family=binomial(link="logit"))
#summary(Log.glmQ2_M2.1COnrm_Mol)
```

# Q2 Evolution of interference is associated with genome wide recombination rate evolution

```{r Q2.IFD.long.short.setup, echo=FALSE}
#long BivData
male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
#show
Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
```

<!-- set up chunks here-->

```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
#subset
male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none", axis.text.x = element_blank())+ylim(c(0,1.1))

```

```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
#t.test for IFD
#mouse.av - the mouse average IFD is from just 2CO
t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )

t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
```

```{r Q2.t.test_IFD,echo=FALSE, eval=FALSE}
#already made
#subset
male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
#t.tests for the Rec group
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
#Curated_BivData <- add_IFD(Curated_BivData)
#THIS SHOULD ALREADY BE ADDED
#single biv level
Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])

Rec.group.BivData_IFD.PER_Ttest$p.value
#p-value < 2.2e-16
#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )

Rec.group_IFD.PER_Ttest$p.value 
#p-value = 6.872e-07
```

<write some kind of intro here>motivation /

A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation. ()

how crossover interference affects the evolution of genome wide recombination rate -- is

prediction for the interference pattern in the high rec strains is that the high rec strains will have weaker interference (ie. less space between foci on the same bivalent, which allows more foci to be placed. GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.

The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^). The pattern is confirmed with t.test comparing groups within the same subepcies .. add these ttests and p values.

<glms> The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.

**Linear model M1 ** $$mouse \ average \ IFD^PER^ ~=~ subsp * strain + \varepsilon $$
**Linear model M2 ** $$mouse \ average \ IFD^PER^ ~=~ strain + \varepsilon $$
```{r Q2.IFD.glm.test, echo=FALSE}
#REPLACE Male.poly with ..
#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
#M1.ABS
Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
#summary(Q2.M1glm_IFD.raw) 
#M1.PER
Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
#summary(Q2.M1glm_IFD.PER) 
#M2.raw
Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
#summary(Q2.M2glm_IFD.raw)
Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
#  MSM strain effect for 2CO PER not significant
#within subspecies
#Dom
Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q2.glm.M2_IFD_Dom)
#Musc
Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q2.glm.M2_IFD_Musc)
#Mol
Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q2.glm.M2_IFD_Mol)
#Dom
Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q2.glm.M2_IFD.PER_Dom) 
#Musc
Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q2.glm.M2_IFD.PER_Musc) 
#Mol
Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q2.glm.M2_IFD.PER_Mol)
```

Within the models, PWD and SKIVE seem to have higher support of longer IFD^PER^ compared to MSM
MSM does have longer IFD^PER^ compared to MOLF (and dom strains)

In M1, the model build on the mouse averages of IFD^PER^ metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]`  for PWD, MSM and SKIVE respectively). In M2, the mouse average for normalzied IFD^PER^ is modeled with only strain effects. Only PWD and SKIVE IFD^PER^ had significant strain effects (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]` for PWD and SKIVE respectively).

Additionally when M2 models (for just strain) was run, they confirmed the within subspecies predictions informed by the variation in mean MLH1 foci number. (IFD is the only single bivalent based trait wcich supported these predictions)

no significant strain effects across the _domesticus_ strains (glm; p = `r`).  
_musculus^PWD^_ and _musculus^SKIVE^_ have significantly longer IFD compared to the other strains (glm p= ?, Table). And _molossinus^MSM^_ has significantly longer IFD^PER^ compared to _molossinus^MOLF^_.

-when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF-->

<!-- we ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length. 
**model BivLen.M1 ** $$Single bivalent\ IFD ~=~ subsp * strain * chromosomeLength + \varepsilon $$
**model BivLen.M2 ** $$Single bivalent\ IFD ~=~ strain * chromosomeLength + \varepsilon $$
```{r Q2.IFD.glm.chrmLength, echo=FALSE}
#repeating chrm models with chromsome length
#FOR BIVALENT LEVEL MODELS
Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
#M.Len.ABS
Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
#summary(Q2.BIV.LEN.glm_IFD.raw)
#M.Len.PER
Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
#summary(Q2.BIV.LEN.glm_IFD.PER)
#M.Len.raw
Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
##M.Len PER
Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
#summary(Q2.BIV.LEN.M2glm_IFD.PER)
#  MSM strain effect for 2CO PER not significant
```
For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs). Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^. The Mol\*chrmLength is negative 
<PWD, MSM, MOLF, SKIVE are the only significant strain effects.
- all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF). All of the high rec strains and MOLF have significantly longer IFD^PER^. There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
-->
```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
#calculate the rates of 'close foci' (<30%)
#examine the changes in the distribution closer.
#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
                         c("mouse"), summarise,
                        N.obs  = length(Obj.ID),
                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
                        var.IFD = var(IFD1_PER, na.rm=TRUE),

                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
                      n.close.foci = length(which(IFD1_PER <= .3)),

                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),

                      rate.close.foci = n.close.foci / N.obs
)
close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
                         c("strain"), summarise,
                        N.obs  = length(Obj.ID),
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
                        var.IFD = var(IFD1_PER),

                      inverse.close.foci = length(which(IFD1_PER > .3)),  
                      n.close.foci = length(which(IFD1_PER < .3)),
                      rate.close.foci = n.close.foci / N.obs,

                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )

)
#the

show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
#add rec.group
close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
#POOL all short and 
#chi square test
high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
#p-value = 0.03, slight dif tween 3 high rec strains
#don't need to run chi square for this pattern, since it's enough to 
```
```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
#Log regression all show significant prediction
#IFD.PER
log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.PER)
#MUSC
log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.PER.Musc)
#Mol
log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.PER.Mol)
#SINGLE BIVALENT LEVEL
Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
#summary(Log.Q2.IFD.per_bivdata)
#Musc
Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
#summary(Log.Q2.IFD.per_bivdata.Musc)
#Mol
Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
#summary(Log.Q2.IFD.per_bivdata.Mol)
```
```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
#IFD.raw
log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.raw)
#MUSC
log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.raw.Musc)
#Mol
log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.reg.Q2.IFD.raw.Mol)
#SINGLE BIV LEVELs
male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
#summary(Log.Q2.IFD.raw.PER_bivdata)
#Musc
Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
#Mol
Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
```
```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
#Full dataset
male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
#single bivalent
Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
#summary(Q2.glm_long.biv_M2.IFD.PER)
#The PER.IFDs don't have significant strain effects
```

Given the differences in chromosomes proportions (Figure X), one might expect that --- the rec landscapes for 2CO bivalents in PWD and MSM would be more variable given that a larger proportion of chromsomes belong to this class (60% to 50%) (there are more opportunities for chromosome specific effect to affect the landscape). Instead the increased variation in the 2CO rec landscape is seen in the lower rec strains (which have on average 30% 2COs. However, we determine that the lower mean IFD^PER^ in these strains is driven by an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC in low rec strains. The rate of these close foci range from 8.2% ( _domesticus^G^_ ) to 16% ( _musculus^KAZ^_ ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_ and _musculus^PWD^_ respectively). <no chi sqaure test see above>

 **(unlike total SC length and 1CO pattern, IFD metreics are a significant predictors of rapid genome wide recombination rate evolution. However, they go agaisnt a standard perdiction for the relationship between the chromosome level recombination landscape and genome wide recombiantion rate.**

**the difference in interference measures is mainly driven by close foci in the 2COs of low rec strains**

<!-- 
Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents. But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
-->

<Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plots excluded for space.

Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correctly mapped)>

<!-- final summary, BE brief-->

BE brief,

final summary, highlight the themes of 1. more variance in female traits (MLH1 foci per cell, rec landscape / interference)

while the male traits -- more focused (less variance in MLH1 counts per cell, shorter chromatin comparction?  bias position, stronger interference)

(should I measure the var of all traits in Q1)

2. rapid evolution of genome wide rec rates can fit into this -

- evolution of more COs requires more DSBs (evidence points to a earlier step which is likely the chromatin compaction (amount of axis))

- the strongest single bivalent predictor/correlate of rapid gwRR for males is interferece strength in 2COs

AND somewhat puzzeling result of (having stronger interference), meaning there is a greater proportion of a bivalent that seperates two foci  (2 foci push to the ends of chrms (maybe decreased centromere supression))

# Blank1

<!-- unused chunks -->

```{r show.long.SC, echo=FALSE, eval=FALSE}
female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
```

<ABOVE plots show the SC lengths for the 'long SC data set'. They are supposed to be the longest 4-5 SC from cells where I could get good measures. These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position). Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>

<HAND measure METRICS, below chunk goes with the above chunk>
```{r long.biv.plots, echo=FALSE, eval=FALSE}
show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  scale_color_manual(values=colors_of_strains)+
  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
show.the.cells_females_dom
#PWD seems to have the biggest variance across cells
almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
  scale_color_manual(values=colors_of_strains)+
  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
almost_cells_DOM_f
#### Musc female
show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+

  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
### Dom Male
show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+

  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
#MUSC male
show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+

   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+

  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+

#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+

#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+

#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

 # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
#show.the.cells_male
#almost_whole_cell_BivData
#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
```

<All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data. It shows the amount of within cell variance across strains. There is a bit of variance across the SC length distributions in the PWD females.
This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures. The full data set is `r length(Curated_BivData$Obj.ID)`. This is the breakdown of bivalent observations by category for this long dataset are

**- Try to merge this DF with the whole.cell manual measures.**  
- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.  
in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.>

```{r SC.distb_manual_Auto, echo=FALSE, eval=FALSE}
#calq the mouse means
manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
                        median_SC = median(SC.length, na.rm = TRUE),
                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)

)
manual_summary_stats <- add_mouse(manual_summary_stats)
manual_summary_stats <- add_strain(manual_summary_stats)
manual_summary_stats <- add_sex(manual_summary_stats)
manual_summary_stats <- add_subsp(manual_summary_stats)
###
#calq the mouse means
Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        median_SC = median(chromosomeLength),
                        Qar.4 = quantile(chromosomeLength)[[4]],

                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)

)
Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
#colors_of_strains
#scale_color_manual(values=colors_of_strains)
#un-order these factors
```

<!-- Whole Cell Manual Comparison 
```{r XX.adj_manual.whole.cell, echo=FALSE, eval=FALSE}
#load and examine the the whole cell hand measures here
female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
data.frame(table(droplevels(female.whole.cell$category)))
# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
#missing some strains, most have 3 cells (min). MSM has 5 full cells
#remove the incomplete LEW female cell
female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
#calq mean SC lengths by rank
mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,

                       mean.sc =   mean(SC.length, na.rm = TRUE),
                         nbiv = length(Obj.ID),
                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
                         se.SC_length   = sd.SC_length / sqrt(nbiv)
)
         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =

top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
                    position=position_dodge(width=0.7), 
                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
show.SC_by.rank_top
bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
                    position=position_dodge(width=0.7), 
                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
#CALQ PERCENT TOTAL SC
CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,

                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
                         nbiv = length(Obj.ID),
                       #rank.by.per

                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
)

#MELT ABOVE
MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
              #leave out the Foci
               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],

            id=c( "fileName","total.sc.cell","nbiv"))
MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
show_long.sc.proportions
#original plot code 
#ggplot(data = df1, aes(x, y, color = Group)) +
#    geom_pointrange(aes(ymin = lb, ymax = ub), 
#                    position=position_jitter(width=0.5), 
#                    linetype='dotted') +
#    theme_bw()
```
The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)). 
The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank. For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
The other figure shows of each single bivalent contributes to the total SC area. Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell. on average, each of the top longest bivalents make up ~10% of the cell's total SC area. So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX, 
- Is the difference between cell averages for males and females less that 10%?
- also interesting, the pwd and msm don't have longer SC, compared to other strains.
<!-- Automated BivData Comparison -->

```{r XX.Automated.biv.stuff, echo=FALSE}
#make automated female data to compare to
female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
#table(female.Bivdata_comp$strain)
#female.Bivdata_comp
#1. calq the rate of chrm segmentation in female categories in automateted data
box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,

                   Nmice = length(unique(mouse)),
                   mx.box.num = max(boxNumber), # need to run this at the cell level
                    N.biv  = length(Obj.ID),
                   pass.biv = N.biv / mx.box.num
)
box.rates <- add_mouse(box.rates)
box.rates <- add_strain(box.rates)
box.rates <- add_category(box.rates)
cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
cell.pass.rates
category.box.rates <- ddply(box.rates, c("category"), summarise,

                   mean.pass.rate = mean(pass.biv)
)
#category.box.rates
#2. pull the 3rd rank biv's calculate the SC means
#calq means by rank (focus on the top 5)
```

For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.

The table above shows the number of bivalents from the same strains as in the manual whole cell data. The plot shows the bivalent passing rate across all of the individual cells from this female data set. For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).

(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level. (it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)

strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*

** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX. But I am not using 'whole cell' summaries to compare female and males.

**What is the effect of an extra XX-autosome on single bivalent means?**

use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents. Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females. Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly. Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.  The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.

# Blank2
<!-- Note on Heterochiasmy Definition
<!-- add outline for new XX adjust section 
adjust the colors for the plot below (standardize)
-->

I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison. In females the XX bivalent is indistinguishable from the autosomes. To the meiotic recombination machinery, it is an autosome and has a similar REC landscape. Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count). (I note if the and Y are paired, which they are at a high rate). The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y). To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.

1. Compile full-cell data from females (all 20 bivalents measured)
2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
3. check how variable the REC is across the 1st,2nd,4th, and 5th are.

According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).

(Put the XX adjustment section here)

There is now MOLF, which has female biased hetC
3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
1 of the musc strains has female biased heterochiasmy, KAZ.

```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
#REMOVE This BLOCK?
#### Distributions by mouse
#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
#remove this block since it has it's own src/script
### to integrate 'pass'/fail
### this should be in setup data
#mice that don't have female data will break the chain-loop
#MOLF female, HMI, CZECH female, CAROLI female, Tom male
#instead of adding fake data, remove the categories from list
#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
#this loop makes a quant_status col for a list of mice
for(i in 1:length(MLH1_data$mouse)){
  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
}
df_list <- split(MLH1_data, (MLH1_data$category))
#remove sections which are in category but have no data
#this is where you remove the non-data data!
df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
#7 PERC, 12 MOLF female, 
#
#df list, is a list of df's which were split() from BIG df by category
df_list 
o=0
i=0
MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
for(i in 1:(length(df_list))){
 # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
                      mouse_nMLH1 = mean(adj_nMLH1.foci)
  )
# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
  for(o in 1:length(df_list[[i]]$Original.Name)){
    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
        }
}
plot_list = list()
for (h in 1:length(df_list)) {
  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
    facet_wrap(~ mouse, drop = TRUE) + 
    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
  #make an empty ggplot? for missing data

  plot_list[[h]] =   pf
}
#orig (I think this is the same as above)
#plot_list = list()
#for (h in 1:length(df_list)) {
#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +

 #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
    #theme for indiviual mice, 
#theme(plot.background = element_rect(fill = "green"))
#  plot_list[[h]] =   pf
#}
#can't change the alpha/transparency of the red rec
#aes must be same length ---   #geom_point(stat = "identity") +
# the last one is correct....?
#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
```

```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
geom_hline(aes(yintercept = med, group = gr), colour = 'red')
#no plot is made that goes in the list -- so it can't be shown
for (gg in 1:length(plot_list) ){
  show(plot_list[gg])
 # invisible(lapply(obj, function(x) plot(x,main="some plot")))
}
#this stops at MSM
```

The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.

Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse. The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.

```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
#add titles
#use multiplot
#remove X axis
multiplot(kuku,bubu, cols=2)
```

```{r F1.by.F2, eval=FALSE}
#try remaking the plot Megan suggested
# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
#isolate 2COs
#facet by sex and subsp
F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
```

<!-- Deleted notes -->

```{r variance.hist, echo=FALSE, eval=FALSE}
#figures of histograms, 
#make series for total dataset?
Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +
  ggtitle("Histograms of within mouse variance, all mice")
Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
Var.Hist.total.mice
#DF.HetC.MixedModel.HQ
Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  + 
  ggtitle("Histograms of within mouse variance, 10 cell min")
Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
Q12.var.hist.plot
#some LEW female still have v high within mouse variance (2 mice, )
```

  -->

# References

### 2020-02-27 doc/Results.rmd
making notes for the Q1 IFD sex difference patterns

@@
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document: default
   html_document:
@@
     toc_float: yes
+  pdf_document: default
+  word_document:
+    toc: yes
+    toc_depth: '5'
 bibliography: refs_Main_Report.bib
@@
 first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-first.IFD.plot_ABS
 
 first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-first.IFD.plot_PER
+
 #+theme(legend.position="none")
@@
 When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
-MSM   `r `
-SKIVE  `r `
-PWD  `r `
-MSM  `r `
-LEW  `r `
-G  `r `
-WSB  `r `
+MSM   `r Q1.IFD.PER_t.test_MSM$p.value`
+SKIVE  `r Q1.IFD.PER_t.test_SKIVE$p.value`
+PWD  `r Q1.IFD.PER_t.test_PWD$p.value`
+MSM  `r Q1.IFD.PER_t.test_MSM$p.value`
+LEW  `r Q1.IFD.PER_t.test_LEW$p.value`
+G  `r Q1.IFD.PER_t.test_G$p.value`
+WSB  `r Q1.IFD.PER_t.test_WSB$p.value`
 
@@
 
-```{r Q1.show.IFD.things, echo=FALSE}
+```{r Q1.show.IFD.things, echo=FALSE, eval=FALSE}
 nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
@@
 
-```
-
-see the reference plots within this dir, --
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+```
+
+strains where males have higher upper threshold,
+
+G - kinda, little weak
+
+WSB - higher upper threshold (female's have many below the 25% lower threshold, males have a few below 25% threshold)
+
+LEW -- upper threshold looks very similar, much fewer low threshold obs in males,
+
+PWD -- female lower threshold extends muchlower .25 (like .15?), male lower threshold very different (higher like .3) the male upper threshold is above .75
+
+MSM -- female thresdholds, upper set around .75(to .65) lower at mostly .25, some cells(mice) have more .1 to .2 IFD observatsions.  Male upper threshold it at .75 while lower is set to .3 or .4.
+
+SKIVE -- female upper threshold is capped at .75, lower threshold is between .25 to .3 but low rate of outliers below .25.
+Male lower threshold is moved up, .4 or .35 and upper threshold moved up aprx .8
+
+KAZ  -- female upper threshold around .75, lower threshold around .25 the distribution is quite uniform and spread out.
+Male pattern; upper threshold slightly higher than .75 (.8) and lower threshold seems to extend beyond .25 -- but maybe low sampling contributes to the spread?
+(the spread in IFD values seems much larger for KAZ males compared to males from other strains.)
 
@@
 
-multiplot(kuku,bubu, cols=2)
+#multiplot(kuku,bubu, cols=2)
 ```

### 2020-02-27 Results_library.Rmd
cleaning up and more notes for 2CO.IFD section

@@
 library(raster)
-library(lme4)
+library(lme4)F
 library(nlme)
@@
 options(digits=2)
-#setwd("~.")#switch over to here()
+setwd("~.")#switch over to here()
 
 #MLH1_data_setup_2.24.20.RData
-load(file = here( "MLH1/MLH1_data_setup_2.24.20.RData") )
-#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
+#load(file = here( "MLH1/MLH1_data_setup_2.24.20.RData") )
+
+#meta.data = read.csv(here("data", "/clean.Meta.Data.txt"),sep = "\t", header=TRUE)
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
 

### 2020-02-27 doc/Results.rmd
cleaning up and more notes for 2CO.IFD section

@@
 # Q1 Sex Differences in SC-AE Lengths
+
 <!-- breif background on longer female SC  (mouse, human) Gruhn-->
@@
 
-## Adjusting for XX
+Adjusting for XX
 
@@
 - rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX>
-
-# short transition
 
@@
 
-Crossover interference is one of the major determinates -- of the CO landscape. We use X number of 2CO bivalents, to test if there are diffferences between males and females in interference. 
-The mean interfocal distance (IFD) for raw and normalized values was calculated from X female and X male mice.
-
-The raw IFD measures -- are usuful for thinking of crossover interference as a mechnical force mediated through the SC. Previous measures found no sex difference in IFD in B6 mice (de Boer et al )
-
-In the abosolute plots, you can start to see clustering/ enrichment of IFD values between 40 and 80 pixels for PWD and SKIVE, 
-(and maybe slight clustering for MSM).
+Crossover interference is one of the major determinates -- of the CO landscape. We isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
+`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males to test discribe sex differences in crossover interference.
+Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
+
+We examined both raw IFD^raw^ and normalized normalized by SC length (IFD^norm^)
+
+The mean IFD^raw^ measures of crossover interference as a mechnical force mediated through the SC.
+--- we compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
+Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al ).
+
+While the comparisons of the IFD^PER^ metrics can reveal more general recpmbination landscape patterns while controling for the underlying differences in chromatin compaction / SC length.
+(Also in a way this metric is closer to measures of interference from genetic (linkage) maps since they are not quantified / measured in phyiscal space either).
 
 <!--t.tests -->
-
-t.tests are performed on the normalized metrics for establish if there are sex differences in the IFD meterics (raw and normalized).
 
@@
 
-The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
-BUT WHEN G IS REMOVE, the pvalue is no longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
-<look closer at G>
-
-The normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`) between sexes, indicating there a
-
-When the t.tests were performed within each strain, all but KAZ show (p-value = 0.3), significant sex differences in the normalized IFD measures
-MSM   `r Q1.IFD.PER_t.test_MSM$p.value`
-SKIVE  `r Q1.IFD.PER_t.test_SKIVE$p.value`
-PWD  `r Q1.IFD.PER_t.test_PWD$p.value`
-MSM  `r Q1.IFD.PER_t.test_MSM$p.value`
-LEW  `r Q1.IFD.PER_t.test_LEW$p.value`
-G  `r Q1.IFD.PER_t.test_G$p.value`
-WSB  `r Q1.IFD.PER_t.test_WSB$p.value`
-
-This indicates a general pattern of two foci are seperated by more area (hance have stronger interference) in males, this pattern is also for genetic maps.
-
-<!-- short and long -->
-
-```{r Q1.IFD.long.short.setup, echo=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-```{r Q1.show.IFD.things, echo=FALSE, eval=FALSE}
+The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`)
+-- However _domesticus^G^_ might be an outlier (by having a larger than average degree of sexual dirmohpsim ).
+When the _domesticus^G^_ observations are removed, the difference is not longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
+ This indicates there is no general pattern of sex differences in crossover interference measured in phyiscal (SC) units.
+(confirming deBoer and other models of crossover interference acting through phyical/mechanical force).
+
+On the other hand, the difference in mean IFD^PER^ are highly significant between sexes.
+The mean IFD^PER^ are significanly long in size of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
+Only _musculus^KAZ^_ did not have a difference in mean IFD^PER^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
+These results indicate a general pattern of two foci are seperated by more area (hence stronger interference) in males.
+<which is supported by the pattern in genetics maps>
+
+<Full dataset (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`).>
+
+<!--  finer distribution patterns -->
+
+```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE}
 nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
@@
 
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
-```
-
-strains where males have higher upper threshold,
-
-G - kinda, little weak
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+
+```
+
+We examined the 2CO observations  // the distributions of 2CO IFD^PER^ to better discribe how the means / distributions of this metric are different between the sexes.
+
+We note that in females there are/ may be signals of upper and lower thresdholds for IFD^PER^.
+
+**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference
+
+the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs.
+There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
+off. Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?
+
+The general pattern of this sex difference is that the IFD^PER^ distribution is pushed higher   (females have mean of .5  IFD^PER^ where as males have a IFD^PER^ mean of ~.6?)
+
+The differences between the sexes is slightly different depending on if the males are in the high rec group.
+
+<INTRODUCED idea of constrain / thresholds>
+
+in the low rec strains (Dom + KAZ)...
+(both males and females IFD distributions have lower and upper thresholds around 25% and 75%) respectively -- however the weight of observations are slightly shifted across the sexes.
+There's an enrichment of lower IFD^PER^ in females relative to males, while the opposite is true for the higher threshold -- where males have an enrichment of higher/long IFD^PER^ compared to females.
+
+In the three high rec strains -- the distributions have a more striking difference of the lower threshold being much higher in males, aproximately 30% to 40% compared to 25% in females.
+Also while the female upper threshold is capped at 75%, in males (some strains have higher upper thresholds -- and it's more likely to observe IFD^PER^ above the threshold.
+
+<!--
+G - the upper and lower thresholds of females look pretty similar, the lower is closer to .2 than .25 -- female has more lower IFDs
 
@@
 
-LEW -- upper threshold looks very similar, much fewer low threshold obs in males,
-
-PWD -- female lower threshold extends muchlower .25 (like .15?), male lower threshold very different (higher like .3) the male upper threshold is above .75
-
-MSM -- female thresdholds, upper set around .75(to .65) lower at mostly .25, some cells(mice) have more .1 to .2 IFD observatsions.  Male upper threshold it at .75 while lower is set to .3 or .4.
-
-SKIVE -- female upper threshold is capped at .75, lower threshold is between .25 to .3 but low rate of outliers below .25.
-Male lower threshold is moved up, .4 or .35 and upper threshold moved up aprx .8
+LEW -- upper threshold looks very similar (.75) in males and females, much fewer low threshold obs in males,
 
@@
 
-this section -- all strains but KAZ, males have a higher mean normalized IFDs
-
-the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs
-
-The KAZ strain is the only strain without significant p value for the normalized IFDs, due to KAZ males having a higher rate (enrichment of short normalized IFDs.
-
-There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
-off.
-Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?
-
-<The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes?
-they were supposed to be 1CO -- but somehow a second CO formed anyways ..
-and very close by
-
-Where as in PWD, SKIVE and MSM there ARE NO 2CO bivalents which have such short IFD
-WHAT espicially werid about this --
-
-whats the expectations following the confonding proportions? 
-high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.>
-
-**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference
-
-<remove the chrm size effects -- or check out model with chromosome length>
-
-to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
-Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)
-
-Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
-These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).
-
-Considering that the short-long bivData is much smaller than the full data set
+PWD -- female lower threshold extends much lower .25 (like .15?),  upper threshold capped at .75
+Male lower threshold very different (higher like .3) the male upper threshold is above .75
+
+MSM -- female thresdholds, upper set around .75(to .65) lower at mostly .25, some cells(mice) have more .1 to .2 IFD observatsions.  
+Male upper threshold it at .75 while lower is set to .3 or .4.
+
+SKIVE -- female upper threshold is capped at .75, lower threshold is between .25 to .3 but low rate of outliers below .25.
+Male lower threshold is moved up, .4 or .35 and upper threshold moved up aprx .8
+-->
 
@@
 
-### Mixed Model Tests, Fixed Effects
-
-Since the most significant t.test were significantly different across sexes for the NORMALIZED effects -- we follow up with the model framework to --check out the potential evolutionary patterns--
-
-We applied the same frame work of models, Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.
+Since the most significant t.test were significantly different across sexes for IFD^PER^ and we follow up with the model framework to.
 
@@
 
-```{r Q1.IFD.M1, echo=FALSE, message=FALSE}
+```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
 #get a list of the old mice to double check
@@
 #MIXED MODEL
-
 #https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
 
+#mouse.avs_4MM -- is also the bivbased mouse average table 
 #Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
@@
 
-#M1_IFD_2CO
+#M1_IFD_2CO_ABS
 lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
@@
 
-sex.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
 
@@
                interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-#don't understand why 2 rows?
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
-#anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)
-
-#these 2 below are LR tests
-#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
-#exactRLRT(lmer.IFD.PER_MM.M1)
-
-```
-
-```{r IFD_HetC_M2, echo=FALSE}
+
+```
+
+```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE}
+
+#mouse.avs_4MM -- is also the bivbased mouse average table 
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
 
 glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-
 summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)
+
+glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
+summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
 
@@
 glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-
 summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r IFD_HetC_M3, echo=FALSE}
+```
+
+```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
 
 glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-
 summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-#round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)
 
@@
 summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
+#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
 
 #write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-
-```
-
-For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.
-
- (IFD.PER)
-sex
-MM p = `r`
-
-subsp
-p = `r`
-
-the interaction effects were slightly significant for both raw and nrm.IFD.
-The most significant value was the sex effect for nrm.IFDs. 
-The the random strain effect was not significant for either model.
+```
+
+For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD.
+
+(IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`) subspecies, interaction are slightly significant.
+<ONLY REPORT FOR PER MEASURES>
+
+glm M2, 
+sex (glm; p= `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`) p = .0080 (this is different in the knited version!! arg)
+
+male*PWD p = 0.02 (glm; p =`r summry.glm.Q1_IFD.PER.M2$coefficients[,4][14]`)
+
+male * SKIVE p = 0.0058  (glm; p=`r summry.glm.Q1_IFD.PER.M2$coefficients[,4][15]`)
+
+for **glm M3**
+
+sex / male p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]` and male*SKIVE just slightly significant (p =  `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][14]`
 
@@
 
+<remove the chrm size effects -- or check out model with chromosome length
+to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
+Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)>
+
+Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
+These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).
+
 The above plots display the raw and normalized area which seperate two foci on the same chromosome.
@@
 
-# transition Q1 to Q2
-
-we tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.
+```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#real.short.bivData
+
+Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.long <- add_sex(Q1.IFD.long)
+Q1.IFD.long <- add_strain(Q1.IFD.long)
+Q1.IFD.long <- add_subsp(Q1.IFD.long)
+
+Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
+Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
+
+Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.short <- add_sex(Q1.IFD.short)
+Q1.IFD.short <- add_strain(Q1.IFD.short)
+Q1.IFD.short <- add_subsp(Q1.IFD.short)
+
+Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
+Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
+
+#t.test. mouse av
+Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+
+Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+#both the short and long norm IFD are NS
+
+#t.test Biv
+Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+
+Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+```
+
+<!-- transition Q1 to Q2 -->
+
+We tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.
 
@@
 ```
+
+<write some kind of intro here>motivation /
+
+A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation. ()
+
+how crossover interference affects the evolution of genome wide recombination rate -- is
+
+prediction for the interference pattern in the high rec strains is that the high rec strains will have weaker interference (ie.
+less space between foci on the same bivalent, which allows more foci to be placed.
+GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.
+
+The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The pattern is confirmed with t.test comparing groups within the same subepcies ..
+add these ttests and p values.
+
+<glms> The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
 

### 2020-02-27 doc/Results.rmd
more editing

@@
 
-# Genome wide recombiantion rate estimates for both sexes
+# Genome wide recombination rate estimates for both sexes
 
@@
 m.
-musculus_.
-and _M.
+musculus_ and _M.
 m.
@@
 The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) that the genome wide recombination rates averages are greater females compared to males (female biaded heterochiasmy) and ii) the degree of heterochiasmy (F:M rates) is generally low in these house mouse strains, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
+Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome wide recombination rates averages are greater females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (F:M rates) is generally low in these house mouse strains, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
 Three notable exceptions of male biased heterochiasmy are from the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
 
-Examining the mouse means of MLH1 foci per cell seperately for each sex reveal disctinct patterns of variation (Figure1 B and C).
-The female means can be discribed as more evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B).
-While in males, the strain specific means seperate more clearly into two groups of high genome wide recombination rates near 30 MLH1 foci per cell and low genome wide recombiantion rates, near 23 MLH1 foci per cell (Figure1 C).
+Examining the mouse means of MLH1 foci per cell separately for each sex reveal distinct patterns of variation (Figure1 B and C).
+The female means are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B).
+While in males, the strain specific means separate more clearly into two groups of high genome wide recombination rates near 30 MLH1 foci per cell and low genome wide recombination rates, near 23 MLH1 foci per cell (Figure1 C).
 
@@
 
-<!--(the figure is made from pooling MLH1 cells)
-the male 
-- the ranges of mouse means are displayed in Sup table X
-- How the means were calculated
--The male gwRR is adjusted for the PAR
--The means for each strains are report in Table 1
--Normality was assessed for the MLH1 counts (include in Supplement histograms)
--->
+<add a transition sentence, about the expectations under different evolutionary scnarios>
 
@@
 
-**M3. LInear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
 
@@
 
-We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework.
-All of the effects within the mixed model M1.
-were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
+We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
+All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
 
@@
  <!-- (random strain effect was meant to reflect sampling of gentic backgrounds/standing genetic variation for a subspecies) -- (significant pvalue indicates that the variance is not 0)>
- indicating, fixed effects -- random effect variance due to each strain nested within subspecies -->
+ indicating, fixed effects -- random effect variance due to each strain nested within subspecies
+ <The linear models, allow us to quantifying the significance in sepecific fixed effects and focus on how specific levels and interactions create these differences.
+ -->
 
@@
 
-<The linear models, allow us to quantifying the significance in sepecific fixed effects and focus on how specific levels and interactions create these differences.>
-
-The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombiantion rates.
+The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombination rates.
 The subspecies fixed effects tested in M2, had weak to no significance (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
@@
 
-In addition to significant strain effects, ( p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and _molossinus^MSM^_ strain p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
+In addition to significant strain effects for _molossinus^MSM^_, (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
 and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.
 
-Combined these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of sexual dimorphism /heterochiasmy in _domesticus^G^_.
-No significant sex effect showing that the sex effect on mean CO number is not uniform across the sample but has synergy with speciefic strains.
+Combined these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of sexual dimorphism/heterochiasmy in _domesticus^G^_.
+
+<The lack of a significant sex effect indicates that the evolution of mean MLH1 foci per cell is not uniform across sexes (but varies for specfic strains)>
 
@@
 
-A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` number of male and `r length(female_mouse.avs_for.analysis$mouse)` female mean MLH1 foci counts were used in each of these models.
-For the male specific data,
-
-_musculus^PWD^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
-(with effect sizes ranging from 5, 7, and 2 foci respectively).
+A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` male and `r length(female_mouse.avs_for.analysis$mouse)` female mean MLH1 foci per cell.
+There were three significant strain effects in the male specific data; _musculus^PWD^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` with effect sizes ranging from 5, 7, and 2 foci respectively.
 Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
 
-three of the strain effects of 13 strains had significant effects on mean MLH1 per cell, _musculus^PWD^_  (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` ).
-(with effect sizes ranging from 5, 7, and 2 foci respectively).
-
-While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`) the effect sizes have a smaller range of four to one, indicateing specific, but small strain effects for the mean number of MLH1 foci per cell.
+While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`), the effect sizes have a smaller range of (4 to 1), indicateing significant, but small strain specific effects on mean MLH1 foci per cell.
+<for this reason differences between females aren't studied further>
 
@@
 
-Variance of a trait can contain information pertinant to it's mode of evolution and mechanistically could reflect the strength of constraints acting on the trait.
-We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependant variable).
-<note about reduced power>
-<in some traits, like gene expression patterns, evolution of the mean -- also has an effect on the variance>
-
-<!-- 
-**M5. Mixed Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
- **M6. Linear Model** $$mouse\ variance\ in \ CO \ per cell ~=~ subsp * sex * strain + \varepsilon$$
-
- **M7. Linear Model** $$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$>
--->
-
 ```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
@@
 ```
-<!--
-These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
-var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
-cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
-5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
--->
-
-**the general pattern is that females have more variance (alsmot twice as much (Figure 1))**
-
+
+<Variance of a trait can contain information pertinent to its mode of evolution and mechanistically could reflect the strength of constraints acting on the trait.> We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
+
+The linear models support the general pattern of females having almost twice as much variance in MLH1 foci per cell compared to males (Figure 1)).
 For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`).
@@
 
-The sex specific effect linear models indicate that there is no significant difference in the amount of within mouse variance in males.
-While in females, one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`).
-However, in the model using the high quality cells, _musculus^PWD^_ has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.
-
-**the general pattern is that females have more variance (alsmot twice as much (Figure 1))**
+There is no significant difference in the amount of within mouse variance of MLH1 foci counts per cell in males according the the linear models.
+In females, the significance of strain effects was not consistent between the full data set and higher quality dataset _domesticus^LEW^_ (glm; p = 0) was a significant effect in the full model and _musculus^PWD^_ (glm; p =0.04) in the higher quality dataset.
+
+<one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`).
+However, in the model using the high quality cells, _musculus^PWD^_ has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.>
 
@@
 This limits the power for detecting significance differences across groups,
-
 the main/general pattern, is that females have significantly higher within mouse variance compared to males.
@@
 
-<!--
-$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
 <!--
@@
 <SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE>
-
-(combined the summary of the MLH1 distributions evolution in the mean and consistant pattern of sex differences)
-
 <The greater amount of within mouse variance in females (almost and, (males have no detecable difference across strains for within mouse variance)
@@
 
-3. -- the striking differences / patterns of variation across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories for male and female rates. 
-For female the pattern of Strain averages distributed around a species wide average - fits a model of stabilizing / relaxed / neutral evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic backgrounds / strains which fits a model of directional selection on genome wide recombination rate.
-
-# Evolution of Genome wide recombination rate associated with evolution of mean precursor number
-
-<transition, >
+The strikingpatterns of variation across male and females suggest that the genome wide recombination rates have distinct evolutionary trajectories.
+For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome wide recombination rate.
+
+# Evolution of Genome wide recombination rate associated with evolution of mean DSB number
 
@@
 
-In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in early prophase spermtaocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes form juvenille mice (12 to 18 days).
-A subset of strains were choosen to test for differences between the high and low recombination groups; _musculus^PWD^_, _molossinus^MSM^_, _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_ (Figure2).
-One mouse was sampled for each strain.
-
-For all strains there were significantly more foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the timing of meiosis.
-Confirming the prediction that evolution of the mean number of crossovers per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
-
-However the the differences in DMC1 foci were not significant for (later prophase) zygotene cells between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
-Indicating that this marker at early prophase is more precitive of the downstream crossover number differences.
-
-<explain the biological significance of the ratios!>
-
-after DSB formation, DSBs are repaired as either NCO or CO (CO designation step)
-The vast majority of DSBs are repaired as NCOs, with a minority (in most species) of approximately 10% are repaired as crossovers.
-
--DSBs are either repaired as crossovers or noncrossovers (NCO), with conversion tract lengths (30 to 100bp long). (this is also known as the CO repair deicison )
-
-thus the ratio of CO:DSB is in part an indicator of the proportion of DSBs which are designated into COs -- (and the others are repaired as NCO).
-
-The ratios, calculated for DMC1 means from both stages of prophase, are not significantly differet between the strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
-These results indicate that the targets of evolution which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).
+In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in prophase spermtaocytes.
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes from juvenile mice (12 to 18 days) of three low rec, ( _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_) and two high rec strains ( _musculus^PWD^_ and _molossinus^MSM^_).
+
+For all strains there were significantly more DMC1 foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the cell stages in prophase.
+The mean number of DMC1 foci per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
+However the differences in DMC1 foci were not significant for zygotene cells (later prophase) between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
+Indicating that this marker at early prophase is more predictive of the downstream crossover number differences.
+
+<explain the biological significance of the ratios!> After DSB formation, DSBs are repaired as either NCO or CO (CO designation step) with the vast majority being repaired as non-crossover NCO.
+Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated into COs.
+The ratios, calculated for DMC1 means from both stages of prophase, are not significantly different between the high and low strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+<These results indicate that the targets of evolution which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).>
 
@@
 ```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-
 #these are histogram plots for testing Normality shape -- move to supplemenatry section
-
 #- check that distributions are close to normal
@@
 
-DMC1.histogram
 #checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
 #bartlet's test, Levine's test
-
-#The counts look close to normal.
-#Potentially figure 2. Using a function to #display the p-vales for mean comparison
-
-```
-
-<!--
-interpertation notes, 
-the number of DMC1 across stages (L and Z stages) are driven by different equations
-Leptoene the number / rate of DMC1 is primarily driven by the substrate -- number of DNA loops -- empty sites are being filled.
-Zygotene the number / rate of DMC1 foci are more variable across cells because it is driven by the rate / success of homolog pairing  / if this varies across cells (via the timing) - the there might be a different total number across cells
-
-*the repair decision can happen at both stages and remove DMC1.
-formation of new DSB -> DMC1 foci happens at both stages, but the are triggered by different events (L: cell wide signal to start the homolog pairing and Z: unpaired bivalent signal)
-
-back to CO number -- variation in CO number is best predicted by metrics/indicators placed BEFORE DSB formation starts.
-
-expectations, predictions if this is the case -- more within mouse/group variation for DMC1 number in Z compared to L?
-OR
-
-variance in L stage for DMC1 per cell would be proportional to the variation in the 'packing ratio' across pachtyene cells
--Baier et al 2014 supports this model (Chmr axis / DNA loop model)
--->
+```
 
@@
 PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.MSM.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "MSM") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "MSM") )$nbiv )
 
@@
 We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per chromosomes (bivalent).
-Because our observations of MLH1 foci are made in pachtyene cells, our chromosome opbservations are fully synapsesed paired homologous chromosomes (i.e.
-4n copies of each chromsomes).
-Additionally we can distinguish the idenetiy of indiviual chromsomes (Chromsome 1 or Chromsome 2)
+Because our observations of MLH1 foci are made in pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
+Additionally we can distinguish the idenetiy of indiviual chromsomes (Chromosome 1 or Chromsome 2).
 for these reasons we use the term bivalent.
 
-Of the pooled dataset of `r total.biv.all` single bivalents, 96% of bivalents had either one or two crossovers (Figure X) with proportion of 1CO:2CO distinguishes the high and low rec males (Figure X).
-THis confirms the intuitive interpertation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents.
+Ninety six percent of the pooled dataset of single bivalents (n=`r total.biv.all`) have either one or two crossovers (Figure X).
 Across the female observations the proportions of chromosome classes are not notably different.
-
-<Most of the low rec strains range from `r min(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` _musculus^KAZ^_ to `r max(chrm.prop.male.low$n2CO / chrm.prop.male.low$nbiv)` ( _molossinus^MOLF^_ ).>
-
+The proportion of 1CO:2CO distinguishes the high and low rec strains (Figure X).
+This confirms the intuitive interpretation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents.
 In the high rec strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
 
-While the proportions are significantly different
-(p = `r`), PWD male and MSM male proportions are more similar to each other than strains from the same subspcies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-
-<SUMMARY the / TRANSITION to single bivalen measures>
+While the proportions are significantly different (chi-square-test; p = `r PWD.MSM.2CO.prop.test$p.value`),Males from _musculus^PWD^_ and _molossinus^MSM^_ have 2CO proportions more similar to each other than strains from the same subspecies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+
+<SUMMARY the TRANSITION to single bivalen measures>
 
 # Single Bivalent Level Results
-
-In order to deconstruct the cell wide MLH1 counts
-deconstruct we 
-look at the recombination landscape at the single bivalent level.
-
-From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent measures were left.
-
-We use these observations to look at the recombination landscapes, here refereing to the i) total length of the SC ii) the position of single crossovers along bivalents and iii) the spacing between multiple COs on the same bivalent.
-<the position is general, not genomic corrdinates>
 
@@
 
-<Table X illustrates the break down of single bivalent observations by category. Error compared to human measures and other details have been discribed in (Peterson_2019).>
-
-A potential downside of the image analysis algorithm we used is that not all bivalents per cell are isolated.
-The ranges of rates of isolating chromosomes per cell are `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to
-`r max(algorithm.stat.category$mean_pass.rate)` ) _musculus^KAZ^_ female, we assume that the isolation process is not biased.
-However, because there are hundreds of observations per category, we assume the set of single bivalents are equally representative of general patterns.
+<In order to deconstruct the cell wide MLH1 counts we look at the recombination landscape at the single bivalent level.>
+<FROM number of crossovers per chromsoomes TO rec lanscape across single chromsomes>
+<A big motivation was to quantifiy recombiantion landscape differences across single bivalents>
+
+The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures (Peterson 2019).
+A limitation of our image analysis algorithm we used is that not all bivalents per cell are isolated.
+The range isolation rates per cell in this data set is `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to
+`r max(algorithm.stat.category$mean_pass.rate)`  _musculus^KAZ^_ female.
+
+Due to overlap of chromosomes most cells do not have all chromsomes measured, but we assume that the isolation of chromsomes within cell images is not biased.
+From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
+After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent observations remained.
+Given the large number of single chromosome observations, we assume that each of the data sets are equally representative of general patterns.
 
@@
 <!-- breif background on longer female SC  (mouse, human) Gruhn-->
-
-(in many mammal species -- SC lengths are longer in females - yet most all of these observations come from female biased heterochiasmy species) -- Our data set provides an opportunity to test (how SC length and chromatin compaction relates to heterochiasmy)
-
-if the canonical female biased heterochiasmy pattern and long SC lengths are uncoupled in instances / our strain with male-biased heterochiasmy ( _musculus^PWD^_  and _molossinus^MSM^_).
-A simple model might predict if SC length is the strongest predictor for number of MLH1 foci er cell the high recombining males would have longer SC length proportional to differences in the number of MLH1 foci per cell.
 
@@
 
-To account of the obvious confounding effect of sex chromosome differences in females and males, we use a reduced bivalent data set made up of chromsomes (below the 2nd quartile for SC length of single cells) to make single bivalent groups which exclude the XX homologs from the female data.
-A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(Q1.real.short_mouse_table$ncells)` oocytes and spermatocytes.
-Even though this smaller data set has decreased power due to redcued sample size, it offers a mors comparable set of SC-AE meterics across the sexes.
-
-This 'short biv' data set has advantages in that it is closer to sampling across a similar set of chromosomes (ideally Chromosomes 15 to 19) and it excludes the XX in females, which is predicted to be the 3rd longest based on the physical (Mb) length.
+In many mammalian species the SC lengths are longer in females, yet most all of these observations come from species with female biased heterochiasmy.
+Our panel of house mice provides an opportunity to test
+if the sex differences in chromatin compaction (SC length) are reversed in the high rec strains (with male-biased heterochiasmy).
+
+< _musculus^PWD^_  and _molossinus^MSM^_).
+A simple model might predict if SC length is the strongest predictor for number of MLH1 foci er cell the high recombining males would have longer SC length proportional to differences in the number of MLH1 foci per cell.>
+
+<chromatin compaction background, process within cells - compaction ratio within cells  >
+
+To account for confounding effect of sex chromosome and compare the sex differences in the process of chromatin compaction during meiosis, we use a reduced bivalent data set made up of bivalents with SC lengths below the 2nd quartile for SC length from a single cell.
+
+This results in a dataset of the shortest 4 to 5 chromosomes and excludes the X bivalent (predicted to be the third longest chromosome) in the oocytes.
+A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Even though this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between sexes.
 
@@
 
-For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer <how much longer the largest difference across sexes by strain are>  
+For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer (Figure X) <how much longer the largest difference across sexes by strain are>  
+(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+
 **`r` the magnitude differences between the mouse mean short bivalent SC range from X to X**
 
-(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
 This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
 
-An additional measure of a smaller subset of single bivalents (which excludes the sex chromsoomes), we tested/compared a measure of the SC lengths for the whole genome (whole cells) using a image nalysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang)
-
-We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads.
-A total of X images were run trhought the image analysis pipeline, and after outliers removed, X remained.
+An additional measure of single bivalents, we tested a measure of the SC lengths for the whole genome (whole cells) using a image nalysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang).
+
+We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads. X cell wide total SC measures remained after and after outliers removed,
 Mouse means were calculated for their total SC area per cell and are displayed in figure X.
@@
 
-```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
+```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE, fig.width = 4, fig.height=2.5}
 
@@
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-Q12.Tot.SC_mouse.av_Mol
-Tot.SC_mouse.av_Mol
-
 ```
@@
 CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-CO.Pos.density_1CO
 
@@
 \nNormalized 1CO Foci pos \n all bivalents")
-Q1.mouse.av.biv_scatr_box
-
 ```
@@
 \nNormalized 1CO Foci pos \nShort Biv")
-Q1.short.biv_scatr_box
 
@@
 
-<!--  plots-->
-
 ```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
@@
 
-<!-- EXPLAIN WHY WE SHOULD CARE ABOUT RAW AND PER IFD values-->
-
-Crossover interference is one of the major determinates -- of the CO landscape. We isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
-`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males to test discribe sex differences in crossover interference.
-Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
-
-We examined both raw IFD^raw^ and normalized normalized by SC length (IFD^norm^)
-
-The mean IFD^raw^ measures of crossover interference as a mechnical force mediated through the SC.
---- we compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
-Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al ).
-
-While the comparisons of the IFD^PER^ metrics can reveal more general recpmbination landscape patterns while controling for the underlying differences in chromatin compaction / SC length.
-(Also in a way this metric is closer to measures of interference from genetic (linkage) maps since they are not quantified / measured in phyiscal space either).
-
 <!--t.tests -->
@@
 
-The raw IFD measures for 2CO bivalents are slightly significant sexes (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`)
--- However _domesticus^G^_ might be an outlier (by having a larger than average degree of sexual dirmohpsim ).
+Crossover interference is one of the major determinates of the recombination landscape, -- shifting it away from random / uniform
+
+shifting the CO positions -- away from a random (distribution)
+
+. We isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
+`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males to test discribe sex differences in crossover interference.
+Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
+We examined both raw IFD^raw^ and normalized normalized by SC length (IFD^PER^).
+The mean IFD^raw^ measures of crossover interference as a mechnical force mediated through the SC.
+We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
+Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al ).
+
+The difference in mean IFD^raw^ are slightly significant sexes from the full data set (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
+However _domesticus^G^_ might be an outlier (by having a larger than average degree of sexual dirmohpsim ).
 When the _domesticus^G^_ observations are removed, the difference is not longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
@@
 
+While the comparisons of the IFD^PER^ metrics can reveal more general recombination landscape patterns while controling for the underlying differences in chromatin compaction and SC length.
+Also this metric is closer to measures of interference from linkage maps which are also removed from physcial scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female mapes in a variety of species (ref).
+
 On the other hand, the difference in mean IFD^PER^ are highly significant between sexes.
-The mean IFD^PER^ are significanly long in size of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
+The mean IFD^PER^ are significanly long in six of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
 Only _musculus^KAZ^_ did not have a difference in mean IFD^PER^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
@@
 
-**THE male female pattern** female close to 50% -- which fits with uniform rec landscape / weak interference
-
-the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs.
-There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
-off. Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?
-
-The general pattern of this sex difference is that the IFD^PER^ distribution is pushed higher   (females have mean of .5  IFD^PER^ where as males have a IFD^PER^ mean of ~.6?)
-
-The differences between the sexes is slightly different depending on if the males are in the high rec group.
-
-<INTRODUCED idea of constrain / thresholds>
-
-in the low rec strains (Dom + KAZ)...
-(both males and females IFD distributions have lower and upper thresholds around 25% and 75%) respectively -- however the weight of observations are slightly shifted across the sexes.
+**THE male female pattern** the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs.
+There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut off.
+ANother way to discribe these results is in terms of thresholds (where the distributions end) with the general pattern of this sex difference is that the IFD^PER^ distribution is pushed higher   (females have mean of .5  IFD^PER^ where as males have a IFD^PER^ mean of ~.6?)
+
+The differences between the sexes is slightly different depending on if the males are in the high rec group. In the low rec strains (Dom + KAZ)...
+(both males and females IFD distributions have lower and upper thresholds around 25% and 75%) respectively.
+however the weight of observations are slightly shifted across the sexes.
 There's an enrichment of lower IFD^PER^ in females relative to males, while the opposite is true for the higher threshold -- where males have an enrichment of higher/long IFD^PER^ compared to females.

### 2020-02-28 doc/Results.rmd
cleaning up doc at the lib

@@
 
-<In order to deconstruct the cell wide MLH1 counts we look at the recombination landscape at the single bivalent level.>
+In order to deconstruct the cell wide MLH1 counts we look at the recombination landscape at the single bivalent level. From our total set of cell images
+
 <FROM number of crossovers per chromsoomes TO rec lanscape across single chromsomes>
@@
 Using this data set, we address two main questions 1) Which traits are sexually dimorphic? and 2) Which traits fit distinguish the high and low recombining strains for males?
+
+In our analysis our first step is to test for measurable differences, between sexes or rec.groups and is the differences are significant, our second step is to apply the same sets of models applied to the mean MLH1 foci counts to test / describe the effects due to subspecies or strains.
 
@@
 
-An additional measure of single bivalents, we tested a measure of the SC lengths for the whole genome (whole cells) using a image nalysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang).
-
-We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads. X cell wide total SC measures remained after and after outliers removed,
+An additional measure of single bivalents, we tested a measure of the SC lengths for the whole genome (whole cells) using an image analysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang).
+
+We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads images. X cell wide total SC measures remained after and after outliers removed,
 Mouse means were calculated for their total SC area per cell and are displayed in figure X.
@@
 
-<We estimate the effect of the sex chromosomes from hand measured data...the XX (3rd longest chromsomes) in females is ~10% of total the summed SC area, the average propotion of the XY in spermatocytes is XX%?
-however the morphology of the XY is highly variable across cells>
-
 <!-- Dom males are shorter than this set of high rec mice
@@
 
-In the mixed models for the sex differences mean **short bivalents** length as the dependant variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+In the mixed models for the sex differences mean **short bivalents** length as the dependaent variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
 The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
 
-For the mixed model run on the mean total SC metric, there are more significant effects for strain and subsp for the mixed model using mean **total sc** per cell as the dependant variable, the three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
+In the mixed model run on the mean total SC metric, there are more significant effects for strain and subsp for the mixed model using mean **total sc** all  of the coefficients were significant, (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
 
@@
 
-In linear models (M2 and M3) for **mean short bivalent length**, male is most significant (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`).
-In the linear model M2 the skive strain and strain by sex effects are slightly significant (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively), however these could be due to low sampling of skive females.
+In linear models (M2 and M3) for **mean short bivalent length**, sex is the most significant coefficient (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`) with slight significant strain and interaction effects (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively).
 
@@
 <In summary, females (oocytes) have different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
-
-<1.
-the longer female sc's is consistant and seen across all strains, even in the high rec strains, suggest a decoupling of sc and gwRR - when comparing across sexes -- or the sc area is used differently for CO in oocytes and spermatocytes.
-2. when the total sc per cell is compared, domesticus males have significantly lower SC arae compared to molossinus and musculus strains.Except in Q2 I Say this isn't true.>
 
@@
 
-Adjusting for XX
-
-Summarize for now, refine this section later.
-
-1. illustrate problem(affects mostly SC length)  
-2. Expected impact on sex comparisons, estimated effect size of the X  
-3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
-4. permutations of 19 female, 19male, 20female 20 male
-
-<Of all female single bivalents observations, 5% are XX (1 of 20). 
-- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
-- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
-- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX>
-
 <transition>  
@@
 <!-- motivating
-
 To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
-In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
 -->
-
-We focus on the foci positions from one crossover (1CO n= `r `) bivalents since the landscape patterns for multi crossover bivalents (2COs) will be highly influenced by crossover interference.
-
-The CO positions are normalized by the SC length and anchored at the centromere with the normalized position values ranging from 0 to 1, (refered here as terminal or telomeric position).
 
@@
 
-In the domesticus strains the mean normalized foci positions were significantly different between sexes and fit the predictions of the 'typical recombination landscape' (with female positions being medially placed and terminally placed in males).
-The p.values across the strains are; (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively).
-
-With the musculus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM.
-Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).
-
-After confirming the significan sex differences for 1CO position, we apply the model frameworks
+We focus on the foci positions from one crossover (1CO n= `r `) bivalents since the landscape patterns for multi crossover bivalents (2COs) will be highly influenced by crossover interference.
+
+The CO positions are normalized by the SC length and anchored at the centromere with the normalized position values ranging from 0 to 1, (refered here as terminal or telomeric position).
+
+In all strains, the 1CO landscape had significant differences between males and female which followed the ‘typical landscape’ of females having central positions and males have terminal foci positions.
+(These results were also supported by the Mixed model which 
+For the mixed model, sex is the most significant effect (LTR; p =1.2610^{-25}).The random effect of strain is also significant (LRT; p =0.01).
+
+After confirming the significant sex differences for 1CO position, we apply the model frameworks
 
@@
 ```
+
+The linear models were used --- to detect if specific strains differed in the sex differences in the 1CO landscape. 
+While sex was the most significant effect for both M2 and M3
 
@@
 
-Crossover interference is one of the major determinates of the recombination landscape, -- shifting it away from random / uniform
-
-shifting the CO positions -- away from a random (distribution)
-
-. We isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
-`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males to test discribe sex differences in crossover interference.
+Crossover interference is one of the major determinates of the recombination landscape.
+It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
+
+ We isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
+`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males to test describe sex differences in crossover interference.
 Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
 We examined both raw IFD^raw^ and normalized normalized by SC length (IFD^PER^).
-The mean IFD^raw^ measures of crossover interference as a mechnical force mediated through the SC.
+The mean IFD^raw^ measures of crossover interference as a mechanical force mediated through the SC.
 We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
@@
 
-We examined the 2CO observations  // the distributions of 2CO IFD^PER^ to better discribe how the means / distributions of this metric are different between the sexes.
-
-We note that in females there are/ may be signals of upper and lower thresdholds for IFD^PER^.
+GENERAL RESTULTS – DIFFERENCES between sexes
+
+In examining the IFD^PER^ distributions, we note that 
+the female norm IFD are centered at ~50% – but they also seem to have slight an enrichment of short (<25%) normalized IFDs
+
+While the general pattern in the male distributions are enrichment longer IFD^PER^.
+(however we note that there are slight differences in the IFD^PER^ distributions between the high and low rec males to be followed up in later section)
 
@@
 
+<!--
 The differences between the sexes is slightly different depending on if the males are in the high rec group. In the low rec strains (Dom + KAZ)...
@@
 
-<!--
 G - the upper and lower thresholds of females look pretty similar, the lower is closer to .2 than .25 -- female has more lower IFDs
@@
 Since the most significant t.test were significantly different across sexes for IFD^PER^ and we follow up with the model framework to.
+
+To test for strain specific effects we applied the model framework
 
@@
 
-glm M2, 
-sex (glm; p= `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`) p = .0080 (this is different in the knited version!! arg)
-
-male*PWD p = 0.02 (glm; p =`r summry.glm.Q1_IFD.PER.M2$coefficients[,4][14]`)
-
-male * SKIVE p = 0.0058  (glm; p=`r summry.glm.Q1_IFD.PER.M2$coefficients[,4][15]`)
-
-for **glm M3**
-
-sex / male p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]` and male*SKIVE just slightly significant (p =  `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][14]`
-
 <For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
@@
 For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
-SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure.
+SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure (p = male-SKIVE p =  )
 
@@
 This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
+
+Our comparison of interference across sexes – supports / applies with general patterns /results from the literature:
+When interference is measured in physical SC units,  – the differences between sexes – is low / only slightly significant 
+ IFD raw is not significantly different (t.tests and models?). – indicating that the 
+While when the variance in chromatin compaction is controlled – males have stronger interference – (like genetic maps) – the impact on the rec landscapes being ….
 
@@
 
-We tested the sex differences -- now we focus on the largest aspect of variation, the high rec males vs the low rec males.
+The results from the previous section -- confirm (known sex differences for the 'typical' 
+recombination landscape (Sardell Kirkpatrick).
+The next section is meant to focus on the greater aspect of variation in mean MLH1 counts per cell the high rec males males strains ( _musculus^PWD^_ , _musculus^SKIVE^_, and _molossinus^MSM^_) and the rest of the males with the main objective of this section is to test for significant correlations with the evolution of mean MLH1 foci per cell.
+
+In comparing the male specific single bivalent based metrics the first task/level is to tests for significant differences between the high and low rec groups.
+The second being to follow up with linear models that test for effects of subspecies and strains.
+
+The literature provides basic predictions for the direction of the relationship the genome wide recombination rate and two aspects of the recombiantion landscape; SC length (positve) and crossover interference (negative).
+Following this logic we predict i) no strain effects for SC length or crossover interference within _domesticus_, ii) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which will have longer SC and weaker interference compared to the other _musculus_ strains and iii) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_.
 
 # Q2 SC Length
-
-<put in plot for total SC and mean gwRR count>
-
-<motivation: the positive correlation, Our prediction for this metric is that the high rec strain will have longer sc lengths (axis lengths)>
 
@@
 
-while testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR
+While testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR
+
+The confounding effects of sex chromosomes are removed in this section that makes comparisons within males.
+Three sc based metrics; total SC, mean short bivalent length and mean long bivalent length.
+With long bivalents isolated in the same manner as the short bivalent.
 
@@
 X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)
-
- place set ups for total SC, short biv and long biv here -->
+`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)-->
 
@@
 
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE}
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE}
 
@@
 
-**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
-`r Q2tot.sc_t.test_High.Low$p.value`).**
-
-Molossinus have significant difference between mouse means for total sc, between the high and low strains (mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
-The total sc mouse means are not significant for the high and low strains in musc (musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+There is a generally positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low rec strains (Figure X).
+
+Within the total data set of mean total SC, high rec strains have significantly more total SC area (ttest; p = `r Q2tot.sc_t.test_High.Low$p.value`).
+However when the high and low rec strains are compared within subspecies only the molossinus strains have significant difference  (ttest mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`), while the musculus strains do not have significant differences (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
 
@@
 
-<across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).>
-
 <!-- models -->
+
+becasue there is weak // some significant differences between the sc metrics of the high and the low rec groups --- we proceed with running the model framework.
+
+**Linear model M1 ** $$mouse \ average \ SC metric ~=~ subsp * strain + \varepsilon $$
+
+**Linear model M2 ** $$mouse \ average \ SC metric ~=~ strain + \varepsilon $$
 
@@
 
-**For the total SC glm**
-
-<Thse are our prediction for the M3 glm (just straint) models based on the variation of the gwRR's
-1. no difference in DOM strains (for SC lengths )
-2. PWD > SKIVE > the rest
-3. Msm > mol>
-
 For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
@@
 
-<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
-
-<For the M2 model analyzing the mouse means short bivalents, all strains have significant effect but LEW, G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.>
-
-<!--24sep14_PWD_m1 is lower than the others, 5jul17_MSM_m1 is the low MSM-->
-
-<just like the sex differences in SC area -- this points to differential usage of the SC for placing CO -- investigate the recombination landscape>
-
-<!--
-blocks of other male Q2 - SC lengths
-what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
-The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
--->
-
-**The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).**
+<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).> --
+
+M3 -- tests within subspecies!!
 
@@
 
-- confirm that the positive correlation of SC and CO number is maintained
-
 - Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.
@@
 - th predictions for M3 (strain) differencs (motivated by the gwrr variation are not met,  (correlation of SC means and CO means?))
+
+**THERE IS a weak signal of the high rec strains having longer SC (having evolved) longer SC (different chromatin compaction) –  While the mouse mean total sc are significantly different between the two groups --- this difference isn’t translated to the reduced bivalent data sets (p = long  p = short)**
+
+<!-- junk chunks below -->
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
+#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+#there's no rec group for this table
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+<These are the plots of the logistic regression.>
 
@@
 For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.>
-
-<!-- junk chunks below -->
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-<These are the plots of the logistic regression.>
 
@@
 
-# brief transition, rec landscape
-
-the lack of predictive power of SC length to predict CO number indicates that there might be different patterns of 'SC usage' / different patterns in the layout of the recombination landscape /  these patterns could stem from placement or spacing of foci.
-
 # Q2.1CO rec landscape evolution is decoupled from gwRR evolution
-
-<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents
-
-Main results:  
-1. WSB has the most terminal 1CO rec landscape and MOLF has the most central.
-2. Evidence for evolution of the 1CO rec landscape, but it is decoupled from the evolution in gwRR we observe in figure 1
-2.5  The normalized 1CO position is not a good metric for distinguishing strains with rapid gwRR evolution
-
--->
-<!-- set up chunks here -->
-<!-- plots /figures below-->
 
@@
 
-The previous section found that the single crossover landscape is significantly different between males and females.
-In this section we focus on differences between the high and the low rec male strains
-In order to test if if there are distinguishing meterics for these two groups.
-**The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**
-
-T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).
-The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
+The normalized 1CO position is not significantly different between the high and low rec groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
+While there were significant strain effects for _domesticus^^WSB^_ and _molossinus^MOLF^_ (figure X), this evolution of the 1CO positioning is decoupled from the total genome wide recombination rate.
+
+<!--
+The high rec musc strains, _musculus^PWD^_ and _musculus^SKIVE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
 While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`).
-This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.
-
-<long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`> 
-<!-- GLMs because there's no difference 
+is that _domesticus^WSB^_ has significantly more terminal positioning,
+and _molossinus^MOLF^_ having significantly more medial position.
+A general pattern of strain specific effects
+we found in looking at the 1CO positions, is that _domesticus^WSB^_ has significantly more terminal positioning, while MOLF has significantly more medial position.-->
+
+<!-- don't report GLMs - since not difference in the ttests, GLMs because there's no difference.
+long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`>
 toal data set, short and long biv. There's only a difference between MSM and MOKF
@@
 <!--
+given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models.
+Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
+<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>
+
 In the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
@@
 **For M3 (all strains)**
-
 strain G (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][2]`).
-
 strain Lew (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][3]`).
-
 strain PWD ( p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][4]`).
-
 slight strain MSM  (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][5]`).
-
 very sig strain MOLF (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][6]`).
-
 strain SKIVE (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][7]`).
-
 strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
@@
 ```
-
-given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models.
-Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
-
-<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>
-
-<!--
-**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$ -->
 
@@
 the mol obj above isn't showing the reight value. 
-
 strain G   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][5]`  
-
 strain LEW   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][6]`  
-
 Strain MSM `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][8]`  
-
 chrm * LEW `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][14]`  
-
 chrm *MSM  `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][16]`
@@
 M2
-
 -chromosome length, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][2]`  
-
 -strain G, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][3]`  
-
 -strain LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][4]`  
-
 -strain MOLF, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][7]`  
-
 -strain SKIVE, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][8]`  
-
 -chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]` -->
@@
 
+<Previous prediction for differences in the long and short bivalent dataset.
+Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+The section on IFD only looks at the 2CO bivalents.>
+
 ```{r Q2.IFD.long.short.setup, echo=FALSE}
@@
 
-<!-- set up chunks here-->
-
 ```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
@@
 
-biv.IFDs_PER_by.group
-
-```
-
-<Previous prediction for differences in the long and short bivalent dataset.
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.>
+```
 
@@
 
-```
-
-```{r Q2.t.test_IFD,echo=FALSE, eval=FALSE}
+t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Dom"))$mean_IFD.2CO_PER   )
+
+t.test_mouse.av.IFD_PER_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_PER   )
+
+t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER   )
+
+```
+
+```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
 
@@
 
-<write some kind of intro here>motivation /
-
-A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation. ()
-
-how crossover interference affects the evolution of genome wide recombination rate -- is
-
-prediction for the interference pattern in the high rec strains is that the high rec strains will have weaker interference (ie.
+<write some kind of A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation.
+how crossover interference affects the evolution of genome wide recombination rate -- is prediction for the interference pattern in the high rec strains is that the high rec strains will have weaker interference (ie.
 less space between foci on the same bivalent, which allows more foci to be placed.
-GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.
-
-The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
-The pattern is confirmed with t.test comparing groups within the same subepcies ..
-add these ttests and p values.
-
-<glms> The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
-
-The mouse averages for IFDs were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
-The pattern is confirmed with t.test comparing groups within the same subepcies.
-
-<glms> The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.
+GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.>
+
+Unlike total SC length and 1CO pattern, IFD metrics are significant predictors of the rapid genome wide recombination rate evolution.
+However, instead of a negative correlation between interference strength and number genome wide recombiantion rate (higher gwRR associated with weaker interference, interference strength is stronger in the high rec strains.
+
+Unexpectedly, the mouse averages for both IFD^raw^ and IFD^PER^ were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The pattern is confirmed with t.test comparing groups within _musculus_ and _molossinus_
+the same subepcies (add these ttests and p values.)
+(t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value`).
+
+We test the within subspecies predictions using the model framework
 
@@
 #summary(Q2.glm.M2_IFD.PER_Mol)
-```
-
-In M1, the model build on the mouse averages of IFD metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]`  for PWD, MSM and SKIVE respectively).
-In M2, the mouse average for normalzied IFD is modeled with only strain effects.
-Only PWD and SKIVE IFD^PER^ had significant strain effects (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]` for PWD and SKIVE respectively).
-
-<!--when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF-->
-
-<!-- we ran similar models using 2CO bivalent observations to examine what effects the SC length might have on IFD^PER^ length.
-
-**model BivLen.M1 ** $$Single bivalent\ IFD ~=~ subsp * strain * chromosomeLength + \varepsilon $$
-
-**model BivLen.M2 ** $$Single bivalent\ IFD ~=~ strain * chromosomeLength + \varepsilon $$
-
-```{r Q2.IFD.glm.chrmLength, echo=FALSE}
+
+```
+
+In M1, the model build on the mouse averages of IFD^PER^ metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
+
+In the M2 model results, _musculus^PWD^_, and _musculus^SKIVE^_, were the only significant strain effects, (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]`).
+
+Running M2 within subspecies – confirms the within subepcies predictions across strains – (no significant strain effects
+
+Q2.glm.M2_IFD.PER_Dom
+
+Q2.glm.M2_IFD.PER_Musc
+
+Q2.glm.M2_IFD.PER_Mol
+
+```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
 
@@
 
+```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
+
+#calculate the rates of 'close foci' (<30%)
+#examine the changes in the distribution closer.
+
+#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
+
+close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                      n.close.foci = length(which(IFD1_PER <= .3)),
+
+                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                      rate.close.foci = n.close.foci / N.obs
+)
+
+close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
+
+close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("strain"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER),
+
+                      inverse.close.foci = length(which(IFD1_PER > .3)),  
+                      n.close.foci = length(which(IFD1_PER < .3)),
+                      rate.close.foci = n.close.foci / N.obs,
+
+                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
+
+)
+#the
+
+show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
+
+#add rec.group
+close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
+
+#POOL all short and 
+#chi square test
+high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
+                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
+#p-value = 0.03, slight dif tween 3 high rec strains
+
+#don't need to run chi square for this pattern, since it's enough to
+
+```
+
+```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
+
+#Log regression all show significant prediction
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
+
+#Full dataset
+
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+```
+
+We determined that the main difference in IFD^PER^ disbribution across the high and low groups is an enrichment of IFD^PER^ observations under 30% in low rec strains.
+
+The rate of IFD^PER^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
+
+<!--
 For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs).
@@
 
-```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
-
-#calculate the rates of 'close foci' (<30%)
-#examine the changes in the distribution closer.
-
-#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
-
-close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                      n.close.foci = length(which(IFD1_PER <= .3)),
-
-                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                      rate.close.foci = n.close.foci / N.obs
-)
-
-close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
-
-close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("strain"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER),
-
-                      inverse.close.foci = length(which(IFD1_PER > .3)),  
-                      n.close.foci = length(which(IFD1_PER < .3)),
-                      rate.close.foci = n.close.foci / N.obs,
-
-                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
-
-)
-#the
-
-show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
-
-#add rec.group
-close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
-
-#POOL all short and 
-#chi square test
-high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
-                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
-#p-value = 0.03, slight dif tween 3 high rec strains
-
-#don't need to run chi square for this pattern, since it's enough to
-
-```
-
-```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
-
-#Log regression all show significant prediction
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-```
-
-We determined that the main difference in IFD disbribution across these groups is an enrichment of 'close foci', IFD^PER^s less than 30% of the total SC length in low rec strains.
-The rate of these close foci range from 8.2% (G) to 16% (KAZ) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for skive, MSM and PWD respectively).
-<no chi sqaure test see above>
-
- **(unlike total SC length and 1CO pattern, IFD metreics are a significant predictors of rapid genome wide recombination rate evolution.
-However, they go agaisnt a standard perdiction for the relationship between the chromosome level recombination landscape and genome wide recombiantion rate.**
-
-<!-- 
-Since the majority of cells from the Automated BivData set do not have the complete complement of bivalents this subset is much smaller and might have 'smaller' bivalents from cells missing the longer bivalents.
-But assuming the algorithm's identification of bivalents in random, with a large enough sample size, hopefully this error would be minimized.
+<!-- final summary, BE brief
+
+BE brief,
+
+final summary, highlight the themes of 1. more variance in female traits (MLH1 foci per cell, rec landscape / interference)
+
+while the male traits -- more focused (less variance in MLH1 counts per cell, shorter chromatin comparction?  bias position, stronger interference)
+
+(should I measure the var of all traits in Q1)
+
+2. rapid evolution of genome wide rec rates can fit into this -
+
+- evolution of more COs requires more DSBs (evidence points to a earlier step which is likely the chromatin compaction (amount of axis))
+
+- the strongest single bivalent predictor/correlate of rapid gwRR for males is interferece strength in 2COs
+
+AND somewhat puzzeling result of (having stronger interference), meaning there is a greater proportion of a bivalent that seperates two foci  (2 foci push to the ends of chrms (maybe decreased centromere supression))
+
 -->
+
+# Blank1
+
+<!-- unused chunks -->
 
@@
 (the extra stats are not correctly mapped)>
-
-<!-- final summary, BE brief-->
-
-BE brief,
-
-final summary, highlight the themes of 1. more variance in female traits (MLH1 foci per cell, rec landscape / interference)
-
-while the male traits -- more focused (less variance in MLH1 counts per cell, shorter chromatin comparction?  bias position, stronger interference)
-
-(should I measure the var of all traits in Q1)
-
-2. rapid evolution of genome wide rec rates can fit into this -
-
-- evolution of more COs requires more DSBs (evidence points to a earlier step which is likely the chromatin compaction (amount of axis))
-
-- the strongest single bivalent predictor/correlate of rapid gwRR for males is interferece strength in 2COs
-
-AND somewhat puzzeling result of (having stronger interference), meaning there is a greater proportion of a bivalent that seperates two foci  (2 foci push to the ends of chrms (maybe decreased centromere supression))
-
-# Blank1
-
-<!-- unused chunks -->
 

### 2020-02-28 doc/Results.rmd
editing lg

@@
 
-This is a draft of the main figure. Outgroups are not included in the analysis below. (add horizontal lines for denoting subspecies, group means and expected minimum)
-
-<!--  insert Details of data set cleaning etc is in SetupData.rmd
-Effects from different factors are reviewed in Effects.Rmd -->
-
 ```{r main.fig1s, echo=FALSE, include=FALSE}
@@
 
-<This striking differences across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
-The pattern of strain averages distributed around the total sample mean in females is supportive of stabilizing selection or neutral evolution.
-In contrast to the pattern from male means which have rapid evolution in a subset of genetic backgrounds this fits a model of directional selection.
-(We attempt to make better evolutionary models by applying an evolutionary framework)>
-
 ```{r female.strain.table, echo=FALSE}
@@
 
-While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`), the effect sizes have a smaller range of (4 to 1), indicateing significant, but small strain specific effects on mean MLH1 foci per cell.
-<for this reason differences between females aren't studied further>
+While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`), the effect sizes have a smaller range of (4 to 1), indicating significant but small strain specific effects on mean MLH1 foci per cell.
 
@@
 
-<Variance of a trait can contain information pertinent to its mode of evolution and mechanistically could reflect the strength of constraints acting on the trait.> We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
+We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
 
@@
 
-There is no significant difference in the amount of within mouse variance of MLH1 foci counts per cell in males according the the linear models.
+There is no significant difference in the amount of within mouse variance of MLH1 foci counts per cell in males according to the linear models.
 In females, the significance of strain effects was not consistent between the full data set and higher quality dataset _domesticus^LEW^_ (glm; p = 0) was a significant effect in the full model and _musculus^PWD^_ (glm; p =0.04) in the higher quality dataset.
@@
 
-The strikingpatterns of variation across male and females suggest that the genome wide recombination rates have distinct evolutionary trajectories.
+The striking patterns of variation across male and females suggest that the genome wide recombination rates have distinct evolutionary trajectories.
 For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome wide recombination rate.
@@
 
-<!--
-explain why male oer female measures were used, -- mention some measures were made -- but the number of cells was so too small.
--->
-
 In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in prophase spermtaocytes.
@@
 
-For all strains there were significantly more DMC1 foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the cell stages in prophase.
+<All strains had significantly more DMC1 foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the cell stages in prophase.>
+
 The mean number of DMC1 foci per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
@@
 
-<explain the biological significance of the ratios!> After DSB formation, DSBs are repaired as either NCO or CO (CO designation step) with the vast majority being repaired as non-crossover NCO.
+ After DSB formation, DSBs are repaired as either non-crossover (NCO) or crossovers (COs).
+with the vast majority being repaired as NCOs.
 Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated into COs.
 The ratios, calculated for DMC1 means from both stages of prophase, are not significantly different between the high and low strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
-<These results indicate that the targets of evolution which lead to differences in crossover numbers may be established before DSB formation (cite Baier et al 2014).>
 
@@
 
-We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per chromosomes (bivalent).
+We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per bivalent.
 Because our observations of MLH1 foci are made in pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
-Additionally we can distinguish the idenetiy of indiviual chromsomes (Chromosome 1 or Chromsome 2).
-for these reasons we use the term bivalent.
+Additionally we can distinguish the identity of chromsomes (e.g.
+Chromosome 1 or Chromosome 2), for these reasons we use the 'term bivalent'.
 
 Ninety six percent of the pooled dataset of single bivalents (n=`r total.biv.all`) have either one or two crossovers (Figure X).
-Across the female observations the proportions of chromosome classes are not notably different.
 The proportion of 1CO:2CO distinguishes the high and low rec strains (Figure X).
@@
 
-While the proportions are significantly different (chi-square-test; p = `r PWD.MSM.2CO.prop.test$p.value`),Males from _musculus^PWD^_ and _molossinus^MSM^_ have 2CO proportions more similar to each other than strains from the same subspecies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-
-<SUMMARY the TRANSITION to single bivalen measures>
+While the proportions are significantly different (chi-square-test; p = `r PWD.MSM.2CO.prop.test$p.value`), males _musculus^PWD^_ and _molossinus^MSM^_ have 2CO proportions more similar to each other than strains from the same subspecies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
+
+<Indicateing, SUMMARY the TRANSITION to single bivalen measures>
 
@@
 
-In order to deconstruct the cell wide MLH1 counts we look at the recombination landscape at the single bivalent level. From our total set of cell images
-
-<FROM number of crossovers per chromsoomes TO rec lanscape across single chromsomes>
-<A big motivation was to quantifiy recombiantion landscape differences across single bivalents>
-
+For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
 The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures (Peterson 2019).
@@
 
-Due to overlap of chromosomes most cells do not have all chromsomes measured, but we assume that the isolation of chromsomes within cell images is not biased.
+Due to overlap of chromosomes most cells do not have measurements for all bivalents, but we assume that the isolation of chromsomes within cell images is not biased.
 From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
@@
 
+<chromatin compaction background, process within cells - compaction ratio within cells>
+
 ```{r Q1.Short.Long_biv.setup, echo=FALSE}
@@
 In many mammalian species the SC lengths are longer in females, yet most all of these observations come from species with female biased heterochiasmy.
-Our panel of house mice provides an opportunity to test
-if the sex differences in chromatin compaction (SC length) are reversed in the high rec strains (with male-biased heterochiasmy).
-
-< _musculus^PWD^_  and _molossinus^MSM^_).
-A simple model might predict if SC length is the strongest predictor for number of MLH1 foci er cell the high recombining males would have longer SC length proportional to differences in the number of MLH1 foci per cell.>
-
-<chromatin compaction background, process within cells - compaction ratio within cells  >
-
-To account for confounding effect of sex chromosome and compare the sex differences in the process of chromatin compaction during meiosis, we use a reduced bivalent data set made up of bivalents with SC lengths below the 2nd quartile for SC length from a single cell.
-
-This results in a dataset of the shortest 4 to 5 chromosomes and excludes the X bivalent (predicted to be the third longest chromosome) in the oocytes.
+Our panel of house mice provides an opportunity to test if the sex differences in chromatin compaction (SC length) are reversed in the high rec strains (with male-biased heterochiasmy; _musculus^PWD^_  and _molossinus^MSM^_).
+A simple model might predict if SC length is the strongest predictor for number of MLH1 foci per cell males of the high recombining strains would have longer SC length leading to the difference in total MLH1 foci per cell.
+
+To account for confounding effects of sex chromosomes on measures of chromatin compaction (SC area) we use a reduced bivalent data set made up of bivalents with SC lengths below the 2nd quartile for SC length from a single cell to compare across sexes.
+This results in a dataset of the shortest 4 to 5 chromosomes and excludes the X bivalent in the oocytes, which is predicted to be the third longest chromosome by Mb.
 A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
@@
 
-t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Dom"))$mean_IFD.2CO_PER   )
+#different test must be sed for the Dom
+
+#t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == #"male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
+#   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == #"Dom"))$mean_IFD.2CO_PER   )
 
@@
 
+<This striking differences across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
+The pattern of strain averages distributed around the total sample mean in females is supportive of stabilizing selection or neutral evolution.
+In contrast to the pattern from male means which have rapid evolution in a subset of genetic backgrounds this fits a model of directional selection.
+(We attempt to make better evolutionary models by applying an evolutionary framework)>
+
 # Blank1

### 2020-02-28 doc/Results.rmd
editing lg, still

@@
 
-MLH1.merge.Skel_2 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
+MLH1.merge.Skel_1 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
 
@@
 
-MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
-MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
-
+MLH1.merge.Skel_2 <- MLH1.merge.Skel_1[!MLH1.merge.Skel_1$fileName %in% remove.mice_round2, ]
+#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
+#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
 all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
+
+MLH1.merge.Skel <- MLH1.merge.Skel_2
+
+MLH1.merge.Skel <- add_species(MLH1.merge.Skel)
+
+```
+
+```{r tot.skel.table, echo=FALSE}
+totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
+
+#unorder the factors
+totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
+totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
+totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
+totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
+
+OG.totSC.mouse.av.table <- totSC.mouse.av.table
+
+#REMOVE JUVI mice
+tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
+total SC area")
+
+mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
+
+```
+
+```{r, tot.SC.setup, echo = FALSE}
+
+totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
+                 .(sex,strain),
+                 summarize, 
+                 ncells = sum(ncells),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(mean.skel, na.rm = TRUE),
+                 mean.bin =  mean(mean.bin, na.rm = TRUE)
+
+)
+
+totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
+
 ```
@@
 -->
-
-```{r tot.skel.table, echo=FALSE}
-
-totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
-
-#unorder the factors
-totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
-totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
-totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
-totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
-
-OG.totSC.mouse.av.table <- totSC.mouse.av.table
-
-#REMOVE JUVI mice
-tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
-total SC area")
-
-mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-```
-
-```{r, tot.SC.setup, echo = FALSE}
-
-totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
-                 .(sex,strain),
-                 summarize, 
-                 ncells = sum(ncells),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(mean.skel, na.rm = TRUE),
-                 mean.bin =  mean(mean.bin, na.rm = TRUE)
-
-)
-
-totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
-
-```
 
@@
 ```{r Q1.SC.show.short.biv, echo=FALSE}
+
 Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
@@
 
-For all but one strain, _musculus^SKIVE^_, the mouse mean for the short bivalents are significantly longer (Figure X) <how much longer the largest difference across sexes by strain are>  
-(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
-
-**`r` the magnitude differences between the mouse mean short bivalent SC range from X to X**
-
-This pattern is also true for the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.
-
-An additional measure of single bivalents, we tested a measure of the SC lengths for the whole genome (whole cells) using an image analysis algorithm which measures the skeletonized bivalent and provides an acurate measure of the lengths of single SC/bivalents (ref RWang).
-
-We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads images. X cell wide total SC measures remained after and after outliers removed,
-Mouse means were calculated for their total SC area per cell and are displayed in figure X.
-
-The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
-Breifly/in effect the oocytes have an additional autosome, and the spermatocytes have a mostly recombinationly inactive unit of SC (the XY).
+```{r Q1SC_short.biv_dif, echo=FALSE}
+
+#calq the different in short.biv means
+
+#WSB
+Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
+
+#LEW
+Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
+
+#G
+Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
+
+#PWD
+Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
+
+#KAZ
+Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
+
+#SKIVE
+Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
+
+#MSM
+Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
+
+#this bivData ratio is the same as mouse average
+Q1SC.BivData.short.biv_DIFF_MSM <- mean( (real.short.bivData %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (real.short.bivData %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
+
+```
+
+All strains have longer mouse mean for short bivalents in females (Figure X) and all but one strain, SKIVE, had significant differences.
+The difference, (F:M), across strains ranges from `r Q1SC.mus.av.short.biv_DIFF_MSM` in MSM to `r Q1SC.mus.av.short.biv_DIFF_WSB` in WSB.
+
+<(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.>
+
+For an addtional metric of chromatin compaction, we tested a whole cell measure of SC length using an second image analysis algorithm which converts the SC signal (into single pixel wide skeleton) and provides accurate estimates of the summed SC length of bivalents for single cells (ref Wang).
+We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads images.
+Mouse means were calculated from `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cell wide total SC measures (figure X).
+(where as `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` number of cells had MLH1 measures ).
+
+<The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
+Breifly/in effect the oocytes have an additional autosome, and the spermatocytes have a mostly recombinationly inactive unit of SC (the XY).>
 
@@
 
-In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; _domesticues^WSB^ _ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^ _  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^ _ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^ _ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, KAZ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^ _ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`
+In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; _domesticues^WSB^_ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^_  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^_ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^_ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, _musculus^KAZ^_ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^_ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`
 
@@
 
-<**t.test results for total.SC**
-all values (F vs M) (p= `r Q1_tot.sc_t.test.all$p.value`)
-Dom (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.DOM$p.value`)
-Musc (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.MUSC$p.value`)  (note, missing SKIVE, not enough female measures)  
-MSM (F vs M) -- (ttest; p= `r Q1_tot.sc_t.test.MSM$p.value`)>
-
-<explain why I'm applying models>
-
-We follow up the general result of females have longer SC lengths -- by applying the same model frameworks which we used to analyze the mean MLH1 foci per cell -- to understsand the evolutionary framework.
+To test for strain and subspecies specific effects on the sex difference in SC lengths we apply the following set of models.
 
@@
 
-In the mixed models for the sex differences mean **short bivalents** length as the dependaent variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
+In the mixed models for the sex differences mean **short bivalents** length as the dependent variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
 The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
 
-In the mixed model run on the mean total SC metric, there are more significant effects for strain and subsp for the mixed model using mean **total sc** all  of the coefficients were significant, (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
+In the mixed model run on the mean total SC metric all of the coefficients were significant, (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
 
@@
 (this means...).
-Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized / quantified at this level.
+Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized at this level.
 
@@
 -->
-
-<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >
-
-<In summary, females (oocytes) have different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
 
@@
 
+<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >
+
+<In summary, females (oocytes) have different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
+
 <Q12 total SC mouse means!
@@
 
-<transition>  
--following up on SC area - with how COs are place - placement and number of COs on single bivalents
-
-1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
--- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).
-
-2.
-We focus on two metrics/ aspects i) the placement of single foci along a bivalent (just 1CO) and ii) the placement of two foci on the same bivalent / the interfocal distance of 2CO bivalents (reflects crossover interference)
+The SC length per chromsomes can be thought of as the 'area / available real-estate for COs.
+These results suggest that this area is 'utilized' in different ways between sexes, espicially in the strains with the high rec males, which have less 'area' for crossovers compared to females, but have more crossovers per cell on average.
+
+We focus on aspects of the recombination landscape for the two major classes of bivalents; i) the placement of single foci along a single crossover bivalent and ii) the spacing of two foci on the same bivalent, or the interfocal distance (IFD).
 
@@
 \nNormalized 1CO Foci pos \n all bivalents")
+
 ```
@@
 #norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
+
+Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
 
@@
 #single.biv.1CO_pos_t.test <- t.test( 
-
  # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-
 #   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
@@
 
-We focus on the foci positions from one crossover (1CO n= `r `) bivalents since the landscape patterns for multi crossover bivalents (2COs) will be highly influenced by crossover interference.
-
+The mean normalized foci positions per mouse from `r length( ( cleanCurated_BivData %>% filter(hand.foci.count == "1") )$Obj.ID )` single crossover bivalents.
 The CO positions are normalized by the SC length and anchored at the centromere with the normalized position values ranging from 0 to 1, (refered here as terminal or telomeric position).
 
-In all strains, the 1CO landscape had significant differences between males and female which followed the ‘typical landscape’ of females having central positions and males have terminal foci positions.
-(These results were also supported by the Mixed model which 
-For the mixed model, sex is the most significant effect (LTR; p =1.2610^{-25}).The random effect of strain is also significant (LRT; p =0.01).
-
-After confirming the significant sex differences for 1CO position, we apply the model frameworks
+In all strains, the 1CO landscape had significant differences between males and female which followed the "typical landscape" of sex difference observed in many other species (Sardell Kirkpatrick).
+The foci position was more medially places in females while males have more terminal normalized foci positions (ttest; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
 
@@
 
-The linear models were used --- to detect if specific strains differed in the sex differences in the 1CO landscape. 
-While sex was the most significant effect for both M2 and M3
-
-For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant
-(LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
-These results confirm the sex difference establish in the t.tests above -- and suggesting variation across the strains (in the sexual dimorphism for this pattern).
+For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant (LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
+
+The linear models were used to detect if specific strains differed in the sex differences in the 1CO landscape.
+While sex was the most significant effect for both M2 and M3 (there were some strain specific effects...)
 
@@
 While the musc subspecies, LEW and KAZ strains had slightly signifcant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
-
-(these results will be explored more fully in the Q2 section).
-The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).
-
-REMOVE THE SISCOTEN PLOTs...
-
-<ONLY SHOW 1 PLOT, we don't explore the effects on the redcued bivalent data sets! The main expected pattern is confirmed>
 
@@
 Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
-We examined both raw IFD^raw^ and normalized normalized by SC length (IFD^PER^).
+We examined both raw IFD^raw^ and normalized by SC length (IFD^PER^).
 The mean IFD^raw^ measures of crossover interference as a mechanical force mediated through the SC.
@@
 The difference in mean IFD^raw^ are slightly significant sexes from the full data set (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
-However _domesticus^G^_ might be an outlier (by having a larger than average degree of sexual dirmohpsim ).
+However _domesticus^G^_ might be an outlier (by having a larger than average degree of sexual dimorphism).
 When the _domesticus^G^_ observations are removed, the difference is not longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
-This indicates there is no general pattern of sex differences in crossover interference measured in phyiscal (SC) units.
-(confirming deBoer and other models of crossover interference acting through phyical/mechanical force).
-
-While the comparisons of the IFD^PER^ metrics can reveal more general recombination landscape patterns while controling for the underlying differences in chromatin compaction and SC length.
-Also this metric is closer to measures of interference from linkage maps which are also removed from physcial scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female mapes in a variety of species (ref).
-
-On the other hand, the difference in mean IFD^PER^ are highly significant between sexes.
-The mean IFD^PER^ are significanly long in six of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
+This indicates there is no general pattern of sex differences in crossover interference measured in physical (SC) units.
+(confirming deBoer and other models of crossover interference acting through physical/mechanical force).
+
+While the comparisons of the IFD^PER^ metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+Also this metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+
+The difference in mean IFD^PER^ are highly significant between sexes.
+The mean IFD^PER^ are significantly long in six of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
 Only _musculus^KAZ^_ did not have a difference in mean IFD^PER^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
 These results indicate a general pattern of two foci are seperated by more area (hence stronger interference) in males.
-<which is supported by the pattern in genetics maps>
+These results indicate a general pattern of two foci are separated by more area (hence stronger interference) in males.
 
@@
 
-GENERAL RESTULTS – DIFFERENCES between sexes
+GENERAL RESTULTS DIFFERENCES between sexes
 
@@
 
-**THE male female pattern** the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs.
-There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut off.
-ANother way to discribe these results is in terms of thresholds (where the distributions end) with the general pattern of this sex difference is that the IFD^PER^ distribution is pushed higher   (females have mean of .5  IFD^PER^ where as males have a IFD^PER^ mean of ~.6?)
+THE male vs female pattern...
+the female norm IFD are centered at ~50% but they also seem to have a higher rate of short (<25%) normalized IFDs.
+There seems to be a 25% norm.IFD cutoff (in females) but some mice seem to have less of this cut off.
+Another way to describe these results is in terms of thresholds (where the distributions end) with the general pattern of this sex difference is that the IFD^PER^ distribution is pushed higher (females have mean of .5  IFD^PER^ where as males have a IFD^PER^ mean of ~.6?)
 
@@
 
-(IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`) subspecies, interaction are slightly significant.
-<ONLY REPORT FOR PER MEASURES>
-
-<For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
-The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
-For the M3, only SKIVE*male effect is close to significant>
+(IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`) subspecies, interaction are slightly significant. <ONLY REPORT FOR PER MEASURES>
 
@@
 
-Our comparison of interference across sexes – supports / applies with general patterns /results from the literature:
-When interference is measured in physical SC units,  – the differences between sexes – is low / only slightly significant 
- IFD raw is not significantly different (t.tests and models?). – indicating that the 
-While when the variance in chromatin compaction is controlled – males have stronger interference – (like genetic maps) – the impact on the rec landscapes being ….
+Our comparison of interference across sexes supports the general patterns from the literature.
+
+IFD raw is not significantly different (t.tests and models?). indicating that the...
+
+When interference is measured in physical SC units, the differences between sexes is low / only slightly significant -leading support that the phyisical measures of interference (in SC units) is conserved within species.
+
+While when the variance in chromatin compaction is controlled males have stronger interference (like genetic maps) the impact on the rec landscapes being.
 
@@
 
+<For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
+The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
+For the M3, only SKIVE*male effect is close to significant>
+
 <remove the chrm size effects -- or check out model with chromosome length
@@
 # Q2 SC Length
+
+While testing the sex differences, there were results (the significant subspecies effect) which suggest that the _musculus_ and _molossinus_ strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may facilitate the rapid male specific evolution in genome wide recombiantion rate.
+
+The confounding effects of sex chromosomes are removed in this section that makes comparisons within males.
+Three sc based metrics; total SC, mean short bivalent length and mean long bivalent length.
+With long bivalents isolated in the same manner as the short bivalent.
+
+<!-- basic stats
+`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
+X observations removed for having extra signal from other cells.
+`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)-->
 
@@
 They have a lower probability of 2COs below a certain SC length threshold.>
-
-While testing the sex differences, there were results (the significant subspecies effect) which suggest that the musculus and molossinus strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may proceed - facilitate the rapid male specific evolution in gwRR
-
-The confounding effects of sex chromosomes are removed in this section that makes comparisons within males.
-Three sc based metrics; total SC, mean short bivalent length and mean long bivalent length.
-With long bivalents isolated in the same manner as the short bivalent.
-
-<!-- basic stats
-`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
-X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)-->
-
 ```{r Q2.SC.short.long.setup, echo=FALSE}
@@
 
-<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).> --
-
-M3 -- tests within subspecies!!
+<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
+
+**M3 -- tests within subspecies!!**
 
@@
 
-> **Summary Points**
-
-- Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.
-
-- (the glms, indecate there are many strain effects), Using the mouse average SC lengths, in the full data set all strain effects are significant.
-This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.
-
-- For the M3 models of mouse av.s across SC meterics, for many of the SC metrics, most all of the strain effects are significant
-
-- for the M2 models of the mouse av, musc - mol are significant fixed effects
-
-- many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging
-
-- th predictions for M3 (strain) differencs (motivated by the gwrr variation are not met,  (correlation of SC means and CO means?))
-
-**THERE IS a weak signal of the high rec strains having longer SC (having evolved) longer SC (different chromatin compaction) –  While the mouse mean total sc are significantly different between the two groups --- this difference isn’t translated to the reduced bivalent data sets (p = long  p = short)**
+**THERE IS a weak signal of the high rec strains having longer SC (having evolved) longer SC (different chromatin compaction)  While the mouse mean total sc are significantly different between the two groups --- this difference --- translated to the reduced bivalent data sets (p = long  p = short)**
 

### 2020-02-28 doc/Results.rmd
getting tired

@@
 
-While testing the sex differences, there were results (the significant subspecies effect) which suggest that the _musculus_ and _molossinus_ strains had longer SC compared to the domesticus strains - these two strains also contain the high rec strains - motivating a model/hypothesis that sc length / chromatin compaction evolution may facilitate the rapid male specific evolution in genome wide recombiantion rate.
-
-The confounding effects of sex chromosomes are removed in this section that makes comparisons within males.
-Three sc based metrics; total SC, mean short bivalent length and mean long bivalent length.
-With long bivalents isolated in the same manner as the short bivalent.
+Our prediction for this section is that the high rec strains will have longer SC lengths proportional to the increase in the mean genome wide recombination rate.
+
+The confounding effects of sex chromosomes are removed for this section that compares across males from different inbred strains.
+We compared mouse means of three SC based metrics; total SC, mean short bivalent length and mean long bivalent length.
+The long bivalent data set was isolated in the same manner as the short bivalent data set.
 
@@
 mouse.av_totsc_male
+
 ```
@@
 X.by.texas
+
 ```
@@
 ```
-
-<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
-remember the pairwise t.test
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-G-WSB is sig, WSB-LEW are slightly sig
--->
 
@@
 
-There is a generally positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low rec strains (Figure X).
-
-Within the total data set of mean total SC, high rec strains have significantly more total SC area (ttest; p = `r Q2tot.sc_t.test_High.Low$p.value`).
-However when the high and low rec strains are compared within subspecies only the molossinus strains have significant difference  (ttest mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`), while the musculus strains do not have significant differences (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-The mouse means for the reduced bivalent datasets (short and long) are not significantly different between the high and low rec strains (for the total pooled data) (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
+There is a positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low rec strains (Figure X).
+While for the total data set of mean total SC, the high rec strains have significantly  (ttest; p = `r Q2tot.sc_t.test_High.Low$p.value`) when the high and low rec strains are compared within subspecies only the _molossinus_ strains have significant difference  (ttest mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`), while the _musculus_ strains do not have significant differences (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+Additionally, the means for the reduced bivalent datasets, short and long bivalents, are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
 
@@
 
-becasue there is weak // some significant differences between the sc metrics of the high and the low rec groups --- we proceed with running the model framework.
+We follow up the general pattern of a weak signal of high rec strains having longer SC lengths by running the model framework.
 
@@
 ```{r Q2.totSC.glm, echo=FALSE}
-#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel
 
 Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-
 sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
@@
 Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-
 sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
 #round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
+
+totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+#within subspecies values
+
+#Dom
+Q2.glm_mus.av_totSC_M2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
+sum.Q2.glm_mus.av_totSC_M2_DOM  <- summary(Q2.glm_mus.av_totSC_M2_DOM)
+#there are significant strain effects in DOM,
+
+#Musc
+Q2.glm_mus.av_totSC_M2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
+sum.Q2.glm_mus.av_totSC_M2_MUSC  <- summary(Q2.glm_mus.av_totSC_M2_MUSC)
+#KAZ -- is the only significant effect
+
+#Mol
+Q2.glm_mus.av_totSC_M2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
+sum.Q2.glm_mus.av_totSC_M2_MOL  <- summary(Q2.glm_mus.av_totSC_M2_MOL)
+
+#these pair-wise t.test aren't working
+#dom
+#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean.skel, 
+#                 ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$strain )
+
+#musc
+#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$mean.skel, 
+#                 ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$strain )
+
+#mol
+#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean.skel, 
+#                 ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$strain )
+
 ```
@@
 
+#this dataset is all males
 glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
-
 sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
+#
 
 glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
-
 sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
-
 #round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+
+glmQ2_long.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_long.biv_M2_DOM <- summary(glmQ2_long.biv_M2_DOM)
+
+#MUSC
+glmQ2_long.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_long.biv_M2_MUSC <- summary(glmQ2_long.biv_M2_MUSC)
+
+#MOL can't be done -- only 1 mouse for each MSM and MOLF
+glmQ2_long.biv_M2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
+sum.glmQ2_long.biv_M2_MOL <- summary(glmQ2_long.biv_M2_MOL)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
+
 ```
@@
 glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-```
-
-```{r Q2.SC.long.glm.bivglm, echo=FALSE}
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)
-
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
-#summary(glm_SC.male_Musc_long.biv)
-
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)
-
-## logistic regression for Rec group
-#log.SC.male_long.SC <- glm(Rec.group ~ mean_SC, 
-#              data=Long.biv_mouse_table,  family=binomial(link="logit"))
-
-#summary(log.SC.male_long.SC)
-
-#log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
-#              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-
-```
-
-For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
-
-In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+
+#DOM
+glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
+
+#MUSC
+glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
+
+#MOL
+#within subspecies
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+```
+
+For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= _Musculus_ `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and _Molossinus_ p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
+
+In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= _Musculus_ `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and _Molossinus_ p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+
+long bivalent set in M1 --
+
+Musc `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`
+
+MSM strain `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`
 
@@
 
-**M3 -- tests within subspecies!!**
+**M3 -- tests within subspecies!!** SO the M3 model can be run within species specific data -- to test the predictions (based on the variation in mean MLH1 counts), (use all 3 metircs) **the within subspecies M3 models don't fully meet the predictions
+
+for total.Sc : there are significant Dom strain effects: in musc kaz is significant negative, several other strains have higher total SC, (MOL in the intercept is significant).
+
+short biv: Dom: Significant G effect, MUSC no significant -- strain effects
+
+long.biv:  Dom: no strain effects, MUSC no significant strain effects, MOL - not enough observations
 

### 2020-02-29 doc/Results.rmd
incorporating hand edits

@@
 molossinus_.
-Mean MLH1 foci counts were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
+Mean MLH1 foci counts for each mouse were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
 <!-- state the average number of mice
@@
 The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome wide recombination rates averages are greater females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (F:M rates) is generally low in these house mouse strains, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
-Three notable exceptions of male biased heterochiasmy are from the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
+Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
+Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
 
@@
 
+We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
+All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
+
 **M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
@@
 
-We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
-All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
-
  <These results allow use to reject the hypothesis of genome wide recombination rates (varying consistantly by sex across subepcies -- and having variance within subspecies due to strains)>
@@
 The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombination rates.
-The subspecies fixed effects tested in M2, had weak to no significance (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
+The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
 m.
@@
 The two most significant strain effects were _domesticus^G^_ and _musculus^MSM^_ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]` for _domesticus^G^_ and _molossinus^MSM^_ respectively).
-
 In addition to significant strain effects for _molossinus^MSM^_, (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
@@
 
-Combined these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of sexual dimorphism/heterochiasmy in _domesticus^G^_.
+Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in _domesticus^G^_.
 
@@
 There were three significant strain effects in the male specific data; _musculus^PWD^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` with effect sizes ranging from 5, 7, and 2 foci respectively.
-Given these results, theses strains are classified as 'high rec' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
+Given these results, theses strains are classified as 'high recombining' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
 
@@
 
-We examine the within animal variance in MLH1 foci count per cell within the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
+We examine the within animal variance in MLH1 foci count per cell using the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
 
@@
 
-# Evolution of Genome wide recombination rate associated with evolution of mean DSB number
+# Evolution of genome wide recombination rate associated with evolution of mean DSB number
 
@@
 
-In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for DSBs, DMC1, in prophase spermtaocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes from juvenile mice (12 to 18 days) of three low rec, ( _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_) and two high rec strains ( _musculus^PWD^_ and _molossinus^MSM^_).
+In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for double strand breaks (DSBs), DMC1, in prophase spermtaocytes.
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes from juvenile mice (12 to 18 days) of three low rec, ( _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_) and two high recombining strains ( _musculus^PWD^_ and _molossinus^MSM^_).
 
@@
 
-The mean number of DMC1 foci per cell is associated with evolution of the number of precursors (DSBs), the two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
+The mean number of DMC1 foci per cell is associated with evolution of the number of precursors (DSBs).
+The two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
 However the differences in DMC1 foci were not significant for zygotene cells (later prophase) between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
-Indicating that this marker at early prophase is more predictive of the downstream crossover number differences.
+Indicating that DMC1 counts at this stage is more predictive of the downstream crossover number differences.
 
@@
 with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated into COs.
-The ratios, calculated for DMC1 means from both stages of prophase, are not significantly different between the high and low strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene based ratios respectively).
+Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated into COs, partially known as the NCO:CO decision.
+The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios respectively).
 
@@
 
+# Single Bivalent Level Results
+
+```{r algorithm_performance_stats, echo=FALSE}
+
+#sort by category, 
+#compare total boxes and rate passed
+#Curated_BivData  -- pass rate?
+
+#number of mice, number of cells
+#mean number of isolate bivs per cell,
+#use these two metrics
+#cell.level.BivData_table$boxs.IDd 
+#cell.level.BivData_table$nboxes_passed
+
+#calq the mean rate by category
+
+cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
+
+algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
+                        n_obs = length( unique(fileName) ),
+                         mean_pass.rate = mean(pass.rate)
+                         )
+
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
+
+algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
+                        n_obs = length(unique(fileName)),
+                        mean_pass.rate = mean(pass.rate)
+                         )
+
+#showing the table..
+#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
+```
+
+For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
+The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures (Peterson 2019).
+
+A limitation of our image analysis algorithm is that not all bivalents per cell are isolated, due to overlaping chromosomts.
+The range isolation rates per cell in this data set is `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to `r max(algorithm.stat.category$mean_pass.rate)`  _musculus^KAZ^_ female.
+However, we assume that the isolation of chromsomes within cell images is not biased.
+From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
+After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent observations remained.
+Given the large number of single chromosome observations, we assume that each of the data sets are equally representative of general patterns.
+
+A limitation of these data from cytological stains are the only chromosome which can be accruately identified is the XY in males.
+The unpaired XY chromosomes have SYCP3 signal along the chromsome axis, but it forms a structure distinct from the synapsed autosomes, which can be excluded in the curationstep.
+Since it has homologous pair in oocytes, the XX bivalent is indistinguishable from the rest of the autosomes.
+By phycisal length in Mb, the X is predictied to be the 3rd longest bivalent.
+
 # Evolution of genome wide recombination rate is reflected at the single chromosome level
@@
 
-#add subspecies to these values
+#add subspecies to these values¬
 melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
@@
 We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per bivalent.
-Because our observations of MLH1 foci are made in pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
-Additionally we can distinguish the identity of chromsomes (e.g.
+W use the term bivalent instead of chromosome because our observations of MLH1 foci are made of pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
+Additionally we can not fully distinguish the identity of chromsomes (e.g.
 Chromosome 1 or Chromosome 2), for these reasons we use the 'term bivalent'.
@@
 Ninety six percent of the pooled dataset of single bivalents (n=`r total.biv.all`) have either one or two crossovers (Figure X).
-The proportion of 1CO:2CO distinguishes the high and low rec strains (Figure X).
+The proportion of 1CO:2CO distinguishes the high and low recombining strains (Figure X).
 This confirms the intuitive interpretation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents.
-In the high rec strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
+In the high recombiningstrains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
 
@@
 
-# Single Bivalent Level Results
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#sort by category, 
-#compare total boxes and rate passed
-#Curated_BivData  -- pass rate?
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-#use these two metrics
-#cell.level.BivData_table$boxs.IDd 
-#cell.level.BivData_table$nboxes_passed
-
-#calq the mean rate by category
-
-cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
-
-algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
-                        n_obs = length( unique(fileName) ),
-                         mean_pass.rate = mean(pass.rate)
-                         )
-
-cell.level.BivData_table <- add_sex(cell.level.BivData_table)
-
-algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
-                        n_obs = length(unique(fileName)),
-                        mean_pass.rate = mean(pass.rate)
-                         )
-
-#showing the table..
-#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
-The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures (Peterson 2019).
-A limitation of our image analysis algorithm we used is that not all bivalents per cell are isolated.
-The range isolation rates per cell in this data set is `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to
-`r max(algorithm.stat.category$mean_pass.rate)`  _musculus^KAZ^_ female.
-
-Due to overlap of chromosomes most cells do not have measurements for all bivalents, but we assume that the isolation of chromsomes within cell images is not biased.
-From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent observations remained.
-Given the large number of single chromosome observations, we assume that each of the data sets are equally representative of general patterns.
-
-Using this data set, we address two main questions 1) Which traits are sexually dimorphic? and 2) Which traits fit distinguish the high and low recombining strains for males?
-
-In our analysis our first step is to test for measurable differences, between sexes or rec.groups and is the differences are significant, our second step is to apply the same sets of models applied to the mean MLH1 foci counts to test / describe the effects due to subspecies or strains.
+Using this data set of single bivalents, we focus on aspects of the recombiation landscape to address two main questions 1) Which traits are sexually dimorphic?
+and 2) Which traits fit distinguish the high and low recombining strains for males?
+
+In our analysis our first step is to test for measurable differences, between sexes or recobination groups in males and if the differences are significant, our second step is to apply the same sets of models applied to the mean MLH1 foci counts to test for effects due to subspecies or strains.
 
@@
 In many mammalian species the SC lengths are longer in females, yet most all of these observations come from species with female biased heterochiasmy.
-Our panel of house mice provides an opportunity to test if the sex differences in chromatin compaction (SC length) are reversed in the high rec strains (with male-biased heterochiasmy; _musculus^PWD^_  and _molossinus^MSM^_).
+Our sample of house mice provides an opportunity to test if the sex differences in chromatin compaction (SC length) are reversed in the high recombining strains (espicially the high recombining strains with male-biased heterochiasmy; _musculus^PWD^_  and _molossinus^MSM^_).
 A simple model might predict if SC length is the strongest predictor for number of MLH1 foci per cell males of the high recombining strains would have longer SC length leading to the difference in total MLH1 foci per cell.
 
-To account for confounding effects of sex chromosomes on measures of chromatin compaction (SC area) we use a reduced bivalent data set made up of bivalents with SC lengths below the 2nd quartile for SC length from a single cell to compare across sexes.
-This results in a dataset of the shortest 4 to 5 chromosomes and excludes the X bivalent in the oocytes, which is predicted to be the third longest chromosome by Mb.
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we use a reduced bivalent data set made up of bivalents with SC lengths below the 2nd quartile for SC length from a single cell to compare across sexes.
+This results in a dataset of the shortest 4 to 5 chromosomes and excludes the X bivalent in the oocytes.
 A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
@@
 
-All strains have longer mouse mean for short bivalents in females (Figure X) and all but one strain, SKIVE, had significant differences.
-The difference, (F:M), across strains ranges from `r Q1SC.mus.av.short.biv_DIFF_MSM` in MSM to `r Q1SC.mus.av.short.biv_DIFF_WSB` in WSB.
-
-<(t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
-the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively.>
+All but one strains have significantly longer mouse mean for short bivalents in females (t.test; p <0.05).
+The differences, (F:M), across strains ranges from `r Q1SC.mus.av.short.biv_DIFF_MSM` in _musculus^MSM^_ to `r Q1SC.mus.av.short.biv_DIFF_WSB` in _domesticus^WSB^_.
+While the female value is longer, it is not significantly different from males in _musculus^SKIVE^_ (t.test; p = `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)`).
+
+<t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively>
 
@@
 
-Tot.SC_mouse.av_Dom <- ggplot(data= totsc.dom_tot.sc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.dom, aes(yintercept=mean.skel) )+
+Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(sex~subsp, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
 
@@
 
-Tot.SC_mouse.av_Dom
+#full dataset ---
 
@@
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
-
-Tot.SC_mouse.av_Musc
 
@@
 
-In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; _domesticues^WSB^_ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^_  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^_ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^_ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, _musculus^KAZ^_ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^_ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`
-
-<!-- Dom males are shorter than this set of high rec mice
-The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>-- we follow up with models (which focus on sex differences).
--->
-
-To test for strain and subspecies specific effects on the sex difference in SC lengths we apply the following set of models.
+In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; p < 0.05).
+To test for strain and subspecies specific effects on the sex difference in SC lengths we apply the model framework.
+
+<_domesticues^WSB^_ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^_  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^_ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^_ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, _musculus^KAZ^_ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^_ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`>
 
@@
 
+<!--
 In the mixed models for the sex differences mean **short bivalents** length as the dependent variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
@@
 In the mixed model run on the mean total SC metric all of the coefficients were significant, (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
+-->
+
+!! SUMARIZE and remove most of these pvalues!!
+
+From the linear models a general pattern that emerges is that sex is a major effect but there are also some subspeices and strain effects, indicating that the SC lengths (degree of chromatin compaction) can evolve in both males and females.
+(weaker support in terms of the p values -- but they hint at the musculus and molossinus subspecies having longer sc and WSB SC being espicailly low)
 
@@
 
+<!--
 In linear models (M2 and M3) for **mean short bivalent length**, sex is the most significant coefficient (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`) with slight significant strain and interaction effects (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively).
@@
 Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized at this level.
+-->
 
@@
 
-<SUMMARY, ttests and models indicate female have strongest effect -- longer AC-A >
-
-<In summary, females (oocytes) have different chromatin compaction patterns compared to males. There are suggestions of evolution for this trait across the domesticus and musculus subspecies.>
-
 <Q12 total SC mouse means!
@@
 
-The SC length per chromsomes can be thought of as the 'area / available real-estate for COs.
-These results suggest that this area is 'utilized' in different ways between sexes, espicially in the strains with the high rec males, which have less 'area' for crossovers compared to females, but have more crossovers per cell on average.
-
-We focus on aspects of the recombination landscape for the two major classes of bivalents; i) the placement of single foci along a single crossover bivalent and ii) the spacing of two foci on the same bivalent, or the interfocal distance (IFD).
+These combined results of chromatoin compaction and SC length, indicate that females have longer SC lengths (chromosome axis), even in strains where males have more MLH1 foci per cell.
+I the means MLH1 foci per cell and SC area per cell, are considered seperatly, it is as if the males in these strains have more crossovers (more 2CO), but generally less 'space' to place them.
+<(this suggests that the landcapes between spermatocytes and oocytes -- are different)>
+
+This leads us to focus on aspects of the recombination landscape for the two major classes of bivalents; i) the placement of single foci along a single crossover bivalent and ii) the spacing of two foci on the same bivalent, or the interfocal distance (IFD).
 
@@
 #main figure of density plots,
-
 MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
@@
 
+focus on the rec landscape of a major class of chromosomes 1CO and how these landscapes can vary between males and females.
+
 The mean normalized foci positions per mouse from `r length( ( cleanCurated_BivData %>% filter(hand.foci.count == "1") )$Obj.ID )` single crossover bivalents.
@@
 This indicates there is no general pattern of sex differences in crossover interference measured in physical (SC) units.
-(confirming deBoer and other models of crossover interference acting through physical/mechanical force).
 
@@
 Only _musculus^KAZ^_ did not have a difference in mean IFD^PER^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-These results indicate a general pattern of two foci are seperated by more area (hence stronger interference) in males.
-These results indicate a general pattern of two foci are separated by more area (hence stronger interference) in males.
 
@@
 While the general pattern in the male distributions are enrichment longer IFD^PER^.
-(however we note that there are slight differences in the IFD^PER^ distributions between the high and low rec males to be followed up in later section)
-
-THE male vs female pattern...
-the female norm IFD are centered at ~50% but they also seem to have a higher rate of short (<25%) normalized IFDs.
-There seems to be a 25% norm.IFD cutoff (in females) but some mice seem to have less of this cut off.
-Another way to describe these results is in terms of thresholds (where the distributions end) with the general pattern of this sex difference is that the IFD^PER^ distribution is pushed higher (females have mean of .5  IFD^PER^ where as males have a IFD^PER^ mean of ~.6?)
+(however we note that there are slight differences in the IFD^PER^ distributions between the high and low recombining males to be followed up in later section)
 
@@
 
-Overall There's a low amount of significant effects across the 2CO IFD measures.
-This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.
-
-Our comparison of interference across sexes supports the general patterns from the literature.
-
-IFD raw is not significantly different (t.tests and models?). indicating that the...
-
 When interference is measured in physical SC units, the differences between sexes is low / only slightly significant -leading support that the phyisical measures of interference (in SC units) is conserved within species.
@@
 
+# Q1 Summary, consequenced of sex differences in the recombination landscape
+
+Our results confirm the typical landscape (sex differences).
+
+the terminal position of single foci and the greater distance between two foci on the same chromsome will have a consequence of
+
+The genetic consequence of such a pattern is that larger 'chuncks' a greater proportion of linked sites will segregate together in males compared to females (Veller).
+(having the effect of fewer novel combinations of site pairs along the total chromosome).
+
 <!-- transition Q1 to Q2 -->
@@
 recombination landscape (Sardell Kirkpatrick).
-The next section is meant to focus on the greater aspect of variation in mean MLH1 counts per cell the high rec males males strains ( _musculus^PWD^_ , _musculus^SKIVE^_, and _molossinus^MSM^_) and the rest of the males with the main objective of this section is to test for significant correlations with the evolution of mean MLH1 foci per cell.
-
-In comparing the male specific single bivalent based metrics the first task/level is to tests for significant differences between the high and low rec groups.
+The next section is meant to focus on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( _musculus^PWD^_ , _musculus^SKIVE^_, and _molossinus^MSM^_) and the rest of the males with the main objective of this section is to test for significant correlations with the evolution of mean MLH1 foci per cell.
+
+In comparing the male specific single bivalent based metrics the first task/level is to tests for significant differences between the high and low recombining groups.
 The second being to follow up with linear models that test for effects of subspecies and strains.
@@
 
-Our prediction for this section is that the high rec strains will have longer SC lengths proportional to the increase in the mean genome wide recombination rate.
+Our prediction for this section is that the high recombiningstrains will have longer SC lengths proportional to the increase in the mean genome wide recombination rate.
 
@@
 <pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high rec strains then low rec strains.
-This is likely due to the fact that more physically longer chromosomes have 1CO in the low rec strains, which pushes the mean SC length up.
-Where as in the High rec strains, physically longer chromosomes are more likely to be in the 2CO group.
-This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high rec group.
+Mainly that 1CO are shorter in the high recombiningstrains then low recombining strains.
+This is likely due to the fact that more physically longer chromosomes have 1CO in the low recombining strains, which pushes the mean SC length up.
+Where as in the high recombiningstrains, physically longer chromosomes are more likely to be in the 2CO group.
+This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high recombininggroup.
 They have a lower probability of 2COs below a certain SC length threshold.>
@@
 
-There is a positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low rec strains (Figure X).
-While for the total data set of mean total SC, the high rec strains have significantly  (ttest; p = `r Q2tot.sc_t.test_High.Low$p.value`) when the high and low rec strains are compared within subspecies only the _molossinus_ strains have significant difference  (ttest mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`), while the _musculus_ strains do not have significant differences (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, the means for the reduced bivalent datasets, short and long bivalents, are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
+There is a positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low recombining strains (Figure X).
+While for the total data set of mean total SC, the high recombiningstrains have significantly  (ttest; p = `r Q2tot.sc_t.test_High.Low$p.value`) when the high and low recombining strains are compared within subspecies only the _molossinus_ strains have significant difference  (ttest mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`), while the _musculus_ strains do not have significant differences (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+Additionally, the means for the reduced bivalent datasets, short and long bivalents, are not significantly different between the high and low recombining strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
 
@@
 
-We follow up the general pattern of a weak signal of high rec strains having longer SC lengths by running the model framework.
+We follow up the general pattern of a weak signal of high recombiningstrains having longer SC lengths by running the model framework.
 
@@
 
+<!--
 **M3 -- tests within subspecies!!** SO the M3 model can be run within species specific data -- to test the predictions (based on the variation in mean MLH1 counts), (use all 3 metircs) **the within subspecies M3 models don't fully meet the predictions
-
 for total.Sc : there are significant Dom strain effects: in musc kaz is significant negative, several other strains have higher total SC, (MOL in the intercept is significant).
-
 short biv: Dom: Significant G effect, MUSC no significant -- strain effects
-
 long.biv:  Dom: no strain effects, MUSC no significant strain effects, MOL - not enough observations
-
 **For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
@@
 
-**THERE IS a weak signal of the high rec strains having longer SC (having evolved) longer SC (different chromatin compaction)  While the mouse mean total sc are significantly different between the two groups --- this difference --- translated to the reduced bivalent data sets (p = long  p = short)**
-
-<!-- junk chunks below -->
+**THERE IS a weak signal of the high recombiningstrains having longer SC (having evolved) longer SC (different chromatin compaction)  While the mouse mean total sc are significantly different between the two groups --- this difference --- translated to the reduced bivalent data sets (p = long  p = short)**   -->
 
@@
 
-The normalized 1CO position is not significantly different between the high and low rec groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
+The normalized 1CO position is not significantly different between the high and low recombining groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
 While there were significant strain effects for _domesticus^^WSB^_ and _molossinus^MOLF^_ (figure X), this evolution of the 1CO positioning is decoupled from the total genome wide recombination rate.
+Hence we don't follow up with the model framework.
 
@@
 
-<Previous prediction for differences in the long and short bivalent dataset.
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.>
-
 ```{r Q2.IFD.long.short.setup, echo=FALSE}
@@
 
-<write some kind of A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation.
-how crossover interference affects the evolution of genome wide recombination rate -- is prediction for the interference pattern in the high rec strains is that the high rec strains will have weaker interference (ie.
-less space between foci on the same bivalent, which allows more foci to be placed.
-GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.>
-
-Unlike total SC length and 1CO pattern, IFD metrics are significant predictors of the rapid genome wide recombination rate evolution.
-However, instead of a negative correlation between interference strength and number genome wide recombiantion rate (higher gwRR associated with weaker interference, interference strength is stronger in the high rec strains.
-
-Unexpectedly, the mouse averages for both IFD^raw^ and IFD^PER^ were significantly longer in the high rec groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of high recombining strains (gwRR evolution).
+
+However, the pattern is in the ooposite direction to our prediction.; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
+The the 2CO landscape, instead of being placed closer toegther, the two foci are spaced further apart on the same chromosome.
+
+The mouse averages for both IFD^raw^ and IFD^PER^ were significantly longer in the high recombining groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
 The pattern is confirmed with t.test comparing groups within _musculus_ and _molossinus_
-the same subepcies (add these ttests and p values.)
-(t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value`).
-
-We test the within subspecies predictions using the model framework
+the same subepcies (t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value`).
+
+insert the raw IFD t.test!!
 
@@
 Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-#summary(Q2.M1glm_IFD.raw)
+sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
 
@@
 
-In M1, the model build on the mouse averages of IFD^PER^ metrics with subspecies and strain as fixed effects, the high rec strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
-
+The pattern is strongly suported by the modeling framework, only the high recombining strains have significant effects (glm; p = ).
+
+RAW IFD M1 results `r sum.Q2.M1glm_IFD.raw`
+
+<!--
+In M1, the model build on the mouse averages of IFD^PER^ metrics with subspecies and strain as fixed effects, the high recombining strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
 In the M2 model results, _musculus^PWD^_, and _musculus^SKIVE^_, were the only significant strain effects, (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]`).
-
 Running M2 within subspecies – confirms the within subepcies predictions across strains – (no significant strain effects 
-
 Q2.glm.M2_IFD.PER_Dom
-
 Q2.glm.M2_IFD.PER_Musc
-
-Q2.glm.M2_IFD.PER_Mol
+Q2.glm.M2_IFD.PER_Mol -->
 
@@
 
-We determined that the main difference in IFD^PER^ disbribution across the high and low groups is an enrichment of IFD^PER^ observations under 30% in low rec strains.
-
-The rate of IFD^PER^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low rec strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
+We determined that the main difference in IFD^PER^ disbribution across the high and low groups is an enrichment of IFD^PER^ observations under 30% in low recombining strains.
+
+<what are the rates in IFD^Raw^?
+
+The rate of IFD^PER^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
+
+the fact that wee see significant differences in both IFD^raw^ and IFD^PER^ align with this result -- if IFD^raw^ was conserved -- the divergence is seen in the IFD^PER^ , that would indicated that the denominator of the normalized measures (SC length) whas the main driver of the difference.
+
+Instead the results indicate that the distributions of both raw and normalized measures -- are significantly different between the high and low recombining strains.
+(examining the single observations -- we see that it's due to close foci (small IFDs)) --
+
+the difference was just observed in the IFD^PER^ measures -- while
 
 <!--
+<write some kind of A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation.
+how crossover interference affects the evolution of genome wide recombination rate -- is prediction for the interference pattern in the high recombiningstrains is that the high recombiningstrains will have weaker interference (ie.
+less space between foci on the same bivalent, which allows more foci to be placed.
+GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.>
+
+<Previous prediction for differences in the long and short bivalent dataset.
+Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
+In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
+The section on IFD only looks at the 2CO bivalents.>
+
 For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs).
@@
 
+# Q2 summary
+
+in the results analysis from the Q1 questions we see that -- the female and male recombination landscapes have distinct consequences/patterns.
+The terminal placement of single foci and the stronger interference means that in males larger chunks of linked sites will segregate together compared to the female landscape (weaker interference and uniform single foci landscape).
+
+Our results focusing on the differences in male landscapes -- this pattern (of stronger interference) further emphasized/hightened/exhasterbates in the high recombining male landscape.
+
+The measures of DSB and some comparisons of SC length between high and low recombining strains suggest that the SC length have evolved to be longer in the high recombining strains, (but this signal is weak).
+It is not clear how this result connects to the result of stronger interference.
+
 # Blank1

### 2020-03-01 doc/Results.rmd
finalizing first draft

@@
 title: "MLH1 Results Outline"
-date: 
+date: '3/1/2020'
 output:
@@
 
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE}
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
 
@@
 musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.male.VAR.plot
 
 Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.male.VAR.plot
 
 Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.male.VAR.plot
+
 ```
@@
 #NS
+
 ```
@@
 
-<to better understand this varation in RR within an evolutionary framework. Make table of models>
-
-We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
-All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
-
-**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
-
-**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
-
-```{r MLH1co.lmer.MM, echo=FALSE, warning = FALSE, digits = 3, message=FALSE}
-#THIS IS THE NEW MIXED MODEL
-#should be able to use the loaded 
+```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}
+
 #DF.HetC.MixedModel.HQ
@@
 MLH1.MM_strain)
-```
+
+```
+
+ <to better understand this varation in RR within an evolutionary framework. Make table of models>
+
+We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
+All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
+
+**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
+
+**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+
+**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
 
@@
 
-```{r MLH1.glmM2, echo=FALSE, warning=FALSE,  digits = 3}
+```{r MLH1.glmM2, echo=FALSE}
 
@@
 
-```{r MLH1.glm.M3, echo=FALSE, digits = 3}
+```{r MLH1.glm.M3, echo=FALSE}
 
@@
 and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.
-
 Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in _domesticus^G^_.
-
-<The lack of a significant sex effect indicates that the evolution of mean MLH1 foci per cell is not uniform across sexes (but varies for specfic strains)>
-
-<domesticus^G^ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
-and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`>
 
@@
 <!--
+
+<domesticus^G^ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
+and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`>
+
 M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
@@
 <The greater amount of within mouse variance in females (almost and, (males have no detecable difference across strains for within mouse variance)
--- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes.>
-
+-- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes.
 The striking patterns of variation across male and females suggest that the genome wide recombination rates have distinct evolutionary trajectories.
-For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome wide recombination rate.
+For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome wide recombination rate.>
 
@@
 #melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
 melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
@@
 #add the subsps
-
 melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
@@
 
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))
+chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))+theme(legend.position="none")
 
@@
 We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per bivalent.
-W use the term bivalent instead of chromosome because our observations of MLH1 foci are made of pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
+Because our observations of MLH1 foci are made of pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
 Additionally we can not fully distinguish the identity of chromsomes (e.g.
-Chromosome 1 or Chromosome 2), for these reasons we use the 'term bivalent'.
+Chromosome 1 or Chromosome 2), for these reasons we use the term 'bivalent' instead of chromosome.
 
@@
 
-<Indicateing, SUMMARY the TRANSITION to single bivalen measures>
-
 Using this data set of single bivalents, we focus on aspects of the recombiation landscape to address two main questions 1) Which traits are sexually dimorphic?
@@
 <!-- breif background on longer female SC  (mouse, human) Gruhn-->
-
-<chromatin compaction background, process within cells - compaction ratio within cells>
 
@@
 
-<t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
-the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively>
+<!-- t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
+the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively -->
 
@@
 
-```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE, fig.width = 4, fig.height=2.5}
+```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
+# fig.width = 4, fig.height=2.5
 
@@
 #totSC.strain.av.table %>% filter
-totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
-
-totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
+#totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
+#totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
 
@@
 
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area")
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area") + theme(legend.position="none")
+
+Tot.SC_mouse.av_ALL
 
@@
 
-Tot.SC_mouse.av_Mol
 ```
@@
 
-<_domesticues^WSB^_ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^_  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^_ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^_ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, _musculus^KAZ^_ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^_ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`>
-
-**M1. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$
-
-**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+<!-- _domesticues^WSB^_ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^_  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^_ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^_ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, _musculus^KAZ^_ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^_ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` -->
 
@@
 <!--
+**M1. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$
+
+**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
+
 In the mixed models for the sex differences mean **short bivalents** length as the dependent variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
@@
 
-!! SUMARIZE and remove most of these pvalues!!
+**!! SUMARIZE and remove most of the models pvalues!!**
 
@@
 Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized at this level.
--->
-
-<!--
+
 (subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
@@
 
-#Q1. 1CO Position Sex Differences
+# Q1. 1CO Position Sex Differences
 
@@
 
-**M1. Mixed Model**
-
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear model**
-
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
-
-**M3. Linear model**
-
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
 ```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
@@
 For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant (LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
-
 The linear models were used to detect if specific strains differed in the sex differences in the 1CO landscape.
@@
 
-<!-- Ignore this section / kind of results for know
+<!--
+**M1. Mixed Model**
+$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear model**
+$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
+
+**M3. Linear model**
+$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
+
+Ignore this section / kind of results for know
 Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite).
@@
 
-#Q1. Sex Differences in CO Interference (IFD)
+# Q1. Sex Differences in CO Interference (IFD)
 
@@
 
-<!-- lmer/ glms-->
-
-Since the most significant t.test were significantly different across sexes for IFD^PER^ and we follow up with the model framework to.
-
-To test for strain specific effects we applied the model framework
-
-**M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
+<!-- lmer/ glms  **M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
@@
 $$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
+-->
 
@@
 
+Since the most significant t.test were significantly different across sexes for IFD^PER^ and we follow up with the model framework.
 For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD.
-
-(IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`) subspecies, interaction are slightly significant. <ONLY REPORT FOR PER MEASURES>
-
+(IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`) subspecies, interaction are slightly significant.
+<ONLY REPORT FOR PER MEASURES>.
 For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
-SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure (p = male-SKIVE p =  )
-
+SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure (p = male-SKIVE p =  ).
 When interference is measured in physical SC units, the differences between sexes is low / only slightly significant -leading support that the phyisical measures of interference (in SC units) is conserved within species.
-
 While when the variance in chromatin compaction is controlled males have stronger interference (like genetic maps) the impact on the rec landscapes being.
@@
 Our results confirm the typical landscape (sex differences).
-
-the terminal position of single foci and the greater distance between two foci on the same chromsome will have a consequence of
-
+The terminal position of single foci and the greater distance between two foci on the same chromsome will have a consequence of ...
 The genetic consequence of such a pattern is that larger 'chuncks' a greater proportion of linked sites will segregate together in males compared to females (Veller).
@@
 They have a lower probability of 2COs below a certain SC length threshold.>
+
 ```{r Q2.SC.short.long.setup, echo=FALSE}
@@
 
-<!-- models -->
-
-We follow up the general pattern of a weak signal of high recombiningstrains having longer SC lengths by running the model framework.
-
+<!-- models 
 **Linear model M1 ** $$mouse \ average \ SC metric ~=~ subsp * strain + \varepsilon $$
-
-**Linear model M2 ** $$mouse \ average \ SC metric ~=~ strain + \varepsilon $$
+**Linear model M2 ** $$mouse \ average \ SC metric ~=~ strain + \varepsilon $$-->
 
@@
 
-For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= _Musculus_ `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and _Molossinus_ p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)
-
+We follow up the general pattern of a weak signal of high recombiningstrains having longer SC lengths by running the model framework.
+For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= _Musculus_ `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and _Molossinus_ p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
+(in the models using the short and long reduced datasets, (significant effects are /range from musc and strain MSM, (glm; p < 0.05).
 In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= _Musculus_ `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and _Molossinus_ p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-
+While the models for the reduced dta set suggest that the subspecies with the high recombining strains have evolved longer SC lengths -- But the differences in // but the longer SC lengths do ot cleanily predict the higher recombining strains (there are longer SC in ).
+To some extent the SC lenth evolution is decoupled from the mean MLH1 foci per cell (since some strains have longer sc - but 'lower' recombination rates)
+
+<!--
 long bivalent set in M1 -- 
-
 Musc `r sum.glmQ2_long.biv_M1$coefficients[,4][2]` 
-
 MSM strain `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`
-
 <And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
 
-<!--
 **M3 -- tests within subspecies!!** SO the M3 model can be run within species specific data -- to test the predictions (based on the variation in mean MLH1 counts), (use all 3 metircs) **the within subspecies M3 models don't fully meet the predictions
@@
 
-```
+#ABS
+
+t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
+
+```
+
+Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of high recombining strains (gwRR evolution).
+
+However, the pattern is in the ooposite direction to our prediction.; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
+The the 2CO landscape, instead of being placed closer toegther, the two foci are spaced further apart on the same chromosome.
+
+The mouse averages for both IFD^raw^ and IFD^PER^ were significantly longer in the high recombining groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The pattern is confirmed with t.test comparing groups within _musculus_ and _molossinus_
+the same subepcies (t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value` for IFD^PER^ and 
+`r t.test_mouse.av.IFD_ABS_MUSC$p.value`  `r t.test_mouse.av.IFD_ABS_MOL$p.value` for IFD^raw^).
+
+**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
+
+**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
+
+```{r Q2.IFD.glm.test, echo=FALSE}
+#REPLACE Male.poly with ..
+#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
+
+#M1.ABS
+Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
+sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
+
+#M1.PER
+Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
+
+sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
+#summary(Q2.M1glm_IFD.PER)
+
+#M2.raw
+Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
+
+Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+#within subspecies
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M2_IFD_Mol)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M2_IFD.PER_Mol)
+
+```
+
+The greater IFD measures in high recombing strains is confirmed by fiting both the IFD^raw^ and IFD^PER^ into (M2 and M3, glm; p < 0.05).
+We determined that the main difference in IFD^PER^ disbribution across the high and low groups is an enrichment of IFD^PER^ observations under 30% in low recombining strains.
+The rate of IFD^PER^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
+If IFD^raw^ was conserved while IFD^PER^ is has evolution this would indicated that the denominator of the normalized measures (SC length) would be the main driver of the difference.
+Instead the results indicate that the distributions of both raw and normalized measures are significantly different between the high and low recombining strains.
 
@@
 
-#already made
-#subset
 male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
@@
 
-Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of high recombining strains (gwRR evolution).
-
-However, the pattern is in the ooposite direction to our prediction.; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
-The the 2CO landscape, instead of being placed closer toegther, the two foci are spaced further apart on the same chromosome.
-
-The mouse averages for both IFD^raw^ and IFD^PER^ were significantly longer in the high recombining groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
-The pattern is confirmed with t.test comparing groups within _musculus_ and _molossinus_
-the same subepcies (t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value`).
-
-insert the raw IFD t.test!!
-
-**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
-
-**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
-
-```{r Q2.IFD.glm.test, echo=FALSE}
-#REPLACE Male.poly with ..
-#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-#summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
+```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
+
+#repeating chrm models with chromsome length
+#FOR BIVALENT LEVEL MODELS
+
+Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
+
+#M.Len.ABS
+Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.raw)
+
+#M.Len.PER
+Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.PER)
+
+#M.Len.raw
+Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
+
+##M.Len PER
+Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.PER)
 #  MSM strain effect for 2CO PER not significant
 
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD_Dom)
+```
+
+```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
+
+#calculate the rates of 'close foci' (<30%)
+#examine the changes in the distribution closer.
+
+#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
+
+close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                      n.close.foci = length(which(IFD1_PER <= .3)),
+
+                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                      rate.close.foci = n.close.foci / N.obs
+)
+
+close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
+
+close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("strain"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER),
+
+                      inverse.close.foci = length(which(IFD1_PER > .3)),  
+                      n.close.foci = length(which(IFD1_PER < .3)),
+                      rate.close.foci = n.close.foci / N.obs,
+
+                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
+
+)
+#the
+
+show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
+
+#add rec.group
+close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
+
+#POOL all short and 
+#chi square test
+high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
+                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
+#p-value = 0.03, slight dif tween 3 high rec strains
+
+#don't need to run chi square for this pattern, since it's enough to
+
+```
+
+```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
+
+#Log regression all show significant prediction
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.per_bivdata)
 
 #Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD_Musc)
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Musc)
 
 #Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD.PER_Dom)
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
 
 #Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD.PER_Musc)
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
 
 #Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
-```
-
-The pattern is strongly suported by the modeling framework, only the high recombining strains have significant effects (glm; p = ).
-
-RAW IFD M1 results `r sum.Q2.M1glm_IFD.raw`
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
+
+#Full dataset
+
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+```
 
 <!--
+RAW IFD M1 results for M1  -- `r sum.Q2.M1glm_IFD.raw$coefficients[,4][6]` PWD, 
+MSM `r sum.Q2.M1glm_IFD.raw$coefficients[,4][7]`
+SKIVE `r sum.Q2.M1glm_IFD.raw$coefficients[,4][8]`
+and M2
+for PWD, `r Sum.Q2.M2glm_IFD.raw$coefficients[,4][4]`
+MSM `r Sum.Q2.M2glm_IFD.raw$coefficients[,4][5]`
+and SKIVE   `r Sum.Q2.M2glm_IFD.raw$coefficients[,4][7]`
+
 In M1, the model build on the mouse averages of IFD^PER^ metrics with subspecies and strain as fixed effects, the high recombining strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
@@
 Q2.glm.M2_IFD.PER_Mol -->
-
-```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
-
-#repeating chrm models with chromsome length
-#FOR BIVALENT LEVEL MODELS
-
-Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
-
-#M.Len.ABS
-Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.raw)
-
-#M.Len.PER
-Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.PER)
-
-#M.Len.raw
-Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
-
-##M.Len PER
-Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-```
-
-```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE}
-
-#calculate the rates of 'close foci' (<30%)
-#examine the changes in the distribution closer.
-
-#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
-
-close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                      n.close.foci = length(which(IFD1_PER <= .3)),
-
-                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                      rate.close.foci = n.close.foci / N.obs
-)
-
-close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
-
-close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("strain"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER),
-
-                      inverse.close.foci = length(which(IFD1_PER > .3)),  
-                      n.close.foci = length(which(IFD1_PER < .3)),
-                      rate.close.foci = n.close.foci / N.obs,
-
-                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
-
-)
-#the
-
-show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
-
-#add rec.group
-close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
-
-#POOL all short and 
-#chi square test
-high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
-                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
-#p-value = 0.03, slight dif tween 3 high rec strains
-
-#don't need to run chi square for this pattern, since it's enough to
-
-```
-
-```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
-
-#Log regression all show significant prediction
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-```
-
-We determined that the main difference in IFD^PER^ disbribution across the high and low groups is an enrichment of IFD^PER^ observations under 30% in low recombining strains.
-
-<what are the rates in IFD^Raw^?
-
-The rate of IFD^PER^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
-
-the fact that wee see significant differences in both IFD^raw^ and IFD^PER^ align with this result -- if IFD^raw^ was conserved -- the divergence is seen in the IFD^PER^ , that would indicated that the denominator of the normalized measures (SC length) whas the main driver of the difference.
-
-Instead the results indicate that the distributions of both raw and normalized measures -- are significantly different between the high and low recombining strains.
-(examining the single observations -- we see that it's due to close foci (small IFDs)) --
-
-the difference was just observed in the IFD^PER^ measures -- while
-
 <!--
-<write some kind of A common/logical prediction for the relationship of crossover number and interference strength is -- a negeative correlation.
-how crossover interference affects the evolution of genome wide recombination rate -- is prediction for the interference pattern in the high recombiningstrains is that the high recombiningstrains will have weaker interference (ie.
-less space between foci on the same bivalent, which allows more foci to be placed.
-GIVEN THAT THERE ARE strain effects for the SC length – we only examine the IFD^PER^ which controls for variation in SC lengths.>
 
@@
 
-<!-- final summary, BE brief
-
-BE brief,
-
-final summary, highlight the themes of 1. more variance in female traits (MLH1 foci per cell, rec landscape / interference)
-
-while the male traits -- more focused (less variance in MLH1 counts per cell, shorter chromatin comparction?  bias position, stronger interference)
-
-(should I measure the var of all traits in Q1)
-
-2. rapid evolution of genome wide rec rates can fit into this -
-
-- evolution of more COs requires more DSBs (evidence points to a earlier step which is likely the chromatin compaction (amount of axis))
-
-- the strongest single bivalent predictor/correlate of rapid gwRR for males is interferece strength in 2COs
-
-AND somewhat puzzeling result of (having stronger interference), meaning there is a greater proportion of a bivalent that seperates two foci  (2 foci push to the ends of chrms (maybe decreased centromere supression))
-
--->
-
 <This striking differences across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
@@
 
-# Q2 summary
-
-in the results analysis from the Q1 questions we see that -- the female and male recombination landscapes have distinct consequences/patterns.
-The terminal placement of single foci and the stronger interference means that in males larger chunks of linked sites will segregate together compared to the female landscape (weaker interference and uniform single foci landscape).
-
-Our results focusing on the differences in male landscapes -- this pattern (of stronger interference) further emphasized/hightened/exhasterbates in the high recombining male landscape.
-
+# Q2 Summary
+
+Our results (focusing on the rapid male specific evolution of ) strongly suggest that there was lso evolution in interference.
+However these results don't meet with some basic predictions of whow the recombination landscape (should adjust to accomidating more crossoveras across the genome.
+In the first section (comparison of landscapes across sex), we note that -- the main differences between sexes in their recombination landscapes -- is that males will have larger chunks of linked sites segregating together compared to females.
+The pattern of stronger interference in the high recombining strains -- further acentuates this consequence of the recombiation landscape.
+The differences in male landscapes -- this pattern (of stronger interference) further emphasized/hightened/exhasterbates in the high recombining male landscape.
 The measures of DSB and some comparisons of SC length between high and low recombining strains suggest that the SC length have evolved to be longer in the high recombining strains, (but this signal is weak).
@@
 
-# Blank1
-
-<!-- unused chunks -->
-
-<Below are examples of plots of SC length distributions across cells.
-The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured.
-Most plots excluded for space.
-
-Each point is a bivalent plotted by cell on the x axis.
-X's are the 4th quartile, big point is the mean and smaller black point is the median.
-I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data.
-(the extra stats are not correctly mapped)>
-
-```{r show.long.SC, echo=FALSE, eval=FALSE}
-
-female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
-
-male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
-
-```
-
-<ABOVE plots show the SC lengths for the 'long SC data set'.
-They are supposed to be the longest 4-5 SC from cells where I could get good measures.
-These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position).
-Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
-
-<HAND measure METRICS, below chunk goes with the above chunk>
-```{r long.biv.plots, echo=FALSE, eval=FALSE}
-
-show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
-
-show.the.cells_females_dom
-#PWD seems to have the biggest variance across cells
-
-almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-  scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
-almost_cells_DOM_f
-
-#### Musc female
-
-show.the.cells_females_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "female") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female Musc whole cell manual measures")
-
-almost_cells_Musc_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Musc female, almost whole cell \nMean, median, 4th Quartille ploted")
-
-### Dom Male
-show.the.cells_male_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male Dom whole cell manual measures")
-
-almost_cells_Dom_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Dom male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#MUSC male
-show.the.cells_male_musc <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "male") & (whole.cell.data_manual$subsp == "Musc"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") &(manual_summary_stats$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-   geom_point(data=manual_summary_stats[(manual_summary_stats$sex == "male") & (manual_summary_stats$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc Dom whole cell manual measures")
-
-almost_cells_Musc_m <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "male") & (almost_whole_cell_BivData$subsp == "Musc"),], aes(x=fileName, y= chromosomeLength, color=strain))+
-
-  geom_point()+ 
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
-  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "male") & (Biv_Data_mean_table$subsp == "Musc"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
-  facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Musc male, almost whole cell \nMean, median, 4th Quartille ploted")
-
-#show.the.cells_male <- ggplot(whole.cell.data_manual[whole.cell.data_manual$sex == "male",], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=mean_SC, x=fileName, col="red", size=2))+
-
-#  geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=median_SC, x=fileName, col="black", size=1.5))+
-
-#   geom_point(data=manual_summary_stats[manual_summary_stats$sex == "male",],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
-
- # facet_wrap(~strain, scales="free")+ylim(c(30,150))+ggtitle("Male whole cell manual measures")
-#show.the.cells_male
-
-#almost_whole_cell_BivData
-#almost_whole_cell_wsb <- almost_whole_cell_BivData[almost_whole_cell_BivData$strain == "WSB",]
-#almost_whole_cell_DOM <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Dom",]
-#almost_whole_cell_MUSC <- almost_whole_cell_BivData[almost_whole_cell_BivData$subsp == "Musc",]
-
-#`r data.frame(table(droplevels(real.long.bivData$category)))`.  
-```
-
-<All the plots above show the distributions of manually whole cell measured SC lengths compared to the SC length distributions from the automated bivalent data.
-It shows the amount of within cell variance across strains.
-There is a bit of variance across the SC length distributions in the PWD females.
-This data set might be noisy, given the amount of variance in the SC length distributions across cells (PWD females, WSB females).
-The DF **real.long.bivData** contains `r length(real.long.bivData$Obj.ID)` bivalent measures.
-The full data set is `r length(Curated_BivData$Obj.ID)`.
-This is the breakdown of bivalent observations by category for this long dataset are
-
-**- Try to merge this DF with the whole.cell manual measures.**  
-- Try estimating which 'loose' bivalent observations might be within the long class of bivalents.  
-in code chunk above I ran the mouse averages for the longest bivalents. (680 bivalents, from 54 mice.  10202 bivalents from 86 mice.>
-
-```{r SC.distb_manual_Auto, echo=FALSE, eval=FALSE}
-
-#calq the mouse means
-manual_summary_stats <- ddply(whole.cell.data_manual, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(SC.length, na.rm = TRUE), 3 ), nsmall=3) ),
-                        median_SC = median(SC.length, na.rm = TRUE),
-                        Qar.4 = quantile(SC.length, na.rm = TRUE)[[4]]
-#                        var_SC = format(round(   as.numeric(var(SC.length)),3), nsmall=3)
-
-)
-
-manual_summary_stats <- add_mouse(manual_summary_stats)
-manual_summary_stats <- add_strain(manual_summary_stats)
-manual_summary_stats <- add_sex(manual_summary_stats)
-manual_summary_stats <- add_subsp(manual_summary_stats)
-
-###
-#calq the mouse means
-Biv_Data_mean_table <- ddply(almost_whole_cell_BivData, c("fileName"), summarise,
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        median_SC = median(chromosomeLength),
-                        Qar.4 = quantile(chromosomeLength)[[4]],
-
-                        var_SC = format(round(   as.numeric(var(chromosomeLength)),3), nsmall=3)
-
-)
-Biv_Data_mean_table <- add_mouse(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_strain(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_sex(Biv_Data_mean_table)
-Biv_Data_mean_table <- add_subsp(Biv_Data_mean_table)
-
-#colors_of_strains
-#scale_color_manual(values=colors_of_strains)
-
-#un-order these factors
-```
-
-<!-- Whole Cell Manual Comparison
-
-```{r XX.adj_manual.whole.cell, echo=FALSE, eval=FALSE}
-
-#load and examine the the whole cell hand measures here
-
-female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
-data.frame(table(droplevels(female.whole.cell$category)))
-
-# cell is 22nov16_13nov16_MSM_f2_sp1_19 is missing a biv
-
-#cell 5aug17_14jul17_LEW_f1_sp1_11_rev has broken SC - could be used to compare broken SC
-
-data.frame(table(droplevels(female.whole.cell$category), female.whole.cell$rank))
-#missing some strains, most have 3 cells (min). MSM has 5 full cells
-
-#remove the incomplete LEW female cell
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "5aug17_14jul17_LEW_f1_sp1_11_rev",]
-
-female.whole.cell <- female.whole.cell[female.whole.cell$fileName != "22nov16_13nov16_MSM_f2_sp1_19",]
-
-#calq mean SC lengths by rank
-
-mean.SC_rank.strain <- ddply(female.whole.cell, c("strain", "rank"), summarise,
-
-                       mean.sc =   mean(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                         se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-         #   mean.SC_rank1 = mean(SC.length[which(rank == 20)], na.rm =
-
-top.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank > 15,]
-
-show.SC_by.rank_top <- ggplot(data = top.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(71,210) )+ggtitle("mean SC Length by ranks \n2SE as line")
-
-show.SC_by.rank_top
-
-bottom.biv <- mean.SC_rank.strain[mean.SC_rank.strain$rank < 6,]
-
-show.SC_by.rank_bottom <- ggplot(data = bottom.biv, aes(x=strain, y=mean.sc, color = as.factor(rank)  )) +
-    geom_pointrange(aes(ymin = mean.sc-(se.SC_length*2), ymax = mean.sc+(se.SC_length*2)), 
-                    position=position_dodge(width=0.7), 
-                    linetype='solid') +    theme_bw() + facet_wrap(~strain, scales="free")+
-          ylim(c(35,99) )+ggtitle("Bottom Bivalents mean SC Length by ranks \n2SE as line")
-
-#CALQ PERCENT TOTAL SC
-
-CELL.SC.proportions <- ddply(female.whole.cell, c("fileName"), summarise,
-
-                       total.sc.cell =   sum(SC.length, na.rm = TRUE),
-                         nbiv = length(Obj.ID),
-                       #rank.by.per
-
-                    rank1_prop = (SC.length[which(rank == 20)] / total.sc.cell),
-                    rank2_prop = (SC.length[which(rank == 19)] / total.sc.cell),
-                    rank3_prop = (SC.length[which(rank == 18)] / total.sc.cell),
-                    rank4_prop = (SC.length[which(rank == 17)] / total.sc.cell),
-                      rank5_prop = (SC.length[which(rank == 16)] / total.sc.cell)
-                       #sd.SC_length   = sd(SC.length, na.rm = TRUE),
-                      #   se.SC_length   = sd.SC_length / sqrt(nbiv)
-)
-
-#MELT ABOVE
-MELT_Biv.SC_proporion <- melt(CELL.SC.proportions[,c( "fileName","total.sc.cell","nbiv",
-              #leave out the Foci
-               "rank1_prop","rank2_prop","rank3_prop","rank4_prop","rank5_prop")],
-
-            id=c( "fileName","total.sc.cell","nbiv"))
-
-MELT_Biv.SC_proporion <- MELT_Biv.SC_proporion[(!is.na(MELT_Biv.SC_proporion$value)),]
-
-MELT_Biv.SC_proporion <- add_mouse(MELT_Biv.SC_proporion)
-MELT_Biv.SC_proporion <- add_strain(MELT_Biv.SC_proporion)
-
-show_long.sc.proportions <- ggplot(data = MELT_Biv.SC_proporion, aes(y = value, x = fileName, fill=as.factor(variable)) ) + geom_col()+ggtitle("Breakdown of total SC per cell")+ylim(c(0,1)) +theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())+facet_wrap(~strain, scales = "free")
-
-show_long.sc.proportions
-
-#original plot code 
-#ggplot(data = df1, aes(x, y, color = Group)) +
-#    geom_pointrange(aes(ymin = lb, ymax = ub), 
-#                    position=position_jitter(width=0.5), 
-#                    linetype='dotted') +
-#    theme_bw()
-```
-
-The plots above show the mean SC lengths and 2SE error bars for single bivalents which have been given within cell rank.
-
-The first plot showing the mean SC lengths by the rank (most all of these cells have 3, MSM has 5 cells (observations)).
-
-The purpose of these plots is to display the variance of single bivalents when they are assigned a within cell rank.
-For the longest bivalents, XX is predicted to be the 3rd longest (according to physical length Mb).
-
-(use the value for the 3rd bivalent to adjust the single bivalent traits for XX -- then compare to males values -- or re-run in the MM).
-
-The other figure shows of each single bivalent contributes to the total SC area.
-Each column is a cell and each color is the percent of total SC area for the longest 5 bivalents in that cell.
-on average, each of the top longest bivalents make up ~10% of the cell's total SC area.
-So for cells all 20 bivalents, of it's total SC area, 5-7% is due to a XX,
-
-- Is the difference between cell averages for males and females less that 10%?
-
-- also interesting, the pwd and msm don't have longer SC, compared to other strains.
-
-<!-- Automated BivData Comparison -->
-
-```{r XX.Automated.biv.stuff, echo=FALSE}
-
-#make automated female data to compare to
-female.Cur.Bivdata <- Curated_BivData[(Curated_BivData$sex == "female"),]
-
-female.Bivdata_comp <- female.Cur.Bivdata[ ( (female.Cur.Bivdata$strain == "WSB") | (female.Cur.Bivdata$strain == "G") | (female.Cur.Bivdata$strain == "LEW") | (female.Cur.Bivdata$strain == "PWD")| (female.Cur.Bivdata$strain == "MSM") ),]
-
-#table(female.Bivdata_comp$strain)
-
-#female.Bivdata_comp
-
-#1. calq the rate of chrm segmentation in female categories in automateted data
-
-box.rates <- ddply(female.Bivdata_comp, c("fileName"), summarise,
-
-                   Nmice = length(unique(mouse)),
-                   mx.box.num = max(boxNumber), # need to run this at the cell level
-                    N.biv  = length(Obj.ID),
-                   pass.biv = N.biv / mx.box.num
-)
-box.rates <- add_mouse(box.rates)
-box.rates <- add_strain(box.rates)
-box.rates <- add_category(box.rates)
-
-cell.pass.rates <- ggplot(data = box.rates, aes(y=pass.biv, x=mouse, color=strain))+geom_jitter()+ggtitle("Auto biv pass rate per cell")+facet_wrap(~strain, scales="free")+ylim(c(0,1.1))+ theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-
-cell.pass.rates
-
-category.box.rates <- ddply(box.rates, c("category"), summarise,
-
-                   mean.pass.rate = mean(pass.biv)
-)
-#category.box.rates
-
-#2. pull the 3rd rank biv's calculate the SC means
-#calq means by rank (focus on the top 5)
-
-```
-
-For the Automated data set, I like to measure the rate of passing bivalent per cell. The mean pass rate will be multiplied to the estimated XX mean_SC.
-
-The table above shows the number of bivalents from the same strains as in the manual whole cell data.
-The plot shows the bivalent passing rate across all of the individual cells from this female data set.
-For each strain, I'll calculate the mean bivalent passing rate (maybe I should look at the mouse levels).
-
-(some of the mice have different ranges of per cell passing rate) - given this ranges, i think the xx adjustment factor should be called on the mouse level.
-(it could even be extended to cell level -- except i don't think the XX SC length estimates wont be good.)
-
-strain.XX.adjustment.factor = per_cell_passing rate \* 1 of 20 random biv will be XX \*
-
-** It might be simplier to compare the male and female means, and test it they are greater than the whole cell proprotion of the XX in females cells.**  The XX in a whole female cell contributes ~ 7% of total SC, if the female means for a type of total SC measure are from XX.
-But I am not using 'whole cell' summaries to compare female and males.
-
-**What is the effect of an extra XX-autosome on single bivalent means?**
-
-use a permutation approach: Make a True data set to start with, same(similar) number of cells, mice and bivalents.
-Make fake data sets which sample 19 bivalents, for 'in silico' cells for males and females.
-Also Run cntrl-female data set, where 20 bivalents are sampled, but randomly.
-Run the same bivalent level summaries for each 'permuted data set'; male avSC, 19Female_avSC, and rand.20_Female_avSC.
-The difference between the rand.20 and rand.19 female -permuted data sets should indicate the influence of having an extra 'XX-autosome' in the total data set.
-
-# Blank2
-<!-- Note on Heterochiasmy Definition
-
-<!-- add outline for new XX adjust section 
-adjust the colors for the plot below (standardize)
--->
-
-I present heterochiasmy as a comparison of oocyte to spermatocyte MLH1 counts, but the sex chromosomes/bivalents complicate this comparison.
-In females the XX bivalent is indistinguishable from the autosomes.
-To the meiotic recombination machinery, it is an autosome and has a similar REC landscape.
-Whereas in spermatocytes the XY bivalent is visually distinct and any MLH1 where not included in the count).
-(I note if the and Y are paired, which they are at a high rate).
-The XY pair triggers a response to un-paired chromosomes and only has MLH1 foci within the PAR (the the tips of X and Y).
-To make a more equivalent comparison I will estimate which bivalent is the XX in oocytes, and subtract that average REC from the category average of each strain.
-
-1. Compile full-cell data from females (all 20 bivalents measured)
-2. Look at the SC length -ranked data, extract the 3rd longest estimate average REC for this bivalent, 
-3. check how variable the REC is across the 1st,2nd,4th, and 5th are.
-
-According to mouse genome website, the X is the 3rd largest chromosome by total amount of DNA (Mb).
-
-(Put the XX adjustment section here)
-
-There is now MOLF, which has female biased hetC
-3 of my Musc strains have male biased patter; SKIVE, PWD and MSM.
-1 of the musc strains has female biased heterochiasmy, KAZ.
-
-```{r mouse.specific.Qplot, echo=FALSE, message=FALSE, results='hide', warning=TRUE, include=FALSE, eval=FALSE}
-
-#REMOVE This BLOCK?
-
-#### Distributions by mouse
-#There are code chunks that make a big dfs of CO count by quality score plot. The #code has been commented out, and moved to it's own doc.
-
-#remove this block since it has it's own src/script
-
-### to integrate 'pass'/fail
-### this should be in setup data
-
-#mice that don't have female data will break the chain-loop
-#MOLF female, HMI, CZECH female, CAROLI female, Tom male
-
-#instead of adding fake data, remove the categories from list
-
-#filler_data_MOLF.f = c("BatchX", "1jan19_fake_MOLF_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_MOLF_f1_sp1_rev.tif",
-#                       "fake_MOLF_f1", "MOLF female", "MOLF","female","Musc", "20","")
-#filler_data_HMI.f = c("BatchX", "1jan19_fake_HMI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_HMI_f1_sp1_rev.tif",
-#                       "fake_HMI_f1", "HMI female", "HMI","female","Cast", "20","")
-#filler_data_CZECH.f = c("BatchX", "1jan19_fake_CZECH_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CZECH_f1_sp1_rev.tif",
-#                       "fake_CZECH_f1", "CZECH female", "CZECH","female","Musc", "20","")
-#filler_data_CAROLI.f = c("BatchX", "1jan19_fake_CAROLI_f1_sp1", "99999", "3","20","20","","","","","","1jan19_fake_CAROLI_f1_sp1_rev.tif",
-#                       "fake_CAROLI_f1", "CAROLI female", "CAROLI","female","Caroli", "20","")
-#MLH1_data_table <- rbind(MLH1_data, filler_data_MOLF.f,filler_data_HMI.f,filler_data_CZECH.f,filler_data_CAROLI.f)#error from dataset thing
-
-pass_mice <- c("28sep18_CZECH_m3","26oct18_MOLF_m1","22jun15_G_m2")
-
-#this loop makes a quant_status col for a list of mice
-for(i in 1:length(MLH1_data$mouse)){
-  MLH1_data$quant_status[i] <- ifelse( is.element(MLH1_data$mouse[i], pass_mice), "pass", NA)
-}
-
-df_list <- split(MLH1_data, (MLH1_data$category))
-#remove sections which are in category but have no data
-#this is where you remove the non-data data!
-df_list <- df_list[ -c(7,12,14,18,20,21,22,24,26,28,30) ]
-
-#7 PERC, 12 MOLF female, 
-#
-
-#df list, is a list of df's which were split() from BIG df by category
-df_list 
-o=0
-i=0
-
-MLH1_data$quant_status <- as.factor(MLH1_data$quant_status)
-
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-for(i in 1:(length(df_list))){
- # only include data with quality values, then create a ddply for specific sub-df that calqs mean by each mouse
-  df_list[[i]] <- df_list[[i]][!(is.na(df_list[[i]]$quality) | df_list[[i]]$quality==""), ]
-  sub_mouse_means <- ddply(df_list[[i]], .(mouse), summarise,
-                      mouse_nMLH1 = mean(adj_nMLH1.foci)
-  )
-# this loop is going though each row of dataframes within the df.list and assigning the mouse mean
-  for(o in 1:length(df_list[[i]]$Original.Name)){
-    df_list[[i]]$mouse_mean_MLH1[o] <- sub_mouse_means$mouse_nMLH1[ (df_list[[i]]$mouse[o] == sub_mouse_means$mouse) ]
-        }
-}
-
-plot_list = list()
-for (h in 1:length(df_list)) {
-  pf <- ggplot(data = df_list[[h]], aes(colour = factor(Batch)), width = 0.25)+ 
-    geom_jitter(aes(y= adj_nMLH1.foci, x= quality)) + 
-    geom_rect(data = df_list[[h]], aes(fill=as.factor(quant_status)), alpha=0.5, xmin=-100, xmax=100, ymin=-100, ymax=100)+
-    facet_wrap(~ mouse, drop = TRUE) + 
-    geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +scale_fill_manual(values=c(NA, "red"), breaks=NULL) +    geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-  #theme(plot.background = element_rect(fill = "green")) #theme for indiviual mice, 
-  #make an empty ggplot? for missing data
-
-  plot_list[[h]] =   pf
-}
-
-#orig (I think this is the same as above)
-#plot_list = list()
-#for (h in 1:length(df_list)) {
-#  pf <- ggplot(df_list[[h]],(aes(y= adj_nMLH1.foci, x= quality)))
-#  pf <- pf +geom_rect(data = df_list[[h]], aes(fill=quant_status), xmin=-100, xmax=100, ymin=-100, ymax=100, alpha=0.005)
-
-#    pf <- pf + geom_jitter(width = 0.25, aes(colour = factor(Batch)) )
-#  pf <- pf + geom_hline(yintercept = c( mean(df_list[[h]]$adj_nMLH1.foci) ), color="black" )
-#  pf <- pf + facet_wrap(~ mouse) + geom_hline(aes(yintercept = mouse_mean_MLH1), color="red") +
-
- #   scale_fill_manual(values=alpha(c("red"), .005), breaks=NULL) +
-
-    #theme for indiviual mice, 
-#theme(plot.background = element_rect(fill = "green"))
-#  plot_list[[h]] =   pf
-#}
-#can't change the alpha/transparency of the red rec
-
-#aes must be same length ---   #geom_point(stat = "identity") +
-# the last one is correct....?
-
-#so plot list is some wride dataframe, but indeces 1:12 in plot list are ggplots
-```
-
-```{r show.plots, echo=FALSE, message=FALSE,results='hide', warning=FALSE, include=FALSE, eval=FALSE}
-geom_hline(aes(yintercept = med, group = gr), colour = 'red')
-#no plot is made that goes in the list -- so it can't be shown
-for (gg in 1:length(plot_list) ){
-  show(plot_list[gg])
- # invisible(lapply(obj, function(x) plot(x,main="some plot")))
-}
-
-#this stops at MSM
-```
-
-The mouse specific scatter plots aren't show here because there are too bulky. These plots are in a different document.
-
-Making all of these scatter plots, allows us to look at the whole distributions of the data for each mouse.
-The distance of the red line from the black could be a indicator of slides or mice with slide specific technical noise.
-
-```{r HetC.plot22, echo=FALSE,  fig.width=10, fig.height=5, fig.caption="Sex-Specific MLH1 average", eval=FALSE}
-
-no_female <- c("HMI", "SKIVE", "CZECH", "MOLF", "CAROLI", "TOM","AST")
-
-MixedModel_CO_data <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$strain.x != "other",]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[ ! CO_data_HetC_w.old$strain.x %in% no_female, ]
-CO_data_HetC_w.old <- CO_data_HetC_w.old[!(is.na(CO_data_HetC_w.old$Nmice) | CO_data_HetC_w.old$Nmice==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[MixedModel_CO_data$age.weeks < 15,]
-
-MixedModel_CO_data <- MixedModel_CO_data[!(is.na(MixedModel_CO_data$mouse) | MixedModel_CO_data$mouse==""), ]
-
-MixedModel_CO_data <- MixedModel_CO_data[ ! MixedModel_CO_data$strain.x %in% no_female, ]
-
-kuku <- ggplot(MixedModel_CO_data, aes(y=mean_co, x=mouse, color=sex.y))+geom_point()+facet_wrap(~strain.x)
-
-bubu <- ggplot(data=CO_data_HetC_w.old)+geom_point(aes(y=mean_co, x=mouse, color=sex.x))+facet_wrap(~strain.x)
-
-#add titles
-#use multiplot
-#remove X axis
-
-#multiplot(kuku,bubu, cols=2)
-```
-
-```{r F1.by.F2, eval=FALSE}
-
-#try remaking the plot Megan suggested
-# for 2CO positions, Foci1, Position  on x and Foci 2 position on y
-
-CurBivData_2CO <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-CurBivData_2CO <- CurBivData_2CO[!(is.na(CurBivData_2CO$Foci2) | CurBivData_2CO$Foci2==""), ]
-
-#isolate 2COs
-#facet by sex and subsp
-
-F1.x.F2 <- ggplot(CurBivData_2CO, aes(x=Foci1,y=Foci2, color=strain) ) + geom_point()+ facet_wrap(~sex)+ggtitle("test plot")
-
-```
-
-<!-- Deleted notes -->
-
-```{r variance.hist, echo=FALSE, eval=FALSE}
-#figures of histograms,
-
-#make series for total dataset?
-Var.Hist.total.mice <- ggplot(DF.HetC.MixedModel, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +
-  ggtitle("Histograms of within mouse variance, all mice")
-Var.Hist.total.mice <- Var.Hist.total.mice   + facet_wrap(~ strain, scales="free")
-Var.Hist.total.mice
-
-#DF.HetC.MixedModel.HQ
-Var.Hist <- ggplot(DF.HetC.MixedModel.HQ, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  + 
-  ggtitle("Histograms of within mouse variance, 10 cell min")
-Var.Hist <- Var.Hist   + facet_wrap(~ strain, scales="free")
-
-Var.Hist.15cells <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$Ncells >14,], aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, 15 cell min")
-Var.Hist.15cells <- Var.Hist.15cells   + facet_wrap(~ strain, scales="free")
-
-Q12.var.hist.plot <- ggplot(Q12_mouse_table_w.Ages, aes(x = var, fill=sex)) + 
-  geom_histogram(binwidth = 1, alpha=.6)  +   ggtitle("Histograms of within mouse variance, Q12")
-
-Q12.var.hist.plot <- Q12.var.hist.plot   + facet_wrap(~ strain, scales="free")
-Q12.var.hist.plot
-
-#some LEW female still have v high within mouse variance (2 mice, )
-
-```
-
-  -->
-
 # References

### 2020-03-01 moreJunk.R

@@ -1,0 +1,599 @@
+
+#All of the Extra Mixed Models..
+
+
+
+
+# Blank1
+
+<!-- unused chunks -->
+  
+  <Below are examples of plots of SC length distributions across cells. The top figure shows whole cell hand measured data and the bottom shows the Automated bivData from cells with at least 15 bivalents measured. Most plots excluded for space.  
+
+Each point is a bivalent plotted by cell on the x axis. X's are the 4th quartile, big point is the mean and smaller black point is the median. I'm using these to compare the patterns of these statistics in the automated data set which is missing some bivalent data. (the extra stats are not correctly mapped)>
+  
+  ```{r show.long.SC, echo=FALSE, eval=FALSE}
+
+female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
+
+
+
+male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
+
+
+```
+
+<ABOVE plots show the SC lengths for the 'long SC data set'. They are supposed to be the longest 4-5 SC from cells where I could get good measures. These longer bivalents are useful because their patterns shouldn't be affect by chromosome size effect (which effects, CO position). Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).>
+  
+  
+  
+  <HAND measure METRICS, below chunk goes with the above chunk>
+  ```{r long.biv.plots, echo=FALSE, eval=FALSE}
+
+show.the.cells_females_dom <- ggplot(whole.cell.data_manual[(whole.cell.data_manual$sex == "female") & (whole.cell.data_manual$subsp == "Dom"),], aes(x=fileName, y= SC.length, color=strain))+geom_point()+
+  
[…]

### 2020-03-02 doc/Results.rmd
cleaning up the final draft

@@
 
-In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for double strand breaks (DSBs), DMC1, in prophase spermtaocytes.
+In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
 DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes from juvenile mice (12 to 18 days) of three low rec, ( _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_) and two high recombining strains ( _musculus^PWD^_ and _molossinus^MSM^_).
@@
 The mean number of DMC1 foci per cell is associated with evolution of the number of precursors (DSBs).
-The two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptoene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
+The two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptotene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
 However the differences in DMC1 foci were not significant for zygotene cells (later prophase) between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
@@
 
-A limitation of our image analysis algorithm is that not all bivalents per cell are isolated, due to overlaping chromosomts.
+A limitation of our image analysis algorithm is that not all bivalents per cell are isolated, due to overlapping  chromosomes.
 The range isolation rates per cell in this data set is `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to `r max(algorithm.stat.category$mean_pass.rate)`  _musculus^KAZ^_ female.
@@
 
-A limitation of these data from cytological stains are the only chromosome which can be accruately identified is the XY in males.
-The unpaired XY chromosomes have SYCP3 signal along the chromsome axis, but it forms a structure distinct from the synapsed autosomes, which can be excluded in the curationstep.
+A limitation of these data from cytological stains are the only chromosome which can be accurately identified is the XY in males.
+The unpaired XY chromosomes have SYCP3 signal along the chromosome axis, but it forms a structure distinct from the synapsed autosomes, which can be excluded in the curation step.
 Since it has homologous pair in oocytes, the XX bivalent is indistinguishable from the rest of the autosomes.
-By phycisal length in Mb, the X is predictied to be the 3rd longest bivalent.
+By physical length in Mb, the X is predicted to be the 3rd longest bivalent.
 
@@
 We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per bivalent.
-Because our observations of MLH1 foci are made of pachtyene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
-Additionally we can not fully distinguish the identity of chromsomes (e.g.
+Because our observations of MLH1 foci are made of pachytene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
+Additionally we can not fully distinguish the identity of chromosomes (e.g.
 Chromosome 1 or Chromosome 2), for these reasons we use the term 'bivalent' instead of chromosome.
@@
 This confirms the intuitive interpretation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents.
-In the high recombiningstrains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
+In the high recombining strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
 
 While the proportions are significantly different (chi-square-test; p = `r PWD.MSM.2CO.prop.test$p.value`), males _musculus^PWD^_ and _molossinus^MSM^_ have 2CO proportions more similar to each other than strains from the same subspecies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-
-Using this data set of single bivalents, we focus on aspects of the recombiation landscape to address two main questions 1) Which traits are sexually dimorphic?
+Using this data set of single bivalents, we focus on aspects of the recombination landscape to address two main questions 1) Which traits are sexually dimorphic?
 and 2) Which traits fit distinguish the high and low recombining strains for males?
 
-In our analysis our first step is to test for measurable differences, between sexes or recobination groups in males and if the differences are significant, our second step is to apply the same sets of models applied to the mean MLH1 foci counts to test for effects due to subspecies or strains.
+In our analysis our first step is to test for measurable differences, between sexes or recombination groups in males and if the differences are significant, our second step is to apply the same sets of models applied to the mean MLH1 foci counts to test for effects due to subspecies or strains.
 
@@
 
-In many mammalian species the SC lengths are longer in females, yet most all of these observations come from species with female biased heterochiasmy.
-Our sample of house mice provides an opportunity to test if the sex differences in chromatin compaction (SC length) are reversed in the high recombining strains (espicially the high recombining strains with male-biased heterochiasmy; _musculus^PWD^_  and _molossinus^MSM^_).
-A simple model might predict if SC length is the strongest predictor for number of MLH1 foci per cell males of the high recombining strains would have longer SC length leading to the difference in total MLH1 foci per cell.
+In many mammalian species the SC lengths are longer in females, yet the majority of these observations come from species with female biased heterochiasmy.
+Our sample of house mice provides an opportunity to test if the sex differences in chromatin compaction (SC length) are reversed in the high recombining strains with male-biased heterochiasmy; _musculus^PWD^_  and _molossinus^MSM^_).
+If SC length is the strongest predictor for number of MLH1 foci per cell, males of the high recombining strains would have longer SC length leading to the difference in total MLH1 foci per cell.
 
@@
 
-For an addtional metric of chromatin compaction, we tested a whole cell measure of SC length using an second image analysis algorithm which converts the SC signal (into single pixel wide skeleton) and provides accurate estimates of the summed SC length of bivalents for single cells (ref Wang).
-We apply a skeletonizing image analysis algorithm to all MLH1 meiocyte spreads images.
+For an additional  metric of chromatin compaction, we tested a whole cell measure of SC length using a second image analysis algorithm which converts the SC signal into single pixel wide trace of the SC shapes and provides accurate estimates of the summed SC length of bivalents for single cells (ref Wang).
+We apply the skeletonizing algorithm to all MLH1 meiocyte images.
 Mouse means were calculated from `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cell wide total SC measures (figure X).
@@
 
-**!! SUMARIZE and remove most of the models pvalues!!**
-
-From the linear models a general pattern that emerges is that sex is a major effect but there are also some subspeices and strain effects, indicating that the SC lengths (degree of chromatin compaction) can evolve in both males and females.
-(weaker support in terms of the p values -- but they hint at the musculus and molossinus subspecies having longer sc and WSB SC being espicailly low)
-
 ```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
@@
 
-These combined results of chromatoin compaction and SC length, indicate that females have longer SC lengths (chromosome axis), even in strains where males have more MLH1 foci per cell.
+Longer Sc compaction in females is supporte by the model framework, with sex being the most consistant significant effect (LTR and glm; p < 0.05).
+Additionally there are some significant subspecies and strain effects indicating that the SC lengths can evolve within strains and subspecies (LTR and glm; p < 0.05).
+
+These combined results of total SC per cell and reduced single bivalent SC metrics, indicate that females have longer SC lengths (chromosome axis), even in strains where males have more MLH1 foci per cell.
 I the means MLH1 foci per cell and SC area per cell, are considered seperatly, it is as if the males in these strains have more crossovers (more 2CO), but generally less 'space' to place them.
-<(this suggests that the landcapes between spermatocytes and oocytes -- are different)>
 
@@
 
-focus on the rec landscape of a major class of chromosomes 1CO and how these landscapes can vary between males and females.
+For most of the house mouse strains, 1COs are the major class of chromosomes observed across cells.
+There are known sex differences in the relative placement along chromsomes this section tests if there has been evolution in this aspect of the recombination landscape.
 
@@
 
-For the Mixed model, sex is the most significant effect (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`).The random effect of strain is also significant (LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
-The linear models were used to detect if specific strains differed in the sex differences in the 1CO landscape.
-While sex was the most significant effect for both M2 and M3 (there were some strain specific effects...)
-
-The male effects were the most significant (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-While the musc subspecies, LEW and KAZ strains had slightly signifcant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
+The differences in 1CO landscapes due to sexes is confirmed by the model framework, with sex is the most significant effect in the mixed model (M1) (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`) and the two linear models (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 
 <!--
+not addressing the strain and subspecies effects
+The random effect of strain is also significant (LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
+
+While the musc subspecies, LEW and KAZ strains had slightly significant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
+
 **M1. Mixed Model**
@@
 #real.long.bivData
-
 Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
@@
 It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
-
- We isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
-`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males to test describe sex differences in crossover interference.
+To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and
+`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males respectively.
 Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
-We examined both raw IFD^raw^ and normalized by SC length (IFD^PER^).
+We examined both raw IFD^raw^ and normalized by SC length (IFD^norm^).
 The mean IFD^raw^ measures of crossover interference as a mechanical force mediated through the SC.
 We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
-Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al ).
+Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al 2004).
 
@@
 
-While the comparisons of the IFD^PER^ metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-Also this metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+While the comparisons of the IFD^norm^ metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
 Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
 
-The difference in mean IFD^PER^ are highly significant between sexes.
-The mean IFD^PER^ are significantly long in six of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
-Only _musculus^KAZ^_ did not have a difference in mean IFD^PER^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-
-<Full dataset (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`).>
-
-<!--  finer distribution patterns -->
+The difference in mean IFD^norm^ are highly significant between sexes.
+The mean IFD^norm^ are significantly long in six of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
+Only _musculus^KAZ^_ did not have a difference in mean IFD^norm^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
+
+<!--  finer distribution patterns <Full dataset (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`).> -->
 
@@
 
-GENERAL RESTULTS DIFFERENCES between sexes
-
-In examining the IFD^PER^ distributions, we note that 
-the female norm IFD are centered at ~50% – but they also seem to have slight an enrichment of short (<25%) normalized IFDs
-
-While the general pattern in the male distributions are enrichment longer IFD^PER^.
-(however we note that there are slight differences in the IFD^PER^ distributions between the high and low recombining males to be followed up in later section)
+In examining the IFD^norm^ distributions, we note that the female IFD^norm^ observations are centered at approximately 50% and have slight an enrichment of short (<25%) IFD^norm^ observations.
+Comparitively, the IDF^norm^ distributions in males are enriched for longer IFD^norm^.
 
@@
 however the weight of observations are slightly shifted across the sexes.
-There's an enrichment of lower IFD^PER^ in females relative to males, while the opposite is true for the higher threshold -- where males have an enrichment of higher/long IFD^PER^ compared to females.
+There's an enrichment of lower IFD^norm^ in females relative to males, while the opposite is true for the higher threshold -- where males have an enrichment of higher/long IFD^norm^ compared to females.
 
 In the three high rec strains -- the distributions have a more striking difference of the lower threshold being much higher in males, aproximately 30% to 40% compared to 25% in females.
-Also while the female upper threshold is capped at 75%, in males (some strains have higher upper thresholds -- and it's more likely to observe IFD^PER^ above the threshold.
+Also while the female upper threshold is capped at 75%, in males (some strains have higher upper thresholds -- and it's more likely to observe IFD^norm^ above the threshold.
 
@@
 
-Since the most significant t.test were significantly different across sexes for IFD^PER^ and we follow up with the model framework.
-For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD.
-(IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`) subspecies, interaction are slightly significant.
-<ONLY REPORT FOR PER MEASURES>.
-For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males.
-SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure (p = male-SKIVE p =  ).
-When interference is measured in physical SC units, the differences between sexes is low / only slightly significant -leading support that the phyisical measures of interference (in SC units) is conserved within species.
-While when the variance in chromatin compaction is controlled males have stronger interference (like genetic maps) the impact on the rec landscapes being.
+The three sets of models support the result of stronger interference in males, sex was the most significant significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`).
+When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
+These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.
 
@@
 
+  For the normalized values in both M2 and M3, sex is a significant effect, increasing 
+with males nrm.IFD in males.
+  SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure (p = male-SKIVE p =  ).
+For the mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. (IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`)
+
 <For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
@@
 
-# Q1 Summary, consequenced of sex differences in the recombination landscape
-
-Our results confirm the typical landscape (sex differences).
-The terminal position of single foci and the greater distance between two foci on the same chromsome will have a consequence of ...
-The genetic consequence of such a pattern is that larger 'chuncks' a greater proportion of linked sites will segregate together in males compared to females (Veller).
-(having the effect of fewer novel combinations of site pairs along the total chromosome).
+# Q1 Summary, consequences of sex differences in the recombination landscape
+
+Our results confirm sex differences in recombination landscapes described in several other species and we note that these differences have distinct consequences on the potential patterns of genetic variation resulting from the distinct recombination landscapes.
+The terminal position of single foci and the greater distance between two foci on the same chromosome will have a consequence of larger sections of linked sites segregating together in male gametes compared to female gametes (Veller).
 
@@
 
-The results from the previous section -- confirm (known sex differences for the 'typical' 
-recombination landscape (Sardell Kirkpatrick).
-The next section is meant to focus on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( _musculus^PWD^_ , _musculus^SKIVE^_, and _molossinus^MSM^_) and the rest of the males with the main objective of this section is to test for significant correlations with the evolution of mean MLH1 foci per cell.
-
-In comparing the male specific single bivalent based metrics the first task/level is to tests for significant differences between the high and low recombining groups.
-The second being to follow up with linear models that test for effects of subspecies and strains.
-
-The literature provides basic predictions for the direction of the relationship the genome wide recombination rate and two aspects of the recombiantion landscape; SC length (positve) and crossover interference (negative).
+The next section is meant to focus on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( _musculus^PWD^_ , _musculus^SKIVE^_, and _molossinus^MSM^_) to the low recombing strains.
+The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
+
+In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and  the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
+
+**Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
+
+**Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
+
+The literature provides basic predictions for the direction of the relationship the genome wide recombination rate and two aspects of the recombination landscape; SC length (positive) and crossover interference (negative).
 Following this logic we predict i) no strain effects for SC length or crossover interference within _domesticus_, ii) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which will have longer SC and weaker interference compared to the other _musculus_ strains and iii) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_.
@@
 
-Our prediction for this section is that the high recombiningstrains will have longer SC lengths proportional to the increase in the mean genome wide recombination rate.
-
-The confounding effects of sex chromosomes are removed for this section that compares across males from different inbred strains.
+Given the differences in the DSB numbers between a subset of the high and low recombining strains we predict that there will also be differences in SC lengths proportional to the increase in the mean genome wide recombination rate.
 We compared mouse means of three SC based metrics; total SC, mean short bivalent length and mean long bivalent length.
@@
 X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)-->
-
-<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high recombiningstrains then low recombining strains.
+`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
+Mainly that 1CO are shorter in the high recombining strains then low recombining strains.
 This is likely due to the fact that more physically longer chromosomes have 1CO in the low recombining strains, which pushes the mean SC length up.
-Where as in the high recombiningstrains, physically longer chromosomes are more likely to be in the 2CO group.
+Where as in the high recombining strains, physically longer chromosomes are more likely to be in the 2CO group.
 This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high recombininggroup.
-They have a lower probability of 2COs below a certain SC length threshold.>
+They have a lower probability of 2COs below a certain SC length threshold.>  -->
 
@@
 
-There is a positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low recombining strains (Figure X).
-While for the total data set of mean total SC, the high recombiningstrains have significantly  (ttest; p = `r Q2tot.sc_t.test_High.Low$p.value`) when the high and low recombining strains are compared within subspecies only the _molossinus_ strains have significant difference  (ttest mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`), while the _musculus_ strains do not have significant differences (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, the means for the reduced bivalent datasets, short and long bivalents, are not significantly different between the high and low recombining strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low recombining strains (Figure X).
+While for the total data set of mean total SC, the high recombining strains have significantly more SC area
+(t.test; p = `r Q2tot.sc_t.test_High.Low$p.value`), only the _molossinus_ strains are significantly different between the high and low recombining strains (t.test mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
+The _musculus_ are not significantly different across the groups (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+Additionally, the means for the reduced bivalent datasets, short and long bivalents, are not significantly different between the high and low recombining strains (t.test; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
 
@@
 ```
-
-<!-- models 
-**Linear model M1 ** $$mouse \ average \ SC metric ~=~ subsp * strain + \varepsilon $$
-**Linear model M2 ** $$mouse \ average \ SC metric ~=~ strain + \varepsilon $$-->
 
@@
 
-We follow up the general pattern of a weak signal of high recombiningstrains having longer SC lengths by running the model framework.
-For total SC in M1 (glm including subsp), the two subspecies effects are significant (glm; p= _Musculus_ `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and _Molossinus_ p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
-(in the models using the short and long reduced datasets, (significant effects are /range from musc and strain MSM, (glm; p < 0.05).
-In the short bivalent data set, analyzed by M1, both subspecies are sig (glm; p= _Musculus_ `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and _Molossinus_ p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-While the models for the reduced dta set suggest that the subspecies with the high recombining strains have evolved longer SC lengths -- But the differences in // but the longer SC lengths do ot cleanily predict the higher recombining strains (there are longer SC in ).
-To some extent the SC lenth evolution is decoupled from the mean MLH1 foci per cell (since some strains have longer sc - but 'lower' recombination rates)
+For the linear model testing subspeices and strain (M1) and total SC as the dependant variable, the two subspecies effects are significant (glm; p= _Musculus_ `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and _Molossinus_ p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
+Whereas the model with reduced bivalent means as the dependant variables, a number of subspeicies and strains effects reach significance, (glm; p < 0.05) they are not consistant across the sets of models tested, indecateing to some extent the chromation compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
 
 <!--
+In the short bivalent data set, analyzed by M1, both subspecies are 
+ (glm; p= _Musculus_ `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and _Molossinus_ p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
+While the models for the reduced data set suggest that the subspecies with the high recombining strains have evolved longer SC lengths --
 long bivalent set in M1 -- 
@@
 
-The normalized 1CO position is not significantly different between the high and low recombining groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
+The normalized 1CO position is not significantly different between the high and low recombining groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t.test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
 While there were significant strain effects for _domesticus^^WSB^_ and _molossinus^MOLF^_ (figure X), this evolution of the 1CO positioning is decoupled from the total genome wide recombination rate.
@@
 
-Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of high recombining strains (gwRR evolution).
-
-However, the pattern is in the ooposite direction to our prediction.; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
-The the 2CO landscape, instead of being placed closer toegther, the two foci are spaced further apart on the same chromosome.
-
-The mouse averages for both IFD^raw^ and IFD^PER^ were significantly longer in the high recombining groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^PER^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
-The pattern is confirmed with t.test comparing groups within _musculus_ and _molossinus_
-the same subepcies (t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value` for IFD^PER^ and 
-`r t.test_mouse.av.IFD_ABS_MUSC$p.value`  `r t.test_mouse.av.IFD_ABS_MOL$p.value` for IFD^raw^).
-
-**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
-
-**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$
+Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of evolution in genome wide recombination rates.
+However, the pattern is in the opposite direction to our prediction; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
+In the 2CO landscape, instead of observing weaker interference strength and two foci placed closer together on a single chromosomes, the two foci are spaced further apart.
+
+The mouse averages for both IFD^raw^ and IFD^norm^ were significantly longer in the high recombining groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^norm^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
+The pattern is confirmed with t.test comparing strains within _musculus_ and _molossinus_ (t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value` for IFD^norm^ and `r t.test_mouse.av.IFD_ABS_MUSC$p.value`  `r t.test_mouse.av.IFD_ABS_MOL$p.value` for IFD^raw^).
 
@@
 
-The greater IFD measures in high recombing strains is confirmed by fiting both the IFD^raw^ and IFD^PER^ into (M2 and M3, glm; p < 0.05).
-We determined that the main difference in IFD^PER^ disbribution across the high and low groups is an enrichment of IFD^PER^ observations under 30% in low recombining strains.
-The rate of IFD^PER^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
-If IFD^raw^ was conserved while IFD^PER^ is has evolution this would indicated that the denominator of the normalized measures (SC length) would be the main driver of the difference.
-Instead the results indicate that the distributions of both raw and normalized measures are significantly different between the high and low recombining strains.
+The greater IFD measures in high recombing strains is confirmed by fitting both the IFD^raw^ and IFD^norm^ into the linear model framework, where only the strain effects for the high recombining strains had significant p values (glm; p < 0.05).
+The similar patterns across the raw and normalized IFD values rule out differences in IFD^raw^ measures, SC lengths and chromosome size as the major driver of this pattern.
+We determined that the main difference in IFDPER distribution across the high and low groups is an enrichment of IFD^norm^ observations under 30% in low recombining strains.
+The rate of IFD^norm^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
+
+<!--
+**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
+
+**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$ -->
 
@@
 
-In M1, the model build on the mouse averages of IFD^PER^ metrics with subspecies and strain as fixed effects, the high recombining strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
+In M1, the model build on the mouse averages of IFD^norm^ metrics with subspecies and strain as fixed effects, the high recombining strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
 In the M2 model results, _musculus^PWD^_, and _musculus^SKIVE^_, were the only significant strain effects, (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]`).
@@
 - all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
-With each unit increase in SC length, the IFD^PER^ also increases (for the high rec strains + MOLF).
-All of the high rec strains and MOLF have significantly longer IFD^PER^.
-There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^PER^ decreases slightly.
+With each unit increase in SC length, the IFD^norm^ also increases (for the high rec strains + MOLF).
+All of the high rec strains and MOLF have significantly longer IFD^norm^.
+There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^norm^ decreases slightly.
 -->
@@
 
-Our results (focusing on the rapid male specific evolution of ) strongly suggest that there was lso evolution in interference.
-However these results don't meet with some basic predictions of whow the recombination landscape (should adjust to accomidating more crossoveras across the genome.
-In the first section (comparison of landscapes across sex), we note that -- the main differences between sexes in their recombination landscapes -- is that males will have larger chunks of linked sites segregating together compared to females.
-The pattern of stronger interference in the high recombining strains -- further acentuates this consequence of the recombiation landscape.
-The differences in male landscapes -- this pattern (of stronger interference) further emphasized/hightened/exhasterbates in the high recombining male landscape.
-The measures of DSB and some comparisons of SC length between high and low recombining strains suggest that the SC length have evolved to be longer in the high recombining strains, (but this signal is weak).
-It is not clear how this result connects to the result of stronger interference.
+Our results show that the greater crossover interference is the strongest single bivalent-based predictor for the observed rapid evolution of mean MLH1 foci per cell.
+While these results do not conform initial predictions for how a recombination landscapes would accommodate more crossovers, the increased strength of interference aligns with our results on sex differences.
+The typical recombination landscapes for males and females results in divergence in the proportion of linked sites along chromosomes which segregate together.
+The stronger interference of the 2CO bivalents in the high recombining strains accentuates this effect.
+The measures of DSB and some comparisons of SC length between high and low recombining strains suggest that the SC length have evolved to be longer in high recombining strains, however this evolution of SC length is partially decoupled from the number of crossovers since similar amounts of SC length evolution are seen in low recombining strains.
 

### 2020-03-02 doc/Results.rmd
getting hang of incorporating ref, use BibTex!

@@
 title: "MLH1 Results Outline"
-date: '3/1/2020'
+date:
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document: default
   html_document:
@@
 <!--
-RW's main qtl paper used the within animal variance,
-
-@article{wang2017genetics,
-  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
-  author={Wang, Richard J and Payseur, Bret A},
-  journal={Genetics},
-  volume={206},
-  number={4},
-  pages={1841--1852},
-  year={2017},
-  publisher={Genetics Soc America}
-}
+
 references for phenotypic variance connected to robustness to environment, increase in developmental consistency
@@
 For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
-The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures (Peterson 2019).
+The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures [@peterson_2019].
 
@@
 
-For an additional  metric of chromatin compaction, we tested a whole cell measure of SC length using a second image analysis algorithm which converts the SC signal into single pixel wide trace of the SC shapes and provides accurate estimates of the summed SC length of bivalents for single cells (ref Wang).
+For an additional  metric of chromatin compaction, we tested a whole cell measure of SC length using a second image analysis algorithm which converts the SC signal into single pixel wide trace of the SC shapes and provides accurate estimates of the summed SC length of bivalents for single cells [@wang_2019_sc].
 We apply the skeletonizing algorithm to all MLH1 meiocyte images.
@@
 
-In all strains, the 1CO landscape had significant differences between males and female which followed the "typical landscape" of sex difference observed in many other species (Sardell Kirkpatrick).
+In all strains, the 1CO landscape had significant differences between males and female which followed the "typical landscape" of sex difference observed in many other species [@sardell_sex_2020].
 The foci position was more medially places in females while males have more terminal normalized foci positions (ttest; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
@@
 We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
-Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al 2004).
+Previous measures found no sex difference in measures of IFD^raw^ in mice [@de_boer_two_2006].
 

### 2020-03-04 doc/MatMeth.Rmd
note to self
notes for results, more work on Methods

@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
   html_document:
@@
     toc_float: yes
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
 bibliography: refs_MatMeth.bib
@@
 ```
+
+Todo
+
+Add the single Bivalent predictions to one of the sections.
 

### 2020-03-04 doc/Results.rmd
notes for results, more work on Methods

@@
 
+nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
+
+raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
 ###look at why KAZ males are weird
@@
 In the M2 model results, _musculus^PWD^_, and _musculus^SKIVE^_, were the only significant strain effects, (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]`).
-Running M2 within subspecies – confirms the within subepcies predictions across strains – (no significant strain effects 
+Running M2 within subspecies confirms the within subepcies predictions across strains – (no significant strain effects 
 Q2.glm.M2_IFD.PER_Dom

### 2020-03-04 doc/MatMeth.Rmd

@@
 
-Todo
-
-Add the single Bivalent predictions to one of the sections.
-
 # Mouse Husbandry
@@
 To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
-The Mus musculus strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
+The _Mus musculus_ strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
 The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
@@
 The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
+We use abriviated superscript notation for the remainder of the paper (e.g.
+_musculus^WSB^_).
 
- All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.
-A breeding colony of wild derived *Mus musculus* mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
+ All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
+A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
  Mice were fed on dry standard breeder chow.
@@
 Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison.
-
 Over the course of data collection some breeding colonies  mice were moved from facilities.
 Additionally the GI strain was kept at a separate facility.
-(evidence for environmental (room) effect on (MLH1 / SC lengths / a variety of meiotic traits was tested for.
-Effects on average MLH1 counts were compared -- for strains which were kept at mulitple institutions.
-((We observed no effect, Supplemental figure))
+We tested for effects on mean MLH1 counts across difference mouse rooms/facilities and found no effect (Supplemental section).
 
@@
 The majority of mice used were between 5 and 12 weeks, (supplemental table).
-However some strains had problems breeding, so testes from older mice were collected.
-We found a (small/no) age effect on MLH1 counts for male mice (Supplemental Figure)).
+for DMC1 spreads we targeted juvinile male mice (X ages).
 
-The only difference between preparation of spermatocyte and oocyte spreads was the volumns of the hypotonic buffer was 300 u l instead of 3ml.
+Because some strains have breeding issues, all mice some strains were kept in breeding pairs (before collecting, and aged out of the desired age range.
+We performed tests for effects on MLH1 counts and found no effects in the older male mice ages.
+(Also we tested other effects, including maternal age...)
 
-The vast majority of oocyte data was collected from neonate litters/mice between 5 to 48 hours old.
-This approach maintain breeding pairing and still result in prophase oocytes (cite timeline of oocytes in neonates).
-(differences were assessed between embryo samples and neonates?).
-
+The vast majority of oocyte data was collected from neonate mice between 5 to 48 hours old.
+This approach maintain breeding pairing and still result in prophase oocytes <(cite timeline of oocytes in neonates)>.
+.
 Precise staging of embryos via copulary plugs was not feasible in many of these wild derived strains due to their behavior.
 Embryonic samples were collected when pregnancy noted in females, embryos were staged based on (X table markers).
+No effect on mlh1 count due to age effects (pooled as embryo and neonate) was found (Supplemental section).
 
-Meiocyte spreads were made following [@peters_1997].
-For testes the tunica was removed and whole testes was incubated in 3ml of hypotonic solution for 45min.
-For ovaries after dissection, the pair of ovaries were decapsulated in cold PBS and both were incubated in 300ul of hypotonic solution for 45 min.
+Meiocyte spreads were made following [@peters_1997].  whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
 
-After incubation, gonads were transferred to sucrose solution -for mastication to make a cell slurry which was transferred to 80ul of 2% PFA solution.
+After incubation, gonads were transferred to sucrose solution for mastication.
+15ul of cell slurry was transferred to 80ul of 2% PFA solution.
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
@@
 
-Staining / Immunohistochemistry
+**Staining / Immunohistochemistry**
 
 SYCP3 was used to mark  visualize the Axial element of the synaptonemal complex.
-Sicne all mouse chromosomes are telocentric/acrocentric, CREST was used to indicate the centromere -- indicate the side of the bivalent with the centromere.
-
-the genome wide recombination rate was charaecterized / estimated with staining of MLH1 antibody, 
-The same staining protocol was applied to spermatocyte and oocyte spreads.
-The staining protocol was based on X with some adaptations, and same as previously described in [@peterson_2019].
+Sicne all mouse chromosomes are telocentric/acrocentric, CREST was used to indicate the centromere <insert the concentrations>
 
 Precursurs to COs, DSBs were quantified with the DMC1 staning antibody.
-Due to resource constraints -- only a subset of strains were stained for this marker.
-Only spermatocyte spreads produced enough good cells for DMC1 quantification.
 
@@
 
+```{r error.repeatbility, echo=FALSE}
+
+```
+
 Images of cells were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
-
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
 
-For MLH1 pachytene characterization, cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
+For MLH1 pachytene quantification, cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
 Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
@@
 
-Hand measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
-
 The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson_2019] respectively.
 
-Foci counts/numbers were manually scored.
+(the data from these scripts differed slightly in that the total sc script returned the skeletonized SC area for the whole cell. <state the script parameters, (morphogy (3) and other parameter)>
+
+to quantify / charaecterize the SC (length) (synapsed chromosomes, )-- two image analysis algorithms were used both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
+
+The DNA CrossOver script isolated single chromosomes shapes from an image, in addition to returning the SC length -- also measured the location of (green and blue signal along the single SC - reflecting MLH1 foci and cetromere signal respectively).
+
+Hand measures of SC/pactyene / meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
 
@@
 
-DMC1 -- Ttest
-
-MLH1 counts per cell are good /decent proxies for genome wide recombination rates estimates (cite).
-
 The distributions of CO counts per cell was assessed for normality (supplemental figure).
@@
 
-The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R). these mouse level summary statistics were analyzed -- within our model frameworks.
+The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).
+
+**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
+
+**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+
+**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
 
@@
 
-We can also compare the size of the sex effects between (subspecies) and within (strain) mouse groups.
-(ie is the male or female strain effect larger?
-is there more variance due to random strain effect in male or females?)
-(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)
-I may have to do more research/reading on testing the random effects.
-
-### Single bivalet quant
-
-Single bivalent chromosomes were isolated from cytology cell images as discribed in (X ). the single bivalents were 'curated' (removed errors in the segmentation algorithm)
-
-to account for effects of CO number and physical chromosome size -- the data was also analyzed after spliting up by chromosome class and in long and short chromosome bins.
-
-- Split up 'classes' of bivalents depending on the number of foci -- to tests/ comparisons more comprable.
+### Single bivalent charaecterization
 
@@
 
-- whole cell bivalent measures (measures of all bivalents ), were compiled by mixed automated bivalent measures with hand measurements. (Since the error in automated measures was so low)
+The single bivalent data set was curated of algorithm mistakes manually>.
+For charaecterizing any of the foci positions or interfocal distances any normalized refers to dividing the raw measures by the total SC for that invidiualy bivalent.
 
-**SC length**
+The proportions of different chromosome classes (seperated by number of foci) was tested with chi-sqaure tests. <in R using the X package/function>
 
-The axial element (AE) of the SC -- for single bivalents (synapsesd homolog pairs), was quantified as the skeletonized (single pixel width line) --
+**chromosome size effects**
 
--comparison to human measures previously in permyscus species -- and in musculuc (shown in the supplement)
+Using the single bivalent dataset, the within cell distribution of bivalent SC lengths from cells with at least 17 isolated chromosome measures.
+The longest and shortest 4 to 5 chromosomes within each cell were identified by qualqulating the fourth and second quartille.
+(these comprised the short and long bivalent datasets).
 
@@
 
-**CO interference**
+< manual measures of whole cell bivalent measures (measures of all bivalents were incorporated into this data set.
+), were compiled by mixed automated bivalent measures with hand measurements.
+(Since the error in automated measures was so low)>
 
-- how we choose to measure interference, the average distance between2 foci on same bivalent.
+**CO Interferecne**
 
-- downside to using the gamma distribution -- not the primary meteric for CO interference.
-
-- Also pooling intervals for categories and fitting a gamma distribution (in supplement)
-
-**rec landscape** **1CO normalized position**
-
-- how why measure the recombination landscape -- coohse to charaecterize the rec landscaoe bby isloating the 1CO bivalents; since there are known sexual dimorphic features of this and since bivalents with more than 1 foci will be primarily influenced by CO interference.
-We look at both the raw and
-
-the distance to the telomere and centromere end of the SC shape were measured -- for all / pooled bivalents
-
--specific R packages
--non-parametric  (tests and measures for dealing with count data)
--hypothesis tested (size, DSB number)
-
--Mixed models and glm models were built
-
--tests for fixed effects
--tests of random effects in mixed models was LTR of exactRLRT function in X package.
-
-single bivalent analysis
-
-# Numbers of mice cells ect for each method?
-
-above belongs more in results
-
-**outline**
-
-1. framework for analyzing the MLH1 evolutionary patterns across mus species complex. (what is the main question we were tackling when designing this framework?)
-
-2. write equation, describe terms (proxies for within and between polymorphism)
-logic / choices for fixed and random effects
-
-3. what are the 2 general / basic questions this framework can be applied to.
-
-$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
-
-4. how can DMC1 measures be added to the mixed model
-
-5. descriptions of mlh1 outside of framework (missing female data)
+Our primary mode of examining interference is to compared the interfocal disatnces (IFDs) (raw and normalized by SC length.)
+Gamma parameter v has historically been used to charaecterize interference and was calculated using <x function>.
 

### 2020-03-06 src/BivData_Setup.Rmd
fixed a bug in the IFD calculation for 3CO

@@
 
+```{r neg.IFD, echo=FALSE}
+
+#fixed a bug for the multi CO IFD
+
+#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+#remove IFD
+#Curated_BivData <- add_IFD(Curated_BivData)
+#t.test.4CO <- Curated_BivData %>% filter(Obj.ID == "8feb17_4jan17_LEW_f1_sp1_3_rev_17")
+
+#neg.IFD <- Curated_BivData[Curated_BivData$IFD1_ABS < 0, ]
+#neg.ifd <- Curated_BivData %>% filter(hand.foci.count > 1) %>% filter(IFD3_ABS < 0 )
+
+```
+
 Main files from this Script:
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_19.2.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_6.3.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-03-06 src/CommonFunc_MLH1repo.R

@@ -389,25 +389,27 @@ add_SisCoTen <- function(DF){
 #Calculate InterFocal Distance (IFD) in Chrm1 dataframe
 #requires Percent foci positions, for the sceiont part
 
+#t.test.4CO, for the 3IFD  #fixed bug
+
 add_IFD <- function(DF){
   
+  DF$IFD3_ABS <- ifelse(  (DF$hand.foci.count >= 4),
+                          #manual.verified.f1, 
+                          as.numeric(DF$Foci4) - as.numeric(DF$Foci3),   "" )
+  #need to include the IFD2
+  DF$IFD3_ABS <- as.numeric(DF$IFD3_ABS)
+
+    
   DF$IFD1_ABS <- ifelse(  (DF$hand.foci.count >= 2),
                           #manual.verified.f1, 
                           as.numeric(DF$Foci2) - as.numeric(DF$Foci1),   "" )
   DF$IFD1_ABS <- as.numeric(DF$IFD1_ABS)
   
   
-  DF$IFD2_ABS <- ifelse(  (DF$hand.foci.count == 3),
+  DF$IFD2_ABS <- ifelse(  (DF$hand.foci.count >= 3),
                           #manual.verified.f1, 
                           as.numeric(DF$Foci3) - as.numeric(DF$Foci2),   "" )
   DF$IFD2_ABS <- as.numeric(DF$IFD2_ABS)
-  
-  
-  DF$IFD3_ABS <- ifelse(  (DF$hand.foci.count == 4),
-                          #manual.verified.f1, 
-                          as.numeric(DF$Foci3) - as.numeric(DF$Foci4),   "" )
-  DF$IFD3_ABS <- as.numeric(DF$IFD3_ABS)
-  
   
  # DF$IFD1_PER <- ifelse(  (DF$hand.foci.count >= 2), 
   #                        as.numeric(DF$Foci2.PER) - as.numeric(DF$Foci1.PER),   "" )

### 2020-03-06 doc/Results.rmd
looking at gamma interference results

@@
 
-load(file="~./MLH1repo/data/CleanBivData_19.2.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
 
@@
 
+```{r Q1.IFD_gamma, echo=FALSE}
+library(EnvStats)
+#cleanCurated_BivData
+
+#all.female.multi <- (cleanCurated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count > 1) %>%
+#                       filter( !is.na( IFD1_ABS) ))  #$IFD1_ABS
+
+#this MELT BIVDATA should be good to go
+MELT.IFD.BIVDATA <- melt(cleanCurated_BivData[,c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
+                            "Obj.ID","hand.foci.count",
+                            "mouse","category","strain", "sex", "Rec.group",
+                            "IFD1_ABS","IFD2_ABS","IFD3_ABS")],
+
+             id=c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
+                            "Obj.ID","hand.foci.count",
+                            "mouse","category","strain", "sex", "Rec.group"))
+#rename the cols and clean up
+MELT.IFD.BIVDATA <- MELT.IFD.BIVDATA[(!is.na(MELT.IFD.BIVDATA$value)),]
+
+colnames(MELT.IFD.BIVDATA) <- c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
+                            "Obj.ID","hand.foci.count",
+                            "mouse","category","strain", "sex", "Rec.group", 
+                               "whichIFD", "IFDval")
+
+MELT.IFD.BIVDATA$Nrm.IFD <- MELT.IFD.BIVDATA$IFDval / MELT.IFD.BIVDATA$chromosomeLength
+
+#START isolating all the types of IFDs
+
+#all.female.multi$mycol <- apply(all.female.multi[,c('IFD1_ABS','IFD2_ABS','IFD3_ABS')], 1, function(x) x[!is.na(x)][1])
+
+all.female.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "female")#1697
+
+all.male.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "male")#1392
+
+#gamma fits
+FULL.female.gamma <- egamma( all.female.IFDs$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.8
+
+#gamma fits
+FULL.male.gamma <- egamma( all.male.IFDs$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#shape  5.4
+
+#male, stronger interference
+
+WSB.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.1, 3.73 (raw, PER) (n=252)
+
+WSB.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.6, 5.10 (raw, PER) (n=186)
+#try the short and long bivalents dataset
+
+WSB.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3, 3.237 (raw, PER) (n=60)
+
+#gamma fits
+WSB.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.5,  5.1 (raw, PER) (n= 109)
+
+WSB.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.9, 3.9 (raw, PER)  (n=107)
+#there are 0 3CO and 2  4CO bivalents -- this causes a greater difference in the PER based value
+
+```
+
+```{r gamma.IFD, echo=FALSE}
+
+female.gamma <- egamma( (all.female.IFDs )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+f2CO.female.gamma <- egamma( (all.female.IFDs  %>% filter(hand.foci.count == 2)  )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+f3CO.female.gamma <- egamma( (all.female.IFDs %>%  filter(hand.foci.count == 3)  )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+f4CO.female.gamma <- egamma( (all.female.IFDs %>%  filter(hand.foci.count == 4)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+
+PWD.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.8, 4.844 (raw, nrm)  (n = 357)
+PWD.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3.8, 6.746 (raw, nrm) (n = 285)
+PWD.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#3, 3.628 (n = 72)
+
+KAZ.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+
+KAZ.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+
+KAZ.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+#####
+SKIVE.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+SKIVE.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+SKIVE.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+##
+MSM.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+MSM.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+MSM.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+MSM.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+######
+G.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+G.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+G.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+G.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+######
+LEW.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+LEW.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+LEW.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+LEW.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+######
+WSB.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+WSB.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+WSB.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+WSB.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+```
+
+```{r Q1.male.IFD, echo=FALSE}
+
+#gamma fits
+male.gamma <- egamma( (all.male.IFDs )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+m2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+m3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE,         ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+#gamma fits
+PWD.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#8.1, 11.74 (n=318) (lots of variation when I apply sample size of 18)
+
+PWD.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#10.3, 16.962, (n=300)
+
+PWD.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#2.9, 3.4, (n=18)
+
+####
+KAZ.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+KAZ.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+KAZ.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+SKIVE.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+SKIVE.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+SKIVE.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+MSM.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+MSM.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE,     ci.type = "two-sided", ci.method = "normal.approx",
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+MSM.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+MOLF.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+MOLF.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+MOLF.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+G.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+G.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+G.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+LEW.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+LEW.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+LEW.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+###
+WSB.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+#
+WSB.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
+                            ci.type = "two-sided", ci.method = "normal.approx", 
+                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
+
+```
+
 <!--

### 2020-03-06 doc/Results.rmd
trying to find pattern for rbar and gamma estimates

@@
 #MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
+
+```
+
+```{r Q1.rbar, echo=FALSE}
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
+
+#read.csv of gamma estimates from above
+gamma_estimates <- read.csv("~./MLH1repo/results/gamma_estimates.csv", header = TRUE)
+
+table_gamma_estimates <- read.csv("~./MLH1repo/table.csv", header = TRUE)
+
+for(e in 1:length(cleanCurated_BivData$fileName)){
+#  mini.IFD <- calq.intra.rbar(mini.IFD)
+  cleanCurated_BivData$rbar[e] <- new.calq.intra.rbar(cleanCurated_BivData[e,])
+  }
+
+#calq the mouse / strain means for rbar
+
+summarize_rbar_table <- ddply(.data=cleanCurated_BivData,
+                 .(sex, strain),
+                 summarize,
+                 mean.rbar.all = mean(rbar, na.rm = TRUE),
+
+                 mean.rbar.2CO = mean(rbar[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean.rbar.3CO = mean(rbar[which(hand.foci.count == 3)], na.rm = TRUE)
+)
+#the 3 is negative...
+
+#how best to merge rbar and gamma estiamtes (which estimate)
+
+raw.gamma_estimates <-  gamma_estimates %>% filter(type.IFD == "raw.IFD") %>% filter( (class == "2CO") | (class == "3CO") )
+
+table <- raw.gamma_estimates
+write.table(table, "~./MLH1repo/table.csv", sep=",",row.names = TRUE)
+
+write.table(summarize_rbar_table, "~./MLH1repo/RBAR.table.csv", sep=",",row.names = TRUE)
+
+#what comparison of IFD and rbar would be logical
+#plot rbar and gamma
+ll <- ggplot(table_gamma_estimates, aes(y=mean.rbar, x = all.biv.gamma, color = class, shape = sex ))+geom_point()+facet_wrap(~strain)+ ggtitle("")+ylim(c(0,.5))
+
+#mean rbars are up around 
+#SKIVE male have extra (very strong interference)
+#there's so much variance across the gamma estimates -- 
+#most all females seem to have stronger interfernce
+#there don't seem to be consistant patterns
 

### 2020-03-06 src/CommonFunc_MLH1repo.R

@@ -444,27 +444,117 @@ addIFD.2 <- function(DF){
 }
 
 
+add_umeter <- function(dat){
+  #how to convert dat$chrm.length to chrm.length
+  #sc unit
+  dat$chrm.length.um  <- dat$chromosomeLength / 9.8152
+  return(dat)
+}
+
+
+
+#calculate Interchromsomal rbar
+#tested with values from Veller et al 2018
+#Chrm row should be full DF
+
+calq.intra.rbar <- function(chrm_row){
+  o = 1
+  # rrbar <- c()
+  for(o in 1:length(chrm_row$fileName) ){
+    #print(chrm_row$fileName[o])
+    if(grepl(0, chrm_row$handFoci[o])){
+      #print("fist has 0 CO")
+      #rrbar[o] <- NA
+      rrbar <- 0
+    }
+    if(grepl(1, chrm_row$handFoci[o])){  #do the block here
+      #print("fist has 1 CO")
+      #rrbar[o] <- "1CO"
+      portion1 <- chrm_row$F1[o]
+      portion2  <- chrm_row$chromosomeLength[o] - portion1
+      portion.sum <- portion1+portion2
+      rel.portion1 <- (portion1 / portion.sum)
+      rel.portion2 <- (portion2 / portion.sum)
+      rel.por.sqrd1 <- rel.portion1^2
+      rel.por.sqrd2 <- rel.portion2^2
+      chrm.sq <- sum(rel.por.sqrd1, rel.por.sqrd2)
+      
+      #rbar[o] <- 0.5*(1-chrm.sq)
+      #print(chrm_row$rbar[o])
+      rrbar <- 0.5*(1-chrm.sq)
+      #o = o +1
+    }
+    if(grepl(2, chrm_row$handFoci[o])){  #do the block here
+      #print("this has 2 CO")
+      portion1 <- chrm_row$F1[o]
+      portion2  <- chrm_row$F2[o] - chrm_row$F1[o]
+      portion3 <- chrm_row$chromosomeLength[o] - chrm_row$F2[o]+1
+      portion.sum <- portion1+portion2+portion3 #in 2CO portions 
+      rel.portion1 <- (portion1 / portion.sum)
+      rel.portion2 <- (portion2 / portion.sum)
+      rel.portion3 <- (portion3 / portion.sum)
+      rel.por.sqrd1 <- rel.portion1^2
+      rel.por.sqrd2 <- rel.portion2^2
[…]

### 2020-03-06 doc/Results.rmd
adding more interference notes

@@
 table <- raw.gamma_estimates
-write.table(table, "~./MLH1repo/table.csv", sep=",",row.names = TRUE)
-
-write.table(summarize_rbar_table, "~./MLH1repo/RBAR.table.csv", sep=",",row.names = TRUE)
+#write.table(table, "~./MLH1repo/table.csv", sep=",",row.names = TRUE)
+
+#write.table(summarize_rbar_table, "~./MLH1repo/RBAR.table.csv", sep=",",row.names = TRUE)
 
@@
 #there don't seem to be consistant patterns
+
+#this plot shows much higher 2CO interference (via the gamma estimat from nrmIFD) in high rec males
+#across other classes of bivalents
+
+#most all other estiamtes of gamma estimates are around 4
+#
+
+#house mouse raw nu is around 11
 

### 2020-03-10 doc/Discussion.rmd
kinda notes for writing/drafting

---
title: "Discussion"
author: "April Peterson"
date: "March 10, 2020"
output: html_document
---

# Conserved Pattern

- COnserved heterochiasmy paterns
    - typical landscape

    - DNA loops and Interference strength

    -

# Evolved Pattern

- DSBs

- interference strength

# Explanation

# Future Steps

### 2020-03-11 doc/Discussion.rmd
slow notes progress for discussion

@@
 
-# Conserved Pattern
+# Conserved Sexually Dimorphic Features of Recombination landscape
 
-- COnserved heterochiasmy paterns
-    - typical landscape
+Conserved heterochiasmy paterns
 
-    - DNA loops and Interference strength
+  - (the canonical pattern has too many exceptions to be the uniform pattern)
 
-    -
+  - our results have exception, and other exceptions are...
+
+- (heterochiasmy refers to different numbers of chiasmata - and recently has been extended to number of crossovers -- this view / charaecterization is slightly incomplete, since it considers just one cell stage and 1 value (a count))
+
+We argue that the sex differences (broaden the definition of heterochiasmy or use a different word) -- which includes more aspects that show conserved sexes differences
+
+Our results indicate, sexes have distinct evolutionary trajectories for genome wide recombinaiotn rate (and landscape) in males nd females).
+
+Conserved features of the landscape may play a role in influencing/keeping the sex specific trajectories.
+In our examination of the panel of mouse strains, we note several features of the recobmination landscape with conserved sexual dirmophism pattern.
+That is the sex within sex these features show no evolution and consistant pattern of variation.
+
+## Typical landscape
+<review the typical landscape, one of the major broad scale patterns,
+exceptions
+consequences, (differences in variation, Sardell paper) >
+
+   1.
+male telocentric and female uniform landscape / general placement of crossovers.
+What is referecened to as typical landscape (Sardell) -- see references sardell kirkpatrick for list of species which show/display this typical phenotype
+
+A difference in the general placement (general recombination landscape), with males having telomere bias and females having more uniform placement of crossovers along chromosomes.
+
+-Exception to landscape (marisupials, birds?)
+
+## Chromatin Structure
+
+< review loop-axis relationship,  review how interference is transmitter throught the axis  -- transformation of genetic area -- results in longer loops having stronger interference,
+
+OUR pattern is that
+
+>
+
+  2. Explain the connection between interference strength in physical space and (abstracted space)
+
+  it boils down to, the relationship with axis length and DNA loop size -- 
+  --females have shorter loops - longer axis and weaker interference (in abstracted sense)
+
+  Chromatin structure -- the size / length of meiotic chromosomes
+
+  (exceptions, birds (fish?) plants?) -- examining this is hard due to chromosome size effecs
+
+  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes
+
+2. Short axis / longer DNA loops will have stronger interference
+This prediction combines the findings of chromatin compaction and interference differences which are intertwined by the transmission of interference along the chromosome axis and the inverse relationship chromatin compaction (axis length and DNA loop sizes).
+In physical units (micron) the counting model fits interference well to the 2D linear space of the axis these measures of interference are conserved across sexes in mouse.
+This is because the counting units are the BASES of DNA loops which is conserved along the 2D length of chromatin axis.
+When this space is transformed to account for all of the DNA base pairs which are sequestered into the DNA loops above the axis, the strength of interference will vary depending the on underlying chromatin structure.
+The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
+Thus the strength of interference will be stronger because with each step (count) along the 2D axis,
+more bases pairs are being skipped above the axis.
+
+## Within mouse variance
+
+ mouse and humans show greater between cell variance 
+ this metric isn't really focused on (not many studies that examine inidivial level and both sexes)
+3. Within mouse variance differences
+In our results we observe more between cell variance in females (across all strains).
+This pattern has been observed in humans (Lynn et al).
+(But results of reporting observations of single meiosis in both sexes).
+
+ Our results confirm many of the typical result fir the sex difference in the recombination landscape.
+
+ For many years the canoncial pattern of heterochiasmy has been dicribed (with linkage maps) as females have more genome wide recombination rates compared to males of the sames speces (ref older hetC refs)
+
+more recntly more nuanced patterns have also been noted/discribed in greater detail; general/broad scale landscape patterns male telomeric and female uniform, and chromatin compaction (in some species)
 
@@
 
+2.Predictors of Recombination rate evolution within (males)
+This section focuses on the largest axis of variation over a short time scale.
+We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.
+
 - DSBs
 
-- interference strength
+A. DSB Differences and moderate support for SC length differences
+1. More DSBs in the high rec strains
+2.
+The conserved ratios of DSB : CO implies that the increase in DSB numbers are accompanied by proportional increase in crossovers.
+This is evidence that shifts support away from the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
+3. DSB number being driven by the number of chromatin loops which is positively correlated with axis length.
+4. These results supported by the moderate support of longer SC length (pachytene stage in the high rec males).
+(Review Baier et al 2014)
+
+B.  INTERFERENCE
+1. Review results
+-The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.
+-Caveats:  chromosome size effects, 
+However, the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).
+
+-Also:
+	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD.
+	-The difference in interference patterns is driven by the difference in the rates of small IFD observations.
+
+2.  Review the assumptions for measuring CO interference in different ways.
+The logical model: Crossover interference is a mechanism of suppressing crossovers, negative relationship between interference strength and gwRR.
+More crossovers can fit along the chromosomes and results in more crossovers overall.
+Examples of empirical results supporting this pattern;  Otto Payseur, (Ruiz-Herrera? Bomblies?)
+Quantifing using the gamma distribution: 
+Caveats,  This doesn't make since for the general recombination landscape:   i) finite chromosome length and ii) limited range of crossovers per chromosomes (1 to 4).
+Alternative interference characterization:  i) the number of crossovers are held constant and ii) transform / generalize the rec landscape, (focus on how the rec landscape would change instead of the genome wide rate) iii) assume crossover assurance.
+	-The Goldstein model, models predicts a positive correlation with interference strength and higher recombination rate.
+A modifier which changes the probability of the recombination landscape having 0 crossovers to 2 crossovers, (driven by crossover assurance).
+
+-Examples of empirical results which show this result / pattern
+	-(cattle) comparison WITHIN sexes
+	- Preliminary results for F2 cross (Hannah's work), (PWD vs CAST)
+	(maybe more once I start looking closer)
+
+4. HYPER-Crossover experiments. (review the effects when crossover number is artificially altered)
 
@@
 
+_Models_
+
+1.Gamete selection
+The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.
+2.SACE
+Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.
+3.Two-locus (protect for drivers)
+Asymmetric division of egg opens them up to meiotic drivers.
+Recombination modifiers for sex specific female landscape will decrease the chances of driving centromeres to segregate to the egg by increasing the number of crossovers and placing them across a larger area of the genome (uniform placement).
+4.Difference in acentromeric spindle and SAC
+A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle).
+(Our hypothesis is that the tension in a centrosome spindle is stronger (or more uniform) since all MT-KT are anchored at single points across cells (the two centrosomes).
+This will also make prediction for the spindle assembly checkpoint (SAC), which is stronger/stricter in smaller gametes with centrosomes.
+(the SAC is more sensitive to achiasmate bivalents on the spindle / better at detecting a lack of tension).
+Whereas in the acentrosome spindle is more diffuse (across a larger area) and the tension across bivalents are anchored by multiple MTOCs, resulting in weaker tension force or less uniform strengths across all the bivalents within the cells.
+The SAC in eggs seems to be weaker/leakier, that is more achiasmate bivalents are required to trigger the SAC (stop the division).
+
+Review of modifier / popgen models
+Indirect forces common assumptions themes across all the modifier models.
+Review of cell biology
+Look at conserved features of gametogenesis that distinguish male and females
+Different Centrosome spindle
+Asymmetrical division
+
+5. Model prediction
+The models above which predict directional selection on the recombination landscape in males: SACE and Spindle differences are most parsimonious for the models review in this paper.
+The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.
+
+Consequences of evolution in the number and placement of crossovers / evolution of interference strength:
+Spindle / SAC model: The number and placement of crossovers changes the resulting amount of sister cohesin involved the tension signal, which affects the tetrad/chiasmata structure at MI when the homologs are pulled apart (Hollis et al 2020, Lee et al 2019).
+This model predicts a difference in the meiotic spindles or division mechanism between high and low rec strains.
+ii. SACE: For the 2CO in the high rec males, larger blocks of chromosomes are kept together in the next generation.
+	This model would predict greater reproductive variance for males.
+
 # Future Steps

### 2020-03-12 doc/Discussion.rmd
draft work

@@
 
-- DSBs
+Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength.
 
-A. DSB Differences and moderate support for SC length differences
+## DSBs
+
+Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al). Likely // evidence suggests this pattern is also a feature of the sex differences.
+
 1. More DSBs in the high rec strains
@@
 This is evidence that shifts support away from the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
+Together these results indicate that (it's not the CO : NCO decision but an earlier step (when the total number of chromatin loops (the substrate for SPO11, are laid down with the establishment of the meiotic chromosome structure (DNA loop size)
+
+The tethered loop model of DSB formation
+
 3. DSB number being driven by the number of chromatin loops which is positively correlated with axis length.
@@
 
-B.  INTERFERENCE
+_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
+during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
+
+## INTERFERENCE evolution
+
 1. Review results
 -The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.
--Caveats:  chromosome size effects, 
-However, the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).
+
+These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol.
+(it also goes against the logical assumptions models of fitting more crossovers along the SC chromosome axis space)
+
+Yet (we are confident that this pattern is not due to errors).
+
+ the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).
 
 -Also:
-	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD.
-	-The difference in interference patterns is driven by the difference in the rates of small IFD observations.
+	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.
+
+	-The difference in interference patterns is driven by the difference in the rates of small IFD observations, (mean vs median). a pattern that is suggestive of selection agisnt short IFDs
+
+There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
+breifly, the model simulates / sets up a scnario where a modifer acts through the probabilty of flanking regions having 0 or 2 crossovers.
+ith crossover assuance, recombination rate increases the further apart the loci are spaced (?)
+
+Additionally there are a handful of empircal results which also show the unexpected pssotive correlation between gwRR and interference strength.
+
+1.
+Cattle (Ma?), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)
+
+2. The IFDs from (Beth's cross) a related house mouse cross show that the PWD (same strain, different mice) had longer IFD compared to Cast (HVR unpublished data). (also gamma?)
+
+3. In Peromyscus, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@Peterson]
+
+<note about the differences in the patterns>  assumptions with gamma dstribution and linkage maps, (closer to infinte genetic space), less weight on crossover assurance - more weigth on 
+(pooled SC/chromosome length gives strong correlation )
+
+<which is much different than when a single chromosome is examined (by itself), finite 2D length, crossover assurance, weaker correlation with ((the closer to chromosome level for SC legnth metric / summary, the weaker the correlation with crossover number))
 
@@
 
-# Explanation
+# Explanations and Model testing
+
+This section refers to the matrix of models and results from this study where we (conceptually test if each model would support our results)
+
+We were motivated to run a comparison of
 
@@
 # Future Steps
+
+be aware of sex-specific and sex average biases
+
+- linkage / R between two foci is sex averaged
+
+- Celegans -- cytlogy focuses on oogenesis  and genetic maps are often from hermaphrodites (selfing -- sperm and egg landscapes are merged in hemaphordites)
+
+-- the value of Chromosome level measures -- incorporate chromosome behavior into models
+
+-  more attention to cell physiology
+
+-- lack of SC length prophase cell charaecterization
+
+-- comparisons within and between sexes (that cattle paper)

### 2020-03-13 doc/Discussion.rmd
notes from Wu thesis

@@
 
-## Typical landscape
+## male telomeric, female centromeric Typical landscape
 <review the typical landscape, one of the major broad scale patterns,
@@
 
-## Chromatin Structure
+The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes.
+we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalents.
+)
 
-< review loop-axis relationship,  review how interference is transmitter throught the axis  -- transformation of genetic area -- results in longer loops having stronger interference,
-
-OUR pattern is that
-
->
+## Chromatin Structure and interference strength
 
   2. Explain the connection between interference strength in physical space and (abstracted space)
+
+Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.
 
@@
 
-1.Gamete selection
+## 1.Gamete selection
 The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.
-2.SACE
+
+<go through the squares>
+
+this is a modifer model it abstracts the genome area.
+
+This model predicts the evolution of heterchiasmy -- through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
+
+this model doesn't predict that the heterochismy direction would reverse.
+
+##2.SACE
 Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.
-3.Two-locus (protect for drivers)
+
+##3.Two-locus (protect for drivers)
 Asymmetric division of egg opens them up to meiotic drivers.
 Recombination modifiers for sex specific female landscape will decrease the chances of driving centromeres to segregate to the egg by increasing the number of crossovers and placing them across a larger area of the genome (uniform placement).
-4.Difference in acentromeric spindle and SAC
+
+## 4. Sindle   // Difference in acentromeric spindle and SAC
 A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle).
@@
 
+<schuh ellenberg 2007 spindle
+
+_review the correct SAC_
+
+bivalent orientation early metaphase -- search and capture model -- when there are mulitple KT-MT attachmments, (lateral (on the side or not on centromere/KT proper), merotelic MT attachemtn from ooposite poles) and amplitelic (MT from same pole)
+then biorientation process starts, --search and capature process (only amplictic Mt-KT attachments get stablize)
+
+Aurora -- detects bad connections,
+
+_evidence that SAC in oogenesis is leaky_
+
+(Lane et al 2012)
+
+whenn going through the attachment correection, bivalents are thought to be rotated via pulling from MT
+
+Akera 2017 -- chromosomes at different ends of spindle have different densites (in reference to the cortical or central side of spindle (PB or egg))
+
+Wu thesis -- chrm 17 with asymetric centromere str, cenromeres with large major stat drove to (polar body (opposite of drive) (lack of re-orientation, once it ended up on that side))
+
+the biased rotation happened before spindle migratgation (from center of egg to near cortex)
+
+when a aurora inhibitor applied -- the biased rotation of chrm 17 went away and anaphase moved up (by passed SAC as in Lane 2010)
+
+(k-fibre the bundle MTs connected to KT Iwata-otsubo 2017) -- larger KT would provide more k0fiber binding sites
+
+the spindle migration (from center of egg to cortext depends on actin (Li albertini 2013))
+cytoplasmic streaming when spindle moves close to the cortex (yi et al 2013)
+
+the KT and centromere seperate when MT's start pulling (trains start moving) (they does seperate whn MT inhibitors are applied)
+
+Coretical KTs undergo greater tension than centrak KTs.
+
 Review of modifier / popgen models
@@
 Asymmetrical division
+
+aMTOC (Cliff and Schuh 2015, Schuh and ellenburg)
+ since there are multiple aMTOC (not 2 centrioles) -- the MTOCs across the sides of the cells could be unven.
+
+after NEBD the aMTOCs are fragmented and coalest into two main clusters (Clift and Schuh 2015 and Luksza 2013). 
+(Kolano et al 2012 maintence of spindle)
+aMTOC migration to pools is dynamic and self organized.
+
+Wu thesis compared th density of spindles on the central and cortical halves
+data suggests that more MTOC are recruited to the cortical side
 

### 2020-03-14 doc/Discussion.rmd
notes for outline draft

@@
 ---
+
+# Differences across models
+
+Explain the model differences;
+
+## 1.can it be applied to broad scale chromosome level landscape?
+
+-genome wide recombination rates are made of the chromosome level patterns
+
+This is important because general patterns have been observed for the recombination landscape at the chromosome level (also small range of crossovers per chromosome and broad scale pattern (Haenel et al ))
+
+## 2. Does the model account for evolution?
+
+given all the variance observed and fast evolution, models have to have some engine for evolution.
+
+(some source of affecting fitness, or shifting the optima)
 
@@
   - our results have exception, and other exceptions are...
+
+These are the metrics for the _conserved sex differences_
 
@@
 The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes.
-we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalents.
-)
+we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalent
+s. )
 
 ## Chromatin Structure and interference strength
+
+1. SC length  2. connection to interference
+
+general pattern for the sexual dimorphsim the relationship between axis length and CO number
+-- axis are longer in females than male, suggesting suport more crossovers (more DSBs)
+but the density is less in females.
+
+Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
+Two strains have have more crossovers in males, but females still have longer sc axis.
+_This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape_
+
+The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov)
 
@@
   it boils down to, the relationship with axis length and DNA loop size --
-  --females have shorter loops - longer axis and weaker interference (in abstracted sense)
+
+  --females have shorter loops - longer axis and weaker interference
 
   Chromatin structure -- the size / length of meiotic chromosomes
+
+  same pattern in humans (Gruhn et al 2013)
 
@@
 
- mouse and humans show greater between cell variance 
- this metric isn't really focused on (not many studies that examine inidivial level and both sexes)
+- Our results confirm previous findings of females having greater between cell variation (for indiviual females compared to males)
+the results -- indicate there is more than can be due to technical noise (for oocytes vs spermatocytes spreads)
+also all ovary collections were from early/ young mice -- not due to the maternal age effects, (which has to do with the stability of bivalents at metaphase -- after sexual maturity and ovulation (ref))
+
+We did not observe
+
+- Because these are inbred strains raised in common / standard lab environment, (an single levels)
+
+_sources of within animal variation_
+
+-age differences
+
+new paper has revealed within nucleus variation for (CO number -- potentially via the chromatin remodeling) (Veller Kleckner)
+
+prev reported sex diffferences in
+
+ Gruhn et al 2013 (the favorite figures)
+
+ Gruhn et al 2016 -- not due to the developmental stage of the oocytes (ovulation) (the production line)
+
+ mouse and humans show greater between cell variance -- this metric isn't really focused on (not many studies that examine inidivial level and both sexes)
+
 3. Within mouse variance differences
@@
 
- Our results confirm many of the typical result fir the sex difference in the recombination landscape.
+(references, more within animal variance in
+(number of crossovers per cell)
+but maybe, rec landscapes in general)
 
- For many years the canoncial pattern of heterochiasmy has been dicribed (with linkage maps) as females have more genome wide recombination rates compared to males of the sames speces (ref older hetC refs)
+(Gruhn et al), Kong et al?
 
-more recntly more nuanced patterns have also been noted/discribed in greater detail; general/broad scale landscape patterns male telomeric and female uniform, and chromatin compaction (in some species)
+cattle,
+
+- age effects?
+
+- Hassold hunt
+
+_Relaxed selection_ leads to
 
@@
 
+Our results find evolution between a subset of strains -- demonstrating that the high rec strains have greater DSB number
+
 Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al). Likely // evidence suggests this pattern is also a feature of the sex differences.
@@
 1. More DSBs in the high rec strains
+
 2.
@@
 4. These results supported by the moderate support of longer SC length (pachytene stage in the high rec males).
+
+_the resuts support / replicate those reported in Baier et al_
+
+a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
+
+-different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures
+
+our results are similar in several ways
+- correlation of final MLH1 and DBS counts, SC measues correlate with the DSB number (even tho the differnces we observe in our study are -- slight / less significant -- we attribute that to chromosome specific noise )
+
+the correlation of DSB marker with MLH1
+
 (Review Baier et al 2014)
@@
 
+-- this model is an extension of the modifer model presented in Lenormand
+
+-- (basics)
+ extends the lenomand focusing on two conditions.
+first that the stength of epistatsis in dependant on the cis or trans nature.
+the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)
+
+ second that the relative interference strengths differnt between male and female.  This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
+
+This new model has _strengths_
+
+  - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)
+
+ - incorporates know cellular / molecular mechanisms with a very common pattern sexual dimorphic recombination rate landscape.
+
+- extends modifer model to explain a chromosome based / level pattern
+
+_weaknesses_, and ways the model diverges from our results
+
+ - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
+
+- Other ways this model departs from our results,
+
+-Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
+
+However / while not explicitaly arameterized or predicted in this model,
+ -- this model predicts larger chuncks of chromosomes will segregate together
+ (as outlined in Veller et al)
+
+ our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle
+
 ##3.Two-locus (protect for drivers)
+
 Asymmetric division of egg opens them up to meiotic drivers.
@@
 
-## 4. Sindle   // Difference in acentromeric spindle and SAC
+_model basics_
+
+'break up driv system consiracies'
+
+_strengths_
+
+a clear fitness effect (the drive)
+
+explains female landscape pattern in reference to the centromere (presummed location of drivers), yet the model doesn't scale down to chromsome level (such as to explain, chromosome size differences)
+
+_weaknesses_
+(modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
+
+## 4. Spindle   // Difference in acentromeric spindle and SAC
+
 A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle).
@@
 
-<schuh ellenberg 2007 spindle
+we like this model because it can be applied to _both our results_ for i) sexual dimorphism variation and ii) the fast male evolution in gwRR and recombination landscape.
+
+for the distinct _evolutionary patterns_ selection on the recombination landscape via the spindle (segregation checkpoint),
+
+_main sex differences in spindle_ are the anchoring of MT at opposite poles
+
+in males there are centromsomes for Mi and MII divisions
+where are in oocytes, a (universal) conserved feature of oogenesis in many species is that they loose centrosomes -- instead the MTs are anchored at mulitple aMTOCs.
+
+There is potential for these aMTOC to be unevenly distributd across the cortical and central side of the spindles (Wu et al, Schuh and ellenberg) --- (could be an origin of different tension / pulling )
+
+The survalence of SAC in oocytes might be decreased due to large volumne difusing the signals.
+
+_leaky SAC_ for oocytes (is known) .. meaning that bivalents / homologs without proper tension force
+
+(potentially this bad checkpoint is adaptive for thwarting centromere drive systems, which work through the SAC and tension signals)
+
+how the SAC spindle works, correctly
+
+evidence that the oocyte spindle is leaky
+
+_precition that the leaky spindle makes relaxes selection_
+
+aTMOC's  (because there are mulitple, these could be uneven across the cell)
+
+_asymetrical division_
+
+this makes egg weak to driving centromeres which (are segregation distortors), that means some centromeres segreation (rotate to the polar body side of the egg)
+
+_number of placement of crossovers influencing spindle_
+
+For this model we propose that the number and placement of crossovers -- will affect the
+
+_strengths of model_
+
+_weakness_
+
+No very few spermatocyte spindle data  (it's assumed to )
 
@@
 
+References:
+Schuh Ellenberg 2007 spindle
+
+## notes for Wu thesis
 bivalent orientation early metaphase -- search and capture model -- when there are mulitple KT-MT attachmments, (lateral (on the side or not on centromere/KT proper), merotelic MT attachemtn from ooposite poles) and amplitelic (MT from same pole)

### 2020-03-14 doc/ThesisIntrOutline.rmd
notes for outline draft

---
title: "Intro Outline"
output: html_notebook
---

# themes across chapters

## 1. recombination rate variation in wild population

natural variation for recombiantion rate,

how Rec evolves, parasites, stress het Environments

genetic variation and environmental effects,

## 2. Heterochiasmy, sexspecific evolution of recombination rates

## 3. between cell variation in recombination (meiotic features)

## Figure / Image ideas

- chrom axis / DNA loop

## aim of the thesis

the aim of this thesis --- is to investigate variation and evolution of genome with recombination rates

within species variation (natural populaiton) and sex

genome wide rates -- broken down (to investigate the chromosome - level landscape / broad scale patterns / meiotic features)

### 2020-03-16 doc/Discussion.rmd
notes on discussion

@@
 title: "Discussion"
-author: "April Peterson"
-date: "March 10, 2020"
-output: html_document
+date: ""
+output:
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '5'
+    toc_float: yes
+  pdf_document: default
+  word_document:
+    toc: yes
+    toc_depth: '5'
+bibliography: refs_Main_Report.bib
 ---
@@
 
-This section refers to the matrix of models and results from this study where we (conceptually test if each model would support our results)
-
-We were motivated to run a comparison of
-
-_Models_
+This section refers to the matrix of models and results from this study where we (conceptually test if each model would support our results). We were motivated to run a comparison of
 
 ## 1.Gamete selection
+
 The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.
@@
 
-##2.SACE
+## 2. SACE
+
 Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.
@@
 
-This new model has _strengths_
+This new model has **strengths**
 
@@
 
-##3.Two-locus (protect for drivers)
+## 3.Two-locus (protect for drivers)
 
@@
 
-_model basics_
+**model basics**
 
@@
 
-_strengths_
+**strengths**
 
@@
 
-_weaknesses_
+**weaknesses**
+
 (modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
 
-## 4. Spindle   // Difference in acentromeric spindle and SAC
+## 4. Segregation and physiological Spindle Differences
+
+### The CONTEXT of MEIOSIS
+
+the study confirmed preivously identified sexual dimorhisc patterns in regard to the recombiantion landscape (in mammals) -- in the goal of unifying all of these results, we sough to investigate physiologycal aspects of gametogenesis which have conserved sexual dimorphic features.
+
+LOOK at the oogenesis and spermatogenesis as the context for meiosis (and how )
+
+lead us to ask, how the different contexts of meiosis could influence recombination rates (the recombinaiton landscape)
+
+(not sex chromosome)
+
+while there are many different forms, flavors of anisogamy (cite that one paper)
+
+symetrical vs asymetrical gamete division
 
@@
 The SAC in eggs seems to be weaker/leakier, that is more achiasmate bivalents are required to trigger the SAC (stop the division).
-
-we like this model because it can be applied to _both our results_ for i) sexual dimorphism variation and ii) the fast male evolution in gwRR and recombination landscape.
-
-for the distinct _evolutionary patterns_ selection on the recombination landscape via the spindle (segregation checkpoint),
 
@@
 The survalence of SAC in oocytes might be decreased due to large volumne difusing the signals.
+
+### CO placement translates to bivalent shape
+
+**still an open question of if and how the CO number and placement of crossovers translates to different fitness values for segregation rates**
+
+the metric / amout of sister cohesion might be an good place to start
+
+the ultimate goal is proper disjunction of chromosomes, the main checkpoint is spindle assembly checkpoint (SAC) which responds to the tension signal across chiasmata
+
+we like this model because it can be applied to **both our results** for i) sexual dimorphism variation and ii) the fast male evolution in gwRR and recombination landscape.
+
+Uses / centers around one of the best known selection factors on meiotic division, anneuploidy.
+Several other facors related to meiotic recombiniation center on **segregation** (obligate chiasmata, lower bound of the genome wide rate) it's not unbelievable that the recombination
+
+for the distinct **evolutionary patterns** selection on the recombination landscape via the spindle (segregation checkpoint),
 
@@
 
-## notes for Wu thesis
+#### notes for Wu thesis
 bivalent orientation early metaphase -- search and capture model -- when there are mulitple KT-MT attachmments, (lateral (on the side or not on centromere/KT proper), merotelic MT attachemtn from ooposite poles) and amplitelic (MT from same pole)
@@
 
--  more attention to cell physiology
+-  more attention to cell physiology and biophysics (conserved nature of meiosis pathway (and mitosis ) -- use some models such as Sc for manipulation) (Murkarjee gardner -- quantify the spindle tension
 

### 2020-03-17 doc/Discussion.rmd
added refs to discussion

@@
 
-## Within mouse variance
+## Greater Within mouse variance in female
 
@@
 new paper has revealed within nucleus variation for (CO number -- potentially via the chromatin remodeling) (Veller Kleckner)
-
-prev reported sex diffferences in
 
@@
 
-# Evolved Pattern
+# Evolved Traits
 
-2.Predictors of Recombination rate evolution within (males)
-This section focuses on the largest axis of variation over a short time scale.
-We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.
+We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
+One of our primary objectives was to identify other traits which correlate (could explain the increase).
 
 Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength.
+<While SC length and the telomere distances are similar / conserved across the mouse strains>
 
-## DSBs
+<We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>
 
-Our results find evolution between a subset of strains -- demonstrating that the high rec strains have greater DSB number
+# Evolved rates of precursors, DSBs
 
-Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al). Likely // evidence suggests this pattern is also a feature of the sex differences.
+<define CO:NCO decision, tethered-loop DSB formation SPO11 (cite, paigen? deMassy
+setup baier study,  explain >
 
-1. More DSBs in the high rec strains
+_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
+during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
 
-2.
-The conserved ratios of DSB : CO implies that the increase in DSB numbers are accompanied by proportional increase in crossovers.
-This is evidence that shifts support away from the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
-Together these results indicate that (it's not the CO : NCO decision but an earlier step (when the total number of chromatin loops (the substrate for SPO11, are laid down with the establishment of the meiotic chromosome structure (DNA loop size)
+Our results find evolution between a subset of strains -- demonstrating that the high rec strains have significant higher DSB number.
 
-The tethered loop model of DSB formation
+Our results mirror those found in Baier et al [@baier2014]
 
-3. DSB number being driven by the number of chromatin loops which is positively correlated with axis length.
-4. These results supported by the moderate support of longer SC length (pachytene stage in the high rec males).
+Our results show that the ratios of DSB : CO 
+**range from X to X which is similar to those presented in [@baier2014] (which report X to X) **
 
-_the resuts support / replicate those reported in Baier et al_
+These conserved ratios suggest that the increase in DSB numbers are accompanied by proportional increase in crossovers.
+
+This shifts support away from evolution at the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
+**increase in DSBs is driven by the number of DNA loops** and due to the inverse relationship of loop and chromatin axis length, this would result in longer sc lengths (which reflect the underlying chromatin axis length).
+
+Baier et al found significant differences in sc legnths for 
+The tests in our study - show moderate to low significance in the differences in SC lengths. (this is because we have more noise-can't distinguish individual chromosomes)
 
 a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
-
 -different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures
@@
 
-the correlation of DSB marker with MLH1
+Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al).
 
-(Review Baier et al 2014)
-
-_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
-during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
-
-## INTERFERENCE evolution
+# INTERFERENCE evolution
 
@@
 -- lack of SC length prophase cell charaecterization
+-- comparisons within and between sexes (that cattle paper)
 
--- comparisons within and between sexes (that cattle paper)
+# Reference

### 2020-03-17 doc/Discussion.rmd
better writing points

@@
 ---
-title: "Discussion"
+title: "Discussion Draft"
 date: ""
@@
 
-# Differences across models
+```{r format.Notes, include=FALSE}
 
-Explain the model differences;
+#table, wite by all of the columns and rows
 
-## 1.can it be applied to broad scale chromosome level landscape?
+#constrast on model types
+#-evolution
+#-models accounting for chromosome based traits
 
--genome wide recombination rates are made of the chromosome level patterns
+#HetC traits
+# 1. typical landscape
+# 2. female variance
+# 3. chromatin organization
 
-This is important because general patterns have been observed for the recombination landscape at the chromosome level (also small range of crossovers per chromosome and broad scale pattern (Haenel et al ))
+#faster male evolution traits
+# 1. Reversed Direction of Heterochiasmy
+# 2. correlated gwRR increase and interference strength
 
-## 2. Does the model account for evolution?
+#go through each traits and point out which models support each of the traits
 
-given all the variance observed and fast evolution, models have to have some engine for evolution.
-
-(some source of affecting fitness, or shifting the optima)
+```
 
@@
 
-## male telomeric, female centromeric Typical landscape
+## Broad Scale crossover placement patterns, male telomeric, female centromeric 'Typical landscape'
+
+```{r typical.landscape.outline, include=FALSE}
+
+#-long observed
+#measured from a variety of approaches,
+#larger variety of species
+
+#-Exception to landscape (marisupials, birds?)
+
+#which models make clear predctions for typical landscape?
+#SACE
+#drivers
+#spindle differences
+
+```
+
 <review the typical landscape, one of the major broad scale patterns,
@@
 
-   1.
-male telocentric and female uniform landscape / general placement of crossovers.
-What is referecened to as typical landscape (Sardell) -- see references sardell kirkpatrick for list of species which show/display this typical phenotype
+The 'typical landscapes' refers to the sex differences in the board scale pattern or general placement of crossovers along the length of chromosomes.
+The primary pattern which has been observed across mny species is that crossovers are shifted to telomereic positions in males while a more unifrom landscape in observed in females [@sardell_sex_2020].
 
-A difference in the general placement (general recombination landscape), with males having telomere bias and females having more uniform placement of crossovers along chromosomes.
+Our results show that the typical landscape is consered across all strains of house mice in this study (for all single chromosome daata and when controlled for crossover number), the broad scale position of crossovers was shifted closer to the telomere end (of the long arm).
 
--Exception to landscape (marisupials, birds?)
-
-The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes.
+<Detangling centromere suppression from telomere bias,  The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes.
 we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalent
-s. )
+s. )>
 
@@
 
-1. SC length  2. connection to interference
+```{r chromatin.organization, include=FALSE}
+
+#1. chromatin organization, female shorter loops, longer axis
+
+#2. A consequence of chromatin organization (loop size) is difference in interference strength in (MB / cM)
+
+#seen in humans, mice (signal comes through even with cell to cell variation in chromatin compaction / organization)
+#
+
+#exceptions to female pattern (fish birds), lack of data
+# interference strength reversal (At and cattle) -- we predict that the loop size will also reverse, -- but there is a lack of cytological data for both sexes.
+
+#lots of data on genetic maps for both sexes, but lack of cytological data to support
+
+#our results, female longer axis, 
+#review caveats -- why we believe the pattern is true
+
+#refs
+#BF klecner?
+#negative relationship of axis and loop size (Lynn? Gruhn?)
+#global re-orgnaiziton of DNA / chromosomes (3D, new Chip-seq pubs)
+```
+
+1. basics and expectations for chromatin organization
+
+2. Well supported model for interference is that it is mediated / communicated through the 2D chromatin axis or SC (depending on the timing)
+
+3. consequence for interference (general patterns) is that interference measured as (mb) will be stronger if DNA loops are stronger
+
+The relationship // molecular organization of chromosome -- undergoes global re-orgzniation during meiosis.
+The basic take away is that eraly stage in meiosis) there is a central axis made up of chromatin and proteins while most of the DNA / chromatin are folded into loops that emenate off of the axis.
+This structure has been called 'lampbrush' by some (cite).
+
+A pattern that has been long noticed --
+
+Merging of two types of data
+
+-total chromatin structure (axis length) (cytology)
+
+-Interfere measured as cM or DNA base pairs Mb  (the spacing between markers from linkage map)
 
@@
 
-The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov)
+**lack of difference in the raw measures of IFD, but significant difference in **
 
-  2. Explain the connection between interference strength in physical space and (abstracted space)
+The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov) **these results mirror those seen in** [@petkov2007]
 

### 2020-03-17 doc/Discussion.rmd
more edits

@@
 
-Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.
+< Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.>
 
-  it boils down to, the relationship with axis length and DNA loop size --
-
-  --females have shorter loops - longer axis and weaker interference
-
-  Chromatin structure -- the size / length of meiotic chromosomes
-
-  same pattern in humans (Gruhn et al 2013)
-
+<same pattern in humans (Gruhn et al 2013)
   (exceptions, birds (fish?) plants?) -- examining this is hard due to chromosome size effecs
-
-  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes
+  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes>
 
@@
 
-## Greater Within mouse variance in female
+## Greater within mouse variance in female
+
+```{r within.mouse.var, include=FALSE}
+
+#our results, female more (greater than tech error) male no significant difference
+
+#-replicate some previous results
+#(Gruhn et al), Kong et al?, cattle,
+
+#between cell variance (Veller Kleckner), mechnism
+#evolution / popgen model from (veller)
+
+#sources of within mouse variance; age effects,  Hassold hunt
+
+#models that predict more within animal variance
+
+##connection on evolution of transcriptional noise
+
+```
 
@@
 
-We did not observe
-
-- Because these are inbred strains raised in common / standard lab environment, (an single levels)
-
 _sources of within animal variation_
 
--age differences
+ mouse and humans show greater between cell variance -- [@]
 
-new paper has revealed within nucleus variation for (CO number -- potentially via the chromatin remodeling) (Veller Kleckner)
-
- Gruhn et al 2013 (the favorite figures)
-
- Gruhn et al 2016 -- not due to the developmental stage of the oocytes (ovulation) (the production line)
-
- mouse and humans show greater between cell variance -- this metric isn't really focused on (not many studies that examine inidivial level and both sexes)
+ this metric isn't really focused on (not many studies that examine inidivial level and both sexes)
 
 3. Within mouse variance differences
-In our results we observe more between cell variance in females (across all strains).
-This pattern has been observed in humans (Lynn et al).
+In our results we observe more between cell variance in females (across all strains). This pattern has been observed in humans (Lynn et al).[@lynn2002]
 (But results of reporting observations of single meiosis in both sexes).
@@
 but maybe, rec landscapes in general)
-
-(Gruhn et al), Kong et al?
-
-cattle,
-
-- age effects?
-
-- Hassold hunt
-
-_Relaxed selection_ leads to
 

### 2020-03-17 doc/Discussion.rmd
small edits

@@
 
-## Greater within mouse variance in female
+# Greater within mouse variance in female
 
@@
 #our results, female more (greater than tech error) male no significant difference
+#between cell variation
 
@@
 
-##connection on evolution of transcriptional noise
+## connection on evolution of transcriptional noise
+# transcriptional noise and complex traits -- both show the same repsonses to different modes of selection
 
@@
 
-- Our results confirm previous findings of females having greater between cell variation (for indiviual females compared to males)
-the results -- indicate there is more than can be due to technical noise (for oocytes vs spermatocytes spreads)
-also all ovary collections were from early/ young mice -- not due to the maternal age effects, (which has to do with the stability of bivalents at metaphase -- after sexual maturity and ovulation (ref))
+**the within animal variation is equvillent to between cell variation**
 
-_sources of within animal variation_
+Our results confirm previous findings from mice [@lynn2002] and humans of females having greater between cell variation compared to males.
 
- mouse and humans show greater between cell variance -- [@]
+While these (age, technical noise, ect) are potential sources of technical noise -- these results indecate this isn't due to technical noise or age effects.
 
- this metric isn't really focused on (not many studies that examine inidivial level and both sexes)
-
-3. Within mouse variance differences
-In our results we observe more between cell variance in females (across all strains). This pattern has been observed in humans (Lynn et al).[@lynn2002]
-(But results of reporting observations of single meiosis in both sexes).
-
-(references, more within animal variance in
-(number of crossovers per cell)
-but maybe, rec landscapes in general)
+Transcriptional noise and complex traits both show the same repsonses to different modes of selection, lower variation under stabilizing selection and greater variaion with relaxed selection.
+The difference in between
 

### 2020-03-17 doc/Discussion.rmd
more notes and edits

@@
 
+```{r Evolve.tratis.outline, include=FALSE}
+
+```
+
 We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
@@
 # Evolved rates of precursors, DSBs
+
+```{r evo.DSB, include=FALSE}
+
+```
 
@@
 
-# INTERFERENCE evolution
+# evolution in Interference
+
+```{r evo.interference, include=FALSE}
+
+#review results,
+
+#why unexpected
+
+#assuptions for predictions
+
+#models that predict this pattern?
+
+```
 
@@
 1.
-Cattle (Ma?), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)
+Cattle (Ma?) [@ma2015cattle] ?
+(or newer one), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)
 
@@
 # Explanations and Model testing
+
+```{r Model.testing, echo=FALSE}
+
+```
 
@@
 ## 4. Segregation and physiological Spindle Differences
+
+```{r Spindle.Model, include=FALSE}
+
+#1. context of meiosis
+
+#2. 
+```
 

### 2020-03-18 doc/Discussion.rmd
trying to force writing

@@
 
+#imortant of testable hypotheses
+
+#model not included in this comparison
+
 ```
 
-This section refers to the matrix of models and results from this study where we (conceptually test if each model would support our results). We were motivated to run a comparison of
+In our analysis / synthesis of results, we decded to test and systematically compared models which have been proposed to explain our results (both for the sex dfferences and the faster male evolution of recombination rates).
+We attempted to test the hypotheses of if our choosen models could produce the main / general patterns oberved in this study (Table X)
 
-## 1.Gamete selection
+We generall focus on models for the evolution of heterochiasmy since that is the primary data/set of results in this study.
+We search for a parsimonius explanation /model that fits both the results from sex differences and the faster male evolution.
 
-The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.
+# 1.Gamete selection
 
-<go through the squares>
+originally proposed/outlined by Trivers [] and refined /formatlized in [@lenormand2003].
+This model discribes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombiaion rates between males and females.
 
-this is a modifer model it abstracts the genome area.
+ This is a modefier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifer).
 
@@
 
+<the sex with the most variable reproductive fitness will be selected to have lower genome wide recombination rates>
+
+From our results, this model would predict -- fit the results of:
+
 this model doesn't predict that the heterochismy direction would reverse.
 
-## 2. SACE
+# 2. SACE
 
@@
 
-## 3.Two-locus (protect for drivers)
+# 3.Two-locus (protect for drivers)
 
@@
 
-## 4. Segregation and physiological Spindle Differences
+# 4. Segregation and physiological Spindle Differences
 

### 2020-03-19 Rmd_Discussion_Notes.Rmd
notes in discussion and reference note doc

--- 2020-03-17 doc/Discussion.rmd
@@
 ---
-title: "Discussion Draft"
-date: ""
-output:
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-  pdf_document: default
-  word_document:
-    toc: yes
-    toc_depth: '5'
-bibliography: refs_Main_Report.bib
+title: "writing notes"
+output: html_document
 ---
 
-```{r format.Notes, include=FALSE}
-
-#table, wite by all of the columns and rows
-
-#constrast on model types
-#-evolution
-#-models accounting for chromosome based traits
-
-#HetC traits
-# 1. typical landscape
-# 2. female variance
-# 3. chromatin organization
-
-#faster male evolution traits
-# 1. Reversed Direction of Heterochiasmy
-# 2. correlated gwRR increase and interference strength
-
-#go through each traits and point out which models support each of the traits
-
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-# Conserved Sexually Dimorphic Features of Recombination landscape
-
-Conserved heterochiasmy paterns
-
-  - (the canonical pattern has too many exceptions to be the uniform pattern)
-
-  - our results have exception, and other exceptions are...
-
-These are the metrics for the _conserved sex differences_
-
-- (heterochiasmy refers to different numbers of chiasmata - and recently has been extended to number of crossovers -- this view / charaecterization is slightly incomplete, since it considers just one cell stage and 1 value (a count))
-
-We argue that the sex differences (broaden the definition of heterochiasmy or use a different word) -- which includes more aspects that show conserved sexes differences
-
-Our results indicate, sexes have distinct evolutionary trajectories for genome wide recombinaiotn rate (and landscape) in males nd females).
-
-Conserved features of the landscape may play a role in influencing/keeping the sex specific trajectories.
-In our examination of the panel of mouse strains, we note several features of the recobmination landscape with conserved sexual dirmophism pattern.
-That is the sex within sex these features show no evolution and consistant pattern of variation.
-
-## Broad Scale crossover placement patterns, male telomeric, female centromeric 'Typical landscape'
-
-```{r typical.landscape.outline, include=FALSE}
-
-#-long observed
-#measured from a variety of approaches,
-#larger variety of species
-
-#-Exception to landscape (marisupials, birds?)
-
-#which models make clear predctions for typical landscape?
-#SACE
-#drivers
-#spindle differences
-
-```
-
-<review the typical landscape, one of the major broad scale patterns,
-exceptions
-consequences, (differences in variation, Sardell paper) >
-
-The 'typical landscapes' refers to the sex differences in the board scale pattern or general placement of crossovers along the length of chromosomes.
-The primary pattern which has been observed across mny species is that crossovers are shifted to telomereic positions in males while a more unifrom landscape in observed in females [@sardell_sex_2020].
-
-Our results show that the typical landscape is consered across all strains of house mice in this study (for all single chromosome daata and when controlled for crossover number), the broad scale position of crossovers was shifted closer to the telomere end (of the long arm).
-
-<Detangling centromere suppression from telomere bias,  The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes.
-we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalent
-s. )>
-
-## Chromatin Structure and interference strength
-
-```{r chromatin.organization, include=FALSE}
-
-#1. chromatin organization, female shorter loops, longer axis
-
-#2. A consequence of chromatin organization (loop size) is difference in interference strength in (MB / cM)
-
-#seen in humans, mice (signal comes through even with cell to cell variation in chromatin compaction / organization)
-#
-
-#exceptions to female pattern (fish birds), lack of data
-# interference strength reversal (At and cattle) -- we predict that the loop size will also reverse, -- but there is a lack of cytological data for both sexes.
-
-#lots of data on genetic maps for both sexes, but lack of cytological data to support
-
-#our results, female longer axis, 
-#review caveats -- why we believe the pattern is true
-
-#refs
-#BF klecner?
-#negative relationship of axis and loop size (Lynn? Gruhn?)
-#global re-orgnaiziton of DNA / chromosomes (3D, new Chip-seq pubs)
-```
-
-1. basics and expectations for chromatin organization
-
-2. Well supported model for interference is that it is mediated / communicated through the 2D chromatin axis or SC (depending on the timing)
-
-3. consequence for interference (general patterns) is that interference measured as (mb) will be stronger if DNA loops are stronger
-
-The relationship // molecular organization of chromosome -- undergoes global re-orgzniation during meiosis.
-The basic take away is that eraly stage in meiosis) there is a central axis made up of chromatin and proteins while most of the DNA / chromatin are folded into loops that emenate off of the axis.
-This structure has been called 'lampbrush' by some (cite).
-
-A pattern that has been long noticed --
-
-Merging of two types of data
-
--total chromatin structure (axis length) (cytology)
-
--Interfere measured as cM or DNA base pairs Mb  (the spacing between markers from linkage map)
-
-general pattern for the sexual dimorphsim the relationship between axis length and CO number
--- axis are longer in females than male, suggesting suport more crossovers (more DSBs)
-but the density is less in females.
-
-Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
-Two strains have have more crossovers in males, but females still have longer sc axis.
-_This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape_
-
-**lack of difference in the raw measures of IFD, but significant difference in **
-
-The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov) **these results mirror those seen in** [@petkov2007]
-
-< Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.>
-
-<same pattern in humans (Gruhn et al 2013)
-  (exceptions, birds (fish?) plants?) -- examining this is hard due to chromosome size effecs
-  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes>
-
-2. Short axis / longer DNA loops will have stronger interference
-This prediction combines the findings of chromatin compaction and interference differences which are intertwined by the transmission of interference along the chromosome axis and the inverse relationship chromatin compaction (axis length and DNA loop sizes).
-In physical units (micron) the counting model fits interference well to the 2D linear space of the axis these measures of interference are conserved across sexes in mouse.
-This is because the counting units are the BASES of DNA loops which is conserved along the 2D length of chromatin axis.
-When this space is transformed to account for all of the DNA base pairs which are sequestered into the DNA loops above the axis, the strength of interference will vary depending the on underlying chromatin structure.
-The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
-Thus the strength of interference will be stronger because with each step (count) along the 2D axis,
-more bases pairs are being skipped above the axis.
-
-# Greater within mouse variance in female
-
-```{r within.mouse.var, include=FALSE}
-
-#our results, female more (greater than tech error) male no significant difference
-#between cell variation
-
-#-replicate some previous results
-#(Gruhn et al), Kong et al?, cattle,
-
-#between cell variance (Veller Kleckner), mechnism
-#evolution / popgen model from (veller)
-
-#sources of within mouse variance; age effects,  Hassold hunt
-
-#models that predict more within animal variance
-
-## connection on evolution of transcriptional noise
-# transcriptional noise and complex traits -- both show the same repsonses to different modes of selection
-
-```
-
-**the within animal variation is equvillent to between cell variation**
-
-Our results confirm previous findings from mice [@lynn2002] and humans of females having greater between cell variation compared to males.
-
-While these (age, technical noise, ect) are potential sources of technical noise -- these results indecate this isn't due to technical noise or age effects.
-
-Transcriptional noise and complex traits both show the same repsonses to different modes of selection, lower variation under stabilizing selection and greater variaion with relaxed selection.
-The difference in between
-
-# Evolved Traits
-
-We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
-One of our primary objectives was to identify other traits which correlate (could explain the increase).
-
-Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength.
-<While SC length and the telomere distances are similar / conserved across the mouse strains>
-
-<We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>
-
-# Evolved rates of precursors, DSBs
-
-<define CO:NCO decision, tethered-loop DSB formation SPO11 (cite, paigen? deMassy
-setup baier study,  explain >
-
-_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
-during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
-
-Our results find evolution between a subset of strains -- demonstrating that the high rec strains have significant higher DSB number.
-
-Our results mirror those found in Baier et al [@baier2014]
-
-Our results show that the ratios of DSB : CO 
-**range from X to X which is similar to those presented in [@baier2014] (which report X to X) **
-
-These conserved ratios suggest that the increase in DSB numbers are accompanied by proportional increase in crossovers.
-
-This shifts support away from evolution at the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
-**increase in DSBs is driven by the number of DNA loops** and due to the inverse relationship of loop and chromatin axis length, this would result in longer sc lengths (which reflect the underlying chromatin axis length).
-
-Baier et al found significant differences in sc legnths for 
-The tests in our study - show moderate to low significance in the differences in SC lengths. (this is because we have more noise-can't distinguish individual chromosomes)
-
-a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
--different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures
-
-our results are similar in several ways
-- correlation of final MLH1 and DBS counts, SC measues correlate with the DSB number (even tho the differnces we observe in our study are -- slight / less significant -- we attribute that to chromosome specific noise )
-
-Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al).
-
-# INTERFERENCE evolution
-
-1. Review results
--The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.
-
-These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol.
-(it also goes against the logical assumptions models of fitting more crossovers along the SC chromosome axis space)
-
-Yet (we are confident that this pattern is not due to errors).
-
- the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).
-
--Also:
-	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.
-
-	-The difference in interference patterns is driven by the difference in the rates of small IFD observations, (mean vs median). a pattern that is suggestive of selection agisnt short IFDs
-
-There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
-breifly, the model simulates / sets up a scnario where a modifer acts through the probabilty of flanking regions having 0 or 2 crossovers.
-ith crossover assuance, recombination rate increases the further apart the loci are spaced (?)
-
-Additionally there are a handful of empircal results which also show the unexpected pssotive correlation between gwRR and interference strength.
-
-1.
-Cattle (Ma?), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)
-
-2. The IFDs from (Beth's cross) a related house mouse cross show that the PWD (same strain, different mice) had longer IFD compared to Cast (HVR unpublished data). (also gamma?)
-
-3. In Peromyscus, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@Peterson]
-
-<note about the differences in the patterns>  assumptions with gamma dstribution and linkage maps, (closer to infinte genetic space), less weight on crossover assurance - more weigth on 
-(pooled SC/chromosome length gives strong correlation )
-
-<which is much different than when a single chromosome is examined (by itself), finite 2D length, crossover assurance, weaker correlation with ((the closer to chromosome level for SC legnth metric / summary, the weaker the correlation with crossover number))
-
-2.  Review the assumptions for measuring CO interference in different ways.
-The logical model: Crossover interference is a mechanism of suppressing crossovers, negative relationship between interference strength and gwRR.
-More crossovers can fit along the chromosomes and results in more crossovers overall.
-Examples of empirical results supporting this pattern;  Otto Payseur, (Ruiz-Herrera? Bomblies?)
-Quantifing using the gamma distribution: 
-Caveats,  This doesn't make since for the general recombination landscape:   i) finite chromosome length and ii) limited range of crossovers per chromosomes (1 to 4).
-Alternative interference characterization:  i) the number of crossovers are held constant and ii) transform / generalize the rec landscape, (focus on how the rec landscape would change instead of the genome wide rate) iii) assume crossover assurance.
-	-The Goldstein model, models predicts a positive correlation with interference strength and higher recombination rate.
-A modifier which changes the probability of the recombination landscape having 0 crossovers to 2 crossovers, (driven by crossover assurance).
-
--Examples of empirical results which show this result / pattern
-	-(cattle) comparison WITHIN sexes
-	- Preliminary results for F2 cross (Hannah's work), (PWD vs CAST)
-	(maybe more once I start looking closer)
-
-4. HYPER-Crossover experiments. (review the effects when crossover number is artificially altered)
-
-# Explanations and Model testing
-
-This section refers to the matrix of models and results from this study where we (conceptually test if each model would support our results). We were motivated to run a comparison of
-
-## 1.Gamete selection
-
-The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.
-
-<go through the squares>
-
-this is a modifer model it abstracts the genome area.
-
-This model predicts the evolution of heterchiasmy -- through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
-
-this model doesn't predict that the heterochismy direction would reverse.
-
-## 2. SACE
-
-Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.
-
--- this model is an extension of the modifer model presented in Lenormand
-
--- (basics)
- extends the lenomand focusing on two conditions.
-first that the stength of epistatsis in dependant on the cis or trans nature.
-the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)
-
- second that the relative interference strengths differnt between male and female.  This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
-
-This new model has **strengths**
-
-  - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)
-
- - incorporates know cellular / molecular mechanisms with a very common pattern sexual dimorphic recombination rate landscape.
-
-- extends modifer model to explain a chromosome based / level pattern
-
-_weaknesses_, and ways the model diverges from our results
-
- - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
-
-- Other ways this model departs from our results,
-
--Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
-
-However / while not explicitaly arameterized or predicted in this model,
- -- this model predicts larger chuncks of chromosomes will segregate together
- (as outlined in Veller et al)
-
- our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle
-
-## 3.Two-locus (protect for drivers)
-
-Asymmetric division of egg opens them up to meiotic drivers.
-Recombination modifiers for sex specific female landscape will decrease the chances of driving centromeres to segregate to the egg by increasing the number of crossovers and placing them across a larger area of the genome (uniform placement).
-
-**model basics**
-
-'break up driv system consiracies'
-
-**strengths**
-
-a clear fitness effect (the drive)
-
-explains female landscape pattern in reference to the centromere (presummed location of drivers), yet the model doesn't scale down to chromsome level (such as to explain, chromosome size differences)
-
-**weaknesses**
-
-(modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
-
-## 4. Segregation and physiological Spindle Differences
-
-### The CONTEXT of MEIOSIS
-
-the study confirmed preivously identified sexual dimorhisc patterns in regard to the recombiantion landscape (in mammals) -- in the goal of unifying all of these results, we sough to investigate physiologycal aspects of gametogenesis which have conserved sexual dimorphic features.
-
-LOOK at the oogenesis and spermatogenesis as the context for meiosis (and how )
-
-lead us to ask, how the different contexts of meiosis could influence recombination rates (the recombinaiton landscape)
-
-(not sex chromosome)
-
-while there are many different forms, flavors of anisogamy (cite that one paper)
-
-symetrical vs asymetrical gamete division
-
-A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle).
-(Our hypothesis is that the tension in a centrosome spindle is stronger (or more uniform) since all MT-KT are anchored at single points across cells (the two centrosomes).
-This will also make prediction for the spindle assembly checkpoint (SAC), which is stronger/stricter in smaller gametes with centrosomes.
-(the SAC is more sensitive to achiasmate bivalents on the spindle / better at detecting a lack of tension).
-Whereas in the acentrosome spindle is more diffuse (across a larger area) and the tension across bivalents are anchored by multiple MTOCs, resulting in weaker tension force or less uniform strengths across all the bivalents within the cells.
-The SAC in eggs seems to be weaker/leakier, that is more achiasmate bivalents are required to trigger the SAC (stop the division).
-
-_main sex differences in spindle_ are the anchoring of MT at opposite poles
-
-in males there are centromsomes for Mi and MII divisions
-where are in oocytes, a (universal) conserved feature of oogenesis in many species is that they loose centrosomes -- instead the MTs are anchored at mulitple aMTOCs.
-
-There is potential for these aMTOC to be unevenly distributd across the cortical and central side of the spindles (Wu et al, Schuh and ellenberg) --- (could be an origin of different tension / pulling )
-
-The survalence of SAC in oocytes might be decreased due to large volumne difusing the signals.
-
-### CO placement translates to bivalent shape
-
-**still an open question of if and how the CO number and placement of crossovers translates to different fitness values for segregation rates**
-
-the metric / amout of sister cohesion might be an good place to start
-
-the ultimate goal is proper disjunction of chromosomes, the main checkpoint is spindle assembly checkpoint (SAC) which responds to the tension signal across chiasmata
-
-we like this model because it can be applied to **both our results** for i) sexual dimorphism variation and ii) the fast male evolution in gwRR and recombination landscape.
-
-Uses / centers around one of the best known selection factors on meiotic division, anneuploidy.
-Several other facors related to meiotic recombiniation center on **segregation** (obligate chiasmata, lower bound of the genome wide rate) it's not unbelievable that the recombination
-
-for the distinct **evolutionary patterns** selection on the recombination landscape via the spindle (segregation checkpoint),
-
-_leaky SAC_ for oocytes (is known) .. meaning that bivalents / homologs without proper tension force
-
-(potentially this bad checkpoint is adaptive for thwarting centromere drive systems, which work through the SAC and tension signals)
-
-how the SAC spindle works, correctly
-
-evidence that the oocyte spindle is leaky
-
-_precition that the leaky spindle makes relaxes selection_
-
-aTMOC's  (because there are mulitple, these could be uneven across the cell)
-
-_asymetrical division_
-
-this makes egg weak to driving centromeres which (are segregation distortors), that means some centromeres segreation (rotate to the polar body side of the egg)
-
-_number of placement of crossovers influencing spindle_
-
-For this model we propose that the number and placement of crossovers -- will affect the
-
-_strengths of model_
-
-_weakness_
-
-No very few spermatocyte spindle data  (it's assumed to )
-
-_review the correct SAC_
-
-References:
-Schuh Ellenberg 2007 spindle
-
-#### notes for Wu thesis
+# notes for Wu thesis
 bivalent orientation early metaphase -- search and capture model -- when there are mulitple KT-MT attachmments, (lateral (on the side or not on centromere/KT proper), merotelic MT attachemtn from ooposite poles) and amplitelic (MT from same pole)
@@
 data suggests that more MTOC are recruited to the cortical side
-
-5. Model prediction
-The models above which predict directional selection on the recombination landscape in males: SACE and Spindle differences are most parsimonious for the models review in this paper.
-The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.
-
-Consequences of evolution in the number and placement of crossovers / evolution of interference strength:
-Spindle / SAC model: The number and placement of crossovers changes the resulting amount of sister cohesin involved the tension signal, which affects the tetrad/chiasmata structure at MI when the homologs are pulled apart (Hollis et al 2020, Lee et al 2019).
-This model predicts a difference in the meiotic spindles or division mechanism between high and low rec strains.
-ii. SACE: For the 2CO in the high rec males, larger blocks of chromosomes are kept together in the next generation.
-	This model would predict greater reproductive variance for males.
-
-# Future Steps
-
-be aware of sex-specific and sex average biases
-
-- linkage / R between two foci is sex averaged
-
-- Celegans -- cytlogy focuses on oogenesis  and genetic maps are often from hermaphrodites (selfing -- sperm and egg landscapes are merged in hemaphordites)
-
--- the value of Chromosome level measures -- incorporate chromosome behavior into models
-
--  more attention to cell physiology and biophysics (conserved nature of meiosis pathway (and mitosis ) -- use some models such as Sc for manipulation) (Murkarjee gardner -- quantify the spindle tension
-
--- lack of SC length prophase cell charaecterization
--- comparisons within and between sexes (that cattle paper)
-
-# Reference

### 2020-03-19 doc/Discussion.rmd
notes in discussion and reference note doc

@@
 ```{r format.Notes, include=FALSE}
+#table, wite by all of the columns and rows
 
-#table, wite by all of the columns and rows
+#over-view of the synthesis of hypothesis testing
+#(this is how we're synthsizing / testing / the framework were using to compare the results to the literature)
 
@@
 #go through each traits and point out which models support each of the traits
+```
+
+# Explanations and Model testing
+
+```{r Model.testing, include=FALSE}
+
+#imortant of testable hypotheses
+
+#model not included in this comparison
 
@@
 
-# Conserved Sexually Dimorphic Features of Recombination landscape
+This is how we are going to test/ synthesize / fit our results into the literuature and test hypothesis for which models best explain our results.
+In our analysis / synthesis of results, we decded to systematically compared models which have been proposed to explain our results (both for the sex dfferences and the faster male evolution of recombination rates).
+We attempted to test the hypotheses of if our choosen models could produce the main / general patterns oberved in this study (Table X)
 
-Conserved heterochiasmy paterns
+We generally focus on models for the evolution of heterochiasmy since that is the primary data/set of results in this study.
+We search for a parsimonius explanation /model that fits both the results from sex differences and the faster male evolution.
+But we also test the male specific traits.
 
-  - (the canonical pattern has too many exceptions to be the uniform pattern)
+**the models**  We examined 3 modifer models; haploid selection, two locus model, and SACE.
+We also examined and propose cell physiology based models: the Chromosome Ocsilation (COM) model and a model focused on the spindle differences between males and females.
 
-  - our results have exception, and other exceptions are...
+<we left out Beam film model for XX reasons>
 
-These are the metrics for the _conserved sex differences_
+**Conserved heterochiasmy paterns**
 
-- (heterochiasmy refers to different numbers of chiasmata - and recently has been extended to number of crossovers -- this view / charaecterization is slightly incomplete, since it considers just one cell stage and 1 value (a count))
+Below are the main traits from our results: more between cell variance in females, chromatin compaction (with consequence for interference strength), and broad scale landscape differences.
 
-We argue that the sex differences (broaden the definition of heterochiasmy or use a different word) -- which includes more aspects that show conserved sexes differences
+<heterochiasmy refers to different numbers of chiasmata - and recently has been extended to number of crossovers -- this view / charaecterization is slightly incomplete, since it considers just one cell stage and 1 value (a count))
+We argue that the sex differences (broaden the definition of heterochiasmy or use a different word) -- which includes more aspects that show conserved sexes differences>
 
-Our results indicate, sexes have distinct evolutionary trajectories for genome wide recombinaiotn rate (and landscape) in males nd females).
+<Our results indicate, sexes have distinct evolutionary trajectories for genome wide recombinaiotn rate (and landscape) in males nd females).
 
@@
 In our examination of the panel of mouse strains, we note several features of the recobmination landscape with conserved sexual dirmophism pattern.
-That is the sex within sex these features show no evolution and consistant pattern of variation.
+That is the sex within sex these features show no evolution and consistant pattern of variation.>
 
-## Broad Scale crossover placement patterns, male telomeric, female centromeric 'Typical landscape'
+# 'Typical landscape'
 
@@
 
-<review the typical landscape, one of the major broad scale patterns,
+<Broad Scale crossover placement patterns, male telomeric, female centromeric,
+review the typical landscape, one of the major broad scale patterns,
 exceptions
-consequences, (differences in variation, Sardell paper) >
+consequences, (differences in variation, Sardell paper)>
 
@@
 
-## Chromatin Structure and interference strength
+# Chromatin Structure and interference strength
 
@@
 
-# Greater within mouse variance in female
+# Greater between cell variation
 
@@
 
+so all of the models could reasonably predict the above results ... examining some of the traits related to the faster evolution in the male specific data indecate different support across the models.
+
 # Evolved Traits
@@
 
+#it's not the sex differences that distinguish
+
+#revrsed heterochiasmy direction
+#faster -- rapid evoluiton in males
+
+#this also goes along with
+
+# interference strength
+
 ```
 
-We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
-One of our primary objectives was to identify other traits which correlate (could explain the increase).
+Some of the traits related to faster male evolution -- distinguish in which models can redict them:
+
+<We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
+One of our primary objectives was to identify other traits which correlate (could explain the increase).>
 
@@
 1.
-Cattle (Ma?) [@ma2015cattle] ?
+Cattle (Ma?) [@ma2015_cattle] ?
 (or newer one), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)
@@
 
-3. In Peromyscus, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@Peterson]
+3. In Peromyscus, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019]
 
@@
 
-# Explanations and Model testing
-
-```{r Model.testing, echo=FALSE}
-
-#imortant of testable hypotheses
-
-#model not included in this comparison
-
-```
-
-In our analysis / synthesis of results, we decded to test and systematically compared models which have been proposed to explain our results (both for the sex dfferences and the faster male evolution of recombination rates).
-We attempted to test the hypotheses of if our choosen models could produce the main / general patterns oberved in this study (Table X)
-
-We generall focus on models for the evolution of heterochiasmy since that is the primary data/set of results in this study.
-We search for a parsimonius explanation /model that fits both the results from sex differences and the faster male evolution.
-
-# 1.Gamete selection
+ 1.Gamete selection
 
@@
 
-# 2. SACE
+2. SACE
 
@@
 
-# 3.Two-locus (protect for drivers)
+ 3.Two-locus (protect for drivers)
 
-Asymmetric division of egg opens them up to meiotic drivers.
-Recombination modifiers for sex specific female landscape will decrease the chances of driving centromeres to segregate to the egg by increasing the number of crossovers and placing them across a larger area of the genome (uniform placement).
+This second modifer model was persented in [@brandvain2012scrambling] the basics of the model are that females will have a longer genetic map and more centromere biased recombination landscape compared to males.
 
-**model basics**
+A strength of this model is that is presents a clear driver / source of differential fitness (ie meiotic drivers can decrease fitness/fertility).
+This model also accounts for the general female pattern,
 
-'break up driv system consiracies'
+However, this model doesn't explain the borad scale pattern for the male pattern.
 
-**strengths**
+**weaknesses**  (modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
 
-a clear fitness effect (the drive)
+this model would predict / fit the results of
 
-explains female landscape pattern in reference to the centromere (presummed location of drivers), yet the model doesn't scale down to chromsome level (such as to explain, chromosome size differences)
+<**model basics**, 'break up driv system consiracies'>
 
-**weaknesses**
-
-(modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
-
-# 4. Segregation and physiological Spindle Differences
+ 4. Segregation and physiological Spindle Differences
 
@@
 
-### The CONTEXT of MEIOSIS
+# The CONTEXT of MEIOSIS
 
@@
 
-### CO placement translates to bivalent shape
+# CO placement translates to bivalent shape
 
@@
 
-#### notes for Wu thesis
-bivalent orientation early metaphase -- search and capture model -- when there are mulitple KT-MT attachmments, (lateral (on the side or not on centromere/KT proper), merotelic MT attachemtn from ooposite poles) and amplitelic (MT from same pole)
-then biorientation process starts, --search and capature process (only amplictic Mt-KT attachments get stablize)
-
-Aurora -- detects bad connections,
-
-_evidence that SAC in oogenesis is leaky_
-
-(Lane et al 2012)
-
-whenn going through the attachment correection, bivalents are thought to be rotated via pulling from MT
-
-Akera 2017 -- chromosomes at different ends of spindle have different densites (in reference to the cortical or central side of spindle (PB or egg))
-
-Wu thesis -- chrm 17 with asymetric centromere str, cenromeres with large major stat drove to (polar body (opposite of drive) (lack of re-orientation, once it ended up on that side))
-
-the biased rotation happened before spindle migratgation (from center of egg to near cortex)
-
-when a aurora inhibitor applied -- the biased rotation of chrm 17 went away and anaphase moved up (by passed SAC as in Lane 2010)
-
-(k-fibre the bundle MTs connected to KT Iwata-otsubo 2017) -- larger KT would provide more k0fiber binding sites
-
-the spindle migration (from center of egg to cortext depends on actin (Li albertini 2013))
-cytoplasmic streaming when spindle moves close to the cortex (yi et al 2013)
-
-the KT and centromere seperate when MT's start pulling (trains start moving) (they does seperate whn MT inhibitors are applied)
-
-Coretical KTs undergo greater tension than centrak KTs.
-
-Review of modifier / popgen models
-Indirect forces common assumptions themes across all the modifier models.
-Review of cell biology
-Look at conserved features of gametogenesis that distinguish male and females
-Different Centrosome spindle
-Asymmetrical division
-
-aMTOC (Cliff and Schuh 2015, Schuh and ellenburg)
- since there are multiple aMTOC (not 2 centrioles) -- the MTOCs across the sides of the cells could be unven.
-
-after NEBD the aMTOCs are fragmented and coalest into two main clusters (Clift and Schuh 2015 and Luksza 2013). 
-(Kolano et al 2012 maintence of spindle)
-aMTOC migration to pools is dynamic and self organized.
-
-Wu thesis compared th density of spindles on the central and cortical halves
-data suggests that more MTOC are recruited to the cortical side
-
 5. Model prediction

### 2020-03-19 doc/Results.rmd
small updates to files
kinda notes for writing/drafting
added notes on short-IFD effects, might need debugging for knitr

@@
 ```{r Q1.rbar, echo=FALSE}
+
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
@@
 all.male.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "male")#1392
+
+#shape and scale parameter
 
@@
 
+#long biv's - short IFD rates
+
+LongBiv_IFD_mus.rate <- ddply(male.long.biv.2CO_IFD %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+
+                   N.obs  = length(Obj.ID),
+
+                  mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
+                  mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                  var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                  #mean_IFD_short.removed
+                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
+                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)] ),
+
+                inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                n.close.foci = length(which(IFD1_PER <= .3)),
+
+                off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                rate.close.foci = n.close.foci / N.obs
+
+                                              )
+
+#run t.tests on long biv
+LongBiv_IFD_mus.rate <- add_strain(LongBiv_IFD_mus.rate)
+LongBiv_IFD_mus.rate <- add_category(LongBiv_IFD_mus.rate)
+LongBiv_IFD_mus.rate <- add_subsp(LongBiv_IFD_mus.rate)
+
+#add rec.group
+LongBiv_IFD_mus.rate$Rec.group <- ifelse(grepl("PWD", LongBiv_IFD_mus.rate$strain), 1, 
+                                ifelse(grepl("MSM", LongBiv_IFD_mus.rate$strain), 1,
+                                 ifelse(grepl("SKIVE", LongBiv_IFD_mus.rate$strain), 1, 0)))
+
+```
+
+```{r Q2.IFD.long.biv_t.test, echo=FALSE}
+
+#short removed PER
+Short.removed_BivData.LongBiv_IFD.PER_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
+                                 LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
+
+Short.removed_BivData.LongBiv_IFD.PER_Ttest$p.value
+#0.0083
+
+#ABS short removed
+Short.removed_BivData.LongBiv_IFD.ABS_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
+                                 LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
+
+Short.removed_BivData.LongBiv_IFD.ABS_Ttest$p.value
+#0.047
+
 ```
@@
 
+```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
+
+#repeating chrm models with chromsome length
+#FOR BIVALENT LEVEL MODELS
+
+Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
+
+#M.Len.ABS
+Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.raw)
+
+#M.Len.PER
+Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.PER)
+
+#M.Len.raw
+Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
+
+##M.Len PER
+Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+```
+
+```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
+
+#calculate the rates of 'close foci' (<30%)
+#examine the changes in the distribution closer.
+
+#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
+
+close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                      mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
+                      mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                  #mean_IFD_short.removed
+                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
+
+                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)], na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                      n.close.foci = length(which(IFD1_PER <= .3)),
+
+                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                      rate.close.foci = n.close.foci / N.obs
+)
+
+#mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+
+close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
+#add rec.group
+close.IFD.rates.mus.rate$Rec.group <- ifelse(grepl("PWD", close.IFD.rates.mus.rate$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates.mus.rate$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates.mus.rate$strain), 1, 0)))
+
+close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("strain"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm = TRUE),
+
+                        #mean_IFD_short.removed
+                      mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER > .3)),  
+                      n.close.foci = length(which(IFD1_PER < .3)),
+                      rate.close.foci = n.close.foci / N.obs,
+
+                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
+
+)
+
+close.IFD.rates_strain <- add_subsp(close.IFD.rates_strain)
+
+show.IFDrm <- ggplot(close.IFD.rates_strain, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences")+ geom_abline(slope = 1, intercept = 0)
+
+#close.IFD.rates.mus.rate
+show.IFDrm_mus <- ggplot(close.IFD.rates.mus.rate, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences") + geom_abline(slope = 1, intercept = 0)
+
+show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
+
+#add rec.group
+close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
+
+#POOL all short and 
+#chi square test
+high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
+                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
+#p-value = 0.03, slight dif tween 3 high rec strains
+
+#don't need to run chi square for this pattern, since it's enough to 
+```
+
+```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
+
+#Log regression all show significant prediction
+
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.per_bivdata.Mol)
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
+
+#Full dataset
+
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+```
+
 The greater IFD measures in high recombing strains is confirmed by fitting both the IFD^raw^ and IFD^norm^ into the linear model framework, where only the strain effects for the high recombining strains had significant p values (glm; p < 0.05).
@@
 The rate of IFD^norm^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
+
+<REDO tests with the short IFDs removed, where do those above values come from >
+
+```{r Q2.IFD_remove.short, echo=FALSE, include=FALSE}
+#mus
+#PER 
+#biv level
+
+#mus
+
+```
+
+When I compared the normalized IFDs with and without the short IFD's included, -- predictably, many mouse averages for 2CO IFD do nnot change for higher rec strains (but some low too).
+Most of the low rec strains show a change in the PER IFD (but some high rec mice also changed)
+
+For the strain averages, the high rec strains also show low change in the normalized IFD -also these are all higher (0.6)
 
@@
 
-```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
-
-#repeating chrm models with chromsome length
-#FOR BIVALENT LEVEL MODELS
-
-Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
-
-#M.Len.ABS
-Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.raw)
-
-#M.Len.PER
-Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.PER)
-
-#M.Len.raw
-Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
-
-##M.Len PER
-Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-```
-
-```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
-
-#calculate the rates of 'close foci' (<30%)
-#examine the changes in the distribution closer.
-
-#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
-
-close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                      n.close.foci = length(which(IFD1_PER <= .3)),
-
-                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                      rate.close.foci = n.close.foci / N.obs
-)
-
-close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
-
-close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("strain"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER),
-
-                      inverse.close.foci = length(which(IFD1_PER > .3)),  
-                      n.close.foci = length(which(IFD1_PER < .3)),
-                      rate.close.foci = n.close.foci / N.obs,
-
-                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
-
-)
-#the
-
-show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
-
-#add rec.group
-close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
-
-#POOL all short and 
-#chi square test
-high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
-                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
-#p-value = 0.03, slight dif tween 3 high rec strains
-
-#don't need to run chi square for this pattern, since it's enough to
-
-```
-
-```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
-
-#Log regression all show significant prediction
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-```
+```{r Q2.IFD_short.removed_t.test, echo=FALSE}
+
+#Curated_BivData_short.IFD.removed
+#still need to calq the mouse averages
+
+Short.removed_BivData_IFD.PER_Ttest <- t.test(Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 1], 
+                                 Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 0])
+
+Short.removed_BivData_IFD.PER_Ttest$p.value
+
+#
+Short.removed_BivData_IFD.PER_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
+                                 close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
+
+Short.removed_BivData_IFD.PER_Ttest_mus.av$p.value
+
+Short.removed_BivData_IFD.ABS_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
+                                 close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
+
+#p.value
+
+```
+
+The p values are still significant
 

### 2020-03-24 doc/Discussion.rmd
re-wrote outline of chromatin org section

@@
 
-# Explanations and Model testing
-
 ```{r Model.testing, include=FALSE}
@@
 
-This is how we are going to test/ synthesize / fit our results into the literuature and test hypothesis for which models best explain our results.
-In our analysis / synthesis of results, we decded to systematically compared models which have been proposed to explain our results (both for the sex dfferences and the faster male evolution of recombination rates).
-We attempted to test the hypotheses of if our choosen models could produce the main / general patterns oberved in this study (Table X)
-
-We generally focus on models for the evolution of heterochiasmy since that is the primary data/set of results in this study.
-We search for a parsimonius explanation /model that fits both the results from sex differences and the faster male evolution.
-But we also test the male specific traits.
+To synthesize our results, we decded to systematically compare models which have been proposed for explaining the evolution of heterochiasmy or meiotic machinery.
+There are two main types of models: modifers (indirect) modifers or functional (cell physiology based models).
+We attempted to test the hypotheses (mental test for if the model) (Table X).
 
@@
 <we left out Beam film model for XX reasons>
-
-**Conserved heterochiasmy paterns**
-
-Below are the main traits from our results: more between cell variance in females, chromatin compaction (with consequence for interference strength), and broad scale landscape differences.
 
@@
 
-# Chromatin Structure and interference strength
-
-```{r chromatin.organization, include=FALSE}
-
-#1. chromatin organization, female shorter loops, longer axis
-
-#2. A consequence of chromatin organization (loop size) is difference in interference strength in (MB / cM)
-
-#seen in humans, mice (signal comes through even with cell to cell variation in chromatin compaction / organization)
-#
-
-#exceptions to female pattern (fish birds), lack of data
-# interference strength reversal (At and cattle) -- we predict that the loop size will also reverse, -- but there is a lack of cytological data for both sexes.
-
-#lots of data on genetic maps for both sexes, but lack of cytological data to support
-
-#our results, female longer axis, 
-#review caveats -- why we believe the pattern is true
-
-#refs
-#BF klecner?
-#negative relationship of axis and loop size (Lynn? Gruhn?)
-#global re-orgnaiziton of DNA / chromosomes (3D, new Chip-seq pubs)
-```
-
-1. basics and expectations for chromatin organization
-
-2. Well supported model for interference is that it is mediated / communicated through the 2D chromatin axis or SC (depending on the timing)
-
-3. consequence for interference (general patterns) is that interference measured as (mb) will be stronger if DNA loops are stronger
-
-The relationship // molecular organization of chromosome -- undergoes global re-orgzniation during meiosis.
-The basic take away is that eraly stage in meiosis) there is a central axis made up of chromatin and proteins while most of the DNA / chromatin are folded into loops that emenate off of the axis.
-This structure has been called 'lampbrush' by some (cite).
-
-A pattern that has been long noticed --
-
-Merging of two types of data
-
--total chromatin structure (axis length) (cytology)
-
--Interfere measured as cM or DNA base pairs Mb  (the spacing between markers from linkage map)
-
-general pattern for the sexual dimorphsim the relationship between axis length and CO number
--- axis are longer in females than male, suggesting suport more crossovers (more DSBs)
-but the density is less in females.
-
-Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
-Two strains have have more crossovers in males, but females still have longer sc axis.
-_This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape_
-
-**lack of difference in the raw measures of IFD, but significant difference in **
-
-The pattern of interference across raw and normalized measures -- also confirms previous results of male having stronger interfere -- due to the longer loops (Petkov) **these results mirror those seen in** [@petkov2007]
-
-< Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.>
-
-<same pattern in humans (Gruhn et al 2013)
-  (exceptions, birds (fish?) plants?) -- examining this is hard due to chromosome size effecs
-  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes>
-
-2. Short axis / longer DNA loops will have stronger interference
-This prediction combines the findings of chromatin compaction and interference differences which are intertwined by the transmission of interference along the chromosome axis and the inverse relationship chromatin compaction (axis length and DNA loop sizes).
-In physical units (micron) the counting model fits interference well to the 2D linear space of the axis these measures of interference are conserved across sexes in mouse.
-This is because the counting units are the BASES of DNA loops which is conserved along the 2D length of chromatin axis.
-When this space is transformed to account for all of the DNA base pairs which are sequestered into the DNA loops above the axis, the strength of interference will vary depending the on underlying chromatin structure.
-The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
-Thus the strength of interference will be stronger because with each step (count) along the 2D axis,
-more bases pairs are being skipped above the axis.
-
 # Greater between cell variation
@@
 
+# Chromatin Structure and interference strength
+
+```{r chromatin.organization, include=FALSE}
+#1. basics and expectations for chromatin organization
+#seen in humans, mice (signal comes through even with cell to cell variation in chromatin compaction / organization)
+
+#2. how the model treat this limiting factor
+
+#3. consequence for interference (general patterns) is that interference measured as (mb) will be stronger if DNA loops are stronger
+#prediction for general -- pattern
+```
+
+0.
+Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
+Two strains have have more crossovers in males, but females still have longer sc axis.
+_This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy_
+
+1. the Chromosome axis length // chromatin organizatio in meiotisis
+
+according to the tethered loop model of DSB formation -- the 'real estate' for COs are the DNA loops - the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
+When the total amount of DNA is kept constant as for a given Chromsome, there is a negative relationship between the loop length and axi length.
+**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**   This structure has been called 'lampbrush' by some (cite).
+
+2. Only one of the models we examined predicts sexual dimorphism for axis length.
+
+The modeifer models do not parameterize this since they do not models / are not built to consider single meiotic events.
+
+**COM** predicts that axis length will be larger in oocytes due to the larger volumne -- this would help in the bouquet stage -- may be connected to the 'the search' during homolog pairing.
+
+the **spindle based selection** model does not specific predictions about this aspect -- assumes that the 2D areas for chromosomes are proportaional across stages (during the decondensing stage between pachtyene and dipl?
+/ metaphase) -- the chromosomes condenses uniformly across the length.
+
+3.
+another thing to point out / regaurding this model -- is the connection to sex difference in interference strength.
+Our results of conserved within strain raw IFD lengths - yet when the IFDs are controled for total SC length, we observed sexual dimorphism with males having longer IFD^normalized^ (stronger interference), mirror those observed in [@petkov2007].
+The model / explaination proposed in that paper is that The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
+Thus the strength of interference will be stronger because with each step (count) along the 2D axis, more bases pairs are being skipped above the axis.
+These findings point out two things, This part of heterochiasmy generates testable prediction: First, the interference strength measured in chromsoome axis units (or SC) is generally conserved between sexes.
+Second the sex with longer axis and shorter loops will have weaker interference compared to the other sex.
+Or conversely, the sex with stronger linkage map based interference woud have chromatin organization with shorter axis compared to the other sex.
+There is a plefora of data of sexual dimorphism in linkage map based interference strength indeicating sexual dimorphism, but comparative cytological data is lacking (difference in perparness of some sexes and complications due to idenification of specific chromosomes between cells.)
+
 # Evolved Traits
@@
 
-#revrsed heterochiasmy direction
+#tricker to fit the models to this, 
+#paradox -- how the broad/general patterns from the sex differences
+
+# while DSBs have evolved between these groups - not integrated into the table for models -- since most models don't integrate this stage of the pathway.
+
+#1.reversed heterochiasmy direction
 #faster -- rapid evoluiton in males
 
-#this also goes along with
+#2 interference strength (intro)
 
-# interference strength
+#SACE could fit ...
 
+#spindle based selection 
 ```
 
-Some of the traits related to faster male evolution -- distinguish in which models can redict them:
+- the male specific differences in meiotic traits are a bit trickier to explain using the above models (since they were designed to explain broad scale pattern).
 
-<We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
-One of our primary objectives was to identify other traits which correlate (could explain the increase).>
+The general patterns observed from the within strain / between sex data --generate predictions for the relationship between chromatin compaction and interference strength.
+(the predictions are).
+Based on the general / braod scale patterns, those predictions are not met -- but we argue that our data are not suited / don't have enough power to -- overcome the weaker signal within the nose of chromosome specific and within nucleus correlation in chromatin compaction.
+(also distingushing the recombination rate evolution at this scale is not the focus of this manuscript.)
 
-Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength.
-<While SC length and the telomere distances are similar / conserved across the mouse strains>
+Some of the traits related to faster male evolution -- distinguish in which models can redict them: the reverseal of heterochiasmy direction and the correlated evolution of gwRR and interference strength.
 
-<We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>
-
-# Evolved rates of precursors, DSBs
-
-```{r evo.DSB, include=FALSE}
-
-```
-
-<define CO:NCO decision, tethered-loop DSB formation SPO11 (cite, paigen? deMassy
-setup baier study,  explain >
-
+**DSB evolved but, don't included as an examined trait, since most evolution based models don't make predictions for this stage of meiosis.**
 _the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
 during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
-
-Our results find evolution between a subset of strains -- demonstrating that the high rec strains have significant higher DSB number.
-
 Our results mirror those found in Baier et al [@baier2014]
@@
 
-Baier et al found significant differences in sc legnths for 
-The tests in our study - show moderate to low significance in the differences in SC lengths. (this is because we have more noise-can't distinguish individual chromosomes)
+# revered heterochiasmy
 
-a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
--different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures
+```{r reverse.HetC, echo=FALSE}
 
-our results are similar in several ways
-- correlation of final MLH1 and DBS counts, SC measues correlate with the DSB number (even tho the differnces we observe in our study are -- slight / less significant -- we attribute that to chromosome specific noise )
+```
 
-Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al).
+models for haploid selection -- this **reversal of the direction of heterochiasmy** would be accompanied by (vast changes in haploid stages or degree of imprinting)
+
+the meiotic drive -- model, females should be predicted to have larger linkage maps/gwRR since they are more exposed / weaker to true meiotic drivers.
+
+A reverselal in heteroschiasmy is not clearly predicted by the sace model (unless with the -- except under the conditions discribed below)
 
@@
 
-#models that predict this pattern?
+#models that predict this pattern
 
@@
 
-1. Review results
+We had an unexpectedd result of correlated evolution of gwRR with interference strength. This is unexpected beacuse...
+
 -The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.
@@
  the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).
+
+**SACE**
+a reversal in heterochiasmy direction would make sense (under the sae model) if it maintained 
+this isn't an explict prediction, but accordding this this model -- stronger interference (such as we observed) would have the consequence of maintaining larger blocks of chromosomes together.
 
@@
 # Reference
+
+<Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.
+
+same pattern in humans (Gruhn et al 2013)
+  (exceptions, birds (fish?) plants?) -- examining this is hard due to chromosome size effecs
+  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes
+
+2. Short axis / longer DNA loops will have stronger interference
+This prediction combines the findings of chromatin compaction and interference differences which are intertwined by the transmission of interference along the chromosome axis and the inverse relationship chromatin compaction (axis length and DNA loop sizes).
+In physical units (micron) the counting model fits interference well to the 2D linear space of the axis these measures of interference are conserved across sexes in mouse.
+This is because the counting units are the BASES of DNA loops which is conserved along the 2D length of chromatin axis.
+When this space is transformed to account for all of the DNA base pairs which are sequestered into the DNA loops above the axis, the strength of interference will vary depending the on underlying chromatin structure.>
+
+<We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
+One of our primary objectives was to identify other traits which correlate (could explain the increase).>
+
+<Baier et al found significant differences in sc legnths for 
+The tests in our study - show moderate to low significance in the differences in SC lengths. (this is because we have more noise-can't distinguish individual chromosomes)
+
+a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
+-different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures
+
+our results are similar in several ways
+- correlation of final MLH1 and DBS counts, SC measues correlate with the DSB number (even tho the differnces we observe in our study are -- slight / less significant -- we attribute that to chromosome specific noise )>
+
+<We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>

### 2020-03-24 doc/Discussion.rmd
edited outline for reversed HetC section

@@
 ```
+
+Advice from Bret:
+
+- Don't know where to fit in DSB evolution results
 
@@
 
-3.
-another thing to point out / regaurding this model -- is the connection to sex difference in interference strength.
+Another thing to point out / regaurding this model -- is the connection to sex difference in interference strength.
 Our results of conserved within strain raw IFD lengths - yet when the IFDs are controled for total SC length, we observed sexual dimorphism with males having longer IFD^normalized^ (stronger interference), mirror those observed in [@petkov2007].
@@
 
-```{r Evolve.tratis.outline, include=FALSE}
-
-#it's not the sex differences that distinguish
+```{r Evolve.traits.outline, include=FALSE}
 
@@
 
-# while DSBs have evolved between these groups - not integrated into the table for models -- since most models don't integrate this stage of the pathway.
-
-#1.reversed heterochiasmy direction
-#faster -- rapid evoluiton in males
-
-#2 interference strength (intro)
-
-#SACE could fit ...
-
-#spindle based selection 
-```
-
-- the male specific differences in meiotic traits are a bit trickier to explain using the above models (since they were designed to explain broad scale pattern).
-
-The general patterns observed from the within strain / between sex data --generate predictions for the relationship between chromatin compaction and interference strength.
-(the predictions are).
-Based on the general / braod scale patterns, those predictions are not met -- but we argue that our data are not suited / don't have enough power to -- overcome the weaker signal within the nose of chromosome specific and within nucleus correlation in chromatin compaction.
-(also distingushing the recombination rate evolution at this scale is not the focus of this manuscript.)
-
-Some of the traits related to faster male evolution -- distinguish in which models can redict them: the reverseal of heterochiasmy direction and the correlated evolution of gwRR and interference strength.
-
-**DSB evolved but, don't included as an examined trait, since most evolution based models don't make predictions for this stage of meiosis.**
-_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
-during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
-Our results mirror those found in Baier et al [@baier2014]
-
-Our results show that the ratios of DSB : CO 
-**range from X to X which is similar to those presented in [@baier2014] (which report X to X) **
-
-These conserved ratios suggest that the increase in DSB numbers are accompanied by proportional increase in crossovers.
-
-This shifts support away from evolution at the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
-**increase in DSBs is driven by the number of DNA loops** and due to the inverse relationship of loop and chromatin axis length, this would result in longer sc lengths (which reflect the underlying chromatin axis length).
-
-# revered heterochiasmy
-
-```{r reverse.HetC, echo=FALSE}
+#models weren't designed to explain within species variation in recombination rates 
+#or evolution
 
@@
 
-models for haploid selection -- this **reversal of the direction of heterochiasmy** would be accompanied by (vast changes in haploid stages or degree of imprinting)
+A thing about this study that we were very interested in was examining the faster evolution in males with the high recombining rates.
 
-the meiotic drive -- model, females should be predicted to have larger linkage maps/gwRR since they are more exposed / weaker to true meiotic drivers.
+BUT applying these above models to the traits connected to the faster male evolution is a bit tricker since none of these moedls were designed to explain within species variation in recombination rates.
+(most were designed to explain the general pattern of sexual dimorphism in meiotic recombination rates.)
 
-A reverselal in heteroschiasmy is not clearly predicted by the sace model (unless with the -- except under the conditions discribed below)
+# Revered heterochiasmy
 
-# evolution in Interference
+```{r reverse.HetC, echo=FALSE}
+# transition connection to the hetC models
+
+#our results - total gwRR isn't the most conserved, (we observed 3 other traits more)
+
+#what does a reversal of HetC imply?  gwRR
+#driven by faster evolution in males
+
+#our model had faster evolution in males resulting in a reversal of the canonical pattern of heterichiasmy ( F gwRR > M gwRR)
+#other examples, cattle, marsupials
+
+#what do models say about faster evolution in one sex? (yes -- but almost always for canonical pattern of heterochiasmy)
+
+```
+
+Our results demonstrate that total gwRR isn't the most conserved, (we observed 3 other traits more conserved sexual dimorphism examined above).
+Also our data shovariyng degrees of heterochiasmy magnitude (0 to gough).
+We the general pattern 2 strains attribute this pattern of reversed heterochiasmy to faster evolution of gwRR in two strains males.
+Which models predict dfferent rates of evolution between the sexes.
+Some of the models could explain a reversal of the direction of sexaul dimrohpsim, however most all of the predictions for faster evolution result in males having reduced recombination rates.
+Under the haploid selection model a  **reversal of the direction of heterochiasmy** would be accompanied by vast changes in haploid stages or degree of imprinting.
+Following the meiotic drive based model, females should be predicted to have larger linkage maps/gwRR since they are more exposed / weaker to true meiotic drivers.
+A reversal in heteroschiasmy is not clearly predicted by the SACE model (unless with the -- except under the conditions discribed below).
+Under the sindle selection model, the male (centrosome/effect SAC) metaphase could drive directional selection (and faster evolution) in males compared to female.
+However the mechanism/aspect of different fitness is not clear.
+Under the COM model, evolutionary predictions are not clear (explicit), but following the logic would come from changes in axis length (to generate another wave point) -- which might be due to meiocyte volumne.
+
+# Evolution in Interference
 
@@
 
+<Some of the traits related to faster male evolution -- distinguish in which models can redict them: the reverseal of heterochiasmy direction and the correlated evolution of gwRR and interference strength.
+
+**DSB evolved but, don't included as an examined trait, since most evolution based models don't make predictions for this stage of meiosis.**
+_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
+during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
+Our results mirror those found in Baier et al [@baier2014]
+
+Our results show that the ratios of DSB : CO 
+**range from X to X which is similar to those presented in [@baier2014] (which report X to X) **
+
+These conserved ratios suggest that the increase in DSB numbers are accompanied by proportional increase in crossovers.
+
+This shifts support away from evolution at the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
+**increase in DSBs is driven by the number of DNA loops** and due to the inverse relationship of loop and chromatin axis length, this would result in longer sc lengths (which reflect the underlying chromatin axis length).>
+
 <Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.

### 2020-03-25 doc/Discussion.rmd
Trimming down the discussion part

@@
 - Don't know where to fit in DSB evolution results
+
+- Keeping interference section breif (background, explaningillustrating spindle based selection model)
 
@@
 A thing about this study that we were very interested in was examining the faster evolution in males with the high recombining rates.
-
 BUT applying these above models to the traits connected to the faster male evolution is a bit tricker since none of these moedls were designed to explain within species variation in recombination rates.
@@
 ```{r evo.interference, include=FALSE}
+#review results,
+#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
+#(we're gonna spend alot of space on this section)
 
-#review results,
+# more convincing that the pattern is real
 
-#why unexpected
+#lil background, for interferences 
+#assuptions for predictions (most modifer models, )
 
-#assuptions for predictions
+#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)
 
-#models that predict this pattern
+#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)
 
@@
 
-We had an unexpectedd result of correlated evolution of gwRR with interference strength. This is unexpected beacuse...
-
--The rapid evolution of the genome wide rate in our high rec strains is due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.
-
+The strongest correlate of faster gwRR in males was intercrease interference strength.
+This is unexpected beacuse most literature and logical models assoicate more dense spacing of crossovers along chromosomes.
+Luckily having unexpected results is good for distingushing the most parimonius model.
 These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol.
@@
 
-Yet (we are confident that this pattern is not due to errors).
+**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
 
- the IFD distributions show less variance in the high rec strains, which would have MORE Chromosomes (Chm1 to Chm7) compared to the low rec strains which have fewer 2Cos (presumably limited to large chromosomes).
+We considered carefully if this pattern was due to error -- and we are confident that this pattern is genuine -- goes aginst expected noise ALSO this is observed from other work in this lab -- and some empiracal studies (cattle) (plant bomblies).
+
+<while these results don't clearly fit the chromatin organization and interference strength prediction... determining the clean pattern is beyond the scope of this paper>
+
+HOW do the models above treat this unexpected correlation?
+
+**typical modifer models -- are not good for predicting interferece (in mode that is relevant to these types of results) since the models don't parameterize for indiviual meiosis events.** We conclude that under the haploid selection and meiotic driver model any prediction for evolution of interference strength follows from the 'logical' expectation of a negative correlation.
+
+originally proposed/outlined by Trivers [] and refined /formatlized in [@lenormand2003].
+This model discribes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombiaion rates between males and females.
+
+ This is a modefier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifer).
+
+This model predicts the evolution of heterchiasmy -- through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
 
 **SACE**
-a reversal in heterochiasmy direction would make sense (under the sae model) if it maintained 
-this isn't an explict prediction, but accordding this this model -- stronger interference (such as we observed) would have the consequence of maintaining larger blocks of chromosomes together.
+While the SACE model doesn't explecitly make predictions aof the out come for interference strength (or averge spacing of crossovers, )
+We note that the main predicted pattern under this model of maintaining larger blocks of chromosomes in linkage can asrise from stronger interference / greater spacing of the distance between two foci (also maximize rbar).
 
--Also:
-	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.
+ extends the lenomand focusing on two conditions.
+first that the stength of epistatsis in dependant on the cis or trans nature.
+the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)
 
-	-The difference in interference patterns is driven by the difference in the rates of small IFD observations, (mean vs median). a pattern that is suggestive of selection agisnt short IFDs
+ second that the relative interference strengths differnt between male and female.  This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
 
+Under the **COM** model -- the placement of crossovers (the recombination landscape) is generated by 'waves' / movement from the bouquet stage.
+It is conceivable that the pattern of stronger interference could be generated from the waves/movements of chrmosomes in the early prophase stage this would come from a change in the movement or extension of the axis length.
+
+**spindle based selection model**
 There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
@@
 
- 1.Gamete selection
-
-originally proposed/outlined by Trivers [] and refined /formatlized in [@lenormand2003].
-This model discribes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombiaion rates between males and females.
-
- This is a modefier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifer).
-
-This model predicts the evolution of heterchiasmy -- through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
-
-<the sex with the most variable reproductive fitness will be selected to have lower genome wide recombination rates>
-
-From our results, this model would predict -- fit the results of:
-
-this model doesn't predict that the heterochismy direction would reverse.
-
-2. SACE
-
-Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.
-
--- this model is an extension of the modifer model presented in Lenormand
-
--- (basics)
- extends the lenomand focusing on two conditions.
-first that the stength of epistatsis in dependant on the cis or trans nature.
-the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)
-
- second that the relative interference strengths differnt between male and female.  This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
-
 This new model has **strengths**
@@
  - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
-
 - Other ways this model departs from our results, 
-
 -Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
-
 However / while not explicitaly arameterized or predicted in this model,
@@
 
- 3.Two-locus (protect for drivers)
+# REVIEW
 
-This second modifer model was persented in [@brandvain2012scrambling] the basics of the model are that females will have a longer genetic map and more centromere biased recombination landscape compared to males.
+```{r final.review.outline, echo=FALSE}
+# comes down to explaning the interference correlation
 
-A strength of this model is that is presents a clear driver / source of differential fitness (ie meiotic drivers can decrease fitness/fertility).
-This model also accounts for the general female pattern,
+# spindel selection -- 
+#good for sex differences because it frames heterochiasmy within the context of meiosis
 
-However, this model doesn't explain the borad scale pattern for the male pattern.
+#can fit the interference evolution due to the (pressumed) effect on bivalent shape
 
-**weaknesses**  (modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
+#caveats and unanswered parts
 
-this model would predict / fit the results of
+# SACE is runner up
 
-<**model basics**, 'break up driv system consiracies'>
-
- 4. Segregation and physiological Spindle Differences
-
-```{r Spindle.Model, include=FALSE}
-
-#1. context of meiosis
-
-#2. 
 ```
@@
 the study confirmed preivously identified sexual dimorhisc patterns in regard to the recombiantion landscape (in mammals) -- in the goal of unifying all of these results, we sough to investigate physiologycal aspects of gametogenesis which have conserved sexual dimorphic features.
-
-LOOK at the oogenesis and spermatogenesis as the context for meiosis (and how )
-
-lead us to ask, how the different contexts of meiosis could influence recombination rates (the recombinaiton landscape)
-
-(not sex chromosome)
-
-while there are many different forms, flavors of anisogamy (cite that one paper)
-
-symetrical vs asymetrical gamete division
 
@@
 
-for the distinct **evolutionary patterns** selection on the recombination landscape via the spindle (segregation checkpoint),
-
-_leaky SAC_ for oocytes (is known) .. meaning that bivalents / homologs without proper tension force
-
-(potentially this bad checkpoint is adaptive for thwarting centromere drive systems, which work through the SAC and tension signals)
-
-how the SAC spindle works, correctly
-
-evidence that the oocyte spindle is leaky
-
-_precition that the leaky spindle makes relaxes selection_
-
-aTMOC's  (because there are mulitple, these could be uneven across the cell)
-
-_asymetrical division_
-
-this makes egg weak to driving centromeres which (are segregation distortors), that means some centromeres segreation (rotate to the polar body side of the egg)
-
-_number of placement of crossovers influencing spindle_
-
-For this model we propose that the number and placement of crossovers -- will affect the
-
-_strengths of model_
-
-_weakness_
-
-No very few spermatocyte spindle data  (it's assumed to )
-
-_review the correct SAC_
-
-References:
-Schuh Ellenberg 2007 spindle
-
-5. Model prediction
-The models above which predict directional selection on the recombination landscape in males: SACE and Spindle differences are most parsimonious for the models review in this paper.
-The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.
-
 Consequences of evolution in the number and placement of crossovers / evolution of interference strength:
@@
 # Reference
+
+<5. Model prediction
+The models above which predict directional selection on the recombination landscape in males: SACE and Spindle based selection  are most parsimonious for the models review in this paper.
+The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.>
+
+<3.Two-locus (protect for drivers)
+This second modifer model was persented in [@brandvain2012scrambling] the basics of the model are that females will have a longer genetic map and more centromere biased recombination landscape compared to males.
+
+A strength of this model is that is presents a clear driver / source of differential fitness (ie meiotic drivers can decrease fitness/fertility).
+This model also accounts for the general female pattern,
+
+However, this model doesn't explain the borad scale pattern for the male pattern.
+
+**weaknesses**  (modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
+
+this model would predict / fit the results of>
+
+<**model basics**, 'break up driv system consiracies'>
+
+<Also:
+	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.>
 

### 2020-03-25 doc/Discussion.rmd
Interference parts

@@
 
+#**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
+
 ```
 
-The strongest correlate of faster gwRR in males was intercrease interference strength.
-This is unexpected beacuse most literature and logical models assoicate more dense spacing of crossovers along chromosomes.
-Luckily having unexpected results is good for distingushing the most parimonius model.
-These results were unexcepted and differ ffrom the major patterns listed in the literature (otto payseur, argawal Korol.
-(it also goes against the logical assumptions models of fitting more crossovers along the SC chromosome axis space)
+Our results indecated that interference strength has evolved with the genome wide recombiantion rates in the three strains with faster male evolution.
 
-**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
+ This is unexpected because most literature and logical models are built / assume that weaker dense spacing of crossovers along chromosomes (ie. weaker interference) is how .
 
-We considered carefully if this pattern was due to error -- and we are confident that this pattern is genuine -- goes aginst expected noise ALSO this is observed from other work in this lab -- and some empiracal studies (cattle) (plant bomblies).
+The history, models and empiracal examples are extensively reviewed in [@ottoPaysuer2019].
+It is notable that borad discriptions / broad scale patterns for the correlation of gwRR and interference strength (gamma parameter) do fit the expected negative pattern [@ottoPaysuer2019].
 
-<while these results don't clearly fit the chromatin organization and interference strength prediction... determining the clean pattern is beyond the scope of this paper>
+(genome wide rate vs single bivalent measures)
 
-HOW do the models above treat this unexpected correlation?
+**Caveats**
+We considered carefully if this pattern was due to error and we are confident that this pattern is genuine.
 
-**typical modifer models -- are not good for predicting interferece (in mode that is relevant to these types of results) since the models don't parameterize for indiviual meiosis events.** We conclude that under the haploid selection and meiotic driver model any prediction for evolution of interference strength follows from the 'logical' expectation of a negative correlation.
+There's no model that fits evolution of heterochiasmy and interference independantly.
+Under our undertsanding of the models, the relatioship of genome wide recombiantion rate and interference strength is assumed to be negative.
+We now review how some of the assumptions of typical modifer models might not fit realities of the meiotic pathway.
 
-originally proposed/outlined by Trivers [] and refined /formatlized in [@lenormand2003].
-This model discribes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombiaion rates between males and females.
+First, the **typical modifer models do not model single meiotic events**. (think about how pooling man meiotic events can skew patterns)
+(crossover interference can be quantified from linkage maps (which are the average of many hundreds to thousands of meiotic events).
+It can also be discribed / measrued /quantified from single bivalets / chromosomes (either cytology of dense genotyping of single cells (cite).
+While it's widely agreed that they both reflect the same biological process, these two approaches are fundamentally different. (they are subject to distinct biases and noise).
 
- This is a modefier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifer).
+Second, most modifer models use abstracted genomic space -- like an infite 2D line between markers.
+(space between markers is abstracted).
+This is useful for fitting the gamma model, but within the cell, the physical chromosomes's axis length is fitnite and crossovers cannot form beyond those bounds.
+
+We conclude that under the haploid selection and meiotic driver model any prediction for evolution of interference strength follows from the 'logical' expectation of a negative correlation.
+
+originally proposed/outlined by Trivers [] and refined /formalized in [@lenormand2003].
+This model describes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombination rates between males and females.
+
+ This is a modifier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifier).
 
@@
 **spindle based selection model**
-There is a model that predicts this positive ocrrelation. The goldstein results (reviewed in ottoPayseur). when the number of crossovers (the obligate crossover assumption is more strict / the
-breifly, the model simulates / sets up a scnario where a modifer acts through the probabilty of flanking regions having 0 or 2 crossovers.
-ith crossover assuance, recombination rate increases the further apart the loci are spaced (?)
+There is a model that predicts this positive correlation between genome wide recombination rate and interference. The goldstein results (reviewed in).
 
-Additionally there are a handful of empircal results which also show the unexpected pssotive correlation between gwRR and interference strength.
+The results from golstein reviewed in [@ottoPaysuer2019], predict a modeifier model (which placed more weight on crossover assurance (holds the number of crossovers constant) and simulates a scnario where a modifer acts on modulating the probabilty of flanking regions having 0 or 2 crossovers.
+Which has the result of increasing the distance between two foci.
 
-1.
-Cattle (Ma?) [@ma2015_cattle] ?
-(or newer one), when comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter)
+Additionally there are a handful of empircal results which also show the unexpected possitive correlation between gwRR and interference strength.
+When comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter) [@ma2015_cattle].
+In *Peromyscus leucopus*, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019].
 
-2. The IFDs from (Beth's cross) a related house mouse cross show that the PWD (same strain, different mice) had longer IFD compared to Cast (HVR unpublished data). (also gamma?)
+Additionaly -- preliminary data from a previous cross using MLH1 cytology in spermatocytes in house mouse replicated these results with IFD measures from  M.
+m.
+musculus^PWD^  being longer compared to M.
+m.
+castenus^Cast^ (HVR unpublished data) (cast has low gwRR).
 
-3. In Peromyscus, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019]
-
-<note about the differences in the patterns>  assumptions with gamma dstribution and linkage maps, (closer to infinte genetic space), less weight on crossover assurance - more weigth on 
+note about the differences in the patterns>  assumptions with gamma dstribution and linkage maps, (closer to infinte genetic space), less weight on crossover assurance - more weigth on 
 (pooled SC/chromosome length gives strong correlation )
 
-<which is much different than when a single chromosome is examined (by itself), finite 2D length, crossover assurance, weaker correlation with ((the closer to chromosome level for SC legnth metric / summary, the weaker the correlation with crossover number))
+which is much different than when a single chromosome is examined (by itself), finite 2D length, crossover assurance, weaker correlation with ((the closer to chromosome level for SC legnth metric / summary, the weaker the correlation with crossover number))
 
@@
 
-4. HYPER-Crossover experiments. (review the effects when crossover number is artificially altered)
-
-This new model has **strengths**
-
+This new model has **strengths**:
   - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)
@@
 - extends modifer model to explain a chromosome based / level pattern
-
  _weaknesses_, and ways the model diverges from our results
@@
  (as outlined in Veller et al)
-
 our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle
@@
 
-# The CONTEXT of MEIOSIS
+## The CONTEXT of MEIOSIS
 
 the study confirmed preivously identified sexual dimorhisc patterns in regard to the recombiantion landscape (in mammals) -- in the goal of unifying all of these results, we sough to investigate physiologycal aspects of gametogenesis which have conserved sexual dimorphic features.
-
 A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle).
@@
 
-_main sex differences in spindle_ are the anchoring of MT at opposite poles
+**main sex differences in spindle** are the anchoring of MT at opposite poles
 
@@
 
-# CO placement translates to bivalent shape
+#CO placement translates to bivalent shape
 
@@
 
-# Future Steps
+#Future Steps
 
-be aware of sex-specific and sex average biases
+Be aware of sex-specific and sex average biases.
+Some cytlogy models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites.
+(lack of data from one sex).
+Recognize the value of Chromosome level measures and incorporating chromosome behavior into models.
 
-- linkage / R between two foci is sex averaged
+Borader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution.
+Utilize mitosis systems given the conservation of the molecular pathway.
+THink of meiosis as a program.
 
-- Celegans -- cytlogy focuses on oogenesis  and genetic maps are often from hermaphrodites (selfing -- sperm and egg landscapes are merged in hemaphordites)
+Dark matter of the genome.
+We can't really explude that some of the recombination rate variation is due to strucutreal variation (more of less genome).
+(since everything is mapped to B6).
+(also maybe there are crossovers in the centromeres and telomeres -- which might escape detected in linkage maps)
 
--- the value of Chromosome level measures -- incorporate chromosome behavior into models
+what are the rates of sister-sister crossovers?
 
--  more attention to cell physiology and biophysics (conserved nature of meiosis pathway (and mitosis ) -- use some models such as Sc for manipulation) (Murkarjee gardner -- quantify the spindle tension
+How does elevated rates of precocious sister centromere seperation effect oogenesis?
 
--- lack of SC length prophase cell charaecterization
--- comparisons within and between sexes (that cattle paper)
+Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).
 
-# Reference
+Keep generating those testable hypotheses (cite).
+
+#blah
 
@@
 
-However, this model doesn't explain the borad scale pattern for the male pattern.
-
-**weaknesses**  (modifers - driver systems everywhere!), impossible to verify, doesn't account for male landscape pattern (supression from centromere to end)
-
-this model would predict / fit the results of>
-
-<**model basics**, 'break up driv system consiracies'>
+However, this model doesn't explain the borad scale pattern for the male pattern.>
 
@@
 <We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>
+
+#References

### 2020-03-27 doc/Discussion.rmd
small notes

@@
 
-#-replicate some previous results
-#(Gruhn et al), Kong et al?, cattle,
-
 #between cell variance (Veller Kleckner), mechnism
@@
 
-#sources of within mouse variance; age effects,  Hassold hunt
-
 #models that predict more within animal variance
+#none of the models expliciltly predict fr traits at the between cell level
+#so any between cell variation -- patterns is due to DIFFERENCES IN THE STRENGTH OF SELECTION
 

### 2020-03-27 doc/Discussion.rmd
small clean up

@@
 
+We focus first on the results for sexually dimorphic traits.
+
 # 'Typical landscape'
@@
 
-<Broad Scale crossover placement patterns, male telomeric, female centromeric,
-review the typical landscape, one of the major broad scale patterns,
-exceptions
-consequences, (differences in variation, Sardell paper)>
-
 The 'typical landscapes' refers to the sex differences in the board scale pattern or general placement of crossovers along the length of chromosomes.
@@
 s. )>
+
+The conclusions from the models -- basically all of the models could account for differences in the general recombination landscape patterns.
 
@@
 
-While these (age, technical noise, ect) are potential sources of technical noise -- these results indecate this isn't due to technical noise or age effects.
-
 Transcriptional noise and complex traits both show the same repsonses to different modes of selection, lower variation under stabilizing selection and greater variaion with relaxed selection.
@@
 so all of the models could reasonably predict the above results ... examining some of the traits related to the faster evolution in the male specific data indecate different support across the models.
+
+None of the models make predictions for between cell variation -- so we draw conclusions from their preciations based on the strength of interference.
 
@@
 
-1. the Chromosome axis length // chromatin organizatio in meiotisis
+**chromatin organization**
 
-according to the tethered loop model of DSB formation -- the 'real estate' for COs are the DNA loops - the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
-When the total amount of DNA is kept constant as for a given Chromsome, there is a negative relationship between the loop length and axi length.
-**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**   This structure has been called 'lampbrush' by some (cite).
+According to the tethered loop model of DSB formation (cite) -- the 'real estate' for COs are the DNA loops - the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
+**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**
 
-2. Only one of the models we examined predicts sexual dimorphism for axis length.
+2. Only one of the models we examined aligns with our observed results for sexual dimorphism for axis length.
 
-The modeifer models do not parameterize this since they do not models / are not built to consider single meiotic events.
+The modifier models do not parameterize this since they do not models / are not built to consider single meiotic events.
 
-**COM** predicts that axis length will be larger in oocytes due to the larger volumne -- this would help in the bouquet stage -- may be connected to the 'the search' during homolog pairing.
+**COM** predicts that axis length will be larger in oocytes due to the larger volume -- this would help in the bouquet stage -- may be connected to the 'the search' during homolog pairing.
 
-the **spindle based selection** model does not specific predictions about this aspect -- assumes that the 2D areas for chromosomes are proportaional across stages (during the decondensing stage between pachtyene and dipl?
+the **spindle based selection** model does not specific predictions about this aspect -- assumes that the 2D areas for chromosomes are proportional across stages (during the decondensing stage between pachtyene and dipl?
 / metaphase) -- the chromosomes condenses uniformly across the length.
 
-Another thing to point out / regaurding this model -- is the connection to sex difference in interference strength.
-Our results of conserved within strain raw IFD lengths - yet when the IFDs are controled for total SC length, we observed sexual dimorphism with males having longer IFD^normalized^ (stronger interference), mirror those observed in [@petkov2007].
-The model / explaination proposed in that paper is that The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
+Another thing to point out / regarding this model -- is the connection to sex difference in interference strength.
+Our results of conserved within strain raw IFD lengths - yet when the IFDs are controlled for total SC length, we observed sexual dimorphism with males having longer IFD^normalized^ (stronger interference), mirror those observed in [@petkov2007].
+The model / explanation proposed in that paper is that The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
 Thus the strength of interference will be stronger because with each step (count) along the 2D axis, more bases pairs are being skipped above the axis.
-These findings point out two things, This part of heterochiasmy generates testable prediction: First, the interference strength measured in chromsoome axis units (or SC) is generally conserved between sexes.
+These findings point out two things, This part of heterochiasmy generates testable prediction: First, the interference strength measured in chromosome axis units (or SC) is generally conserved between sexes.
 Second the sex with longer axis and shorter loops will have weaker interference compared to the other sex.
-Or conversely, the sex with stronger linkage map based interference woud have chromatin organization with shorter axis compared to the other sex.
-There is a plefora of data of sexual dimorphism in linkage map based interference strength indeicating sexual dimorphism, but comparative cytological data is lacking (difference in perparness of some sexes and complications due to idenification of specific chromosomes between cells.)
+Or conversely, the sex with stronger linkage map based interference would have chromatin organization with shorter axis compared to the other sex.
+There is a plethora of data of sexual dimorphism in linkage map based interference strength indicating sexual dimorphism, but comparative cytological data is lacking (difference in perparness of some sexes and complications due to identification of specific chromosomes between cells.)
 
@@
 
-A thing about this study that we were very interested in was examining the faster evolution in males with the high recombining rates.
-BUT applying these above models to the traits connected to the faster male evolution is a bit tricker since none of these moedls were designed to explain within species variation in recombination rates.
+The faster evolution in males with the high recombining rates.
+BUT applying these above models to the traits connected to the faster male evolution is a bit trickier since none of these models were designed to explain within species variation in recombination rates.
 (most were designed to explain the general pattern of sexual dimorphism in meiotic recombination rates.)
@@
 
-Our results demonstrate that total gwRR isn't the most conserved, (we observed 3 other traits more conserved sexual dimorphism examined above).
-Also our data shovariyng degrees of heterochiasmy magnitude (0 to gough).
+Our results demonstrate that gwRR is decoupled from other sexually dimorphic meotic traits, (we observed 3 other traits more conserved sexual dimorphism examined above).
 We the general pattern 2 strains attribute this pattern of reversed heterochiasmy to faster evolution of gwRR in two strains males.
-Which models predict dfferent rates of evolution between the sexes.
-Some of the models could explain a reversal of the direction of sexaul dimrohpsim, however most all of the predictions for faster evolution result in males having reduced recombination rates.
+
+Which models predict different rates of evolution between the sexes.
+Some of the models could explain a reversal of the direction of sexual dimrohpsim, however most all of the predictions for faster evolution result in males having reduced recombination rates.
 Under the haploid selection model a  **reversal of the direction of heterochiasmy** would be accompanied by vast changes in haploid stages or degree of imprinting.
 Following the meiotic drive based model, females should be predicted to have larger linkage maps/gwRR since they are more exposed / weaker to true meiotic drivers.
-A reversal in heteroschiasmy is not clearly predicted by the SACE model (unless with the -- except under the conditions discribed below).
-Under the sindle selection model, the male (centrosome/effect SAC) metaphase could drive directional selection (and faster evolution) in males compared to female.
+A reversal in heteroschiasmy is not clearly predicted by the SACE model (unless with the -- except under the conditions described below).
+Under the spindle selection model, the male (centrosome/effect SAC) metaphase could drive directional selection (and faster evolution) in males compared to female.
 However the mechanism/aspect of different fitness is not clear.
-Under the COM model, evolutionary predictions are not clear (explicit), but following the logic would come from changes in axis length (to generate another wave point) -- which might be due to meiocyte volumne.
+Under the COM model, evolutionary predictions are not clear (explicit), but following the logic would come from changes in axis length (to generate another wave point) -- which might be due to meiocyte volume.
 
-# Evolution in Interference
+# Evolution of Interference
 
@@
 
-Our results indecated that interference strength has evolved with the genome wide recombiantion rates in the three strains with faster male evolution.
+Our results indicated that interference strength has evolved with the genome wide recombination rates in the three strains with faster male evolution.
 
- This is unexpected because most literature and logical models are built / assume that weaker dense spacing of crossovers along chromosomes (ie. weaker interference) is how .
+This is unexpected because most literature and logical models are built / assume that weaker dense spacing of crossovers along chromosomes (ie.
+weaker interference) is how.
+The history, models and empirical examples are extensively reviewed in [@ottoPaysuer2019].
+Broad scale patterns for the correlation of gwRR and interference strength (gamma parameter) do fit the expected negative pattern [@ottoPaysuer2019].
 
-The history, models and empiracal examples are extensively reviewed in [@ottoPaysuer2019].
-It is notable that borad discriptions / broad scale patterns for the correlation of gwRR and interference strength (gamma parameter) do fit the expected negative pattern [@ottoPaysuer2019].
-
-(genome wide rate vs single bivalent measures)
-
-**Caveats**
-We considered carefully if this pattern was due to error and we are confident that this pattern is genuine.
-
-There's no model that fits evolution of heterochiasmy and interference independantly.
+There's no model that fits evolution of heterochiasmy and interference together.
 Under our undertsanding of the models, the relatioship of genome wide recombiantion rate and interference strength is assumed to be negative.
@@
 
-First, the **typical modifer models do not model single meiotic events**. (think about how pooling man meiotic events can skew patterns)
-(crossover interference can be quantified from linkage maps (which are the average of many hundreds to thousands of meiotic events).
-It can also be discribed / measrued /quantified from single bivalets / chromosomes (either cytology of dense genotyping of single cells (cite).
-While it's widely agreed that they both reflect the same biological process, these two approaches are fundamentally different. (they are subject to distinct biases and noise).
-
-Second, most modifer models use abstracted genomic space -- like an infite 2D line between markers.
-(space between markers is abstracted).
-This is useful for fitting the gamma model, but within the cell, the physical chromosomes's axis length is fitnite and crossovers cannot form beyond those bounds.
-
+First, the **typical modifer models do not model single meiotic events**.
 We conclude that under the haploid selection and meiotic driver model any prediction for evolution of interference strength follows from the 'logical' expectation of a negative correlation.
 
-originally proposed/outlined by Trivers [] and refined /formalized in [@lenormand2003].
-This model describes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombination rates between males and females.
-
- This is a modifier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifier).
-
-This model predicts the evolution of heterchiasmy -- through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
+This model predicts the evolution of heterchiasmy -through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
 
 **SACE**
-While the SACE model doesn't explecitly make predictions aof the out come for interference strength (or averge spacing of crossovers, )
+While the SACE model doesn't explecitly make predictions aof the out come for interference strength (or averge spacing of crossovers)
 We note that the main predicted pattern under this model of maintaining larger blocks of chromosomes in linkage can asrise from stronger interference / greater spacing of the distance between two foci (also maximize rbar).
-
- extends the lenomand focusing on two conditions.
-first that the stength of epistatsis in dependant on the cis or trans nature.
-the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together)
-
- second that the relative interference strengths differnt between male and female.  This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
 
@@
 It is conceivable that the pattern of stronger interference could be generated from the waves/movements of chrmosomes in the early prophase stage this would come from a change in the movement or extension of the axis length.
-
-**spindle based selection model**
-There is a model that predicts this positive correlation between genome wide recombination rate and interference. The goldstein results (reviewed in).
 
@@
 
-note about the differences in the patterns>  assumptions with gamma dstribution and linkage maps, (closer to infinte genetic space), less weight on crossover assurance - more weigth on 
-(pooled SC/chromosome length gives strong correlation )
+# REVIEW
 
-which is much different than when a single chromosome is examined (by itself), finite 2D length, crossover assurance, weaker correlation with ((the closer to chromosome level for SC legnth metric / summary, the weaker the correlation with crossover number))
+We propose that the spindle based selection model is the most parsimounious with our results and known recombination patterns from the literautre.
 
-2.  Review the assumptions for measuring CO interference in different ways.
-The logical model: Crossover interference is a mechanism of suppressing crossovers, negative relationship between interference strength and gwRR.
-More crossovers can fit along the chromosomes and results in more crossovers overall.
-Examples of empirical results supporting this pattern;  Otto Payseur, (Ruiz-Herrera? Bomblies?)
-Quantifing using the gamma distribution: 
-Caveats,  This doesn't make since for the general recombination landscape:   i) finite chromosome length and ii) limited range of crossovers per chromosomes (1 to 4).
-Alternative interference characterization:  i) the number of crossovers are held constant and ii) transform / generalize the rec landscape, (focus on how the rec landscape would change instead of the genome wide rate) iii) assume crossover assurance.
-	-The Goldstein model, models predicts a positive correlation with interference strength and higher recombination rate.
-A modifier which changes the probability of the recombination landscape having 0 crossovers to 2 crossovers, (driven by crossover assurance).
-
--Examples of empirical results which show this result / pattern
-	-(cattle) comparison WITHIN sexes
-	- Preliminary results for F2 cross (Hannah's work), (PWD vs CAST)
-	(maybe more once I start looking closer)
-
-This new model has **strengths**:
   - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)
@@
 our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle
-
-# REVIEW
 
@@
 <We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>
+<extends the lenomand focusing on two conditions.
+first that the stength of epistatsis in dependant on the cis or trans nature.
+the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together).
+second that the relative interference strengths differnt between male and female.
+This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
+originally proposed/outlined by Trivers [] and refined /formalized in [@lenormand2003].
+This model describes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombination rates between males and females.
+This is a modifier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifier)>
+<2.  Review the assumptions for measuring CO interference in different ways.
+The logical model: Crossover interference is a mechanism of suppressing crossovers, negative relationship between interference strength and gwRR.
+More crossovers can fit along the chromosomes and results in more crossovers overall.
+Examples of empirical results supporting this pattern;  Otto Payseur, (Ruiz-Herrera? Bomblies?)
+Quantifing using the gamma distribution: 
+Caveats,  This doesn't make since for the general recombination landscape:   i) finite chromosome length and ii) limited range of crossovers per chromosomes (1 to 4).
+Alternative interference characterization:  i) the number of crossovers are held constant and ii) transform / generalize the rec landscape, (focus on how the rec landscape would change instead of the genome wide rate) iii) assume crossover assurance.
+	-The Goldstein model, models predicts a positive correlation with interference strength and higher recombination rate.
+A modifier which changes the probability of the recombination landscape having 0 crossovers to 2 crossovers, (driven by crossover assurance)>
 

### 2020-03-31 doc/Discussion.rmd
incorporating thoughts from new discussion outline

@@
 title: "Discussion Draft"
-date: ""
+date: ''
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document: default
   html_document:
@@
 
+#reversed hetC direction
+
+Reversed heterochaiasmy (briefly noted in the first part)
+The common direction of heterochiasmy is female biased, however male biased heterochiasmy has (significant number of observations.
+-- has been seen before (cattle, marsupial)-- Our results are novel, because the pattern is indicative of rapid evolution (male specific) -- polymorphism of low and high group.
+
+models
+the prediction from two of the modifier models, haploid selection and SACE are that males will generally evolve lower recombination rates overall compared to females.
+(cases with exceptions.?
+we don't have evidence or reason to predict these exception cases are held)
+The two locus modifier model  does make predictions / parameterize cases for (reversal of the common pattern of female biased heterochiasmy).
+Which are dependant on the stage of action for the driver and the linkage status (see table 1 Brandvain and Coop).
+There are no prediction for reversal of heterochiasmy direction  under the COM (pairing model, but it should be noted that this model wasn't designed to incorporate this type of data.
+
+Under the spindle based model faster male specific evolution of the genome wide recombination rate is primarily driven by differences in the strength of the SAC.
+ leading to different evolutionary trajectories for the genome wide recombination rate.
+
+The SAC 'read's that all tetrads have tension -- (each set of centromeres have sinlge and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
+
+Efficient normal SAC -- works by pausing entry into anaphase or triggers apoptosis -- if a unpaired / tension-lacking tetrad is detected
+(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher treshold for being triggered  (4-5?)
+(Land Kauppi,
+
+(relaxed selection - / inefficient SAC -- in females -- fits with higher rate of achiasmata bivalents (there might be alternative mechanisms facilitating achiasmate segregation)
+
+  Mechanisms for sex differences in the strength of the SAC are the cell volumne (diffusion of signal molecules) and the centrosome being present at the spindle have been proposed in the literature.
+(testing / proving the mechanism for the new optima is for the high recombining group in males is out of the scope of this paper, but we propose some hypothesis are suggested below.
+ but the evolutionary pattern fits that of directional.
+
 # 'Typical landscape'
@@
 
-The 'typical landscapes' refers to the sex differences in the board scale pattern or general placement of crossovers along the length of chromosomes.
-The primary pattern which has been observed across mny species is that crossovers are shifted to telomereic positions in males while a more unifrom landscape in observed in females [@sardell_sex_2020].
+Our results on the broad scale pattern of crossover placement are generally supported by the literature.
+As reviewed in (Sardel Kirkpatrick), sexual dirmorohism in the broad scale landscape is a highly conserved trait.
+However there are exceptions, (marsupial, bird?)
+the popgen models
+The two locus modifier model and SACE model make predictions for (sexual dimorphism ) in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
+functional models
+The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons.
+The chromosome pairing based model (COM), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the RPMs.
+The spindle based model proposed that sexual dimorphism in the recombination landscape is driven by difference strength in the spindle assembly checkpoint (Lane Kauppi).
+The relaxed selection imposed by an ineffective SAC results in a uniform distribution of positions (while maintaining positive interference).
+Whereas,  in males, the directional selection could cause decreased variance in the recombination landscape via telomeric positioning of crossovers.
+Under the telomeric bias  recombination landscape, entry into anaphase may be more synchoriznes (and faster) given that the reductional separation of homologs is dependent on a minimal amount of sister cohesion connecting the homologs (Figure X).
 
-Our results show that the typical landscape is consered across all strains of house mice in this study (for all single chromosome daata and when controlled for crossover number), the broad scale position of crossovers was shifted closer to the telomere end (of the long arm).
-
-<Detangling centromere suppression from telomere bias,  The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes.
-we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalent
-s. )>
-
-The conclusions from the models -- basically all of the models could account for differences in the general recombination landscape patterns.
+[@sardell_sex_2020].
 
@@
 **the within animal variation is equvillent to between cell variation**
+[@lynn2002]
 
-Our results confirm previous findings from mice [@lynn2002] and humans of females having greater between cell variation compared to males.
+Our results of greater between cell variation .
+these results are supported by human and mouse (Gruhn (Kong)?
+(Zwel -- Cole).
+These data are also scant, between cell variance isn't often measured especially for both sexes.
+(given they require single cell observations for both sexes).
 
-Transcriptional noise and complex traits both show the same repsonses to different modes of selection, lower variation under stabilizing selection and greater variaion with relaxed selection.
-The difference in between
+While the modifier models parametrize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
+(selection on penetrance).
+In the spindle based selection model,  the source of between cell variance is attributed to difference strength of checkpoints.
+(Sexual dimorphism in the stringency of the spindle assembly checkpoint is a well established result in the literature.
+Most empirical results come from mammals (is this true?) yet, some of the hypotheses for the mechanism of inefficient SAC in oocytes are attributed to acentrosomal spindles (cite) and cell volume (cite), which are conserved sexual dirmohpic features of gametogenesis (cite).
 
-so all of the models could reasonably predict the above results ... examining some of the traits related to the faster evolution in the male specific data indecate different support across the models.
-
-None of the models make predictions for between cell variation -- so we draw conclusions from their preciations based on the strength of interference.
-
-# Chromatin Structure and interference strength
+# Chromatin Structure
 
@@
 
+The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals (Lynn, Gruhn, ). 
+models
+Very few models make predictions for how chromatin organization relates to variation in recombination rates / the recombination landscape.
+(the chromatin organization is how chromosomes with the same number of base pairs) are transformed into bottle brush structures with a central axis and most of the genetic material arranged as loops emanating outwards (figure).
+Only the COM model states that chromatin structure will be longer in oocytes due to greater cell volume.
+This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) (cahoon libuda).
+Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes.
+We argue that the chromatin structure (and axis length) should be considered as an integral parameter when studying recombination landscape.
+The sex differences in chromatin strucutre are closely connected to sex differences  in the interference strength and our results mirror those in Petkov et al.
+(prediction, shorter-loop will have stronger interference).
+
 According to the tethered loop model of DSB formation (cite) -- the 'real estate' for COs are the DNA loops - the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
@@
 / metaphase) -- the chromosomes condenses uniformly across the length.
-
-Another thing to point out / regarding this model -- is the connection to sex difference in interference strength.
-Our results of conserved within strain raw IFD lengths - yet when the IFDs are controlled for total SC length, we observed sexual dimorphism with males having longer IFD^normalized^ (stronger interference), mirror those observed in [@petkov2007].
-The model / explanation proposed in that paper is that The chromatin structure which has longer loops (and shorter axis) will have more DNA bps folded / sequestered into the DNA loops.
-Thus the strength of interference will be stronger because with each step (count) along the 2D axis, more bases pairs are being skipped above the axis.
-These findings point out two things, This part of heterochiasmy generates testable prediction: First, the interference strength measured in chromosome axis units (or SC) is generally conserved between sexes.
-Second the sex with longer axis and shorter loops will have weaker interference compared to the other sex.
-Or conversely, the sex with stronger linkage map based interference would have chromatin organization with shorter axis compared to the other sex.
-There is a plethora of data of sexual dimorphism in linkage map based interference strength indicating sexual dimorphism, but comparative cytological data is lacking (difference in perparness of some sexes and complications due to identification of specific chromosomes between cells.)
 
@@
 
-# Revered heterochiasmy
+# Evolution of Interference
 
-```{r reverse.HetC, echo=FALSE}
-# transition connection to the hetC models
-
-#our results - total gwRR isn't the most conserved, (we observed 3 other traits more)
-
-#what does a reversal of HetC imply?  gwRR
-#driven by faster evolution in males
-
-#our model had faster evolution in males resulting in a reversal of the canonical pattern of heterichiasmy ( F gwRR > M gwRR)
-#other examples, cattle, marsupials
-
-#what do models say about faster evolution in one sex? (yes -- but almost always for canonical pattern of heterochiasmy)
-
-```
-
-Our results demonstrate that gwRR is decoupled from other sexually dimorphic meotic traits, (we observed 3 other traits more conserved sexual dimorphism examined above).
-We the general pattern 2 strains attribute this pattern of reversed heterochiasmy to faster evolution of gwRR in two strains males.
-
-Which models predict different rates of evolution between the sexes.
-Some of the models could explain a reversal of the direction of sexual dimrohpsim, however most all of the predictions for faster evolution result in males having reduced recombination rates.
-Under the haploid selection model a  **reversal of the direction of heterochiasmy** would be accompanied by vast changes in haploid stages or degree of imprinting.
-Following the meiotic drive based model, females should be predicted to have larger linkage maps/gwRR since they are more exposed / weaker to true meiotic drivers.
-A reversal in heteroschiasmy is not clearly predicted by the SACE model (unless with the -- except under the conditions described below).
-Under the spindle selection model, the male (centrosome/effect SAC) metaphase could drive directional selection (and faster evolution) in males compared to female.
-However the mechanism/aspect of different fitness is not clear.
-Under the COM model, evolutionary predictions are not clear (explicit), but following the logic would come from changes in axis length (to generate another wave point) -- which might be due to meiocyte volume.
-
-# Evolution of Interference
+[@ottoPaysuer2019].
 
@@
 
-Our results indicated that interference strength has evolved with the genome wide recombination rates in the three strains with faster male evolution.
+[@ma2015_cattle]. In *Peromyscus leucopus*, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019].
 
-This is unexpected because most literature and logical models are built / assume that weaker dense spacing of crossovers along chromosomes (ie.
-weaker interference) is how.
-The history, models and empirical examples are extensively reviewed in [@ottoPaysuer2019].
-Broad scale patterns for the correlation of gwRR and interference strength (gamma parameter) do fit the expected negative pattern [@ottoPaysuer2019].
+Some basics of the constraints / bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength. 
+ The non-random spacing of crossovers along the 2D length of chromosomes is the first level of proof /meeting the rejection of the (null model of random placement (cite gamma COI papers).
+The physical limit of how far apart crossovers can be placed on a chromosome is a function of the length of the meiotic chromosome axis.
+Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength.
+This pattern has empirical support -- and describes  -- the fundamental relationship / of SC area of crossover number which is observed in much of the literature (which describes large/broad scale patterns).
 
-There's no model that fits evolution of heterochiasmy and interference together.
-Under our undertsanding of the models, the relatioship of genome wide recombiantion rate and interference strength is assumed to be negative.
-We now review how some of the assumptions of typical modifer models might not fit realities of the meiotic pathway.
+Our results support an opposite pattern, a positive correlation between interference strength and genome wide recombination rates; the strains which show rapid evolution of the gwRR also show stronger interference strength.
+We've found a number of examples of positive correlation between genome wide recombination rate and interference.
+The within sex comparison of two breeds of cattle (Ma et al), the Peromyscus species in (Peterson et al ), and preliminary data (HVR un published).
 
-First, the **typical modifer models do not model single meiotic events**.
-We conclude that under the haploid selection and meiotic driver model any prediction for evolution of interference strength follows from the 'logical' expectation of a negative correlation.
+Theoretical models haven't really considered the evolution of interference strength -- While the COM model predicts that interference (and recombination landscape) arises from known oscillatory movements during prophase it lacks an evolutionary based prediction.
+While the SACE modifier model does not explicitly model evolution of interference strength -- we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
+Veller et al rbar -- and predictions / for this - found a similar pattern (positive correlation in their simulations of male human data.
 
-This model predicts the evolution of heterchiasmy -through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)
-
-**SACE**
-While the SACE model doesn't explecitly make predictions aof the out come for interference strength (or averge spacing of crossovers)
-We note that the main predicted pattern under this model of maintaining larger blocks of chromosomes in linkage can asrise from stronger interference / greater spacing of the distance between two foci (also maximize rbar).
-
-Under the **COM** model -- the placement of crossovers (the recombination landscape) is generated by 'waves' / movement from the bouquet stage.
-It is conceivable that the pattern of stronger interference could be generated from the waves/movements of chrmosomes in the early prophase stage this would come from a change in the movement or extension of the axis length.
-
-The results from golstein reviewed in [@ottoPaysuer2019], predict a modeifier model (which placed more weight on crossover assurance (holds the number of crossovers constant) and simulates a scnario where a modifer acts on modulating the probabilty of flanking regions having 0 or 2 crossovers.
-Which has the result of increasing the distance between two foci.
-
-Additionally there are a handful of empircal results which also show the unexpected possitive correlation between gwRR and interference strength.
-When comparisons of genetic map length and interference strength are made across strengths within sex, the strain with the higher genome wide recombinaiton rate also has stronger interference (measured as nu gamma parameter) [@ma2015_cattle].
-In *Peromyscus leucopus*, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019].
-
-Additionaly -- preliminary data from a previous cross using MLH1 cytology in spermatocytes in house mouse replicated these results with IFD measures from  M.
-m.
-musculus^PWD^  being longer compared to M.
-m.
-castenus^Cast^ (HVR unpublished data) (cast has low gwRR).
+Models from Goldstein et al (review in Otto and Payseur) suggest that if this pattern is widespread --- interference evolves in X manner.
 
@@
 
-## The CONTEXT of MEIOSIS
-
-the study confirmed preivously identified sexual dimorhisc patterns in regard to the recombiantion landscape (in mammals) -- in the goal of unifying all of these results, we sough to investigate physiologycal aspects of gametogenesis which have conserved sexual dimorphic features.
-A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle).
-(Our hypothesis is that the tension in a centrosome spindle is stronger (or more uniform) since all MT-KT are anchored at single points across cells (the two centrosomes).
-This will also make prediction for the spindle assembly checkpoint (SAC), which is stronger/stricter in smaller gametes with centrosomes.
-(the SAC is more sensitive to achiasmate bivalents on the spindle / better at detecting a lack of tension).
-Whereas in the acentrosome spindle is more diffuse (across a larger area) and the tension across bivalents are anchored by multiple MTOCs, resulting in weaker tension force or less uniform strengths across all the bivalents within the cells.
-The SAC in eggs seems to be weaker/leakier, that is more achiasmate bivalents are required to trigger the SAC (stop the division).
-
-**main sex differences in spindle** are the anchoring of MT at opposite poles
-
-in males there are centromsomes for Mi and MII divisions
-where are in oocytes, a (universal) conserved feature of oogenesis in many species is that they loose centrosomes -- instead the MTs are anchored at mulitple aMTOCs.
-
-There is potential for these aMTOC to be unevenly distributd across the cortical and central side of the spindles (Wu et al, Schuh and ellenberg) --- (could be an origin of different tension / pulling )
-
-The survalence of SAC in oocytes might be decreased due to large volumne difusing the signals.
-
-#CO placement translates to bivalent shape
-
-**still an open question of if and how the CO number and placement of crossovers translates to different fitness values for segregation rates**
-
-the metric / amout of sister cohesion might be an good place to start
-
-the ultimate goal is proper disjunction of chromosomes, the main checkpoint is spindle assembly checkpoint (SAC) which responds to the tension signal across chiasmata
-
-we like this model because it can be applied to **both our results** for i) sexual dimorphism variation and ii) the fast male evolution in gwRR and recombination landscape.
-
-Uses / centers around one of the best known selection factors on meiotic division, anneuploidy.
-Several other facors related to meiotic recombiniation center on **segregation** (obligate chiasmata, lower bound of the genome wide rate) it's not unbelievable that the recombination
-
-Consequences of evolution in the number and placement of crossovers / evolution of interference strength:
-Spindle / SAC model: The number and placement of crossovers changes the resulting amount of sister cohesin involved the tension signal, which affects the tetrad/chiasmata structure at MI when the homologs are pulled apart (Hollis et al 2020, Lee et al 2019).
-This model predicts a difference in the meiotic spindles or division mechanism between high and low rec strains.
-ii. SACE: For the 2CO in the high rec males, larger blocks of chromosomes are kept together in the next generation.
-This model would predict greater reproductive variance for males.
-
 #Future Steps
@@
 
-#blah
-
-<5. Model prediction
-The models above which predict directional selection on the recombination landscape in males: SACE and Spindle based selection  are most parsimonious for the models review in this paper.
-The reduction principle wouldn't predict males diverge in genome wide rates and two locus modifier doesn't predict males gwRR would be higher than females.>
-
-<3.Two-locus (protect for drivers)
-This second modifer model was persented in [@brandvain2012scrambling] the basics of the model are that females will have a longer genetic map and more centromere biased recombination landscape compared to males.
-
-A strength of this model is that is presents a clear driver / source of differential fitness (ie meiotic drivers can decrease fitness/fertility).
-This model also accounts for the general female pattern,
-
-However, this model doesn't explain the borad scale pattern for the male pattern.>
-
-<Also:
-	-These results are different from the sex differences because the difference is seen in both the raw and normalized IFD, indicateing the mechanism  / physcial strength (interference strength (signal) mediated through the SC has evolved.>
-
-<Some of the traits related to faster male evolution -- distinguish in which models can redict them: the reverseal of heterochiasmy direction and the correlated evolution of gwRR and interference strength.
-
-**DSB evolved but, don't included as an examined trait, since most evolution based models don't make predictions for this stage of meiosis.**
-_the propotion of DSB repaired into crossovers hasn't evolved, the total number of DSBs has evolved_  (following the tethered lop model of DSBs formation -- this would suggest that in the high rec strains, assuming no large difference in the Mb size of the chromosomes,
-during chromatin remodeling in prophase, in the high rec strains the chromosomes have longer axis, allowing a higher number but shorter DNA loops (for SPO11 to make the DSBs).
-Our results mirror those found in Baier et al [@baier2014]
-
-Our results show that the ratios of DSB : CO 
-**range from X to X which is similar to those presented in [@baier2014] (which report X to X) **
-
-These conserved ratios suggest that the increase in DSB numbers are accompanied by proportional increase in crossovers.
-
-This shifts support away from evolution at the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up.
-**increase in DSBs is driven by the number of DNA loops** and due to the inverse relationship of loop and chromatin axis length, this would result in longer sc lengths (which reflect the underlying chromatin axis length).>
-
-<Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.
-
-same pattern in humans (Gruhn et al 2013)
-  (exceptions, birds (fish?) plants?) -- examining this is hard due to chromosome size effecs
-  - This also assumes that the DNA loop sizes have uniform size along and between chromosomes
-
-2. Short axis / longer DNA loops will have stronger interference
-This prediction combines the findings of chromatin compaction and interference differences which are intertwined by the transmission of interference along the chromosome axis and the inverse relationship chromatin compaction (axis length and DNA loop sizes).
-In physical units (micron) the counting model fits interference well to the 2D linear space of the axis these measures of interference are conserved across sexes in mouse.
-This is because the counting units are the BASES of DNA loops which is conserved along the 2D length of chromatin axis.
-When this space is transformed to account for all of the DNA base pairs which are sequestered into the DNA loops above the axis, the strength of interference will vary depending the on underlying chromatin structure.>
-
-<We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males.
-One of our primary objectives was to identify other traits which correlate (could explain the increase).>
-
-<Baier et al found significant differences in sc legnths for 
-The tests in our study - show moderate to low significance in the differences in SC lengths. (this is because we have more noise-can't distinguish individual chromosomes)
-
-a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
--different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures
-
-our results are similar in several ways
-- correlation of final MLH1 and DBS counts, SC measues correlate with the DSB number (even tho the differnces we observe in our study are -- slight / less significant -- we attribute that to chromosome specific noise )>
-
-<We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>
-<extends the lenomand focusing on two conditions.
-first that the stength of epistatsis in dependant on the cis or trans nature.
-the authors note that this would apply to gene archtechture (and increase with in cis ) -- where the regulatroy region and coding region of a gene evolve (together).
-second that the relative interference strengths differnt between male and female.
-This can apply to (genes acting in sexual antagonism) or can be mediated via imprinted genes.
-originally proposed/outlined by Trivers [] and refined /formalized in [@lenormand2003].
-This model describes a system where gametes are competing or selection is acting in the haploid stage to cause difference in genome wide recombination rates between males and females.
-This is a modifier model of three loci, where the recombination rate between two loci is determined by a third locus (or modifier)>
-<2.  Review the assumptions for measuring CO interference in different ways.
-The logical model: Crossover interference is a mechanism of suppressing crossovers, negative relationship between interference strength and gwRR.
-More crossovers can fit along the chromosomes and results in more crossovers overall.
-Examples of empirical results supporting this pattern;  Otto Payseur, (Ruiz-Herrera? Bomblies?)
-Quantifing using the gamma distribution: 
-Caveats,  This doesn't make since for the general recombination landscape:   i) finite chromosome length and ii) limited range of crossovers per chromosomes (1 to 4).
-Alternative interference characterization:  i) the number of crossovers are held constant and ii) transform / generalize the rec landscape, (focus on how the rec landscape would change instead of the genome wide rate) iii) assume crossover assurance.
-	-The Goldstein model, models predicts a positive correlation with interference strength and higher recombination rate.
-A modifier which changes the probability of the recombination landscape having 0 crossovers to 2 crossovers, (driven by crossover assurance)>
-
 #References

### 2020-03-31 doc/Discussion.rmd
added some sources

@@
 
-Advice from Bret:
-
-- Don't know where to fit in DSB evolution results
-
-- Keeping interference section breif (background, explaningillustrating spindle based selection model)
-
 To synthesize our results, we decded to systematically compare models which have been proposed for explaining the evolution of heterochiasmy or meiotic machinery.
@@
 
-Reversed heterochaiasmy (briefly noted in the first part)
-The common direction of heterochiasmy is female biased, however male biased heterochiasmy has (significant number of observations.
--- has been seen before (cattle, marsupial)-- Our results are novel, because the pattern is indicative of rapid evolution (male specific) -- polymorphism of low and high group.
+The common direction of heterochiasmy is female biased, however male biased heterochiasmy has significant number of observations, seen in cattle, marsupial, and (more to the list).
+Our results are novel in the sense that the evolutionary scale of the male-specific is short.
 
-models
-the prediction from two of the modifier models, haploid selection and SACE are that males will generally evolve lower recombination rates overall compared to females.
-(cases with exceptions.?
-we don't have evidence or reason to predict these exception cases are held)
-The two locus modifier model  does make predictions / parameterize cases for (reversal of the common pattern of female biased heterochiasmy).
-Which are dependant on the stage of action for the driver and the linkage status (see table 1 Brandvain and Coop).
-There are no prediction for reversal of heterochiasmy direction  under the COM (pairing model, but it should be noted that this model wasn't designed to incorporate this type of data.
+The prediction from two of the modifier models, haploid selection and SACE are that males will generally evolve lower recombination rates overall compared to females.
+<cases with exceptions.?
+we don't have evidence or reason to predict these exception cases are held>
+The two locus modifier model  does parameterize cases for male biased heterochiasmy, dependant on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).
+
+There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model wasn't designed to incorporate this type of data.
+This model were design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.
 
 Under the spindle based model faster male specific evolution of the genome wide recombination rate is primarily driven by differences in the strength of the SAC.
- leading to different evolutionary trajectories for the genome wide recombination rate.
 
-The SAC 'read's that all tetrads have tension -- (each set of centromeres have sinlge and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
+(Lane and Kauppi) [@laneKauppi2019]
+
+The SAC 'read's that kinetochores (the movement machinery built on top of the centromere) are connected to the MT of the spindle originating from the correct pole of the cell. (tension is )
+
+centromeres have sinlge and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
 
@@
 Our results on the broad scale pattern of crossover placement are generally supported by the literature.
-As reviewed in (Sardel Kirkpatrick), sexual dirmorohism in the broad scale landscape is a highly conserved trait.
+As reviewed in [@sardell_sex_2020], sexual dirmorohism in the broad scale landscape is a highly conserved trait.
 However there are exceptions, (marsupial, bird?)
@@
 
-The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals (Lynn, Gruhn, ). 
+The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002], [@gruhn2013]. 
 models
@@
 Only the COM model states that chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) (cahoon libuda).
+This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) (cahoon libuda) [@cahoonLibuda2019].
 Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes.
@@
 / metaphase) -- the chromosomes condenses uniformly across the length.
-
-# Evolved Traits
-
-```{r Evolve.traits.outline, include=FALSE}
-
-#tricker to fit the models to this, 
-#paradox -- how the broad/general patterns from the sex differences
-
-#models weren't designed to explain within species variation in recombination rates 
-#or evolution
-
-```
-
-The faster evolution in males with the high recombining rates.
-BUT applying these above models to the traits connected to the faster male evolution is a bit trickier since none of these models were designed to explain within species variation in recombination rates.
-(most were designed to explain the general pattern of sexual dimorphism in meiotic recombination rates.)
 
@@
 
-Some basics of the constraints / bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength. 
- The non-random spacing of crossovers along the 2D length of chromosomes is the first level of proof /meeting the rejection of the (null model of random placement (cite gamma COI papers).
-The physical limit of how far apart crossovers can be placed on a chromosome is a function of the length of the meiotic chromosome axis.
+Evolved traits, Evolution of interference (hardest section)
+(interference is a fundamental aspect of the recombination landscape)
+Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.
+The non-random spacing of crossovers along the 2D length of chromosomes is the first level of quantification of crossover interference (cite gamma COI papers).
 Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength.
-This pattern has empirical support -- and describes  -- the fundamental relationship / of SC area of crossover number which is observed in much of the literature (which describes large/broad scale patterns).
+This pattern has empirical support from the most species (Otto and Payseur) and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
+Our results support an opposite pattern, a positive correlation between interference strength and genome wide recombination rates; we find support that interference strength has evolved in the two groups of male strains.
+There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.
+The within sex comparison of two breeds of cattle with different genome wide recombintion rates (Ma et al), between lab and wild mice of Peromyscus leucopus, and in a previous house mouse cross (Dumont,  preliminary data (HVR unpublished).
+Theoretical models haven't really considered the evolution of interference strength, neither the haploid selection or two locus modifier model can not be applied to evolution of interference strength.
+While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
+The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions.
+We propose the spindle based selection model would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).
+Models from Goldstein et al (review in Otto and Payseur) suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve.
+Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant.
+The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates.
+Given that the empirical range of crossovers per chromosome is quite small (1-3  (Otto Payseur 2019, Stapley et al 207) and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.
 
-Our results support an opposite pattern, a positive correlation between interference strength and genome wide recombination rates; the strains which show rapid evolution of the gwRR also show stronger interference strength.
-We've found a number of examples of positive correlation between genome wide recombination rate and interference.
-The within sex comparison of two breeds of cattle (Ma et al), the Peromyscus species in (Peterson et al ), and preliminary data (HVR un published).
-
-Theoretical models haven't really considered the evolution of interference strength -- While the COM model predicts that interference (and recombination landscape) arises from known oscillatory movements during prophase it lacks an evolutionary based prediction.
-While the SACE modifier model does not explicitly model evolution of interference strength -- we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
-Veller et al rbar -- and predictions / for this - found a similar pattern (positive correlation in their simulations of male human data.
-
-Models from Goldstein et al (review in Otto and Payseur) suggest that if this pattern is widespread --- interference evolves in X manner.
+Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.
 

### 2020-03-31 doc/Discussion.rmd
small updates

@@
 
-To synthesize our results, we decded to systematically compare models which have been proposed for explaining the evolution of heterochiasmy or meiotic machinery.
-There are two main types of models: modifers (indirect) modifers or functional (cell physiology based models).
-We attempted to test the hypotheses (mental test for if the model) (Table X).
+We performed hypothesis testing of models from the literature and adapted here fit the main results from this study.
+We choose to examine three modifier models and two functional and cell physiology models (table X).
+It is a bit challenging to test all of these models together since they were built with difference parameters and built to describe different patterns of recombination variation.
+The modifier models were built to explain heterochiasmy variation between sexes and the COM functional model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
+In cases where the models can't be extended to predicting results, we list as 'NA'.
 
-**the models**  We examined 3 modifer models; haploid selection, two locus model, and SACE.
-We also examined and propose cell physiology based models: the Chromosome Ocsilation (COM) model and a model focused on the spindle differences between males and females.
+# Reversed heterochiasmy direction
 
-<we left out Beam film model for XX reasons>
+```{r reversed.HetC, echo=FALSE}
 
-<heterochiasmy refers to different numbers of chiasmata - and recently has been extended to number of crossovers -- this view / charaecterization is slightly incomplete, since it considers just one cell stage and 1 value (a count))
-We argue that the sex differences (broaden the definition of heterochiasmy or use a different word) -- which includes more aspects that show conserved sexes differences>
+```
 
-<Our results indicate, sexes have distinct evolutionary trajectories for genome wide recombinaiotn rate (and landscape) in males nd females).
-
-Conserved features of the landscape may play a role in influencing/keeping the sex specific trajectories.
-In our examination of the panel of mouse strains, we note several features of the recobmination landscape with conserved sexual dirmophism pattern.
-That is the sex within sex these features show no evolution and consistant pattern of variation.>
-
-We focus first on the results for sexually dimorphic traits.
-
-#reversed hetC direction
-
-The common direction of heterochiasmy is female biased, however male biased heterochiasmy has significant number of observations, seen in cattle, marsupial, and (more to the list).
+The common direction of heterochiasmy is female biased, however male biased heterochiasmy has significant number of observations (list).
 Our results are novel in the sense that the evolutionary scale of the male-specific is short.
@@
 we don't have evidence or reason to predict these exception cases are held>
-The two locus modifier model  does parameterize cases for male biased heterochiasmy, dependant on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).
+The two locus modifier model  does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).
 
-There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model wasn't designed to incorporate this type of data.
+There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model was not't designed to incorporate this type of data.
 This model were design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.
@@
 
-(Lane and Kauppi) [@laneKauppi2019]
+[@laneKauppi2019]
 
@@
 
-centromeres have sinlge and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
+centromeres have single and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
 
 Efficient normal SAC -- works by pausing entry into anaphase or triggers apoptosis -- if a unpaired / tension-lacking tetrad is detected
-(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher treshold for being triggered  (4-5?)
+(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher threshold for being triggered  (4-5?)
 (Land Kauppi,
@@
 
-# 'Typical landscape'
-
-```{r typical.landscape.outline, include=FALSE}
-
-#-long observed
-#measured from a variety of approaches,
-#larger variety of species
-
-#-Exception to landscape (marisupials, birds?)
-
-#which models make clear predctions for typical landscape?
-#SACE
-#drivers
-#spindle differences
-
-```
-
-Our results on the broad scale pattern of crossover placement are generally supported by the literature.
-As reviewed in [@sardell_sex_2020], sexual dirmorohism in the broad scale landscape is a highly conserved trait.
-However there are exceptions, (marsupial, bird?)
-the popgen models
-The two locus modifier model and SACE model make predictions for (sexual dimorphism ) in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
-functional models
-The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons.
-The chromosome pairing based model (COM), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the RPMs.
-The spindle based model proposed that sexual dimorphism in the recombination landscape is driven by difference strength in the spindle assembly checkpoint (Lane Kauppi).
-The relaxed selection imposed by an ineffective SAC results in a uniform distribution of positions (while maintaining positive interference).
-Whereas,  in males, the directional selection could cause decreased variance in the recombination landscape via telomeric positioning of crossovers.
-Under the telomeric bias  recombination landscape, entry into anaphase may be more synchoriznes (and faster) given that the reductional separation of homologs is dependent on a minimal amount of sister cohesion connecting the homologs (Figure X).
-
-[@sardell_sex_2020].
-
 # Greater between cell variation
@@
 ```{r within.mouse.var, include=FALSE}
-
-#our results, female more (greater than tech error) male no significant difference
-#between cell variation
-
-#between cell variance (Veller Kleckner), mechnism
-#evolution / popgen model from (veller)
-
-#models that predict more within animal variance
-#none of the models expliciltly predict fr traits at the between cell level
-#so any between cell variation -- patterns is due to DIFFERENCES IN THE STRENGTH OF SELECTION
-
-## connection on evolution of transcriptional noise
-# transcriptional noise and complex traits -- both show the same repsonses to different modes of selection
 
@@
 
+# Typical landscape
+
+```{r typical.landscape.outline, include=FALSE}
+
+# typical landscape is conserved -- sexually dimorphic meiotic trait (Sardell)
+
+# models
+
+#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
+
+```
+
+Our results on the broad scale pattern of crossover placement are generally supported by the literature.
+As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale landscape is a highly conserved trait.
+
+The two locus modifier model and SACE model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
+
+The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons.
+The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) (hulten, Rubin et al 2020).
+
+The spindle based model proposed that sexual dimorphism in the recombination landscape is driven by difference strength in the spindle assembly checkpoint (Lane Kauppi).
+The relaxed selection imposed by an ineffective SAC results in a uniform distribution of positions (while maintaining positive interference).
+
+Just as the difference in sac (checkpoint) strength can lead to disinct evolutionary trajectories of the genome wide recombiation rate for oocytes and spermatocytes, the other side of the same coin is that there is increased variance between cells for crossover number.
+(weaker / leaky checkpoint )
+
+The telomeric placement of crossovers in the male recombiation landscape, minimizes the amount of sister cohesion connecting homologs.
+This conserved feature of the male recombination landscape, might act to synchornize the seperation homologs after alignment at metaphase, (and in the degargation of the sister cohesion connecting the homologs.
+Another conserved feature of gametogenesis is that the timing of spermatocyte metapphase is much shorter of time compared to oocytes.
+
+Whereas, in males, the directional selection could cause decreased variance in the recombination landscape via telomeric positioning of crossovers.
+Under the telomeric bias  recombination landscape, entry into anaphase may be more synchronizes (and faster) given that the reductional separation of homologs is dependent on a minimal amount of sister cohesion connecting the homologs (Figure X).
+
 # Chromatin Structure
@@
 ```{r chromatin.organization, include=FALSE}
-#1. basics and expectations for chromatin organization
-#seen in humans, mice (signal comes through even with cell to cell variation in chromatin compaction / organization)
 
-#2. how the model treat this limiting factor
+#supporting results from lit
+#chromatin organization
 
-#3. consequence for interference (general patterns) is that interference measured as (mb) will be stronger if DNA loops are stronger
-#prediction for general -- pattern
+#models
+# none of the models make predictions for the axis length (chromatin organization)
+# only COM predicts that sexual dimorphism in axis length is due to cell vol
+
 ```
@@
 
-**chromatin organization**
+The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002], [@gruhn2013].
 
-The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002], [@gruhn2013]. 
-models
 Very few models make predictions for how chromatin organization relates to variation in recombination rates / the recombination landscape.
@@
 Only the COM model states that chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) (cahoon libuda) [@cahoonLibuda2019].
+This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].
 Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes.
@@
 According to the tethered loop model of DSB formation (cite) -- the 'real estate' for COs are the DNA loops - the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
-**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**
-
-2. Only one of the models we examined aligns with our observed results for sexual dimorphism for axis length.
-
-The modifier models do not parameterize this since they do not models / are not built to consider single meiotic events.
-
-**COM** predicts that axis length will be larger in oocytes due to the larger volume -- this would help in the bouquet stage -- may be connected to the 'the search' during homolog pairing.
-
-the **spindle based selection** model does not specific predictions about this aspect -- assumes that the 2D areas for chromosomes are proportional across stages (during the decondensing stage between pachtyene and dipl?
-/ metaphase) -- the chromosomes condenses uniformly across the length.
+**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.
+Only one of the models COM aligns with our observed results for sexual dimorphism for axis length.
+**COM** predicts that axis length will be larger in oocytes due to the larger volume.
 

### 2020-04-01 doc/Discussion.rmd
added spindle points

@@
 ```{r reversed.HetC, echo=FALSE}
+1c. sex differences in the SAC strength connected to conserved features of gametogenesis
+- Sperm SAC strong (can detect 1 unstable bivalent) and oocyte SAC weak / 'leaky' achiasmate chromosomes pass through and higher rate of annueploidy
+- leaky oocyte SAC linked to acentrosome spindle (mulitple aMTOCs) and cell volumne, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)
 
+1d. directional selection could be seen under strict SAC if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
 ```
@@
 ```{r within.mouse.var, include=FALSE}
+2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
+2a.
+relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
+The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.
+
+?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
 
@@
 #<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
+3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
+3a.
+There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.
+In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle.
+(required to start with more since the cohesion isn't replenished.
+The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.
+
+3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
+-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs.
+This area reduced when crossover position shifted to telomere (each end of chromosomes).
+Two crossover acrocentric chromosome the area is distal from the first until the second crossover.
+
+3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
+- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.
+
+*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.
 
@@
 #**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
+
+5. Interference evolution shows a positive correlation with gwRR
+5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
+5b. Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
 

### 2020-04-01 doc/Discussion.rmd
first shot at writing the spindle logic for reversed hetc

@@
 ```{r reversed.HetC, echo=FALSE}
-1c. sex differences in the SAC strength connected to conserved features of gametogenesis
-- Sperm SAC strong (can detect 1 unstable bivalent) and oocyte SAC weak / 'leaky' achiasmate chromosomes pass through and higher rate of annueploidy
-- leaky oocyte SAC linked to acentrosome spindle (mulitple aMTOCs) and cell volumne, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)
 
-1d. directional selection could be seen under strict SAC if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
+#spindle selection modle points
+
 ```
+
+[@laneKauppi2019]
 
@@
 
-The prediction from two of the modifier models, haploid selection and SACE are that males will generally evolve lower recombination rates overall compared to females.
-<cases with exceptions.?
-we don't have evidence or reason to predict these exception cases are held>
-The two locus modifier model  does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).
+The prediction from two of the modifier models, haploid selection and SACE are that males will generally evolve lower recombination rates.
+<Male biased heterochiasmy predicted XX> The two locus modifier model  does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).
 
@@
 
-Under the spindle based model faster male specific evolution of the genome wide recombination rate is primarily driven by differences in the strength of the SAC.
+Under the spindle based model faster male specific evolution of the genome wide recombination rate is primarily driven by differences in the strength of the spindle assemby checkpoint (SAC).
+(we won't go deeply into the molecular details (but the genes and pathways are will known, we refer the reader to [@laneKauppi2019]).
 
-[@laneKauppi2019]
+This is our logic for how a effective SAC will caus puriying selection on the bivalent / chromosome structure at meiotisis I division / reuctional division (MI).
 
-The SAC 'read's that kinetochores (the movement machinery built on top of the centromere) are connected to the MT of the spindle originating from the correct pole of the cell. (tension is )
+The transition from metaphse to analphase is irreversable, so everything (all bivalents in the cell) has to be set up correctly before anaphase to prevent lower fitness later in the gametogenesis pathway.
+If unstable bivalents are detected on the MI spindle, the SAC delays the transition and buys time for the lose bivalent to correctly attach to the MT spindle.
+If this is delayed long enough, apoptosis will be triggered.
+If there are genetic variants which increase the number of unstable bivalents, they selected against in favor to genetic variants causing faster stabilization of bivalents on the MI spindle.
+Thus the SAC can lead to stabilizing selection on the chromosome structure.
 
-centromeres have single and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
+It is known that the strength of the SAC varies between the forms of gametogenesis.
+(via the rate of anueploidy or achiasmate bivalents in mature gametes). 
+Sperm have lower anueploidy rates compared to oocytes (other lines of evidence) indcating that
 
-Efficient normal SAC -- works by pausing entry into anaphase or triggers apoptosis -- if a unpaired / tension-lacking tetrad is detected
-(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher threshold for being triggered  (4-5?)
-(Land Kauppi,
+the SAC in sperm is more strict compared to oocytes (which have been discribed as 'leaky').
+In sperm single unpaired bivalents can trigger the SAC.
+While in females more oocytes progress through meiosis even with unpaired bivalents.
 
-(relaxed selection - / inefficient SAC -- in females -- fits with higher rate of achiasmata bivalents (there might be alternative mechanisms facilitating achiasmate segregation)
+These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volumne (cite) -- which result in differing morphology of the metaphase MT spindle.
 
-  Mechanisms for sex differences in the strength of the SAC are the cell volumne (diffusion of signal molecules) and the centrosome being present at the spindle have been proposed in the literature.
-(testing / proving the mechanism for the new optima is for the high recombining group in males is out of the scope of this paper, but we propose some hypothesis are suggested below.
- but the evolutionary pattern fits that of directional.
+<leaky oocyte SAC linked to acentrosome spindle (mulitple aMTOCs) and cell volumne, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>
+
+Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
+(in oogenesis the leaky SAC has weak selection coefficient)
 
@@
 
+#blah
+
+The SAC 'read's that kinetochores (the movement machinery built on top of the centromere) are connected to the MT of the spindle originating from the correct pole of the cell. (tension is )
+
+centromeres have single and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
+
+Efficient normal SAC -- works by pausing entry into anaphase or triggers apoptosis -- if a unpaired / tension-lacking tetrad is detected
+(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher threshold for being triggered  (4-5?)
+(Land Kauppi,
+
+(relaxed selection - / inefficient SAC -- in females -- fits with higher rate of achiasmata bivalents (there might be alternative mechanisms facilitating achiasmate segregation)
+
+  Mechanisms for sex differences in the strength of the SAC are the cell volumne (diffusion of signal molecules) and the centrosome being present at the spindle have been proposed in the literature.
+(testing / proving the mechanism for the new optima is for the high recombining group in males is out of the scope of this paper, but we propose some hypothesis are suggested below.
+ but the evolutionary pattern fits that of directional.
+
 #References

### 2020-04-01 doc/Discussion.rmd
knitted

@@
 ```{r within.mouse.var, include=FALSE}
-2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
-2a.
+#2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
+#2a.
 relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
@@
 
-?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
+#?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
 
@@
 #<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
-3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
-3a.
+#3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
+#3a.
 There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.
@@
 
-3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
--For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs.
+#3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
+#-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs.
 This area reduced when crossover position shifted to telomere (each end of chromosomes).
@@
 
-3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
-- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.
+#3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
+#- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.
 
-*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.
+#*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.
 
@@
 
-5. Interference evolution shows a positive correlation with gwRR
-5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
-5b. Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
+#5. Interference evolution shows a positive correlation with gwRR
+#5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
+#5b.
+Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
 

### 2020-04-01 doc/Discussion.rmd
more thoughts

@@
 (we won't go deeply into the molecular details (but the genes and pathways are will known, we refer the reader to [@laneKauppi2019]).
-
-This is our logic for how a effective SAC will caus puriying selection on the bivalent / chromosome structure at meiotisis I division / reuctional division (MI).
-
-The transition from metaphse to analphase is irreversable, so everything (all bivalents in the cell) has to be set up correctly before anaphase to prevent lower fitness later in the gametogenesis pathway.
+The transition from metaphse to analphase is irreversable, so all bivalents in the cell has to be set up correctly before anaphase to prevent lower fitness later in the gametogenesis pathway.
 If unstable bivalents are detected on the MI spindle, the SAC delays the transition and buys time for the lose bivalent to correctly attach to the MT spindle.
@@
 (selection on penetrance).
+
 In the spindle based selection model, the source of between cell variance is attributed to difference strength of checkpoints.
-(Sexual dimorphism in the stringency of the spindle assembly checkpoint is a well established result in the literature.
-Most empirical results come from mammals (is this true?) yet, some of the hypotheses for the mechanism of inefficient SAC in oocytes are attributed to acentrosomal spindles (cite) and cell volume (cite), which are conserved sexual dirmohpic features of gametogenesis (cite).
+As mentioned above, there is sexual dimorphism in the stringency of the spindle assembly checkpoint (SAC).  And these are linked to conserved features of gametogenesis.
+
+As mentioned above (fundamental/conserved) sex differences in the MI spindle presence of centrioles in males anchors spindle at two cell poles.
+While oocytes lack centrioles and so the MT spindle is more difuse and made of mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
+This or larger cell volumne might -- lead to anaphase promoting signal being more difuse (or tension signal?)
+
+2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
+
+2a.
+relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
+The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.
+
+2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
 

### 2020-04-02 doc/Discussion.rmd
small clean up

@@
 
+#our results of reversed hetC are novel due to the short evolutionary disances across the panel of mouse strains
+
 #spindle selection modle points
@@
 
-[@laneKauppi2019]
+The common direction of heterochiasmy is female biased, however male biased heterochiasmy examples are not especially rare.
+Our results are novel from previous examples of male biased heterochiasmy species because the panel of strains have much shorter evolutionary distances.
+This context gives us insights to sex-specific evolution of recombination rates.
 
-The common direction of heterochiasmy is female biased, however male biased heterochiasmy has significant number of observations (list).
-Our results are novel in the sense that the evolutionary scale of the male-specific is short.
+The prediction from two of the modifier models, haploid selection and SACE, state that males will generally evolve lower recombination rates.
+Male biased heterochiasmy is predicted under the two locus modifier model when XXXX.
+<does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).>
 
-The prediction from two of the modifier models, haploid selection and SACE are that males will generally evolve lower recombination rates.
-<Male biased heterochiasmy predicted XX> The two locus modifier model  does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).
+There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model was not designed to incorporate this type of data.
+This model was design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.
 
-There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model was not't designed to incorporate this type of data.
-This model were design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.
+Under the spindle based model faster male evolution in male genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) <acting on the chromosome structure>.
 
-Under the spindle based model faster male specific evolution of the genome wide recombination rate is primarily driven by differences in the strength of the spindle assemby checkpoint (SAC).
-(we won't go deeply into the molecular details (but the genes and pathways are will known, we refer the reader to [@laneKauppi2019]).
-The transition from metaphse to analphase is irreversable, so all bivalents in the cell has to be set up correctly before anaphase to prevent lower fitness later in the gametogenesis pathway.
-If unstable bivalents are detected on the MI spindle, the SAC delays the transition and buys time for the lose bivalent to correctly attach to the MT spindle.
-If this is delayed long enough, apoptosis will be triggered.
-If there are genetic variants which increase the number of unstable bivalents, they selected against in favor to genetic variants causing faster stabilization of bivalents on the MI spindle.
+The genes and pathways are will known, we refer the reader to [@laneKauppi2019].
+
+The transition from metaphase to anaphase is irreversible, so all bivalents in the cell have to be aligned correctly to prevent anueploidy.
+
+Unstable bivalents are detected on the MI spindle, the SAC delays the cell cycle which potentially allows time for lose bivalents to correctly attach to the MT spindle.
+If the 'unattached' persists apoptosis will be triggered.
+Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
 Thus the SAC can lead to stabilizing selection on the chromosome structure.
 
-It is known that the strength of the SAC varies between the forms of gametogenesis.
-(via the rate of anueploidy or achiasmate bivalents in mature gametes). 
-Sperm have lower anueploidy rates compared to oocytes (other lines of evidence) indcating that
+Sperm have lower anueploidy rates compared to oocytes (other lines of evidence) indicating that the 'strength' of the SAC varies between oogenesis and spermatogenesis.
+In sperm single unpaired bivalents can trigger the SAC.
+While in females more oocytes progress through meiosis even with unpaired bivalents (refereed to as leaky SAC).
 
-the SAC in sperm is more strict compared to oocytes (which have been discribed as 'leaky').
-In sperm single unpaired bivalents can trigger the SAC.
-While in females more oocytes progress through meiosis even with unpaired bivalents.
+These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite) which result in differing morphology of the metaphase microtubule spindle.
 
-These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volumne (cite) -- which result in differing morphology of the metaphase MT spindle.
-
-<leaky oocyte SAC linked to acentrosome spindle (mulitple aMTOCs) and cell volumne, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>
+<leaky oocyte SAC linked to acentrosome spindle (multiple aMTOCs) and cell volume, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>
 
@@
 ```{r within.mouse.var, include=FALSE}
+#review results, other data
+
 #2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
+
 #2a.
@@
 
-**the within animal variation is equvillent to between cell variation**
-[@lynn2002]
+Our results of greater between cell variation in females, these results are supported by human and mouse (Gruhn, Lynn Kong). [@lynn2002].
 
-Our results of greater between cell variation .
-these results are supported by human and mouse (Gruhn (Kong)?
-(Zwel -- Cole).
-These data are also scant, between cell variance isn't often measured especially for both sexes.
-(given they require single cell observations for both sexes).
+These data are also scant, between cell variance isn't often measured especially for both sexes. (given they require single cell observations for both sexes).
 
-While the modifier models parametrize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
+While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
 (selection on penetrance).
@@
 
-As mentioned above (fundamental/conserved) sex differences in the MI spindle presence of centrioles in males anchors spindle at two cell poles.
-While oocytes lack centrioles and so the MT spindle is more difuse and made of mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
-This or larger cell volumne might -- lead to anaphase promoting signal being more difuse (or tension signal?)
-
-2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
-
-2a.
-relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
-The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.
-
-2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
+The opposite dynamics of the purifying selection for the 'strict'  SAC in spermatocytes is relaxed selection from a 'permissive' SAC.
+As mentioned above a fundamental sex difference in the metaphase spindle is the presence of centrioles in spermatocytes which anchors and nucleates the microtubulue spindle at the two cell poles.
+While oocytes lack centrioles which results in a difuse network of microtubules with mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
+This or larger cell volumne may lead/be a potnetial source of the leaky SAC in oocytes.
+Leaky SAC, relaxed selection on the anaphase to metaphase transition could increase the between cell variance for crossover number
 
@@
 
-Our results on the broad scale pattern of crossover placement are generally supported by the literature.
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale landscape is a highly conserved trait.
+Our results on the broad scale pattern of crossover placement are supported by the literature.
+As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
 
-The two locus modifier model and SACE model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
+The **two locus modifier** and **SACE** model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
 
@@
 
-Just as the difference in sac (checkpoint) strength can lead to disinct evolutionary trajectories of the genome wide recombiation rate for oocytes and spermatocytes, the other side of the same coin is that there is increased variance between cells for crossover number.
+Just as the difference in SAC (checkpoint) strength can lead to disinct evolutionary trajectories of the genome wide recombination rate for oocytes and spermatocytes, the other side of the same coin is that there is increased variance between cells for crossover number.
 (weaker / leaky checkpoint )
@@
 
-0.
 Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
 Two strains have have more crossovers in males, but females still have longer sc axis.
-_This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy_
+**This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**
 
-The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002], [@gruhn2013].
+The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013].
 
@@
 (the chromatin organization is how chromosomes with the same number of base pairs) are transformed into bottle brush structures with a central axis and most of the genetic material arranged as loops emanating outwards (figure).
-Only the COM model states that chromatin structure will be longer in oocytes due to greater cell volume.
+Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
 This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].
+
 Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes.
 We argue that the chromatin structure (and axis length) should be considered as an integral parameter when studying recombination landscape.
-The sex differences in chromatin strucutre are closely connected to sex differences  in the interference strength and our results mirror those in Petkov et al.
-(prediction, shorter-loop will have stronger interference).
 
-According to the tethered loop model of DSB formation (cite) -- the 'real estate' for COs are the DNA loops - the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
+We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
+
+According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
 **The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.
@@
 
-Evolved traits, Evolution of interference (hardest section)
-(interference is a fundamental aspect of the recombination landscape)
+Interference is a fundamental aspect of the recombination landscape.
 Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.
@@
 This pattern has empirical support from the most species (Otto and Payseur) and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
+
 Our results support an opposite pattern, a positive correlation between interference strength and genome wide recombination rates; we find support that interference strength has evolved in the two groups of male strains.
 There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.
-The within sex comparison of two breeds of cattle with different genome wide recombintion rates (Ma et al), between lab and wild mice of Peromyscus leucopus, and in a previous house mouse cross (Dumont,  preliminary data (HVR unpublished).
+The within sex comparison of two breeds of cattle with different genome wide recombintion rates (Ma et al) [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+
 Theoretical models haven't really considered the evolution of interference strength, neither the haploid selection or two locus modifier model can not be applied to evolution of interference strength.
@@
 We propose the spindle based selection model would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).
+
 Models from Goldstein et al (review in Otto and Payseur) suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve.

### 2020-04-03 doc/Discussion.rmd
some work on rec landscape

@@
 
-Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
-(in oogenesis the leaky SAC has weak selection coefficient)
+<Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
+(in oogenesis the leaky SAC has weak selection coefficient) ...
+bret says remove>
 
@@
 
-The spindle based model proposed that sexual dimorphism in the recombination landscape is driven by difference strength in the spindle assembly checkpoint (Lane Kauppi).
-The relaxed selection imposed by an ineffective SAC results in a uniform distribution of positions (while maintaining positive interference).
+**sister cohesin, chromosome structure altered by the number and placement of crossovers** synchornize timing of the decay of sister cohesion connecting homologs at metaphase I.
+conserved gametogensis differences, spermatogenesis faster than oogenesis -- The sex difference in requirements for sister cohesion could hinge on the timing differences between spermatogenesis and oogenesis.
 
-Just as the difference in SAC (checkpoint) strength can lead to disinct evolutionary trajectories of the genome wide recombination rate for oocytes and spermatocytes, the other side of the same coin is that there is increased variance between cells for crossover number.
-(weaker / leaky checkpoint )
+The connection of the **The spindle based model** centers on sex differences on the requirements for chromosome cohesion in late meiosis I.
+In oocytes the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.
 
-The telomeric placement of crossovers in the male recombiation landscape, minimizes the amount of sister cohesion connecting homologs.
-This conserved feature of the male recombination landscape, might act to synchornize the seperation homologs after alignment at metaphase, (and in the degargation of the sister cohesion connecting the homologs.
+The number and placement of crossovers alters the lay-out / distribution of sister cohesion and the 'chromosome stucture' once kinetochores attached to MT-spindle and tension applied.
+For single acrocentric chromsoomes the sister cohesion (area of the chromosome) distal to the crossover connects the homologs.
+(known as cruciform tetrad is some systems).
+(this is the part of the tension force read by the SAC).
+
+This area is reduced when crossover position is shifted towards the telomere (ends of the chromsome).
+The transition of metaphase to anaphase proceeds by the (irreversable) proteolytic decay of the sister cohesion on these tetrad arms.
+The male telomere bias locatlization of crossovers could be adaptive in order to synchornize (or minimize) the timing of homologs sepearting at metaphase I.
+
  Another conserved feature of gametogenesis is that the timing of spermatocyte metapphase is much shorter of time compared to oocytes.

### 2020-04-03 doc/Discussion.rmd
bad merge issue

--- 2020-04-02 doc/Discussion.rmd
@@
 
-[@ottoPaysuer2019].
-
 ```{r evo.interference, include=FALSE}
@@
 
-[@ma2015_cattle]. In *Peromyscus leucopus*, the wild mice had higher gwRR and stronger (gamma nu estimates) compared to lab raised mice of the same speicies [@peterson2019].
+**we find support that interference strength has evolved in the two groups of male strains.
+There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.**
 
-Interference is a fundamental aspect of the recombination landscape.
+Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromaint compaction
+
 Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.
-The non-random spacing of crossovers along the 2D length of chromosomes is the first level of quantification of crossover interference (cite gamma COI papers).
+
+Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombintion rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+
 Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength.
-This pattern has empirical support from the most species (Otto and Payseur) and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
-
-Our results support an opposite pattern, a positive correlation between interference strength and genome wide recombination rates; we find support that interference strength has evolved in the two groups of male strains.
-There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.
-The within sex comparison of two breeds of cattle with different genome wide recombintion rates (Ma et al) [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
 
@@
 
-Models from Goldstein et al (review in Otto and Payseur) suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve.
+Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve.
 Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant.
 The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates.
-Given that the empirical range of crossovers per chromosome is quite small (1-3  (Otto Payseur 2019, Stapley et al 207) and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.
+Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.
 

### 2020-04-03 doc/Discussion.rmd
clean up chromatin org section

@@
 
-Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
+Our results from MSM and PWD show that the sc length chromatin compaction is decoupled from the direction of heterochiasmy.
 Two strains have have more crossovers in males, but females still have longer sc axis.
-**This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**
+**This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part **.
+These departs from predictions based on empiracal data which nominate chromatin organization (axis length) as the primary driver of heterochiasmy (Petkov).
 
-The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013].
+while sexual dimorphism in chromatin structure is supported by data from mammals literature, especially for mammals [@lynn2002, @gruhn2013].
+(but there's generally a lack of data for other systems see cahoon libuda).
 
-Very few models make predictions for how chromatin organization relates to variation in recombination rates / the recombination landscape.
-(the chromatin organization is how chromosomes with the same number of base pairs) are transformed into bottle brush structures with a central axis and most of the genetic material arranged as loops emanating outwards (figure).
+There are few models evolutionary models for sex differences in chromatin organization.
 Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].
-
-Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes.
+This cell volumne prediction model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].
 We argue that the chromatin structure (and axis length) should be considered as an integral parameter when studying recombination landscape.
@@
 We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
-
 According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.

### 2020-04-03 doc/Discussion.rmd
hopefully tiny change
stil small changes

--- 2020-04-03 doc/Discussion.rmd
@@
 
-Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes.
-We argue that the chromatin structure (and axis length) should be considered as an integral parameter when studying recombination landscape.
-
-We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
-
-According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
-**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.
-Only one of the models COM aligns with our observed results for sexual dimorphism for axis length.
-**COM** predicts that axis length will be larger in oocytes due to the larger volume.
-
 # Evolution of Interference
@@
 This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
+**although it is not particulary strong in our results** We argue that this is due to difference in scale (genome wide vs single chromosome)
 

### 2020-04-03 doc/Discussion.rmd
reorganizing typical landscape section

@@
 
-The spindle based model proposed that sexual dimorphism in the recombination landscape is driven by difference strength in the spindle assembly checkpoint (Lane Kauppi).
-The relaxed selection imposed by an ineffective SAC results in a uniform distribution of positions (while maintaining positive interference).
+Under **spindle based model** we hypothesize that the telomeric placement of crossovers in the male recombination landscape, minimizes the amount of sister cohesion connecting homologs.
+the irreversable process of the metaphase to anaphase transistion is initiated by the protyltic decay of the sister cohesion connecting homologs (cite).
 
-Just as the difference in SAC (checkpoint) strength can lead to disinct evolutionary trajectories of the genome wide recombination rate for oocytes and spermatocytes, the other side of the same coin is that there is increased variance between cells for crossover number.
-(weaker / leaky checkpoint )
+<2> The number and placement of crossovers alters the lay-out / distribution of sister cohesion and the 'chromosome stucture' once kinetochores attached to MT-spindle and tension applied.
+For single acrocentric chromsoomes the sister cohesion (area of the chromosome) distal to the crossover connects the homologs.
+(known as cruciform tetrad is some systems).
+(this is the part of the tension force read by the SAC).
 
-The telomeric placement of crossovers in the male recombiation landscape, minimizes the amount of sister cohesion connecting homologs.
-This conserved feature of the male recombination landscape, might act to synchornize the seperation homologs after alignment at metaphase, (and in the degargation of the sister cohesion connecting the homologs.
+This area is reduced when crossover position is shifted towards the telomere (ends of the chromsome).
+The transition of metaphase to anaphase proceeds by the (irreversable) proteolytic decay of the sister cohesion on these tetrad arms.
+The male telomere bias locatlization of crossovers could be adaptive in order to synchornize (or minimize) the timing of homologs sepearting at metaphase I.
+
+<3>
+This conserved feature of the male recombination landscape, might act to synchronize the separation homologs after alignment at metaphase, (and in the degargation of the sister cohesion connecting the homologs.
 Another conserved feature of gametogenesis is that the timing of spermatocyte metapphase is much shorter of time compared to oocytes.
@@
 Whereas, in males, the directional selection could cause decreased variance in the recombination landscape via telomeric positioning of crossovers.
+
 Under the telomeric bias  recombination landscape, entry into anaphase may be more synchronizes (and faster) given that the reductional separation of homologs is dependent on a minimal amount of sister cohesion connecting the homologs (Figure X).
+
+The connection of the **The spindle based model** centers on sex differences on the requirements for chromosome cohesion in late meiosis I.
+In oocytes the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.
+
+ Another conserved feature of gametogenesis is that the timing of spermatocyte metapphase is much shorter of time compared to oocytes.
 

### 2020-04-03 doc/Discussion.rmd
cleaning up typical landscape section

--- 2020-04-03 doc/Discussion.rmd
@@
 
-The spindle based model proposed that sexual dimorphism in the recombination landscape is driven by difference strength in the spindle assembly checkpoint (Lane Kauppi).
-The relaxed selection imposed by an ineffective SAC results in a uniform distribution of positions (while maintaining positive interference).
+Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landsacpes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
+<1> the irreversable process of the metaphase to anaphase transistion is initiated by the protyltic decay of the sister cohesion connecting homologs --along the tetrad arms (cite).
 
-Just as the difference in SAC (checkpoint) strength can lead to disinct evolutionary trajectories of the genome wide recombination rate for oocytes and spermatocytes, the other side of the same coin is that there is increased variance between cells for crossover number.
-(weaker / leaky checkpoint )
+<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome stucture' once kinetochores attached to MT-spindle and tension applied.
+For single acrocentric chromosomes the sister cohesion distal to the crossover connects the homologs (when aligned on metaphase plate).
+This is known as cruciform tetrad is some systems.
 
-The telomeric placement of crossovers in the male recombiation landscape, minimizes the amount of sister cohesion connecting homologs.
-This conserved feature of the male recombination landscape, might act to synchornize the seperation homologs after alignment at metaphase, (and in the degargation of the sister cohesion connecting the homologs.
-Another conserved feature of gametogenesis is that the timing of spermatocyte metapphase is much shorter of time compared to oocytes.
+This area is reduced when crossover position is shifted towards the telomere (ends of the chromsome).
 
-Whereas, in males, the directional selection could cause decreased variance in the recombination landscape via telomeric positioning of crossovers.
-Under the telomeric bias  recombination landscape, entry into anaphase may be more synchronizes (and faster) given that the reductional separation of homologs is dependent on a minimal amount of sister cohesion connecting the homologs (Figure X).
+The male crossover telomere bias could be adaptive in order to synchornize (or minimize) the timing of homologs sepearting at metaphase I.
+
+<3> Conserved feature of gametogenesis -- is the length of timing.
+faster spermatogenesis -- may select for synchronize the separation homologs or faster seperate after alignment at metaphase.
+While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog strucutres)
+
+<In oocytes the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.>
 

### 2020-04-04 doc/Discussion.rmd
spell check

@@
 
-Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landsacpes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-<1> the irreversable process of the metaphase to anaphase transistion is initiated by the protyltic decay of the sister cohesion connecting homologs --along the tetrad arms (cite).
+Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
+<1> the irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (referred to as along the tetrad arms (cite)).
 
-<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome stucture' once kinetochores attached to MT-spindle and tension applied.
-For single acrocentric chromosomes the sister cohesion distal to the crossover connects the homologs (when aligned on metaphase plate).
-This is known as cruciform tetrad is some systems.
+<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome structure' which is a consequence of kinetochores attaching to the MT-spindle and tension applied.
+Single crossover bivalents form what is known as cruciform tetrad in some systems (figure x).
+This area is reduced when crossover positions are shifted towards the telomere ends of the chromosome.
+The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.
 
-This area is reduced when crossover position is shifted towards the telomere (ends of the chromsome).
-
-The male crossover telomere bias could be adaptive in order to synchornize (or minimize) the timing of homologs sepearting at metaphase I.
-
-<3> Conserved feature of gametogenesis -- is the length of timing.
-faster spermatogenesis -- may select for synchronize the separation homologs or faster seperate after alignment at metaphase.
-While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog strucutres)
+<3> Conserved feature of gametogenesis -- is the length of timing of the cell cycle (difference between spermat.
+faster spermatogenesis -- may select for synchronize the separation homologs or faster separate after alignment at metaphase.
+While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog structures)
 
@@
 
-Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy.
-Two strains have have more crossovers in males, but females still have longer sc axis.
+Our results from MSM and PWD show that the SC length compaction is independent (of the direction of heterochiasmy.
+Two strains have have more crossovers in males, but females still have longer SC axis.
 **This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**
@@
 
-Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromaint compaction
+Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromatin compaction
 
@@
 
-Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombintion rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
 

### 2020-04-04 doc/Discussion.rmd
cleaning up chromatin str section, add refs

@@
 The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons.
-The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) (hulten, Rubin et al 2020).
+The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) [@hulten2011_COM, @rubin2020mixing].
 
-Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-<1> the irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (referred to as along the tetrad arms (cite)).
-
-<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome structure' which is a consequence of kinetochores attaching to the MT-spindle and tension applied.
-Single crossover bivalents form what is known as cruciform tetrad in some systems (figure x).
-This area is reduced when crossover positions are shifted towards the telomere ends of the chromosome.
-The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.
-
-<3> Conserved feature of gametogenesis -- is the length of timing of the cell cycle (difference between spermat.
-faster spermatogenesis -- may select for synchronize the separation homologs or faster separate after alignment at metaphase.
-While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog structures)
-
-<In oocytes the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.>
+Under the **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
+The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite).
+The number and placement of crossovers alter the distribution of sister cohesion and the resulting 'chromosome structure' when bivalents are aligned on the metaphase plate (cite).
+The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.	Differing lengths of the timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures.
+Faster spermatogenesis may select for synchronization of the separation homologs or faster separatation at the metaphase to anaphase transition.
+While in oogenesis, the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.
 
@@
 
-Our results from MSM and PWD show that the SC length compaction is independent (of the direction of heterochiasmy.
-Two strains have have more crossovers in males, but females still have longer SC axis.
-**This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**
+Our results from musculus^MSM^ and musculus^PWD^ show that the SC length and chromatin compaction is uncoupled from the direction of heterochiasmy.
+These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
+While the relationship may hold for more heterochiasmy patterns, but these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction).
 
-The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013].
+<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>
 
-Very few models make predictions for how chromatin organization relates to variation in recombination rates / the recombination landscape.
-(the chromatin organization is how chromosomes with the same number of base pairs) are transformed into bottle brush structures with a central axis and most of the genetic material arranged as loops emanating outwards (figure).
 Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].
+This prediction / model might fit broader pattern such as in At where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
 

### 2020-04-04 doc/Discussion.tex
cleaning up chromatin str section, add refs

--- 2020-04-04 doc/Discussion.rmd
@@
----
-title: "Discussion Draft"
-date: ''
-output:
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-  pdf_document: default
-  word_document:
-    toc: yes
-    toc_depth: '5'
-bibliography: refs_Main_Report.bib
----
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Discussion Draft},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-```{r format.Notes, include=FALSE}
-#table, wite by all of the columns and rows
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-#over-view of the synthesis of hypothesis testing
-#(this is how we're synthsizing / testing / the framework were using to compare the results to the literature)
+%%% Change title format to be more compact
+\usepackage{titling}
 
-#constrast on model types
-#-evolution
-#-models accounting for chromosome based traits
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
-#HetC traits
-# 1. typical landscape
-# 2. female variance
-# 3. chromatin organization
+\setlength{\droptitle}{-2em}
 
-#faster male evolution traits
-# 1. Reversed Direction of Heterochiasmy
-# 2. correlated gwRR increase and interference strength
+  \title{Discussion Draft}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{}
+    \preauthor{}\postauthor{}
+    \date{}
+    \predate{}\postdate{}
 
-#go through each traits and point out which models support each of the traits
-```
+\begin{document}
+\maketitle
 
-```{r Model.testing, include=FALSE}
+We performed hypothesis testing of models from the literature and
+adapted here fit the main results from this study. We choose to examine
+three modifier models and two functional and cell physiology models
+(table X). It is a bit challenging to test all of these models together
+since they were built with difference parameters and built to describe
+different patterns of recombination variation. The modifier models were
+built to explain heterochiasmy variation between sexes and the COM
+functional model was built to describe the interference pattern and
+difference in the recombination landscape between the sexes. In cases
+where the models can't be extended to predicting results, we list as
+`NA'.
 
-#imortant of testable hypotheses
+\section{Reversed heterochiasmy
+direction}\label{reversed-heterochiasmy-direction}
 
-#model not included in this comparison
+The common direction of heterochiasmy is female biased, however male
+biased heterochiasmy examples are not especially rare. Our results are
+novel from previous examples of male biased heterochiasmy species
+because the panel of strains have much shorter evolutionary distances.
+This context gives us insights to sex-specific evolution of
+recombination rates.
 
-```
+The prediction from two of the modifier models, haploid selection and
+SACE, state that males will generally evolve lower recombination rates.
+Male biased heterochiasmy is predicted under the two locus modifier
+model when XXXX.
 
-We performed hypothesis testing of models from the literature and adapted here fit the main results from this study.
-We choose to examine three modifier models and two functional and cell physiology models (table X).
-It is a bit challenging to test all of these models together since they were built with difference parameters and built to describe different patterns of recombination variation.
-The modifier models were built to explain heterochiasmy variation between sexes and the COM functional model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
-In cases where the models can't be extended to predicting results, we list as 'NA'.
+There are no prediction for male biased heterochiasmy evolution
+direction under the COM , but it should be noted that this model was not
+designed to incorporate this type of data. This model was design to
+explain the positive interference pattern for of the recombination
+landscape by incorporating a cell mechanics of rapid prophase movements
+(RPM) that are thought to facilitate homolog pairing in many species.
 
-# Reversed heterochiasmy direction
+Under the spindle based model faster male evolution in male genome wide
+recombination rates is primarily driven the strength of the spindle
+assembly checkpoint (SAC) .
 
-```{r reversed.HetC, echo=FALSE}
+The genes and pathways are will known, we refer the reader to (Lane and
+Kauppi 2019).
 
-#our results of reversed hetC are novel due to the short evolutionary disances across the panel of mouse strains
+The transition from metaphase to anaphase is irreversible, so all
+bivalents in the cell have to be aligned correctly to prevent
+anueploidy.
 
-#spindle selection modle points
+Unstable bivalents are detected on the MI spindle, the SAC delays the
+cell cycle which potentially allows time for lose bivalents to correctly
+attach to the MT spindle. If the `unattached' persists apoptosis will be
+triggered. Genetic variants which increase the number of unstable
+bivalents, will be selected against in favor for genetic variants
+causing faster stabilization of bivalents on the MI spindle. Thus the
+SAC can lead to stabilizing selection on the chromosome structure.
 
-```
+Sperm have lower anueploidy rates compared to oocytes (other lines of
+evidence) indicating that the `strength' of the SAC varies between
+oogenesis and spermatogenesis. In sperm single unpaired bivalents can
+trigger the SAC. While in females more oocytes progress through meiosis
+even with unpaired bivalents (refereed to as leaky SAC).
 
-The common direction of heterochiasmy is female biased, however male biased heterochiasmy examples are not especially rare.
-Our results are novel from previous examples of male biased heterochiasmy species because the panel of strains have much shorter evolutionary distances.
-This context gives us insights to sex-specific evolution of recombination rates.
+These differences in SAC strength have been connected to conserved
+features of gametogenesis; centrosome spindle and cell volume (cite)
+which result in differing morphology of the metaphase microtubule
+spindle.
 
-The prediction from two of the modifier models, haploid selection and SACE, state that males will generally evolve lower recombination rates.
-Male biased heterochiasmy is predicted under the two locus modifier model when XXXX.
-<does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).>
+Directional selection could be the result in spermatogenesis if the
+fitness landscape for `chromosome structure' is altered via genetic
+architecture changes or GxE effects. (in oogenesis the leaky SAC has
+weak selection coefficient)
 
-There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model was not designed to incorporate this type of data.
-This model was design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.
+\section{Greater between cell
+variation}\label{greater-between-cell-variation}
 
-Under the spindle based model faster male evolution in male genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) <acting on the chromosome structure>.
+Our results of greater between cell variation in females, these results
+are supported by human and mouse (Gruhn, Lynn Kong). (Lynn et al. 2002).
 
-The genes and pathways are will known, we refer the reader to [@laneKauppi2019].
+These data are also scant, between cell variance isn't often measured
+especially for both sexes. (given they require single cell observations
+for both sexes).
 
-The transition from metaphase to anaphase is irreversible, so all bivalents in the cell have to be aligned correctly to prevent anueploidy.
+While the modifier models parameterize sexual dimorphism in the strength
+of selection which can shape the variance in a trait these types of
+models make predictions for variance between individuals but not within
+individuals where the genetics are constant. (selection on penetrance).
 
-Unstable bivalents are detected on the MI spindle, the SAC delays the cell cycle which potentially allows time for lose bivalents to correctly attach to the MT spindle.
-If the 'unattached' persists apoptosis will be triggered.
-Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-Thus the SAC can lead to stabilizing selection on the chromosome structure.
+In the spindle based selection model, the source of between cell
+variance is attributed to difference strength of checkpoints. As
+mentioned above, there is sexual dimorphism in the stringency of the
+spindle assembly checkpoint (SAC). And these are linked to conserved
+features of gametogenesis.
 
-Sperm have lower anueploidy rates compared to oocytes (other lines of evidence) indicating that the 'strength' of the SAC varies between oogenesis and spermatogenesis.
-In sperm single unpaired bivalents can trigger the SAC.
-While in females more oocytes progress through meiosis even with unpaired bivalents (refereed to as leaky SAC).
+The opposite dynamics of the purifying selection for the `strict' SAC in
+spermatocytes is relaxed selection from a `permissive' SAC. As mentioned
+above a fundamental sex difference in the metaphase spindle is the
+presence of centrioles in spermatocytes which anchors and nucleates the
+microtubulue spindle at the two cell poles. While oocytes lack
+centrioles which results in a difuse network of microtubules with
+mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
+This or larger cell volumne may lead/be a potnetial source of the leaky
+SAC in oocytes. Leaky SAC, relaxed selection on the anaphase to
+metaphase transition could increase the between cell variance for
+crossover number
 
-These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite) which result in differing morphology of the metaphase microtubule spindle.
+\section{Typical landscape}\label{typical-landscape}
 
-<leaky oocyte SAC linked to acentrosome spindle (multiple aMTOCs) and cell volume, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>
+Our results on the broad scale pattern of crossover placement are
+supported by the literature. As reviewed in (Sardell and Kirkpatrick
+2020), sexual dimorphism in the broad scale recombination landscape is a
+highly conserved trait.
 
-Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
-(in oogenesis the leaky SAC has weak selection coefficient)
+The \textbf{two locus modifier} and \textbf{SACE} model make predictions
+for sexual dimorphism in the broad scale recombination landscape, for
+diminishing the effect of drive systems and maintaining larger
+chromosome blocks respectively.
 
-# Greater between cell variation
+The two mechanical models also both predict sexual dimorphism in the
+recombination landscape for distinct reasons. The COM model (chromosome
+pairing based model), predicts the sex differences in the positioning of
+crossovers is due to a combination of the length of the axis and
+intensity of the rapid prophase movements (RPMs) (Hultén 2011, Rubin,
+Macaisne, and Huynh (2020)).
 
-```{r within.mouse.var, include=FALSE}
-#review results, other data
+Under the \textbf{spindle based model} we hypothesize that the sexual
+dimorphism in the recombination landscapes hinges on sex differences on
+the requirements for chromosome cohesion in late meiosis I. The
+irreversible process of the metaphase to anaphase transition is
+initiated by the protyltic decay of the sister cohesion connecting
+homologs (cite). The number and placement of crossovers alter the
+distribution of sister cohesion and the resulting `chromosome structure'
+when bivalents are aligned on the metaphase plate (cite). The male
+crossover telomere bias could be adaptive in order to synchronize (or
+minimize) the timing of homologs separating at metaphase I. Differing
+lengths of the timing of cell cycle between oogenesis and
+spermatogenesis may impose the the different selective pressures. Faster
+spermatogenesis may select for synchronization of the separation
+homologs or faster separatation at the metaphase to anaphase transition.
+While in oogenesis, the slower cell cycle and prolonged arrest may
+require chromosome structures with greater stability on the MI spindle.
 
-#2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
+\section{Chromatin Structure}\label{chromatin-structure}
 
-#2a.
-relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
-The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.
+Our results from musculus\textsuperscript{MSM} and
+musculus\textsuperscript{PWD} show that the SC length and chromatin
+compaction is uncoupled from the direction of heterochiasmy. These
+results slightly depart from predictions which nominate chromatin
+organization (axis length) as the primary driver of heterochiasmy
+(Petkov et al. 2007). While the relationship may hold for more
+heterochiasmy patterns, but these results indicate that rapid male
+specific evolution in gwRR, is through a different trait than chromatin
+compaction).
 
-#?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
+Only the COM model predicts sexual dimorphism in chromatin structure
+will be longer in oocytes due to greater cell volume. This prediction /
+model might fit broader pattern such as in At where pollen is the larger
+cell and has longer axis length (Cahoon and Libuda 2019).
 
-```
+\section{Evolution of Interference}\label{evolution-of-interference}
 
-Our results of greater between cell variation in females, these results are supported by human and mouse (Gruhn, Lynn Kong). [@lynn2002].
+\textbf{we find support that interference strength has evolved in the
+two groups of male strains. There is a small number of positive
+correlations between genome wide recombination rate and interference
+strength in the literature.}
 
-These data are also scant, between cell variance isn't often measured especially for both sexes. (given they require single cell observations for both sexes).
+Interference is a fundamental aspect of the recombination landscape. --
+genome-wide recombination rate and chromatin compaction
 
-While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
-(selection on penetrance).
+Some basics of the constraints bounds of interference are known, yet
+there are still many unknowns regarding the relationship of genome wide
+recombination rate and interference strength.
 
-In the spindle based selection model, the source of between cell variance is attributed to difference strength of checkpoints.
-As mentioned above, there is sexual dimorphism in the stringency of the spindle assembly checkpoint (SAC). And these are linked to conserved features of gametogenesis.
+Our results support an opposite pattern predicted from the `logical
+model', a positive correlation between interference strength and genome
+wide recombination rates; The within sex comparison of two breeds of
+cattle with different genome wide recombination rates (Ma et al. 2015),
+between lab and wild mice of Peromyscus leucopus, (Peterson, Miller, and
+Payseur 2019).
 
-The opposite dynamics of the purifying selection for the 'strict'  SAC in spermatocytes is relaxed selection from a 'permissive' SAC.
-As mentioned above a fundamental sex difference in the metaphase spindle is the presence of centrioles in spermatocytes which anchors and nucleates the microtubulue spindle at the two cell poles.
-While oocytes lack centrioles which results in a difuse network of microtubules with mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
-This or larger cell volumne may lead/be a potnetial source of the leaky SAC in oocytes.
-Leaky SAC, relaxed selection on the anaphase to metaphase transition could increase the between cell variance for crossover number
+Logically a negative correlation is expected; increasing the number of
+crossovers across chromosomes would most logically be done by more
+densely spacing crossovers along chromosomes and decreasing interference
+strength. This pattern has empirical support from the most species (Otto
+and Payseur 2019) and fits well with the fundamental relationship
+between the SC area or axis length, the physical upper limit for the
+number of crossovers. \textbf{although it is not particulary strong in
+our results} We argue that this is due to difference in scale (genome
+wide vs single chromosome)
 
-# Typical landscape
+Theoretical models haven't really considered the evolution of
+interference strength, neither the haploid selection or two locus
+modifier model can not be applied to evolution of interference strength.
+While the SACE modifier model does not explicitly model evolution of
+interference strength we note that a logical outcome of the main
+prediction of maintaining larger chromosome blocks in males, would be a
+landscape with stronger interference strength. The COM model predicts
+that interference and the recombination landscape arises from known
+oscillatory movements during prophase, it lacks evolutionary based
+predictions. We propose the spindle based selection model would support
+the evolution of interference strength in the positive direction via
+modulation of the amount of sister cohesion connecting homologs
+(figure).
 
-```{r typical.landscape.outline, include=FALSE}
+Models from Goldstein et al (review in (Otto and Payseur 2019) suggest
+that if this pattern is widespread interference evolves whenever
+increased recombination rates evolve. Perhaps a distinguishing feature
+of models which come to this finding is that the number of crossovers is
+kept constant. The space across multiple loci (veller?) or between
+multiple crossovers increases in a positive manner with genome wide
+recombination rates. Given that the empirical range of crossovers per
+chromosome is quite small (1-3 (Otto and Payseur 2019, Stapley et al.
+(2017)) and the obligate crossover rule, the assumption of constraining
+the number of crossovers per chromosome fits well with empirical data.
 
-# typical landscape is conserved -- sexually dimorphic meiotic trait (Sardell)
+Biological exceptions to interference: experimentally increasing
+crossover number through mutants or fusion chromosomes (Celegans,
+(plants, fungi with negative interference.
 
-# models
+\section{REVIEW}\label{review}
 
-#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
-#3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
-#3a.
-There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.
-In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle.
-(required to start with more since the cohesion isn't replenished.
-The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.
+We propose that the spindle based selection model is the most
+parsimounious with our results and known recombination patterns from the
+literautre.
 
-#3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
-#-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs.
-This area reduced when crossover position shifted to telomere (each end of chromosomes).
-Two crossover acrocentric chromosome the area is distal from the first until the second crossover.
+\begin{itemize}
+\item
+  It matches fine scale differences in how DSB hotspots had different
+  sex specific strengths with varied due to sex, epigenetc landscape and
+  the gene architecture (Brick et al)
+\item
+  incorporates know cellular / molecular mechanisms with a very common
+  pattern sexual dimorphic recombination rate landscape.
+\item
+  extends modifer model to explain a chromosome based / level pattern
+  \emph{weaknesses}, and ways the model diverges from our results
+\item
+  the simulation present in the paper are too weak to support the rapid
+  spread of such a modifer system (selection coefficients are too weak
+  to support the evolution -- spread of such a modifer.)
+\item
+  Other ways this model departs from our results, -Always predicting
+  male will be lower recombination rates (or risk breaking up the
+  epistatsis blocks) -- which doesn't fit with our results of faster
+  male specific evolution in gwRR. However / while not explicitaly
+  arameterized or predicted in this model, -- this model predicts larger
+  chuncks of chromosomes will segregate together (as outlined in Veller
+  et al) our results of stronger male interference and stronger
+  interference in high rec strains have the same consequences as
+  predited from this modle
+\end{itemize}
 
-#3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
-#- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.
+\section{Future Steps}\label{future-steps}
 
-#*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.
+Be aware of sex-specific and sex average biases. Some cytlogy models
+(Celegans) mostly focus on oogenesis and genetic maps are often from
+hermaphrodites. (lack of data from one sex). Recognize the value of
+Chromosome level measures and incorporating chromosome behavior into
+models.
 
-```
+Borader communication across sub-fields of cell physiology and
+biophysics for understanding recombination rate evolution. Utilize
+mitosis systems given the conservation of the molecular pathway. THink
+of meiosis as a program.
 
-Our results on the broad scale pattern of crossover placement are supported by the literature.
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
-
-The **two locus modifier** and **SACE** model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
-
-The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons.
-The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) [@hulten2011_COM, @rubin2020mixing].
-
-Under the **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting 'chromosome structure' when bivalents are aligned on the metaphase plate (cite).
-The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.	Differing lengths of the timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures.
-Faster spermatogenesis may select for synchronization of the separation homologs or faster separatation at the metaphase to anaphase transition.
-While in oogenesis, the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.
-
-# Chromatin Structure
-
-```{r chromatin.organization, include=FALSE}
-
-#supporting results from lit
-#chromatin organization
-
-#models
-# none of the models make predictions for the axis length (chromatin organization)
-# only COM predicts that sexual dimorphism in axis length is due to cell vol
-
-```
-
-Our results from musculus^MSM^ and musculus^PWD^ show that the SC length and chromatin compaction is uncoupled from the direction of heterochiasmy.
-These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
-While the relationship may hold for more heterochiasmy patterns, but these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction).
-
-<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>
-
-Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction / model might fit broader pattern such as in At where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
-
-# Evolution of Interference
-
-```{r evo.interference, include=FALSE}
-#review results,
-#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
-#(we're gonna spend alot of space on this section)
-
-# more convincing that the pattern is real
-
-#lil background, for interferences 
-#assuptions for predictions (most modifer models, )
-
-#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)
-
-#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)
-
-#**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
-
-#5. Interference evolution shows a positive correlation with gwRR
-#5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
-#5b.
-Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
-
-```
-
-**we find support that interference strength has evolved in the two groups of male strains.
-There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.**
-
-Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromatin compaction
-
-Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.
-
-Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
-
-Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength.
-This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
-**although it is not particulary strong in our results** We argue that this is due to difference in scale (genome wide vs single chromosome)
-
-Theoretical models haven't really considered the evolution of interference strength, neither the haploid selection or two locus modifier model can not be applied to evolution of interference strength.
-While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
-The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions.
-We propose the spindle based selection model would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).
-
-Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve.
-Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant.
-The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates.
-Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.
-
-Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.
-
-# REVIEW
-
-We propose that the spindle based selection model is the most parsimounious with our results and known recombination patterns from the literautre.
-
-  - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)
-
- - incorporates know cellular / molecular mechanisms with a very common pattern sexual dimorphic recombination rate landscape.
-
-- extends modifer model to explain a chromosome based / level pattern
- _weaknesses_, and ways the model diverges from our results
-
- - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
-- Other ways this model departs from our results, 
--Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
-However / while not explicitaly arameterized or predicted in this model,
- -- this model predicts larger chuncks of chromosomes will segregate together
- (as outlined in Veller et al)
-our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle
-
-```{r final.review.outline, echo=FALSE}
-# comes down to explaning the interference correlation
-
-# spindel selection -- 
-#good for sex differences because it frames heterochiasmy within the context of meiosis
-
-#can fit the interference evolution due to the (pressumed) effect on bivalent shape
-
-#caveats and unanswered parts
-
-# SACE is runner up
-
-```
-
-#Future Steps
-
-Be aware of sex-specific and sex average biases.
-Some cytlogy models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites.
-(lack of data from one sex).
-Recognize the value of Chromosome level measures and incorporating chromosome behavior into models.
-
-Borader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution.
-Utilize mitosis systems given the conservation of the molecular pathway.
-THink of meiosis as a program.
-
-Dark matter of the genome.
-We can't really explude that some of the recombination rate variation is due to strucutreal variation (more of less genome).
-(since everything is mapped to B6).
-(also maybe there are crossovers in the centromeres and telomeres -- which might escape detected in linkage maps)
+Dark matter of the genome. We can't really explude that some of the
+recombination rate variation is due to strucutreal variation (more of
+less genome). (since everything is mapped to B6). (also maybe there are
+crossovers in the centromeres and telomeres -- which might escape
+detected in linkage maps)
 
@@
 
-How does elevated rates of precocious sister centromere seperation effect oogenesis?
+How does elevated rates of precocious sister centromere seperation
+effect oogenesis?
 
-Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).
+Are selfish elements affecting the evolution of meiosis -- and thus
+recombination rates (from a functional standpoint).
 
@@
 
-#blah
+\section{blah}\label{blah}
 
-The SAC 'read's that kinetochores (the movement machinery built on top of the centromere) are connected to the MT of the spindle originating from the correct pole of the cell. (tension is )
+The SAC 'read's that kinetochores (the movement machinery built on top
+of the centromere) are connected to the MT of the spindle originating
+from the correct pole of the cell. (tension is )
 
-centromeres have single and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
+centromeres have single and opposite pole connection (synthellic?) and
+that the sister cohesion connecting the homologs -- sets up the tension
+-- opposes the pull from each of the poles.
 
-Efficient normal SAC -- works by pausing entry into anaphase or triggers apoptosis -- if a unpaired / tension-lacking tetrad is detected
-(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher threshold for being triggered  (4-5?)
-(Land Kauppi,
+Efficient normal SAC -- works by pausing entry into anaphase or triggers
+apoptosis -- if a unpaired / tension-lacking tetrad is detected (sperm
+SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while
+the female SAC (has a higher threshold for being triggered (4-5?) (Land
+Kauppi,
 
-(relaxed selection - / inefficient SAC -- in females -- fits with higher rate of achiasmata bivalents (there might be alternative mechanisms facilitating achiasmate segregation)
+(relaxed selection - / inefficient SAC -- in females -- fits with higher
+rate of achiasmata bivalents (there might be alternative mechanisms
+facilitating achiasmate segregation)
 
-  Mechanisms for sex differences in the strength of the SAC are the cell volumne (diffusion of signal molecules) and the centrosome being present at the spindle have been proposed in the literature.
-(testing / proving the mechanism for the new optima is for the high recombining group in males is out of the scope of this paper, but we propose some hypothesis are suggested below.
- but the evolutionary pattern fits that of directional.
+Mechanisms for sex differences in the strength of the SAC are the cell
+volumne (diffusion of signal molecules) and the centrosome being present
+at the spindle have been proposed in the literature. (testing / proving
+the mechanism for the new optima is for the high recombining group in
+males is out of the scope of this paper, but we propose some hypothesis
+are suggested below. but the evolutionary pattern fits that of
+directional.
 
-#References
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-cahoonLibuda2019}{}
+Cahoon, Cori K, and Diana E Libuda. 2019. ``Leagues of Their Own:
+Sexually Dimorphic Features of Meiotic Prophase I.'' \emph{Chromosoma}.
+Springer, 1--16.
+
+\hypertarget{ref-hulten2011_COM}{}
+Hultén, Maj A. 2011. ``On the Origin of Crossover Interference: A
+Chromosome Oscillatory Movement (Com) Model.'' \emph{Molecular
+Cytogenetics} 4 (1). Springer: 10.
+
+\hypertarget{ref-laneKauppi2019}{}
+Lane, Simon, and Liisa Kauppi. 2019. ``Meiotic Spindle Assembly
+Checkpoint and Aneuploidy in Males Versus Females.'' \emph{Cellular and
+Molecular Life Sciences} 76 (6). Springer: 1135--50.
+
+\hypertarget{ref-lynn2002}{}
+Lynn, Audrey, Kara E Koehler, LuAnn Judis, Ernest R Chan, Jonathan P
+Cherry, Stuart Schwartz, Allen Seftel, Patricia A Hunt, and Terry J
+Hassold. 2002. ``Covariation of Synaptonemal Complex Length and
+Mammalian Meiotic Exchange Rates.'' \emph{Science} 296 (5576). American
+Association for the Advancement of Science: 2222--5.
+
+\hypertarget{ref-ma2015_cattle}{}
+Ma, Li, Jeffrey R O'Connell, Paul M VanRaden, Botong Shen, Abinash
+Padhi, Chuanyu Sun, Derek M Bickhart, et al. 2015. ``Cattle Sex-Specific
+Recombination and Genetic Control from a Large Pedigree Analysis.''
+\emph{PLoS Genetics} 11 (11). Public Library of Science.
+
+\hypertarget{ref-ottoPaysuer2019}{}
+Otto, Sarah P, and Bret A Payseur. 2019. ``Crossover Interference:
+Shedding Light on the Evolution of Recombination.'' \emph{Annual Review
+of Genetics} 53. Annual Reviews: 19--44.
+
+\hypertarget{ref-peterson2019}{}
+Peterson, April L, Nathan D Miller, and Bret A Payseur. 2019.
+``Conservation of the Genome-Wide Recombination Rate in White-Footed
+Mice.'' \emph{Heredity} 123 (4). Nature Publishing Group: 442--57.
+
+\hypertarget{ref-petkov2007}{}
+Petkov, Petko M, Karl W Broman, Jin P Szatkiewicz, and Kenneth Paigen.
+2007. ``Crossover Interference Underlies Sex Differences in
+Recombination Rates.'' \emph{Trends in Genetics} 23 (11). Elsevier:
+539--42.
+
+\hypertarget{ref-rubin2020mixing}{}
+Rubin, Thomas, Nicolas Macaisne, and Jean-René Huynh. 2020. ``Mixing and
+Matching Chromosomes During Female Meiosis.'' \emph{Cells} 9 (3).
+Multidisciplinary Digital Publishing Institute: 696.
+
+\hypertarget{ref-sardell_sex_2020}{}
+Sardell, Jason M., and Mark Kirkpatrick. 2020. ``Sex Differences in the
+Recombination Landscape.'' \emph{The American Naturalist} 195 (2):
+361--79. doi:\href{https://doi.org/10.1086/704943}{10.1086/704943}.
+
+\hypertarget{ref-stapley_variation_2017}{}
+Stapley, Jessica, Philine G. D. Feulner, Susan E. Johnston, Anna W.
+Santure, and Carole M. Smadja. 2017. ``Variation in Recombination
+Frequency and Distribution Across Eukaryotes: Patterns and Processes.''
+\emph{Philosophical Transactions of the Royal Society B: Biological
+Sciences} 372 (1736): 20160455.
+doi:\href{https://doi.org/10.1098/rstb.2016.0455}{10.1098/rstb.2016.0455}.
+
+\end{document}

### 2020-04-04 doc/Results.rmd
starting a figure for IFDs

@@
 Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized IFD")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
+
+#each point is a 2CO, with the F1 and F2 position
+
+P1.by.P2_plot_raw <- ggplot(Curated_BivData.22, aes(y = Foci2, x= Foci1, color=strain))+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized IFD")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")
 
@@
 
+```{r Q1.IFD.plot, echo=FALSE}
+
+#make interference plots of IFD, position p1 by p2
+
+```
+
 ```{r Comp_Ratios, echo=FALSE}

### 2020-04-04 doc/Results.rmd
thinking of IFD plots

@@
 #each point is a 2CO, with the F1 and F2 position 
-
-P1.by.P2_plot_raw <- ggplot(Curated_BivData.22, aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")
+#based on plot from Dourad et al (Mercier)
+
+P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "WSB"),
+    aes(y = Foci2, x= Foci1, color=strain))+
+  geom_jitter(alpha=.5)+
+  ggtitle("Raw IFD WSB")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#trying to move x axis to top
+#cowplot might be the easier way -- since kinda diffecult in ggplot
+#or ggdraw
+
+P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
 

### 2020-04-04 doc/Discussion.rmd
more small changes

@@
 This context gives us insights to sex-specific evolution of recombination rates.
-
-The prediction from two of the modifier models, haploid selection and SACE, state that males will generally evolve lower recombination rates.
 Male biased heterochiasmy is predicted under the two locus modifier model when XXXX.
@@
 
-There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model was not designed to incorporate this type of data.
-This model was design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.
+**strict SAC drives directional selection**,  
+**genetic variants which lead to 'unstable chromosome structure will be selected agaisnt**  
+**sex differences in SAC strength are connected to conserved differences in gametogensis**
 
-Under the spindle based model faster male evolution in male genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) <acting on the chromosome structure>.
-
-The genes and pathways are will known, we refer the reader to [@laneKauppi2019].
+Under the **spindle based** model faster male evolution in male genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) acting on the chromosome structure at metaphase alignment.
+We will not go into much detail on the mechanism of the SAC, The genes and pathways are will known, we refer the reader to [@laneKauppi2019].
 
@@
 Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-Thus the SAC can lead to stabilizing selection on the chromosome structure.
 
@@
 
-Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromatin compaction
-
-Some basics of the constraints bounds of interference are known, yet there are still many unknowns regarding the relationship of genome wide recombination rate and interference strength.
-
-Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+Interference is a fundamental aspect genome-wide recombination rate the, chromatin compaction and the resulting recombination landscape.
 
@@
 **although it is not particulary strong in our results** We argue that this is due to difference in scale (genome wide vs single chromosome)
+
+Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
 
@@
 
-# REVIEW
+#Future Steps
 
-We propose that the spindle based selection model is the most parsimounious with our results and known recombination patterns from the literautre.
-
-  - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)
-
- - incorporates know cellular / molecular mechanisms with a very common pattern sexual dimorphic recombination rate landscape.
-
-- extends modifer model to explain a chromosome based / level pattern
- _weaknesses_, and ways the model diverges from our results
-
- - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
-- Other ways this model departs from our results, 
--Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
-However / while not explicitaly arameterized or predicted in this model,
- -- this model predicts larger chuncks of chromosomes will segregate together
- (as outlined in Veller et al)
-our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle
-
-```{r final.review.outline, echo=FALSE}
-# comes down to explaning the interference correlation
-
-# spindel selection -- 
-#good for sex differences because it frames heterochiasmy within the context of meiosis
-
-#can fit the interference evolution due to the (pressumed) effect on bivalent shape
-
-#caveats and unanswered parts
-
-# SACE is runner up
-
+```{r future.step.plan, echo=FALSE}
+#still preliminary
 ```
 
-#Future Steps
-
-Be aware of sex-specific and sex average biases.
-Some cytlogy models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites.
+- sex-specific and sex average biases.
+Some cytology models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites.
 (lack of data from one sex).
@@
 
-Borader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution.
+Broader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution.
 Utilize mitosis systems given the conservation of the molecular pathway.
-THink of meiosis as a program.
+Think of meiosis as a program.
 
 Dark matter of the genome.
-We can't really explude that some of the recombination rate variation is due to strucutreal variation (more of less genome).
+We can't really exclude that some of the recombination rate variation is due to structural variation (more of less genome).
 (since everything is mapped to B6).
@@
 
-How does elevated rates of precocious sister centromere seperation effect oogenesis?
+How does elevated rates of precocious sister centromere separation effect oogenesis?
 
@@
 
-#blah
-
-The SAC 'read's that kinetochores (the movement machinery built on top of the centromere) are connected to the MT of the spindle originating from the correct pole of the cell. (tension is )
-
-centromeres have single and opposite pole connection (synthellic?) and that the sister cohesion connecting the homologs -- sets up the tension -- opposes the pull from each of the poles.
-
-Efficient normal SAC -- works by pausing entry into anaphase or triggers apoptosis -- if a unpaired / tension-lacking tetrad is detected
-(sperm SAC is finer tuned (can detect 1-2 tetrads lacking tension -- ), while the female SAC  (has a higher threshold for being triggered  (4-5?)
-(Land Kauppi,
-
-(relaxed selection - / inefficient SAC -- in females -- fits with higher rate of achiasmata bivalents (there might be alternative mechanisms facilitating achiasmate segregation)
-
-  Mechanisms for sex differences in the strength of the SAC are the cell volumne (diffusion of signal molecules) and the centrosome being present at the spindle have been proposed in the literature.
-(testing / proving the mechanism for the new optima is for the high recombining group in males is out of the scope of this paper, but we propose some hypothesis are suggested below.
- but the evolutionary pattern fits that of directional.
+- chromatin interference
 

### 2020-04-05 doc/Discussion.rmd
kinda fixed 2 locus part in hetC reversal

@@
 
-The common direction of heterochiasmy is female biased, however male biased heterochiasmy examples are not especially rare.
-Our results are novel from previous examples of male biased heterochiasmy species because the panel of strains have much shorter evolutionary distances.
-This context gives us insights to sex-specific evolution of recombination rates.
-Male biased heterochiasmy is predicted under the two locus modifier model when XXXX.
-<does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).>
+Female biased heterochiasmy is the most common pattern observed, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
+Our results of male biased heterochiasmy is novel due to the shorter evolutionary distances between the examined strains.
+
+The direction of heterochiasmy under the **two locus modfier model** is dependant on female specific recombination modifer supresses or increases recombination, if the modifer is linked to the driving locus, and the meiotic stage (reduction or equalatoral division) that the driver acts (see Table 1 of @brandvain2012scrambling).
 
@@
 
-Under the **spindle based** model faster male evolution in male genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) acting on the chromosome structure at metaphase alignment.
-We will not go into much detail on the mechanism of the SAC, The genes and pathways are will known, we refer the reader to [@laneKauppi2019].
+Under the **spindle based** model, faster male evolution of the genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) acting on the chromosome structure at metaphase alignment.
+We refer the reader to [@laneKauppi2019] for more details on the sAC pathway.
 
-The transition from metaphase to anaphase is irreversible, so all bivalents in the cell have to be aligned correctly to prevent anueploidy.
+The transition from metaphase to anaphase is irreversible,
+
+so all bivalents in the cell have to be aligned correctly to prevent anueploidy.
 

### 2020-04-05 doc/Discussion.rmd
more additions / clean up of reversed hetC section

@@
 
-The direction of heterochiasmy under the **two locus modfier model** is dependant on female specific recombination modifer supresses or increases recombination, if the modifer is linked to the driving locus, and the meiotic stage (reduction or equalatoral division) that the driver acts (see Table 1 of @brandvain2012scrambling).
+The direction of heterochiasmy under the **two locus modfier model** is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) that the driver acts (see Table 1 of @brandvain2012scrambling).
 
@@
 Under the **spindle based** model, faster male evolution of the genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) acting on the chromosome structure at metaphase alignment.
-We refer the reader to [@laneKauppi2019] for more details on the sAC pathway.
+We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.
 
-The transition from metaphase to anaphase is irreversible,
-
-so all bivalents in the cell have to be aligned correctly to prevent anueploidy.
-
-Unstable bivalents are detected on the MI spindle, the SAC delays the cell cycle which potentially allows time for lose bivalents to correctly attach to the MT spindle.
-If the 'unattached' persists apoptosis will be triggered.
+The transition from metaphase to anaphase is irreversible, so prevention of annueploidy requires that all bivalents be correctly aligned before meiosis can progress (cite).
+Unstable bivalents detected on the MI spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize / connect to the microtubule spindle.
+If the 'unattached' / unstable bivalents fail to attach, the SAC triggers apoptosis.
 Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
 
-Sperm have lower anueploidy rates compared to oocytes (other lines of evidence) indicating that the 'strength' of the SAC varies between oogenesis and spermatogenesis.
-In sperm single unpaired bivalents can trigger the SAC.
-While in females more oocytes progress through meiosis even with unpaired bivalents (refereed to as leaky SAC).
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, cite).
+The different the detection of achiasmated bivalents (), the rates of annuploidy in mature gametes, and stronger infertility phenotypes in mutants of key SAC genes (cite).
 
-These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite) which result in differing morphology of the metaphase microtubule spindle.
+These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite).
+
+-- Supporting this stage/cell physiology based mechanism  influencing broad sexual dimorphic phenotypes in recombination landscape.
+(leading to distinct evolutionary trajectories) **directional selection** in males and relaxed selection in females.
 
@@
 
-Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
-(in oogenesis the leaky SAC has weak selection coefficient)
+<Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
+(in oogenesis the leaky SAC has weak selection coefficient)>
 

### 2020-04-05 doc/Discussion.rmd
little more

@@
 
+**connecting reverse HetC to faster male evolution**
+
 **strict SAC drives directional selection**,  
@@
 
-Under the **spindle based** model, faster male evolution of the genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) acting on the chromosome structure at metaphase alignment.
-We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.
+Under the **spindle based** model, faster male evolution of the genome wide recombination rate, (and instances of reversed heterochiasmy) is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversable metaphase-to-anaphase transition (cite).
+A stable chromosome structure requires, i) kinetochores attched to opposite poles of the cells and ii) at least one crossover to create tension across a propotion of sister cohesion (cite).
+Unstable bivalents on the MI spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle.
+If unstable bivalents fail to attach, the SAC triggers apoptosis.
+Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
 
-The transition from metaphase to anaphase is irreversible, so prevention of annueploidy requires that all bivalents be correctly aligned before meiosis can progress (cite).
-Unstable bivalents detected on the MI spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize / connect to the microtubule spindle.
-If the 'unattached' / unstable bivalents fail to attach, the SAC triggers apoptosis.
-Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, cite).
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, Sub cite).
 The different the detection of achiasmated bivalents (), the rates of annuploidy in mature gametes, and stronger infertility phenotypes in mutants of key SAC genes (cite).
@@
 These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite).
-
--- Supporting this stage/cell physiology based mechanism  influencing broad sexual dimorphic phenotypes in recombination landscape.
+These results // Supporting this stage/cell physiology based mechanism  influencing broad sexual dimorphic phenotypes in recombination landscape.
 (leading to distinct evolutionary trajectories) **directional selection** in males and relaxed selection in females.

### 2020-04-05 doc/Discussion.rmd
distracted by papers

@@
 
-Our results of greater between cell variation in females, these results are supported by human and mouse (Gruhn, Lynn Kong). [@lynn2002].
+Our results of greater between cell variation in females, these results are supported by human and mouse [@lynn2002, @gruhn2013].
 
-These data are also scant, between cell variance isn't often measured especially for both sexes. (given they require single cell observations for both sexes).
+While the **modifier models** parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
 
-While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
-(selection on penetrance).
+In the **spindle based** selection model, the less efficient SAC in oocytes could lead to relaxed selection on the chromosome structure.
+If the SAC/checkpoint which normally acts in a stabilizing selection on the chromosome structure has a reduction in efficiency (relaxed selection), will result in increased variance in (these associated traits), not just across strains (populations), but across cells within individuals (quant-gen ref?).
+Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007) which result in
 
-In the spindle based selection model, the source of between cell variance is attributed to difference strength of checkpoints.
-As mentioned above, there is sexual dimorphism in the stringency of the spindle assembly checkpoint (SAC). And these are linked to conserved features of gametogenesis.
+<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and biorientation of bivalents in the metaphase cell> -- spindle dynamics much different due to asymetrical diviation Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle.
 
-The opposite dynamics of the purifying selection for the 'strict'  SAC in spermatocytes is relaxed selection from a 'permissive' SAC.
-As mentioned above a fundamental sex difference in the metaphase spindle is the presence of centrioles in spermatocytes which anchors and nucleates the microtubulue spindle at the two cell poles.
-While oocytes lack centrioles which results in a difuse network of microtubules with mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
-This or larger cell volumne may lead/be a potnetial source of the leaky SAC in oocytes.
-Leaky SAC, relaxed selection on the anaphase to metaphase transition could increase the between cell variance for crossover number
+- which might have a consequence of reducing the (reaction/signal of the SAC (anaphase promoting complexes (cite) )).
+(this is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite))
+
+<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromsomes>
+Mogessie et al 2018 -- meiotic spindle (big)review>
+(what generates MT nucleation is unclear (aMTOC or chromsoomes))
 

### 2020-04-06 doc/Discussion.rmd
futial effort from last night

@@
 Our results of greater between cell variation in females, these results are supported by human and mouse [@lynn2002, @gruhn2013].
-
 While the **modifier models** parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
 
-In the **spindle based** selection model, the less efficient SAC in oocytes could lead to relaxed selection on the chromosome structure.
-If the SAC/checkpoint which normally acts in a stabilizing selection on the chromosome structure has a reduction in efficiency (relaxed selection), will result in increased variance in (these associated traits), not just across strains (populations), but across cells within individuals (quant-gen ref?).
-Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007) which result in
+In the **spindle based** selection model, the less efficient SAC in oocytes leads to relaxed selection on the chromosome structure due to meiotic progression even with with unstable bivalents.
+This will result in increased variance in chromosome structure, and crossover number, not just across strains, but across cells within individuals.
+Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
+This is thought to reduce the triggering signal for the SAC.
+This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite)).
 
-<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and biorientation of bivalents in the metaphase cell> -- spindle dynamics much different due to asymetrical diviation Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle.
+<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and biorientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymetrical diviation Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>
 
-- which might have a consequence of reducing the (reaction/signal of the SAC (anaphase promoting complexes (cite) )).
-(this is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite))
-
-<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromsomes>
-Mogessie et al 2018 -- meiotic spindle (big)review>
-(what generates MT nucleation is unclear (aMTOC or chromsoomes))
+<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromsomes, Mogessie et al 2018 -- meiotic spindle (big)review>
 
@@
 As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
+The **two locus modifier** and **SACE** model predicts sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
 
-The **two locus modifier** and **SACE** model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.
-
-The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons.
-The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) [@hulten2011_COM, @rubin2020mixing].
+The two mechanical models both predict sexual dimorphism in the recombination landscape for distinct reasons.
+The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differening strengths of chromosome movements during the pairing process [@hulten2011_COM, @rubin2020mixing].
 
@@
 The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting 'chromosome structure' when bivalents are aligned on the metaphase plate (cite).
-The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.	Differing lengths of the timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures.
-Faster spermatogenesis may select for synchronization of the separation homologs or faster separatation at the metaphase to anaphase transition.
-While in oogenesis, the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.
+The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate (cite).
+Different timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures on sister cohesion.
+Faster spermatogenesis may select for synchronization of the separation homologs.
+While in oogenesis, the slower cell cycle and mulitple arrest stages may require chromosome structures with greater stability on the MI spindle.
+
+<The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.>
 
@@
 
-**we find support that interference strength has evolved in the two groups of male strains.
-There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.**
-
 Interference is a fundamental aspect genome-wide recombination rate the, chromatin compaction and the resulting recombination landscape.
@@
 Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength.
+
 This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
-**although it is not particulary strong in our results** We argue that this is due to difference in scale (genome wide vs single chromosome)
+(**although it is not particulary strong in our results**).
 
-Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+**interference strength evolve**
+**other examples of interferene strength evolution, tricky due to chromosome specific effects and complications from large evolutionary distance (change in underlying karyotype**
 
-Theoretical models haven't really considered the evolution of interference strength, neither the haploid selection or two locus modifier model can not be applied to evolution of interference strength.
-While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
-The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions.
-We propose the spindle based selection model would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).
+Our results find that interference strength has evolved in males from 3 strains from our panel of house mice.
+
+Examples of evolution of interference strength are rare and complicated by changes in the underlying karyotype structure (Segura et al 2014).
+
+The general patern of the relationship
+ -- most findings / the general pattern of interference evolution is that it has a negative correlation to genome wide recombination rate.
+(Review in otto payseur), X examples -- correlation of X across x data.
+--This fit logic of XX -- and also can be applied to the known mechanism of DSB formation / and the relationship with chromosome axis.
+
+While examples of positive correlation of interference strength and genome-wide recombination rate are less common; the within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
+
+we propose that --- this is due to difference in scale (resolution)
+and how these two metrics are measured at the genome wide vs single chromosome (bivalent) level.
+
+<explain this difference>
+
+<**the models**>
+
+The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions.
+
+While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
+
+ We propose the **spindle based selection model** would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).
+
+<**other models**>
 
@@
 
-Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.
+<Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.>
 

### 2020-04-06 doc/Discussion.rmd
better lines for interference part

@@
 
-The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions.
-
+The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, however it lacks evolutionary based predictions.
 While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
 
- We propose the **spindle based selection model** would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).
+ We propose that the **spindle based selection model** support the evolution of stronger interference strength via selection onthe amount of sister cohesion connecting homologs (figure).
+-- and this can evolve with increased crossover number
 
-<**other models**>
+<we don't know the chromosome identities, but based on the propotions of chromosome classes this is the logical conclusion.
 
-Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve.
-Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant.
-The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates.
-Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.
+The evolution of gwRR in our study is driven by a transition of recombination landscapes in a propotion of chromosomes, from single crossovers to two crossovers --
 
-<Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.>
+All else being equal, axis length) -- interference strength would be expected to be equal or even weaker
 
-#Future Steps
+Our model suggests that an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs and could change the chromsome structure of these bivalents when they are aligned and bi-oriented.
+This model gives rise to the possibility that the chromosome structure at metaphase is the target of selection and changing the recombination landscape is an outcome  pleiotropic.
+This interpertation has potential to fit with hypotheses mentioned above for explaining other results regarding the evolution of the recombination landscape.
 
-```{r future.step.plan, echo=FALSE}
+We have focused on the model involving direct selection on the recombination pathway, that is the impact on the meotic process on an indiviual, but we make note on models centering on indirect selection or fitness outcomes of offspring which could also explain our results.
+Models from Goldstein et al, reviewed in [@ottoPaysuer2019] and Veller 2017?, predict that interference strength evolves whenever increased recombination rates evolve.
+The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
+A distinguishing feature of their model is that the number of crossovers per chromosome (genetic unit) is kept constant.
+The range of crossovers per chromosome is quite small, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017].
+Constraining the number of crossovers per chromosome in models may fit empiracal data better than those where recombiantion rate across an abstracted genetic space is unconstrained.
+
+```{r future.step.plan, echo=FALSE, include=FALSE}
 #still preliminary
-```
-
-- sex-specific and sex average biases.
+#sex-specific and sex average biases.
 Some cytology models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites.
@@
 Recognize the value of Chromosome level measures and incorporating chromosome behavior into models.
-
-Broader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution.
+#Broader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution.
 Utilize mitosis systems given the conservation of the molecular pathway.
@@
 
-Dark matter of the genome.
+#Dark matter of the genome.
 We can't really exclude that some of the recombination rate variation is due to structural variation (more of less genome).
@@
 
-what are the rates of sister-sister crossovers?
+#what are the rates of sister-sister crossovers?
+#How does elevated rates of precocious sister centromere separation effect oogenesis?
 
-How does elevated rates of precocious sister centromere separation effect oogenesis?
+#Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).
 
-Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).
-
-Keep generating those testable hypotheses (cite).
-
-- chromatin interference
+#Keep generating those testable hypotheses (cite).  
+#- chromatin interference
+```
 

### 2020-04-06 doc/Discussion.rmd
wraping up draft

@@
 We choose to examine three modifier models and two functional and cell physiology models (table X).
-It is a bit challenging to test all of these models together since they were built with difference parameters and built to describe different patterns of recombination variation.
-The modifier models were built to explain heterochiasmy variation between sexes and the COM functional model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
-In cases where the models can't be extended to predicting results, we list as 'NA'.
+It is a bit challenging to test all of these models together since they were built to describe different patterns of recombination variation.
+The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the COM model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
+In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
 
@@
 
-The direction of heterochiasmy under the **two locus modfier model** is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) that the driver acts (see Table 1 of @brandvain2012scrambling).
+The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) that the driver acts (see Table 1 of @brandvain2012scrambling).
 
-**connecting reverse HetC to faster male evolution**
-
-**strict SAC drives directional selection**,  
-**genetic variants which lead to 'unstable chromosome structure will be selected agaisnt**  
-**sex differences in SAC strength are connected to conserved differences in gametogensis**
-
-Under the **spindle based** model, faster male evolution of the genome wide recombination rate, (and instances of reversed heterochiasmy) is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversable metaphase-to-anaphase transition (cite).
-A stable chromosome structure requires, i) kinetochores attched to opposite poles of the cells and ii) at least one crossover to create tension across a propotion of sister cohesion (cite).
-Unstable bivalents on the MI spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle.
-If unstable bivalents fail to attach, the SAC triggers apoptosis.
-Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, Sub cite).
-The different the detection of achiasmated bivalents (), the rates of annuploidy in mature gametes, and stronger infertility phenotypes in mutants of key SAC genes (cite).
-
+Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversible metaphase-to-anaphase transition (cite).
+A stable chromosome structure, known as bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite).
+Unstable bivalents on the meiosis I spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle.
+If they fail to attach, the SAC triggers apoptosis.
+Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate or cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, Subrahman).
 These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite).
-These results // Supporting this stage/cell physiology based mechanism  influencing broad sexual dimorphic phenotypes in recombination landscape.
-(leading to distinct evolutionary trajectories) **directional selection** in males and relaxed selection in females.
+This model nominates conserved differences in in the context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
 
@@
 
-Our results of greater between cell variation in females, these results are supported by human and mouse [@lynn2002, @gruhn2013].
-While the **modifier models** parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
+Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013]. <More details on the empirarcal pattern>
 
-In the **spindle based** selection model, the less efficient SAC in oocytes leads to relaxed selection on the chromosome structure due to meiotic progression even with with unstable bivalents.
-This will result in increased variance in chromosome structure, and crossover number, not just across strains, but across cells within individuals.
-Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
-This is thought to reduce the triggering signal for the SAC.
-This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite)).
+While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
 
-<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and biorientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymetrical diviation Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>
+In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
+This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals.
+Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (Schuh Ellenburg 2007).
+This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite).
 
-<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromsomes, Mogessie et al 2018 -- meiotic spindle (big)review>
+<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>
+
+<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big)review>
 
@@
 As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
-The **two locus modifier** and **SACE** model predicts sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
+The two locus modifier and SACE model predicts sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
 
 The two mechanical models both predict sexual dimorphism in the recombination landscape for distinct reasons.
-The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differening strengths of chromosome movements during the pairing process [@hulten2011_COM, @rubin2020mixing].
+The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM, @rubin2020mixing].
 
-Under the **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
+Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
 The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite).
@@
 Faster spermatogenesis may select for synchronization of the separation homologs.
-While in oogenesis, the slower cell cycle and mulitple arrest stages may require chromosome structures with greater stability on the MI spindle.
+While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle.
 
@@
 These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
-While the relationship may hold for more heterochiasmy patterns, but these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction).
+While the relationship may hold for more general heterochiasmy patterns across broad scale comparisons, these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction.
 
 <The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>
-
 Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction / model might fit broader pattern such as in At where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+This prediction model might fit broader pattern such as in *Arabidopsis* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
 
@@
 
-Interference is a fundamental aspect genome-wide recombination rate the, chromatin compaction and the resulting recombination landscape.
+Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
+The general pastern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
+This fits the logic of the chromosome axis acting as the 'real estate' for DSBs and aligns with known molecular mechanisms (cite).
+Two examples of the opposite direction, a positive correlation of interference strength and genome-wide recombination rate, involve descriptions of observations at the single bivalent level; the between lab-raised and wild mice of *Peromyscus leucopus* from [@peterson2019] or large effects on genome wide recombination rates across a short evolutionary differences; the within-sex comparison of two breeds of cattle with different genome wide recombination rates [@ma2015_cattle].
+We propose that there is a difference in resolution and power between these two groups of empirical results with the positive and negative relationship between genome wide recombination rates and interference strength.
 
-Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength.
-
-This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers.
-(**although it is not particulary strong in our results**).
-
-**interference strength evolve**
-**other examples of interferene strength evolution, tricky due to chromosome specific effects and complications from large evolutionary distance (change in underlying karyotype**
-
-Our results find that interference strength has evolved in males from 3 strains from our panel of house mice.
-
-Examples of evolution of interference strength are rare and complicated by changes in the underlying karyotype structure (Segura et al 2014).
-
-The general patern of the relationship
- -- most findings / the general pattern of interference evolution is that it has a negative correlation to genome wide recombination rate.
-(Review in otto payseur), X examples -- correlation of X across x data.
---This fit logic of XX -- and also can be applied to the known mechanism of DSB formation / and the relationship with chromosome axis.
-
-While examples of positive correlation of interference strength and genome-wide recombination rate are less common; the within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
-
-we propose that --- this is due to difference in scale (resolution)
-and how these two metrics are measured at the genome wide vs single chromosome (bivalent) level.
-
-<explain this difference>
-
-<**the models**>
+We propose that the **spindle based selection model** can explain the positive correlation and evolution of stronger interference strength via selection on the amount of sister cohesion connecting homologs (figure X, table X).
+The evolution of genome wide recombination rates in our study is driven by a transition of recombination landscapes from single crossovers to two crossovers.
+All else being equal, interference strength would be expected to be equal or even weaker.
+Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain with more two crossover chromosomes.
+Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
+This model gives rise to the possibility that the chromosome structure at metaphase is the target of selection  which is achieved by changing the average recombination landscape across bivalents within the cell.
+This interpretation has potential to fit with hypotheses mentioned above for explaining other results regarding the evolution of the recombination landscape.
 
@@
 
- We propose that the **spindle based selection model** support the evolution of stronger interference strength via selection onthe amount of sister cohesion connecting homologs (figure).
--- and this can evolve with increased crossover number
+We have focused on the model involving direct selection on the recombination pathway, that is the impact on the meiotic process on an individual, but we make note on models centering on indirect selection or fitness outcomes of offspring which could also explain our results.
+Models from Goldstein et al, reviewed in [@ottoPaysuer2019] and Veller 2017, predict that interference strength evolves whenever increased recombination rates evolve.
+The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
+A distinguishing feature of their model is that the number of crossovers per chromosome is kept constant.
+The range of crossovers per chromosome is quite small, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017].
+Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
 
-<we don't know the chromosome identities, but based on the propotions of chromosome classes this is the logical conclusion.
-
-The evolution of gwRR in our study is driven by a transition of recombination landscapes in a propotion of chromosomes, from single crossovers to two crossovers --
-
-All else being equal, axis length) -- interference strength would be expected to be equal or even weaker
-
-Our model suggests that an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs and could change the chromsome structure of these bivalents when they are aligned and bi-oriented.
-This model gives rise to the possibility that the chromosome structure at metaphase is the target of selection and changing the recombination landscape is an outcome  pleiotropic.
-This interpertation has potential to fit with hypotheses mentioned above for explaining other results regarding the evolution of the recombination landscape.
-
-We have focused on the model involving direct selection on the recombination pathway, that is the impact on the meotic process on an indiviual, but we make note on models centering on indirect selection or fitness outcomes of offspring which could also explain our results.
-Models from Goldstein et al, reviewed in [@ottoPaysuer2019] and Veller 2017?, predict that interference strength evolves whenever increased recombination rates evolve.
-The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
-A distinguishing feature of their model is that the number of crossovers per chromosome (genetic unit) is kept constant.
-The range of crossovers per chromosome is quite small, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017].
-Constraining the number of crossovers per chromosome in models may fit empiracal data better than those where recombiantion rate across an abstracted genetic space is unconstrained.
+<Still working on Future Steps>
 

### 2020-04-07 doc/Results_v2.rmd
starting to modify results sectino

--- 2020-04-04 doc/Results.rmd
@@
 
-for(e in 1:length(cleanCurated_BivData$fileName)){
+#for(e in 1:length(cleanCurated_BivData$fileName)){
 #  mini.IFD <- calq.intra.rbar(mini.IFD)
-  cleanCurated_BivData$rbar[e] <- new.calq.intra.rbar(cleanCurated_BivData[e,])
-  }
+#  cleanCurated_BivData$rbar[e] <- new.calq.intra.rbar(cleanCurated_BivData[e,])
+#  }
 
@@
 
-summarize_rbar_table <- ddply(.data=cleanCurated_BivData,
-                 .(sex, strain),
-                 summarize,
-                 mean.rbar.all = mean(rbar, na.rm = TRUE),
-
-                 mean.rbar.2CO = mean(rbar[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean.rbar.3CO = mean(rbar[which(hand.foci.count == 3)], na.rm = TRUE)
-)
+#summarize_rbar_table <- ddply(.data=cleanCurated_BivData,
+#                 .(sex, strain),
+#                 summarize,
+#                 mean.rbar.all = mean(rbar, na.rm = TRUE),
+
+#                 mean.rbar.2CO = mean(rbar[which(hand.foci.count == 2)], na.rm = TRUE),
+#                 mean.rbar.3CO = mean(rbar[which(hand.foci.count == 3)], na.rm = TRUE)
+#)
 #the 3 is negative...
@@
 ```
+
+We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in 14 wild-derived inbred strains sampled from three subspecies of house mice: M.
+m.
+domesticus, M.
+m.
+musculus and M.
+m.
+molossinus.
+Mean MLH1 foci counts for XX mice were quantified from an average of XX spermatocytes per male (for a total of 1,867 spermatocytes) and XX oocytes per female (for a total of 1,409 oocytes).
+Graphical comparisons between the two sexes reveal several patterns (Figure 1A).
+First, recombination rate differs between females and males in most strains.
+Second, the direction and magnitude of heterochiasmy varies among strains.
+Although the majority of strains show higher recombination rates in females (following the pattern in laboratory mice), two musculus strains and one molossinus strain exhibit male-biased heterochiasmy.
+Hence, relative recombination rates in the two sexes are evolutionarily labile.
+Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C).
+Female recombination rates are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure 1B).
+In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).
 
@@
 
-```{r Q2.IFD_short.removed_t.test, echo=FALSE}
+```{r Q2.IFD_short.removed_t.test, echo=FALSE, include=FALSE, eval=FALSE}
 
@@
 ```
-
-The p values are still significant
 

### 2020-04-07 src/CommonFunc_MLH1repo.R

@@ -542,7 +542,7 @@ calq.intra.rbar <- function(chrm_row){
 #changed SC. and chrm length
 new.calq.intra.rbar <- function(chrm_row){
   o = 1
-  # rrbar <- c()
+  rrbar <- c()
   for(o in 1:length(chrm_row$Obj.ID) ){
     #print(chrm_row$fileName[o])
     if(grepl(0, chrm_row$hand.foci.count)){

### 2020-04-07 doc/Results_v2.rmd
incorporating bret's comments

@@
 library(reshape2)
-library(patchwork)
+#library(patchwork)
 
@@
 #use the above line for when order is out of wak
-
-#count the non-quality measures,  #remove non qualit
-#length(MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ] )  #15 rows with out quality scores, remove.
-#MLH1_data <- MLH1_data[ !(is.na(MLH1_data$quality) | MLH1_data$quality==""), ]
-
-#MLH1_by_F_strain <- MLH1_data[MLH1_data$sex == "female", ]
-#MLH1_by_M_strain <- MLH1_data[MLH1_data$sex == "male", ]
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(category) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
 
@@
 
-```{r, tot.SC.setup, echo = FALSE}
+```{r tot.SC.setup, echo = FALSE}
 
@@
 
-str(AP_mouse_table.HQ)
+#str(AP_mouse_table.HQ)
 
@@
 
-# Genome wide recombination rate estimates for both sexes
+# genome-wide recombination rate estimates for both sexes
 
@@
 
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in 14 wild-derived inbred strains sampled from three subspecies of house mice: M.
+```{r MLH1.sum.stats.Bret.edits, echo=FALSE}
+
+AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
+
+#total mice
+length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )
+
+#average cell number for spermatocytes and oocytes
+
+#mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
+
+#mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
+
+```
+
+We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M.
 m.
-domesticus, M.
+domesticus*, *M.
 m.
-musculus and M.
+musculus* and *M.
 m.
-molossinus.
-Mean MLH1 foci counts for XX mice were quantified from an average of XX spermatocytes per male (for a total of 1,867 spermatocytes) and XX oocytes per female (for a total of 1,409 oocytes).
+molossinus*.
+Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )` mice were quantified from an average of `r mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes) and `mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes).
+
 Graphical comparisons between the two sexes reveal several patterns (Figure 1A).
@@
 
-We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M.
-m.
-domesticus,_ _M.
-m.
-musculus_ and _M.
-m.
-molossinus_.
-Mean MLH1 foci counts for each mouse were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes.
-<!-- state the average number of mice
-The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises. -->
-
-<!--
-1. Fig1A  magnitude and direction of heterochiasmy;
--general patterns i) female higher and ii) low degree of hetC  (range of 1 to 1.2) (human is 1.6)
--All strain means are above the expected minimum of 20 (1CO per chromosome)
--there are exceptions they are followed up in the sex-specific plots  (male higher)
-
-2. Fig1B and C mouse means and sex specific patterns; 
--female mouse mean distributions are have more spread across the range of means
--male distributions are more clustered, low and high (~30% evolution largest-rapid evolution of gwRR)
--sex-specific evolution, both sexes don't show equivalent elevation of gwRR
--Dom vs Molossinus and Musculus -->
-
 The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
+Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
 Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
 
-Examining the mouse means of MLH1 foci per cell separately for each sex reveal distinct patterns of variation (Figure1 B and C).
-The female means are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B).
-While in males, the strain specific means separate more clearly into two groups of high genome wide recombination rates near 30 MLH1 foci per cell and low genome wide recombination rates, near 23 MLH1 foci per cell (Figure1 C).
-
-<illustrate sex differences female mean `r mean((main.table %>% filter(sex=="female") )$mean_co )`>
+Separtely examining the mouse means of MLH1 foci per cell separately for each sex reveal distinct patterns of variation (Figure1 B and C).
+Female recombination rate means are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B).
+In stark contrast, males specific means separate more clearly into two groups of strains with high genome-wide recombination rates (near 30 MLH1 foci per cell) and low genome-wide recombination rates, (near 23 MLH1 foci per cell) (Figure1 C).
 
@@
 
-<!--
-The variance across mouse means for each strain is also variable across strains. With _domesticus^LEW^_ having the most significant strain effect in the second model (glm; p= ).
- -->
-
 ```{r FEMALE.VAR.MODELS, echo=FALSE}
@@
 
- <These results allow use to reject the hypothesis of genome wide recombination rates (varying consistantly by sex across subepcies -- and having variance within subspecies due to strains)>
+ <These results allow use to reject the hypothesis of genome-wide recombination rates (varying consistantly by sex across subepcies -- and having variance within subspecies due to strains)>
 
@@
 
-The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombination rates.
+The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome-wide recombination rates.
 The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
@@
 <The greater amount of within mouse variance in females (almost and, (males have no detecable difference across strains for within mouse variance)
--- supports the distinct evolutionary pressures acting on the genome wide recombination rate in oocytes and spermatocytes.
-The striking patterns of variation across male and females suggest that the genome wide recombination rates have distinct evolutionary trajectories.
-For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome wide recombination rate.>
-
-# Evolution of genome wide recombination rate associated with evolution of mean DSB number
+-- supports the distinct evolutionary pressures acting on the genome-wide recombination rate in oocytes and spermatocytes.
+The striking patterns of variation across male and females suggest that the genome-wide recombination rates have distinct evolutionary trajectories.
+For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome-wide recombination rate.>
+
+# Evolution of genome-wide recombination rate associated with evolution of mean DSB number
 
@@
 
-For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
+For each strain, we deconstructed the estimates of genome-wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
 The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures [@peterson_2019].
@@
 
-# Evolution of genome wide recombination rate is reflected at the single chromosome level
+# Evolution of genome-wide recombination rate is reflected at the single chromosome level
 
@@
 
-The literature provides basic predictions for the direction of the relationship the genome wide recombination rate and two aspects of the recombination landscape; SC length (positive) and crossover interference (negative).
+The literature provides basic predictions for the direction of the relationship the genome-wide recombination rate and two aspects of the recombination landscape; SC length (positive) and crossover interference (negative).
 Following this logic we predict i) no strain effects for SC length or crossover interference within _domesticus_, ii) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which will have longer SC and weaker interference compared to the other _musculus_ strains and iii) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_.
@@
 
-Given the differences in the DSB numbers between a subset of the high and low recombining strains we predict that there will also be differences in SC lengths proportional to the increase in the mean genome wide recombination rate.
+Given the differences in the DSB numbers between a subset of the high and low recombining strains we predict that there will also be differences in SC lengths proportional to the increase in the mean genome-wide recombination rate.
 We compared mouse means of three SC based metrics; total SC, mean short bivalent length and mean long bivalent length.
@@
 The normalized 1CO position is not significantly different between the high and low recombining groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t.test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
-While there were significant strain effects for _domesticus^^WSB^_ and _molossinus^MOLF^_ (figure X), this evolution of the 1CO positioning is decoupled from the total genome wide recombination rate.
+While there were significant strain effects for _domesticus^^WSB^_ and _molossinus^MOLF^_ (figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
 Hence we don't follow up with the model framework.
@@
 given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models.
-Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation
+Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome-wide recombination rate variation
 <I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>
@@
 
-# Q2 Evolution of interference is associated with genome wide recombination rate evolution
+# Q2 Evolution of interference is associated with genome-wide recombination rate evolution
 
@@
 
-Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of evolution in genome wide recombination rates.
+Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of evolution in genome-wide recombination rates.
 However, the pattern is in the opposite direction to our prediction; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
@@
 
-<This striking differences across sexes suggest that the genome wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
+<This striking differences across sexes suggest that the genome-wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
 The pattern of strain averages distributed around the total sample mean in females is supportive of stabilizing selection or neutral evolution.

### 2020-04-07 doc/Results_v2.rmd
adding Bret's re-writes and removing commented out sections

@@
 
-# Evolutionary Framework
+# Partitioning variation in recombination rate
+
+To situate variation in recombination rate within an evolutionary framework, we
+We applied fit a series of models including subspecies, strain, and sex, to to our dataset of mean MLH1 foci counts from 187 mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
+We began with a full mixed model (M1, see Methods), which showed that strain (random effect p < someSmallNumber), sex (p = 3.9710^{-5}), subspecies (p= 1.7210^{-4}), and subspecies * sex (p = 3.110^{-5}) each significantly affect recombination rate.
+After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
+A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculusMSM (p= 3.9910^{-6}) and domesticusG (p= 1.0410^{-6}).
+In addition, two strains exhibit strain-by-sex interactions:  molossinusMSM  (p = 1.2610^{-4}) and musculusPWD (p = 3.8610^{-4}).
+We next fit general linear models separately for 192 males and 144 females (M4).
+In the male dataset, three strains significantly affect recombination rate: molossinusMSM (p = 2.2310^{-14}; effect = X MLH1 foci), All of the effects within the mixed model M1 were highly significant (LRT: subsp p= 1.7210^{-4}, sex p = 3.9710^{-5}, subspecies * sex p = 3.110^{-5}) and random LTR; strain p= 0 ).
+M1. Mixed Model 
+mouse av. MLH1 foci per cell = subsp*sex+rand(strain)+??
+M2. Linear Model
+mouse av. MLH1 foci per cell = subsp*sex*strain+??
+M3. Linear Model
+mouse av. MLH1 foci per cell = sex*strain+??
+M4. Linear Model
+sex specific mouse av. MLH1 foci per cell = subsp*strain+??
+
+The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombination rates.
+The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=0.22 and p= 0.09 for M.
+m.
+musculus and M.
+m.
+molossinus respectively), indicating that most of the variance is due to strain effects.
+As such, we focus on results from M3.
+The two most significant strain effects were domesticusG and musculusMSM (p= 1.0410^{-6} and p= 3.9910^{-6} for domesticusG and molossinusMSM respectively).
+In addition to significant strain effects for molossinusMSM, (glm; p = 0.06 and p = 3.9910^{-6}) both musculusPWD and musculusMSM had significant strain by sex effects (glm; p = 3.8610^{-4} and p = 1.2610^{-4}) musculusPWD by male and musculusMSM by male.
+Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in domesticusG.
+A linear model was run on the sex specific data sets (M4), with 192 male and 144 female mean MLH1 foci per cell.
+There were three significant strain effects in the male specific data; musculusPWD ((glm; p = 7.3710^{-10}; effect = X foci), and musculusSKIVE (glm; p = 0.01; effect = X foci).
+These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
+Analysis of the female dataset points to four strains with significant effects on recombination rate:  and molossinusMSM (glm; p=2.2310^{-14} with effect sizes ranging from 5, 7, and 2 foci respectively.
+Given these results, theses strains are classified as 'high recombining' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
+While the female specific linear model has many more had significant strain effects; domesticusG (glm; p = 2.510^{-6}), molossinusMSM (p = 6.2410^{-6}), domesticusLEW (glm; p = 0.01), and domesticusPWD musculusPWD (glm; p=0.02).
+Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males., domesticusMSM (glm; p=6.2410^{-6}), domesticusMOLF(glm; p=0.08), and domesticusKAZ(glm; p=0.1), the effect sizes have a smaller range of (4 to 1), indicating significant but small strain specific effects on mean MLH1 foci per cell.
+Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
 
@@
 
-## Within Mouse Variance in CO Count per Cell
+## Within mouse variance in CO count per cell
+
+Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
+To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
+Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber) and general linear model (M2) (p = 2.310^{-4}).
+In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
+Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (criteria?).
+The results are similar: sex is the strongest effect (p < smallNumber, M1; p = 2.310^{-4}, M2).
+When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
+These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
 
@@
 # Evolution of genome-wide recombination rate associated with evolution of mean DSB number
+
+Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
+
+In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
+DMC1 foci were scored from a total of 76 leptotene-stage and 75 zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculusKAZ , domesticusWSB , and domesticusG) and two high-recombination strains (musculusPWD and molossinusMSM).
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<someSmallNumber; one-way ANOVA, p = 0.00027).
+In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66; one-way-anova, p = 0.15).
+After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
+Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
+The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = 0.94 and p = 0.11 for leptotene and zygotene ratios, respectively).
+This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
 
@@
 
+Evolution of genome-wide recombination rate is reflected at the single chromosome level
+
+To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents (Peterson et al.
+2019).
+This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
+In this dataset, isolation rates per cell range from 0.51 (molossinusMSM male) to 0.72 (musculusKAZ female).
+From the total set of cell images, 10,458 bivalent objects were isolated by the image analysis software.
+After a human curation step (Peterson et al.
+2019), 9,829 single-bivalent observations remained.
+We assume that the isolation of bivalents within cells is unbiased.
+Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns.
+An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
+
+<figure>
+
+Ninety-six percent of single bivalents in our pooled dataset (n = 34,982) have either one or two crossovers (Figure X).
+The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
+low recombining strains (Figure X).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents  are 0.33 in musculusSKIVE , 0.44 in musculusPWD , and 0.53 in molossinusMSM .
+Following patterns in the genome-wide recombination rate, male musculusPWD and male molossinusMSM have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; musculusPWD vs.
+musculusKAZ p = 3.1510^{-33}; molossinusMSM vs.
+molossinusMOLF p = 4.7210^{-13}).
+Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
+We address two main questions.
+First, which traits are sexually dimorphic?
+Second, which traits differ between males from high-recombination vs.
+low-recombination strains?
+
 ```{r algorithm_performance_stats, echo=FALSE}
@@
 
-<!--
-- This is motivation to investigate more traits at the chromosome level
-1.	Basic stats - bivalent data numbers 
-2.	Rang of co per chrm, mostly 1co 2co
-3.	High rec strains have evolved a higher proportion of 2CO
-For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
-MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
-4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
--->
-
 <from the means the expected number of COs per chroosomes ranges from X to X we confirm these numbers Can't have fractions of COs>
@@
 
-<!-- breif background on longer female SC  (mouse, human) Gruhn-->
+Q1 Sex Differences in the Length of the Synaptonemal Complex
+
+In many mammalian species, the synaptonemal complex (SC) is longer in females.
+Yet, the majority of these observations come from species with female-biased heterochiasmy.
+Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
+In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male musculusPWD and male molossinusMSM should have longer SC lengths than the other strains.
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
+This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
+A total of 678 'short' bivalents were isolated from 103 oocytes and 37 spermatocytes.
+Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+
+<figure>
+
+Short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculusSKIVE.
+The female:male ratio ranges across strains from 1.15 (musculusMSM) to 1.49 (domesticusWSB).
+As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of Wang et al.
+(2019).
+Mouse means were calculated from cell-wide total SC lengths in 2,984 out of 3,680 cells with MLH1 counts (Figure X).
+Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
+That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
+Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies.
+
+<figure>
+
+In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
+Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers.
 
@@
 
-<!--
-In linear models (M2 and M3) for **mean short bivalent length**, sex is the most significant coefficient (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`) with slight significant strain and interaction effects (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively).
-
-The following effects had significant p values for the **total SC** from the above model (for the linear models M2 and M3), subspecies _musculus_ (glm M2; p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`), subspecies _molossinus_ (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`), sex male (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`), strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
-
-For M3 in the mean total SC lengths there were many strain effects are significant, but the largest are sex (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`)
-but also some strain effects, G strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][2]`), LEW strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][3]`), PWD strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][4]`).
-(this means...).
-Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel, when chromatin condensation is summarized at this level.
-
-(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
-for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
--->
-
 ```{r Q1.SC.Q12, echo=FALSE}
@@
 # Q1. 1CO Position Sex Differences
+Q1. Sex Differences in the Positions of Single Crossovers
+
+In most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus).
+In all strains, the landscape across 1CO bivalents is significantly different in females and males.
+Normalized foci positions tend to be more central in females (mean focus position = X) and closer to the telomere in males (mean focus position = Y) (t-test; p = 2.9210^{-22}).
+Sex is also the most significant effect on focus position in a mixed model (M1: p =1.2610^{-25}) and in general linear models (M2: p =1.3310^{-7}; M3: p = 1.3310^{-7}).
+These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species (Sardell Kirkpatrick).
 
@@
 
-<!-- 
-Brief lit review for IFD / interference expectations  
-**Petkov et al 2007,  CO interference underlies sex differences in RR**
-
-"Here we show that in mice, this is because of a shorter genomic interference distance in females than in males, measured in Mb.
-However, the interference distance is the same in terms of bivalent length.
-We propose a model in which the interference distance in the two sexes reflects the compaction of chromosomes at the pachytene stage of meiosis."
-
-Chrm1 genetic map, 
-(other human and mice refs 
-Tease n Hulten 2004: -- no difference between MLH1 foci in males and females
-
-DeBoer 2006, 13 -- measured chrm 1 in males and females, both were 2.8 microns -->
-
-<!--  set up -->
+Q1. Sex Differences in CO Interference (IFD)
+Crossover interference is one of the major determinates of the recombination landscape.
+It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
+To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated 1360 and 1272 2CO bivalents in females and males respectively.
+Mean interfocal distances (IFD) were calculated from 42 female and 45 male mice.
+We examined both raw IFDraw and normalized by SC length (IFDnorm).
+The mean IFDraw measures of crossover interference as a mechanical force mediated through the SC.
+We compared the IFDraw to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
+Previous measures found no sex difference in measures of IFDraw in mice (de Boer et al 2004).
+There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
+A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticusG; p = 0.27).
+The difference in mean IFDraw are slightly significant sexes from the full data set (t.test; p = 0.07).
+However domesticusG might be an outlier (by having a larger than average degree of sexual dimorphism).
+When the domesticusG observations are removed, the difference is not longer significant (t.test; p = 0.27).
+This result indicates there is no general pattern of sex differences in crossover interferencethat females and males exhibit a similar level of interference when it is measured in physical (SC) units.
+In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculusKAZ showing no difference (t-test; p = 0.33).
+While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+The difference in mean IFDnorm are highly significant between sexes.
+The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculusSKIVE p-value = 4.310^{-5}), musculusPWD p-value = 2.2810^{-4}), molossinusMSM p-value =0.01), domesticusLEW p-value = 3.5710^{-5}), domesticusG p-value = 0.05), and domesticusWSB (t.test; p-value =0.03) ).
+Only musculusKAZ did not have a difference in mean IFDnorm (t.test; p-value = 0.33).
+In examining theExamination of IFDnorm distributions, we note indicates that  the female IFDnorm observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFDnorm observationsvalues, whereas males are enriched for higher values.
+Comparitively, the IDFnorm distributions in males are enriched for longer IFDnorm.
+The three sets of modelsMixed models and general linear models of IFDnorm support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
+When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
+In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+
+Q1 Summary, consequences of sex differences in the recombination landscape
+Our results confirm sex differences in recombination landscapes described in several other species and we note that these differences have distinct consequences on the potential patterns of genetic variation resulting from the distinct recombination landscapes.
+The terminal position of single foci and the greater distance between two foci on the same chromosome will have a consequence of larger sections of linked sites segregating together in male gametes compared to female gametes (Veller).
+
+Additional Determinants of Genome-wide Recombination Rate Evolution in Males
+The next section is meant to focusNext, we used the contrast between high-recombination strains (musculusPWD , musculusSKIVE, and molossinusMSM) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( musculusPWD , musculusSKIVE, and molossinusMSM) to the low recombing strains.
+The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
+In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
+
+<equations>
+
+**Linear model M1**
+
+mouse average bivalentmetric = subsp*strain+??
+Linear model M2 
+mouse average bivalentmetric = strain+?? 
+The literature provides Previous empirical work suggests basic predictions for the direction of the relationship between the genome wide recombination rate and two aspects of the recombination landscape.; SC length is expected to be positively associated with genome-wide recombination rate because X.
+Crossover interference is expected to be negatively associated with genome-wide recombination rate because X.
+(positive) and crossover interference (negative).
+Following this logic we predict i) no strain effects for SC length or crossover interference within domesticus, ii(1) musculusPWD will have greater SC length and weaker interference than musculusSKIVE, which in turn will have longer SC and weaker interference compared to the other musculus strains, and iii(2) molossinusMSM will have longer SC and weaker interference compared to molossinusMOLF, and 3) domesticus strains will have similar SC length and crossover interference..
 
@@
 
-<!--
-The differences between the sexes is slightly different depending on if the males are in the high rec group. In the low rec strains (Dom + KAZ)...
-(both males and females IFD distributions have lower and upper thresholds around 25% and 75%) respectively.
-however the weight of observations are slightly shifted across the sexes.
-There's an enrichment of lower IFD^norm^ in females relative to males, while the opposite is true for the higher threshold -- where males have an enrichment of higher/long IFD^norm^ compared to females.
-
-In the three high rec strains -- the distributions have a more striking difference of the lower threshold being much higher in males, aproximately 30% to 40% compared to 25% in females.
-Also while the female upper threshold is capped at 75%, in males (some strains have higher upper thresholds -- and it's more likely to observe IFD^norm^ above the threshold.
-
-G - the upper and lower thresholds of females look pretty similar, the lower is closer to .2 than .25 -- female has more lower IFDs
-
-WSB - higher upper threshold (female's have many below the 25% lower threshold, males have a few below 25% threshold)
-
-LEW -- upper threshold looks very similar (.75) in males and females, much fewer low threshold obs in males,
-
-KAZ  -- female upper threshold around .75, lower threshold around .25 the distribution is quite uniform and spread out.  
-Male pattern; upper threshold slightly higher than .75 (.8) and lower threshold seems to extend beyond .25 -- but maybe low sampling contributes to the spread?
-(the spread in IFD values seems much larger for KAZ males compared to males from other strains.)
-
-PWD -- female lower threshold extends much lower .25 (like .15?),  upper threshold capped at .75
-Male lower threshold very different (higher like .3) the male upper threshold is above .75
-
-MSM -- female thresdholds, upper set around .75(to .65) lower at mostly .25, some cells(mice) have more .1 to .2 IFD observatsions.  
-Male upper threshold it at .75 while lower is set to .3 or .4.
-
-SKIVE -- female upper threshold is capped at .75, lower threshold is between .25 to .3 but low rate of outliers below .25.
-Male lower threshold is moved up, .4 or .35 and upper threshold moved up aprx .8
--->
-
-<!-- lmer/ glms  **M1. Mixed Model** $$mouse \ average \ IFD ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear Model** $$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$
-
-**M3. Linear model**
-
-$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$
--->
-
 ```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-<!-- junk for Q1.IFD
-
-  For the normalized values in both M2 and M3, sex is a significant effect, increasing 
-with males nrm.IFD in males.
-  SKIVE*male is the only other consistently significant effect, which also increases the nrm.IFD measure (p = male-SKIVE p =  ).
-For the mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. (IFD.PER), sex is most significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`)
-
-<For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw.
-The other effects are for PWD and SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance.
-For the M3, only SKIVE*male effect is close to significant>
-
-<remove the chrm size effects -- or check out model with chromosome length
-to rule out chromosome size effects, (espically in KAZ) I re-run the ttests using the short and long biv data sets.
-Our predictions for these are that longer chromsomes will show more of the 'typical male pattern' while short chromosomes often have a more uniform positing (hence tighter)>
-
-Plotting the single bivalent data for the short and the long data set show in KAZ -- there are two observations near the lowerest IFD1.PER, (even lower than female).
-These come from 2 different males and I looked up the images to confirm that the measures are accurate (by the bivalent).
-
-The above plots display the raw and normalized area which seperate two foci on the same chromosome.
-In the normalized plot, 9 observations fall above the length threshold of 150.
-The female measures fall more uniformly across the range of (0 to 1).
--->
-
 ```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
@@
 
+We compared mouse means of three SC-based metrics: total SC length, mean short bivalent length, and mean long bivalent length.
+The long bivalent data set was isolated by choosing bivalents in the X quartile of the SC length distribution within cells.
+The short bivalent dataset was isolated as described above.
+
+figure
+
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ([report spearman's rho and p-value]).
+Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
+While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01), separate tests by strain show that only within molossinus are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03; musculus: p = 0.87).
+Additionally, the means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = 0.88; long: p = 0.18).
+In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6}, molossinus p = 10^{-6}).
+In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
+
+<bret's version above>
+
 Given the differences in the DSB numbers between a subset of the high and low recombining strains we predict that there will also be differences in SC lengths proportional to the increase in the mean genome-wide recombination rate.
@@
 The long bivalent data set was isolated in the same manner as the short bivalent data set.
-
-<!-- basic stats
-`r length(OG.MLH1.merge.Skel$fileName)` total skelentonize measures were made from all of the MLH1 pachtyene cell images.
-X observations removed for having extra signal from other cells.
-`r length(unique( all.remove ) )` measures removed as potential outliers (by examining the category specific plots)<pattern for the single bivalent, pos correlation, The predicted pattern for SC length becomes more nuanced when the data are split up by Chromosome class.
-Mainly that 1CO are shorter in the high recombining strains then low recombining strains.
-This is likely due to the fact that more physically longer chromosomes have 1CO in the low recombining strains, which pushes the mean SC length up.
-Where as in the high recombining strains, physically longer chromosomes are more likely to be in the 2CO group.
-This supports a general pattern of tighter clustering of SC lengths across chromosome classes in the high recombininggroup.
-They have a lower probability of 2COs below a certain SC length threshold.>  -->
 
@@
 
-<!--
-In the short bivalent data set, analyzed by M1, both subspecies are 
- (glm; p= _Musculus_ `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and _Molossinus_ p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`).
-While the models for the reduced data set suggest that the subspecies with the high recombining strains have evolved longer SC lengths --
-long bivalent set in M1 -- 
-Musc `r sum.glmQ2_long.biv_M1$coefficients[,4][2]` 
-MSM strain `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`
-<And the strain effect for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).>
-
-**M3 -- tests within subspecies!!** SO the M3 model can be run within species specific data -- to test the predictions (based on the variation in mean MLH1 counts), (use all 3 metircs) **the within subspecies M3 models don't fully meet the predictions
-for total.Sc : there are significant Dom strain effects: in musc kaz is significant negative, several other strains have higher total SC, (MOL in the intercept is significant).
-short biv: Dom: Significant G effect, MUSC no significant -- strain effects
-long.biv:  Dom: no strain effects, MUSC no significant strain effects, MOL - not enough observations
-**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`).
-The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**
-
-**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**
-
-**THERE IS a weak signal of the high recombiningstrains having longer SC (having evolved) longer SC (different chromatin compaction)  While the mouse mean total sc are significantly different between the two groups --- this difference --- translated to the reduced bivalent data sets (p = long  p = short)**   -->
-
 ```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
@@
 
+The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = 0.24) and also when examined within subspecies (t-test; p = 0.41 and p = 0.07 for musculus and molossinus, respectively).
+While there are significant strain effects for domesticusWSB and molossinusMOLF in a general linear model with normalized position as the dependent variable (Figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
+
 ```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
@@
 
-<!--
-The high rec musc strains, _musculus^PWD^_ and _musculus^SKIVE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).
-While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`).
-is that _domesticus^WSB^_ has significantly more terminal positioning,
-and _molossinus^MOLF^_ having significantly more medial position.
-A general pattern of strain specific effects
-we found in looking at the 1CO positions, is that _domesticus^WSB^_ has significantly more terminal positioning, while MOLF has significantly more medial position.-->
-
-<!-- don't report GLMs - since not difference in the ttests, GLMs because there's no difference.
-long p = `r Rec.group_LONG.BivData_F1post_Ttest$p.value` short 1CO, male, p = `r Q2.1COpos.PER_short_ttest_HL$p.value`>
-toal data set, short and long biv. There's only a difference between MSM and MOKF
-test ppredictions, Dom = NULL,  -->
-
 ```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
@@
 
-<!--
-given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models.
-Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome-wide recombination rate variation
-<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>
-
-In the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower.
-..
-the chromosome length has the most significant effect, LEW by chrm effect are significant
-
-In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects -->
-
-<!-- Just hold these for reference
-**For the M2 (subsp)**
-There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).
-
-LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).  
-But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).  
-and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)  
-**For M3 (all strains)**
-strain G (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][2]`).
-strain Lew (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][3]`).
-strain PWD ( p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][4]`).
-slight strain MSM  (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][5]`).
-very sig strain MOLF (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][6]`).
-strain SKIVE (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][7]`).
-strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
--->
-
 ```{r Q2.1COpos.show.MOL, echo=FALSE}
@@
 ```
-
-<!-- hide these linear model results for now
-for the first model, 
-chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`
-
-subsp MOL `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][4]`  
-the mol obj above isn't showing the reight value. 
-strain G   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][5]`  
-strain LEW   `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][6]`  
-Strain MSM `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][8]`  
-chrm * LEW `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][14]`  
-chrm *MSM  `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][16]`
-
-M2
--chromosome length, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][2]`  
--strain G, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][3]`  
--strain LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][4]`  
--strain MOLF, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][7]`  
--strain SKIVE, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][8]`  
--chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]` -->
 
@@
 # Q2 Evolution of interference is associated with genome-wide recombination rate evolution
+
+Mouse averages for both IFDraw and IFDnorm are significantly longer in high-recombination strains (t-test; IFDnorm: p = 7.7410^{-7}; IFDraw: p = 8.7810^{-6}).
+This pattern is confirmed by separate comparisons within musculus (t-test; IFDnorm: p =2.0410^{-5}; IFDraw: p = 1.9410^{-4}) and within molossinus (IFDnorm: p= 0.17; IFDraw: p = 0.08).
+Similar results are seen with general linear models for both IFDraw and IFDnorm: only effects associated with high-recombination strains are significant (p < 0.05).
+That IFDraw and IFDnorm show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
+We determined that the main difference in IFDnorm distributions between high-recombination and low-recombination strains is an enrichment of IFDnorm values under 30% in low-recombination strains.
+The frequency of IFDnorm values that fall below 30% ranges from 8.2% (domesticusG) to 16% (musculusKAZ) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculusSKIVE, molossinusMSM, and musculusPWD, respectively).
+In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
+However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
+At least on 2CO bivalents, crossovers are spaced further apart when the genome-wide recombination rate is higher.
 
@@
 
-<!--
-**Linear model M1 ** $$mouse \ average \ IFD ~=~ subsp * strain + \varepsilon $$
-
-**Linear model M2 ** $$mouse \ average \ IFD ~=~ strain + \varepsilon $$ -->
-
 ```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
@@
 
-<!--
-RAW IFD M1 results for M1  -- `r sum.Q2.M1glm_IFD.raw$coefficients[,4][6]` PWD, 
-MSM `r sum.Q2.M1glm_IFD.raw$coefficients[,4][7]`
-SKIVE `r sum.Q2.M1glm_IFD.raw$coefficients[,4][8]`
-and M2
-for PWD, `r Sum.Q2.M2glm_IFD.raw$coefficients[,4][4]`
-MSM `r Sum.Q2.M2glm_IFD.raw$coefficients[,4][5]`
-and SKIVE   `r Sum.Q2.M2glm_IFD.raw$coefficients[,4][7]`
-
-In M1, the model build on the mouse averages of IFD^norm^ metrics with subspecies and strain as fixed effects, the high recombining strains had significant strain effects for (glm; p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][6]`, p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][7]` and p = `r sum.Q2.M1glm_IFD.PER$coefficients[,4][8]` for _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, respectively).
-In the M2 model results, _musculus^PWD^_, and _musculus^SKIVE^_, were the only significant strain effects, (glm; p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][4]` and p = `r sum.Q2.M2glm_IFD.PER$coefficients[,4][7]`).
-Running M2 within subspecies confirms the within subepcies predictions across strains – (no significant strain effects 
-Q2.glm.M2_IFD.PER_Dom
-Q2.glm.M2_IFD.PER_Musc
-Q2.glm.M2_IFD.PER_Mol -->
-<!--
-
-<Previous prediction for differences in the long and short bivalent dataset.
-Examine the pattern in the Long bivalent data set to get a feel for if the chromosome size effect skews the pattern.
-In the long bivalent data set, the longest bivalents (top 25%) from each cell are isolated, so they are more likely to be the same chromosome identities (Chrm1 to Chrm5).
-The section on IFD only looks at the 2CO bivalents.>
-
-For the first version of the bivalent level model (which has subspecies effect), the most significant effect is chromosome length, is slight positive direction (both for raw and normalized IFDs).
-Molossinus subspecies and Mol\*chrm length effect for IFD^Normalized^ and IFD^raw^.
-The Mol\*chrmLength is negative
-<PWD, MSM, MOLF, SKIVE are the only significant strain effects.
-- all of the strin\*chrm lengths effects are negative, PWD, MSM, MOLF, SKIVE significant effects (for IFD.ABS and IFD.PER)
-With each unit increase in SC length, the IFD^norm^ also increases (for the high rec strains + MOLF).
-All of the high rec strains and MOLF have significantly longer IFD^norm^.
-There are also significant interaction effect of strain and chromsome length, in the high rec strains and MOLF, with each increase in bivalent length, the IFD^norm^ decreases slightly.
--->
-
-<This striking differences across sexes suggest that the genome-wide recombination rates have distinct evolutionary trajectories in males and females of the same species.
-The pattern of strain averages distributed around the total sample mean in females is supportive of stabilizing selection or neutral evolution.
-In contrast to the pattern from male means which have rapid evolution in a subset of genetic backgrounds this fits a model of directional selection.
-(We attempt to make better evolutionary models by applying an evolutionary framework)>
-
 # Q2 Summary
@@
 
+<Bret: Are you sure this summary is necessary? Maybe cover this material in the Discussion instead?>
+
 # References

### 2020-04-07 doc/Results_v2.rmd
result work

@@
 To situate variation in recombination rate within an evolutionary framework, we
-We applied fit a series of models including subspecies, strain, and sex, to to our dataset of mean MLH1 foci counts from 187 mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
-We began with a full mixed model (M1, see Methods), which showed that strain (random effect p < someSmallNumber), sex (p = 3.9710^{-5}), subspecies (p= 1.7210^{-4}), and subspecies * sex (p = 3.110^{-5}) each significantly affect recombination rate.
+We applied fit a series of models including subspecies, strain, and sex, to to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
+We began with a full mixed model (M1, see Methods), which showed that strain
+(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results` -- 3.9710^{-5}), subspecies (p=`r MLH1.MMlmer_results$interaction.pval`--  1.7210^{-4}), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval` -- 3.110^{-5}) each significantly affect recombination rate.
+
+<for values> The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome-wide recombination rates.
+The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
+m.
+musculus_ and _M.
+m.
+molossinus_ respectively), indicating that most of the variance is due to strain effects.
+
+As such, we focus on results from M3.
+The two most significant strain effects were _domesticus^G^_ and _musculus^MSM^_ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]` for _domesticus^G^_ and _molossinus^MSM^_ respectively).
+In addition to significant strain effects for _molossinus^MSM^_, (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
+and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.
+Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in _domesticus^G^_.
+
 After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculusMSM (p= 3.9910^{-6}) and domesticusG (p= 1.0410^{-6}).
-In addition, two strains exhibit strain-by-sex interactions:  molossinusMSM  (p = 1.2610^{-4}) and musculusPWD (p = 3.8610^{-4}).
+A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p= 3.9910^{-6} p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
+and domesticus^G^ (p= 1.0410^{-6}  -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
+In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
+
 We next fit general linear models separately for 192 males and 144 females (M4).
-In the male dataset, three strains significantly affect recombination rate: molossinusMSM (p = 2.2310^{-14}; effect = X MLH1 foci), All of the effects within the mixed model M1 were highly significant (LRT: subsp p= 1.7210^{-4}, sex p = 3.9710^{-5}, subspecies * sex p = 3.110^{-5}) and random LTR; strain p= 0 ).
-M1. Mixed Model 
-mouse av. MLH1 foci per cell = subsp*sex+rand(strain)+??
-M2. Linear Model
-mouse av. MLH1 foci per cell = subsp*sex*strain+??
-M3. Linear Model
-mouse av. MLH1 foci per cell = sex*strain+??
-M4. Linear Model
-sex specific mouse av. MLH1 foci per cell = subsp*strain+??
+In the male dataset, three strains significantly affect recombination rate: molossinus^MSM^ (p = 2.2310^{-14}; effect = X MLH1 foci), All of the effects within the mixed model M1 were highly significant (LRT: subsp p= 1.7210^{-4}, sex p = 3.9710^{-5}, subspecies * sex p = 3.110^{-5}) and random LTR; strain p= 0 ).
+
+**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
+
+**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
+
+**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
 
 The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombination rates.
-The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=0.22 and p= 0.09 for M.
+The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=0.22 and p= 0.09 for *M.
 m.
-musculus and M.
+musculus* and *M.
 m.
-molossinus respectively), indicating that most of the variance is due to strain effects.
+molossinus* respectively), indicating that most of the variance is due to strain effects.
 As such, we focus on results from M3.
-The two most significant strain effects were domesticusG and musculusMSM (p= 1.0410^{-6} and p= 3.9910^{-6} for domesticusG and molossinusMSM respectively).
-In addition to significant strain effects for molossinusMSM, (glm; p = 0.06 and p = 3.9910^{-6}) both musculusPWD and musculusMSM had significant strain by sex effects (glm; p = 3.8610^{-4} and p = 1.2610^{-4}) musculusPWD by male and musculusMSM by male.
+The two most significant strain effects were domesticusG and musculus^MSM^ (p= 1.0410^{-6} and p= 3.9910^{-6} for domesticusG and molossinusMSM respectively).
+In addition to significant strain effects for molossinus^MSM^, (glm; p = 0.06 and p = 3.9910^{-6}) both musculusPWD and musculus^MSM^ had significant strain by sex effects (glm; p = 3.8610^{-4} and p = 1.2610^{-4}) musculus^PWD^ by male and musculusMSM by male.
 Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in domesticusG.
-A linear model was run on the sex specific data sets (M4), with 192 male and 144 female mean MLH1 foci per cell.
-There were three significant strain effects in the male specific data; musculusPWD ((glm; p = 7.3710^{-10}; effect = X foci), and musculusSKIVE (glm; p = 0.01; effect = X foci).
+
+A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` -- 192 male and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 female mean MLH1 foci per cell.
+There were three significant strain effects in the male specific data; musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = ~ 5 X foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect ~ 2 foci).
+
 These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate:  and molossinusMSM (glm; p=2.2310^{-14} with effect sizes ranging from 5, 7, and 2 foci respectively.
+Analysis of the female dataset points to four strains with significant effects on recombination rate:  and molossinus^MSM^ (glm; p=2.2310^{-14} with effect sizes ranging from 5, 7, and 2 foci respectively.
 Given these results, theses strains are classified as 'high recombining' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
-While the female specific linear model has many more had significant strain effects; domesticusG (glm; p = 2.510^{-6}), molossinusMSM (p = 6.2410^{-6}), domesticusLEW (glm; p = 0.01), and domesticusPWD musculusPWD (glm; p=0.02).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males., domesticusMSM (glm; p=6.2410^{-6}), domesticusMOLF(glm; p=0.08), and domesticusKAZ(glm; p=0.1), the effect sizes have a smaller range of (4 to 1), indicating significant but small strain specific effects on mean MLH1 foci per cell.
+While the female specific linear model has many more had significant strain effects; domesticus^G^ (glm; p = 2.510^{-6}), molossinus^MSM^ (p = 6.2410^{-6}), domesticus^LEW^ (glm; p = 0.01), and domesticusPWD musculusPWD (glm; p=0.02).
+Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males., domesticus^MSM^ (glm; p=6.2410^{-6}), domesticus^MOLF^(glm; p=0.08), and domesticus^KAZ^ (glm; p=0.1), the effect sizes have a smaller range of (4 to 1), indicating significant but small strain specific effects on mean MLH1 foci per cell.
 Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
+
+<bret above>
 
@@
 
- <to better understand this varation in RR within an evolutionary framework. Make table of models>
-
 We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
 All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
-
-**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
-
-**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
-
- <These results allow use to reject the hypothesis of genome-wide recombination rates (varying consistantly by sex across subepcies -- and having variance within subspecies due to strains)>
-
- <!-- (random strain effect was meant to reflect sampling of gentic backgrounds/standing genetic variation for a subspecies) -- (significant pvalue indicates that the variance is not 0)>
- indicating, fixed effects -- random effect variance due to each strain nested within subspecies
- <The linear models, allow us to quantifying the significance in sepecific fixed effects and focus on how specific levels and interactions create these differences.
- -->
 

### 2020-04-07 doc/Results_v2.rmd
updating values

@@
 #Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
-Fig1_v1_strain / (female.specific.plot + male.plot )
+#Fig1_v1_strain / (female.specific.plot + male.plot )
 
 #(p1 | p2 | p3)/p4
-
 ```
@@
 
-To situate variation in recombination rate within an evolutionary framework, we
-We applied fit a series of models including subspecies, strain, and sex, to to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
+```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}
+
+#DF.HetC.MixedModel.HQ
+
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1.MM.M1)$coefficients
+
+MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+
+MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
+sex.results =  MLH1.MM_sex, 
+               interaction.pval = MLH1.MM_interaction,  rand.pval = 
+MLH1.MM_strain)
+
+```
+
+```{r MLH1.glmM2, echo=FALSE}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
+#strain = fixed
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
+
+summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
+
+#table of just coeff anf pvalue
+#round(summ$coefficients[,c(1,4)],5 )
+
+#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
+```
+
+```{r MLH1.glm.M3, echo=FALSE}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
+
+summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
+
+#round(summ.strain.fixed$coefficients[,c(1,4)],5)
+
+#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+```
+
+To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
+
 We began with a full mixed model (M1, see Methods), which showed that strain 
-(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results` -- 3.9710^{-5}), subspecies (p=`r MLH1.MMlmer_results$interaction.pval`--  1.7210^{-4}), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval` -- 3.110^{-5}) each significantly affect recombination rate.
-
-<for values> The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome-wide recombination rates.
-The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
-m.
-musculus_ and _M.
-m.
-molossinus_ respectively), indicating that most of the variance is due to strain effects.
-
-As such, we focus on results from M3.
-The two most significant strain effects were _domesticus^G^_ and _musculus^MSM^_ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]` for _domesticus^G^_ and _molossinus^MSM^_ respectively).
-In addition to significant strain effects for _molossinus^MSM^_, (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
-and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.
-Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in _domesticus^G^_.
+(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results` -- 3.9710^{-5}), subspecies (p=`r MLH1.MMlmer_results$sub.sp`--  1.7210^{-4}), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval` -- 3.110^{-5}) each significantly affect recombination rate.
 
@@
 
-We next fit general linear models separately for 192 males and 144 females (M4).
-In the male dataset, three strains significantly affect recombination rate: molossinus^MSM^ (p = 2.2310^{-14}; effect = X MLH1 foci), All of the effects within the mixed model M1 were highly significant (LRT: subsp p= 1.7210^{-4}, sex p = 3.9710^{-5}, subspecies * sex p = 3.110^{-5}) and random LTR; strain p= 0 ).
-
-**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
-
-**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
-
-The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome wide recombination rates.
-The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=0.22 and p= 0.09 for *M.
-m.
-musculus* and *M.
-m.
-molossinus* respectively), indicating that most of the variance is due to strain effects.
-As such, we focus on results from M3.
-The two most significant strain effects were domesticusG and musculus^MSM^ (p= 1.0410^{-6} and p= 3.9910^{-6} for domesticusG and molossinusMSM respectively).
-In addition to significant strain effects for molossinus^MSM^, (glm; p = 0.06 and p = 3.9910^{-6}) both musculusPWD and musculus^MSM^ had significant strain by sex effects (glm; p = 3.8610^{-4} and p = 1.2610^{-4}) musculus^PWD^ by male and musculusMSM by male.
-Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in domesticusG.
-
-A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` -- 192 male and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 female mean MLH1 foci per cell.
-There were three significant strain effects in the male specific data; musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = ~ 5 X foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect ~ 2 foci).
-
+We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4).
+In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = ~ 5 X foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect ~ 2 foci).
 These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate:  and molossinus^MSM^ (glm; p=2.2310^{-14} with effect sizes ranging from 5, 7, and 2 foci respectively.
-Given these results, theses strains are classified as 'high recombining' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
-While the female specific linear model has many more had significant strain effects; domesticus^G^ (glm; p = 2.510^{-6}), molossinus^MSM^ (p = 6.2410^{-6}), domesticus^LEW^ (glm; p = 0.01), and domesticusPWD musculusPWD (glm; p=0.02).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males., domesticus^MSM^ (glm; p=6.2410^{-6}), domesticus^MOLF^(glm; p=0.08), and domesticus^KAZ^ (glm; p=0.1), the effect sizes have a smaller range of (4 to 1), indicating significant but small strain specific effects on mean MLH1 foci per cell.
+Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p=0.02  -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
+Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
+
 Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
@@
 <bret above>
-
-```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}
-
-#DF.HetC.MixedModel.HQ
-
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
-
-MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
-sex.results =  MLH1.MM_sex, 
-               interaction.pval = MLH1.MM_interaction,  rand.pval = 
-MLH1.MM_strain)
-
-```
-
-We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice to fit the patterns of variation across sex, subspecies and strain into an evolutionary framework.
-All of the effects within the mixed model M1 were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).
-
-```{r MLH1.glmM2, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
-
-#strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
-
-summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
-
-#table of just coeff anf pvalue
-#round(summ$coefficients[,c(1,4)],5 )
-
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-```
-
-```{r MLH1.glm.M3, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
-
-summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
-
-#round(summ.strain.fixed$coefficients[,c(1,4)],5)
-
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-```
-
-The linear model results confirm the qualitative observations of sex-specific evolution and significant difference in the degree of sexual dimorphism in genome-wide recombination rates.
-The subspecies fixed effects tested in M2, were either weak or not significant (glm; p=`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` and p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` for _M.
-m.
-musculus_ and _M.
-m.
-molossinus_ respectively), indicating that most of the variance is due to strain effects.
-As such, we focus on results from M3.
-The two most significant strain effects were _domesticus^G^_ and _musculus^MSM^_ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]` for _domesticus^G^_ and _molossinus^MSM^_ respectively).
-In addition to significant strain effects for _molossinus^MSM^_, (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][5]` and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)  both _musculus^PWD^_ and _musculus^MSM^_ had significant strain by sex effects (glm; p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`
-and p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) _musculus^PWD^_ by male and _musculus^MSM^_ by male.
-Combined, these results demonstrate that i) the significant sex-specific evolution in MSM and PWD and ii) the larger magnitude of heterochiasmy in _domesticus^G^_.
-
-A linear model was run on the sex specific data sets (M4), with `r length(fomo.male_mouse.av$mouse)` male and `r length(female_mouse.avs_for.analysis$mouse)` female mean MLH1 foci per cell.
-There were three significant strain effects in the male specific data; _musculus^PWD^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`), _musculus^SKIVE^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]`) and _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]` with effect sizes ranging from 5, 7, and 2 foci respectively.
-Given these results, theses strains are classified as 'high recombining' strains and point to rapid evolution in the recombiantion rate for spermatocytes.
-
-While the female specific linear model has many more had significant strain effects; _domesticus^G^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _domesticus^LEW^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), _domesticus^PWD^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`), _domesticus^MSM^_ (glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^MOLF^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[6,][4]`), and _domesticus^KAZ^_(glm; p=`r sum_MLH1.female.spc.glm.M2$coefficients[8,][4]`), the effect sizes have a smaller range of (4 to 1), indicating significant but small strain specific effects on mean MLH1 foci per cell.
-
-<!--
-
-<domesticus^G^ had significant strain (p= `r summ.MLH1.M3.strain.fixed$coefficients[4,][3]`)
-and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][10]`>
-
-M2(glm; PWD*male p= `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][14]` and MSM-male p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][15]`)>
-<Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.MLH1.M3.strain.fixed$coefficients[,4][3]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect).
-These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.
-
-In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.
-(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )
-Check the DF of the number cited above
-This is the break down of mice used within the Mixed model
-`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`
-add number of cells
-`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`
- $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$
-
-- only Mus musculus strains with sex matched observations  
-- Male gwRR have been adjusted for the PAR
-- Q5 excluded
-- All the effects are un-ordered factors  
-- male ages within 5 - 20 weeks
-- Molossinus is coded as a subsp level
-also g strain effect/ dom^G^ strain effect 
-`r summ.MLH1.M2.glm.strain.fixed.full$coefficients[12,][4]` 
-In both models strain by sex interaction effects were significant for (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` and `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively.
--->
 

### 2020-04-08 doc/Results_v2.rmd
removed non-text, accepted all changes

@@
 AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
-
 #total mice
-length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )
+#length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )
 
 #average cell number for spermatocytes and oocytes
-
 #mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
-
 #mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
-
 ```
@@
 molossinus*.
-Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )` mice were quantified from an average of `r mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes) and `mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes).
+Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )` mice were quantified from an average of `r mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes).
 
@@
 To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
-
 We began with a full mixed model (M1, see Methods), which showed that strain
@@
 ## Within mouse variance in CO count per cell
-
-Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
-To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber) and general linear model (M2) (p = 2.310^{-4}).
-In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (criteria?).
-The results are similar: sex is the strongest effect (p < smallNumber, M1; p = 2.310^{-4}, M2).
-When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
-These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
 
@@
 
-We examine the within animal variance in MLH1 foci count per cell using the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
-
-The linear models support the general pattern of females having almost twice as much variance in MLH1 foci per cell compared to males (Figure 1)).
-For both the mixed and linear models, sex was the only significant effect (LRT; p= `r round(LmerMLH1_VAR_results$sex.results, 5)`, glm; p= `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`).
-Since the measures for within mouse variance may be more susceptible to technical error effects from the staining protocol (i.e.
-background noise), we replicated the model analysis using a subset of cells with higher quality scores.
-These models replicated the results of the full data set with sex being the most significant effect (LTR; p = `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, and glm; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)` ).
-
-There is no significant difference in the amount of within mouse variance of MLH1 foci counts per cell in males according to the linear models.
-In females, the significance of strain effects was not consistent between the full data set and higher quality dataset _domesticus^LEW^_ (glm; p = 0) was a significant effect in the full model and _musculus^PWD^_ (glm; p =0.04) in the higher quality dataset.
-
-<one strain, _domesticus^LEW^_ has significantly higher within mouse variance (glm; p = `r lmVAR.female.strain_sum$coefficients[3,][4]`).
-However, in the model using the high quality cells, _musculus^PWD^_ has the only signicant strain effect (glm; p =`r lm.m2.Q12VAR.female_sum$coefficients[4,][4]`) suggesting a lack of a consistant signal for differences in the amount of within mouse variance in oocytes.>
-
-<there are some caveats, but the general pattern is that females have more variance ...
-For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell.
-This limits the power for detecting significance differences across groups,
-the main/general pattern, is that females have significantly higher within mouse variance compared to males.
-This general pattern is replicated in higher quality dataset and using the coefficient of variance (cv) as a dependant variable.>
-
-<!--
-
-references for phenotypic variance connected to robustness to environment, increase in developmental consistency
-
-Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
-phenotypic variance leading to adaptation - 
-(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
-Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
-meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
--->
-
-<SUMMERIZE BOTH THE MLH1 evolution and within mouse variance HERE>
-<The greater amount of within mouse variance in females (almost and, (males have no detecable difference across strains for within mouse variance)
--- supports the distinct evolutionary pressures acting on the genome-wide recombination rate in oocytes and spermatocytes.
-The striking patterns of variation across male and females suggest that the genome-wide recombination rates have distinct evolutionary trajectories.
-For female the pattern of Strain averages distributed around a species wide average fits a model of stabilizing or relaxed evolution in contrast to the male pattern where there is rapid evolution in a subset of genetic strains that is close to a model of directional selection on genome-wide recombination rate.>
-
-# Evolution of genome-wide recombination rate associated with evolution of mean DSB number
-
-Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
-
-In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of 76 leptotene-stage and 75 zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculusKAZ , domesticusWSB , and domesticusG) and two high-recombination strains (musculusPWD and molossinusMSM).
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<someSmallNumber; one-way ANOVA, p = 0.00027).
-In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66; one-way-anova, p = 0.15).
-After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = 0.94 and p = 0.11 for leptotene and zygotene ratios, respectively).
-This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
+Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
+To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
+Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`).
+In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
+Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (**criteria?**).
+The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2).
+When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
+These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
+
+# Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
 
@@
 
-In an attempt to localize  the above male specific crossover number evolution within the meiotic pathway we quantified a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene staged spermatocytes from juvenile mice (12 to 18 days) of three low rec, ( _musculus^KAZ^_ ,  _domesticus^WSB^_ , and _domesticus^G^_) and two high recombining strains ( _musculus^PWD^_ and _molossinus^MSM^_).
-
-<All strains had significantly more DMC1 foci from leptotene cells compared to zygotene (t.test; p = `r ttest.ZtoL$p.value`), which reflects the expected patterns across the cell stages in prophase.>
-
-The mean number of DMC1 foci per cell is associated with evolution of the number of precursors (DSBs).
-The two high recombining strains tested have significantly more DMC1 foci compared to low recombining strains in leptotene cells (t.test; p=`r ttest.HighLow.L.pval`, one-way-anova; p=`r format(l.anova.pval, scientific=F)`).
-However the differences in DMC1 foci were not significant for zygotene cells (later prophase) between the two groups (t.test; p=`r ttest.HighLow.Z.pval`, one-way-anova; p=`r z.anova.pval`).
-Indicating that DMC1 counts at this stage is more predictive of the downstream crossover number differences.
-
- After DSB formation, DSBs are repaired as either non-crossover (NCO) or crossovers (COs).
-with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated into COs, partially known as the NCO:CO decision.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios respectively).
-
-<!--
-The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.
-The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.
-`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M.
-m.
-domesticus^WSB^_ , _M.
-m.
-domesticus^G^_, _M.
-m.
-molossinus^MSM^_, and  _M.
-m.
-musculus^PWD^_ respecitively.
--->
+In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` -- 76 leptotene-stage and 75 --`r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
+
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`; one-way ANOVA, p = 0.00027 -- `r format(l.anova.pval, scientific=F)`).
+In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`; one-way-anova, p = 0.15 -- `r z.anova.pval`).
+After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
+Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
+The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = 0.94 -- `r DMC1.Lratio_t.test$p.value` and p = 0.11 -- `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios, respectively).
+This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
 
@@
 
-# Single Bivalent Level Results
-
-Evolution of genome-wide recombination rate is reflected at the single chromosome level
-
-To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents (Peterson et al.
-2019).
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from 0.51 (molossinusMSM male) to 0.72 (musculusKAZ female).
-From the total set of cell images, 10,458 bivalent objects were isolated by the image analysis software.
-After a human curation step (Peterson et al.
-2019), 9,829 single-bivalent observations remained.
-We assume that the isolation of bivalents within cells is unbiased.
-Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns.
-An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
-
-<figure>
-
-Ninety-six percent of single bivalents in our pooled dataset (n = 34,982) have either one or two crossovers (Figure X).
-The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
-low recombining strains (Figure X).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents  are 0.33 in musculusSKIVE , 0.44 in musculusPWD , and 0.53 in molossinusMSM .
-Following patterns in the genome-wide recombination rate, male musculusPWD and male molossinusMSM have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; musculusPWD vs.
-musculusKAZ p = 3.1510^{-33}; molossinusMSM vs.
-molossinusMOLF p = 4.7210^{-13}).
-Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
-We address two main questions.
-First, which traits are sexually dimorphic?
-Second, which traits differ between males from high-recombination vs.
-low-recombination strains?
+# Evolution of genome-wide recombination rate is reflected at the single chromosome level
 
@@
 ```
-
-For each strain, we deconstructed the estimates of genome-wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline.
-The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures [@peterson_2019].
-
-A limitation of our image analysis algorithm is that not all bivalents per cell are isolated, due to overlapping  chromosomes.
-The range isolation rates per cell in this data set is `r min(algorithm.stat.category$mean_pass.rate)` in _molossinus^MSM^_ male to `r max(algorithm.stat.category$mean_pass.rate)`  _musculus^KAZ^_ female.
-However, we assume that the isolation of chromsomes within cell images is not biased.
-From our total set of cell images `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` chromosome objects were isolated by the image analysis software.
-After the human curation step, `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single bivalent observations remained.
-Given the large number of single chromosome observations, we assume that each of the data sets are equally representative of general patterns.
-
-A limitation of these data from cytological stains are the only chromosome which can be accurately identified is the XY in males.
-The unpaired XY chromosomes have SYCP3 signal along the chromosome axis, but it forms a structure distinct from the synapsed autosomes, which can be excluded in the curation step.
-Since it has homologous pair in oocytes, the XX bivalent is indistinguishable from the rest of the autosomes.
-By physical length in Mb, the X is predicted to be the 3rd longest bivalent.
-
-# Evolution of genome-wide recombination rate is reflected at the single chromosome level
 
@@
 
-<from the means the expected number of COs per chroosomes ranges from X to X we confirm these numbers Can't have fractions of COs>
-
-We distill the variance observed in the mean MLH1 foci counts per cell to number of MLH1 foci per bivalent.
-Because our observations of MLH1 foci are made of pachytene cells, our chromosome observations are fully synapsed paired homologous chromosomes (4n copies of each chromosomes).
-Additionally we can not fully distinguish the identity of chromosomes (e.g.
-Chromosome 1 or Chromosome 2), for these reasons we use the term 'bivalent' instead of chromosome.
-
-Ninety six percent of the pooled dataset of single bivalents (n=`r total.biv.all`) have either one or two crossovers (Figure X).
-The proportion of 1CO:2CO distinguishes the high and low recombining strains (Figure X).
-This confirms the intuitive interpretation that high recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents.
-In the high recombining strains the 2CO proportions are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ), `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` ( _musculus^PWD^_ ), and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` (_molossinus^MSM^_ ).
-
-While the proportions are significantly different (chi-square-test; p = `r PWD.MSM.2CO.prop.test$p.value`), males _musculus^PWD^_ and _molossinus^MSM^_ have 2CO proportions more similar to each other than strains from the same subspecies ( _musculus^PWD^_ and _musculus^KAZ^_ chi-square test; 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`) and _molossinus^MSM^_ and _molossinus^MOLF^_ chi-square test; 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).
-Using this data set of single bivalents, we focus on aspects of the recombination landscape to address two main questions 1) Which traits are sexually dimorphic?
-and 2) Which traits fit distinguish the high and low recombining strains for males?
-
-In our analysis our first step is to test for measurable differences, between sexes or recombination groups in males and if the differences are significant, our second step is to apply the same sets of models applied to the mean MLH1 foci counts to test for effects due to subspecies or strains.
-
-# Q1 Sex Differences in SC-AE Lengths
-
-Q1 Sex Differences in the Length of the Synaptonemal Complex
-
-In many mammalian species, the synaptonemal complex (SC) is longer in females.
-Yet, the majority of these observations come from species with female-biased heterochiasmy.
-Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male musculusPWD and male molossinusMSM should have longer SC lengths than the other strains.
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
-This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
-A total of 678 'short' bivalents were isolated from 103 oocytes and 37 spermatocytes.
-Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents ([@peterson2019]).
+This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
+In this dataset, isolation rates per cell range from
+`r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
+From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
+After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.
+We assume that the isolation of bivalents within cells is unbiased.
+Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns.
+An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
 
@@
 
-Short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculusSKIVE.
-The female:male ratio ranges across strains from 1.15 (musculusMSM) to 1.49 (domesticusWSB).
-As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of Wang et al.
-(2019).
-Mouse means were calculated from cell-wide total SC lengths in 2,984 out of 3,680 cells with MLH1 counts (Figure X).
-Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
-That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies.
-
-<figure>
-
-In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers.
+Ninety-six percent of single bivalents in our pooled dataset (n = 34,982 -- `r total.biv.all`) have either one or two crossovers (Figure X).
+The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
+low recombining strains (Figure X).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in molossinus^MSM^.
+
+Following patterns in the genome-wide recombination rate, male musculus^PWD^ and male molossinus^MSM^ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; musculus^PWD^ vs.
+musculus^KAZ^ p = 3.1510^{-33} -- `r PWD.KAZ.2CO.prop.test$p.value`; molossinus^MSM^ vs.
+molossinus^MOLF^ p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
+
+Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
+We address two main questions.
+First, which traits are sexually dimorphic?
+Second, which traits differ between males from high-recombination vs.
+low-recombination strains?
+
+# Q1 Sex Differences in the Length of the Synaptonemal Complex
 
@@
 ```
-
-In many mammalian species the SC lengths are longer in females, yet the majority of these observations come from species with female biased heterochiasmy.
-Our sample of house mice provides an opportunity to test if the sex differences in chromatin compaction (SC length) are reversed in the high recombining strains with male-biased heterochiasmy; _musculus^PWD^_  and _molossinus^MSM^_).
-If SC length is the strongest predictor for number of MLH1 foci per cell, males of the high recombining strains would have longer SC length leading to the difference in total MLH1 foci per cell.
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we use a reduced bivalent data set made up of bivalents with SC lengths below the 2nd quartile for SC length from a single cell to compare across sexes.
-This results in a dataset of the shortest 4 to 5 chromosomes and excludes the X bivalent in the oocytes.
-A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Even though this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between sexes.
 
@@
 
-All but one strains have significantly longer mouse mean for short bivalents in females (t.test; p <0.05).
-The differences, (F:M), across strains ranges from `r Q1SC.mus.av.short.biv_DIFF_MSM` in _musculus^MSM^_ to `r Q1SC.mus.av.short.biv_DIFF_WSB` in _domesticus^WSB^_.
-While the female value is longer, it is not significantly different from males in _musculus^SKIVE^_ (t.test; p = `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)`).
-
-<!-- t.test; p = `r Q1_mus.av_shortSC_t.test.WSB$p.value`, p =`r format(Q1_mus.av_shortSC_t.test.G$p.value, scientific = F)`, and p=`r format(Q1_mus.av_shortSC_t.test.LEW$p.value, scientific = F)` for _domesticus^WSB^_, _domesticus^G^_ and _domesticus^LEW^_ respectively.
-the musculus strins (t.test; p= `r format(Q1_mus.av_shortSC_t.test.PWD$p.value, scientific = F)`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r  format(Q1_mus.av_shortSC_t.test.SKIVE$p.value, scientific = F)` for _musculus^PWD^_ , _musculus^KAZ^_ , _musculus^SKIVE^_  respectively -->
-
-For an additional  metric of chromatin compaction, we tested a whole cell measure of SC length using a second image analysis algorithm which converts the SC signal into single pixel wide trace of the SC shapes and provides accurate estimates of the summed SC length of bivalents for single cells [@wang_2019_sc].
-We apply the skeletonizing algorithm to all MLH1 meiocyte images.
-Mouse means were calculated from `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cell wide total SC measures (figure X).
-(where as `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` number of cells had MLH1 measures ).
-
-<The comparison of this metric is somewhat complicated across sexes due to the sex chromosomes.
-Breifly/in effect the oocytes have an additional autosome, and the spermatocytes have a mostly recombinationly inactive unit of SC (the XY).>
-
 ```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
@@
 ```
-
-<these are the number of mice for each of the categories r as.data.frame( table(droplevels( totSC.mouse.av.table$category) ) )>
 
@@
 
-In all of the strains the means of total SC per cell was significantly greater in females than males (ttest; p < 0.05).
-To test for strain and subspecies specific effects on the sex difference in SC lengths we apply the model framework.
-
-<!-- _domesticues^WSB^_ p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, _domesticues^G^_  p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, _domesticues^LEW^_ p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`, _musculus^PWD^_ p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, _musculus^KAZ^_ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, _molossinus^MSM^_ `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` -->
-
 ```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
@@
 
-<!--
-**M1. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$
-
-**M3. Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$
-
-In the mixed models for the sex differences mean **short bivalents** length as the dependent variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`).
-The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).
-
-In the mixed model run on the mean total SC metric all of the coefficients were significant, (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).
--->
+In many mammalian species, the synaptonemal complex (SC) is longer in females.
+Yet, the majority of these observations come from species with female-biased heterochiasmy.
+Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
+In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male musculus^PWD^ and male molossinus^MSM^ should have longer SC lengths than the other strains.
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
+This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
+A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+
+<figure>
+
+Short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculus^SKIVE^.
+The female:male ratio ranges across strains from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
+
+As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
+
+Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
+Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
+That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
+Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies.
+
+<figure>
+
+In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
+Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers.
 
@@
 
-<Q12 total SC mouse means!
-In Dom when running the Q12 SC lengths, the sex difference seems to get larger (male mean total SC gets lower).
-In Musc the female mean total sc gets lower (maybe fewer cells (SKIVE is removed)).
-For Mol / MSM female mean stays about the same, MSM and MOLF means converge around 1600.>
-
 ```{r show.long.biv, echo=FALSE, include=FALSE}
@@
 
-Longer Sc compaction in females is supporte by the model framework, with sex being the most consistant significant effect (LTR and glm; p < 0.05).
-Additionally there are some significant subspecies and strain effects indicating that the SC lengths can evolve within strains and subspecies (LTR and glm; p < 0.05).
-
-These combined results of total SC per cell and reduced single bivalent SC metrics, indicate that females have longer SC lengths (chromosome axis), even in strains where males have more MLH1 foci per cell.
-I the means MLH1 foci per cell and SC area per cell, are considered seperatly, it is as if the males in these strains have more crossovers (more 2CO), but generally less 'space' to place them.
-
-This leads us to focus on aspects of the recombination landscape for the two major classes of bivalents; i) the placement of single foci along a single crossover bivalent and ii) the spacing of two foci on the same bivalent, or the interfocal distance (IFD).
-
-# Q1. 1CO Position Sex Differences
-Q1. Sex Differences in the Positions of Single Crossovers
-
-In most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus).
-In all strains, the landscape across 1CO bivalents is significantly different in females and males.
-Normalized foci positions tend to be more central in females (mean focus position = X) and closer to the telomere in males (mean focus position = Y) (t-test; p = 2.9210^{-22}).
-Sex is also the most significant effect on focus position in a mixed model (M1: p =1.2610^{-25}) and in general linear models (M2: p =1.3310^{-7}; M3: p = 1.3310^{-7}).
-These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species (Sardell Kirkpatrick).
+# Q1. Sex Differences in the Positions of Single Crossovers
 
@@
 ```
-
-<!-- motivating
-To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
--->
 
@@
 
-For most of the house mouse strains, 1COs are the major class of chromosomes observed across cells.
-There are known sex differences in the relative placement along chromsomes this section tests if there has been evolution in this aspect of the recombination landscape.
-
-The mean normalized foci positions per mouse from `r length( ( cleanCurated_BivData %>% filter(hand.foci.count == "1") )$Obj.ID )` single crossover bivalents.
-The CO positions are normalized by the SC length and anchored at the centromere with the normalized position values ranging from 0 to 1, (refered here as terminal or telomeric position).
-
-In all strains, the 1CO landscape had significant differences between males and female which followed the "typical landscape" of sex difference observed in many other species [@sardell_sex_2020].
-The foci position was more medially places in females while males have more terminal normalized foci positions (ttest; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
-
 ```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
@@
 
-The differences in 1CO landscapes due to sexes is confirmed by the model framework, with sex is the most significant effect in the mixed model (M1) (LTR; p =`r lmer.Q1.nrm1CO_M1_results$sex.results`) and the two linear models (glm M2; p =`r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`, glm M3; p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-
-<!--
-not addressing the strain and subspecies effects
-The random effect of strain is also significant (LRT; p =`r lmer.Q1.nrm1CO_M1_results$rand.pval`).
-
-While the musc subspecies, LEW and KAZ strains had slightly significant effects (glm M2; Musc subsp p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][2]`, LEW strain p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][5]`) and (glm M3; LEW strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][3]`, KAZ strain p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][7]`).
-
-**M1. Mixed Model**
-$$mouse \ average \ F1 position ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear model**
-$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$
-
-**M3. Linear model**
-$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$
-
-Ignore this section / kind of results for know
-Chromosome size effects are a confounding factor for CO position, with the expectation that shorter chromosomes will have a more uniform landscape compared to larger/longer chromosomes (cite).
-To account for this we again use the reduced bivalent sets (long.biv and short.biv) -->
+In most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus).
+In all strains, the landscape across 1CO bivalents is significantly different in females and males.
+Normalized foci positions tend to be more central in females **(mean focus position = X)** and closer to the telomere in males **(mean focus position = Y)** (t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value`).
+Sex is also the most significant effect on focus position in a mixed model (M1: p =1.2610^{-25} -- `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p =1.3310^{-7} -- `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 1.3310^{-7} -- `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species [@sardell_sex_2020].
 
@@
 
-<!--
-### Siscoten
-The metric Sis-co-ten measures the amount of sister cohesion connected to the other pole.
-The logic of how the sis-co-ten metric is outlined in the figure below.
-The goal is to use this metric to model different tension active cohesion amounts as a consequence of different numbers and placements of chiasmata/CO.
-This metric is calculated using SC area as a proxy to the amount of cohesion at metaphase.
-
-![from (Lee, J.
-(2019).
-Is age-related increase of chromosome segregation errors in mammalian oocytes caused by cohesin deterioration?.
-Reproductive Medicine and Biology.)](figureFiles/Lee2019_Fig2.png)
-
 ```{r show.siscoten, echo=FALSE, eval=FALSE}
@@
 
-Males have much clearer separation of siscoten across chrm classes.
-This is emphasized when SC length is also plotted.
-It seems like musc males have higher amounts of this metric compared to Dom males.
-To formally test the differences in sis-co-ten I plan to write a sub sampling / permutation loop to compare the mean(sis.co.ten) of the same numbers of bivalents of the same class.
-BUT females have a greater range -- so maybe it's just a scale issue.
-
 ```{r show.norm.siscoten, echo=FALSE, eval=FALSE}
@@
 ```
-
-I think the the normalized sis.co.ten plots also show that the there is more clustering of the sis.co.ten for the males.
 
@@
 
-All the sis.co.ten tests are highly significant. Maybe I should consider running a normalized sis.co.ten? I think nrm_siscoten would still reflect the differing cohesion structure/outcome.
-
-### Telomere and centromere Distance
-
-My metric for telomere and centromere distance measure the distance of the nearest foci to the ends of the bivalent (SC).
-In the plots below each point is a single bivalent.
-I choose not to use the mark for centromere because it seems noisy and inconsistent...
-
 ```{r telo.dist, echo=FALSE, eval=FALSE}
@@
 ```
-
-Males on average have much lower raw telomere distance (reflects the telomere bias) compared to females.
-In Males, 2CO bivalents have very low telomere distances, while the 1CO bivalents have a greater range.
-In females the ranges of telomere distances have much more overlap.
 
@@
 
-Mixed model result summary:
-
 ```{r log.reg_telo.dist, echo=FALSE, eval=FALSE}
@@
 ```
-
-The normalized centromere plots show that in Musc males, on 2CO bivalents the 1st CO is closer to the centromere end than in Dom males.
-Females have more overlap in the distributions of centromere distances across chromosome class compared to males.-->
 
@@
 
-Q1. Sex Differences in CO Interference (IFD)
-Crossover interference is one of the major determinates of the recombination landscape.
-It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
-To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated 1360 and 1272 2CO bivalents in females and males respectively.
-Mean interfocal distances (IFD) were calculated from 42 female and 45 male mice.
-We examined both raw IFDraw and normalized by SC length (IFDnorm).
-The mean IFDraw measures of crossover interference as a mechanical force mediated through the SC.
-We compared the IFDraw to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
-Previous measures found no sex difference in measures of IFDraw in mice (de Boer et al 2004).
-There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticusG; p = 0.27).
-The difference in mean IFDraw are slightly significant sexes from the full data set (t.test; p = 0.07).
-However domesticusG might be an outlier (by having a larger than average degree of sexual dimorphism).
-When the domesticusG observations are removed, the difference is not longer significant (t.test; p = 0.27).
-This result indicates there is no general pattern of sex differences in crossover interferencethat females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculusKAZ showing no difference (t-test; p = 0.33).
-While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-The difference in mean IFDnorm are highly significant between sexes.
-The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculusSKIVE p-value = 4.310^{-5}), musculusPWD p-value = 2.2810^{-4}), molossinusMSM p-value =0.01), domesticusLEW p-value = 3.5710^{-5}), domesticusG p-value = 0.05), and domesticusWSB (t.test; p-value =0.03) ).
-Only musculusKAZ did not have a difference in mean IFDnorm (t.test; p-value = 0.33).
-In examining theExamination of IFDnorm distributions, we note indicates that  the female IFDnorm observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFDnorm observationsvalues, whereas males are enriched for higher values.
-Comparitively, the IDFnorm distributions in males are enriched for longer IFDnorm.
-The three sets of modelsMixed models and general linear models of IFDnorm support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
-When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
 Q1 Summary, consequences of sex differences in the recombination landscape
@@
 The terminal position of single foci and the greater distance between two foci on the same chromosome will have a consequence of larger sections of linked sites segregating together in male gametes compared to female gametes (Veller).
-
-Additional Determinants of Genome-wide Recombination Rate Evolution in Males
-The next section is meant to focusNext, we used the contrast between high-recombination strains (musculusPWD , musculusSKIVE, and molossinusMSM) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( musculusPWD , musculusSKIVE, and molossinusMSM) to the low recombing strains.
-The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
-
-<equations>
-
-**Linear model M1**
-
-mouse average bivalentmetric = subsp*strain+??
-Linear model M2 
-mouse average bivalentmetric = strain+?? 
-The literature provides Previous empirical work suggests basic predictions for the direction of the relationship between the genome wide recombination rate and two aspects of the recombination landscape.; SC length is expected to be positively associated with genome-wide recombination rate because X.
-Crossover interference is expected to be negatively associated with genome-wide recombination rate because X.
-(positive) and crossover interference (negative).
-Following this logic we predict i) no strain effects for SC length or crossover interference within domesticus, ii(1) musculusPWD will have greater SC length and weaker interference than musculusSKIVE, which in turn will have longer SC and weaker interference compared to the other musculus strains, and iii(2) molossinusMSM will have longer SC and weaker interference compared to molossinusMOLF, and 3) domesticus strains will have similar SC length and crossover interference..
 
@@
 ```
+
+There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
+A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticusG; p = 0.27).
+This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
+In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculusKAZ showing no difference (t-test; p = 0.33).
+Examination of IFDnorm distributions indicates that female IFDnorm values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Mixed models and general linear models of IFDnorm support the inference of stronger interference in males: sex is the most significant variable (LRT: p = 6.7410^{-14}).
+When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
+In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+
+--- not sure what below is from --
+
+Crossover interference is one of the major determinates of the recombination landscape.
+It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
+To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated 1360 -- `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 1272 -- `r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males respectively.
+Mean interfocal distances (IFD) were calculated from 42 -- `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )` female and 45 -- `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
+We examined both raw IFDraw and normalized by SC length (IFDnorm).
+The mean IFDraw measures of crossover interference as a mechanical force mediated through the SC.
+We compared the IFDraw to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
+Previous measures found no sex difference in measures of IFDraw in mice [@de_boer_two_2006].
+
+----
+
+There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
+A marginally significant difference between the sexes (t-test; p = 0.07 -- `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain (t-test without domesticus^G^; p = 0.27 -- `r Q1.IFD.raw_sansG_t.test$p.value`).
+The difference in mean IFDraw are slightly significant sexes from the full data set (t.test; **p = 0.07)**.
+However domesticusG might be an outlier (by having a larger than average degree of sexual dimorphism).
+
+When the domesticusG observations are removed, the difference is not longer significant (t.test; p = 0.27) <already stated?>.
+
+This result indicates there is no general pattern of sex differences in crossover interferencethat females and males exhibit a similar level of interference when it is measured in physical (SC) units.
+In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculus^KAZ^ showing no difference (t-test; p = 0.33).
+While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+
+The difference in mean IFDnorm are highly significant between sexes.
+The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculus^SKIVE^ p-value = 4.310^{-5}), musculus^PWD^ p-value = 2.2810^{-4}), molossinus^MSM^ p-value =0.01), domesticus^LEW^ p-value = 3.5710^{-5}), domesticusG p-value = 0.05), and domesticus^WSB^ (t.test; p-value =0.03) ).
+Only musculus^KAZ^ did not have a difference in mean IFDnorm (t.test; p-value = 0.33).
+In examining theExamination of IFDnorm distributions, we note indicates that  the female IFDnorm observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFDnorm observationsvalues, whereas males are enriched for higher values.
+Comparitively, the IDFnorm distributions in males are enriched for longer IFDnorm.
+The three sets of models Mixed models and general linear models of IFDnorm support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
+When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
+In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
 
@@
 
-# Q1 Summary, consequences of sex differences in the recombination landscape
-
-Our results confirm sex differences in recombination landscapes described in several other species and we note that these differences have distinct consequences on the potential patterns of genetic variation resulting from the distinct recombination landscapes.
-The terminal position of single foci and the greater distance between two foci on the same chromosome will have a consequence of larger sections of linked sites segregating together in male gametes compared to female gametes (Veller).
-
-<!-- transition Q1 to Q2 -->
-
-The next section is meant to focus on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( _musculus^PWD^_ , _musculus^SKIVE^_, and _molossinus^MSM^_) to the low recombing strains.
+## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
+
+The next section is meant to focusNext, we used the contrast between high-recombination strains (musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) to the low recombing strains.
 The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-
 In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
 
+<equations>
+
+**Linear model M1**
+
+$$mouse \ average \ bivalent \ metric ~=~ subsp * rand(strain) + \varepsilon $$
+
+**Linear model M2**
+
+mouse average bivalent metric = strain+??
+
+$$mouse \ average \ bivalent \ metric ~=~ rand(strain) + \varepsilon $$
+
 **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
@@
 
-The literature provides basic predictions for the direction of the relationship the genome-wide recombination rate and two aspects of the recombination landscape; SC length (positive) and crossover interference (negative).
-Following this logic we predict i) no strain effects for SC length or crossover interference within _domesticus_, ii) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which will have longer SC and weaker interference compared to the other _musculus_ strains and iii) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_.
+Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
+**SC length is expected to be positively associated with genome-wide recombination rate because X**.
+Crossover interference is expected to be negatively **associated with genome-wide recombination rate because X**.
+Following this logic we predict (1) musculus^PWD^ will have greater SC length and weaker interference than musculus^SKIVE^, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) molossinus^MSM^ will have longer SC and weaker interference compared to molossinus^MOLF^, and 3) _domesticus_ strains will have similar SC length and crossover interference.
 
@@
 
-We compared mouse means of three SC-based metrics: total SC length, mean short bivalent length, and mean long bivalent length.
-The long bivalent data set was isolated by choosing bivalents in the X quartile of the SC length distribution within cells.
-The short bivalent dataset was isolated as described above.
-
-figure
-
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ([report spearman's rho and p-value]).
-Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
-While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01), separate tests by strain show that only within molossinus are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03; musculus: p = 0.87).
-Additionally, the means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = 0.88; long: p = 0.18).
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6}, molossinus p = 10^{-6}).
-In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-
-<bret's version above>
-
-Given the differences in the DSB numbers between a subset of the high and low recombining strains we predict that there will also be differences in SC lengths proportional to the increase in the mean genome-wide recombination rate.
-We compared mouse means of three SC based metrics; total SC, mean short bivalent length and mean long bivalent length.
-The long bivalent data set was isolated in the same manner as the short bivalent data set.
+```{r Q2.SC.length.bret.comments, echo=FALSE}
+#correlation
+
+```
 
@@
 
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC, however the variation in mean total SC does not completely predict the high and low recombining strains (Figure X).
-While for the total data set of mean total SC, the high recombining strains have significantly more SC area
-(t.test; p = `r Q2tot.sc_t.test_High.Low$p.value`), only the _molossinus_ strains are significantly different between the high and low recombining strains (t.test mol p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
-The _musculus_ are not significantly different across the groups (ttest; musc p= `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-Additionally, the means for the reduced bivalent datasets, short and long bivalents, are not significantly different between the high and low recombining strains (t.test; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`).
-
 ```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
@@
 
-For the linear model testing subspeices and strain (M1) and total SC as the dependant variable, the two subspecies effects are significant (glm; p= _Musculus_ `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and _Molossinus_ p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
-Whereas the model with reduced bivalent means as the dependant variables, a number of subspeicies and strains effects reach significance, (glm; p < 0.05) they are not consistant across the sets of models tested, indecateing to some extent the chromation compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-
 ```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
@@
 ```
-
-<These are the plots of the logistic regression.>
-
-<Each point is a mouse average SC.
-When all bivalents are pooled, the logistic regression SC lengths predict the groups.
-When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
-For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.>
 
@@
 
+We compared mouse means of three SC-based metrics: total SC length, mean short bivalent length, and mean long bivalent length.
+The long bivalent data set was isolated by choosing bivalents in the X quartile of the SC length distribution within cells.
+The short bivalent dataset was isolated as described above.
+
+figure
+
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC **([report spearman's rho and p-value])**.
+Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
+
+While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+Additionally, the means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value`; long: p = 0.18 -- `r Q2long.biv_t.test_High.Low$p.value`).
+
+In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
+In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
+
+<bret's version above>
+
 # Q2.1CO rec landscape evolution is decoupled from gwRR evolution
-
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = 0.24) and also when examined within subspecies (t-test; p = 0.41 and p = 0.07 for musculus and molossinus, respectively).
-While there are significant strain effects for domesticusWSB and molossinusMOLF in a general linear model with normalized position as the dependent variable (Figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
 
@@
 
-The normalized 1CO position is not significantly different between the high and low recombining groups for the total pooled data (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t.test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_ respectively).
-While there were significant strain effects for _domesticus^^WSB^_ and _molossinus^MOLF^_ (figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
-Hence we don't follow up with the model framework.
-
 ```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
@@
 
-# Q2 Evolution of interference is associated with genome-wide recombination rate evolution
-
-Mouse averages for both IFDraw and IFDnorm are significantly longer in high-recombination strains (t-test; IFDnorm: p = 7.7410^{-7}; IFDraw: p = 8.7810^{-6}).
-This pattern is confirmed by separate comparisons within musculus (t-test; IFDnorm: p =2.0410^{-5}; IFDraw: p = 1.9410^{-4}) and within molossinus (IFDnorm: p= 0.17; IFDraw: p = 0.08).
-Similar results are seen with general linear models for both IFDraw and IFDnorm: only effects associated with high-recombination strains are significant (p < 0.05).
-That IFDraw and IFDnorm show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFDnorm distributions between high-recombination and low-recombination strains is an enrichment of IFDnorm values under 30% in low-recombination strains.
-The frequency of IFDnorm values that fall below 30% ranges from 8.2% (domesticusG) to 16% (musculusKAZ) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculusSKIVE, molossinusMSM, and musculusPWD, respectively).
-In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
-However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
-At least on 2CO bivalents, crossovers are spaced further apart when the genome-wide recombination rate is higher.
+The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = 0.24 -- `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = 0.41 -- `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = 0.07 -- `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
+While there are significant strain effects for domesticus^WSB^ and molossinus^MOLF^ in a general linear model with normalized position as the dependent variable (Figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
+
+# Q2 Evolution of interference is associated with genome wide recombination rate evolution
 
@@
 ```
-
-Unlike total SC length and single foci landscapes, our metrics of interference from 2COs are significant predictors of evolution in genome-wide recombination rates.
-However, the pattern is in the opposite direction to our prediction; the high recombining strains have significantly stronger interference (instead of being predicted weaker interference).
-In the 2CO landscape, instead of observing weaker interference strength and two foci placed closer together on a single chromosomes, the two foci are spaced further apart.
-
-The mouse averages for both IFD^raw^ and IFD^norm^ were significantly longer in the high recombining groups (t.test; p = `r t.test_mouse.av.IFD_PER$p.value` for IFD^norm^ and p = `r t.test_mouse.av.IFD_ABS$p.value` for IFD^raw^).
-The pattern is confirmed with t.test comparing strains within _musculus_ and _molossinus_ (t.tests p =`r t.test_mouse.av.IFD_PER_MUSC$p.value` and p= `r t.test_mouse.av.IFD_PER_MOL$p.value` for IFD^norm^ and `r t.test_mouse.av.IFD_ABS_MUSC$p.value`  `r t.test_mouse.av.IFD_ABS_MOL$p.value` for IFD^raw^).
 
@@
 
-The greater IFD measures in high recombing strains is confirmed by fitting both the IFD^raw^ and IFD^norm^ into the linear model framework, where only the strain effects for the high recombining strains had significant p values (glm; p < 0.05).
-The similar patterns across the raw and normalized IFD values rule out differences in IFD^raw^ measures, SC lengths and chromosome size as the major driver of this pattern.
-We determined that the main difference in IFDPER distribution across the high and low groups is an enrichment of IFD^norm^ observations under 30% in low recombining strains.
-The rate of IFD^norm^ below 30% range from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in the low recombining strains, while the high strains all had rates under 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_ respectively).
-
-<REDO tests with the short IFDs removed, where do those above values come from >
-
 ```{r Q2.IFD_remove.short, echo=FALSE, include=FALSE}
@@
 ```
-
-When I compared the normalized IFDs with and without the short IFD's included, -- predictably, many mouse averages for 2CO IFD do nnot change for higher rec strains (but some low too).
-Most of the low rec strains show a change in the PER IFD (but some high rec mice also changed)
-
-For the strain averages, the high rec strains also show low change in the normalized IFD -also these are all higher (0.6)
 
@@
 
+Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
+This pattern is confirmed by separate comparisons within _musculus_ (t-test; IFD^norm^: p =2.0410^{-5} -- `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = 1.9410^{-4} -- `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
+Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
+
+That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
+We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
+The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (domesticus^G^) to 16% (musculus^KAZ^) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculus^SKIVE^, molossinus^MSM^, and musculus^PWD^, respectively).
+
+In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
+However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
+At least on 2CO bivalents, crossovers are spaced further apart when the genome-wide recombination rate is higher.
+
 # Q2 Summary

### 2020-04-08 doc/MatMeth.Rmd
checking all code chuncks after integrating brets comments

@@
 title: "Material Methods"
-date: "`r Sys.Date()`"
+date: '`r Sys.Date()`'
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
   html_document:
@@
     toc_float: yes
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
+  word_document:
+    toc: yes
+    toc_depth: '5'
 bibliography: refs_MatMeth.bib
@@
 ```
+
+```{r data.sets, echo=FALSE}
+
+#Q1.real.short_mouse_table
+
+#
+
+```
+
+ToDo; Bret's suggestion for moving thing to Mat and Methods
+
+**For SC based metric**
+
+- To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
+This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
+A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+
+- As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
+Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
+
+**SC Length calq**
+
+-  **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
+
+- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
+
+- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+**IFD^norm^**
 

### 2020-04-08 doc/Results_v2.rmd
checking all code chuncks after integrating brets comments

@@
 title: "MLH1 Results Outline"
-date:
+date: "null"
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
+  pdf_document: default
   html_document:
@@
 -->
-# Main Figure
 
@@
 Hence, relative recombination rates in the two sexes are evolutionarily labile.
+
+The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
+Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
+Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
+
 Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C).
@@
 In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).
-
-The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
-Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
-
-Separtely examining the mouse means of MLH1 foci per cell separately for each sex reveal distinct patterns of variation (Figure1 B and C).
-Female recombination rate means are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure1 B).
-In stark contrast, males specific means separate more clearly into two groups of strains with high genome-wide recombination rates (near 30 MLH1 foci per cell) and low genome-wide recombination rates, (near 23 MLH1 foci per cell) (Figure1 C).
 
@@
 ```
-
-<add a transition sentence, about the expectations under different evolutionary scnarios>
 
@@
 
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
-This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
-A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-
 <figure>
@@
 
-As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
-
-Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
 Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
@@
 
-Q1 Summary, consequences of sex differences in the recombination landscape
-Our results confirm sex differences in recombination landscapes described in several other species and we note that these differences have distinct consequences on the potential patterns of genetic variation resulting from the distinct recombination landscapes.
-The terminal position of single foci and the greater distance between two foci on the same chromosome will have a consequence of larger sections of linked sites segregating together in male gametes compared to female gametes (Veller).
-
 ```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
@@
 While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
---- not sure what below is from --
-
-Crossover interference is one of the major determinates of the recombination landscape.
-It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
-To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated 1360 -- `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 1272 -- `r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males respectively.
-Mean interfocal distances (IFD) were calculated from 42 -- `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )` female and 45 -- `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
-We examined both raw IFDraw and normalized by SC length (IFDnorm).
-The mean IFDraw measures of crossover interference as a mechanical force mediated through the SC.
-We compared the IFDraw to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
-Previous measures found no sex difference in measures of IFDraw in mice [@de_boer_two_2006].
-
-----
-
-There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = 0.07 -- `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain (t-test without domesticus^G^; p = 0.27 -- `r Q1.IFD.raw_sansG_t.test$p.value`).
-The difference in mean IFDraw are slightly significant sexes from the full data set (t.test; **p = 0.07)**.
-However domesticusG might be an outlier (by having a larger than average degree of sexual dimorphism).
-
-When the domesticusG observations are removed, the difference is not longer significant (t.test; p = 0.27) <already stated?>.
 
@@
 
-Crossover interference is one of the major determinates of the recombination landscape.
-It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution.
-To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and
-`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males respectively.
-Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice.
-We examined both raw IFD^raw^ and normalized by SC length (IFD^norm^).
-The mean IFD^raw^ measures of crossover interference as a mechanical force mediated through the SC.
-We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC.
-Previous measures found no sex difference in measures of IFD^raw^ in mice [@de_boer_two_2006].
-
-The difference in mean IFD^raw^ are slightly significant sexes from the full data set (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`).
-However _domesticus^G^_ might be an outlier (by having a larger than average degree of sexual dimorphism).
-When the _domesticus^G^_ observations are removed, the difference is not longer significant (t.test; p = `r Q1.IFD.raw_sansG_t.test$p.value`).
-This indicates there is no general pattern of sex differences in crossover interference measured in physical (SC) units.
-
-While the comparisons of the IFD^norm^ metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
-The difference in mean IFD^norm^ are highly significant between sexes.
-The mean IFD^norm^ are significantly long in six of seven strains (MSM t.test; p-value = `r Q1.IFD.PER_t.test_MSM$p.value`), _musculus^SKIVE^_ p-value = `r Q1.IFD.PER_t.test_SKIVE$p.value`), _musculus^PWD^_ p-value = `r Q1.IFD.PER_t.test_PWD$p.value`), _molossinus^MSM^_ p-value =`r Q1.IFD.PER_t.test_MSM$p.value`), _domesticus^LEW^_ p-value = `r Q1.IFD.PER_t.test_LEW$p.value`), _domesticus^G^_ p-value = `r Q1.IFD.PER_t.test_G$p.value`), and _domesticus^WSB^_ (t.test; p-value =`r Q1.IFD.PER_t.test_WSB$p.value`) ).
-Only _musculus^KAZ^_ did not have a difference in mean IFD^norm^ (t.test; p-value = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-
 <!--  finer distribution patterns <Full dataset (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`).> -->
@@
 ```
-
-In examining the IFD^norm^ distributions, we note that the female IFD^norm^ observations are centered at approximately 50% and have slight an enrichment of short (<25%) IFD^norm^ observations.
-Comparitively, the IDF^norm^ distributions in males are enriched for longer IFD^norm^.
 
@@
 
-## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
+# Additional Determinants of Genome-wide Recombination Rate Evolution in Males
 
@@
 In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
-
-<equations>
-
-**Linear model M1**
-
-$$mouse \ average \ bivalent \ metric ~=~ subsp * rand(strain) + \varepsilon $$
-
-**Linear model M2**
-
-mouse average bivalent metric = strain+??
-
-$$mouse \ average \ bivalent \ metric ~=~ rand(strain) + \varepsilon $$
-
-**Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
-
-**Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
 

### 2020-04-08 src/Result_Setup.Rmd
generating file for organizing DF and other results

--- 2020-04-08 doc/Results_v2.rmd
@@
 ---
-title: "MLH1 Results Outline"
-date: "null"
-output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document: default
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-bibliography: refs_Main_Report.bib
+title: "Results_setup"
+author: "April Peterson"
+date: "April 7, 2020"
+output: html_document
 ---
@@
 
-<!-- 
-two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
--->
-
-<!--
-From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
-
-the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
-
-INSERT TABLE FOR THE TOTAL SC means ect
-
-The difference between
-
-bin is short for binary, it reports the binarized area of the SC
-
-skel -- takes the skeletonized midline of the bin objects
--->
-
-```{r table.for.plot, echo=FALSE, warning=FALSE}
-#why do I add in a different table?
-
-#this move this to the setup file
-#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed
-
-#pooled
-strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  =length(nMLH1.foci),
-                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(nMLH1.foci, na.rm = TRUE),
-                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
-                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
-                       #Standard error (mice vs cells)
-                        se  = round(sd / sqrt(Ncells), 3)
-
-                       #range_mouse_means = range()
-                   #consider adding crhm proportions?
-    )
-
-strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
-                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")
-
-#adjust order of plot here
-strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
-     "MSM", "MOLF",   
-  "PWD", "SKIVE",  "KAZ",
-  "WSB","G","LEW","PERC",
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-```
-
-```{r Fig1, echo=FALSE, warning=FALSE}
-
-#might need to nudge the X values
-
-Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-#remove y labfor male
-
-   #segments
-#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
-  #Musc
-#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
-  #mol
-#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
-  #cast
-#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
-  #outgroups
- # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
-  #label
-
- #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
-   #        y = c(18.5,18.5,33,18.5, 18.5),
-
-  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
-    #       fontface = 'italic', hjust = 0.5, size=3.5)
-
-```
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-#table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
-  theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-#female.specific.plot
-
-```
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
-#try customize the plot size
-
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")
-
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-#fomo.male_mouse.av
-
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-male.plot
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-
-```
-
-```{r three.plots, echo=FALSE, warning=FALSE}
-#merge the 3 plots
-#try patchwork
-
-#Fig1_v1_strain + female.specific.plot + male.plot
-
-#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
-#Fig1_v1_strain / (female.specific.plot + male.plot )
-
-#(p1 | p2 | p3)/p4
-```
-
-```{r main.fig1s, echo=FALSE, include=FALSE}
-strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
-
-Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point( size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-
-#strain.table.HQ_Fig.og
-
-strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")
-
-strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
-     "MSM", "MOLF",   
-  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
-  "WSB","G","LEW"), order=T )
-
-Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-
-```
-
-```{r basic.numbers, echo=FALSE, eval=FALSE}
-
-#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-
-#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-```
-
-```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
-#str(AP_mouse_table_w.Ages)
-
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-#remove sex.x == na, and strain.x=na or F1
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
-
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
-#plot variance and mean (facet sex)
-
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
-
-VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
-
-cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
-
-#str(AP_mouse_table.HQ)
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
-#[1] 0.5750399
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
-#0.321601
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
-# 0.3515544
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
-#-0.03713536
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
-# p-value = 8.623e-09
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
-
-#remove mice with fewer than 10 cells
-
-AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
-
-VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
-
-cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
-
-cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
-    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-
-#why did I male just a female table?
-female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-
-female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-                        #quality?
-)
-```
-
-# genome-wide recombination rate estimates for both sexes
-
-```{r MLH1.summ.stats, echo=FALSE}
-
-#strain.table.HQ_Fig.og  all strains
-#strain.table.HQ_Fig  extra strains removed
-
-main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
-
-main.table <- add_subsp(main.table)
-main.table <- add_species(main.table)
-
-#main.table %>% filter(sex == "female")
-#main.table %>% filter(sex == "male")
-
-#G.HETC.val #1.2
-#LEW.HETC.val  #1
-#WSB.HETC.val  #1
-#KAZ.HETC.val  #1.1
-
-#SKIVE.HETC.val # 0.96
-#PWD.HETC.val  #0.88
-#MSM.HETC.val #0.9
-
-#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
-
-#strain.table.HQ_Fig
-
-#sum( (main.table  %>% filter(sex == "male"))$Ncells)
-#the kable version is bulkier than the other versaion
-
-```
-
-```{r MLH1.sum.stats.Bret.edits, echo=FALSE}
-
-AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
-#total mice
-#length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )
-
-#average cell number for spermatocytes and oocytes
-#mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
-#mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
-```
-
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M.
-m.
-domesticus*, *M.
-m.
-musculus* and *M.
-m.
-molossinus*.
-Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )` mice were quantified from an average of `r mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes).
-
-Graphical comparisons between the two sexes reveal several patterns (Figure 1A).
-First, recombination rate differs between females and males in most strains.
-Second, the direction and magnitude of heterochiasmy varies among strains.
-Although the majority of strains show higher recombination rates in females (following the pattern in laboratory mice), two musculus strains and one molossinus strain exhibit male-biased heterochiasmy.
-Hence, relative recombination rates in the two sexes are evolutionarily labile.
-
-The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
-Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
-
-Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C).
-Female recombination rates are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure 1B).
-In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).
-
-```{r female.strain.table, echo=FALSE}
-female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-
-                        Nmice = length(unique(mouse)),
-
-                        total.cells  = sum(Ncells, na.rm = TRUE),
-
-                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#F.strain.table
-
-#F.strain.table
-
-# I don't know why most of the table is blank
-```
-
-```{r female.dif, echo=FALSE}
-
-#how much greater are the high strains?
-G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
-
-LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
-mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
-
-#how much greater is MSM female
-MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
-
-sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
-mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
-
-G.female.diff <- (mean.G.female / mean.all.other)
-LEW.female.diff <- (mean.LEW.female / mean.all.other)
-MSM.female.diff <- (mean.MSM.female / mean.all.other)
-```
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-#variance plots
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-#Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
-
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-```
-
-```{r FEMALE.VAR.MODELS, echo=FALSE}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
-
-lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
-lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
-<!-- male section -->
-
-```{r male.strain.table, echo=FALSE}
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#M.strain.table
-#same issue as the female table .. ugh
-```
-
-```{r male.diff, echo=FALSE}
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /  
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-```
-
-```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-#summry.male.strain.effects <- summary(glm.male.strain.subsp)
-#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#1. determine which strain effects are sig
-
-#M2
-MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
-sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-```
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-```
-
-# Partitioning variation in recombination rate
-
-```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}
-
-#DF.HetC.MixedModel.HQ
-
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
-
-MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
-sex.results =  MLH1.MM_sex, 
-               interaction.pval = MLH1.MM_interaction,  rand.pval = 
-MLH1.MM_strain)
-
-```
-
-```{r MLH1.glmM2, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
-
-#strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
-
-summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
-
-#table of just coeff anf pvalue
-#round(summ$coefficients[,c(1,4)],5 )
-
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-```
-
-```{r MLH1.glm.M3, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
-
-summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
-
-#round(summ.strain.fixed$coefficients[,c(1,4)],5)
-
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-```
-
-To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
-We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results` -- 3.9710^{-5}), subspecies (p=`r MLH1.MMlmer_results$sub.sp`--  1.7210^{-4}), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval` -- 3.110^{-5}) each significantly affect recombination rate.
-
-After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p= 3.9910^{-6} p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and domesticus^G^ (p= 1.0410^{-6}  -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
-
-We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4).
-In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = ~ 5 X foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect ~ 2 foci).
-These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p=0.02  -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
-
-Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
-
-<bret above>
-
-<!-- transition gwRR to within mouse variance -->
-
-## Within mouse variance in CO count per cell
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
-
-#Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-
-Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
-To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`).
-In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (**criteria?**).
-The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2).
-When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
-These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
-
-# Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
-
-```{r DMC1.setup, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "L"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "Z"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
-#            row.names = FALSE)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-```
-
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
-#think of better color scheme
-#make the jitter col thinner or make the stacked dots.
-#remember that box-plots add outlier dots.
-#change y labels
-
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
-#change order
-
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.boxplot
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-  ggsignif::geom_signif(
-    data=anno_df_stage, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
-```
-
-```{r plot.t.test, echo=FALSE, include=FALSE}
-
-  #this will add neat bars, but still need to figure out what groups ect it's actually using
-#THE stat_compare function adds a arm thing and pvalue for 
-#  stat_compare_means(comparisons = my_comparisons)
-
-#status - legend, expression y-axis, facet, facet_wrap SYMBOL
-#legend - stage. y=total count, FACet strain
-
-#add column to distinguish high vs low rec strains
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
-DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
-
-# annotation table with adjusted pvals and y-position of the labels
-
-anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=stage), position=position_jitterdodge()) + 
-  facet_wrap(~stage, nrow=1) + 
-  ggsignif::geom_signif(
-    data=anno_df, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#example for ggplot+ test from here
-#https://github.com/kassambara/ggpubr/issues/65
-
-#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
-
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_anova, echo=FALSE, warning=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-
-#unorder strain
-DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
-
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
-z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
-```
-
-```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-#attach this mini DF to the DMC1 data
-
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-
-#DMC1.mean.table
-DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
-
-colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
-                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
-
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
-In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` -- 76 leptotene-stage and 75 --`r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
-
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`; one-way ANOVA, p = 0.00027 -- `r format(l.anova.pval, scientific=F)`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`; one-way-anova, p = 0.15 -- `r z.anova.pval`).
-After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = 0.94 -- `r DMC1.Lratio_t.test$p.value` and p = 0.11 -- `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios, respectively).
-This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
-
-```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
-
-#Subsampling code -- don't report
-
-#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
-Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
-Permutations were done for both cell stages.
-The actual p value is in red.
-
-#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
-
-#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
-# the 2 groups
-
-#make the 2 pools
-High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-
-Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-
-#permute drawing / subsampling and performing the t-test
-#put the p.value results into a DF!
-
-#permut.SC.cv.diff  <- as.data.frame(
-#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
-
-#permuting samples -- and re-licating t-tests
-
-#nrep <- 1000
-
-#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
-bunchOpvals.L <- replicate(1000, 
-    # t.test
-    t.test( sample(High.L.pool, 25, replace = FALSE),
-            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
-          )
-bunchOpvals.L <- as.data.frame(bunchOpvals.L)
-bunchOpvals.L$stage <- "L"
-
-bunchOpvals.Z <- replicate(1000, 
-    # t.test
-    t.test( sample(High.Z.pool, 25, replace = FALSE),
-            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
-          )
-
-bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
-bunchOpvals.Z$stage <- "Z"
-
-#could merge the lists of p-values
-replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
-#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
-
-#ASSUMPTIONS; normality, equal variance
-
-#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-#plot.pvals.L
-#plot.pvals.Z
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
-```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-#these are histogram plots for testing Normality shape -- move to supplemenatry section
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-```
-
-# Evolution of genome-wide recombination rate is reflected at the single chromosome level
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#sort by category, 
-#compare total boxes and rate passed
-#Curated_BivData  -- pass rate?
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-#use these two metrics
-#cell.level.BivData_table$boxs.IDd 
-#cell.level.BivData_table$nboxes_passed
-
-#calq the mean rate by category
-
-cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
-
-algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
-                        n_obs = length( unique(fileName) ),
-                         mean_pass.rate = mean(pass.rate)
-                         )
-
-cell.level.BivData_table <- add_sex(cell.level.BivData_table)
-
-algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
-                        n_obs = length(unique(fileName)),
-                        mean_pass.rate = mean(pass.rate)
-                         )
-
-#showing the table..
-#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-```{r chrm_prop_data, echo=FALSE, warning=FALSE}
-#shut off eval for now because of wd / path error
-#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#try adding the BivData version of the proportions
-
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
-
-                 n0CO = sum(n0CO, na.rm = TRUE),
-                  n1CO = sum(n1CO,na.rm = TRUE),
-                  n2CO = sum(n2CO,na.rm = TRUE),
-                  n3CO = sum(n3CO,na.rm = TRUE),
-                  n4CO = sum(n4CO,na.rm = TRUE),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#this is the table that needs melting
-#but is also easy to redo by hand
-
-#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
-```
-
-```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
-
-#add subspecies to these values¬
-melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
-
-#add the subsps
-melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
-
-melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
-
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))+theme(legend.position="none")
-
-#need a second DF / the first df
-#+ annotate() text 
-#annotate()text
-
-#proportion_stats_category,  tot_biv, ncells
-#add number of cells and bivalents in the corner
-
-#chrm.class.barplot
-```
-
-```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
-
-#prop.test(x, y)  x number of successes, y total  
-#strain sums of n
-#look up continuity correction
-
-#mouse.level.BivData_table <-- has more male strains
-#Mouse.Table_BivData_4MM <- only sex matched strains
-
-#perform test across strain 
-#removed all lines with no hand.foci.count entry
-Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
-
-                  nbiv = sum(tot.biv ),              
-                 n0CO = sum(n0CO ),
-                  n1CO = sum(n1CO ),
-                 n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO ),
-                 n4CO = sum(n4CO )
-                          )
-
-Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
-
-for(g in 1:length(Chrm.prop_strain_table$nbiv)){
-
-  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
-                                     Chrm.prop_strain_table$n1CO[g],
-                                     Chrm.prop_strain_table$n2CO[g],
-                                     Chrm.prop_strain_table$n3CO[g] ))  
-}
-
-Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
-#there are still some wacky entries -- but much fewer than before
-
-chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
-
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
-#the pvalues for 0 are not sig dif
-#the p values for 1 and 2CO are sig diff
-
-#seperate by subsp
-chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
-chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
-chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
-
-Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
-Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
-Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
-Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
-#WSB has fewer 2COs than other 3 strains
-
-Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
-Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
-Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
-#KAZ has slightly more 1CO than the other two
-
-#high.female_G.LEW.
-chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
-
-high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
-high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
-high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
-
-#Low.female.G.LEW.
-chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
-
-low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
-low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
-low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
-
-#all females
-
-chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
-
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
-#the pvals for these 3 class are sig dif
-
-#seperate by subsp
-chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
-chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
-chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
-
-Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
-Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
-Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
-#non of the Dom proportions are different
-
-Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
-Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
-Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
-#all v different
-
-Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
-Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
-Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
-
-#high vs low
-chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
-                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
-                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
-
-chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
-chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
-
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
-
-Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
-Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
-Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
-#the low strains do vary slightly in their 1CO and 2CO proportions
-
-#includes SKIVE
-Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
-Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
-Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
-#all very sig p values
-
-High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
-#these proportions are v different
-
-#no SKIVE
-High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
-High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
-High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
-#MSM and PWD only slightly sig dif 2CO
-
-#MSM-MOLF
-MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-#PWD-KAZ
-PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.MSM.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "MSM") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "MSM") )$nbiv )
-
-```
-
-```{r Chrm.props.plot2, echo=FALSE}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
-
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
-
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
-
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-#no filename for add mouse, strain subsp
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#can't apply function use this
-chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
-             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
-                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
-                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
-
-          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
-              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
-
-                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
-                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
-            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
-                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
-                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
-                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
-                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
-                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
-
-             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
-               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
-
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
-                                                                 "Musc",
-                                                                 "Mol",
-                                                                 "Spretus","Spic",
-                                                                 "other") )
-#ggplot
-
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
-#chrm.class.barplot
-chrm.class.barplot
-
-```
-
-```{r Chrm.Prop.misc.calulations, echo=FALSE}
-
-total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-
-#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
-#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
-
-#MELT_Chrm.Proportion$Rec.group <- ifelse(grepl("PWD male", MELT_Chrm.Proportion$category), 1, 
-#                              ifelse(grepl("MSM male", MELT_Chrm.Proportion$category), 1,
-#                                ifelse(grepl("SKIVE male", MELT_Chrm.Proportion$category), 1, 0)))
-
-#Chrm.prop.male.low <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 0)
-#Chrm.prop.male.HIGH <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 1)
-
-#strain av tables
-#chrm.prop.male.high1
-#chrm.prop.male.low
-```
-
-To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents ([@peterson2019]).
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from
-`r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
-From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
-After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.
-We assume that the isolation of bivalents within cells is unbiased.
-Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns.
-An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
-
-<figure>
-
-Ninety-six percent of single bivalents in our pooled dataset (n = 34,982 -- `r total.biv.all`) have either one or two crossovers (Figure X).
-The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
-low recombining strains (Figure X).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in molossinus^MSM^.
-
-Following patterns in the genome-wide recombination rate, male musculus^PWD^ and male molossinus^MSM^ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; musculus^PWD^ vs.
-musculus^KAZ^ p = 3.1510^{-33} -- `r PWD.KAZ.2CO.prop.test$p.value`; molossinus^MSM^ vs.
-molossinus^MOLF^ p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
-
-Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
-We address two main questions.
-First, which traits are sexually dimorphic?
-Second, which traits differ between males from high-recombination vs.
-low-recombination strains?
-
-# Q1 Sex Differences in the Length of the Synaptonemal Complex
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#CLEAN UP THE FULL DATA SET
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-#nested boxplots, pooled Bivalents
-
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
-  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
-```{r Q1.SC.show.short.biv, echo=FALSE}
-
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
-Q1.SC.short.mouse_scatter_boxplot
-
-#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
-skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-#skiveQ1.SC.short.mouse_scatter_boxplot
-#there are 2 lowish female means,
-
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
-#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
-#1 male, 3 females measures
-#try to merge the manual - whole cell observations (at least 2 males)
-
-```
-
-```{r Q1SC_short.biv_dif, echo=FALSE}
-
-#calq the different in short.biv means
-
-#WSB
-Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
-
-#LEW
-Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
-
-#G
-Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
-
-#PWD
-Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
-
-#KAZ
-Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
-
-#SKIVE
-Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
-
-#MSM
-Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
-
-#this bivData ratio is the same as mouse average
-Q1SC.BivData.short.biv_DIFF_MSM <- mean( (real.short.bivData %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (real.short.bivData %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
-
-```
-
-```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
-# fig.width = 4, fig.height=2.5
-
-Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
-
-Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
-   "WSB","G","LEW","PERC", 
-  "PWD", "SKIVE",  "KAZ",
-  "MSM", "MOLF"), order=T )
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#order the strains
-Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
-  arrange(strain) %>%
-  mutate(mouse = factor(mouse))
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#add category mean
-#totSC.strain.av.table %>% filter
-#totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
-#totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
-
-Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(sex~subsp, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area") + theme(legend.position="none")
-
-Tot.SC_mouse.av_ALL
-
-#full dataset ---
-
-totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
-totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
-
-Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
-
-totsc.mol <- Q1.totSC.mouse.av.table %>% filter( subsp == "Mol" )
-totsc_strain.av.mol <- totSC.strain.av.table %>% filter( subsp == "Mol" )
-
-Tot.SC_mouse.av_Mol <- ggplot(data= totsc.mol, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.mol, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
-
-```
-
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#ALL
-Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)
-
-#dom
-Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#Musc
-Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
-#cell level total SC
-#MLH1.merge.Skel   $skel_size
-#error not enought female observations, ONLY 1 female SKIVE mouse
-
-Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
-           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-```{r Q1.totSC_male.first.comp, echo=FALSE}
-
-Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
-```
-
-```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA!!
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-#Q1.lmer.totSC_results
-
-#G
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-
-```
-
-In many mammalian species, the synaptonemal complex (SC) is longer in females.
-Yet, the majority of these observations come from species with female-biased heterochiasmy.
-Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male musculus^PWD^ and male molossinus^MSM^ should have longer SC lengths than the other strains.
-
-<figure>
-
-Short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculus^SKIVE^.
-The female:male ratio ranges across strains from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
-
-Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
-That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies.
-
-<figure>
-
-In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers.
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-#subsp effect, many strain, #male
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-```{r Q1.SC.Q12, echo=FALSE}
-
-#make comparisons for the high quality cells
-
-#get Q12 cells from MLH1.merge.Skel
-#calq the mouse averages
-
-MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
-
-#unorder the factors
-totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
-totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
-totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
-totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
-
-#REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(strain, sex),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
-
-#unorder the factors
-totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
-totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
-totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
-totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
-
-```
-
-```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
-
-Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Dom
-#Tot.SC_mouse.av_Dom
-
-Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Musc
-#Tot.SC_mouse.av_Musc
-
-Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-```
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-# Q1. Sex Differences in the Positions of Single Crossovers
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE}
-
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
-#main figure of density plots,
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
-           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test( 
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-#lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-
-```
-
-In most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus).
-In all strains, the landscape across 1CO bivalents is significantly different in females and males.
-Normalized foci positions tend to be more central in females **(mean focus position = X)** and closer to the telomere in males **(mean focus position = Y)** (t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value`).
-Sex is also the most significant effect on focus position in a mixed model (M1: p =1.2610^{-25} -- `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p =1.3310^{-7} -- `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 1.3310^{-7} -- `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species [@sardell_sex_2020].
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-
-#real.long.bivData
-Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-
-Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
-
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-#mouse.av.long.biv.table
-
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-```
-
-```{r show.siscoten, echo=FALSE, eval=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-```{r show.norm.siscoten, echo=FALSE, eval=FALSE}
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-```
-
-```{r log.reg_siscoten, echo=FALSE, eval=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-```
-
-```{r telo.dist, echo=FALSE, eval=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-```
-
-```{r MM.telo, echo=FALSE, eval=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-drop1(lme_MM_telo.dist, test="Chisq")
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-```
-
-```{r log.reg_telo.dist, echo=FALSE, eval=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE, eval=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE, eval=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE, eval=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE, eval=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, eval=FALSE, include=FALSE}
-#IGNORE THIS FOR KNOW, THIS
-
-#filter all Curated biv data for 1 strain
-#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
-
-# the stardell paper sums the number of COs for bins (10)
-
-#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
-#but then the model would be close to logistic
-
-#summing the number of COs per bin, losses the chrom specific data....
-#!! need to use a melted Data frame to capture all nrm.foci obs
-#wsb.ttest<- mouse.avs_4MM %>% filter()
-#start with same chrm classes
-#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
-
-WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
-
-PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
-
-anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
-
-#or AIC(model1,model2)
-
-```
-
-```{r Q1.mean.nrmF1_table, echo=FALSE, eval=FALSE}
-
-#make a table for the sex specific values of the normalized F1 positions
-
-#for strain, sex -- Curated_BivData
-#or (Mouse.Table_BivData_4MM)
-#norm.first.foci.pos
-Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
-
-                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
-                  n1CO = sum(n1CO )
-
-)
-
-```
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#MELT_Curated_BivData
-
-#these isn't long.biv data
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-real.long.bivData
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
-
-```
-
-# Q1. Sex Differences in CO Interference (IFD)
-
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
-
-#each point is a 2CO, with the F1 and F2 position 
-#based on plot from Dourad et al (Mercier)
-
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "WSB"),
-    aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.5)+
-  ggtitle("Raw IFD WSB")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#trying to move x axis to top
-#cowplot might be the easier way -- since kinda diffecult in ggplot
-#or ggdraw
-
-P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
-<!--t.tests -->
-
-```{r Q1.IFD_Ttests, echo=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
-#SKIVE
-Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-
-```
-
-There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticusG; p = 0.27).
-This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculusKAZ showing no difference (t-test; p = 0.33).
-Examination of IFDnorm distributions indicates that female IFDnorm values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Mixed models and general linear models of IFDnorm support the inference of stronger interference in males: sex is the most significant variable (LRT: p = 6.7410^{-14}).
-When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-
-This result indicates there is no general pattern of sex differences in crossover interferencethat females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculus^KAZ^ showing no difference (t-test; p = 0.33).
-While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
-The difference in mean IFDnorm are highly significant between sexes.
-The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculus^SKIVE^ p-value = 4.310^{-5}), musculus^PWD^ p-value = 2.2810^{-4}), molossinus^MSM^ p-value =0.01), domesticus^LEW^ p-value = 3.5710^{-5}), domesticusG p-value = 0.05), and domesticus^WSB^ (t.test; p-value =0.03) ).
-Only musculus^KAZ^ did not have a difference in mean IFDnorm (t.test; p-value = 0.33).
-In examining theExamination of IFDnorm distributions, we note indicates that  the female IFDnorm observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFDnorm observationsvalues, whereas males are enriched for higher values.
-Comparitively, the IDFnorm distributions in males are enriched for longer IFDnorm.
-The three sets of models Mixed models and general linear models of IFDnorm support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
-When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
-<!--  finer distribution patterns <Full dataset (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`).> -->
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE}
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#this might
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
-
-```
-
-```{r Q1.IFD_gamma, echo=FALSE}
-library(EnvStats)
-#cleanCurated_BivData
-
-#all.female.multi <- (cleanCurated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count > 1) %>%
-#                       filter( !is.na( IFD1_ABS) ))  #$IFD1_ABS
-
-#this MELT BIVDATA should be good to go
-MELT.IFD.BIVDATA <- melt(cleanCurated_BivData[,c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
-                            "Obj.ID","hand.foci.count",
-                            "mouse","category","strain", "sex", "Rec.group",
-                            "IFD1_ABS","IFD2_ABS","IFD3_ABS")],
-
-             id=c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
-                            "Obj.ID","hand.foci.count",
-                            "mouse","category","strain", "sex", "Rec.group"))
-#rename the cols and clean up
-MELT.IFD.BIVDATA <- MELT.IFD.BIVDATA[(!is.na(MELT.IFD.BIVDATA$value)),]
-
-colnames(MELT.IFD.BIVDATA) <- c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
-                            "Obj.ID","hand.foci.count",
-                            "mouse","category","strain", "sex", "Rec.group", 
-                               "whichIFD", "IFDval")
-
-MELT.IFD.BIVDATA$Nrm.IFD <- MELT.IFD.BIVDATA$IFDval / MELT.IFD.BIVDATA$chromosomeLength
-
-#START isolating all the types of IFDs
-
-#all.female.multi$mycol <- apply(all.female.multi[,c('IFD1_ABS','IFD2_ABS','IFD3_ABS')], 1, function(x) x[!is.na(x)][1])
-
-all.female.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "female")#1697
-
-all.male.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "male")#1392
-
-#shape and scale parameter
-
-#gamma fits
-FULL.female.gamma <- egamma( all.female.IFDs$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.8
-
-#gamma fits
-FULL.male.gamma <- egamma( all.male.IFDs$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#shape  5.4
-
-#male, stronger interference
-
-WSB.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.1, 3.73 (raw, PER) (n=252)
-
-WSB.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.6, 5.10 (raw, PER) (n=186)
-#try the short and long bivalents dataset
-
-WSB.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3, 3.237 (raw, PER) (n=60)
-
-#gamma fits
-WSB.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.5,  5.1 (raw, PER) (n= 109)
-
-WSB.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.9, 3.9 (raw, PER)  (n=107)
-#there are 0 3CO and 2  4CO bivalents -- this causes a greater difference in the PER based value
-
-```
-
-```{r gamma.IFD, echo=FALSE}
-
-female.gamma <- egamma( (all.female.IFDs )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-f2CO.female.gamma <- egamma( (all.female.IFDs  %>% filter(hand.foci.count == 2)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-f3CO.female.gamma <- egamma( (all.female.IFDs %>%  filter(hand.foci.count == 3)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-f4CO.female.gamma <- egamma( (all.female.IFDs %>%  filter(hand.foci.count == 4)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-
-PWD.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.8, 4.844 (raw, nrm)  (n = 357)
-PWD.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.8, 6.746 (raw, nrm) (n = 285)
-PWD.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3, 3.628 (n = 72)
-
-KAZ.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-
-KAZ.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-
-KAZ.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-#####
-SKIVE.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-##
-MSM.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-MSM.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-######
-G.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-G.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-######
-LEW.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-LEW.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-######
-WSB.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-WSB.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-WSB.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-WSB.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-```
-
-```{r Q1.male.IFD, echo=FALSE}
-
-#gamma fits
-male.gamma <- egamma( (all.male.IFDs )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-m2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-m3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE,         ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-#gamma fits
-PWD.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#8.1, 11.74 (n=318) (lots of variation when I apply sample size of 18)
-
-PWD.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#10.3, 16.962, (n=300)
-
-PWD.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#2.9, 3.4, (n=18)
-
-####
-KAZ.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-KAZ.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-KAZ.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-SKIVE.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-MSM.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE,     ci.type = "two-sided", ci.method = "normal.approx",
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-MOLF.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MOLF.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MOLF.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-G.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-LEW.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-WSB.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-WSB.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-```
-
-```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
-
-#M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-
-```
-
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE}
-
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
-summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-```
-
-```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-```
-
-The three sets of models support the result of stronger interference in males, sex was the most significant significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`).
-When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.
-
-```{r Q1.IFD.plot, echo=FALSE}
-
-#make interference plots of IFD, position p1 by p2
-
-```
-
-```{r Comp_Ratios, echo=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-# Additional Determinants of Genome-wide Recombination Rate Evolution in Males
-
-The next section is meant to focusNext, we used the contrast between high-recombination strains (musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) to the low recombing strains.
-The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
-
-Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-**SC length is expected to be positively associated with genome-wide recombination rate because X**.
-Crossover interference is expected to be negatively **associated with genome-wide recombination rate because X**.
-Following this logic we predict (1) musculus^PWD^ will have greater SC length and weaker interference than musculus^SKIVE^, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) molossinus^MSM^ will have longer SC and weaker interference compared to molossinus^MOLF^, and 3) _domesticus_ strains will have similar SC length and crossover interference.
-
-# Q2 SC Length
-
-```{r Q2.SC.length.bret.comments, echo=FALSE}
-#correlation
-
-```
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
-```
-
-<!-- plots -->
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-
-#+theme(legend.position="none") 
-```
-
-```{r Q2.SC.long.biv_plots, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, mouse) %>%
-#  mutate(file.name = factor(file.name))
-
-mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-mouse.av_totsc_male
-
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE}
-
-#make plot of mouse averages for MLH1 and total SC
-
-bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
-
-X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-
-X.by.texas
-
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-```{r Q2.totSC.ttest, echo=FALSE}
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol
-Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.09
-
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-```{r Q2.totSC.glm, echo=FALSE}
-
-Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
-
-#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
-#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
-#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)
-
-Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
-#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
-
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
-#within subspecies values
-
-#Dom
-Q2.glm_mus.av_totSC_M2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
-sum.Q2.glm_mus.av_totSC_M2_DOM  <- summary(Q2.glm_mus.av_totSC_M2_DOM)
-#there are significant strain effects in DOM,
-
-#Musc
-Q2.glm_mus.av_totSC_M2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
-sum.Q2.glm_mus.av_totSC_M2_MUSC  <- summary(Q2.glm_mus.av_totSC_M2_MUSC)
-#KAZ -- is the only significant effect
-
-#Mol
-Q2.glm_mus.av_totSC_M2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
-sum.Q2.glm_mus.av_totSC_M2_MOL  <- summary(Q2.glm_mus.av_totSC_M2_MOL)
-
-#these pair-wise t.test aren't working
-#dom
-#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean.skel, 
-#                 ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$strain )
-
-#musc
-#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$mean.skel, 
-#                 ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$strain )
-
-#mol
-#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean.skel, 
-#                 ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$strain )
-
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-
-#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
-#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
-#table(mouse.av.long.biv.table$mouse)
-#Q2mouse.av.long.biv.table
-
-#this dataset is all males
-glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
-#
-
-glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_long.biv_M2_DOM <- summary(glmQ2_long.biv_M2_DOM)
-
-#MUSC
-glmQ2_long.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_long.biv_M2_MUSC <- summary(glmQ2_long.biv_M2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-glmQ2_long.biv_M2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
-sum.glmQ2_long.biv_M2_MOL <- summary(glmQ2_long.biv_M2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
-#DOM
-glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
-
-#MUSC
-glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
-
-#MOL
-#within subspecies
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
-#COMPARISONS ACROSS ALL CLASSES
-
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
-
-```
-
-```{r log.reg_SC.length, echo=FALSE, include=FALSE}
-#REMOVE THIS COMPARES FROM FULL BIV DATA
-
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
-#REMOVE THIS, IT uses total bivalent data
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
-```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-table( droplevels(fomo.male_mouse.av$strain) )
-#there are cast strains too
-
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
-
-#make the Mol and
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
-
-#Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-```
-
-```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
-## run the tests to check following expectations
-
-#1.Dom strains not different from ech other
-#2. PWD and SKIVE are different from the other strains
-#3.MSM and MOLF are different from each other
-
-#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
-
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
-#the subsp labels get messed up, unless they are designated chr first
-Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
-
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
-Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
-#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
-#total sc sig, 1CO sig, 2CO sig
-
-Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
-#0.000369 ***
-Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
-#0.000816 ***
-Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
-#0.0148 *
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
-#making plots above show that mouse average SC lengths are lower could there be microscope issue?
-
-#make sure to remove some of the juvi mice!! from the AP_mouse table
-show.Dom.male.CO <- ggplot(
-male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
-#P not sig for CO mean
-
-#seems like there are more mice in the MLH1 data set than bivData
-#
-
-#WSB all dissected and imaged in '15
-#G imaged 15 - 16
-#LEW imaged 15 to 17
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "1CO SC Length", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "2CO SC Length", xlab = "")
-```
-
-```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}
-
-#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
-#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
-
-male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
-male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
-
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
-#PWD and MSM are most diff from SKIVE?
-
-Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
-#the pooled SC value tests are sig
-
-#the mol comparisons are different due to adj p values
-
-Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
-#no sig (only 3) -- does include CZECCH
-
-#non-sig, with mol pooled with musc
-
-Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
-
-Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
-
-Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
-#non of these sig with the mulitple adjust
-
-Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
-
-#pairwise.t.test(Ozone, Month, p.adj = "bonf")
-```
-
-We compared mouse means of three SC-based metrics: total SC length, mean short bivalent length, and mean long bivalent length.
-The long bivalent data set was isolated by choosing bivalents in the X quartile of the SC length distribution within cells.
-The short bivalent dataset was isolated as described above.
-
-figure
-
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC **([report spearman's rho and p-value])**.
-Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
-
-While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, the means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value`; long: p = 0.18 -- `r Q2long.biv_t.test_High.Low$p.value`).
-
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
-In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-
-<bret's version above>
-
-# Q2.1CO rec landscape evolution is decoupled from gwRR evolution
-
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos All bivalents")
-Q2.show.nrmF1
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
-
-```
-
-```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->
-
-```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
-#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
-
-#High = MSM, PWD, SKIVE
-Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-#high = PWD and MSM
-Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-```
-
-```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
-
-#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
-
-#Musc
-Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
-# p-value = 0.4
-
-#Mol
-Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
-# p-value = 0.07
-# exclusive rec group not sig
-```
-
-```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
-#these don't have mean.first.foci
-
-#short -- real.long.bivData -- REAL.short_mouse_table
-Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
-
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
-#p-value = 0.3
-
-Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
-# p-value = 0.1
-
-#remove SKIVE??
-```
-
-```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
-
-#real.long.bivData
-#single biv level
-Rec.group_LONG.BivData_F1post_Ttest <- t.test(
-  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.5
-
-Rec.group_BivData_F1post_Ttest <- t.test(
-  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.2
-
-#
-#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
-#p-value = 0.2
-
-#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
-#p-value = 0.4
-
-```
-
-```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
-
-Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
-#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-#with.chrm.legth  ?
-Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-
-#M2
-Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
-
-```
-
-```{r Q2.1COpos.show.MOL, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
-
-mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
-                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-
-#+facet_wrap(~mouse, scales = "free")
-
-single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
-                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
-
-#+scale_color_manual(values=colors_of_strains)
-
-single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
-                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-
-#CHRM.length is very significant-- some of the same subsp and strains are sig     
-```
-
-```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
-#remove this cell,
-
-Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
-
-Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
-
-#making new models
-
-Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
-Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
-
-chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
-
-sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
-#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
-
-#summary(chrm.1CO.pos.chrmLength_strain)
-
-```
-
-```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
-```{r Q2.1COpos.glm, echo=FALSE, eval=FALSE}
-male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
-
-Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.Q2.F1_PER_Musc)
-
-#Mol
-Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
-              family=binomial(link="logit"))
-
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
-```
-
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = 0.24 -- `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = 0.41 -- `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = 0.07 -- `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
-While there are significant strain effects for domesticus^WSB^ and molossinus^MOLF^ in a general linear model with normalized position as the dependent variable (Figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
-
-# Q2 Evolution of interference is associated with genome wide recombination rate evolution
-
-```{r Q2.IFD.long.short.setup, echo=FALSE}
-#long BivData
-male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
-
-#show
-Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-
-#long biv's - short IFD rates
-
-LongBiv_IFD_mus.rate <- ddply(male.long.biv.2CO_IFD %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-
-                   N.obs  = length(Obj.ID),
-
-                  mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
-                  mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                  var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                  #mean_IFD_short.removed
-                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
-                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)] ),
-
-                inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                n.close.foci = length(which(IFD1_PER <= .3)),
-
-                off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                rate.close.foci = n.close.foci / N.obs
-
-                                              )
-
-#run t.tests on long biv
-LongBiv_IFD_mus.rate <- add_strain(LongBiv_IFD_mus.rate)
-LongBiv_IFD_mus.rate <- add_category(LongBiv_IFD_mus.rate)
-LongBiv_IFD_mus.rate <- add_subsp(LongBiv_IFD_mus.rate)
-
-#add rec.group
-LongBiv_IFD_mus.rate$Rec.group <- ifelse(grepl("PWD", LongBiv_IFD_mus.rate$strain), 1, 
-                                ifelse(grepl("MSM", LongBiv_IFD_mus.rate$strain), 1,
-                                 ifelse(grepl("SKIVE", LongBiv_IFD_mus.rate$strain), 1, 0)))
-
-```
-
-```{r Q2.IFD.long.biv_t.test, echo=FALSE}
-
-#short removed PER
-Short.removed_BivData.LongBiv_IFD.PER_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
-                                 LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
-
-Short.removed_BivData.LongBiv_IFD.PER_Ttest$p.value
-#0.0083
-
-#ABS short removed
-Short.removed_BivData.LongBiv_IFD.ABS_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
-                                 LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
-
-Short.removed_BivData.LongBiv_IFD.ABS_Ttest$p.value
-#0.047
-
-```
-
-```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
-
-#subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
-
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none", axis.text.x = element_blank())+ylim(c(0,1.1))
-
-```
-
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
-
-#t.test for IFD
-#mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
-
-#different test must be sed for the Dom
-
-#t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == #"male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
-#   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == #"Dom"))$mean_IFD.2CO_PER   )
-
-t.test_mouse.av.IFD_PER_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_PER   )
-
-t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER   )
-
-#ABS
-
-t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
-
-```
-
-```{r Q2.IFD.glm.test, echo=FALSE}
-#REPLACE Male.poly with ..
-#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
-```
-
-```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
-
-#repeating chrm models with chromsome length
-#FOR BIVALENT LEVEL MODELS
-
-Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
-
-#M.Len.ABS
-Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.raw)
-
-#M.Len.PER
-Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.PER)
-
-#M.Len.raw
-Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
-
-##M.Len PER
-Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-```
-
-```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
-
-#calculate the rates of 'close foci' (<30%)
-#examine the changes in the distribution closer.
-
-#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
-
-close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                      mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
-                      mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                  #mean_IFD_short.removed
-                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
-
-                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)], na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                      n.close.foci = length(which(IFD1_PER <= .3)),
-
-                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                      rate.close.foci = n.close.foci / N.obs
-)
-
-#mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-
-close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
-#add rec.group
-close.IFD.rates.mus.rate$Rec.group <- ifelse(grepl("PWD", close.IFD.rates.mus.rate$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates.mus.rate$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates.mus.rate$strain), 1, 0)))
-
-close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("strain"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm = TRUE),
-
-                        #mean_IFD_short.removed
-                      mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER > .3)),  
-                      n.close.foci = length(which(IFD1_PER < .3)),
-                      rate.close.foci = n.close.foci / N.obs,
-
-                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
-
-)
-
-close.IFD.rates_strain <- add_subsp(close.IFD.rates_strain)
-
-show.IFDrm <- ggplot(close.IFD.rates_strain, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences")+ geom_abline(slope = 1, intercept = 0)
-
-#close.IFD.rates.mus.rate
-show.IFDrm_mus <- ggplot(close.IFD.rates.mus.rate, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences") + geom_abline(slope = 1, intercept = 0)
-
-show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
-
-#add rec.group
-close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
-
-#POOL all short and 
-#chi square test
-high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
-                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
-#p-value = 0.03, slight dif tween 3 high rec strains
-
-#don't need to run chi square for this pattern, since it's enough to 
-```
-
-```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
-
-#Log regression all show significant prediction
-
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Mol)
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-
-#Full dataset
-
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-```
-
-```{r Q2.IFD_remove.short, echo=FALSE, include=FALSE}
-#mus
-#PER 
-#biv level
-
-#mus
-
-```
-
-```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
-
-male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-#t.tests for the Rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
-                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
-
-Rec.group.BivData_IFD.PER_Ttest$p.value
-#p-value < 2.2e-16
-
-#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
-Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
-                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
-
-Rec.group_IFD.PER_Ttest$p.value 
-#p-value = 6.872e-07
-
-```
-
-```{r Q2.IFD_short.removed_t.test, echo=FALSE, include=FALSE, eval=FALSE}
-
-#Curated_BivData_short.IFD.removed
-#still need to calq the mouse averages
-
-Short.removed_BivData_IFD.PER_Ttest <- t.test(Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 1], 
-                                 Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 0])
-
-Short.removed_BivData_IFD.PER_Ttest$p.value
-
-#
-Short.removed_BivData_IFD.PER_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
-                                 close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
-
-Short.removed_BivData_IFD.PER_Ttest_mus.av$p.value
-
-Short.removed_BivData_IFD.ABS_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
-                                 close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
-
-#p.value
-
-```
-
-Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
-This pattern is confirmed by separate comparisons within _musculus_ (t-test; IFD^norm^: p =2.0410^{-5} -- `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = 1.9410^{-4} -- `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
-
-That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
-The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (domesticus^G^) to 16% (musculus^KAZ^) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculus^SKIVE^, molossinus^MSM^, and musculus^PWD^, respectively).
-
-In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
-However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
-At least on 2CO bivalents, crossovers are spaced further apart when the genome-wide recombination rate is higher.
-
-# Q2 Summary
-
-Our results show that the greater crossover interference is the strongest single bivalent-based predictor for the observed rapid evolution of mean MLH1 foci per cell.
-While these results do not conform initial predictions for how a recombination landscapes would accommodate more crossovers, the increased strength of interference aligns with our results on sex differences.
-The typical recombination landscapes for males and females results in divergence in the proportion of linked sites along chromosomes which segregate together.
-The stronger interference of the 2CO bivalents in the high recombining strains accentuates this effect.
-The measures of DSB and some comparisons of SC length between high and low recombining strains suggest that the SC length have evolved to be longer in high recombining strains, however this evolution of SC length is partially decoupled from the number of crossovers since similar amounts of SC length evolution are seen in low recombining strains.
-
-<Bret: Are you sure this summary is necessary? Maybe cover this material in the Discussion instead?>
-
-# References
+# SAVE the data

### 2020-04-08 doc/Discussion.rmd
notes for consequences

@@
 
+# -- counter points, whats the consequence of reverse heterochiasmy (rapid male evolution in gwRR), 
+# -- sex-average / species average of gwRR is higher (for these strains)
+# --
+
 ```
@@
 #?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
+
+# consequence of greater female variance -- (between cell)
+#
 
@@
 
+# consequences of conserved typical landscape
+# (sardel kirkptarick - the rec landscape -- diviation from uniform is due to male rec landscape
+# (female CO suppresion -- DSB (repaired as NCO -- elevated DsB (halldorsson?))  )
+#
+
 ```
@@
 # only COM predicts that sexual dimorphism in axis length is due to cell vol
+
+# consequences of longer female axis - 
+# (normalzed interference strength due to axis differences / Chromtain loop length)
+#
 
@@
 Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
+
+# consequence of rapid evolution of normalized interference strength,
+# in the 3 strains -- larger linkage blocks?
+# but what if the rec pattern is repeated?
 

### 2020-04-09 doc/Results_v2.rmd
incorporated bret's commnets into rmd

@@
 
+<notes about figure; make sure the order highlights the difference (mol, musc, dom). Clean up the axis and titles. Resizing might be an issue too.>
+
 ```{r MLH1.summ.stats, echo=FALSE}
@@
 
+<in the above section reorder the strains by the effect sizes>
+
 Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
@@
 
-## Within mouse variance in CO count per cell
+# Within mouse variance in CO count per cell
 
@@
 In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (**criteria?**).
+Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (**criteria?
+Q1, see methods for certeria**).
+
+<Bret's comment for below, are there differences between the curated and the full dataset? he propbably means, Q1 and full data set.>
+
 The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2).
@@
 
-# Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
+<strain specific effect section was deleted for not highlighting main pattern. I agree>
+
+# DSB. Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
 
@@
 In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`; one-way-anova, p = 0.15 -- `r z.anova.pval`).
+**note that this interpertation depends on which stage is most relevant to CO number**
+
 After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
@@
 This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
+
+<Bret's Q, ANOVA is not appropriate for 2 samples unless it's specilaized somehow>
+
+<This result, combined with the big difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
+**It depends on whether leptonene or zygotene DSBs are viewed as more relevant for crossover numbers**.>
 
@@
 To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents ([@peterson2019]).
+
+<bret suggests moving this section to Mat and Methods>
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from
-`r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
+In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
+
 From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
@@
 We assume that the isolation of bivalents within cells is unbiased.
-Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns.
+<might be the only section to leave in Results>
+
+Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns. <Bret says this statment should be more clear>
+
 An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
+<Bret suggests moving to Methods>
 
@@
 In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male musculus^PWD^ and male molossinus^MSM^ should have longer SC lengths than the other strains.
+<moved section on reduced bivalent data set to methods>
 
@@
 The female:male ratio ranges across strains from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
+<moved section on total SC to methods>
 
@@
 
-There is no strong signal of sex differences in raw mean inter-focal distances (IFDraw) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticusG; p = 0.27).
+There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
+A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticus^G^; p = 0.27).
 This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculusKAZ showing no difference (t-test; p = 0.33).
-Examination of IFDnorm distributions indicates that female IFDnorm values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Mixed models and general linear models of IFDnorm support the inference of stronger interference in males: sex is the most significant variable (LRT: p = 6.7410^{-14}).
+In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculus^KAZ^ showing no difference (t-test; p = 0.33).
+Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (LRT: p = 6.7410^{-14}).
 When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+
 While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+<moved a section on relationship of linkage maps to Discussion (or Methods)>
 
@@
 
-The difference in mean IFDnorm are highly significant between sexes.
-The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculus^SKIVE^ p-value = 4.310^{-5}), musculus^PWD^ p-value = 2.2810^{-4}), molossinus^MSM^ p-value =0.01), domesticus^LEW^ p-value = 3.5710^{-5}), domesticusG p-value = 0.05), and domesticus^WSB^ (t.test; p-value =0.03) ).
-Only musculus^KAZ^ did not have a difference in mean IFDnorm (t.test; p-value = 0.33).
-In examining theExamination of IFDnorm distributions, we note indicates that  the female IFDnorm observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFDnorm observationsvalues, whereas males are enriched for higher values.
-Comparitively, the IDFnorm distributions in males are enriched for longer IFDnorm.
-The three sets of models Mixed models and general linear models of IFDnorm support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
-When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
+The difference in mean IFD^norm^ are highly significant between sexes.
+The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculus^SKIVE^ p-value = 4.310^{-5}), musculus^PWD^ p-value = 2.2810^{-4}), molossinus^MSM^ p-value =0.01), domesticus^LEW^ p-value = 3.5710^{-5}), domesticus^G^ p-value = 0.05), and domesticus^WSB^ (t.test; p-value =0.03) ).
+Only musculus^KAZ^ did not have a difference in mean IFD^norm^ (t.test; p-value = 0.33).
+In examining theExamination of IFDnorm distributions, we note indicates that  the female IFD^norm^ observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFD^norm^ observationsvalues, whereas males are enriched for higher values.
+Comparitively, the IDF^norm^ distributions in males are enriched for longer IFD^norm^.
+The three sets of models Mixed models and general linear models of IFD^norm^ support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
+When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
 In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
@@
 Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
+<Bret's note on positive correlation, 'A correlation across mice, across strains, or both?'>
 
@@
 In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-
-<bret's version above>
+<Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?>
 
@@
 The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = 0.24 -- `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = 0.41 -- `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = 0.07 -- `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
-While there are significant strain effects for domesticus^WSB^ and molossinus^MOLF^ in a general linear model with normalized position as the dependent variable (Figure X), this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.
+**While there are significant strain effects for domesticus^WSB^ and molossinus^MOLF^ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.* <bret asks if the bolded section is correct.
+are there sig strain effects for WSB and MOLF?> <bret asks if italised statement has been tested statisticaly or in what sense am I discribing this>
 
@@
 This pattern is confirmed by separate comparisons within _musculus_ (t-test; IFD^norm^: p =2.0410^{-5} -- `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = 1.9410^{-4} -- `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
+<Bret's comment, these p values are not very significant.> Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
 

### 2020-04-09 doc/Results_v2.rmd
working on results and shifting code to the src file

@@
 title: "MLH1 Results Outline"
-date: "null"
+date: 'null'
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document: default
   html_document:
@@
 
-<!-- 
-two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
--->
-
-<!--
-From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
-
-the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
-
-INSERT TABLE FOR THE TOTAL SC means ect
-
-The difference between
-
-bin is short for binary, it reports the binarized area of the SC
-
-skel -- takes the skeletonized midline of the bin objects
--->
-
 ```{r table.for.plot, echo=FALSE, warning=FALSE}
@@
 
+```{r bret.comment.var, echo=FALSE}
+
+```
+
 To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
@@
 
-We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4).
-In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = ~ 5 X foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect ~ 2 foci).
+We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4  **male specific, M2?
+rename the code for these models**).
+In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect =
+`r MLH1.male.specific.glm.M2$coefficients[5]` -- ~ 5 foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = `r MLH1.male.specific.glm.M2$coefficients[8]` -- ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect  `r MLH1.male.specific.glm.M2$coefficients[6]` ~ 2 foci).
 These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
@@
 
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
+```{r M2.MLH1.var, echo=FALSE, include=FALSE}
 #not including this model for now / brivity
@@
 
-MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M3)
-anova(MLH1var_M3)  
-as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))
-
-MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+#why are these named M3? -- rename to M2
+
+MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+
+MLH1var_M2_sum <- summary(MLH1var_M2)
+
+#summary(MLH1var_M2)
+#anova(MLH1var_M2)  
+#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
+
+MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
 
@@
 
-anova(MLH1cv_M3)
-as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))
+#anova(MLH1cv_M2)
+#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
 
 #Q12 var
-MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
 
 #summary(MLH1var_M3_Q12)
-anova(MLH1var_M3_Q12)  
-as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))
+#anova(MLH1var_M2_Q12)  
+#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
 
 #Q12 cv
-MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M3_Q12)
-anova(MLH1cv_M3_Q12)#sex large F
-
-as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))
-
+MLH1cv_M2_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M2_Q12)
+#anova(MLH1cv_M2_Q12)#sex large F
+#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
 ```
@@
 ```{r M3.MLH1.var, echo=FALSE}
-
 glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
@@
 
+#use the values from this model below
+
 MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
@@
 To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`).
+Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` (old) `r round(MLH1var_M2_sum$coefficients[,4][4], 5)`  correct).
+**the pland full M2 model is not significant**.
+M3 `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`.
+
 In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (**criteria?
-Q1, see methods for certeria**).
+Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
 
@@
 
-The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2).
+The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2 -- also M3).
 When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
@@
 z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
+
+#anova's aren't appropriate for 2 samples... re-do the anova
+#so that it's compareing more than just 2 categories -- or delete
+
 ```
@@
 The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`; one-way ANOVA, p = 0.00027 -- `r format(l.anova.pval, scientific=F)`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`; one-way-anova, p = 0.15 -- `r z.anova.pval`).
+In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`).
 **note that this interpertation depends on which stage is most relevant to CO number**

### 2020-04-09 src/Result_Setup.Rmd
working on results and shifting code to the src file

@@
 
+# MLH1 models
+
+M1 - subsp, sex, (1 | strain)
+M2 - subsp, sex, strain
+M3 - sex, strain
+M4 - strain (sex specific)
+
+#MLH1 variance models
+
+M1 - subsp, sex, (1 | strain)
+M2 - subsp, sex, strain
+M3 - sex, strain
+M4 - strain (sex specific)
+
+```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M1)
+LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
+
+LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
+sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
+
+```
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+MLH1var_M2_sum <- summary(MLH1var_M2)
+
+#summary(MLH1var_M2)
+#anova(MLH1var_M2)  
+#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
+
+MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+#summary(MLH1cv_M3)
+#anova(MLH1cv_M2)
+#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
+
+#Q12 var
+MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1var_M3_Q12)
+#anova(MLH1var_M2_Q12)  
+#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M2_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M2_Q12)
+#anova(MLH1cv_M2_Q12)#sex large F
+#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
+```
+
+```{r M3.MLH1.var, echo=FALSE}
+
+glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
+#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
+
+MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+#summary(MLH1cv_M4)
+#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
+
+### Q12 var
+MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
+
+#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
+#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
+
+#Q12 cv
+MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
+```
+
+```{r MLH1.var.M4}
+
+#male
+
+#female
+```
+
 # SAVE the data

### 2020-04-09 src/Result_Setup.Rmd
copying code chuncks for DMC1

@@
 
+# DMC1 Results
+
+main table (DMC1.mean.table), t.tests  , ratios
+
+set up, t.tests, correlations
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize,
+
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
+
+#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
+#            row.names = FALSE)
+
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+
+#make category averages from AP_mouse_table.HQ
+mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+#attach this mini DF to the DMC1 data
+
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+```{r DMC1.rand.calq, echo=FALSE}
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+```
+
+```{r DMC1.ratios, echo=FALSE}
+
+##RATIOS
+DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
+DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+
+#DMC1.mean.table
+DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
+
+colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
+                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
+
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
+
+kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
 # SAVE the data
+
+```{r Save.File, echo=FALSE}
+
+```

### 2020-04-09 doc/Discussion.rmd

@@
 
-<leaky oocyte SAC linked to acentrosome spindle (multiple aMTOCs) and cell volume, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>
+**if none of the models turn out to be appropriate,** it 
+could be due to assumptions we made regaurding....
 
-<Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
-(in oogenesis the leaky SAC has weak selection coefficient)>
+**We also note that** The musculus and molossinus subspecies have more shared ancestry (compared to domesticus) (cite Geraldes?), so we  cant determine if the rapid male evolution in gwrr are independant (mulitple events) or due to incomplete lineage sorting (single event).
+
+If these hypotheses turn out to not be true, (a prediction that can be made, due to the rapid male specific
+
+The consequences of this reversed heterochismy in house mice is that the sex-average / species average of gwRR is higher. meaning/resulting lower ld overall.
 
@@
 <Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big)review>
+
+If these hypotheses turn out to not be true, (a prediction that can be made, due to
+
+The consequences of this difference in meitic events -- (it fits follows from some basic differences in reproductive stategies between males and females) -- females diversify offspring, males --
+
+-- predictions for adaptations to changing environments -- but applied to short timescales -- since all the genes go back in the other sex the fllowing generation.
 
@@
 
+If these hypotheses concerning evolution of these patterns are not true -- there's still predictions that can be made given the consequences of these patterns.
+The consequences of this difference in the typical landscape as reviewed in [@sardell_sex_2020].
+For the sex averaged landscape along chromosomes, any diviation from uniform placement is generally due to male specific patterns.
+
+-- There's also the prediction that females have active CO supression near telomeres 
+(variation) -- geiven the evidence in (female mice) that show a higher elevation of DSBs at the telomere (halldorsson?).
+
 # Chromatin Structure
@@
 # (normalzed interference strength due to axis differences / Chromtain loop length)
-#
 
@@
 This prediction model might fit broader pattern such as in *Arabidopsis* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+
+ consequences of longer female axis - 
+ (normalzed interference strength due to axis differences / Chromtain loop length)
+
+We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
+
+According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
+**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.
+Only one of the models COM aligns with our observed results for sexual dimorphism for axis length.
+**COM** predicts that axis length will be larger in oocytes due to the larger volume.
 
@@
 
-<Still working on Future Steps>
+None of these models may be appropriate due to X Y and  Z caveats. The consequence of this pattern of greater interference strength is ---
+ consequence of rapid evolution of normalized interference strength, larger linkage blocks.
+
+#Future steps
+
+<importance of evolutionary distinct trajectories, measure both sexes -- do evolutionary comparisons within sex.
+
+wild samples, short times scales.
+
+single cell measures
+
+chromatid interference
+
+integrate physical models -- centromere and kinetochore
+
+>
 

### 2020-04-10 doc/Discussion.rmd
nots while reading morelli cohen

@@
 
-Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013]. <More details on the empirarcal pattern>
+Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013]. <More details on the empirarcal pattern> (cite morelli cohen)
 
@@
 This prediction model might fit broader pattern such as in *Arabidopsis* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+
+It also could be that the -- greater error in regulation of crossovers may have driven the longer female axis -- to esure / increase the probabilty of succesful maturation of DSB into crossover / chiasmata (cite morelli cohen)
+
+longer female axis has evol
 

### 2020-04-13 doc/Discussion.rmd
small changes for adding to discussion
rough rough draft for conseq of chromatin str

@@
 
+# significance, consequences
+# weak evidence of sex differences in chromatin organization, mice-humans have a similar direction -- but At opposite -- not much empiracal evidence. (more data needed)
+
+# a consequence of these sex differences for the chromatin organization is -- has an outcome of different effects on interference strenth measured in genetic map space -- for a given chromosome, chromatin organization of longer axis will result in shorter loops (emenating off the central axis)
+
+#1. interference signal travels through the SC / axis
+#since interference strength is mediated through the SC / chromatid axis
+# in both sexes interference strength discribed in units SC (axis) -- will be more conserved within and between species.
+
+#2. the tethering of loops to the axis is highly conserved - so variable chromatin organization for a given chromosome of constant size stems from varying the size of the loops and length of the axis.
+
+#3 when interference strength of a given unit of SC/axis / physical area -- is translated into base pairs (Mb), the chromosome organization with larger loops (smaller axis length) will have stronger interference -since more base pairs are passed/sequestered in the DNA loops per unit of SC/axis.
+
+#4.ppl should care -- this is an important aspect of -- how sexual dimorphism in recombination /linkage maps has been charaecterized -- but due to a lack of cytological evidence in both sexes /linking this to chromatin organization has only been done for a handful of species (mice, human, At?
+Dourad)
+
 ```
@@
 **COM** predicts that axis length will be larger in oocytes due to the larger volume.
+
+1. interference signal travels through the SC / axis
+since interference strength is mediated through the SC / chromatid axis
+ in both sexes interference strength discribed in units SC (axis) -- will be more conserved within and between species.
+
+2. the tethering of loops to the axis is highly conserved - so variable chromatin organization for a given chromosome of constant size stems from varying the size of the loops and length of the axis.
+
+3 when interference strength of a given unit of SC/axis / physical area -- is translated into base pairs (Mb), the chromosome organization with larger loops (smaller axis length) will have stronger interference -since more base pairs are passed/sequestered in the DNA loops per unit of SC/axis.
+
+4.ppl should care -- this is an important aspect of -- how sexual dimorphism in recombination /linkage maps has been charaecterized -- but due to a lack of cytological evidence in both sexes /linking this to chromatin organization has only been done for a handful of species (mice, human, At?
+Dourad)
 

### 2020-04-13 doc/Discussion.rmd
rough draft of caveats for typical landscape

@@
 # consequences of conserved typical landscape
-# (sardel kirkptarick - the rec landscape -- diviation from uniform is due to male rec landscape
-# (female CO suppresion -- DSB (repaired as NCO -- elevated DsB (halldorsson?))  )
+# most exceptions to the typical landscape, marsupials (pig and tomoato)
+
+#(sardel kirkptarick -
+
+# diviation from uniform is due to male rec landscape
+# patterns of Fst (Hanel et al, but see Kartej?)
+# source could be DSB formation along chromosome - and /or biases in DSB repair 
+#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick))  )
 #

### 2020-04-14 doc/Discussion.rmd
more work on chromatin str

--- 2020-04-10 doc/Discussion.rmd
@@
 
-**if none of the models turn out to be appropriate,** it 
-could be due to assumptions we made regaurding....
+0. Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.
 
-**We also note that** The musculus and molossinus subspecies have more shared ancestry (compared to domesticus) (cite Geraldes?), so we  cant determine if the rapid male evolution in gwrr are independant (mulitple events) or due to incomplete lineage sorting (single event).
+1. whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.
 
-If these hypotheses turn out to not be true, (a prediction that can be made, due to the rapid male specific
-
-The consequences of this reversed heterochismy in house mice is that the sex-average / species average of gwRR is higher. meaning/resulting lower ld overall.
+2.
+These results may hold importance to animal breeding programs -- may suggest that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most efficient way to increase the sex-averaged recombination rates.
+This could be important application for animal breeding programs (Battagin et al 2016).
 
@@
 
-If these hypotheses turn out to not be true, (a prediction that can be made, due to
+0. Reguardless of whether these models and hypothese are true ... if the pattern of females having more across cell variation in gwRR
 
-The consequences of this difference in meitic events -- (it fits follows from some basic differences in reproductive stategies between males and females) -- females diversify offspring, males --
-
--- predictions for adaptations to changing environments -- but applied to short timescales -- since all the genes go back in the other sex the fllowing generation.
+1.
+It has a consequence of fitting with distinct evolutionary streategies between the sexes of females diversifying the genetics their offspring and males limiting the shuffling of successful combinations (Trivers).
 
@@
 
-If these hypotheses concerning evolution of these patterns are not true -- there's still predictions that can be made given the consequences of these patterns.
+If these hypotheses concerning evolution of these patterns **are not true** -- there's still predictions that can be made given the consequences of these patterns.
 The consequences of this difference in the typical landscape as reviewed in [@sardell_sex_2020].
@@
 
+# significance, consequences
+# weak evidence of sex differences in chromatin organization, mice-humans have a similar direction -- but At opposite -- not much empiracal evidence. (more data needed)
+
+# a consequence of these sex differences for the chromatin organization is -- has an outcome of different effects on interference strenth measured in genetic map space -- for a given chromosome, chromatin organization of longer axis will result in shorter loops (emenating off the central axis)
+
+#4.ppl should care -- this is an important aspect of -- how sexual dimorphism in recombination /linkage maps has been charaecterized -- but due to a lack of cytological evidence in both sexes /linking this to chromatin organization has only been done for a handful of species (mice, human, At?
+Dourad)
+
 ```
@@
 
-longer female axis has evol
+0. There is weak evidence that female chromatin organization having a longer axis is a conserved trait. (only mice and humans, but there is a notable lack of data)
 
- consequences of longer female axis - 
- (normalzed interference strength due to axis differences / Chromtain loop length)
+1.
+We note that A consequence of sexual dmorphism is the chromatin axis (chromatin organization) is that interference strength^norm^ / (general patterns of sex specific genetic maps) will be sexually dimorphic.
 
-We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
+2. The density of tethering of loops to the axis is highly conserved (Zand Kleck) - 
+variablibilty of the axis length of a given chromosome stems from varing the loop sizes (Z n K)).
 
+3 When interference strength is translated into base pairs (Mb), the chromatin organization with larger loops will have stronger interference (Mb) since more base pairs are sequestered in the DNA loops per unit of SC/axis.
+
+5.
+while this prediction / the general pattern -- relationhsip of chromatin organization and interference strength is complicated -- the biology of chromatin (3D chromatin organization, heterogeneity of loop size within (cite) and among chromosomes (cite) and (limitations in most / for identifying indiviual chromosomes across single cells)
+
+(this prediction was outlined in Petkov, -- Lynn and Gruhn)
+
+It holds for several model organzism (mice, human, At) and sexual dimorphism in linkage maps (has been one of the oldest measurements of heterochiasmy --  This is an important aspect of how sexual dimorphism in recombination linkage maps has been charaecterized but a lack of cytological evidence for both sexes --
+
+--(so get the cytological data to test if this prediction holds)
+
+<We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
 According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
@@
 Only one of the models COM aligns with our observed results for sexual dimorphism for axis length.
-**COM** predicts that axis length will be larger in oocytes due to the larger volume.
+**COM** predicts that axis length will be larger in oocytes due to the larger volume.>
 
@@
 integrate physical models -- centromere and kinetochore
+>
 
->
+Diversity of systems for overcoming limitations
+
+1.
+identifying indiviual (autosomes) Chromosomes across cells: FISH (diffecult) or karyotypes with identifyable chromosomes (either into groups or largest)(birds, zfish?, peromyscus, Rbt mice, bovids?, humans, ), translocations (robertsonian translocations),
+
+2. non-model orgs can be used to study meiosis -- look at the weird ones (protezoa)
 

### 2020-04-14 doc/Discussion.rmd
more discussion writing. weird hyper link bug

@@
 <The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>
+
 Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
-This prediction model might fit broader pattern such as in *Arabidopsis* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+This prediction model might fit broader pattern such as in *Arabidopsis thalania* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
 
-It also could be that the -- greater error in regulation of crossovers may have driven the longer female axis -- to esure / increase the probabilty of succesful maturation of DSB into crossover / chiasmata (cite morelli cohen)
+A consequence of sexual dimorphism chromatin organization is that interference strength quantified in base pairs will be sexually dimorphic.
+There are two aspects that define this relationship.
+First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K)).
+Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg).
+Thus the chromatin organization with larger loops (and shorter axis) will have stronger crossover interference when quantified in base pairs, since more base pairs are sequestered in the chromatin loops per unit of axis or SC.
+Empirical data confirms that interference strength measured in micrometer along the axis is conserved between sexes while interference strength quantified from genetic map is sexually dimorphic, comes from mice [@petkov2007], humans (Gruhn, Lynn).
+However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.
 
-0. There is weak evidence that female chromatin organization having a longer axis is a conserved trait. (only mice and humans, but there is a notable lack of data)
-
-1.
-We note that A consequence of sexual dmorphism is the chromatin axis (chromatin organization) is that interference strength^norm^ / (general patterns of sex specific genetic maps) will be sexually dimorphic.
-
-2. The density of tethering of loops to the axis is highly conserved (Zand Kleck) - 
-variablibilty of the axis length of a given chromosome stems from varing the loop sizes (Z n K)).
-
-3 When interference strength is translated into base pairs (Mb), the chromatin organization with larger loops will have stronger interference (Mb) since more base pairs are sequestered in the DNA loops per unit of SC/axis.
-
-5.
-while this prediction / the general pattern -- relationhsip of chromatin organization and interference strength is complicated -- the biology of chromatin (3D chromatin organization, heterogeneity of loop size within (cite) and among chromosomes (cite) and (limitations in most / for identifying indiviual chromosomes across single cells)
-
-(this prediction was outlined in Petkov, -- Lynn and Gruhn)
-
-It holds for several model organzism (mice, human, At) and sexual dimorphism in linkage maps (has been one of the oldest measurements of heterochiasmy --  This is an important aspect of how sexual dimorphism in recombination linkage maps has been charaecterized but a lack of cytological evidence for both sexes --
-
---(so get the cytological data to test if this prediction holds)
-
-<We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.
-According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis.
-**The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.
-Only one of the models COM aligns with our observed results for sexual dimorphism for axis length.
-**COM** predicts that axis length will be larger in oocytes due to the larger volume.>
+<a partial pattern seen in Arabidopsis thalinia (Drouaud et al 2007) however there is no SC based measure of interference and low number of double crossovers in the female specific linkage map.
+It also could be that the greater error in regulation of crossovers may have driven the longer female axis to ensure increase the probability of successful maturation of DSB into crossover chiasmata>
 
@@
 
-```{r evo.interference, include=FALSE}
-#review results,
-#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
-#(we're gonna spend alot of space on this section)
-
-# more convincing that the pattern is real
-
-#lil background, for interferences 
-#assuptions for predictions (most modifer models, )
+```{r evo.interference, include=FALSE, echo=FALSE}
 
 #models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)
-
 #the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)
@@
 # but what if the rec pattern is repeated?
-
 ```

### 2020-04-15 doc/Results_v2.rmd
starting clean results section

@@
 
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`; one-way ANOVA, p = 0.00027 -- `r format(l.anova.pval, scientific=F)`).
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`).
 In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`).

### 2020-04-15 doc/Discussion.rmd
fixed wierd underline big, moved a comment out section

@@
     toc: yes
-    toc_depth: '5'
+    toc_depth: '3'
     toc_float: yes
@@
     toc: yes
-    toc_depth: '5'
+    toc_depth: '3'
 bibliography: refs_Main_Report.bib
@@
 #go through each traits and point out which models support each of the traits
-```
-
-```{r Model.testing, include=FALSE}
-
-#imortant of testable hypotheses
-
-#model not included in this comparison
-
 ```
@@
 
-0. Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.
+Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.
 
-1. whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.
+whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.
 
-2.
 These results may hold importance to animal breeding programs -- may suggest that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most efficient way to increase the sex-averaged recombination rates.
@@
 # consequence of greater female variance -- (between cell)
-#
-
 ```
 
-Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013]. <More details on the empirarcal pattern> (cite morelli cohen)
+Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013].
 
@@
 
-<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big)review>
+<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
 
-0. Reguardless of whether these models and hypothese are true ... if the pattern of females having more across cell variation in gwRR
+0. Reguardless of whether these models and hypothese are true if the pattern of females having more across cell variation in gwRR
 
@@
 
-#(sardel kirkptarick -
+#sardel kirkptarick -
 
@@
 # source could be DSB formation along chromosome - and /or biases in DSB repair 
-#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick))  )
-#
+#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick)
 
@@
 
-<The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.>
+<The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I>
 
-If these hypotheses concerning evolution of these patterns **are not true** -- there's still predictions that can be made given the consequences of these patterns.
+If these hypotheses concerning evolution of these patterns **are not true**  there's still predictions that can be made given the consequences of these patterns.
 The consequences of this difference in the typical landscape as reviewed in [@sardell_sex_2020].
@@
 
--- There's also the prediction that females have active CO supression near telomeres 
-(variation) -- geiven the evidence in (female mice) that show a higher elevation of DSBs at the telomere (halldorsson?).
+There's also the prediction that females have active CO supression near telomeres (variation) -- geiven the evidence in (female mice) that show a higher elevation of DSBs at the telomere (halldorsson?).
 
@@
 
-<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>
+<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]>
 
@@
 There are two aspects that define this relationship.
-First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K)).
+First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K).
 Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg).
@@
 However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.
-
-<a partial pattern seen in Arabidopsis thalinia (Drouaud et al 2007) however there is no SC based measure of interference and low number of double crossovers in the female specific linkage map.
-It also could be that the greater error in regulation of crossovers may have driven the longer female axis to ensure increase the probability of successful maturation of DSB into crossover chiasmata>
 
@@
 
-<importance of evolutionary distinct trajectories, measure both sexes -- do evolutionary comparisons within sex.
+importance of evolutionary distinct trajectories, measure both sexes -- do evolutionary comparisons within sex.
 
@@
 integrate physical models -- centromere and kinetochore
->
 
@@
 
+<a partial pattern seen in Arabidopsis thalinia (Drouaud et al 2007) however there is no SC based measure of interference and low number of double crossovers in the female specific linkage map.
+It also could be that the greater error in regulation of crossovers may have driven the longer female axis to ensure increase the probability of successful maturation of DSB into crossover chiasmata>
+
 #References

### 2020-04-15 doc/Discussion.rmd
small changes, spell check

@@
 ```
+
+other COM, / nucleus membrane citations (Harper, Golubovskaya, & Cande, 2004; Naranjo & Corredor, 2008; Scherthan et al., 1996; Zickler & Kleckner, 2016).
+
+telomere guided movements  citations: (Bass et al.,
+2000; Chacon, Delivani, & Tolic, 2016; Curtis, Lukaszewski, & Chrzastek, 1991; Ding, Yamamoto, Haraguchi, & Hiraoka, 2004; Gerton &
+Hawley, 2005; Lee, Conrad, & Dresser, 2012; Lefrancois, Rockmill,
+Xie, Roeder, & Snyder, 2016; Page & Hawley, 2003)
+
+Repair of DSB - CO
+(Anderson & Stack, 2005; Bass et al., 2000; Brown et al., 2005; Croft & Jones, 1989; Higgins, Osman, Jones, & Franklin, 2014; Klutstein & Cooper, 2014, Lukaszewski, 1997; Pratto et al., 2014; Viera, Santos, & Rufas, 2009;
+Xiang, Miller, Ross, Alvarado, & Hawley, 2014).
 
@@
 
-# typical landscape is conserved -- sexually dimorphic meiotic trait (Sardell)
-
-# models
-
 #<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
@@
 
-#*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.
+#0. we find the conserved pattern of the typical landscape 
+#this trait is a highly conserved aspect of heterochiasmy 
+#the exceptions include marsupials which form sex chromosome chains during meiosis -- indicating a potential change to the meiotic program (functional selection)
+#it's not known if a similar -- evolutionary force in the other exceptions to the typical landscape, pig and tomatoe.
 
-# consequences of conserved typical landscape
-# most exceptions to the typical landscape, marsupials (pig and tomoato)
+#regaurdless
+#As reviewed in sardel kirkptarick - diviation for a uniform background is generally due to the male recombination landscape.
 
-#sardel kirkptarick -
+# these pattern due to (different recombination landscapes along chromosomes can have an impact on the pattern of Fst (Hanel et al 2018)
+# additionally other patterns of nucleotide and recombination rates could vary across
 
-# diviation from uniform is due to male rec landscape
+#co rate defines the physical scale of linked selection around a locus
+
+# evolutionary genomics from the broad scale heterogenity
+#linked selection, (maynarid smith and haig 1974)
+#background selection (only with negtive allele), genetic draft (for positive selective mutations)
+#gene flow barrier (directional selection is keeping two populations from merging)
 # patterns of Fst (Hanel et al, but see Kartej?)
+
+#
+
 # source could be DSB formation along chromosome - and /or biases in DSB repair
+
 #CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick)
@@
 This prediction model might fit broader pattern such as in *Arabidopsis thalania* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+
+(also outlined in Hanel et al 2018 -- which connects to typical landscape)
 

### 2020-04-15 doc/Discussion.rmd
few more thoughts

@@
 
-#
+# Hanel et al predicts -- lower genetic diversity near centers -- away from chromosome ends (this is supported )
+#finds that the strenght of telomere is connected to chrm size (positive stronger) -- independant of cenromere
+
+#as first outlined in petkov et al and restated in Haenel et al, -- chromatin organization
 
@@
 
-(also outlined in Hanel et al 2018 -- which connects to typical landscape)
+As outlined in Hanel et al 2018, sexual dimorphism in chromatin organization translates to interference strength in base pairs.
+Since more DNA is sequestered into the radial loops per unit of the central axis, in the chromosome organization with longer DNA loops, as seen in males for mice (Lynn, gruhn) and humans (petkov), interference strength measured in base pairs will be stronger compared to female linkage maps.
+The signal for interference is mediated though the chromosome axis (Zickler Kleckner), interference strength measured in micrometers of is conserved between sexes [@petkov2007].
 
-A consequence of sexual dimorphism chromatin organization is that interference strength quantified in base pairs will be sexually dimorphic.
-There are two aspects that define this relationship.
-First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K).
-Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg).
-Thus the chromatin organization with larger loops (and shorter axis) will have stronger crossover interference when quantified in base pairs, since more base pairs are sequestered in the chromatin loops per unit of axis or SC.
-Empirical data confirms that interference strength measured in micrometer along the axis is conserved between sexes while interference strength quantified from genetic map is sexually dimorphic, comes from mice [@petkov2007], humans (Gruhn, Lynn).
-However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.
+(hanel makes connections to the typical (telomere-bias) --(stronger interference pushes the crossovers to the ends)
+
+<just outlined a consequence of sexual dimorphism in chromatin organization... but this isn't conserved as a
+
+-- not confirmed - if this is a conserved pattern / 
+pattern - (general pattern for linkage maps, but there's a lack of cytological data for both sexes)
+
+-- the role of chromatin organization in recombination variation landscapes -- should be studied/ (needs more support)
 

### 2020-04-15 doc/Discussion.rmd
added to bib

@@
 
-As outlined in Hanel et al 2018, sexual dimorphism in chromatin organization translates to interference strength in base pairs.
-Since more DNA is sequestered into the radial loops per unit of the central axis, in the chromosome organization with longer DNA loops, as seen in males for mice (Lynn, gruhn) and humans (petkov), interference strength measured in base pairs will be stronger compared to female linkage maps.
-The signal for interference is mediated though the chromosome axis (Zickler Kleckner), interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-
-(hanel makes connections to the typical (telomere-bias) --(stronger interference pushes the crossovers to the ends)
-
-<just outlined a consequence of sexual dimorphism in chromatin organization... but this isn't conserved as a
-
--- not confirmed - if this is a conserved pattern / 
-pattern - (general pattern for linkage maps, but there's a lack of cytological data for both sexes)
-
--- the role of chromatin organization in recombination variation landscapes -- should be studied/ (needs more support)
+As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength in base pairs.
+Since more DNA is sequestered into the radial loops per unit of the central axis, in the chromosome organization with longer DNA loops, as seen in males for mice [@petkov2007, @lynn2002], and humans [@gruhn2013], interference strength measured in base pairs will be stronger compared to female linkage maps.
+The signal for interference is mediated though the chromosome axis [@zicklerKleckner2015], interference strength measured in micrometers of is conserved between sexes [@petkov2007].
+The wide spread pattern of stronger interference strength in Mb (and cm/Mb) for male specfic linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
 

### 2020-04-15 doc/Discussion.rmd
cleaning up the writing

@@
 
-other COM, / nucleus membrane citations (Harper, Golubovskaya, & Cande, 2004; Naranjo & Corredor, 2008; Scherthan et al., 1996; Zickler & Kleckner, 2016).
-
-telomere guided movements  citations: (Bass et al.,
-2000; Chacon, Delivani, & Tolic, 2016; Curtis, Lukaszewski, & Chrzastek, 1991; Ding, Yamamoto, Haraguchi, & Hiraoka, 2004; Gerton &
-Hawley, 2005; Lee, Conrad, & Dresser, 2012; Lefrancois, Rockmill,
-Xie, Roeder, & Snyder, 2016; Page & Hawley, 2003)
-
-Repair of DSB - CO
-(Anderson & Stack, 2005; Bass et al., 2000; Brown et al., 2005; Croft & Jones, 1989; Higgins, Osman, Jones, & Franklin, 2014; Klutstein & Cooper, 2014, Lukaszewski, 1997; Pratto et al., 2014; Viera, Santos, & Rufas, 2009;
-Xiang, Miller, Ross, Alvarado, & Hawley, 2014).
-
 We performed hypothesis testing of models from the literature and adapted here fit the main results from this study.
@@
 <We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversible metaphase-to-anaphase transition (cite).
-A stable chromosome structure, known as bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite).
+A stable chromosome structure, through bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite).
 Unstable bivalents on the meiosis I spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle.
@@
 
-Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.
-
-whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.
-
-These results may hold importance to animal breeding programs -- may suggest that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most efficient way to increase the sex-averaged recombination rates.
-This could be important application for animal breeding programs (Battagin et al 2016).
+Regardless of if these hypotheses turn out to be true, our results extend previous reports [@dumont2011] of rapid male-specific evolution in multiple subspecies and strains of house mouse.
+Whether these instances of evolution are due to separate independent events or segregating genetic variation within the population requires further study.
+The faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
+This insight could be important for animal breeding programs [@battagin2016].
 
@@
 
-Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013].
+Our results of greater between cell variation in females are supported by empirical results from human and mouse [@lynn2002, @gruhn2013].
 

### 2020-04-15 doc/Discussion.rmd
cleaning up sentences

@@
 Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate or cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, Subrahman).
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis ([@laneKauppi2019], Subrahman).
 These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite).
@@
 
-Regardless of if these hypotheses turn out to be true, our results extend previous reports [@dumont2011] of rapid male-specific evolution in multiple subspecies and strains of house mouse.
-Whether these instances of evolution are due to separate independent events or segregating genetic variation within the population requires further study.
-The faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
+Our results extend previous reports [@dumont2011] of rapid male-specific evolution in multiple subspecies and strains of house mouse.
+Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within the population requires further study.
+Regaurdless, the faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
 This insight could be important for animal breeding programs [@battagin2016].
@@
 Our results of greater between cell variation in females are supported by empirical results from human and mouse [@lynn2002, @gruhn2013].
-
 While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
@@
 This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals.
-Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (Schuh Ellenburg 2007).
+Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
 This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite).
@@
 
-0. Reguardless of whether these models and hypothese are true if the pattern of females having more across cell variation in gwRR
-
-1.
-It has a consequence of fitting with distinct evolutionary streategies between the sexes of females diversifying the genetics their offspring and males limiting the shuffling of successful combinations (Trivers).
+Regardless of whether these models and hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells.
+This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
 
@@
 
-#as first outlined in petkov et al and restated in Haenel et al, -- chromatin organization
-
 # source could be DSB formation along chromosome - and /or biases in DSB repair 
-
 #CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick)
-
 ```
 
-Our results on the broad scale pattern of crossover placement are supported by the literature.
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
-The two locus modifier and SACE model predicts sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
+As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait, which our results support.
+The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
 
-The two mechanical models both predict sexual dimorphism in the recombination landscape for distinct reasons.
 The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM, @rubin2020mixing].
@@
 
-If these hypotheses concerning evolution of these patterns **are not true**  there's still predictions that can be made given the consequences of these patterns.
-The consequences of this difference in the typical landscape as reviewed in [@sardell_sex_2020].
-For the sex averaged landscape along chromosomes, any diviation from uniform placement is generally due to male specific patterns.
-
-There's also the prediction that females have active CO supression near telomeres (variation) -- geiven the evidence in (female mice) that show a higher elevation of DSBs at the telomere (halldorsson?).
+A consequence of the sexual dimorphism in broad scale recombination landscapes is that the male recombination pattern drives divation from an uniform distribution in any sex averaged recombination landscape [@sardell_sex_2020].
+Elevation at the chromosome ends is a conserved trait across animals and plants and has a consequence of higher genetic diversity near chromosome ends [@haenel2018].
 
@@
 As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength in base pairs.
-Since more DNA is sequestered into the radial loops per unit of the central axis, in the chromosome organization with longer DNA loops, as seen in males for mice [@petkov2007, @lynn2002], and humans [@gruhn2013], interference strength measured in base pairs will be stronger compared to female linkage maps.
+Since more DNA is sequestered into the radial loops per unit of the central axis, chromatin organization with longer DNA loops, as seen in males for mice [@petkov2007, @lynn2002] and humans [@gruhn2013], interference strength measured in base pairs will be stronger compared to female linkage maps.
 The signal for interference is mediated though the chromosome axis [@zicklerKleckner2015], interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-The wide spread pattern of stronger interference strength in Mb (and cm/Mb) for male specfic linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
+Stronger interference strength in basepairs for male specfic linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
 
@@
 
-None of these models may be appropriate due to X Y and  Z caveats. The consequence of this pattern of greater interference strength is ---
- consequence of rapid evolution of normalized interference strength, larger linkage blocks.
+An expected consequence of our observed elevated genome-wide recombination rate is lower LD, which may increase the efficiency of selection and impact the dynamics of intrgresion.
+However given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogenous signal along chromosomes.
 
@@
 
-importance of evolutionary distinct trajectories, measure both sexes -- do evolutionary comparisons within sex.
+We make three suggestions for future steps in studying recombination variation.
+First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
+<while comparisons of traits across sex -are interesting - -- some comparisons within sex are more appropriate for some evolutionary comparisons.>
 
-wild samples, short times scales.
+Second recognize that certain systems can be leveraged to overcome limitations inherent in some approached.
+For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
+FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian translocations.
 
-single cell measures
-
-chromatid interference
-
-integrate physical models -- centromere and kinetochore
-
-Diversity of systems for overcoming limitations
-
-1.
-identifying indiviual (autosomes) Chromosomes across cells: FISH (diffecult) or karyotypes with identifyable chromosomes (either into groups or largest)(birds, zfish?, peromyscus, Rbt mice, bovids?, humans, ), translocations (robertsonian translocations),
-
-2. non-model orgs can be used to study meiosis -- look at the weird ones (protezoa)
+Third we promote the cross-pollination of physiology based models and more abstract model such as population genetic models with testablle hypotheses (Dapper Paysuer).
+A goal of such merging of models could be to connect empiracal findings across scales -- for example fine scale hotspot data to chromosome wide variation in recombination patterns (Keeney).
 
@@
 
-<a partial pattern seen in Arabidopsis thalinia (Drouaud et al 2007) however there is no SC based measure of interference and low number of double crossovers in the female specific linkage map.
-It also could be that the greater error in regulation of crossovers may have driven the longer female axis to ensure increase the probability of successful maturation of DSB into crossover chiasmata>
-
 #References

### 2020-04-15 doc/Discussion.rmd
doc/Discussion.html

@@
 
-We performed hypothesis testing of models from the literature and adapted here fit the main results from this study.
-We choose to examine three modifier models and two functional and cell physiology models (table X).
-It is a bit challenging to test all of these models together since they were built to describe different patterns of recombination variation.
-The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the COM model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
+We examine three modifier models and two functional and cell physiology models (table X) to test how well existing models fit out results.
+While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns across models.The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the COM model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
 In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
@@
 
-Female biased heterochiasmy is the most common pattern observed, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
-Our results of male biased heterochiasmy is novel due to the shorter evolutionary distances between the examined strains.
-
+In our results and more generally in the literature, female biased heterochiasmy is the most common pattern observed, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
+Our results of male biased heterochiasmy in 2 strains is a novel finding due to the shorter evolutionary distances between the examined strains.
 The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) that the driver acts (see Table 1 of @brandvain2012scrambling).
@@
 Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversible metaphase-to-anaphase transition (cite).
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue, (bi-orientated) spindle before starting the irreversible metaphase-to-anaphase transition (cite).
 A stable chromosome structure, through bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite).
@@
 Our results extend previous reports [@dumont2011] of rapid male-specific evolution in multiple subspecies and strains of house mouse.
-Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within the population requires further study.
-Regaurdless, the faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
+Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within house mouse subspecies requires further study.
+Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
 This insight could be important for animal breeding programs [@battagin2016].
@@
 
-A consequence of the sexual dimorphism in broad scale recombination landscapes is that the male recombination pattern drives divation from an uniform distribution in any sex averaged recombination landscape [@sardell_sex_2020].
+A consequence of the sexual dimorphism in broad scale recombination landscapes is that the male recombination pattern drives deviation from an uniform distribution in any sex averaged recombination landscape [@sardell_sex_2020].
 Elevation at the chromosome ends is a conserved trait across animals and plants and has a consequence of higher genetic diversity near chromosome ends [@haenel2018].
@@
 The signal for interference is mediated though the chromosome axis [@zicklerKleckner2015], interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-Stronger interference strength in basepairs for male specfic linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
+Stronger interference strength in base pairs for male specific linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
 
@@
 
-An expected consequence of our observed elevated genome-wide recombination rate is lower LD, which may increase the efficiency of selection and impact the dynamics of intrgresion.
-However given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogenous signal along chromosomes.
+An expected consequence of our observed elevated genome-wide recombination rate is lower LD, which may increase the efficiency of selection and impact the dynamics of introgression.
+However given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
 
@@
 For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
-FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian translocations.
+FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian trans-locations.
 
-Third we promote the cross-pollination of physiology based models and more abstract model such as population genetic models with testablle hypotheses (Dapper Paysuer).
-A goal of such merging of models could be to connect empiracal findings across scales -- for example fine scale hotspot data to chromosome wide variation in recombination patterns (Keeney).
+Third we promote the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses (Dapper Paysuer).
+A goal of such merging of models could be to connect empirical findings across scales.
+For example fine scale hotspot data to chromosome wide variation in recombination patterns (Keeney).
 

### 2020-04-15 doc/Discussion.rmd
more sentence changes and adding citations

@@
 
-Our results of greater between cell variation in females are supported by empirical results from human and mouse [@lynn2002, @gruhn2013].
-While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.
+<While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.>
 
@@
 
-Regardless of whether these models and hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells.
+Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013].
 This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
@@
 
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait, which our results support.
+As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
 The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
 
-The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM, @rubin2020mixing].
+The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM] <@rubin2020mixing>.
 
@@
 These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
-While the relationship may hold for more general heterochiasmy patterns across broad scale comparisons, these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction.
+While the relationship may hold for more general heterochiasmy patterns across broad scale comparisons, our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
+females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation.
 
@@
 Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
-The general pastern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
+The general pattern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
 This fits the logic of the chromosome axis acting as the 'real estate' for DSBs and aligns with known molecular mechanisms (cite).
@@
 
-We propose that the **spindle based selection model** can explain the positive correlation and evolution of stronger interference strength via selection on the amount of sister cohesion connecting homologs (figure X, table X).
-The evolution of genome wide recombination rates in our study is driven by a transition of recombination landscapes from single crossovers to two crossovers.
+We propose that the **spindle based selection model** can explain the positive correlation via selection on the amount of sister cohesion connecting homologs.
+The evolution of genome wide recombination rates in our study is driven by a transition of the majority of
+bivalent having single crossovers to an enrichment of bivalents with two crossovers.
 All else being equal, interference strength would be expected to be equal or even weaker.
-Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain with more two crossover chromosomes.
+Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain enriched for two crossover bivalents.
 Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
 This model gives rise to the possibility that the chromosome structure at metaphase is the target of selection  which is achieved by changing the average recombination landscape across bivalents within the cell.
-This interpretation has potential to fit with hypotheses mentioned above for explaining other results regarding the evolution of the recombination landscape.
 
 The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, however it lacks evolutionary based predictions.
-While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.
+While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
 
-We have focused on the model involving direct selection on the recombination pathway, that is the impact on the meiotic process on an individual, but we make note on models centering on indirect selection or fitness outcomes of offspring which could also explain our results.
-Models from Goldstein et al, reviewed in [@ottoPaysuer2019] and Veller 2017, predict that interference strength evolves whenever increased recombination rates evolve.
+We have focused on the model involving direct selection on the recombination pathway, which has an impact on the fitness of an individual, but we acknowledge connections to other models involving indirect selection on recombination rates which involve the fitness outcomes of offspring.
+A model from Goldstein et al, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
 The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
-A distinguishing feature of their model is that the number of crossovers per chromosome is kept constant.
-The range of crossovers per chromosome is quite small, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017].
+A distinguishing feature of this model from other modifer models is that the number of crossovers per chromosome is kept constant.
+Empiracally, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017] across most taxa.
 Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.

### 2020-04-16 doc/Discussion.rmd
small changes

@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '3'
+  pdf_document: default
   html_document:
@@
     toc_float: yes
-  pdf_document: default
-  word_document:
-    toc: yes
-    toc_depth: '3'
 bibliography: refs_Main_Report.bib

### 2020-04-16 doc/Discussion.rmd
small changes

--- 2020-04-15 doc/Discussion.rmd
@@
 
-<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
+<Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
 

### 2020-04-16 doc/Discussion.rmd
adding in edits

--- 2020-04-16 doc/Discussion.rmd
@@
 
-We examine three modifier models and two functional and cell physiology models (table X) to test how well existing models fit out results.
-While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns across models.The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the COM model was built to describe the interference pattern and difference in the recombination landscape between the sexes.
+To place our results in the broader context of meiotic variation we examine three modifier models and two functional and cell physiology models (table X).
+While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across exsisting models.
+The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome ocillatory movement (COM) model was built to describe positive interference.
+This spindle based selection model was developed to explain the results resented in this manuscript.
 In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
@@
 
-In our results and more generally in the literature, female biased heterochiasmy is the most common pattern observed, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
-Our results of male biased heterochiasmy in 2 strains is a novel finding due to the shorter evolutionary distances between the examined strains.
-The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) that the driver acts (see Table 1 of @brandvain2012scrambling).
+The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) when the driver acts (see Table 1 of @brandvain2012scrambling).
 
 Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue, (bi-orientated) spindle before starting the irreversible metaphase-to-anaphase transition (cite).
-A stable chromosome structure, through bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite).
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue, (bi-orientated) spindle before starting the metaphase-to-anaphase transition ([@laneKauppi2019], subramanian and Hochwagen , @dumontDesai2012).
+
+A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion [@laneKauppi2019].
 Unstable bivalents on the meiosis I spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle.
@@
 Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis ([@laneKauppi2019], Subrahman).
-These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite).
+These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
 This model nominates conserved differences in in the context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
 
-Our results extend previous reports [@dumont2011] of rapid male-specific evolution in multiple subspecies and strains of house mouse.
+In our results and more generally in the literature, female biased heterochiasmy is the most common pattern, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
+Our results extend previous reports of male biased heterochiasmy [@dumont2011] to  new subspecies of house mouse (*molossinus*) and emphasize the short evolutionary period.
 Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within house mouse subspecies requires further study.
-Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
+Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species in female biased heterochiasmy species, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
 This insight could be important for animal breeding programs [@battagin2016].
@@
 
-<While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.>
-
 In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
@@
 Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite).
+Centrosomes spindles are in spermatogensis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization (@manandhar2005centrosome).
+This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
 
@@
 The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
-
-The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM] <@rubin2020mixing>.
+The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM].
 
 Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate (cite).
-Different timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures on sister cohesion.
-Faster spermatogenesis may select for synchronization of the separation homologs.
-While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle.
-
-<The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I>
+The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs [@laneKauppi2019], subramanian and Hochwagen , @dumontDesai2012].
+The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
+Differences in timing of cell cycle between oogenesis and spermatogenesis imposes the the different selective pressures on how sister cohesion affects chromosome structure.
+Faster spermatogenesis may select for synchronization of the separation homologs (cite?).
+While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, espicially for those with dicyate arrest [@lee2019].
 
@@
 
-Our results from musculus^MSM^ and musculus^PWD^ show that the SC length and chromatin compaction is uncoupled from the direction of heterochiasmy.
-These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
-While the relationship may hold for more general heterochiasmy patterns across broad scale comparisons, our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
+Our results from musculus^MSM^ and musculus^PWD^ demonstrate that the SC length/chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
+These results slightly depart from predictions which nominate chromatin compaction/organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
+
+The relationship may hold for more general heterochiasmy patterns across sexes but fail to fully explain rapid evolution in the sex specific context.
+(this can be attributed to distinct evolutionary trajectories due to sex specific meiotic context)
+
+Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
 females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation.
-
-<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]>
 
@@
 Stronger interference strength in base pairs for male specific linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
+
+<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]>
 

### 2020-04-16 doc/Discussion.rmd
fixing a few more sentences

@@
 
-Our results from musculus^MSM^ and musculus^PWD^ demonstrate that the SC length/chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
-These results slightly depart from predictions which nominate chromatin compaction/organization (axis length) as the primary driver of heterochiasmy [@petkov2007].
+Our results from musculus^MSM^ and musculus^PWD^ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
+These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
+Chromatin compaction could explain variation in heterochiasmy, which is driven by the disinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
+Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
+females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation (see below).
 
-The relationship may hold for more general heterochiasmy patterns across sexes but fail to fully explain rapid evolution in the sex specific context.
-(this can be attributed to distinct evolutionary trajectories due to sex specific meiotic context)
-
-Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
-females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation.
-
-Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.
+Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
 This prediction model might fit broader pattern such as in *Arabidopsis thalania* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
 
-As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength in base pairs.
-Since more DNA is sequestered into the radial loops per unit of the central axis, chromatin organization with longer DNA loops, as seen in males for mice [@petkov2007, @lynn2002] and humans [@gruhn2013], interference strength measured in base pairs will be stronger compared to female linkage maps.
+As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
+Since more DNA is sequestered into the radial loops per unit of the central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males for mice [@petkov2007, @lynn2002] and humans [@gruhn2013], interference strength measured in base pairs will be stronger compared to female linkage maps.
 The signal for interference is mediated though the chromosome axis [@zicklerKleckner2015], interference strength measured in micrometers of is conserved between sexes [@petkov2007].
@@
 
-We propose that the **spindle based selection model** can explain the positive correlation via selection on the amount of sister cohesion connecting homologs.
+We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphse.
 The evolution of genome wide recombination rates in our study is driven by a transition of the majority of
@@
 Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
-This model gives rise to the possibility that the chromosome structure at metaphase is the target of selection  which is achieved by changing the average recombination landscape across bivalents within the cell.
 
-The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, however it lacks evolutionary based predictions.
-While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
+The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, however it lacks a mechanism for a evolutionary based predictions.
+While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
 
@@
 A distinguishing feature of this model from other modifer models is that the number of crossovers per chromosome is kept constant.
-Empiracally, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017] across most taxa.
+Empiracally, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
 Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
@@
 An expected consequence of our observed elevated genome-wide recombination rate is lower LD, which may increase the efficiency of selection and impact the dynamics of introgression.
-However given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
+Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
 
@@
 First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
-<while comparisons of traits across sex -are interesting - -- some comparisons within sex are more appropriate for some evolutionary comparisons.>
-
-Second recognize that certain systems can be leveraged to overcome limitations inherent in some approached.
+Second certain systems can be leveraged to overcome limitations inherent in some approached.
 For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
@@
 
-Third we promote the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses (Dapper Paysuer).
+Third we encourage the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses (Dapper Paysuer).
 A goal of such merging of models could be to connect empirical findings across scales.
-For example fine scale hotspot data to chromosome wide variation in recombination patterns (Keeney).
 

### 2020-04-16 doc/Discussion.rmd
changed headers

@@
 ---
-title: "Discussion Draft"
+title: "Discussion Draft v1"
 date: ''
@@
 
-To place our results in the broader context of meiotic variation we examine three modifier models and two functional and cell physiology models (table X).
+To place our results in the broader context of recombination rate variation we examine three modifier models and two functional and cell physiology models (table X).
 While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across exsisting models.
@@
 
-# Reversed heterochiasmy direction
+### Evolution of reversed heterochiasmy direction
 
@@
 
-# Greater between cell variation
+### Conservation in sexual dimorphism in between cell variation in crossover number
 
@@
 
-# Typical landscape
+### Conservation in sex-specific recombination landscape
 
@@
 
-# Chromatin Structure
+### Conservation of sex differences in chromatin structure
 
@@
 
-# Evolution of Interference
+### Evolution of interference strength in males
 
@@
 
-#Future steps
+### Future steps
 
@@
 
-#References
+# References

### 2020-04-16 doc/Discussion.rmd
slight change of headers

@@
 ---
-title: "Discussion Draft v1"
+title: ""
 date: ''
@@
 ---
+
+# Discussion Draft v1
 

### 2020-04-16 doc/Discussion.rmd
small edits

@@
 
-To place our results in the broader context of recombination rate variation we examine three modifier models and two functional and cell physiology models (table X).
+To place our results in the broader context of recombination rate evolution we compare our main resuts to expectionat under three modifier models and two functional and cell physiology models (table X).
 While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across exsisting models.
@@
 
-The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction or equatorial division) when the driver acts (see Table 1 of @brandvain2012scrambling).
+The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction (MI) or equatorial division (MII)) when the driver acts (see Table 1 of @brandvain2012scrambling).
 
 Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue, (bi-orientated) spindle before starting the metaphase-to-anaphase transition ([@laneKauppi2019], subramanian and Hochwagen , @dumontDesai2012).
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
 
@@
 Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate or cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis ([@laneKauppi2019], Subrahman).
-These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019], @subramanian2014].
+These differences in SAC strength have been connected to conserved features of gametogenesis; presence of a centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
 This model nominates conserved differences in in the context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
@@
 # consequence of greater female variance -- (between cell)
+
 ```
@@
 
+Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013].
+This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
+
 <stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>
@@
 <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
-
-Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013].
-This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
 
@@
 
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.
+As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a conserved trait.
 The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
@@
 Differences in timing of cell cycle between oogenesis and spermatogenesis imposes the the different selective pressures on how sister cohesion affects chromosome structure.
-Faster spermatogenesis may select for synchronization of the separation homologs (cite?).
+Faster spermatogenesis may select for synchronization of the separation homologs (cite).
 While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, espicially for those with dicyate arrest [@lee2019].
@@
 
-The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, however it lacks a mechanism for a evolutionary based predictions.
+The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase [@rubin2020mixing], however it lacks a mechanism for a evolutionary based predictions.
 While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
@@
 We have focused on the model involving direct selection on the recombination pathway, which has an impact on the fitness of an individual, but we acknowledge connections to other models involving indirect selection on recombination rates which involve the fitness outcomes of offspring.
-A model from Goldstein et al, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
+A model from @goldstein1993, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
 The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
@@
 First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
+
 Second certain systems can be leveraged to overcome limitations inherent in some approached.

### 2020-04-16 doc/Discussion.rmd
small changes

--- 2020-04-16 doc/Discussion.rmd
@@
 ---
-title: ""
+title: ''
 date: ''
 output:
+  word_document:
+    toc: yes
+    toc_depth: '3'
+  pdf_document: default
   html_document:

### 2020-04-16 doc/Results.rmd
small changes I think

@@
 
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
+P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "G") %>% filter(sex == "male"),
     aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
@@
   scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+    ggtitle("Nrm IFD G, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.background = element_blank() )

### 2020-04-17 doc/MatMeth.Rmd
starting to change matmethods outline

@@
 ---
-title: "Material Methods"
+title: ''
 date: '`r Sys.Date()`'
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
   html_document:
@@
     toc_float: yes
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
-  word_document:
-    toc: yes
-    toc_depth: '5'
-bibliography: refs_MatMeth.bib
+bibliography: refs_Main_Report.bib
 ---
 
-```{r set.up, echo=FALSE}
+#Material Methods Draft v1
+
+```{r set.up, echo=FALSE, include=FALSE}
+library(knitr)
+library(ggplot2)
+#library(pwr)
+library(plyr)
+library(lattice)
+library(dplyr)
+library(raster)
+library(lme4)
+library(nlme)
+library(RLRsim)
+library(ggpubr)
+library(reshape2)
+#library(patchwork)
+
 #set wd
 #load datafile
+load(file="~./MLH1repo/data/Results_setup_4.16.20.RData")
 
@@
 
-**For SC based metric**
-
-- To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
-This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
-A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-
-- As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
-Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
-
-**SC Length calq**
-
--  **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
-
-- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
-
-- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-**IFD^norm^**
-
-# Mouse Husbandry
+## Mouse Husbandry
 
@@
 
-# Tissue Collection and Immunohistochemistry
+## Tissue Collection and Immunohistochemistry
 
-Spermatocyte spreads were collected and perpared as described in [@peterson_2019] .
+Spermatocyte spreads were collected and perpared as described in [@peterson2019] .
 The majority of mice used were between 5 and 12 weeks, (supplemental table).
@@
 
-# Image Processing
+## Image Processing
 
@@
 
-The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson_2019] respectively.
+The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson2019] respectively.
 
@@
 
-Single bivalent features were measured both manual and with an automated image analysis software [@Peterson_2019]
+Single bivalent features were measured both manual and with an automated image analysis software [@peterson2019]
 
@@
 
-# Statisical Analysis
+## Statisical Analysis
 
@@
 The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).
+
+### Models
 
@@
 
-### Single bivalent charaecterization
+## Single bivalent charaecterization
+
+**For SC based metric**
+
+- To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
+This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
+A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+
+- As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
+Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
 
@@
 
+**SC Length model**
+
+-  **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
+
+- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
+
+- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+**IFD^norm^**
+
 **CO Interferecne**

### 2020-04-17 src/Result_Setup.Rmd
starting to update outline

@@
 
+# Datasubsets
+
+```{r Q1.Short.Long_biv.setup, echo=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#make the Q1 strain list
+Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+```
+
 # MLH1 models
@@
 ```{r Save.File, echo=FALSE}
-
-```
+#save Rdata
+save.image("~./MLH1repo/data/Results_setup_4.16.20.RData")
+#save.image("data/BivData_4_PCA.RData")
+```

### 2020-04-17 doc/MatMeth.Rmd
started outline and more section

@@
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
   html_document:
@@
     toc_float: yes
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
+  word_document:
+    toc: yes
+    toc_depth: '5'
 bibliography: refs_Main_Report.bib
@@
 
-```{r error.repeatbility, echo=FALSE}
+```{r Image.profile, echo=FALSE}
+
+# microscope
+# cell certeria
 
@@
 
-Repeatability was assessed by X (comparing mulitple counted images)
+## Analysis
 
-The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson2019] respectively.
+### DMC1
 
-(the data from these scripts differed slightly in that the total sc script returned the skeletonized SC area for the whole cell. <state the script parameters, (morphogy (3) and other parameter)>
+```{r DMC1.section.outline, echo=FALSE, include=FALSE}
+#cell certeria, number, t.test
+#DMC1 -- leptotene and sygotene cells were identified by SC morphology (by degree of synapsis) (cite another paper?).
+foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
+```
 
-to quantify / charaecterize the SC (length) (synapsed chromosomes, )-- two image analysis algorithms were used both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
+<certeria for DMC1 stained cells (single cells) SC morphology used for staging.>
+<subset of males from strains were used>
 
-The DNA CrossOver script isolated single chromosomes shapes from an image, in addition to returning the SC length -- also measured the location of (green and blue signal along the single SC - reflecting MLH1 foci and cetromere signal respectively).
+A subset of strains, *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were quantified for DMC1.
+A total of `r length( (DMC1.DF %>% filter(stage == "Z"))$Original.Name)` and `r length( (DMC1.DF %>% filter(stage == "L"))$Original.Name)` of zygotene and leptotene cells were quantified respectively.
 
-Hand measures of SC/pactyene / meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
-
-Single bivalent features were measured both manual and with an automated image analysis software [@peterson2019]
-
-DMC1 -- leptotene and sygotene cells were identified by SC morphology (by degree of synapsis) (cite another paper?).
-foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
-
-total number of DMC1 foci per cell were reported.
-
-## Statisical Analysis
+### MLH1
 
@@
 
+```{r MLH1.analysis, echo=FALSE, include=FALSE}
+
+```
+
 The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).
-
-### Models
 
@@
 
-## Single bivalent charaecterization
+### Single bivalent charaecterization
+
+```{r single.biv.chr, echo=FALSE, include=FALSE}
+
+```
+
+The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson2019] respectively.
+
+Hand measures of SC/pactyene / meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
+
+to quantify / charaecterize the SC (length) (synapsed chromosomes, )-- two image analysis algorithms were used both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
+
+The DNA CrossOver script isolated single chromosomes shapes from an image, in addition to returning the SC length -- also measured the location of (green and blue signal along the single SC - reflecting MLH1 foci and cetromere signal respectively).
 

### 2020-04-17 doc/MatMeth.Rmd
starting new outline for material and methods

@@
 
+#MLH1 + Bivalent
+
+#-models (how several were used)
+#the code of models
+#the tests / analysis of models
+
 ```
 
-The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).
-
-**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3. Linear Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ sex * strain + \varepsilon$$
-
-**M4. Linear Model** $$sex\ specific\ mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * strain + \varepsilon$$
-
-We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio]. (lme)
+<The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).>
 
@@
 
+We applied the following models to analze several dependant variables: mouse average mlh1 count per cell, the mouse average of variance in crossover count per cell, mouse mean SC length, mouse mean IFD^norm^ and IFD^raw^, and X.
+
+**M1. Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2. Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
+
+**M3. Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
+
+**M4. Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
+
 This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
@@
 
+We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
+<We used the lme package - for estimating random variabkes>
+<We used x package for testing significance of fixed effects and V package for testing significance of (strain) random variables (strain)>
+
 ### Single bivalent charaecterization
@@
 
+# image analysis programs
+# Richard total sc
+
+# DNA.CrossOver (site, efficiency, curation, chromosome size effects-reduced dataset,
+
+# chromosome proportions
+
+# sc bivalent used in models (aldready above)
+
+# IFD / interference
 ```
 
-The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson2019] respectively.
+The proportions of different chromosome classes (seperated by number of foci) was tested with chi-sqaure tests. <in R using the X package/function>
 
+**CO Interferecne**  -- Our primary mode of examining interference is to compared the interfocal disatnces (IFDs) (raw and normalized by SC length.)  -- Gamma parameter v has historically been used to charaecterize interference and was calculated using <x function>.
+
+<to quantify / charaecterize the SC (length) (synapsed chromosomes, )-- two image analysis algorithms were used both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.>
+
+To quantify the SC area of indiviual cells and single chromosomes we applied the approaches discribed in [@wang_2019_sc] and [@peterson2019] respectively.
+Both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
 Hand measures of SC/pactyene / meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
 
-to quantify / charaecterize the SC (length) (synapsed chromosomes, )-- two image analysis algorithms were used both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
+##### total SC - Wang
 
-The DNA CrossOver script isolated single chromosomes shapes from an image, in addition to returning the SC length -- also measured the location of (green and blue signal along the single SC - reflecting MLH1 foci and cetromere signal respectively).
+As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
+Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
+the mouse average total sc was applied to the model framework.
+<how else was this analyzed?!>
 
-**For SC based metric**
+##### DNACrossover
 
-- To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
+The DNACrossover algorithm is hosted on CyVerse, more detailed description in @peterson2019.
+The DNA CrossOver algorithm isolated single chromosomes shapes from an image, in addition to returning the SC length and also measured the location of (green and blue signal along the single SC - reflecting MLH1 foci and cetromere signal respectively).
+
+**CURATION data ect** (i think these are still in results)
+
+- The performance of the segmentaion algorithm was assessed by counting the number of objects segmented per single cell image (expected 19 or 20 in males and females respectively).
+All automated bivalent data was manually curated before analysis.
+
+**REDUCED DATASET**
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
 This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
@@
 
-- As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
-Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
-
-- The performance of the segmentaion algorithm was assessed by counting the number of objects segmented per single cell image (expected 19 or 20 in males and females respectively).
-All automated bivalent data was manually curated before analysis.
-
-The single bivalent data set was curated of algorithm mistakes manually>.
-For charaecterizing any of the foci positions or interfocal distances any normalized refers to dividing the raw measures by the total SC for that invidiualy bivalent.
-
-The proportions of different chromosome classes (seperated by number of foci) was tested with chi-sqaure tests. <in R using the X package/function>
-
-**chromosome size effects**
+- The long bivalent and short bivalent data sets were compiled by calculating the 1st and 4th quartille of SC lengths for each cell with at least 16 bivalent measures.
+Bivalents from each cell were assigned into the long and short bins based on these SC Length cut offs.
 
@@
 
-- The long bivalent and short bivalent data sets were compiled by calculating the 1st and 4th quartille of SC lengths for each cell with at least 16 bivalent measures.
-Bivalents from each cell were assigned into the long and short bins based on these SC Length cut offs.
-
-< manual measures of whole cell bivalent measures (measures of all bivalents were incorporated into this data set.
-), were compiled by mixed automated bivalent measures with hand measurements.
-(Since the error in automated measures was so low)>
+**chromosome proportion**
 
@@
 
+**IFD, Interference**
+
+In order to examine interference, Interfocal distance was measured from the single bivalent data sets.
 - This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
@@
 
-**CO Interferecne**
-
-Our primary mode of examining interference is to compared the interfocal disatnces (IFDs) (raw and normalized by SC length.)
-Gamma parameter v has historically been used to charaecterize interference and was calculated using <x function>.
-
 # References

### 2020-04-17 doc/MatMeth.Rmd
adding, improving methods section

@@
 ### DMC1
-
 ```{r DMC1.section.outline, echo=FALSE, include=FALSE}
@@
 foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
+
+#<A subset of strains, *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were quantified for DMC1.
+A total of `r length( (DMC1.DF %>% filter(stage == "Z"))$Original.Name)` and `r length( (DMC1.DF %>% filter(stage == "L"))$Original.Name)` of zygotene and leptotene cells were quantified respectively.>
 ```
+<certeria for DMC1 stained cells (single cells) SC morphology used for staging.>
 
-<certeria for DMC1 stained cells (single cells) SC morphology used for staging.>
-<subset of males from strains were used>
-
-A subset of strains, *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were quantified for DMC1.
-A total of `r length( (DMC1.DF %>% filter(stage == "Z"))$Original.Name)` and `r length( (DMC1.DF %>% filter(stage == "L"))$Original.Name)` of zygotene and leptotene cells were quantified respectively.
+The means of DMC1 foci per cell from a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were tested by t-tests.
 
@@
 
-# chromosome proportions
+# chromosome proportions, number of CO per chrm, proportions perpared by chi-sq tests
 
-# sc bivalent used in models (aldready above)
+#chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
+#these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups
 
+# sc bivalent used in models (aldready above), the length of single SC compared by t-tests -- and fit into models
+# Foci Position -- t-test  and put into models
 # IFD / interference
 ```
-
-The proportions of different chromosome classes (seperated by number of foci) was tested with chi-sqaure tests. <in R using the X package/function>
-
-**CO Interferecne**  -- Our primary mode of examining interference is to compared the interfocal disatnces (IFDs) (raw and normalized by SC length.)  -- Gamma parameter v has historically been used to charaecterize interference and was calculated using <x function>.
 
@@
 
-**CURATION data ect** (i think these are still in results)
+This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
+In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
 
-- The performance of the segmentaion algorithm was assessed by counting the number of objects segmented per single cell image (expected 19 or 20 in males and females respectively).
-All automated bivalent data was manually curated before analysis.
+**curation step**  From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
+After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.
+
+To better understand the whole genome rec landscap -- we charecterizaed the number of crossover/MLH1 foci on single chromosomes (from the curated bivalent dataset).
+
+We defined/classified each bivalent into a 'class' based on the number of MLH1 foci (range X to X).
+The **chromosome class proportions** different chromosome classes per strains by sex were tested with chi-sqaure tests with the prop.test function.
+
+#main traits
+
+chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
+these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups
+
+**CO Interferecne**  -- Our primary mode of examining interference is to compared the interfocal disatnces (IFDs) (raw and normalized by SC length.)  -- Gamma parameter v has historically been used to charaecterize interference and was calculated using <x function>.
+
+**SC Length model**
+
+-  **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
+
+- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
+
+**IFD, Interference**
+
+In order to examine interference, Interfocal distance was measured from the single bivalent data sets.
+- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+**IFD^norm^**
 
@@
 
-**chromosome proportion**
-
-**SC Length model**
-
--  **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
-
-- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
-
-**IFD, Interference**
-
-In order to examine interference, Interfocal distance was measured from the single bivalent data sets.
-- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-**IFD^norm^**
-
 # References

### 2020-04-17 src/Result_Setup.Rmd
added in another code chunk

@@
 #the 3 is negative...
-
 #how best to merge rbar and gamma estiamtes (which estimate)
@@
 
+```
+
+```{r algorithm_performance_stats, echo=FALSE}
+
+#sort by category, 
+#compare total boxes and rate passed
+#Curated_BivData  -- pass rate?
+
+#number of mice, number of cells
+#mean number of isolate bivs per cell,
+#use these two metrics
+#cell.level.BivData_table$boxs.IDd 
+#cell.level.BivData_table$nboxes_passed
+
+#calq the mean rate by category
+
+cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
+
+algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
+                        n_obs = length( unique(fileName) ),
+                         mean_pass.rate = mean(pass.rate)
+                         )
+
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
+
+algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
+                        n_obs = length(unique(fileName)),
+                        mean_pass.rate = mean(pass.rate)
+                         )
+
+#showing the table..
+#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
 ```
@@
 #save Rdata
-save.image("~./MLH1repo/data/Results_setup_4.16.20.RData")
+save.image("~./MLH1repo/data/Results_setup_4.17.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-04-17 doc/MatMeth.Rmd
added in another code chunk
more edits

@@
 #load datafile
-load(file="~./MLH1repo/data/Results_setup_4.16.20.RData")
+load(file="~./MLH1repo/data/Results_setup_4.17.20.RData")
 
@@
 
-## Mouse Husbandry
+### Mouse Husbandry
 
@@
  All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
-A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
- Mice were fed on dry standard breeder chow.
+A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (G) is maintained at UW Veterinary school facilities.
+
+<Mice were fed on dry standard breeder chow.
 Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
@@
 Over the course of data collection some breeding colonies  mice were moved from facilities.
-Additionally the GI strain was kept at a separate facility.
-We tested for effects on mean MLH1 counts across difference mouse rooms/facilities and found no effect (Supplemental section).
+Additionally the GI strain was kept at a separate facility.>
 
-## Tissue Collection and Immunohistochemistry
+### Tissue Collection and Immunohistochemistry
 
 Spermatocyte spreads were collected and perpared as described in [@peterson2019] .
-The majority of mice used were between 5 and 12 weeks, (supplemental table).
-for DMC1 spreads we targeted juvinile male mice (X ages).
-
-Because some strains have breeding issues, all mice some strains were kept in breeding pairs (before collecting, and aged out of the desired age range.
-We performed tests for effects on MLH1 counts and found no effects in the older male mice ages.
-(Also we tested other effects, including maternal age...)
-
-The vast majority of oocyte data was collected from neonate mice between 5 to 48 hours old.
-This approach maintain breeding pairing and still result in prophase oocytes <(cite timeline of oocytes in neonates)>.
-.
-Precise staging of embryos via copulary plugs was not feasible in many of these wild derived strains due to their behavior.
-Embryonic samples were collected when pregnancy noted in females, embryos were staged based on (X table markers).
-No effect on mlh1 count due to age effects (pooled as embryo and neonate) was found (Supplemental section).
-
-Meiocyte spreads were made following [@peters_1997].  whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
-
+The majority of mice used were between 5 and 12 weeks for MLH1 and juvinille mice (12 to 15 days) for DMC1 spreads.
+The majority of oocyte tissue was collected from neonate mice between 5 to 48 hours old.
+Meiocyte spreads were made following [@peters_1997] - and was similar for both gonads -- adjusted for volumes.
+Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
 After incubation, gonads were transferred to sucrose solution for mastication.
@@
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The next morning, slides were treated with a photoflow (Kodak, cite) wash.
+The next morning, slides were treated with a photoflow wash (Kodak, cite).
+
+We tested for **effects** on mean MLH1 counts per cell; across mouse rooms, ages of mature males -- and differnces between embryos and neonates.
+No effects were found  (Supplemental section).
+(Supplemental table).
+
+<Because some strains have breeding issues, all mice some strains were kept in breeding pairs (before collecting, and aged out of the desired age range.
+We performed tests for effects on MLH1 counts and found no effects in the older male mice ages.
+(Also we tested other effects, including maternal age...)>
+<This approach maintain breeding pairing and still result in prophase oocytes <(cite timeline of oocytes in neonates)>.
 
@@
 
-SYCP3 was used to mark  visualize the Axial element of the synaptonemal complex.
-Sicne all mouse chromosomes are telocentric/acrocentric, CREST was used to indicate the centromere <insert the concentrations>
+Meiocyte spreads were incubated with antibody protocol.
 
-Precursurs to COs, DSBs were quantified with the DMC1 staning antibody.
+To visualize chromosomes/chromatin in the meiocytes -- we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
+Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
 
-## Image Processing
+The master mix was made --
+
+### Image Processing
 
@@
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
+Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop (cite).
 
-For MLH1 pachytene quantification, cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci.
-
-## Analysis
+### Analysis
 
@@
 
-The means of DMC1 foci per cell from a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were tested by t-tests.
+SC morphology was used to stage spermatocytes (as zygotene, leptotene, or pachtyene) (following X source).
+The means of **DMC1** foci per cell from a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were tested by t-tests.
 
@@
 
-The distributions of CO counts per cell was assessed for normality (supplemental figure).
-Mouse mean MLH1 count was used to get around the bad discreet nature of count data.
-(normality was confirmed with X distribution, supplemental figure).
+For **MLH1** pachytene quantification, cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
+(Quality score 1 to 5 )  based on foci brightness /staining and spreading of bivalents.
+The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
 
 ```{r MLH1.analysis, echo=FALSE, include=FALSE}
-
 #MLH1 + Bivalent
@@
 #the tests / analysis of models
-
 ```

### 2020-04-17 doc/MatMeth.Rmd
added text from antibody mix

@@
 The master mix was made --
+
+ Immunohistochemistry was performed following Anderson et al.
+(1999) and Koehler et al.
+(2002).
+Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
+Each slide was blocked for 30 minutes in ADB before sixty ul of a primary antibody mix containing anti rabbit anti-MLH1 polyclonal antibody to MLH1 (Calbiochem; diluted 1:50), anti goat polyclonal antibody to SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) was in ADB and incubated for 48 hours at 37 degrees.
+Slides were washed twice in 50ml ADB between primary and secondary antibody incubations.
+Slides were incubated overnight at 37 degrees in Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200).
+Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
+Slides were washed in 1x PBS, dried, and fixed with Prolong Gold Antifade (Invitrogen) for at least 24 hours.
+Mice with at least 10 cells with good staining were included in our analysis.
 

### 2020-04-17 doc/MatMeth.Rmd
hopefully small changes

@@
     toc_float: yes
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
-  word_document:
-    toc: yes
-    toc_depth: '5'
 bibliography: refs_Main_Report.bib

### 2020-04-17 doc/MatMeth.Rmd
more updates and changes

--- 2020-04-17 doc/MatMeth.Rmd
@@
 
-Meiocyte spreads were incubated with antibody protocol.
+Meiocyte spreads were incubated with antibody protocol /  Immunohistochemistry was based on that in Anderson et al. (1999) and Koehler et al. (2002).
 
@@
 
-The master mix was made --
+Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
+
+Each slide was blocked for 30 minutes in ADB incubating with sixty ul of primary antibody master mix (60ul) 
+was in ADB and incubated for 48 hours at 37 degrees.
+The master mix was made (with a primary with polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) -- (**mix of DMC1**), anti goat polyclonal antibody to SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200)
+
+Slides were washed twice in 50ml ADB between primary and secondary antibody incubations.
+
+secondary antibodies master mix was made -- Slides were incubated overnight at 37 degrees in Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200).
+
+Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
+Slides were washed in 1x PBS, dried, and fixed with Prolong Gold Antifade (Invitrogen) for at least 24 hours.
+Mice with at least 10 cells with good staining were included in our analysis.
 

### 2020-04-17 doc/MatMeth.Rmd
small changes

@@
 _musculus^WSB^_).
-
 All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
@@
 
-Spermatocyte spreads were collected and perpared as described in [@peterson2019] .
-The majority of mice used were between 5 and 12 weeks for MLH1 and juvinille mice (12 to 15 days) for DMC1 spreads.
-The majority of oocyte tissue was collected from neonate mice between 5 to 48 hours old.
-Meiocyte spreads were made following [@peters_1997] - and was similar for both gonads -- adjusted for volumes.
+Spermatocyte spreads were collected and prepared as described in [@peterson2019] .
+The majority of mice used for testes tissues were between 5 and 12 weeks for MLH1 and juvinile mice, 12 to 15 days, for DMC1 spreads.
+The majority of oocyte tissue was collected from neonate mice between 0 to 48 hours old.
+
+Meiocyte spreads for both sexes were made following [@peters_1997] with ajustments for volumes.
 Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
-After incubation, gonads were transferred to sucrose solution for mastication.
-15ul of cell slurry was transferred to 80ul of 2% PFA solution.
+Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The next morning, slides were treated with a photoflow wash (Kodak, cite).
+The following morning, slides were treated with a photoflow wash (Kodak, cite).
 
-We tested for **effects** on mean MLH1 counts per cell; across mouse rooms, ages of mature males -- and differnces between embryos and neonates.
-No effects were found  (Supplemental section).
-(Supplemental table).
+We tested for **effects** on mean MLH1 counts per cell; across mouse rooms, ages of mature males -- and differnces between embryos and neonates. No effects were found (Supplemental section).
 
@@
 (Also we tested other effects, including maternal age...)>
-<This approach maintain breeding pairing and still result in prophase oocytes <(cite timeline of oocytes in neonates)>.
+
+<This approach maintain breeding pairing and still result in prophase oocytes <(cite timeline of oocytes in neonates)>
 

### 2020-04-17 doc/MatMeth.Rmd
adding in more citations

@@
 
-Meiocyte spreads were incubated with antibody protocol /  Immunohistochemistry was based on that in Anderson et al. (1999) and Koehler et al. (2002).
+Meiocyte spreads were incubated with antibody protocol /  Immunohistochemistry was based on that in [@anderson1999] and [@koehler2002].
 
@@
 
-Each slide was blocked for 30 minutes in ADB incubating with sixty ul of primary antibody master mix (60ul) 
-was in ADB and incubated for 48 hours at 37 degrees.
-The master mix was made (with a primary with polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) -- (**mix of DMC1**), anti goat polyclonal antibody to SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200)
-
-Slides were washed twice in 50ml ADB between primary and secondary antibody incubations.
-
-secondary antibodies master mix was made -- Slides were incubated overnight at 37 degrees in Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200).
-
-Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
-Slides were washed in 1x PBS, dried, and fixed with Prolong Gold Antifade (Invitrogen) for at least 24 hours.
-Mice with at least 10 cells with good staining were included in our analysis.
+Each slide was blocked for 30 minutes in ADB then incubated with  60ul of primary antibody master mix for 48 hours at 37 degrees.
+For each master mix polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) were suspended in ADB.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
+Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200)  were suspended in ADB.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
+Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
 
@@
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop (cite).
+Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.
 
@@
 ```
+
 <certeria for DMC1 stained cells (single cells) SC morphology used for staging.>

### 2020-04-18 doc/MatMeth.Rmd
more edits

@@
 
-<This approach maintain breeding pairing and still result in prophase oocytes <(cite timeline of oocytes in neonates)>
-
 **Staining / Immunohistochemistry**
 
-Meiocyte spreads were incubated with antibody protocol /  Immunohistochemistry was based on that in [@anderson1999] and [@koehler2002].
-
-To visualize chromosomes/chromatin in the meiocytes -- we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
+The immunohistochemistry was based on that in [@anderson1999] and [@koehler2002].
+To visualize chromosomes in the meiocytes we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
 Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
-
 Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
@@
 
-Images of cells were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
+Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
@@
 ### DMC1
+
 ```{r DMC1.section.outline, echo=FALSE, include=FALSE}
@@
 foci counted manually, error ccalulated from a subset of cells counted twice (supplement)
-
-#<A subset of strains, *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were quantified for DMC1.
-A total of `r length( (DMC1.DF %>% filter(stage == "Z"))$Original.Name)` and `r length( (DMC1.DF %>% filter(stage == "L"))$Original.Name)` of zygotene and leptotene cells were quantified respectively.>
 ```
 
-<certeria for DMC1 stained cells (single cells) SC morphology used for staging.>
-
-SC morphology was used to stage spermatocytes (as zygotene, leptotene, or pachtyene) (following X source).
-The means of **DMC1** foci per cell from a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were tested by t-tests.
+SC morphology was used to stage spermatocytes in the zygotene, leptotene, or pachtyene cell stage.
+Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were compared by t-tests.
 
@@
 
-For **MLH1** pachytene quantification, cells with a full karyotype (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes), distinct foci, and intact bivalents were included for quantification.
-(Quality score 1 to 5 )  based on foci brightness /staining and spreading of bivalents.
-The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
+MLH1 foci were quantified from cells with intact and compelete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score, between 1 to 5, was asigned to each cell based on staining quality and general spread of bivalents.
+<cell with small defects were included -- and coded as a score of 5 - but these cell weren't included in the final analysis.>  The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
 
@@
 
-The use of wild derived inbred strains enabled us to design an utilize a mixed model for our analytical framework.
-We built our model to predict the strain average MLH1 count per cell.
+The use of wild derived inbred strains enabled us to design an utilize a mixed model for our analytical framework. We built our model to predict the strain average MLH1 count per cell.
+
 The main effects and their interaction of subspecies and sex were coded as fixed.
@@
 
-We applied the following models to analze several dependant variables: mouse average mlh1 count per cell, the mouse average of variance in crossover count per cell, mouse mean SC length, mouse mean IFD^norm^ and IFD^raw^, and X.
+We applied the following models to analyze several dependent variables: mouse average mlh1 count per cell, the mouse average of variance in crossover count per cell, mouse mean total SC per cell, mouse mean single SC length, mouse mean IFD^norm^ and IFD^raw^, and position of crossover for single crossover bivalents.
+
+X package used to test the fixed effects, Z package used to test the random effects.
 
@@
 We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
-<We used the lme package - for estimating random variabkes>
-<We used x package for testing significance of fixed effects and V package for testing significance of (strain) random variables (strain)>
+
+We used x package for testing significance of fixed effects and V package for testing significance of (strain) random variables (strain).
 
@@
 
-<to quantify / charaecterize the SC (length) (synapsed chromosomes, )-- two image analysis algorithms were used both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.>
+To charaecterize the SC length of individual chromosomes which is a marker of chromatin compaction <from MLH1 stained cells>, two image analysis algorithms were used to specifically (measure indiviual cells) [@wang_2019_sc] and indiviaul chromosomes [@peterson2019].
+Both apply a 'skeletonizing' transformation to the synapsed chromosomes which transforms the chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
+Hand measures of SC/pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
 
-To quantify the SC area of indiviual cells and single chromosomes we applied the approaches discribed in [@wang_2019_sc] and [@peterson2019] respectively.
-Both which apply a 'skeletonizing' transformation to the synapsed chromosomes,  which transforms the elongatde chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
-Hand measures of SC/pactyene / meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].
-
-##### total SC - Wang
-
-As an additional metric of chromatin compaction, we computed the total (summed) SC length of all bivalents for single cells, using the image analysis algorithm of [@wang_2019_sc].
+The **total SC** per cell (all bivalents) was quantified across single cell images.
 Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
@@
 
-The DNACrossover algorithm is hosted on CyVerse, more detailed description in @peterson2019.
-The DNA CrossOver algorithm isolated single chromosomes shapes from an image, in addition to returning the SC length and also measured the location of (green and blue signal along the single SC - reflecting MLH1 foci and cetromere signal respectively).
-
+The **DNA CrossOver** algorithm isolates single straightened bivalents/chromosomes shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
@@
 
-**curation step**  From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
+A curation step is applied to the total single bivalent dataset to remove poor isolation/segmentation instances.
+From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
 After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.
 
-To better understand the whole genome rec landscap -- we charecterizaed the number of crossover/MLH1 foci on single chromosomes (from the curated bivalent dataset).
+**This dataset is used to charaecterize single bivalent traits**
 
-We defined/classified each bivalent into a 'class' based on the number of MLH1 foci (range X to X).
-The **chromosome class proportions** different chromosome classes per strains by sex were tested with chi-sqaure tests with the prop.test function.
+The number of MLH1 foci on single chromosomes was quantified from the curated bivalent dataset to better understand the total recombination landscape.
+The **proportions** of single and double crossover bivalents were tested across mouse categories by chi-sqaure tests with the prop.test function.
 

### 2020-04-18 doc/MatMeth.Rmd
more edits to methods

@@
 
+#main traits
+
 **This dataset is used to charaecterize single bivalent traits**
@@
 
-#main traits
+**other traits** were also charaecterized -- SC length (a marker for chromatin compaction), foci position (raw and normalized), and interfocal distance a marker for interference, these traits were fit into the main models (as mentioned above).
+In addition to fiting the mouse means into the models discribed above, t.tests were used to test the means between sexes and groups.
 
-chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
-these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups
-
-**CO Interferecne**  -- Our primary mode of examining interference is to compared the interfocal disatnces (IFDs) (raw and normalized by SC length.)  -- Gamma parameter v has historically been used to charaecterize interference and was calculated using <x function>.
+**IFD, Interference** In order to examine interference, Interfocal distance was measured from the single bivalent data sets.
+- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
+Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
+**IFD^norm^**
 
@@
 - **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
-
-**IFD, Interference**
-
-In order to examine interference, Interfocal distance was measured from the single bivalent data sets.
-- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-**IFD^norm^**
 
@@
 Using the single bivalent dataset, the within cell distribution of bivalent SC lengths from cells with at least 17 isolated chromosome measures.
-The longest and shortest 4 to 5 chromosomes within each cell were identified by qualqulating the fourth and second quartille.
-(these comprised the short and long bivalent datasets).
+The longest and shortest 4 to 5 chromosomes within each cell were identified by calculating the fourth and second quartille.
 

### 2020-04-20 doc/MatMeth.Rmd
adding in some calqs to results

--- 2020-04-18 doc/MatMeth.Rmd
@@
 
-**Staining / Immunohistochemistry**
-
-The immunohistochemistry was based on that in [@anderson1999] and [@koehler2002].
+The **immunohistochemistry** was based on that in [@anderson1999] and [@koehler2002].
 To visualize chromosomes in the meiocytes we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
@@
 
+Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were compared by t-tests.
 SC morphology was used to stage spermatocytes in the zygotene, leptotene, or pachtyene cell stage.
-Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were compared by t-tests.
 
-### MLH1
+### MLH1 (Analytical Framework)
+
+-Background for setting up/organizing models -- (heiarchal nature of multiple strains within subspecies -- ) - hierarchical nature of categories for inviual mouse averages ()
+
+-(primary focus is mean MLH1 count per cell - at the mouse observation level - apply to more meiocyte traits)
+
+-Our goal -- design an utilize a mixed model for our analytical framework - which leverages / accounts for the hierarchical/nested nature of categories across all the individual mice (e.g.
+sex, strain subspecies).
 
@@
 
+**X package used to test the fixed effects, Z package used to test the random effects.**
+
 The main effects and their interaction of subspecies and sex were coded as fixed.
@@
 We applied the following models to analyze several dependent variables: mouse average mlh1 count per cell, the mouse average of variance in crossover count per cell, mouse mean total SC per cell, mouse mean single SC length, mouse mean IFD^norm^ and IFD^raw^, and position of crossover for single crossover bivalents.
-
-X package used to test the fixed effects, Z package used to test the random effects.
 
@@
 ### Single bivalent charaecterization
+
+**combine any X and XY issue with reduced dataset** An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
 

### 2020-04-20 doc/Results.rmd
adding in some calqs to results

@@
 
-<!--
-From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.
-
-the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1
-
-INSERT TABLE FOR THE TOTAL SC means ect
-
-The difference between
-
-bin is short for binary, it reports the binarized area of the SC
-
-skel -- takes the skeletonized midline of the bin objects
--->
 # Main Figure
@@
 
+**re-read Lenzi et al -- heterogeneity in oogenesis**
+
 We examine the within animal variance in MLH1 foci count per cell using the same models applied to mean MLH1 foci count (replacing mean MLH1 foci count with variance of MLH1 foci across cells as the dependent variable).
@@
 The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t.test; p = `r DMC1.Lratio_t.test$p.value` and p=`r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios respectively).
+
+**we should note that these interpertations are dependant on which count of DSBs is most reflective for the final number of matur crossovers.** A number of factors make this determintion less clear - including the potential for mulitple rounds of DSB formation (CO homeostasis) (cite) and difference in rates for repair (cite)
+
+(also the data is very good)
 

### 2020-04-20 doc/Results_v2.rmd
adding in some calqs to results

@@
 #load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
-#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
-#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
-#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
-#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
-#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
 
@@
 
+#adjust 
+MLH1.merge.Skel_2$mouse <- as.factor(MLH1.merge.Skel_2$mouse)
+MLH1.merge.Skel_2$nMLH1.foci <- as.numeric(MLH1.merge.Skel_2$nMLH1.foci)
+MLH1.merge.Skel_2$quality <- as.numeric(MLH1.merge.Skel_2$quality)
+
 ```
@@
 ```{r tot.skel.table, echo=FALSE}
-totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
+#calq the correlation for tota.SC and mlh1 - after mouse averages
+#use MLH1.merge.Skel_2 -- quality 1-4
+
+totSC.mouse.av.table <- ddply(.data= ( MLH1.merge.Skel_2 %>% filter(quality < 5) ), 
                  .(mouse),
@@
                  nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                 # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
@@
 
+#ADD CORRELATION
+for(g in length(totSC.mouse.av.table$mouse) ){
+totSC.mouse.av.table$cor.SCskel.MLH1[g] <- cor(totSC.mouse.av.table$mean.MLH1[g], totSC.mouse.av.table$mean.skel[g])
+}
+
 mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
@@
 ```{r tot.SC.setup, echo = FALSE}
+#calc=q correlation, total.sc and mean MLH1
 
@@
 # Within mouse variance in CO count per cell
+
+**re-read Lenzi et al -- heterogeneity in oogenesis**
 
@@
 
-```{r Q2.SC.length.bret.comments, echo=FALSE}
-#correlation
-
-```
-
 ```{r Q2.SC.short.long.setup, echo=FALSE}
@@
 
+```{r Q2.SC.length.bret.comments, echo=FALSE}
+#correlation, total sc and Mean MLH1
+
+```
+
 ```{r Q2.short.long.ttest, echo=FALSE}
@@
 
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC **([report spearman's rho and p-value])**.
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC  `r cor(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1)`.
 Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).

### 2020-04-20 doc/Results_v2.rmd
updating values

@@
 We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results` -- 3.9710^{-5}), subspecies (p=`r MLH1.MMlmer_results$sub.sp`--  1.7210^{-4}), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval` -- 3.110^{-5}) each significantly affect recombination rate.
+(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
 
 After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p= 3.9910^{-6} p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and domesticus^G^ (p= 1.0410^{-6}  -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
-
-We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4  **male specific, M2?
+A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
+and domesticus^G^ (`r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
+In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (`r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (`r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
+
+We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4  **male specific, M2?
 rename the code for these models**).
-In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect =
+In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect =
 `r MLH1.male.specific.glm.M2$coefficients[5]` -- ~ 5 foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = `r MLH1.male.specific.glm.M2$coefficients[8]` -- ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect  `r MLH1.male.specific.glm.M2$coefficients[6]` ~ 2 foci).
 These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p=0.02  -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
+Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p= `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
@@
 To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` (old) `r round(MLH1var_M2_sum$coefficients[,4][4], 5)`  correct).
+Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < **smallNumber** -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` (old) `r round(MLH1var_M2_sum$coefficients[,4][4], 5)`  correct).
 **the pland full M2 model is not significant**.
@@
 
-The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2 -- also M3).
+The results are similar: sex is the strongest effect (p < **smallNumber** `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2 -- also M3).
 When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
@@
 In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` -- 76 leptotene-stage and 75 --`r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene-stage and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
 
 The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`).
+In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 `r ttest.HighLow.Z.pval`).
 **note that this interpertation depends on which stage is most relevant to CO number**
@@
 Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = 0.94 -- `r DMC1.Lratio_t.test$p.value` and p = 0.11 -- `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios, respectively).
+The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = 0.11 `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios, respectively).
 This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
@@
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
-
-From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
-After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.
+In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)` (musculus^KAZ^ female).
+
+From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` bivalent objects were isolated by the image analysis software.
+After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )`  single-bivalent observations remained.
 We assume that the isolation of bivalents within cells is unbiased.
@@
 
-Ninety-six percent of single bivalents in our pooled dataset (n = 34,982 -- `r total.biv.all`) have either one or two crossovers (Figure X).
+Ninety-six percent of single bivalents in our pooled dataset (n = `r total.biv.all`) have either one or two crossovers (Figure X).
 The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
 low recombining strains (Figure X).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in molossinus^MSM^.
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in molossinus^MSM^.
 
 Following patterns in the genome-wide recombination rate, male musculus^PWD^ and male molossinus^MSM^ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; musculus^PWD^ vs.
-musculus^KAZ^ p = 3.1510^{-33} -- `r PWD.KAZ.2CO.prop.test$p.value`; molossinus^MSM^ vs.
-molossinus^MOLF^ p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
+musculus^KAZ^ p = `r PWD.KAZ.2CO.prop.test$p.value`; molossinus^MSM^ vs.
+molossinus^MOLF^ p = `r MSM.MOLF.2CO.prop.test$p.value`).
 
@@
 Short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculus^SKIVE^.
-The female:male ratio ranges across strains from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
+The female:male ratio ranges across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
 <moved section on total SC to methods>
@@
 ```{r Q1.F1pos_long.short_setup, echo=FALSE}
-
-#REAL.short_mouse_table
-#REAL.long_mouse_table
 
@@
 In all strains, the landscape across 1CO bivalents is significantly different in females and males.
-Normalized foci positions tend to be more central in females **(mean focus position = X)** and closer to the telomere in males **(mean focus position = Y)** (t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value`).
-Sex is also the most significant effect on focus position in a mixed model (M1: p =1.2610^{-25} -- `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p =1.3310^{-7} -- `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 1.3310^{-7} -- `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+Normalized foci positions tend to be more central in females `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` and closer to the telomere in males `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )` (t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
+Sex is also the most significant effect on focus position in a mixed model (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species [@sardell_sex_2020].
@@
 
-```{r Q1.IFD_Ttests, echo=FALSE}
+```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE}
 #below are calq just from 2COs, have no fear
@@
 #p-value = 0.3
-
-```
-
-There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = 0.07) is driven by one strain (t-test without domesticus^G^; p = 0.27).
-This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculus^KAZ^ showing no difference (t-test; p = 0.33).
-Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (LRT: p = 6.7410^{-14}).
-When interference is measured in physical SC units (IFDraw), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-
-While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-<moved a section on relationship of linkage maps to Discussion (or Methods)>
-
-This result indicates there is no general pattern of sex differences in crossover interferencethat females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFDnorm) than females in seven out of eight strains (t-tests; p < 0.02), with only musculus^KAZ^ showing no difference (t-test; p = 0.33).
-While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
-The difference in mean IFD^norm^ are highly significant between sexes.
-The mean IFDnorm are significantly long in six of seven strains (MSM t.test; p-value = 0.01), musculus^SKIVE^ p-value = 4.310^{-5}), musculus^PWD^ p-value = 2.2810^{-4}), molossinus^MSM^ p-value =0.01), domesticus^LEW^ p-value = 3.5710^{-5}), domesticus^G^ p-value = 0.05), and domesticus^WSB^ (t.test; p-value =0.03) ).
-Only musculus^KAZ^ did not have a difference in mean IFD^norm^ (t.test; p-value = 0.33).
-In examining theExamination of IFDnorm distributions, we note indicates that  the female IFD^norm^ observations values are centered at approximately 50% and have show a slight an enrichment of short low (<25%) IFD^norm^ observationsvalues, whereas males are enriched for higher values.
-Comparitively, the IDF^norm^ distributions in males are enriched for longer IFD^norm^.
-The three sets of models Mixed models and general linear models of IFD^norm^ support the result inference of stronger interference in males:, sex was theis the most significant significant effectvariable (LRT: p = 6.7410^{-14}).
-When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFDraw shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.While the comparisons of the IFDnorm metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-
-<!--  finer distribution patterns <Full dataset (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`).> -->
+```
 
@@
 
-```{r Q1.IFD_gamma, echo=FALSE}
+```{r Q1.IFD_gamma, echo=FALSE, message=FALSE}
 library(EnvStats)
@@
 #shape  5.4
-
 #male, stronger interference
@@
 
-```{r gamma.IFD, echo=FALSE}
-
+```{r Q1.gamma.IFD, echo=FALSE}
 female.gamma <- egamma( (all.female.IFDs )$IFDval, method = "mle", ci = TRUE, 
@@
 ###
-
 PWD.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
@@
                             normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-```
-
-```{r Q1.male.IFD, echo=FALSE}
-
+```
+
+```{r Q1.male.IFD_gamma, echo=FALSE}
 #gamma fits
@@
                             normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
 ```
@@
 ```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-#get a list of the old mice to double check
-#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#MIXED MODEL
-#https://rdrr.io/cran/lme4/man/pvalues.html notes of fixed effect p values
-
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
 
@@
 ```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE}
-
 #mouse.avs_4MM -- is also the bivbased mouse average table 
@@
 
-The three sets of models support the result of stronger interference in males, sex was the most significant significant effect (LRT: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`).
+There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
+A marginally significant difference between the sexes (t-test; p = 0.07 -- `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain (t-test without domesticus^G^; p = 0.27 -- `r Q1.IFD.raw_t.test_G$p.value`).
+This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
+In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < 0.02 -- `r Q1.IFD.PER_FULL_t.test$p.value`), with only musculus^KAZ^ showing no difference (t-test; p = 0.33 --  `r Q1.IFD.PER_t.test_KAZ$p.value`).
+Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1 - LRT **random**: p = 6.7410^{-14} -- `r Q1.IFD.PER_M1_lmer.results$sex.results`  glm *M2* `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`  *M3* `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
 When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-These results support the model of the physical measures of interference (in SC units) is conserved between sexes inwithin the same species.
+In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+While the comparisons of the IFD^norm^ metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
+
+<moved a section on relationship of linkage maps to Discussion (or Methods)>
 
@@
 Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-**SC length is expected to be positively associated with genome-wide recombination rate because X**.
-Crossover interference is expected to be negatively **associated with genome-wide recombination rate because X**.
+**SC length is expected to be positively associated with genome-wide recombination rate because of the loop-axis strucutre** (zickler Kleckner 1999, Merier 2015?).
+Crossover interference strength is expected to be negatively **associated with genome-wide recombination rate since interference is defined as the non-random spacing between crossovers on the same chromosome**.
+
 Following this logic we predict (1) musculus^PWD^ will have greater SC length and weaker interference than musculus^SKIVE^, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) molossinus^MSM^ will have longer SC and weaker interference compared to molossinus^MOLF^, and 3) _domesticus_ strains will have similar SC length and crossover interference.
@@
 ```{r Q2.IFD.long.biv_t.test, echo=FALSE}
-
 #short removed PER
@@
 
-Short.removed_BivData.LongBiv_IFD.PER_Ttest$p.value
+#Short.removed_BivData.LongBiv_IFD.PER_Ttest$p.value
 #0.0083
@@
 
-Short.removed_BivData.LongBiv_IFD.ABS_Ttest$p.value
+#Short.removed_BivData.LongBiv_IFD.ABS_Ttest$p.value
 #0.047
-
 ```
@@
 ```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
-
 #subset
@@
 ```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
-
 #t.test for IFD
@@
 
-#different test must be sed for the Dom
-
 #t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == #"male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
@@
    (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
-
 ```
@@
 ```{r Q2.IFD.glm.test, echo=FALSE}
-#REPLACE Male.poly with ..
-#(mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1")
 
@@
 
-#repeating chrm models with chromsome length
-#FOR BIVALENT LEVEL MODELS
-
 Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
@@
 ```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
-
-#calculate the rates of 'close foci' (<30%)
-#examine the changes in the distribution closer.
 
@@
 #p-value = 0.03, slight dif tween 3 high rec strains
-
-#don't need to run chi square for this pattern, since it's enough to 
 ```
@@
 #Log regression all show significant prediction
-
 #IFD.PER
@@
 #SINGLE BIVALENT LEVEL
-
 Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
@@
               data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
 #summary(Log.Q2.IFD.per_bivdata.Musc)
@@
               data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(Log.Q2.IFD.per_bivdata.Mol)
 
@@
               data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-
 #summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
@@
               data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
 #summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
@@
 ```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-
 #Full dataset
-
 male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
@@
 #The PER.IFDs don't have significant strain effects
-```
-
-```{r Q2.IFD_remove.short, echo=FALSE, include=FALSE}
-#mus
-#PER 
-#biv level
-
-#mus
-
 ```
@@
 
-Rec.group.BivData_IFD.PER_Ttest$p.value
+#Rec.group.BivData_IFD.PER_Ttest$p.value
 #p-value < 2.2e-16
@@
 
-Rec.group_IFD.PER_Ttest$p.value 
+#Rec.group_IFD.PER_Ttest$p.value 
 #p-value = 6.872e-07
-
 ```
@@
 ```{r Q2.IFD_short.removed_t.test, echo=FALSE, include=FALSE, eval=FALSE}
-
-#Curated_BivData_short.IFD.removed
-#still need to calq the mouse averages
-
 Short.removed_BivData_IFD.PER_Ttest <- t.test(Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 1], 
@@
                        close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
-
 Short.removed_BivData_IFD.PER_Ttest_mus.av$p.value
@@
                                  close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
-
-#p.value
-
 ```
@@
 That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
+We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% --
+
+in low-recombination strains.
 The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (domesticus^G^) to 16% (musculus^KAZ^) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculus^SKIVE^, molossinus^MSM^, and musculus^PWD^, respectively).
@@
 # Q2 Summary
+
+**move main themes to conclusion**
 

### 2020-04-20 doc/MatMeth.Rmd
updating values

--- 2020-04-18 doc/MatMeth.Rmd
@@
 <how else was this analyzed?!>
+
+**walk through how outliers are removed**
 

### 2020-04-20 doc/Results_v2.rmd
working on the DMC1 part

@@
 
-kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+#MAIN TABLE, headers need to be fixed, remove var.MLH1
+#try adding double decker colume names / lines for early Z and late Z
+#kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
 
@@
 In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` leptotene-stage and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
-
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 `r ttest.HighLow.Z.pval`).
-**note that this interpertation depends on which stage is most relevant to CO number**
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` early zygotene-stage and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` late zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
+
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`).
+In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = 0.66 --- `r ttest.HighLow.Z.pval`).
+
+[@cole2012] suggest that early Z counts for DMC1 are most relevant for predicting CO number.
+(CO homeostasis -- robustness )
 
@@
 Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = 0.11 `r DMC1.Zratio_t.test$p.value` for leptotene and zygotene ratios, respectively).
+**what does the ratio tell us? which stage of DMC1 counts is most relevant for CO count**
+
+The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = 0.11 -- `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
 This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
+
+**note that this interpertation depends on which stage is most relevant to CO number**
+
+the difference in DMC1 counts across staes of zygotene may be related to the findings related to (crossover homeostasis).
+
+The number of DSBs increase in zygotene, then deminish as the effects of crossover homeostasis are implemented (Cole). (co homeostasis also indecates dsb )
+
+'Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links/ connects the CO number variation to 'pre' DSB formation -- fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
+(this study only examined one stage of early prophase for DMC1 counts)
+
+'does the number of DSBs per se regulate CO number?'
+
+bairer-- DSB number is not a primary driver of variation in CO number -- these authors suggest that chromatin organizaiton (loop-axis) structure is the primary determinant of CO variation.
+
+the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 c3h, 7.3 B6 )
 
@@
 
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
+```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE, warning=FALSE, message=FALSE}
 #long SC Long_biv_mouse.avs_4MM
@@
 
-<!--t.tests -->
-
-```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE}
+```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
 #below are calq just from 2COs, have no fear
@@
 
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE}
+```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
 nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
@@
 
-```{r Q1.gamma.IFD, echo=FALSE}
+```{r Q1.gamma.IFD, echo=FALSE, message=FALSE}
 female.gamma <- egamma( (all.female.IFDs )$IFDval, method = "mle", ci = TRUE, 
@@
 
-```{r Q1.male.IFD_gamma, echo=FALSE}
+```{r Q1.male.IFD_gamma, echo=FALSE, message=FALSE}
 #gamma fits
@@
 
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE}
+```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
 #mouse.avs_4MM -- is also the bivbased mouse average table 
@@
 
-We compared mouse means of three SC-based metrics: total SC length, mean short bivalent length, and mean long bivalent length.
-The long bivalent data set was isolated by choosing bivalents in the X quartile of the SC length distribution within cells.
-The short bivalent dataset was isolated as described above.
-
-figure
-
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC  `r cor(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1)`.
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
 Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
-<Bret's note on positive correlation, 'A correlation across mice, across strains, or both?'>
-
-While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+**Bret's note on positive correlation, 'A correlation across mice, across strains, or both?** While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+**the values above are diferent between editions**
 
@@
 
+**the values between versions are slightly different**
 In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
 In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-<Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?>
+**Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?**
 
@@
                                            ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_by_group <- ggplot(male.biv.2CO, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS \nSeperated by Rec Groups")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+ylim(c(0,110))
 

### 2020-04-21 src/SetupData.Rmd
switched adj_nMLH1 to MLH1

@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
                         cV = cv(adj_nMLH1.foci, na.rm=TRUE),
-                        var = as.numeric(format(round(   var(adj_nMLH1.foci, na.rm=TRUE),3), nsmall=3)),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
+
+                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm=TRUE),3), nsmall=3)),
+                        sd   = round(sd(nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
 

### 2020-04-21 doc/Results_v2.rmd
better job at removing adj_nMLH1 variables

@@
 
-#MLH1_data_setup_2.24.20.RData
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
-
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
 #load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
-
 #2.2.20
@@
 
-#What are the main DFs for this code chunk
-#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
-#DF.HetC.MixedModel   <- only Mmusculus
-#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
-
-#variance is still chr
-#AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
-
-#use the merged DF to filter by age 
-#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-#AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
 DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
@@
 
-#REMOVE EXTRA STRAINS
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-#DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
-
 #AP_mouse_table excluide Qual5 cells
 DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
-
 DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)

### 2020-04-21 src/Result_Setup.Rmd
better job at removing adj_nMLH1 variables

@@
 
-#MLH1_data_setup_2.24.20.RData
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
-
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
+#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
 #load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")

### 2020-04-21 src/SetupData.Rmd
better job at removing adj_nMLH1 variables

@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci),
+                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
@@
                         mean_co = as.numeric(format(round(  mean(nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci, na.rm=TRUE),
+                        cV = cv(nMLH1.foci, na.rm=TRUE),
 

### 2020-04-21 doc/MatMeth.Rmd
working on the DMC1 part
changing the Rdata files across scripts

@@
 #load datafile
-load(file="~./MLH1repo/data/Results_setup_4.17.20.RData")
+load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
 
@@
 ```
-
-ToDo; Bret's suggestion for moving thing to Mat and Methods
 
@@
 
-Spermatocyte spreads were collected and prepared as described in [@peterson2019] .
-The majority of mice used for testes tissues were between 5 and 12 weeks for MLH1 and juvinile mice, 12 to 15 days, for DMC1 spreads.
-The majority of oocyte tissue was collected from neonate mice between 0 to 48 hours old.
+The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
+Spermatocyte spreads were collected and prepared as described in [@peterson2019].
+The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
+Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
 
-Meiocyte spreads for both sexes were made following [@peters_1997] with ajustments for volumes.
 Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
@@
 The following morning, slides were treated with a photoflow wash (Kodak, cite).
-
-We tested for **effects** on mean MLH1 counts per cell; across mouse rooms, ages of mature males -- and differnces between embryos and neonates. No effects were found (Supplemental section).
+**slides were stored at -20*C if not stained imeediantly**
 
@@
 
-The **immunohistochemistry** was based on that in [@anderson1999] and [@koehler2002].
-To visualize chromosomes in the meiocytes we used markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
+The **immunohistochemistry**
+
+ To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
 Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
+The staining protocol was based on that in [@anderson1999] and [@koehler2002].
 Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
 
-Each slide was blocked for 30 minutes in ADB then incubated with 60ul of primary antibody master mix for 48 hours at 37 degrees.
-For each master mix polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) were suspended in ADB.
+Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
+The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
 Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
@@
 
+We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).
+
 ### DMC1
@@
 
-Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* were compared by t-tests.
-SC morphology was used to stage spermatocytes in the zygotene, leptotene, or pachtyene cell stage.
+Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
+
+SC morphology and CREST/centromere number was used to stage spermatocytes **Certeria for early and late zygotene** - (bright staining - )
+early zygotene defined as seperate centromere pairing -- mostly complete axis signal.
+late zygotene judged as -- low degree of centromere pairing -- and more homologous pairing (but not fully synapsis).
+Differences in DMC1 counts per cell were tested by ttests.
+<numbers of cell?>
 
 ### MLH1 (Analytical Framework)
+
+```{r MLH1.analysis, echo=FALSE}
+
+#1.certeria
+#2. model framework 
+#2A (logic of fixed and random effects) -- 
+#  2b. packages used
+
+```
+
+**certeria**  MLH1 foci were quantified from cells with intact and compelete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score, between 1 to 5, was asigned to each cell based on staining quality and general spread of bivalents.
+<cell with small defects were included -- and coded as a score of 5 - but these cell weren't included in the final analysis.>  The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
+
+**the XY foci is not counted / analyzed in this study**
 
@@
 
+**partioning variation in recombination rate**
+
 -Our goal -- design an utilize a mixed model for our analytical framework - which leverages / accounts for the hierarchical/nested nature of categories across all the individual mice (e.g.
 sex, strain subspecies).
-
-MLH1 foci were quantified from cells with intact and compelete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score, between 1 to 5, was asigned to each cell based on staining quality and general spread of bivalents.
-<cell with small defects were included -- and coded as a score of 5 - but these cell weren't included in the final analysis.>  The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
-
-```{r MLH1.analysis, echo=FALSE, include=FALSE}
-#MLH1 + Bivalent
-
-#-models (how several were used)
-#the code of models
-#the tests / analysis of models
-```
 
@@
 The main effects and their interaction of subspecies and sex were coded as fixed.
-While, strain (or genetic background) was coded as a random effect to reflect a random sample of natural genetic variation from each subspecies.
-A strains to have a specific sex effect was also included.
 
-We applied the following models to analyze several dependent variables: mouse average mlh1 count per cell, the mouse average of variance in crossover count per cell, mouse mean total SC per cell, mouse mean single SC length, mouse mean IFD^norm^ and IFD^raw^, and position of crossover for single crossover bivalents.
+strain (or genetic background) was coded as a **random** effect to reflect a random sample of natural genetic variation from each subspecies.
+
+In addition to mouse mean of MLH1 count per cell
+**to expand the analysis of meiotic features** We applied the following models to several dependant variables (mouse averages) <consider a figure/diagram>
+
+ the mouse average of variance in crossover count per cell,
+
+ mouse mean total SC per cell, mouse mean single SC length, mouse mean IFD^norm^ and IFD^raw^, and position of crossover for single crossover bivalents.
 

### 2020-04-21 doc/Results_v2.rmd
changing the Rdata files across scripts

@@
                  nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
+                 mean.MLH1 = mean( nMLH1.foci),
+                 var.MLH1 = var( nMLH1.foci)
 )
@@
                          Nmice = length(unique(mouse)),
-                         Ncells  = length(adj_nMLH1.foci),
-                         mean_co = format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3),
+                         Ncells  = length( nMLH1.foci),
+                         mean_co = format(round(  mean( nMLH1.foci), 3 ), nsmall=3),
                          var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd(adj_nMLH1.foci), 3),
+                         sd   = round(sd( nMLH1.foci), 3),
                          se   = round(sd / sqrt(Ncells), 3),
@@
                  mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
 
@@
                  nmice = length(unique(mouse)),
-                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
 
@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(adj_nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(adj_nMLH1.foci),
-                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd(adj_nMLH1.foci), 3),
+                        Ncells  = length( nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean( nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv( nMLH1.foci),
+                        var = format(round(   var( nMLH1.foci),3), nsmall=3),
+                        sd   = round(sd( nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3)
@@
 #        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +
+#ylim( c((min(MLH1_data$ nMLH1.foci)-3) ), (max(MLH1_data$ nMLH1.foci)+3)) +
 

### 2020-04-21 src/Result_Setup.Rmd
changing the Rdata files across scripts

@@
                  nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci),
-                 var.MLH1 = var(adj_nMLH1.foci)
+                 mean.MLH1 = mean(nMLH1.foci),
+                 var.MLH1 = var(nMLH1.foci)
 )
@@
 #save Rdata
-save.image("~./MLH1repo/data/Results_setup_4.17.20.RData")
+save.image("~./MLH1repo/data/Results_setup_4.21.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-04-22 doc/Results_v2.rmd
slimming down results v2

@@
 
-```{r organizing.data, echo=FALSE }
+```{r FOR.PLOT.organizing.data, echo=FALSE }
 #set the order of another column, based on another variable. consistant order for ploting
-
-#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
-#-ene endings refer to the cells/units within the cell stage? (I think)
 
@@
 
-AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
-                 .(strain,sex),
-                 summarize,
-                 sum.ncells = sum(Ncells),
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co),
-                 AV.var_MLH1 = mean(var, na.rm=TRUE)
-
-)
-
-#there might be an error for changing wd
-#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
-
 AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
@@
 
-```{r misc.setup, echo=FALSE}
-
-colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
-                       #blues
-
-                       'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4',
-
-                       'TOM'= 'indianred1','AST' = 'tan2',
-                       'CZECH'= 'brown1',
-
-                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
-
-                    'CAST'='green','HMI'= 'forestgreen', #greens
-
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
-                    'CAROLI'='gold3',
-
-                    "F1"='grey', "other" = 'grey'
-                    )
-
-color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
-#4!!
-
-#show_col(hue_pal()(5))
-
-#add Rec group,
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
-#apply this before ploting, to display strains in desired order
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, category, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-
-```
-
-```{r Q1.rbar, echo=FALSE}
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-
-#read.csv of gamma estimates from above
-gamma_estimates <- read.csv("~./MLH1repo/results/gamma_estimates.csv", header = TRUE)
-
-table_gamma_estimates <- read.csv("~./MLH1repo/table.csv", header = TRUE)
-
-#for(e in 1:length(cleanCurated_BivData$fileName)){
-#  mini.IFD <- calq.intra.rbar(mini.IFD)
-#  cleanCurated_BivData$rbar[e] <- new.calq.intra.rbar(cleanCurated_BivData[e,])
-#  }
-
-#calq the mouse / strain means for rbar
-
-#summarize_rbar_table <- ddply(.data=cleanCurated_BivData,
-#                 .(sex, strain),
-#                 summarize,
-#                 mean.rbar.all = mean(rbar, na.rm = TRUE),
-
-#                 mean.rbar.2CO = mean(rbar[which(hand.foci.count == 2)], na.rm = TRUE),
-#                 mean.rbar.3CO = mean(rbar[which(hand.foci.count == 3)], na.rm = TRUE)
-#)
-#the 3 is negative...
-
-#how best to merge rbar and gamma estiamtes (which estimate)
-
-raw.gamma_estimates <-  gamma_estimates %>% filter(type.IFD == "raw.IFD") %>% filter( (class == "2CO") | (class == "3CO") )
-
-table <- raw.gamma_estimates
-#write.table(table, "~./MLH1repo/table.csv", sep=",",row.names = TRUE)
-
-#write.table(summarize_rbar_table, "~./MLH1repo/RBAR.table.csv", sep=",",row.names = TRUE)
-
-#what comparison of IFD and rbar would be logical
-#plot rbar and gamma
-ll <- ggplot(table_gamma_estimates, aes(y=mean.rbar, x = all.biv.gamma, color = class, shape = sex ))+geom_point()+facet_wrap(~strain)+ ggtitle("")+ylim(c(0,.5))
-
-#mean rbars are up around 
-#SKIVE male have extra (very strong interference)
-#there's so much variance across the gamma estimates -- 
-#most all females seem to have stronger interfernce
-#there don't seem to be consistant patterns
-
-#this plot shows much higher 2CO interference (via the gamma estimat from nrmIFD) in high rec males
-#across other classes of bivalents
-
-#most all other estiamtes of gamma estimates are around 4
-#
-
-#house mouse raw nu is around 11
-
-```
-
-```{r MLH1.FEMALE.TABLE, echo=FALSE}
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-#Mol
-#Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
-
-#Dom 
-Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
-
-```
-
-```{r MLH1.MALE.TABLE, echo=FALSE}
-#AP_mouse_table.MLH1_HQ
-
-#find the table that includes the las image data
-male_MLH1mouse.av <- AP_mouse_table_w.Ages
-
-male_MLH1mouse.av <- add_sex(male_MLH1mouse.av)
-male_MLH1mouse.av <- add_strain(male_MLH1mouse.av)
-male_MLH1mouse.av <- add_subsp(male_MLH1mouse.av)
-male_MLH1mouse.av <- add_species(male_MLH1mouse.av)
-
-fomo.male_mouse.av <-  male_MLH1mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-#Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-```
-
 ```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
 #DF.HetC.MixedModel.HQ
-
 #this blck is duplicated in setupData.rm
@@
 DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#remove the extra merged cols
-#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-#                            subsp.y,subsp.x,
-#                    category.y,category.x, strain.x,
-#                    strain.y, mouse.1  )  )
 
@@
 
-```
-
-```{r MLH1.HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
-#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
-#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
-#remove quality 5 cells
-
-#outgroups
-#SPIC
-HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
-#unique(HetC.SPIC$mouse)
-#non-passing mice
-SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
-HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
-#calq sex mean
-SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
-
-#SPRET
-HetC.SPRET <- MLH1_data[(MLH1_data$strain == "SPRET")&(MLH1_data$quality < 5),]
-#unique(HetC.SPRET$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-SPRET.mean.MLH1.obs <- ddply(.data=HetC.SPRET, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-SPRET.HETC.val <- SPRET.mean.MLH1.obs$mean.MLH1[1] / SPRET.mean.MLH1.obs$mean.MLH1[2]
-
-#CAROLI
-HetC.CAROLI <- MLH1_data[(MLH1_data$strain == "CAROLI")&(MLH1_data$quality < 5),]
-#unique(HetC.CAROLI$mouse)
-#non-passing mice
-#SPRET.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAROLI.mean.MLH1.obs <- ddply(.data=HetC.CAROLI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-#cast
-HetC.CAST <- MLH1_data[(MLH1_data$strain == "CAST")&(MLH1_data$quality < 5),]
-#unique(HetC.CAST$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-CAST.mean.MLH1.obs <- ddply(.data=HetC.CAST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-CAST.HETC.val <- CAST.mean.MLH1.obs$mean.MLH1[1] / CAST.mean.MLH1.obs$mean.MLH1[2]
-
-#HMI
-HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
-#unique(HetC.HMI$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
-HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-#dom
-
-#WSB
-HetC.WSB <- MLH1_data[(MLH1_data$strain == "WSB")&(MLH1_data$quality < 5),]
-#unique(HetC.WSB$mouse)
-#non-passing mice
-WSB.remove.mice <- c("8oct14_WSB_f1", "18nov17_WSB_f5", "16jun15_WSB_f3","10mar15_WSB_m2")
-HetC.WSB <- HetC.WSB[ ! HetC.WSB$mouse %in% WSB.remove.mice, ]
-#calq sex mean
-WSB.mean.MLH1.obs <- ddply(.data=HetC.WSB, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-WSB.HETC.val <- WSB.mean.MLH1.obs$mean.MLH1[1] / WSB.mean.MLH1.obs$mean.MLH1[2]
-
-#LEW
-HetC.LEW <- MLH1_data[(MLH1_data$strain == "LEW")&(MLH1_data$quality < 5),]
-unique(HetC.LEW$mouse)
-#non-passing mice
-LEW.remove.mice <- c("4jan17_LEW_f2")
-HetC.LEW <- HetC.WSB[ ! HetC.LEW$mouse %in% LEW.remove.mice, ]
-
-#calq sex mean
-LEW.mean.MLH1.obs <- ddply(.data=HetC.LEW, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-LEW.HETC.val <- LEW.mean.MLH1.obs$mean.MLH1[1] / LEW.mean.MLH1.obs$mean.MLH1[2]
-
-#G
-HetC.G <- MLH1_data[(MLH1_data$strain == "G")&(MLH1_data$quality < 5),]
-#unique(HetC.G$mouse)
-#non-passing mice
-G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-HetC.G <- HetC.G[ ! HetC.G$mouse %in% G.remove.mice, ]
-
-#calq sex mean
-G.mean.MLH1.obs <- ddply(.data=HetC.G, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#PERC
-HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
-#unique(HetC.PERC$mouse)
-#non-passing mice
-#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
-
-#calq sex mean
-PERC.mean.MLH1.obs <- ddply(.data=HetC.PERC, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-G.HETC.val <- G.mean.MLH1.obs$mean.MLH1[1] / G.mean.MLH1.obs$mean.MLH1[2]
-
-#MUSC
-
-#PWD
-HetC.PWD <- MLH1_data[(MLH1_data$strain == "PWD")&(MLH1_data$quality < 5),]
-#unique(HetC.PWD$mouse)
-#non-passing mice
-PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-PWD.mean.MLH1.obs <- ddply(.data=HetC.PWD, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-PWD.HETC.val <- PWD.mean.MLH1.obs$mean.MLH1[1] / PWD.mean.MLH1.obs$mean.MLH1[2]
-
-# CZECH
-HetC.CZECH <- MLH1_data[(MLH1_data$strain == "CZECH")&(MLH1_data$quality < 5),]
-#unique(HetC.CZECH$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
-
-# SKIVE
-HetC.SKIVE <- MLH1_data[(MLH1_data$strain == "SKIVE")&(MLH1_data$quality < 5),]
-#unique(HetC.SKIVE$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-SKIVE.mean.MLH1.obs <- ddply(.data=HetC.SKIVE, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-SKIVE.HETC.val <- SKIVE.mean.MLH1.obs$mean.MLH1[1] / SKIVE.mean.MLH1.obs$mean.MLH1[2]
-
-# KAZ
-HetC.KAZ <- MLH1_data[(MLH1_data$strain == "KAZ")&(MLH1_data$quality < 5),]
-#unique(HetC.KAZ$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-KAZ.mean.MLH1.obs <- ddply(.data=HetC.KAZ, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-KAZ.HETC.val <- KAZ.mean.MLH1.obs$mean.MLH1[1] / KAZ.mean.MLH1.obs$mean.MLH1[2]
-
-#TOM
-HetC.TOM <- MLH1_data[(MLH1_data$strain == "TOM")&(MLH1_data$quality < 5),]
-#unique(HetC.TOM$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-TOM.mean.MLH1.obs <- ddply(.data=HetC.TOM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-TOM.HETC.val <- TOM.mean.MLH1.obs$mean.MLH1[1] / TOM.mean.MLH1.obs$mean.MLH1[2]
-
-#AST
-HetC.AST <- MLH1_data[(MLH1_data$strain == "AST")&(MLH1_data$quality < 5),]
-#unique(HetC.AST$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
-#calq sex mean
-AST.mean.MLH1.obs <- ddply(.data=HetC.AST, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
-
-#Molosinus
-
-#MSM
-HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
-unique(HetC.MSM$mouse)
-
-#non-passing mice
-MSM.remove.mice <- c("12sep16_MSM_f3")
-HetC.MSM <- HetC.MSM[ ! HetC.MSM$mouse %in% MSM.remove.mice, ]
-
-#calq sex mean
-MSM.mean.MLH1.obs <- ddply(.data=HetC.MSM, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-MSM.HETC.val <- MSM.mean.MLH1.obs$mean.MLH1[1] / MSM.mean.MLH1.obs$mean.MLH1[2]
-
-#MOLF
-HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
-unique(HetC.MOLF$mouse)
-
-#non-passing mice
-MOLF.remove.mice <- c("27oct18_MOLF_m3")
-HetC.MOLF <- HetC.MOLF[ ! HetC.MOLF$mouse %in% MOLF.remove.mice, ]
-
-#calq sex mean
-MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean( nMLH1.foci),
-                 var.MLH1 = var( nMLH1.foci)
-)
-
-MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
-
-```
-
-```{r MLH1.HetC.plot, echo=FALSE, warning=FALSE}
-#HETC plot
-#change the main data?
-
-HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length( nMLH1.foci),
-                         mean_co = format(round(  mean( nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd( nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#
-#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]
-
-HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
-                         Nmice = length(unique(mouse)),
-                         Ncells  = length( nMLH1.foci),
-                         mean_co = format(round(  mean( nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd( nMLH1.foci), 3),
-                         se   = round(sd / sqrt(Ncells), 3),
-                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
-)
-#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
-
-#remove the 'other'
-
-HetC_strain_table.M$sex <- "male"
-HetC_strain_table.F$sex <- "female"
-
-HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
-
-HetC_strain_table <- add_subsp(HetC_strain_table)
-
-#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
-
-HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
-HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
-#str(HetC_strain_table)
-
-Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
-  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
-
-```
-
-```{r juvimice.list, echo=FALSE}
-juvi_mice <- c(
-  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
-  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
-  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
-
-  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
-   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
-  )
-
-#REMOVE JUVI mice
-#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
-#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
-
-```
-
-```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
-
-#this block should be in the data set up
-#the mouse averages will be used
-
-#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
-cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
-
-#cleanCurated_BivData
-#removed the no hand foci count
-
-mouse.avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-mouse.avs_4MM <- add_sex(mouse.avs_4MM)
-mouse.avs_4MM <- add_strain(mouse.avs_4MM)
-mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-mouse.avs_4MM <- add_category(mouse.avs_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
-mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
-mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
-
-#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-#code if males are high or low?
-
-#REMOVE JUVI mice
-juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
-mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
-
-#assign high and low
-#divide musc males into groups
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
-
-mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
-
-```
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#add this to setup
-
-#remove female, 
-#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
-
-#Curated_BivData_male
-
-male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-male.av_4MM <- add_sex(male.av_4MM)
-male.av_4MM <- add_strain(male.av_4MM)
-male.av_4MM <- add_subsp(male.av_4MM)
-male.av_4MM <- add_category(male.av_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
-male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
-male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-
-#REMOVE JUVI mice
-juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
-mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
-
-```
-
-```{r totSC_setup, echo=FALSE}
-#SC.skel
-#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included
-
-SC.skel.OG <- SC.skel #5683
-
-#MERGEING MLH1 with SKEL
-MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-
-MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
-#5110,
-
-OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
-
-#check for DUPs, #check for duplicatations of the random name
-MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
-tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
-
-MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
-
-MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
-MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
-#table(MLH1.merge.Skel$category)
-
-#remove outliers
-MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
-
-```
-
-```{r totSC.viz4outliers, echo=FALSE}
-
-# -- EXAMINE the ggplot scatter plots for outliers
-tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
-
-#round 1
-SC.skel.outliers_round1 <- c(
-  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
-  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
-            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
-
-    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
-                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
-              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
-          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",
-
-  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
-          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",
-
-       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
-                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",
-
-   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
-   "6may15_17apr15_PWD_f1_sp1_7_rev",
-
-      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
-            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",
-
-    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
-    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
-  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
- "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
- "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
- "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
-"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,
-
-  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
-    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
-
-# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
-"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
-
-#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
-"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
-
-#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
-"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
-"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
-"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
-
-#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
-"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
- "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
- "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
- "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
- "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
- "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
-"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
-"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
- "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
- "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
- "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
- "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
-
-MLH1.merge.Skel_1 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
-
-remove.mice_round2 <- c(
-    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
-    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
-
-    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
-  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
-"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
-  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",
-
-  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
-  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
-"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
-"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
-
-# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
-"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
-"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
-
-#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
-"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
-#SPIC male  x
-#SPIC female  x
-#TOM x
-#AST x
-#CZECH male x
-
-#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
- "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
-#SPRET female X
-#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
-"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
-)
-#3remove.mice_round2
-
-MLH1.merge.Skel_2 <- MLH1.merge.Skel_1[!MLH1.merge.Skel_1$fileName %in% remove.mice_round2, ]
-#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
-#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
-all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
-
-MLH1.merge.Skel <- MLH1.merge.Skel_2
-
-MLH1.merge.Skel <- add_species(MLH1.merge.Skel)
-
-#adjust 
-MLH1.merge.Skel_2$mouse <- as.factor(MLH1.merge.Skel_2$mouse)
-MLH1.merge.Skel_2$nMLH1.foci <- as.numeric(MLH1.merge.Skel_2$nMLH1.foci)
-MLH1.merge.Skel_2$quality <- as.numeric(MLH1.merge.Skel_2$quality)
-
-```
-
-```{r tot.skel.table, echo=FALSE}
-#calq the correlation for tota.SC and mlh1 - after mouse averages
-#use MLH1.merge.Skel_2 -- quality 1-4
-
-totSC.mouse.av.table <- ddply(.data= ( MLH1.merge.Skel_2 %>% filter(quality < 5) ), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
-
-#unorder the factors
-totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
-totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
-totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
-totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
-
-OG.totSC.mouse.av.table <- totSC.mouse.av.table
-
-#REMOVE JUVI mice
-tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#ADD CORRELATION
-for(g in length(totSC.mouse.av.table$mouse) ){
-totSC.mouse.av.table$cor.SCskel.MLH1[g] <- cor(totSC.mouse.av.table$mean.MLH1[g], totSC.mouse.av.table$mean.skel[g])
-}
-
-mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
-total SC area")
-
-mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-```
-
-```{r tot.SC.setup, echo = FALSE}
-#calc=q correlation, total.sc and mean MLH1
-
-totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
-                 .(sex,strain),
-                 summarize, 
-                 ncells = sum(ncells),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(mean.skel, na.rm = TRUE),
-                 mean.bin =  mean(mean.bin, na.rm = TRUE)
-
-)
-
-totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
-
-```
-
-```{r LONG.SHORT.biv_setup, echo=FALSE}
-
-#FROM q1
-#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#from Q2 IFD
-#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-##long biv
-#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-#FROM q1 
-REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO  
-                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                         )
-
-REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
-
-REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
-REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
-REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
-                                           0)))
-
-REAL.short_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1, 0))
-
-###LONG BIV
-
-REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                       n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-                         )
-
-REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
-
-REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
-REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
-REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
-                                           0)))
-
-REAL.long_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1, 0))
-
-#find which one 
 ```
@@
 
-```{r basic.numbers, echo=FALSE, eval=FALSE}
+```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
 

### 2020-04-22 src/Result_Setup.Rmd
slimming down results v2

@@
 
-#this blck is duplicated in setupData.rm
-
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-
-#What are the main DFs for this code chunk
-#AP_mouse_table_w.Ages   <- used to remove males which are too old (includes outgroups) 
-#DF.HetC.MixedModel   <- only Mmusculus
-#DF.HetC.MixedModel.HQ <- only mice of certain age, Q5 excluded
-
-#variance is still chr
-#AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
-
-#use the merged DF to filter by age 
-#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-#AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
 DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
@@
 
-#REMOVE EXTRA STRAINS
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-#DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
-
 #AP_mouse_table excluide Qual5 cells
@@
 
+DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
+
+#main df with variables coded as chr
+DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+
+#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+
+all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
+                 .(sex, strain),
+                 summarize,
+                 all.cells = sum(Ncells)
+)
+
+```
+
+```{r MLH1.MIXED.MODEL.setup.2, echo=FALSE, warning=FALSE, eval=FALSE}
+#DF.HetC.MixedModel.HQ
+#this blck is duplicated in setupData.rm
+
+#https://rcompanion.org/handbook/G_03.html
+#testing random effects within lme
+
+DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+
+#AP_mouse_table excluide Qual5 cells
+DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
 DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)

### 2020-04-22 src/SetupData.Rmd
slimming down results v2

@@
 MLH1_data <- add_species(MLH1_data)
+```
+
+```{r misc.setup, echo=FALSE}
+
+colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
+                       #blues
+
+                       'PWD'= 'red2',
+                       'SKIVE'= 'coral1','KAZ'= 'coral4',
+
+                       'TOM'= 'indianred1','AST' = 'tan2',
+                       'CZECH'= 'brown1',
+
+                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
+
+                    'CAST'='green','HMI'= 'forestgreen', #greens
+
+                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'CAROLI'='gold3',
+
+                    "F1"='grey', "other" = 'grey'
+                    )
+
+color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
+#4!!
+
+#show_col(hue_pal()(5))
+
+#add Rec group,
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
 ```

### 2020-04-22 doc/MatMeth.Rmd
slimming down results v2
transfering code to setup file

@@
 
+**M1s code**
+
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+summary(MLH1var_M1)
+LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
+
+LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
+sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
+
+tested coefficients from the mixed models (using the lmer() function) with chisquare test between full and reduced models for mixed. the random effects were tested with the exactRLRT()
+
+M2 to M4 are glm -- use lm()
+
+MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+MLH1var_M2_sum <- summary(MLH1var_M2)
+
 ```{r MLH1.analysis, echo=FALSE}
@@
 
+#
+#
+#partitioning variance
 ```
+
+Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
+-ene endings refer to the cells/units within the cell stage? (I think)
 
@@
 **M4. Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
+
+**M4.2** removed the subspecies coefficient for
 

### 2020-04-22 doc/Results_v2.rmd
transfering code to setup file

@@
 
+#write the table out
+
 ```
@@
 
-```{r female.dif, echo=FALSE}
-
-#how much greater are the high strains?
-G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
-mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
-
-LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
-mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
-
-#how much greater is MSM female
-MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
-mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
-
-sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
-mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
-
-G.female.diff <- (mean.G.female / mean.all.other)
-LEW.female.diff <- (mean.LEW.female / mean.all.other)
-MSM.female.diff <- (mean.MSM.female / mean.all.other)
-```
-
 ```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
 
 #variance plots
-musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
 musc.female.VAR.plot
 
-#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-#Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.female.VAR.plot
+
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
 Dom.female.VAR.plot
@@
 
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
-
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-```
-
-```{r FEMALE.VAR.MODELS, echo=FALSE}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
-
-lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
-lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
 <!-- male section -->
@@
 
-```{r male.diff, echo=FALSE}
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
-#mean(sans_LewRec_male$mean_co)
-
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
-#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
-#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
-
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
-#mean(SKIVE_male$mean_co) /  
-```
-
 ```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
@@
 
-```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-#summry.male.strain.effects <- summary(glm.male.strain.subsp)
-#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#1. determine which strain effects are sig
-
-#M2
-MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
-sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
-```
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-```
-
 # Partitioning variation in recombination rate
-
-```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}
-
-#DF.HetC.MixedModel.HQ
-
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
-
-MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
-sex.results =  MLH1.MM_sex, 
-               interaction.pval = MLH1.MM_interaction,  rand.pval = 
-MLH1.MM_strain)
-
-```
-
-```{r MLH1.glmM2, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
-
-#strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
-
-summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
-
-#table of just coeff anf pvalue
-#round(summ$coefficients[,c(1,4)],5 )
-
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-```
-
-```{r MLH1.glm.M3, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
-
-summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
-
-#round(summ.strain.fixed$coefficients[,c(1,4)],5)
-
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-```
-
-```{r bret.comment.var, echo=FALSE}
-
-```
 
@@
 
-# Within mouse variance in CO count per cell
+# Within mouse **variance** in CO count per cell
 
@@
 ```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-
 MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
@@
 ```{r M2.MLH1.var, echo=FALSE, include=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-#why are these named M3? -- rename to M2
 
 MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-
 MLH1var_M2_sum <- summary(MLH1var_M2)
 
-#summary(MLH1var_M2)
 #anova(MLH1var_M2)  
@@
 MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-
+#summary(MLH1cv_M2)
 #anova(MLH1cv_M2)
@@
 MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
 #summary(MLH1var_M3_Q12)
@@
 #Q12 cv
-MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+#MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
 #round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
@@
 
-# Q1 Sex Differences in the Length of the Synaptonemal Complex
+# Q1. Sex Differences in the Length of the Synaptonemal Complex
 
@@
 
-#TOTAL SC AREA!!
+#TOTAL SC AREA
 lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
@@
 
-#Q1.lmer.totSC_results
-
-#G
 ```
@@
 
-```{r Q2.SC.length.bret.comments, echo=FALSE}
-#correlation, total sc and Mean MLH1
-
-```
-
 ```{r Q2.short.long.ttest, echo=FALSE}
@@
 
-Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
-
-#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
-#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
-#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)
-
-Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
+Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
+
+#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
+#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
+#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)
+
+#M4.2? (removed subspecies?)
+
+Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
 #round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
@@
 #Dom
-Q2.glm_mus.av_totSC_M2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
-sum.Q2.glm_mus.av_totSC_M2_DOM  <- summary(Q2.glm_mus.av_totSC_M2_DOM)
+Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
+sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
 #there are significant strain effects in DOM,
@@
 #Musc
-Q2.glm_mus.av_totSC_M2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
-sum.Q2.glm_mus.av_totSC_M2_MUSC  <- summary(Q2.glm_mus.av_totSC_M2_MUSC)
+Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
+sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
 #KAZ -- is the only significant effect
@@
 #Mol
-Q2.glm_mus.av_totSC_M2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
-sum.Q2.glm_mus.av_totSC_M2_MOL  <- summary(Q2.glm_mus.av_totSC_M2_MOL)
-
-#these pair-wise t.test aren't working
-#dom
-#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean.skel, 
-#                 ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$strain )
-
-#musc
-#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$mean.skel, 
-#                 ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$strain )
-
-#mol
-#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean.skel, 
-#                 ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$strain )
+Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
+sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)
 
@@
 
-#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
-#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
-#table(mouse.av.long.biv.table$mouse)
-#Q2mouse.av.long.biv.table
-
 #this dataset is all males
-glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
-#
-
-glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
+glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+
+glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
 #round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
 
-glmQ2_long.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_long.biv_M2_DOM <- summary(glmQ2_long.biv_M2_DOM)
+glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
 
 #MUSC
-glmQ2_long.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_long.biv_M2_MUSC <- summary(glmQ2_long.biv_M2_MUSC)
+glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
 
 #MOL can't be done -- only 1 mouse for each MSM and MOLF
-glmQ2_long.biv_M2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
-sum.glmQ2_long.biv_M2_MOL <- summary(glmQ2_long.biv_M2_MOL)
+glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
+sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
 
@@
 
-```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
-## run the tests to check following expectations
-
-#1.Dom strains not different from ech other
-#2. PWD and SKIVE are different from the other strains
-#3.MSM and MOLF are different from each other
-
-#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]
-
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
-#the subsp labels get messed up, unless they are designated chr first
-Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )
-
-Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
-Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
-#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
-#total sc sig, 1CO sig, 2CO sig
-
-Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
-#0.000369 ***
-Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
-#0.000816 ***
-Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
-#0.0148 *
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
-#making plots above show that mouse average SC lengths are lower could there be microscope issue?
-
-#make sure to remove some of the juvi mice!! from the AP_mouse table
-show.Dom.male.CO <- ggplot(
-male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()
-
-Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
-#P not sig for CO mean
-
-#seems like there are more mice in the MLH1 data set than bivData
-#
-
-#WSB all dissected and imaged in '15
-#G imaged 15 - 16
-#LEW imaged 15 to 17
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "1CO SC Length", xlab = "")
-
-ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
-          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-          ylab = "2CO SC Length", xlab = "")
-```
-
-```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}
-
-#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
-#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av
-
-male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
-male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
-male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]
-
-Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
-#PWD and MSM are most diff from SKIVE?
-
-Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
-#the pooled SC value tests are sig
-
-#the mol comparisons are different due to adj p values
-
-Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
-#G-WSB is sig, WSB-LEW are slightly sig
-
-Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
-#no sig (only 3) -- does include CZECCH
-
-#non-sig, with mol pooled with musc
-
-Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)
-
-Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)
-
-Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
-#non of these sig with the mulitple adjust
-
-Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)
-
-#pairwise.t.test(Ozone, Month, p.adj = "bonf")
-```
-
 Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
@@
 
-<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->
-
-```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
-#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
-
-#High = MSM, PWD, SKIVE
-Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-#high = PWD and MSM
-Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-```
-
-```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
-
-#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
-
-#Musc
-Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
-# p-value = 0.4
-
-#Mol
-Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
-# p-value = 0.07
-# exclusive rec group not sig
-```
-
-```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
-#these don't have mean.first.foci
-
-#short -- real.long.bivData -- REAL.short_mouse_table
-Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
-
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
-#p-value = 0.3
-
-Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
-# p-value = 0.1
-
-#remove SKIVE??
-```
-
-```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
-
-#real.long.bivData
-#single biv level
-Rec.group_LONG.BivData_F1post_Ttest <- t.test(
-  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.5
-
-Rec.group_BivData_F1post_Ttest <- t.test(
-  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.2
-
-#
-#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
-#p-value = 0.2
-
-#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
-#p-value = 0.4
-
-```
-
-```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
-
-Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
-#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-#with.chrm.legth  ?
-Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-
-#M2
-Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients
-
-```
-
 ```{r Q2.1COpos.show.MOL, echo=FALSE}
@@
 #CHRM.length is very significant-- some of the same subsp and strains are sig     
-```
-
-```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
-#remove this cell,
-
-Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
-
-Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
-
-#making new models
-
-Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
-Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
-
-chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
-
-sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
-#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
-
-#summary(chrm.1CO.pos.chrmLength_strain)
-
-```
-
-```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
-#isolate all musc
-#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
-
-##long biv
-male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean_F1 = mean(Foci1, na.rm=TRUE),
-                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
-
-                        mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
-
-                                               )
-#add subsp
-
-mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
-mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
-
-mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
-mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#add rec group
-mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
-                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
-```
-
-```{r Q2.1COpos.glm, echo=FALSE, eval=FALSE}
-male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
-
-Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO,  family=binomial(link="logit"))
-#summary(Log.glmQ2_M2.1COnrm)
-
-#Musc
-Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
-              family=binomial(link="logit"))
-#summary(Log.Q2.F1_PER_Musc)
-
-#Mol
-Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
-              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
-              family=binomial(link="logit"))
-
-#summary(Log.glmQ2_M2.1COnrm_Mol)
-
 ```
@@
 # Q2 Evolution of interference is associated with genome wide recombination rate evolution
-
-```{r Q2.IFD.long.short.setup, echo=FALSE}
-#long BivData
-male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
-
-#show
-Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
-
-#long biv's - short IFD rates
-
-LongBiv_IFD_mus.rate <- ddply(male.long.biv.2CO_IFD %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-
-                   N.obs  = length(Obj.ID),
-
-                  mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
-                  mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                  var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                  #mean_IFD_short.removed
-                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
-                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)] ),
-
-                inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                n.close.foci = length(which(IFD1_PER <= .3)),
-
-                off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                rate.close.foci = n.close.foci / N.obs
-
-                                              )
-
-#run t.tests on long biv
-LongBiv_IFD_mus.rate <- add_strain(LongBiv_IFD_mus.rate)
-LongBiv_IFD_mus.rate <- add_category(LongBiv_IFD_mus.rate)
-LongBiv_IFD_mus.rate <- add_subsp(LongBiv_IFD_mus.rate)
-
-#add rec.group
-LongBiv_IFD_mus.rate$Rec.group <- ifelse(grepl("PWD", LongBiv_IFD_mus.rate$strain), 1, 
-                                ifelse(grepl("MSM", LongBiv_IFD_mus.rate$strain), 1,
-                                 ifelse(grepl("SKIVE", LongBiv_IFD_mus.rate$strain), 1, 0)))
-
-```
-
-```{r Q2.IFD.long.biv_t.test, echo=FALSE}
-#short removed PER
-Short.removed_BivData.LongBiv_IFD.PER_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
-                                 LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
-
-#Short.removed_BivData.LongBiv_IFD.PER_Ttest$p.value
-#0.0083
-
-#ABS short removed
-Short.removed_BivData.LongBiv_IFD.ABS_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
-                                 LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
-
-#Short.removed_BivData.LongBiv_IFD.ABS_Ttest$p.value
-#0.047
-```
 
@@
 
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
-#t.test for IFD
-#mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
-
-#t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == #"male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
-#   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == #"Dom"))$mean_IFD.2CO_PER   )
-
-t.test_mouse.av.IFD_PER_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_PER   )
-
-t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER   )
-
-#ABS
-
-t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
-```
-
 ```{r Q2.IFD.glm.test, echo=FALSE}
@@
 
-```
-
-```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
-
-Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
-
-#M.Len.ABS
-Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.raw)
-
-#M.Len.PER
-Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
-#summary(Q2.BIV.LEN.glm_IFD.PER)
-
-#M.Len.raw
-Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
-
-##M.Len PER
-Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
-#summary(Q2.BIV.LEN.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-```
-
-```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
-
-#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
-
-close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("mouse"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                      mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
-                      mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm=TRUE),
-
-                  #mean_IFD_short.removed
-                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
-
-                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)], na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
-                      n.close.foci = length(which(IFD1_PER <= .3)),
-
-                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
-
-                      rate.close.foci = n.close.foci / N.obs
-)
-
-#mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-
-close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
-close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
-#add rec.group
-close.IFD.rates.mus.rate$Rec.group <- ifelse(grepl("PWD", close.IFD.rates.mus.rate$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates.mus.rate$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates.mus.rate$strain), 1, 0)))
-
-close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
-                         c("strain"), summarise,
-                        N.obs  = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
-                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
-                        var.IFD = var(IFD1_PER, na.rm = TRUE),
-
-                        #mean_IFD_short.removed
-                      mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
-
-                      inverse.close.foci = length(which(IFD1_PER > .3)),  
-                      n.close.foci = length(which(IFD1_PER < .3)),
-                      rate.close.foci = n.close.foci / N.obs,
-
-                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
-
-)
-
-close.IFD.rates_strain <- add_subsp(close.IFD.rates_strain)
-
-show.IFDrm <- ggplot(close.IFD.rates_strain, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences")+ geom_abline(slope = 1, intercept = 0)
-
-#close.IFD.rates.mus.rate
-show.IFDrm_mus <- ggplot(close.IFD.rates.mus.rate, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences") + geom_abline(slope = 1, intercept = 0)
-
-show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
-
-#add rec.group
-close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
-                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
-                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
-
-#POOL all short and 
-#chi square test
-high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
-                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
-#p-value = 0.03, slight dif tween 3 high rec strains
-```
-
-```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
-
-#Log regression all show significant prediction
-#IFD.PER
-log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER)
-
-#MUSC
-log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Musc)
-
-#Mol
-log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.PER.Mol)
-
-#SINGLE BIVALENT LEVEL
-Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata)
-
-#Musc
-Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.per_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-
-```
-
-```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
-
-#IFD.raw
-log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw)
-
-#MUSC
-log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Musc)
-
-#Mol
-log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.reg.Q2.IFD.raw.Mol)
-
-#SINGLE BIV LEVELs
-male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
-                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
-                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
-
-Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata)
-
-#Musc
-Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
-
-#Mol
-Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
-              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
-```
-
-```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
-#Full dataset
-male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
-
-#single bivalent
-Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
-#summary(Q2.glm_long.biv_M2.IFD.PER)
-#The PER.IFDs don't have significant strain effects
-```
-
-```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
-
-male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-#t.tests for the Rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
-                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
-
-#Rec.group.BivData_IFD.PER_Ttest$p.value
-#p-value < 2.2e-16
-
-#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
-Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
-                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
-
-#Rec.group_IFD.PER_Ttest$p.value 
-#p-value = 6.872e-07
-```
-
-```{r Q2.IFD_short.removed_t.test, echo=FALSE, include=FALSE, eval=FALSE}
-Short.removed_BivData_IFD.PER_Ttest <- t.test(Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 1], 
-  Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 0])
-
-Short.removed_BivData_IFD.PER_Ttest$p.value
-
-#
-Short.removed_BivData_IFD.PER_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
-                       close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
-Short.removed_BivData_IFD.PER_Ttest_mus.av$p.value
-
-Short.removed_BivData_IFD.ABS_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
-                                 close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
 ```

### 2020-04-22 src/Result_Setup.Rmd
transfering code to setup file

@@
 
+```
+
+```{r female.diference, echo=FALSE}
+
+#how much greater are the high strains?
+G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
+mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)
+
+LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
+mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)
+
+#how much greater is MSM female
+MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
+mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)
+
+sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
+mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)
+
+G.female.diff <- (mean.G.female / mean.all.other)
+LEW.female.diff <- (mean.LEW.female / mean.all.other)
+MSM.female.diff <- (mean.MSM.female / mean.all.other)
+```
+
+```{r male.difference, echo=FALSE}
+just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+
+#how much greater are the High Rec gwRRs
+just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
+                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
+                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
+
+#how much bigger are the high rec male means?
+sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+#mean(sans_LewRec_male$mean_co)
+
+PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
+
+SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+#mean(SKIVE_male$mean_co) /  
 ```
@@
 
-Het.C.plot <- ggplot(data= HetC_strain_table, aes(x= mean_co.x, y=mean_co.y, color = subsp, label=strain ))+ geom_point(aes(size=2)) + xlim(c(20,35)) + ylim(c(20, 35)) + geom_abline(slope = 1, intercept = 0)+labs(x="male MLH1 mean",y="female MLH1 mean")+ggtitle("Heterochiasmy Plot")+
-  scale_color_manual(values=c("royalblue", "limegreen",  "tomato", "slategray2", "plum", "grey")) + geom_text(data = HetC_strain_table, aes(label=HetC_strain_table$strain, x=mean_co.x, y= mean_co.y,  hjust = 0, nudge_x = 0.09, nudge_y = 0.09) )
-
 ```
@@
 
+RENAME the curate bivdata below
+
 ```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
@@
 
-#Curated_BivData_male
+#Curated_BivData_male  -- RE
 
@@
 
-mouse.av_sc_female <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "female",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1500,3000 ))+ggtitle("Female Mouse av.
-total SC area")
-
-mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
 ```
@@
 
+```{r MLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE}
+
+#DF.HetC.MixedModel.HQ
+
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1.MM.M1)$coefficients
+
+MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+
+MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
+sex.results =  MLH1.MM_sex, 
+               interaction.pval = MLH1.MM_interaction,  rand.pval = 
+MLH1.MM_strain)
+
+```
+
+```{r MLH1.glmM2, echo=FALSE}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
+#strain = fixed
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
+
+summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
+
+#table of just coeff anf pvalue
+#round(summ$coefficients[,c(1,4)],5 )
+
+#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
+```
+
+```{r MLH1.glm.M3, echo=FALSE}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
+
+summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
+
+#round(summ.strain.fixed$coefficients[,c(1,4)],5)
+
+#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+```
+
+```{r MLH1.M4, echo=FALSE}
+
+```
+
+```{r Female.specific.glm, echo=FALSE}
+
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
+
+#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
+
+#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+```
+
+```{r Male.specific.glm, echo=FALSE}
+# 1 way anova, is at least 1 strain different from the rest?
+# given a lm, how sig is the strain effect
+glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+
+#summry.male.strain.effects <- summary(glm.male.strain.subsp)
+#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
+
+#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+
+#1. determine which strain effects are sig
+
+#M2
+MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
+sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
+
+#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+
+#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+
+#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
+```
+
+```{r MALE.SPECIFIC.variance.model, echo=FALSE}
+## within mouse variance models 
+lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
+
+#all p values kinda high, MSM and TOM are slightly sig
+lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
+#non sig
+
+lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+```
+
 #MLH1 variance models
 
-M1 - subsp, sex, (1 | strain)
-M2 - subsp, sex, strain
+M1 - subsp, sex, (1 | strain)  (lmer)
+M2 - subsp, sex, strain  (glm / lm?)
 M3 - sex, strain
@@
 
+standardize the lm's and glms!
+
 ```{r M2.MLH1.var, echo=FALSE, include=FALSE}
@@
 #male
+glmMLH1Var_M4.male <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "male") ) )
 
 #female
+glmMLH1Var_M4.female <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "female") ) )
+
+glmMLH1Var_M4.femaleQ12 <- glm( var ~ subsp * strain,  data= (Q12_mouse_table_w.Ages %>% filter(sex == "female") ) )
+
+```
+
+```{r FEMALE.VAR.MODELS, echo=FALSE}
+## within mouse variance models 
+lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#round( #round isn't working for this sumary
+#lmVAR.female.strain_sum$coefficients[,c(1,4)]
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+
+lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
+
+lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
+lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
+
+#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
+#LEW sig
+
 ```
@@
 
+#Q2 section
+
+#Q2 SC
+
+```{r Q2.SC.short.long.setup, echo=FALSE}
+
+#table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+
+Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                         #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+                         )
+
+Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
+Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
+
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                        #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
+                         ) )
+
+Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
+
+Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
+Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
+
+Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
+```
+
+<!-- plots -->
+
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
+
+#nested boxplots, pooled Bivalents
+#x axis for hand.focicount -- good for seperating the points --- but it pools
+
+#prediction for Dom strains: low variation 
+Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
+
+#+theme(legend.position="none") 
+```
+
+```{r Q2.SC.long.biv_plots, echo=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#real.long.bivData
+#real.short.bivData
+
+shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+```
+
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
+
+#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+#MLH1_data <- MLH1_data %>%
+#  arrange(strain, sex, mouse) %>%
+#  mutate(file.name = factor(file.name))
+
+mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
+total SC area")
+
+mouse.av_totsc_male
+
+```
+
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE}
+
+#make plot of mouse averages for MLH1 and total SC
+
+bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
+
+X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+
+X.by.texas
+
+```
+
+```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
+
+#investigating potential skew / technical effects in the algorithm
+
+Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+#not sure what my expectations for this plot are,
+
+median(bep$oldest.image.date) #"2017-08-03"
+#there might be an issue with the algorithm performance 
+#given the parrelle diagonal lines -- 
+#this might be due to the dates imaged  2017-02-07
+
+bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
+
+Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
+
+#there's evidence that the algorithm might have some biased effects -- 
+#that will be investigated later.
+
+```
+
+```{r Q2.totSC.ttest, echo=FALSE}
+#totalSC high and low rec group
+
+Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
+                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+#p-value = 0.007
+
+#mol
+Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+
+#musc
+Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+
+```
+
+```{r Q2.short.long.ttest, echo=FALSE}
+
+#long and short 
+#mouse.av.short.biv.table
+#Q2mouse.av.short.biv.table
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
+
+#p-value = 0.9
+
+#mouse.av.long.biv.table
+Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
+#p-value = 0.2
+
+#mUSC-short
+Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-short
+#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
+
+#musc-long
+Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-long
+#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
+
+#not enought observations
+
+```
+
+```{r Q2.SC_HvL_allBiv, echo=FALSE}
+
+#High-Low_All-BivData
+
+Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#short
+Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#long
+Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+```
+
+```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
+
+#Dom-Musc
+Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
+                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
+#p-value = 0.5
+
+#Dom-Mol
+Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
+                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
+#p-value = 0.1
+```
+
+```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
+
+#DOM-musc-long
+Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.08
+
+#DOM-musc-short
+Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.09
+
+```
+
+```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
+
+#DOM-MUSC-ALL
+Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 4e-13
+
+#DOM-MOL-ALL
+Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 6e-10
+
+#DOM-musc-long
+Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.2
+
+#DOM-musc-short
+Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.06
+
+```
+
+```{r Q2.totSC.glm, echo=FALSE}
+
+Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
+
+#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
+#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
+#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)
+
+#M4.2? (removed subspecies?)
+
+Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
+#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
+
+totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+#within subspecies values
+
+#Dom
+Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
+sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
+#there are significant strain effects in DOM,
+
+#Musc
+Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
+sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
+#KAZ -- is the only significant effect
+
+#Mol
+Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
+sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)
+
+```
+
+```{r Q2.glm.SC.long, echo=FALSE}
+
+#this dataset is all males
+glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+
+glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+
+glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
+
+#MUSC
+glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
+
+#MOL can't be done -- only 1 mouse for each MSM and MOLF
+glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
+sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
+
+```
+
+```{r Q2.glm.SC.short, echo=FALSE}
+#Q2mouse.av.short.biv.table
+#mouse.av.short.biv.table
+
+glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
+sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
+
+#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
+
+glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
+glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
+
+#DOM
+glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
+
+#MUSC
+glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
+
+#MOL
+#within subspecies
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+```
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
+#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+#there's no rec group for this table
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
+
+plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
+#COMPARISONS ACROSS ALL CLASSES
+
+#MOUSE AVE, all
+Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
+
+#mouse.av 1CO
+Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
+
+#MOUSE.av 2CO
+Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
+
+#single biv level
+#male.biv?
+
+#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value
+
+```
+
+```{r log.reg_SC.length, echo=FALSE, include=FALSE}
+#REMOVE THIS COMPARES FROM FULL BIV DATA
+
+#male_mouse.av
+#Male.poly.Mouse.Table_BivData_4MM
+
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+
+#male poly shouldn't be here
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
+                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
+                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
+
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
+
+#1CO
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_1CO)
+
+#2CO
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_mouse.av_2CO)
+
+#3CO
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_3CO)
+
+# bivalent level log regression
+
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV)
+
+#1CO  (BIVDATA)
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_1CO)
+
+#2CO
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_2CO)
+
+#3CO
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV_3CO)
+```
+
+```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
+#REMOVE THIS, IT uses total bivalent data
+
+#rerun the models for datasets -- deviding by chrm class and rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
+
+Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
+Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
+
+#calq_mouse av
+mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        n_cells = length(unique(fileName) ),
+                        mean.SC.total = mean(chromosomeLength),
+
+                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+                         )
+
+mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
+mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
+
+mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
+mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
+
+#calq_mouse av
+mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
+
+mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
+mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
+```
+
+#Q2 1CO postion
+
+```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
+#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
+
+#High = MSM, PWD, SKIVE
+Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+#p-value = 1
+
+#high = PWD and MSM
+Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
+#p-value = 1
+
+```
+
+```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
+
+#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
+
+#Musc
+Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+# p-value = 0.4
+
+#Mol
+Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+# p-value = 0.07
+# exclusive rec group not sig
+```
+
+```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
+#these don't have mean.first.foci
+
+#short -- real.long.bivData -- REAL.short_mouse_table
+Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
+
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
+#p-value = 0.3
+
+Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
+# p-value = 0.1
+
+#remove SKIVE??
+```
+
+```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}
+
+#real.long.bivData
+#single biv level
+Rec.group_LONG.BivData_F1post_Ttest <- t.test(
+  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
+(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
+#p-value = 0.5
+
+Rec.group_BivData_F1post_Ttest <- t.test(
+  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
+(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
+#p-value = 0.2
+
+#
+#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
+#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
+#p-value = 0.2
+
+#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
+#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
+#p-value = 0.4
+
+```
+
+```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
+
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
+
+#none of these are M1 -- since they are male specific 
+#all of these are M4
+
+Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
+#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
+
+#with.chrm.legth  ?
+Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+
+#M4.2 (just strain)
+Q2.glm_mus.av_1COpos_M4.2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.glm_mus.av_1COpos_M4.2 <- summary(Q2.glm_mus.av_1COpos_M4.2)$coefficients
+
+```
+
+```{r Q2.1COpos.show.MOL, echo=FALSE}
+
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
+
+mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
+                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
+
+#+facet_wrap(~mouse, scales = "free")
+
+single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
+                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
+
+#+scale_color_manual(values=colors_of_strains)
+
+single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
+                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
+
+#CHRM.length is very significant-- some of the same subsp and strains are sig     
+```
+
+```{r Q2.1CO.pos_glm.chrm.Length, echo=FALSE}
+#remove this cell,
+
+Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
+
+Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
+
+#making new models
+
+Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
+Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
+
+chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+
+sum.chrm.1CO.pos.chrmLength <- summary(chrm.1CO.pos.chrmLength)$coefficients
+#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
+
+chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+
+#summary(chrm.1CO.pos.chrmLength_strain)
+
+```
+
+```{r Q2.1COpos_strain, echo=FALSE, eval=FALSE}
+#isolate all musc
+#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+##long biv
+male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean_F1 = mean(Foci1, na.rm=TRUE),
+                        mean_F1.nrm = mean(PER_Foci_1,na.rm=TRUE),
+
+                        mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE)
+
+                                               )
+#add subsp
+
+mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
+mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)
+
+mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
+mouse.av.long.biv.table$subsp <- factor(mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#add rec group
+mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", mouse.av.long.biv.table$strain), 1, 
+                            ifelse(grepl("MSM", mouse.av.long.biv.table$strain), 1,
+                             ifelse(grepl("SKIVE", mouse.av.long.biv.table$strain), 1, 0)) )
+```
+
+```{r Q2.1COpos.LOG.glm, echo=FALSE, eval=FALSE}
+
+#logistic regression, reverse variables
+
+male.bivdata.1CO$Rec.group <- ifelse(grepl("PWD", male.bivdata.1CO$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.1CO$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.1CO$strain), 1, 0)))
+
+Log.Q2.1COnrm.F1.PER <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO,  family=binomial(link="logit"))
+#summary(Log.glmQ2_M2.1COnrm)
+
+#Musc
+Log.Q2.F1_PER_Musc <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Musc",], 
+              family=binomial(link="logit"))
+#summary(Log.Q2.F1_PER_Musc)
+
+#Mol
+Log.Q2.F1_PER_Mol <- glm(Rec.group ~ PER_Foci_1,
+              data=male.bivdata.1CO[male.bivdata.1CO$subsp == "Mol",],
+              family=binomial(link="logit"))
+
+#summary(Log.glmQ2_M2.1COnrm_Mol)
+
+```
+
+#Q2 IFD
+
+```{r Q2.IFD.long.short.setup, echo=FALSE}
+#long BivData
+male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.long.biv.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.long.biv.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.long.biv.2CO_IFD$strain), 1, 
+                                           ifelse(grepl("SKIVE", male.long.biv.2CO_IFD$strain), 1, 0)))
+
+#show
+Long.biv.IFDs_PER_by.group <- ggplot(male.long.biv.2CO_IFD, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized Interfocal Distances, Long BivData\nLow Rec group =0, High Rec group =1")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,1.1))
+
+#long biv's - short IFD rates
+
+LongBiv_IFD_mus.rate <- ddply(male.long.biv.2CO_IFD %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+
+                   N.obs  = length(Obj.ID),
+
+                  mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
+                  mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                  var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                  #mean_IFD_short.removed
+                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
+                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)] ),
+
+                inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                n.close.foci = length(which(IFD1_PER <= .3)),
+
+                off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                rate.close.foci = n.close.foci / N.obs
+
+                                              )
+
+#run t.tests on long biv
+LongBiv_IFD_mus.rate <- add_strain(LongBiv_IFD_mus.rate)
+LongBiv_IFD_mus.rate <- add_category(LongBiv_IFD_mus.rate)
+LongBiv_IFD_mus.rate <- add_subsp(LongBiv_IFD_mus.rate)
+
+#add rec.group
+LongBiv_IFD_mus.rate$Rec.group <- ifelse(grepl("PWD", LongBiv_IFD_mus.rate$strain), 1, 
+                                ifelse(grepl("MSM", LongBiv_IFD_mus.rate$strain), 1,
+                                 ifelse(grepl("SKIVE", LongBiv_IFD_mus.rate$strain), 1, 0)))
+
+```
+
+```{r Q2.IFD.long.biv_t.test, echo=FALSE}
+#short removed PER
+Short.removed_BivData.LongBiv_IFD.PER_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
+                                 LongBiv_IFD_mus.rate$mean.perIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
+
+#Short.removed_BivData.LongBiv_IFD.PER_Ttest$p.value
+#0.0083
+
+#ABS short removed
+Short.removed_BivData.LongBiv_IFD.ABS_Ttest <- t.test(LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 1], 
+                                 LongBiv_IFD_mus.rate$mean.absIFD_short.rm[LongBiv_IFD_mus.rate$Rec.group == 0])
+
+#Short.removed_BivData.LongBiv_IFD.ABS_Ttest$p.value
+#0.047
+```
+
+```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
+#t.test for IFD
+#mouse.av - the mouse average IFD is from just 2CO
+t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
+
+#t.test_mouse.av.IFD_PER_DOM <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == #"male") %>% filter(subsp == "Dom") )$mean_IFD.2CO_PER,
+#   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == #"Dom"))$mean_IFD.2CO_PER   )
+
+t.test_mouse.av.IFD_PER_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_PER   )
+
+t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER   )
+
+#ABS
+
+t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS,
+   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
+```
+
+```{r Q2.IFD.glm.test, echo=FALSE}
+
+#M4.ABS
+Q2.M4glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
+sum.Q2.M1glm_IFD.raw <- summary(Q2.M4glm_IFD.raw)
+
+#M4.PER
+Q2.M4glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
+
+sum.Q2.M4glm_IFD.PER <- summary(Q2.M4glm_IFD.PER)
+#summary(Q2.M1glm_IFD.PER)
+
+#M4.2.raw
+Q2.M4.2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
+
+Q2.M4.2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+sum.Q2.M4.2glm_IFD.PER <- summary(Q2.M4.2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+#within subspecies
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M2_IFD_Mol)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M2_IFD.PER_Mol)
+
+```
+
+```{r Q2.IFD.glm.chrmLength, echo=FALSE, eval=FALSE}
+
+Curated_BivData$subsp  <- factor(Curated_BivData$subsp, ordered = FALSE)
+
+#M.Len.ABS
+Q2.BIV.LEN.glm_IFD.raw <- glm(IFD1_ABS ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.raw)
+
+#M.Len.PER
+Q2.BIV.LEN.glm_IFD.PER <- glm(IFD1_PER ~ subsp * strain * chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2")    )
+#summary(Q2.BIV.LEN.glm_IFD.PER)
+
+#M.Len.raw
+Q2.BIV.LEN.M2glm_IFD.ABS <- glm(IFD1_ABS ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.ABS)
+
+##M.Len PER
+Q2.BIV.LEN.M2glm_IFD.PER <- glm(IFD1_PER ~ strain* chromosomeLength, data= Curated_BivData %>% filter(sex == "male")%>% filter(hand.foci.count == "2"))
+#summary(Q2.BIV.LEN.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+```
+
+```{r Q2.IFD_rate.shortIFD, echo=FALSE, warning = FALSE, eval=FALSE}
+
+#for the relevant data, by strain, calcu the number of observations of 2CO above and below 30%
+
+close.IFD.rates.mus.rate <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("mouse"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                      #  mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                      mean.ABS.IFD = mean(IFD1_ABS, na.rm = TRUE),  
+                      mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm=TRUE),
+
+                  #mean_IFD_short.removed
+                  mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
+
+                mean.absIFD_short.rm = mean(  IFD1_ABS[which(IFD1_PER >= .3)], na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER >= .31)),  
+                      n.close.foci = length(which(IFD1_PER <= .3)),
+
+                      off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) ),
+
+                      rate.close.foci = n.close.foci / N.obs
+)
+
+#mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+
+close.IFD.rates.mus.rate <- add_strain(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_category(close.IFD.rates.mus.rate)
+close.IFD.rates.mus.rate <- add_subsp(close.IFD.rates.mus.rate)
+#add rec.group
+close.IFD.rates.mus.rate$Rec.group <- ifelse(grepl("PWD", close.IFD.rates.mus.rate$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates.mus.rate$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates.mus.rate$strain), 1, 0)))
+
+close.IFD.rates_strain <- ddply(Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2), 
+                         c("strain"), summarise,
+                        N.obs  = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength, na.rm = TRUE), 3 ), nsmall=3) ),
+                        mean.per.IFD = mean(IFD1_PER, na.rm=TRUE),
+                        var.IFD = var(IFD1_PER, na.rm = TRUE),
+
+                        #mean_IFD_short.removed
+                      mean.perIFD_short.rm = mean(  IFD1_PER[which(IFD1_PER >= .3)], na.rm=TRUE),
+
+                      inverse.close.foci = length(which(IFD1_PER > .3)),  
+                      n.close.foci = length(which(IFD1_PER < .3)),
+                      rate.close.foci = n.close.foci / N.obs,
+
+                       off.by = (N.obs - sum(c(inverse.close.foci,n.close.foci)) )
+
+)
+
+close.IFD.rates_strain <- add_subsp(close.IFD.rates_strain)
+
+show.IFDrm <- ggplot(close.IFD.rates_strain, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences")+ geom_abline(slope = 1, intercept = 0)
+
+#close.IFD.rates.mus.rate
+show.IFDrm_mus <- ggplot(close.IFD.rates.mus.rate, aes(y=mean.per.IFD, x=mean.perIFD_short.rm, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)+ylim(c(.4,.7))+xlim(c(.4,.7))+ggtitle("2CO interference differences") + geom_abline(slope = 1, intercept = 0)
+
+show.close.IFD.rate <- ggplot(close.IFD.rates_strain, aes(y=rate.close.foci, x=strain, color=strain))+geom_jitter()+scale_color_manual(values=colors_of_strains)
+
+#add rec.group
+close.IFD.rates_strain$Rec.group <- ifelse(grepl("PWD", close.IFD.rates_strain$strain), 1, 
+                                ifelse(grepl("MSM", close.IFD.rates_strain$strain), 1,
+                                 ifelse(grepl("SKIVE", close.IFD.rates_strain$strain), 1, 0)))
+
+#POOL all short and 
+#chi square test
+high.rec.props <- prop.test( ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$n.close.foci,
+                              ( close.IFD.rates_strain %>% filter(Rec.group == 1) )$N.obs)     
+#p-value = 0.03, slight dif tween 3 high rec strains
+```
+
+```{r Q2.IFD.PER.logreg, echo=FALSE, eval=FALSE}
+
+#Log regression all show significant prediction
+#IFD.PER
+log.reg.Q2.IFD.PER <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER)
+
+#MUSC
+log.reg.Q2.IFD.PER.Musc <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Musc)
+
+#Mol
+log.reg.Q2.IFD.PER.Mol <- glm(Rec.group ~ mean_IFD.2CO_PER,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.PER.Mol)
+
+#SINGLE BIVALENT LEVEL
+Log.Q2.IFD.per_bivdata <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.per_bivdata)
+
+#Musc
+Log.Q2.IFD.per_bivdata.Musc <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.per_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.per_bivdata.Mol <- glm(Rec.group ~ IFD1_PER,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+
+```
+
+```{r Q2IFD_logreg, echo=FALSE, eval=FALSE}
+
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                    ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
+                                           ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)))
+
+#IFD.raw
+log.reg.Q2.IFD.raw <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw)
+
+#MUSC
+log.reg.Q2.IFD.raw.Musc <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Musc)
+
+#Mol
+log.reg.Q2.IFD.raw.Mol <- glm(Rec.group ~ mean_IFD.2CO_ABS,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.reg.Q2.IFD.raw.Mol)
+
+#SINGLE BIV LEVELs
+male.bivdata.2CO_IFD <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+
+male.bivdata.2CO_IFD$Rec.group <- ifelse(grepl("PWD", male.bivdata.2CO_IFD$strain), 1, 
+                                    ifelse(grepl("MSM", male.bivdata.2CO_IFD$strain), 1, 
+                                            ifelse(grepl("SKIVE", male.bivdata.2CO_IFD$strain), 1, 0)))
+
+Log.Q2.IFD.raw_bivdata <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD,  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata)
+
+#Musc
+Log.Q2.IFD.raw_bivdata.Musc <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata.Musc)
+
+#Mol
+Log.Q2.IFD.raw_bivdata.Mol <- glm(Rec.group ~ IFD1_ABS,
+              data=male.bivdata.2CO_IFD[male.bivdata.2CO_IFD$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(Log.Q2.IFD.raw.PER_bivdata.Mol)
+```
+
+```{r Q2.IFD.longbiv_glm, echo=FALSE, eval=FALSE}
+#Full dataset
+male.long.biv.2CO_IFD$strain <- factor(male.long.biv.2CO_IFD$strain, ordered = FALSE)
+
+#single bivalent
+Q2.glm_long.biv_M2.IFD.PER <- glm(IFD1_PER ~ strain, data= male.long.biv.2CO_IFD)
+#summary(Q2.glm_long.biv_M2.IFD.PER)
+#The PER.IFDs don't have significant strain effects
+```
+
+```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
+
+male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+#t.tests for the Rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#THIS SHOULD ALREADY BE ADDED
+
+#single biv level
+Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
+                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
+
+#Rec.group.BivData_IFD.PER_Ttest$p.value
+#p-value < 2.2e-16
+
+#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
+Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
+                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
+
+#Rec.group_IFD.PER_Ttest$p.value 
+#p-value = 6.872e-07
+```
+
+```{r Q2.IFD_short.removed_t.test, echo=FALSE, include=FALSE, eval=FALSE}
+Short.removed_BivData_IFD.PER_Ttest <- t.test(Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 1], 
+  Curated_BivData_short.IFD.removed$IFD1_PER[Curated_BivData_short.IFD.removed$Rec.group == 0])
+
+Short.removed_BivData_IFD.PER_Ttest$p.value
+
+#
+Short.removed_BivData_IFD.PER_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
+                       close.IFD.rates.mus.rate$mean.perIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
+Short.removed_BivData_IFD.PER_Ttest_mus.av$p.value
+
+Short.removed_BivData_IFD.ABS_Ttest_mus.av <- t.test(close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 1], 
+                                 close.IFD.rates.mus.rate$mean.absIFD_short.rm[close.IFD.rates.mus.rate$Rec.group == 0])
+```
+
 # SAVE the data

### 2020-04-22 doc/Results_v2.rmd
fixing bugs and playing with triangle plots

@@
 
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
-#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
-#2.2.20
-#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
+#load(file = here("data", "MLH1_data_setup_4.21.20.RData") )
+
+load(file = "~./MLH1repo/data/Results_setup_4.21.20.RData")
 
@@
 #load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
-
-#this MLH1 only goes up to batch16!
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
 
@@
 
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
 #try customize the plot size
@@
 #just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus") 
-
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
+#just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+#just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
+#just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
+#just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
+#just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
 
@@
 
-Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+#Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
 
@@
 
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-male.plot
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+#male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
+#  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+#  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
+#male.plot
+
+#musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
 
@@
 
+```{r Q1.IFD.triangle, echo=FALSE}
+
+#make triangle for the IFD measures (try doing female top male bottom)
+#Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+
+#remove grey background
+#add diagonal
+#
+
+IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
+  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.4)+
+  ggtitle("Normalized IFD\nDom")+
+  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
+  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)
+
+IFD.Male.DOM.tri.plot
+
+basey.base <- plot( y = (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") )$PER_Foci_2, x= (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") )$PER_Foci_1 )
+
+```
+
+```{r Q1.show.IFD_female.male.comparison, echo=FALSE}
+
+#compare female IFD patterns which have telomeric bias (more similar to males)
+
+```
+
 ```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE}
@@
 
-```{r Q2.totSC.glm, echo=FALSE}
-
-Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
-
-#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
-#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
-#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)
-
-#M4.2? (removed subspecies?)
-
-Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
-#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
-
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
-#within subspecies values
-
-#Dom
-Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
-sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
-#there are significant strain effects in DOM,
-
-#Musc
-Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
-sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
-#KAZ -- is the only significant effect
-
-#Mol
-Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
-sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)
-
-```
-
 ```{r Q2.glm.SC.long, echo=FALSE}
@@
 
-table( droplevels(fomo.male_mouse.av$strain) )
+#table( droplevels(fomo.male_mouse.av$strain) )
 #there are cast strains too
 
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-musc.male.plot
+#Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
+
+#musc.male.plot <- ggplot( ( male_mouse.av %>% filter(species == "M.musculus") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+#musc.male.plot
 
@@
 
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`;
+<no mean.MLH1>
+
+p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
 Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
@@
 **the values between versions are slightly different**
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`).
+In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
 In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
@@
 
+```{r Q2.triangle.IFD, echo=FALSE}
+
+```
+
 ```{r Q2.show.IFD, echo=FALSE, warning = FALSE}

### 2020-04-22 src/Result_Setup.Rmd
fixing bugs and playing with triangle plots

@@
 
-```{r MLH1.HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
+```{r MLH1.HetC.rates, echo=FALSE, warning=FALSE}
 #calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
@@
 ```{r tot.skel.table, echo=FALSE}
+
+MLH1.merge.Skel$nMLH1.foci <- as.numeric(MLH1.merge.Skel$nMLH1.foci)
+
 totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
@@
                  nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
+                 var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
 
@@
                  nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+
+                av.MLH1 = mean( mean.MLH1, na.rm = TRUE),
                 # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
@@
 Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients
+sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M4)$coefficients
 
@@
 Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
+sum.Q2.glm_mus.av_1COpos_M4  <- summary(Q2.glm_mus.av_1COpos_M4)$coefficients
 #summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
@@
 #with.chrm.legth  ?
-Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+#Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
 
@@
 
-```
-
-```{r Q2.1COpos.show.MOL, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
-
-mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
-                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-
-#+facet_wrap(~mouse, scales = "free")
-
-single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
-                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
-
-#+scale_color_manual(values=colors_of_strains)
-
-single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
-                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-
-#CHRM.length is very significant-- some of the same subsp and strains are sig     
 ```
@@
 Q2.M4glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M4glm_IFD.raw)
+sum.Q2.M4glm_IFD.raw <- summary(Q2.M4glm_IFD.raw)
 
@@
 Q2.M4.2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
+Sum.Q2.M4.2glm_IFD.raw <- summary(Q2.M4.2glm_IFD.raw)
 
@@
 #Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD_Dom)
+Q2.glm.M4.2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M4.2_IFD_Dom)
 
 #Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD_Musc)
+Q2.glm.M4.2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M4.2_IFD_Musc)
 
 #Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD_Mol)
+Q2.glm.M4.2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M4.2_IFD_Mol)
 
 #Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD.PER_Dom)
+Q2.glm.M4.2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q2.glm.M4.2_IFD.PER_Dom)
 
 #Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD.PER_Musc)
+Q2.glm.M4.2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q2.glm.M4.2_IFD.PER_Musc)
 
 #Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
+Q2.glm.M4.2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q2.glm.M4.2_IFD.PER_Mol)
 ```
@@
 save.image("~./MLH1repo/data/Results_setup_4.21.20.RData")
-#save.image("data/BivData_4_PCA.RData")
-```
+
+```

### 2020-04-23 doc/Results_v2.rmd
working on the triangle plots

@@
 
+```{r Q1.ggplot_IFD.triangle, echo=FALSE}
+#IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
+#  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
+#  geom_jitter(alpha=.4)+
+#  ggtitle("Normalized IFD\nDom")+
+#  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
+#  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)
+
+```
+
+```{r with.cow, echo=FALSE, warning=FALSE}
+library(cowplot)
+
+top.triangle <- function() {
+  plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
+      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
+      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
+  axis(side = 2, las = 1, pos=0)
+  axis(side = 3, las = 1, pos=1)
+  abline(coef = c(0,1))
+
+}
+
+bottom.triangle <- function() {
+  plot( PER_Foci_1 ~ PER_Foci_2, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
+      axes = FALSE, xlab = NA, ylab = NA, xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
+  axis(side = 1, las = 1, pos=0)
+  axis(4, las = 1, pos=1) #flip label to right side
+  abline(coef = c(0,1))
+}
+
+#plot_grid(top.triangle, bottom.triangle, ncol=1)
+
+plot_grid(top.triangle, bottom.triangle, align = "v", rel_widths = c(1.3, 1.3))
+
+```
+
 ```{r Q1.IFD.triangle, echo=FALSE}
@@
 #remove grey background
-#add diagonal
-#
-
-IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
-  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.4)+
-  ggtitle("Normalized IFD\nDom")+
-  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
-  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)
-
-IFD.Male.DOM.tri.plot
-
-basey.base <- plot( y = (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") )$PER_Foci_2, x= (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") )$PER_Foci_1 )
+#add diagona
+
+#trying with base R
+
+plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
+      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
+      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
+
+axis(side = 2, las = 1, pos=0)
+axis(side = 3, las = 1, pos=1)
+abline(coef = c(0,1))
+
+#bottom triangle, Female
+plot( PER_Foci_1 ~ PER_Foci_2, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
+      axes = FALSE, xlab = NA, ylab = NA, xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
+axis(side = 1, las = 1, pos=0)
+axis(4, las = 1, pos=1) #flip label to right side
+abline(coef = c(0,1))
+
+#use cowplot or another way to paste these together
 

### 2020-04-23 doc/Results_v2.rmd
making table, changing L and Z  to early and late

@@
 library(knitr)
+library(kableExtra)
 library(ggplot2)
@@
 
+- update figure
+
+- make table, save in figure folder, print head here
+
+- remove cody blocks
+
 ```{r DMC1.setup, echo=FALSE}
@@
 )
-DMC1.mean.obs.L$stage <- "L"
+DMC1.mean.obs.L$stage <- "early Z"
 
@@
 )
-DMC1.mean.obs.Z$stage <- "Z"
+DMC1.mean.obs.Z$stage <- "late Z"
 
 DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
-#            row.names = FALSE)
 
@@
 ```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
-#think of better color scheme
-#make the jitter col thinner or make the stacked dots.
-#remember that box-plots add outlier dots.
-#change y labels
+DMC1.DF$stage
 
@@
 
-```{r plot.t.test, echo=FALSE, include=FALSE}
-
-  #this will add neat bars, but still need to figure out what groups ect it's actually using
-#THE stat_compare function adds a arm thing and pvalue for 
-#  stat_compare_means(comparisons = my_comparisons)
-
-#status - legend, expression y-axis, facet, facet_wrap SYMBOL
-#legend - stage. y=total count, FACet strain
-
-#add column to distinguish high vs low rec strains
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
-DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)
-
-# annotation table with adjusted pvals and y-position of the labels
-
-anno_df = compare_means(total.count ~ rec_group, group.by = "stage", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_High.to.Low <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=rec_group, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=stage), position=position_jitterdodge()) + 
-  facet_wrap(~stage, nrow=1) + 
-  ggsignif::geom_signif(
-    data=anno_df, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#example for ggplot+ test from here
-#https://github.com/kassambara/ggpubr/issues/65
-
-#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/
-
-```
-
 ```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
@@
 
-```{r DMC1_anova, echo=FALSE, warning=FALSE}
-#1. ANOVA, is there a difference in DMC1 counts across strains?
-
-#unorder strain
-DMC1.DF$strain <- factor(DMC1.DF$strain, levels = c("G","WSB","KAZ","PWD","MSM"), ordered = FALSE )
-
-mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
-mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])
-
-l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
-z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
-
-#anova's aren't appropriate for 2 samples... re-do the anova
-#so that it's compareing more than just 2 categories -- or delete
-
-```
-
 ```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+##RATIOS
+DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
+DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
+```{r DMC1.table, echo=FALSE}
+
 #make category averages from AP_mouse_table.HQ
@@
 #attach this mini DF to the DMC1 data
-
 DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
@@
 
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
 
 #DMC1.mean.table
-DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
-
-colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
-                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
-
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-#MAIN TABLE, headers need to be fixed, remove var.MLH1
-#try adding double decker colume names / lines for early Z and late Z
-#kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
+DMC1table.main99 <- subset(DMC1table.main, select= -c(stage.x, var.MLH1, stage.y, mean.count.x, mean.count.y))
+
+DMC1table.main99 <-  DMC1table.main[, c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
+
+colnames(DMC1table.main99) <- c("category", "mean MLH1",    "ncells",     "mean DMC1", "MLH1:DMC1 ratio", "ncells",     "mean DMC1", "MLH1:DMC1 ratio")
+
+kable(DMC1table.main99, row.names = F) %>%
+  kable_styling() %>%
+  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
+  scroll_box(width = "800px", height = "200px")
 ```
@@
 
-```{r with.cow, echo=FALSE, warning=FALSE}
+```{r with.cow, echo=FALSE, warning=FALSE, message=FALSE}
 library(cowplot)
@@
 
-plot_grid(top.triangle, bottom.triangle, align = "v", rel_widths = c(1.3, 1.3))
-
-```
+#plot_grid(top.triangle, bottom.triangle, align = "v", rel_widths = c(1.3, 1.3))
+
+```
+
+```{r another.triangle, echo=FALSE, fig.width = 5, fig.height=5}
+
+Curated_BivData <- Curated_BivData %>% filter(PER_Foci_2 < 1.01)
+
+with( (Curated_BivData %>% filter(strain == "G") ), { 
+  plot(  c( (Curated_BivData %>% filter(sex == "male") %>% filter(strain == "G") %>% filter(hand.foci.count == 2))$PER_Foci_2,
+              (Curated_BivData %>% filter(strain == "G") %>% filter(sex == "female") %>% filter(hand.foci.count == 2))$PER_Foci_1),
+
+                     c(   (Curated_BivData %>% filter(sex == "male") %>% filter(strain == "G") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
+
+                       (Curated_BivData %>% filter(strain == "G") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2), 
+          xlab = "Position1", ylab = "Position2", asp = 1, main="G IFD distributions"); 
+  axis(4, las = 1, pos=1); #right side
+  axis(side = 3, las = 1, pos=1); #
+  segments(0, 0, 1, 1); abline(v = 0:1, h = 0:1) })
+
+```
+
+reverse the axis
 

### 2020-04-23 src/Result_Setup.Rmd
renaming stages

@@
 #make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
                  .(category),
@@
 )
-DMC1.mean.obs.L$stage <- "L"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+DMC1.mean.obs.L$stage <- "Early"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
                  .(category),
@@
 )
-DMC1.mean.obs.Z$stage <- "Z"
+DMC1.mean.obs.Z$stage <- "Late"
 
@@
 #high vs low group (L)
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+```{r DMC1.rand.calq, echo=FALSE}
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+```
+
+```{r DMC1.ratios, echo=FALSE}
+
+##RATIOS
+DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
+DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+
+#DMC1.mean.table
+DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
+
+colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
+                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
+
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
+
+kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
+-----
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize,
+
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
+
+#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
+#            row.names = FALSE)
+
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+```
+
+```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+#change order
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
+                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.boxplot
+```
+
+```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
+
+anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#
+
+DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
+
+  #facet_wrap(~stage, nrow=1) +
+
+  ggsignif::geom_signif(
+    data=anno_df_stage, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
+
+ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
+               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
+
+DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
+
+ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
+               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
+
+DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
+
+ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
+               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
 ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
@@
 #Kaz Z are higher than Dom
-```
-
-```{r DMC1.rand.calq, echo=FALSE}
-
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-```
-
-```{r DMC1.ratios, echo=FALSE}
-
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-
-#DMC1.mean.table
-DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
-
-colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
-                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
-
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
 ```

### 2020-04-23 doc/Results_v2.rmd
wrote the big loop for triangle plots

@@
 
-- update figure
-
-- make table, save in figure folder, print head here
-
-- remove cody blocks
-
 ```{r DMC1.setup, echo=FALSE}
@@
 
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
                  .(category),
@@
 ```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
-DMC1.DF$stage
 
@@
 DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice\nZygotene spermatogcytes")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black"))
@@
 anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
 
@@
 
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
+ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
+               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )
 
@@
 
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
+ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
+               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )
 
@@
 
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
+ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
+               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
 ```
@@
 #high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
 # p-value =  0.002145
@@
 #high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
 # p-value = 0.6628
@@
 #dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
 # p-value = 0.003016
@@
 #dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
 #p-value = 0.09588
@@
 #KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
 
@@
 
-```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}
+```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}
 
@@
 
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
+```{r DMC1.table, echo=FALSE, message=FALSE}
 ##RATIOS
@@
 DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
-```{r DMC1.table, echo=FALSE}
 
@@
                     category == "PWD male")
+
 #attach this mini DF to the DMC1 data
@@
 
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+DMC1table.main88 <- merge(DMC1.target.strains, DMC1table, by="category")
 
@@
 
-DMC1table.main99 <-  DMC1table.main[, c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
+DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
 
@@
 
-kable(DMC1table.main99, row.names = F) %>%
+kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
   kable_styling() %>%
@@
   scroll_box(width = "800px", height = "200px")
+
 ```
@@
 
-```{r with.cow, echo=FALSE, warning=FALSE, message=FALSE}
+```{r with.cow, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
 library(cowplot)
@@
   abline(coef = c(0,1))
-
 }
@@
 
-#plot_grid(top.triangle, bottom.triangle, ncol=1)
-
-#plot_grid(top.triangle, bottom.triangle, align = "v", rel_widths = c(1.3, 1.3))
-
-```
-
-```{r another.triangle, echo=FALSE, fig.width = 5, fig.height=5}
+```
+
+```{r another.triangle, echo=FALSE, fig.width = 5, fig.height=5, eval=FALSE}
 
@@
 
-```{r Q1.IFD.triangle, echo=FALSE}
-
-#make triangle for the IFD measures (try doing female top male bottom)
-#Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-
-#remove grey background
-#add diagona
-
-#trying with base R
-
-plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
-      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
-      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
-
-axis(side = 2, las = 1, pos=0)
-axis(side = 3, las = 1, pos=1)
-abline(coef = c(0,1))
-
-#bottom triangle, Female
-plot( PER_Foci_1 ~ PER_Foci_2, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
-      axes = FALSE, xlab = NA, ylab = NA, xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
-axis(side = 1, las = 1, pos=0)
-axis(4, las = 1, pos=1) #flip label to right side
-abline(coef = c(0,1))
-
-#use cowplot or another way to paste these together
-
-```
-
-```{r Q1.show.IFD_female.male.comparison, echo=FALSE}
-
-#compare female IFD patterns which have telomeric bias (more similar to males)
-
+```{r updated.triangl.plots, echo=FALSE, message=FALSE, warning=FALSE}
+my.data <- data.frame(my.x=c(.2, .4, .6, .1), my.y=c(.3, .5, .7, .9))
+x.dist <- .5
+my.data.2 <- transform(my.data, my.y=my.y + x.dist)
+
+#x_dist is a value for spacing the trianlges, remeber that the data values need
+#to be adjusted as well
+#the rest of the positions are slightly off
+
+#make a loop to enter/push the merge triangles into a empty list
+
+#Top is female
+top.triangle2 <- function() {
+  #top-y position is male Pos2
+  plot(  (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
+    #top-x position is female Pos1     
+    (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
+
+       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2) ), 
+       col =  "coral4",
+       axes=FALSE, ylab='', xlab="", 
+       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
+  mtext("KAZ IFD Distributions", 3, 5, font=2, cex=1.3, adj=.5)
+
+  mtext("Female Position.2", 2, .75)
+  mtext("Female Position.1", 3, 2)
+  axis(side=2, las=1, pos=0)
+  axis(side=3, las=1, pos=1)
+  lines(0:1, 0:1)
+}
+
+#bottom is male
+bottom.triangle2 <- function() {
+    #bottom-y position is male Pos1
+  points(  (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
+    #bottom-x position is male Pos2     
+    (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,
+
+         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2) ),
+
+    col =  "coral4",
+         xpd=TRUE)
+  mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
+  mtext("Male Position.1", 4, 3)#, padj=par()$usr[1]
+  x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
+  axis(side=1, las=1, pos=0, at=x.at, 
+       labels=F, xpd=TRUE)
+  mtext(seq(0, 1, .2), 1, 0, at=x.at)
+  axis(4, las=1, pos=1) # + x.dist
+  lines(0:1, 0:1, xpd=TRUE)# + x.dist
+
+  }
+
+png("~./MLH1repo/doc/figureFiles/KAZ.triangle.png", width=650, height=500)
+op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
+top.triangle2()
+bottom.triangle2()
+par(op)
+dev.off()
+
+```
+
+```{r auto.loop.ifd, echo=FALSE}
+
+#don't think I need this loop
+for(i in 1:length(auto_curate_DF$strain )){
+  auto_curate_DF$quant_status[i] <- ifelse( is.element(auto_curate_DF$strain[i], exclude_mice_list), "exclude", NA)
+  }
+
+#big loop for automating the IFD triangles
+only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
+auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
+
+df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
+df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
+#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
+
+for(i in 1:(length(df_strain_list))){
+  current.strain <- as.character(df_strain_list[[i]]$strain[1])
+  #declare file
+  png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)
+
+  #Top is female
+top.triangle2 <- function() {
+  #top-y position is male Pos2
+  plot(  (Curated_BivData %>% filter(strain == current.strain  ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
+    #top-x position is female Pos1     
+    (Curated_BivData %>% filter(strain == current.strain   ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
+
+       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) ), 
+       col =  colors_of_strains[current.strain],
+       axes=FALSE, ylab='', xlab="", 
+       main= "IFD Distribution",  xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
+
+  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.5)
+
+  mtext("Female Position.2", 2, .75)
+  mtext("Female Position.1", 3, 2)
+  axis(side=2, las=1, pos=0)
+  axis(side=3, las=1, pos=1)
+  lines(0:1, 0:1)
+  }
+
+bottom.triangle2 <- function() {
+    #bottom-y position is male Pos1
+  points(  (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
+    #bottom-x position is male Pos2     
+    (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,
+
+         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) ),
+
+    col =  colors_of_strains[current.strain], #this might not be working right
+         xpd=TRUE)
+  mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
+  mtext("Male Position.1", 4, 3)#, padj=par()$usr[1]
+  x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
+  axis(side=1, las=1, pos=0, at=x.at, 
+       labels=F, xpd=TRUE)
+  mtext(seq(0, 1, .2), 1, 0, at=x.at)
+  axis(4, las=1, pos=1) # + x.dist
+  lines(0:1, 0:1, xpd=TRUE)# + x.dist
+
+  }
+
+  op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
+  top.triangle2()
+  bottom.triangle2() #not working
+  par(op)
+  dev.off()
+  }
+
+```
+
+```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE}
+
+library(png)
+
+G.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/G.triangle.png")
+WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/WSB.triangle.png")
+LEW.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/LEW.triangle.png")
+
+#make empty DF for ggplot object
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+#make an empty ggplot thing, then background ontop
+WSB.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(WSB.triangle.png) + theme(axis.title.x=element_blank(),
+ axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+G.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(G.triangle.png) + theme(axis.title.x=element_blank(),
+ axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+LEW.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(LEW.triangle.png) + theme(axis.title.x=element_blank(),
+ axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+library("cowplot")
+fig1  <- ggdraw() +
+  draw_plot(WSB.triangle.ggplot, x=0, y=0, width=.45, height=.45) +
+  draw_plot(G.triangle.ggplot,   x=0, y=.4, width=.45, height=.45) +
+  draw_plot(LEW.triangle.ggplot, x=0, y=.6, width=.45, height=.45)
+
+fig1
+
+#ggsave("~./WildMiceMadison/results/Figure1.png")
 ```
@@
 
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "WSB"),
+P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
     aes(y = Foci2, x= Foci1, color=strain))+

### 2020-04-24 doc/Results_v2.rmd
adjusting values, cleaning up rDatafile

@@
 library(reshape2)
-#library(patchwork)
+library(patchwork)
 
@@
 
-load(file = "~./MLH1repo/data/Results_setup_4.21.20.RData")
+load(file = "~./MLH1repo/data/Results_setup_4.24.20.RData")
 
@@
 
-#table( droplevels(female_mouse.avs_for.analysis$strain) )
-#there are cast strains too
-
 Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
@@
   theme(legend.position="none",
+      axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-#female.specific.plot
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))
+
+#+theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 #there are cast strains too
-
 #Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
 #fomo.male_mouse.av
 
-#male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
-#  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-#  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-#male.plot
-
-#musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
+male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+  theme(legend.position="none", axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), 
+  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))
+
+#+theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 #Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
-#Fig1_v1_strain / (female.specific.plot + male.plot )
-
-#(p1 | p2 | p3)/p4
-```
-
-```{r main.fig1s, echo=FALSE, include=FALSE}
+Fig1_v1_strain / (female.specific.plot + male.plot )
+
+```
+
+```{r main.fig1s, echo=FALSE}
 strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
@@
 We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r MLH1.MMlmer_results$rand.pval`   **someSmallNumber**), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
+(random effect p < `r MLH1.MMlmer_results$rand.pval` -- `r 1/1000000` someSmallNumber), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
 
@@
 rename the code for these models**).
-In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect =
-`r MLH1.male.specific.glm.M2$coefficients[5]` -- ~ 5 foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = `r MLH1.male.specific.glm.M2$coefficients[8]` -- ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect  `r MLH1.male.specific.glm.M2$coefficients[6]` ~ 2 foci).
+
+In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[5,][8]`; effect = 
+`r sum_MLH1.male.glm.M4$coefficients[8]` -- ~ 5 foci), and 
+musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.glm.M4$coefficients[5,][10]` -- effect = `r sum_MLH1.male.glm.M4$coefficients[10]` -- ~2 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[5,][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]` ~ 7 foci).
+
 These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
@@
 
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`).
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p< someSmallNumber -- `r 1/1000000` -- `r ttest.HighLow.L.pval`).
 In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = 0.66 --- `r ttest.HighLow.Z.pval`).
@@
 
-```{r with.cow, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-library(cowplot)
-
-top.triangle <- function() {
-  plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
-      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
-      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
-  axis(side = 2, las = 1, pos=0)
-  axis(side = 3, las = 1, pos=1)
-  abline(coef = c(0,1))
-}
-
-bottom.triangle <- function() {
-  plot( PER_Foci_1 ~ PER_Foci_2, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
-      axes = FALSE, xlab = NA, ylab = NA, xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
-  axis(side = 1, las = 1, pos=0)
-  axis(4, las = 1, pos=1) #flip label to right side
-  abline(coef = c(0,1))
-}
-
-```
-
-```{r another.triangle, echo=FALSE, fig.width = 5, fig.height=5, eval=FALSE}
-
-Curated_BivData <- Curated_BivData %>% filter(PER_Foci_2 < 1.01)
-
-with( (Curated_BivData %>% filter(strain == "G") ), { 
-  plot(  c( (Curated_BivData %>% filter(sex == "male") %>% filter(strain == "G") %>% filter(hand.foci.count == 2))$PER_Foci_2,
-              (Curated_BivData %>% filter(strain == "G") %>% filter(sex == "female") %>% filter(hand.foci.count == 2))$PER_Foci_1),
-
-                     c(   (Curated_BivData %>% filter(sex == "male") %>% filter(strain == "G") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
-
-                       (Curated_BivData %>% filter(strain == "G") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2), 
-          xlab = "Position1", ylab = "Position2", asp = 1, main="G IFD distributions"); 
-  axis(4, las = 1, pos=1); #right side
-  axis(side = 3, las = 1, pos=1); #
-  segments(0, 0, 1, 1); abline(v = 0:1, h = 0:1) })
-
-```
-
-reverse the axis
-
-```{r updated.triangl.plots, echo=FALSE, message=FALSE, warning=FALSE}
-my.data <- data.frame(my.x=c(.2, .4, .6, .1), my.y=c(.3, .5, .7, .9))
-x.dist <- .5
-my.data.2 <- transform(my.data, my.y=my.y + x.dist)
-
-#x_dist is a value for spacing the trianlges, remeber that the data values need
-#to be adjusted as well
-#the rest of the positions are slightly off
-
-#make a loop to enter/push the merge triangles into a empty list
+```{r triangle.IFD.auto.loop.ifd, echo=FALSE, eval=FALSE}
 
@@
   #top-y position is male Pos2
-  plot(  (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
+  plot(  (Curated_BivData %>% filter(strain == "WSB"  ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
     #top-x position is female Pos1     
-    (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
-
-       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2) ), 
-       col =  "coral4",
+    (Curated_BivData %>% filter(strain == "WSB"   ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
+
+       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2) ), 
+       col =  colors_of_strains["WSB"],
        axes=FALSE, ylab='', xlab="", 
-       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
-  mtext("KAZ IFD Distributions", 3, 5, font=2, cex=1.3, adj=.5)
-
+       main= "IFD Distribution",  xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
+
+  mtext("WSB", 3, 5, font=2, cex=1.3, adj=.5)
   mtext("Female Position.2", 2, .75)
@@
 
-#bottom is male
 bottom.triangle2 <- function() {
     #bottom-y position is male Pos1
-  points(  (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
+  points(  (Curated_BivData %>% filter(strain == "WSB") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
     #bottom-x position is male Pos2     
-    (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,
-
-         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2) ),
-
-    col =  "coral4",
-         xpd=TRUE)
+    (Curated_BivData %>% filter(strain == "WSB") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,
+
+         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2) ),
+
+    col =  colors_of_strains["WSB"], xpd=TRUE)
   mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
@@
   x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
+  axis(side=1, las=1, pos=0, at=x.at, labels=F, xpd=TRUE)
   mtext(seq(0, 1, .2), 1, 0, at=x.at)
@@
   lines(0:1, 0:1, xpd=TRUE)# + x.dist
-
   }
 
-png("~./MLH1repo/doc/figureFiles/KAZ.triangle.png", width=650, height=500)
+#par(mar=c(1,1,1,1))
 op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
-top.triangle2()
-bottom.triangle2()
+plot(top.triangle2(), bottom.triangle2() ) #not working
 par(op)
@@
 
-```{r auto.loop.ifd, echo=FALSE}
-
-#don't think I need this loop
-for(i in 1:length(auto_curate_DF$strain )){
-  auto_curate_DF$quant_status[i] <- ifelse( is.element(auto_curate_DF$strain[i], exclude_mice_list), "exclude", NA)
-  }
+```{r double.tri.rewrite, echo=FALSE, eval=FALSE}
+#good function, saves to file, can't reproduce in rmd
 
@@
 
-for(i in 1:(length(df_strain_list))){
-  current.strain <- as.character(df_strain_list[[i]]$strain[1])
-  #declare file
-  png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)
-
-  #Top is female
-top.triangle2 <- function() {
-  #top-y position is male Pos2
-  plot(  (Curated_BivData %>% filter(strain == current.strain  ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
-    #top-x position is female Pos1     
-    (Curated_BivData %>% filter(strain == current.strain   ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
-
-       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) ), 
-       col =  colors_of_strains[current.strain],
-       axes=FALSE, ylab='', xlab="", 
-       main= "IFD Distribution",  xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
-
-  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.5)
-
-  mtext("Female Position.2", 2, .75)
-  mtext("Female Position.1", 3, 2)
+#this works
+top.triangle2 <- function(my.y, my.x, my.data) {
+  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
+       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
+       col =  colors_of_strains[current.strain])
+  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
+  mtext("Position.2.Female", 2, .75)
+  mtext("Position.1.Female", 3, 2)
   axis(side=2, las=1, pos=0)
@@
 
-bottom.triangle2 <- function() {
-    #bottom-y position is male Pos1
-  points(  (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
-    #bottom-x position is male Pos2     
-    (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,
-
-         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) ),
-
-    col =  colors_of_strains[current.strain], #this might not be working right
-         xpd=TRUE)
-  mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
-  mtext("Male Position.1", 4, 3)#, padj=par()$usr[1]
-  x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
+bottom.triangle2 <- function(my.y, my.x, my.data){
+  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
+          col =  colors_of_strains[current.strain])
+
+  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
+  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] + 10)
+  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
   axis(side=1, las=1, pos=0, at=x.at, 
@@
   mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1) # + x.dist
-  lines(0:1, 0:1, xpd=TRUE)# + x.dist
-
+  axis(4, las=1, pos=1 + x.dist)
+  lines(0:1 + x.dist, 0:1, xpd=TRUE)
 }
 
+x.dist <- .5
+
+for(i in 1:(length(df_strain_list))){
+  print(i)
+  current.strain <- as.character(df_strain_list[[i]]$strain[1])
+
+  #declare file
+  png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)
+
+  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
+
+  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
+
+  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
+
+  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
   op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
-  top.triangle2()
-  bottom.triangle2() #not working
+  top.triangle2(my.y.top, my.x.top, my.data)
+  bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
   par(op)
@@
 
-```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE}
+```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Example of Sex differences in IFD distributions"}
 
@@
 
-G.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/G.triangle.png")
-WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/WSB.triangle.png")
-LEW.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/LEW.triangle.png")
+#load all the files / arrange the plots from above 
+WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
 
@@
 
-G.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(G.triangle.png) + theme(axis.title.x=element_blank(),
- axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-LEW.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(LEW.triangle.png) + theme(axis.title.x=element_blank(),
- axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-library("cowplot")
-fig1  <- ggdraw() +
-  draw_plot(WSB.triangle.ggplot, x=0, y=0, width=.45, height=.45) +
-  draw_plot(G.triangle.ggplot,   x=0, y=.4, width=.45, height=.45) +
-  draw_plot(LEW.triangle.ggplot, x=0, y=.6, width=.45, height=.45)
-
-fig1
+#show wsb as example
+WSB.triangle.ggplot
+
+#library("cowplot")
+#fig1  <- ggdraw() +
+#  draw_plot(WSB.triangle.ggplot, x=0, y=0, width=.45, height=.45) +
+#  draw_plot(G.triangle.ggplot,   x=0, y=.4, width=.45, height=.45) +
+#  draw_plot(LEW.triangle.ggplot, x=0, y=.6, width=.45, height=.45)
 

### 2020-04-24 src/Result_Setup.Rmd
adjusting values, cleaning up rDatafile

@@
 ```{r Male.specific.glm, echo=FALSE}
-# 1 way anova, is at least 1 strain different from the rest?
-# given a lm, how sig is the strain effect
-glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-
-#summry.male.strain.effects <- summary(glm.male.strain.subsp)
+
+#these are M4
+#M4
+MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+sum_MLH1.male.glm.M4 <- summary(MLH1.male.glm.M4)
 #round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
 
+#sum_MLH1.male.specific.glm.M2$coefficients[5,][4]
+
+#PWD, SKIVE, MSM
+#8,9,10
+
 #write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
 
-#1. determine which strain effects are sig
-
-#M2
-MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
-sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)
+#M4.2
+MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
+sum_MLH1.male.glm.M4.2 <- summary(MLH1.male.glm.M4.2)
 
@@
 
-#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
 ```
@@
 ## within mouse variance models 
-lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
+lmVAR.male.strain.subsp <- glm(var ~ subsp * strain, data= just.male_mouse.av)
 summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
@@
 #all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain <- glm(var ~ strain, data= just.male_mouse.av)
 lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
@@
 
-lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain <- glm(cV ~ strain, data= just.male_mouse.av)
 lmcv.male.strain_sum <- summary(lmcv.male.strain)
@@
 
-lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male <- glm(var ~ subsp * strain, data= Q12_male_table)
 lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
@@
 
-lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12_male <- glm(var ~ strain, data= Q12_male_table)
 lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
@@
 
-set up, t.tests, correlations
-
-```{r DMC1.setup, echo=FALSE}
+set up, t.tests, correlations, says there are duplicate titles
+
+```{r DMC1.setup1, echo=FALSE}
 #load DMC1 csv
@@
 #attach this mini DF to the DMC1 data
-
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+```
+
+```{r DMC1_t.tests1, echo=FALSE, warning=FALSE}
 #2. t-tests for the DMC1 counts 
@@
 
-```{r DMC1.rand.calq, echo=FALSE}
+```{r DMC1.rand.calq1, echo=FALSE}
 
@@
 
-```{r DMC1.ratios, echo=FALSE}
+```{r DMC1.ratios1, echo=FALSE}
 
@@
 
-```{r DMC1.setup, echo=FALSE}
+```{r DMC1.setup, echo=FALSE, eval=FALSE}
 #load DMC1 csv
@@
 
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
+```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE, eval=FALSE}
 #think of better color scheme
@@
 
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
+```{r plot.stage.ttest, echo=FALSE, warning=FALSE, eval=FALSE}
 
@@
 
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE, eval=FALSE}
 #2. t-tests for the DMC1 counts 
@@
 #save Rdata
-save.image("~./MLH1repo/data/Results_setup_4.21.20.RData")
-
-```
+save.image("~./MLH1repo/data/Results_setup_4.24.20.RData")
+
+```

### 2020-04-24 doc/MatMeth.Rmd
mostly edits to mat methods

--- 2020-04-21 doc/MatMeth.Rmd
@@
 
-### Analysis
+### Statistical Analysis
 
 We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).
-
-### DMC1
 
@@
 
-### MLH1 (Analytical Framework)
-
 ```{r MLH1.analysis, echo=FALSE}
@@
 
+#
+#
+#partitioning variance
 ```
 
-**certeria**  MLH1 foci were quantified from cells with intact and compelete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score, between 1 to 5, was asigned to each cell based on staining quality and general spread of bivalents.
-<cell with small defects were included -- and coded as a score of 5 - but these cell weren't included in the final analysis.>  The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
+<Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.   -ene endings refer to the cells/units within the cell stage? (I think)>
 
-**the XY foci is not counted / analyzed in this study**
+**certeria**  MLH1 foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
+With scores of 1 and 2 having the highest quality.
+Cells and coded as a score of 5 were excluded from the final analysis.
+The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
+MLH1 foci located on the XY in spermatocytes were excluded from this study.
 
--Background for setting up/organizing models -- (heiarchal nature of multiple strains within subspecies -- ) - hierarchical nature of categories for inviual mouse averages ()
+**M1 Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
 
--(primary focus is mean MLH1 count per cell - at the mouse observation level - apply to more meiocyte traits)
+**M2 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
 
-**partioning variation in recombination rate**
+**M3 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
 
--Our goal -- design an utilize a mixed model for our analytical framework - which leverages / accounts for the hierarchical/nested nature of categories across all the individual mice (e.g.
-sex, strain subspecies).
+**M4 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
 
-<The mean, variance and coefficient of variance for MLH1 counts per cell were calculated for each mouse (in R).>
+**M4.2 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
 
-The use of wild derived inbred strains enabled us to design an utilize a mixed model for our analytical framework. We built our model to predict the strain average MLH1 count per cell.
+Straight forward comparisons of groups (strain, sex) were done through t.tests().
 
-**X package used to test the fixed effects, Z package used to test the random effects.**
-
-The main effects and their interaction of subspecies and sex were coded as fixed.
-
-strain (or genetic background) was coded as a **random** effect to reflect a random sample of natural genetic variation from each subspecies.
-
-In addition to mouse mean of MLH1 count per cell
-**to expand the analysis of meiotic features** We applied the following models to several dependant variables (mouse averages) <consider a figure/diagram>
-
- the mouse average of variance in crossover count per cell,
-
- mouse mean total SC per cell, mouse mean single SC length, mouse mean IFD^norm^ and IFD^raw^, and position of crossover for single crossover bivalents.
-
-**M1. Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2. Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3. Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
-
-**M4. Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
+We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
+Strain was coded as a random effect and tested using exactRLRT() to reflect a random sample of natural genetic variation from each subspecies.
+The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
 
 This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism).
-(in a sex specific manner).
-From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.
+We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
+The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
+The general linear models, **M4** were used to analyze sex specific data sets.
 
-We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
+Our goal was to design an analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
+sex, strain subspecies) and can be applied to multiple (meiotic traits) dependent variables.
+In addition to the MLH1 count variables for individual mice, mean MLH1 count per cell, variance and coefficient of variance for MLH1 count per cell, we analyzed SC based features: mouse average of total SC per cell, mouse average of single SC length, mouse mean IFD^norm^ and IFD^raw^ for double crossover chromosomes, and the average position of crossover for single crossover bivalents.
 
-We used x package for testing significance of fixed effects and V package for testing significance of (strain) random variables (strain).
-
-### Single bivalent charaecterization
-
-**combine any X and XY issue with reduced dataset** An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
+These traits were choosen because they are markers of phenomena in the recombinaiton pathway and provide more detailed information about the recombination landscape.
+In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
+Normalized measures were calculated by the position divided by the length of the SC for that single bivalent.
 
@@
 
-To charaecterize the SC length of individual chromosomes which is a marker of chromatin compaction <from MLH1 stained cells>, two image analysis algorithms were used to specifically (measure indiviual cells) [@wang_2019_sc] and indiviaul chromosomes [@peterson2019].
-Both apply a 'skeletonizing' transformation to the synapsed chromosomes which transforms the chromosome space into a single pixel wide 'trace' of the originial chromosom shape.
+To characterize the SC lengths across bivalents, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
+Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
 Hand measures of SC/pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
 
-The **total SC** per cell (all bivalents) was quantified across single cell images.
-Mouse means were calculated from cell-wide total SC lengths in 2,984 -- `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of 3,680 -- `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
-the mouse average total sc was applied to the model framework.
-<how else was this analyzed?!>
-
-**walk through how outliers are removed**
-
-##### DNACrossover
+The **total SC** per cell was quantified from the pachtyene cell images.
+To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
+Mouse means were calculated from cell-wide total SC lengths in `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
 
@@
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)`-- 0.72 (musculus^KAZ^ female).
+In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)` (musculus^KAZ^ female).
 
-A curation step is applied to the total single bivalent dataset to remove poor isolation/segmentation instances.
-From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` -- 10,458 bivalent objects were isolated by the image analysis software.
-After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` -- 9,829 single-bivalent observations remained.
+A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
+From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` bivalent objects were isolated by the image analysis software.
+After a human curation step , `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step ([@peterson2019])).
 
-#main traits
-
-**This dataset is used to charaecterize single bivalent traits**
-
-The number of MLH1 foci on single chromosomes was quantified from the curated bivalent dataset to better understand the total recombination landscape.
-The **proportions** of single and double crossover bivalents were tested across mouse categories by chi-sqaure tests with the prop.test function.
-
-**other traits** were also charaecterized -- SC length (a marker for chromatin compaction), foci position (raw and normalized), and interfocal distance a marker for interference, these traits were fit into the main models (as mentioned above).
-In addition to fiting the mouse means into the models discribed above, t.tests were used to test the means between sexes and groups.
-
-**IFD, Interference** In order to examine interference, Interfocal distance was measured from the single bivalent data sets.
-- This metric is closer to measures of interference from linkage maps which are also removed from physical scales in that they measure frequencies of crossovers.
-Crossover interference is stronger in male specific linkage maps compared to female maps in a variety of species (ref).
-**IFD^norm^**
-
-**SC Length model**
-
-- **Linear model M1 ** $$mouse \ average \ bivalent metric ~=~ subsp * strain + \varepsilon $$
-
-- **Linear model M2 ** $$mouse \ average \ bivalent metric ~=~ strain + \varepsilon $$
+From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
 
@@
 
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced dataset including only bivalents with SC lengths below the 2nd quartile (for SC length) in each cell.
-This dataset included the four or five shortest bivalents  and excluded the X bivalent in oocytes.
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
 A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller dataset has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-
-- The long bivalent and short bivalent data sets were compiled by calculating the 1st and 4th quartille of SC lengths for each cell with at least 16 bivalent measures.
-Bivalents from each cell were assigned into the long and short bins based on these SC Length cut offs.
-
-Using the single bivalent dataset, the within cell distribution of bivalent SC lengths from cells with at least 17 isolated chromosome measures.
-The longest and shortest 4 to 5 chromosomes within each cell were identified by calculating the fourth and second quartille.
+Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
 

### 2020-04-24 doc/Results_v2.rmd
mostly edits to mat methods

@@
 title: "MLH1 Results Outline"
-date: 'null'
+date: 
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
   html_document:
@@
     toc_float: yes
-  pdf_document: default
-  word_document:
-    toc: yes
-    toc_depth: '5'
+- \linenumbers
 bibliography: refs_Main_Report.bib
 ---
+
+**sex differences in Celegan meiosis**
+
+Kelly, William G., Sara M. Fielder, and Rieke Kempfer. "Multiple Sex-Specific Differences in the Regulation of Meiotic Progression in C. elegans." bioRxiv (2020).
 
@@
 library(patchwork)
+library(pander)
 
@@
 
-```{r main.fig1s, echo=FALSE}
-strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )
-
-Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point( size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-
-#strain.table.HQ_Fig.og
-
-strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")
-
-strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
-     "MSM", "MOLF",   
-  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
-  "WSB","G","LEW"), order=T )
-
-Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
-
-```
-
-```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
+```{r MAIN.TABLE1, echo=FALSE, eval=TRUE }
+#DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
+
+#colnames(DMC1table.main99) <- c("category", "mean MLH1",    "ncells",     "mean DMC1", "MLH1:DMC1 ratio", "ncells",     "mean DMC1", "MLH1:DMC1 ratio")
+
+#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
+#  kable_styling() %>%
+#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
+#  scroll_box(width = "800px", height = "200px")
+
+#write the full table to a supplemental table
+#show the head (or shortened)
 
@@
 
+#kable(strain.table.HQ_Fig, "latex")
+
+df.strain.table <- strain.table.HQ_Fig
+
+#kable(df.strain.table)
+
+pander(df.strain.table)
+
+#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
+#  kable_styling() %>%
+#  scroll_box(width = "800px", height = "200px")
+
 #kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-#write the table out
-
 ```
@@
 
-We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4  **male specific, M2?
-rename the code for these models**).
-
-In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[5,][8]`; effect = 
-`r sum_MLH1.male.glm.M4$coefficients[8]` -- ~ 5 foci), and 
-musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.glm.M4$coefficients[5,][10]` -- effect = `r sum_MLH1.male.glm.M4$coefficients[10]` -- ~2 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[5,][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]` ~ 7 foci).
+We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4; see methods).
+In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect =
+`r sum_MLH1.male.glm.M4$coefficients[8]` foci), and 
+musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
 
@@
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
-
-<in the above section reorder the strains by the effect sizes>
-
 Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
 
-<bret above>
-
 <!-- transition gwRR to within mouse variance -->
 
-# Within mouse **variance** in CO count per cell
-
-**re-read Lenzi et al -- heterogeneity in oogenesis**
+# Within mouse variance in CO count per cell
 
@@
 
-MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
 MLH1var_M2_sum <- summary(MLH1var_M2)
@@
 
-MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
 #summary(MLH1cv_M2)
@@
 #Q12 var
-MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-#summary(MLH1var_M3_Q12)
+MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+sum.MLH1var_M2_Q12 <- summary(MLH1var_M2_Q12)
 #anova(MLH1var_M2_Q12)  
@@
 #Q12 cv
-MLH1cv_M2_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
 
@@
 ```{r M3.MLH1.var, echo=FALSE}
+
 glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
+sum.glmMLH1Var_M3 <- summary(glmMLH1Var_M3)
 #round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-#use the values from this model below
 
@@
 To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < **smallNumber** -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` (old) `r round(MLH1var_M2_sum$coefficients[,4][4], 5)`  correct).
-**the pland full M2 model is not significant**.
-M3 `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`.
+Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < `r 1/1000000` -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear models (M2) (p = `r MLH1var_M2_sum$coefficients[,4][9]`) and M3 (p = `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`).
 
@@
 Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
-
-<Bret's comment for below, are there differences between the curated and the full dataset? he propbably means, Q1 and full data set.>
-
-The results are similar: sex is the strongest effect (p < **smallNumber** `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2 -- also M3).
+The results are similar: sex is the strongest effect (M1 p < `r 1/1000000` `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
 When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
@@
 
-<strain specific effect section was deleted for not highlighting main pattern. I agree>
+<strain specific effect section was deleted for not highlighting main pattern. I agree> **re-read Lenzi et al -- heterogeneity in oogenesis**
+<Bret's comment for below, are there differences between the curated and the full dataset? he propbably means, Q1 and full data set.>
 
@@
 
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
-
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
-#change order
-
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice\nZygotene spermatogcytes")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.boxplot
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-
-DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-  ggsignif::geom_signif(
-    data=anno_df_stage, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
 ```{r DMC1.table, echo=FALSE, message=FALSE}
@@
 
-kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-  kable_styling() %>%
-  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-  scroll_box(width = "800px", height = "200px")
+#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
+#  kable_styling() %>%
+#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
+#  scroll_box(width = "800px", height = "200px")
+
+pander(DMC1table.main99)
+
+```
+
+```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+#change order
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
+                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
+
+DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice\nZygotene spermatogcytes")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+DMC1.boxplot
+```
+
+```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
+
+anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+
+DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
+
+  #facet_wrap(~stage, nrow=1) +
+
+  ggsignif::geom_signif(
+    data=anno_df_stage, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
+
+ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
+               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )
+
+DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
+
+ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
+               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )
+
+DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
+
+ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
+               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
 
@@
 In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "L") )$ncells)` early zygotene-stage and `r sum( (DMC1.mean.table %>%  filter(stage == "Z") )$ncells)` late zygotene-stage spermatocytes from juvenile mice (12 to 18 days) from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
-
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p< someSmallNumber -- `r 1/1000000` -- `r ttest.HighLow.L.pval`).
+DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "Early") )$ncells)` early zygotene-stage and `r sum( (DMC1.mean.table %>%  filter(stage == "Late") )$ncells)` late zygotene-stage spermatocytes from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
+
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p<  -- `r 1/1000000` -- `r ttest.HighLow.L.pval`).
 In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = 0.66 --- `r ttest.HighLow.Z.pval`).
-
-[@cole2012] suggest that early Z counts for DMC1 are most relevant for predicting CO number.
-(CO homeostasis -- robustness )
 
@@
 Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-**what does the ratio tell us? which stage of DMC1 counts is most relevant for CO count**
-
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = 0.11 -- `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
+The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
+
 This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
-
-**note that this interpertation depends on which stage is most relevant to CO number**
-
-the difference in DMC1 counts across staes of zygotene may be related to the findings related to (crossover homeostasis).
-
-The number of DSBs increase in zygotene, then deminish as the effects of crossover homeostasis are implemented (Cole). (co homeostasis also indecates dsb )
-
-'Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links/ connects the CO number variation to 'pre' DSB formation -- fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
-(this study only examined one stage of early prophase for DMC1 counts)
-
-'does the number of DSBs per se regulate CO number?'
-
-bairer-- DSB number is not a primary driver of variation in CO number -- these authors suggest that chromatin organizaiton (loop-axis) structure is the primary determinant of CO variation.
-
-the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 c3h, 7.3 B6 )
+[@cole2012] suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
+This result, combined with the big difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
+
+<The number of DSBs increase in zygotene, then deminish as the effects of crossover homeostasis are implemented (Cole).
+(co homeostasis also indecates dsb ) Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links connects the CO number variation to 'pre' DSB formation fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
+(this study only examined one stage of early prophase for DMC1 counts)>
+
+<bairer-- DSB number is not a primary driver of variation in CO number -- these authors suggest that chromatin organizaiton (loop-axis) structure is the primary determinant of CO variation.>
+
+<the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 C3H, 7.3 B6>
 
 <Bret's Q, ANOVA is not appropriate for 2 samples unless it's specilaized somehow>
-
-<This result, combined with the big difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
-**It depends on whether leptonene or zygotene DSBs are viewed as more relevant for crossover numbers**.>
 
@@
 
-```{r Q2.triangle.IFD, echo=FALSE}
-
-```
-
 ```{r Q2.show.IFD, echo=FALSE, warning = FALSE}

### 2020-04-24 doc/Results_v2.tex
mostly edits to mat methods

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={MLH1 Results Outline},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{MLH1 Results Outline}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{null}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\maketitle

\includegraphics{Results_v2_files/figure-latex/three.plots-1.pdf}

\begin{tabular}{l|l|l|r|r|r|r|r|r|r}
\hline
subsp & strain & sex & Nmice & Ncells & mean\_co & cV & var & sd & se\\
\hline
Dom & WSB & female & 14 & 184 & 25 & 15 & 13.1 & 3.6 & 0.27\\
\hline
Dom & WSB & male & 11 & 222 & 23 & 11 & 7.2 & 2.7 & 0.18\\
\hline
Dom & G & female & 12 & 318 & 28 & 15 & 17.5 & 4.2 & 0.24\\
\hline
Dom & G & male & 18 & 355 & 23 & 11 & 6.9 & 2.6 & 0.14\\
\hline
Dom & LEW & female & 9 & 147 & 27 & 18 & 23.3 & 4.8 & 0.40\\
\hline
Dom & LEW & male & 10 & 253 & 24 & 13 & 9.6 & 3.1 & 0.20\\
\hline
Cast & CAST & female & 1 & 1 & 26 & NA & NA & NaN & NaN\\
\hline
Cast & CAST & male & 2 & 44 & 22 & 10 & 5.2 & 2.3 & 0.34\\
\hline
Musc & PWD & female & 15 & 222 & 26 & 14 & 14.0 & 3.7 & 0.25\\
\hline
Musc & PWD & male & 8 & 161 & 29 & 11 & 9.8 & 3.1 & 0.25\\
\hline
Musc & MSM & female & 14 & 300 & 28 & 16 & 19.4 & 4.4 & 0.25\\
\hline
Musc & MSM & male & 7 & 166 & 30 & 10 & 9.7 & 3.1 & 0.24\\
\hline
Musc & MOLF & female & 1 & 21 & 28 & 15 & 17.9 & 4.2 & 0.92\\
\hline
Musc & MOLF & male & 6 & 119 & 23 & 11 & 6.4 & 2.5 & 0.23\\
\hline
Musc & SKIVE & female & 1 & 32 & 26 & 12 & 9.8 & 3.1 & 0.55\\
\hline
Musc & SKIVE & male & 3 & 86 & 26 & 10 & 7.4 & 2.7 & 0.29\\
\hline
Musc & KAZ & female & 9 & 184 & 26 & 16 & 16.0 & 4.0 & 0.30\\
\hline
Musc & KAZ & male & 13 & 264 & 23 & 13 & 9.2 & 3.0 & 0.19\\
\hline
Spretus & SPRET & female & 2 & 2 & 26 & 11 & 8.0 & 2.8 & 2.00\\
\hline
Spretus & SPRET & male & 5 & 103 & 24 & 10 & 6.2 & 2.5 & 0.25\\
\hline
Spic & SPIC & female & 6 & 97 & 28 & 16 & 19.5 & 4.4 & 0.45\\
\hline
Spic & SPIC & male & 4 & 133 & 26 & 11 & 7.7 & 2.8 & 0.24\\
\hline
\end{tabular}

\section{genome-wide recombination rate estimates for both
sexes}\label{genome-wide-recombination-rate-estimates-for-both-sexes}

We used counts of MLH1 foci per cell to estimate genome-wide
recombination rates in 14 wild-derived inbred strains sampled from three
subspecies of house mice: \emph{M. m. domesticus}, \emph{M. m. musculus}
and \emph{M. m. molossinus}. Mean MLH1 foci counts for 166 mice were
quantified from an average of 19.73 spermatocytes per male (for a total
of 1867 spermatocytes) and 18.54 oocytes per female (for a total of 1409
oocytes).

Graphical comparisons between the two sexes reveal several patterns
(Figure 1A). First, recombination rate differs between females and males
in most strains. Second, the direction and magnitude of heterochiasmy
varies among strains. Although the majority of strains show higher
recombination rates in females (following the pattern in laboratory
mice), two musculus strains and one molossinus strain exhibit
male-biased heterochiasmy. Hence, relative recombination rates in the
two sexes are evolutionarily labile.

The general patterns of heterochiasmy in house mouse are displayed in
Figure 1A. Taking note of the direction and the magnitude of the sex
differences, our results confirm two general patterns: i) genome-wide
recombination rates averages are greater in females compared to males
(female biased heterochiasmy) and ii) the degree of heterochiasmy
(Female:Male ratio) is generally low, ranging from 1.22 in
\emph{domesticus\textsuperscript{G}} to 1.06 in
\emph{domesticus\textsuperscript{LEW}}. Three notable exceptions of male
biased heterochiasmy are the strains
\emph{musculus\textsuperscript{PWD}},
\emph{molossinus\textsuperscript{MSM}} and
\emph{musculus\textsuperscript{SKIVE}}, with heterochiasmy values of
0.91, 0.93 and 0.99 respectively.

Separately examining the mouse means of MLH1 foci per cell for each sex
points to distinct patterns of variation (Figure 1B-C). Female
recombination rates are evenly distributed around the sex-wide mean of
approximately 25 MLH1 foci per cell (Figure 1B). In stark contrast,
males separate more clearly into two groups of strains with high rates
(near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell)
(Figure 1C).

\section{Partitioning variation in recombination
rate}\label{partitioning-variation-in-recombination-rate}

To situate variation in recombination rate within an evolutionary
framework, we fit a series of models including subspecies, strain, and
sex, to mean MLH1 foci counts from 187 mice. We began with a full mixed
model (M1, see Methods), which showed that strain (random effect p
\textless{} 0 -- 10\^{}\{-6\} someSmallNumber), sex (p =
1.55\times 10\^{}\{-8\}), subspecies (p=1.72\times 10\^{}\{-4\}), and
subspecies sex (p = 3.1\times 10\^{}\{-5\}) each significantly affect
recombination rate.

After a general linear model including all factors as fixed effects (M2)
revealed only weak contributions of subspecies, we focused on additional
models designed to illuminate the role of strain and sex. A general
linear model with these two variables (M3) identified two strains with
particularly strong effects on recombination rate:
musculus\textsuperscript{MSM} (p= 3.99\times 10\^{}\{-6\}) and
domesticus\textsuperscript{G} (1.04\times 10\^{}\{-6\}). In addition,
two strains exhibit strain-by-sex interactions:
molossinus\textsuperscript{MSM} (1.26\times 10\^{}\{-4\}) and
musculus\textsuperscript{PWD} (3.86\times 10\^{}\{-4\}).

We next fit general linear models separately for 192 males and 144
females (M4; see methods). In the male dataset, three strains
significantly affect recombination rate: musculus\textsuperscript{PWD}
((glm; p = 6.31\times 10\^{}\{-8\}; effect = 6.11 foci), and
musculus\textsuperscript{SKIVE} (glm; p = 0.01; 0; effect = 3.8), and
\emph{molossinus\textsuperscript{MSM}} (glm; p=2.42\times 10\^{}\{-12\};
effect 6.99).

These three strains point to rapid evolution in recombination rate in
spermatocytes; we subsequently refer to them collectively as
``high-recombination'' strains. Analysis of the female dataset points to
four strains with significant effects on recombination rate:
domesticus\textsuperscript{G} (p = 2.5\times 10\^{}\{-6\}),
molossinus\textsuperscript{MSM} (p = 6.24\times 10\^{}\{-6\}),
domesticus\textsuperscript{LEW} (p = 0.01), and
musculus\textsuperscript{PWD} (p= 0.02). Strain effect sizes in females
are modest in magnitude (ranging from 1 to 4 foci) compared to those in
males. Together, these results demonstrate heritable differences in the
genome-wide recombination rate evolving in a highly sex-specific manner
over short evolutionary timescales.

\section{Within mouse variance in CO count per
cell}\label{within-mouse-variance-in-co-count-per-cell}

Counting MLH1 foci in multiple oocytes for each female and multiple
spermatocytes for each male allowed us to examine determinants of the
within-mouse variance in recombination rate. To do this, we considered
the same models as above, but replaced mean MLH1 foci count with
within-mouse variance in MLH1 foci count as the dependent variable. Sex
is the only variable that significantly affects recombination rate in
both the mixed model (M1) (p \textless{} 10\^{}\{-6\} -- 0) and general
linear models (M2) (p = 0.03) and M3 (p = 0.03).

In general, females have almost twice as much variance in MLH1 foci per
cell compared to males (Figure 1). Since estimates of within-mouse
variance may be more susceptible to technical error from the staining
protocol, we repeated the analyses using a subset of cells with higher
quality scores (quality score 1 and 2, see Materials and Methods). The
results are similar: sex is the strongest effect (M1 p \textless{}
10\^{}\{-6\} 0; M2 p = 2.3\times 10\^{}\{-4\}; also M3 p =
2.28\times 10\^{}\{-4\}). When both quality-curated and full datasets
are considered, strain does not significantly and consistently affect
variance in MLH1 foci count per cell in either sex. These results
suggest that within-mouse variance in recombination rate evolves
independently of mean recombination rate.

 \textbf{re-read Lenzi et al -- heterogeneity in oogenesis}

\section{DSB. Evolution of genome-wide recombination rate is associated
with evolution of double strand
breaks}\label{dsb.-evolution-of-genome-wide-recombination-rate-is-associated-with-evolution-of-double-strand-breaks}

\includegraphics{Results_v2_files/figure-latex/MAIN.DMC1.plot-1.pdf}

In an attempt to localize the male-specific evolution of crossover
number to steps of the meiotic pathway, we counted foci from a marker
for double strand breaks (DSBs), DMC1, in prophase spermatocytes. DMC1
foci were scored from a total of 0 early zygotene-stage and 0 late
zygotene-stage spermatocytes from three low-recombination strains
(musculus\textsuperscript{KAZ} , domesticus\textsuperscript{WSB} , and
domesticus\textsuperscript{G}) and two high-recombination strains
(musculus\textsuperscript{PWD} and molossinus\textsuperscript{MSM}).

The high-recombination strains have significantly more DMC1 foci than
the low-recombination strains in early zygotene cells (t-test,
p\textless{} -- 10\^{}\{-6\} -- 0). In contrast, the two strain groups
do not differ in DMC1 foci counted in late zygotene cells (t-test, p =
0.66 --- 0.66).

After DSB formation, DSBs are repaired as either non-crossovers (NCO) or
as crossovers (COs), with the vast majority being repaired as NCOs. Thus
the ratio of CO:DSB is a partial indicator of the proportion of DSBs
which are designated as COs. The ratios, calculated for DMC1 means from
both stages, are not significantly different between the high and low
strain groups (t-test, p = 0.94 and p = 0.11 for early zygotene and late
zygotene ratios, respectively).

This comparison raises the possibility that the evolution of crossover
number is primarily due to processes that precede the
crossover/non-crossover decision. (Cole et al. 2012) suggest that early
zygotene counts for DMC1 are most relevant for predicting the final CO
number due to the crossover homeostasis process. This result, combined
with the big difference in MLH1 count between high and low strains,
might indicate that the CO/NCO decision contributes to strain
differences in recombination rate.

\section{Evolution of genome-wide recombination rate is reflected at the
single chromosome
level}\label{evolution-of-genome-wide-recombination-rate-is-reflected-at-the-single-chromosome-level}

\includegraphics{Results_v2_files/figure-latex/Chrm.props.plot2-1.pdf}

To examine the connection between evolution of the genome-wide
recombination rate and changes to the recombination landscape, we used
an image analysis pipeline to measure properties of single bivalents
((Peterson, Miller, and Payseur 2019)).

 This algorithm substantially speeds the accurate measurement of
bivalents, but has the limitation that not all bivalents per cell can be
isolated due to overlapping bivalents. In this dataset, isolation rates
per cell range from 0.51 (molossinus\textsuperscript{MSM} male) to 0.72
(musculus\textsuperscript{KAZ} female).

From the total set of cell images, 10458 bivalent objects were isolated
by the image analysis software. After a human curation step ((Peterson,
Miller, and Payseur 2019)), 9829 single-bivalent observations remained.
We assume that the isolation of bivalents within cells is unbiased.

Given the large number of single-bivalent observations, we assume that
each of the datasets are equally representative of general patterns.

An additional challenge of the MLH1 framework is that the identities of
individual autosomes and the XX in females cannot be easily obtained
(the male XY is distinct).

Ninety-six percent of single bivalents in our pooled dataset (n = 34982)
have either one or two crossovers (Figure X). The proportion of
one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the
high vs.~low recombining strains (Figure X). High-recombination strains
are enriched for 2CO bivalents at the expense of 1CO bivalents:
proportions of 2CO bivalents are 0.33 -- 0.33 (
\emph{musculus\textsuperscript{SKIVE}} ) in
musculus\textsuperscript{SKIVE}, 0.44 -- 0.44 in
musculus\textsuperscript{PWD}, and 0.53 in
molossinus\textsuperscript{MSM}.

Following patterns in the genome-wide recombination rate, male
musculus\textsuperscript{PWD} and male molossinus\textsuperscript{MSM}
have 2CO proportions that are more similar to each other than to strains
from their own subspecies (chi-square tests;
musculus\textsuperscript{PWD} vs.~musculus\textsuperscript{KAZ} p =
3.15\times 10\^{}\{-33\}; molossinus\textsuperscript{MSM}
vs.~molossinus\textsuperscript{MOLF} p = 4.72\times 10\^{}\{-13\}).

Next, we use this single-bivalent dataset to focus on aspects of the
recombination landscape along chromosomes. We address two main
questions. First, which traits are sexually dimorphic? Second, which
traits differ between males from high-recombination
vs.~low-recombination strains?

\section{Q1. Sex Differences in the Length of the Synaptonemal
Complex}\label{q1.-sex-differences-in-the-length-of-the-synaptonemal-complex}

\includegraphics{Results_v2_files/figure-latex/Q1.SC.show.short.biv-1.pdf}

\includegraphics{Results_v2_files/figure-latex/Q1.tot.SC_show-1.pdf}

In many mammalian species, the synaptonemal complex (SC) is longer in
females. Yet, the majority of these observations come from species with
female-biased heterochiasmy. Our wider survey of recombination in house
mice provides an opportunity to determine whether sex differences in
chromatin compaction (SC length) are reversed when heterochiasmy is
male-biased. In addition, if SC length is a strong determinant of the
genome-wide recombination rate, male musculus\textsuperscript{PWD} and
male molossinus\textsuperscript{MSM} should have longer SC lengths than
the other strains.

Short bivalents are significantly longer in females than males in all
strains (t-test; p \textless{} 0.05) except
musculus\textsuperscript{SKIVE}. The female:male ratio ranges across
strains from 1.15 (musculus\textsuperscript{MSM}) to 1.49
(domesticus\textsuperscript{WSB}).

Females have significantly longer total SC than males in each strain
tested separately (t-test; p \textless{} 0.05). That females have longer
SCs is also supported by mixed models and general linear models with
covariates, which identify sex as the most consistently significant
effect (p \textless{} 0.05). Additionally, there are some significant
subspecies and strain effects (p \textless{} 0.05), indicating that SC
length has evolved among strains and subspecies.

In summary, two approaches for measuring and analyzing SC length
indicate that females have longer SCs (chromosome axes), even in strains
where males have more MLH1 foci per cell. Furthermore, joint
consideration of MLH1 foci and total SC length suggests that males from
high-recombination strains have less ``space'' in which to place their
additional crossovers.

\section{Q1. Sex Differences in the Positions of Single
Crossovers}\label{q1.-sex-differences-in-the-positions-of-single-crossovers}

In most of the strains we surveyed, the majority of bivalents are
observed to contain one crossover (focus). In all strains, the landscape
across 1CO bivalents is significantly different in females and males.
Normalized foci positions tend to be more central in females 0.56 and
closer to the telomere in males 0.68 (t-test; p =
2.92\times 10\^{}\{-22\}). Sex is also the most significant effect on
focus position in a mixed model (M1: p = 1.26\times 10\^{}\{-25\}) and
in general linear models (M2: p = 1.33\times 10\^{}\{-7\}; M3: p =
1.33\times 10\^{}\{-7\}). These sex differences in the placement of foci
on 1CO bivalents follow a pattern observed across a variety of mammalian
species (Sardell and Kirkpatrick 2020).

\section{Q1. Sex Differences in CO Interference
(IFD)}\label{q1.-sex-differences-in-co-interference-ifd}

\begin{figure}
\centering
\includegraphics{Results_v2_files/figure-latex/Q1.IFD_load.triangles-1.pdf}
\caption{Example of Sex differences in IFD distributions}
\end{figure}

There is no strong signal of sex differences in raw mean inter-focal
distances (IFD\textsuperscript{raw}) across the full set of strains. A
marginally significant difference between the sexes (t-test; p = 0.07 --
0.07) is driven by one strain (t-test without
domesticus\textsuperscript{G}; p = 0.27 -- 0.02). This result indicates
that females and males exhibit a similar level of interference when it
is measured in physical (SC) units. In contrast, males have
significantly longer normalized mean inter-focal distances
(IFD\textsuperscript{norm}) than females in seven out of eight strains
(t-tests; p \textless{} 0.02 -- 1.49\times 10\^{}\{-12\}), with only
musculus\textsuperscript{KAZ} showing no difference (t-test; p = 0.33 --
0.33). Examination of IFD\textsuperscript{norm} distributions indicates
that female IFD\textsuperscript{norm} values are centered at
approximately 50\% and show a slight enrichment of low (\textless{}25\%)
values, whereas males are enriched for higher values. Mixed models and
general linear models of IFD\textsuperscript{norm} support the inference
of stronger interference in males: sex is the most significant variable
(M1 - LRT \textbf{random}: p = 6.7410\^{}\{-14\} --
6.74\times 10\^{}\{-14\} glm \emph{M2} 0.01 \emph{M3} 0.01). When
interference is measured in physical SC units
(IFD\textsuperscript{raw}), the differences between sexes is low and
only slightly significant (data not shown). In summary, controlling for
differences in chromatin compaction (SC length) using IFDnorm indicates
that interference is stronger in males, whereas consideration of
IFD\textsuperscript{raw} shows that the sexes exhibit a similar level of
interference on the physical (SC) scale. While the comparisons of the
IFD\textsuperscript{norm} metrics can reveal more general recombination
landscape patterns while controlling for the underlying differences in
chromatin compaction and SC length.

\section{Additional Determinants of Genome-wide Recombination Rate
Evolution in
Males}\label{additional-determinants-of-genome-wide-recombination-rate-evolution-in-males}

The next section is meant to focusNext, we used the contrast between
high-recombination strains (musculus\textsuperscript{PWD} ,
musculus\textsuperscript{SKIVE}, and molossinus\textsuperscript{MSM})
and low-recombination strains to identify features of the recombination
landscape associated with evolutionary transitions in the genome-wide
recombination rate. on the greater aspect of variation in mean MLH1
counts per cell the high recombining strains (
musculus\textsuperscript{PWD} , musculus\textsuperscript{SKIVE}, and
molossinus\textsuperscript{MSM}) to the low recombing strains. The main
objective of this section is to test for significant correlations
between features of the recombination landscape and the evolution of
mean MLH1 foci per cell. In comparing the male specific single bivalent
based metrics the first creteria for analysis is a significant
differences between the high and low recombining groups and the second
step is testing for significant subspecies and strain effects in the
same models which used mean MLH1 count per cell.

Previous empirical work suggests basic predictions for the relationship
between the genome wide recombination rate and two aspects of the
recombination landscape. \textbf{SC length is expected to be positively
associated with genome-wide recombination rate because of the loop-axis
strucutre} (zickler Kleckner 1999, Merier 2015?). Crossover interference
strength is expected to be negatively \textbf{associated with
genome-wide recombination rate since interference is defined as the
non-random spacing between crossovers on the same chromosome}.

Following this logic we predict (1) musculus\textsuperscript{PWD} will
have greater SC length and weaker interference than
musculus\textsuperscript{SKIVE}, which in turn will have longer SC and
weaker interference compared to the other \emph{musculus} strains, (2)
molossinus\textsuperscript{MSM} will have longer SC and weaker
interference compared to molossinus\textsuperscript{MOLF}, and 3)
\emph{domesticus} strains will have similar SC length and crossover
interference.

\section{Q2 SC Length}\label{q2-sc-length}

\includegraphics{Results_v2_files/figure-latex/Q2_tot.sc_plot-1.pdf}

\includegraphics{Results_v2_files/figure-latex/Q2.SC_MLH1.by.totalSC-1.pdf}

Confirming the basic predictions, there is a positive correlation
between mean MLH1 foci per cell and total SC ( Spearmans' r = 0.48;

p = 2.24\times 10\^{}\{-10\}). Nevertheless, mean total SC only
partially differentiates high-recombination and low-recombination
strains (Figure X). \textbf{Bret's note on positive correlation, 'A
correlation across mice, across strains, or both?} While
high-recombination strains have significantly more SC area in the total
dataset (t-test; p = 0.01 -- 0.01), separate tests by strain show that
only within \emph{molossinus} are high- and low-recombination strains
significantly different (t-test; molossinus: p = 0.03 -- 0.03; musculus:
p = 0.87 -- 0.87). \textbf{the values above are diferent between
editions}

Additionally, the means for the reduced (short and long) bivalent
datasets are not significantly different between high-recombination and
low-recombination strains (t-test; short: p = 0.88 -- 0.88; long: p =
0.18 -- 0.18).

\textbf{the values between versions are slightly different} In a general
linear model with total SC as the dependent variable, two subspecies
effects are significant (p = musculus 1.2410\^{}\{-6\} --
1.24\times 10\^{}\{-6\}, molossinus p = 10\^{}\{-6\} -- 10\^{}\{-6\}).
In general linear models with reduced bivalent means as dependent
variables, several subspecies and strain effects reach significance (p
\textless{} 0.05) but they are not consistent across models, indicating
to some extent the chromatin compaction evolution is decoupled from
evolution in mean MLH1 foci per cell. \textbf{Bret's comment; How can
you differentiate between this biological conclusion and uncertainty due
to the size of the datasets and statistical approaches?}

\section{Q2.1CO rec landscape evolution is decoupled from gwRR
evolution}\label{q2.1co-rec-landscape-evolution-is-decoupled-from-gwrr-evolution}

\includegraphics{Results_v2_files/figure-latex/Q2.1CO.show_plots-1.pdf}

The normalized 1CO position is not significantly different between
high-recombination and low-recombination strains for the total pooled
data (t-test; p = 0.24 -- 0.24) and also when examined within subspecies
(t-test; p = 0.41 -- 0.41 and p = 0.07 -- 0.07 for \emph{musculus} and
\emph{molossinus}, respectively). \textbf{While there are significant
strain effects for domesticus\textsuperscript{WSB} and
molossinus\textsuperscript{MOLF} in a general linear model with
normalized position as the dependent variable (Figure X)}, \emph{this
evolution of the 1CO positioning is decoupled from the total genome-wide
recombination rate.}

\section{Q2 Evolution of interference is associated with genome wide
recombination rate
evolution}\label{q2-evolution-of-interference-is-associated-with-genome-wide-recombination-rate-evolution}

Mouse averages for both IFD\textsuperscript{raw} and
IFD\textsuperscript{norm} are significantly longer in high-recombination
strains (t-test; IFD\textsuperscript{norm}: p = 7.7410\^{}\{-7\} --
7.74\times 10\^{}\{-7\}; IFD\textsuperscript{raw}: p = 8.7810\^{}\{-6\}
-- 8.78\times 10\^{}\{-6\}). This pattern is confirmed by separate
comparisons within \emph{musculus} (t-test; IFD\textsuperscript{norm}: p
=2.0410\^{}\{-5\} -- 2.04\times 10\^{}\{-5\}; IFD\textsuperscript{raw}:
p = 1.9410\^{}\{-4\} -- 1.94\times 10\^{}\{-4\}) and within
\emph{molossinus} (IFD\textsuperscript{norm}: p= 0.17 -- 0.17;
IFD\textsuperscript{raw}: p = 0.08 -- 0.08). Similar results are seen
with general linear models for both IFD\textsuperscript{raw} and
IFD\textsuperscript{norm}: only effects associated with
high-recombination strains are significant (p \textless{} 0.05).

That IFD\textsuperscript{raw} and IFD\textsuperscript{norm} show similar
patterns eliminates variation in SC lengths and bivalent sizes as
primary explanations. We determined that the main difference in
IFD\textsuperscript{norm} distributions between high-recombination and
low-recombination strains is an enrichment of IFD\textsuperscript{norm}
values under 30\% --

in low-recombination strains. The frequency of IFD\textsuperscript{norm}
values that fall below 30\% ranges from 8.2\%
(domesticus\textsuperscript{G}) to 16\% (musculus\textsuperscript{KAZ})
in low-recombination strains, whereas high-recombination strains all
show such frequencies below 5\% (0\%, 1.3\%, and 3.3\% for
musculus\textsuperscript{SKIVE}, molossinus\textsuperscript{MSM}, and
musculus\textsuperscript{PWD}, respectively).

In summary, the level of interference is a significant predictor of
evolution in the genome-wide recombination rate, but SC length and
crossover position on 1CO bivalents are not. However, the pattern is in
the opposite direction to our prediction: high-recombination strains
have stronger interference. At least on 2CO bivalents, crossovers are
spaced further apart when the genome-wide recombination rate is higher.

\section{Q2 Summary}\label{q2-summary}

\textbf{move main themes to conclusion}

Our results show that the greater crossover interference is the
strongest single bivalent-based predictor for the observed rapid
evolution of mean MLH1 foci per cell. While these results do not conform
initial predictions for how a recombination landscapes would accommodate
more crossovers, the increased strength of interference aligns with our
results on sex differences. The typical recombination landscapes for
males and females results in divergence in the proportion of linked
sites along chromosomes which segregate together. The stronger
interference of the 2CO bivalents in the high recombining strains
accentuates this effect. The measures of DSB and some comparisons of SC
length between high and low recombining strains suggest that the SC
length have evolved to be longer in high recombining strains, however
this evolution of SC length is partially decoupled from the number of
crossovers since similar amounts of SC length evolution are seen in low
recombining strains.

\textless{}Bret: Are you sure this summary is necessary? Maybe cover
this material in the Discussion instead?\textgreater{}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-cole2012}{}
Cole, Francesca, Liisa Kauppi, Julian Lange, Ignasi Roig, Raymond Wang,
Scott Keeney, and Maria Jasin. 2012. ``Homeostatic Control of
Recombination Is Implemented Progressively in Mouse Meiosis.''
\emph{Nature Cell Biology} 14 (4). Nature Publishing Group: 424--30.

\hypertarget{ref-peterson2019}{}
Peterson, April L, Nathan D Miller, and Bret A Payseur. 2019.
``Conservation of the Genome-Wide Recombination Rate in White-Footed
Mice.'' \emph{Heredity} 123 (4). Nature Publishing Group: 442--57.

\hypertarget{ref-sardell_sex_2020}{}
Sardell, Jason M., and Mark Kirkpatrick. 2020. ``Sex Differences in the
Recombination Landscape.'' \emph{The American Naturalist} 195 (2):
361--79. doi:\href{https://doi.org/10.1086/704943}{10.1086/704943}.

\end{document}

### 2020-04-24 src/Result_Setup.Rmd
mostly edits to mat methods

@@
 
-MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
+MLH1.M3.strain.fixed <- glm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
 
@@
 
-MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
 
@@
 #Q12 var
-MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
 
@@
 
-```{r MLH1.var.M4}
-
+```{r MLH1.var.M4, echo=FALSE}
 #male
@@
 glmMLH1Var_M4.female <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "female") ) )
+sum.glmMLH1Var_M4.female <- summary(glmMLH1Var_M4.female)
 

### 2020-04-26 doc/MatMeth.Rmd
adding in edits

@@
 
-```{r data.sets, echo=FALSE}
-
-#Q1.real.short_mouse_table
-
-#
-
-```
-
 ### Mouse Husbandry
@@
 
-<Mice were fed on dry standard breeder chow.
+Mice were fed on dry standard breeder chow.
 Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
@@
 Over the course of data collection some breeding colonies  mice were moved from facilities.
-Additionally the GI strain was kept at a separate facility.>
+Additionally the GI strain was kept at a separate facility.
 
@@
 
-The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
+The dry-down spread protocol [@peters_1997] was used for both spermatocytes and oocytes with adjustment for volumes.
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
+The majority of mice used for MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
 Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
 
-Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
-Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
+Whole testes or decapsulated ovaries were incubated in 3ml or 300ul of hypotonic solution for 45min.
+Fifteen microliters of cell slurry were transferred to 80ul of 2% PFA solution.
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a photoflow wash (Kodak, cite).
-**slides were stored at -20*C if not stained imeediantly**
+The following morning, slides were treated with a photoflow wash (Kodak, cite) and stored at -20*C if not stained immediantly.
 
@@
 
-The **immunohistochemistry**
+**Immunohistochemistry**
 
 To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
-Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
-The staining protocol was based on that in [@anderson1999] and [@koehler2002].
-Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
-
-Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
-The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
+Double strand breaks (DSB) and crossovers were visualized with DMC1 and MLH1 respectively.
+The staining protocol was based on  [@anderson1999]  and  [@koehler2002].
+Following a 30 minute blocking wash in 1X ABD, (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), and  Triton X-100 (Sigma) ), each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
+The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1 (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
 Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
@@
 
-Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
+Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
@@
 
-Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
-
-SC morphology and CREST/centromere number was used to stage spermatocytes **Certeria for early and late zygotene** - (bright staining - )
-early zygotene defined as seperate centromere pairing -- mostly complete axis signal.
-late zygotene judged as -- low degree of centromere pairing -- and more homologous pairing (but not fully synapsis).
+Mean DMC1 foci per cell for a single juvenile mouse from a subset of strains; _musculus^PWD^_, _musculus^MSM^_, _domesticus^WSB^_, and _domesticus^G^_ was quantified.
 Differences in DMC1 counts per cell were tested by ttests.
-<numbers of cell?>
+(lateral elements of) Synaptonemal complex morphology and CREST/centromere number was used to stage spermatocytes.
+Early zygotene cells were defined as having a larger degree of indiviual centromeres and mostly complete axis signal.
+Late zygotene cells were identified as more paired centromeres and homolog axis (but below full synapsis).
 
@@
 
-**certeria**  MLH1 foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+Mean **MLH1** foci counts were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct, clear MLH1 foci.
 A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
-With scores of 1 and 2 having the highest quality.
-Cells and coded as a score of 5 were excluded from the final analysis.
+Scores of 1 and 2 have the highest quality while those coded as a score of 5 had the lowest quality and were excluded from the final analysis.
 The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
 MLH1 foci located on the XY in spermatocytes were excluded from this study.
+**(mouse quality) --- average of X cells quantified per female and male mice respectively - only included mice with at least 12?
+10 cells quantified.**
+
+To characterize the SC length and chromatin compaction, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
+Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
+Hand measures of pachtyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
+To remove erroneous SC isolation from the total SC algorithm, outliers were visually assessed at the mouse level and removed from the data set.
+Mouse means were calculated from cell-wide total SC lengths in 2984 --  `r  formatC( length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark="," )` out of 3680 -- `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name),format="d", big.mark=",")` cells with MLH1 counts (Figure X).
+
+The DNA Cross Over algorithm isolates single straightened bivalents from an image, and returns the SC length, and locations of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
+This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
+
+In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r round( min(algorithm.stat.category$mean_pass.rate), 4 )` (molossinus^MSM^ male) to `r round( max(algorithm.stat.category$mean_pass.rate), 4)` (_musculus^KAZ^_ female).
+A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
+
+From the total set of cell images, `r  formatC(length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10458 bivalent objects were isolated by the image analysis software.
+After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` --- 9829 single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
+From the curated single bivalent data set the proportions of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+
+The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
+A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 - `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
+A total of `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$n_obs)` and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$n_obs)` were isolated from `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$ncells)` spermatoctyes and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes respectively.
+
+-MLH1 traits, SC based traits, CO position along single bivalents
 
@@
 
+All statistical analyses were done in R (v3.5.2) [@Rstudio].
 Straight forward comparisons of groups (strain, sex) were done through t.tests().
+**(outgroup species were not included for the analyses).**
+The full mixed model (**M1**) was built to interpret the patterns of variation using the lmer() from the lmer4 package (cite).
+To reflect a random sample of natural genetic variation from each subspecies, strain was coded as a random effect and tested using exactRLRT().
+The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
+This model allows us to estimate and test the fixed effects (sex and subspecies), and variation in mean MLH1 counts due to strain effect.
+We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
+The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model. The general linear models, **M4** were used to analyze sex specific data sets. 
+We used this analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
+sex, strain subspecies) and can be applied to multiple dependent variables.
 
-We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
-Strain was coded as a random effect and tested using exactRLRT() to reflect a random sample of natural genetic variation from each subspecies.
-The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
+**slightly expand on the other traits**
 
-This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
-The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
-The general linear models, **M4** were used to analyze sex specific data sets.
+For indiviual mice the MLH1 mean, variance and coefficicent of variance per cell, five other related meiotic traits were investigated.
 
-Our goal was to design an analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
-sex, strain subspecies) and can be applied to multiple (meiotic traits) dependent variables.
-In addition to the MLH1 count variables for individual mice, mean MLH1 count per cell, variance and coefficient of variance for MLH1 count per cell, we analyzed SC based features: mouse average of total SC per cell, mouse average of single SC length, mouse mean IFD^norm^ and IFD^raw^ for double crossover chromosomes, and the average position of crossover for single crossover bivalents.
+We quantified mouse averages of SC based features including the 
+mean total SC area per cell and the mean SC lengths for single bivalents.
 
-These traits were choosen because they are markers of phenomena in the recombinaiton pathway and provide more detailed information about the recombination landscape.
+mouse average of inter-focal distances IFD^norm^ and IFD^raw^ for double crossover chromosomes, and average position of crossover for single crossover bivalents.
+
+These traits were choosen because they are markers of processes in the recombinaiton pathway and provide more detailed information about the recombination landscape.
 In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
@@
 
-```{r single.biv.chr, echo=FALSE, include=FALSE}
-
-# image analysis programs
-# Richard total sc
-
-# DNA.CrossOver (site, efficiency, curation, chromosome size effects-reduced dataset,
-
-# chromosome proportions, number of CO per chrm, proportions perpared by chi-sq tests
-
-#chrm chrteriztions -- SC length (chromatin compaction), foci position(raw and normalized), and interfocal distance, 
-#these traits were fit into the main models (as mentioned above) ... t.tests were used to test between sexes and groups
-
-# sc bivalent used in models (aldready above), the length of single SC compared by t-tests -- and fit into models
-# Foci Position -- t-test  and put into models
-# IFD / interference
-```
-
-To characterize the SC lengths across bivalents, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
-Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
-Hand measures of SC/pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
-
-The **total SC** per cell was quantified from the pachtyene cell images.
-To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` out of `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
-
-The **DNA CrossOver** algorithm isolates single straightened bivalents/chromosomes shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)` (musculus^KAZ^ female).
-
-A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
-From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` bivalent objects were isolated by the image analysis software.
-After a human curation step , `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step ([@peterson2019])).
-
-From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
-
-**REDUCED DATASET**
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
-
 # References

### 2020-04-26 src/SetupData.Rmd
wrote main column names to the main MLH1 file

@@
 to Do
-
-- rm Peromycus data BivData!!
-
-- code Mol as discint subsp
-
-- make sure to remove some of the juvi mice!! from the AP_mouse table
 
@@
 
-MLH1_data = read.csv(here("data", "/MLH1/AnonData.csv"), header=TRUE)
+MLH1_data = read.csv(here("data", "/AnonData_formated.csv"), header=TRUE)
 
@@
 
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
+#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+#                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
+#                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
 
-Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
+#Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+#                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
 ```
@@
 #make the table with ALL MLH1 obs
-
 FULL_mouse_table_MLH1 <- ddply(MLH1_data, c("mouse"), summarise,
@@
 
+Number of cells per mouse, (were any mice excluded for low numbers of cells?) (were all cells pooled?)
+
 ## Mouse Ages
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_4.26.20.RData")
 

### 2020-04-26 doc/MatMeth.Rmd
more edits to material methods

@@
 #load datafile
-load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
+load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")
+#MLH1_data_setup_4.26.20.RData
 
@@
 
-We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).
+<We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).>
 
@@
 Differences in DMC1 counts per cell were tested by ttests.
-(lateral elements of) Synaptonemal complex morphology and CREST/centromere number was used to stage spermatocytes.
+The lateral elements of the synaptonemal complex morphology and CREST/centromere number was used to stage spermatocytes.
 Early zygotene cells were defined as having a larger degree of indiviual centromeres and mostly complete axis signal.
@@
 MLH1 foci located on the XY in spermatocytes were excluded from this study.
-**(mouse quality) --- average of X cells quantified per female and male mice respectively - only included mice with at least 12?
-10 cells quantified.**
+An average of `r round( mean( ( AP_mouse_table.HQ %>% filter(sex == "female") )$Ncells), 4)` and `r round( mean( ( AP_mouse_table.HQ %>% filter(sex == "male") )$Ncells), 4)` cells were quantified for female and male mice respectively.
 
@@
 To remove erroneous SC isolation from the total SC algorithm, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in 2984 --  `r  formatC( length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark="," )` out of 3680 -- `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name),format="d", big.mark=",")` cells with MLH1 counts (Figure X).
+Mouse means were calculated from cell-wide total SC lengths in  `r  formatC( length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark="," )` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name),format="d", big.mark=",")` cells with MLH1 counts (Figure X).
 
@@
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r round( min(algorithm.stat.category$mean_pass.rate), 4 )` (molossinus^MSM^ male) to `r round( max(algorithm.stat.category$mean_pass.rate), 4)` (_musculus^KAZ^_ female).
+In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r round( min(algorithm.stat.category$mean_pass.rate), 3)` (molossinus^MSM^ male) to `r round( max(algorithm.stat.category$mean_pass.rate), 3)` (_musculus^KAZ^_ female).
 A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
-
-From the total set of cell images, `r  formatC(length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10458 bivalent objects were isolated by the image analysis software.
-After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` --- 9829 single-bivalent observations remained.
+From the total set of cell images, `r  formatC(length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` bivalent objects were isolated by the image analysis software.
+After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
 The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
@@
 To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
+A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+A 'long bivalent' data set was calculated as above, using the 4th quartille for SC lengths per cell.
+A total of `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$n_obs)` and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$n_obs)` bivalents were isolated from `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$ncells)` spermatoctyes and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes respectively.
 
-The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 - `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
-A total of `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$n_obs)` and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$n_obs)` were isolated from `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$ncells)` spermatoctyes and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes respectively.
+For indiviual mice the MLH1 mean, variance and coefficicent of variance per cell were investigated.
+We quantified mouse averages of SC based features including the mean total SC area per cell and the mean SC lengths for single bivalents.
+The crossover position traits crossover position for single crossover bivalents and crossover interference from double crossover interference.
+For double crossover bivalents the distance between two MLH1 foci, inter-focal distances (IFD), was examined in IFD^raw^ or the normalized by the SC length IFD^norm^.
+These traits were examined using a our model framework because they are markers of processes in the recombinaiton pathway and provide more detailed information about the recombination landscape.
 
--MLH1 traits, SC based traits, CO position along single bivalents
+All statistical analyses were done in R (v3.5.2) [@Rstudio].
+Straight forward comparisons of groups (strain, sex) were done through t.tests().
+**(outgroup species were not included for the analyses).**
+The full mixed model (**M1**) was built to interpret the patterns of variation using the lmer() from the lmer4 package (cite).
+To reflect a random sample of natural genetic variation from each subspecies, strain was coded as a random effect and tested using exactRLRT().
+The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
+This model allows us to estimate and test the fixed effects (sex and subspecies), and variation in mean MLH1 counts due to strain effect.
+We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
+The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
+The general linear models, **M4** were used to analyze sex specific data sets.
+We used this analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
+sex, strain subspecies) and can be applied to multiple dependent variables.
 
@@
 
-All statistical analyses were done in R (v3.5.2) [@Rstudio].
-Straight forward comparisons of groups (strain, sex) were done through t.tests().
-**(outgroup species were not included for the analyses).**
-The full mixed model (**M1**) was built to interpret the patterns of variation using the lmer() from the lmer4 package (cite).
-To reflect a random sample of natural genetic variation from each subspecies, strain was coded as a random effect and tested using exactRLRT().
-The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
-This model allows us to estimate and test the fixed effects (sex and subspecies), and variation in mean MLH1 counts due to strain effect.
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
-The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model. The general linear models, **M4** were used to analyze sex specific data sets. 
-We used this analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
-sex, strain subspecies) and can be applied to multiple dependent variables.
-
-**slightly expand on the other traits**
-
-For indiviual mice the MLH1 mean, variance and coefficicent of variance per cell, five other related meiotic traits were investigated.
-
-We quantified mouse averages of SC based features including the 
-mean total SC area per cell and the mean SC lengths for single bivalents.
-
-mouse average of inter-focal distances IFD^norm^ and IFD^raw^ for double crossover chromosomes, and average position of crossover for single crossover bivalents.
-
-These traits were choosen because they are markers of processes in the recombinaiton pathway and provide more detailed information about the recombination landscape.
-In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
-Normalized measures were calculated by the position divided by the length of the SC for that single bivalent.
-
 # References

### 2020-04-26 src/Result_Setup.Rmd
more edits to material methods

@@
 
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.26.20.RData")
 #load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
@@
 
-```{r Q1.rbar, echo=FALSE}
+```{r Q1.rbar, echo=FALSE, eval=FALSE}
 
@@
 #DF.HetC.MixedModel.HQ
+
+#no mice removed for low cell numbers?
 
@@
 AST.HETC.val <- AST.mean.MLH1.obs$mean.MLH1[1] / AST.mean.MLH1.obs$mean.MLH1[2]
-
 #Molosinus
@@
 HetC.MSM <- MLH1_data[(MLH1_data$strain == "MSM")&(MLH1_data$quality < 5),]
-unique(HetC.MSM$mouse)
+#unique(HetC.MSM$mouse)
 

### 2020-04-27 src/Result_Setup.Rmd
re-formated the strain level table

@@
 
-#this block should be in the data set up
-#the mouse averages will be used
-
 #18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        Ncells  = length(unique( fileName) ),
+                        nbiv = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
                         cV = cv(chromosomeLength),
@@
                         se   = round(sd / sqrt(Ncells), 3),
+
+              #can't figure out how to get this format to work  
+                       n0CO = sum(length( which(hand.foci.count == 0) )  ),
+                       n1CO = sum(length( which(hand.foci.count == 1) )  ),
+                      n2CO = sum(length( which(hand.foci.count == 2) )  ),
+                      n3CO = sum(length( which(hand.foci.count == 3) )  ),
+              #     
+                  proportion_0CO = (n0CO / nbiv),
+                  proportion_1CO = (n1CO / nbiv),
+                 proportion_2CO = (n2CO / nbiv),
+                     proportion_3CO = (n3CO / nbiv),
 
@@
 
+                #mean 1CO.pos
+                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
+                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
+                #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
                 # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+            #  mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
+            # mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),
 
@@
 
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
               #sis.co.ten!!
               mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
               mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
@@
 
+mouse.BivData.Table <- mouse.avs_4MM
+
+```
+
+```{r strain.curated.biv, echo=FALSE}
+
+strain.BivData.Table <- ddply(mouse.BivData.Table, c("strain", "sex"), summarise,
+
+                        Nmice = length(Nmice),
+                        Ncells  = sum(Ncells), #might not to be sum
+                        mean.SC = mean(mean_SC),
+                        var.SC = mean(var), #mouse average variance of bivalent per mouse
+
+                        #chrm classes
+                        av.proportion_0CO = mean(proportion_0CO),
+                        av.proportion_1CO = mean(proportion_1CO),
+                        av.proportion_2CO = mean(proportion_2CO),
+                        av.proportion_3CO = mean(proportion_3CO),
+
+                        #1CO pos
+                        mean.1CO.pos = mean(mean_1CO.pos),
+
+                        #mean.IFD
+                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS),
+                        mean.IFD.PER = mean(mean_IFD.2CO_PER)
+
+                        )
+
+#add
+strain.BivData.Table <- add_subsp(strain.BivData.Table)
+strain.BivData.Table <- add_species(strain.BivData.Table)
+
 ```
@@
 
+```{r Q1.Short.Long_biv.setup, echo=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#make the Q1 strain list
+Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+```
+
 ```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#add this to setup
-
-#remove female, 
-#add back in the BIVDATA from strains which were  removed for the sex-specific comparisons
-
-#Curated_BivData_male  -- RE
 
@@
                  mean.skel = mean(skel_size, na.rm = TRUE),
+                 var.skel = var(skel_size, na.rm = TRUE),
                 #se for skel
@@
 
+#
 totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
@@
 
-```{r tot.SC.setup, echo = FALSE}
+```{r tot.SC_strain.setup, echo = FALSE}
 
@@
                  mean.skel = mean(mean.skel, na.rm = TRUE),
+               # var.skel = var(mean.skel, na.rm = TRUE), #average mouse variance
                  mean.bin =  mean(mean.bin, na.rm = TRUE)
@@
 
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
 # MLH1 models
@@
 #Kaz Z are higher than Dom
+```
+
+# Q1 single bivalent section
+
+```{r Q1.single.biv_position_table, echo=FALSE}
+#strain averages for single bivalent 
+#
+colnames(Curated_BivData99)
+
 ```

### 2020-04-28 doc/Results_v2.rmd
commit before restart

@@
 title: "MLH1 Results Outline"
-date: 
+date: 'null'
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
   html_document:
@@
     toc_float: yes
-- \linenumbers
+  word_document:
+    toc: yes
+    toc_depth: '5'
 bibliography: refs_Main_Report.bib
@@
 library(pander)
+library("xlsx")
 
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE }
-#DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
-
-#colnames(DMC1table.main99) <- c("category", "mean MLH1",    "ncells",     "mean DMC1", "MLH1:DMC1 ratio", "ncells",     "mean DMC1", "MLH1:DMC1 ratio")
-
-#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-#  kable_styling() %>%
-#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#write the full table to a supplemental table
-#show the head (or shortened)
-
-#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-
-#kable(strain.table.HQ_Fig, "latex")
-
-df.strain.table <- strain.table.HQ_Fig
-
-#kable(df.strain.table)
-
-pander(df.strain.table)
-
-#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-```
-
 ```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
@@
 
-# genome-wide recombination rate estimates for both sexes
+# Genome-wide recombination rate estimates for both sexes
 
@@
 
+```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
+#write the full table to a supplemental table
+#show the head (or shortened)
+#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+#kable(strain.table.HQ_Fig, "latex")
+
+df.strain.table <- strain.table.HQ_Fig
+#kable(df.strain.table)
+#pander(df.strain.table)
+
+#reorder the df.strain.table
+#species, subspecies, strain
+
+#code for reording column names...
+#DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", #"ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
+
+#write.xlsx(df.strain.table, file = "Results_Table.xlsx", sheetName = "MLH1.results", append = T, row.names = F)
+
+kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
+  kable_styling() %>%
+  scroll_box(width = "800px", height = "200px")
+
+#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+```
+
 # Partitioning variation in recombination rate
@@
 
-#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-#  kable_styling() %>%
-#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-#  scroll_box(width = "800px", height = "200px")
-
-pander(DMC1table.main99)
-
+kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
+  kable_styling() %>%
+  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
+  scroll_box(width = "800px", height = "200px")
+
+#pander is better for word
+#pander(DMC1table.main99, row.names = F)
+
+#write table
+#write.xlsx(DMC1table.main99,file = "Results_Table.xlsx",sheetName = "DMC1.results", append = FALSE, row.names=F)
 ```
@@
 
-<Bret's Q, ANOVA is not appropriate for 2 samples unless it's specilaized somehow>
-
-```{r DMC1.subsampling, echo=FALSE, eval = FALSE, include = FALSE}
-
-#Subsampling code -- don't report
-
-#The number of observations are not quite equal across strains and cell stages, so I ran permutations sampling 25 cells from the high and low recombining groups of cells.
-Below are the distributions of p values for permuted t-tests from sampling 25 cells from the pooled high and low recombining groups.
-Permutations were done for both cell stages.
-The actual p value is in red.
-
-#The observed p value does not fall outside of the permuted distribution with assumes equal cell sampling.
-
-#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
-# the 2 groups
-
-#make the 2 pools
-High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
-
-Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
-
-#permute drawing / subsampling and performing the t-test
-#put the p.value results into a DF!
-
-#permut.SC.cv.diff  <- as.data.frame(
-#sample(THE.DF, SAMPLE.SIZE, REPLACE?)
-
-#permuting samples -- and re-licating t-tests
-
-#nrep <- 1000
-
-#permuting samples to replicate the t.test across cell stages from high and low with equal sample sizes.
-bunchOpvals.L <- replicate(1000, 
-    # t.test
-    t.test( sample(High.L.pool, 25, replace = FALSE),
-            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
-          )
-bunchOpvals.L <- as.data.frame(bunchOpvals.L)
-bunchOpvals.L$stage <- "L"
-
-bunchOpvals.Z <- replicate(1000, 
-    # t.test
-    t.test( sample(High.Z.pool, 25, replace = FALSE),
-            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
-          )
-
-bunchOpvals.Z <- as.data.frame(bunchOpvals.Z)
-bunchOpvals.Z$stage <- "Z"
-
-#could merge the lists of p-values
-replicated.Ttests <- cbind(bunchOpvals.L, bunchOpvals.Z)
-#have to melt, the rbinded dataframe to get the cols 'stage' and 'pvale'
-
-#ASSUMPTIONS; normality, equal variance
-
-#plot the lists of pvals in histogram? compare to the observed pvalue to get a sense as of how much sample size shifts the pvalues.
-
-#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")
-
-#plot.pvals.Z <- ggplot(bunchOpvals.Z, aes(bunchOpvals.Z) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Zygotene")+geom_vline(xintercept=ttest.HighLow.Z$p.value, linetype="dotted", color="red")
-#plot.pvals.L
-#plot.pvals.Z
-
-#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
-#n use the var.equal = TRUE
-#https://www.statmethods.net/stats/ttest.html
-```
-
-```{r normality, echo=FALSE, warning=FALSE, eval=FALSE}
-#these are histogram plots for testing Normality shape -- move to supplemenatry section
-#- check that distributions are close to normal
-#- check for equal variance.
-
-DMC1.histogram <- ggplot(DMC1.DF, aes(total.count, fill = stage)) + geom_histogram(alpha = 0.4)+facet_wrap(~strain)+ ggtitle("DMC1 histogram")
-
-#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
-#bartlet's test, Levine's test
-```
-
 # Evolution of genome-wide recombination rate is reflected at the single chromosome level
@@
 ```{r algorithm_performance_stats, echo=FALSE}
-
-#sort by category, 
-#compare total boxes and rate passed
-#Curated_BivData  -- pass rate?
 
@@
 
-#add category mean
-#totSC.strain.av.table %>% filter
-#totsc.dom_tot.sc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Dom" )
-#totsc_strain.av.dom <- totSC.strain.av.table %>% filter( subsp == "Dom" )
-
-Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(sex~subsp, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area") + theme(legend.position="none")
+#COPY THIS CODE TO BELOW
+#Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
+
+#Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
+
+#change this plot to look like the short bivalent plot -- boxplots + mouse points
+Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
+  scale_color_manual(values=colors_of_strains)+
+
+  facet_wrap(~subsp, scales="free", ncol = 3)+ 
+  geom_jitter()+ geom_boxplot(alpha=.2)+ 
+  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area") + 
+  #change the white themes
+  theme(legend.position="none")
 
 Tot.SC_mouse.av_ALL
-
-#full dataset ---
 
@@
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
+
+```
+
+```{r Q1.total_SC.table, echo=FALSE}
+
+totSC.strain.av.table <-  totSC.strain.av.table[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
+
+#head of table
+#see the DMC1 code
+
+kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
+  kable_styling() %>%
+  scroll_box(width = "800px", height = "200px")
+
+#add variance
+#write.xlsx(totSC.strain.av.table, file = "Results_Table.xlsx", sheetName = "Total.SC.results", append = TRUE, row.names = F)
 
@@
 ```{r show.long.biv, echo=FALSE, include=FALSE}
+
 #mouse av
@@
 
-```{r show.siscoten, echo=FALSE, eval=FALSE}
-#think of a good way to display this
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_female
-
-sis.co.ten.plot_WSB_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-sis.co.ten.plot_WSB_male
-
-sis.co.ten.plot_LEW_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_female
-
-sis.co.ten.plot_LEW_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "LEW male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_LEW_male
-
-sis.co.ten.plot_PWD_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_female
-
-sis.co.ten.plot_PWD_male <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-#sis.co.ten.plot_PWD_male
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-jitter_siscoten_PWDf <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ ggtitle("PWD female Sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDf
-
-jitter_siscoten_PWDm <- ggplot(data = Curated_BivData[Curated_BivData$category == "PWD male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("PWD male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_PWDm
-
-jitter_siscoten_WSBf <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB female sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBf
-
-jitter_siscoten_WSBm <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB male",], aes(y=SisCoTen, x=hand.foci.count, color = as.factor(hand.foci.count) ))+ggtitle("WSB male sis.co.ten")+
-  geom_jitter()+facet_wrap(sex~strain)
-#jitter_siscoten_WSBm
-```
-
-```{r show.norm.siscoten, echo=FALSE, eval=FALSE}
-Curated_BivData$nrm.siscoten <- Curated_BivData$SisCoTen/Curated_BivData$chromosomeLength
-
-Nrm.sis.co.ten.plot_WSB <- ggplot(data = Curated_BivData[Curated_BivData$strain == "WSB",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n WSB")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_WSB
-
-Nrm.sis.co.ten.plot_G <- ggplot(data = Curated_BivData[Curated_BivData$strain == "G",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n G")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_G
-
-Nrm.sis.co.ten.plot_LEW <- ggplot(data = Curated_BivData[Curated_BivData$strain == "LEW",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten\nLEW")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_LEW
-
-Nrm.sis.co.ten.plot_PWD <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n PWD")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_PWD
-
-Nrm.sis.co.ten.plot_KAZ <- ggplot(data = Curated_BivData[Curated_BivData$strain == "KAZ",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n KAZ")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_KAZ
-
-Nrm.sis.co.ten.plot_SKIVE <- ggplot(data = Curated_BivData[Curated_BivData$strain == "PWD",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n SKIVE")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_SKIVE
-
-Nrm.sis.co.ten.plot_MSM <- ggplot(data = Curated_BivData[Curated_BivData$strain == "MSM",], aes(y=SisCoTen/chromosomeLength, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(~sex)+ggtitle("Normalized siscoten \n MSM")+theme(legend.position="none")
-Nrm.sis.co.ten.plot_MSM
-
-```
-
-```{r sis.co.ten.test_HetC, echo=FALSE, eval=FALSE}
-
-#glm for siscoten
-#mouse.avs_4MM  - mouse averages with mean sis.co.ten
-#MM
-lme_MM_siscoten <- lme(mean.siscoten ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_siscoten)$tTable, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-lmer_MM_siscoten <- lmer(mean.siscoten ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-#drop1(lmer_MM_siscoten, test="Chisq")
-
-rand_siscoten_MM <- exactRLRT(lmer_MM_siscoten)
-rand_siscoten_MM
-#slightly sig
-```
-
-```{r siscoten.test.poly, echo=FALSE, eval=FALSE}
-
-#move this to the set up blocks
-
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#test high vs low Musc strains
-#mouse.avs_4MM
-```
-
-```{r log.reg_siscoten, echo=FALSE, eval=FALSE}
-
-#replace the table
-
-male_mouse.av$Rec.group <- ifelse(grepl("PWD", male_mouse.av$strain), 1, 
-                            ifelse(grepl("MSM", male_mouse.av$strain), 1,
-                             ifelse(grepl("SKIVE", male_mouse.av$strain), 1, 0)) )
-
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 
-                            ifelse(grepl("MSM", Male.poly.Mouse.Table_BivData_4MM$strain), 1,
-                             ifelse(grepl("SKIVE", Male.poly.Mouse.Table_BivData_4MM$strain), 1, 0)) )
-
-#SISCOTEN
-#mouse av, norm.first foci is from 1CO bivalents
-
-log.reg_SIS.co.ten_mouse.av <- glm(Rec.group ~ mean.siscoten,
-
-                                 data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-summary(log.reg_SIS.co.ten_mouse.av)
-
-#single biv level
-log.reg_Sis.co.ten_biv <- glm(Rec.group ~ SisCoTen, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Sis.co.ten_biv)
-
-```
-
-```{r t.test_sis.co.ten,echo=FALSE, eval=FALSE}
-
-#bivData --
-
-t.test_mouse.av_siscoten <- t.test(
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(mouse.avs_4MM$Rec.group == 1)],
-
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc") &(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#1CO
-t.test_mouse.av_siscoten_1CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_1CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#2CO
-t.test_mouse.av_siscoten_2CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_2CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-#3CO
-t.test_mouse.av_siscoten_3CO <- t.test( 
-  Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")& (Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-   Male.poly.Mouse.Table_BivData_4MM$mean.siscoten_3CO[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
-  )
-
-####
-#Single Biv Level -- from the main 
-#pull the single BivData -- MUSC
-
-#check that these have the extra male strains
-
-t.test_single.biv_siscoten <- t.test(
-
-  Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-#1CO
-mouse.avs_4MM$mean.siscoten_2CO
-
-```
-
-```{r telo.dist, echo=FALSE, eval=FALSE}
-
-sis.co.ten.plot_WSB_female <- ggplot(data = Curated_BivData[Curated_BivData$category == "WSB female",], aes(y=SisCoTen, x=chromosomeLength, color = as.factor(hand.foci.count) ))+
-  geom_point()+facet_wrap(sex~strain)
-
-Curated_BivData <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$hand.foci.count) | Curated_BivData$hand.foci.count==""), ]
-
-telo.dist_boxplots <- ggplot(Curated_BivData, aes(x= hand.foci.count, y=telo_dist ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+ ylim(c(0, 150))
-
-telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw telomere distance")+ ylim(c(0, 150))
-telo_jitter
-
-nrm.telo_jitter <- ggplot(Curated_BivData, aes(x= strain, y=telo_dist_PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized telomere distance")+ylim(c(0,1.2))
-nrm.telo_jitter
-
-#why do a different number of rows removed each time??
-```
-
-```{r MM.telo, echo=FALSE, eval=FALSE}
-#HetC
-
-lme_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data= Mouse.Table_BivData_4MM )
-drop1(lme_MM_telo.dist, test="Chisq")
-as.data.frame( round(  summary(lme_MM_telo.dist)$coefficients, 6 ) )
-
-#sex and sex*subsp interaction effect are significant for mean sis.co.ten
-
-lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-#anova(lmer_MM_siscoten)
-
-#thing1  <- summary(lmer_MM_siscoten)
-drop1(lmer_MM_telo.dist, test="Chisq") 
-rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
-rand_telo.dist_MM
-```
-
-```{r log.reg_telo.dist, echo=FALSE, eval=FALSE}
-
-Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
-                            ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
-                             ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)) )
-
-#TELO.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_Telo.Dist_mouse.av <- glm(Rec.group ~ mean.telo.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_Telo.Dist_mouse.av)
-
-#single biv level
-log.reg_Telo.Dist_biv <- glm(Rec.group ~ telo_dist, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist_biv)
-
-#PER versions
-log.reg_Telo.Dist.PER_biv <- glm(Rec.group ~ telo_dist_PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_Telo.Dist.PER_biv)
-```
-
-```{r t.test_Telo.Dist, echo=FALSE, eval=FALSE}
-
-#mouse av
-t.test_mouse.av_telo.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.telo.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Telo.Dist <- t.test(
-
-  Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Telo Dist PER
-single.biv_t.test_Telo.Dist_PER <- t.test(
-
-  Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$telo_dist_PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
-```{r cent.dist_show, echo=FALSE, eval=FALSE}
-cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))
-
-cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere dist")+ ylim(c(0, 200))
-cent.dist_jitter
-
-norm.cent.dist_jitter <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent.PER, color=as.factor(hand.foci.count) ) )+ geom_jitter(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Normalized centromere distance")+ylim(c(0,1.2))
-norm.cent.dist_jitter
-
-```
-
-```{r MM.cent.dist, echo=FALSE, eval=FALSE}
-
-lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= Mouse.Table_BivData_4MM, random=list(strain=pdDiag(~sex) ) )
-as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )
-
-lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-exactRLRT(lmer_MM_cent.dist)
-```
-
-```{r log.reg_CENT.dist, echo=FALSE, eval=FALSE}
-
-#CENT.DIST
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_CENT.Dist_mouse.av <- glm(Rec.group ~ mean.cent.dist,
-
-                                 data=Mouse.Table_BivData_4MM[(Mouse.Table_BivData_4MM$subsp == "Musc")&(Mouse.Table_BivData_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_CENT.Dist_mouse.av)
-
-#single biv level
-log.reg_CENT.Dist_biv <- glm(Rec.group ~ dis.cent, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist_biv)
-
-#PER versions
-log.reg_CENT.Dist.PER_biv <- glm(Rec.group ~ dis.cent.PER, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-summary(log.reg_CENT.Dist.PER_biv)
-```
-
-```{r t.test_CENT.Dist, echo=FALSE, eval=FALSE}
-#CENT.DIST
-#mouse av
-t.test_mouse.av_Cent.dist <- t.test(
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$Rec.group == 1)],
-
-  Mouse.Table_BivData_4MM$mean.cent.dist[(Mouse.Table_BivData_4MM$sex == "male") & (Mouse.Table_BivData_4MM$subsp == "Musc") &(Mouse.Table_BivData_4MM$Rec.group == 0)] 
-  )
-
-#Single Biv Level
-single.biv_t.test_Cent.Dist <- t.test(
-
-  Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 0)]
-  )
-
-#Cent Dist PER
-single.biv_t.test_Cent.Dist_PER <- t.test(
-
-  Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male") & (Curated_BivData$Rec.group == 1)],
-
-   Curated_BivData$dis.cent.PER[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
-  )
-
-```
-
 ```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
@@
 #or AIC(model1,model2)
-
-```
-
-```{r Q1.mean.nrmF1_table, echo=FALSE, eval=FALSE}
-
-#make a table for the sex specific values of the normalized F1 positions
-
-#for strain, sex -- Curated_BivData
-#or (Mouse.Table_BivData_4MM)
-#norm.first.foci.pos
-Q1.normF1.pos_table <- ddply(Mouse.Table_BivData_4MM, c("strain","sex"), summarise,
-
-                    mean_1CO_nrm.pos = mean(norm.first.foci.pos, na.rm = TRUE), 
-                  n1CO = sum(n1CO )
-
-)
-
-```
-
-```{r norm.CO.pos_longBiv, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#MELT_Curated_BivData
-
-#these isn't long.biv data
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-real.long.bivData
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-MELT_long_biv <- add_sex(MELT_long_biv)
-MELT_long_biv <- add_strain(MELT_long_biv)
-MELT_long_biv <- add_subsp(MELT_long_biv)
-MELT_long_biv <- add_category(MELT_long_biv)
-
-long.biv_Dom.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Dom") & (MELT_long_biv$sex == "male"), ]
-
-Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Dom.male_MELT_long
-
-long.biv_Musc.male <- MELT_long_biv[ (MELT_long_biv$subsp == "Musc") & (MELT_long_biv$sex == "male"), ]
-
-Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
-Musc_male_MELT_long
-
-#there are v few observations in this dataset
-```
-
-```{r Q1.log.reg.CO1_pos, echo=FALSE, eval=FALSE}
-
-#LOGISTIC REGRESSION for FOCI POSITION
-
-#mouse av, norm.first foci is from 1CO bivalents
-log.reg_1CO_pos_mouse.av <- glm(Rec.group ~ norm.first.foci.pos, 
-                            data=mouse.avs_4MM[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male"),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_mouse.av)
-
-#single biv level
-log.reg_1CO_pos_biv <- glm(Rec.group ~ PER_Foci_1, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
-summary(log.reg_1CO_pos_biv)
 
@@
 
-```{r double.tri.rewrite, echo=FALSE, eval=FALSE}
-#good function, saves to file, can't reproduce in rmd
-
+```{r double.tri.rewrite, fig.width = 6.5, fig.height=5}
+
+#good function, saves to file, can't reproduce in rmd ..width=650, height=500
 #big loop for automating the IFD triangles
@@
   mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] + 10)
+  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
   x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
@@
 x.dist <- .5
+RP <- list()
 
 for(i in 1:(length(df_strain_list))){
-  print(i)
+  #print(i)
   current.strain <- as.character(df_strain_list[[i]]$strain[1])
 
-  #declare file
-  png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)
+  #declare file for saving png
+  #png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)
+  plot.new()
 
@@
   op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
+
   top.triangle2(my.y.top, my.x.top, my.data)
+
   bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
+
+  #try recordPlot(),
+
+  #x.plot.list[[i]] <- function(){top.triangle2(my.y.top, my.x.top, my.data)}
+  #x.plot.list.b[[i]] <- function(){top.triangle2(my.y.top, my.x.top, my.data), bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data)}
+  # p <- recordPlot()
+  #grid.echo()
+  #a <- grid.grab()
+
   par(op)
+  RP[[i]] <- recordPlot()
   dev.off()
 }
+
+#for(ee in 1:7){
+#  #print(ee)
+#  RP[[ee]]
+#  plot.new()
+#}
 
@@
 
-```
-
-```{r Q1.IFD_gamma, echo=FALSE, message=FALSE}
-library(EnvStats)
-#cleanCurated_BivData
-
-#all.female.multi <- (cleanCurated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count > 1) %>%
-#                       filter( !is.na( IFD1_ABS) ))  #$IFD1_ABS
-
-#this MELT BIVDATA should be good to go
-MELT.IFD.BIVDATA <- melt(cleanCurated_BivData[,c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
-                            "Obj.ID","hand.foci.count",
-                            "mouse","category","strain", "sex", "Rec.group",
-                            "IFD1_ABS","IFD2_ABS","IFD3_ABS")],
-
-             id=c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
-                            "Obj.ID","hand.foci.count",
-                            "mouse","category","strain", "sex", "Rec.group"))
-#rename the cols and clean up
-MELT.IFD.BIVDATA <- MELT.IFD.BIVDATA[(!is.na(MELT.IFD.BIVDATA$value)),]
-
-colnames(MELT.IFD.BIVDATA) <- c("fileName", "chromosomeLength","centromere_ABS_Position", "centromere_PER_Position",
-                            "Obj.ID","hand.foci.count",
-                            "mouse","category","strain", "sex", "Rec.group", 
-                               "whichIFD", "IFDval")
-
-MELT.IFD.BIVDATA$Nrm.IFD <- MELT.IFD.BIVDATA$IFDval / MELT.IFD.BIVDATA$chromosomeLength
-
-#START isolating all the types of IFDs
-
-#all.female.multi$mycol <- apply(all.female.multi[,c('IFD1_ABS','IFD2_ABS','IFD3_ABS')], 1, function(x) x[!is.na(x)][1])
-
-all.female.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "female")#1697
-
-all.male.IFDs <- MELT.IFD.BIVDATA %>% filter(sex == "male")#1392
-
-#shape and scale parameter
-
-#gamma fits
-FULL.female.gamma <- egamma( all.female.IFDs$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.8
-
-#gamma fits
-FULL.male.gamma <- egamma( all.male.IFDs$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#shape  5.4
-#male, stronger interference
-
-WSB.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.1, 3.73 (raw, PER) (n=252)
-
-WSB.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.6, 5.10 (raw, PER) (n=186)
-#try the short and long bivalents dataset
-
-WSB.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3, 3.237 (raw, PER) (n=60)
-
-#gamma fits
-WSB.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.5,  5.1 (raw, PER) (n= 109)
-
-WSB.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.9, 3.9 (raw, PER)  (n=107)
-#there are 0 3CO and 2  4CO bivalents -- this causes a greater difference in the PER based value
-
-```
-
-```{r Q1.gamma.IFD, echo=FALSE, message=FALSE}
-female.gamma <- egamma( (all.female.IFDs )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-f2CO.female.gamma <- egamma( (all.female.IFDs  %>% filter(hand.foci.count == 2)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-f3CO.female.gamma <- egamma( (all.female.IFDs %>%  filter(hand.foci.count == 3)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-f4CO.female.gamma <- egamma( (all.female.IFDs %>%  filter(hand.foci.count == 4)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-PWD.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.8, 4.844 (raw, nrm)  (n = 357)
-PWD.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3.8, 6.746 (raw, nrm) (n = 285)
-PWD.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#3, 3.628 (n = 72)
-
-KAZ.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-
-KAZ.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-
-KAZ.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-#####
-SKIVE.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-##
-MSM.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-MSM.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-######
-G.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-G.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-######
-LEW.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-LEW.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-######
-WSB.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-WSB.2CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-WSB.3CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-WSB.4CO.female.gamma <- egamma( (all.female.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 4)  )$IFDval, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-```
-
-```{r Q1.male.IFD_gamma, echo=FALSE, message=FALSE}
-#gamma fits
-male.gamma <- egamma( (all.male.IFDs )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-m2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-m3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE,         ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-#gamma fits
-PWD.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#8.1, 11.74 (n=318) (lots of variation when I apply sample size of 18)
-
-PWD.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#10.3, 16.962, (n=300)
-
-PWD.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "PWD") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#2.9, 3.4, (n=18)
-
-####
-KAZ.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-KAZ.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-KAZ.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-SKIVE.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-SKIVE.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "SKIVE") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-MSM.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE,     ci.type = "two-sided", ci.method = "normal.approx",
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MSM.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MSM") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-MOLF.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MOLF.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-MOLF.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "MOLF") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-G.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-G.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "G") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-LEW.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-LEW.3CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "LEW") %>% filter(hand.foci.count == 3)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-
-###
-WSB.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB")  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
-#
-WSB.2CO.male.gamma <- egamma( (all.male.IFDs %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2)  )$Nrm.IFD, method = "mle", ci = TRUE, 
-                            ci.type = "two-sided", ci.method = "normal.approx", 
-                            normal.approx.transform = "kulkarni.powar", conf.level = 0.95)
 ```
@@
 
+```{r Q1.single.biv.table, echo=FALSE}
+
+#strain.BivData.Table
+
+```
+
 ```{r Q1.IFD.plot, echo=FALSE}
@@
 #pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
 

### 2020-04-28 doc/Results_v2.rmd
IFD figure drafts knited into html results draft

@@
 library("xlsx")
+library(png)
+#library(tidyverse)
+library(png)
+library(grid)
+library(gridBase)
+library(gridExtra)
+library(dplyr)
+library(cowplot)
 
@@
 
-```{r triangle.IFD.auto.loop.ifd, echo=FALSE, eval=FALSE}
-
-#Top is female
-top.triangle2 <- function() {
-  #top-y position is male Pos2
-  plot(  (Curated_BivData %>% filter(strain == "WSB"  ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
-    #top-x position is female Pos1     
-    (Curated_BivData %>% filter(strain == "WSB"   ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,
-
-       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2) ), 
-       col =  colors_of_strains["WSB"],
-       axes=FALSE, ylab='', xlab="", 
-       main= "IFD Distribution",  xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
-
-  mtext("WSB", 3, 5, font=2, cex=1.3, adj=.5)
-  mtext("Female Position.2", 2, .75)
-  mtext("Female Position.1", 3, 2)
-  axis(side=2, las=1, pos=0)
-  axis(side=3, las=1, pos=1)
-  lines(0:1, 0:1)
-  }
-
-bottom.triangle2 <- function() {
-    #bottom-y position is male Pos1
-  points(  (Curated_BivData %>% filter(strain == "WSB") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
-    #bottom-x position is male Pos2     
-    (Curated_BivData %>% filter(strain == "WSB") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,
-
-         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == "WSB") %>% filter(hand.foci.count == 2) ),
-
-    col =  colors_of_strains["WSB"], xpd=TRUE)
-  mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
-  mtext("Male Position.1", 4, 3)#, padj=par()$usr[1]
-  x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
-  axis(side=1, las=1, pos=0, at=x.at, labels=F, xpd=TRUE)
-  mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1) # + x.dist
-  lines(0:1, 0:1, xpd=TRUE)# + x.dist
-  }
-
-#par(mar=c(1,1,1,1))
-op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
-plot(top.triangle2(), bottom.triangle2() ) #not working
-par(op)
-dev.off()
-
-```
-
-```{r double.tri.rewrite, fig.width = 6.5, fig.height=5}
+```{r double.tri.rewrite, echo=FALSE}
 
@@
        col =  colors_of_strains[current.strain])
+
   mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
@@
   mtext("Position.1.Female", 3, 2)
+
   axis(side=2, las=1, pos=0)
@@
 
-x.dist <- .5
+x.dist <- .25
 RP <- list()
@@
   #declare file for saving png
-  #png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)
+  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
+
   plot.new()
@@
 
-  op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
+  #adjusting this mar lowers the margins
+  op <- par(mar=c(1.5, 2, 4, 6) + 0.1, oma=c(2, 0, 0, 2))
 
@@
   bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
-
-  #try recordPlot(),
-
-  #x.plot.list[[i]] <- function(){top.triangle2(my.y.top, my.x.top, my.data)}
-  #x.plot.list.b[[i]] <- function(){top.triangle2(my.y.top, my.x.top, my.data), bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data)}
-  # p <- recordPlot()
-  #grid.echo()
-  #a <- grid.grab()
 
@@
 
-#for(ee in 1:7){
-#  #print(ee)
-#  RP[[ee]]
-#  plot.new()
-#}
+#testing <- plot_grid( rasterGrob(RP[[1]]), rasterGrob(RP[[2]]), rasterGrob(RP[[3]]),
+#                   rasterGrob(RP[[4]]), rasterGrob(RP[[5]]), rasterGrob(RP[[6]]),
+#                   rasterGrob(RP[[7]]),                     labels="AUTO")
+#ggsave(filename = "~./MLH1repo/doc/figureFiles/test.triangle_all.png", plot = testing, 
+#       width = 10, height = 10)
 
@@
 
-library(png)
+plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
+plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
+plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
+
+plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
+plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
+plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
+
+plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
+
+grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
+             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
+             rasterGrob(plot7.MSM), 
+             ncol=3)
 
 #load all the files / arrange the plots from above 
-WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
-
-#make empty DF for ggplot object
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-#make an empty ggplot thing, then background ontop
-WSB.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(WSB.triangle.png) + theme(axis.title.x=element_blank(),
- axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-#show wsb as example
-WSB.triangle.ggplot
-
-#library("cowplot")
-#fig1  <- ggdraw() +
-#  draw_plot(WSB.triangle.ggplot, x=0, y=0, width=.45, height=.45) +
-#  draw_plot(G.triangle.ggplot,   x=0, y=.4, width=.45, height=.45) +
-#  draw_plot(LEW.triangle.ggplot, x=0, y=.6, width=.45, height=.45)
-
-#ggsave("~./WildMiceMadison/results/Figure1.png")
+#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
+
 ```
@@
 
-```{r Q1.IFD.plot, echo=FALSE}
-
-#make interference plots of IFD, position p1 by p2
-
-```
-
 ```{r Comp_Ratios, echo=FALSE}
@@
 
+```{r Q2.double.tri.rewrite, echo=FALSE}
+#choose the categories to compare 
+#all male strains
+#just bottom triangles?
+
+#good function, saves to file, can't reproduce in rmd ..width=650, height=500
+#big loop for automating the IFD triangles
+Q2.males.IFD <- c("F1", "other")
+Q2.auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% Q2.males.IFD, ]
+
+Q2.df_strain_list <- split(Q2.auto_curate_DF, (Q2.auto_curate_DF$strain))
+Q2.df_strain_list <- Q2.df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
+
+#this works
+Q2.bottom.triangle2 <- function(my.y, my.x, my.data){
+  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
+          col =  colors_of_strains[current.strain])
+
+  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]))
+  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +4)
+  x.at <- axisTicks(par()$usr[1:2], 0)
+  axis(side=1, las=1, pos=0, at=x.at, 
+       labels=F, xpd=TRUE)
+  mtext(seq(0, 1, .2), 1, 0, at=x.at)
+  axis(4, las=1, pos=1)
+  lines(0:1, 0:1, xpd=TRUE)
+}
+
+#figure out to move up the bottom axis text
+
+x.dist <- .25
+RP <- list()
+
+for(i in 1:(length(Q2.df_strain_list))){
+  #print(i)
+  current.strain <- as.character(Q2.df_strain_list[[i]]$strain[1])
+
+  #declare file for saving png
+  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
+  plot.new()
+
+  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
+  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
+  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  #adjusting this mar lowers the margins
+  op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
+
+  Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
+
+  par(op)
+  RP[[i]] <- recordPlot()
+  dev.off()
+}
+
+```
+
+```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE}
+
+Q2.plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_WSB.png')
+Q2.plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
+Q2.plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_G.png')
+
+Q2.plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
+Q2.plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
+Q2.plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
+
+Q2.plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
+Q2.plot8.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
+
+grid.arrange(rasterGrob(Q2.plot1.WSB), rasterGrob(Q2.plot2.LEW), rasterGrob(Q2.plot3.G),
+             rasterGrob(Q2.plot4.KAZ), rasterGrob(Q2.plot5.SKIVE), rasterGrob(Q2.plot6.PWD),
+             rasterGrob(Q2.plot7.MSM), rasterGrob(Q2.plot8.MOLF), 
+             ncol=3)
+
+```
+
 Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).

### 2020-05-01 doc/Results_v2.rmd
updating the figures

@@
 title: "MLH1 Results Outline"
+bibliography: refs_Main_Report.bib
 date: 'null'
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
   html_document:
@@
     toc_float: yes
-  word_document:
-    toc: yes
-    toc_depth: '5'
-bibliography: refs_Main_Report.bib
+always_allow_html: yes
 ---
@@
 
-#setwd("~./MLH1repo/doc/")
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
-#setwd("C:/Users/April/Documents/MLH1repo")
-#load(file="data/MLH1/MLH1_data_setup.RData")
-
-#load main data file
-#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
-#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
-
-#load(file="data/MetaData.RData") #this was replacing most updated data file
-
 #load functions
@@
 #why do I add in a different table?
-
 #this move this to the setup file
@@
 
-```{r Fig1, echo=FALSE, warning=FALSE}
-
-#might need to nudge the X values
+```{r Fig1.top, echo=FALSE, warning=FALSE}
 
@@
   panel.background = element_blank(), axis.line=element_line(colour = "black"),
+  axis.title.y=element_blank(),
   axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
@@
 
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("Female mouse av")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
   theme(legend.position="none",
-      axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
+
+        axis.title.y =element_blank(), 
+      axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
 
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus") 
-#just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-#just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-#just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-#just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-#just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#table( droplevels(just.male_mouse.av$strain) )
-#there are cast strains too
-#Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-#fomo.male_mouse.av
-
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
-  theme(legend.position="none", axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), 
+male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+
+   theme(legend.position="none", axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),  axis.title.y = element_blank(), 
   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
 
-#+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-
-```{r three.plots, echo=FALSE, warning=FALSE}
-#merge the 3 plots
-#try patchwork
-
-#Fig1_v1_strain + female.specific.plot + male.plot
-
-#Fig1_v1_strain + female.specific.plot - male.plot + plot_layout(ncol=1)
-Fig1_v1_strain / (female.specific.plot + male.plot )
+```
+
+```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7}
+
+Figure1.main.plot <- Fig1_v1_strain / (female.specific.plot + male.plot )
+
+Figure1.main.plot + plot_annotation(
+  title = '',
+  subtitle = '',
+  caption = 'caption'
+) + plot_annotation(tag_levels = 'A')
 
@@
 molossinus*.
-Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )` mice were quantified from an average of `r mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes).
+Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ_target)$mouse )` mice were quantified from an average of `r mean( (AP_mouse_table.HQ_target %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
 
@@
 
-#reorder the df.strain.table
-#species, subspecies, strain
-
-#code for reording column names...
-#DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", #"ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
-
-#write.xlsx(df.strain.table, file = "Results_Table.xlsx", sheetName = "MLH1.results", append = T, row.names = F)
-
 kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
@@
 We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r MLH1.MMlmer_results$rand.pval` -- `r 1/1000000` someSmallNumber), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
+(random effect p < <r MLH1.MMlmer_results\$rand.pval> `r 1/1000000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
 
 After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p= `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and domesticus^G^ (`r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (`r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (`r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
+A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
+and domesticus^G^ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
+In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
 
@@
 Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
-To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < `r 1/1000000` -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear models (M2) (p = `r MLH1var_M2_sum$coefficients[,4][9]`) and M3 (p = `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`).
+To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count per cell as the dependent variable.
+Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < `r 1/1000000`) and general linear models (M2) (p = `r MLH1var_M2_sum$coefficients[,4][9]`) and M3 (p = `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`).
 
@@
 Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
-The results are similar: sex is the strongest effect (M1 p < `r 1/1000000` `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
+The results are similar: sex is the strongest effect (M1 p < `r 1/1000000` <round(LmerMLH1_M1_Q12_VAR_results\$sex.results, 5)>; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
 When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
 These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
+
+**(sex specific results - no efficient strain effect for males. In females LEW - but this is due to cell quality difference, what about quality score and cell number within the models?**
 
@@
 
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
+```{r DMC1.boxplot.plot, echo=FALSE, warning=FALSE}
 
@@
 #change order
-
 DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
@@
 DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice\nZygotene spermatogcytes")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.boxplot
+  geom_boxplot(alpha=.5) +geom_jitter() +
+
+    labs(y = "DMC1 count per Cell")+
+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.x = element_blank())
+
+```
+
+```{r Main.plot.DMC1, echo=FALSE, fig.width=6.5, fig.height=5}
+
+png.DMC1.early.cell <- readPNG('~/MLH1repo/doc/figureFiles/cell.examples/DMC1.early/good.early.cell.png')
+png.DMC1.late.cell <- readPNG('~/MLH1repo/doc/figureFiles/cell.examples/DMC1.late/good.late.cell.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+#EMPTY GGPLOT
+Early.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(png.DMC1.early.cell) + theme(axis.title.x=element_blank(),
+                         axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+Late.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(png.DMC1.late.cell) + theme(axis.title.x=element_blank(),
+                         axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+three.DMC1.plot <- { Early.cell.obj + Late.cell.obj  } / DMC1.boxplot + plot_layout(widths = 1)
+
+#+ plot_layout(ncol = 2, widths = c(.65, .65), heights = c(.65, .65))
+#plot_layout(nrow = 2, widths = c(1), heights = c(.45) )
+
+#plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or #labels="AUTO"
+
+three.DMC1.plot + plot_annotation(
+  title = '',
+  subtitle = '',
+  caption = 'caption'
+) + plot_annotation(tag_levels = 'A')
+
 ```
@@
 In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1.mean.table %>%  filter(stage == "Early") )$ncells)` early zygotene-stage and `r sum( (DMC1.mean.table %>%  filter(stage == "Late") )$ncells)` late zygotene-stage spermatocytes from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
-
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p<  -- `r 1/1000000` -- `r ttest.HighLow.L.pval`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = 0.66 --- `r ttest.HighLow.Z.pval`).
+DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
+
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p < `r 1/1000000` <ttest.HighLow.L.pval>).
+In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = `r ttest.HighLow.Z.pval`).
 
@@
 The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
+(connection to barier)
 
 This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
-[@cole2012] suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
-This result, combined with the big difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
+@cole2012 suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
+
+This result, combined with the large difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
 
@@
 (this study only examined one stage of early prophase for DMC1 counts)>
-
-<bairer-- DSB number is not a primary driver of variation in CO number -- these authors suggest that chromatin organizaiton (loop-axis) structure is the primary determinant of CO variation.>
 
@@
 
-```{r Chrm.props.plot2, echo=FALSE}
+```{r Chrm.props.plot2, echo=FALSE, fig.width=8}
 
@@
 
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
+chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(legend.position="none", axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )
 #chrm.class.barplot
@@
 
-From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` bivalent objects were isolated by the image analysis software.
-After a human curation step ([@peterson2019]), `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )`  single-bivalent observations remained.
+From the total set of cell images, `r formatC(sum( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ) ), , format="d", big.mark="," )` bivalent objects were isolated by the image analysis software.
+After a human curation step ([@peterson2019]), `r formatC(sum(  length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ) ), , format="d", big.mark=",")`  single-bivalent observations remained.
 We assume that the isolation of bivalents within cells is unbiased.
-<might be the only section to leave in Results>
-
-Given the large number of single-bivalent observations, we assume that each of the datasets are equally representative of general patterns. <Bret says this statment should be more clear>
-
-An additional challenge of the MLH1 framework is that the identities of individual autosomes and the XX in females cannot be easily obtained (the male XY is distinct).
-<Bret suggests moving to Methods>
-
-<figure>
-
-Ninety-six percent of single bivalents in our pooled dataset (n = `r total.biv.all`) have either one or two crossovers (Figure X).
+
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure X).
 The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
 low recombining strains (Figure X).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in molossinus^MSM^.
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in molossinus^MSM^.
 
@@
 Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#CLEAN UP THE FULL DATA SET
 
@@
 #nested boxplots, pooled Bivalents
-
 Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
 
-Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
-  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
 ```
@@
 
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
-Q1.SC.short.mouse_scatter_boxplot
-
-#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
-skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-#skiveQ1.SC.short.mouse_scatter_boxplot
-#there are 2 lowish female means,
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  theme(legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=9), axis.text.y = element_text(size=9))+labs(y = "Mouse Mean Short SC Length")
 
@@
 ```{r Q1SC_short.biv_dif, echo=FALSE}
-
-#calq the different in short.biv means
+#calq the different in short.biv means across sexes
 
@@
 
-```{r Q1.tot.SC_show, echo=FALSE, warning=FALSE}
+```{r Q1.tot.SC.plot, echo=FALSE, warning=FALSE}
 # fig.width = 4, fig.height=2.5
@@
 
-#COPY THIS CODE TO BELOW
-#Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-#Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
-
 #change this plot to look like the short bivalent plot -- boxplots + mouse points
@@
 
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. total SC area") + 
-  #change the white themes
-  theme(legend.position="none")
-
-Tot.SC_mouse.av_ALL
-
-totsc.musc <- Q1.totSC.mouse.av.table %>% filter( subsp == "Musc" )
-totsc_strain.av.msc <- totSC.strain.av.table %>% filter( subsp == "Musc" )
-
-Tot.SC_mouse.av_Musc <- ggplot(data= totsc.musc, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.msc, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ggtitle("Mouse av. total SC area")+  ylim(c(1000,3000 ))
-
-totsc.mol <- Q1.totSC.mouse.av.table %>% filter( subsp == "Mol" )
-totsc_strain.av.mol <- totSC.strain.av.table %>% filter( subsp == "Mol" )
-
-Tot.SC_mouse.av_Mol <- ggplot(data= totsc.mol, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totsc_strain.av.mol, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. total SC area")
+  theme(axis.title.x=element_blank(), axis.ticks.x=element_blank(),axis.text.x = element_text(size=9), legend.position="none", axis.text.y = element_text(size=9))+  
+  labs(y = "Mouse Mean Short SC Length")+
+  ylim(c(1000,2600 ))+ggtitle("")
+
+```
+
+```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8}
+
+two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(1,1)) }
+
+two.Q1.SC.plots + plot_annotation(
+  title = '',
+  subtitle = '',
+  caption = 'caption'
+) + plot_annotation(tag_levels = 'A')
 
@@
 
-In many mammalian species, the synaptonemal complex (SC) is longer in females.
+In many mammalian species, the synaptonemal complex (SC) is longer in females (@cahoonLibuda2019, @sardell_sex_2020).
 Yet, the majority of these observations come from species with female-biased heterochiasmy.
@@
 
-Short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculus^SKIVE^.
-The female:male ratio ranges across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
-<moved section on total SC to methods>
+We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
+The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculus^SKIVE^, which my be attributed to low sample sizes (p = `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`).
+The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
 
@@
 That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies.
+Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
 
@@
 In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers.
+Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers compared to females.
 
@@
 
-In most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus).
+We used the single crossover bivalent (1CO) normalized position to compared general location of crossovers while controlling for differences in total SC length.
+Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus) -- thus justifying the use of single crossover bivalents -- to discribe this metric / trait.
+
 In all strains, the landscape across 1CO bivalents is significantly different in females and males.
 Normalized foci positions tend to be more central in females `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` and closer to the telomere in males `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )` (t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
+
 Sex is also the most significant effect on focus position in a mixed model (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
@@
 
-```{r double.tri.rewrite, echo=FALSE}
+```{r double.IFD.triangle.plots, echo=FALSE}
 
@@
 
-x.dist <- .25
+x.dist <- .1
 RP <- list()
@@
   #adjusting this mar lowers the margins
-  op <- par(mar=c(1.5, 2, 4, 6) + 0.1, oma=c(2, 0, 0, 2))
+  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
 
@@
 }
-
-#testing <- plot_grid( rasterGrob(RP[[1]]), rasterGrob(RP[[2]]), rasterGrob(RP[[3]]),
-#                   rasterGrob(RP[[4]]), rasterGrob(RP[[5]]), rasterGrob(RP[[6]]),
-#                   rasterGrob(RP[[7]]),                     labels="AUTO")
-#ggsave(filename = "~./MLH1repo/doc/figureFiles/test.triangle_all.png", plot = testing, 
-#       width = 10, height = 10)
 
@@
 
-```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE, warning=FALSE, message=FALSE}
-#long SC Long_biv_mouse.avs_4MM
-#real.long.bivData
-#Only use 2CO bivalents
-
-real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
-real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]
-
-long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-
-long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
-```
-
 ```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
@@
 ```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
+
+#these are the scatter plots
+
 nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
@@
 ###look at why KAZ males are weird
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
 
@@
 #the difference the the Nrm IFD cut off is striking with these plots
-
 nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
 
-#this might
-
 nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
@@
 
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none")
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
 
@@
 There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = 0.07 -- `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain (t-test without domesticus^G^; p = 0.27 -- `r Q1.IFD.raw_t.test_G$p.value`).
+A marginally significant difference between the sexes (t-test; p = `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain, *domesticus^G^* (t-test without domesticus^G^; p = `r Q1.IFD.raw_t.test_G$p.value`).
 This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
-In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < 0.02 -- `r Q1.IFD.PER_FULL_t.test$p.value`), with only musculus^KAZ^ showing no difference (t-test; p = 0.33 --  `r Q1.IFD.PER_t.test_KAZ$p.value`).
+
+In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < `r Q1.IFD.PER_FULL_t.test$p.value`),
+
+with only musculus^KAZ^ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
 Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1 - LRT **random**: p = 6.7410^{-14} -- `r Q1.IFD.PER_M1_lmer.results$sex.results`  glm *M2* `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`  *M3* `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1 - LRT **random**: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`  glm *M2* `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`  *M3* `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
 When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in chromatin compaction (SC length) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-While the comparisons of the IFD^norm^ metrics can reveal more general recombination landscape patterns while controlling for the underlying differences in chromatin compaction and SC length.
-
-<moved a section on relationship of linkage maps to Discussion (or Methods)>
+In summary, controlling for differences in SC length (chromatin compaction) using IFD^norm^ indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
+
+**address the difference is pattern signal for IFD triangles and boxplots** (make boxplots?)
 
@@
 
-The next section is meant to focusNext, we used the contrast between high-recombination strains (musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-on the greater aspect of variation in mean MLH1 counts per cell the high recombining strains ( musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) to the low recombing strains.
+Next, we used the contrast between males in the high-recombination strains (musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
 The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first creteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects in the same models which used mean MLH1 count per cell.
+In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M4.2).
 
 Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-**SC length is expected to be positively associated with genome-wide recombination rate because of the loop-axis strucutre** (zickler Kleckner 1999, Merier 2015?).
-Crossover interference strength is expected to be negatively **associated with genome-wide recombination rate since interference is defined as the non-random spacing between crossovers on the same chromosome**.
+**SC length is expected to be positively associated with genome-wide recombination rate because of the loop-axis structure** (zickler Kleckner 1999, Merier 2015?).
+Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome (on chromosomes of  a finite length, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
 
@@
 
-#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, mouse) %>%
-#  mutate(file.name = factor(file.name))
-
 mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-mouse.av_totsc_male
-
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE}
+  facet_wrap(Rec.group~subsp, scales="free")+
+  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
+
+```
+
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8}
 
 #make plot of mouse averages for MLH1 and total SC
-
 bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
@@
 X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-
-X.by.texas
+  scale_color_manual(values=colors_of_strains)+theme(legend.position="none")+labs(x="Mouse Mean Total SC", y="Mouse Mean MLH1 count")
+
+plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
 
@@
 ```{r log.reg_SC.length, echo=FALSE, include=FALSE}
-#REMOVE THIS COMPARES FROM FULL BIV DATA
-
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
 
@@
 
-#table( droplevels(fomo.male_mouse.av$strain) )
-#there are cast strains too
-
-#Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-
-#musc.male.plot <- ggplot( ( male_mouse.av %>% filter(species == "M.musculus") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
-#musc.male.plot
-
-# + theme(axis.title.x=element_blank(),
-#        axis.text.x=element_blank(),
-#        axis.ticks.x=element_blank()) +
-#ylim( c((min(MLH1_data$ nMLH1.foci)-3) ), (max(MLH1_data$ nMLH1.foci)+3)) +
-
 #make the Mol and
@@
 Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Mol.male.plot
 
@@
 Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-Dom.male.plot
-```
-
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`;
-<no mean.MLH1>
-
-p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
+
+```
+
+Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
 Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
-**Bret's note on positive correlation, 'A correlation across mice, across strains, or both?** While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+**Bret's note on positive correlation, 'A correlation across mice, across strains, or both?**
+
+While high-recombination strains have significantly greater mouse mean total SC area in the total dataset (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p =  `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
 **the values above are diferent between editions**
 
-Additionally, the means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value`; long: p = 0.18 -- `r Q2long.biv_t.test_High.Low$p.value`).
+Additionally, the mouse means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
 
 **the values between versions are slightly different**
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
+In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, molossinus p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
 In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
@@
 
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos All bivalents")
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
+
 Q2.show.nrmF1
@@
 
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = 0.24 -- `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = 0.41 -- `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = 0.07 -- `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
+The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
 **While there are significant strain effects for domesticus^WSB^ and molossinus^MOLF^ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.* <bret asks if the bolded section is correct.
@@
 
-```{r Q2.IFD.glm.test, echo=FALSE}
+```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE}
 
@@
 
-```{r Q2.double.tri.rewrite, echo=FALSE}
-#choose the categories to compare 
-#all male strains
-#just bottom triangles?
+```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE}
 
@@
 Q2.df_strain_list <- split(Q2.auto_curate_DF, (Q2.auto_curate_DF$strain))
-Q2.df_strain_list <- Q2.df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
 
@@
 
-  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]))
+  mtext("Position.2.Male", 1, 1.5, at= mean(par()$usr[1:2])) #at should be changing the position
+  #above is just changing the text not the numbers
+  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)
+
+  #second number - on which MARgin line, starting at 0 counting outwards.
+  #where text should be,  at = 0, 1 (bad)
+
   mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +4)
-  x.at <- axisTicks(par()$usr[1:2], 0)
+
+  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
   axis(side=1, las=1, pos=0, at=x.at, 
        labels=F, xpd=TRUE)
-  mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1)
-  lines(0:1, 0:1, xpd=TRUE)
+#las, orientation, at location of line
+  #labels automatic labels
+  #pos
+  #xpd - clipping
+
+   mtext(seq(0, 1, .2), 1, 1, at=x.at)
+  #second / third val negative moves the valu up
+
+    axis(4, las=1, pos=1) #this is the y axis
+ #   axis(4, las=1, pos=1 + x.dist)
+  #this is calling axis twice, not sure why
+
+    lines(0:1, 0:1, xpd=TRUE) #diagonal line
 }
-
-#figure out to move up the bottom axis text
 
@@
 
+  #par(mar = c(2, 2, 2, 2))
+  #dev.new(width = 8, height = 4, unit = "in")
   #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
+  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
   plot.new()
@@
 
-  #adjusting this mar lowers the margins
-  op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
-
+  #adjusting this mar lowers the margins (the edges)
+  op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))
+
+  #   op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
   Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
@@
 
-```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE}
+```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6}
 
@@
 
-Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
-This pattern is confirmed by separate comparisons within _musculus_ (t-test; IFD^norm^: p =2.0410^{-5} -- `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = 1.9410^{-4} -- `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
+```{r double.IFD.triangle_copy, echo=FALSE, eval=FALSE}
+
+#good function, saves to file, can't reproduce in rmd ..width=650, height=500
+#big loop for automating the IFD triangles
+only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
+auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
+
+df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
+df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
+#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
+
+#this works
+top.triangle2 <- function(my.y, my.x, my.data) {
+  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
+       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
+       col =  colors_of_strains[current.strain])
+
+  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
+  mtext("Position.2.Female", 2, .75)
+  mtext("Position.1.Female", 3, 2)
+
+  axis(side=2, las=1, pos=0)
+  axis(side=3, las=1, pos=1)
+  lines(0:1, 0:1)
+}
+
+bottom.triangle2 <- function(my.y, my.x, my.data){
+  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
+          col =  colors_of_strains[current.strain])
+
+  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
+  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
+  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
+  axis(side=1, las=1, pos=0, at=x.at, 
+       labels=F, xpd=TRUE)
+  mtext(seq(0, 1, .2), 1, 0, at=x.at)
+  axis(4, las=1, pos=1 + x.dist)
+  lines(0:1 + x.dist, 0:1, xpd=TRUE)
+}
+
+x.dist <- .1
+RP <- list()
+
+for(i in 1:(length(df_strain_list))){
+  #print(i)
+  current.strain <- as.character(df_strain_list[[i]]$strain[1])
+
+  #declare file for saving png
+  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
+
+  plot.new()
+
+  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
+
+  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
+
+  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
+
+  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  #adjusting this mar lowers the margins
+  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
+
+  top.triangle2(my.y.top, my.x.top, my.data)
+
+  bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
+
+  par(op)
+  RP[[i]] <- recordPlot()
+  dev.off()
+}
+
+```
+
+```{r Q1.IFD_load.triangles.copy, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Example of Sex differences in IFD distributions", eval=FALSE}
+
+plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
+plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
+plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
+
+plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
+plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
+plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
+
+plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
+
+grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
+             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
+             rasterGrob(plot7.MSM), 
+             ncol=3)
+
+#load all the files / arrange the plots from above 
+#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
+
+```
+
+Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS$p.value`).
+This pattern is confirmed by separate comparisons within _musculus_ (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
 <Bret's comment, these p values are not very significant.> Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
@@
 That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% --
-
-in low-recombination strains.
+We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
 The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (domesticus^G^) to 16% (musculus^KAZ^) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculus^SKIVE^, molossinus^MSM^, and musculus^PWD^, respectively).
@@
 However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
-At least on 2CO bivalents, crossovers are spaced further apart when the genome-wide recombination rate is higher.
-
-# Q2 Summary
-
-**move main themes to conclusion**
-
-Our results show that the greater crossover interference is the strongest single bivalent-based predictor for the observed rapid evolution of mean MLH1 foci per cell.
-While these results do not conform initial predictions for how a recombination landscapes would accommodate more crossovers, the increased strength of interference aligns with our results on sex differences.
-The typical recombination landscapes for males and females results in divergence in the proportion of linked sites along chromosomes which segregate together.
-The stronger interference of the 2CO bivalents in the high recombining strains accentuates this effect.
-The measures of DSB and some comparisons of SC length between high and low recombining strains suggest that the SC length have evolved to be longer in high recombining strains, however this evolution of SC length is partially decoupled from the number of crossovers since similar amounts of SC length evolution are seen in low recombining strains.
-
-<Bret: Are you sure this summary is necessary? Maybe cover this material in the Discussion instead?>
 

### 2020-05-01 doc/Results_v2.rmd
making triangle plots, and icon drafts

@@
 output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
   html_document:
@@
     toc_float: yes
+  word_document:
+    toc: yes
+    toc_depth: '5'
 always_allow_html: yes
@@
 
-two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(1,1)) }
+two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(1.5,1.5)) }
 

### 2020-05-01 src/CommonFunc_MLH1repo.R
added make small legend function

@@ -727,3 +727,13 @@ multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
     }
   }
 }
+
+#from https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.25, textSize = 7, spaceLegend = 0.3) {
+  myPlot +
+    guides(shape = guide_legend(override.aes = list(size = pointSize)),
+           color = guide_legend(override.aes = list(size = pointSize))) +
+    theme(legend.title = element_text(size = textSize), 
+          legend.text  = element_text(size = textSize),
+          legend.key.size = unit(spaceLegend, "lines"))
+}

### 2020-05-01 doc/Results_v2.rmd
notes for editing results

@@
 
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+
+    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1)+
 
@@
   axis.title.y=element_blank(),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
+  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
 #remove y labfor male
@@
    #segments
-#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
+  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
   #Musc
-#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)
   #mol
@@
 
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
   theme(legend.position="none",
@@
   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))+geom_hline(yintercept = 20)
 
@@
 
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+
+male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
 
@@
   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))
-
-```
-
-```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7}
+  panel.background = element_blank(), axis.line=element_line(colour = "black"))+geom_hline(yintercept = 19)
+
+#horizontal line = 20
+
+```
+
+```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
+B) Female specific MLH1 count distributions.
+C)	Male specific MLH1 counts per cell.'}
 
@@
   subtitle = '',
-  caption = 'caption'
+  caption = ''
 ) + plot_annotation(tag_levels = 'A')
@@
 
-```{r Main.plot.DMC1, echo=FALSE, fig.width=6.5, fig.height=5}
+```{r Main.plot.DMC1, echo=FALSE, fig.width=6.5, fig.height=5, fig.cap="*Figure 2 male DSB estimates* A) Example early zygotene spermatocyte spread, B) Example late zygotene spermatocyte spread.
+Green foci show DMC1.
+B)	Distribution of DMC1 counts per cell  by strains.
+(1 mouse was used for each strain)"}
 
@@
   title = '',
-  subtitle = '',
-  caption = 'caption'
-) + plot_annotation(tag_levels = 'A')
+  subtitle = '') + plot_annotation(tag_levels = 'A')
 
@@
 
-```{r Chrm.props.plot2, echo=FALSE, fig.width=8}
+```{r Chrm.props.plot2, echo=FALSE, fig.width=8, fig.cap="*Figure 3* Chromosome Class Proportions. proportions calculated from pooled single bivalent data by strain."}
 
@@
 
-chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(legend.position="none", axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )
-#chrm.class.barplot
+chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),
+                             aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(legend.position="top", legend.title = element_blank(),
+                                                                                                           axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
+
 chrm.class.barplot
+kk <- addSmallLegend(chrm.class.barplot)
 
@@
 
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  theme(legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=9), axis.text.y = element_text(size=9))+labs(y = "Mouse Mean Short SC Length")
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  theme(legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=9), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9))+labs(y = "Mouse Mean Short SC Length")
 
@@
 
-  theme(axis.title.x=element_blank(), axis.ticks.x=element_blank(),axis.text.x = element_text(size=9), legend.position="none", axis.text.y = element_text(size=9))+  
-  labs(y = "Mouse Mean Short SC Length")+
+  theme(axis.title.x=element_blank(), axis.ticks.x=element_blank(),axis.text.x = element_text(size=9), legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10)  )+
+  labs(y = "Mouse Mean Total SC Length")+
   ylim(c(1000,2600 ))+ggtitle("")
@@
 
-```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8}
-
-two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(1.5,1.5)) }
+```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8, fig.cap= 'caption'}
+
+two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }
 
@@
 ```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-
 #real.long.bivData
@@
 ```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
-
 #REAL.short_mouse_table
@@
 \nNormalized 1CO Foci pos \nShort Biv")
-```
-
-```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}
-
-MELT_long_biv <- MELT_Curated_BivData[ MELT_Curated_BivData$Obj.ID %in% real.long.bivData$Obj.ID, ]
-
-CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("Percent Foci Position")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
-CO.Pos.boxplots_1CO
-
-```
-
-```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, eval=FALSE, include=FALSE}
-#IGNORE THIS FOR KNOW, THIS
-
-#filter all Curated biv data for 1 strain
-#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs
-
-# the stardell paper sums the number of COs for bins (10)
-
-#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
-#but then the model would be close to logistic
-
-#summing the number of COs per bin, losses the chrom specific data....
-#!! need to use a melted Data frame to capture all nrm.foci obs
-#wsb.ttest<- mouse.avs_4MM %>% filter()
-#start with same chrm classes
-#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci
-
-WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)
-
-PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)
-
-Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig
-
-anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low
-
-anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1
-
-#or AIC(model1,model2)
-
 ```
@@
 
-```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Example of Sex differences in IFD distributions"}
+```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Distribution of normalized foci positions (IFDs) for double crossover bivalents across sexes.
+Female and male patterns are displayed in the top and bottom triangle respectively.
+Each point reflects a single double crossover bivalent, with the normalized positions of foci reflected on each axis."}
 
@@
 
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8}
+mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
+
+```
+
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice.
+A)	 Mean total SC area per mouse across subspecies and recombination groups.
+High- and low-recombining groups are indicated by 1 and 0 respectively.
+(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell"}
 
@@
 
-```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
 ```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
@@
 SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
-#COMPARISONS ACROSS ALL CLASSES
-
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
 
@@
 
-```{r male.show.plotsEXample, echo=FALSE, eval=FALSE}
-male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-male_mouse.av <- add_sex(male_mouse.av)
-male_mouse.av <- add_strain(male_mouse.av)
-male_mouse.av <- add_subsp(male_mouse.av)
-male_mouse.av <- add_species(male_mouse.av)
-
-fomo.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-#make the Mol and
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
-
-Mol.male.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-#Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-Dom.male.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,33))
-
-```
-
 Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
@@
 
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE}
+```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4}
 #Curated_BivData_male
@@
 
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.23 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
 
@@
 
-```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6}
+```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
 
@@
 
-```{r double.IFD.triangle_copy, echo=FALSE, eval=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
-auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
-
-df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
-df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-
-#this works
-top.triangle2 <- function(my.y, my.x, my.data) {
-  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
-       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
-       col =  colors_of_strains[current.strain])
-
-  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
-  mtext("Position.2.Female", 2, .75)
-  mtext("Position.1.Female", 3, 2)
-
-  axis(side=2, las=1, pos=0)
-  axis(side=3, las=1, pos=1)
-  lines(0:1, 0:1)
-}
-
-bottom.triangle2 <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[current.strain])
-
-  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
-  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-  mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1 + x.dist)
-  lines(0:1 + x.dist, 0:1, xpd=TRUE)
-}
-
-x.dist <- .1
-RP <- list()
-
-for(i in 1:(length(df_strain_list))){
-  #print(i)
-  current.strain <- as.character(df_strain_list[[i]]$strain[1])
-
-  #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
-
-  plot.new()
-
-  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-
-  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins
-  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
-
-  top.triangle2(my.y.top, my.x.top, my.data)
-
-  bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-
-```
-
-```{r Q1.IFD_load.triangles.copy, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Example of Sex differences in IFD distributions", eval=FALSE}
-
-plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
-plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
-plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
-
-plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
-plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
-plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
-
-plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
-
-grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
-             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
-             rasterGrob(plot7.MSM), 
-             ncol=3)
-
-#load all the files / arrange the plots from above 
-#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
-
-```
-
 Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS$p.value`).

### 2020-05-04 doc/Discussion.rmd
spell check and fixing typos

@@
 
-# Discussion Draft v1
+# Discussion Draft v2
 
@@
 
+Despite quantifying a similar number of wild derived inbred strains from *M.
+m.
+domesticus* and *M.
+m.
+musculus*, we only observed rapid male specific evolution for the genome wide recombination rate for *M.
+m.
+musculus.* Estimates of effective population size for the three main subspecies of the house mouse found *M.
+m.
+musculus* had the lowest [@geraldes2011].
+Selection becomes less efficient at removing deleterious mutations in small populations outside of mutation-drift balance.
+In such populations, modifiers for increased recombination rates will be selected for as they increase the efficiency of selection.
+
+An additional possibility is that the different rates in recombination rate variation across subspecies is related to history of hybridization across the populations sampled in this panel of inbred strains.
+Theoretical work as connected recombination that rate of purging deleterious introgressed DNA in populations after hybridization (Caballero et al 2019, Schumer 2018).
+The three high-recombining strains identified in this study have connections to hybridization.
+Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the european house mouse hybrid zone.
+Notably *musculus^SKIVE^* comes from a yonger section of the hybrid zone with a nuclear genome from *M.
+m.
+musculus* and a mitochondria genome from *M.
+m.
+domesticus* [@gyllensten1987_SKIVE].
+The subspecies *M.
+m.
+molossinus* is a natural hybrid of *M.
+m.
+musculus* and *M.
+m.
+cataneus* (Geraldes et al 2008, [@geraldes2011]).
+Models for the aggregate recombination metics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019) -- motivating increased study of recombination rates and natural populations.
+4
+<transition to looking at the number and placement of crossovers>
+
 To place our results in the broader context of recombination rate evolution we compare our main resuts to expectionat under three modifier models and two functional and cell physiology models (table X).
@@
 
-The direction of heterochiasmy under the two locus modifier model is dependent on female specific recombination modifier suppresses or increases recombination, if the modifier is linked to the driving locus, and the meiotic stage (reduction (MI) or equatorial division (MII)) when the driver acts (see Table 1 of @brandvain2012scrambling).
-
-Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
-
-A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion [@laneKauppi2019].
-Unstable bivalents on the meiosis I spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle.
-If they fail to attach, the SAC triggers apoptosis.
-Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate or cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019], @subramanian2014].
-These differences in SAC strength have been connected to conserved features of gametogenesis; presence of a centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
-This model nominates conserved differences in in the context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
-
 In our results and more generally in the literature, female biased heterochiasmy is the most common pattern, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
@@
 This insight could be important for animal breeding programs [@battagin2016].
+
+Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
+
+A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion ([@laneKauppi2019]).
+Genetic variants will be selected in a manner to optimize the process of bi-orientation and prevent SAC from delaying cell cycle or triggering apoptosis.
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014]), which are connected to conserved features of gametogenesis; centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
+The more stringent SAC in spermatogenesis will be more effective at removing genetic variants which interfere with bi-orientation compared to females.
+This dynamic can give rise to sex-specific genetic variants which in turn can result in  context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
 
@@
 
+Regardless of whether these hypotheses are true, Our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
+Across all strains, including the high-recombining strains, musc^PWD^, musc^SKIVE^ and mol^MSM^, females consistently have higher between cell variance for crossover count.
+These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005]?, compared to spermatocytes.
+This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
+
 In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
@@
 
-<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>
-
-<Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
+<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
 
@@
 
+Our results extend (the mountain / plethora) or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] - data comprised mostly of highly divergence species / large evolutionary distances.
+In contrast our results demonstrate that this pattern is maintained across much shorter evolutionary distances and even with rapid evolution in the genome-wide recombination rate.
+We argue that this conservation of the male chromosome end bias // recombination landscape is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
+
 As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a conserved trait.
@@
 
-A consequence of the sexual dimorphism in broad scale recombination landscapes is that the male recombination pattern drives deviation from an uniform distribution in any sex averaged recombination landscape [@sardell_sex_2020].
-Elevation at the chromosome ends is a conserved trait across animals and plants and has a consequence of higher genetic diversity near chromosome ends [@haenel2018].
+The higher levels /increased genetic diversity and recombination rates at chromosome ends @haenel2018 - suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020].
+(giving rise to interesting evolutionary predictions).
 
@@
 Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
-This prediction model might fit broader pattern such as in *Arabidopsis thalania* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier paper?
+Sera?) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
 
 As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
-Since more DNA is sequestered into the radial loops per unit of the central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males for mice [@petkov2007, @lynn2002] and humans [@gruhn2013], interference strength measured in base pairs will be stronger compared to female linkage maps.
-The signal for interference is mediated though the chromosome axis [@zicklerKleckner2015], interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-Stronger interference strength in base pairs for male specific linkage maps is indirect evidence that sexual dimorphism in chromatin organization however this has yet to be confirmed in cytological data from both sexes in species beyond mice and humans.
+From empirical data in mice [@petkov2007, @lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
+The signal for interference is thought to be mediated though the SC, which is built on top of the chromosome axis [@zicklerKleckner2015] regardless of the loop size.
+Supporting this is the observation that interference strength measured in micrometers of is conserved between sexes [@petkov2007].
+As a consequence, since more base pairs are sequestered into the radial loops per unit of the linear central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males,  will have a signal of stronger interference when measured in base pairs compared to 'the typical female chromatin configuration.
+Stronger interference in males, has been noted across many sex-specific linkage maps (cite) strengthening the hypothesis that sexual dimorphism in the chromatin organization is widespread, however this has only been confirmed by cytological data in mice and humans.
+(Also large difficulties in observing sperm cells through meiosis).
 
@@
 
+An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (cite, Schumer).
+Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
+**(new, veller paper -- in addition to the total gwRR, the location of crossovers (rbar), should be considered in this dynamic)**
+
 Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
@@
 Two examples of the opposite direction, a positive correlation of interference strength and genome-wide recombination rate, involve descriptions of observations at the single bivalent level; the between lab-raised and wild mice of *Peromyscus leucopus* from [@peterson2019] or large effects on genome wide recombination rates across a short evolutionary differences; the within-sex comparison of two breeds of cattle with different genome wide recombination rates [@ma2015_cattle].
-We propose that there is a difference in resolution and power between these two groups of empirical results with the positive and negative relationship between genome wide recombination rates and interference strength.
 
@@
 The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
-A distinguishing feature of this model from other modifer models is that the number of crossovers per chromosome is kept constant.
-Empiracally, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
+A distinguishing feature of this model from other modifier  models is that the number of crossovers per chromosome is kept constant.
+Empirically, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
 Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
 
-An expected consequence of our observed elevated genome-wide recombination rate is lower LD, which may increase the efficiency of selection and impact the dynamics of introgression.
-Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
+### Future steps
 
-### Future steps
+There are still many un-known aspects regarding heterochiasmy (Lenormand) , these novel results suggest - we hope ppl we follow up .  (fitness - evolutionary dynamics - role in linked selection)
 
@@
 
-Third we encourage the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses (Dapper Paysuer).
+Third we encourage the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
 A goal of such merging of models could be to connect empirical findings across scales.

### 2020-05-04 doc/MatMeth.Rmd
spell check and fixing typos

@@
 The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
-We use abriviated superscript notation for the remainder of the paper (e.g.
+We use abbreviated superscript notation for the remainder of the paper (e.g.
 _musculus^WSB^_).
@@
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used for MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
-Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
+The majority of mice used for MLH1 quantification were from mice between 5 and 12 weeks, while juvenile mice aged between 12 to 15 days were used for DMC1 quantification.
+Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embryonic days and 0 to 48 hours old respectively.
 
@@
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a photoflow wash (Kodak, cite) and stored at -20*C if not stained immediantly.
+The following morning, slides were treated with a Photoflow wash (Kodak, cite) and stored at -20*C if not stained immediately.
 
@@
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.
+Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using Photoshop.
 
@@
 Mean DMC1 foci per cell for a single juvenile mouse from a subset of strains; _musculus^PWD^_, _musculus^MSM^_, _domesticus^WSB^_, and _domesticus^G^_ was quantified.
-Differences in DMC1 counts per cell were tested by ttests.
+Differences in DMC1 counts per cell were tested by t-tests.
 The lateral elements of the synaptonemal complex morphology and CREST/centromere number was used to stage spermatocytes.
-Early zygotene cells were defined as having a larger degree of indiviual centromeres and mostly complete axis signal.
+Early zygotene cells were defined as having a larger degree of individual centromeres and mostly complete axis signal.
 Late zygotene cells were identified as more paired centromeres and homolog axis (but below full synapsis).
@@
 
-The DNA Cross Over algorithm isolates single straightened bivalents from an image, and returns the SC length, and locations of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
+The DNA Cross Over algorithm isolates single straightened bivalents from an image, and returns the SC length, and locations of green and blue signal, reflecting MLH1 foci and centromere signal respectively.
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r round( min(algorithm.stat.category$mean_pass.rate), 3)` (molossinus^MSM^ male) to `r round( max(algorithm.stat.category$mean_pass.rate), 3)` (_musculus^KAZ^_ female).
+In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r round( min(algorithm.stat.category$mean_pass.rate), 3)` (_molossinus^MSM^_ male) to `r round( max(algorithm.stat.category$mean_pass.rate), 3)` (_musculus^KAZ^_ female).
 A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
 From the total set of cell images, `r  formatC(length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` bivalent objects were isolated by the image analysis software.
-After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
+After a human curation step, `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
 The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
@@
 Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A 'long bivalent' data set was calculated as above, using the 4th quartille for SC lengths per cell.
+A 'long bivalent' data set was calculated as above, using the 4th quartile for SC lengths per cell.
 A total of `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$n_obs)` and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$n_obs)` bivalents were isolated from `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="male") )$ncells)` spermatoctyes and `r sum(  (Q1.real.long_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes respectively.
 
-For indiviual mice the MLH1 mean, variance and coefficicent of variance per cell were investigated.
+For individual mice the MLH1 mean, variance and coefficient of variance per cell were investigated.
 We quantified mouse averages of SC based features including the mean total SC area per cell and the mean SC lengths for single bivalents.
@@
 For double crossover bivalents the distance between two MLH1 foci, inter-focal distances (IFD), was examined in IFD^raw^ or the normalized by the SC length IFD^norm^.
-These traits were examined using a our model framework because they are markers of processes in the recombinaiton pathway and provide more detailed information about the recombination landscape.
+These traits were examined using a our model framework because they are markers of processes in the recombination pathway and provide more detailed information about the recombination landscape.
 
@@
 Straight forward comparisons of groups (strain, sex) were done through t.tests().
-**(outgroup species were not included for the analyses).**
 The full mixed model (**M1**) was built to interpret the patterns of variation using the lmer() from the lmer4 package (cite).

### 2020-05-04 doc/Results_v2.rmd
spell check and fixing typos

@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '5'
   html_document:
@@
     toc_float: yes
-  word_document:
-    toc: yes
-    toc_depth: '5'
 always_allow_html: yes
@@
 After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: musculus^MSM^ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and domesticus^G^ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  molossinus^MSM^  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and musculus^PWD^ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
+A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
+and _domesticus^G^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
+In addition, two strains exhibit strain-by-sex interactions:  _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
 
 We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4; see methods).
-In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect =
+In the male dataset, three strains significantly affect recombination rate: _musculus^PWD^_ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect =
 `r sum_MLH1.male.glm.M4$coefficients[8]` foci), and 
-musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
+_musculus^SKIVE^_ (glm; p = 0.01; `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
 
 These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p= `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
+Analysis of the female dataset points to four strains with significant effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and _musculus^PWD^_ (p= `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
@@
 In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains (musculus^KAZ^ , domesticus^WSB^ , and domesticus^G^) and two high-recombination strains (musculus^PWD^ and molossinus^MSM^).
+DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains ( _musculus^KAZ^_ , _domesticus^WSB^_ , and _domesticus^G^_ ) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_ ).
 
@@
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)` (musculus^KAZ^ female).
+In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
 
@@
 low recombining strains (Figure X).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in musculus^PWD^, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in molossinus^MSM^.
-
-Following patterns in the genome-wide recombination rate, male musculus^PWD^ and male molossinus^MSM^ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; musculus^PWD^ vs.
-musculus^KAZ^ p = `r PWD.KAZ.2CO.prop.test$p.value`; molossinus^MSM^ vs.
-molossinus^MOLF^ p = `r MSM.MOLF.2CO.prop.test$p.value`).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
+
+Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
+_musculus^KAZ^_ p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_ p = `r MSM.MOLF.2CO.prop.test$p.value`).
 
@@
 Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male musculus^PWD^ and male molossinus^MSM^ should have longer SC lengths than the other strains.
+In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
 <moved section on reduced bivalent data set to methods>
@@
 We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
-The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except musculus^SKIVE^, which my be attributed to low sample sizes (p = `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`).
-The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` (musculus^MSM^) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (domesticus^WSB^).
+The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`).
+The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
 
@@
 Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
-
-<figure>
 
@@
 We used the single crossover bivalent (1CO) normalized position to compared general location of crossovers while controlling for differences in total SC length.
-Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus) -- thus justifying the use of single crossover bivalents -- to discribe this metric / trait.
+Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.
 
@@
 
-In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < `r Q1.IFD.PER_FULL_t.test$p.value`),
-
-with only musculus^KAZ^ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
+In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < `r Q1.IFD.PER_FULL_t.test$p.value`), with only _musculus^KAZ^_ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
 Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+
 Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1 - LRT **random**: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`  glm *M2* `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`  *M3* `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
@@
 
-**address the difference is pattern signal for IFD triangles and boxplots** (make boxplots?)
+Visual comparison of interference pattern is aided by Figure X, which combines the total length of distance between two foci, but also their normalized locations along the chromosome.
+Two main patterns of sexual dimorphism emerge across all strains, (Supplemental figure X).
+First, the female patterns (top triangle), are more uniformly distributed compared to males.
+Suggesting a wider range of normalized IFD lengths compared to males (bottom triangle).
+Male IFDs display a stronger chromosome end-locatization of the second foci, (clustered near the 90* corner).
 
@@
 
-Next, we used the contrast between males in the high-recombination strains (musculus^PWD^ , musculus^SKIVE^, and molossinus^MSM^) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+Next, we used the contrast between males in the high-recombination strains ( _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ ) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
 The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
@@
 
-Following this logic we predict (1) musculus^PWD^ will have greater SC length and weaker interference than musculus^SKIVE^, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) molossinus^MSM^ will have longer SC and weaker interference compared to molossinus^MOLF^, and 3) _domesticus_ strains will have similar SC length and crossover interference.
+Following this logic we predict (1) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_, and 3) _domesticus_ strains will have similar SC length and crossover interference.
 
@@
 **the values between versions are slightly different**
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, molossinus p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
+In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = _musculus_ `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, _molossinus_ p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
 In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
@@
 The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
-**While there are significant strain effects for domesticus^WSB^ and molossinus^MOLF^ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.* <bret asks if the bolded section is correct.
+**While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.* <bret asks if the bolded section is correct.
 are there sig strain effects for WSB and MOLF?> <bret asks if italised statement has been tested statisticaly or in what sense am I discribing this>
@@
 
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= strain, color=strain))+geom_boxplot(alpha=.2)+
   geom_jitter(alpha=.2)+
-  ggtitle("Normalized Interfocal Distances,\n Low Rec group =0, High Rec group =1")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(subsp~Rec.group, scales = "free")+theme(legend.position="none", axis.text.x = element_blank())+ylim(c(0,1.1))
-
-```
+  ggtitle("")+
+  scale_color_manual(values=colors_of_strains)+ facet_wrap(Rec.group ~ subsp, scales = "free")+
+  theme(legend.position="none", axis.text.x = element_blank(), axis.title.x = element_blank() )  + labs("Normalized IFD ") +
+  ylim(c(0,1.1))
+
+#consider adding horizontal line
+
+#to adjust the boxplot width
+#https://stackoverflow.com/questions/46174222/ggplot2-incorrect-boxplot-width-when-facetting-with-facets-of-different-scales
+#I don't think there's a straightforward way to achieve that in ggplot.
+You may have better luck plotting each facet as a separate ggplot object (on a subset of the data), then arrange them on a single page, using grid.arrange() from gridExtra for example.
+
+```
+
+Merge Dom together,   merge Musc - low, show mol seperately
 
@@
 #big loop for automating the IFD triangles
+
 Q2.males.IFD <- c("F1", "other")
-Q2.auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% Q2.males.IFD, ]
+Q2.auto_curate_DF <- cleanCurated_BivData[ ! cleanCurated_BivData$strain %in% Q2.males.IFD, ]
 
@@
   points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[current.strain])
+          col =  colors_of_strains[unique(my.data$strain)])
 
@@
   #this is calling axis twice, not sure why
-
   lines(0:1, 0:1, xpd=TRUE) #diagonal line
@@
   #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
+  png(paste0("~./MLH1repo/doc/figureFiles/fff.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
   plot.new()
 
-  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
-  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+  my.data = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
+  my.x.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
+  my.y.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
 
@@
 
+```{r fuck, eval=FALSE}
+
+Q2.Dom_DF <- Q2.auto_curate_DF %>% filter(sex == "male") %>% filter(subsp == "Dom")
+
+Dom_DF.Q2.df_strain_list <- split(Q2.Dom_DF, (Q2.Dom_DF$strain))
+Q2.df_strain_list
+
+Dom_DF.Q2.df_strain_list
+
+for(i in 1:(length(Q2.df_strain_list))){
+  #print(i)
+  current.strain <- as.character(Dom_DF.Q2.df_strain_list[[i]]$strain[1])
+  print(current.strain)
+}
+
+my.data.Dom = (Q2.Dom_DF %>% filter(hand.foci.count == 2) %>% filter(sex == "male") )
+my.x.bottom.Dom = (Q2.Dom_DF %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
+my.y.bottom.Dom = (Q2.Dom_DF %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+png(paste0("~./MLH1repo/doc/figureFiles/Dom.IFD.triangle_", "wtf", ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
+plot.new()
+
+#adjusting this mar lowers the margins (the edges)
+op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))
+
+Q2.bottom.triangle2(my.y.bottom.Dom, my.x.bottom.Dom, my.data.Dom) #x.dist
+par(op)
+dev.off()
+
+```
+
 ```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
@@
 We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
-The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (domesticus^G^) to 16% (musculus^KAZ^) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for musculus^SKIVE^, molossinus^MSM^, and musculus^PWD^, respectively).
+The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (_domesticus^G^_) to 16% (_musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
+
+Differnces in IFD can also be noted from the triangle IFD plots (Figure X).
+We note that the three high-recombining strains show more clustering of IFD observations near the area of the plot which indicates longer IFDs.
+The remaining strains have more observations short IFD observations near the diagonal.
 

### 2020-05-04 doc/MS_v1.Rmd
merging all thress sections together to one file

--- 2020-05-04 doc/Results_v2.rmd
@@
 ---
-title: "MLH1 Results Outline"
-bibliography: refs_Main_Report.bib
-date: 'null'
+title: 'MS draft'
+date: '`r Sys.Date()`'
 output:
@@
     toc_depth: '5'
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
   html_document:
-    df_print: paged
     toc: yes
@@
 always_allow_html: yes
+bibliography: refs_Main_Report.bib
 ---
 
-**sex differences in Celegan meiosis**
-
-Kelly, William G., Sara M. Fielder, and Rieke Kempfer. "Multiple Sex-Specific Differences in the Regulation of Meiotic Progression in C. elegans." bioRxiv (2020).
-
-```{r setup, echo=FALSE, include = FALSE}
+```{r set.up, echo=FALSE, include=FALSE}
+
+#library(tidyverse)
 library(knitr)
-library(kableExtra)
 library(ggplot2)
 #library(pwr)
+
+#library(dplyr)
 library(plyr)
 library(lattice)
-library(dplyr)
+
 library(raster)
@@
 library(patchwork)
-library(pander)
-library("xlsx")
-library(png)
-#library(tidyverse)
 library(png)
@@
 library(cowplot)
+library(kableExtra)
+
+#set wd
+#load datafile
 
@@
 
-setwd("~.")#switch over to here()
-
-#load(file = here("data", "MLH1_data_setup_4.21.20.RData") )
-
-load(file = "~./MLH1repo/data/Results_setup_4.24.20.RData")
+#setwd("~.")#switch over to here()
+
+load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")
+
+load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
 
 load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
-
 load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
@@
 
-#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
-
 #load functions
@@
 
-<!-- Data Set Up Notes -->
+#Intro
+
+#Material Methods Draft v1
+
+### Mouse Husbandry
+
+To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
+The _Mus musculus_ strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
+The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
+(https://en.brc.riken.jp) and rederived by the UW-Biotech centre (Madison UW).
+The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
+We use abriviated superscript notation for the remainder of the paper (e.g.
+_musculus^WSB^_).
+All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
+A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (G) is maintained at UW Veterinary school facilities.
+
+<Mice were fed on dry standard breeder chow.
+Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
+Adult mice were euthanized by CO asphyxiation.
+Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison.
+Over the course of data collection some breeding colonies  mice were moved from facilities.
+Additionally the GI strain was kept at a separate facility.>
+
+### Tissue Collection and Immunohistochemistry
+
+The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
+Spermatocyte spreads were collected and prepared as described in [@peterson2019].
+The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
+Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
+
+Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
+Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
+Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
+The following morning, slides were treated with a photoflow wash (Kodak, cite).
+**slides were stored at -20*C if not stained imeediantly**
+
+<Because some strains have breeding issues, all mice some strains were kept in breeding pairs (before collecting, and aged out of the desired age range.
+We performed tests for effects on MLH1 counts and found no effects in the older male mice ages.
+(Also we tested other effects, including maternal age...)>
+
+The **immunohistochemistry**
+
+ To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
+Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
+The staining protocol was based on that in [@anderson1999] and [@koehler2002].
+Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
+
+Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
+The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
+Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200)  were suspended in ADB.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
+Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
+
+### Image Processing
+
+Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
+Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
+Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.
+
+### Analysis
+
+We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).
+
+### DMC1
+
+Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
+
+SC morphology and CREST/centromere number was used to stage spermatocytes **Certeria for early and late zygotene** - (bright staining - )
+early zygotene defined as seperate centromere pairing -- mostly complete axis signal.
+late zygotene judged as -- low degree of centromere pairing -- and more homologous pairing (but not fully synapsis).
+Differences in DMC1 counts per cell were tested by ttests.
+<numbers of cell?>
+
+### MLH1 (Analytical Framework)
+
+<Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.   -ene endings refer to the cells/units within the cell stage? (I think)>
+
+**certeria**  MLH1 foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
+With scores of 1 and 2 having the highest quality.
+Cells and coded as a score of 5 were excluded from the final analysis.
+The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
+MLH1 foci located on the XY in spermatocytes were excluded from this study.
+
+**M1 Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
+
+**M2 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
+
+**M3 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
+
+**M4 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
+
+**M4.2 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
+
+Straight forward comparisons of groups (strain, sex) were done through t.tests().
+
+We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
+Strain was coded as a random effect and tested using exactRLRT() to reflect a random sample of natural genetic variation from each subspecies.
+The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
+
+This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
+We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
+The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
+The general linear models, **M4** were used to analyze sex specific data sets.
+
+Our goal was to design an analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
+sex, strain subspecies) and can be applied to multiple (meiotic traits) dependent variables.
+In addition to the MLH1 count variables for individual mice, mean MLH1 count per cell, variance and coefficient of variance for MLH1 count per cell, we analyzed SC based features: mouse average of total SC per cell, mouse average of single SC length, mouse mean IFD^norm^ and IFD^raw^ for double crossover chromosomes, and the average position of crossover for single crossover bivalents.
+
+These traits were choosen because they are markers of phenomena in the recombinaiton pathway and provide more detailed information about the recombination landscape.
+In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
+Normalized measures were calculated by the position divided by the length of the SC for that single bivalent.
+
+### Single bivalent charaecterization
+
+To characterize the SC length of individual chromosomes, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
+Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
+Hand measures of SC/pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
+
+The **total SC** per cell was quantified from the pachtyene cell images.
+To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
+Mouse means were calculated from cell-wide total SC lengths in `r length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name)` out of `r length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
+
+The **DNA CrossOver** algorithm isolates single straightened bivalents/chromosomes shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
+This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
+In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)` (musculus^KAZ^ female).
+
+A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
+From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` bivalent objects were isolated by the image analysis software.
+After a human curation step , `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step ([@peterson2019])).
+
+From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
+
+**REDUCED DATASET**
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
+A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
+
+#Results
 
@@
 
-```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
+```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=TRUE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
 B) Female specific MLH1 count distributions.
@@
 
-# Genome-wide recombination rate estimates for both sexes
+## Genome-wide recombination rate estimates for both sexes
 
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
+```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
 #write the full table to a supplemental table
@@
 
-kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
-  kable_styling() %>%
-  scroll_box(width = "800px", height = "200px")
+#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
+#  kable_styling() %>%
+#  scroll_box(width = "800px", height = "200px")
 
@@
 
-# Partitioning variation in recombination rate
+## Partitioning variation in recombination rate
 
@@
 
-# Within mouse variance in CO count per cell
+## Within mouse variance in CO count per cell
 
@@
 
-# DSB. Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
+## DSB. Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
 
@@
 
-kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-  kable_styling() %>%
-  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-  scroll_box(width = "800px", height = "200px")
+#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
+#  kable_styling() %>%
+#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
+#  scroll_box(width = "800px", height = "200px")
 
@@
 
-DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
+#DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
+#  geom_boxplot(position=position_dodge()) + 
+#  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
 
@@
 
-  ggsignif::geom_signif(
-    data=anno_df_stage, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+#  ggsignif::geom_signif(
+#    data=anno_df_stage, 
+#    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+#    manual=TRUE
+#  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+#        panel.background = element_blank(), axis.line = element_line(colour = "black"))
 
@@
 
-# Evolution of genome-wide recombination rate is reflected at the single chromosome level
+## Evolution of genome-wide recombination rate is reflected at the single chromosome level
 
@@
 
-# Q1. Sex Differences in the Length of the Synaptonemal Complex
+## Q1. Sex Differences in the Length of the Synaptonemal Complex
 
@@
 
-kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
-  kable_styling() %>%
-  scroll_box(width = "800px", height = "200px")
+#kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
+#  kable_styling() %>%
+#  scroll_box(width = "800px", height = "200px")
 
@@
 
-# Q1. Sex Differences in the Positions of Single Crossovers
+## Q1. Sex Differences in the Positions of Single Crossovers
 
@@
 
-# Q1. Sex Differences in CO Interference (IFD)
+## Q1. Sex Differences in CO Interference (IFD)
 
@@
 
-# Additional Determinants of Genome-wide Recombination Rate Evolution in Males
+## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
 
@@
 
-# Q2 SC Length
+## Q2 SC Length
 
@@
 
-# Q2.1CO rec landscape evolution is decoupled from gwRR evolution
+## Q2.1CO rec landscape evolution is decoupled from gwRR evolution
 
@@
 
-# Q2 Evolution of interference is associated with genome wide recombination rate evolution
+## Q2 Evolution of interference is associated with genome wide recombination rate evolution
 
@@
 
-```{r fuck, eval=FALSE}
-
-Q2.Dom_DF <- Q2.auto_curate_DF %>% filter(sex == "male") %>% filter(subsp == "Dom")
-
-Dom_DF.Q2.df_strain_list <- split(Q2.Dom_DF, (Q2.Dom_DF$strain))
-Q2.df_strain_list
-
-Dom_DF.Q2.df_strain_list
-
-for(i in 1:(length(Q2.df_strain_list))){
-  #print(i)
-  current.strain <- as.character(Dom_DF.Q2.df_strain_list[[i]]$strain[1])
-  print(current.strain)
-}
-
-my.data.Dom = (Q2.Dom_DF %>% filter(hand.foci.count == 2) %>% filter(sex == "male") )
-my.x.bottom.Dom = (Q2.Dom_DF %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
-my.y.bottom.Dom = (Q2.Dom_DF %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-png(paste0("~./MLH1repo/doc/figureFiles/Dom.IFD.triangle_", "wtf", ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
-plot.new()
-
-#adjusting this mar lowers the margins (the edges)
-op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))
-
-Q2.bottom.triangle2(my.y.bottom.Dom, my.x.bottom.Dom, my.data.Dom) #x.dist
-par(op)
-dev.off()
-
-```
-
 ```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
@@
 
+# Discussion Draft v2
+
+Despite quantifying a similar number of wild derived inbred strains from *M.
+m.
+domesticus* and *M.
+m.
+musculus*, we only observed rapid male specific evolution for the genome wide recombination rate for *M.
+m.
+musculus.* Estimates of effective population size for the three main subspecies of the house mouse found *M.
+m.
+musculus* had the lowest [@geraldes2011].
+Selection becomes less efficient at removing deleterious mutations in small populations outside of mutation-drift balance.
+In such populations, modifiers for increased recombination rates will be selected for as they increase the efficiency of selection.
+
+An additional possibility is that the different rates in recombination rate variation across subspecies is related to history of hybridization across the populations sampled in this panel of inbred strains.
+Theoretical work as connected recombination that rate of purging deleterious introgressed DNA in populations after hybridization (Caballero et al 2019, Schumer 2018).
+The three high-recombining strains identified in this study have connections to hybridization.
+Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the european house mouse hybrid zone.
+Notably *musculus^SKIVE^* comes from a yonger section of the hybrid zone with a nuclear genome from *M.
+m.
+musculus* and a mitochondria genome from *M.
+m.
+domesticus* [@gyllensten1987_SKIVE].
+The subspecies *M.
+m.
+molossinus* is a natural hybrid of *M.
+m.
+musculus* and *M.
+m.
+cataneus* (Geraldes et al 2008, [@geraldes2011]).
+Models for the aggregate recombination metics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019) -- motivating increased study of recombination rates and natural populations.
+4
+<transition to looking at the number and placement of crossovers>
+
+To place our results in the broader context of recombination rate evolution we compare our main resuts to expectionat under three modifier models and two functional and cell physiology models (table X).
+While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across exsisting models.
+The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome ocillatory movement (COM) model was built to describe positive interference.
+This spindle based selection model was developed to explain the results resented in this manuscript.
+In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
+
+### Evolution of reversed heterochiasmy direction
+
+```{r reversed.HetC, echo=FALSE}
+
+#our results of reversed hetC are novel due to the short evolutionary disances across the panel of mouse strains
+
+#spindle selection modle points
+
+# -- counter points, whats the consequence of reverse heterochiasmy (rapid male evolution in gwRR), 
+# -- sex-average / species average of gwRR is higher (for these strains)
+# --
+
+```
+
+In our results and more generally in the literature, female biased heterochiasmy is the most common pattern, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
+Our results extend previous reports of male biased heterochiasmy [@dumont2011] to  new subspecies of house mouse (*molossinus*) and emphasize the short evolutionary period.
+Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within house mouse subspecies requires further study.
+Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species in female biased heterochiasmy species, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
+This insight could be important for animal breeding programs [@battagin2016].
+
+Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
+<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
+
+A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion ([@laneKauppi2019]).
+Genetic variants will be selected in a manner to optimize the process of bi-orientation and prevent SAC from delaying cell cycle or triggering apoptosis.
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014]), which are connected to conserved features of gametogenesis; centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
+The more stringent SAC in spermatogenesis will be more effective at removing genetic variants which interfere with bi-orientation compared to females.
+This dynamic can give rise to sex-specific genetic variants which in turn can result in  context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
+
+### Conservation in sexual dimorphism in between cell variation in crossover number
+
+```{r within.mouse.var, include=FALSE}
+#review results, other data
+
+#2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
+
+#2a.
+relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
+The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.
+
+#?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
+
+# consequence of greater female variance -- (between cell)
+
+```
+
+Regardless of whether these hypotheses are true, Our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
+Across all strains, including the high-recombining strains, musc^PWD^, musc^SKIVE^ and mol^MSM^, females consistently have higher between cell variance for crossover count.
+These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005]?, compared to spermatocytes.
+This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
+
+In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
+This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals.
+Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
+Centrosomes spindles are in spermatogensis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization (@manandhar2005centrosome).
+This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
+
+Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013].
+This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
+
+<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
+
+### Conservation in sex-specific recombination landscape
+
+```{r typical.landscape.outline, include=FALSE}
+
+#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
+#3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
+#3a.
+There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.
+In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle.
+(required to start with more since the cohesion isn't replenished.
+The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.
+
+#3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
+#-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs.
+This area reduced when crossover position shifted to telomere (each end of chromosomes).
+Two crossover acrocentric chromosome the area is distal from the first until the second crossover.
+
+#3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
+#- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.
+
+#0. we find the conserved pattern of the typical landscape 
+#this trait is a highly conserved aspect of heterochiasmy 
+#the exceptions include marsupials which form sex chromosome chains during meiosis -- indicating a potential change to the meiotic program (functional selection)
+#it's not known if a similar -- evolutionary force in the other exceptions to the typical landscape, pig and tomatoe.
+
+#regaurdless
+#As reviewed in sardel kirkptarick - diviation for a uniform background is generally due to the male recombination landscape.
+
+# these pattern due to (different recombination landscapes along chromosomes can have an impact on the pattern of Fst (Hanel et al 2018)
+# additionally other patterns of nucleotide and recombination rates could vary across
+
+#co rate defines the physical scale of linked selection around a locus
+
+# evolutionary genomics from the broad scale heterogenity
+#linked selection, (maynarid smith and haig 1974)
+#background selection (only with negtive allele), genetic draft (for positive selective mutations)
+#gene flow barrier (directional selection is keeping two populations from merging)
+# patterns of Fst (Hanel et al, but see Kartej?)
+
+# Hanel et al predicts -- lower genetic diversity near centers -- away from chromosome ends (this is supported )
+#finds that the strenght of telomere is connected to chrm size (positive stronger) -- independant of cenromere
+
+# source could be DSB formation along chromosome - and /or biases in DSB repair 
+#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick)
+```
+
+Our results extend (the mountain / plethora) or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] - data comprised mostly of highly divergence species / large evolutionary distances.
+In contrast our results demonstrate that this pattern is maintained across much shorter evolutionary distances and even with rapid evolution in the genome-wide recombination rate.
+We argue that this conservation of the male chromosome end bias // recombination landscape is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
+
+As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a conserved trait.
+The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
+The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM].
+
+Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
+The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs [@laneKauppi2019], subramanian and Hochwagen , @dumontDesai2012].
+The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
+Differences in timing of cell cycle between oogenesis and spermatogenesis imposes the the different selective pressures on how sister cohesion affects chromosome structure.
+Faster spermatogenesis may select for synchronization of the separation homologs (cite).
+While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, espicially for those with dicyate arrest [@lee2019].
+
+The higher levels /increased genetic diversity and recombination rates at chromosome ends @haenel2018 - suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020].
+(giving rise to interesting evolutionary predictions).
+
+### Conservation of sex differences in chromatin structure
+
+```{r chromatin.organization, include=FALSE}
+
+#supporting results from lit
+#chromatin organization
+
+#models
+# none of the models make predictions for the axis length (chromatin organization)
+# only COM predicts that sexual dimorphism in axis length is due to cell vol
+
+# consequences of longer female axis - 
+# (normalzed interference strength due to axis differences / Chromtain loop length)
+
+# significance, consequences
+# weak evidence of sex differences in chromatin organization, mice-humans have a similar direction -- but At opposite -- not much empiracal evidence. (more data needed)
+
+# a consequence of these sex differences for the chromatin organization is -- has an outcome of different effects on interference strenth measured in genetic map space -- for a given chromosome, chromatin organization of longer axis will result in shorter loops (emenating off the central axis)
+
+#4.ppl should care -- this is an important aspect of -- how sexual dimorphism in recombination /linkage maps has been charaecterized -- but due to a lack of cytological evidence in both sexes /linking this to chromatin organization has only been done for a handful of species (mice, human, At?
+Dourad)
+
+```
+
+Our results from musculus^MSM^ and musculus^PWD^ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
+These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
+Chromatin compaction could explain variation in heterochiasmy, which is driven by the disinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
+Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
+females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation (see below).
+
+Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
+This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier paper?
+Sera?) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+
+As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
+From empirical data in mice [@petkov2007, @lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
+The signal for interference is thought to be mediated though the SC, which is built on top of the chromosome axis [@zicklerKleckner2015] regardless of the loop size.
+Supporting this is the observation that interference strength measured in micrometers of is conserved between sexes [@petkov2007].
+As a consequence, since more base pairs are sequestered into the radial loops per unit of the linear central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males,  will have a signal of stronger interference when measured in base pairs compared to 'the typical female chromatin configuration.
+Stronger interference in males, has been noted across many sex-specific linkage maps (cite) strengthening the hypothesis that sexual dimorphism in the chromatin organization is widespread, however this has only been confirmed by cytological data in mice and humans.
+(Also large difficulties in observing sperm cells through meiosis).
+
+<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]>
+
+### Evolution of interference strength in males
+
+```{r evo.interference, include=FALSE, echo=FALSE}
+
+#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)
+#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)
+
+#**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
+
+#5. Interference evolution shows a positive correlation with gwRR
+#5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
+#5b.
+Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
+
+# consequence of rapid evolution of normalized interference strength,
+# in the 3 strains -- larger linkage blocks?
+# but what if the rec pattern is repeated?
+```
+
+An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (cite, Schumer).
+Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
+**(new, veller paper -- in addition to the total gwRR, the location of crossovers (rbar), should be considered in this dynamic)**
+
+Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
+The general pattern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
+This fits the logic of the chromosome axis acting as the 'real estate' for DSBs and aligns with known molecular mechanisms (cite).
+Two examples of the opposite direction, a positive correlation of interference strength and genome-wide recombination rate, involve descriptions of observations at the single bivalent level; the between lab-raised and wild mice of *Peromyscus leucopus* from [@peterson2019] or large effects on genome wide recombination rates across a short evolutionary differences; the within-sex comparison of two breeds of cattle with different genome wide recombination rates [@ma2015_cattle].
+
+We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphse.
+The evolution of genome wide recombination rates in our study is driven by a transition of the majority of
+bivalent having single crossovers to an enrichment of bivalents with two crossovers.
+All else being equal, interference strength would be expected to be equal or even weaker.
+Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain enriched for two crossover bivalents.
+Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
+
+The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase [@rubin2020mixing], however it lacks a mechanism for a evolutionary based predictions.
+While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
+
+We have focused on the model involving direct selection on the recombination pathway, which has an impact on the fitness of an individual, but we acknowledge connections to other models involving indirect selection on recombination rates which involve the fitness outcomes of offspring.
+A model from @goldstein1993, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
+The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
+A distinguishing feature of this model from other modifier  models is that the number of crossovers per chromosome is kept constant.
+Empirically, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
+Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
+
+### Future steps
+
+There are still many un-known aspects regarding heterochiasmy (Lenormand) , these novel results suggest - we hope ppl we follow up .  (fitness - evolutionary dynamics - role in linked selection)
+
+We make three suggestions for future steps in studying recombination variation.
+First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
+
+Second certain systems can be leveraged to overcome limitations inherent in some approached.
+For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
+FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian trans-locations.
+
+Third we encourage the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
+A goal of such merging of models could be to connect empirical findings across scales.
+
 # References

### 2020-05-04 doc/Discussion.rmd
commting during debugging

@@
 # Discussion Draft v2
-
-```{r format.Notes, include=FALSE}
-#table, wite by all of the columns and rows
-
-#over-view of the synthesis of hypothesis testing
-#(this is how we're synthsizing / testing / the framework were using to compare the results to the literature)
-
-#constrast on model types
-#-evolution
-#-models accounting for chromosome based traits
-
-#HetC traits
-# 1. typical landscape
-# 2. female variance
-# 3. chromatin organization
-
-#faster male evolution traits
-# 1. Reversed Direction of Heterochiasmy
-# 2. correlated gwRR increase and interference strength
-
-#go through each traits and point out which models support each of the traits
-```
 

### 2020-05-04 doc/MS_v1.Rmd
commting during debugging
finializing edits

@@
 date: '`r Sys.Date()`'
+author: "April L. Peterson, Bret Payseur"
 output:
   word_document:
-    toc: yes
-    toc_depth: '5'
   pdf_document:
@@
 library(RLRsim)
-library(ggpubr)
+#library(ggpubr)
 library(reshape2)
@@
 
-<Mice were fed on dry standard breeder chow.
-Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival.
-Adult mice were euthanized by CO asphyxiation.
-Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison.
-Over the course of data collection some breeding colonies  mice were moved from facilities.
-Additionally the GI strain was kept at a separate facility.>
-
 ### Tissue Collection and Immunohistochemistry
@@
 Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
-
 Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
@@
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a photoflow wash (Kodak, cite).
-**slides were stored at -20*C if not stained imeediantly**
-
-<Because some strains have breeding issues, all mice some strains were kept in breeding pairs (before collecting, and aged out of the desired age range.
-We performed tests for effects on MLH1 counts and found no effects in the older male mice ages.
-(Also we tested other effects, including maternal age...)>
-
-The **immunohistochemistry**
+The following morning, slides were treated with a Photoflow wash (Kodak, cite).
+Slides were stored at -20*C if not stained immediately.
 
@@
 
-We tested for **effects** on mean MLH1 counts per cell; across mulitple mouse rooms, ages of mature males and differnces between embryos and neonates. No effects were found (Supplemental section).
-
-### DMC1
-
-Mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
-
-SC morphology and CREST/centromere number was used to stage spermatocytes **Certeria for early and late zygotene** - (bright staining - )
-early zygotene defined as seperate centromere pairing -- mostly complete axis signal.
-late zygotene judged as -- low degree of centromere pairing -- and more homologous pairing (but not fully synapsis).
-Differences in DMC1 counts per cell were tested by ttests.
-<numbers of cell?>
-
-### MLH1 (Analytical Framework)
-
-<Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.   -ene endings refer to the cells/units within the cell stage? (I think)>
-
-**certeria**  MLH1 foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+To estimate the numbers of percursor sites which may lead to crossovers, the mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
+SC morphology and CREST/centromere foci number was used to stage spermatocytes, in early and late zygotema stages.
+
+**MLH1** foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
 A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
@@
 
-Straight forward comparisons of groups (strain, sex) were done through t.tests().
-
+A series of mixed models and general linear models were used as an analetical framework.
 We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
@@
 
-### Single bivalent charaecterization
+### Single bivalent characterization
 
@@
 
-The **total SC** per cell was quantified from the pachtyene cell images.
+The **total SC** area per cell was quantified from the pachtyene cell images.
 To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in `r length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name)` out of `r length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name)` cells with MLH1 counts (Figure X).
+Mouse means were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts (Figure X).
 
@@
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` (molossinus^MSM^ male) to `r max(algorithm.stat.category$mean_pass.rate)` (musculus^KAZ^ female).
+In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
 
 A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
-From the total set of cell images, `r length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName )` bivalent objects were isolated by the image analysis software.
-After a human curation step , `r length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID )` single-bivalent observations remained.
+From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` bivalent objects were isolated by the image analysis software.
+After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
 The accuracy of the algorithm is high compared to hand measures after this curation step ([@peterson2019])).
-
 From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
 
-**REDUCED DATASET**
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a **reduced data** set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
 The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of 678 -- `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from 103 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and 37 -- `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
 Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
@@
 
-#Results
-
-```{r FOR.PLOT.organizing.data, echo=FALSE }
+# Results
+
+```{r FOR.PLOT.organizing.data, echo=FALSE, warning=FALSE}
 #set the order of another column, based on another variable. consistant order for ploting
@@
 
-```{r Fig1.top, echo=FALSE, warning=FALSE}
+```{r Fig1.top, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
 
@@
 
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
 #try customize the plot size
@@
 
-```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=TRUE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
+```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
 B) Female specific MLH1 count distributions.
-C)	Male specific MLH1 counts per cell.'}
+C)	Male specific MLH1 counts per cell.', message=FALSE}
 
@@
 
-<notes about figure; make sure the order highlights the difference (mol, musc, dom). Clean up the axis and titles. Resizing might be an issue too.>
-
 ```{r MLH1.summ.stats, echo=FALSE}
@@
 We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < <r MLH1.MMlmer_results\$rand.pval> `r 1/1000000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
+(random effect p < `r 1/1000000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
 
@@
 `r sum_MLH1.male.glm.M4$coefficients[8]` foci), and 
-_musculus^SKIVE^_ (glm; p = 0.01; `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
+_musculus^SKIVE^_ (glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
 
@@
 Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
-
-<!-- transition gwRR to within mouse variance -->
 
@@
 Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
-The results are similar: sex is the strongest effect (M1 p < `r 1/1000000` <round(LmerMLH1_M1_Q12_VAR_results\$sex.results, 5)>; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
+The results are similar: sex is the strongest effect (M1 p < `r 1/1000000`; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
 When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
@@
 
-**(sex specific results - no efficient strain effect for males. In females LEW - but this is due to cell quality difference, what about quality score and cell number within the models?**
-
-<strain specific effect section was deleted for not highlighting main pattern. I agree> **re-read Lenzi et al -- heterogeneity in oogenesis**
-<Bret's comment for below, are there differences between the curated and the full dataset? he propbably means, Q1 and full data set.>
-
-## DSB. Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
+<LmerMLH1_M1_Q12_VAR_results> <Bret's comment for below, are there differences between the curated and the full dataset? -- Q1 and full data set>
+
+## Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
 
@@
 
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
 #change order
@@
         panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.x = element_blank())
-
 ```
@@
 B)	Distribution of DMC1 counts per cell  by strains.
-(1 mouse was used for each strain)"}
+(1 mouse was used for each strain)", eval=FALSE}
 
@@
   subtitle = '') + plot_annotation(tag_levels = 'A')
-
 ```
@@
 
-anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage #!= "P",],method = "t.test") %>% mutate(y_pos = 300)
 
@@
 #2. t-tests for the DMC1 counts
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
 
@@
 DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains ( _musculus^KAZ^_ , _domesticus^WSB^_ , and _domesticus^G^_ ) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_ ).
-
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p < `r 1/1000000` <ttest.HighLow.L.pval>).
+The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p < `r 1/1000000`).
 In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = `r ttest.HighLow.Z.pval`).
+
+<ttest.HighLow.L.pval>
 
@@
 The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
-(connection to barier)
-
 This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
 @cole2012 suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
-
 This result, combined with the large difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
 
-<The number of DSBs increase in zygotene, then deminish as the effects of crossover homeostasis are implemented (Cole).
+<!-- The number of DSBs increase in zygotema, then deminish as the effects of crossover homeostasis are implemented (Cole).
 (co homeostasis also indecates dsb ) Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links connects the CO number variation to 'pre' DSB formation fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
-(this study only examined one stage of early prophase for DMC1 counts)>
-
-<the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 C3H, 7.3 B6>
+(this study only examined one stage of early prophase for DMC1 counts -->
+
+<!-- the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 C3H, 7.3 B6 -->
 
@@
 To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents ([@peterson2019]).
-
-<bret suggests moving this section to Mat and Methods>
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
 In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
-
 From the total set of cell images, `r formatC(sum( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ) ), , format="d", big.mark="," )` bivalent objects were isolated by the image analysis software.
@@
 We assume that the isolation of bivalents within cells is unbiased.
-
 Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure X).
@@
 High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
-
 Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
@@
 _molossinus^MOLF^_ p = `r MSM.MOLF.2CO.prop.test$p.value`).
-
 Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
@@
 
-## Q1. Sex Differences in the Length of the Synaptonemal Complex
+## Sex Differences in Meiotic Traits
 
@@
 
-two.Q1.SC.plots + plot_annotation(
-  title = '',
-  subtitle = '') + plot_annotation(tag_levels = 'A')
+#two.Q1.SC.plots + plot_annotation(
+#  title = '',
+#  subtitle = '') + plot_annotation(tag_levels = 'A')
 
@@
 In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
-<moved section on reduced bivalent data set to methods>
-
-<figure>
 
@@
 
-## Q1. Sex Differences in the Positions of Single Crossovers
+### Positions of Single Crossovers
 
@@
 plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
-
 plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
@@
 plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
-
 plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
 
-grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
-             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
-             rasterGrob(plot7.MSM), 
-             ncol=3)
+# grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
+#             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
+#             rasterGrob(plot7.MSM), 
+#             ncol=3)
 
@@
 #WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
-
 ```
@@
 
-```{r Q1.single.biv.table, echo=FALSE}
-
-#strain.BivData.Table
-
-```
-
 ```{r Comp_Ratios, echo=FALSE}
@@
 Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-**SC length is expected to be positively associated with genome-wide recombination rate because of the loop-axis structure** (zickler Kleckner 1999, Merier 2015?).
+**SC length is expected to be positively associated with genome-wide recombination rate because of the loop-axis structure** (zickler Kleckner 1999, Mercier 2015).
 Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome (on chromosomes of  a finite length, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
@@
 ```
-
-<!-- plots -->
 
@@
 ```{r Q2.glm.SC.long, echo=FALSE}
-
 #this dataset is all males
@@
 The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
-**While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.* <bret asks if the bolded section is correct.
-are there sig strain effects for WSB and MOLF?> <bret asks if italised statement has been tested statisticaly or in what sense am I discribing this>
+**While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.*
+<!-- bret asks if the bolded section is correct. are there sig strain effects for WSB and MOLF?, bret asks if italised statement has been tested statisticaly or in what sense am I discribing this -->
 
@@
 ```
-
-Merge Dom together,   merge Musc - low, show mol seperately
 
@@
   #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/fff.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
+  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
   plot.new()
@@
 
-# Discussion Draft v2
+# Discussion
 
@@
 4
-<transition to looking at the number and placement of crossovers>
+
+<!-- transition to looking at the number and placement of crossovers -->
 
@@
 
-### Evolution of reversed heterochiasmy direction
-
-```{r reversed.HetC, echo=FALSE}
-
-#our results of reversed hetC are novel due to the short evolutionary disances across the panel of mouse strains
-
-#spindle selection modle points
-
-# -- counter points, whats the consequence of reverse heterochiasmy (rapid male evolution in gwRR), 
-# -- sex-average / species average of gwRR is higher (for these strains)
-# --
-
-```
+## Evolution of reversed heterochiasmy direction
 
@@
 
-### Conservation in sexual dimorphism in between cell variation in crossover number
-
-```{r within.mouse.var, include=FALSE}
-#review results, other data
-
-#2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
-
-#2a.
-relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells.
-The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.
-
-#?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?
-
-# consequence of greater female variance -- (between cell)
-
-```
+## Conservation in sexual dimorphism in between cell variation in crossover number
 
 Regardless of whether these hypotheses are true, Our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
-Across all strains, including the high-recombining strains, musc^PWD^, musc^SKIVE^ and mol^MSM^, females consistently have higher between cell variance for crossover count.
+Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
 These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005]?, compared to spermatocytes.
@@
 
-<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>
-
-### Conservation in sex-specific recombination landscape
-
-```{r typical.landscape.outline, include=FALSE}
-
-#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
-#3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
-#3a.
-There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.
-In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle.
-(required to start with more since the cohesion isn't replenished.
-The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.
-
-#3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
-#-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs.
-This area reduced when crossover position shifted to telomere (each end of chromosomes).
-Two crossover acrocentric chromosome the area is distal from the first until the second crossover.
-
-#3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
-#- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.
-
-#0. we find the conserved pattern of the typical landscape 
-#this trait is a highly conserved aspect of heterochiasmy 
-#the exceptions include marsupials which form sex chromosome chains during meiosis -- indicating a potential change to the meiotic program (functional selection)
-#it's not known if a similar -- evolutionary force in the other exceptions to the typical landscape, pig and tomatoe.
-
-#regaurdless
-#As reviewed in sardel kirkptarick - diviation for a uniform background is generally due to the male recombination landscape.
-
-# these pattern due to (different recombination landscapes along chromosomes can have an impact on the pattern of Fst (Hanel et al 2018)
-# additionally other patterns of nucleotide and recombination rates could vary across
-
-#co rate defines the physical scale of linked selection around a locus
-
-# evolutionary genomics from the broad scale heterogenity
-#linked selection, (maynarid smith and haig 1974)
-#background selection (only with negtive allele), genetic draft (for positive selective mutations)
-#gene flow barrier (directional selection is keeping two populations from merging)
-# patterns of Fst (Hanel et al, but see Kartej?)
-
-# Hanel et al predicts -- lower genetic diversity near centers -- away from chromosome ends (this is supported )
-#finds that the strenght of telomere is connected to chrm size (positive stronger) -- independant of cenromere
-
-# source could be DSB formation along chromosome - and /or biases in DSB repair 
-#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick)
-```
-
-Our results extend (the mountain / plethora) or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] - data comprised mostly of highly divergence species / large evolutionary distances.
+<!-- stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review -->
+
+## Conservation in sex-specific recombination landscape
+
+Our results extend the plethora or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] are data comprised mostly of species with large evolutionary distances.
 In contrast our results demonstrate that this pattern is maintained across much shorter evolutionary distances and even with rapid evolution in the genome-wide recombination rate.
-We argue that this conservation of the male chromosome end bias // recombination landscape is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
+We argue that this conservation of the male chromosome end bias is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
 
@@
 
-The higher levels /increased genetic diversity and recombination rates at chromosome ends @haenel2018 - suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020].
-(giving rise to interesting evolutionary predictions).
-
-### Conservation of sex differences in chromatin structure
-
-```{r chromatin.organization, include=FALSE}
-
-#supporting results from lit
-#chromatin organization
-
-#models
-# none of the models make predictions for the axis length (chromatin organization)
-# only COM predicts that sexual dimorphism in axis length is due to cell vol
-
-# consequences of longer female axis - 
-# (normalzed interference strength due to axis differences / Chromtain loop length)
-
-# significance, consequences
-# weak evidence of sex differences in chromatin organization, mice-humans have a similar direction -- but At opposite -- not much empiracal evidence. (more data needed)
-
-# a consequence of these sex differences for the chromatin organization is -- has an outcome of different effects on interference strenth measured in genetic map space -- for a given chromosome, chromatin organization of longer axis will result in shorter loops (emenating off the central axis)
-
-#4.ppl should care -- this is an important aspect of -- how sexual dimorphism in recombination /linkage maps has been charaecterized -- but due to a lack of cytological evidence in both sexes /linking this to chromatin organization has only been done for a handful of species (mice, human, At?
-Dourad)
-
-```
-
-Our results from musculus^MSM^ and musculus^PWD^ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
+The increased genetic diversity and recombination rates at chromosome ends @haenel2018 suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020], which provides some interesting evolutionary predictions.
+
+## Conservation of sex differences in chromatin structure
+
+Our results from _musculus^MSM^_ and _musculus^PWD^_ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
 These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
@@
 Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
-This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier paper?
+This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier?
 Sera?) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
@@
 As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
-From empirical data in mice [@petkov2007, @lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
+From empirical data in mice [@petkov2007], [@lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
 The signal for interference is thought to be mediated though the SC, which is built on top of the chromosome axis [@zicklerKleckner2015] regardless of the loop size.
@@
 
-<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]>
-
-### Evolution of interference strength in males
-
-```{r evo.interference, include=FALSE, echo=FALSE}
-
-#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)
-#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)
-
-#**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**
-
-#5. Interference evolution shows a positive correlation with gwRR
-#5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
-#5b.
-Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
-
-# consequence of rapid evolution of normalized interference strength,
-# in the 3 strains -- larger linkage blocks?
-# but what if the rec pattern is repeated?
-```
-
-An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (cite, Schumer).
+<!-- The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019] -->
+
+## Evolution of interference strength in males
+
+An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (Schumer).
 Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
-**(new, veller paper -- in addition to the total gwRR, the location of crossovers (rbar), should be considered in this dynamic)**
 
@@
 
-### Future steps
-
-There are still many un-known aspects regarding heterochiasmy (Lenormand) , these novel results suggest - we hope ppl we follow up .  (fitness - evolutionary dynamics - role in linked selection)
-
-We make three suggestions for future steps in studying recombination variation.
+## Future steps
+
+There are still many un-known aspects regarding heterochiasmy (@lenormand2016) --- that would benefit from the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
+A goal of such merging of models could be to connect empirical findings across scales.
+We make two suggestions for future steps in studying recombination variation.
 First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
-
 Second certain systems can be leveraged to overcome limitations inherent in some approached.
@@
 
-Third we encourage the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
-A goal of such merging of models could be to connect empirical findings across scales.
-
 # References

### 2020-05-05 doc/MS_v1.Rmd
work on drafts before sending to bret

@@
 ---
-title: 'MS draft'
+title: 'Evolution of Sex-Specific Recombination Rate'
 date: '`r Sys.Date()`'
@@
   word_document:
+    toc: yes
+    toc_depth: '5'
   pdf_document:
@@
 
-#library(dplyr)
+library(dplyr)
 library(plyr)
@@
 library(kableExtra)
+library(xlsx)
 
@@
 
-To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of of the SC (SYCP3).
+To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
 Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
@@
 
-**M4.2 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
+**M5 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
 
@@
 Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
-Hand measures of SC/pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
+Hand measures of pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
 
@@
 To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts (Figure X).
-
-The **DNA CrossOver** algorithm isolates single straightened bivalents/chromosomes shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
+Mouse means were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
+
+The **DNA CrossOver** algorithm isolates single straightened bivalents shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
 This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
@@
 After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step ([@peterson2019])).
+The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
 From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
@@
 
-<!-- male section -->
-
 ```{r male.strain.table, echo=FALSE}
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
+```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
 #write the full table to a supplemental table
@@
 #kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+
+#write excel
+#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table.xlsx", sheetName = "Sheet1", 
+#  col.names = TRUE, row.names = FALSE, append = FALSE)
+
 ```
@@
 #write table
-#write.xlsx(DMC1table.main99,file = "Results_Table.xlsx",sheetName = "DMC1.results", append = FALSE, row.names=F)
+write.xlsx2(DMC1table.main99, file = "DMC1.Results_Table.xlsx",sheetName = "DMC1.results", append = FALSE, row.names=F)
+
 ```
@@
 Green foci show DMC1.
-B)	Distribution of DMC1 counts per cell  by strains.
-(1 mouse was used for each strain)", eval=FALSE}
+C)	Distribution of DMC1 counts per cell  by strains.", eval=FALSE}
 
@@
 
-To examine the connection between evolution of the genome-wide recombination rate and changes to the recombination landscape, we used an image analysis pipeline to measure properties of single bivalents ([@peterson2019]).
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this dataset, isolation rates per cell range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
-From the total set of cell images, `r formatC(sum( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ) ), , format="d", big.mark="," )` bivalent objects were isolated by the image analysis software.
-After a human curation step ([@peterson2019]), `r formatC(sum(  length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ) ), , format="d", big.mark=",")`  single-bivalent observations remained.
-We assume that the isolation of bivalents within cells is unbiased.
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure X).
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure 3).
 The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
-low recombining strains (Figure X).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` ( _musculus^SKIVE^_ ) in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
+low recombining strains (Figure 3).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
 Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
@@
 
-In many mammalian species, the synaptonemal complex (SC) is longer in females (@cahoonLibuda2019, @sardell_sex_2020).
+In many mammalian species, the synaptonemal complex (SC) is longer in females [@cahoonLibuda2019, @sardell_sex_2020].
 Yet, the majority of these observations come from species with female-biased heterochiasmy.
@@
 
-## Q1. Sex Differences in CO Interference (IFD)
+## Sex Differences in CO Interference
 
@@
 A marginally significant difference between the sexes (t-test; p = `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain, *domesticus^G^* (t-test without domesticus^G^; p = `r Q1.IFD.raw_t.test_G$p.value`).
-This result indicates that females and males exhibit a similar level of interference when it is measured in physical (SC) units.
+This result indicates that females and males exhibit a similar level of interference when it is measured in physical SC units.
 
@@
 
-Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1 - LRT **random**: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`  glm *M2* `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`  *M3* `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` glm *M2*, p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`,  *M3* p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
 When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in SC length (chromatin compaction) using IFD^norm^ indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical (SC) scale.
-
-Visual comparison of interference pattern is aided by Figure X, which combines the total length of distance between two foci, but also their normalized locations along the chromosome.
+In summary, controlling for differences in SC length (chromatin compaction) using IFD^norm^ indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical scale.
+
+Visual comparison of interference pattern is aided by Figure 4C, which combines the total length of distance between two foci, but also their normalized locations along the chromosome.
 Two main patterns of sexual dimorphism emerge across all strains, (Supplemental figure X).
@@
 Suggesting a wider range of normalized IFD lengths compared to males (bottom triangle).
-Male IFDs display a stronger chromosome end-locatization of the second foci, (clustered near the 90* corner).
+Male IFDs display a stronger chromosome end-locatization of the second foci (Figure 4).
 
@@
 The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M4.2).
+In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M5).
 
 Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-**SC length is expected to be positively associated with genome-wide recombination rate because of the loop-axis structure** (zickler Kleckner 1999, Mercier 2015).
-Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome (on chromosomes of  a finite length, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
+SC length is expected to be positively associated with genome-wide recombination rate because of the connection between DSB formation and the loop-axis structure [@zickler1999, @mercier2015].
+Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome with a finite length.
+Thus, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
 
@@
 
-## Q2 SC Length
+## Conservation of Mean SC length
 
@@
 #pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-
 ```
@@
 
-Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
-Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).
-**Bret's note on positive correlation, 'A correlation across mice, across strains, or both?**
-
+Confirming the basic predictions, there is a positive correlation between mouse mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
+Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure 5).
 While high-recombination strains have significantly greater mouse mean total SC area in the total dataset (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p =  `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-**the values above are diferent between editions**
 
@@
 
-**the values between versions are slightly different**
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = _musculus_ `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, _molossinus_ p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
+In a general linear model with total SC as the dependent variable, two subspecies effects are significant (_musculus_ p =  `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, _molossinus_, p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
 In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
@@
 
-## Q2.1CO rec landscape evolution is decoupled from gwRR evolution
+## Single Crossover position is decoupled from genome wide recombination rate evolution
 
@@
 The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
-**While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable (Figure X)**, *this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate.*
-<!-- bret asks if the bolded section is correct. are there sig strain effects for WSB and MOLF?, bret asks if italised statement has been tested statisticaly or in what sense am I discribing this -->
-
-## Q2 Evolution of interference is associated with genome wide recombination rate evolution
+While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.
+
+## Evolution of interference is associated with genome wide recombination rate evolution
 
@@
 Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS$p.value`).
-This pattern is confirmed by separate comparisons within _musculus_ (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-<Bret's comment, these p values are not very significant.> Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
+
+This pattern is supported by slightly significant p values from comparisons within _musculus_ (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
+Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
 
@@
 
-Differnces in IFD can also be noted from the triangle IFD plots (Figure X).
-We note that the three high-recombining strains show more clustering of IFD observations near the area of the plot which indicates longer IFDs.
+Differnces in IFD can also be noted from the triangle IFD plots (Figure 6).
+The three high-recombining strains show more clustering of IFD observations near the area of the plot which indicates longer IFDs.
 The remaining strains have more observations short IFD observations near the diagonal.
@@
 cataneus* (Geraldes et al 2008, [@geraldes2011]).
-Models for the aggregate recombination metics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019) -- motivating increased study of recombination rates and natural populations.
-4
+Models for the aggregate recombination metics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019), motivating increased study of recombination rates and natural populations.
 
@@
 Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-<We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
-
-A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion ([@laneKauppi2019]).
+The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
+
+A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion [@laneKauppi2019].
 Genetic variants will be selected in a manner to optimize the process of bi-orientation and prevent SAC from delaying cell cycle or triggering apoptosis.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014]), which are connected to conserved features of gametogenesis; centrosome spindle (@So2019, @dumontDesai2012) and cell volume [@kyogoku2017].
+Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014], which are connected to conserved features of gametogenesis; centrosome spindle [@So2019, @dumontDesai2012] and cell volume [@kyogoku2017].
 The more stringent SAC in spermatogenesis will be more effective at removing genetic variants which interfere with bi-orientation compared to females.
@@
 Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
-These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005]?, compared to spermatocytes.
+These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes.
 This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
@@
 Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-Centrosomes spindles are in spermatogensis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization (@manandhar2005centrosome).
+Centrosomes spindles are in spermatogensis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization [@manandhar2005centrosome].
 This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
@@
 Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs [@laneKauppi2019], subramanian and Hochwagen , @dumontDesai2012].
+The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (@laneKauppi2019, @subramanian2014, @dumontDesai2012).
 The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
@@
 Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
-This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier?
-Sera?) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
+This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier, Sera) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
 

### 2020-05-06 doc/MS_v1.Rmd
fixed annonying typos

@@
 The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp) and rederived by the UW-Biotech centre (Madison UW).
+(https://en.brc.riken.jp) and rederived by the UW-Biotech center (Madison UW).
 The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
-We use abriviated superscript notation for the remainder of the paper (e.g.
+We use abbreviated  superscript notation for the remainder of the paper (e.g.
 _musculus^WSB^_).
@@
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
-Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
-Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
+The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvenile mice aged between 12 to 15 days were used for DMC1 quantification.
+Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embryonic days and 0 to 48 hours old respectively.
+Whole testes and decapsulated ovaries were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
 Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
@@
 
-To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
+To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
 Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
@@
 The three high-recombining strains identified in this study have connections to hybridization.
-Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the european house mouse hybrid zone.
-Notably *musculus^SKIVE^* comes from a yonger section of the hybrid zone with a nuclear genome from *M.
+Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the European house mouse hybrid zone.
+Notably *musculus^SKIVE^* comes from a younger section of the hybrid zone with a nuclear genome from *M.
 m.
@@
 cataneus* (Geraldes et al 2008, [@geraldes2011]).
-Models for the aggregate recombination metics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019), motivating increased study of recombination rates and natural populations.
+Models for the aggregate recombination metrics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019), motivating increased study of recombination rates and natural populations.
 
@@
 
-To place our results in the broader context of recombination rate evolution we compare our main resuts to expectionat under three modifier models and two functional and cell physiology models (table X).
-While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across exsisting models.
-The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome ocillatory movement (COM) model was built to describe positive interference.
+To place our results in the broader context of recombination rate evolution we compare our main results to expectations under three modifier models and two functional and cell physiology models (table X).
+While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across existing models.
+The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome oscillatory movement (COM) model was built to describe positive interference.
 This spindle based selection model was developed to explain the results resented in this manuscript.
@@
 Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
+The SAC prevents aneuploidy by ensuring all bivalents are correctly attached to the microtubule spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
 
@@
 
-Regardless of whether these hypotheses are true, Our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
+Regardless of whether these hypotheses are true, our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
 Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
@@
 Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-Centrosomes spindles are in spermatogensis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization [@manandhar2005centrosome].
+Centrosomes spindles are in spermatogenesis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization [@manandhar2005centrosome].
 This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
@@
 Faster spermatogenesis may select for synchronization of the separation homologs (cite).
-While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, espicially for those with dicyate arrest [@lee2019].
+While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dicyate arrest [@lee2019].
 
@@
 These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
-Chromatin compaction could explain variation in heterochiasmy, which is driven by the disinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
+Chromatin compaction could explain variation in heterochiasmy, which is driven by the distinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
 Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
@@
 
-We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphse.
+We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphase.
 The evolution of genome wide recombination rates in our study is driven by a transition of the majority of

### 2020-05-08 src/SetupData.Rmd
saving while debugginh
formating the setupData files

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-output: html_document
+output:
+  html_document: default
+  pdf_document: default
 ---
@@
 
+- try calq correlation of MLH1 counts across duplicates
+
 - format the DF for applying to Mixed models
-
-- rm all of BD data?
 
@@
 library(here)
-
+library(xlsx)
 #knitr::opts_chunk$set(echo = FALSE)
@@
 
-meta.data = read.csv(here("data", "/clean.Meta.Data.txt"),sep = "\t", header=TRUE)
-
+meta.data = read.csv(here("data", "/clean.Meta.Data.txt"), sep = "\t", header=TRUE)
 #remove all MAD (peromyscus) from meta.data
-meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]
-
-MLH1_data = read.csv(here("data", "/AnonData_formated.csv"), header=TRUE)
+#meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]
 
@@
 
+#MLH1_data = read.csv(here("data", "/AnonData.csv"), sep="\t", header=TRUE)
+#read.csv was breaking things
+MLH1_data = read.xlsx2(here("data", "/AnonData_formated.xlsx"), header=TRUE, 1)
+
+#formating the original anon file
+#MLH1_data = read.xlsx2(here("data", "/AnonData.xlsx"), header=TRUE, 1)
+#MLH1_data$fileName <- (MLH1_data$Original.Name)
+
+#MLH1_data <- add_mouse(MLH1_data)
+#MLH1_data <- add_category(MLH1_data)
+#MLH1_data <- add_strain(MLH1_data)
+#MLH1_data <- add_sex(MLH1_data)
+#MLH1_data <- add_subsp(MLH1_data)
+#MLH1_data <- add_species(MLH1_data)
+#write csv to MLH1 file
+#write.xlsx2(MLH1_data, "AnonData_formated.xlsx", sheetName="Sheet1",
+#  col.names=TRUE, row.names=FALSE, append=FALSE)
 ```
@@
 original_DF = MLH1_data
-original_DF$fileName <- (original_DF$Original.Name)
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-original_DF <- add_mouse(original_DF)
-original_DF <- add_strain(original_DF)
-original_DF <- add_category(original_DF)
-original_DF <- add_sex(original_DF)
-
-#check for duplicated data, make duplicated data DF for repeatability
-#anyDuplicated(original_DF$fileName)#219
-#orig_DF <- original_DF[duplicated(original_DF$Original.Name),]
-#these just produce 1 set of the duplicated
 
@@
 original_DF_duplicates <- original_DF[original_DF$fileName %in% n_occur$Var1[n_occur$Freq > 1],]#2 is the max freq of duplicates,
+
 #use the above DF for quantifying repeatability in the Main_report.Rmd
@@
 original_DF_duplicates$unique.key <-  paste(original_DF_duplicates$fileName, original_DF_duplicates$Batch, sep = "_")
-#anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch ... really 2?
+
+anyDuplicated(original_DF_duplicates$unique.key)#31 duplicated fileName_batch and batch number --
 
@@
 
+#calculate the correlation across duplicates. First issue is to match up and organize the matching 
+#for each fileName - make cols, for MLH1-1 and MLH1-2
+#remove Xs, for fileName with 
+original_DF_duplicates <- original_DF_duplicates[!grepl("X", original_DF_duplicates$X),]
+#make a new DF -- with one fileName and the MLH1-1 and MLH1-2
+#  (duplicate / dostinct fileName )
+
+#keep just 1 of duplicates from MLH1_data
+#MLH1_data <- MLH1_data[duplicated(MLH1_data$fileName),]
+MLH1_data <- MLH1_data %>% distinct(fileName, .keep_all = TRUE)
+
 ```
+
+`r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.
 
@@
 MLH1_data$fileName <- (MLH1_data$Original.Name)
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
 
@@
 
-MLH1_data <- add_mouse(MLH1_data)
-MLH1_data <- add_category(MLH1_data)
-MLH1_data <- add_strain(MLH1_data)
-MLH1_data <- add_sex(MLH1_data)
-MLH1_data <- add_subsp(MLH1_data)
-#add species
-MLH1_data <- add_species(MLH1_data)
 ```
@@
 
-#show_col(hue_pal()(5))
-
-#add Rec group,
-
-#Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-#                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
-#                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-#                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
 ```
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_4.26.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.5.20.RData")
 

### 2020-05-08 Chap3_MLH1var.rmd
starting work on last ms outline

--- 2020-05-05 doc/MS_v1.Rmd
@@
 ---
-title: 'Evolution of Sex-Specific Recombination Rate'
-date: '`r Sys.Date()`'
-author: "April L. Peterson, Bret Payseur"
-output:
-  word_document:
-    toc: yes
-    toc_depth: '5'
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
-  html_document:
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-always_allow_html: yes
-bibliography: refs_Main_Report.bib
+title: "betweenCell.Var"
+author: "April Peterson"
+date: "5/5/2020"
+output: html_document
 ---
 
-```{r set.up, echo=FALSE, include=FALSE}
+load all the file set ups
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 
@@
 
-#Intro
-
-#Material Methods Draft v1
-
-### Mouse Husbandry
-
-To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
-The _Mus musculus_ strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp) and rederived by the UW-Biotech centre (Madison UW).
-The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
-We use abriviated superscript notation for the remainder of the paper (e.g.
-_musculus^WSB^_).
-All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
-A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (G) is maintained at UW Veterinary school facilities.
-
-### Tissue Collection and Immunohistochemistry
-
-The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvillne mice ageed between 12 to 15 days were used for DMC1 quantification.
-Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embyronic days and 0 to 48 hours old respctively.
-Whole testes and decapsulated overaires were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
-Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
-Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a Photoflow wash (Kodak, cite).
-Slides were stored at -20*C if not stained immediately.
-
-To visualize the strucutre of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
-Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
-The staining protocol was based on that in [@anderson1999] and [@koehler2002].
-Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
-
-Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
-The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
-Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200)  were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
-Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
-
-### Image Processing
-
-Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
-Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.
-
-### Analysis
-
-To estimate the numbers of percursor sites which may lead to crossovers, the mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
-SC morphology and CREST/centromere foci number was used to stage spermatocytes, in early and late zygotema stages.
-
-**MLH1** foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
-With scores of 1 and 2 having the highest quality.
-Cells and coded as a score of 5 were excluded from the final analysis.
-The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
-MLH1 foci located on the XY in spermatocytes were excluded from this study.
-
-**M1 Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
-
-**M4 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
-
-**M5 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
-
-A series of mixed models and general linear models were used as an analetical framework.
-We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
-Strain was coded as a random effect and tested using exactRLRT() to reflect a random sample of natural genetic variation from each subspecies.
-The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
-
-This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
-The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
-The general linear models, **M4** were used to analyze sex specific data sets.
-
-Our goal was to design an analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
-sex, strain subspecies) and can be applied to multiple (meiotic traits) dependent variables.
-In addition to the MLH1 count variables for individual mice, mean MLH1 count per cell, variance and coefficient of variance for MLH1 count per cell, we analyzed SC based features: mouse average of total SC per cell, mouse average of single SC length, mouse mean IFD^norm^ and IFD^raw^ for double crossover chromosomes, and the average position of crossover for single crossover bivalents.
-
-These traits were choosen because they are markers of phenomena in the recombinaiton pathway and provide more detailed information about the recombination landscape.
-In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
-Normalized measures were calculated by the position divided by the length of the SC for that single bivalent.
-
-### Single bivalent characterization
-
-To characterize the SC length of individual chromosomes, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
-Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
-Hand measures of pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
-
-The **total SC** area per cell was quantified from the pachtyene cell images.
-To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-
-The **DNA CrossOver** algorithm isolates single straightened bivalents shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
-
-A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
-From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` bivalent objects were isolated by the image analysis software.
-After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a **reduced data** set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
-
-# Results
-
-```{r FOR.PLOT.organizing.data, echo=FALSE, warning=FALSE}
-#set the order of another column, based on another variable. consistant order for ploting
-
-#apply this before ploting
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, category, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-#use the above line for when order is out of wak
-
-#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
-AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-
-var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
-  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
-
-```
-
-```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
-#DF.HetC.MixedModel.HQ
-#this blck is duplicated in setupData.rm
-
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#AP_mouse_table excluide Qual5 cells
-DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
-DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
-
-#main df with variables coded as chr
-DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
-
-all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
-                 .(sex, strain),
-                 summarize,
-                 all.cells = sum(Ncells)
-)
-
-```
-
-```{r table.for.plot, echo=FALSE, warning=FALSE}
-#why do I add in a different table?
-#this move this to the setup file
-#MLH1_data  %>% filter(quality < 5)  -- low quality cells removed
-
-#pooled
-strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
-                        Nmice =length(unique(mouse)),
-                        Ncells  =length(nMLH1.foci),
-                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(nMLH1.foci, na.rm = TRUE),
-                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
-                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
-                       #Standard error (mice vs cells)
-                        se  = round(sd / sqrt(Ncells), 3)
-
-                       #range_mouse_means = range()
-                   #consider adding crhm proportions?
-    )
-
-strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
-                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")
-
-#adjust order of plot here
-strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
-     "MSM", "MOLF",   
-  "PWD", "SKIVE",  "KAZ",
-  "WSB","G","LEW","PERC",
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-```
-
-```{r Fig1.top, echo=FALSE, warning=FALSE, message=FALSE}
-
-Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.y=element_blank(),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
-#remove y labfor male
-
-   #segments
-  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
-  #Musc
-  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)
-  #mol
-#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
-  #cast
-#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
-  #outgroups
- # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
-  #label
-
- #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
-   #        y = c(18.5,18.5,33,18.5, 18.5),
-
-  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
-    #       fontface = 'italic', hjust = 0.5, size=3.5)
-
-```
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
-
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
-  theme(legend.position="none",
-
-        axis.title.y =element_blank(), 
-      axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+geom_hline(yintercept = 20)
-
-#+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
-#try customize the plot size
-
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
-
-   theme(legend.position="none", axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),  axis.title.y = element_blank(), 
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+geom_hline(yintercept = 19)
-
-#horizontal line = 20
-
-```
-
-```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
-B) Female specific MLH1 count distributions.
-C)	Male specific MLH1 counts per cell.', message=FALSE}
-
-Figure1.main.plot <- Fig1_v1_strain / (female.specific.plot + male.plot )
-
-Figure1.main.plot + plot_annotation(
-  title = '',
-  subtitle = '',
-  caption = ''
-) + plot_annotation(tag_levels = 'A')
-
-```
-
-```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
-#str(AP_mouse_table_w.Ages)
-
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-#remove sex.x == na, and strain.x=na or F1
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
-
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
-#plot variance and mean (facet sex)
-
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
-
-VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
-
-cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
-
-#str(AP_mouse_table.HQ)
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
-#[1] 0.5750399
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
-#0.321601
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
-# 0.3515544
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
-#-0.03713536
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
-# p-value = 8.623e-09
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
-
-#remove mice with fewer than 10 cells
-
-AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
-
-VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
-
-cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
-
-cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
-    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-
-#why did I male just a female table?
-female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-
-female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length( nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean( nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv( nMLH1.foci),
-                        var = format(round(   var( nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd( nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-                        #quality?
-)
-```
-
-## Genome-wide recombination rate estimates for both sexes
-
-```{r MLH1.summ.stats, echo=FALSE}
-
-#strain.table.HQ_Fig.og  all strains
-#strain.table.HQ_Fig  extra strains removed
-
-main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
-
-main.table <- add_subsp(main.table)
-main.table <- add_species(main.table)
-
-#main.table %>% filter(sex == "female")
-#main.table %>% filter(sex == "male")
-
-#G.HETC.val #1.2
-#LEW.HETC.val  #1
-#WSB.HETC.val  #1
-#KAZ.HETC.val  #1.1
-
-#SKIVE.HETC.val # 0.96
-#PWD.HETC.val  #0.88
-#MSM.HETC.val #0.9
-
-#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
-
-#strain.table.HQ_Fig
-
-#sum( (main.table  %>% filter(sex == "male"))$Ncells)
-#the kable version is bulkier than the other versaion
-
-```
-
-```{r MLH1.sum.stats.Bret.edits, echo=FALSE}
-
-AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
-#total mice
-#length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )
-
-#average cell number for spermatocytes and oocytes
-#mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
-#mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
-```
-
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M.
-m.
-domesticus*, *M.
-m.
-musculus* and *M.
-m.
-molossinus*.
-Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ_target)$mouse )` mice were quantified from an average of `r mean( (AP_mouse_table.HQ_target %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
-
-Graphical comparisons between the two sexes reveal several patterns (Figure 1A).
-First, recombination rate differs between females and males in most strains.
-Second, the direction and magnitude of heterochiasmy varies among strains.
-Although the majority of strains show higher recombination rates in females (following the pattern in laboratory mice), two musculus strains and one molossinus strain exhibit male-biased heterochiasmy.
-Hence, relative recombination rates in the two sexes are evolutionarily labile.
-
-The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
-Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
-
-Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C).
-Female recombination rates are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure 1B).
-In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).
+plot the variances
 
@@
     )
-#F.strain.table
 
-#F.strain.table
-
-# I don't know why most of the table is blank
 ```
 
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=TRUE}
 
@@
 ```{r male.strain.table, echo=FALSE}
+
 M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
@@
 musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.male.VAR.plot
 
 Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.male.VAR.plot
 
 Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.male.VAR.plot
+```
+
+go over the main patterns (from the model framework)
+
+Bret's suggestions
+
+On the question of variance in MLH1 count, I agree that it is crucial to reduce experimental noise to find a signal.
+
+I would suggest the following approach:
+1.  Consider only high-quality cells.
+2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.
+3. Randomly permute cells among strains, keeping the structure of the dataset otherwise the same.
+4. For each permutation, compute the among-strain variance in the MLH1 count variance (separately for each sex).
+Compute the among-strain variance in the MLH1 count variance (separately for each sex) in your original data.
+Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).
+
+```{r bret.filters, echo=FALSE}
 
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
-#write the full table to a supplemental table
-#show the head (or shortened)
-#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-#kable(strain.table.HQ_Fig, "latex")
+# Identifying mice with largest sample sizes
 
-df.strain.table <- strain.table.HQ_Fig
-#kable(df.strain.table)
-#pander(df.strain.table)
-
-#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-#write excel
-#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table.xlsx", sheetName = "Sheet1", 
-#  col.names = TRUE, row.names = FALSE, append = FALSE)
-
-```
-
-## Partitioning variation in recombination rate
-
-To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
-We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r 1/1000000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
-
-After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and _domesticus^G^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
-
-We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4; see methods).
-In the male dataset, three strains significantly affect recombination rate: _musculus^PWD^_ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect =
-`r sum_MLH1.male.glm.M4$coefficients[8]` foci), and 
-_musculus^SKIVE^_ (glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
-
-These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and _musculus^PWD^_ (p= `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
-Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
-
-## Within mouse variance in CO count per cell
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE}
-
-MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-MLH1var_M2_sum <- summary(MLH1var_M2)
-
-#anova(MLH1var_M2)  
-#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
-
-MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M2)
-#anova(MLH1cv_M2)
-#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
-
-#Q12 var
-MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-sum.MLH1var_M2_Q12 <- summary(MLH1var_M2_Q12)
-#anova(MLH1var_M2_Q12)  
-#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M2_Q12)
-#anova(MLH1cv_M2_Q12)#sex large F
-#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-sum.glmMLH1Var_M3 <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-#MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-
-Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
-To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count per cell as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < `r 1/1000000`) and general linear models (M2) (p = `r MLH1var_M2_sum$coefficients[,4][9]`) and M3 (p = `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`).
-
-In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
-The results are similar: sex is the strongest effect (M1 p < `r 1/1000000`; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
-When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
-These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
-
-<LmerMLH1_M1_Q12_VAR_results> <Bret's comment for below, are there differences between the curated and the full dataset? -- Q1 and full data set>
-
-## Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
-
-```{r DMC1.setup, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "early Z"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "late Z"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-```
-
-```{r DMC1.table, echo=FALSE, message=FALSE}
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-
-#attach this mini DF to the DMC1 data
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main88 <- merge(DMC1.target.strains, DMC1table, by="category")
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-#DMC1.mean.table
-DMC1table.main99 <- subset(DMC1table.main, select= -c(stage.x, var.MLH1, stage.y, mean.count.x, mean.count.y))
-
-DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
-
-colnames(DMC1table.main99) <- c("category", "mean MLH1",    "ncells",     "mean DMC1", "MLH1:DMC1 ratio", "ncells",     "mean DMC1", "MLH1:DMC1 ratio")
-
-#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-#  kable_styling() %>%
-#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#pander is better for word
-#pander(DMC1table.main99, row.names = F)
-
-#write table
-write.xlsx2(DMC1table.main99, file = "DMC1.Results_Table.xlsx",sheetName = "DMC1.results", append = FALSE, row.names=F)
-
-```
-
-```{r DMC1.boxplot.plot, echo=FALSE, warning=FALSE}
-
-#change order
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter() +
-
-    labs(y = "DMC1 count per Cell")+
-
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.x = element_blank())
-```
-
-```{r Main.plot.DMC1, echo=FALSE, fig.width=6.5, fig.height=5, fig.cap="*Figure 2 male DSB estimates* A) Example early zygotene spermatocyte spread, B) Example late zygotene spermatocyte spread.
-Green foci show DMC1.
-C)	Distribution of DMC1 counts per cell  by strains.", eval=FALSE}
-
-png.DMC1.early.cell <- readPNG('~/MLH1repo/doc/figureFiles/cell.examples/DMC1.early/good.early.cell.png')
-png.DMC1.late.cell <- readPNG('~/MLH1repo/doc/figureFiles/cell.examples/DMC1.late/good.late.cell.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-#EMPTY GGPLOT
-Early.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(png.DMC1.early.cell) + theme(axis.title.x=element_blank(),
-                         axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-Late.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(png.DMC1.late.cell) + theme(axis.title.x=element_blank(),
-                         axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-three.DMC1.plot <- { Early.cell.obj + Late.cell.obj  } / DMC1.boxplot + plot_layout(widths = 1)
-
-#+ plot_layout(ncol = 2, widths = c(.65, .65), heights = c(.65, .65))
-#plot_layout(nrow = 2, widths = c(1), heights = c(.45) )
-
-#plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or #labels="AUTO"
-
-three.DMC1.plot + plot_annotation(
-  title = '',
-  subtitle = '') + plot_annotation(tag_levels = 'A')
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-#anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage #!= "P",],method = "t.test") %>% mutate(y_pos = 300)
-
-#DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-#  geom_boxplot(position=position_dodge()) + 
-#  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-#  ggsignif::geom_signif(
-#    data=anno_df_stage, 
-#    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-#    manual=TRUE
-#  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-#        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
-In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains ( _musculus^KAZ^_ , _domesticus^WSB^_ , and _domesticus^G^_ ) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_ ).
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p < `r 1/1000000`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = `r ttest.HighLow.Z.pval`).
-
-<ttest.HighLow.L.pval>
-
-After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
-This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
-@cole2012 suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
-This result, combined with the large difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
-
-<!-- The number of DSBs increase in zygotema, then deminish as the effects of crossover homeostasis are implemented (Cole).
-(co homeostasis also indecates dsb ) Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links connects the CO number variation to 'pre' DSB formation fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
-(this study only examined one stage of early prophase for DMC1 counts -->
-
-<!-- the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 C3H, 7.3 B6 -->
-
-## Evolution of genome-wide recombination rate is reflected at the single chromosome level
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-#use these two metrics
-#cell.level.BivData_table$boxs.IDd 
-#cell.level.BivData_table$nboxes_passed
-
-#calq the mean rate by category
-
-cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
-
-algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
-                        n_obs = length( unique(fileName) ),
-                         mean_pass.rate = mean(pass.rate)
-                         )
-
-cell.level.BivData_table <- add_sex(cell.level.BivData_table)
-
-algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
-                        n_obs = length(unique(fileName)),
-                        mean_pass.rate = mean(pass.rate)
-                         )
-
-#showing the table..
-#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-```{r chrm_prop_data, echo=FALSE, warning=FALSE}
-#shut off eval for now because of wd / path error
-#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#try adding the BivData version of the proportions
-
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
-
-                 n0CO = sum(n0CO, na.rm = TRUE),
-                  n1CO = sum(n1CO,na.rm = TRUE),
-                  n2CO = sum(n2CO,na.rm = TRUE),
-                  n3CO = sum(n3CO,na.rm = TRUE),
-                  n4CO = sum(n4CO,na.rm = TRUE),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#this is the table that needs melting
-#but is also easy to redo by hand
-
-#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
-```
-
-```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
-
-#add subspecies to these values¬
-melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
-
-#add the subsps
-melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
-
-melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
-
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))+theme(legend.position="none")
-
-#need a second DF / the first df
-#+ annotate() text 
-#annotate()text
-
-#proportion_stats_category,  tot_biv, ncells
-#add number of cells and bivalents in the corner
-
-#chrm.class.barplot
-```
-
-```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
-
-#prop.test(x, y)  x number of successes, y total  
-#strain sums of n
-#look up continuity correction
-
-#mouse.level.BivData_table <-- has more male strains
-#Mouse.Table_BivData_4MM <- only sex matched strains
-
-#perform test across strain 
-#removed all lines with no hand.foci.count entry
-Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
-
-                  nbiv = sum(tot.biv ),              
-                 n0CO = sum(n0CO ),
-                  n1CO = sum(n1CO ),
-                 n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO ),
-                 n4CO = sum(n4CO )
-                          )
-
-Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
-
-for(g in 1:length(Chrm.prop_strain_table$nbiv)){
-
-  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
-                                     Chrm.prop_strain_table$n1CO[g],
-                                     Chrm.prop_strain_table$n2CO[g],
-                                     Chrm.prop_strain_table$n3CO[g] ))  
-}
-
-Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
-#there are still some wacky entries -- but much fewer than before
-
-chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
-
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
-#the pvalues for 0 are not sig dif
-#the p values for 1 and 2CO are sig diff
-
-#seperate by subsp
-chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
-chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
-chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
-
-Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
-Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
-Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
-Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
-#WSB has fewer 2COs than other 3 strains
-
-Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
-Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
-Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
-#KAZ has slightly more 1CO than the other two
-
-#high.female_G.LEW.
-chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
-
-high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
-high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
-high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
-
-#Low.female.G.LEW.
-chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
-
-low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
-low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
-low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
-
-#all females
-
-chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
-
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
-#the pvals for these 3 class are sig dif
-
-#seperate by subsp
-chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
-chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
-chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
-
-Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
-Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
-Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
-#non of the Dom proportions are different
-
-Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
-Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
-Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
-#all v different
-
-Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
-Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
-Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
-
-#high vs low
-chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
-                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
-                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
-
-chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
-chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
-
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
-
-Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
-Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
-Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
-#the low strains do vary slightly in their 1CO and 2CO proportions
-
-#includes SKIVE
-Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
-Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
-Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
-#all very sig p values
-
-High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
-#these proportions are v different
-
-#no SKIVE
-High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
-High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
-High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
-#MSM and PWD only slightly sig dif 2CO
-
-#MSM-MOLF
-MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-#PWD-KAZ
-PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.MSM.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "MSM") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "MSM") )$nbiv )
-
-```
-
-```{r Chrm.props.plot2, echo=FALSE, fig.width=8, fig.cap="*Figure 3* Chromosome Class Proportions. proportions calculated from pooled single bivalent data by strain."}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
-
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
-
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
-
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-#no filename for add mouse, strain subsp
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#can't apply function use this
-chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
-             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
-                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
-                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
-
-          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
-              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
-
-                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
-                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
-            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
-                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
-                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
-                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
-                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
-                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
-
-             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
-               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
-
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
-                                                                 "Musc",
-                                                                 "Mol",
-                                                                 "Spretus","Spic",
-                                                                 "other") )
-#ggplot
-
-chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),
-                             aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(legend.position="top", legend.title = element_blank(),
-                                                                                                           axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
-
-chrm.class.barplot
-kk <- addSmallLegend(chrm.class.barplot)
-
-```
-
-```{r Chrm.Prop.misc.calulations, echo=FALSE}
-
-total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-
-#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
-#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
-
-#MELT_Chrm.Proportion$Rec.group <- ifelse(grepl("PWD male", MELT_Chrm.Proportion$category), 1, 
-#                              ifelse(grepl("MSM male", MELT_Chrm.Proportion$category), 1,
-#                                ifelse(grepl("SKIVE male", MELT_Chrm.Proportion$category), 1, 0)))
-
-#Chrm.prop.male.low <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 0)
-#Chrm.prop.male.HIGH <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 1)
-
-#strain av tables
-#chrm.prop.male.high1
-#chrm.prop.male.low
-```
-
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure 3).
-The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
-low recombining strains (Figure 3).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
-Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-_musculus^KAZ^_ p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_ p = `r MSM.MOLF.2CO.prop.test$p.value`).
-Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
-We address two main questions.
-First, which traits are sexually dimorphic?
-Second, which traits differ between males from high-recombination vs.
-low-recombination strains?
-
-## Sex Differences in Meiotic Traits
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-#nested boxplots, pooled Bivalents
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
-```{r Q1.SC.show.short.biv, echo=FALSE}
-
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  theme(legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=9), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9))+labs(y = "Mouse Mean Short SC Length")
-
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
-#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
-#1 male, 3 females measures
-#try to merge the manual - whole cell observations (at least 2 males)
-
-```
-
-```{r Q1SC_short.biv_dif, echo=FALSE}
-#calq the different in short.biv means across sexes
-
-#WSB
-Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
-
-#LEW
-Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
-
-#G
-Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
-
-#PWD
-Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
-
-#KAZ
-Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
-
-#SKIVE
-Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
-
-#MSM
-Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
-
-#this bivData ratio is the same as mouse average
-Q1SC.BivData.short.biv_DIFF_MSM <- mean( (real.short.bivData %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (real.short.bivData %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
-
-```
-
-```{r Q1.tot.SC.plot, echo=FALSE, warning=FALSE}
-# fig.width = 4, fig.height=2.5
-
-Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
-
-Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
-   "WSB","G","LEW","PERC", 
-  "PWD", "SKIVE",  "KAZ",
-  "MSM", "MOLF"), order=T )
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#order the strains
-Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
-  arrange(strain) %>%
-  mutate(mouse = factor(mouse))
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#change this plot to look like the short bivalent plot -- boxplots + mouse points
-Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
-  scale_color_manual(values=colors_of_strains)+
-
-  facet_wrap(~subsp, scales="free", ncol = 3)+ 
-  geom_jitter()+ geom_boxplot(alpha=.2)+ 
-  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.ticks.x=element_blank(),axis.text.x = element_text(size=9), legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10)  )+
-  labs(y = "Mouse Mean Total SC Length")+
-  ylim(c(1000,2600 ))+ggtitle("")
-
-```
-
-```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8, fig.cap= 'caption'}
-
-two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }
-
-#two.Q1.SC.plots + plot_annotation(
-#  title = '',
-#  subtitle = '') + plot_annotation(tag_levels = 'A')
-
-```
-
-```{r Q1.total_SC.table, echo=FALSE}
-
-totSC.strain.av.table <-  totSC.strain.av.table[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
-
-#head of table
-#see the DMC1 code
-
-#kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#add variance
-#write.xlsx(totSC.strain.av.table, file = "Results_Table.xlsx", sheetName = "Total.SC.results", append = TRUE, row.names = F)
-
-```
-
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#ALL
-Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)
-
-#dom
-Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#Musc
-Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
-#cell level total SC
-#MLH1.merge.Skel   $skel_size
-#error not enought female observations, ONLY 1 female SKIVE mouse
-
-Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
-           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-```{r Q1.totSC_male.first.comp, echo=FALSE}
-
-Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
-```
-
-```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-
-```
-
-In many mammalian species, the synaptonemal complex (SC) is longer in females [@cahoonLibuda2019, @sardell_sex_2020].
-Yet, the majority of these observations come from species with female-biased heterochiasmy.
-Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
-
-We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
-The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`).
-The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
-
-Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
-That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
-
-In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers compared to females.
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-#subsp effect, many strain, #male
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-```{r Q1.SC.Q12, echo=FALSE}
-
-#make comparisons for the high quality cells
-
-#get Q12 cells from MLH1.merge.Skel
-#calq the mouse averages
-
-MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
-
-#unorder the factors
-totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
-totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
-totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
-totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
-
-#REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(strain, sex),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
-
-#unorder the factors
-totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
-totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
-totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
-totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
-
-```
-
-```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
-
-Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Dom
-#Tot.SC_mouse.av_Dom
-
-Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Musc
-#Tot.SC_mouse.av_Musc
-
-Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-```
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE
-Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-### Positions of Single Crossovers
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE}
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
-#main figure of density plots,
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
-           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test( 
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-#lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-
-```
-
-We used the single crossover bivalent (1CO) normalized position to compared general location of crossovers while controlling for differences in total SC length.
-Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.
-
-In all strains, the landscape across 1CO bivalents is significantly different in females and males.
-Normalized foci positions tend to be more central in females `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` and closer to the telomere in males `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )` (t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
-
-Sex is also the most significant effect on focus position in a mixed model (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species [@sardell_sex_2020].
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-#real.long.bivData
-Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-
-Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-#mouse.av.long.biv.table
-
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-```
-
-## Sex Differences in CO Interference
-
-```{r Q1.ggplot_IFD.triangle, echo=FALSE}
-#IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
-#  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-#  geom_jitter(alpha=.4)+
-#  ggtitle("Normalized IFD\nDom")+
-#  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
-#  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)
-
-```
-
-```{r double.IFD.triangle.plots, echo=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
-auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
-
-df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
-df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-
-#this works
-top.triangle2 <- function(my.y, my.x, my.data) {
-  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
-       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
-       col =  colors_of_strains[current.strain])
-
-  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
-  mtext("Position.2.Female", 2, .75)
-  mtext("Position.1.Female", 3, 2)
-
-  axis(side=2, las=1, pos=0)
-  axis(side=3, las=1, pos=1)
-  lines(0:1, 0:1)
-}
-
-bottom.triangle2 <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[current.strain])
-
-  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
-  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-  mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1 + x.dist)
-  lines(0:1 + x.dist, 0:1, xpd=TRUE)
-}
-
-x.dist <- .1
-RP <- list()
-
-for(i in 1:(length(df_strain_list))){
-  #print(i)
-  current.strain <- as.character(df_strain_list[[i]]$strain[1])
-
-  #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
-
-  plot.new()
-
-  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-
-  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins
-  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
-
-  top.triangle2(my.y.top, my.x.top, my.data)
-
-  bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-
-```
-
-```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Distribution of normalized foci positions (IFDs) for double crossover bivalents across sexes.
-Female and male patterns are displayed in the top and bottom triangle respectively.
-Each point reflects a single double crossover bivalent, with the normalized positions of foci reflected on each axis."}
-
-plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
-plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
-plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
-plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
-plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
-plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
-plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
-
-# grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
-#             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
-#             rasterGrob(plot7.MSM), 
-#             ncol=3)
-
-#load all the files / arrange the plots from above 
-#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
-```
-
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
-
-#each point is a 2CO, with the F1 and F2 position 
-#based on plot from Dourad et al (Mercier)
-
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
-    aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.5)+
-  ggtitle("Raw IFD WSB")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#trying to move x axis to top
-#cowplot might be the easier way -- since kinda diffecult in ggplot
-#or ggdraw
-
-P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
-#SKIVE
-Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-```
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
-
-#these are the scatter plots
-
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
-
-```
-
-```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-
-#M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-
-```
-
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
-summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-```
-
-```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-```
-
-There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain, *domesticus^G^* (t-test without domesticus^G^; p = `r Q1.IFD.raw_t.test_G$p.value`).
-This result indicates that females and males exhibit a similar level of interference when it is measured in physical SC units.
-
-In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < `r Q1.IFD.PER_FULL_t.test$p.value`), with only _musculus^KAZ^_ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-
-Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` glm *M2*, p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`,  *M3* p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
-When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in SC length (chromatin compaction) using IFD^norm^ indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical scale.
-
-Visual comparison of interference pattern is aided by Figure 4C, which combines the total length of distance between two foci, but also their normalized locations along the chromosome.
-Two main patterns of sexual dimorphism emerge across all strains, (Supplemental figure X).
-First, the female patterns (top triangle), are more uniformly distributed compared to males.
-Suggesting a wider range of normalized IFD lengths compared to males (bottom triangle).
-Male IFDs display a stronger chromosome end-locatization of the second foci (Figure 4).
-
-```{r Comp_Ratios, echo=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
-
-Next, we used the contrast between males in the high-recombination strains ( _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ ) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M5).
-
-Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-SC length is expected to be positively associated with genome-wide recombination rate because of the connection between DSB formation and the loop-axis structure [@zickler1999, @mercier2015].
-Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome with a finite length.
-Thus, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
-
-Following this logic we predict (1) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_, and 3) _domesticus_ strains will have similar SC length and crossover interference.
-
-## Conservation of Mean SC length
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
-```
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-
-#+theme(legend.position="none") 
-```
-
-```{r Q2.SC.long.biv_plots, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+
-  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
-
-mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
-
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice.
-A)	 Mean total SC area per mouse across subspecies and recombination groups.
-High- and low-recombining groups are indicated by 1 and 0 respectively.
-(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell"}
-
-#make plot of mouse averages for MLH1 and total SC
-bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
-
-X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+theme(legend.position="none")+labs(x="Mouse Mean Total SC", y="Mouse Mean MLH1 count")
-
-plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
-
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-```{r Q2.totSC.ttest, echo=FALSE}
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol
-Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.09
-
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-#this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-
-glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
-
-#MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
-sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
-#DOM
-glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
-
-#MUSC
-glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
-
-#MOL
-#within subspecies
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r log.reg_SC.length, echo=FALSE, include=FALSE}
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
-#REMOVE THIS, IT uses total bivalent data
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
-Confirming the basic predictions, there is a positive correlation between mouse mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
-Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure 5).
-While high-recombination strains have significantly greater mouse mean total SC area in the total dataset (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p =  `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, the mouse means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
-
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (_musculus_ p =  `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, _molossinus_, p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
-In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-**Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?**
-
-## Single Crossover position is decoupled from genome wide recombination rate evolution
-
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.23 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
-
-Q2.show.nrmF1
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
-
-```
-
-```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q2.1COpos.show.MOL, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
-
-mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
-                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-
-#+facet_wrap(~mouse, scales = "free")
-
-single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
-                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
-
-#+scale_color_manual(values=colors_of_strains)
-
-single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
-                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-
-#CHRM.length is very significant-- some of the same subsp and strains are sig     
-```
-
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
-While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.
-
-## Evolution of interference is associated with genome wide recombination rate evolution
-
-```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
-#subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= strain, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(Rec.group ~ subsp, scales = "free")+
-  theme(legend.position="none", axis.text.x = element_blank(), axis.title.x = element_blank() )  + labs("Normalized IFD ") +
-  ylim(c(0,1.1))
-
-#consider adding horizontal line
-
-#to adjust the boxplot width
-#https://stackoverflow.com/questions/46174222/ggplot2-incorrect-boxplot-width-when-facetting-with-facets-of-different-scales
-#I don't think there's a straightforward way to achieve that in ggplot.
-You may have better luck plotting each facet as a separate ggplot object (on a subset of the data), then arrange them on a single page, using grid.arrange() from gridExtra for example.
-
-```
-
-```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE}
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
-```
-
-```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-
-Q2.males.IFD <- c("F1", "other")
-Q2.auto_curate_DF <- cleanCurated_BivData[ ! cleanCurated_BivData$strain %in% Q2.males.IFD, ]
-
-Q2.df_strain_list <- split(Q2.auto_curate_DF, (Q2.auto_curate_DF$strain))
-
-#this works
-Q2.bottom.triangle2 <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[unique(my.data$strain)])
-
-  mtext("Position.2.Male", 1, 1.5, at= mean(par()$usr[1:2])) #at should be changing the position
-  #above is just changing the text not the numbers
-  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)
-
-  #second number - on which MARgin line, starting at 0 counting outwards.
-  #where text should be,  at = 0, 1 (bad)
-
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +4)
-
-  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-#las, orientation, at location of line
-  #labels automatic labels
-  #pos
-  #xpd - clipping
-
-   mtext(seq(0, 1, .2), 1, 1, at=x.at)
-  #second / third val negative moves the valu up
-
-  axis(4, las=1, pos=1) #this is the y axis
- #   axis(4, las=1, pos=1 + x.dist)
-  #this is calling axis twice, not sure why
-  lines(0:1, 0:1, xpd=TRUE) #diagonal line
-}
-
-x.dist <- .25
-RP <- list()
-
-for(i in 1:(length(Q2.df_strain_list))){
-  #print(i)
-  current.strain <- as.character(Q2.df_strain_list[[i]]$strain[1])
-
-  #par(mar = c(2, 2, 2, 2))
-  #dev.new(width = 8, height = 4, unit = "in")
-  #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
-  plot.new()
-
-  my.data = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-  my.x.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
-  my.y.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins (the edges)
-  op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))
-
-  #   op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
-  Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-
-```
-
-```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
-
-Q2.plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_WSB.png')
-Q2.plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
-Q2.plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_G.png')
-
-Q2.plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
-Q2.plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
-Q2.plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
-
-Q2.plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
-Q2.plot8.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
-
-grid.arrange(rasterGrob(Q2.plot1.WSB), rasterGrob(Q2.plot2.LEW), rasterGrob(Q2.plot3.G),
-             rasterGrob(Q2.plot4.KAZ), rasterGrob(Q2.plot5.SKIVE), rasterGrob(Q2.plot6.PWD),
-             rasterGrob(Q2.plot7.MSM), rasterGrob(Q2.plot8.MOLF), 
-             ncol=3)
-
-```
-
-Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS$p.value`).
-
-This pattern is supported by slightly significant p values from comparisons within _musculus_ (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
-
-That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
-The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (_domesticus^G^_) to 16% (_musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
-
-Differnces in IFD can also be noted from the triangle IFD plots (Figure 6).
-The three high-recombining strains show more clustering of IFD observations near the area of the plot which indicates longer IFDs.
-The remaining strains have more observations short IFD observations near the diagonal.
-
-In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
-However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
-
-# Discussion
-
-Despite quantifying a similar number of wild derived inbred strains from *M.
-m.
-domesticus* and *M.
-m.
-musculus*, we only observed rapid male specific evolution for the genome wide recombination rate for *M.
-m.
-musculus.* Estimates of effective population size for the three main subspecies of the house mouse found *M.
-m.
-musculus* had the lowest [@geraldes2011].
-Selection becomes less efficient at removing deleterious mutations in small populations outside of mutation-drift balance.
-In such populations, modifiers for increased recombination rates will be selected for as they increase the efficiency of selection.
-
-An additional possibility is that the different rates in recombination rate variation across subspecies is related to history of hybridization across the populations sampled in this panel of inbred strains.
-Theoretical work as connected recombination that rate of purging deleterious introgressed DNA in populations after hybridization (Caballero et al 2019, Schumer 2018).
-The three high-recombining strains identified in this study have connections to hybridization.
-Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the european house mouse hybrid zone.
-Notably *musculus^SKIVE^* comes from a yonger section of the hybrid zone with a nuclear genome from *M.
-m.
-musculus* and a mitochondria genome from *M.
-m.
-domesticus* [@gyllensten1987_SKIVE].
-The subspecies *M.
-m.
-molossinus* is a natural hybrid of *M.
-m.
-musculus* and *M.
-m.
-cataneus* (Geraldes et al 2008, [@geraldes2011]).
-Models for the aggregate recombination metics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019), motivating increased study of recombination rates and natural populations.
-
-<!-- transition to looking at the number and placement of crossovers -->
-
-To place our results in the broader context of recombination rate evolution we compare our main resuts to expectionat under three modifier models and two functional and cell physiology models (table X).
-While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across exsisting models.
-The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome ocillatory movement (COM) model was built to describe positive interference.
-This spindle based selection model was developed to explain the results resented in this manuscript.
-In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
-
-## Evolution of reversed heterochiasmy direction
-
-In our results and more generally in the literature, female biased heterochiasmy is the most common pattern, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
-Our results extend previous reports of male biased heterochiasmy [@dumont2011] to  new subspecies of house mouse (*molossinus*) and emphasize the short evolutionary period.
-Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within house mouse subspecies requires further study.
-Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species in female biased heterochiasmy species, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
-This insight could be important for animal breeding programs [@battagin2016].
-
-Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
-
-A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion [@laneKauppi2019].
-Genetic variants will be selected in a manner to optimize the process of bi-orientation and prevent SAC from delaying cell cycle or triggering apoptosis.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014], which are connected to conserved features of gametogenesis; centrosome spindle [@So2019, @dumontDesai2012] and cell volume [@kyogoku2017].
-The more stringent SAC in spermatogenesis will be more effective at removing genetic variants which interfere with bi-orientation compared to females.
-This dynamic can give rise to sex-specific genetic variants which in turn can result in  context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
-
-## Conservation in sexual dimorphism in between cell variation in crossover number
-
-Regardless of whether these hypotheses are true, Our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
-Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
-These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes.
-This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
-
-In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
-This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals.
-Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-Centrosomes spindles are in spermatogensis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization [@manandhar2005centrosome].
-This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
-
-Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013].
-This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
-
-<!-- stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review -->
-
-## Conservation in sex-specific recombination landscape
-
-Our results extend the plethora or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] are data comprised mostly of species with large evolutionary distances.
-In contrast our results demonstrate that this pattern is maintained across much shorter evolutionary distances and even with rapid evolution in the genome-wide recombination rate.
-We argue that this conservation of the male chromosome end bias is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
-
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a conserved trait.
-The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
-The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM].
-
-Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (@laneKauppi2019, @subramanian2014, @dumontDesai2012).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
-Differences in timing of cell cycle between oogenesis and spermatogenesis imposes the the different selective pressures on how sister cohesion affects chromosome structure.
-Faster spermatogenesis may select for synchronization of the separation homologs (cite).
-While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, espicially for those with dicyate arrest [@lee2019].
-
-The increased genetic diversity and recombination rates at chromosome ends @haenel2018 suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020], which provides some interesting evolutionary predictions.
-
-## Conservation of sex differences in chromatin structure
-
-Our results from _musculus^MSM^_ and _musculus^PWD^_ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
-These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
-Chromatin compaction could explain variation in heterochiasmy, which is driven by the disinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
-Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
-females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation (see below).
-
-Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
-This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier, Sera) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
-
-As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
-From empirical data in mice [@petkov2007], [@lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
-The signal for interference is thought to be mediated though the SC, which is built on top of the chromosome axis [@zicklerKleckner2015] regardless of the loop size.
-Supporting this is the observation that interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-As a consequence, since more base pairs are sequestered into the radial loops per unit of the linear central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males,  will have a signal of stronger interference when measured in base pairs compared to 'the typical female chromatin configuration.
-Stronger interference in males, has been noted across many sex-specific linkage maps (cite) strengthening the hypothesis that sexual dimorphism in the chromatin organization is widespread, however this has only been confirmed by cytological data in mice and humans.
-(Also large difficulties in observing sperm cells through meiosis).
-
-<!-- The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019] -->
-
-## Evolution of interference strength in males
-
-An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (Schumer).
-Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
-
-Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
-The general pattern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
-This fits the logic of the chromosome axis acting as the 'real estate' for DSBs and aligns with known molecular mechanisms (cite).
-Two examples of the opposite direction, a positive correlation of interference strength and genome-wide recombination rate, involve descriptions of observations at the single bivalent level; the between lab-raised and wild mice of *Peromyscus leucopus* from [@peterson2019] or large effects on genome wide recombination rates across a short evolutionary differences; the within-sex comparison of two breeds of cattle with different genome wide recombination rates [@ma2015_cattle].
-
-We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphse.
-The evolution of genome wide recombination rates in our study is driven by a transition of the majority of
-bivalent having single crossovers to an enrichment of bivalents with two crossovers.
-All else being equal, interference strength would be expected to be equal or even weaker.
-Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain enriched for two crossover bivalents.
-Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
-
-The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase [@rubin2020mixing], however it lacks a mechanism for a evolutionary based predictions.
-While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
-
-We have focused on the model involving direct selection on the recombination pathway, which has an impact on the fitness of an individual, but we acknowledge connections to other models involving indirect selection on recombination rates which involve the fitness outcomes of offspring.
-A model from @goldstein1993, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
-The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
-A distinguishing feature of this model from other modifier  models is that the number of crossovers per chromosome is kept constant.
-Empirically, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
-Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
-
-## Future steps
-
-There are still many un-known aspects regarding heterochiasmy (@lenormand2016) --- that would benefit from the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
-A goal of such merging of models could be to connect empirical findings across scales.
-We make two suggestions for future steps in studying recombination variation.
-First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
-Second certain systems can be leveraged to overcome limitations inherent in some approached.
-For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
-FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian trans-locations.
-
-# References
+I have a hunch that using these mice might be the best (least risky way forward)

### 2020-05-08 doc/MS_v1.Rmd
fixing typos and updating the draft

@@
 
-We used the single crossover bivalent (1CO) normalized position to compared general location of crossovers while controlling for differences in total SC length.
+We used the single crossover bivalent (1CO) position to compared general location of crossovers while controlling for differences in total SC length.
 Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.
@@
 
-Regardless of whether these hypotheses are true, our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance.
-Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
-These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes.
-This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
-
 In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
@@
 Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-Centrosomes spindles are in spermatogenesis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization [@manandhar2005centrosome].
-This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
-
-Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013].
+Centrosomes act as the two MTOCs
+for the spindle during spermatogenesis and mitosis.
+The delivery of centrioles (most mammals) or centriole material (rodents and rabbits) to the egg is thought to be a barrier to evolution of parthenogenesis [@manandhar2005centrosome].This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
+
+Regardless of whether these hypotheses are true, our observations of levels of between cell variation fail to support a basic prediction of that a higher mean will be associated with a higher variance.
+Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
+These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002], [@gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes.
 This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].

### 2020-05-12 doc/MS_v1.Rmd
improve Figure 1 and 2

@@
    #segments
-  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
-  #Musc
-  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)
+  annotate("segment", x=.75, xend=2.3, y=19.5, yend=19.5, colour="black", size=1)+
+
+  #Musc, 30
+  annotate("segment", x= 3, xend=5.2, y=30, yend=30, colour="black", size=1)+
+
   #mol
-#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
-  #cast
-#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
-  #outgroups
- # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
+  annotate("segment", x=5.5, xend=8.1, y=19.5, yend=19.5, colour="black", size=1)+
+
+  annotate("segment", x=8.5, xend=8.8, y=30, yend=30, colour="black", size=1)+
+
+  annotate("segment",x=9.5,xend=11.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #spret
+
+  annotate("segment",x=9.5,xend=11.2, y=30, yend=30, colour="black", size=1)+
+  #spic
+
   #label
-
- #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
-   #        y = c(18.5,18.5,33,18.5, 18.5),
-
-  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
-    #       fontface = 'italic', hjust = 0.5, size=3.5)
+    annotate("text", x = c(1.2,3.9,6.3,8,10, 11), #list of all x points 
+           y = c(18.5, 31, 18.5, 31, 18.5, 31),
+
+   label = c("M. m. molossinus","M. m. musculus", "M. m. domesticus", "M. m. castaneus", "M. spretus", "M. spicilegus"),
+           fontface = 'italic', hjust = 0.5, size=3.5)
 
@@
 
+```{r MLH1.cell.examples, echo=FALSE}
+
+male.MLH1.cell.ex <- readPNG('~/MLH1repo/doc/figureFiles/MLH1.male.cell_ex.png')
+
+female.MLH1.cell.ex <- readPNG('~/MLH1repo/doc/figureFiles/MLH1.female_cell.ex.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+#Cell GGPLOT
+library(ggpubr)
+
+maleMLH1.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(male.MLH1.cell.ex) + theme(axis.title.x=element_blank(),
+     axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+femaleMLH1.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(female.MLH1.cell.ex) + theme(axis.title.x=element_blank(),
+     axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
 ```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
@@
 
-Figure1.main.plot <- Fig1_v1_strain / (female.specific.plot + male.plot )
-
-Figure1.main.plot + plot_annotation(
-  title = '',
-  subtitle = '',
-  caption = ''
-) + plot_annotation(tag_levels = 'A')
-
-```
-
-```{r rand.misc, echo=FALSE, include=FALSE, eval=FALSE}
-#str(AP_mouse_table_w.Ages)
-
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-#remove sex.x == na, and strain.x=na or F1
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
-
-AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
-#plot variance and mean (facet sex)
-
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]
-
-VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")
-
-cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")
-
-#str(AP_mouse_table.HQ)
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
-#[1] 0.5750399
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
-#0.321601
-
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
-# 0.3515544
-cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
-#-0.03713536
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
-# p-value = 8.623e-09
-
-cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
-    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )
-
-#remove mice with fewer than 10 cells
-
-AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]
-
-VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")
-
-cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])
-
-cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
-    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
-
-#why did I male just a female table?
-female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
-
-female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length( nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean( nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv( nMLH1.foci),
-                        var = format(round(   var( nMLH1.foci),3), nsmall=3),
-                        sd   = round(sd( nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-                        #quality?
-)
-```
+#try cowplot for more control of sizing
+library("cowplot")
+
+MLH1.Figure1  <- ggdraw() +
+  draw_plot(Fig1_v1_strain, x = 0, y = .55, width=1, height=.45)+
+
+  draw_plot(female.specific.plot, x=0, y=0, width=.5,height= .55)+
+  draw_plot(femaleMLH1.cell.obj, x=.32, y=.32, width=.19,height= .19)+
+
+  draw_plot(male.plot, x=.5, y = 0,  width = .5, height = .55)+
+  draw_plot(maleMLH1.cell.obj, x=.85, y = .32,  width = .19, height=.19)+
+
+  draw_plot_label(label = c("A", "B", "C"), size = 14,
+                  x = c(0, 0, .49), y = c(1, .55, 0.55) )
+
+MLH1.Figure1
+#add the cell images
+
+ggsave("~/MLH1repo/doc/Figure1.png")
+```
+
+updates for Figure 1, try shifting the lines up for B C
 
@@
 
-## Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
+## DMC1 Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
 
@@
 #change order
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c(
+                      "WSB","G", "KAZ","PWD","MSM"), order=T )
 
@@
 
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  scale_x_discrete( labels = c( expression( italic( domesticus^WSB) ), expression(  italic( domesticus^G) ), expression( italic( musculus^KAZ) ),
+
+     expression( italic(musculus^PWD)), expression( italic( molossinus^MSM) ) ) ) +
+
+  scale_color_discrete(name = "", labels = c("Late Zygotene", "Early Zygotene"))+
+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="top",
         panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.x = element_blank())
+
 ```
@@
 empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
 #EMPTY GGPLOT
@@
 
-three.DMC1.plot <- { Early.cell.obj + Late.cell.obj  } / DMC1.boxplot + plot_layout(widths = 1)
-
-#+ plot_layout(ncol = 2, widths = c(.65, .65), heights = c(.65, .65))
-#plot_layout(nrow = 2, widths = c(1), heights = c(.45) )
-
-#plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or #labels="AUTO"
-
-three.DMC1.plot + plot_annotation(
-  title = '',
-  subtitle = '') + plot_annotation(tag_levels = 'A')
+#cowplot
+
+DMC1.Figure2  <- ggdraw() +
+  draw_plot(Early.cell.obj, x=0, y=.5, width=.4,height= .45)+
+  draw_plot(Late.cell.obj, x=.5, y=.5,  width = .4, height = .45)+
+  draw_plot(DMC1.boxplot, x=0, y = 0,  width =1, height=.55)+
+
+  draw_plot_label(label = c("A", "B", "C"), size = 14,
+                  x = c(0, .49, 0), y = c(1,1, 0.55) )+
+#WSB
+  draw_line(
+    x = c(0.12, .23),
+    y = c( 0.01, .01),
+    color = colors_of_strains['WSB'], size=1
+  ) +
+  #G
+  draw_line(
+    x = c( 0.31, .42),
+    y = c(0.01, 0.01),
+    color = colors_of_strains['G'],size=1
+  ) +
+  #KAZ
+  draw_line(
+    x = c(0.49, .57),
+    y = c(0.01, 0.01),
+    color = colors_of_strains['KAZ'], size=1
+  ) +
+  #PWD
+  draw_line(
+    x = c(0.64, .73),
+    y = c(0.01, .01),
+    color = colors_of_strains['PWD'], size=1
+  ) +
+  #MSM
+  draw_line(
+    x = c(0.80, .91),
+    y = c(0.01, .01),
+    color = colors_of_strains['MSM'], size=1
+  )
+DMC1.Figure2
+#three.DMC1.plot <- { Early.cell.obj + Late.cell.obj  } / DMC1.boxplot + plot_layout(widths = 1)
+
+#three.DMC1.plot + plot_annotation(
+#  title = '',
+#  subtitle = '') + plot_annotation(tag_levels = 'A')
+
+ggsave("~/MLH1repo/doc/Figure2.png")
+
 ```

### 2020-05-12 src/SetupData.Rmd
updating the setupscript, move DMC1 DF setup

@@
 
-#opts_knit$set(root.dir = "~./MLH1repo/")
-#setwd("~./MLH1repo/")
+opts_knit$set(root.dir = "~./MLH1repo/")
+setwd("~./MLH1repo/")
 
-meta.data = read.csv(here("data", "/clean.Meta.Data.txt"), sep = "\t", header=TRUE)
+meta.data = read.csv( "~./MLH1repo/data/clean.Meta.Data.txt", sep = "\t", header=TRUE)
 #remove all MAD (peromyscus) from meta.data
@@
 
-source(here("src", "/CommonFunc_MLH1repo.R") )
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )
 
@@
 #read.csv was breaking things
-MLH1_data = read.xlsx2(here("data", "/AnonData_formated.xlsx"), header=TRUE, 1)
+MLH1_data = read.xlsx2("~./MLH1repo/data/AnonData_formated.xlsx", header=TRUE, 1)
 
-#formating the original anon file
-#MLH1_data = read.xlsx2(here("data", "/AnonData.xlsx"), header=TRUE, 1)
-#MLH1_data$fileName <- (MLH1_data$Original.Name)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
 
-#MLH1_data <- add_mouse(MLH1_data)
-#MLH1_data <- add_category(MLH1_data)
-#MLH1_data <- add_strain(MLH1_data)
-#MLH1_data <- add_sex(MLH1_data)
-#MLH1_data <- add_subsp(MLH1_data)
-#MLH1_data <- add_species(MLH1_data)
-#write csv to MLH1 file
-#write.xlsx2(MLH1_data, "AnonData_formated.xlsx", sheetName="Sheet1",
-#  col.names=TRUE, row.names=FALSE, append=FALSE)
 ```
@@
 
+#DMC1 Sections
+
+move the chunk below to setup
+
+```{r DMC1.setup, echo=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
+                 .(category),
+                 summarize,
+
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "early Z"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
+                 .(category),
+                 summarize, 
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "late Z"
+
+DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
+
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+```
+
 # Saving Data file
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.5.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
 

### 2020-05-12 src/BivData_Setup.Rmd
updating and refining the BivData setup script. goal write final xcel file

@@
 
-```{r setup, include=FALSE}
+```{r BivData.setup, include=FALSE}
 library(plyr)
@@
 
+#add_mouse ect.
+
 #write files for basic stats to csv files
@@
 #            row.names = TRUE)
+
 ```
@@
 
+#write.xlsx2(MLH1.merge.Skel, "Total.SC_merge.MLH1.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
 ```
@@
 
-```{r neg.IFD, echo=FALSE}
+```{r neg.IFD, echo=FALSE, eval=FALSE}
 
@@
 
-#mice to remove from the Curated_BivData -- should be similar to the mice exclude from the MLH1
+#mice to remove from the Curated_BivData -- should be similar to the mice exclude from the MLH1 (why?)
 
@@
 
+after all of these chunks are run, save the excel file for Curated_BivData
+
+DUPs, negative IFDs, remove juvi and MLH1 exclude mice,
+
+There are 3 bivalents with Per.Foci2 > 1 (remove or go correct)
+
+```{r write.total.SC, echo=FALSE}
+
+#write.xlsx2(MLH1.merge.Skel, "Total.SC_merge.MLH1.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
+```
+
 I'm considering removing the same mice from the MLH1 counts, from the BivData, but this might -- limit the comparisons -- might be too stict.
@@
 #can I plot these stats
-
 ```
@@
 )
-#can I plot these stats
 
@@
 Male.poly.Mouse.Table_BivData_4MM$sex <- droplevels(Male.poly.Mouse.Table_BivData_4MM$sex)
-
-```
-
-```{r Q1Q2.mouse.table, echo=FALSE}
-
-#How Do I apply quality score things to BivData?
-#make the same table but
 

### 2020-05-14 doc/MS_v1.Rmd
extensive debugging and organizing the datascripts

@@
     toc_float: yes
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
+  word_document:
+    toc: yes
+    toc_depth: '5'
 always_allow_html: yes
@@
 
-load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")
-
-load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
-
-load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
-load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
-
-#skeletonize
-SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData
+
+#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
+
+#load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
+
+#
+#load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
+
+#skeletonize - this is the total sc file
+#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
 
@@
 The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of `r sum(Q1.real.short_mouse_table$n_obs)` 'short' bivalents were isolated from `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Q1.real.short_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+A total of `r sum(Long.biv_mouse_table$n_obs)`   'short' bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
 Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
@@
 #MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
+Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)
+
+AP_mouse_table.HQ_target <- Mouse.Table_HQ[!(is.na(Mouse.Table_HQ$sex) | Mouse.Table_HQ$sex==""), ]
 
@@
 ```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
+
 #DF.HetC.MixedModel.HQ
 #this blck is duplicated in setupData.rm
-
 #https://rcompanion.org/handbook/G_03.html
@@
 
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
+DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]
 
 #AP_mouse_table excluide Qual5 cells
-DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
+DF.HetC.MixedModel.HQ <- Mouse.Table_HQ
 DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
@@
 
+#
+strain.table.HQ_Fig.og <- add_subsp(strain.table.HQ_Fig.og)
+strain.table.HQ_Fig.og <- add_species(strain.table.HQ_Fig.og)
+
 strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
@@
 
+sperman.corr.FM <- cor( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
+                        method = c("spearman") )
+
+sperman.corr.FM.pval <- cor.test( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
+                        method = c("spearman") )
+
+#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table2.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
+#qk.plot <- plot( ( strain.table.HQ_Fig %>% filter(sex == "female") )$mean_co,
+#      ( strain.table.HQ_Fig %>% filter(sex == "male") )$mean_co)
+
 ```
@@
 
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
+female_mouse.av <- Mouse.Table_HQ %>% filter(sex == "female")
 
@@
 #+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
 ```
@@
 
-```
-
-```{r MLH1.sum.stats.Bret.edits, echo=FALSE}
-
-AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
-#total mice
-#length( unique(AP_mouse_table.HQ %>% filter(species == "M.musculus") )$mouse )
-
-#average cell number for spermatocytes and oocytes
-#mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)
-#mean( (male_mouse.av %>% filter(species == "M.musculus") )$Ncells, na.rm = T)
 ```
@@
 molossinus*.
-Mean MLH1 foci counts for `r length( unique(AP_mouse_table.HQ_target)$mouse )` mice were quantified from an average of `r mean( (AP_mouse_table.HQ_target %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
+Mean MLH1 foci counts for `r length( unique(Mouse.Table_HQ)$mouse )` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
 
@@
 
-```{r DMC1.setup, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "early Z"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "late Z"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-```
-
 ```{r DMC1.table, echo=FALSE, message=FALSE}
@@
 #make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
+mean.MLH1.obs <- ddply(.data=Mouse.Table_HQ,
                  .(category),
@@
 
-chrm.class.barplot
 kk <- addSmallLegend(chrm.class.barplot)
@@
 
-#strain av tables
-#chrm.prop.male.high1
-#chrm.prop.male.low
 ```
@@
 
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
+```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
+#moved this to setupdata
 
@@
 #REAL.short_mouse_table
+#Long.biv_mouse_table
 
 #make the Q1 strain list
-Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}
-
+Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+```
+
+```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
 Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
@@
 
-```{r Q1.SC.show.short.biv, echo=FALSE}
-
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  theme(legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=9), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9))+labs(y = "Mouse Mean Short SC Length")
+```{r Q1.SC.show.short.biv, echo=FALSE, eval=TRUE}
+
+#Short.biv_mouse_table
+
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(), axis.line = element_line(colour = "black"),  legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=7), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9) )+
+
+  labs(y = "Short SC Length")
 
@@
 #run t.tests for the mouse av. for reduced single bivalent datasets
+
 #WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
 
 #G
-Q1_mus.av_shortSC_t.test.G <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
 
 #LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
 
 #PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
 
 #KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE - this value used in the text
+#only one SKVIE female, this is breaking the test
+#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
+
+#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
+#      )
+
+```
+
+```{r Q1SC_short.biv_dif, echo=FALSE, eval=TRUE}
+
+#calq the different in short.biv means across sexes
+#Long.biv_mouse_table
+
+#WSB
+Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
+
+#LEW
+Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
+
+#G
+#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
+
+#PWD
+Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
+
+#KAZ
+Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
 
 #SKIVE
-Q1_mus.av_shortSC_t.test.SKIVE <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
 
 #MSM
-#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
-#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
-#1 male, 3 females measures
-#try to merge the manual - whole cell observations (at least 2 males)
-
-```
-
-```{r Q1SC_short.biv_dif, echo=FALSE}
-#calq the different in short.biv means across sexes
-
-#WSB
-Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
-
-#LEW
-Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
-
-#G
-Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
-
-#PWD
-Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
-
-#KAZ
-Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
-
-#SKIVE
-Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
-
-#MSM
-Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
+Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
 
 #this bivData ratio is the same as mouse average
-Q1SC.BivData.short.biv_DIFF_MSM <- mean( (real.short.bivData %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (real.short.bivData %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
+Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
 
@@
 
+Q1.totSC.mouse.av.table$subsp2 <- 
+   ifelse(grepl("Dom", Q1.totSC.mouse.av.table$subsp), "M. m. domesticus", 
+        ifelse(grepl("Musc",Q1.totSC.mouse.av.table$subsp), "M. m. musculus",
+            ifelse(grepl("Mol",Q1.totSC.mouse.av.table$subsp), "M. m. molossinus",   
+               "other")))
+
 #change this plot to look like the short bivalent plot -- boxplots + mouse points
+
 Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
   scale_color_manual(values=colors_of_strains)+
-
-  facet_wrap(~subsp, scales="free", ncol = 3)+ 
+    facet_wrap(~subsp, scales="free", ncol = 3) +
+
   geom_jitter()+ geom_boxplot(alpha=.2)+ 
@@
 
-  theme(axis.title.x=element_blank(), axis.ticks.x=element_blank(),axis.text.x = element_text(size=9), legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10)  )+
-  labs(y = "Mouse Mean Total SC Length")+
-  ylim(c(1000,2600 ))+ggtitle("")
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
+    axis.title.x=element_blank(), axis.ticks.x=element_blank(),
+axis.text.x = element_text(size=7), 
+legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10) ) +
+
+  labs(y = "Total SC Length")+ylim(c(1000,2600 ))+ggtitle("")
+
+#italizing the facet label semms too hard -- do it in the final AI version
+#levels(powerDF$alph) <- c("italic('p=')*0.001", "italic('p=')*0.01", "italic('p=')*0.05")
+
+```
+
+```{r Fig3.load.icons, echo=FALSE}
+# load in the icons
+
+all.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/all_yellow_binary.png')
+single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+#EMPTY GGPLOT
+all.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(all.chrm.yellow) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+single.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(single.chrm.yellow) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+    axis.title.x=element_blank(),
+   axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
+```{r Fig3.example.double.triangle, echo=FALSE}
+#load an example two triangle plot
+
+single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
 
@@
 
-two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }
+#use cowplot
+library(cowplot)
+
+Q2.SC_dif.Figure3  <- ggdraw() +
+
+  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
+  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
+
+  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
+  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
+
+  #drawing triangles in the corner
+  #top right-triangle
+  draw_line(
+    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
+    y = c( .25, .25,  .15, .25),
+    color = "blue", size=.75
+  ) +
+
+  #bottom right-triangle
+  draw_line(
+    x = c( .1, .2, .2, .1 ),
+    y = c( .1,  .2, .1, .1),
+    color = "black", size=.75
+  ) +
+
+  draw_plot_label(label = c("A", "B", "C"), size = 14,
+                  x = c(0, .49, 0), y = c(1,1, 0.33) )
+
+#two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }
 
@@
 ```
+
+Make a figure for the diagram.  Import the straightened bivalents.. Load in the wsb double triangle.
 
@@
 ```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
 
@@
 We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
-The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`).
+The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = r Q1_mus.av_shortSC_t.test.SKIVE$p.value).
 The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
@@
 
-Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
 
@@
 
-#round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)
-
 ```
@@
 
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
 
@@
 #make comparisons for the high quality cells
-
 #get Q12 cells from MLH1.merge.Skel
@@
 
-totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
+totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
                  .(mouse),
@@
 
+#Long.biv_mouse_table
+
 #mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}
+Q1.SC.long.mouse_scatter_boxplot <- ggplot( Long.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
+
+```
+
+```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
 
@@
 
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
 
 #G
-Q1_mus.av_longSC_t.test.G <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
 
 #LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
 
 #PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
 
 #KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
+Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#not enough y observations
 #SKIVE
-Q1_mus.av_longSC_t.test.SKIVE <- t.test((Q1.real.long_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-                           (Q1.real.long_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
+#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
 
@@
 
-```{r Q1.F1pos_long.short_setup, echo=FALSE}
+```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
+#move to setup
 
@@
 #SKIVE
-Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
-           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )
+#Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain == "SKIVE") )$norm.first.foci.pos,
+ #          (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$norm.first.foci.pos )
 
@@
 
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE}
+```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
 #REAL.short_mouse_table
@@
 #SKIVE
-Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
 
@@
 
-```{r Comp_Ratios, echo=FALSE}
+```{r Comp_Ratios, echo=FALSE, eval=FALSE}
 
@@
 
-mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+mouse.shoo.long.biv <- ggplot( Long.biv_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
 \nMale Long Biv SC Lengths")
@@
 
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
+```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
+
+# theses t.tests not working
+#mouse average
 
 #DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+Q2.long_t.test_DomMusc <- t.test((real.long.bivData  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean_SC,
+                          (real.long.bivData %>% filter(subsp == "Musc") %>% filter(sex == "male"))$mean_SC)
 #p-value = 0.08
@@
 #DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+Q2.short_t.test_DomMusc <- t.test((Short.biv_mouse_table  %>% filter(subsp == "Dom")%>% filter(sex == "male") )$mean_SC,
+                          (Short.biv_mouse_table %>% filter(subsp == "Musc")%>% filter(sex == "male"))$mean_SC)
 #p-value = 0.8
@@
 #DOM-mol-long
-Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+Q2.long_t.test_DomMol <- t.test( (Long.biv_mouse_table  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean_SC,
+                          (Long.biv_mouse_table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean_SC )
 #p-value = 0.4
@@
 #DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+Q2.short_t.test_DomMol <- t.test((Short.biv_mouse_table  %>% filter(subsp == "Dom") %>% filter(sex == "male"))$mean_SC,
+                          (Short.biv_mouse_table %>% filter(subsp == "Mol")%>% filter(sex == "male"))$mean_SC)
 #p-value = 0.09
@@
 ```{r Q2.glm.SC.long, echo=FALSE}
+
 #this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
 sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
 
-glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
 sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
@@
 
-glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
+glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
+
 sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
@@
 #MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
+glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
+  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
 sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
@@
 #MOL can't be done -- only 1 mouse for each MSM and MOLF
-glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
-sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
+#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
+
+#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
 
@@
 
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` for _musculus_ and _molossinus_, respectively).
+The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively).
 While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.

### 2020-05-14 doc/totalSC.Rmd
extensive debugging and organizing the datascripts

@@
 
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#batch18
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
 #load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent
@@
 #these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
-SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
+SC.skel = read.csv("~./MLH1repo/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
 
 #load BivData  
-load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
+#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")
 
@@
 
-#strain specific xlsx sheets
+#strain specific xlsx sheets (why?)
 
@@
 
+# write table
+
 ```{r write.table, echo=FALSE, eval=FALSE}
+
+#add categories and levels
+
+#above threshold is bad,
 #write out these cell's info so I can double check them.
@@
 
+#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table2.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
 #write.table(MLH1.merge.Skel, "~./MLH1repo/data/BivData/total_SC_below_threshold.csv", sep=",", row.names = FALSE)
+
+#Below threshold is are the values to use
+
+#write.xlsx2(MLH1.merge.Skel, "Total.SC_merge.MLH1.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
 ```

### 2020-05-14 src/BivData_Setup.Rmd
extensive debugging and organizing the datascripts

--- 2020-03-06 src/BivData_Setup.Rmd
@@
 
+  8. **Strain DFs** for Het models **Mouse.Table_BivData_4MM**
+
+  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
+
+files that will be made in setup
   5. CELL level table (full all mice) (use this to make long biv dataset)
@@
  7. mouse table (full all mice)
-
-  8. **Strain DFs** for Het models **Mouse.Table_BivData_4MM**
-
-  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
 
@@
 
+#add_mouse ect.
 #write files for basic stats to csv files
@@
 
-```{r split.up.DF, echo=FALSE}
+```{r mk.Curate_BivData, echo=FALSE}
 #new.Curate.FULL to Curated_BivData
@@
 #Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
-```
-
-```{r set.type_Curated_BivData, echo=FALSE}
+
 ##change factor cols to numbers (factor -> chr -> number)
@@
 
-Curated_BivData$Rec.group.exclu <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
+#remove bivalent entries without hand foci count values
+cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
+                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
 ```
@@
 
-```{r no.Foci.Pos, echo=FALSE}
+```{r no.Foci.Pos, echo=FALSE, eval=FALSE}
 
@@
 #just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
+
+#not sure why these are different DFs
+just.SC.length2 <- Curated_BivData %>% filter(Foci1 == " ")
 just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]
-
-#table(just.SC.length$category)
-
-#decide what to do with these lines
-#how will they be represented in Mat Methods?
-```
-
-```{r input.check,echo=FALSE}
-
-#10jan18_13jan17_LEW_f1_sp1_14_rev_12  '?'
-
-##calq the nromalized.siscoten, double check the negative values;  #13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14, #27aug15_12jun15_WSB_m2_sp1_18_rev_18, #21nov17_16nov17_MSM_f3_sp1_3_rev_1
-
-#foci.pass = 2  #5aug17_14jul17_MSM_f1_sp1_13_rev_3
-#hand foci =0 but there is a foci position
-#too high hand foci count
-
-#check for mistakes that might have been made when inputing / curation process
-
-#hand foci > 4
-#foci pos > SC lengths
-#negative PER.foci.pos
-
-#it hink most (ALL) of these came from bug in my function (ugh)
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#negative.IFD <- Curated_BivData[which(Curated_BivData$IFD1_ABS < 0),]
-#caught 2
-#15mar16_16jan16_G_f1_sp1_4_rev
-#27mar16_6mar16_G_m3_sp1_12_rev
-
-#neg Nrm F1
-#negative.IFD2 <- Curated_BivData[which(Curated_BivData$PER_Foci_1 < 0),]
-#0 really?
-
-```
-
-```{r neg.IFD, echo=FALSE}
+#these are different than hand.foci.count
+
+table(Curated_BivData$hand.foci.count, exclude = NULL)
+#there are ~600 rows with SC and no hand foci count (blah)
+```
+
+```{r neg.IFD, echo=FALSE, eval=FALSE}
 
@@
 
-#neg.IFD <- Curated_BivData[Curated_BivData$IFD1_ABS < 0, ]
-#neg.ifd <- Curated_BivData %>% filter(hand.foci.count > 1) %>% filter(IFD3_ABS < 0 )
-
-```
-
-Main files from this Script:
-
-  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
-
-  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
-
-  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
-
-  4. **MELT_ABS**: make melt version of foci positions for abs measuers
-
-  5. CELL level table (full all mice) (use this to make long biv dataset)
-
-  6. Long Biv Dataset
-
-  7. mouse table (full all mice)
-
-  8. **Strain DFs** for Het models (summary strain DFs and tables?)
-
-  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models
-
-  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
-
-```{r CLEAN.UP.1, echo=FALSE, include=FALSE}
+neg.IFD <- Curated_BivData[Curated_BivData$IFD1_ABS < 0, ]
+neg.ifd <- Curated_BivData %>% filter(hand.foci.count > 1) %>% filter(IFD3_ABS < 0 )
+
+#same deal as above
+```
+
+```{r remove.DUP, echo=FALSE}
 
@@
 
-dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
+dup.curated <- Curated_BivData[duplicated(Curated_BivData$Obj.ID),] #784
 dup.curated <- add_mouse(dup.curated)
 
-#table(dup.curated$mouse)
-#1mar15_PWD_m1
+#table(dup.curated$category)#MSM male, PWD male
+#obj.ID is not complete for all rows
+
+Curated_BivData_w.dup <- Curated_BivData
+
+Curated_BivData <- Curated_BivData %>% distinct(Obj.ID, .keep_all = TRUE)
 
@@
 
-```{r remove.mice, echo=FALSE, eval=FALSE}
-
-#mice to remove from the Curated_BivData -- should be similar to the mice exclude from the MLH1
-
-#poor quality?
+Duplicates removed, `r length(Curated_BivData_w.dup$Obj.ID) - length(Curated_BivData$Obj.ID)`. most were MSM and PWD males.
+
+```{r remove.mice, echo=FALSE, eval=TRUE}
 
 #reaxamine these mice for if they should be included in 
+#this list is the same as setupData
 exclude_mice_list_MLH1 <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
@@
     "3jul18_TOM_m1","3sep18_TOM_m1",
-
                 "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
@@
 
-exmice.Curated_BivData <- Curated_BivData[ Curated_BivData$mouse %in% exclude_mice_list, ]
+exmice.Curated_BivData <- Curated_BivData[ Curated_BivData$mouse %in% exclude_mice_list_MLH1, ]
+
+Curated_BivData <- Curated_BivData[ ! Curated_BivData$mouse %in% exclude_mice_list_MLH1, ]
 
@@
 
-#old males
-
-```
+```
+
+`r length(exmice.Curated_BivData$Obj.ID)` rows removed for excluded mice.
+
+after all of these chunks are run, save the excel file for Curated_BivData
+
+DUPs, negative IFDs, remove juvi and MLH1 exclude mice,
+
+There are 3 bivalents with Per.Foci2 > 1 (remove or go correct)
+
+```{r write.Curated.BivData, echo=FALSE}
+
+#this file should have all the category columns, and pass the curation levels above
+
+#write.xlsx2(Curated_BivData, "Curated_BivData.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+#this excel causes an error
+
+#write.csv(Curated_BivData, "Curated_BivData.csv",   row.names = FALSE)
+
+```
+
+Main files from this Script:
+
+  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
+
+  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
+
+  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  4. **MELT_ABS**: make melt version of foci positions for abs measuers
+
+  5. CELL level table (full all mice) (use this to make long biv dataset)
+
+  6. Long Biv Dataset
+
+  7. mouse table (full all mice)
+
+  8. **Strain DFs** for Het models (summary strain DFs and tables?)
+
+  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models
+
+  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
 
@@
 #correct and remove, 26dec17_18nov17_WSB_f4_sp1_12_rev_28, 26dec17_18nov17_WSB_f4_sp1_12_rev_27
-table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count)
+table(WSB.Bivalents.DF$sex, WSB.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()
+wsb.ploty <- ggplot(WSB.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(WSB.Bivalents.DF$strain)
 wsb.ploty
@@
 LEW.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "LEW",]
-table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count)
+table(LEW.Bivalents.DF$sex, LEW.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+LEW.ploty <- ggplot(LEW.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(LEW.Bivalents.DF$strain)
 
@@
 G.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "G",]
-table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count)
+table(G.Bivalents.DF$sex, G.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+G.ploty <- ggplot(G.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(G.Bivalents.DF$strain)
 
@@
 PWD.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "PWD",]
-table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count)
+table(PWD.Bivalents.DF$sex, PWD.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+PWD.ploty <- ggplot(PWD.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(PWD.Bivalents.DF$strain)
 
@@
 KAZ.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "KAZ",]
-table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count)
+table(KAZ.Bivalents.DF$sex, KAZ.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+KAZ.ploty <- ggplot(KAZ.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(KAZ.Bivalents.DF$strain)
 
@@
 SKIVE.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "SKIVE",]
-table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count)
+table(SKIVE.Bivalents.DF$sex, SKIVE.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+SKIVE.ploty <- ggplot(SKIVE.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(SKIVE.Bivalents.DF$strain)
 
@@
 MSM.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MSM",]
-table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count)
+table(MSM.Bivalents.DF$sex, MSM.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+MSM.ploty <- ggplot(MSM.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(MSM.Bivalents.DF$strain)
 
@@
 
-```{r MOLF.SC, eval=FALSE}
+```{r MOLF.SC, echo=FALSE, eval=FALSE}
 #make the dataset
 MOLF.Bivalents.DF <- Curated_BivData[Curated_BivData$strain == "MOLF",]
-table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count)
+table(MOLF.Bivalents.DF$sex, MOLF.Bivalents.DF$hand.foci.count, exclude = NULL)
 
 #plot the differences
-MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("")
+MOLF.ploty <- ggplot(MOLF.Bivalents.DF,  aes(y=chromosomeLength, x = sex, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle(MOLF.Bivalents.DF$strain)
 
@@
 ## Male specific BivData set
+
+Redo the plots from above, just male data with strain on the x
+
+```{r BivLengths_by_strain, echo=FALSE}
+
+#set the order of strains
+
+BivLen.Male <- ggplot(
+  (Curated_BivData %>% filter(sex == "male") %>% filter(species == "M.musculus")),  aes(y=chromosomeLength, x = strain, color= as.factor(hand.foci.count)))+geom_jitter()+ggtitle("Male Single Bivalent Lengths")
+BivLen.Male
+```
 
@@
 
-# Long and Short BivData Set
+# reduced Long and Short BivData Set
 
@@
 #1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
+
 over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
@@
 
-```{r short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}
+```{r short.biv.data, echo=FALSE, include=FALSE, eval=TRUE}
 
@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
+                        nobs  = length(Obj.ID),
 
@@
 #save Rdata
-save.image("~./MLH1repo/data/CleanBivData_6.3.20.RData")
+save.image("~./MLH1repo/data/CleanBivData_5.12.20.RData")
 #save.image("data/BivData_4_PCA.RData")

### 2020-05-14 src/Result_Setup.Rmd
extensive debugging and organizing the datascripts

@@
 options(digits=2)
-
-setwd("~.")#switch over to here()
-
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.26.20.RData")
+#setwd("~.")
+
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
 #load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
@@
 
-load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
+load(file="~./MLH1repo/data/CleanBivData_5.12.20.RData") #CleanBivData_5.12.20.RData
 
@@
 
-```{r organizing.data, echo=FALSE }
+```{r organizing.data, echo=FALSE}
 #set the order of another column, based on another variable. consistant order for ploting
@@
 
-#there might be an error for changing wd
-#write.table(AP_category_table, "~./MLH1repo/strain_table_variances.csv", sep=",", row.names = FALSE)
-
 AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
@@
   scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
-
-```
-
-```{r misc.setup, echo=FALSE}
-
-colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
-                       #blues
-
-                       'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4',
-
-                       'TOM'= 'indianred1','AST' = 'tan2',
-                       'CZECH'= 'brown1',
-
-                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
-
-                    'CAST'='green','HMI'= 'forestgreen', #greens
-
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
-                    'CAROLI'='gold3',
-
-                    "F1"='grey', "other" = 'grey'
-                    )
-
-color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF", '4'= "#F564E3")
-#4!!
-
-#show_col(hue_pal()(5))
-
-#add Rec group,
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-Curated_BivData$Rec.group_alt <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                              ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))
-
-#apply this before ploting, to display strains in desired order
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, category, mouse) %>%
-#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-#MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
 
@@
 
-table <- raw.gamma_estimates
 #write.table(table, "~./MLH1repo/table.csv", sep=",",row.names = TRUE)
-
 #write.table(summarize_rbar_table, "~./MLH1repo/RBAR.table.csv", sep=",",row.names = TRUE)
@@
 
-#mean rbars are up around 
-#SKIVE male have extra (very strong interference)
-#there's so much variance across the gamma estimates -- 
-#most all females seem to have stronger interfernce
-#there don't seem to be consistant patterns
-
-#this plot shows much higher 2CO interference (via the gamma estimat from nrmIFD) in high rec males
-#across other classes of bivalents
-
-#most all other estiamtes of gamma estimates are around 4
-#
-
-#house mouse raw nu is around 11
-
-```
-
-```{r MLH1.FEMALE.TABLE, echo=FALSE}
+```
+
+```{r MLH1.FEMALE.TABLE, echo=FALSE, eval=TRUE}
+
 female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
@@
 
-```{r MLH1.MALE.TABLE, echo=FALSE}
+```{r MLH1.MALE.TABLE, echo=FALSE, eval=TRUE}
+
 #AP_mouse_table.MLH1_HQ
-
 #find the table that includes the las image data
@@
 Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+
 ```
@@
 #DF.HetC.MixedModel.HQ
-
 #no mice removed for low cell numbers?
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#remove the extra merged cols
-#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-#                            subsp.y,subsp.x,
-#                    category.y,category.x, strain.x,
-#                    strain.y, mouse.1  )  )
-
-#AP_mouse_table excluide Qual5 cells
-DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
-
-DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
-
-#main df with variables coded as chr
+AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
+
+#DF.HetC.MixedModel <- add_species(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- AP_mouse_table.HQ[AP_mouse_table.HQ$species == "M.musculus",]
+#DF.HetC.MixedModel
+
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel
+
+#main df with variables coded as chr, set false for the model
 DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
@@
 DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
+DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
 
@@
 #mean(SKIVE_male$mean_co) /  
-```
-
-```{r MLH1.MIXED.MODEL.setup.2, echo=FALSE, warning=FALSE, eval=FALSE}
-#DF.HetC.MixedModel.HQ
-#this blck is duplicated in setupData.rm
-
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-#AP_mouse_table excluide Qual5 cells
-DF.HetC.MixedModel.HQ <- AP_mouse_table.MLH1_HQ
-DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
-
-#main df with variables coded as chr
-DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
-
-all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
-                 .(sex, strain),
-                 summarize,
-                 all.cells = sum(Ncells)
-)
-
 ```
@@
 
-```{r juvimice.list, echo=FALSE}
-juvi_mice <- c(
-  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
-  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
-  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",
-
-  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
-   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
-  )
-
-#REMOVE JUVI mice
-#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
-#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
-
-```
-
 ```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
@@
 mouse.avs_4MM <- add_category(mouse.avs_4MM)
+mouse.avs_4MM <- add_species(mouse.avs_4MM)
 
@@
 
-#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
-#code if males are high or low?
-
 #REMOVE JUVI mice
@@
 
-mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))
-
 mouse.BivData.Table <- mouse.avs_4MM
-
 ```
@@
                         mean.IFD.PER = mean(mean_IFD.2CO_PER)
-
                         )
@@
 strain.BivData.Table <- add_species(strain.BivData.Table)
-
 ```
@@
 
-```{r Q1.Short.Long_biv.setup, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- REAL.short_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- REAL.long_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=TRUE}
 
@@
 juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
-mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
-
-```
+#mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
+
+```
+
+# MLH1 models
+
+M1 - subsp, sex, (1 | strain)
+M2 - subsp, sex, strain
+M3 - sex, strain
+M4 - strain (sex specific)
+
+```{r MLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
+
+#this dataframe doesn't work
+
+#don't run here, run in the MS file
+#DF.HetC.MixedModel.HQ
+#DF.HetC.MixedModel.HQ <- subset()
+
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1.MM.M1)$coefficients
+
+MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
+MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+
+MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
+sex.results =  MLH1.MM_sex, 
+               interaction.pval = MLH1.MM_interaction,  rand.pval = 
+MLH1.MM_strain)
+
+```
+
+```{r MLH1.glmM2, echo=FALSE, eval=TRUE}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
+#strain = fixed
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
+
+summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
+
+#table of just coeff anf pvalue
+#round(summ$coefficients[,c(1,4)],5 )
+
+#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
+```
+
+```{r MLH1.glm.M3, echo=FALSE, eval=TRUE}
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+MLH1.M3.strain.fixed <- glm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
+
+summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
+
+#round(summ.strain.fixed$coefficients[,c(1,4)],5)
+
+#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+```
+
+```{r Female.specific.glm, echo=FALSE, eval=TRUE}
+
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
+
+#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
+
+#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+```
+
+```{r Male.specific.glm, echo=FALSE, eval=TRUE}
+
+#these are M4
+#M4
+MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+sum_MLH1.male.glm.M4 <- summary(MLH1.male.glm.M4)
+#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
+
+#sum_MLH1.male.specific.glm.M2$coefficients[5,][4]
+
+#PWD, SKIVE, MSM
+#8,9,10
+
+#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+
+#M4.2
+MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
+sum_MLH1.male.glm.M4.2 <- summary(MLH1.male.glm.M4.2)
+
+#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
+
+#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+
+```
+
+#MLH1 variance models
+
+M1 - subsp, sex, (1 | strain)  (lmer)
+M2 - subsp, sex, strain  (glm / lm?)
+M3 - sex, strain
+M4 - strain (sex specific)
+
+```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+#summary(MLH1var_M1)
+LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
+
+LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
+sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
+
+```
+
+standardize the lm's and glms!
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=TRUE}
+#not including this model for now / brivity
+#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
+
+MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+MLH1var_M2_sum <- summary(MLH1var_M2)
+
+#summary(MLH1var_M2)
+#anova(MLH1var_M2)  
+#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
+
+MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
+
+#summary(MLH1cv_M3)
+#anova(MLH1cv_M2)
+#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
+
+#Q12 var
+MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1var_M3_Q12)
+#anova(MLH1var_M2_Q12)  
+#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
+
+#Q12 cv
+MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
+
+#summary(MLH1cv_M2_Q12)
+#anova(MLH1cv_M2_Q12)#sex large F
+#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
+```
+
+```{r M3.MLH1.var, echo=FALSE, eval=TRUE}
+glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
+#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
+
+MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
+#summary(MLH1cv_M4)
+#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
+
+### Q12 var
+MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
+
+#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
+#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
+
+#Q12 cv
+MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
+#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
+```
+
+```{r MLH1.var.M4, echo=FALSE, eval=TRUE}
+#male
+glmMLH1Var_M4.male <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "male") ) )
+
+#female
+glmMLH1Var_M4.female <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "female") ) )
+sum.glmMLH1Var_M4.female <- summary(glmMLH1Var_M4.female)
+
+glmMLH1Var_M4.femaleQ12 <- glm( var ~ subsp * strain,  data= (Q12_mouse_table_w.Ages %>% filter(sex == "female") ) )
+
+```
+
+```{r FEMALE.VAR.MODELS, echo=FALSE, eval=TRUE}
+## within mouse variance models 
+lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
+summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
+
+lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
+lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
+
+#round( #round isn't working for this sumary
+#lmVAR.female.strain_sum$coefficients[,c(1,4)]
+#Q12
+Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
+
+lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
+lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
+
+lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
+lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
+
+lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
+lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
+
+#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
+#LEW sig
+
+```
+
+```{r MALE.SPECIFIC.variance.model, echo=FALSE, eval=TRUE}
+## within mouse variance models 
+lmVAR.male.strain.subsp <- glm(var ~ subsp * strain, data= just.male_mouse.av)
+summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
+#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
+
+#all p values kinda high, MSM and TOM are slightly sig
+lmVAR.male.strain <- glm(var ~ strain, data= just.male_mouse.av)
+lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
+#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
+#non sig
+
+lmcv.male.strain <- glm(cV ~ strain, data= just.male_mouse.av)
+lmcv.male.strain_sum <- summary(lmcv.male.strain)
+#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
+#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
+
+#Q12
+Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
+
+lm.m1.Q12_male <- glm(var ~ subsp * strain, data= Q12_male_table)
+lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
+#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+lm.m2.Q12_male <- glm(var ~ strain, data= Q12_male_table)
+lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
+#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
+#NS
+
+```
+
+# DMC1 Results
+
+main table (DMC1.mean.table), t.tests  , ratios
+
+set up, t.tests, correlations, says there are duplicate titles
+
+```{r DMC1_t.tests1, echo=FALSE, warning=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+```{r DMC1.rand.calq1, echo=FALSE}
+
+DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
+DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
+
+mean.MLH1.obs <- ddply(.data=Mouse.Table_HQ,
+                 .(category),
+                 summarize, 
+                 mean.MLH1 = mean(mean_co),
+                 var.MLH1 = var(mean_co)
+)
+
+DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
+                   category == "WSB male" |
+                     category == "KAZ male" |
+                  category == "MSM male" |
+                    category == "PWD male")
+
+DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
+
+DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
+
+MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
+
+MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
+```
+
+```{r DMC1.ratios1, echo=FALSE}
+
+##RATIOS
+DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
+DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
+
+#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
+
+#DMC1.mean.table
+DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
+
+colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
+                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
+
+WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
+
+G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
+
+MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
+
+PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
+
+DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
+
+DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
+
+kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+
+DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
+
+DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
+                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+
+```
+
+-----
+
+```{r DMC1.setup, echo=FALSE, eval=FALSE}
+#load DMC1 csv
+#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
+DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+DMC1.DF$fileName <- DMC1.DF$Original.Name
+DMC1.DF <- add_mouse(DMC1.DF)
+DMC1.DF <- add_strain(DMC1.DF)
+DMC1.DF <- add_category(DMC1.DF)
+DMC1.DF <- add_subsp(DMC1.DF)
+DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
+
+#make DMC1 table
+DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
+                 .(category),
+                 summarize,
+
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.L$stage <- "L"
+
+DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
+                 .(category),
+                 summarize, 
+                 ncells = length(fileName),
+                 mean.count = mean( total.count)
+)
+DMC1.mean.obs.Z$stage <- "Z"
+
+DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
+
+#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
+#            row.names = FALSE)
+
+DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
+                          ifelse(grepl("G",DMC1.DF$strain), "Low",
+
+              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
+            ifelse(grepl("PWD", DMC1.DF$strain), "High",
+               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
+```
+
+```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE, eval=FALSE}
+#think of better color scheme
+#make the jitter col thinner or make the stacked dots.
+#remember that box-plots add outlier dots.
+#change y labels
+
+#experimenting with ggboplot
+everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
+          color = "stage", palette = "jco",
+          add = "jitter", short.panel.labs = FALSE)
+  #library gpubr has functions to add in sig/test bars.
+
+#change order
+
+DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
+                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
+
+```
+
+```{r plot.stage.ttest, echo=FALSE, warning=FALSE, eval=FALSE}
+
+anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
+#
+
+DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
+  geom_boxplot(position=position_dodge()) + 
+  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
+
+  #facet_wrap(~stage, nrow=1) +
+
+  ggsignif::geom_signif(
+    data=anno_df_stage, 
+    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
+    manual=TRUE
+  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
+
+ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
+               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
+
+DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
+
+ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
+               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
+
+DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
+
+ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
+               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
+```
+
+```{r DMC1_t.tests, echo=FALSE, warning=FALSE, eval=FALSE}
+#2. t-tests for the DMC1 counts 
+#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
+#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
+
+#high vs low group (L)
+ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value =  0.002145
+
+#high vs low group (Z)
+ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
+               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
+# p-value = 0.6628
+
+#dom vs musc (L) (2 subsp vs)
+ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
+# p-value = 0.003016
+
+#dom vs musc (Z)
+ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
+#p-value = 0.09588
+
+#KAZ vs Dom (Z) (only 1 L cell)
+ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
+      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
+
+#p-value = 0.004148
+#Kaz Z are higher than Dom
+```
+
+# toal SC
 
@@
 ```{r tot.skel.table, echo=FALSE}
+#move to setup file
 
@@
 ```{r tot.SC_strain.setup, echo = FALSE}
+#move to setup
 
@@
 
-```{r LONG.SHORT.biv_setup, echo=FALSE}
-
-#FROM q1
-#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#from Q2 IFD
-#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-##long biv
-#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-#FROM q1 
-REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO  
-                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                         )
-
-REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
-
-REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
-REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
-REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
-                                           0)))
-
-REAL.short_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1, 0))
-
-###LONG BIV
-
-REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                       n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-                         )
-
-REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
-
-REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
-REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
-REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
-                                           0)))
-
-REAL.long_mouse_table$Rec.group.exclu <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1, 0))
-
-#find which one 
-```
-
-# Datasubsets
-
-# MLH1 models
-
-M1 - subsp, sex, (1 | strain)
-M2 - subsp, sex, strain
-M3 - sex, strain
-M4 - strain (sex specific)
-
-```{r MLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE}
-
-#DF.HetC.MixedModel.HQ
-
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
-MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
-
-MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
-MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
-
-MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
-sex.results =  MLH1.MM_sex, 
-               interaction.pval = MLH1.MM_interaction,  rand.pval = 
-MLH1.MM_strain)
-
-```
-
-```{r MLH1.glmM2, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
-
-#strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
-
-summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)
-
-#table of just coeff anf pvalue
-#round(summ$coefficients[,c(1,4)],5 )
-
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
-```
-
-```{r MLH1.glm.M3, echo=FALSE}
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-MLH1.M3.strain.fixed <- glm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
-
-summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
-
-#round(summ.strain.fixed$coefficients[,c(1,4)],5)
-
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
-```
-
-```{r MLH1.M4, echo=FALSE}
-
-```
-
-```{r Female.specific.glm, echo=FALSE}
-
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
-
-MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
-sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
-
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
-```
-
-```{r Male.specific.glm, echo=FALSE}
-
-#these are M4
-#M4
-MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
-sum_MLH1.male.glm.M4 <- summary(MLH1.male.glm.M4)
-#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#sum_MLH1.male.specific.glm.M2$coefficients[5,][4]
-
-#PWD, SKIVE, MSM
-#8,9,10
-
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
-
-#M4.2
-MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
-sum_MLH1.male.glm.M4.2 <- summary(MLH1.male.glm.M4.2)
-
-#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
-
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- glm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- glm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- glm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- glm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- glm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-```
-
-#MLH1 variance models
-
-M1 - subsp, sex, (1 | strain)  (lmer)
-M2 - subsp, sex, strain  (glm / lm?)
-M3 - sex, strain
-M4 - strain (sex specific)
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-```
-
-standardize the lm's and glms!
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE}
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-MLH1var_M2_sum <- summary(MLH1var_M2)
-
-#summary(MLH1var_M2)
-#anova(MLH1var_M2)  
-#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
-
-MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
-#anova(MLH1cv_M2)
-#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
-
-#Q12 var
-MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-#anova(MLH1var_M2_Q12)  
-#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M2_Q12)
-#anova(MLH1cv_M2_Q12)#sex large F
-#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-
-```{r MLH1.var.M4, echo=FALSE}
-#male
-glmMLH1Var_M4.male <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "male") ) )
-
-#female
-glmMLH1Var_M4.female <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "female") ) )
-sum.glmMLH1Var_M4.female <- summary(glmMLH1Var_M4.female)
-
-glmMLH1Var_M4.femaleQ12 <- glm( var ~ subsp * strain,  data= (Q12_mouse_table_w.Ages %>% filter(sex == "female") ) )
-
-```
-
-```{r FEMALE.VAR.MODELS, echo=FALSE}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
-
-lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
-lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
-
-# DMC1 Results
-
-main table (DMC1.mean.table), t.tests  , ratios
-
-set up, t.tests, correlations, says there are duplicate titles
-
-```{r DMC1.setup1, echo=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "Early"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "Late"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
-#            row.names = FALSE)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-#attach this mini DF to the DMC1 data
-```
-
-```{r DMC1_t.tests1, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1.rand.calq1, echo=FALSE}
-
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-```
-
-```{r DMC1.ratios1, echo=FALSE}
-
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-
-#DMC1.mean.table
-DMC1.mean.table.99 <- subset(DMC1table.main, select= -c(var.MLH1,  mean.count.x,  mean.count.y, stage.x, stage.y))
-
-colnames(DMC1.mean.table.99) <- c("category","mean.MLH1","ncells.L","mean.L",
-                  "ncells.Z", "mean.Z", "MLH1.L_ratio", "MLH1.Z_ratio")
-
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
------
-
-```{r DMC1.setup, echo=FALSE, eval=FALSE}
-#load DMC1 csv
-#DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
-DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
-
-DMC1.DF$fileName <- DMC1.DF$Original.Name
-DMC1.DF <- add_mouse(DMC1.DF)
-DMC1.DF <- add_strain(DMC1.DF)
-DMC1.DF <- add_category(DMC1.DF)
-DMC1.DF <- add_subsp(DMC1.DF)
-DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]
-
-#make DMC1 table
-DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
-                 .(category),
-                 summarize,
-
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.L$stage <- "L"
-
-DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
-                 .(category),
-                 summarize, 
-                 ncells = length(fileName),
-                 mean.count = mean( total.count)
-)
-DMC1.mean.obs.Z$stage <- "Z"
-
-DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)
-
-#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
-#            row.names = FALSE)
-
-DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
-                          ifelse(grepl("G",DMC1.DF$strain), "Low",
-
-              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
-            ifelse(grepl("PWD", DMC1.DF$strain), "High",
-               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
-```
-
-```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE, eval=FALSE}
-#think of better color scheme
-#make the jitter col thinner or make the stacked dots.
-#remember that box-plots add outlier dots.
-#change y labels
-
-#experimenting with ggboplot
-everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
-          color = "stage", palette = "jco",
-          add = "jitter", short.panel.labs = FALSE)
-  #library gpubr has functions to add in sig/test bars.
-
-#change order
-
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
-                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-DMC1.boxplot
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE, eval=FALSE}
-
-anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
-#
-
-DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-  geom_boxplot(position=position_dodge()) + 
-  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-  ggsignif::geom_signif(
-    data=anno_df_stage, 
-    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-    manual=TRUE
-  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE, eval=FALSE}
-#2. t-tests for the DMC1 counts 
-#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
-#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
+```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=TRUE}
+
+Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
 ```
@@
 
-```{r Q1.single.biv_position_table, echo=FALSE}
+```{r Q1.single.biv_position_table, echo=FALSE, eval=FALSE}
 #strain averages for single bivalent 
 #
-colnames(Curated_BivData99)
-
-```
+#colnames(Curated_BivData99)
+
+```
+
+```{r table.long.biv, echo=FALSE}
+
+# make a table for the Mixed Models -- from the long.BivData
+# Calq mouse averages
+
+#add in all the other stats there are in the other tables
+
+Long.biv_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells = length( unique(fileName) ),
+                        n_obs  = length(Obj.ID),
+
+                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                        #SC lengths by class
+                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
+                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+        #          mean_siscoten = mean(siscoten)
+        #
+)
+#mouse averages for long.biv
+
+Long.biv_mouse_table <- add_sex(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_strain(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_subsp(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_category(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_species(Long.biv_mouse_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Long.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
+Long.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
+Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)
+
+#rec group
+Long.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Long.biv_mouse_table$category), 1, 
+                            ifelse(grepl("MSM", Long.biv_mouse_table$category), 1,
+                             ifelse(grepl("SKIVE", Long.biv_mouse_table$category), 1, 0)) )
+
+#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)
+
+#1) Long Biv for 
+#remove all the non-used strains!!
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table
+
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
+Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
+
+#remove na's
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]
+
+#table(droplevels( Long.biv_mouse_table_HetC$category) )  #for double checking
+
+#2) Long Biv for male.poly
+#just males, just Musculus
+Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
+
+Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
+
+#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
+```
+
+```{r table.short.biv, echo=FALSE}
+
+#real.short.bivData
+
+#same chunk as above
+Short.biv_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells = length( unique(fileName) ),
+                        n_obs  = length(Obj.ID),
+
+                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                        #SC lengths by class
+                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
+                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+        #          mean_siscoten = mean(siscoten)
+        #
+)
+#mouse averages for
+
+Short.biv_mouse_table <- add_sex(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_strain(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_subsp(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_category(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_species(Short.biv_mouse_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Short.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)
+
+#rec group
+Short.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Short.biv_mouse_table$category), 1, 
+                            ifelse(grepl("MSM", Short.biv_mouse_table$category), 1,
+                             ifelse(grepl("SKIVE", Short.biv_mouse_table$category), 1, 0)) )
+
+#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)
+
+#1) Long Biv for 
+#remove all the non-used strains!!
+#Short.biv_mouse_table_HetC <- Short.biv_mouse_table
+
+#Short.biv_mouse_table_HetC <- Short.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
+
+#Short.biv_mouse_table_HetC <- Short.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
+
+#Short.biv_mouse_table_HetC <- Short.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
+
+#Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
+
+#remove na's
+#Long.biv_mouse_table_HetC <- #Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | #Long.biv_mouse_table_HetC$mouse==""), ]
+
+```
+
+need to save this reduced dataset, Q1.real.short_mouse_table.
 
@@
 
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
 male.short.biv <- real.short.bivData %>% filter(sex== "male")
@@
 
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
 ```
@@
 
-```{r Q2.SC.long.biv_plots, echo=FALSE}
+```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
 
@@
 
-mouse.av_totsc_male
-
 ```
@@
 #make plot of mouse averages for MLH1 and total SC
-
 bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
@@
   scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-
-X.by.texas
 
@@
 
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE}
+```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
 
@@
 
-```{r Q2.glm.SC.long, echo=FALSE}
+```{r Q2.glm.SC.long, echo=FALSE, eval=FALSE}
 
@@
 
-```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
-#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos
+```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE, eval=FALSE}
 
 #High = MSM, PWD, SKIVE
-Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == 0))$norm.first.foci.pos,    
+        (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == 1))$norm.first.foci.pos )
+
+#high = PWD and MSM
+#Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
+#            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == #"1"))$norm.first.foci.pos )
 #p-value = 1
 
-#high = PWD and MSM
-Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
-            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
-#p-value = 1
-
-```
-
-```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
-
+```
+
+```{r Q2.1COpos_within.subsp_ttest, echo=FALSE, eval=FALSE}
 #Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
 
+#mouse.av_4MM -- n
+
 #Musc
-Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
+Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == 0))$norm.first.foci.pos,
+
+            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == 1))$norm.first.foci.pos )
 # p-value = 0.4
@@
 
-```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
+```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE, eval=FALSE}
 #these don't have mean.first.foci
@@
 #short -- real.long.bivData -- REAL.short_mouse_table
-Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,
-
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
+Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0" ))$mean_F1pos_ABS,
+
+      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == 1))$mean_F1pos_ABS   )
 #p-value = 0.3
@@
 
-#whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% #filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )
-#p-value = 0.4
-
 ```
@@
 
-```{r Q2.t.test_IFD_mouse.av, echo=FALSE}
+```{r Q2.t.test_IFD_mouse.av, echo=FALSE, eval=TRUE}
 #t.test for IFD
@@
 #save Rdata
-save.image("~./MLH1repo/data/Results_setup_4.24.20.RData")
-
-```
+save.image("~./MLH1repo/data/Results_setup_5.14.20.RData")
+
+```

### 2020-05-14 src/SetupData.Rmd
extensive debugging and organizing the datascripts

@@
 ---
-
-to Do
-
-- try calq correlation of MLH1 counts across duplicates
-
-- format the DF for applying to Mixed models
-
-- try coding i) date image for each cell image, ii) most recent image date for each mouse
 
@@
 library(xlsx)
+library(reshape2)
 #knitr::opts_chunk$set(echo = FALSE)
@@
 DMC1.DF <- read.csv("~./MLH1repo/data/DMC1newJuviQuant.csv", header = TRUE)
+
+#total SC -- the one file should be all I need?
+Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)
+
+#Curated Bivdata - this excel file is too big to work with read.xlsx2
+#Curated_BivData = read.xlsx2("~./MLH1repo/data/Curated_BivData.xlsx", header=TRUE, 1)
+
+Curated_BivData <-  read.csv("~./MLH1repo/data/Curated_BivData.csv", header=TRUE)
 
@@
 
+#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
 colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
@@
 
-```{r adj_nMLH1, echo=FALSE}
+```{r adj_nMLH1, echo=FALSE, eval=FALSE}
 #add a column with male adjusted MLH1 values (+1 to all males)
@@
 MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
+
+#adjusted MLH1 no longer used
 ```
@@
 #order the batches in MLH1 and original
+#apply this ordering to all other DF?
+
 ```
 
-```{r kable.table, echo=FALSE}
-#table comparing number of images in quantification batches
+```{r Batch.Stats, echo=FALSE, eval=TRUE}
+#plot / table comparing number of images in quantification batches
+#basic visualization for number of images in batches
+
 #ToDo:  correct the case of all batches, better names
@@
 
-```{r all.data.table, echo=FALSE}
+```{r all.data.table, echo=FALSE, eval=FALSE}
 
@@
 
+#NA cause by CAST female
+#this table isn't used much
 ```
 
-```{r mouse.table, echo=FALSE}
+```{r HQ.mouse.table, echo=FALSE}
 #Make a mouse level table
 
-AP_mouse_table.MLH1_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
+#change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
+#
+
+Mouse.Table_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
 
-#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
-
-AP_mouse_table.MLH1_HQ <- add_strain(AP_mouse_table.MLH1_HQ)
-AP_mouse_table.MLH1_HQ <- add_subsp(AP_mouse_table.MLH1_HQ)
-AP_mouse_table.MLH1_HQ <- add_sex(AP_mouse_table.MLH1_HQ)
-AP_mouse_table.MLH1_HQ <- add_category(AP_mouse_table.MLH1_HQ)
-
-#add DOB to this table?
-mice.frm.meta <- AP_mouse_table.MLH1_HQ[as.character(AP_mouse_table.MLH1_HQ$mouse) %in% as.character(meta.data$mouse),]
+Mouse.Table_HQ <- add_strain(Mouse.Table_HQ)
+Mouse.Table_HQ <- add_subsp(Mouse.Table_HQ)
+Mouse.Table_HQ <- add_sex(Mouse.Table_HQ)
+Mouse.Table_HQ <- add_category(Mouse.Table_HQ)
+Mouse.Table_HQ <- add_species(Mouse.Table_HQ)
 
@@
 meta.data <- add_category(meta.data)
+meta.data <- add_species(meta.data)
+
+#add DOB to this table?
+mice.frm.meta <- Mouse.Table_HQ[as.character(Mouse.Table_HQ$mouse) %in% as.character(meta.data$mouse),]
+
+Mouse.Table_HQ.OG <- Mouse.Table_HQ
+
+Mouse.Table_HQ <- Mouse.Table_HQ.OG
 
 #during merging -- try to just merge a few cols: 
-AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data)
+Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
+#this removes ~28 mice
 
-#this should be transfering the image dataes!
-
-#don't use by, that makes duplicate cols
-# by.y = "mouse", by.x = "mouse", all = TRUE)
-#what are the differences between these two
-
-#remove mice without MLH1 obs
-AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice == ""), ]
-
-#remove cols with bad suffix
-#AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-#                            subsp.y,subsp.x,
-#                    category.y,category.x, strain.x, strain.y,
-#                    mouse.1  )  )
-
-#add all the cols back
-AP_mouse_table_w.Ages <- add_strain(AP_mouse_table_w.Ages)
-AP_mouse_table_w.Ages <- add_subsp(AP_mouse_table_w.Ages)
-AP_mouse_table_w.Ages <- add_sex(AP_mouse_table_w.Ages)
-AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-#make sure they are all factors - this was breaking the kniting
-#AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
-#AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
-#AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
-#AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)
-
-old.mice.df <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.weeks > 25),]
+old.mice.df <- Mouse.Table_HQ[(Mouse.Table_HQ$age.weeks > 25),]
 #remove the NAs
@@
 
+Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]
+
 #remove juvi mice via the list
+#remove bad ages
+#this is removing female
 
-#remove bad ages
-#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[( (AP_mouse_table_w.Ages$age.weeks < 25) &  (AP_mouse_table_w.Ages$age.days != 13) ) | #is.na(AP_mouse_table_w.Ages$age.weeks), ]
+#this removes everything but males in these categories
+#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]
+
+#i don't understand why so many WSB female are removed..
+# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)
+
 ```
@@
 
-#REMOVE EXTRA STRAINS
-Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
+#REMOVE EXTRA STRAINS -- why?
+#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
 #remove musc strains
-Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
-Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
+#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
+#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
 #remove PERC, MOLF, CZECH, (no females)
-Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
-Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]
+#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
+#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]
 
@@
 ```{r HetC.DF.4MM, echo=FALSE}
-#this block has so many fing errors
-
-#for the DF for use in the HetC Mixed Model
-#Which Start with the AP_mouse_table
 
 #variance is still chr
-AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
+Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)
 
-#use the merged DF to filter by age 
-#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-
-#setwd("~./MLH1repo/src/")
-
-source(here("src","CommonFunc_MLH1repo.R") )
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages %>% filter(species == "M.musculus")
+DF.HetC.MixedModel <- Mouse.Table_HQ %>% filter(species == "M.musculus")
 
@@
 #use the above table to pull the male mouse levels
-mlh1male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")
+mlh1male_mouse.av <-  Mouse.Table_HQ %>% filter(sex== "male")
 
 #add the Rec.group markers
-
 #mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
@@
 
-male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex == "male") &(AP_mouse_table_w.Ages$category != "other"),]
+male_mouse_table_w.Ages <- Mouse.Table_HQ[(Mouse.Table_HQ$sex == "male") &(Mouse.Table_HQ$category != "other"),]
 
-all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
+all.ages_plot <- ggplot(Mouse.Table_HQ[Mouse.Table_HQ$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
 all.ages_plot
@@
 
-```{r color.scheme}
-#color scheme, #mouse colors
-#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
-colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
-                       #blues
-
-                       'PWD'= 'red2',
-                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
-                       'CZECH'= 'indianred4',
-
-                       'MSM'= 'deeppink', 'MOLF'= 'hotpink', #mol are pinks, bc they are fancy mice
-
-                    'CAST'='green','HMI'= 'forestgreen', #greens
-
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
-                    'CAROLI'='gold3',
-
-                    "F1"='grey', "other" = 'grey'
-                    )
-
-```
-
-#DMC1 Sections
+# DMC1 Section
 
@@
 
+# BivData
+
+for these full curated, add all the labels and save them as excel files in the BivData setup file.
+
+full curated, reduced datasets
+
+Are the strain levels plots needed?
+
+Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.
+
+```{r no.Foci.Pos, echo=FALSE}
+#no posittion
+#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
+just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]
+
+#table(just.SC.length$category)
+```
+
+Main files from this Script:
+
+  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures
+
+  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information
+
+  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)
+
+  4. **MELT_ABS**: make melt version of foci positions for abs measuers
+
+  5. CELL level table (full all mice) (use this to make long biv dataset)
+
+  6. Long Biv Dataset
+
+  7. mouse table (full all mice)
+
+  8. **Strain DFs** for Het models (summary strain DFs and tables?)
+
+  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models
+
+  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**
+
+# Male specific BivData set
+
+not sure if SC.Length.male_table is used for anything.
+
+```{r male.sc.length.strain, echo=FALSE}
+
+#these measure ect should be in the male - mouse 
+#isolate all musc and male
+Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "M.musculus"),]
+
+#make strain average table
+SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
+                             mean_SC = mean(chromosomeLength),
+                              n_SC = length(Obj.ID),
+
+                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+)
+
+```
+
+# Cell Level Table
+
+```{r cell.level.table, echo=FALSE}
+
+cell.level.BivData_table <- ddply(.data=Curated_BivData, 
+                       .(fileName),
+                       summarize, 
+                       boxs.IDd = max(boxNumber),
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                       ncells = length(unique(fileName)),
+                       nbivs = length(unique(Obj.ID)),
+
+                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       var.SC_length = var(chromosomeLength),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number
+
+                       cv.SC = cv(chromosomeLength),
+
+#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
+                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),
+
+                #this below might not be correct
+                     mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
+                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
+                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)
+
+#cent_dis
+                      mean_cent_dist = mean(dis.cent, na.rm=TRUE),
+                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]], #calq threshold for all cells, pull out 'almost.whole.cells' later
+                      Qar.3 = quantile(chromosomeLength, na.rm = TRUE)[[3]],
+                      Qar.2 = quantile(chromosomeLength, na.rm = TRUE)[[2]],
+                      Qar.1 = quantile(chromosomeLength, na.rm = TRUE)[[1]]
+
+                        )
+
+cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
+cell.level.BivData_table <- add_strain(cell.level.BivData_table)
+cell.level.BivData_table <- add_category(cell.level.BivData_table)
+cell.level.BivData_table <- add_species(cell.level.BivData_table)
+
+#as.data.frame( head(cell.level.BivData_table) )
+#use quartille 2 as a min, on average it grabs 4 
+```
+
+```{r show.pass.rate, echo=FALSE}
+
+pass.rate.hist <- ggplot(cell.level.BivData_table, aes(x = nboxes_passed) )+geom_histogram(binwidth = 1, alpha=.6) + facet_wrap(~category)+ggtitle("number of objects ID'd per cell")
+pass.rate.hist
+
+```
+
+make histogram of the pass rate per cell.
+
+3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.  
+8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev
+
+# Reduced Bivalent Dataset (Long and Short)
+
+I want to make a dataset of 'longest' 25% chromosomes from cells with at least 16 segmented bivalents from a single cell.
+
+```{r Reduced.Biv.DataSet.chrm.dataset, echo=FALSE}
+
+#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
+over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
+BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]
+
+#check for blank obj.id
+BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")
+
+#check for dup
+#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
+#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
+#dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
+#no duplicates
+
+#remove duplicate obj.ids
+#BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
+#BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]
+
+#make sure obj.id and fileName are charcter for below
+BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
+BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)
+
+missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ] 
+#all these files stem from
+
+#these are bivalents which don't appear the in cell level table, which can break the loop
+BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
+#final cleaned BivData table, 2428
+
+#2. calq the Q4 threshold
+#loop for greping the mouse from file name ect
+#each Q is a quartille
+
+#mini.now  <- BivData_from_15biv_cells[1:300,]
+
+for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){
+
+  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_" )[[1]][1:7]
+  new.egg <- paste0(newfile.str, collapse = "_")
+
+  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
+  the.Q3 <-cell.level.BivData_table$Qar.3[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
+  the.Q2 <-cell.level.BivData_table$Qar.2[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
+  the.Q1 <-cell.level.BivData_table$Qar.1[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]
+
+  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
+    BivData_from_15biv_cells$cell.Q3[i] <- the.Q3
+  BivData_from_15biv_cells$cell.Q2[i] <- the.Q2
+  BivData_from_15biv_cells$cell.Q1[i] <- the.Q1
+
+  the.FN <- as.character(cell.level.BivData_table$fileName[which(grepl(new.egg, cell.level.BivData_table$fileName ) )] )
+
+  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)
+  BivData_from_15biv_cells$Q1.FN[i] <- as.character(the.FN)
+}
+
+#3.mark which bivalents are above the threshold
+
+BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= as.numeric(BivData_from_15biv_cells$cell.Q4)  ), "yes", "no" )
+
+real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]#713 observations
+
+#isolate the short bivdata
+BivData_from_15biv_cells$short.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength <= as.numeric(BivData_from_15biv_cells$cell.Q2)  ), "yes", "no" )
+
+real.short.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$short.Biv == "yes",]# 708 much few than
+
+```
+
+```{r Look.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}
+
+#showing SC lengths per cell, with lines for the quartille values.
+
+mini_G.15_15biv_FN <- BivData_from_15biv_cells %>% filter(category == "G male")
+#ploting short biv data
+G.15_15biv_male <- BivData_from_15biv_cells %>% filter(category == "G male")
+mini_G.15_15biv_FN <- sample( G.15_15biv_male$fileName, 10)
+mini_G.15_15biv  <- BivData_from_15biv_cells[BivData_from_15biv_cells$fileName %in% mini_G.15_15biv_FN, ]
+
+ppll <- ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
+  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
+  geom_hline(aes(yintercept= cell.Q2), color='red') +
+ geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
+ geom_hline(aes(yintercept= cell.Q4), color='black')
+
+  #this plot shows that the Q1 is capturing only single values -- I should try Q2
+#use Q2 as a 
+```
+
+```{r table.long.biv, echo=FALSE}
+
+# make a table for the Mixed Models -- from the long.BivData
+# Calq mouse averages
+
+#add in all the other stats there are in the other tables
+
+Long.biv_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                        #SC lengths by class
+                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
+                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+        #          mean_siscoten = mean(siscoten)
+        #
+)
+#mouse averages for long.biv
+
+Long.biv_mouse_table <- add_sex(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_strain(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_subsp(Long.biv_mouse_table)
+Long.biv_mouse_table <- add_category(Long.biv_mouse_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Long.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
+Long.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
+Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)
+
+#rec group
+Long.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Long.biv_mouse_table$category), 1, 
+                            ifelse(grepl("MSM", Long.biv_mouse_table$category), 1,
+                             ifelse(grepl("SKIVE", Long.biv_mouse_table$category), 1, 0)) )
+
+#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)
+
+#1) Long Biv for 
+#remove all the non-used strains!!
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table
+
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
+Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
+
+#remove na's
+Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]
+
+#table(droplevels( Long.biv_mouse_table_HetC$category) )  #for double checking
+
+#2) Long Biv for male.poly
+#just males, just Musculus
+Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
+
+Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
+
+#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
+```
+
+The above code chunk makes 2 mouse level tables; **Long.biv_mouse_table_HetC** and **Long.biv_mouse_table_male.poly**. Run these through the model systems.
+
+# Mouse Level Single BivData Table
+
+Mouse level tables are the main meat of the models. There will be several for use in different models. Format these tables for run the mixed models (unordered factors, ect).
+
+The block below creates clean.BivData, which removes ~600 rows if just SC measures
+(This might be important for the proportion code)
+
+```{r FULL.mouse.BivData.table, echo=FALSE}
+
+#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36
+
+table(Curated_BivData$hand.foci.count, exclude = NULL)
+#there are ~600 rows with SC and no hand foci count (blah)
+
+#remove bivalent entries without hand foci count values
+cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
+                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
+
+mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
+                       .(mouse),
+                       summarize,
+
+                       ncells = length(unique(fileName)),
+
+                       boxs.IDd = max(boxNumber),
+                       nboxes_passed = length(boxNumber),
+
+                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
+
+                      tot.biv = length( !is.na(hand.foci.count) ),
+
+                       n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
+                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(Nbiv), 3),
+
+                       #sis.co.ten
+                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
+)
+
+mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_category(mouse.level.BivData_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
+
+mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
+                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
+```
+
+```{r mouse.table.HetC.MM, echo=FALSE}
+
+#this mouse level table -- should not have strains without female obs
+#only Mmusculus
+
+Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(  unique(fileName) ),
+                      tot.biv = length(Obj.ID),
+
+                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#
+Mouse.Table_BivData_4MM <- add_sex(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_strain(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_subsp(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_category(Mouse.Table_BivData_4MM)
+Mouse.Table_BivData_4MM <- add_species(Mouse.Table_BivData_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Mouse.Table_BivData_4MM$subsp <- factor(Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
+Mouse.Table_BivData_4MM$sex <- factor(Mouse.Table_BivData_4MM$sex, ordered = FALSE)
+Mouse.Table_BivData_4MM$strain <- factor(Mouse.Table_BivData_4MM$strain, ordered = FALSE)
+
+#remove all the non-used strains!!
+
+Mouse.Table_BivData_4MM <-  Mouse.Table_BivData_4MM %>%  filter(strain != "AST")  %>% filter(strain != "TOM")  %>% filter(strain != "CZECH")
+
+#Mouse.Table_BivData_4MM$strain <- droplevels(Mouse.Table_BivData_4MM$strain)
+#Mouse.Table_BivData_4MM <- Mouse.Table_BivData_4MM[(!is.na(Mouse.Table_BivData_4MM$mouse)),] #so many NAs still
+
+Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Mouse.Table_BivData_4MM$category), 1, 
+                              ifelse(grepl("MSM male", Mouse.Table_BivData_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", Mouse.Table_BivData_4MM$category), 1, 0)))
+
+```
+
+```{r mouse.table.Male.Poly, echo=FALSE}
+
+#this table will have male BivData
+#it will include Musc strains
+
+Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]
+
+Male.poly.Mouse.Table_BivData_4MM <- ddply(Curated_BivData_male, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
+
+                       #calq mean.sc by chrm class
+                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                # IFD things
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+Male.poly.Mouse.Table_BivData_4MM <- add_sex(Male.poly.Mouse.Table_BivData_4MM)
+Male.poly.Mouse.Table_BivData_4MM <- add_strain(Male.poly.Mouse.Table_BivData_4MM)
+Male.poly.Mouse.Table_BivData_4MM <- add_subsp(Male.poly.Mouse.Table_BivData_4MM)
+Male.poly.Mouse.Table_BivData_4MM <- add_category(Male.poly.Mouse.Table_BivData_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Male.poly.Mouse.Table_BivData_4MM$subsp <- factor(Male.poly.Mouse.Table_BivData_4MM$subsp, ordered = FALSE)
+Male.poly.Mouse.Table_BivData_4MM$sex <- factor(Male.poly.Mouse.Table_BivData_4MM$sex, ordered = FALSE)
+Male.poly.Mouse.Table_BivData_4MM$strain <- factor(Male.poly.Mouse.Table_BivData_4MM$strain, ordered = FALSE)
+
+#remove females
+Male.poly.Mouse.Table_BivData_4MM$sex <- droplevels(Male.poly.Mouse.Table_BivData_4MM$sex)
+
+```
+
+analysis using the mouse level table -- to justify that using mouse averages is a good idea.
+
+# Melt Position Table
+
+The BivData MELT DF is primarily for the nrmF1.pos plots/analysis. It is a Melted DF for raw and normalized foci positions from the Curated_BivData.
+
+```{r MELT.block1, echo=FALSE}
+
+#not sure what the melt table is needed for
+col.names.list <- colnames(Curated_BivData)
+
+#all of the cols to use
+MELT_Curated_BivData <-   melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse", "category",      "strain",        "sex",           "subsp",         "species", "IFD3_ABS", "IFD1_ABS","IFD2_ABS","IFD1_PER",
+              #remoevd PER.Foci positions
+"dis.cent", "dis.cent.PER",  "telo_dist", "telo_dist_PER", "SisCoTen", "Rec.group",
+              #leave out the Foci
+                "Foci1","Foci2","Foci3", "Foci4")],
+
+            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
+       "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
+              "curated.1pass.0fail","curation.notes", "hand.foci.count",
+
+       "notes", "mouse", 
+      "category","strain","sex","subsp","species", "IFD3_ABS","IFD1_ABS","IFD2_ABS","IFD1_PER",
+              #remoevd PER.Foci positions
+"dis.cent", "dis.cent.PER",  "telo_dist", "telo_dist_PER", "SisCoTen", "Rec.group"))
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]
+
+#this might not be right
+colnames(MELT_Curated_BivData) <- c( col.names.list[c(1:11, 16:26, 31:36)], "Foci_Type", "Pos")
+
+#
+MELT_Curated_BivData$Pos <- as.numeric(MELT_Curated_BivData$Pos)
+
+#calq normalized melt pos
+MELT_Curated_BivData$PER.Foci.Pos <- MELT_Curated_BivData$Pos / MELT_Curated_BivData$chromosomeLength
+
+MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$Pos)),]
+
+```
+
+```{r SHORT.biv_setup, echo=FALSE, eval=TRUE}
+
+#move to Data.setup
+
+#FROM q1
+#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+#from Q2 IFD
+#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+##long biv
+#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
+#table( droplevels( male.long.biv$strain)  )
+
+#FROM q1 
+REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
+                         mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                      #make sure 1CO  
+                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
+                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
+                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                         )
+
+REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
+REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
+
+REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
+REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
+REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
+                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
+                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
+                                           0)))
+
+###LONG BIV
+
+REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
+                       n_obs = length(Obj.ID),
+                        ncells = length(unique(fileName)),
+                         mean_SC = mean(chromosomeLength),
+
+                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
+                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
+                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+                         )
+
+REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
+REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
+
+REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
+REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
+REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
+                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
+                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
+                                           0)))
+
+#find which one
+
+#REAL.short_mouse_table
+```
+
+# Total SC
+
+total sc is processed in the results setup.
+
 # Saving Data file

### 2020-05-16 Chap3_MLH1var.rmd
notes and updates of the chapter 3 variances

@@
 ---
-title: "betweenCell.Var"
+title: "Between Cell Variance"
 author: "April Peterson"
-date: "5/5/2020"
+date: 
 output: html_document
 ---
-
-load all the file set ups
 
@@
 #library(pwr)
-
+library(purrr) #for replicate
 library(dplyr)
@@
 
-plot the variances
+```{r female.strain.table, echo=FALSE}
 
-```{r female.strain.table, echo=FALSE}
 female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
@@
 
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=TRUE}
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
 
 #variance plots
-musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
+musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse \nMLH1 variance")+scale_color_manual(values=colors_of_strains)
 
-Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.female.VAR.plot
+Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female\nwithin mouse MLH1 variance")+scale_color_manual(values=colors_of_strains)
 
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse \nMLH1 variance") + scale_color_manual(values=colors_of_strains)
 
@@
 #same issue as the female table .. ugh
+
 ```
 
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
 
 #variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.male.VAR.plot
+OG.male.VAR.plot <- ggplot(M.strain.table, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
 
 Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.male.VAR.plot
 
 Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Dom.male.VAR.plot
+
 ```
 
-go over the main patterns (from the model framework)
+> How much do CO counts vary among cells from the same individual?
 
-Bret's suggestions
+> What factors shape between-cell variance in CO count?
+
+Answering these will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution
+
+Below is the table for the number of cells per category by quality score. For these between cell variance will just use those in quality 1-2.
+
+```{r OG.table.qual.num, echo=FALSE}
+
+table(MLH1_data$quality, MLH1_data$category, exclude = NULL)
+
+```
+
+# Q12 Varince Plots
+
+The plots below show each point as a mouse-level CO count variance, calculated from Quality 1-2 cells. The number of cells per mouse is indicated by size.
+
+```{r Q12.table.qual.num, echo=FALSE, warning=FALSE}
+
+#high quality cells
+Q12.MLH1_data <- MLH1_data %>% filter(quality < 3 )
+
+Q12.MLH1_data$fileName <- Q12.MLH1_data$Original.Name
+
+#table(Q12.MLH1_data$quality, Q12.MLH1_data$category, exclude = NULL)
+#all this information is in the top table
+
+#make mouse level table
+FULL_mouse_table_Q12.MLH1 <- ddply(Q12.MLH1_data, c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci),
+                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+)
+
+FULL_mouse_table_Q12.MLH1 <- add_strain(FULL_mouse_table_Q12.MLH1)
+FULL_mouse_table_Q12.MLH1 <- add_subsp(FULL_mouse_table_Q12.MLH1)
+FULL_mouse_table_Q12.MLH1 <- add_sex(FULL_mouse_table_Q12.MLH1)
+FULL_mouse_table_Q12.MLH1 <- add_category(FULL_mouse_table_Q12.MLH1)
+FULL_mouse_table_Q12.MLH1 <- add_species(FULL_mouse_table_Q12.MLH1)
+
+```
+
+```{r Q12.mouse.var.plots.male, echo=FALSE, fig.width = 6, fig.height=6, warning=FALSE}
+
+#variance plots
+#musc.male.VAR.plot  #%>% filter(subsp == "Musc")
+
+FULL.male.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus") ),
+                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))
+
+Dom.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Dom")  ),
+                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+mol.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Mol")  ),
+                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+FULL.male.Var.plot
+```
+
+```{r Q12.mouse.var.plots.female, echo=FALSE, fig.width = 5, fig.height=5, warning=FALSE}
+
+FULL.female.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") ),
+                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))
+
+#variance plots
+musc.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Musc")  ),
+                             aes(color= strain, size = Ncells))+
+  geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+Dom.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Dom")  ),
+                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+mol.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Mol")  ),
+                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+FULL.female.Var.plot
+```
+
+The figure above nominates MOLF should be removed from the full dataset. I think all cells can be included for the sex specific data.
+
+The female LEW strains have a larger distribution of variance.
+(or it could just be the 2 mice).
+Some of the low cell number mouse.
+(note on the dot size, small dots could be mice with few high Q cells.
+Med size cells have a higher number of cells.)
+
+# Models (var)
+
+The same glm models are being applied to variance and cV (M1 to M3) and sex specific models (M4 and M5). Out groups and strains with just one sex have been removed for M1-M3. Al factors are unordered.
+
+```{r model.set.up, echo=FALSE}
+
+#FULL_mouse_table_Q12.MLH1
+#sex-specific
+
+FULL_mouse_table_Q12.MLH1$subsp <- factor(FULL_mouse_table_Q12.MLH1$subsp, ordered = FALSE)
+FULL_mouse_table_Q12.MLH1$sex <- factor(FULL_mouse_table_Q12.MLH1$sex, ordered = FALSE)
+FULL_mouse_table_Q12.MLH1$strain <- factor(FULL_mouse_table_Q12.MLH1$strain, ordered = FALSE)
+
+#remove all the mice with fewer than 3cells
+One.cell.mice <- FULL_mouse_table_Q12.MLH1 %>% filter(Ncells == 1)
+oo <- FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$mouse %in% One.cell.mice$mouse, ]
+```
+
+```{r varMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
+
+#FULL_mouse_table_Q12.MLH1, remove molf not enough female
+
+#DF.for.model <- ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )
+
+#these are all mouse level values
+
+varMLH1.M1 <- lmer(var ~ subsp * sex + (1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )  )
+
+varMLH1.MM.M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )
+
+varMLH1.MM.M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )
+
+#summary(varMLH1.M1)$coefficients
+
+varMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
+varMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+varMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+varMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value
+
+varMLH1.MMlmer_results <- data.frame(sub.sp = varMLH1.MM_subp, 
+sex.results =  varMLH1.MM_sex, 
+               interaction.pval = varMLH1.MM_interaction,  rand.pval = 
+varMLH1.MM_strain)
+
+```
+
+**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain `r varMLH1.MMlmer_results`. (Sex is the most significant).
+
+**M2**
+```{r varMLH1.glm.M2, echo=FALSE, eval=TRUE}
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
+#strain = fixed, 
+varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )
+
+summ.varMLH1.M2.glm.strain.fixed.full <- summary(varMLH1.M2.glm.strain.fixed.full)
+round(summ.varMLH1.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
+#adding cell number has some significant parts
+```
+
+**M3**
+```{r varMLH1.glm.M3, echo=FALSE, eval=TRUE}
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+varMLH1.M3.strain.fixed <- glm(var ~ sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ] )
+
+summ.MLH1.M3.strain.fixed <- summary(varMLH1.M3.strain.fixed)
+
+round(summ.MLH1.M3.strain.fixed$coefficients[,c(1,4)],5)
+#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+```
+
+LEW and G are slightly significant strain p values. Sex is also slightly significant. Intercept is also significant.
+
+**M4-M5 female**
+```{r Female.specific.MLHvar.glm, echo=FALSE, eval=TRUE}
+
+Var.female.strain.subsp <- glm(var ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )
+
+Var.summry.female.strain.effects <- summary(Var.female.strain.subsp)
+round(Var.summry.female.strain.effects$coefficients[,c(1,4)],5)
+
+varMLH1.female.glm.M2 <- glm(var ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )
+
+sum_varMLH1.female.spc.glm.M2 <- summary(varMLH1.female.glm.M2)
+round(sum_varMLH1.female.spc.glm.M2$coefficients[,c(1,4)],5)
+```
+
+Only the intercept's are significant in the female specific models. None of the subsp or strain effects are significant.
+
+**M4-M5 male**
+```{r Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}
+
+#M4
+varMLH1.male.glm.M4 <- glm(var ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
+sum_varMLH1.male.glm.M4 <- summary(varMLH1.male.glm.M4)
+round(sum_varMLH1.male.glm.M4$coefficients[,c(1,4)],5)
+
+#M4.2
+varMLH1.male.glm.M4.2 <- glm(var ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
+
+sum_varMLH1.male.glm.M4.2 <- summary(varMLH1.male.glm.M4.2)
+round(sum_varMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)
+```
+
+There are no signifcant subspecies + strain effects for the male specific data.
+
+# Models (cV)
+```{r cVMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
+
+#FULL_mouse_table_Q12.MLH1, remove molf not enough female
+
+#DF.for.model <- ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )
+
+#these are all mouse level values
+
+cVMLH1.M1 <- lmer(cV ~ subsp * sex + (1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )  )
+
+cVMLH1.MM.M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data= ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )
+
+cVMLH1.MM.M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  ) )
+
+#summary(varMLH1.M1)$coefficients
+
+cVMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
+cVMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+cVMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value
+
+cVMLH1.MMlmer_results <- data.frame(sub.sp = cVMLH1.MM_subp, 
+sex.results =  varMLH1.MM_sex, 
+               interaction.pval = cVMLH1.MM_interaction,  rand.pval = 
+cVMLH1.MM_strain)
+
+```
+
+**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain `r cVMLH1.MMlmer_results`. (Sex is the most significant).
+
+**M2**
+```{r cVMLH1.glm.M2, echo=FALSE, eval=TRUE}
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
+#strain = fixed, 
+cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )
+
+summ.cvMLH1.M2.glm.strain.fixed.full <- summary(cVMLH1.M2.glm.strain.fixed.full)
+round(summ.cvMLH1.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
+
+#adding cell number has some significant parts
+```
+
+**M3**
+```{r cVMLH1.glm.M3, echo=FALSE, eval=TRUE}
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+cVMLH1.M3.strain.fixed <- glm(cV ~ sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ] )
+
+summ.cV.MLH1.M3.strain.fixed <- summary(cVMLH1.M3.strain.fixed)
+
+round(summ.cV.MLH1.M3.strain.fixed$coefficients[,c(1,4)],5)
+
+```
+
+LEW and G are slightly significant strain p values. Sex is also slightly significant. Intercept is also significant.
+
+**M4-M5 female**
+```{r Female.specific.MLHcV.glm, echo=FALSE, eval=TRUE}
+
+cV.female.strain.subsp <- glm(cV ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )
+
+cV.summry.female.strain.effects <- summary(cV.female.strain.subsp)
+round(cV.summry.female.strain.effects$coefficients[,c(1,4)],5)
+
+cVMLH1.female.glm.M2 <- glm(cV ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )
+
+sum_cVMLH1.female.spc.glm.M2 <- summary(cVMLH1.female.glm.M2)
+round(sum_cVMLH1.female.spc.glm.M2$coefficients[,c(1,4)],5)
+
+```
+
+Only the intercept's are significant in the female specific models. None of the subsp or strain effects are significant.
+
+**M4-M5 male**
+```{r Male.specific.MLHcV.glm, echo=FALSE, eval=TRUE}
+
+#M4
+cVMLH1.male.glm.M4 <- glm(cV ~ subsp * strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
+sum_varMLH1.male.glm.M4 <- summary(varMLH1.male.glm.M4)
+round(sum_varMLH1.male.glm.M4$coefficients[,c(1,4)],5)
+
+#M4.2
+cVMLH1.male.glm.M4.2 <- glm(cV ~ strain, data= (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
+
+sum_cVMLH1.male.glm.M4.2 <- summary(cVMLH1.male.glm.M4.2)
+round(sum_cVMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)
+
+```
+
+<!-- Bret's suggestions / Permutations
 
@@
 2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.
+
 3. Randomly permute cells among strains, keeping the structure of the dataset otherwise the same.
+
 4. For each permutation, compute the among-strain variance in the MLH1 count variance (separately for each sex).
@@
 Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).
+-->
 
-```{r bret.filters, echo=FALSE}
+# Permutations
+
+Calculate the category level between cell variance (and cV).
+
+```{r pool.cell.calq, echo=FALSE}
+
+OG.vals_Q12.MLH1 <- ddply(Q12.MLH1_data, c("strain", "sex"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        cV = cv(nMLH1.foci),
+                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+)
+
+OG.vals_Q12.MLH1 <- add_subsp(OG.vals_Q12.MLH1)
+OG.vals_Q12.MLH1 <- add_species(OG.vals_Q12.MLH1)
 
@@
 
-# Identifying mice with largest sample sizes
+Made a table with the cell-pooled calculations of var and cV.
+Next step is to i) record the data structure ii) then permute within sex, then calculate the cell-pooled values.
 
-I have a hunch that using these mice might be the best (least risky way forward)
+```{r permutations, echo=FALSE, eval=FALSE}
+
+Q12.permute.DF <- Q12.MLH1_data %>% filter()
+
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
+Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]
+
+#Data structure
+len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
+len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )
+
+len.G.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
+len.G.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )
+
+len.LEW.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
+len.LEW.male <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "male") )$fileName )
+
+len.PWD.female <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "female") )$fileName )
+len.PWD.male <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "male") )$fileName )
+
+len.KAZ.female <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "female") )$fileName )
+len.KAZ.male <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "male") )$fileName )
+
+len.SKIVE.female <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$fileName )
+len.SKIVE.male <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$fileName )
+
+len.MSM.female <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "female") )$fileName )
+len.MSM.male <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "male") )$fileName )
+
+#these are the bones of the structure -- for permuting a new DF -- then calculating
+
+#this DF is kinda hard, I'm not sure how to make a full randomized DF
+
+#replicate(n = 3, rnorm(5, 0, 1), simplify = FALSE )
+
+nrepeat = 50
+
+permute.var <- data.frame( "wsb.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.WSB.female, replace = FALSE) ) ),
+
+                            "wsb.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
+  "lew.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.LEW.female, replace = FALSE) ) ),
+
+  "lew.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.LEW.male, replace = FALSE) ) ),
+
+  "g.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.G.female, replace = FALSE) ) ), 
+  "g.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.G.male, replace = FALSE) ) ),
+
+  "pwd.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
+  "pwd.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.PWD.male, replace = FALSE))),
+
+  "kaz.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.KAZ.female,replace = FALSE) ) ),
+
+  "kaz.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.KAZ.male, replace = FALSE)) ),
+
+  "skive.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),
+
+  "skive.m.var" =  replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),
+  "msm.f.var" = replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.MSM.male, replace = FALSE) )   )  )
+
+#now that these is a dataframe of new/permuted variance (specific to each sex) ... compare them to the origincal values?
+#my prediction is that there would be many significant differences from the true values
+
+#to visualize, I can plot the variances .. I guess with a histogram
+
+#now automate making these plots -- or calculate the propotion of permutations which are below the observed
+
+pp <- ggplot(permute.var, aes(x = wsb.f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )
+
+#randomized data frame, sample mouse means from all P.leucopus mice
+
+#calculate pvalue as proportion of differences greater than observed difference
+p.val_permut_mean.lab.wild <- length(permute.dif.meanOmeans$permuted.diff[permute.dif.meanOmeans$permuted.diff >= obs.dif.mean.lab.wild]) / permut.x
+var.pval <- paste( "p-value ", p.val_permut_mean.lab.wild)
+
+#visualize the permuted distribution
+permute_hist_means  <-  ggplot(permute.dif.meanOmeans,aes(x=permuted.diff))+
+  geom_histogram()+geom_vline(xintercept=obs.dif.mean.lab.wild,color='red') +
+  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of mouse means, wild to lab")
+
+#answer: The difference in CO means between wild and LL Pleucopus, is greater than if mice are randomized.
+
+#Same permutations, at the cell level
+Pl.cells <- MLH1_data.hq[MLH1_data.hq$species == "Pleuc",]
+
+obs.status.cell.diff <- mean(Pl.cells$nMLH1.foci[Pl.cells$status == "wild"])-
+    mean(Pl.cells$nMLH1.foci[Pl.cells$status == "lab"])
+
+permute.dif.cell.status  <- as.data.frame( replicate(permut.x, 
+                                  (  mean(sample(Pl.cells$nMLH1.foci,  243, replace = FALSE) ) - 
+                                   mean(sample(Pl.cells$nMLH1.foci, 221, replace = FALSE) ) )  
+) )
+
+colnames(permute.dif.cell.status) <- "permuted.diff"
+p.val_permut_cell.diff.lab.wild <- length(permute.dif.cell.status$permuted.diff[permute.dif.cell.status$permuted.diff >= obs.status.cell.diff]) / permut.x
+var.pval <- paste( "p-value ", p.val_permut_cell.diff.lab.wild)
+
+permute_hist_means.cell_level  <-  ggplot(permute.dif.cell.status, aes(x=permuted.diff))+
+  geom_histogram()+geom_vline(xintercept=obs.status.cell.diff,color='red') +
+  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of cell level means, wild to lab")
+
+```

### 2020-05-18 Chap3_MLH1var.rmd
corrected permutation and replicate number

@@
 library(lattice)
-
+library(tidyr)
 library(raster)
@@
 library(xlsx)
+
+library(ggpubr) #for arranging ggplots
 
@@
 
-Answering these will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution
+The approach will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution.
 
@@
 
-# Q12 Varince Plots
+# Q12 Variance Plots
 
@@
 
-The figure above nominates MOLF should be removed from the full dataset. I think all cells can be included for the sex specific data.
-
-The female LEW strains have a larger distribution of variance.
-(or it could just be the 2 mice).
-Some of the low cell number mouse.
-(note on the dot size, small dots could be mice with few high Q cells.
-Med size cells have a higher number of cells.)
-
 # Models (var)
@@
 #strain = fixed, 
-varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )
+varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )
 
@@
 **M4-M5 male**
+
 ```{r Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}
@@
 **M2**
+
 ```{r cVMLH1.glm.M2, echo=FALSE, eval=TRUE}
@@
 #strain = fixed, 
-cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )
+cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )
 
@@
 **M3**
+
 ```{r cVMLH1.glm.M3, echo=FALSE, eval=TRUE}
@@
 **M4-M5 female**
+
 ```{r Female.specific.MLHcV.glm, echo=FALSE, eval=TRUE}
@@
 **M4-M5 male**
+
 ```{r Male.specific.MLHcV.glm, echo=FALSE, eval=TRUE}
@@
 sum_cVMLH1.male.glm.M4.2 <- summary(cVMLH1.male.glm.M4.2)
+
 round(sum_cVMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)
@@
 1. Consider only high-quality cells.
+
 2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.
@@
 Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).
+
 -->
@@
 
-Calculate the category level between cell variance (and cV).
+We'd like to know if the set of strains has bigger differences in between-cell variance than expected by chance.
+Your liner models provide one test, but they could suffer from small sample sizes per mouse.
+We can conduct a permutation test that enables you to pool all cells in each strain to increase the sample size.
+
+#Permutation Step Up
+
+Here's the permutation test I was imagining:
+1. Separate the sexes.
+
+2. Pool all cells within a strain and compute the **variance for each strain**.
+
+3. Compute the **variance across strains** (VAS) of the variance within strains. This will leave you with a single estimate for the set of strains.
+
+4. Randomly permute cells across strains. (pool all male cells, pool all female cells)
+
+5. For the permuted data set, pool all cells within a strain and compute the variance for each strain. (compute a variance for a strain, var( sample( all.female.cells, N.WSB.f ) ) ).
+
+6. For the permuted data set, compute the variance across strains of the variance within strains. This will leave you with a single estimate for the set of strains.
+
+7. Repeat 4-6 10,000 times to establish a null distribution for the variance in variances.
+
+8. Compare observed variance in variances to null distribution.
+
+#observed values
 
@@
                         Ncells  = length(nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
+                        mean_co = as.numeric(format(round(mean(nMLH1.foci), 3), nsmall=3)),
                         cV = cv(nMLH1.foci),
@@
 
+#remove extra strains
+remove.strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
+vals_Q12.MLH1 <- OG.vals_Q12.MLH1[!OG.vals_Q12.MLH1$strain %in% remove.strains, ]
+
+#for males and females, calq the total variance across the strain-specific var
+Obs.f.VAS <- var( ( vals_Q12.MLH1 %>% filter(sex == "female") )$var )
+Obs.m.VAS <- var( ( vals_Q12.MLH1 %>%  filter(sex == "male") )$var )
+
 ```
 
-Made a table with the cell-pooled calculations of var and cV.
-Next step is to i) record the data structure ii) then permute within sex, then calculate the cell-pooled values.
+These are the female strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$var`, the variance across strains (VAS) is `r Obs.f.VAS`.
 
-```{r permutations, echo=FALSE, eval=FALSE}
+These are the male strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "male") )$var`, the variance across strains (VAS) is `r Obs.m.VAS`.
+(For `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$strain` respectively)
 
-Q12.permute.DF <- Q12.MLH1_data %>% filter()
+```{r str.of.DF, echo=FALSE, eval=TRUE}
+#redundant
+#remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
+#Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]
 
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
-Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]
-
-#Data structure
+#Data structure (from the cell level data)
 len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
@@
 
-#these are the bones of the structure -- for permuting a new DF -- then calculating
+```
 
-#this DF is kinda hard, I'm not sure how to make a full randomized DF
+<!--
+random permutations below, (COMPUTE the VAS)
 
-#replicate(n = 3, rnorm(5, 0, 1), simplify = FALSE )
+For the permuted data set, **pool all cells within a strain (permuted strain ie defined by sample size?) and compute the variance for each strain**.
 
-nrepeat = 50
+6. For the permuted data set, compute the variance across strains of the variance within strains. This will leave you with a single estimate for the set of strains.
 
-permute.var <- data.frame( "wsb.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.WSB.female, replace = FALSE) ) ),
+7. Repeat 4-6 10,000 times to establish a null distribution for the variance in variances.
 
-                            "wsb.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
-  "lew.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.LEW.female, replace = FALSE) ) ),
+8. Compare observed variance in variances to null distribution.  -->
 
-  "lew.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.LEW.male, replace = FALSE) ) ),
+# Permutation Output
 
-  "g.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.G.female, replace = FALSE) ) ), 
-  "g.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.G.male, replace = FALSE) ) ),
+```{r permutations2, echo=FALSE}
+#I made a DF of strain X sex specific variance
 
-  "pwd.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
-  "pwd.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.PWD.male, replace = FALSE))),
+nrepeat = 10000
 
-  "kaz.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.KAZ.female,replace = FALSE) ) ),
+#I need to sample within strain? -- then I feel like I need to change the sample size -- otherwise I'm just sampling all the same episodes
 
-  "kaz.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.KAZ.male, replace = FALSE)) ),
+permuted.var <- data.frame( "wsb_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$sex == "female")], len.WSB.female, replace = FALSE) ) ),
+                            "wsb_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
+  "lew_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.LEW.female, replace = FALSE) ) ),
 
-  "skive.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),
+  "lew_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.LEW.male, replace = FALSE) ) ),
 
-  "skive.m.var" =  replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),
-  "msm.f.var" = replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.MSM.male, replace = FALSE) )   )  )
+  "g_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.G.female, replace = FALSE) ) ), 
+  "g_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.G.male, replace = FALSE) ) ),
 
-#now that these is a dataframe of new/permuted variance (specific to each sex) ... compare them to the origincal values?
-#my prediction is that there would be many significant differences from the true values
+  "pwd_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
+  "pwd_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.PWD.male, replace = FALSE))),
 
-#to visualize, I can plot the variances .. I guess with a histogram
+  "kaz_f.var" = replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"],len.KAZ.female,replace = FALSE) ) ),
 
-#now automate making these plots -- or calculate the propotion of permutations which are below the observed
+  "kaz_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.KAZ.male, replace = FALSE)) ),
 
-pp <- ggplot(permute.var, aes(x = wsb.f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )
+  "skive_f.var" = replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),
 
-#randomized data frame, sample mouse means from all P.leucopus mice
+  "skive_m.var" =  replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),
 
-#calculate pvalue as proportion of differences greater than observed difference
-p.val_permut_mean.lab.wild <- length(permute.dif.meanOmeans$permuted.diff[permute.dif.meanOmeans$permuted.diff >= obs.dif.mean.lab.wild]) / permut.x
-var.pval <- paste( "p-value ", p.val_permut_mean.lab.wild)
+  "msm_f.var" = replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.MSM.male, replace = FALSE) )   )
+  )
 
-#visualize the permuted distribution
-permute_hist_means  <-  ggplot(permute.dif.meanOmeans,aes(x=permuted.diff))+
-  geom_histogram()+geom_vline(xintercept=obs.dif.mean.lab.wild,color='red') +
-  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of mouse means, wild to lab")
+```
+
+```{r functions.Chap3, echo=FALSE}
+#formate the permuted DF, adding strain and sex to the DF
+
+add_sex_chp3 <- function(dat){
+  dframe <- dat
+  dframe$sex <- "other"
+  dframe$sex <- ifelse(grepl("f.", dframe$names), "female", 
+                       ifelse(grepl("m.",dframe$names), "male", "other"))
+  dframe$sex<- factor(dframe$sex, ordered = TRUE, levels =c( "female", "male"))
+  return(dframe)
+}
+
+add_strain_Chap3 <- function(dat){
+  dframe <- dat
+  dframe$strain <- "other"
+  dframe$strain <- ifelse(grepl("wsb", dframe$names), "WSB",
+                    ifelse(grepl("g",dframe$names), "G",
+                   ifelse(grepl("lew", dframe$names), "LEW",
+                     ifelse(grepl("perc", dframe$names), "PERC",
+                     ifelse(grepl("msm", dframe$names), "MSM",
+                 ifelse(grepl("molf", dframe$names), "MOLF",
+                     ifelse(grepl("skive", dframe$names), "SKIVE",      
+                      ifelse(grepl("pwd", dframe$names), "PWD",
+                  ifelse(grepl("kaz", dframe$names), "KAZ",
+                   ifelse(grepl("tom", dframe$names), "TOM",
+                  ifelse(grepl("ast", dframe$names), "AST",
+                   ifelse(grepl("czech", dframe$names), "CZECH",
+                    ifelse(grepl("cast", dframe$names), "CAST",
+                   ifelse(grepl("hmi", dframe$names), "HMI",
+                       ifelse(grepl("_SPRET_", dframe$names), "SPRET",
+                       ifelse(grepl("_SPI_", dframe$names), "SPIC",
+                                                                                                                                          ifelse(grepl("spic", dframe$names), "SPIC",
+                                                                                                                                                 ifelse(grepl("_CAROLI_", dframe$names), "CAROLI",
+                                 ifelse(grepl("F1_", dframe$names), "F1",         
+                                   "other")))))))))))))))))))
+
+  #the ordering factor below deletes all strain entries
+  dframe$strain<- factor(dframe$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW", "PERC",
+                                                                   "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
+                                                                   "CAST", "HMI",
+                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
+  return(dframe)
+}
+
+```
+
+```{r pvals.VAS, echo=FALSE, warning=F, message=FALSE,  fig.width = 5, fig.height=5}
+
+#need to merge all the pvales into a DF
+#for(u in unique( OG.vals_Q12.MLH1$strain[ ( (OG.vals_Q12.MLH1$sex == "female") & (OG.vals_Q12.MLH1$species == "M.musculus") )] ) ){
+#  print(u)
+#  }
+
+transposed.permuted <- t(permuted.var)
+
+permuted.DF <- data.frame(names = row.names(transposed.permuted), transposed.permuted); rownames(transposed.permuted) <- NULL
+
+permuted.DF <- add_sex_chp3(permuted.DF)
+permuted.DF <- add_strain_Chap3(permuted.DF)
+
+# 
+VAS.F.permuted.DF <- permuted.DF %>% filter(sex == "female")
+VAS.M.permuted.DF <- permuted.DF %>% filter(sex == "male")
+
+#sampling.size <- nrepeat-3
+#careful aboout matching the size of the dataframe!!
+
+list.VAS_F <- c()
+for(ff in 1:length(VAS.F.permuted.DF[,2:9997]) ){
+  list.VAS_F[ff]  <- var( VAS.F.permuted.DF[,ff] )
+}
+#list of VAS
+list.VAS_F.DF <- data.frame( "female.VAS" =  list.VAS_F)
+
+list.VAS_M <- c()
+for(mm in 1:length(VAS.M.permuted.DF[,2:9997]) ){
+
+  perm.VAS.mm <- var( VAS.M.permuted.DF[,mm] )
+  list.VAS_M[mm]  <- perm.VAS.mm
+
+}
+#p value is calculated for the distribution.
+
+list.VAS_M.DF <- data.frame("male.VAS" = list.VAS_M)
+#cbind the 2 VAS DF
+Full.VAS <- cbind( list.VAS_F.DF, list.VAS_M.DF)
+
+male.plot <- ggplot(Full.VAS, aes(x = male.VAS)) + geom_histogram() + geom_vline(xintercept = Obs.m.VAS ) + ggtitle("Male Variance Across Species")
+
+female.plot <- ggplot(Full.VAS, aes(x = female.VAS)) + geom_histogram() + geom_vline(xintercept = Obs.f.VAS ) + ggtitle("Female Variance Across Species")
+
+male.plot
+female.plot
+```
+
+The plots above show the distributions of permuted VAS for 10,000 permutations, the vertical line shows the observed VAS.
+
+The proportion of permutations which are less than the oberserved VAS are `r sum(  Full.VAS$male.VAS <= Obs.m.VAS ) / nrepeat` and `r sum(  Full.VAS$female.VAS <= Obs.f.VAS ) / nrepeat`  for male and female respectively
+
+```{r old.pval, echo=FALSE, eval=F}
+
+F.strain.var <- c(0)
+counter = 1
+for(u in unique(permuted.DF$strain[permuted.DF$sex == "female"])){
+  # print(u)
+
+  F.strain.var[counter] <-  sum( ( permuted.DF %>% filter(sex == "female")  %>% filter(strain == u )  )[,2:1000]  <= ( OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "female") )$var  ) / nrepeat
+  counter = counter + 1
+  #print(counter)
+}
+names(F.strain.var) <- unique( permuted.DF$strain[permuted.DF$sex == "female"]  )
+
+#male loop
+M.strain.var <- c(0)
+counter = 1
+for(u in unique(permuted.DF$strain[permuted.DF$sex == "male"])){
+  # print(u)
+  # print( sum( ( permuted.DF %>% filter(sex == "male")  %>% filter(strain == u )  )[,2:1000]  <= (    OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "male") )$var  ) / nrepeat  )
+
+  M.strain.var[counter] <-  sum( ( permuted.DF %>% filter(sex == "male")  %>% filter(strain == u )  )[,2:1000]  <= ( OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "male") )$var  ) / nrepeat
+
+  counter = counter + 1
+  #print(counter)
+}
+names(M.strain.var) <- unique( permuted.DF$strain[permuted.DF$sex == "male"]  )
+
+df.M.strain.var <- data.frame( vals = M.strain.var, strain = names(M.strain.var) )
+df.F.strain.var <- data.frame( vals = F.strain.var, strain = names(F.strain.var) )
+
+df.M.strain.var$sex <- "male"
+df.F.strain.var$sex <- "female"
+
+row.names(df.M.strain.var) <- c()
+row.names(df.F.strain.var) <- c()
+
+Strain.var.full <- rbind(df.M.strain.var, df.F.strain.var)
+
+```
+
+```{r p.vals, echo=FALSE, eval=FALSE}
+
+p.val.WSB.f <- sum( (permuted.var$wsb.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )) / nrepeat
+
+p.val.WSB.m <- sum( (permuted.var$wsb.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "male") )$var )) / nrepeat
+
+p.val.G.f <- sum( (permuted.var$g.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "female") )$var )) / nrepeat
+p.val.G.m <- sum( (permuted.var$g.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "male") )$var )) / nrepeat
+
+p.val.LEW.f <- sum( (permuted.var$lew.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "female") )$var )) / nrepeat
+p.val.LEW.m <- sum( (permuted.var$lew.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "male") )$var )) / nrepeat
+
+p.val.KAZ.f <- sum( (permuted.var$kaz.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "female") )$var )) / nrepeat
+p.val.KAZ.m <- sum( (permuted.var$kaz.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "male") )$var )) / nrepeat
+
+p.val.SKIVE.f <- sum( (permuted.var$skive.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$var )) / nrepeat
+p.val.SKIVE.m <- sum( (permuted.var$skive.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$var )) / nrepeat
+
+p.val.PWD.f <- sum( (permuted.var$pwd.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "female") )$var )) / nrepeat
+p.val.PWD.m <- sum( (permuted.var$pwd.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "male") )$var )) / nrepeat
+
+#calculate p values
+p.val.MSM.f <- sum( (permuted.var$msm.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "female") )$var )) / nrepeat
+
+p.val.MSM.m <- sum( (permuted.var$msm.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "male") )$var )) / nrepeat
+
+```
+
+```{r plots.var.dist, echo=FALSE, eval=F, warning=FALSE, message=FALSE}
+
+#good ref
+#http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/
+
+wsb.fff <- ggplot(permuted.var, aes(x = wsb_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var ) +  
+  geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "WSB") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+wsb.mmm <- ggplot(permuted.var, aes(x = wsb_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "WSB") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+lew.fff <- ggplot(permuted.var, aes(x = lew_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "female") )$var ) + geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "LEW") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+#lew female observed var is greater than randomized
+lew.mmm <- ggplot(permuted.var, aes(x = lew_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "male") )$var ) + geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "LEW") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+g.fff <- ggplot(permuted.var, aes(x = g_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "G") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+#lew female observed var is greater than randomized
+g.mmm <- ggplot(permuted.var, aes(x = g_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "G") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+kaz.fff <- ggplot(permuted.var, aes(x = kaz_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "KAZ") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+kaz.mmm <- ggplot(permuted.var, aes(x = kaz_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "KAZ") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+pwd.fff <- ggplot(permuted.var, aes(x = pwd_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "PWD") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+#pwd female oberserved variance is lower than the permutations.
+pwd.mmm <- ggplot(permuted.var, aes(x = pwd_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "PWD") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+skive.fff <- ggplot(permuted.var, aes(x = skive_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "SKIVE") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+skive.mmm <- ggplot(permuted.var, aes(x = skive_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "SKIVE") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+msm.mmm <- ggplot(permuted.var, aes(x = msm_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "MSM") ) , aes(x=18, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
+
+msm.fff <- ggplot(permuted.var, aes(x = msm_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "MSM") ) , aes(x=18, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)
 
@@
 
-#Same permutations, at the cell level
-Pl.cells <- MLH1_data.hq[MLH1_data.hq$species == "Pleuc",]
+permutation.results <- ggarrange(wsb.fff, wsb.mmm,
+          lew.fff, lew.mmm,
+          g.fff, g.mmm, 
+          kaz.fff, kaz.mmm,
+          pwd.fff, pwd.mmm,
+          skive.fff, skive.mmm, 
+           msm.fff, msm.mmm + rremove("x.text"), 
+          ncol = 4, nrow = 4)
 
-obs.status.cell.diff <- mean(Pl.cells$nMLH1.foci[Pl.cells$status == "wild"])-
-    mean(Pl.cells$nMLH1.foci[Pl.cells$status == "lab"])
+annotate_figure(
 
-permute.dif.cell.status  <- as.data.frame( replicate(permut.x, 
-                                  (  mean(sample(Pl.cells$nMLH1.foci,  243, replace = FALSE) ) - 
-                                   mean(sample(Pl.cells$nMLH1.foci, 221, replace = FALSE) ) )  
-) )
-
-colnames(permute.dif.cell.status) <- "permuted.diff"
-p.val_permut_cell.diff.lab.wild <- length(permute.dif.cell.status$permuted.diff[permute.dif.cell.status$permuted.diff >= obs.status.cell.diff]) / permut.x
-var.pval <- paste( "p-value ", p.val_permut_cell.diff.lab.wild)
-
-permute_hist_means.cell_level  <-  ggplot(permute.dif.cell.status, aes(x=permuted.diff))+
-  geom_histogram()+geom_vline(xintercept=obs.status.cell.diff,color='red') +
-  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of cell level means, wild to lab")
+  permutation.results, top = text_grob("permutations", color = "Black", face = "bold", size = 14) )
 

### 2020-05-19 doc/MS_v2.Rmd
start integrating Bret's new edits

--- 2020-05-14 doc/MS_v1.Rmd
@@
 ---
+
+<v2 - Bret's edits, v1>
 
@@
 
-```{r Fig3.example.double.triangle, echo=FALSE}
-#load an example two triangle plot
-
-single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-```
-
 ```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8, fig.cap= 'caption'}
@@
 
-```{r Q1.ggplot_IFD.triangle, echo=FALSE}
-#IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
-#  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-#  geom_jitter(alpha=.4)+
-#  ggtitle("Normalized IFD\nDom")+
-#  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
-#  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)
-
-```
+# IFD double triangle
+
+# Figure 1
 
@@
 #this works
-top.triangle2 <- function(my.y, my.x, my.data) {
+top.triangle <- function(my.y, my.x, my.data) {
   plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
@@
 
-bottom.triangle2 <- function(my.y, my.x, my.data){
+bottom.triangle <- function(my.y, my.x, my.data){
   points(my.x ~ my.y, data=my.data, xpd=TRUE, 
@@
 }
-
 ```
@@
 
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
 
@@
 
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
 
@@
 High- and low-recombining groups are indicated by 1 and 0 respectively.
-(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell"}
+(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell", echo=FALSE}
 
@@
 
-Q2.show.nrmF1
-
 nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
@@
 
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively).
+The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively).
 While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.
@@
 
+# rewrite double triangle
+
 ```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE}
@@
 
-  mtext("Position.2.Male", 1, 1.5, at= mean(par()$usr[1:2])) #at should be changing the position
+  mtext("Position.2.Male", 1, -2, at= mean(par()$usr[1:2]) + 0)
+
+  #the second val will move the x lable when negative
+
+  #at should be changing the position
   #above is just changing the text not the numbers
@@
 
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +4)
+  mtext("Position.1.Male", 4, 0, padj=par()$usr[1] + .5) #last value at .5 is good for moving y label
 
   x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
+
   axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-#las, orientation, at location of line
-  #labels automatic labels
-  #pos
-  #xpd - clipping
-
-   mtext(seq(0, 1, .2), 1, 1, at=x.at)
-  #second / third val negative moves the valu up
-
-  axis(4, las=1, pos=1) #this is the y axis
- #   axis(4, las=1, pos=1 + x.dist)
-  #this is calling axis twice, not sure why
+       labels=T, xpd=TRUE)
+#changing labels=T, puts the ddefault labels back
+
+  #second value should be around .75 (moves label text closer)
+  axis(4, las=1, pos=1, at=x.at) #this is the y axis
   lines(0:1, 0:1, xpd=TRUE) #diagonal line
@@
   #adjusting this mar lowers the margins (the edges)
-  op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))
-
-  #   op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
+  op <- par(mar=c(.25, .15, .25, .75), oma=c(2, 0, 0, 2))
+#(1=bottom, 2=left, 3=top, 4=right)
+
 Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
@@
 
+```{r Q2.load.bottom.tri, echo=FALSE}
+
+#only load, SKIVE, PWD, MSM
+#MOLF, KAZ, LEW
+Q2.bottom.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
+Q2.bottom.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
+Q2.bottom.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
+Q2.bottom.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
+Q2.bottom.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
+Q2.bottom.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+#EMPTY GGPLOT
+Q2.LEW.bottom.tri <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(Q2.bottom.LEW) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
 ```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
 
-Q2.plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_WSB.png')
-Q2.plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
-Q2.plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_G.png')
-
-Q2.plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
-Q2.plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
-Q2.plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
-
-Q2.plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
-Q2.plot8.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
-
-grid.arrange(rasterGrob(Q2.plot1.WSB), rasterGrob(Q2.plot2.LEW), rasterGrob(Q2.plot3.G),
-             rasterGrob(Q2.plot4.KAZ), rasterGrob(Q2.plot5.SKIVE), rasterGrob(Q2.plot6.PWD),
-             rasterGrob(Q2.plot7.MSM), rasterGrob(Q2.plot8.MOLF), 
-             ncol=3)
+Q2.SC_dif.Figure3  <- ggdraw() +
+
+  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
+  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
+
+  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
+  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
+
+  #drawing triangles in the corner
+  #top right-triangle
+  draw_line(
+    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
+    y = c( .25, .25,  .15, .25),
+    color = "blue", size=.75
+  )
+
+#draw boxes around the
 
@@
 
+```{r Fig6.IFD.square, echo=FALSE}
+#new figure, triangle squares
+
+```
+
 An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (Schumer).
@@
 Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
+
+new figure,
 

### 2020-05-20 doc/MS_v2.Rmd
first round of incorporating edits for MS_v2

@@
 ---
-title: 'Evolution of Sex-Specific Recombination Rate'
+title: 'Sex-Specific Evolution of the Meiotic Recombination Rate'
 date: '`r Sys.Date()`'
@@
 
-<v2 - Bret's edits, v1>
-
 ```{r set.up, echo=FALSE, include=FALSE}
@@
 
-#Intro
-
-#Material Methods Draft v1
-
-### Mouse Husbandry
-
-To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
-The _Mus musculus_ strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp) and rederived by the UW-Biotech center (Madison UW).
-The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
-We use abbreviated  superscript notation for the remainder of the paper (e.g.
-_musculus^WSB^_).
-All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
-A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (G) is maintained at UW Veterinary school facilities.
+# INTRODUCTION
+
+# MATERIALS AND METHODS
+
+### MICE
+
+We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murids to profile natural genetic variation in recombination (Table 1).
+Our survey included 5 strains from **M.
+m.
+musculus**, 4 strains from _M.
+m.
+domesticus_, 2 strains from _M.
+m.
+molossinus_, 2 strains from **M.
+m.
+castaneus**, and 1 strain each from **M.
+caroli**, **M.
+spicilegus**, and **M.
+spretus**.
+We subsequently denote strains by their abbreviated subspecies and name (e.g.
+**domesticus^WSB^**).
+Mice were housed at dedicated, temperature controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
+Mice from an inbred strain of Gough Island mice were sampled after XX generations of brother-sister mating.
+All mice were provided with ad libitum food and water.
+All procedures followed protocols approved by IACUC.
 
@@
 
-The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
+*new* - The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvenile mice aged between 12 to 15 days were used for DMC1 quantification.
-Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embryonic days and 0 to 48 hours old respectively.
-Whole testes and decapsulated ovaries were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
+The majority of mice used MLH1 quantification were between 5 and 12 weeks of age.
+Juvenile mice between 12 and 15 days of age were used for DMC1 quantification.
+Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
+Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
+Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
 Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
@@
 The following morning, slides were treated with a Photoflow wash (Kodak, cite).
-Slides were stored at -20*C if not stained immediately.
-
+Slides were stored at -20 \* C if not stained immediately.
 To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
-Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
-The staining protocol was based on that in [@anderson1999] and [@koehler2002].
-Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
-
-Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
-The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
+Crossovers (COs) were visualized as MLH1 foci.
+Double strand breaks (DSBs) were visualized as DMC1 foci.
+The staining protocol followed [@anderson1999] and [@koehler2002].
+Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
+Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\* C.
+The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\* C.
 Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
 Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
@@
 
-Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
+new - Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
@@
 
-To estimate the numbers of percursor sites which may lead to crossovers, the mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
-SC morphology and CREST/centromere foci number was used to stage spermatocytes, in early and late zygotema stages.
-
-**MLH1** foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
-With scores of 1 and 2 having the highest quality.
-Cells and coded as a score of 5 were excluded from the final analysis.
-The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
-MLH1 foci located on the XY in spermatocytes were excluded from this study.
-
-**M1 Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
-
-**M4 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
-
-**M5 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
-
-A series of mixed models and general linear models were used as an analetical framework.
-We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
-Strain was coded as a random effect and tested using exactRLRT() to reflect a random sample of natural genetic variation from each subspecies.
-The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
-
-This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
-The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
-The general linear models, **M4** were used to analyze sex specific data sets.
-
-Our goal was to design an analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
-sex, strain subspecies) and can be applied to multiple (meiotic traits) dependent variables.
-In addition to the MLH1 count variables for individual mice, mean MLH1 count per cell, variance and coefficient of variance for MLH1 count per cell, we analyzed SC based features: mouse average of total SC per cell, mouse average of single SC length, mouse mean IFD^norm^ and IFD^raw^ for double crossover chromosomes, and the average position of crossover for single crossover bivalents.
-
-These traits were choosen because they are markers of phenomena in the recombinaiton pathway and provide more detailed information about the recombination landscape.
-In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
-Normalized measures were calculated by the position divided by the length of the SC for that single bivalent.
-
-### Single bivalent characterization
-
-To characterize the SC length of individual chromosomes, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
-Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
-Hand measures of pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
-
-The **total SC** area per cell was quantified from the pachtyene cell images.
-To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-
-The **DNA CrossOver** algorithm isolates single straightened bivalents shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
-
-A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
-From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` bivalent objects were isolated by the image analysis software.
-After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a **reduced data** set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of `r sum(Long.biv_mouse_table$n_obs)`   'short' bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+*new-* To estimate the number of crossovers across the genome, we counted MLH1 foci.
+MLH1 foci were counted in cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score ranging from 1 (best) to 5 (worst) was assigned to each cell based on visual appearance of staining and spread of bivalents.
+Cells with a score of 5 were excluded from the final analysis.
+Distributions of MLH1 count per cell were visually inspected for normality (supplemental figure ).
+MLH1 foci located on the XY in spermatocytes were excluded from counts.
+In addition to MLH1 counts, we measured several traits to further characterize the recombination landscape.
+To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from a subset of strains ( _musculus^PWD^_, _musculus^MSM^_, _domesticus^WSB^_, and _domesticus^G^_ ).
+SC morphology and CREST foci number was used to stage spermatocytes as early zygotene or late zygotene.
+
+To measure bivalent SC length, two image analysis algorithms were used.
+The first algorithm estimates the total (summed) SC length across bivalents for individual cells [@wang2019_sc].
+The second algorithm estimates the SC length of individual bivalents [@peterson2020].
+Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
+
+Total SC length per cell was quantified from pachtyene cell images [@wang2019_sc].
+To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
+Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
+
+SC length of individual bivalents was quantified in pachytene cell images [@peterson2020].
+The DNA CrossOver algorithm [@peterson2020] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
+The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
+In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female).
+From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
+Following a manual curation, `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2020].
+The curated single bivalent data supplements our cell wide MLH1 count data with MLH1 foci counts for single bivalents.
+The proportions of bivalents with the same number of MLH1 foci where compared across strains by a chi-squared test using prop.test().
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+This “short bivalent” data set included the four or five shortest bivalents, and excluded the -- X bivalent in oocytes.
+A total of `r sum(Long.biv_mouse_table$n_obs)` -- 678 short bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` -- 103 oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` -- 37 spermatocytes.
 Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
-
-# Results
-
-```{r FOR.PLOT.organizing.data, echo=FALSE, warning=FALSE}
+A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
+To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD^raw^) was measured for those single bivalents containing two MLH1 foci.
+A normalized measure of interference (IFD^norm^) was computed by dividing IFDraw by SC length on a per-bivalent basis.
+
+We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
+We first constructed a mixed model (M1) using the lmer() from the lmer4 package (cite) in R (v3.5.2)(Team 2015).
+In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT()).
+Subspecies, strain, and their interaction were coded as fixed effects,  with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()).
+We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
+In separate analyses, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, variance in MLH1 count per cell, coefficient of variation in MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD^raw^, IFD^norm^, and average MLH1 position (for single-focus bivalents).
+Four additional models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from the initial mixed model.
+
+# RESULTS
+
+```{r FOR.PLOT.organizing.data, echo=FALSE, message=FALSE, warning=FALSE}
 #set the order of another column, based on another variable. consistant order for ploting
@@
 
-```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
+```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE}
 
@@
 
-```{r table.for.plot, echo=FALSE, warning=FALSE}
+```{r table.for.plot, echo=FALSE, warning=FALSE, message=FALSE}
 #why do I add in a different table?
@@
 
-updates for Figure 1, try shifting the lines up for B C
-
-## Genome-wide recombination rate estimates for both sexes
+## B Genome-wide recombination rate evolves differently in females and males
 
@@
 
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M.
+We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` -- 14 wild-derived inbred strains sampled from three subspecies of house mice ( _M.
 m.
-domesticus*, *M.
+domesticus_, _M.
 m.
-musculus* and *M.
+musculus_ and _M.
 m.
-molossinus*.
-Mean MLH1 foci counts for `r length( unique(Mouse.Table_HQ)$mouse )` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
-
-Graphical comparisons between the two sexes reveal several patterns (Figure 1A).
-First, recombination rate differs between females and males in most strains.
-Second, the direction and magnitude of heterochiasmy varies among strains.
-Although the majority of strains show higher recombination rates in females (following the pattern in laboratory mice), two musculus strains and one molossinus strain exhibit male-biased heterochiasmy.
-Hence, relative recombination rates in the two sexes are evolutionarily labile.
-
-The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
-Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
-
-Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C).
-Female recombination rates are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure 1B).
-In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).
+molossinus_ ) and two additional species of Mus ( _M.
+spretus_ and _M.
+spicilegus_ ).
+Mean MLH1 foci counts for -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 1,867 spermatocytes) and -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
+Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
+First, MLH1 focus counts differ between females and males in most strains.
+Second, the relationship between MLH1 foci in the sexes varies among strains.
+Although most strains show more MLH1 foci in females, three strains - _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ - exhibit higher MLH1 focus counts in males.
+In females, numbers of MLH1 foci are evenly distributed around the sex-wide mean of approximately 25 (Figure 1B).
+In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23 MLH1 foci per cell) of foci (Figure 1C).
+
+To further situate variation in recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
+We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < 10^{-6} -- p < `r 1/1000000`), sex (p = 1.5510^{-8} -- `r MLH1.MMlmer_results$sex.results`), subspecies (p=1.7210^{-4} -- `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = 3.110^{-5} -- `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
+After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
+A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = 1.0410^{-6} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
+In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
+
+We next fit models separately for  `r length(fomo.male_mouse.av$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4).
+In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = 0--`r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).
+
+These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
+Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p= 0.02 -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
+Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
+Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.
 
@@
 
-## Partitioning variation in recombination rate
-
-To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
-We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r 1/1000000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
-
-After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and _domesticus^G^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
-
-We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4; see methods).
-In the male dataset, three strains significantly affect recombination rate: _musculus^PWD^_ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect =
-`r sum_MLH1.male.glm.M4$coefficients[8]` foci), and 
-_musculus^SKIVE^_ (glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
-
-These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and _musculus^PWD^_ (p= `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
-Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
-
-## Within mouse variance in CO count per cell
+## B Synaptonemal complexes are longer in females
 
@@
 
-## DMC1 Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
+## Sex Differences in Meiotic Traits
+
+```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
+#total-SC
+#Q1.lmer_totSC.mouse.av.table -- #can't find this table
+#Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>%
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
+
+#subsp effect, many strain, #male
+#G * male v.sig and male*LEW
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
+#LOT OF SIGNIFICant strain effects in this model
+```
+
+```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
+
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+
+```
+
+```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
+
+#summary(Q1.glm_short.SC_M3)
+
+summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
+
+#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#sex male very sig -- only sig
+```
+
+```{r Q1.SC.Q12, echo=FALSE}
+
+#make comparisons for the high quality cells
+#get Q12 cells from MLH1.merge.Skel
+#calq the mouse averages
+
+MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
+
+totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
+
+#unorder the factors
+totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
+totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
+totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
+totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
+
+#REMOVE JUVI mice
+tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
+totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
+                 .(strain, sex),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
+
+#unorder the factors
+totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
+totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
+totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
+totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
+
+```
+
+```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
+
+Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Dom
+#Tot.SC_mouse.av_Dom
+
+Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Musc
+#Tot.SC_mouse.av_Musc
+
+Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+```
+
+```{r show.long.biv, echo=FALSE, include=FALSE}
+
+#Long.biv_mouse_table
+
+#mouse av
+Q1.SC.long.mouse_scatter_boxplot <- ggplot( Long.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
+
+```
+
+```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
+
+#Q1.real.long_mouse_table
+
+Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#not enough y observations
+#SKIVE
+#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
+#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+```
+
+```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
+
+#removing these plots for space
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+## B Females and males differ in crossover positions and crossover interference
+
+```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
+#move to setup
+
+Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+```
+
+```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
+#main figure of density plots,
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+```
+
+```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
+
+Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \n all bivalents")
+
+```
+
+```{r Q1_1COpos.t.test, echo=FALSE}
+#testing,
+#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
+
+Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
+
+#WSB
+Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
+#G
+Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
+#LEW
+Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
+
+#PWD
+Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
+
+#KAZ
+Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
+
+#SKIVE
+#Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain == "SKIVE") )$norm.first.foci.pos,
+ #          (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$norm.first.foci.pos )
+
+#MSM
+Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
+                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
+
+#Single Biv Level
+#single.biv.1CO_pos_t.test <- t.test( 
+ # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
+#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+#  )
+
+#t.tests for mous.av_nrm.1CO.pos 
+```
+
+```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
+
+#make sure this dataset has all the right strains
+Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
+lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
+
+lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
+lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
+
+lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
+sex.results = lmer.Q1.nrm1CO_M1.sex , 
+               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
+
+#lmer.Q1.nrm1CO_M1_results
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+```{r Q1.CO1pos_glm.M2, echo=FALSE}
+
+#in case factors aren't ordered
+#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
+
+#summary(Q1.glm_F1pos_M2)
+
+summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
+#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
+
+Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
+
+#summary(Q1.glm_1CO.pos_M3)
+
+summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
+#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
+#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+
+```
+
+```{r double.IFD.triangle.plots, echo=FALSE}
+
+#good function, saves to file, can't reproduce in rmd ..width=650, height=500
+#big loop for automating the IFD triangles
+only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
+auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
+
+df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
+df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
+#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
+
+#this works
+top.triangle <- function(my.y, my.x, my.data) {
+  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
+       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
+       col =  colors_of_strains[current.strain])
+
+  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
+  mtext("Position.2.Female", 2, .75)
+  mtext("Position.1.Female", 3, 2)
+
+  axis(side=2, las=1, pos=0)
+  axis(side=3, las=1, pos=1)
+  lines(0:1, 0:1)
+}
+
+bottom.triangle <- function(my.y, my.x, my.data){
+  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
+          col =  colors_of_strains[current.strain])
+
+  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
+  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
+  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
+  axis(side=1, las=1, pos=0, at=x.at, 
+       labels=F, xpd=TRUE)
+  mtext(seq(0, 1, .2), 1, 0, at=x.at)
+  axis(4, las=1, pos=1 + x.dist)
+  lines(0:1 + x.dist, 0:1, xpd=TRUE)
+}
+
+x.dist <- .1
+RP <- list()
+
+for(i in 1:(length(df_strain_list))){
+  #print(i)
+  current.strain <- as.character(df_strain_list[[i]]$strain[1])
+
+  #declare file for saving png
+  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
+
+  plot.new()
+
+  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
+
+  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
+
+  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
+
+  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  #adjusting this mar lowers the margins
+  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
+
+  top.triangle(my.y.top, my.x.top, my.data)
+
+  bottom.triangle(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
+
+  par(op)
+  RP[[i]] <- recordPlot()
+  dev.off()
+}
+```
+
+```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Distribution of normalized foci positions (IFDs) for double crossover bivalents across sexes.
+Female and male patterns are displayed in the top and bottom triangle respectively.
+Each point reflects a single double crossover bivalent, with the normalized positions of foci reflected on each axis."}
+
+plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
+plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
+plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
+plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
+plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
+plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
+plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
+
+# grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
+#             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
+#             rasterGrob(plot7.MSM), 
+#             ncol=3)
+
+#load all the files / arrange the plots from above 
+#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
+```
+
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized IFD")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
+
+#each point is a 2CO, with the F1 and F2 position 
+#based on plot from Dourad et al (Mercier)
+
+P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
+    aes(y = Foci2, x= Foci1, color=strain))+
+  geom_jitter(alpha=.5)+
+  ggtitle("Raw IFD WSB")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#trying to move x axis to top
+#cowplot might be the easier way -- since kinda diffecult in ggplot
+#or ggdraw
+
+P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+
+#+theme(legend.position="none")
+```
+
+```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
+#below are calq just from 2COs, have no fear
+#mean_IFD.2CO_ABS
+#mean_IFD.2CO_PER
+
+#FULL all strains
+Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
+#p-value = 0.07
+
+Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
+#p-value = 1e-12
+
+#WSB
+Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
+                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
+
+#LEW
+Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
+
+#G
+Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
+
+#PWD
+Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
+
+#KAZ
+Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
+
+#SKIVE
+#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+
+#MSM
+Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
+
+#FULL all strains
+Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
+#p-value = 0.3
+```
+
+```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
+
+#these are the scatter plots
+
+nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
+#looked up image, measures confirmed
+
+####
+nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#21may18_KAZ_m2 -- seems like it might be an outlier
+#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
+#second foci slightly smaller, kinda fainter -- but also much brighter than background
+
+#check all KAZ males
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ.female_BIV.data
+
+nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD.female_BIV.data
+
+nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_LEW.female_BIV.data
+
+nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G.female_BIV.data
+
+nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
+
+raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+###look at why KAZ males are weird
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#the difference the the Nrm IFD cut off is striking with these plots
+nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
+
+```
+
+```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
+
+#M1_IFD_2CO_ABS
+lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
+               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
+
+Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
+               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
+
+```
+
+```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
+#mouse.avs_4MM -- is also the bivbased mouse average table 
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
+
+glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
+summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
+
+#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
+```
+
+```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
+
+glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
+
+#PER
+glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
+#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+```
+
+```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
+#real.long.bivData
+Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+
+Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+
+```
+
+```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
+#REAL.short_mouse_table
+#REAL.long_mouse_table
+#mouse.av.long.biv.table
+
+Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+
+Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+```
+
+*new* -- We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68 -- `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.56; t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value` ).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = 1.2610^{-25} -- `r lmer.Q1.nrm1CO_M1_results$sex.results` ; M2: p = 1.3310^{-7} -- `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 1.3310^{-7} -- `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+
+Males have longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < 1.4910^{-12} --), with only _musculus^KAZ^_ showing no difference (t-test; p = 0.33 -- `r Q1.IFD.raw_FULL_t.test$p.value` ).
+Examination of IFD^norm^ distributions indicates that female values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Analysis of models treating IFD^norm^ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = 6.7410^{-14} --`r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains, whether they are considered separately or together.
+Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males and that males display a stronger telomeric bias in the placement of the distal crossover.
+In summary, controlling for differences in SC length (chromatin compaction) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical scale.
+
+## B Evolution of genome-wide recombination rate is dispersed across bivalents, associated with double-strand break number, and connected to crossover interference
 
@@
 
-In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains ( _musculus^KAZ^_ , _domesticus^WSB^_ , and _domesticus^G^_ ) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_ ).
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p < `r 1/1000000`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = `r ttest.HighLow.Z.pval`).
-
-<ttest.HighLow.L.pval>
-
-After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
-This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
-@cole2012 suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
-This result, combined with the large difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
-
-<!-- The number of DSBs increase in zygotema, then deminish as the effects of crossover homeostasis are implemented (Cole).
-(co homeostasis also indecates dsb ) Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links connects the CO number variation to 'pre' DSB formation fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
-(this study only examined one stage of early prophase for DMC1 counts -->
-
-<!-- the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 C3H, 7.3 B6 -->
-
-## Evolution of genome-wide recombination rate is reflected at the single chromosome level
-
 ```{r algorithm_performance_stats, echo=FALSE}
@@
 ```
-
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure 3).
-The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
-low recombining strains (Figure 3).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
-Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-_musculus^KAZ^_ p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_ p = `r MSM.MOLF.2CO.prop.test$p.value`).
-Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
-We address two main questions.
-First, which traits are sexually dimorphic?
-Second, which traits differ between males from high-recombination vs.
-low-recombination strains?
-
-## Sex Differences in Meiotic Traits
 
@@
 
-In many mammalian species, the synaptonemal complex (SC) is longer in females [@cahoonLibuda2019, @sardell_sex_2020].
-Yet, the majority of these observations come from species with female-biased heterochiasmy.
-Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
-
-We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
-The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = r Q1_mus.av_shortSC_t.test.SKIVE$p.value).
-The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
-
-Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
-That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
-
-In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers compared to females.
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-#subsp effect, many strain, #male
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
-Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
-Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-```{r Q1.SC.Q12, echo=FALSE}
-
-#make comparisons for the high quality cells
-#get Q12 cells from MLH1.merge.Skel
-#calq the mouse averages
-
-MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
-
-#unorder the factors
-totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
-totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
-totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
-totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
-
-#REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(strain, sex),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
-
-#unorder the factors
-totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
-totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
-totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
-totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
-
-```
-
-```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
-
-Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Dom
-#Tot.SC_mouse.av_Dom
-
-Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Musc
-#Tot.SC_mouse.av_Musc
-
-Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-```
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-
-#Long.biv_mouse_table
-
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Long.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#not enough y observations
-#SKIVE
-#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
-#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-### Positions of Single Crossovers
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
-#move to setup
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
-#main figure of density plots,
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-#Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain == "SKIVE") )$norm.first.foci.pos,
- #          (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$norm.first.foci.pos )
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test( 
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-#lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-
-```
-
-We used the single crossover bivalent (1CO) position to compared general location of crossovers while controlling for differences in total SC length.
-Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.
-
-In all strains, the landscape across 1CO bivalents is significantly different in females and males.
-Normalized foci positions tend to be more central in females `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` and closer to the telomere in males `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )` (t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
-
-Sex is also the most significant effect on focus position in a mixed model (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species [@sardell_sex_2020].
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-#real.long.bivData
-Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-
-Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-#mouse.av.long.biv.table
-
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-```
-
-## Sex Differences in CO Interference
-
-# IFD double triangle
-
-# Figure 1
-
-```{r double.IFD.triangle.plots, echo=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
-auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
-
-df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
-df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-
-#this works
-top.triangle <- function(my.y, my.x, my.data) {
-  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
-       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
-       col =  colors_of_strains[current.strain])
-
-  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
-  mtext("Position.2.Female", 2, .75)
-  mtext("Position.1.Female", 3, 2)
-
-  axis(side=2, las=1, pos=0)
-  axis(side=3, las=1, pos=1)
-  lines(0:1, 0:1)
-}
-
-bottom.triangle <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[current.strain])
-
-  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
-  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-  mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1 + x.dist)
-  lines(0:1 + x.dist, 0:1, xpd=TRUE)
-}
-
-x.dist <- .1
-RP <- list()
-
-for(i in 1:(length(df_strain_list))){
-  #print(i)
-  current.strain <- as.character(df_strain_list[[i]]$strain[1])
-
-  #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
-
-  plot.new()
-
-  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-
-  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins
-  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
-
-  top.triangle2(my.y.top, my.x.top, my.data)
-
-  bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-```
-
-```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Distribution of normalized foci positions (IFDs) for double crossover bivalents across sexes.
-Female and male patterns are displayed in the top and bottom triangle respectively.
-Each point reflects a single double crossover bivalent, with the normalized positions of foci reflected on each axis."}
-
-plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
-plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
-plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
-plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
-plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
-plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
-plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
-
-# grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
-#             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
-#             rasterGrob(plot7.MSM), 
-#             ncol=3)
-
-#load all the files / arrange the plots from above 
-#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
-```
-
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
-
-#each point is a 2CO, with the F1 and F2 position 
-#based on plot from Dourad et al (Mercier)
-
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
-    aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.5)+
-  ggtitle("Raw IFD WSB")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#trying to move x axis to top
-#cowplot might be the easier way -- since kinda diffecult in ggplot
-#or ggdraw
-
-P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
-#SKIVE
-#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-```
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
-
-#these are the scatter plots
-
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
-
-```
-
-```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-
-#M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-
-```
-
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
-summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-```
-
-```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-```
-
-There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain, *domesticus^G^* (t-test without domesticus^G^; p = `r Q1.IFD.raw_t.test_G$p.value`).
-This result indicates that females and males exhibit a similar level of interference when it is measured in physical SC units.
-
-In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < `r Q1.IFD.PER_FULL_t.test$p.value`), with only _musculus^KAZ^_ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-
-Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` glm *M2*, p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`,  *M3* p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
-When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in SC length (chromatin compaction) using IFD^norm^ indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical scale.
-
-Visual comparison of interference pattern is aided by Figure 4C, which combines the total length of distance between two foci, but also their normalized locations along the chromosome.
-Two main patterns of sexual dimorphism emerge across all strains, (Supplemental figure X).
-First, the female patterns (top triangle), are more uniformly distributed compared to males.
-Suggesting a wider range of normalized IFD lengths compared to males (bottom triangle).
-Male IFDs display a stronger chromosome end-locatization of the second foci (Figure 4).
-
-```{r Comp_Ratios, echo=FALSE, eval=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
-
-Next, we used the contrast between males in the high-recombination strains ( _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ ) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M5).
-
-Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-SC length is expected to be positively associated with genome-wide recombination rate because of the connection between DSB formation and the loop-axis structure [@zickler1999, @mercier2015].
-Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome with a finite length.
-Thus, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
-
-Following this logic we predict (1) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_, and 3) _domesticus_ strains will have similar SC length and crossover interference.
-
-## Conservation of Mean SC length
-
 ```{r Q2.SC.short.long.setup, echo=FALSE}
@@
 
-Confirming the basic predictions, there is a positive correlation between mouse mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
-Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure 5).
-While high-recombination strains have significantly greater mouse mean total SC area in the total dataset (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p =  `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, the mouse means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
-
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (_musculus_ p =  `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, _molossinus_, p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
-In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-**Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?**
+*new* -- We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+We considered proportions of bivalents with different numbers of crossovers, double-strand breaks, SC length, and crossover positions, including crossover interference.
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")` -- 34,982) have either one or two MLH foci (Figure 3).
+The proportions of single-focus (1CO) bivalents vs.
+double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Figure 3).
+
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in _molossinus^MSM^_ (Figure 3).
+Following patterns in the genome-wide recombination rate, male musculusPWD and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
+_musculus^KAZ^_: p = 3.1510^{-33} -- `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_: p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
+These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
+To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
+DMC1 foci were counted in a total of 76 -- `r sum( (DMC1table.main99$ncells.x) )`  early zygotene and 76 -- `r sum( (DMC1table.main99$ncells.y) )` late zygotene spermatocytes from three low-recombination strains ( _musculus^KAZ^_, _domesticus^WSB^_ , and  _domesticus^G^_) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_).
+
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < 10^{-6} --  p < `r 1/1000000` ).
+In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = 0.66 -- `r ttest.HighLow.Z.pval` ).
+Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
+High-recombination and low-recombination strains do not differ in the MLH1/DMC1 ratio, whether DMC1 foci were counted in early zygotene cells or late zygotene cells (t-test; p > 0.05).
+These results raise the possibility that the evolution of genome-wide recombination rate is primarily determined by processes that precede the CO/NCO decision.
+
+Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
+
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within M.
+m.
+molossinus (p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`) but not within M.
+m.
+musculus (p = 0.87 -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- ; long: p = 0.18 -- ).
+In a model with total SC length as the dependent variable, two subspecies effects are significant ( _M.
+m.
+musculus_ p = 1.2410^{-6} -- , _M.
+m.
+molossinus_, p = 10^{-6} -- ).
+In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05), but they are not consistent across models.
+Collectively, these observations reveal that evolution of the genome-wide recombination rate is not strongly associated with evolution of SC length.
+High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = 0.24 -- ) or within subspecies (t-test; _M.
+m.
+musculus_: p = 0.41 -- ; _M.
+m.
+molossinus_ : p = 0.07 -- ).
+While domesticusWSB and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable (), these two strains do not differ in MLH1 count.
+These results show that evolution of crossover position on chromosomes with single crossovers is decoupled from evolution of the genome-wide recombination rate, at least at this scale of resolution.
+
+High-recombination strains have greater inter-focal distances than low-recombination strains (t-test; IFD^norm^: p = 7.7410^{-7} -- ; IFD^raw^: p = 8.7810^{-6} -- ).
+This pattern holds within M.
+m.
+musculus (t-test; IFD^norm^: p = 2.0410^{-5} -- ; IFD^raw^: p = 1.9410^{-4}), with a trend in the same direction within M.
+m.
+molossinus (IFD^norm^: p= 0.17 -- ; IFD^raw^: p = 0.08 -- ).
+Similar results are observed with models treating IFDraw and IFDnorm as dependent variables: only effects associated with high-recombination strains are significant (p < 0.05).
+That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths (and bivalent sizes) as the primary explanation.
+The main distinction in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains (Figure 6).
+The frequency of IFDnorm values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and musculusPWD, respectively).
+These results indicate that evolution of the genome-wide recombination rate is accompanied by the evolution of crossover interference.
+In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not (consistently) to SC length and crossover position (on single-crossover bivalents).
+
+*old* -- In many mammalian species, the synaptonemal complex (SC) is longer in females [@cahoonLibuda2019, @sardell_sex_2020].
+Yet, the majority of these observations come from species with female-biased heterochiasmy.
+Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
+In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
+
+We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
+The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = r Q1_mus.av_shortSC_t.test.SKIVE$p.value).
+The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
+Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
+That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
+Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
+In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
+Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers compared to females.
+
+# IFD double triangle
+
+# Figure 1
+
+```{r Comp_Ratios, echo=FALSE, eval=FALSE}
+
+#calq ratios of belo 30
+
+top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
+
+#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
+MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
+
+DF.strain.list <- list(
+  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
+  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
+  MSM.Bivalents.DF, MOLF.Bivalents.DF
+  )
+
+entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
+        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
+          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
+
+i = 1
+for(x in (DF.strain.list) ){
+
+  #print(head())
+
+  #these filters won't apply to the indexed DF
+  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
+
+  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
+  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
+  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
+
+  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
+
+  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
+
+  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
+  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
+  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
+
+  i = i+1
+
+  }
+#entr_DF
+
+entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
+
+```
+
+```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#real.short.bivData
+
+Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.long <- add_sex(Q1.IFD.long)
+Q1.IFD.long <- add_strain(Q1.IFD.long)
+Q1.IFD.long <- add_subsp(Q1.IFD.long)
+
+Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
+Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
+
+Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.short <- add_sex(Q1.IFD.short)
+Q1.IFD.short <- add_strain(Q1.IFD.short)
+Q1.IFD.short <- add_subsp(Q1.IFD.short)
+
+Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
+Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
+
+#t.test. mouse av
+Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+
+Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+#both the short and long norm IFD are NS
+
+#t.test Biv
+Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+
+Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+```
+
+## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
+
+Next, we used the contrast between males in the high-recombination strains ( _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ ) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
+In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M5).
+
+Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
+SC length is expected to be positively associated with genome-wide recombination rate because of the connection between DSB formation and the loop-axis structure [@zickler1999, @mercier2015].
+Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome with a finite length.
+Thus, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
+
+Following this logic we predict (1) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_, and 3) _domesticus_ strains will have similar SC length and crossover interference.
+
+## Conservation of Mean SC length
 
@@
 
-Despite quantifying a similar number of wild derived inbred strains from *M.
-m.
-domesticus* and *M.
-m.
-musculus*, we only observed rapid male specific evolution for the genome wide recombination rate for *M.
-m.
-musculus.* Estimates of effective population size for the three main subspecies of the house mouse found *M.
-m.
-musculus* had the lowest [@geraldes2011].
-Selection becomes less efficient at removing deleterious mutations in small populations outside of mutation-drift balance.
-In such populations, modifiers for increased recombination rates will be selected for as they increase the efficiency of selection.
-
-An additional possibility is that the different rates in recombination rate variation across subspecies is related to history of hybridization across the populations sampled in this panel of inbred strains.
-Theoretical work as connected recombination that rate of purging deleterious introgressed DNA in populations after hybridization (Caballero et al 2019, Schumer 2018).
-The three high-recombining strains identified in this study have connections to hybridization.
-Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the European house mouse hybrid zone.
-Notably *musculus^SKIVE^* comes from a younger section of the hybrid zone with a nuclear genome from *M.
-m.
-musculus* and a mitochondria genome from *M.
-m.
-domesticus* [@gyllensten1987_SKIVE].
-The subspecies *M.
-m.
-molossinus* is a natural hybrid of *M.
-m.
-musculus* and *M.
-m.
-cataneus* (Geraldes et al 2008, [@geraldes2011]).
-Models for the aggregate recombination metrics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019), motivating increased study of recombination rates and natural populations.
-
-<!-- transition to looking at the number and placement of crossovers -->
-
-To place our results in the broader context of recombination rate evolution we compare our main results to expectations under three modifier models and two functional and cell physiology models (table X).
-While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across existing models.
-The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome oscillatory movement (COM) model was built to describe positive interference.
-This spindle based selection model was developed to explain the results resented in this manuscript.
-In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
-
-## Evolution of reversed heterochiasmy direction
-
-In our results and more generally in the literature, female biased heterochiasmy is the most common pattern, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
-Our results extend previous reports of male biased heterochiasmy [@dumont2011] to  new subspecies of house mouse (*molossinus*) and emphasize the short evolutionary period.
-Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within house mouse subspecies requires further study.
-Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species in female biased heterochiasmy species, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
-This insight could be important for animal breeding programs [@battagin2016].
-
-Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-The SAC prevents aneuploidy by ensuring all bivalents are correctly attached to the microtubule spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
-
-A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion [@laneKauppi2019].
-Genetic variants will be selected in a manner to optimize the process of bi-orientation and prevent SAC from delaying cell cycle or triggering apoptosis.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014], which are connected to conserved features of gametogenesis; centrosome spindle [@So2019, @dumontDesai2012] and cell volume [@kyogoku2017].
-The more stringent SAC in spermatogenesis will be more effective at removing genetic variants which interfere with bi-orientation compared to females.
-This dynamic can give rise to sex-specific genetic variants which in turn can result in  context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
-
-## Conservation in sexual dimorphism in between cell variation in crossover number
-
-In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
-This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals.
-Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-Centrosomes act as the two MTOCs
-for the spindle during spermatogenesis and mitosis.
-The delivery of centrioles (most mammals) or centriole material (rodents and rabbits) to the egg is thought to be a barrier to evolution of parthenogenesis [@manandhar2005centrosome].This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
-
-Regardless of whether these hypotheses are true, our observations of levels of between cell variation fail to support a basic prediction of that a higher mean will be associated with a higher variance.
-Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
-These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002], [@gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes.
-This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
-
-<!-- stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review -->
-
-## Conservation in sex-specific recombination landscape
-
-Our results extend the plethora or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] are data comprised mostly of species with large evolutionary distances.
-In contrast our results demonstrate that this pattern is maintained across much shorter evolutionary distances and even with rapid evolution in the genome-wide recombination rate.
-We argue that this conservation of the male chromosome end bias is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
-
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a conserved trait.
-The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
-The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM].
-
-Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (@laneKauppi2019, @subramanian2014, @dumontDesai2012).
+By comparing recombination rates in females and males from the same diverse set of genetic backgrounds, we isolated sex as a primary factor in the evolution of this fundamental meiotic trait.
+Recombination rate differences are more pronounced in males than females.
+Because inter-strain divergence times are identical for the two sexes, this observation demonstrates that the genome-wide recombination rate evolves faster in males, at least in house mice.
+More generally, recombination rate divergence is decoupled in males and females (Spearman’s correlation test comparing female and male strain mean MLH1 values for house mice; rho = 0.18; p = 0.60).
+These disparities are remarkable given that recombination rates for the two sexes were measured in the same genomic backgrounds.
+Our results provide the strongest evidence yet that the genome-wide recombination rate follows distinct evolutionary trajectories in males and females.
+Next, we consider several explanations for this phenomenon.
+At the genetic level, the sex-specific evolution we documented indicates that some mutations responsible for divergence in recombination rate have dissimilar phenotypic effects in females and males.
+A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@kong2004, @kong2008, @kong2014, @halldorsson2019]), Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle, @shen2018_cattle] appear to show sex-specific properties, including antagonistic effects in females and males.
+Inter-sexual correlations are weak for recombination rate in humans [@fledel2011] and Soay sheep [@johnston2016_soay].
+Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011, @wang2017island].
+These variants could differentially affect females and males at any step in the recombination pathway.
+Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contributed to sex-specific evolution in the recombination rate.
+Our results also raise the prospect that the connection between recombination rate and fitness differs between males and females.
+Little is known about whether and how natural selection shapes recombination in nature [@DapperPayseur2017, @ritz2017]).
+Using a quantitative genetic test, [@samuk2020]  recently discovered that genome-wide recombination rates in females from two populations of _Drosophila pseudoobscura_ diverged in a manner consistent with local adaptation.
+Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
+Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling, @sardell_sex_2020].
+Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020].
+Another modifier model showed that meiotic drive could stimulate female-specific evolution of the recombination rate [@brandvain2012scrambling].
+Although these models fit the conserved pattern of sex differences in crossover positions, they do not readily explain our observations of sex-specific evolution in the genome-wide recombination rate.
+In particular, the alternation across strains in which sex has more crossovers is unexpected.
+We propose an alternative interpretation based on the cell biology of gametogenesis.
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across a proportion of sister cohesion [@LaneKauppi2019].
+The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition [@LaneKauppi2019].
+Hence, selection seems likely to favor mutations that optimize the process of bi-orientation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019], perhaps due to the presence of the centrosome spindle [@So2019] and higher cell volume [@kyogoku2017] in oocytes.
+The higher stringency of the SAC during spermatogenesis suggests that selection will be better at removing mutations that interfere with bi-orientation in males than in females.
+Therefore, faster male evolution of the genome-wide recombination rate could be driven by the stronger SAC acting on chromosome structures at the metaphase I alignment.
+Our SAC model is consistent with other features of our data.
+We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
+We hypothesize that these sex differences hinge on disparate requirements for chromosome cohesion in late meiosis I.
+The irreversible process of the metaphase-to-anaphase transition is initiated by the proteolytic decay of the sister cohesion proteins that connect homologs [@LaneKauppi2019, @subramanian2014, @dumontDesai2012).
 The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
-Differences in timing of cell cycle between oogenesis and spermatogenesis imposes the the different selective pressures on how sister cohesion affects chromosome structure.
+Differences in timing of the cell cycle between oogenesis and spermatogenesis could impose contrasting selective pressures on how sister cohesion affects chromosome structure.
 Faster spermatogenesis may select for synchronization of the separation homologs (cite).
-While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dicyate arrest [@lee2019].
-
-The increased genetic diversity and recombination rates at chromosome ends @haenel2018 suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020], which provides some interesting evolutionary predictions.
-
-## Conservation of sex differences in chromatin structure
-
-Our results from _musculus^MSM^_ and _musculus^PWD^_ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
-These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
-Chromatin compaction could explain variation in heterochiasmy, which is driven by the distinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
-Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
-females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation (see below).
-
-Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
-This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier, Sera) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
-
-As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
-From empirical data in mice [@petkov2007], [@lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
-The signal for interference is thought to be mediated though the SC, which is built on top of the chromosome axis [@zicklerKleckner2015] regardless of the loop size.
-Supporting this is the observation that interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-As a consequence, since more base pairs are sequestered into the radial loops per unit of the linear central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males,  will have a signal of stronger interference when measured in base pairs compared to 'the typical female chromatin configuration.
-Stronger interference in males, has been noted across many sex-specific linkage maps (cite) strengthening the hypothesis that sexual dimorphism in the chromatin organization is widespread, however this has only been confirmed by cytological data in mice and humans.
-(Also large difficulties in observing sperm cells through meiosis).
-
-<!-- The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019] -->
-
-## Evolution of interference strength in males
-
-```{r Fig6.IFD.square, echo=FALSE}
-#new figure, triangle squares
-
-```
-
-An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (Schumer).
-Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
-
-Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
-The general pattern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
-This fits the logic of the chromosome axis acting as the 'real estate' for DSBs and aligns with known molecular mechanisms (cite).
-Two examples of the opposite direction, a positive correlation of interference strength and genome-wide recombination rate, involve descriptions of observations at the single bivalent level; the between lab-raised and wild mice of *Peromyscus leucopus* from [@peterson2019] or large effects on genome wide recombination rates across a short evolutionary differences; the within-sex comparison of two breeds of cattle with different genome wide recombination rates [@ma2015_cattle].
-
-We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphase.
-The evolution of genome wide recombination rates in our study is driven by a transition of the majority of
-bivalent having single crossovers to an enrichment of bivalents with two crossovers.
-All else being equal, interference strength would be expected to be equal or even weaker.
-Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain enriched for two crossover bivalents.
-Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
-
-The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase [@rubin2020mixing], however it lacks a mechanism for a evolutionary based predictions.
-While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
-
-We have focused on the model involving direct selection on the recombination pathway, which has an impact on the fitness of an individual, but we acknowledge connections to other models involving indirect selection on recombination rates which involve the fitness outcomes of offspring.
-A model from @goldstein1993, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
-The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
-A distinguishing feature of this model from other modifier  models is that the number of crossovers per chromosome is kept constant.
-Empirically, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
-Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
-
-new figure,
-
-## Future steps
-
-There are still many un-known aspects regarding heterochiasmy (@lenormand2016) --- that would benefit from the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
-A goal of such merging of models could be to connect empirical findings across scales.
-We make two suggestions for future steps in studying recombination variation.
-First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
-Second certain systems can be leveraged to overcome limitations inherent in some approached.
-For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
-FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian trans-locations.
-
-# References
+While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dictyate arrest [@Lee2019].
+We propose that the SAC model also can explain the correlated evolution of stronger crossover interference and higher genome-wide recombination rate in male house mice.
+Our results show that crossovers are spaced further apart in strains enriched for double-crossover bivalents when SC length is taken into account and chromosome size effects are minimized.
+Assuming chromatin compaction between (prophase) pachtyene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and improve the fidelity of chromosomal segregation.
+Although the SAC model postulates direct fitness effects of interference, a modifier model predicts that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
+Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
+The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle @ma2015_cattle and differences between wild-born and laboratory-raised white-footed mice @peterson2019.
+In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@segura2013, @ottoPaysuer2019].
+Collectively, these patterns suggest that inferences about the evolutionary dynamics of interference depend on the timescale under consideration.
+Our findings further reveal that evolution of the genome-wide recombination rate does not require changes in the degree of chromatin compaction.
+Females consistently showed longer SCs, even in strains with more recombination in males.
+Studies in mice [@lynn2002, @petkov2007] and humans [@gruhn2013] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes @hulten2011_COM.
+Our conclusions are accompanied by several caveats.
+First, MLH1 foci only identify interfering crossovers (REF).
+Although most crossovers (XX%; REF) belong to this class, our approach likely underestimated genome-wide recombination rates.
+Evolution of the number of non-interfering crossovers is a topic worth examining.
+A second limitation is that our investigation of crossover positions was confined to the relatively low resolution possible with immunofluorescent cytology.
+Positioning crossovers with higher resolution could reveal additional evolutionary patterns.
+Finally, the panel of inbred lines we surveyed may not be representative of recombination rate variation within and between subspecies of house mice.
+We considered most available wild-derived inbred lines, but house mice have a broad geographic distribution.
+Nevertheless, we expect our conclusion that recombination evolves in a sex-specific manner to be robust to geographic sampling because differences between females and males exist for the same set of inbred strains.
+The causes of sex differences in recombination remain mysterious @lenormand2016.
+Our conclusions have implications for a wide range of recombination research.
+For biologists uncovering the cellular and molecular determinants of recombination, our results suggest that mechanistic differences between the sexes could vary by genetic background.
+For researchers charting the evolutionary trajectory of recombination, our findings indicate that sex-specific comparisons are crucial.
+For theoreticians building evolutionary models of recombination, different fitness regimes and genetic architectures in females and males should be considered.
+Elevating sex as a primary determinant of recombination would be a promising step toward integrating knowledge of cellular mechanisms with evolutionary patterns to understand recombination rate variation in nature.
+
+# REFERENCES

### 2020-05-21 Chap3_MLH1var.rmd
notes and preliminary results for Chap3

@@
 
-load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
+#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
 
@@
 #skeletonize
-SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+
+Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
 
@@
 
+#Writing Outline
+
+I find these results encouraging. The higher variance in females is visible across a variety of genetic backgrounds, including those in which females have lower means than males.
+
+I feel like this result solidifies a plan for Chapter 3.
+
+It could take the following basic form:
+Lower stringency of checkpoints during oogenesis predicts higher variance in crossover number within females
+Despite acceptance of this prediction, it has rarely been tested
+We tested the prediction across a wide array of genetic backgrounds
+Females consistently have higher variance than males in the same inbred line, with little evidence for line effects
+[Perhaps something about variance in SC length]
+In terms of recombination mechanisms, our results mean ...
+In terms of evolutionary mechanisms, our results imply ...
+
 > How much do CO counts vary among cells from the same individual?
@@
 
-Below is the table for the number of cells per category by quality score. For these between cell variance will just use those in quality 1-2.
+Below is the table for the number of cells per category by quality score. For this document just the cells with quality 1-2 will be used.
 
@@
 FULL.male.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus") ),
-                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))
-
-Dom.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Dom")  ),
-                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
-
-mol.male.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(subsp == "Mol")  ),
-                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain), width = .2, alpha=0.5 )+ ggtitle("Q12 VAR male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))
 
@@
 FULL.female.Var.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") ),
-                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Musc male of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ylim(c(0,55))
+                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain), width = .2, alpha=0.5 )+ ggtitle("Q12 VAR female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free", ncol = 2)+ylim(c(0,55))
+
+FULL.female.Var.plot
 
@@
                              aes(color= strain, size = Ncells))+
-  geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
+  geom_jitter(aes(x = as.factor(strain), y =var, color= strain), alpha=0.5, width = .2, guide = FALSE )+ ggtitle("Q12 VAR Musc female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
 
-Dom.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Dom")  ),
-                             aes(color= strain, size = Ncells) ) +geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Dom female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
-
-mol.female.VAR.plot <- ggplot( ( FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(subsp == "Mol")  ),
-                             aes(color= strain, size = Ncells))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Q12 VAR Mol female of within mouse\nMLH1 variance")+scale_color_manual(values=colors_of_strains, guide = FALSE)
-
-FULL.female.Var.plot
 ```
 
-# Models (var)
+# observed values
+
+```{r pool.cell.calq, echo=FALSE}
+
+OG.vals_Q12.MLH1 <- ddply(Q12.MLH1_data, c("strain", "sex"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(nMLH1.foci),
+                        mean_co = as.numeric(format(round(mean(nMLH1.foci), 3), nsmall=3)),
+                        cV = cv(nMLH1.foci),
+                        var = as.numeric(format(round( var(nMLH1.foci),3), nsmall=3)),
+                        sd   = round(sd(nMLH1.foci), 3),
+                        se   = round(sd / sqrt(Ncells), 3)
+)
+
+OG.vals_Q12.MLH1 <- add_subsp(OG.vals_Q12.MLH1)
+OG.vals_Q12.MLH1 <- add_species(OG.vals_Q12.MLH1)
+
+#remove extra strains
+remove.strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
+vals_Q12.MLH1 <- OG.vals_Q12.MLH1[!OG.vals_Q12.MLH1$strain %in% remove.strains, ]
+
+#for males and females, calq the total variance across the strain-specific var
+Obs.f.VAS <- var( ( vals_Q12.MLH1 %>% filter(sex == "female") )$var )
+Obs.m.VAS <- var( ( vals_Q12.MLH1 %>%  filter(sex == "male") )$var )
+
+```
+
+These are the female strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$var`, the variance across strains (VAS) is `r Obs.f.VAS`.
+
+These are the male strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "male") )$var`, the variance across strains (VAS) is `r Obs.m.VAS`.
+(For `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$strain` respectively)
+
+```{r str.of.DF, echo=FALSE, eval=TRUE}
+#redundant
+#remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
+#Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]
+
+#Data structure (from the cell level data)
+len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
+len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )
+
+len.G.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
+len.G.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )
+
+len.LEW.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
+len.LEW.male <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "male") )$fileName )
+
+len.PWD.female <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "female") )$fileName )
+len.PWD.male <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "male") )$fileName )
+
+len.KAZ.female <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "female") )$fileName )
+len.KAZ.male <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "male") )$fileName )
+
+len.SKIVE.female <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$fileName )
+len.SKIVE.male <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$fileName )
+
+len.MSM.female <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "female") )$fileName )
+len.MSM.male <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "male") )$fileName )
+
+```
+
+# Mean and Variance correlations
+
+```{r var.x.mean, echo=FALSE, warning=FALSE, message=FALSE}
+
+#FULL_mouse_table_Q12.MLH1
+Fcor.test <- cor.test( (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_co,
+  (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "female")%>% filter(species == "M.musculus") )$var)
+
+Mcor.test <- cor.test( (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus")%>% filter(species == "M.musculus"))$mean_co,
+  (FULL_mouse_table_Q12.MLH1 %>% filter(sex == "male") %>% filter(species == "M.musculus"))$var)
+
+small.DF <- data.frame( corr = paste("corr =", c(Fcor.test$estimate, Mcor.test$estimate) ), 
+                     p.val = paste("p.val =", c(Fcor.test$p.value, Mcor.test$p.value) ), sex = c("female", "male") )
+
+Cor.plot <- ggplot(( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus") ), aes(x=mean_co, y=var, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= small.DF, aes(x=27, y=80, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= small.DF, aes(x=27, y=75, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by CO variance per Mouse")
+
+Cor.plot
+```
+
+# Models MLH1(var)
 
@@
 
-# Models (cV)
+# Models MLH1(cV)
+
 ```{r cVMLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
-
-#FULL_mouse_table_Q12.MLH1, remove molf not enough female
-
-#DF.for.model <- ( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus")  %>% filter(strain != "MOLF") %>% filter(strain != "AST") %>% filter(strain != "HMI") %>% filter(strain != "CAST") %>% filter(strain != "CZECH")  )
-
 #these are all mouse level values
@@
 
-cVMLH1.MM_interaction <- drop1(varMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
-cVMLH1.MM_sex <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-cVMLH1.MM_subp <- anova(varMLH1.M1, varMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(varMLH1.M1)$p.value
+cVMLH1.MM_interaction <- drop1(cVMLH1.M1, test="Chisq")$`Pr(Chi)`[2]
+cVMLH1.MM_sex <- anova(cVMLH1.M1, cVMLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+cVMLH1.MM_subp <- anova(cVMLH1.M1, cVMLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+cVMLH1.MM_strain<- MM.rand.pval <- exactRLRT(cVMLH1.M1)$p.value
 
 cVMLH1.MMlmer_results <- data.frame(sub.sp = cVMLH1.MM_subp, 
-sex.results =  varMLH1.MM_sex, 
+sex.results =  cVMLH1.MM_sex, 
                interaction.pval = cVMLH1.MM_interaction,  rand.pval = 
@@
 
-# Permutations
+# Permutations; MLH1
 
@@
 
-#Permutation Step Up
-
+<!--
 Here's the permutation test I was imagining:
@@
 8. Compare observed variance in variances to null distribution.  
+-->
 
-#observed values
-
-```{r pool.cell.calq, echo=FALSE}
-
-OG.vals_Q12.MLH1 <- ddply(Q12.MLH1_data, c("strain", "sex"), summarise,
-
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(nMLH1.foci),
-                        mean_co = as.numeric(format(round(mean(nMLH1.foci), 3), nsmall=3)),
-                        cV = cv(nMLH1.foci),
-                        var = as.numeric(format(round( var(nMLH1.foci),3), nsmall=3)),
-                        sd   = round(sd(nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-)
-
-OG.vals_Q12.MLH1 <- add_subsp(OG.vals_Q12.MLH1)
-OG.vals_Q12.MLH1 <- add_species(OG.vals_Q12.MLH1)
-
-#remove extra strains
-remove.strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
-vals_Q12.MLH1 <- OG.vals_Q12.MLH1[!OG.vals_Q12.MLH1$strain %in% remove.strains, ]
-
-#for males and females, calq the total variance across the strain-specific var
-Obs.f.VAS <- var( ( vals_Q12.MLH1 %>% filter(sex == "female") )$var )
-Obs.m.VAS <- var( ( vals_Q12.MLH1 %>%  filter(sex == "male") )$var )
-
-```
-
-These are the female strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$var`, the variance across strains (VAS) is `r Obs.f.VAS`.
-
-These are the male strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "male") )$var`, the variance across strains (VAS) is `r Obs.m.VAS`.
-(For `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$strain` respectively)
-
-```{r str.of.DF, echo=FALSE, eval=TRUE}
-#redundant
-#remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
-#Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]
-
-#Data structure (from the cell level data)
-len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
-len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )
-
-len.G.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
-len.G.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )
-
-len.LEW.female <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "female") )$fileName )
-len.LEW.male <- length( ( Q12.MLH1_data %>% filter(strain == "LEW") %>% filter(sex == "male") )$fileName )
-
-len.PWD.female <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "female") )$fileName )
-len.PWD.male <- length( ( Q12.MLH1_data %>% filter(strain == "PWD") %>% filter(sex == "male") )$fileName )
-
-len.KAZ.female <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "female") )$fileName )
-len.KAZ.male <- length( ( Q12.MLH1_data %>% filter(strain == "KAZ") %>% filter(sex == "male") )$fileName )
-
-len.SKIVE.female <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$fileName )
-len.SKIVE.male <- length( ( Q12.MLH1_data %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$fileName )
-
-len.MSM.female <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "female") )$fileName )
-len.MSM.male <- length( ( Q12.MLH1_data %>% filter(strain == "MSM") %>% filter(sex == "male") )$fileName )
-
-```
-
-<!--
-random permutations below, (COMPUTE the VAS)
-
-For the permuted data set, **pool all cells within a strain (permuted strain ie defined by sample size?) and compute the variance for each strain**.
-
-6. For the permuted data set, compute the variance across strains of the variance within strains. This will leave you with a single estimate for the set of strains.
-
-7. Repeat 4-6 10,000 times to establish a null distribution for the variance in variances.
-
-8. Compare observed variance in variances to null distribution.  -->
-
-# Permutation Output
-
-```{r permutations2, echo=FALSE}
+```{r permutations, echo=FALSE}
 #I made a DF of strain X sex specific variance
@@
 
-#sampling.size <- nrepeat-3
 #careful aboout matching the size of the dataframe!!
-
+#calculate the variance across permutations
 list.VAS_F <- c()
@@
 ```
+
+# Permutations: var(F) - var(M)
+
+1. Choose 1 strain.
+2. Compute femaleVar-maleVar (pooling all high-quality cells within sexes for that strain).
+3. Permute cells between sexes.
+4. Compute femaleVar-maleVar for the permuted dataset.
+5. Repeat 3-4 10,000 times.
+6. Compare the observed value to the distribution of permuted values to compute a p-value.
+7. Repeat 1-6 separately for each strain.
+
+```{r Obs.SDvar, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
+
+#observed sex difference in varince, by strain
+#Q12.MLH1_data by strain
+
+Diff.Var  <- ddply(Q12.MLH1_data, c("strain"), summarise,
+
+                  N.F.mice = length( unique( mouse[which(sex == "female")])  ),
+                  N.F.cells  = length( unique(fileName[which(sex == "female")]) ),
+
+                  N.M.mice = length(unique(mouse[which(sex == "male")]) ),
+                  N.M.cells = length(unique(fileName[which(sex == "male")]) ),
+
+                F.mean.MLH1 = mean( nMLH1.foci[which(sex == "female")]),
+                F.var.MLH1 = var(nMLH1.foci[which(sex == "female")]),
+
+                M.mean.MLH1 = mean( nMLH1.foci[which(sex == "male")]),
+                M.var.MLH1 = var(nMLH1.foci[which(sex == "male")]),
+
+               Fvar.Mvar.dif = F.var.MLH1 - M.var.MLH1
+
+)
+
+Diff.Var <- add_subsp(Diff.Var)
+Diff.Var <- add_species(Diff.Var)
+
+#plot the values of difference
+
+Var.axis.plot <- ggplot( data = (Diff.Var %>% filter(species == "M.musculus") ), aes(x=F.var.MLH1, y=M.var.MLH1, color=strain))+geom_point()+scale_color_manual(values=colors_of_strains) + ylim(c(0,30))+xlim(c(0,30))+ ggtitle("MLH1 variance by Sex")
+Var.axis.plot
+
+Dif.Var.plot <- ggplot( data = (Diff.Var %>% filter(species == "M.musculus") ), aes(x=strain, y=Fvar.Mvar.dif, color=strain))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE) + ggtitle("Degree of Sex Difference in MLH1 Variance")
+
+Dif.Var.plot
+
+```
+
+Above are the plots for observed sex differences for pooled cell variance.
+The first plot shows that all female values are higher than male, but PWD and KAZ have the lowest female variance and close to the highest male variance.
+
+```{r permute.by.strain, echo=FALSE}
+
+#pool.cells within strain, calc a Fvar and Mvar, calc the difference, plug into a lis / DF
+
+nrepeat = 10000
+
+FM.var.diff_permuted <- data.frame( "pwd.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "PWD")], Diff.Var$N.F.cells[Diff.Var$strain == "PWD"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "PWD")], Diff.Var$N.M.cells[Diff.Var$strain == "PWD"], replace = FALSE)  )
+                        )  ),
+
+             "msm.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MSM")], Diff.Var$N.F.cells[Diff.Var$strain == "MSM"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MSM")], Diff.Var$N.M.cells[Diff.Var$strain == "MSM"], replace = FALSE)  )
+                        ) ),
+
+             "skive.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "SKIVE")], Diff.Var$N.F.cells[Diff.Var$strain == "SKIVE"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "SKIVE")], Diff.Var$N.M.cells[Diff.Var$strain == "SKIVE"], replace = FALSE)  )
+                        ) ),
+             "wsb.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "WSB")], Diff.Var$N.F.cells[Diff.Var$strain == "WSB"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "WSB")], Diff.Var$N.M.cells[Diff.Var$strain == "WSB"], replace = FALSE)  )
+                        ) ),
+             "lew.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "LEW")], Diff.Var$N.F.cells[Diff.Var$strain == "LEW"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "LEW")], Diff.Var$N.M.cells[Diff.Var$strain == "LEW"], replace = FALSE)  )
+                        ) ),
+
+             "g.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "G")], Diff.Var$N.F.cells[Diff.Var$strain == "G"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "G")], Diff.Var$N.M.cells[Diff.Var$strain == "G"], replace = FALSE)  )
+                        ) ),
+
+              "kaz.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "KAZ")], Diff.Var$N.F.cells[Diff.Var$strain == "KAZ"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "KAZ")], Diff.Var$N.M.cells[Diff.Var$strain == "KAZ"], replace = FALSE)  )
+                        ) ),
+
+             "molf.FM.dif" = replicate(n=nrepeat, 
+              #female var         
+             (  var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MOLF")], Diff.Var$N.F.cells[Diff.Var$strain == "MOLF"], replace = FALSE) ) -
+              #male var
+              var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$strain == "MOLF")], Diff.Var$N.M.cells[Diff.Var$strain == "MOLF"], replace = FALSE)  )
+                        ) )
+
+)#dataframe, FM.var.diff_permuted
+
+#transpose the perrmuted DF, add a strain col  
+#t_FM.var.diff_permuted <- t(FM.var.diff_permuted)
+#Perm2.DF <- data.frame(names = row.names(t_FM.var.diff_permuted), #t_FM.var.diff_permuted); rownames(t_FM.var.diff_permuted) <- NULL
+#Perm2.DF <- add_sex_chp3(Perm2.DF) #sex doesn't make sense for this
+#Perm2.DF <- add_strain_Chap3(Perm2.DF)
+
+#calc p - values  (permuted > observed )
+#FM.var.diff_permuted
+
+```
+
+```{r plot.perm2, echo=FALSE, fig.width = 4, fig.height=2.5, message=FALSE}
+
+#plot the first permuted DF-- don't need the transposed
+pwd.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = pwd.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) + ggtitle("PWD difference in variance across sexes")  + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$pwd.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) / nrepeat )
+pwd.perm2.plot
+
+skive.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = skive.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) + ggtitle("SKIVE difference in variance across sexes") + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$skive.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) / nrepeat )
+skive.perm2.plot
+
+msm.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = msm.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) + ggtitle("MSM difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$msm.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) / nrepeat )
+msm.perm2.plot
+
+g.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = g.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) + ggtitle("G difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$g.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) / nrepeat )
+g.perm2.plot
+
+wsb.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = wsb.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) + ggtitle("WSB difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$wsb.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) / nrepeat )
+wsb.perm2.plot
+
+lew.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = lew.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) + ggtitle("LEW difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$lew.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) / nrepeat )
+lew.perm2.plot
+
+kaz.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = kaz.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "KAZ"] ) + ggtitle("KAZ difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=5, y=750, label= sum(FM.var.diff_permuted$kaz.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "KAZ"] ) / nrepeat )
+kaz.perm2.plot
+
+molf.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = molf.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MOLF"] ) + ggtitle("MOLF difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=9, y=750, label= sum(FM.var.diff_permuted$molf.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MOLF"] ) / nrepeat )
+molf.perm2.plot
+
+```
+
+The above plots show the difference of Female.var - Male.var for n=10,000 permutations.
+
+For all strains the female cell variance is larger than if cells are randomized across sex. But the degree of the difference varies.  PWD and KAZ females have smaller between cell variance.
+
+I should assess how these strain differences might be affected by the number of mice and cells, (espically MOLF).
+
+# Total SC
+
+I isolated cells with Q12 with MLH1 count and Skel measure. I removed rows that had a 'yes' for redo crop.
+
+```{r mouse.total.SC, echo=FALSE}
+
+# factor -> chr
+Total.SC_DF$n <- as.character(Total.SC_DF$n)
+Total.SC_DF$nMLH1.foci<- as.character(Total.SC_DF$nMLH1.foci)
+Total.SC_DF$quality<- as.character(Total.SC_DF$quality)
+
+Total.SC_DF$bin_size<- as.character(Total.SC_DF$bin_size)
+Total.SC_DF$skel_size<- as.character(Total.SC_DF$skel_size)
+
+#chr -> num
+Total.SC_DF$n <- as.numeric(Total.SC_DF$n)
+Total.SC_DF$nMLH1.foci<- as.numeric(Total.SC_DF$nMLH1.foci)
+Total.SC_DF$quality<- as.numeric(Total.SC_DF$quality)
+
+Total.SC_DF$bin_size<- as.numeric(Total.SC_DF$bin_size)
+Total.SC_DF$skel_size<- as.numeric(Total.SC_DF$skel_size)
+
+Q12.mouse.level.skel <- ddply(.data= (Total.SC_DF %>% filter(quality < 3) ),
+                          .(mouse),
+                          summarize, 
+                          ncells = length(unique(Original.Name)),
+                          nmice = length(unique(mouse)),
+                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
+                          var.MLH1 = var(nMLH1.foci),
+
+                          mean.skel = mean(skel_size),
+                          var.skel = var(skel_size),
+
+                          mean.bin = mean(bin_size)
+)
+
+Q12.mouse.level.skel <- add_strain(Q12.mouse.level.skel)
+Q12.mouse.level.skel <- add_sex(Q12.mouse.level.skel)
+Q12.mouse.level.skel <- add_category(Q12.mouse.level.skel)
+Q12.mouse.level.skel <- add_subsp(Q12.mouse.level.skel)
+Q12.mouse.level.skel <- add_species(Q12.mouse.level.skel)
+
+mouse.level.skel <- ddply(.data= (Total.SC_DF %>% filter(quality < 5) ),
+                          .(mouse),
+                          summarize, 
+                          ncells = length(unique(Original.Name)),
+                          nmice = length(unique(mouse)),
+
+                          Q12MLH1.mean = mean( nMLH1.foci[which(quality < 3)], na.rm = TRUE ),
+                          Q12MLH1.var = var( nMLH1.foci[which(quality < 3)], na.rm = TRUE ),
+            #  mean(chromosomeLength[which(hand.foci.count == 1)]
+                          Q12skel.mean = mean( skel_size[which(quality < 3)], na.rm = TRUE ),
+                          Q12skel.var = var( skel_size[which(quality < 3)], na.rm = TRUE ),
+
+                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
+                          var.MLH1 = var(nMLH1.foci),
+
+                          mean.skel = mean(skel_size),
+                          var.skel = var(skel_size),
+
+                          mean.bin = mean(bin_size)
+)
+
+mouse.level.skel <- add_strain(mouse.level.skel)
+mouse.level.skel <- add_sex(mouse.level.skel)
+mouse.level.skel <- add_category(mouse.level.skel)
+mouse.level.skel <- add_subsp(mouse.level.skel)
+mouse.level.skel <- add_species(mouse.level.skel)
+
+```
+
+```{r matching, echo=FALSE}
+
+#plot the differences between the larger dataset and Q12
+
+ii <- ggplot(mouse.level.skel, aes(x=Q12MLH1.mean, y=mean.MLH1, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+skel.skel <- ggplot(mouse.level.skel, aes(x=Q12skel.mean, y=mean.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)
+
+skel.skel.var.F <- ggplot( data = (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(sex == "female") ), aes(x=Q12skel.var, y=var.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)+facet_wrap(~ subsp, scales = "free", ncol=2)+ylim(c(0, 310000))+xlim(c(0, 310000))
+#for some reason, Q12 has a greater range -- but might just be a outlier
+
+#woah, sex differences are large. there is one female PWD with really large variance, 17apr15_PWD_f4 (11 cells)
+
+skel.skel.var.all <- ggplot( data = (mouse.level.skel %>% filter(species == "M.musculus") ), aes(x=Q12skel.var, y=var.skel, color=strain, size = ncells))+geom_point()+scale_color_manual(values=colors_of_strains, guide = FALSE)+xlim(c(0, 500000))+ylim(c(0, 500000))
+
+```
+
+```{r tot.sc.plot, echo=FALSE, warning=FALSE, message=FALSE}
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+#mouse.level.skel  #mouse level mlh1 and skel
+#table(mouse.level.skel$sex)
+
+#
+total.SC.plots.F <- ggplot( ( mouse.level.skel %>% filter(sex == "female")%>% filter(species == "M.musculus") %>% filter(strain != "CAST" )  ),  aes(color= strain, size = ncells))+geom_jitter(aes(x = as.factor(strain), y = var.skel, color= strain, alpha=0.5), width = .2 )+ggtitle("Variance of Total Skel Female by Mouse")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free", ncol=2)+ ylim(c(0, 350000 ) )
+
+total.SC.plots.M <- ggplot( ( mouse.level.skel %>% filter(sex == "male")%>% filter(species == "M.musculus") ),  aes(color= strain, size = ncells))+geom_jitter(aes(x = as.factor(strain), y = var.skel, color= strain, alpha=0.5), width = .2 )+ggtitle("Variance of Total Skel Male by Mouse")+scale_color_manual(values=colors_of_strains, guide = FALSE) + facet_wrap(~ subsp, scales = "free")+ ylim(c(0, 180000 ) )
+
+total.SC.plots.F
+total.SC.plots.M
+
+```
+
+This is the number of cells per category
+
+`r kable(table((Total.SC_DF %>% filter(quality < 3) )$sex, (Total.SC_DF %>% filter(quality < 3) )$strain ))`
+
+These plots are for the number of mice `r kable(table(Q12.mouse.level.skel$sex, Q12.mouse.level.skel$strain))`
+
+```{r corr.total.skel, echo=FALSE, warning=FALSE, message=FALSE}
+
+#FULL_mouse_table_Q12.MLH1
+tot.SC.Fcor.test <- cor.test( (Q12.mouse.level.skel %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean.skel,
+  (Q12.mouse.level.skel %>% filter(sex == "female")%>% filter(species == "M.musculus") )$var.skel)
+
+tot.SC.Mcor.test <- cor.test( (Q12.mouse.level.skel %>% filter(sex == "male") %>% filter(species == "M.musculus")%>% filter(species == "M.musculus"))$mean.skel,
+  (Q12.mouse.level.skel %>% filter(sex == "male") %>% filter(species == "M.musculus"))$var.skel)
+
+total.SC.small.DF <- data.frame( corr = paste("corr =", c(tot.SC.Fcor.test$estimate, tot.SC.Mcor.test$estimate) ), 
+                     p.val = paste("p.val =", c(tot.SC.Fcor.test$p.value, tot.SC.Mcor.test$p.value) ), sex = c("female", "male") )
+
+total.SC.Cor.plot <- ggplot(( Q12.mouse.level.skel %>% filter(species == "M.musculus") ), aes(x=mean.skel, y=var.skel, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= total.SC.small.DF, aes(x=2000, y=300000, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= total.SC.small.DF, aes(x=2000, y=250000, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by total SC variance per Mouse")
+
+```
+
+# Total SC Models of mouse-level variance
+
+The same glm models are being applied to variance and cV (M1 to M3) and sex specific models (M4 and M5). Out groups and strains with just one sex have been removed for M1-M3. Al factors are unordered.
+
+Changing the quality cut off I apply for the total SC dataset changes the significance of the sex coefficient.
+
+```{r total.SC.model.set.up, echo=FALSE, eval=T, warning=FALSE}
+
+#Total.SC_DF - cells
+#sex-specific   Q12.mouse.level.skel
+
+Q12.mouse.level.skel$subsp <- factor(Q12.mouse.level.skel$subsp, ordered = FALSE)
+
+Q12.mouse.level.skel$sex <- factor(Q12.mouse.level.skel$sex, ordered = FALSE)
+
+Q12.mouse.level.skel$strain <- factor(Q12.mouse.level.skel$strain, ordered = FALSE)
+
+mouse.level.skel$subsp <- factor(mouse.level.skel$subsp, ordered = FALSE)
+mouse.level.skel$sex <- factor(mouse.level.skel$sex, ordered = FALSE)
+mouse.level.skel$strain <- factor(mouse.level.skel$strain, ordered = FALSE)
+
+```
+
+```{r total.SC.var.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=T}
+
+#Q12.mouse.level.skel
+varTot.SC.M1 <- lmer(var.skel ~ subsp * sex + (1|strain), data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")  ) )
+
+varTot.SC.M1.reduce.sex <- lmer( var.skel ~ subsp  +(1|strain), data= 
+                                   ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")  )
+                                 )
+
+varTot.SC.M1.reduce.subp <- lmer( var.skel ~ sex  +(1|strain), data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC") ) )
+
+#summary(varTot.SC.M1)$coefficients
+
+varTot.SC.MM_interaction <- drop1(varTot.SC.M1, test="Chisq")$`Pr(Chi)`[2]
+
+varTot.SC.MM_sex <- anova(varTot.SC.M1, varTot.SC.M1.reduce.sex)$`Pr(>Chisq)`[2]
+
+varTot.SC.MM_subp <- anova(varTot.SC.M1, varTot.SC.M1.reduce.subp)$`Pr(>Chisq)`[2]
+
+varTot.SC.MM_strain <- exactRLRT(varTot.SC.M1)$p.value
+
+varTot.SC.MMlmer_results <- data.frame(  sub.sp = varTot.SC.MM_subp, 
+        sex.results =  varTot.SC.MM_sex, 
+               interaction.pval = varTot.SC.MM_interaction,  rand.pval = 
+varTot.SC.MM_strain)
+
+```
+
+**M1** Mixed model result for the mouse level for: subspecies, sex, interaction and strain.
+The p values for the total sc variance are as follow, `r varTot.SC.MMlmer_results`.
+There are significant subspecies, sex and interaction effects, (but no significant strain effects).
+This uses the larger dataset for total SC.
+
+**M2**
+
+```{r total.SC.varMLH1.glm.M2, echo=FALSE, eval=T}
+
+#strain = fixed, 
+totalSC.M2.glm.strain.fixed.full <- glm(var.skel ~ subsp * sex * strain, data= ( Q12.mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )
+
+summ.totalSC.M2.glm.strain.fixed.full <- summary(totalSC.M2.glm.strain.fixed.full)
+
+#round(summ.totalSC.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
+#adding cell number has some significant parts
+
+totalSC.M2.glm.strain.fixed.full <- glm(var.skel ~ subsp * sex * strain, data= ( mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )
+
+summ.totalSC.M2.glm.strain.fixed.full <- summary(totalSC.M2.glm.strain.fixed.full)
+
+round(summ.totalSC.M2.glm.strain.fixed.full$coefficients[,c(1,4)],5)
+```
+
+The sex effect for total SC variance isn't significant, only in interactions.
+
+**M3**
+
+```{r total.SC.varMLH1.glm.M3, echo=FALSE, eval=T}
+
+totalSC.M3.strain.fixed <- glm(var.skel ~ sex * strain, data= (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )
+
+summ.totalSC.M3.strain.fixed <- summary(totalSC.M3.strain.fixed)
+round(summ.totalSC.M3.strain.fixed$coefficients[,c(1,4)],5)
+
+```
+
+Again most strain effects than sex effects.
+
+**M4-M5 female**
+
+```{r total.SC.Female.specific.MLHvar.glm, echo=FALSE, eval=TRUE}
+
+totalSC.female.strain.subsp <- glm(var.skel ~ subsp * strain, data= (mouse.level.skel %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )
+
+summry.totalSC.female.strain.subsp<- summary(totalSC.female.strain.subsp)
+round(summry.totalSC.female.strain.subsp$coefficients[,c(1,4)],5)
+
+totalSC.female.glm.M2 <- glm(var.skel ~ strain, data= (mouse.level.skel %>% filter(sex == "female" ) %>% filter(species == "M.musculus") ) )
+
+sum_totalSC.female.glm.M2 <- summary(totalSC.female.glm.M2)
+round(sum_totalSC.female.glm.M2$coefficients[,c(1,4)],5)
+```
+
+**M4-M5 male**
+
+```{r total.SC.Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}
+
+#M4
+totalSC.male.glm.M4 <- glm(var.skel ~ subsp * strain, data= (mouse.level.skel %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
+sum_totalSC.male.glm.M4 <- summary(totalSC.male.glm.M4)
+round(sum_totalSC.male.glm.M4$coefficients[,c(1,4)],5)
+
+#M4.2
+totalSC.male.glm.M4.2 <- glm(var.skel ~ strain, data= (mouse.level.skel %>% filter(sex == "male" ) %>% filter(species == "M.musculus") ) )
+
+sum_totalSC.male.glm.M4.2 <- summary(totalSC.male.glm.M4.2)
+round(sum_totalSC.male.glm.M4.2$coefficients[,c(1,4)],5)
+
+```
+
+<!-- Permutations; Total SC
+
+1. structure of data, 2. permute all the data
+
+```{r}
+
+```
+-->
+
+# Direct tests of MLH1 and SC compaction
+
+The Q12 dataset for MLH1 and the larger dataset for total SC are used in these models.
+
+**MLH1variance = SCvariance * sex**
+
+```{r M.SC.models, echo=FALSE}
+#glm
+#Q12.mouse.level.skel[  ! Q12.mouse.level.skel$strain %in% remove_strains, ]
+
+the.two.sex.glm1 <- glm(Q12MLH1.var ~ var.skel + sex, data= (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   )  )
+
+summ.the.two.sex.glm1 <- summary(the.two.sex.glm1)
+round(summ.the.two.sex.glm1$coefficients[,c(1,4)],5)
+
+```
+
+**MLH1variance = SCmean * sex**
+
+```{r M2SC.skel, echo=FALSE}
+
+the.two.sex.glm <- glm(Q12MLH1.var ~ mean.skel + sex, data= (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )
+summ.the.two.sex.glm <- summary(the.two.sex.glm)
+round(summ.the.two.sex.glm$coefficients[,c(1,4)],5)
+
+```
+
+**MLH1variance = SCvariance * SCmean * sex**
+
+```{r M2SC.skel3, echo=FALSE}
+
+the.three.sex.glm <- glm(Q12MLH1.var ~ mean.skel + var.skel + sex, data= 
+                           (mouse.level.skel %>% filter(species == "M.musculus") %>% filter(strain != "AST") %>% filter(strain != "MOLF") %>% filter(strain != "CZECH") %>% filter(strain != "CAST") %>% filter(strain != "HMI") %>% filter(strain != "PERC")   ) )
+
+summ.the.three.sex.glm <- summary(the.three.sex.glm)
+round(summ.the.three.sex.glm$coefficients[,c(1,4)],5)
+
+```

### 2020-05-21 doc/MS_v2.Rmd
DEBUGGED and CLEANED up the setup rmd and MS_v2 scripts

@@
 library(RLRsim)
-#library(ggpubr)
+library(ggpubr)
 library(reshape2)
@@
 
-load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData
+load(file="~./MLH1repo/data/Results_setup_5.21.20.RData") #Results_setup_5.14.20.RData
+
+#load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData
 
@@
 
-We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murids to profile natural genetic variation in recombination (Table 1).
-Our survey included 5 strains from **M.
+We used a panel of wild-derived inbred strains of house mice ( _Mus musculus_ ) and related murids to profile natural genetic variation in recombination (Table 1).
+Our survey included 5 strains from _M.
 m.
-musculus**, 4 strains from _M.
+musculus_, 4 strains from _M.
 m.
@@
 m.
-molossinus_, 2 strains from **M.
+molossinus_, 2 strains from _M.
 m.
-castaneus**, and 1 strain each from **M.
-caroli**, **M.
-spicilegus**, and **M.
-spretus**.
+castaneus_, and 1 strain each from _M.
+caroli_, _M.
+spicilegus_, and _M.
+spretus_.
 We subsequently denote strains by their abbreviated subspecies and name (e.g.
-**domesticus^WSB^**).
+_domesticus^WSB^_).
 Mice were housed at dedicated, temperature controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
@@
 
-*new* - The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
+*new* - The same dry-down spread technique was applied to both spermatocytes and oocytes based on (@peters_1997), with adjustment for volumes.
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
@@
 To measure bivalent SC length, two image analysis algorithms were used.
-The first algorithm estimates the total (summed) SC length across bivalents for individual cells [@wang2019_sc].
-The second algorithm estimates the SC length of individual bivalents [@peterson2020].
+The first algorithm estimates the total (summed) SC length across bivalents for individual cells (@wang2019_SC).
+The second algorithm estimates the SC length of individual bivalents (@peterson2019).
 Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
 
-Total SC length per cell was quantified from pachtyene cell images [@wang2019_sc].
+Total SC length per cell was quantified from pachtyene cell images [@wang2019_SC].
 To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
-Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
-
-SC length of individual bivalents was quantified in pachytene cell images [@peterson2020].
-The DNA CrossOver algorithm [@peterson2020] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
+Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
+
+SC length of individual bivalents was quantified in pachytene cell images [@peterson2019].
+The DNA CrossOver algorithm [@peterson2019] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
 The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
 In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female).
-From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
-Following a manual curation, `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2020].
+From the total set of cell images, `r formatC( length(unique(cleanCurated_BivData %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
+Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
 The curated single bivalent data supplements our cell wide MLH1 count data with MLH1 foci counts for single bivalents.
@@
 
+#change the segements and text!!
+
 Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
@@
 
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
+male.plot <- ggplot( male_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
 
@@
 
-```{r MLH1.cell.examples, echo=FALSE}
+```{r MLH1.cell.examples, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 #Cell GGPLOT
-library(ggpubr)
 
@@
 C)	Male specific MLH1 counts per cell.', message=FALSE}
-
-#try cowplot for more control of sizing
-library("cowplot")
 
@@
 
-We next fit models separately for  `r length(fomo.male_mouse.av$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4).
+We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4).
 In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = 0--`r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).
@@
 ```{r male.strain.table, echo=FALSE}
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
+
+M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
                         Nmice =length(unique(mouse)),
@@
 glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-sum.glmMLH1Var_M3 <- summary(glmMLH1Var_M3)
+glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
 #round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
@@
 #Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>% 
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
+mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
+mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
+
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data= (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") )  )
 sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
 
-#subsp effect, many strain, #male
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") ) )
 sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
@@
 summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
 ```
@@
 
-MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
+Total.SC_DF$quality <- as.numeric(Total.SC_DF$quality)
+
+totSC_Q12_mouse.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
                  .(mouse),
@@
 #REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
+tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
+totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
+
+totSC_Q12_strain.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
                  .(strain, sex),
@@
 ```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+#
+
+Q1.mouse.av.biv_scatr_box   <- ggplot( (mouse.level.BivData_table %>% filter(species == "M.musculus") ), aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \n all bivalents")
@@
 
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.FULL <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos)
 
 #WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.WSB <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
 #G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.G <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
 #LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.LEW <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
 
 #PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.PWD <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
 
 #KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-#Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain == "SKIVE") )$norm.first.foci.pos,
- #          (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$norm.first.foci.pos )
+Q1_1CO.pos_t.test.KAZ <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
 
 #MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.MSM <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
 
@@
 #FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
+Q1.IFD.raw_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" ) %>% filter(species == "M.musculus")  )$mean_IFD.2CO_ABS,
+
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
 #p-value = 0.07
 
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
+Q1.IFD.PER_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>% filter(species == "M.musculus")    )$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER)
 #p-value = 1e-12
@@
 #WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
+Q1.IFD.raw_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
+                                (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
 
 #LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
+Q1.IFD.raw_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
 
 #G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
+Q1.IFD.raw_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
 
 #PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
+Q1.IFD.raw_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
 
 #KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
+Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
 
@@
 #MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
+Q1.IFD.raw_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
 
 #FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
+Q1.IFD.raw_sansG_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
 #p-value = 0.3
@@
 
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM  %>% filter(species == "M.musculus")  )
 summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
 
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
+glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.level.BivData_table  %>% filter(species == "M.musculus")  )
 summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
@@
 #PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= (mouse.level.BivData_table %>% filter(species == "M.musculus") ) )
 summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
@@
 *new* -- We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68 -- `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.56; t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value` ).
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68 -- `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.56; t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value` ).
 Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = 1.2610^{-25} -- `r lmer.Q1.nrm1CO_M1_results$sex.results` ; M2: p = 1.3310^{-7} -- `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 1.3310^{-7} -- `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
@@
 
-#DF needs to be melted -- so they all in 1 col
-#per code, I've done this by hand
-
-#subset DF for ease
-skinny.mouse.level <- mouse.level.BivData_table[-c(12:21) ]
-col.names.list.99 <- colnames(skinny.mouse.level)
-
-#all of the cols to use
-MELT_Chrm.Proportion <- melt(skinny.mouse.level[,c(  "mouse", "ncells","boxs.IDd", "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO",
+MELT_Chrm.Proportion <- melt(mouse.level.BivData_table[,c(
+        "mouse", "Nmice", "Ncells", "boxs.IDd", "nboxes_passed",  "nbiv",               "tot.biv", "mean_SC", "var.SC_length", "sd.SC_length", "se.SC_length",        "mean.SC_1CO", "mean.SC_2CO", "mean.SC_3CO", "mean.SC_4CO", "mean_1CO.pos",  "mean_1CO.pos.PER", "norm.first.foci.pos", "mean_IFD.2CO_ABS", "mean_IFD.2CO_PER", "mean_IFD_ABS", "mean_IFD_PER", "mean.siscoten", "mean.siscoten_1CO", "mean.siscoten_2CO", "mean.siscoten_3CO", "mean.telo.dist", "mean.cent.dist",     "mean.sis.co.ten", "sex", "strain", "subsp", "category", "Rec.group", "species",
 
@@
 
-  id=c("mouse", "ncells","boxs.IDd",    "nboxes_passed","Nbiv","tot.biv","n0CO","n1CO","n2CO","n3CO","n4CO"))
+  id=c("mouse", "Nmice", "Ncells", "boxs.IDd", "nboxes_passed",  "nbiv",               "tot.biv", "mean_SC", "var.SC_length", "sd.SC_length", "se.SC_length",        "mean.SC_1CO", "mean.SC_2CO", "mean.SC_3CO", "mean.SC_4CO", "mean_1CO.pos",  "mean_1CO.pos.PER", "norm.first.foci.pos", "mean_IFD.2CO_ABS", "mean_IFD.2CO_PER", "mean_IFD_ABS", "mean_IFD_PER", "mean.siscoten", "mean.siscoten_1CO", "mean.siscoten_2CO", "mean.siscoten_3CO", "mean.telo.dist", "mean.cent.dist",     "mean.sis.co.ten", "sex", "strain", "subsp", "category", "Rec.group", "species"))
 
@@
 MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
+MELT_Chrm.Proportion <- add_species(MELT_Chrm.Proportion)
 
@@
 
-#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
-#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
-
-#MELT_Chrm.Proportion$Rec.group <- ifelse(grepl("PWD male", MELT_Chrm.Proportion$category), 1, 
-#                              ifelse(grepl("MSM male", MELT_Chrm.Proportion$category), 1,
-#                                ifelse(grepl("SKIVE male", MELT_Chrm.Proportion$category), 1, 0)))
-
-#Chrm.prop.male.low <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 0)
-#Chrm.prop.male.HIGH <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 1)
-
 ```
@@
 
-Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
+Q1.totSC.mouse.av.table <-  mouse.total.skel %>% filter(species == "M.musculus")
 
@@
 
-totSC.strain.av.table <-  totSC.strain.av.table[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
+totSC.strain.av.table <-  mouse.total.skel[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
 
@@
 #ALL
-Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)
+Q1_tot.sc_t.test.all <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(species == "M.musculus")  )$mean.skel)
 
 #dom
-Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
+Q1_tot.sc_t.test.DOM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
 
 #WSB
-Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
+Q1_tot.sc_t.test.WSB <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
 #G
-Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
+Q1_tot.sc_t.test.G <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
 #LEW
-Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
+Q1_tot.sc_t.test.LEW <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
 
 #Musc
-Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
+Q1_tot.sc_t.test.MUSC <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
 
 #PWD
-Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
+Q1_tot.sc_t.test.PWD <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
 
 #KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
+Q1_tot.sc_t.test.KAZ <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
 
@@
 
-Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
-           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
+Q1_tot.scBIV_t.test.SKIVE <- t.test((Total.SC_DF %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
+           (Total.SC_DF %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
 
 #MSM
-Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
+Q1_tot.sc_t.test.MSM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
 
@@
 
-Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
+Q1.male.Dom.Musc.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
+
+Q1.male.Dom.Mol.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
 ```
@@
 
-Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
+Q1.lmer_totSC.mouse.av.table <- mouse.total.skel %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
 
@@
 
-mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
   facet_wrap(Rec.group~subsp, scales="free")+
@@
 #make plot of mouse averages for MLH1 and total SC
-bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
+bep <- merge(  male_mouse.avs_for.analysis,   mouse.total.skel)
 
@@
 
-Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+Q2tot.sc_t.test_High.Low <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
 #p-value = 0.007
 
-#mol
-Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+#mol -- not enough Mol observations
+#Q2tot.sc_t.test_High.Low_Mol <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
 
 #musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
 
@@
 #Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
+Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
 #p-value = 0.5
@@
 #Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
+Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
 #p-value = 0.1
@@
 
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+Male.poly.Mouse.Table_BivData_4MM <- mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")
+
+plot.blah.mouse.av <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
                           aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
@@
 
-Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
+Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
 
@@
 m.
-molossinus (p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`) but not within M.
+molossinus (p = 0.03 -- r Q2tot.sc_t.test_High.Low_Mol -- p.value ) but not within M.
 m.
@@
 #boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+male.bivdata.1CO <- cleanCurated_BivData %>% filter(sex == "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
 
@@
 
-mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
+mus.av1COpos.plot <- ggplot(data= mouse.level.BivData_table %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
                       aes(x= mouse, y = norm.first.foci.pos))+geom_point()
@@
 #subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+male.biv.2CO <- cleanCurated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
 
@@
 #M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
+Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")  )
 sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
@@
 #M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
+Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 
@@
 #M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
 
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
@@
 #Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
 #summary(Q2.glm.M2_IFD_Dom)
@@
 #Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
 #summary(Q2.glm.M2_IFD_Musc)
@@
 #Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol") )
 #summary(Q2.glm.M2_IFD_Mol)
@@
 #Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
 #summary(Q2.glm.M2_IFD.PER_Dom)
@@
 #Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
 #summary(Q2.glm.M2_IFD.PER_Musc)
@@
 #Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol")  )
 #summary(Q2.glm.M2_IFD.PER_Mol)

### 2020-05-21 doc/totalSC.Rmd
DEBUGGED and CLEANED up the setup rmd and MS_v2 scripts

@@
 library(dplyr)
+library(tidyr)
+library(xlsx)
+
 #load main MLH1 files to merge MLH1 cell data with the sc data 
@@
 
-#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
-SC.skel = read.csv("~./MLH1repo/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
-
-#load BivData  
-load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")
-
-#Curated list
-
-#strain specific xlsx sheets (why?)
-
 # Total.SC
@@
 # output: DFs of merged MLH1 and Skel data, cell and mouse levels
+
+#these data generated by running "python script --, it should be re-run when new cells are added to the anon.csv file
+#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)
+
+#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+
+Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)
+
+#this most recent version -- is missing most measures for MOLF_m
 ```
 
-# Goals
+Read in most recent output, remove the outliers (by strain), WRITE out the excel file for the nice clean total sc file, output from RW SC_skel run over Anon_Data images.
+The file names are random  The Main dataframe is the Merge MLH1 + skel data.
 
-Why include these results in the Main MLH1 results?
-
- - test if **greater SC area in females is a conserved trait of heterochiasmy**
-
- - test how average total SC correlates / contributes to mean CO
-
-## Discribe these Data
-
-output from RW SC_skel run over Anon_Data images. The file names are random
-
-The Main dataframe is the Merge MLH1 + skel data.
-
-## Running the Python Script
+## Running the Python Script Notes
 
@@
 (should I update all the crops? How many images need cropping)
-
-### ToDo
-
-- How will I test these / and adjust for XX?
-
-- Look up the python output
-
-- Check for outliers in the skel data
-
-- rerun the Python script
-
-- check that the mice from mixed models are in these
-
-ToDo re-run the pipeline, to get measues for SKIVE and MOLF females
-molf has almost eq SC -- but this is just 1 female
 
@@
 
-```{r clean.data, echo=FALSE, warning=FALSE}
+# Clean Skel data
+
+```{r clean.skel.data, echo=FALSE, warning=FALSE}
+
 #format all factors to numbers
@@
 #3 duplicate rows in original DF -- because Random name was blank -- how will these be matched to total.SC?
+
 ```
 
-```{r merge.DF, echo=FALSE}
+```{r merge.MLH1.skel, echo=FALSE}
 
-#MERGEING MLH! with SKEL
+#MERGEING MLH1 with SKEL by Random.Name
 MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
@@
 
-```
-
-```{r grab.nice.mice.setup, echo=FALSE, warning=FALSE}
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-#notes for testing random effects
-
-#variance is still chr
-AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)
-
-#use the merged DF to filter by age 
-AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
-AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
-AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
-#main file, make everything unordered factors
-AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
-
-DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]
-
-DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
-                            subsp.y,subsp.x,
-                    category.y, category.x,
-                        strain.x,strain.y, mouse.1  )  )
-
-#REMOVE EXTRA STRAINS
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-
-DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
-
-#mean_co.HQ  (this might remove some of the MOLF males)
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
-#above should only apply to males
-
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse)|DF.HetC.MixedModel.HQ$mouse==""),]
+#no filename in the skel output
+#MLH1.merge.Skel.FF <- merge(original_DF, SC.skel, by.x = "fileName", by.y = "fileName", all= F)
+#OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
 
@@
 
-```{r check.lists, echo=FALSE}
+#clean total.DF
+
+Warning most of the number levels are factors, so to chanage to numbers -- they need to follow the order, factor -> chr -> number.
+
+str(Total.SC_DF)#quality is a factor. must go factor - chr - number
+
+```{r clean.total.DF, echo=FALSE}
+
+#this is a nice merged excel file, still has X, and might have some 'outlier' images
+Total.SC_DF.og <- Total.SC_DF
+#quality is mess up here
+
+Total.SC_DF$quality <- as.character(Total.SC_DF$quality)
+Total.SC_DF$X <- as.character(Total.SC_DF$X)
+
+#Total.SC_DF <- Total.SC_DF %>% filter(X != "X")
+#remvove Xs
+Total.SC_DF <- Total.SC_DF[ !grepl("X", Total.SC_DF$X) , ]
+#Total.SC_DF$X <- droplevels(Total.SC_DF$X)
+
+str(Total.SC_DF)
+#check that there are only numbers in the quality
+table(Total.SC_DF$quality)
+
+#jj <- Total.SC_DF[ grepl("X", Total.SC_DF$X) , ]
+
+#factor -> chr
+Total.SC_DF$n <- as.character(Total.SC_DF$n)
+Total.SC_DF$nMLH1.foci<- as.character(Total.SC_DF$nMLH1.foci)
+
+Total.SC_DF$skel_size<- as.character(Total.SC_DF$skel_size)
+Total.SC_DF$bin_size<- as.character(Total.SC_DF$bin_size)
+
+#chr -> num
+Total.SC_DF$quality<- as.numeric(Total.SC_DF$quality)
+Total.SC_DF$n <- as.numeric(Total.SC_DF$n)
+Total.SC_DF$nMLH1.foci<- as.numeric(Total.SC_DF$nMLH1.foci)
+
+Total.SC_DF$skel_size<- as.numeric(Total.SC_DF$skel_size)
+Total.SC_DF$bin_size<- as.numeric(Total.SC_DF$bin_size)
+
+#table(Total.SC_DF$quality)
+
+```
+
+Total.SC has fewer cells than MLH1.merge (maybe the outliers have been removed). MLH1.merge has 'X'.
+
+```{r check.DFs, echo=FALSE, eval=FALSE}
+
+#compare the total.SC_DF
+
 #MLH1.merge.Skel
@@
 #this loop puts list in fileName2
-
-library(tidyr)
 
@@
 
-```{r more.blahs, echo=FALSE}
-#WSB -- WSB_curation, WSB_female output on big External
-#not sure why Curated bivData catching more categories than this above
+```{r cleaning.up.MLH1.merge, echo=FALSE, eval=FALSE}
 
-#this table now has merged bin/skel, and have been curated (but it doesn't have data on the curated bivalents)
-#the curated list is the current bottle neck
-
-#I should start inventory  how many mice are in each curated category
-
-Skel.summary.table <- ddply(.data=skel.mice_from.MM.list, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                 mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
-                 var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-#add strain, category, etc
-
-```
-
-```{r cleaning.up, echo=FALSE}
 #clean up this DataFrame REMOVE 'X'
@@
 MLH1.merge.Skel <- MLH1.merge.Skel[ !grepl("x", MLH1.merge.Skel$X) , ]
+MLH1.merge.Skel <- MLH1.merge.Skel[ !grepl("X", MLH1.merge.Skel$X) , ]
 MLH1.merge.Skel <- MLH1.merge.Skel[!(is.na(MLH1.merge.Skel$quality) | MLH1.merge.Skel$quality==""), ]
@@
 
-#it seems like this has duplicates
-MLH1.merge.Skel.DUPs <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$Random.Name),]
-#with new dataframe, there are 0 duplicates..
-#make sure to seperate mice based on adult and juvinille
-
-#remove for now
-#MLH1.merge.Skel <- MLH1.merge.Skel %>% distinct()
-
 #keep list of outliers (need to adjust threshold with lastest batch)
@@
 MLH1.merge.Skel <- MLH1.merge.Skel[MLH1.merge.Skel$skel_size < 25000,]
+
+```
+
+# Check Duplicates
+
+```{r Duplicates, echo=FALSE, eval=FALSE}
+
+Total.SC_DF_DUP.list <- Total.SC_DF[duplicated(Total.SC_DF$Random.Name),]
+Total.SC_DF_DUP.list <- Total.SC_DF[duplicated(Total.SC_DF$fileName),]
+#no Duplicates in total.SC
+
+MLH1.merge.Skel_DUP.list <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$Random.Name),]
+MLH1.merge.Skel_DUP.list <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$fileName),]
+
+kk <- MLH1.merge.Skel$fileName[duplicated(MLH1.merge.Skel$fileName)]
+```
+
+total.SC doesn't have duplicates
+
+# Remove Outliers
+
+below are the list of cell images which have been removed because they look like outliers compared to other cells from that mouse.
+
+Total.SC_DF  vs  MLH1.merge.Skel,  I think these both have outliers.
+
+```{r totSC.viz4outliers, echo=FALSE}
+
+# -- EXAMINE the ggplot scatter plots for outliers
+#tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ #geom_jitter()+facet_wrap(~category, scales = "free")
+
+#round 1
+SC.skel.outliers_round1 <- c(
+  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
+  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
+            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
+
+    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
+                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
+              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",
+
+  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
+          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",
+
+  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
+          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",
+
+       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
+                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",
+
+   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
+   "6may15_17apr15_PWD_f1_sp1_7_rev",
+
+      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
+            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",
+
+    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
+    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
+  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
+ "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
+ "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
+ "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
+"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,
+
+  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
+    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
+
+# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
+"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
+
+#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
+"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
+
+#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
+"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
+"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
+"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
+
+#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
+"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
+ "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
+ "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
+ "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
+ "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
+ "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
+"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
+"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
+ "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
+ "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
+ "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
+ "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
+
+MLH1.merge.Skel <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
+
+remove.mice_round2 <- c(
+    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
+    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
+
+    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
+  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
+"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
+  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",
+
+  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
+  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
+"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
+"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
+
+# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
+"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
+
+#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
+"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
+
+#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
+"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
+#SPIC male  x
+#SPIC female  x
+#TOM x
+#AST x
+#CZECH male x
+
+#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
+ "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
+#SPRET female X
+#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
+"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
+)
+#3remove.mice_round2
+
+MLH1.merge.Skel <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% remove.mice_round2, ]
+
+#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
+#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
+all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
+
+Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% SC.skel.outliers_round1, ]
+Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% remove.mice_round2, ]
+#sum( length(SC.skel.outliers_round1), length(remove.mice_round2) )
+```
+
+`r sum( length(SC.skel.outliers_round1), length(remove.mice_round2) )` images were removed as probable outliers.  ~20 outliers were removed from total.sc.skel.
+
+```{r category.table, echo=FALSE, eval=FALSE}
+
+#I should start inventory  how many mice are in each curated category
+Skel.summary.table <- ddply(.data=skel.mice_from.MM.list, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                 mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
+                 var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+#add strain, category, etc
 
@@
 #Below threshold is are the values to use
+write.xlsx2(Total.SC_DF, "Total.SC_DF.xlsx", sheetName = "Sheet1", 
+            col.names = TRUE, row.names = FALSE, append = FALSE)
 
-#write.xlsx2(MLH1.merge.Skel, "Total.SC_merge.MLH1.xlsx", sheetName = "Sheet1", 
+#merge skel has more observations (are they dups?)
+#write.xlsx2(MLH1.merge.Skel, "Merge.MLH1.xlsx", sheetName = "Sheet1", 
 #            col.names = TRUE, row.names = FALSE, append = FALSE)
@@
 
-```{r ploting, echo=FALSE}
+```{r ploting, echo=FALSE, eval=FALSE}
 Total.SC.box <- ggplot(data = MLH1.merge.Skel, aes(y=skel_size, x=strain, color=sex) ) + geom_boxplot()+ggtitle("Total SC Measures")
@@
 
-```{r crop.stuff, echo=FALSE}
+```{r crop.stuff, echo=FALSE, eval=FALSE}
 
@@
 
-table of clean crop stat
-table of total (minus clean crop, stat)
- difference in means is the amount of error
+table of clean crop stat, table of total (minus clean crop, stat, difference in means is the amount of error
 
-for some categories the difference is ~20% between the means
-usuall difference means the croped version has less compared to other
+For some categories the difference is ~20% between the means
+.. usuall difference means the croped version has less compared to other.
 
-```{r more.plot, echo=FALSE}
+```{r more.plot, echo=FALSE, eval=FALSE}
 
@@
 
-```{r table.making, echo=FALSE}
+```{r table.making, echo=FALSE, eval=FALSE}
 #mouse level
@@
 
-```{r mouse.level.plot, echo=FALSE, warning=FALSE}
+```{r mouse.level.plot, echo=FALSE, warning=FALSE, eval=FALSE}
 

### 2020-05-21 src/Result_Setup.Rmd
DEBUGGED and CLEANED up the setup rmd and MS_v2 scripts

@@
 author: "April Peterson"
-date: "April 7, 2020"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 output: html_document
@@
 
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
+#just load the data_setup, since I load most everything else
+
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")
+#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
 #load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
@@
 
-load(file="~./MLH1repo/data/CleanBivData_5.12.20.RData") #CleanBivData_5.12.20.RData
-
-load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
+#load(file="~./MLH1repo/data/CleanBivData_5.12.20.RData") #CleanBivData_5.12.20.RData
+#load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
 
 #skeletonize
-SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
+#Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)
+#remove X -- this table isn't used
 
@@
 
-```{r organizing.data, echo=FALSE}
+```{r organizing.data, echo=FALSE, eval=FALSE}
 #set the order of another column, based on another variable. consistant order for ploting
@@
 AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-
-var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
-  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
-
-```
-
-```{r Q1.rbar, echo=FALSE, eval=FALSE}
+```
+
+```{r Q1.rbar.gamma, echo=FALSE, eval=FALSE}
 
@@
 #plot rbar and gamma
-ll <- ggplot(table_gamma_estimates, aes(y=mean.rbar, x = all.biv.gamma, color = class, shape = sex ))+geom_point()+facet_wrap(~strain)+ ggtitle("")+ylim(c(0,.5))
+#ll <- ggplot(table_gamma_estimates, aes(y=mean.rbar, x = all.biv.gamma, color = class, shape = sex ))+geom_point()+facet_wrap(~strain)+ ggtitle("")+ylim(c(0,.5))
 
@@
 
-female_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
-                            subsp.y,subsp.x,
-                    category.y,category.x, strain.x,
-                    strain.y, mouse.1  )  )
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")
+female_mouse.avs_for.analysis <-  Mouse.Table_HQ %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")
 
@@
 
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-#Mol
-#Mol.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Mol")
-
-#Dom 
-Dom.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Dom")
-
 ```
@@
 #find the table that includes the las image data
-male_MLH1mouse.av <- AP_mouse_table_w.Ages
-
-male_MLH1mouse.av <- add_sex(male_MLH1mouse.av)
-male_MLH1mouse.av <- add_strain(male_MLH1mouse.av)
-male_MLH1mouse.av <- add_subsp(male_MLH1mouse.av)
-male_MLH1mouse.av <- add_species(male_MLH1mouse.av)
-
-fomo.male_mouse.av <-  male_MLH1mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-fomo.male_mouse.av$subsp <- factor(fomo.male_mouse.av$subsp, ordered = FALSE)
-fomo.male_mouse.av$sex <- factor(fomo.male_mouse.av$sex, ordered = FALSE)
-fomo.male_mouse.av$strain <- factor(fomo.male_mouse.av$strain, ordered = FALSE)
-fomo.male_mouse.av$species <- factor(fomo.male_mouse.av$species, ordered = FALSE)
-
-Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")
-Mol.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Mol")
+male_mouse.avs_for.analysis <-  Mouse.Table_HQ %>% filter(sex== "male") %>% filter(species == "M.musculus")
+
+male_mouse.avs_for.analysis$subsp <- factor(male_mouse.avs_for.analysis$subsp, ordered = FALSE)
+male_mouse.avs_for.analysis$sex <- factor(male_mouse.avs_for.analysis$sex, ordered = FALSE)
+male_mouse.avs_for.analysis$strain <- factor(male_mouse.avs_for.analysis$strain, ordered = FALSE)
+male_mouse.avs_for.analysis$species <- factor(male_mouse.avs_for.analysis$species, ordered = FALSE)
+
+#how much greater are the High Rec gwRRs
+male_mouse.avs_for.analysis$Rec.group <- ifelse(grepl("PWD", male_mouse.avs_for.analysis$strain), 1, 
+                              ifelse(grepl("MSM", male_mouse.avs_for.analysis$strain), 1,
+                                ifelse(grepl("SKIVE", male_mouse.avs_for.analysis$strain), 1, 0)))
+
+Musc.male_mouse.av <-  male_mouse.avs_for.analysis %>% filter(sex== "male") %>% filter(subsp == "Musc")
+Mol.male_mouse.av <-  male_mouse.avs_for.analysis %>% filter(sex== "male") %>% filter(subsp == "Mol")
 #Dom version 
-Dom.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Dom")
+Dom.male_mouse.av <-  male_mouse.avs_for.analysis %>% filter(sex== "male") %>% filter(subsp == "Dom")
 
@@
 ```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
-#DF.HetC.MixedModel.HQ
-#no mice removed for low cell numbers?
-AP_mouse_table.HQ <- add_species(AP_mouse_table.HQ)
 
 #DF.HetC.MixedModel <- add_species(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- AP_mouse_table.HQ[AP_mouse_table.HQ$species == "M.musculus",]
+DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]
 #DF.HetC.MixedModel
@@
 
-all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
-                 .(sex, strain),
-                 summarize,
-                 all.cells = sum(Ncells)
-)
+#what was this for
+#all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
+#                 .(sex, strain),
+#                 summarize,
+#                 all.cells = sum(Ncells)
+#)
 ```
@@
 ```{r male.difference, echo=FALSE}
-just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")
-
-just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
-just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
-just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
-just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)
-
-#how much greater are the High Rec gwRRs
-just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
-                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
-                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))
-
-#how much bigger are the high rec male means?
-sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
+
+sans_LowRec_male <- male_mouse.avs_for.analysis %>% filter(species == "M.musculus")  %>%   filter(Rec.group == 0)
 #mean(sans_LewRec_male$mean_co)
 
-PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
+PWD_male <-  male_mouse.avs_for.analysis %>% filter(strain == "PWD")
 #mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)
 
-MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
+MSM_male <- male_mouse.avs_for.analysis %>% filter(strain == "MSM")
 #mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)
 
-SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
+SKIVE_male <-   male_mouse.avs_for.analysis %>% filter(strain == "SKIVE")
 #mean(SKIVE_male$mean_co) /  
@@
 
-```{r CurBiv.Mouse.av.TABLE, echo=FALSE}
+cleanCurated_BivData  vs Curated_BivData, clean removed the few rows with bad handmeasure values and removed rows which are missing values for the hand foci count.
+
+```{r CurBiv.Mouse.av.TABLE, echo=FALSE, eval=FALSE}
 
@@
 
-mouse.avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
+#there's alread a Mouse.Table_BivData --mouse.level.BivData_table
+
+mouse.BivData_avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
 #REMOVE JUVI mice
-juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
-mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]
+juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice.list, ]  
+mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice.list, ]
 
@@
 
-strain.BivData.Table <- ddply(mouse.BivData.Table, c("strain", "sex"), summarise,
+strain.BivData.Table <- ddply(mouse.level.BivData_table, c("strain", "sex"), summarise,
 
@@
 
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=TRUE}
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+
+#this is redundant
 
@@
 #M4
-MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)
+MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data= male_mouse.avs_for.analysis)
 sum_MLH1.male.glm.M4 <- summary(MLH1.male.glm.M4)
@@
 #M4.2
-MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
+MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= male_mouse.avs_for.analysis)
 sum_MLH1.male.glm.M4.2 <- summary(MLH1.male.glm.M4.2)
@@
 ```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+#why is this block false? are the models in the MS.RMD file?
+
 MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
@@
 
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=TRUE}
+```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=F}
+
 #not including this model for now / brivity
@@
 
-```{r M3.MLH1.var, echo=FALSE, eval=TRUE}
+```{r M3.MLH1.var, echo=FALSE, eval=F}
 glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
@@
 
-```{r MLH1.var.M4, echo=FALSE, eval=TRUE}
+```{r MLH1.var.M4, echo=FALSE, eval=F}
 #male
@@
 
-```{r FEMALE.VAR.MODELS, echo=FALSE, eval=TRUE}
+```{r FEMALE.VAR.MODELS, echo=FALSE, eval=F}
 ## within mouse variance models 
@@
 
-```{r MALE.SPECIFIC.variance.model, echo=FALSE, eval=TRUE}
+```{r MALE.SPECIFIC.variance.model, echo=FALSE, eval=F}
 ## within mouse variance models 
@@
 
-```{r totSC_setup, echo=FALSE}
-#SC.skel
-#"anon.batch" "rand_name"  "bin_size"   "skel_size" , all batches included
-
-SC.skel.OG <- SC.skel #5683
-
-#MERGEING MLH1 with SKEL
-MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
-
-MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
-MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
-#5110,
+```{r totSC_setup, echo=FALSE, eval = F}
+
+#Total.SC_DF is the main dataframe and should be set up cleanly in the SetUpData.Rmd
+#SC.skel.OG <- SC.skel #5683
+
+tot.sc.x.list <- Total.SC_DF %>% filter(X == "X")
+
+Total.SC_DF.og <- Total.SC_DF[! Total.SC_DF$Random.Name %in% tot.sc.x.list$Random.Name, ]
+#quality is still a factor
+str(Total.SC_DF.og)
+Total.SC_DF.og$quality <- as.numeric(Total.SC_DF.og$quality)
+
+Total.SC_DF.og <- Total.SC_DF.og %>% filter(quality < 5)
 
@@
 #check for DUPs, #check for duplicatations of the random name
-MLH1.merge.Skel$ fileName  <- as.character(MLH1.merge.Skel$fileName)
+MLH1.merge.Skel$fileName  <- as.character(MLH1.merge.Skel$fileName)
 tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
@@
 #remove outliers
-MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
-
-```
+#MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
+
+```
+
+-- The total_SC outliers should be removed in the total.SC RData script
 
@@
 
+#Total.SC_DF
+#mouse.total.skel
+
 # -- EXAMINE the ggplot scatter plots for outliers
-tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
-
-#round 1
-SC.skel.outliers_round1 <- c(
-  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
-  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
-            "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
-
-    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 20000) 
-                      "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000) 
-              "1mar16_20feb16_G_f3_sp1_8_rev.tif","21nov16_16jan16_G_f2_sp1_15_rev.tif",
-
-  #MLH1.merge.Skel  %>% filter(category == "LEW male") %>% filter(skel_size < 1000)          
-          "6mar18_1feb18_LEW_m1_sp1_21.5_rev",
-
-  #MLH1.merge.Skel  %>% filter(category == "LEW female") %>% filter(skel_size > 20000)   
-          "12jan17_4jan17_LEW_f1_sp1_32_rev.tif","20mar17_13jan17_LEW_f2_sp1_12_rev.tif","9apr18_1feb18_LEW_f1_sp1_23_rev.tif","9apr18_1feb18_LEW_f1_sp1_12_rev.tif",
-
-       #MLH1.merge.Skel  %>% filter(category == "PWD male") %>% filter(skel_size > 20000)  
-                      "6mar15_1mar15_PWD_m1_sp4_6.1_rev.tif","7mar15_1mar15_PWD_m1_sp4_2_rev.tif","20jul15_18may15_PWD_m2_sp1_10.2_rev.tif",
-
-   #MLH1.merge.Skel  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
-   "6may15_17apr15_PWD_f1_sp1_7_rev",
-
-      #MLH1.merge.Skel  %>% filter(category == "KAZ male") %>% filter(skel_size < 1000)
-            "6aug19_10jul19_KAZ_m1_sp1_7.1_rev.tif",
-
-    # MLH1.merge.Skel  %>% filter(category == "KAZ female") %>% filter(skel_size > 30000)     
-    "16feb18_1feb18_KAZ_f4_sp1_20_rev.tif",  "8feb18_1feb18_KAZ_f1_sp1_24_rev.tif",  
-  "16feb18_1feb18_KAZ_f4_sp1_1_rev.tif",   "8feb18_1feb18_KAZ_f1_sp1_5_rev.tif",   
- "16feb18_1feb18_KAZ_f4_sp1_17_rev.tif",  "1mar18_1feb18_KAZ_f3_sp1_8_rev.tif",   
- "7feb18_11jan18_KAZ_f3_sp1_27_rev.tif",  "20apr18_11jan18_KAZ_f1_sp1_6_rev.tif", 
- "16feb18_1feb18_KAZ_f2_sp1_5.2_rev.tif", "16feb18_1feb18_KAZ_f2_sp1_15_rev.tif", 
-"7feb18_11jan18_KAZ_f3_sp1_13_rev.tif",  "13apr18_1feb18_KAZ_f6_sp1_25_rev.tif" ,
-
-  # MLH1.merge.Skel  %>% filter(category == "SKIVE female") %>% filter(skel_size > 10000)
-    "10jun19_31may19_SKIVE_f6_sp1_47_rev.tif","10jun19_31may19_SKIVE_f6_sp1_7_rev.tif",
-
-# MLH1.merge.Skel  %>% filter(category == "MOLF male") %>% filter(skel_size < 1200)
-"4jun19_28sep18_MOLF_m1_sp1_10.1_rev.tif",
-
-#MLH1.merge.Skel  %>% filter(category == "AST male") %>% filter(skel_size > 2100)
-"16aug19_27aug18_AST_m1_sp1_23_rev.tif",
-
-#MLH1.merge.Skel  %>% filter(category == "MSM female") %>% filter(skel_size > 20000)
-"21nov17_16nov17_MSM_f3_sp1_24_rev.tif","21nov17_16nov17_MSM_f3_sp1_3_rev.tif", "21nov17_16nov17_MSM_f3_sp1_27_rev.tif","4dec17_16nov17_MSM_f2_sp1_10_rev.tif",
-"5aug17_14jul17_MSM_f1_sp1_25_rev.tif","24jan18_31dec17_MSM_f3_sp1_3_rev.tif", 
-"24jan18_31dec17_MSM_f3_sp1_22_rev.tif","21nov17_16nov17_MSM_f3_sp1_17_rev.tif",
-
-#tt <- MLH1.merge.Skel  %>% filter(category == "SPIC female") %>% filter(skel_size > 20000)
-"27dec17_4dec17_SPIC_f1_sp1_12_rev.tif",  "5apr18_31dec17_SPIC_f2_sp1_1_rev.tif",
- "26sep18_30may18_SPIC_f4_sp1_14_rev.tif", "4feb18_31dec17_SPIC_f1_sp1_21_rev.tif", 
- "5apr18_31dec17_SPIC_f2_sp1_22_rev.tif",  "27dec17_4dec17_SPIC_f1_sp1_40_rev.tif", 
- "4jul18_30may18_SPIC_f1_sp1_6_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_6_rev.tif", 
- "29dec17_4dec17_SPIC_f2_sp1_10_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_37_rev.tif" , 
- "5sep17_21aug17_SPI_f2_sp1_15_rev.tif",   "26sep18_30may18_SPIC_f4_sp1_24_rev.tif",
-"29dec17_4dec17_SPIC_f2_sp1_16_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_30_rev.tif",  
-"4feb18_31dec17_SPIC_f1_sp1_5_rev.tif",   "5apr18_31dec17_SPIC_f2_sp1_25_rev.tif", 
- "26sep18_30may18_SPIC_f4_sp1_4_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_22_rev.tif",  
- "5sep17_21aug17_SPI_f2_sp1_19_rev.tif",   "5sep17_21aug17_SPI_f2_sp1_29_rev.tif",  
- "26sep18_30may18_SPIC_f4_sp1_3_rev.tif",  "5sep17_21aug17_SPI_f2_sp1_5.2_rev.tif", 
- "29dec17_4dec17_SPIC_f2_sp1_6_rev.tif",   "13oct17_21aug17_SPI_f1_sp1_13_rev.tif"  )
-
-MLH1.merge.Skel_1 <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
-
-remove.mice_round2 <- c(
-    #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
-    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
-
-    #MLH1.merge.Skel  %>% filter(category == "G female") %>% filter(skel_size > 20000)
-  "19oct15_12oct15_G_m1_sp1_17.1_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
-"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "PWD female") %>% filter(skel_size > 20000)
-  "6may15_17apr15_PWD_f1_sp1_7_rev.tif",
-
-  #MLH1.merge.Skel_2  %>% filter(category == "KAZ female") %>% filter(skel_size < 1000)
-  "8feb18_1feb18_KAZ_f1_sp1_19_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "LEW male") %>% filter(skel_size < 1000 | skel_size > 1750)
-"6mar18_1feb18_LEW_m1_sp1_10_rev.tif","3mar15_13jan17_LEW_m1_sp1_15_rev.tif", "4feb17_4jan17_LEW_m1_sp1_13_rev.tif","6mar18_1feb18_LEW_m1_sp1_21.5_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "LEW female") %>% filter(skel_size > 2750 ) 
-"20mar17_13jan17_LEW_f2_sp1_10_rev.tif",
-
-# MLH1.merge.Skel_2  %>% filter(category == "CAROLI male") %>% filter(skel_size < 1400 )
-"31aug18_1mar18_CAROLI_m1_sp3_29.2_rev.tif",
-
-#MLH1.merge.Skel_2  %>% filter(category == "MSM male") %>% filter(skel_size < 1200 )
-"16aug17_5jul17_MSM_m1_sp1_6.1_rev.tif",
-
-#tt <- MLH1.merge.Skel_2  %>% filter(category == "MSM female") %>% filter(skel_size > 2850 ) 
-"29nov17_16nov17_MSM_f1_sp1_10_rev.tif", "4dec17_16nov17_MSM_f2_sp1_24_rev.tif", "21nov17_16nov17_MSM_f3_sp1_26_rev.tif",
-#SPIC male  x
-#SPIC female  x
-#TOM x
-#AST x
-#CZECH male x
-
-#MLH1.merge.Skel_2  %>% filter(category == "SPRET male") %>% filter(skel_size < 1900 ) 
- "24may17_8may17_SPRET_m1_sp1_21.1_rev.tif",
-#SPRET female X
-#MLH1.merge.Skel_2  %>% filter(category == "HMI male") %>% filter(skel_size < 1000 ) 
-"29dec17_4dec17_HMI_m1_sp1_4.1_rev.tif"
-)
-#3remove.mice_round2
-
-MLH1.merge.Skel_2 <- MLH1.merge.Skel_1[!MLH1.merge.Skel_1$fileName %in% remove.mice_round2, ]
-#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1),]
-#MLH1.merge.Skel <- MLH1.merge.Skel[! (MLH1.merge.Skel$fileName %in% remove.mice_round2),]
-all.remove <- paste( c(SC.skel.outliers_round1, remove.mice_round2 ) )
-
-MLH1.merge.Skel <- MLH1.merge.Skel_2
-
-MLH1.merge.Skel <- add_species(MLH1.merge.Skel)
-
-```
-
-```{r tot.skel.table, echo=FALSE}
+show.tot.sc_plot <- ggplot(data= ( Total.SC_DF %>% filter(sex == "male") ), aes(y = skel_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
+```
+
+```{r tot.skel.table, echo=FALSE, eval=FALSE}
 #move to setup file
+
+#this is redundant, setup data made -- mouse.total.skel
+
+#make sure the table is clean -- no 5
 
@@
 
-```{r tot.SC_strain.setup, echo = FALSE}
-#move to setup
-
-totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
+```{r tot.SC_strain.setup, echo = FALSE, eval=FALSE}
+
+#not sure if this is used
+
+totSC.strain.av.table <- ddply(.data=mouse.total.skel, 
                  .(sex,strain),
@@
 totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
-
-```
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=TRUE}
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
 
@@
 
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE, eval=FALSE}
 
@@
 total SC area")
-
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE}
-
-#make plot of mouse averages for MLH1 and total SC
-bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
-
-X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
 
@@
 ```{r Q2.totSC.ttest, echo=FALSE}
+
 #totalSC high and low rec group
 
-Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol
-Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+Q2tot.sc_t.test_High.Low <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(species == "M.musculus") )$mean.skel,
+                (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+
+#mol -- why is there only 1 MOLF male now?
+#Q2tot.sc_t.test_High.Low_Mol <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% #filter(subsp == "Mol") )$mean.skel)
 
 #musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
+Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+
+```
+The molf male t-tests don't work because I only have total Sc measures for 1 MOLF male. to of the todo list if I get back to lab.
 
@@
 
-Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M4)$coefficients
@@
 
-Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
+Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
@@
 
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+mouse.total.skel <- add_strain(mouse.total.skel)
+mouse.total.skel <- add_sex(mouse.total.skel)
+mouse.total.skel <- add_subsp(mouse.total.skel)
 #within subspecies values
@@
 #Dom
-Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
+Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") %>% filter(species == "M.musculus") )
 sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
@@
 #Musc
-Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
+Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(species == "M.musculus") )
 sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
@@
 #Mol
-Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
+Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )
 sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)
@@
 
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+plot.blah.mouse.av <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
                           aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
@@
 
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+plot.logReg.mouse.av_1CO <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
                           aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
@@
 
-```{r log.reg_SC.length, echo=FALSE, include=FALSE}
+```{r log.reg_SC.length, echo=FALSE, include=FALSE, eval=FALSE}
 #REMOVE THIS COMPARES FROM FULL BIV DATA
-
-#male_mouse.av
-#Male.poly.Mouse.Table_BivData_4MM
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
 
@@
 #REMOVE THIS, IT uses total bivalent data
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
 
@@
 
+mouse.level.BivData_table <- add_species(mouse.level.BivData_table)
+
 #(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
@@
 #all of these are M4
-
-Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+#mouse table SC
+#mouse.level.BivData_table .. mouse.avs_4MM
+
+#why won't this work
+Q2.glm_mus.av_1COpos_M4 <- glm(mean_1CO.pos.PER ~ subsp * strain, data = (mouse.level.BivData_table  %>% filter(sex == "male") %>% filter(species == "M.musculus") ) )
+
 sum.Q2.glm_mus.av_1COpos_M4  <- summary(Q2.glm_mus.av_1COpos_M4)$coefficients
@@
 #M4.2 (just strain)
-Q2.glm_mus.av_1COpos_M4.2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Q2.glm_mus.av_1COpos_M4.2 <- glm(norm.first.foci.pos ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 sum.Q2.glm_mus.av_1COpos_M4.2 <- summary(Q2.glm_mus.av_1COpos_M4.2)$coefficients
@@
 #mouse.av - the mouse average IFD is from just 2CO
-t.test_mouse.av.IFD_ABS <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_ABS, 
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_PER <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") )$mean_IFD.2CO_PER   )
+t.test_mouse.av.IFD_ABS <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS, 
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_PER <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") )$mean_IFD.2CO_PER,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER   )
 
@@
 
-t.test_mouse.av.IFD_PER_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_PER   )
-
-t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_PER   )
+t.test_mouse.av.IFD_PER_MUSC <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER   )
+
+t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER   )
 
@@
 
-t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc"))$mean_IFD.2CO_ABS   )
-
-t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.avs_4MM %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS,
-   (mouse.avs_4MM %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
+t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(species == "M.musculus"))$mean_IFD.2CO_ABS,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male")  %>% filter(subsp == "Musc")%>% filter(species == "M.musculus"))$mean_IFD.2CO_ABS   )
+
+t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus"))$mean_IFD.2CO_ABS,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS  )
 ```
@@
 #M4.ABS
-Q2.M4glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
+Q2.M4glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male")  %>% filter(species == "M.musculus")  )
 sum.Q2.M4glm_IFD.raw <- summary(Q2.M4glm_IFD.raw)
@@
 #M4.PER
-Q2.M4glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
+Q2.M4glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")  )
 
@@
 #M4.2.raw
-Q2.M4.2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Q2.M4.2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 Sum.Q2.M4.2glm_IFD.raw <- summary(Q2.M4.2glm_IFD.raw)
 
-Q2.M4.2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
+Q2.M4.2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
 sum.Q2.M4.2glm_IFD.PER <- summary(Q2.M4.2glm_IFD.PER)
@@
 #Dom
-Q2.glm.M4.2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#Q2.glm.M4.2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
 #summary(Q2.glm.M4.2_IFD_Dom)
@@
 #Musc
-Q2.glm.M4.2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#Q2.glm.M4.2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
 #summary(Q2.glm.M4.2_IFD_Musc)
@@
 #Mol
-Q2.glm.M4.2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#Q2.glm.M4.2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
 #summary(Q2.glm.M4.2_IFD_Mol)
@@
 #Dom
-Q2.glm.M4.2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
+#Q2.glm.M4.2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
 #summary(Q2.glm.M4.2_IFD.PER_Dom)
@@
 #Musc
-Q2.glm.M4.2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
+#Q2.glm.M4.2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
 #summary(Q2.glm.M4.2_IFD.PER_Musc)
@@
 #Mol
-Q2.glm.M4.2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
+#Q2.glm.M4.2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
 #summary(Q2.glm.M4.2_IFD.PER_Mol)
@@
 #save Rdata
-save.image("~./MLH1repo/data/Results_setup_5.14.20.RData")
-
-```
+save.image("~./MLH1repo/data/Results_setup_5.21.20.RData")
+
+```

### 2020-05-21 src/SetupData.Rmd
DEBUGGED and CLEANED up the setup rmd and MS_v2 scripts

@@
 meta.data = read.csv( "~./MLH1repo/data/clean.Meta.Data.txt", sep = "\t", header=TRUE)
-#remove all MAD (peromyscus) from meta.data
-#meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]
-
-source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )
 
@@
 
-#total SC -- the one file should be all I need?
-Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)
-
-#Curated Bivdata - this excel file is too big to work with read.xlsx2
-#Curated_BivData = read.xlsx2("~./MLH1repo/data/Curated_BivData.xlsx", header=TRUE, 1)
+#total SC 
+Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
 
@@
 
+source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )
 ```
@@
 original_DF_duplicates <- original_DF_duplicates[!grepl("X", original_DF_duplicates$X),]
-#make a new DF -- with one fileName and the MLH1-1 and MLH1-2
-#  (duplicate / dostinct fileName )
+#these are images counted across mulitple batches
 
@@
 `r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.
+`r length(original_DF_duplicates$fileName)` rows (images) have been counted across mulitple batches.
+These are removed in the MLH1_data DF.
 
@@
 
+`r length(original_DF$fileName[ grepl("X", original_DF$X)] )` images (rows) have X's, they are removed from the main dataset.
+
 ```{r misc.setup, echo=FALSE}
@@
 ```
+
+# Mouse Data MLH1
 
@@
 #juvi mice ~ 2 weeks old
-juvi_mice <- c(
+juvi_mice.list <- c(
   "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
@@
 #make juvi.mice DF
-juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice),]
+juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice.list),]
 
 #remove juvi mice from big MLH1 -- but Then i need to remake the mouse_table
-MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice, ]
+MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice.list, ]
 
@@
 
-old.mice <- c(
+old.mice.list <- c(
   "12oct15_WSB_m1","9oct17_WSB_m1","1mar18_WSB_m2","21dec14_WSB_m2","21dec14_WSB_m3", "1mar18_WSB_m1",
@@
 
-I have made a list of the old mice, but from the effect document the older mice might not be that different.
-
-the excluded TOM, had low quality cells, only 4.
-The exclude AST, had low poor quality cells, there are 3 other mice.
-The excluded CZECH males are poor quality (but near the mean).
-There are 3 other males.
+`r length(old.mice.list)` old mice (> 25 weeks) were dissected, some of the RIKEN musc strains measures only have old mouse observations.
+
+`r length(juvi_mice.list)` juvinile mice were dissected, mostly only for DMC1 counts. Few had good MLH1 cells so they were removed from the main MLH1 DF.
 
@@
 #this is most important for plotting
-
 #reorder dataframe
@@
 
-```{r Date.imaged, echo=FALSE}
-
-#for issues with the total SC metrics, code in date imaged
-#use the add_mouse function
-
-#mini <- MLH1_data[1:500,]
-
+## MLH1 mouse level tables**
+
+```{r HQ.mouse.table, echo=FALSE}
+#Make a mouse level table
+
+#add date image
 count=1
@@
 
-#MLH1_data <- add_mouse2(MLH1_data)
-
-#mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
-#                        Nmice = length(unique(mouse)),
-#                        Ncells  = length(Random.Name),
-#                        #most recent date
-#                        oldest = min(date.imaged),
-#                        newest = max(date.imaged)
-#)
-
-```
-
-# Mouse Data
-
-Make mouse level tables  (re think this name, AP to distinguish BD -- but this isn't really an issue)
-
-```{r all.data.table, echo=FALSE, eval=FALSE}
-
-#make the table with ALL MLH1 obs
-FULL_mouse_table_MLH1 <- ddply(MLH1_data, c("mouse"), summarise,
-
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(nMLH1.foci),
-                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
-                        cV = cv(nMLH1.foci),
-                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
-                        sd   = round(sd(nMLH1.foci), 3),
-                        se   = round(sd / sqrt(Ncells), 3)
-)
-
-#NA cause by CAST female
-#this table isn't used much
-```
-
-```{r HQ.mouse.table, echo=FALSE}
-#Make a mouse level table
-
 #change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
@@
 
+                        #dates Imaged not in the DF
                         oldest.image.date = min(date.imaged),
@@
 Mouse.Table_HQ <- add_species(Mouse.Table_HQ)
+
+#remove juvi mice via the list
+#remove bad ages
+#this is removing female
+
+#this removes everything but males in these categories
+#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]
+
+#i don't understand why so many WSB female are removed..
+# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)
+
+```
+
+```{r merge.meta_MLH1, echo=FALSE}
+
+#merge - MLH1 with metdata
 
@@
 Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]
-
-#remove juvi mice via the list
-#remove bad ages
-#this is removing female
-
-#this removes everything but males in these categories
-#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]
-
-#i don't understand why so many WSB female are removed..
-# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)
 
@@
 
+Set up the dataframes in results or MS.
+
 Below I format the DF's for the models;
 
-i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus. start with **AP_mouse_table**
-
-```{r HetC.DF.4MM, echo=FALSE}
-
-#variance is still chr
-Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)
-
-DF.HetC.MixedModel <- Mouse.Table_HQ %>% filter(species == "M.musculus")
-
-#remove the extra merged cols
-#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
-#                            subsp.y,subsp.x,
-#                    category.y,category.x, strain.x,
-#                    strain.y, mouse.1  )  )
-
-#REMOVE STRAINS WITHOUT FEMALES
-#DF.HetC.MixedModel <- DF.HetC.MixedModel %>% filter( subsp != "Cast") %>% filter(strain != "AST")
-
-# These are removed in the analysis
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
-#remove musc strains
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
-#remove PERC, MOLF, CZECH, (no females)
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
-#F.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used
-
-#Only us the righ age window
-#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles
-
-#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)
-
-#remove blank rows (which are these?)
-#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
-#151 mice, no rows with nas for values
-
-#main df with variables coded as chr
-#DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
-#only female male reps
-
-```
-
-```{r male.specific.table,echo=FALSE}
-
-#AP_mouse_table_w.Ages
-#use the above table to pull the male mouse levels
-mlh1male_mouse.av <-  Mouse.Table_HQ %>% filter(sex== "male")
-
-#add the Rec.group markers
-#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
-#                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
-#                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )
-
-#formate the table for pluggin into models
-```
+i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus.
 
@@
 
-## Number of mice per category
+### Number of mice per category
 
@@
 
+```{r FULL.mouse.BivData.table, echo=FALSE}
+
+#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36
+
+table(Curated_BivData$hand.foci.count, exclude = NULL)
+#there are ~600 rows with SC and no hand foci count (blah)
+#Remove bivalent entries without hand foci count values
+cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
+                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
+
+length( (Curated_BivData  %>%  filter(  is.na(hand.foci.count) == T ) )$fileName )
+
+table(cleanCurated_BivData$hand.foci.count, exclude = NULL)
+
+mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
+                       .(mouse),
+                       summarize,
+
+                      Nmice = length(unique(mouse)),
+                       Ncells = length(unique(fileName)),
+                       boxs.IDd = max(boxNumber),
+                       nboxes_passed = length(boxNumber),
+                      nbiv = length(Obj.ID),
+
+                      tot.biv = length( !is.na(hand.foci.count) ),
+                      # mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
+                       mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
+                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
+                       sd.SC_length   = round(sd(chromosomeLength), 3),
+                       se.SC_length   = round(sd.SC_length / sqrt(nbiv), 3),
+                        SC.cV = cv(chromosomeLength),
+
+                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
+                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
+                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
+                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
+                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                  proportion_0CO = (n0CO / nbiv),
+                  proportion_1CO = (n1CO / nbiv),
+                 proportion_2CO = (n2CO / nbiv),
+                     proportion_3CO = (n3CO / nbiv),
+
+                  #calq mean.sc by chrm class
+                 mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                #mean 1CO.pos
+                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
+                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
+                #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+                # IFD things
+             mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
+             mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),
+
+              #sis.co.ten
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE),
+                 #sis.co.ten
+                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
+)
+
+mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
+mouse.level.BivData_table <- add_species(mouse.level.BivData_table)
+
+mouse.level.BivData_table <- add_category(mouse.level.BivData_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
+
+mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
+                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
+
+```
+
+```{r CurBiv.Mouse.av.TABLE, echo=FALSE, eval=FALSE}
+
+#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
+cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
+
+#cleanCurated_BivData
+#removed the no hand foci count
+
+#there's alread a Mouse.Table_BivData
+
+mouse.BivData_avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(unique( fileName) ),
+                        nbiv = length(Obj.ID),
+
+                        mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+              #can't figure out how to get this format to work  
+                       n0CO = sum(length( which(hand.foci.count == 0) )  ),
+                       n1CO = sum(length( which(hand.foci.count == 1) )  ),
+                      n2CO = sum(length( which(hand.foci.count == 2) )  ),
+                      n3CO = sum(length( which(hand.foci.count == 3) )  ),
+              #     
+                  proportion_0CO = (n0CO / nbiv),
+                  proportion_1CO = (n1CO / nbiv),
+                 proportion_2CO = (n2CO / nbiv),
+                     proportion_3CO = (n3CO / nbiv),
+
+                  #calq mean.sc by chrm class
+                 mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
+
+                #mean 1CO.pos
+                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
+                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
+                #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
+
+                # IFD things
+            #  mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
+            # mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+              #sis.co.ten!!
+              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
+              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
+              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
+)
+
+#vari
+mouse.avs_4MM <- add_sex(mouse.avs_4MM)
+mouse.avs_4MM <- add_strain(mouse.avs_4MM)
+mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
+mouse.avs_4MM <- add_category(mouse.avs_4MM)
+mouse.avs_4MM <- add_species(mouse.avs_4MM)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+#keep order for the plots?
+mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
+mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
+mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
+
+#REMOVE JUVI mice
+juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice.list, ]  
+mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice.list, ]
+
+#assign high and low
+#divide musc males into groups
+mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
+                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
+
+mouse.BivData.Table <- mouse.avs_4MM
+```
+
 ```{r no.Foci.Pos, echo=FALSE}
@@
 
-# Male specific BivData set
-
-not sure if SC.Length.male_table is used for anything.
-
-```{r male.sc.length.strain, echo=FALSE}
-
-#these measure ect should be in the male - mouse 
-#isolate all musc and male
-Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "M.musculus"),]
-
-#make strain average table
-SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
-                             mean_SC = mean(chromosomeLength),
-                              n_SC = length(Obj.ID),
-
-                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-)
-
-```
-
-# Cell Level Table
+## Cell Level Table BivData
 
@@
 
-# Reduced Bivalent Dataset (Long and Short)
+## Reduced Bivalent Dataset (Long and Short)
 
@@
 #each Q is a quartille
-
-#mini.now  <- BivData_from_15biv_cells[1:300,]
-
 for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){
@@
 
-```{r Look.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}
+The above chunk makes the *real.short.bivData* and *real.long.BivData* BivData.DFs.
+
+```{r plot.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}
 
 #showing SC lengths per cell, with lines for the quartille values.
-
-mini_G.15_15biv_FN <- BivData_from_15biv_cells %>% filter(category == "G male")
-#ploting short biv data
-G.15_15biv_male <- BivData_from_15biv_cells %>% filter(category == "G male")
-mini_G.15_15biv_FN <- sample( G.15_15biv_male$fileName, 10)
-mini_G.15_15biv  <- BivData_from_15biv_cells[BivData_from_15biv_cells$fileName %in% mini_G.15_15biv_FN, ]
-
-ppll <- ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
+ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
   geom_hline(aes(yintercept= cell.Q1), color='coral') + 
@@
 
-```{r table.long.biv, echo=FALSE}
-
+The chunck below makes the mouse level tabeles for the 2 reduced DFs.
+
+```{r Mouse.Long.Biv_table, echo=FALSE}
 # make a table for the Mixed Models -- from the long.BivData
-# Calq mouse averages
-
-#add in all the other stats there are in the other tables
 
@@
 #1) Long Biv for
+
 #remove all the non-used strains!!
-Long.biv_mouse_table_HetC <- Long.biv_mouse_table
-
-Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
-Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
-Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
-Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
+#Long.biv_mouse_table_HetC <- Long.biv_mouse_table
+#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
+#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
+#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
+#Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
 
 #remove na's
-Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]
+#Long.biv_mouse_table_HetC <- #Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | #Long.biv_mouse_table_HetC$mouse==""), ]
 
@@
 #just males, just Musculus
-Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
-
-Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
-
+#Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
+#Long.biv_mouse_table_male.poly <- #Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
 #table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
@@
 
-The above code chunk makes 2 mouse level tables; **Long.biv_mouse_table_HetC** and **Long.biv_mouse_table_male.poly**. Run these through the model systems.
-
-# Mouse Level Single BivData Table
+```{r Mouse.Short.Biv_table, echo=FALSE}
+# make a table for the Mixed Models -- from the long.BivData
+
+Short.biv_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        Nmice = length(unique(mouse)),
+                        Ncells  = length(Obj.ID),
+
+                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
+                        cV = cv(chromosomeLength),
+                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
+                        sd   = round(sd(chromosomeLength), 3),
+                        se   = round(sd / sqrt(Ncells), 3),
+
+                        #SC lengths by class
+                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
+                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
+                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
+                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
+                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),
+
+                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
+                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
+                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
+                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
+
+                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
+                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
+
+                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
+
+                       #CO position things, requires the melt?
+              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+
+        #          mean_siscoten = mean(siscoten)
+        #
+)
+#mouse averages for long.biv
+
+Short.biv_mouse_table <- add_sex(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_strain(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_subsp(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_category(Short.biv_mouse_table)
+
+#unordered for the mixed models (move this to the MM blocks...?)
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+
+#rec group
+Short.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Short.biv_mouse_table$category), 1, 
+                            ifelse(grepl("MSM", Short.biv_mouse_table$category), 1,
+                             ifelse(grepl("SKIVE", Short.biv_mouse_table$category), 1, 0)) )
+
+```
+
+The above 2 plots make mouse level tables for the reduced single biv datasets.
+
+## Mouse Level Single BivData Table
 
@@
 
-The block below creates clean.BivData, which removes ~600 rows if just SC measures
-(This might be important for the proportion code)
-
-```{r FULL.mouse.BivData.table, echo=FALSE}
-
-#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36
-
-table(Curated_BivData$hand.foci.count, exclude = NULL)
-#there are ~600 rows with SC and no hand foci count (blah)
-
-#remove bivalent entries without hand foci count values
-cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
-                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
-
-mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
-                       .(mouse),
-                       summarize,
-
-                       ncells = length(unique(fileName)),
-
-                       boxs.IDd = max(boxNumber),
-                       nboxes_passed = length(boxNumber),
-
-                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
-
-                      tot.biv = length( !is.na(hand.foci.count) ),
-
-                       n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
-                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
-                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
-                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
-                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),
-
-                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
-                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
-                       sd.SC_length   = round(sd(chromosomeLength), 3),
-                       se.SC_length   = round(sd.SC_length / sqrt(Nbiv), 3),
-
-                       #sis.co.ten
-                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
-)
-
-mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
-mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
-mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
-mouse.level.BivData_table <- add_category(mouse.level.BivData_table)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
-mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
-mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
-
-mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
-                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
-```
+The block below creates clean.BivData, which removes ~600 rows with just SC measures and no position data.
 
@@
 
-```{r mouse.table.Male.Poly, echo=FALSE}
+```{r mouse.table.Male.Poly, echo=FALSE, eval=FALSE}
+
+#this makes a redundant table
 
@@
 
-# Melt Position Table
+## BivData Melt Position Table
 
@@
 
-```{r MELT.block1, echo=FALSE}
+```{r BivData.MELT.block1, echo=FALSE}
 
@@
 
-```{r SHORT.biv_setup, echo=FALSE, eval=TRUE}
-
-#move to Data.setup
-
-#FROM q1
-#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-#from Q2 IFD
-#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-##long biv
-#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
-#table( droplevels( male.long.biv$strain)  )
-
-#FROM q1 
-REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO  
-                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                         )
-
-REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
-REAL.short_mouse_table <- add_category(REAL.short_mouse_table)
-
-REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
-REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
-REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
-                                           0)))
-
-###LONG BIV
-
-REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
-                       n_obs = length(Obj.ID),
-                        ncells = length(unique(fileName)),
-                         mean_SC = mean(chromosomeLength),
-
-                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
-                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
-                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-                         )
-
-REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
-REAL.long_mouse_table <- add_category(REAL.long_mouse_table)
-
-REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
-REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
-REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
-                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
-                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
-                                           0)))
-
-#find which one
-
-#REAL.short_mouse_table
-```
+The above chunk makes the MELT table (*MELT_Curated_BivData*), which can be used for Foci position comparison -- But I might not really use this for that.
 
@@
 
-total sc is processed in the results setup.
+The Total_SC excel file is made from total_SC.rmd, it removes 'X' cells and unclear quality scores.
+
+```{r clean.total.sc, echo=FALSE}
+#everything are factors again
+#clean
+Total.SC_DF$notes <- as.character(Total.SC_DF$notes)
+Total.SC_DF$REDO.crop <- as.character(Total.SC_DF$REDO.crop)
+
+redo.crops.rows <- Total.SC_DF[ grepl("yes", Total.SC_DF$REDO.crop) , ]#this catches most of the notes
+
+#remove 
+Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% redo.crops.rows$fileName, ]
+#284 rows removed
+
+# factor -> chr
+Total.SC_DF$n <- as.character(Total.SC_DF$n)
+Total.SC_DF$nMLH1.foci<- as.character(Total.SC_DF$nMLH1.foci)
+Total.SC_DF$quality<- as.character(Total.SC_DF$quality)
+
+Total.SC_DF$bin_size<- as.character(Total.SC_DF$bin_size)
+Total.SC_DF$skel_size<- as.character(Total.SC_DF$skel_size)
+
+#chr -> num
+Total.SC_DF$n <- as.numeric(Total.SC_DF$n)
+Total.SC_DF$nMLH1.foci<- as.numeric(Total.SC_DF$nMLH1.foci)
+Total.SC_DF$quality<- as.numeric(Total.SC_DF$quality)
+
+Total.SC_DF$bin_size<- as.numeric(Total.SC_DF$bin_size)
+Total.SC_DF$skel_size<- as.numeric(Total.SC_DF$skel_size)
+
+```
+
+```{r Mouse.total.sc.table, echo=FALSE}
+
+#Total.SC_DF
+
+#mouse level
+mouse.total.skel <- ddply(.data=Total.SC_DF,
+                          .(mouse),
+                          summarize, 
+                          ncells = length(unique(Original.Name)),
+                          nmice = length(unique(mouse)),
+                          mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
+                          var.MLH1 = var(nMLH1.foci),
+
+                          mean.skel = mean(skel_size),
+                          var.skel = var(skel_size),
+
+                          mean.bin = mean(bin_size)
+)
+
+mouse.total.skel <- add_strain(mouse.total.skel)
+mouse.total.skel <- add_sex(mouse.total.skel)
+mouse.total.skel <- add_category(mouse.total.skel)
+mouse.total.skel <- add_subsp(mouse.total.skel)
+mouse.total.skel <- add_species(mouse.total.skel)
+
+#add rec group
+mouse.total.skel$Rec.group <- ifelse(grepl("PWD male", mouse.total.skel$category), 1, 
+                              ifelse(grepl("MSM male", mouse.total.skel$category), 1,
+                                 ifelse(grepl("SKIVE male", mouse.total.skel$category), 1, 0)))
+
+```
+
+The above chunk makes mouse level *mouse.total.skel*, it uses the whole DF.
 
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")
 
@@
 ```
-
-# Deciding on Mice for Final Dataset
-
-Use the above information to make the final dataset for analysis
-
-- age (for males)
-- enough cells
-- both male and female

### 2020-05-25 doc/MS_v2.Rmd
working on finalizing figures

@@
 library(lattice)
-
 library(raster)
@@
 To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
-Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
+Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
 
@@
 The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
-In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female).
+In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from
+`r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female).
 From the total set of cell images, `r formatC( length(unique(cleanCurated_BivData %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
@@
 This “short bivalent” data set included the four or five shortest bivalents, and excluded the -- X bivalent in oocytes.
-A total of `r sum(Long.biv_mouse_table$n_obs)` -- 678 short bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` -- 103 oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` -- 37 spermatocytes.
+A total of `r sum(Short.biv_mouse_table$nobs )` -- 678 short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$ncells)` -- 103 oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$ncells)` -- 37 spermatocytes.
 Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
 A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
-To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD^raw^) was measured for those single bivalents containing two MLH1 foci.
-A normalized measure of interference (IFD^norm^) was computed by dividing IFDraw by SC length on a per-bivalent basis.
+`r sum(Long.biv_mouse_table$n_obs)`  number of long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and xx `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+
+To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
+A normalized measure of interference (IFD~norm~) was computed by dividing IFDraw by SC length on a per-bivalent basis.
 
 We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
-We first constructed a mixed model (M1) using the lmer() from the lmer4 package (cite) in R (v3.5.2)(Team 2015).
+We first constructed a mixed model (M1) using the lmer() from the lmer4 package (cite) in R (v3.5.2) [@Rstudio].
 In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT()).
 Subspecies, strain, and their interaction were coded as fixed effects,  with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()).
+*discribe the hiearchal nature*.
+strains nested within subspecies -- nesting by default,
+
 We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
-In separate analyses, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, variance in MLH1 count per cell, coefficient of variation in MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD^raw^, IFD^norm^, and average MLH1 position (for single-focus bivalents).
+In separate analyses, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, variance in MLH1 count per cell, coefficient of variation in MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
 Four additional models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from the initial mixed model.
@@
 
+# Figure 1
+
 ```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE}
@@
 #pooled
-strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
+strain.table.HQ_Fig.og <- ddply( MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                         Nmice =length(unique(mouse)),
@@
                          "SPRET","SPIC" ), order=T )
-
-sperman.corr.FM <- cor( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
-                        method = c("spearman") )
-
-sperman.corr.FM.pval <- cor.test( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
-                        method = c("spearman") )
-
-#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table2.xlsx", sheetName = "Sheet1", 
-#            col.names = TRUE, row.names = FALSE, append = FALSE)
-
-#qk.plot <- plot( ( strain.table.HQ_Fig %>% filter(sex == "female") )$mean_co,
-#      ( strain.table.HQ_Fig %>% filter(sex == "male") )$mean_co)
-
 ```
@@
 ```{r Fig1.top, echo=FALSE, warning=FALSE, message=FALSE}
-
+#Mouse.Table_HQ
 #change the segements and text!!
@@
 
-   #segments
+   #mol
   annotate("segment", x=.75, xend=2.3, y=19.5, yend=19.5, colour="black", size=1)+
 
-  #Musc, 30
-  annotate("segment", x= 3, xend=5.2, y=30, yend=30, colour="black", size=1)+
-
-  #mol
-  annotate("segment", x=5.5, xend=8.1, y=19.5, yend=19.5, colour="black", size=1)+
-
-  annotate("segment", x=8.5, xend=8.8, y=30, yend=30, colour="black", size=1)+
-
-  annotate("segment",x=9.5,xend=11.2, y=19.5, yend=19.5, colour="black", size=1)+
+  #Musc,
+  annotate("segment", x= 2.75, xend=5.2, y=30, yend=30, colour="black", size=1)+
+
+  #dom
+  annotate("segment", x=5.8, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
+
+  #cast
+  annotate("segment", x=8.8, xend= 9.3, y=30, yend=30, colour="black", size=1)+
+
     #spret
-
-  annotate("segment",x=9.5,xend=11.2, y=30, yend=30, colour="black", size=1)+
+  annotate("segment",x=9.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
+
   #spic
+  annotate("segment",x=10.5, xend=11.2, y=30, yend=30, colour="black", size=1)+
 
   #label
-    annotate("text", x = c(1.2,3.9,6.3,8,10, 11), #list of all x points 
+    annotate("text", x = c(1.6,3.9,6.3,8,10, 10.3), #list of all x points 
            y = c(18.5, 31, 18.5, 31, 18.5, 31),
@@
 
-female_mouse.av <- Mouse.Table_HQ %>% filter(sex == "female")
+#MLH1_data  %>% filter(quality < 5)
+#mouse avs
+
+female_mouse.av <- Mouse.Table_HQ.OG %>% filter(sex == "female") %>% filter(species == "M.musculus")
+
+female_mouse.av99 <- Mouse.Table_HQ %>% filter(sex == "female") %>% filter(species == "M.musculus")
 
@@
 
-female_mouse.avs_for.analysis <-  female_mouse.av %>% filter(sex== "female") %>% filter(species == "M.musculus")
-
-female_mouse.avs_for.analysis$subsp <- factor(female_mouse.avs_for.analysis$subsp, ordered = FALSE)
-female_mouse.avs_for.analysis$sex <- factor(female_mouse.avs_for.analysis$sex, ordered = FALSE)
-female_mouse.avs_for.analysis$strain <- factor(female_mouse.avs_for.analysis$strain, ordered = FALSE)
-female_mouse.avs_for.analysis$species <- factor(female_mouse.avs_for.analysis$species, ordered = FALSE)
-
-Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")
-
-female.specific.plot <- ggplot(female_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
+female_mouse.av$strain <- factor(female_mouse.av$strain, ordered = FALSE, levels =c( "MSM", "MOLF",
+  "PWD", "SKIVE", "KAZ",
+
+  "WSB", "G", "LEW",
+                                                                         "CAST", 
+                                                                   "SPRET", "SPIC") )
+
+female.specific.plot <- ggplot(female_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
   theme(legend.position="none",
@@
 
+male_mouse.avs_for.analysis$strain <- factor(male_mouse.avs_for.analysis$strain, ordered = FALSE, levels =c( "MSM", "MOLF",
+  "PWD", "SKIVE", "KAZ","TOM","AST","CZECH",
+
+  "WSB", "G", "LEW","PERC",
+                                                                         "CAST", "HMI",
+                                                                   "SPRET", "SPIC") )
+
 male.plot <- ggplot( male_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
@@
 MLH1.Figure1  <- ggdraw() +
-  draw_plot(Fig1_v1_strain, x = 0, y = .55, width=1, height=.45)+
+  draw_plot(Fig1_v1_strain, x = .02, y = .55, width=1, height=.45)+
 
   draw_plot(female.specific.plot, x=0, y=0, width=.5,height= .55)+
-  draw_plot(femaleMLH1.cell.obj, x=.32, y=.32, width=.19,height= .19)+
+  draw_plot(femaleMLH1.cell.obj, x=.35, y=.35, width=.18, height= .18)+
 
   draw_plot(male.plot, x=.5, y = 0,  width = .5, height = .55)+
-  draw_plot(maleMLH1.cell.obj, x=.85, y = .32,  width = .19, height=.19)+
+  draw_plot(maleMLH1.cell.obj, x=.85, y = .35,  width = .19, height=.19)+
 
   draw_plot_label(label = c("A", "B", "C"), size = 14,
-                  x = c(0, 0, .49), y = c(1, .55, 0.55) )
+                  x = c(0, .01, .51), y = c(.99, .55, 0.55) )
 
@@
 
-ggsave("~/MLH1repo/doc/Figure1.png")
+#ggsave("~/MLH1repo/doc/Figure1.png")
 ```
@@
 
-#main.table %>% filter(sex == "female")
-#main.table %>% filter(sex == "male")
-
-#G.HETC.val #1.2
-#LEW.HETC.val  #1
-#WSB.HETC.val  #1
-#KAZ.HETC.val  #1.1
-
-#SKIVE.HETC.val # 0.96
-#PWD.HETC.val  #0.88
-#MSM.HETC.val #0.9
-
-#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
-
-#strain.table.HQ_Fig
-
-#sum( (main.table  %>% filter(sex == "male"))$Ncells)
-#the kable version is bulkier than the other versaion
+```
+
+```{r female.strain.table, echo=FALSE}
+female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
+#female strain table
+#female_mouse.avs_for.analysis
+F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
+
+                        Nmice = length(unique(mouse)),
+
+                        total.cells  = sum(Ncells, na.rm = TRUE),
+
+                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#F.strain.table
+
+#F.strain.table
+
+# I don't know why most of the table is blank
+```
+
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
+
+#variance plots
+musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot
+
+Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.female.VAR.plot
+
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
+Dom.female.VAR.plot
+
+```
+
+```{r male.strain.table, echo=FALSE}
+
+M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+
+                        totNcells  = sum(Ncells, na.rm=TRUE),
+
+                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
+
+                        #Standard error (mice . cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#M.strain.table
+#same issue as the female table .. ugh
+```
+
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
+
+#variance plots
+musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+```
+
+```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
+#write the full table to a supplemental table
+#show the head (or shortened)
+#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+#kable(strain.table.HQ_Fig, "latex")
+
+df.strain.table <- strain.table.HQ_Fig
+#kable(df.strain.table)
+#pander(df.strain.table)
+
+#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
+#  kable_styling() %>%
+#  scroll_box(width = "800px", height = "200px")
+
+#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+
+#write excel
+#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table.xlsx", sheetName = "Sheet1", 
+#  col.names = TRUE, row.names = FALSE, append = FALSE)
 
@@
 spicilegus_ ).
-Mean MLH1 foci counts for -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 1,867 spermatocytes) and -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
+Mean MLH1 foci counts for `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$mouse) )` -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 1,867 spermatocytes) and -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
 Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
@@
 To further situate variation in recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
-We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < 10^{-6} -- p < `r 1/1000000`), sex (p = 1.5510^{-8} -- `r MLH1.MMlmer_results$sex.results`), subspecies (p=1.7210^{-4} -- `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = 3.110^{-5} -- `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
+We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < 10^{-6} -- p < `r 1/10000`), sex (p = 1.5510^{-8} -- `r MLH1.MMlmer_results$sex.results`), subspecies (p=1.7210^{-4} -- `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = 3.110^{-5} -- `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
 After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
@@
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p= 0.02 -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
+Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`),  *why is mol so different*  _molossinus^MSM^_  (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p= 0.02 -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
@@
 
-```{r female.strain.table, echo=FALSE}
-female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-
-                        Nmice = length(unique(mouse)),
-
-                        total.cells  = sum(Ncells, na.rm = TRUE),
-
-                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#F.strain.table
-
-#F.strain.table
-
-# I don't know why most of the table is blank
-```
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r male.strain.table, echo=FALSE}
-
-M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#M.strain.table
-#same issue as the female table .. ugh
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-```
-
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
-#write the full table to a supplemental table
-#show the head (or shortened)
-#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-#kable(strain.table.HQ_Fig, "latex")
-
-df.strain.table <- strain.table.HQ_Fig
-#kable(df.strain.table)
-#pander(df.strain.table)
-
-#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-#write excel
-#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table.xlsx", sheetName = "Sheet1", 
-#  col.names = TRUE, row.names = FALSE, append = FALSE)
-
-```
-
 ## B Synaptonemal complexes are longer in females
 
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE}
-
-MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-MLH1var_M2_sum <- summary(MLH1var_M2)
-
-#anova(MLH1var_M2)  
-#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
-
-MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M2)
-#anova(MLH1cv_M2)
-#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
-
-#Q12 var
-MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-sum.MLH1var_M2_Q12 <- summary(MLH1var_M2_Q12)
-#anova(MLH1var_M2_Q12)  
-#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M2_Q12)
-#anova(MLH1cv_M2_Q12)#sex large F
-#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-#MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-
-Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
+*new* -- Our survey of recombination rate across house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is reversed.
+In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-test; p < 0.05 -- ).
+In the dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.15 -- ( _musculus^MSM^_) to 1.49 -- ( _domesticus^WSB^_) across strains.
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect (p < 0.05 -- ).
+The existence of some subspecies and strain effects (p < 0.05 -- ) further indicates that SC length has evolved among strains and among subspecies.
+In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males have more MLH1 foci per cell.
+This pattern further suggests that in high-recombination strains, spermatocytes have less space in which to position additional crossovers than do oocytes.
+
+*old?
+-- * Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
 To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count per cell as the dependent variable.
@@
 
-## Sex Differences in Meiotic Traits
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table -- #can't find this table
-#Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>% 
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
-mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
-mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
-mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
-
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data= (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") )  )
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") ) )
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
-Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
-Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-```{r Q1.SC.Q12, echo=FALSE}
-
-#make comparisons for the high quality cells
-#get Q12 cells from MLH1.merge.Skel
-#calq the mouse averages
-
-Total.SC_DF$quality <- as.numeric(Total.SC_DF$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
-
-#unorder the factors
-totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
-totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
-totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
-totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
-
-#REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
-
-totSC_Q12_strain.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
-                 .(strain, sex),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
-
-#unorder the factors
-totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
-totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
-totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
-totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
-
-```
-
-```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
-
-Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Dom
-#Tot.SC_mouse.av_Dom
-
-Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Musc
-#Tot.SC_mouse.av_Musc
-
-Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-```
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-
-#Long.biv_mouse_table
-
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Long.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#not enough y observations
-#SKIVE
-#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
-#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
 ## B Females and males differ in crossover positions and crossover interference
 
-```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
-#move to setup
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
-#main figure of density plots,
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-#
-
-Q1.mouse.av.biv_scatr_box   <- ggplot( (mouse.level.BivData_table %>% filter(species == "M.musculus") ), aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos)
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test( 
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-#lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-
-```
+# triangles
 
@@
   #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
+  #png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
 
@@
 
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
-
-#each point is a 2CO, with the F1 and F2 position 
-#based on plot from Dourad et al (Mercier)
-
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
-    aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.5)+
-  ggtitle("Raw IFD WSB")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#trying to move x axis to top
-#cowplot might be the easier way -- since kinda diffecult in ggplot
-#or ggdraw
-
-P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" ) %>% filter(species == "M.musculus")  )$mean_IFD.2CO_ABS,
-
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>% filter(species == "M.musculus")    )$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
-#SKIVE
-#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-```
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
-
-#these are the scatter plots
-
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
-
-```
-
-```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-
-#M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-
-```
-
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM  %>% filter(species == "M.musculus")  )
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.level.BivData_table  %>% filter(species == "M.musculus")  )
-summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= (mouse.level.BivData_table %>% filter(species == "M.musculus") ) )
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-```
-
-```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-```
-
 ```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
@@
 
-Males have longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < 1.4910^{-12} --), with only _musculus^KAZ^_ showing no difference (t-test; p = 0.33 -- `r Q1.IFD.raw_FULL_t.test$p.value` ).
-Examination of IFD^norm^ distributions indicates that female values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Analysis of models treating IFD^norm^ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = 6.7410^{-14} --`r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
-In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains, whether they are considered separately or together.
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 1.4910^{-12} --), with only _musculus^KAZ^_ showing no difference (t-test; p = 0.33 -- `r Q1.IFD.raw_FULL_t.test$p.value` ).
+Examination of IFD~norm~ distributions indicates that female values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Analysis of models treating IFD~norm~ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = 6.7410^{-14} --`r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) across the full set of strains, whether they are considered separately or together.
 Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males and that males display a stronger telomeric bias in the placement of the distal crossover.
-In summary, controlling for differences in SC length (chromatin compaction) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical scale.
+In summary, controlling for differences in SC length (chromatin compaction) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD~raw~ shows that the sexes exhibit a similar level of interference on the physical scale.
 
@@
 
-```{r DMC1.table, echo=FALSE, message=FALSE}
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=Mouse.Table_HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-
-#attach this mini DF to the DMC1 data
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main88 <- merge(DMC1.target.strains, DMC1table, by="category")
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-#DMC1.mean.table
-DMC1table.main99 <- subset(DMC1table.main, select= -c(stage.x, var.MLH1, stage.y, mean.count.x, mean.count.y))
-
-DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
-
-colnames(DMC1table.main99) <- c("category", "mean MLH1",    "ncells",     "mean DMC1", "MLH1:DMC1 ratio", "ncells",     "mean DMC1", "MLH1:DMC1 ratio")
-
-#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-#  kable_styling() %>%
-#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#pander is better for word
-#pander(DMC1table.main99, row.names = F)
-
-#write table
-write.xlsx2(DMC1table.main99, file = "DMC1.Results_Table.xlsx",sheetName = "DMC1.results", append = FALSE, row.names=F)
-
-```
+turn off all these DMC1 tables -- to makes sure the right DFs are loaded in from the setup scripts.
 
@@
 
-  scale_color_discrete(name = "", labels = c("Late Zygotene", "Early Zygotene"))+
+  scale_color_discrete(name = "", labels = c("Early Zygotene", "Late Zygotene"))+scale_color_manual(values=c("#E69F00", "#0072B2"))+
 
@@
 
+#"#E69F00" orange
+#"#0072B2" darker blue
+
 ```
@@
 Green foci show DMC1.
-C)	Distribution of DMC1 counts per cell  by strains.", eval=FALSE}
+C)	Distribution of DMC1 counts per cell  by strains.", message=F, warning=F}
 
@@
 
-ggsave("~/MLH1repo/doc/Figure2.png")
-
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-#anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage #!= "P",],method = "t.test") %>% mutate(y_pos = 300)
-
-#DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-#  geom_boxplot(position=position_dodge()) + 
-#  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-#  ggsignif::geom_signif(
-#    data=anno_df_stage, 
-#    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-#    manual=TRUE
-#  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-#        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
+#ggsave("~/MLH1repo/doc/Figure2.png")
 
@@
 #data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-```{r chrm_prop_data, echo=FALSE, warning=FALSE}
-#shut off eval for now because of wd / path error
-#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#try adding the BivData version of the proportions
-
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
-
-                 n0CO = sum(n0CO, na.rm = TRUE),
-                  n1CO = sum(n1CO,na.rm = TRUE),
-                  n2CO = sum(n2CO,na.rm = TRUE),
-                  n3CO = sum(n3CO,na.rm = TRUE),
-                  n4CO = sum(n4CO,na.rm = TRUE),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#this is the table that needs melting
-#but is also easy to redo by hand
-
-#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
-```
-
-```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
-
-#add subspecies to these values¬
-melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
-
-#add the subsps
-melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
-
-melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
-
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))+theme(legend.position="none")
-
-#need a second DF / the first df
-#+ annotate() text 
-#annotate()text
-
-#proportion_stats_category,  tot_biv, ncells
-#add number of cells and bivalents in the corner
-
-#chrm.class.barplot
-```
-
-```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
-
-#prop.test(x, y)  x number of successes, y total  
-#strain sums of n
-#look up continuity correction
-
-#mouse.level.BivData_table <-- has more male strains
-#Mouse.Table_BivData_4MM <- only sex matched strains
-
-#perform test across strain 
-#removed all lines with no hand.foci.count entry
-Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
-
-                  nbiv = sum(tot.biv ),              
-                 n0CO = sum(n0CO ),
-                  n1CO = sum(n1CO ),
-                 n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO ),
-                 n4CO = sum(n4CO )
-                          )
-
-Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
-
-for(g in 1:length(Chrm.prop_strain_table$nbiv)){
-
-  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
-                                     Chrm.prop_strain_table$n1CO[g],
-                                     Chrm.prop_strain_table$n2CO[g],
-                                     Chrm.prop_strain_table$n3CO[g] ))  
-}
-
-Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
-#there are still some wacky entries -- but much fewer than before
-
-chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
-
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
-#the pvalues for 0 are not sig dif
-#the p values for 1 and 2CO are sig diff
-
-#seperate by subsp
-chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
-chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
-chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
-
-Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
-Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
-Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
-Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
-#WSB has fewer 2COs than other 3 strains
-
-Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
-Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
-Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
-#KAZ has slightly more 1CO than the other two
-
-#high.female_G.LEW.
-chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
-
-high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
-high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
-high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
-
-#Low.female.G.LEW.
-chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
-
-low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
-low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
-low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
-
-#all females
-
-chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
-
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
-#the pvals for these 3 class are sig dif
-
-#seperate by subsp
-chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
-chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
-chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
-
-Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
-Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
-Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
-#non of the Dom proportions are different
-
-Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
-Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
-Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
-#all v different
-
-Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
-Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
-Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
-
-#high vs low
-chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
-                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
-                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
-
-chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
-chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
-
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
-
-Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
-Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
-Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
-#the low strains do vary slightly in their 1CO and 2CO proportions
-
-#includes SKIVE
-Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
-Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
-Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
-#all very sig p values
-
-High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
-#these proportions are v different
-
-#no SKIVE
-High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
-High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
-High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
-#MSM and PWD only slightly sig dif 2CO
-
-#MSM-MOLF
-MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-#PWD-KAZ
-PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.MSM.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "MSM") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "MSM") )$nbiv )
-
 ```
@@
 
-```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
-#moved this to setupdata
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
+```{r Q1_SC.biv.short_ttest, echo=FALSE}
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+
+#WSB
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE - this value used in the text
+#only one SKVIE female, this is breaking the test
+#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
+
+#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
+#      )
+
+```
+
+```{r Q1SC_short.biv_dif, echo=FALSE, eval=TRUE}
+
+#calq the different in short.biv means across sexes
 #Long.biv_mouse_table
 
-#make the Q1 strain list
-Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-#nested boxplots, pooled Bivalents
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
+#WSB
+Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
+
+#LEW
+Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
+
+#G
+#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
+
+#PWD
+Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
+
+#KAZ
+Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
+
+#SKIVE
+Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
+
+#MSM
+Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
+
+#this bivData ratio is the same as mouse average
+Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
+
+```
+
+# Fig 3
 
@@
 
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE - this value used in the text
-#only one SKVIE female, this is breaking the test
-#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
-
-#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
-#      )
-
-```
-
-```{r Q1SC_short.biv_dif, echo=FALSE, eval=TRUE}
-
-#calq the different in short.biv means across sexes
-#Long.biv_mouse_table
-
-#WSB
-Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
-
-#LEW
-Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
-
-#G
-#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
-
-#PWD
-Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
-
-#KAZ
-Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
-
-#SKIVE
-Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
-
-#MSM
-Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
-
-#this bivData ratio is the same as mouse average
-Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
+ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")
 
@@
 
-Q1.totSC.mouse.av.table <-  mouse.total.skel %>% filter(species == "M.musculus")
+Q1.totSC.mouse.av.table <-  mouse.total.skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")
 
@@
 
-#italizing the facet label semms too hard -- do it in the final AI version
-#levels(powerDF$alph) <- c("italic('p=')*0.001", "italic('p=')*0.01", "italic('p=')*0.05")
+ggsave("~/MLH1repo/doc/Figure3_Tot.SC_mouse.av_ALL.png")
 
@@
 
-#use cowplot
-library(cowplot)
-
 Q2.SC_dif.Figure3  <- ggdraw() +
@@
                   x = c(0, .49, 0), y = c(1,1, 0.33) )
-
-#two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }
-
-#two.Q1.SC.plots + plot_annotation(
-#  title = '',
-#  subtitle = '') + plot_annotation(tag_levels = 'A')
 
@@
 #mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+Male.cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
 
@@
                          )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+Male.cell.av.long.biv.table <- add_mouse(Male.cell.av.long.biv.table)
+Male.cell.av.long.biv.table <- add_strain(Male.cell.av.long.biv.table)
+Male.cell.av.long.biv.table <- add_subsp(Male.cell.av.long.biv.table)
+
+Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
+Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
+Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
 
@@
 
-plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
+#plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
 
@@
 #p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
-
-# theses t.tests not working
-#mouse average
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((real.long.bivData  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean_SC,
-                          (real.long.bivData %>% filter(subsp == "Musc") %>% filter(sex == "male"))$mean_SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Short.biv_mouse_table  %>% filter(subsp == "Dom")%>% filter(sex == "male") )$mean_SC,
-                          (Short.biv_mouse_table %>% filter(subsp == "Musc")%>% filter(sex == "male"))$mean_SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test( (Long.biv_mouse_table  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean_SC,
-                          (Long.biv_mouse_table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean_SC )
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Short.biv_mouse_table  %>% filter(subsp == "Dom") %>% filter(sex == "male"))$mean_SC,
-                          (Short.biv_mouse_table %>% filter(subsp == "Mol")%>% filter(sex == "male"))$mean_SC)
-#p-value = 0.09
-
 ```
@@
 
-```
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
-
-Male.poly.Mouse.Table_BivData_4MM <- mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")
-
-plot.blah.mouse.av <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r log.reg_SC.length, echo=FALSE, include=FALSE}
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
 ```
@@
 We considered proportions of bivalents with different numbers of crossovers, double-strand breaks, SC length, and crossover positions, including crossover interference.
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")` -- 34,982) have either one or two MLH foci (Figure 3).
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` -- 34,982) have either one or two MLH foci (Figure 3).
 The proportions of single-focus (1CO) bivalents vs.
@@
 
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in _molossinus^MSM^_ (Figure 3).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in _molossinus^MSM^_ (*SUP* Figure 3).
 Following patterns in the genome-wide recombination rate, male musculusPWD and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
@@
 _molossinus^MOLF^_: p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
-These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
+PWD vs MS: p = `r PWD.MSM.2CO.prop.test$p.value`.
+**Add the extra test PWD-MSM**  These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
+
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of 76 -- `r sum( (DMC1table.main99$ncells.x) )`  early zygotene and 76 -- `r sum( (DMC1table.main99$ncells.y) )` late zygotene spermatocytes from three low-recombination strains ( _musculus^KAZ^_, _domesticus^WSB^_ , and  _domesticus^G^_) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_).
-
+DMC1 foci were counted in a total of 76 -- `r sum( (DMC1table.main$ncells.x) )`  early zygotene and 76 -- `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from three low-recombination strains ( _musculus^KAZ^_, _domesticus^WSB^_ , and  _domesticus^G^_) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_).
 High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < 10^{-6} --  p < `r 1/1000000` ).
@@
 Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
-
-Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within M.
+**Check if this is correct for all strains**.
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within _M.
 m.
-molossinus (p = 0.03 -- r Q2tot.sc_t.test_High.Low_Mol -- p.value ) but not within M.
+molossinus_ (p = 0.03 -- r Q2tot.sc_t.test_High.Low_Mol -- p.value ) but not within M.
 m.
@@
 
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- ; long: p = 0.18 -- ).
-In a model with total SC length as the dependent variable, two subspecies effects are significant ( _M.
+**Q2.SC.reduced** Q2short.biv_t.test_High.Low
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value` ; long: p = 0.18 -- `r Q2long.biv_t.test_High.Low$p.value`).
+In a model with **total SC length** as the dependent variable, two subspecies effects are significant ( _M.
 m.
-musculus_ p = 1.2410^{-6} -- , _M.
+musculus_ p = 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[2,4]` , _M.
 m.
-molossinus_, p = 10^{-6} -- ).
-In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05), but they are not consistent across models.
+molossinus_, p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[3,4]` ).
+In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05 -- ), but they are not consistent across models.
 Collectively, these observations reveal that evolution of the genome-wide recombination rate is not strongly associated with evolution of SC length.
-High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = 0.24 -- ) or within subspecies (t-test; _M.
+
+**Q2-CO.pos**  Q2.CO.pos t-test
+High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = 0.24 -- `r Q2.1COpos_Ttest$p.value` ) or within subspecies (t-test; _M.
 m.
-musculus_: p = 0.41 -- ; _M.
+musculus_: p = 0.41 -- `r Q2.1COpos_mus.av_Musc_ttest$p.value` ; _M.
 m.
-molossinus_ : p = 0.07 -- ).
-While domesticusWSB and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable (), these two strains do not differ in MLH1 count.
+molossinus_ : p = 0.07 -- r Q2.1COpos_mus.av_Mol_ttest$p.value ).
+**While domesticusWSB and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable** (Q2.COpos), these two strains do not differ in MLH1 count.
 These results show that evolution of crossover position on chromosomes with single crossovers is decoupled from evolution of the genome-wide recombination rate, at least at this scale of resolution.
 
-High-recombination strains have greater inter-focal distances than low-recombination strains (t-test; IFD^norm^: p = 7.7410^{-7} -- ; IFD^raw^: p = 8.7810^{-6} -- ).
-This pattern holds within M.
+**Q2_IFD**
+High-recombination strains have greater inter-focal distances than low-recombination strains (t-test; IFD~norm~: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value` ; IFD~raw~: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
+This pattern holds within _M.
 m.
-musculus (t-test; IFD^norm^: p = 2.0410^{-5} -- ; IFD^raw^: p = 1.9410^{-4}), with a trend in the same direction within M.
+musculus_ (t-test; IFD~norm~: p = 2.0410^{-5} -- `r  t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD~raw~: p = 1.9410^{-4} -- `r  t.test_mouse.av.IFD_ABS_MUSC$p.value`), with a trend in the same direction within _M.
 m.
-molossinus (IFD^norm^: p= 0.17 -- ; IFD^raw^: p = 0.08 -- ).
-Similar results are observed with models treating IFDraw and IFDnorm as dependent variables: only effects associated with high-recombination strains are significant (p < 0.05).
-That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths (and bivalent sizes) as the primary explanation.
-The main distinction in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains (Figure 6).
-The frequency of IFDnorm values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and musculusPWD, respectively).
+molossinus_ (IFD~norm~: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value` ; IFD~raw~: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
+Similar results are observed with models treating IFD~raw~ and IFD~norm~ as dependent variables: only effects associated with high-recombination strains are significant (p < 0.05).
+That IFD~raw~ and IFD~norm~ show similar patterns eliminates variation in SC lengths (and bivalent sizes) as the primary explanation.
+The main distinction in IFD~norm~ distributions between high-recombination and low-recombination strains is an enrichment of IFD~norm~ values under 30% in low-recombination strains (Figure 6).
+The frequency of IFDnorm values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
 These results indicate that evolution of the genome-wide recombination rate is accompanied by the evolution of crossover interference.
+
 In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not (consistently) to SC length and crossover position (on single-crossover bivalents).
 
-*old* -- In many mammalian species, the synaptonemal complex (SC) is longer in females [@cahoonLibuda2019, @sardell_sex_2020].
-Yet, the majority of these observations come from species with female-biased heterochiasmy.
-Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
-
-We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
-The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = r Q1_mus.av_shortSC_t.test.SKIVE$p.value).
-The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
-Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
-That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
-In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers compared to females.
+# B[Reserve for chapter 3?] Within-individual variation in the genome-wide recombination rate is higher in females
+
+round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)
+**MLH1var**  M1 - LmerMLH1_CV_results, M2 MLH1var_M2_sum_coE, M3  M4 glmMLH1Var_M3_sum.Coef
+
+Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
+To do this, we considered the same models as above, but replaced mean MLH1 focus count (for a mouse) with within-mouse variance in MLH1 focus count (for a mouse) as the dependent variable.
+Sex is the only variable that affects MLH1 focus count in all models (M1: p < 10^{-6} --`r LmerMLH1_CV_results$sex.results` ; M2: p = 0.03 -- `r` ; M3: p = 0.03 -- ``).
+
+In general, females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1).
+Since estimates of within-mouse variance may be more susceptible to technical error from staining, we repeated the analyses using a subset of cells with high quality scores (1 or 2; Materials and Methods).
+The results are similar: sex is the strongest effect (M1 p < 10^{-6}; M2 p = 2.310^{-4} -- ; M3 p = 2.2810^{-4} -- ).
+When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 focus count in either sex.
+These results show that sex is a primary determinant of inter-cellular variance in the genome-wide recombination rate across diverse genetic backgrounds.
+
+,
+ Q12 MLH1.var -- M1 -LmerMLH1_M1_Q12_VAR_results -- LmerMLH1_M1_Q12_VAR_results,  M2 -sum.MLH1var_M2_Q12.co_eff,  M3 -MLH1var_M3_Q12_sum
 
 # IFD double triangle
-
-# Figure 1
 
@@
 
-## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
-
-Next, we used the contrast between males in the high-recombination strains ( _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ ) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M5).
-
-Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-SC length is expected to be positively associated with genome-wide recombination rate because of the connection between DSB formation and the loop-axis structure [@zickler1999, @mercier2015].
-Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome with a finite length.
-Thus, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
-
-Following this logic we predict (1) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_, and 3) _domesticus_ strains will have similar SC length and crossover interference.
-
-## Conservation of Mean SC length
-
-## Single Crossover position is decoupled from genome wide recombination rate evolution
-
 ```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4}
@@
 Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.23 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
 
@@
                       aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-
 #+facet_wrap(~mouse, scales = "free")
@@
 ```
-
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = `r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively).
-While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.
-
-## Evolution of interference is associated with genome wide recombination rate evolution
 
@@
 
-# rewrite double triangle
+# Single rewrite IFD triangle
 
@@
 
-Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS$p.value`).
-
-This pattern is supported by slightly significant p values from comparisons within _musculus_ (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
-
-That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
-The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (_domesticus^G^_) to 16% (_musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
-
-Differnces in IFD can also be noted from the triangle IFD plots (Figure 6).
-The three high-recombining strains show more clustering of IFD observations near the area of the plot which indicates longer IFDs.
-The remaining strains have more observations short IFD observations near the diagonal.
-
-In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
-However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
-
 # Discussion
+
+```{r corMLH1.FM, echo=FALSE}
+
+#needs to be strain
+sperman.corr.FM <- cor( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
+                        method = c("spearman") )
+
+sperman.corr.FM.pval <- cor.test( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
+                        method = c("spearman") )
+```
 
@@
 Because inter-strain divergence times are identical for the two sexes, this observation demonstrates that the genome-wide recombination rate evolves faster in males, at least in house mice.
-More generally, recombination rate divergence is decoupled in males and females (Spearman’s correlation test comparing female and male strain mean MLH1 values for house mice; rho = 0.18; p = 0.60).
+More generally, recombination rate divergence is decoupled in males and females (Spearman’s correlation test comparing female and male strain mean MLH1 values for house mice; rho = 0.18 -- `r sperman.corr.FM.pval$estimate`; p = 0.60 -- `r sperman.corr.FM.pval$p.value` ).
 These disparities are remarkable given that recombination rates for the two sexes were measured in the same genomic backgrounds.
@@
 At the genetic level, the sex-specific evolution we documented indicates that some mutations responsible for divergence in recombination rate have dissimilar phenotypic effects in females and males.
-A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@kong2004, @kong2008, @kong2014, @halldorsson2019]), Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle, @shen2018_cattle] appear to show sex-specific properties, including antagonistic effects in females and males.
-Inter-sexual correlations are weak for recombination rate in humans [@fledel2011] and Soay sheep [@johnston2016_soay].
-Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011, @wang2017island].
+A subset of the genetic variants associated with genome-wide recombination rate within populations of humans (@Kong2004, @Kong2008, @Kong2014, @halldorsson2019), Soay sheep [@johnston2016_soay], and cattle (@ma2015_cattle, @Shen2018_cattle) appear to show sex-specific properties, including antagonistic effects in females and males.
+Inter-sexual correlations are weak for recombination rate in humans [@fledel2011] and Soay sheep (@johnston2016_soay).
+Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice (@dumont2011, @Wang2017island).
 These variants could differentially affect females and males at any step in the recombination pathway.
 Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contributed to sex-specific evolution in the recombination rate.
+
 Our results also raise the prospect that the connection between recombination rate and fitness differs between males and females.
-Little is known about whether and how natural selection shapes recombination in nature [@DapperPayseur2017, @ritz2017]).
+Little is known about whether and how natural selection shapes recombination in nature (@DapperPayseur2017, @Ritz2017).
 Using a quantitative genetic test, [@samuk2020]  recently discovered that genome-wide recombination rates in females from two populations of _Drosophila pseudoobscura_ diverged in a manner consistent with local adaptation.
 Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
-Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling, @sardell_sex_2020].
+Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon (@brandvain2012scrambling, @sardell_sex_2020).
 Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020].
@@
 Hence, selection seems likely to favor mutations that optimize the process of bi-orientation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+
 Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019], perhaps due to the presence of the centrosome spindle [@So2019] and higher cell volume [@kyogoku2017] in oocytes.
@@
 Therefore, faster male evolution of the genome-wide recombination rate could be driven by the stronger SAC acting on chromosome structures at the metaphase I alignment.
+
 Our SAC model is consistent with other features of our data.
@@
 We hypothesize that these sex differences hinge on disparate requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase-to-anaphase transition is initiated by the proteolytic decay of the sister cohesion proteins that connect homologs [@LaneKauppi2019, @subramanian2014, @dumontDesai2012).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
+The irreversible process of the metaphase-to-anaphase transition is initiated by the proteolytic decay of the sister cohesion proteins that connect homologs (@LaneKauppi2019, @subramanian2014, @dumontDesai2012).
+The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate (@vanVeen2003, @altendorfer2020).
 Differences in timing of the cell cycle between oogenesis and spermatogenesis could impose contrasting selective pressures on how sister cohesion affects chromosome structure.
@@
 Although the SAC model postulates direct fitness effects of interference, a modifier model predicts that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
+
 Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
 The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle @ma2015_cattle and differences between wild-born and laboratory-raised white-footed mice @peterson2019.
-In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@segura2013, @ottoPaysuer2019].
+In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates (@segura2013, @ottoPaysuer2019).
 Collectively, these patterns suggest that inferences about the evolutionary dynamics of interference depend on the timescale under consideration.
+
 Our findings further reveal that evolution of the genome-wide recombination rate does not require changes in the degree of chromatin compaction.
 Females consistently showed longer SCs, even in strains with more recombination in males.
-Studies in mice [@lynn2002, @petkov2007] and humans [@gruhn2013] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
-A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes @hulten2011_COM.
+Studies in mice (@lynn2002, @petkov2007) and humans [@gruhn2013] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes (@hulten2011_COM).
+
 Our conclusions are accompanied by several caveats.
-First, MLH1 foci only identify interfering crossovers (REF).
+First, MLH1 foci only identify interfering crossovers **(REF)**.
 Although most crossovers (XX%; REF) belong to this class, our approach likely underestimated genome-wide recombination rates.
@@
 Nevertheless, we expect our conclusion that recombination evolves in a sex-specific manner to be robust to geographic sampling because differences between females and males exist for the same set of inbred strains.
+
 The causes of sex differences in recombination remain mysterious @lenormand2016.

### 2020-05-25 fig3_double_tri.R

@@ -1,0 +1,82 @@
+#load data
+remove_strains.1sex <- c("CZECH", "TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
+
+#Curated_BivData.FM <- Curated_BivData[ ! Curated_BivData$strain %in% remove_strains.1sex, ]
+
+Curated_BivData
+
+my.data = (Curated_BivData %>% filter(hand.foci.count == 2) )
+#remove_strains
+
+#reorder dataframe
+my.data$strain <- factor(my.data$strain, ordered = TRUE, 
+              levels =c( "WSB", "G", "LEW", 
+              "PWD","SKIVE", "KAZ",
+              "MSM", "MOLF") )
+
+
+top.triangle <- function(my.data) {
+  #current.strain = 
+  for(i in 1:(length( unique(my.data$strain) ))){
+    print(i)  
+    current.strain <- as.character( unique(my.data$strain)[i] )
+
+    plot( (my.data %>% filter(strain == current.strain) %>% filter(sex == "female") )$PER_Foci_2 ~
+     (my.data %>% filter(strain == current.strain) %>% filter(sex == "female") )$PER_Foci_1, 
+         data= ( my.data %>% filter(sex == "female") %>% filter(strain == current.strain) ), 
+       axes=FALSE, ylab="", xlab="", main='', xlim=c(0, 1), ylim=c(0, 1), 
+       xaxt="n", yaxt="n", asp=1, col =  colors_of_strains[current.strain] )
+    
+    #mtext(current.strain, 3, 5, font=1, cex=1, adj=.95)
+  
+     # mtext("Female Position.2", 2, .75)
+  #  mtext("Female Position.1", 3, 2)
+    axis(side=2, las=1, pos=0)
+    axis(side=3, las=1, pos=1)
+    lines(0:1, 0:1)
+      }
+  }
+
+bottom.triangle <- function(my.data) {
+  x.dist <- .1
+  #my.data.2 <- transform(my.data, my.y.top=my.y.bottom + x.dist)
+
+  my.data.2 <- transform( ( my.data %>% filter(sex == "male")  ) , 
+                            PER_Foci_2=PER_Foci_2 + x.dist)
+  
+  for(i in 1:(length( unique(my.data.2$strain) ) ) ){
+    print(i)
[…]

### 2020-05-25 fig4_bottom_tri.R

@@ -1,0 +1,83 @@
+#load data
+remove_strains.1sex <- c("CZECH", "TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
+
+#Curated_BivData.FM <- Curated_BivData[ ! Curated_BivData$strain %in% remove_strains.1sex, ]
+
+Curated_BivData
+
+my.data = (Curated_BivData %>% filter(hand.foci.count == 2) )
+#remove_strains
+
+#reorder dataframe
+my.data$strain <- factor(my.data$strain, ordered = TRUE, 
+              levels =c( "WSB", "G", "LEW", 
+              "PWD","SKIVE", "KAZ",
+              "MSM", "MOLF") )
+
+
+top.triangle <- function(my.data) {
+  #current.strain = 
+  for(i in 1:(length( unique(my.data$strain) ))){
+    print(i)  
+    current.strain <- as.character( unique(my.data$strain)[i] )
+
+    plot( (my.data %>% filter(strain == current.strain) %>% filter(sex == "female") )$PER_Foci_2 ~
+     (my.data %>% filter(strain == current.strain) %>% filter(sex == "female") )$PER_Foci_1, 
+         data= ( my.data %>% filter(sex == "female") %>% filter(strain == current.strain) ), 
+       axes=FALSE, ylab="", xlab="", main='', xlim=c(0, 1), ylim=c(0, 1), 
+       xaxt="n", yaxt="n", asp=1, col =  colors_of_strains[current.strain] )
+    
+    #mtext(current.strain, 3, 5, font=1, cex=1, adj=.95)
+  
+     # mtext("Female Position.2", 2, .75)
+  #  mtext("Female Position.1", 3, 2)
+    axis(side=2, las=1, pos=0)
+    axis(side=3, las=1, pos=1)
+    lines(0:1, 0:1)
+      }
+  }
+
+bottom.triangle.only <- function(my.data) {
+  x.dist <- .1
+  #my.data.2 <- transform(my.data, my.y.top=my.y.bottom + x.dist)
+
+  my.data.2 <- transform( ( my.data %>% filter(sex == "male")  ) , 
+                            PER_Foci_2=PER_Foci_2 + x.dist)
+  
+  for(i in 1:(length( unique(my.data.2$strain) ) ) ){
+    print(i)
[…]

### 2020-05-25 src/Result_Setup.Rmd
updating things

@@
 
-female_mouse.avs_for.analysis <-  Mouse.Table_HQ %>% filter(sex== "female") %>% filter(species == "M.musculus") %>% filter(strain != "CZECH") %>% filter(strain != "CAST")
+female_mouse.avs_for.analysis <-  Mouse.Table_HQ %>% filter(sex== "female") %>% filter(species == "M.musculus")
 
@@
 
-```{r CurBiv.Mouse.av.TABLE, echo=FALSE, eval=FALSE}
-
-#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
-cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")
-
-#cleanCurated_BivData
-#removed the no hand foci count
-
-#there's alread a Mouse.Table_BivData --mouse.level.BivData_table
-
-mouse.BivData_avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
+```{r strain.curated.biv, echo=FALSE}
+
+strain.BivData.Table <- ddply(mouse.level.BivData_table, c("strain", "sex"), summarise,
+
+                        Nmice = length(Nmice),
+                        Ncells  = sum(Ncells), #might not to be sum
+                        mean.SC = mean(mean_SC),
+                        var.SC = mean(var), #mouse average variance of bivalent per mouse
+
+                        #chrm classes
+                        av.proportion_0CO = mean(proportion_0CO),
+                        av.proportion_1CO = mean(proportion_1CO),
+                        av.proportion_2CO = mean(proportion_2CO),
+                        av.proportion_3CO = mean(proportion_3CO),
+
+                        #1CO pos
+                        mean.1CO.pos = mean(mean_1CO.pos),
+
+                        #mean.IFD
+                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS),
+                        mean.IFD.PER = mean(mean_IFD.2CO_PER)
+                        )
+
+#add
+strain.BivData.Table <- add_subsp(strain.BivData.Table)
+strain.BivData.Table <- add_species(strain.BivData.Table)
+```
+
+```{r algorithm_performance_stats, echo=FALSE}
+
+#sort by category, 
+#compare total boxes and rate passed
+#Curated_BivData  -- pass rate?
+
+#number of mice, number of cells
+#mean number of isolate bivs per cell,
+#use these two metrics
+#cell.level.BivData_table$boxs.IDd 
+#cell.level.BivData_table$nboxes_passed
+
+#calq the mean rate by category
+
+cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
+
+algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
+                        n_obs = length( unique(fileName) ),
+                         mean_pass.rate = mean(pass.rate)
+                         )
+
+cell.level.BivData_table <- add_sex(cell.level.BivData_table)
+
+algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
+                        n_obs = length(unique(fileName)),
+                        mean_pass.rate = mean(pass.rate)
+                         )
+
+#showing the table..
+#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
+```
+
+RENAME the curate bivdata below
+
+```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
+
+#this is redundant
+
+male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(unique( fileName) ),
-                        nbiv = length(Obj.ID),
-
-                        mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
+                        Ncells  = length(Obj.ID),
+
+                      # mean_CO = mean()
+
+                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                         cV = cv(chromosomeLength),
@@
                         se   = round(sd / sqrt(Ncells), 3),
-
-              #can't figure out how to get this format to work  
-                       n0CO = sum(length( which(hand.foci.count == 0) )  ),
-                       n1CO = sum(length( which(hand.foci.count == 1) )  ),
-                      n2CO = sum(length( which(hand.foci.count == 2) )  ),
-                      n3CO = sum(length( which(hand.foci.count == 3) )  ),
-              #     
-                  proportion_0CO = (n0CO / nbiv),
-                  proportion_1CO = (n1CO / nbiv),
-                 proportion_2CO = (n2CO / nbiv),
-                     proportion_3CO = (n3CO / nbiv),
 
@@
 
-                #mean 1CO.pos
-                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
-                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
-                #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-                # IFD things
-            #  mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
-            # mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-mouse.avs_4MM <- add_sex(mouse.avs_4MM)
-mouse.avs_4MM <- add_strain(mouse.avs_4MM)
-mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
-mouse.avs_4MM <- add_category(mouse.avs_4MM)
-mouse.avs_4MM <- add_species(mouse.avs_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
-mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
-mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)
-
-#REMOVE JUVI mice
-juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice.list, ]  
-mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice.list, ]
-
-#assign high and low
-#divide musc males into groups
-mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
-                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
-                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))
-
-mouse.BivData.Table <- mouse.avs_4MM
-```
-
-```{r strain.curated.biv, echo=FALSE}
-
-strain.BivData.Table <- ddply(mouse.level.BivData_table, c("strain", "sex"), summarise,
-
-                        Nmice = length(Nmice),
-                        Ncells  = sum(Ncells), #might not to be sum
-                        mean.SC = mean(mean_SC),
-                        var.SC = mean(var), #mouse average variance of bivalent per mouse
-
-                        #chrm classes
-                        av.proportion_0CO = mean(proportion_0CO),
-                        av.proportion_1CO = mean(proportion_1CO),
-                        av.proportion_2CO = mean(proportion_2CO),
-                        av.proportion_3CO = mean(proportion_3CO),
-
-                        #1CO pos
-                        mean.1CO.pos = mean(mean_1CO.pos),
-
-                        #mean.IFD
-                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS),
-                        mean.IFD.PER = mean(mean_IFD.2CO_PER)
-                        )
-
-#add
-strain.BivData.Table <- add_subsp(strain.BivData.Table)
-strain.BivData.Table <- add_species(strain.BivData.Table)
-```
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#sort by category, 
-#compare total boxes and rate passed
-#Curated_BivData  -- pass rate?
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-#use these two metrics
-#cell.level.BivData_table$boxs.IDd 
-#cell.level.BivData_table$nboxes_passed
-
-#calq the mean rate by category
-
-cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
-
-algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
-                        n_obs = length( unique(fileName) ),
-                         mean_pass.rate = mean(pass.rate)
-                         )
-
-cell.level.BivData_table <- add_sex(cell.level.BivData_table)
-
-algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
-                        n_obs = length(unique(fileName)),
-                        mean_pass.rate = mean(pass.rate)
-                         )
-
-#showing the table..
-#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-RENAME the curate bivdata below
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#this is redundant
-
-male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
                 # IFD things
@@
 
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-#why is this block false? are the models in the MS.RMD file?
-
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
+MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus"))   )
+MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus")) )
+MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus")) )
 
@@
 
-```
-
-standardize the lm's and glms!
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=F}
-
-#not including this model for now / brivity
-#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
-
-MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
+#CV
+MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
+MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
+
+LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
+
+LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
+sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
+
+#Q12
+Q12_mouse_table_MLH1$subsp <- factor(Q12_mouse_table_MLH1$subsp, ordered = FALSE)
+Q12_mouse_table_MLH1$sex <- factor(Q12_mouse_table_MLH1$sex, ordered = FALSE)
+Q12_mouse_table_MLH1$strain <- factor(Q12_mouse_table_MLH1$strain, ordered = FALSE)
+Q12_mouse_table_MLH1$species <- factor(Q12_mouse_table_MLH1$species, ordered = FALSE)
+
+#USE Q12 (Q123) table
+#Q12_mouse_table_w.Ages
+#Q12 var
+MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
+MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
+MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
+
+LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
+
+LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
+sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
+
+#Q12 cv
+MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
+MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
+
+LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
+LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
+LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
+
+LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
+sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
+```
+
+```{r M2.MLH1.var, echo=FALSE, include=FALSE}
+
+MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data= ( DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus") )  )
 MLH1var_M2_sum <- summary(MLH1var_M2)
 
-#summary(MLH1var_M2)
-#anova(MLH1var_M2)  
-#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
+MLH1var_M2_sum_coE <- as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
 
 MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-
-#summary(MLH1cv_M3)
+#summary(MLH1cv_M2)
 #anova(MLH1cv_M2)
@@
 #Q12 var
-MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1var_M3_Q12)
-#anova(MLH1var_M2_Q12)  
-#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
+MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data= (Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
+sum.MLH1var_M2_Q12 <- summary(MLH1var_M2_Q12)
+sum.MLH1var_M2_Q12.co_eff <- as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
 
@@
 MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
 #summary(MLH1cv_M2_Q12)
@@
 
-```{r M3.MLH1.var, echo=FALSE, eval=F}
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
+```{r M3.MLH1.var, echo=FALSE}
+
+glmMLH1Var_M3 <- glm( var ~ sex * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus") ) )
+
 glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
+
+glmMLH1Var_M3_sum.Coef <- round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
 
@@
 ### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
+MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
 MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
@@
 
-# toal SC
+# toal SC setup
 
@@
 
-#Q2 section
-
-#Q2 SC
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-```
-
-<!-- plots -->
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-
+# Proportions
+
+```{r chrm_prop_data, echo=FALSE, warning=FALSE}
+#shut off eval for now because of wd / path error
+#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
+
+chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
+
+#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
+
+chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
+chrm.class.DF <- add_mouse(chrm.class.DF)
+chrm.class.DF <- add_strain(chrm.class.DF)
+chrm.class.DF <- add_sex(chrm.class.DF)
+chrm.class.DF <- add_category(chrm.class.DF)
+#chrm.class.DF <- add_euth_date(meta.data.file2)
+#chrm.class.DF <- add_age(meta.data.file2)
+
+#table(chrm.class.DF$category)
+#str(chrm.class.DF)
+
+#summarize by category
+proportion_stats_category  <- ddply(.data=chrm.class.DF, 
+                 .(category),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
+
+                 n0CO = sum(n0CO, na.rm = TRUE),
+                  n1CO = sum(n1CO,na.rm = TRUE),
+                  n2CO = sum(n2CO,na.rm = TRUE),
+                  n3CO = sum(n3CO,na.rm = TRUE),
+                  n4CO = sum(n4CO,na.rm = TRUE),
+
+                 per0CO = sum(n0CO) / tot_biv,                 
+                 per1CO= sum(n1CO)/ tot_biv,
+                 per2CO= sum(n2CO)/ tot_biv,
+                 per3CO= sum(n3CO)/ tot_biv,
+                 per4CO= sum(n4CO)/ tot_biv,
+
+                 mean.MLH1.foci = mean(MLH1.sum.check),
+                 mean.MLH1.foci2 = mean(nMLH1.foci),
+                 meanQ = mean(quality)
+
+)
+#this is the table that needs melting
+#but is also easy to redo by hand
+
+#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
+```
+
+```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
+
+#add subspecies to these values¬
+melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+
+#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
+melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
+
+#add the subsps
+melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
+                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
+
+melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
+
+chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))+theme(legend.position="none")
+
+#need a second DF / the first df
+#+ annotate() text 
+#annotate()text
+
+#proportion_stats_category,  tot_biv, ncells
+#add number of cells and bivalents in the corner
+
+#chrm.class.barplot
+```
+
+```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
+
+#prop.test(x, y)  x number of successes, y total  
+#strain sums of n
+#look up continuity correction
+
+#mouse.level.BivData_table <-- has more male strains
+#Mouse.Table_BivData_4MM <- only sex matched strains
+
+#perform test across strain 
+#removed all lines with no hand.foci.count entry
+Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
+
+                  nbiv = sum(tot.biv ),              
+                 n0CO = sum(n0CO ),
+                  n1CO = sum(n1CO ),
+                 n2CO = sum(n2CO ),
+                  n3CO = sum(n3CO ),
+                 n4CO = sum(n4CO )
+                          )
+
+Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
+
+for(g in 1:length(Chrm.prop_strain_table$nbiv)){
+
+  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
+                                     Chrm.prop_strain_table$n1CO[g],
+                                     Chrm.prop_strain_table$n2CO[g],
+                                     Chrm.prop_strain_table$n3CO[g] ))  
+}
+
+Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
+#there are still some wacky entries -- but much fewer than before
+
+chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
+
+female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
+female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
+female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
+#the pvalues for 0 are not sig dif
+#the p values for 1 and 2CO are sig diff
+
+#seperate by subsp
+chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
+chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
+chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
+
+Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
+Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
+Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
+Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
+#WSB has fewer 2COs than other 3 strains
+
+Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
+Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
+Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
+#KAZ has slightly more 1CO than the other two
+
+#high.female_G.LEW.
+chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
+
+high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
+high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
+high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
+
+#Low.female.G.LEW.
+chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
+
+low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
+low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
+low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
+
+#all females
+chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
+male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
+
+male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
+male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
+#the pvals for these 3 class are sig dif
+
+#seperate by subsp
+chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
+chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
+chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
+
+Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
+Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
+Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
+#non of the Dom proportions are different
+
+Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
+Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
+Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
+#all v different
+
+Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
+Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
+Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
+
+#high vs low
+chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
+                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
+                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
+
+chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
+chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
+
+chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
+chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
+
+Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
+Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
+Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
+#the low strains do vary slightly in their 1CO and 2CO proportions
+
+#includes SKIVE
+Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
+Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
+Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
+#all very sig p values
+
+High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
+High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
+High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
+#these proportions are v different
+
+#no SKIVE
+High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
+High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
+High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
+#MSM and PWD only slightly sig dif 2CO
+
+#MSM-MOLF
+MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
+
+#PWD-KAZ
+PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
+
+#PWD-MSM
+PWD.MSM.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "MSM") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "MSM") )$nbiv )
+
+```
+
+# Q1 CO pos
+
+```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
+#move to setup
+
+Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
+
+```
+
+```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
+#main figure of density plots,
+MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
+MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
+
+#clean up all the non 1 to 4 hand.foci counts
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
+
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
+
+# remove the na's
+MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
+
+CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
+
+#category specific plots, don't run for now.
+Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
+#Musc male
+Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
+
+CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
+#Dom male
+Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
+
+CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
+```
+
+```{r Q1_1COpos.t.test, echo=FALSE}
+#testing,
+#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
+
+Q1_1CO.pos_t.test.FULL <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos)
+
+#WSB
+Q1_1CO.pos_t.test.WSB <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
+#G
+Q1_1CO.pos_t.test.G <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
+#LEW
+Q1_1CO.pos_t.test.LEW <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
+
+#PWD
+Q1_1CO.pos_t.test.PWD <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
+
+#KAZ
+Q1_1CO.pos_t.test.KAZ <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
+
+#MSM
+Q1_1CO.pos_t.test.MSM <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
+
+#Single Biv Level
+#single.biv.1CO_pos_t.test <- t.test( 
+ # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
+#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
+#  )
+
+#t.tests for mous.av_nrm.1CO.pos 
+```
+
+```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
+
+#make sure this dataset has all the right strains
+Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
+
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
+
+lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
+lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
+
+lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
+lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
+
+lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
+sex.results = lmer.Q1.nrm1CO_M1.sex , 
+               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
+
+#lmer.Q1.nrm1CO_M1_results
+
+#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+
+```
+
+```{r Q1.CO1pos_glm.M2, echo=FALSE}
+
+#in case factors aren't ordered
+#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
+#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
+#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
+
+#summary(Q1.glm_F1pos_M2)
+
+summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
+#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
+
+#DOM
+Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
+#summary(Q1.glm_nrmF1pos_M2_Dom)
+
+#MUSC
+Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
+#summary(Q1.glm_nrmF1pos_M2_Musc)
+
+#MOL
+Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
+#summary(Q1.glm_nrmF1pos_M2_Mol)
+
+```
+
+```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
+
+Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
+
+#summary(Q1.glm_1CO.pos_M3)
+
+summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
+#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
+#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+
+```
+
+# Q1 IFD
+
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized IFD")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
+
+#each point is a 2CO, with the F1 and F2 position 
+#based on plot from Dourad et al (Mercier)
+
+P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
+    aes(y = Foci2, x= Foci1, color=strain))+
+  geom_jitter(alpha=.5)+
+  ggtitle("Raw IFD WSB")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#trying to move x axis to top
+#cowplot might be the easier way -- since kinda diffecult in ggplot
+#or ggdraw
+
+P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+
+#+theme(legend.position="none")
+```
+
+```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
+#below are calq just from 2COs, have no fear
+#mean_IFD.2CO_ABS
+#mean_IFD.2CO_PER
+
+#FULL all strains
+Q1.IFD.raw_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" ) %>% filter(species == "M.musculus")  )$mean_IFD.2CO_ABS,
+
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
+#p-value = 0.07
+
+Q1.IFD.PER_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>% filter(species == "M.musculus")    )$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER)
+#p-value = 1e-12
+
+#WSB
+Q1.IFD.raw_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
+                                (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
+
+#LEW
+Q1.IFD.raw_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
+
+#G
+Q1.IFD.raw_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
+
+#PWD
+Q1.IFD.raw_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
+
+#KAZ
+Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
+
+#SKIVE
+#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+
+#MSM
+Q1.IFD.raw_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
+
+Q1.IFD.PER_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
+
+#FULL all strains
+Q1.IFD.raw_sansG_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
+#p-value = 0.3
+```
+
+```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
+
+#these are the scatter plots
+
+nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
+#looked up image, measures confirmed
+
+####
+nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#21may18_KAZ_m2 -- seems like it might be an outlier
+#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
+#second foci slightly smaller, kinda fainter -- but also much brighter than background
+
+#check all KAZ males
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ.female_BIV.data
+
+nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD.female_BIV.data
+
+nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_LEW.female_BIV.data
+
+nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G.female_BIV.data
+
+nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
+
+raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+###look at why KAZ males are weird
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#the difference the the Nrm IFD cut off is striking with these plots
+nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
+
+```
+
+```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
+
+#M1_IFD_2CO_ABS
+lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
+               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
+
+sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
+
+Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
+               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
+
+```
+
+```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
+#mouse.avs_4MM -- is also the bivbased mouse average table 
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM  %>% filter(species == "M.musculus")  )
+summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
+
+glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.level.BivData_table  %>% filter(species == "M.musculus")  )
+summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
+
+#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= (mouse.level.BivData_table %>% filter(species == "M.musculus") ) )
+summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
+```
+
+```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
+
+glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
+
+#PER
+glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
+summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
+#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+```
+
+# Q1 total SC
+
+```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
+#total-SC
+#Q1.lmer_totSC.mouse.av.table -- #can't find this table
+#Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>% 
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
+mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
+mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
+
+#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data= (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") )  )
+sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") ) )
+sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
+#LOT OF SIGNIFICant strain effects in this model
+```
+
+```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
+
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
+
+#summary(Q1.glm_short.SC_M2)
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+```
+
+```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
+
+#summary(Q1.glm_short.SC_M3)
+
+summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
+
+#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#sex male very sig -- only sig
+```
+
+```{r Q1.SC.Q12, echo=FALSE}
+
+#make comparisons for the high quality cells
+#get Q12 cells from MLH1.merge.Skel
+#calq the mouse averages
+
+Total.SC_DF$quality <- as.numeric(Total.SC_DF$quality)
+
+totSC_Q12_mouse.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
+
+#unorder the factors
+totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
+totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
+totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
+totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
+
+#REMOVE JUVI mice
+tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
+totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
+
+totSC_Q12_strain.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
+                 .(strain, sex),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
+
+#unorder the factors
+totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
+totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
+totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
+totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
+
+```
+
+```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
+
+Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Dom
+#Tot.SC_mouse.av_Dom
+
+Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Musc
+#Tot.SC_mouse.av_Musc
+
+Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+```
+
+```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
+
+#Q1.real.long_mouse_table
+
+Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#not enough y observations
+#SKIVE
+#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
+#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
+
+#MSM
+#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+```
+
+```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
+
+#removing these plots for space
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
+#moved this to setupdata
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+#Long.biv_mouse_table
+
+#make the Q1 strain list
+Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+```
+
+```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
 Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
@@
 
+#nested boxplots, pooled Bivalents
+Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+
+```
+
+#Q2 section
+
+#Q2 SC
+
+```{r Q2.SC.short.long.setup, echo=FALSE}
+
+#table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+
+Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                         #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+                         )
+
+Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
+Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
+
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
+                                           0)))
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                        #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
+                         ) )
+
+Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
+
+Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
+Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
+
+Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
+                                           0)))
+
+```
+
+<!-- plots -->
+
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
 Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
@@
 #Q2mouse.av.short.biv.table
+
 Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
@@
 
-Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = F)
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = F)
+
+Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
 sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M4)$coefficients
@@
 
-Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
 sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
 #round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
-
-mouse.total.skel <- add_strain(mouse.total.skel)
-mouse.total.skel <- add_sex(mouse.total.skel)
-mouse.total.skel <- add_subsp(mouse.total.skel)
-#within subspecies values
 
@@
 
-```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE, eval=FALSE}
+```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
 
 #High = MSM, PWD, SKIVE
-Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == 0))$norm.first.foci.pos,    
-        (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == 1))$norm.first.foci.pos )
+Q2.1COpos_Ttest <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(Rec.group == 0))$norm.first.foci.pos,    
+        (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(Rec.group == 1))$norm.first.foci.pos )
 
@@
 
-```{r Q2.1COpos_within.subsp_ttest, echo=FALSE, eval=FALSE}
-#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..
-
-#mouse.av_4MM -- n
-
+```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}
 #Musc
-Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == 0))$norm.first.foci.pos,
-
-            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == 1))$norm.first.foci.pos )
+Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.level.BivData_table %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == 0))$norm.first.foci.pos,
+
+            (mouse.level.BivData_table  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == 1))$norm.first.foci.pos )
 # p-value = 0.4
@@
 #Mol
-Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
-
-              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
+Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.level.BivData_table %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,
+
+              (mouse.level.BivData_table  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
 # p-value = 0.07
@@
 
+- mouse.level.BivData, t-tests
+
 ```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE, eval=FALSE}
-#these don't have mean.first.foci
+#these t.tests are causing issues not sure if they are used in final MS
 
 #short -- real.long.bivData -- REAL.short_mouse_table
-Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0" ))$mean_F1pos_ABS,
-
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == 1))$mean_F1pos_ABS   )
+Q2.1COposABS_short_ttest_HL <- t.test( (Short.biv_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0" ) %>% filter(species == "M.musculus") )$ABS.first.foci.pos,
+
+      (Short.biv_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == 1) %>% filter(species == "M.musculus") )$ABS.first.foci.pos   )
 #p-value = 0.3
 
-Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
-      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
+Q2.1COpos.PER_short_ttest_HL <- t.test( (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(species == "M.musculus") )$norm.first.foci.pos, 
+      (Short.biv_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1") %>% filter(species == "M.musculus") )$norm.first.foci.pos   )
 # p-value = 0.1
 
-#remove SKIVE??
 ```
@@
 
+- COpos short long t.tests
+
 ```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
@@
 #mouse.av - the mouse average IFD is from just 2CO
+
+#high, low
 t.test_mouse.av.IFD_ABS <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS, 
@@
 #ABS
-
 t.test_mouse.av.IFD_ABS_MUSC <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(species == "M.musculus"))$mean_IFD.2CO_ABS,

### 2020-05-25 src/SetupData.Rmd
updating things

@@
 #remove bad ages
-#this is removing female
+#this is removing female MOLF?
 
@@
 Q12_mouse_table_MLH1 <- add_category(Q12_mouse_table_MLH1)
+Q12_mouse_table_MLH1 <- add_species(Q12_mouse_table_MLH1)
+
+#main df with variables coded as chr
+Q12_mouse_table_MLH1$subsp <- factor(Q12_mouse_table_MLH1$subsp, ordered = FALSE)
+Q12_mouse_table_MLH1$sex <- factor(Q12_mouse_table_MLH1$sex, ordered = FALSE)
+Q12_mouse_table_MLH1$strain <- factor(Q12_mouse_table_MLH1$strain, ordered = FALSE)
+Q12_mouse_table_MLH1$species <- factor(Q12_mouse_table_MLH1$species, ordered = FALSE)
 
@@
                        cv.SC = cv(chromosomeLength),
-
-#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
                       mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),
@@
                      mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-
                        var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
@@
 
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
                  mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
@@
                        #CO position things, requires the melt?
+              ABS.first.foci.pos =  mean(Foci1[which(hand.foci.count == 1)]),
                  norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+              #norm first foci is mis normer
 
@@
                        #CO position things, requires the melt?
+               ABS.first.foci.pos =  mean(Foci1[which(hand.foci.count == 1)]), 
               norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
@@
 Short.biv_mouse_table <- add_subsp(Short.biv_mouse_table)
+Short.biv_mouse_table <- add_species(Short.biv_mouse_table)
 Short.biv_mouse_table <- add_category(Short.biv_mouse_table)

### 2020-05-31 src/Result_Setup.Rmd
realized total SC dataset missing some batches

@@
 
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")
+load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.30.20.RData")#added SKIVE_f BiveData
+#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")
 #load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
@@
 
-# toal SC setup
-
-```{r totSC_setup, echo=FALSE, eval = F}
-
-#Total.SC_DF is the main dataframe and should be set up cleanly in the SetUpData.Rmd
-#SC.skel.OG <- SC.skel #5683
-
-tot.sc.x.list <- Total.SC_DF %>% filter(X == "X")
-
-Total.SC_DF.og <- Total.SC_DF[! Total.SC_DF$Random.Name %in% tot.sc.x.list$Random.Name, ]
-#quality is still a factor
-str(Total.SC_DF.og)
-Total.SC_DF.og$quality <- as.numeric(Total.SC_DF.og$quality)
-
-Total.SC_DF.og <- Total.SC_DF.og %>% filter(quality < 5)
-
-OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
-
-#check for DUPs, #check for duplicatations of the random name
-MLH1.merge.Skel$fileName  <- as.character(MLH1.merge.Skel$fileName)
-tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
-
-MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
-
-MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
-MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
-#table(MLH1.merge.Skel$category)
-
-#remove outliers
-#MLH1.merge.Skel <- MLH1.merge.Skel  %>% filter(bin_size < 60000)
-
-```
-
--- The total_SC outliers should be removed in the total.SC RData script
-
-```{r totSC.viz4outliers, echo=FALSE}
-
-#Total.SC_DF
-#mouse.total.skel
-
-# -- EXAMINE the ggplot scatter plots for outliers
-show.tot.sc_plot <- ggplot(data= ( Total.SC_DF %>% filter(sex == "male") ), aes(y = skel_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
-```
-
-```{r tot.skel.table, echo=FALSE, eval=FALSE}
-#move to setup file
-
-#this is redundant, setup data made -- mouse.total.skel
-
-#make sure the table is clean -- no 5
-
-MLH1.merge.Skel$nMLH1.foci <- as.numeric(MLH1.merge.Skel$nMLH1.foci)
-
-totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-
-                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
-                 var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                 var.skel = var(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
-totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
-
-#unorder the factors
-totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
-totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
-totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
-totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
-
-OG.totSC.mouse.av.table <- totSC.mouse.av.table
-
-#REMOVE JUVI mice
-tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
-
-#
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-```
-
-```{r tot.SC_strain.setup, echo = FALSE, eval=FALSE}
-
-#not sure if this is used
-
-totSC.strain.av.table <- ddply(.data=mouse.total.skel, 
-                 .(sex,strain),
-                 summarize, 
-                 ncells = sum(ncells),
-                 nmice = length(unique(mouse)),
-
-                av.MLH1 = mean( mean.MLH1, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(mean.skel, na.rm = TRUE),
-               # var.skel = var(mean.skel, na.rm = TRUE), #average mouse variance
-                 mean.bin =  mean(mean.bin, na.rm = TRUE)
-
-)
-
-totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
-
-```
-
-# Q1 single bivalent section
-
-```{r Q1.single.biv_position_table, echo=FALSE, eval=FALSE}
-#strain averages for single bivalent 
-#
-#colnames(Curated_BivData99)
-
-```
+# Q1 single bivalent section and total SC
 
@@
 #Long.biv_mouse_table_HetC <- #Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | #Long.biv_mouse_table_HetC$mouse==""), ]
+
+```
+
+```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
+#moved this to setupdata
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+#Long.biv_mouse_table
+
+#make the Q1 strain list
+Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
+
+```
+
+# toal SC setup
+
+mouse total sc t.tests male female (Q1)
+
+```{r Q1.total.SC_ttests, echo=FALSE}
+
+#mouse.total.skel
+
+#full group
+Q1.tot.SC_t.test_FM <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean.skel,
+                    (mouse.total.skel %>% filter(sex == "female") %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ))$mean.skel  ) #<2e-16
+
+#WSB
+Q1.tot.SC_t.test_WSB <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean.skel)
+
+#LEW
+Q1.tot.SC_t.test_LEW <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean.skel)
+
+#G
+Q1.tot.SC_t.test_G <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="G" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="G"))$mean.skel)
+
+#PWD
+Q1.tot.SC_t.test_PWD <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean.skel)
+
+#KAZ
+Q1.tot.SC_t.test_KAZ <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean.skel)
+
+#SKIVE
+Q1.tot.SC_t.test_SKIVE <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean.skel)
+
+#MSM
+Q1.tot.SC_t.test_MSM <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean.skel,
+                                (mouse.total.skel %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean.skel)
+
+```
+
+long bivs shouldn't be compared for Q1
+
+```{r Q1_SC.biv.short_ttest, echo=FALSE, include=TRUE}
+
+#Q1.real.long_mouse_table -- change to Short.biv_mouse_table
+
+#all strains -#full group
+Q1_short.SC_t.test.FM <- t.test( ( Short.biv_mouse_table %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_SC,
+                    (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ))$mean_SC  ) #<2e-16
+
+# WSB p-value = 0.02 (highest p val)
+Q1_short.SC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G  p-value = 5e-04
+Q1_short.SC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW  p-value = 0.002
+Q1_short.SC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD p-value = 0.008
+Q1_short.SC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ p-value = 0.06
+Q1_short.SC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#not enough y observations
+#SKIVE -- not significant (kinda few )
+Q1_short.SC_t.test.SKIVE <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+                   (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC)
+
+#MSM -- no enough male measures
+#Q1_short.SC_t.test.MSM <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC,
+#                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC)
+
+#run t.tests for the mouse av. for reduced single bivalent datasets
+```
+
+```{r Q1_SC.biv.short_ttest99, echo=FALSE, eval=FALSE}
+#this show is re-dundant
+#
+#run t.tests for the mouse av. for reduced single bivalent datasets
+
+#WSB
+Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
+
+#G
+Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
+
+#LEW
+Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
+
+#PWD
+Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
+
+#KAZ
+Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
+                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
+
+#SKIVE - this value used in the text
+#only one SKVIE female, this is breaking the test
+#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
+
+#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
+#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
+#      )
+
+```
+
+```{r Q1SC_short_ratios, echo=FALSE, eval=TRUE}
+
+#calq the different in short.biv means across sexes
+#Long.biv_mouse_table
+
+#WSB
+Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
+
+#LEW
+Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
+
+#G
+#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
+
+#PWD
+Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
+
+#KAZ
+Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
+
+#SKIVE
+Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
+
+#MSM
+Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
+
+#this bivData ratio is the same as mouse average
+Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
+
+```
+
+```{r totSC_setup, echo=FALSE, eval = F}
+
+#Total.SC_DF is the main dataframe and should be set up cleanly in the SetUpData.Rmd
+#SC.skel.OG <- SC.skel #5683
+
+tot.sc.x.list <- Total.SC_DF %>% filter(X == "X")
+
+Total.SC_DF.og <- Total.SC_DF[! Total.SC_DF$Random.Name %in% tot.sc.x.list$Random.Name, ]
+#quality is still a factor
+str(Total.SC_DF.og)
+Total.SC_DF.og$quality <- as.numeric(Total.SC_DF.og$quality)
+
+Total.SC_DF.og <- Total.SC_DF.og %>% filter(quality < 5)
+
+OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
+
+#check for DUPs, #check for duplicatations of the random name
+MLH1.merge.Skel$fileName  <- as.character(MLH1.merge.Skel$fileName)
+tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
+
+MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
+
+MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
+MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
+#table(MLH1.merge.Skel$category)
+
+```
+
+-- The total_SC outliers should be removed in the total.SC RData script
+
+```{r totSC.viz4outliers, echo=FALSE, eval=FALSE}
+
+#Total.SC_DF
+#mouse.total.skel
+
+# -- EXAMINE the ggplot scatter plots for outliers
+show.tot.sc_plot <- ggplot(data= ( Total.SC_DF %>% filter(sex == "male") ), aes(y = skel_size, x=strain))+ geom_jitter()+facet_wrap(~category, scales = "free")
+
+show.mus.tot.sc_plot <- ggplot(data= ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") ), aes(y = mean.skel, x=strain, color = strain))+ geom_jitter()+ scale_color_manual(values=colors_of_strains)+ facet_wrap(~subsp, scales = "free")
+
+```
+
+```{r tot.skel.table, echo=FALSE, eval=FALSE}
+#move to setup file
+#this is redundant, setup data made -- mouse.total.skel
+#make sure the table is clean -- no 5
+
+MLH1.merge.Skel$nMLH1.foci <- as.numeric(MLH1.merge.Skel$nMLH1.foci)
+
+totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+
+                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
+                 var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                 var.skel = var(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
+totSC.mouse.av.table <- add_species(totSC.mouse.av.table)
+
+#unorder the factors
+totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
+totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
+totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)
+totSC.mouse.av.table$species <- factor(totSC.mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))
+
+OG.totSC.mouse.av.table <- totSC.mouse.av.table
+
+#REMOVE JUVI mice
+tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+#
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+
+```
+
+```{r tot.SC_strain.setup, echo = FALSE, eval=FALSE}
+
+#not sure if this is used
+
+totSC.strain.av.table <- ddply(.data=MLH1.merge.Skel, 
+                 .(sex,strain),
+                 summarize, 
+                 ncells = sum(ncells),
+                 nmice = length(unique(mouse)),
+
+                av.MLH1 = mean( mean.MLH1, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(mean.skel, na.rm = TRUE),
+               # var.skel = var(mean.skel, na.rm = TRUE), #average mouse variance
+                 mean.bin =  mean(mean.bin, na.rm = TRUE)
+
+)
+
+totSC.strain.av.table <- add_subsp(totSC.strain.av.table)
+
+```
+
+## total sc models
+
+where is total_SC M1?
+
+```{r Q1.total.SC_M1, echo=FALSE}
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
+mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
+mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
+
+#mouse.total.skel
+Q1.lmer_totSC_M1 <- lmer(mean.skel ~ subsp * sex + (1|strain), data= ( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+
+Q1.lmer_totSC_M1_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ))
+
+Q1.lmer_totSC_M1_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ))
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_totSC_M1_sex <- anova(Q1.lmer_totSC_M1, Q1.lmer_totSC_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_totSC_M1_subsp <- anova(Q1.lmer_totSC_M1, Q1.lmer_totSC_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_totSC_M1_int <- drop1(Q1.lmer_totSC_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_totSC_M1_strain <- exactRLRT(Q1.lmer_totSC_M1)$p.value
+
+Q1.lmer_totSC_M1_results <- data.frame(sub.sp = Q1.lmer_totSC_M1_subsp, 
+sex.results =  Q1.lmer_totSC_M1_sex, 
+               interaction.pval = Q1.lmer_totSC_M1_int,
+rand.pval = Q1.lmer_totSC_M1_strain)
+
+```
+
+```{r Q1.glm_totalSC_M2, echo=FALSE, message=FALSE}
+#mouse.total.skel
+
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
+mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
+mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
+
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=  (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+
+sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
+
+#sex most sig
+#subsp
+
+#straing G, LEW
+#mol * sex
+#G * sex
+```
+
+```{r Q1.glm_totalSC_M3, echo=FALSE, message=FALSE}
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+
+sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
+#sex most significant, 
+#PWD more sig
+
+#G, LEW, MSM, SKIVE, KAZ,  
+#G*male, MSM*male
+
+#LOT OF SIGNIFICant strain effects in this model
+
+```
+
+```{r Q1_single.Biv_M1.lmer, echo=FALSE, include=FALSE, eval=FALSE}
+
+#1CO  lmer.Q1_SCLength_M1
+Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
+
+Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
+sex.results =  Q1.lmer_SCLength.1CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
+Q1.lmer_SCLength.1CO_strain)
+
+#2CO
+Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
+Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
+Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
+
+#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
+
+Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
+Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
+
+Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
+sex.results =  Q1.lmer_SCLength.2CO_sex, 
+               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
+Q1.lmer_SCLength.2CO_strain)
+#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
+```
+
+```{r Q1_short.M1, echo=FALSE}
+
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+Short.biv_mouse_table$species <- factor(Short.biv_mouse_table$species, ordered = FALSE)
+
+Q1.lmer_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+
+Q1.lmer_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data= ( Short.biv_mouse_table %>%  filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+
+Q1.lmer_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+
+Q1.lmer_short_M1_sex <- anova(Q1.lmer_short_M1, Q1.lmer_short_M1_R.sex)$`Pr(>Chisq)`[2]
+
+Q1.lmer_short_M1_subsp <- anova(Q1.lmer_short_M1, Q1.lmer_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+
+Q1.lmer_short_M1_int <- drop1(Q1.lmer_short_M1, test="Chisq")$`Pr(Chi)`[2]
+
+Q1.lmer_short_M1_strain <- exactRLRT(Q1.lmer_short_M1)$p.value
+
+Q1.lmer_short_M1_results <- data.frame(sub.sp = Q1.lmer_short_M1_subsp, 
+sex.results =  Q1.lmer_short_M1_sex, 
+               interaction.pval = Q1.lmer_short_M1_int,
+rand.pval = Q1.lmer_short_M1_strain)
+
+```
+
+```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
+
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+
+summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
+#summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,]
+
+#sex most sig
+#skive, slightly sig
+```
+
+```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+
+#summary(Q1.glm_short.SC_M3)
+
+summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
+
+#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
+#sex male very sig -- only sig
+#sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,]
+
+#SKIVE * male
+```
+
+```{r Q1.SC.M2plot, echo=FALSE, include=FALSE, eval=FALSE}
+
+#removing these plots for space
+
+Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
+#Musc.all
+Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
+
+#BivData Level
+Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
+
+Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
+SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+#BivData Level
+Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
+
+Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
+
+```
+
+```{r Q1.SC.Q12, echo=FALSE}
+
+#make comparisons for the high quality cells
+#get Q12 cells from MLH1.merge.Skel
+#calq the mouse averages
+
+Total.SC_DF$quality <- as.numeric(Total.SC_DF$quality)
+
+totSC_Q12_mouse.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
+                 .(mouse),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+
+totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
+totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
+
+#unorder the factors
+totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
+totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
+totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
+totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
+
+#add rec group
+totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
+                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
+                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
+
+#REMOVE JUVI mice
+tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
+totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
+
+totSC_Q12_strain.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
+                 .(strain, sex),
+                 summarize, 
+                 ncells = length(unique(Original.Name)),
+                 nmice = length(unique(mouse)),
+                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
+                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
+
+                 mean.skel = mean(skel_size, na.rm = TRUE),
+                #se for skel
+
+                 mean.bin =  mean(bin_size, na.rm = TRUE)
+
+)
+#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
+#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
+totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
+
+#unorder the factors
+totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
+totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
+totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
+totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
+
+```
+
+```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
+
+Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Dom
+#Tot.SC_mouse.av_Dom
+
+Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+
+#Q12.Tot.SC_mouse.av_Musc
+#Tot.SC_mouse.av_Musc
+
+Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(~sex, scales="free", ncol = 3)+
+
+  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
+
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
+```
+
+```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
+#switch the cleanCurated_BivData
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+#nested boxplots, pooled Bivalents
+Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
 
@@
 )
-#this is the table that needs melting
-#but is also easy to redo by hand
 
@@
 
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos,
-                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos)
+Q1_1CO.pos_t.test.FULL <- t.test( ( mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$norm.first.foci.pos,
+(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$norm.first.foci.pos ) #<2e-16
 
@@
                            (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
+
 #G
@@
 
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test( 
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
 ```
@@
 
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
+lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), 
+                          data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
 
@@
 #in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
+
+Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
 
@@
 
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
+Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ))
 
@@
 
-```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Q1.IFD_Ttests_ABS, echo=FALSE, warning=FALSE, message=FALSE}
 #below are calq just from 2COs, have no fear
@@
 #FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" ) %>% filter(species == "M.musculus")  )$mean_IFD.2CO_ABS,
-
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>% filter(species == "M.musculus")    )$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
+Q1.IFD.ABS_FULL_t.test <- t.test(   (mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_ABS,
+
+(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_ABS    ) #0.03 (female larger)
+#this comparison should be within subspecies due to strain, SC differences
 
@@
 
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
 #LEW
@@
 
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
 #G
@@
 
-Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
 #PWD
@@
 
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
+#KAZ - low KAZ male curation (3 mice)
 Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
@@
 
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
 #SKIVE
-#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
+Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+#MSM - 3 male 4 female measures
+#only 3 MSM males
+#could be missing MSM measures, (4 females, 3 males)
 Q1.IFD.raw_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
@@
 
+#FULL all strains
+Q1.IFD.raw_all <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
+#p-value = 0.3
+```
+
+```{r Q1.IFD_ttest_PER, echo=FALSE}
+
+Q1.IFD.PER_FULL_t.test <- t.test(   (mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_PER,
+
+(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_PER    ) #p-value = 1e-10
+
+#WSB
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER )#p-value = 0.03
+
+#LEW
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)#p-value = 7e-05
+
+#G
+Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER )#p-value = 0.03
+
+#PWD
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)#p-value = 9e-04
+
+#KAZ
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)#p-value = 0.3
+
+#MSM
 Q1.IFD.PER_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
@@
 
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
+#SKIVE
+Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+
 ```
@@
 
-# Q1 total SC
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table -- #can't find this table
-#Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>% 
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
-mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
-mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
-mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
-
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data= (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") )  )
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") ) )
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
-Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
-Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-```{r Q1.SC.Q12, echo=FALSE}
-
-#make comparisons for the high quality cells
-#get Q12 cells from MLH1.merge.Skel
-#calq the mouse averages
-
-Total.SC_DF$quality <- as.numeric(Total.SC_DF$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
-
-#unorder the factors
-totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
-totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
-totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
-totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
-
-#REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
-
-totSC_Q12_strain.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
-                 .(strain, sex),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
-
-#unorder the factors
-totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
-totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
-totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
-totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
-
-```
-
-```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
-
-Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Dom
-#Tot.SC_mouse.av_Dom
-
-Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Musc
-#Tot.SC_mouse.av_Musc
-
-Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#not enough y observations
-#SKIVE
-#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
-#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
-#moved this to setupdata
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-#Long.biv_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
+#Q2 section
+
+#Q2 SC
+
+```{r Q2.SC.short.long.setup, echo=FALSE}
+
+#table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
+cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
+
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
+
+Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                         #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+                         )
+
+Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
+Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
+
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
+                                           0)))
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                        #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
+                         ) )
+
+Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
+
+Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
+Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
+
+Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
+                                           0)))
+
+```
+
+<!-- plots -->
+
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
+
 Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
@@
 
-#nested boxplots, pooled Bivalents
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
-#Q2 section
-
-#Q2 SC
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-```
-
-<!-- plots -->
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
 Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
@@
 #mol -- why is there only 1 MOLF male now?
-#Q2tot.sc_t.test_High.Low_Mol <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% #filter(subsp == "Mol") )$mean.skel)
+#there are 3 MOLF males in bivdata
+Q2tot.sc_t.test_High.Low_Mol <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
 
@@
 
-Curated_BivData.Mol <- Curated_BivData %>% filter(Foci1 != "")
-
 #making new models
@@
 
-chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= ( Curated_BivData %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") ) )
 
@@
 
-chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData.Mol %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
+chrm.1CO.pos.chrmLength_strain <- glm(PER_Foci_1 ~ chromosomeLength *  strain, data= Curated_BivData %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") )
 
@@
 #save Rdata
-save.image("~./MLH1repo/data/Results_setup_5.21.20.RData")
-
-```
+save.image("~./MLH1repo/data/Results_setup_5.30.20.RData")
+
+```

### 2020-05-31 src/SetupData.Rmd
realized total SC dataset missing some batches

@@
 library(reshape2)
+
 #knitr::opts_chunk$set(echo = FALSE)
@@
 #total SC 
-Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
-
-Curated_BivData <-  read.csv("~./MLH1repo/data/Curated_BivData.csv", header=TRUE)
+#new total_SC == will have all of the Batches
+
+Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF_5.31.20.xlsx", header=TRUE, 1) #this new total_SC has the additional batches and outliers have been removed
+
+#Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
+
+#original Curate_bivata + missing SKIVE female
+Curated_BivData <-  read.csv2("~./MLH1repo/data/Curated_BivData_5.30.20.csv", sep=" ", header=TRUE)
+#why they fuck was this space deliminated? I hate excel uggg
 
 source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )
+```
+
+```{r missing.curated.data, echo=FALSE, include=FALSE, eval=FALSE}
+
+#Curated.missing.SKIVE.female = read.xlsx2("Curated_missing.SKIVE.female.xlsx", header=TRUE, 1)
+Curated.missing.SKIVE.female$Foci4 <- ''
+Curated.missing.SKIVE.female$PER_Foci_4 <- ''
+
+Curated.missing.SKIVE.female$dis.cent <- ''
+Curated.missing.SKIVE.female$dis.cent.PER <- ''
+
+Curated_BivData_new.skive <- rbind(Curated_BivData, Curated.missing.SKIVE.female)
+
+#write.xlsx2(Curated_BivData_new.skive, "Curated_BivData_5.30.20.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
 ```
@@
 
-```{r FULL.mouse.BivData.table, echo=FALSE}
-
+```{r remove.non.foci, echo=FALSE}
 #length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36
-
+str(Curated_BivData)
 table(Curated_BivData$hand.foci.count, exclude = NULL)
-#there are ~600 rows with SC and no hand foci count (blah)
+#there are ~600 rows with SC and no hand foci count
+table( (Curated_BivData %>% filter(  is.na(hand.foci.count) ) )$category )
+#KAZ male and WSB female have >100 observations without hand.foci.counts
+
 #Remove bivalent entries without hand foci count values
@@
 table(cleanCurated_BivData$hand.foci.count, exclude = NULL)
+#600 removed
+```
+
+```{r format.str.Curate_BivData, echo=FALSE}
+#new.Curate.FULL to Curated_BivData
+
+#missing.SKIVE.female
+
+#missing.SKIVE.female=read.xlsx2("~./MLH1repo/data/missing_SKIVE_female.xlsx", header=TRUE, 1)
+
+#
+Curated_BivData <- Curated_BivData[!(is.na(Curated_BivData$fileName) | Curated_BivData$fileName==""), ]
+#this takes a while to run
+
+#new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
+#new.Curate.FULL.org$Obj.ID <- paste(new.Curate.FULL.org$fileName, new.Curate.FULL.org$boxNumber, sep = "_")
+
+#SC PASS
+Curated_BivData <- Curated_BivData[Curated_BivData$SC.pass == '1',]#
+#10480 #9819, #9769
+
+#fix these duplicated ones
+#check duplicates
+#anyDuplicated(missing.SKIVE.female$Obj.ID)
+#0 duplicates for missing SKIVE.f
+#remove duplicates
+Curated_BivData <- Curated_BivData %>% distinct()#from 8933 to 8151
+
+Curated_BivData <- add_mouse(Curated_BivData)
+Curated_BivData <- add_category(Curated_BivData)
+Curated_BivData <- add_strain(Curated_BivData)
+Curated_BivData <- add_sex(Curated_BivData)
+Curated_BivData <- add_subsp(Curated_BivData)
+Curated_BivData <- add_species(Curated_BivData)
+
+#don't think I need this one
+#Foci_TRUE_Curate_BivData <-   Curated_BivData[(!is.na(Curated_BivData$hand.foci.count)),]
+
+##change factor cols to numbers (factor -> chr -> number)
+Curated_BivData$fileName <- as.character(Curated_BivData$fileName)
+Curated_BivData$Obj.ID <- as.character(Curated_BivData$Obj.ID)
+
+Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
+
+##Numeric
+Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
+Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
+Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
+
+Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
+
+Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
+Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
+Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
+Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
+#I think the number of sig figs and rounding changes slightly from the blocks above
+
+#adding more feature calculations (IFD, ect)
+Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
+Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+
+#IFD functions
+#  - IFD: currently I think just does 2CO bivalents 
+#  - IFD.2: for mulitple IFDs (3CO ect.)
+
+#IFD.2 does mulitple IFDs (just absolute)
+#Curated_BivData <- add_IFD.2(Curated_BivData)
+
+Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
+Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
+
+Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
+Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
+Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)
+Curated_BivData$PER_Foci_4 <- as.numeric(Curated_BivData$PER_Foci_4)
+
+#not working
+Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
+Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
+
+#add telomere.dist
+Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
+                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
+                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
+                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
+
+Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
+
+#sis.co.ten
+Curated_BivData <- add_SisCoTen(Curated_BivData)
+
+#add rec.group
+#divide musc males into groups
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                  ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                  ifelse(grepl("SKIVE male", Curated_BivData$category), 1,
+                                           0)))
+
+#remove bivalent entries without hand foci count values
+cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
+                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
+
+#write excel file
+#write.xlsx2(missing.SKIVE.female, "Curated_missing.SKIVE.female.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
+```
+
+```{r FULL.mouse.BivData.table, echo=FALSE}
 
@@
                       n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
-                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),
+                      n4CO = sum( (hand.foci.count == 4), na.rm=TRUE),
 
@@
                      proportion_3CO = (n3CO / nbiv),
+                 proportion_4CO = (n4CO / nbiv),
 
@@
                                  ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
-
 ```
@@
 
+#look at date of imaging affects on total SC
+#count=1
+#for(i in MLH1_data$Original.Name){
+   # print(i)
+#    templist= strsplit(i, split="_")[[1]]
+#    c = (templist[1])
+#    MLH1_data$image.string[count] <-  c
+    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
+#    count= count +1
+#}
+#works better if dat is converted out of image
+#MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')
+
 ```
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")
+save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.30.20.RData")
 

### 2020-05-31 doc/totalSC.Rmd
extra updates for total SC

@@
 
-load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
+load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_5.30.20.RData")
+#load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
 #load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#most recent
@@
 
-Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)
+#remake THIS FILE IT IS MISSING BATCHES 16 and 15
+Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF_2.20.xlsx", header=TRUE, 1)
 
+#Total.SC_DF.2.20 = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
 #this most recent version -- is missing most measures for MOLF_m
@@
 
+#Total.SC_DF.og <- Total.SC_DF.2.20
+SC.skel <- Total.SC_DF.2.20
+
 #format all factors to numbers
@@
 #3 duplicate rows in original DF -- because Random name was blank -- how will these be matched to total.SC?
-
 ```
@@
 
+#write the new SC excel file
+#write.xlsx2(MLH1.merge.Skel, "Total.SC_DF_2.20.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
 ```
@@
 
+#total SC_DF is the new file, dataframe
 #this is a nice merged excel file, still has X, and might have some 'outlier' images
 Total.SC_DF.og <- Total.SC_DF
+Total.SC_DF <- Total.SC_DF
 #quality is mess up here
@@
 Total.SC_DF <- Total.SC_DF[ !grepl("X", Total.SC_DF$X) , ]
+Total.SC_DF <- Total.SC_DF[ !grepl("x", Total.SC_DF$X) , ]
 #Total.SC_DF$X <- droplevels(Total.SC_DF$X)
@@
 
-#table(Total.SC_DF$quality)
-
+str(Total.SC_DF)
 ```
@@
 this.plot_WSB.female <- ggplot(DF.2[DF.2$category == "WSB female",], aes(n.SC.pacc.bivs)) + geom_histogram(bins = 40)+ facet_wrap(~mouse)+ggtitle("Number of curated bivs per image")
-
 ```
@@
 
-# -- EXAMINE the ggplot scatter plots for outliers
-#tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=strain))+ #geom_jitter()+facet_wrap(~category, scales = "free")
+#EXAMINE the ggplot scatter plots for outliers
+tot.sc.ggplot <- ggplot(data= (Total.SC_DF %>% filter(category == "CAST male")  ),
+                        aes(y = bin_size, x=mouse))+ geom_jitter()+facet_wrap(~category, scales = "free")
+
+#MOLF male bin_size skel_size
+#noe there's a molf female
+
+#need round 3 for the batch 16 15
+SC.skel.outliers_round3 <- c("9sep16_30jun16_CAST_m1_sp1_13_rev.tif",
+  "1may16_3nov15_WSB_f1_sp2_1_rev.tif", "30jul15_12jun15_WSB_m2_sp1_1_rev.tif",                            "6mar18_1feb18_LEW_m1_sp1_18_rev.tif", "2jul17_20dec16_LEW_m2_sp1_8_rev.tif", "9jun18_8may17_LEW_m1_sp1_16_rev.tif",
+  "12jan17_4jan17_LEW_f3_sp1_1_rev.tif","9apr18_1feb18_LEW_f1_sp1_7_rev.tif",      "13jan15_30sep14_G_m1_sp1_8_rev.tif", "28jul15_22jun15_G_m2_sp1_7.2_rev.tif", "1oct16_1apr16_G_m4_sp1_5_rev.tif", "27jul15_22jun15_G_m2_sp1_6_rev.tif",        "22jun18_24aug17_PERC_m1_sp1_11_rev.tif",
+  "8apr15_24mar15_PWD_m2_sp1_4_rev.tif", "11mar15_28feb15_PWD_f1_sp1_13_rev.tif",  "8jun15_23apr15_PWD_f2_sp1_21_rev.tif", "11mar15_28feb15_PWD_f6_sp1_4_rev.tif",
+  "5apr18_11jan18_KAZ_f2_sp1_17_rev.tif",
+  "6aug19_10jul19_KAZ_m3_sp1_8.3_rev.tif",  "6aug19_10jul19_KAZ_m2_sp1_5.3_rev.tif", "6aug19_10jul19_KAZ_m3_sp1_5.2_rev.tif  6aug19_10jul19_KAZ_m3_sp1_9_rev.tif",  "6aug19_10jul19_KAZ_m2_sp1_1.2_rev.tif","6aug19_10jul19_KAZ_m2_sp1_4.1_rev.tif", "6aug19_10jul19_KAZ_m1_sp1_11.3_rev.tif", "6aug19_10jul19_KAZ_m1_sp1_11.1_rev.tif", "6aug19_10jul19_KAZ_m1_sp1_11.6_rev.tif", "27jul19_26oct18_SKIVE_f1_sp1_2_rev.tif", "18jul19_31may19_SKIVE_m2_sp1_22.2_rev.tif", "4jul18_8jun18_CZECH_m1_sp1_17.1_rev.tif", "29nov16_13nov16_MSM_f1_sp1_10_rev.tif", "28jul17_30may17_SPRET_m1_sp1_3_rev.tif","24may17_8may17_SPRET_m1_sp1_14_rev.tif", "26sep18_19may18_SPIC_m2_sp1_18_rev.tif", "26sep18_19may18_SPIC_m2_sp1_20_rev.tif", "6jun19_28jun18_SPIC_m2_sp1_8.1_rev.tif", "29dec17_4dec17_SPIC_f2_sp1_11_rev.tif", "29dec17_4dec17_SPIC_f2_sp1_3_rev.tif")
+
+# ( Total.SC_DF  %>% filter(category == "SPIC male") %>% filter(bin_size > 32000) )$fileName
+
+#KAZ -- has lower outliers?!
+#MSM also has kinda strange pattern
+
+Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% SC.skel.outliers_round3, ]
 
@@
 SC.skel.outliers_round1 <- c(
-  #MLH1.merge.Skel  %>% filter(category == "WSB male") %>% filter(skel_size > 2000)
-  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif",
 
-  #MLH1.merge.Skel  %>% filter(category == "WSB female") %>% filter(skel_size > 2700)       
+  "10apr15_31mar15_WSB_m3_sp1_4_rev.tif", "1jul15_12jun15_WSB_f2_sp1_1_rev.tif",
+
             "17nov16_3nov15_WSB_f1_sp3_4_rev.tif","1may16_3nov15_WSB_f1_sp2_6_rev.tif",
@@
 
-MLH1.merge.Skel <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% SC.skel.outliers_round1, ]#62 removed
+Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% SC.skel.outliers_round1, ]#62 removed
 
@@
     #MLH1.merge.Skel  %>% filter(category == "G male") %>% filter(skel_size > 2500)
-    "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
+
+  #Total.SC_DF %>% filter(category == "WSB male") %>% filter(bin_size > 25000) )$fileName
+    " 30jul15_12jun15_WSB_m2_sp1_1_rev.tif", "10apr15_31mar15_WSB_m3_sp1_11_rev.tif", "10apr15_31mar15_WSB_m3_sp1_3.1_rev.tif", "2mar15_30dec14_WSB_m4_sp3_32_rev.tif", "10apr15_31mar15_WSB_m3_sp1_4_rev.tif", "28jul15_22jun15_G_m2_sp1_5_rev.tif","12may15_30sep14_G_m5_sp1_26_rev.tif",
 
@@
 #MLH1.merge.Skel_2  %>% filter(category == "PWD male") %>% filter(skel_size > 2100)
-"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",
+"20jul15_18may15_PWD_m1_sp1_10_rev.tif","22mar15_10mar15_PWD_m2_sp1_11.2_rev.tif","22mar15_10mar15_PWD_m1_sp1_4.1_rev.tif","20jul15_18may15_PWD_m1_sp1_7.2_rev.tif",  "6mar15_1mar15_PWD_m1_sp4_1_rev.tif",
 
@@
 )
-#3remove.mice_round2
 
-MLH1.merge.Skel <- MLH1.merge.Skel[!MLH1.merge.Skel$fileName %in% remove.mice_round2, ]
+Total.SC_DF <- Total.SC_DF[!Total.SC_DF$fileName %in% remove.mice_round2, ]
 
@@
 
-`r sum( length(SC.skel.outliers_round1), length(remove.mice_round2) )` images were removed as probable outliers.  ~20 outliers were removed from total.sc.skel.
+`r sum( length(SC.skel.outliers_round1), length(remove.mice_round2), length(SC.skel.outliers_round3) )` images were removed as probable outliers.  ~20 outliers were removed from total.sc.skel.
 
@@
 #Below threshold is are the values to use
-write.xlsx2(Total.SC_DF, "Total.SC_DF.xlsx", sheetName = "Sheet1", 
+write.xlsx2(Total.SC_DF, "Total.SC_DF_5.31.20.xlsx", sheetName = "Sheet1", 
             col.names = TRUE, row.names = FALSE, append = FALSE)

### 2020-05-31 src/BivData_Setup.Rmd
more updates to figure scripts ect

@@
 new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
+
+#found that one of the data sheets is missin bunch of SKIVE female measures
+#read in missing SKIVE -- then
+
+#original Curate_bivata + missing SKIVE female
+Curated_BivData <-  read.csv2("~./MLH1repo/data/Curated_BivData_5.30.20.csv", sep=" ", header=TRUE)
+#reformate this
+
 #83975
@@
 #Clean away all the empty rows
-new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
+missing.SKIVE.female <- missing.SKIVE.female[!(is.na(missing.SKIVE.female$fileName) | missing.SKIVE.female$fileName==""), ]
 #this takes a while to run
 
-new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
+#new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
+
+missing.SKIVE.female <- add_mouse(missing.SKIVE.female)
 #make OG copy
@@
 
+#missing.SKIVE.female
+
+missing.SKIVE.female=read.xlsx2("~./MLH1repo/data/missing_SKIVE_female.xlsx", header=TRUE, 1)
+str(missing.SKIVE.female)
+
+missing.SKIVE.female$fileName <- as.character(missing.SKIVE.female$fileName)
+
+missing.SKIVE.female <- missing.SKIVE.female[!(is.na(missing.SKIVE.female$fileName) | missing.SKIVE.female$fileName==""), ]
+#this takes a while to run
+
+#new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
+
+new.Curate.FULL.org$Obj.ID <- paste(new.Curate.FULL.org$fileName, new.Curate.FULL.org$boxNumber, sep = "_")
+
 #SC PASS
-new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#
+missing.SKIVE.female <- missing.SKIVE.female[missing.SKIVE.female$SC.pass == '1',]#
 #10480 #9819, #9769
@@
 #fix these duplicated ones
+#check duplicates
+#anyDuplicated(missing.SKIVE.female$Obj.ID)
+#0 duplicates for missing SKIVE.f
 #remove duplicates
@@
 
-Curated_BivData <- add_mouse(Curated_BivData)
-Curated_BivData <- add_category(Curated_BivData)
-Curated_BivData <- add_strain(Curated_BivData)
-Curated_BivData <- add_sex(Curated_BivData)
-Curated_BivData <- add_subsp(Curated_BivData)
-Curated_BivData <- add_species(Curated_BivData)
+missing.SKIVE.female <- add_mouse(missing.SKIVE.female)
+missing.SKIVE.female <- add_category(missing.SKIVE.female)
+missing.SKIVE.female <- add_strain(missing.SKIVE.female)
+missing.SKIVE.female <- add_sex(missing.SKIVE.female)
+missing.SKIVE.female <- add_subsp(missing.SKIVE.female)
+missing.SKIVE.female <- add_species(missing.SKIVE.female)
 
@@
 ##change factor cols to numbers (factor -> chr -> number)
-Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
-Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
-Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)
-
-Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)
-
-Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
-Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
-Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
-Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)
+missing.SKIVE.female$fileName <- as.character(missing.SKIVE.female$fileName)
+missing.SKIVE.female$Obj.ID <- as.character(missing.SKIVE.female$Obj.ID)
+
+missing.SKIVE.female$centromere_ABS_Position <- as.character(missing.SKIVE.female$centromere_ABS_Position)
+missing.SKIVE.female$centromere_PER_Position <- as.character(missing.SKIVE.female$centromere_PER_Position)
+missing.SKIVE.female$chromosomeLength <- as.character(missing.SKIVE.female$chromosomeLength)
+
+missing.SKIVE.female$hand.foci.count <- as.character(missing.SKIVE.female$hand.foci.count)
+
+missing.SKIVE.female$Foci1 <- as.character(missing.SKIVE.female$Foci1)
+missing.SKIVE.female$Foci2 <- as.character(missing.SKIVE.female$Foci2)
+missing.SKIVE.female$Foci3 <- as.character(missing.SKIVE.female$Foci3)
+#missing.SKIVE.female$Foci4 <- as.character(missing.SKIVE.female$Foci4)
 
 ##Numeric
-Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
-Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
-Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)
-
-Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)
-
-Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
-Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
-Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
-Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
+missing.SKIVE.female$centromere_ABS_Position <- as.numeric(missing.SKIVE.female$centromere_ABS_Position)
+missing.SKIVE.female$centromere_PER_Position <- as.numeric(missing.SKIVE.female$centromere_PER_Position)
+missing.SKIVE.female$chromosomeLength <- as.numeric(missing.SKIVE.female$chromosomeLength)
+
+missing.SKIVE.female$hand.foci.count <- as.numeric(missing.SKIVE.female$hand.foci.count)
+
+missing.SKIVE.female$Foci1 <- as.numeric(missing.SKIVE.female$Foci1)
+missing.SKIVE.female$Foci2 <- as.numeric(missing.SKIVE.female$Foci2)
+missing.SKIVE.female$Foci3 <- as.numeric(missing.SKIVE.female$Foci3)
+#Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
 #I think the number of sig figs and rounding changes slightly from the blocks above
@@
 #adding more feature calculations (IFD, ect)
-Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
-Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
+missing.SKIVE.female <- add_IFD(missing.SKIVE.female)#fixed a bug
+missing.SKIVE.female$IFD1_PER = missing.SKIVE.female$IFD1_ABS / missing.SKIVE.female$chromosomeLength
 
@@
 
-Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
-Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength
+missing.SKIVE.female$PER_Foci_1 <- missing.SKIVE.female$Foci1 / missing.SKIVE.female$chromosomeLength
+missing.SKIVE.female$PER_Foci_2 <- missing.SKIVE.female$Foci2 / missing.SKIVE.female$chromosomeLength
+missing.SKIVE.female$PER_Foci_3 <- missing.SKIVE.female$Foci3 / missing.SKIVE.female$chromosomeLength
+missing.SKIVE.female$PER_Foci_4 <- missing.SKIVE.female$Foci4 / missing.SKIVE.female$chromosomeLength
 
@@
 #not working
-Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
-Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength
+missing.SKIVE.female$dis.cent <- missing.SKIVE.female$Foci1 - missing.SKIVE.female$centromere_ABS_Position
+missing.SKIVE.female$dis.cent.PER  <- missing.SKIVE.female$dis.cent / missing.SKIVE.female$chromosomeLength
 
 #add telomere.dist
-Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
-                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
-                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
-                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )
-
-Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength
+missing.SKIVE.female$telo_dist <- ifelse(grepl(1, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength - missing.SKIVE.female$Foci1), 
+                     ifelse(grepl(2, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength -  missing.SKIVE.female$Foci2), 
+                     ifelse(grepl(3, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength - missing.SKIVE.female$Foci3),
+                    ifelse(grepl(4, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength - missing.SKIVE.female$Foci4),   0))) )
+
+missing.SKIVE.female$telo_dist_PER <- missing.SKIVE.female$telo_dist / missing.SKIVE.female$chromosomeLength
 
 #sis.co.ten
-Curated_BivData <- add_SisCoTen(Curated_BivData)
+missing.SKIVE.female <- add_SisCoTen(missing.SKIVE.female)
 
@@
 #divide musc males into groups
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                  ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                  ifelse(grepl("SKIVE male", Curated_BivData$category), 1,
+missing.SKIVE.female$Rec.group <- ifelse(grepl("PWD male", missing.SKIVE.female$category), 1, 
+                                  ifelse(grepl("MSM male", missing.SKIVE.female$category), 1,
+                                  ifelse(grepl("SKIVE male", missing.SKIVE.female$category), 1,
                                            0)))
@@
                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
+
+#write excel file
+#write.xlsx2(missing.SKIVE.female, "Curated_missing.SKIVE.female.xlsx", sheetName = "Sheet1", 
+#            col.names = TRUE, row.names = FALSE, append = FALSE)
+
 ```

### 2020-05-31 fig3_double_tri.R

@@ -3,7 +3,6 @@ remove_strains.1sex <- c("CZECH", "TOM","AST","CZECH","CAST","HMI","SPRET","CARO
 
 #Curated_BivData.FM <- Curated_BivData[ ! Curated_BivData$strain %in% remove_strains.1sex, ]
 
-Curated_BivData
 
 my.data = (Curated_BivData %>% filter(hand.foci.count == 2) )
 #remove_strains
@@ -67,12 +66,17 @@ bottom.triangle <- function(my.data) {
   }
 
 
-my.data.ast <- my.data %>% filter(strain == "AST")
+#try isolating 2COs with terminal 2nd foci -- compare the sex difference
+my.data.ends <- my.data %>% filter(PER_Foci_2 > 0.8)
+
+my.data = (Curated_BivData %>% filter(hand.foci.count == 2) )
+#remove_strains
+
 
 dev.off()
-par(mar=c(2.5, 4, 2, 3), oma=c(1, 1, 1, 1), mfrow=c(1,1))
+par(mar=c(2.5, 4, 2, 3), oma=c(1, 1, 1, 1), mfrow=c(3,3))
 
-by(my.data.ast, my.data.ast$strain, function(sub){
+by(my.data, my.data$strain, function(sub){
   #print(unique(strain) )
   top.triangle(sub)
   bottom.triangle(sub)

### 2020-05-31 fig4_bottom_tri.R

@@ -70,7 +70,7 @@ bottom.triangle.only <- function(my.data) {
   }
 
 
-my.data.wsb <- my.data %>% filter(strain == "WSB")
+
 
 dev.off()
 par(mar=c(2.5, 4, 2, 3), oma=c(1, 1, 1, 1), mfrow=c(1,1))

### 2020-05-31 src/CommonFunc_MLH1repo.R

@@ -249,7 +249,7 @@ add_subsp <- function(oldframe) {
                                                                                                ifelse(grepl("PWD", dframe$strain), "Musc", 
                                                                                                       
                                                                                                       ifelse(grepl("CZECH", dframe$strain), "Musc", 
-                                                                                                             ifelse(grepl("AST", dframe$strain), "Musc",     
+                                     ifelse(grepl("AST", dframe$strain), "Musc",     
                                                                                                                     ifelse(grepl("TOM", dframe$strain), "Musc",  
                                                                                                                            ifelse(grepl("SKIVE", dframe$strain), "Musc", 
                                                                                                                                   
@@ -346,7 +346,6 @@ add_mouse2 <- function(dat){
   return(dframe)
 }
 
-#make sis-co-ten
 #metric for calculating relative / predicted about of sister-cohesion tension at metaphase
 #requires handfocicount, Foci positions, and IFD
 #start with ABS measures

### 2020-05-31 Chap3_MLH1var.rmd
working on the preliminary results for chapter 3

@@
 
-Cor.plot <- ggplot(( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus") ), aes(x=mean_co, y=var, color = strain)) + geom_point() + facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= small.DF, aes(x=27, y=80, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= small.DF, aes(x=27, y=75, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by CO variance per Mouse")
+Cor.plot <- ggplot(( FULL_mouse_table_Q12.MLH1 %>% filter(species == "M.musculus") ), aes(x=mean_co, y=var, color = strain)) + 
+  geom_point() +   facet_wrap(~sex)+scale_color_manual(values=colors_of_strains, guide = FALSE) +  geom_text(data= small.DF, aes(x=27, y=40, label=corr),  colour="black", inherit.aes=FALSE, parse=FALSE ) +  geom_text(data= small.DF, aes(x=27, y=35, label=p.val),  colour="black", inherit.aes=FALSE, parse=FALSE )+ggtitle("Mean by CO variance per Mouse")+ylim(c(0,40))
 
@@
 #plot the first permuted DF-- don't need the transposed
-pwd.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = pwd.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) + ggtitle("PWD difference in variance across sexes")  + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$pwd.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) / nrepeat )
+pwd.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = pwd.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) + ggtitle("PWD difference in variance across sexes")  + annotate("text", x=3.6, y=750, label= "p=" ) + annotate("text", x=5.1, y=750, label= sum(FM.var.diff_permuted$pwd.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "PWD"] ) / nrepeat )
 pwd.perm2.plot
 
-skive.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = skive.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) + ggtitle("SKIVE difference in variance across sexes") + annotate("text", x=3, y=800, label= "p=" ) + annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$skive.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) / nrepeat )
+skive.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = skive.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) + ggtitle("SKIVE difference in variance across sexes") + annotate("text", x=3, y=750, label= "p=" ) + annotate("text", x=7, y=750, label= sum(FM.var.diff_permuted$skive.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "SKIVE"] ) / nrepeat )
 skive.perm2.plot
 
-msm.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = msm.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) + ggtitle("MSM difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$msm.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) / nrepeat )
+msm.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = msm.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) + ggtitle("MSM difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$msm.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "MSM"] ) / nrepeat )
 msm.perm2.plot
 
-g.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = g.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) + ggtitle("G difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$g.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) / nrepeat )
+g.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = g.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) + ggtitle("G difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$g.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "G"] ) / nrepeat )
 g.perm2.plot
 
-wsb.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = wsb.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) + ggtitle("WSB difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$wsb.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) / nrepeat )
+wsb.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = wsb.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) + ggtitle("WSB difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$wsb.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "WSB"] ) / nrepeat )
 wsb.perm2.plot
 
-lew.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = lew.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) + ggtitle("LEW difference in variance across sexes")+ annotate("text", x=3, y=800, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$lew.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) / nrepeat )
+lew.perm2.plot <- ggplot(FM.var.diff_permuted, aes(x = lew.FM.dif)) + geom_histogram() + geom_vline(xintercept = Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) + ggtitle("LEW difference in variance across sexes")+ annotate("text", x=3, y=750, label= "p=" ) +annotate("text", x=4, y=750, label= sum(FM.var.diff_permuted$lew.FM.dif >= Diff.Var$Fvar.Mvar.dif[Diff.Var$strain == "LEW"] ) / nrepeat )
 lew.perm2.plot

### 2020-06-01 doc/MS_v2.Rmd
shortening some longer sections

@@
     toc_float: yes
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
-  word_document:
-    toc: yes
-    toc_depth: '5'
 always_allow_html: yes
@@
 
-```{r set.up, echo=FALSE, include=FALSE}
+```{r set.up, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 library(ggpubr)
-library(reshape2)
-library(patchwork)
+#library(reshape2)
+#library(patchwork)
 library(png)
@@
 library(gridExtra)
-library(dplyr)
 library(cowplot)
@@
 library(xlsx)
-
-#set wd
-#load datafile
+library(reshape2)
 
@@
 
-load(file="~./MLH1repo/data/Results_setup_5.21.20.RData") #Results_setup_5.14.20.RData
-
-#load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData
-
-#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
+load(file="~./MLH1repo/data/Results_setup_5.30.20.RData")
+#load(file="~./MLH1repo/data/Results_setup_5.21.20.RData") #Results_setup_5.14.20.RData
 
@@
 
-### MICE
+### Mice
 
@@
 The following morning, slides were treated with a Photoflow wash (Kodak, cite).
-Slides were stored at -20 \* C if not stained immediately.
+Slides were stored at -20 * C if not stained immediately.
 To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
@@
 Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
-Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\* C.
+Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
 The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\* C.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37* C.
 Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37* C.
 Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
@@
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.
+Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using Photoshop.
 
 ### Analysis
+
+```{r normal.curve, echo=FALSE, include=FALSE, eval=FALSE}
+
+#this block is throwing an error..not sure why
+#normality of MLH1 distributions
+
+#there are a bunch of strains that should be removed
+#remove grey background
+
+#subsp x sex, color by strain
+
+gaus.curve.remove.strains <- c("F1")
+
+gaus.MLH1 <- MLH1_data
+#gaus.MLH1 <- gaus.MLH1[!gaus.MLH1$strain %in% gaus.curve.remove.strains, ]
+
+gaus.MLH1$strain <-  factor(gaus.MLH1$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW","PERC",
+      "PWD","SKIVE","KAZ", "CZECH", "AST", "TOM",
+      "MSM", "MOLF",
+      "CAST", "HMI",
+      "SPIC","SPRET","CAROLI"))
+
+#combine outgroups
+#change subsp for spic, caroli and spret to outgroup
+gaus.MLH1$subsp2[gaus.MLH1$subsp == "Caroli"] <- "Outgroup"
+gaus.MLH1$subsp2[gaus.MLH1$subsp == "Spretus"] <- "Outgroup"
+gaus.MLH1$subsp2[gaus.MLH1$subsp == "Spic"] <- "Outgroup"
+
+gaus.MLH1$subsp2[gaus.MLH1$species == "M.musculus"] <- "M. m. musculus"
+#make seperate curves / plots for outgroups and house mouse
+
+#MLH1_data
+gaussian.curve.out <- ggplot(data = (gaus.MLH1 %>% filter(quality < 5)  %>% filter(subsp2 == "Outgroup") ),
+                         aes(x=nMLH1.foci, fill=strain) ) + geom_histogram( binwidth = 1)+facet_wrap( ~ sex)  +  scale_fill_manual(values=colors_of_strains)+ 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+gaussian.curve.Mm <- ggplot(data = (gaus.MLH1 %>% filter(quality < 5)  %>% filter(subsp == "Cast") ),
+                         aes(x=nMLH1.foci, fill=strain) ) + geom_histogram( binwidth = 1)+facet_wrap(subsp ~ sex) +  scale_fill_manual(values=colors_of_strains)+ 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#aes( alpha = .7)
+
+```
 
@@
 Cells with a score of 5 were excluded from the final analysis.
-Distributions of MLH1 count per cell were visually inspected for normality (supplemental figure ).
+**Distributions of MLH1 count per cell were visually inspected for normality (supplemental figure )**.
 MLH1 foci located on the XY in spermatocytes were excluded from counts.
@@
 Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
+
+<!-- what should be included or removed from these large numbers?? Q5, only house mouse?-->
 
@@
 *discribe the hiearchal nature*.
-strains nested within subspecies -- nesting by default,
+As observations were at the level of single mouse (mouse average) was uniquely coded within the dataset, nesting was implict, each strain only occurs within one strain, and was not explictly coded in our linear and mixed models.
 
@@
 
-# Figure 1
+### Figure 1
 
@@
                         Ncells  =length(nMLH1.foci),
-                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
+
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 4 ), nsmall=3) ),
+                        mean_co_long = mean(nMLH1.foci),
                         cV = cv(nMLH1.foci, na.rm = TRUE),
                         var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
-                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 4),
                        #Standard error (mice vs cells)
-                        se  = round(sd / sqrt(Ncells), 3)
+                        se  = round(sd / sqrt(Ncells), 4)
 
@@
 
-female_mouse.av99 <- Mouse.Table_HQ %>% filter(sex == "female") %>% filter(species == "M.musculus")
-
 female_mouse.av <- add_sex(female_mouse.av)
@@
 
+#### Mouse variance plots
+
 ```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
+```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
+
 #write the full table to a supplemental table
@@
 
-#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
+#kable(strain.table.HQ_Fig.og, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", #"Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain #Level Summary Statistics" )  %>%
 #  kable_styling() %>%
@@
 #write excel
-#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table.xlsx", sheetName = "Sheet1", 
+#write.xlsx2(strain.table.HQ_Fig.og, "second.MLH1.table.xlsx", sheetName = "Sheet1", 
 #  col.names = TRUE, row.names = FALSE, append = FALSE)
@@
 ```
+
+Main table is strain level, (n)
 
@@
 spicilegus_ ).
-Mean MLH1 foci counts for `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$mouse) )` -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 1,867 spermatocytes) and -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
+Mean MLH1 foci counts for `r sum(main.table$Nmice)` -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 1,867 spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
 Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
@@
 
-To further situate variation in recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
+To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
+**to best investigate sex, will limited our analysis to strains with sufficient obersevations from both sexes (Table 1).**
+
 We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < 10^{-6} -- p < `r 1/10000`), sex (p = 1.5510^{-8} -- `r MLH1.MMlmer_results$sex.results`), subspecies (p=1.7210^{-4} -- `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = 3.110^{-5} -- `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
@@
 *new* -- Our survey of recombination rate across house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is reversed.
-In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-test; p < 0.05 -- ).
-In the dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.15 -- ( _musculus^MSM^_) to 1.49 -- ( _domesticus^WSB^_) across strains.
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect (p < 0.05 -- ).
-The existence of some subspecies and strain effects (p < 0.05 -- ) further indicates that SC length has evolved among strains and among subspecies.
+In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (all t-test below; p < 0.05; except short bivalents for SKIVE p = `r Q1_short.SC_t.test.SKIVE$p.value` ).
+
+In the dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_) across strains.
+
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p=`r Q1.lmer_totSC_M1_results$sex.results`; M2: p=`r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p= `r Q1.lmer_short_M1_results$sex.results`; M2: p= `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,]`; M3: `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2]`).
+
+The existence of some subspecies and strain effects (M1 interaction: `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][7,]`,  SKIVE * male `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][13,]`;  M3: SKIVE * male  `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][13,]`) further indicates that SC length has evolved among strains and among subspecies.
+
 In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males have more MLH1 foci per cell.
@@
 
-*old?
--- * Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
-To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count per cell as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < `r 1/1000000`) and general linear models (M2) (p = `r MLH1var_M2_sum$coefficients[,4][9]`) and M3 (p = `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`).
-
-In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
-The results are similar: sex is the strongest effect (M1 p < `r 1/1000000`; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
-When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
-These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
-
-<LmerMLH1_M1_Q12_VAR_results> <Bret's comment for below, are there differences between the curated and the full dataset? -- Q1 and full data set>
-
 ## B Females and males differ in crossover positions and crossover interference
 
-# triangles
-
-```{r double.IFD.triangle.plots, echo=FALSE}
+### triangles
+
+double triangle plots made with different scripts
+
+```{r double.IFD.triangle.plots, echo=FALSE, eval=FALSE}
 
@@
 
-Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 1.4910^{-12} --), with only _musculus^KAZ^_ showing no difference (t-test; p = 0.33 -- `r Q1.IFD.raw_FULL_t.test$p.value` ).
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 1.4910^{-12} -- p < 0.05), with only _musculus^KAZ^_ showing no difference (t-test; p = 0.33 -- `r Q1.IFD.PER_t.test_KAZ$p.value` ).
 Examination of IFD~norm~ distributions indicates that female values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
 Analysis of models treating IFD~norm~ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = 6.7410^{-14} --`r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+
 In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) across the full set of strains, whether they are considered separately or together.
 Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males and that males display a stronger telomeric bias in the placement of the distal crossover.
-In summary, controlling for differences in SC length (chromatin compaction) using IFDnorm indicates that interference is stronger in males, whereas consideration of IFD~raw~ shows that the sexes exhibit a similar level of interference on the physical scale.
+In summary, controlling for differences in SC length (chromatin compaction) using IFD~norm~ indicates that interference is stronger in males, whereas consideration of IFD~raw~ shows that the sexes exhibit a similar level of interference on the physical scale.
 
@@
 turn off all these DMC1 tables -- to makes sure the right DFs are loaded in from the setup scripts.
+
+### Figure 2 DMC1
 
@@
   )
-DMC1.Figure2
+
+#DMC1.Figure2
 #three.DMC1.plot <- { Early.cell.obj + Late.cell.obj  } / DMC1.boxplot + plot_layout(widths = 1)
@@
 ```
+
+# Fig Chrm Proportion plot
 
@@
 
-chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),
-                             aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(legend.position="top", legend.title = element_blank(),
-                                                                                                           axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
-
-kk <- addSmallLegend(chrm.class.barplot)
+chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position="top", legend.title = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+ scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
+
+chrm.class.barplot.final <- addSmallLegend(chrm.class.barplot)
 
@@
 ```{r Chrm.Prop.misc.calulations, echo=FALSE}
-
 total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE - this value used in the text
-#only one SKVIE female, this is breaking the test
-#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
-
-#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
-#      )
-
-```
-
-```{r Q1SC_short.biv_dif, echo=FALSE, eval=TRUE}
-
-#calq the different in short.biv means across sexes
-#Long.biv_mouse_table
-
-#WSB
-Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
-
-#LEW
-Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
-
-#G
-#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
-
-#PWD
-Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
-
-#KAZ
-Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
-
-#SKIVE
-Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
-
-#MSM
-Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
-
-#this bivData ratio is the same as mouse average
-Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
-
 ```
@@
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
-    axis.title.x=element_blank(), axis.ticks.x=element_blank(),
+    axis.title.x=element_blank(),
 axis.text.x = element_text(size=7), 
-legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10) ) +
+legend.position="none",axis.text.y = element_text(size=9), # axis.text.y = element_blank()
+axis.title.y = element_text(size = 10) ) +
 
@@
 ggsave("~/MLH1repo/doc/Figure3_Tot.SC_mouse.av_ALL.png")
+
+#ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")
 
@@
 ```
-
-Make a figure for the diagram.  Import the straightened bivalents.. Load in the wsb double triangle.
 
@@
 We considered proportions of bivalents with different numbers of crossovers, double-strand breaks, SC length, and crossover positions, including crossover interference.
+
+**why chrm proportion table for the sum numbers?
+- because I am summing all of the chrm classes** The number is much smaller (redo the proportions table?) - part of the difference is, these are from the new are from the Clean_Curated_BivData.
+(i'm not sure if the big number if right)
+
 Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` -- 34,982) have either one or two MLH foci (Figure 3).
@@
 _molossinus^MOLF^_: p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
-PWD vs MS: p = `r PWD.MSM.2CO.prop.test$p.value`.
-**Add the extra test PWD-MSM**  These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
+PWD vs MS: p = `r PWD.MSM.2CO.prop.test$p.value`  and `r PWD.MSM.2CO.prop.test$p.value`.
+These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
 
@@
 
+**differences in total SC correlation likely due to more total SC measures**
+
 Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
 **Check if this is correct for all strains**.
+
 Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within _M.
 m.
-molossinus_ (p = 0.03 -- r Q2tot.sc_t.test_High.Low_Mol -- p.value ) but not within M.
+molossinus_ (p = 0.03 -- r Q2tot.sc_t.test_High.Low_Mol$p.value -- p.value ) but not within _M.
 m.
-musculus (p = 0.87 -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+musculus_ (p = 0.87 -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
 
@@
 molossinus_ (IFD~norm~: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value` ; IFD~raw~: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
+**these pvalues aren't the same anymore**
+
 Similar results are observed with models treating IFD~raw~ and IFD~norm~ as dependent variables: only effects associated with high-recombination strains are significant (p < 0.05).
@@
 
+Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
+While this trait will be examined in greater detail in future manuscript, we not that the general pattern is that females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1 and Table 2).
+
+<!-- Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
+While this trait will be examined in greater detail in future manuscript, we not that the general pattern is that females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1 and Table 2).
+
 round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)
@@
 
-,
- Q12 MLH1.var -- M1 -LmerMLH1_M1_Q12_VAR_results -- LmerMLH1_M1_Q12_VAR_results,  M2 -sum.MLH1var_M2_Q12.co_eff,  M3 -MLH1var_M3_Q12_sum
+ Q12 MLH1.var -- M1 -LmerMLH1_M1_Q12_VAR_results -- LmerMLH1_M1_Q12_VAR_results,  M2 -sum.MLH1var_M2_Q12.co_eff,  M3 -MLH1var_M3_Q12_sum -->
 
@@
 
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4}
+```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4, eval=FALSE}
 #Curated_BivData_male
@@
 
-```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
+```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE, eval=FALSE}
 #make the same plots as above but with the long biv data set
@@
 
-```{r Q2.1COpos.show.MOL, echo=FALSE}
+```{r Q2.1COpos.show.MOL, echo=FALSE, eval=FALSE}
 
@@
 
-```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
+```{r Q2.show.IFD, echo=FALSE, warning = FALSE, eval=FALSE}
 #subset
@@
 
-```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE}
+```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE, eval=FALSE}
 
@@
 
-```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE}
+figure script below
+
+```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE, eval=FALSE}
 
@@
 
-```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
+```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups", eval=FALSE}
 
@@
 #draw boxes around the
-
 ```
@@
 ```
+
+correlations are from the fig strain table estimates
 
@@
 In particular, the alternation across strains in which sex has more crossovers is unexpected.
+
 We propose an alternative interpretation based on the cell biology of gametogenesis.
-During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across a proportion of sister cohesion [@LaneKauppi2019].
-The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition [@LaneKauppi2019].
-Hence, selection seems likely to favor mutations that optimize the process of bi-orientation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister cohesin distal to chiasmata [@LaneKauppi2019, @vanveen].
+The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesin holding homologs together [@LaneKauppi2019].
+Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and smooth chromosome seperation , thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
 
@@
 We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
-We hypothesize that these sex differences hinge on disparate requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase-to-anaphase transition is initiated by the proteolytic decay of the sister cohesion proteins that connect homologs (@LaneKauppi2019, @subramanian2014, @dumontDesai2012).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate (@vanVeen2003, @altendorfer2020).
-Differences in timing of the cell cycle between oogenesis and spermatogenesis could impose contrasting selective pressures on how sister cohesion affects chromosome structure.
-Faster spermatogenesis may select for synchronization of the separation homologs (cite).
+
+A source for the evolution of sexual dimorphism in meiotic traits is the spindle assembly checkpoint, which is more strict and effective in spermatogenesis and permissive (or 'leaky') in oogenesis (cite).
+This difference can result in different abilities to fiter sex-specific genetic variants.
+The spindle assembly checkpoint acts to regulate the reductional chromosome division, which is facilitated by the proteolytic decay of sister cohesin distal to chiasmata (that is holding homologs together (during biorientation)).
+The number and placement of crossovers alter the distribution of sister cohesin distal to chiasmata which could in theory affect this metaphase to anaphase transition (@vanVeen2003, @LaneKauppi2019, @subramanian2014, @dumontDesai2012).
+(additional features of gametogenesis may contribute to sex differences) Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell (Kudo ).
 While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dictyate arrest [@Lee2019].
+
 We propose that the SAC model also can explain the correlated evolution of stronger crossover interference and higher genome-wide recombination rate in male house mice.

### 2020-06-02 doc/MS_v2.Rmd
updating manuscript

@@
   word_document:
-    toc: yes
-    toc_depth: '5'
   pdf_document:
@@
 
-*new* - The same dry-down spread technique was applied to both spermatocytes and oocytes based on (@peters_1997), with adjustment for volumes.
+The same dry-down spread technique was applied to both spermatocytes and oocytes based on (@peters_1997), with adjustment for volumes.
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
@@
 The staining protocol followed [@anderson1999] and [@koehler2002].
-Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
+Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
 Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
@@
 Cells with a score of 5 were excluded from the final analysis.
-**Distributions of MLH1 count per cell were visually inspected for normality (supplemental figure )**.
+Distributions of MLH1 count per cell were visually inspected for normality (Supplemental figure 1).
 MLH1 foci located on the XY in spermatocytes were excluded from counts.
@@
 
-Total SC length per cell was quantified from pachtyene cell images [@wang2019_SC].
+Total SC length per cell was quantified from pachytene cell images [@wang2019_SC].
 To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
@@
 To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
-A normalized measure of interference (IFD~norm~) was computed by dividing IFDraw by SC length on a per-bivalent basis.
+A normalized measure of interference (IFD~norm~) was computed by dividing IFD~raw~ by SC length on a per-bivalent basis.
 
@@
 Subspecies, strain, and their interaction were coded as fixed effects,  with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()).
-*discribe the hiearchal nature*.
-As observations were at the level of single mouse (mouse average) was uniquely coded within the dataset, nesting was implict, each strain only occurs within one strain, and was not explictly coded in our linear and mixed models.
+As observations were at the level of single mouse (mouse average) was uniquely coded within the dataset, nesting was implicit, each strain only occurs within one strain, and was not explicitly coded in our linear and mixed models.
 
@@
 # RESULTS
-
-```{r FOR.PLOT.organizing.data, echo=FALSE, message=FALSE, warning=FALSE}
-#set the order of another column, based on another variable. consistant order for ploting
-
-#apply this before ploting
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, category, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-#use the above line for when order is out of wak
-
-#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)
-
-AP_mouse_table.HQ_target <- Mouse.Table_HQ[!(is.na(Mouse.Table_HQ$sex) | Mouse.Table_HQ$sex==""), ]
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
-AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-
-var.plot <- ggplot(AP_mouse_table.HQ_target, aes(y=var, x = sex, color=sex))+geom_jitter()+facet_wrap(~strain)+ggtitle("Plots of mouse Variances for MLH1 counts")+
-  scale_color_manual(values=c("royalblue", "limegreen"))+theme_bw()
-
-```
-
-### Figure 1
-
-```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE}
-
-#DF.HetC.MixedModel.HQ
-#this blck is duplicated in setupData.rm
-#https://rcompanion.org/handbook/G_03.html
-#testing random effects within lme
-
-DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]
-
-#AP_mouse_table excluide Qual5 cells
-DF.HetC.MixedModel.HQ <- Mouse.Table_HQ
-DF.HetC.MixedModel.HQ <- add_species(DF.HetC.MixedModel.HQ)
-
-#main df with variables coded as chr
-DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
-
-all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
-                 .(sex, strain),
-                 summarize,
-                 all.cells = sum(Ncells)
-)
-
-```
 
@@
 
-```{r Fig1.top, echo=FALSE, warning=FALSE, message=FALSE}
-#Mouse.Table_HQ
-#change the segements and text!!
-
-Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
-  ylab("Average MLH1 Foci per Cell")+
-
-  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
-
-    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1)+
-
-  ylim(c(18, 35))+
- scale_color_manual(values=colors_of_strains)+
-
-    theme(legend.position="none",
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"),
-  axis.title.y=element_blank(),
-  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
-#remove y labfor male
-
-   #mol
-  annotate("segment", x=.75, xend=2.3, y=19.5, yend=19.5, colour="black", size=1)+
-
-  #Musc,
-  annotate("segment", x= 2.75, xend=5.2, y=30, yend=30, colour="black", size=1)+
-
-  #dom
-  annotate("segment", x=5.8, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
-
-  #cast
-  annotate("segment", x=8.8, xend= 9.3, y=30, yend=30, colour="black", size=1)+
-
-    #spret
-  annotate("segment",x=9.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
-
-  #spic
-  annotate("segment",x=10.5, xend=11.2, y=30, yend=30, colour="black", size=1)+
-
-  #label
-    annotate("text", x = c(1.6,3.9,6.3,8,10, 10.3), #list of all x points 
-           y = c(18.5, 31, 18.5, 31, 18.5, 31),
-
-   label = c("M. m. molossinus","M. m. musculus", "M. m. domesticus", "M. m. castaneus", "M. spretus", "M. spicilegus"),
-           fontface = 'italic', hjust = 0.5, size=3.5)
-
-```
-
-```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
-
-#MLH1_data  %>% filter(quality < 5)
-#mouse avs
-
-female_mouse.av <- Mouse.Table_HQ.OG %>% filter(sex == "female") %>% filter(species == "M.musculus")
-
-female_mouse.av <- add_sex(female_mouse.av)
-female_mouse.av <- add_strain(female_mouse.av)
-female_mouse.av <- add_subsp(female_mouse.av)
-female_mouse.av <- add_species(female_mouse.av)
-
-female_mouse.av$strain <- factor(female_mouse.av$strain, ordered = FALSE, levels =c( "MSM", "MOLF",
-  "PWD", "SKIVE", "KAZ",
-
-  "WSB", "G", "LEW",
-                                                                         "CAST", 
-                                                                   "SPRET", "SPIC") )
-
-female.specific.plot <- ggplot(female_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+ geom_boxplot()+  geom_jitter(aes( alpha=0.5), width = .2 )+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
-  theme(legend.position="none",
-
-        axis.title.y =element_blank(), 
-      axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+geom_hline(yintercept = 20)
-
-#+theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, message=FALSE}
-#try customize the plot size
-
-male_mouse.avs_for.analysis$strain <- factor(male_mouse.avs_for.analysis$strain, ordered = FALSE, levels =c( "MSM", "MOLF",
-  "PWD", "SKIVE", "KAZ","TOM","AST","CZECH",
-
-  "WSB", "G", "LEW","PERC",
-                                                                         "CAST", "HMI",
-                                                                   "SPRET", "SPIC") )
-
-male.plot <- ggplot( male_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
-
-   theme(legend.position="none", axis.text.x=element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),  axis.title.y = element_blank(), 
-  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(), axis.line=element_line(colour = "black"))+geom_hline(yintercept = 19)
-
-#horizontal line = 20
-
-```
-
-```{r MLH1.cell.examples, echo=FALSE, message=FALSE, warning=FALSE}
-
-male.MLH1.cell.ex <- readPNG('~/MLH1repo/doc/figureFiles/MLH1.male.cell_ex.png')
-
-female.MLH1.cell.ex <- readPNG('~/MLH1repo/doc/figureFiles/MLH1.female_cell.ex.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-#Cell GGPLOT
-
-maleMLH1.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(male.MLH1.cell.ex) + theme(axis.title.x=element_blank(),
-     axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-femaleMLH1.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(female.MLH1.cell.ex) + theme(axis.title.x=element_blank(),
-     axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-```
-
-```{r Figure1.Main.plot.three.plots, echo=FALSE, warning=FALSE, fig.width=7, fig.cap='*Figure 1 Mean MLH1 count distributions by strain.* A) Strain averages of MLH1 counts per cell.
-B) Female specific MLH1 count distributions.
-C)	Male specific MLH1 counts per cell.', message=FALSE}
-
-MLH1.Figure1  <- ggdraw() +
-  draw_plot(Fig1_v1_strain, x = .02, y = .55, width=1, height=.45)+
-
-  draw_plot(female.specific.plot, x=0, y=0, width=.5,height= .55)+
-  draw_plot(femaleMLH1.cell.obj, x=.35, y=.35, width=.18, height= .18)+
-
-  draw_plot(male.plot, x=.5, y = 0,  width = .5, height = .55)+
-  draw_plot(maleMLH1.cell.obj, x=.85, y = .35,  width = .19, height=.19)+
-
-  draw_plot_label(label = c("A", "B", "C"), size = 14,
-                  x = c(0, .01, .51), y = c(.99, .55, 0.55) )
-
-MLH1.Figure1
-#add the cell images
-
-#ggsave("~/MLH1repo/doc/Figure1.png")
-```
-
-## B Genome-wide recombination rate evolves differently in females and males
-
 ```{r MLH1.summ.stats, echo=FALSE}
@@
 main.table <- add_species(main.table)
-
-```
-
-```{r female.strain.table, echo=FALSE}
-female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-
-                        Nmice = length(unique(mouse)),
-
-                        total.cells  = sum(Ncells, na.rm = TRUE),
-
-                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#F.strain.table
-
-#F.strain.table
-
-# I don't know why most of the table is blank
-```
-
-#### Mouse variance plots
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r male.strain.table, echo=FALSE}
-
-M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#M.strain.table
-#same issue as the female table .. ugh
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
 
@@
 
-Main table is strain level, (n)
+## Genome-wide recombination rate evolves differently in females and males
 
@@
 To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
-**to best investigate sex, will limited our analysis to strains with sufficient obersevations from both sexes (Table 1).**
+**to best investigate sex, will limited our analysis to strains with sufficient observations from both sexes (Table 1).**
 
@@
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`),  *why is mol so different*  _molossinus^MSM^_  (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p= 0.02 -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
+Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p= 0.02 -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
@@
 
-## B Synaptonemal complexes are longer in females
-
-*new* -- Our survey of recombination rate across house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is reversed.
+## Synaptonemal complexes are longer in females
+
+Our survey of recombination rate across house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is reversed.
 In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (all t-test below; p < 0.05; except short bivalents for SKIVE p = `r Q1_short.SC_t.test.SKIVE$p.value` ).
 
-In the dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_) across strains.
-
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p=`r Q1.lmer_totSC_M1_results$sex.results`; M2: p=`r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p= `r Q1.lmer_short_M1_results$sex.results`; M2: p= `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,]`; M3: `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2]`).
+In the dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_PWD` ( _musculus^PWD^_) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_) across strains.
+
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p=`r Q1.lmer_totSC_M1_results$sex.results`; M2: p=`r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p= `r Q1.lmer_short_M1_results$sex.results`; M2: p= `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2]`; M3: `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2]`).
 
@@
 
-## B Females and males differ in crossover positions and crossover interference
-
-### triangles
-
-double triangle plots made with different scripts
-
-```{r double.IFD.triangle.plots, echo=FALSE, eval=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
-auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]
-
-df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
-df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
-#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)
-
-#this works
-top.triangle <- function(my.y, my.x, my.data) {
-  plot(my.y ~ my.x, data= my.data, axes=FALSE, ylab='', xlab="", 
-       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1,
-       col =  colors_of_strains[current.strain])
-
-  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.95)
-  mtext("Position.2.Female", 2, .75)
-  mtext("Position.1.Female", 3, 2)
-
-  axis(side=2, las=1, pos=0)
-  axis(side=3, las=1, pos=1)
-  lines(0:1, 0:1)
-}
-
-bottom.triangle <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[current.strain])
-
-  mtext("Position.2.Male", 1, 1.5, at=mean(par()$usr[1:2]) + x.dist)
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +  7)
-  x.at <- axisTicks(par()$usr[1:2], 0) + x.dist
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-  mtext(seq(0, 1, .2), 1, 0, at=x.at)
-  axis(4, las=1, pos=1 + x.dist)
-  lines(0:1 + x.dist, 0:1, xpd=TRUE)
-}
-
-x.dist <- .1
-RP <- list()
-
-for(i in 1:(length(df_strain_list))){
-  #print(i)
-  current.strain <- as.character(df_strain_list[[i]]$strain[1])
-
-  #declare file for saving png
-  #png(paste0("~./MLH1repo/doc/figureFiles/IFD.triangle_", current.strain, ".png"),  width=650, height=500, res=100)
-
-  plot.new()
-
-  my.data = (Curated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-
-  my.x.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.top = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  my.x.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1
-
-  my.y.bottom = (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins
-  op <- par(mar=c(.6,.6, .6, .6) + 0.1, oma=c(2, 0, 0, 2))
-
-  top.triangle(my.y.top, my.x.top, my.data)
-
-  bottom.triangle(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-```
-
-```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Distribution of normalized foci positions (IFDs) for double crossover bivalents across sexes.
-Female and male patterns are displayed in the top and bottom triangle respectively.
-Each point reflects a single double crossover bivalent, with the normalized positions of foci reflected on each axis."}
-
-plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_WSB.png')
-plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_LEW.png')
-plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_G.png')
-plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_KAZ.png')
-plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_SKIVE.png')
-plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_PWD.png')
-plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/IFD.triangle_MSM.png')
-
-# grid.arrange(rasterGrob(plot1.WSB), rasterGrob(plot2.LEW), rasterGrob(plot3.G),
-#             rasterGrob(plot4.KAZ), rasterGrob(plot5.SKIVE), rasterGrob(plot6.PWD),
-#             rasterGrob(plot7.MSM), 
-#             ncol=3)
-
-#load all the files / arrange the plots from above 
-#WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/test.triangle_WSB.png")
-```
+## Females and males differ in crossover positions and crossover interference
 
@@
 
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-#mouse.av.long.biv.table
-
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-```
-
-*new* -- We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
+We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
 In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68 -- `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.56; t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value` ).
@@
 
-## B Evolution of genome-wide recombination rate is dispersed across bivalents, associated with double-strand break number, and connected to crossover interference
-
-turn off all these DMC1 tables -- to makes sure the right DFs are loaded in from the setup scripts.
-
-### Figure 2 DMC1
-
-```{r DMC1.boxplot.plot, echo=FALSE, warning=FALSE}
-
-#change order
-
-DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c(
-                      "WSB","G", "KAZ","PWD","MSM"), order=T )
-
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter() +
-
-    labs(y = "DMC1 count per Cell")+
-
-  scale_x_discrete( labels = c( expression( italic( domesticus^WSB) ), expression(  italic( domesticus^G) ), expression( italic( musculus^KAZ) ),
-
-     expression( italic(musculus^PWD)), expression( italic( molossinus^MSM) ) ) ) +
-
-  scale_color_discrete(name = "", labels = c("Early Zygotene", "Late Zygotene"))+scale_color_manual(values=c("#E69F00", "#0072B2"))+
-
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position="top",
-        panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title.x = element_blank())
-
-#"#E69F00" orange
-#"#0072B2" darker blue
-
-```
-
-```{r Main.plot.DMC1, echo=FALSE, fig.width=6.5, fig.height=5, fig.cap="*Figure 2 male DSB estimates* A) Example early zygotene spermatocyte spread, B) Example late zygotene spermatocyte spread.
-Green foci show DMC1.
-C)	Distribution of DMC1 counts per cell  by strains.", message=F, warning=F}
-
-png.DMC1.early.cell <- readPNG('~/MLH1repo/doc/figureFiles/cell.examples/DMC1.early/good.early.cell.png')
-png.DMC1.late.cell <- readPNG('~/MLH1repo/doc/figureFiles/cell.examples/DMC1.late/good.late.cell.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-#EMPTY GGPLOT
-Early.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(png.DMC1.early.cell) + theme(axis.title.x=element_blank(),
-                         axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-Late.cell.obj <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(png.DMC1.late.cell) + theme(axis.title.x=element_blank(),
-                         axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-#cowplot
-
-DMC1.Figure2  <- ggdraw() +
-  draw_plot(Early.cell.obj, x=0, y=.5, width=.4,height= .45)+
-  draw_plot(Late.cell.obj, x=.5, y=.5,  width = .4, height = .45)+
-  draw_plot(DMC1.boxplot, x=0, y = 0,  width =1, height=.55)+
-
-  draw_plot_label(label = c("A", "B", "C"), size = 14,
-                  x = c(0, .49, 0), y = c(1,1, 0.55) )+
-#WSB
-  draw_line(
-    x = c(0.12, .23),
-    y = c( 0.01, .01),
-    color = colors_of_strains['WSB'], size=1
-  ) +
-  #G
-  draw_line(
-    x = c( 0.31, .42),
-    y = c(0.01, 0.01),
-    color = colors_of_strains['G'],size=1
-  ) +
-  #KAZ
-  draw_line(
-    x = c(0.49, .57),
-    y = c(0.01, 0.01),
-    color = colors_of_strains['KAZ'], size=1
-  ) +
-  #PWD
-  draw_line(
-    x = c(0.64, .73),
-    y = c(0.01, .01),
-    color = colors_of_strains['PWD'], size=1
-  ) +
-  #MSM
-  draw_line(
-    x = c(0.80, .91),
-    y = c(0.01, .01),
-    color = colors_of_strains['MSM'], size=1
-  )
-
-#DMC1.Figure2
-#three.DMC1.plot <- { Early.cell.obj + Late.cell.obj  } / DMC1.boxplot + plot_layout(widths = 1)
-
-#three.DMC1.plot + plot_annotation(
-#  title = '',
-#  subtitle = '') + plot_annotation(tag_levels = 'A')
-
-#ggsave("~/MLH1repo/doc/Figure2.png")
-
-```
+## Evolution of genome-wide recombination rate is dispersed across bivalents, associated with double-strand break number, and connected to crossover interference
 
@@
 
-# Fig Chrm Proportion plot
-
-```{r Chrm.props.plot2, echo=FALSE, fig.width=8, fig.cap="*Figure 3* Chromosome Class Proportions. proportions calculated from pooled single bivalent data by strain."}
-
-#barplots for proportions of chrm classes
-
-#consider adding this the the setup block (could have done this to category level table)
-mouse.level.BivData_table$per.0CO <- mouse.level.BivData_table$n0CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.1CO <- mouse.level.BivData_table$n1CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.2CO <- mouse.level.BivData_table$n2CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.3CO <- mouse.level.BivData_table$n3CO / mouse.level.BivData_table$tot.biv
-
-mouse.level.BivData_table$per.4CO <- mouse.level.BivData_table$n4CO / mouse.level.BivData_table$tot.biv
-
-MELT_Chrm.Proportion <- melt(mouse.level.BivData_table[,c(
-        "mouse", "Nmice", "Ncells", "boxs.IDd", "nboxes_passed",  "nbiv",               "tot.biv", "mean_SC", "var.SC_length", "sd.SC_length", "se.SC_length",        "mean.SC_1CO", "mean.SC_2CO", "mean.SC_3CO", "mean.SC_4CO", "mean_1CO.pos",  "mean_1CO.pos.PER", "norm.first.foci.pos", "mean_IFD.2CO_ABS", "mean_IFD.2CO_PER", "mean_IFD_ABS", "mean_IFD_PER", "mean.siscoten", "mean.siscoten_1CO", "mean.siscoten_2CO", "mean.siscoten_3CO", "mean.telo.dist", "mean.cent.dist",     "mean.sis.co.ten", "sex", "strain", "subsp", "category", "Rec.group", "species",
-
-               #leave out the Foci
-                "per.0CO", "per.1CO", "per.2CO", "per.3CO", "per.4CO")],
-
-  id=c("mouse", "Nmice", "Ncells", "boxs.IDd", "nboxes_passed",  "nbiv",               "tot.biv", "mean_SC", "var.SC_length", "sd.SC_length", "se.SC_length",        "mean.SC_1CO", "mean.SC_2CO", "mean.SC_3CO", "mean.SC_4CO", "mean_1CO.pos",  "mean_1CO.pos.PER", "norm.first.foci.pos", "mean_IFD.2CO_ABS", "mean_IFD.2CO_PER", "mean_IFD_ABS", "mean_IFD_PER", "mean.siscoten", "mean.siscoten_1CO", "mean.siscoten_2CO", "mean.siscoten_3CO", "mean.telo.dist", "mean.cent.dist",     "mean.sis.co.ten", "sex", "strain", "subsp", "category", "Rec.group", "species"))
-
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[(!is.na(MELT_Chrm.Proportion$value)),]
-MELT_Chrm.Proportion <- MELT_Chrm.Proportion[!(is.na(MELT_Chrm.Proportion$value) | MELT_Chrm.Proportion$value==0), ]
-
-#no filename for add mouse, strain subsp
-MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)
-MELT_Chrm.Proportion <- add_species(MELT_Chrm.Proportion)
-
-#then this DF has to be summarized into categories
-chrm.class.category_MELT <- ddply(MELT_Chrm.Proportion, c("strain","sex", "variable"), summarise,
-                             mean.percent.type = mean(value)
-                             #sum is the wrong!
-                             #used average percentage, might consider a different summary
-)
-
-#can't apply function use this
-chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
-             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
-                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
-                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",
-
-          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
-              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",
-
-                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
-                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
-            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
-                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
-                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
-                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
-                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
-                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",
-
-             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
-               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))
-
-chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",                    "Cast",
-                                                                 "Musc",
-                                                                 "Mol",
-                                                                 "Spretus","Spic",
-                                                                 "other") )
-#ggplot
-
-chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position="top", legend.title = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+ scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
-
-chrm.class.barplot.final <- addSmallLegend(chrm.class.barplot)
-
-```
-
-```{r Chrm.Prop.misc.calulations, echo=FALSE}
+```{r Chrm.Prop.misc.calulations, echo=FALSE, eval=FALSE}
+
 total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-```
-
-# Fig 3
-
-```{r Q1.SC.show.short.biv, echo=FALSE, eval=TRUE}
-
-#Short.biv_mouse_table
-
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(), axis.line = element_line(colour = "black"),  legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=7), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9) )+
-
-  labs(y = "Short SC Length")
-
-ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")
-
-```
-
-```{r Q1.tot.SC.plot, echo=FALSE, warning=FALSE}
-# fig.width = 4, fig.height=2.5
-
-Q1.totSC.mouse.av.table <-  mouse.total.skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")
-
-Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
-   "WSB","G","LEW","PERC", 
-  "PWD", "SKIVE",  "KAZ",
-  "MSM", "MOLF"), order=T )
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#order the strains
-Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
-  arrange(strain) %>%
-  mutate(mouse = factor(mouse))
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-Q1.totSC.mouse.av.table$subsp2 <- 
-   ifelse(grepl("Dom", Q1.totSC.mouse.av.table$subsp), "M. m. domesticus", 
-        ifelse(grepl("Musc",Q1.totSC.mouse.av.table$subsp), "M. m. musculus",
-            ifelse(grepl("Mol",Q1.totSC.mouse.av.table$subsp), "M. m. molossinus",   
-               "other")))
-
-#change this plot to look like the short bivalent plot -- boxplots + mouse points
-
-Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
-  scale_color_manual(values=colors_of_strains)+
-    facet_wrap(~subsp, scales="free", ncol = 3) +
-
-  geom_jitter()+ geom_boxplot(alpha=.2)+ 
-  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
-
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
-    axis.title.x=element_blank(),
-axis.text.x = element_text(size=7), 
-legend.position="none",axis.text.y = element_text(size=9), # axis.text.y = element_blank()
-axis.title.y = element_text(size = 10) ) +
-
-  labs(y = "Total SC Length")+ylim(c(1000,2600 ))+ggtitle("")
-
-ggsave("~/MLH1repo/doc/Figure3_Tot.SC_mouse.av_ALL.png")
-
-#ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")
-
-```
-
-```{r Fig3.load.icons, echo=FALSE}
-# load in the icons
-
-all.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/all_yellow_binary.png')
-single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-#EMPTY GGPLOT
-all.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(all.chrm.yellow) +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-single.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(single.chrm.yellow) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-    axis.title.x=element_blank(),
-   axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-```
-
-```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8, fig.cap= 'caption'}
-
-Q2.SC_dif.Figure3  <- ggdraw() +
-
-  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
-  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
-
-  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
-  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
-
-  #drawing triangles in the corner
-  #top right-triangle
-  draw_line(
-    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
-    y = c( .25, .25,  .15, .25),
-    color = "blue", size=.75
-  ) +
-
-  #bottom right-triangle
-  draw_line(
-    x = c( .1, .2, .2, .1 ),
-    y = c( .1,  .2, .1, .1),
-    color = "black", size=.75
-  ) +
-
-  draw_plot_label(label = c("A", "B", "C"), size = 14,
-                  x = c(0, .49, 0), y = c(1,1, 0.33) )
-
-```
-
-```{r Q1.total_SC.table, echo=FALSE}
-
-totSC.strain.av.table <-  mouse.total.skel[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
-
-#head of table
-#see the DMC1 code
-
-#kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#add variance
-#write.xlsx(totSC.strain.av.table, file = "Results_Table.xlsx", sheetName = "Total.SC.results", append = TRUE, row.names = F)
-
-```
-
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#ALL
-Q1_tot.sc_t.test.all <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(species == "M.musculus")  )$mean.skel)
-
-#dom
-Q1_tot.sc_t.test.DOM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#Musc
-Q1_tot.sc_t.test.MUSC <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
-#cell level total SC
-#MLH1.merge.Skel   $skel_size
-#error not enought female observations, ONLY 1 female SKIVE mouse
-
-Q1_tot.scBIV_t.test.SKIVE <- t.test((Total.SC_DF %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
-           (Total.SC_DF %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-```{r Q1.totSC_male.first.comp, echo=FALSE}
-
-Q1.male.Dom.Musc.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
-```
-
-```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- mouse.total.skel %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-
-```
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-Male.cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-Male.cell.av.long.biv.table <- add_mouse(Male.cell.av.long.biv.table)
-Male.cell.av.long.biv.table <- add_strain(Male.cell.av.long.biv.table)
-Male.cell.av.long.biv.table <- add_subsp(Male.cell.av.long.biv.table)
-
-Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
-Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
-Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
-```
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-
-#+theme(legend.position="none") 
-```
-
-```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( Long.biv_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+
-  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
-
-mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
-
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice.
-A)	 Mean total SC area per mouse across subspecies and recombination groups.
-High- and low-recombining groups are indicated by 1 and 0 respectively.
-(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell", echo=FALSE}
-
-#make plot of mouse averages for MLH1 and total SC
-bep <- merge(  male_mouse.avs_for.analysis,   mouse.total.skel)
-
-X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+theme(legend.position="none")+labs(x="Mouse Mean Total SC", y="Mouse Mean MLH1 count")
-
-#plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
-
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-```{r Q2.totSC.ttest, echo=FALSE}
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol -- not enough Mol observations
-#Q2tot.sc_t.test_High.Low_Mol <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-
-#this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-
-glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
-
-sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
-
-#MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
-  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
-sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
-
-#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
-#DOM
-glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
-
-#MUSC
-glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
-
-#MOL
-#within subspecies
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
+
+```
+
+```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE, eval=FALSE}
 #REMOVE THIS, IT uses total bivalent data
@@
 
-*new* -- We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
 We considered proportions of bivalents with different numbers of crossovers, double-strand breaks, SC length, and crossover positions, including crossover interference.
-
-**why chrm proportion table for the sum numbers?
-- because I am summing all of the chrm classes** The number is much smaller (redo the proportions table?) - part of the difference is, these are from the new are from the Clean_Curated_BivData.
-(i'm not sure if the big number if right)
 
@@
 
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in _molossinus^MSM^_ (*SUP* Figure 3).
-Following patterns in the genome-wide recombination rate, male musculusPWD and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in _molossinus^MSM^_ (Supplemental Figure 3).
+Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
 _musculus^KAZ^_: p = 3.1510^{-33} -- `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_: p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`).
-PWD vs MS: p = `r PWD.MSM.2CO.prop.test$p.value`  and `r PWD.MSM.2CO.prop.test$p.value`.
+_molossinus^MOLF^_: p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`) and _musculus^PWD^_ vs.
+_molossinus^MSM^_: p = `r PWD.MSM.2CO.prop.test$p.value`.
 These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
@@
 
-**differences in total SC correlation likely due to more total SC measures**
-
 Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
-**Check if this is correct for all strains**.
 
@@
 m.
-molossinus_ (p = 0.03 -- r Q2tot.sc_t.test_High.Low_Mol$p.value -- p.value ) but not within _M.
+molossinus_ (p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value` ) but not within _M.
 m.
@@
 
-**Q2.SC.reduced** Q2short.biv_t.test_High.Low
 Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value` ; long: p = 0.18 -- `r Q2long.biv_t.test_High.Low$p.value`).
@@
 molossinus_, p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[3,4]` ).
-In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05 -- ), but they are not consistent across models.
+In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05), but they are not consistent across models.
 Collectively, these observations reveal that evolution of the genome-wide recombination rate is not strongly associated with evolution of SC length.
 
-**Q2-CO.pos**  Q2.CO.pos t-test
 High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = 0.24 -- `r Q2.1COpos_Ttest$p.value` ) or within subspecies (t-test; _M.
@@
 m.
-molossinus_ : p = 0.07 -- r Q2.1COpos_mus.av_Mol_ttest$p.value ).
-**While domesticusWSB and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable** (Q2.COpos), these two strains do not differ in MLH1 count.
+molossinus_ : p = 0.07 -- `r Q2.1COpos_mus.av_Mol_ttest$p.value` ).
+While _domesticus^WSB^_ and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable, these two strains do not differ in MLH1 count.
 These results show that evolution of crossover position on chromosomes with single crossovers is decoupled from evolution of the genome-wide recombination rate, at least at this scale of resolution.
@@
 **Q2_IFD**
-High-recombination strains have greater inter-focal distances than low-recombination strains (t-test; IFD~norm~: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value` ; IFD~raw~: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
+High-recombination strains have greater inter-focal distances (in both raw and normalized measured) than low-recombination strains (t-test; IFD~norm~: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value` ; IFD~raw~: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
 This pattern holds within _M.
@@
 molossinus_ (IFD~norm~: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value` ; IFD~raw~: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-**these pvalues aren't the same anymore**
+**these pvalues aren't the same anymore**  is this cause there aren't many molf measures?
 
@@
 The main distinction in IFD~norm~ distributions between high-recombination and low-recombination strains is an enrichment of IFD~norm~ values under 30% in low-recombination strains (Figure 6).
-The frequency of IFDnorm values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
+The frequency of IFD~norm~ values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
 These results indicate that evolution of the genome-wide recombination rate is accompanied by the evolution of crossover interference.
@@
 
-# IFD double triangle
-
-```{r Comp_Ratios, echo=FALSE, eval=FALSE}
-
-#calq ratios of belo 30
-
-top.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-bottom.male <-  G.Bivalents.DF %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-ratio.male <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-top.female <-  G.Bivalents.DF %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-bottom.female <-  G.Bivalents.DF %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-ratio.female <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-#what is the fastest way to calq all these ratios?  also what's the fastst way to get them into a DF / table?
-
-#put all the DFs in a list, for each DF in the list -- apply the filters for the nom and denominator
-MOLF.Bivalents.DF <- Curated_BivData %>% filter(strain == "MOLF")
-
-DF.strain.list <- list(
-  WSB.Bivalents.DF, G.Bivalents.DF, LEW.Bivalents.DF, 
-  PWD.Bivalents.DF, KAZ.Bivalents.DF, SKIVE.Bivalents.DF,
-  MSM.Bivalents.DF, MOLF.Bivalents.DF
-  )
-
-entr_DF <- data.frame(strain=c("WSB", "G", "LEW", "PWD", "KAZ", "SKIVE", "MSM", "MOLF"), 
-        total.2CO_female=c(0,0,0,0,0,0,0,0), total.2CO_male=c(0,0,0,0,0,0,0,0), below30_male = c(0,0,0,0,0,0,0,0), below30_female = c(0,0,0,0,0,0,0,0),     
-          female.ratio = c(0,0,0,0,0,0,0,0), male.ratio = c(0,0,0,0,0,0,0,0))
-
-i = 1
-for(x in (DF.strain.list) ){
-
-  #print(head())
-
-  #these filters won't apply to the indexed DF
-  top.male =  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2) %>%filter(IFD1_PER <= .3)
-
-  bottom.male <-  x %>% filter(sex== "male") %>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_male[i] <- length(bottom.male$Obj.ID)
-  entr_DF$below30_male[i] <- length(top.male$Obj.ID)
-  entr_DF$male.ratio[i] <- length(top.male$Obj.ID)/ length(bottom.male$Obj.ID)
-
-  top.female <-  x %>% filter(sex== "female") %>%  filter(hand.foci.count== 2) %>% filter(IFD1_PER <= .3)
-
-  bottom.female <-  x %>% filter(sex== "female")%>%  filter(hand.foci.count== 2)
-
-  entr_DF$total.2CO_female[i] <- length(bottom.female$Obj.ID)
-  entr_DF$below30_female[i] <- length(top.female$Obj.ID)
-  entr_DF$female.ratio[i] <- length(top.female$Obj.ID)/ length(bottom.female$Obj.ID)
-
-  i = i+1
-
-  }
-#entr_DF
-
-entr_DF$sex.dif <- entr_DF$female.ratio - entr_DF$male.ratio
-
-```
-
-```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4, eval=FALSE}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- cleanCurated_BivData %>% filter(sex == "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.23 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
-
-```
-
-```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE, eval=FALSE}
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q2.1COpos.show.MOL, echo=FALSE, eval=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
-
-mus.av1COpos.plot <- ggplot(data= mouse.level.BivData_table %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
-                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-#+facet_wrap(~mouse, scales = "free")
-
-single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
-                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
-
-#+scale_color_manual(values=colors_of_strains)
-
-single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
-                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-
-#CHRM.length is very significant-- some of the same subsp and strains are sig     
-```
-
-```{r Q2.show.IFD, echo=FALSE, warning = FALSE, eval=FALSE}
-#subset
-male.biv.2CO <- cleanCurated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-
-male.biv.2CO$Rec.group <- ifelse(grepl("PWD", male.biv.2CO$strain), 1, 
-                                    ifelse(grepl("MSM", male.biv.2CO$strain), 1, 
-                                           ifelse(grepl("SKIVE", male.biv.2CO$strain), 1, 0)))
-
-biv.IFDs_PER_by.group <- ggplot(male.biv.2CO, aes(y = IFD1_PER, x= strain, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("")+
-  scale_color_manual(values=colors_of_strains)+ facet_wrap(Rec.group ~ subsp, scales = "free")+
-  theme(legend.position="none", axis.text.x = element_blank(), axis.title.x = element_blank() )  + labs("Normalized IFD ") +
-  ylim(c(0,1.1))
-
-#consider adding horizontal line
-
-#to adjust the boxplot width
-#https://stackoverflow.com/questions/46174222/ggplot2-incorrect-boxplot-width-when-facetting-with-facets-of-different-scales
-#I don't think there's a straightforward way to achieve that in ggplot.
-You may have better luck plotting each facet as a separate ggplot object (on a subset of the data), then arrange them on a single page, using grid.arrange() from gridExtra for example.
-
-```
-
-```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")  )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
-#summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
-#summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol") )
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
-#summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
-#summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol")  )
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
-```
-
-# Single rewrite IFD triangle
-
-figure script below
-
-```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-
-Q2.males.IFD <- c("F1", "other")
-Q2.auto_curate_DF <- cleanCurated_BivData[ ! cleanCurated_BivData$strain %in% Q2.males.IFD, ]
-
-Q2.df_strain_list <- split(Q2.auto_curate_DF, (Q2.auto_curate_DF$strain))
-
-#this works
-Q2.bottom.triangle2 <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[unique(my.data$strain)])
-
-  mtext("Position.2.Male", 1, -2, at= mean(par()$usr[1:2]) + 0)
-
-  #the second val will move the x lable when negative
-
-  #at should be changing the position
-  #above is just changing the text not the numbers
-  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)
-
-  #second number - on which MARgin line, starting at 0 counting outwards.
-  #where text should be,  at = 0, 1 (bad)
-
-  mtext("Position.1.Male", 4, 0, padj=par()$usr[1] + .5) #last value at .5 is good for moving y label
-
-  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
-
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=T, xpd=TRUE)
-#changing labels=T, puts the ddefault labels back
-
-  #second value should be around .75 (moves label text closer)
-  axis(4, las=1, pos=1, at=x.at) #this is the y axis
-  lines(0:1, 0:1, xpd=TRUE) #diagonal line
-}
-
-x.dist <- .25
-RP <- list()
-
-for(i in 1:(length(Q2.df_strain_list))){
-  #print(i)
-  current.strain <- as.character(Q2.df_strain_list[[i]]$strain[1])
-
-  #par(mar = c(2, 2, 2, 2))
-  #dev.new(width = 8, height = 4, unit = "in")
-  #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
-  plot.new()
-
-  my.data = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-  my.x.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
-  my.y.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins (the edges)
-  op <- par(mar=c(.25, .15, .25, .75), oma=c(2, 0, 0, 2))
-#(1=bottom, 2=left, 3=top, 4=right)
-
-Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-
-```
-
-```{r Q2.load.bottom.tri, echo=FALSE}
-
-#only load, SKIVE, PWD, MSM
-#MOLF, KAZ, LEW
-Q2.bottom.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
-Q2.bottom.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
-Q2.bottom.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
-Q2.bottom.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
-Q2.bottom.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
-Q2.bottom.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-#EMPTY GGPLOT
-Q2.LEW.bottom.tri <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(Q2.bottom.LEW) +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-```
-
-```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups", eval=FALSE}
-
-Q2.SC_dif.Figure3  <- ggdraw() +
-
-  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
-  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
-
-  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
-  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
-
-  #drawing triangles in the corner
-  #top right-triangle
-  draw_line(
-    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
-    y = c( .25, .25,  .15, .25),
-    color = "blue", size=.75
-  )
-
-#draw boxes around the
-```
-
 # Discussion
@@
 ```
-
-correlations are from the fig strain table estimates
 
@@
 Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice (@dumont2011, @Wang2017island).
-These variants could differentially affect females and males at any step in the recombination pathway.
+These variants could differently affect females and males at any step in the recombination pathway.
 Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contributed to sex-specific evolution in the recombination rate.
@@
 The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesin holding homologs together [@LaneKauppi2019].
-Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and smooth chromosome seperation , thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and smooth chromosome separation , thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
 
@@
 A source for the evolution of sexual dimorphism in meiotic traits is the spindle assembly checkpoint, which is more strict and effective in spermatogenesis and permissive (or 'leaky') in oogenesis (cite).
-This difference can result in different abilities to fiter sex-specific genetic variants.
-The spindle assembly checkpoint acts to regulate the reductional chromosome division, which is facilitated by the proteolytic decay of sister cohesin distal to chiasmata (that is holding homologs together (during biorientation)).
+This difference can result in different abilities to filter sex-specific genetic variants.
+The spindle assembly checkpoint acts to regulate the reductional chromosome division, which is facilitated by the proteolytic decay of sister cohesin distal to chiasmata (that is holding homologs together (during bi-orientation)).
 The number and placement of crossovers alter the distribution of sister cohesin distal to chiasmata which could in theory affect this metaphase to anaphase transition (@vanVeen2003, @LaneKauppi2019, @subramanian2014, @dumontDesai2012).
@@
 Our results show that crossovers are spaced further apart in strains enriched for double-crossover bivalents when SC length is taken into account and chromosome size effects are minimized.
-Assuming chromatin compaction between (prophase) pachtyene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and improve the fidelity of chromosomal segregation.
+Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and improve the fidelity of chromosomal segregation.
 Although the SAC model postulates direct fitness effects of interference, a modifier model predicts that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].

### 2020-06-02 src/Result_Setup.Rmd
updating manuscript

@@
 
+```{r Q1.total_SC.table, echo=FALSE, eval=FALSE}
+
+totSC.strain.av.table <-  mouse.total.skel[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
+
+#head of table
+#see the DMC1 code
+
+#kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
+#  kable_styling() %>%
+#  scroll_box(width = "800px", height = "200px")
+
+#add variance
+#write.xlsx(totSC.strain.av.table, file = "Results_Table.xlsx", sheetName = "Total.SC.results", append = TRUE, row.names = F)
+
+```
+
+```{r Q1.total.SC_t.test, echo=FALSE}
+
+#run t-test for male female avergage for each strain
+
+#ALL
+Q1_tot.sc_t.test.all <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(species == "M.musculus")  )$mean.skel)
+
+#dom
+Q1_tot.sc_t.test.DOM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
+
+#WSB
+Q1_tot.sc_t.test.WSB <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
+#G
+Q1_tot.sc_t.test.G <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
+#LEW
+Q1_tot.sc_t.test.LEW <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
+
+#Musc
+Q1_tot.sc_t.test.MUSC <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
+
+#PWD
+Q1_tot.sc_t.test.PWD <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
+
+#KAZ
+Q1_tot.sc_t.test.KAZ <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
+
+#SKIVE
+#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
+#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
+#cell level total SC
+#MLH1.merge.Skel   $skel_size
+#error not enought female observations, ONLY 1 female SKIVE mouse
+
+Q1_tot.scBIV_t.test.SKIVE <- t.test((Total.SC_DF %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
+           (Total.SC_DF %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
+
+#MSM
+Q1_tot.sc_t.test.MSM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
+                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
+
+```
+
+```{r Q1.totSC_male.first.comp, echo=FALSE}
+
+Q1.male.Dom.Musc.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
+
+Q1.male.Dom.Mol.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
+```
+
+```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
+
+Q1.lmer_totSC.mouse.av.table <- mouse.total.skel %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
+
+#str(Q1.lmer_totSC.mouse.av.table)
+
+#unorder factors
+Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
+Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
+
+#TOTAL SC AREA
+lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
+
+lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
+
+lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
+
+lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
+
+lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
+sex.results =  lmer.Q1_mus.av_totSC_sex, 
+               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
+lmer.Q1_mus.av_totSC_strain)
+
+```
+
+```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
+
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+
+#Q1lmer.MM_SC_short_results
+
+```
+
 mouse total sc t.tests male female (Q1)
@@
 
-```{r Q1SC_short_ratios, echo=FALSE, eval=TRUE}
-
-#calq the different in short.biv means across sexes
-#Long.biv_mouse_table
-
+```{r Q1_short_ratios, echo=FALSE, eval=TRUE}
 #WSB
@@
 #G
-#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
+Q1SC.mus.av.short.biv_DIFF_G <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
 
@@
 sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
+sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,]
 
@@
 
+```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
+
+#real.long.bivData
+#real.short.bivData
+
+Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.long <- add_sex(Q1.IFD.long)
+Q1.IFD.long <- add_strain(Q1.IFD.long)
+Q1.IFD.long <- add_subsp(Q1.IFD.long)
+
+Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
+Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
+
+Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
+                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
+
+                     )
+
+Q1.IFD.short <- add_sex(Q1.IFD.short)
+Q1.IFD.short <- add_strain(Q1.IFD.short)
+Q1.IFD.short <- add_subsp(Q1.IFD.short)
+
+Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
+Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
+
+#t.test. mouse av
+Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+
+Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
+                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
+#both the short and long norm IFD are NS
+
+#t.test Biv
+Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+
+Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
+                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
+```
+
 #Q2 section
@@
 
+```{r Q2.SC.short.long.setup, echo=FALSE}
+
+#table(male.long.biv$strain)
+#the Long biv dataset is quite unbalance (esp MSM)
+male.long.biv <- real.long.bivData %>% filter(sex== "male")
+
+#mouse av for long biv
+Male.cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+Male.cell.av.long.biv.table <- add_mouse(Male.cell.av.long.biv.table)
+Male.cell.av.long.biv.table <- add_strain(Male.cell.av.long.biv.table)
+Male.cell.av.long.biv.table <- add_subsp(Male.cell.av.long.biv.table)
+
+Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
+Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
+Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
+
+Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                         #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
+                         )
+
+Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
+Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
+
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
+
+#divide musc males into groups
+Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
+
+male.short.biv <- real.short.bivData %>% filter(sex== "male")
+
+Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
+
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength),
+
+                        #mean.first.1CO
+                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
+                         ) )
+
+Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
+Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
+
+Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
+Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
+
+Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
+                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
+                                           0)))
+
+Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
+                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
+```
+
+```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
+
+Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
+Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
+
+#adjust order of plot here
+Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
+  "WSB","G","LEW","PERC",  
+"PWD", "SKIVE",  "KAZ",
+     "MSM", "MOLF",   
+                     "CAST",
+                         "SPRET","SPIC" ), order=T )
+
+Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
+
+#nested boxplots, pooled Bivalents
+#x axis for hand.focicount -- good for seperating the points --- but it pools
+
+#prediction for Dom strains: low variation 
+Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
+  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
+
+Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
+
+#+theme(legend.position="none") 
+```
+
+```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#real.long.bivData
+#real.short.bivData
+
+shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot( Long.biv_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+```
+
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
+
+mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+
+  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
+
+mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
+
+```
+
+```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
+
+#investigating potential skew / technical effects in the algorithm
+
+Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+#not sure what my expectations for this plot are,
+
+median(bep$oldest.image.date) #"2017-08-03"
+#there might be an issue with the algorithm performance 
+#given the parrelle diagonal lines -- 
+#this might be due to the dates imaged  2017-02-07
+
+bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
+
+Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
+
+#there's evidence that the algorithm might have some biased effects -- 
+#that will be investigated later.
+
+```
+
+```{r Q2.totSC.ttest, echo=FALSE}
+
+#move this to MS_v2?
+
+#totalSC high and low rec group
+
+Q2tot.sc_t.test_High.Low <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+#p-value = 0.007
+
+#mol -- not enough Mol observations
+#Q2tot.sc_t.test_High.Low_Mol <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+
+#musc
+Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+
+```
+
+```{r Q2.short.long.ttest, echo=FALSE}
+
+#long and short 
+#mouse.av.short.biv.table
+#Q2mouse.av.short.biv.table
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
+
+#p-value = 0.9
+
+#mouse.av.long.biv.table
+Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
+#p-value = 0.2
+
+#mUSC-short
+Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-short
+#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
+
+#musc-long
+Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-long
+#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
+
+#not enought observations
+
+```
+
+```{r Q2.SC_HvL_allBiv, echo=FALSE}
+
+#High-Low_All-BivData
+
+Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#short
+Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#long
+Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+```
+
+```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
+
+#Dom-Musc
+Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
+#p-value = 0.5
+
+#Dom-Mol
+Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
+#p-value = 0.1
+```
+
+```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
+
+#DOM-MUSC-ALL
+Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 4e-13
+
+#DOM-MOL-ALL
+Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 6e-10
+
+#DOM-musc-long
+Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.2
+
+#DOM-musc-short
+Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.06
+
+```
+
+```{r Q2.glm.SC.long, echo=FALSE}
+
+#this dataset is all males
+glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+
+glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+
+glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
+
+sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
+
+#MUSC
+glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
+  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
+sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
+
+#MOL can't be done -- only 1 mouse for each MSM and MOLF
+#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
+
+#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
+```
+
+```{r Q2.glm.SC.short, echo=FALSE}
+#Q2mouse.av.short.biv.table
+#mouse.av.short.biv.table
+
+glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
+sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
+
+#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
+
+glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
+glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
+
+#DOM
+glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
+
+#MUSC
+glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
+
+#MOL
+#within subspecies
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+```
+
 #Q2 1CO postion
@@
 
+```
+
+```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE, eval=FALSE}
+#make the same plots as above but with the long biv data set
+male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+#mouse.av.long.biv.table
+
+Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
+
+Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
+                        Nmice = length(unique(mouse)),
+                        Nbiv  = length(Obj.ID),
+                       mean.1CO_PER = mean(PER_Foci_1)
+                      # mean_CO = mean()
+                       )
+
+Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
+Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
+
+Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
+
+```
+
+```{r Q2.1COpos.show.MOL, echo=FALSE, eval=FALSE}
+
+#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
+
+mus.av1COpos.plot <- ggplot(data= mouse.level.BivData_table %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
+                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
+#+facet_wrap(~mouse, scales = "free")
+
+single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
+                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
+
+#+scale_color_manual(values=colors_of_strains)
+
+single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
+                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
+
+#CHRM.length is very significant-- some of the same subsp and strains are sig     
 ```
@@
 
+```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#M1.ABS
+Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")  )
+sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
+
+#M1.PER
+Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+
+sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
+#summary(Q2.M1glm_IFD.PER)
+
+#M2.raw
+Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
+
+Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+#within subspecies
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
+#summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
+#summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol") )
+#summary(Q2.glm.M2_IFD_Mol)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
+#summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
+#summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol")  )
+#summary(Q2.glm.M2_IFD.PER_Mol)
+
+```
+
 # SAVE the data
@@
 ```
+
+```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
+#REAL.short_mouse_table
+#REAL.long_mouse_table
+#mouse.av.long.biv.table
+
+Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+
+Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+```

### 2020-06-02 doc/MS_v2.Rmd
updating the Rdata files, reduce redundancyes

@@
 
-new - Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
+Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
@@
 
-*new-* To estimate the number of crossovers across the genome, we counted MLH1 foci.
+To estimate the number of crossovers across the genome, we counted MLH1 foci.
 MLH1 foci were counted in cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
@@
 Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
-
 Total SC length per cell was quantified from pachytene cell images [@wang2019_SC].
@@
 `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female).
-From the total set of cell images, `r formatC( length(unique(cleanCurated_BivData %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
+From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
 Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained.
@@
 A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
-`r sum(Long.biv_mouse_table$n_obs)`  number of long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and xx `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
+A total of `r sum(Long.biv_mouse_table$n_obs)`  long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
 
@@
 We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
-We first constructed a mixed model (M1) using the lmer() from the lmer4 package (cite) in R (v3.5.2) [@Rstudio].
-In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT()).
-Subspecies, strain, and their interaction were coded as fixed effects,  with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()).
+We first constructed a mixed model (M1) using the lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
+In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT() from RLRsim [@RLRsim]).
+Subspecies, strain, and their interaction were coded as fixed effects,  with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
 As observations were at the level of single mouse (mouse average) was uniquely coded within the dataset, nesting was implicit, each strain only occurs within one strain, and was not explicitly coded in our linear and mixed models.
@@
 We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
-In separate analyses, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, variance in MLH1 count per cell, coefficient of variation in MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
+In separate analyses, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
 Four additional models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from the initial mixed model.
@@
 spicilegus_ ).
-Mean MLH1 foci counts for `r sum(main.table$Nmice)` -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 1,867 spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
+Mean MLH1 foci counts for `r sum(main.table$Nmice)` -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` -- 1,867 spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
 Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
@@
 Second, the relationship between MLH1 foci in the sexes varies among strains.
-Although most strains show more MLH1 foci in females, three strains - _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ - exhibit higher MLH1 focus counts in males.
+Although most strains show more MLH1 foci in females, two strains - _musculus^PWD^_ and _molossinus^MSM^_ - exhibit higher MLH1 focus counts in males.
+While _musculus^SKIVE^_ has a very low degree of sex differences in MLH1 focus counts.
 In females, numbers of MLH1 foci are evenly distributed around the sex-wide mean of approximately 25 (Figure 1B).
@@
 
-To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
-**to best investigate sex, will limited our analysis to strains with sufficient observations from both sexes (Table 1).**
+To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
+To investigate the effects of sex, we limited our analysis to `r length( unique( DF.HetC.MixedModel.HQ$strain) )` strains with sufficient observations from both sexes (Table 1).
 
@@
 After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = 1.0410^{-6} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
+A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = 1.0410^{-6} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
 In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
@@
 We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4).
-In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = 0--`r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).
+In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).
 
@@
 In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (all t-test below; p < 0.05; except short bivalents for SKIVE p = `r Q1_short.SC_t.test.SKIVE$p.value` ).
+In the reduced dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from  `r Q1SC.mus.av.short.biv_DIFF_PWD` ( _musculus^PWD^_) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_) across strains.
 
-In the dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.15 -- `r Q1SC.mus.av.short.biv_DIFF_PWD` ( _musculus^PWD^_) to 1.49 -- `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_) across strains.
-
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p=`r Q1.lmer_totSC_M1_results$sex.results`; M2: p=`r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p= `r Q1.lmer_short_M1_results$sex.results`; M2: p= `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2]`; M3: `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2]`).
-
-The existence of some subspecies and strain effects (M1 interaction: `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][7,]`,  SKIVE * male `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][13,]`;  M3: SKIVE * male  `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][13,]`) further indicates that SC length has evolved among strains and among subspecies.
-
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p=`r Q1.lmer_totSC_M1_results$sex.results`; M2: p=`r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p= `r Q1.lmer_short_M1_results$sex.results`; M2: p= `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
+The existence of some subspecies and strain effects (M1 interaction: p = `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]`,  SKIVE * male p= `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.
 In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males have more MLH1 foci per cell.
@@
 ## Females and males differ in crossover positions and crossover interference
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-#real.long.bivData
-Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-
-Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-```
 
@@
 Our conclusions are accompanied by several caveats.
-First, MLH1 foci only identify interfering crossovers **(REF)**.
+First, MLH1 foci only identify interfering crossovers [@holloway2008mus81].
 Although most crossovers (XX%; REF) belong to this class, our approach likely underestimated genome-wide recombination rates.

### 2020-06-02 src/Result_Setup.Rmd
updating the Rdata files, reduce redundancyes

@@
 
-#what was this for
-#all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
-#                 .(sex, strain),
-#                 summarize,
-#                 all.cells = sum(Ncells)
-#)
+#remove male only data
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
 ```
@@
 
-```{r table.long.biv, echo=FALSE}
+```{r table.long.biv, echo=FALSE, eval =FALSE}
+#This table is already made in setupData
 
@@
 
-#2) Long Biv for male.poly
-#just males, just Musculus
-Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
-
-Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
-
-#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
-```
-
-```{r table.short.biv, echo=FALSE}
-
+```
+
+```{r table.short.biv, echo=FALSE, eval=FALSE}
+
+#this table is made in SetUpData
 #real.short.bivData
@@
 #unordered for the mixed models (move this to the MM blocks...?)
-Short.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
-Short.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
-Short.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
 
@@
                              ifelse(grepl("SKIVE", Short.biv_mouse_table$category), 1, 0)) )
-
-#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)
-
-#1) Long Biv for 
-#remove all the non-used strains!!
-#Short.biv_mouse_table_HetC <- Short.biv_mouse_table
-
-#Short.biv_mouse_table_HetC <- Short.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
-
-#Short.biv_mouse_table_HetC <- Short.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
-
-#Short.biv_mouse_table_HetC <- Short.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
-
-#Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)
-
-#remove na's
-#Long.biv_mouse_table_HetC <- #Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | #Long.biv_mouse_table_HetC$mouse==""), ]
-
-```
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
-#moved this to setupdata
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-#Long.biv_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
 
@@
 
-<!-- plots -->
+```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#real.long.bivData
+#real.short.bivData
+
+shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+```
+
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
+
+mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+
+  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
+
+mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
+
+```
+
+```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
+
+#investigating potential skew / technical effects in the algorithm
+
+Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
+#not sure what my expectations for this plot are,
+
+median(bep$oldest.image.date) #"2017-08-03"
+#there might be an issue with the algorithm performance 
+#given the parrelle diagonal lines -- 
+#this might be due to the dates imaged  2017-02-07
+
+bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
+
+Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
+
+#there's evidence that the algorithm might have some biased effects -- 
+#that will be investigated later.
+
+```
+
+```{r Q2.totSC.ttest, echo=FALSE}
+
+#totalSC high and low rec group
+
+Q2tot.sc_t.test_High.Low <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(species == "M.musculus") )$mean.skel,
+                (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+
+#mol -- why is there only 1 MOLF male now?
+#there are 3 MOLF males in bivdata
+Q2tot.sc_t.test_High.Low_Mol <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+
+#musc
+Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+
+```
+
+The molf male t-tests don't work because I only have total Sc measures for 1 MOLF male. to of the
+
+```{r Q2.totSC.ttest_old, echo=FALSE, eval=FALSE}
+
+#move this to MS_v2?
+
+#totalSC high and low rec group
+
+Q2tot.sc_t.test_High.Low <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
+#p-value = 0.007
+
+#mol -- not enough Mol observations
+#Q2tot.sc_t.test_High.Low_Mol <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
+#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+
+#musc
+Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
+                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
+
+```
+
+todo list if I get back to lab.
+
+```{r Q2.short.long.ttest, echo=FALSE}
+
+#long and short 
+#mouse.av.short.biv.table
+#Q2mouse.av.short.biv.table
+
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
+
+#p-value = 0.9
+
+#mouse.av.long.biv.table
+Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
+#p-value = 0.2
+
+#mUSC-short
+Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-short
+#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
+
+#musc-long
+Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-long
+#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
+
+#not enought observations
+
+```
+
+```{r Q2.short.long.ttest.old, echo=FALSE, eval=FALSE}
+
+#long and short 
+#mouse.av.short.biv.table
+#Q2mouse.av.short.biv.table
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
+
+#p-value = 0.9
+
+#mouse.av.long.biv.table
+Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
+#p-value = 0.2
+
+#mUSC-short
+Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-short
+#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
+
+#musc-long
+Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
+                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
+
+#mol-long
+#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
+#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
+
+#not enought observations
+
+```
+
+```{r Q2.SC_HvL_allBiv, echo=FALSE}
+
+#High-Low_All-BivData
+
+Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#short
+Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#long
+Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+```
+
+```{r Q2.SC_HvL_allBiv.old, echo=FALSE, eval=FALSE}
+
+#High-Low_All-BivData
+
+Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#short
+Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+#long
+Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+
+```
+
+```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
+
+#Dom-Musc
+Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
+#p-value = 0.5
+
+#Dom-Mol
+Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
+#p-value = 0.1
+```
+
+```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
+
+#DOM-musc-long
+Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.08
+
+#DOM-musc-short
+Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
+                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
+#p-value = 0.09
+
+```
+
+```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
+
+#DOM-MUSC-ALL
+Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 4e-13
+
+#DOM-MOL-ALL
+Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 6e-10
+
+#DOM-musc-long
+Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.2
+
+#DOM-musc-short
+Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.06
+
+```
+
+```{r Q2.SC_across.subsp_short.long_biv.old, echo=FALSE, eval=FALSE}
+
+#DOM-MUSC-ALL
+Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 4e-13
+
+#DOM-MOL-ALL
+Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 6e-10
+
+#DOM-musc-long
+Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.2
+
+#DOM-musc-short
+Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.8
+
+#DOM-mol-long
+Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.4
+
+#DOM-mol-short
+Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+#p-value = 0.06
+
+```
+
+```{r Q2.totSC.glm, echo=FALSE}
+
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = F)
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = F)
+
+Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
+sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M4)$coefficients
+
+#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
+#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
+#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)
+
+#M4.2? (removed subspecies?)
+
+Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
+sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
+#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
+
+#Dom
+Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") %>% filter(species == "M.musculus") )
+sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
+#there are significant strain effects in DOM,
+
+#Musc
+Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(species == "M.musculus") )
+sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
+#KAZ -- is the only significant effect
+
+#Mol
+Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )
+sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)
+
+```
+
+```{r Q2.glm.SC.long, echo=FALSE, eval=FALSE}
+
+#this dataset is all males
+glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+
+glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+
+glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
+
+#MUSC
+glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
+
+#MOL can't be done -- only 1 mouse for each MSM and MOLF
+glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
+sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
+
+```
+
+```{r Q2.glm.SC.long.old, echo=FALSE, eval=FALSE}
+
+#this dataset is all males
+glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+
+glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
+sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
+#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
+
+glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
+
+sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
+
+#MUSC
+glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
+  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
+sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
+
+#MOL can't be done -- only 1 mouse for each MSM and MOLF
+#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
+
+#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
+
+#Dom
+glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
+#Musc
+glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
+#summary(glm_SC.male_Musc_long.biv)
+#Mol
+glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
+```
+
+```{r Q2.glm.SC.short, echo=FALSE}
+#Q2mouse.av.short.biv.table
+#mouse.av.short.biv.table
+
+glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
+sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
+
+#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
+
+glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
+glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
+
+#DOM
+glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
+
+#MUSC
+glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
+
+#MOL
+#within subspecies
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+```
+
+```{r Q2.glm.SC.short.old, echo=FALSE, eval=FALSE}
+#Q2mouse.av.short.biv.table
+#mouse.av.short.biv.table
+
+glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
+sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
+
+#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
+
+glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
+glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
+
+#DOM
+glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
+sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
+
+#MUSC
+glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
+sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
+
+#MOL
+#within subspecies
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
+
+```
+
+```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class_all)
+
+#use the 1CO or 2CO DF
+Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
+
+log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class.WSB)
+
+Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
+log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_G,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.G)
+
+Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
+log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.LEW)
+
+Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
+log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.PWD)
+
+Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
+log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.SKIVE)
+
+Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
+log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.KAZ)
+
+Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
+log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.CZECH)
+
+Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
+log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MSM)
+
+Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
+log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
+              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
+summary(log.SC_chrm.class.MOLF)
+
+```
+
+```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
+log.reg.chrm.class_all.strains
+
+WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
+
+MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
+                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
+#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
+
+#1CO
+log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.poly)
+
+log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_1CO.Musc)
+#double check these directions
+
+log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+
+#summary(log.SC_male_1CO.Mol)
+
+#2CO
+log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.poly)
+
+log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Musc)
+
+log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_2CO.Mol)
+
+#3CO
+log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.poly)
+
+log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Musc)
+
+#there's no rec group for this table
+log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
+              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
+#summary(log.SC_male_3CO.Mol)
+
+#reverse logistic regression (single bivalent level)
+#how well does SC length predict 1 or 2 COs?
+
+Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
+
+#need to switch the co counts to 1 - 0
+Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
+                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
+
+Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
+Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
+
+log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
+              data=Bivalent_1o2,  family=binomial(link="logit"))
+
+summary(log.SC_male_chrm.class)
+#not sure if this is the equivilent of running each strain seperatly
+#the strain effect is probably also pulling in the different proportion of chrm class
+#useing the chrm.length*strain gives fewer significant factors.
+
+#i'd be more convinced by running all those log regression plots for each strain
+#come back later
+```
+
+```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE, eval=FALSE}
+
+plot.blah.mouse.av <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+plot.logReg.mouse.av_1CO <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
+                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
+SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
+
+```
+
+```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
+#COMPARISONS ACROSS ALL CLASSES
+
+#MOUSE AVE, all
+Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
+
+#mouse.av 1CO
+Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
+
+#MOUSE.av 2CO
+Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
+                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
+
+Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
+
+#single biv level
+#male.biv?
+
+#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
+#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
+
+#Rec.group_BivData_F1post_Ttest$p.value
+
+```
+
+```{r log.reg_SC.length, echo=FALSE, include=FALSE, eval=FALSE}
+#REMOVE THIS COMPARES FROM FULL BIV DATA
+
+#male poly shouldn't be here
+Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
+                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
+                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
+
+#All chrm
+log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av)
+
+#considering seperating out by hand foci counts
+
+#1CO
+log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_1CO)
+
+#2CO
+log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_mouse.av_2CO)
+
+#3CO
+log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
+                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_mouse.av_3CO)
+
+# bivalent level log regression
+
+log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV)
+
+#1CO  (BIVDATA)
+log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
+
+                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_1CO)
+
+#2CO
+log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+#summary(log.reg_SC.length_BIV_2CO)
+
+#3CO
+log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
+                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
+
+#summary(log.reg_SC.length_BIV_3CO)
+```
+
+```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
+#REMOVE THIS, IT uses total bivalent data
+
+Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
+Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
+
+#calq_mouse av
+mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                        n_cells = length(unique(fileName) ),
+                        mean.SC.total = mean(chromosomeLength),
+
+                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
+                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
+
+                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
+                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
+
+                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
+                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
+
+                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
+                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
+
+                         )
+
+mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
+mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
+
+mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
+mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
+
+#calq_mouse av
+mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
+                        n_obs = length(Obj.ID),
+                         mean.SC = mean(chromosomeLength)
+                         )
+mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
+mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
+
+mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
+mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
+```
 
@@
 
-```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
-
-#MLH1_data <- MLH1_data %>%
-#  arrange(strain, sex, mouse) %>%
-#  mutate(file.name = factor(file.name))
-
-mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av.
-total SC area")
-
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-```{r Q2.totSC.ttest, echo=FALSE}
-
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(species == "M.musculus") )$mean.skel,
-                (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-
-#mol -- why is there only 1 MOLF male now?
-#there are 3 MOLF males in bivdata
-Q2tot.sc_t.test_High.Low_Mol <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-The molf male t-tests don't work because I only have total Sc measures for 1 MOLF male. to of the todo list if I get back to lab.
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.09
-
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-```{r Q2.totSC.glm, echo=FALSE}
-
-mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = F)
-mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = F)
-
-Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
-sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M4)$coefficients
-
-#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
-#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
-#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)
-
-#M4.2? (removed subspecies?)
-
-Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
-sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
-#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
-
-#Dom
-Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") %>% filter(species == "M.musculus") )
-sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
-#there are significant strain effects in DOM,
-
-#Musc
-Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(species == "M.musculus") )
-sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
-#KAZ -- is the only significant effect
-
-#Mol
-Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )
-sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)
-
-```
-
-```{r Q2.glm.SC.long, echo=FALSE, eval=FALSE}
-
-#this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-
-glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
-
-#MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
-sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
-#DOM
-glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
-
-#MUSC
-glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
-
-#MOL
-#within subspecies
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
-```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.SC_male_chrm.class_all <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class_all)
-
-#use the 1CO or 2CO DF
-Bivalent_1o2_WSB <-  Bivalent_1o2 %>% filter(strain == "WSB")
-
-log.SC_male_chrm.class.WSB <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_WSB,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class.WSB)
-
-Bivalent_1o2_G <-  Bivalent_1o2 %>% filter(strain == "G") 
-log.SC_chrm.class.G <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_G,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.G)
-
-Bivalent_1o2_LEW <-  Bivalent_1o2 %>% filter(strain == "LEW") 
-log.SC_chrm.class.LEW <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_LEW,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.LEW)
-
-Bivalent_1o2_PWD <-  Bivalent_1o2 %>% filter(strain == "PWD") 
-log.SC_chrm.class.PWD <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_PWD,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.PWD)
-
-Bivalent_1o2_SKIVE <-  Bivalent_1o2 %>% filter(strain == "SKIVE") 
-log.SC_chrm.class.SKIVE <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_SKIVE,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.SKIVE)
-
-Bivalent_1o2_KAZ <-  Bivalent_1o2 %>% filter(strain == "KAZ") 
-log.SC_chrm.class.KAZ <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_KAZ,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.KAZ)
-
-Bivalent_1o2_CZECH <-  Bivalent_1o2 %>% filter(strain == "CZECH") 
-log.SC_chrm.class.CZECH <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_CZECH,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.CZECH)
-
-Bivalent_1o2_MSM <-  Bivalent_1o2 %>% filter(strain == "MSM") 
-log.SC_chrm.class.MSM <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MSM,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MSM)
-
-Bivalent_1o2_MOLF <-  Bivalent_1o2 %>% filter(strain == "MOLF")
-log.SC_chrm.class.MOLF <- glm(chrm.class ~ chromosomeLength,
-              data=Bivalent_1o2_MOLF,  family=binomial(link="logit"))
-summary(log.SC_chrm.class.MOLF)
-
-```
-
-```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}
-
-log.reg.chrm.class_all.strains <- ggplot(Bivalent_1o2,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO)")+facet_wrap(subsp~ strain)
-log.reg.chrm.class_all.strains
-
-WSB.log.reg.chrm.class <- ggplot(Bivalent_1o2_WSB,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Chrm Class prediction (1CO or 2CO, WSB)")+facet_wrap(~subsp)
-
-MSMlog.reg.chrm.class <- ggplot(Bivalent_1o2_MSM,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("MSM Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-PWDlog.reg.chrm.class <- ggplot(Bivalent_1o2_PWD,
-                          aes(x=chromosomeLength, y=chrm.class)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("PWD Chrm Class prediction (1CO or 2CO)")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE COMPARES ACROSS FROM FULL BIVDAAT
-
-#1CO
-log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.poly)
-
-log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_1CO.Musc)
-#double check these directions
-
-log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-
-#summary(log.SC_male_1CO.Mol)
-
-#2CO
-log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.poly)
-
-log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Musc)
-
-log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_2CO.Mol)
-
-#3CO
-log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.poly)
-
-log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Musc)
-
-#there's no rec group for this table
-log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
-              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
-#summary(log.SC_male_3CO.Mol)
-
-#reverse logistic regression (single bivalent level)
-#how well does SC length predict 1 or 2 COs?
-
-Bivalent_1o2 <-  Curated_BivData_male %>% filter(species== "M.musculus") %>% filter(hand.foci.count > 0) %>% filter(hand.foci.count < 3)
-
-#need to switch the co counts to 1 - 0
-Bivalent_1o2$chrm.class <- ifelse(grepl(1, Bivalent_1o2$hand.foci.count), 0, 
-                                    ifelse(grepl(2, Bivalent_1o2$hand.foci.count), 1, ""))
-
-Bivalent_1o2$chrm.class <- as.numeric(Bivalent_1o2$chrm.class)
-Bivalent_1o2$strain <- factor(Bivalent_1o2$strain, ordered = FALSE)
-
-log.SC_male_chrm.class <- glm(chrm.class ~ chromosomeLength*strain,
-              data=Bivalent_1o2,  family=binomial(link="logit"))
-
-summary(log.SC_male_chrm.class)
-#not sure if this is the equivilent of running each strain seperatly
-#the strain effect is probably also pulling in the different proportion of chrm class
-#useing the chrm.length*strain gives fewer significant factors.
-
-#i'd be more convinced by running all those log regression plots for each strain
-#come back later
-```
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE, eval=FALSE}
-
-plot.blah.mouse.av <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot( Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r Q2.SClength.ttest, echo=FALSE, include=FALSE, eval=FALSE}
-#COMPARISONS ACROSS ALL CLASSES
-
-#MOUSE AVE, all
-Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC_Ttest$p.value  #p-value = 0.00238
-
-#mouse.av 1CO
-Rec.group.mouse.SC.1CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_1CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.1CO_Ttest$p.value  #p-value = 0.9993
-
-#MOUSE.av 2CO
-Rec.group.mouse.SC.2CO_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
-                                 Male.poly.Mouse.Table_BivData_4MM$mean.SC_2CO[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])
-
-Rec.group.mouse.SC.2CO_Ttest$p.value  #p-value = 0.03428
-
-#single biv level
-#male.biv?
-
-#Rec.group_BivData_F1post_Ttest <- t.test(male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 1], 
-#                                 male.bivdata.1CO$PER_Foci_1[male.bivdata.1CO$Rec.group == 0])
-
-#Rec.group_BivData_F1post_Ttest$p.value
-
-```
-
-```{r log.reg_SC.length, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE THIS COMPARES FROM FULL BIV DATA
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
-#REMOVE THIS, IT uses total bivalent data
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-Male.cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-Male.cell.av.long.biv.table <- add_mouse(Male.cell.av.long.biv.table)
-Male.cell.av.long.biv.table <- add_strain(Male.cell.av.long.biv.table)
-Male.cell.av.long.biv.table <- add_subsp(Male.cell.av.long.biv.table)
-
-Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
-Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
-Male.cell.av.long.biv.table$strain <- factor(Male.cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
-```
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-
-#+theme(legend.position="none") 
-```
-
-```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( Long.biv_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+
-  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
-
-mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
-
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-```{r Q2.totSC.ttest, echo=FALSE}
-
-#move this to MS_v2?
-
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol -- not enough Mol observations
-#Q2tot.sc_t.test_High.Low_Mol <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-
-#this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-
-glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
-
-sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
-
-#MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
-  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
-sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
-
-#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
-#DOM
-glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
-
-#MUSC
-glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
-
-#MOL
-#within subspecies
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
 #Q2 1CO postion
@@
 #summary(Q2.glm.M4.2_IFD.PER_Mol)
+```
+
+```{r Q2.IFD.glm.test.old, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#M1.ABS
+Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")  )
+sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
+
+#M1.PER
+Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+
+sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
+#summary(Q2.M1glm_IFD.PER)
+
+#M2.raw
+Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
+
+Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
+#  MSM strain effect for 2CO PER not significant
+
+#within subspecies
+#Dom
+Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
+#summary(Q2.glm.M2_IFD_Dom)
+
+#Musc
+Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
+#summary(Q2.glm.M2_IFD_Musc)
+
+#Mol
+Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol") )
+#summary(Q2.glm.M2_IFD_Mol)
+
+#Dom
+Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
+#summary(Q2.glm.M2_IFD.PER_Dom)
+
+#Musc
+Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
+#summary(Q2.glm.M2_IFD.PER_Musc)
+
+#Mol
+Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol")  )
+#summary(Q2.glm.M2_IFD.PER_Mol)
+
 ```
@@
 
-```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE, eval=FALSE}
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus")  )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
-#summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
-#summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol") )
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Dom") )
-#summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Musc") )
-#summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.level.BivData_table %>% filter(subsp == "Mol")  )
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
-```
-
 # SAVE the data

### 2020-06-02 src/SetupData.Rmd
updating the Rdata files, reduce redundancyes

@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
+                        Ncells  = length( unique(fileName)),
+                      n_obs  = length(Obj.ID),
                       mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
@@
                         Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
+                        Ncells  = length( unique(fileName) ),
+                        n_obs  = length(Obj.ID),
 

### 2020-06-04 doc/MS_Figures_v2.rmd
updates from final v2 files

--- 2020-05-14 doc/MS_v1.Rmd
@@
 ---
-title: 'Evolution of Sex-Specific Recombination Rate'
-date: '`r Sys.Date()`'
-author: "April L. Peterson, Bret Payseur"
-output:
-  html_document:
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
-  word_document:
-    toc: yes
-    toc_depth: '5'
-always_allow_html: yes
-bibliography: refs_Main_Report.bib
+title: "Manuscript Figures"
+author: "April Peterson"
+date: "6/2/2020"
+output: html_document
 ---
 
-```{r set.up, echo=FALSE, include=FALSE}
+# setup
+
+```{r set.up, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 library(RLRsim)
-#library(ggpubr)
-library(reshape2)
-library(patchwork)
+library(ggpubr)
+#library(reshape2)
+#library(patchwork)
 library(png)
@@
 
-load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData
-
-#load(file="~./MLH1repo/data/Results_setup_4.21.20.RData")
+load(file="~./MLH1repo/data/Results_setup_5.30.20.RData")
+#load(file="~./MLH1repo/data/Results_setup_5.21.20.RData") #Results_setup_5.14.20.RData
 
@@
 
-#Intro
-
-#Material Methods Draft v1
-
-### Mouse Husbandry
-
-To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used.
-The _Mus musculus_ strains PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA).
-The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were generated from frozen embryos purchased  from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).
-(https://en.brc.riken.jp) and rederived by the UW-Biotech center (Madison UW).
-The related murid species *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jackson labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.
-We use abbreviated  superscript notation for the remainder of the paper (e.g.
-_musculus^WSB^_).
-All mice were housed UW-Madison Biotech and MSC facilities following RARC approved protocols.
-A breeding colony of wild derived _Mus musculus_ mice sampled from Gough Island (G) is maintained at UW Veterinary school facilities.
-
-### Tissue Collection and Immunohistochemistry
-
-The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used MLH1 quantification were from mice between 5 and 12 weeks, while juvenile mice aged between 12 to 15 days were used for DMC1 quantification.
-Both pairs of ovaries were collected from embryo or neonate mice between 16 to 21 embryonic days and 0 to 48 hours old respectively.
-Whole testes and decapsulated ovaries were incubated in 3ml or 300ul of hypotonic solution for 45min respectively.
-Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
-Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a Photoflow wash (Kodak, cite).
-Slides were stored at -20*C if not stained immediately.
-
-To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
-Double strand breaks (DSB) and COs were visualized with DMC1 and MLH1 respectively.
-The staining protocol was based on that in [@anderson1999] and [@koehler2002].
-Antibodies and slide blocking was performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), Triton X-100 (Sigma) ).
-
-Following a 30 minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
-The master mix recipe contained polyclonal anti rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37 degrees.
-Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200)  were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37 degrees.
-Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
-
-### Image Processing
-
-Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and Axio Vision software (Zeiss, Cambridge, UK).
-Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using photoshop.
-
-### Analysis
-
-To estimate the numbers of percursor sites which may lead to crossovers, the mean **DMC1** foci per cell for a single mouse from a subset of srains *musculus^PWD^*, *musculus^MSM^*, *domesticus^WSB^*, and *domesticus^G^* was quantified.
-SC morphology and CREST/centromere foci number was used to stage spermatocytes, in early and late zygotema stages.
-
-**MLH1** foci were quantified from cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes or 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score, between 1 to 5, was assigned to each cell based on staining quality and general spread of bivalents.
-With scores of 1 and 2 having the highest quality.
-Cells and coded as a score of 5 were excluded from the final analysis.
-The distributions of MLH1 counts per cell was assessed for normality (supplemental figure).
-MLH1 foci located on the XY in spermatocytes were excluded from this study.
-
-**M1 Mixed Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex + rand(strain) + \varepsilon $$
-
-**M2 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ subsp * sex * strain + \varepsilon$$
-
-**M3 Linear Model** $$mouse \ av.\ dependant\ variable\ ~=~ sex * strain + \varepsilon$$
-
-**M4 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~ subsp * strain + \varepsilon$$
-
-**M5 Linear Model** $$sex\ specific\ mouse \ av.\ dependant\ variable\ ~=~  strain + \varepsilon$$
-
-A series of mixed models and general linear models were used as an analetical framework.
-We built the full mixed model (**M1**) to interpret the patterns of variation using the lmer() from the lmer4 package (cite) in R (v3.5.2)[@Rstudio].
-Strain was coded as a random effect and tested using exactRLRT() to reflect a random sample of natural genetic variation from each subspecies.
-The main effects, subspecies, strain, and interaction were coded as fixed and tested by a chi-square test comparing the full and reduced models using drop1() and anova().
-
-This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.
-We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism) in a sex specific manner.
-The remaining general linear models, (**M2-M3**), were used to follow up on results from the full mixed model.
-The general linear models, **M4** were used to analyze sex specific data sets.
-
-Our goal was to design an analytical framework which accounts for the nested nature across categories across all the individual mice (e.g.
-sex, strain subspecies) and can be applied to multiple (meiotic traits) dependent variables.
-In addition to the MLH1 count variables for individual mice, mean MLH1 count per cell, variance and coefficient of variance for MLH1 count per cell, we analyzed SC based features: mouse average of total SC per cell, mouse average of single SC length, mouse mean IFD^norm^ and IFD^raw^ for double crossover chromosomes, and the average position of crossover for single crossover bivalents.
-
-These traits were choosen because they are markers of phenomena in the recombinaiton pathway and provide more detailed information about the recombination landscape.
-In order to examine crossover interference, inter-focal distance (IFD) was measured from the single bivalents containing two crossovers.
-Normalized measures were calculated by the position divided by the length of the SC for that single bivalent.
-
-### Single bivalent characterization
-
-To characterize the SC length of individual chromosomes, two image analysis algorithms were used to quantify SC area for individual cells [@wang_2019_sc] and for individual bivalents [@peterson2019].
-Both apply a 'skeletonizing' transformation to the synapsed chromosomes which produces a single pixel wide 'trace' of the original chromosome shape.
-Hand measures of pactyene chromosomes were performed by using ImageJ/Fiji (v1.52) [@fiji].
-
-The **total SC** area per cell was quantified from the pachtyene cell images.
-To remove erroneous SC isolation, outliers were visually assessed at the mouse level and removed from the data set.
-Mouse means were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-
-The **DNA CrossOver** algorithm isolates single straightened bivalents shapes from an image, and returns SC length, and location of green and blue signal, reflecting MLH1 foci and cetromere signal respectively.
-This algorithm substantially speeds the accurate measurement of bivalents, but has the limitation that not all bivalents per cell can be isolated due to overlapping bivalents.
-In this data set, isolation rates per cell (the number of all bivalents per cell) range from `r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
-
-A curation step was applied to the total single bivalent data set to remove poor measures from the algorithm.
-From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` bivalent objects were isolated by the image analysis software.
-After a human curation step , `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-From the curated single bivalent data set the **proportions** of bivalents by crossover number were quantified and tested by chi-square tests with the prop.test().
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we considered a **reduced data** set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-The 'short bivalent' data set included the four or five shortest bivalents, and excluded the X bivalent in oocytes.
-A total of `r sum(Long.biv_mouse_table$n_obs)`   'short' bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$ncells)` spermatocytes.
-Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A 'long bivalent' data set was calculated as above the 4th quartille in SC lengths per cell.
-
-# Results
-
-```{r FOR.PLOT.organizing.data, echo=FALSE, warning=FALSE}
+### Figure 1
+
+```{r female.strain.table, echo=FALSE}
+female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
+#female strain table
+#female_mouse.avs_for.analysis
+F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
+
+                        Nmice = length(unique(mouse)),
+
+                        total.cells  = sum(Ncells, na.rm = TRUE),
+
+                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
+
+                        #Standard error (mice vs cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#F.strain.table
+
+#F.strain.table
+
+# I don't know why most of the table is blank
+```
+
+```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
+
+#variance plots
+musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+musc.female.VAR.plot
+
+Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.female.VAR.plot
+
+Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
+Dom.female.VAR.plot
+
+```
+
+```{r male.strain.table, echo=FALSE}
+
+M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
+                        Nmice =length(unique(mouse)),
+
+                        totNcells  = sum(Ncells, na.rm=TRUE),
+
+                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
+
+                        mean.mouse_cV = mean(cV, na.rm = TRUE),
+
+                        mean.mouse_var = mean(var, na.rm = TRUE ),
+
+                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
+
+                        #Standard error (mice . cells)
+                        mean.mouse_se  = mean(se, na.rm=TRUE)
+    )
+#M.strain.table
+#same issue as the female table .. ugh
+```
+
+```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
+
+#variance plots
+musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+
+```
+
+```{r FOR.PLOT.organizing.data, echo=FALSE, message=FALSE, warning=FALSE}
 #set the order of another column, based on another variable. consistant order for ploting
@@
 #pooled
-strain.table.HQ_Fig.og <- ddply(MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
+strain.table.HQ_Fig.og <- ddply( MLH1_data  %>% filter(quality < 5), c("subsp","strain", "sex"), summarise,
                         Nmice =length(unique(mouse)),
                         Ncells  =length(nMLH1.foci),
-                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
+
+                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 4 ), nsmall=3) ),
+                        mean_co_long = mean(nMLH1.foci),
                         cV = cv(nMLH1.foci, na.rm = TRUE),
                         var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
-                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
+                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 4),
                        #Standard error (mice vs cells)
-                        se  = round(sd / sqrt(Ncells), 3)
+                        se  = round(sd / sqrt(Ncells), 4)
 
@@
 ```{r Fig1.top, echo=FALSE, warning=FALSE, message=FALSE}
+#Mouse.Table_HQ
+#change the segements and text!!
 
@@
 
-female_mouse.av <- Mouse.Table_HQ %>% filter(sex == "female")
+#MLH1_data  %>% filter(quality < 5)
+#mouse avs
+
+female_mouse.av <- Mouse.Table_HQ.OG %>% filter(sex == "female") %>% filter(species == "M.musculus")
 
@@
 
-male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
+male_mouse.avs_for.analysis$strain <- factor(male_mouse.avs_for.analysis$strain, ordered = FALSE, levels =c( "MSM", "MOLF",
+  "PWD", "SKIVE", "KAZ","TOM","AST","CZECH",
+
+  "WSB", "G", "LEW","PERC",
+                                                                         "CAST", "HMI",
+                                                                   "SPRET", "SPIC") )
+
+male.plot <- ggplot( male_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+
 
@@
 B) Female specific MLH1 count distributions.
-C)	Male specific MLH1 counts per cell.', message=FALSE}
-
-#try cowplot for more control of sizing
-library("cowplot")
+C)	Male specific MLH1 counts per cell.', message=FALSE, eval=FALSE}
 
 MLH1.Figure1  <- ggdraw() +
-  draw_plot(Fig1_v1_strain, x = 0, y = .55, width=1, height=.45)+
+  draw_plot(Fig1_v1_strain, x = .02, y = .55, width=1, height=.45)+
 
   draw_plot(female.specific.plot, x=0, y=0, width=.5,height= .55)+
-  draw_plot(femaleMLH1.cell.obj, x=.32, y=.32, width=.19,height= .19)+
+  draw_plot(femaleMLH1.cell.obj, x=.35, y=.35, width=.18, height= .18)+
 
   draw_plot(male.plot, x=.5, y = 0,  width = .5, height = .55)+
-  draw_plot(maleMLH1.cell.obj, x=.85, y = .32,  width = .19, height=.19)+
+  draw_plot(maleMLH1.cell.obj, x=.85, y = .35,  width = .19, height=.19)+
 
   draw_plot_label(label = c("A", "B", "C"), size = 14,
-                  x = c(0, 0, .49), y = c(1, .55, 0.55) )
+                  x = c(0, .01, .51), y = c(.99, .55, 0.55) )
 
@@
 
-ggsave("~/MLH1repo/doc/Figure1.png")
-```
-
-updates for Figure 1, try shifting the lines up for B C
-
-## Genome-wide recombination rate estimates for both sexes
-
-```{r MLH1.summ.stats, echo=FALSE}
-
-#strain.table.HQ_Fig.og  all strains
-#strain.table.HQ_Fig  extra strains removed
-
-main.table <- strain.table.HQ_Fig.og  %>% filter(strain != "F1")
-
-main.table <- add_subsp(main.table)
-main.table <- add_species(main.table)
-
-#main.table %>% filter(sex == "female")
-#main.table %>% filter(sex == "male")
-
-#G.HETC.val #1.2
-#LEW.HETC.val  #1
-#WSB.HETC.val  #1
-#KAZ.HETC.val  #1.1
-
-#SKIVE.HETC.val # 0.96
-#PWD.HETC.val  #0.88
-#MSM.HETC.val #0.9
-
-#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")
-
-#strain.table.HQ_Fig
-
-#sum( (main.table  %>% filter(sex == "male"))$Ncells)
-#the kable version is bulkier than the other versaion
-
-```
-
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice: *M.
-m.
-domesticus*, *M.
-m.
-musculus* and *M.
-m.
-molossinus*.
-Mean MLH1 foci counts for `r length( unique(Mouse.Table_HQ)$mouse )` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells),  format="d", big.mark=",")` spermatocytes) and `r mean(female_mouse.avs_for.analysis$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
-
-Graphical comparisons between the two sexes reveal several patterns (Figure 1A).
-First, recombination rate differs between females and males in most strains.
-Second, the direction and magnitude of heterochiasmy varies among strains.
-Although the majority of strains show higher recombination rates in females (following the pattern in laboratory mice), two musculus strains and one molossinus strain exhibit male-biased heterochiasmy.
-Hence, relative recombination rates in the two sexes are evolutionarily labile.
-
-The general patterns of heterochiasmy in house mouse are displayed in Figure 1A.
-Taking note of the direction and the magnitude of the sex differences, our results confirm two general patterns: i) genome-wide recombination rates averages are greater in females compared to males (female biased heterochiasmy) and ii) the degree of heterochiasmy (Female:Male ratio) is generally low, ranging from `r G.HETC.val` in _domesticus^G^_ to `r LEW.HETC.val` in _domesticus^LEW^_.
-Three notable exceptions of male biased heterochiasmy are the strains _musculus^PWD^_, _molossinus^MSM^_ and _musculus^SKIVE^_, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` respectively.
-
-Separately examining the mouse means of MLH1 foci per cell for each sex points to distinct patterns of variation (Figure 1B-C).
-Female recombination rates are evenly distributed around the sex-wide mean of approximately 25 MLH1 foci per cell (Figure 1B).
-In stark contrast, males separate more clearly into two groups of strains with high rates (near 30 MLH1 foci per cell) and low rates (near 23 MLH1 foci per cell) (Figure 1C).
-
-```{r female.strain.table, echo=FALSE}
-female_mouse.avs_for.analysis <- female_mouse.avs_for.analysis %>% filter(strain != "CAST")
-#female strain table
-#female_mouse.avs_for.analysis
-F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
-
-                        Nmice = length(unique(mouse)),
-
-                        total.cells  = sum(Ncells, na.rm = TRUE),
-
-                        mouse.mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),
-
-                        #Standard error (mice vs cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#F.strain.table
-
-#F.strain.table
-
-# I don't know why most of the table is blank
-```
-
-```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-musc.female.VAR.plot
-
-Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-Mol.female.VAR.plot
-
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
-Dom.female.VAR.plot
-
-```
-
-```{r male.strain.table, echo=FALSE}
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
-                        Nmice =length(unique(mouse)),
-
-                        totNcells  = sum(Ncells, na.rm=TRUE),
-
-                        mean.mouse_mean_co = mean(mean_co, na.rm=TRUE),
-
-                        mean.mouse_cV = mean(cV, na.rm = TRUE),
-
-                        mean.mouse_var = mean(var, na.rm = TRUE ),
-
-                        mean.mouse_sd   = mean(sd, na.rm=TRUE),
-
-                        #Standard error (mice . cells)
-                        mean.mouse_se  = mean(se, na.rm=TRUE)
-    )
-#M.strain.table
-#same issue as the female table .. ugh
-```
-
-```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, include=FALSE, eval=FALSE}
-
-#variance plots
-musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
-
-```
-
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
-#write the full table to a supplemental table
-#show the head (or shortened)
-#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-#kable(strain.table.HQ_Fig, "latex")
-
-df.strain.table <- strain.table.HQ_Fig
-#kable(df.strain.table)
-#pander(df.strain.table)
-
-#kable(strain.table.HQ_Fig, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", "Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain Level Summary Statistics" )  %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-#write excel
-#write.xlsx2(strain.table.HQ_Fig, "Main_MLH1_table.xlsx", sheetName = "Sheet1", 
-#  col.names = TRUE, row.names = FALSE, append = FALSE)
-
-```
-
-## Partitioning variation in recombination rate
-
-To situate variation in recombination rate within an evolutionary framework, we fit a series of models including subspecies, strain, and sex, to mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co)` mice.
-We began with a full mixed model (M1, see Methods), which showed that strain
-(random effect p < `r 1/1000000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p=`r MLH1.MMlmer_results$sub.sp`), and subspecies sex (p = `r MLH1.MMlmer_results$interaction.pval`) each significantly affect recombination rate.
-
-After a general linear model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A general linear model with these two variables (M3) identified two strains with particularly strong effects on recombination rate: _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`)
-and _domesticus^G^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`).
-In addition, two strains exhibit strain-by-sex interactions:  _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
-
-We next fit general linear models separately for `r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4; see methods).
-In the male dataset, three strains significantly affect recombination rate: _musculus^PWD^_ ((glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect =
-`r sum_MLH1.male.glm.M4$coefficients[8]` foci), and 
-_musculus^SKIVE^_ (glm; p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]`; effect = `r sum_MLH1.male.glm.M4$coefficients[10]`), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect  `r sum_MLH1.male.glm.M4$coefficients[9]`).
-
-These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.
-Analysis of the female dataset points to four strains with significant effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and _musculus^PWD^_ (p= `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
-Together, these results demonstrate heritable differences in the genome-wide recombination rate evolving in a highly sex-specific manner over short evolutionary timescales.
-
-## Within mouse variance in CO count per cell
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE}
-
-MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
-MLH1var_M2_sum <- summary(MLH1var_M2)
-
-#anova(MLH1var_M2)  
-#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
-
-MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M2)
-#anova(MLH1cv_M2)
-#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
-
-#Q12 var
-MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
-
-sum.MLH1var_M2_Q12 <- summary(MLH1var_M2_Q12)
-#anova(MLH1var_M2_Q12)  
-#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-
-#summary(MLH1cv_M2_Q12)
-#anova(MLH1cv_M2_Q12)#sex large F
-#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
-sum.glmMLH1Var_M3 <- summary(glmMLH1Var_M3)
-#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-#MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-
-Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate.
-To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count per cell as the dependent variable.
-Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < `r 1/1000000`) and general linear models (M2) (p = `r MLH1var_M2_sum$coefficients[,4][9]`) and M3 (p = `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`).
-
-In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).
-The results are similar: sex is the strongest effect (M1 p < `r 1/1000000`; M2 p = `r  round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)`;  also M3 p = `r MLH1var_M3_Q12_sum$coefficients[,4][2]`).
-When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex.
-These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.
-
-<LmerMLH1_M1_Q12_VAR_results> <Bret's comment for below, are there differences between the curated and the full dataset? -- Q1 and full data set>
-
-## DMC1 Evolution of genome-wide recombination rate is associated with evolution of double strand breaks
-
-```{r DMC1.table, echo=FALSE, message=FALSE}
-##RATIOS
-DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
-DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z
-
-#make category averages from AP_mouse_table.HQ
-mean.MLH1.obs <- ddply(.data=Mouse.Table_HQ,
-                 .(category),
-                 summarize, 
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co)
-)
-
-DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
-                   category == "WSB male" |
-                     category == "KAZ male" |
-                  category == "MSM male" |
-                    category == "PWD male")
-
-#attach this mini DF to the DMC1 data
-DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
-DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count
-
-DMC1table <- merge(DMC1.mean.obs.L, DMC1.mean.obs.Z, by="category")
-
-DMC1table.main88 <- merge(DMC1.target.strains, DMC1table, by="category")
-
-DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
-
-#DMC1.mean.table
-DMC1table.main99 <- subset(DMC1table.main, select= -c(stage.x, var.MLH1, stage.y, mean.count.x, mean.count.y))
-
-DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]
-
-colnames(DMC1table.main99) <- c("category", "mean MLH1",    "ncells",     "mean DMC1", "MLH1:DMC1 ratio", "ncells",     "mean DMC1", "MLH1:DMC1 ratio")
-
-#kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
-#  kable_styling() %>%
-#  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#pander is better for word
-#pander(DMC1table.main99, row.names = F)
-
-#write table
-write.xlsx2(DMC1table.main99, file = "DMC1.Results_Table.xlsx",sheetName = "DMC1.results", append = FALSE, row.names=F)
-
-```
+#ggsave("~/MLH1repo/doc/Figure1.png")
+```
+
+#Figure 2 DCM1
 
@@
 Green foci show DMC1.
-C)	Distribution of DMC1 counts per cell  by strains.", eval=FALSE}
+C)	Distribution of DMC1 counts per cell  by strains.", message=F, warning=F, eval=FALSE}
 
@@
 
-ggsave("~/MLH1repo/doc/Figure2.png")
-
-```
-
-```{r plot.stage.ttest, echo=FALSE, warning=FALSE}
-
-#anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage #!= "P",],method = "t.test") %>% mutate(y_pos = 300)
-
-#DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
-#  geom_boxplot(position=position_dodge()) + 
-#  geom_point(aes(color=rec_group), position=position_jitterdodge()) +
-
-  #facet_wrap(~stage, nrow=1) +
-
-#  ggsignif::geom_signif(
-#    data=anno_df_stage, 
-#    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
-#    manual=TRUE
-#  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-#        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")
-
-ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
-               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )
-
-DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")
-
-ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
-               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )
-
-DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")
-
-ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
-               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
-```
-
-```{r DMC1_t.tests, echo=FALSE, warning=FALSE}
-#2. t-tests for the DMC1 counts
-
-#high vs low group (L)
-ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value =  0.002145
-
-#high vs low group (Z)
-ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
-               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
-# p-value = 0.6628
-
-#dom vs musc (L) (2 subsp vs)
-ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
-# p-value = 0.003016
-
-#dom vs musc (Z)
-ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
-#p-value = 0.09588
-
-#KAZ vs Dom (Z) (only 1 L cell)
-ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
-      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )
-
-#p-value = 0.004148
-#Kaz Z are higher than Dom
-```
-
-```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}
-
-MLH1.DMC1.L_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.L, method = c("pearson"))#MLH1.DMC1.Z_cor
-
-MLH1.DMC1.Z_cor <- cor(DMC1table.main$mean.MLH1, DMC1table.main$mean.count.Z, method = c("pearson"))#0.2793237
-
-#colnames(DMC1table.main) <- c("category", "mean.MLH1", "", "", "")
-WSB_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "WSB male") )$mean.count.L / (DMC1table.main %>% filter(category == "WSB male"))$mean.MLH1
-
-G_DMC1.Z_CO_ratio <-   ( DMC1table.main %>% filter(category == "G male") )$mean.count.L / (DMC1table.main %>% filter(category == "G male"))$mean.MLH1
-
-MSM_DMC1.Z_CO_ratio <- ( DMC1table.main %>% filter(category == "MSM male") )$mean.count.L / (DMC1table.main %>% filter(category == "MSM male"))$mean.MLH1
-
-PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
-
-DMC1table.main$category <- factor(DMC1table.main$category, levels = c("WSB male","G male","KAZ male",               "PWD male","MSM male"), ordered = TRUE )
-
-DMC1.Lratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.L_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.L_ratio        )
-
-DMC1.Zratio_t.test <- t.test(  (DMC1table.main %>% filter(  ( category == "G male" | category == "WSB male" | category == "KAZ male" ) ) )$MLH1.Z_ratio, 
-                (DMC1table.main %>% filter(  ( category == "PWD male" | category == "MSM male") ) )$MLH1.Z_ratio        )
-
-```
-
-In an attempt to localize the male-specific evolution of crossover number to steps of the meiotic pathway, we counted foci from a marker for double strand breaks (DSBs), DMC1, in prophase spermatocytes.
-DMC1 foci were scored from a total of `r sum( (DMC1table.main99$ncells) )` early zygotene-stage and `r sum( (DMC1table.main99$ncells) )` late zygotene-stage spermatocytes from three low-recombination strains ( _musculus^KAZ^_ , _domesticus^WSB^_ , and _domesticus^G^_ ) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_ ).
-The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in early zygotene cells (t-test, p < `r 1/1000000`).
-In contrast, the two strain groups do not differ in DMC1 foci counted in late zygotene cells (t-test, p = `r ttest.HighLow.Z.pval`).
-
-<ttest.HighLow.L.pval>
-
-After DSB formation, DSBs are repaired as either non-crossovers (NCO) or as crossovers (COs), with the vast majority being repaired as NCOs.
-Thus the ratio of CO:DSB is a partial indicator of the proportion of DSBs which are designated as COs.
-The ratios, calculated for DMC1 means from both stages, are not significantly different between the high and low strain groups (t-test, p = `r DMC1.Lratio_t.test$p.value` and p = `r DMC1.Zratio_t.test$p.value` for early zygotene and late zygotene ratios, respectively).
-This comparison raises the possibility that the evolution of crossover number is primarily due to processes that precede the crossover/non-crossover decision.
-@cole2012 suggest that early zygotene counts for DMC1 are most relevant for predicting the final CO number due to the crossover homeostasis process.
-This result, combined with the large difference in MLH1 count between high and low strains, might indicate that the CO/NCO decision contributes to strain differences in recombination rate.
-
-<!-- The number of DSBs increase in zygotema, then deminish as the effects of crossover homeostasis are implemented (Cole).
-(co homeostasis also indecates dsb ) Natural' variation in precursor number (the relationship of DSB and CO number) was analyzed in Barier et al  and links connects the CO number variation to 'pre' DSB formation fitting the model that DSB formation is dependant on the axis-loop structure of chromatin.
-(this study only examined one stage of early prophase for DMC1 counts -->
-
-<!-- the birer et al ratios for strain average DMC1: MLH1 were 5.6 cast. 6.6 C3H, 7.3 B6 -->
-
-## Evolution of genome-wide recombination rate is reflected at the single chromosome level
-
-```{r algorithm_performance_stats, echo=FALSE}
-
-#number of mice, number of cells
-#mean number of isolate bivs per cell,
-#use these two metrics
-#cell.level.BivData_table$boxs.IDd 
-#cell.level.BivData_table$nboxes_passed
-
-#calq the mean rate by category
-
-cell.level.BivData_table$pass.rate <- cell.level.BivData_table$nboxes_passed / cell.level.BivData_table$boxs.IDd
-
-algorithm.stat.mouse <- ddply(cell.level.BivData_table, c("mouse"), summarise,
-                        n_obs = length( unique(fileName) ),
-                         mean_pass.rate = mean(pass.rate)
-                         )
-
-cell.level.BivData_table <- add_sex(cell.level.BivData_table)
-
-algorithm.stat.category <- ddply(cell.level.BivData_table, c("sex", "strain"), summarise,
-                        n_obs = length(unique(fileName)),
-                        mean_pass.rate = mean(pass.rate)
-                         )
-
-#showing the table..
-#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-```{r chrm_prop_data, echo=FALSE, warning=FALSE}
-#shut off eval for now because of wd / path error
-#chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
-
-chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#try adding the BivData version of the proportions
-
-#chrm.class.DF = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/ChrmClassProportions2.csv", header = TRUE)
-
-#this excel file is ~100 cells (10 per category) where I manually verified the number of chrom classes. (this file can also be used to quantify error for CO counts)
-
-chrm.class.DF$fileName <- (chrm.class.DF$Original.Name)
-chrm.class.DF <- add_mouse(chrm.class.DF)
-chrm.class.DF <- add_strain(chrm.class.DF)
-chrm.class.DF <- add_sex(chrm.class.DF)
-chrm.class.DF <- add_category(chrm.class.DF)
-#chrm.class.DF <- add_euth_date(meta.data.file2)
-#chrm.class.DF <- add_age(meta.data.file2)
-
-#table(chrm.class.DF$category)
-#str(chrm.class.DF)
-
-#summarize by category
-proportion_stats_category  <- ddply(.data=chrm.class.DF, 
-                 .(category),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                  tot_biv  = sum(biv.sum.check,  na.rm = TRUE),
-
-                 n0CO = sum(n0CO, na.rm = TRUE),
-                  n1CO = sum(n1CO,na.rm = TRUE),
-                  n2CO = sum(n2CO,na.rm = TRUE),
-                  n3CO = sum(n3CO,na.rm = TRUE),
-                  n4CO = sum(n4CO,na.rm = TRUE),
-
-                 per0CO = sum(n0CO) / tot_biv,                 
-                 per1CO= sum(n1CO)/ tot_biv,
-                 per2CO= sum(n2CO)/ tot_biv,
-                 per3CO= sum(n3CO)/ tot_biv,
-                 per4CO= sum(n4CO)/ tot_biv,
-
-                 mean.MLH1.foci = mean(MLH1.sum.check),
-                 mean.MLH1.foci2 = mean(nMLH1.foci),
-                 meanQ = mean(quality)
-
-)
-#this is the table that needs melting
-#but is also easy to redo by hand
-
-#write.table(proportion_stats_category, "~./MLH1repo/prop_category.csv", sep=",", row.names = FALSE)
-```
-
-```{r plot.chrm.prop.melt, echo=FALSE, warning=FALSE}
-
-#add subspecies to these values¬
-melt.Chrm.Prop.table = read.csv("~./MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-
-#melt.Chrm.Prop.table = read.csv("MLH1repo/data/ChrmClassProportions2_melt.csv", header = TRUE)
-melt.Chrm.Prop.table <- melt.Chrm.Prop.table[c(1:5)]
-
-#add the subsps
-melt.Chrm.Prop.table$strain<- factor(melt.Chrm.Prop.table$strain, levels =c("LEW",
-                      "WSB","G", "TOM","AST","KAZ", "SKIVE","PWD","MSM", "MOLF"), order=T )
-
-melt.Chrm.Prop.table <- with(melt.Chrm.Prop.table, melt.Chrm.Prop.table[order(strain),])
-
-chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))+theme(legend.position="none")
-
-#need a second DF / the first df
-#+ annotate() text 
-#annotate()text
-
-#proportion_stats_category,  tot_biv, ncells
-#add number of cells and bivalents in the corner
-
-#chrm.class.barplot
-```
-
-```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}
-
-#prop.test(x, y)  x number of successes, y total  
-#strain sums of n
-#look up continuity correction
-
-#mouse.level.BivData_table <-- has more male strains
-#Mouse.Table_BivData_4MM <- only sex matched strains
-
-#perform test across strain 
-#removed all lines with no hand.foci.count entry
-Chrm.prop_strain_table <- ddply(mouse.level.BivData_table, c("strain","sex"), summarise,
-
-                  nbiv = sum(tot.biv ),              
-                 n0CO = sum(n0CO ),
-                  n1CO = sum(n1CO ),
-                 n2CO = sum(n2CO ),
-                  n3CO = sum(n3CO ),
-                 n4CO = sum(n4CO )
-                          )
-
-Chrm.prop_strain_table <- add_subsp(Chrm.prop_strain_table)
-
-for(g in 1:length(Chrm.prop_strain_table$nbiv)){
-
-  Chrm.prop_strain_table$sum2[g] <- sum(c(Chrm.prop_strain_table$n0CO[g],
-                                     Chrm.prop_strain_table$n1CO[g],
-                                     Chrm.prop_strain_table$n2CO[g],
-                                     Chrm.prop_strain_table$n3CO[g] ))  
-}
-
-Chrm.prop_strain_table$dcheck <- Chrm.prop_strain_table$nbiv - Chrm.prop_strain_table$sum2
-#there are still some wacky entries -- but much fewer than before
-
-chrm.prop.strain.female <-  Chrm.prop_strain_table %>%  filter(sex == "female")
-
-female.0CO.prop.test <- prop.test(chrm.prop.strain.female$n0CO, chrm.prop.strain.female$sum2 )
-female.1CO.prop.test <- prop.test(chrm.prop.strain.female$n1CO, chrm.prop.strain.female$sum2 )
-female.2CO.prop.test <- prop.test(chrm.prop.strain.female$n2CO, chrm.prop.strain.female$sum2 )
-#the pvalues for 0 are not sig dif
-#the p values for 1 and 2CO are sig diff
-
-#seperate by subsp
-chrm.prop.female.dom <-  chrm.prop.strain.female %>%  filter(subsp == "Dom")
-chrm.prop.female.musc <-  chrm.prop.strain.female %>%  filter(subsp == "Musc")
-chrm.prop.female.mol <-  chrm.prop.strain.female %>%  filter(subsp == "Mol")
-
-Dom_female.0CO.prop.test <- prop.test(chrm.prop.female.dom$n0CO, chrm.prop.female.dom$sum2 )
-Dom_female.1CO.prop.test <- prop.test(chrm.prop.female.dom$n1CO, chrm.prop.female.dom$sum2 )
-Dom_female.2CO.prop.test <- prop.test(chrm.prop.female.dom$n2CO, chrm.prop.female.dom$sum2 )
-Dom_female.3CO.prop.test <- prop.test(chrm.prop.female.dom$n3CO, chrm.prop.female.dom$sum2 )
-#WSB has fewer 2COs than other 3 strains
-
-Musc_female.0CO.prop.test <- prop.test(chrm.prop.female.musc$n0CO, chrm.prop.female.musc$sum2 )
-Musc_female.1CO.prop.test <- prop.test(chrm.prop.female.musc$n1CO, chrm.prop.female.musc$sum2 )
-Musc_female.2CO.prop.test <- prop.test(chrm.prop.female.musc$n2CO, chrm.prop.female.musc$sum2 )
-#KAZ has slightly more 1CO than the other two
-
-#high.female_G.LEW.
-chrm.prop.female.high <-  chrm.prop.strain.female %>%  filter(strain == "G" | strain == "LEW")
-
-high_female.0CO.prop.test <- prop.test(chrm.prop.female.high$n0CO, chrm.prop.female.high$sum2 )
-high_female.1CO.prop.test <- prop.test(chrm.prop.female.high$n1CO, chrm.prop.female.high$sum2 )
-high_female.2CO.prop.test <- prop.test(chrm.prop.female.high$n2CO, chrm.prop.female.high$sum2 )
-
-#Low.female.G.LEW.
-chrm.prop.female.low <-  chrm.prop.strain.female %>%  filter(strain != "G" | strain != "LEW")
-
-low_female.0CO.prop.test <- prop.test(chrm.prop.female.low$n0CO, chrm.prop.female.low$sum2 )
-low_female.1CO.prop.test <- prop.test(chrm.prop.female.low$n1CO, chrm.prop.female.low$sum2 )
-low_female.2CO.prop.test <- prop.test(chrm.prop.female.low$n2CO, chrm.prop.female.low$sum2 )
-
-#all females
-
-chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
-male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
-
-male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
-male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
-#the pvals for these 3 class are sig dif
-
-#seperate by subsp
-chrm.prop.male.dom <-  chrm.prop.strain.male %>%  filter(subsp == "Dom")
-chrm.prop.male.musc <-  chrm.prop.strain.male %>%  filter(subsp == "Musc")
-chrm.prop.male.mol <-  chrm.prop.strain.male %>%  filter(subsp == "Mol")
-
-Dom_male.0CO.prop.test <- prop.test(chrm.prop.male.dom$n0CO, chrm.prop.male.dom$nbiv )
-Dom_male.1CO.prop.test <- prop.test(chrm.prop.male.dom$n1CO, chrm.prop.male.dom$nbiv )
-Dom_male.2CO.prop.test <- prop.test(chrm.prop.male.dom$n2CO, chrm.prop.male.dom$nbiv )
-#non of the Dom proportions are different
-
-Musc_male.0CO.prop.test <- prop.test(chrm.prop.male.musc$n0CO, chrm.prop.male.musc$nbiv )
-Musc_male.1CO.prop.test <- prop.test(chrm.prop.male.musc$n1CO, chrm.prop.male.musc$nbiv )
-Musc_male.2CO.prop.test <- prop.test(chrm.prop.male.musc$n2CO, chrm.prop.male.musc$nbiv )
-#all v different
-
-Mol_male.0CO.prop.test <- prop.test(chrm.prop.male.mol$n0CO, chrm.prop.male.mol$nbiv )
-Mol_male.1CO.prop.test <- prop.test(chrm.prop.male.mol$n1CO, chrm.prop.male.mol$nbiv )
-Mol_male.2CO.prop.test <- prop.test(chrm.prop.male.mol$n2CO, chrm.prop.male.mol$nbiv )
-
-#high vs low
-chrm.prop.strain.male$Rec.group <- ifelse(grepl("PWD", chrm.prop.strain.male$strain), 1, 
-                              ifelse(grepl("MSM", chrm.prop.strain.male$strain), 1,
-                                 ifelse(grepl("SKIVE", chrm.prop.strain.male$strain), 1, 0)))
-
-chrm.prop.male.low <-  chrm.prop.strain.male %>%  filter(Rec.group == 0)
-chrm.prop.male.low2 <-  chrm.prop.strain.male  %>%  filter(Rec.group == 0 | strain == "SKIVE") #cool this worked
-
-chrm.prop.male.high2 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1) %>%  filter(strain != "SKIVE")
-chrm.prop.male.high1 <-  chrm.prop.strain.male %>%  filter(Rec.group == 1)
-
-Low_male.0CO.prop.test <- prop.test(chrm.prop.male.low$n0CO, chrm.prop.male.low$nbiv )
-Low_male.1CO.prop.test <- prop.test(chrm.prop.male.low$n1CO, chrm.prop.male.low$nbiv )
-Low_male.2CO.prop.test <- prop.test(chrm.prop.male.low$n2CO, chrm.prop.male.low$nbiv )
-#the low strains do vary slightly in their 1CO and 2CO proportions
-
-#includes SKIVE
-Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
-Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
-Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
-#all very sig p values
-
-High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
-High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
-High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
-#these proportions are v different
-
-#no SKIVE
-High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
-High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
-High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
-#MSM and PWD only slightly sig dif 2CO
-
-#MSM-MOLF
-MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )
-
-#PWD-KAZ
-PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )
-
-PWD.MSM.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "MSM") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "MSM") )$nbiv )
-
-```
+#ggsave("~/MLH1repo/doc/Figure2.png")
+
+```
+
+# Figure 3
+
+# Fig 3
+
+```{r Q1.SC.show.short.biv, echo=FALSE, eval=TRUE}
+
+#Short.biv_mouse_table
+
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(), axis.line = element_line(colour = "black"),  legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=7), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9) )+
+
+  labs(y = "Short SC Length")
+
+ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")
+
+```
+
+```{r Q1.tot.SC.plot, echo=FALSE, warning=FALSE}
+# fig.width = 4, fig.height=2.5
+
+Q1.totSC.mouse.av.table <-  mouse.total.skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")
+
+Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
+   "WSB","G","LEW","PERC", 
+  "PWD", "SKIVE",  "KAZ",
+  "MSM", "MOLF"), order=T )
+
+Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
+
+#order the strains
+Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
+  arrange(strain) %>%
+  mutate(mouse = factor(mouse))
+
+Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
+
+Q1.totSC.mouse.av.table$subsp2 <- 
+   ifelse(grepl("Dom", Q1.totSC.mouse.av.table$subsp), "M. m. domesticus", 
+        ifelse(grepl("Musc",Q1.totSC.mouse.av.table$subsp), "M. m. musculus",
+            ifelse(grepl("Mol",Q1.totSC.mouse.av.table$subsp), "M. m. molossinus",   
+               "other")))
+
+#change this plot to look like the short bivalent plot -- boxplots + mouse points
+
+Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
+  scale_color_manual(values=colors_of_strains)+
+    facet_wrap(~subsp, scales="free", ncol = 3) +
+
+  geom_jitter()+ geom_boxplot(alpha=.2)+ 
+  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
+    axis.title.x=element_blank(),
+axis.text.x = element_text(size=7), 
+legend.position="none",axis.text.y = element_text(size=9), # axis.text.y = element_blank()
+axis.title.y = element_text(size = 10) ) +
+
+  labs(y = "Total SC Length")+ylim(c(1000,2600 ))+ggtitle("")
+
+ggsave("~/MLH1repo/doc/Figure3_Tot.SC_mouse.av_ALL.png")
+
+#ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")
+
+```
+
+```{r Fig3.load.icons, echo=FALSE}
+# load in the icons
+
+all.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/all_yellow_binary.png')
+single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+#EMPTY GGPLOT
+all.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(all.chrm.yellow) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+single.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(single.chrm.yellow) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+    axis.title.x=element_blank(),
+   axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
+```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8, fig.cap= 'caption', eval=FALSE}
+
+Q2.SC_dif.Figure3  <- ggdraw() +
+
+  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
+  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
+
+  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
+  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
+
+  #drawing triangles in the corner
+  #top right-triangle
+  draw_line(
+    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
+    y = c( .25, .25,  .15, .25),
+    color = "blue", size=.75
+  ) +
+
+  #bottom right-triangle
+  draw_line(
+    x = c( .1, .2, .2, .1 ),
+    y = c( .1,  .2, .1, .1),
+    color = "black", size=.75
+  ) +
+
+  draw_plot_label(label = c("A", "B", "C"), size = 14,
+                  x = c(0, .49, 0), y = c(1,1, 0.33) )
+
+```
+
+```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice.
+A)	 Mean total SC area per mouse across subspecies and recombination groups.
+High- and low-recombining groups are indicated by 1 and 0 respectively.
+(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell", echo=FALSE}
+
+#make plot of mouse averages for MLH1 and total SC
+bep <- merge(  male_mouse.avs_for.analysis,   mouse.total.skel)
+
+X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
+  scale_color_manual(values=colors_of_strains)+theme(legend.position="none")+labs(x="Mouse Mean Total SC", y="Mouse Mean MLH1 count")
+
+#plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
+
+```
+
+# Supplemental Figures
+
+## Fig Chrm Proportion plot
 
@@
 
-chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),
-                             aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(legend.position="top", legend.title = element_blank(),
-                                                                                                           axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
-
-kk <- addSmallLegend(chrm.class.barplot)
-
-```
-
-```{r Chrm.Prop.misc.calulations, echo=FALSE}
-
-total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-
-#sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
-#sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
-
-#MELT_Chrm.Proportion$Rec.group <- ifelse(grepl("PWD male", MELT_Chrm.Proportion$category), 1, 
-#                              ifelse(grepl("MSM male", MELT_Chrm.Proportion$category), 1,
-#                                ifelse(grepl("SKIVE male", MELT_Chrm.Proportion$category), 1, 0)))
-
-#Chrm.prop.male.low <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 0)
-#Chrm.prop.male.HIGH <- MELT_Chrm.Proportion %>% filter(sex=="male") %>% filter(Rec.group == 1)
-
-```
-
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( total.biv.all, format="d", big.mark=",")`) have either one or two crossovers (Figure 3).
-The proportion of one-crossover (1CO) to two-crossover (2CO) bivalents distinguishes the high vs.
-low recombining strains (Figure 3).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in musculus^SKIVE^, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_.
-Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-_musculus^KAZ^_ p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_ p = `r MSM.MOLF.2CO.prop.test$p.value`).
-Next, we use this single-bivalent dataset to focus on aspects of the recombination landscape along chromosomes.
-We address two main questions.
-First, which traits are sexually dimorphic?
-Second, which traits differ between males from high-recombination vs.
-low-recombination strains?
-
-## Sex Differences in Meiotic Traits
-
-```{r Q1.Short.Long_biv.setup, echo=FALSE, eval=FALSE}
-#moved this to setupdata
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-#Long.biv_mouse_table
-
-#make the Q1 strain list
-Q1.real.short_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-Q1.real.long_mouse_table <- Long.biv_mouse_table %>% filter( (strain != "CZECH") & (strain != "MOLF") )
-
-```
-
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-#nested boxplots, pooled Bivalents
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
-```{r Q1.SC.show.short.biv, echo=FALSE, eval=TRUE}
-
-#Short.biv_mouse_table
-
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(), axis.line = element_line(colour = "black"),  legend.position="none", axis.title.x = element_blank(), axis.text.x = element_text(size=7), axis.title.y = element_text(size = 10), axis.text.y = element_text(size=9) )+
-
-  labs(y = "Short SC Length")
-
-```
-
-```{r Q1_SC.biv.short_ttest, echo=FALSE}
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE - this value used in the text
-#only one SKVIE female, this is breaking the test
-#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
-
-#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
-#      )
-
-```
-
-```{r Q1SC_short.biv_dif, echo=FALSE, eval=TRUE}
-
-#calq the different in short.biv means across sexes
-#Long.biv_mouse_table
-
-#WSB
-Q1SC.mus.av.short.biv_DIFF_WSB <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC )
-
-#LEW
-Q1SC.mus.av.short.biv_DIFF_LEW <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC )
-
-#G
-#Q1SC.mus.av.short.biv_DIFF_G <- mean( (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC) / mean(  (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC )
-
-#PWD
-Q1SC.mus.av.short.biv_DIFF_PWD <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC )
-
-#KAZ
-Q1SC.mus.av.short.biv_DIFF_KAZ <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC )
-
-#SKIVE
-Q1SC.mus.av.short.biv_DIFF_SKIVE <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC )
-
-#MSM
-Q1SC.mus.av.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean_SC) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC )
-
-#this bivData ratio is the same as mouse average
-Q1SC.BivData.short.biv_DIFF_MSM <- mean( (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$chromosomeLength) / mean(  (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$chromosomeLength )
-
-```
-
-```{r Q1.tot.SC.plot, echo=FALSE, warning=FALSE}
-# fig.width = 4, fig.height=2.5
-
-Q1.totSC.mouse.av.table <-  totSC.mouse.av.table ##%>% filter(species == "")
-
-Q1.totSC.mouse.av.table$strain <- factor(Q1.totSC.mouse.av.table$strain,levels =c(
-   "WSB","G","LEW","PERC", 
-  "PWD", "SKIVE",  "KAZ",
-  "MSM", "MOLF"), order=T )
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-#order the strains
-Q1.totSC.mouse.av.table <- Q1.totSC.mouse.av.table %>%
-  arrange(strain) %>%
-  mutate(mouse = factor(mouse))
-
-Q1.totSC.mouse.av.table <- with(Q1.totSC.mouse.av.table, Q1.totSC.mouse.av.table[order(sex, strain),])
-
-Q1.totSC.mouse.av.table$subsp2 <- 
-   ifelse(grepl("Dom", Q1.totSC.mouse.av.table$subsp), "M. m. domesticus", 
-        ifelse(grepl("Musc",Q1.totSC.mouse.av.table$subsp), "M. m. musculus",
-            ifelse(grepl("Mol",Q1.totSC.mouse.av.table$subsp), "M. m. molossinus",   
-               "other")))
-
-#change this plot to look like the short bivalent plot -- boxplots + mouse points
-
-Tot.SC_mouse.av_ALL <- ggplot(data= Q1.totSC.mouse.av.table, aes(y = mean.skel, x=sex, color=strain))+
-  scale_color_manual(values=colors_of_strains)+
-    facet_wrap(~subsp, scales="free", ncol = 3) +
-
-  geom_jitter()+ geom_boxplot(alpha=.2)+ 
-  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
-
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
-    axis.title.x=element_blank(), axis.ticks.x=element_blank(),
-axis.text.x = element_text(size=7), 
-legend.position="none", axis.text.y = element_text(size=9), axis.title.y = element_text(size = 10) ) +
-
-  labs(y = "Total SC Length")+ylim(c(1000,2600 ))+ggtitle("")
-
-#italizing the facet label semms too hard -- do it in the final AI version
-#levels(powerDF$alph) <- c("italic('p=')*0.001", "italic('p=')*0.01", "italic('p=')*0.05")
-
-```
-
-```{r Fig3.load.icons, echo=FALSE}
-# load in the icons
-
-all.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/all_yellow_binary.png')
-single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-#EMPTY GGPLOT
-all.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(all.chrm.yellow) +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-single.chrm.icon <- ggplot(empty.DF, aes(x = col1, y = col2)) +
-  background_image(single.chrm.yellow) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-    axis.title.x=element_blank(),
-   axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
-                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
-
-```
-
-```{r Fig3.example.double.triangle, echo=FALSE}
-#load an example two triangle plot
-
-single.chrm.yellow <- readPNG('~/MLH1repo/doc/figureFiles/single_chrm_yellow.png')
-
-empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
-
-```
-
-```{r Q1.Two.SC.plots, echo=FALSE, warning=FALSE, fig.width=8, fig.cap= 'caption'}
-
-#use cowplot
-library(cowplot)
-
-Q2.SC_dif.Figure3  <- ggdraw() +
-
-  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
-  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
-
-  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
-  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
-
-  #drawing triangles in the corner
-  #top right-triangle
-  draw_line(
-    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
-    y = c( .25, .25,  .15, .25),
-    color = "blue", size=.75
-  ) +
-
-  #bottom right-triangle
-  draw_line(
-    x = c( .1, .2, .2, .1 ),
-    y = c( .1,  .2, .1, .1),
-    color = "black", size=.75
-  ) +
-
-  draw_plot_label(label = c("A", "B", "C"), size = 14,
-                  x = c(0, .49, 0), y = c(1,1, 0.33) )
-
-#two.Q1.SC.plots <- {Q1.SC.short.mouse_scatter_boxplot + {Tot.SC_mouse.av_ALL} + plot_layout(ncol = 2, widths = c(1,1), heights = c(2,2)) }
-
-#two.Q1.SC.plots + plot_annotation(
-#  title = '',
-#  subtitle = '') + plot_annotation(tag_levels = 'A')
-
-```
-
-Make a figure for the diagram.  Import the straightened bivalents.. Load in the wsb double triangle.
-
-```{r Q1.total_SC.table, echo=FALSE}
-
-totSC.strain.av.table <-  totSC.strain.av.table[,c("subsp","strain", "sex","nmice","ncells","mean.skel"),  drop=T]
-
-#head of table
-#see the DMC1 code
-
-#kable(totSC.strain.av.table, row.names = F, caption = "caption") %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#add variance
-#write.xlsx(totSC.strain.av.table, file = "Results_Table.xlsx", sheetName = "Total.SC.results", append = TRUE, row.names = F)
-
-```
-
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#ALL
-Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)
-
-#dom
-Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#Musc
-Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
-#cell level total SC
-#MLH1.merge.Skel   $skel_size
-#error not enought female observations, ONLY 1 female SKIVE mouse
-
-Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
-           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-```{r Q1.totSC_male.first.comp, echo=FALSE}
-
-Q1.male.Dom.Musc.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
-```
-
-```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-
-```
-
-In many mammalian species, the synaptonemal complex (SC) is longer in females [@cahoonLibuda2019, @sardell_sex_2020].
-Yet, the majority of these observations come from species with female-biased heterochiasmy.
-Our wider survey of recombination in house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is male-biased.
-In addition, if SC length is a strong determinant of the genome-wide recombination rate,  male _musculus^PWD^_ and male _molossinus^MSM^_ should have longer SC lengths than the other strains.
-
-We use the reduce data set of short bivalents to remove the effects of the paired X-X bivalent in females.
-The mouse average of short bivalents are significantly longer in females than males in all strains (t-test; p < 0.05) except _musculus^SKIVE^_, which my be attributed to low sample sizes (p = r Q1_mus.av_shortSC_t.test.SKIVE$p.value).
-The ratios of the mouse mean SC across sexes range across strains from `r Q1SC.mus.av.short.biv_DIFF_MSM` ( _musculus^MSM^_ ) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_ ).
-
-Females have significantly longer total SC than males in each strain tested separately (t-test; p < 0.05).
-That females have longer SCs is also supported by mixed models and general linear models with covariates, which identify sex as the most consistently significant effect (p < 0.05).
-Additionally, there are some significant subspecies and strain effects (p < 0.05), indicating that SC length has evolved among strains and subspecies, however further investigation is beyond the scope of this paper.
-
-In summary, two approaches for measuring and analyzing SC length indicate that females have longer SCs (chromosome axes), even in strains where males have more MLH1 foci per cell.
-Furthermore, joint consideration of MLH1 foci and total SC length suggests that males from high-recombination strains have less "space" in which to place their additional crossovers compared to females.
-
-```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
-#total-SC
-#Q1.lmer_totSC.mouse.av.table
-
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-
-#subsp effect, many strain, #male
-#G * male v.sig and male*LEW
-
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
-sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
-#LOT OF SIGNIFICant strain effects in this model
-```
-
-```{r Q1.SC_short_M2, echo=FALSE, warning=FALSE, message=FALSE}
-
-Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
-Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
-Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M2)
-
-summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
-
-```
-
-```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Short.biv_mouse_table)
-
-#summary(Q1.glm_short.SC_M3)
-
-summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
-
-#round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
-#sex male very sig -- only sig
-```
-
-```{r Q1.SC.Q12, echo=FALSE}
-
-#make comparisons for the high quality cells
-#get Q12 cells from MLH1.merge.Skel
-#calq the mouse averages
-
-MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)
-
-totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(mouse),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-
-totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
-totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)
-
-#unorder the factors
-totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
-totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
-totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
-totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)
-
-#add rec group
-totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
-                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
-                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))
-
-#REMOVE JUVI mice
-tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
-
-totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
-                 .(strain, sex),
-                 summarize, 
-                 ncells = length(unique(Original.Name)),
-                 nmice = length(unique(mouse)),
-                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
-                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),
-
-                 mean.skel = mean(skel_size, na.rm = TRUE),
-                #se for skel
-
-                 mean.bin =  mean(bin_size, na.rm = TRUE)
-
-)
-#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
-#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
-totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)
-
-#unorder the factors
-totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
-totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
-totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
-totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)
-
-```
-
-```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}
-
-Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Dom
-#Tot.SC_mouse.av_Dom
-
-Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-
-#Q12.Tot.SC_mouse.av_Musc
-#Tot.SC_mouse.av_Musc
-
-Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(~sex, scales="free", ncol = 3)+
-
-  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+
-
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")
-```
-
-```{r show.long.biv, echo=FALSE, include=FALSE}
-
-#Long.biv_mouse_table
-
-#mouse av
-Q1.SC.long.mouse_scatter_boxplot <- ggplot( Long.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")
-
-```
-
-```{r Q1_SC.biv.long_ttest, echo=FALSE, include=TRUE}
-
-#Q1.real.long_mouse_table
-
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_longSC_t.test.G <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_longSC_t.test.LEW <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_longSC_t.test.PWD <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_longSC_t.test.WSB <- t.test((Long.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Long.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#not enough y observations
-#SKIVE
-#Q1_mus.av_longSC_t.test.SKIVE <- t.test((Long.biv_mouse_table %>% filter(sex == #"male") %>% filter(strain == "SKIVE") )$mean_SC,
-#                           (Long.biv_mouse_table %>% filter(sex == "female") %>% #filter(strain == "SKIVE") )$mean_SC)
-
-#MSM
-#Q1_mus.av_longSC_t.test.MSM <- t.test((Q1.real.long_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
-#                           (Q1.real.long_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)
-
-#run t.tests for the mouse av. for reduced single bivalent datasets
-```
-
-```{r Q1other_SC_lmer, echo=FALSE, include=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}
-
-#removing these plots for space
-
-Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc") 
-#Musc.all
-Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data = Musc.mouse.av)
-
-#BivData Level
-Musc_BivData <-  Curated_BivData %>%  filter(subsp == "Musc")
-
-Q1.SClength_Musc.BivData_plot <- ggplot( Musc_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Musc av.
-SC Length \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-#BivData Level
-Dom_BivData <-  Curated_BivData %>%  filter(subsp == "Dom")
-
-Q1.SClength_Dom.BivData_plot <- ggplot( Dom_BivData, aes(y = chromosomeLength, x = strain, color=strain)) + geom_jitter(alpha=.3)+ geom_boxplot(alpha=.3)+scale_color_manual(values=colors_of_strains)+ggtitle("Dom SC Lengths \nBivData Level")+facet_wrap(~sex, scales="free")+ylim(c(30,225))
-
-```
-
-### Positions of Single Crossovers
-
-```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
-#move to setup
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
-```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
-#main figure of density plots,
-MELT_Curated_BivData <- add_sex(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_strain(MELT_Curated_BivData)
-MELT_Curated_BivData <- add_subsp(MELT_Curated_BivData)#subsp, needs strain
-
-#clean up all the non 1 to 4 hand.foci counts
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData[MELT_Curated_BivData$hand.foci.count < 2, ]
-
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[MELT_Curated_BivData.1CO$PER.Foci.Pos < 1.01, ]
-
-# remove the na's
-MELT_Curated_BivData.1CO <- MELT_Curated_BivData.1CO[!(is.na(MELT_Curated_BivData.1CO$hand.foci.count) | MELT_Curated_BivData.1CO$hand.foci.count==""), ]
-
-CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
-
-#category specific plots, don't run for now.
-Dom.female_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "female") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-CO.Pos.density_Dom_female <- ggplot(Dom.female_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom female \nNormalized Foci Positions")
-#Musc male
-Musc.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Musc") ), ]
-
-CO.Pos.density_Musc_male <- ggplot(Musc.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male \nNormalized Foci Positions")
-#Dom male
-Dom.male_MELT_curBivData <- MELT_Curated_BivData[ ((MELT_Curated_BivData$sex == "male") & (MELT_Curated_BivData$subsp == "Dom") ), ]
-
-CO.Pos.density_Dom_male <- ggplot(Dom.male_MELT_curBivData, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male \nNormalized Foci Positions")
-```
-
-```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
-
-Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \n all bivalents")
-
-```
-
-```{r Q1_1COpos.t.test, echo=FALSE}
-#testing,
-#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents
-
-Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)
-
-#WSB
-Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
-#G
-Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
-#LEW
-Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)
-
-#PWD
-Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)
-
-#KAZ
-Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)
-
-#SKIVE
-#Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain == "SKIVE") )$norm.first.foci.pos,
- #          (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$norm.first.foci.pos )
-
-#MSM
-Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
-                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)
-
-#Single Biv Level
-#single.biv.1CO_pos_t.test <- t.test( 
- # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)], 
-#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
-#  )
-
-#t.tests for mous.av_nrm.1CO.pos 
-```
-
-```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
-
-lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )
-
-lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
-lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]
-
-lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
-lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value
-
-lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
-sex.results = lmer.Q1.nrm1CO_M1.sex , 
-               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)
-
-#lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
-
-```
-
-```{r Q1.CO1pos_glm.M2, echo=FALSE}
-
-#in case factors aren't ordered
-#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
-#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
-#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
-
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_F1pos_M2)
-
-summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
-#round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
-
-#DOM
-Q1.glm_nrmF1pos_M2_Dom <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q1.glm_nrmF1pos_M2_Dom)
-
-#MUSC
-Q1.glm_nrmF1pos_M2_Musc <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q1.glm_nrmF1pos_M2_Musc)
-
-#MOL
-Q1.glm_nrmF1pos_M2_Mol <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q1.glm_nrmF1pos_M2_Mol)
-
-```
-
-```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
-
-#summary(Q1.glm_1CO.pos_M3)
-
-summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
-#round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)
-#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
-
-```
-
-We used the single crossover bivalent (1CO) position to compared general location of crossovers while controlling for differences in total SC length.
-Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.
-
-In all strains, the landscape across 1CO bivalents is significantly different in females and males.
-Normalized foci positions tend to be more central in females `r mean( (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos, na.rm = TRUE )` and closer to the telomere in males `r mean( (mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos, na.rm = TRUE )` (t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
-
-Sex is also the most significant effect on focus position in a mixed model (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`) and in general linear models (M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-These sex differences in the placement of foci on 1CO bivalents follow a pattern observed across a variety of mammalian species [@sardell_sex_2020].
-
-```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
-#real.long.bivData
-Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-
-Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
-#REAL.short_mouse_table
-#REAL.long_mouse_table
-#mouse.av.long.biv.table
-
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nLong Biv")
-
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
-\nNormalized 1CO Foci pos \nShort Biv")
-```
-
-## Sex Differences in CO Interference
-
-```{r Q1.ggplot_IFD.triangle, echo=FALSE}
-#IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
-#  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-#  geom_jitter(alpha=.4)+
-#  ggtitle("Normalized IFD\nDom")+
-#  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
-#  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)
-
-```
-
-```{r double.IFD.triangle.plots, echo=FALSE}
+chrm.class.barplot <- ggplot( data = (chrm.class.category_MELT %>% filter(strain != "CZECH")),aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position="top", legend.title = element_blank(),  axis.title.y = element_blank(), axis.text.x = element_text(size=9), axis.title.x = element_blank() )+ scale_fill_discrete(name = "Chromosome Class", labels =c('0CO','1CO','2CO','3CO','4CO') )
+
+chrm.class.barplot.final <- addSmallLegend(chrm.class.barplot)
+
+```
+
+```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4, eval=FALSE}
+#Curated_BivData_male
+
+#boxplots of nrm F1
+male.bivdata.1CO <- cleanCurated_BivData %>% filter(sex == "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
+
+Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.23 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
+
+```
+
+## IFD Single
+
+## Double triangle
+
+```{r double.IFD.triangle.plots, echo=FALSE, eval=FALSE}
 
@@
 
-  top.triangle2(my.y.top, my.x.top, my.data)
-
-  bottom.triangle2(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
+  top.triangle(my.y.top, my.x.top, my.data)
+
+  bottom.triangle(my.y.bottom+x.dist, my.x.bottom, my.data) #x.dist
 
@@
 
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
-
-#each point is a 2CO, with the F1 and F2 position 
-#based on plot from Dourad et al (Mercier)
-
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
-    aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.5)+
-  ggtitle("Raw IFD WSB")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#trying to move x axis to top
-#cowplot might be the easier way -- since kinda diffecult in ggplot
-#or ggdraw
-
-P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
-```
-
-```{r Q1.IFD_Ttests, echo=FALSE, warning=FALSE, message=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
-#p-value = 0.07
-
-Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
-#p-value = 1e-12
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)
-
-#KAZ
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)
-
-#SKIVE
-#Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-#Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-#                                ( mouse.avs_4MM %>% filter(sex == "female") %>% #filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-#MSM
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#FULL all strains
-Q1.IFD.raw_sansG_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.avs_4MM %>% filter(sex == "female") %>%  filter(strain !="G") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-```
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
-
-#these are the scatter plots
-
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
-
-```
-
-```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-
-#M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-
-```
-
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.avs_4MM)
-summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-```
-
-```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-```
-
-There is no strong signal of sex differences in raw mean inter-focal distances (IFD^raw^) across the full set of strains.
-A marginally significant difference between the sexes (t-test; p = `r Q1.IFD.raw_FULL_t.test$p.value`) is driven by one strain, *domesticus^G^* (t-test without domesticus^G^; p = `r Q1.IFD.raw_t.test_G$p.value`).
-This result indicates that females and males exhibit a similar level of interference when it is measured in physical SC units.
-
-In contrast, males have significantly longer normalized mean inter-focal distances (IFD^norm^) than females in seven out of eight strains (t-tests; p < `r Q1.IFD.PER_FULL_t.test$p.value`), with only _musculus^KAZ^_ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-Examination of IFD^norm^ distributions indicates that female IFD^norm^ values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-
-Mixed models and general linear models of IFD^norm^ support the inference of stronger interference in males: sex is the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` glm *M2*, p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`,  *M3* p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
-When interference is measured in physical SC units (IFD^raw^), the differences between sexes is low and only slightly significant (data not shown).
-In summary, controlling for differences in SC length (chromatin compaction) using IFD^norm^ indicates that interference is stronger in males, whereas consideration of IFD^raw^ shows that the sexes exhibit a similar level of interference on the physical scale.
-
-Visual comparison of interference pattern is aided by Figure 4C, which combines the total length of distance between two foci, but also their normalized locations along the chromosome.
-Two main patterns of sexual dimorphism emerge across all strains, (Supplemental figure X).
-First, the female patterns (top triangle), are more uniformly distributed compared to males.
-Suggesting a wider range of normalized IFD lengths compared to males (bottom triangle).
-Male IFDs display a stronger chromosome end-locatization of the second foci (Figure 4).
-
-```{r Comp_Ratios, echo=FALSE, eval=FALSE}
+## Single rewrite IFD triangle
+
+figure script below
+
+```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#good function, saves to file, can't reproduce in rmd ..width=650, height=500
+#big loop for automating the IFD triangles
+
+Q2.males.IFD <- c("F1", "other")
+Q2.auto_curate_DF <- cleanCurated_BivData[ ! cleanCurated_BivData$strain %in% Q2.males.IFD, ]
+
+Q2.df_strain_list <- split(Q2.auto_curate_DF, (Q2.auto_curate_DF$strain))
+
+#this works
+Q2.bottom.triangle2 <- function(my.y, my.x, my.data){
+  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
+          col =  colors_of_strains[unique(my.data$strain)])
+
+  mtext("Position.2.Male", 1, -2, at= mean(par()$usr[1:2]) + 0)
+
+  #the second val will move the x lable when negative
+
+  #at should be changing the position
+  #above is just changing the text not the numbers
+  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)
+
+  #second number - on which MARgin line, starting at 0 counting outwards.
+  #where text should be,  at = 0, 1 (bad)
+
+  mtext("Position.1.Male", 4, 0, padj=par()$usr[1] + .5) #last value at .5 is good for moving y label
+
+  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
+
+  axis(side=1, las=1, pos=0, at=x.at, 
+       labels=T, xpd=TRUE)
+#changing labels=T, puts the ddefault labels back
+
+  #second value should be around .75 (moves label text closer)
+  axis(4, las=1, pos=1, at=x.at) #this is the y axis
+  lines(0:1, 0:1, xpd=TRUE) #diagonal line
+}
+
+x.dist <- .25
+RP <- list()
+
+for(i in 1:(length(Q2.df_strain_list))){
+  #print(i)
+  current.strain <- as.character(Q2.df_strain_list[[i]]$strain[1])
+
+  #par(mar = c(2, 2, 2, 2))
+  #dev.new(width = 8, height = 4, unit = "in")
+  #declare file for saving png
+  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
+  plot.new()
+
+  my.data = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
+  my.x.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
+  my.y.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
+
+  #adjusting this mar lowers the margins (the edges)
+  op <- par(mar=c(.25, .15, .25, .75), oma=c(2, 0, 0, 2))
+#(1=bottom, 2=left, 3=top, 4=right)
+
+Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
+
+  par(op)
+  RP[[i]] <- recordPlot()
+  dev.off()
+}
+
+```
+
+```{r Q2.load.bottom.tri, echo=FALSE}
+
+#only load, SKIVE, PWD, MSM
+#MOLF, KAZ, LEW
+Q2.bottom.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
+Q2.bottom.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
+Q2.bottom.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
+Q2.bottom.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
+Q2.bottom.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
+Q2.bottom.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
+
+empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )
+
+#EMPTY GGPLOT
+Q2.LEW.bottom.tri <- ggplot(empty.DF, aes(x = col1, y = col2)) +
+  background_image(Q2.bottom.LEW) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+    axis.title.x=element_blank(),                       axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
+                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )
+
+```
+
+```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups", eval=FALSE}
+
+Q2.SC_dif.Figure3  <- ggdraw() +
+
+  draw_plot(Q1.SC.short.mouse_scatter_boxplot, x=0, y=.38, width=.5, height= .6)+
+  draw_plot(Tot.SC_mouse.av_ALL, x=.475, y=.38,  width = .5, height = .6)+
+
+  draw_plot(single.chrm.icon, x=.44, y=.79,  width = .1, height = .1)+   
+  draw_plot(all.chrm.icon, x=.88, y=.79,  width = .1, height = .1)+
+
+  #drawing triangles in the corner
+  #top right-triangle
+  draw_line(
+    x = c(.1, .2,  .1, .1 ), #keep the x values consistant
+    y = c( .25, .25,  .15, .25),
+    color = "blue", size=.75
+  )
+
+#draw boxes around the
+```
+
+# IFD double triangle
+
+```{r double.tri.Comp_Ratios, echo=FALSE, eval=FALSE}
 
@@
 
-```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
-## Additional Determinants of Genome-wide Recombination Rate Evolution in Males
-
-Next, we used the contrast between males in the high-recombination strains ( _musculus^PWD^_, _musculus^SKIVE^_, and _molossinus^MSM^_ ) and low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-The main objective of this section is to test for significant correlations between features of the recombination landscape and the evolution of mean MLH1 foci per cell.
-In comparing the male specific single bivalent based metrics the first criteria for analysis is a significant differences between the high and low recombining groups and the second step is testing for significant subspecies and strain effects (M4 and M5).
-
-Previous empirical work suggests basic predictions for the relationship between the genome wide recombination rate and two aspects of the recombination landscape.
-SC length is expected to be positively associated with genome-wide recombination rate because of the connection between DSB formation and the loop-axis structure [@zickler1999, @mercier2015].
-Crossover interference strength is expected to be negatively associated with genome-wide recombination rate as interference strength is reflected as the space between crossovers on the same chromosome with a finite length.
-Thus, increasing interference strength can result in 'pushing' additional crossovers off of the chromosomes.
-
-Following this logic we predict (1) _musculus^PWD^_ will have greater SC length and weaker interference than _musculus^SKIVE^_, which in turn will have longer SC and weaker interference compared to the other _musculus_ strains, (2) _molossinus^MSM^_ will have longer SC and weaker interference compared to _molossinus^MOLF^_, and 3) _domesticus_ strains will have similar SC length and crossover interference.
-
-## Conservation of Mean SC length
-
-```{r Q2.SC.short.long.setup, echo=FALSE}
-
-#table(male.long.biv$strain)
-#the Long biv dataset is quite unbalance (esp MSM)
-male.long.biv <- real.long.bivData %>% filter(sex== "male")
-
-#mouse av for long biv
-cell.av.long.biv.table <- ddply(male.long.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
-cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)
-
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
-
-Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                         #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
-                         )
-
-Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
-Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)
-
-Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
-Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)
-
-#divide musc males into groups
-Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))
-
-male.short.biv <- real.short.bivData %>% filter(sex== "male")
-
-Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                        #mean.first.1CO
-                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]
-                         ) )
-
-Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
-Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)
-
-Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
-Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)
-
-Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
-                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
-                                           0)))
-
-Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
-                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
-```
-
-```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}
-
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
-
-#nested boxplots, pooled Bivalents
-#x axis for hand.focicount -- good for seperating the points --- but it pools
-
-#prediction for Dom strains: low variation 
-Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
-  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
-
-Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
-
-#+theme(legend.position="none") 
-```
-
-```{r Q2.SC.long.biv_plots, echo=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( Long.biv_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+
-  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
-
-mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
-
-```
-
-```{r Q2.SC_MLH1.by.totalSC, echo=FALSE, warning=FALSE, fig.height=3, fig.width=8, fig.cap="Variation in total SC area per cell across strains and Relationship between total sc and mean MLH1 count per cell across male mice.
-A)	 Mean total SC area per mouse across subspecies and recombination groups.
-High- and low-recombining groups are indicated by 1 and 0 respectively.
-(Horizontal lines indicate strain averages for total SC area.)B) Mouse averaged rates of MLH1 and total SC per cell"}
-
-#make plot of mouse averages for MLH1 and total SC
-bep <- merge(  fomo.male_mouse.av,   totSC.mouse.av.table)
-
-X.by.texas <- ggplot(bep, aes(x=mean.skel, y=mean_co, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+theme(legend.position="none")+labs(x="Mouse Mean Total SC", y="Mouse Mean MLH1 count")
-
-plot_grid(mouse.av_totsc_male, X.by.texas, nrow=1, labels=c('A', 'B')) #Or labels="AUTO"
-
-```
-
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
-```{r Q2.totSC.ttest, echo=FALSE}
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol
-Q2tot.sc_t.test_High.Low_Mol <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-```{r Q2.short.long.ttest, echo=FALSE}
-
-#long and short 
-#mouse.av.short.biv.table
-#Q2mouse.av.short.biv.table
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
-                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
-
-#p-value = 0.9
-
-#mouse.av.long.biv.table
-Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
-#p-value = 0.2
-
-#mUSC-short
-Q2.short_t.test_High.Low.musc <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-short
-#Q2.short_t.test_High.Low.mol <- t.test((Q2mouse.av.short.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.short.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Mol"))$mean.SC)
-
-#musc-long
-Q2.long_t.test_High.Low.musc <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Musc") )$mean.SC,
-                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp == "Musc"))$mean.SC)
-
-#mol-long
-#Q2.long_t.test_High.Low.mol <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.SC,
-#                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") %>% filter(subsp =="Mol"))$mean.SC)
-
-#not enought observations
-
-```
-
-```{r Q2.SC_HvL_allBiv, echo=FALSE}
-
-#High-Low_All-BivData
-
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#short
-Q2.BIV_short_t.test_HL <- t.test((real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.short.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-#long
-Q2.BIV_long_t.test_HL <- t.test((real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (real.long.bivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
-
-```
-
-```{r Q2.SC_across.subsp_tot.SC, echo=FALSE}
-
-#Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
-
-#Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((totSC.mouse.av.table  %>% filter(subsp == "Dom") )$mean.skel,
-                          (totSC.mouse.av.table  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
-```
-
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
-
-# theses t.tests not working
-#mouse average
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((real.long.bivData  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean_SC,
-                          (real.long.bivData %>% filter(subsp == "Musc") %>% filter(sex == "male"))$mean_SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Short.biv_mouse_table  %>% filter(subsp == "Dom")%>% filter(sex == "male") )$mean_SC,
-                          (Short.biv_mouse_table %>% filter(subsp == "Musc")%>% filter(sex == "male"))$mean_SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test( (Long.biv_mouse_table  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean_SC,
-                          (Long.biv_mouse_table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean_SC )
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Short.biv_mouse_table  %>% filter(subsp == "Dom") %>% filter(sex == "male"))$mean_SC,
-                          (Short.biv_mouse_table %>% filter(subsp == "Mol")%>% filter(sex == "male"))$mean_SC)
-#p-value = 0.09
-
-```
-
-```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
-
-#DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 4e-13
-
-#DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 6e-10
-
-#DOM-musc-long
-Q2.longBIV_t.test_DomMusc <- t.test((real.long.bivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.long.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.2
-
-#DOM-musc-short
-Q2.shortBIV_t.test_DomMusc <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (real.short.bivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.longBIV_t.test_DomMol <- t.test((real.long.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.long.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.shortBIV_t.test_DomMol <- t.test((real.short.bivData  %>% filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-                          (real.short.bivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
-#p-value = 0.06
-
-```
-
-```{r Q2.glm.SC.long, echo=FALSE}
-
-#this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-
-glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
-
-sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
-
-#MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
-  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
-sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
-
-#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-```
-
-```{r Q2.glm.SC.short, echo=FALSE}
-#Q2mouse.av.short.biv.table
-#mouse.av.short.biv.table
-
-glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
-sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)
-
-#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)
-
-glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
-glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
-
-#DOM
-glmQ2_short.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )  )
-sum.glmQ2_short.biv_M2_DOM <- summary(glmQ2_short.biv_M2_DOM)
-
-#MUSC
-glmQ2_short.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )  )
-sum.glmQ2_short.biv_M2_MUSC <- summary(glmQ2_short.biv_M2_MUSC)
-
-#MOL
-#within subspecies
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.short.biv.table %>% filter(subsp == "Musc") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-```
-
-```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}
-
-plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
-                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av.
-SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
-
-```
-
-```{r log.reg_SC.length, echo=FALSE, include=FALSE}
-
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)))
-
-#male poly shouldn't be here
-Male.poly.Mouse.Table_BivData_4MM$Rec.group <- ifelse(grepl("PWD male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 
-                                ifelse(grepl("MSM male", Male.poly.Mouse.Table_BivData_4MM$category), 1,
-                                ifelse(grepl("SKIVE male", Male.poly.Mouse.Table_BivData_4MM$category), 1, 0)))
-
-#All chrm
-log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av)
-
-#considering seperating out by hand foci counts
-
-#1CO
-log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_1CO)
-
-#2CO
-log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_mouse.av_2CO)
-
-#3CO
-log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
-                            data=Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_mouse.av_3CO)
-
-# bivalent level log regression
-
-log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV)
-
-#1CO  (BIVDATA)
-log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,
-
-                data=Curated_BivData[(Curated_BivData$subsp == "Musc") & (Curated_BivData$hand.foci.count == 1) &(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_1CO)
-
-#2CO
-log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 2) & (Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-#summary(log.reg_SC.length_BIV_2CO)
-
-#3CO
-log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
-                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")&(Curated_BivData$hand.foci.count == 3)&(Curated_BivData$sex == "male"),], family=binomial(link="logit"))
-
-#summary(log.reg_SC.length_BIV_3CO)
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
-#REMOVE THIS, IT uses total bivalent data
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
-Confirming the basic predictions, there is a positive correlation between mouse mean MLH1 foci per cell and total SC (Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`).
-Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure 5).
-While high-recombination strains have significantly greater mouse mean total SC area in the total dataset (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p =  `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, the mouse means for the reduced (short and long) bivalent datasets are not significantly different between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
-
-In a general linear model with total SC as the dependent variable, two subspecies effects are significant (_musculus_ p =  `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, _molossinus_, p = `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`).
-In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.
-**Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?**
-
-## Single Crossover position is decoupled from genome wide recombination rate evolution
-
-```{r Q2.1CO.show_plots, echo=FALSE,  warning=FALSE, fig.cap="Male distributions of single crossover normalized positions.", fig.width=4}
-#Curated_BivData_male
-
-#boxplots of nrm F1
-male.bivdata.1CO <- Curated_BivData_male %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-Q2.show.nrmF1 <- ggplot(male.bivdata.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter(alpha=.23 )+ geom_boxplot(aes(y= PER_Foci_1), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+theme(legend.position="none", axis.text.x = element_text(size=9), axis.title.x = element_blank())+labs(y="Foci 1 Relative Position")
-
-Q2.show.nrmF1
-
-nrmF1_mouse_level <- ggplot(Male.poly.Mouse.Table_BivData_4MM, aes(x=strain, y=norm.first.foci.pos, color=strain) )+geom_jitter(alpha=.5 )+ geom_boxplot(aes(y= norm.first.foci.pos), alpha=.5 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse Level, \nNormalized 1CO Foci pos")
-
-```
-
-```{r Q2.1COpos.LONG.BIV_plot, echo=FALSE, warning=FALSE}
-#make the same plots as above but with the long biv data set
-male.long.biv.1CO <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)
-
-#mouse.av.long.biv.table
-
-Q2.show.nrmF1_long.biv <- ggplot(male.long.biv.1CO, aes(x=strain, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Normalized 1CO Foci pos \nLong Biv")
-
-Mouse.av.Long.biv <- ddply(male.long.biv.1CO, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Nbiv  = length(Obj.ID),
-                       mean.1CO_PER = mean(PER_Foci_1)
-                      # mean_CO = mean()
-                       )
-
-Mouse.av.Long.biv <- add_strain(Mouse.av.Long.biv)
-Mouse.av.Long.biv <- add_subsp(Mouse.av.Long.biv)
-
-Q2.shoooo   <- ggplot(Mouse.av.Long.biv, aes(x=strain, y=mean.1CO_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean.1CO_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("mouse ave \nNormalized 1CO Foci pos \nLong Biv")
-
-```
-
-```{r Q2.1COpos.show.MOL, echo=FALSE}
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos
-
-mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
-                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()
-
-#+facet_wrap(~mouse, scales = "free")
-
-single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
-                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))
-
-#+scale_color_manual(values=colors_of_strains)
-
-single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
-                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
-
-#CHRM.length is very significant-- some of the same subsp and strains are sig     
-```
-
-The normalized 1CO position is not significantly different between high-recombination and low-recombination strains for the total pooled data (t-test; p = `r Rec.group_BivData_F1post_Ttest$p.value`) and also when examined within subspecies (t-test; p = r Q2.1COpos_mus.av_Musc_ttest$p.value and p = r Q2.1COpos_mus.av_Mol_ttest$p.value for _musculus_ and _molossinus_, respectively).
-While there are significant strain effects for _domesticus^WSB^_ and _molossinus^MOLF^_ in a general linear model with normalized position as the dependent variable, this evolution of the 1CO positioning is decoupled from the total genome-wide recombination rate since these two strains do not vary significantly from the male-specific rate.
-
-## Evolution of interference is associated with genome wide recombination rate evolution
-
-```{r Q2.show.IFD, echo=FALSE, warning = FALSE}
+```{r Q2.show.IFD, echo=FALSE, warning = FALSE, eval=FALSE}
 #subset
-male.biv.2CO <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+male.biv.2CO <- cleanCurated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
 
@@
 
-```{r Q2.IFD.glm.test, echo=FALSE, warning=FALSE}
-
-#M1.ABS
-Q2.M1glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
-sum.Q2.M1glm_IFD.raw <- summary(Q2.M1glm_IFD.raw)
-
-#M1.PER
-Q2.M1glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male") )
-
-sum.Q2.M1glm_IFD.PER <- summary(Q2.M1glm_IFD.PER)
-#summary(Q2.M1glm_IFD.PER)
-
-#M2.raw
-Q2.M2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)
-
-Q2.M2glm_IFD.PER <- glm(mean_IFD.2CO_PER ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
-sum.Q2.M2glm_IFD.PER <- summary(Q2.M2glm_IFD.PER)
-#  MSM strain effect for 2CO PER not significant
-
-#within subspecies
-#Dom
-Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD_Dom)
-
-#Musc
-Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD_Musc)
-
-#Mol
-Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD_Mol)
-
-#Dom
-Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
-#summary(Q2.glm.M2_IFD.PER_Dom)
-
-#Musc
-Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
-#summary(Q2.glm.M2_IFD.PER_Musc)
-
-#Mol
-Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
-#summary(Q2.glm.M2_IFD.PER_Mol)
-
-```
-
-```{r Q2.double.tri.rewrite, echo=FALSE, warning=FALSE}
-
-#good function, saves to file, can't reproduce in rmd ..width=650, height=500
-#big loop for automating the IFD triangles
-
-Q2.males.IFD <- c("F1", "other")
-Q2.auto_curate_DF <- cleanCurated_BivData[ ! cleanCurated_BivData$strain %in% Q2.males.IFD, ]
-
-Q2.df_strain_list <- split(Q2.auto_curate_DF, (Q2.auto_curate_DF$strain))
-
-#this works
-Q2.bottom.triangle2 <- function(my.y, my.x, my.data){
-  points(my.x ~ my.y, data=my.data, xpd=TRUE, 
-          col =  colors_of_strains[unique(my.data$strain)])
-
-  mtext("Position.2.Male", 1, 1.5, at= mean(par()$usr[1:2])) #at should be changing the position
-  #above is just changing the text not the numbers
-  #first number, on which side of the plot (1=bottom, 2=left, 3=top, 4=right)
-
-  #second number - on which MARgin line, starting at 0 counting outwards.
-  #where text should be,  at = 0, 1 (bad)
-
-  mtext("Position.1.Male", 4, 3, padj=par()$usr[1] +4)
-
-  x.at <- axisTicks(par()$usr[1:2], 0) #where the ticks should go
-  axis(side=1, las=1, pos=0, at=x.at, 
-       labels=F, xpd=TRUE)
-#las, orientation, at location of line
-  #labels automatic labels
-  #pos
-  #xpd - clipping
-
-   mtext(seq(0, 1, .2), 1, 1, at=x.at)
-  #second / third val negative moves the valu up
-
-  axis(4, las=1, pos=1) #this is the y axis
- #   axis(4, las=1, pos=1 + x.dist)
-  #this is calling axis twice, not sure why
-  lines(0:1, 0:1, xpd=TRUE) #diagonal line
-}
-
-x.dist <- .25
-RP <- list()
-
-for(i in 1:(length(Q2.df_strain_list))){
-  #print(i)
-  current.strain <- as.character(Q2.df_strain_list[[i]]$strain[1])
-
-  #par(mar = c(2, 2, 2, 2))
-  #dev.new(width = 8, height = 4, unit = "in")
-  #declare file for saving png
-  png(paste0("~./MLH1repo/doc/figureFiles/Q2.IFD.triangle_", current.strain, ".png"),  width= 650, height=500, res = 100) #500 seems like a good size
-  plot.new()
-
-  my.data = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) )
-  my.x.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 
-  my.y.bottom = (cleanCurated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2
-
-  #adjusting this mar lowers the margins (the edges)
-  op <- par(mar=c(1.5, 1.5, 1.5, 1.5), oma=c(2, 0, 0, 2))
-
-  #   op <- par(mar=c(1.5, 2, 4, 6), oma=c(2, 0, 0, 2))
-  Q2.bottom.triangle2(my.y.bottom, my.x.bottom, my.data) #x.dist
-
-  par(op)
-  RP[[i]] <- recordPlot()
-  dev.off()
-}
-
-```
-
-```{r Q2.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.cap="male IFD distributions highlighting differences across recombining groups"}
-
-Q2.plot1.WSB <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_WSB.png')
-Q2.plot2.LEW <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_LEW.png')
-Q2.plot3.G <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_G.png')
-
-Q2.plot4.KAZ <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_KAZ.png')
-Q2.plot5.SKIVE <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_SKIVE.png')
-Q2.plot6.PWD <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_PWD.png')
-
-Q2.plot7.MSM <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MSM.png')
-Q2.plot8.MOLF <- readPNG('~/MLH1repo/doc/figureFiles/Q2.IFD.triangle_MOLF.png')
-
-grid.arrange(rasterGrob(Q2.plot1.WSB), rasterGrob(Q2.plot2.LEW), rasterGrob(Q2.plot3.G),
-             rasterGrob(Q2.plot4.KAZ), rasterGrob(Q2.plot5.SKIVE), rasterGrob(Q2.plot6.PWD),
-             rasterGrob(Q2.plot7.MSM), rasterGrob(Q2.plot8.MOLF), 
-             ncol=3)
-
-```
-
-Mouse averages for both IFD^raw^ and IFD^norm^ are significantly longer in high-recombination strains (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS$p.value`).
-
-This pattern is supported by slightly significant p values from comparisons within _musculus_ (t-test; IFD^norm^: p = `r t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MUSC$p.value`) and within _molossinus_ (IFD^norm^: p= `r t.test_mouse.av.IFD_PER_MOL$p.value`; IFD^raw^: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-Similar results are seen with general linear models for both IFD^raw^ and IFD^norm^: only effects associated with high-recombination strains are significant (p < 0.05).
-
-That IFD^raw^ and IFD^norm^ show similar patterns eliminates variation in SC lengths and bivalent sizes as primary explanations.
-We determined that the main difference in IFD^norm^ distributions between high-recombination and low-recombination strains is an enrichment of IFD^norm^ values under 30% in low-recombination strains.
-The frequency of IFD^norm^ values that fall below 30% ranges from 8.2% (_domesticus^G^_) to 16% (_musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show such frequencies below 5% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
-
-Differnces in IFD can also be noted from the triangle IFD plots (Figure 6).
-The three high-recombining strains show more clustering of IFD observations near the area of the plot which indicates longer IFDs.
-The remaining strains have more observations short IFD observations near the diagonal.
-
-In summary, the level of interference is a significant predictor of evolution in the genome-wide recombination rate, but SC length and crossover position on 1CO bivalents are not.
-However, the pattern is in the opposite direction to our prediction: high-recombination strains have stronger interference.
-
-# Discussion
-
-Despite quantifying a similar number of wild derived inbred strains from *M.
-m.
-domesticus* and *M.
-m.
-musculus*, we only observed rapid male specific evolution for the genome wide recombination rate for *M.
-m.
-musculus.* Estimates of effective population size for the three main subspecies of the house mouse found *M.
-m.
-musculus* had the lowest [@geraldes2011].
-Selection becomes less efficient at removing deleterious mutations in small populations outside of mutation-drift balance.
-In such populations, modifiers for increased recombination rates will be selected for as they increase the efficiency of selection.
-
-An additional possibility is that the different rates in recombination rate variation across subspecies is related to history of hybridization across the populations sampled in this panel of inbred strains.
-Theoretical work as connected recombination that rate of purging deleterious introgressed DNA in populations after hybridization (Caballero et al 2019, Schumer 2018).
-The three high-recombining strains identified in this study have connections to hybridization.
-Both *Musculus^PWD^* and *musculus^SKIVE^* strains were established from samples from the European house mouse hybrid zone.
-Notably *musculus^SKIVE^* comes from a younger section of the hybrid zone with a nuclear genome from *M.
-m.
-musculus* and a mitochondria genome from *M.
-m.
-domesticus* [@gyllensten1987_SKIVE].
-The subspecies *M.
-m.
-molossinus* is a natural hybrid of *M.
-m.
-musculus* and *M.
-m.
-cataneus* (Geraldes et al 2008, [@geraldes2011]).
-Models for the aggregate recombination metrics, suggest that the location of recombination events and the total rate influence the dynamics of purging of deleterious introgress DNA (Veller et al 2019), motivating increased study of recombination rates and natural populations.
-
-<!-- transition to looking at the number and placement of crossovers -->
-
-To place our results in the broader context of recombination rate evolution we compare our main results to expectations under three modifier models and two functional and cell physiology models (table X).
-While there are challenges to fitting models built under different assumptions, we feel it's a useful exercise to highlight any unifying patterns for our specific traits across existing models.
-The three modifier models were built to explain variation in genome wide recombination rates between sexes (heterochiasmy), while the chromosome oscillatory movement (COM) model was built to describe positive interference.
-This spindle based selection model was developed to explain the results resented in this manuscript.
-In cases where the models can't be extended to predicting results, we list a 'NA' for the prediction.
-
-## Evolution of reversed heterochiasmy direction
-
-In our results and more generally in the literature, female biased heterochiasmy is the most common pattern, however male biased heterochiasmy is not especially rare [@brandvain2012scrambling].
-Our results extend previous reports of male biased heterochiasmy [@dumont2011] to  new subspecies of house mouse (*molossinus*) and emphasize the short evolutionary period.
-Whether these instances of evolution of genome-wide recombination rates are due to independent events or segregating genetic variation within house mouse subspecies requires further study.
-Regardless, the faster male evolution in genome-wide recombination rates suggests that even for species in female biased heterochiasmy species, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background.
-This insight could be important for animal breeding programs [@battagin2016].
-
-Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment.
-The SAC prevents aneuploidy by ensuring all bivalents are correctly attached to the microtubule spindle (bi-orientated) before starting the metaphase-to-anaphase transition [@laneKauppi2019, @subramanian2014, @dumontDesai2012].
-
-A stable chromosome structure requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion [@laneKauppi2019].
-Genetic variants will be selected in a manner to optimize the process of bi-orientation and prevent SAC from delaying cell cycle or triggering apoptosis.
-Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis [@laneKauppi2019,  @subramanian2014], which are connected to conserved features of gametogenesis; centrosome spindle [@So2019, @dumontDesai2012] and cell volume [@kyogoku2017].
-The more stringent SAC in spermatogenesis will be more effective at removing genetic variants which interfere with bi-orientation compared to females.
-This dynamic can give rise to sex-specific genetic variants which in turn can result in  context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.
-
-## Conservation in sexual dimorphism in between cell variation in crossover number
-
-In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure.
-This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals.
-Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007].
-Centrosomes act as the two MTOCs
-for the spindle during spermatogenesis and mitosis.
-The delivery of centrioles (most mammals) or centriole material (rodents and rabbits) to the egg is thought to be a barrier to evolution of parthenogenesis [@manandhar2005centrosome].This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].
-
-Regardless of whether these hypotheses are true, our observations of levels of between cell variation fail to support a basic prediction of that a higher mean will be associated with a higher variance.
-Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count.
-These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002], [@gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes.
-This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
-
-<!-- stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and bi-orientation of bivalents in the metaphase cell, -- spindle dynamics much different due to asymmetrical division Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle>   <Bennabi et al 2016 --spindle assembly in oocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review -->
-
-## Conservation in sex-specific recombination landscape
-
-Our results extend the plethora or previous data reports for broad scale sex difference in the recombination landscape [@sardell_sex_2020] are data comprised mostly of species with large evolutionary distances.
-In contrast our results demonstrate that this pattern is maintained across much shorter evolutionary distances and even with rapid evolution in the genome-wide recombination rate.
-We argue that this conservation of the male chromosome end bias is one of the most conserved features of heterochiasmy and could be connected to fundamental aspects of gametogenesis.
-
-As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a conserved trait.
-The two locus modifier and SACE models predict sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks under linkage in males respectively.
-The COM model, predicts the sex differences in the recombination landscape via interference is due to a combination of the length of the axis and differing strengths of chromosome movements during the pairing process [@hulten2011_COM].
-
-Under the spindle based model we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I.
-The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (@laneKauppi2019, @subramanian2014, @dumontDesai2012).
-The number and placement of crossovers alter the distribution of sister cohesion and the resulting chromosome structure when bivalents are aligned and bi-oriented on the metaphase plate [@vanVeen2003, @altendorfer2020].
-Differences in timing of cell cycle between oogenesis and spermatogenesis imposes the the different selective pressures on how sister cohesion affects chromosome structure.
-Faster spermatogenesis may select for synchronization of the separation homologs (cite).
-While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dicyate arrest [@lee2019].
-
-The increased genetic diversity and recombination rates at chromosome ends @haenel2018 suggests that this is driven by the male-specific recombination landscape [@sardell_sex_2020], which provides some interesting evolutionary predictions.
-
-## Conservation of sex differences in chromatin structure
-
-Our results from _musculus^MSM^_ and _musculus^PWD^_ demonstrate that chromatin compaction are uncoupled from the direction of heterochiasmy in house mice.
-These results slightly depart from predictions which nominate chromatin compaction as the primary driver of recombination rate variation [@petkov2007].
-Chromatin compaction could explain variation in heterochiasmy, which is driven by the distinct meiotic contexts across sexes but is a weak predictor for recombination rate variation within the sexes.
-Our results indicate the sexual dimorphism in chromatin compaction is conserved in house mouse (e.g.
-females have longer SC than males) and that the rapid male-specific evolution in crossover number per cell proceeded through another aspect of crossover regulation (see below).
-
-Only the COM model predicts sexual dimorphism in chromatin structure and attributes the source to greater cell volume in oocytes.
-This prediction model might fit broader pattern such as in *Arabidopsis thalania* (Mercier, Sera) where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].
-
-As outlined in @haenel2018, sexual dimorphism in chromatin organization translates to interference strength measured in base pairs.
-From empirical data in mice [@petkov2007], [@lynn2002] and humans [@gruhn2013], there seem to be to general configurations for chromatin structure: longer axis and shorter loops in females; shorter axis and longer loops in males.
-The signal for interference is thought to be mediated though the SC, which is built on top of the chromosome axis [@zicklerKleckner2015] regardless of the loop size.
-Supporting this is the observation that interference strength measured in micrometers of is conserved between sexes [@petkov2007].
-As a consequence, since more base pairs are sequestered into the radial loops per unit of the linear central axis, chromatin organization with longer DNA loops (and shorter axis), as seen in males,  will have a signal of stronger interference when measured in base pairs compared to 'the typical female chromatin configuration.
-Stronger interference in males, has been noted across many sex-specific linkage maps (cite) strengthening the hypothesis that sexual dimorphism in the chromatin organization is widespread, however this has only been confirmed by cytological data in mice and humans.
-(Also large difficulties in observing sperm cells through meiosis).
-
-<!-- The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019] -->
-
-## Evolution of interference strength in males
-
-An expected consequence of our observed elevated genome-wide recombination rate is lower LD across the genome, which may increase the efficiency of selection and impact the dynamics of introgression (Schumer).
-Given that the higher genome-wide recombination rates are paired with stronger interference these predicted pattern may have heterogeneous signal along chromosomes.
-
-Examples of evolution of interference strength are rare and may be complicated by changes in the underlying karyotype structure (Segura et al 2014).
-The general pattern of empirical measures of interference strength is that it has a negative correlation to the genome wide recombination rate [@ottoPaysuer2019].
-This fits the logic of the chromosome axis acting as the 'real estate' for DSBs and aligns with known molecular mechanisms (cite).
-Two examples of the opposite direction, a positive correlation of interference strength and genome-wide recombination rate, involve descriptions of observations at the single bivalent level; the between lab-raised and wild mice of *Peromyscus leucopus* from [@peterson2019] or large effects on genome wide recombination rates across a short evolutionary differences; the within-sex comparison of two breeds of cattle with different genome wide recombination rates [@ma2015_cattle].
-
-We propose that the spindle based selection model can explain the positive correlation via selection on the amount of sister cohesion connecting homologs at metaphase.
-The evolution of genome wide recombination rates in our study is driven by a transition of the majority of
-bivalent having single crossovers to an enrichment of bivalents with two crossovers.
-All else being equal, interference strength would be expected to be equal or even weaker.
-Our results show that the two crossovers are spaced further apart when SC length is controlled for and chromosome size effects are minimized, in the strain enriched for two crossover bivalents.
-Assuming that chromatin compaction between pachtyene and metaphase is uniform along chromosomes, an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs (Figure X).
-
-The COM model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase [@rubin2020mixing], however it lacks a mechanism for a evolutionary based predictions.
-While the SACE modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a recombination landscape with stronger interference strength.
-
-We have focused on the model involving direct selection on the recombination pathway, which has an impact on the fitness of an individual, but we acknowledge connections to other models involving indirect selection on recombination rates which involve the fitness outcomes of offspring.
-A model from @goldstein1993, reviewed in [@ottoPaysuer2019] and @veller2019shuffle, predict that interference strength evolves whenever increased recombination rates evolve.
-The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate.
-A distinguishing feature of this model from other modifier  models is that the number of crossovers per chromosome is kept constant.
-Empirically, the range of crossovers per chromosome is quite limited, 1-3, for the majority of chromosomes across most taxa [@ottoPaysuer2019, @stapley_variation_2017].
-Constraining the number of crossovers per chromosome in models may fit empirical data better than those where recombination rate across an abstracted genetic space is unconstrained.
-
-## Future steps
-
-There are still many un-known aspects regarding heterochiasmy (@lenormand2016) --- that would benefit from the cross-pollination of physiology based models and more abstract model such as population genetic models with testable hypotheses  @dapperPayseur2017.
-A goal of such merging of models could be to connect empirical findings across scales.
-We make two suggestions for future steps in studying recombination variation.
-First consider the when comparing sex specific meiotic traits consider that evolutionary distinct trajectories due to the fundamental differences in gametogenesis.
-Second certain systems can be leveraged to overcome limitations inherent in some approached.
-For example, identifying chromosomes across cells in cytological data requires chromosome specific probes, (e.g.
-FISH) but chromosome specific data can be collected from organisms with diverse karyotypes (birds, peromyscus,  humans) or backgrounds with Robertsonian trans-locations.
-
-# References
+```{r Q1.F1.pos.LONG.SHORT.BIV_scatter, echo=FALSE}
+#real.long.bivData
+Q1.short.biv_scatr_box   <- ggplot( (real.short.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nShort Biv")
+
+Q1.long.biv_scatr_box   <- ggplot((real.long.bivData %>% filter(hand.foci.count == 1) ), aes(x=sex, y=PER_Foci_1, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= PER_Foci_1 ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+\nNormalized 1CO Foci pos \nLong Biv")
+
+```

### 2020-06-04 doc/MS_v2.Rmd
updates from final v2 files

@@
 To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
-Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.
+Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
 
@@
 In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from
-`r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female).
-From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm.
-Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained.
+`r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
+From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` bivalent objects were isolated by the algorithm.
+Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
 The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
@@
 This “short bivalent” data set included the four or five shortest bivalents, and excluded the -- X bivalent in oocytes.
-A total of `r sum(Short.biv_mouse_table$nobs )` -- 678 short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$ncells)` -- 103 oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$ncells)` -- 37 spermatocytes.
+A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
 Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
@@
 
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` -- 14 wild-derived inbred strains sampled from three subspecies of house mice ( _M.
+We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice ( _M.
 m.
@@
 spicilegus_ ).
-Mean MLH1 foci counts for `r sum(main.table$Nmice)` -- 161 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 22.0 spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` -- 1,867 spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` -- 18.5 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,409 oocytes).
+Mean MLH1 foci counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
+
 Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
@@
 
-To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` -- 187 mice.
+To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` mice.
 To investigate the effects of sex, we limited our analysis to `r length( unique( DF.HetC.MixedModel.HQ$strain) )` strains with sufficient observations from both sexes (Table 1).
 
-We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < 10^{-6} -- p < `r 1/10000`), sex (p = 1.5510^{-8} -- `r MLH1.MMlmer_results$sex.results`), subspecies (p=1.7210^{-4} -- `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = 3.110^{-5} -- `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
+We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < `r 1/10000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p= `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
 After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = 1.0410^{-6} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
-In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = 1.2610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = 3.8610^{-4} -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
+A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
+In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
 
-We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4).
-In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).
+We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4).
+In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).
 
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p= 0.02 -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
+Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
@@
 
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p=`r Q1.lmer_totSC_M1_results$sex.results`; M2: p=`r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p= `r Q1.lmer_short_M1_results$sex.results`; M2: p= `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
-The existence of some subspecies and strain effects (M1 interaction: p = `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]`,  SKIVE * male p= `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
+The existence of some subspecies and strain effects (M1 interaction: p = `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]`,  SKIVE * male p = `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.
 In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males have more MLH1 foci per cell.
@@
 We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68 -- `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.56; t-test; p = 2.9210^{-22} -- `r Q1_1CO.pos_t.test.FULL$p.value` ).
-Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = 1.2610^{-25} -- `r lmer.Q1.nrm1CO_M1_results$sex.results` ; M2: p = 1.3310^{-7} -- `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 1.3310^{-7} -- `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value` ).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results` ; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 
-Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 1.4910^{-12} -- p < 0.05), with only _musculus^KAZ^_ showing no difference (t-test; p = 0.33 -- `r Q1.IFD.PER_t.test_KAZ$p.value` ).
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (all t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value` ).
 Examination of IFD~norm~ distributions indicates that female values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Analysis of models treating IFD~norm~ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = 6.7410^{-14} --`r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = 0.01 -- `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+Analysis of models treating IFD~norm~ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
 
@@
 
-```{r algorithm_performance_stats, echo=FALSE}
+```{r algorithm_performance_stats, echo=FALSE, eval=FALSE}
 
@@
 
-```{r Chrm.Prop.misc.calulations, echo=FALSE, eval=FALSE}
-
-total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)
-
-```
-
-```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE, eval=FALSE}
-#REMOVE THIS, IT uses total bivalent data
-
-#rerun the models for datasets -- deviding by chrm class and rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )
-
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
-
-#calq_mouse av
-mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                        n_cells = length(unique(fileName) ),
-                        mean.SC.total = mean(chromosomeLength),
-
-                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
-                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
-
-                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),
-
-                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),
-
-                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)
-
-                         )
-
-mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
-mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)
-
-mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
-mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)
-
-#calq_mouse av
-mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength)
-                         )
-mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
-mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)
-
-mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
-mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
-```
-
 We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
 We considered proportions of bivalents with different numbers of crossovers, double-strand breaks, SC length, and crossover positions, including crossover interference.
-
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` -- 34,982) have either one or two MLH foci (Figure 3).
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")`) have either one or two MLH focus (Figure 3).
 The proportions of single-focus (1CO) bivalents vs.
@@
 
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.53 in _molossinus^MSM^_ (Supplemental Figure 3).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
 Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-_musculus^KAZ^_: p = 3.1510^{-33} -- `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_: p = 4.7210^{-13} -- `r MSM.MOLF.2CO.prop.test$p.value`) and _musculus^PWD^_ vs.
+_musculus^KAZ^_: p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value`) and _musculus^PWD^_ vs.
 _molossinus^MSM^_: p = `r PWD.MSM.2CO.prop.test$p.value`.
@@
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of 76 -- `r sum( (DMC1table.main$ncells.x) )`  early zygotene and 76 -- `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from three low-recombination strains ( _musculus^KAZ^_, _domesticus^WSB^_ , and  _domesticus^G^_) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_).
-High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < 10^{-6} --  p < `r 1/1000000` ).
-In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = 0.66 -- `r ttest.HighLow.Z.pval` ).
+DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )`  early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from three low-recombination strains ( _musculus^KAZ^_, _domesticus^WSB^_ , and  _domesticus^G^_) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_).
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test;  p < `r 1/1000000` ).
+In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval` ).
 Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
@@
 
-Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = 0.48 --`r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = 2.2410^{-10} -- `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
+Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = `r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within _M.
+m.
+molossinus_ (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` ) but not within _M.
+m.
+musculus_ (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
 
-Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within _M.
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value` ; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
+In a model with total SC length as the dependent variable, two subspecies effects are significant ( _M.
 m.
-molossinus_ (p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value` ) but not within _M.
+musculus_ p = `r sum.Q2.glm_mus.av_totSC_M4[2,4]` , _M.
 m.
-musculus_ (p = 0.87 -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value` ; long: p = 0.18 -- `r Q2long.biv_t.test_High.Low$p.value`).
-In a model with **total SC length** as the dependent variable, two subspecies effects are significant ( _M.
-m.
-musculus_ p = 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[2,4]` , _M.
-m.
-molossinus_, p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[3,4]` ).
+molossinus_, p = `r sum.Q2.glm_mus.av_totSC_M4[3,][4]` ).
 In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05), but they are not consistent across models.
@@
 
-High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = 0.24 -- `r Q2.1COpos_Ttest$p.value` ) or within subspecies (t-test; _M.
+High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = `r Q2.1COpos_Ttest$p.value` ) or within subspecies (t-test; _M.
 m.
-musculus_: p = 0.41 -- `r Q2.1COpos_mus.av_Musc_ttest$p.value` ; _M.
+musculus_: p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` ; _M.
 m.
-molossinus_ : p = 0.07 -- `r Q2.1COpos_mus.av_Mol_ttest$p.value` ).
+molossinus_ : p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` ).
 While _domesticus^WSB^_ and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable, these two strains do not differ in MLH1 count.
@@
 
-**Q2_IFD**
-High-recombination strains have greater inter-focal distances (in both raw and normalized measured) than low-recombination strains (t-test; IFD~norm~: p = 7.7410^{-7} -- `r t.test_mouse.av.IFD_PER$p.value` ; IFD~raw~: p = 8.7810^{-6} -- `r t.test_mouse.av.IFD_ABS$p.value`).
+High-recombination strains have greater inter-focal distances (in both raw and normalized measured) than low-recombination strains (t-test; IFD~norm~: p = `r t.test_mouse.av.IFD_PER$p.value` ; IFD~raw~: p = `r t.test_mouse.av.IFD_ABS$p.value`).
 This pattern holds within _M.
 m.
-musculus_ (t-test; IFD~norm~: p = 2.0410^{-5} -- `r  t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD~raw~: p = 1.9410^{-4} -- `r  t.test_mouse.av.IFD_ABS_MUSC$p.value`), with a trend in the same direction within _M.
+musculus_ (t-test; IFD~norm~: p = `r  t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD~raw~: p = `r  t.test_mouse.av.IFD_ABS_MUSC$p.value`), with a trend in the same direction within _M.
 m.
-molossinus_ (IFD~norm~: p= 0.17 -- `r t.test_mouse.av.IFD_PER_MOL$p.value` ; IFD~raw~: p = 0.08 -- `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
-**these pvalues aren't the same anymore**  is this cause there aren't many molf measures?
+molossinus_ (IFD~norm~: p= `r t.test_mouse.av.IFD_PER_MOL$p.value` ; IFD~raw~: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
 
@@
 That IFD~raw~ and IFD~norm~ show similar patterns eliminates variation in SC lengths (and bivalent sizes) as the primary explanation.
-The main distinction in IFD~norm~ distributions between high-recombination and low-recombination strains is an enrichment of IFD~norm~ values under 30% in low-recombination strains (Figure 6).
+The main distinction in IFD~norm~ distributions between high-recombination and low-recombination strains is an enrichment of IFD~norm~ values under 30% in low-recombination strains (Supplemental Figure 3).
 The frequency of IFD~norm~ values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
@@
 
+*not sure where to insert this*
+
 Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
-While this trait will be examined in greater detail in future manuscript, we not that the general pattern is that females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1 and Table 2).
+While this trait will be examined in greater detail in future manuscript, we not that the general pattern is that females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1 and Table 3).
 
@@
 Because inter-strain divergence times are identical for the two sexes, this observation demonstrates that the genome-wide recombination rate evolves faster in males, at least in house mice.
-More generally, recombination rate divergence is decoupled in males and females (Spearman’s correlation test comparing female and male strain mean MLH1 values for house mice; rho = 0.18 -- `r sperman.corr.FM.pval$estimate`; p = 0.60 -- `r sperman.corr.FM.pval$p.value` ).
+More generally, recombination rate divergence is decoupled in males and females (Spearman’s correlation test comparing female and male strain mean MLH1 values for house mice; rho = `r sperman.corr.FM.pval$estimate`; p = `r sperman.corr.FM.pval$p.value` ).
 These disparities are remarkable given that recombination rates for the two sexes were measured in the same genomic backgrounds.
@@
 These variants could differently affect females and males at any step in the recombination pathway.
-Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contributed to sex-specific evolution in the recombination rate.
+Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons) and the significance of SC length distinguishing high and low recombining strains, our findings suggest that mutations that determine the chromatin structure and number of double-strand breaks contributed to sex-specific evolution in the recombination rate [@baier2014].
 
@@
 We propose an alternative interpretation based on the cell biology of gametogenesis.
-During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister cohesin distal to chiasmata [@LaneKauppi2019, @vanveen].
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister cohesin distal to chiasmata (@LaneKauppi2019, @vanVeen2003).
 The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesin holding homologs together [@LaneKauppi2019].
-Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and smooth chromosome separation , thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and chromosome separation , thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
 
@@
 The higher stringency of the SAC during spermatogenesis suggests that selection will be better at removing mutations that interfere with bi-orientation in males than in females.
-Therefore, faster male evolution of the genome-wide recombination rate could be driven by the stronger SAC acting on chromosome structures at the metaphase I alignment.
+Therefore, faster male evolution of the genome-wide recombination rate could be driven by the more stringent SAC acting on chromosome structures at the metaphase I alignment.
 
@@
 We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
-
-A source for the evolution of sexual dimorphism in meiotic traits is the spindle assembly checkpoint, which is more strict and effective in spermatogenesis and permissive (or 'leaky') in oogenesis (cite).
-This difference can result in different abilities to filter sex-specific genetic variants.
-The spindle assembly checkpoint acts to regulate the reductional chromosome division, which is facilitated by the proteolytic decay of sister cohesin distal to chiasmata (that is holding homologs together (during bi-orientation)).
-The number and placement of crossovers alter the distribution of sister cohesin distal to chiasmata which could in theory affect this metaphase to anaphase transition (@vanVeen2003, @LaneKauppi2019, @subramanian2014, @dumontDesai2012).
-(additional features of gametogenesis may contribute to sex differences) Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell (Kudo ).
+The number and placement of crossovers affects the area of sister chromosome cohesion distal to crossovers which needs to be released for the first reductional chromosome segregation (@vanVeen2003, @LaneKauppi2019, @subramanian2014, @dumontDesai2012).
+Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell (Kudo ).
 While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dictyate arrest [@Lee2019].
@@
 Our results show that crossovers are spaced further apart in strains enriched for double-crossover bivalents when SC length is taken into account and chromosome size effects are minimized.
-Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and improve the fidelity of chromosomal segregation.
+Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and may improve the fidelity of chromosomal segregation.
 Although the SAC model postulates direct fitness effects of interference, a modifier model predicts that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
@@
 Females consistently showed longer SCs, even in strains with more recombination in males.
-Studies in mice (@lynn2002, @petkov2007) and humans [@gruhn2013] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+Studies in mice (@lynn2002, @petkov2007) and humans (@gruhn2013, @tease2004) suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+Some authors have suggested that the conserved sex differences in the crossover position (male telomere bias, female uniform placement) and genome levels of crossover interference strength (male stronger interference, female weaker interference), could be due to the loser vs tighter chromatin packing of the meiotic chromosome structure in females and males respectively (@haenel2018,  @petkov2007).
 A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes (@hulten2011_COM).
+More recent work has connected the sparser recombination landscape has been attributed to sex differences in the crossover repair pathway [@wang2017inefficient].
 
@@
 First, MLH1 foci only identify interfering crossovers [@holloway2008mus81].
-Although most crossovers (XX%; REF) belong to this class, our approach likely underestimated genome-wide recombination rates.
+Although most crossovers (80%; [@@holloway2008mus81]) belong to this class, our approach likely underestimated genome-wide recombination rates.
 Evolution of the number of non-interfering crossovers is a topic worth examining.

### 2020-06-04 src/Result_Setup.Rmd
updates from final v2 files

@@
 
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
 
@@
 
-MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
+MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
 sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
@@
 ```{r Male.specific.glm, echo=FALSE, eval=TRUE}
-
-#these are M4
 #M4
-MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data= male_mouse.avs_for.analysis)
+MLH1.male.glm.M4 <- glm(mean_co ~ subsp * strain, data=(male_mouse.avs_for.analysis  %>% filter(subsp != "Cast") )   )
+
 sum_MLH1.male.glm.M4 <- summary(MLH1.male.glm.M4)
-#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)
-
-#sum_MLH1.male.specific.glm.M2$coefficients[5,][4]
-
 #PWD, SKIVE, MSM
@@
 #M4.2
-MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= male_mouse.avs_for.analysis)
+MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= (male_mouse.avs_for.analysis  %>% filter(subsp != "Cast") ))
 sum_MLH1.male.glm.M4.2 <- summary(MLH1.male.glm.M4.2)
@@
 
+```{r Q2.IFD.setup, echo=FALSE}
+
+#identify and remove mice with 1 bivalent observation (MSM:  1feb18_MSM_m1)
+#some cells just have 1 cell, but the MSM_m is the only with just 1 bivlent
+
+```
+
 ```{r Q2.IFD.long.short.setup, echo=FALSE}
@@
 
-t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER,
-   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER   )
-
 #ABS
@@
 
-t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus"))$mean_IFD.2CO_ABS,
-   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS  )
+#niether PER or ABS is sig, 1 mouse, #1feb18_MSM_m1 has only 1 observation
+
+t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(mouse != "1feb18_MSM_m1")   )$mean_IFD.2CO_PER,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER   )
+
+t.test_mouse.av.IFD_ABS_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(mouse != "1feb18_MSM_m1") )$mean_IFD.2CO_ABS,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean_IFD.2CO_ABS  )
+
+```
+
+```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
+
+male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
+#t.tests for the Rec group
+Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
+                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
+                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
+
+#Curated_BivData <- add_IFD(Curated_BivData)
+#THIS SHOULD ALREADY BE ADDED
+
+#single biv level
+Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
+                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
+
+#Rec.group.BivData_IFD.PER_Ttest$p.value
+#p-value < 2.2e-16
+
+#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
+Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
+                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
+
+#Rec.group_IFD.PER_Ttest$p.value 
+#p-value = 6.872e-07
+```
+
+```{r Q2.show.IFD.Mol, echo=FALSE}
+
+#mouse.level.BivData_table
+#cleanCurated_BivData
+
+#molossinus does not show
+
+Musc.IFD.Biv <- ggplot(data = (cleanCurated_BivData %>% filter(subsp == "Musc") %>% filter(sex == "male") %>% filter(hand.foci.count == 2)  ), aes(x = strain, y = IFD1_PER) ) + geom_jitter()
+
+Mol.IFD <- ggplot(data = (cleanCurated_BivData %>% filter(subsp == "Mol") %>% filter(sex == "male") %>% filter(hand.foci.count == 2)  ), aes(x = strain, y = IFD1_PER) ) + geom_jitter() +facet_wrap(~mouse)
+
+t.test_biv.av.IFD_PER_MOL <- t.test(  (cleanCurated_BivData %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$IFD1_PER,
+   (cleanCurated_BivData %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$IFD1_PER   )
+#p value for the bivalent level IFD1 norm positions are significant
+
+#mouse level
+t.test_mouse.av.IFD_PER_MOL <- t.test(  (mouse.level.BivData_table %>% filter(Rec.group == "1") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER,
+   (mouse.level.BivData_table %>% filter(Rec.group == "0") %>% filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER   )
+
+#1feb18_MSM_m1
+
+Mol.mus.IFD <- ggplot(data = (mouse.level.BivData_table %>% filter(subsp == "Mol") %>% filter(sex == "male")   ), aes(x = strain, y = mean_IFD_PER, color = mouse) ) + geom_jitter()
+#there is mouse MSM with v low, 1feb18_MSM_m1 -- it only has 1 measure
+
 ```
@@
 
-```{r Q2.t.test_IFD, echo=FALSE, eval=FALSE}
-
-male.bivdata.2CO_IFD <- Curated_BivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
-#t.tests for the Rec group
-Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
-                                    ifelse(grepl("MSM male", Curated_BivData$category), 1,
-                                           ifelse(grepl("SKIVE male",Curated_BivData$category),1, 0)))
-
-#Curated_BivData <- add_IFD(Curated_BivData)
-#THIS SHOULD ALREADY BE ADDED
-
-#single biv level
-Rec.group.BivData_IFD.PER_Ttest <- t.test(male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 1], 
-                                 male.bivdata.2CO_IFD$IFD1_PER[male.bivdata.2CO_IFD$Rec.group == 0])
-
-#Rec.group.BivData_IFD.PER_Ttest$p.value
-#p-value < 2.2e-16
-
-#MOUSE AVE  -- replace with mouse    mouse.avs_4MM,   male.av_4MM
-Rec.group_IFD.PER_Ttest <- t.test(  (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$mean_IFD.2CO_PER, 
-                                (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_IFD.2CO_PER )
-
-#Rec.group_IFD.PER_Ttest$p.value 
-#p-value = 6.872e-07
-```
-
 ```{r Q2.IFD_short.removed_t.test, echo=FALSE, include=FALSE, eval=FALSE}

### 2020-06-05 doc/Intro.Rmd
small notes
starting the intro outline re-write

@@
 ---
-title: "Intro"
+title: "Intro Outline"
 author: "April Peterson"
@@
 
+**0.   Importance of recombination rate**
+•	The total number of COs per (4n meiotic) cell = genome wide recombination rate (gwRR)
+•	The gwRR regulates populations responses to selection, and determine the fate of novel mutations.
+•	This process shapes the genomic patterns of genetic variation.
+•	It is an integral part for proper chromosome segregation. With an obligate crossover per bivalent may act as the lower bound for gwRR.
+
+**1. Understanding levels of variation in gwRR**
+1a. There is substantial in gwRR variation within species (across populations) and (across individuals)
+•	While there are fewer measures from closely related species, variation at this level is more restricted to finer scales (hotspot level / recombination landscape level).
+1b. Sex is one of the most notable axes along which individuals vary 
+•	Long history since the discovery.
+•	Not due to sex chromosomes, but pattern of achiasmy evolution is different case.
+•	Most species have more recombination in females than males, but there are exceptions.
+•	Crossover placement is sexually dimorphic: male telomeric, female uniform placement
+•	SC length / meiotic chromosome length is longer in females of a few mammal species, but there are exceptions, Celegean (not much different) and A.thalnia (opposite direction) (Cahoon and Libdua).
+1c. There are still gaps missing from the field
+•	An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
+Comprehensive comparisons of female and male recombination rates usually come from outbred populations (human, dog, cattle, sheep, mouse collaborative cross REFS), in which the role of sex is confounded with the contributions of genetic variation.
+Although it is clear that the relationship between female and male recombination rates can differ among species, comparisons between and within closely related species are missing.
+•	Direct contrasts between females and males across a common, diverse set of genomic backgrounds would reveal whether the recombination rate evolves differently in the sexes.
+
+**2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale.**
+•	House mouse complex comes from a recent radiation providing an opportunity to interrogate variation at short evolutionary scales.
+•	Wild derived inbred strains generate the best comparison of females and males, besides the sex chromosomes, the mouse for each genome is highly similar.
+•	Unlike some house mouse strains, all strains have the same karyotypes, 20 pairs of acrocentric chromosomes.
+•	Classical lab strains of mice have generated a mountain of knowledge regarding the genetics and molecular pathway of meiosis.
+•	House mouse is suited for single cell cytology approaches.
+
+**3. What we accomplished in this paper**
+•	We report a rare, direct, evolutionary comparison of recombination rate in females and males (or something like this).
+•	We quantify gwRR of both sexes, from 3 subspecies and outgroups.
+•	We use rare strains with a recent origin from multiple geographic locations of the species territory.
+•	We quantified meiotic chromosome morphology (SC length) and placement of crossovers to comprise an approximate picture of the recombination landscape.
+•	Our results indicate rapid male specific evolution of gwRR.
+
 ## Background:
@@
 ### prev hetC lit:
+
 (Lenormand, BurtBellOther, Shapley,)

### 2020-06-08 Chap3_MLH1var.rmd
slow updates -revisions to intro

@@
 ---
+
+todo
+
+-update the variable names
+
+- re-run analyses with sd
+
+- Cecile's comment on transformations...
 
@@
 
-load(file="~./MLH1repo/data/Results_setup_4.24.20.RData")
+load(file="~./MLH1repo/data/Results_setup_5.30.20.RData")
 
@@
 
-load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
-load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
+#load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
+#load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
 
@@
 
-Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
+#TOTAL SC should already be loaded into Results
+#XXXXTotal.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
+
+#Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF_5.31.20.xlsx", header=TRUE, 1) #this new total_SC has the additional batches and outliers have been removed
 
@@
 
-Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse \nMLH1 variance") + scale_color_manual(values=colors_of_strains)
+Dom.female.VAR.plot <- ggplot( data = (female_mouse.avs_for.analysis %>% filter(subsp == "Dom") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse \nMLH1 variance") + scale_color_manual(values=colors_of_strains)
+
+```
+
+```{r mouse.SD.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+
+#variance plots
+musc.female.SD.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = sd, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse \nMLH1 s.d.")+scale_color_manual(values=colors_of_strains)
+musc.female.SD.plot
+
+Mol.female.SD.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = sd, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female\nwithin mouse MLH1 s.d.")+scale_color_manual(values=colors_of_strains)
+Mol.female.SD.plot
+
+Dom.female.SD.plot <- ggplot( data = (female_mouse.avs_for.analysis %>% filter(subsp == "Dom") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = sd, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse \nMLH1 s.d.") + scale_color_manual(values=colors_of_strains)
+Dom.female.SD.plot
 
@@
 
-M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
+M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
                         Nmice =length(unique(mouse)),
@@
 #same issue as the female table .. ugh
-
 ```
@@
 
+```
+
+```{r mouse.sd.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
+#variance plots
+OG.male.sd.plot <- ggplot(data = (male_mouse.avs_for.analysis %>% filter(subsp == "Dom")  ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =sd, color= strain, alpha=0.5), width = .2 )+ ggtitle("SD Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+OG.male.sd.plot
+
+Musc.male.sd.plot <- ggplot(data = (male_mouse.avs_for.analysis %>% filter(subsp == "Musc")  ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = sd, color= strain, alpha=0.5), width = .2 )+ ggtitle("SD Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Musc.male.sd.plot
+
+Mol.male.sd.plot <- ggplot(data = (male_mouse.avs_for.analysis %>% filter(subsp == "Mol")  ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = sd, color= strain, alpha=0.5), width = .2 )+ ggtitle("SD Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Mol.male.sd.plot
+
+Dom.male.sd.plot <- ggplot(data = (male_mouse.avs_for.analysis %>% filter(subsp == "Dom")  ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = sd, color= strain, alpha=0.5), width = .2 )+ ggtitle("SD Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
+Dom.male.sd.plot
 ```

### 2020-06-08 doc/Intro.Rmd
slow updates -revisions to intro

@@
     toc_float: yes
-bibliography: refs_Intro.bib
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
+bibliography: refs_Main_Report.bib
 ---
 
-# Outline
+### **0.   Importance of recombination rate**
 
-**0.   Importance of recombination rate**
-•	The total number of COs per (4n meiotic) cell = genome wide recombination rate (gwRR)
-•	The gwRR regulates populations responses to selection, and determine the fate of novel mutations.
-•	This process shapes the genomic patterns of genetic variation.
-•	It is an integral part for proper chromosome segregation. With an obligate crossover per bivalent may act as the lower bound for gwRR.
+(1) Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
+The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
+The total number of COs per cell (at the 4n stage) is equal to the genome wide recombination rate (gwRR).
+The gwRR regulates populations responses to selection, and determine the fate of novel mutations by transfering beneficial mutations onto novel genetic backgrounds or by breaking linkage of bad mutations from benefical backgrounds.
 
-**1. Understanding levels of variation in gwRR**
+This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation (Begun Aquadro, Nachmann, Payseur).
+
+(mention fitness more explicitly).
+The lower bound/threshold is thought to be an obligate crossover per homolog pair -- is a checkpoint for proper chromosome segregation (Hunt Nagaoka, Hassold).
+While too much recombination (ectopoic recombination) and DNA damage preceeding crossover formation (may be a constraint for an upper threshold.
+(cite other than Inoue, OttoPaysuer?).
+**As the gwRR influences the evolution of other traits, it is important to document and understand the evolutionary patterns for the genome wide recombination rate itself.**
+
+# **1. Understanding levels of variation in gwRR**
+
 1a. There is substantial in gwRR variation within species (across populations) and (across individuals)
 •	While there are fewer measures from closely related species, variation at this level is more restricted to finer scales (hotspot level / recombination landscape level).
+
 1b. Sex is one of the most notable axes along which individuals vary
-•	Long history since the discovery.
-•	Not due to sex chromosomes, but pattern of achiasmy evolution is different case.
-•	Most species have more recombination in females than males, but there are exceptions.
+
+•	Long history since the discovery -- in fruit flies and silk moths
+Heterochiasmy was first documented ---
+
+(2) One forms of sexual reproduction, anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates.
+Yet sexual dimorphism for the gwRR, called heterochiasmy, is commonly observed, suggesting that other meiotic traits which distinguish the gametes may impose selection for sexually dimorphic recombination rates.
+
+As detailed in Sardell Kirkpatrick, (for a species with hetreochiasmy, rarely do they have a single recombination landscape. Instead there are broad scale differences recombination rate along
+
+•	Most species have more recombination in females than males,
+
 •	Crossover placement is sexually dimorphic: male telomeric, female uniform placement
+
 •	SC length / meiotic chromosome length is longer in females of a few mammal species, but there are exceptions, Celegean (not much different) and A.thalnia (opposite direction) (Cahoon and Libdua).
+
+The cellular process of meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
+The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
+**Under one of the most common forms of sexual reproduction, anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates.** Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
+
+As reviewed in Cahoon Libuda, there are fundamental differences for the two forms of gametogenesis: length of timing (spermatogenetsis faster), sensativeity to errors, asymetrical cell division vs symetrical cell devision.
+The physiological consequences of these cell volumne may have an impact meiotic machenery.
+
+the authors track the sexual dimorphism patterns in the recombination landscape for three species.
+
 1c. There are still gaps missing from the field
@@
 
-**2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale.**
+# **2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale.**
 •	House mouse complex comes from a recent radiation providing an opportunity to interrogate variation at short evolutionary scales.
@@
 
-**3. What we accomplished in this paper**
+# **3. What we accomplished in this paper**
 •	We report a rare, direct, evolutionary comparison of recombination rate in females and males (or something like this).
@@
 •	Our results indicate rapid male specific evolution of gwRR.
-
-## Background:
-
-### prev hetC lit:
 
@@
 
-## gap, angle
-
 # text
@@
 Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
-
-# Notes and Ideas
-
-# Background
-
-## natural variation in RR
-
-empiracal measures -- important / useful for understanding how rec evolves with narutal selection
-
-sources from Ritz Noor,
-Samuk et al 2019, (selection,)
-(barley in north africa),  Dreissig et al 2019
-
-humans (not very natural population)
-
-## empiracal measures for hetC / theories for evo
-
-Lenormand Duthiel,
-Shapley / stapley et al
-Sardell Kirkpatrick  
-Gruhn + hassold
-
-Hoffman papers -- strange inverted meiosis
-Uniparental disomy
--- evidence that RR-constraint connection may be not straight forward.
-
-### lack of female data / sig of female data
-
-human annuploidy / fertility
-
-# Bibliography

### 2020-06-08 doc/Intro.Rmd
small changes to v2

@@
 
-(1) Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
+Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
 The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
@@
 
-(mention fitness more explicitly).
-The lower bound/threshold is thought to be an obligate crossover per homolog pair -- is a checkpoint for proper chromosome segregation (Hunt Nagaoka, Hassold).
+<(mention fitness more explicitly?)> The lower bound/threshold is thought to be an obligate crossover per homolog pair -- is a checkpoint for proper chromosome segregation (Hunt Nagaoka, Hassold).
 While too much recombination (ectopoic recombination) and DNA damage preceeding crossover formation (may be a constraint for an upper threshold.
@@
 1a. There is substantial in gwRR variation within species (across populations) and (across individuals)
+
 •	While there are fewer measures from closely related species, variation at this level is more restricted to finer scales (hotspot level / recombination landscape level).
+**find a good citation for this**
 
@@
 
-•	Long history since the discovery -- in fruit flies and silk moths
+•	Long history since the discovery -- in fruit flies and silk moths at the turn of the 20th century (cite sturavant)
 Heterochiasmy was first documented ---
@@
 
-As detailed in Sardell Kirkpatrick, (for a species with hetreochiasmy, rarely do they have a single recombination landscape. Instead there are broad scale differences recombination rate along
+A early pattern of heterochiasmy that emerged was that in many species females have more recombination compared to males, considered female bias-heterochiasmy.
+However as observations have expanded to more species, more examples of male biased heterochiasmy have been observed; including cattle, marsupials, (fish)  (while birds have a smaller overall degree of heterochiasmy)
 
-•	Most species have more recombination in females than males,
+A more conserved feature of heterochiasmy As discussed in Sardell Kirkpatrick, are the broad scle differences in the placement of crossovers along chromosomes.
 
-•	Crossover placement is sexually dimorphic: male telomeric, female uniform placement
+In male recombination landscapes, there is a bias towards the ends near the telomeres while in females crossovers are more uniformly placed (and occur closer to the centromere compared to males).
 
-•	SC length / meiotic chromosome length is longer in females of a few mammal species, but there are exceptions, Celegean (not much different) and A.thalnia (opposite direction) (Cahoon and Libdua).
-
-The cellular process of meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
-The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
-**Under one of the most common forms of sexual reproduction, anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates.** Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
+(the degree of this pattern varies across species (some have female biased telomere-bias, some have chromosome specific patterns and others have no significant sex difference.
+It's worth pointing out that the exceptions may require modifed assumptions as to how the meiotic program evolves; marsupials form chromosome chains during meiosis, some fish species has sequential hermaphrodism, and birds have many microchromosomes chromosomes.
+<grasshoppers?>
 
@@
 the authors track the sexual dimorphism patterns in the recombination landscape for three species.
+
+<SC length / meiotic chromosome length is longer in females of a few mammal species, but there are exceptions, Celegean (not much different) and A.thalnia (opposite direction) (Cahoon and Libdua).>
 
@@
 •	House mouse is suited for single cell cytology approaches.
+
+Overall house mouse -- using the house mouse model -- will better allow narrowing in on the causes // developing testable hypotheses for explaining any (particular patterns) variation discovered.
 
@@
 Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
+
+The cellular process of meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
+The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
+**Under one of the most common forms of sexual reproduction, anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates.** Yet sexual dimorphism for this trait, called heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.

### 2020-06-09 src/Result_Setup.Rmd
reducing redundancies standardizing DFs

@@
 
-#this dataframe doesn't work
-
-#don't run here, run in the MS file
-#DF.HetC.MixedModel.HQ
-#DF.HetC.MixedModel.HQ <- subset()
-
 MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
@@
 MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value
+MLH1.MM_strain <- exactRLRT(MLH1.MM.M1)$p.value
 
@@
 
-```{r Female.specific.glm, echo=FALSE, eval=TRUE}
+```{r Female.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
 
@@
 
-MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
-sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)
+MLH1.female.spc.glm.M5 <- glm(mean_co ~ strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
+sum_MLH1.female.spc.glm.M5 <- summary(MLH1.female.spc.glm.M5)
 
 #round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-
 #write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
@@
 
-```{r Male.specific.glm, echo=FALSE, eval=TRUE}
+```{r Male.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
 #M4
@@
 
-# Q1 single bivalent section and total SC
+# Q1 single bivalent
 
@@
 
-```{r Q1.total.SC_t.test, echo=FALSE}
-
-#run t-test for male female avergage for each strain
-
-#ALL
-Q1_tot.sc_t.test.all <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(species == "M.musculus")  )$mean.skel)
-
-#dom
-Q1_tot.sc_t.test.DOM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)
-
-#WSB
-Q1_tot.sc_t.test.WSB <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean.skel)
-#G
-Q1_tot.sc_t.test.G <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "G") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "G") )$mean.skel)
-#LEW
-Q1_tot.sc_t.test.LEW <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean.skel)
-
-#Musc
-Q1_tot.sc_t.test.MUSC <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)
-
-#PWD
-Q1_tot.sc_t.test.PWD <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean.skel)
-
-#KAZ
-Q1_tot.sc_t.test.KAZ <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean.skel)
-
-#SKIVE
-#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
-#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
-#cell level total SC
-#MLH1.merge.Skel   $skel_size
-#error not enought female observations, ONLY 1 female SKIVE mouse
-
-Q1_tot.scBIV_t.test.SKIVE <- t.test((Total.SC_DF %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
-           (Total.SC_DF %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )
-
-#MSM
-Q1_tot.sc_t.test.MSM <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean.skel,
-                           (mouse.total.skel %>% filter(sex == "female") %>% filter(strain == "MSM") )$mean.skel)
-
-```
-
-```{r Q1.totSC_male.first.comp, echo=FALSE}
-
-Q1.male.Dom.Musc.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
-
-Q1.male.Dom.Mol.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
-(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
-```
-
-```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}
-
-Q1.lmer_totSC.mouse.av.table <- mouse.total.skel %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
-
-#str(Q1.lmer_totSC.mouse.av.table)
-
-#unorder factors
-Q1.lmer_totSC.mouse.av.table$subsp <- factor(Q1.lmer_totSC.mouse.av.table$subsp, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$sex <- factor(Q1.lmer_totSC.mouse.av.table$sex, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$strain <- factor(Q1.lmer_totSC.mouse.av.table$strain, ordered = FALSE)
-Q1.lmer_totSC.mouse.av.table$species <- factor(Q1.lmer_totSC.mouse.av.table$species, ordered = FALSE)
-
-#TOTAL SC AREA
-lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)
-
-lmer.Q1_mus.av_totSC_interaction  <-drop1(lmer.Q1_mus.av_totSC, test="Chisq" )$`Pr(Chi)`[2]
-
-lmer.Q1_mus.av_totSC_sex  <-anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.sex)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_subp  <- anova(lmer.Q1_mus.av_totSC, lmer.Q1_mus.av_totSC_R.subsp)$`Pr(>Chisq)`[2]
-
-lmer.Q1_mus.av_totSC_strain  <- exactRLRT(lmer.Q1_mus.av_totSC)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1.lmer.totSC_results <- data.frame(sub.sp = lmer.Q1_mus.av_totSC_subp, 
-sex.results =  lmer.Q1_mus.av_totSC_sex, 
-               interaction.pval = lmer.Q1_mus.av_totSC_interaction,  rand.pval = 
-lmer.Q1_mus.av_totSC_strain)
-
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
-
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-
-```
-
 mouse total sc t.tests male female (Q1)
@@
 
-```{r Q1_SC.biv.short_ttest99, echo=FALSE, eval=FALSE}
-#this show is re-dundant
-#
-#run t.tests for the mouse av. for reduced single bivalent datasets
-
-#WSB
-Q1_mus.av_shortSC_t.test.WSB <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
-
-#G
-Q1_mus.av_shortSC_t.test.G <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "G") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "G") )$mean_SC)
-
-#LEW
-Q1_mus.av_shortSC_t.test.LEW <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$mean_SC)
-
-#PWD
-Q1_mus.av_shortSC_t.test.PWD <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$mean_SC)
-
-#KAZ
-Q1_mus.av_shortSC_t.test.KAZ <- t.test((Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$mean_SC,
-                           (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$mean_SC)
-
-#SKIVE - this value used in the text
-#only one SKVIE female, this is breaking the test
-#Q1_mus.av_shortSC_t.test.SKIVE <- t.test(
-
-#      (Short.biv_mouse_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean_SC,
-#      (Short.biv_mouse_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean_SC
-#      )
-
-```
-
 ```{r Q1_short_ratios, echo=FALSE, eval=TRUE}
@@
 ```
-
-```{r totSC_setup, echo=FALSE, eval = F}
-
-#Total.SC_DF is the main dataframe and should be set up cleanly in the SetUpData.Rmd
-#SC.skel.OG <- SC.skel #5683
-
-tot.sc.x.list <- Total.SC_DF %>% filter(X == "X")
-
-Total.SC_DF.og <- Total.SC_DF[! Total.SC_DF$Random.Name %in% tot.sc.x.list$Random.Name, ]
-#quality is still a factor
-str(Total.SC_DF.og)
-Total.SC_DF.og$quality <- as.numeric(Total.SC_DF.og$quality)
-
-Total.SC_DF.og <- Total.SC_DF.og %>% filter(quality < 5)
-
-OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF
-
-#check for DUPs, #check for duplicatations of the random name
-MLH1.merge.Skel$fileName  <- as.character(MLH1.merge.Skel$fileName)
-tot.SC.dupies <- MLH1.merge.Skel[duplicated(MLH1.merge.Skel$ fileName ),]
-
-MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")
-
-MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
-MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)
-#table(MLH1.merge.Skel$category)
-
-```
-
--- The total_SC outliers should be removed in the total.SC RData script
 
@@
 
-where is total_SC M1?
-
 ```{r Q1.total.SC_M1, echo=FALSE}
@@
 sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)
-sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,]
+#sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,]
 
@@
 
-```{r Q1_short.M1, echo=FALSE}
+```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
+
+lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
+
+Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
+Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
+Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
+
+#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
+
+Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
+sex.results =  Q1lmer.MM_SC_short_sex, 
+               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
+Q1lmer.MM_SC_short_strain)
+
+#Q1lmer.MM_SC_short_results
+
+```
+
+```{r Q1_SC_short.M1, echo=FALSE}
 
@@
 ```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-
 Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
@@
 #summary(Q1.glm_short.SC_M3)
-
 summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)
@@
 #sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][8,]
-
 #SKIVE * male
@@
 
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
 #nested boxplots, pooled Bivalents
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+Q1.SC.length_Nested_boxplots_Dom <- ggplot(cleanCurated_BivData[cleanCurated_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
 
@@
 
-```{r Q1.F1pos_long.short_setup, echo=FALSE, eval=FALSE}
-#move to setup
-
-Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
-
-```
-
 ```{r CO.positions.figs, echo=FALSE, include=FALSE, warning=FALSE}
@@
 
+#SKIVE
+Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
+                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos)
+
 #MSM
@@
 ```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
-
-#make sure this dataset has all the right strains
-Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")
 
@@
 # Q1 IFD
+
+```{r Q1.IFD_Ttests_ABS, echo=FALSE, warning=FALSE, message=FALSE}
+#below are calq just from 2COs, have no fear
+#mean_IFD.2CO_ABS
+#mean_IFD.2CO_PER
+
+#FULL all strains
+Q1.IFD.ABS_FULL_t.test <- t.test(   (mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_ABS,
+
+(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_ABS    ) #0.03 (female larger)
+#this comparison should be within subspecies due to strain, SC differences
+
+#WSB
+Q1.IFD.raw_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
+                                (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
+
+#LEW
+Q1.IFD.raw_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
+
+#G
+Q1.IFD.raw_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
+
+#PWD
+Q1.IFD.raw_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
+
+#KAZ - low KAZ male curation (3 mice)
+Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
+
+#SKIVE
+Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
+
+#MSM - 3 male 4 female measures
+#only 3 MSM males
+#could be missing MSM measures, (4 females, 3 males)
+Q1.IFD.raw_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
+
+#FULL all strains
+Q1.IFD.raw_all <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
+
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
+#p-value = 0.3
+```
+
+```{r Q1.IFD_ttest_PER, echo=FALSE}
+
+Q1.IFD.PER_FULL_t.test <- t.test(   (mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_PER,
+
+(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_PER    ) #p-value = 1e-10
+
+#WSB
+Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER )#p-value = 0.03
+
+#LEW
+Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)#p-value = 7e-05
+
+#G
+Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER )#p-value = 0.03
+
+#PWD
+Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)#p-value = 9e-04
+
+#KAZ
+Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)#p-value = 0.3
+
+#MSM
+Q1.IFD.PER_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
+
+#SKIVE
+Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
+                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
+
+```
+
+```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
+
+#M1_IFD_2CO_ABS
+lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") )) 
+lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+
+ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
+
+Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
+               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
+
+#M1_IFD_2CO
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) ) 
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") )  )
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+
+sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
+subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
+subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
+strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
+
+Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
+               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
+
+```
+
+```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
+#mouse.avs_4MM -- is also the bivbased mouse average table 
+#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
+
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") )  )
+summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
+
+#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
+
+#PER
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
+
+```
+
+```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
+
+glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ))
+summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
+
+#PER
+glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
+#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
+
+#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+```
+
+```{r Q1.IFD_glm.M4.M5_female, echo=FALSE}
+
+glm.Q1_IFD.ABS.M4.female  <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(sex == "female") ) )
+summry.glm.Q1_IFD.ABS.M4.female  <- summary(glm.Q1_IFD.ABS.M4.female)
+
+glm.Q1_IFD.PER.M4.female  <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(sex == "female") ) )
+summry.glm.Q1_IFD.PER.M4.female  <- summary(glm.Q1_IFD.PER.M4.female)
+
+#M5
+glm.Q1_IFD.ABS.M5.female  <- glm(mean_IFD.2CO_PER ~  strain, data= (mouse.level.BivData_table %>% filter(sex == "female") ) )
+summry.glm.Q1_IFD.ABS.M5.female  <- summary(glm.Q1_IFD.ABS.M4.female)
+
+glm.Q1_IFD.PER.M5.female  <- glm(mean_IFD.2CO_PER ~  strain, data= (mouse.level.BivData_table %>% filter(sex == "female") ) )
+summry.glm.Q1_IFD.PER.M5.female  <- summary(glm.Q1_IFD.PER.M5.female)
+
+```
+
+```{r Q1.IFD_glm.M4.M5_male, echo=FALSE}
+
+glm.Q1_IFD.ABS.M4.male  <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(sex == "male") ) )
+summry.glm.Q1_IFD.ABS.M4.male  <- summary(glm.Q1_IFD.ABS.M4.male)
+
+glm.Q1_IFD.PER.M4.male  <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(sex == "male") ) )
+summry.glm.Q1_IFD.PER.M4.male  <- summary(glm.Q1_IFD.PER.M4.male)
+
+#M5
+glm.Q1_IFD.ABS.M5.male  <- glm(mean_IFD.2CO_PER ~  strain, data= (mouse.level.BivData_table %>% filter(sex == "male") ) )
+summry.glm.Q1_IFD.ABS.M5.male  <- summary(glm.Q1_IFD.ABS.M4.male)
+
+glm.Q1_IFD.PER.M5.male  <- glm(mean_IFD.2CO_PER ~  strain, data= (mouse.level.BivData_table %>% filter(sex == "male") ) )
+summry.glm.Q1_IFD.PER.M5.male  <- summary(glm.Q1_IFD.PER.M5.male)
+
+```
+
+```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
+
+#these are the scatter plots
+
+nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
+#looked up image, measures confirmed
+
+####
+nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#21may18_KAZ_m2 -- seems like it might be an outlier
+#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
+#second foci slightly smaller, kinda fainter -- but also much brighter than background
+
+#check all KAZ males
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (cleanCurated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_KAZ.female_BIV.data <- ggplot( (cleanCurated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ.female_BIV.data
+
+nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD.female_BIV.data
+
+nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_LEW.female_BIV.data
+
+nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G.female_BIV.data
+
+nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
+
+raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+###look at why KAZ males are weird
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#the difference the the Nrm IFD cut off is striking with these plots
+nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
+
+```
 
@@
 
-```{r Q1.IFD_Ttests_ABS, echo=FALSE, warning=FALSE, message=FALSE}
-#below are calq just from 2COs, have no fear
-#mean_IFD.2CO_ABS
-#mean_IFD.2CO_PER
-
-#FULL all strains
-Q1.IFD.ABS_FULL_t.test <- t.test(   (mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_ABS,
-
-(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_ABS    ) #0.03 (female larger)
-#this comparison should be within subspecies due to strain, SC differences
-
-#WSB
-Q1.IFD.raw_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
-                                (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)
-
-#LEW
-Q1.IFD.raw_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)
-
-#G
-Q1.IFD.raw_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)
-
-#PWD
-Q1.IFD.raw_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)
-
-#KAZ - low KAZ male curation (3 mice)
-Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)
-
-#SKIVE
-Q1.IFD.raw_t.test_SKIVE <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_ABS)
-
-#MSM - 3 male 4 female measures
-#only 3 MSM males
-#could be missing MSM measures, (4 females, 3 males)
-Q1.IFD.raw_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_ABS,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_ABS)
-
-#FULL all strains
-Q1.IFD.raw_all <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>%  filter(strain !="G") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS,
-
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>%  filter(strain !="G")%>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
-#p-value = 0.3
-```
-
-```{r Q1.IFD_ttest_PER, echo=FALSE}
-
-Q1.IFD.PER_FULL_t.test <- t.test(   (mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "male") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_PER,
-
-(mouse.level.BivData_table %>% filter(species == "M.musculus")  %>%  filter(sex == "female") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )$mean_IFD.2CO_PER    ) #p-value = 1e-10
-
-#WSB
-Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER )#p-value = 0.03
-
-#LEW
-Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)#p-value = 7e-05
-
-#G
-Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER )#p-value = 0.03
-
-#PWD
-Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)#p-value = 9e-04
-
-#KAZ
-Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)#p-value = 0.3
-
-#MSM
-Q1.IFD.PER_t.test_MSM <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="MSM" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="MSM"))$mean_IFD.2CO_PER)
-
-#SKIVE
-Q1.IFD.PER_t.test_SKIVE <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="SKIVE" ))$mean_IFD.2CO_PER,
-                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="SKIVE"))$mean_IFD.2CO_PER)
-
-```
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
-
-#these are the scatter plots
-
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
-
-```
-
-```{r Q1.IFD.lmer.M1, echo=FALSE, message=FALSE, warning=FALSE}
-
-#M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-ABS.IFD_Q1.M1_sex  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_subsp  <- anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-ABS.IFD_Q1.M1_intr  <- drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-ABS.IFD_Q1.M1_strain  <- exactRLRT(lmer.IFD.ABS_MM.M1)$p.value
-
-Q1.IFD.ABS_M1_lmer.results <- data.frame(sub.sp = ABS.IFD_Q1.M1_subsp, sex.results =  ABS.IFD_Q1.M1_sex, 
-               interaction.pval = ABS.IFD_Q1.M1_intr,  rand.pval = ABS.IFD_Q1.M1_strain)
-
-#M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )
-
-sex.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
-subsp.pval <- anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
-subsp.sex.pval <- drop1(lmer.IFD.PER_MM.M1, test="Chisq" )$`Pr(Chi)`[2]
-strain.pval <- exactRLRT(lmer.IFD.PER_MM.M1)$p.value
-
-Q1.IFD.PER_M1_lmer.results <- data.frame(sub.sp = subsp.pval, sex.results =  sex.pval, 
-               interaction.pval = subsp.sex.pval,  rand.pval = strain.pval)
-
-```
-
-```{r Q1.IFD_glm.M2, echo=FALSE, warning=FALSE, message=FALSE}
-#mouse.avs_4MM -- is also the bivbased mouse average table 
-#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM  %>% filter(species == "M.musculus")  )
-summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
-
-glm.Q1_IFD.ABS.M2_kindaFU  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=mouse.level.BivData_table  %>% filter(species == "M.musculus")  )
-summry.glm.Q1_IFD.ABS.M2_kindaFU <- summary(glm.Q1_IFD.ABS.M2_kindaFU)
-
-#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
-
-#PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= (mouse.level.BivData_table %>% filter(species == "M.musculus") ) )
-summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
-```
-
-```{r Q1.IFD_glmM3, echo=FALSE, warning=FALSE, message=FALSE}
-
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
-
-#PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
-summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
-#round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)
-
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
-```
-
-```{r Q1.IFD.long.short.setup, echo=FALSE, eval=FALSE}
-
-#real.long.bivData
-#real.short.bivData
-
-Q1.IFD.long <- ddply(real.long.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.long <- add_sex(Q1.IFD.long)
-Q1.IFD.long <- add_strain(Q1.IFD.long)
-Q1.IFD.long <- add_subsp(Q1.IFD.long)
-
-Q1.IFD.long$strain <- factor(Q1.IFD.long$strain, ordered = FALSE)
-Q1.IFD.long$subsp <- factor(Q1.IFD.long$subsp, ordered = FALSE)
-
-Q1.IFD.short <- ddply(real.short.bivData, c("mouse"), summarise,
-
-                        n_obs = length(Obj.ID),
-                         mean.SC = mean(chromosomeLength),
-
-                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
-
-                     )
-
-Q1.IFD.short <- add_sex(Q1.IFD.short)
-Q1.IFD.short <- add_strain(Q1.IFD.short)
-Q1.IFD.short <- add_subsp(Q1.IFD.short)
-
-Q1.IFD.short$strain <- factor(Q1.IFD.short$strain, ordered = FALSE)
-Q1.IFD.short$subsp <- factor(Q1.IFD.short$subsp, ordered = FALSE)
-
-#t.test. mouse av
-Q1.IFD.raw_short.ttest_KAZ <-  t.test( (Q1.IFD.short %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.short %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-
-Q1.IFD.raw_long.ttest_KAZ <- t.test( (Q1.IFD.long %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$meanIFD2CO_norm,
-                                ( Q1.IFD.long %>% filter(sex == "female") %>% filter(strain =="KAZ"))$meanIFD2CO_norm)
-#both the short and long norm IFD are NS
-
-#t.test Biv
-Q1.IFD.raw_BIV.long.ttest_KAZ <-  t.test( (real.long.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.long.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-
-Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
-                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
-```
-
 #Q2 section
@@
 #Q2 SC
+
+```{r Q2.totSC_male.first.comp, echo=FALSE}
+#move to Q2
+Q1.male.Dom.Musc.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel)
+
+Q1.male.Dom.Mol.t.test <- t.test( ( mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
+(mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Mol") )$mean.skel)
+```
 
@@
 
-```{r Q2totalSC_checkdates, echo=FALSE, eval=FALSE}
-
-#investigating potential skew / technical effects in the algorithm
-
-Neg.control <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")
-#not sure what my expectations for this plot are,
-
-median(bep$oldest.image.date) #"2017-08-03"
-#there might be an issue with the algorithm performance 
-#given the parrelle diagonal lines -- 
-#this might be due to the dates imaged  2017-02-07
-
-bep$date.cut.off <- ifelse( bep$oldest.image.date < "2017-02-07", 0, 1 )
-
-Neg.control.find.date <- ggplot(bep, aes(x=mean.skel, y=mean.bin, color=strain))+geom_point()+
-  scale_color_manual(values=colors_of_strains)+ggtitle("Total SC by mean CO\nMouse Means")+facet_wrap(~date.cut.off)
-
-#there's evidence that the algorithm might have some biased effects -- 
-#that will be investigated later.
-
-```
-
 ```{r Q2.totSC.ttest, echo=FALSE}
@@
 
-The molf male t-tests don't work because I only have total Sc measures for 1 MOLF male. to of the
-
-```{r Q2.totSC.ttest_old, echo=FALSE, eval=FALSE}
-
-#move this to MS_v2?
-
-#totalSC high and low rec group
-
-Q2tot.sc_t.test_High.Low <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
-#p-value = 0.007
-
-#mol -- not enough Mol observations
-#Q2tot.sc_t.test_High.Low_Mol <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-#         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
-
-#musc
-Q2tot.sc_t.test_High.Low_Musc <- t.test((mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Musc"))$mean.skel,
-                (mouse.total.skel %>% filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1") )$mean.skel)
-
-```
-
-todo list if I get back to lab.
+The molf male t-tests don't work because I only have total Sc measures for 1 MOLF male.
 
@@
 
-Q2.all.BIV_t.test_HL <- t.test((Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
-              (Curated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
+Q2.all.BIV_t.test_HL <- t.test((cleanCurated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "0") )$chromosomeLength,
+              (cleanCurated_BivData %>%filter(sex == "male")%>%filter(Rec.group == "1"))$chromosomeLength)
 
@@
 #DOM-MUSC-ALL
-Q2.all.BIV_t.test_DomMusc <- t.test((Curated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
+Q2.all.BIV_t.test_DomMusc <- t.test((cleanCurated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (cleanCurated_BivData %>% filter(subsp == "Musc")%>%filter(sex == "male"))$chromosomeLength)
 #p-value = 4e-13
@@
 #DOM-MOL-ALL
-Q2.all.BIV_t.test_DomMol <- t.test((Curated_BivData%>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
-              (Curated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
+Q2.all.BIV_t.test_DomMol <- t.test((cleanCurated_BivData %>%filter(subsp == "Dom")%>%filter(sex == "male") )$chromosomeLength,
+              (cleanCurated_BivData %>% filter(subsp == "Mol")%>%filter(sex == "male"))$chromosomeLength)
 #p-value = 6e-10
@@
 
-Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
-Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
+Q2.SC_BivData_low.rec <- cleanCurated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
+Q2.SC_BivData_high.rec <- cleanCurated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )
 
@@
 
-Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
-Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]
-
-#adjust order of plot here
-Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
-  "WSB","G","LEW","PERC",  
-"PWD", "SKIVE",  "KAZ",
-     "MSM", "MOLF",   
-                     "CAST",
-                         "SPRET","SPIC" ), order=T )
-
-Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
+Q2.SC_BivData <- cleanCurated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)
 
@@
 Rec.group_BivData_F1post_Ttest <- t.test(
-  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
-(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
-#p-value = 0.2
-
-#
-#whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count #== 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
-#            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   #%>% filter(Rec.group == "1"))$PER_Foci_1   )
-#p-value = 0.2
-
-```
-
-- COpos short long t.tests
+  (cleanCurated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
+(cleanCurated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
+#p-value = 0.3
+
+```
 
@@
 
-mouse.level.BivData_table <- add_species(mouse.level.BivData_table)
-
-#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos
-
-#none of these are M1 -- since they are male specific 
-#all of these are M4
-#mouse table SC
-#mouse.level.BivData_table .. mouse.avs_4MM
-
-#why won't this work
 Q2.glm_mus.av_1COpos_M4 <- glm(mean_1CO.pos.PER ~ subsp * strain, data = (mouse.level.BivData_table  %>% filter(sex == "male") %>% filter(species == "M.musculus") ) )
@@
 #summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
-
-#with.chrm.legth  ?
-#Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
 

### 2020-06-09 src/SetupData.Rmd
reducing redundancies standardizing DFs

@@
 
-```{r remove.non.foci, echo=FALSE}
-#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36
-str(Curated_BivData)
-table(Curated_BivData$hand.foci.count, exclude = NULL)
-#there are ~600 rows with SC and no hand foci count
-table( (Curated_BivData %>% filter(  is.na(hand.foci.count) ) )$category )
-#KAZ male and WSB female have >100 observations without hand.foci.counts
-
-#Remove bivalent entries without hand foci count values
-cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
-                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
-
-length( (Curated_BivData  %>%  filter(  is.na(hand.foci.count) == T ) )$fileName )
-
-table(cleanCurated_BivData$hand.foci.count, exclude = NULL)
-#600 removed
-```
-
 ```{r format.str.Curate_BivData, echo=FALSE}
@@
                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
+
+cleanCurated_BivData <- cleanCurated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14") #this object has wrongg position data
 
@@
 
-Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
+Mouse.Table_BivData_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                         Nmice = length(unique(mouse)),
@@
                           mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
-                          var.MLH1 = var(nMLH1.foci),
-
-                          mean.skel = mean(skel_size),
-                          var.skel = var(skel_size),
-
-                          mean.bin = mean(bin_size)
+                          var.MLH1 = var(nMLH1.foci, na.rm = TRUE),
+
+                          mean.skel = mean(skel_size, na.rm = TRUE),
+                          var.skel = var(skel_size, na.rm = TRUE),
+
+                          mean.bin = mean(bin_size, na.rm = TRUE)
 )

### 2020-06-11 doc/MS_v2.Rmd
tracked down difference due to M1 p values

@@
 output:
-  word_document:
-  pdf_document:
-    toc: yes
-    toc_depth: '5'
   html_document:
@@
     toc_float: yes
+  pdf_document:
+    toc: yes
+    toc_depth: '5'
+  word_document:
 always_allow_html: yes
@@
 After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
+A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
 In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
@@
 We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4).
-In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).
+In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).
 
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ).
+Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`), _molossinus^MSM^_  (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ), and _musculus^PWD^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]` ).
 Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.

### 2020-06-11 src/Result_Setup.Rmd
tracked down difference due to M1 p values

@@
 Dom.male_mouse.av <-  male_mouse.avs_for.analysis %>% filter(sex== "male") %>% filter(subsp == "Dom")
-
-```
-
-```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
-
-#DF.HetC.MixedModel <- add_species(DF.HetC.MixedModel)
-DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]
-#DF.HetC.MixedModel
-
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel
-
-#main df with variables coded as chr, set false for the model
-DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
-DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
-DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
-DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
-
-#remove male only data
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
 
@@
 HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
-#unique(HetC.HMI$mouse)
-#non-passing mice
-#X.remove.mice <- c("", "")
-#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
-#calq sex mean
 HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
@@
 HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
-#unique(HetC.PERC$mouse)
-#non-passing mice
-#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
-#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]
 
@@
 
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
-
 #calq sex mean
@@
 )
-
 CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]
@@
 #unique(HetC.SKIVE$mouse)
-
-#non-passing mice
-#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
-#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]
 
@@
 HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
-unique(HetC.MOLF$mouse)
 
@@
 )
-
 MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
-
-```
-
-```{r MLH1.HetC.plot, echo=FALSE, warning=FALSE}
+```
+
+```{r DF.MLH1.HetC.plot, echo=FALSE, warning=FALSE, eval=FALSE}
+
+#I don't think this block is needed.
+
 #HETC plot
@@
                          Ncells  = length( nMLH1.foci),
-                         mean_co = format(round(  mean( nMLH1.foci), 3 ), nsmall=3),
-                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
-                         sd   = round(sd( nMLH1.foci), 3),
+                         mean_co = format(round(  mean( nMLH1.foci, na.rm = TRUE), 3 ), nsmall=3),
+                         var = format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3),
+                         sd   = round(sd( nMLH1.foci, na.rm = TRUE), 3),
                          se   = round(sd / sqrt(Ncells), 3),
@@
 #HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
-
 #remove the 'other'
-
 HetC_strain_table.M$sex <- "male"
@@
 HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
-
 HetC_strain_table <- add_subsp(HetC_strain_table)
-
-#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")
 
@@
 #str(HetC_strain_table)
-
 ```
@@
 
-                        Nmice = length(Nmice),
-                        Ncells  = sum(Ncells), #might not to be sum
-                        mean.SC = mean(mean_SC),
-                        var.SC = mean(var), #mouse average variance of bivalent per mouse
-
+                       # Nmice = length(Nmice, na.rm = TRUE),
+                        Ncells  = sum(Ncells, na.rm = TRUE), #might not to be sum
+                        mean.SC = mean(mean_SC, na.rm =TRUE),
+                        var.SC = mean(var, na.rm = TRUE), #mouse average variance of bivalent per mouse
                         #chrm classes
-                        av.proportion_0CO = mean(proportion_0CO),
-                        av.proportion_1CO = mean(proportion_1CO),
-                        av.proportion_2CO = mean(proportion_2CO),
-                        av.proportion_3CO = mean(proportion_3CO),
-
+                        av.proportion_0CO = mean(proportion_0CO, na.rm = TRUE),
+                        av.proportion_1CO = mean(proportion_1CO, na.rm = TRUE),
+                        av.proportion_2CO = mean(proportion_2CO, na.rm = TRUE),
+                        av.proportion_3CO = mean(proportion_3CO, na.rm = TRUE),
                         #1CO pos
-                        mean.1CO.pos = mean(mean_1CO.pos),
+                        mean.1CO.pos = mean(mean_1CO.pos, na.rm = TRUE),
 
                         #mean.IFD
-                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS),
-                        mean.IFD.PER = mean(mean_IFD.2CO_PER)
-                        )
-
-#add
+                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS, na.rm = TRUE),
+                        mean.IFD.PER = mean(mean_IFD.2CO_PER, na.rm = TRUE)
+                        )
+
 strain.BivData.Table <- add_subsp(strain.BivData.Table)
 strain.BivData.Table <- add_species(strain.BivData.Table)
+#not sure the source of the NAs
 ```
@@
 
-#showing the table..
-#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
-```
-
-RENAME the curate bivdata below
-
-```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}
-
-#this is redundant
-
-male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
-                        Nmice = length(unique(mouse)),
-                        Ncells  = length(Obj.ID),
-
-                      # mean_CO = mean()
-
-                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
-                        cV = cv(chromosomeLength),
-                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
-                        sd   = round(sd(chromosomeLength), 3),
-                        se   = round(sd / sqrt(Ncells), 3),
-
-                       #calq mean.sc by chrm class
-                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
-                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
-                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
-                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
-
-                # IFD things
-                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
-                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),
-
-                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
-                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
-
-                       #CO position things, requires the melt?
-              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),
-
-              #sis.co.ten!!
-              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
-
-              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
-              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
-              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),
-
-              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
-              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
-)
-
-#vari
-male.av_4MM <- add_sex(male.av_4MM)
-male.av_4MM <- add_strain(male.av_4MM)
-male.av_4MM <- add_subsp(male.av_4MM)
-male.av_4MM <- add_category(male.av_4MM)
-
-#unordered for the mixed models (move this to the MM blocks...?)
-#keep order for the plots?
-male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
-male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
-male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
-
-#REMOVE JUVI mice
-juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
-#mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]
-
 ```
@@
 # MLH1 models
+
+```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
+
+#make sure that MOLF female is included!
+
+#DF.HetC.MixedModel <- add_species(DF.HetC.MixedModel)
+DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]
+
+DF.HetC.MixedModel.3 <- Mouse.Table_HQ %>% filter(species == "M.musculus")
+#DF.HetC.MixedModel
+
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel
+
+#main df with variables coded as chr, set false for the model
+DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
+DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
+DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
+DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)
+
+#remove male only data
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+
+DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
+
+```
+
+127 mice, 8 strains (correct), 0 MOLF females!
+
+comes from Mouse.Table_HQ, this has 0 females,
 
@@
 
+there is another set of of this DF at line 200, this DF isn't defined in the setup script
+
+DF.HetC.MixedModel.HQ should have only the relevant strains, but adding in
+
 ```{r MLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
 
-MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
+MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data= DF.HetC.MixedModel.HQ ) 
 MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
 MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-#summary(MLH1.MM.M1)$coefficients
 
@@
 MLH1.MM_strain)
-
-```
+#I get the better values: when there is MOLF female and the old mice are removed
+#consider removing these from the start
+MLH1.MMlmer_results
+```
+
+**these should be the M1 MLH1 results**
+sub.sp      sex.results      interaction.pval     rand.pval
+1 0.0011     4.9e-06          0.00021         0
 
@@
 
-Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
+cleanCurated_BivData <- cleanCurated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")
 
@@
 
-Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
-Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)
-
-chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= ( Curated_BivData %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") ) )
+cleanCurated_BivData$subsp <- factor(cleanCurated_BivData$subsp , ordered = F)
+cleanCurated_BivData$strain <- factor(cleanCurated_BivData$strain , ordered = F)
+
+chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= ( cleanCurated_BivData %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") ) )
 
@@
 ```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
-#REAL.short_mouse_table
-#REAL.long_mouse_table
+#REAL.short_mouse_table -- Short.biv_mouse_table
+#REAL.long_mouse_table -- Long.biv_mouse_table
 #mouse.av.long.biv.table
 
-Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+Q1.av.long.biv_scatr_box   <- ggplot(Long.biv_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nLong Biv")
 
-Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
+Q1.av.short.biv_scatr_box   <- ggplot(Short.biv_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av.
 \nNormalized 1CO Foci pos \nShort Biv")

### 2020-06-11 src/SetupData.Rmd
tracked down difference due to M1 p values

@@
 MLH1_data = read.xlsx2("~./MLH1repo/data/AnonData_formated.xlsx", header=TRUE, 1)
+
+#remove all X's from MLH1
+#add date imaged
 
@@
 
-#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
-#juvi.mice_DF <- juvi.mice_DF[!(is.na(juvi.mice_DF$mouse) | juvi.mice_DF$mouse==""), ]
-
 #juvi mice ~ 2 weeks old
@@
 
-#change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
-#
-
 Mouse.Table_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
@@
 
-#remove juvi mice via the list
-#remove bad ages
-#this is removing female MOLF?
-
-#this removes everything but males in these categories
-#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]
-
-#i don't understand why so many WSB female are removed..
-# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)
+#there are ~26 old mice
+
+Mouse.Table_HQ.OG <- Mouse.Table_HQ
+
+#this doesn't remove
+Mouse.Table_HQ <- Mouse.Table_HQ.OG[ ! Mouse.Table_HQ.OG$mouse %in% old.mice.list, ]
+
+#MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice.list, ]
 
@@
 
-Mouse.Table_HQ.OG <- Mouse.Table_HQ
-
-Mouse.Table_HQ <- Mouse.Table_HQ.OG
+#Mouse.Table_HQ.OG <- Mouse.Table_HQ
 
 #during merging -- try to just merge a few cols: 
-Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
-#this removes ~28 mice
+#don't merge
+#Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
 
@@
 old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]
-
-Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]
 
@@
 
-```{r Age.Hists, echo=FALSE, warning=FALSE}
+```{r Age.Hists, echo=FALSE, warning=FALSE, eval=FALSE}
+#age. weeks was removed from the Mouse.Table.HQ for now
+
 #age.weeks col is missing

### 2020-06-11 doc/MS_v3.Rmd
Set the citation formate for elife, single author

--- 2020-06-11 doc/MS_v2.Rmd
@@
 bibliography: refs_Main_Report.bib
+csl: elife.csl
 ---
@@
 
-We used a panel of wild-derived inbred strains of house mice ( _Mus musculus_ ) and related murids to profile natural genetic variation in recombination (Table 1).
-Our survey included 5 strains from _M.
+<new>
+We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 1).
+Our survey included 5 strains from *M.
 m.
-musculus_, 4 strains from _M.
+musculus*, 4 strains from *M.
 m.
-domesticus_, 2 strains from _M.
+domesticus*, 2 strains from *M.
 m.
-molossinus_, 2 strains from _M.
+molossinus*, 2 strains from *M.
 m.
-castaneus_, and 1 strain each from _M.
-caroli_, _M.
-spicilegus_, and _M.
-spretus_.
-We subsequently denote strains by their abbreviated subspecies and name (e.g.
-_domesticus^WSB^_).
-Mice were housed at dedicated, temperature controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice from an inbred strain of Gough Island mice were sampled after XX generations of brother-sister mating.
+castaneus*, and 1 strain each from *M.
+spicilegus* and *M.
+spretus*.
+<removed _M.
+caroli_> We subsequently denote strains by their abbreviated subspecies and name (e.g.
+domesticusWSB).
+Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
+Mice from an inbred strain of Gough Island mice were sampled after **XX** <trck down the generation of Gough inbred levels> generations of brother-sister mating.
 All mice were provided with ad libitum food and water.
-All procedures followed protocols approved by IACUC.
+Procedures followed protocols approved by IACUC.
 
@@
 
+<old>
 The same dry-down spread technique was applied to both spermatocytes and oocytes based on (@peters_1997), with adjustment for volumes.
@@
 Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37* C.
+Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37* C.
+Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
+
+<new>
+The same dry-down spread technique was applied to both spermatocytes and oocytes, following (@peters_1997), with adjustment for volumes.
+Spermatocyte spreads were collected and prepared as described in Peterson et al.
+(2019).
+The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
+Juvenile mice between 12 and 15 days of age were used for DMC1 counts.
+Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
+Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
+Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
+Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
+Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
+The following morning, slides were treated with a Photoflow wash (Kodak, cite).
+Slides were stored at -20*C if not stained immediately.
+To visualize the structure of meiotic chromosomes, we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
+Crossovers (COs) were visualized as MLH1 foci.
+Double strand breaks (DSBs) were visualized as DMC1 foci.
+The staining protocol followed (Anderson et al.
+1999) and (Koehler et al.
+2002).
+Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
+Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37*C.
+The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37*C.
 Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.

### 2020-06-11 doc/MS_Figures_v2.rmd
more edits, learned greek letters

@@
 
+## Normal Distributions
+
+```{r normal.curve, echo=FALSE, include=FALSE, eval=FALSE}
+
+#this block is throwing an error..not sure why
+#normality of MLH1 distributions
+
+#there are a bunch of strains that should be removed
+#remove grey background
+
+#subsp x sex, color by strain
+
+gaus.curve.remove.strains <- c("F1")
+
+gaus.MLH1 <- MLH1_data
+#gaus.MLH1 <- gaus.MLH1[!gaus.MLH1$strain %in% gaus.curve.remove.strains, ]
+
+gaus.MLH1$strain <-  factor(gaus.MLH1$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW","PERC",
+      "PWD","SKIVE","KAZ", "CZECH", "AST", "TOM",
+      "MSM", "MOLF",
+      "CAST", "HMI",
+      "SPIC","SPRET","CAROLI"))
+
+#combine outgroups
+#change subsp for spic, caroli and spret to outgroup
+gaus.MLH1$subsp2[gaus.MLH1$subsp == "Caroli"] <- "Outgroup"
+gaus.MLH1$subsp2[gaus.MLH1$subsp == "Spretus"] <- "Outgroup"
+gaus.MLH1$subsp2[gaus.MLH1$subsp == "Spic"] <- "Outgroup"
+
+gaus.MLH1$subsp2[gaus.MLH1$species == "M.musculus"] <- "M. m. musculus"
+#make seperate curves / plots for outgroups and house mouse
+
+#MLH1_data
+gaussian.curve.out <- ggplot(data = (gaus.MLH1 %>% filter(quality < 5)  %>% filter(subsp2 == "Outgroup") ),
+                         aes(x=nMLH1.foci, fill=strain) ) + geom_histogram( binwidth = 1)+facet_wrap( ~ sex)  +  scale_fill_manual(values=colors_of_strains)+ 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+gaussian.curve.Mm <- ggplot(data = (gaus.MLH1 %>% filter(quality < 5)  %>% filter(subsp == "Cast") ),
+                         aes(x=nMLH1.foci, fill=strain) ) + geom_histogram( binwidth = 1)+facet_wrap(subsp ~ sex) +  scale_fill_manual(values=colors_of_strains)+ 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black"))
+
+#aes( alpha = .7)
+
+```
+
 ## Fig Chrm Proportion plot

### 2020-06-11 doc/MS_v3.Rmd
more edits, learned greek letters

@@
 
-<old>
-The same dry-down spread technique was applied to both spermatocytes and oocytes based on (@peters_1997), with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used MLH1 quantification were between 5 and 12 weeks of age.
-Juvenile mice between 12 and 15 days of age were used for DMC1 quantification.
-Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
-Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
-Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
-Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
-Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a Photoflow wash (Kodak, cite).
-Slides were stored at -20 * C if not stained immediately.
-To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
-Crossovers (COs) were visualized as MLH1 foci.
-Double strand breaks (DSBs) were visualized as DMC1 foci.
-The staining protocol followed [@anderson1999] and [@koehler2002].
-Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
-Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37* C.
-The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37* C.
-Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37* C.
-Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
-
 <new>
 The same dry-down spread technique was applied to both spermatocytes and oocytes, following (@peters_1997), with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in Peterson et al.
-(2019).
+Spermatocyte spreads were collected and prepared as described in [@peterson2019].
 The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
@@
 The following morning, slides were treated with a Photoflow wash (Kodak, cite).
-Slides were stored at -20*C if not stained immediately.
+Slides were stored at -20\*C if not stained immediately.
 To visualize the structure of meiotic chromosomes, we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
@@
 Double strand breaks (DSBs) were visualized as DMC1 foci.
-The staining protocol followed (Anderson et al.
-1999) and (Koehler et al.
-2002).
+The staining protocol followed [@anderson1999] and [@koehler2002].
 Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
-Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37*C.
+Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\*C.
 The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37*C.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\*C.
 Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37* C.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
 Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
@@
 
-Images were capture using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
+<new>
+Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 or DMC1 foci using Photoshop.
+Image file names were anonymized before manual scoring of MLH1 foci or DMC1 foci using Photoshop.
 
@@
 
-```{r normal.curve, echo=FALSE, include=FALSE, eval=FALSE}
-
-#this block is throwing an error..not sure why
-#normality of MLH1 distributions
-
-#there are a bunch of strains that should be removed
-#remove grey background
-
-#subsp x sex, color by strain
-
-gaus.curve.remove.strains <- c("F1")
-
-gaus.MLH1 <- MLH1_data
-#gaus.MLH1 <- gaus.MLH1[!gaus.MLH1$strain %in% gaus.curve.remove.strains, ]
-
-gaus.MLH1$strain <-  factor(gaus.MLH1$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW","PERC",
-      "PWD","SKIVE","KAZ", "CZECH", "AST", "TOM",
-      "MSM", "MOLF",
-      "CAST", "HMI",
-      "SPIC","SPRET","CAROLI"))
-
-#combine outgroups
-#change subsp for spic, caroli and spret to outgroup
-gaus.MLH1$subsp2[gaus.MLH1$subsp == "Caroli"] <- "Outgroup"
-gaus.MLH1$subsp2[gaus.MLH1$subsp == "Spretus"] <- "Outgroup"
-gaus.MLH1$subsp2[gaus.MLH1$subsp == "Spic"] <- "Outgroup"
-
-gaus.MLH1$subsp2[gaus.MLH1$species == "M.musculus"] <- "M. m. musculus"
-#make seperate curves / plots for outgroups and house mouse
-
-#MLH1_data
-gaussian.curve.out <- ggplot(data = (gaus.MLH1 %>% filter(quality < 5)  %>% filter(subsp2 == "Outgroup") ),
-                         aes(x=nMLH1.foci, fill=strain) ) + geom_histogram( binwidth = 1)+facet_wrap( ~ sex)  +  scale_fill_manual(values=colors_of_strains)+ 
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-gaussian.curve.Mm <- ggplot(data = (gaus.MLH1 %>% filter(quality < 5)  %>% filter(subsp == "Cast") ),
-                         aes(x=nMLH1.foci, fill=strain) ) + geom_histogram( binwidth = 1)+facet_wrap(subsp ~ sex) +  scale_fill_manual(values=colors_of_strains)+ 
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black"))
-
-#aes( alpha = .7)
-
-```
+<new 1/n>
 
@@
 Cells with a score of 5 were excluded from the final analysis.
-Distributions of MLH1 count per cell were visually inspected for normality (Supplemental figure 1).
+Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
 MLH1 foci located on the XY in spermatocytes were excluded from counts.
 In addition to MLH1 counts, we measured several traits to further characterize the recombination landscape.
-To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from a subset of strains ( _musculus^PWD^_, _musculus^MSM^_, _domesticus^WSB^_, and _domesticus^G^_ ).
-SC morphology and CREST foci number was used to stage spermatocytes as early zygotene or late zygotene.
+To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from each of a subset of strains (*molossinus^MSM^*, _musculus^PWD^_,  _domesticus^WSB^_, and _domesticus^G^_ ).
+SC morphology and CREST foci number were used to stage spermatocytes as early zygotene or late zygotene.
 
+<new 2/n>
 To measure bivalent SC length, two image analysis algorithms were used.
@@
 Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
-Total SC length per cell was quantified from pachytene cell images [@wang2019_SC].
+Total SC length per cell was quantified from pachytene cell images (@wang2019_SC).
+
 To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
 Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-
-<!-- what should be included or removed from these large numbers?? Q5, only house mouse?-->
-
-SC length of individual bivalents was quantified in pachytene cell images [@peterson2019].
-The DNA CrossOver algorithm [@peterson2019] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
+SC length of individual bivalents was quantified in pachytene cell images (@peterson2019).
+The DNA CrossOver algorithm (@peterson2019) isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
 The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
-In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from
-`r min(algorithm.stat.category$mean_pass.rate)` ( _molossinus^MSM^_ male) to `r max(algorithm.stat.category$mean_pass.rate)` ( _musculus^KAZ^_ female).
+In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` (*molossinus^MSM^* male) to `r max(algorithm.stat.category$mean_pass.rate)` (*musculus^KAZ^* female).
 From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` bivalent objects were isolated by the algorithm.
@@
 The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-The curated single bivalent data supplements our cell wide MLH1 count data with MLH1 foci counts for single bivalents.
-The proportions of bivalents with the same number of MLH1 foci where compared across strains by a chi-squared test using prop.test().
+The curated single bivalent data supplements our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
+Proportions of bivalents with the same number of MLH1 foci were compared across strains using a chi-square test.
 
+<new 3/n>
 To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-This “short bivalent” data set included the four or five shortest bivalents, and excluded the -- X bivalent in oocytes.
+This “short bivalent” data set included the four or five shortest bivalents and excluded the X bivalent in oocytes.
 A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
@@
 
+<new 4/n>
 To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
@@
 
+<new 5/n>
 We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
-We first constructed a mixed model (M1) using the lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
+We used mouse average as the dependent variable in all analyses.
+We first constructed a linear mixed model (M1) using lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
 In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT() from RLRsim [@RLRsim]).
-Subspecies, strain, and their interaction were coded as fixed effects,  with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
-As observations were at the level of single mouse (mouse average) was uniquely coded within the dataset, nesting was implicit, each strain only occurs within one strain, and was not explicitly coded in our linear and mixed models.
-
+Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
+As each observation was at the level of a single mouse (mouse average), it was uniquely coded within the dataset, nesting was implicit, each strain only occurs within one strain, and was not explicitly coded in our linear and mixed models.
 We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
-In separate analyses, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
-Four additional models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from the initial mixed model.
+In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFDraw, IFDnorm, and average MLH1 position (for single-focus bivalents).
+Four additional lnear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
 
@@
 
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice ( _M.
+<new 1/n>
+
+We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice (*M.
 m.
-domesticus_, _M.
+domesticus*, _M.
 m.
@@
 m.
-molossinus_ ) and two additional species of Mus ( _M.
-spretus_ and _M.
-spicilegus_ ).
-Mean MLH1 foci counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
+molossinus_ ) and three additional species of Mus (*M.
+spretus*, _M.
+spicilegus_ , and *M.
+caroli*).
+Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
+
+<new 2/n>
 
@@
 First, MLH1 focus counts differ between females and males in most strains.
-Second, the relationship between MLH1 foci in the sexes varies among strains.
-Although most strains show more MLH1 foci in females, two strains - _musculus^PWD^_ and _molossinus^MSM^_ - exhibit higher MLH1 focus counts in males.
-While _musculus^SKIVE^_ has a very low degree of sex differences in MLH1 focus counts.
+Second, the difference in counts between the sexes varies among strains.
+Although most strains show more MLH1 foci in females, two strains (*musculus^PWD^* and _molossinus^MSM^_) exhibit higher counts in males.
 In females, numbers of MLH1 foci are evenly distributed around the sex-wide mean of approximately 25 (Figure 1B).
-In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23 MLH1 foci per cell) of foci (Figure 1C).
+In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23) of foci (Figure 1C).
+Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = 0.08; p = 0.84) across the set of strains.
+**track down the correlation value**
 
+<new 3/n>
+
+To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (Table 2; detailed results available in Supplemental Table X).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < 10-4; sex: p = 6.2x10-4; subspecies: p = 1.1x10-3; subspecies\*sex: p = 2x10-4).
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; p = 0.09), highlighting strain and sex as salient variables.
+Two strains exhibit particularly strong effects on MLH1 focus count (M3; domesticusG p = ; domesticusLEW p =), with sex-strain interactions involving three strains (M3; domesticusG p = ; molossinusMSM p = ; musculusPWD p = ).
+In separate analyses of males (M4; n = 80), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = ; effect = ), *molossinus^MSM^* (p = ; effect = ), and *musculus^SKIVE^* (p = ; effect = ).
+These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
+In females (M4; n= 76), four strains affect MLH1 focus count: domesticusG (p = ; effect = ), molossinusMSM (p = ; effect = ), domesticusLEW (p = ; effect = ), and musculusPWD (p = ; effect = ).
+Strain effect sizes in females are modest in magnitude compared to those in males.
+
+-----
+
+<old 3/n>
 To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` mice.

### 2020-06-11 src/Result_Setup.Rmd
more edits, learned greek letters

@@
 
+FOR new supplemental table write out all of the model coeffiecnts (into raw -- then reformat), use these to double check the text values.
+
 ```{r MLH1co.lmer.MM1, echo=FALSE, waring=FALSE, message=FALSE, eval=TRUE}
@@
 
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]
-
 #strain = fixed
@@
 #MLH1 variance models
+
+**THIS IS BEING MOVED TO CHAPTER 3**
 

### 2020-06-11 src/SetupData.Rmd
more edits, learned greek letters

@@
 MLH1_data = read.xlsx2("~./MLH1repo/data/AnonData_formated.xlsx", header=TRUE, 1)
-
-#remove all X's from MLH1
+#remove all X's from MLH1, remove all old mice?
 #add date imaged
+#remove the Remove mice, (juvi, old, bad quality)
 

### 2020-06-11 doc/MS_v3.Rmd
save often

@@
 To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (Table 2; detailed results available in Supplemental Table X).
-Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < 10-4; sex: p = 6.2x10-4; subspecies: p = 1.1x10-3; subspecies\*sex: p = 2x10-4).
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; p = 0.09), highlighting strain and sex as salient variables.
-Two strains exhibit particularly strong effects on MLH1 focus count (M3; domesticusG p = ; domesticusLEW p =), with sex-strain interactions involving three strains (M3; domesticusG p = ; molossinusMSM p = ; musculusPWD p = ).
-In separate analyses of males (M4; n = 80), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = ; effect = ), *molossinus^MSM^* (p = ; effect = ), and *musculus^SKIVE^* (p = ; effect = ).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < 10-4 -- `r 1/10000`; sex: p = 6.2x10-4 -- `r MLH1.MMlmer_results$sex.results`; subspecies: p = 1.1x10-3 -- `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = 2x10-4 -- `r MLH1.MMlmer_results$interaction.pval`).
+
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects **(M2; p = 0.09)**, highlighting strain and sex as salient variables.
+Two strains exhibit particularly strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
+
+In separate analyses of males (M4; n = 80 - `r length(male_mouse.avs_for.analysis$mouse)`), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` ), *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`), and *musculus^SKIVE^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` ).
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-In females (M4; n= 76), four strains affect MLH1 focus count: domesticusG (p = ; effect = ), molossinusMSM (p = ; effect = ), domesticusLEW (p = ; effect = ), and musculusPWD (p = ; effect = ).
+In females (M4; n= 76), four strains affect MLH1 focus count: domesticusG (p = ; effect = ), molossinusMSM (p = ; effect = ), domesticusLEW (p = ; effect = ), and *musculus^PWD^* (p = ; effect = ).
 Strain effect sizes in females are modest in magnitude compared to those in males.
@@
 -----
-
 <old 3/n>
-To further situate variation in house mouse recombination rate within an evolutionary framework, we fit a series of statistical models including subspecies, strain, and sex, to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` mice.
-To investigate the effects of sex, we limited our analysis to `r length( unique( DF.HetC.MixedModel.HQ$strain) )` strains with sufficient observations from both sexes (Table 1).
-
-We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < `r 1/10000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p= `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count.
 After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
+A model with only these two variables
+(M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
 In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
@@
 We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4).
+
 In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).

### 2020-06-11 doc/MS_v3.Rmd
removing old files, Model results

@@
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-In females (M4; n= 76), four strains affect MLH1 focus count: domesticusG (p = ; effect = ), molossinusMSM (p = ; effect = ), domesticusLEW (p = ; effect = ), and *musculus^PWD^* (p = ; effect = ).
+
+In females (M4; n= 76 - `r length(female_mouse.avs_for.analysis$mouse)`), four strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]`), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]`), *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]`), and *musculus^PWD^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[4]`).
 Strain effect sizes in females are modest in magnitude compared to those in males.
 
------
-<old 3/n>
-After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex.
-A model with only these two variables
-(M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ).
-In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).
-
-We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4).
-
-In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).
-
-These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`), _molossinus^MSM^_  (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ), and _musculus^PWD^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]` ).
-Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.
 Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.
@@
 
-Our survey of recombination rate across house mice provides an opportunity to determine whether sex differences in chromatin compaction (SC length) are reversed when heterochiasmy is reversed.
-In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (all t-test below; p < 0.05; except short bivalents for SKIVE p = `r Q1_short.SC_t.test.SKIVE$p.value` ).
-In the reduced dataset of short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from  `r Q1SC.mus.av.short.biv_DIFF_PWD` ( _musculus^PWD^_) to `r Q1SC.mus.av.short.biv_DIFF_WSB` ( _domesticus^WSB^_) across strains.
+<new 1/n>
+The variation in sex differences in recombination we discovered provided an opportunity to determine whether sex differences in chromatin compaction, as measured by the length of the synaptonemal complex (SC), are reversed when heterochiasmy is reversed.
+In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = 0.11 --  `r Q1_short.SC_t.test.SKIVE$p.value``).
 
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and short bivalent measures (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
+Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.26 - `r Q1SC.mus.av.short.biv_DIFF_PWD`  (*musculus^PWD^*) to 1.49 - `r Q1SC.mus.av.short.biv_DIFF_WSB` (*domesticus^WSB^*) across strains.
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = 7.16^{-20} -- `r Q1.lmer_totSC_M1_results$sex.results` ; M2: p = 5.33^{-4} -- `r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =0.05 -- `r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]` ) and short bivalent length (M1: p = 1.27^{-11 -- `r Q1.lmer_short_M1_results$sex.results`}; M2: p = 1.98^{-7} -- `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = 1.98^{-7} -- `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
+
+The existence of some subspecies and strain effects (M1 subspecies\*sex: p = 0.03 -- `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-*musculus^SKIVE^* : p=0.02 -- `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]` ; _musculus^SKIVE^_\*male: p = 0.08 -- `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.
+
+**double check that these are the only sig factors for all the models**
+
+--------
+
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and
+short bivalent measures (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
+
 The existence of some subspecies and strain effects (M1 interaction: p = `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]`,  SKIVE * male p = `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.

### 2020-06-12 src/Result_Setup.Rmd
removing more outliers for total SC

@@
 MLH1.MMlmer_results
+
+write.table(MLH1.MMlmer_results, "~./MLH1repo/MLH1_M1_glm.csv", sep=",",row.names = TRUE)
 ```
@@
 #strain = fixed
-MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )
+MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
 
@@
 
-#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
+write.table(summ.MLH1.M2.glm.strain.fixed.full$coefficients , "~./MLH1repo/MLH1_M2_glm.csv", sep=",",row.names = TRUE)
 ```
@@
 
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
-
-MLH1.M3.strain.fixed <- glm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])
+MLH1.M3.strain.fixed <- glm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ )
 
@@
 
-#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
+write.table(summ.MLH1.M3.strain.fixed$coefficients , "~./MLH1repo/MLH1_M3_glm.csv", sep=",",row.names = TRUE)
 ```
@@
 #round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)
+
+write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/MLH1_M4_female_glm.csv", sep=",",row.names = TRUE)
 
@@
 
-#round(lm.female.strain_sum$coefficients[,c(1,4)],5)
-#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
+write.table(sum_MLH1.female.spc.glm.M5$coefficients, "~./MLH1repo/MLH1_M5_female_glm.csv", sep=",",row.names = TRUE)
 ```
@@
 
-#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)
+write.table(sum_MLH1.male.glm.M4$coefficients , "~./MLH1repo/MLH1_M4_male_glm.csv", sep=",",row.names = TRUE)
 
@@
 
-#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)
+write.table(sum_MLH1.male.glm.M4.2$coefficients , "~./MLH1repo/MLH1_M5_male_glm.csv", sep=",",row.names = TRUE)
 
@@
 tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+tot.SC_old.mice <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% old.mice.list, ]
 
 totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]
+
+totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% old.mice.list, ]
+
+tot.SC_old.mice.DF <- Total.SC_DF[Total.SC_DF$mouse  %in% old.mice.list, ]
 
@@
 
-## total sc models
+## Total sc models
 
@@
 
+write.table(Q1.lmer_totSC_M1_results, "~./MLH1repo/TotalSC_Q1_M1.csv", sep=",",row.names = TRUE)
+
 ```
@@
 #G * sex
+
+write.table(sum.Q1.glm_totSC_M2$coefficients, "~./MLH1repo/TotalSC_Q1_M2.csv", sep=",",row.names = TRUE)
+
 ```
@@
 #LOT OF SIGNIFICant strain effects in this model
-
-```
-
-```{r Q1_single.Biv_M1.lmer, echo=FALSE, include=FALSE, eval=FALSE}
-
-#1CO  lmer.Q1_SCLength_M1
-Q1.lmer_SCLength_1CO_M1 <- lmer(mean.SC_1CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_1CO_M1_R.sex <- lmer(mean.SC_1CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-Q1.lmer_SCLength_1CO_M1_R.subsp <- lmer(mean.SC_1CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_1CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.1CO_sex <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_subsp <- anova(Q1.lmer_SCLength_1CO_M1, Q1.lmer_SCLength_1CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.1CO_int <- drop1(Q1.lmer_SCLength_1CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.1CO_strain <- exactRLRT(Q1.lmer_SCLength_1CO_M1)$p.value
-
-Q1.lmer_SCLength.1CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.1CO_subsp, 
-sex.results =  Q1.lmer_SCLength.1CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.1CO_int,  rand.pval = 
-Q1.lmer_SCLength.1CO_strain)
-
-#2CO
-Q1.lmer_SCLength_2CO_M1 <- lmer(mean.SC_2CO ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM)
-Q1.lmer_SCLength_2CO_M1_R.sex <- lmer(mean.SC_2CO ~ subsp + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE) 
-Q1.lmer_SCLength_2CO_M1_R.subsp <- lmer(mean.SC_2CO ~ sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)
-
-#as.data.frame(round(summary(Q1.lmer_SCLength_2CO_M1)$coefficients, 6))
-
-Q1.lmer_SCLength.2CO_sex <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_subsp <- anova(Q1.lmer_SCLength_2CO_M1, Q1.lmer_SCLength_2CO_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1.lmer_SCLength.2CO_int <- drop1(Q1.lmer_SCLength_2CO_M1, test="Chisq")$`Pr(Chi)`[2]
-Q1.lmer_SCLength.2CO_strain <- exactRLRT(Q1.lmer_SCLength_2CO_M1)$p.value
-
-Q1.lmer_SCLength.2CO_results <- data.frame(sub.sp = Q1.lmer_SCLength.2CO_subsp, 
-sex.results =  Q1.lmer_SCLength.2CO_sex, 
-               interaction.pval = Q1.lmer_SCLength.2CO_int,  rand.pval = 
-Q1.lmer_SCLength.2CO_strain)
-#as.data.frame(round(summary(SCLength_long.biv_MM.M1)$coefficients, 6))
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
+write.table(sum.Q1.glm_totSC_M3$coefficients, "~./MLH1repo/TotalSC_Q1_M3.csv", sep=",",row.names = TRUE)
+```
+
+```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
 
 lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE) 
-
 lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE) 
-
 lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
@@
 #Q1lmer.MM_SC_short_results
-
+write.table(Q1lmer.MM_SC_short_results, "~./MLH1repo/ShortSC_Q1_M1.csv", sep=",",row.names = TRUE)
 ```
@@
 
+write.table(Q1.lmer_short_M1_results, "~./MLH1repo/ShortSC_Q1_M1.csv", sep=",",row.names = TRUE)
+
 ```
@@
 #skive, slightly sig
+
+write.table(summ.Q1.glm_short.SC_M2$coefficients, "~./MLH1repo/ShortSC_Q1_M2.csv", sep=",",row.names = TRUE)
 ```
@@
 #SKIVE * male
+
+write.table(summ.Q1.glm_short.SC_M3$coefficients, "~./MLH1repo/ShortSC_Q1_M3.csv", sep=",",row.names = TRUE)
 ```
@@
 
-```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
-#switch the cleanCurated_BivData
-
-#nested boxplots, pooled Bivalents
-Q1.SC.length_Nested_boxplots_Dom <- ggplot(cleanCurated_BivData[cleanCurated_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
-
-```
-
 need to save this reduced dataset, Q1.real.short_mouse_table.
@@
 #lmer.Q1.nrm1CO_M1_results
-
-#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)
+write.table(lmer.Q1.nrm1CO_M1_results, "~./MLH1repo/Nrm1CO_Pos_Q1_M1.csv", sep=",",row.names = TRUE)
 
@@
 
-#summary(Q1.glm_F1pos_M2)
-
 summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
@@
 
-#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)
+write.table(summry.Q1.glm_F1pos_M2$coefficients, "~./MLH1repo/Nrm1CO_Pos_Q1_M2.csv", sep=",",row.names = TRUE)
 
@@
 #write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
+
+write.table(summry.Q1.glm_1CO.pos_M3$coefficients, "~./MLH1repo/Nrm1CO_Pos_Q1_M3.csv", sep=",",row.names = TRUE)
 
@@
 
+write.table(Q1.IFD.ABS_M1_lmer.results, "~./MLH1repo/ABS.IFD_Q1_M1.csv", sep=",",row.names = TRUE)
+
 #M1_IFD_2CO
@@
 
+write.table(Q1.IFD.PER_M1_lmer.results, "~./MLH1repo/PER.IFD_Q1_M1.csv", sep=",",row.names = TRUE)
+
 ```
@@
 #Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-
 glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") )  )
@@
 #write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)
+write.table(summry.glm.Q1_IFD.ABS.M2$coefficients, "~./MLH1repo/ABS.IFD_Q1_M2.csv", sep=",",row.names = TRUE)
 
@@
 
+write.table(summry.glm.Q1_IFD.PER.M2$coefficients, "~./MLH1repo/PER.IFD_Q1_M2.csv", sep=",",row.names = TRUE)
 ```
@@
 summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
+
+write.table(summry.glm.Q1_IFD.ABS.M3$coefficients, "~./MLH1repo/ABS.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
 
@@
 
-#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)
+write.table(summry.glm.Q1_IFD.PER.M3$coefficients, "~./MLH1repo/PER.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
+
 ```
@@
 summry.glm.Q1_IFD.ABS.M4.female  <- summary(glm.Q1_IFD.ABS.M4.female)
+#write.table(summry.glm.Q1_IFD.PER.M3, "~./MLH1repo/PER.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
 
@@
 summry.glm.Q1_IFD.PER.M4.female  <- summary(glm.Q1_IFD.PER.M4.female)
+#write.table(summry.glm.Q1_IFD.PER.M3, "~./MLH1repo/PER.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
 
@@
 summry.glm.Q1_IFD.ABS.M5.female  <- summary(glm.Q1_IFD.ABS.M4.female)
+#write.table(summry.glm.Q1_IFD.PER.M3, "~./MLH1repo/PER.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
 
@@
 summry.glm.Q1_IFD.PER.M5.female  <- summary(glm.Q1_IFD.PER.M5.female)
+#write.table(summry.glm.Q1_IFD.PER.M3, "~./MLH1repo/PER.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
 
@@
 
-```
-
-```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
-
-#these are the scatter plots
-
-nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
-#looked up image, measures confirmed
-
-####
-nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-
-nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
-#21may18_KAZ_m2 -- seems like it might be an outlier
-#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
-#second foci slightly smaller, kinda fainter -- but also much brighter than background
-
-#check all KAZ males
-
-nrm.IFD_KAZ_BIV.data <- ggplot( (cleanCurated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_KAZ.female_BIV.data <- ggplot( (cleanCurated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_KAZ.female_BIV.data
-
-nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_PWD.female_BIV.data
-
-nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_LEW.female_BIV.data
-
-nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_G.female_BIV.data
-
-nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
-
-raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
-
-Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
-
-###look at why KAZ males are weird
-nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
-
-nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-#the difference the the Nrm IFD cut off is striking with these plots
-nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
-
-nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
-
-```
-
-```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-
-#first draft of the IFD comparisons
-#make the mouse boxplots -- instead of geom_points
-
-#Only use 2CO bivalents
-Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
-Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
-
-P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.2)+
-  ggtitle("Normalized IFD")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
-
-#each point is a 2CO, with the F1 and F2 position 
-#based on plot from Dourad et al (Mercier)
-
-P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
-    aes(y = Foci2, x= Foci1, color=strain))+
-  geom_jitter(alpha=.5)+
-  ggtitle("Raw IFD WSB")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#trying to move x axis to top
-#cowplot might be the easier way -- since kinda diffecult in ggplot
-#or ggdraw
-
-P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
-    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
-  geom_jitter(alpha=.5)+
-#  scale_x_continuous(position = "top") +
-  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
-    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank() )
-
-#boxplots
-boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, ABS")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
-#order is out of wack, but -- good for now
-
-boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
-  geom_jitter(alpha=.2)+
-  ggtitle("Interfocal Distances, PER")+
-  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
-
-first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
-
-first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
-
-#+theme(legend.position="none")
 ```
@@
 
-```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
-
-#REAL.long_mouse_table
-#REAL.short_mouse_table
-
-#real.long.bivData
-#real.short.bivData
-
-shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
-\nMale Long Biv SC Lengths")
-
-#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
-\nMale Long Biv SC Lengths")
-
-```
-
-```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
-
-mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
-  facet_wrap(Rec.group~subsp, scales="free")+
-  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
-
-mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
-
-```
-
 ```{r Q2.totSC.ttest, echo=FALSE}
@@
 #Dom-Musc
-Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Musc"))$mean.skel)
-#p-value = 0.5
+Q2.totSC_t.test_DomMusc <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean.skel,
+                          (mouse.total.skel  %>%  filter(subsp == "Musc") %>% filter(sex == "male") )$mean.skel)
+# = p-value = 8e-06
 
 #Dom-Mol
-Q2.totSC_t.test_DomMol <- t.test((mouse.total.skel  %>% filter(subsp == "Dom") )$mean.skel,
-                          (mouse.total.skel  %>%  filter(subsp == "Mol"))$mean.skel)
-#p-value = 0.1
+Q2.totSC_t.test_DomMol <- t.test( (mouse.total.skel  %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean.skel,
+    (mouse.total.skel  %>%  filter(subsp == "Mol") %>% filter(sex == "male") )$mean.skel )
+#p-value = p-value = 2e-05
+
+#Musc-Mol
+Q2.totSC_t.test_MuscMol <- t.test( (mouse.total.skel  %>% filter(subsp == "Musc") %>% filter(sex == "male") )$mean.skel,
+    (mouse.total.skel  %>%  filter(subsp == "Mol") %>% filter(sex == "male") )$mean.skel )
+#p-value = 0.02
+
 ```
@@
 
-```{r Q2.totSC.glm, echo=FALSE}
+```{r Q2.totSC.glm, echo=FALSE, eval=FALSE}
+#this model shouln't be used -- switch to M4 and M5 males
 
@@
 
+```
+
+```{r totalSC.M4M5, echo=FALSE}
+
+#Q2.totl.SC <- (mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast") )
+
+#M4
+Q2.total.SC.male.glm.M4 <- glm(mean.skel ~ subsp * strain, data= (mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast") ) )
+
+sum_Q2.total.SC.male.glm.M4 <- summary(Q2.total.SC.male.glm.M4)
+
+#write.table(sum_MLH1.male.glm.M4$coefficients , "~./MLH1repo/MLH1_M4_male_glm.csv", sep=",",row.names = TRUE)
+
+#M5
+Q2.total.SC.male.glm.M5 <- glm( mean.skel ~ strain, data= (mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast") ) )
+
+sum_Q2.total.SC.male.glm.M5 <- summary(Q2.total.SC.male.glm.M5)
+
+#write.table(sum_MLH1.male.glm.M4.2$coefficients , "~./MLH1repo/MLH1_M5_male_glm.csv", sep=",",row.names = TRUE)
+
+#write.table(MLH1.MMlmer_results, "~./MLH1repo/MLH1_M1_glm.csv", sep=",",row.names = TRUE)
 ```

### 2020-06-12 src/SetupData.Rmd
removing more outliers for total SC

@@
 
+Total.SC_DF <- Total.SC_DF[Total.SC_DF$subsp != "other", ]
+
 # factor -> chr
@@
 
+#last outliers
+#Total.SC_DF <- Total.SC_DF[Total.SC_DF$skel_size < 10000, ]#one very high female measure
+#Total.SC_DF <- Total.SC_DF[Total.SC_DF$skel_size < 1900, ]
+
+Total.SC_DF.OG <- Total.SC_DF
+
+#Total.SC_DF
+tot.SC_old.mice.DF <- Total.SC_DF[Total.SC_DF$mouse  %in% old.mice.list, ]#581 lines from old mice
+tot.SC_juvi.mice.DF <- Total.SC_DF[ Total.SC_DF$mouse  %in% juvi_mice.list, ]#118 lines from juvi mice
+
+Total.SC_DF <- Total.SC_DF[! Total.SC_DF$mouse  %in% old.mice.list, ] 
+Total.SC_DF <- Total.SC_DF[! Total.SC_DF$mouse  %in% juvi_mice.list, ]
+
+```
+
+```{r total.SC.outliers, echo=FALSE, eval=FALSE}
+
+#yy <- Total.SC_DF %>% filter(subsp == "Dom") %>% filter(sex == "male")
+Total.SC_DF %>% filter(subsp == "Dom") %>% filter(sex == "male")
+
+show.total.SC <- ggplot(data= Total.SC_DF %>% filter(strain == "SKIVE"), aes(y = skel_size, x=sex, color=strain))+
+  scale_color_manual(values=colors_of_strains)+
+  geom_jitter() +ylim(c(950,3500 ))+ facet_wrap(~mouse, scales="free")
+
+ # geom_boxplot(alpha=.2)+   facet_wrap(~REDO.crop, scales="free", ncol = 3) +  #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
+
+  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),
+   # axis.title.x=element_blank(),
+#axis.text.x = element_text(size=7), 
+#legend.position="none",axis.text.y = element_text(size=9), # axis.text.y = element_blank()
+#axis.title.y = element_text(size = 10) ) +
+
+ # labs(y = "Total SC Length")+ylim(c(1000,5000 ))+ggtitle("")
+
 ```
@@
 ```{r Mouse.total.sc.table, echo=FALSE}
-
-#Total.SC_DF
 
@@
 
-#look at date of imaging affects on total SC
-#count=1
-#for(i in MLH1_data$Original.Name){
-   # print(i)
-#    templist= strsplit(i, split="_")[[1]]
-#    c = (templist[1])
-#    MLH1_data$image.string[count] <-  c
-    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
-#    count= count +1
-#}
-#works better if dat is converted out of image
-#MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')
-
 ```

### 2020-06-13 doc/MS_v3.Rmd
first draft of abstract

@@
   html_document:
-    toc: yes
-    toc_depth: '5'
-    toc_float: yes
+  word_document:
   pdf_document:
-    toc: yes
-    toc_depth: '5'
-  word_document:
+    includes:
+      in_header: header.tex
 always_allow_html: yes
@@
 
+# ABSTRACT
+
+Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
+Differences in recombination rate between females and males are perhaps the most striking form of this variation.
+Existing data fail to directly address the extent to which recombination experiences similar evolutionary pressures.
+To fill this gap, we measured meiotic recombination in both sexes for a panel of house mouse cross three subspecies.
+Using inbred strains and single cell immunohistochemistry allowed us to place sex-specific observations within the same genetic background and meiotic context.
+Our results indicate highly discordant evolutionary patterns in the two sexes.
+Whereas male recombination rates show evidence of rapid evolution over short evolutionary timescales, female recombination rates measured in the same strains are mostly static.
+These results strongly indicate that house mouse has two genome wide recombination rates which display distinct evolutionary trajectories.
+
 # INTRODUCTION
+
+### 0.   Importance of recombination rate
+
+### 1. Understanding levels of variation in gwRR
+
+### 2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale
+
+### 3. What we accomplished in this paper
 
@@
 caroli_> We subsequently denote strains by their abbreviated subspecies and name (e.g.
-domesticusWSB).
+*domesticus^WSB^*).
 Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice from an inbred strain of Gough Island mice were sampled after **XX** <trck down the generation of Gough inbred levels> generations of brother-sister mating.
+Mice from an inbred strain of Gough Island mice were sampled after **XX** <track down the generation of Gough inbred levels> generations of brother-sister mating.
 All mice were provided with ad libitum food and water.
@@
 <new>
-The same dry-down spread technique was applied to both spermatocytes and oocytes, following (@peters_1997), with adjustment for volumes.
+The same dry-down spread technique was applied to both spermatocytes and oocytes, following [@peters_1997], with adjustment for volumes.
 Spermatocyte spreads were collected and prepared as described in [@peterson2019].
@@
 We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
-In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFDraw, IFDnorm, and average MLH1 position (for single-focus bivalents).
+In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
 Four additional lnear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
@@
                          "SPRET","SPIC" ), order=T )
+
 ```
@@
 
+```{r corMLH1.FM, echo=FALSE}
+
+#needs to be strain
+sperman.corr.FM <- cor( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
+                        method = c("spearman") )
+
+sperman.corr.FM.pval <- cor.test( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
+                        method = c("spearman") )
+
+```
+
 ## Genome-wide recombination rate evolves differently in females and males
@@
 In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23) of foci (Figure 1C).
-Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = 0.08; p = 0.84) across the set of strains.
-**track down the correlation value**
+Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = 0.08 - `r sperman.corr.FM`; p = 0.84 -- `r sperman.corr.FM.pval$p.value`) across the set of strains.
 
@@
 
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects **(M2; p = 0.09)**, highlighting strain and sex as salient variables.
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects **(M2; p = 0.09 -- )**, highlighting strain and sex as salient variables.
 Two strains exhibit particularly strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
@@
 
+In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males recombine more.
+This pattern implies that in high-recombination strains, spermatocytes have less space than oocytes in which to position additional crossovers.
+
 **double check that these are the only sig factors for all the models**
@@
 
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC  (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]`) and
-short bivalent measures (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
-
-The existence of some subspecies and strain effects (M1 interaction: p = `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-SKIVE `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]`,  SKIVE * male p = `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.
-In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males have more MLH1 foci per cell.
-This pattern further suggests that in high-recombination strains, spermatocytes have less space in which to position additional crossovers than do oocytes.
-
 ## Females and males differ in crossover positions and crossover interference
 
+<new 1/n>
 We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value` ).
-Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results` ; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68--`r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: 0.56 --  `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; t-test; p = 3.84^{-23} -- `r Q1_1CO.pos_t.test.FULL$p.value`).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = 7.79^{-26 --  `r lmer.Q1.nrm1CO_M1_results$sex.results`}; M2: p = 5.39^{-8} --  `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 5.39^{-8} --  `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 
-Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (all t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (t-test; p = `r Q1.IFD.PER_t.test_KAZ$p.value` ).
-Examination of IFD~norm~ distributions indicates that female values are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Analysis of models treating IFD~norm~ as the dependent variable supports the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`, M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+<new 2/n>
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = 0.33 -- `r Q1.IFD.PER_t.test_KAZ$p.value`).
+Examination of IFD~norm~ distributions indicates that females are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Models treating IFDnorm as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = 3.11^{-12} -- `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01  --  `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = 0.01  --  `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) across the full set of strains, whether they are considered separately or together.
+Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males, and that males display a stronger telomeric bias in the placement of the distal crossover.
 
-In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) across the full set of strains, whether they are considered separately or together.
-Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males and that males display a stronger telomeric bias in the placement of the distal crossover.
-In summary, controlling for differences in SC length (chromatin compaction) using IFD~norm~ indicates that interference is stronger in males, whereas consideration of IFD~raw~ shows that the sexes exhibit a similar level of interference on the physical scale.
+In summary, controlling for differences in SC length (chromatin compaction) indicates that interference is stronger in males, whereas interference on the physical scale is similar in the sexes.
+
+--------
 
@@
 
+<new 1/n>
 We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-We considered proportions of bivalents with different numbers of crossovers, double-strand breaks, SC length, and crossover positions, including crossover interference.
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")`) have either one or two MLH focus (Figure 3).
+We considered proportions of bivalents with different numbers of crossovers, double-strand break number, SC length, and crossover positioning.
+
+<new 2/n>
+Ninety-six percent of single bivalents in our pooled dataset (n = 9,576 -- `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` ) have either one or two MLH foci (Figure 3).
 The proportions of single-focus (1CO) bivalents vs.
 double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Figure 3).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_ , and 0.51 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
+Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
+*molossinus^MSM^*: p = 0.37 -- `r PWD.MSM.2CO.prop.test$p.value` ; _musculus^PWD^_ vs.
+_musculus^KAZ^_: p = 1.23^{-31} -- `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_: p = 2.65^{-6} -- `r MSM.MOLF.2CO.prop.test$p.value` ).
+These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number across multiple bivalents.
 
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
-Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-_musculus^KAZ^_: p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value`) and _musculus^PWD^_ vs.
-_molossinus^MSM^_: p = `r PWD.MSM.2CO.prop.test$p.value`.
-These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number on multiple bivalents.
-
+<new 3/n>
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )`  early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from three low-recombination strains ( _musculus^KAZ^_, _domesticus^WSB^_ , and  _domesticus^G^_) and two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_).
-High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test;  p < `r 1/1000000` ).
-In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval` ).
+DMC1 foci were counted in a total of 76 -- `r sum( (DMC1table.main$ncells.x) )` early zygotene and 75 -- `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains (*musculus^PWD^* and *molossinus^MSM^*) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < 10^{-6} -- `r 1/1000000` ).
+In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = 0.66 -- `r ttest.HighLow.Z.pval` ).
 Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
 High-recombination and low-recombination strains do not differ in the MLH1/DMC1 ratio, whether DMC1 foci were counted in early zygotene cells or late zygotene cells (t-test; p > 0.05).
-These results raise the possibility that the evolution of genome-wide recombination rate is primarily determined by processes that precede the CO/NCO decision.
+These results raise the possibility that the evolution of genome-wide recombination rate is primarily determined by processes that precede the CO/NCO decision, at least in house mice.
 
-Although there is a positive correlation between total SC length and MLH1 foci when both traits are measured as mouse means (Spearman’s r = `r cor( mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method = c("spearman") )`; p = `r cor.test(mouse.total.skel$mean.skel, mouse.total.skel$mean.MLH1, method=c("spearman") )$p.value` ), total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
-Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within _M.
+<new 4/n>
+Total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value` ), separate tests within subspecies show that the two strain categories differ within *M.
 m.
-molossinus_ (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` ) but not within _M.
+molossinus* (p = 0.02 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value` ) but not within *M.
 m.
-musculus_ (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+musculus* (p = 0.4 -- `r Q2tot.sc_t.test_High.Low_Musc$p.value` ).
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value`  ; long: p = 0.19 --  `r Q2long.biv_t.test_High.Low$p.value` ).
 
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value` ; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
-In a model with total SC length as the dependent variable, two subspecies effects are significant ( _M.
+Write results for M4 and M5 here
+total SC M4 results
+
+Musc subspecies very ***, Mol *** (both ++)
+G, LEW, 
+MSM
+SKIVE, KAZ (- -)
+
+M5 (almost all strains are significant)
+G, LEW, PWD, MSM
+MOLF, 
+SKIVE, KAZ, TOM, AST, CZECH
+
+In a model with total SC length as the dependent variable (*M2?*), two subspecies effects are significant (*M.
 m.
-musculus_ p = `r sum.Q2.glm_mus.av_totSC_M4[2,4]` , _M.
+musculus* p = 2.33^{-6} -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`, *M.
 m.
-molossinus_, p = `r sum.Q2.glm_mus.av_totSC_M4[3,][4]` ).
+molossinus*, p = 21^{-6} -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
 In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05), but they are not consistent across models.
-Collectively, these observations reveal that evolution of the genome-wide recombination rate is not strongly associated with evolution of SC length.
+Collectively, these results reveal that evolution of SC length is not strongly associated with evolution of genome-wide recombination rate in house mice.
 
-High-recombination and low-recombination strains do not differ in MLH1 focus positions on bivalents with a single focus, either for the full dataset (t-test; p = `r Q2.1COpos_Ttest$p.value` ) or within subspecies (t-test; _M.
-m.
-musculus_: p = `r Q2.1COpos_mus.av_Musc_ttest$p.value` ; _M.
-m.
-molossinus_ : p = `r Q2.1COpos_mus.av_Mol_ttest$p.value` ).
-While _domesticus^WSB^_ and _molossinus^MOLF^_ exhibit strain effects in a model with normalized crossover position as the dependent variable, these two strains do not differ in MLH1 count.
-These results show that evolution of crossover position on chromosomes with single crossovers is decoupled from evolution of the genome-wide recombination rate, at least at this scale of resolution.
+<new 5/n>
+In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not to SC length and crossover position (on single-crossover bivalents).
 
-High-recombination strains have greater inter-focal distances (in both raw and normalized measured) than low-recombination strains (t-test; IFD~norm~: p = `r t.test_mouse.av.IFD_PER$p.value` ; IFD~raw~: p = `r t.test_mouse.av.IFD_ABS$p.value`).
-This pattern holds within _M.
-m.
-musculus_ (t-test; IFD~norm~: p = `r  t.test_mouse.av.IFD_PER_MUSC$p.value`; IFD~raw~: p = `r  t.test_mouse.av.IFD_ABS_MUSC$p.value`), with a trend in the same direction within _M.
-m.
-molossinus_ (IFD~norm~: p= `r t.test_mouse.av.IFD_PER_MOL$p.value` ; IFD~raw~: p = `r t.test_mouse.av.IFD_ABS_MOL$p.value`).
+--------------------------
 
-Similar results are observed with models treating IFD~raw~ and IFD~norm~ as dependent variables: only effects associated with high-recombination strains are significant (p < 0.05).
-That IFD~raw~ and IFD~norm~ show similar patterns eliminates variation in SC lengths (and bivalent sizes) as the primary explanation.
-The main distinction in IFD~norm~ distributions between high-recombination and low-recombination strains is an enrichment of IFD~norm~ values under 30% in low-recombination strains (Supplemental Figure 3).
-The frequency of IFD~norm~ values that fall below 30% ranges from 8.2% ( _domesticus^G^_) to 16% ( _musculus^KAZ^_) in low-recombination strains, whereas high-recombination strains all show frequencies below 4% (0%, 1.3%, and 3.3% for _musculus^SKIVE^_, _molossinus^MSM^_, and _musculus^PWD^_, respectively).
-These results indicate that evolution of the genome-wide recombination rate is accompanied by the evolution of crossover interference.
+# DISCUSSION
 
-In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not (consistently) to SC length and crossover position (on single-crossover bivalents).
-
-# B[Reserve for chapter 3?] Within-individual variation in the genome-wide recombination rate is higher in females
-
-*not sure where to insert this*
-
-Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
-While this trait will be examined in greater detail in future manuscript, we not that the general pattern is that females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1 and Table 3).
-
-<!-- Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
-While this trait will be examined in greater detail in future manuscript, we not that the general pattern is that females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1 and Table 2).
-
-round(sum.MLH1var_M2_Q12$coefficients[,4][4], 5)
-**MLH1var**  M1 - LmerMLH1_CV_results, M2 MLH1var_M2_sum_coE, M3  M4 glmMLH1Var_M3_sum.Coef
-
-Counting MLH1 foci in multiple oocytes for each female and in multiple spermatocytes for each male allowed us to examine determinants of variation in recombination rate within mice.
-To do this, we considered the same models as above, but replaced mean MLH1 focus count (for a mouse) with within-mouse variance in MLH1 focus count (for a mouse) as the dependent variable.
-Sex is the only variable that affects MLH1 focus count in all models (M1: p < 10^{-6} --`r LmerMLH1_CV_results$sex.results` ; M2: p = 0.03 -- `r` ; M3: p = 0.03 -- ``).
-
-In general, females have almost twice as much inter-cellular variance in MLH1 foci compared to males (Figure 1).
-Since estimates of within-mouse variance may be more susceptible to technical error from staining, we repeated the analyses using a subset of cells with high quality scores (1 or 2; Materials and Methods).
-The results are similar: sex is the strongest effect (M1 p < 10^{-6}; M2 p = 2.310^{-4} -- ; M3 p = 2.2810^{-4} -- ).
-When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 focus count in either sex.
-These results show that sex is a primary determinant of inter-cellular variance in the genome-wide recombination rate across diverse genetic backgrounds.
-
- Q12 MLH1.var -- M1 -LmerMLH1_M1_Q12_VAR_results -- LmerMLH1_M1_Q12_VAR_results,  M2 -sum.MLH1var_M2_Q12.co_eff,  M3 -MLH1var_M3_Q12_sum -->
-
-# Discussion
-
-```{r corMLH1.FM, echo=FALSE}
-
-#needs to be strain
-sperman.corr.FM <- cor( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
-                        method = c("spearman") )
-
-sperman.corr.FM.pval <- cor.test( ( strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(species == "M.musculus" ) )$mean_co, ( strain.table.HQ_Fig %>% filter(sex == "male") %>% filter(species == "M.musculus" ) )$mean_co,
-                        method = c("spearman") )
-```
-
+<new 1/n>
 By comparing recombination rates in females and males from the same diverse set of genetic backgrounds, we isolated sex as a primary factor in the evolution of this fundamental meiotic trait.
@@
 Because inter-strain divergence times are identical for the two sexes, this observation demonstrates that the genome-wide recombination rate evolves faster in males, at least in house mice.
-More generally, recombination rate divergence is decoupled in males and females (Spearman’s correlation test comparing female and male strain mean MLH1 values for house mice; rho = `r sperman.corr.FM.pval$estimate`; p = `r sperman.corr.FM.pval$p.value` ).
-These disparities are remarkable given that recombination rates for the two sexes were measured in the same genomic backgrounds.
+More generally, recombination rate divergence is decoupled in females and males.
+These disparities are remarkable given that recombination rates for the two sexes were measured in identical genomic backgrounds (other than the presence/absence of the Y chromosome).
 Our results provide the strongest evidence yet that the genome-wide recombination rate follows distinct evolutionary trajectories in males and females.
-Next, we consider several explanations for this phenomenon.
+
+<new 2/n>
 At the genetic level, the sex-specific evolution we documented indicates that some mutations responsible for divergence in recombination rate have dissimilar phenotypic effects in females and males.
-A subset of the genetic variants associated with genome-wide recombination rate within populations of humans (@Kong2004, @Kong2008, @Kong2014, @halldorsson2019), Soay sheep [@johnston2016_soay], and cattle (@ma2015_cattle, @Shen2018_cattle) appear to show sex-specific properties, including antagonistic effects in females and males.
-Inter-sexual correlations are weak for recombination rate in humans [@fledel2011] and Soay sheep (@johnston2016_soay).
-Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice (@dumont2011, @Wang2017island).
-These variants could differently affect females and males at any step in the recombination pathway.
-Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons) and the significance of SC length distinguishing high and low recombining strains, our findings suggest that mutations that determine the chromatin structure and number of double-strand breaks contributed to sex-specific evolution in the recombination rate [@baier2014].
+A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle; @Shen2018_cattle] appear to show sex-specific properties, including opposite effects in females and males.
+Furthermore, inter-sexual correlations for recombination rate are weak in humans [@fledel2011] and Soay sheep [@johnston2016_soay].
+Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011; @Wang2017island; @wang2019_SC].
+These variants could differentially affect females and males at any step in the recombination pathway.
+Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contribute to sex-specific evolution in the recombination rate.
+A study of two classical inbred strains and one wild-derived inbred strain of house mice also found a positive association between crossover number and double-strand break number in males [@baier2014].
 
-Our results also raise the prospect that the connection between recombination rate and fitness differs between males and females.
-Little is known about whether and how natural selection shapes recombination in nature (@DapperPayseur2017, @Ritz2017).
-Using a quantitative genetic test, [@samuk2020]  recently discovered that genome-wide recombination rates in females from two populations of _Drosophila pseudoobscura_ diverged in a manner consistent with local adaptation.
+<new 3/n>
+Another implication of our results is that the connection between recombination rate and fitness differs between males and females.
+Little is known about whether and how natural selection shapes recombination rate in nature [@DapperPayseur2017; @Ritz2017].
+@samuk2020 recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of *Drosophila pseudoobscura* was caused by natural selection.
 Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
-Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon (@brandvain2012scrambling, @sardell_sex_2020).
+
+<new 4/n>
+Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling; @sardell_sex_2020].
 Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020].
@@
 
-We propose an alternative interpretation based on the cell biology of gametogenesis.
-During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister cohesin distal to chiasmata (@LaneKauppi2019, @vanVeen2003).
-The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesin holding homologs together [@LaneKauppi2019].
-Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and chromosome separation , thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
-
-Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019], perhaps due to the presence of the centrosome spindle [@So2019] and higher cell volume [@kyogoku2017] in oocytes.
+<new 5/n>
+We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@LaneKauppi2019; @vanVeen2003].
+The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesion holding homologs together [@LaneKauppi2019].
+Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and chromosome separation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019], perhaps due to the presence of the centrosome spindle [@So2019] and larger cell volume [@kyogoku2017] in oocytes.
 The higher stringency of the SAC during spermatogenesis suggests that selection will be better at removing mutations that interfere with bi-orientation in males than in females.
@@
 
+<new 5/n>
 Our SAC model is consistent with other features of our data.
 We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
-The number and placement of crossovers affects the area of sister chromosome cohesion distal to crossovers which needs to be released for the first reductional chromosome segregation (@vanVeen2003, @LaneKauppi2019, @subramanian2014, @dumontDesai2012).
-Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell (Kudo ).
-While in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those with dictyate arrest [@Lee2019].
+The number and placement of crossovers affects the area of sister chromosome cohesion distal to crossovers which needs to be released for the first reductional chromosome segregation [@vanVeen2003; @LaneKauppi2019; @subramanian2014; @dumontDesai2012].
+Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell (**Kudo** ), whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those organisms that undergo dictyate arrest [@Lee2019].
 
+<new 6/n>
 We propose that the SAC model also can explain the correlated evolution of stronger crossover interference and higher genome-wide recombination rate in male house mice.
@@
 Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and may improve the fidelity of chromosomal segregation.
-Although the SAC model postulates direct fitness effects of interference, a modifier model predicts that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
+While the SAC model postulates direct fitness effects of interference, a modifier model predicted that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
 
+<new 7/n>
 Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
-The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle @ma2015_cattle and differences between wild-born and laboratory-raised white-footed mice @peterson2019.
+The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle (@ma2015_cattle) and differences between populations of white-footed mice [@peterson2019].
 In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates (@segura2013, @ottoPaysuer2019).
@@
 
-Our findings further reveal that evolution of the genome-wide recombination rate does not require changes in the degree of chromatin compaction.
-Females consistently showed longer SCs, even in strains with more recombination in males.
-Studies in mice (@lynn2002, @petkov2007) and humans (@gruhn2013, @tease2004) suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
-Some authors have suggested that the conserved sex differences in the crossover position (male telomere bias, female uniform placement) and genome levels of crossover interference strength (male stronger interference, female weaker interference), could be due to the loser vs tighter chromatin packing of the meiotic chromosome structure in females and males respectively (@haenel2018,  @petkov2007).
-A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes (@hulten2011_COM).
-More recent work has connected the sparser recombination landscape has been attributed to sex differences in the crossover repair pathway [@wang2017inefficient].
+<new 8/n>
+Our findings further reveal that evolution of the genome-wide recombination rate does not require major changes in the degree of chromatin compaction.
+Female house mice consistently show longer SCs, even in strains with more recombination in males.
+Studies in mice [@lynn2002, @petkov2007] and humans [@gruhn2013, @tease2004] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+Some authors have suggested that conserved sex differences in crossover positioning (more uniform placement in females) and interference strength (stronger interference in males) could be due to looser chromatin packing of the meiotic chromosome structure in females [@haenel2018; @petkov2007].
+A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM].
+More recent work connects the sparser recombination landscape to sex differences in the crossover repair pathway [@wang2017inefficient].
 
+<new 9/n>
 Our conclusions are accompanied by several caveats.
 First, MLH1 foci only identify interfering crossovers [@holloway2008mus81].
-Although most crossovers (80%; [@@holloway2008mus81]) belong to this class, our approach likely underestimated genome-wide recombination rates.
+Although most crossovers belong to this class **(REF)**, our approach likely underestimated genome-wide recombination rates.
 Evolution of the number of non-interfering crossovers is a topic worth examining.
-A second limitation is that our investigation of crossover positions was confined to the relatively low resolution possible with immunofluorescent cytology.
+A second limitation is that our investigation of crossover locations was confined to the relatively low resolution possible with immunofluorescent cytology.
 Positioning crossovers with higher resolution could reveal additional evolutionary patterns.
@@
 We considered most available wild-derived inbred lines, but house mice have a broad geographic distribution.
-Nevertheless, we expect our conclusion that recombination evolves in a sex-specific manner to be robust to geographic sampling because differences between females and males exist for the same set of inbred strains.
+Nevertheless, we expect our primary conclusion that recombination rate evolves in a sex-specific manner to be robust to geographic sampling because differences between females and males exist for the same set of inbred strains.
 
-The causes of sex differences in recombination remain mysterious @lenormand2016.
-Our conclusions have implications for a wide range of recombination research.
+<new 10/n>
+While the causes of sex differences in recombination remain mysterious @lenormand2016, our conclusions have implications for a wide range of recombination research.
 For biologists uncovering the cellular and molecular determinants of recombination, our results suggest that mechanistic differences between the sexes could vary by genetic background.
@@
 
+--------
+
 # REFERENCES

### 2020-06-14 doc/MS_v3.Rmd
learning trying to convert to latex

@@
 output:
-  html_document:
-  word_document:
   pdf_document:
@@
       in_header: header.tex
+  word_document:
+  html_document:
 always_allow_html: yes
@@
 
+```{r global_options, include=FALSE}
+
+knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'figures/', echo = TRUE, warning = FALSE, message = FALSE)
+
+```
+
 # ABSTRACT
@@
 Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
+**define genome-wide recombination rate**.
+
 Differences in recombination rate between females and males are perhaps the most striking form of this variation.
@@
 
+Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
+The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
+The total number of COs per cell (at the 4n stage) is equal to the genome wide recombination rate (gwRR).
+**The genome wide recombination rate is a cell based metric for genome wide recombination rate is tightly connected to organisms fitness through correct chromosome segregation and fertility.**
+
+**at the fine scale / in regards to genetic variation**, In the indriect manner, the recombination rate regulates populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (cite).
+This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation (Begun Aquadro, Nachman, Payseur).
+
+**but back to the genome-wide recombination rate**, The process of chromosome segregation is integral to the process of producing haploid cells and can be considered a constraint in regards to the evolution.
+When the genome wide recombination rate is framed within the single cell context it is closely connected to the  contraints which are imposed at the single cell level.
+Reflecting the importance of it's role in chromosome segreation is that the  number of haploid pairs is strong predictor for genome wide RR across large species(taxa) / divergence scales [@ottoPaysuer2019; @stapley_variation_2017].
+(the number of crossovers per chromosome are highly conserved)<However, whie the rate of recombination is highly variable, the count of crossovers per chromosome is conserved between 1 to 3 (cite), suggesting that such thresholds/constraints may be behind the overall patterns.>
+
+**connection of single cell based gwRR to thresholds imposed at the single cell level** (why is gwRR the better scale for evolutionary patterns?)
+
 ### 1. Understanding levels of variation in gwRR
+
+Because we know the meiotic programs, we have expectations for where the thresholds(constraints) of the gwRR 
+the rate will fall.
+First, the lower threshold is defined as an obligate crossover per homolog pair which promotes proper chromosome segregation [@nagaoka2012].
+second (while there is less evidence,), the upper threshold maybe regulated/shaped/affected due to rates of ectopic recombination or the programmed DNA damage preceding crossovers act as constraints to the upper threshold of meiotic recombination (cite).
+Yet, Within these thresholds the genome wide recombination rate still varies across species and indiviuals.
+which is the subject of an open question in the field.  **the gwRR is the best metric for since both of these constraints are thought to act at the cell wide level.**
+
+**Sex is one of the most notable axes along which individuals vary**.
+Sexual dimorphism in recombination rates is called heterochiasmy [@lenormand2003].
+An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
+Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog,  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep) sheep [@johnston2016_soay], and mouse (CC cite) in which the role of sex is confounded with the contributions of genetic variation.
+Under anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates.
+Yet heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
+Although it is clear that the relationship between female and male recombination rates can differ among species, comparisons between and within closely related species are missing.
+Direct contrasts between females and males across a common, diverse set of genomic backgrounds would reveal whether the recombination rate evolves differently in the sexes.
+
+<It’s important to note, – that the genome wide recombination rate can be decomposed – and connected to related traits (which also show conserved patterns of sexual dimorphism>
+
+**understanding heterochiasmy and sexual dimorphism in meiosis have advanced since first being documented (morgan 1914) -- more than just the direction -- more evidence is accumulating that there are conserved patterns of sexual dimorphism in meiotic recombinatio - related to recombination rate**.
+Recently two specific features have been highlighted as conserved patterns of heterochiasmy:
+
+@CahoonLibuda2019 show In species with meiotic chromatin (SC axis length) quantified for both sexes, longer chromosome axis generally has higher recombination rates. 
+@sardell_sex_2020, A survey of 51 species shows there are conserved broad scale recombination landscape differences.
+Generally males have telomere-bias crossover placement and females have more uniform placement.
+
+These observations raise the question of conservation in sexually patterns are maintained with the gwRR and it’s decomposed traits
+
+**remember these points kinda distract from gwRR aspect**
 
@@
 
+The House mouse complex arose from a recent radiation providing an opportunity to interrogate natural variation at short evolutionary scales, potentially narrowing the mutational space for identifying variants which would cause gwRR variation.
+Add divergence ranges.
+divergence spanning 0.5 mya to 5 mya.
+
+Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway (cite) and previous crosses for understanding the genetic architecture of recombination rate variation @dumont2011house; @murdoch2010; @Wang2017island].
+
+**Wild derived inbred strains generate the best comparison of females and males, besides the sex chromosomes, the mouse for each genome are almost identical**.
+Inbred strains enable one of the most direct comparison of both female and male --  versions of meiosis --enable use to conclude that sex is a primary factor --
+
+Additionally due to their global distribution allows us to compare samples of natural diversity from a broad geographic range.
+(wild derived inbred strains – will capture genetic diversity not represented in classical lab strains.)
+
+While there are strains with Robertsonian translocations, in house mouse and related murid species it is possible to assemble sets of strains with identical karyotypes, 20 pairs of acrocentric chromosomes.
+
+House mouse is well suited for single cell cytology approaches.
+The single cell level allows integration of data at a closer connection to the molecular pathway and meiotic program.
+In addition to the quantification of gwRR.
+This study quantifies precursors to crossovers, double strand breaks (DSBs), and meiotic chromosome morphology across stages of meiosis which would not be accessible to with genetic linkage data and crosses.
+
 ### 3. What we accomplished in this paper
+
+  For the first time in many of these strains, We report a rare, direct, evolutionary comparison of recombination rate in females and males for this short evolutionary scale.
+We use rare strains from a broad range of geographic locations of the species territory.
+
+The gwRR is made up of crossovers occurring on individual chromosomes within cells.
+Thus we quantified meiotic chromosome morphology (SC length) and placement of crossovers to comprise an approximate picture of the recombination landscape.
+
+Our results indicate rapid male specific evolution of gwRR occurred in a subset of house mouse lines.
+We observed up to a 30% difference (translating into approximately ~7 more crossovers per cell), a surprising amount considering the short evolutionary time scale.
 

### 2020-06-14 doc/MS_v3.tex
learning trying to convert to latex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Sex-Specific Evolution of the Meiotic Recombination Rate},
  pdfauthor={April L. Peterson, Bret Payseur},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{titling}

\pretitle{%
  \begin{center}
  \LARGE
  \includegraphics[width=4cm,height=6cm]{logo.png}\\[\bigskipamount]
}
\posttitle{\end{center}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Sex-Specific Evolution of the Meiotic Recombination Rate}
\author{April L. Peterson, Bret Payseur}
\date{2020-06-14}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{ABSTRACT}\label{abstract}}

Although meiotic recombination is required for successful gametogenesis
in most species that reproduce sexually, the rate of crossing over
varies among individuals. \textbf{define genome-wide recombination
rate}.

Differences in recombination rate between females and males are perhaps
the most striking form of this variation. Existing data fail to directly
address the extent to which recombination experiences similar
evolutionary pressures. To fill this gap, we measured meiotic
recombination in both sexes for a panel of house mouse cross three
subspecies. Using inbred strains and single cell immunohistochemistry
allowed us to place sex-specific observations within the same genetic
background and meiotic context. Our results indicate highly discordant
evolutionary patterns in the two sexes. Whereas male recombination rates
show evidence of rapid evolution over short evolutionary timescales,
female recombination rates measured in the same strains are mostly
static. These results strongly indicate that house mouse has two genome
wide recombination rates which display distinct evolutionary
trajectories.

\hypertarget{introduction}{%
\section{INTRODUCTION}\label{introduction}}

\hypertarget{importance-of-recombination-rate}{%
\subsubsection{0. Importance of recombination
rate}\label{importance-of-recombination-rate}}

Meiosis can be reduced to the expression of (2n -\textgreater{} 4n
-\textgreater{} 2n -\textgreater{} 1n) which tracks the duplication of a
diploid genome into haploid cell products. The meiotic program relies on
crossovers and the process of recombination to ensure the correct
separation of chromosomes. The total number of COs per cell (at the 4n
stage) is equal to the genome wide recombination rate (gwRR).
\textbf{The genome wide recombination rate is a cell based metric for
genome wide recombination rate is tightly connected to organisms fitness
through correct chromosome segregation and fertility.}

\textbf{at the fine scale / in regards to genetic variation}, In the
indriect manner, the recombination rate regulates populations' response
to selection, and determine the fate of novel mutations by transferring
beneficial mutations onto novel genetic backgrounds or by breaking
linkage of negative mutations from beneficial genetic backgrounds
(cite). This process shapes the genomic patterns of genetic variation,
with high recombination areas of genome having more nucleotide variation
while areas of low recombination have decreased genetic variation (Begun
Aquadro, Nachman, Payseur).

\textbf{but back to the genome-wide recombination rate}, The process of
chromosome segregation is integral to the process of producing haploid
cells and can be considered a constraint in regards to the evolution.
When the genome wide recombination rate is framed within the single cell
context it is closely connected to the contraints which are imposed at
the single cell level. Reflecting the importance of it's role in
chromosome segreation is that the number of haploid pairs is strong
predictor for genome wide RR across large species(taxa) / divergence
scales (Otto and Payseur, 2019; Stapley et al., 2017). (the number of
crossovers per chromosome are highly conserved)\textless However, whie
the rate of recombination is highly variable, the count of crossovers
per chromosome is conserved between 1 to 3 (cite), suggesting that such
thresholds/constraints may be behind the overall patterns.\textgreater{}

\textbf{connection of single cell based gwRR to thresholds imposed at
the single cell level} (why is gwRR the better scale for evolutionary
patterns?)

\hypertarget{understanding-levels-of-variation-in-gwrr}{%
\subsubsection{1. Understanding levels of variation in
gwRR}\label{understanding-levels-of-variation-in-gwrr}}

Because we know the meiotic programs, we have expectations for where the
thresholds(constraints) of the gwRR the rate will fall. First, the lower
threshold is defined as an obligate crossover per homolog pair which
promotes proper chromosome segregation (Nagaoka et al., 2012). second
(while there is less evidence,), the upper threshold maybe
regulated/shaped/affected due to rates of ectopic recombination or the
programmed DNA damage preceding crossovers act as constraints to the
upper threshold of meiotic recombination (cite). Yet, Within these
thresholds the genome wide recombination rate still varies across
species and indiviuals. which is the subject of an open question in the
field. \textbf{the gwRR is the best metric for since both of these
constraints are thought to act at the cell wide level.}

\textbf{Sex is one of the most notable axes along which individuals
vary}. Sexual dimorphism in recombination rates is called heterochiasmy
(Lenormand, 2003). An understanding of how sex shapes the evolution of
recombination cannot be achieved with available data. Comprehensive
comparisons of female and male recombination rates usually come from
outbred populations humans (Kong et al., 2004,, 2008, 2014; Halldorsson
et al., 2019), dog, cattle (Ma et al., 2015; Shen et al., 2018), (other
sheep) sheep (Johnston et al., 2016), and mouse (CC cite) in which the
role of sex is confounded with the contributions of genetic variation.
Under anisogamy where a species has distinct gametes, there is no first
principle which would predict the evolution of sexually dimorphic
recombination rates. Yet heterochiasmy, is commonly observed in
dioecious species, suggesting that other meiotic traits which
distinguish the gametes and their meiotic programs, for example
symmetrical vs asymmetrical cell division, may impose selection for
sexually dimorphic recombination rates. Although it is clear that the
relationship between female and male recombination rates can differ
among species, comparisons between and within closely related species
are missing. Direct contrasts between females and males across a common,
diverse set of genomic backgrounds would reveal whether the
recombination rate evolves differently in the sexes.

\textless It's important to note, -- that the genome wide recombination
rate can be decomposed -- and connected to related traits (which also
show conserved patterns of sexual dimorphism\textgreater{}

\textbf{understanding heterochiasmy and sexual dimorphism in meiosis
have advanced since first being documented (morgan 1914) -- more than
just the direction -- more evidence is accumulating that there are
conserved patterns of sexual dimorphism in meiotic recombinatio -
related to recombination rate}. Recently two specific features have been
highlighted as conserved patterns of heterochiasmy:

Cahoon and Libuda (2019) show In species with meiotic chromatin (SC axis
length) quantified for both sexes, longer chromosome axis generally has
higher recombination rates. Sardell and Kirkpatrick (2020), A survey of
51 species shows there are conserved broad scale recombination landscape
differences. Generally males have telomere-bias crossover placement and
females have more uniform placement.

These observations raise the question of conservation in sexually
patterns are maintained with the gwRR and it's decomposed traits

\textbf{remember these points kinda distract from gwRR aspect}

\hypertarget{the-house-mouse-is-a-great-model-for-uncovering-evolutionary-patterns-at-a-short-timescale}{%
\subsubsection{2. The House Mouse is a great model for uncovering
evolutionary patterns at a short
timescale}\label{the-house-mouse-is-a-great-model-for-uncovering-evolutionary-patterns-at-a-short-timescale}}

The House mouse complex arose from a recent radiation providing an
opportunity to interrogate natural variation at short evolutionary
scales, potentially narrowing the mutational space for identifying
variants which would cause gwRR variation. Add divergence ranges.
divergence spanning 0.5 mya to 5 mya.

Classical lab strains of mice have generated extensive knowledge central
to meiosis and outcomes on recombination through extensive studies on
the genes involved in the meiotic recombination pathway (cite) and
previous crosses for understanding the genetic architecture of
recombination rate variation ({\textbf{???}}); Murdoch et al. (2010);
Wang and Payseur (2017){]}.

\textbf{Wild derived inbred strains generate the best comparison of
females and males, besides the sex chromosomes, the mouse for each
genome are almost identical}. Inbred strains enable one of the most
direct comparison of both female and male -- versions of meiosis
--enable use to conclude that sex is a primary factor --

Additionally due to their global distribution allows us to compare
samples of natural diversity from a broad geographic range. (wild
derived inbred strains -- will capture genetic diversity not represented
in classical lab strains.)

While there are strains with Robertsonian translocations, in house mouse
and related murid species it is possible to assemble sets of strains
with identical karyotypes, 20 pairs of acrocentric chromosomes.

House mouse is well suited for single cell cytology approaches. The
single cell level allows integration of data at a closer connection to
the molecular pathway and meiotic program. In addition to the
quantification of gwRR. This study quantifies precursors to crossovers,
double strand breaks (DSBs), and meiotic chromosome morphology across
stages of meiosis which would not be accessible to with genetic linkage
data and crosses.

\hypertarget{what-we-accomplished-in-this-paper}{%
\subsubsection{3. What we accomplished in this
paper}\label{what-we-accomplished-in-this-paper}}

For the first time in many of these strains, We report a rare, direct,
evolutionary comparison of recombination rate in females and males for
this short evolutionary scale. We use rare strains from a broad range of
geographic locations of the species territory.

The gwRR is made up of crossovers occurring on individual chromosomes
within cells. Thus we quantified meiotic chromosome morphology (SC
length) and placement of crossovers to comprise an approximate picture
of the recombination landscape.

Our results indicate rapid male specific evolution of gwRR occurred in a
subset of house mouse lines. We observed up to a 30\% difference
(translating into approximately \textasciitilde7 more crossovers per
cell), a surprising amount considering the short evolutionary time
scale.

\hypertarget{materials-and-methods}{%
\section{MATERIALS AND METHODS}\label{materials-and-methods}}

\hypertarget{mice}{%
\subsubsection{Mice}\label{mice}}

We used a panel of wild-derived inbred strains of house mice (\emph{Mus
musculus}) and related murid species to profile natural genetic
variation in recombination (Table 1). Our survey included 5 strains from
\emph{M. m. musculus}, 4 strains from \emph{M. m. domesticus}, 2 strains
from \emph{M. m. molossinus}, 2 strains from \emph{M. m. castaneus}, and
1 strain each from \emph{M. spicilegus} and \emph{M. spretus}.
\textless removed \emph{M. caroli}\textgreater{} We subsequently denote
strains by their abbreviated subspecies and name
(e.g.~\emph{domesticus\textsuperscript{WSB}}). Mice were housed at
dedicated, temperature-controlled facilities in the UW-Madison School of
Medicine and Public Health, with the exception of mice from Gough
Island, which were housed in a temperature-controlled facility in the
UW-Madison School of Veterinary Medicine. Mice from an inbred strain of
Gough Island mice were sampled after \textbf{XX} generations of
brother-sister mating. All mice were provided with ad libitum food and
water. Procedures followed protocols approved by IACUC.

\hypertarget{tissue-collection-and-immunohistochemistry}{%
\subsubsection{Tissue Collection and
Immunohistochemistry}\label{tissue-collection-and-immunohistochemistry}}

The same dry-down spread technique was applied to both spermatocytes and
oocytes, following (Peters et al., 1997), with adjustment for volumes.
Spermatocyte spreads were collected and prepared as described in
(Peterson et al., 2019). The majority of mice used for MLH1 counts were
between 5 and 12 weeks of age. Juvenile mice between 12 and 15 days of
age were used for DMC1 counts. Both ovaries were collected from embryos
(16-21 embryonic days) or neonates (0-48 hours after birth). Whole
testes were incubated in 3ml of hypotonic solution for 45 minutes.
Decapsulated ovaries were incubated in 300ul of hypotonic solution for
45 minutes. Fifteen microliters of cell slurry (masticated gonads) were
transferred to 80ul of 2\% PFA solution. Cells were fixed in this
solution and dried in a humid chamber at room temperature overnight. The
following morning, slides were treated with a Photoflow wash (Kodak,
cite). Slides were stored at -20*C if not stained immediately. To
visualize the structure of meiotic chromosomes, we used antibody markers
for the centromere (CREST) and lateral element of the synaptonemal
complex (SC) (SYCP3). Crossovers (COs) were visualized as MLH1 foci.
Double strand breaks (DSBs) were visualized as DMC1 foci. The staining
protocol followed (Anderson et al., 1999) and (Koehler et al., 2002).
Antibody staining and slide blocking were performed in 1X antibody
dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X
PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)). Following
a 30-minute blocking wash in ABD, each slide was incubated with 60ul of
a primary antibody master mix for 48 hours at 37*C. The master mix
recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted
1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal
anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to
CREST (Antibodies, Inc; diluted 1:200) suspended in ADB. Slides were
washed twice in 50ml ADB before the first round of secondary antibody
incubation for 12 hours at 37*C. Alexa Fluor 488 donkey anti-rabbit IgG
(Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey
anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended
in ADB. The last incubation of Alexa Fluor 568 donkey anti-goat
(Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37* C.
Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours
after a final wash in 1x PBS.

\hypertarget{image-processing}{%
\subsubsection{Image Processing}\label{image-processing}}

Images were captured using a Zeiss Axioplan 2 microscope with AxioLab
camera and AxioVision software (Zeiss, Cambridge, UK). Preprocessing,
including cropping, noise reduction, and histogram adjustments, was
performed using Photoshop (v13.0). Image file names were anonymized
before manual scoring of MLH1 foci or DMC1 foci using Photoshop.

\hypertarget{analysis}{%
\subsubsection{Analysis}\label{analysis}}

\textless new 1/n\textgreater{}

To estimate the number of crossovers across the genome, we counted MLH1
foci. MLH1 foci were counted in cells with intact and complete
karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20
acrocentric bivalents for oocytes) and distinct MLH1 foci. A quality
score ranging from 1 (best) to 5 (worst) was assigned to each cell based
on visual appearance of staining and spread of bivalents. Cells with a
score of 5 were excluded from the final analysis. Distributions of MLH1
count per cell were visually inspected for normality (Supplemental
Figure 1). MLH1 foci located on the XY in spermatocytes were excluded
from counts. In addition to MLH1 counts, we measured several traits to
further characterize the recombination landscape. To estimate the number
of double-strand breaks, a minority of which lead to crossovers, mean
DMC1 foci per cell was quantified for a single male from each of a
subset of strains (\emph{molossinus\textsuperscript{MSM}},
\emph{musculus\textsuperscript{PWD}},
\emph{domesticus\textsuperscript{WSB}}, and
\emph{domesticus\textsuperscript{G}} ). SC morphology and CREST foci
number were used to stage spermatocytes as early zygotene or late
zygotene.

\textless new 2/n\textgreater{} To measure bivalent SC length, two image
analysis algorithms were used. The first algorithm estimates the total
(summed) SC length across bivalents for individual cells (Wang et al.
(2019)). The second algorithm estimates the SC length of individual
bivalents (Peterson et al. (2019)). Both algorithms apply a
`skeletonizing' transformation to synapsed chromosomes that produces a
single, pixel-wide `trace' of the bivalent shape. Total SC length per
cell was quantified from pachytene cell images (Wang et al. (2019)).

To reduce algorithmic errors in SC isolation, outliers were visually
identified at the mouse level and removed from the data set. Mouse
averages were calculated from cell-wide total SC lengths in 3,204 out of
3,884 cells with MLH1 counts. SC length of individual bivalents was
quantified in pachytene cell images (Peterson et al. (2019)). The DNA
CrossOver algorithm (Peterson et al. (2019)) isolates single,
straightened bivalent shapes, returning SC length, location of MLH1
foci, and location of CREST (centromere) foci. The algorithm
substantially speeds the accurate measurement of bivalents, but it
sometimes interprets overlapping bivalents as single bivalents. In our
data set, average proportions of bivalents per cell isolated by the
algorithm ranged from 0.48 (\emph{molossinus\textsuperscript{MSM}} male)
to 0.72 (\emph{musculus\textsuperscript{KAZ}} female). From the total
set of pachytene cell images, 10,222 bivalent objects were isolated by
the algorithm. Following a manual curation, 9,575 single-bivalent
observations remained. The accuracy of the algorithm is high compared to
hand measures after this curation step (Peterson et al., 2019). The
curated single bivalent data supplements our cell-wide MLH1 count data
with MLH1 foci counts for single bivalents. Proportions of bivalents
with the same number of MLH1 foci were compared across strains using a
chi-square test.

\textless new 3/n\textgreater{} To account for confounding effects of
sex chromosomes from pooled samples of bivalents, we also considered a
reduced data set including only bivalents with SC lengths below the 2nd
quartile in cells with at least 17 of 20 single bivalent measures. This
``short bivalent'' data set included the four or five shortest bivalents
and excluded the X bivalent in oocytes. A total of 704 short bivalents
were isolated from 102 oocytes and 43 spermatocytes. Although this
smaller data set has decreased power, it offers a more comparable set of
single bivalents to compare between the sexes. A ``long bivalent'' data
set was formed from those bivalents above the 4th quartile in SC lengths
per cell. A total of 709 long bivalents were isolated from 102 oocytes
and 43 spermatocytes.

\textless new 4/n\textgreater{} To examine crossover interference, the
distance (in SC units) between MLH1 foci (inter-focal distance;
IFD\textsubscript{raw}) was measured for those single bivalents
containing two MLH1 foci. A normalized measure of interference
(IFD\textsubscript{norm}) was computed by dividing
IFD\textsubscript{raw} by SC length on a per-bivalent basis.

\textless new 5/n\textgreater{} We used a series of statistical models
to interpret patterns of variation in the recombination traits we
measured (Table 2). We used mouse average as the dependent variable in
all analyses. We first constructed a linear mixed model (M1) using
lmer() from the lmer4 package (Bates et al., 2015) in R (v3.5.2) (Team,
2015). In this model, strain was coded as a random effect, with
significance evaluated using a likelihood ratio test (using exactRLRT()
from RLRsim (Scheipl et al., 2008)). Subspecies, sex, and their
interaction were coded as fixed effects, with significance evaluated
using a chi-square test comparing the full and reduced models (drop1()
and anova()) (Bates et al., 2015). As each observation was at the level
of a single mouse (mouse average), it was uniquely coded within the
dataset, nesting was implicit, each strain only occurs within one
strain, and was not explicitly coded in our linear and mixed models. We
used the subspecies effect to quantify divergence between subspecies and
the (random) strain effect to quantify variation within subspecies in a
sex-specific manner. In separate analyses using model M1, we considered
mouse averages as dependent variables for each of the following traits:
MLH1 count per cell, total SC length per cell, single bivalent SC length
per cell, IFD\textsubscript{raw}, IFD\textsubscript{norm}, and average
MLH1 position (for single-focus bivalents). Four additional lnear models
containing only fixed effects (M2-M5) (Table 2) were used to further
investigate results obtained from model M1.

\hypertarget{results}{%
\section{RESULTS}\label{results}}

\hypertarget{genome-wide-recombination-rate-evolves-differently-in-females-and-males}{%
\subsection{Genome-wide recombination rate evolves differently in
females and
males}\label{genome-wide-recombination-rate-evolves-differently-in-females-and-males}}

\textless new 1/n\textgreater{}

We used counts of MLH1 foci per cell to estimate genome-wide
recombination rates in 14 wild-derived inbred strains sampled from three
subspecies of house mice (\emph{M. m. domesticus}, \emph{M. m. musculus}
and \emph{M. m. molossinus} ) and three additional species of Mus
(\emph{M. spretus}, \emph{M. spicilegus} , and \emph{M. caroli}). Mean
MLH1 focus counts for 186 mice were quantified from an average of 21.97
spermatocytes per male (for a total of 1,714 spermatocytes) and 18
oocytes per female (for a total of 1,427 oocytes).

\textless new 2/n\textgreater{}

Graphical comparisons reveal sex-specific dynamics to the evolution of
genome-wide recombination rate (Figure 1A). First, MLH1 focus counts
differ between females and males in most strains. Second, the difference
in counts between the sexes varies among strains. Although most strains
show more MLH1 foci in females, two strains
(\emph{musculus\textsuperscript{PWD}} and
\emph{molossinus\textsuperscript{MSM}}) exhibit higher counts in males.
In females, numbers of MLH1 foci are evenly distributed around the
sex-wide mean of approximately 25 (Figure 1B). In stark contrast, males
largely separate into two groups of strains with high numbers (near 30)
and low numbers (near 23) of foci (Figure 1C). Strain mean MLH1 focus
counts from females and males are uncorrelated (Spearman's \(\rho\) =
0.08 - 0.08; p = 0.84 -- 0.84) across the set of strains.

\textless new 3/n\textgreater{}

To further partition variation in recombination rate, we fit a series of
linear models to mean MLH1 focus counts from 137 house mice (Table 2;
detailed results available in Supplemental Table X). Strain, sex,
subspecies, and sex*subspecies each affect MLH1 focus count in a linear
mixed model (M1; strain (random effect): p \textless{} 10-4 --
\ensuremath{10^{-4}}; sex: p = 6.2x10-4 --
\ensuremath{4.95\times 10^{-6}}; subspecies: p = 1.1x10-3 -- 0;
subspecies*sex: p = 2x10-4 -- \ensuremath{2.1\times 10^{-4}}).

The effect of subspecies is no longer significant in a model treating
all factors as fixed effects \textbf{(M2; p = 0.09 -- )}, highlighting
strain and sex as salient variables. Two strains exhibit particularly
strong effects on MLH1 focus count (M3;
\emph{domesticus\textsuperscript{G}} p = \ensuremath{2.19\times 10^{-6}}
; \emph{domesticus\textsuperscript{LEW}} p = 0.02), with sex-strain
interactions involving three strains (M3;
\emph{domesticus\textsuperscript{G}} p = 0;
\emph{molossinus\textsuperscript{MSM}} p = 0 ;
\emph{musculus\textsuperscript{PWD}} p =
\ensuremath{5.31\times 10^{-4}}).

In separate analyses of males (M4; n = 80 - 71), three strains
disproportionately shape MLH1 focus count (as observed in Figure 1C):
\emph{musculus\textsuperscript{PWD}} (p =
\ensuremath{3.08\times 10^{-7}} ; effect = 6.11 ),
\emph{molossinus\textsuperscript{MSM}} (p =
\ensuremath{3.52\times 10^{-9}}; effect = 7.02), and
\emph{musculus\textsuperscript{SKIVE}} (p =
\ensuremath{7.59\times 10^{-4}}; effect = 4.04 ). These three strains
point to substantial evolution in the genome-wide recombination rate in
spermatocytes; we subsequently refer to them as ``high-recombination''
strains.

In females (M4; n= 76 - 76), four strains affect MLH1 focus count:
\emph{domesticus\textsuperscript{G}} (p =
\ensuremath{1.16\times 10^{-5}}; effect = 3.28),
\emph{molossinus\textsuperscript{MSM}} (p =
\ensuremath{2.97\times 10^{-5}}; effect = 2.98),
\emph{domesticus\textsuperscript{LEW}} (p = 0.03 ; effect = 1.7), and
\emph{musculus\textsuperscript{PWD}} (p = 0.08; effect = 1.16). Strain
effect sizes in females are modest in magnitude compared to those in
males.

Together, these results demonstrate that the genome-wide recombination
rate evolves in a highly sex-specific manner.

\hypertarget{synaptonemal-complexes-are-longer-in-females}{%
\subsection{Synaptonemal complexes are longer in
females}\label{synaptonemal-complexes-are-longer-in-females}}

\textless new 1/n\textgreater{} The variation in sex differences in
recombination we discovered provided an opportunity to determine whether
sex differences in chromatin compaction, as measured by the length of
the synaptonemal complex (SC), are reversed when heterochiasmy is
reversed. In all strains except \emph{musculus\textsuperscript{SKIVE}},
females have longer SCs than males, whether SC length was estimated as
the total length across bivalents or as the length of short bivalents
(t-tests; all p \textless{} 0.05, except short bivalents in
\emph{musculus\textsuperscript{SKIVE}}, p = 0.11 -- 0.11`).

Among short bivalents (to which the female X bivalent does not
contribute), female to male ratios of mouse mean SC length range from
1.26 - 1.26 (\emph{musculus\textsuperscript{PWD}}) to 1.49 - 1.49
(\emph{domesticus\textsuperscript{WSB}}) across strains. That females
have longer SCs is further supported by models that include covariates,
which identify sex as the most consistently significant effect for total
SC length (M1: p = 7.16\^{}\{-20\} -- \ensuremath{1.2\times 10^{-31}} ;
M2: p = 5.33\^{}\{-4\} -- \ensuremath{1.95\times 10^{-8}}; M3: p =0.05
-- 0.01 ) and short bivalent length (M1: p = 1.27\^{}\{-11 --
\ensuremath{1.27\times 10^{-11}}\}; M2: p = 1.98\^{}\{-7\} --
\ensuremath{1.98\times 10^{-7}}; M3: p = 1.98\^{}\{-7\} --
\ensuremath{1.98\times 10^{-7}}).

The existence of some subspecies and strain effects (M1 subspecies*sex:
p = 0.03 -- 0.03; M2: strain-\emph{musculus\textsuperscript{SKIVE}} :
p=0.02 -- 0.02 ; \emph{musculus\textsuperscript{SKIVE}}*male: p = 0.08
-- 0.08) further indicates that SC length has evolved among strains and
among subspecies.

In summary, two approaches for measuring SC length demonstrate that
females have longer SCs (chromosome axes), even in strains in which
males recombine more. This pattern implies that in high-recombination
strains, spermatocytes have less space than oocytes in which to position
additional crossovers.

\textbf{double check that these are the only sig factors for all the
models}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{females-and-males-differ-in-crossover-positions-and-crossover-interference}{%
\subsection{Females and males differ in crossover positions and
crossover
interference}\label{females-and-males-differ-in-crossover-positions-and-crossover-interference}}

\textless new 1/n\textgreater{} We used normalized positions of MLH1
foci along bivalents with a single focus to compare crossover location
while controlling for differences in SC length. In all strains, MLH1
foci tend to be closer to the telomere in males (mean normalized
position in males: 0.68--0.68; mean normalized position in females: 0.56
-- 0.56; t-test; p = 3.84\^{}\{-23\} --
\ensuremath{3.84\times 10^{-23}}). Sex is also the strongest determinant
of MLH1 focus position in the models we tested (M1: p = 7.79\^{}\{-26 --
\ensuremath{7.79\times 10^{-26}}\}; M2: p = 5.39\^{}\{-8\} --
\ensuremath{5.39\times 10^{-8}}; M3: p = 5.39\^{}\{-8\} --
\ensuremath{5.39\times 10^{-8}}).

\textless new 2/n\textgreater{} Males have longer normalized mean
inter-focal distances (IFD\textsubscript{norm}) than females in seven
out of eight strains (t-tests; p \textless{} 0.05), with only
\emph{musculus\textsuperscript{KAZ}} showing no difference (p = 0.33 --
0.33). Examination of IFD\textsubscript{norm} distributions indicates
that females are centered at approximately 50\% and show a slight
enrichment of low (\textless25\%) values, whereas males are enriched for
higher values. Models treating IFDnorm as the dependent variable support
the inference of stronger interference in males, with sex being the most
significant variable (M1: p = 3.11\^{}\{-12\} --
\ensuremath{3.11\times 10^{-12}} ; M2: p = 0.01 -- 0.01; M3: p = 0.01 --
0.01). In contrast, there is no clear signal of sex differences in raw
mean inter-focal distances (IFD\textsubscript{raw}) across the full set
of strains, whether they are considered separately or together.
Visualization of normalized MLH1 foci positions on bivalents with two
crossovers (Figure 4C) further suggests that interference distances vary
more in females than in males, and that males display a stronger
telomeric bias in the placement of the distal crossover.

In summary, controlling for differences in SC length (chromatin
compaction) indicates that interference is stronger in males, whereas
interference on the physical scale is similar in the sexes.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{evolution-of-genome-wide-recombination-rate-is-dispersed-across-bivalents-associated-with-double-strand-break-number-and-connected-to-crossover-interference}{%
\subsection{Evolution of genome-wide recombination rate is dispersed
across bivalents, associated with double-strand break number, and
connected to crossover
interference}\label{evolution-of-genome-wide-recombination-rate-is-dispersed-across-bivalents-associated-with-double-strand-break-number-and-connected-to-crossover-interference}}

\textless new 1/n\textgreater{} We used the contrast between males from
high-recombination strains and males from low-recombination strains to
identify features of the recombination landscape associated with
evolutionary transitions in the genome-wide recombination rate. We
considered proportions of bivalents with different numbers of
crossovers, double-strand break number, SC length, and crossover
positioning.

\textless new 2/n\textgreater{} Ninety-six percent of single bivalents
in our pooled dataset (n = 9,576 -- 9,575 ) have either one or two MLH
foci (Figure 3). The proportions of single-focus (1CO) bivalents
vs.~double-focus (2CO) bivalents distinguish high-recombination strains
from low-recombination strains (Figure 3). High-recombination strains
are enriched for 2CO bivalents at the expense of 1CO bivalents:
proportions of 2CO bivalents are 0.33 -- 0.33 in
\emph{musculus\textsuperscript{SKIVE}}, 0.44 -- 0.44 in
\emph{musculus\textsuperscript{PWD}} , and 0.51 -- 0.51 in
\emph{molossinus\textsuperscript{MSM}} (Supplemental Figure 3).
Following patterns in the genome-wide recombination rate, male
\emph{musculus\textsuperscript{PWD}} and male
\emph{molossinus\textsuperscript{MSM}} have 2CO proportions that are
more similar to each other than to strains from their own subspecies
(chi-square tests; \emph{musculus\textsuperscript{PWD}}
vs.~\emph{molossinus\textsuperscript{MSM}}: p = 0.37 -- 0.37 ;
\emph{musculus\textsuperscript{PWD}}
vs.~\emph{musculus\textsuperscript{KAZ}}: p = 1.23\^{}\{-31\} --
\ensuremath{1.23\times 10^{-31}}; \emph{molossinus\textsuperscript{MSM}}
vs.~\emph{molossinus\textsuperscript{MOLF}}: p = 2.65\^{}\{-6\} --
\ensuremath{2.82\times 10^{-6}} ). These results demonstrate that
evolution of the genome-wide recombination rate reflects changes in
crossover number across multiple bivalents.

\textless new 3/n\textgreater{} To begin to localize evolution of
genome-wide recombination rate to steps of the recombination pathway, we
counted DMC1 foci in prophase spermatocytes as markers for double-strand
breaks (DSBs). DMC1 foci were counted in a total of 76 -- 76 early
zygotene and 75 -- 75 late zygotene spermatocytes from two
high-recombination strains (\emph{musculus\textsuperscript{PWD}} and
\emph{molossinus\textsuperscript{MSM}}) and three low-recombination
strains (\emph{musculus\textsuperscript{KAZ}},
\emph{domesticus\textsuperscript{WSB}}, and
\emph{domesticus\textsuperscript{G}}). High-recombination strains have
significantly more DMC1 foci than low-recombination strains in early
zygotene cells (t-test; p \textless{} 10\^{}\{-6\} --
\ensuremath{10^{-6}} ). In contrast, the two strain groups do not differ
in DMC1 foci in late zygotene cells (t-test; p = 0.66 -- 0.66 ). Since
DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of
MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs
designated as COs. High-recombination and low-recombination strains do
not differ in the MLH1/DMC1 ratio, whether DMC1 foci were counted in
early zygotene cells or late zygotene cells (t-test; p \textgreater{}
0.05). These results raise the possibility that the evolution of
genome-wide recombination rate is primarily determined by processes that
precede the CO/NCO decision, at least in house mice.

\textless new 4/n\textgreater{} Total SC length only partially
differentiates high-recombination strains from low-recombination strains
(Figure 5). Whereas high-recombination strains as a group have
significantly greater total SC length than low-recombination strains
(t-test; p = 0.01 -- 0.02 ), separate tests within subspecies show that
the two strain categories differ within \emph{M. m. molossinus} (p =
0.02 -- \ensuremath{4.71\times 10^{-4}} ) but not within \emph{M. m.
musculus} (p = 0.4 -- 0.65 ). Additionally, mouse means for the reduced
(short and long) bivalent datasets do not differ between
high-recombination and low-recombination strains (t-test; short: p =
0.88 -- 0.88 ; long: p = 0.19 -- 0.19 ).

Write results for M4 and M5 here total SC M4 results

Musc subspecies very \textbf{\emph{, Mol }} (both ++) G, LEW, MSM SKIVE,
KAZ (- -)

M5 (almost all strains are significant) G, LEW, PWD, MSM MOLF, SKIVE,
KAZ, TOM, AST, CZECH

In a model with total SC length as the dependent variable (\emph{M2?}),
two subspecies effects are significant (\emph{M. m. musculus} p =
2.33\^{}\{-6\} -- 0.65, \emph{M. m. molossinus}, p = 21\^{}\{-6\} --
\ensuremath{4.71\times 10^{-4}}). In models with SC lengths of short and
long bivalents as dependent variables, several subspecies and strain
effects reach significance (p \textless{} 0.05), but they are not
consistent across models. Collectively, these results reveal that
evolution of SC length is not strongly associated with evolution of
genome-wide recombination rate in house mice.

\textless new 5/n\textgreater{} In summary, evolution of the genome-wide
recombination rate in males is connected to double-strand break number
and crossover interference, but not to SC length and crossover position
(on single-crossover bivalents).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{discussion}{%
\section{DISCUSSION}\label{discussion}}

\textless new 1/n\textgreater{} By comparing recombination rates in
females and males from the same diverse set of genetic backgrounds, we
isolated sex as a primary factor in the evolution of this fundamental
meiotic trait. Recombination rate differences are more pronounced in
males than females. Because inter-strain divergence times are identical
for the two sexes, this observation demonstrates that the genome-wide
recombination rate evolves faster in males, at least in house mice. More
generally, recombination rate divergence is decoupled in females and
males. These disparities are remarkable given that recombination rates
for the two sexes were measured in identical genomic backgrounds (other
than the presence/absence of the Y chromosome). Our results provide the
strongest evidence yet that the genome-wide recombination rate follows
distinct evolutionary trajectories in males and females.

\textless new 2/n\textgreater{} At the genetic level, the sex-specific
evolution we documented indicates that some mutations responsible for
divergence in recombination rate have dissimilar phenotypic effects in
females and males. A subset of the genetic variants associated with
genome-wide recombination rate within populations of humans (Kong et
al., 2004,, 2008, 2014; Halldorsson et al., 2019), Soay sheep (Johnston
et al., 2016), and cattle (Ma et al., 2015; Shen et al., 2018) appear to
show sex-specific properties, including opposite effects in females and
males. Furthermore, inter-sexual correlations for recombination rate are
weak in humans (Fledel-Alon et al., 2011) and Soay sheep (Johnston et
al., 2016). Crosses between the strains we surveyed could be used to
identify and characterize the genetic variants responsible for
recombination rate evolution in house mice (Dumont and Payseur, 2011;
Wang et al., 2019; Wang and Payseur, 2017). These variants could
differentially affect females and males at any step in the recombination
pathway. Although our DMC1 profiling was limited to males from a small
number of strains (for practical reasons), our findings suggest that
mutations that determine the number of double-strand breaks contribute
to sex-specific evolution in the recombination rate. A study of two
classical inbred strains and one wild-derived inbred strain of house
mice also found a positive association between crossover number and
double-strand break number in males (Baier et al., 2014).

\textless new 3/n\textgreater{} Another implication of our results is
that the connection between recombination rate and fitness differs
between males and females. Little is known about whether and how natural
selection shapes recombination rate in nature (Dapper and Payseur, 2017;
Ritz et al., 2017). Samuk et al. (2020) recently used a quantitative
genetic test to conclude that an 8\% difference in genome-wide
recombination rate between females from two populations of
\emph{Drosophila pseudoobscura} was caused by natural selection.
Applying similar strategies to species in which both sexes recombine,
including house mice, would be a logical next step to understanding the
sex-specific evolution of recombination rate.

\textless new 4/n\textgreater{} Population genetic models have been
built to explain sexual dimorphism in the number and placement of
crossovers, which is a common phenomenon (Brandvain and Coop, 2012;
Sardell and Kirkpatrick, 2020). Modifier models predicted that lower
recombination rates in males will result from haploid selection
(Lenormand, 2003) or sexually antagonistic selection on coding and
cis-regulatory regions of genes (Sardell and Kirkpatrick, 2020). Another
modifier model showed that meiotic drive could stimulate female-specific
evolution of the recombination rate (Brandvain and Coop, 2012). Although
these models fit the conserved pattern of sex differences in crossover
positions, they do not readily explain our observations of sex-specific
evolution in the genome-wide recombination rate. In particular, the
alternation across strains in which sex has more crossovers is
unexpected.

\textless new 5/n\textgreater{} We propose an alternative interpretation
of our findings based on the cell biology of gametogenesis. During
meiosis, achieving a stable chromosome structure requires the attachment
of kinetochores to opposite poles of the cell and at least one crossover
to create tension across the sister chromosome cohesion distal to
chiasmata (Lane and Kauppi, 2019; VanVeen and Hawley, 2003). The spindle
assembly checkpoint (SAC) prevents aneuploidy by ensuring that all
bivalents are correctly attached to the microtubule spindle
(``bi-oriented'') before starting the metaphase-to-anaphase transition
via the release of the sister cohesion holding homologs together (Lane
and Kauppi, 2019). Hence, selection seems likely to favor mutations that
optimize the process of bi-orientation and chromosome separation,
thereby prohibiting the SAC from delaying the cell cycle or triggering
apoptosis. Multiple lines of evidence indicate that the SAC is more
effective in spermatogenesis than in oogenesis (Lane and Kauppi, 2019),
perhaps due to the presence of the centrosome spindle (So et al., 2019)
and larger cell volume (Kyogoku and Kitajima, 2017) in oocytes. The
higher stringency of the SAC during spermatogenesis suggests that
selection will be better at removing mutations that interfere with
bi-orientation in males than in females. Therefore, faster male
evolution of the genome-wide recombination rate could be driven by the
more stringent SAC acting on chromosome structures at the metaphase I
alignment.

\textless new 5/n\textgreater{} Our SAC model is consistent with other
features of our data. We showed that widespread sex differences in
broad-scale crossover positioning (Sardell and Kirkpatrick, 2020) apply
across house mice, even in lineages where the direction of heterochiasmy
is reversed. The number and placement of crossovers affects the area of
sister chromosome cohesion distal to crossovers which needs to be
released for the first reductional chromosome segregation (Dumont and
Desai, 2012; Lane and Kauppi, 2019; Subramanian and Hochwagen, 2014;
VanVeen and Hawley, 2003). Faster spermatogenesis may select for
synchronization of the separation across all homologs within the cell
(\textbf{Kudo} ), whereas in oogenesis, the slower cell cycle and
multiple arrest stages may require chromosome structures with greater
stability on the MI spindle, especially for those organisms that undergo
dictyate arrest (Lee, 2019).

\textless new 6/n\textgreater{} We propose that the SAC model also can
explain the correlated evolution of stronger crossover interference and
higher genome-wide recombination rate in male house mice. Our results
show that crossovers are spaced further apart in strains enriched for
double-crossover bivalents when SC length is taken into account and
chromosome size effects are minimized. Assuming chromatin compaction
between (prophase) pachytene and metaphase is uniform along bivalents,
this increased spacing is expected to expand the area for sister
cohesion to connect homologs and may improve the fidelity of chromosomal
segregation. While the SAC model postulates direct fitness effects of
interference, a modifier model predicted that indirect selection on
recombination rate -- via its modulation of offspring genotypes -- can
strengthen interference as well (Goldstein et al., 1993).

\textless new 7/n\textgreater{} Regardless of the underlying mechanism,
our results provide a rare demonstration that crossover interference can
diverge over short evolutionary timescales. The notion that stronger
interference can co-evolve with higher genome-wide recombination rate is
supported by differences between breeds of cattle (Ma et al. (2015)) and
differences between populations of white-footed mice (Peterson et al.,
2019). In contrast, mammalian species with stronger interference tend to
exhibit lower genome-wide recombination rates (Segura et al. (2013),
Otto and Payseur (2019)). Collectively, these patterns suggest that
inferences about the evolutionary dynamics of interference depend on the
timescale under consideration.

\textless new 8/n\textgreater{} Our findings further reveal that
evolution of the genome-wide recombination rate does not require major
changes in the degree of chromatin compaction. Female house mice
consistently show longer SCs, even in strains with more recombination in
males. Studies in mice (Lynn et al., 2002, p. @petkov2007) and humans
(Gruhn et al., 2013, p. @tease2004) suggest that chromosomal axes are
longer (and DNA loops are shorter) in females than males. Some authors
have suggested that conserved sex differences in crossover positioning
(more uniform placement in females) and interference strength (stronger
interference in males) could be due to looser chromatin packing of the
meiotic chromosome structure in females (Haenel et al., 2018; Petkov et
al., 2007). A cellular model designed to explain interference attributes
sexual dimorphism in chromatin structure to greater cell volumes and
oscillatory movements of telomeres and kinetochores in oocytes (Hultén,
2011). More recent work connects the sparser recombination landscape to
sex differences in the crossover repair pathway (Wang et al., 2017).

\textless new 9/n\textgreater{} Our conclusions are accompanied by
several caveats. First, MLH1 foci only identify interfering crossovers
(Holloway et al., 2008). Although most crossovers belong to this class
\textbf{(REF)}, our approach likely underestimated genome-wide
recombination rates. Evolution of the number of non-interfering
crossovers is a topic worth examining. A second limitation is that our
investigation of crossover locations was confined to the relatively low
resolution possible with immunofluorescent cytology. Positioning
crossovers with higher resolution could reveal additional evolutionary
patterns. Finally, the panel of inbred lines we surveyed may not be
representative of recombination rate variation within and between
subspecies of house mice. We considered most available wild-derived
inbred lines, but house mice have a broad geographic distribution.
Nevertheless, we expect our primary conclusion that recombination rate
evolves in a sex-specific manner to be robust to geographic sampling
because differences between females and males exist for the same set of
inbred strains.

\textless new 10/n\textgreater{} While the causes of sex differences in
recombination remain mysterious Lenormand et al. (2016), our conclusions
have implications for a wide range of recombination research. For
biologists uncovering the cellular and molecular determinants of
recombination, our results suggest that mechanistic differences between
the sexes could vary by genetic background. For researchers charting the
evolutionary trajectory of recombination, our findings indicate that
sex-specific comparisons are crucial. For theoreticians building
evolutionary models of recombination, different fitness regimes and
genetic architectures in females and males should be considered.
Elevating sex as a primary determinant of recombination would be a
promising step toward integrating knowledge of cellular mechanisms with
evolutionary patterns to understand recombination rate variation in
nature.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{references}{%
\section*{REFERENCES}\label{references}}
\addcontentsline{toc}{section}{REFERENCES}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-anderson1999}{}%
Anderson LK, Reeves A, Webb LM, Ashley T. 1999. Distribution of crossing
over on mouse synaptonemal complexes using immunofluorescent
localization of mlh1 protein. \emph{Genetics} \textbf{151}:1569--1579.

\leavevmode\hypertarget{ref-baier2014}{}%
Baier B, Hunt P, Broman KW, Hassold T. 2014. Variation in genome-wide
levels of meiotic recombination is established at the onset of prophase
in mammalian males. \emph{PLoS genetics} \textbf{10}.

\leavevmode\hypertarget{ref-lme4}{}%
Bates D, Mächler M, Bolker B, Walker S. 2015. Fitting linear
mixed-effects models using lme4. \emph{Journal of Statistical Software}
\textbf{67}:1--48.
doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}

\leavevmode\hypertarget{ref-brandvain2012scrambling}{}%
Brandvain Y, Coop G. 2012. Scrambling eggs: Meiotic drive and the
evolution of female recombination rates. \emph{Genetics}
\textbf{190}:709--723.

\leavevmode\hypertarget{ref-CahoonLibuda2019}{}%
Cahoon CK, Libuda DE. 2019. Leagues of their own: Sexually dimorphic
features of meiotic prophase i. \emph{Chromosoma} 1--16.

\leavevmode\hypertarget{ref-DapperPayseur2017}{}%
Dapper AL, Payseur BA. 2017. Connecting theory and data to understand
recombination rate evolution. \emph{Philosophical Transactions of the
Royal Society B: Biological Sciences} \textbf{372}:20160469.

\leavevmode\hypertarget{ref-dumont2011}{}%
Dumont BL, Payseur BA. 2011. Genetic analysis of genome-scale
recombination rate evolution in house mice. \emph{PLoS genetics}
\textbf{7}.

\leavevmode\hypertarget{ref-dumontDesai2012}{}%
Dumont J, Desai A. 2012. Acentrosomal spindle assembly and chromosome
segregation during oocyte meiosis. \emph{Trends in cell biology}
\textbf{22}:241--249.

\leavevmode\hypertarget{ref-fledel2011}{}%
Fledel-Alon A, Leffler EM, Guan Y, Stephens M, Coop G, Przeworski M.
2011. Variation in human recombination rates and its genetic
determinants. \emph{PloS one} \textbf{6}.

\leavevmode\hypertarget{ref-goldstein1993}{}%
Goldstein DB, Bergman A, Feldman MW. 1993. The evolution of
interference: Reduction of recombination among three loci.
\emph{Theoretical population biology} \textbf{44}:246--259.

\leavevmode\hypertarget{ref-gruhn2013}{}%
Gruhn JR, Rubio C, Broman KW, Hunt PA, Hassold T. 2013. Cytological
studies of human meiosis: Sex-specific differences in recombination
originate at, or prior to, establishment of double-strand breaks.
\emph{PloS one} \textbf{8}.

\leavevmode\hypertarget{ref-haenel2018}{}%
Haenel Q, Laurentino TG, Roesti M, Berner D. 2018. Meta-analysis of
chromosome-scale crossover rate variation in eukaryotes and its
significance to evolutionary genomics. \emph{Molecular ecology}
\textbf{27}:2477--2497.

\leavevmode\hypertarget{ref-halldorsson2019}{}%
Halldorsson BV, Palsson G, Stefansson OA, Jonsson H, Hardarson MT,
Eggertsson HP, Gunnarsson B, Oddsson A, Halldorsson GH, Zink F, others.
2019. Characterizing mutagenic effects of recombination through a
sequence-level genetic map. \emph{Science} \textbf{363}:eaau1043.

\leavevmode\hypertarget{ref-holloway2008mus81}{}%
Holloway JK, Booth J, Edelmann W, McGowan CH, Cohen PE. 2008. MUS81
generates a subset of mlh1-mlh3--independent crossovers in mammalian
meiosis. \emph{PLoS genetics} \textbf{4}.

\leavevmode\hypertarget{ref-hulten2011_COM}{}%
Hultén MA. 2011. On the origin of crossover interference: A chromosome
oscillatory movement (com) model. \emph{Molecular cytogenetics}
\textbf{4}:10.

\leavevmode\hypertarget{ref-johnston2016_soay}{}%
Johnston SE, Bérénos C, Slate J, Pemberton JM. 2016. Conserved genetic
architecture underlying individual recombination rate variation in a
wild population of soay sheep (ovis aries). \emph{Genetics}
\textbf{203}:583--598.

\leavevmode\hypertarget{ref-koehler2002}{}%
Koehler KE, Cherry JP, Lynn A, Hunt PA, Hassold TJ. 2002. Genetic
control of mammalian meiotic recombination. I. Variation in exchange
frequencies among males from inbred mouse strains. \emph{Genetics}
\textbf{162}:297--306.

\leavevmode\hypertarget{ref-Kong2004}{}%
Kong A, Barnard J, Gudbjartsson DF, Thorleifsson G, Jonsdottir G,
Sigurdardottir S, Richardsson B, Jonsdottir J, Thorgeirsson T, Frigge
ML, others. 2004. Recombination rate and reproductive success in humans.
\emph{Nature genetics} \textbf{36}:1203--1206.

\leavevmode\hypertarget{ref-Kong2014}{}%
Kong A, Thorleifsson G, Frigge ML, Masson G, Gudbjartsson DF, Villemoes
R, Magnusdottir E, Olafsdottir SB, Thorsteinsdottir U, Stefansson K.
2014. Common and low-frequency variants associated with genome-wide
recombination rate. \emph{Nature genetics} \textbf{46}:11.

\leavevmode\hypertarget{ref-Kong2008}{}%
Kong A, Thorleifsson G, Stefansson H, Masson G, Helgason A, Gudbjartsson
DF, Jonsdottir GM, Gudjonsson SA, Sverrisson S, Thorlacius T, others.
2008. Sequence variants in the rnf212 gene associate with genome-wide
recombination rate. \emph{Science} \textbf{319}:1398--1401.

\leavevmode\hypertarget{ref-kyogoku2017}{}%
Kyogoku H, Kitajima TS. 2017. Large cytoplasm is linked to the
error-prone nature of oocytes. \emph{Developmental cell}
\textbf{41}:287--298.

\leavevmode\hypertarget{ref-LaneKauppi2019}{}%
Lane S, Kauppi L. 2019. Meiotic spindle assembly checkpoint and
aneuploidy in males versus females. \emph{Cellular and molecular life
sciences} \textbf{76}:1135--1150.

\leavevmode\hypertarget{ref-Lee2019}{}%
Lee J. 2019. Is age-related increase of chromosome segregation errors in
mammalian oocytes caused by cohesin deterioration? \emph{Reproductive
Medicine and Biology}.

\leavevmode\hypertarget{ref-lenormand2003}{}%
Lenormand T. 2003. The evolution of sex dimorphism in recombination.
\emph{Genetics} \textbf{163}:811--822.

\leavevmode\hypertarget{ref-lenormand2016}{}%
Lenormand T, Engelstädter J, Johnston SE, Wijnker E, Haag CR. 2016.
Evolutionary mysteries in meiosis. \emph{Philosophical Transactions of
the Royal Society B: Biological Sciences} \textbf{371}:20160001.

\leavevmode\hypertarget{ref-lynn2002}{}%
Lynn A, Koehler KE, Judis L, Chan ER, Cherry JP, Schwartz S, Seftel A,
Hunt PA, Hassold TJ. 2002. Covariation of synaptonemal complex length
and mammalian meiotic exchange rates. \emph{Science}
\textbf{296}:2222--2225.

\leavevmode\hypertarget{ref-ma2015_cattle}{}%
Ma L, O'Connell JR, VanRaden PM, Shen B, Padhi A, Sun C, Bickhart DM,
Cole JB, Null DJ, Liu GE, others. 2015. Cattle sex-specific
recombination and genetic control from a large pedigree analysis.
\emph{PLoS genetics} \textbf{11}.

\leavevmode\hypertarget{ref-murdoch2010}{}%
Murdoch B, Owen N, Shirley S, Crumb S, Broman KW, Hassold T. 2010.
Multiple loci contribute to genome-wide recombination levels in male
mice. \emph{Mammalian Genome} \textbf{21}:550--555.

\leavevmode\hypertarget{ref-nagaoka2012}{}%
Nagaoka SI, Hassold TJ, Hunt PA. 2012. Human aneuploidy: Mechanisms and
new insights into an age-old problem. \emph{Nature Reviews Genetics}
\textbf{13}:493--504.

\leavevmode\hypertarget{ref-ottoPaysuer2019}{}%
Otto SP, Payseur BA. 2019. Crossover interference: Shedding light on the
evolution of recombination. \emph{Annual review of genetics}
\textbf{53}:19--44.

\leavevmode\hypertarget{ref-peters_1997}{}%
Peters AH, Plug AW, Vugt MJ van, De Boer P. 1997. SHORT COMMUNICATIONS A
drying-down technique for the spreading of mammalian meiocytes from the
male and female germline. \emph{Chromosome research} \textbf{5}:66--68.

\leavevmode\hypertarget{ref-peterson2019}{}%
Peterson AL, Miller ND, Payseur BA. 2019. Conservation of the
genome-wide recombination rate in white-footed mice. \emph{Heredity}
\textbf{123}:442--457.

\leavevmode\hypertarget{ref-petkov2007}{}%
Petkov PM, Broman KW, Szatkiewicz JP, Paigen K. 2007. Crossover
interference underlies sex differences in recombination rates.
\emph{Trends in Genetics} \textbf{23}:539--542.

\leavevmode\hypertarget{ref-Ritz2017}{}%
Ritz KR, Noor MA, Singh ND. 2017. Variation in recombination rate:
Adaptive or not? \emph{Trends in Genetics} \textbf{33}:364--374.

\leavevmode\hypertarget{ref-samuk2020}{}%
Samuk K, Manzano-Winkler B, Ritz KR, Noor MA. 2020. Natural selection
shapes variation in genome-wide recombination rate in drosophila
pseudoobscura. \emph{Current Biology}.

\leavevmode\hypertarget{ref-sardell_sex_2020}{}%
Sardell JM, Kirkpatrick M. 2020. Sex differences in the recombination
landscape. \emph{The American Naturalist} \textbf{195}:361--379.
doi:\href{https://doi.org/10.1086/704943}{10.1086/704943}

\leavevmode\hypertarget{ref-RLRsim}{}%
Scheipl F, Greven S, Kuechenhoff H. 2008. Size and power of tests for a
zero random effect variance or polynomial regression in additive and
linear mixed models. \emph{Computational Statistics \& Data Analysis}
\textbf{52}:3283--3299.

\leavevmode\hypertarget{ref-segura2013}{}%
Segura J, Ferretti L, Ramos-Onsins S, Capilla L, Farré M, Reis F,
Oliver-Bonet M, Fernández-Bellón H, Garcia F, Garcia-Caldés M, others.
2013. Evolution of recombination in eutherian mammals: Insights into
mechanisms that affect recombination rates and crossover interference.
\emph{Proceedings of the Royal Society B: Biological Sciences}
\textbf{280}:20131945.

\leavevmode\hypertarget{ref-Shen2018_cattle}{}%
Shen B, Jiang J, Seroussi E, Liu GE, Ma L. 2018. Characterization of
recombination features and the genetic basis in multiple cattle breeds.
\emph{BMC genomics} \textbf{19}:304.

\leavevmode\hypertarget{ref-So2019}{}%
So C, Seres KB, Steyer AM, Mönnich E, Clift D, Pejkovska A, Möbius W,
Schuh M. 2019. A liquid-like spindle domain promotes acentrosomal
spindle assembly in mammalian oocytes. \emph{Science}
\textbf{364}:eaat9557.

\leavevmode\hypertarget{ref-stapley_variation_2017}{}%
Stapley J, Feulner PGD, Johnston SE, Santure AW, Smadja CM. 2017.
Variation in recombination frequency and distribution across eukaryotes:
Patterns and processes. \emph{Philosophical Transactions of the Royal
Society B: Biological Sciences} \textbf{372}:20160455.
doi:\href{https://doi.org/10.1098/rstb.2016.0455}{10.1098/rstb.2016.0455}

\leavevmode\hypertarget{ref-subramanian2014}{}%
Subramanian VV, Hochwagen A. 2014. The meiotic checkpoint network:
Step-by-step through meiotic prophase. \emph{Cold Spring Harbor
perspectives in biology} \textbf{6}:a016675.

\leavevmode\hypertarget{ref-Rstudio}{}%
Team R. 2015. RStudio: Integrated Development Environment for R.

\leavevmode\hypertarget{ref-tease2004}{}%
Tease C, Hulten M. 2004. Inter-sex variation in synaptonemal complex
lengths largely determine the different recombination rates in male and
female germ cells. \emph{Cytogenetic and genome research}
\textbf{107}:208--215.

\leavevmode\hypertarget{ref-vanVeen2003}{}%
VanVeen JE, Hawley RS. 2003. Meiosis: When even two is a crowd.
\emph{Current Biology} \textbf{13}:R831--R833.

\leavevmode\hypertarget{ref-wang2019_SC}{}%
Wang RJ, Dumont BL, Jing P, Payseur BA. 2019. A first genetic portrait
of synaptonemal complex variation. \emph{PLoS genetics}
\textbf{15}:e1008337.

\leavevmode\hypertarget{ref-Wang2017island}{}%
Wang RJ, Payseur BA. 2017. Genetics of genome-wide recombination rate
evolution in mice from an isolated island. \emph{Genetics}
\textbf{206}:1841--1852.

\leavevmode\hypertarget{ref-wang2017inefficient}{}%
Wang S, Hassold T, Hunt P, White MA, Zickler D, Kleckner N, Zhang L.
2017. Inefficient crossover maturation underlies elevated aneuploidy in
human female meiosis. \emph{Cell} \textbf{168}:977--989.

\end{document}

### 2020-06-17 doc/MS_v3.Rmd
work on intro

@@
 output:
+  html_document:
   pdf_document:
@@
   word_document:
-  html_document:
 always_allow_html: yes
@@
 For the first time in many of these strains, We report a rare, direct, evolutionary comparison of recombination rate in females and males for this short evolutionary scale.
-We use rare strains from a broad range of geographic locations of the species territory.
-
-The gwRR is made up of crossovers occurring on individual chromosomes within cells.
-Thus we quantified meiotic chromosome morphology (SC length) and placement of crossovers to comprise an approximate picture of the recombination landscape.
-
+We choose Mus musculus because of the available strains across subspecies.
+Our design of using multiple inbred strains within subspecies  and multiple subspecies across the species --  (can provide information anaglogous to polymorphism within and divergence within subspecies of the house mouse.
 Our results indicate rapid male specific evolution of gwRR occurred in a subset of house mouse lines.
 We observed up to a 30% difference (translating into approximately ~7 more crossovers per cell), a surprising amount considering the short evolutionary time scale.
+(found previously for PWK (forjt) (PWD beth) – but mostly measures from males.
+
+Brief mouse background and (touch on polymorphism 
+-musculus subspecies – have mostly been samples from central Europe – we extend with samples from more eastern extent of the subspecies range in Russia and kazakstan. 
+-molossinus is endemic to the japan – a hybrid that arose with humans brining musculus and castaneus to the islands.
+(AGAIN the divergence between mus musculus subspecies is ~500K – and males and females share the same phylogeny)
+
+OUR results for genome wide recombination rate (variation) we observed highly distinct evolutionary trajectories between males and females.
+This is due to rapid male specific evolution in the gwRR from two strains in musculus and 1 strain in molossinus.
+(this pattern of rapid evolution was not observed in females)
+
+The metric of the genome wide recombination rate reflects a non-random process of crossover formation on individual chromosomes // the recombination landscape at the single chromosome level.
+The recombination landscape for single chromosome scale can be general summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
+
+We quantified the general recombination landscape because patterns in the recombination landscape are important to document because they can shed light on constraints and regulation of the processes ( could end up leading to variation in the genome wide recombination rates.
+Our results confirm (well established) previously reported patterns for sex differences in the general recombination landscape – and add by demonstrating that these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
+
+-------------------
 
@@
 Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a Photoflow wash (Kodak, cite).
+The following morning, slides were treated with a Photoflow wash (Kodak, diluted 1:200).
 Slides were stored at -20\*C if not stained immediately.
@@
 In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
-Four additional lnear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
+Four additional linear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
 

### 2020-06-17 doc/MS_v3.Rmd
continued work on intro

@@
 
-Meiosis can be reduced to the expression of (2n -> 4n -> 2n -> 1n) which tracks the duplication of a diploid genome into haploid cell products.
-The meiotic program relies on crossovers and the process of recombination to ensure the correct separation of chromosomes.
+Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process diploid genome producing haploid cell products.
+In most species the meiotic program relies on crossovers to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [Cahoon].
 The total number of COs per cell (at the 4n stage) is equal to the genome wide recombination rate (gwRR).
-**The genome wide recombination rate is a cell based metric for genome wide recombination rate is tightly connected to organisms fitness through correct chromosome segregation and fertility.**
-
-**at the fine scale / in regards to genetic variation**, In the indriect manner, the recombination rate regulates populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (cite).
-This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation (Begun Aquadro, Nachman, Payseur).
-
-**but back to the genome-wide recombination rate**, The process of chromosome segregation is integral to the process of producing haploid cells and can be considered a constraint in regards to the evolution.
-When the genome wide recombination rate is framed within the single cell context it is closely connected to the  contraints which are imposed at the single cell level.
-Reflecting the importance of it's role in chromosome segreation is that the  number of haploid pairs is strong predictor for genome wide RR across large species(taxa) / divergence scales [@ottoPaysuer2019; @stapley_variation_2017].
-(the number of crossovers per chromosome are highly conserved)<However, whie the rate of recombination is highly variable, the count of crossovers per chromosome is conserved between 1 to 3 (cite), suggesting that such thresholds/constraints may be behind the overall patterns.>
-
-**connection of single cell based gwRR to thresholds imposed at the single cell level** (why is gwRR the better scale for evolutionary patterns?)
+Additioanlly, the gwRR can regulate a populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (cite).
+This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
+**Because of it's role in facilitating chromosomes disjunction, the genome wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**
 
 ### 1. Understanding levels of variation in gwRR
+
+The process of chromosome segregation is integral to the process of producing haploid gametes and can be considered a constraint in regards to the evolution of the genome-wide recombination rate.
+Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination (cite) are considered to be the lower and upper thresholds for a genome-wide recombiantion rate [@nagaoka2012].
+The number of haploid pairs of chromosomes is strong predictor for genome wide recombination rate across large divergence scales [@ottoPaysuer2019; @stapley_variation_2017].
+Yet, within these thresholds the genome wide recombination rate still varies across species and individuals.
+Sex is one of the most notable axes of this form of variation.
+(examples, Humans 1.7 female to male
+alligators female to male ratio is X, temperature dependent sex determination)
+
+<2.	Gap is due to Lack of comparable measures>
+An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
+Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog,  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay], and mouse (CC cite) in which the role of sex is confounded with the contributions of genetic variation.
+(other gaps in data )
+
+<3.	Open Q, no first principle>
+Under anisogamy where gametogenesis is modified to produce distinct gametes (cite Gorelick?), the meiotic program is the same and there are no first principles which would predict the evolution of sexually dimorphic recombination rates.
+Yet heterochiasmy is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
+Although it is clear that the relationship between female and male recombination rates can differ among species, comparisons between and within closely related species are missing.
+Direct contrasts between females and males across a common, diverse set of genomic backgrounds would reveal whether the recombination rate evolves differently in the sexes.
+
+<4.	Additional metrics (more nuanced)>
+Initial focus/evidence on heterochiasmy was primarily focused on which sex had more recombination – the field has advanced since it’s first documentation (cite morgan 1914).
+Recently two specific features have been highlighted as conserved patterns of heterochiasmy: @CahoonLibuda2019 show that in species with meiotic chromatin (SC axis length) quantified for both sexes, longer chromosome axis generally has higher recombination rates.
+In @sardell_sex_2020, a survey of 51 species shows there are conserved broad scale recombination landscape differences.
+Generally males have telomere-bias crossover placement and females have more uniform placement.
+
+These observations raise the question of how conservation in sexually dimorphic patterns are maintained with the gwRR and it’s decomposed traits.
+
+-----------
 
@@
 
+The House mouse species complex arose from a recent radiation providing an opportunity to interrogate natural variation at short evolutionary scales, potentially narrowing the mutational space for identifying variants which would cause gwRR variation.
+Add divergence ranges.
+divergence spanning 0.5 mya to 5 mya  (cite / Geraldes et al 2008 and Geraldes et al 2010).
+
+Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway (cite) and previous crosses for understanding the genetic architecture of recombination rate variation @dumont2011house; @murdoch2010; @Wang2017island].
+Additionally due to their global distribution allows us to compare samples of natural diversity from a broad geographic range.
+(wild derived inbred strains – will capture genetic diversity not represented in classical lab strains.)
+
+**Wild derived inbred strains generate the best comparison of females and males, besides the sex chromosomes, the mouse for each genome are almost identical**.
+Inbred strains enable one of the most direct comparison of both female and male gametogenesis and enable us to conclude that sex is a primary factor.
+
+While there are strains with Robertsonian translocations, in house mouse and related murid species it is possible to assemble sets of strains with identical karyotypes, 20 pairs of acrocentric chromosomes.
+
+House mouse is well suited for single cell cytology approaches (due to large cell sizes).
+The single cell level allows integration of data at a closer connection to the molecular pathway and meiotic program.
+In addition to the quantification of gwRR.
+This study quantifies precursors to crossovers, double strand breaks (DSBs), and meiotic chromosome morphology across stages of meiosis which would not be accessible to with genetic linkage data and crosses.
+
+--------
+
 The House mouse complex arose from a recent radiation providing an opportunity to interrogate natural variation at short evolutionary scales, potentially narrowing the mutational space for identifying variants which would cause gwRR variation.
@@
 ### 3. What we accomplished in this paper
+
+Main messages: these results are centered/focused on a unique evo comparison 
+Advantages / importance of using multiple strains from and multiple subspecies
+-importance / advantage of sampling for multiple geographic locations.
+Chromosome morphology / general recombination landscape
+-results for general landscape – conserved sex differences  and
+
+---------------
+OUR results for genome wide recombination rate (variation) we observed highly distinct evolutionary trajectories between males and females.
+This is due to rapid male specific evolution in the gwRR from two strains in musculus and 1 strain in molossinus.
+(this pattern of rapid evolution was not observed in females).
+We observed up to a 30% difference (translating into approximately ~7 more crossovers per cell), a surprising amount considering the short evolutionary time scale.
+
+The metric of the genome wide recombination rate reflects a non-random process of crossover formation on individual chromosomes // the recombination landscape at the single chromosome level.
+The recombination landscape for single chromosome scale can be general summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
+
+We quantified the general recombination landscape because patterns in the recombination landscape are important to document because they can shed light on constraints and regulation of the processes ( could end up leading to variation in the genome wide recombination rates.
+Our results confirm (well established) previously reported patterns for sex differences in the general recombination landscape – and add by demonstrating that these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
+
+Conserved patterns of the recombination landscape of the single chromosome level – reflect constraints  At the single chromosome level – (crossover number is conserved – suggesting processes regulating the rec landscape at this scale might face similar constraints
+
+------
 

### 2020-06-17 doc/MS_v3.Rmd
more work on Intro

@@
 output:
-  html_document:
+  word_document:
   pdf_document:
@@
       in_header: header.tex
-  word_document:
+  html_document:
 always_allow_html: yes
@@
 Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
-**define genome-wide recombination rate**.
-
+The number of crossovers per meiotic cell is the  genome-wide recombination rate.
 Differences in recombination rate between females and males are perhaps the most striking form of this variation.
@@
 Our results indicate highly discordant evolutionary patterns in the two sexes.
-Whereas male recombination rates show evidence of rapid evolution over short evolutionary timescales, female recombination rates measured in the same strains are mostly static.
-These results strongly indicate that house mouse has two genome wide recombination rates which display distinct evolutionary trajectories.
+Whereas male recombination rates show evidence of rapid evolution over short timescales, female recombination rates measured in the same strains are mostly static.
+These results strongly indicate that house mouse has two, one for each sex, genome wide recombination rates which display distinct evolutionary trajectories.
 
@@
 
-### 0.   Importance of recombination rate
+<!--### 0.   Importance of recombination rate-->
 
-Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process diploid genome producing haploid cell products.
-In most species the meiotic program relies on crossovers to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [Cahoon].
-The total number of COs per cell (at the 4n stage) is equal to the genome wide recombination rate (gwRR).
-Additioanlly, the gwRR can regulate a populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (cite).
+  Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process of a diploid genome producing haploid gametes.
+In most species the meiotic program relies on crossovers to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [@CahoonLibuda2019].
+The total number of COs at this stage of meiosis is equal to the genome wide recombination rate (gwRR).
+In addition to this functional purpose, the gwRR can regulate a populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (Charlesworth 1993 cite).
 This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
-**Because of it's role in facilitating chromosomes disjunction, the genome wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**
 
-### 1. Understanding levels of variation in gwRR
+<!-- ### 1. Understanding levels of variation in gwRR -->
 
-The process of chromosome segregation is integral to the process of producing haploid gametes and can be considered a constraint in regards to the evolution of the genome-wide recombination rate.
-Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination (cite) are considered to be the lower and upper thresholds for a genome-wide recombiantion rate [@nagaoka2012].
-The number of haploid pairs of chromosomes is strong predictor for genome wide recombination rate across large divergence scales [@ottoPaysuer2019; @stapley_variation_2017].
+  **Due to it's role in facilitating chromosomes disjunction, the genome wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**  Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombiantion rate [@nagaoka2012] (Inoue).
 Yet, within these thresholds the genome wide recombination rate still varies across species and individuals.
 Sex is one of the most notable axes of this form of variation.
-(examples, Humans 1.7 female to male
-alligators female to male ratio is X, temperature dependent sex determination)
 
-<2.	Gap is due to Lack of comparable measures>
+<!-- (examples, Humans 1.7 female to male
+alligators female to male ratio is X, temperature dependent sex determination) -->
+
+<!-- The number of haploid pairs of chromosomes is strong predictor for genome wide recombination rate across large divergence scales [@ottoPaysuer2019; @stapley_variation_2017]. -->
+
+<!-- 2.Gap is due to Lack of comparable measures -->
+
 An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
 Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog,  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay], and mouse (CC cite) in which the role of sex is confounded with the contributions of genetic variation.
-(other gaps in data )
 
-<3.	Open Q, no first principle>
-Under anisogamy where gametogenesis is modified to produce distinct gametes (cite Gorelick?), the meiotic program is the same and there are no first principles which would predict the evolution of sexually dimorphic recombination rates.
+<!-- 3.	Open Q, no first principle -->
+
+Even under anisogamy, where gametogenesis is modified to produce distinct gametes across sexes, the meiotic program is the same.
+There are no first principles which would predict the evolution of sexually dimorphic recombination rates.
 Yet heterochiasmy is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
@@
 
-<4.	Additional metrics (more nuanced)>
-Initial focus/evidence on heterochiasmy was primarily focused on which sex had more recombination – the field has advanced since it’s first documentation (cite morgan 1914).
-Recently two specific features have been highlighted as conserved patterns of heterochiasmy: @CahoonLibuda2019 show that in species with meiotic chromatin (SC axis length) quantified for both sexes, longer chromosome axis generally has higher recombination rates.
-In @sardell_sex_2020, a survey of 51 species shows there are conserved broad scale recombination landscape differences.
-Generally males have telomere-bias crossover placement and females have more uniform placement.
-
+<!-- 4.	Additional metrics (more nuanced) -->  
+  Heterochiasimy was defined as sexual dimorphism in the total counts measures of recombination (Lenormand 2003).
+It's important to note that the genome-wide recombiantion rate can be decomposed into single chromosome patterns, refered to as recombination landscapes.
+Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
+In a review of data from mouse, celegans and worms with meiotic chromosome morphilogy (axis-loop structure) quantified for both sexes, @CahoonLibuda2019 find that the sex with longer chromosome axis generally has higher recombination rates.
+A survey of 51 species shows there are conserved broad scale recombination landscape differences.
+Generally males have telomere-bias crossover placement and females have more uniform placement [@sardell_sex_2020].
 These observations raise the question of how conservation in sexually dimorphic patterns are maintained with the gwRR and it’s decomposed traits.
@@
 
-Because we know the meiotic programs, we have expectations for where the thresholds(constraints) of the gwRR 
-the rate will fall.
-First, the lower threshold is defined as an obligate crossover per homolog pair which promotes proper chromosome segregation [@nagaoka2012].
-second (while there is less evidence,), the upper threshold maybe regulated/shaped/affected due to rates of ectopic recombination or the programmed DNA damage preceding crossovers act as constraints to the upper threshold of meiotic recombination (cite).
-Yet, Within these thresholds the genome wide recombination rate still varies across species and indiviuals.
-which is the subject of an open question in the field.  **the gwRR is the best metric for since both of these constraints are thought to act at the cell wide level.**
+<!-- ### 2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale -->
 
-**Sex is one of the most notable axes along which individuals vary**.
-Sexual dimorphism in recombination rates is called heterochiasmy [@lenormand2003].
-An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
-Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog,  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep) sheep [@johnston2016_soay], and mouse (CC cite) in which the role of sex is confounded with the contributions of genetic variation.
-Under anisogamy where a species has distinct gametes, there is no first principle which would predict the evolution of sexually dimorphic recombination rates.
-Yet heterochiasmy, is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
-Although it is clear that the relationship between female and male recombination rates can differ among species, comparisons between and within closely related species are missing.
-Direct contrasts between females and males across a common, diverse set of genomic backgrounds would reveal whether the recombination rate evolves differently in the sexes.
-
-<It’s important to note, – that the genome wide recombination rate can be decomposed – and connected to related traits (which also show conserved patterns of sexual dimorphism>
-
-**understanding heterochiasmy and sexual dimorphism in meiosis have advanced since first being documented (morgan 1914) -- more than just the direction -- more evidence is accumulating that there are conserved patterns of sexual dimorphism in meiotic recombinatio - related to recombination rate**.
-Recently two specific features have been highlighted as conserved patterns of heterochiasmy:
-
-@CahoonLibuda2019 show In species with meiotic chromatin (SC axis length) quantified for both sexes, longer chromosome axis generally has higher recombination rates. 
-@sardell_sex_2020, A survey of 51 species shows there are conserved broad scale recombination landscape differences.
-Generally males have telomere-bias crossover placement and females have more uniform placement.
-
-These observations raise the question of conservation in sexually patterns are maintained with the gwRR and it’s decomposed traits
-
-**remember these points kinda distract from gwRR aspect**
-
-### 2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale
-
-The House mouse species complex arose from a recent radiation providing an opportunity to interrogate natural variation at short evolutionary scales, potentially narrowing the mutational space for identifying variants which would cause gwRR variation.
-Add divergence ranges.
-divergence spanning 0.5 mya to 5 mya  (cite / Geraldes et al 2008 and Geraldes et al 2010).
-
-Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway (cite) and previous crosses for understanding the genetic architecture of recombination rate variation @dumont2011house; @murdoch2010; @Wang2017island].
+  The house mouse species, (_Mus musculus_), complex arose from a recent radiation (0.5 mya to 5 mya) (@geraldes2011; @geraldes2008) providing an opportunity to interrogate natural variation at short evolutionary scales.
+This feature narrows the mutational space for identifying variants which would cause gwRR variation.
+Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway **(cite)** and previous crosses for understanding the genetic architecture of recombination rate variation [@dumont2011; @murdoch2010; @Wang2017island].
 Additionally due to their global distribution allows us to compare samples of natural diversity from a broad geographic range.
-(wild derived inbred strains – will capture genetic diversity not represented in classical lab strains.)
 
@@
 
-While there are strains with Robertsonian translocations, in house mouse and related murid species it is possible to assemble sets of strains with identical karyotypes, 20 pairs of acrocentric chromosomes.
-
   House mouse is well suited for single cell cytology approaches (due to large cell sizes).
 The single cell level allows integration of data at a closer connection to the molecular pathway and meiotic program.
-In addition to the quantification of gwRR.
-This study quantifies precursors to crossovers, double strand breaks (DSBs), and meiotic chromosome morphology across stages of meiosis which would not be accessible to with genetic linkage data and crosses.
+(Allows us to compare counts from cells at is the same meiotic (molecular pathway context)).
 
@@
 
-The House mouse complex arose from a recent radiation providing an opportunity to interrogate natural variation at short evolutionary scales, potentially narrowing the mutational space for identifying variants which would cause gwRR variation.
-Add divergence ranges.
-divergence spanning 0.5 mya to 5 mya.
+<!-- ### 3. What we accomplished in this paper -->
 
-Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway (cite) and previous crosses for understanding the genetic architecture of recombination rate variation @dumont2011house; @murdoch2010; @Wang2017island].
-
-**Wild derived inbred strains generate the best comparison of females and males, besides the sex chromosomes, the mouse for each genome are almost identical**.
-Inbred strains enable one of the most direct comparison of both female and male --  versions of meiosis --enable use to conclude that sex is a primary factor --
-
-Additionally due to their global distribution allows us to compare samples of natural diversity from a broad geographic range.
-(wild derived inbred strains – will capture genetic diversity not represented in classical lab strains.)
-
-While there are strains with Robertsonian translocations, in house mouse and related murid species it is possible to assemble sets of strains with identical karyotypes, 20 pairs of acrocentric chromosomes.
-
-House mouse is well suited for single cell cytology approaches.
-The single cell level allows integration of data at a closer connection to the molecular pathway and meiotic program.
-In addition to the quantification of gwRR.
-This study quantifies precursors to crossovers, double strand breaks (DSBs), and meiotic chromosome morphology across stages of meiosis which would not be accessible to with genetic linkage data and crosses.
-
-### 3. What we accomplished in this paper
-
-Main messages: these results are centered/focused on a unique evo comparison 
+<!-- Main messages: these results are centered/focused on a unique evo comparison 
 Advantages / importance of using multiple strains from and multiple subspecies
@@
 Chromosome morphology / general recombination landscape
--results for general landscape – conserved sex differences  and
+-results for general landscape 
+– conserved sex differences  -->
 
 ---------------
-OUR results for genome wide recombination rate (variation) we observed highly distinct evolutionary trajectories between males and females.
+
+  (we quantify gwRR from both sexes, using house mouse inbred strains as a common genetic background).
+We observed highly distinct evolutionary trajectories between males and females.
 This is due to rapid male specific evolution in the gwRR from two strains in musculus and 1 strain in molossinus.
@@
 
-The metric of the genome wide recombination rate reflects a non-random process of crossover formation on individual chromosomes // the recombination landscape at the single chromosome level.
+<!-- we decompsed the gwRR into single chromosome features to access data which  reflects the non-random processes of crossover formation on individual chromosomes.-->
+
   The recombination landscape for single chromosome scale can be general summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
-
 We quantified the general recombination landscape because patterns in the recombination landscape are important to document because they can shed light on constraints and regulation of the processes ( could end up leading to variation in the genome wide recombination rates.
 Our results confirm (well established) previously reported patterns for sex differences in the general recombination landscape – and add by demonstrating that these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
-
 Conserved patterns of the recombination landscape of the single chromosome level – reflect constraints  At the single chromosome level – (crossover number is conserved – suggesting processes regulating the rec landscape at this scale might face similar constraints
@@
 
-For the first time in many of these strains, We report a rare, direct, evolutionary comparison of recombination rate in females and males for this short evolutionary scale.
+<!-- For the first time in many of these strains, We report a rare, direct, evolutionary comparison of recombination rate in females and males for this short evolutionary scale.
 We choose Mus musculus because of the available strains across subspecies.
@@
 Our results confirm (well established) previously reported patterns for sex differences in the general recombination landscape – and add by demonstrating that these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
+-->
 

### 2020-06-17 doc/MS_v3.Rmd
sorta almost complete intro draft

@@
 Whereas male recombination rates show evidence of rapid evolution over short timescales, female recombination rates measured in the same strains are mostly static.
-These results strongly indicate that house mouse has two, one for each sex, genome wide recombination rates which display distinct evolutionary trajectories.
+These results strongly indicate that house mouse has two, one for each sex, genome-wide recombination rates which display distinct evolutionary trajectories.
 
@@
   Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process of a diploid genome producing haploid gametes.
-In most species the meiotic program relies on crossovers to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [@CahoonLibuda2019].
-The total number of COs at this stage of meiosis is equal to the genome wide recombination rate (gwRR).
-In addition to this functional purpose, the gwRR can regulate a populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (Charlesworth 1993 cite).
+The total number of COs at this stage of meiosis is equal to the genome-wide recombination rate (gwRR).
+Meiosis (mixing of haploid gametes) is an important process for evolution.
+Meiotic recombination can regulate a populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (Charlesworth 1993 cite).
 This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
+At the individual level (fertility of a individual), the meiotic program relies on crossovers (the products of meiotic recombination), to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [@CahoonLibuda2019].
 
-<!-- ### 1. Understanding levels of variation in gwRR -->
+<!-- In most species the meiotic program relies on crossovers to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [@CahoonLibuda2019].
+### 1.
+Understanding levels of variation in gwRR -->
 
-  **Due to it's role in facilitating chromosomes disjunction, the genome wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**  Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombiantion rate [@nagaoka2012] (Inoue).
-Yet, within these thresholds the genome wide recombination rate still varies across species and individuals.
+  **Due to it's role in facilitating chromosomes disjunction, the genome-wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**  (thus clear target of evolution).
+Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombiantion rate [@nagaoka2012] (Inoue).
+Yet, within these thresholds the genome-wide recombination rate still varies across species and individuals.
 Sex is one of the most notable axes of this form of variation.
-
-<!-- (examples, Humans 1.7 female to male
-alligators female to male ratio is X, temperature dependent sex determination) -->
-
-<!-- The number of haploid pairs of chromosomes is strong predictor for genome wide recombination rate across large divergence scales [@ottoPaysuer2019; @stapley_variation_2017]. -->
-
-<!-- 2.Gap is due to Lack of comparable measures -->
-
+  There are two classes of sexual dimorphism in recombiantion rates; the first is called achiasmy and refered to where one sex recombines (Bell 1982 – Huxley 1928, Lenormand Duthiel 2005).
+The second, which is the focus of this manuscript, is the instance where both sexes recombine (Lenormand and Duthiel 2003).
 An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
-Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog,  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay], and mouse (CC cite) in which the role of sex is confounded with the contributions of genetic variation.
+Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog],  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay], and mouse [@srivastava2017CC] in which the role of sex is confounded with the contributions of genetic variation.
 
 <!-- 3.	Open Q, no first principle -->  
-
-Even under anisogamy, where gametogenesis is modified to produce distinct gametes across sexes, the meiotic program is the same.
-There are no first principles which would predict the evolution of sexually dimorphic recombination rates.
+In anisogamy ist systems where gametogenesis is modified to produce distinct gametes across sexes, the meiotic program is the same (gorelick?).
+There are no first principles which would predict the evolution of sexually dimorphic genome-wide recombination rates.
 Yet heterochiasmy is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
@@
 <!-- 4.	Additional metrics (more nuanced) -->
-  Heterochiasimy was defined as sexual dimorphism in the total counts measures of recombination (Lenormand 2003).
-It's important to note that the genome-wide recombiantion rate can be decomposed into single chromosome patterns, refered to as recombination landscapes.
-Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
+
+  The genome-wide recombiantion rate can be decomposed into single chromosome patterns, refered to as recombination landscapes.
+**(at this scale –there are also relevant features of heterochiasmy).** Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
 In a review of data from mouse, celegans and worms with meiotic chromosome morphilogy (axis-loop structure) quantified for both sexes, @CahoonLibuda2019 find that the sex with longer chromosome axis generally has higher recombination rates.
@@
 -----------
-
 <!-- ### 2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale -->
@@
   The house mouse species, (_Mus musculus_), complex arose from a recent radiation (0.5 mya to 5 mya) (@geraldes2011; @geraldes2008) providing an opportunity to interrogate natural variation at short evolutionary scales.
-This feature narrows the mutational space for identifying variants which would cause gwRR variation.
-Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway **(cite)** and previous crosses for understanding the genetic architecture of recombination rate variation [@dumont2011; @murdoch2010; @Wang2017island].
-Additionally due to their global distribution allows us to compare samples of natural diversity from a broad geographic range.
-
-  **Wild derived inbred strains generate the best comparison of females and males, besides the sex chromosomes, the mouse for each genome are almost identical**.
-Inbred strains enable one of the most direct comparison of both female and male gametogenesis and enable us to conclude that sex is a primary factor.
-
-  House mouse is well suited for single cell cytology approaches (due to large cell sizes).
-The single cell level allows integration of data at a closer connection to the molecular pathway and meiotic program.
-(Allows us to compare counts from cells at is the same meiotic (molecular pathway context)).
+This feature narrows the mutational space for identifying variants which lead to genome-wide recombination rate variation.
+Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway **(Shementi?
+Hassold?
+cite)** and previous crosses for understanding the genetic architecture of recombination rate variation [@dumont2011; @murdoch2010; @Wang2017island].
+Additionally due to their global distribution allows comparisons of from a broad geographic range that might be missed with classical inbred strains.
+Wild derived inbred strains of house mouse, enable one of the most direct comparison of female and male gametogenesis, since besides the sex chromosomes, the mouse for each genome are identical.
+Additionally, House mouse is well suited for single cell cytology approaches, enabling integration of data at a closer connection to the molecular pathway and meiotic program.
 
@@
 
----------------
-
-  (we quantify gwRR from both sexes, using house mouse inbred strains as a common genetic background).
+  In this study / paper, We quantify gwRRs from both sexes, in a panel of inbred mouse strains.
 We observed highly distinct evolutionary trajectories between males and females.
 This is due to rapid male specific evolution in the gwRR from two strains in musculus and 1 strain in molossinus.
-(this pattern of rapid evolution was not observed in females).
 We observed up to a 30% difference (translating into approximately ~7 more crossovers per cell), a surprising amount considering the short evolutionary time scale.
+This pattern of rapid evolution was not observed in females.
 
-<!-- we decompsed the gwRR into single chromosome features to access data which  reflects the non-random processes of crossover formation on individual chromosomes.-->
+<!-- we decompsed the gwRR into single chromosome features to access data which  reflects the non-random processes of crossover formation on individual chromosomes.
+The recombination landscape for single chromosome scale can be summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
+-->
 
-  The recombination landscape for single chromosome scale can be general summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
-We quantified the general recombination landscape because patterns in the recombination landscape are important to document because they can shed light on constraints and regulation of the processes ( could end up leading to variation in the genome wide recombination rates.
-Our results confirm (well established) previously reported patterns for sex differences in the general recombination landscape – and add by demonstrating that these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
-Conserved patterns of the recombination landscape of the single chromosome level – reflect constraints  At the single chromosome level – (crossover number is conserved – suggesting processes regulating the rec landscape at this scale might face similar constraints
+  In our comparisons of recombiantion landscapes across sex, our results confirm previously reported patterns for sex differences in the general recombination landscape [@sardell_sex_2020].
+Intriguingly,  these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
+
+  Our comparison of recombination landscapes framed within the rapid male specific evolution of gwRR revealed unexpected patterns - <**put something interesting here**>.
 

### 2020-06-17 doc/MS_Figures_v2.rmd
updating some Figures

@@
 
-  labs(y = "Total SC Length")+ylim(c(1000,2600 ))+ggtitle("")
+  labs(y = "Total SC Length")+ylim(c(1000,5000 ))+ggtitle("")
 
@@
 
+#Fig 4 Proportions
 ## Fig Chrm Proportion plot
@@
 chrm.class.barplot.final <- addSmallLegend(chrm.class.barplot)
-
+#this legend is too small and fuzzy in figures
 ```
@@
 ```
+
+```{r Q1.show.SC.length, echo=FALSE, warning=FALSE, eval=FALSE}
+#switch the cleanCurated_BivData
+
+#nested boxplots, pooled Bivalents
+Q1.SC.length_Nested_boxplots_Dom <- ggplot(cleanCurated_BivData[cleanCurated_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none")
+
+```
+
+```{r Q1.show.IFD.scatter, echo=FALSE, eval=FALSE, message=FALSE}
+
+#these are the scatter plots
+
+nrm.IFD_KAZ_long <- ggplot( (Q1.IFD.long %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.long <- ggplot( (real.long.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#11jan18_KAZ_m3, 26jan18_11jan18_KAZ_m3_sp1_8_rev_19
+#looked up image, measures confirmed
+
+####
+nrm.IFD_KAZ_short <- ggplot( (Q1.IFD.short %>% filter(strain =="KAZ" ) ) , aes(y = meanIFD2CO_norm, x= sex,color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+
+nrm.IFD_KAZ_BIV.short <- ggplot( (real.short.bivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2")  ) , aes(y = IFD1_PER, x= sex, color=mouse))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")
+#21may18_KAZ_m2 -- seems like it might be an outlier
+#5jun18_21may18_KAZ_m2_sp1_19.2_rev_16
+#second foci slightly smaller, kinda fainter -- but also much brighter than background
+
+#check all KAZ males
+
+nrm.IFD_KAZ_BIV.data <- ggplot( (cleanCurated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_KAZ.female_BIV.data <- ggplot( (cleanCurated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_KAZ.female_BIV.data
+
+nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_PWD.female_BIV.data
+
+nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_LEW.female_BIV.data
+
+nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_G.female_BIV.data
+
+nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank())
+
+raw.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MOLF")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+raw.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_ABS, x= mouse, color=chromosomeLength))+geom_point()+ggtitle("Interfocal Distances, ABS MSM")+theme(axis.text.x = element_blank())+ylim(c(0, 100))
+
+Fby.F.IFD_MOLF_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MOLF" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("FOce Pos by Fos Pos MOLF")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+Fby.F.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = Foci1, x=Foci2, color=chromosomeLength))+geom_point()+ggtitle("Foci Pos by Fos Pos MSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,75))+xlim(c(0,120))+theme(legend.position="none")
+
+###look at why KAZ males are weird
+nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))+theme(axis.text.x = element_blank(), legend.position = "none")
+
+nrm.IFD_LEW_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+#the difference the the Nrm IFD cut off is striking with these plots
+nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
+
+nrm.IFD_sex.comp_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, ")+facet_wrap(~sex, scales = "free")+ylim(c(0,1))+theme(legend.position="none",axis.text.x = element_blank())
+
+```
+
+```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
+
+#first draft of the IFD comparisons
+#make the mouse boxplots -- instead of geom_points
+
+#Only use 2CO bivalents
+Curated_BivData.22 <- Curated_BivData[Curated_BivData$hand.foci.count == 2,]
+Curated_BivData.22 <- Curated_BivData.22[(!is.na(Curated_BivData.22$hand.foci.count)),]
+
+P1.by.P2_plot <- ggplot(Curated_BivData.22, aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.2)+
+  ggtitle("Normalized IFD")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+xlim(c(0,1))+ylim(c(0,1))
+
+#each point is a 2CO, with the F1 and F2 position 
+#based on plot from Dourad et al (Mercier)
+
+P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
+    aes(y = Foci2, x= Foci1, color=strain))+
+  geom_jitter(alpha=.5)+
+  ggtitle("Raw IFD WSB")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#trying to move x axis to top
+#cowplot might be the easier way -- since kinda diffecult in ggplot
+#or ggdraw
+
+P1.by.P2_plot_nrm.PWD.f <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "female"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, female")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+P1.by.P2_plot_nrm.PWD.male <- ggplot(Curated_BivData.22 %>% filter(strain == "PWD") %>% filter(sex == "male"),
+    aes(y = PER_Foci_2, x= PER_Foci_1, color=strain))+
+  geom_jitter(alpha=.5)+
+#  scale_x_continuous(position = "top") +
+  scale_x_continuous(sec.axis = dup_axis(), position = "top")+
+    ggtitle("Nrm IFD PWD, male")+scale_color_manual(values=colors_of_strains)+ ylim(c(0,1))+xlim(c(0,1))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank() )
+
+#boxplots
+boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, ABS")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
+#order is out of wack, but -- good for now
+
+boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
+  geom_jitter(alpha=.2)+
+  ggtitle("Interfocal Distances, PER")+
+  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
+
+first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)
+
+first.IFD.plot_PER <- ggplot(Curated_BivData.22 %>% filter(chromosomeLength < 150), aes(y = IFD1_PER, x= strain, color=chromosomeLength))+geom_jitter()+ggtitle("Interfocal Distances, PER \ncut off of chrm lengths, 150")+facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))
+
+#+theme(legend.position="none")
+```
+
+```{r Q2.SC.long.biv_plots, echo=FALSE, eval=FALSE}
+
+#REAL.long_mouse_table
+#REAL.short_mouse_table
+
+#real.long.bivData
+#real.short.bivData
+
+shoo.long.biv <- ggplot( (real.long.bivData %>% filter(sex == "male") ), aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+mouse.shoo.long.biv <- ggplot( REAL.long_mouse_table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av.
+\nMale Long Biv SC Lengths")
+
+#cell.avlong.biv <- ggplot( cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + #ggtitle("Cell av.
+\nMale Long Biv SC Lengths")
+
+```
+
+```{r Q2_tot.sc_plot, echo=FALSE, warning=FALSE}
+
+mouse.av_totsc_male <- ggplot(data= mouse.total.skel %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
+  facet_wrap(Rec.group~subsp, scales="free")+
+  theme(axis.title.x=element_blank(),  axis.text.x=element_blank(), axis.text.y = element_text(size=9), axis.ticks.x=element_blank(), legend.position = c(0.8, 0.2) ) + ylim(c(1200,1850 ) )+labs(y="Mouse Mean Total SC")
+
+mouse.av_totsc_male <- addSmallLegend(mouse.av_totsc_male)
+
+```

### 2020-06-18 src/Result_Setup.Rmd
Updating some data files

@@
 
-load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.30.20.RData")#added SKIVE_f BiveData
+load(file = "~./MLH1repo/data/MLH1_data_setup_5.30.20.RData")#added SKIVE_f BiveData
 #load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.21.20.RData")
-#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_5.12.20.RData")
-#load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
-#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
-
-#2.2.20
-#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
-
-#load(file="~./MLH1repo/data/CleanBivData_5.12.20.RData") #CleanBivData_5.12.20.RData
-#load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
-
-#skeletonize
-#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
-#Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)
-#remove X -- this table isn't used
-
-#load(file="~./MLH1repo/data/MLH1_data_setup_1.5.20.RData")
-
-#setwd("~./MLH1repo/doc/")
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo/doc/")
-#setwd("C:/Users/April/Documents/MLH1repo")
-#load(file="data/MLH1/MLH1_data_setup.RData")
-
-#load main data file
-#load(file="C:/Users/April/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18
-#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.28.20.RData")
-
-#this MLH1 only goes up to batch16!
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
-#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
-#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
-#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
-#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
-#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
-#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
-
-#load(file="data/MetaData.RData") #this was replacing most updated data file
 
@@
 #outgroups
-#SPIC
-HetC.SPIC <- MLH1_data[(MLH1_data$strain == "SPIC")&(MLH1_data$quality < 5),]
-#unique(HetC.SPIC$mouse)
+#SPI
+HetC.SPI <- MLH1_data[(MLH1_data$strain == "SPI")&(MLH1_data$quality < 5),]
+#unique(HetC.SPI$mouse)
 #non-passing mice
-SPIC.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
-HetC.SPIC <- HetC.SPIC[ ! HetC.SPIC$mouse %in% SPIC.remove.mice, ]
+SPI.remove.mice <- c("30may18_SPIC_f1", "8jun18_SPIC_m2")
+HetC.SPI <- HetC.SPI[ ! HetC.SPI$mouse %in% SPI.remove.mice, ]
 #calq sex mean
-SPIC.mean.MLH1.obs <- ddply(.data=HetC.SPIC, 
+SPI.mean.MLH1.obs <- ddply(.data=HetC.SPI, 
                  .(category),
@@
 )
-SPIC.HETC.val <- SPIC.mean.MLH1.obs$mean.MLH1[1] / SPIC.mean.MLH1.obs$mean.MLH1[2]
+SPI.HETC.val <- SPI.mean.MLH1.obs$mean.MLH1[1] / SPI.mean.MLH1.obs$mean.MLH1[2]
 
@@
 #remove male only data
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
+remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPI")
 
@@
 write.table(MLH1.MMlmer_results, "~./MLH1repo/MLH1_M1_glm.csv", sep=",",row.names = TRUE)
+#Write to excel?
+
 ```
@@
 #
-totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
+totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1" & strain !="SPI" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")
 

### 2020-06-19 doc/MS_Figures.rmd
working on figures

--- 2020-06-17 doc/MS_Figures_v2.rmd
@@
 
-#load(file="~./MLH1repo/data/CleanBivData_6.3.20.RData")
-
-#
-#load(file="~./MLH1repo/data/total_SC_30.12.19.RData")
-
-#skeletonize - this is the total sc file
-#SC.skel = read.csv("~./MLH1repo/data/SCskel_output_feb20.csv", header = TRUE, strip.white = TRUE)
-
 #load functions
@@
 
-DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter() +
+#aes(x=strain, y=total.count, color=stage
+
+DMC1.boxplot <- ggplot( aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
+
+  #stat_boxplot( aes(x=strain, y=total.count, color=stage), 
+ #   geom='errorbar', linetype=1, width=0.5)+
+
+  geom_boxplot( aes(x=strain, y=total.count),outlier.shape = NA, alpha=.5 )+
+
+  geom_jitter() +
 

### 2020-06-19 doc/MS_Figures_v2.rmd
working on figures

@@
 
+#remoe outliers, with outlier shape
+
 DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
-  geom_boxplot(alpha=.5) +geom_jitter() +
-
-    labs(y = "DMC1 count per Cell")+
+
+  geom_boxplot(alpha=.5, outlier.shape = NA) +
+
+  #Yeah, I can dodge the points!!
+  geom_point(pch = 19, position = position_jitterdodge()) +
+  #geom_jitter() +
+
+    labs(y = "DMC1 Count per Cell")+
 
@@
 #"#0072B2" darker blue
-
 ```

### 2020-06-19 data/SetupData.Rmd
updating data files after removing outlier, fix fig3

--- 2020-06-12 src/SetupData.Rmd
@@
 
-opts_knit$set(root.dir = "~./MLH1repo/")
-setwd("~./MLH1repo/")
+opts_knit$set(root.dir = "~./MLH1repo/data/")
+setwd("~./MLH1repo/data/")
 
@@
 #read.csv was breaking things
-MLH1_data = read.xlsx2("~./MLH1repo/data/AnonData_formated.xlsx", header=TRUE, 1)
+MLH1_data = read.xlsx2("~./MLH1repo/data/MLH1.AnonData_formated.xlsx", header=TRUE, 1)
 #remove all X's from MLH1, remove all old mice?
@@
 
-#total SC 
-#new total_SC == will have all of the Batches
-
 Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF_5.31.20.xlsx", header=TRUE, 1) #this new total_SC has the additional batches and outliers have been removed
 
-#Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)
-
-#original Curate_bivata + missing SKIVE female
+#this file is missing
 Curated_BivData <-  read.csv2("~./MLH1repo/data/Curated_BivData_5.30.20.csv", sep=" ", header=TRUE)
@@
 
-```{r MLH1.DF.setup, warning=FALSE, echo=FALSE}
+```{r MLH1.DF.setup, warning=FALSE, echo=FALSE, eval=FALSE}
+
+#duplicate file names and rows have been removed
+#they can be found in older files for accuracy checks
+
+#check that these should all be redundant
 
@@
 
-`r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.
-`r length(original_DF_duplicates$fileName)` rows (images) have been counted across mulitple batches.
-These are removed in the MLH1_data DF.
-
 ```{r clean_MLH1_data, echo=FALSE}
@@
 #https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/
+
+original_DF = MLH1_data
+
 #use distinct() in dlpyr()
@@
 
-`r length(original_DF$fileName[ grepl("X", original_DF$X)] )` images (rows) have X's, they are removed from the main dataset.
-
 ```{r misc.setup, echo=FALSE}
@@
 
-                    'SPRET'='gold','SPIC'= 'goldenrod1', #browns? color of mounds?
+                    'SPRET'='gold','SPI'= 'goldenrod1', #browns? color of mounds?
                     'CAROLI'='gold3',
@@
 Make the raw original DF and main working DF with bad cells exlcuede.
-The original dataset had `r length(original_DF_duplicates$fileName)` duplicated file names.
+The original dataset had a number of duplicated file names.
 These were removed and can be used to quantify repeatability or repair.
@@
 #write excel file
-#write.xlsx2(missing.SKIVE.female, "Curated_missing.SKIVE.female.xlsx", sheetName = "Sheet1", 
+#write.xlsx2(XX, "XX.xlsx", sheetName = "Sheet1", 
 #            col.names = TRUE, row.names = FALSE, append = FALSE)
@@
 #clean
-Total.SC_DF$notes <- as.character(Total.SC_DF$notes)
+#Total.SC_DF$notes <- as.character(Total.SC_DF$notes)
 Total.SC_DF$REDO.crop <- as.character(Total.SC_DF$REDO.crop)
@@
 ```{r Mouse.total.sc.table, echo=FALSE}
+
+#20feb16_G_f2
 
@@
 
-save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_5.30.20.RData")
+save.image("~./MLH1repo/data/MLH1_data_setup_5.30.20.RData")
 

### 2020-06-19 src/Result_Setup.Rmd
updating data files after removing outlier, fix fig3

@@
 library(reshape2)
+library(writexl)
+library(xlsx)
 #library(patchwork)
@@
 #MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
+#AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
 
@@
 ```{r MLH1.FEMALE.TABLE, echo=FALSE, eval=TRUE}
-
 female_mouse.avs_for.analysis <-  Mouse.Table_HQ %>% filter(sex== "female") %>% filter(species == "M.musculus")
@@
 
-127 mice, 8 strains (correct), 0 MOLF females!
+127 mice, 8 strains (correct), 0 MOLF females! -- this was fixed
 
@@
 
+#goal get all the results below to a .tex format
+#write_xlsx(MLH1.MMlmer_results, "~./MLH1repo/MLH1_M1_glm.xlsx", col_names = TRUE)
+
+write.xlsx(MLH1.MMlmer_results, "MLH1_M1_glm.xlsx", col.names = TRUE)
 ```
@@
 
-write.table(summ.MLH1.M2.glm.strain.fixed.full$coefficients , "~./MLH1repo/MLH1_M2_glm.csv", sep=",",row.names = TRUE)
+write.table(summ.MLH1.M2.glm.strain.fixed.full$coefficients, "~./MLH1repo/MLH1_M2_glm.csv", sep=",",row.names = TRUE)
+
+summary.glm(MLH1.M2.glm.strain.fixed.full)$coefficients
+
+write.xlsx(as.data.frame(summary.glm(MLH1.M2.glm.strain.fixed.full)$coefficients), "MLH1_M2_glm.xlsx", col.names = TRUE, row.names = TRUE)
+
 ```
@@
 ```{r MLH1.glm.M3, echo=FALSE, eval=TRUE}
-
 MLH1.M3.strain.fixed <- glm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ )
-
 summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)
@@
 
-write.table(summ.MLH1.M3.strain.fixed$coefficients , "~./MLH1repo/MLH1_M3_glm.csv", sep=",",row.names = TRUE)
+write.table(summ.MLH1.M3.strain.fixed$coefficients, "~./MLH1repo/MLH1_M3_glm.csv", sep=",",row.names = TRUE)
+
+#this table wont write!!
+write.xlsx(as.data.frame(summary(MLH1.M3.strain.fixed)$coefficients), "MLH1_M3_glm.xlsx", col.names = TRUE, row.names = TRUE)
 ```
@@
 
+write.xlsx(as.data.frame( summry.female.strain.effects$coefficients), "MLH1_M4_female_glm.xlsx", col.names = TRUE, row.names = TRUE)
+
 MLH1.female.spc.glm.M5 <- glm(mean_co ~ strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
@@
 write.table(sum_MLH1.female.spc.glm.M5$coefficients, "~./MLH1repo/MLH1_M5_female_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_MLH1.female.spc.glm.M5$coefficients), "MLH1_M5_female_glm.xlsx", col.names = TRUE, row.names = TRUE)
 ```
@@
 
-write.table(sum_MLH1.male.glm.M4$coefficients , "~./MLH1repo/MLH1_M4_male_glm.csv", sep=",",row.names = TRUE)
-
-#M4.2
-MLH1.male.glm.M4.2 <- glm(mean_co ~ strain, data= (male_mouse.avs_for.analysis  %>% filter(subsp != "Cast") ))
-sum_MLH1.male.glm.M4.2 <- summary(MLH1.male.glm.M4.2)
+write.table(sum_MLH1.male.glm.M4$coefficients, "~./MLH1repo/MLH1_M4_male_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame(sum_MLH1.male.glm.M4$coefficients), "MLH1_M4_male_glm.xlsx", col.names = TRUE, row.names = TRUE)
+
+#write_xlsx(sum_MLH1.female.spc.glm.M5$coefficients, #"MLH1_M5_female_glm.xlsx", col_names = TRUE)
+
+#M5
+MLH1.male.glm.M5 <- glm(mean_co ~ strain, data= (male_mouse.avs_for.analysis  %>% filter(subsp != "Cast") ))
+sum_MLH1.male.glm.M5 <- summary(MLH1.male.glm.M5)
 
@@
 
-write.table(sum_MLH1.male.glm.M4.2$coefficients , "~./MLH1repo/MLH1_M5_male_glm.csv", sep=",",row.names = TRUE)
-
-```
-
-#MLH1 variance models
-
-**THIS IS BEING MOVED TO CHAPTER 3**
-
-M1 - subsp, sex, (1 | strain)  (lmer)
-M2 - subsp, sex, strain  (glm / lm?)
-M3 - sex, strain
-M4 - strain (sex specific)
-
-```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
-MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus"))   )
-MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus")) )
-MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus")) )
-
-#summary(MLH1var_M1)
-LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value
-
-LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
-sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)
-
-#CV
-MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
-MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )
-
-LmerMLH1_CV_int <- drop1(MLH1cV_M1, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_CV_sex <- anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_subsp <- anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_CV_strain <- exactRLRT(MLH1cV_M1)$p.value
-
-LmerMLH1_CV_results <- data.frame(sub.sp = LmerMLH1_CV_subsp,
-sex.results =  LmerMLH1_CV_sex, interaction.pval = LmerMLH1_CV_int,  rand.pval = LmerMLH1_CV_strain)
-
-#Q12
-Q12_mouse_table_MLH1$subsp <- factor(Q12_mouse_table_MLH1$subsp, ordered = FALSE)
-Q12_mouse_table_MLH1$sex <- factor(Q12_mouse_table_MLH1$sex, ordered = FALSE)
-Q12_mouse_table_MLH1$strain <- factor(Q12_mouse_table_MLH1$strain, ordered = FALSE)
-Q12_mouse_table_MLH1$species <- factor(Q12_mouse_table_MLH1$species, ordered = FALSE)
-
-#USE Q12 (Q123) table
-#Q12_mouse_table_w.Ages
-#Q12 var
-MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
-MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
-MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
-
-LmerMLH1_M1_Q12VAR_int <- drop1(MLH1var_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_VAR_sex <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_subsp <- anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_VAR_strain <- exactRLRT(MLH1var_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_VAR_subsp,
-sex.results =  LmerMLH1_M1_Q12_VAR_sex, interaction.pval = LmerMLH1_M1_Q12VAR_int,  rand.pval = LmerMLH1_M1_Q12_VAR_strain)
-
-#Q12 cv
-MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
-MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )
-
-LmerMLH1_M1_Q12_CV_int <- drop1(MLH1cv_M1_Q12, test="Chisq")$`Pr(Chi)`[2]
-LmerMLH1_M1_Q12_CV_sex <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_subsp <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)$`Pr(>Chisq)`[2]
-LmerMLH1_M1_Q12_CV_strain <- exactRLRT(MLH1cv_M1_Q12)$p.value
-
-LmerMLH1_M1_Q12_VAR_results <- data.frame(sub.sp = LmerMLH1_M1_Q12_CV_subsp,
-sex.results =  LmerMLH1_M1_Q12_CV_sex, interaction.pval = LmerMLH1_M1_Q12_CV_int,  rand.pval = LmerMLH1_M1_Q12_CV_strain)
-```
-
-```{r M2.MLH1.var, echo=FALSE, include=FALSE}
-
-MLH1var_M2 <- glm( var ~ subsp * sex * strain,  data= ( DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus") )  )
-MLH1var_M2_sum <- summary(MLH1var_M2)
-
-MLH1var_M2_sum_coE <- as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.
-
-MLH1cv_M2 <- glm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M2)
-#anova(MLH1cv_M2)
-#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))
-
-#Q12 var
-MLH1var_M2_Q12 <- glm( var ~ subsp*sex*strain,  data= (Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
-sum.MLH1var_M2_Q12 <- summary(MLH1var_M2_Q12)
-sum.MLH1var_M2_Q12.co_eff <- as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))
-
-#Q12 cv
-MLH1cv_M2_Q12 <- glm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )
-#summary(MLH1cv_M2_Q12)
-#anova(MLH1cv_M2_Q12)#sex large F
-#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
-```
-
-```{r M3.MLH1.var, echo=FALSE}
-
-glmMLH1Var_M3 <- glm( var ~ sex * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(species == "M.musculus") ) )
-
-glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
-
-glmMLH1Var_M3_sum.Coef <- round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)
-
-MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
-#summary(MLH1cv_M4)
-#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)
-
-### Q12 var
-MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=(Q12_mouse_table_MLH1 %>% filter(species == "M.musculus")) )
-MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)
-
-#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
-#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)
-
-#Q12 cv
-MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
-#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
-```
-
-```{r MLH1.var.M4, echo=FALSE, eval=F}
-#male
-glmMLH1Var_M4.male <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "male") ) )
-
-#female
-glmMLH1Var_M4.female <- glm( var ~ subsp * strain,  data= (DF.HetC.MixedModel.HQ %>% filter(sex == "female") ) )
-sum.glmMLH1Var_M4.female <- summary(glmMLH1Var_M4.female)
-
-glmMLH1Var_M4.femaleQ12 <- glm( var ~ subsp * strain,  data= (Q12_mouse_table_w.Ages %>% filter(sex == "female") ) )
-
-```
-
-```{r FEMALE.VAR.MODELS, echo=FALSE, eval=F}
-## within mouse variance models 
-lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
-summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)
-
-lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
-lmVAR.female.strain_sum <- summary(lmVAR.female.strain)
-
-#round( #round isn't working for this sumary
-#lmVAR.female.strain_sum$coefficients[,c(1,4)]
-#Q12
-Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")
-
-lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
-lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)
-
-lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
-lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD
-
-lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
-lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)
-
-#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
-#LEW sig
-
-```
-
-```{r MALE.SPECIFIC.variance.model, echo=FALSE, eval=F}
-## within mouse variance models 
-lmVAR.male.strain.subsp <- glm(var ~ subsp * strain, data= just.male_mouse.av)
-summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
-#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)
-
-#all p values kinda high, MSM and TOM are slightly sig
-lmVAR.male.strain <- glm(var ~ strain, data= just.male_mouse.av)
-lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
-#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
-#non sig
-
-lmcv.male.strain <- glm(cV ~ strain, data= just.male_mouse.av)
-lmcv.male.strain_sum <- summary(lmcv.male.strain)
-#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
-#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant
-
-#Q12
-Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")
-
-lm.m1.Q12_male <- glm(var ~ subsp * strain, data= Q12_male_table)
-lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
-#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
-#NS
-
-lm.m2.Q12_male <- glm(var ~ strain, data= Q12_male_table)
-lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
-#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
-#NS
-
+write.table(sum_MLH1.male.glm.M5$coefficients, "~./MLH1repo/MLH1_M5_male_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_MLH1.male.glm.M5$coefficients), "MLH1_M5_male_glm.xlsx", col.names = TRUE, row.names = TRUE)
+
+#write_xlsx(sum_MLH1.male.glm.M4.2$coefficients, "~./MLH1repo/MLH1_M5_male_glm.xlsx", col_names = TRUE)
 ```
@@
 
+#write_xlsx(Q1.lmer_totSC_M1_results, "~./MLH1repo/TotalSC_Q1_M1.xlsx", col_names = TRUE)
+
+write.xlsx(Q1.lmer_totSC_M1_results, "TotalSC_Q1_M1.xlsx", col.names = TRUE)
 ```
@@
 
-#sex most sig
-#subsp
-
+#sex most sig  #subsp
 #straing G, LEW
@@
 
+write.xlsx(as.data.frame(summary.glm(Q1.glm_totSC_M2)$coefficients), "TotalSC_Q1_M2.xlsx", col.names = TRUE, row.names = TRUE)
+
+#write_xlsx(sum.Q1.glm_totSC_M2$coefficients, "~./MLH1repo/TotalSC_Q1_M2.xlsx", col_names = TRUE)
 ```
@@
 write.table(sum.Q1.glm_totSC_M3$coefficients, "~./MLH1repo/TotalSC_Q1_M3.csv", sep=",",row.names = TRUE)
+
+#not working currently.. might be same 'ol error
+#write.xlsx(  as.data.frame( summary(sum.Q1.glm_totSC_M3)$coefficients ), "TotalSC_Q1_M3.xlsx", col.names = TRUE, row.names = TRUE)
+
+#write_xlsx(sum.Q1.glm_totSC_M3$coefficients, "~./MLH1repo/TotalSC_Q1_M3.xlsx", col_names = TRUE)
 ```
@@
 write.table(Q1lmer.MM_SC_short_results, "~./MLH1repo/ShortSC_Q1_M1.csv", sep=",",row.names = TRUE)
+
+#write_xlsx(Q1lmer.MM_SC_short_results, "~./MLH1repo/ShortSC_Q1_M1.xlsx", col_names = TRUE)
+
+write.xlsx(Q1lmer.MM_SC_short_results, "ShortSC_Q1_M1.xlsx", col.names = TRUE)
+
 ```
@@
 
+write_xlsx(Q1.lmer_short_M1_results, "~./MLH1repo/ShortSC_Q1_M1.xlsx", col_names = TRUE)
 ```
@@
 write.table(summ.Q1.glm_short.SC_M2$coefficients, "~./MLH1repo/ShortSC_Q1_M2.csv", sep=",",row.names = TRUE)
+
+#write_xlsx(summ.Q1.glm_short.SC_M2$coefficients, "~./MLH1repo/ShortSC_Q1_M2.xlsx", col_names = TRUE)
+
+write.xlsx( as.data.frame( summ.Q1.glm_short.SC_M2$coefficients) , "ShortSC_Q1_M2.xlsx", col.names = TRUE, row.names = TRUE)
 ```
@@
 write.table(summ.Q1.glm_short.SC_M3$coefficients, "~./MLH1repo/ShortSC_Q1_M3.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( summ.Q1.glm_short.SC_M3$coefficients), "ShortSC_Q1_M3.xlsx", col.names = TRUE, row.names = TRUE)
+
 ```
@@
 
+#write_xlsx(lmer.Q1.nrm1CO_M1_results, "~./MLH1repo/Nrm1CO_Pos_Q1_M1.xlsx", col_names = TRUE)
+
+write.xlsx(lmer.Q1.nrm1CO_M1_results, "Nrm1CO_Pos_Q1_M1.xlsx", col.names = TRUE, row.names = TRUE)
 ```
@@
 
+write.xlsx(summry.Q1.glm_F1pos_M2$coefficients, "Nrm1CO_Pos_Q1_M2.xlsx", col.names = TRUE)
+
+#write_xlsx(summry.Q1.glm_F1pos_M2$coefficients, "~./MLH1repo/Nrm1CO_Pos_Q1_M2.xlsx", col_names = TRUE)
+
 #DOM
@@
 
+write.xlsx(summry.Q1.glm_1CO.pos_M3$coefficients, "Nrm1CO_Pos_Q1_M3.xlsx", col.names = TRUE)
+
+#write_xlsx(summry.Q1.glm_1CO.pos_M3$coefficients, "~./MLH1repo/Nrm1CO_Pos_Q1_M3.xlsx", col_names = TRUE)
 ```
@@
 
+write_xlsx(Q1.IFD.ABS_M1_lmer.results, "~./MLH1repo/ABS.IFD_Q1_M1.xlsx", col_names = TRUE)
+
 #M1_IFD_2CO
@@
 write.table(Q1.IFD.PER_M1_lmer.results, "~./MLH1repo/PER.IFD_Q1_M1.csv", sep=",",row.names = TRUE)
+
+write_xlsx(Q1.IFD.PER_M1_lmer.results, "~./MLH1repo/PER.IFD_Q1_M1.xlsx", col_names = TRUE)
 
@@
 
+write.xlsx(summry.glm.Q1_IFD.ABS.M2$coefficients, "ABS.IFD_Q1_M2.xlsx", col.names = TRUE)
+
+#write_xlsx(summry.glm.Q1_IFD.ABS.M2$coefficients, "~./MLH1repo/ABS.IFD_Q1_M2.xlsx", col_names = TRUE)
+
 #PER
@@
 write.table(summry.glm.Q1_IFD.PER.M2$coefficients, "~./MLH1repo/PER.IFD_Q1_M2.csv", sep=",",row.names = TRUE)
+
+write.xlsx(summry.glm.Q1_IFD.PER.M2$coefficients, "PER.IFD_Q1_M2.xlsx", col.names = TRUE)
+
+#write_xlsx(summry.glm.Q1_IFD.PER.M2$coefficients, "~./MLH1repo/PER.IFD_Q1_M2.xlsx", col_names = TRUE)
 ```
@@
 write.table(summry.glm.Q1_IFD.ABS.M3$coefficients, "~./MLH1repo/ABS.IFD_Q1_M3.csv", sep=",",row.names = TRUE)
+
+write.xlsx(summry.glm.Q1_IFD.ABS.M3$coefficients, "ABS.IFD_Q1_M3.xlsx", col.names = TRUE)
+
+#write_xlsx(summry.glm.Q1_IFD.ABS.M3$coefficients, "~./MLH1repo/ABS.IFD_Q1_M3.xlsx", col_names = TRUE)
 
@@
 
+write.xlsx(summry.glm.Q1_IFD.PER.M3$coefficients, "PER.IFD_Q1_M3.xlsx", col.names = TRUE)
+
+#write_xlsx(summry.glm.Q1_IFD.PER.M3$coefficients, "~./MLH1repo/PER.IFD_Q1_M3.xlsx", col_names = TRUE)
+
 ```
@@
 summry.glm.Q1_IFD.PER.M5.male  <- summary(glm.Q1_IFD.PER.M5.male)
-
 ```

### 2020-06-19 src/setupData_MLH1repo.R

@@ -1,155 +1,0 @@
-##R script for making R.data files 
-## read in data files and adjust with additional columns
-## input: csv file of anon MLH1 counts
-## output: .rdata saved workspace with dataframe and other stat tables (save tables)
-## df of MLH1 should go from 12cols to 17cols
-#import MLH1 measures across data sets; AP, BD's and Cast female Koehler
-# 1. MLH1_data, tables: AP strains
-#updated meta.data.file
-
-#library(plyr)
-library(dplyr)
-library(raster)#for cV
-library(ggplot2)
-
-########################
-# Setup main dataframe #
-########################
-
-#TODO: remove observations without quality scores (why?)
-#TODO integrate newest batch
-
-#images from slides
-#12jan17_4jan17_LEW_f1_sp1_34_rev, 8feb17_4jan17_LEW_f1_sp1_3_rev, potential outliers
-
-#Write a make file that will merge all of the batch files and record the batch
-#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
-setwd("~./MLH1repo/")
-
-#forgot why I'm reading in this metadata file
-meta.data = read.csv("data/clean.Meta.Data.txt",sep = "\t", header=TRUE )
-
-MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE ) #7.30.19, 4382
-
-
-source("src/CommonFunc_MLH1repo.R")
-original_DF = MLH1_data
-original_DF$fileName <- (original_DF$Original.Name)
-original_DF <- add_mouse(original_DF)
-#original_length <- length(MLH1_data$Original.Name)
-
-#check that all batches are included
-#table(original_DF$Batch) #if I want this I should make this Rmd
-
-
-MLH1_data$Original.Name <- as.character(MLH1_data$Original.Name)
-MLH1_data$fileName <- (MLH1_data$Original.Name)
-
-#these introduce NAs
-MLH1_data$n <- as.character(MLH1_data$n)
-MLH1_data$nMLH1.foci<- as.character(MLH1_data$nMLH1.foci)
-MLH1_data$quality<- as.character(MLH1_data$quality)
-
-MLH1_data$n <- as.numeric(MLH1_data$n)
-MLH1_data$nMLH1.foci<- as.numeric(MLH1_data$nMLH1.foci)
[…]

### 2020-06-19 doc/MS_Figures_v2.rmd
new figure3, updated data files

@@
 
-# Fig 3
-
 ```{r Q1.SC.show.short.biv, echo=FALSE, eval=TRUE}
@@
 
-Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+
+Q1.SC.short.mouse_scatter_boxplot <- ggplot( Short.biv_mouse_table, aes(y = mean_SC, x = sex, color=strain)) +
+
+     geom_boxplot(alpha=.5, outlier.shape = NA) +
+  #Yeah, I can dodge the points!!
+  geom_point(pch = 19, position = position_jitterdodge()) +
+
+  scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("")+  
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
 
-  labs(y = "Short SC Length")
+  labs(y = "Short Bivalent Length")
 
@@
 # fig.width = 4, fig.height=2.5
+
+#20feb16_G_f2
 
@@
     facet_wrap(~subsp, scales="free", ncol = 3) + 
-
-  geom_jitter()+ geom_boxplot(alpha=.2)+ 
+    geom_boxplot(alpha=.5, outlier.shape = NA) +
+  #Yeah, I can dodge the points!!
+  geom_point(pch = 19, position = position_jitterdodge()) +
+
   #geom_hline(data =totSC.strain.av.table, aes(yintercept=mean.skel) )+
@@
 legend.position="none",axis.text.y = element_text(size=9), # axis.text.y = element_blank()
-axis.title.y = element_text(size = 10) ) +
-
-  labs(y = "Total SC Length")+ylim(c(1000,5000 ))+ggtitle("")
-
-ggsave("~/MLH1repo/doc/Figure3_Tot.SC_mouse.av_ALL.png")
-
+axis.title.y = element_text(size = 10) ) + labs(y = "Total SC Length")+ylim(c(1000,2700 ))+ggtitle("")
+
+#ggsave("~/MLH1repo/doc/Figure3_Tot.SC_mouse.av_ALL.png")
 #ggsave("~/MLH1repo/doc/Figure3_Q1.SC.short.mouse_scatter_boxplot.png")

### 2020-06-22 doc/MS_v3.Rmd
fixed a bug from Common functions

@@
 date: '`r Sys.Date()`'
-author: "April L. Peterson, Bret Payseur"
+author: "April L. Peterson, Bret A. Payseur"
 output:
-  word_document:
+  html_document:
   pdf_document:
@@
       in_header: header.tex
-  html_document:
+  word_document:
 always_allow_html: yes
@@
 -----------
+
 <!-- ### 2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale -->
@@
                      "CAST",
-                         "SPRET","SPIC" ), order=T )
+                         "SPRET","SPI" ), order=T )
 
@@
 
-<new 1/n>
-
 We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice (*M.
@@
 
-<new 2/n>
-
 Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
@@
 In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23) of foci (Figure 1C).
-Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = 0.08 - `r sperman.corr.FM`; p = 0.84 -- `r sperman.corr.FM.pval$p.value`) across the set of strains.
-
-<new 3/n>
+Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = `r sperman.corr.FM`; p = `r sperman.corr.FM.pval$p.value`) across the set of strains.
 
 To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (Table 2; detailed results available in Supplemental Table X).
-Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < 10-4 -- `r 1/10000`; sex: p = 6.2x10-4 -- `r MLH1.MMlmer_results$sex.results`; subspecies: p = 1.1x10-3 -- `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = 2x10-4 -- `r MLH1.MMlmer_results$interaction.pval`).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`; sex: p = `r MLH1.MMlmer_results$sex.results`; subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
 
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects **(M2; p = 0.09 -- )**, highlighting strain and sex as salient variables.
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects **(M2; Musc p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, Mol p =  `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`)**, highlighting strain and sex as salient variables.
 Two strains exhibit particularly strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
@@
 
-<new 1/n>
 We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68--`r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: 0.56 --  `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; t-test; p = 3.84^{-23} -- `r Q1_1CO.pos_t.test.FULL$p.value`).
-Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = 7.79^{-26 --  `r lmer.Q1.nrm1CO_M1_results$sex.results`}; M2: p = 5.39^{-8} --  `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = 5.39^{-8} --  `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68--`r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`}; M2: p =  `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p =  `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 
-<new 2/n>
-Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = 0.33 -- `r Q1.IFD.PER_t.test_KAZ$p.value`).
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
 Examination of IFD~norm~ distributions indicates that females are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Models treating IFDnorm as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = 3.11^{-12} -- `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = 0.01  --  `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = 0.01  --  `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+Models treating IFDnorm as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
 In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) across the full set of strains, whether they are considered separately or together.
@@
 
-Write results for M4 and M5 here
-total SC M4 results
+Write results for M4 
+total SC M4 results -- many strain specific effects for the M4 and M5 models, but two subspecies are the largest, but MSM strain not significant, and SKIVE is in oposite direction.
+
+and M5 here
 
@@
 molossinus*, p = 21^{-6} -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
-In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05), but they are not consistent across models.
+In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) <what are these referencing?>, but they are not consistent across models.
 Collectively, these results reveal that evolution of SC length is not strongly associated with evolution of genome-wide recombination rate in house mice.
@@
 
-<new 1/n>
 By comparing recombination rates in females and males from the same diverse set of genetic backgrounds, we isolated sex as a primary factor in the evolution of this fundamental meiotic trait.
@@
 
-<new 2/n>
 At the genetic level, the sex-specific evolution we documented indicates that some mutations responsible for divergence in recombination rate have dissimilar phenotypic effects in females and males.
@@
 
-<new 3/n>
 Another implication of our results is that the connection between recombination rate and fitness differs between males and females.
@@
 
-<new 4/n>
 Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling; @sardell_sex_2020].
@@
 
-<new 5/n>
 We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
@@
 
-<new 5/n>
 Our SAC model is consistent with other features of our data.
@@
 
-<new 6/n>
 We propose that the SAC model also can explain the correlated evolution of stronger crossover interference and higher genome-wide recombination rate in male house mice.
@@
 
-<new 7/n>
 Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
 The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle (@ma2015_cattle) and differences between populations of white-footed mice [@peterson2019].
-In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates (@segura2013, @ottoPaysuer2019).
+In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@segura2013; @ottoPaysuer2019].
 Collectively, these patterns suggest that inferences about the evolutionary dynamics of interference depend on the timescale under consideration.
 
-<new 8/n>
 Our findings further reveal that evolution of the genome-wide recombination rate does not require major changes in the degree of chromatin compaction.
 Female house mice consistently show longer SCs, even in strains with more recombination in males.
-Studies in mice [@lynn2002, @petkov2007] and humans [@gruhn2013, @tease2004] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+Studies in mice [@lynn2002; @petkov2007] and humans [@gruhn2013; @tease2004] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
 Some authors have suggested that conserved sex differences in crossover positioning (more uniform placement in females) and interference strength (stronger interference in males) could be due to looser chromatin packing of the meiotic chromosome structure in females [@haenel2018; @petkov2007].
@@
 
-<new 9/n>
 Our conclusions are accompanied by several caveats.
 First, MLH1 foci only identify interfering crossovers [@holloway2008mus81].
-Although most crossovers belong to this class **(REF)**, our approach likely underestimated genome-wide recombination rates.
+Although most crossovers belong to this class [@holloway2008mus81], our approach likely underestimated genome-wide recombination rates.
 Evolution of the number of non-interfering crossovers is a topic worth examining.

### 2020-06-22 src/Result_Setup.Rmd
fixed a bug from Common functions

@@
 
-summary.glm(MLH1.M2.glm.strain.fixed.full)$coefficients
+#summary.glm(MLH1.M2.glm.strain.fixed.full)$coefficients
 
@@
 #mouse.total.skel
-Q1.lmer_totSC_M1 <- lmer(mean.skel ~ subsp * sex + (1|strain), data= ( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
-
-Q1.lmer_totSC_M1_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ))
-
-Q1.lmer_totSC_M1_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ))
+Q1.lmer_totSC_M1 <- lmer(mean.skel ~ subsp * sex + (1|strain), data= ( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ))  )
+
+Q1.lmer_totSC_M1_R.sex <- lmer(mean.skel ~ subsp + (1|strain), data=( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ))
+
+Q1.lmer_totSC_M1_R.subsp <- lmer(mean.skel ~ sex + (1|strain), data=( mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
 
@@
 
-Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=  (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=  (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )  )
 
@@
 
-Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
+mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
+mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)
+
+Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
 
@@
 
+write.xlsx(as.data.frame(summary.glm(Q1.glm_totSC_M3)$coefficients), "TotalSC_Q1_M3.xlsx", col.names = TRUE, row.names = TRUE)
+
 #write_xlsx(sum.Q1.glm_totSC_M3$coefficients, "~./MLH1repo/TotalSC_Q1_M3.xlsx", col_names = TRUE)
-```
-
-```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-
-lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Short.biv_mouse_table, REML = TRUE) 
-lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Short.biv_mouse_table, REML = TRUE)
-
-Q1lmer.MM_SC_short_interaction  <-drop1(lmer.Q1_SCLength_short_M1, test="Chisq" )$`Pr(Chi)`[2]
-Q1lmer.MM_SC_short_sex  <-anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.sex)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_subp  <- anova(lmer.Q1_SCLength_short_M1, lmer.Q1_SCLength_short_M1_R.subsp)$`Pr(>Chisq)`[2]
-Q1lmer.MM_SC_short_strain  <- exactRLRT(lmer.Q1_SCLength_short_M1)$p.value
-
-#as.data.frame(round(summary(lmer.SCLength_all_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
-
-Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
-sex.results =  Q1lmer.MM_SC_short_sex, 
-               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
-Q1lmer.MM_SC_short_strain)
-
-#Q1lmer.MM_SC_short_results
-write.table(Q1lmer.MM_SC_short_results, "~./MLH1repo/ShortSC_Q1_M1.csv", sep=",",row.names = TRUE)
-
-#write_xlsx(Q1lmer.MM_SC_short_results, "~./MLH1repo/ShortSC_Q1_M1.xlsx", col_names = TRUE)
-
-write.xlsx(Q1lmer.MM_SC_short_results, "ShortSC_Q1_M1.xlsx", col.names = TRUE)
-
 ```
@@
 
-Q1.lmer_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
-
-Q1.lmer_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data= ( Short.biv_mouse_table %>%  filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
-
-Q1.lmer_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+Q1.lmer_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
+
+Q1.lmer_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data= ( Short.biv_mouse_table %>%  filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
+
+Q1.lmer_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" )  %>% filter(strain !=  "TOM" ) )  )
 
@@
 
-write_xlsx(Q1.lmer_short_M1_results, "~./MLH1repo/ShortSC_Q1_M1.xlsx", col_names = TRUE)
+#write_xlsx(Q1.lmer_short_M1_results, "~./MLH1repo/ShortSC_Q1_M1.xlsx", col_names = TRUE)
+write.xlsx(Q1.lmer_short_M1_results, "ShortSC_Q1_M1.xlsx", col.names = TRUE)
+
 ```
@@
 
-Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
 
@@
 ```{r Q1.glm_short.SC_M3, echo=FALSE, message=FALSE}
-Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) )  )
+
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+
+Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data= ( Short.biv_mouse_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )  )
 
@@
 ```{r Q1.CO1pos_lmer.M1, echo=FALSE, message=FALSE}
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
+mouse.level.BivData_table$species <- factor(mouse.level.BivData_table$species, ordered = FALSE)
 
 lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), 
-                          data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
-
-lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
-
-lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+                          data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
+
+lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" )%>% filter(strain !=  "TOM" ) ) )
+
+lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" )) )
 
@@
 
-Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ) )
+Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
 
@@
 ```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=TRUE}
-
-Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) ))
-
-#summary(Q1.glm_1CO.pos_M3)
+#in case factors aren't ordered
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
+
+Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ))
 
@@
 
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
+mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)
+mouse.level.BivData_table$species <- factor(mouse.level.BivData_table$species, ordered = FALSE)
+
+#No MOLF female for curated biv data
+
 #M1_IFD_2CO_ABS
-lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") )) 
-lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
-lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                           )
+
+lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
+
+lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) ) )
 
@@
 
-write_xlsx(Q1.IFD.ABS_M1_lmer.results, "~./MLH1repo/ABS.IFD_Q1_M1.xlsx", col_names = TRUE)
+#write_xlsx(Q1.IFD.ABS_M1_lmer.results, "~./MLH1repo/ABS.IFD_Q1_M1.xlsx", col_names = TRUE)
+
+write.xlsx(Q1.IFD.ABS_M1_lmer.results, "ABS.IFD_Q1_M1.xlsx", col.names = TRUE)
+
+oo <- ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
 
 #M1_IFD_2CO
-lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) ) 
-lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") )  )
-lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+lmer.IFD.PER_MM.M1 <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=
+                           ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                           )
+
+lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )  )
+
+lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                                        )
 
@@
 
-write_xlsx(Q1.IFD.PER_M1_lmer.results, "~./MLH1repo/PER.IFD_Q1_M1.xlsx", col_names = TRUE)
+#write_xlsx(Q1.IFD.PER_M1_lmer.results, "~./MLH1repo/PER.IFD_Q1_M1.xlsx", col_names = TRUE)
+
+write.xlsx(Q1.IFD.PER_M1_lmer.results, "PER.IFD_Q1_M1.xlsx", col.names = TRUE)
 
@@
 #Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
-glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") )  )
+
+glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                          )
+
 summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
@@
 #PER
-glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data= ( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                         )
+
 summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
@@
 
-glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ))
+glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                          )
 summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
@@
 #PER
-glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=( mouse.level.BivData_table %>% filter(strain != "CZECH") ) )
+glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=( mouse.level.BivData_table %>% filter(species == "M.musculus") %>% filter(strain !=  "CAST" ) %>% filter(strain !=  "MOLF" ) %>% filter(strain !=  "HMI" ) %>% filter(strain !=  "AST" ) %>% filter(strain !=  "PERC" ) %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "TOM" ) )
+                          )
 summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
@@
 
+re-write the blocks for M4 M5 MLH1 to total SC. to check results
+
+```{r Female.specific.glm.M4.M5-temp, echo=FALSE, eval=FALSE}
+#Change MLH1 to mean_skel
+
+#mean.skel
+#mouse.total.skel
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = F)
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = F)
+
+glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
+
+summry.female.strain.effects <- summary(glm.female.strain.subsp)
+#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
+
+write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/MLH1_M4_female_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( summry.female.strain.effects$coefficients), "MLH1_M4_female_glm.xlsx", col.names = TRUE, row.names = TRUE)
+
+MLH1.female.spc.glm.M5 <- glm(mean_co ~ strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
+sum_MLH1.female.spc.glm.M5 <- summary(MLH1.female.spc.glm.M5)
+
+write.table(sum_MLH1.female.spc.glm.M5$coefficients, "~./MLH1repo/MLH1_M5_female_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_MLH1.female.spc.glm.M5$coefficients), "MLH1_M5_female_glm.xlsx", col.names = TRUE, row.names = TRUE)
+```
+
+```{r tot.SC.Male.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
+#mean.skel
+#mouse.total.skel
+
+mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = F)
+mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = F)
+
+#M4
+totSC.male.glm.M4 <- glm(mean.skel ~ subsp * strain, data=(mouse.total.skel %>% filter(sex == "male")  %>% filter(subsp != "Cast") ) %>% filter(species == "M.musculus")   )
+
+sum_totSC.male.glm.M4 <- summary(totSC.male.glm.M4)
+
+#write.table(sum_MLH1.male.glm.M4$coefficients, "~./MLH1repo/MLH1_M4_male_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame(sum_totSC.male.glm.M4$coefficients), "TotSC.male.glm.M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+#M5
+totSC.male.glm.M5 <- glm(mean.skel ~ strain, data= 
+    (mouse.total.skel %>% filter(sex == "male")  %>% filter(subsp != "Cast") %>% filter(species == "M.musculus")) )
+
+sum_totSC.male.glm.M5 <- summary(totSC.male.glm.M5)
+
+#write.table(sum_MLH1.male.glm.M5$coefficients, "~./MLH1repo/MLH1_M5_male_glm.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_totSC.male.glm.M5$coefficients), "TotSC.male.glm.M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+#write_xlsx(sum_MLH1.male.glm.M4.2$coefficients, "~./MLH1repo/MLH1_M5_male_glm.xlsx", col_names = TRUE)
+```
+
 ```{r Q2.totSC.glm, echo=FALSE, eval=FALSE}
@@
 Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= mouse.total.skel %>% filter(sex == "male") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast"))
+
 sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)

### 2020-06-22 src/CommonFunc_MLH1repo.R

@@ -130,12 +130,12 @@ add_strain <- function(dat){
                                         ifelse(grepl("_LEWES_", dframe$mouse), "LEW",
                                                ifelse(grepl("PERC", dframe$mouse), "PERC",
                                                       
-                                                      ifelse(grepl("_MSM_", dframe$mouse), "MSM",
-                                                             ifelse(grepl("_MOLF_", dframe$mouse), "MOLF",
-                                                                    ifelse(grepl("_SKIVE_", dframe$mouse), "SKIVE",      
-                                                                           ifelse(grepl("_PWD_", dframe$mouse), "PWD",
-                                                                                  ifelse(grepl("_KAZ_", dframe$mouse), "KAZ",
-                                                                                         ifelse(grepl("_TOM_", dframe$mouse), "TOM",
+              ifelse(grepl("_MSM_", dframe$mouse), "MSM",
+             ifelse(grepl("_MOLF_", dframe$mouse), "MOLF",
+                  ifelse(grepl("_SKIVE_", dframe$mouse), "SKIVE",      
+                     ifelse(grepl("_PWD_", dframe$mouse), "PWD",
+                   ifelse(grepl("_KAZ_", dframe$mouse), "KAZ",
+                    ifelse(grepl("_TOM_", dframe$mouse), "TOM",
                            ifelse(grepl("_AST_", dframe$mouse), "AST",
               ifelse(grepl("_CZECH_", dframe$mouse), "CZECH",
                                                                                                               
@@ -237,34 +237,35 @@ add_subsp <- function(oldframe) {
                                 ifelse(grepl("LEW", dframe$strain), "Dom", 
                                        ifelse(grepl("LEWES", dframe$strain), "Dom",    
                                               ifelse(grepl("PERA", dframe$strain), "Dom",
-                                                     ifelse(grepl("PERC", dframe$strain), "Dom",
+                                     ifelse(grepl("PERC", dframe$strain), "Dom",
                                                             
-                                                            ifelse(grepl("CAST", dframe$strain), "Cast",
-                                                                   ifelse(grepl("CIM", dframe$strain), "Cast",
-                                                                          ifelse(grepl("HMI", dframe$strain), "Cast",
+                                    ifelse(grepl("CAST", dframe$strain), "Cast",
+                                      ifelse(grepl("CIM", dframe$strain), "Cast",
+                                     ifelse(grepl("HMI", dframe$strain), "Cast",
                                                                                  
[…]

### 2020-06-22 doc/MS_v3.Rmd
some more updates to intro and edits

@@
 The total number of COs at this stage of meiosis is equal to the genome-wide recombination rate (gwRR).
+Meiosis, and the mixing of haploid gametes, is an important process for evolution, because it can determine a populations’ response to selection by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds [@charlesworth1993].
+This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
+
+  Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process of a diploid genome producing haploid gametes.
+The total number of COs at this stage of meiosis is equal to the genome-wide recombination rate (gwRR).
 Meiosis (mixing of haploid gametes) is an important process for evolution.
@@
   **Due to it's role in facilitating chromosomes disjunction, the genome-wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**  (thus clear target of evolution).
-Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombiantion rate [@nagaoka2012] (Inoue).
+Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombination rate [@nagaoka2012] (Inoue).
 Yet, within these thresholds the genome-wide recombination rate still varies across species and individuals.
 Sex is one of the most notable axes of this form of variation.
-  There are two classes of sexual dimorphism in recombiantion rates; the first is called achiasmy and refered to where one sex recombines (Bell 1982 – Huxley 1928, Lenormand Duthiel 2005).
+  There are two classes of sexual dimorphism in recombination rates; the first is called achiasmy and refered to where one sex recombines (Bell 1982 – Huxley 1928, Lenormand Duthiel 2005).
 The second, which is the focus of this manuscript, is the instance where both sexes recombine (Lenormand and Duthiel 2003).
@@
 
-  The genome-wide recombiantion rate can be decomposed into single chromosome patterns, refered to as recombination landscapes.
+  The genome-wide recombination rate can be decomposed into single chromosome patterns, refered to as recombination landscapes.
 **(at this scale –there are also relevant features of heterochiasmy).** Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
@@
 
-  In our comparisons of recombiantion landscapes across sex, our results confirm previously reported patterns for sex differences in the general recombination landscape [@sardell_sex_2020].
+  In our comparisons of recombination landscapes across sex, our results confirm previously reported patterns for sex differences in the general recombination landscape [@sardell_sex_2020].
 Intriguingly,  these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
@@
 
-<new 1/n>
-
 To estimate the number of crossovers across the genome, we counted MLH1 foci.
@@
 
-<new 2/n>
 To measure bivalent SC length, two image analysis algorithms were used.
@@
 
-<new 3/n>
 To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
@@
 
-<new 4/n>
 To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
@@
 
-<new 5/n>
 We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
@@
 
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects **(M2; Musc p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, Mol p =  `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`)**, highlighting strain and sex as salient variables.
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; Musc p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, Mol p =  `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`), highlighting strain and sex as salient variables.
 Two strains exhibit particularly strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
 
-In separate analyses of males (M4; n = 80 - `r length(male_mouse.avs_for.analysis$mouse)`), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` ), *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`), and *musculus^SKIVE^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` ).
+In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)`), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` ), *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`), and *musculus^SKIVE^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` ).
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-
-In females (M4; n= 76 - `r length(female_mouse.avs_for.analysis$mouse)`), four strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]`), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]`), *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]`), and *musculus^PWD^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[4]`).
+In females (M4; n= `r length(female_mouse.avs_for.analysis$mouse)`), four strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]`), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]`), *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]`), and *musculus^PWD^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[4]`).
 Strain effect sizes in females are modest in magnitude compared to those in males.
-
 Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.

### 2020-06-22 src/Result_Setup.Rmd
added M4 and M5 to other traits

@@
 
+M4 and M5 for short bivalent
+
+```{r Female.shortBivalent.M4.M5, echo=FALSE}
+#these M4 and M5 could be run for short and Long
+Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
+Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
+Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)
+#mean_SC
+#Short.biv_mouse_table
+
+#M4
+ShortBiv.female.glm_M4 <- glm(mean_SC ~ subsp * strain, data= (Short.biv_mouse_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "female") ) )
+sum.ShortBiv.female.glm_M4 <- summary(ShortBiv.female.glm_M4)
+write.xlsx(as.data.frame( sum.ShortBiv.female.glm_M4$coefficients), "ShortBiv.female.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+#M5
+ShortBiv.female.glm_M5 <- glm(mean_SC ~ strain, data= (Short.biv_mouse_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "female") ) )
+sum.ShortBiv.female.glm_M5 <- summary(ShortBiv.female.glm_M5)
+write.xlsx(as.data.frame( sum.ShortBiv.female.glm_M5$coefficients), "ShortBiv.female.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+```
+
+```{r male.ShortBiv.M4.M5, echo=FALSE}
+#M4
+ShortBiv.male.glm_M4 <- glm(mean_SC ~ subsp * strain, data= (Short.biv_mouse_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "male") ) )
+sum.ShortBiv.male.glm_M4 <- summary(ShortBiv.male.glm_M4)
+write.xlsx(as.data.frame( sum.ShortBiv.male.glm_M4$coefficients), "ShortBiv.male.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+#M5
+ShortBiv.male.glm_M5 <- glm(mean_SC ~ strain, data= (Short.biv_mouse_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "male") ) )
+sum.ShortBiv.male.glm_M5 <- summary(ShortBiv.male.glm_M5)
+write.xlsx(as.data.frame( sum.ShortBiv.male.glm_M5$coefficients), "ShortBiv.male.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+```
+
 ```{r Q1.SC.M2plot, echo=FALSE, include=FALSE, eval=FALSE}
@@
 
-```{r Female.specific.glm.M4.M5-temp, echo=FALSE, eval=FALSE}
+```{r Female.specific.glm.M4.M5-temp, echo=FALSE, eval=TRUE}
 #Change MLH1 to mean_skel
@@
 
-glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
-
-summry.female.strain.effects <- summary(glm.female.strain.subsp)
-#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
-
-write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/MLH1_M4_female_glm.csv", sep=",",row.names = TRUE)
-
-write.xlsx(as.data.frame( summry.female.strain.effects$coefficients), "MLH1_M4_female_glm.xlsx", col.names = TRUE, row.names = TRUE)
-
-MLH1.female.spc.glm.M5 <- glm(mean_co ~ strain, data= (female_mouse.avs_for.analysis %>% filter(subsp != "Cast") ) )
-sum_MLH1.female.spc.glm.M5 <- summary(MLH1.female.spc.glm.M5)
-
-write.table(sum_MLH1.female.spc.glm.M5$coefficients, "~./MLH1repo/MLH1_M5_female_glm.csv", sep=",",row.names = TRUE)
-
-write.xlsx(as.data.frame( sum_MLH1.female.spc.glm.M5$coefficients), "MLH1_M5_female_glm.xlsx", col.names = TRUE, row.names = TRUE)
+totSC.female.glm_M4 <- glm(mean.skel ~ subsp * strain, data= (mouse.total.skel %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "female") ) )
+
+sum.totSC.female.glm_M4 <- summary(totSC.female.glm_M4)
+
+write.xlsx(as.data.frame( sum.totSC.female.glm_M4$coefficients), "totSC.female.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+totalSC.female.glm.M5 <- glm(mean.skel ~ strain, data= (mouse.total.skel %>% filter(subsp != "Cast")  %>% filter(species == "M.musculus")  %>% filter(sex == "female") ) )
+sum_totalSC.female.glm.M5 <- summary(totalSC.female.glm.M5)
+
+#write.table(sum_MLH1.female.spc.glm.M5$coefficients, "~./MLH1repo/sum_MLH1.female.spc.glm.M5.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_totalSC.female.glm.M5$coefficients), "totalSC.female.glm.M5.xlsx", col.names = TRUE, row.names = TRUE)
 ```
@@
 
-```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}
-
-Q2.glm_mus.av_1COpos_M4 <- glm(mean_1CO.pos.PER ~ subsp * strain, data = (mouse.level.BivData_table  %>% filter(sex == "male") %>% filter(species == "M.musculus") ) )
-
-sum.Q2.glm_mus.av_1COpos_M4  <- summary(Q2.glm_mus.av_1COpos_M4)$coefficients
+```{r Female.1COpos.M4.M5, echo=FALSE}
+#Change MLH1 to mean_skel
+
+#Pos  PER_Foci_1
+#cleanCurated_BivData
+
+cleanCurated_BivData$strain <- factor(cleanCurated_BivData$strain, ordered = F)
+cleanCurated_BivData$subsp <- factor(cleanCurated_BivData$subsp, ordered = F)
+
+chrm.1CO.pos.female.glm_M4 <- glm(PER_Foci_1 ~ subsp * strain, data= (cleanCurated_BivData %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "female") ) )
+
+sum.chrm.1CO.pos.female.glm_M4 <- summary(chrm.1CO.pos.female.glm_M4)
+
+write.xlsx(as.data.frame( sum.chrm.1CO.pos.female.glm_M4$coefficients), "chrm.1CO.pos.female.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+chrm.1CO.pos.female.glm_M5 <- glm(PER_Foci_1 ~ strain, data= (cleanCurated_BivData %>% filter(subsp != "Cast")  %>% filter(species == "M.musculus")  %>% filter(sex == "female") ) )
+sum_chrm.1CO.pos.female.glm_M5 <- summary(chrm.1CO.pos.female.glm_M5)
+
+#write.table(sum_MLH1.female.spc.glm.M5$coefficients, "~./MLH1repo/sum_MLH1.female.spc.glm.M5.csv", sep=",",row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_chrm.1CO.pos.female.glm_M5$coefficients), "chrm.1CO.pos.female.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+```
+
+```{r Male.Q2.1COpos_male_M4.M5, echo=FALSE}
+
+#norm.first.foci.pos AND mean_1CO.pos.PER  are the same
+
+male_1COpos.glm_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data = (mouse.level.BivData_table  %>% filter(sex == "male") %>% filter(species == "M.musculus")  ) )
+
+sum.male_1COpos.glm_M4  <- summary(male_1COpos.glm_M4)
 #summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]
 
-#M4.2 (just strain)
-Q2.glm_mus.av_1COpos_M4.2 <- glm(norm.first.foci.pos ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
-sum.Q2.glm_mus.av_1COpos_M4.2 <- summary(Q2.glm_mus.av_1COpos_M4.2)$coefficients
+write.xlsx(as.data.frame( sum.male_1COpos.glm_M4$coefficients), "male_1COpos.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+#M5 (just strain)
+male_1COpos.glm_M5 <- glm(norm.first.foci.pos ~ strain, data= mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )
+
+sum.male_1COpos.glm_M5 <- summary(male_1COpos.glm_M5)
+
+write.xlsx(as.data.frame( sum.male_1COpos.glm_M5$coefficients), "male_1COpos.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
 
@@
 
-```{r Q2.IFD.setup, echo=FALSE}
-
-#identify and remove mice with 1 bivalent observation (MSM:  1feb18_MSM_m1)
-#some cells just have 1 cell, but the MSM_m is the only with just 1 bivlent
-
-```
-
 ```{r Q2.IFD.long.short.setup, echo=FALSE}
@@
 #there is mouse MSM with v low, 1feb18_MSM_m1 -- it only has 1 measure
+
+```
+
+```{r Female.IFD.M4.M5, echo=FALSE}
+#mean_IFD.2CO_ABS
+#mouse.level.BivData_table
+
+mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = F)
+mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = F)
+
+#RAW IFD
+IFD.raw.female.glm_M4 <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "female") ) )
+sum.IFD.raw.female.glm_M4 <- summary(IFD.raw.female.glm_M4)
+write.xlsx(as.data.frame( sum.IFD.raw.female.glm_M4$coefficients), "IFD.raw.female.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+IFD.raw.female.glm_M5 <- glm(mean_IFD.2CO_ABS ~ strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast")  %>% filter(species == "M.musculus")  %>% filter(sex == "female") ) )
+sum_IFD.raw.female.glm_M5 <- summary(IFD.raw.female.glm_M5)
+write.xlsx(as.data.frame( sum_IFD.raw.female.glm_M5$coefficients), "IFD.raw.female.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+#PER IFD
+IFD.PER.female.glm_M4 <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "female") ) )
+
+sum.IFD.PER.female.glm_M4 <- summary(IFD.PER.female.glm_M4)
+write.xlsx(as.data.frame( sum.IFD.PER.female.glm_M4$coefficients), "IFD.PER.female.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+IFD.PER.female.glm_M5 <- glm(mean_IFD.2CO_PER ~ strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast")  %>% filter(species == "M.musculus")  %>% filter(sex == "female") ) )
+sum_IFD.PER.female.glm_M5 <- summary(IFD.PER.female.glm_M5)
+
+write.xlsx(as.data.frame( sum_IFD.PER.female.glm_M5$coefficients), "IFD.PER.female.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+```
+
+```{r Male.Q2.IFD_male_M4.M5, echo=FALSE}
+
+#RAW IFD
+IFD.raw.male.glm_M4 <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "male") ) )
+sum.IFD.raw.male.glm_M4 <- summary(IFD.raw.male.glm_M4)
+write.xlsx(as.data.frame( sum.IFD.raw.male.glm_M4$coefficients), "IFD.raw.male.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+IFD.raw.male.glm_M5 <- glm(mean_IFD.2CO_ABS ~ strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast")  %>% filter(species == "M.musculus")  %>% filter(sex == "male") ) )
+sum_IFD.raw.male.glm_M5 <- summary(IFD.raw.male.glm_M5)
+write.xlsx(as.data.frame( sum_IFD.raw.male.glm_M5$coefficients), "IFD.raw.male.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+#PER IFD
+IFD.PER.male.glm_M4 <- glm(mean_IFD.2CO_PER ~ subsp * strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast") %>% filter(species == "M.musculus") %>% filter(sex == "male") ) )
+sum.IFD.PER.male.glm_M4 <- summary(IFD.PER.male.glm_M4)
+write.xlsx(as.data.frame( sum.IFD.PER.male.glm_M4$coefficients), "IFD.PER.male.glm_M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+IFD.PER.male.glm_M5 <- glm(mean_IFD.2CO_PER ~ strain, data= (mouse.level.BivData_table %>% filter(subsp != "Cast")  %>% filter(species == "M.musculus")  %>% filter(sex == "male") ) )
+sum_IFD.PER.male.glm_M5 <- summary(IFD.PER.male.glm_M5)
+write.xlsx(as.data.frame( sum_IFD.PER.female.glm_M5$coefficients), "IFD.PER.male.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
 

### 2020-06-23 doc/MS_v3.Rmd
finalizing the v3 draft

@@
 This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
+Due to it’s role in facilitating chromosomes disjunction, the genome-wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.
+Ensuring at least one crossover per chromosome while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombination rate [@nagaoka2012] (Inoue).
+Yet, within these thresholds the genome-wide recombination rate still varies across species and individuals.
 
-  Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process of a diploid genome producing haploid gametes.
-The total number of COs at this stage of meiosis is equal to the genome-wide recombination rate (gwRR).
-Meiosis (mixing of haploid gametes) is an important process for evolution.
-Meiotic recombination can regulate a populations’ response to selection, and determine the fate of novel mutations by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds (Charlesworth 1993 cite).
-This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
-At the individual level (fertility of a individual), the meiotic program relies on crossovers (the products of meiotic recombination), to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [@CahoonLibuda2019].
+Sex, or the version of gametogenesis, is one of the most notable axes of this form of variation.
+There are two classes of sexual dimorphism in recombination rates; the first is called achiasmy and referred to where one sex recombines [@huxley1928; @lenormandDuthiel_2005].
+The second, which is the focus of this manuscript, is the instance where both sexes recombine [@lenormandDuthiel_2005].
+An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
+Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay] petite?
+and mouse [@srivastava2017CC] in which the role of sex is confounded with the contributions of genetic variation.
 
-<!-- In most species the meiotic program relies on crossovers to ensure the correct separation of chromosomes at the reductional division (4n to 2n) [@CahoonLibuda2019].
-### 1.
-Understanding levels of variation in gwRR -->
+**<evolutionary patterns of sexual dimorphism in traits such as body size, coloration, and – can be traced to sexual antagonism / directional selection due to --- differences in.
+not sure what I meant here**
 
-  **Due to it's role in facilitating chromosomes disjunction, the genome-wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.**  (thus clear target of evolution).
-Ensuring at least one crossover per chrm while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombination rate [@nagaoka2012] (Inoue).
-Yet, within these thresholds the genome-wide recombination rate still varies across species and individuals.
-Sex is one of the most notable axes of this form of variation.
-  There are two classes of sexual dimorphism in recombination rates; the first is called achiasmy and refered to where one sex recombines (Bell 1982 – Huxley 1928, Lenormand Duthiel 2005).
-The second, which is the focus of this manuscript, is the instance where both sexes recombine (Lenormand and Duthiel 2003).
-An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
-Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog],  cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay], and mouse [@srivastava2017CC] in which the role of sex is confounded with the contributions of genetic variation.
-
-<!-- 3.	Open Q, no first principle -->  
-In anisogamy ist systems where gametogenesis is modified to produce distinct gametes across sexes, the meiotic program is the same (gorelick?).
-There are no first principles which would predict the evolution of sexually dimorphic genome-wide recombination rates.
+For many secondary sexually dimorphic traits, the mode of selection can be traced to the divergent targets of selection.
+In anisogamy systems where gametogenesis is modified to produce distinct gametes across sexes, the meiotic program is the same and there is no first principles which would predict the evolution of sexually dimorphism in crossover number per cell.
 Yet heterochiasmy is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
-Although it is clear that the relationship between female and male recombination rates can differ among species, comparisons between and within closely related species are missing.
+Asymmetrical cell division is one of the best distinguishers for sex in metazoans, however there are exceptions, including scarid flies and brown algae [@gorelick2017], yet empirical data on meiotic recombination from such organisms is missing.
+In the organisms for which there are empirical data on meiotic recombination rates, it is clear that the relationship between female and male recombination rates can differ among species, however comparisons between and within closely related species are missing.
 Direct contrasts between females and males across a common, diverse set of genomic backgrounds would reveal whether the recombination rate evolves differently in the sexes.
@@
 
-  The genome-wide recombination rate can be decomposed into single chromosome patterns, refered to as recombination landscapes.
-**(at this scale –there are also relevant features of heterochiasmy).** Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
-In a review of data from mouse, celegans and worms with meiotic chromosome morphilogy (axis-loop structure) quantified for both sexes, @CahoonLibuda2019 find that the sex with longer chromosome axis generally has higher recombination rates.
+When considering the evolutionary patterns for genome based metric is to also consider patterns at the individual chromosome scale.
+The genome-wide recombination rate can be decomposed into single chromosome patterns, referred to as recombination landscapes.
+There are conserved patterns relevant to heterochiasmy at this scale.
+Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
+In a review of data from mus musculus, C.
+elegans and arabidopsis thalania with meiotic chromosome morphology (axis-loop structure) quantified in both sexes, @CahoonLibuda2019 find that the sex with longer chromosome axis generally has higher recombination rates.
 A survey of 51 species shows there are conserved broad scale recombination landscape differences.
 Generally males have telomere-bias crossover placement and females have more uniform placement [@sardell_sex_2020].
-These observations raise the question of how conservation in sexually dimorphic patterns are maintained with the gwRR and it’s decomposed traits.
-
------------
+These observations raise the question of how conservation of sexually dimorphic genome-wide recombination rates and its’ decomposed traits.
 
@@
 
-  The house mouse species, (_Mus musculus_), complex arose from a recent radiation (0.5 mya to 5 mya) (@geraldes2011; @geraldes2008) providing an opportunity to interrogate natural variation at short evolutionary scales.
-This feature narrows the mutational space for identifying variants which lead to genome-wide recombination rate variation.
-Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway **(Shementi?
-Hassold?
-cite)** and previous crosses for understanding the genetic architecture of recombination rate variation [@dumont2011; @murdoch2010; @Wang2017island].
+Integrating patterns across multiple scales – into a unified evolutionary pattern requires a good model system with multiple evolutionary levels for gauging the robustness of patterns across multiple scales.
+The house mouse, (Mus musculus), arose from a recent radiation (0.5 mya to 5 mya) (@geraldes2011; @geraldes2008) into several subspecies providing an opportunity to interrogate natural variation at short evolutionary scales.
+This short evolutionary scale is valuable for increasing heterochiasmy observations, since recombination rates are known to evolve rapidly (cite).
+Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway [@baudat2013; @handel2010],  and previous crosses for understanding the genetic architecture of recombination rate variation [@dumont2011; @murdoch2010; @Wang2017island].
 Additionally due to their global distribution allows comparisons of from a broad geographic range that might be missed with classical inbred strains.
 Wild derived inbred strains of house mouse, enable one of the most direct comparison of female and male gametogenesis, since besides the sex chromosomes, the mouse for each genome are identical.
-Additionally, House mouse is well suited for single cell cytology approaches, enabling integration of data at a closer connection to the molecular pathway and meiotic program.
-
---------
+Additionally, this model has advantages for single cell cytology approaches, enabling integration of data at a closer connection to the molecular pathway and meiotic program.
 
@@
 
-  In this study / paper, We quantify gwRRs from both sexes, in a panel of inbred mouse strains.
+In this study, we quantify gwRRs from both sexes, in a panel of inbred mouse strains.
 We observed highly distinct evolutionary trajectories between males and females.
@@
 This pattern of rapid evolution was not observed in females.
-
-<!-- we decompsed the gwRR into single chromosome features to access data which  reflects the non-random processes of crossover formation on individual chromosomes.
-The recombination landscape for single chromosome scale can be summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
--->
-
-  In our comparisons of recombination landscapes across sex, our results confirm previously reported patterns for sex differences in the general recombination landscape [@sardell_sex_2020].
+In our comparisons of recombination landscapes across sex, our results confirm previously reported patterns for sex differences in the general recombination landscape (Sardell and Kirkpatrick, 2020).
 Intriguingly, these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
-
-  Our comparison of recombination landscapes framed within the rapid male specific evolution of gwRR revealed unexpected patterns - <**put something interesting here**>.
-
-------
 
@@
 -->
-
--------------------
-
-# MATERIALS AND METHODS
-
-### Mice
-
-<new>
-We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 1).
-Our survey included 5 strains from *M.
-m.
-musculus*, 4 strains from *M.
-m.
-domesticus*, 2 strains from *M.
-m.
-molossinus*, 2 strains from *M.
-m.
-castaneus*, and 1 strain each from *M.
-spicilegus* and *M.
-spretus*.
-<removed _M.
-caroli_> We subsequently denote strains by their abbreviated subspecies and name (e.g.
-*domesticus^WSB^*).
-Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice from an inbred strain of Gough Island mice were sampled after **XX** <track down the generation of Gough inbred levels> generations of brother-sister mating.
-All mice were provided with ad libitum food and water.
-Procedures followed protocols approved by IACUC.
-
-### Tissue Collection and Immunohistochemistry
-
-<new>
-The same dry-down spread technique was applied to both spermatocytes and oocytes, following [@peters_1997], with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in [@peterson2019].
-The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
-Juvenile mice between 12 and 15 days of age were used for DMC1 counts.
-Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
-Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
-Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
-Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
-Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a Photoflow wash (Kodak, diluted 1:200).
-Slides were stored at -20\*C if not stained immediately.
-To visualize the structure of meiotic chromosomes, we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
-Crossovers (COs) were visualized as MLH1 foci.
-Double strand breaks (DSBs) were visualized as DMC1 foci.
-The staining protocol followed [@anderson1999] and [@koehler2002].
-Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
-Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\*C.
-The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\*C.
-Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
-Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
-
-### Image Processing
-
-<new>
-Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
-Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 foci or DMC1 foci using Photoshop.
-
-### Analysis
-
-To estimate the number of crossovers across the genome, we counted MLH1 foci.
-MLH1 foci were counted in cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score ranging from 1 (best) to 5 (worst) was assigned to each cell based on visual appearance of staining and spread of bivalents.
-Cells with a score of 5 were excluded from the final analysis.
-Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
-MLH1 foci located on the XY in spermatocytes were excluded from counts.
-In addition to MLH1 counts, we measured several traits to further characterize the recombination landscape.
-To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from each of a subset of strains (*molossinus^MSM^*, _musculus^PWD^_,  _domesticus^WSB^_, and _domesticus^G^_ ).
-SC morphology and CREST foci number were used to stage spermatocytes as early zygotene or late zygotene.
-
-To measure bivalent SC length, two image analysis algorithms were used.
-The first algorithm estimates the total (summed) SC length across bivalents for individual cells (@wang2019_SC).
-The second algorithm estimates the SC length of individual bivalents (@peterson2019).
-Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
-Total SC length per cell was quantified from pachytene cell images (@wang2019_SC).
-
-To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
-Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-SC length of individual bivalents was quantified in pachytene cell images (@peterson2019).
-The DNA CrossOver algorithm (@peterson2019) isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
-The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
-In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` (*molossinus^MSM^* male) to `r max(algorithm.stat.category$mean_pass.rate)` (*musculus^KAZ^* female).
-From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` bivalent objects were isolated by the algorithm.
-Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-The curated single bivalent data supplements our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
-Proportions of bivalents with the same number of MLH1 foci were compared across strains using a chi-square test.
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-This “short bivalent” data set included the four or five shortest bivalents and excluded the X bivalent in oocytes.
-A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
-Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
-A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
-A total of `r sum(Long.biv_mouse_table$n_obs)` long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
-
-To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
-A normalized measure of interference (IFD~norm~) was computed by dividing IFD~raw~ by SC length on a per-bivalent basis.
-
-We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
-We used mouse average as the dependent variable in all analyses.
-We first constructed a linear mixed model (M1) using lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
-In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT() from RLRsim [@RLRsim]).
-Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
-As each observation was at the level of a single mouse (mouse average), it was uniquely coded within the dataset, nesting was implicit, each strain only occurs within one strain, and was not explicitly coded in our linear and mixed models.
-We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
-In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
-Four additional linear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
 
@@
 
-<new 1/n>
 The variation in sex differences in recombination we discovered provided an opportunity to determine whether sex differences in chromatin compaction, as measured by the length of the synaptonemal complex (SC), are reversed when heterochiasmy is reversed.
-In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = 0.11 --  `r Q1_short.SC_t.test.SKIVE$p.value``).
+In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = `r Q1_short.SC_t.test.SKIVE$p.value`).
 
-Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from 1.26 - `r Q1SC.mus.av.short.biv_DIFF_PWD`  (*musculus^PWD^*) to 1.49 - `r Q1SC.mus.av.short.biv_DIFF_WSB` (*domesticus^WSB^*) across strains.
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = 7.16^{-20} -- `r Q1.lmer_totSC_M1_results$sex.results` ; M2: p = 5.33^{-4} -- `r sum.Q1.glm_totSC_M2$coefficients[,c(1,4)][9,2]`; M3: p =0.05 -- `r sum.Q1.glm_totSC_M3$coefficients[,c(1,4)][8,2]` ) and short bivalent length (M1: p = 1.27^{-11 -- `r Q1.lmer_short_M1_results$sex.results`}; M2: p = 1.98^{-7} -- `r summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)][8,2] / 10000`; M3: p = 1.98^{-7} -- `r summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)][8,2] / 10000`).
-
-The existence of some subspecies and strain effects (M1 subspecies\*sex: p = 0.03 -- `r Q1.lmer_short_M1_results$interaction.pval`; M2: strain-*musculus^SKIVE^* : p=0.02 -- `r summ.Q1.glm_short.SC_M2$coefficients[7,c(1,4)][2]` ; _musculus^SKIVE^_\*male: p = 0.08 -- `r summ.Q1.glm_short.SC_M2$coefficients[13,c(1,4)][2]`) further indicates that SC length has evolved among strains and among subspecies.
-
+Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from `r Q1SC.mus.av.short.biv_DIFF_PWD`  (*musculus^PWD^*) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (*domesticus^WSB^*) across strains.
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results` ; M2: p =
+`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,4][8]` ) and short bivalent length (M1: p = `r Q1.lmer_short_M1_results$sex.results`}; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]/10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]/10000`).
+The existence of some subspecies and strain effects on total SC and short bivalent SC Length further indicates that SC length has evolved among strains and among subspecies (See Supplemental Tables 5 and 8).
 In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males recombine more.
 This pattern implies that in high-recombination strains, spermatocytes have less space than oocytes in which to position additional crossovers.
-
-**double check that these are the only sig factors for all the models**
-
---------
 
@@
 We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: 0.68--`r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; t-test; p = `r Q1_1CO.pos_t.test.FULL$p.value`).
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; paired t-test; p = `r Q1.strain.F1.Pos$p.value`).
 Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`}; M2: p =  `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p =  `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
@@
 Examination of IFD~norm~ distributions indicates that females are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Models treating IFDnorm as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
-In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) across the full set of strains, whether they are considered separately or together.
+Models treating IFD~norm~ as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
+In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) (Supplemental Table 13 and 14) across the full set of strains, whether they are considered separately or together.
 Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males, and that males display a stronger telomeric bias in the placement of the distal crossover.
-
 In summary, controlling for differences in SC length (chromatin compaction) indicates that interference is stronger in males, whereas interference on the physical scale is similar in the sexes.
-
---------
 
@@
 
-<new 1/n>
 We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
@@
 
-<new 2/n>
-Ninety-six percent of single bivalents in our pooled dataset (n = 9,576 -- `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` ) have either one or two MLH foci (Figure 3).
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` ) have either one or two MLH foci (Supplemental Figure 2).
 The proportions of single-focus (1CO) bivalents vs.
-double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Figure 3).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are 0.33 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, 0.44 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_ , and 0.51 -- `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
+double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Supplemental Figure 2).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
 Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-*molossinus^MSM^*: p = 0.37 -- `r PWD.MSM.2CO.prop.test$p.value` ; _musculus^PWD^_ vs.
-_musculus^KAZ^_: p = 1.23^{-31} -- `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_: p = 2.65^{-6} -- `r MSM.MOLF.2CO.prop.test$p.value` ).
+*molossinus^MSM^*: p = `r PWD.MSM.2CO.prop.test$p.value` ; _musculus^PWD^_ vs.
+_musculus^KAZ^_: p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value` ).
 These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number across multiple bivalents.
 
-<new 3/n>
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of 76 -- `r sum( (DMC1table.main$ncells.x) )` early zygotene and 75 -- `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains (*musculus^PWD^* and *molossinus^MSM^*) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
-High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < 10^{-6} -- `r 1/1000000` ).
-In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = 0.66 -- `r ttest.HighLow.Z.pval` ).
+DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains (*musculus^PWD^* and *molossinus^MSM^*) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < `r 1/1000000` ).
+In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval` ).
 Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
@@
 
-<new 4/n>
-Total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 5).
-Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value` ), separate tests within subspecies show that the two strain categories differ within *M.
+Total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 3).
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within *M.
 m.
-molossinus* (p = 0.02 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value` ) but not within *M.
+molossinus* (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`) but not within *M.
 m.
-musculus* (p = 0.4 -- `r Q2tot.sc_t.test_High.Low_Musc$p.value` ).
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = 0.88 -- `r Q2short.biv_t.test_High.Low$p.value`  ; long: p = 0.19 --  `r Q2long.biv_t.test_High.Low$p.value` ).
+musculus* (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`  ; long: p = `r Q2long.biv_t.test_High.Low$p.value` ).
+In a model with total SC length as the dependent variable (M4), the two subspecies effects are significant (*M.
+m.
+musculus* p = `r sum_totSC.male.glm.M4$coefficients[,4][2]` and *M.
+m.
+molossinus*, p = `r sum_totSC.male.glm.M4$coefficients[,4][3]`), however there are also strain specific effects (Supplemental Table 6).
 
-Write results for M4 
-total SC M4 results -- many strain specific effects for the M4 and M5 models, but two subspecies are the largest, but MSM strain not significant, and SKIVE is in oposite direction.
-
-and M5 here
-
-Musc subspecies very ***, Mol *** (both ++)
-G, LEW, 
-MSM
-SKIVE, KAZ (- -)
-
-M5 (almost all strains are significant)
-G, LEW, PWD, MSM
-MOLF, 
-SKIVE, KAZ, TOM, AST, CZECH
-
-In a model with total SC length as the dependent variable (*M2?*), two subspecies effects are significant (*M.
-m.
-musculus* p = 2.33^{-6} -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`, *M.
-m.
-molossinus*, p = 21^{-6} -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`).
-In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) <what are these referencing?>, but they are not consistent across models.
+In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) (Supplemental Table 9), but they are not consistent across models.
 Collectively, these results reveal that evolution of SC length is not strongly associated with evolution of genome-wide recombination rate in house mice.
-
-<new 5/n>
 In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not to SC length and crossover position (on single-crossover bivalents).
-
---------------------------
 
@@
 We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
-During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@LaneKauppi2019; @vanVeen2003].
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@LaneKauppi2019; @vanVeen2003; @subramanian2014; @dumontDesai2012].
 The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesion holding homologs together [@LaneKauppi2019].
@@
 We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
-The number and placement of crossovers affects the area of sister chromosome cohesion distal to crossovers which needs to be released for the first reductional chromosome segregation [@vanVeen2003; @LaneKauppi2019; @subramanian2014; @dumontDesai2012].
-Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell (**Kudo** ), whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those organisms that undergo dictyate arrest [@Lee2019].
+Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell [@kudo2009], whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those organisms that undergo dictyate arrest [@Lee2019].
 
@@
 Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
-The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle (@ma2015_cattle) and differences between populations of white-footed mice [@peterson2019].
+The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle (@ma2015_cattle) and differences between wild and captive white-footed mice [@peterson2019].
 In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@segura2013; @ottoPaysuer2019].
@@
 A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM].
-More recent work connects the sparser recombination landscape to sex differences in the crossover repair pathway [@wang2017inefficient].
+Recent work in mice connects the sparser recombination landscape in females to sex differences in the crossover maturation efficiency [@wang2017inefficient].
 
@@
 
-<new 10/n>
 While the causes of sex differences in recombination remain mysterious @lenormand2016, our conclusions have implications for a wide range of recombination research.
@@
 
+# MATERIALS AND METHODS
+
+### Mice
+
+<new>
+We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 1).
+Our survey included 5 strains from *M.
+m.
+musculus*, 4 strains from *M.
+m.
+domesticus*, 2 strains from *M.
+m.
+molossinus*, 2 strains from *M.
+m.
+castaneus*, and 1 strain each from *M.
+spicilegus*, *M.
+spretus* and _M.
+caroli_.
+We subsequently denote strains by their abbreviated subspecies and name (e.g.
+*domesticus^WSB^*).
+Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
+Mice from an inbred strain of Gough Island mice were sampled after 17 generations of brother-sister mating.
+All mice were provided with ad libitum food and water.
+Procedures followed protocols approved by IACUC.
+
+### Tissue Collection and Immunohistochemistry
+
+The same dry-down spread technique was applied to both spermatocytes and oocytes, following [@peters_1997], with adjustment for volumes.
+Spermatocyte spreads were collected and prepared as described in [@peterson2019].
+The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
+Juvenile mice between 12 and 15 days of age were used for DMC1 counts.
+Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
+Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
+Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
+Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
+Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
+The following morning, slides were treated with a Photoflow wash (Kodak, diluted 1:200).
+Slides were stored at -20\*C if not stained immediately.
+To visualize the structure of meiotic chromosomes, we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
+Crossovers (COs) were visualized as MLH1 foci.
+Double strand breaks (DSBs) were visualized as DMC1 foci.
+The staining protocol followed [@anderson1999] and [@koehler2002].
+Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
+Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\*C.
+The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\*C.
+Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
+Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
+
+### Image Processing
+
+Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
+Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
+Image file names were anonymized before manual scoring of MLH1 foci or DMC1 foci using Photoshop.
+
+### Analysis
+
+To estimate the number of crossovers across the genome, we counted MLH1 foci.
+MLH1 foci were counted in cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score ranging from 1 (best) to 5 (worst) was assigned to each cell based on visual appearance of staining and spread of bivalents.
+Cells with a score of 5 were excluded from the final analysis.
+Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
+MLH1 foci located on the XY in spermatocytes were excluded from counts.
+In addition to MLH1 counts, we measured several traits to further characterize the recombination landscape.
+To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from each of a subset of strains (*molossinus^MSM^*, _musculus^PWD^_,  _domesticus^WSB^_, and _domesticus^G^_ ).
+SC morphology and CREST foci number were used to stage spermatocytes as early zygotene or late zygotene.
+
+To measure bivalent SC length, two image analysis algorithms were used.
+The first algorithm estimates the total (summed) SC length across bivalents for individual cells (@wang2019_SC).
+The second algorithm estimates the SC length of individual bivalents (@peterson2019).
+Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
+Total SC length per cell was quantified from pachytene cell images (@wang2019_SC).
+
+To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
+Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
+SC length of individual bivalents was quantified in pachytene cell images (@peterson2019).
+The DNA CrossOver algorithm (@peterson2019) isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
+The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
+In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` (*molossinus^MSM^* male) to `r max(algorithm.stat.category$mean_pass.rate)` (*musculus^KAZ^* female).
+From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` bivalent objects were isolated by the algorithm.
+Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
+The curated single bivalent data supplements our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
+Proportions of bivalents with the same number of MLH1 foci were compared across strains using a chi-square test.
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+This “short bivalent” data set included the four or five shortest bivalents and excluded the X bivalent in oocytes.
+A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
+Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
+A total of `r sum(Long.biv_mouse_table$n_obs)` long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
+
+To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
+A normalized measure of interference (IFD~norm~) was computed by dividing IFD~raw~ by SC length on a per-bivalent basis.
+
+We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
+We used mouse average as the dependent variable in all analyses.
+We first constructed a linear mixed model (M1) using lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
+In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT() from RLRsim [@RLRsim]).
+Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
+The hierarchal nature of the data, meant that nesting of levels across observations was implicit (ie.
+mouse within strain, within subspecies) and not explicitly coded.
+We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
+In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
+Four additional linear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
+
+# Acknowledgements
+
+# Competing interests
+
 # REFERENCES

### 2020-06-23 src/Result_Setup.Rmd
finalizing the v3 draft

@@
                         av.proportion_3CO = mean(proportion_3CO, na.rm = TRUE),
+
                         #1CO pos
@@
 
-```{r Female.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
+```{r MLH1.Female.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
 
@@
 
-```{r Male.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
+```{r MLH1.Male.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
 #M4
@@
 
+#rerun as strain means
+Q1.strain.F1.Pos <- t.test(
+( strain.BivData.Table %>% filter(species == "M.musculus") %>% filter(sex == "male") %>% filter(strain != "CZECH") %>% filter(strain != "MOLF") )$mean.1CO.pos,
+
+  ( strain.BivData.Table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$mean.1CO.pos, paired = TRUE  )
+
 #WSB
@@
 
-Q2tot.sc_t.test_High.Low <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(species == "M.musculus") )$mean.skel,
+Q2tot.sc_t.test_High.Low <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(species == "M.musculus") %>% filter(subsp != "Cast") )$mean.skel,
                 (mouse.total.skel %>% filter(species == "M.musculus") %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
 
-#mol -- why is there only 1 MOLF male now?
-#there are 3 MOLF males in bivdata
 Q2tot.sc_t.test_High.Low_Mol <- t.test( (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "0") %>% filter(subsp == "Mol") )$mean.skel,
-         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel)
+         (mouse.total.skel %>% filter(sex == "male") %>% filter(Rec.group == "1") %>% filter(subsp == "Mol") )$mean.skel )
 
@@
 
-The molf male t-tests don't work because I only have total Sc measures for 1 MOLF male.
-
 ```{r Q2.short.long.ttest, echo=FALSE}
@@
 
-Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
+Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") %>% filter(subsp != "Cast") )$mean.SC,
                                       (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)
@@
 
-```{r Q2.SC_across.subsp_short.long.mus.av, echo=FALSE, eval=FALSE}
-
-#DOM-musc-long
-Q2.long_t.test_DomMusc <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.08
-
-#DOM-musc-short
-Q2.short_t.test_DomMusc <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Musc"))$mean.SC)
-#p-value = 0.8
-
-#DOM-mol-long
-Q2.long_t.test_DomMol <- t.test((Q2mouse.av.long.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.long.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.4
-
-#DOM-mol-short
-Q2.short_t.test_DomMol <- t.test((Q2mouse.av.short.biv.table  %>% filter(subsp == "Dom") )$mean.SC,
-                          (Q2mouse.av.short.biv.table %>% filter(subsp == "Mol"))$mean.SC)
-#p-value = 0.09
-
-```
-
 ```{r Q2.SC_across.subsp_short.long_biv, echo=FALSE}
@@
 
-```{r Female.specific.glm.M4.M5-temp, echo=FALSE, eval=TRUE}
+```{r Female.specific.glm.M4.M5, echo=FALSE, eval=TRUE}
 #Change MLH1 to mean_skel
@@
 
-```{r Q2.glm.SC.long, echo=FALSE, eval=FALSE}
+```{r Q2.glm.SC.long, echo=FALSE}
+Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = F)
+Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = F)
 
 #this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+male.Long_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
+sum.male.Long_M4 <- summary(male.Long_M4)
+
+write.xlsx(as.data.frame( sum.male.Long_M4$coefficients), "male.long.M4.xlsx", col.names = TRUE, row.names = TRUE)
+
+#Q2mouse.av.long.biv.table$um.SC <- add_umeter(Q2mouse.av.long.biv.table$mean.SC)
+
+male.Long_M5 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+sum.male.Long_M5 <- summary(male.Long_M5)
+
+write.xlsx(as.data.frame( sum.male.Long_M5$coefficients), "male.Long_M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+```
+
+```{r Q2.glm.SC.long.old, echo=FALSE, eval=FALSE}
+
+#this dataset is all males
+glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
 sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
 
-glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
+glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
 sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
@@
 #pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
-
-```
-
-```{r Q2.glm.SC.long.old, echo=FALSE, eval=FALSE}
-
-#this dataset is all males
-glmQ2_long.biv_M4 <- glm(mean_SC ~ subsp * strain, data= (Long.biv_mouse_table %>% filter(sex == "male") )  )
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-
-glmQ2_long.biv_M4 <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") ))
-sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
-#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
-
-glmQ2_long.biv_M4.2_DOM <- glm(mean_SC ~ strain, data= (Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Dom") )  )
-
-sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)
-
-#MUSC
-glmQ2_long.biv_M4.2_MUSC <- glm(mean_SC ~ strain, data= 
-  ( Long.biv_mouse_table %>% filter(sex == "male") %>% filter(subsp == "Musc"))  )
-sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)
-
-#MOL can't be done -- only 1 mouse for each MSM and MOLF
-#glmQ2_long.biv_M4.2_MOL <- glm(mean_SC ~ strain, data=  ( Long.biv_mouse_table %>% filter(sex == "male") %>%  filter(subsp == "Mol") )  )
-
-#sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)
-
-#Dom
-glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
-#Musc
-glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
-#summary(glm_SC.male_Musc_long.biv)
-#Mol
-glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom")%>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain )
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )$strain, p.adj = "bonf", pool.sd = FALSE)
-
-#pairwise.t.test( ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$mean.SC, ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )$strain )
 ```

### 2020-06-23 doc/MS_v3.Rmd
finalized v3 of MS, sent to bret

@@
 output:
-  word_document:
+  html_document:
   pdf_document:

### 2020-06-23 src/CommonFunc_MLH1repo.R

@@ -445,6 +445,7 @@ addIFD.2 <- function(DF){
 }
 
 
+#convert pixel to micrometer. Conversion factor found from image metadata files
 add_umeter <- function(dat){
   #how to convert dat$chrm.length to chrm.length
   #sc unit

### 2020-06-30 doc/MS_v4.Rmd
starting to load in new edits, changing citations

--- 2020-06-23 doc/MS_v3.Rmd
@@
 #library(tidyverse)
-library(knitr)
+#library(knitr)
 library(ggplot2)
@@
 
+-new-
+Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
+Differences in recombination rate between females and males are perhaps the most striking form of this variation.
+Despite the recognition that the sexes recombine at different rates, existing data fail to address the extent to which recombination experiences similar evolutionary pressures in the two sexes.
+To fill this gap, we measured the genome-wide recombination rate in both sexes from a panel of wild-derived inbred strains sampled from across three subspecies of house mice.
+We applied single-cell immunohistochemistry to inbred strains to directly compare females and males from the same genetic background and meiotic context, thereby isolating the effect of sex on evolutionary differences in recombination.
+Our results reveal highly discordant evolutionary trajectories in the two sexes.
+Whereas male recombination rates show rapid evolution over short timescales, female recombination rates measured in the same strains are mostly static.
+Strains with high recombination in males have more double-strand breaks and stronger crossover interference than strains with low recombination in males, suggesting that these factors contribute to the sex-specific evolution we document.
+These results provide the strongest evidence yet that sex is a primary driver of recombination rate evolution.
+
+-old-
 Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
@@
 
-<!--### 0.   Importance of recombination rate-->
+-new-
+
+Meiosis converts diploid germ cells into haploid gametes.
+During meiosis I, DNA crossovers aid the separation of homologous chromosomes by physically linking them and establishing tension between them on the spindle (REF?).
+The wrong number of recombination events can disrupt chromosomal segregation, leading to infertility, miscarriage, and birth defects (Hassold and Hunt, 2001).
+Recombination also shapes evolution by shuffling the combinations of genetic variants offspring inherit.
+Recombination affects the fates of beneficial and deleterious mutations (Felsenstein, 1974; Fisher, 1930; Hill and Robertson, 1966) and interacts with natural selection to leave gradients in genomic patterns of diversity [@begun1992; @nachman2012; @charlesworth1993; @cutter2013; @smith1974] (Begun and Aquadro, 1992; Charlesworth et al., 1993; Cutter and Payseur, 2013; Maynard Smith and Haigh 1974).
+The role of recombination in facilitating meiotic chromosome assortment suggests that the total number of crossovers in a cell – the genome-wide recombination rate – is an important cellular characteristic connected to organismal fitness.
+The dual pressures of ensuring at least one crossover per chromosome and minimizing levels of DNA damage and ectopic exchange are thought to impose lower and upper thresholds on the genome-wide recombination rate (Inoue; Nagaoka et al., 2012).
+Yet, within these bounds, individuals from the same species can vary substantially in crossover number (Kong et al., 2008; Gruhn et al., 2013; Ma et al., 2015; Johnston et al., 2016).
+Sex is perhaps the most notable axis along which recombination rate varies.
+Broadly speaking, sexual dimorphism in the genome-wide recombination rate assumes two forms.
+In species such as Drosophila melanogaster, one sex completes meiosis without forming crossovers (“achiasmy”), while the other sex recombines (Haldane, 1922; Huxley, 1928; Burt et al., 1991).
+Alternatively, in most species with recombination, crossovers occur in both sexes but at different rates (“heterochiasmy”).
+In these species, females tend to recombine more than males (Bell, 1982; Brandvain and Coop, 2012; Burt et al., 1991; Lenormand and Dutheil, 2005; Lorch, 2005).
+In plants, heterochiasmy is correlated with the opportunity for haploid selection (Lenormand and Dutheil, 2005).
+Despite the establishment of these interspecific trends, an understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
+Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans (Kong et al., 2004, 2008, 2014; Halldorsson et al., 2019; Gruhn et al., 2013), dogs (Campbell et al.
+2014), cattle (Ma et al., 2015; Shen et al., 2018), and Soay sheep (Johnston et al., 2016), in which the role of sex is confounded with the contributions of genetic variation.
+Although it is known that the level and direction of heterochiasmy can differ among species (Brandvain and Coop 2012; Lenormand and Dutheil, 2005), the correlation between female and male recombination rates among closely related species remains poorly documented.
+Direct contrasts between the two sexes across a common, diverse set of genomic backgrounds that represent recent timescales would reveal whether the genome-wide recombination rate evolves differently in males and females.
+Examining variation in the total number of crossovers in a sex-specific manner could also illuminate evolutionary connections between recombination rate and crossover positioning.
+Analyses of meiotic chromosome morphology in Arabidopsis thaliana, Caenorhabditis elegans, and Mus musculus suggest that the sex with more recombination usually has longer chromosome axes (Cahoon and Libuda, 2019).
+A survey of 51 species found conserved sex differences in the recombination landscape, including telomere-biased placement of crossovers in males but not in females (Sardell and Kirkpatrick, 2020).
+The degree to which a crossover reduces the probability of another crossover nearby (crossover interference) also differs between females and males (Otto and Payseur 2019).
+The house mouse, Mus musculus, provides a compelling system for understanding how sex determines the evolution of recombination.
+Multiple subspecies share a most recent common ancestor approximately 0.5 million years ago (Geraldes et al.
+2011), providing the opportunity to examine natural variation on recent evolutionary timescales.
+Wild Mus musculus belong to the same species as classical inbred strains of mice, where the molecular and cellular pathways that lead to crossovers have been studied extensively (Baudat et al., 2013; Bolcun-Filas and Schimenti, 2012; Handel and Schimenti, 2010).
+Single-cell immunofluorescent approaches make it possible to estimate genome-wide recombination rates in individual males and females (Koehler et al.
+2002; Peters et al., 1997).
+A collection of wild-derived inbred strains founded from a variety of geographic locations is available, enabling genetic variation in recombination to be profiled across the species range.
+Most importantly, by measuring recombination rates in females and males from the same set of wild-derived inbred strains, the evolutionary dynamics of recombination can be directly compared in the two sexes.
+In this paper, we report genome-wide recombination rates from both sexes in a diverse panel of wild-derived inbred strains of house mice and their close relatives.
+We demonstrate that recombination rate evolves differently in females and males, even over short timescales.
+
+-old--
 
@@
 
-<!-- ### 2.  The House Mouse is a great model for uncovering evolutionary patterns at a short timescale -->
-
 Integrating patterns across multiple scales – into a unified evolutionary pattern requires a good model system with multiple evolutionary levels for gauging the robustness of patterns across multiple scales.
@@
 
-<!-- ### 3. What we accomplished in this paper -->
-
-<!-- Main messages: these results are centered/focused on a unique evo comparison 
-Advantages / importance of using multiple strains from and multiple subspecies
--importance / advantage of sampling for multiple geographic locations.
-Chromosome morphology / general recombination landscape
--results for general landscape 
-– conserved sex differences  -->
-
 In this study, we quantify gwRRs from both sexes, in a panel of inbred mouse strains.
@@
 Intriguingly, these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
-
-<!-- For the first time in many of these strains, We report a rare, direct, evolutionary comparison of recombination rate in females and males for this short evolutionary scale.
-We choose Mus musculus because of the available strains across subspecies.
-Our design of using multiple inbred strains within subspecies  and multiple subspecies across the species --  (can provide information anaglogous to polymorphism within and divergence within subspecies of the house mouse.
-Our results indicate rapid male specific evolution of gwRR occurred in a subset of house mouse lines.
-We observed up to a 30% difference (translating into approximately ~7 more crossovers per cell), a surprising amount considering the short evolutionary time scale.
-(found previously for PWK (forjt) (PWD beth) – but mostly measures from males.
-
-Brief mouse background and (touch on polymorphism 
--musculus subspecies – have mostly been samples from central Europe – we extend with samples from more eastern extent of the subspecies range in Russia and kazakstan. 
--molossinus is endemic to the japan – a hybrid that arose with humans brining musculus and castaneus to the islands.
-(AGAIN the divergence between mus musculus subspecies is ~500K – and males and females share the same phylogeny)
-
-OUR results for genome wide recombination rate (variation) we observed highly distinct evolutionary trajectories between males and females.
-This is due to rapid male specific evolution in the gwRR from two strains in musculus and 1 strain in molossinus.
-(this pattern of rapid evolution was not observed in females)
-
-The metric of the genome wide recombination rate reflects a non-random process of crossover formation on individual chromosomes // the recombination landscape at the single chromosome level.
-The recombination landscape for single chromosome scale can be general summarized by i) spacing of crossovers and ii) the linear area of synapsed homologous chromosomes where crossovers can form.
-
-We quantified the general recombination landscape because patterns in the recombination landscape are important to document because they can shed light on constraints and regulation of the processes ( could end up leading to variation in the genome wide recombination rates.
-Our results confirm (well established) previously reported patterns for sex differences in the general recombination landscape – and add by demonstrating that these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
--->
 
@@
 To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (Table 2; detailed results available in Supplemental Table X).
-Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`; sex: p = `r MLH1.MMlmer_results$sex.results`; subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`; sex: p = **`r MLH1.MMlmer_results$sex.results`;** subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
+
+<tried scientific_10 to make the scientific format of p values look better, haven't gotten it to work>
 

### 2020-06-30 doc/MS_v4.Rmd
more MS_v4 edits

@@
 
--old-
-Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
-The number of crossovers per meiotic cell is the  genome-wide recombination rate.
-Differences in recombination rate between females and males are perhaps the most striking form of this variation.
-Existing data fail to directly address the extent to which recombination experiences similar evolutionary pressures.
-To fill this gap, we measured meiotic recombination in both sexes for a panel of house mouse cross three subspecies.
-Using inbred strains and single cell immunohistochemistry allowed us to place sex-specific observations within the same genetic background and meiotic context.
-Our results indicate highly discordant evolutionary patterns in the two sexes.
-Whereas male recombination rates show evidence of rapid evolution over short timescales, female recombination rates measured in the same strains are mostly static.
-These results strongly indicate that house mouse has two, one for each sex, genome-wide recombination rates which display distinct evolutionary trajectories.
-
 # INTRODUCTION
@@
 Meiosis converts diploid germ cells into haploid gametes.
-During meiosis I, DNA crossovers aid the separation of homologous chromosomes by physically linking them and establishing tension between them on the spindle (REF?).
-The wrong number of recombination events can disrupt chromosomal segregation, leading to infertility, miscarriage, and birth defects (Hassold and Hunt, 2001).
+During meiosis I, DNA crossovers aid the separation of homologous chromosomes by physically linking them and establishing tension between them on the spindle **(REF?)**.
+The wrong number of recombination events can disrupt chromosomal segregation, leading to infertility, miscarriage, and birth defects [@hassold2001].
 Recombination also shapes evolution by shuffling the combinations of genetic variants offspring inherit.
-Recombination affects the fates of beneficial and deleterious mutations (Felsenstein, 1974; Fisher, 1930; Hill and Robertson, 1966) and interacts with natural selection to leave gradients in genomic patterns of diversity [@begun1992; @nachman2012; @charlesworth1993; @cutter2013; @smith1974] (Begun and Aquadro, 1992; Charlesworth et al., 1993; Cutter and Payseur, 2013; Maynard Smith and Haigh 1974).
+Recombination affects the fates of beneficial and deleterious mutations [@felsenstein1974; @fisher1930; @hill1966] and interacts with natural selection to leave gradients in genomic patterns of diversity [@begun1992; @nachman2012; @charlesworth1993; @cutter2013; @smith1974].
+
 The role of recombination in facilitating meiotic chromosome assortment suggests that the total number of crossovers in a cell – the genome-wide recombination rate – is an important cellular characteristic connected to organismal fitness.
-The dual pressures of ensuring at least one crossover per chromosome and minimizing levels of DNA damage and ectopic exchange are thought to impose lower and upper thresholds on the genome-wide recombination rate (Inoue; Nagaoka et al., 2012).
-Yet, within these bounds, individuals from the same species can vary substantially in crossover number (Kong et al., 2008; Gruhn et al., 2013; Ma et al., 2015; Johnston et al., 2016).
+The dual pressures of ensuring at least one crossover per chromosome and minimizing levels of DNA damage and ectopic exchange are thought to impose lower and upper thresholds on the genome-wide recombination rate [@nagaoka2012; @inoue2002].
+Yet, within these bounds, individuals from the same species can vary substantially in crossover number [@Kong2008; @gruhn2013; @ma2015_cattle; @johnston2016_soay].
+
 Sex is perhaps the most notable axis along which recombination rate varies.
 Broadly speaking, sexual dimorphism in the genome-wide recombination rate assumes two forms.
-In species such as Drosophila melanogaster, one sex completes meiosis without forming crossovers (“achiasmy”), while the other sex recombines (Haldane, 1922; Huxley, 1928; Burt et al., 1991).
+In species such as Drosophila melanogaster, one sex completes meiosis without forming crossovers (“achiasmy”), while the other sex recombines [@huxley1928; @haldane1922sex; @burt1991].
 Alternatively, in most species with recombination, crossovers occur in both sexes but at different rates (“heterochiasmy”).
 In these species, females tend to recombine more than males (Bell, 1982; Brandvain and Coop, 2012; Burt et al., 1991; Lenormand and Dutheil, 2005; Lorch, 2005).
-In plants, heterochiasmy is correlated with the opportunity for haploid selection (Lenormand and Dutheil, 2005).
+[@lorch2005; @brandvain2012scrambling; @burt1991; @lenormandDuthiel_2005 ] (**can't find Bell 1982 citation**).
+In plants, heterochiasmy is correlated with the opportunity for haploid selection [@lenormandDuthiel_2005].
 Despite the establishment of these interspecific trends, an understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
-Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans (Kong et al., 2004, 2008, 2014; Halldorsson et al., 2019; Gruhn et al., 2013), dogs (Campbell et al.
+Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle],  sheep [@johnston2016_soay] and mouse [@srivastava2017CC] (Kong et al., 2004, 2008, 2014; Halldorsson et al., 2019; Gruhn et al., 2013), dogs (Campbell et al.
 2014), cattle (Ma et al., 2015; Shen et al., 2018), and Soay sheep (Johnston et al., 2016), in which the role of sex is confounded with the contributions of genetic variation.
-Although it is known that the level and direction of heterochiasmy can differ among species (Brandvain and Coop 2012; Lenormand and Dutheil, 2005), the correlation between female and male recombination rates among closely related species remains poorly documented.
+Although it is known that the level and direction of heterochiasmy can differ among species [@brandvain2012scrambling; @lenormandDuthiel_2005], the correlation between female and male recombination rates among closely related species remains poorly documented.
 Direct contrasts between the two sexes across a common, diverse set of genomic backgrounds that represent recent timescales would reveal whether the genome-wide recombination rate evolves differently in males and females.
 Examining variation in the total number of crossovers in a sex-specific manner could also illuminate evolutionary connections between recombination rate and crossover positioning.
-Analyses of meiotic chromosome morphology in Arabidopsis thaliana, Caenorhabditis elegans, and Mus musculus suggest that the sex with more recombination usually has longer chromosome axes (Cahoon and Libuda, 2019).
-A survey of 51 species found conserved sex differences in the recombination landscape, including telomere-biased placement of crossovers in males but not in females (Sardell and Kirkpatrick, 2020).
-The degree to which a crossover reduces the probability of another crossover nearby (crossover interference) also differs between females and males (Otto and Payseur 2019).
+Analyses of meiotic chromosome morphology in Arabidopsis thaliana, Caenorhabditis elegans, and Mus musculus suggest that the sex with more recombination usually has longer chromosome axes [@CahoonLibuda2019].
+A survey of 51 species found conserved sex differences in the recombination landscape, including telomere-biased placement of crossovers in males but not in females [@sardell_sex_2020].
+The degree to which a crossover reduces the probability of another crossover nearby (crossover interference) also differs between females and males [@ottoPaysuer2019].
 The house mouse, Mus musculus, provides a compelling system for understanding how sex determines the evolution of recombination.
-Multiple subspecies share a most recent common ancestor approximately 0.5 million years ago (Geraldes et al.
-2011), providing the opportunity to examine natural variation on recent evolutionary timescales.
-Wild Mus musculus belong to the same species as classical inbred strains of mice, where the molecular and cellular pathways that lead to crossovers have been studied extensively (Baudat et al., 2013; Bolcun-Filas and Schimenti, 2012; Handel and Schimenti, 2010).
-Single-cell immunofluorescent approaches make it possible to estimate genome-wide recombination rates in individual males and females (Koehler et al.
-2002; Peters et al., 1997).
+Multiple subspecies share a most recent common ancestor approximately 0.5 million years ago [@geraldes2011], providing the opportunity to examine natural variation on recent evolutionary timescales.
+Wild Mus musculus belong to the same species as classical inbred strains of mice, where the molecular and cellular pathways that lead to crossovers have been studied extensively [@baudat2013; @handel2010; @bolcun2012].
+Single-cell immunofluorescent approaches make it possible to estimate genome-wide recombination rates in individual males and females [@koehler2002; @peters_1997].
 A collection of wild-derived inbred strains founded from a variety of geographic locations is available, enabling genetic variation in recombination to be profiled across the species range.
 Most importantly, by measuring recombination rates in females and males from the same set of wild-derived inbred strains, the evolutionary dynamics of recombination can be directly compared in the two sexes.
+
 In this paper, we report genome-wide recombination rates from both sexes in a diverse panel of wild-derived inbred strains of house mice and their close relatives.
 We demonstrate that recombination rate evolves differently in females and males, even over short timescales.
-
--old--
-
-Meiosis can be reduced to the expression (2n -> 4n -> 2n -> 1n) which reflects the process of a diploid genome producing haploid gametes.
-The total number of COs at this stage of meiosis is equal to the genome-wide recombination rate (gwRR).
-Meiosis, and the mixing of haploid gametes, is an important process for evolution, because it can determine a populations’ response to selection by transferring beneficial mutations onto novel genetic backgrounds or by breaking linkage of negative mutations from beneficial genetic backgrounds [@charlesworth1993].
-This process shapes the genomic patterns of genetic variation, with high recombination areas of genome having more nucleotide variation while areas of low recombination have decreased genetic variation [@begun1992; @nachman2012].
-Due to it’s role in facilitating chromosomes disjunction, the genome-wide recombination rate can be thought of as a cell-based metric which is tightly connected to organisms fitness and fertility.
-Ensuring at least one crossover per chromosome while minimizing the rates of DNA damage and ectopic recombination are thought to regulate the lower and upper thresholds for a genome-wide recombination rate [@nagaoka2012] (Inoue).
-Yet, within these thresholds the genome-wide recombination rate still varies across species and individuals.
-
-Sex, or the version of gametogenesis, is one of the most notable axes of this form of variation.
-There are two classes of sexual dimorphism in recombination rates; the first is called achiasmy and referred to where one sex recombines [@huxley1928; @lenormandDuthiel_2005].
-The second, which is the focus of this manuscript, is the instance where both sexes recombine [@lenormandDuthiel_2005].
-An understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
-Comprehensive comparisons of female and male recombination rates usually come from outbred populations humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle], (other sheep?) sheep [@johnston2016_soay] petite?
-and mouse [@srivastava2017CC] in which the role of sex is confounded with the contributions of genetic variation.
-
-**<evolutionary patterns of sexual dimorphism in traits such as body size, coloration, and – can be traced to sexual antagonism / directional selection due to --- differences in.
-not sure what I meant here**
-
-For many secondary sexually dimorphic traits, the mode of selection can be traced to the divergent targets of selection.
-In anisogamy systems where gametogenesis is modified to produce distinct gametes across sexes, the meiotic program is the same and there is no first principles which would predict the evolution of sexually dimorphism in crossover number per cell.
-Yet heterochiasmy is commonly observed in dioecious species, suggesting that other meiotic traits which distinguish the gametes and their meiotic programs, for example symmetrical vs asymmetrical cell division, may impose selection for sexually dimorphic recombination rates.
-Asymmetrical cell division is one of the best distinguishers for sex in metazoans, however there are exceptions, including scarid flies and brown algae [@gorelick2017], yet empirical data on meiotic recombination from such organisms is missing.
-In the organisms for which there are empirical data on meiotic recombination rates, it is clear that the relationship between female and male recombination rates can differ among species, however comparisons between and within closely related species are missing.
-Direct contrasts between females and males across a common, diverse set of genomic backgrounds would reveal whether the recombination rate evolves differently in the sexes.
-
-<!-- 4.	Additional metrics (more nuanced) -->
-
-When considering the evolutionary patterns for genome based metric is to also consider patterns at the individual chromosome scale.
-The genome-wide recombination rate can be decomposed into single chromosome patterns, referred to as recombination landscapes.
-There are conserved patterns relevant to heterochiasmy at this scale.
-Recently two specific features have been highlighted as conserved patterns related heterochiasmy, specifically the recombination landscape.
-In a review of data from mus musculus, C.
-elegans and arabidopsis thalania with meiotic chromosome morphology (axis-loop structure) quantified in both sexes, @CahoonLibuda2019 find that the sex with longer chromosome axis generally has higher recombination rates.
-A survey of 51 species shows there are conserved broad scale recombination landscape differences.
-Generally males have telomere-bias crossover placement and females have more uniform placement [@sardell_sex_2020].
-These observations raise the question of how conservation of sexually dimorphic genome-wide recombination rates and its’ decomposed traits.
-
-Integrating patterns across multiple scales – into a unified evolutionary pattern requires a good model system with multiple evolutionary levels for gauging the robustness of patterns across multiple scales.
-The house mouse, (Mus musculus), arose from a recent radiation (0.5 mya to 5 mya) (@geraldes2011; @geraldes2008) into several subspecies providing an opportunity to interrogate natural variation at short evolutionary scales.
-This short evolutionary scale is valuable for increasing heterochiasmy observations, since recombination rates are known to evolve rapidly (cite).
-Classical lab strains of mice have generated extensive knowledge central to meiosis and outcomes on recombination through extensive studies on the genes involved in the meiotic recombination pathway [@baudat2013; @handel2010],  and previous crosses for understanding the genetic architecture of recombination rate variation [@dumont2011; @murdoch2010; @Wang2017island].
-Additionally due to their global distribution allows comparisons of from a broad geographic range that might be missed with classical inbred strains.
-Wild derived inbred strains of house mouse, enable one of the most direct comparison of female and male gametogenesis, since besides the sex chromosomes, the mouse for each genome are identical.
-Additionally, this model has advantages for single cell cytology approaches, enabling integration of data at a closer connection to the molecular pathway and meiotic program.
-
-In this study, we quantify gwRRs from both sexes, in a panel of inbred mouse strains.
-We observed highly distinct evolutionary trajectories between males and females.
-This is due to rapid male specific evolution in the gwRR from two strains in musculus and 1 strain in molossinus.
-We observed up to a 30% difference (translating into approximately ~7 more crossovers per cell), a surprising amount considering the short evolutionary time scale.
-This pattern of rapid evolution was not observed in females.
-In our comparisons of recombination landscapes across sex, our results confirm previously reported patterns for sex differences in the general recombination landscape (Sardell and Kirkpatrick, 2020).
-Intriguingly, these patterns hold even in the case of genome more crossovers, but general landscape patterns still maintain sex-specific patterns.
 
@@
 
+-new-
+
+We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` -- 14 wild-derived inbred strains sampled from three subspecies of house mice ( _M.
+m.
+domesticus_, _M.
+m.
+musculus_ and _M.
+m.
+molossinus_ ) and three other species of Mus ( _M.
+spretus_, _M.
+spicilegus_ , and _M.
+caroli_ ).
+Mean MLH1 focus counts for `r sum(main.table$Nmice)` -- 188 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 21.77 spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` -- 1,742 spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` -- 17.85 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,427 oocytes).
+**<insert statement for cell sample size number>**
+
+Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
+First, MLH1 focus counts differ between females and males in most strains.
+Second, the difference in counts between the sexes varies among strains.
+Although most strains show more MLH1 foci in females, two strains (*musculus^PWD^* and _molossinus^MSM^_) exhibit higher counts in males.
+In females, numbers of MLH1 foci are evenly distributed around the sex-wide mean of approximately 25 (Figure 1B).
+In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23) of foci (Figure 1C).
+Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = `r sperman.corr.FM` -- 0.08; p = `r sperman.corr.FM.pval$p.value` -- 0.84) across the set of strains.
+
+To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` -- 137 house mice (from three subspecies; _M.
+m.
+domesticus_, _M.
+m.
+musculus_ and _M.
+m.
+molossinus_) (Table 2; detailed results available in Supplemental Table X).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`-- 10-4; sex: p = `r MLH1.MMlmer_results$sex.results` -- 3.64 x 10-6; subspecies: p = `r MLH1.MMlmer_results$sub.sp` -- 9.69 x 10-4; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval` -- 1.8 x 10-4).
+
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; musculus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` -- 0.24, molossinus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` -- 0.1), highlighting strain and sex as salient variables.
+Two strains exhibit strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` -- 1.78 x 10-6; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]` -- 0.02), with sex-strain interactions involving three strains (M3; *domesticus^G^* p < x 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p < x 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` -- 3.87 x 10-4).
+
+In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)` -- 71), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` -- 3.6 x 10-7; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` -- 6.11 foci **(add CI’s here)**, *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` -- 6.3 x 10-9; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`-- 6.91), and *musculus^SKIVE^* (p = -- 8.22 x 10-4; effect = -- 4.04).
+These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
+In females (M4; n= `r length(female_mouse.avs_for.analysis$mouse)` --76), three strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]` -- 8.7 x 10-6; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]` -- 3.3), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]` -- 2.43 x 10-5; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]` -- 2.99), and *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]`  -- 0.03; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]` -- 1.69).
+Strain effect sizes in females are modest in magnitude compared to those in males.
+Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.
+
+-old-
+
 We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice (*M.
@@
 To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (Table 2; detailed results available in Supplemental Table X).
-Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`; sex: p = **`r MLH1.MMlmer_results$sex.results`;** subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`; sex: p = `r MLH1.MMlmer_results$sex.results`; subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
 
@@
 
+-new-
+
+The variation in sex differences in recombination we discovered provided an opportunity to determine whether sex differences in chromatin compaction, as measured by the length of the synaptonemal complex (SC), are reversed when heterochiasmy is reversed.
+In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = `r Q1_short.SC_t.test.SKIVE$p.value` -- 0.11).
+Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from `r Q1SC.mus.av.short.biv_DIFF_PWD` -- 1.26 (*musculus^PWD^*) to `r Q1SC.mus.av.short.biv_DIFF_WSB` -- 1.52 (*domesticus^WSB^*) across strains.
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results` -- 2.56 x 10-31; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,4][8]` -- 2.56 x 10-8; M3: p =  `r sum.Q1.glm_totSC_M3$coefficients[,4][8]` -- 2.56 x 10-8) and short bivalent SC length (M1: p = `r Q1.lmer_short_M1_results$sex.results` -- 1.12 x 10-11; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]/10000` -- 1.33 x 10-7; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]/10000` -- 1.33 x 10-7).
+The existence of some subspecies and strain effects on total SC length and short bivalent SC length (Supplemental Tables 5 and 8) further indicates that SC length has evolved among strains and among subspecies.
+In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males recombine more.
+This pattern implies that in high-recombination strains, spermatocytes have less space than oocytes in which to position additional crossovers.
+
+-old-
+
 The variation in sex differences in recombination we discovered provided an opportunity to determine whether sex differences in chromatin compaction, as measured by the length of the synaptonemal complex (SC), are reversed when heterochiasmy is reversed.
@@
 That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results` ; M2: p =
-`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,4][8]` ) and short bivalent length (M1: p = `r Q1.lmer_short_M1_results$sex.results`}; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]/10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]/10000`).
+`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,4][8]` ) and short bivalent length (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]/10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]/10000`).
 The existence of some subspecies and strain effects on total SC and short bivalent SC Length further indicates that SC length has evolved among strains and among subspecies (See Supplemental Tables 5 and 8).
@@
 
+-new-
+
+We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.68 **add s.d here?**; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` --  0.56; paired t-test; p = `r Q1.strain.F1.Pos$p.value` -- 8.49 x 10-4).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results` -- 2.82 x 10-26; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]` -- 3.96 x 10-8; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]` -- 3.96 x 10-8).
+
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = `r Q1.IFD.PER_t.test_KAZ$p.value` -- 0.33).
+Examination of IFD~norm~ distributions indicates that females are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
+Models treating IFD~norm~ as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` -- 9.08 x 10-12; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]` -- 0.01; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]` -- 0.01).
+In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) (Supplemental Table 13 and 14) across the full set of strains, whether they are considered separately or together.
+Visualization of normalized MLH1 foci positions on bivalents with two crossovers **(Figure 4C and Supplemental Figure X)** further suggests that interference distances vary more in females than in males, and that males display a stronger telomeric bias in the placement of the distal crossover.
+
+In summary, controlling for differences in SC length (chromatin compaction) indicates that interference is consistently stronger in males, whereas interference on the physical scale is similar in the two sexes.
+
+-old-
+
 We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
 In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; paired t-test; p = `r Q1.strain.F1.Pos$p.value`).
-Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`}; M2: p =  `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p =  `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 
@@
 ```
+
+-new-
+
+We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
+We considered proportions of bivalents with different numbers of crossovers, double-strand break number, SC length, and crossover positioning.
+
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")`-- 9,569) have either one or two MLH foci (Supplemental Figure 2).
+The proportions of single-focus (1CO) bivalents vs.
+double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Supplemental Figure 2).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.51 in _molossinus^MSM^_ (Supplemental Figure 3).
+Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
+_molossinus^MSM^_: p = `r PWD.MSM.2CO.prop.test$p.value` -- 0.37; _musculus^PWD^_ vs.
+*musculus^KAZ^*: p = `r PWD.KAZ.2CO.prop.test$p.value` -- 1.23 x 10-31; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value` -- 2.34 x  10-6).
+These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number across multiple bivalents.
+
+To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
+DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` -- 76 early zygotene and `r sum( (DMC1table.main$ncells.y) )` -- 75 late zygotene spermatocytes from two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; -- `r 1/1000000`p < 10-6).
+In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval` -- 0.66).
+Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
+High-recombination and low-recombination strains do not differ in the MLH1/DMC1 ratio, whether DMC1 foci were counted in early zygotene cells or late zygotene cells (t-test; p > 0.05).
+These results raise the possibility that the evolution of genome-wide recombination rate is primarily determined by processes that precede the CO/NCO decision, at least in house mice.
+
+Total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 3).
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value` -- 0.01), separate tests within subspecies show that the two strain categories differ within *M.
+m.
+molossinus* (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` -- 2.59 x 10-4) but not within *M.
+m.
+musculus* (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value` -- 0.65).
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value` -- 0.84; long: p = -- 0.19).
+In a model with total SC length as the dependent variable (M4), the two subspecies effects are significant (*M.
+m.
+musculus* p = -- 3.95 x 10-7; *M.
+m.
+molossinus* p = `r Q2long.biv_t.test_High.Low$p.value` -- 3.33 x 10-7), but there are also strain-specific effects (Supplemental Table 6).
+In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) (Supplemental Table 9), but they are not consistent across models.
+Collectively, these results reveal that evolution of SC length is not strongly associated with evolution of genome-wide recombination rate in house mice.
+
+In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not to SC length and crossover position (on single-crossover bivalents).
+
+-old-
 
@@
 # DISCUSSION
+
+-new-
+
+By comparing recombination rates in females and males from the same diverse set of genetic backgrounds, we isolated sex as a primary factor in the evolution of this fundamental meiotic trait.
+Recombination rate differences are more pronounced in males than females.
+Because inter-strain divergence times are identical for the two sexes, this observation demonstrates that the genome-wide recombination rate evolves faster in males.
+More generally, recombination rate divergence is decoupled in females and males.
+These disparities are remarkable given that recombination rates for the two sexes were measured in identical genomic backgrounds (other than the number and identity of sex chromosomes).
+Our results provide the strongest evidence yet that the genome-wide recombination rate follows distinct evolutionary trajectories in males and females.
+
+At the genetic level, the sex-specific patterns we documented indicate that some mutations responsible for the evolution of recombination rate have dissimilar phenotypic effects in the two sexes.
+A subset of the genetic variants associated with genome-wide recombination rate within populations of humans (Kong et al., 2004, 2008, 2014; Halldorsson et al., 2019), Soay sheep (Johnston et al., 2016), and cattle (Ma et al., 2015; Shen et al., 2018) [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle; @Shen2018_cattle] appear to show sex-specific properties, including opposite effects in females and males.
+Furthermore, inter-sexual correlations for recombination rate are weak in humans [@fledel2011] (Fledel-Alon et al., 2011) and Soay sheep [@johnston2016_soay] (Johnston et al., 2016).
+Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011; @Wang2017island; @wang2019_SC] (Dumont and Payseur, 2011; Wang et al., 2019; Wang and Payseur, 2017).
+These variants could differentially affect females and males at any step in the recombination pathway.
+Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contribute to sex-specific evolution in the recombination rate.
+A study of two classical inbred strains and one wild-derived inbred strain of house mice also found a positive association between crossover number and double-strand break number in males [@baier2014] (Baier et al., 2014).
+
+Another implication of our results is that the connection between recombination rate and fitness differs between males and females.
+Little is known about whether and how natural selection shapes recombination rate in nature [@DapperPayseur2017; @Ritz2017] (Dapper and Payseur, 2017; Ritz et al., 2017).
+@samuk2020 Samuk et al.
+(2020) recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of Drosophila pseudoobscura was caused by natural selection.
+Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
+
+Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling; @sardell_sex_2020] (Brandvain and Coop, 2012; Sardell and Kirkpatrick, 2020).
+Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] (Lenormand, 2003) or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020] (Sardell and Kirkpatrick, 2020).
+Another modifier model showed that meiotic drive could stimulate female-specific evolution of the recombination rate [@brandvain2012scrambling] (Brandvain and Coop, 2012).
+Although these models fit the conserved pattern of sex differences in crossover positions, they do not readily explain our observations of sex-specific evolution in the genome-wide recombination rate.
+In particular, the alternation across strains in which sex has more crossovers is unexpected.
+
+We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@dumontDesai2012; @LaneKauppi2019; @vanVeen2003; @subramanian2014] (Dumont and Desai, 2012; Lane and Kauppi, 2019; Subramanian and Hochwagen, 2014; VanVeen and Hawley, 2003).
+The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesion holding homologs together [@LaneKauppi2019] (Lane and Kauppi, 2019).
+Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and chromosome separation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
+Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019] (Lane and Kauppi, 2019), perhaps due to the presence of the acentrosome spindle [@So2019] (So et al., 2019) and larger cell volume [@kyogoku2017] (Kyogoku and Kitajima, 2017) in oocytes.
+The higher stringency of the SAC during spermatogenesis suggests that selection will be better at removing mutations that interfere with bi-orientation in males than in females.
+Therefore, faster male evolution of the genome-wide recombination rate could be driven by the more stringent SAC acting on chromosome structures at the metaphase I alignment.
+
+Our SAC model is consistent with other features of our data.
+We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] (Sardell and Kirkpatrick, 2020) apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
+Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell [@kudo2009] (Kudo et al., 2009), whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the meiosis I spindle, especially for those organisms that undergo dictyate arrest [@Lee2019] (Lee, 2019).
+
+We propose that the SAC model also can explain the correlated evolution of stronger crossover interference and higher genome-wide recombination rate in male house mice.
+Our results show that crossovers are spaced further apart in strains enriched for double-crossover bivalents when SC length is considered and bivalent size effects are minimized.
+Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and may improve the fidelity of chromosomal segregation.
+While the SAC model postulates direct fitness effects of interference, a modifier model predicted that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993] (Goldstein et al., 1993).
+
+Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
+The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle [@ma2015_cattle] (Ma et al., 2015).
+In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@ottoPaysuer2019; @segura2013] (Otto and Payseur, 2019; Segura et al., 2013).
+The evolution of crossover interference and its relationship to changes in crossover number on the genomic scale is a topic deserving of more empirical and theoretical work.
+Our findings further reveal that evolution of the genome-wide recombination rate does not require major changes in the degree of chromatin compaction.
+Female house mice consistently show longer SCs, even in strains with more recombination in males.
+Studies in mice [@lynn2002; @petkov2007]  (Lynn et al., 2002; Petkov et al., 2007) and humans [@gruhn2013; @tease2004] (Gruhn et al., 2013; Tease and Hulten, 2004) suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+Some authors have suggested that conserved sex differences in crossover positioning (more uniform placement in females) and interference strength (stronger interference in males) could be due to looser chromatin packing of the meiotic chromosome structure in females [@haenel2018; @petkov2007] (Haenel et al., 2018; Petkov et al., 2007).
+A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM] (Hultén, 2011).
+Recent work in mice connects the sparser recombination landscape in females to sex differences in the crossover maturation efficiency [@wang2017inefficient] (Wang S.
+et al., 2017).
+
+Our conclusions are accompanied by several caveats.
+First, MLH1 foci only identify interfering crossovers [@holloway2008mus81] (Holloway et al., 2008).
+Although most crossovers belong to this class [@holloway2008mus81] (Holloway et al., 2008), our approach likely underestimated genome-wide recombination rates.
+Evolution of the number of non-interfering crossovers is a subject worth examining.
+A second limitation is that our investigation of crossover locations was confined to the relatively low resolution possible with immunofluorescent cytology.
+Positioning crossovers with higher resolution and maintaining single chromosome scale could reveal additional evolutionary patterns.
+Finally, the panel of inbred lines we surveyed may not be representative of recombination rate variation within and between subspecies of house mice.
+We considered most available wild-derived inbred lines, but house mice have a broad geographic distribution.
+Nevertheless, we expect our primary conclusion that recombination rate evolves in a sex-specific manner to be robust to geographic sampling because differences between females and males exist for the same set of inbred strains.
+
+While the causes of sex differences in recombination remain mysterious  @lenormand2016 (Lenormand et al., 2016), our conclusions have implications for a wide range of recombination research.
+For biologists uncovering the cellular and molecular determinants of recombination, our results suggest that mechanistic differences between the sexes could vary by genetic background.
+For researchers charting the evolutionary trajectory of recombination, our findings indicate that sex-specific comparisons are crucial.
+For theoreticians building evolutionary models of recombination, different fitness regimes and genetic architectures in females and males should be considered.
+Elevating sex as a primary determinant of recombination would be a promising step toward integrating knowledge of cellular mechanisms with evolutionary patterns to understand recombination rate variation in nature.
+
+-old-
 
@@
 
-<new>
+<not many edits to replace>
+
 We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 1).
@@
 Procedures followed protocols approved by IACUC.
+**statement on biological replicates**
 
@@
 To measure bivalent SC length, two image analysis algorithms were used.
-The first algorithm estimates the total (summed) SC length across bivalents for individual cells (@wang2019_SC).
-The second algorithm estimates the SC length of individual bivalents (@peterson2019).
+The first algorithm estimates the total (summed) SC length across bivalents for individual cells [@wang2019_SC].
+The second algorithm estimates the SC length of individual bivalents [@peterson2019].
 Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
-Total SC length per cell was quantified from pachytene cell images (@wang2019_SC).
+Total SC length per cell was quantified from pachytene cell images [@wang2019_SC].
 
@@
 Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-SC length of individual bivalents was quantified in pachytene cell images (@peterson2019).
-The DNA CrossOver algorithm (@peterson2019) isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
+SC length of individual bivalents was quantified in pachytene cell images [@peterson2019].
+The DNA CrossOver algorithm [@peterson2019] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
 The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.

### 2020-06-30 doc/MS_v4.Rmd
going through more edits

@@
 Alternatively, in most species with recombination, crossovers occur in both sexes but at different rates (“heterochiasmy”).
-In these species, females tend to recombine more than males (Bell, 1982; Brandvain and Coop, 2012; Burt et al., 1991; Lenormand and Dutheil, 2005; Lorch, 2005).
-[@lorch2005; @brandvain2012scrambling; @burt1991; @lenormandDuthiel_2005 ] (**can't find Bell 1982 citation**).
+In these species, females tend to recombine more than males [@lorch2005; @brandvain2012scrambling; @burt1991; @lenormandDuthiel_2005; @bell1982].
 In plants, heterochiasmy is correlated with the opportunity for haploid selection [@lenormandDuthiel_2005].
 Despite the establishment of these interspecific trends, an understanding of how sex shapes the evolution of recombination cannot be achieved with available data.
-Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle],  sheep [@johnston2016_soay] and mouse [@srivastava2017CC] (Kong et al., 2004, 2008, 2014; Halldorsson et al., 2019; Gruhn et al., 2013), dogs (Campbell et al.
-2014), cattle (Ma et al., 2015; Shen et al., 2018), and Soay sheep (Johnston et al., 2016), in which the role of sex is confounded with the contributions of genetic variation.
+
+Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle], and Soay sheep [@johnston2016_soay], in which the role of sex is confounded with the contributions of genetic variation.
 Although it is known that the level and direction of heterochiasmy can differ among species [@brandvain2012scrambling; @lenormandDuthiel_2005], the correlation between female and male recombination rates among closely related species remains poorly documented.
 Direct contrasts between the two sexes across a common, diverse set of genomic backgrounds that represent recent timescales would reveal whether the genome-wide recombination rate evolves differently in males and females.
+
 Examining variation in the total number of crossovers in a sex-specific manner could also illuminate evolutionary connections between recombination rate and crossover positioning.
@@
 The degree to which a crossover reduces the probability of another crossover nearby (crossover interference) also differs between females and males [@ottoPaysuer2019].
+
 The house mouse, Mus musculus, provides a compelling system for understanding how sex determines the evolution of recombination.
@@
 
-<tried scientific_10 to make the scientific format of p values look better, haven't gotten it to work>
-
 The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; Musc p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, Mol p =  `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`), highlighting strain and sex as salient variables.
@@
 musculus* (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value` -- 0.65).
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value` -- 0.84; long: p = -- 0.19).
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value` -- 0.84; long: p = `r Q2long.biv_t.test_High.Low$p.value` -- 0.19).
 In a model with total SC length as the dependent variable (M4), the two subspecies effects are significant (*M.
 m.
-musculus* p = -- 3.95 x 10-7; *M.
+musculus* p = `r sum_totSC.male.glm.M4$coefficients[,4][2]` -- 3.95 x 10-7; *M.
 m.
-molossinus* p = `r Q2long.biv_t.test_High.Low$p.value` -- 3.33 x 10-7), but there are also strain-specific effects (Supplemental Table 6).
+molossinus* p = `r sum_totSC.male.glm.M4$coefficients[,4][3]` -- 3.33 x 10-7), but there are also strain-specific effects (Supplemental Table 6).
 In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) (Supplemental Table 9), but they are not consistent across models.
@@
 At the genetic level, the sex-specific patterns we documented indicate that some mutations responsible for the evolution of recombination rate have dissimilar phenotypic effects in the two sexes.
-A subset of the genetic variants associated with genome-wide recombination rate within populations of humans (Kong et al., 2004, 2008, 2014; Halldorsson et al., 2019), Soay sheep (Johnston et al., 2016), and cattle (Ma et al., 2015; Shen et al., 2018) [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle; @Shen2018_cattle] appear to show sex-specific properties, including opposite effects in females and males.
-Furthermore, inter-sexual correlations for recombination rate are weak in humans [@fledel2011] (Fledel-Alon et al., 2011) and Soay sheep [@johnston2016_soay] (Johnston et al., 2016).
-Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011; @Wang2017island; @wang2019_SC] (Dumont and Payseur, 2011; Wang et al., 2019; Wang and Payseur, 2017).
+A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle; @Shen2018_cattle] appear to show sex-specific properties, including opposite effects in females and males.
+Furthermore, inter-sexual correlations for recombination rate are weak in humans [@fledel2011] and Soay sheep [@johnston2016_soay].
+Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011murid; @Wang2017island; @wang2019_SC].
 These variants could differentially affect females and males at any step in the recombination pathway.
 Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contribute to sex-specific evolution in the recombination rate.
-A study of two classical inbred strains and one wild-derived inbred strain of house mice also found a positive association between crossover number and double-strand break number in males [@baier2014] (Baier et al., 2014).
+A study of two classical inbred strains and one wild-derived inbred strain of house mice also found a positive association between crossover number and double-strand break number in males [@baier2014].
 
 Another implication of our results is that the connection between recombination rate and fitness differs between males and females.
-Little is known about whether and how natural selection shapes recombination rate in nature [@DapperPayseur2017; @Ritz2017] (Dapper and Payseur, 2017; Ritz et al., 2017).
-@samuk2020 Samuk et al.
-(2020) recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of Drosophila pseudoobscura was caused by natural selection.
+Little is known about whether and how natural selection shapes recombination rate in nature [@DapperPayseur2017; @Ritz2017].
+@samuk2020 recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of Drosophila pseudoobscura was caused by natural selection.
 Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
 
-Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling; @sardell_sex_2020] (Brandvain and Coop, 2012; Sardell and Kirkpatrick, 2020).
-Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] (Lenormand, 2003) or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020] (Sardell and Kirkpatrick, 2020).
-Another modifier model showed that meiotic drive could stimulate female-specific evolution of the recombination rate [@brandvain2012scrambling] (Brandvain and Coop, 2012).
+Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling; @sardell_sex_2020].
+Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020].
+Another modifier model showed that meiotic drive could stimulate female-specific evolution of the recombination rate [@brandvain2012scrambling].
 Although these models fit the conserved pattern of sex differences in crossover positions, they do not readily explain our observations of sex-specific evolution in the genome-wide recombination rate.
@@
 We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
-During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@dumontDesai2012; @LaneKauppi2019; @vanVeen2003; @subramanian2014] (Dumont and Desai, 2012; Lane and Kauppi, 2019; Subramanian and Hochwagen, 2014; VanVeen and Hawley, 2003).
-The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesion holding homologs together [@LaneKauppi2019] (Lane and Kauppi, 2019).
+During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@dumontDesai2012; @LaneKauppi2019; @vanVeen2003; @subramanian2014].
+The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesion holding homologs together [@LaneKauppi2019].
 Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and chromosome separation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
-Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019] (Lane and Kauppi, 2019), perhaps due to the presence of the acentrosome spindle [@So2019] (So et al., 2019) and larger cell volume [@kyogoku2017] (Kyogoku and Kitajima, 2017) in oocytes.
+Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019], perhaps due to the presence of the acentrosome spindle [@So2019] and larger cell volume [@kyogoku2017] in oocytes.
 The higher stringency of the SAC during spermatogenesis suggests that selection will be better at removing mutations that interfere with bi-orientation in males than in females.
@@
 Our SAC model is consistent with other features of our data.
-We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] (Sardell and Kirkpatrick, 2020) apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
-Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell [@kudo2009] (Kudo et al., 2009), whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the meiosis I spindle, especially for those organisms that undergo dictyate arrest [@Lee2019] (Lee, 2019).
+We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
+Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell [@kudo2009], whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the meiosis I spindle, especially for those organisms that undergo dictyate arrest [@Lee2019].
 
@@
 Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and may improve the fidelity of chromosomal segregation.
-While the SAC model postulates direct fitness effects of interference, a modifier model predicted that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993] (Goldstein et al., 1993).
+While the SAC model postulates direct fitness effects of interference, a modifier model predicted that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
 
 Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
-The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle [@ma2015_cattle] (Ma et al., 2015).
-In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@ottoPaysuer2019; @segura2013] (Otto and Payseur, 2019; Segura et al., 2013).
+The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle [@ma2015_cattle].
+In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@ottoPaysuer2019; @segura2013].
 The evolution of crossover interference and its relationship to changes in crossover number on the genomic scale is a topic deserving of more empirical and theoretical work.
@@
 Female house mice consistently show longer SCs, even in strains with more recombination in males.
-Studies in mice [@lynn2002; @petkov2007]  (Lynn et al., 2002; Petkov et al., 2007) and humans [@gruhn2013; @tease2004] (Gruhn et al., 2013; Tease and Hulten, 2004) suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
-Some authors have suggested that conserved sex differences in crossover positioning (more uniform placement in females) and interference strength (stronger interference in males) could be due to looser chromatin packing of the meiotic chromosome structure in females [@haenel2018; @petkov2007] (Haenel et al., 2018; Petkov et al., 2007).
-A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM] (Hultén, 2011).
-Recent work in mice connects the sparser recombination landscape in females to sex differences in the crossover maturation efficiency [@wang2017inefficient] (Wang S.
-et al., 2017).
+Studies in mice [@lynn2002; @petkov2007] and humans [@gruhn2013; @tease2004] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
+Some authors have suggested that conserved sex differences in crossover positioning (more uniform placement in females) and interference strength (stronger interference in males) could be due to looser chromatin packing of the meiotic chromosome structure in females [@haenel2018; @petkov2007].
+A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM].
+Recent work in mice connects the sparser recombination landscape in females to sex differences in the crossover maturation efficiency [@wang2017inefficient].
 
 Our conclusions are accompanied by several caveats.
-First, MLH1 foci only identify interfering crossovers [@holloway2008mus81] (Holloway et al., 2008).
-Although most crossovers belong to this class [@holloway2008mus81] (Holloway et al., 2008), our approach likely underestimated genome-wide recombination rates.
+First, MLH1 foci only identify interfering crossovers [@holloway2008mus81].
+Although most crossovers belong to this class [@holloway2008mus81], our approach likely underestimated genome-wide recombination rates.
 Evolution of the number of non-interfering crossovers is a subject worth examining.
@@
 
-While the causes of sex differences in recombination remain mysterious  @lenormand2016 (Lenormand et al., 2016), our conclusions have implications for a wide range of recombination research.
+While the causes of sex differences in recombination remain mysterious  @lenormand2016, our conclusions have implications for a wide range of recombination research.
 For biologists uncovering the cellular and molecular determinants of recombination, our results suggest that mechanistic differences between the sexes could vary by genetic background.
@@
 
+This research was funded by NIH grants R01GM120051 and R01GM100426 to B. A. P.. A. L. P. was partly supported by NIH T32GM007133. Generous Franscico Peligri for microscope.
+
 # Competing interests
 
+The authors declare that there are no competing interests.
+
 # REFERENCES

### 2020-07-01 doc/MS_v4.Rmd
some older updates

@@
 
--new-
 Although meiotic recombination is required for successful gametogenesis in most species that reproduce sexually, the rate of crossing over varies among individuals.
@@
 # INTRODUCTION
-
--new-
 
@@
 
-Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle], and Soay sheep [@johnston2016_soay], in which the role of sex is confounded with the contributions of genetic variation.
+Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004; @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle], and Soay sheep [@johnston2016_soay], in which the role of sex is confounded with the contributions of genetic variation.
 Although it is known that the level and direction of heterochiasmy can differ among species [@brandvain2012scrambling; @lenormandDuthiel_2005], the correlation between female and male recombination rates among closely related species remains poorly documented.
@@
 Examining variation in the total number of crossovers in a sex-specific manner could also illuminate evolutionary connections between recombination rate and crossover positioning.
-Analyses of meiotic chromosome morphology in Arabidopsis thaliana, Caenorhabditis elegans, and Mus musculus suggest that the sex with more recombination usually has longer chromosome axes [@CahoonLibuda2019].
+Analyses of meiotic chromosome morphology in _Arabidopsis thaliana_, _Caenorhabditis elegans_, and _Mus musculus_ suggest that the sex with more recombination usually has longer chromosome axes [@CahoonLibuda2019].
 A survey of 51 species found conserved sex differences in the recombination landscape, including telomere-biased placement of crossovers in males but not in females [@sardell_sex_2020].
@@
 
--new-
-
 We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` -- 14 wild-derived inbred strains sampled from three subspecies of house mice ( _M.
@@
 The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; musculus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` -- 0.24, molossinus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` -- 0.1), highlighting strain and sex as salient variables.
-Two strains exhibit strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` -- 1.78 x 10-6; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]` -- 0.02), with sex-strain interactions involving three strains (M3; *domesticus^G^* p < x 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p < x 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` -- 3.87 x 10-4).
+Two strains exhibit strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` -- 1.78 x 10-6; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]` -- 0.02), with sex-strain interactions involving three strains (M3; *domesticus^G^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` -- 3.87 x 10-4).
+**fix the pvalue above, change the 0 to p < 10^-6**
 
-In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)` -- 71), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` -- 3.6 x 10-7; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` -- 6.11 foci **(add CI’s here)**, *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` -- 6.3 x 10-9; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`-- 6.91), and *musculus^SKIVE^* (p = -- 8.22 x 10-4; effect = -- 4.04).
+In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)` -- 71), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` -- 3.6 x 10-7; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` -- 6.11 foci **No CI's for these effect values**, *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` -- 6.3 x 10-9; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`-- 6.91), and *musculus^SKIVE^* (p = -- 8.22 x 10-4; effect = -- 4.04).
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
@@
 
--old-
-
-We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice (*M.
-m.
-domesticus*, _M.
-m.
-musculus_ and _M.
-m.
-molossinus_ ) and three additional species of Mus (*M.
-spretus*, _M.
-spicilegus_ , and *M.
-caroli*).
-Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
-
-Graphical comparisons reveal sex-specific dynamics to the evolution of genome-wide recombination rate (Figure 1A).
-First, MLH1 focus counts differ between females and males in most strains.
-Second, the difference in counts between the sexes varies among strains.
-Although most strains show more MLH1 foci in females, two strains (*musculus^PWD^* and _molossinus^MSM^_) exhibit higher counts in males.
-In females, numbers of MLH1 foci are evenly distributed around the sex-wide mean of approximately 25 (Figure 1B).
-In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23) of foci (Figure 1C).
-Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = `r sperman.corr.FM`; p = `r sperman.corr.FM.pval$p.value`) across the set of strains.
-
-To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (Table 2; detailed results available in Supplemental Table X).
-Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`; sex: p = `r MLH1.MMlmer_results$sex.results`; subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
-
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; Musc p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, Mol p =  `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`), highlighting strain and sex as salient variables.
-Two strains exhibit particularly strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
-
-In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)`), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` ), *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`), and *musculus^SKIVE^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` ).
-These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-In females (M4; n= `r length(female_mouse.avs_for.analysis$mouse)`), four strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]`), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]`), *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]`), and *musculus^PWD^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[4]`).
-Strain effect sizes in females are modest in magnitude compared to those in males.
-Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.
-
 ## Synaptonemal complexes are longer in females
-
--new-
 
@@
 
--old-
-
-The variation in sex differences in recombination we discovered provided an opportunity to determine whether sex differences in chromatin compaction, as measured by the length of the synaptonemal complex (SC), are reversed when heterochiasmy is reversed.
-In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = `r Q1_short.SC_t.test.SKIVE$p.value`).
-
-Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from `r Q1SC.mus.av.short.biv_DIFF_PWD`  (*musculus^PWD^*) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (*domesticus^WSB^*) across strains.
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results` ; M2: p =
-`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =`r sum.Q1.glm_totSC_M3$coefficients[,4][8]` ) and short bivalent length (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]/10000`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]/10000`).
-The existence of some subspecies and strain effects on total SC and short bivalent SC Length further indicates that SC length has evolved among strains and among subspecies (See Supplemental Tables 5 and 8).
-In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males recombine more.
-This pattern implies that in high-recombination strains, spermatocytes have less space than oocytes in which to position additional crossovers.
-
 ## Females and males differ in crossover positions and crossover interference
-
--new-
 
@@
 In summary, controlling for differences in SC length (chromatin compaction) indicates that interference is consistently stronger in males, whereas interference on the physical scale is similar in the two sexes.
-
--old-
-
-We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; paired t-test; p = `r Q1.strain.F1.Pos$p.value`).
-Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
-
-Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
-Examination of IFD~norm~ distributions indicates that females are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Models treating IFD~norm~ as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` ; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
-In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) (Supplemental Table 13 and 14) across the full set of strains, whether they are considered separately or together.
-Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 4C) further suggests that interference distances vary more in females than in males, and that males display a stronger telomeric bias in the placement of the distal crossover.
-In summary, controlling for differences in SC length (chromatin compaction) indicates that interference is stronger in males, whereas interference on the physical scale is similar in the sexes.
 
@@
 ```
-
--new-
 
@@
 
--old-
-
-We used the contrast between males from high-recombination strains and males from low-recombination strains to identify features of the recombination landscape associated with evolutionary transitions in the genome-wide recombination rate.
-We considered proportions of bivalents with different numbers of crossovers, double-strand break number, SC length, and crossover positioning.
-
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")` ) have either one or two MLH foci (Supplemental Figure 2).
-The proportions of single-focus (1CO) bivalents vs.
-double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Supplemental Figure 2).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_ , and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
-Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-*molossinus^MSM^*: p = `r PWD.MSM.2CO.prop.test$p.value` ; _musculus^PWD^_ vs.
-_musculus^KAZ^_: p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value` ).
-These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number across multiple bivalents.
-
-To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains (*musculus^PWD^* and *molossinus^MSM^*) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
-High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < `r 1/1000000` ).
-In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval` ).
-Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
-High-recombination and low-recombination strains do not differ in the MLH1/DMC1 ratio, whether DMC1 foci were counted in early zygotene cells or late zygotene cells (t-test; p > 0.05).
-These results raise the possibility that the evolution of genome-wide recombination rate is primarily determined by processes that precede the CO/NCO decision, at least in house mice.
-
-Total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 3).
-Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within *M.
-m.
-molossinus* (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`) but not within *M.
-m.
-musculus* (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`  ; long: p = `r Q2long.biv_t.test_High.Low$p.value` ).
-In a model with total SC length as the dependent variable (M4), the two subspecies effects are significant (*M.
-m.
-musculus* p = `r sum_totSC.male.glm.M4$coefficients[,4][2]` and *M.
-m.
-molossinus*, p = `r sum_totSC.male.glm.M4$coefficients[,4][3]`), however there are also strain specific effects (Supplemental Table 6).
-
-In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) (Supplemental Table 9), but they are not consistent across models.
-Collectively, these results reveal that evolution of SC length is not strongly associated with evolution of genome-wide recombination rate in house mice.
-In summary, evolution of the genome-wide recombination rate in males is connected to double-strand break number and crossover interference, but not to SC length and crossover position (on single-crossover bivalents).
-
 # DISCUSSION
-
--new-
 
@@
 
--old-
-
-By comparing recombination rates in females and males from the same diverse set of genetic backgrounds, we isolated sex as a primary factor in the evolution of this fundamental meiotic trait.
-Recombination rate differences are more pronounced in males than females.
-Because inter-strain divergence times are identical for the two sexes, this observation demonstrates that the genome-wide recombination rate evolves faster in males, at least in house mice.
-More generally, recombination rate divergence is decoupled in females and males.
-These disparities are remarkable given that recombination rates for the two sexes were measured in identical genomic backgrounds (other than the presence/absence of the Y chromosome).
-Our results provide the strongest evidence yet that the genome-wide recombination rate follows distinct evolutionary trajectories in males and females.
-
-At the genetic level, the sex-specific evolution we documented indicates that some mutations responsible for divergence in recombination rate have dissimilar phenotypic effects in females and males.
-A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle; @Shen2018_cattle] appear to show sex-specific properties, including opposite effects in females and males.
-Furthermore, inter-sexual correlations for recombination rate are weak in humans [@fledel2011] and Soay sheep [@johnston2016_soay].
-Crosses between the strains we surveyed could be used to identify and characterize the genetic variants responsible for recombination rate evolution in house mice [@dumont2011; @Wang2017island; @wang2019_SC].
-These variants could differentially affect females and males at any step in the recombination pathway.
-Although our DMC1 profiling was limited to males from a small number of strains (for practical reasons), our findings suggest that mutations that determine the number of double-strand breaks contribute to sex-specific evolution in the recombination rate.
-A study of two classical inbred strains and one wild-derived inbred strain of house mice also found a positive association between crossover number and double-strand break number in males [@baier2014].
-
-Another implication of our results is that the connection between recombination rate and fitness differs between males and females.
-Little is known about whether and how natural selection shapes recombination rate in nature [@DapperPayseur2017; @Ritz2017].
-@samuk2020 recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of *Drosophila pseudoobscura* was caused by natural selection.
-Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
-
-Population genetic models have been built to explain sexual dimorphism in the number and placement of crossovers, which is a common phenomenon [@brandvain2012scrambling; @sardell_sex_2020].
-Modifier models predicted that lower recombination rates in males will result from haploid selection [@lenormand2003] or sexually antagonistic selection on coding and cis-regulatory regions of genes [@sardell_sex_2020].
-Another modifier model showed that meiotic drive could stimulate female-specific evolution of the recombination rate [@brandvain2012scrambling].
-Although these models fit the conserved pattern of sex differences in crossover positions, they do not readily explain our observations of sex-specific evolution in the genome-wide recombination rate.
-In particular, the alternation across strains in which sex has more crossovers is unexpected.
-
-We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
-During meiosis, achieving a stable chromosome structure requires the attachment of kinetochores to opposite poles of the cell and at least one crossover to create tension across the sister chromosome cohesion distal to chiasmata [@LaneKauppi2019; @vanVeen2003; @subramanian2014; @dumontDesai2012].
-The spindle assembly checkpoint (SAC) prevents aneuploidy by ensuring that all bivalents are correctly attached to the microtubule spindle (“bi-oriented”) before starting the metaphase-to-anaphase transition via the release of the sister cohesion holding homologs together [@LaneKauppi2019].
-Hence, selection seems likely to favor mutations that optimize the process of bi-orientation and chromosome separation, thereby prohibiting the SAC from delaying the cell cycle or triggering apoptosis.
-Multiple lines of evidence indicate that the SAC is more effective in spermatogenesis than in oogenesis [@LaneKauppi2019], perhaps due to the presence of the centrosome spindle [@So2019] and larger cell volume [@kyogoku2017] in oocytes.
-The higher stringency of the SAC during spermatogenesis suggests that selection will be better at removing mutations that interfere with bi-orientation in males than in females.
-Therefore, faster male evolution of the genome-wide recombination rate could be driven by the more stringent SAC acting on chromosome structures at the metaphase I alignment.
-
-Our SAC model is consistent with other features of our data.
-We showed that widespread sex differences in broad-scale crossover positioning [@sardell_sex_2020] apply across house mice, even in lineages where the direction of heterochiasmy is reversed.
-Faster spermatogenesis may select for synchronization of the separation across all homologs within the cell [@kudo2009], whereas in oogenesis, the slower cell cycle and multiple arrest stages may require chromosome structures with greater stability on the MI spindle, especially for those organisms that undergo dictyate arrest [@Lee2019].
-
-We propose that the SAC model also can explain the correlated evolution of stronger crossover interference and higher genome-wide recombination rate in male house mice.
-Our results show that crossovers are spaced further apart in strains enriched for double-crossover bivalents when SC length is taken into account and chromosome size effects are minimized.
-Assuming chromatin compaction between (prophase) pachytene and metaphase is uniform along bivalents, this increased spacing is expected to expand the area for sister cohesion to connect homologs and may improve the fidelity of chromosomal segregation.
-While the SAC model postulates direct fitness effects of interference, a modifier model predicted that indirect selection on recombination rate – via its modulation of offspring genotypes – can strengthen interference as well [@goldstein1993].
-
-Regardless of the underlying mechanism, our results provide a rare demonstration that crossover interference can diverge over short evolutionary timescales.
-The notion that stronger interference can co-evolve with higher genome-wide recombination rate is supported by differences between breeds of cattle (@ma2015_cattle) and differences between wild and captive white-footed mice [@peterson2019].
-In contrast, mammalian species with stronger interference tend to exhibit lower genome-wide recombination rates [@segura2013; @ottoPaysuer2019].
-Collectively, these patterns suggest that inferences about the evolutionary dynamics of interference depend on the timescale under consideration.
-
-Our findings further reveal that evolution of the genome-wide recombination rate does not require major changes in the degree of chromatin compaction.
-Female house mice consistently show longer SCs, even in strains with more recombination in males.
-Studies in mice [@lynn2002; @petkov2007] and humans [@gruhn2013; @tease2004] suggest that chromosomal axes are longer (and DNA loops are shorter) in females than males.
-Some authors have suggested that conserved sex differences in crossover positioning (more uniform placement in females) and interference strength (stronger interference in males) could be due to looser chromatin packing of the meiotic chromosome structure in females [@haenel2018; @petkov2007].
-A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM].
-Recent work in mice connects the sparser recombination landscape in females to sex differences in the crossover maturation efficiency [@wang2017inefficient].
-
-Our conclusions are accompanied by several caveats.
-First, MLH1 foci only identify interfering crossovers [@holloway2008mus81].
-Although most crossovers belong to this class [@holloway2008mus81], our approach likely underestimated genome-wide recombination rates.
-Evolution of the number of non-interfering crossovers is a topic worth examining.
-A second limitation is that our investigation of crossover locations was confined to the relatively low resolution possible with immunofluorescent cytology.
-Positioning crossovers with higher resolution could reveal additional evolutionary patterns.
-Finally, the panel of inbred lines we surveyed may not be representative of recombination rate variation within and between subspecies of house mice.
-We considered most available wild-derived inbred lines, but house mice have a broad geographic distribution.
-Nevertheless, we expect our primary conclusion that recombination rate evolves in a sex-specific manner to be robust to geographic sampling because differences between females and males exist for the same set of inbred strains.
-
-While the causes of sex differences in recombination remain mysterious @lenormand2016, our conclusions have implications for a wide range of recombination research.
-For biologists uncovering the cellular and molecular determinants of recombination, our results suggest that mechanistic differences between the sexes could vary by genetic background.
-For researchers charting the evolutionary trajectory of recombination, our findings indicate that sex-specific comparisons are crucial.
-For theoreticians building evolutionary models of recombination, different fitness regimes and genetic architectures in females and males should be considered.
-Elevating sex as a primary determinant of recombination would be a promising step toward integrating knowledge of cellular mechanisms with evolutionary patterns to understand recombination rate variation in nature.
-
---------
-
 # MATERIALS AND METHODS
@@
 
-<not many edits to replace>
+**not many edits to replace, bbut double check**
 

### 2020-07-01 doc/Tables.Rmd
some older updates

---
title: "MS_Tables"
author: "April Peterson"
date: "6/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xlsx)
library(knitr)
library(kableExtra)
#read in excel plates
Table1 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 1)

Table2 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 2)

Table3 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 3)

Table4 = read.xlsx2("~./MLH1repo/doc/MainTables.xlsx", header=TRUE, 4)

#Supplemental Table
```

```{r table1, echo=FALSE}

#change the headers

colnames(Table1) <- c("Species","Official name","abbreviation for this paper","Geographic Origin","Source")
print.tex <- kable(Table1, "latex")
```

```{r tab2, echo=FALSE}
print.tex2 <- kable(Table2, "latex")
```

```{r table3, echo=FALSE}

#arg tha numbers are factors!!!!
Table3$Number.of.Mice <- as.character(Table3$Number.of.Mice)
Table3$Number.of..Cells <- as.character(Table3$Number.of..Cells)
Table3$Mean.MLH1.Count <- as.character(Table3$Mean.MLH1.Count)
Table3$se <- as.character(Table3$se)
Table3$cV <- as.character(Table3$cV)
Table3$var <- as.character(Table3$var)

Table3$Number.of.Mice <- as.numeric(Table3$Number.of.Mice)
Table3$Number.of..Cells <- as.numeric(Table3$Number.of..Cells)
Table3$Mean.MLH1.Count <- as.numeric(Table3$Mean.MLH1.Count)
Table3$se <- as.numeric(Table3$se)
Table3$cV <- as.numeric(Table3$cV)
Table3$var <- as.numeric(Table3$var)

#MLH1 summary stats
print.tex3 <- kable(Table3, "latex",booktabs = T, digits = 2)

write.table(kable(Table3, "latex",booktabs = T, digits = 2), "~./MLH1repo/Table3.mean.table.tex", sep=",",
            row.names = FALSE)

```

```{r table4, echo=FALSE}
#DMC1 table

colnames(Table4) <- c("Recombination Group","Strain",
                      "Mean DMC1 Foci per Cell","MLH1:DMC1   ratio","MLH1:DMC1   ratio",
                     "Mean DMC1 Foci per Cell","MLH1:DMC1   ratio","MLH1:DMC1   ratio")

#DMC1
print.tex4 <- kable(Table4, "latex", booktabs = T) %>%
kable_styling() %>%
add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3) )

write.table(print.tex4, "~./MLH1repo/DMC1.mean.table.tex")

save_kable(print.tex4, "~./MLH1repo/kable.DMC1.mean.table.html", keep_tex = TRUE)

#kable(dt, "latex", booktabs = T) %>%
#kable_styling() %>%
#add_header_above(c(" " = 1, "Group 1" = 2, "Group 2" = 2, "Group 3" = 2))

#write the kable file

```

\begin{tabular}{S l l l r}
\toprule
{Speed (mph)} & Driver          & Car                        & Engine    & Date     \\
\midrule
407.447     & Craig Breedlove & Spirit of America          & GE J47    & 8/5/63   \\
413.199     & Tom Green       & Wingfoot Express           & WE J46    & 10/2/64  \\
434.22      & Art Arfons      & Green Monster              & GE J79    & 10/5/64  \\
468.719     & Craig Breedlove & Spirit of America          & GE J79    & 10/13/64 \\
526.277     & Craig Breedlove & Spirit of America          & GE J79    & 10/15/65 \\
536.712     & Art Arfons      & Green Monster              & GE J79    & 10/27/65 \\
555.127     & Craig Breedlove & Spirit of America, Sonic 1 & GE J79    & 11/2/65  \\
576.553     & Art Arfons      & Green Monster              & GE J79    & 11/7/65  \\
600.601     & Craig Breedlove & Spirit of America, Sonic 1 & GE J79    & 11/15/65 \\
622.407     & Gary Gabelich   & Blue Flame                 & Rocket    & 10/23/70 \\
633.468     & Richard Noble   & Thrust 2                   & RR RG 146 & 10/4/83  \\
763.035     & Andy Green      & Thrust SSC                 & RR Spey   & 10/15/97\\
\bottomrule
\end{tabular}

### 2020-07-01 src/Result_Setup.Rmd
some older updates

@@
 
-```{r organizing.data, echo=FALSE, eval=FALSE}
-#set the order of another column, based on another variable. consistant order for ploting
-
-#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
-#-ene endings refer to the cells/units within the cell stage? (I think)
-
-#apply this before ploting
-MLH1_data <- MLH1_data %>%
-  arrange(strain, sex, category, mouse) %>%
-  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
-MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])
-
-# sort your dataframe, by the focal categories
-MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))
-#use the above line for when order is out of wak
-
-#MLH1_data <- with(MLH1_data, MLH1_data[order(category),])
-#AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)
-
-AP_category_table <-  ddply(.data=AP_mouse_table.HQ,
-                 .(strain,sex),
-                 summarize,
-                 sum.ncells = sum(Ncells),
-                 mean.MLH1 = mean(mean_co),
-                 var.MLH1 = var(mean_co),
-                 AV.var_MLH1 = mean(var, na.rm=TRUE)
-
-)
-
-AP_mouse_table.HQ_target <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]
-
-remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC")
-AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
-```
-
 ```{r Q1.rbar.gamma, echo=FALSE, eval=FALSE}
@@
 
-kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+#kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")
+
+#write
 
@@
 sum_IFD.PER.male.glm_M5 <- summary(IFD.PER.male.glm_M5)
-write.xlsx(as.data.frame( sum_IFD.PER.female.glm_M5$coefficients), "IFD.PER.male.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
+
+write.xlsx(as.data.frame( sum_IFD.PER.male.glm_M5$coefficients), "IFD.PER.male.glm_M5.xlsx", col.names = TRUE, row.names = TRUE)
 

### 2020-07-01 src/CommonFunc_MLH1repo.R

@@ -454,6 +454,10 @@ add_umeter <- function(dat){
 }
 
 
+scientific_10 <- function(x) {
+  parse(text=gsub("e", " %*% 10^", scales::scientific_format()(x)))
+}
+
 
 #calculate Interchromsomal rbar
 #tested with values from Veller et al 2018

### 2020-07-01 doc/MS_v4.Rmd
citation for physical linkage

@@
 Meiosis converts diploid germ cells into haploid gametes.
-During meiosis I, DNA crossovers aid the separation of homologous chromosomes by physically linking them and establishing tension between them on the spindle **(REF?)**.
+During meiosis I, DNA crossovers aid the separation of homologous chromosomes by physically linking them and establishing tension between them on the spindle [@petronczki2003].
 The wrong number of recombination events can disrupt chromosomal segregation, leading to infertility, miscarriage, and birth defects [@hassold2001].
@@
 caroli_ ).
-Mean MLH1 focus counts for `r sum(main.table$Nmice)` -- 188 mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` -- 21.77 spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` -- 1,742 spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` -- 17.85 oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` -- 1,427 oocytes).
-**<insert statement for cell sample size number>**
+Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
+**insert statement for cell sample size number**
 
@@
 In stark contrast, males largely separate into two groups of strains with high numbers (near 30) and low numbers (near 23) of foci (Figure 1C).
-Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = `r sperman.corr.FM` -- 0.08; p = `r sperman.corr.FM.pval$p.value` -- 0.84) across the set of strains.
+Strain mean MLH1 focus counts from females and males are uncorrelated (Spearman’s $\rho$ = `r sperman.corr.FM`; p = `r sperman.corr.FM.pval$p.value`) across the set of strains.
 
-To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` -- 137 house mice (from three subspecies; _M.
+To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (from three subspecies; _M.
 m.
@@
 molossinus_) (Table 2; detailed results available in Supplemental Table X).
-Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`-- 10-4; sex: p = `r MLH1.MMlmer_results$sex.results` -- 3.64 x 10-6; subspecies: p = `r MLH1.MMlmer_results$sub.sp` -- 9.69 x 10-4; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval` -- 1.8 x 10-4).
+Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`-- 10-4; sex: p = `r MLH1.MMlmer_results$sex.results`; subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
 
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; musculus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]` -- 0.24, molossinus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]` -- 0.1), highlighting strain and sex as salient variables.
-Two strains exhibit strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` -- 1.78 x 10-6; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]` -- 0.02), with sex-strain interactions involving three strains (M3; *domesticus^G^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` -- 3.87 x 10-4).
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; musculus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, molossinus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`), highlighting strain and sex as salient variables.
+Two strains exhibit strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
 **fix the pvalue above, change the 0 to p < 10^-6**
 
-In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)` -- 71), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` -- 3.6 x 10-7; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` -- 6.11 foci **No CI's for these effect values**, *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` -- 6.3 x 10-9; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`-- 6.91), and *musculus^SKIVE^* (p = -- 8.22 x 10-4; effect = -- 4.04).
+In separate analyses of males (M4; n = `r length(male_mouse.avs_for.analysis$mouse)`), three strains disproportionately shape MLH1 focus count (as observed in Figure 1C): *musculus^PWD^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]`; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` foci, *molossinus^MSM^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[8]`), and *musculus^SKIVE^* (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).
 These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains.
-In females (M4; n= `r length(female_mouse.avs_for.analysis$mouse)` --76), three strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]` -- 8.7 x 10-6; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]` -- 3.3), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]` -- 2.43 x 10-5; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]` -- 2.99), and *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]`  -- 0.03; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]` -- 1.69).
+In females (M4; n= `r length(female_mouse.avs_for.analysis$mouse)` ), three strains affect MLH1 focus count: *domesticus^G^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[2]`), *molossinus^MSM^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[5]`), and *domesticus^LEW^* (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]`; effect = `r sum_MLH1.female.spc.glm.M5$coefficients[3]`).
 Strain effect sizes in females are modest in magnitude compared to those in males.
@@
 The variation in sex differences in recombination we discovered provided an opportunity to determine whether sex differences in chromatin compaction, as measured by the length of the synaptonemal complex (SC), are reversed when heterochiasmy is reversed.
-In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = `r Q1_short.SC_t.test.SKIVE$p.value` -- 0.11).
-Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from `r Q1SC.mus.av.short.biv_DIFF_PWD` -- 1.26 (*musculus^PWD^*) to `r Q1SC.mus.av.short.biv_DIFF_WSB` -- 1.52 (*domesticus^WSB^*) across strains.
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results` -- 2.56 x 10-31; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,4][8]` -- 2.56 x 10-8; M3: p =  `r sum.Q1.glm_totSC_M3$coefficients[,4][8]` -- 2.56 x 10-8) and short bivalent SC length (M1: p = `r Q1.lmer_short_M1_results$sex.results` -- 1.12 x 10-11; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]/10000` -- 1.33 x 10-7; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]/10000` -- 1.33 x 10-7).
+In all strains except _musculus^SKIVE^_, females have longer SCs than males, whether SC length was estimated as the total length across bivalents or as the length of short bivalents (t-tests; all p < 0.05, except short bivalents in _musculus^SKIVE^_, p = `r Q1_short.SC_t.test.SKIVE$p.value`).
+Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from `r Q1SC.mus.av.short.biv_DIFF_PWD` (*musculus^PWD^*) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (*domesticus^WSB^*) across strains.
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =  `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`) and short bivalent SC length (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`).
 The existence of some subspecies and strain effects on total SC length and short bivalent SC length (Supplemental Tables 5 and 8) further indicates that SC length has evolved among strains and among subspecies.
+
 In summary, two approaches for measuring SC length demonstrate that females have longer SCs (chromosome axes), even in strains in which males recombine more.
@@
 We used normalized positions of MLH1 foci along bivalents with a single focus to compare crossover location while controlling for differences in SC length.
-In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` -- 0.68 **add s.d here?**; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )` --  0.56; paired t-test; p = `r Q1.strain.F1.Pos$p.value` -- 8.49 x 10-4).
-Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results` -- 2.82 x 10-26; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]` -- 3.96 x 10-8; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]` -- 3.96 x 10-8).
+In all strains, MLH1 foci tend to be closer to the telomere in males (mean normalized position in males: `r mean( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; mean normalized position in females: `r mean( (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos, na.rm = TRUE )`; paired t-test; p = `r Q1.strain.F1.Pos$p.value`).
+Sex is also the strongest determinant of MLH1 focus position in the models we tested (M1: p = `r lmer.Q1.nrm1CO_M1_results$sex.results`; M2: p = `r summry.Q1.glm_F1pos_M2$coefficients[,4][8]`; M3: p = `r summry.Q1.glm_1CO.pos_M3$coefficients[,4][8]`).
 
-Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = `r Q1.IFD.PER_t.test_KAZ$p.value` -- 0.33).
+Males have longer normalized mean inter-focal distances (IFD~norm~) than females in seven out of eight strains (t-tests; p < 0.05), with only _musculus^KAZ^_ showing no difference (p = `r Q1.IFD.PER_t.test_KAZ$p.value`).
 Examination of IFD~norm~ distributions indicates that females are centered at approximately 50% and show a slight enrichment of low (<25%) values, whereas males are enriched for higher values.
-Models treating IFD~norm~ as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results` -- 9.08 x 10-12; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]` -- 0.01; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]` -- 0.01).
+Models treating IFD~norm~ as the dependent variable support the inference of stronger interference in males, with sex being the most significant variable (M1: p = `r Q1.IFD.PER_M1_lmer.results$sex.results`; M2: p = `r summry.glm.Q1_IFD.PER.M2$coefficients[,4][4]`; M3: p = `r summry.glm.Q1_IFD.PER.M3$coefficients[,4][2]`).
 In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) (Supplemental Table 13 and 14) across the full set of strains, whether they are considered separately or together.
@@
 
-Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")`-- 9,569) have either one or two MLH foci (Supplemental Figure 2).
+Ninety-six percent of single bivalents in our pooled dataset (n = `r formatC( sum( Chrm.prop_strain_table$n0CO, Chrm.prop_strain_table$n1CO, Chrm.prop_strain_table$n2CO, Chrm.prop_strain_table$n3CO, Chrm.prop_strain_table$n4CO ), format="d", big.mark=",")`) have either one or two MLH foci (Supplemental Figure 2).
 The proportions of single-focus (1CO) bivalents vs.
 double-focus (2CO) bivalents distinguish high-recombination strains from low-recombination strains (Supplemental Figure 2).
-High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` -- 0.33 in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` -- 0.44 in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` -- 0.51 in _molossinus^MSM^_ (Supplemental Figure 3).
+High-recombination strains are enriched for 2CO bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents are `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "SKIVE"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "SKIVE"]` in _musculus^SKIVE^_, `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "PWD"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "PWD"]` in _musculus^PWD^_, and `r chrm.prop.male.high1$n2CO[chrm.prop.male.high1$strain == "MSM"] / chrm.prop.male.high1$nbiv[chrm.prop.male.high1$strain == "MSM"]` in _molossinus^MSM^_ (Supplemental Figure 3).
 Following patterns in the genome-wide recombination rate, male _musculus^PWD^_ and male _molossinus^MSM^_ have 2CO proportions that are more similar to each other than to strains from their own subspecies (chi-square tests; _musculus^PWD^_ vs.
-_molossinus^MSM^_: p = `r PWD.MSM.2CO.prop.test$p.value` -- 0.37; _musculus^PWD^_ vs.
-*musculus^KAZ^*: p = `r PWD.KAZ.2CO.prop.test$p.value` -- 1.23 x 10-31; _molossinus^MSM^_ vs.
-_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value` -- 2.34 x  10-6).
+_molossinus^MSM^_: p = `r PWD.MSM.2CO.prop.test$p.value`; _musculus^PWD^_ vs.
+*musculus^KAZ^*: p = `r PWD.KAZ.2CO.prop.test$p.value`; _molossinus^MSM^_ vs.
+_molossinus^MOLF^_: p = `r MSM.MOLF.2CO.prop.test$p.value`).
 These results demonstrate that evolution of the genome-wide recombination rate reflects changes in crossover number across multiple bivalents.
@@
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` -- 76 early zygotene and `r sum( (DMC1table.main$ncells.y) )` -- 75 late zygotene spermatocytes from two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
-High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; -- `r 1/1000000`p < 10-6).
-In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval` -- 0.66).
+DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < `r 1/1000000` -- 10-6).
+In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval`).
 Since DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs designated as COs.
@@
 Total SC length only partially differentiates high-recombination strains from low-recombination strains (Figure 3).
-Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value` -- 0.01), separate tests within subspecies show that the two strain categories differ within *M.
+Whereas high-recombination strains as a group have significantly greater total SC length than low-recombination strains (t-test; p = `r Q2tot.sc_t.test_High.Low$p.value`), separate tests within subspecies show that the two strain categories differ within *M.
 m.
-molossinus* (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value` -- 2.59 x 10-4) but not within *M.
+molossinus* (p = `r Q2tot.sc_t.test_High.Low_Mol$p.value`) but not within *M.
 m.
-musculus* (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value` -- 0.65).
-Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value` -- 0.84; long: p = `r Q2long.biv_t.test_High.Low$p.value` -- 0.19).
+musculus* (p = `r Q2tot.sc_t.test_High.Low_Musc$p.value`).
+Additionally, mouse means for the reduced (short and long) bivalent datasets do not differ between high-recombination and low-recombination strains (t-test; short: p = `r Q2short.biv_t.test_High.Low$p.value`; long: p = `r Q2long.biv_t.test_High.Low$p.value`).
 In a model with total SC length as the dependent variable (M4), the two subspecies effects are significant (*M.
 m.
-musculus* p = `r sum_totSC.male.glm.M4$coefficients[,4][2]` -- 3.95 x 10-7; *M.
+musculus* p = `r sum_totSC.male.glm.M4$coefficients[,4][2]`; *M.
 m.
-molossinus* p = `r sum_totSC.male.glm.M4$coefficients[,4][3]` -- 3.33 x 10-7), but there are also strain-specific effects (Supplemental Table 6).
+molossinus* p = `r sum_totSC.male.glm.M4$coefficients[,4][3]`), but there are also strain-specific effects (Supplemental Table 6).
 In models with SC lengths of short and long bivalents as dependent variables, several subspecies and strain effects reach significance (p < 0.05) (Supplemental Table 9), but they are not consistent across models.
@@
 
-**not many edits to replace, bbut double check**
+**not many edits to replace, but double check**
 
@@
 *domesticus^WSB^*).
+**Each mouse from a inbred group was treated as a biological replicate**.
+
 Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice from an inbred strain of Gough Island mice were sampled after 17 generations of brother-sister mating.
+Mice were sampled from an partially-inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
 All mice were provided with ad libitum food and water.
 Procedures followed protocols approved by IACUC.
-**statement on biological replicates**
 
@@
 
+**technical replicate definition** 
+Three slides of cell spreads per mouse were prepared to serve as technical replicates for the staining protocol.
+Comparisons of multiple stained slides from the same mouse, showed no difference in mean MLH1 cell counts and mean cell quality.
+
 ### Image Processing
 
+The number of cells imaged per individual mouse is based on previous studies [Dumont, Wang].
 Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
@@
 Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
+**During preliminary analysis, cells found to be outliers for MLH1 count had the original image was examined for issues.**
+
 MLH1 foci located on the XY in spermatocytes were excluded from counts.
@@
 Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
-The hierarchal nature of the data, meant that nesting of levels across observations was implicit (ie.
+The hierarchical nature of the data, meant that nesting of levels across observations was implicit (ie.
 mouse within strain, within subspecies) and not explicitly coded.
@@
 
-This research was funded by NIH grants R01GM120051 and R01GM100426 to B. A. P.. A. L. P. was partly supported by NIH T32GM007133. Generous Franscico Peligri for microscope.
+This research was funded by NIH grants R01GM120051 and R01GM100426 to B.
+A.
+P..
+A.
+L.
+P.
+was partly supported by NIH T32GM007133.
+We are grateful to Francisco Pelegri for generous assistance with microscopy.
 

### 2020-07-03 doc/MS_v4.Rmd
updates DOIs to ref. Moved Tables around

@@
 caroli_ ).
-Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes).
-**insert statement for cell sample size number**
+Approximately 25 cells were sampled per cell were sampled given time constraints and based on previous studies [@Wang2017island; @dumont2011murid; @murdoch2010].
+Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes) (Table 1).
+<cite table 1 for MLH2 stats>
 
@@
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_).
+DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_) (Table 3).
 High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < `r 1/1000000` -- 10-6).
@@
 
-We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 1).
+We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 4).
 Our survey included 5 strains from *M.
@@
 Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice were sampled from an partially-inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
+Mice were sampled from a partially-inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
 All mice were provided with ad libitum food and water.
@@
 
-The number of cells imaged per individual mouse is based on previous studies [Dumont, Wang].
 Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
+
+**A 
+The number of cells imaged per individual mouse is based on previous studies [@Wang2017island; @dumont2011murid; @murdoch2010].
+
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
@@
 Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
-**During preliminary analysis, cells found to be outliers for MLH1 count had the original image was examined for issues.**
+When Outliers for MLH1 count were found during preliminary analysis, the original images were inspected and the counts confirmed.
 

### 2020-07-05 doc/MS_v4.Rmd
more small edits

@@
 Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes) (Table 1).
-<cite table 1 for MLH2 stats>
 
@@
 
-**not many edits to replace, but double check**
-
 We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 4).
+Since each mouse was from an inbred strain, they serve as a biological replicate.
 Our survey included 5 strains from *M.
@@
 *domesticus^WSB^*).
-**Each mouse from a inbred group was treated as a biological replicate**.
 
@@
 Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
-
-**technical replicate definition** 
 Three slides of cell spreads per mouse were prepared to serve as technical replicates for the staining protocol.
@@
 Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
-
-**A 
 The number of cells imaged per individual mouse is based on previous studies [@Wang2017island; @dumont2011murid; @murdoch2010].
-
 Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
@@
 Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
-When Outliers for MLH1 count were found during preliminary analysis, the original images were inspected and the counts confirmed.
+When outliers for MLH1 count were found during preliminary analysis, the original images were inspected and the counts confirmed.
 

### 2020-07-30 doc/MS_v5.Rmd
overdue commits

--- 2020-07-05 doc/MS_v4.Rmd
@@
   pdf_document:
-    includes:
-      in_header: header.tex
+    keep_tex:  true
   html_document:
@@
 Broadly speaking, sexual dimorphism in the genome-wide recombination rate assumes two forms.
-In species such as Drosophila melanogaster, one sex completes meiosis without forming crossovers (“achiasmy”), while the other sex recombines [@huxley1928; @haldane1922sex; @burt1991].
+In species such as _Drosophila melanogaster_, one sex completes meiosis without forming crossovers (“achiasmy”), while the other sex recombines [@huxley1928; @haldane1922sex; @burt1991].
 Alternatively, in most species with recombination, crossovers occur in both sexes but at different rates (“heterochiasmy”).
@@
 
-Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004; @Kong2008; @Kong2014; @halldorsson2019], dog [@wong2010dog], cattle [@ma2015_cattle; @Shen2018_cattle], and Soay sheep [@johnston2016_soay], in which the role of sex is confounded with the contributions of genetic variation.
+Comprehensive comparisons of variation in female and male recombination rates within species have come from outbred populations of humans [@Kong2004; @Kong2008; @Kong2014; @halldorsson2019; @gruhn2013], dog [@campbell2016], cattle [@ma2015_cattle; @Shen2018_cattle], and Soay sheep [@johnston2016_soay], in which the role of sex is confounded with the contributions of genetic variation.
 Although it is known that the level and direction of heterochiasmy can differ among species [@brandvain2012scrambling; @lenormandDuthiel_2005], the correlation between female and male recombination rates among closely related species remains poorly documented.
@@
 
-The house mouse, Mus musculus, provides a compelling system for understanding how sex determines the evolution of recombination.
+The house mouse, _Mus musculus_, is a compelling system for understanding how sex affects the evolution of recombination.
 Multiple subspecies share a most recent common ancestor approximately 0.5 million years ago [@geraldes2011], providing the opportunity to examine natural variation on recent evolutionary timescales.
-Wild Mus musculus belong to the same species as classical inbred strains of mice, where the molecular and cellular pathways that lead to crossovers have been studied extensively [@baudat2013; @handel2010; @bolcun2012].
+Wild _Mus musculus_ belong to the same species as classical inbred strains of mice, where the molecular and cellular pathways that lead to crossovers have been studied extensively [@baudat2013; @handel2010; @bolcun2012].
 Single-cell immunofluorescent approaches make it possible to estimate genome-wide recombination rates in individual males and females [@koehler2002; @peters_1997].
@@
 We demonstrate that recombination rate evolves differently in females and males, even over short timescales.
+
+![Caption for the picture.](Figure1.jpg){ width=50% }
 
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
-
-#write the full table to a supplemental table
-#show the head (or shortened)
-#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
-#kable(strain.table.HQ_Fig, "latex")
-
-df.strain.table <- strain.table.HQ_Fig
-#kable(df.strain.table)
-#pander(df.strain.table)
-
-#kable(strain.table.HQ_Fig.og, col.names = c("Subspecies", "Strain","Sex", "Number of Mice", #"Number of Cells", "Mean MLH1 count per cell", "cV", "Var", "sd", "se"), caption = "Strain #Level Summary Statistics" )  %>%
-#  kable_styling() %>%
-#  scroll_box(width = "800px", height = "200px")
-
-#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
-
-#write excel
-#write.xlsx2(strain.table.HQ_Fig.og, "second.MLH1.table.xlsx", sheetName = "Sheet1", 
-#  col.names = TRUE, row.names = FALSE, append = FALSE)
-
-```
-
 ```{r corMLH1.FM, echo=FALSE}
@@
 We used counts of MLH1 foci per cell to estimate genome-wide recombination rates in `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild-derived inbred strains sampled from three subspecies of house mice ( _M.
-m.
-domesticus_, _M.
+musculus domesticus_, _M.
 m.
@@
 caroli_ ).
-Approximately 25 cells were sampled per cell were sampled given time constraints and based on previous studies [@Wang2017island; @dumont2011murid; @murdoch2010].
 Mean MLH1 focus counts for `r sum(main.table$Nmice)` mice were quantified from an average of `r mean( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T)` spermatocytes per male (for a total of `r formatC( sum( (Mouse.Table_HQ %>% filter(sex == "male") )$Ncells, na.rm = T), format="d", big.mark=",")` spermatocytes) and `r mean( (Mouse.Table_HQ %>% filter(sex == "female") )$Ncells, na.rm = T)` oocytes per female (for a total of `r formatC(sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells), format="d", big.mark=",")` oocytes) (Table 1).
@@
 
-To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice (from three subspecies; _M.
+```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
+
+#write the full table to a supplemental table
+#show the head (or shortened)
+#kable(table(AP_mouse_table.HQ$category), col.names = c("Category", "mice"), label = "Number of mice by category" )
+#kable(strain.table.HQ_Fig, "latex")
+
+#df.strain.table <- strain.table.HQ_Fig
+#kable(df.strain.table)
+#pander(df.strain.table)
+
+#kable(table(MLH1_data$category), col.names = c("Category","Cells"), label = "Number of Cells by category" )
+
+#write excel
+#write.xlsx2(strain.table.HQ_Fig.og, "second.MLH1.table.xlsx", sheetName = "Sheet1", 
+#  col.names = TRUE, row.names = FALSE, append = FALSE)
+
+#read excel -- then kable
+Table1_MLH1 = read.xlsx2("~/MLH1repo/doc/MainTables.xlsx", header=TRUE, 1)
+
+#might have to change DF
+
+kable(Table1_MLH1, digits=2) %>%
+  kable_styling(latex_options="scale_down") %>%
+  collapse_rows(columns = 1:3)
+
+```
+
+To further partition variation in recombination rate, we fit a series of linear models to mean MLH1 focus counts from `r length( DF.HetC.MixedModel.HQ$mean_co)` house mice from three subspecies; _M.
 m.
@@
 m.
-molossinus_) (Table 2; detailed results available in Supplemental Table X).
+molossinus_ (Table 2; detailed results available in Supplemental Table X).
 Strain, sex, subspecies, and sex\*subspecies each affect MLH1 focus count in a linear mixed model (M1; strain (random effect): p < `r 1/10000`-- 10-4; sex: p = `r MLH1.MMlmer_results$sex.results`; subspecies: p = `r MLH1.MMlmer_results$sub.sp`; subspecies\*sex: p = `r MLH1.MMlmer_results$interaction.pval`).
 
-The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; musculus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, molossinus p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`), highlighting strain and sex as salient variables.
+The effect of subspecies is no longer significant in a model treating all factors as fixed effects (M2; _musculus_ p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][2]`, _molossinus_ p = `r summ.MLH1.M2.glm.strain.fixed.full$coefficients[,4][3]`), highlighting strain and sex as salient variables.
 Two strains exhibit strong effects on MLH1 focus count (M3; *domesticus^G^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]`; *domesticus^LEW^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][4]`), with sex-strain interactions involving three strains (M3; *domesticus^G^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][10]`; *molossinus^MSM^* p < 10-6 -- `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]`; *musculus^PWD^* p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]`).
 **fix the pvalue above, change the 0 to p < 10^-6**
+
+![Caption for the picture.](Figure2.jpg){ width=50% }
 
@@
 Among short bivalents (to which the female X bivalent does not contribute), female to male ratios of mouse mean SC length range from `r Q1SC.mus.av.short.biv_DIFF_PWD` (*musculus^PWD^*) to `r Q1SC.mus.av.short.biv_DIFF_WSB` (*domesticus^WSB^*) across strains.
-That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =  `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`) and short bivalent SC length (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p = `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`; M3: p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`).
+That females have longer SCs is further supported by models that include covariates, which identify sex as the most consistently significant effect for total SC length (M1: p = `r Q1.lmer_totSC_M1_results$sex.results`; M2: p = `r sum.Q1.glm_totSC_M2$coefficients[,4][8]`; M3: p =  `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`) and short bivalent SC length (M1: p = `r Q1.lmer_short_M1_results$sex.results`; M2: p < `r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`; M3: p < `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`).
 The existence of some subspecies and strain effects on total SC length and short bivalent SC length (Supplemental Tables 5 and 8) further indicates that SC length has evolved among strains and among subspecies.
@@
 This pattern implies that in high-recombination strains, spermatocytes have less space than oocytes in which to position additional crossovers.
+
+![Caption for the picture.](Fig3_SC.jpg){ width=50% }
 
@@
 In contrast, there is no clear signal of sex differences in raw mean inter-focal distances (IFD~raw~) (Supplemental Table 13 and 14) across the full set of strains, whether they are considered separately or together.
-Visualization of normalized MLH1 foci positions on bivalents with two crossovers **(Figure 4C and Supplemental Figure X)** further suggests that interference distances vary more in females than in males, and that males display a stronger telomeric bias in the placement of the distal crossover.
+Visualization of normalized MLH1 foci positions on bivalents with two crossovers (Figure 3C and Supplemental Figure 3) further suggests that interference distances vary more in females than in males, and that males display a stronger telomeric bias in the placement of the distal crossover.
 
@@
 ## Evolution of genome-wide recombination rate is dispersed across bivalents, associated with double-strand break number, and connected to crossover interference
+
+```{r DMC1.Table, echo=FALSE}
+
+#insert and read
+
+```
 
@@
 To begin to localize evolution of genome-wide recombination rate to steps of the recombination pathway, we counted DMC1 foci in prophase spermatocytes as markers for double-strand breaks (DSBs).
-DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains ( _musculus^PWD^_ and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_) (Table 3).
-High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < `r 1/1000000` -- 10-6).
+DMC1 foci were counted in a total of `r sum( (DMC1table.main$ncells.x) )` early zygotene and `r sum( (DMC1table.main$ncells.y) )` late zygotene spermatocytes from two high-recombination strains (*musculus^PWD^* and _molossinus^MSM^_) and three low-recombination strains (*musculus^KAZ^*, _domesticus^WSB^_, and _domesticus^G^_) (Table 3).
+High-recombination strains have significantly more DMC1 foci than low-recombination strains in early zygotene cells (t-test; p < `r 1/1000000`).
 In contrast, the two strain groups do not differ in DMC1 foci in late zygotene cells (t-test; p = `r ttest.HighLow.Z.pval`).
@@
 At the genetic level, the sex-specific patterns we documented indicate that some mutations responsible for the evolution of recombination rate have dissimilar phenotypic effects in the two sexes.
-A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@Kong2004;, @Kong2008; @Kong2014; @halldorsson2019], Soay sheep [@johnston2016_soay], and cattle [@ma2015_cattle; @Shen2018_cattle] appear to show sex-specific properties, including opposite effects in females and males.
+A subset of the genetic variants associated with genome-wide recombination rate within populations of humans [@Kong2004; @Kong2008; @Kong2014; @halldorsson2019], cattle [@ma2015_cattle; @Shen2018_cattle], and Soay sheep [@johnston2016_soay] appear to show sex-specific properties, including opposite effects in females and males.
 Furthermore, inter-sexual correlations for recombination rate are weak in humans [@fledel2011] and Soay sheep [@johnston2016_soay].
@@
 Little is known about whether and how natural selection shapes recombination rate in nature [@DapperPayseur2017; @Ritz2017].
-@samuk2020 recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of Drosophila pseudoobscura was caused by natural selection.
+@samuk2020 recently used a quantitative genetic test to conclude that an 8% difference in genome-wide recombination rate between females from two populations of_Drosophila pseudoobscura_ was caused by natural selection.
 Applying similar strategies to species in which both sexes recombine, including house mice, would be a logical next step to understanding the sex-specific evolution of recombination rate.
@@
 A cellular model designed to explain interference attributes sexual dimorphism in chromatin structure to greater cell volumes and oscillatory movements of telomeres and kinetochores in oocytes [@hulten2011_COM].
-Recent work in mice connects the sparser recombination landscape in females to sex differences in the crossover maturation efficiency [@wang2017inefficient].
+Recent work in mice connects the sparser recombination landscape in females to sex differences in crossover maturation efficiency [@wang2017inefficient].
 
@@
 A second limitation is that our investigation of crossover locations was confined to the relatively low resolution possible with immunofluorescent cytology.
-Positioning crossovers with higher resolution and maintaining single chromosome scale could reveal additional evolutionary patterns.
+Positioning crossovers with higher resolution could reveal additional evolutionary patterns.
 Finally, the panel of inbred lines we surveyed may not be representative of recombination rate variation within and between subspecies of house mice.
@@
 We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 4).
-Since each mouse was from an inbred strain, they serve as a biological replicate.
+Mice from the same inbred strain served as serve as a biological replicates.
 Our survey included 5 strains from *M.
@@
 caroli_.
-We subsequently denote strains by their abbreviated subspecies and name (e.g.
-*domesticus^WSB^*).
-
+We subsequently denote strains by their abbreviated subspecies and name (*e.g.
+domesticus^WSB^*).
 Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice were sampled from a partially-inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
-All mice were provided with ad libitum food and water.
+Mice were sampled from a partially inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
+All mice were provided with *ad libitum* food and water.
 Procedures followed protocols approved by IACUC.
@@
 
-The same dry-down spread technique was applied to both spermatocytes and oocytes, following [@peters_1997], with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in [@peterson2019].
+The same dry-down spread technique was applied to both spermatocytes and oocytes, following @peters_1997, with adjustment for volumes.
+Spermatocyte spreads were collected and prepared as described in @peterson2019.
 The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
-Juvenile mice between 12 and 15 days of age were used for DMC1 counts.
+Juvenile males between 12 and 15 days of age were used for DMC1 counts.
 Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
@@
 Three slides of cell spreads per mouse were prepared to serve as technical replicates for the staining protocol.
-Comparisons of multiple stained slides from the same mouse, showed no difference in mean MLH1 cell counts and mean cell quality.
+Comparisons of multiple stained slides from the same mouse showed no difference in mean MLH1 cell counts and mean cell quality.
+Sampled numbers of mice and cells per mouse were maximized to the extent possible given constraints on breeding and time.
 
@@
 
-### Analysis
+### Analyses
 
-To estimate the number of crossovers across the genome, we counted MLH1 foci.
-MLH1 foci were counted in cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+To estimate the number of crossovers across the genome, we counted MLH1 foci within bivalents, synapsed homologous chromosomes.
+MLH1 foci were counted in pachytene cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
 A quality score ranging from 1 (best) to 5 (worst) was assigned to each cell based on visual appearance of staining and spread of bivalents.
@@
 The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-The curated single bivalent data supplements our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
+The curated single bivalent data supplemented our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
 Proportions of bivalents with the same number of MLH1 foci were compared across strains using a chi-square test.
@@
 To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-This “short bivalent” data set included the four or five shortest bivalents and excluded the X bivalent in oocytes.
+This “short bivalent” data set included the four or five shortest bivalents within a cell, thus excluding the X bivalent in oocytes.
 A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
-Although this smaller data set has decreased power, it offers a more comparable set of single bivalents to compare between the sexes.
+Although this smaller data set had decreased power, it offered a more comparable set of single bivalents to compare between the sexes.
 A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
@@
 We first constructed a linear mixed model (M1) using lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
-In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test (using exactRLRT() from RLRsim [@RLRsim]).
+In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test using exactRLRT() from RLRsim [@RLRsim].
 Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
-The hierarchical nature of the data, meant that nesting of levels across observations was implicit (ie.
-mouse within strain, within subspecies) and not explicitly coded.
+The hierarchical nature of the data, meant that nesting of levels across observations was implicit (*i.e.* mouse within strain, within subspecies) and not explicitly coded.
 We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.

### 2020-07-30 doc/MS_v6.Rmd
overdue commits

--- 2020-07-30 doc/MS_v5.Rmd
@@
 title: 'Sex-Specific Evolution of the Meiotic Recombination Rate'
-date: '`r Sys.Date()`'
 author: "April L. Peterson, Bret A. Payseur"
@@
     keep_tex:  true
+  header-includes:
+    - \usepackage{setspace}
+    - \doublespacing
+    - \usepackage{lineno}
+    - \linenumbers
   html_document:
 always_allow_html: yes
+header-includes:
+    - \usepackage{setspace}
+    - \doublespacing
+    - \usepackage{lineno}
+    - \linenumbers
 bibliography: refs_Main_Report.bib
@@
 
+# MATERIALS AND METHODS
+
+### Mice
+
+We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 4).
+Mice from the same inbred strain served as serve as a biological replicates.
+Our survey included 5 strains from *M.
+m.
+musculus*, 4 strains from *M.
+m.
+domesticus*, 2 strains from *M.
+m.
+molossinus*, 2 strains from *M.
+m.
+castaneus*, and 1 strain each from *M.
+spicilegus*, *M.
+spretus* and _M.
+caroli_.
+We subsequently denote strains by their abbreviated subspecies and name (*e.g.
+domesticus^WSB^*).
+Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
+Mice were sampled from a partially inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
+All mice were provided with *ad libitum* food and water.
+Procedures followed protocols approved by IACUC.
+
+### Tissue Collection and Immunohistochemistry
+
+The same dry-down spread technique was applied to both spermatocytes and oocytes, following @peters_1997, with adjustment for volumes.
+Spermatocyte spreads were collected and prepared as described in @peterson2019.
+The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
+Juvenile males between 12 and 15 days of age were used for DMC1 counts.
+Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
+Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
+Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
+Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
+Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
+The following morning, slides were treated with a Photoflow wash (Kodak, diluted 1:200).
+Slides were stored at -20\*C if not stained immediately.
+To visualize the structure of meiotic chromosomes, we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
+Crossovers (COs) were visualized as MLH1 foci.
+Double strand breaks (DSBs) were visualized as DMC1 foci.
+The staining protocol followed [@anderson1999] and [@koehler2002].
+Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
+Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\*C.
+The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
+Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\*C.
+Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
+The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
+Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
+Three slides of cell spreads per mouse were prepared to serve as technical replicates for the staining protocol.
+Comparisons of multiple stained slides from the same mouse showed no difference in mean MLH1 cell counts and mean cell quality.
+Sampled numbers of mice and cells per mouse were maximized to the extent possible given constraints on breeding and time.
+
+### Image Processing
+
+Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
+The number of cells imaged per individual mouse is based on previous studies [@Wang2017island; @dumont2011murid; @murdoch2010].
+Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
+Image file names were anonymized before manual scoring of MLH1 foci or DMC1 foci using Photoshop.
+
+### Analyses
+
+To estimate the number of crossovers across the genome, we counted MLH1 foci within bivalents, synapsed homologous chromosomes.
+MLH1 foci were counted in pachytene cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
+A quality score ranging from 1 (best) to 5 (worst) was assigned to each cell based on visual appearance of staining and spread of bivalents.
+Cells with a score of 5 were excluded from the final analysis.
+Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
+When outliers for MLH1 count were found during preliminary analysis, the original images were inspected and the counts confirmed.
+
+MLH1 foci located on the XY in spermatocytes were excluded from counts.
+In addition to MLH1 counts, we measured several traits to further characterize the recombination landscape.
+To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from each of a subset of strains (*molossinus^MSM^*, _musculus^PWD^_,  _domesticus^WSB^_, and _domesticus^G^_ ).
+SC morphology and CREST foci number were used to stage spermatocytes as early zygotene or late zygotene.
+
+To measure bivalent SC length, two image analysis algorithms were used.
+The first algorithm estimates the total (summed) SC length across bivalents for individual cells [@wang2019_SC].
+The second algorithm estimates the SC length of individual bivalents [@peterson2019].
+Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
+Total SC length per cell was quantified from pachytene cell images [@wang2019_SC].
+
+To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
+Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
+SC length of individual bivalents was quantified in pachytene cell images [@peterson2019].
+The DNA CrossOver algorithm [@peterson2019] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
+The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
+In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` (*molossinus^MSM^* male) to `r max(algorithm.stat.category$mean_pass.rate)` (*musculus^KAZ^* female).
+From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` bivalent objects were isolated by the algorithm.
+Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
+The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
+The curated single bivalent data supplemented our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
+Proportions of bivalents with the same number of MLH1 foci were compared across strains using a chi-square test.
+
+To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
+This “short bivalent” data set included the four or five shortest bivalents within a cell, thus excluding the X bivalent in oocytes.
+A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
+Although this smaller data set had decreased power, it offered a more comparable set of single bivalents to compare between the sexes.
+A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
+A total of `r sum(Long.biv_mouse_table$n_obs)` long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
+
+To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
+A normalized measure of interference (IFD~norm~) was computed by dividing IFD~raw~ by SC length on a per-bivalent basis.
+
+We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
+We used mouse average as the dependent variable in all analyses.
+We first constructed a linear mixed model (M1) using lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
+In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test using exactRLRT() from RLRsim [@RLRsim].
+Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
+The hierarchical nature of the data, meant that nesting of levels across observations was implicit (*i.e.* mouse within strain, within subspecies) and not explicitly coded.
+We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
+In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
+Four additional linear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
+
 # RESULTS
@@
 
-```{r MAIN.TABLE1, echo=FALSE, eval=TRUE}
+```{r MAIN.TABLE1, echo=FALSE, eval=FALSE}
 
@@
 
+<SAC section 1>
+
 We propose an alternative interpretation of our findings based on the cell biology of gametogenesis.
@@
 
-# MATERIALS AND METHODS
-
-### Mice
-
-We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murid species to profile natural genetic variation in recombination (Table 4).
-Mice from the same inbred strain served as serve as a biological replicates.
-Our survey included 5 strains from *M.
-m.
-musculus*, 4 strains from *M.
-m.
-domesticus*, 2 strains from *M.
-m.
-molossinus*, 2 strains from *M.
-m.
-castaneus*, and 1 strain each from *M.
-spicilegus*, *M.
-spretus* and _M.
-caroli_.
-We subsequently denote strains by their abbreviated subspecies and name (*e.g.
-domesticus^WSB^*).
-Mice were housed at dedicated, temperature-controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine.
-Mice were sampled from a partially inbred strain of Gough Island mice, approximately 6 generations of brother-sister matting.
-All mice were provided with *ad libitum* food and water.
-Procedures followed protocols approved by IACUC.
-
-### Tissue Collection and Immunohistochemistry
-
-The same dry-down spread technique was applied to both spermatocytes and oocytes, following @peters_1997, with adjustment for volumes.
-Spermatocyte spreads were collected and prepared as described in @peterson2019.
-The majority of mice used for MLH1 counts were between 5 and 12 weeks of age.
-Juvenile males between 12 and 15 days of age were used for DMC1 counts.
-Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth).
-Whole testes were incubated in 3ml of hypotonic solution for 45 minutes.
-Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes.
-Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution.
-Cells were fixed in this solution and dried in a humid chamber at room temperature overnight.
-The following morning, slides were treated with a Photoflow wash (Kodak, diluted 1:200).
-Slides were stored at -20\*C if not stained immediately.
-To visualize the structure of meiotic chromosomes, we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3).
-Crossovers (COs) were visualized as MLH1 foci.
-Double strand breaks (DSBs) were visualized as DMC1 foci.
-The staining protocol followed [@anderson1999] and [@koehler2002].
-Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
-Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\*C.
-The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB.
-Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\*C.
-Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB.
-The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C.
-Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.
-Three slides of cell spreads per mouse were prepared to serve as technical replicates for the staining protocol.
-Comparisons of multiple stained slides from the same mouse showed no difference in mean MLH1 cell counts and mean cell quality.
-Sampled numbers of mice and cells per mouse were maximized to the extent possible given constraints on breeding and time.
-
-### Image Processing
-
-Images were captured using a Zeiss Axioplan 2 microscope with AxioLab camera and AxioVision software (Zeiss, Cambridge, UK).
-The number of cells imaged per individual mouse is based on previous studies [@Wang2017island; @dumont2011murid; @murdoch2010].
-Preprocessing, including cropping, noise reduction, and histogram adjustments, was performed using Photoshop (v13.0).
-Image file names were anonymized before manual scoring of MLH1 foci or DMC1 foci using Photoshop.
-
-### Analyses
-
-To estimate the number of crossovers across the genome, we counted MLH1 foci within bivalents, synapsed homologous chromosomes.
-MLH1 foci were counted in pachytene cells with intact and complete karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents for oocytes) and distinct MLH1 foci.
-A quality score ranging from 1 (best) to 5 (worst) was assigned to each cell based on visual appearance of staining and spread of bivalents.
-Cells with a score of 5 were excluded from the final analysis.
-Distributions of MLH1 count per cell were visually inspected for normality (Supplemental Figure 1).
-When outliers for MLH1 count were found during preliminary analysis, the original images were inspected and the counts confirmed.
-
-MLH1 foci located on the XY in spermatocytes were excluded from counts.
-In addition to MLH1 counts, we measured several traits to further characterize the recombination landscape.
-To estimate the number of double-strand breaks, a minority of which lead to crossovers, mean DMC1 foci per cell was quantified for a single male from each of a subset of strains (*molossinus^MSM^*, _musculus^PWD^_,  _domesticus^WSB^_, and _domesticus^G^_ ).
-SC morphology and CREST foci number were used to stage spermatocytes as early zygotene or late zygotene.
-
-To measure bivalent SC length, two image analysis algorithms were used.
-The first algorithm estimates the total (summed) SC length across bivalents for individual cells [@wang2019_SC].
-The second algorithm estimates the SC length of individual bivalents [@peterson2019].
-Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.
-Total SC length per cell was quantified from pachytene cell images [@wang2019_SC].
-
-To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set.
-Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus")  )$Random.Name), format="d", big.mark=",")` out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") )$Random.Name), format="d", big.mark=",")` cells with MLH1 counts.
-SC length of individual bivalents was quantified in pachytene cell images [@peterson2019].
-The DNA CrossOver algorithm [@peterson2019] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci.
-The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents.
-In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` (*molossinus^MSM^* male) to `r max(algorithm.stat.category$mean_pass.rate)` (*musculus^KAZ^* female).
-From the total set of pachytene cell images, `r formatC(  length(unique(Curated_BivData$Obj.ID ) ) , format="d", big.mark=",")` bivalent objects were isolated by the algorithm.
-Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` single-bivalent observations remained.
-The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019].
-The curated single bivalent data supplemented our cell-wide MLH1 count data with MLH1 foci counts for single bivalents.
-Proportions of bivalents with the same number of MLH1 foci were compared across strains using a chi-square test.
-
-To account for confounding effects of sex chromosomes from pooled samples of bivalents, we also considered a reduced data set including only bivalents with SC lengths below the 2nd quartile in cells with at least 17 of 20 single bivalent measures.
-This “short bivalent” data set included the four or five shortest bivalents within a cell, thus excluding the X bivalent in oocytes.
-A total of `r sum(Short.biv_mouse_table$n_obs )` short bivalents were isolated from `r sum(  (Short.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Short.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
-Although this smaller data set had decreased power, it offered a more comparable set of single bivalents to compare between the sexes.
-A “long bivalent” data set was formed from those bivalents above the 4th quartile in SC lengths per cell.
-A total of `r sum(Long.biv_mouse_table$n_obs)` long bivalents were isolated from `r sum(  (Long.biv_mouse_table  %>% filter(sex=="female") )$Ncells)` oocytes and `r sum(  (Long.biv_mouse_table  %>% filter(sex=="male") )$Ncells)` spermatocytes.
-
-To examine crossover interference, the distance (in SC units) between MLH1 foci (inter-focal distance; IFD~raw~) was measured for those single bivalents containing two MLH1 foci.
-A normalized measure of interference (IFD~norm~) was computed by dividing IFD~raw~ by SC length on a per-bivalent basis.
-
-We used a series of statistical models to interpret patterns of variation in the recombination traits we measured (Table 2).
-We used mouse average as the dependent variable in all analyses.
-We first constructed a linear mixed model (M1) using lmer() from the lmer4 package [@lme4] in R (v3.5.2) [@Rstudio].
-In this model, strain was coded as a random effect, with significance evaluated using a likelihood ratio test using exactRLRT() from RLRsim [@RLRsim].
-Subspecies, sex, and their interaction were coded as fixed effects, with significance evaluated using a chi-square test comparing the full and reduced models (drop1() and anova()) [@lme4].
-The hierarchical nature of the data, meant that nesting of levels across observations was implicit (*i.e.* mouse within strain, within subspecies) and not explicitly coded.
-We used the subspecies effect to quantify divergence between subspecies and the (random) strain effect to quantify variation within subspecies in a sex-specific manner.
-In separate analyses using model M1, we considered mouse averages as dependent variables for each of the following traits: MLH1 count per cell, total SC length per cell, single bivalent SC length per cell, IFD~raw~, IFD~norm~, and average MLH1 position (for single-focus bivalents).
-Four additional linear models containing only fixed effects (M2-M5) (Table 2) were used to further investigate results obtained from model M1.
-
 # Acknowledgements

### 2020-07-30 doc/MS_v6.tex
overdue commits

--- 2020-06-14 doc/MS_v3.tex
@@
   pdftitle={Sex-Specific Evolution of the Meiotic Recombination Rate},
-  pdfauthor={April L. Peterson, Bret Payseur},
+  pdfauthor={April L. Peterson, Bret A. Payseur},
   hidelinks,
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{titling}
-
-\pretitle{%
-  \begin{center}
-  \LARGE
-  \includegraphics[width=4cm,height=6cm]{logo.png}\\[\bigskipamount]
-}
-\posttitle{\end{center}}
+\usepackage{setspace}
+\doublespacing
+\usepackage{lineno}
+\linenumbers
 \usepackage{booktabs}
@@
 \title{Sex-Specific Evolution of the Meiotic Recombination Rate}
-\author{April L. Peterson, Bret Payseur}
-\date{2020-06-14}
+\author{April L. Peterson, Bret A. Payseur}
+\date{2020-07-29}
 
@@
 in most species that reproduce sexually, the rate of crossing over
-varies among individuals. \textbf{define genome-wide recombination
-rate}.
-
-Differences in recombination rate between females and males are perhaps
-the most striking form of this variation. Existing data fail to directly
-address the extent to which recombination experiences similar
-evolutionary pressures. To fill this gap, we measured meiotic
-recombination in both sexes for a panel of house mouse cross three
-subspecies. Using inbred strains and single cell immunohistochemistry
-allowed us to place sex-specific observations within the same genetic
-background and meiotic context. Our results indicate highly discordant
-evolutionary patterns in the two sexes. Whereas male recombination rates
-show evidence of rapid evolution over short evolutionary timescales,
+varies among individuals. Differences in recombination rate between
+females and males are perhaps the most striking form of this variation.
+Despite the recognition that the sexes recombine at different rates,
+existing data fail to address the extent to which recombination
+experiences similar evolutionary pressures in the two sexes. To fill
+this gap, we measured the genome-wide recombination rate in both sexes
+from a panel of wild-derived inbred strains sampled from across three
+subspecies of house mice. We applied single-cell immunohistochemistry to
+inbred strains to directly compare females and males from the same
+genetic background and meiotic context, thereby isolating the effect of
+sex on evolutionary differences in recombination. Our results reveal
+highly discordant evolutionary trajectories in the two sexes. Whereas
+male recombination rates show rapid evolution over short timescales,
 female recombination rates measured in the same strains are mostly
-static. These results strongly indicate that house mouse has two genome
-wide recombination rates which display distinct evolutionary
-trajectories.
+static. Strains with high recombination in males have more double-strand
+breaks and stronger crossover interference than strains with low
+recombination in males, suggesting that these factors contribute to the
+sex-specific evolution we document. These results provide the strongest
+evidence yet that sex is a primary driver of recombination rate
+evolution.
 
@@
 
-\hypertarget{importance-of-recombination-rate}{%
-\subsubsection{0. Importance of recombination
-rate}\label{importance-of-recombination-rate}}
-
-Meiosis can be reduced to the expression of (2n -\textgreater{} 4n
--\textgreater{} 2n -\textgreater{} 1n) which tracks the duplication of a
-diploid genome into haploid cell products. The meiotic program relies on
-crossovers and the process of recombination to ensure the correct
-separation of chromosomes. The total number of COs per cell (at the 4n
-stage) is equal to the genome wide recombination rate (gwRR).
-\textbf{The genome wide recombination rate is a cell based metric for
-genome wide recombination rate is tightly connected to organisms fitness
-through correct chromosome segregation and fertility.}
-
-\textbf{at the fine scale / in regards to genetic variation}, In the
-indriect manner, the recombination rate regulates populations' response
-to selection, and determine the fate of novel mutations by transferring
-beneficial mutations onto novel genetic backgrounds or by breaking
-linkage of negative mutations from beneficial genetic backgrounds
-(cite). This process shapes the genomic patterns of genetic variation,
-with high recombination areas of genome having more nucleotide variation
-while areas of low recombination have decreased genetic variation (Begun
-Aquadro, Nachman, Payseur).
-
-\textbf{but back to the genome-wide recombination rate}, The process of
-chromosome segregation is integral to the process of producing haploid
-cells and can be considered a constraint in regards to the evolution.
-When the genome wide recombination rate is framed within the single cell
-context it is closely connected to the contraints which are imposed at
-the single cell level. Reflecting the importance of it's role in
-chromosome segreation is that the number of haploid pairs is strong
-predictor for genome wide RR across large species(taxa) / divergence
-scales (Otto and Payseur, 2019; Stapley et al., 2017). (the number of
-crossovers per chromosome are highly conserved)\textless However, whie
-the rate of recombination is highly variable, the count of crossovers
-per chromosome is conserved between 1 to 3 (cite), suggesting that such
-thresholds/constraints may be behind the overall patterns.\textgreater{}
-
-\textbf{connection of single cell based gwRR to thresholds imposed at
-the single cell level} (why is gwRR the better scale for evolutionary
-patterns?)
-
-\hypertarget{understanding-levels-of-variation-in-gwrr}{%
-\subsubsection{1. Understanding levels of variation in
-gwRR}\label{understanding-levels-of-variation-in-gwrr}}
-
-Because we know the meiotic programs, we have expectations for where the
-thresholds(constraints) of the gwRR the rate will fall. First, the lower
-threshold is defined as an obligate crossover per homolog pair which
-promotes proper chromosome segregation (Nagaoka et al., 2012). second
-(while there is less evidence,), the upper threshold maybe
-regulated/shaped/affected due to rates of ectopic recombination or the
-programmed DNA damage preceding crossovers act as constraints to the
-upper threshold of meiotic recombination (cite). Yet, Within these
-thresholds the genome wide recombination rate still varies across
-species and indiviuals. which is the subject of an open question in the
-field. \textbf{the gwRR is the best metric for since both of these
-constraints are thought to act at the cell wide level.}
-
-\textbf{Sex is one of the most notable axes along which individuals
-vary}. Sexual dimorphism in recombination rates is called heterochiasmy
-(Lenormand, 2003). An understanding of how sex shapes the evolution of
-recombination cannot be achieved with available data. Comprehensive
-comparisons of female and male recombination rates usually come from
-outbred populations humans (Kong et al., 2004,, 2008, 2014; Halldorsson
-et al., 2019), dog, cattle (Ma et al., 2015; Shen et al., 2018), (other
-sheep) sheep (Johnston et al., 2016), and mouse (CC cite) in which the
-role of sex is confounded with the contributions of genetic variation.
-Under anisogamy where a species has distinct gametes, there is no first
-principle which would predict the evolution of sexually dimorphic
-recombination rates. Yet heterochiasmy, is commonly observed in
-dioecious species, suggesting that other meiotic traits which
-distinguish the gametes and their meiotic programs, for example
-symmetrical vs asymmetrical cell division, may impose selection for
-sexually dimorphic recombination rates. Although it is clear that the
-relationship between female and male recombination rates can differ
-among species, comparisons between and within closely related species
-are missing. Direct contrasts between females and males across a common,
-diverse set of genomic backgrounds would reveal whether the
-recombination rate evolves differently in the sexes.
-
-\textless It's important to note, -- that the genome wide recombination
-rate can be decomposed -- and connected to related traits (which also
-show conserved patterns of sexual dimorphism\textgreater{}
-
-\textbf{understanding heterochiasmy and sexual dimorphism in meiosis
-have advanced since first being documented (morgan 1914) -- more than
-just the direction -- more evidence is accumulating that there are
-conserved patterns of sexual dimorphism in meiotic recombinatio -
-related to recombination rate}. Recently two specific features have been
-highlighted as conserved patterns of heterochiasmy:
-
-Cahoon and Libuda (2019) show In species with meiotic chromatin (SC axis
-length) quantified for both sexes, longer chromosome axis generally has
-higher recombination rates. Sardell and Kirkpatrick (2020), A survey of
-51 species shows there are conserved broad scale recombination landscape
-differences. Generally males have telomere-bias crossover placement and
-females have more uniform placement.
-
-These observations raise the question of conservation in sexually
-patterns are maintained with the gwRR and it's decomposed traits
-
-\textbf{remember these points kinda distract from gwRR aspect}
-
-\hypertarget{the-house-mouse-is-a-great-model-for-uncovering-evolutionary-patterns-at-a-short-timescale}{%
-\subsubsection{2. The House Mouse is a great model for uncovering
-evolutionary patterns at a short
-timescale}\label{the-house-mouse-is-a-great-model-for-uncovering-evolutionary-patterns-at-a-short-timescale}}
-
-The House mouse complex arose from a recent radiation providing an
-opportunity to interrogate natural variation at short evolutionary
-scales, potentially narrowing the mutational space for identifying
-variants which would cause gwRR variation. Add divergence ranges.
-divergence spanning 0.5 mya to 5 mya.
-
-Classical lab strains of mice have generated extensive knowledge central
-to meiosis and outcomes on recombination through extensive studies on
-the genes involved in the meiotic recombination pathway (cite) and
-previous crosses for understanding the genetic architecture of
-recombination rate variation ({\textbf{???}}); Murdoch et al. (2010);
-Wang and Payseur (2017){]}.
-
-\textbf{Wild derived inbred strains generate the best comparison of
-females and males, besides the sex chromosomes, the mouse for each
-genome are almost identical}. Inbred strains enable one of the most
-direct comparison of both female and male -- versions of meiosis
---enable use to conclude that sex is a primary factor --
-
-Additionally due to their global distribution allows us to compare
-samples of natural diversity from a broad geographic range. (wild
-derived inbred strains -- will capture genetic diversity not represented
-in classical lab strains.)
-
-While there are strains with Robertsonian translocations, in house mouse
-and related murid species it is possible to assemble sets of strains
-with identical karyotypes, 20 pairs of acrocentric chromosomes.
-
-House mouse is well suited for single cell cytology approaches. The
-single cell level allows integration of data at a closer connection to
-the molecular pathway and meiotic program. In addition to the
-quantification of gwRR. This study quantifies precursors to crossovers,
-double strand breaks (DSBs), and meiotic chromosome morphology across
-stages of meiosis which would not be accessible to with genetic linkage
-data and crosses.
-
-\hypertarget{what-we-accomplished-in-this-paper}{%
-\subsubsection{3. What we accomplished in this
-paper}\label{what-we-accomplished-in-this-paper}}
-
-For the first time in many of these strains, We report a rare, direct,
-evolutionary comparison of recombination rate in females and males for
-this short evolutionary scale. We use rare strains from a broad range of
-geographic locations of the species territory.
-
-The gwRR is made up of crossovers occurring on individual chromosomes
-within cells. Thus we quantified meiotic chromosome morphology (SC
-length) and placement of crossovers to comprise an approximate picture
-of the recombination landscape.
-
-Our results indicate rapid male specific evolution of gwRR occurred in a
-subset of house mouse lines. We observed up to a 30\% difference
-(translating into approximately \textasciitilde7 more crossovers per
-cell), a surprising amount considering the short evolutionary time
-scale.
+Meiosis converts diploid germ cells into haploid gametes. During meiosis
+I, DNA crossovers aid the separation of homologous chromosomes by
+physically linking them and establishing tension between them on the
+spindle (Petronczki et al., 2003). The wrong number of recombination
+events can disrupt chromosomal segregation, leading to infertility,
+miscarriage, and birth defects (Hassold and Hunt, 2001). Recombination
+also shapes evolution by shuffling the combinations of genetic variants
+offspring inherit. Recombination affects the fates of beneficial and
+deleterious mutations (Felsenstein, 1974; Fisher, 1930; Hill and
+Robertson, 1966) and interacts with natural selection to leave gradients
+in genomic patterns of diversity (Begun and Aquadro, 1992; Charlesworth
+et al., 1993; Cutter and Payseur, 2013; Nachman and Payseur, 2012; Smith
+and Haigh, 1974).
+
+The role of recombination in facilitating meiotic chromosome assortment
+suggests that the total number of crossovers in a cell -- the
+genome-wide recombination rate -- is an important cellular
+characteristic connected to organismal fitness. The dual pressures of
+ensuring at least one crossover per chromosome and minimizing levels of
+DNA damage and ectopic exchange are thought to impose lower and upper
+thresholds on the genome-wide recombination rate (Inoue and Lupski,
+2002; Nagaoka et al., 2012). Yet, within these bounds, individuals from
+the same species can vary substantially in crossover number (Gruhn et
+al., 2013; Johnston et al., 2016; Kong et al., 2008; Ma et al., 2015).
+
+Sex is perhaps the most notable axis along which recombination rate
+varies. Broadly speaking, sexual dimorphism in the genome-wide
+recombination rate assumes two forms. In species such as
+\emph{Drosophila melanogaster}, one sex completes meiosis without
+forming crossovers (``achiasmy''), while the other sex recombines (Burt
+et al., 1991; Haldane, 1922; Huxley, 1928). Alternatively, in most
+species with recombination, crossovers occur in both sexes but at
+different rates (``heterochiasmy''). In these species, females tend to
+recombine more than males (Bell, 1982; Brandvain and Coop, 2012; Burt et
+al., 1991; Lenormand and Dutheil, 2005; Lorch, 2005). In plants,
+heterochiasmy is correlated with the opportunity for haploid selection
+(Lenormand and Dutheil, 2005). Despite the establishment of these
+interspecific trends, an understanding of how sex shapes the evolution
+of recombination cannot be achieved with available data.
+
+Comprehensive comparisons of variation in female and male recombination
+rates within species have come from outbred populations of humans (Gruhn
+et al., 2013; Halldorsson et al., 2019; Kong et al., 2004, 2014, 2008),
+dog (Campbell et al., 2016), cattle (Ma et al., 2015; Shen et al.,
+2018), and Soay sheep (Johnston et al., 2016), in which the role of sex
+is confounded with the contributions of genetic variation. Although it
+is known that the level and direction of heterochiasmy can differ among
+species (Brandvain and Coop, 2012; Lenormand and Dutheil, 2005), the
+correlation between female and male recombination rates among closely
+related species remains poorly documented. Direct contrasts between the
+two sexes across a common, diverse set of genomic backgrounds that
+represent recent timescales would reveal whether the genome-wide
+recombination rate evolves differently in males and females.
+
+Examining variation in the total number of crossovers in a sex-specific
+manner could also illuminate evolutionary connections between
+recombination rate and crossover positioning. Analyses of meiotic
+chromosome morphology in \emph{Arabidopsis thaliana},
+\emph{Caenorhabditis elegans}, and \emph{Mus musculus} suggest that the
+sex with more recombination usually has longer chromosome axes (Cahoon
+and Libuda, 2019). A survey of 51 species found conserved sex
+differences in the recombination landscape, including telomere-biased
+placement of crossovers in males but not in females (Sardell and
+Kirkpatrick, 2020). The degree to which a crossover reduces the
+probability of another crossover nearby (crossover interference) also
+differs between females and males (Otto and Payseur, 2019).
+
+The house mouse, \emph{Mus musculus}, is a compelling system for
+understanding how sex affects the evolution of recombination. Multiple
+subspecies share a most recent common ancestor approximately 0.5 million
+years ago (Geraldes et al., 2011), providing the opportunity to examine
+natural variation on recent evolutionary timescales. Wild \emph{Mus
+musculus} belong to the same species as classical inbred strains of
+mice, where the molecular and cellular pathways that lead to crossovers
+have been studied extensively (Baudat et al., 2013; Bolcun-Filas and
+Schimenti, 2012; Handel and Schimenti, 2010). Single-cell
+immunofluorescent approaches make it possible to estimate genome-wide
+recombination rates in individual males and females (Koehler et al.,
+2002; Peters et al., 1997). A collection of wild-derived inbred strains
+founded from a variety of geographic locations is available, enabling
+genetic variation in recombination to be profiled across the species
+range. Most importantly, by measuring recombination rates in females and
+males from the same set of wild-derived inbred strains, the evolutionary
+dynamics of recombination can be directly compared in the two sexes.
+
+In this paper, we report genome-wide recombination rates from both sexes
+in a diverse panel of wild-derived inbred strains of house mice and
+their close relatives. We demonstrate that recombination rate evolves
+differently in females and males, even over short timescales.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth,height=\textheight]{Figure1.jpg}
+\caption{Caption for the picture.}
+\end{figure}
 
@@
 musculus}) and related murid species to profile natural genetic
-variation in recombination (Table 1). Our survey included 5 strains from
-\emph{M. m. musculus}, 4 strains from \emph{M. m. domesticus}, 2 strains
-from \emph{M. m. molossinus}, 2 strains from \emph{M. m. castaneus}, and
-1 strain each from \emph{M. spicilegus} and \emph{M. spretus}.
-\textless removed \emph{M. caroli}\textgreater{} We subsequently denote
-strains by their abbreviated subspecies and name
-(e.g.~\emph{domesticus\textsuperscript{WSB}}). Mice were housed at
+variation in recombination (Table 4). Mice from the same inbred strain
+served as serve as a biological replicates. Our survey included 5
+strains from \emph{M. m. musculus}, 4 strains from \emph{M. m.
+domesticus}, 2 strains from \emph{M. m. molossinus}, 2 strains from
+\emph{M. m. castaneus}, and 1 strain each from \emph{M. spicilegus},
+\emph{M. spretus} and \emph{M. caroli}. We subsequently denote strains
+by their abbreviated subspecies and name
+(\emph{e.g.~domesticus\textsuperscript{WSB}}). Mice were housed at
 dedicated, temperature-controlled facilities in the UW-Madison School of
@@
 Island, which were housed in a temperature-controlled facility in the
-UW-Madison School of Veterinary Medicine. Mice from an inbred strain of
-Gough Island mice were sampled after \textbf{XX} generations of
-brother-sister mating. All mice were provided with ad libitum food and
-water. Procedures followed protocols approved by IACUC.
+UW-Madison School of Veterinary Medicine. Mice were sampled from a
+partially inbred strain of Gough Island mice, approximately 6
+generations of brother-sister matting. All mice were provided with
+\emph{ad libitum} food and water. Procedures followed protocols approved
+by IACUC.
 
@@
 The same dry-down spread technique was applied to both spermatocytes and
-oocytes, following (Peters et al., 1997), with adjustment for volumes.
+oocytes, following Peters et al. (1997), with adjustment for volumes.
 Spermatocyte spreads were collected and prepared as described in
-(Peterson et al., 2019). The majority of mice used for MLH1 counts were
-between 5 and 12 weeks of age. Juvenile mice between 12 and 15 days of
+Peterson et al. (2019). The majority of mice used for MLH1 counts were
+between 5 and 12 weeks of age. Juvenile males between 12 and 15 days of
 age were used for DMC1 counts. Both ovaries were collected from embryos
@@
 following morning, slides were treated with a Photoflow wash (Kodak,
-cite). Slides were stored at -20*C if not stained immediately. To
-visualize the structure of meiotic chromosomes, we used antibody markers
-for the centromere (CREST) and lateral element of the synaptonemal
-complex (SC) (SYCP3). Crossovers (COs) were visualized as MLH1 foci.
-Double strand breaks (DSBs) were visualized as DMC1 foci. The staining
-protocol followed (Anderson et al., 1999) and (Koehler et al., 2002).
-Antibody staining and slide blocking were performed in 1X antibody
-dilution buffer (ADB) (normal donkey serum (Jackson ImmunoResearch), 1X
-PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)). Following
-a 30-minute blocking wash in ABD, each slide was incubated with 60ul of
-a primary antibody master mix for 48 hours at 37*C. The master mix
-recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted
-1:50) or anti-rabbit anti-DMC1 (mix of DMC1), anti-goat polyclonal
-anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to
-CREST (Antibodies, Inc; diluted 1:200) suspended in ADB. Slides were
-washed twice in 50ml ADB before the first round of secondary antibody
-incubation for 12 hours at 37*C. Alexa Fluor 488 donkey anti-rabbit IgG
-(Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey
-anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended
-in ADB. The last incubation of Alexa Fluor 568 donkey anti-goat
-(Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37* C.
-Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours
-after a final wash in 1x PBS.
+diluted 1:200). Slides were stored at -20*C if not stained immediately.
+To visualize the structure of meiotic chromosomes, we used antibody
+markers for the centromere (CREST) and lateral element of the
+synaptonemal complex (SC) (SYCP3). Crossovers (COs) were visualized as
+MLH1 foci. Double strand breaks (DSBs) were visualized as DMC1 foci. The
+staining protocol followed (Anderson et al., 1999) and (Koehler et al.,
+2002). Antibody staining and slide blocking were performed in 1X
+antibody dilution buffer (ADB) (normal donkey serum (Jackson
+ImmunoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100
+(Sigma)). Following a 30-minute blocking wash in ABD, each slide was
+incubated with 60ul of a primary antibody master mix for 48 hours at
+37*C. The master mix recipe contained polyclonal anti-rabbit anti-MLH1
+(Calbiochem; diluted 1:50) or anti-rabbit anti-DMC1 (mix of DMC1),
+anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human
+polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended
+in ADB. Slides were washed twice in 50ml ADB before the first round of
+secondary antibody incubation for 12 hours at 37*C. Alexa Fluor 488
+donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and
+Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to
+1:200) were suspended in ADB. The last incubation of Alexa Fluor 568
+donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for
+2 hours at 37* C. Slides were fixed with Prolong Gold Antifade
+(Invitrogen) for 24 hours after a final wash in 1x PBS. Three slides of
+cell spreads per mouse were prepared to serve as technical replicates
+for the staining protocol. Comparisons of multiple stained slides from
+the same mouse showed no difference in mean MLH1 cell counts and mean
+cell quality. Sampled numbers of mice and cells per mouse were maximized
+to the extent possible given constraints on breeding and time.
 
@@
 Images were captured using a Zeiss Axioplan 2 microscope with AxioLab
-camera and AxioVision software (Zeiss, Cambridge, UK). Preprocessing,
-including cropping, noise reduction, and histogram adjustments, was
-performed using Photoshop (v13.0). Image file names were anonymized
-before manual scoring of MLH1 foci or DMC1 foci using Photoshop.
-
-\hypertarget{analysis}{%
-\subsubsection{Analysis}\label{analysis}}
-
-\textless new 1/n\textgreater{}
+camera and AxioVision software (Zeiss, Cambridge, UK). The number of
+cells imaged per individual mouse is based on previous studies (Dumont
+and Payseur, 2011; Murdoch et al., 2010; Wang and Payseur, 2017).
+Preprocessing, including cropping, noise reduction, and histogram
+adjustments, was performed using Photoshop (v13.0). Image file names
+were anonymized before manual scoring of MLH1 foci or DMC1 foci using
+Photoshop.
+
+\hypertarget{analyses}{%
+\subsubsection{Analyses}\label{analyses}}
 
 To estimate the number of crossovers across the genome, we counted MLH1
-foci. MLH1 foci were counted in cells with intact and complete
-karyotypes (19 acrocentric bivalents and XY for spermatocytes; 20
-acrocentric bivalents for oocytes) and distinct MLH1 foci. A quality
-score ranging from 1 (best) to 5 (worst) was assigned to each cell based
-on visual appearance of staining and spread of bivalents. Cells with a
-score of 5 were excluded from the final analysis. Distributions of MLH1
-count per cell were visually inspected for normality (Supplemental
-Figure 1). MLH1 foci located on the XY in spermatocytes were excluded
-from counts. In addition to MLH1 counts, we measured several traits to
-further characterize the recombination landscape. To estimate the number
-of double-strand breaks, a minority of which lead to crossovers, mean
-DMC1 foci per cell was quantified for a single male from each of a
-subset of strains (\emph{molossinus\textsuperscript{MSM}},
+foci within bivalents, synapsed homologous chromosomes. MLH1 foci were
+counted in pachytene cells with intact and complete karyotypes (19
+acrocentric bivalents and XY for spermatocytes; 20 acrocentric bivalents
+for oocytes) and distinct MLH1 foci. A quality score ranging from 1
+(best) to 5 (worst) was assigned to each cell based on visual appearance
+of staining and spread of bivalents. Cells with a score of 5 were
+excluded from the final analysis. Distributions of MLH1 count per cell
+were visually inspected for normality (Supplemental Figure 1). When
+outliers for MLH1 count were found during preliminary analysis, the
+original images were inspected and the counts confirmed.
+
+MLH1 foci located on the XY in spermatocytes were excluded from counts.
+In addition to MLH1 counts, we measured several traits to further
+characterize the recombination landscape. To estimate the number of
+double-strand breaks, a minority of which lead to crossovers, mean DMC1
+foci per cell was quantified for a single male from each of a subset of
+strains (\emph{molossinus\textsuperscript{MSM}},
 \emph{musculus\textsuperscript{PWD}},
@@
 
-\textless new 2/n\textgreater{} To measure bivalent SC length, two image
-analysis algorithms were used. The first algorithm estimates the total
-(summed) SC length across bivalents for individual cells (Wang et al.
-(2019)). The second algorithm estimates the SC length of individual
-bivalents (Peterson et al. (2019)). Both algorithms apply a
-`skeletonizing' transformation to synapsed chromosomes that produces a
-single, pixel-wide `trace' of the bivalent shape. Total SC length per
-cell was quantified from pachytene cell images (Wang et al. (2019)).
+To measure bivalent SC length, two image analysis algorithms were used.
+The first algorithm estimates the total (summed) SC length across
+bivalents for individual cells (Wang et al., 2019). The second algorithm
+estimates the SC length of individual bivalents (Peterson et al., 2019).
+Both algorithms apply a `skeletonizing' transformation to synapsed
+chromosomes that produces a single, pixel-wide `trace' of the bivalent
+shape. Total SC length per cell was quantified from pachytene cell
+images (Wang et al., 2019).
 
@@
 identified at the mouse level and removed from the data set. Mouse
-averages were calculated from cell-wide total SC lengths in 3,204 out of
-3,884 cells with MLH1 counts. SC length of individual bivalents was
-quantified in pachytene cell images (Peterson et al. (2019)). The DNA
-CrossOver algorithm (Peterson et al. (2019)) isolates single,
+averages were calculated from cell-wide total SC lengths in 3,195 out of
+3,871 cells with MLH1 counts. SC length of individual bivalents was
+quantified in pachytene cell images (Peterson et al., 2019). The DNA
+CrossOver algorithm (Peterson et al., 2019) isolates single,
 straightened bivalent shapes, returning SC length, location of MLH1
@@
 to 0.72 (\emph{musculus\textsuperscript{KAZ}} female). From the total
-set of pachytene cell images, 10,222 bivalent objects were isolated by
-the algorithm. Following a manual curation, 9,575 single-bivalent
+set of pachytene cell images, 10,213 bivalent objects were isolated by
+the algorithm. Following a manual curation, 9,569 single-bivalent
 observations remained. The accuracy of the algorithm is high compared to
 hand measures after this curation step (Peterson et al., 2019). The
-curated single bivalent data supplements our cell-wide MLH1 count data
+curated single bivalent data supplemented our cell-wide MLH1 count data
 with MLH1 foci counts for single bivalents. Proportions of bivalents
@@
 
-\textless new 3/n\textgreater{} To account for confounding effects of
-sex chromosomes from pooled samples of bivalents, we also considered a
-reduced data set including only bivalents with SC lengths below the 2nd
-quartile in cells with at least 17 of 20 single bivalent measures. This
-``short bivalent'' data set included the four or five shortest bivalents
-and excluded the X bivalent in oocytes. A total of 704 short bivalents
-were isolated from 102 oocytes and 43 spermatocytes. Although this
-smaller data set has decreased power, it offers a more comparable set of
-single bivalents to compare between the sexes. A ``long bivalent'' data
-set was formed from those bivalents above the 4th quartile in SC lengths
-per cell. A total of 709 long bivalents were isolated from 102 oocytes
-and 43 spermatocytes.
-
-\textless new 4/n\textgreater{} To examine crossover interference, the
-distance (in SC units) between MLH1 foci (inter-focal distance;
-IFD\textsubscript{raw}) was measured for those single bivalents
-containing two MLH1 foci. A normalized measure of interference
-(IFD\textsubscript{norm}) was computed by dividing
-IFD\textsubscript{raw} by SC length on a per-bivalent basis.
-
-\textless new 5/n\textgreater{} We used a series of statistical models
-to interpret patterns of variation in the recombination traits we
-measured (Table 2). We used mouse average as the dependent variable in
-all analyses. We first constructed a linear mixed model (M1) using
-lmer() from the lmer4 package (Bates et al., 2015) in R (v3.5.2) (Team,
-2015). In this model, strain was coded as a random effect, with
-significance evaluated using a likelihood ratio test (using exactRLRT()
-from RLRsim (Scheipl et al., 2008)). Subspecies, sex, and their
-interaction were coded as fixed effects, with significance evaluated
-using a chi-square test comparing the full and reduced models (drop1()
-and anova()) (Bates et al., 2015). As each observation was at the level
-of a single mouse (mouse average), it was uniquely coded within the
-dataset, nesting was implicit, each strain only occurs within one
-strain, and was not explicitly coded in our linear and mixed models. We
-used the subspecies effect to quantify divergence between subspecies and
-the (random) strain effect to quantify variation within subspecies in a
-sex-specific manner. In separate analyses using model M1, we considered
-mouse averages as dependent variables for each of the following traits:
-MLH1 count per cell, total SC length per cell, single bivalent SC length
-per cell, IFD\textsubscript{raw}, IFD\textsubscript{norm}, and average
-MLH1 position (for single-focus bivalents). Four additional lnear models
+To account for confounding effects of sex chromosomes from pooled
+samples of bivalents, we also considered a reduced data set including
+only bivalents with SC lengths below the 2nd quartile in cells with at
+least 17 of 20 single bivalent measures. This ``short bivalent'' data
+set included the four or five shortest bivalents within a cell, thus
+excluding the X bivalent in oocytes. A total of 699 short bivalents were
+isolated from 102 oocytes and 42 spermatocytes. Although this smaller
+data set had decreased power, it offered a more comparable set of single
+bivalents to compare between the sexes. A ``long bivalent'' data set was
+formed from those bivalents above the 4th quartile in SC lengths per
+cell. A total of 703 long bivalents were isolated from 102 oocytes and
+42 spermatocytes.
+
+To examine crossover interference, the distance (in SC units) between
+MLH1 foci (inter-focal distance; IFD\textsubscript{raw}) was measured
+for those single bivalents containing two MLH1 foci. A normalized
+measure of interference (IFD\textsubscript{norm}) was computed by
+dividing IFD\textsubscript{raw} by SC length on a per-bivalent basis.
+
+We used a series of statistical models to interpret patterns of
+variation in the recombination traits we measured (Table 2). We used
+mouse average as the dependent variable in all analyses. We first
+constructed a linear mixed model (M1) using lmer() from the lmer4
+package (Bates et al., 2015) in R (v3.5.2) (Team, 2015). In this model,
+strain was coded as a random effect, with significance evaluated using a
+likelihood ratio test using exactRLRT() from RLRsim (Scheipl et al.,
+2008). Subspecies, sex, and their interaction were coded as fixed
+effects, with significance evaluated using a chi-square test comparing
+the full and reduced models (drop1() and anova()) (Bates et al., 2015).
+The hierarchical nature of the data, meant that nesting of levels across
+observations was implicit (\emph{i.e.} mouse within strain, within
+subspecies) and not explicitly coded. We used the subspecies effect to
+quantify divergence between subspecies and the (random) strain effect to
+quantify variation within subspecies in a sex-specific manner. In
+separate analyses using model M1, we considered mouse averages as
+dependent variables for each of the following traits: MLH1 count per
+cell, total SC length per cell, single bivalent SC length per cell,
+IFD\textsubscript{raw}, IFD\textsubscript{norm}, and average MLH1
+position (for single-focus bivalents). Four additional linear models
 containing only fixed effects (M2-M5) (Table 2) were used to further
@@
 
-\textless new 1/n\textgreater{}
-
 We used counts of MLH1 foci per cell to estimate genome-wide
 recombination rates in 14 wild-derived inbred strains sampled from three
-subspecies of house mice (\emph{M. m. domesticus}, \emph{M. m. musculus}
-and \emph{M. m. molossinus} ) and three additional species of Mus
-(\emph{M. spretus}, \emph{M. spicilegus} , and \emph{M. caroli}). Mean
-MLH1 focus counts for 186 mice were quantified from an average of 21.97
-spermatocytes per male (for a total of 1,714 spermatocytes) and 18
-oocytes per female (for a total of 1,427 oocytes).
-
-\textless new 2/n\textgreater{}
+subspecies of house mice ( \emph{M. musculus domesticus}, \emph{M. m.
+musculus} and \emph{M. m. molossinus} ) and three other species of Mus (
+\emph{M. spretus}, \emph{M. spicilegus} , and \emph{M. caroli} ). Mean
+MLH1 focus counts for 188 mice were quantified from an average of 21.77
+spermatocytes per male (for a total of 1,742 spermatocytes) and 17.85
+oocytes per female (for a total of 1,427 oocytes) (Table 1).
 
@@
 counts from females and males are uncorrelated (Spearman's \(\rho\) =
-0.08 - 0.08; p = 0.84 -- 0.84) across the set of strains.
-
-\textless new 3/n\textgreater{}
+0.08; p = 0.84) across the set of strains.
+
+\begin{table}[H]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{l|l|l|l|l|l|l|l|l|l}
+\hline
+Species & Subspecies & Strain & Sex & Number.of.Mice & Number.of..Cells & Mean.MLH1.Count & SE & cV & Variance\\
+\hline
+M. musculus & M. m. domesticus & WSB & female & 14 & 184 & 24.701 & 0.267 & 14.638271224837814 & 13.074\\
+\cline{1-10}
+ &  &  & male & 11 & 222 & 23.383 & 0.18 & 11.479933887268658 & 7.206\\
+\cline{3-10}
+ &  & G & female & 12 & 318 & 28.211 & 0.235 & 14.835969061582837 & 17.517\\
+\cline{3-10}
+ &  &  & male & 18 & 355 & 23.161 & 0.14 & 11.35380583435758 & 6.915\\
+\cline{3-10}
+ &  & LEW & female & 9 & 147 & 26.585 & 0.398 & 18.16189516714001 & 23.313\\
+\cline{3-10}
+ &  &  & male & 10 & 253 & 24.162 & 0.195 & 12.83701437670674 & 9.62\\
+\cline{3-10}
+ & \multirow{-6}{*}{\raggedright\arraybackslash } & PERC & male & 1 & 26 & 21.808 & 0.41 & 9.71 & 4.48\\
+\cline{2-10}
+ & M. m. musculus & PWD & female & 15 & 222 & 25.977 & 0.251 & 14.41018621027294 & 14.013\\
+\cline{2-10}
+ &  &  & male & 8 & 161 & 28.671 & 0.246 & 10.896284161204306 & 9.76\\
+\cline{3-10}
+ &  & SKIVE & female & 1 & 32 & 25.938 & 0.553 & 12.070868672504577 & 9.802\\
+\cline{3-10}
+ &  &  & male & 3 & 86 & 26.081 & 0.293 & 10.408689886052935 & 7.37\\
+\cline{3-10}
+ &  & KAZ & female & 9 & 184 & 25.625 & 0.295 & 15.62873696060005 & 16.039\\
+\cline{3-10}
+ &  &  & male & 13 & 264 & 22.989 & 0.186 & 13.159636243359548 & 9.152\\
+\cline{3-10}
+ &  & CZECH & male & 3 & 62 & 22.3 & 0.32 & 11.21 & 6.25\\
+\cline{3-10}
+ &  & AST & male & 3 & 63 & 24.41 & 0.33 & 10.65 & 6.76\\
+\cline{3-10}
+ & \multirow{-8}{*}{\raggedright\arraybackslash } & TOM & male & 2 & 10 & 23.7 & 1.18 & 15.79 & 14.01\\
+\cline{2-10}
+ & M. m. castaneus & CAST & female & 1 & 1 & 26 & NA &  & \\
+\cline{2-10}
+ &  &  & male & 2 & 44 & 22 & 0.34 & 10 & 5.2\\
+\cline{3-10}
+ & \multirow{-2}{*}{\raggedright\arraybackslash } & HMI & male & 4 & 44 & 24 & 0.41 & 11 & 7.5\\
+\cline{2-10}
+ & M. m. molossinus & MSM & female & 14 & 300 & 28.123 & 0.254 & 15.64239374084605 & 19.353\\
+\cline{2-10}
+ &  &  & male & 7 & 166 & 30.367 & 0.242 & 10.262658213499735 & 9.713\\
+\cline{3-10}
+ &  & MOLF & female & 1 & 21 & 27.619 & 0.924 & 15.33891788564294 & 17.948\\
+\cline{3-10}
+\multirow{-22}{*}{\raggedright\arraybackslash } &  &  & male & 6 & 119 & 23.42 & 0.232 & 10.80040516644884 & 6.398\\
+\cline{1-1}
+\cline{3-10}
+Mus spretus &  & SPRET & female & 2 & 2 & 26 & 2 & 10.878565864408424 & 8\\
+\cline{1-1}
+\cline{3-10}
+ &  &  & male & 5 & 103 & 24.427 & 0.246 & 10.232121984584271 & 6.247\\
+\cline{1-1}
+\cline{3-10}
+Mus spicilegus &  & SPIC & female & 6 & 97 & 28.237 & 0.448 & 15.628316890383017 & 19.474\\
+\cline{1-1}
+\cline{3-10}
+ &  &  & male & 4 & 133 & 25.774 & 0.241 & 10.781039022222362 & 7.721\\
+\cline{1-1}
+\cline{3-10}
+Mus caroli & \multirow{-8}{*}{\raggedright\arraybackslash } & CAROLI & male & 2 & 57 & 27 & 0.4 & 11 & 8.9\\
+\hline
+\end{tabular}}
+\end{table}
 
 To further partition variation in recombination rate, we fit a series of
-linear models to mean MLH1 focus counts from 137 house mice (Table 2;
-detailed results available in Supplemental Table X). Strain, sex,
-subspecies, and sex*subspecies each affect MLH1 focus count in a linear
-mixed model (M1; strain (random effect): p \textless{} 10-4 --
-\ensuremath{10^{-4}}; sex: p = 6.2x10-4 --
-\ensuremath{4.95\times 10^{-6}}; subspecies: p = 1.1x10-3 -- 0;
-subspecies*sex: p = 2x10-4 -- \ensuremath{2.1\times 10^{-4}}).
+linear models to mean MLH1 focus counts from 137 house mice from three
+subspecies; \emph{M. m. domesticus}, \emph{M. m. musculus} and \emph{M.
+m. molossinus} (Table 2; detailed results available in Supplemental
+Table X). Strain, sex, subspecies, and sex*subspecies each affect MLH1
+focus count in a linear mixed model (M1; strain (random effect): p
+\textless{} \ensuremath{10^{-4}}-- 10-4; sex: p =
+\ensuremath{3.64\times 10^{-6}}; subspecies: p =
+\ensuremath{9.69\times 10^{-4}}; subspecies*sex: p =
+\ensuremath{1.8\times 10^{-4}}).
 
 The effect of subspecies is no longer significant in a model treating
-all factors as fixed effects \textbf{(M2; p = 0.09 -- )}, highlighting
-strain and sex as salient variables. Two strains exhibit particularly
-strong effects on MLH1 focus count (M3;
-\emph{domesticus\textsuperscript{G}} p = \ensuremath{2.19\times 10^{-6}}
-; \emph{domesticus\textsuperscript{LEW}} p = 0.02), with sex-strain
-interactions involving three strains (M3;
-\emph{domesticus\textsuperscript{G}} p = 0;
-\emph{molossinus\textsuperscript{MSM}} p = 0 ;
+all factors as fixed effects (M2; \emph{musculus} p = 0.24,
+\emph{molossinus} p = 0.1), highlighting strain and sex as salient
+variables. Two strains exhibit strong effects on MLH1 focus count (M3;
+\emph{domesticus\textsuperscript{G}} p =
+\ensuremath{1.78\times 10^{-6}}; \emph{domesticus\textsuperscript{LEW}}
+p = 0.02), with sex-strain interactions involving three strains (M3;
+\emph{domesticus\textsuperscript{G}} p \textless{} 10-6 -- 0;
+\emph{molossinus\textsuperscript{MSM}} p \textless{} 10-6 -- 0;
 \emph{musculus\textsuperscript{PWD}} p =
-\ensuremath{5.31\times 10^{-4}}).
-
-In separate analyses of males (M4; n = 80 - 71), three strains
+\ensuremath{3.87\times 10^{-4}}). \textbf{fix the pvalue above, change
+the 0 to p \textless{} 10\^{}-6}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth,height=\textheight]{Figure2.jpg}
+\caption{Caption for the picture.}
+\end{figure}
+
+In separate analyses of males (M4; n = 71), three strains
 disproportionately shape MLH1 focus count (as observed in Figure 1C):
 \emph{musculus\textsuperscript{PWD}} (p =
-\ensuremath{3.08\times 10^{-7}} ; effect = 6.11 ),
+\ensuremath{3.6\times 10^{-7}}; effect = 6.11 foci,
 \emph{molossinus\textsuperscript{MSM}} (p =
-\ensuremath{3.52\times 10^{-9}}; effect = 7.02), and
+\ensuremath{6.3\times 10^{-9}}; effect = 6.91), and
 \emph{musculus\textsuperscript{SKIVE}} (p =
-\ensuremath{7.59\times 10^{-4}}; effect = 4.04 ). These three strains
+\ensuremath{8.22\times 10^{-4}}; effect = 4.04). These three strains
 point to substantial evolution in the genome-wide recombination rate in
 spermatocytes; we subsequently refer to them as ``high-recombination''
-strains.
-
-In females (M4; n= 76 - 76), four strains affect MLH1 focus count:
+strains. In females (M4; n= 76 ), three strains affect MLH1 focus count:
 \emph{domesticus\textsuperscript{G}} (p =
-\ensuremath{1.16\times 10^{-5}}; effect = 3.28),
+\ensuremath{8.7\times 10^{-6}}; effect = 3.3),
 \emph{molossinus\textsuperscript{MSM}} (p =
-\ensuremath{2.97\times 10^{-5}}; effect = 2.98),
-\emph{domesticus\textsuperscript{LEW}} (p = 0.03 ; effect = 1.7), and
-\emph{musculus\textsuperscript{PWD}} (p = 0.08; effect = 1.16). Strain
+\ensuremath{2.43\times 10^{-5}}; effect = 2.99), and
+\emph{domesticus\textsuperscript{LEW}} (p = 0.03; effect = 1.69). Strain
 effect sizes in females are modest in magnitude compared to those in
-males.
-
-Together, these results demonstrate that the genome-wide recombination
-rate evolves in a highly sex-specific manner.
+males. Together, these results demonstrate that the genome-wide
+recombination rate evolves in a highly sex-specific manner.
 
@@
 
-\textless new 1/n\textgreater{} The variation in sex differences in
-recombination we discovered provided an opportunity to determine whether
-sex differences in chromatin compaction, as measured by the length of
-the synaptonemal complex (SC), are reversed when heterochiasmy is
-reversed. In all strains except \emph{musculus\textsuperscript{SKIVE}},
-females have longer SCs than males, whether SC length was estimated as
-the total length across bivalents or as the length of short bivalents
-(t-tests; all p \textless{} 0.05, except short bivalents in
-\emph{musculus\textsuperscript{SKIVE}}, p = 0.11 -- 0.11`).
-
-Among short bivalents (to which the female X bivalent does not
-contribute), female to male ratios of mouse mean SC length range from
-1.26 - 1.26 (\emph{musculus\textsuperscript{PWD}}) to 1.49 - 1.49
+The variation in sex differences in recombination we discovered provided
+an opportunity to determine whether sex differences in chromatin
+compaction, as measured by the length of the synaptonemal complex (SC),
+are reversed when heterochiasmy is reversed. In all strains except
+\emph{musculus\textsuperscript{SKIVE}}, females have longer SCs than
+males, whether SC length was estimated as the total length across
+bivalents or as the length of short bivalents (t-tests; all p
+\textless{} 0.05, except short bivalents in
+\emph{musculus\textsuperscript{SKIVE}}, p = 0.11). Among short bivalents
+(to which the female X bivalent does not contribute), female to male
+ratios of mouse mean SC length range from 1.26
+(\emph{musculus\textsuperscript{PWD}}) to 1.52
 (\emph{domesticus\textsuperscript{WSB}}) across strains. That females
@@
 which identify sex as the most consistently significant effect for total
-SC length (M1: p = 7.16\^{}\{-20\} -- \ensuremath{1.2\times 10^{-31}} ;
-M2: p = 5.33\^{}\{-4\} -- \ensuremath{1.95\times 10^{-8}}; M3: p =0.05
--- 0.01 ) and short bivalent length (M1: p = 1.27\^{}\{-11 --
-\ensuremath{1.27\times 10^{-11}}\}; M2: p = 1.98\^{}\{-7\} --
-\ensuremath{1.98\times 10^{-7}}; M3: p = 1.98\^{}\{-7\} --
-\ensuremath{1.98\times 10^{-7}}).
-
-The existence of some subspecies and strain effects (M1 subspecies*sex:
-p = 0.03 -- 0.03; M2: strain-\emph{musculus\textsuperscript{SKIVE}} :
-p=0.02 -- 0.02 ; \emph{musculus\textsuperscript{SKIVE}}*male: p = 0.08
--- 0.08) further indicates that SC length has evolved among strains and
-among subspecies.
+SC length (M1: p = \ensuremath{2.56\times 10^{-31}}; M2: p =
+\ensuremath{2.56\times 10^{-8}}; M3: p =
+\ensuremath{2.56\times 10^{-8}}) and short bivalent SC length (M1: p =
+\ensuremath{1.12\times 10^{-11}}; M2: p \textless{} 0; M3: p \textless{}
+0). The existence of some subspecies and strain effects on total SC
+length and short bivalent SC length (Supplemental Tables 5 and 8)
+further indicates that SC length has evolved among strains and among
+subspecies.
 
@@
 
-\textbf{double check that these are the only sig factors for all the
-models}
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth,height=\textheight]{Fig3_SC.jpg}
+\caption{Caption for the picture.}
+\end{figure}
 
@@
 
-\textless new 1/n\textgreater{} We used normalized positions of MLH1
-foci along bivalents with a single focus to compare crossover location
-while controlling for differences in SC length. In all strains, MLH1
-foci tend to be closer to the telomere in males (mean normalized
-position in males: 0.68--0.68; mean normalized position in females: 0.56
--- 0.56; t-test; p = 3.84\^{}\{-23\} --
-\ensuremath{3.84\times 10^{-23}}). Sex is also the strongest determinant
-of MLH1 focus position in the models we tested (M1: p = 7.79\^{}\{-26 --
-\ensuremath{7.79\times 10^{-26}}\}; M2: p = 5.39\^{}\{-8\} --
-\ensuremath{5.39\times 10^{-8}}; M3: p = 5.39\^{}\{-8\} --
-\ensuremath{5.39\times 10^{-8}}).
-
-\textless new 2/n\textgreater{} Males have longer normalized mean
-inter-focal distances (IFD\textsubscript{norm}) than females in seven
-out of eight strains (t-tests; p \textless{} 0.05), with only
-\emph{musculus\textsuperscript{KAZ}} showing no difference (p = 0.33 --
-0.33). Examination of IFD\textsubscript{norm} distributions indicates
-that females are centered at approximately 50\% and show a slight
-enrichment of low (\textless25\%) values, whereas males are enriched for
-higher values. Models treating IFDnorm as the dependent variable support
-the inference of stronger interference in males, with sex being the most
-significant variable (M1: p = 3.11\^{}\{-12\} --
-\ensuremath{3.11\times 10^{-12}} ; M2: p = 0.01 -- 0.01; M3: p = 0.01 --
-0.01). In contrast, there is no clear signal of sex differences in raw
-mean inter-focal distances (IFD\textsubscript{raw}) across the full set
-of strains, whether they are considered separately or together.
-Visualization of normalized MLH1 foci positions on bivalents with two
-crossovers (Figure 4C) further suggests that interference distances vary
-more in females than in males, and that males display a stronger
-telomeric bias in the placement of the distal crossover.
+We used normalized positions of MLH1 foci along bivalents with a single
+focus to compare crossover location while controlling for differences in
+SC length. In all strains, MLH1 foci tend to be closer to the telomere
+in males (mean normalized position in males: 0.68; mean normalized
+position in females: 0.56; paired t-test; p =
+\ensuremath{8.49\times 10^{-4}}). Sex is also the strongest determinant
+of MLH1 focus position in the models we tested (M1: p =
+\ensuremath{2.82\times 10^{-26}}; M2: p =
+\ensuremath{3.96\times 10^{-8}}; M3: p =
+\ensuremath{3.96\times 10^{-8}}).
+
+Males have longer normalized mean inter-focal distances
+(IFD\textsubscript{norm}) than females in seven out of eight strains
+(t-tests; p \textless{} 0.05), with only
+\emph{musculus\textsuperscript{KAZ}} showing no difference (p = 0.33).
+Examination of IFD\textsubscript{norm} distributions indicates that
+females are centered at approximately 50\% and show a slight enrichment
+of low (\textless25\%) values, whereas males are enriched for higher
+values. Models treating IFD\textsubscript{norm} as the dependent
+variable support the inference of stronger interference in males, with
+sex being the most significant variable (M1: p =
+\ensuremath{9.08\times 10^{-12}}; M2: p = 0.01; M3: p = 0.01). In
+contrast, there is no clear signal of sex differences in raw mean
+inter-focal distances (IFD\textsubscript{raw}) (Supplemental Table 13
+and 14) across the full set of strains, whether they are considered
+separately or together. Visualization of normalized MLH1 foci positions
+on bivalents with two crossovers (Figure 3C and Supplemental Figure 3)
+further suggests that interference distances vary more in females than
+in males, and that males display a stronger telomeric bias in the
+placement of the distal crossover.
 
 In summary, controlling for differences in SC length (chromatin
-compaction) indicates that interference is stronger in males, whereas
-interference on the physical scale is similar in the sexes.
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+compaction) indicates that interference is consistently stronger in
+males, whereas interference on the physical scale is similar in the two
+sexes.
 
@@
 
-\textless new 1/n\textgreater{} We used the contrast between males from
-high-recombination strains and males from low-recombination strains to
-identify features of the recombination landscape associated with
-evolutionary transitions in the genome-wide recombination rate. We
-considered proportions of bivalents with different numbers of
-crossovers, double-strand break number, SC length, and crossover
-positioning.
-
-\textless new 2/n\textgreater{} Ninety-six percent of single bivalents
-in our pooled dataset (n = 9,576 -- 9,575 ) have either one or two MLH
-foci (Figure 3). The proportions of single-focus (1CO) bivalents
-vs.~double-focus (2CO) bivalents distinguish high-recombination strains
-from low-recombination strains (Figure 3). High-recombination strains
-are enriched for 2CO bivalents at the expense of 1CO bivalents:
-proportions of 2CO bivalents are 0.33 -- 0.33 in
-\emph{musculus\textsuperscript{SKIVE}}, 0.44 -- 0.44 in
-\emph{musculus\textsuperscript{PWD}} , and 0.51 -- 0.51 in
+We used the contrast between males from high-recombination strains and
+males from low-recombination strains to identify features of the
+recombination landscape associated with evolutionary transitions in the
+genome-wide recombination rate. We considered proportions of bivalents
+with different numbers of crossovers, double-strand break number, SC
+length, and crossover positioning.
+
+Ninety-six percent of single bivalents in our pooled dataset (n = 9,569)
+have either one or two MLH foci (Supplemental Figure 2). The proportions
+of single-focus (1CO) bivalents vs.~double-focus (2CO) bivalents
+distinguish high-recombination strains from low-recombination strains
+(Supplemental Figure 2). High-recombination strains are enriched for 2CO
+bivalents at the expense of 1CO bivalents: proportions of 2CO bivalents
+are 0.33 in \emph{musculus\textsuperscript{SKIVE}}, 0.44 in
+\emph{musculus\textsuperscript{PWD}}, and 0.51 in
 \emph{molossinus\textsuperscript{MSM}} (Supplemental Figure 3).
@@
 (chi-square tests; \emph{musculus\textsuperscript{PWD}}
-vs.~\emph{molossinus\textsuperscript{MSM}}: p = 0.37 -- 0.37 ;
+vs.~\emph{molossinus\textsuperscript{MSM}}: p = 0.37;
 \emph{musculus\textsuperscript{PWD}}
-vs.~\emph{musculus\textsuperscript{KAZ}}: p = 1.23\^{}\{-31\} --
+vs.~\emph{musculus\textsuperscript{KAZ}}: p =
 \ensuremath{1.23\times 10^{-31}}; \emph{molossinus\textsuperscript{MSM}}
-vs.~\emph{molossinus\textsuperscript{MOLF}}: p = 2.65\^{}\{-6\} --
-\ensuremath{2.82\times 10^{-6}} ). These results demonstrate that
+vs.~\emph{molossinus\textsuperscript{MOLF}}: p =
+\ensuremath{2.34\times 10^{-6}}). These results demonstrate that
 evolution of the genome-wide recombination rate reflects changes in
@@
 
-\textless new 3/n\textgreater{} To begin to localize evolution of
-genome-wide recombination rate to steps of the recombination pathway, we
-counted DMC1 foci in prophase spermatocytes as markers for double-strand
-breaks (DSBs). DMC1 foci were counted in a total of 76 -- 76 early
-zygotene and 75 -- 75 late zygotene spermatocytes from two
-high-recombination strains (\emph{musculus\textsuperscript{PWD}} and
+To begin to localize evolution of genome-wide recombination rate to
+steps of the recombination pathway, we counted DMC1 foci in prophase
+spermatocytes as markers for double-strand breaks (DSBs). DMC1 foci were
+counted in a total of 76 early zygotene and 75 late zygotene
+spermatocytes from two high-recombination strains
+(\emph{musculus\textsuperscript{PWD}} and
 \emph{molossinus\textsuperscript{MSM}}) and three low-recombination
@@
 \emph{domesticus\textsuperscript{WSB}}, and
-\emph{domesticus\textsuperscript{G}}). High-recombination strains have
-significantly more DMC1 foci than low-recombination strains in early
-zygotene cells (t-test; p \textless{} 10\^{}\{-6\} --
-\ensuremath{10^{-6}} ). In contrast, the two strain groups do not differ
-in DMC1 foci in late zygotene cells (t-test; p = 0.66 -- 0.66 ). Since
-DSBs are repaired as either COs or non-crossovers (NCOs), the ratio of
-MLH1 foci to DMC1 foci can be used to estimate the proportion of DSBs
-designated as COs. High-recombination and low-recombination strains do
-not differ in the MLH1/DMC1 ratio, whether DMC1 foci were counted in
-early zygotene cells or late zygotene cells (t-test; p \textgreater{}
-0.05). These results raise the possibility that the evolution of
-genome-wide recombination rate is primarily determined by processes that
-precede the CO/NCO decision, at least in house mice.
-
-\textless new 4/n\textgreater{} Total SC length only partially
-differentiates high-recombination strains from low-recombination strains
-(Figure 5). Whereas high-recombination strains as a group have
-significantly greater total SC length than low-recombination strains
-(t-test; p = 0.01 -- 0.02 ), separate tests within subspecies show that
-the two strain categories differ within \emph{M. m. molossinus} (p =
-0.02 -- \ensuremath{4.71\times 10^{-4}} ) but not within \emph{M. m.
-musculus} (p = 0.4 -- 0.65 ). Additionally, mouse means for the reduced
-(short and long) bivalent datasets do not differ between
+\emph{domesticus\textsuperscript{G}}) (Table 3). High-recombination
+strains have significantly more DMC1 foci than low-recombination strains
+in early zygotene cells (t-test; p \textless{} \ensuremath{10^{-6}}). In
+contrast, the two strain groups do not differ in DMC1 foci in late
+zygotene cells (t-test; p = 0.66). Since DSBs are repaired as either COs
+or non-crossovers (NCOs), the ratio of MLH1 foci to DMC1 foci can be
+used to estimate the proportion of DSBs designated as COs.
+High-recombination and low-recombination strains do not differ in the
+MLH1/DMC1 ratio, whether DMC1 foci were counted in early zygotene cells
+or late zygotene cells (t-test; p \textgreater{} 0.05). These results
+raise the possibility that the evolution of genome-wide recombination
+rate is primarily determined by processes that precede the CO/NCO
+decision, at least in house mice.
+
+Total SC length only partially differentiates high-recombination strains
+from low-recombination strains (Figure 3). Whereas high-recombination
+strains as a group have significantly greater total SC length than
+low-recombination strains (t-test; p = 0.01), separate tests within
+subspecies show that the two strain categories differ within \emph{M. m.
+molossinus} (p = \ensuremath{2.59\times 10^{-4}}) but not within
+\emph{M. m. musculus} (p = 0.65). Additionally, mouse means for the
+reduced (short and long) bivalent datasets do not differ between
 high-recombination and low-recombination strains (t-test; short: p =
-0.88 -- 0.88 ; long: p = 0.19 -- 0.19 ).
-
-Write results for M4 and M5 here total SC M4 results
-
-Musc subspecies very \textbf{\emph{, Mol }} (both ++) G, LEW, MSM SKIVE,
-KAZ (- -)
-
-M5 (almost all strains are significant) G, LEW, PWD, MSM MOLF, SKIVE,
-KAZ, TOM, AST, CZECH
-
-In a model with total SC length as the dependent variable (\emph{M2?}),
-two subspecies effects are significant (\emph{M. m. musculus} p =
-2.33\^{}\{-6\} -- 0.65, \emph{M. m. molossinus}, p = 21\^{}\{-6\} --
-\ensuremath{4.71\times 10^{-4}}). In models with SC lengths of short and
+0.84; long: p = 0.19). In a model with total SC length as the dependent
+variable (M4), the two subspecies effects are significant (\emph{M. m.
+musculus} p = \ensuremath{3.95\times 10^{-7}}; \emph{M. m. molossinus} p
+= \ensuremath{3.33\times 10^{-7}}), but there are also strain-specific
+effects (Supplemental Table 6). In models with SC lengths of short and
 long bivalents as dependent variables, several subspecies and strain
-effects reach significance (p \textless{} 0.05), but they are not
-consistent across models. Collectively, these results reveal that
-evolution of SC length is not strongly associated with evolution of
-genome-wide recombination rate in house mice.
-
-\textless new 5/n\textgreater{} In summary, evolution of the genome-wide
-recombination rate in males is connected to double-strand break number
-and crossover interference, but not to SC length and crossover position
-(on single-crossover bivalents).
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+effects reach significance (p \textless{} 0.05) (Supplemental Table 9),
+but they are not consistent across models. Collectively, these results
+reveal that evolution of SC length is not strongly associated with
+evolution of genome-wide recombination rate in house mice.
+
+In summary, evolution of the genome-wide recombination rate in males is
+connected to double-strand break number and crossover interference, but
+not to SC length and crossover position (on single-crossover bivalents).
 
@@
 
-\textless new 1/n\textgreater{} By comparing recombination rates in
-females and males from the same diverse set of genetic backgrounds, we
-isolated sex as a primary factor in the evolution of this fundamental
-meiotic trait. Recombination rate differences are more pronounced in
-males than females. Because inter-strain divergence times are identical
-for the two sexes, this observation demonstrates that the genome-wide
-recombination rate evolves faster in males, at least in house mice. More
-generally, recombination rate divergence is decoupled in females and
-males. These disparities are remarkable given that recombination rates
-for the two sexes were measured in identical genomic backgrounds (other
-than the presence/absence of the Y chromosome). Our results provide the
-strongest evidence yet that the genome-wide recombination rate follows
-distinct evolutionary trajectories in males and females.
-
-\textless new 2/n\textgreater{} At the genetic level, the sex-specific
-evolution we documented indicates that some mutations responsible for
-divergence in recombination rate have dissimilar phenotypic effects in
-females and males. A subset of the genetic variants associated with
-genome-wide recombination rate within populations of humans (Kong et
-al., 2004,, 2008, 2014; Halldorsson et al., 2019), Soay sheep (Johnston
-et al., 2016), and cattle (Ma et al., 2015; Shen et al., 2018) appear to
-show sex-specific properties, including opposite effects in females and
-males. Furthermore, inter-sexual correlations for recombination rate are
-weak in humans (Fledel-Alon et al., 2011) and Soay sheep (Johnston et
-al., 2016). Crosses between the strains we surveyed could be used to
-identify and characterize the genetic variants responsible for
-recombination rate evolution in house mice (Dumont and Payseur, 2011;
-Wang et al., 2019; Wang and Payseur, 2017). These variants could
-differentially affect females and males at any step in the recombination
-pathway. Although our DMC1 profiling was limited to males from a small
-number of strains (for practical reasons), our findings suggest that
-mutations that determine the number of double-strand breaks contribute
-to sex-specific evolution in the recombination rate. A study of two
-classical inbred strains and one wild-derived inbred strain of house
-mice also found a positive association between crossover number and
-double-strand break number in males (Baier et al., 2014).
-
-\textless new 3/n\textgreater{} Another implication of our results is
-that the connection between recombination rate and fitness differs
-between males and females. Little is known about whether and how natural
-selection shapes recombination rate in nature (Dapper and Payseur, 2017;
-Ritz et al., 2017). Samuk et al. (2020) recently used a quantitative
-genetic test to conclude that an 8\% difference in genome-wide
-recombination rate between females from two populations of
-\emph{Drosophila pseudoobscura} was caused by natural selection.
-Applying similar strategies to species in which both sexes recombine,
-including house mice, would be a logical next step to understanding the
-sex-specific evolution of recombination rate.
-
-\textless new 4/n\textgreater{} Population genetic models have been
-built to explain sexual dimorphism in the number and placement of
-crossovers, which is a common phenomenon (Brandvain and Coop, 2012;
-Sardell and Kirkpatrick, 2020). Modifier models predicted that lower
-recombination rates in males will result from haploid selection
-(Lenormand, 2003) or sexually antagonistic selection on coding and
-cis-regulatory regions of genes (Sardell and Kirkpatrick, 2020). Another
-modifier model showed that meiotic drive could stimulate female-specific
-evolution of the recombination rate (Brandvain and Coop, 2012). Although
-these models fit the conserved pattern of sex differences in crossover
-positions, they do not readily explain our observations of sex-specific
-evolution in the genome-wide recombination rate. In particular, the
-alternation across strains in which sex has more crossovers is
-unexpected.
-
-\textless new 5/n\textgreater{} We propose an alternative interpretation
-of our findings based on the cell biology of gametogenesis. During
-meiosis, achieving a stable chromosome structure requires the attachment
-of kinetochores to opposite poles of the cell and at least one crossover
-to create tension across the sister chromosome cohesion distal to
-chiasmata (Lane and Kauppi, 2019; VanVeen and Hawley, 2003). The spindle
-assembly checkpoint (SAC) prevents aneuploidy by ensuring that all
-bivalents are correctly attached to the microtubule spindle
-(``bi-oriented'') before starting the metaphase-to-anaphase transition
-via the release of the sister cohesion holding homologs together (Lane
-and Kauppi, 2019). Hence, selection seems likely to favor mutations that
-optimize the process of bi-orientation and chromosome separation,
-thereby prohibiting the SAC from delaying the cell cycle or triggering
-apoptosis. Multiple lines of evidence indicate that the SAC is more
-effective in spermatogenesis than in oogenesis (Lane and Kauppi, 2019),
-perhaps due to the presence of the centrosome spindle (So et al., 2019)
-and larger cell volume (Kyogoku and Kitajima, 2017) in oocytes. The
-higher stringency of the SAC during spermatogenesis suggests that
-selection will be better at removing mutations that interfere with
-bi-orientation in males than in females. Therefore, faster male
-evolution of the genome-wide recombination rate could be driven by the
-more stringent SAC acting on chromosome structures at the metaphase I
-alignment.
-
-\textless new 5/n\textgreater{} Our SAC model is consistent with other
-features of our data. We showed that widespread sex differences in
-broad-scale crossover positioning (Sardell and Kirkpatrick, 2020) apply
-across house mice, even in lineages where the direction of heterochiasmy
-is reversed. The number and placement of crossovers affects the area of
-sister chromosome cohesion distal to crossovers which needs to be
-released for the first reductional chromosome segregation (Dumont and
-Desai, 2012; Lane and Kauppi, 2019; Subramanian and Hochwagen, 2014;
-VanVeen and Hawley, 2003). Faster spermatogenesis may select for
-synchronization of the separation across all homologs within the cell
-(\textbf{Kudo} ), whereas in oogenesis, the slower cell cycle and
-multiple arrest stages may require chromosome structures with greater
-stability on the MI spindle, especially for those organisms that undergo
-dictyate arrest (Lee, 2019).
-
-\textless new 6/n\textgreater{} We propose that the SAC model also can
-explain the correlated evolution of stronger crossover interference and
-higher genome-wide recombination rate in male house mice. Our results
-show that crossovers are spaced further apart in strains enriched for
-double-crossover bivalents when SC length is taken into account and
-chromosome size effects are minimized. Assuming chromatin compaction
-between (prophase) pachytene and metaphase is uniform along bivalents,
-this increased spacing is expected to expand the area for sister
-cohesion to connect homologs and may improve the fidelity of chromosomal
-segregation. While the SAC model postulates direct fitness effects of
-interference, a modifier model predicted that indirect selection on
-recombination rate -- via its modulation of offspring genotypes -- can
-strengthen interference as well (Goldstein et al., 1993).
-
-\textless new 7/n\textgreater{} Regardless of the underlying mechanism,
-our results provide a rare demonstration that crossover interference can
-diverge over short evolutionary timescales. The notion that stronger
-interference can co-evolve with higher genome-wide recombination rate is
-supported by differences between breeds of cattle (Ma et al. (2015)) and
-differences between populations of white-footed mice (Peterson et al.,
-2019). In contrast, mammalian species with stronger interference tend to
-exhibit lower genome-wide recombination rates (Segura et al. (2013),
-Otto and Payseur (2019)). Collectively, these patterns suggest that
-inferences about the evolutionary dynamics of interference depend on the
-timescale under consideration.
-
-\textless new 8/n\textgreater{} Our findings further reveal that
+By comparing recombination rates in females and males from the same
+diverse set of genetic backgrounds, we isolated sex as a primary factor
+in the evolution of this fundamental meiotic trait. Recombination rate
+differences are more pronounced in males than females. Because
+inter-strain divergence times are identical for the two sexes, this
+observation demonstrates that the genome-wide recombination rate evolves
+faster in males. More generally, recombination rate divergence is
+decoupled in females and males. These disparities are remarkable given
+that recombination rates for the two sexes were measured in identical
+genomic backgrounds (other than the number and identity of sex
+chromosomes). Our results provide the strongest evidence yet that the
+genome-wide recombination rate follows distinct evolutionary
+trajectories in males and females.
+
+At the genetic level, the sex-specific patterns we documented indicate
+that some mutations responsible for the evolution of recombination rate
+have dissimilar phenotypic effects in the two sexes. A subset of the
+genetic variants associated with genome-wide recombination rate within
+populations of humans (Halldorsson et al., 2019; Kong et al., 2004,
+2014, 2008), cattle (Ma et al., 2015; Shen et al., 2018), and Soay sheep
+(Johnston et al., 2016) appear to show sex-specific properties,
+including opposite effects in females and males. Furthermore,
+inter-sexual correlations for recombination rate are weak in humans
+(Fledel-Alon et al., 2011) and Soay sheep (Johnston et al., 2016).
+Crosses between the strains we surveyed could be used to identify and
+characterize the genetic variants responsible for recombination rate
+evolution in house mice (Dumont and Payseur, 2011; Wang et al., 2019;
+Wang and Payseur, 2017). These variants could differentially affect
+females and males at any step in the recombination pathway. Although our
+DMC1 profiling was limited to males from a small number of strains (for
+practical reasons), our findings suggest that mutations that determine
+the number of double-strand breaks contribute to sex-specific evolution
+in the recombination rate. A study of two classical inbred strains and
+one wild-derived inbred strain of house mice also found a positive
+association between crossover number and double-strand break number in
+males (Baier et al., 2014).
+
+Another implication of our results is that the connection between
+recombination rate and fitness differs between males and females. Little
+is known about whether and how natural selection shapes recombination
+rate in nature (Dapper and Payseur, 2017; Ritz et al., 2017). Samuk et
+al. (2020) recently used a quantitative genetic test to conclude that an
+8\% difference in genome-wide recombination rate between females from
+two populations of\_Drosophila pseudoobscura\_ was caused by natural
+selection. Applying similar strategies to species in which both sexes
+recombine, including house mice, would be a logical next step to
+understanding the sex-specific evolution of recombination rate.
+
+Population genetic models have been built to explain sexual dimorphism
+in the number and placement of crossovers, which is a common phenomenon
+(Brandvain and Coop, 2012; Sardell and Kirkpatrick, 2020). Modifier
+models predicted that lower recombination rates in males will result
+from haploid selection (Lenormand, 2003) or sexually antagonistic
+selection on coding and cis-regulatory regions of genes (Sardell and
+Kirkpatrick, 2020). Another modifier model showed that meiotic drive
+could stimulate female-specific evolution of the recombination rate
+(Brandvain and Coop, 2012). Although these models fit the conserved
+pattern of sex differences in crossover positions, they do not readily
+explain our observations of sex-specific evolution in the genome-wide
+recombination rate. In particular, the alternation across strains in
+which sex has more crossovers is unexpected.
+
+\textless SAC section 1\textgreater{}
+
+We propose an alternative interpretation of our findings based on the
+cell biology of gametogenesis. During meiosis, achieving a stable
+chromosome structure requires the attachment of kinetochores to opposite
+poles of the cell and at least one crossover to create tension across
+the sister chromosome cohesion distal to chiasmata (Dumont and Desai,
+2012; Lane and Kauppi, 2019; Subramanian and Hochwagen, 2014; VanVeen
+and Hawley, 2003). The spindle assembly checkpoint (SAC) prevents
+aneuploidy by ensuring that all bivalents are correctly attached to the
+microtubule spindle (``bi-oriented'') before starting the
+metaphase-to-anaphase transition via the release of the sister cohesion
+holding homologs together (Lane and Kauppi, 2019). Hence, selection
+seems likely to favor mutations that optimize the process of
+bi-orientation and chromosome separation, thereby prohibiting the SAC
+from delaying the cell cycle or triggering apoptosis. Multiple lines of
+evidence indicate that the SAC is more effective in spermatogenesis than
+in oogenesis (Lane and Kauppi, 2019), perhaps due to the presence of the
+acentrosome spindle (So et al., 2019) and larger cell volume (Kyogoku
+and Kitajima, 2017) in oocytes. The higher stringency of the SAC during
+spermatogenesis suggests that selection will be better at removing
+mutations that interfere with bi-orientation in males than in females.
+Therefore, faster male evolution of the genome-wide recombination rate
+could be driven by the more stringent SAC acting on chromosome
+structures at the metaphase I alignment.
+
+Our SAC model is consistent with other features of our data. We showed
+that widespread sex differences in broad-scale crossover positioning
+(Sardell and Kirkpatrick, 2020) apply across house mice, even in
+lineages where the direction of heterochiasmy is reversed. Faster
+spermatogenesis may select for synchronization of the separation across
+all homologs within the cell (Kudo et al., 2009), whereas in oogenesis,
+the slower cell cycle and multiple arrest stages may require chromosome
+structures with greater stability on the meiosis I spindle, especially
+for those organisms that undergo dictyate arrest (Lee, 2019).
+
+We propose that the SAC model also can explain the correlated evolution
+of stronger crossover interference and higher genome-wide recombination
+rate in male house mice. Our results show that crossovers are spaced
+further apart in strains enriched for double-crossover bivalents when SC
+length is considered and bivalent size effects are minimized. Assuming
+chromatin compaction between (prophase) pachytene and metaphase is
+uniform along bivalents, this increased spacing is expected to expand
+the area for sister cohesion to connect homologs and may improve the
+fidelity of chromosomal segregation. While the SAC model postulates
+direct fitness effects of interference, a modifier model predicted that
+indirect selection on recombination rate -- via its modulation of
+offspring genotypes -- can strengthen interference as well (Goldstein et
+al., 1993).
+
+Regardless of the underlying mechanism, our results provide a rare
+demonstration that crossover interference can diverge over short
+evolutionary timescales. The notion that stronger interference can
+co-evolve with higher genome-wide recombination rate is supported by
+differences between breeds of cattle (Ma et al., 2015). In contrast,
+mammalian species with stronger interference tend to exhibit lower
+genome-wide recombination rates (Otto and Payseur, 2019; Segura et al.,
+2013). The evolution of crossover interference and its relationship to
+changes in crossover number on the genomic scale is a topic deserving of
+more empirical and theoretical work. Our findings further reveal that
 evolution of the genome-wide recombination rate does not require major
@@
 consistently show longer SCs, even in strains with more recombination in
-males. Studies in mice (Lynn et al., 2002, p. @petkov2007) and humans
-(Gruhn et al., 2013, p. @tease2004) suggest that chromosomal axes are
-longer (and DNA loops are shorter) in females than males. Some authors
-have suggested that conserved sex differences in crossover positioning
-(more uniform placement in females) and interference strength (stronger
-interference in males) could be due to looser chromatin packing of the
-meiotic chromosome structure in females (Haenel et al., 2018; Petkov et
-al., 2007). A cellular model designed to explain interference attributes
-sexual dimorphism in chromatin structure to greater cell volumes and
-oscillatory movements of telomeres and kinetochores in oocytes (Hultén,
-2011). More recent work connects the sparser recombination landscape to
-sex differences in the crossover repair pathway (Wang et al., 2017).
-
-\textless new 9/n\textgreater{} Our conclusions are accompanied by
-several caveats. First, MLH1 foci only identify interfering crossovers
-(Holloway et al., 2008). Although most crossovers belong to this class
-\textbf{(REF)}, our approach likely underestimated genome-wide
-recombination rates. Evolution of the number of non-interfering
-crossovers is a topic worth examining. A second limitation is that our
-investigation of crossover locations was confined to the relatively low
-resolution possible with immunofluorescent cytology. Positioning
-crossovers with higher resolution could reveal additional evolutionary
-patterns. Finally, the panel of inbred lines we surveyed may not be
-representative of recombination rate variation within and between
-subspecies of house mice. We considered most available wild-derived
-inbred lines, but house mice have a broad geographic distribution.
-Nevertheless, we expect our primary conclusion that recombination rate
-evolves in a sex-specific manner to be robust to geographic sampling
-because differences between females and males exist for the same set of
-inbred strains.
-
-\textless new 10/n\textgreater{} While the causes of sex differences in
-recombination remain mysterious Lenormand et al. (2016), our conclusions
-have implications for a wide range of recombination research. For
-biologists uncovering the cellular and molecular determinants of
-recombination, our results suggest that mechanistic differences between
-the sexes could vary by genetic background. For researchers charting the
-evolutionary trajectory of recombination, our findings indicate that
-sex-specific comparisons are crucial. For theoreticians building
-evolutionary models of recombination, different fitness regimes and
-genetic architectures in females and males should be considered.
-Elevating sex as a primary determinant of recombination would be a
-promising step toward integrating knowledge of cellular mechanisms with
-evolutionary patterns to understand recombination rate variation in
-nature.
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+males. Studies in mice (Lynn et al., 2002; Petkov et al., 2007) and
+humans (Gruhn et al., 2013; Tease and Hulten, 2004) suggest that
+chromosomal axes are longer (and DNA loops are shorter) in females than
+males. Some authors have suggested that conserved sex differences in
+crossover positioning (more uniform placement in females) and
+interference strength (stronger interference in males) could be due to
+looser chromatin packing of the meiotic chromosome structure in females
+(Haenel et al., 2018; Petkov et al., 2007). A cellular model designed to
+explain interference attributes sexual dimorphism in chromatin structure
+to greater cell volumes and oscillatory movements of telomeres and
+kinetochores in oocytes (Hultén, 2011). Recent work in mice connects the
+sparser recombination landscape in females to sex differences in
+crossover maturation efficiency (Wang et al., 2017).
+
+Our conclusions are accompanied by several caveats. First, MLH1 foci
+only identify interfering crossovers (Holloway et al., 2008). Although
+most crossovers belong to this class (Holloway et al., 2008), our
+approach likely underestimated genome-wide recombination rates.
+Evolution of the number of non-interfering crossovers is a subject worth
+examining. A second limitation is that our investigation of crossover
+locations was confined to the relatively low resolution possible with
+immunofluorescent cytology. Positioning crossovers with higher
+resolution could reveal additional evolutionary patterns. Finally, the
+panel of inbred lines we surveyed may not be representative of
+recombination rate variation within and between subspecies of house
+mice. We considered most available wild-derived inbred lines, but house
+mice have a broad geographic distribution. Nevertheless, we expect our
+primary conclusion that recombination rate evolves in a sex-specific
+manner to be robust to geographic sampling because differences between
+females and males exist for the same set of inbred strains.
+
+While the causes of sex differences in recombination remain mysterious
+Lenormand et al. (2016), our conclusions have implications for a wide
+range of recombination research. For biologists uncovering the cellular
+and molecular determinants of recombination, our results suggest that
+mechanistic differences between the sexes could vary by genetic
+background. For researchers charting the evolutionary trajectory of
+recombination, our findings indicate that sex-specific comparisons are
+crucial. For theoreticians building evolutionary models of
+recombination, different fitness regimes and genetic architectures in
+females and males should be considered. Elevating sex as a primary
+determinant of recombination would be a promising step toward
+integrating knowledge of cellular mechanisms with evolutionary patterns
+to understand recombination rate variation in nature.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+This research was funded by NIH grants R01GM120051 and R01GM100426 to B.
+A. P.. A. L. P. was partly supported by NIH T32GM007133. We are grateful
+to Francisco Pelegri for generous assistance with microscopy.
+
+\hypertarget{competing-interests}{%
+\section{Competing interests}\label{competing-interests}}
+
+The authors declare that there are no competing interests.
 
@@
 in mammalian males. \emph{PLoS genetics} \textbf{10}.
+doi:\href{https://doi.org/10.1371/journal.pgen.1004125}{10.1371/journal.pgen.1004125}
 
@@
 
+\leavevmode\hypertarget{ref-baudat2013}{}%
+Baudat F, Imai Y, De Massy B. 2013. Meiotic recombination in mammals:
+Localization and regulation. \emph{Nature Reviews Genetics}
+\textbf{14}:794--806.
+doi:\href{https://doi.org/10.1038/nrg3573}{10.1038/nrg3573}
+
+\leavevmode\hypertarget{ref-begun1992}{}%
+Begun DJ, Aquadro CF. 1992. Levels of naturally occurring dna
+polymorphism correlate with recombination rates in d. Melanogaster.
+\emph{Nature} \textbf{356}:519--520.
+doi:\href{https://doi.org/10.1038/356519a0}{10.1038/356519a0}
+
+\leavevmode\hypertarget{ref-bell1982}{}%
+Bell G. 1982. The masterpiece of nature: The evolution and genetics of
+sexuality. Berkeley, CA.: University of California Press.
+
+\leavevmode\hypertarget{ref-bolcun2012}{}%
+Bolcun-Filas E, Schimenti J. 2012. Genetics of meiosis and recombination
+in mice. \emph{International review of cell and molecular biology}
+\textbf{298}:179.
+doi:\href{https://doi.org/https://doi.org/10.1016/B978-0-12-394309-5.00005-5}{https://doi.org/10.1016/B978-0-12-394309-5.00005-5}
+
 \leavevmode\hypertarget{ref-brandvain2012scrambling}{}%
@@
 \textbf{190}:709--723.
+doi:\href{https://doi.org/10.1534/genetics.111.136721}{10.1534/genetics.111.136721}
+
+\leavevmode\hypertarget{ref-burt1991}{}%
+Burt A, Bell G, Harvey PH. 1991. Sex differences in recombination.
+\emph{Journal of evolutionary biology} \textbf{4}:259--277.
+doi:\href{https://doi.org/https://doi.org/10.1046/j.1420-9101.1991.4020259.x}{https://doi.org/10.1046/j.1420-9101.1991.4020259.x}
 
@@
 features of meiotic prophase i. \emph{Chromosoma} 1--16.
+doi:\href{https://doi.org/10.1007/s00412-019-00692-x}{10.1007/s00412-019-00692-x}
+
+\leavevmode\hypertarget{ref-campbell2016}{}%
+Campbell CL, Bhérer C, Morrow BE, Boyko AR, Auton A. 2016. A
+pedigree-based map of recombination in the domestic dog genome.
+\emph{G3: Genes, Genomes, Genetics} \textbf{6}:3517--3524.
+
+\leavevmode\hypertarget{ref-charlesworth1993}{}%
+Charlesworth B, Morgan M, Charlesworth D. 1993. The effect of
+deleterious mutations on neutral molecular variation. \emph{Genetics}
+\textbf{134}:1289--1303.
+
+\leavevmode\hypertarget{ref-cutter2013}{}%
+Cutter AD, Payseur BA. 2013. Genomic signatures of selection at linked
+sites: Unifying the disparity among species. \emph{Nature Reviews
+Genetics} \textbf{14}:262--274.
+doi:\href{https://doi.org/https://doi.org/10.1038/nrg3425}{https://doi.org/10.1038/nrg3425}
 
@@
 Royal Society B: Biological Sciences} \textbf{372}:20160469.
-
-\leavevmode\hypertarget{ref-dumont2011}{}%
-Dumont BL, Payseur BA. 2011. Genetic analysis of genome-scale
-recombination rate evolution in house mice. \emph{PLoS genetics}
-\textbf{7}.
+doi:\href{https://doi.org/10.1098/rstb.2016.0469}{10.1098/rstb.2016.0469}
+
+\leavevmode\hypertarget{ref-dumont2011murid}{}%
+Dumont BL, Payseur BA. 2011. Evolution of the genomic recombination rate
+in murid rodents. \emph{Genetics} \textbf{187}:643--657.
+doi:\href{https://doi.org/10.1534/genetics.110.123851}{10.1534/genetics.110.123851}
 
@@
 \textbf{22}:241--249.
+doi:\href{https://doi.org/10.1016/j.tcb.2012.02.007}{10.1016/j.tcb.2012.02.007}
+
+\leavevmode\hypertarget{ref-felsenstein1974}{}%
+Felsenstein J. 1974. The evolutionary advantage of recombination.
+\emph{Genetics} \textbf{78}:737--756.
+
+\leavevmode\hypertarget{ref-fisher1930}{}%
+Fisher RA. 1930. The genetical theory of natural selection. Oxford
+University Press.
 
@@
 determinants. \emph{PloS one} \textbf{6}.
+doi:\href{https://doi.org/10.1371/journal.pone.0020321}{10.1371/journal.pone.0020321}
+
+\leavevmode\hypertarget{ref-geraldes2011}{}%
+Geraldes A, Basset P, Smith KL, Nachman MW. 2011. Higher differentiation
+among subspecies of the house mouse (mus musculus) in genomic regions
+with low recombination. \emph{Molecular ecology} \textbf{20}:4722--4736.
+doi:\href{https://doi.org/10.1111/j.1365-294X.2011.05285.x}{10.1111/j.1365-294X.2011.05285.x}
 
@@
 \emph{Theoretical population biology} \textbf{44}:246--259.
+doi:\href{https://doi.org/10.1006/tpbi.1993.1028}{10.1006/tpbi.1993.1028}
 
@@
 \emph{PloS one} \textbf{8}.
+doi:\href{https://doi.org/10.1371/journal.pone.0085075}{10.1371/journal.pone.0085075}
 
@@
 \textbf{27}:2477--2497.
+doi:\href{https://doi.org/10.1111/mec.14699}{10.1111/mec.14699}
+
+\leavevmode\hypertarget{ref-haldane1922sex}{}%
+Haldane J. 1922. Sex ratio and unisexual sterility in hybrid animals.
+\emph{Journal of genetics} \textbf{12}:101--109.
 
@@
 sequence-level genetic map. \emph{Science} \textbf{363}:eaau1043.
+doi:\href{https://doi.org/10.1126/science.aau1043}{10.1126/science.aau1043}
+
+\leavevmode\hypertarget{ref-handel2010}{}%
+Handel MA, Schimenti JC. 2010. Genetics of mammalian meiosis:
+Regulation, dynamics and impact on fertility. \emph{Nature Reviews
+Genetics} \textbf{11}:124--136.
+doi:\href{https://doi.org/10.1038/nrg2723}{10.1038/nrg2723}
+
+\leavevmode\hypertarget{ref-hassold2001}{}%
+Hassold T, Hunt P. 2001. To err (meiotically) is human: The genesis of
+human aneuploidy. \emph{Nature Reviews Genetics} \textbf{2}:280--291.
+doi:\href{https://doi.org/https://doi.org/10.1038/35066065}{https://doi.org/10.1038/35066065}
+
+\leavevmode\hypertarget{ref-hill1966}{}%
+Hill WG, Robertson A. 1966. The effect of linkage on limits to
+artificial selection. \emph{Genetics Research} \textbf{8}:269--294.
+doi:\href{https://doi.org/https://doi.org/10.1017/S0016672300010156}{https://doi.org/10.1017/S0016672300010156}
 
@@
 meiosis. \emph{PLoS genetics} \textbf{4}.
+doi:\href{https://doi.org/10.1371/journal.pgen.1000186}{10.1371/journal.pgen.1000186}
 
@@
 \textbf{4}:10.
+doi:\href{https://doi.org/10.1186/1755-8166-4-10}{10.1186/1755-8166-4-10}
+
+\leavevmode\hypertarget{ref-huxley1928}{}%
+Huxley J. 1928. Sexual difference of linkage in gammarus chevreuxi.
+\emph{Journal of Genetics} \textbf{20}:145--156.
+
+\leavevmode\hypertarget{ref-inoue2002}{}%
+Inoue K, Lupski JR. 2002. Molecular mechanisms for genomic disorders.
+\emph{Annual review of genomics and human genetics} \textbf{3}:199--242.
+doi:\href{https://doi.org/10.1146/annurev.genom.3.032802.120023}{10.1146/annurev.genom.3.032802.120023}
 
@@
 \textbf{203}:583--598.
+doi:\href{https://doi.org/10.1534/genetics.115.185553}{10.1534/genetics.115.185553}
 
@@
 \emph{Nature genetics} \textbf{36}:1203--1206.
+doi:\href{https://doi.org/10.1038/ng1445}{10.1038/ng1445}
 
@@
 recombination rate. \emph{Nature genetics} \textbf{46}:11.
+doi:\href{https://doi.org/10.1038/ng.2833}{10.1038/ng.2833}
 
@@
 recombination rate. \emph{Science} \textbf{319}:1398--1401.
+doi:\href{https://doi.org/10.1126/science.1152422}{10.1126/science.1152422}
+
+\leavevmode\hypertarget{ref-kudo2009}{}%
+Kudo NR, Anger M, Peters AH, Stemmann O, Theussl H-C, Helmhart W, Kudo
+H, Heyting C, Nasmyth K. 2009. Role of cleavage by separase of the rec8
+kleisin subunit of cohesin during mammalian meiosis i. \emph{Journal of
+cell science} \textbf{122}:2686--2698.
+doi:\href{https://doi.org/10.1242/jcs.035287}{10.1242/jcs.035287}
 
@@
 \textbf{41}:287--298.
+doi:\href{https://doi.org/10.1016/j.devcel.2017.04.009}{10.1016/j.devcel.2017.04.009}
 
@@
 sciences} \textbf{76}:1135--1150.
+doi:\href{https://doi.org/10.1007/s00018-018-2986-6}{10.1007/s00018-018-2986-6}
 
@@
 Medicine and Biology}.
+doi:\href{https://doi.org/10.1002/rmb2.12299}{10.1002/rmb2.12299}
 
@@
 \emph{Genetics} \textbf{163}:811--822.
+
+\leavevmode\hypertarget{ref-lenormandDuthiel_2005}{}%
+Lenormand T, Dutheil J. 2005. Recombination difference between sexes: A
+role for haploid selection. \emph{PLoS biology} \textbf{3}.
+doi:\href{https://doi.org/10.1371/journal.pbio.0030063}{10.1371/journal.pbio.0030063}
 
@@
 the Royal Society B: Biological Sciences} \textbf{371}:20160001.
+doi:\href{https://doi.org/10.1098/rstb.2016.0001}{10.1098/rstb.2016.0001}
+
+\leavevmode\hypertarget{ref-lorch2005}{}%
+Lorch P. 2005. Sex differences in recombination and mapping adaptations.
+\emph{Genetica} \textbf{123}:39.
+doi:\href{https://doi.org/10.1007/s10709-003-2706-4}{10.1007/s10709-003-2706-4}
 
@@
 \emph{PLoS genetics} \textbf{11}.
+doi:\href{https://doi.org/10.1371/journal.pgen.1005387}{10.1371/journal.pgen.1005387}
 
@@
 mice. \emph{Mammalian Genome} \textbf{21}:550--555.
+doi:\href{https://doi.org/https://doi.org/10.1007/s00335-010-9303-5}{https://doi.org/10.1007/s00335-010-9303-5}
+
+\leavevmode\hypertarget{ref-nachman2012}{}%
+Nachman MW, Payseur BA. 2012. Recombination rate variation and
+speciation: Theoretical predictions and empirical results from rabbits
+and mice. \emph{Philosophical Transactions of the Royal Society B:
+Biological Sciences} \textbf{367}:409--421.
+doi:\href{https://doi.org/10.1098/rstb.2011.0249}{10.1098/rstb.2011.0249}
 
@@
 \textbf{13}:493--504.
+doi:\href{https://doi.org/10.1038/nrg3245}{10.1038/nrg3245}
 
@@
 \textbf{53}:19--44.
+doi:\href{https://doi.org/10.1146/annurev-genet-040119-093957}{10.1146/annurev-genet-040119-093957}
 
@@
 male and female germline. \emph{Chromosome research} \textbf{5}:66--68.
+doi:\href{https://doi.org/10.1023/A:1018445520117}{10.1023/A:1018445520117}
 
@@
 \textbf{123}:442--457.
+doi:\href{https://doi.org/10.1038/s41437-019-0252-9}{10.1038/s41437-019-0252-9}
 
@@
 \emph{Trends in Genetics} \textbf{23}:539--542.
+doi:\href{https://doi.org/10.1016/j.tig.2007.08.015}{10.1016/j.tig.2007.08.015}
+
+\leavevmode\hypertarget{ref-petronczki2003}{}%
+Petronczki M, Siomos MF, Nasmyth K. 2003. Un menage a quatre: The
+molecular biology of chromosome segregation in meiosis. \emph{Cell}
+\textbf{112}:423--440.
+doi:\href{https://doi.org/https://doi.org/10.1016/S0092-8674(03)00083-7}{https://doi.org/10.1016/S0092-8674(03)00083-7}
 
@@
 Adaptive or not? \emph{Trends in Genetics} \textbf{33}:364--374.
+doi:\href{https://doi.org/10.1016/j.tig.2017.03.003}{10.1016/j.tig.2017.03.003}
 
@@
 pseudoobscura. \emph{Current Biology}.
+doi:\href{https://doi.org/10.1016/j.cub.2020.03.053}{10.1016/j.cub.2020.03.053}
 
@@
 \textbf{52}:3283--3299.
+doi:\href{https://doi.org/10.1016/j.csda.2007.10.022}{10.1016/j.csda.2007.10.022}
 
@@
 \textbf{280}:20131945.
+doi:\href{https://doi.org/10.1098/rspb.2013.1945}{10.1098/rspb.2013.1945}
 
@@
 \emph{BMC genomics} \textbf{19}:304.
+doi:\href{https://doi.org/10.1186/s12864-018-4705-y}{10.1186/s12864-018-4705-y}
+
+\leavevmode\hypertarget{ref-smith1974}{}%
+Smith JM, Haigh J. 1974. The hitch-hiking effect of a favourable gene.
+\emph{Genetics Research} \textbf{23}:23--35.
+doi:\href{https://doi.org/https://doi.org/10.1017/S0016672300014634}{https://doi.org/10.1017/S0016672300014634}
 
@@
 \textbf{364}:eaat9557.
-
-\leavevmode\hypertarget{ref-stapley_variation_2017}{}%
-Stapley J, Feulner PGD, Johnston SE, Santure AW, Smadja CM. 2017.
-Variation in recombination frequency and distribution across eukaryotes:
-Patterns and processes. \emph{Philosophical Transactions of the Royal
-Society B: Biological Sciences} \textbf{372}:20160455.
-doi:\href{https://doi.org/10.1098/rstb.2016.0455}{10.1098/rstb.2016.0455}
+doi:\href{https://doi.org/10.1126/science.aat9557}{10.1126/science.aat9557}
 
@@
 perspectives in biology} \textbf{6}:a016675.
+doi:\href{https://doi.org/10.1101/cshperspect.a016675}{10.1101/cshperspect.a016675}
 
@@
 \textbf{107}:208--215.
+doi:\href{https://doi.org/10.1159/000080599}{10.1159/000080599}
 
@@
 \emph{Current Biology} \textbf{13}:R831--R833.
+doi:\href{https://doi.org/10.1016/j.cub.2003.12.004}{10.1016/j.cub.2003.12.004}
 
@@
 \textbf{15}:e1008337.
+doi:\href{https://doi.org/10.1371/journal.pgen.1008337}{10.1371/journal.pgen.1008337}
 
@@
 \textbf{206}:1841--1852.
+doi:\href{https://doi.org/10.1534/genetics.117.202382}{10.1534/genetics.117.202382}
 
@@
 human female meiosis. \emph{Cell} \textbf{168}:977--989.
+doi:\href{https://doi.org/10.1016/j.cell.2017.02.002}{10.1016/j.cell.2017.02.002}
 
